### 2007-04-24 doc/hed/counter.tex
Counter library requirements and API proposal

\documentclass[a4paper,11pt]{article}

\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[small]{caption}
\usepackage{url}

\title{ARC Counter Library\\
Requirements and Design}
\author{Markus Nord\'en}

\begin{document}

\maketitle

\begin{abstract}
  This is a draft summary of requirements and a proposition of an API
  for the counter library of the ARC HED. Comments are encouraged and
  expected. In particular, feedback is needed on globality
  (section~\ref{globality}), persistency (section~\ref{persistency})
  and the suggested API (section~\ref{api}).

  Please, send any feedback to: \url{nordugrid-discuss@nordugrid.org}
\end{abstract}

\section{Introduction}

During the ``Uppsala HED Implementation F2F Meeting'', it was decided
that a counter library shall be provided for services and other
components to use for e.g. allocation and housekeeping of resources.

This document contains requirements on that counter library based on
e-mail discussions on the \url{nordugrid-discuss} e-mail list.

\section{Requirements}

\subsection{Purpose}

The purpose of a counter is to provide housekeeping of a resource that
can be allocated and deallocated such as e.g. disk space or network
bandwidth. The counter itself will not be aware of what kind of
resource it limits the use of. Neither will it be aware of what unit
is being used to measure that resource. Counters are thus very similar
to semaphores.

The users of the resource must thus first call the counter in order to
allocate (make a reservation of) an appropriate amount of the
resource, then use the resource and finally call the counter again to
deallocate what was previously allocated.

Furthermore, the users of the resource must agree on what unit they
use to measure the resource, e.g. if disk space is measured in GB, MB,
kB or single bytes. The counter itself will thus only contain a number.

\subsection{Counter types}

Integers provide sufficient resolution for internal representation of
available resources. Floating point numbers could be an alternative,
but no such requirement has been expressed.

If there should arise a need to represent quantities of a resource
that are smaller than one unit, the choice of unit for the
corresponding counter can be reconsidered.

\subsection{Atomicity and Synchronization}

Counters must be able to handle concurrent operations in a consistent
manner. This implies that all operations must be atomic. Furthermore,
no operation may be performed until it is guaranteed that it will
succeed, i.e. a call to the allocation method must block until there
is a sufficiently large amount of the resource free.

No requirements for simultaneous operations on sets of counters have
been requested.

\subsection{Globality}
\label{globality}

Counters for different purposes (limiting different kinds of
resources) may need to have different scopes:

\begin{description}
\item[Single thread] The counter can only be accessed from within a
single thread. There is no concurrency or need for synchronization and
a special counter class is superfluos.
\item[Single process] (several threads) The counter can be accessed
from all threads within a single process, but is not accessible from
outside that process.
\item[Single computer] (several processes) The counter can be accessed
from all processes running on CPUs with access to a common shared
memory. This requires inter-process communication.
\item[Single cluster] (several computers) The counter can be accessed
from any process running on any computer within a cluster.
\item[Entire grid] The counter can be accessed from any process
running on any computer connected to a certain grid -- probably by
means of a counter service.
\end{description}

It deserves to be mentioned here that the scope of the counter itself
is not necessarily the same as the scope of the resource it limits the
use of. For example, there may be counter available only on the front
end machine of a cluster that keeps track of the number of free CPUs
in the entire cluster. This counter can be used by a process that runs
on the front end machine but starts jobs on other computers in the
cluster.

Requests have been made for single process counters, single computer
counters and single cluster counters.

\subsection{Persistency}
\label{persistency}

Different kinds of counters have different lifetimes. The lifetime of
single computer counters is an open question.

\begin{description}
\item[Process lifetime] The counter is initialized at process start up
  and vanishes when the process finishes. Applicable only for single
  process counters.
\item[System lifetime] The counter ts initialized at system start up
  and is reset upon system restart. Possibly applicable for single
  computer counters.
\item[Eternal lifetime] The counter is initialized once and is then
  available until it is explicitly removed. Applicable for single
  cluster, entire grid and possibly also for single computer counters.
\end{description}

\subsection{Identification of counters}

Counters are identified by unique names. The form of a name will
depend of the globality of a counter.

Single process counters may, for example, be implemented as global
objects accessible from all threads of the process. The ``name'' of a
counter will in this case be the address of the corresponding counter
object. This address can be referred to by an arbitrary number of
pointers and references as well as one or zero ``ordinary variables'',
i.e. just as any other object.

A single computer counter or a single cluster counter may be accessed
through an arbitrary number of counter objects in different processes.
There may thus be many counter objects that correspond to the same
counter and manipulates it in a synchronized way. Upon construction of
such a counter object, the name of the counter shall be provided. Such
names can follow the same rules as e.g. identifiers in the C
programming language. Furthermore, the name of some global object,
e.g. a file, will probably also be required as an ``environment'' for
the counters.

The name of an entire grid counter will probably consist of the URI of
a counter service (the ``environment'' part) and a counter name as
described above.

\subsection{Priority}

In some cases it may be desirable to let important tasks exceed the
limit of a counter to a certain extent. This may be implemented by
means of a credit. That an allocation is prioritized may be expressed
by means of an extra, optional, parameter to the allocation method.

\subsection{Time Awareness}

The basic use of a counter is that a certain amount of it is acquired
by a call to an allocation method and subsequently returned by a call
to a deallocation method.

There has also been requests for self expiring allocations, i.e. that
it is specified at allocation for how long that acquisition shall
last. After that time has elapsed, deallocation will occur
automatically.

It has also been mentioned that reservation of resources in advance
would be a nice feature, but this is considered to be less important.

\section{Design and API}
\label{api}

There will be an abstract base class that defines the interface of a
counter, i.e. what operations can be performed on it. The different
kinds of counters (different globality) will be implemented as
subclasses of that base class.

The signature of the constructors of the different classes will be
determined later based on the needs of the implementation.

The following operations will be available for counters:

\subsection{void wait(int amount, bool prioritized, int duration)}

Blocks until the requested amount is free and allocates it.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to be allocated.
\item[prioritized] True if this is a prioritized request, false
  otherwise. Optional, default value is false.
\item[duration] Zero for allocations that subsequently will be
  explicitly deallocated. If positive, duration of a self expiring
  allocation. Optional, default value zero.
\end{description}

\subsection{bool tryWait(int amount, bool prioritized, int duration)}

Tries to allocate the requested amount and returns immediately.
Returns true if the allocation succeded and false if it did not
succeed.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to be allocated.
\item[prioritized] True if this is a prioritized request, false
  otherwise. Optional, default value is false.
\item[duration] Zero for allocations that subsequently will be
  explicitly deallocated. If positive, duration of a self expiring
  allocation. Optional, default value zero.
\end{description}

\subsection{void post(bool amount)}

Deallocates a certain amount.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to be deallocated.
\end{description}

\subsection{int getLimit()}

Returns the current limit of the counter.

\subsection{int setLimit(int limit)}

Sets the limit of the counter. Returns the new limit.

\paragraph{Parameters:}
\begin{description}
\item[limit] The new limit.
\end{description}

\subsection{int changeLimit(void amount)}

Changes the limit of the counter by a certain amount. Returns the new
limit.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to add to the the limit. May be positive or
  negative.
\end{description}

\subsection{int getCredit()}

Returns the current credit of the counter.

\subsection{int setCredit(void credit)}

Sets the limit of the counter. Returns the new limit.

\paragraph{Parameters:}
\begin{description}
\item[credit] The new credit.
\end{description}

\subsection{int changeCredit(void amount)}

Changes the credit of the counter by a certain amount. Returns the new
credit.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to add to the the credit. May be positive or
  negative.
\end{description}

\subsection{int getValue()}

Returns the current value of the counter, i.e. the number of units
currently available for allocation.

\section{Implementation}

The implementation will be based on ``lower level'' libraries. Posix
and System V semaphores have been considered so far but have been
rejected due to limited functionality and portability issues. The
current candidate is glibmm, possibly combined with file storage for
counters that cannot reside in memory.

\end{document}

### 2007-08-24 doc/a-rex/arex_tech_doc.tex
Starting A-Rex technical doc based on Grid Manager's one.

%% LyX 1.4.5 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside,english]{article}
\usepackage{pslatex}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\usepackage{array}
\usepackage{amsmath}
\usepackage{setspace}
\onehalfspacing
\usepackage{amssymb}
\IfFileExists{url.sty}{\usepackage{url}}
                      {\newcommand{\url}{\texttt}}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\newenvironment{LyXParagraphLeftIndent}[1]%
{
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \addtolength{\leftmargin}{#1}
    \setlength{\parsep}{0pt plus 1pt}%
  }
  \item[]
}
{\end{list}}
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newenvironment{lyxlist}[1]
{\begin{list}{}
{\settowidth{\labelwidth}{#1}
 \setlength{\leftmargin}{\labelwidth}
 \addtolength{\leftmargin}{\labelsep}
 \renewcommand{\makelabel}[1]{##1\hfil}}}
{\end{list}}
\newenvironment{lyxcode}
{\begin{list}{}{
\setlength{\rightmargin}{\leftmargin}
\setlength{\listparindent}{0pt}% needed for AMS classes
\raggedright
\setlength{\itemsep}{0pt}
\setlength{\parsep}{0pt}
\normalfont\ttfamily}%
 \item[]}
{\end{list}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{multirow}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {A.Konstantinov},
  pdftitle = {The NorduGrid Grid Manager and GridFTP Server.},
  pdfsubject = {},
  pdfkeywords = {},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%

\usepackage{babel}
\makeatother
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}

\resizebox*{3cm}{!}{\includegraphics{ng-logo}}

&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}

\end{tabular}

\hrulefill

{\raggedleft NORDUGRID-TECH-2\par}

{\raggedleft \today\par}

\vspace*{2cm}

\begin{center}
\textsc{\large The ARC Computational Job Management Module - A-Rex}
\par\end{center}{\large \par}

\vspace*{0.5cm}

\begin{center}
\textit{\large Description and Administrator's Manual}
\par\end{center}{\large \par}

\vspace*{1.5cm}

{\centering \large A.Konstantinov\footnote{aleks@fys.uio.no} \large \par}

\end{titlepage}

\tableofcontents{}

\newpage

\section{Introduction\label{sec:intro}}

A-Rex is one of ARC middleware componets implementing functions of
so called \emph{Computing Element} (CE). It's made of few components
- A-Rex Web Service (WS) Interface and Grid Manager (GM). Alternative
and currently only stable interface to A-Rex is GridFTP Server (GFS).

The aim of the Grid Manager (GM) is to take care of job pre- and post-processing.
It provides functionality to stage-in files containing input data
and program modules from wide range of sources and transfer or store
output results.

Both WS Interface and GridFTP Interface components provide a way to
submit and control computational tasks (jobs) to be executed by the
GM and undelying Local Resource Management System. While WS interface
is currently being developed the GridFTP is considered to be stable
and prefered one.

\textbf{You should use this document for advanced configuration purposes
and understanding of internals of the aforementioned tools. For general
instalaltion and configuration refer to other documents avaialble
at \url{http://www.nordugrid.org/papers.html}. }

\section{Main concepts\label{sec:main concepts}}

A job is a set of input files (which may or may not include executables),
a main executable and a set of output files. The process of gathering
input files, executing a job, and transferring/storing output files
is called a \emph{session}.

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files also in the SD. GM does not guarantee the availability
of any other places accessible by the job other than SD (unless such
place is part of requested Runtime Environement). The SD is also the
only place which is controlled by the GM. It is is accessible by the
user from outside through HTTPS and GridFTP protocols. Any file created
outside the SD is not controlled by the GM. Any exchange of data between
client and GM (including also program modules) is done through GridFTP
\cite{gridftp} or HTTPS protocols depending on selected interface.
A URL for accessing input/output files is obtained through the Information
System or WS Information Interface of A-Rex.

Each job gets an identifier (\textit{jobid}). This is a handle which
identifies the job in the GM and the Information System \cite{is}.

Each job is initiated and controlled through GFS or WS. All job parameters
(not data) are passed to the GM through GFS in RSL \cite{rsl} or
through WS in JSDL-coded \cite{jsdl} description (job description
- JD).

\section{Input/output data\label{sec:input/output data}}

The main task of the GM is to take care of processing input and output
data (files) of the job. Input files are gathered in SD. There are
2 ways to put file into the SD:

\begin{itemize}
\item Downloads initiated by the GM. Such files (name and source) are defined
in the JD. It is a sole responsibility of the GM to make sure that
a file will be available in the SD.\\
The supported sources are at the moment: GridFTP, FTP, HTTP, HTTPS
(HTTP over SSLv3) and HTTPg (HTTP over GSI). Also some nonstandard
sources are supported. Those are described below.
\item Upload initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the GM. An example of such files
can be the main executable of the job, files containing job's options/parameters,
etc. These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE belonging to NorduGrid. Access to AFS, NFS, FTP, HTTP
and any other remote data transport during execution of the job is
not guaranteed (at least not by GM).

Job stores output files in the SD. Those files also belong to 2 groups:

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like \emph{Replica
Catalog} (RC). The GM takes care of those files. They have to be specified
in the JD. If job fails during any stage of processing no attept is
taken to transfer those files to their final destination, unless there
is option \emph{preserve=yes} specified in their URLs.
\item Files which are supposed to be fetched by the user. The user runs
UI to obtain those files. They \textbf{must} also be specified in
the JD.
\end{itemize}

\section{\label{Section: Job Flow}Job flow}

From the point of view of the GM a job passes through various states.
Picture \ref{job states diagram} presents a diagram of the possible
states of a job.%
\begin{figure}[tbh]
\begin{centering}
\includegraphics{pic1}
\par\end{centering}

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the NorduGrid Information
System (IS) using the UI or any other tool. Please remember that IS
can manipulate with state names to make them more user friendly and
to combine them with states introduced by other parts of whole setup.
Another way is to access virtual informational files through GridFTP
interface or to use query metod of JCS.

Configuration can put limits on amout of simultaneous jobs at some
states. If such limit is reached job stays in it's current state waiting
for free slot. This situation is presented by prepending current state
name with \textbf{PENDING:} status mark.

Below is description of all actions taken by the GM at every state:

\begin{itemize}
\item \textbf{Accepted} - At this state the job has been submitted to a
CE but not processed yet. The GM will analyze the JD and move to the
next stage. If JD can not be processed the job will be canceled and
moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD. The
GM is downloading files specified in the JD and waiting for files
which are supposed to be downloaded by the UI. If all files are successfully
gathered the job moves to the next state. If \textbf{any} file can't
be downloaded or it takes UI too long to upload a file - the job moves
to \textbf{Finishing} state. It is possible to put limit on number
of simultaneous \textbf{Preparing} jobs. Those jobs out of limit will
stay in previous \textbf{Accepted} state with PENDING mark. Exceptions
are jobs which has no files to be downloaded. Those are processed
out of limits.
\item \textbf{Submitting} - This is a point of interaction with \emph{Local
Resource Management System} (LRMS). At the moment PBS is supported
best and correspoding backends are provided with default installation.
The job is being submitted for execution. If the local job submission
is successful the job moves to the next state. Otherwise it moves
to \textbf{Finishing}. It is possible to limit number of jobs in \textbf{Submitting}
and following \textbf{InLRMS} states.
\item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
The GM takes no actions except waiting until job finishes.
\item \textbf{Cancelling} - Necessary action to cancel job in the LRMS is
being taken.
\item \textbf{Finishing} - The output data is being processed. Specified
data files are moved to the specified SEs and are optionally registered
at RC. The user can download data files from the SD by using UI or
any other tool. All the files not specified as output files are removed
from the SD at very beginning of this state. It is possible to limit
number of simultaneous jobs in this state.
\item \textbf{Finished} - No more processing is performed by the GM. The
user can continue to download data files from the SD. The SD is kept
available for some time (default is 1 week). After that it is moved
to the state \textbf{Deleted}. The 'deletion' time can be queried
at NorduGrid Information System as attribute \texttt{nordugrid-pbs-job-sessiondirerasetime}
of \texttt{nordugrid-pbs-job}. If job was moved to \textbf{Finished}
because of failure, it may be restarted on request of client. Job
is moved to a state previous to one which failed and is assigned mark
PENDING. This is needed in order to not break the configuration limits.
Exception is a job failed in \textbf{InLRMS} state and lacking input
files specified in JD. Such job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} - Job is moved to this state if user does not request
job to be cleaned. Only minimal subset of information about such job
is kept.
\end{itemize}
In the case of the failure special processing is applied to output
files. All specified output files are treated as \textbf{downloadable
by user}. No files will be moved to the SE.

\section{URLs}

The GM and it's components support following data transfer protocols
and corresponding URLs: \emph{ftp, gsiftp, http, httpg, https, se,
rc} and \emph{rls.} For more information please see {}``Protocols,
Uniform Resource Locators (URL) and extensions supported in ARC''
document \cite{urls}.

\section{\label{section: internals}Internals}

\subsection{Internal Files of the Grid Manager}

For each local UNIX user listed in the GM configuration a \textit{control
directory} exists. In this directory the GM stores information about
jobs belonging to that user. Multiple users can share the same \textit{control
directory}. To make it easier to recover in the case of failure, the
GM stores most information in files rather than in memory. All files
belonging to same job have names starting with \textbf{job.ID.} here
ID is the job identifier.

The files in the control directory and their formats are described
below:

\begin{itemize}
\item \textit{job.ID.status} - current state of the job. It contains one
word of text representing the current state of the job. Possible values
are :

\begin{itemize}
\item ACCEPTED
\item PREPARING
\item SUBMITTING
\item INLRMS
\item FINISHING
\item FINISHED
\item CANCELING
\item DELETED
\end{itemize}
\end{itemize}
See section \ref{Section: Job Flow} for a description of the various
states. Additionally each value can be prepended with a prefix {}``PENDING:''
(like PENDING:ACCEPTED, see section \ref{Section: Job Flow}). That
is used to show that the job is \emph{ready} to be moved to a next
state and it stays in a current state \emph{only} because some limits
set in configuration are exceeded.

\begin{itemize}
\item \textit{job.ID.description} - contains the RSL description of the
job.
\item \textit{job.ID.local} - information about job used by the GM. It consists
of lines of format \textit{{}``name = value''} . Not all of them
are always available. The following names are defined:

\begin{itemize}
\item \textit{subject} - user subject also known as the distinguished name
(DN)
\item \textit{starttime} - the GMT time when the job was accepted represented
in Generalized Time format of LDAP 
\item \textit{lifetime} - time period to live for the SD after job finished
in seconds
\item \textit{cleanuptime} - the GMT time when job to be removed from cluster
and SD deleted in Generalized Time format
\item \textit{notify} - email addresses and flags to send mail to about
job specified status changes
\item \textit{processtime} - the GMT time when to start processing the job
in Generalized Time format
\item \textit{exectime} - the GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} - the GMT time when credentials delegated to job
expire in Generalized Time format
\item \textit{rerun} - number of retries left to run the job
\item \textit{jobname} - name of the job as supplied by the user
\item \textit{lrms} - name of LRMS to run the job at
\item \textit{queue} - name of the queue to run the job at
\item \textit{localid} - job id in LRMS (appears only then the job is at
state \textbf{InLRMS})
\item \textit{args} - list of command-line arguments including the executable
\item \textit{downloads} - number of files to download into SD before execution
\item \textit{uploads} - number of files to upload from SD after execution
\item \textit{gmlog} - directory name which holds files containing information
about job when accessed through GridFTP interface
\item \textit{clientname} - name and ip address:port of client machine (name
is provided by user interface)
\item \textit{clientsoftware} - version of software used to submit job
\item \textit{sessiondir} - SD of job
\item \textit{failedstate} - state at which job failed (available only if
it is possible to restart job)
\item \textit{jobreport} - URL of \emph{logger service} used to keep track
of executed jobs (one requested by user)
\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

\begin{itemize}
\item \textit{job.ID.input} - list of input files. Each line contains 2
values separated by a space. First value contains name of the file
relative to the SD. Second value is an URL or a file description.
Example:

\hspace*{1cm}\textit{input.dat gsiftp://grid.domain.org/dir/input\_12378.dat}

\textit{url} - ordinary URL for gsiftp, ftp, http, https or httpg
protocols with the addition of '\textbf{replica catalog url}' (RC
URL) and '\textbf{replica location service url}' (RLS URL).\\
Each URL can contain additional options.

\textit{file description} - {[}size]{[}.checksum].

\hspace*{1cm}\textit{size} - size of the file in bytes.

\hspace*{1cm}\textit{checksum} - checksum of the file identical to
the one produced by \textbf{\textit{cksum}} (1).

Both size and checksum can be left out. Special kind of file description
{*}.{*} is used to specify files which are \textbf{not} required to
exist.

\end{itemize}
This file is used by the '\textbf{\textit{downloader}}' utility. Files
with 'url' will be downloaded to the SD and files with 'file description'
will simply be checked to exist. Each time a new \textbf{valid} file
appears in the SD it is removed from the list and \textit{job.ID.input}
is updated. Any external tool can thus track the process of collecting
input files by checking \textit{job.ID.input.}

\begin{itemize}
\item \textit{job.ID.output} - list of output files. Each line contains
1 or 2 values separated by a space. First value is the name of the
file relative to the SD. The second value, if present, is a URL. Supported
URLs are the same as those supported by job.ID.input.
\end{itemize}
This file is used by the '\textbf{\textit{uploader}}' utility. Files
with \textit{url} will be uploaded to SE and remaining files will
be left in the SD. Each time a file is uploaded it is removed from
the list and \textit{job.ID.output} is updated. Files not mentioned
as output files are removed from the SD at the the beginning of the
\textbf{Finishing} state.

\begin{itemize}
\item \textit{job.ID.failed} - the existence of this file marks the failure
of the job. It can also contain one or more lines of text describing
the reason of failure. Failure includes the return code different
from zero of the job itself.
\item \textit{job.ID.errors} - this file contains the output produced by
external utilities like \textbf{\textit{downloader}}, \textbf{\textit{uploader}},
script for job submission to LRMS, etc on their stderr handle. Those
are not necessarily errors, but can be just useful information about
actions taken during the job processing. In case of problem include
content of that file while asking for help.
\item \textit{job.ID.diag} - information about resources used during execution
of job and other information suitable for diagnostics and statistics.
It's format is similar to that of \textit{job.ID.local}. The following
names are at least defined:

\begin{itemize}
\item \textit{nodename} - name of computing node which was used to execute
job,
\item \textit{runtimeenvironments} - used runtime environments separated
by ';',
\item \textit{exitcode} - numerical exit code of job,
\item \textit{frontend\_distribution} - name and version of operating system
distribution on frontend computer,
\item \textit{frontend\_system} - name of operating on frontend computer,
\item \textit{frontend\_subject} - subject (DN) of certificate representing
frontend computer,
\item \textit{frontend\_ca} - subject (DN) of issuer of certificate representing
frontend computer,
\end{itemize}
and other information provided by GNU \emph{time} utility. Note that
some implementation of \emph{time} insert unrequested information
in their output. Hence some lines can have broken format.

\item \textit{job.ID.proxy} -delegated GSI proxy.
\item \textit{job.ID.proxy.tmp} - temporary GSI proxy with different unx
ownership used by processes run with effective \emph{user id} different
form job owner's \emph{id}.
\end{itemize}
There are other files with names like job.ID.{*} which are created
and used by different parts of the GM. Their presence in the \textit{control
directory} can not be guaranteed and can change depending on changes
in the GM code.

\subsection{GridFTP Interface}

The GridFTP Interface to job management is implemented through GridFTP
server. The GFS is made of layer implementing GridFTP protocol and
plugins for accessing actual data. Local file access in the GFS is
implemented through plugins (shared libraries). There are 3 plugins
provided with the GFS: \textit{fileplugin.so, gaclplugin.so} and \textit{jobplugin.so}
.

The \textit{fileplugin.so} is intended to be uses for plain file access
with the configuration senitive to user subject and is not needed
for job management interface.

The \textit{gaclplugin.so} uses GACL (\url{http://www.gridpp.ac.uk/authz/gacl/})
to control access to local file system instead of Unix user identity.
It's also not needed for job management.

The \textit{jobplugin.so} is using information about jobs being controlled
by GM and provides access to session directories of the jobs owned
by user. It also provides an interface (virtual directory and virtual
operations) to submit, cancel, clean, renew credentials and obtain
information about the job.

\subsection{Web Service Interface}

A-REX Web Service Specific Interface provides means to submit a description
of a computational job to a computing resource, to stage-in additional
data, to monitor and control processing of jobs, and obtain data produced
during the execution of a job.

\subsubsection{Basic Execution Service Interface}

The job submission and control interface is based on document produced
by OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
The exchange of SOAP messages is performed over HTTPS protocol. The
BES interface is represented by two port-types - BES-Management and
BES-Factory. The former is made to control the A-REX service itself
and thus defines operations to start and stop the functionality of
the BES service. A-REX does not implement remote control of service
functionality. Hence the BES-Management port-type is not implemented.
The BES-Factory port-type provides operations to submit new jobs (to
create an activity in terms of BES) and to monitor its state. It also
has an ability to provide information about the service. A-REX fully
implements the functionality of this port-type.

For job descriptions A-REX accepts the Job Submission Description
Language (JSDL)\cite{jsdl} as defined by the OGF Job Submission Description
Language Working Group. Supported elements and extensions are described
below.

\subsubsection{Extensions to OGSA BES interface}

A-REX introduces a new operation in addition to those provided by
BES. It does that by defining its own port-type A-REX with the single
operation \emph{ChangeActivityStatus} (see Appendix D). This operation
provides a way to request simple transfers between states of jobs
being processed and corresponding actions.

\begin{itemize}
\item \emph{ChangeActivityStatus}

\begin{itemize}
\item Input

\begin{itemize}
\item \emph{ActivityStatusType OldStatus}: Description of the state which
the job is supposed to be in during execution of this request. If
the current state of the job is different from the one having been
given, the operation is aborted and a fault is returned. This parameter
is optional.
\item \emph{ActivityStatusType NewStatus}: Description of the state that
the job is to be put into.
\end{itemize}
\item Output

\begin{itemize}
\item \emph{ActivityStatusType NewStatus}: Description of the current state
of the job.
\end{itemize}
\item Fault(s)

\begin{itemize}
\item \emph{NotAuthorizedFault}: Indicates that client is not allowed to
do this operation.
\item \emph{InvalidActivityIdentifierFault}: There is no such job/activity.
\item \emph{CantApplyOperationToCurrentStateFault}: Requested transition
is not possible.
\end{itemize}
\end{itemize}
\end{itemize}
On result of this command, the job should be put into the requested
state. If such a procedure cannot be performed immediately then the
corresponding sequence is initiated and fault OperationWillBeAppliedEventuallyFault
will be returned.

Since BES allows implementators to extend their initial activity states
with additional sub-states, A-REX defines a set of sub-states of activity
processing in addition to those defined by the BES, as listed in Table
\ref{tab:Job-states-definitions}.

%
\begin{table}

\caption{\label{tab:Job-states-definitions}Job states definitions and mappings}

\begin{tabular}{|c|c|>{\raggedright}p{10cm}|}
\hline 
Applicable BES State&
ARC Sub-state&
\multicolumn{1}{c|}{Description}\tabularnewline
\hline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{2}{2cm}{Pending}&
Accepting&
Job is in the process of being submitted\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Accepted&
Job was submitted\tabularnewline
\hline 
\renewcommand{\multirowsetup}{\centering}\multirow{7}{2cm}{Running}&
Preparing&
Stage-in process is going on\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Prepared&
Stage-in process has finished\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Submitting&
Communication with local batch system is in process\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Executing&
Job is being executed in local batch system\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Killing&
Communication with local batch system to terminate execution is in
process\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Executed&
Job execution in local batch system has finished\tabularnewline
\cline{2-2} \cline{3-3} 
\multicolumn{1}{|c|}{}&
Finishing&
Stage-out process is going on\tabularnewline
\hline 
Cancelled&
Failed&
There was a failure during execution\tabularnewline
\cline{1-1} 
Failed&
\multicolumn{1}{c|}{}&
\multicolumn{1}{l|}{}\tabularnewline
\hline 
All&
Pending&
Job is prevented from going to the next state due to some internal
limits; this sub-state appears in parallel with other sub-states\tabularnewline
\hline 
All&
Held&
Job processing is suspended on client request; this sub-state appears
in parallel with other sub-states\tabularnewline
\hline
\end{tabular}
\end{table}

\subsubsection{Delegation Interface}

ARC service interfaces optionally offer a sub-interface, called the
Delegation Interface (see Appendix E). This is a common purpose interface
to be used by ARC services which accept delegated credentials from
clients. The Delegation Interface implements two operations: initiatialization
of credentials delegation (DelegateCredentialsInit) and update/renewal
of credentials (UpdateCredentials).

\begin{itemize}
\item \emph{DelegateCredentialsInit} operation - this operation performs
the first half of credentials delegation sequence.

\begin{itemize}
\item Input

\begin{itemize}
\item None. On this request the service generates a pair of \emph{public}
and private keys. The public key is then sent to the client in response.
\end{itemize}
\item Output(s)

\begin{itemize}
\item \emph{TokenRequestType TokenRequest}: Contains the public key generated
by the service as a Value element. It also provides an identifier
in the Id element which should be used to refer to the corresponding
private key.
\end{itemize}
\item Fault(s)

\begin{itemize}
\item \emph{UnsupportedFault}: Indicates that the service does not support
this operation despite supporting the port-type.
\item \emph{ProcessingFault}: Internal problems during generation of the
token.
\end{itemize}
\end{itemize}
\item \emph{UpdateCredentials} operation - this operation makes it possible
to update the content of delegated credentials (like in the case of
credentials being renewed) unrelated to other operations of the service.

\begin{itemize}
\item Input

\begin{itemize}
\item \emph{DelegatedTokenType DelegatedToken}: Contains an X509 proxy certificate
based on the public key from the DelegateCredentialsInit signed by
the user?s proxy certificate. Also includes the Id element which identifies
the private key stored at the service side associated with these credentials.
Reference element refers to the object which these credentials should
be applied to in a way specific to the service. The same element must
also be used for delegating credentials as part of other operations
on service.
\end{itemize}
\item Output(s)

\begin{itemize}
\item None.
\end{itemize}
\item Fault(s)

\begin{itemize}
\item \emph{UnsupportedFault}: Indicates that service does not support this
operation despite supporting the port-type.
\item \emph{ProcessingFault}: Internal problems during generation of the
token.
\end{itemize}
\end{itemize}
\end{itemize}
Additionally, A-Rex Web Service Interface allows delegation to be
performed as part of the \emph{CreateActivity} operation of the BES-Factory
port-type. For this it accepts the element \emph{DelegatedCredentials}
inside the \emph{CreateActivity} element. The \emph{Id} element of
\emph{DelegatedCredentials} must contain an identifier obtained in
response to the previous \emph{DelegateCredentialsInit} operation.

\subsubsection{Local Information Description Interface}

The A-REX implements the Local Information Description Interface (LIDI)
interface being common for all ARC services. This interface is based
on OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
Information about resources and maintained activities/jobs are represented
in a \emph{WS-Resource Properties} informational XML document. The
document type is defined in the A-Rex WSDL as a \emph{ResourceInformationDocumentType}.
It contains the following elements/resources:

\begin{lyxlist}{00.00.0000}
\item [{\emph{BESFactory}}] collection of BES Factory attributes as defined
in the BES specifications.
\item [{\emph{Glue2Resource}}] description of a computation resource that
uses Glue2 schema. This one is going to be specified after the Glue2
XML bindings are be available.
\item [{\emph{Activities}}] list of maintained activities. Each activity
contains an identifier (\emph{ActivityIdentifier}), BES (\emph{ActivityDocument})
and Glue2 (\emph{Glue2Job}) descriptions of the activity.
\end{lyxlist}
All information can be accessed either through requests on a particular
resources or through XPath queries using WS-Resource Properties operations.

\subsubsection{Supported JSDL elements}

A-REX supports the following elements from the JSDL version 1.0 specification\cite{jsdl}
including POSIX Applications extension:

\begin{lyxlist}{00.00.0000}
\item [{\emph{JobName}}] name of the job as assigned by the user.
\item [{\emph{Executable}\ (POSIX)}] name of the executable file.
\item [{\emph{Argument}\ (POSIX)}] arguments the executable will be launched
with.
\item [{\emph{DataStaging}}]~

\begin{lyxlist}{00.00.0000}
\item [{Filename}] name of the data file on the executing node.
\item [{\emph{Source}}] source where the file will be taken from before
execution.
\item [{\emph{Target}}] destination the file will be delivered to after
execution.
\end{lyxlist}
\item [{\emph{Input}\,\ (POSIX)}] file to be used as standard input for
the executable.
\item [{\emph{Output}\ (POSIX)}] file to be used as standard output for
the executable.
\item [{\emph{Error}\ (POSIX)}] file to be used as standard error for
the executable.
\item [{\emph{MemoryLimit}\ (POSIX)}] amount of physical memory needed
for execution.
\item [{\emph{TotalPhysicalMemory}}] same as MemoryLimit.
\item [{\emph{IndividualPhysicalMemory}}] same as MemoryLimit.
\item [{\emph{CPUTimeLimit}\ (POSIX)}] maximal amount of CPU time needed
for execution.
\item [{\emph{TotalCPUTime}}] same as CPUTimeLimit.
\item [{\emph{IndividualCPUTime}}] same as CPUTimeLimit.
\item [{\emph{WallTimeLimit}\ (POSIX)}] amount of clock time needed for
execution.
\item [{\emph{TotalCPUCount}}] number of CPUs needed for execution.
\item [{\emph{IndividualCPUCount}}] same as TotalCPUCount.
\end{lyxlist}

\subsubsection{ARC-specific JSDL Extensions}

A-REX accepts JSDL documents having the following additional elements
(see Appendix F):

\begin{lyxlist}{00.00.0000}
\item [{\emph{IsExecutable}}] marks file to become executable after being
delivered to the computing resource.
\item [{\emph{RunTimeEnvironment}}] specifies the name of the Runtime Environment
needed for job execution.
\item [{\emph{Middleware}}] request for specific middleware on the computing
resource frontend.
\item [{\emph{RemoteLogging}}] destination for the usage record report
of the executed job.
\item [{\emph{LocalLogging}}] name for the virtual directory available
through job interface and containing various debug information about
job execution.
\item [{\emph{AccessControl}}] ACL expression which describes the identities
of those clients who are allowed to perform operations on this job.
\item [{\emph{Notify}}] Email destination for notification of job state
changes.
\item [{\emph{SessionLifeTime}}] duration for the directory containing
job-related files to exist after the job finished executing.
\item [{\emph{JoinOutputs}}] specifies if standard output and standard
error channels must be merged.
\item [{\emph{Reruns}}] defines how many times a job is allowed to rerun
in case of failure.
\item [{\emph{CredentialServer}}] URL of MyProxy service which may be used
for renewing the expired delegated job credentials.
\item [{\emph{CandidateTarget}}] specifies host name and queue of a computing
resource.
\end{lyxlist}

\section{Cache\label{sec:cache}}

The GM can cache input files. Caching is enabled if corresponding
command is present in configuration file. The GM does not cache files
marked as executable in job. Caching can also be explicitly turned
off by user for each file by using \textit{cache=no} option in URL
(for URL options read {}``Protocols, Uniform Resource Locators (URL)
and extensions supported in ARC'' \cite{urls}). The disc space occupied
by cache is controlled by removing unused files. For more information
look in section \ref{SubSection: ConfigFile}.

\subsection{Structure}

Cache directory contains plain files. Those are

\begin{itemize}
\item \textit{list} - stores names of the files (8 digit numbers) and corresponding
URLs delimited by blank space. Each pair is delimited by some amount
of \textbackslash{}0 codes. Also creation and expiration times are
stored if available
\item \textit{old} - stores URLs which have been removed from cache. Records
are delimited by some amount of \textbackslash{}0 codes and are meant
to be removed by some external routine.
\item \textit{new} - stores URLs which have been added to cache. Records
are delimited by some amount of \textbackslash{}0 codes and are removed
when corresponding files are removed from cache. They can also be
handled by some external routines. Every time record is added to \emph{old}
it is removed from \emph{new}.
\item \textit{statistics} - consists of strings containing \emph{name=value}
. Following names are defined:

\begin{itemize}
\item \textit{hardsize} -size of file system for storing cached data
\item \textit{hardfree} - amount of disc space available on that file system
\item \textit{softsize} - if cache exceeds this size files are started being
removed 
\item \textit{softfree} - space left till softsize (can be negative)
\item \textit{claimed} - space used by files claimed by running jobs
\item \textit{unclaimed} - space used by files not being currently used
by any job
\end{itemize}
\item \textit{\#\#\#\#\#\#\#\#.info} - stores state of file (\textit{\#\#\#\#\#\#\#\#}
stands for 8 digits). State is represented by one character:\\
\textit{c} - just created, content is empty.\\
\textit{f} - failed to download (treated same as 'c').\\
\textit{r} - ready to be used, content is valid.\\
\textit{d} - being downloaded. 'd' is followed by identifier of application/job
downloading that file. During content's download this file has write
lock set.
\item \textit{\#\#\#\#\#\#\#\#.claim} - stores list of identifiers of applications/jobs
using this file. Identifiers are stored one per line.
\item \textit{\#\#\#\#\#\#\#\#} - files storing content of corresponding
URL. These can be stored in separate directory.
\end{itemize}
Files \textit{list, old , new} and \textit{\#\#\#\#\#\#\#\#.info}
has to be stored on filesystem which has support for files' locking.

\subsection{How it works}

If job requests input file, which can/allowed to be cached, it is
stored in cache directory instead and soft-link is created in the
SD, pointing to that file. Or file can be stored in cache and then
copied to the SD. Last option is more secure and hence advised.

Before downloading file the GM tries to determine it's size and to
preallocate space in cache directory, by writing file of same size.
If that fails (file system has no more space), it tries to remove
oldest cache files, which are not being used by any job. That means
\textbf{hard limit of cache size is space available at file-system}.
In case cache gets full and it is impossible to free any space, download
fails and is retried without using cache.

Before giving access to already cached file the GM contacts initial
file source to check if user is allowed to do that if protocols allows
to do that.

Also file creation or validity times are checked to make sure cached
file is fresh enough. If it is impossible to obtain creation and invalidation
times for file it is invalidated 24 hours after downloaded.

Also the GM checks cache periodically. If used space exceeds high
water-mark given in configuration file (\emph{softsize}) it tries
to remove oldest unused files to reduce size to low water-mark. This
sets soft limit of cache size.

There are 2 kinds of caches supported. Files in \textit{private} cache
are owned by Unix user to which grid user is mapped. Those files are
readable only by that particular Unix user. Another kind of cache
is \textit{shared}. Files are owned by Unix user who started GM and
are readable by everyone.

\section{Files and directories\label{sec:files and directories}}

\subsection{Modules}

The GM consists of few separate executable modules. Those are:

\begin{itemize}
\item \textit{grid-manager} - Main module. It is responsible for processing
the job, moving it through states, running other modules.
\item \textit{downloader} - This is a module responsible for gathering input
files in the SD. It processes the \textit{job.ID.input} file and updates
it.
\item \textit{uploader} - This module is responsible for delivering output
files to the specified SEs and registration at the RC. It processes
and updates the \textit{job.ID.output} file.
\item \emph{cache-register} - Utility to register cached data into Indexing
Services like RC and RLS. It reads and modifies cache informational
files \emph{old} and \emph{new} \ref{sec:cache}. Configuration is
read directly from the GM's configuration file \ref{SubSection: ConfigFile}.
It is run by the GM every 5 minutes.
\item \emph{frontend-info-collector} - Utility to gather information about
frontend and to put it into \emph{job.ID.diag} file.
\item \emph{gm-kick} - Sends signal to the GM though FIFO file to wake it
up. It's used to increase responsiveness of GM.
\end{itemize}
Following modules are always run under Unix account to which user
is mapped.

\begin{itemize}
\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
responsible for sending e-mail notifications to the user. The format
of the mail messages can be easily changed by editing the simple shell
script \textit{smtp-send.sh}. 
\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
Curently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
pseudo-LRMS is supported for testing purposes. This module is responsible
for the job submission to the LRMS.
\item \textit{cancel-{*}-job} - This one is for canceling the job, which
was submitted to LRMS.
\item \textit{scan-{*}-job} -This shell script is responsible for notifying
the GM about completion of the job. It's implementation for PBS system
uses server logs to extract information about jobs. If logs are not
available it uses less reliable \emph{qstat} command for that. other
backends use different techniques.
\end{itemize}
Also there is administrator utility:

\begin{itemize}
\item \textit{gm-jobs} - prints list of jobs available on cluster and amount
of jobs in every state.\\
\hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name]\\
\hspace*{0.5cm}-l-l - print more information about each job,\\
\hspace*{0.5cm}-l-u - pretend utility is run by user with id \emph{uid},\\
\hspace*{0.5cm}-l-l - pretend utility is run by user with name \emph{name}.
\end{itemize}
GM comes with plugins useable for various authorization purposes (see
for example description of \emph{authplugin} command below):

\begin{itemize}
\item \textit{inputcheck} - checks if all input files specified in job description
are downloadable.\\
\hspace*{0.5cm}inputcheck {[}-h] {[}-d debug\_level] RSL\_fle {[}proxy\_file]\\
\hspace*{0.5cm}-lRSL\_file -file with job description,\\
\hspace*{0.5cm}-lproxy\_file - credentials proxy.
\item \textit{lcas} - executes LCAS plugins on credentials and returns 0
if authorization passed.\\
\hspace*{0.5cm} lcas credentials description {[}library {[}db {[}directory]]]\\
\hspace*{0.5cm}-lcredentials - path to file with credentials to authorize,\\
\hspace*{0.5cm}-ldescription - path to file with job description,\\
\hspace*{0.5cm}-llibrary - path to LCAS library (full or relative
to LCAS directory),\\
\hspace*{0.5cm}-ldb - path to LCAS DB file (full or relative to LCAS
directory),\\
\hspace*{0.5cm}-ldirectory - LCAS directory.
\end{itemize}

\subsection{Directories}

The GM is installed into a single installation point referred as \$NORDUGRID\_LOCATION
and following sub-directories are used:

\begin{LyXParagraphLeftIndent}{1cm}
\$NORDUGRID\_LOCATION/bin - tools

\$NORDUGRID\_LOCATION/libexec - program modules used by GM

\$NORDUGRID\_LOCATION/etc - configuration files, deprecated, central
configuration file is used by deault

\$NORDUGRID\_LOCATION/sbin - daemons

\$NORDUGRID\_LOCATION/lib - gridftp server's plugins and API libraries

The GM also uses following directories:

\end{LyXParagraphLeftIndent}
\begin{itemize}
\item \textit{session root directory} - In this directory the SD is created.
It can be multiple directories for the various users specified in
the configuration file.\\
There are 2 processes which need to have a permissions to create new
files and directories in it. Those are GM and GFS. \\
If any of those processes processes are run under dedicated user account,
that account need full permissions in the \textit{session root directory}.\\
If those processes are run under \emph{root} account make sure \textit{session
root directory} is not on filesystem which limits capabilities of
\emph{root} user. For example NFS with \emph{root\_squash} option.\\
If there is need to run processes under \emph{root} account (to run
jobs in LRMS under different users' accounts) but there is no way
to provide suitable \textit{session root directory} use \emph{norootpower}
command in configuration of the GM. In that case GM and GFS will use
identity of local user to which Grid identity is mapped to access
\textit{session root directory}. Hence those users will need full
access there.\\
The GM creates SD with proper ownership and permissions for local
identity used to run job. Some filesystems require \emph{executable}
permissions on \textit{session root directory} to be set for local
identity in order to access any file or subdirectory there.\\
This directory should also be shared among cluster nodes in order
for job to access input files. Or internal means of LRMS must be used
to transfer files to executing node. For more see section \ref{sub:LRMS}.
\item \textit{control directory} - In this directory the SD stores an information
about the accepted jobs. Both GM and GFS processes must have full
permissions there.\\
Also subdirectory called \textit{log} is created there. It is used
to accumulate information about started and finished jobs. This information
is periodically sent to the \emph{logger service}.
\end{itemize}

\section{Configuration}

\subsection{\label{SubSection: ConfigFile}Configuration of the Grid Manager}

The GM configuration is done through single configuration file. Historically
GM supports 2 kinds of configuration files. For old one it looks at
following places:

\begin{itemize}
\item \textit{\$NORDUGRID\_LOCATION/etc/grid-manager.conf}
\item \textit{/etc/grid-manager.conf}
\end{itemize}
And for new one in

\begin{itemize}
\item \textit{/etc/arc.conf}
\end{itemize}
The old configuration file consists of empty lines, lines containing
comment (line starts from \#) or configuration commands. Blank spaces
in arguments must be escaped using '\textbackslash{}' or arguments
must be enclosed in '\char`\"{}'. Command line starts from command
followed by arguments separated from command and between them by spaces.

The new configuration file can also contain empty lines and comments
starting from \#. It is separated into sections. Each sections starts
from string containing \emph{}

\begin{itemize}
\item \emph{{[}section name/subsection name/subsubsection name]}. 
\end{itemize}
Each section continues till next section of end of file. One configuration
file can have commands for multiple services/modules/programs. Each
service get it's own section named after it. The GM uses section \emph{{[}grid-manager]}.
Some services can make use of multiple subsections to reflect their
internal modular structure. Commands in section \emph{{[}common]}
apply to all services. Command lines have format

\begin{itemize}
\item \emph{name=''arguments string''.}
\end{itemize}
Names are same as in old configuration file. The \emph{argument string}
consists of same arguments as in old format. And they must obey same
rules.

Both files support almost same commands. Following commands are defined
(examples are given for new format):

\noindent Global commads (those which affect global parameters of
the GM and affect all serviced users, also described in \cite{config}):

\begin{itemize}
\item \textbf{\textit{daemon}}\textit{=yes|no} - specifies whether the GM
should go to background after started. Defaults to \emph{yes}.
\item \textbf{\textit{logfile}}\textit{={[}path]} - specifies name of file
for logging debug/informational output. Defaults to \emph{/dev/null}
for daemon mode and \emph{stderr} for foreground mode.
\item \textbf{\textit{user}}\textit{={[}uid{[}:gid]]} - specifies user id
(and optionally group id) to which the GM must switch after reading
configuration. Defaults to \emph{not switch}.
\item \textbf{\textit{pidfile}}\textit{={[}path]} - specifies file where
id if GM process will be stored. Defaults to \emph{not write}.
\item \textbf{\textit{debug}}\textit{=number} - specifies level of debug
information. More information is printed for higher levels. Currently
highest effective number is 3 and lowest 0. Defaults to 2.
\end{itemize}
All commands above are generic for every daemon-enabled server in
ARC NorduGrid toolkit (like GFS and HTTPSD).

\begin{itemize}
\item \textbf{\textit{joblog}}\textit{={[}path]} - specifies where to store
log file containing information about started and finished jobs.
\item \textbf{\textit{jobreport}}\textit{={[}URL ... number]} - specifies
that GM has to report information about jobs being processed (started,
finished) to centralized service running at given \textit{URL}. Multiple
entries and multiple URLs are allowed. \textit{number} specifies how
long old records have to be kept if failed to be reported. That time
is in days. Last specified value becomes effective.
\item \textbf{\textit{securetransfer}}\textit{=yes|no} - specifies whether
to use encryption while transferring data. Currently works for GridFTP
only. Default is no. It is overridden by value specified in URL options.
\item \textbf{\textit{localtransfer}}\textit{=yes|no} - specifies whether
to pass file downloading/uploading task to computing node. If set
to yes the GM won't download/upload files. Instead it composes script
submitted to LRMS in way to make it do that. This requires installation
of GM and Globus to be accessible from computing nodes and environment
variables GLOBUS\_LOCATION and NORDUGRID\_LOCATION to be set accordingly.
Default is no.
\item \textbf{\textit{maxjobs}}\textit{={[}max\_processed\_jobs {[}max\_running\_jobs]]}
- specifies maximum number of jobs being processed by the GM at different
stages:\\
\textit{max\_processed\_jobs} - maximal amount of jobs being processed
by GM. This does not limit amount of jobs, which can be suNOTE:bmitted
to cluster\\
\textit{max\_running jobs} - maximal amount of jobs passed to Local
Resource Management System\\
Missing value or -1 means no limit.
\item \textbf{\textit{maxlod}}\textit{={[}max\_frontend\_jobs {[}emergency\_frontend\_jobs
{[}max\_transferred\_files]]]} - specifies maximum load caused by
jobs being processed on frontend:\\
\textit{max\_frontend\_jobs} - maximal amount of jobs heavily using
resources of frontend (applied before moving job to PREPARING and
FINISHING states)\textit{}\\
\textit{emergency\_frontend\_jobs} - if limit of \textit{max\_frontend\_jobs}
is used only by PREPARING or by FINISHING jobs aforementioned number
of jobs can be moved to another state .This is used to avoid case
then jobs can't finish due to big amount of recently submitted jobs.\textit{}\\
\textit{max\_transfered\_files} - maximal number of files being transfered
in parallel by every job. Used to decrease load on not so powerful
frontends.\\
Missing value or -1 means no limit.
\item \textbf{\textit{wakeupperiod}}\textit{=time} - specifies how often
for external changes are performed (like new arrived job, job finished
in LRMS, etc.). \textit{time} is a minimal time period specified in
seconds. Default is 3 minutes.
\item \textbf{\textit{cacheregistration}}\textit{=yes|no} - enables or disables
registration of cache data into Indexing Services like RC or RLS.
The default is \emph{no}. Only files dowmloaded through \emph{meta-url}
are registred. Registration is done to same service used for obtaining
information about file. For this opeartion credentials of the GM (host
key and certificate) are used. If required new files storage location
is registered at Indexing Service with quasi-url \emph{cache://hostname/}
and name \emph{hostname:cache} .
\item \textbf{\textit{authplugin}}\textit{=state options plugin} - specifies
\emph{plugin} (external executable) to be run every time job is going
to switch to \emph{state}. Following states are allowed: ACCEPTED,
PREPARING, SUBMIT, FINISHING, FINISHED and DELETED. If exit code is
not 0 job is canceled by default. \textit{Options} consist of \textit{name}=\textit{value}
pairs separated by a comma. Following \textit{name}s are supported:\\
\textit{timeout} - specifies how long in seconds execution of the
plugin allowed to last (mandatory, {}``\textit{timeout=}{}`` can
be skipped for backward compatibility).\\
\textit{onsuccess}, \textit{onfailure} and \textit{ontimeout} - defines
action taken in each case (\textit{onsuccess} happens if exit code
is 0). Possible actions are:\\
\textit{pass} - continue execution,\\
\textit{log} - write information about result into logfile and continue
execution,\\
\textit{fail} - write information about result into logfile and cancel
job.
\item \textbf{\textit{localcred}}\textit{=timeout plugin} - specifies \emph{plugin}
(external executable or function in shared library) to be run every
time job has to do something on behalf of local user. Execution of
\emph{plugin} may not last longer than \emph{timeout} seconds. If
\emph{plugin} looks like \emph{function@path} then function \emph{int
function(char{*},char{*},char{*},...)} from shared library \emph{path}
is called (\emph{timeout} is not functional in that case). If exit
code is not 0 current operation will fail.
\item \textbf{\textit{norootpower}}\textit{=yes/no} - if set to yes all
processes involved in job management will use local identity of a
user to which Grid identity is mapped in order to access filesystem
at path specified in \textbf{\textit{session}} command (see below).
Sometimes this may involve running temporary external process.
\item \textbf{\textit{allowsubmit}}\textit{={[}group ...]} - list of authorization
groups of users allowed to submit new jobs while \char`\"{}allownew=no\char`\"{}
is active in \emph{jobplugin.so} configuration (see below in section
\ref{subsec: configfile of the gfs}). Multiple commands are allowed.
\item \textbf{\textit{speedcontrol}}\textit{=min\_speed min\_time min\_average\_speed
max\_inactivity} - specifies how long/slow data fransfer is allowed
to take place. Transfer is canceled if transfer rate (bytes per second)
is lower than \emph{min\_speed} for at least \emph{min\_time} seconds,
or if average rate is lower than \emph{min\_average\_speed}, or no
data is receved for longer than \textit{max\_inactivity} seconds.
\item \textbf{\textit{copyurl}}\textit{=template replacement} - specifies
that URLs, starting from template should be accessed in a different
way (most probably Unix open). The \textit{template} part of the URL
will be replaced with \textit{replacement.} \textit{replacement} can
be either URL or local path starting from '/'. It is advisable to
end template with '/'.
\item \textbf{\textit{linkurl}}\textit{=template replacement {[}node\_path]}
- mostly identical to \textit{copyurl} but file won't be copied. Instead
soft-link will be created. \textit{replacement} specifies the way
to access the file from the frontend, and is used to check permissions.
The \textit{node\_path} specifies how the file can be accessed from
computing nodes, and will be used for soft-link creation. If \textit{node\_path}
is missing - \textit{local\_path} will be used instead. Both \textit{node\_path}
and \textit{replacement} should not be URLs.

NOTE: URLs which fit into \textit{copyurl} or \textit{linkurl} are
treated as more easily accessible than other URLs. That means if GM
has to choose between few URLs from which should it download input
file, these will be tried first.

\end{itemize}
\noindent Per UNIX user commands:

\begin{itemize}
\item \textbf{\textit{mail}}\textit{=e-mail\_address} - specifies an email
address \textbf{from} which the notification mails are sent.
\item \textbf{\textit{defaultttl}}\textit{=ttl {[}ttr]} - specifies the
time in seconds for the SD to be available after job finished (\emph{ttl})
and after job was deleted (\emph{ttr}) due to \emph{ttl}. Defaults
are 7days for \emph{ttl} and 30 days for \emph{ttr}.
\item \textbf{\textit{defaultlrms}}\textit{=default\_lrms\_name default\_queue\_name}
- specifies names for the LRMS and queue. Queue name can also be specified
in the JD (currently it is not allowed to override used LRMS by using
JD). In new configuration file this command is called \textbf{\textit{lrms}}.
\item \textbf{\textit{session}}\textit{=path} - specifies path to the directory
in which the SD is created. If the path is {*} the default one is
used - \textit{\$HOME/.jobs} . In new configuration file this command
is called \textbf{\textit{sessiondir}}.
\item \textbf{\textit{cache}}\textit{dir=path {[}link\_path]} - specifies
the directory to store cached data. Empty \emph{path} disables caching.
Default is not to cache data. Optional \textit{link\_path} specifies
the path at which cache is accessible at computing nodes. If \textit{link\_path}
is set to '.' files are not soft-linked, but copied to session directory.
In old configuration file this command is called \textbf{\textit{cache}}.
\item \textbf{\textit{privatecache}}\textit{=path {[}link\_path]} - same
as \textit{cache} command, but cache belongs (owned) to user. For
shared caches use 'cache'.
\item \textbf{\textit{cachedata}}\textit{=path} - allows to specify separate
place to store cache files containing data itself. This can be useful
in case of big data storage available only on NSF server which does
not support file locking. If command or \textit{path} is missing -
default is to store data at place specified in \textit{cache} or \textit{privatecache}
command, together with control files. 
\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} - specifies
high and low water-mark for space used by cache. Values are specified
in bytes. Both \textit{high\_mark} and \textit{low\_mark} can be negative
values. In that case corresponding positive value means space left
on filesystem. If \textit{low\_mark} is omitted it becomes equal to
\textit{high\_mark}. By default this feature is turned off. To turn
it off explicitly \textit{cachesize} without parameters should be
specified. If turned off cache will grow up till it fills whole file
system.
\item \textbf{\textit{maxrerun}}\textit{=number} - specifies maximal number
of times job will be allowed to rerun after it failed in LRMS. Default
value is 2. This only specifies a upper limit. Actual number is provided
in job description and defaults to 0.
\end{itemize}
All per-user commands should be put before \textit{control} command
which initiates serviced user.

\begin{itemize}
\item \textbf{\textit{control}}\textit{=path username {[}username {[}...]]}
- This option initiates UNIX user as being serviced by the GM. \textit{path}
refers to the control directory (see section \ref{section: internals}
for the description of control directory). If the path is {*} the
default one is used - \$HOME/.jobstatus . \textit{username} stands
for UNIX name of the local user. Multiple names can be specified.
If the name is {*} it is substituted by all names found in file /etc/grid-security/grid-mapfile
(for the format of this file one should study the Globus project \cite{globus}).\\
Also the special name '.'(dot) can be used. Corresponding control
directory will be used for \textbf{any} user. This option should be
the last one in the configuration file. In new configuration file
command \textbf{\textit{controldir}}\textit{=path} is also available.
It uses special username '.' and is always executed last independent
of placement in file.
\item \textbf{\textit{helper}}\textit{=username command {[}argument {[}argument
{[}...]]]} - associates external program with the local UNIX user.
This program will be kept running under account of the specified user.
\textit{username} stands for the name of the user. Special names can
be used: '{*}' - all names from /etc/grid-security/grid-mapfile, '.'
- root user. The user should be already configured with \textit{control}
option (except root, who is always configured). \textit{command} is
an executable and \textit{argument}s are passed as arguments to it.
\end{itemize}
Following are global commands supported only in new configuration
file. Most of them are specific to underlying LRMS (PBS in this case)
and are passed in environment variables if old configuration file
is used.

\begin{itemize}
\item \textbf{\textit{pbs\_bin\_path}}\textit{=path} - path to directory
which contains PBS commands.
\item \textbf{\textit{pbs\_log\_path}}\textit{=path} - path to directory
with PBS server's log files.
\item \textbf{\textit{gnu\_time}}\textit{=path} - path to \emph{time} utility.
\item \textbf{\textit{tmpdir}}\textit{=path} - path to directory for temporary
files.
\item \textbf{\textit{runtimedir}}\textit{=path} - path to directory which
contains \emph{runtimenvironment} scripts.
\item \textbf{\textit{shared\_filesystem}}\textit{=yes|no} - if compiting
nodes have an access to session directory through a shared filesystem
like NFS. Corresponds to an environement variable RUNTIME\_NODE\_SEES\_FRONTEND.
\item \textbf{\textit{nodename}}\textit{=command} - command to obtain hostname
of computing node.
\item \textbf{\textit{scratchdir}}\textit{=path} - path on computing node
where to move session directory before execution.
\item \textbf{\textit{shared\_scratch}}\textit{=path} - path on frontend
where \textbf{\textit{scratchdir}} can be found.
\item \textbf{\textit{nodename}}\textit{=command} - command to obtain hostname
of computing node.
\end{itemize}
In the command arguments (paths, executables, ...) following substitutions
can be used:

\begin{lyxlist}{00.00.0000}
\item [{\%R}] - session root - see command \emph{session}
\item [{\%C}] - control dir - see command \emph{control}
\item [{\%U}] - username
\item [{\%u}] - userid - numerical
\item [{\%g}] - groupid - numerical
\item [{\%H}] - home dir - home specified in /etc/passwd
\item [{\%Q}] - default queue - look command 'defaultlrms'
\item [{\%L}] - default lrms - look command 'defaultlrms'
\item [{\%W}] - installation path - \$\{NORDUGRID\_LOCATION\}
\item [{\%G}] - globus path - \$\{GLOBUS\_LOCATION\}
\item [{\%c}] - list of all control directories
\item [{\%I}] - job's ID (for plugins only, substituted in runtime)
\item [{\%S}] - job's state (for \emph{authplugin} plugins only, substituted
in runtime)
\item [{\%O}] - reason (for \emph{localcred} plugins only, substituted
in runtime).\\
Possible reasons are:

\begin{lyxlist}{00.00.0000}
\item [{new}] - new job, new credentials
\item [{renew}] - old job, new credentials
\item [{write}] - write/delete file, create/delete directory (through gridftp)
\item [{read}] - read file, directory, etc. (through gridftp)
\item [{extern}] - call external program (grid-manager)
\end{lyxlist}
\end{lyxlist}
Some configuration parameters can be specified from command line while
starting the GM:

\emph{grid-manager {[}-h] {[}-C level] {[}-d level] {[}-c path] {[}-F]
{[}-U uid{[}:gid]] {[}-L path] {[}-P path]}\\
\hspace*{0.5cm}\emph{-h} - short help,\\
\hspace*{0.5cm}\emph{-d} - debug level,\\
\hspace*{0.5cm}\emph{-L} - name log file (overwrites value in configuration
file),\\
\hspace*{0.5cm}\emph{-P} - name for file containing process id (overwrites
value in configuration file),\\
\hspace*{0.5cm}\emph{-U} - user and gropu id to use for running daemon,\\
\hspace*{0.5cm}\emph{-F} - do not make process daemon,\\
\hspace*{0.5cm}\emph{-c} - name od configuration file,\\
\hspace*{0.5cm}\emph{-C} - remove old information before starting:
1- remove finished jobs, 2 - remove active jobs too, 3- also remove
everything that looks like junk.\\

\subsection{\label{subsec: configfile of the gfs}Configuration of the GridFTP
Server}

Default location of the GFS configuration file is \textit{/etc/arc.conf}
or \textit{\$NORDUGRID\_LOCATION/etc/gridftpd.conf}. Format of these
configuration files is similar to that of the GM. It also supports
generic commands described at the beginning of previous section \ref{SubSection: ConfigFile}.
In the new format sections {[}common] and {[}gridftpd] are used. Commands
specific to the GFS are described below.

\begin{itemize}
\item \textbf{\textit{port}}\textit{=number} - specifies TCP/IP port number.
Default is 2811.
\item \textbf{\textit{include}}\textit{=path} - include contents of another
file. Generic commands can't be specified there.
\item \textbf{\textit{encryption}}\textit{=yes|no} - specifies if server
will allow data transfer to be encrypted. Default is yes.
\item \textbf{\textit{pluginpath}}\textit{=path} - specifies the path where
plugin libraries are installed
\item \textbf{\textit{allowunknown}}\textit{=yes|no} - if set to \emph{yes}
clients are not checked agains grid-mapfile. Hence only access rules
specified in this configuration file will be applied.
\item \textbf{\textit{firewall}}\textit{=hostname} - use IP address of the
\textit{hostname} in response to PASV command instead of IP address
of a network interface of computer. You can write IP address directly
instead of \textit{hostname}. This command may be if server is situated
behind NAT.
\item \textbf{\textit{unixgroup}}\textit{=group rule} - define local UNIX
user and optionally UNIX group to which user belonging to specified
authorization \textit{group} is mapped (see Section \ref{sub:Authorization}
for definition of group). Local names are obtained from specified
\textit{rule}. If specified rule could not produce any mapping, next
command is used. Mapping stops at first matched rule. Following rules
are supported:

\begin{itemize}
\item \textbf{\textit{mapfile}} \textit{file} - user's subject is matched
against list of subjects stored inspecified file, one per line followed
by local UNIX name.
\item \textbf{\textit{simplepool}} \textit{directory} - user is assigned
one of local UNIX names stored in a file \textit{directory/pool},
one per line. Used names are stored in other files placed in the same
\textit{directory}. If UNIX name was not used for 10 days, it may
be reassigned to another user.
\item \textbf{\textit{lcmaps}} \textit{library directory database} - call
LCMAPS functions to do mapping. Here \textit{library} is path to shared
library of LCMAPS, either absolute or relative to \textit{directory};
\textit{directory} is path to LCMAPS installation directory, equivalent
of LCMAPS\_DIR variable; \textit{database} is path to LCMAPS database,
equivalent to LCMAPS\_DB\_FILE variable. Each arguments except \textit{library}
is optional and may be either skiped or replaced with '{*}'.
\item \textbf{\textit{mapplugin}} \textit{timeout plugin} {[}\textit{arg1}
{[}\textit{arg2} {[}...]]] - run external \emph{plugin} executable
with specified arguments. Execution of \emph{plugin} may not last
longer than \emph{timeout} seconds. Rule matches if exit code is 0
and there is UNIX name printed on \emph{stdout}. Name may be optionaly
followed by UNIX group separated by ':'. In arguments following substitions
are applied before plugin is started:

\begin{itemize}
\item \%D - subject of users's cerificate,
\item \%P - name of credentials' proxy file.
\end{itemize}
\end{itemize}
\item \textbf{\textit{unixvo}}\textit{=vo rule} - same as \textbf{\textit{unixgroup}}
for users belonging to Virtual Organization (VO) \textit{vo}.
\item \textbf{\textit{unixmap}}\textit{={[}unixname]{[}:unixgroup] rule}
- define local UNIX user and optionally group used to represent connected
client. \textit{rule} is one of those allowed for \textbf{\textit{authorization
groups}} (see Section \ref{sub:Authorization}) and for \textbf{\textit{unixgroup}}/\textbf{\textit{unixvo}}.
In case of mapping rule username is one, provided by rule. Otherwise
specified \textit{unixname:unixgroup} is taken. Both \textit{unixname}
and \textit{unixgroup} may be either omited or set to '{*}' to specify
missing value.
\item \textbf{\textit{groupcfg}}\textit{=name} - is put into subsections
representnig plugin or {[}group] section and defines if that section
is effective. In old format it selects the group to which all following
lines apply. Only unaffected option is \textbf{\textit{groupcfg}}.
If name is empty (or no groupcfg is used at all) following lines apply
to all users.
\end{itemize}
Subsections of \emph{gridftpd} section specifies plugins which serve
virtual FTP path (similar to mount command of UNIX). Name of subsection
is irrelevant. In old format this section starts with command \textbf{\textit{plugin}}
\textit{path library\_name} and ends with keyword \textbf{\textit{end}}.
Inside subsection following commands are supported

\begin{itemize}
\item \textbf{\textit{plugin}}\textit{=library\_name} - use plugin \textit{library\_name}
to serve virtual path.
\item \textbf{\textit{path}}\textit{=path} - virtual path to serve.

GFS comes with 3 plugins: \textit{filepligin.so, gaclplugin.so} and
\textit{jobplugin.so}.

\begin{itemize}
\item \textit{jobplugin.so} does not require any specific options in case
of old configuration format. It reads the configuration file of the
GM located at the standard place as specified in the section \ref{SubSection: ConfigFile}.
Following options are supported:

\begin{itemize}
\item \textbf{\textit{configfile}}\textit{=path} - defines non-standard
place for GM's configuration file,
\item \textbf{\textit{allownew}}\textit{=yes|no} - specifies if new jobs
can be submitted. Default is \emph{yes}.
\item \textbf{\textit{unixgroup}}/\textbf{\textit{unixvo}}/\textbf{\textit{unixmap}}
- same options like in top-level GFS configuration. If mapping succeeds
obtained local user will be used to run submitted job.
\end{itemize}
\item \textit{filepligin.so} supports following options:

\begin{itemize}
\item \textbf{\textit{mount}}\textit{=path} - defines the place on local
filesystem to which file access operations apply
\item \textbf{\textit{dir}}\textit{=path options} - specifies access rules
for accessing files in \textit{path} (relative to virtual and real
path) and all the files below. \\
\textit{options} is the list of the following keywords:

\begin{itemize}
\item \textbf{\textit{nouser}} - do not use local file system rights, only
use those specifies in this line
\item \textbf{\textit{owner}} - check only file owner access rights
\item \textbf{\textit{group}} - check only group access rights
\item \textbf{\textit{other}} - check only \char`\"{}others\char`\"{} access
rights
\end{itemize}
The options above are exclusive. If none of the above specified usual
Unix access rights are applied.

\begin{itemize}
\item \textbf{\textit{read}} - allow reading files
\item \textbf{\textit{delete}} - allow deleting files
\item \textbf{\textit{append}} - allow appending files (does not allow creation)
\item \textbf{\textit{overwrite}} - allow overwriting already existing files
(does not allow creation, file attributes are not changed)
\item \textbf{\textit{dirlist}} - allow obtaining list of the files
\item \textbf{\textit{cd}} - allow to make this directory current
\item \textbf{\textit{create}} \textit{owner:group permissions\_or:permissions\_and}
- allow creating new files. File will be owned by \textit{owner} and
owning group will be \textit{group}. If '{*}' is used, the user/group
to which connected user is mapped will be used. The permissions will
be set to \textit{permissions\_or} \& \textit{permissions\_and} (second
number is reserved for the future usage). 
\item \textbf{\textit{mkdir}} \textit{owner:group permissions\_or:permissions\_and}
- allow creating new directories.
\end{itemize}
\end{itemize}
\item \textit{gaclpligin.so} does not have options in case of the old configuration.
First line of it's configuration contains local path (root directory)
served by it. Rest till keyword \textbf{\textit{end}} contains GACL
\cite{gacl} XML used to setup initial access rules for every newly
created file and directory. If GACL XML is empty then there will be
no default ACLs created for new files and directories. That means
ACL of parent directory will be used. \\
For the new configuration format following options are supported:
\textbf{\textit{gacl=}}\textit{gacl} - GACL XML, \textbf{\textit{mount}}\textit{=path}
- local path server by plugin. \\
XML may contain variables which are replaced with values taken from
client's credentials. Following variables are supported:

\begin{lyxlist}{00.00.0000}
\item [{\emph{\$subject}}] - subject of user's certificate (DN),
\item [{\emph{\$voms}}] - subject of VOMS\cite{voms} server (DN),
\item [{\emph{\$vo}}] - name of VO (from VOMS certificate),
\item [{\emph{\$role}}] - role (from VOMS certificate),
\item [{\emph{\$capability}}] - capabilities (from VOMS certificate),
\item [{\emph{\$group}}] - name of group (from VOMS certificate) .
\end{lyxlist}
Additionally root directory must contain \emph{.gacl} file with initial
ACL. Otherwise rule will be {}``deny all for everyone''.

\end{itemize}
\end{itemize}
Some configuration parameters can be specified from command line while
starting the GFS:

\emph{gridftpd {[}-h] {[}-p number] {[}-n number] {[}-b number] {[}-B
number] {[}-d level] {[}-c path] {[}-F] {[}-U uid{[}:gid]] {[}-L path]
{[}-P path]}\\
\hspace*{0.5cm}\emph{-h} - short help,\\
\hspace*{0.5cm}\emph{-d} - debug level,\\
\hspace*{0.5cm}\emph{-L} - name log file (overwrites value in configuration
file),\\
\hspace*{0.5cm}\emph{-P} - name for file containing process id (overwrites
value in configuration file),\\
\hspace*{0.5cm}\emph{-U} - user and gropu id to use for running daemon,\\
\hspace*{0.5cm}\emph{-F} - do not make process daemon,\\
\hspace*{0.5cm}\emph{-c} - name od configuration file,\\
\hspace*{0.5cm}\emph{-p} - TCP/IP port number,\\
\hspace*{0.5cm}\emph{-n} - maximal number of simultaneously served
connection,\\
\hspace*{0.5cm}\emph{-b} - default size of buffer used for data transfer
(default is 64kB),\\
\hspace*{0.5cm}\emph{-B} - maximal size of buffer used for data transfer
(default is 640kB).

\subsection{Authorization\label{sub:Authorization}}

Authorization is performed at GFS by applying set of rules. Each rule
takes one line in the \emph{group} section. For information about
supported rules please read {}``Configuration and authorisation of
ARC (NorduGrid) Services'' \cite{config}.

\subsection{\label{sub:LRMS}LRMS support}

The GM comes with support for several LRMS. And this number is slowly
growing. Features explained below are for \textbf{PBS} backend. This
support is provided through \textit{submit-pbs-job}, \textit{cancel-pbs-job},
\textit{scan-pbs-job} scripts. \textit{submit-pbs-job} creates job's
script and submits it to PBS. Created job's script is responsible
for moving data between frontend machine and cluster node (if required)
and execution of actual job. Alternatively it can download input files
and upload output if \emph{{}``localtransfer no''} is specified
in the configuration file.

Behavior of submission script is mostly controlled using environment
variables. Most of them can be specified on frontend in GM's environment
and overwritten on cluster's node through PBS configuration. Some
of them may be set in configuration file too.

\textbf{\textit{PBS\_BIN\_PATH}} - path to PBS executables. Like \emph{/usr/local/bin}
for example. \emph{pbs\_bin\_path} configuration command.

\textbf{\textit{PBS\_LOG\_PATH}} - path to PBS server logs. \emph{pbs\_log\_path}
configuration command.

\textbf{\textit{TMP\_DIR}} - path to directory to store temporary
files. Default value is \emph{/tmp}. \emph{tmpdir} configuration command.

\textbf{\textit{RUNTIME\_CONFIG\_DIR}} - path where runtime setup
scripts can be found. \emph{runtimedir} configuration command.

\textbf{\textit{GNU\_TIME}} - path to GNU time utility. It is important
to path to utility compatible with GNU time. If such utility is not
available, modify \textit{submit-pbs-job} to either reset this variable
or change usage of available utility. \emph{gnu\_time} configuration
command.

\textbf{\textit{NODENAME}} - command to obtain name of cluster's node.
Default is \emph{/bin/hostname -f}. \emph{nodename} configuration
command.

\textbf{\textit{RUNTIME\_LOCAL\_SCRATCH\_DIR}} - if defined should
contain path to the directory on computing node, which can be used
to store job's files during execution. \emph{scratchdir} configuration
command.

\textbf{\textit{RUNTIME\_FRONTEND\_SEES\_NODE}} - if defined should
contain path corresponding to \textit{RUNTIME\_LOCAL\_SCRATCH\_DIR}
as \textbf{}seen on \textbf{frontend} machine. \emph{shared\_scratch}
configuration command.

\textbf{\textit{RUNTIME\_NODE\_SEES\_FRONTEND}} - if set to {}``\emph{no}''
means computing node does not share filesystem with frontend. In that
case content of the SD is moved to computing node by using means provided
by the LRMS. Results are moved back after job's execution in a same
way. \emph{shared\_filesystem} configuration command.

Figures \ref{fig:no node scratch},\ref{fig:node scratch, not vis on front},\ref{fig:node scratch, vis on front}
present some possible combinations for RUNTIME\_LOCAL\_SCRATCH\_DIR
and RUNTIME\_FRONTEND\_SEES\_NODE and explain how data movement is
performed. Pictures a) correspond to situation right after all input
files are gathered in session directory and actions taken right after
job's script starts. Pictures b) show how it looks while job is running
and actions which are taken right after it finished. Pictures c) stand
for final situation, when job files are ready to be uploaded to external
storage element or be downloaded by user.

%
\begin{figure}[tbh]
\begin{centering}
\includegraphics{pic2}
\par\end{centering}

\caption{\label{fig:no node scratch}Both RUNTIME\_LOCAL\_SCRATCH\_DIR and
RUNTIME\_FRONTEND\_SEES\_NODE undefined. Job is executed in session
directory placed on frontend.}
\end{figure}
%
\begin{figure}[htbp]
\begin{centering}
\includegraphics{pic3}
\par\end{centering}

\caption{\label{fig:node scratch, not vis on front}RUNTIME\_LOCAL\_SCRATCH\_DIR
is set to value representing sratch directory on computing node, RUNTIME\_FRONTEND\_SEES\_NODE
undefined.}

\begin{lyxlist}{00.00.0000}
\item [{a)}] After job script starts all input files are moved to 'scratch
directory' on computing node.
\item [{b)}] Job runs in separate directory in 'scratch directory'. Only
files representing job's \textit{stdout} and \textit{stderr} are placed
in original 'session directory' and soft-linked in 'scratch'. After
execution all files from 'scratch' are moved back to original 'session
directory'.
\item [{c)}] All output files are in 'session directory' and are ready
to be uploaded/downloaded.
\end{lyxlist}

\end{figure}
%
\begin{figure}[htbp]
\begin{centering}
\includegraphics{pic4}
\par\end{centering}

\caption{\label{fig:node scratch, vis on front}Both RUNTIME\_LOCAL\_SCRATCH\_DIR
and RUNTIME\_FRONTEND\_SEES\_NODE are set to valuea representing sratch
directory on computing node and way to access that scratch from frontend
correspondingly.}

\begin{lyxlist}{00.00.0000}
\item [{a)}] After job script starts all input files are moved to 'scratch
directory' on computing node. Original 'session directory' is removed
and replaced with soft-link to copy of session directory in 'scratch'
as seen on frontend.
\item [{b)}] Job runs in separate directory in 'scratch directory'. All
files are also available on frontend through soft-link. After execution
soft-link is replaced with directory and all files from 'scratch'
are moved back to original 'session directory'.
\item [{c)}] All output files are in 'session directory' and are ready
to be uploaded/downloaded. 
\end{lyxlist}

\end{figure}

\subsection{Runtime environment}

The GM can run specially prepared \emph{BASH} scripts prior creation
of job's script, before and after executing job's main executable.
Those scripts are requested by user through \emph{runtimeenvironment}
attribute in RSL and are run with only argument set equal to '0',
'1' or '2' during creation of job's script, before execution of main
executable and after main executable finished accordingly. They all
are run through BASH's 'source' command, and hence can manipulate
with shell variables. With argument '0' scripts are run by the GM
on frontend. Some environment variables are defined in that case and
can be changed to influence job's execution later:

\begin{itemize}
\item joboption\_directory - session directory.
\item joboption\_args - command to be executed as specified in RSL.
\item joboption\_env\_\# - array of 'NAME=VALUE' environment variables (\textbf{not}
bash array).
\item joboption\_runtime\_\# - array of requested \emph{runtimeenvironment}
names (\textbf{not} bash array).
\item joboption\_num - \emph{runtimeenvironment} currently beeing processed
(number starting from 0).
\item joboption\_stdin - name of file to be attached to stdin handle.
\item joboption\_stdout - same for stdout.
\item joboption\_stderr - same for stderr.
\item joboption\_maxcputime - amout of CPU time requested (minutes).
\item joboption\_maxmemory - amout of memory requested (megabytes).
\item joboption\_count - number of processors requested.
\item joboption\_lrms - LRMS to be used to run job.
\item joboption\_queue - name of a queue of LRMS to put job into.
\item joboption\_nodeproperty\_\# - array of properties of computing nodes
(LRMS specific, \textbf{not} bash array).
\item joboption\_jobname - name of the job as given by user.
\item joboption\_rsl - whole RSL for very clever submission scripts.
\item joboption\_rsl\_\emph{name} - RSL attributes and values (like joboption\_rsl\_executable=''/bin/echo'')
\end{itemize}
For example \emph{joboption\_args} could be changed to wrap main executable.
Or \emph{joboption\_runtime} could be expanded if current one depends
on others.

With argument '1' scripts are run just before main executable is run.
They are executed on computing node. Such script can prepare environment
for some third-party software package. A current directory in that
case is one which would be used for execution of job. Variable HOME
also points to that directory.

With argument '2' scripts are executed after main executable finished.
Main purpose is to clean possible changes done by scripts run with
'1' (like removing temporary files). Execution of scripts at that
stage also happens on computing node and is not reliable. If the job
is killed by LRMS they most probably won't be executed.

\section{Installation\label{sec:installation}}

To install GM as part of ARC-enabled site please read {}``NorduGrid
ARC server installation instructions'' at \url{http://www.nordugrid.org/documents/ng-server-install.html}.

\subsection{Requirements}

The GM is mostly written using C++. It was tested and should compile
on recent enough \emph{Linux} systems using \emph{gcc} compiler and
\emph{GNU make} (gcc versions 2.95, 2.96, 3.2, 3.4 were tested). You
will also need \emph{Globus Toolkit$^{TM}$} of version higher than
2.2 \emph{}installed \url{http://www-unix.globus.org/toolkit/}.

\subsection{Setup of the Grid Manager}

For in-depth information about how to properly setup the GM and related
software please read {}``NorduGrid ARC server installation instructions''
at \url{http://www.nordugrid.org/documents/ng-server-install.html}.
Follow that manual to install GM, configure and run it. Additional
tips are described here.

The GM is designed to be able to run both as root and as ordinary
user. You can chose the name of the user by using corresponding command
in configuration file. It is better run GM as root if You want to
serve few users.

The GM writes debug information into a file /var/log/grid-manager.log
by default. . Also file /var/log/gm-jobs.log (default path in configuration
template, turned off by default) contains information about all started
and finished jobs, 2 lines per job (1 when job is started and 1 after
it finished).

\subsection{Setup of the GridFTP Server}

For in-depth information about how to properly setup the GM and related
software please read {}``NorduGrid ARC server installation instructions''
at \url{http://www.nordugrid.org/documents/ng-server-install.html}.
Follow that manual to install GM, configure and run it. Additional
tips are described here.

To make GFS to interoperate with other parts of the ARC only one \emph{jobplugin.so}
needs to be configured. It is advisable to use the template configuration
file. You can leave only part which configures \textit{jobplugin.so}
plugin.

\subsection{Usage}

Refer to the description of the \emph{User Interface} part \cite{ui}
and extensions to RSL \cite{xrsl} for using the GM.

\subsection{Unix accounts}

Bot GM and GFS are designed to be run by \emph{root} UNIX account
and serve multiple local UNIX and global Grid identities. Nevertheless
it is possible to use \emph{non-root} accounts to run those services.
Although this means some functionality loss described below.

There are no implication on running GFS with \emph{gaclplugin} or
\emph{fileplugin} under \emph{non-root} account as long as only Grid
identity of user is used and all served files and directories are
owned by server's account.

For combination of GM and GFS with \emph{jobplugin} both services
must be run either by same account or one of services must be run
under \emph{root} account. That is needed because services communicate
over local filesystem, hence must have \emph{full} access to same
set of files.

As long as GFS with \emph{jobplugin} is run under non-root account
there is no mapping from Grid identity to local UNIX account taking
place. All alowed Grid users are assigned server's account and are
then processes by GM using same account. Only way to overcome this
limitation is to run one GFS per local account with proper access
control configured.

Because GM has to represent user's local account while communication
with LRMS, it can serve only account it is run under (unless it is
run under \emph{root} account, of course). Like in case of GFS, multiple
instances of GM may be run, one per local account. That solution causes
another implications. The GM looses possibility to share cached files
among serviced users. It is not also possible to control load on a
frontend by limiting number of simultatenuosly running \emph{downloader}
and \emph{uploader} modules.

One has also take into account that private part of GSI infrastructure
(private key of a host at least) has to be duplicated for every account
used to run GFS.

\section*{Appendix A. Job control over jobplugin.so}

\subsection*{Virtual tree}

Under mount point of jobplugin gridftp client can see directories
representing job belonging to user, who started client. Directory
per job. Directory names are same as jobs' identifiers. Those directories
are directly connected to session directories of jobs and contain
same files and subdirectories. Except if jobs session directory is
moved to computing node. In that case directories only contain files
with redirected stdout and strderr as specified in xRSL.

If job's xRSL has \emph{gmlog} specified job's directory also contains
virtual subdirectory with same name, which contains files with information
about job as created by GM. The most important are 'errors' and 'status'.
'errors' contains stderr of separate modules run by GM in order to
process job (downloader, uploader, job's submission to LRMS). 'status'
contains one word representing state of job.

Also under mount point there is additional directory named \char`\"{}new\char`\"{}
used to submit new jobs. And another directory {}``info'' with subdirectories
named after job ids. Those subdirectories contain files with information
about job identical to those in subdirectory specified through \emph{gmlog}.

\subsection*{Submission}

Each xRSL put into directory \char`\"{}new\char`\"{} is accepted as
job's description. jobplugin parses it and client gets positive response
if there are no errors in request.

Job gets identifier and directory with corresponding name appears.
If job's description contains input files which should be delivered
from client's machine, client must upload them to that directory under
specified names.

Because each job gets identifier there should be a way for client
to obtain it. For that prior to providing xRSL client sends command
CWD to change current directory to \char`\"{}new\char`\"{}. In this
way job's identifier is reserved, new directory corresponding to that
identifier is created and client is redirected to it (as specified
in FTP protocol). Job's description put into \char`\"{}new\char`\"{}
will get reserved identifier.

\subsection*{Actions}

Various actions to affect processing of existing job are performed
by uploading xRSL files into directory {}``new''. Content of xRSL
may consist of only 2 parameters - action for \emph{action} to be
performed, and \emph{jobid} to identify job to be affected. Rest of
parameters are ignored.

Currently supported actions are:

\begin{lyxlist}{00.00.0000}
\item [{\emph{cancel}}] to cancel job
\item [{\emph{clean}}] to remove job from computing resource
\item [{\emph{renew}}] to renew credentials delegated to job
\item [{\emph{restart}}] to restart job after failure at some phases
\end{lyxlist}
It is also possible to perform some actions by using shortcut FTP
operations described below.

\subsubsection*{Cancel}

Job is canceled by performing DELE (delete file) command on directory
representing job. It can take some time (few minutes) before job is
actually canceled. Nevertheless client gets response immediately.

\subsubsection*{Clean}

Job's content is cleaned by performing RMD (remove directory) command
on directory representing job. If job is in \char`\"{}FINISHED\char`\"{}
state it will be cleaned immediately. Otherwise it will be cleaned
after it reaches state \char`\"{}FINISHED\char`\"{}.

\subsubsection*{Renew}

If client requests CWD to session directory credentials passed during
authentication are compared to current credentials of the job. If
validity time of the new credentials is longer job's credentials are
replaced with new.

\section*{Appendix B. Library \textit{libarcdata}}

\emph{libarcdata} is now part of \emph{libngui} library. It's functions
are declared in a header file \emph{arcdata.h}. They correspond to
ng{*} utilities meant for data handling - \emph{arcacl}, \emph{arccp},
\emph{arcls}, \emph{arcrm}, \emph{arctransfer}. It consists of following
functions:

\begin{lyxcode}
void~arcacl(const~std::string\&~file\_url,~const~std::string\&~command,~int~timeout~=~0);

void~arcregister~(const~std::string\&~source\_url,~const~std::string\&~destination\_url,~bool~secure~=~false,~bool~passive~=~true,~bool~force\_meta~=~false,~int~timeout~=~0);

void~arccp~(const~std::string\&~source\_url,~const~std::string\&~destination\_url,~bool~secure~=~false,~bool~passive~=~true,~bool~force\_meta~=~false,~int~recursion~=~0,~bool~verbose~=~false,~int~timeout~=~0);

void~arcls(const~std::string\&~dir\_url,~bool~show\_details~=~false,~bool~show\_urls~=~false,~int~recursion~=~0,~int~timeout~=~0);

void~arcrm(const~std::string\&~file\_url,~bool~errcont~=~false,~int~timeout~=~0);

void~arctransfer(const~std::string\&~destination,~std::list<std::string>\&~sources,~int~timeout~=~0);
\end{lyxcode}
Additionally this library contains C++ classes used by \emph{ng{*}}
data management utilities. Those are described in {}``ARC::DataMove
Reference Manual''.

\section*{Appendix C. Error messages of GM}

If job has not finished successfully the GM put one or more lines
into \textit{job.ID.failed}. Possible valuesinclude those generated
by the GM itself:

\begin{tabular}{|p{2.5in}|p{4in}|}
\hline 
\emph{Error string}&
\emph{Reason/description}\tabularnewline
\hline 
Internal error&
Error in internal algorithm\tabularnewline
\hline 
Internal error: can't read local file&
Error manipulating files in the control directory\tabularnewline
\hline 
Failed reading local job information&
-//-\tabularnewline
\hline 
Failed reading status of the job&
-//-\tabularnewline
\hline 
Failed writing job status&
-//-\tabularnewline
\hline 
Failed during processing failure&
-//-\tabularnewline
\hline 
Serious troubles (problems during processing problems)&
-//-\tabularnewline
\hline 
Failed initiating job submission to LRMS&
Could not run backend executable to pass job to LRMS\tabularnewline
\hline 
Job submission to LRMS failed&
Backend executable supposed to pass job to LRMs returned non-zero
exit code\tabularnewline
\hline 
Failed extracting LRMS ID due to some internal error&
Output of Backend executable supposed to contain local ID of passed
job could not be parsed\tabularnewline
\hline 
Failed in files upload (post-processing)&
Failed to upload some or all output files\tabularnewline
\hline 
Failed in files upload due to expired credentials - try to renew&
Failed to upload some or all output files most probably due to expired
credentials (proxy certificate)\tabularnewline
\hline 
Failed to run uploader (post-processing)&
Could not run \emph{uploader} executable\tabularnewline
\hline 
uploader failed (postprocessing)&
Generic error related to \emph{uploader} component\tabularnewline
\hline 
Failed in files download (pre-processing)&
Failed to upload some or all input files\tabularnewline
\hline 
Failed in files download due to expired credentials - try to renew&
Failed to download some or all input files most probably due to expired
credentials (proxy certificate)\tabularnewline
\hline 
Failed to run downloader (pre-processing)&
Could not run \emph{downloader} executable\tabularnewline
\hline
downloader failed (preprocessing)&
Generic error related to \emph{downloader} component\tabularnewline
\hline
User requested to cancel the job&
GM detected external request to cancel this job, most probably issued
by user\tabularnewline
\hline
Could not process RSL&
Job description could not be processed to syntax errors or missing
elements\tabularnewline
\hline
User requested dryrun. Job skiped.&
Job description contains request not to process this job\tabularnewline
\hline
LRMS error: (CODE) DESCRIPTION&
LRMS returned error. CODE is replaced with numeric code of LRMS, and
DESCRIPTION with textual description\tabularnewline
\hline
Plugin at state STATE failed: OUTPUT&
External plugin specified in GM's configuration returned non-zero
exit code. STATE is replcaced by name of state to which job was going
to be passed, OUTPUT by textual output generated by plugin.\tabularnewline
\hline
Failed running plugin at state STATE&
External plugin specified in GM's configuration could not be executed.\tabularnewline
\hline
\end{tabular}

\medskip{}
Provided by downloader component (URL is replcaced by source of input
file, FILE by name of file):

\begin{tabular}{|p{2.5in}|p{4in}|}
\hline 
\emph{Error string}&
\emph{Reason/description}\tabularnewline
\hline 
Internal error in downloader&
Generic error \tabularnewline
\hline 
Input file: URL - unknown error&
Generic error \tabularnewline
\hline 
Input file: URL - unexpected error&
Generic error \tabularnewline
\hline 
Input file: URL - bad source URL&
Source URL is either malformed or not supported\tabularnewline
\hline 
Input file: URL - bad destination URL&
Shouldn't happen\tabularnewline
\hline 
Input file: URL - failed to resolve source locations&
File either not registred or other problems related to Data Indexing
service.\tabularnewline
\hline 
Input file: URL - failed to resolve destination locations&
Shouldn't happen\tabularnewline
\hline 
Input file: URL - failed to register new destination file&
Shouldn't happen\tabularnewline
\hline 
Input file: URL - can't start reading from source&
Problems related to accessing instance of file at Data Storing service.\tabularnewline
\hline 
Input file: URL - can't read from source&
-//-\tabularnewline
\hline 
Input file: URL - can't start writing to destination&
Access problems in a session directory\tabularnewline
\hline 
Input file: URL - can't write to destination&
-//-\tabularnewline
\hline 
Input file: URL - data transfer was too slow&
Timeouted while trying to download file\tabularnewline
\hline 
Input file: URL - failed while closing connection to source&
Shouldn't happen\tabularnewline
\hline 
Input file: URL - failed while closing connection to destination&
Shouldn't happen\tabularnewline
\hline 
Input file: URL - failed to register new location&
Shouldn't happen\tabularnewline
\hline 
Input file: URL - can't use local cache&
Problems with GM cache \tabularnewline
\hline 
Input file: URL - system error&
Operating System returned error code where unexpected\tabularnewline
\hline 
Input file: URL - delegated credentials expired&
Access to source requires credententials and they are either outdated
or missing (not delegated).\tabularnewline
\hline 
User file: FILENAME - Bad information about file: checksum can't be
parsed.&
In job description there is a checksum provided for file uploadable
by user interface and this record can't be interpreted.\tabularnewline
\hline 
User file: FILENAME - Bad information about file: size can't be parsed.&
In job description there is a size provided for file uploadable by
user interface and this record can't be interpreted.\tabularnewline
\hline 
User file: FILENAME - Expected file. Directory found.&
Instead of file uploadable by user interface GM found directory with
same name in a session directory.\tabularnewline
\hline 
User file: FILENAME - Expected ordinary file. Special object found.&
Instead of file uploadable by user interface GM found special object
with same name in a session directory.\tabularnewline
\hline 
User file: FILENAME - Delivered file is bigger than specified.&
The size of file uploadable by user interface is bigger than specified
in job description.\tabularnewline
\hline
User file: FILENAME - Delivered file is unreadable.&
GM can't check user uploadable file due to some internal error. Most
probably due to improperly configured local permissions.\tabularnewline
\hline
User file: FILENAME - Could not read file to compute checksum.&
GM can't read user uploadable file due to some internal error. Most
probably due to improperly configured local permissions.\tabularnewline
\hline
User file: FILENAME - Timeout waiting&
GM waited for user uploadable file too long.\tabularnewline
\hline
\end{tabular}

\medskip{}
Provided by uploader component (URL is replcaced by destination of
output file) :

\begin{tabular}{|p{2.5in}|p{4in}|}
\hline 
\emph{Error string}&
\emph{Reason/description}\tabularnewline
\hline 
Internal error in uploader&
Generic error \tabularnewline
\hline 
Output file: URL - unknown error&
Generic error \tabularnewline
\hline 
Output file: URL - unexpected error&
Generic error \tabularnewline
\hline 
User requested to store output locally URL&
Destination is URL of type \emph{file}.\tabularnewline
\hline 
Output file: URL - bad source URL&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - bad destination URL&
Destination URL is either malformed or not supported\tabularnewline
\hline 
Output file: URL - failed to resolve source locations&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed to resolve destination locations&
Problems related to Data Indexing service.\tabularnewline
\hline 
Output file: URL - failed to register new destination file&
-//-\tabularnewline
\hline 
Output file: URL - can't start reading from source&
User request to store output file, but there is no such file or there
are problems accessing session directory\tabularnewline
\hline 
Output file: URL - can't start writing to destination&
Problems with Data Storing services\tabularnewline
\hline 
Output file: URL - can't read from source&
Problems accessing session directory\tabularnewline
\hline 
Output file: URL - can't write to destination&
Problems with Data Storing services\tabularnewline
\hline 
Output file: URL - data transfer was too slow&
Timeout during transfer\tabularnewline
\hline 
Output file: URL - failed while closing connection to source&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed while closing connection to destination&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - failed to register new location&
Problems related to Data Indexing service.\tabularnewline
\hline 
Output file: URL - can't use local cache&
Shouldn't happen\tabularnewline
\hline 
Output file: URL - system error&
Operating System returned error code where unexpected\tabularnewline
\hline 
Output file: URL - delegated credentials expired&
Access to destination requires credententials and they are either
outdated or missing (not delegated).\tabularnewline
\hline 
&
\tabularnewline
\hline
&
\tabularnewline
\hline
\end{tabular}

\medskip{}
Coming from LRMS (PBS) backend:

\begin{tabular}{|p{2.5in}|p{4in}|}
\hline 
\emph{Error string}&
\emph{Reason/description}\tabularnewline
\hline 
Submission: Configuration error.&
\tabularnewline
\hline 
Submission: System error.&
\tabularnewline
\hline 
Submission: Job description error.&
\tabularnewline
\hline 
Submission: Local submission client behaved unexpectedly.&
\tabularnewline
\hline 
Submission: Local submission client failed.&
\tabularnewline
\hline
\end{tabular}

\begin{lyxcode}
~
\end{lyxcode}

\section*{Appendix D. A-Rex WSDL}

\begin{lyxcode}
<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>

<wsdl:definitions~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}

~xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}

~xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}

~xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}

~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}

~xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}

~xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}

~xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}

~xmlns:bes-factory=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}

~xmlns:bes-mgmt=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}

~xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}

~xmlns:wsrf-rpw=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}

~xmlns:a-rex=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>

~~<wsdl:import~namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}~location=\char`\"{}./bes-factory.wsdl\char`\"{}/>

~~<wsdl:import~namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}~location=\char`\"{}./bes-management.wsdl\char`\"{}/>

~~<wsdl:import~namespase=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}~location=\char`\"{}../schemas/delegation.wsdl\char`\"{}/>

~~<wsdl:import~namespase=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}~location=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2.wsdl\char`\"{}/>

~~<wsdl:types>

~~~~<xsd:schema~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>

~~~~~~<xsd:import~namespace=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}~schemaLocation=\char`\"{}./ws-addr.xsd\char`\"{}/>

~~~~~~<xsd:simpleType~name=\char`\"{}ActivitySubStateType\char`\"{}>

~~~~~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Accepting\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Accepted\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Preparing\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Prepared\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Submiting\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Executing\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Killing\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Executed\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Finishing\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Finished\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Failed\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Deleted\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Pending\char`\"{}/>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Held\char`\"{}/>

~~~~~~~~</xsd:restriction>

~~~~~~</xsd:simpleType>

~~~~~~<xsd:element~name=\char`\"{}State\char`\"{}~type=\char`\"{}a-rex:ActivitySubStateType\char`\"{}/>

~~~~~~<xsd:complexType~name=\char`\"{}ResourceInformationDocumentType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~~<xsd:element~name=\char`\"{}BESFactory\char`\"{}~type=\char`\"{}bes-factory:FactoryResourceAttributesDocumentType\char`\"{}/>

~~~~~~~~~~<xsd:complexType~name=?Glue2Resource?~minOccurs='0'>

~~~~~~~~~~~~<xsd:sequence>

~~~~~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>

~~~~~~~~~~~~</xsd:sequence>

~~~~~~~~~~</xsd:complexType>

~~~~~~~~~~<xsd:complexType~name=\char`\"{}Activities\char`\"{}~minOccurs='0'>

~~~~~~~~~~~~<xsd:sequence>

~~~~~~~~~~~~~~<xsd:complexType~name=\char`\"{}Activity\char`\"{}~minOccurs='0'~maxOccurs='unbounded'>

~~~~~~~~~~~~~~~~<xsd:sequence>

~~~~~~~~~~~~~~~~~~<xsd:element~name=\char`\"{}ActivityIdentifier\char`\"{}~type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>

~~~~~~~~~~~~~~~~~~<xsd:element~ref=?bes-factory:ActivityDocument?~minOccurs='0'/>

~~~~~~~~~~~~~~~~~~<xsd:complexType~name=?Glue2Job?~minOccurs='0'>

~~~~~~~~~~~~~~~~~~~~<xsd:sequence>

~~~~~~~~~~~~~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>

~~~~~~~~~~~~~~~~~~~~</xsd:sequence>

~~~~~~~~~~~~~~~~~~</xsd:complexType>

~~~~~~~~~~~~~~~~</xsd:sequence>

~~~~~~~~~~~~~~</xsd:complexType>

~~~~~~~~~~~~</xsd:sequence>

~~~~~~~~~~</xsd:complexType>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:complexType~name=\char`\"{}ChangeActivityStatusRequestType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}ActivityIdentifier\char`\"{}~type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>

~~~~~~~~~~<xsd:element~name=\char`\"{}OldStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~~~~~~~~<xsd:element~name=\char`\"{}NewStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}ChangeActivityStatus\char`\"{}~type=\char`\"{}a-rex:ChangeActivityStatusRequestType\char`\"{}/>

~~~~~~<xsd:complexType~name=\char`\"{}ChangeActivityStatusResponseType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}NewStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}~type=\char`\"{}a-rex:ChangeActivityStatusResponseType\char`\"{}/>

~~~~</xsd:schema>

~~</wsdl:types>

~~<wsdl:message~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}~element=\char`\"{}a-rex:ChangeActivityStatus\char`\"{}/>

~~</wsdl:message>

~~<wsdl:message~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}~element=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>

~~</wsdl:message>

~~<wsdl:portType~name=\char`\"{}a-rex\char`\"{}>

~~~~<wsdl:operation~name=\char`\"{}ChangeActivityStatus\char`\"{}>

~~~~~~<wsdl:documentation>

~~~~~~~~This~operation~allows~any~simple~status~change~request

~~~~~~~~which~involves~no~additional~parameters.~It~should~be

~~~~~~~~used~to~modify~job/activity~execution~flow:

~~~~~~~~~~-~To~put~job~on~hold

~~~~~~~~~~-~To~rerun~job~in~case~of~failure

~~~~~~~~~~-~To~cancel~job~(same~as~TerminateActivity~of~BESFActory)

~~~~~~~~~~-~To~remove/release~job~-~as~long~as~non-existence~is~a~state

~~~~~~~~~~-~Any~other~status~change~no~supported~by~BES

~~~~~~</wsdl:documentation>

~~~~~~<wsdl:input~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}

~~~~~~~~message=\char`\"{}a-rex:ChangeActivityStatusRequest\char`\"{}/>

~~~~~~<wsdl:output~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}

~~~~~~~~message=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}

~~~~~~~~message=\char`\"{}bes-factory:NotAuthorizedFault\char`\"{}

~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}

~~~~~~~~message=\char`\"{}bes-factory:InvalidActivityIdentifierFault\char`\"{}

~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}

~~~~~~~~~message=\char`\"{}bes-factory:CantApplyOperationToCurrentStateFault\char`\"{}

~~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}

~~~~~~~~~message=\char`\"{}bes-factory:OperationWillBeAppliedEventuallyFault\char`\"{}

~~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>

~~~~</wsdl:operation>

~~</wsdl:portType>

~~<wsdl:binding~name=\char`\"{}a-rex\char`\"{}~type=\char`\"{}a-rex:a-rex\char`\"{}>

~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>

~~~~<wsdl:operation~name=\char`\"{}ChangeActivityStatus\char`\"{}>

~~~~~~<soap:operation~soapAction=\char`\"{}ChangeActivityStatus\char`\"{}/>

~~~~~~<wsdl:input~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>

~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:input>

~~~~~~<wsdl:output~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>

~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:output>

~~~~</wsdl:operation>

~~</wsdl:binding>

~~<wsdl:binding~name=\char`\"{}GetResourcePropertyDocument\char`\"{}~type=\char`\"{}wsrf-rpw:GetResourcePropertyDocument\char`\"{}>

~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>

~~~~<wsdl:operation~name=\char`\"{}GetResourcePropertyDocument\char`\"{}>

~~~~~~<soap:operation~soapAction=\char`\"{}GetResourcePropertyDocument\char`\"{}/>

~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentRequest\char`\"{}>

~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:input>

~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentResponse\char`\"{}>

~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:output>

~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~</wsdl:operation>

~~</wsdl:binding>

~~<wsdl:binding~name=\char`\"{}GetResourceProperty\char`\"{}~type=\char`\"{}wsrf-rpw:GetResourceProperty\char`\"{}>

~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>

~~~~<wsdl:operation~name=\char`\"{}GetResourceProperty\char`\"{}>

~~~~~~<soap:operation~soapAction=\char`\"{}GetResourceProperty\char`\"{}/>

~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:GetResourcePropertyRequest\char`\"{}>

~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:input>

~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:GetResourcePropertyResponse\char`\"{}>

~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:output>

~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~~~<wsdl:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~</wsdl:operation>

~~</wsdl:binding>

~~<wsdl:binding~name=\char`\"{}QueryResourceProperties\char`\"{}~type=\char`\"{}wsrf:QueryResourceProperties\char`\"{}>

~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>

~~~~<wsdl:operation~name=\char`\"{}QueryResourceProperties\char`\"{}>

~~~~~~<soap:operation~soapAction=\char`\"{}QueryResourceProperties\char`\"{}/>

~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:QueryResourcePropertiesRequest\char`\"{}>

~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:input>

~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:QueryResourcePropertiesResponse\char`\"{}>

~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:output>

~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~~~<wsdl:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~~~<wsdl:fault~name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~~~<wsdl:fault~name=\char`\"{}InvalidQueryExpressionFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}InvalidQueryExpressionFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~~~<wsdl:fault~name=\char`\"{}QueryEvaluationErrorFault\char`\"{}>

~~~~~~~~<soap:fault~name=\char`\"{}QueryEvaluationErrorFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>

~~~~~~</wsdl:fault>

~~~~</wsdl:operation>

~~</wsdl:binding>

~~<wsdl:service~name=\char`\"{}a-rex\char`\"{}>

~~~~<wsdl:port~name=\char`\"{}delegation\char`\"{}~binding=\char`\"{}deleg:DelegationBinding\char`\"{}>

~~~~</wsdl:port>

~~~~<wsdl:port~name=\char`\"{}bes-factory\char`\"{}~binding=\char`\"{}bes-factory:BESFactoryBinding\char`\"{}>

~~~~</wsdl:port>

~~~~<wsdl:port~name=\char`\"{}bes-mgmt\char`\"{}~binding=\char`\"{}bes-mgmt:BESManagementBinding\char`\"{}>

~~~~</wsdl:port>

~~~~<wsdl:port~name=\char`\"{}GetResourcePropertyDocument\char`\"{}~binding=\char`\"{}a-rex:GetResourcePropertyDocument\char`\"{}>

~~~~</wsdl:port>

~~~~<wsdl:port~name=\char`\"{}GetResourceProperty\char`\"{}~binding=\char`\"{}a-rex:GetResourceProperty\char`\"{}>

~~~~</wsdl:port>

~~~~<wsdl:port~name=\char`\"{}QueryResourceProperties\char`\"{}~binding=\char`\"{}a-rex:QueryResourceProperties\char`\"{}>

~~~~</wsdl:port>

~~~~<wsdl:port~name=\char`\"{}a-rex\char`\"{}~binding=\char`\"{}a-rex:a-rex\char`\"{}>

~~~~</wsdl:port>

~~</wsdl:service>

</wsdl:definitions>
\end{lyxcode}

\section*{Appendix E. Delegation WSDL}

\begin{lyxcode}
<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>

<wsdl:definitions~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}

~xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}

~xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}

~xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}

~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}

~xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}

~xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}

~xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}

~xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>

~~<wsdl:types>

~~~~<xsd:schema~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>

~~~~~~<!-{}-~Common~types~-{}->

~~~~~~<xsd:simpleType~name=\char`\"{}TokenFormatType\char`\"{}>

~~~~~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>

~~~~~~~~~~<xsd:enumeration~value=\char`\"{}x509\char`\"{}/>

~~~~~~~~</xsd:restriction>

~~~~~~</xsd:simpleType>

~~~~~~<xsd:complexType~name=\char`\"{}ReferenceType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:complexType~name=\char`\"{}DelegatedTokenType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}Id\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~~~~~~~~~<xsd:element~name=\char`\"{}Value\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~~~~~~~~~<xsd:element~name=\char`\"{}Reference\char`\"{}~type=\char`\"{}deleg:ReferenceType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~~~<xsd:attribute~name=\char`\"{}Format\char`\"{}~type=\char`\"{}deleg:TokenFormatType\char`\"{}~use=\char`\"{}required\char`\"{}/>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}DelegatedToken\char`\"{}~type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>

~~~~~~<xsd:complexType~name=\char`\"{}TokenRequestType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}Id\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~~~~~~~~~<xsd:element~name=\char`\"{}Value\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~~~<xsd:attribute~name=\char`\"{}Format\char`\"{}~type=\char`\"{}deleg:TokenFormatType\char`\"{}~use=\char`\"{}required\char`\"{}/>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}TokenRequest\char`\"{}~type=\char`\"{}deleg:TokenRequestType\char`\"{}/>

~~~~~~<!-{}-~Types~for~messages~-{}->

~~~~~~<xsd:complexType~name=\char`\"{}DelegateCredentialsInitRequestType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}DelegateCredentialsInit\char`\"{}~type=\char`\"{}deleg:DelegateCredentialsInitRequestType\char`\"{}/>

~~~~~~<xsd:complexType~name=\char`\"{}DelegateCredentialsInitResponseType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}TokenRequest\char`\"{}~type=\char`\"{}deleg:TokenRequestType\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}~type=\char`\"{}deleg:DelegateCredentialsInitResponseType\char`\"{}/>

~~~~~~<xsd:complexType~name=\char`\"{}UpdateCredentialsRequestType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}DelegatedToken\char`\"{}~type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}UpdateCredentials\char`\"{}~type=\char`\"{}deleg:UpdateCredentialsRequestType\char`\"{}/>

~~~~~~<xsd:complexType~name=\char`\"{}UpdateCredentialsResponseType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}UpdateCredentialsResponse\char`\"{}~type=\char`\"{}deleg:UpdateCredentialsResponseType\char`\"{}/>

~~~~~~<!-{}-~Faults~-{}->

~~~~~~<xsd:complexType~name=\char`\"{}UnsupportedFaultType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}UnsupportedFault\char`\"{}~type=\char`\"{}deleg:UnsupportedFaultType\char`\"{}/>

~~~~~~<xsd:complexType~name=\char`\"{}ProcessingFaultType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}ProcessingFault\char`\"{}~type=\char`\"{}deleg:ProcessingFaultType\char`\"{}/>

~~~~~~<xsd:complexType~name=\char`\"{}WrongReferenceFaultType\char`\"{}>

~~~~~~~~<xsd:sequence>

~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~~~~~~</xsd:sequence>

~~~~~~</xsd:complexType>

~~~~~~<xsd:element~name=\char`\"{}WrongReferenceFault\char`\"{}~type=\char`\"{}deleg:WrongReferenceFaultType\char`\"{}/>

~~~~</xsd:schema>

~~</wsdl:types>

~~<wsdl:message~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}~element=\char`\"{}deleg:DelegateCredentialsInit\char`\"{}/>

~~</wsdl:message>

~~<wsdl:message~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}~element=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>

~~</wsdl:message>

~~<wsdl:message~name=\char`\"{}UpdateCredentialsRequest\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}UpdateCredentialsRequest\char`\"{}~element=\char`\"{}deleg:UpdateCredentials\char`\"{}/>

~~</wsdl:message>

~~<wsdl:message~name=\char`\"{}UpdateCredentialsResponse\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}UpdateCredentialsResponse\char`\"{}~element=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>

~~</wsdl:message>

~~<wsdl:message~name=\char`\"{}UnsupportedFault\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:UnsupportedFault\char`\"{}/>

~~</wsdl:message>

~~<wsdl:message~name=\char`\"{}ProcessingFault\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:ProcessingFault\char`\"{}/>

~~</wsdl:message>

~~<wsdl:message~name=\char`\"{}WrongReferenceFault\char`\"{}>

~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>

~~</wsdl:message>

~~<wsdl:portType~name=\char`\"{}DelegationPortType\char`\"{}>

~~~<wsdl:operation~name=\char`\"{}DelegateCredentialsInit\char`\"{}>

~~~~~~<wsdl:documentation>

~~~~~~</wsdl:documentation>

~~~~~~<wsdl:input~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}

~~~~~~~~message=\char`\"{}deleg:DelegateCredentialsInitRequest\char`\"{}/>

~~~~~~<wsdl:output~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}

~~~~~~~~message=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}UnsupportedFault\char`\"{}

~~~~~~~~message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}ProcessingFault\char`\"{}

~~~~~~~~message=\char`\"{}deleg:ProcessingFault\char`\"{}/>

~~~~</wsdl:operation>

~~~<wsdl:operation~name=\char`\"{}UpdateCredentials\char`\"{}>

~~~~~~<wsdl:documentation>

~~~~~~</wsdl:documentation>

~~~~~~<wsdl:input~name=\char`\"{}UpdateCredentialsRequest\char`\"{}

~~~~~~~~message=\char`\"{}deleg:UpdateCredentialsRequest\char`\"{}/>

~~~~~~<wsdl:output~name=\char`\"{}UpdateCredentialsResponse\char`\"{}

~~~~~~~~message=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}UnsupportedFault\char`\"{}

~~~~~~~~message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}ProcessingFault\char`\"{}

~~~~~~~~message=\char`\"{}deleg:ProcessingFault\char`\"{}/>

~~~~~~<wsdl:fault~name=\char`\"{}WrongReferenceFault\char`\"{}

~~~~~~~~message=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>

~~~~</wsdl:operation>

~~</wsdl:portType>

~~<wsdl:binding~name=\char`\"{}DelegationBinding\char`\"{}~type=\char`\"{}deleg:DelegationPortType\char`\"{}>

~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>

~~~~<wsdl:operation~name=\char`\"{}DelegateCredentialsInit\char`\"{}>

~~~~~~<soap:operation~soapAction=\char`\"{}DelegateCredentialsInit\char`\"{}/>

~~~~~~<wsdl:input~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>

~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:input>

~~~~~~<wsdl:output~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>

~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:output>

~~~~</wsdl:operation>

~~~~<wsdl:operation~name=\char`\"{}UpdateCredentials\char`\"{}>

~~~~~~<soap:operation~soapAction=\char`\"{}UpdateCredentials\char`\"{}/>

~~~~~~<wsdl:input~name=\char`\"{}UpdateCredentialsRequest\char`\"{}>

~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:input>

~~~~~~<wsdl:output~name=\char`\"{}UpdateCredentialsResponse\char`\"{}>

~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>

~~~~~~</wsdl:output>

~~~~</wsdl:operation>

~~</wsdl:binding>

</wsdl:definitions>
\end{lyxcode}

\section*{Appendix F. ARC extensions for JSDL schema}

\begin{lyxcode}
<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>

<xsd:schema~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}

~~~~~~~~~~~~xmlns=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}

~~~~~~~~~~~~xmlns:jsdl-arc=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}

~~~~~~~~~~~~targetNamespace=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}>

~<xsd:simpleType~name=\char`\"{}GMState\_Type\char`\"{}>

~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>

~~~<xsd:enumeration~value=\char`\"{}ACCEPTED\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}PREPARING\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}SUBMIT\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}INLRMS\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}FINISHING\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}FINISHED\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}DELETED\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}CANCELING\char`\"{}/>

~~</xsd:restriction>

~</xsd:simpleType>

~<xsd:complexType~name=\char`\"{}Version\_Type\char`\"{}>

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}UpperExclusive\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~<xsd:element~name=\char`\"{}LowerExclusive\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~<xsd:element~name=\char`\"{}Exact\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>

~~~<xsd:element~name=\char`\"{}Exclusive\char`\"{}~type=\char`\"{}xsd:boolean\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:simpleType~name=\char`\"{}SessionType\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}SessionType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->

~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>

~~~<xsd:enumeration~value=\char`\"{}INTERNAL\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}LIMITED\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}READONLY\char`\"{}/>

~~~<xsd:enumeration~value=\char`\"{}FULL\char`\"{}/>

~~</xsd:restriction>

~</xsd:simpleType>

~<xsd:simpleType~name=\char`\"{}IsExecutable\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:DataStaging\_Type~(default:~false)~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}IsExecutable\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->

~~<xsd:restriction~base=\char`\"{}xsd:boolean\char`\"{}/>

~</xsd:simpleType>

~<xsd:simpleType~name=\char`\"{}FileParameters\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:DataStaging\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}IsExecutable\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->

~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}/>

~</xsd:simpleType>

~<xsd:simpleType~name=\char`\"{}JoinOutputs\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:JobDescription\_Type~(default:~false)~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}JoinOutputs\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->

~~<xsd:restriction~base=\char`\"{}xsd:boolean\char`\"{}/>

~</xsd:simpleType>

~<xsd:simpleType~name=\char`\"{}Reruns\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:JobDescription\_Type~(default:~false)~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}Reruns\char`\"{}~minOccurs=\char`\"{}0\char`\"{}//~-{}->

~~<xsd:restriction~base=\char`\"{}xsd:integer\char`\"{}/>

~</xsd:simpleType>

~<xsd:complexType~name=\char`\"{}RunTimeEnvironment\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}RunTimeEnvironment\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/~-{}->

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}Name\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~~<xsd:element~name=\char`\"{}Version\char`\"{}~type=\char`\"{}Version\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:complexType~name=\char`\"{}Middleware\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}Middleware\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/~-{}->

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}Name\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~~<xsd:element~name=\char`\"{}Version\char`\"{}~type=\char`\"{}Version\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:complexType~name=\char`\"{}RemoteLogging\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}RemoteLogging\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}3\char`\"{}/~-{}->

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}URL\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:anyURI\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:complexType~name=\char`\"{}CredentialServer\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}CredentialServer\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}URL\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:anyURI\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:complexType~name=\char`\"{}LocalLogging\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}LocalLogging\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}Directory\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:simpleType~name=\char`\"{}AccessControlType\_Type\char`\"{}>

~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>

~~~<xsd:enumeration~value=\char`\"{}GACL\char`\"{}/>

~~</xsd:restriction>

~</xsd:simpleType>

~<xsd:complexType~name=\char`\"{}AccessControl\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}AccessControl\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}OwnerAlwaysAllowed\char`\"{}~type=\char`\"{}xsd:boolean\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~<xsd:element~name=\char`\"{}Type\char`\"{}~type=\char`\"{}AccessControlType\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~<xsd:element~name=\char`\"{}Content\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:simpleType~name=\char`\"{}NotificationType\_Type\char`\"{}>

~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>

~~~<xsd:enumeration~value=\char`\"{}Email\char`\"{}/>

~~</xsd:restriction>

~</xsd:simpleType>

~<xsd:complexType~name=\char`\"{}Notify\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}Notify\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}3\char`\"{}/~-{}->

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}Type\char`\"{}~type=\char`\"{}NotificationType\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>

~~~<xsd:element~name=\char`\"{}Endpoint\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~~<xsd:element~name=\char`\"{}State\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}~type=\char`\"{}GMState\_Type\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:simpleType~name=\char`\"{}SessionLifeTime\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}SessionLifeTime\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->

~~<xsd:restriction~base=\char`\"{}xsd:long\char`\"{}/>

~</xsd:simpleType>

~<xsd:simpleType~name=\char`\"{}GridTimeLimit\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}GridTimeLimit\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->

~~<xsd:restriction~base=\char`\"{}xsd:positiveInteger\char`\"{}/>

~</xsd:simpleType>

~<xsd:complexType~name=\char`\"{}CandidateTarget\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}jsdl-arc:CandidateTarget\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->

~~<xsd:sequence>

~~~<xsd:element~name=\char`\"{}HostName\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~~<xsd:element~name=\char`\"{}QueueName\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>

~~</xsd:sequence>

~</xsd:complexType>

~<xsd:simpleType~name=\char`\"{}Time\_Type\char`\"{}>

~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>

~~<!-{}-~xsd:element~ref=\char`\"{}ProcessingStartTime\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->

~~<xsd:restriction~base=\char`\"{}xsd:dateTime\char`\"{}/>

~</xsd:simpleType>

~<!-{}-=========================================================-{}->

~<xsd:element~name=\char`\"{}IsExecutable\char`\"{}~type=\char`\"{}IsExecutable\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}FileParameters\char`\"{}~type=\char`\"{}FileParameters\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}RunTimeEnvironment\char`\"{}~type=\char`\"{}RunTimeEnvironment\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}Middleware\char`\"{}~type=\char`\"{}Middleware\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}RemoteLogging\char`\"{}~type=\char`\"{}RemoteLogging\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}LocalLogging\char`\"{}~type=\char`\"{}LocalLogging\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}AccessControl\char`\"{}~type=\char`\"{}AccessControl\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}Notify\char`\"{}~type=\char`\"{}Notify\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}SessionLifeTime\char`\"{}~type=\char`\"{}SessionLifeTime\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}SessionType\char`\"{}~type=\char`\"{}SessionType\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}JoinOutputs\char`\"{}~type=\char`\"{}JoinOutputs\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}Reruns\char`\"{}~type=\char`\"{}Reruns\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}CredentialServer\char`\"{}~type=\char`\"{}CredentialServer\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}GridTimeLimit\char`\"{}~type=\char`\"{}GridTimeLimit\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}CandidateTarget\char`\"{}~type=\char`\"{}CandidateTarget\_Type\char`\"{}/>

~<xsd:element~name=\char`\"{}ProcessingStartTime\char`\"{}~type=\char`\"{}Time\_Type\char`\"{}/>

</xsd:schema>

\end{lyxcode}
\bibliography{grid}
\end{document}

### 2007-09-06 doc/a-rex/arex_tech_doc.tex
Updating a-rex document according to some of comments from Adrian.

@@
 
-{\raggedleft NORDUGRID-TECH-2\par}
+{\raggedleft NORDUGRID-TECH-14\par}
 
@@
 
-Each job is initiated and controlled through GFS or WS. All job parameters
-(not data) are passed to the GM through GFS in RSL \cite{rsl} or
-through WS in JSDL-coded \cite{jsdl} description (job description
+Each job is initiated and controlled through GFS or WS. Complete description
+of job (not data) is passed to the GM through GFS in RSL \cite{rsl}
+or through WS in JSDL-coded \cite{jsdl} description (job description
 - JD).
@@
 
-A-REX Web Service Specific Interface provides means to submit a description
+A-REX Web Service Interface provides means to submit a description
 of a computational job to a computing resource, to stage-in additional
 data, to monitor and control processing of jobs, and obtain data produced
-during the execution of a job.
+during the execution of a job. The WS Interface is built and deployed
+inside Hosting Envirnoment Daemon {[}HED] infrastructure\cite{hed}.
 
@@
 \begin{itemize}
-\item \emph{ActivityStatusType OldStatus}: Description of the state which
-the job is supposed to be in during execution of this request. If
-the current state of the job is different from the one having been
-given, the operation is aborted and a fault is returned. This parameter
-is optional.
-\item \emph{ActivityStatusType NewStatus}: Description of the state that
-the job is to be put into.
+\item \emph{ActivityStatusType OldStatus}: Description of the state the
+job is supposed to be in during execution of this request. If the
+current state of the job is different from the one having been given,
+the operation is aborted and a fault is returned. This parameter is
+optional.
+\item \emph{ActivityStatusType NewStatus}: Description of the state the
+job is to be put into.
 \end{itemize}
@@
 \item \emph{DelegateCredentialsInit} operation - this operation performs
-the first half of credentials delegation sequence.
+the first half of the credentials delegation sequence.
 
@@
 based on the public key from the DelegateCredentialsInit signed by
-the user?s proxy certificate. Also includes the Id element which identifies
+the user's proxy certificate. Also includes the Id element which identifies
 the private key stored at the service side associated with these credentials.
-Reference element refers to the object which these credentials should
-be applied to in a way specific to the service. The same element must
-also be used for delegating credentials as part of other operations
+The reference element refers to the object which these credentials
+should be applied to in a way specific to the service. The same element
+must also be used for delegating credentials as part of other operations
 on service.
@@
 \begin{lyxlist}{00.00.0000}
-\item [{\emph{BESFactory}}] collection of BES Factory attributes as defined
+\item [{\emph{BESFactory}}] - collection of BES Factory attributes as defined
 in the BES specifications.
-\item [{\emph{Glue2Resource}}] description of a computation resource that
-uses Glue2 schema. This one is going to be specified after the Glue2
-XML bindings are be available.
-\item [{\emph{Activities}}] list of maintained activities. Each activity
-contains an identifier (\emph{ActivityIdentifier}), BES (\emph{ActivityDocument})
+\item [{\emph{Glue2Resource}}] \emph{-} description of a computation resource
+that uses Glue2 schema. This one is going to be specified after the
+Glue2 XML bindings will be available.
+\item [{\emph{Activities}}] \emph{-} list of maintained activities. Each
+activity contains an identifier (\emph{ActivityIdentifier}), BES (\emph{ActivityDocument})
 and Glue2 (\emph{Glue2Job}) descriptions of the activity.
 \end{lyxlist}
-All information can be accessed either through requests on a particular
+All information can be accessed either through requests on particular
 resources or through XPath queries using WS-Resource Properties operations.
@@
 \begin{lyxlist}{00.00.0000}
-\item [{\emph{JobName}}] name of the job as assigned by the user.
-\item [{\emph{Executable}\ (POSIX)}] name of the executable file.
-\item [{\emph{Argument}\ (POSIX)}] arguments the executable will be launched
-with.
+\item [{\emph{JobName}}] \emph{-} name of the job as assigned by the user.
+\item [{\emph{Executable}\ (POSIX)}] - name of the executable file.
+\item [{\emph{Argument}\ (POSIX)}] - arguments the executable will be
+launched with.
 \item [{\emph{DataStaging}}]~
@@
 \begin{lyxlist}{00.00.0000}
-\item [{Filename}] name of the data file on the executing node.
-\item [{\emph{Source}}] source where the file will be taken from before
+\item [{Filename}] - name of the data file on the executing node.
+\item [{\emph{Source}}] - source where the file will be taken from before
 execution.
-\item [{\emph{Target}}] destination the file will be delivered to after
-execution.
+\item [{\emph{Target}}] \emph{-} destination the file will be delivered
+to after execution.
 \end{lyxlist}
-\item [{\emph{Input}\,\ (POSIX)}] file to be used as standard input for
+\item [{\emph{Input}\,\ (POSIX)}] - file to be used as standard input
+for the executable.
+\item [{\emph{Output}\ (POSIX)}] - file to be used as standard output
+for the executable.
+\item [{\emph{Error}\ (POSIX)}] - file to be used as standard error for
 the executable.
-\item [{\emph{Output}\ (POSIX)}] file to be used as standard output for
-the executable.
-\item [{\emph{Error}\ (POSIX)}] file to be used as standard error for
-the executable.
-\item [{\emph{MemoryLimit}\ (POSIX)}] amount of physical memory needed
+\item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
 for execution.
-\item [{\emph{TotalPhysicalMemory}}] same as MemoryLimit.
-\item [{\emph{IndividualPhysicalMemory}}] same as MemoryLimit.
-\item [{\emph{CPUTimeLimit}\ (POSIX)}] maximal amount of CPU time needed
+\item [{\emph{TotalPhysicalMemory}}] - same as MemoryLimit.
+\item [{\emph{IndividualPhysicalMemory}}] - same as MemoryLimit.
+\item [{\emph{CPUTimeLimit}\ (POSIX)}] - maximal amount of CPU time needed
 for execution.
-\item [{\emph{TotalCPUTime}}] same as CPUTimeLimit.
-\item [{\emph{IndividualCPUTime}}] same as CPUTimeLimit.
-\item [{\emph{WallTimeLimit}\ (POSIX)}] amount of clock time needed for
-execution.
-\item [{\emph{TotalCPUCount}}] number of CPUs needed for execution.
-\item [{\emph{IndividualCPUCount}}] same as TotalCPUCount.
+\item [{\emph{TotalCPUTime}}] - same as CPUTimeLimit.
+\item [{\emph{IndividualCPUTime}}] - same as CPUTimeLimit.
+\item [{\emph{WallTimeLimit}\ (POSIX)}] - amount of clock time needed
+for execution.
+\item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
+\item [{\emph{IndividualCPUCount}}] - same as TotalCPUCount.
 \end{lyxlist}
@@
 \begin{lyxlist}{00.00.0000}
-\item [{\emph{IsExecutable}}] marks file to become executable after being
-delivered to the computing resource.
-\item [{\emph{RunTimeEnvironment}}] specifies the name of the Runtime Environment
-needed for job execution.
-\item [{\emph{Middleware}}] request for specific middleware on the computing
+\item [{\emph{IsExecutable}}] \emph{-} marks file to become executable
+after being delivered to the computing resource.
+\item [{\emph{RunTimeEnvironment}}] - specifies the name of the Runtime
+Environment needed for job execution.
+\item [{\emph{Middleware}}] - request for specific middleware on the computing
 resource frontend.
-\item [{\emph{RemoteLogging}}] destination for the usage record report
-of the executed job.
-\item [{\emph{LocalLogging}}] name for the virtual directory available
+\item [{\emph{RemoteLogging}}] \emph{-} destination for the usage record
+report of the executed job.
+\item [{\emph{LocalLogging}}] - name for the virtual directory available
 through job interface and containing various debug information about
 job execution.
-\item [{\emph{AccessControl}}] ACL expression which describes the identities
+\item [{\emph{AccessControl}}] - ACL expression which describes the identities
 of those clients who are allowed to perform operations on this job.
-\item [{\emph{Notify}}] Email destination for notification of job state
+\item [{\emph{Notify}}] - Email destination for notification of job state
 changes.
-\item [{\emph{SessionLifeTime}}] duration for the directory containing
+\item [{\emph{SessionLifeTime}}] - duration for the directory containing
 job-related files to exist after the job finished executing.
-\item [{\emph{JoinOutputs}}] specifies if standard output and standard
+\item [{\emph{JoinOutputs}}] - specifies if standard output and standard
 error channels must be merged.
-\item [{\emph{Reruns}}] defines how many times a job is allowed to rerun
-in case of failure.
-\item [{\emph{CredentialServer}}] URL of MyProxy service which may be used
-for renewing the expired delegated job credentials.
-\item [{\emph{CandidateTarget}}] specifies host name and queue of a computing
-resource.
+\item [{\emph{Reruns}}] \emph{-} defines how many times a job is allowed
+to rerun in case of failure.
+\item [{\emph{CredentialServer}}] \emph{-} URL of MyProxy service which
+may be used for renewing the expired delegated job credentials.
+\item [{\emph{CandidateTarget}}] \emph{-} specifies host name and queue
+of a computing resource.
 \end{lyxlist}
@@
 
-~~~~~~~~~~<xsd:complexType~name=?Glue2Resource?~minOccurs='0'>
+~~~~~~~~~~<xsd:complexType~name=\char`\"{}Glue2Resource\char`\"{}~minOccurs='0'>
 
@@
 
-~~~~~~~~~~~~~~~~~~<xsd:element~ref=?bes-factory:ActivityDocument?~minOccurs='0'/>
-
-~~~~~~~~~~~~~~~~~~<xsd:complexType~name=?Glue2Job?~minOccurs='0'>
+~~~~~~~~~~~~~~~~~~<xsd:element~ref=\char`\"{}bes-factory:ActivityDocument\char`\"{}~minOccurs='0'/>
+
+~~~~~~~~~~~~~~~~~~<xsd:complexType~name=\char`\"{}Glue2Job\char`\"{}~minOccurs='0'>
 
@@
 ~~~~~~</wsdl:output>
+
+~~~~~~<wsdl:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
 

### 2007-09-12 doc/a-rex/arex_tech_doc.tex
Changed title on request of Balazs. Language fixes from Adrian.

@@
 \begin{center}
-\textsc{\large The ARC Computational Job Management Module - A-Rex}
+\textsc{\large The ARC Computational Job Management Component
+- A-Rex}
 \par\end{center}{\large \par}
@@
 (SD). Input files are gathered in the SD. The job is supposed to produce
-new data files also in the SD. GM does not guarantee the availability
-of any other places accessible by the job other than SD (unless such
-place is part of requested Runtime Environement). The SD is also the
-only place which is controlled by the GM. It is is accessible by the
-user from outside through HTTPS and GridFTP protocols. Any file created
-outside the SD is not controlled by the GM. Any exchange of data between
-client and GM (including also program modules) is done through GridFTP
-\cite{gridftp} or HTTPS protocols depending on selected interface.
-A URL for accessing input/output files is obtained through the Information
-System or WS Information Interface of A-Rex.
+new data files  in the  SD as well. GM does not guarantee the
+availability of any other places accessible by the job other than
+SD (unless such a place is part of a requested Runtime Environement).
+The SD is also the only place which is controlled by the GM. It is
+is accessible by the user from outside through HTTPS and GridFTP protocols.
+Any file created outside the SD is not controlled by the GM. Any exchange
+of data between client and GM (including also program modules) is
+done through GridFTP \cite{gridftp} or HTTPS protocols depending
+on the selected interface. A URL for accessing input/output files
+is obtained through the Information System or the WS Information Interface
+of A-Rex.
 
@@
 
-Each job is initiated and controlled through GFS or WS. Complete description
-of job (not data) is passed to the GM through GFS in RSL \cite{rsl}
-or through WS in JSDL-coded \cite{jsdl} description (job description
-- JD).
+  Jobs are initiated and controlled through GFS or WS. Complete
+job descriptions   (JD)   are passed to the GM either
+through GFS in RSL \cite{rsl} or through WS in JSDL \cite{jsdl}
+coded description     . Input data files
+and job executables are transferred separately through the same interface,
+as described in the next section.
 
@@
 The main task of the GM is to take care of processing input and output
-data (files) of the job. Input files are gathered in SD. There are
-2 ways to put file into the SD:
-
-\begin{itemize}
-\item Downloads initiated by the GM. Such files (name and source) are defined
-in the JD. It is a sole responsibility of the GM to make sure that
-a file will be available in the SD.\\
-The supported sources are at the moment: GridFTP, FTP, HTTP, HTTPS
-(HTTP over SSLv3) and HTTPg (HTTP over GSI). Also some nonstandard
-sources are supported. Those are described below.
-\item Upload initiated by the user directly or through the User Interface
+data (files) of the job. Input files are gathered in the SD. There
+are two ways to put a file into the SD:
+
+\begin{itemize}
+\item Download is initiated by the GM --  This is the case for
+files     defined in the JD (with name and source).
+      The GM alone is responsible to make
+sure that all required files will be available in the SD.\\
+The supported protocols for sources  at the moment are: GridFTP,
+FTP, HTTP, HTTPS (HTTP over SSLv3) and HTTPg (HTTP over GSI). Also
+some nonstandard sources are supported. Those are described below.
+\item Upload is initiated by the user directly or through the User Interface
 (UI). Because the SD becomes available immediately at the time of
 submission of JD, UI can (and should) use that to upload data files
-which are not otherwise accessible by the GM. An example of such files
-can be the main executable of the job, files containing job's options/parameters,
-etc. These files can (and should) also be specified in the JD.
+which are not otherwise accessible by the GM.  Examples of such
+files  are the main executable of the job,   the
+job's input files, etc. These files can (and should)
+also be specified in the JD.
 \end{itemize}
@@
 data on the CE belonging to NorduGrid. Access to AFS, NFS, FTP, HTTP
-and any other remote data transport during execution of the job is
+and any other remote data transport during execution of a job is
 not guaranteed (at least not by GM).
 
-Job stores output files in the SD. Those files also belong to 2 groups:
+Jobs should store output files in thir SD. Like input files,
+output files  belong into two groups:
 
@@
 \item Files which are supposed to be moved to a \emph{Storage Element} (SE)
-and optionally registered in some \emph{Indexing Service} like \emph{Replica
-Catalog} (RC). The GM takes care of those files. They have to be specified
-in the JD. If job fails during any stage of processing no attept is
-taken to transfer those files to their final destination, unless there
-is option \emph{preserve=yes} specified in their URLs.
-\item Files which are supposed to be fetched by the user. The user runs
-UI to obtain those files. They \textbf{must} also be specified in
-the JD.
+and optionally registered in some \emph{Indexing Service} like the
+Globus \emph{Replica Location Service } (RLS)  --
+The GM takes care of these files. They have to be specified in the
+JD. If the job fails during any stage of processing, no attept is
+made to transfer those files to their final destination, unless
+the  option \emph{preserve=yes} is specified in their URLs.
+\item Files which are supposed to be fetched by the user -- The user has
+to use a tool like the UI to obtain these files. They \textbf{must}
+also be specified in the JD.
 \end{itemize}
@@
 A user can examine the state of a job by querying the NorduGrid Information
-System (IS) using the UI or any other tool. Please remember that IS
-can manipulate with state names to make them more user friendly and
-to combine them with states introduced by other parts of whole setup.
-Another way is to access virtual informational files through GridFTP
-interface or to use query metod of JCS.
-
-Configuration can put limits on amout of simultaneous jobs at some
-states. If such limit is reached job stays in it's current state waiting
-for free slot. This situation is presented by prepending current state
-name with \textbf{PENDING:} status mark.
-
-Below is description of all actions taken by the GM at every state:
-
-\begin{itemize}
-\item \textbf{Accepted} - At this state the job has been submitted to a
-CE but not processed yet. The GM will analyze the JD and move to the
-next stage. If JD can not be processed the job will be canceled and
-moved to the state \textbf{Finishing}.
-\item \textbf{Preparing} - The input data is being gathered in the SD. The
-GM is downloading files specified in the JD and waiting for files
-which are supposed to be downloaded by the UI. If all files are successfully
-gathered the job moves to the next state. If \textbf{any} file can't
-be downloaded or it takes UI too long to upload a file - the job moves
-to \textbf{Finishing} state. It is possible to put limit on number
-of simultaneous \textbf{Preparing} jobs. Those jobs out of limit will
-stay in previous \textbf{Accepted} state with PENDING mark. Exceptions
-are jobs which has no files to be downloaded. Those are processed
-out of limits.
-\item \textbf{Submitting} - This is a point of interaction with \emph{Local
-Resource Management System} (LRMS). At the moment PBS is supported
-best and correspoding backends are provided with default installation.
-The job is being submitted for execution. If the local job submission
-is successful the job moves to the next state. Otherwise it moves
-to \textbf{Finishing}. It is possible to limit number of jobs in \textbf{Submitting}
-and following \textbf{InLRMS} states.
+System (IS) using the UI or any other suitable tool. Please remember
+that the IS can manipulate  state names to make them more user
+friendly and to combine them with states introduced by other parts
+of the whole setup. Another way is to access virtual informational
+files through GridFTP interface or to use query method of WS interface.
+
+Configuration can put limits on the amout of simultaneous jobs in
+some states. If such a limit is reached, a job ready to enter into
+the state in question will stay in it's current state waiting for
+a free slot. This situation is presented by prepending the prefix
+\textbf{PENDING:} to the current state name  \textbf{}
+in the job's status mark.
+
+Below is the description of all actions taken by the GM at every state:
+
+\begin{itemize}
+\item \textbf{Accepted} - In this state the job has been submitted to
+a CE but is not processed yet. The GM will analyze the JD and move
+to the next stage. If the JD can not be processed the job will be
+canceled and moved to the state \textbf{Finishing}.
+\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
+The GM is downloading the files specified in the JD and is waiting
+for files which are supposed to be uploaded by the UI. If
+all files are successfully gathered the job moves to the next state.
+If \textbf{any} file can't be downloaded or it takes the UI too long
+to upload a file,  the job moves to \textbf{Finishing} state. It
+is possible to put a limit on the number of simultaneous \textbf{Preparing}
+jobs. If this limit is exceeded, jobs ready to enter the \textbf{Preparing}
+state      will stay in the \textbf{Accepted}
+state, but prefixed with the PENDING: mark. Exceptions are jobs which
+have no files to be downloaded. These are processed out of limits.
+\item \textbf{Submitting} -      The job is being
+passed for execution to the \emph{Local Resource Management System}
+(LRMS). At the moment PBS is supported best but correspoding backends
+for many other LRMSs are provided with the default installation. 
+      If the local job submission
+is successful the job moves to the \textbf{InLRMS} state. Otherwise
+it moves to \textbf{Finishing}. It is possible to limit the agregate
+number of jobs in \textbf{Submitting} and  \textbf{InLRMS}
+states.
 \item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
-The GM takes no actions except waiting until job finishes.
-\item \textbf{Cancelling} - Necessary action to cancel job in the LRMS is
-being taken.
-\item \textbf{Finishing} - The output data is being processed. Specified
-data files are moved to the specified SEs and are optionally registered
-at RC. The user can download data files from the SD by using UI or
-any other tool. All the files not specified as output files are removed
-from the SD at very beginning of this state. It is possible to limit
-number of simultaneous jobs in this state.
+The GM takes no actions except waiting until the job finishes.
+\item \textbf{Cancelling} - Necessary action to cancel the job in the LRMS
+is being taken.
+\item \textbf{Finishing} - The output data is being processed (stage-out).
+Specified data files are moved to the specified SEs and are optionally
+registered at  an Indexing Service. The user can download data files
+from the SD by using the UI or  other adequate tool. All the files
+not specified as output files are removed from the SD at very beginning
+of this state. It is possible to limit number of simultaneous jobs
+in this state.
 \item \textbf{Finished} - No more processing is performed by the GM. The
 user can continue to download data files from the SD. The SD is kept
-available for some time (default is 1 week). After that it is moved
-to the state \textbf{Deleted}. The 'deletion' time can be queried
-at NorduGrid Information System as attribute \texttt{nordugrid-pbs-job-sessiondirerasetime}
-of \texttt{nordugrid-pbs-job}. If job was moved to \textbf{Finished}
-because of failure, it may be restarted on request of client. Job
-is moved to a state previous to one which failed and is assigned mark
-PENDING. This is needed in order to not break the configuration limits.
-Exception is a job failed in \textbf{InLRMS} state and lacking input
-files specified in JD. Such job is treated like failed in \textbf{Preparing}
-state.
-\item \textbf{Deleted} - Job is moved to this state if user does not request
-job to be cleaned. Only minimal subset of information about such job
-is kept.
-\end{itemize}
-In the case of the failure special processing is applied to output
+available for some time (default is 1 week). After that the job
+is moved to the state \textbf{Deleted}. The 'deletion' time can be
+obtained by querying the NorduGrid Information System  attribute
+\texttt{nordugrid-pbs-job-sessiondirerasetime} of a \texttt{nordugrid-pbs-job}
+entry. If a job was moved to \textbf{Finished} because of failure,
+it may be restarted on request of a client. When restarted, a job
+is moved to the state previous to the one in which it failed and
+is assigned mark PENDING. This is needed in order to not break the
+configuration limits. Exception is a job failed in \textbf{InLRMS}
+state and lacking input files specified in JD. Such a job is treated
+like failed in \textbf{Preparing} state.
+\item \textbf{Deleted} - The job is moved to this state if the user have
+not requested job to be cleaned before the SD's lifetime expires.
+Only minimal subset of information about such job is kept. The SD
+is not available anymore.
+\end{itemize}
+In  case of a failure special processing is applied to output
 files. All specified output files are treated as \textbf{downloadable
-by user}. No files will be moved to the SE.
+by the user}. No files will be moved to their destination SE.
 
@@
 
-The GM and it's components support following data transfer protocols
+The GM and it's components support the following data transfer protocols
 and corresponding URLs: \emph{ftp, gsiftp, http, httpg, https, se,
@@
 jobs belonging to that user. Multiple users can share the same \textit{control
-directory}. To make it easier to recover in the case of failure, the
+directory}. To make it easier to recover in  case of failure, the
 GM stores most information in files rather than in memory. All files
-belonging to same job have names starting with \textbf{job.ID.} here
-ID is the job identifier.
+belonging to same the job have names starting with \textbf{job.ID.,}
+where ID is the job identifier.
 
@@
 \begin{itemize}
-\item \textit{job.ID.status} - current state of the job. It contains one
-word of text representing the current state of the job. Possible values
-are :
+\item \textit{job.ID.status} - current state of the job. This is a plain
+text file containing a single word   representing the current
+state of the job. Possible values are :
 
@@
 See section \ref{Section: Job Flow} for a description of the various
-states. Additionally each value can be prepended with a prefix {}``PENDING:''
-(like PENDING:ACCEPTED, see section \ref{Section: Job Flow}). That
-is used to show that the job is \emph{ready} to be moved to a next
-state and it stays in a current state \emph{only} because some limits
-set in configuration are exceeded.
+states. Additionally each value can be prepended  the prefix
+{}``PENDING:'' (like PENDING:ACCEPTED, see section \ref{Section: Job Flow}).
+This is used to show that a job  is \emph{ready} to be moved
+to the next state but it has to stay in it's current state \emph{only}
+because otherwise some limits set in the configuration would be
+exceeded.
 
@@
 job.
-\item \textit{job.ID.local} - information about job used by the GM. It consists
-of lines of format \textit{{}``name = value''} . Not all of them
-are always available. The following names are defined:
-
-\begin{itemize}
-\item \textit{subject} - user subject also known as the distinguished name
-(DN)
-\item \textit{starttime} - the GMT time when the job was accepted represented
-in Generalized Time format of LDAP 
-\item \textit{lifetime} - time period to live for the SD after job finished
-in seconds
-\item \textit{cleanuptime} - the GMT time when job to be removed from cluster
-and SD deleted in Generalized Time format
+\item \textit{job.ID.local} - information about the job used by the GM.
+It consists of lines of format \textit{{}``name = value''} . Not
+all of them are always available. The following names are defined:
+
+\begin{itemize}
+\item \textit{subject} - user certificate's subject, also known as the distinguished
+name (DN)
+\item \textit{starttime} -  GMT time when the job was accepted represented
+in the Generalized Time format of LDAP 
+\item \textit{lifetime} - time period to  preserve the SD after the
+job has finished in seconds
+\item \textit{cleanuptime} -  GMT time when the job should be removed
+from the cluster and it's SD deleted in Generalized Time format
 \item \textit{notify} - email addresses and flags to send mail to about
-job specified status changes
-\item \textit{processtime} - the GMT time when to start processing the job
+the job specified status changes
+\item \textit{processtime} -  GMT time when to start processing the job
 in Generalized Time format
-\item \textit{exectime} - the GMT time when to start job execution in Generalized
+\item \textit{exectime} -  GMT time when to start job execution in Generalized
 Time format
-\item \textit{expiretime} - the GMT time when credentials delegated to job
-expire in Generalized Time format
-\item \textit{rerun} - number of retries left to run the job
+\item \textit{expiretime} -  GMT time when the credentials delegated
+to the job expire in Generalized Time format
+\item \textit{rerun} - number of retries left to rerun the job
 \item \textit{jobname} - name of the job as supplied by the user
-\item \textit{lrms} - name of LRMS to run the job at
+\item \textit{lrms} - name of the LRMS backend     to be used
+for local submission
 \item \textit{queue} - name of the queue to run the job at
-\item \textit{localid} - job id in LRMS (appears only then the job is at
-state \textbf{InLRMS})
-\item \textit{args} - list of command-line arguments including the executable
-\item \textit{downloads} - number of files to download into SD before execution
-\item \textit{uploads} - number of files to upload from SD after execution
+\item \textit{localid} - job id in LRMS (appears only after the job
+ has reached state \textbf{InLRMS})
+\item \textit{args} - executable name followed by a list of command-line
+arguments   
+\item \textit{downloads} - number of files to download into the SD before
+execution
+\item \textit{uploads} - number of files to upload from the SD after execution
 \item \textit{gmlog} - directory name which holds files containing information
-about job when accessed through GridFTP interface
-\item \textit{clientname} - name and ip address:port of client machine (name
-is provided by user interface)
-\item \textit{clientsoftware} - version of software used to submit job
-\item \textit{sessiondir} - SD of job
-\item \textit{failedstate} - state at which job failed (available only if
-it is possible to restart job)
-\item \textit{jobreport} - URL of \emph{logger service} used to keep track
-of executed jobs (one requested by user)
+about the job when accessed through GridFTP interface
+\item \textit{clientname} - name (as provided by the user interface) and
+ip address:port of the submitting client machine
+
+\item \textit{clientsoftware} - version of software used to submit the job
+\item \textit{sessiondir} - the job's SD  
+\item \textit{failedstate} - state in which job failed (available only
+if it is possible to restart the job)
+\item \textit{jobreport} - URL of a user requested \emph{logger service}.
+The GM will also send job records to this service in addition to the
+default logger service configured in the configuration
+
 \end{itemize}
@@
 The job submission and control interface is based on document produced
-by OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
-The exchange of SOAP messages is performed over HTTPS protocol. The
-BES interface is represented by two port-types - BES-Management and
-BES-Factory. The former is made to control the A-REX service itself
+by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
+The exchange of SOAP messages is performed over the HTTPS protocol.
+The BES interface is represented by two port-types - BES-Management
+and BES-Factory. The former is made to control the A-REX service itself
 and thus defines operations to start and stop the functionality of
@@
 \begin{itemize}
-\item \emph{NotAuthorizedFault}: Indicates that client is not allowed to
-do this operation.
+\item \emph{NotAuthorizedFault}: Indicates that the client is not allowed
+to do this operation.
 \item \emph{InvalidActivityIdentifierFault}: There is no such job/activity.
-\item \emph{CantApplyOperationToCurrentStateFault}: Requested transition
+\item \emph{CantApplyOperationToCurrentStateFault}: The requested transition
 is not possible.
@@
 the private key stored at the service side associated with these credentials.
-The reference element refers to the object which these credentials
-should be applied to in a way specific to the service. The same element
+The reference element refers to the object to which these credentials
+should be applied  in a way specific to the service. The same element
 must also be used for delegating credentials as part of other operations
@@
 \end{lyxlist}
-\item [{\emph{Input}\,\ (POSIX)}] - file to be used as standard input
-for the executable.
+\item [{\emph{Input}\ (POSIX)}] - file to be used as standard input for
+the executable.
 \item [{\emph{Output}\ (POSIX)}] - file to be used as standard output
@@
 \item \textit{statistics} - consists of strings containing \emph{name=value}
-. Following names are defined:
+pairs. Following names are defined:
 
@@
 
-If job requests input file, which can/allowed to be cached, it is
-stored in cache directory instead and soft-link is created in the
-SD, pointing to that file. Or file can be stored in cache and then
-copied to the SD. Last option is more secure and hence advised.
-
-Before downloading file the GM tries to determine it's size and to
-preallocate space in cache directory, by writing file of same size.
-If that fails (file system has no more space), it tries to remove
-oldest cache files, which are not being used by any job. That means
-\textbf{hard limit of cache size is space available at file-system}.
-In case cache gets full and it is impossible to free any space, download
-fails and is retried without using cache.
-
-Before giving access to already cached file the GM contacts initial
-file source to check if user is allowed to do that if protocols allows
-to do that.
-
-Also file creation or validity times are checked to make sure cached
-file is fresh enough. If it is impossible to obtain creation and invalidation
-times for file it is invalidated 24 hours after downloaded.
-
-Also the GM checks cache periodically. If used space exceeds high
-water-mark given in configuration file (\emph{softsize}) it tries
-to remove oldest unused files to reduce size to low water-mark. This
-sets soft limit of cache size.
-
-There are 2 kinds of caches supported. Files in \textit{private} cache
-are owned by Unix user to which grid user is mapped. Those files are
-readable only by that particular Unix user. Another kind of cache
-is \textit{shared}. Files are owned by Unix user who started GM and
-are readable by everyone.
+If a job requests an input file which is subject  for
+cacheing, it is stored in the cache directory instead of the SD.
+The file is made available to the job by either  soft-linking it
+  in the SD         
+    or copying it to the SD. The latter option
+is more secure and hence advised.
+
+Before downloading a file the GM tries to determine it's size and
+then tries to preallocate space in the cache directory by writing
+a file of the same size. If this fails (possibly because the file
+system has no more space), it tries to the remove oldest cached files
+which are not being used by any job. This means that \textbf{the
+hard limit of cache size is the space available in the file-system}.
+In case cache gets full and it is impossible to free enough space,
+the download fails and then is retried without using cache.
+
+Before giving access to an already cached file, the GM contacts the
+original file source to check if the user     has
+suficient access rights to the file. Not all protocols  
+ make this check possible.
+
+Also, file creation or validity times are checked to make sure the
+cached file is fresh . If it is impossible to obtain creation
+and invalidation times  , the file is invalidated 24 hours
+after it was downloaded.
+
+ The GM checks the cache periodically. If  the space used
+by cache exceeds the high water-mark given in the configuration file
+(\emph{softsize}), it tries to remove the oldest unused files until
+cache size drops below the  low water-mark. This sets the
+soft limit of cache size.
+
+There are two kinds of caches supported. Files in \textit{private}
+cache are owned by the Unix user to which a grid user is mapped. Those
+files are readable only by that particular Unix user. Another kind
+of cache is \textit{shared}. Files in the shared cache are owned by
+the Unix user who started GM and are readable by everyone.
 
@@
 
-The GM consists of few separate executable modules. Those are:
-
-\begin{itemize}
-\item \textit{grid-manager} - Main module. It is responsible for processing
-the job, moving it through states, running other modules.
+The GM consists of several separate executable modules. These
+are:
+
+\begin{itemize}
+\item \textit{grid-manager} - The main module. It is responsible for processing
+ jobs, moving them through states and running other modules.
 \item \textit{downloader} - This is a module responsible for gathering input
@@
 \item \textit{uploader} - This module is responsible for delivering output
-files to the specified SEs and registration at the RC. It processes
-and updates the \textit{job.ID.output} file.
-\item \emph{cache-register} - Utility to register cached data into Indexing
-Services like RC and RLS. It reads and modifies cache informational
-files \emph{old} and \emph{new} \ref{sec:cache}. Configuration is
-read directly from the GM's configuration file \ref{SubSection: ConfigFile}.
-It is run by the GM every 5 minutes.
+files to the specified SEs and registration at  an Indexding
+Service (like RLS) as needed. It processes and updates the \textit{job.ID.output}
+file.
+\item \emph{cache-register} - Utility to register cached data into an Indexing
+Service    . It reads and modifies cache informational
+files \emph{old} and \emph{new} (as described in section \ref{sec:cache}).
+Configuration is read directly from the GM's configuration file (see
+section \ref{SubSection: ConfigFile}). It is run by the GM every
+5 minutes.
 \item \emph{frontend-info-collector} - Utility to gather information about
-frontend and to put it into \emph{job.ID.diag} file.
-\item \emph{gm-kick} - Sends signal to the GM though FIFO file to wake it
-up. It's used to increase responsiveness of GM.
-\end{itemize}
-Following modules are always run under Unix account to which user
-is mapped.
+the frontend  . It puts  collected information into the \emph{job.ID.diag}
+file.
+\item \emph{gm-kick} - Sends a signal to the GM though a FIFO file to wake
+it up. It's used to increase responsiveness of GM.
+\end{itemize}
+The following modules are always run under the Unix account to which
+a grid user is mapped.
 
@@
 pseudo-LRMS is supported for testing purposes. This module is responsible
-for the job submission to the LRMS.
-\item \textit{cancel-{*}-job} - This one is for canceling the job, which
-was submitted to LRMS.
+for  job submission to the LRMS.
+\item \textit{cancel-{*}-job} - This script is for canceling  jobs
+which have been already submitted to the LRMS.
 \item \textit{scan-{*}-job} -This shell script is responsible for notifying
-the GM about completion of the job. It's implementation for PBS system
+the GM about completion of  jobs. It's implementation for PBS 
 uses server logs to extract information about jobs. If logs are not
-available it uses less reliable \emph{qstat} command for that. other
-backends use different techniques.
-\end{itemize}
-Also there is administrator utility:
-
-\begin{itemize}
-\item \textit{gm-jobs} - prints list of jobs available on cluster and amount
-of jobs in every state.\\
+available it uses the less reliable \emph{qstat} command for that.
+Other backends use different techniques.
+\end{itemize}
+ In addition, there is also an administration utility:
+
+\begin{itemize}
+\item \textit{gm-jobs} - prints a list of jobs available on the cluster
+and the number of jobs in each state.\\
 \hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name]\\
-\hspace*{0.5cm}-l-l - print more information about each job,\\
-\hspace*{0.5cm}-l-u - pretend utility is run by user with id \emph{uid},\\
-\hspace*{0.5cm}-l-l - pretend utility is run by user with name \emph{name}.
+\hspace*{0.5cm}-l -- print more information about each job,\\
+\hspace*{0.5cm}-u -- pretend utility is run by user with id \emph{uid},\\
+\hspace*{0.5cm}-U -- pretend utility is run by user with name \emph{name}.
 \end{itemize}
 GM comes with plugins useable for various authorization purposes (see
-for example description of \emph{authplugin} command below):
+for example the description of \emph{authplugin} command below):
 
@@
 \hspace*{0.5cm}inputcheck {[}-h] {[}-d debug\_level] RSL\_fle {[}proxy\_file]\\
-\hspace*{0.5cm}-lRSL\_file -file with job description,\\
-\hspace*{0.5cm}-lproxy\_file - credentials proxy.
+\hspace*{0.5cm}- RSL\_file -- file with job description,\\
+\hspace*{0.5cm}- proxy\_file -- credentials proxy.
 \item \textit{lcas} - executes LCAS plugins on credentials and returns 0
@@
 \hspace*{0.5cm} lcas credentials description {[}library {[}db {[}directory]]]\\
-\hspace*{0.5cm}-lcredentials - path to file with credentials to authorize,\\
-\hspace*{0.5cm}-ldescription - path to file with job description,\\
-\hspace*{0.5cm}-llibrary - path to LCAS library (full or relative
+\hspace*{0.5cm}- credentials -- path to file with credentials to
+authorize,\\
+\hspace*{0.5cm}- description -- path to file with job description,\\
+\hspace*{0.5cm}- library -- path to LCAS library (full or relative
 to LCAS directory),\\
-\hspace*{0.5cm}-ldb - path to LCAS DB file (full or relative to LCAS
-directory),\\
-\hspace*{0.5cm}-ldirectory - LCAS directory.
+\hspace*{0.5cm}- db -- path to LCAS DB file (full or relative to
+LCAS directory),\\
+\hspace*{0.5cm}- directory -- LCAS directory.
 \end{itemize}
@@
 The GM is installed into a single installation point referred as \$NORDUGRID\_LOCATION
-and following sub-directories are used:
+and the following sub-directories are used:
 
 \begin{LyXParagraphLeftIndent}{1cm}
-\$NORDUGRID\_LOCATION/bin - tools
-
-\$NORDUGRID\_LOCATION/libexec - program modules used by GM
-
-\$NORDUGRID\_LOCATION/etc - configuration files, deprecated, central
-configuration file is used by deault
-
-\$NORDUGRID\_LOCATION/sbin - daemons
-
-\$NORDUGRID\_LOCATION/lib - gridftp server's plugins and API libraries
-
+\$NORDUGRID\_LOCATION/bin -- tools
+
+\$NORDUGRID\_LOCATION/libexec -- program modules used by GM
+
+\$NORDUGRID\_LOCATION/etc -- configuration files, deprecated, central
+configuration file is used by default
+
+\$NORDUGRID\_LOCATION/sbin -- daemons
+
+\$NORDUGRID\_LOCATION/lib -- gridftp server's plugins and API libraries
+
+\end{LyXParagraphLeftIndent}
 The GM also uses following directories:
 
-\end{LyXParagraphLeftIndent}
-\begin{itemize}
-\item \textit{session root directory} - In this directory the SD is created.
-It can be multiple directories for the various users specified in
-the configuration file.\\
-There are 2 processes which need to have a permissions to create new
-files and directories in it. Those are GM and GFS. \\
-If any of those processes processes are run under dedicated user account,
-that account need full permissions in the \textit{session root directory}.\\
-If those processes are run under \emph{root} account make sure \textit{session
-root directory} is not on filesystem which limits capabilities of
-\emph{root} user. For example NFS with \emph{root\_squash} option.\\
-If there is need to run processes under \emph{root} account (to run
-jobs in LRMS under different users' accounts) but there is no way
-to provide suitable \textit{session root directory} use \emph{norootpower}
-command in configuration of the GM. In that case GM and GFS will use
-identity of local user to which Grid identity is mapped to access
-\textit{session root directory}. Hence those users will need full
-access there.\\
-The GM creates SD with proper ownership and permissions for local
-identity used to run job. Some filesystems require \emph{executable}
-permissions on \textit{session root directory} to be set for local
-identity in order to access any file or subdirectory there.\\
-This directory should also be shared among cluster nodes in order
-for job to access input files. Or internal means of LRMS must be used
-to transfer files to executing node. For more see section \ref{sub:LRMS}.
-\item \textit{control directory} - In this directory the SD stores an information
-about the accepted jobs. Both GM and GFS processes must have full
-permissions there.\\
-Also subdirectory called \textit{log} is created there. It is used
+\begin{itemize}
+\item \textit{session root directory} -  Tthis is the directory under
+which a user's SDs are created. It's location is configurable
+per UNIX user. Several (or even all) users may share the same session
+root directory.        
+    .\\
+There are 2 processes which need to have  permissions to create new
+files and directories in it  : GM and GFS. \\
+If any of these processes  are run under a dedicated user
+account, that account needs full permissions in the \textit{session
+root directory}.\\
+If these processes are run under the \emph{root} account, make sure
+\textit{session root directory}  resides on a filesystem which
+does not limit the capabilities of the \emph{root} user  (as does
+for example NFS with \emph{root\_squash} option).\\
+If there is a need to run processes under the \emph{root} account
+(to be able to run jobs in LRMS under different users' accounts, for
+example) but there is no way to provide a suitable \textit{session
+root directory,} use the \emph{norootpower} command in GM's configuration
+  file. In that case GM and GFS will use the identity of the
+local user to which a Grid identity is mapped to access the \textit{session
+root directory}. Hence those users will need full access there.\\
+The GM creates SDs with proper ownership and permissions for the local
+identity used to run a job. Some filesystems require users to have
+\emph{execute} permission on the \textit{session root directory}
+      in order to access any file or subdirectory
+there.\\
+         In order
+for jobs to access their input files, session root directories should
+be shared across cluster nodes. Otherwise,    LRMS-specific
+methods must be used to transfer files to execution nodes. For more
+information see section \ref{sub:LRMS}.
+\item \textit{control directory} - In this directory  GM and GFS store
+an information about  accepted jobs. Both    processes
+must have full permissions there.\\
+A subdirectory called \textit{log} is created there. It is used
 to accumulate information about started and finished jobs. This information
-is periodically sent to the \emph{logger service}.
+is periodically sent to the desired \emph{logger service}(s). For
+each job start and stop event, and for each logger service where that
+event must be sent, a separate file is written. Once an event is sent,
+the corresponding file is deleted.
 \end{itemize}
@@
 
-The GM configuration is done through single configuration file. Historically
-GM supports 2 kinds of configuration files. For old one it looks at
-following places:
+The GM configuration is done through a single configuration file.
+Historically GM supports 2 kinds of configuration files. For old one
+(deprecated) it looks at following places:
 
@@
 starting from \#. It is separated into sections. Each sections starts
-from string containing \emph{}
+from string containing
 
@@
 contain path corresponding to \textit{RUNTIME\_LOCAL\_SCRATCH\_DIR}
-as \textbf{}seen on \textbf{frontend} machine. \emph{shared\_scratch}
-configuration command.
+as seen on \textbf{frontend} machine. \emph{shared\_scratch} configuration
+command.
 
@@
 and RUNTIME\_FRONTEND\_SEES\_NODE and explain how data movement is
-performed. Pictures a) correspond to situation right after all input
-files are gathered in session directory and actions taken right after
-job's script starts. Pictures b) show how it looks while job is running
-and actions which are taken right after it finished. Pictures c) stand
-for final situation, when job files are ready to be uploaded to external
-storage element or be downloaded by user.
+performed. Pictures a) corresponds to the situation right after all
+input files have been gathered in the session directory and show
+the actions taken right after the job's script starts. Pictures b)
+shows   the situation while the job is running and the actions
+which are taken right after it has finished. Pictures c)  illustrates
+the final situation, when the job's output files are ready to be uploaded
+to an external storage element or be downloaded by the user.
 
@@
 \caption{\label{fig:no node scratch}Both RUNTIME\_LOCAL\_SCRATCH\_DIR and
-RUNTIME\_FRONTEND\_SEES\_NODE undefined. Job is executed in session
-directory placed on frontend.}
+RUNTIME\_FRONTEND\_SEES\_NODE are undefined. It is assumed that session
+directories are visible from computing nodes. The job is executed
+directly in the session directory prepared by GM on the frontend.}
 \end{figure}
@@
 \caption{\label{fig:node scratch, not vis on front}RUNTIME\_LOCAL\_SCRATCH\_DIR
-is set to value representing sratch directory on computing node, RUNTIME\_FRONTEND\_SEES\_NODE
-undefined.}
+is set to  the location of the scratch directory
+on the computing node, RUNTIME\_FRONTEND\_SEES\_NODE is undefined.}
 
 \begin{lyxlist}{00.00.0000}
-\item [{a)}] After job script starts all input files are moved to 'scratch
-directory' on computing node.
-\item [{b)}] Job runs in separate directory in 'scratch directory'. Only
-files representing job's \textit{stdout} and \textit{stderr} are placed
-in original 'session directory' and soft-linked in 'scratch'. After
-execution all files from 'scratch' are moved back to original 'session
-directory'.
+\item [{a)}] After the job script starts all input files are moved to 'scratch
+directory' on the computing node.
+\item [{b)}] The job runs in a separate directory in 'scratch directory'.
+Only the files representing the job's \textit{stdout} and \textit{stderr}
+are placed in the original 'session directory' and soft-linked in
+'scratch'. After execution all files from 'scratch' are moved back
+to the original 'session directory'.
 \item [{c)}] All output files are in 'session directory' and are ready
@@
 \caption{\label{fig:node scratch, vis on front}Both RUNTIME\_LOCAL\_SCRATCH\_DIR
-and RUNTIME\_FRONTEND\_SEES\_NODE are set to valuea representing sratch
-directory on computing node and way to access that scratch from frontend
-correspondingly.}
+and RUNTIME\_FRONTEND\_SEES\_NODE are set to the  location
+of the scratch directory on the computing node and    the
+location where this scratch directory is accessible from the frontend,
+respectively .}
 
 \begin{lyxlist}{00.00.0000}
-\item [{a)}] After job script starts all input files are moved to 'scratch
-directory' on computing node. Original 'session directory' is removed
-and replaced with soft-link to copy of session directory in 'scratch'
-as seen on frontend.
-\item [{b)}] Job runs in separate directory in 'scratch directory'. All
-files are also available on frontend through soft-link. After execution
-soft-link is replaced with directory and all files from 'scratch'
-are moved back to original 'session directory'.
+\item [{a)}] After the job script starts all input files are moved to 'scratch
+directory' on the computing node. The original 'session directory'
+is removed and replaced with a soft-link to copy of session directory
+in 'scratch directory' as seen on the frontend.
+\item [{b)}] The job runs in a separate directory in 'scratch directory'.
+All files are also available on frontend through the soft-link. After
+execution the soft-link is replaced with a directory and all files
+from 'scratch' are moved back to the original 'session directory'.
 \item [{c)}] All output files are in 'session directory' and are ready
@@
 will also need \emph{Globus Toolkit$^{TM}$} of version higher than
-2.2 \emph{}installed \url{http://www-unix.globus.org/toolkit/}.
+2.2 installed \url{http://www-unix.globus.org/toolkit/}.
 
@@
 
-\subsection{Unix accounts}
-
-Bot GM and GFS are designed to be run by \emph{root} UNIX account
-and serve multiple local UNIX and global Grid identities. Nevertheless
-it is possible to use \emph{non-root} accounts to run those services.
-Although this means some functionality loss described below.
-
-There are no implication on running GFS with \emph{gaclplugin} or
-\emph{fileplugin} under \emph{non-root} account as long as only Grid
-identity of user is used and all served files and directories are
-owned by server's account.
-
-For combination of GM and GFS with \emph{jobplugin} both services
-must be run either by same account or one of services must be run
-under \emph{root} account. That is needed because services communicate
-over local filesystem, hence must have \emph{full} access to same
-set of files.
-
-As long as GFS with \emph{jobplugin} is run under non-root account
-there is no mapping from Grid identity to local UNIX account taking
-place. All alowed Grid users are assigned server's account and are
-then processes by GM using same account. Only way to overcome this
-limitation is to run one GFS per local account with proper access
+\subsection{ Running as non-root}
+
+Bot GM and GFS are primarily designed to be run by the \emph{root}
+UNIX account and serve multiple    global Grid identities
+mapped to several UNIX accounts. Nevertheless it is possible to use
+\emph{non-root} accounts to run these services   
+at the cost of some functionality loss as described below.
+
+There are no drawbacks of running GFS with \emph{gaclplugin}
+or \emph{fileplugin} under a \emph{non-root} account as long as the
+only Grid identity used is that of the user   who runs the services
+and all served files and directories are owned by the server's account.
+
+  In order for GM and GFS with \emph{jobplugin} to
+be able to cooperate, both services must be run either by the same
+non-root account or one of the services must be run under the \emph{root}
+account. This is needed because the two services communicate over
+a local filesystem, hence they must have \emph{full} access to the
+same set of files.
+
+As long as GFS with \emph{jobplugin} is run under a non-root account
+ , no mapping from Grid identity to local UNIX account takes
+place. All alowed Grid users are assigned the server's account and
+are then processes by GM using same account. The only way to overcome
+this limitation is to run one GFS per local account with proper access
 control configured.
 
-Because GM has to represent user's local account while communication
-with LRMS, it can serve only account it is run under (unless it is
-run under \emph{root} account, of course). Like in case of GFS, multiple
-instances of GM may be run, one per local account. That solution causes
-another implications. The GM looses possibility to share cached files
-among serviced users. It is not also possible to control load on a
-frontend by limiting number of simultatenuosly running \emph{downloader}
-and \emph{uploader} modules.
-
-One has also take into account that private part of GSI infrastructure
-(private key of a host at least) has to be duplicated for every account
-used to run GFS.
+Because GM has to impersonate a user's local account while
+communicating with the LRMS, it can serve only the account it is
+run under (unless it is run under the \emph{root} account, of course).
+As in the case of GFS, multiple instances of GM may be run, one
+per local account. This solution   has
+some drawbacks. The GM looses possibility to share cached files among
+serviced users. In addition, it is not  possible to control the
+load on a frontend by limiting the number of simultatenuosly running
+\emph{downloader} and \emph{uploader} modules.
+
+One  also has to take into account that the private part of the
+GSI infrastructure (private key of a host at least) has to be duplicated
+for every account used to run GFS.
 
@@
 
-Under mount point of jobplugin gridftp client can see directories
-representing job belonging to user, who started client. Directory
-per job. Directory names are same as jobs' identifiers. Those directories
-are directly connected to session directories of jobs and contain
-same files and subdirectories. Except if jobs session directory is
-moved to computing node. In that case directories only contain files
-with redirected stdout and strderr as specified in xRSL.
-
-If job's xRSL has \emph{gmlog} specified job's directory also contains
-virtual subdirectory with same name, which contains files with information
-about job as created by GM. The most important are 'errors' and 'status'.
-'errors' contains stderr of separate modules run by GM in order to
-process job (downloader, uploader, job's submission to LRMS). 'status'
-contains one word representing state of job.
-
-Also under mount point there is additional directory named \char`\"{}new\char`\"{}
-used to submit new jobs. And another directory {}``info'' with subdirectories
-named after job ids. Those subdirectories contain files with information
-about job identical to those in subdirectory specified through \emph{gmlog}.
+Under the mount point of the jobplugin, a user connecting with a gridftp
+client will see virtual directories representing the jobs belonging
+to    him/her. Directory names are job identifiers,
+each representing one job      
+  . These directories are directly connected to
+session directories of jobs and contain the  files and subdirectories
+that are visible on the frontend.
+
+If a job's xRSL has \emph{gmlog} specified, then the job's directory
+also contains a virtual subdirectory  of that name 
+ holding files with information about the job as created by
+GM. The    'errors' and 'status' files are especially
+usefull for troubleshooting. 'errors' contains the stderr output of
+the various modules run by GM   during job processing
+stages (downloader, uploader, job's submission to LRMS). 'status'
+contains one word representing the job's state  .
+
+Also under the jobplugin's mount point there is an additional directory
+named \char`\"{}new\char`\"{} -- used to submit new jobs -- and
+another directory named ''info''. The latter has subdirectories
+named after job ids  , each of which contains
+files with information about a job. These are the same files 
+ that can be accessed in the subdirectory specified through
+\emph{gmlog}.
 
@@
 
-Each xRSL put into directory \char`\"{}new\char`\"{} is accepted as
-job's description. jobplugin parses it and client gets positive response
-if there are no errors in request.
-
-Job gets identifier and directory with corresponding name appears.
-If job's description contains input files which should be delivered
-from client's machine, client must upload them to that directory under
-specified names.
-
-Because each job gets identifier there should be a way for client
-to obtain it. For that prior to providing xRSL client sends command
-CWD to change current directory to \char`\"{}new\char`\"{}. In this
-way job's identifier is reserved, new directory corresponding to that
-identifier is created and client is redirected to it (as specified
-in FTP protocol). Job's description put into \char`\"{}new\char`\"{}
-will get reserved identifier.
+Each xRSL put into the {}``new'' directory by a gridftp client
+is treated as a new job's description. GFS's jobplugin parses
+the xRSL and returns the client a positive response if there
+were no errors in the request.
+
+The new job gets an identifier and a directory with the corresponding
+name appears. If the job's description contains input files which
+should be transferred from the client's machine, the client
+must upload them to that directory under specified names.
+
+   The job identifier reserved by GFS to a new job must
+be somehow communicated back to the client. Within the bounds of the
+FTP protocol, this is achieved in the following way. Prior to uploading
+the xRSL, the client issues a CWD command to change the current directory
+to {}``new''. In turn, the server responds with a redirect to the
+new session directory named after the reserved identifier. The client
+has now found out the job's id, and proceeds with uploading the xRLS.
 
@@
 
-Various actions to affect processing of existing job are performed
-by uploading xRSL files into directory {}``new''. Content of xRSL
-may consist of only 2 parameters - action for \emph{action} to be
-performed, and \emph{jobid} to identify job to be affected. Rest of
-parameters are ignored.
-
-Currently supported actions are:
+Various actions to affect processing of an existing job are requested
+by uploading special xRSL files into directory {}``new''. Such an
+  xRSL may consist of only 2 parameters - \emph{action} for
+action to be performed, and \emph{jobid} to identify the job to be
+affected.  All other parameters are ignored.
+
+The currently supported actions are:
 
 \begin{lyxlist}{00.00.0000}
-\item [{\emph{cancel}}] to cancel job
-\item [{\emph{clean}}] to remove job from computing resource
-\item [{\emph{renew}}] to renew credentials delegated to job
-\item [{\emph{restart}}] to restart job after failure at some phases
+\item [{\emph{cancel}}] to cancel a job
+\item [{\emph{clean}}] to remove a job from computing resource
+\item [{\emph{renew}}] to renew credentials delegated to a job
+\item [{\emph{restart}}] to restart a job after failure at some phases
 \end{lyxlist}
-It is also possible to perform some actions by using shortcut FTP
-operations described below.
+Alternatively, it is also possible to perform some of these actions
+by using the shortcut FTP operations described below.
 
@@
 
-Job is canceled by performing DELE (delete file) command on directory
-representing job. It can take some time (few minutes) before job is
-actually canceled. Nevertheless client gets response immediately.
+A job is canceled by performing a DELE (delete file) command on the
+directory representing the job. It can take some time (a few minutes)
+before the job is actually canceled. Nevertheless, the client gets
+a response immediately.
 
@@
 
-Job's content is cleaned by performing RMD (remove directory) command
-on directory representing job. If job is in \char`\"{}FINISHED\char`\"{}
+A job's content is cleaned by performing a RMD (remove directory)
+command on the directory representing the job. If the job is in \char`\"{}FINISHED\char`\"{}
 state it will be cleaned immediately. Otherwise it will be cleaned
@@
 
-If client requests CWD to session directory credentials passed during
-authentication are compared to current credentials of the job. If
-validity time of the new credentials is longer job's credentials are
-replaced with new.
+If a client requests CWD to a job's session directory, credentials
+passed during authentication are compared to the current credentials
+of the job. If the validity time of the new credentials is longer,
+the job's current credentials are replaced with the new ones.
 

### 2008-01-22 doc/dmc/dmc.tex
DMC documentation (incomplete)

\documentclass{book}
\usepackage{graphicx}                              %for PNG images
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{makeidx}                               %for index page
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\makeindex
\hypersetup{
  pdfauthor = {Mattias Ellert},
  pdftitle = {The ARC Data Manager Component (DMC) Implementation Guide},
  pdfsubject = {DMC Implemenatation Guide},
  pdfkeywords = {Grid,NorduGrid,ARC,DMC},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
\begin{document}

\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

  \begin{tabular}{rl}
    \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
    &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
  \end{tabular}

  \hrulefill

  {\raggedleft NORDUGRID-TECH-16\par}

  {\raggedleft \today\par}

  \vspace*{2cm}

  {\centering \textsc{\Large ARC Data Manager Component (DMC)}\Large \par}

  \vspace*{0.5cm}

  {\centering \textit{\large Implementation Guide}\large \par}

  \vspace*{1.5cm}

  {\centering \large Mattias Ellert\footnote{mattias.ellert@tsl.uu.se} \large \par}

\end{titlepage}

\tableofcontents
%\listoffigures
\newpage
\chapter{Introduction}
\label{sec:intro}

The libarcdata2 library in the new Advanced Resource Connector (ARC)
is based on libarcdata library in the old ARC. Unlike its predecessor
the libarcdata2 library takes advantage of the plugin mechanism of the
ARC Loader. The core libarcdata2 library does not introduce any
additional external dependencies. Plugins for specific data access
protocols can however have various external dependencies. This
separation of external dependencies from the core library helps reduce
the minimum set of requirements for ARC while allowing the support for
additional access protocols requiring special dependencies to be
installed by those who need it.

The plugins used by the libarcdata2 library are called Data Manager
Components (DMCs). Each DMC should provide a class that inherits from
the DMC base class. Like other ARC Loader plugins the DMCs should
provide a descriptor that maps the name of the plugin to an function
that creates an instance of the DMC. In order for the default DMC
client configuration to automatically find the DMC the name of the DMC
in the descriptor and the name of the loadable module containing the
DMC should match. A DMC called X should reside in a module called
libdmcX.

The constructor and destructor of a DMC should call the Register and
Unregister method of respectively.

The most central class in the arcdata2 library is the DataPoint
class. This class describes a source or destination of a file
transfer.

A DMC should provide an iGetDataPoint method that returns a pointer to
an instance of a specialisation of the DataPoint class when called
with a URL. If the DMC does not support the provided URL a NULL
pointer should be returned instead.

The libarcdata2 library contains a wrapper class for the DataPoint
class called DataHandle. Calling the constructor of the DataHandle
class with a URL will trigger the iGetDataPoint class of all
registered DMCs to be called until a non-NULL pointer is returned.
By dereferencing the DataHandle object the underlying DataPoint can be
accessed.

Two partial specialisations of the DataPoint class are available in
the libarcdata2 library. These are the DataPointIndex and
DataPointDirect classes. The DataPointIndex class implements common
functionality for DMCs that support file catalogues, while the
DataPointDirect implement common functionality for DMCs that support
direct file access. The specialisation of the DataPoint class used by
a DMC should normally inherit from one of these classes rather than
from the DataPoint base class directly.

\chapter{The DMC class}
\label{sec:DMC}

The DMC class describes a loadable module for arcdata2 library. A
typical specialization of this class for a custom DMC plugin would
look like this:

\begin{verbatim}
#ifndef __ARC_DMCCUSTOM_H__
#define __ARC_DMCCUSTOM_H__

#include <arc/data/DMC.h>

namespace Arc {

  class DMCCustom : public DMC {
   public:
    DMCCustom(Config *cfg);
    ~DMCCustom();
    static DMC* Instance(Config *cfg, ChainContext *ctx);
    DataPoint* iGetDataPoint(const URL& url);
   protected:
    static Logger logger;
  };

} // namespace Arc

#endif // __ARC_DMCCUSTOM_H__
\end{verbatim}

The corresponding implementation should look like this:

\begin{verbatim}
#include <arc/Logger.h>
#include <arc/URL.h>
#include <arc/loader/DMCLoader.h>

#include "DMCCustom.h"
#include "DataPointCustom.h"

namespace Arc {

  Logger DMCCustom::logger(DMC::logger, "Custom");

  DMCCustom::DMCCustom(Config *cfg) : DMC(cfg) {
    Register(this);
  }

  DMCCustom::~DMCCustom() {
    Unregister(this);
  }

  DMC* DMCCustom::Instance(Arc::Config *cfg, Arc::ChainContext*) {
    return new DMCCustom(cfg);
  }

  DataPoint* DMCCustom::iGetDataPoint(const URL& url) {
    if (url.Protocol() != "custom") return NULL;
    return new DataPointCustom(url);
  }

} // namespace Arc

dmc_descriptors ARC_DMC_LOADER = {
  { "custom", 0, &Arc::DMCCustom::Instance },
  { NULL, 0, NULL }
};
\end{verbatim}

The DataPointCustom class in the above example is the specialization
of the DataPoint class for this DMC (see chapter \ref{sec:DataPoint}).

The DMC can support more than one protocol in which case the
iGetDataPoint should be changed accordingly.

\chapter{The DataPoint class}
\label{sec:DataPoint}

The DataPoint class is the most central class in the arcdata2
library. This is an abstract base class that must be specialized for
each DMC. Two partial specialisations of the DataPoint class are
available in the libarcdata2 library. These are the DataPointIndex and
DataPointDirect classes. The DataPointIndex class implements common
functionality for DMCs that support file catalogues, while the
DataPointDirect implement common functionality for DMCs that support
direct file access. The specialisation of the DataPoint class used by
a DMC should normally inherit from one of these classes rather than
from the DataPoint base class directly.

\section{DataPointDirect}

To create a DataPoint class for a direct access protocol, the
following methods should be provided:

\begin{verbatim}
namespace Arc {

  class DataPointCustomDirect : public DataPointDirect {
   public:
    DataPointCustomDirect(const URL& url);
    virtual ~DataPointCustomDirect();
    virtual bool start_reading(DataBufferPar& buffer);
    virtual bool start_writing(DataBufferPar& buffer,
                               DataCallback *space_cb = NULL);
    virtual bool stop_reading(void);
    virtual bool stop_writing(void);
    virtual bool analyze(analyze_t& arg);
    virtual bool check(void);
    virtual bool remove(void);
    virtual bool list_files(std::list<FileInfo>& files, bool resolve = true);
  };

} // namespace Arc
\end{verbatim}

\section{DataPointIndex}

To create a DataPoint class for an index protocol, the following
methods should be provided:

\begin{verbatim}
namespace Arc {

  class DataPointCustom2 : public DataPointIndex {
   public:
    DataPointCustom2(const URL& url);
    ~DataPointCustom2() {};
    virtual bool meta_resolve(bool source);
    virtual bool meta_preregister(bool replication, bool force = false);
    virtual bool meta_postregister(bool replication);
    virtual bool meta_preunregister(bool replication);
    virtual bool meta_unregister(bool all);
    virtual bool list_files(std::list<FileInfo> &files, bool resolve = true);
  };

} // namespace Arc
\end{verbatim}

\bibliography{grid}
\printindex

\end{document}

### 2008-04-29 doc/hed/ARCHED_article.tex
Extremely skeletal current version of text about HED.

\documentclass[a4paper,10pt]{article}

%opening
\title{The Hosting Environment Daemon of the Advanced Resource Connector middleware}

\author{J. J\"{o}nemo, et al}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

The Hosting Environment Daemon (HED) is the container of all the functional components of the new generation of the Advanced Resource Connector (ARC) middleware on the server side. It is the central part in a new very lightweight incarnation of ARC that is aimed at - but not limited to - providing Web Service.

\section{Architecture}

\subsection{Requirements}
% Perhaps this can be an introductory part of the section without any subsection, we'll see
In the design of the HED, several goals and requirements were considered. These were weighed against each other and the factual context.

The implementation language needed to be object oriented, efficient and provide easy access to system functionality. This eventually lead to the adoption of C++. but languages such as Java and Python were also considered at an early stage.

External dependencies needed to be kept to a minimum while also taking into consideration their ubiquity or relative rarity and license related concerns. Software of this level of complexity must of course depend on many external libraries and components but each such dependenscy has been introduced only after due consideration.

Conservation of resources was an important goal. The present design enables many services sharing both the same process and the same network ports or even port while at the same time exhibiting a remarkably low memory footprint.

%\subsection{Functional design}

%firewalls, modularity, portability, flexibility

\subsection{Technical design}

%How was this implemented, dynamic loading, mcc...
In the technical design it turned out that the endeavours to provide dynamic loading, portability and a well tested high level memory management could all be greatly assisted by introducing glibmm - the C++ interface to the gnome projects library for memory management and related functionality. This enables the daemon process to dynamically load the components it is going to use in a portable way.

Most of the configuration

\subsubsection{MCC}

In the HED all data channels to the outside world are set up by chains of small processing units called Message Chain Components (MCCs). These work on units called messages which represent data going in to or out of the HED. The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general attributes and security attributes. Each MCC typically implements one level in the Internet Protocol suite by transforming a message to an input suitable to propagate to the next component and then performs the corresponding transformation of the response on the way back. The components are all dynamically loaded to provide maximum flexibility and extensibility. Each instance of these MCC's can be individually configuered.

As the data is passed through the individual MCCs, they each populate structures with both general attributes and special security attributes that are available at that particular protocol level.

Each MCC can also be configured to have loadable modules called security handlers attached to it in order to enforce security policies such as authentication and authorization or to assist such activities by gathering specialized scurity information.

\paragraph{TCP MCC}

The TCP MCC in the HED is special in that it produces messages by listening on a socket rather than passing on messages from other MCCs. As such it spawns new threads to handle the message and its response throughout the message chain. One could envision other parallel MCCs having these properties but producing messages from other sources such as e g unix sockets.

This MCC can be configured as to what port to listen to.

\paragraph{TLS MCC}

\paragraph{HTTP MCC}

\paragraph{SOAP MCC}

\subsubsection{Services}

The services are dynamically loaded on start up just like the MCCs. They can even be regarded as special cases of MCCs in that they constitute the last link in the so called Message Chain and are attached in much the same manner as other MCCs. After being loaded, instanciated and configured by the HED, services are provided with a range of

The services are typically expected to receive parsed SOAP messages which they service - possibly by invoking or communicating with external processes.

\subsubsection{Alternative implementation languages}

In order to facilitate e g the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done 
%Not much aboút the actual services, more about what they need to implement and what 
\section{Conclusion}

\end{document}

### 2008-05-07 doc/hed/ARCHED_article.tex
paragrpahs on MCC in general

@@
 %opening
-\title{The Hosting Environment Daemon of the Advanced Resource Connector middleware}
+\title{The Hosting Environment of the Advanced Resource Connector middleware}
 
@@
 
+The HED itself means two things:
+\begin{enumerate}
+\item the daemon (called \texttt{arched}) which hooks up the system and initialize components the way as it is described by the configuration files.
+This configuration describes the components and their relations to each other.
+In optimal cases these single services runs on any node where ARC1 is deployed and started.
+Without loadable components the daemon itself does nothing at all.
+\item sometimes using the HED terms to refer to collection of libraries which is used by service or other component developers.
+These libraries implements some common classes which may simplify the life of service and component developes however only few of these classes are mandatory to use to make the components and services loadable and hookable by the daemon.
+\end{enumerate}
+
 \subsubsection{MCC}
@@
 The components are all dynamically loaded to provide maximum flexibility and extensibility.
-Each instance of these MCC's can be individually configuered.
+Each instance of these MCC's can be individually configured.
 
@@
 
-Each MCC can also be configured to have loadable modules called security handlers attached to it in order to enforce security policies such as authentication and authorization or to assist such activities by gathering specialized scurity information.
+Each MCC can also be configured to have loadable modules called security handlers attached to it in order to enforce security policies such as authentication and authorization or to assist such activities by gathering specialized security information.
+
+The chain formulation using MCCs is up to the administrator.
+The chain is an ordered list the MCCs and their interconection can be described in the configuration file.
+Based on the configuration information the Loader components formulates the actual chain in the memory of the daemon.
+The MCC may implement some routing algorithm which means one MCC may connection to multiple other MCCs.
+Tipical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to SOAP MCC but the messages with GET operation are goes to for example a simple and limited HTTPD service component.
+
+Most of the MCCs has client and server version because the behavior of MCCs should be different depending on wheater it is sitting on server or client part.
+The tipical scenarion here the TCP MCC which should listen and wait for incomming messages on socket on server part but call connect method on client part.
+
+In general every MCC have optimal and natural places in certain chain and this place is not always can be modified.
+For example in the server part the TCP MCC must be the first MCC in any chain where it is used and the TLS MCC should be right after of the TCP MCC.
+
+The developer whoes write an MCC is free to choose any 3rd party library and comonent to implement functionalty of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by libarcloader and libarcmessage libraries of ARC1.
 
@@
 
-The TCP MCC in the HED is special in that it produces messages by listening on a socket rather than passing on messages from other MCCs.
+The server TCP MCC in the HED is special in that it produces messages by listening on a socket rather than passing on messages from other MCCs.
 As such it spawns new threads to handle the message and its response throughout the message chain.
-One could envision other parallel MCCs having these properties but producing messages from other sources such as e g unix sockets.
+One could envision other parallel MCCs having these properties but producing messages from other sources such as e.g.
+unix sockets.
 
@@
 \paragraph{SOAP MCC}
+
+\paragraph{Plexer}
+
+In general case multiple service living in the HED so the incomming message should route to the proper service.
+The Plexer MCC does this job.
+It takes the the ENDPOINT attributes of the message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forward the message to the all matching service.
+It acts as a dispacher.
+The Plexer is special in a sence that it is not a plugin but part of the libarcloader library.
 
@@
 
-In order to facilitate e g the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done 
+In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done 
 %Not much aboút the actual services, more about what they need to implement and what 

### 2008-05-30 doc/sec/security_infrastructure.tex
Skeleton with a bit of meat for security documentation for deliverable. You may ignore .lyx and use .tex file directly.

%% LyX 1.4.5.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{book}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{calc}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

%\documentclass{article}                            %for shorter notes
                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref}
 %for \url tag
\usepackage{color}
                                 %for defining custom colors
\usepackage{framed}
                                %for shaded and framed paragraphs
\usepackage{textcomp}
                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}
                              %for defining page size
\usepackage{longtable}
                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER

\usepackage{babel}
\makeatother
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{tex/ng-logo}} &
%
\parbox[b][0pt]{2cm}{%
\textit{{\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}
}}%
}%
 \tabularnewline
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-XXXXXXX-NN

}

{\raggedleft \today

}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large Security Infrastructure of ARC1}{\Large \par}

} \vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Paper subtitle}{\large \par}

}

\vspace*{1.5cm}
 {\centering {\large Weizhong Qiang}%
\footnote{weizhongqiang@gmail.com%
}{\large , Aleksandr Konstantinov}%
\footnote{aleks@fys.uio.no%
} {\large }{\large \par}

}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

\tableofcontents{}%Comment if use article style
\newpage{}

\chapter{Introduction}

\label{sec:intro}

This document describes Security Infractructure of ARC1 middleware.

\chapter{Security architecture in HED: SecHandlers and PDP}

\chapter{Policy Evaluation engine}

\section{ArcPDP Policy Evaluator}

\chapter{Secuirty Attributes}

\section{Infrastructure}

Security Attributes represent security related information inside

HED framework and store information representing various aspects

needed to perform authorization decison - identity of client, requested

action, targeted resource, constraint policies. Each kind of Security

Attribute is represented by own class inherited from parent SecAttr
class.

Each Security Attribute stores it's information in internal format
and

provides it to external code in one of predefined formats. Currently

only supported format is Arc Policy/Request.

Collectors of Security Attributes instantiate corresponding classes
and

link them to Secuirity Attributes containers - MessageAuth and

MessageAuthContext representing attributes per request and per session

correspondingly. Current implementations of Security Attributes Collectors

are either integratd into existing MCCs or implemented as separate
SecHandler

plugins. See Appendix A for available collectors and corresponding
Security

Attributes.

Note for service developers: Services may implement own authorization

algorithms. But they may use Security Atributes as well by providing

instances of classes inherited from SecAttr and running them through

either configured or hardcoded processors/PDPs (see below).

Processors of Security Attributes are implemented as Policy Decision
Point

components. Currently there are 2 PDP components available:

{*} Arc PDP makes use of Security Attributes containing identities
of client,

resource and requested action. It evaluates either all or selected
set

of attributes against specified Policy documents thus making it possible

to enforce policies defined/selected by service providers.

{*} Delegation PDP is similar to Arc PDP except that it takes it's
Policy

documents directly from Security Attributes. Differently from Arc
PDP it

is meant to be used for enforcing policies defined by client.

Typical service example - TODO

\section{Available collectors}

{*} TCP

Information is collected inside TCP MCC. Security Attribute exports

ARC Request with following attributes

Element AttributeId Content

Resource http://www.nordugrid.org/schemas/policy-arc/types/localendpoint
service\_ip{[}:service\_port]

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/remoteendpoint
client\_ip{[}:client\_port]

{*} TLS

Information is collected inside TLS MCC. Security Attribute exports

ARC Request with following attributes

Element AttributeId Content

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/ca
Subject of signer of first certificate in client's chain

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/chain
Subject of certificate in client's chain - multiple items

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/subject
Subject of last certificate in client's chain

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/identity
Subject of last non-proxy certificate in client's chain

{*} HTTP

Information is collected inside TLS MCC. Security Attribute exports

ARC Request with following attributes

Element AttributeId Content

Resource http://www.nordugrid.org/schemas/policy-arc/types/http/path
HTTP path without host and port part

Action http://www.nordugrid.org/schemas/policy-arc/types/http/method
HTTP method

{*} SOAP

Information is collected inside TLS MCC. Security Attribute exports

ARC Request with following attributes

Element AttributeId Content

Resource http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint
WS-Addr To element

Action http://www.nordugrid.org/schemas/policy-arc/types/soap/operation
SOAP top level element name without namespace prefix

Context http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace
Namespace of SOAP top level element

{*} Delegation

This Security Attribute is collected by dedicated Security Handler
plugin.

It extracts policy document stored inside X509 certificate proxy extension

as defined in RFC3820. All certificates in a chain provided by client
are

examined. Proxy certificates with id-ppl-inheritAll property are passed

through and no policy is generated for them. From those with id-ppl-anyLanguage

string content of ProxyPolicy element is extracted and converted into
XML

document. Then document is check to be of Arc Policy kind. If policy
is

not recognized as Arc Policy procedure fails and that causes failure
of

comunication. Proxies with other type of policies including id-ppl-independent

are not accepted and generate immediate failure.

\chapter{Delegation restrictions}

\section{Delegation architecture}

\section{Delegation PDP}

\chapter*{Schemas}

\section*{Policy}

\section*{Request}

\chapter*{Examples:}

\section*{Policy}

\section*{Request}

\section*{Delegation Policy}

\chapter*{Template examples}

In expressions, the following operands are allowed: \begin{shaded}
\verb#=   !=   >   <   >=   <=# \end{shaded}

\begin{framed} Examples of URLs are:\\
 \\
 \verb#http://grid.domain.org/dir/script.sh#\\
 \verb#gsiftp://grid.domain.org:2811;threads=10/dir/input_12378.dat#\\
 \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
 \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#
\verb#file:///home/auser/griddir/steer.cra#\\
 \end{framed}

%\subsection{Subsection}
%\label{sec:subsection}

%
\begin{figure}[ht]
 \centering{{\scalebox{0.9}{\includegraphics{tex/ng-logo}}}

\caption{\label{fig:myfigure1}The figure shows a logo.}

} 
\end{figure}

\bibliography{grid}

\end{document}

### 2008-05-30 doc/sec/security_infrastructure.tex
Nice formating

@@
 
-{\raggedleft NORDUGRID-XXXXXXX-NN
+{\raggedleft NORDUGRID-TECH-16
 
@@
 
-\chapter{Introduction}
-
-\label{sec:intro}
+\chapter{Introduction\label{sec:introduction}}
 
@@
 
-\chapter{Security architecture in HED: SecHandlers and PDP}
+\chapter{Security architecture in HED: SecHandlers and PDP\label{sec:architecture}}
 
-\chapter{Policy Evaluation engine}
+\chapter{Policy Evaluation engine\label{sec:policy_engine}}
 
-\section{ArcPDP Policy Evaluator}
+\section{ArcPDP Policy Evaluator\label{sec:policy_evaluator}}
 
-\chapter{Secuirty Attributes}
+\chapter{Secuirty Attributes\label{sec:sec_attr}}
 
-\section{Infrastructure}
+\section{Infrastructure\label{sec:sec_attr_infrastructure}}
 
 Security Attributes represent security related information inside
+HED framework and store information representing various aspects needed
+to perform authorization decison - identity of client, requested action,
+targeted resource, constraint policies.
 
-HED framework and store information representing various aspects
-
-needed to perform authorization decison - identity of client, requested
-
-action, targeted resource, constraint policies. Each kind of Security
-
-Attribute is represented by own class inherited from parent SecAttr
-class.
-
-Each Security Attribute stores it's information in internal format
-and
-
-provides it to external code in one of predefined formats. Currently
-
-only supported format is Arc Policy/Request.
+Each kind of Security Attribute is represented by own class inherited
+from parent SecAttr class <arc/message/SecAttr.h>. Each Security Attribute
+stores it's information in internal format and is capable to export
+it to one of predefined formats using Export() method. Currently only
+supported format is Arc Policy/Request XML document.
 
 Collectors of Security Attributes instantiate corresponding classes
-and
+and link them to Secuirity Attributes containers - MessageAuth <arc/message/MessageAuth.h>
+and MessageAuthContext <arc/message/Message.h> storing collected attributes
+per request and per session correspondingly. Current implementations
+of Security Attributes Collectors are either integrated into existing
+MCCs or implemented as separate SecHandler plugins. See Section \ref{sec:sec_attr_collectors}
+for available Collectors and corresponding Security Attributes.
 
-link them to Secuirity Attributes containers - MessageAuth and
-
-MessageAuthContext representing attributes per request and per session
-
-correspondingly. Current implementations of Security Attributes Collectors
-
-are either integratd into existing MCCs or implemented as separate
-SecHandler
-
-plugins. See Appendix A for available collectors and corresponding
-Security
-
-Attributes.
-
-Note for service developers: Services may implement own authorization
-
-algorithms. But they may use Security Atributes as well by providing
-
-instances of classes inherited from SecAttr and running them through
-
-either configured or hardcoded processors/PDPs (see below).
+\begin{framed}Note for service developers: Services may implement
+own authorization algorithms. But they may use Security Atributes
+as well by providing instances of classes inherited from SecAttr and
+running them through either configured or hardcoded processors/PDPs.\end{framed}
 
 Processors of Security Attributes are implemented as Policy Decision
-Point
+Point components. Currently there are 2 PDP components available:
 
-components. Currently there are 2 PDP components available:
+\begin{itemize}
+\item ArcPDP makes use of Security Attributes containing identities of client,
+resource and requested action. It evaluates either all or selected
+set of attributes against specified Policy documents thus making it
+possible to enforce policies defined/selected by service providers. 
+\item (to move to Delegation section) DelegationPDP is similar to ArcPDP
+except that it takes it's Policy documents directly from Security
+Attributes. Differently from Arc PDP it is meant to be used for enforcing
+policies defined by client.
+\end{itemize}
 
-{*} Arc PDP makes use of Security Attributes containing identities
-of client,
+\section{Available collectors\label{sec:sec_attr_collectors}}
 
-resource and requested action. It evaluates either all or selected
-set
-
-of attributes against specified Policy documents thus making it possible
-
-to enforce policies defined/selected by service providers.
-
-{*} Delegation PDP is similar to Arc PDP except that it takes it's
-Policy
-
-documents directly from Security Attributes. Differently from Arc
-PDP it
-
-is meant to be used for enforcing policies defined by client.
-
-Typical service example - TODO
-
-\section{Available collectors}
-
-{*} TCP
+\subsection{TCP}
 
@@
 
-{*} TLS
+\subsection{TLS}
+
+Information is collected inside TLS MCC. Generated Security Attribute
+class exports ARC Request with following attributes:
+
+\begin{tabular}{|c|c|c|}
+\hline 
+\multicolumn{1}{|c|}{Element}&
+AttributeId&
+Content\tabularnewline
+\hline
+\hline 
+SubjectAttribute&
+http://www.nordugrid.org/schemas/policy-arc/types/tls/ca&
+Subject of signer of first certificate in client's chain\tabularnewline
+\hline 
+SubjectAttribute&
+http://www.nordugrid.org/schemas/policy-arc/types/tls/chain&
+Subject of certificate in client's chain - multiple items\tabularnewline
+\hline 
+SubjectAttribute&
+http://www.nordugrid.org/schemas/policy-arc/types/tls/subject&
+Subject of last certificate in client's chain\tabularnewline
+\hline 
+SubjectAttribute&
+http://www.nordugrid.org/schemas/policy-arc/types/tls/identity&
+Subject of last non-proxy certificate in client's chain\tabularnewline
+\hline
+\end{tabular}
+
+\subsection{HTTP}
 
 Information is collected inside TLS MCC. Security Attribute exports
+ARC Request with following attributes:
 
+\begin{tabular}{|c|c|c|}
+\hline 
+Element&
+AttributeId&
+Content\tabularnewline
+\hline
+\hline 
+Resource&
+http://www.nordugrid.org/schemas/policy-arc/types/http/path&
+HTTP path without host and port part\tabularnewline
+\hline 
+Action&
+http://www.nordugrid.org/schemas/policy-arc/types/http/method&
+HTTP method\tabularnewline
+\hline
+\end{tabular}
+
+\subsection{SOAP}
+
+Information is collected inside TLS MCC. Security Attribute exports
 ARC Request with following attributes
 
-Element AttributeId Content
+\begin{tabular}{|c|c|c|}
+\hline 
+Element&
+AttributeId&
+Content\tabularnewline
+\hline
+\hline 
+Resource&
+http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint&
+To element of WS-Addressing structure\tabularnewline
+\hline 
+Action&
+http://www.nordugrid.org/schemas/policy-arc/types/soap/operation&
+SOAP top level element name without namespace prefix\tabularnewline
+\hline 
+Context&
+http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace&
+Namespace of SOAP top level element\tabularnewline
+\hline
+\end{tabular}
 
-SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/ca
-Subject of signer of first certificate in client's chain
+\subsection{Delegation}
 
-SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/chain
-Subject of certificate in client's chain - multiple items
+(to move to Delegation section) This Security Attribute is collected
+by dedicated Security Handler plugin named {}``\emph{delegation.pdp}''
+avaialble in \emph{arcpdc} plugins' library. It extracts policy document
+stored inside X509 certificate proxy extension as defined in RFC3820.
+All proxy certificates in a chain provided by client are examined
+and all available policies are extracted.
 
-SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/subject
-Subject of last certificate in client's chain
+Policies are stored in string content of \emph{ProxyPolicy} element
+with \emph{id-ppl-anyLanguage}. Extracted content is converted into
+XML document. Then document is checked to be of Arc Policy kind. If
+policy is not recognized as Arc Policy procedure fails and that causes
+failure of comunication.
 
-SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/identity
-Subject of last non-proxy certificate in client's chain
+Proxy certificates with \emph{id-ppl-inheritAll} property are passed
+through and no policy document is generated for them. Proxies with
+other type of policies including \emph{id-ppl-independent} are not
+accepted and generate immediate failure.
 
-{*} HTTP
+\chapter{Delegation restrictions\label{sec:delegation}}
 
-Information is collected inside TLS MCC. Security Attribute exports
+\section{Delegation architecture\label{sec:delegation_architectire}}
 
-ARC Request with following attributes
+\section{Delegation PDP\label{sec:delegation_pdp}}
 
-Element AttributeId Content
-
-Resource http://www.nordugrid.org/schemas/policy-arc/types/http/path
-HTTP path without host and port part
-
-Action http://www.nordugrid.org/schemas/policy-arc/types/http/method
-HTTP method
-
-{*} SOAP
-
-Information is collected inside TLS MCC. Security Attribute exports
-
-ARC Request with following attributes
-
-Element AttributeId Content
-
-Resource http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint
-WS-Addr To element
-
-Action http://www.nordugrid.org/schemas/policy-arc/types/soap/operation
-SOAP top level element name without namespace prefix
-
-Context http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace
-Namespace of SOAP top level element
-
-{*} Delegation
-
-This Security Attribute is collected by dedicated Security Handler
-plugin.
-
-It extracts policy document stored inside X509 certificate proxy extension
-
-as defined in RFC3820. All certificates in a chain provided by client
-are
-
-examined. Proxy certificates with id-ppl-inheritAll property are passed
-
-through and no policy is generated for them. From those with id-ppl-anyLanguage
-
-string content of ProxyPolicy element is extracted and converted into
-XML
-
-document. Then document is check to be of Arc Policy kind. If policy
-is
-
-not recognized as Arc Policy procedure fails and that causes failure
-of
-
-comunication. Proxies with other type of policies including id-ppl-independent
-
-are not accepted and generate immediate failure.
-
-\chapter{Delegation restrictions}
-
-\section{Delegation architecture}
-
-\section{Delegation PDP}
-
-\chapter*{Schemas}
+\chapter*{Schemas and comments\label{sec:schemas}}
 
@@
 \section*{Request}
+
+\section*{ArcPDP configuration}
+
+\section*{DelegationPDP configuration}
 

### 2008-05-30 doc/sec/security_infrastructure.tex
Delegation part alsmost done. Added few schemas and description.

@@
 \setcounter{tocdepth}{3}
+\usepackage{array}
 \usepackage{calc}
@@
 and MessageAuthContext <arc/message/Message.h> storing collected attributes
-per request and per session correspondingly. Current implementations
-of Security Attributes Collectors are either integrated into existing
-MCCs or implemented as separate SecHandler plugins. See Section \ref{sec:sec_attr_collectors}
-for available Collectors and corresponding Security Attributes.
+per request and per session correspondingly. Each attribute is assigned
+a name. Current implementations of Security Attributes Collectors
+are either integrated into existing MCCs or implemented as separate
+SecHandler plugins. See Section \ref{sec:sec_attr_collectors} for
+available Collectors and corresponding Security Attributes.
 
@@
 possible to enforce policies defined/selected by service providers. 
-\item (to move to Delegation section) DelegationPDP is similar to ArcPDP
-except that it takes it's Policy documents directly from Security
-Attributes. Differently from Arc PDP it is meant to be used for enforcing
-policies defined by client.
+\item DelegationPDP is described below in Section \ref{sec:delegation_pdp}.
 \end{itemize}
@@
 
+Here Security Attribute collectors distributed as part of ARC1 are
+described except those used for Delegation Restrictions. Those are
+described in Section \ref{sec:delegation}.
+
 \subsection{TCP}
 
-Information is collected inside TCP MCC. Security Attribute exports
+Information is collected inside TCP MCC. Security Attribute is stored
+under name 'TCP' and exports ARC Request with following attributes:
 
-ARC Request with following attributes
-
-Element AttributeId Content
-
-Resource http://www.nordugrid.org/schemas/policy-arc/types/localendpoint
-service\_ip{[}:service\_port]
-
-SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/remoteendpoint
-client\_ip{[}:client\_port]
+\begin{tabular}{|>{\centering}p{0.2\textwidth}|c|>{\centering}p{0.2\textwidth}|}
+\hline 
+Element&
+AttributeId&
+Content\tabularnewline
+\hline
+\hline 
+Resource&
+http://www.nordugrid.org/schemas/policy-arc/types/localendpoint&
+service\_ip{[}:service\_port] \tabularnewline
+\hline 
+SubjectAttribute&
+http://www.nordugrid.org/schemas/policy-arc/types/remoteendpoint&
+client\_ip{[}:client\_port] \tabularnewline
+\hline
+\end{tabular}
 
@@
 
-\begin{tabular}{|c|c|c|}
+\begin{tabular}{|>{\centering}p{0.2\textwidth}|c|>{\centering}p{0.2\textwidth}|}
 \hline 
-\multicolumn{1}{|c|}{Element}&
+Element&
 AttributeId&
@@
 
-Information is collected inside TLS MCC. Security Attribute exports
-ARC Request with following attributes:
+Information is collected inside TLS MCC. Security Attribute is stored
+under name 'HTTP' and exports ARC Request with following attributes:
 
-\begin{tabular}{|c|c|c|}
+\begin{tabular}{|>{\centering}p{0.2\textwidth}|c|>{\centering}p{0.2\textwidth}|}
 \hline 
@@
 
-Information is collected inside TLS MCC. Security Attribute exports
-ARC Request with following attributes
+Information is collected inside TLS MCC. Security Attribute is stored
+under name 'SOAP' and exports ARC Request with following attributes:
 
-\begin{tabular}{|c|c|c|}
+\begin{tabular}{|>{\centering}p{0.2\textwidth}|c|>{\centering}p{0.2\textwidth}|}
 \hline 
@@
 
-\subsection{Delegation}
+\chapter{Delegation Restrictions\label{sec:delegation}}
 
-(to move to Delegation section) This Security Attribute is collected
-by dedicated Security Handler plugin named {}``\emph{delegation.pdp}''
-avaialble in \emph{arcpdc} plugins' library. It extracts policy document
-stored inside X509 certificate proxy extension as defined in RFC3820.
+\section{Delegation Architecture\label{sec:delegation_architectire}}
+
+In current implementation delegation is achieved through Identity
+Delegation implemented using X509 Proxy Certificates as defined in
+RFC 3820. Client wishing to allow service to act on it's behalf provides
+Proxy Certificate to the service using Web Service based Delegation
+interface described in \ref{sec:delegation_interface}.
+
+For limitinng the scope of delegated credentials along with usually
+used time constraints it is possible to attach Policy document to
+Proxy Certificate. According to RFC 3820 Policy is stored in \emph{ProxyPolicy}
+extension. In order not to introduce new type of object Policy is
+assigned \emph{id-ppl-anyLanguage} identifier. RFC 3820 allows any
+octet string associated with such object. We are using textual representation
+of ARC Policy XML document.
+
+Each deployment implementing Delegation Restrictions must use dedicated
+Security Handler plugin (see section \ref{sec:delegation_collector})
+to collect all Policy documents from Proxy Certificates used for establishing
+secure connection. Then those documents must be processed by dedicated
+Policy Decision Point plugin (see section \ref{sec:delegation_pdp})
+to make a final decision based on collected Policies and various information
+about client's identity and requested operation. Service or MCC chain
+supporting Delegation Restrictions must accept negative decision of
+this PDP as final and do not override it with any other decision based
+on other policies.
+
+\section{Delegation Collector\label{sec:delegation_collector}}
+
+This Security Attribute is collected by dedicated Security Handler
+plugin named {}``\emph{delegation.pdp}'' avaialble in \emph{arcpdc}
+plugins' library. It extracts policy document stored inside X509 certificate
+proxy extension as defined in RFC3820 and described in Section \ref{sec:delegation_architectire}.
 All proxy certificates in a chain provided by client are examined
@@
 
-Policies are stored in string content of \emph{ProxyPolicy} element
-with \emph{id-ppl-anyLanguage}. Extracted content is converted into
-XML document. Then document is checked to be of Arc Policy kind. If
-policy is not recognized as Arc Policy procedure fails and that causes
-failure of comunication.
+Extracted content is converted into XML document. Then document is
+checked to be of Arc Policy kind. If policy is not recognized as Arc
+Policy procedure fails and that causes failure of comunication.
 
@@
 
-\chapter{Delegation restrictions\label{sec:delegation}}
-
-\section{Delegation architecture\label{sec:delegation_architectire}}
-
 \section{Delegation PDP\label{sec:delegation_pdp}}
 
-\chapter*{Schemas and comments\label{sec:schemas}}
+DelegationPDP is similar to ArcPDP described above except that it
+takes it's Policy documents directly from Security Attributes. Differently
+from Arc PDP it is meant to be used for enforcing policies defined
+by client.
 
-\section*{Policy}
+\chapter{Schemas and descriptions\label{sec:schemas}}
 
-\section*{Request}
+\section{Policy}
 
-\section*{ArcPDP configuration}
+\section{Request}
 
-\section*{DelegationPDP configuration}
+\section{ArcPDP configuration}
+
+\begin{verbatim}
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns="http://www.nordugrid.org/schemas/ArcPDP"
+  targetNamespace="http://www.nordugrid.org/schemas/ArcPDP"
+  elementFormDefault="qualified">
+
+    <xsd:complexType name="FilterType">
+        <!-- This element defines Security Attributes to select and reject.
+            If there are no Select elements all Attributes are used except
+            those listed in Reject elements.  -->
+        <xsd:sequence>
+            <xsd:element name="Select" type="xsd:string" minOccurs="0" maxOccurs="unbounded"/>
+            <xsd:element name="Reject" type="xsd:string" minOccurs="0" maxOccurs="unbounded"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="Filter" type="FilterType"/>
+
+    <!--
+        This element specified file containing policy document.
+        There may be multiple such elements.
+    -->
+    <xsd:element name="PolicyStore" type="xsd:string"/>
+
+</xsd:schema>
+\end{verbatim}
+
+\section{DelegationPDP configuration}
+
+\begin{verbatim}
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns="http://www.nordugrid.org/schemas/DelegationPDP"
+  targetNamespace="http://www.nordugrid.org/schemas/DelegationPDP"
+  elementFormDefault="qualified">
+
+    <xsd:complexType name="FilterType">
+        <!-- This element defines Security Attributes to select and reject.
+            If there are no Select elements all Attributes are used except
+            those listed in Reject elements.  -->
+        <xsd:sequence>
+            <xsd:element name="Select" type="xsd:string" minOccurs="0" maxOccurs="unbounded"/>
+            <xsd:element name="Reject" type="xsd:string" minOccurs="0" maxOccurs="unbounded"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="Filter" type="FilterType"/>
+
+</xsd:schema>
+\end{verbatim}
+
+\section{Delegation interface\label{sec:delegation_interface}}
+
+Web Service delegation interface. Each ARC1 service wishing to accept
+delegated credentials implements this interface. Here is how delegation
+procedure works:
+
+\begin{itemize}
+\item Step 1
+
+\begin{itemize}
+\item Client contacts service requesting operation DelegateCredentialsInit.
+This operation has no arguments.
+\item Service responds with \emph{DelegateCredentialsInitResponse} message
+with element \emph{TokenRequest}. That element contains credentials
+request generated by service in \emph{Value}. Type of request is defined
+by attribute \emph{Format}. Currently only supported format is \emph{x509}.
+Along with \emph{Value} service provides identifier \emph{Id} which
+is used in second step.
+\end{itemize}
+\item Step 2
+
+\begin{itemize}
+\item Client requests \emph{UpdateCredentials} operation with \emph{DelegatedToken}
+argument. This element contains \emph{Value} with serialized delegated
+credentials and \emph{Id} which links it to first step. Delegated
+token element may also contain multiple \emph{Reference} elements.
+\emph{Reference} refers to the object which these credentials should
+be applied to in a way specific to the service. \\
+The \emph{DelegatedToken} element must also be used for delegating
+credentials when Step 2 is combined with other operations on service.
+\item Service responds with empty \emph{UpdateCredentialsResponse} message.
+\end{itemize}
+\end{itemize}
+\begin{verbatim}<?xml version="1.0" encoding="UTF-8"?>
+<wsdl:definitions targetNamespace="http://www.nordugrid.org/schemas/delegation"
+ xmlns:SOAP-ENV="http://schemas.xmlsoap.org/soap/envelope/"
+ xmlns:SOAP-ENC="http://schemas.xmlsoap.org/soap/encoding/"
+ xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"
+ xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+ xmlns:soap="http://schemas.xmlsoap.org/wsdl/soap/"
+ xmlns:wsdl="http://schemas.xmlsoap.org/wsdl/"
+ xmlns:wsa="http://www.w3.org/2005/08/addressing"
+ xmlns:deleg="http://www.nordugrid.org/schemas/delegation">
+
+  <wsdl:types>
+    <xsd:schema targetNamespace="http://www.nordugrid.org/schemas/delegation">
+
+      <!-- Common types -->
+
+      <xsd:simpleType name="TokenFormatType">
+        <xsd:restriction base="xsd:string">
+          <xsd:enumeration value="x509"/>
+        </xsd:restriction>
+      </xsd:simpleType>
+
+      <xsd:complexType name="ReferenceType">
+        <xsd:sequence>
+          <xsd:any namespace="##other" processContents="lax" minOccurs="0" maxOccurs="unbounded"/>
+        </xsd:sequence>
+      </xsd:complexType>
+
+      <xsd:complexType name="DelegatedTokenType">
+        <xsd:sequence>
+          <xsd:element name="Id" type="xsd:string"/>
+          <xsd:element name="Value" type="xsd:string"/>
+          <xsd:element name="Reference" type="deleg:ReferenceType" minOccurs="0" maxOccurs="unbounded"/>
+        </xsd:sequence>
+        <xsd:attribute name="Format" type="deleg:TokenFormatType" use="required"/>
+      </xsd:complexType>
+      <xsd:element name="DelegatedToken" type="deleg:DelegatedTokenType"/>
+
+      <xsd:complexType name="TokenRequestType">
+        <xsd:sequence>
+          <xsd:element name="Id" type="xsd:string"/>
+          <xsd:element name="Value" type="xsd:string"/>
+        </xsd:sequence>
+        <xsd:attribute name="Format" type="deleg:TokenFormatType" use="required"/>
+      </xsd:complexType>
+      <xsd:element name="TokenRequest" type="deleg:TokenRequestType"/>
+
+      <!-- Types for messages -->
+
+      <xsd:complexType name="DelegateCredentialsInitRequestType">
+        <xsd:sequence>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="DelegateCredentialsInit" type="deleg:DelegateCredentialsInitRequestType"/>
+
+      <xsd:complexType name="DelegateCredentialsInitResponseType">
+        <xsd:sequence>
+          <xsd:element name="TokenRequest" type="deleg:TokenRequestType"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="DelegateCredentialsInitResponse" type="deleg:DelegateCredentialsInitResponseType"/>
+
+      <xsd:complexType name="UpdateCredentialsRequestType">
+        <xsd:sequence>
+          <xsd:element name="DelegatedToken" type="deleg:DelegatedTokenType"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="UpdateCredentials" type="deleg:UpdateCredentialsRequestType"/>
+
+      <xsd:complexType name="UpdateCredentialsResponseType">
+        <xsd:sequence>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="UpdateCredentialsResponse" type="deleg:UpdateCredentialsResponseType"/>
+
+      <!-- Faults -->
+
+      <xsd:complexType name="UnsupportedFaultType">
+        <xsd:sequence>
+          <xsd:element name="Description" type="xsd:string" minOccurs="0"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="UnsupportedFault" type="deleg:UnsupportedFaultType"/>
+
+      <xsd:complexType name="ProcessingFaultType">
+        <xsd:sequence>
+          <xsd:element name="Description" type="xsd:string" minOccurs="0"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="ProcessingFault" type="deleg:ProcessingFaultType"/>
+
+      <xsd:complexType name="WrongReferenceFaultType">
+        <xsd:sequence>
+          <xsd:element name="Description" type="xsd:string" minOccurs="0"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="WrongReferenceFault" type="deleg:WrongReferenceFaultType"/>
+
+    </xsd:schema>
+  </wsdl:types>
+
+  <wsdl:message name="DelegateCredentialsInitRequest">
+    <wsdl:part name="DelegateCredentialsInitRequest" element="deleg:DelegateCredentialsInit"/>
+  </wsdl:message>
+
+  <wsdl:message name="DelegateCredentialsInitResponse">
+    <wsdl:part name="DelegateCredentialsInitResponse" element="deleg:DelegateCredentialsInitResponse"/>
+  </wsdl:message>
+
+  <wsdl:message name="UpdateCredentialsRequest">
+    <wsdl:part name="UpdateCredentialsRequest" element="deleg:UpdateCredentials"/>
+  </wsdl:message>
+
+  <wsdl:message name="UpdateCredentialsResponse">
+    <wsdl:part name="UpdateCredentialsResponse" element="deleg:UpdateCredentialsResponse"/>
+  </wsdl:message>
+
+  <wsdl:message name="UnsupportedFault">
+    <wsdl:part name="Detail" element="deleg:UnsupportedFault"/>
+  </wsdl:message>
+
+  <wsdl:message name="ProcessingFault">
+    <wsdl:part name="Detail" element="deleg:ProcessingFault"/>
+  </wsdl:message>
+
+  <wsdl:message name="WrongReferenceFault">
+    <wsdl:part name="Detail" element="deleg:WrongReferenceFault"/>
+  </wsdl:message>
+
+  <wsdl:portType name="DelegationPortType">
+
+   <wsdl:operation name="DelegateCredentialsInit">
+      <wsdl:documentation>
+      </wsdl:documentation>
+      <wsdl:input name="DelegateCredentialsInitRequest"
+        message="deleg:DelegateCredentialsInitRequest"/>
+      <wsdl:output name="DelegateCredentialsInitResponse"
+        message="deleg:DelegateCredentialsInitResponse"/>
+      <wsdl:fault name="UnsupportedFault"
+        message="deleg:UnsupportedFault"/>
+      <wsdl:fault name="ProcessingFault"
+        message="deleg:ProcessingFault"/>
+    </wsdl:operation>
+
+   <wsdl:operation name="UpdateCredentials">
+      <wsdl:documentation>
+      </wsdl:documentation>
+      <wsdl:input name="UpdateCredentialsRequest"
+        message="deleg:UpdateCredentialsRequest"/>
+      <wsdl:output name="UpdateCredentialsResponse"
+        message="deleg:UpdateCredentialsResponse"/>
+      <wsdl:fault name="UnsupportedFault"
+        message="deleg:UnsupportedFault"/>
+      <wsdl:fault name="ProcessingFault"
+        message="deleg:ProcessingFault"/>
+      <wsdl:fault name="WrongReferenceFault"
+        message="deleg:WrongReferenceFault"/>
+    </wsdl:operation>
+
+  </wsdl:portType>
+
+  <wsdl:binding name="DelegationBinding" type="deleg:DelegationPortType">
+
+    <soap:binding style="document" transport="http://schemas.xmlsoap.org/soap/http"/>
+
+    <wsdl:operation name="DelegateCredentialsInit">
+      <soap:operation soapAction="DelegateCredentialsInit"/>
+      <wsdl:input name="DelegateCredentialsInitRequest">
+         <soap:body use="literal"/>
+      </wsdl:input>
+      <wsdl:output name="DelegateCredentialsInitResponse">
+         <soap:body use="literal"/>
+      </wsdl:output>
+    </wsdl:operation>
+
+    <wsdl:operation name="UpdateCredentials">
+      <soap:operation soapAction="UpdateCredentials"/>
+      <wsdl:input name="UpdateCredentialsRequest">
+        <soap:body use="literal"/>
+      </wsdl:input>
+      <wsdl:output name="UpdateCredentialsResponse">
+        <soap:body use="literal"/>
+      </wsdl:output>
+    </wsdl:operation>
+
+  </wsdl:binding>
+
+</wsdl:definitions>
+\end{verbatim}
 

### 2008-07-22 doc/storage_system/arc1-storage-documentation.tex
evolving technical documentation for the ARC1 storage

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {Documentation of the ARC1 storage system},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

% \begin{titlepage}
% 
% \begin{tabular}{rl}
% \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
% &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
% \end{tabular}
% 
% \hrulefill
% 
% %-------- Change this to NORDUGRID-XXXXXXX-NN
% 
% {\raggedleft NORDUGRID-XXXXXXX-NN\par}
% 
% {\raggedleft \today\par}
% 
% \vspace*{2cm}
% 
% %%%%---- The title ----
% {\centering \textsc{\Large Documentation of the ARC1 storage system}\Large \par}
% \vspace*{0.5cm}
%     
% %%%%---- A subtitle, if necessary ----
% {\centering \textit{\large First prototype status and plans}\large \par}
%     
% \vspace*{1.5cm}
% %%%%---- A list of authors ----
%     {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
%     
% \end{titlepage}
% 
% \tableofcontents                          %Comment if use article style
% \newpage
\chapter{Design Overview}
\label{sec:overview}

% NorduGrid~\cite{nordugrid} Advanced Resource Connector (ARC)
% \textit{middleware}\index{middleware} uses \GT~2 by Globus.
% 
% In expressions, the following operands are allowed:
% \begin{shaded}
%   \verb#=   !=   >   <   >=   <=#
% \end{shaded}
% 
% \begin{framed}
%   Examples of URLs are:\\
%   \\
%   \verb#http://grid.domain.org/dir/script.sh#\\
%   \verb#gsiftp://grid.domain.org:2811;threads=10/dir/input_12378.dat#\\
%   \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
%   \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#
%   \verb#file:///home/auser/griddir/steer.cra#\\
% \end{framed}

The ARC1 storage system is a distributed system for storing replicated \emph{files} on several file storage nodes and manage them in a global namespace. The files can be grouped into collections, and a \emph{collection} can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Names} (LN). The users can use this global namespace  as they were using a local filesystem. Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node should have one or more data transfer service running (e.g. HTTP(S), FTP(S), GridFTP, ByteIO, etc.), and one of the ARC1 storage system’s services is needed to manage each storage node and integrating it into the system. And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system. The main services are shown on Figure~\ref{fig:services}.

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{arc1-storage-services.pdf}}}
\caption{\label{fig:services}The components of the ARC1 storage;
A-H: the \textbf{A-Hash}, which is a currently centralized, later distributed database;
L: the \textbf{Librarian}, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
B: the \textbf{Bartender}, which provides a high-level interface for the users and for other services;
S: the \textbf{Shepherd}, which provides a simple interface for storing files on storage nodes.
} }
\end{figure}

\section{IDs used in the system}
\label{sec:IDs}

There are a number of IDs used in the ARC storage system, such as:
\begin{itemize}
    \item Each file and collection has a globally unique ID called \textbf{GUID}.
    \item The files and collections are organized into a global hierarchical namespace and can be referred to using paths of this namespace called \textbf{Logical Names (LN)}.
    \item Each service has a unique \textbf{serviceID} which can be used to get an endpoint reference from the information system. We need an endpoint reference which is an address (URL) which we could connect to.
    \item The Shepherds in the system identify their files using a \textbf{referenceID}.
    \item The \textbf{location} of a replica consists of two IDs: the ID of the Shepherd and the ID of the file within the Shepherd: (serviceID, referenceID).
\end{itemize}

% section IDs (end)

\section{GUIDs and Logical Names (LN)} % (fold)
\label{sec:GUIDsLNs}

The syntax of Logical Names (LN): \verb#/[path]# or \verb#<GUID>[/<path>]# where [...] indicates optional parts.

Each file and collection has a GUID which is globally unique, so they can be unambiguously referred using this GUID, that's why a single GUID is a Logical Name itself.

In a collection each entry has a name, and this entry can be a sub-collection, in which there are files and sub-sub-collections, etc.

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-namespace.pdf}}}
\caption{\label{fig:namespace}Example of the global namespace's hierarchy} }
\end{figure}

Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file as \verb#/green/orange/huge#, if we know the GUID of the starting collection, so let's prefix the path with it: \verb#1234/green/orange/huge#. This is the Logical Name of that file. If there is a well-known system-wide root collection (its GUID could be e.g. \verb#0#), then if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g. \verb#/why/blue# means \verb#0/why/blue#.

If a client wants to find the file called \verb#/why/blue#, the system knows where to start the search: the GUID of the root collection. The root collection knows the GUID of \verb#why#, and the (sub-)collection \verb#why# knows the GUID of \verb#blue#. If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, so it's a hard link. To count the references (the number of hardlinks) of a file or collection, the GUID of the parent collection(s) is stored as a metadata for each file and collection.

Each VO should create a VO-wide root collection, and put it in the generic root collection, e.g. if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#. Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g. \verb#/vo1/file1#, etc. Or sub-collections, e.g. \verb#/vo1/col1#, \verb#/vo1/col2/file3#, etc. For this the VO does not need the install any service. These files and collections can be created using a Bartender service.

% section GUIDsLNs (end)

\section{The Bartenders} % (fold)
\label{sec:the_bartenders}
Clients can access the storage system through a Bartender service. If a client wants to create a collection, upload or download a file, the first step is to connect a Bartender. The Bartender then asks a Librarian to resolve Logical Names and return metadata, then initiates file transfers on some storage node by asking the Shepherd service on the node, then returns the transfer URL to the client, which allows the client to directly transfer from/to the storage node. So the data transfer itself is not going through the Bartender, it is performed over a direct link between a storage node and the client.

There could be any number of independent Bartender services in the system which provides high-availability and load-balancing.

% section the_bartenders (end)

\section{The Librarians} % (fold)
\label{sec:the_librarians}
The Librarian is capable of managing the hierarchy and metadata of files and collections, and health information of the Shepherd services. Each file and collection in the Librarian has a globally unique ID (GUID). A collection contains files and other collections, and each of these entries has a name unique within the collection very much like entries in a usual directory on a local filesystem. Besides files and collections the Librarian stores a third type of entries called Mount Points which are references to external services creating the capability to mount the namespace of third-party storages to our global namespace and make the files on a third-party storage available through the interface of the ARC storage system.

The Librarian also manages information about registered Shepherd services which are associated with a storage node, and receives heartbeat messages from them and change replica states automatically if needed.

The Librarian uses the A-Hash as database, that’s why there could be any number of independent Librarian services (all using the same A-Hash) which provides high-availability and load-balancing.

% section the_librarians (end)

\section{The A-Hash} % (fold)
\label{sec:the_a_hash}
The A-Hash is a currently centralized, but later distributed service capable of consistently storing objects containing property-value pairs organized in sections. All metadata about files and collections are stored in the A-Hash, and some information about Shepherd services is stored in it as well.

% section the_a_hash (end)

\section{The Shepherds} % (fold)
\label{sec:the_shepherds}
When a new file is put into the system the number of needed replicas is given for the file. The file replicas are stored on different storage nodes, for each storage node there is Shepherd service which provides the interface for initiating file transfer.

The file-naming used by a Shepherd has nothing to do with the the hierarchy of collections, or Logical Names. When a replica is given to a Shepherd it gets an ID which refers to it within that Shepherd. Each Shepherd has a unique ID itself, so with these IDs the replica can be unambiguously referenced, this is called a Location. The namespace of these Locations has nothing to do with the namespace of GUIDs or the namespace of Logical Names. It consists of two IDs: the ID of the Shepherd and the ID of the file within the Shepherd: \verb!(serviceID, referenceID)!

% section the_shepherds (end)

\section{Heartbeats and replication} % (fold)
\label{sec:heartbeats_and_replication}
Each Shepherd should periodically send heartbeats to a Librarian service with information about replicas whose state changed since the last heartbeat, the Librarian stores these file lists (which contains the GUIDs of the files as well), and if it doesn’t receive a heartbeat for a Shepherd in a given time, it invalidates all the replicas the Shepherd stores. This invalidating means that the state of that location will be \verb!offline!.

If a Shepherd finds out that a file is missing or has a bad checksum, it reports that the file is \verb#invalid# to the Librarian immediately, and the Librarian alters the state of the given replica of the file.

The Shepherd tries to recover its replica by downloading it from another Shepherd. In order to do this the Shepherd contacts a Bartender and asks for the file. The Bartender chooses a valid replica, initiate file transfer by a Shepherd having a valid replica, and returns the TURL to the Shepherd with the invalid replica. The Shepherd with the invalid replica downloads the file from the other Shepherd, and if everything is OK, signals to the Librarian that the replica is \verb!alive! again.

The Shepherds periodically ask the Librarian whether the files they store have enough replicas. If a Shepherd finds that one of the files has not enough replica it turns to a Bartender offering replication. The Bartender chooses a Shepherd, initiates a put request then returns the TURL to the offering Shepherd which could upload the replica. The Shepherd who now has the new replica notifies the Librarian that the file is \verb!alive!. The Librarian sets the state of this new replica.

% section heartbeats_and_replication (end)

\chapter{Use cases} % (fold)
\label{cha:use_cases}

\section{Downloading a file} % (fold)
\label{sec:downloading_a_file}
\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-downloading.pdf}}}
\caption{\label{fig:downloading}Downloading a file} }
\end{figure}

We want to download a file about which we know that it is somewhere in our home collection on the storage (see Figure~\ref{fig:downloading}). The LN of our home collection is e.g. \verb#/ourvo/users/we#. We can get a list of entries in this collection from any Bartender.

\begin{enumerate}
\item We need to find a Bartender. Maybe we have a cached list of recently used Bartenders or we can get one from the information system. When we have an endpoint reference of a Bartender, we could call its list method with the LN \verb#/ourvo/users/we#.
\item The Bartender has to find a Librarian service, again using its cache of recently used Librarian services or get a new one from the information system. When rhe Bartender has an endpoint reference of a Librarian service, it could ask the Librarian to traverse the LN \verb#/ourvo/users/we#.
\item The Librarian needs the A-Hash service to access the stored data, when it has the endpoint reference of the A-Hash service, it could get the information about the root collection, which contains the GUID of the \verb#ourvo# sub-collection. Then the Librarian gets the entries of this \verb#ourvo# collection, and in it it can find the GUID of \verb#users#, and in the entries of \verb#users# there is the GUID of \verb#we#, which the Librarian returns to the Bartender with all the metadata.
\item The Bartender now has the GUID and the metadata of the collection \verb#/ourvo/users/we#, including the list of its entries. This is returned to us.
\item So we get the list of our \verb#/ourvo/users/we# collection, and now we realize that the file we want has the LN \verb#/ourvo/users/we/thefilewewant# and we know the GUID of it as well: e.g. \verb#a4b2e#. (Of course we know the GUID of the \verb#/ourvo/users/we# collection too, which is e.g. \verb#13245# and using this we could refer to our file as \verb#13245/thefilewewant# which means the entry called \verb#thefilewewant# in the collection with a GUID \verb#13245#.) We connect a Bartender again (the same one or maybe another one) to get the file with any of these LNs, the \verb#a4b2e# is the fastest solution because the Bartender need not to look up the whole LN again in the Librarian, a well-written client API should use this. With the get request we give the Bartender the list of transfer protocols we are able to use.
\item The Bartender contacts the Librarian to get the locations of the replicas of this file.
\item The Librarian returns all the metadata of the requested LN.
\item The Bartender chooses a ‘location’ which consists of the ID of a Shepherd, and the referenceID of the file within the Shepherd. Using the information system or its local cache it could get the endpoint reference of the Shepherd. The Bartender initiates a transfer by the Shepherd, if the Shepherd supports one of the transfer protocols we give, it can create a transfer URL (TURL) with a protocol we can download.
\item The Shepherd returns the TURL to the Bartender.
\item The Bartender returns the TURL to us.
\item Now we have a TURL from which we can download it.
\end{enumerate}

% section downloading_a_file (end)

\section{Uploading a file} % (fold)
\label{sec:uploading_a_file}

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-uploading.pdf}}}
\caption{\label{fig:uploading}Uploading a file} }
\end{figure}

We have a file on our local disk we want to upload to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
\begin{enumerate}
\item We contact a Bartender to put the file, we give the size and checksum and other metadata, and the transfer protocols we want to use. And of course we give the Logical Name we want to be the name of the file, which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
\item The Bartender ask a Librarian to traverse this LN.
\item If the Librarian can traverse the whole LN then this LN is already exists, but if the LN is still available and the parent exists, the Librarian’s response contains the GUID and metadata of the parent collection.
\item Then the Bartender creates a new file entry within the Librarian with all the information we gave. 
\item The Librarian returns the GUID of this new entry.
\item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf# which points to a file which has no replica at all. If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message ‘try again later’.
\item The Bartender (using the information system) chooses a Shepherd and gets its endpoint reference. Then the Bartender initiates uploading of the file to the Shepherd: the request includes the size and checksum of the file, the GUID, and the protocols we are able to use.
\item The Shepherd creates a transfer URL and a referenceID for this file and registers the GUID of the file in its own database and reports to the Librarian that there is a new replica with state \verb#creating#. The Librarian gets the message from the Shepherd and creates a new entry in the locations list of the given file with the serviceID and the referenceID the Shepherd have just reported. If someone tries to download this file now, still gets a ‘try again later’ error message, because this new replica is still not \verb#alive#.
\item The Shepherd returns the the TURL to the Bartender.
\item The Bartender returns the TURL to us.
\item Then we can upload the file to this TURL.
\item The Shepherd detects that the file is arrived and reports the change of state to \verb#alive# to the  Librarian who alters the state in the given file-entry. At this point the file has only one replica. 
\end{enumerate}
\begin{itemize}
\item The Shepherd periodically checks the Librarian if this is less than the needed replica number, and if it is then it initiates creating a new replica by a Bartender.
\item The Bartender chooses another Shepherd, initiates the transfer then returns the TURL to the  first Shepherd which uploads the file to the new Shepherd.
\item All the Shepherds check periodically whether their files have enough replica, and if any of them find that there is more replica needed, it initiates creating a new. If more than one Shepperd of course could cause that there will be more replicas than needed. If a Shepherd finds out that a file has more replicas than needed it notifies a Bartender about it.
\item The Bartender ask the Librarian about all Shepherds this file has replicas on, flags this file as ‘removing a replica’ which prevents other Bartenders to remove an other replica accidentally, then make a decision of which one is to be removed, then contacts the chosen Shepherd and asks it to remove the replica. The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag ‘removing a replica’ as well.
\item If we cannot upload the file to the given TURL for some reason, we should remove the file entry from the collection, or we should call \verb#addReplica# to get a new TURL without removing and recreating the file.
\end{itemize}

% section uploading_a_file (end)

\section{Removing a file} % (fold)
\label{sec:removing_a_file}

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-removing.pdf}}}
\caption{\label{fig:removing}Removing a file} }
\end{figure}
\begin{enumerate}
    \item If we want to remove a file, we should connect to a Bartender with the LN of the file we want to remove.
    \item The Bartender asks the Librarian to traverse the LN and return the list of parent collections of this file.
    \item The Librarian returns the data.
    \item If the file has only parent, it asks the Librarian to remove the file entry itself, and the entry from the parent collection. If the file has more parent collections (hardlinks) the Bartender only removes the entry from the parent collection and removes the parent from the list of parent  collections of this file.
    \item Next time the Shepherd does its periodic check, it asks the Librarian about each of its stored replicas, and finds out that one of them no longer exists, so it removes the replica from the storage node.
\end{enumerate}

% section removing_a_file (end)

% chapter use_cases (end)

\chapter{Technical description and implementation status} % (fold)
\label{cha:technical_description_and_implementation_status}

The current version of the prototype has no information system and no security, these are soon to be integrated to the system.
The information system is needed to discover services, and to translate serviceIDs to endpoint references (URLs). Currently the URLs are written in the configuration files, and the Shepherd services are reporting their URLs to the Librarian, so the Bartender could ask for all alive Shepherds.

The security is needed to do proper authorization of the users, and to manage access policies of files and collections. ARC has its own policy language, for each file and collection there will be a policy XML document stored as a metadata. The storage services will use the properties extracted from the communication channel and these policies to make authorization decisions. If the properties and the policies are present, the decision is actually made by the security framework of HED.
These are the planned actions which can be used for access control:
\begin{itemize}
    \item \emph{read}: user can get the list of entries in the collection; user can download the file
    \item \emph{addEntry}: user can add a new entry to the collection;
    \item \emph{removeEntry}: user can remove any entry from the collection 
    \item \emph{delete}: user can delete the collection if it is empty; user can delete a file (if you want to remove a file/collection, then the Bartender needs to remove the entry from the parent collection, and then delete the file/collection itself, so you need to have both permissions)
    \item \emph{modifyPolicy}: user can modify the policy of the file/collection
    \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are key-value pairs)
\end{itemize}
When a user has the permission in the Librarian to download a file then the user should have permission to access at least one of the file's replica, so there should be a Shepherd which allows the user the get the file. If the Bartender has permission to access the Shepherd, then the Bartender should create an assertion which allows the user to access the file. This could be a signed token which contain a policy defining access to particular file. But all these are currently just plans.

Further prototype statuses and plans can be found below within each section about the services.

% chapter technical_description_and_implementation_status (end)

\section{A-Hash} % (fold)
\label{sec:a_hash}

\subsection{Functionality} % (fold)

The A-Hash is a distributed service capable of storing objects containing property-value pairs grouped in sections in a scalable manner. Each object has an arbitrary string ID, and contains any number of property-value pairs grouped in sections, where property, value and section are arbitrary strings. There could only be one value for a property in a section.

If you have an ID, you can get all property-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties you need. You can add or remove property-value pairs of an object or delete all occurrences of a property or create a new object with the \emph{change} method, and you can specify conditions, which means the change is only applied if the given conditions are met.

% subsection functionality (end)

\subsection{Prototype status and plans} % (fold)

The A-Hash service currently implemented as a single central service, which stores the data on disk in separate files per object. In the fall of 2008 it will be reimplemented on a distributed hash table (DHT) algorithm, for example the Chord algorithm with a consistency solution called Etna on top of it. This reimplementation hopefully won’t change the interface of the service.

% subsection prototype_status_and_plans (end)

\subsection{Data model} % (fold)

\begin{itemize}
    \item \emph{ID} is an arbitrary string
    \item \emph{object} contains property-value pairs in sections, technically it is a list of key-value pairs where the key is a (section, property) tuple
\end{itemize}

% subsection data_model (end)

\subsection{Interface} % (fold)

\begin{description}
\item [get(ids, neededMetadata)] returns \emph{getResponse} which is a list of (\emph{ID}, \emph{object}) pairs.

The \emph{ids} is a list of string \emph{ID}s, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs. For each \emph{ID} it returns all the \emph{value}s for each \emph{property} in each \emph{section} (filtered by \emph{neededMetadata}), so \emph{object} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.

\item [change(changeRequest)] returns \emph{changeResponse} which is a list of (\emph{changeID}, \emph{success}, \emph{failedConditionID}) tuples.

\emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}, \emph{conditions}), where \emph{changeID} is an arbitrary ID to identify in the response which change was successful; \emph{ID} points to the object we want to change; \emph{changeType} can be `\textbf{set}' (set the property within the section to value), `\textbf{unset}' (remove the property from the section regardless of the value), `\textbf{delete}' (removes the whole object), conditions is a list of (\emph{conditionID}, \emph{type}, \emph{section}, \emph{property}, \emph{value}) tuples, where \emph{type} could be `\textbf{is}' (the property in the section is set to the value), `\textbf{isnot}' (the property in the section is not set to the value), `\textbf{isset}' (the property of the section is set to any value), `\textbf{unset}' (the property of the section is not set at all).
If all conditions are met, tries to apply changes to the objects, creates a new object if a previously non-existent ID is given. If one of the conditions is not met, returns the ID of the failed condition.
\end{description}

% subsection interface (end)

% section a_hash (end)

\section{Librarian} % (fold)
\label{sec:librarian}

\subsection{Functionality} % (fold)
% 
The Librarian manages a tree-hierarchy of files, grouping them into collections. There is a root collection with a well-known GUID which can be used as starting point when resolving Logical Names. If you create a new collection with the method \emph{new}, the Librarian generates a new GUID, but does not insert it into the tree-hierarchy which can be done by adding this GUID as a new entry to one of the existing collection using the \emph{modifyMetadata} method of the existing collection which makes it the parent of the new collection. A collection can be closed via metadata modification which cannot be undone and prevents files to be added or removed from this collection. A new file  also can be created with the \emph{new} method which returns the newly generated GUID of the new file entry which should be added to a parent collection to insert it into the global namespace. A file has a list of locations where its replicas are stored, this list too can be manipulated with \emph{modifyMetadata}. The access policies of the files and collections are also stored as metadata. The \emph{remove} method deletes an entry from the Librarian. The \emph{traverseLN} method try to traverse Logical Names by walking the hierarchy of the namespace and to return the GUID of the entry pointed by the LN. After you have a GUID of file, collection or mount point, you can get all the information using the \emph{get} method.

% subsection functionality (end)

\subsection{Prototype status and plans} % (fold)

The Librarian service currently implements all the methods below, but doesn’t do very much error checking. This should be changed, the Librarian should check the validity of metadata, and forbid some cases, e.g. reopen a closed collection.

% subsection prototype_status_and_plans (end)

\subsection{Data model} % (fold)
\label{sub:data_model}
Each librarian entry has a unique ID called \emph{GUID}.

The Librarian uses the A-Hash to store all the data about the files and collections. The A-Hash is capable of storing property-value pairs organized in sections, which actually means that it stores (\emph{section}, \emph{property}, \emph{value}) tuples where each member is simply a string, e.g. (‘entry’, ’type’, ’collection’) or (‘ACL’, ’johnsmith’, ’owner’) or (‘timestamps’, ’created’, ’1196265901’) or (‘locations’, ’64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB:9A293F27DC86’, ’sentenced’). There could be only one value for a (\emph{section}, \emph{property}) pair.

\begin{itemize}
    \item A \textbf{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-hierarchy. Each entry has a name which is only valid within this collection, and it is unique within the collection. Each entry is referenced by its GUID. So the metadata sections of a collection are as follows: 
    \begin{description}
        \item [entry] section 
        \begin{itemize}
            \item \emph{type}: `collection' 
        \end{itemize}
        \item [entries] section 
        \begin{itemize}
            \item \emph{(name, GUID) pairs}: a Collection is basically a list of name-GUID pairs. 
        \end{itemize}
        \item [timestamps] section 
        \begin{itemize}
            \item \emph{created}: timestamp of creation 
            \item \emph{modified}: timestamp of last modification 
        \end{itemize}
        \item [states] section 
        \begin{itemize}
            \item closed: if the collection is closed, then nothing can be added to its contents 
        \end{itemize}
        \item [policies] section 
        \begin{itemize}
            \item XML representations of access policies 
        \end{itemize}
        \item [metadata] section 
        \begin{itemize}
            \item any other arbitrary metadata 
        \end{itemize}
    \end{description}
    \item A \textbf{file} entry contains the following sections: 
    \begin{description}
        \item [entry] section 
        \begin{itemize}
            \item \emph{type}: `file' 
        \end{itemize}
        \item [locations] section 
        \begin{itemize}
            \item \emph{(location, state) pairs}, where a location is a (\emph{serviceID}, \emph{referenceID}) pair serialized as a  string, where \emph{serviceID} is the ID of the Shepherd service storing this replica, \emph{referenceID} is the ID of the file within that Shepherd service, and state could be `\textbf{alive}' (if the replica passed the checksum test, and the storage element storing it is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the storage element claims it has no such file), `\textbf{offline}' (if the storage element is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{sentenced}' (if the replica is marked for deletion) 
        \end{itemize}
        \item [timestamps] section 
        \begin{itemize}
            \item \emph{created}: timestamp of creation 
            \item \emph{modified}: timestamp of last modification (e.g. modification of metadata)
        \end{itemize}
        \item [states] section 
        \begin{itemize}
            \item \emph{size}: the file size in bytes
            \item \emph{checksum}: checksum of the file
            \item \emph{checksumType}: the name of the checksum method
            \item \emph{neededReplicas}: how many valid replicas should this file have 
        \end{itemize}
        \item [policies] section 
        \begin{itemize}
            \item XML representations of access policies 
        \end{itemize}
        \item [metadata] section 
        \begin{itemize}
            \item any other arbitrary metadata 
        \end{itemize}
    \end{description}
    \item There is one more type of Librarian entries called \textbf{mount point} which is a reference to a service which is capable of handling a subtree of the namespace. The properties of a mount point in sections:
    \begin{description}
        \item [entry] section 
        \begin{itemize}
            \item \emph{type}: `mountpoint' 
        \end{itemize}
        \item [mount] section 
        \begin{itemize}
            \item \emph{target}: the ID of the service
            \item \emph{ID}: an ID within the service (optional)
        \end{itemize}
        \item [timestamps] section 
        \begin{itemize}
            \item \emph{created}: timestamp of creation 
            \item \emph{modified}: timestamp of last modification (e.g. modification of metadata)
        \end{itemize}
        \item [policies] section 
        \begin{itemize}
            \item XML representations of access policies 
        \end{itemize}
        \item [metadata] section 
        \begin{itemize}
            \item any other arbitrary metadata 
        \end{itemize}
    \end{description}

    \item The Librarian stores information about the Shepherds, so each Shepherd has a GUID as well. There is an entry (with GUID `1’ by default) which contains the GUID and the timestamp of the last heartbeat for each registered Shepherd:
    \begin{description}
    	\item [nextHeartBeat] section
    	\begin{itemize}
    		\item (ID, timestamp) pairs
    	\end{itemize}
    	\item [serviceGUID] section
        \begin{itemize}
            \item (ID, GUID) pairs
        \end{itemize}
    \end{description}

    \item For each Shepherd there is a separate entry with the list of files:
    \begin{description}
    	\item [entry] section
        \begin{itemize}
            \item \emph{type}: `shepherd'
        \end{itemize}
    	\item [file] section
    	\begin{itemize}
    	    \item \emph{(referenceID, GUID) pairs} for each replica stored on the Shepherd
    	\end{itemize}
    \end{description}
\end{itemize}

% subsection data_model (end)

% section librarian (end)

\bibliography{grid}
\end{document}

### 2008-07-24 doc/storage_system/arc1-storage-documentation.tex
almost done with the technical documentation of the ARC1 storage

@@
 \newcommand{\dothis}{\colorbox{shadecolor}}
-\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
-\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
 \newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
@@
 
-% \begin{titlepage}
-% 
-% \begin{tabular}{rl}
-% \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
-% &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
-% \end{tabular}
-% 
-% \hrulefill
-% 
-% %-------- Change this to NORDUGRID-XXXXXXX-NN
-% 
-% {\raggedleft NORDUGRID-XXXXXXX-NN\par}
-% 
-% {\raggedleft \today\par}
-% 
-% \vspace*{2cm}
-% 
-% %%%%---- The title ----
-% {\centering \textsc{\Large Documentation of the ARC1 storage system}\Large \par}
-% \vspace*{0.5cm}
-%     
-% %%%%---- A subtitle, if necessary ----
-% {\centering \textit{\large First prototype status and plans}\large \par}
-%     
-% \vspace*{1.5cm}
-% %%%%---- A list of authors ----
-%     {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
-%     
-% \end{titlepage}
-% 
-% \tableofcontents                          %Comment if use article style
-% \newpage
+\begin{titlepage}
+
+\begin{tabular}{rl}
+\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
+&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
+\end{tabular}
+
+\hrulefill
+
+%-------- Change this to NORDUGRID-XXXXXXX-NN
+
+{\raggedleft NORDUGRID-XXXXXXX-NN\par}
+
+{\raggedleft \today\par}
+
+\vspace*{2cm}
+
+%%%%---- The title ----
+{\centering \textsc{\Large Documentation of the ARC1 storage system}\Large \par}
+\vspace*{0.5cm}
+
+%%%%---- A subtitle, if necessary ----
+{\centering \textit{\large First prototype status and plans}\large \par}
+
+\vspace*{1.5cm}
+%%%%---- A list of authors ----
+    {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
+
+\end{titlepage}
+
+\tableofcontents                          %Comment if use article style
+\newpage
+
+\renewcommand{\thefootnote}{\arabic{footnote}}
+
 \chapter{Design Overview}
@@
 
-% NorduGrid~\cite{nordugrid} Advanced Resource Connector (ARC)
-% \textit{middleware}\index{middleware} uses \GT~2 by Globus.
-% 
-% In expressions, the following operands are allowed:
-% \begin{shaded}
-%   \verb#=   !=   >   <   >=   <=#
-% \end{shaded}
-% 
-% \begin{framed}
-%   Examples of URLs are:\\
-%   \\
-%   \verb#http://grid.domain.org/dir/script.sh#\\
-%   \verb#gsiftp://grid.domain.org:2811;threads=10/dir/input_12378.dat#\\
-%   \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
-%   \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#
-%   \verb#file:///home/auser/griddir/steer.cra#\\
-% \end{framed}
-
-The ARC1 storage system is a distributed system for storing replicated \emph{files} on several file storage nodes and manage them in a global namespace.
-The files can be grouped into collections, and a \emph{collection} can contain sub-collections and sub-sub-collections in any depth.
+The ARC1 storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.
+The files can be grouped into \emph{collection}s, and a collection can contain sub-collections and sub-sub-collections in any depth.
 There is a dedicated \emph{root collection} to gather all collections to the global namespace.
-This hierarchy of collections and files can be referenced using \emph{Logical Names} (LN).
+This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}s).
 The users can use this global namespace  as they were using a local filesystem.
@@
 A storage node should have one or more data transfer service running (e.g.
-HTTP(S), FTP(S), GridFTP, ByteIO, etc.), and one of the ARC1 storage system’s services is needed to manage each storage node and integrating it into the system.
+HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.), and one of the services of the ARC1 storage system is needed to manage each storage node and to integrate it into the system.
 And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system.
@@
 \caption{\label{fig:services}The components of the ARC1 storage;
-A-H: the \textbf{A-Hash}, which is a currently centralized, later distributed database;
-L: the \textbf{Librarian}, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
-B: the \textbf{Bartender}, which provides a high-level interface for the users and for other services;
-S: the \textbf{Shepherd}, which provides a simple interface for storing files on storage nodes.
+\emph{A-H}: the \textbf{A-Hash}, which is a currently centralized, later distributed database;
+\emph{L}: the \textbf{Librarian}, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
+\emph{B}: the \textbf{Bartender}, which provides a high-level interface for the users and for other services;
+\emph{S}: the \textbf{Shepherd}, which provides a simple interface for storing files on storage nodes.
 } }
@@
     \item Each file and collection has a globally unique ID called \textbf{GUID}.
-    \item The files and collections are organized into a global hierarchical namespace and can be referred to using paths of this namespace called \textbf{Logical Names (LN)}.
+    \item The files and collections are organized into a global hierarchical namespace and can be referred to using paths of this namespace called \textbf{Logical Name}s (\textbf{LN}s).
     \item Each service has a unique \textbf{serviceID} which can be used to get an endpoint reference from the information system.
@@
 
-\section{GUIDs and Logical Names (LN)} % (fold)
+\section{GUIDs and Logical Names (LNs)} % (fold)
 \label{sec:GUIDsLNs}
 
-The syntax of Logical Names (LN): \verb#/[path]# or \verb#<GUID>[/<path>]# where [...] indicates optional parts.
+The syntax of Logical Names is \verb#/[path]# or \verb#<GUID>[/<path>]# where [...] indicates optional parts.
 
@@
 \centering{{\scalebox{0.7}{\includegraphics{arc1-storage-namespace.pdf}}}
-\caption{\label{fig:namespace}Example of the global namespace's hierarchy} }
+\caption{\label{fig:namespace}Example of the hierarchy of the global namespace} }
 \end{figure}
 
-Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file as \verb#/green/orange/huge#, if we know the GUID of the starting collection, so let's prefix the path with it: \verb#1234/green/orange/huge#.
+Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file as \verb#/green/orange/huge# if we know the GUID of the starting collection, so let's prefix the path with it: \verb#1234/green/orange/huge#.
 This is the Logical Name of that file.
@@
 If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, so it's a hard link.
-To count the references (the number of hardlinks) of a file or collection, the GUID of the parent collection(s) is stored as a metadata for each file and collection.
+To count the references (the number of hard links) of a file or collection, the GUID of the parent collection(s) is stored as a metadata for each file and collection.
 
-Each VO should create a VO-wide root collection, and put it in the generic root collection, e.g.
+Each VO\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} should create a VO-wide root collection, and put it in the generic root collection, e.g.
 if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#.
@@
 Clients can access the storage system through a Bartender service.
-If a client wants to create a collection, upload or download a file, the first step is to connect a Bartender.
-The Bartender then asks a Librarian to resolve Logical Names and return metadata, then initiates file transfers on some storage node by asking the Shepherd service on the node, then returns the transfer URL to the client, which allows the client to directly transfer from/to the storage node.
+If a client wants to create a collection, upload or download a file, the first step is to connect to a Bartender.
+The Bartender then asks a Librarian to resolve Logical Names and return metadata, then initiates file transfers on some storage node by asking the Shepherd service of that node, then returns the transfer URL (\emph{TURL}) to the client, which allows the client to directly transfer from/to the storage node.
 So the data transfer itself is not going through the Bartender, it is performed over a direct link between a storage node and the client.
@@
 A collection contains files and other collections, and each of these entries has a name unique within the collection very much like entries in a usual directory on a local filesystem.
-Besides files and collections the Librarian stores a third type of entries called Mount Points which are references to external services creating the capability to mount the namespace of third-party storages to our global namespace and make the files on a third-party storage available through the interface of the ARC storage system.
+Besides files and collections the Librarian stores a third type of entries called Mount Points which are references to external services which will be in the future used to mount the namespace of third-party storages to our global namespace and make the files on a third-party storage available through the interface of the ARC storage system.
 
@@
 All metadata about files and collections are stored in the A-Hash, and some information about Shepherd services is stored in it as well.
+The A-Hash itself does not interpret the data, it just stores tuples of strings.
 
@@
 When a new file is put into the system the number of needed replicas is given for the file.
-The file replicas are stored on different storage nodes, for each storage node there is Shepherd service which provides the interface for initiating file transfer.
+The file replicas are stored on different storage nodes, for each storage node there is Shepherd service which manages the storage node, reports its health state to the Librarian and provides the interface for initiating file transfer.
 
 The file-naming used by a Shepherd has nothing to do with the the hierarchy of collections, or Logical Names.
-When a replica is given to a Shepherd it gets an ID which refers to it within that Shepherd.
-Each Shepherd has a unique ID itself, so with these IDs the replica can be unambiguously referenced, this is called a Location.
+When a new replica upload is initiated, the Shepherd generates an ID which refers to it within that Shepherd.
+Each Shepherd has a unique ID itself, so with these IDs the replica can be unambiguously referenced, this is called a \emph{Location}.
 The namespace of these Locations has nothing to do with the namespace of GUIDs or the namespace of Logical Names.
@@
 If a Shepherd finds out that a file is missing or has a bad checksum, it reports that the file is \verb#invalid# to the Librarian immediately, and the Librarian alters the state of the given replica of the file.
-
 The Shepherd tries to recover its replica by downloading it from another Shepherd.
@@
 The LN of our home collection is e.g.
-\verb#/ourvo/users/we#.
+\verb#/ourvo/users/we# (here a `home collection' could be just a collection which is given to us by our VO).
 We can get a list of entries in this collection from any Bartender.
@@
 \item The Bartender has to find a Librarian service, again using its cache of recently used Librarian services or get a new one from the information system.
-When rhe Bartender has an endpoint reference of a Librarian service, it could ask the Librarian to traverse the LN \verb#/ourvo/users/we#.
+When the Bartender has an endpoint reference of a Librarian service, it could ask the Librarian to traverse the LN \verb#/ourvo/users/we#.
 \item The Librarian needs the A-Hash service to access the stored data, when it has the endpoint reference of the A-Hash service, it could get the information about the root collection, which contains the GUID of the \verb#ourvo# sub-collection.
@@
 \item The Librarian returns all the metadata of the requested LN.
-\item The Bartender chooses a ‘location’ which consists of the ID of a Shepherd, and the referenceID of the file within the Shepherd.
+\item The Bartender chooses a `location' which consists of the ID of a Shepherd, and the referenceID of the file within the Shepherd.
 Using the information system or its local cache it could get the endpoint reference of the Shepherd.
@@
 \item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf# which points to a file which has no replica at all.
-If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message ‘try again later’.
+If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message `try again later'.
 \item The Bartender (using the information system) chooses a Shepherd and gets its endpoint reference.
@@
 The Librarian gets the message from the Shepherd and creates a new entry in the locations list of the given file with the serviceID and the referenceID the Shepherd have just reported.
-If someone tries to download this file now, still gets a ‘try again later’ error message, because this new replica is still not \verb#alive#.
+If someone tries to download this file now, still gets a `try again later' error message, because this new replica is still not \verb#alive#.
 \item The Shepherd returns the the TURL to the Bartender.
@@
 \item All the Shepherds check periodically whether their files have enough replica, and if any of them find that there is more replica needed, it initiates creating a new.
-If more than one Shepperd of course could cause that there will be more replicas than needed.
+If more than one Shepherd of course could cause that there will be more replicas than needed.
 If a Shepherd finds out that a file has more replicas than needed it notifies a Bartender about it.
-\item The Bartender ask the Librarian about all Shepherds this file has replicas on, flags this file as ‘removing a replica’ which prevents other Bartenders to remove an other replica accidentally, then make a decision of which one is to be removed, then contacts the chosen Shepherd and asks it to remove the replica.
-The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag ‘removing a replica’ as well.
+\item The Bartender ask the Librarian about all Shepherds this file has replicas on, flags this file as `removing a replica' which prevents other Bartenders to remove an other replica accidentally, then make a decision of which one is to be removed, then contacts the chosen Shepherd and asks it to remove the replica.
+The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag `removing a replica' as well.
 \item If we cannot upload the file to the given TURL for some reason, we should remove the file entry from the collection, or we should call \verb#addReplica# to get a new TURL without removing and recreating the file.
@@
     \item If the file has only parent, it asks the Librarian to remove the file entry itself, and the entry from the parent collection.
-If the file has more parent collections (hardlinks) the Bartender only removes the entry from the parent collection and removes the parent from the list of parent  collections of this file.
+If the file has more parent collections (hard links) the Bartender only removes the entry from the parent collection and removes the parent from the list of parent  collections of this file.
     \item Next time the Shepherd does its periodic check, it asks the Librarian about each of its stored replicas, and finds out that one of them no longer exists, so it removes the replica from the storage node.
@@
 
+The services are written in Python and running in the HED\footnote{The ARC container - https://www.knowarc.eu/documents/Knowarc\_D1.2-2\_07.pdf} hosting environment.
+The HED itself is written in C++, but there are language bindings which allow us to write services in other languages, e.g.
+in Python or Java.
+The source tree of the storage services are in the NorduGrid Subversion repository\footnote{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage}.
+
 The current version of the prototype has no information system and no security, these are soon to be integrated to the system.
-The information system is needed to discover services, and to translate serviceIDs to endpoint references (URLs).
+
+The information system is needed to discover services, and to translate \emph{serviceID}s to endpoint references (URLs).
 Currently the URLs are written in the configuration files, and the Shepherd services are reporting their URLs to the Librarian, so the Bartender could ask for all alive Shepherds.
@@
 If the properties and the policies are present, the decision is actually made by the security framework of HED.
-These are the planned actions which can be used for access control:
-\begin{itemize}
-    \item \emph{read}: user can get the list of entries in the collection; user can download the file
-    \item \emph{addEntry}: user can add a new entry to the collection;
-    \item \emph{removeEntry}: user can remove any entry from the collection 
-    \item \emph{delete}: user can delete the collection if it is empty; user can delete a file (if you want to remove a file/collection, then the Bartender needs to remove the entry from the parent collection, and then delete the file/collection itself, so you need to have both permissions)
-    \item \emph{modifyPolicy}: user can modify the policy of the file/collection
-    \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
-    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are key-value pairs)
-\end{itemize}
-When a user has the permission in the Librarian to download a file then the user should have permission to access at least one of the file's replica, so there should be a Shepherd which allows the user the get the file.
-If the Bartender has permission to access the Shepherd, then the Bartender should create an assertion which allows the user to access the file.
-This could be a signed token which contain a policy defining access to particular file.
-But all these are currently just plans.
 
@@
 % chapter technical_description_and_implementation_status (end)
+\newpage
 
@@
 % section a_hash (end)
+\newpage
 
-\section{Librarian} % (fold)
+\section{Librarians} % (fold)
 \label{sec:librarian}
@@
 The A-Hash is capable of storing property-value pairs organized in sections, which actually means that it stores (\emph{section}, \emph{property}, \emph{value}) tuples where each member is simply a string, e.g.
-(‘entry’, ’type’, ’collection’) or (‘ACL’, ’johnsmith’, ’owner’) or (‘timestamps’, ’created’, ’1196265901’) or (‘locations’, ’64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB:9A293F27DC86’, ’sentenced’).
-There could be only one value for a (\emph{section}, \emph{property}) pair.
+(`entry', `type', `collection') or (`ACL', `johnsmith', `owner') or (`timestamps', `created', `1196265901') or (`locations', `64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB 9A293F27DC86', `sentenced').
+There could be only one \emph{value} for a (\emph{section}, \emph{property}) pair.
 
@@
     \item The Librarian stores information about the Shepherds, so each Shepherd has a GUID as well.
-There is an entry (with GUID `1’ by default) which contains the GUID and the timestamp of the last heartbeat for each registered Shepherd:
+There is an entry (with GUID `1' by default) which contains the GUID and the timestamp of the last heartbeat for each registered Shepherd:
     \begin{description}
@@
 
+\subsection{Interface} % (fold)
+
+ \begin{description}
+    \item[new(newRequestList)] returns a list of (\emph{requestID}, \emph{GUID}, \emph{success})
+
+    \emph{newRequestList} is a list of (\emph{requestID}, \emph{metadata}) where \emph{requestID} is an arbitrary ID used to identify this request in the list of responses; \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+    This method generates a \emph{GUID} for each request, and inserts the new entry into the A-Hash, then returns the GUIDs of the newly created entries.
+
+    \item[modifyMetadata(modifyMetadataRequestList)] returns a list of (\emph{changeID}, \emph{success})
+
+    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}’ (set the property in the section to the value), `\textbf{unset}’ (remove the property-value pair from a section), `\textbf{add}’ (set the property in the section to the value only if it is not exists already).
+
+    \item [get(GUIDs, neededMetadata)] returns \emph{getResponse}
+
+    \emph{GUIDs} is a list of GUIDs, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs indicating only which properties we need, \emph{getResponse} is a list of (\emph{GUID}, \emph{metadata}) where metadata is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+
+    \item [remove(removeRequestList)] returns a list of (\emph{requestID}, \emph{success}) pairs
+
+    \emph{removeRequestList} is a list of (\emph{requestID}, \emph{GUID}) pairs. \emph{success} could be `\textbf{removed}' or `\textbf{failed}: reason'.
+
+    \item [traverseLN(traverseRequestList)] returns \emph{traverseResponseList}
+
+    \emph{traverseRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names to be traversed
+
+    \emph{traverseResponseList} is a list of (\emph{requestID}, \emph{metadata}, \emph{GUID}, \emph{traversedLN}, \emph{restLN}, \emph{wasComplete}, \emph{traversedList}) where:
+    \begin{description}
+        \item[metadata] is all the metadata of the of traversedLN in the form of (\emph{section}, \emph{property}, \emph{value}) tuples
+        \item[GUID] is the \emph{GUID} of the \emph{traversedLN}
+        \item[traversedLN] is the part of the \emph{LN} which was traversed, if \emph{wasComplete} is true, this should be the full \emph{LN}
+        \item[restLN] is the postfix of the \emph{LN} which was not traversed for some reason, if \emph{wasComplete} is true, this should be an empty string
+        \item[wasComplete] indicates whether the full \emph{LN} was traversed
+        \item[traversedList] is a list of (\emph{LNpart}, \emph{GUID}) pairs, where \emph{LNpart} is a part of the \emph{LN}, \emph{GUID} is the GUID of the Librarian-entry referenced by that part of the \emph{LN}, the first element of this list is the shortest prefix of the \emph{LN}, the last element is the \emph{LN} without its last part
+    \end{description}
+
+    \item [report(serviceID, filelist)] returns in \emph{nextReportTime} a number of seconds, which is the timeframe within the Librarian expects the next heartbeat from the Shepherd
+
+    \emph{filelist} is a list of (\emph{GUID}, \emph{referenceID}, \emph{state}) tuples containing the state of changed or new files, where \emph{state} could be `\textbf{invalid}' (if the periodic self-check of the Shepherd found a non-matching checksum or missing file), `\textbf{creating}' (if this is a new file not uploaded yet) or `\textbf{alive}' (if the file is uploaded and the checksum is OK).
+
+\end{description}
+
+% subsection interface (end)
+
 % section librarian (end)
+\newpage
 
-\bibliography{grid}
+\section{Shepherds} % (fold)
+\label{sec:shepherds}
+
+\subsection{Functionality} % (fold)
+
+The Shepherd service is capable of managing a storage node.
+It keeps track all the files it stores with their GUIDs and checksums.
+The Shepherds periodically send reports to a Librarian indicating that the Shepherd is up and running, and whether some file's state has been changed.
+The Shepherd periodically checks each file to detect corruption.
+If a file goes missing or has a bad checksum the Shepherd notify the Librarian about the error referring the file with its GUID.
+The Shepherd periodically asks the Librarian how many replicas its files have, and if a file has fewer replicas than needed, the Shepherd offers its copy for replication by calling the Bartender.
+
+A Shepherd service is always connected to a file transfer service (e.g.
+`HTTP(S)', `FTP(S)', `ByteIO', `GridFTP', etc.).
+For each supported file transfer service we need a backend module which makes the Shepherd able to communicate with the file transfer service to initiate file transfers, to detect whether a transfer was successful or not, to generate local IDs and checksum, etc.
+
+A file in a Shepherd could be identified with a \emph{referenceID} which is unique within that Shepherd.
+If we know the \emph{location} of a file, which is the ID of the Shepherd service (\emph{serviceID}) and the \emph{referenceID}, we could get the endpoint reference (URL) of the Shepherd from the information system, then we could call its \emph{get} method with the \emph{referenceID} and a list of transfer protocols we are able to handle (e.g.
+`HTTP', `FTP'), the Shepherd chooses a protocol from this list which it can provide, and create a transfer URL (\emph{TURL}) and returns it along with the \emph{checksum} of the file.
+We could download the file from this \emph{TURL}, and verify it with the \emph{checksum}.
+An end user of the storage system does not need to call this \emph{get} method, because the Bartender service will do it, the user just ask the Bartender and get the TURL.
+
+Storing a file starts with initiating the transfer with the \emph{put} method of the Shepherd, we should give the \emph{size} and \emph{checksum} of the file and its \emph{GUID} as well.
+We also specify a list of transfer protocols we are able to use, and the Shepherd chooses a \emph{protocol}, creates a \emph{TURL} for uploading and generates a \emph{referenceID}, then we can upload the file to the TURL.
+Again, the end user just asks the Bartender, and get the TURL, the user does not need to call the \emph{put} method of the Shepherd directly.
+
+These \emph{TURL}s are one-time URLs which means that after the client uploads or downloads the file these \emph{TURL}s cannot be used again to do the same.
+If we want to download the same file twice, we have to initiate the transfer twice, and will get two different \emph{TURL}s.
+
+With the \emph{stat} method we can get some information about a replica, e.g. checksum, GUID, state (`creating', `alive' or `invalid'), etc. The \emph{delete} method removes the file.
+
+In normal operation the \emph{put} and \emph{get} calls is made by a Bartender but the actual uploading and downloading is done by the user’s client.
+In case of replication a Shepherd with a valid replica initiates the replication, this Shepherd asks the Bartender to choose a new Shepherd, the Bartender initiates putting the new replica on a chosen Shepherd and receives the TURL, then the Bartender returns the TURL to the initiator Shepherd, which uploads its replica to the given TURL.
+
+% subsection functionality (end)
+
+\subsection{Prototype status and plans} % (fold)
+
+The current implementation of the Shepherd service has a working \emph{get}, \emph{put}, \emph{stat}, \emph{delete} methods, and a method called \emph{toggleReport} which can be used the simulate storage node failure with the Shepherd not reporting to a Librarian.
+There is a separate service which provide a subset of the ByteIO interface, and there is an other separate service which is a basic HTTP server, these are both could be used as file transfer services, both have its backend module for the Shepherd.
+Currently both file transfer services have the problem of using too much memory while transferring files.
+Further plans include better file transfer services and backend modules for third-party file transfer services.
+
+% subsection prototype_status_and_plans (end)
+
+\subsection{Data model} % (fold)
+
+A file of a Shepherd service is referenced by its \emph{referenceID}.
+Each file has a \emph{state} which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, or `\textbf{invalid}' if it does not exists anymore or has a bad checksum.
+Each file has a \emph{localID} which is used in the backend modules..
+
+% subsection data_model (end)
+
+\subsection{Interface} % (fold)
+
+\begin{description}
+    \item[get(getRequestList)] returns list of (\emph{requestID}, \emph{getResponseData})
+
+    \emph{getRequestList} is a list of (\emph{requestID}, \emph{getRequestData}) where \emph{requestID} is an arbitrary ID used in the reply
+
+    \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client is able to use, there could be multiple protocols in getRequestData.
+
+    \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.
+
+    \item[put(putRequestList)] returns a list of (\emph{requestID}, \emph{putResponseData})
+
+    \emph{putRequestList} is a list of (\emph{requestID}, \emph{putRequestData}) where \emph{requestID} is an ID used for the response
+
+    \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (for additional access policy).
+
+    \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.
+
+    \item[delete(deleteRequestList)] returns a list of (\emph{requestID}, \emph{status})
+
+    \emph{deleteRequestList} is a list of (\emph{requestID}, \emph{referenceID}) pairs selecting the files to remove. The status could be `\textbf{deleted}' or `\textbf{nosuchfile}'
+    .
+    \item[stat(statRequestList)] returns a list of (\emph{requestID}, \emph{referenceID}, \emph{state}, \emph{checksumType}, \emph{checksum}, \emph{acl}, \emph{size}, \emph{GUID}, \emph{localID})
+
+    \emph{statRequestList} is a list of (\emph{requestID}, \emph{referenceID}) where \emph{referenceID} points to the file whose data we want to get.
+The method returns all the data the Shepherd know about the replica.
+
+\end{description}
+
+% subsection interface (end)
+
+\subsection{Backend modules} % (fold)
+\label{sub:backend_modules}
+
+The Shepherd could communicate with the file transfer services via backend modules.
+Currently there are two kinds of backend modules, one for the byteio service (which is a simple implementation of a subset of the ByteIO interface) and one for the httpd service (which is a simple HED-based HTTP server).
+
+In both cases the Shepherd service uses two separate directories: one for storing all the files (e.g.
+\verb!./store!) and one for the file transfers (e.g.
+\verb!./transfer!).
+The store directory always contains all the files the Shepherd manages, the transfer directory is empty at the beginning.
+
+Let's see the scenario for the httpd service which should be in a special `slave' mode for this kind of operation: if a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), then the Shepherd service creates a hardlink into the transfer directory (e.g.
+\verb!./transfer/abc!) and sets this file read-only.
+If the httpd service is configured that way that it handles the HTTP path \verb!/prb!
+and it is serving files from the directory \verb!./transfer!
+then after the hardlink is created, we have this URL for this file: \verb!http://localhost:60000/prb/abc!.
+Now we can give this URL to the client.
+Then the client \verb!GET!s this URL and gets the file.
+The httpd service removes (unlinks) this file immediately after the \verb!GET!
+request arrived, which makes this \verb!http://localhost:60000/prb/abc!
+URL invalid (so this is a one-time URL), but because of the hardlink the file is still there in the store directory, it is just removed from the transfer directory.
+Now if some other user wants this file, the Shepherd creates an other hardlink, e.g.
+\verb!./transfer/qwe!
+and now we have an URL \verb!http://localhost:60000/prb/qwe!.
+
+If a client wants to upload a new file, then the Shepherd creates an empty file in the store directory, e.g.
+\verb!./store/file2!
+and creates a hardlink into the transfer directory, e.g.
+\verb!./transfer/oiu!
+and makes it writable, and now we have a URL \verb!http://localhost:60000/prb/oiu!, and the client is able to do a \verb!PUT!
+to this URL.
+When the client \verb!PUT!s the file there, the httpd service immediately removes the uploaded file from the transfer directory, but because it has a hardlink in the store directory, the file is stored there as \verb!./store/file2!.
+The backend module for the httpd service periodically checks whether a new file has two or just one hard links.
+If it has only one that means that a file is uploaded, so it could notify the Shepherd that the file is arrived.
+In order to do that, all the backend modules get a callback method `file\_arrived' from the Shepherd.
+
+In case of the byteio service, there is some small differences.
+The byteio service does not removes the files from the transfer directory, but it calls the backend module via SOAP to notify it that something is happened.
+The byteio backend has one SOAP method called `\textbf{notify}':
+\begin{description}
+    \item[notify(subject, state)] returns `\textbf{OK}' in \emph{notifyResponse}.
+\end{description}
+When this method is called, the byteio backend module notifies the Shepherd that the file is arrived.
+
+All the backend modules should have this common interface which the Shepherd can use to communicate with the file transfer service:
+\begin{description}
+    \item[prepareToGet(referenceID, localID, protocol)] returns the \emph{TURL}.
+
+    Initiate transfer with \emph{protocol} for the file which has these IDs: \emph{localID} and \emph{referenceID}.
+The reason for including here the referenceID as well is that this information could be used by the backend module later, e.g.
+when the transfer finished and the state of the file needs to be changed.
+
+    \item[prepareToPut(referenceID, localID, protocol)] returns the \emph{TURL}.
+
+    Initiate transfer with \emph{protocol} for the file which has these IDs: \emph{localID} and \emph{referenceID}.
+
+    \item[copyTo(localID, turl, protocol)] returns \emph{success}.
+
+    Upload the file referenced by \emph{localID} to the given \emph{TURL} with the given \emph{protocol}.
+
+    \item[copyFrom(localID, turl, protocol)] returns \emph{success}.
+
+    Download the file from the given \emph{TURL} with the given \emph{protocol}, and store it as \emph{localID}.
+
+    \item[list()] returns a list of \emph{localID}s currently in the store directory.
+    \item[getAvailableSpace()] returns the available disk space in bytes.
+    \item[generateLocalID()] returns a new unique \emph{localID}.
+    \item[matchProtocols(protocols)] only leave that protocols in the list \emph{protocols} which are supported by this file transfer service.
+    \item[checksum(localID, checksumType)] returns the checksum of the file referenced by \emph{localID}, which checksum is generated by the method \emph{checksumType}.
+
+\end{description}
+
+% subsection backend_modules (end)
+
+% section shepherds (end)
+\newpage
+
+\section{Bartenders} % (fold)
+\label{sec:bartenders}
+
+\subsection{Functionality} % (fold)
+
+The Bartender provides an easy to use interface of the ARC1 storage system to the users.
+You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}.
+The metadata of a file or collection (e.g.
+whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}.
+A \emph{stat} call gives all the information about a file or collection, and you can move (or hardlink) collections and files within the namespace with \emph{move}.
+You can upload an entirely new replica to a file (e.g.
+if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.
+
+% subsection functionality (end)
+
+\subsection{Prototype status and plans} % (fold)
+
+The methods mentioned in the above section are all implemented, but need more error-checking and metadata-checking.
+There are plans of adding new methods, e.g.
+a \emph{copy} method or a \emph{glob} method for file pattern matching.
+The current version does not support closed (unmodifiable) collections yet.
+
+% subsection protoype_status_and_plans (end)
+
+\subsection{Data model} % (fold)
+
+The Bartender interface uses mostly Logical Names (\emph{LN}s), which have the syntax of: \verb!<GUID>/<path>!
+where both sides can be omitted (and in the case of a sole GUID we don't need the slash either), e.g.
+\verb!afg342/foo!
+is an entry called \verb!foo!
+in the collection with GUID \verb!afg342!; the LN `\verb!f36a7481!' refers to the a file or collection with GUID \verb!f36a7481!; \verb!/vo/dir/stg!
+points to the entry which is reachable from the root collection using the given path; and \verb!/!
+simply refers to the root collection.
+The term `\emph{metadata}' here refers to a list of property-value pairs organized in sections, see the data model description in Section~\ref{sub:data_model}.
+
+% subsection data_model (end)
+
+\subsection{Interface} % (fold)
+
+\begin{description}
+    \item[putFile(putFileRequest)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})
+
+    \emph{putFileRequest} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other property-value pairs in the `\textbf{metadata}' section.
+The returned \emph{TURL} is a URL with a chosen \emph{protocol} to upload the file itself, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.
+
+    \item[getFile(getFileRequest)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})
+
+    \emph{getFileRequest} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the file we want to get, \emph{protocols} is a list of transfer protocols the client supports.
+    In the response \emph{TURL} is the transfer URL using \emph{protocol}, with which we can download the file, \emph{success} could be `\textbf{done}', `\textbf{not found}', `\textbf{is not a file}', `\textbf{file has no valid replica}', `\textbf{error while getting TURL:} reason', etc.
+
+    \item[delFile(delFileRequest)] returns a list of (\emph{requestID}, \emph{status})
+
+    \emph{delFileRequest} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file we want to delete. The status in response could be `\textbf{deleted}' or `\textbf{nosuchLN}'.
+
+    \item[stat(statRequest)] returns a list of (\emph{requestID}, \emph{metadata})
+
+    \emph{statRequest} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sub:data_model})
+
+    \item[makeCollection(makeCollectionRequest)] returns a list of (\emph{requestID}, \emph{success})
+
+    \emph{makeCollectionRequest} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in key-value pairs.
+The \emph{success} in the response could be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.
+    \emph[list(listRequest, neededMetadata)] returns \emph{listResponse}.
+
+    \emph{listRequest} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
+
+    \emph{listResponse} is a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (Section~\ref{sub:data_model}), the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}' (because only collections can be listed).
+
+    \item[move(moveRequest)] returns a list of (\emph{requestID}, \emph{status}).
+
+    \emph{moveRequest} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) where \emph{sourceLN} is the Logical Name referring to the file or collection we want to move (or just rename) and \emph{targetLN} is the new path, and if \emph{preserveOriginal} is true the \emph{sourceLN} would not be removed, so with \emph{preserveOriginal} we actually creating a hard link.
+The status could be `\textbf{moved}', `\textbf{nosuchLN}', `\textbf{targetexists}', `\textbf{invalidtarget}', `\textbf{failed adding child to parent}', `\textbf{failed removing child from parent}'
+
+    \item[unmakeCollection(unmakeRequest)] returns a list of (\emph{requestID}, \emph{success}).
+
+    \emph{unmakeRequest} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names of the collections we want to remove.
+\emph{success} could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}', `\textbf{failed: reason}'.
+
+    \item[modify(modifyRequest)] returns a list of (\emph{changeID}, \emph{success})
+    \emph{modifyRequest} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}’ (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}’ (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}’ (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).
+\emph{success} could be `\textbf{no such LN}', `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}', `\textbf{failed}: reason'.
+
+\end{description}
+
+% subsection interface (end)
+
+% section bartenders (end)
+\newpage
+
+\section{Client tools} % (fold)
+\label{sec:client_tools}
+
+In the first prototype release there is one client tool called \verb!arc_storage_cli.py!, which is written in Python, and only need a basic Python installation to run.
+It is capable of communicating with a given Bartender service, and uploading and downloading TURLs via HTTP.
+
+The methods can be listed with:
+\begin{verbatim}
+$ arc_storage_cli.py
+Usage:
+  arc_storage_cli.py <method> [<arguments>]
+Supported methods: stat, make[Collection], unmake[Collection], list, move,
+    put[File], get[File], del[File]
+\end{verbatim}
+Without arguments, each method prints its own help:
+\begin{verbatim}
+$ arc_storage_cli.py move
+Usage: move <sourceLN> <targetLN>
+\end{verbatim}
+
+Uploading, downloading and stat files:
+
+\begin{verbatim}
+$ cat testfile 
+This is a testfile.
+$ arc_storage_cli.py put testfile /tmp/
+- The size of the file is 20 bytes
+- The md5 checksum of the file is 9a9dffa22d227afe0f1959f936993a80
+- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
+- Calling the Bartender's putFile method...
+- done in 0.08 seconds.
+- Got transfer URL: http://localhost:60000/httpd/d15900f5-34ee-4bba-bb10-73d60d1c0d75
+- Uploading from 'testfile'
+    to 'http://localhost:60000/httpd/d15900f5-34ee-4bba-bb10-73d60d1c0d75' with http...
+Uploading 20 bytes... data sent, waiting... done.
+- done in 0.0042 seconds.
+'testfile' (20 bytes) uploaded as '/tmp/testfile'.
+$ arc_storage_cli.py stat /tmp/testfile
+- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
+- Calling the Bartender's stat method...
+- done in 0.05 seconds.
+'/tmp/testfile': found
+  states
+    checksumType: md5
+    neededReplicas: 1
+    size: 20
+    checksum: 9a9dffa22d227afe0f1959f936993a80
+  timestamps
+    created: 1210232135.57
+  parents
+    51e12fab-fd3d-43ec-9bc5-17041da3f0b2/testfile: parent
+  locations
+    http://localhost:60000/Shepherd fc0d3d99-6406-4c43-b2eb-c7ec6d6ab7fe: alive
+  entry
+    type: file
+$ arc_storage_cli.py get /tmp/testfile newfile
+- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
+- Calling the Bartender's getFile method...
+- done in 0.05 seconds.
+- Got transfer URL: http://localhost:60000/httpd/dab911d0-110f-468e-b0c3-627af6e3af31
+- Downloading from 'http://localhost:60000/httpd/dab911d0-110f-468e-b0c3-627af6e3af31'
+    to 'newfile' with http...
+Downloading 20 bytes... done.
+- done in 0.0035 seconds.
+'/tmp/testfile' (20 bytes) downloaded as 'newfile'.
+$ cat newfile 
+This is a testfile.
+\end{verbatim}
+
+You can find more examples in the SVN\footnote{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage/README}.
+
+There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, http://fuse.sourceforge.net/} module and Windows Shell Extensions\footnote{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx} to be able to mount the ARC1 storage namespace into the local filesystem namespace, and use it with the commands of the operating system.
+
+% section client_tools (end)
+
+\section{Other plans} % (fold)
+
+\subsection{Integrating third-party storage solutions} % (fold)
+\label{sub:integrating_third_party_storage_solutions}
+
+To intagrate existing files on a third-party storage to our namespace thus make them accessable through the interface of the ARC1 storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
+It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC1 storage (Section~\ref{sub:data_model}).
+The files here would be referenced by a path which is local in the namespace of the third-party storage.
+The interface could be something like this:
+\begin{description}
+    \item[get(getRequest)] returns list of (\emph{requestID}, \emph{getResponseData})
+    very similar to the \emph{get} method of the Shepherd
+    \item[put(putRequest)] returns a list of (\emph{requestID}, \emph{putResponseData})
+    very similar to the \emph{put} method of the Shepherd
+    \item[delete(deleteRequest)] returns a list of (\emph{requestID}, \emph{success})
+    very similar to the \emph{delete} method of the Shepherd
+    \item[stat(statRequest)] returns a list of (\emph{requestID}, \emph{statResponse})
+    very similar to the \emph{stat} method of the Shepherd
+    \item[list(listRequest)] returns \emph{listResponse}
+    very similar to the \emph{list} method of the Bartender
+    \item[move(moveRequest)] returns a list of (\emph{requestID}, \emph{status})
+    very similar to the \emph{move} method of the Bartender
+\end{description}
+This interface has some methods similar to the Shepherd and some other methods similar to the Bartender.
+% subsection integrating_third_party_storage_solutions (end)
+
+\subsection{Security} % (fold)
+\label{sub:security}
+
+These are the planned actions which can be used for access control:
+\begin{itemize}
+    \item \emph{read}: user can get the list of entries in the collection; user can download the file
+    \item \emph{addEntry}: user can add a new entry to the collection;
+    \item \emph{removeEntry}: user can remove any entry from the collection 
+    \item \emph{delete}: user can delete the collection if it is empty; user can delete a file (if you want to remove a file/collection, then the Bartender needs to remove the entry from the parent collection, and then delete the file/collection itself, so you need to have both permissions)
+    \item \emph{modifyPolicy}: user can modify the policy of the file/collection
+    \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
+    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are key-value pairs)
+\end{itemize}
+When a user has the permission in the Librarian to download a file then the user should have permission to access at least one of the file's replica, so there should be a Shepherd which allows the user the get the file.
+If the Bartender has permission to access the Shepherd, then the Bartender should create an assertion which allows the user to access the file.
+This could be a signed token which contain a policy defining access to particular file.
+But all these are currently just plans.
+
+% subsection security (end)
+
+% section other_plans (end)
+
+% \bibliography{grid}
 \end{document}

### 2008-07-24 doc/storage_system/arc1-storage-documentation.tex
arc1 storage documentation - D2.1-1 technical documentation for the first review

@@
 %----- DON'T CHANGE HEADER MATTER
+\hyphenation{preserve-Original}
 \begin{document}
@@
 The replicas of the files are stored on different storage nodes.
-A storage node should have one or more data transfer service running (e.g.
-HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.), and one of the services of the ARC1 storage system is needed to manage each storage node and to integrate it into the system.
+A storage node should have one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.), and one of the services of the ARC1 storage system is needed to manage each storage node and to integrate it into the system.
 And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system.
@@
 This is the Logical Name of that file.
-If there is a well-known system-wide root collection (its GUID could be e.g.
-\verb#0#), then if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g.
-\verb#/why/blue# means \verb#0/why/blue#.
+If there is a well-known system-wide root collection (its GUID could be e.g.~\verb#0#), then if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g.~\verb#/why/blue# means \verb#0/why/blue#.
 
@@
 
-Each VO\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} should create a VO-wide root collection, and put it in the generic root collection, e.g.
-if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#.
-Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g.
-\verb#/vo1/file1#, etc.
-Or sub-collections, e.g.
-\verb#/vo1/col1#, \verb#/vo1/col2/file3#, etc.
+Each VO\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} should create a VO-wide root collection, and put it in the generic root collection, e.g.~if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#.
+Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g.~\verb#/vo1/file1#, etc.
+Or sub-collections, e.g.~\verb#/vo1/col1#, \verb#/vo1/col2/file3#, etc.
 For this the VO does not need the install any service.
@@
 We want to download a file about which we know that it is somewhere in our home collection on the storage (see Figure~\ref{fig:downloading}).
-The LN of our home collection is e.g.
-\verb#/ourvo/users/we# (here a `home collection' could be just a collection which is given to us by our VO).
+The LN of our home collection is e.g.~\verb#/ourvo/users/we# (here a `home collection' could be just a collection which is given to us by our VO).
 We can get a list of entries in this collection from any Bartender.
@@
 \item The Bartender now has the GUID and the metadata of the collection \verb#/ourvo/users/we#, including the list of its entries. This is returned to us.
-\item So we get the list of our \verb#/ourvo/users/we# collection, and now we realize that the file we want has the LN \verb#/ourvo/users/we/thefilewewant# and we know the GUID of it as well: e.g.
-\verb#a4b2e#.
-(Of course we know the GUID of the \verb#/ourvo/users/we# collection too, which is e.g.
-\verb#13245# and using this we could refer to our file as \verb#13245/thefilewewant# which means the entry called \verb#thefilewewant# in the collection with a GUID \verb#13245#.) We connect a Bartender again (the same one or maybe another one) to get the file with any of these LNs, the \verb#a4b2e# is the fastest solution because the Bartender need not to look up the whole LN again in the Librarian, a well-written client API should use this.
+\item So we get the list of our \verb#/ourvo/users/we# collection, and now we realize that the file we want has the LN \verb#/ourvo/users/we/thefilewewant# and we know the GUID of it as well: e.g.~\verb#a4b2e#.
+(Of course we know the GUID of the \verb#/ourvo/users/we# collection too, which is e.g.~\verb#13245# and using this we could refer to our file as \verb#13245/thefilewewant# which means the entry called \verb#thefilewewant# in the collection with a GUID \verb#13245#.) We connect a Bartender again (the same one or maybe another one) to get the file with any of these LNs, the \verb#a4b2e# is the fastest solution because the Bartender need not to look up the whole LN again in the Librarian, a well-written client API should use this.
 With the get request we give the Bartender the list of transfer protocols we are able to use.
@@
 The services are written in Python and running in the HED\footnote{The ARC container - https://www.knowarc.eu/documents/Knowarc\_D1.2-2\_07.pdf} hosting environment.
-The HED itself is written in C++, but there are language bindings which allow us to write services in other languages, e.g.
-in Python or Java.
-The source tree of the storage services are in the NorduGrid Subversion repository\footnote{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage}.
+The HED itself is written in C++, but there are language bindings which allow us to write services in other languages, e.g.~in Python or Java.
+The source code of the storage services are in the NorduGrid Subversion repository\footnote{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage}.
 
@@
 
-The A-Hash is a distributed service capable of storing objects containing property-value pairs grouped in sections in a scalable manner.
-Each object has an arbitrary string ID, and contains any number of property-value pairs grouped in sections, where property, value and section are arbitrary strings.
-There could only be one value for a property in a section.
+The A-Hash will be a distributed service capable of storing tuples of strings in a scalable manner.
+Currently it only has a centralized implementation.
+It stores \emph{objects}, where each object has an arbitrary string \emph{ID}, and contains any number of \emph{property}-\emph{value} pairs grouped in \emph{section}s, where \emph{property}, \emph{value} and \emph{section} are arbitrary strings.
+There could only be a single \emph{value} for a \emph{property} in a \emph{section}.
 
-If you have an ID, you can get all property-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties you need.
+If you have an ID, you can get all property-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties do you need.
 You can add or remove property-value pairs of an object or delete all occurrences of a property or create a new object with the \emph{change} method, and you can specify conditions, which means the change is only applied if the given conditions are met.
@@
 
-The A-Hash service currently implemented as a single central service, which stores the data on disk in separate files per object.
-In the fall of 2008 it will be reimplemented on a distributed hash table (DHT) algorithm, for example the Chord algorithm with a consistency solution called Etna on top of it.
+The A-Hash service currently implemented as a single central service, which stores the data on disk in separate files per \emph{object}.
+Fall of 2008 it will be reimplemented using a distributed hash table (DHT) algorithm, one possible candidate is the Chord\footnote{The Chord project - http://pdos.csail.mit.edu/chord/} algorithm with a consistency solution called Etna\footnote{Etna: a Fault-tolerant Algorithm for Atomic Mutable DHT Data - http://pdos.csail.mit.edu/\~athicha/papers/etna.ps} on top of it.
 This reimplementation hopefully won’t change the interface of the service.
@@
     \item \emph{ID} is an arbitrary string
-    \item \emph{object} contains property-value pairs in sections, technically it is a list of key-value pairs where the key is a (section, property) tuple
+    \item \emph{object} contains property-value pairs in sections, technically it is a list of key-\emph{value} pairs where the key is a (\emph{section}, \emph{property}) tuple
 \end{itemize}
@@
 
-The Librarian service currently implements all the methods below, but doesn’t do very much error checking.
-This should be changed, the Librarian should check the validity of metadata, and forbid some cases, e.g.
-reopen a closed collection.
+The Librarian service currently implements all the methods below, but doesn't do very much error checking.
+This should be changed, the Librarian should check the validity of metadata, and forbid some cases, e.g.~reopen a closed collection.
 
@@
 
-The Librarian uses the A-Hash to store all the data about the files and collections.
-The A-Hash is capable of storing property-value pairs organized in sections, which actually means that it stores (\emph{section}, \emph{property}, \emph{value}) tuples where each member is simply a string, e.g.
-(`entry', `type', `collection') or (`ACL', `johnsmith', `owner') or (`timestamps', `created', `1196265901') or (`locations', `64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB 9A293F27DC86', `sentenced').
+The Librarian uses the A-Hash to store all the data about files and collections.
+The A-Hash is capable of storing property-value pairs organized in sections, which actually means that it stores (\emph{section}, \emph{property}, \emph{value}) tuples where each member is simply a string, e.g.~(`entry', `type', `collection') or (`ACL', `johnsmith', `owner') or (`timestamps', `created', `1196265901') or (`locations', `64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB:9A293F27DC86', `sentenced').
 There could be only one \emph{value} for a (\emph{section}, \emph{property}) pair.
@@
         \begin{itemize}
-            \item closed: if the collection is closed, then nothing can be added to its contents 
+            \item \emph{closed}: if the collection is closed, then nothing can be added to its contents 
         \end{itemize}
@@
         \begin{itemize}
-            \item \emph{(location, state) pairs}, where a location is a (\emph{serviceID}, \emph{referenceID}) pair serialized as a  string, where \emph{serviceID} is the ID of the Shepherd service storing this replica, \emph{referenceID} is the ID of the file within that Shepherd service, and state could be `\textbf{alive}' (if the replica passed the checksum test, and the storage element storing it is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the storage element claims it has no such file), `\textbf{offline}' (if the storage element is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{sentenced}' (if the replica is marked for deletion)
+            \item \emph{(location, state) pairs}, where a location is a (\emph{serviceID}, \emph{referenceID}) pair serialized as a  string, where \emph{serviceID} is the ID of the Shepherd service storing this replica, \emph{referenceID} is the ID of the file within that Shepherd service, and state could be `\textbf{alive}' (if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the Shepherd claims it has no such file), `\textbf{offline}' (if the Shepherd is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{sentenced}' (if the replica is marked for deletion)
         \end{itemize}
@@
             \item \emph{created}: timestamp of creation 
-            \item \emph{modified}: timestamp of last modification (e.g. modification of metadata)
+            \item \emph{modified}: timestamp of last modification (e.g.~modification of metadata)
         \end{itemize}
@@
         \end{itemize}
-    	\item [file] section
+    	\item [files] section
     	\begin{itemize}
@@
     \emph{newRequestList} is a list of (\emph{requestID}, \emph{metadata}) where \emph{requestID} is an arbitrary ID used to identify this request in the list of responses; \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
-    This method generates a \emph{GUID} for each request, and inserts the new entry into the A-Hash, then returns the GUIDs of the newly created entries.
+    This method generates a \emph{GUID} for each request, and inserts the new entry (with the given metadata) into the A-Hash, then returns the GUIDs of the newly created entries.
+The (`entry', `type') property of the metadata contains whether it is a file or a collection.
 
@@
 
-    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}’ (set the property in the section to the value), `\textbf{unset}’ (remove the property-value pair from a section), `\textbf{add}’ (set the property in the section to the value only if it is not exists already).
+    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the property in the section to value), `\textbf{unset}' (remove the property-value pair from the section), `\textbf{add}' (set the property in the section to value only if it is not exists already).
 
@@
     \emph{GUIDs} is a list of GUIDs, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs indicating only which properties we need, \emph{getResponse} is a list of (\emph{GUID}, \emph{metadata}) where metadata is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+This method returns the metadata of all the \emph{GUID}s filtered with \emph{neededMetadata}.
 
@@
 
-The Shepherd service is capable of managing a storage node.
+A Shepherd service is capable of managing a storage node.
 It keeps track all the files it stores with their GUIDs and checksums.
-The Shepherds periodically send reports to a Librarian indicating that the Shepherd is up and running, and whether some file's state has been changed.
-The Shepherd periodically checks each file to detect corruption.
-If a file goes missing or has a bad checksum the Shepherd notify the Librarian about the error referring the file with its GUID.
-The Shepherd periodically asks the Librarian how many replicas its files have, and if a file has fewer replicas than needed, the Shepherd offers its copy for replication by calling the Bartender.
+It periodically checks each file to detect corruption, and send reports to a Librarian indicating that the storage node is up and running, and whether some file's state has been changed.
+If a file goes missing or has a bad checksum then the Librarian is notified about the error (here the Shepherd refers to the file with its GUID, that's why it needs to store the GUIDs of its files).
+It periodically asks the Librarian how many replicas its files have, and if a file has fewer replicas than needed, the Shepherd offers its copy for replication by calling the Bartender.
 
-A Shepherd service is always connected to a file transfer service (e.g.
-`HTTP(S)', `FTP(S)', `ByteIO', `GridFTP', etc.).
-For each supported file transfer service we need a backend module which makes the Shepherd able to communicate with the file transfer service to initiate file transfers, to detect whether a transfer was successful or not, to generate local IDs and checksum, etc.
+A Shepherd service is always connected to a file transfer service (e.g.~`HTTP(S)', `FTP(S)', `ByteIO', `GridFTP', etc.).
+For each supported file transfer service we need a backend module which makes the Shepherd capable of communicating with the file transfer service to initiate file transfers, to detect whether a transfer was successful or not, to generate local IDs and checksums, etc.
 
-A file in a Shepherd could be identified with a \emph{referenceID} which is unique within that Shepherd.
-If we know the \emph{location} of a file, which is the ID of the Shepherd service (\emph{serviceID}) and the \emph{referenceID}, we could get the endpoint reference (URL) of the Shepherd from the information system, then we could call its \emph{get} method with the \emph{referenceID} and a list of transfer protocols we are able to handle (e.g.
-`HTTP', `FTP'), the Shepherd chooses a protocol from this list which it can provide, and create a transfer URL (\emph{TURL}) and returns it along with the \emph{checksum} of the file.
+A file in a storage node could be identified with a \emph{referenceID} which is unique within that node.
+If we know the \emph{location} of a file, which is the ID of the Shepherd service (\emph{serviceID}) and the \emph{referenceID}, we could get the endpoint reference (URL) of the Shepherd from the information system, then we could call its \emph{get} method with the \emph{referenceID} and a list of transfer protocols we can use (e.g.~`HTTP', `FTP'), the Shepherd chooses a protocol from this list which it can provide, and create a transfer URL (\emph{TURL}) and returns it along with the \emph{checksum} of the file.
 We could download the file from this \emph{TURL}, and verify it with the \emph{checksum}.
-An end user of the storage system does not need to call this \emph{get} method, because the Bartender service will do it, the user just ask the Bartender and get the TURL.
+An end user of the storage system does not need to call this \emph{get} method, because the Bartender service will do it, the user just asks the Bartender and gets the TURL.
 
@@
 We also specify a list of transfer protocols we are able to use, and the Shepherd chooses a \emph{protocol}, creates a \emph{TURL} for uploading and generates a \emph{referenceID}, then we can upload the file to the TURL.
-Again, the end user just asks the Bartender, and get the TURL, the user does not need to call the \emph{put} method of the Shepherd directly.
+Again, the end user just asks the Bartender, and gets the TURL, the user does not need to call the \emph{put} method of the Shepherd directly.
 
@@
 
-With the \emph{stat} method we can get some information about a replica, e.g. checksum, GUID, state (`creating', `alive' or `invalid'), etc. The \emph{delete} method removes the file.
+With the \emph{stat} method we can get some information about a replica, e.g.~checksum, GUID, state (`creating', `alive' or `invalid'), etc. The \emph{delete} method removes the file.
 
-In normal operation the \emph{put} and \emph{get} calls is made by a Bartender but the actual uploading and downloading is done by the user’s client.
-In case of replication a Shepherd with a valid replica initiates the replication, this Shepherd asks the Bartender to choose a new Shepherd, the Bartender initiates putting the new replica on a chosen Shepherd and receives the TURL, then the Bartender returns the TURL to the initiator Shepherd, which uploads its replica to the given TURL.
+In normal operation the \emph{put} and \emph{get} calls is made by a Bartender but the actual uploading and downloading is done by the user's client.
+In the case of replication a Shepherd with a valid replica initiates the replication, this Shepherd asks the Bartender to choose a new Shepherd, the Bartender initiates putting the new replica on a chosen Shepherd and receives the TURL, then the Bartender returns the TURL to the initiator Shepherd, which uploads its replica to the given TURL.
 
@@
 Each file has a \emph{state} which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, or `\textbf{invalid}' if it does not exists anymore or has a bad checksum.
-Each file has a \emph{localID} which is used in the backend modules..
+Each file has a \emph{localID} which is used in the backend modules.
 
@@
 
-    \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client is able to use, there could be multiple protocols in getRequestData.
+    \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).
 
@@
 The Shepherd could communicate with the file transfer services via backend modules.
-Currently there are two kinds of backend modules, one for the byteio service (which is a simple implementation of a subset of the ByteIO interface) and one for the httpd service (which is a simple HED-based HTTP server).
+Currently there are two kinds of backend modules, one for the \emph{byteio} service (which is a simple implementation of a subset of the ByteIO interface) and one for the \emph{httpd} service (which is a simple HED-based HTTP server).
 
-In both cases the Shepherd service uses two separate directories: one for storing all the files (e.g.
-\verb!./store!) and one for the file transfers (e.g.
-\verb!./transfer!).
+In both cases the Shepherd service uses two separate directories: one for storing all the files (e.g.~\verb!./store!) and one for the file transfers (e.g.~\verb!./transfer!).
 The store directory always contains all the files the Shepherd manages, the transfer directory is empty at the beginning.
 
-Let's see the scenario for the httpd service which should be in a special `slave' mode for this kind of operation: if a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), then the Shepherd service creates a hardlink into the transfer directory (e.g.
-\verb!./transfer/abc!) and sets this file read-only.
+Let's see the scenario for the httpd service which should be in a special `slave' mode for this kind of operation: if a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), then the Shepherd service creates a hardlink into the transfer directory (e.g.~\verb!./transfer/abc!) and sets this file read-only.
 If the httpd service is configured that way that it handles the HTTP path \verb!/prb!
@@
 URL invalid (so this is a one-time URL), but because of the hardlink the file is still there in the store directory, it is just removed from the transfer directory.
-Now if some other user wants this file, the Shepherd creates an other hardlink, e.g.
-\verb!./transfer/qwe!
+Now if some other user wants this file, the Shepherd creates an other hardlink, e.g.~\verb!./transfer/qwe!
 and now we have an URL \verb!http://localhost:60000/prb/qwe!.
 
-If a client wants to upload a new file, then the Shepherd creates an empty file in the store directory, e.g.
-\verb!./store/file2!
-and creates a hardlink into the transfer directory, e.g.
-\verb!./transfer/oiu!
+If a client wants to upload a new file, then the Shepherd creates an empty file in the store directory, e.g.~\verb!./store/file2!
+and creates a hardlink into the transfer directory, e.g.~\verb!./transfer/oiu!
 and makes it writable, and now we have a URL \verb!http://localhost:60000/prb/oiu!, and the client is able to do a \verb!PUT!
@@
     Initiate transfer with \emph{protocol} for the file which has these IDs: \emph{localID} and \emph{referenceID}.
-The reason for including here the referenceID as well is that this information could be used by the backend module later, e.g.
-when the transfer finished and the state of the file needs to be changed.
+The reason for including here the referenceID as well is that this information could be used by the backend module later, e.g.~when the transfer finished and the state of the file needs to be changed.
 
@@
 You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}.
-The metadata of a file or collection (e.g.
-whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}.
+The metadata of a file or collection (e.g.~whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}.
 A \emph{stat} call gives all the information about a file or collection, and you can move (or hardlink) collections and files within the namespace with \emph{move}.
-You can upload an entirely new replica to a file (e.g.
-if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.
+You can upload an entirely new replica to a file (e.g.~if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.
 
@@
 The methods mentioned in the above section are all implemented, but need more error-checking and metadata-checking.
-There are plans of adding new methods, e.g.
-a \emph{copy} method or a \emph{glob} method for file pattern matching.
+There are plans of adding new methods, e.g.~a \emph{copy} method or a \emph{glob} method for file pattern matching.
 The current version does not support closed (unmodifiable) collections yet.
@@
 The Bartender interface uses mostly Logical Names (\emph{LN}s), which have the syntax of: \verb!<GUID>/<path>!
-where both sides can be omitted (and in the case of a sole GUID we don't need the slash either), e.g.
-\verb!afg342/foo!
+where both sides can be omitted (and in the case of a sole GUID we don't need the slash either), e.g.~\verb!afg342/foo!
 is an entry called \verb!foo!
-in the collection with GUID \verb!afg342!; the LN `\verb!f36a7481!' refers to the a file or collection with GUID \verb!f36a7481!; \verb!/vo/dir/stg!
+in the collection with GUID \verb!afg342!; the LN \verb!f36a7481!
+refers to the a file or collection with GUID \verb!f36a7481!; \verb!/vo/dir/stg!
 points to the entry which is reachable from the root collection using the given path; and \verb!/!
@@
 \begin{description}
-    \item[putFile(putFileRequest)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})
+    \item[putFile(putFileRequestList)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})
 
-    \emph{putFileRequest} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other property-value pairs in the `\textbf{metadata}' section.
+    \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other property-value pairs in the `\textbf{metadata}' section.
 The returned \emph{TURL} is a URL with a chosen \emph{protocol} to upload the file itself, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.
 
-    \item[getFile(getFileRequest)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})
+    \item[getFile(getFileRequestList)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})
 
-    \emph{getFileRequest} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the file we want to get, \emph{protocols} is a list of transfer protocols the client supports.
+    \emph{getFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the file we want to get, \emph{protocols} is a list of transfer protocols the client supports.
     In the response \emph{TURL} is the transfer URL using \emph{protocol}, with which we can download the file, \emph{success} could be `\textbf{done}', `\textbf{not found}', `\textbf{is not a file}', `\textbf{file has no valid replica}', `\textbf{error while getting TURL:} reason', etc.
 
-    \item[delFile(delFileRequest)] returns a list of (\emph{requestID}, \emph{status})
+    \item[delFile(delFileRequestList)] returns a list of (\emph{requestID}, \emph{status})
 
-    \emph{delFileRequest} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file we want to delete. The status in response could be `\textbf{deleted}' or `\textbf{nosuchLN}'.
+    \emph{delFileRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file we want to delete. The status in response could be `\textbf{deleted}' or `\textbf{nosuchLN}'.
 
-    \item[stat(statRequest)] returns a list of (\emph{requestID}, \emph{metadata})
+    \item[stat(statRequestList)] returns a list of (\emph{requestID}, \emph{metadata})
 
-    \emph{statRequest} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sub:data_model})
+    \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sub:data_model})
 
-    \item[makeCollection(makeCollectionRequest)] returns a list of (\emph{requestID}, \emph{success})
+    \item[makeCollection(makeCollectionRequestList)] returns a list of (\emph{requestID}, \emph{success})
 
-    \emph{makeCollectionRequest} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in key-value pairs.
+    \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in key-value pairs.
 The \emph{success} in the response could be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.
-    \emph[list(listRequest, neededMetadata)] returns \emph{listResponse}.
+    \item[unmakeCollection(unmakeCollectionRequestList)] returns a list of (\emph{requestID}, \emph{success}).
 
-    \emph{listRequest} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
+    \emph{unmakeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names of the collections we want to remove.
+\emph{success} could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}', `\textbf{failed}: reason'.
+
+    \item[list(listRequestList, neededMetadata)] returns \emph{listResponse}.
+
+    \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
 
@@
 
-    \item[move(moveRequest)] returns a list of (\emph{requestID}, \emph{status}).
+    \item[move(moveRequestList)] returns a list of (\emph{requestID}, \emph{status}).
 
-    \emph{moveRequest} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) where \emph{sourceLN} is the Logical Name referring to the file or collection we want to move (or just rename) and \emph{targetLN} is the new path, and if \emph{preserveOriginal} is true the \emph{sourceLN} would not be removed, so with \emph{preserveOriginal} we actually creating a hard link.
+    \emph{moveRequestList} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) where \emph{sourceLN} is the Logical Name referring to the file or collection we want to move (or just rename) and \emph{targetLN} is the new path, and if \emph{preserveOriginal} is true the \emph{sourceLN} would not be removed, so with \emph{preserveOriginal} we actually creating a hard link.
 The status could be `\textbf{moved}', `\textbf{nosuchLN}', `\textbf{targetexists}', `\textbf{invalidtarget}', `\textbf{failed adding child to parent}', `\textbf{failed removing child from parent}'
 
-    \item[unmakeCollection(unmakeRequest)] returns a list of (\emph{requestID}, \emph{success}).
+    \item[modify(modifyRequestList)] returns a list of (\emph{changeID}, \emph{success})
 
-    \emph{unmakeRequest} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names of the collections we want to remove.
-\emph{success} could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}', `\textbf{failed: reason}'.
-
-    \item[modify(modifyRequest)] returns a list of (\emph{changeID}, \emph{success})
-    \emph{modifyRequest} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}’ (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}’ (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}’ (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).
+    \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).
 \emph{success} could be `\textbf{no such LN}', `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}', `\textbf{failed}: reason'.
@@
 
-To intagrate existing files on a third-party storage to our namespace thus make them accessable through the interface of the ARC1 storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
+To intagrate existing files on a third-party storage to our namespace thus make them accessible through the interface of the ARC1 storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
 It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC1 storage (Section~\ref{sub:data_model}).
@@
     \item[get(getRequest)] returns list of (\emph{requestID}, \emph{getResponseData})
+
     very similar to the \emph{get} method of the Shepherd
+
     \item[put(putRequest)] returns a list of (\emph{requestID}, \emph{putResponseData})
+
     very similar to the \emph{put} method of the Shepherd
+
     \item[delete(deleteRequest)] returns a list of (\emph{requestID}, \emph{success})
+
     very similar to the \emph{delete} method of the Shepherd
+
     \item[stat(statRequest)] returns a list of (\emph{requestID}, \emph{statResponse})
+
     very similar to the \emph{stat} method of the Shepherd
+
     \item[list(listRequest)] returns \emph{listResponse}
+
     very similar to the \emph{list} method of the Bartender
+
     \item[move(moveRequest)] returns a list of (\emph{requestID}, \emph{status})
+
     very similar to the \emph{move} method of the Bartender
+
 \end{description}

### 2008-07-25 doc/ws-quick-guide/ws-quick-guide.tex
Start writing ARC WS Quick Usage guide (development of the top README)

%\documentclass{book}
\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{listings}                                %for code listings
\lstloadlanguages{bash,XML}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {NorduGrid},
  pdftitle = {ARC Web Services Quick Usage Guide},
  pdfsubject = {ARC Grid middleware},
  pdfkeywords = {Grid,middleware,ARC,KnowARC,NorduGrid},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~4~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://download.nordugrid.org/}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
\definecolor{lightblue}{cmyk}{0.05,0,0,0.05}
\lstset{
frame=single,
backgroundcolor=\color{shadecolor},
fillcolor=\color{shadecolor},
rulecolor=\color{lightblue},
basicstyle=\ttfamily,
xleftmargin=0.05\linewidth,
aboveskip=2\medskipamount
}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-XXXXXXX-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large ARC Web Services Quick Usage Guide}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large Paper subtitle}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
%    {\centering \large A. Konstantinov\footnote{const@takas.lt} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage
%\chapter{Preface}

\section{ARC Middleware}                    %Use Sections for articles
\label{sec:arc-intro}

The Advanced Resource Connector (ARC) middleware~\cite{arc}, introduced by
NorduGrid~\cite{nordugrid}, is an open source software solution
enabling production quality computational and data grids.

Since the first release (May 2002) the middleware has been deployed and
been used in production environments. Emphasis is put on scalability,
stability, reliability and performance of the middleware. A growing
number of grid projects, like NDGF, Swegrid or Swiss National Grid has
chosen ARC as their middleware.

This document provides a quick usage guide to the new Web Service (WS)
components of ARC, often referred to as ARC1 components. WS design of ARC is
based on a service container -- the Hosting Environment Daemon (HED), and
different grid capabilities are implemented as Web Services residing in HED.

Currently included components are:
\begin{itemize}
\item an OGSA BES~\cite{ogsa-bes} compliant execution service -- \em{ARC
Resource-coupled EXecution service (A-REX)}
\item an echo service (for testing purposes)
\end{itemize}

\section{Dependencies}
\label{sec:deps}

The core part of middleware is written in C/C++. Building the software
from source or installing a precompiled binary requires different external
packages, furthermore the client and server packages have different dependencies
too. Below a list of the explicit requirements is shown:

\begin{itemize}
\item Mandatory (on client as well as server side):
\begin{itemize}
\item gnu make, autotools (autoconf $\geq$ 2.56, automake $\geq$ 1.8) (build)
\item C++ compiler and library (build)
\item libtool (build)
\item pkg-config (build)
\item gthread-2.0 $\geq$2 .4.7 (build, run)
\item glibmm-2.4 $\geq$ 2.4.7 (build, run)
\item libxml-2.0 $\geq$ 2.4.0 (build, run)
\item openssl $\geq$ 0.9.7a (build, run)
\item e2fsprogs (build, run)
\item doxygen (build)
\end{itemize}
\item Optional (mainly applicable on server side):
\begin{itemize}
\item Swig $\geq$ 1.3.28 (build)
\item Java SDK $\geq$ 1.4 for Java bindings (build, run)
\item Python for Python bindings (build, run)
\item Grid Packaging Tools (GPT)~\cite{gpt} (build)
\item \globus~\cite{globus} which contains (build, run):
\begin{itemize}
\item Globus RLS client
\item Globus FTP client
\item Globus RSL
\end{itemize}
\item LHC File Catalog (LFC)~\cite{lfc} (build, run)
\item CppUnit for unit testing (build)
\item Berkeley DB C++ interface (build, run)
\end{itemize}
\end{itemize}

\begin{framed}
Please note that depending on operating system distribution in order to build
ARC1 you may need to install development versions of mentioned packages.
\end{framed}

\section{Getting the software}
\label{sec:download}

The software is free to deploy anywhere by anybody. Pre-built binaries for a
dozen of Linux platforms can be downloaded from the NorduGrid software
repository at \ngdl (``arc1'' repository).

The software is released under the GNU General Public License (GPL) (see the
COPYING file distributed with the software).

You can get the latest source code for ARC1 from the Subversion repository. See
\url{http://svn.nordugrid.org} for more details.

The NorduGrid software repository hosts the source code, and provides most of
the required external software which are not part of a standard Linux
distribution.

There are also nightly code snapshots available at:

\begin{shaded}
\url{http://download.nordugrid.org/software/nordugrid-arc1/nightly/}
\end{shaded}

Choose the latest date available and download snapshot tarball -- for example
\verb|nordugrid-arc1-200802201038-snapshot.tar.gz|.

\section{Building and Installation}
\label{sec:build}

Building from source is currently the recommended way to install ARC Web
Services. If you downloaded the tarball, unpack it and traverse into the created
directory:

\begin{lstlisting}
tar -zxvf nordugrid-arc1-200802201038-snapshot.tar.gz
cd nordugrid-arc1-200802201038
\end{lstlisting}

If you obtained the code from the Subversion repository, use the ``trunk''
directory:

\begin{lstlisting}
cd trunk
\end{lstlisting}

Now configure the obtained code with:

\begin{lstlisting}
./autogen.sh
./configure --prefix=PLACE_TO_INSTALL_ARC
\end{lstlisting}

Choose installation prefix wisely and according to the requirements of your OS
and personal preferences. ARC1 services should function properly from any
location. By default installation goes into \texttt{/usr/local} if you omit the
``\texttt{--prefix}'' option. For some modules of ARC1 to work properly you may
need to set up the environment variable after installation:

\begin{lstlisting}
export ARC_LOCATION=PLACE_TO_INSTALL_ARC
\end{lstlisting}

On some systems ``\texttt{autogen.sh}'' may produce few warnings. Ignore them as
long as ``\texttt{configure}'' passes without errors. But in case of problems
during configure or compilation, collect them and present while reporting
problems.

If the previous commands finish without errors, compile and install ARC1
services:

\begin{lstlisting}
make
make install
\end{lstlisting}

Depending on chosen installation location you may need to run the last command
from \texttt{root} account. That should install the following components:

\begin{itemize}
\item \texttt{sbin/arched} - server executable
\item \texttt{bin/} - user tools and command line clients
\item \texttt{lib/} - common libraries used by clients, server and plugins
\item \texttt{lib/arc/} - plugins implementing Message Chain, Service and
Security components
\item \texttt{include/arc/} - C++ headers for application development
\item \texttt{libexec/} - additional modules used by ARC1 services - currently
only A-REX
\item \texttt{share/doc/arc} - configuration examples/templates and
documentation
\item \texttt{share/locale} - internationalization files - curently very limited
support
\item \texttt{share/man} - manual pages for various utilities
\end{itemize}

\section{Security and Authorization}
\label{sec:security}

ARC1 sevices implement security related features through set of Security Handler
and Policy Decision Point components. Security Handler components are attached
to message processing components. Each Security Handler takes care of processing
own part of security information. Currently ARC comes with the following
Security Handlers:

\begin{itemize}
 \item \texttt{identity.map} -- associates client's identity with local (UNIX)
identity. It uses PDP components to choose local isentity and/or identity
mapping algorithm.
\item \texttt{arc.authz} -- calls PDP components and combines
obtained authorization decisions.
\item \texttt{delegation.collector} -- parses proxy policy from remote proxy
certificate. This Security Handler should be configured under TLS MCC component.
\item \texttt{usernametoken.handler} -- implement the functioanlity
of WS-Security Usernametoken profile. It will generate Usernametoken
into SOAP header, or extract Usernametoken out of SOAP header and
do authentication based on the extracted Usernametoken.
\end{itemize}

Among available PDP components there are:
\begin{itemize}
\item \texttt{allow} -- always returns positive result
\item \texttt{deny} -- always returns negative result 
\item  \texttt{simplelist.pdp} -- compares DN of user to those stored in a file.
\item  \texttt{arc.pdp} -- compares request information parsed from message and
policy information specified in this PDP.
\item  \texttt{pdpservice.invoker} -- composes the request, puts request into
SOAP message, and invokes the remote PDP service to get the response SOAP
which includes authorization decision. The PDP service has similar
functionality with \texttt{arc.pdp}.
 \item \texttt{delegation.pdp} -- compares request information parsed from
message and policy information specified in proxy certificate from remote side.
\end{itemize}

There are examples of A-REX service and Echo service with Security Handlers
being used. They may be found in
\verb|$ARC_LOCATION/share/doc/arc/arex_secure.xml| and
\verb|$ARC_LOCATION/share/doc/arc/echo.xml|.

There is also a PDP service which implements the same functionality as
\texttt{arc.pdp}. See \texttt{src/service/pdp/README}.

Specifically for \texttt{arc.pdp} and \texttt{pdpservice}, a formatted policy
with specific schema should be managed, see
\verb|$ARC_LOCATION/share/doc/arc/pdp_policy.xml.example| and 
\verb|$ARC_LOCATION/share/doc/arc/Policy.xsd| for details.

For usernametoken handler, there is example about configuration on service side
in \verb|$ARC_LOCATION/share/doc/arc/echo.xml|, you can run Echo service by
using this configuration file with usernametoken sechandler configuered. For
client side, the echo client (\texttt{src/client/echo})can use usernametoken
sechandler to authenticate against echo service (see README under
\verb|src/client/echo|); there is also a test program in
\verb|src/tests/echo/test_clientinterface.cpp| which can be compiled and
tested against Echo service with usernametoken sechandler configured.

\subsection{X509 Certificates}
\label{sec:x509}

Most of  planned and existing ARC1 services use HTTPS as transport protocol so
they require proper setup of X509 security infrastructure~\cite{x509}. Minimal
requirements are:

\begin{itemize}
\item Host certificate aka public key in PEM format
\item Corresponding private key
\item Certificate of the Certification Authority (CA) which was used to sign
the host certificate
\item Certificates of CAs of clients which are going to send requests to
services (unless of course clients use the same CA as the server).
\end{itemize}

More information about X509 certificates and their usage in Grid environement
can be found in:

\begin{shaded}
\url{http://www.nordugrid.org/documents/certificate_howto.html}\\
\url{http://www.nordugrid.org/documents/ng-server-install.html#security}.
\end{shaded}

For testing purposes you can use the pre-generated certificates and keys
available in the code repository:

\begin{shaded}
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/sec/TestCA}
\end{shaded}

Alternatively you may choose to use KnowARC Instant CA service available at:

\begin{shaded}
\url{https://vls.grid.upjs.sk/CA/instantCA}
\end{shaded}

The latter is especially usefull if you want to test installation consisting of
multiple hosts.

Please remember that it is not safe to use these keys in publicly accessible
insallations of ARC. Make sure that even the CA certificate is removed before
making your services available to the outside world.

You can put host certificates and private keys anywhere, their location is
configurable. Common locations for servers running from root account are:

\texttt{/etc/grid-security/hostcert.pem} \\and\\
\texttt{/etc/grid-security/hostkey.pem}

respectively.

\begin{framed}
Since services have no way to ask for passwords, the content of private key
must not be encrypted, neither it should be protected by password. So make sure
it is properly protected by means of the file system.
\end{framed}

It is possible to configure ARC1 server to accept either a single CA certificate
or multiple CA certificates located in the specified directory. The latter
option is recommended. The common location is
\texttt{/etc/grid-security/certificates/} . In that case names of certificate
files have to follow hash values of the certificates. These are obtainable by
running the command:

\begin{lstlisting}
openssl x509 -hash -noout -in path_to_certificate
\end{lstlisting}

The corresponding file name for the certificate should be
\texttt{<hash\_value>.0}. The hash value for the pre-generated CA certificate is
\texttt{4457e417}.

Please make sure the chosen location of certificates is correctly configured in
the service configuration file. The configuration for the certificate for TLS
MCC should look like this:

\begin{lstlisting}
<KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
<CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
<CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
\end{lstlisting}

or

\begin{lstlisting}
<CACertificatePath>/etc/grid-security/ca.pem</CACertificatePath>
\end{lstlisting}

The same requirements are valid for the client tools for ARC1. You may use the
pre-generated user certificate and key located at the same place. Locations of
the credentials are configurable, or can be provided to the client tools from
the command line.

The set of pre-generated keys and certificates also includes a user certificate
in PKCS12 format which you can import into your browser for accessing ARC1
services capable of producing HTML output.

ARC comes with utility \texttt{approxy} which generates proxy credentials from
certificate/private key pair. It provides only basic functionality and is meant
for testing purposes only.

\begin{framed}
\textbf{IMPORTANT}: If during configuration stage you see a message

 \textit{``OpenSSL contains no support for proxy credentials''}

then you won't be able to use proxy credentials generated by utilities
like \texttt{grid-proxy-init}, \texttt{voms-proxy-init} or \texttt{approxy}.
Because of that all user private keys will have to be kept unencrypted.
\end{framed}

To avoid providing credential information on command line it is possible to have
user configuration file with predefined values. Default location for this file
is \texttt{~/.arc/client.xml}. There is an example installed in
\verb|$ARC_LOCATION/share/doc/arc/client.xml|. Please edit before copying it
into your home directory.

\subsection{Proxy Certificate Generation and Usage}
\label{sec:proxy}

As metioned above, ARC comes with proxy generation utility
-- \texttt{approxy}, installed in \verb|$ARC_LOCATION/bin|. The usage
of \texttt{approxy} is like:

\begin{lstlisting}
$ARC_LOCATION/bin/approxy -P proxy.pem -C cert.pem -K key.pem 
  -c validityStart=2008-05-29T10:20:30Z
  -c validityEnd=2008-06-29T10:20:30Z
  -c proxyPolicyFile=delegation_policy.xml
\end{lstlisting}

By using argument ``-c'', some constraints can be specified for proxy
certificate. Currently, the life-time can be specified by using ``\texttt{-c
validityStart=...}'' and ``\texttt{-c validityEnd=...}'', or ``\texttt{-c
validityStart=...}'' and ``\texttt{-c validityPeriod=...}''; the proxy policy
can be specified by using ``\texttt{-c proxyPolicyFile=...}'' or ``\texttt{-c
proxyPolicy=...}''.

If proxy certificate is used, in the configuration file for service side or
client side, the configuration for the certificate for TLS MCC should look like
this:

\begin{lstlisting}
<KeyPath>./proxy.pem</KeyPath>
<CertificatePath>./proxy.pem</CertificatePath>
<CACertificatePath>./ca.pem</CACertificatePath>
\end{lstlisting}

Because normally a proxy certificate file includes the proxy certificate and
private key corresponding to the proxy certificate, \texttt{<KeyPath/>} and
\texttt{<CertificatePath/>} are configured to be the same.

Proxy policy can be spefified as constraint. Proxy policy is for constraining
identity delegation. Currently, the supported policy is ARC specific policy.
Proxy policy is inserted into proxy certificate's ``proxy cert info'' extenstion
in RFC3820's policy language ``\verb|NID_id_ppl_anyLanguage|''.

\section{ARC Server Setup and Configuration}
\label{sec:server-setup}

The configuration of the ARC server is specified in an XML file, the location of
which is specified as a command line argument with the -c option of
``\texttt{arched}'' daemon. Examples of configuration files with comments
describing various elements are available in directory
\texttt{/usr/share/doc/nordugrid-arc<version>} corresponding to the ARC1
installation.

\section{The Echo Service}
\label{sec:echo-service}

The \textit{Echo} service is offered purely for testing purposes. It is
``atomic'' and has no additional dependencies other than what is provided by the
Hosting Environment Daemon (HED). An example of an Echo service configuration
can be found in \texttt{/usr/share/doc/nordugrid-arc<version>/echo.xml}.

\subsection{The Echo Client}
\label{sec:echo-client}

The configuration of the ARC Echo client is specified in an XML
file. The location of the configuration file is specified by the
environment variable \verb|$ARC_ECHO_CONFIG|. If there is no such
environment
variable, the configuration file is assumed to be \verb|echo_client.xml| in
the current working directory.

To use the Echo client, type

\begin{lstlisting}
$ARC_LOCATION/bin/apecho <message>
\end{lstlisting}

where \verb|<message>| is the message which the Echo service should return.

\section{The A-REX Service}
\label{sec:a-rex}

ARC1 comes with OGSA BES compliant Grid job management service called A-REX.
To deploy A-REX use example configuration files available in
\verb|/usr/share/doc/nordugrid-arc<version>|:

\begin{itemize}
 \item \texttt{arex.xml} -- configuration for arched server. Read comments
inside this file.
 \item \texttt{arc-arex.conf} -- legacy configuration for Grid Manager part of
A-REX. This
file defines how jobs are managed by A-REX locally. Read and edit it.
and edit it to fit your installation. This file defines WS interface of A-REX.
\end{itemize}

For more detailed information please read Grid Manager
documentation~\cite{arex}. Grid Manager runs as part of A-REX service. There is
no need to run any additional executable. But you still need to setup its
infrastructure as long as you are going to have anything more sophisticated than
described in the example configuration. For more information read the previously
mentioned document.

Currently, a proper functioning A-REX requires environment variable
\verb|$ARC_LOCATION| to be set to the installation prefix of ARC.

A-REX uses HTTPS as transport protocol (although one can reconfigure it to use
plain HTTP) so it requires proper setup of X509 security infrastructure. See
Section~\ref{sec:x509} for instructions.

Copy example configuration files to some location and edit them. Make sure all
paths to X509 certificates and Grid Manager configuration are set correctly.
Start server with command:

\begin{lstlisting}
$ARC_LOCATION/sbin/arched -c path_to_edited_arex.xml
\end{lstlisting}

Look into log file specified in \texttt{arex.xml} for possible errors. You can
safely ignore messages like \textit{``Not a '...' type plugin''} and
\textit{``Unknown element ... - ignoring''}.

If you compiled ARC1 services with Globus support and you see complaints about
``libglobus...'' and that it cannot open a shared object file, try to add
\texttt{/opt/globus/lib} to your \verb|$LD_LIBRARY_PATH|:

\begin{lstlisting}
export LD_LIBRARY_PATH=$LD_LIBRARY_PATH:/opt/globus/lib
\end{lstlisting}

\subsection{Testing and Using A-REX (clients)}
\label{sec:testing}

Now you may use the command line utility ``\texttt{apinfo}'' to obtain a service
description. Make sure user key and certificate and CA certificate(s) are
readable from account you use, and run something like:

\begin{lstlisting}
./apinfo -d VERBOSE -C $HOME/usercert.pem -K $HOME/userkey.pem 
 -A /etc/grid-security/certificates https://localhost:60000/arex 2>error.log
\end{lstlisting}

This should produce XML output describing resource A-REX represents.
Listing~\ref{lst:apinfo} shows an example of proper output. In case of error see
\texttt{error.log} file and report problems to ARC development team. Include
content of \texttt{error.log} and server log.

\begin{lstlisting}[caption={Example of a good apinfo output},label=lst:apinfo]
<n:nordugrid xmlns:n="urn:nordugrid" xmlns:M0="urn:Mds"
xmlns:nc0="urn:nordugrid-cluster">
  <nc0:name>mobile2.home.net</nc0:name>
  <nc0:aliasname>MINIMAL Computing Element</nc0:aliasname>
  <nc0:comment>This is a minimal out-of-box CE setup</nc0:comment>
  <nc0:issuerca>/O=Grid/O=Test/CN=CA</nc0:issuerca>
  <nc0:issuerca-hash>4457e417</nc0:issuerca-hash>
  <nc0:trustedca>/O=Grid/O=NorduGrid/CN=NorduGrid Certification
Authority</nc0:trustedca>
  <nc0:trustedca>/O=Grid/O=Test/CN=CA</nc0:trustedca>
  <nc0:contactstring>gsiftp://mobile2.home.net:2811/jobs</nc0:contactstring>
  <nc0:lrms-type>fork</nc0:lrms-type>
  <nc0:architecture>i686</nc0:architecture>
  <nc0:homogeneity>TRUE</nc0:homogeneity>
  <nc0:nodeaccess>inbound</nc0:nodeaccess>
  <nc0:nodeaccess>outbound</nc0:nodeaccess>
  <nc0:totalcpus>1</nc0:totalcpus>
  <nc0:usedcpus>0</nc0:usedcpus>
  <nc0:cpudistribution>1cpu:1</nc0:cpudistribution>
  <nc0:prelrmsqueued>0</nc0:prelrmsqueued>
  <nc0:totaljobs>0</nc0:totaljobs>
  <nc0:sessiondir-free>8469</nc0:sessiondir-free>
  <nc0:sessiondir-total>43865</nc0:sessiondir-total>
  <nc0:sessiondir-lifetime>10080</nc0:sessiondir-lifetime>
  <M0:validfrom>20080220142836Z</M0:validfrom>
  <M0:validto>20080220143036Z</M0:validto>
  <nq0:name xmlns:nq0="urn:nordugrid-queue" name="fork">
    <nq0:name>fork</nq0:name>
    <nq0:status>inactive, grid-manager is down</nq0:status>
    <nq0:comment>This queue is nothing more than a fork host</nq0:comment>
    <nq0:schedulingpolicy>FIFO</nq0:schedulingpolicy>
    <nq0:homogeneity>TRUE</nq0:homogeneity>
    <nq0:nodecpu>Intel(R) Celeron(R) M CPU        420  @ 1.60GHz @ 1596.048
MHz</nq0:nodecpu>
    <nq0:architecture>i686</nq0:architecture>
    <nq0:maxcputime>unlimited</nq0:maxcputime>
    <nq0:maxwalltime>unlimited</nq0:maxwalltime>
    <nq0:running>0</nq0:running>
    <nq0:gridrunning>0</nq0:gridrunning>
    <nq0:localqueued>0</nq0:localqueued>
    <nq0:gridqueued>0</nq0:gridqueued>
    <nq0:prelrmsqueued>0</nq0:prelrmsqueued>
    <nq0:totalcpus>1</nq0:totalcpus>
    <IM0:validfrom>20080220142836Z</M0:validfrom>
    <M0:validto>20080220143036Z</M0:validto>
    <nig0:name xmlns:nig0="urn:nordugrid-info-group" name="jobs">
      <nig0:name>jobs</nig0:name>
      <M0:validfrom>20080220142836Z</M0:validfrom>
      <M0:validto>20080220143036Z</M0:validto>
      <nj0:globalid xmlns:nj0="urn:nordugrid-job"
name="gsiftp://mobile2.home.net:2811/jobs/636312034687631804
289383">

<nj0:globalid>gsiftp://mobile2.home.net:2811/jobs/636312034687631804289383</nj0:
globalid>
        <nj0:globalowner>/O=Grid/O=Test/CN=user</nj0:globalowner>
        <nj0:jobname>TEST</nj0:jobname>
        <nj0:execcluster>mobile2.home.net</nj0:execcluster>
        <nj0:execqueue>fork</nj0:execqueue>
        <nj0:submissionui>127.0.0.1:59790</nj0:submissionui>
        <nj0:submissiontime>20080220005243Z</nj0:submissiontime>
        <nj0:sessiondirerasetime>20080227005843Z</nj0:sessiondirerasetime>
        <nj0:completiontime>20080220005843Z</nj0:completiontime>
        <nj0:cpucount>1</nj0:cpucount>
        <nj0:comment>GM: The grid-manager is down</nj0:comment>
        <nj0:usedcputime>0</nj0:usedcputime>
        <nj0:usedwalltime>0</nj0:usedwalltime>
        <nj0:errors>Job submission to LRMS failed</nj0:errors>
        <nj0:status>FAILED</nj0:status>
        <nj0:rerunable>SUBMIT</nj0:rerunable>
        <M0:validfrom>20080220142836Z</M0:validfrom>
        <M0:validto>20080220143036Z</M0:validto>
      </nj0:globalid>
    </nig0:name>
    <nig0:name xmlns:nig0="urn:nordugrid-info-group" name="users">
      <nig0:name>users</nig0:name>
      <M0:validfrom>20080220142836Z</M0:validfrom>
      <M0:validto>20080220143036Z</M0:validto>
    </nig0:name>
  </nq0:name>
</n:nordugrid>
\end{lstlisting}

Please note that you can run similar \texttt{apinfo} request against any ARC1
service except Echo service. Output should always be XML description of service.

A-REX accepts jobs described in JSDL language. Example JSDL jobs are provided in
\verb|$ARC_LOCATION/share/doc/| in files \verb|jsdl_simple.xml| and
\verb|jsdl_stage.xml|. To submit job to A-REX service one may use
\texttt{apsub} command:

\begin{lstlisting}
$ARC_LOCATION/bin/apsub https://localhost:60000/arex
 /usr/local/share/doc/arc/jsdl_simple.xml id.xml
\end{lstlisting}

If everything goes properly somewhere in it's output there should be a message
``Submitted the job!''. Obtained job identifier is an XML document and is stored
in id.xml file. It may then be used to query job state with \texttt{apstat}
utility:

\begin{lstlisting}
$ARC_LOCATION/bin/apstat id.xml 2>/dev/null
Job status: Running/Submiting

$ARC_LOCATION/bin/apstat id.xml 2>/dev/null
Job status: Running/Finishing

$ARC_LOCATION/bin/apstat id.xml 2>/dev/null
Job status: Finished/Finished
\end{lstlisting}

Full set of A-REX client tools consists of \texttt{apsub}, \texttt{apstat},
\texttt{apkill}, \texttt{asstat}, \texttt{apget} and \texttt{apclean} commands.
For more information please see man pages ot those utilities.

\section{Finding more information}
\label{sec:more-info}

Many information about functionality and configuration of various components
may be found inside corresponding configuratrion XML schemas.

There are several API documents available as well.

\section{Contributing}
\label{sec:contributing}

The open source development of the ARC middleware is coordinated by the
NorduGrid Collaboration. Currently, the main contributor is the KnowARC project
(\url{www.knowarc.eu}), but the Collaboration is open to new members.
Contributions from the community to the software and the documentation is
welcomed. Sources can be downloaded from the software repository at
\url{download.nordugrid.org} or the Subversion code repository at
\url{svn.nordugrid.org}.

The technical coordination group defines outstanding issues that have
to be addressed in the framework of the ARC development. Feature
requests and enhancement proposals are recorded in the Bugzilla bug
tracking system at \url{bugzilla.nordugrid.org}. For a more detailed
description, write access to the code repository and further
questions, write to the \texttt{nordugrid-discuss} mailing list (see
\url{www.nordugrid.org} for details). Ongoing and completed Grid Research
projects and student assignments related to the middleware are listed
on the NorduGrid Web site as well.

\section{Support, documentation, mailing lists, contact}
\label{sec:support}

User support and site installation assistance is provided via the
request tracking system available at \texttt{nordugrid-support@nordugrid.org}.
In addition, NorduGrid runs a couple of mailing lists, among which the
nordugrid-discuss mailing list is a general forum for all kind of
issues related to the ARC middleware.

Feature and enhancement requests, as well as discovered problems, should be
reported in the Bugzilla problem tracking system \url{bugzilla.nordugrid.org}.

Research papers, overview talks, reference manuals, user guides,
installation instructions, conference presentations, FAQ and even
tutorial materials can be fetched from the documentation section of
\url{www.nordugrid.org}.

Contact information is kept updated on the \url{www.nordugrid.org} Web site.

\bibliography{ws-quick-guide}
\end{document}

### 2008-07-28 doc/storage_system/arc1-storage-documentation.tex
modifications based on the first internal review

@@
 \chapter{Design Overview}
-\label{sec:overview}
+\label{cha:overview}
 
 The ARC1 storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.
-The files can be grouped into \emph{collection}s, and a collection can contain sub-collections and sub-sub-collections in any depth.
+The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems), and a collection can contain sub-collections and sub-sub-collections in any depth.
 There is a dedicated \emph{root collection} to gather all collections to the global namespace.
@@
 The replicas of the files are stored on different storage nodes.
-A storage node should have one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.), and one of the services of the ARC1 storage system is needed to manage each storage node and to integrate it into the system.
+A storage node here is a network-accessible computer having storage space to share, and one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.).
+For each storage node one of the services of the ARC1 storage system is needed to manage it and to integrate it into the system.
 And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system.
-The main services are shown on Figure~\ref{fig:services}.
+The main services of the storage system are the following (see Figure~\ref{fig:services}):
+\begin{itemize}
+    \item the \textbf{A-Hash} service, which is a currently centralized, later distributed database;
+    \item the \textbf{Librarian} service, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
+    \item the \textbf{Bartender} service, which provides a high-level interface for the users and for other services;
+    \item the \textbf{Shepherd} service, which provides a simple interface for storing files on storage nodes.
+\end{itemize}
 
@@
 \centering{{\scalebox{0.9}{\includegraphics{arc1-storage-services.pdf}}}
-\caption{\label{fig:services}The components of the ARC1 storage;
-\emph{A-H}: the \textbf{A-Hash}, which is a currently centralized, later distributed database;
-\emph{L}: the \textbf{Librarian}, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
-\emph{B}: the \textbf{Bartender}, which provides a high-level interface for the users and for other services;
-\emph{S}: the \textbf{Shepherd}, which provides a simple interface for storing files on storage nodes.
-} }
+\caption{\label{fig:services}The components of the ARC1 storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
 \end{figure}
+
+The descriptions in Chapter~\ref{cha:overview} and Chapter~\ref{cha:use_cases} are containing some elements which are not implemented yet in the first prototype.
+See Chapter~\ref{cha:technical_description_and_implementation_status} for the details about what is implemented and what is missing.
 
@@
 
-In a collection each entry has a name, and this entry can be a sub-collection, in which there are files and sub-sub-collections, etc.
+A collection is basically just a list of entries, where each entry has a name unique within this collection, and an entry can be a file or a sub-collection, in which there are files and sub-sub-collections, etc.
 
@@
 
-Each VO\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} should create a VO-wide root collection, and put it in the generic root collection, e.g.~if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#.
+One possible usage scenario would be if each VO (Virtual Organization)\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} would create a VO-wide root collection, and put it in the generic root collection, e.g.~if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#.
 Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g.~\verb#/vo1/file1#, etc.
@@
 \label{sec:the_bartenders}
-Clients can access the storage system through a Bartender service.
-If a client wants to create a collection, upload or download a file, the first step is to connect to a Bartender.
-The Bartender then asks a Librarian to resolve Logical Names and return metadata, then initiates file transfers on some storage node by asking the Shepherd service of that node, then returns the transfer URL (\emph{TURL}) to the client, which allows the client to directly transfer from/to the storage node.
-So the data transfer itself is not going through the Bartender, it is performed over a direct link between a storage node and the client.
 
+The Bartender service provides a high-level interface for the storage system to the clients.
+You can create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names.
+The Bartender communicates with the Librarian and Shepherd services to accomplish the client’s requests.
+The actual file data does not go through the Bartender; file transfers are directly performed between the storage nodes and the clients.
 There could be any number of independent Bartender services in the system which provides high-availability and load-balancing.
@@
 The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag `removing a replica' as well.
-\item If we cannot upload the file to the given TURL for some reason, we should remove the file entry from the collection, or we should call \verb#addReplica# to get a new TURL without removing and recreating the file.
+\item If we cannot upload the file to the given TURL for some reason, we could call \verb#addReplica# to get a new TURL without removing and recreating the file.
 \end{itemize}
@@
 
-In both cases the Shepherd service uses two separate directories: one for storing all the files (e.g.~\verb!./store!) and one for the file transfers (e.g.~\verb!./transfer!).
+In both cases the Shepherd and the transfer services should have access to the same local filesystem where the Shepherd creates two separate directories: one for storing all the files (e.g.~\verb!./store!) and one for the file transfers (e.g.~\verb!./transfer!).
 The store directory always contains all the files the Shepherd manages, the transfer directory is empty at the beginning.

### 2008-07-29 doc/storage_system/arc1-storage-documentation.tex
small changes based on the last comments of Aleksandr

@@
 The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag `removing a replica' as well.
-\item If we cannot upload the file to the given TURL for some reason, we could call \verb#addReplica# to get a new TURL without removing and recreating the file.
+\item If the client cannot upload the file to the given TURL for some reason, it is possible to call \verb#addReplica# to get a new TURL without removing and recreating the file.
 \end{itemize}
@@
 If the properties and the policies are present, the decision is actually made by the security framework of HED.
+
+Some self-healing mechanisms are not implemented yet, e.g.
+the files and collections contains their parent-collections as their metadata, and of course a collection contains a list of its files and sub-collections, and if somehow this information became inconsistent, the system should detect and repair it.
+This is not implemented yet.
 

### 2008-07-30 doc/backend_interface/Backends.tex
Back-ends interface documentation - Initial version

%% LyX 1.4.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside,english]{article}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{amsmath}
\usepackage{amssymb}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {C.U.Soettrup},
  pdftitle ={Arc backend infrastructure},
  pdfsubject = {Description and developer guide},
  pdfkeywords = {ARC,NorduGrid,manual,Backend,LRMS, Infosystem},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%

\usepackage{babel}
\makeatother
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}\resizebox*{3cm}{!}{\includegraphics[width=3cm,height=2cm]{ng-logo.png}} &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}} \end{tabular}

\hrulefill

{\raggedleft NORDUGRID-TECH-13\par}

{\raggedleft \today\par}

\vspace*{2cm}

\begin{center}
\textsc{\Large Arc batch system back-end interface guide with support for GLUE 2}
\par\end{center}{\Large \par}

\vspace*{0.5cm}

\begin{center}
\textit{\large Description and developer's guide}
\par\end{center}{\large \par}

\vspace*{1.5cm}

{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomasf@fys.uio.no}, C.U.Søttrup, Bal\'azs K\'onya, G\'abor R\H oczei \par}

\end{titlepage}

\thispagestyle{empty}

$ $

\newpage

\tableofcontents{}

\newpage

\section{Introduction\label{sec:intro}}

This document describes the Advanced Resource Connector\cite{ARC}
batch system back-end infrastrucure. It aims to describe the infrastructure
in enough detail that a developer can add a new local resource management
system (LRMS) to the ARC middleware. It also intends to serve as a
reference manual describing existing batch system interfaces. Note
that the current solution described in this manual is not a real interface,
it is too diffuse to be that.

The batch system back-ends are what tie the ARC grid middleware (through
the Grid Manager (GM\cite{GM}) and the Information System layer\cite{INFOSYS})
to the underlying cluster management system or LRMS. The back-ends
consist of set of shell and Perl scripts whose role are twofold:

\begin{enumerate}
\item to allow the GM to manipulate jobs on the LRMS including job submit,
cancel operations and scanning for completed jobs
\item to collect information about jobs, users, batch system and the cluster
itself for the Information System. 
\end{enumerate}
The former will be referred to as the job control back-end interface
while the latter is the information system interface of the batch
system back-ends. These two will be treated separately in the following
sections.

\subsection{ARC Classic}
As of ARC version 0.6 the the two sets of scripts are located in different
directories within the NorduGrid subversion tree\cite{subversion}.
The job control interface scripts can be found under \texttt{arc\_development\_line/grid-manager/LRMS/}
while the information collectors are located at \texttt{arc\_development\_line/infosystem/.}

Section 2, 3, and 4 gives a description of the classic ARC back-ends scripts.

\subsection{ARC1}
As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories within the NorduGrid subversion tree\cite{subversion} as well. 
The architecture of ARC1 is different from Classic ARC so the job control interface scripts can be found under
\texttt{arc1/trunk/src/services/a-rex/lrms/}
while the information collectors are located in \texttt{arc1/trunk/src/services/a-rex/infoproviders/}.

Furthermore, the next generation of ARC supports both the classic NorduGrid information schema\cite{INFOSYS} and
currently a minimal set of the GLUE specification version 2.0 schema\cite{GLUE2}.

The ARC1 back-end architecture and GLUE2 information model implementation is presented in Section 5.

\section{Job control interface}

Job control part of the LRMS interface is handled by the Grid Manager\cite{GM}.
It takes care of preparing a native batch system submission script,
managing the actual submission of the batch system job, cancellation
of job on request and scanning for completed batch jobs. Besides the
LRMS job control interface it is also the GM which provides e.g. the
data staging and communication with the grid client, provides RTE
environments, arranges file staging (to the node via LRMS capability),
dealing with stdout/stderr, etc. The job control batch system interface
of the GM requires three programs. These programs can be implemented
any way the designer sees fit, but all the existing back-end interfaces
use shell scripting for portability and ease of tailoring to a specific
site. The GM will call the following programs: cancel-LRMS-job, submit-LRMS-job,
and scan-LRMS-job. LRMS is replaced with the short hand name for the
LRMS e.g. cancel-pbs-job. The scripts are described one by one in
the following subsections. Useful information can also be found in
the Section \char`\"{}8.6 LRMS Support\char`\"{} and Section \char`\"{}8.7
Runtime Environment\char`\"{} of the Grid-Manager guide\cite{GM}.

\subsection{Cancel-LRMS-job}

If a grid user cancels his job, the message will reach the grid-manager.
The manager will then call the cancel-LRMS-job for the suitable back-end.
The cancel script is called with a text file containing information
about the job. It is called the GRAMi file. It contains such information
as the original xRSL option and the job id in the local LRMS. Cancel-LRMS-job
must then use that information to find the job and remove it from
the queue or actually cancel it if it is running in the LRMS.

\subsection{Submit-LRMS-job}

The submit program is the most involved. It is called by the GM once
a new job arrives and needs to be submitted to the LRMS. Like cancel-LRMS-job
it is given the grami file as argument on execution. Submit-LRMS-job
then has to set up the session directories, run-time environment and
anything else needed. Then it should submit the job to the local LRMS.
This is normally done by generating a native job script for the LRMS
and then running the local submit command, but it could also be done
through an API if the LRMS supports it.

\subsection{Scan-LRMS-job}

Scan-LRMS-job is run periodically. Its job is to scan the LRMS for
jobs that have finished. Once it has found a finished job it should
write the exit-code of that job to the file job.\{gridid\}.lrms\_done
in the ARC jobstatus directory. Then it should call the gm-kick program
the GM will then notice that the job has finished and start finalizing
the job.

Generally two approaches are taken to find jobs that are done. One
is asking the LRMS. First all started grid jobs are found in the jobstatus
directory%
\footnote{normally /var/spool/nordugrid/jobstatus/, but can be set via the controldir
variable of arc.conf%
}. This can be done by checking if the status is INLRMS in the jobs
status file%
\footnote{job.\{gridid\}.status%
}. Then the LRMS is asked for the status of those jobs. If they are
done they are marked as such and the GM is activated. The problem
with this approach is that for most LRMSs the information about finished
jobs are only available for a short period after the job finished.
Therefore appropriate steps have to be taken if the job is known to
have been started but is no longer present in the LRMS. The normal
approach is to analyze the jobs status output in the session directory.

The second approach is to parse the LRMSs log files. This method has
some drawbacks e.g. the GM has to be allowed read access to the logs.
The back-end will then have to remember where in the log it was last
time it ran. This information will have to be stored in a file some
where on the front-end.

\subsection{Configuration parser}

Some of the back-ends will need information from the configuration
file. Since this functionality can be shared  among the back-ends,
a configuration file parser written in bash has been provided separately
for easy maintenance and extendability.

\subsubsection{Functions}

\texttt{config\_parse\_file <config\_file>}

\begin{quote}
Parses a config file. It returns exit status 1 if the file cannot
be read, or 0 otherwise. Badly formed lines are silently ignored.
Option values can be surrounded by sigle quotes or double quotes.
Values without quotes are also accepted. Multi-valued options are
not supported currently. Only the last defined value is retained. 
\end{quote}
\texttt{config\_update\_from\_section <section\_name>}

\begin{quote}
Imports options from section <section\_name> of the config file into
environment variables of the form 'CONFIG\_optionname'. Already existing
environment variables are overwritten.
\end{quote}
Example:

\texttt{source \$ARC\_LOCATION/libexec/config\_parser.sh}

\texttt{config\_parse\_file /etc/arc.conf || exit 1}

\texttt{config\_update\_from\_section common}

\texttt{config\_update\_from\_section grid-manager}

\texttt{config\_update\_from\_section infosys}

\texttt{echo \$CONFIG\_pbs\_bin\_path }

\section{Information System interface}

The main purpose of the information system batch interface is to populate
the nordugrid information model with locally collected information
obtained from the batch system. Important to recall that the locally
collected information is broader than what the batch-system can offer,
information taken from the grid-layer (mostly grid manager), from
the front-end machine and from the \texttt{arc.conf} configuration
file are also needed and used to populate the nordugrid information
model\cite{INFOSYS}. Below only the batch system related information
collection is covered. Note that the separation of the different sources
is not always that straightforward (e.g. the grid-manager information
has some overlap with the batch system info).

The information system interface consists of set of Perl scripts which
populate the local lDAP database by generating ldif files. Historically
two information provider \char`\"{}frameworks\char`\"{} have been
coexisted. As of ARC version 0.6 the old framework is being phased
out by the new framework. Briefly about the two frameworks:

\begin{itemize}
\item the old framework separated the non-batch system related functionality
into the \texttt{InfosysCluster.pm, InfosysQJU.pm} modules and kept
the batch system specific interface within the \texttt{cluster-condor.pl,
queue+jobs+users-condor.pl} (replace condor with your scheduler name)
files. There are two templates provided for batch system interface
developers: \texttt{cluster-gm.pl, queue+jobs+users-gm.pl}
\item the new framework originally proposed by Juha Lento%
\footnote{the original posting of the new framework proposal: http://staff.csc.fi/juha.lento/grid/rewrite.html%
} has been brought up-to-date by implementing all the features present
in the old framework. The new framework consists of two main scripts,
the \texttt{cluster.pl} and the \texttt{qju.pl}. The non-batch system
functionality is implemented by these script (plus some auxiliary
scripts). The batch system specific interface is moved to LRMS modules
such as \texttt{SGE.pm, PBS.pm, Fork.pm}, a template module \texttt{Template.pm}
is also provided. The main advantage of the new framework is the separation
of output generation from the information collection (this will ease
the migration to XML output).
\end{itemize}
The following subsection describes the LRMS interface part of the
information system as implemented via the new framework. To support
a particular LRMS a dedicated Perl module should be written that implements
the interface presented in the next subsections. Furthermore hooks
should be added to the \texttt{LRMS.pm} module, so that it uses the
correct Perl module depending on the LRMS type. Short description
of the files of the new framework:

\begin{itemize}
\item \texttt{cluster.pl and qju.pl}: main scripts called to populate the
nordugrid information model
\item \texttt{Shared.pm:} subroutines shared by cluster.pl and cluster.pl
\item \texttt{LogUtils.pm}: logging utilities
\item \texttt{LRMS.pm:} defines the LRMS interface (described in the next
subsection)
\item \texttt{Template.pm:} module template for implementing a new LRMS
interface
\item \texttt{SGE.pm, PBS.pm, LSF.pm, Fork.pm, LL.pm, Condor.pm}: actual
LRMS modules
\end{itemize}

\subsection{The LRMS Perl module interface}

The LRMS module should implement four functions as defined by the
\texttt{LRMS.pm}: \texttt{Cluster\_info}, that provides general information
about the local cluster, \texttt{Queue\_info}, that provides information
about a specified queue, \texttt{Jobs\_info} that gives information
about a list of specified jobs, and finally \texttt{user\_info} that
gives information about a grid user mapped to a local UNIX account.
All the functions are called with a hash containing the contents of
arc.conf.

\subsubsection{cluster\_info}

\texttt{Cluster\_info} is called only with the config hash. It should
then return an associative array with the following keys and the appropriate
value.

\begin{tabular}{|c|c|}
\hline 
key&
value\tabularnewline
\hline
\hline 
lrms\_type&
the type of LRMS e.g PBS\tabularnewline
\hline 
lrms\_version&
the version of the LRMS\tabularnewline
\hline 
totalcpus&
total number of CPUs in the cluster\tabularnewline
\hline 
queuedcpus&
total number of CPUs requested by LRMS queuing jobs (both grid and
non-grid)\tabularnewline
\hline 
usedcpus&
CPUs in the LRMS that are currently in use either by grid or non-grid
jobs\tabularnewline
\hline 
cpudistribution&
number of CPUs in a node and number of each type e.g. {}``8cpu:5
2cpu:100''\tabularnewline
\hline
\end{tabular}

\subsubsection{queue\_info}

In addition to the config \texttt{queue\_info} is called with the
name of the queue to get information on. The result should take the
following form.

\begin{tabular}{|c|c|}
\hline 
key&
value\tabularnewline
\hline
\hline 
status&
available slots in queue, negative number signals error\tabularnewline
\hline 
maxrunning&
limit on number of running jobs\tabularnewline
\hline 
maxqueuable&
limit on number of jobs queued on this queue\tabularnewline
\hline 
maxuserrun&
limit on number of running jobs per user\tabularnewline
\hline 
maxcputime&
limit on maximum CPU time for a job in this queue\tabularnewline
\hline 
mincputime&
limit on minimum CPU time for a job in this queue\tabularnewline
\hline 
defaultcput&
default CPU time limit for a job in this queue\tabularnewline
\hline 
maxwalltime&
limit on maximum wall time for a job in this queue\tabularnewline
\hline 
minwalltime&
limit on minimum wall time for a job in this queue\tabularnewline
\hline 
defaultwallt&
default wall time limit for a job in this queue\tabularnewline
\hline 
running&
number of CPUs used by running jobs (both grid and non-grid) in the
queue\tabularnewline
\hline 
queued&
CPUs requested by queuing jobs in the queue\tabularnewline
\hline 
totalcpus&
number of CPUs available to the queue\tabularnewline
\hline
\end{tabular}

\subsubsection{jobs\_info}

Like the rest the first argument is the config hash, then comes the
queue name and finally a list of job ids. Once again an associative
mapping should be made for each job. The key value pairs are:

\begin{tabular}{|c|c|}
\hline 
key&
value\tabularnewline
\hline
\hline 
status&
LRMS jobstatus mapping%
\footnote{LRMS job states mapping is described in\cite{INFOSYS} %
}:Running->'R', Queued->'Q', Suspended->'S', Exiting->'E', Other->'O'\tabularnewline
\hline 
rank&
the job's position in the LRMS queue\tabularnewline
\hline 
mem&
the memory usage of the job in kBs\tabularnewline
\hline 
walltime&
consumed wall-time in minutes\tabularnewline
\hline 
cputime&
consumed CPU-time in minutes\tabularnewline
\hline 
reqwalltime&
wall-time requested by job in minutes\tabularnewline
\hline 
reqcputime&
CPU-time requested by job in minutes\tabularnewline
\hline 
nodes&
list of execution hosts\tabularnewline
\hline 
comment&
array of string comments about the job in LRMS, can be an empty array\tabularnewline
\hline
\end{tabular}

\subsubsection{user\_info}

\texttt{Users\_info} is given the config hash, a queue name and a
list of local UNIX accounts (UIDs). It should then return an associative
array containing for each UID the number of free CPUs available for
that local account on the given queue and the number of queued jobs
on that queue.

\begin{tabular}{|c|c|}
\hline 
key&
value\tabularnewline
\hline
\hline 
freecpus&
number of freely available CPUs with their time limits in minutes
(see nordugrid-authuser-freecpus in\cite{INFOSYS})\tabularnewline
\hline 
queuelength&
estimated queue length for the specified user\tabularnewline
\hline
\end{tabular}

\section{Batch system specifics}

This section presents the batch system specific implementation details
including information on supported versions, constraints on batch
system configuration, known limitations, \texttt{arc.conf} parameters
and list of batch system features being utilized within the interface
are described.

\subsection{PBS}

The Portable Batch System (PBS) is one of the most popular batch systems.
PBS comes in many flavours such as OpenPBS (unsupported), Terascale
Open-Source Resource and QUEue Manager (TORQUE) and PBSPro (currently
owned by Altair Engineering). ARC supports all the flavours and versions
of PBS.

\subsubsection{Recommended batch system configuration}

PBS is a very powerful Local Resource Manager System with dozens of
configurable options. Server, queue and node attributes can be used
to configure the cluster's behaviour. In order to correctly interface
PBS to ARC (mainly the information provider scripts) there are a couple
of configuration REQUIREMENTS asked to be implemented by the local
system administrator:

\begin{enumerate}
\item The computing nodes MUST be declared as cluster nodes (job-exclusive),
at the moment time-shared nodes are not supported by the ARC setup.
If you intend to run more than one job on a single processor then
you can use the virtual processor feature of PBS. 
\item For each queue, you MUST set one of the max\_user\_run or max\_running
attributes and its value SHOULD BE IN AGREEMENT with the number of
available resources (i.e. don't set the max\_running = 10 if you have
only six (virtual) processors in your system). If you set both max\_running
and max\_user\_run then obviously max\_user\_run has to be less equal
than max\_running. 
\item For the time being, do NOT set server limits like max\_running, please
use queue-based limits instead. 
\item Avoid using the max\_load and the ideal\_load directives. The node's
mom config file (<PBS home on the node>/mom\_priv/config) should not
contain any max\_load or ideal\_load directives. PBS closes down a
node (no jobs are allocated to it) when the load on the node reaches
the max\_load value. The max\_load value is meant for controlling
time-shared nodes. In case of job-exclusive nodes there is no need
for setting these directives, moreover incorrectly set values can
close down your node. 
\item Routing queues are not supported, those can't be used within ARC.
\end{enumerate}
Additional useful configuration hints:

\begin{itemize}
\item If possible, please use queue-based attributes instead of server level
ones (for the time being, do not use server level attributes at all). 
\item You may use the \char`\"{}acl\_user\_enable = True\char`\"{} with
\char`\"{}acl\_users = user1,user2\char`\"{} attribute to enable user
access control for the queue. 
\item It is advisory to set the max\_queuable attribute in order to avoid
a painfully long dead queue. 
\item You can use node properties from the \texttt{<PBS home on the server>/server\_priv/nodes}
file together with the \texttt{resources\_default.neednodes} to assign
a queue to a certain type of node.
\end{itemize}
Checking your PBS configuration:

\begin{itemize}
\item The node definition can be checked by \texttt{<PBS installation path>/bin/pbsnodes
-a.} All the nodes MUST have ntype=cluster. 
\item The required queue attributes can be checked as \texttt{<PBS installation
path>/bin/qstat -f -Q queuename.} There MUST be a max\_user\_run or
a max\_running attribute listed with a REASONABLE value. 
\end{itemize}

\subsubsection{Relevant \texttt{arc.conf} options}

The ARC configuration manual contains detailed information on all
the ARC configuration parameters\cite{config}and should be consulted
as primary source. Below the PBS specific variables are collected.

\begin{itemize}
\item The PBS batch system back-end is enabled via setting the \texttt{lrms=\char`\"{}pbs\char`\"{}}
in the \texttt{{[}common]} configuration block. No need to specify
the flavour or the version number of the PBS, simply use the \texttt{{}``pbs''}
keyword as lrms configuration value.
\item \texttt{pbs\_bin\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path to the qstat,pbsnodes,qmgr etc PBS
binaries.
\item \texttt{pbs\_log\_path} configuration variable of the \texttt{{[}common]}
block should be set to the path of the PBS server logfiles which are
used by the Grid-Manager to determine whether a PBS job is completed.
If not specified, the Grid-Manager will use the \texttt{qstat} command
to find completed jobs. 
\item \texttt{lrmsconfig} from the \texttt{{[}cluster]} block can be used
as an optional free text field to describe further details about the
PBS configuration (e.g. \texttt{lrmsconfig=\char`\"{}single job per
processor\char`\"{}}).
\item \texttt{dedicated\_node\_string} from the \texttt{{[}cluster]} block
specifies the string which is used in the PBS node config to distinguish
the grid nodes from the rest. Suppose only a subset of nodes are available
for grid jobs, and these nodes have a common \texttt{node property}
string, this case the \texttt{dedicated\_node\_string} should be set
to this value and only the nodes with the corresponding PBS \texttt{node
property} are counted as grid enabled nodes. Setting the \texttt{dedicated\_node\_string}
to the value of the PBS \texttt{node property} of the grid-enabled
nodes will influence how the totalcpus, user freecpus is calculated.
No need to set this attribute if the cluster is fully available for
the grid and the PBS configuration does not use the \texttt{node property}
method to assign certain nodes to grid queues.
\item \texttt{{[}queue/queuename]} block. For each grid-enabled (or grid
visible) PBS queue a corresponding \texttt{{[}queue]} block must be
defined. \texttt{queuename} should be the PBS queue name.
\item \texttt{scheduling\_policy} from the \texttt{{[}queue/queuename]}
block describes the scheduling policy of the queue. PBS by default
offers the FIFO scheduler, many sites run the MAUI. At the moment
\texttt{FIFO} \& \texttt{MAUI} are supported values. If you have a
MAUI scheduler you should specify the \char`\"{}MAUI\char`\"{} value
since it modifies the way the queue resources are calculated. By default
the \char`\"{}FIFO\char`\"{} scheduler type is assumed.
\item \texttt{maui\_bin\_path} from the \texttt{{[}queue/queuename]} block
sets the path of the maui commands like \texttt{showbf} when \char`\"{}MAUI\char`\"{}
is specified as \texttt{scheduling\_policy} value. This parameter
can be set in the \texttt{{[}common]} block as well. 
\item \texttt{queue\_node\_string} of the \texttt{{[}queue/queuename]} block
can be used similar to the \texttt{dedicated\_node\_string}. In PBS
you can assign nodes to a queue (or a queue to nodes) by using the
\texttt{node property} PBS node configuration method and assigning
the marked nodes to the queue (setting the \texttt{resources\_default.neednodes}
= \texttt{queue\_node\_string} for that queue). This parameter should
contain the \texttt{node property} string of the queue-assigned nodes.
Setting the \texttt{queue\_node\_string} changes how the queue-totalcpus,
user freecpus are determined for this queue. 
\end{itemize}

\subsubsection{Implementation details}

The job control batch interface makes use of the \texttt{qsub} command
to submit native PBS jobscripts to the batch system. The following
options are used:

-l nodes, cput, walltime, pvmem, pmem,

-W stagein, stageout

-e, -j eo

-q

-A

-N

For job cancellation the \texttt{qdel} command is used. To find completed
jobs, i.e. to scan for finished jobs the \texttt{qstat} command or
the \texttt{PBS server log file} is used.

The information system interface utilizes the \texttt{qstat -f -Q
queuename} and \texttt{qstat -f queuename} commands to obtain detailed
job and queue information. \texttt{qmgr -c \char`\"{}list server\char`\"{}}
is used to determine PBS flavour and version. The \texttt{pbsnodes}
command is used to calculate total/used/free cpus within the cluster.
In case of a Maui scheduller the \texttt{showbf} command is used to
determine user freecpu values. All these external PBS commands are
interfaced via parsing the commands' output.

\subsubsection{Known limitations}

Some of the limitations are already mentioned under the PBS deployment
requirements. No support for routing queues, difficulty of treating
overlapping queues, the complexity of node string specifications for
parallel jobs are the main shortcomings.

\subsection{Condor}

The Condor \cite{Condor} system, developed at the University of Wisconsin-Madison,
was initially used to harness free cpu cycles of workstations. Over
time it has evolved into a complex system with many grid-oriented
features. Condor is available on a large variety of platforms.

\subsubsection{Recommended batch system configuration}

Install Condor on the Grid Manager (GM) node and configure it as a
submit machine. Next, you must add the following to the node's Condor
configuration (\texttt{CONDOR\_IDS} can also be an environment variable):

\texttt{MAIL = <ARC\_install\_prefix>/libexec/finish-condor-job}

\texttt{CONDOR\_IDS = 0.0}

The \texttt{MAIL} attribute will instruct Condor to run the specified
program on job completion. The default on Condor is to run /bin/mail
to notify the user, but in this case, it is the GM that needs the
notification. Therefore, /bin/mail is replaced with a program especially
written for talking to the GM.

\texttt{CONDOR\_IDS} has to be 0.0, so that the above notification
program can access the Grid job's session directories (needed to extract
the job exit code from the Condor log).

Make sure that no normal users are allowed to submit Condor jobs from
this node. For one thing, it would not work for the user, since Condor
will try to notify the GM instead of the job owner on job completion.
If you don't allow normal user logins on the GM machine, then you
don't have to do anything. If you for some reason want to allow users
to log into the GM machine, simply don't allow them to execute the
condor\_submit program. This can be done by putting all local Unix
users allocated to the Grid in a single group, e.g. 'griduser', and
then setting the file ownership and permissions on condor\_submit
like this:

\texttt{chgrp griduser \$condor\_location/bin/condor\_submit}

\texttt{chmod 750 \$condor\_location/bin/condor\_submit }

\subsubsection{Relevant arc.conf options}

\begin{itemize}
\item The Condor batch system back-end is enabled by setting \texttt{lrms=\char`\"{}condor\char`\"{}}
in the \texttt{{[}common]} configuration block. 
\item \texttt{condor\_location} configuration variable of the \texttt{{[}common]}
block should be set to the Condor install prefix (i.e., the directory
containing Condor's bin, sbin, etc).
\item \texttt{condor\_config} configuration variable of the \texttt{{[}common]}
block should be set to the value the environment variable \texttt{CONDOR\_CONFIG}
should have (but don't try to use the environment variable directly
as \texttt{\$CONDOR\_CONFIG}, since it will probably not be defined
when arc.conf is parsed!)
\item \texttt{condor\_rank} configuration variable of the \texttt{{[}common]}
block, if defined, will cause the \texttt{Rank} attribute to be set
in each job description submitted to Condor. Use this option if you
are not happy with the way Condor picks out nodes when running jobs
and want to define your own ranking algorithm. \texttt{condor\_rank}
should be set to a ClassAd float expression that you could use in
the \texttt{Rank} attribute in a Condor job description. For example:
\end{itemize}
\begin{quote}
\texttt{condor\_rank=\char`\"{}(1-LoadAvg/2){*}(1-LoadAvg/2){*}Memory/1000{*}KFlops/1000000\char`\"{} }
\end{quote}
\begin{itemize}
\item \texttt{condor\_requirements} configuration variable of the \texttt{{[}queue]}
block defined a subpool of condor nodes. See next section for details.
\end{itemize}

\subsubsection{Implementation details}

The job control part of the interface uses the \texttt{condor\_submit}
command to submit jobs. Some of the options used in the job's ClassAd
are:

\texttt{Requirements} -- is used to select the nodes that may run
the job. This is how ARC queues are implemented for Condor.

\texttt{Periodic\_remove} -- is used to enforce cputime and walltime
limits.

\texttt{Log} -- the job's condor log file is parsed by the information
scripts to find out whether the job was suspended.

The information system component uses the following Condor commands:

\texttt{condor\_status -long} -- for collecting information about
nodes

\texttt{condor\_status -format \char`\"{}\%s\textbackslash{}n\char`\"{}
Machine -constraint '...'} -- for listing nodes that make up an ARC
queue.

\texttt{condor\_q -long -global} -- for monitoring running jobs.

\texttt{condor\_history -l jobid} -- for collecting information about
finished jobs. Further cues are taken from the job's condor log file
and the body of the email sent by Condor when a job completes.

\subsubsection{Configuring Queues}

Condor does not support queues in the classical sense. It is possible,
however, to divide the Condor pool in several sub-pools. An ARC 'queue'
is then nothing more than a subset of nodes from the Condor pool.

Which nodes go into which queue is defined using the \texttt{condor\_requirements}
configuration option in the corresponding \texttt{{[}queue]} section.
It's value must be a well-formed constraint string that is accepted
by a \texttt{condor\_status -constraint '...'} command. Internally,
this constraint string is used to determine the list of nodes belonging
to a queue. This string can get quite long, so, for readability reasons
it is allowed to split it up into pieces by using multiple \texttt{condor\_requirements}
options. The full constrains string will be reconstructed by concatenating
all pieces.

Queues should be defined in such a way that their nodes all match
the information available in ARC about the queue. A good start is
for the \texttt{condor\_requirements} attribute to contain restrictions
on the following: \texttt{Opsys}, \texttt{Arch}, \texttt{Memory} and
\texttt{Disk}. If you wish to configure more than one queue, it's
good to have queues defined in such a way that they do not overlap.
In the following example disjoint memory ranges are used to ensure
this:

\begin{quote}
\texttt{{[}queue/large] condor\_requirements=\char`\"{}(Opsys == \char`\"{}linux\char`\"{}
\&\& (Arch == \char`\"{}intel\char`\"{} || Arch == \char`\"{}x86\_64\char`\"{})\char`\"{}
condor\_requirements=\char`\"{} \&\& (Disk > 30000000 \&\& Memory
> 2000)\char`\"{}}

\texttt{{[}queue/small] condor\_requirements=\char`\"{}(Opsys == \char`\"{}linux\char`\"{}
\&\& (Arch == \char`\"{}intel\char`\"{} || Arch == \char`\"{}x86\_64\char`\"{})\char`\"{}
condor\_requirements=\char`\"{} \&\& (Disk > 30000000 \&\& Memory
<= 2000 \&\& Memory > 1000)\char`\"{}}
\end{quote}
Note that '\texttt{nodememory}' attribute in arc.conf means the maximum
memory available for jobs, while the Memory attribute in Condor is
the physical memory of the machine. To avoid swapping (and these are
probably not dedicated machines!), make sure that '\texttt{nodememory}'
is smaller than the minimum physical memory of the machines in that
queue. If for example the smallest node in a queue has 1Gb memory,
then it would be sensible to use \texttt{nodememory=\char`\"{}850\char`\"{}}
for the maximum job size.

In case you want more precise control over which nodes are available
for grid jobs, using pre-defined ClassAds attributes (like in the
example above) might not be sufficient. Fortunately, it's possible
to mark nodes by using some custom attribute, say \texttt{NORDUGRID\_RESOURCE}.
This is accomplished by adding a parameter to the node's local Condor
configuration file, and then adding that parameter to \texttt{STARTD\_EXPRS}:

\begin{quote}
\texttt{NORDUGRID\_RESOURCE = True }

\texttt{STARTD\_EXPRS = NORDUGRID\_RESOURCE, \$(STARTD\_EXPRS)}
\end{quote}
Now queues can be restricted to contain only 'good' nodes. Just add
to each \texttt{{[}queue]} section in arc.conf:

\begin{quote}
\texttt{condor\_requirements=\char`\"{} \&\& NORDUGRID\_RESOURCE\char`\"{} }
\end{quote}

\subsubsection{Known limitations}

Only Vanilla universe is supported. MPI universe (for multi-CPU jobs)
is not supported. Neither is Java universe (for running Java executables).
ARC can only send jobs to Linux machines in the Condor pool, therefore
excluding other unixes and Windows destinations. The session directory
must be on a network shared directory, visible from all worker nodes.

\subsection{LoadLeveler}

LoadLeveler(LL), or Tivoli Workload Scheduler LoadLeveler in full,
is a parallel job scheduling system developed by IBM.

\subsubsection{Recommended batch system configuration}

The back-end should work fine with a standard installation of LoadLeveler.
For the back-end to report the correct memory usage and cputime spent,
while running. LoadLeveler has to be set-up to show this data in the
llq command. Normally this is turned off for performance reasons.
It is up to the cluster administrator to decide whether or not to
publish this information. The back-end will work whether or not this
is turned on.

\subsubsection{Relevant arc.conf options}

Only the two basic LRMS config options are relevant for LoadLeveler:

\begin{itemize}
\item The LoadLeveler batch system back-end is enabled by setting \texttt{lrms=\char`\"{}ll\char`\"{}}
in the \texttt{{[}common]} configuration block. 
\item \texttt{ll\_bin\_path} configuration variable of the \texttt{{[}common]}
block must be set to the path of the LoadLeveler binaries.
\end{itemize}

\subsubsection{Implementation details}

The LoadLeveler back-end uses LoadLeveler's command line interface(CLI)
commands to submit and cancel jobs. All information in the information
system is similarly parsed from the output of CLI commands. It does
not parse any log files, nor does it use the binary APIs. The reason
that the back-end is completely based on the CLI is that the log files
are normally kept on another machine than the front end and that the
binary API for LL changes quite often. Often with each new version
of LL.

\subsubsection{Known limitations}

There is at the moment no support for parallel jobs on the LoadLeveler
back-end.

\subsection{Fork}

The Fork back-end is a simple back-end that interfaces to the local
machine i.e. there is no batch system underneath. It simply forks
the job, hence the name. The back-end then uses standard posix commands(e.g.
ps or kill) to manage the job.

\subsubsection{Recommended batch system configuration}

Since fork is a simple back-end and does not use any batch system,
there is no specific configuration needed of the underlying system.

\subsubsection{Relevant arc.conf options}

\begin{itemize}
\item The Fork back-end is enabled by setting \texttt{lrms=\char`\"{}fork\char`\"{}}
in the \texttt{{[}common]} configuration block. 
\item The queue must be named {}``fork'' in the queue section.
\item fork\_job\_limit=\char`\"{}cpunumber\char`\"{}, this option is used
to set the number of running grid jobs on the fork machine, allowing
a multi core machine to use some or all of its cores for Grid jobs.
The default value is 1. 
\end{itemize}

\subsubsection{Implementation details and known limitations}

The Fork backend implements an interface to the {}``fork'' unix
command which is not a batch system. Therefore the backend should
rather be seen as an interface to the operating system itself. Most
of the {}``batch system values'' are determined from the operating
system (e.g. cpu load) or manually set in the configuration file.

Fork is not a batch system, therefore many of the queue specific attributes
or detailed job information is not available. The support for the
{}``Fork batch system'' was introduced so that quick deployments
and testing of the middleware can be possible without dealing with
deployment of a real batch system since fork is available on every
unix box. The Fork backend is not recommended to be used in production.
The backend by its nature, has lots of limitations, for example does
not support parallel jobs.

\subsection{LSF}

\subsubsection{Recommended batch system configuration}

Set up one or more LSF queues dedicated for access by grid users.
All nodes in these queues should have a resource type which corresponds
to the one of the the frontend and which is reported to the outside.
The resource type needs to be set properly in the 'lsb.queues' configuration
file. Be aware that LSF distinguishes between 32 and 64 bit for Linux.
For a homogeneous cluster, the 'type==any' option is convenient alternative.

Example: In lsb.queues set either of the following:

\begin{itemize}
\item RES\_REQ = type==X86\_64 
\item RES\_REQ = type==any 
\end{itemize}
See the '-R' option of the bsub command man page for more explanation.

\subsubsection{Relevant arc.conf options}

The LSF back-end requires that the following options are specified:

\begin{itemize}
\item \texttt{lrms=\char`\"{}lsf\char`\"{}} in the \texttt{{[}common]} configuration
block. 
\item \texttt{lsf\_bin\_path} configuration variable of the \texttt{{[}common]}
block must be set to the path of the LSF binaries.
\item lsf\_profile\_path must be set to the filename of the LSF profile
that the back-end should use. 
\end{itemize}
Furthermore it is very important to specify the correct architecture
for a given queue in arc.conf. Because the architecture flag is rarely
set in the xRSL file the LSF back-end will automatically set the architecture
to match the chosen queue. LSF's standard behaviour is to assume the
same architecture as the frontend. This will fail for instance if
the frontend is a 32 bit machine and all the cluster resources are
64 bit.If this is not done the result will be jobs eing rejected by
LSF because LSF believes there are no useful resources available.

\subsubsection{Implementation details}

The LSF implementation of the back-end are based solely on parsing
and running LSF's command line interface commands. No log files or
other methods are used. To get the correct output o any output at
all the back-end needs to have an apropriate LSF profile. The path
to this profile must be set in arc.conf. It will then be executed
by the back-end before running any of LSF's CLI commands.

\subsubsection{Known limitations}

Parallel jobs have not been tested on the LSF back-end.

The back-end does not at present support reporting different number
of free CPUs per user.

\subsection{SGE}

Sun Grid Engine, formely known as Codine, is an open source batch
system maintained by Sun. Runs on Linux, Solaris.

\subsubsection{Recommended batch system configuration}

Set up one or more SGE queues for access by grid users. Queues can
be shared by normal and grid users. In case you want to set up nore
than one ARC queue, make sure that the corresponding SGE queues have
no shared nodes among them. Otherwise the counts of free and occupied
CPUs might be wrong. Only SGE versions 6 and above are supported.

\subsubsection{Relevant arc.conf options}

The SGE back-end requires that the following options are specified:

\begin{itemize}
\item The SGE batch system back-end is enabled by setting \texttt{lrms=\char`\"{}sge\char`\"{}}
in the \texttt{{[}common]} configuration block. 
\item \texttt{sge\_root} must be set to SGE's install root.
\item \texttt{sge\_bin\_path} configuration variable of the \texttt{{[}common]}
block must be set to the path of the SGE binaries.
\item \texttt{sge\_cell}, \texttt{sge\_qmaster\_port} and \texttt{sge\_execd\_port}
options might be necessary to set in special cases. See the \texttt{arc.conf(5)}
man page for more details.
\item \texttt{sge\_jobopts} configuration variable of the \texttt{{[}queue]}
block can be used to add custom SGE options to job scripts submitted
to SGE. Consult SGE documentation for possible options.
\end{itemize}
Example:

\texttt{lrms=\char`\"{}sge\char`\"{}}

\texttt{sge\_root=\char`\"{}/opt/n1ge6\char`\"{}}

\texttt{sge\_bin\_path=\char`\"{}/opt/n1ge6/bin/lx24-x86\char`\"{}}

\texttt{...}

\texttt{{[}queue/long]}

\texttt{sge\_jobopts=\char`\"{}-P atlas -r yes\char`\"{}}

\subsubsection{Implementation details}

The SGE backend's commands are similar to the PBS commands. These
commands are used in the code:

Submit job:

\begin{itemize}
\item qsub -S /bin/sh (specifies the interpreting shell for the job)
\end{itemize}
Get jobs status:

If the job state is not suspended, running or pending then its state
is failed.

\begin{itemize}
\item qstat -u '{*}' -s rs (show the running and suspended jobs status)
\item qstat -u '{*}' -s p (show the pending jobs status)
\item qstat -j job\_id (long job information)
\item qacct -j job\_id (finished job report)
\end{itemize}
Job terminating:

\begin{itemize}
\item qdel job\_id (delete Sun Grid Engine job from the queue)
\end{itemize}
Queue commands:

\begin{itemize}
\item qconf -spl (show a list of all currently defined parallel environments)
\item qconf -sql (show a list of all queues)
\item qconf -sep (show a list of all licensed processors/slots)
\item qstat -g c (display cluster queue summary)
\item qconf -sconf global (show global configuration)
\item qconf -sq queue\_name (show the given queue cofiguration)
\end{itemize}
Other:

\begin{itemize}
\item qstat -help (show Sun Grid Engine's version and type)
\end{itemize}

\subsubsection{Known limitations}

Multi-CPU support is not well tested. All users are shown with the
same quotas in infosys, even if they are mapped to different local
users. The requirement that one ARC queue maps to one SGE queue is
too restrictive, as the SGE's notion of a queue differs widely from
ARC's definition. The flexibility available in SGE for defining policies
is difficult to accurately translate into nordugrid's information
schema. The closest equivalent of nordugrid-queue-maxqueuable is a
per-cluster limit in SGE, and the value of nordugrid-queue-localqueued
is not well defined if pending jobs can have multiple destination
queues.

\section{ARC1}
This section contains the current implementation of the ARC1 back-ends.

\subsection{The Information System}
ARC1 provides support for both the classic NorduGrid information schema\cite{INFOSYS} as well as an early minimal GLUE 2 implementation. The functionality is divided between several information collecting scripts and the output is rendered in XML format. Further details about each script are given in the subsections below.
The current GLUE 2 implementation is given in section Section 5.2.

\subsubsection{The information Collector Scripts}
The information collector scripts are divided between several seperate scripts:
\begin{itemize}
\item \texttt{cluster+qju.pl} - driver for information collection. It calls all other infomation collectors and prints results in XML. The information collection is done by one single invocation of this script
\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm})
\item \texttt{InfoChecker.pm} - used by InfoCollector to validate options and results against a simple 'schema' (not XML Schema)
\item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments ...)
\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...)
\item \texttt{BATCH\_SYSTEM\_NAME.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework
\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NG schema
\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
\item \texttt{ARC0ClusterSchema.pm} - description of the info produced by ARC0ClusterInfo
\item \texttt{ARC1ClusterSchema.pm} - description of the info produced by ARC1ClusterInfo
\end{itemize}

All information about the hosting environment and queues is collected by one invocation of a perl script called \texttt{cluster+qju.pl}. The XML output is printed to \texttt{STDOUT} and it includes two representations of the output data, namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.

\subsection{The ARC GLUE 2 implementation}
Currently, the the following GLUE2 classes are included in ARC1:
\begin{itemize}
\item ComputingService
\item ComputingEndpoint
\item ComputingShare
\end{itemize}

While the following classes are still missing (TODO: According to our table in wiki, there should be some other sections as well):
\begin{itemize}
\item ApplicationEnvironment
\item ExecutionEnvironment
\item ComputingActivity
\item Policy including MappingPolicy / AccessPolicy
\end{itemize}

The different classes with their attributes are described in their respective sections below.

\subsubsection{ApplicationEnvironment}

\phantomsection
%\addcontentsline{toc}{subsubsection}{LocalID}
\hspace*{0.5cm}
\begin{shaded}
\textbf{LocalID}
\end{shaded}
\begin{tabular}{lp{10cm}}  
Attribute value:& LocalID\_t\\
Example:&  \verb#TODO#\\
Related xRSL:& \verb#NA#\\
UI role:& TODO \\

\end{tabular}

Description: An opaque identifier local to the Computing Service.

\phantomsection
%\addcontentsline{toc}{subsubsection}{Name}
\hspace*{0.5cm}
\begin{shaded}
\textbf{Name}
\end{shaded}
\begin{tabular}{lp{10cm}}  
Attribute value:& String\\
Example:& \verb#Name: TODO#\\
Related xRSL:& TODO\\
UI role:& ?\\
\end{tabular}

Description: Name of the application environment.

\phantomsection
%\addcontentsline{toc}{subsubsection}{Version}
\hspace*{0.5cm}
\begin{shaded}
\textbf{Version}
\end{shaded}
\begin{tabular}{lp{10cm}}  
Attribute value:& String\\
Example:& \verb#Version: TODO#\\
Related xRSL:& TODO\\
UI role:& TODO\\
\end{tabular}

Description: Version of the application environment.

\begin{thebibliography}{1}
\bibitem{ARC}\char`\"{}Advanced Resource Connector middleware for
lightweight computational Grids\char`\"{}. M.Ellert et al., Future
Generation Computer Systems 23 (2007) 219-240.

\bibitem{GM}The NorduGrid Grid Manager And GridFTP Server: Description
And Administrator's Manual {[}NORDUGRID-TECH-2], A.Konstantinov

\bibitem{INFOSYS}\char`\"{}The NorduGrid/ARC Information System\char`\"{}
{[}NORDUGRID-TECH-4], B.Kónya

\bibitem{subversion}NorduGrid subversion: http://svn.nordugrid.org/trac/nordugrid/browser

\bibitem{config}ARC configuration document: currently the \texttt{arc.conf.template}
file.

\bibitem{Condor}http://www.cs.wisc.edu/condor/

\bibitem{GLUE2} TODO: ADD REFERENCE
\end{thebibliography}

\end{document}

### 2008-08-01 doc/ws-quick-guide/ws-quick-guide.tex
Remove executable properties from files (windows feature?) and mark binaries as ... binaries

@@
 stability, reliability and performance of the middleware. A growing
-number of grid projects, like NDGF, Swegrid or Swiss National Grid has
-chosen ARC as their middleware.
+number of grid projects, like NDGF~\cite{ndgf}, Swegrid~\cite{swegrid} or Swiss
+National Grid Association~\cite{swing} has chosen ARC as their middleware.
 
@@
 components of ARC, often referred to as ARC1 components. WS design of ARC is
-based on a service container -- the Hosting Environment Daemon (HED), and
-different grid capabilities are implemented as Web Services residing in HED.
+based on a service container -- the \textit{Hosting Environment Daemon (HED)},
+and different grid capabilities are implemented as Web Services residing in HED.
 
@@
 \begin{itemize}
-\item an OGSA BES~\cite{ogsa-bes} compliant execution service -- \em{ARC
-Resource-coupled EXecution service (A-REX)}
-\item an echo service (for testing purposes)
+\item an OGSA BES~\cite{ogsa-bes} compliant execution service --
+\textit{ARC Resource-coupled EXecution service (A-REX)}~\cite{arex}
+\item an \textit{Echo} service (for testing purposes)
 \end{itemize}
@@
 
-You can get the latest source code for ARC1 from the Subversion repository. See
-\url{http://svn.nordugrid.org} for more details.
+You can get the latest source code for ARC1 from the Subversion repository.
+For details, see:
+
+\url{http://svn.nordugrid.org}
 
@@
 
-\begin{shaded}
 \url{http://download.nordugrid.org/software/nordugrid-arc1/nightly/}
-\end{shaded}
 
-Choose the latest date available and download snapshot tarball -- for example
+Choose the latest date available and download snapshot tarball -- for example\\
 \verb|nordugrid-arc1-200802201038-snapshot.tar.gz|.
@@
 There are examples of A-REX service and Echo service with Security Handlers
-being used. They may be found in
-\verb|$ARC_LOCATION/share/doc/arc/arex_secure.xml| and
+being used. They may be found in:
+
+\verb|$ARC_LOCATION/share/doc/arc/arex_secure.xml| \\and\\
 \verb|$ARC_LOCATION/share/doc/arc/echo.xml|.
@@
 Specifically for \texttt{arc.pdp} and \texttt{pdpservice}, a formatted policy
-with specific schema should be managed, see
-\verb|$ARC_LOCATION/share/doc/arc/pdp_policy.xml.example| and 
-\verb|$ARC_LOCATION/share/doc/arc/Policy.xsd| for details.
+with specific schema should be managed, for details see:
+
+\verb|$ARC_LOCATION/share/doc/arc/pdp_policy.xml.example| \\and\\
+\verb|$ARC_LOCATION/share/doc/arc/Policy.xsd|
 
@@
 
-\begin{shaded}
 \url{http://www.nordugrid.org/documents/certificate_howto.html}\\
-\url{http://www.nordugrid.org/documents/ng-server-install.html#security}.
-\end{shaded}
+\url{http://www.nordugrid.org/documents/ng-server-install.html#security}
 
@@
 
-\begin{shaded}
 \url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/sec/TestCA}
-\end{shaded}
 
@@
 
-\begin{shaded}
 \url{https://vls.grid.upjs.sk/CA/instantCA}
-\end{shaded}
 
@@
 
-\texttt{/etc/grid-security/hostcert.pem} \\and\\
-\texttt{/etc/grid-security/hostkey.pem}
+\verb|/etc/grid-security/hostcert.pem| \\and\\
+\verb|/etc/grid-security/hostkey.pem|
 
@@
 The open source development of the ARC middleware is coordinated by the
-NorduGrid Collaboration. Currently, the main contributor is the KnowARC project
-(\url{www.knowarc.eu}), but the Collaboration is open to new members.
+NorduGrid Collaboration. Currently, the main contributor is the KnowARC
+project~\cite{knowarc}, but the Collaboration is open to new members.
 Contributions from the community to the software and the documentation is
-welcomed. Sources can be downloaded from the software repository at
-\url{download.nordugrid.org} or the Subversion code repository at
-\url{svn.nordugrid.org}.
+welcomed. Sources can be downloaded from the software repository at:
+
+\url{http://download.nordugrid.org}
+
+or the Subversion code repository at
+
+\url{http://svn.nordugrid.org}.
 
@@
 to be addressed in the framework of the ARC development. Feature
-requests and enhancement proposals are recorded in the Bugzilla bug
-tracking system at \url{bugzilla.nordugrid.org}. For a more detailed
-description, write access to the code repository and further
-questions, write to the \texttt{nordugrid-discuss} mailing list (see
-\url{www.nordugrid.org} for details). Ongoing and completed Grid Research
-projects and student assignments related to the middleware are listed
+requests and enhancement proposals are recorded in the Bugzilla problem
+tracking system at:
+
+ \url{http://bugzilla.nordugrid.org}
+
+For a more detailed description, write access to the code repository and
+further questions, write to the \texttt{nordugrid-discuss} mailing list (see
+NorduGrid Web site~\cite{nordugrid} for details). Ongoing and completed Grid
+Research projects and student assignments related to the middleware are listed
 on the NorduGrid Web site as well.
@@
 User support and site installation assistance is provided via the
-request tracking system available at \texttt{nordugrid-support@nordugrid.org}.
+request tracking system available by e-mail to
+\texttt{nordugrid-support@nordugrid.org}.
 In addition, NorduGrid runs a couple of mailing lists, among which the
@@
 Feature and enhancement requests, as well as discovered problems, should be
-reported in the Bugzilla problem tracking system \url{bugzilla.nordugrid.org}.
+reported in the Bugzilla problem tracking system
+(\url{http://bugzilla.nordugrid.org}).
 
@@
 installation instructions, conference presentations, FAQ and even
-tutorial materials can be fetched from the documentation section of
-\url{www.nordugrid.org}.
+tutorial materials can be fetched from the documentation section of the
+NorduGrid Web site.
 
-Contact information is kept updated on the \url{www.nordugrid.org} Web site.
+Contact information is kept updated on the NorduGrid Web site.
 
-\bibliography{ws-quick-guide}
+\bibliography{grid}
 \end{document}

### 2008-08-01 doc/backend_interface/Backends.tex
Initial version

@@
 \hypersetup{
-  pdfauthor = {C.U.Soettrup},
-  pdftitle ={Arc backend infrastructure},
-  pdfsubject = {Description and developer guide},
-  pdfkeywords = {ARC,NorduGrid,manual,Backend,LRMS, Infosystem},
+  pdfauthor = {T.Frågåt},
+  pdftitle ={Arc batch system back-end interface guide with support for GLUE 2},
+  pdfsubject = {Description and developer's guide for ARC1},
+  pdfkeywords = {ARC,ARC1,NorduGrid,manual,Backend,LRMS,Infosystem,GLUE2},
   pdfcreator = {PDFLaTeX with hyperref package},
@@
 
-{\raggedleft NORDUGRID-TECH-13\par}
+{\raggedleft NORDUGRID-TECH-XX\par}
 
@@
 \begin{center}
-\textit{\large Description and developer's guide}
+\textit{\large Description and developer's guide for ARC1}
 \par\end{center}{\large \par}
@@
 
-{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomasf@fys.uio.no}, C.U.Søttrup, Bal\'azs K\'onya, G\'abor R\H oczei \par}
+{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomasf@fys.uio.no} \par}
 
@@
 
-This document describes the Advanced Resource Connector\cite{ARC}
+This document describes the Advanced Resource Connector~\cite{arc1}
 batch system back-end infrastrucure. It aims to describe the infrastructure
@@
 The batch system back-ends are what tie the ARC grid middleware (through
-the Grid Manager (GM\cite{GM}) and the Information System layer\cite{INFOSYS})
-to the underlying cluster management system or LRMS. The back-ends
-consist of set of shell and Perl scripts whose role are twofold:
+(ARC Resource-coupled EXecution servcie, A-REX~\cite{arex}) to the underlying
+cluster management system or LRMS. The back-ends consist of set of shell and
+Perl scripts whose role are twofold:
 
 \begin{enumerate}
-\item to allow the GM to manipulate jobs on the LRMS including job submit,
-cancel operations and scanning for completed jobs
+\item to allow the GM, which is resided in A-REX, to manipulate jobs on the LRMS including job submit,
+cancel operations etc.
 \item to collect information about jobs, users, batch system and the cluster
@@
 \subsection{ARC Classic}
-As of ARC version 0.6 the the two sets of scripts are located in different
-directories within the NorduGrid subversion tree\cite{subversion}.
-The job control interface scripts can be found under \texttt{arc\_development\_line/grid-manager/LRMS/}
-while the information collectors are located at \texttt{arc\_development\_line/infosystem/.}
-
-Section 2, 3, and 4 gives a description of the classic ARC back-ends scripts.
+The ARC Classic has its own solution, please refer to ~\cite{backends-arc0}.
 
 \subsection{ARC1}
-As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories within the NorduGrid subversion tree\cite{subversion} as well. 
-The architecture of ARC1 is different from Classic ARC so the job control interface scripts can be found under
-\texttt{arc1/trunk/src/services/a-rex/lrms/}
+As of ARC version 0.9, which is to be the next generation ARC, the scripts are located in different directories
+within the NorduGrid subversion tree~\cite{svn}. 
+The job control interface scripts can be found under \texttt{arc1/trunk/src/services/a-rex/lrms/}
 while the information collectors are located in \texttt{arc1/trunk/src/services/a-rex/infoproviders/}.
 
-Furthermore, the next generation of ARC supports both the classic NorduGrid information schema\cite{INFOSYS} and
-currently a minimal set of the GLUE specification version 2.0 schema\cite{GLUE2}.
-
-The ARC1 back-end architecture and GLUE2 information model implementation is presented in Section 5.
+The next generation of ARC supports both the classic NorduGrid information schema~\cite{is} and
+currently a minimal set of the GLUE specification version 2.0 schema~\cite{glue2}.
 
@@
 
-Job control part of the LRMS interface is handled by the Grid Manager\cite{GM}.
+Job control part of the LRMS interface is handled by the Grid Manager~\cite{gm}.
 It takes care of preparing a native batch system submission script,
@@
 the Section \char`\"{}8.6 LRMS Support\char`\"{} and Section \char`\"{}8.7
-Runtime Environment\char`\"{} of the Grid-Manager guide\cite{GM}.
+Runtime Environment\char`\"{} of the Grid-Manager guide~\cite{gm}.
 
@@
 The main purpose of the information system batch interface is to populate
-the nordugrid information model with locally collected information
-obtained from the batch system. Important to recall that the locally
-collected information is broader than what the batch-system can offer,
-information taken from the grid-layer (mostly grid manager), from
+the nordugrid information model and the GLUE2 model with locally collected
+information obtained from the batch system. Important to recall that the
+locally collected information is broader than what the batch-system can
+offer, information taken from the grid-layer (mostly grid manager), from
 the front-end machine and from the \texttt{arc.conf} configuration
 file are also needed and used to populate the nordugrid information
-model\cite{INFOSYS}. Below only the batch system related information
-collection is covered. Note that the separation of the different sources
-is not always that straightforward (e.g. the grid-manager information
-has some overlap with the batch system info).
+model~\cite{is} and the GLUE2 schema~\cite{glue2}.
 
 The information system interface consists of set of Perl scripts which
-populate the local lDAP database by generating ldif files. Historically
-two information provider \char`\"{}frameworks\char`\"{} have been
-coexisted. As of ARC version 0.6 the old framework is being phased
-out by the new framework. Briefly about the two frameworks:
+generates an XML output to \texttt{STDOUT}.
 
+The following subsection describes the LRMS interface part of the
+information system. To support a particular LRMS a dedicated Perl module
+should be written that implements the interface presented in the next
+subsections. Furthermore hooks should be added to the \texttt{LRMS.pm} module,
+so that it uses the correct Perl module depending on the LRMS type.
+
+The information collector scripts are divided between several seperate scripts:
 \begin{itemize}
-\item the old framework separated the non-batch system related functionality
-into the \texttt{InfosysCluster.pm, InfosysQJU.pm} modules and kept
-the batch system specific interface within the \texttt{cluster-condor.pl,
-queue+jobs+users-condor.pl} (replace condor with your scheduler name)
-files. There are two templates provided for batch system interface
-developers: \texttt{cluster-gm.pl, queue+jobs+users-gm.pl}
-\item the new framework originally proposed by Juha Lento%
-\footnote{the original posting of the new framework proposal: http://staff.csc.fi/juha.lento/grid/rewrite.html%
-} has been brought up-to-date by implementing all the features present
-in the old framework. The new framework consists of two main scripts,
-the \texttt{cluster.pl} and the \texttt{qju.pl}. The non-batch system
-functionality is implemented by these script (plus some auxiliary
-scripts). The batch system specific interface is moved to LRMS modules
-such as \texttt{SGE.pm, PBS.pm, Fork.pm}, a template module \texttt{Template.pm}
-is also provided. The main advantage of the new framework is the separation
-of output generation from the information collection (this will ease
-the migration to XML output).
+\item \texttt{cluster+qju.pl} - driver for information collection.
+It calls all other infomation collectors and prints results in XML.
+The information collection is done by one single invocation of this script
+\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm})
+\item \texttt{InfoChecker.pm} - used by InfoCollector to validate options and results against a simple 'schema' (not XML Schema)
+\item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
+\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments ...)
+\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...)
+\item \texttt{BATCH\_SYSTEM\_NAME.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework
+\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NG schema
+\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
+\item \texttt{ARC0ClusterSchema.pm} - description of the info produced by ARC0ClusterInfo
+\item \texttt{ARC1ClusterSchema.pm} - description of the info produced by ARC1ClusterInfo
 \end{itemize}
-The following subsection describes the LRMS interface part of the
-information system as implemented via the new framework. To support
-a particular LRMS a dedicated Perl module should be written that implements
-the interface presented in the next subsections. Furthermore hooks
-should be added to the \texttt{LRMS.pm} module, so that it uses the
-correct Perl module depending on the LRMS type. Short description
-of the files of the new framework:
 
-\begin{itemize}
-\item \texttt{cluster.pl and qju.pl}: main scripts called to populate the
-nordugrid information model
-\item \texttt{Shared.pm:} subroutines shared by cluster.pl and cluster.pl
-\item \texttt{LogUtils.pm}: logging utilities
-\item \texttt{LRMS.pm:} defines the LRMS interface (described in the next
-subsection)
-\item \texttt{Template.pm:} module template for implementing a new LRMS
-interface
-\item \texttt{SGE.pm, PBS.pm, LSF.pm, Fork.pm, LL.pm, Condor.pm}: actual
-LRMS modules
-\end{itemize}
+All information about the hosting environment and queues is collected by one invocation of the perl script named
+\texttt{cluster+qju.pl}. The XML output is printed to \texttt{STDOUT} and it includes
+two representations of the output data, namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.
 
@@
 LRMS jobstatus mapping%
-\footnote{LRMS job states mapping is described in\cite{INFOSYS} %
+\footnote{LRMS job states mapping is described in~\cite{is} %
 }:Running->'R', Queued->'Q', Suspended->'S', Exiting->'E', Other->'O'\tabularnewline
@@
 number of freely available CPUs with their time limits in minutes
-(see nordugrid-authuser-freecpus in\cite{INFOSYS})\tabularnewline
+(see nordugrid-authuser-freecpus in~\cite{is})\tabularnewline
 \hline 
@@
 
-The ARC configuration manual contains detailed information on all
-the ARC configuration parameters\cite{config}and should be consulted
-as primary source. Below the PBS specific variables are collected.
+Below the PBS specific configuration variables are collected.
 
@@
 
-\subsection{Condor}
-
-The Condor \cite{Condor} system, developed at the University of Wisconsin-Madison,
-was initially used to harness free cpu cycles of workstations. Over
-time it has evolved into a complex system with many grid-oriented
-features. Condor is available on a large variety of platforms.
-
-\subsubsection{Recommended batch system configuration}
-
-Install Condor on the Grid Manager (GM) node and configure it as a
-submit machine. Next, you must add the following to the node's Condor
-configuration (\texttt{CONDOR\_IDS} can also be an environment variable):
-
-\texttt{MAIL = <ARC\_install\_prefix>/libexec/finish-condor-job}
-
-\texttt{CONDOR\_IDS = 0.0}
-
-The \texttt{MAIL} attribute will instruct Condor to run the specified
-program on job completion. The default on Condor is to run /bin/mail
-to notify the user, but in this case, it is the GM that needs the
-notification. Therefore, /bin/mail is replaced with a program especially
-written for talking to the GM.
-
-\texttt{CONDOR\_IDS} has to be 0.0, so that the above notification
-program can access the Grid job's session directories (needed to extract
-the job exit code from the Condor log).
-
-Make sure that no normal users are allowed to submit Condor jobs from
-this node. For one thing, it would not work for the user, since Condor
-will try to notify the GM instead of the job owner on job completion.
-If you don't allow normal user logins on the GM machine, then you
-don't have to do anything. If you for some reason want to allow users
-to log into the GM machine, simply don't allow them to execute the
-condor\_submit program. This can be done by putting all local Unix
-users allocated to the Grid in a single group, e.g. 'griduser', and
-then setting the file ownership and permissions on condor\_submit
-like this:
-
-\texttt{chgrp griduser \$condor\_location/bin/condor\_submit}
-
-\texttt{chmod 750 \$condor\_location/bin/condor\_submit }
-
-\subsubsection{Relevant arc.conf options}
-
-\begin{itemize}
-\item The Condor batch system back-end is enabled by setting \texttt{lrms=\char`\"{}condor\char`\"{}}
-in the \texttt{{[}common]} configuration block. 
-\item \texttt{condor\_location} configuration variable of the \texttt{{[}common]}
-block should be set to the Condor install prefix (i.e., the directory
-containing Condor's bin, sbin, etc).
-\item \texttt{condor\_config} configuration variable of the \texttt{{[}common]}
-block should be set to the value the environment variable \texttt{CONDOR\_CONFIG}
-should have (but don't try to use the environment variable directly
-as \texttt{\$CONDOR\_CONFIG}, since it will probably not be defined
-when arc.conf is parsed!)
-\item \texttt{condor\_rank} configuration variable of the \texttt{{[}common]}
-block, if defined, will cause the \texttt{Rank} attribute to be set
-in each job description submitted to Condor. Use this option if you
-are not happy with the way Condor picks out nodes when running jobs
-and want to define your own ranking algorithm. \texttt{condor\_rank}
-should be set to a ClassAd float expression that you could use in
-the \texttt{Rank} attribute in a Condor job description. For example:
-\end{itemize}
-\begin{quote}
-\texttt{condor\_rank=\char`\"{}(1-LoadAvg/2){*}(1-LoadAvg/2){*}Memory/1000{*}KFlops/1000000\char`\"{} }
-\end{quote}
-\begin{itemize}
-\item \texttt{condor\_requirements} configuration variable of the \texttt{{[}queue]}
-block defined a subpool of condor nodes. See next section for details.
-\end{itemize}
-
-\subsubsection{Implementation details}
-
-The job control part of the interface uses the \texttt{condor\_submit}
-command to submit jobs. Some of the options used in the job's ClassAd
-are:
-
-\texttt{Requirements} -- is used to select the nodes that may run
-the job. This is how ARC queues are implemented for Condor.
-
-\texttt{Periodic\_remove} -- is used to enforce cputime and walltime
-limits.
-
-\texttt{Log} -- the job's condor log file is parsed by the information
-scripts to find out whether the job was suspended.
-
-The information system component uses the following Condor commands:
-
-\texttt{condor\_status -long} -- for collecting information about
-nodes
-
-\texttt{condor\_status -format \char`\"{}\%s\textbackslash{}n\char`\"{}
-Machine -constraint '...'} -- for listing nodes that make up an ARC
-queue.
-
-\texttt{condor\_q -long -global} -- for monitoring running jobs.
-
-\texttt{condor\_history -l jobid} -- for collecting information about
-finished jobs. Further cues are taken from the job's condor log file
-and the body of the email sent by Condor when a job completes.
-
-\subsubsection{Configuring Queues}
-
-Condor does not support queues in the classical sense. It is possible,
-however, to divide the Condor pool in several sub-pools. An ARC 'queue'
-is then nothing more than a subset of nodes from the Condor pool.
-
-Which nodes go into which queue is defined using the \texttt{condor\_requirements}
-configuration option in the corresponding \texttt{{[}queue]} section.
-It's value must be a well-formed constraint string that is accepted
-by a \texttt{condor\_status -constraint '...'} command. Internally,
-this constraint string is used to determine the list of nodes belonging
-to a queue. This string can get quite long, so, for readability reasons
-it is allowed to split it up into pieces by using multiple \texttt{condor\_requirements}
-options. The full constrains string will be reconstructed by concatenating
-all pieces.
-
-Queues should be defined in such a way that their nodes all match
-the information available in ARC about the queue. A good start is
-for the \texttt{condor\_requirements} attribute to contain restrictions
-on the following: \texttt{Opsys}, \texttt{Arch}, \texttt{Memory} and
-\texttt{Disk}. If you wish to configure more than one queue, it's
-good to have queues defined in such a way that they do not overlap.
-In the following example disjoint memory ranges are used to ensure
-this:
-
-\begin{quote}
-\texttt{{[}queue/large] condor\_requirements=\char`\"{}(Opsys == \char`\"{}linux\char`\"{}
-\&\& (Arch == \char`\"{}intel\char`\"{} || Arch == \char`\"{}x86\_64\char`\"{})\char`\"{}
-condor\_requirements=\char`\"{} \&\& (Disk > 30000000 \&\& Memory
-> 2000)\char`\"{}}
-
-\texttt{{[}queue/small] condor\_requirements=\char`\"{}(Opsys == \char`\"{}linux\char`\"{}
-\&\& (Arch == \char`\"{}intel\char`\"{} || Arch == \char`\"{}x86\_64\char`\"{})\char`\"{}
-condor\_requirements=\char`\"{} \&\& (Disk > 30000000 \&\& Memory
-<= 2000 \&\& Memory > 1000)\char`\"{}}
-\end{quote}
-Note that '\texttt{nodememory}' attribute in arc.conf means the maximum
-memory available for jobs, while the Memory attribute in Condor is
-the physical memory of the machine. To avoid swapping (and these are
-probably not dedicated machines!), make sure that '\texttt{nodememory}'
-is smaller than the minimum physical memory of the machines in that
-queue. If for example the smallest node in a queue has 1Gb memory,
-then it would be sensible to use \texttt{nodememory=\char`\"{}850\char`\"{}}
-for the maximum job size.
-
-In case you want more precise control over which nodes are available
-for grid jobs, using pre-defined ClassAds attributes (like in the
-example above) might not be sufficient. Fortunately, it's possible
-to mark nodes by using some custom attribute, say \texttt{NORDUGRID\_RESOURCE}.
-This is accomplished by adding a parameter to the node's local Condor
-configuration file, and then adding that parameter to \texttt{STARTD\_EXPRS}:
-
-\begin{quote}
-\texttt{NORDUGRID\_RESOURCE = True }
-
-\texttt{STARTD\_EXPRS = NORDUGRID\_RESOURCE, \$(STARTD\_EXPRS)}
-\end{quote}
-Now queues can be restricted to contain only 'good' nodes. Just add
-to each \texttt{{[}queue]} section in arc.conf:
-
-\begin{quote}
-\texttt{condor\_requirements=\char`\"{} \&\& NORDUGRID\_RESOURCE\char`\"{} }
-\end{quote}
-
-\subsubsection{Known limitations}
-
-Only Vanilla universe is supported. MPI universe (for multi-CPU jobs)
-is not supported. Neither is Java universe (for running Java executables).
-ARC can only send jobs to Linux machines in the Condor pool, therefore
-excluding other unixes and Windows destinations. The session directory
-must be on a network shared directory, visible from all worker nodes.
-
-\subsection{LoadLeveler}
-
-LoadLeveler(LL), or Tivoli Workload Scheduler LoadLeveler in full,
-is a parallel job scheduling system developed by IBM.
-
-\subsubsection{Recommended batch system configuration}
-
-The back-end should work fine with a standard installation of LoadLeveler.
-For the back-end to report the correct memory usage and cputime spent,
-while running. LoadLeveler has to be set-up to show this data in the
-llq command. Normally this is turned off for performance reasons.
-It is up to the cluster administrator to decide whether or not to
-publish this information. The back-end will work whether or not this
-is turned on.
-
-\subsubsection{Relevant arc.conf options}
-
-Only the two basic LRMS config options are relevant for LoadLeveler:
-
-\begin{itemize}
-\item The LoadLeveler batch system back-end is enabled by setting \texttt{lrms=\char`\"{}ll\char`\"{}}
-in the \texttt{{[}common]} configuration block. 
-\item \texttt{ll\_bin\_path} configuration variable of the \texttt{{[}common]}
-block must be set to the path of the LoadLeveler binaries.
-\end{itemize}
-
-\subsubsection{Implementation details}
-
-The LoadLeveler back-end uses LoadLeveler's command line interface(CLI)
-commands to submit and cancel jobs. All information in the information
-system is similarly parsed from the output of CLI commands. It does
-not parse any log files, nor does it use the binary APIs. The reason
-that the back-end is completely based on the CLI is that the log files
-are normally kept on another machine than the front end and that the
-binary API for LL changes quite often. Often with each new version
-of LL.
-
-\subsubsection{Known limitations}
-
-There is at the moment no support for parallel jobs on the LoadLeveler
-back-end.
-
 \subsection{Fork}
@@
 not support parallel jobs.
-
-\subsection{LSF}
-
-\subsubsection{Recommended batch system configuration}
-
-Set up one or more LSF queues dedicated for access by grid users.
-All nodes in these queues should have a resource type which corresponds
-to the one of the the frontend and which is reported to the outside.
-The resource type needs to be set properly in the 'lsb.queues' configuration
-file. Be aware that LSF distinguishes between 32 and 64 bit for Linux.
-For a homogeneous cluster, the 'type==any' option is convenient alternative.
-
-Example: In lsb.queues set either of the following:
-
-\begin{itemize}
-\item RES\_REQ = type==X86\_64 
-\item RES\_REQ = type==any 
-\end{itemize}
-See the '-R' option of the bsub command man page for more explanation.
-
-\subsubsection{Relevant arc.conf options}
-
-The LSF back-end requires that the following options are specified:
-
-\begin{itemize}
-\item \texttt{lrms=\char`\"{}lsf\char`\"{}} in the \texttt{{[}common]} configuration
-block. 
-\item \texttt{lsf\_bin\_path} configuration variable of the \texttt{{[}common]}
-block must be set to the path of the LSF binaries.
-\item lsf\_profile\_path must be set to the filename of the LSF profile
-that the back-end should use. 
-\end{itemize}
-Furthermore it is very important to specify the correct architecture
-for a given queue in arc.conf. Because the architecture flag is rarely
-set in the xRSL file the LSF back-end will automatically set the architecture
-to match the chosen queue. LSF's standard behaviour is to assume the
-same architecture as the frontend. This will fail for instance if
-the frontend is a 32 bit machine and all the cluster resources are
-64 bit.If this is not done the result will be jobs eing rejected by
-LSF because LSF believes there are no useful resources available.
-
-\subsubsection{Implementation details}
-
-The LSF implementation of the back-end are based solely on parsing
-and running LSF's command line interface commands. No log files or
-other methods are used. To get the correct output o any output at
-all the back-end needs to have an apropriate LSF profile. The path
-to this profile must be set in arc.conf. It will then be executed
-by the back-end before running any of LSF's CLI commands.
-
-\subsubsection{Known limitations}
-
-Parallel jobs have not been tested on the LSF back-end.
-
-The back-end does not at present support reporting different number
-of free CPUs per user.
 
@@
 
-\section{ARC1}
-This section contains the current implementation of the ARC1 back-ends.
+\section{The ARC GLUE 2 implementation}
+ARC1 provides support for both the classic NorduGrid information
+schema~\cite{is} as well as an early minimal GLUE 2 implementation.
+The output is rendered in XML format.
 
-\subsection{The Information System}
-ARC1 provides support for both the classic NorduGrid information schema\cite{INFOSYS} as well as an early minimal GLUE 2 implementation.
-The functionality is divided between several information collecting scripts and the output is rendered in XML format.
-Further details about each script are given in the subsections below.
-The current GLUE 2 implementation is given in section Section 5.2.
-
-\subsubsection{The information Collector Scripts}
-The information collector scripts are divided between several seperate scripts:
-\begin{itemize}
-\item \texttt{cluster+qju.pl} - driver for information collection.
-It calls all other infomation collectors and prints results in XML.
-The information collection is done by one single invocation of this script
-\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm})
-\item \texttt{InfoChecker.pm} - used by InfoCollector to validate options and results against a simple 'schema' (not XML Schema)
-\item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
-\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments ...)
-\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...)
-\item \texttt{BATCH\_SYSTEM\_NAME.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework
-\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NG schema
-\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
-\item \texttt{ARC0ClusterSchema.pm} - description of the info produced by ARC0ClusterInfo
-\item \texttt{ARC1ClusterSchema.pm} - description of the info produced by ARC1ClusterInfo
-\end{itemize}
-
-All information about the hosting environment and queues is collected by one invocation of a perl script called \texttt{cluster+qju.pl}.
-The XML output is printed to \texttt{STDOUT} and it includes two representations of the output data, namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.
-
-\subsection{The ARC GLUE 2 implementation}
-Currently, the the following GLUE2 classes are included in ARC1:
+Currently, the the following GLUE2 entities are included in ARC1:
 \begin{itemize}
@@
 
-While the following classes are still missing (TODO: According to our table in wiki, there should be some other sections as well):
+While the following entities are still missing (TODO: According to our table in wiki, there should be some other sections as well):
 \begin{itemize}
@@
 
-The different classes with their attributes are described in their respective sections below.
+The different entities with their attributes are described in their respective sections below.
 
-\subsubsection{ApplicationEnvironment}
+\subsection{ComputingService}
 
@@
 
-\begin{thebibliography}{1}
-\bibitem{ARC}\char`\"{}Advanced Resource Connector middleware for
-lightweight computational Grids\char`\"{}. M.Ellert et al., Future
-Generation Computer Systems 23 (2007) 219-240.
+\subsection{ComputingEndpoint}
 
-\bibitem{GM}The NorduGrid Grid Manager And GridFTP Server: Description
-And Administrator's Manual {[}NORDUGRID-TECH-2], A.Konstantinov
+\subsection{ComputingShare}
 
-\bibitem{INFOSYS}\char`\"{}The NorduGrid/ARC Information System\char`\"{}
-{[}NORDUGRID-TECH-4], B.Kónya
-
-\bibitem{subversion}NorduGrid subversion: http://svn.nordugrid.org/trac/nordugrid/browser
-
-\bibitem{config}ARC configuration document: currently the \texttt{arc.conf.template}
-file.
-
-\bibitem{Condor}http://www.cs.wisc.edu/condor/
-
-\bibitem{GLUE2} TODO: ADD REFERENCE
-\end{thebibliography}
+\bibliography{grid}
 

### 2008-08-01 doc/backend_interface/Backends.tex
Major restructuring

@@
 This document describes the Advanced Resource Connector~\cite{arc1}
-batch system back-end infrastrucure. It aims to describe the infrastructure
+batch system back-end infrastructure. It aims to describe the infrastructure
 in enough detail that a developer can add a new local resource management
@@
 The batch system back-ends are what tie the ARC grid middleware (through
-(ARC Resource-coupled EXecution servcie, A-REX~\cite{arex}) to the underlying
+(ARC Resource-coupled EXecution service, A-REX~\cite{arex}) to the underlying
 cluster management system or LRMS. The back-ends consist of set of shell and
@@
 be read, or 0 otherwise. Badly formed lines are silently ignored.
-Option values can be surrounded by sigle quotes or double quotes.
+Option values can be surrounded by single quotes or double quotes.
 Values without quotes are also accepted. Multi-valued options are
@@
 
-The information collector scripts are divided between several seperate scripts:
+The information collector scripts are divided between several separate scripts:
 \begin{itemize}
 \item \texttt{cluster+qju.pl} - driver for information collection.
-It calls all other infomation collectors and prints results in XML.
+It calls all other information collectors and prints results in XML.
 The information collection is done by one single invocation of this script
@@
 
-All information about the hosting environment and queues is collected by one invocation of the perl script named
+All information about the hosting environment and queues is collected by one invocation of the Perl script named
 \texttt{cluster+qju.pl}. The XML output is printed to \texttt{STDOUT} and it includes
@@
 
-Sun Grid Engine, formely known as Codine, is an open source batch
+Sun Grid Engine, formerly known as Codine, is an open source batch
 system maintained by Sun. Runs on Linux, Solaris.
@@
 Set up one or more SGE queues for access by grid users. Queues can
-be shared by normal and grid users. In case you want to set up nore
+be shared by normal and grid users. In case you want to set up more
 than one ARC queue, make sure that the corresponding SGE queues have

### 2008-08-03 doc/backend_interface/Backends.tex
backends doc update

@@
 \end{quote}
-\texttt{config\_update\_from\_section <section\_name>}
+\texttt{config\_import\_section <section\_name>}
 
@@
 
-\texttt{config\_update\_from\_section common}
+\texttt{config\_import\_section common}
 
-\texttt{config\_update\_from\_section grid-manager}
+\texttt{config\_import\_section grid-manager}
 
-\texttt{config\_update\_from\_section infosys}
+\texttt{config\_import\_section infosys}
 
@@
 The information collection is done by one single invocation of this script
-\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm})
-\item \texttt{InfoChecker.pm} - used by InfoCollector to validate options and results against a simple 'schema' (not XML Schema)
+\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm}).
+\item \texttt{InfoChecker.pm} - used by InfoCollector to validate options and results against a sort of ``schema'', as implemented by \texttt{InfoChecker.pm}.
 \item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
 \item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments ...)
-\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...)
+\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...) by calling the appropriate LRMS plugin.
+It also does validation of input options to the plugin and of the data returned by the plugin.
 \item \texttt{BATCH\_SYSTEM\_NAME.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework
@@
 \item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
-\item \texttt{ARC0ClusterSchema.pm} - description of the info produced by ARC0ClusterInfo
-\item \texttt{ARC1ClusterSchema.pm} - description of the info produced by ARC1ClusterInfo
+\item \texttt{ARC0ClusterSchema.pm} - ``schema'' of the info produced by ARC0ClusterInfo
+\item \texttt{ARC1ClusterSchema.pm} - ``schema'' of the info produced by ARC1ClusterInfo
 \end{itemize}
@@
 
-The LRMS module should implement four functions as defined by the
-\texttt{LRMS.pm}: \texttt{Cluster\_info}, that provides general information
-about the local cluster, \texttt{Queue\_info}, that provides information
-about a specified queue, \texttt{Jobs\_info} that gives information
-about a list of specified jobs, and finally \texttt{user\_info} that
-gives information about a grid user mapped to a local UNIX account.
-All the functions are called with a hash containing the contents of
-arc.conf.
+Each LRMS module should implement two functions: \texttt{lrms\_info} and
+\texttt{get\_lrms\_options\_schema}.
 
-\subsubsection{cluster\_info}
+\subsubsection{get\_lrms\_options\_schema}
 
-\texttt{Cluster\_info} is called only with the config hash. It should
-then return an associative array with the following keys and the appropriate
-value.
+This function is called without arguments and should return a ``schema'' hash
+declaring the configuration options specific for this LRMS plugin. This
+``schema'' conforms to the format understood by \texttt{InfoChecker.pm} and is
+used to validate input options by \texttt{LRMSInfo.pm}.
 
+\subsubsection{lrms\_info}
+
+\texttt{lrms\_info} returns all LRMS specific information. This
+includes information about the cluster, queues, jobs and local users. The
+function is called with an options hash of hashes. It should then return an
+hash of hashes with the following keys and the appropriate value.
+
+\begin{table}
+\caption{The \texttt{lrms\_info} hash} \label{t:lrms_info}
+\begin{center}
 \begin{tabular}{|c|c|}
 \hline
-key&
-value\tabularnewline
+key & value\\
+\hline \hline
+lrms\_type & the type of LRMS e.g PBS\\
 \hline
+lrms\_version & the version of the LRMS\\
 \hline
-lrms\_type&
-the type of LRMS e.g PBS\tabularnewline
+totalcpus & total number of CPUs in the cluster\\
 \hline
-lrms\_version&
-the version of the LRMS\tabularnewline
+queuedcpus & total number of CPUs requested by LRMS queuing jobs (both grid and
+non-grid)\\
 \hline
-totalcpus&
-total number of CPUs in the cluster\tabularnewline
+usedcpus & CPUs in the LRMS that are currently in use either by grid or
+non-grid jobs\\
 \hline
-queuedcpus&
-total number of CPUs requested by LRMS queuing jobs (both grid and
-non-grid)\tabularnewline
+cpudistribution & number of CPUs in a node and number of each type e.g.
+{}``8cpu:5 2cpu:100''\\
 \hline
-usedcpus&
-CPUs in the LRMS that are currently in use either by grid or non-grid
-jobs\tabularnewline
+queues & a hash with keys being queue names and values being
+\texttt{queue\_info} hashes as described in table \ref{t:queue_info} \\
 \hline
-cpudistribution&
-number of CPUs in a node and number of each type e.g. {}``8cpu:5
-2cpu:100''\tabularnewline
+jobs & a hash with keys being LRMS job IDs and values being
+\texttt{jobs\_info} hashes as described in table \ref{t:jobs_info} \\
 \hline
 \end{tabular}
+\end{center}
+\end{table}
 
-\subsubsection{queue\_info}
-
-In addition to the config \texttt{queue\_info} is called with the
-name of the queue to get information on. The result should take the
-following form.
-
+\begin{table}
+\caption{The \texttt{queue\_info} hash}
+\label{t:queue_info}
+\begin{center}
 \begin{tabular}{|c|c|}
 \hline
-key&
-value\tabularnewline
+key & value\\
+\hline \hline
+status & available slots in queue, negative number signals error\\
 \hline
+maxrunning & limit on number of running jobs\\
 \hline
-status&
-available slots in queue, negative number signals error\tabularnewline
+maxqueuable & limit on number of jobs queued on this queue\\
 \hline
-maxrunning&
-limit on number of running jobs\tabularnewline
+maxuserrun & limit on number of running jobs per user\\
 \hline
-maxqueuable&
-limit on number of jobs queued on this queue\tabularnewline
+maxcputime & limit on maximum CPU time for a job in this queue in minutes\\
 \hline
-maxuserrun&
-limit on number of running jobs per user\tabularnewline
+mincputime & limit on minimum CPU time for a job in this queue in minutes\\
 \hline
-maxcputime&
-limit on maximum CPU time for a job in this queue\tabularnewline
+efaultcput & default CPU time limit for a job in this queue in minutes\\
 \hline
-mincputime&
-limit on minimum CPU time for a job in this queue\tabularnewline
+maxwalltime & limit on maximum wall time for a job in this queue in minutes\\
 \hline
-defaultcput&
-default CPU time limit for a job in this queue\tabularnewline
+minwalltime & limit on minimum wall time for a job in this queue in minutes\\
 \hline
-maxwalltime&
-limit on maximum wall time for a job in this queue\tabularnewline
+defaultwallt & default wall time limit for a job in this queue in minutes\\
 \hline
-minwalltime&
-limit on minimum wall time for a job in this queue\tabularnewline
+running & number of CPUs used by running jobs (both grid and non-grid) in the
+queue\\
 \hline
-defaultwallt&
-default wall time limit for a job in this queue\tabularnewline
+queued & number of CPUs requested by queuing jobs in the queue\\
 \hline
-running&
-number of CPUs used by running jobs (both grid and non-grid) in the
-queue\tabularnewline
+totalcpus & number of CPUs available to the queue\\
 \hline
-queued&
-CPUs requested by queuing jobs in the queue\tabularnewline
-\hline 
-totalcpus&
-number of CPUs available to the queue\tabularnewline
+users & a hash with keys being local UNIX user names and values being
+\texttt{users\_info} hashes as described in table \ref{t:users_info} \\
 \hline
 \end{tabular}
+\end{center}
+\end{table}
 
-\subsubsection{jobs\_info}
-
-Like the rest the first argument is the config hash, then comes the
-queue name and finally a list of job ids. Once again an associative
-mapping should be made for each job. The key value pairs are:
-
+\begin{table}
+\caption{The \texttt{users\_info} hash}
+\label{t:users_info}
+\begin{center}
 \begin{tabular}{|c|c|}
 \hline
-key&
-value\tabularnewline
+key & value\\
+\hline \hline
+freecpus & number of freely available CPUs with their time limits in minutes
+(see ComputingService.FreeSlotsWithDuration in\cite{glue2})\\
 \hline
-\hline 
-status&
-LRMS jobstatus mapping%
-\footnote{LRMS job states mapping is described in~\cite{is} %
-}:Running->'R', Queued->'Q', Suspended->'S', Exiting->'E', Other->'O'\tabularnewline
-\hline 
-rank&
-the job's position in the LRMS queue\tabularnewline
-\hline 
-mem&
-the memory usage of the job in kBs\tabularnewline
-\hline 
-walltime&
-consumed wall-time in minutes\tabularnewline
-\hline 
-cputime&
-consumed CPU-time in minutes\tabularnewline
-\hline 
-reqwalltime&
-wall-time requested by job in minutes\tabularnewline
-\hline 
-reqcputime&
-CPU-time requested by job in minutes\tabularnewline
-\hline 
-nodes&
-list of execution hosts\tabularnewline
-\hline 
-comment&
-array of string comments about the job in LRMS, can be an empty array\tabularnewline
+queuelength & estimated queue length for the specified user\\
 \hline
 \end{tabular}
+\end{center}
+\end{table}
 
-\subsubsection{user\_info}
-
-\texttt{Users\_info} is given the config hash, a queue name and a
-list of local UNIX accounts (UIDs). It should then return an associative
-array containing for each UID the number of free CPUs available for
-that local account on the given queue and the number of queued jobs
-on that queue.
-
+\begin{table}
+\caption{The \texttt{jobs\_info} hash}
+\label{t:jobs_info}
+\begin{center}
 \begin{tabular}{|c|c|}
 \hline
-key&
-value\tabularnewline
+key & value\\
+\hline \hline
+status & LRMS jobstatus mapping\footnotemark[1]:
+Running->'R', Queued->'Q', Suspended->'S', Exiting->'E', Other->'O'\\
 \hline
+rank & the job's position in the LRMS queue\\
 \hline
-freecpus&
-number of freely available CPUs with their time limits in minutes
-(see nordugrid-authuser-freecpus in~\cite{is})\tabularnewline
+mem & the memory usage of the job in kBs\\
 \hline
-queuelength&
-estimated queue length for the specified user\tabularnewline
+walltime & consumed wall-time in minutes\\
+\hline
+cputime & consumed CPU-time in minutes\\
+\hline
+reqwalltime & wall-time requested by job in minutes\\
+\hline
+reqcputime & CPU-time requested by job in minutes\\
+\hline
+nodes & list of execution hosts\\
+\hline
+comment & array of string comments about the job in LRMS, can be an empty
+array\\
 \hline
 \end{tabular}
+\footnotetext[1]{LRMS job states mapping is described in\cite{is}}
+\end{center}
+\end{table}
 

### 2008-08-04 doc/hed/ARCHED_article.tex
Extending introduction of HED document by adding some design ideas - everybody is welcome to add own.

@@
 It is the central part in a new very lightweight incarnation of ARC that is aimed at - but not limited to - providing Web Service.
+
+Whole design of the HED is built around idea of flexibility and modularity.
+Inside HED developer or deployer is supposed to use only as much as needed.
+This is why the HED mostly consists of pluggable modules with some glue between them.
+
+Because in it's current state it mostly provides modules for building SOAP based Web Services there is common misunderstanding that HED is just yet another implementation of Web Services development framework like Axis, gSOAP, XFire or any other of numerous implementations.
+But instead the idea of HED is to provide framework for gluing functionalities and not a re-implementation of various standards.
+Effectively that means if Apache 2 web server is considered by developers as necessary for serving as frontend to services there could be plugin written which puts Apache 2 into a chain of other plugins of the HED.
+
+In current implementation you will not find Apache or Axis plugins.
+That is because developers of HED were concerned very much about lightweigtness of solution and needed rather simple and light despite simplified implementation of supported protocols.
+As result some basic things like SOAP and HTTP are implemented inside HED.
+While external software is used whenever that is found to be appropriate - like in case of TLS, (Grid)FTP, LDAP and some other features.
+That does not exclude possibility to have plugins using entirely external solutions either developed or accepted from third parties.
+
+The HED is relatively young framework and there are quite many rough corners and non-flexible solutions.
+The situation will hopefully improve with time.
+We would be grateful for any suggestions how to improve architecture and code of the HED.
+Statements like "you are doing crap, i'll better use Axis" are understood but not welcome and are usually reacted adequately.
 
@@
 
+\subsubsection{PDC}
+
+\subsubsection{Security Handlers}
+
 \subsubsection{Alternative implementation languages}

### 2008-08-04 doc/glite_gateway/glite_gateway.tex
gLite gateway technical documentation

\documentclass{article}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{caption}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Ivan Marton, Peter Stefan},
  pdftitle = {The ARC gLite Gateway},
  pdfsubject = {ARC gLite gateway manual},
  pdfkeywords = {cream2, ARC1, Advanced Resource Connector, manual, users's manual, developers guide, gateway, gLite},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
\bibliographystyle{IEEEtran}                       %a nice bibliography style
\def\efill{\hfill\nopagebreak}
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0pt}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
%\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}

\def\documenttitle{The ARC gLite Gateway}
\captionsetup{margin=20pt,font={footnotesize, it},labelfont=bf}
\begin{document}
\def\today{\number\day/\number\month/\number\year}
\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{logo-knowarc.png}}
\end{tabular}

\hrulefill

%-------- Change this to Knowarc_D<w>.<d>-<n>_<yy>

{\raggedleft Knowarc\_D3.2-1\_08\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large \documenttitle}\Large \par}
%\vspace*{0.5cm}
%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large Paper subtitle}\large \par}
\vspace*{1.5cm}
%%%%---- A list of authors ----
{\centering \large Ivan Marton\footnote{martoni@niif.hu} \large \par}
{\centering \large Peter Stefan\footnote{stefan@niif.hu} \large \par}
%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}
\tableofcontents
\newpage
\section{Introduction}
\label{Introduction}
The purpose of this document is to describe how Advanced Resource Connector version 1 (ARC1\footnote{\url{http://www.nordugrid.org/}}) users can execute grid jobs on EGEE CREAM2\footnote{\url{http://grid.pd.infn.it/cream/}} resources.\par In sections \ref{Introduction} and \ref{Interoperability} a brief introduction on interoperability issues will be given. Section \ref{Building} describes how to build the gateway library, section \ref{Users guide} shows how it can be used. Section \ref{Developers notes} gives a tiny developer-oriented introduction on how the gateway library is built up, while section \ref{Outstanding issues} addresses the current issues with the code. Section \ref{Conclusions} concludes the technical document.
\section{Interoperability}
\label{Interoperability}
Interoperability has been one of the most important disciplines in recent grid middleware development. This term is often used in the sense of accessing resources operated by one kind of grid middleware (e.g. gLite) from a user interface operated by another kind (e.g. ARC1). There are two main possibilities to achieve this result:
\begin{itemize}
\item either to use a special infrastructure element, the gateway, which performs a full data structure and protocol translation between the two inter-operating grid middleware solutions,
\item or to use standard or close-to-standard interfaces on both sides.
\end{itemize}
The latter solution which is preferred from standardization point of view and also has been chosen by ARC1 developers is illustrated in figure \ref{fig:Chart 1}.\par
\begin{figure}[ht]
\centering{\resizebox{\textwidth}{!}{\includegraphics{D3-2-1-figure.png}}
\caption{\label{fig:Chart 1}This chart illustrates how the ARC1$\rightarrow$gLite gateway library works in general. After successful authentication and credential delegation to the CREAM2 resource, users submit grid jobs through it.  A wrapper script at the computing element may pull input files from different storage elements using storage protocol, say, GridFTP. Users get job information through CREAM2 and may send management commands, like kill a job or undelegate credentials.} }
\end{figure}
\section{CREAM2}
The main purpose of this project was to provide access to gLite CREAM2-enabled resources from ARC1 user interface, but not the other way around: i.e. accessing ARC resources from gLite is addressed in the EGEE project\footnote{\url{http://www.eu-egee.org/}}.\par
CREAM2 was chosen for the following reasons:
\begin{itemize}
\item It is a web service interface that fits to the built-in web service implementation of ARC1.
\item It contains numerous improvements compared to the original CREAM interface. (Unfortunately CREAM2 is not backwards compatible with CREAM.)
\item It supports job status queries compared to the other web service interfaces of gLite, like WMProxy.
\item It is an official interface to gLite computing elements and as EGEE evolves it is expected that more and more computing elements will have CREAM2 interfaces.
\end{itemize}
\section{Building, dependencies}
\label{Building}
The ARC$\rightarrow$gLite gateway software is incorporated into ARC1 releases, so you only need to download either the freshest ARC1 from the source tree\footnote{\url{http://svn.nordugrid.org/}} and compile it or to install it from package\footnote{\url{http://download.nordugrid.org/software/nordugrid-arc1/}}.\par
The gateway library code has two dependencies:
\begin{itemize}
\item Globus Toolkit is needed to enable GridFTP functionality to be able to upload input files from the user interface to the CREAM computing node.
\item VOMS is needed to allow users authentication on a CREAM resource.
\end{itemize}
In both cases it is highly recommended to download these pieces of software along with their NorduGrid-specific Globus\footnote{\url{http://download.nordugrid.org/software/globus/}} and VOMS\footnote{\url{http://download.nordugrid.org/software/voms/}} patches from the NorduGrid repository. In case you are interested in the details of patching, then consult the NorduGrid wiki page\footnote{\url{http://wiki.nordugrid.org/images/4/46/NorduGrid-globus.patch.txt}}. It might also be interesting to read how to compile Globus Toolkit. Some more details on this topic can be found on the NorduGrid Globus compilation page\footnote{\url{http://wiki.nordugrid.org/index.php/Globus\_Libraries}}. There are also some details on how to compile VOMS from source at NorduGrid VOMS compilation page\footnote{\url{http://wiki.nordugrid.org/index.php/VOMS\_compiling\_detailes}}.
\section{Using the ARC${\rightarrow}$gLite Gateway from the Command Line}
\label{Users guide}
There are seven different command line tools for using the gLite gateway functions. In this section these tools will be presented in detail through examples. The job descriptions and services shown are just simple examples, they might be different in real-life usage.\par
Every tool has a short manual page being accessible by using the \textbf{man} on-line manual reader command or some more help can be achieved by using the $<$command$>$ -? command line option.\par
Before getting into the details of the gatewy toolkit, the authentication and authorization processes need clarification.\par
In the different grid systems the most widespread authentication method is the certificate based authentication. GLite also requires VOMS-enabled proxy X509 certificates to access resources. It means that you should be a member of the Virtual Organisation that is valid on the remote site. If the VOMS package has been correctly installed then the following command provides the necessary VOMS proxy certificate.\par
\begin{shaded}\verb#$ voms-proxy-init -voms <Your Virtual Organisation>#\end{shaded}
For example:
\begin{shaded}\verb#$ voms-proxy-init -voms knowarc.eu#\end{shaded}
\subsection{glitedelegate}
\label{glitedelegate}
If you want to submit a job to a gLite server, first you have to own a valid delegation on the remote site. You can either use an old, but still valid delegation identifier, or register a new one. There is a command line tool to register a delegation on the resource to be used.\par
The usage is very easy and straighforward. All you need is a working CREAM2 service URL, and a locally unique arbitrary delegation ID. This delegation ID will be associated to the remote resource. If you are not sure wether your favourite delegation ID is occupied or not, then try to delete it before registering it again. (See Section \ref{gliteundelegate} for further details!) The delegation command has the following syntax:
\begin{shaded}\verb#$ glitedelegate <delegation ID> <service URL>#\end{shaded}
For example:
\begin{shaded}\verb#$ glitedelegate test_delegation //#\\*
\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}
\subsection{glitesub}
\label{glitesub}
In the previous section it was shown how to create delegation. (See Section \ref{glitedelegate}!) This function, the job submission, is the only one which needs the delegation ID.\par The next step is to prepare the necessary input files and the job description!\par
The glitesub command syntax is the following:
\begin{shaded}\verb#$ glitesub -D <delegation ID> <service URL> <job description> <info file>#\end{shaded}
Here the delegation ID is the previously registered identification. The service URL will be different than in the delegation example above, because this is not the URL of the delegation, but the execution service. The job description must be in JDL format like in the example below. The info file stores information about job submitted. It should either be a non-existing file or it is overwritten. This file contains the job ID, the input sandbox URL, the output sandbox URL and the service URL. The submission command will take care of uploading the necessary input files up to the execution node.\par
It is important that the \textit{VirtualOrganisation} and \textit{QueueName} attributes should be the same as your virtual organisation name and the corresponding queue name on the server side.\par

An examle JDL description:
\begin{framed}
\verb#[#\\*
\verb#Executable = "/bin/hostname";#\\*
\verb#StdInput = "std.in";#\\*
\verb#StdOutput = "std.out";#\\*
\verb#StdError = "std.err";#\\*
\verb#BatchSystem = "pbs";#\\*
\verb#VirtualOrganisation = "knowarc.eu";#\\*
\verb#InputSandbox = {"std.in"};#\\*
\verb#OutputSandbox = {"std.out","std.err"};#\\*
\verb#QueueName = "knowarc.eu";#\\*
\verb#OutputSandboxDestURI = { "gsiftp://localhost/std.out", "gsiftp://localhost/std.err" };#\\*
\verb#]#
\end{framed}
An example how to use this command:
\begin{shaded}\verb#$ glitesub -D test_delegation //#\\*
\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/CREAM2 description.jdl job.info#\end{shaded}
\subsection{glitestat}
\label{glitestat}
The \textbf{glitestat} tool extracts job information about a CREAM2 job identified by the information file. This file made by command \textbf{glitesub} is the only argument needed here.\par
Usage:
\begin{shaded}\verb#glitestat <info file>#\end{shaded}
For example:
\begin{shaded}\verb#glitestat job.info#\end{shaded}
The possible responses and their meanings are the following:
\begin{itemize}
\item ACCEPTING - The job submission is comleted but the job is not yet scheduled
\item SUBMITTING - Scheduling in progress
\item INLRMS:Q - The job is already at the local resource manager system and it is queued
\item INLRMS:R - The job is waiting at the local resource manager system for running
\item INLRMS:S - The job is actually running
\item KILLED - The job was terminated
\item FINISHED - The job has finished
\item FAILED - The job had some failure
\item FAILES - The job had some failure at LRMS level
\item EXECUTED - The job has been finished and there is no stating information available
\end{itemize}
\subsection{glitekill}
\label{glitekill}
The \textbf{glitekill} command can be used for killing a remote job. The only necessary argument is the information file made by \textbf{glitesub}. (See Section \ref{glitesub} for further details!) This command initiates stopping the job. You can check the effect by using \textbf{glitestat}. (See Section \ref{glitestat} for its usage!)\par
Usage:
\begin{shaded}\verb#glitekill <info file>#\end{shaded}
For example:
\begin{shaded}\verb#glitekill job.info#\end{shaded}
\subsection{gliteclean}
\label{gliteclean}
Either after killing a job or because of some remote error occurs garbage files may remain on the server. These files can be purged by using command \textbf{gliteclean}. If you intentionally kill a job on the computing node, the job related stuff might also remain as garbage in the remote queue. This command also removes the job from the gLite server queue. It is important to note that this command removes the local information file as well. It is very useful to run this command after killing a job. (See Section \ref{glitekill}!)\par
Usage:
\begin{shaded}\verb#gliteclean <info file>#\end{shaded}
For example:
\begin{shaded}\verb#gliteclean job.info#\end{shaded}
\subsection{gliteundelegate}
\label{gliteundelegate}
After your work is complete it is possible to unregister and delete the delegation entry from the remote site by using command \textbf{gliteundelegate}. You can also use it, if you want to register a delegation ID but are unsure whether it is used or not. Put the delegation service URL into the argument!\par
Usage:
\begin{shaded}\verb#$ gliteundelegate <delegation ID> <service URL>#\end{shaded}
For example:
\begin{shaded}\verb#$ gliteundelegate test_delegation //#\\*
\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}
\section{Developers notes}
\label{Developers notes}
This section describes how to use the library on your own client. The base class of the gLite client is written as an ARC Client Component (ACC) but also can be used as a standalone library. There is nothing more to be done just to link the \textbf{CREAMClient}, \textbf{arcclient} and \textbf{arccommon} libraries installed with the ARC1 sofware and to create an instance of the \textit{CREAMClient} class.\par
This client object has a very easy-to-use and intuitive interface. It uses the VOMS proxy certificate to build the secure channel to the server and to sign that of the server's. Actually the client looks for this certificate just at its default location. If you want to store it somewhere else, you have to modify the source code itself.\par
The different functions throw \textit{CREAMClientError} exceptions that should be caught in the application.\par
The following sections present the set of functions provided by the library.
\subsection{CREAMClient(Arc::URL, Arc::MCCConfig)}
The constructor of the class performs the necessary initialization work. It receives two arguments: the service URL and the ARC message chain component configuration file to establish the communication channel to the server. These pieces of information are indispensable to create your client and to communicate with the remote site.
\subsection{setDelegationId(std::string)}
Besides the constructor this is the other generic-purpose function that can be used in CREAM2 client applications. This function sets the previously registered and referred delegation ID on the client. The function is very simple: just set a private variable and returns with no value.
\subsection{createDelegation(std::string)}
The \textbf{createDelegation} function is about to perform the whole delegation registration process. It sends a \textit{getProxyReq} message having the requested delegation ID, signs the received certificate and sends it back in a \textit{putProxy} SOAP message to the server. These three steps constitute the delegation registration process.
As almost every funtion, this one has no return value either. If there are any problems during the communication, either locally in the channel or at the remote site, the function throws \textit{CREAMClientError} described previously. If there are no exceptions thrown, then the command is considered to have been successfully completed.
\subsection{destroyDelegation(std::string)}
This method has the reverse functionality of the \textbf{createDelegation} function. It sends the CREAM2 \textit{destroy} message to the remote server.
\subsection{submit(std::string)}
The \textbf{submit} function is the most complex part of the class. It translates the job description received as the argument, registers the job with a \textit{JobRegisterRequest} message, uploads the locally stored files, if necessary, then enables the job execution on the server by sending a \textit{JobStartRequest} message. Finally it returns with a \textbf{creamJobInfo} object that contains a \textit{jobId} (job indetifier), a \textit{creamURL} (as the service URL), an \textit{ISB} (reference to the Input Sandbox) and an \textit{OSB} (Output Sandbox) member to describe the registered job. In the \textbf{glitesub} command (See Section \ref{glitesub} for details) these pieces of information are stored in the information file.
\subsection{stat(std::string)}
This method queries the job status from the gLite CREAM2 server. The return value is the job status translated to the ARC terminology. The possible values are presented in the section of the \textbf{glitestat} command line tool. (See Section \ref{glitestat}!)
\subsection{cancel(std::string)}
The \textbf{cancel} function sends a \textit{JobCancelRequest} SOAP message to the server and handles the emerging exceptions. It can be used for canceling a remotely registered and possibly running job.
\subsection{purge(std::string)}
The \textbf{purge} function sends a \textit{JobPurgeRequest} message to server and throws an exception in case of any emerging problems.
\subsection{An example gLite client}
Finally, here is an example code to present how easy and simple is to write a new client. This code sample requests the job status of a previously submitted remote job, and writes it to the standard output. The client is written in C++.
\begin{framed}
\verb?#include "CREAMClient.h"?\\*
\verb?#include <iostream>?\\*
\\*
\verb?int main(int argc, char* argv[]){?\\*
\verb?  Arc::URL url( SERVICE_URL );?\\*
\verb?  Arc::MCCConfig cfg;?\\*
\\*
\verb?    Arc::Cream::CREAMClient gLiteClient(url,cfg);?\\*
\verb?    try {?\\*
\verb?        std::cout << "Job status: " << gLiteClient.stat( JOBID ) << std::endl;?\\*
\verb?    } catch (CREAMClienException& cce) {?\\*
\verb?        std::cerr << "ERROR: " << cce.what() << std::endl;?\\*
\verb?        return 1;?\\*
\verb?    }?\\*
\verb?    return 0;?\\*
\verb?}?
\end{framed}
It is of course mandatory to define \textit{SERVICE\_URL} and \textit{JOBID} strings to their real values to reach the proper functionality.
\section{Outstanding issues}
\label{Outstanding issues}
Even though the gateway library code is ready for production use, there are a couple of issues to be handled in the upcoming development.
\subsection{Full ARC1 client integration}
Full ARC1 client integration basically means that the previously presented commands will disappear and their functionality will be available in the standard ARC1 client toolkit. However there are two dependencies to be met to achieve this stage:
\begin{itemize}
\item On one hand ARC1 broker along with the BDII-capable target object should be finalized.
\item On the other hand a conflicting Globus Toolkit OpenSSL and the standard OpenSSL issue is to be investigated.
\end{itemize}
The latter issue was revealed when the VOMS proxy certificate delegation method was implemented in the gateway library. Since in the ARC1 client-side library there is a built-in Message Chaining Component (MCC) which already uses Globus Toolkit OpenSSL functions to enable establishing secure communication. Linking glite* tools directly against the standard OpenSSL libs caused congesting SSL functions which led to the breakdown of the whole library. The solution had been not linking the standard OpenSSL libs, just the patched OpenSSL in the Globus Toolkit.
\subsection{xRSL${\rightarrow}$JDL translation and related problems}
It is possible to express jobs in XRSL which have no equivalent in JDL. Speciﬁcally, in XRSL it is possible for input files downloaded from storage servers to have their names changed. This is not possible in JDL. So a user might in XRSL ask for this:
\begin{shaded}
\verb#(inputfiles=(gsiftp://interop.dcgc.dk/storage/datafile1.txtinput.txt))#
\end{shaded}
That is not possible to express in JDL. It only allows the file to be downloaded as datafile1.txt. This could be solved by a wrapper script that renames files according to the XRSL speciﬁcation. This does not solve our problem completely however. For example if a user wants to concatenate the stdout from two jobs, he would write something like the following in XRSL:
\begin{shaded}
\verb#(inputfiles=(gsiftp://interop.dcgc.dk/storage/job1/output.txtinput1.txt)#\\
\verb#            (gsiftp://interop.dcgc.dk/storage/job2/output.txtinput2.txt))#
\end{shaded}
This cannot be solved just using the renaming method, because both input files would be retrieved before renaming is done and they would therefore overwrite each other.\par
The solution is to recognize jobs of this type and simply barring them from using gLite resources or to let such jobs fail with an error message explaining that the user has to express their job in a different way.
\subsection{RunTime Environments related issues}
Many grid jobs rely on software being available at the execution location. The requirement as well as the initialization 
of such software is expressed through the use of Run-Time Environments (RTEs). While the library is capable of 
translating the RTE request itself it is not capable of translating the name of the RTE. No standardized cross Grid 
nomenclature has been adopted, nor are the authors aware of any such activity. A temporary solution would be for each 
VO to have a remotely accessible list of RTEs and their names in each job description language. This could then be used for translating jobs.
\section{Conclusions}
\label{Conclusions}
The document has described an ARC1 library extension that allows ARC1 grid users to access gLite CREAM2 computing resources. The current version of the code allows direct access to resources but as the new ARC1 brokering solution comes out the gateway functionality will be integrated fully into the ARC1 client code. The same version will also be able to access the gLite information system to extract CREAM2 resource information.\par
A built-in xRSL${\rightarrow}$JDL translator will also enable to use purely xRSL job description on ARC1 user interfaces.
\end{document}

### 2008-08-04 doc/backend_interface/Backends.tex
backends doc update

@@
 
-The information system interface consists of set of Perl scripts which
+The information system interface consists of a set of Perl scripts which
 generates an XML output to \texttt{STDOUT}.
@@
 \item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm}).
-\item \texttt{InfoChecker.pm} - used by InfoCollector to validate options and results against a sort of ``schema'', as implemented by \texttt{InfoChecker.pm}.
+\item \texttt{InfoChecker.pm} - used by \texttt{InfoCollector} to validate options and results against a ``schema''.
 \item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
@@
 \item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
-\item \texttt{ARC0ClusterSchema.pm} - ``schema'' of the info produced by ARC0ClusterInfo
-\item \texttt{ARC1ClusterSchema.pm} - ``schema'' of the info produced by ARC1ClusterInfo
 \end{itemize}
@@
 ``schema'' conforms to the format understood by \texttt{InfoChecker.pm} and is
-used to validate input options by \texttt{LRMSInfo.pm}.
+used to validate input options to the plugin.
 
@@
 
-\texttt{lrms\_info} returns all LRMS specific information. This
-includes information about the cluster, queues, jobs and local users. The
-function is called with an options hash of hashes. It should then return an
-hash of hashes with the following keys and the appropriate value.
+The \texttt{lrms\_info} function returns all LRMS specific information. This
+includes information about the cluster, queues, jobs and local user limits. The
+function is called with a hash of hashes containing all options required by the
+LRMS plugin. Some options are common to all plugins, and are described in table
+\ref{t:lrms_opts}.
 
 \begin{table}
-\caption{The \texttt{lrms\_info} hash} \label{t:lrms_info}
+\caption{The \texttt{lrms\_options} hash used as input for the \texttt{lrms\_info} function} \label{t:lrms_opts}
 \begin{center}
-\begin{tabular}{|c|c|}
+\begin{tabular}{|c|p{12cm}|}
+\hline
+key & value\\
+\hline \hline
+jobs & array with local job IDs \\
+\hline
+queues & hash with queue names as keys and values being
+\texttt{queue\_options} hashes (table \ref{t:queue_opts}) \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\caption{The \texttt{queue\_options} hash} \label{t:queue_opts}
+\begin{center}
+\begin{tabular}{|c|p{12cm}|}
+\hline
+key & value\\
+\hline \hline
+users & array with local UNIX user names \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+The \texttt{lrms\_info} function should then return a
+hash of hashes with the sctructure described in table \ref{t:lrms_info}.
+
+\begin{table}
+\caption{The \texttt{lrms\_info} hash returned by the \texttt{lrms\_info} function} \label{t:lrms_info}
+\begin{center}
+\begin{tabular}{|c|p{12cm}|}
 \hline
@@
 \begin{center}
-\begin{tabular}{|c|c|}
+\begin{tabular}{|c|p{12cm}|}
 \hline
@@
 status & LRMS jobstatus mapping\footnotemark[1]:
-Running->'R', Queued->'Q', Suspended->'S', Exiting->'E', Other->'O'\\
+Running $\Rightarrow$ 'R',
+Queued $\Rightarrow$ 'Q',
+Suspended $\Rightarrow$ 'S',
+Exiting $\Rightarrow$ 'E',
+Other $\Rightarrow$ 'O'\\
 \hline
@@
 \item The required queue attributes can be checked as \texttt{<PBS installation
-path>/bin/qstat -f -Q queuename.} There MUST be a max\_user\_run or
+path>/bin/qstat -f -Q queuename}. There MUST be a max\_user\_run or
 a max\_running attribute listed with a REASONABLE value. 
@@
 
-\phantomsection
-%\addcontentsline{toc}{subsubsection}{LocalID}
-\hspace*{0.5cm}
-\begin{shaded}
-\textbf{LocalID}
-\end{shaded}
-\begin{tabular}{lp{10cm}}  
-Attribute value:& LocalID\_t\\
-Example:&  \verb#TODO#\\
-Related xRSL:& \verb#NA#\\
-UI role:& TODO \\
+Hardcoded Properties:
 
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & Value \\
+\hline \hline
+Capability & \texttt{executionmanagement.jobexecution} \\
+\hline
+Type & \texttt{org.nordugrid.arex} \\
+\hline
+QualityLevel & \texttt{development} \\
+\hline
+Complexity & \texttt{endpoint=1,share=N,resource=1}, where N is the number of queues in the configuration file\\
+\hline
 \end{tabular}
 
-Description: An opaque identifier local to the Computing Service.
+\vspace*{3mm}
+The following properties referring to job counts are obtained from GridManager's status files:
+
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & Meaning \\
+\hline \hline
+TotalJobs & Number of grid jobs which are not yet finished \\
+\hline
+RunningJobs & Number of grid jobs running \\
+\hline
+WaitingJobs & Number of grid jobs queued in the LRMS and waiting to be submitted to the LRMS \\
+\hline
+\end{tabular}
+
+\subsection{ComputingEndpoint}
+
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & Value\\
+\hline \hline
+URL & \texttt{https://\emph{hostname}:\emph{port}/arex} \\
+\hline
+Technology & \texttt{webservice} \\
+\hline
+Interface & \texttt{OGSA-BES} \\
+\hline
+WSDL & \texttt{https://\emph{hostname}:\emph{port}/arex/?wsdl} \\
+\hline
+Semantics & \texttt{http://www.nordugrid.org/documents/arex.pdf} \\
+\hline
+SupportedProfile & \texttt{WS-I 1.0} and \texttt{HPC-BP} \\
+\hline
+Implementor & \texttt{NorduGrid} \\
+\hline
+ImplementationName & \texttt{ARC1} \\
+\hline
+ImplementationVersion & \texttt{0.9} \\
+\hline
+QualityLevel & \texttt{development} \\
+\hline
+HealthState & \texttt{ok} -- no detection of problems yet\\
+\hline
+ServingState & \texttt{draining} if the configuration has allownew=no, \texttt{production} otherwise \\
+\hline
+IssuerCA & from the \emph{hostcert.pem} file \\
+\hline
+TrustedCA & from scanning the \emph{certificates} directory \\
+\hline
+Staging & \texttt{staginginout} \\
+\hline
+JobDescription & \texttt{ogf:jsdl:1.0}\\
+\hline
+\end{tabular}
+
+\subsection{ComputingShare}
+
+A ComputingShare element is generated for each 'queue' section in the configuration.
+
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & Value\\
+\hline \hline
+BaseType & \texttt{Share} \\
+\hline
+Name & from the configuration \\
+\hline
+Description & ``comment'' configuration option \\
+\hline
+MappingQueue & from the configuration \\
+\hline
+\end{tabular}
+
+\vspace*{3mm}
+The following properties are taken from the corresponding values returned by
+the LRMS plugin. Units are converted as necessary.
+
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & LRMS plugin property \\
+\hline \hline
+MaxCPUTime & maxcputime \\
+\hline
+MinCPUTime & mincputime \\
+\hline
+DefaultCPUTime & defaultcput \\
+\hline
+MaxWallTime & maxwalltime \\
+\hline
+MinWallTime & minwalltime \\
+\hline
+DefaultWallTime & defaultwallt \\
+\hline
+\end{tabular}
+
+\vspace*{3mm}
+
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & Value\\
+\hline \hline
+MaxTotalJobs & maxjobs from configuration\\
+MaxRunningJobs & \\
+MaxWaitingJobs & \\
+\hline
+\end{tabular}
 
@@
 UI role:& ?\\
+Description:& Name of the application environment\\
 \end{tabular}
-
-Description: Name of the application environment.
 
@@
 UI role:& TODO\\
+Description:& Name of the application environment\\
 \end{tabular}
-
-Description: Version of the application environment.
-
-\subsection{ComputingEndpoint}
-
-\subsection{ComputingShare}
 

### 2008-08-04 doc/hed/ARCHED_article.tex
Small adjustments, mainly on a linguistical level.

--- 2008-05-07 doc/hed/ARCHED_article.tex
@@
 
+The whole design of the HED is built around the idea of flexibility and modularity.
+Inside HED the developer or deployer is supposed to use only as much as needed.
+This is why the HED mostly consists of pluggable modules with some glue between them.
+
+Because in it's current state it mostly provides modules for building SOAP based Web Services, it is easy to think that HED is just another Web Services development framework like Axis, gSOAP, XFire or any other of numerous implementations.
+But instead the idea of HED is to provide framework for gluing functionalities and not a re-implementation of various standards.
+Effectively that means if Apache 2 web server is considered by developers as necessary for serving as frontend to services there could be plugin written which puts Apache 2 into a chain of other plugins of the HED.
+
+In the current implementation there are no Apache or Axis plugins.
+That is because the developers of HED were very much concerned about making the solution lightweight and needed an implementation of the supported protocols that was both simple and lightweight.
+As a result essentials like SOAP and HTTP are implemented inside HED, while external software is used whenever that is found to be appropriate - as in the case of TLS, (Grid)FTP, LDAP and some other cases.
+?That does not exclude possibility to have plugins using entirely external solutions either developed or accepted from third parties.?
+
+The HED is a relatively young framework and there are quite a few rough edges and non-flexible solutions.
+The situation will hopefully improve with time.
+We would be grateful for any suggestions how to improve architecture and code of the HED.
+Statements like "you are doing crap, I'll better use Axis" are understood but not welcome and are usually reacted to adequately.
+
 \section{Architecture}
@@
 
-External dependencies needed to be kept to a minimum while also taking into consideration their ubiquity or relative rarity and license related concerns.
+External dependencies needed to be kept to a minimum while also taking into consideration their ubiquity or relative rarity as well as license related concerns.
 Software of this level of complexity must of course depend on many external libraries and components but each such dependenscy has been introduced only after due consideration.

### 2008-08-04 doc/glite_gateway/glite_gateway.tex
service URL query LDAP command added
New ldap query command added.

@@
 \begin{shaded}\verb#$ voms-proxy-init -voms knowarc.eu#\end{shaded}
+CREAM2 service URLs can directly be downloaded from any BDII database of gLite using the following LDAP command:
+\begin{shaded}
+\verb#$ ldapsearch -x -h lcg-bdii.cern.ch -p 2170 -b o=grid '(GlueCEUniqueID=*blah*)'#
+\end{shaded}
+or
+\begin{shaded}
+\verb#$ ldapsearch -x -h lxbra2305.cern.ch -p 2170 -b mds-vo-name=local,o=grid '(GlueCEUniqueID=*cream*)'#
+\end{shaded}
+
 \subsection{glitedelegate}

### 2008-08-04 doc/glite_gateway/glite_gateway.tex
Minor modifications

@@
 \begin{shaded}
-\verb#$ ldapsearch -x -h lxbra2305.cern.ch -p 2170 -b mds-vo-name=local,o=grid '(GlueCEUniqueID=*cream*)'#
+\verb#$ ldapsearch -x -h lxbra2305.cern.ch -p 2170 -b mds-vo-name=local,o=grid //#\\*
+\verb#        '(GlueCEUniqueID=*cream*)'#
 \end{shaded}
@@
 
-An examle JDL description:
+An example JDL description:\\
+\begin{minipage}{\textwidth}
 \begin{framed}
-\verb#[#\\*
-\verb#Executable = "/bin/hostname";#\\*
-\verb#StdInput = "std.in";#\\*
-\verb#StdOutput = "std.out";#\\*
-\verb#StdError = "std.err";#\\*
-\verb#BatchSystem = "pbs";#\\*
-\verb#VirtualOrganisation = "knowarc.eu";#\\*
-\verb#InputSandbox = {"std.in"};#\\*
-\verb#OutputSandbox = {"std.out","std.err"};#\\*
-\verb#QueueName = "knowarc.eu";#\\*
-\verb#OutputSandboxDestURI = { "gsiftp://localhost/std.out", "gsiftp://localhost/std.err" };#\\*
+\verb#[#\\
+\verb#Executable = "/bin/hostname";#\\
+\verb#StdInput = "std.in";#\\
+\verb#StdOutput = "std.out";#\\
+\verb#StdError = "std.err";#\\
+\verb#BatchSystem = "pbs";#\\
+\verb#VirtualOrganisation = "knowarc.eu";#\\
+\verb#InputSandbox = {"std.in"};#\\
+\verb#OutputSandbox = {"std.out","std.err"};#\\
+\verb#QueueName = "knowarc.eu";#\\
+\verb#OutputSandboxDestURI = { "gsiftp://localhost/std.out", "gsiftp://localhost/std.err" };#\\
 \verb#]#
 \end{framed}
+\end{minipage}
 An example how to use this command:
@@
 \begin{shaded}
-\verb#(inputfiles=(gsiftp://interop.dcgc.dk/storage/job1/output.txtinput1.txt)#\\
+\verb#(inputfiles=(gsiftp://interop.dcgc.dk/storage/job1/output.txtinput1.txt)#\\*
 \verb#            (gsiftp://interop.dcgc.dk/storage/job2/output.txtinput2.txt))#

### 2008-08-04 doc/storage_system/arc1-storage-documentation.tex
making security high priority

@@
 
+\section{Security} % (fold)
+\label{sec:security}
+
+The current version of the ARC1 storage system has no implemented solution for security issues.
+Designing and implementing a powerful, flexible, fine-grained solution for authentication and authorization is very important.
+Currently we have plans how to do that, and the design of the storage services makes it easy to integrate the chosen solution into the system.
+
+For additional details, see Section~\ref{sec:plans_for_security}.
+
+% section security (end)
+
 \chapter{Use cases} % (fold)
@@
 
-The security is needed to do proper authorization of the users, and to manage access policies of files and collections.
-ARC has its own policy language, for each file and collection there will be a policy XML document stored as a metadata.
-The storage services will use the properties extracted from the communication channel and these policies to make authorization decisions.
-If the properties and the policies are present, the decision is actually made by the security framework of HED.
-
 Some self-healing mechanisms are not implemented yet, e.g.
@@
 This is not implemented yet.
+
+\section{Plans for security} % (fold)
+\label{sec:plans_for_security}
+
+Security is needed to do proper authorization of the users, and to manage access policies of files and collections.
+ARC has its own policy language, for each file and collection there will be a policy XML document stored as a metadata.
+The storage services will use these policies and the properties extracted from the communication channel to make authorization decisions.
+If the properties and the policies are present, the decision will be actually made by the security framework of HED.
+
+These are the planned actions which can be used for access control:
+\begin{itemize}
+    \item \emph{read}: user can get the list of entries in the collection; user can download the file
+    \item \emph{addEntry}: user can add a new entry to the collection;
+    \item \emph{removeEntry}: user can remove any entry from the collection 
+    \item \emph{delete}: user can delete the collection if it is empty; user can delete a file (if you want to remove a file/collection, then the Bartender needs to remove the entry from the parent collection, and then delete the file/collection itself, so you need to have both permissions)
+    \item \emph{modifyPolicy}: user can modify the policy of the file/collection
+    \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
+    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are key-value pairs)
+\end{itemize}
+When a user has the permission in the Librarian to download a file then the user should have permission to access at least one of the file's replica, so there should be a Shepherd which allows the user the get the file.
+If the Bartender has permission to access the Shepherd, then the Bartender should create an assertion which allows the user to access the file.
+This could be a signed token which contain a policy defining access to particular file.
+But all these are currently just plans.
+
+% section plans_for_security (end)
 
@@
 
-\section{Other plans} % (fold)
+\section{Integrating third-party storage solutions} % (fold)
+\label{sec:integrating_third_party_storage_solutions}
 
-\subsection{Integrating third-party storage solutions} % (fold)
-\label{sub:integrating_third_party_storage_solutions}
-
-To intagrate existing files on a third-party storage to our namespace thus make them accessible through the interface of the ARC1 storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
+To integrate existing files on a third-party storage system to our namespace thus make them accessible through the interface of the ARC1 storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
 It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC1 storage (Section~\ref{sub:data_model}).
@@
 This interface has some methods similar to the Shepherd and some other methods similar to the Bartender.
-% subsection integrating_third_party_storage_solutions (end)
-
-\subsection{Security} % (fold)
-\label{sub:security}
-
-These are the planned actions which can be used for access control:
-\begin{itemize}
-    \item \emph{read}: user can get the list of entries in the collection; user can download the file
-    \item \emph{addEntry}: user can add a new entry to the collection;
-    \item \emph{removeEntry}: user can remove any entry from the collection 
-    \item \emph{delete}: user can delete the collection if it is empty; user can delete a file (if you want to remove a file/collection, then the Bartender needs to remove the entry from the parent collection, and then delete the file/collection itself, so you need to have both permissions)
-    \item \emph{modifyPolicy}: user can modify the policy of the file/collection
-    \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
-    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are key-value pairs)
-\end{itemize}
-When a user has the permission in the Librarian to download a file then the user should have permission to access at least one of the file's replica, so there should be a Shepherd which allows the user the get the file.
-If the Bartender has permission to access the Shepherd, then the Bartender should create an assertion which allows the user to access the file.
-This could be a signed token which contain a policy defining access to particular file.
-But all these are currently just plans.
-
-% subsection security (end)
-
-% section other_plans (end)
+% section integrating_third_party_storage_solutions (end)
 

### 2008-08-04 doc/a-rex/arex_tech_doc.tex
Started removing gridftp and grid-manager from a-rex documentation

@@
-%% LyX 1.4.5 created this file.  For more info, see http://www.lyx.org/.
+%% LyX 1.4.5.1 created this file.  For more info, see http://www.lyx.org/.
 %% Do not edit unless you really know what you are doing.
@@
 A-Rex is one of ARC middleware componets implementing functions of
-so called \emph{Computing Element} (CE). It's made of few components
-- A-Rex Web Service (WS) Interface and Grid Manager (GM). Alternative
-and currently only stable interface to A-Rex is GridFTP Server (GFS).
-
-The aim of the Grid Manager (GM) is to take care of job pre- and post-processing.
-It provides functionality to stage-in files containing input data
-and program modules from wide range of sources and transfer or store
-output results.
-
-Both WS Interface and GridFTP Interface components provide a way to
-submit and control computational tasks (jobs) to be executed by the
-GM and undelying Local Resource Management System. While WS interface
-is currently being developed the GridFTP is considered to be stable
-and prefered one.
+so called \emph{Computing Element} (CE).
+
+The aim of the   A-Rex is to take care of job pre-
+and post-processing. It provides functionality to stage-in files containing
+input data and program modules from wide range of sources and transfer
+or store output results.
+
+The A-Rex implements Web Service (WS) interface which   
+    provides a way to submit and control
+computational tasks (jobs) to be executed by the A-Rex and undelying
+Local Resource Management System.
 
@@
 (SD). Input files are gathered in the SD. The job is supposed to produce
-new data files  in the  SD as well. GM does not guarantee the
-availability of any other places accessible by the job other than
+new data files  in the SD as well. The A-Rex does not guarantee
+the availability of any other places accessible by the job other than
 SD (unless such a place is part of a requested Runtime Environement).
-The SD is also the only place which is controlled by the GM. It is
-is accessible by the user from outside through HTTPS and GridFTP protocols.
-Any file created outside the SD is not controlled by the GM. Any exchange
-of data between client and GM (including also program modules) is
-done through GridFTP \cite{gridftp} or HTTPS protocols depending
-on the selected interface. A URL for accessing input/output files
-is obtained through the Information System or the WS Information Interface
-of A-Rex.
+The SD is also the only place which is controlled by the A-Rex.
+It is  accessible by the user from outside through HTTP(S)  
+protocol. Any file created outside the SD is not controlled by the
+A-Rex. Any exchange of data between client and A-Rex (including
+also program modules) is done through    HTTP(S) protocol
+   . A URL for accessing input/output
+files is       obtained through
+the WS Local Information Description Interface (LIDI) of A-Rex.
 
 Each job gets an identifier (\textit{jobid}). This is a handle which
-identifies the job in the GM and the Information System \cite{is}.
-
-  Jobs are initiated and controlled through GFS or WS. Complete
-job descriptions   (JD)   are passed to the GM either
-through GFS in RSL \cite{rsl} or through WS in JSDL \cite{jsdl}
+identifies the job in the A-Rex and the Information  Interface.
+
+  Jobs are initiated and controlled through   WS interface.
+Complete job descriptions   (JD)   are passed to the
+A-Rex  through       WS in JSDL \cite{jsdl}
 coded description     . Input data files
@@
 
-The main task of the GM is to take care of processing input and output
-data (files) of the job. Input files are gathered in the SD. There
-are two ways to put a file into the SD:
-
-\begin{itemize}
-\item Download is initiated by the GM --  This is the case for
-files     defined in the JD (with name and source).
-      The GM alone is responsible to make
-sure that all required files will be available in the SD.\\
+One of the most important tasks   of the A-Rex is to
+take care of processing input and output data (files) of the job.
+Input files are gathered in the SD. There are two ways to put a file
+into the SD:
+
+\begin{itemize}
+\item Download is initiated by the A-Rex --  This is the case
+for files     defined in the JD (with name and source).
+      The A-Rex alone is responsible to
+make sure that all required files will be available in the SD.\\
 The supported protocols for sources  at the moment are: GridFTP,
@@
 some nonstandard sources are supported. Those are described below.
+The A-Rex fully relies on HED framework \cite{hed} for data transfering
+capabilities. Hence actual set of supported protocols depends on installed
+Data Management Components of the HED.
 \item Upload is initiated by the user directly or through the User Interface
@@
 submission of JD, UI can (and should) use that to upload data files
-which are not otherwise accessible by the GM.  Examples of such
-files  are the main executable of the job,   the
-job's input files, etc. These files can (and should)
-also be specified in the JD.
+which are not otherwise accessible by the A-Rex.  Examples of
+such files  are the main executable of the job,  
+the job's input files, etc. These files can (and
+should) also be specified in the JD.
 \end{itemize}
 \textbf{There is no} other reliable way for a job to obtain input
-data on the CE belonging to NorduGrid. Access to AFS, NFS, FTP, HTTP
-and any other remote data transport during execution of a job is
-not guaranteed (at least not by GM).
-
-Jobs should store output files in thir SD. Like input files,
+data on the CE based on the A-Rex  . Access to
+AFS, NFS, FTP, HTTP and any other remote data transport during execution
+of a job is not guaranteed (at least not by A-Rex).
+
+Jobs should store output files in their SD. Like input files,
 output files  belong into two groups:
@@
 Globus \emph{Replica Location Service } (RLS)  --
-The GM takes care of these files. They have to be specified in the
-JD. If the job fails during any stage of processing, no attept is
-made to transfer those files to their final destination, unless
+The A-Rex takes care of these files. They have to be specified
+in the JD. If the job fails during any stage of processing, no attept
+is made to transfer those files to their final destination, unless
 the  option \emph{preserve=yes} is specified in their URLs.
@@
 
-From the point of view of the GM a job passes through various states.
-Picture \ref{job states diagram} presents a diagram of the possible
-states of a job.%
+From the point of view of the A-Rex a job passes through various
+states. Picture \ref{job states diagram} presents a diagram of the
+possible states of a job.%
 \begin{figure}[tbh]
@@
 \end{figure}
-A user can examine the state of a job by querying the NorduGrid Information
-System (IS) using the UI or any other suitable tool. Please remember
-that the IS can manipulate  state names to make them more user
-friendly and to combine them with states introduced by other parts
-of the whole setup. Another way is to access virtual informational
-files through GridFTP interface or to use query method of WS interface.
+A user can examine the state of a job by querying the dedicated Local
+Information Description Interface of A-Rex  
+ using the UI or any other suitable tool
+
+     or through query method of WS interface.
 
@@
 
-Below is the description of all actions taken by the GM at every state:
+Below is the description of all actions taken by the A-Rex at every
+state:
 
@@
 \item \textbf{Accepted} - In this state the job has been submitted to
-a CE but is not processed yet. The GM will analyze the JD and move
-to the next stage. If the JD can not be processed the job will be
-canceled and moved to the state \textbf{Finishing}.
+a CE but is not processed yet. The A-Rex will analyze the JD and
+move to the next stage. If the JD can not be processed the job will
+be canceled and moved to the state \textbf{Finishing}.
 \item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
-The GM is downloading the files specified in the JD and is waiting
+The A-Rex is downloading the files specified in the JD and is waiting
 for files which are supposed to be uploaded by the UI. If
@@
 passed for execution to the \emph{Local Resource Management System}
-(LRMS). At the moment PBS is supported best but correspoding backends
-for many other LRMSs are provided with the default installation. 
-      If the local job submission
-is successful the job moves to the \textbf{InLRMS} state. Otherwise
-it moves to \textbf{Finishing}. It is possible to limit the agregate
-number of jobs in \textbf{Submitting} and  \textbf{InLRMS}
-states.
+(LRMS).         The correspoding backends
+for many LRMSs are provided with the default installation.  
+     If the local job submission is successful
+the job moves to the \textbf{InLRMS} state. Otherwise it moves
+to \textbf{Finishing}. It is possible to limit the agregate number
+of jobs in \textbf{Submitting} and  \textbf{InLRMS} states.
 \item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
-The GM takes no actions except waiting until the job finishes.
+The A-Rex takes no actions except waiting until the job finishes.
 \item \textbf{Cancelling} - Necessary action to cancel the job in the LRMS
@@
 in this state.
-\item \textbf{Finished} - No more processing is performed by the GM. The
-user can continue to download data files from the SD. The SD is kept
-available for some time (default is 1 week). After that the job
-is moved to the state \textbf{Deleted}. The 'deletion' time can be
-obtained by querying the NorduGrid Information System  attribute
-\texttt{nordugrid-pbs-job-sessiondirerasetime} of a \texttt{nordugrid-pbs-job}
-entry. If a job was moved to \textbf{Finished} because of failure,
-it may be restarted on request of a client. When restarted, a job
-is moved to the state previous to the one in which it failed and
-is assigned mark PENDING. This is needed in order to not break the
-configuration limits. Exception is a job failed in \textbf{InLRMS}
+\item \textbf{Finished} - No more processing is performed by the A-Rex.
+The user can continue to download data files from the SD. The SD is
+kept available for some time (default is 1 week). After that the
+job is moved to the state \textbf{Deleted}. The 'deletion' time can
+be obtained by querying the Informatrion Interface of the A-Rex
+      . If a job was moved to \textbf{Finished}
+because of failure, it may be restarted on request of a client. When
+restarted, a job is moved to the state previous to the one in which
+it failed and is assigned mark PENDING. This is needed in order to
+not break the configuration limits. Exception is a job failed in \textbf{InLRMS}
 state and lacking input files specified in JD. Such a job is treated
@@
 
-The GM and it's components support the following data transfer protocols
-and corresponding URLs: \emph{ftp, gsiftp, http, httpg, https, se,
-rc} and \emph{rls.} For more information please see {}``Protocols,
-Uniform Resource Locators (URL) and extensions supported in ARC''
-document \cite{urls}.
+In full installation the A-Rex and it's components support the
+following data transfer protocols and corresponding URLs: \emph{ftp,
+gsiftp, http, httpg, https, se, rc} and \emph{rls.} For more information
+please see {}``The Hosting Environment of the Advanced Resource Connector
+middleware      
+  '' document \cite{hed}.
 
@@
 
-\subsection{Internal Files of the Grid Manager}
-
-For each local UNIX user listed in the GM configuration a \textit{control
-directory} exists. In this directory the GM stores information about
-jobs belonging to that user. Multiple users can share the same \textit{control
-directory}. To make it easier to recover in  case of failure, the
-GM stores most information in files rather than in memory. All files
-belonging to same the job have names starting with \textbf{job.ID.,}
-where ID is the job identifier.
+\subsection{Internal Files of the A-Rex }
+
+For each local UNIX user listed in the A-Rex configuration - including
+a generic one which covers all local user identities - a \textit{control
+directory} exists. In this directory the A-Rex stores information
+about jobs belonging to that user. Multiple users can share the same
+\textit{control directory}. To make it easier to recover in  case
+of failure, the A-Rex stores most information in files rather than
+in memory. All files belonging to the same job have names starting
+with \textbf{job.ID.,} where ID is the job identifier.
 
@@
 \begin{itemize}
-\item \textit{job.ID.description} - contains the RSL description of the
-job.
-\item \textit{job.ID.local} - information about the job used by the GM.
+\item \textit{job.ID.description} - contains the  description of the
+job (JD).
+\item \textit{job.ID.local} - information about the job used by the A-Rex.
 It consists of lines of format \textit{{}``name = value''} . Not
@@
 default logger service configured in the configuration  
-
+        (This feature does
+not work in current implementation).
 \end{itemize}
@@
 
-\item \textit{job.ID.proxy} -delegated GSI proxy.
-\item \textit{job.ID.proxy.tmp} - temporary GSI proxy with different unx
-ownership used by processes run with effective \emph{user id} different
-form job owner's \emph{id}.
+\item \textit{job.ID.proxy} -delegated X509 credentials .
+\item \textit{job.ID.proxy.tmp} - temporary X509 credentials  with
+different UNIX ownership used by processes run with effective \emph{user
+id} different form job owner's \emph{id}.
 \end{itemize}
 There are other files with names like job.ID.{*} which are created
-and used by different parts of the GM. Their presence in the \textit{control
-directory} can not be guaranteed and can change depending on changes
-in the GM code.
-
-\subsection{GridFTP Interface}
-
-The GridFTP Interface to job management is implemented through GridFTP
-server. The GFS is made of layer implementing GridFTP protocol and
-plugins for accessing actual data. Local file access in the GFS is
-implemented through plugins (shared libraries). There are 3 plugins
-provided with the GFS: \textit{fileplugin.so, gaclplugin.so} and \textit{jobplugin.so}
-.
-
-The \textit{fileplugin.so} is intended to be uses for plain file access
-with the configuration senitive to user subject and is not needed
-for job management interface.
-
-The \textit{gaclplugin.so} uses GACL (\url{http://www.gridpp.ac.uk/authz/gacl/})
-to control access to local file system instead of Unix user identity.
-It's also not needed for job management.
-
-The \textit{jobplugin.so} is using information about jobs being controlled
-by GM and provides access to session directories of the jobs owned
-by user. It also provides an interface (virtual directory and virtual
-operations) to submit, cancel, clean, renew credentials and obtain
-information about the job.
+and used by different parts of the A-Rex. Their presence in the
+\textit{control directory} can not be guaranteed and can change depending
+on changes in the A-Rex code.
+
+\subsection{ }
+
+    \textit{ }  \textit{}
+
+ \textit{}
+
+ \textit{}
+
+ \textit{}
 
@@
 during the execution of a job. The WS Interface is built and deployed
-inside Hosting Envirnoment Daemon {[}HED] infrastructure\cite{hed}.
+inside Hosting Envirnoment Daemon (HED) infrastructure\cite{hed}.
 
@@
 by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
-The exchange of SOAP messages is performed over the HTTPS protocol.
+The exchange of SOAP messages is performed over the HTTP(S) protocol.
 The BES interface is represented by two port-types - BES-Management
@@
 and thus defines operations to start and stop the functionality of
-the BES service. A-REX does not implement remote control of service
-functionality. Hence the BES-Management port-type is not implemented.
+the BES service. The A-Rex does not implement remote control of
+service functionality. Hence the BES-Management port-type is not functional.
 The BES-Factory port-type provides operations to submit new jobs (to
 create an activity in terms of BES) and to monitor its state. It also
-has an ability to provide information about the service. A-REX fully
+has an ability to provide information about the service. A-Rex fully
 implements the functionality of this port-type.
 
-For job descriptions A-REX accepts the Job Submission Description
-Language (JSDL)\cite{jsdl} as defined by the OGF Job Submission Description
-Language Working Group. Supported elements and extensions are described
-below.
+For job descriptions A-Rex accepts the Job Submission Description
+Language (JSDL)\cite{jsdl} documents as defined by the OGF Job Submission
+Description Language Working Group. Supported elements and extensions
+are described below.
 
@@
 
-A-REX introduces a new operation in addition to those provided by
-BES. It does that by defining its own port-type A-REX with the single
+A-R introduces a new operation in addition to those provided by
+BES. It does that by defining its own port-type  with the single
 operation \emph{ChangeActivityStatus} (see Appendix D). This operation
 provides a way to request simple transfers between states of jobs
-being processed and corresponding actions.
+  and corresponding actions.
 
@@
 Since BES allows implementators to extend their initial activity states
-with additional sub-states, A-REX defines a set of sub-states of activity
-processing in addition to those defined by the BES, as listed in Table
-\ref{tab:Job-states-definitions}.
+with additional sub-states, A-Rex defines a set of sub-states of
+activity processing in addition to those defined by the BES, as listed
+in Table \ref{tab:Job-states-definitions}.
 
@@
 
-ARC service interfaces optionally offer a sub-interface, called the
-Delegation Interface (see Appendix E). This is a common purpose interface
-to be used by ARC services which accept delegated credentials from
-clients. The Delegation Interface implements two operations: initiatialization
-of credentials delegation (DelegateCredentialsInit) and update/renewal
-of credentials (UpdateCredentials).
+       The A-Rex
+also supportd  the Delegation Interface (see Appendix E). This
+is a common purpose interface to be used by ARC services which accepts
+delegated credentials from clients. The Delegation Interface implements
+two operations: initiatialization of credentials delegation (DelegateCredentialsInit)
+and update/renewal of credentials (UpdateCredentials).
 
@@
 The A-REX implements the Local Information Description Interface (LIDI)
-interface being common for all ARC services. This interface is based
+interface  common for all ARC services. This interface is based
 on OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
@@
 1- remove finished jobs, 2 - remove active jobs too, 3- also remove
-everything that looks like junk.\\
-
-\subsection{\label{subsec: configfile of the gfs}Configuration of the GridFTP
-Server}
-
-Default location of the GFS configuration file is \textit{/etc/arc.conf}
-or \textit{\$NORDUGRID\_LOCATION/etc/gridftpd.conf}. Format of these
-configuration files is similar to that of the GM. It also supports
-generic commands described at the beginning of previous section \ref{SubSection: ConfigFile}.
-In the new format sections {[}common] and {[}gridftpd] are used. Commands
-specific to the GFS are described below.
-
-\begin{itemize}
-\item \textbf{\textit{port}}\textit{=number} - specifies TCP/IP port number.
-Default is 2811.
-\item \textbf{\textit{include}}\textit{=path} - include contents of another
-file. Generic commands can't be specified there.
-\item \textbf{\textit{encryption}}\textit{=yes|no} - specifies if server
-will allow data transfer to be encrypted. Default is yes.
-\item \textbf{\textit{pluginpath}}\textit{=path} - specifies the path where
-plugin libraries are installed
-\item \textbf{\textit{allowunknown}}\textit{=yes|no} - if set to \emph{yes}
-clients are not checked agains grid-mapfile. Hence only access rules
-specified in this configuration file will be applied.
-\item \textbf{\textit{firewall}}\textit{=hostname} - use IP address of the
-\textit{hostname} in response to PASV command instead of IP address
-of a network interface of computer. You can write IP address directly
-instead of \textit{hostname}. This command may be if server is situated
-behind NAT.
-\item \textbf{\textit{unixgroup}}\textit{=group rule} - define local UNIX
-user and optionally UNIX group to which user belonging to specified
-authorization \textit{group} is mapped (see Section \ref{sub:Authorization}
-for definition of group). Local names are obtained from specified
-\textit{rule}. If specified rule could not produce any mapping, next
-command is used. Mapping stops at first matched rule. Following rules
-are supported:
-
-\begin{itemize}
-\item \textbf{\textit{mapfile}} \textit{file} - user's subject is matched
-against list of subjects stored inspecified file, one per line followed
-by local UNIX name.
-\item \textbf{\textit{simplepool}} \textit{directory} - user is assigned
-one of local UNIX names stored in a file \textit{directory/pool},
-one per line. Used names are stored in other files placed in the same
-\textit{directory}. If UNIX name was not used for 10 days, it may
-be reassigned to another user.
-\item \textbf{\textit{lcmaps}} \textit{library directory database} - call
-LCMAPS functions to do mapping. Here \textit{library} is path to shared
-library of LCMAPS, either absolute or relative to \textit{directory};
-\textit{directory} is path to LCMAPS installation directory, equivalent
-of LCMAPS\_DIR variable; \textit{database} is path to LCMAPS database,
-equivalent to LCMAPS\_DB\_FILE variable. Each arguments except \textit{library}
-is optional and may be either skiped or replaced with '{*}'.
-\item \textbf{\textit{mapplugin}} \textit{timeout plugin} {[}\textit{arg1}
-{[}\textit{arg2} {[}...]]] - run external \emph{plugin} executable
-with specified arguments. Execution of \emph{plugin} may not last
-longer than \emph{timeout} seconds. Rule matches if exit code is 0
-and there is UNIX name printed on \emph{stdout}. Name may be optionaly
-followed by UNIX group separated by ':'. In arguments following substitions
-are applied before plugin is started:
-
-\begin{itemize}
-\item \%D - subject of users's cerificate,
-\item \%P - name of credentials' proxy file.
-\end{itemize}
-\end{itemize}
-\item \textbf{\textit{unixvo}}\textit{=vo rule} - same as \textbf{\textit{unixgroup}}
-for users belonging to Virtual Organization (VO) \textit{vo}.
-\item \textbf{\textit{unixmap}}\textit{={[}unixname]{[}:unixgroup] rule}
-- define local UNIX user and optionally group used to represent connected
-client. \textit{rule} is one of those allowed for \textbf{\textit{authorization
-groups}} (see Section \ref{sub:Authorization}) and for \textbf{\textit{unixgroup}}/\textbf{\textit{unixvo}}.
-In case of mapping rule username is one, provided by rule. Otherwise
-specified \textit{unixname:unixgroup} is taken. Both \textit{unixname}
-and \textit{unixgroup} may be either omited or set to '{*}' to specify
-missing value.
-\item \textbf{\textit{groupcfg}}\textit{=name} - is put into subsections
-representnig plugin or {[}group] section and defines if that section
-is effective. In old format it selects the group to which all following
-lines apply. Only unaffected option is \textbf{\textit{groupcfg}}.
-If name is empty (or no groupcfg is used at all) following lines apply
-to all users.
-\end{itemize}
-Subsections of \emph{gridftpd} section specifies plugins which serve
-virtual FTP path (similar to mount command of UNIX). Name of subsection
-is irrelevant. In old format this section starts with command \textbf{\textit{plugin}}
-\textit{path library\_name} and ends with keyword \textbf{\textit{end}}.
-Inside subsection following commands are supported
-
-\begin{itemize}
-\item \textbf{\textit{plugin}}\textit{=library\_name} - use plugin \textit{library\_name}
-to serve virtual path.
-\item \textbf{\textit{path}}\textit{=path} - virtual path to serve.
-
-GFS comes with 3 plugins: \textit{filepligin.so, gaclplugin.so} and
-\textit{jobplugin.so}.
-
-\begin{itemize}
-\item \textit{jobplugin.so} does not require any specific options in case
-of old configuration format. It reads the configuration file of the
-GM located at the standard place as specified in the section \ref{SubSection: ConfigFile}.
-Following options are supported:
-
-\begin{itemize}
-\item \textbf{\textit{configfile}}\textit{=path} - defines non-standard
-place for GM's configuration file,
-\item \textbf{\textit{allownew}}\textit{=yes|no} - specifies if new jobs
-can be submitted. Default is \emph{yes}.
-\item \textbf{\textit{unixgroup}}/\textbf{\textit{unixvo}}/\textbf{\textit{unixmap}}
-- same options like in top-level GFS configuration. If mapping succeeds
-obtained local user will be used to run submitted job.
-\end{itemize}
-\item \textit{filepligin.so} supports following options:
-
-\begin{itemize}
-\item \textbf{\textit{mount}}\textit{=path} - defines the place on local
-filesystem to which file access operations apply
-\item \textbf{\textit{dir}}\textit{=path options} - specifies access rules
-for accessing files in \textit{path} (relative to virtual and real
-path) and all the files below. \\
-\textit{options} is the list of the following keywords:
-
-\begin{itemize}
-\item \textbf{\textit{nouser}} - do not use local file system rights, only
-use those specifies in this line
-\item \textbf{\textit{owner}} - check only file owner access rights
-\item \textbf{\textit{group}} - check only group access rights
-\item \textbf{\textit{other}} - check only \char`\"{}others\char`\"{} access
-rights
-\end{itemize}
-The options above are exclusive. If none of the above specified usual
-Unix access rights are applied.
-
-\begin{itemize}
-\item \textbf{\textit{read}} - allow reading files
-\item \textbf{\textit{delete}} - allow deleting files
-\item \textbf{\textit{append}} - allow appending files (does not allow creation)
-\item \textbf{\textit{overwrite}} - allow overwriting already existing files
-(does not allow creation, file attributes are not changed)
-\item \textbf{\textit{dirlist}} - allow obtaining list of the files
-\item \textbf{\textit{cd}} - allow to make this directory current
-\item \textbf{\textit{create}} \textit{owner:group permissions\_or:permissions\_and}
-- allow creating new files. File will be owned by \textit{owner} and
-owning group will be \textit{group}. If '{*}' is used, the user/group
-to which connected user is mapped will be used. The permissions will
-be set to \textit{permissions\_or} \& \textit{permissions\_and} (second
-number is reserved for the future usage). 
-\item \textbf{\textit{mkdir}} \textit{owner:group permissions\_or:permissions\_and}
-- allow creating new directories.
-\end{itemize}
-\end{itemize}
-\item \textit{gaclpligin.so} does not have options in case of the old configuration.
-First line of it's configuration contains local path (root directory)
-served by it. Rest till keyword \textbf{\textit{end}} contains GACL
-\cite{gacl} XML used to setup initial access rules for every newly
-created file and directory. If GACL XML is empty then there will be
-no default ACLs created for new files and directories. That means
-ACL of parent directory will be used. \\
-For the new configuration format following options are supported:
-\textbf{\textit{gacl=}}\textit{gacl} - GACL XML, \textbf{\textit{mount}}\textit{=path}
-- local path server by plugin. \\
-XML may contain variables which are replaced with values taken from
-client's credentials. Following variables are supported:
+everything that looks like junk.
+
+\subsection{    }
+
+        \textit{}
+ \textit{}
+
+\begin{itemize}
+\item \textbf{\textit{}}\textit{}
+
+\item \textbf{\textit{}}\textit{}
+
+\item \textbf{\textit{}}\textit{}
+
+\item \textbf{\textit{}}\textit{}
+
+\item \textbf{\textit{}}\textit{}     \emph{}
+
+\item \textbf{\textit{}}\textit{}      
+\textit{}
+
+  \textit{}
+
+\item \textbf{\textit{}}\textit{ }
+
+ \textit{}       
+        \textit{}
+
+\begin{itemize}
+\item \textbf{\textit{}} \textit{}
+
+\item \textbf{\textit{}} \textit{}    
+         \textit{}
+
+\textit{}
+
+\item \textbf{\textit{}} \textit{  }  
+      \textit{}    
+        \textit{}
+\textit{}       
+   \textit{}     
+       \textit{}
+
+\item \textbf{\textit{}} \textit{ } \textit{}
+\textit{}     \emph{}  
+    \emph{}    
+ \emph{}          
+     \emph{}
+
+\begin{itemize}
+\item      
+\item       
+\end{itemize}
+\end{itemize}
+\item \textbf{\textit{}}\textit{ }    \textbf{\textit{}}
+       \textit{}
+\item \textbf{\textit{}}\textit{ }
+
+\textit{}       \textbf{\textit{
+}}      \textbf{\textit{}}\textbf{\textit{}}
+
+ \textit{}    \textit{}
+ \textit{}
+
+\item \textbf{\textit{}}\textit{}
+
+      \textbf{\textit{}}
+
+\end{itemize}
+  \emph{}
+
+          \textbf{\textit{}}
+\textit{ }     \textbf{\textit{}}
+
+\begin{itemize}
+\item \textbf{\textit{}}\textit{}    \textit{}
+
+\item \textbf{\textit{}}\textit{}
+
+     \textit{ } 
+\textit{}
+
+\begin{itemize}
+\item \textit{}
+
+\begin{itemize}
+\item \textbf{\textit{}}\textit{}
+
+\item \textbf{\textit{}}\textit{}     
+     \emph{}
+\item \textbf{\textit{}}\textbf{\textit{}}\textbf{\textit{}}
+
+\end{itemize}
+\item \textit{}
+
+\begin{itemize}
+\item \textbf{\textit{}}\textit{}
+
+\item \textbf{\textit{}}\textit{ }    
+    \textit{}     
+      \textit{}
+
+\begin{itemize}
+\item \textbf{\textit{}}
+
+\item \textbf{\textit{}}       
+\item \textbf{\textit{}}      
+\item \textbf{\textit{}}      
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{\textit{}}    
+\item \textbf{\textit{}}    
+\item \textbf{\textit{}}        
+\item \textbf{\textit{}}
+
+\item \textbf{\textit{}}       
+\item \textbf{\textit{}}       
+\item \textbf{\textit{}} \textit{ }
+          \textit{} 
+    \textit{}
+
+   \textit{}  \textit{}
+
+\item \textbf{\textit{}} \textit{ }
+
+\end{itemize}
+\end{itemize}
+\item \textit{}
+
+      \textbf{\textit{}}
+
+   \textbf{\textit{}}\textit{}  
+ \textbf{\textit{}}\textit{}
 
 \begin{lyxlist}{00.00.0000}
-\item [{\emph{\$subject}}] - subject of user's certificate (DN),
-\item [{\emph{\$voms}}] - subject of VOMS\cite{voms} server (DN),
-\item [{\emph{\$vo}}] - name of VO (from VOMS certificate),
-\item [{\emph{\$role}}] - role (from VOMS certificate),
-\item [{\emph{\$capability}}] - capabilities (from VOMS certificate),
-\item [{\emph{\$group}}] - name of group (from VOMS certificate) .
+\item [{\emph{}}]      
+\item [{\emph{}}]      
+\item [{\emph{}}]       
+\item [{\emph{}}]     
+\item [{\emph{}}]     
+\item [{\emph{}}]        
 \end{lyxlist}
-Additionally root directory must contain \emph{.gacl} file with initial
-ACL. Otherwise rule will be {}``deny all for everyone''.
-
-\end{itemize}
-\end{itemize}
-Some configuration parameters can be specified from command line while
-starting the GFS:
-
-\emph{gridftpd {[}-h] {[}-p number] {[}-n number] {[}-b number] {[}-B
-number] {[}-d level] {[}-c path] {[}-F] {[}-U uid{[}:gid]] {[}-L path]
-{[}-P path]}\\
-\hspace*{0.5cm}\emph{-h} - short help,\\
-\hspace*{0.5cm}\emph{-d} - debug level,\\
-\hspace*{0.5cm}\emph{-L} - name log file (overwrites value in configuration
-file),\\
-\hspace*{0.5cm}\emph{-P} - name for file containing process id (overwrites
-value in configuration file),\\
-\hspace*{0.5cm}\emph{-U} - user and gropu id to use for running daemon,\\
-\hspace*{0.5cm}\emph{-F} - do not make process daemon,\\
-\hspace*{0.5cm}\emph{-c} - name od configuration file,\\
-\hspace*{0.5cm}\emph{-p} - TCP/IP port number,\\
-\hspace*{0.5cm}\emph{-n} - maximal number of simultaneously served
-connection,\\
-\hspace*{0.5cm}\emph{-b} - default size of buffer used for data transfer
-(default is 64kB),\\
-\hspace*{0.5cm}\emph{-B} - maximal size of buffer used for data transfer
-(default is 640kB).
+     \emph{}
+
+\end{itemize}
+\end{itemize}
+
+\emph{         
+          }\emph{}
+  \emph{}   \emph{}    
+    \emph{}    
+       \emph{}
+         \emph{}   
+  \emph{}     \emph{}
+   \emph{}     
+ \emph{}        
+   \emph{}
 

### 2008-08-05 doc/hed/ARCHED_article.tex
Added conditionals to be able to have different versions of some sections.

@@
 \documentclass[a4paper,10pt]{article}
+
+\usepackage{ifthen}
+\newboolean{brieftechdesc}
+\newboolean{verbtechdesc}
+\setboolean{brieftechdesc}{true}
+\setboolean{verbtechdesc}{false}
 
@@
 These work on units called messages which represent data going in to or out of the HED.
-The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general attributes and security attributes.
+The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general attributes and security attributes where information relevant to each protocol is accumulated as the message progresses.
 Each MCC typically implements one level in the Internet Protocol suite by transforming a message to an input suitable to propagate to the next component and then performs the corresponding transformation of the response on the way back.
@@
 
-As the data is passed through the individual MCCs, they each populate structures with both general attributes and special security attributes that are available at that particular protocol level.
+%As the data is passed through the individual MCCs, they each populate structures with both general attributes and special security attributes that are available at that particular protocol level.
 
@@
 The chain formulation using MCCs is up to the administrator.
-The chain is an ordered list the MCCs and their interconection can be described in the configuration file.
-Based on the configuration information the Loader components formulates the actual chain in the memory of the daemon.
-The MCC may implement some routing algorithm which means one MCC may connection to multiple other MCCs.
-Tipical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to SOAP MCC but the messages with GET operation are goes to for example a simple and limited HTTPD service component.
+The chain is an ordered list of MCCs and their interconection can be described in the configuration file.
+Based on the configuration information the Loader components formulate the actual chain in the memory of the daemon.
+The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs.
+Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component.
 
-Most of the MCCs has client and server version because the behavior of MCCs should be different depending on wheater it is sitting on server or client part.
-The tipical scenarion here the TCP MCC which should listen and wait for incomming messages on socket on server part but call connect method on client part.
+Most of the MCCs has a client and a server version because the behavior of an MCC should be different depending on whether it is sitting on the server or client side.
+The typical scenario here is illustrated by the TCP MCC which should listen and wait for incomming messages on a socket on the server side but call \texttt{connect} on the client side.
 
-In general every MCC have optimal and natural places in certain chain and this place is not always can be modified.
-For example in the server part the TCP MCC must be the first MCC in any chain where it is used and the TLS MCC should be right after of the TCP MCC.
+In general every MCC have optimal and natural places in certain chains and this place cannot always be modified.
+For example on the server side the TCP MCC must be the first MCC in any chain where it is used and the TLS MCC should be right after the TCP MCC.
 
-The developer whoes write an MCC is free to choose any 3rd party library and comonent to implement functionalty of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by libarcloader and libarcmessage libraries of ARC1.
+The developer who writes an MCC is free to choose any 3rd party library and component to implement the functionalty of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by libarcloader and libarcmessage libraries of ARC1.
 
@@
 
+\ifthenelse{\boolean{brieftechdesc}}{
 This MCC can be configured as to what port to listen to.
+}{}
+
+\ifthenelse{\boolean{verbtechdesc}}{
+This is a verbose test text. It might not seem verbose but it is.
+}{}
 
@@
 They can even be regarded as special cases of MCCs in that they constitute the last link in the so called Message Chain and are attached in much the same manner as other MCCs.
-After being loaded, instanciated and configured by the HED, services are provided with a range of
+After being loaded, instanciated and configured by the HED, services are provided with a range of...
 

### 2008-08-05 doc/backend_interface/Backends.tex
backends doc update

@@
 \vspace*{3mm}
-The following properties referring to job counts are obtained from GridManager's status files:
+The following properties referring to job counts are calculated based on information in GridManager's status files:
 
@@
 \hline \hline
-TotalJobs & Number of grid jobs which are not yet finished \\
+TotalJobs & number of grid jobs in states other than FINISHED \\
 \hline
-RunningJobs & Number of grid jobs running \\
+RunningJobs & number of grid jobs in INLRMS:R state \\
 \hline
-WaitingJobs & Number of grid jobs queued in the LRMS and waiting to be submitted to the LRMS \\
+WaitingJobs & number of grid jobs in INLRMS:Q, ACCEPTED, PREPARING and SUBMIT* states \\
 \hline
@@
 
-\begin{tabular}{|l|p{12cm}|}
+\begin{tabular}{|l|p{11cm}|}
 \hline
@@
 \hline
-IssuerCA & from the \emph{hostcert.pem} file \\
+IssuerCA & obtained from the \emph{hostcert.pem} file \\
 \hline
-TrustedCA & from scanning the \emph{certificates} directory \\
+TrustedCA & obtained by scanning the \emph{certificates} directory \\
 \hline
@@
 \hline
-Property & Value\\
+Property & Comments \\
 \hline \hline
@@
 \hline
+SchedulingPolicy & if maui is used, set to \texttt{fairshare} \\
+\hline
+ServingState & has the same value as Computingendpoint.ServingState \\
+\hline
 \end{tabular}
@@
 \vspace*{3mm}
-The following properties are taken from the corresponding values returned by
+The following properties are taken from the corresponding attributes returned by
 the LRMS plugin. Units are converted as necessary.
@@
 \hline
-Property & LRMS plugin property \\
+Property & LRMS plugin attribute \\
 \hline \hline
@@
 
-\begin{tabular}{|l|p{12cm}|}
+The following properties are directly taken from configuration options.
+
+\begin{tabular}{|l|l|p{7cm}|}
 \hline
-Property & Value\\
+Property & Configuration option  & Comments \\
 \hline \hline
-MaxTotalJobs & maxjobs from configuration\\
-MaxRunningJobs & \\
-MaxWaitingJobs & \\
+MaxTotalJobs & maxjobs & maxjobs is a per cluster limit, grid-manager has no equivalent limit per share \\
+\hline
+MaxPreLRMSWaitingJobs & maxjobs & \\
+\hline
+MaxStageInStreams & maxload & maxload is a per cluster limit, grid-manager has no equivalent limit per share \\
+\hline
+MaxStageOutStreams & maxload & \\
+\hline
+MaxMemory & nodememory & \\
+\hline
+MaxDiskSpace & maxdiskperjob & \\
 \hline
@@
 
-\phantomsection
-%\addcontentsline{toc}{subsubsection}{Name}
-\hspace*{0.5cm}
-\begin{shaded}
-\textbf{Name}
-\end{shaded}
-\begin{tabular}{lp{10cm}}  
-Attribute value:& String\\
-Example:& \verb#Name: TODO#\\
-Related xRSL:& TODO\\
-UI role:& ?\\
-Description:& Name of the application environment\\
+\vspace*{3mm}
+The following properties are taken from the corresponding LRMS plugin attributes when available:
+
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & LRMS plugin attribute \\
+\hline \hline
+MaxRunningJobs & maxrunning \\
+\hline
+MaxWaitingJobs & maxqueuable - maxrunning \\
+\hline
+MaxUserRunningJobs & maxuserrun \\
+\hline
+MaxSlotsPerJob & maxslotsperjob \\
+\hline
 \end{tabular}
 
-\phantomsection
-%\addcontentsline{toc}{subsubsection}{Version}
-\hspace*{0.5cm}
-\begin{shaded}
-\textbf{Version}
-\end{shaded}
-\begin{tabular}{lp{10cm}}  
-Attribute value:& String\\
-Example:& \verb#Version: TODO#\\
-Related xRSL:& TODO\\
-UI role:& TODO\\
-Description:& Name of the application environment\\
+\vspace*{3mm}
+The following properties referring to job counts are calculated based on information in GridManager's status files:
+
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & Meaning \\
+\hline \hline
+TotalJobs &  number of grid jobs in this share in states other than FINISHED \\
+\hline
+RunningJobs & number of grid jobs in this share in INLRMS:R state \\
+\hline
+SuspendedJobs & number of grid jobs in this share in INLRMS:S state \\
+\hline
+WaitingJobs & number of grid jobs in this share in INLRMS:Q and INLRMS:O states \\
+\hline
+LocalRunningJobs & The total number of jobs running in the associated LRMS queue  minus the number of grid-running grid jobs (RunningJobs) \\
+\hline
+LocalWaitingJobs & The total number of jobs queued in the associated LRMS queue minus the number of grid-queued grid jobs (WaitingJobs) \\
+\hline
+StagingJobs & number of grid jobs in this share in PREPARING and FINISHING states \\
+\hline
+PreLRMSWaitingJobs & number of grid jobs in this share in ACCEPTED, PREPARING and SUBMIT* states\\
+\hline
 \end{tabular}
+
+\vspace*{3mm}
+The following properties are derived from the corresponding LRMS plugin attributes:
+
+\begin{tabular}{|l|p{12cm}|}
+\hline
+Property & Related LRMS properly \\
+\hline \hline
+UsedSlots & running \\
+\hline
+RequestedSlots & queued \\
+\hline
+FreeSlots & totalcpus - running \\
+\hline
+FreeSlotsWithDuration & freecpus \\
+\hline
+\end{tabular}
+
+\subsection{ExecutionEnvironment}
+
+\subsection{ApplicationEnvironment}
+
+\subsection{ComputingActivity}
+
+\subsection{MappingPolicy and AccessPolicy}
+
+%\phantomsection
+%%\addcontentsline{toc}{subsubsection}{Name}
+%\hspace*{0.5cm}
+%\begin{shaded}
+%\textbf{Name}
+%\end{shaded}
+%\begin{tabular}{lp{10cm}}  
+%Attribute value:& String\\
+%Example:& \verb#Name: TODO#\\
+%Related xRSL:& TODO\\
+%UI role:& ?\\
+%Description:& Name of the application environment\\
+%\end{tabular}
 

### 2008-08-05 doc/hed/ARCHED_article.tex
End of day check in.

@@
 \newboolean{verbtechdesc}
-\setboolean{brieftechdesc}{true}
-\setboolean{verbtechdesc}{false}
+\setboolean{brieftechdesc}{false}
+\setboolean{verbtechdesc}{true}
 
@@
 \ifthenelse{\boolean{verbtechdesc}}{
-This is a verbose test text. It might not seem verbose but it is.
+This MCC can be configuered with one or more \texttt{<tcp:Listen>} elements which in turn contain the elements \texttt{<tcp:Port>}, \texttt{<tcp:Interface>} and \texttt{<tcp:Version>}.
+The \texttt{<tcp:Port>} element is mandatory and should contain an integer corresponding to the TCP port to listen to.
+The \texttt{<tcp:Interface>} element is optional and is meant to identify the network interface to bind to.
+It is currently not used.
+The \texttt{<tcp:Version>} element is used to specify IP version.
+It is optional and should if present contain the single digit 4 or 6.
 }{}
@@
 \paragraph{TLS MCC}
+The server TLS MCC enables ...
+
+\ifthenelse{\boolean{brieftechdesc}}{
+This MCC needs to be configured with a host certificate as well as trusted CA certificates.
+}{}
+
+\ifthenelse{\boolean{verbtechdesc}}{
+The
+}{}
 

### 2008-08-05 doc/backend_interface/Backends-arc1.tex
Renamed files from Backends.* to Backends-arc1.* to be able to distinguish the back-ends' documentation of ARC0 from ARC1

--- 2008-08-05 doc/backend_interface/Backends.tex
@@
   pdfsubject = {Description and developer's guide for ARC1},
-  pdfkeywords = {ARC,ARC1,NorduGrid,manual,Backend,LRMS,Infosystem,GLUE2},
+  pdfkeywords = {ARC,ARC1,NorduGrid,manual,Back-end,LRMS,Infosystem,GLUE2},
   pdfcreator = {PDFLaTeX with hyperref package},
@@
 
-{\raggedleft NORDUGRID-TECH-XX\par}
+{\raggedleft NORDUGRID-TECH-18\par}
 
@@
 \begin{center}
-\textsc{\Large Arc batch system back-end interface guide with support for GLUE 2}
+\textsc{\Large Arc batch system back-end interface guide with support for Glue 2}
 \par\end{center}{\Large \par}
@@
 
-{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomasf@fys.uio.no} \par}
+{\centering \large Adrian Taga\footnote{v.a.taga@fys.uio.no}, Thomas Frågåt\footnote{thomas.fragat@fys.uio.no} \par}
 
@@
 
-This document describes the Advanced Resource Connector~\cite{arc1}
+This document describes the next generation Advanced Resource Connector~\cite{arc1} (ARC)
 batch system back-end infrastructure. It aims to describe the infrastructure
@@
 
-The batch system back-ends are what tie the ARC grid middleware (through
-(ARC Resource-coupled EXecution service, A-REX~\cite{arex}) to the underlying
-cluster management system or LRMS. The back-ends consist of set of shell and
+The batch system back-ends are what tie the ARC grid middleware (through the
+ARC Resource-coupled EXecution service, A-REX~\cite{arex}) to the underlying
+cluster management system or LRMS. The back-ends consist of set of a shell and
 Perl scripts whose role are twofold:
@@
 \begin{enumerate}
-\item to allow the GM, which is resided in A-REX, to manipulate jobs on the LRMS including job submit,
-cancel operations etc.
-\item to collect information about jobs, users, batch system and the cluster
+\item to allow the GridManager (GM), which is resided in A-REX, to manipulate
+jobs in the LRMS including job submit, cancel operations etc.
+\item to collect information about jobs, users, the batch system and the cluster
 itself for the Information System. 
@@
 within the NorduGrid subversion tree~\cite{svn}. 
-The job control interface scripts can be found under \texttt{arc1/trunk/src/services/a-rex/lrms/}
-while the information collectors are located in \texttt{arc1/trunk/src/services/a-rex/infoproviders/}.
-
+The job control interface scripts can be found under 
+\begin{quote}
+\texttt{arc1/trunk/src/services/a-rex/lrms/}
+\end{quote}
+while the information collectors are located in
+\begin{quote}
+\texttt{arc1/trunk/src/services/a-rex/infoproviders/}.
+\end{quote}
 The next generation of ARC supports both the classic NorduGrid information schema~\cite{is} and
@@
 
-Job control part of the LRMS interface is handled by the Grid Manager~\cite{gm}.
+The job control part of the LRMS interface is handled by the Grid Manager~\cite{gm}.
 It takes care of preparing a native batch system submission script,
@@
 of the GM requires three programs. These programs can be implemented
-any way the designer sees fit, but all the existing back-end interfaces
+any way the designer sees it fits, but all the existing back-end interfaces
 use shell scripting for portability and ease of tailoring to a specific
 site. The GM will call the following programs: cancel-LRMS-job, submit-LRMS-job,
-and scan-LRMS-job. LRMS is replaced with the short hand name for the
-LRMS e.g. cancel-pbs-job. The scripts are described one by one in
+and scan-LRMS-job where LRMS is replaced with the short hand name for the
+LRMS; e.g. cancel-pbs-job. The scripts are described one by one in
 the following subsections. Useful information can also be found in
@@
 a new job arrives and needs to be submitted to the LRMS. Like cancel-LRMS-job
-it is given the grami file as argument on execution. Submit-LRMS-job
+it is given the GRAMi file as argument on execution. Submit-LRMS-job
 then has to set up the session directories, run-time environment and
-anything else needed. Then it should submit the job to the local LRMS.
+anything else needed. Then it submits the job to the local LRMS.
 This is normally done by generating a native job script for the LRMS
-and then running the local submit command, but it could also be done
+and then running the local submit command, but it can also be done
 through an API if the LRMS supports it.
@@
 Scan-LRMS-job is run periodically. Its job is to scan the LRMS for
-jobs that have finished. Once it has found a finished job it should
+jobs that have finished. Once it has found a finished job it will
 write the exit-code of that job to the file job.\{gridid\}.lrms\_done
-in the ARC jobstatus directory. Then it should call the gm-kick program
-the GM will then notice that the job has finished and start finalizing
+in the ARC jobstatus directory. Then it will call the gm-kick program
+to notify GM about the finished job. Subsequently, the GM starts finalizing
 the job.
 
-Generally two approaches are taken to find jobs that are done. One
-is asking the LRMS. First all started grid jobs are found in the jobstatus
+Generally, two approaches are taken to find jobs that are done. One
+is to ask the LRMS. First all started grid jobs are found in the jobstatus
 directory%
@@
 variable of arc.conf%
-}. This can be done by checking if the status is INLRMS in the jobs
+}. This can be done by checking if the status is "INLRMS" in the jobs
 status file%
@@
 The second approach is to parse the LRMSs log files. This method has
-some drawbacks e.g. the GM has to be allowed read access to the logs.
+some drawbacks, e.g. the GM has to be allowed read access to the logs.
 The back-end will then have to remember where in the log it was last
-time it ran. This information will have to be stored in a file some
-where on the front-end.
+time it ran. This information will have to be stored in a file somewhere
+on the front-end.
 
@@
 Option values can be surrounded by single quotes or double quotes.
-Values without quotes are also accepted. Multi-valued options are
-not supported currently. Only the last defined value is retained. 
+Values without quotes are also accepted. Currently, multi-valued options are
+not supported. Only the last defined value is retained. 
 \end{quote}
@@
 The main purpose of the information system batch interface is to populate
-the nordugrid information model and the GLUE2 model with locally collected
-information obtained from the batch system. Important to recall that the
+the NorduGrid information model and the GLUE2 model with locally collected
+information obtained from the batch system. It is important to recall that the
 locally collected information is broader than what the batch-system can
-offer, information taken from the grid-layer (mostly grid manager), from
-the front-end machine and from the \texttt{arc.conf} configuration
-file are also needed and used to populate the nordugrid information
+offer, information taken from the grid-layer (mostly A-REX), from
+the front-end machine, and from the \texttt{arc.conf} configuration
+file are also needed and used to populate the NorduGrid information
 model~\cite{is} and the GLUE2 schema~\cite{glue2}.
@@
 The information system interface consists of a set of Perl scripts which
-generates an XML output to \texttt{STDOUT}.
+generates an XML output to \texttt{STDOUT} by one invocation of the Perl script named
+\texttt{cluster+qju.pl}. The XML output includes two representations of the output data,
+namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.
+
+The information collector scripts are divided between several separate scripts:
+\begin{itemize}
+\item \texttt{cluster+qju.pl} - driver for information collection.
+It calls all other information collectors and prints the results in XML.
+The information collection is done by one single invocation of this script.
+\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm}).
+\item \texttt{InfoChecker.pm} - used by \texttt{InfoCollector} to validate options and results against a ``schema''.
+\item \texttt{GMJobsInfo.pm} - collects information about jobs from GM status files.
+\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments, etc.)
+\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits, etc.) by calling the appropriate LRMS plugin.
+It also does validation of input options to the plugin and of the data returned by the plugin.
+\item \texttt{<BATCH\_SYSTEM\_NAME>.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework.
+\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NorduGrid schema.
+\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema.
+\end{itemize}
 
@@
 should be written that implements the interface presented in the next
-subsections. Furthermore hooks should be added to the \texttt{LRMS.pm} module,
+subsections. Furthermore, hooks should be added to the \texttt{<BATCH\_SYSTEM\_NAME>.pm} module,
 so that it uses the correct Perl module depending on the LRMS type.
-
-The information collector scripts are divided between several separate scripts:
-\begin{itemize}
-\item \texttt{cluster+qju.pl} - driver for information collection.
-It calls all other information collectors and prints results in XML.
-The information collection is done by one single invocation of this script
-\item \texttt{InfoCollector.pm} - base class for all information collectors (i.e.: all files \texttt{*Info.pm}).
-\item \texttt{InfoChecker.pm} - used by \texttt{InfoCollector} to validate options and results against a ``schema''.
-\item \texttt{GMJobsInfo.pm} - collects information about jobs from grid manager status files
-\item \texttt{HostInfo.pm} - collects other information that can be collected on the front end (hostname, software version, disk space for users, installed certificates, Runtime environments ...)
-\item \texttt{LRMSInfo.pm} - collects information that is LRMS specific (queues, jobs, local user limits ...) by calling the appropriate LRMS plugin.
-It also does validation of input options to the plugin and of the data returned by the plugin.
-\item \texttt{BATCH\_SYSTEM\_NAME.pm} - plugins for LRMSInfo. Only Fork, SGE and PBS are updated to the new framework
-\item \texttt{ARC0ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the classic NG schema
-\item \texttt{ARC1ClusterInfo.pm} - combines all information about A-REX and produces information structured according to the GLUE2 schema
-\end{itemize}
-
-All information about the hosting environment and queues is collected by one invocation of the Perl script named
-\texttt{cluster+qju.pl}. The XML output is printed to \texttt{STDOUT} and it includes
-two representations of the output data, namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.
 
@@
 The \texttt{lrms\_info} function should then return a
-hash of hashes with the sctructure described in table \ref{t:lrms_info}.
+hash of hashes with the structure described in table \ref{t:lrms_info}.
 
@@
 freecpus & number of freely available CPUs with their time limits in minutes
-(see ComputingService.FreeSlotsWithDuration in\cite{glue2})\\
+(see ComputingService.FreeSlotsWithDuration in~\cite{glue2})\\
 \hline
@@
 including information on supported versions, constraints on batch
-system configuration, known limitations, \texttt{arc.conf} parameters
-and list of batch system features being utilized within the interface
+system configuration, known limitations, \texttt{arc.conf} parameters,
+and a list of batch system features being utilized within the interface
 are described.
@@
 
-PBS is a very powerful Local Resource Manager System with dozens of
+PBS is a very powerful LRMS with dozens of
 configurable options. Server, queue and node attributes can be used
@@
 only six (virtual) processors in your system). If you set both max\_running
-and max\_user\_run then obviously max\_user\_run has to be less equal
-than max\_running. 
+and max\_user\_run then obviously max\_user\_run has to be less or equal
+to max\_running.
 \item For the time being, do NOT set server limits like max\_running, please
@@
 \item Avoid using the max\_load and the ideal\_load directives. The node's
-mom config file (<PBS home on the node>/mom\_priv/config) should not
+mom configuration file (<PBS home on the node>/mom\_priv/config) should not
 contain any max\_load or ideal\_load directives. PBS closes down a
@@
 close down your node. 
-\item Routing queues are not supported, those can't be used within ARC.
+\item Routing queues are not supported, those cannot be used within ARC.
 \end{enumerate}
@@
 the flavour or the version number of the PBS, simply use the \texttt{{}``pbs''}
-keyword as lrms configuration value.
+keyword as LRMS configuration value.
 \item \texttt{pbs\_bin\_path} configuration variable of the \texttt{{[}common]}
@@
 block should be set to the path of the PBS server logfiles which are
-used by the Grid-Manager to determine whether a PBS job is completed.
-If not specified, the Grid-Manager will use the \texttt{qstat} command
+used by the GM to determine whether a PBS job is completed.
+If not specified, the GM will use the \texttt{qstat} command
 to find completed jobs. 
@@
 The job control batch interface makes use of the \texttt{qsub} command
-to submit native PBS jobscripts to the batch system. The following
+to submit native PBS job scripts to the batch system. The following
 options are used:
@@
 command is used to calculate total/used/free cpus within the cluster.
-In case of a Maui scheduller the \texttt{showbf} command is used to
+In case of a MAUI scheduler the \texttt{showbf} command is used to
 determine user freecpu values. All these external PBS commands are
@@
 The Fork back-end is a simple back-end that interfaces to the local
-machine i.e. there is no batch system underneath. It simply forks
-the job, hence the name. The back-end then uses standard posix commands(e.g.
-ps or kill) to manage the job.
+machine, i.e.: there is no batch system underneath. It simply forks
+the job, hence the name. The back-end then uses standard posix commands
+(e.g. ps or kill) to manage the job.
 
@@
 Since fork is a simple back-end and does not use any batch system,
-there is no specific configuration needed of the underlying system.
+there is no specific configuration needed for the underlying system.
 
@@
 to set the number of running grid jobs on the fork machine, allowing
-a multi core machine to use some or all of its cores for Grid jobs.
+a multi-core machine to use some or all of its cores for Grid jobs.
 The default value is 1. 
@@
 
-The Fork backend implements an interface to the {}``fork'' unix
-command which is not a batch system. Therefore the backend should
+The Fork back-end implements an interface to the {}``fork'' UNIX
+command which is not a batch system. Therefore the back-end should
 rather be seen as an interface to the operating system itself. Most
@@
 
-Fork is not a batch system, therefore many of the queue specific attributes
+Since Fork is not a batch system, many of the queue specific attributes
 or detailed job information is not available. The support for the
@@
 deployment of a real batch system since fork is available on every
-unix box. The Fork backend is not recommended to be used in production.
-The backend by its nature, has lots of limitations, for example does
+UNIX box. The "Fork back-end" is not recommended to be used in production.
+The back-end by its nature, has lots of limitations, for example it does
 not support parallel jobs.
@@
 Sun Grid Engine, formerly known as Codine, is an open source batch
-system maintained by Sun. Runs on Linux, Solaris.
+system maintained by Sun. It runs on Linux, and Solaris.
 
@@
 
-The SGE backend's commands are similar to the PBS commands. These
+The SGE back-end's commands are similar to the PBS commands. These
 commands are used in the code:
@@
 \item qconf -sconf global (show global configuration)
-\item qconf -sq queue\_name (show the given queue cofiguration)
+\item qconf -sq queue\_name (show the given queue configuration)
 \end{itemize}
@@
 Multi-CPU support is not well tested. All users are shown with the
-same quotas in infosys, even if they are mapped to different local
+same quotas in the information system, even if they are mapped to different local
 users. The requirement that one ARC queue maps to one SGE queue is
@@
 ARC's definition. The flexibility available in SGE for defining policies
-is difficult to accurately translate into nordugrid's information
+is difficult to accurately translate into NorduGrid's information
 schema. The closest equivalent of nordugrid-queue-maxqueuable is a
@@
 
-\section{The ARC GLUE 2 implementation}
+\section{The ARC GLUE2 implementation}
 ARC1 provides support for both the classic NorduGrid information
@@
 
-While the following entities are still missing (TODO: According to our table in wiki, there should be some other sections as well):
+While the following entities are still missing:
 \begin{itemize}
@@
 \vspace*{3mm}
-The following properties referring to job counts are calculated based on information in GridManager's status files:
+The following properties referring to job counts are calculated based on information in GM's status files:
 
@@
 \hline
-ServingState & has the same value as Computingendpoint.ServingState \\
+ServingState & has the same value as ComputingEndpoint.ServingState \\
 \hline
@@
 \subsection{ExecutionEnvironment}
+Not yet implemented.
 
 \subsection{ApplicationEnvironment}
+Not yet implemented.
 
 \subsection{ComputingActivity}
+Not yet implemented.
 
 \subsection{MappingPolicy and AccessPolicy}
-
-%\phantomsection
-%%\addcontentsline{toc}{subsubsection}{Name}
-%\hspace*{0.5cm}
-%\begin{shaded}
-%\textbf{Name}
-%\end{shaded}
-%\begin{tabular}{lp{10cm}}  
-%Attribute value:& String\\
-%Example:& \verb#Name: TODO#\\
-%Related xRSL:& TODO\\
-%UI role:& ?\\
-%Description:& Name of the application environment\\
-%\end{tabular}
+Not yet implemented.
 

### 2008-08-05 doc/client/arc-ws-clients.tex
Copied the old ap* clients documentation from workarea; modified title page to accommodate new content.

%\documentclass{book}
\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%\usepackage{bibtex}
\usepackage{moreverb}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {A.U. Thors},
  pdftitle = {WS-Based ARC Clients},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\newcommand{\uicommand}[1]{\textbf{\sffamily #1}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-TECH-20\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large WS-Based ARC Clients}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large Subtitile}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
{\centering \large A.U. Thors\footnote{authors@gmail.com} \large \par}

%%%%---- An abstract - if style is article ----
\begin{abstract}
This NorduGrid Technical document gives an introduction on how to use
the client tools developed for WS-based ARC solution.
\end{abstract}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage
%\chapter{Preface}
\section{Introduction}                    %Use Sections for articles
\label{sec:intro}

This document describes a set of simple client tools for the new Web
Service~\cite{ws} (WS) based Advanced Resource Connector~\cite{arc}
(ARC) middleware. These tools are prototypes and are intended to be a
proof-of-concept for the new ARC.

So far, the only service that is available for the new ARC is the ARC
Resource-Coupled Execution Service~\cite{arex} (A-REX). Therefore,
this prototype release of WS clients consists of job handling tools
with the following functionality:
\begin{itemize}
\item submission of jobs;
\item querying the status of jobs; and
\item termination of jobs.
\end{itemize}

The names of the command line tools begin with ``ap'', which shall be
interpreted as ``ARC Prototype'' to distinguish them from the
full-fledged production tools.

Support for staging of data would also have been desirable, but this
is not yet fully implemented on the server side. Therefore it could
not be included in this client prototype either. Thus, it is not
possible to run any jobs that depend on local files on the computer on
which the submitting client runs. Neither is it possible to retrieve
any form of output (not even stdout/stderr) by using the new client.

Behind the command line tools, there is a minimalistic library which
is documented in Appendix~\ref{app:library}. The library will serve as
a basis for the development of a generalized ARClib. This will be used
by the new client tools of the production releases of the next
generation of ARC.

\section{Configuration}
\label{sec:config}

The tools described here do not perform any brokering, but submit jobs
to a specific A-REX service. Which A-REX service to use is specified
in a configuration file. The configuration file also specifies how to
set up the communication chain for the client. The location of the
configuration file is specified by the environment variable
\verb:ARC_AREX_CONFIG:. If there is no such environment variable, the
configuration file is assumed to be \verb:arex_client.xml: in the
current working directory. An example configuration file is shown in
Appendix~\ref{app:config}.

\section{Job Submission}
\label{sec:apsub}

The \verb#apsub# command is used to submit jobs to an A-REX service.

\begin{shaded}
\paragraph{Usage} \verb:apsub <JSDL-file> <JobID-file>:
\end{shaded}

\paragraph{Arguments}
\begin{description}
\item[\texttt{<JSDL-file>}] The name of a file that contains
the job specification in JSDL format.
\item[\texttt{<JobID-file>}] The name of a file in which the
Job ID will be stored. This file is created when the command is
used. The reason for storing the Job ID in a file is its length
(approximately 500 characters) which makes it unsuitable to use as a
command line argument for the other tools.
\end{description}

\paragraph{Output} A message that the job was submitted and in which
file the Job ID is stored if the submission succeded. If the
submission failed for some reason, an error message is displayed.

\section{Job Status Query}
\label{sec:apstat}

The \verb#apstat# command is used to query the status of a job that
has been submitted to an A-REX service.

\begin{shaded}
\paragraph{Usage} \verb:apstat <JobID-file>:
\end{shaded}

\paragraph{Arguments}
\begin{description}
\item[\texttt{<JobID-file>}] The name of a file in which the
Job ID is stored. This file was created by \verb:apsub: when the job
was submitted.
\end{description}

\paragraph{Output} The status of the job, e.g. ``Pending/Accepted'',
or an error message if the status request failed for some reason.

\section{Job Termination}
\label{sec:apkill}

The \verb#apkill# command is used to terminate a job that has been
submitted to an A-REX service.

\begin{shaded}
\paragraph{Usage} \verb:apkill <JobID-file>:
\end{shaded}

\paragraph{Arguments}
\begin{description}
\item[\texttt{<JobID-file>}] The name of a file in which the
Job ID is stored. This file was created by \verb:apsub: when the job
was submitted.
\end{description}

\paragraph{Output} A message that the job was terminated, or an error
message if the termination failed for some reason.

\appendix

\include{ws_client_lib}

\section{Configuration Example}
\label{app:config}
Figure~\ref{fig:config} shows an example configuration file, which is
an XML file. All configuration information is enclosed within an
\verb:ArcConfig: element. That element also specifies which XML
manespaces to use.

The \verb:ModuleManager: element contains paths to directories where
loadable plugins can be found. In this case, there is only one path --
\verb:/usr/lib/arc/: -- but there may be many.

The \verb:Plugins: elements specify which plugins to load -- in this
case message chain components (MCCs) for TCP, TLS, HTTP and SOAP.

The \verb:Chain: element, finally, defines how to combine the MCCs
that have been loaded into a communication chain. In this case, a TCP
MCC that connects to port 6000 on the local host is connected to a TLS
MCC, which is connected to a HTTP that uses the POST method, and
finally a SOAP MCC is connected to the HTTP MCC.

The client tools will use the SOAP MCC as an entry point when sending
messages. The messages will then propagate through the MCCs in the
chain as specified in the configuration file.

\begin{figure}
\begin{shaded}
\verbatiminput{arex_client.xml}
\end{shaded}
\caption{An examle of an \texttt{arex\_client.xml} configuration file.}
\label{fig:config}
\end{figure}

\clearpage
\bibliography{grid}

\end{document}

### 2008-08-05 doc/client/ws_client_lib.tex
Copied the old ap* clients documentation from workarea; modified title page to accommodate new content.

\section{The A-REX Client Library}
\label{app:library}

Besides the three command line tools described earlier in this
document, there is also a small library on which the command line
tools are based. The library consists of two classes; the
\verb:AREXClient: class and the \verb:AREXClientError: class.

\subsection{The \texttt{AREXClient} class}

This class is a client class for the A-REX service. It provides
methods for three operations on an A-REX service:
\begin{itemize}
\item Job submission
\item Job status query
\item Job termination
\end{itemize}

\subsubsection{Constructor}

\begin{shaded}
\verb#AREXClient(std::string configFile="")#
\end{shaded}

This is the constructor for the AREXClient class. It creates an A-REX
client that corresponds to a specific A-REX service, which is
specified in a configuration file. The configuration file also
specifies how to set up the communication chain for the client. The
location of the configuration file can be provided as a parameter to
the method. If no such parameter is given, the environment variable
\verb:ARC_AREX_CONFIG: is assumed to contain the location. If there is
no such environment variable, the configuration file is assumed to be
\verb:arex_client.xml: in the current working directory.

\paragraph{Parameters}
\begin{description}
\item[\texttt{configFile}] The location of the configuration file.
\end{description}

\paragraph{Throws}

An \verb:AREXClientError: object if an error occurs.

%\subsubsection{Destructor}

%\begin{shaded}
%\verb#~AREXClient()#
%\end{shaded}

%This is the destructor. It does what destructors usually do, cleans up.

\subsubsection{Job submission}

\begin{shaded}
\verb#std::string submit(std::istream& jsdl_file)#
\end{shaded}

This method submits a job to the A-REX service corresponding to this
client instance.

\paragraph{Parameters}
\begin{description}
\item[\texttt{jsdl\_file}] An input stream from which the JSDL file for
the job can be read.
\end{description}

\paragraph{Returns}
The Job ID of the the submitted job.

\paragraph{Throws}
An \verb:AREXClientError: object if an error occurs.

\subsubsection{Job status query}

\begin{shaded}
\verb#std::string stat(const std::string& jobid)#
\end{shaded}

This method queries the A-REX service about the status of a
job.

\paragraph{Parameters}
\begin{description}
\item[\texttt{jobid}] The Job ID of the job.
\end{description}

\paragraph{Returns}
The status of the job.

\paragraph{Throws}
An \verb:AREXClientError: object if an error occurs.

\subsubsection{Job termination}

\begin{shaded}
\verb#void kill(const std::string& jobid)#
\end{shaded}

This method sends a request to the A-REX service to terminate a job.

\paragraph{Parameters}
\begin{description}
\item[\texttt{jobid}] The Job ID of the job.
\end{description}

\paragraph{Throws}
An \verb:AREXClientError: object if an error occurs.

\subsection{The \texttt{AREXClientError} class}

This is an exception class that is used to handle runtime errors
discovered in the AREXClient class.

\paragraph{Inherits}
This class inherits \verb#std::runtime_error#

\subsubsection{Constructor}

\begin{shaded}
\verb#AREXClientError(const std::string& what="")#
\end{shaded}

This is the constructor of the AREXClientError class.

\paragraph{Parameters}
\begin{description}
\item[\texttt{what}] An explanation of the error.
\end{description}

### 2008-08-07 doc/storage_system/arc1-storage-documentation.tex
added "NORDUGRID-TECH-17"
added Jon and Salman as authors
use \url{}

@@
 
-{\raggedleft NORDUGRID-XXXXXXX-NN\par}
+{\raggedleft NORDUGRID-TECH-17\par}
 
@@
     {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
-
+    {\centering \large Jon Nilsen\footnote{j.k.nilsen@usit.uio.no} \large \par}
+    {\centering \large Salman Zubair Toor \footnote{salman.toor@it.uu.se} \large \par}
 \end{titlepage}
@@
 The replicas of the files are stored on different storage nodes.
-A storage node here is a network-accessible computer having storage space to share, and one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), https://forge.gridforum.org/projects/byteio-wg/}, etc.).
+A storage node here is a network-accessible computer having storage space to share, and one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.).
 For each storage node one of the services of the ARC1 storage system is needed to manage it and to integrate it into the system.
@@
 
-One possible usage scenario would be if each VO (Virtual Organization)\footnote{Virtual Organization, http://www.globus.org/grid\_software/security/voms.php} would create a VO-wide root collection, and put it in the generic root collection, e.g.~if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#.
+One possible usage scenario would be if each VO (Virtual Organization)\footnote{Virtual Organization, \url{http://www.globus.org/grid\_software/security/voms.php}} would create a VO-wide root collection, and put it in the generic root collection, e.g.~if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#.
 Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g.~\verb#/vo1/file1#, etc.
@@
 
-The services are written in Python and running in the HED\footnote{The ARC container - https://www.knowarc.eu/documents/Knowarc\_D1.2-2\_07.pdf} hosting environment.
+The services are written in Python and running in the HED\footnote{The ARC container - \url{https://www.knowarc.eu/documents/Knowarc\_D1.2-2\_07.pdf}} hosting environment.
 The HED itself is written in C++, but there are language bindings which allow us to write services in other languages, e.g.~in Python or Java.
-The source code of the storage services are in the NorduGrid Subversion repository\footnote{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage}.
+The source code of the storage services are in the NorduGrid Subversion repository\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage}}.
 
@@
 The A-Hash service currently implemented as a single central service, which stores the data on disk in separate files per \emph{object}.
-Fall of 2008 it will be reimplemented using a distributed hash table (DHT) algorithm, one possible candidate is the Chord\footnote{The Chord project - http://pdos.csail.mit.edu/chord/} algorithm with a consistency solution called Etna\footnote{Etna: a Fault-tolerant Algorithm for Atomic Mutable DHT Data - http://pdos.csail.mit.edu/\~athicha/papers/etna.ps} on top of it.
+Fall of 2008 it will be reimplemented using a distributed hash table (DHT) algorithm, one possible candidate is the Chord\footnote{The Chord project - \url{http://pdos.csail.mit.edu/chord/}} algorithm with a consistency solution called Etna\footnote{Etna: a Fault-tolerant Algorithm for Atomic Mutable DHT Data - \url{http://pdos.csail.mit.edu/\~athicha/papers/etna.ps}} on top of it.
 This reimplementation hopefully won’t change the interface of the service.
@@
 
-You can find more examples in the SVN\footnote{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage/README}.
+You can find more examples in the SVN\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage/README}}.
 
-There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, http://fuse.sourceforge.net/} module and Windows Shell Extensions\footnote{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx} to be able to mount the ARC1 storage namespace into the local filesystem namespace, and use it with the commands of the operating system.
+There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, \url{http://fuse.sourceforge.net/}} module and Windows Shell Extensions\footnote{\url{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx}} to be able to mount the ARC1 storage namespace into the local filesystem namespace, and use it with the commands of the operating system.
 

### 2008-08-07 doc/backend_interface/Backends-arc1.tex
Updated the document according to the KnowARC review

@@
 \end{quote}
+
+The backend scripts install along with the ARC1 code.
+
 The next generation of ARC supports both the classic NorduGrid information schema~\cite{is} and
@@
 
+\subsection{Submit-LRMS-job}
+
+The submit program is the most involved. It is called by the GM once
+a new job arrives and needs to be submitted to the LRMS. It is given
+the GRAMi file as argument on execution. The GRAMi file is a file in
+the job control directory containing the job description in a flat list
+of key-value pairs. This file is created by GM and is based on the JSDL
+job description. Submit-LRMS-job then has to set up the session
+directories, run-time environment and anything else needed. Then it
+submits the job to the local LRMS. This is normally done by generating
+a native job script for the LRMS and then running the local submit
+command, but it can also be done through an API if the LRMS supports it.
+
 \subsection{Cancel-LRMS-job}
@@
 The manager will then call the cancel-LRMS-job for the suitable back-end.
-The cancel script is called with a text file containing information
-about the job. It is called the GRAMi file. It contains such information
-as the original xRSL option and the job id in the local LRMS. Cancel-LRMS-job
-must then use that information to find the job and remove it from
-the queue or actually cancel it if it is running in the LRMS.
-
-\subsection{Submit-LRMS-job}
-
-The submit program is the most involved. It is called by the GM once
-a new job arrives and needs to be submitted to the LRMS. Like cancel-LRMS-job
-it is given the GRAMi file as argument on execution. Submit-LRMS-job
-then has to set up the session directories, run-time environment and
-anything else needed. Then it submits the job to the local LRMS.
-This is normally done by generating a native job script for the LRMS
-and then running the local submit command, but it can also be done
-through an API if the LRMS supports it.
+The cancel script is called with the GRAMi file containing information
+about the job such as the job id in the LRMS. Cancel-LRMS-job must then
+use that information to find the job and remove it from the queue or
+actually cancel it if it is running in the LRMS.
 
@@
 
-Scan-LRMS-job is run periodically. Its job is to scan the LRMS for
+The scan-LRMS-job is run periodically. Its job is to scan the LRMS for
 jobs that have finished. Once it has found a finished job it will
 write the exit-code of that job to the file job.\{gridid\}.lrms\_done
-in the ARC jobstatus directory. Then it will call the gm-kick program
-to notify GM about the finished job. Subsequently, the GM starts finalizing
-the job.
-
-Generally, two approaches are taken to find jobs that are done. One
-is to ask the LRMS. First all started grid jobs are found in the jobstatus
-directory%
-\footnote{normally /var/spool/nordugrid/jobstatus/, but can be set via the controldir
-variable of arc.conf%
-}. This can be done by checking if the status is "INLRMS" in the jobs
-status file%
-\footnote{job.\{gridid\}.status%
-}. Then the LRMS is asked for the status of those jobs. If they are
-done they are marked as such and the GM is activated. The problem
-with this approach is that for most LRMSs the information about finished
-jobs are only available for a short period after the job finished.
-Therefore appropriate steps have to be taken if the job is known to
-have been started but is no longer present in the LRMS. The normal
-approach is to analyze the jobs status output in the session directory.
+in the ARC job status directory\footnote{normally /var/spool/nordugrid/jobstatus/,
+but can be set via the controldir variable of arc.conf}. Then it will call the
+gm-kick program to notify GM about the finished job. Subsequently, the GM starts
+finalizing the job.
+
+Generally, two approaches are taken to find jobs which are finished in LRMS. One
+is to directly ask the LRMS. Since all started grid jobs have its own status
+file\footnote{job.\{gridid\}.status} found in the job status directory,
+this can be done by checking if the status is "INLRMS" in this file. If so, a
+call to the LRMS is made asking for the status of the job (or jobs if several
+jobs have status "INLRMS"). If it is finished, it is marked as such in the
+job status directory, and the gm-kick program is activated. For most LRMSs the
+information about finished jobs are only available for a short period of time after
+the job finished. Therefore appropriate steps have to be taken if the job has the
+status "INLRMS" in the job status directory, but is no longer present in the LRMS.
+The normal approach is to analyze the job's status output in the session directory.
 
 The second approach is to parse the LRMSs log files. This method has
-some drawbacks, e.g. the GM has to be allowed read access to the logs.
+some drawbacks like e.g.: the GM has to be allowed read access to the logs.
 The back-end will then have to remember where in the log it was last
@@
 \label{t:queue_info}
+\begin{minipage}{\textwidth}
 \begin{center}
@@
 \hline
-maxwalltime & limit on maximum wall time for a job in this queue in minutes\\
+maxwalltime & limit on maximum wall time\footnote{Also known as Wall clock
+time, \url{http://en.wikipedia.org/wiki/Wall_clock_time}} for a job in this queue in minutes\\
 \hline
@@
 \end{center}
+\end{minipage}
 \end{table}
@@
 \label{t:jobs_info}
+\begin{minipage}{\textwidth}
 \begin{center}
@@
 \hline \hline
-status & LRMS jobstatus mapping\footnotemark[1]:
+status & LRMS jobstatus mapping\footnote{LRMS job states mapping is described in\cite{is}}:
 Running $\Rightarrow$ 'R',
@@
 \hline
-walltime & consumed wall-time in minutes\\
+walltime & consumed wall time in minutes\\
 \hline
@@
 \hline
-reqwalltime & wall-time requested by job in minutes\\
+reqwalltime & wall time requested by job in minutes\\
 \hline
@@
 \end{tabular}
-\footnotetext[1]{LRMS job states mapping is described in\cite{is}}
 \end{center}
+\end{minipage}
 \end{table}
@@
 use queue-based limits instead. 
-\item Avoid using the max\_load and the ideal\_load directives. The node's
-mom configuration file (<PBS home on the node>/mom\_priv/config) should not
+\item Avoid using the max\_load and the ideal\_load directives. The Node Manager
+(MOM) configuration file (<PBS home on the node>/mom\_priv/config) should not
 contain any max\_load or ideal\_load directives. PBS closes down a
@@
 \item \texttt{maui\_bin\_path} from the \texttt{{[}queue/queuename]} block
-sets the path of the maui commands like \texttt{showbf} when \char`\"{}MAUI\char`\"{}
+sets the path of the MAUI commands like \texttt{showbf} when \char`\"{}MAUI\char`\"{}
 is specified as \texttt{scheduling\_policy} value. This parameter
@@
 Sun Grid Engine, formerly known as Codine, is an open source batch
-system maintained by Sun. It runs on Linux, and Solaris.
+system maintained by Sun. It is supported on Linux, and Solaris in
+addition to a numerous other systems.
 
@@
 \hline
-SchedulingPolicy & if maui is used, set to \texttt{fairshare} \\
+SchedulingPolicy & if MAUI is used, set to \texttt{fairshare} \\
 \hline

### 2008-08-11 doc/glite_gateway/glite_gateway.tex
ldap query made more confortable
Installation guide extended

@@
 \section{Building, dependencies}
+\subsection{Generic installation disciplines}
 \label{Building}
@@
 There are also some details on how to compile VOMS from source at NorduGrid VOMS compilation page\footnote{\url{http://wiki.nordugrid.org/index.php/VOMS\_compiling\_detailes}}.
+\subsection{Software installation on Debian Linux}
+This section contains instructions how to install the software onto Debian Linux.
+\begin{enumerate}
+\item Put the Debian repository into the APT source files (/etc/apt/sources.list).
+\begin{shaded}\verb#$ echo "deb http://pc02.inb.uni-luebeck.de:8080/~moeller/debian/stable ./" \#\\*
+\verb#        >> /etc/apt/sources.list #\end{shaded}
+\item Update the package list and install the necessary dependencies.
+\begin{shaded}\verb#$ aptitude update#\\*
+\verb#$ aptitude install voms globus globus-dev globus-doc#\end{shaded}
+\item Configure the Globus libraries.
+\begin{shaded}\verb#$ echo "/opt/globus/lib" > /etc/ld.so.conf.d/globus.conf#\\*
+\verb#$ ldconfig#\end{shaded}
+\item Install ARC1 client. There is a broken dependency that installs separately.
+\begin{shaded}\verb#$ aptitude install nordugrid-arc1-client#\end{shaded}
+\item Install Globus and VOMS specific files.
+\begin{shaded}\verb#$ mkdir -p /opt/glite/etc/#\\*\end{shaded}
+and copy these lines into the vomses file located /opt/glite/etc/vomses
+\begin{footnotesize}
+\begin{shaded}
+\verb#"gin.ggf.org" "kuiken.nikhef.nl" "15050" "/O=dutchgrid/O=hosts/OU=nikhef.nl/CN=kuiken.nikhef.nl" "gin.ggf.org"#\\*
+\verb#"knowarc.eu" "arthur.hep.lu.se" "15001" "/O=Grid/O=NorduGrid/CN=host/arthur.hep.lu.se" "knowarc.eu"#\\*
+\verb#"atlas" "voms.cern.ch" "15001" "/DC=ch/DC=cern/OU=computers/CN=voms.cern.ch" "atlas"#\\*
+\verb#"nordugrid.org" "voms.uninett.no" "15015" "/O=Grid/O=NorduGrid/CN=host/voms.ndgf.org" "nordugrid.org"#\\*
+\end{shaded}
+\end{footnotesize}
+\item Create \$HOME/.globus directory and move the user certificate-key pair into there.
+\end{enumerate}
 \section{Using the ARC${\rightarrow}$gLite Gateway from the Command Line}
@@
 \begin{shaded}
-\verb#$ ldapsearch -x -h lxbra2305.cern.ch -p 2170 -b mds-vo-name=local,o=grid //#\\*
-\verb#        '(GlueCEUniqueID=*cream*)'#
+\verb#$ ldapsearch -x -h lxbra2305.cern.ch -p 2170 -b mds-vo-name=local,o=grid \#\\*
+\verb#        '(GlueCEUniqueID=*cream*)' | grep GlueCEInfoContactString | \#\\*
+\verb#        awk '{print $2}' | sort | uniq#
 \end{shaded}
@@
 For example:
-\begin{shaded}\verb#$ glitedelegate test_delegation //#\\*
+\begin{shaded}\verb#$ glitedelegate test_delegation \#\\*
 \verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}
@@
 An example how to use this command:
-\begin{shaded}\verb#$ glitesub -D test_delegation //#\\*
+\begin{shaded}\verb#$ glitesub -D test_delegation \#\\*
 \verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/CREAM2 description.jdl job.info#\end{shaded}
@@
 For example:
-\begin{shaded}\verb#$ gliteundelegate test_delegation //#\\*
+\begin{shaded}\verb#$ gliteundelegate test_delegation \#\\*
 \verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}

### 2008-08-12 doc/glite_gateway/glite_gateway.tex
Major modification on the document

@@
 
-\def\documenttitle{The ARC gLite Gateway}
+\def\documenttitle{NOT\_READY\_The glite gateway plugin of ARCLIB and related command line tools}
 \captionsetup{margin=20pt,font={footnotesize, it},labelfont=bf}
@@
 
-{\raggedleft Knowarc\_D3.2-1\_08\par}
+{\raggedleft NORDUGRID-TECH-22\par}
 
@@
 %%%%---- A list of authors ----
+{\centering \large Mattias Ellert\footnote{mattias.ellert@fysast.uu.se} \large \par}
 {\centering \large Ivan Marton\footnote{martoni@niif.hu} \large \par}
@@
 \label{Introduction}
-The purpose of this document is to describe how Advanced Resource Connector version 1 (ARC1\footnote{\url{http://www.nordugrid.org/}}) users can execute grid jobs on EGEE CREAM2\footnote{\url{http://grid.pd.infn.it/cream/}} resources.\par In sections \ref{Introduction} and \ref{Interoperability} a brief introduction on interoperability issues will be given.
-Section \ref{Building} describes how to build the gateway library, section \ref{Users guide} shows how it can be used.
-Section \ref{Developers notes} gives a tiny developer-oriented introduction on how the gateway library is built up, while section \ref{Outstanding issues} addresses the current issues with the code.
-Section \ref{Conclusions} concludes the technical document.
-\section{Interoperability}
-\label{Interoperability}
+The purpose of this document is to describe this gLite specific plugin of the ARCLIB (Hivatkozni: jövőbeli Deliverable) to present the related command line utilities built on top of these libraries.
+The gateway library together with the tools enables seemless interoparability with gLite computing elements.
+It is possible, that relevant sections of this document will be moved into ARC1 userguide and into ARCLIB technical documentation.
+Interoperability with gLite storage element is provided by an other set of ARC library.
+This data library provides native interfaces to GridFTP SRM which is the main bulding blocks of a gLite storage element.
+A storage interopability is out of scope of this library.
+\section{Overview}
+\label{Overview}
 Interoperability has been one of the most important disciplines in recent grid middleware development.
@@
 \item either to use a special infrastructure element, the gateway, which performs a full data structure and protocol translation between the two inter-operating grid middleware solutions,
-\item or to use standard or close-to-standard interfaces on both sides.
+\item or to explicitly implement the interface of the other side in the client.
 \end{itemize}
-The latter solution which is preferred from standardization point of view and also has been chosen by ARC1 developers is illustrated in figure \ref{fig:Chart 1}.\par
+The latter solution was preferred by ARC1 developers, the gLite gateway library module has been implemented in the ARCLIB and the gateway functionality will be transparently offered via the usual Nordugrid command line tools.
+An overview of the preferred soulution is presented in figure \ref{fig:Chart 1}.\par
 \begin{figure}[ht]
 \centering{\resizebox{\textwidth}{!}{\includegraphics{D3-2-1-figure.png}}
-\caption{\label{fig:Chart 1}This chart illustrates how the ARC1$\rightarrow$gLite gateway library works in general.
-After successful authentication and credential delegation to the CREAM2 resource, users submit grid jobs through it.
-A wrapper script at the computing element may pull input files from different storage elements using storage protocol, say, GridFTP.
-Users get job information through CREAM2 and may send management commands, like kill a job or undelegate credentials.} }
+\caption{\label{fig:Chart 1}This chart illustrates how the ARC1$\rightarrow$gLite gateway library module based approach works.
+The user will be able to express his requirements in the job description document using any of the xRSL, JDL or JSDL languages.
+The ARC client library will perform an automatic transformation.
+The gLite gateway functionality integrated into the ARCLIB, thus into the ARC client enables the user to transparently access gLite computing resources via the usual ARC command line tools.
+The gLite modules perform authentication, credential delagation and job submission to the CREAM2 CE.
+The CREAM2 CE may perform the necessary input file download from the different storage element.
+The library module gets job information through CREAM2 and sends management commands, like kill a job or undelegate credentials.
+Note, that not all the functionality described on the overview figure is implemented yet, for the current limitations see section \ref{Current limitations}!}}
 \end{figure}
-\section{CREAM2}
-The main purpose of this project was to provide access to gLite CREAM2-enabled resources from ARC1 user interface, but not the other way around: i.e.
-accessing ARC resources from gLite is addressed in the EGEE project\footnote{\url{http://www.eu-egee.org/}}.\par
+\subsection{The supported gLite CE interface: CREAM2}
+The gLite computing element (CE) offers several interfaces.
+Some of them are only suitabe for the WMS (glite resource broker) and was not meant to be used by 3rd party clients.
+Fortunately gLite comes with a webservice based computing element interface, known as the CREAM CE.
+The CREAM has undergone several major revisions.
+We can only hope that some stability will soon arise.
+The current CREAM version has entered the EGEE certification process newertheless it is a full scale deplomeyment on the EGEE worldwide grid has not taken place yet.
+CREAM2 is available on several preproduction and testing grid sites.\par
+Nevertheless the KnowARC project chose the CREAM version 2\cite{cream} as the gLite computing element interface.\par ... Hivatkozás a cikkre, mint CREAM-re...
+In particular, the gateway library was developed against version glite-ce-cream\_R\_1\_8\_3\_0\footnote{\url{http://grid.pd.infn.it/cream/field.php?n=Main.ReleaseNotes}}.\par
 CREAM2 was chosen for the following reasons:
 \begin{itemize}
-\item It is a web service interface that fits to the built-in web service implementation of ARC1.
-\item It contains numerous improvements compared to the original CREAM interface. (Unfortunately CREAM2 is not backwards compatible with CREAM.)
-\item It supports job status queries compared to the other web service interfaces of gLite, like WMProxy.
-\item It is an official interface to gLite computing elements and as EGEE evolves it is expected that more and more computing elements will have CREAM2 interfaces.
+\item CREAM2 has a web service interface that fits the web service based ARC1.
+\item CREAM2 enables direct access to the gLite computing element without the need of going through the WMS.
+\item CREAM2 contains numerous improvements compared to the earlier CREAM versions.
+\item Unlike the other gLite CE interfaces it supports direct job status queries 
+\item It is an official interface to gLite computing elements and as EGEE evolves it is expected that more and more computing elements will be deployed with CREAM2 interface.
+\item CREAM2 offers a convenient way of handling input and output files through accessing the input and output sandbox via GridFTP.
 \end{itemize}
-\section{Building, dependencies}
-\subsection{Generic installation disciplines}
-\label{Building}
-The ARC$\rightarrow$gLite gateway software is incorporated into ARC1 releases, so you only need to download either the freshest ARC1 from the source tree\footnote{\url{http://svn.nordugrid.org/}} and compile it or to install it from package\footnote{\url{http://download.nordugrid.org/software/nordugrid-arc1/}}.\par
-The gateway library code has two dependencies:
-\begin{itemize}
-\item Globus Toolkit is needed to enable GridFTP functionality to be able to upload input files from the user interface to the CREAM computing node.
-\item VOMS is needed to allow users authentication on a CREAM resource.
-\end{itemize}
-In both cases it is highly recommended to download these pieces of software along with their NorduGrid-specific Globus\footnote{\url{http://download.nordugrid.org/software/globus/}} and VOMS\footnote{\url{http://download.nordugrid.org/software/voms/}} patches from the NorduGrid repository.
-In case you are interested in the details of patching, then consult the NorduGrid wiki page\footnote{\url{http://wiki.nordugrid.org/images/4/46/NorduGrid-globus.patch.txt}}.
-It might also be interesting to read how to compile Globus Toolkit.
-Some more details on this topic can be found on the NorduGrid Globus compilation page\footnote{\url{http://wiki.nordugrid.org/index.php/Globus\_Libraries}}.
-There are also some details on how to compile VOMS from source at NorduGrid VOMS compilation page\footnote{\url{http://wiki.nordugrid.org/index.php/VOMS\_compiling\_detailes}}.
-\subsection{Software installation on Debian Linux}
-This section contains instructions how to install the software onto Debian Linux.
-\begin{enumerate}
-\item Put the Debian repository into the APT source files (/etc/apt/sources.list).
-\begin{shaded}\verb#$ echo "deb http://pc02.inb.uni-luebeck.de:8080/~moeller/debian/stable ./" \#\\*
-\verb#        >> /etc/apt/sources.list #\end{shaded}
-\item Update the package list and install the necessary dependencies.
-\begin{shaded}\verb#$ aptitude update#\\*
-\verb#$ aptitude install voms globus globus-dev globus-doc#\end{shaded}
-\item Configure the Globus libraries.
-\begin{shaded}\verb#$ echo "/opt/globus/lib" > /etc/ld.so.conf.d/globus.conf#\\*
-\verb#$ ldconfig#\end{shaded}
-\item Install ARC1 client. There is a broken dependency that installs separately.
-\begin{shaded}\verb#$ aptitude install nordugrid-arc1-client#\end{shaded}
-\item Install Globus and VOMS specific files.
-\begin{shaded}\verb#$ mkdir -p /opt/glite/etc/#\\*\end{shaded}
-and copy these lines into the vomses file located /opt/glite/etc/vomses
-\begin{footnotesize}
-\begin{shaded}
-\verb#"gin.ggf.org" "kuiken.nikhef.nl" "15050" "/O=dutchgrid/O=hosts/OU=nikhef.nl/CN=kuiken.nikhef.nl" "gin.ggf.org"#\\*
-\verb#"knowarc.eu" "arthur.hep.lu.se" "15001" "/O=Grid/O=NorduGrid/CN=host/arthur.hep.lu.se" "knowarc.eu"#\\*
-\verb#"atlas" "voms.cern.ch" "15001" "/DC=ch/DC=cern/OU=computers/CN=voms.cern.ch" "atlas"#\\*
-\verb#"nordugrid.org" "voms.uninett.no" "15015" "/O=Grid/O=NorduGrid/CN=host/voms.ndgf.org" "nordugrid.org"#\\*
-\end{shaded}
-\end{footnotesize}
-\item Create \$HOME/.globus directory and move the user certificate-key pair into there.
-\end{enumerate}
-\section{Using the ARC${\rightarrow}$gLite Gateway from the Command Line}
+In the near future CREAM will offer yet another interface which will be based on the BES/JSDL/GLUE OGF standards. This CREAM BES interface will coexists with the CREAM2 too.\par
+\section{External dependencies and installation instructions}
+\label{External dependencies and installation instructions}
+The ARCLIB gLite plugin have only one dependency.
+GridFTP components of the Globus Toolkit are needed to enable data staging functionality to those CREAM2 computing elements which offer input and output sandbox over a GridFTP enabled storage.
+Unfortunately the gateway library only works together with a strongly patched Globus verion.
+The Nordugrid patched Globus can be dowloaded from the Nordugrid download area\footnote{\url{http://download.nordugrid.org/software/globus/}}.
+Further information about the Nordugrid Globus patches can be read on the Nordugrid page\footnote{\url{http://wiki.nordugrid.org/index.php/Globus\_Libraries}}.
+It is worth to mention that the user will need a VOMS proxy certificate in order to use the command line tools against a CREAM2 service.
+This VOMS proxy can be generated by using the VOMS client tools.
+It is only the VOMS enabled proxy certificate that is needed here, appart from this no VOMS component needed.\par
+The ARC gLite gateway plugin library and the temporary command line utilities are incorporated into ARC1 releases.
+The source code is available from the ARC1 SVN repository\footnote{\url{http://svn.nordugrid.org/repos/nordugrid/arc1/}} and binary packages are provided at the Nordugrid download area.
+The build and installation instructions are given in the README file distributed together with the software.
+A step-by-step guide for Debian binary installation can be found in the appendix \ref{Debian install} Installation from Debian binary packages.
+\section{ARCLIB gLite gateway plugin}
+\label{ARCLIB gLite gateway plugin}
+The ARCLIB gLite gateway plugin constist of the wrapper classes, the job description handling facility and the gLite specific CREAMClient base class.
+The wrapper classes coveys the function provided by the CREAMClient to ARCLIB.
+The job description handling framework makes a general job request representation available for the other classes.
+CREAMClient base class implements the interface towards CREAM2 computing elements.
+This client object has an easy-to-use and intuitive interface. It uses the VOMS proxy certificate to build the secure channel to the server and to sign that of the server's.
+The different functions throw \textit{CREAMClientError} exceptions that should be caught in the application.\par
+The following sections present the set of functions provided by the library.\par
+\subsection{Wrapper classes}
+... mik ezek a wrapper osztályok ... milyen base class, mit csinál, mi a funckiójuk ... viszonylag részletesen ...
+\subsection{Job description translators}
+... Bárhol meghívható ... objektumot példányosítani ... milyen transzformációk állank rendelkezésre ...
+\subsection{CREAMClient base class}
+\label{CREAMClient}
+The client object uses the VOMS proxy certificate to build the secure channel to the server and to sign that of the server's.\par
+...
+1 bekezdés arról, hogy SOAP a hed(Hivatkozni!!!) SOAP framework-jét használja 'fully integrated'...
+az előző deliverable-ben a kommunikáció gsoap-ot használt, ez az egyik nagy előrelépés, hogy ezt lecseréltük...
+\subsubsection*{CREAMClient(Arc::URL, Arc::MCCConfig)}
+The constructor of the class performs the necessary initialization work.
+It receives two arguments: the service URL and the ARC message chain component configuration file to establish the communication channel to the server.
+These pieces of information are indispensable to create your client and to communicate with the remote site.
+\subsubsection*{setDelegationId(std::string)}
+Besides the constructor this is the other generic-purpose function that can be used in CREAM2 client applications.
+This function sets the previously registered and referred delegation ID on the client.
+The function is very simple: just set a private variable and returns with no value.
+\subsubsection*{createDelegation(std::string)}
+The \textbf{createDelegation} function is about to perform the whole delegation registration process.
+It sends a \textit{getProxyReq} message having the requested delegation ID, signs the received certificate and sends it back in a \textit{putProxy} SOAP message to the server.
+These three steps constitute the delegation registration process.
+As almost every funtion, this one has no return value either.
+If there are any problems during the communication, either locally in the channel or at the remote site, the function throws \textit{CREAMClientError} described previously.
+If there are no exceptions thrown, then the command is considered to have successfully completed.
+\subsubsection*{destroyDelegation(std::string)}
+This method has the reverse functionality of the \textbf{createDelegation} function. It sends the \textit{destroy} message of CREAM2 to the remote server.
+\subsubsection*{submit(std::string)}
+The \textbf{submit} function is the most complex part of the class.
+It translates the job description received as the argument, registers the job with a \textit{JobRegisterRequest} message, uploads the locally stored files, if necessary, then enables the job execution on the server by sending a \textit{JobStartRequest} message.
+Finally it returns with a \textbf{creamJobInfo} object that contains a \textit{jobId} (job indetifier), a \textit{creamURL} (as the service URL), an \textit{ISB} (reference to the Input Sandbox) and an \textit{OSB} (Output Sandbox) member to describe the registered job.
+In the \textbf{glitesub} command (See Section \ref{glitesub} for details) these pieces of information are stored in the information file.
+\subsubsection*{stat(std::string)}
+This method queries the job status from the gLite CREAM2 server.
+The return value is the job status translated to the ARC terminology.
+The possible values are presented in the section of the \textbf{glitestat} command line tool.
+(See Section \ref{glitestat}!)
+\subsubsection*{cancel(std::string)}
+The \textbf{cancel} function sends a \textit{JobCancelRequest} SOAP message to the server and handles the emerging exceptions.
+It can be used for canceling a remotely registered and possibly running job.
+\subsubsection*{purge(std::string)}
+The \textbf{purge} function sends a \textit{JobPurgeRequest} message to server and throws an exception in case of any emerging problems.
+\section{ARC command line tools for gLite CE}
 \label{Users guide}
+The well known ARC commands (ngsub, ngget, ...
+See: The Nordugrid/ARC User Guide\cite{userguide}!) will provide seemless access and integration with gLite computing elements.
+Full transparency will be achieved through the gLite gateway ARCLIB plugin.
+An ordinary user will use the same set of commands regardless of what type of CE to access.
+The user will be able to specify grid jobs described both in the gLite (JDL) and the Nordugrid (JSDL, XRSL) job description languages.
+...
++++Kipofozni.
+Hivatkozások...
+The client will perform automatic conversion between the different job description languages.\par
+Unfortunately the interoperability features offered by the gLite gateway library are not yet integrated into the Nordugrid commands.
+Also the current version of ARCLIB does not offer the full spectrum of language transformations.
+However only a set of temporary command line tools exposing the available features of library are provided.
+These temporary commands make the already existing interoparabilily features usable for ordinary users.
 There are seven different command line tools for using the gLite gateway functions.
@@
 Every tool has a short manual page being accessible by using the \textbf{man} on-line manual reader command or some more help can be achieved by using the $<$command$>$ -? command line option.\par
-Before getting into the details of the gatewy toolkit, the authentication and authorization processes need clarification.\par
-In the different grid systems the most widespread authentication method is the certificate based authentication.
-GLite also requires VOMS-enabled proxy X509 certificates to access resources.
-It means that you should be a member of the Virtual Organisation that is valid on the remote site.
-If the VOMS package has been correctly installed then the following command provides the necessary VOMS proxy certificate.\par
-\begin{shaded}\verb#$ voms-proxy-init -voms <Your Virtual Organisation>#\end{shaded}
+\subsection{glitedelegate}
+\label{glitedelegate}
+In order to use gLite resources user must have a VOMS proxy. This proxy can be generated by command \textit{voms-proxy-init}.\par
 For example:
 \begin{shaded}\verb#$ voms-proxy-init -voms knowarc.eu#\end{shaded}
-CREAM2 service URLs can directly be downloaded from any BDII database of gLite using the following LDAP command:
-\begin{shaded}
-\verb#$ ldapsearch -x -h lcg-bdii.cern.ch -p 2170 -b o=grid '(GlueCEUniqueID=*blah*)'#
-\end{shaded}
-or
-\begin{shaded}
-\verb#$ ldapsearch -x -h lxbra2305.cern.ch -p 2170 -b mds-vo-name=local,o=grid \#\\*
-\verb#        '(GlueCEUniqueID=*cream*)' | grep GlueCEInfoContactString | \#\\*
-\verb#        awk '{print $2}' | sort | uniq#
-\end{shaded}
-
-\subsection{glitedelegate}
-\label{glitedelegate}
-If you want to submit a job to a gLite server, first you have to own a valid delegation on the remote site.
-You can either use an old, but still valid delegation identifier, or register a new one.
-There is a command line tool to register a delegation on the resource to be used.\par
-The usage is very easy and straighforward.
-All you need is a working CREAM2 service URL, and a locally unique arbitrary delegation ID.
+Once a VOMS proxy has been generated everything is in place to submit a new job.
+If one wants submit a job to a gLite server, first has to own a valid delegation on the remote site.
+This can either be an old, but still valid delegation identifier, or can be a new one.
+There is a command line tool to register a delegation on the CE to be used.\par
+The usage of command \textbf{glitedelegate} is straighforward.
+All needed is a working CREAM2 delegation service URL, and a locally unique arbitrary delegation ID.
 This delegation ID will be associated to the remote resource.
@@
 (See Section \ref{gliteundelegate} for further details!) The delegation command has the following syntax:
-\begin{shaded}\verb#$ glitedelegate <delegation ID> <service URL>#\end{shaded}
+\begin{shaded}\verb#$ glitedelegate <delegation ID> <delegation service URL>#\end{shaded}
 For example:
@@
 \label{glitesub}
-In the previous section it was shown how to create delegation.
-(See Section \ref{glitedelegate}!) This function, the job submission, is the only one which needs the delegation ID.\par The next step is to prepare the necessary input files and the job description!\par
-The glitesub command syntax is the following:
+The current command only supports direct job submission that is the user must know service endpoints.
+These kind of endpoints usually stored in LDAP databases.
+Ordinary ldapsearch can be used to find such an endpoint.
+Below an example:\par
+... ldap parancs jön ide ...\par
+The \textbf{glitesub} command implements the actual job submission in three phases: registration of a job, uploading the local files to the input sandbox URL received during the registration and starting the job.\par
+The \textbf{glitesub} command requires a valid delegation.
+See Section \ref{glitedelegate} how to create one.
+By the way this is the only command which needs the delegation ID.\par The next step is to prepare the necessary input files and the job description!\par
+The glitesub command syntax is the following (please note that the service URL, job description, info file are mandatory and can only be given in the specified order):
 \begin{shaded}\verb#$ glitesub -D <delegation ID> <service URL> <job description> <info file>#\end{shaded}
-Here the delegation ID is the previously registered identification.
+Here the delegation ID is the identification registered previously.
 The service URL will be different than in the delegation example above, because this is not the URL of the delegation, but the execution service.
 The job description must be in JDL format like in the example below.
-The info file stores information about job submitted.
-It should either be a non-existing file or it is overwritten.
-This file contains the job ID, the input sandbox URL, the output sandbox URL and the service URL.
-The submission command will take care of uploading the necessary input files up to the execution node.\par
-It is important that the \textit{VirtualOrganisation} and \textit{QueueName} attributes should be the same as your virtual organisation name and the corresponding queue name on the server side.\par
+The info file stores information about the job submitted.
+It should either be a non-existing file otherwise it is overwritten.
+This file contains the job ID, the input sandbox URL, the output sandbox URL and the service URL.\par
+Below I present a JDL description example.
+Note that the \textit{VirtualOrganisation} and \textit{QueueName} attributes should be the same as your virtual organisation name and the corresponding queue name on the server side.\par
 
-An example JDL description:\\
 \begin{minipage}{\textwidth}
@@
 \end{minipage}
+
 An example how to use this command:
@@
 \verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/CREAM2 description.jdl job.info#\end{shaded}
+... mit csinál a parancs sikeres és sikertelen futás esetén... \par
 \subsection{glitestat}
 \label{glitestat}
-The \textbf{glitestat} tool extracts job information about a CREAM2 job identified by the information file. This file made by command \textbf{glitesub} is the only argument needed here.\par
+The \textbf{glitestat} command contacts to the CREAM2 service on the service URL and queries job status of a job specified by the job ID given in the info file.\par
 Usage:
@@
 \begin{shaded}\verb#glitestat job.info#\end{shaded}
-The possible responses and their meanings are the following:
+The glitesub command implements a job status mapping between gLite and ARC job status semantics. The possible responses and their meanings are the following:
 \begin{itemize}
-\item ACCEPTING - The job submission is comleted but the job is not yet scheduled
-\item SUBMITTING - Scheduling in progress
-\item INLRMS:Q - The job is already at the local resource manager system and it is queued
-\item INLRMS:R - The job is waiting at the local resource manager system for running
-\item INLRMS:S - The job is actually running
-\item KILLED - The job was terminated
-\item FINISHED - The job has finished
-\item FAILED - The job had some failure
-\item FAILES - The job had some failure at LRMS level
-\item EXECUTED - The job has been finished and there is no stating information available
+\item ACCEPTING - the job submission is completed but the job is not yet scheduled
+\item SUBMITTING - scheduling in progress
+\item INLRMS:Q - the job is already at the local resource manager system and it is queued
+\item INLRMS:R - the job is waiting at the local resource manager system for running
+\item INLRMS:S - the job is running
+\item KILLED - the job was terminated
+\item FINISHED - the job has finished
+\item FAILED - the job had some failure
+\item FAILES - the job had some failure at LRMS level
+\item EXECUTED - the job has been finished and there is no state information available
 \end{itemize}
@@
 \label{glitekill}
-The \textbf{glitekill} command can be used for killing a remote job.
-The only necessary argument is the information file made by \textbf{glitesub}.
-(See Section \ref{glitesub} for further details!) This command initiates stopping the job.
+The \textbf{glitekill} command can be used for killing a remote job running in the underlying batch system.
+The only necessary argument needed is the information file made by \textbf{glitesub}.
+(See Section \ref{glitesub} for further details!) This command initiates stopping the job in the batch system.
 You can check the effect by using \textbf{glitestat}.
@@
 If you intentionally kill a job on the computing node, the job related stuff might also remain as garbage in the remote queue.
-This command also removes the job from the gLite server queue.
+This command also removes the job from the underlying batch system on the gLite computing element.
 It is important to note that this command removes the local information file as well.
-It is very useful to run this command after killing a job.
+It might be useful to run this command after killing a job.
 (See Section \ref{glitekill}!)\par
@@
 \label{gliteundelegate}
-After your work is complete it is possible to unregister and delete the delegation entry from the remote site by using command \textbf{gliteundelegate}.
-You can also use it, if you want to register a delegation ID but are unsure whether it is used or not.
-Put the delegation service URL into the argument!\par
+After the work is complete it is possible to unregister and delete the delegation entry from the remote site by using command \textbf{gliteundelegate}.
+It can also be used when one needs to register a delegation ID but are unsure whether it is already used or not.
+Put the delegation delegation service URL into the argument!\par
 Usage:
-\begin{shaded}\verb#$ gliteundelegate <delegation ID> <service URL>#\end{shaded}
+\begin{shaded}\verb#$ gliteundelegate <delegation ID> <delegation service URL>#\end{shaded}
 For example:
@@
 \verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}
-\section{Developers notes}
-\label{Developers notes}
-This section describes how to use the library on your own client.
-The base class of the gLite client is written as an ARC Client Component (ACC) but also can be used as a standalone library.
-There is nothing more to be done just to link the \textbf{CREAMClient}, \textbf{arcclient} and \textbf{arccommon} libraries installed with the ARC1 sofware and to create an instance of the \textit{CREAMClient} class.\par
-This client object has a very easy-to-use and intuitive interface.
-It uses the VOMS proxy certificate to build the secure channel to the server and to sign that of the server's.
-Actually the client looks for this certificate just at its default location.
-If you want to store it somewhere else, you have to modify the source code itself.\par
-The different functions throw \textit{CREAMClientError} exceptions that should be caught in the application.\par
-The following sections present the set of functions provided by the library.
-\subsection{CREAMClient(Arc::URL, Arc::MCCConfig)}
-The constructor of the class performs the necessary initialization work.
-It receives two arguments: the service URL and the ARC message chain component configuration file to establish the communication channel to the server.
-These pieces of information are indispensable to create your client and to communicate with the remote site.
-\subsection{setDelegationId(std::string)}
-Besides the constructor this is the other generic-purpose function that can be used in CREAM2 client applications.
-This function sets the previously registered and referred delegation ID on the client.
-The function is very simple: just set a private variable and returns with no value.
-\subsection{createDelegation(std::string)}
-The \textbf{createDelegation} function is about to perform the whole delegation registration process.
-It sends a \textit{getProxyReq} message having the requested delegation ID, signs the received certificate and sends it back in a \textit{putProxy} SOAP message to the server.
-These three steps constitute the delegation registration process.
-As almost every funtion, this one has no return value either.
-If there are any problems during the communication, either locally in the channel or at the remote site, the function throws \textit{CREAMClientError} described previously.
-If there are no exceptions thrown, then the command is considered to have been successfully completed.
-\subsection{destroyDelegation(std::string)}
-This method has the reverse functionality of the \textbf{createDelegation} function. It sends the CREAM2 \textit{destroy} message to the remote server.
-\subsection{submit(std::string)}
-The \textbf{submit} function is the most complex part of the class.
-It translates the job description received as the argument, registers the job with a \textit{JobRegisterRequest} message, uploads the locally stored files, if necessary, then enables the job execution on the server by sending a \textit{JobStartRequest} message.
-Finally it returns with a \textbf{creamJobInfo} object that contains a \textit{jobId} (job indetifier), a \textit{creamURL} (as the service URL), an \textit{ISB} (reference to the Input Sandbox) and an \textit{OSB} (Output Sandbox) member to describe the registered job.
-In the \textbf{glitesub} command (See Section \ref{glitesub} for details) these pieces of information are stored in the information file.
-\subsection{stat(std::string)}
-This method queries the job status from the gLite CREAM2 server.
-The return value is the job status translated to the ARC terminology.
-The possible values are presented in the section of the \textbf{glitestat} command line tool.
-(See Section \ref{glitestat}!)
-\subsection{cancel(std::string)}
-The \textbf{cancel} function sends a \textit{JobCancelRequest} SOAP message to the server and handles the emerging exceptions.
-It can be used for canceling a remotely registered and possibly running job.
-\subsection{purge(std::string)}
-The \textbf{purge} function sends a \textit{JobPurgeRequest} message to server and throws an exception in case of any emerging problems.
-\subsection{An example gLite client}
-Finally, here is an example code to present how easy and simple is to write a new client.
+\subsection{Skeleton of a temporary example gLite client}
+Finally, here is an example code to present how easy and simple it is to write a new client using the CREAMClient class (See section \ref{CREAMClient}!).
 This code sample requests the job status of a previously submitted remote job, and writes it to the standard output.
@@
 It is of course mandatory to define \textit{SERVICE\_URL} and \textit{JOBID} strings to their real values to reach the proper functionality.
-\section{Outstanding issues}
-\label{Outstanding issues}
+\section{Current limitations}
+\label{Current limitations}
 Even though the gateway library code is ready for production use, there are a couple of issues to be handled in the upcoming development.
+\subsection{Unsupported CREAM features}
 \subsection{Full ARC1 client integration}
@@
 A built-in xRSL${\rightarrow}$JDL translator will also enable to use purely xRSL job description on ARC1 user interfaces.
+\appendix
+\section{Installation from Debian binary packages}
+\label{Debian install}
+This section contains instructions how to install the gateway libraries and glite* command line tools on Debian Linux.
+\begin{enumerate}
+\item Put the Debian repository into the APT source files (/etc/apt/sources.list). This repository in Luebeck is maintained as long as the KnowARC procject continues.
+\begin{shaded}\verb#$ echo "deb http://pc02.inb.uni-luebeck.de:8080/~moeller/debian/stable ./" \#\\*
+\verb#        >> /etc/apt/sources.list #\end{shaded}
+\item Update the package list and install the necessary dependencies:
+\begin{shaded}\verb#$ aptitude update#\\*
+\verb#$ aptitude install voms globus globus-dev globus-doc#\end{shaded}
+\item Add the Globus libraries to the system path:
+\begin{shaded}\verb#$ echo "/opt/globus/lib" > /etc/ld.so.conf.d/globus.conf#\\*
+\verb#$ ldconfig#\end{shaded}
+\item Install ARC1 client.
+\begin{shaded}\verb#$ aptitude install nordugrid-arc1-client#\end{shaded}
+\item Create the vomses file:
+\begin{shaded}\verb#$ mkdir -p /opt/glite/etc/#\end{shaded}
+and copy these lines into the vomses file located at /opt/glite/etc/vomses
+\begin{footnotesize}
+\begin{shaded}
+\verb#"gin.ggf.org" "kuiken.nikhef.nl" "15050" "/O=dutchgrid/O=hosts/OU=nikhef.nl/CN=kuiken.nikhef.nl" "gin.ggf.org"#\\*
+\verb#"knowarc.eu" "arthur.hep.lu.se" "15001" "/O=Grid/O=NorduGrid/CN=host/arthur.hep.lu.se" "knowarc.eu"#\\*
+\verb#"atlas" "voms.cern.ch" "15001" "/DC=ch/DC=cern/OU=computers/CN=voms.cern.ch" "atlas"#\\*
+\verb#"nordugrid.org" "voms.uninett.no" "15015" "/O=Grid/O=NorduGrid/CN=host/voms.ndgf.org" "nordugrid.org"#\\*
+\end{shaded}
+\end{footnotesize}
+\item Create \$HOME/.globus directory and move the user certificate-key pair into there.
+\end{enumerate}
+\bibliography{grid}
 \end{document}

### 2008-08-12 doc/glite_gateway/glite_gateway.tex
Added citations

@@
 %%%%---- A subtitle, if necessary ----
-%{\centering \textit{\large Paper subtitle}\large \par}
+{\centering \textit{\large Technical description and user manual}\large \par}
 \vspace*{1.5cm}
@@
 \label{Introduction}
-The purpose of this document is to describe this gLite specific plugin of the ARCLIB (Hivatkozni: jövőbeli Deliverable) to present the related command line utilities built on top of these libraries.
+The purpose of this document is to describe this gLite\cite{glite} specific plugin of the ARCLIB (...Hivatkozni: jövőbeli Deliverable) to present the related command line utilities built on top of these libraries.
 The gateway library together with the tools enables seemless interoparability with gLite computing elements.
@@
 Interoperability with gLite storage element is provided by an other set of ARC library.
-This data library provides native interfaces to GridFTP SRM which is the main bulding blocks of a gLite storage element.
+This data library provides native interfaces to GridFTP\cite{gridftp} SRM which is the main bulding blocks of a gLite storage element.
 A storage interopability is out of scope of this library.
@@
 \end{itemize}
-The latter solution was preferred by ARC1 developers, the gLite gateway library module has been implemented in the ARCLIB and the gateway functionality will be transparently offered via the usual Nordugrid command line tools.
+The latter solution was preferred by ARC1 developers, the gLite gateway library module has been implemented in the ARCLIB and the gateway functionality will be transparently offered via the usual NorduGrid command line tools.
 An overview of the preferred soulution is presented in figure \ref{fig:Chart 1}.\par
@@
 \caption{\label{fig:Chart 1}This chart illustrates how the ARC1$\rightarrow$gLite gateway library module based approach works.
-The user will be able to express his requirements in the job description document using any of the xRSL, JDL or JSDL languages.
+The user will be able to express his requirements in the job description document using any of the xRSL\cite{xrsl}, JDL or JSDL\cite{jsdl} languages.
 The ARC client library will perform an automatic transformation.
@@
 CREAM2 is available on several preproduction and testing grid sites.\par
-Nevertheless the KnowARC project chose the CREAM version 2\cite{cream} as the gLite computing element interface.\par ... Hivatkozás a cikkre, mint CREAM-re...
+Nevertheless the KnowARC project chose the CREAM version 2\cite{cream} as the gLite computing element interface.\par
 In particular, the gateway library was developed against version glite-ce-cream\_R\_1\_8\_3\_0\footnote{\url{http://grid.pd.infn.it/cream/field.php?n=Main.ReleaseNotes}}.\par
@@
 The ARCLIB gLite plugin have only one dependency.
-GridFTP components of the Globus Toolkit are needed to enable data staging functionality to those CREAM2 computing elements which offer input and output sandbox over a GridFTP enabled storage.
+GridFTP components of the Globus Toolkit\cite{globus} are needed to enable data staging functionality to those CREAM2 computing elements which offer input and output sandbox over a GridFTP enabled storage.
 Unfortunately the gateway library only works together with a strongly patched Globus verion.
-The Nordugrid patched Globus can be dowloaded from the Nordugrid download area\footnote{\url{http://download.nordugrid.org/software/globus/}}.
-Further information about the Nordugrid Globus patches can be read on the Nordugrid page\footnote{\url{http://wiki.nordugrid.org/index.php/Globus\_Libraries}}.
-It is worth to mention that the user will need a VOMS proxy certificate in order to use the command line tools against a CREAM2 service.
+The NorduGrid patched Globus can be dowloaded from the NorduGrid download area\footnote{\url{http://download.nordugrid.org/software/globus/}}.
+Further information about the NorduGrid Globus patches can be read on the NorduGrid page\footnote{\url{http://wiki.nordugrid.org/index.php/Globus\_Libraries}}.
+It is worth to mention that the user will need a VOMS\cite{voms} proxy certificate in order to use the command line tools against a CREAM2 service.
 This VOMS proxy can be generated by using the VOMS client tools.
@@
 The ARC gLite gateway plugin library and the temporary command line utilities are incorporated into ARC1 releases.
-The source code is available from the ARC1 SVN repository\footnote{\url{http://svn.nordugrid.org/repos/nordugrid/arc1/}} and binary packages are provided at the Nordugrid download area.
+The source code is available from the ARC1 SVN repository\footnote{\url{http://svn.nordugrid.org/repos/nordugrid/arc1/}} and binary packages are provided at the NorduGrid download area.
 The build and installation instructions are given in the README file distributed together with the software.
@@
 1 bekezdés arról, hogy SOAP a hed(Hivatkozni!!!) SOAP framework-jét használja 'fully integrated'...
-az előző deliverable-ben a kommunikáció gsoap-ot használt, ez az egyik nagy előrelépés, hogy ezt lecseréltük...
+az előző deliverable-ben a kommunikáció gSOAP-ot használt, ez az egyik nagy előrelépés, hogy ezt lecseréltük...\cite{gsoap}...
 \subsubsection*{CREAMClient(Arc::URL, Arc::MCCConfig)}
@@
 \label{Users guide}
-The well known ARC commands (ngsub, ngget, ...
-See: The Nordugrid/ARC User Guide\cite{userguide}!) will provide seemless access and integration with gLite computing elements.
+The well known ARC commands (ngsub, ngget, See: The NorduGrid/ARC User Guide\cite{userguide}!) will provide seemless access and integration with gLite computing elements.
 Full transparency will be achieved through the gLite gateway ARCLIB plugin.
 An ordinary user will use the same set of commands regardless of what type of CE to access.
-The user will be able to specify grid jobs described both in the gLite (JDL) and the Nordugrid (JSDL, XRSL) job description languages.
-...
-+++Kipofozni.
-Hivatkozások...
+The user will be able to specify grid jobs described both in the gLite (JDL) and the NorduGrid (JSDL, xRSL) job description languages.
 The client will perform automatic conversion between the different job description languages.\par
-Unfortunately the interoperability features offered by the gLite gateway library are not yet integrated into the Nordugrid commands.
+Unfortunately the interoperability features offered by the gLite gateway library are not yet integrated into the NorduGrid commands.
 Also the current version of ARCLIB does not offer the full spectrum of language transformations.
@@
 Below an example:\par
-... ldap parancs jön ide ...\par
+\begin{shaded}
+\verb#$ ldapsearch -x -h lxbra2305.cern.ch -p 2170 -b mds-vo-name=local,o=grid \#\\*
+\verb#        '(GlueCEUniqueID=*cream*)' | grep GlueCEInfoContactString | \#\\*
+\verb#        awk '{print $2}' | sort | uniq#
+\end{shaded}
 The \textbf{glitesub} command implements the actual job submission in three phases: registration of a job, uploading the local files to the input sandbox URL received during the registration and starting the job.\par
@@
 \subsection{xRSL${\rightarrow}$JDL translation and related problems}
-It is possible to express jobs in XRSL which have no equivalent in JDL.
-Speciﬁcally, in XRSL it is possible for input files downloaded from storage servers to have their names changed.
+It is possible to express jobs in xRSL which have no equivalent in JDL.
+Speciﬁcally, in xRSL it is possible for input files downloaded from storage servers to have their names changed.
 This is not possible in JDL.
-So a user might in XRSL ask for this:
+So a user might in xRSL ask for this:
 \begin{shaded}
@@
 It only allows the file to be downloaded as datafile1.txt.
-This could be solved by a wrapper script that renames files according to the XRSL speciﬁcation.
+This could be solved by a wrapper script that renames files according to the xRSL speciﬁcation.
 This does not solve our problem completely however.
-For example if a user wants to concatenate the stdout from two jobs, he would write something like the following in XRSL:
+For example if a user wants to concatenate the stdout from two jobs, he would write something like the following in xRSL:
 \begin{shaded}

### 2008-08-12 doc/glite_gateway/glite_gateway.tex
Open issues updated

@@
 
-\def\documenttitle{NOT\_READY\_The glite gateway plugin of ARCLIB and related command line tools}
+\def\documenttitle{The glite gateway plugins of ARCLIB and related command line tools}
 \captionsetup{margin=20pt,font={footnotesize, it},labelfont=bf}
@@
 
-{\raggedleft NORDUGRID-TECH-22\par}
+{\raggedleft DRAFT\_NORDUGRID-TECH-22\par}
 
@@
 \label{Introduction}
-The purpose of this document is to describe this gLite\cite{glite} specific plugin of the ARCLIB (...Hivatkozni: jövőbeli Deliverable) to present the related command line utilities built on top of these libraries.
+The purpose of this document is to describe this gLite\cite{glite} specific plugins of the ARCLIB (...Hivatkozni: jövőbeli Deliverable) to present the related command line utilities built on top of these libraries.
 The gateway library together with the tools enables seemless interoparability with gLite computing elements.
@@
 \end{itemize}
-The latter solution was preferred by ARC1 developers, the gLite gateway library module has been implemented in the ARCLIB and the gateway functionality will be transparently offered via the usual NorduGrid command line tools.
-An overview of the preferred soulution is presented in figure \ref{fig:Chart 1}.\par
-\begin{figure}[ht]
-\centering{\resizebox{\textwidth}{!}{\includegraphics{D3-2-1-figure.png}}
-\caption{\label{fig:Chart 1}This chart illustrates how the ARC1$\rightarrow$gLite gateway library module based approach works.
+The latter solution was preferred by ARC1 developers, the gLite gateway library module has been implemented in the ARCLIB and the gateway functionality will be transparently offered via the usual NorduGrid command line tools.\par
+An overview of the preferred soulution is presented in figure \ref{fig:Chart 1}.
 The user will be able to express his requirements in the job description document using any of the xRSL\cite{xrsl}, JDL or JSDL\cite{jsdl} languages.
@@
 The library module gets job information through CREAM2 and sends management commands, like kill a job or undelegate credentials.
-Note, that not all the functionality described on the overview figure is implemented yet, for the current limitations see section \ref{Current limitations}!}}
+Note, that not all the functionality described on the overview figure is implemented yet, for the current limitations see section \ref{Current limitations}!\par
+\begin{figure}[ht]
+\centering{\resizebox{\textwidth}{!}{\includegraphics{D3-2-1-figure.png}}
+\caption{\label{fig:Chart 1}This chart illustrates how the ARC1$\rightarrow$gLite gateway library module based approach works.}}
 \end{figure}
@@
 In the near future CREAM will offer yet another interface which will be based on the BES/JSDL/GLUE OGF standards. This CREAM BES interface will coexists with the CREAM2 too.\par
-\section{External dependencies and installation instructions}
-\label{External dependencies and installation instructions}
-The ARCLIB gLite plugin have only one dependency.
+\section{External dependencies and installation instruction}
+\label{External dependencies and installation instruction}
+The ARCLIB gLite plugins have only one dependency.
 GridFTP components of the Globus Toolkit\cite{globus} are needed to enable data staging functionality to those CREAM2 computing elements which offer input and output sandbox over a GridFTP enabled storage.
@@
 It is only the VOMS enabled proxy certificate that is needed here, appart from this no VOMS component needed.\par
-The ARC gLite gateway plugin library and the temporary command line utilities are incorporated into ARC1 releases.
+The ARC gLite gateway plugins library and the temporary command line utilities are incorporated into ARC1 releases.
 The source code is available from the ARC1 SVN repository\footnote{\url{http://svn.nordugrid.org/repos/nordugrid/arc1/}} and binary packages are provided at the NorduGrid download area.
 The build and installation instructions are given in the README file distributed together with the software.
-A step-by-step guide for Debian binary installation can be found in the appendix \ref{Debian install} Installation from Debian binary packages.
-\section{ARCLIB gLite gateway plugin}
-\label{ARCLIB gLite gateway plugin}
-The ARCLIB gLite gateway plugin constist of the wrapper classes, the job description handling facility and the gLite specific CREAMClient base class.
+A step-by-step guide for Debian binary installation can be found in the appendix \ref{Debian install} \textit{Installation from Debian binary packages}.
+\section{ARCLIB gLite gateway plugins}
+\label{ARCLIB gLite gateway plugins}
+The ARCLIB gLite gateway plugins constist of the wrapper classes, the job description handling facility and the gLite specific CREAMClient base class.
 The wrapper classes coveys the function provided by the CREAMClient to ARCLIB.
@@
 The \textbf{purge} function sends a \textit{JobPurgeRequest} message to server and throws an exception in case of any emerging problems.
-\section{ARC command line tools for gLite CE}
+\section{ARC command line tools for gLite CREAM2 CE}
 \label{Users guide}
 The well known ARC commands (ngsub, ngget, See: The NorduGrid/ARC User Guide\cite{userguide}!) will provide seemless access and integration with gLite computing elements.
-Full transparency will be achieved through the gLite gateway ARCLIB plugin.
+Full transparency will be achieved through the gLite gateway ARCLIB plugins.
 An ordinary user will use the same set of commands regardless of what type of CE to access.
@@
 It is of course mandatory to define \textit{SERVICE\_URL} and \textit{JOBID} strings to their real values to reach the proper functionality.
-\section{Current limitations}
-\label{Current limitations}
-Even though the gateway library code is ready for production use, there are a couple of issues to be handled in the upcoming development.
-\subsection{Unsupported CREAM features}
-\subsection{Full ARC1 client integration}
-Full ARC1 client integration basically means that the previously presented commands will disappear and their functionality will be available in the standard ARC1 client toolkit.
-However there are two dependencies to be met to achieve this stage:
-\begin{itemize}
-\item On one hand ARC1 broker along with the BDII-capable target object should be finalized.
-\item On the other hand a conflicting Globus Toolkit OpenSSL and the standard OpenSSL issue is to be investigated.
-\end{itemize}
-The latter issue was revealed when the VOMS proxy certificate delegation method was implemented in the gateway library.
-Since in the ARC1 client-side library there is a built-in Message Chaining Component (MCC) which already uses Globus Toolkit OpenSSL functions to enable establishing secure communication.
-Linking glite* tools directly against the standard OpenSSL libs caused congesting SSL functions which led to the breakdown of the whole library.
-The solution had been not linking the standard OpenSSL libs, just the patched OpenSSL in the Globus Toolkit.
-\subsection{xRSL${\rightarrow}$JDL translation and related problems}
-It is possible to express jobs in xRSL which have no equivalent in JDL.
+\section{Open issues}
+In order to achieve the full scale transparent interoperability width the gLite computing element transparently offered through the NorduGrid client a couple of issues still need to be resolved.
+Most of these are related to the incomplete integration of the glite gateway library into the ARCLIB framework, even though the gateway library code is ready for production use.
+There are also some conceptual chanlenges due to the incompatible nature of the two systems which require development of some workarounds.
+\subsection{Current limitations}
+Full ARC1 client integration would mean that the temporary commands developed on top of the CREAMClient\ref{CREAMClient} presented in the previous section will be obsoleted by the ordinary ARC commands built on top of the ARCLIB.
+To achieve this the Wrapper class integration of the gLite gateway plugins library should finalized and exposed to extended testing.\par
+Some of the Wrapper class methods and data structures should be extracted from the information available in the CREAMClient class.
+In particular, the detailed job representation needed for job monitoring is an open area.\par
+The CREAMClient class will be extended to support operations such as SUSPEND available both in the CREAM2 interface and the ARCLIB's jobcontrol component.\par
+Due to the incomplete ARCLIB integration currently only direct job submission to a manually selected CE is possible.
+The ARCLIB integration will combine the existing resource discovery (REF ...
+a Wrapperre) and job submission functionality (REF ...
+a Wrapperre)  of the glite gateway plugins with the brokering module of ARCLIB, which is  currently under development.
+The combination of these ARCLIB modules will provide intelligent brokered job submission over glite and nordugrid computing element.\par
+For a fully transparent integration the automatic translation among JSDL/XRSL and JDL should also be provided.
+A user should be able to specify his/her job request in any of these languages and the gateway do the necessary translations.
+The current language translator module currently is not capable generating JDL output.
+Therefore, the user still needs to use JDL if he/she wants to submit jobs to CREAM2.\par
+Further investigation is needed how a CREAM2 service handles data staging and whether the data movement related ARC libraries can be used to manage data staging in the client side or development of special data staging wrapper scripts to be executed on the computing element are needed.
+A set of temporary workaround command line clients had to be developed (the glite* commands [...
+REF]) due to a discovered OpenSSL version conflict.
+The conflict appeared when we tried to integrate the CREAMClient gateway library with the rest of the ARCLIB.
+A tedious and time consuming debugging process took place.
+The debugging revealed the following:  A VOMS proxy certificate delegation method was implemented in the CREAMClient gateway library.
+During this process it was needed to sign a certificate and the original OpenSSL was used for this purpose.
+In the used communication library taken from the ARC1 framework there is a built-in Message Chain Component (MCC) which already utilizes the Globus Toolkit OpenSSL functions to enable establishing secure communication.
+Linking the CREAMClient library directly against the standard OpenSSL libs caused congesting SSL functions.
+In order to resolve the OpenSSL conflict and to make possible the complete ARCLIB integration  a globus patch has been being worked on.
+Once the OpenSSL conflict is resolved the ARCLIB integration will be completed this way obsoleting the glite* temporary commands.\par
+\subsection{Conceptional problems}
+During the development of the gateway library a few conceptual incompatibilities were found.
+These are the typical cases when one of the middlewares offers richer capabilities not available in the other one.
+These conceptual gaps can only be bridged by developing customized case by case workarounds, nevertheless  the interoperation always would result in some information loss in these cases.\par
+Below we collected the most important challenges for which we still have to find some solution:
+\subsubsection*{Expressing file staging in job description}
+XRSL  offers reacher capability to describe input and output date of a job than the JDL.
 Speciﬁcally, in xRSL it is possible for input files downloaded from storage servers to have their names changed.
@@
 This cannot be solved just using the renaming method, because both input files would be retrieved before renaming is done and they would therefore overwrite each other.\par
-The solution is to recognize jobs of this type and simply barring them from using gLite resources or to let such jobs fail with an error message explaining that the user has to express their job in a different way.
-\subsection{RunTime Environments related issues}
-Many grid jobs rely on software being available at the execution location. The requirement as well as the initialization 
-of such software is expressed through the use of Run-Time Environments (RTEs). While the library is capable of 
-translating the RTE request itself it is not capable of translating the name of the RTE. No standardized cross Grid 
-nomenclature has been adopted, nor are the authors aware of any such activity. A temporary solution would be for each 
-VO to have a remotely accessible list of RTEs and their names in each job description language. This could then be used for translating jobs.
-\section{Conclusions}
-\label{Conclusions}
-The document has described an ARC1 library extension that allows ARC1 grid users to access gLite CREAM2 computing resources.
-The current version of the code allows direct access to resources but as the new ARC1 brokering solution comes out the gateway functionality will be integrated fully into the ARC1 client code.
-The same version will also be able to access the gLite information system to extract CREAM2 resource information.\par
-A built-in xRSL${\rightarrow}$JDL translator will also enable to use purely xRSL job description on ARC1 user interfaces.
+A possible  solution is to recognize jobs of this type and simply barring them from using gLite resources or to let such jobs fail with an error message explaining that the user has to express their job in a different way.
+\subsubsection*{RunTime Environments related issues}
+Many grid jobs rely on software being available at the execution location.
+The requirement as well as the initialization of such software is expressed through the use of Run-Time Environments (RTEs).
+While the library is capable of translating the RTE request itself it is not capable of translating the name of the RTE.
+No standardized cross Grid nomenclature has been adopted, nor are the authors aware of any such activity.
+A temporary solution would be to have a remotely accessible list of RTEs and their names in each job description language.
+This could then be used for translating jobs.
 \appendix

### 2008-08-12 doc/backend_interface/Backends-arc1.tex
backends docs update

@@
 system (LRMS) to the ARC middleware. It also intends to serve as a
-reference manual describing existing batch system interfaces. Note
-that the current solution described in this manual is not a real interface,
-it is too diffuse to be that.
+reference manual describing existing batch system interfaces.  Note that
+certain details of the interface which are not yet finalized are not described,
+but will be included in a future revision of this document.
 
@@
 \begin{enumerate}
-\item to allow the GridManager (GM), which is resided in A-REX, to manipulate
+\item to allow the GridManager (GM), which is resided in A-REX, to control
 jobs in the LRMS including job submit, cancel operations etc.
@@
 \hline
-Name & from the configuration \\
+Name & `name'' configuration option \\
 \hline
@@
 \hline
-MappingQueue & from the configuration \\
-\hline
-SchedulingPolicy & if MAUI is used, set to \texttt{fairshare} \\
+MappingQueue & ``lrms\_queue'' configuration option \\
+\hline
+SchedulingPolicy & if MAUI is used, it is set to \texttt{fairshare} \\
 \hline

### 2008-08-13 doc/client/arcdata.tex
Files from ARC0 imported and rudimentarily changed to provide a skeleton for the corresponding documentation for ARC1.

\subsection{arcls}\label{sec:arcls}
The \texttt{arcls} \index{arcls}\index{commands:arcls} is a simple
utility that allows to list contents and view some attributes of
objects of a specified (by an URL) remote directory.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcls [options] $<$URL$>$}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
    Options:&&\\
    \texttt{-h} && short help\\
    \texttt{-v} && print version information\\
    \texttt{-d} & \textit{debuglevel} &debug level: 0 = some, 1 = more, 2 = a lot \\
    \texttt{-l} &  & detailed listing\\
    \texttt{-L} &  & detailed listing including URLs from which file can 
    be downloaded and\\ & & temporary cached locations\\
    Arguments:&&\\
    \texttt{URL} && file or directory URL\\
\end{longtable}

This tool is very convenient not only because it allows to list files
at a Storage Element or records in an indexing service, but also
because it can give a quick overview of a job's working directory,
which is explicitly given by job ID.

Usage examples can be as follows:

\begin{verbatim}
    ngls rls://rc.host:38203/logical_file_name
    ngls -l gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535
    ngls -L srm://grid.uio.no:58000/srm/managerv1/johndoe/log2
\end{verbatim}

Examples of URLs accepted by this tool can be found in
Section~\ref{sec:urls}, though \texttt{arcls} won't be able to list a
directory at an HTTP server, as they normally do not return directory
listings.

\subsection{arccp}\label{sec:arccp}

The \texttt{arccp}\footnote{In ARC $\leq$ 0.5.28 was called
  \texttt{ngcopy}} \index{arccp}\index{commands:arccp} is a powerful
tool to copy files over the Grid. It is a part of the Grid Manager,
but can be used by the User Interface as well.
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arccp [options] $<$source$>$ $<$destination$>$}
\end{shaded}
\versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
\begin{longtable}{llp{8cm}}
    Options:&&\\
    \texttt{-h} && short help\\
    \texttt{-v} && print version information\\
    \texttt{-d} & \textit{debuglevel} &debug level: from -3 = quiet to 3 = verbose\\
    \texttt{-y} & \textit{cache\_path} & path to local cache (use to put file into cache)\\
    \texttt{-Y} & \textit{cache\_data\_path} & path for cache data (if different from -y)\\
    \texttt{-p} && use passive transfer (does not work if secure is on, default if secure is not requested)\\
    \texttt{-n} && do not try to force passive transfer\\
    \texttt{-i} && show progress indicator\\
    \texttt{-u} && use secure transfer (insecure by default)\\
    \texttt{-r} & \textit{recursion\_level} & operate recursively (if possible) up to specified level (0 - no recursion)\\
    \texttt{-R} & \textit{number} & how many times to retry transfer of every file before failing\\
    \texttt{-t} & \textit{time} & timeout in seconds (default 20)\\
    \texttt{-f} && if the destination is an indexing service and not the same as the source and the destination is already registered, then the copy is normally not done. However, if this option is specified the source is assumed to be a replica of the destination created in an uncontrolled way and the copy is done like in case of replication\\
    \texttt{-T} && do not transfer file, just register it - destination must be non-existing meta-url\\
    Arguments:&&\\
    \texttt{source} && source URL\\
    \texttt{destination} && destination URL\\
\end{longtable}

This command transfers contents of a file between 2 end-points.
End-points are represented by URLs or meta-URLs. For supported
endpoints please refer to Section~\ref{sec:urls}.

\texttt{ngcp} can perform multi-stream transfers if \texttt{threads}
URL option is specified and server supports it.

Source URL can end with \verb#"/"#. In that case, the whole fileset
(directory) will be copied. Also, if the destination ends with
\verb#"/"#, it is extended with part of source URL after last
\verb#"/"#, thus allowing users to skip the destination file or
directory name if it is meant to be identical to the source.

\begin{framed}
   Since the job ID is in fact a \verb#gsiftp://# URL of the job
   top directory, you can use \texttt{ngcp} to copy files from the
   job directory at any time.
\end{framed}

Usage examples of \texttt{ngcp} are:

\begin{verbatim}
    ngcp gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535/job.out \
              file:///home/myname/job2.out
    ngcp gsiftp://aftpexp.bnl.gov;threads=10/rep/my.file \
              rc://;threads=4@grid.uio.no/lc=Collection,rc=Catalog/zebra4.f
    ngcp http://www.nordugrid.org/data/somefile gsiftp://hathi.hep.lu.se/data/
\end{verbatim}

\subsection{ngrm}\label{sec:ngrm}

The \texttt{ngrm}\footnote{In ARC $\leq$ 0.5.28 was called
  \texttt{ngremove}} \index{ngrm}\index{commands:ngrm}
command allows users to erase files at any location specified by a
valid URL.
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{ngrm [options] $<$source$>$}
\end{shaded}
\versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
\begin{longtable}{llp{8cm}}
    Options:&&\\
    \texttt{-h} && short help\\
    \texttt{-v} && print version information\\
    \texttt{-d} & \textit{debuglevel} &debug level: 0 = some, 1 = more, 2 = a lot \\
    \texttt{-c} & &continue with meta-data even if it failed to delete real file\\
    \texttt{-C} & \textit{cache\_data\_path} & store cached data\\
    Arguments:&&\\
    \texttt{source} && source URL\\
\end{longtable}

\begin{framed}
   A convenient use for \texttt{ngrm} is to erase the files in a data
   indexing catalog (RC, RLS or such), as it will not only remove the
   physical instance, but also will clean up the database record.
\end{framed}

Here is an \texttt{ngrm} example:

\begin{verbatim}
    ngrm rc://grid.uio.no/lc=Collection,rc=Catalog/badfile#\\
\end{verbatim}

\subsection{ngacl}\label{sec:ngacl}

\index{ngacl}\index{commands:ngacl}This command retrieves or modifies
access control information associated with a stored object if service
supports GridSite GACL language~\cite{gacl} for access control.
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{ngacl [options] get$|$put $<$URL$>$}
\end{shaded}
\versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
\begin{longtable}{llp{8cm}}
   Options:&&\\
    \texttt{-d, -debug} & \textit{debuglevel} &debug level: 0 = some, 1 = more, 2 = a lot \\
    \texttt{-v} && print version information\\
    \texttt{-h} && short help\\
   Arguments:&&\\
    \texttt{get} &\textit{URL}& get Grid ACL for the object\\
    \texttt{put} &\textit{URL}& set Grid ACL for the object\\
    \texttt{URL} && object URL; curently only gsiftp and sse URLs are supported\\
\end{longtable}

ACL document (an XML file) is printed to standard output when
\texttt{get} is requested, and is acquired from standard input when
\texttt{set} is specified\footnote{In ARC $\leq$ 0.5.28, \texttt{set}
  shoud be used instead of \texttt{put}}. Usage examples are:
\begin{verbatim}
    ngacl get gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename
    ngacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename $<$ myacl
\end{verbatim}

\subsection{ngtransfer}\label{sec:ngtransfer}
\index{ngtransfer}\index{commands:ngtransfer}The \texttt{ngtransfer}
command\footnote{In ARC $\leq$ 0.5.28, was called \texttt{ngrequest}}
initiates direct transfer of data between 2 servers (known as
\textit{third-party transfer}).  
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{ngtransfer [options] $<$destination$>$}
\end{shaded}
\versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
\begin{longtable}{llp{8cm}}
   Options:&&\\
    \texttt{-s, -source} & \textit{URL} & source file URL\\
    \texttt{-d, -debug} & \textit{debuglevel} &debug level: 0 = some, 1 = more, 2 = a lot \\
    \texttt{-v} && print version information\\
    \texttt{-h} && short help\\
   Arguments:&&\\
    \texttt{destination} && destination URL; currently only se:// and
    (gsi)ftp:// are supported\\
\end{longtable}

This command initiates file copy from multiple source instances to a
destination URL. Destination of \texttt{(gsi)ftp} type accepts only
similar kinds of sources.  Destination can also be URL of an Indexing
Service. In such a case, real destinations with suitable protocols are
chosen, when available. Requests are sent to the corresponding
services/servers to initiate file transfer from one of the sources.

\begin{framed}
   Absence of \verb#-s# option is treated as a file replication
   request. In this case, destination must be an Indexing service or
   an SRM.
\end{framed}

Following source URL types are supported: \texttt{http},
\texttt{https}, \texttt{httpg}, \texttt{ftp}, \texttt{gsiftp},
\texttt{rc}, \texttt{rls}, \texttt{se}, \texttt{srm}, \texttt{fireman}
(see Section~\ref{sec:urls} for URL details).

Example:
\begin{verbatim}
    ngtransfer -s http://www.host1.org/dat1  -s gsiftp://host2.org/dir/dat1 \
               se://se.host.org/se_service?new_file_lfn
\end{verbatim}

### 2008-08-13 doc/client/arcjob.tex
Files from ARC0 imported and rudimentarily changed to provide a skeleton for the corresponding documentation for ARC1.

\subsection{arcsub}\label{sec:arcsub}
The \texttt{arcsub}\index{arcsub}\index{commands:arcsub} command is the
most essential one, as it is used for submitting jobs to the Grid
resources. \index{submit job}. \texttt{arcsub} matches user's job
description to the information collected from the Grid, and the
optimal site is being selected for job submission. The job description
is then being forwarded to that site, in order to be submitted to the
Local Resource Management System (LRMS), which can be, e.g., PBS or
Condor or SGE etc.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcsub [options] $<$task ...$>$}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{ -c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
   \texttt{ -C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{ -g, -giisurl}&\textit{url}&URL of a central Information System server (GIIS)\\
   \texttt{ -G, -giislist}&\textit{filename}&list of GIIS URLs\\
   \texttt{ -e, -xrsl}&\textit{filename}&string describing the job to be submitted\\
   \texttt{ -f, -file}&\textit{filename}&file describing the job to be submitted\\
   \texttt{ -o, -joblist}&\textit{filename}&file where the job IDs will be stored\\
   \texttt{ ~~~~-dryrun}&&add dryrun option to the job description\\
   \texttt{ ~~~~-dumpxrsl}&&do not submit -- dump transformed job description to stdout\\
   \texttt{ -t, -timeout}&\textit{time}&timeout for queries (default 40 sec)\\
   \texttt{ -d, -debug}&\textit{debuglevel}&debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{ -x, -anonymous}&&use anonymous bind for queries (default)\\
   \texttt{ -X, -gsi}&&use GSI-GSSAPI bind for queries\\
   \texttt{ -v, -version}&&print version information\\
   \texttt{ -h, -help}&&print help page\\
   Arguments:&&\\
   \texttt{ task ...}&&strings or files describing the jobs to be submitted\\
\end{longtable}

\begin{framed}
Some text can be in frames
\end{framed}

A simple \textit{"Hello World"} job would look like

\begin{shaded}
   \uicommand{arcstat [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, -all}& & all jobs\\
   \texttt{-i, -joblist}& \textit{filename}& file containing a list of jobIDs\\
   \texttt{-c, -clusters}& & show information about sites (clusters)\\
   \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-g, -giisurl}& \textit{url} &URL of a central Information System server\\
   \texttt{-G, -giislist}&\textit{filename}&list of GIIS URLs\\
   \texttt{-q, -queues}&&show information about clusters and queues\\
   \texttt{-l, -long}& & long format (extended information)\\
   \texttt{-t, -timeout}& \textit{time}& timeout for queries (default 40 sec)\\
   \texttt{-d, -debug}& \textit{debuglevel}&debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{-x, -anonymous}& &use anonymous bind for queries (default)\\
   \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
   \texttt{-v, -version}& & print version information\\
   \texttt{-h, -help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

The arcstat command returns the status of jobs submitted to the Grid.
A job can be referred to either by the

Diferent sites may report slightly diferent job states, depending on
the installed software version. A summary of essential job
states\index{job:states} is:

\begin{longtable}{l|l|p{8cm}}
ARC 0.3, ARC 0.4 & ARC 0.5, ARC0.6 & Description\\ \hline
                   &\texttt{ACCEPTING} &job has reached the site\\
\texttt{ACCEPTED}  &\texttt{ACCEPTED}  &job submitted but not yet processed\\
\texttt{PREPARING} &\texttt{PREPARING} &input files are being retreived\\
                   &\texttt{PREPARED}  &input files are retreived\\
\texttt{SUBMITTING}&\texttt{SUBMITTING}&interaction with LRMS ongoing\\
\texttt{INLRMS: Q} &\texttt{INLRMS:Q}  &job is queued by LRMS\\
\texttt{INLRMS: R} &\texttt{INLRMS:R}  &job is running\\
                   &\texttt{INLRMS:S}  &job is suspended\\
                   &\texttt{INLRMS:E}  &job is finishing in LRMS\\
                   &\texttt{INLRMS:O}  &job is in any other LRMS state\\
\texttt{CANCELING} &\texttt{KILLING}   &job is being cancelled by user request\\
                   &\texttt{EXECUTED}  &job is completed in LRMS\\
\texttt{FINISHING} &\texttt{FINISHING} &output files are being transferred\\
\texttt{FINISHED}  &\texttt{FINISHED}  &job is finished\\
                   &\texttt{FAILED}    &job is finished with an error\\
                   &\texttt{KILLED}    &job is cancelled by user request\\
\texttt{DELETED}   &\texttt{DELETED}   &job is removed due to expiration time\\
\end{longtable}

\subsection{arccat}
\label{sec:arccat}

It is often useful to monitor the job progress by checking what it
prints on the standard output or error. The command \texttt{arccat}
\index{arccat}\index{commands:arccat} assists here, extracting the
corresponding information from the execution cluster and pasting it
on the user's screen. It works both for running tasks and for the
finished ones. This allows a user to check the output of the
finished task without actually retreiving it.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arccat [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, -all}& & all jobs\\
   \texttt{-i, -joblist}& \textit{filename} & file containing a list of job IDs\\
   \texttt{-c, -clusters}& & show information about clusters\\
   \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-o, -stdout}& & show the stdout of the job (default)\\
   \texttt{-e, -stderr}& & show the stderr of the job\\
   \texttt{-f, -follow}& & show tail of the requested file and follow its changes\\
   \texttt{-l, -gridlog}& & show the grid error log of the job\\
   \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
   \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
   \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
   \texttt{-v, -version}& & print version information\\
   \texttt{-h, -help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

The \texttt{arccat} command can return the standard output of a job
(\texttt{-o} option), the standard error (\texttt{-e} option) and the
errors reported by the Grid Manager (\texttt{-l} option).

\begin{framed}

\end{framed}

\subsection{arcget}
\label{sec:arcget}

To retrieve the results of a finished job, the \texttt{arcget}
\index{arcget}\index{commands:arcget} command should be used. It
will download the files specified by the \texttt{outputfiles}
attribute of job description to the user's computer.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcget [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, -all}& & all jobs\\
   \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
   \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
   \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{~~~~-dir} & \textit{dirname} & download directory (the job directory will be created in this directory)\\
   \texttt{-j, -usejobname}& & use the jobname instead of the digital ID as the job directory name\\
   \texttt{~~~~-keep}& & keep files on gatekeeper (do not clean)\\
   \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
   \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
   \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
   \texttt{-v, -version}& & print version information\\
   \texttt{-h, -help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

Only the results of jobs that have finished can be downloaded. The job
can be referred to either by the \texttt{jobID} that was returned by
\texttt{arcsub} at submission time, or by its name, if the job
description contained a job name attribute.

\subsection{arckill}
\label{sec:arckill}

It happens that a user may wish to cancel a job. This is done by using
the \texttt{arckill} \index{arckill}\index{commands:arckill} command. A
job can be killed amost on any stage of processing through the Grid.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arckill [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, -all}& & all jobs\\
   \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
   \texttt{-c, -clusters}& & show information about clusters\\
   \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{~~~~-keep}& & keep files on gatekeeper (do not clean)\\
   \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
   \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
   \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
   \texttt{-v, -version}& & print version information\\
   \texttt{-h, -help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

When option \verb#-j# is used, and several jobs have the same name,
\textbf{ALL such jobs will be cancelled!}.

\begin{framed}
   Job cancellation is an asynchronous process, such that it
   may take a few minutes before the job is actually cancelled.
\end{framed}

\subsection{arcresub}
\label{sec:arcresub}

Quite often it happens that a user would like to re-submit a job, but
has difficulties recovering the original job description xRSL file.
This happens when xRSL files are created by scripts on-fly, and
matching of xRSL to the job ID is not straightforward. The utility
called \texttt{arcresub}\index{arcresub}\index{commands:arcresub} helps
in such situations, allowing users to resubmit jobs known only by
their job IDs.

\begin{framed}
   Only jobs where the \verb#gmlog#\index{gmlog} attribute was
   specified in the job description can be resubmitted.
\end{framed}

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcresub [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, -all}& & all jobs\\
   \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
   \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
   \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-k, -kluster}&\verb#[-]#textem{name}&explicitly select or
   reject a specific site (cluster) as re-submission target\\
   \texttt{-K, -Klustlist}&\verb#[-]#textem{filename}&list of
   clusters to select or reject as re-submission target\\
   \texttt{-g, -giisurl}& \textit{url} &URL of a central Information System server\\
   \texttt{-G, -giislist}&\textit{filename}&list of GIIS URLs\\
   \texttt{-o, -joblist}&\textit{filename}&file where the job IDs will be stored\\
   \texttt{~~~~-dryrun}&&add dryrun option to the xRSL\\
   \texttt{~~~~-dumpxrsl}&&do not submit -- dump transformed xRSL to stdout\\
   \texttt{~~~~-keep}& & keep files on gatekeeper (do not clean)\\
   \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
   \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
   \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
   \texttt{-v, -version}& & print version information\\
   \texttt{-h, -help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

\subsection{arcclean}
\label{sec:arcclean}

If a job fails, or you are not willing to retrieve the results for
some reasons, a good practice for users is not to wait for the Grid
Manager to clean up the job leftovers, but to use \texttt{arcclean}
\index{arcclean}\index{commands:arcclean} to release the disk space
and to remove the job ID from the list of submitted jobs and from
the Information System.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcclean [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, -all}& & all jobs\\
   \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
   \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
   \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-f, -force} & & removes the job ID from the local list\\
   & &~~~even if the job is not found on the Grid\\
   \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
   \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
   \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
   \texttt{-v, -version}& & print version information\\
   \texttt{-h, -help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

Only jobs that have finished can be cleaned.

\subsection{arcrenew}
\label{sec:arcrenew}

Quite often, the user proxy expires while the job is still running (or
waiting in a queue). In case such job has to upload output files to a
Grid location (Storage Element), it will fail. By using the \texttt{arcrenew}
\index{arcrenew}\index{commands:arcrenew} command, users can upload
a new proxy to the job. This can be done while a job is still running,
thus preventing it from failing, or whithin 24 hours (or whatever is
the expiration time set by the site) after the job
end. In the latter case, the Grid Manager will attempt to finalize
the job by uploading the output files to the desired location.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcrenew [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, -all}& & all jobs\\
   \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
   \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
   \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
   \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
   \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
   \texttt{-v, -version}& & print version information\\
   \texttt{-h, -help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

Prior to using \texttt{arcrenew}, be sure to actually create the
new proxy:

\subsection{arcsync}
\label{sec:arcsync}

If you are using User Interface installations on different machines,
your local lists of submitted jobs will be different. To synchronise
these lists with the information in the Information System, use the
\texttt{arcsync} \index{arcsync}\index{commands:arcsync} command.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcsync [options]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
   \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
   \texttt{-g, -giisurl}& \textit{url} &URL of a central Information System server\\
   \texttt{-G, -giislist}&\textit{filename}&list of GIIS URLs\\
   \texttt{-f, -force}&&don't ask for confirmation\\
   \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
   \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
   \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
   \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
   \texttt{-v, -version}& & print version information\\
   \texttt{-h, -help}& & print help page\\
\end{longtable}

The ARC User Interface keeps a local list of jobs in the user's home
directory (see section~\ref{sec:joblists}). If this file is lost,
corrupt, or the user wants to recreate the file on a different
workstation, the \texttt{arcsync} command will recreate this file from
the information available in the Information System.

\subsection{Auxilliary files}\label{sec:joblists}

User Interface keeps local job lists in two files:
{\$}\verb#HOME/.ngjobs# and {\$}\verb#HOME/.arc/history#\footnote{In
  ARC $\leq$ 0.5.48, {\$}\texttt{HOME/.nghistory}}.

\textbf{{\$}\texttt{HOME/.ngjobs}} is a local list of the user's active
jobs. When a job is successfully submitted, it is added to this list,
and when it is removed from the remote site, it is removed from this
list. This list is used as the list of all active jobs when the user
specifies \verb#-a# option to the various ARC user interface commands.
For information about how to reconstruct this file in case it is
damaged or you relocate to a different workstation, see
section~\ref{sec:arcsync} about the \texttt{arcsync} command.

\textbf{{\$}\texttt{HOME/.arc/history}} contains the \texttt{jobID}s of
the jobs the user has submitted together with the time of
submission. This file is purely informational.

### 2008-08-13 doc/glite_gateway/glite_gateway.tex
Reference fix
Propagated version

@@
 
-\def\documenttitle{The glite gateway plugins of ARCLIB and related command line tools}
+\def\documenttitle{The gLite gateway plugins of ARCLIB and related command line tools}
 \captionsetup{margin=20pt,font={footnotesize, it},labelfont=bf}
@@
 
-{\raggedleft DRAFT\_NORDUGRID-TECH-22\par}
+{\raggedleft NORDUGRID-TECH-22\par}
 
@@
 \label{Introduction}
-The purpose of this document is to describe this gLite\cite{glite} specific plugins of the ARCLIB (...Hivatkozni: jövőbeli Deliverable) to present the related command line utilities built on top of these libraries.
+The purpose of this document is to describe this gLite\cite{glite} specific plugins of the ARCLIB\cite{arclib} to present the related command line utilities built on top of these libraries.
 The gateway library together with the tools enables seemless interoparability with gLite computing elements.
@@
 An overview of the preferred soulution is presented in figure \ref{fig:Chart 1}.
-The user will be able to express his requirements in the job description document using any of the xRSL\cite{xrsl}, JDL or JSDL\cite{jsdl} languages.
+The user will be able to express the job requirements in the job description document using any of the xRSL\cite{xrsl}, JDL or JSDL\cite{jsdl} languages.
 The ARC client library will perform an automatic transformation.
-The gLite gateway functionality integrated into the ARCLIB, thus into the ARC client enables the user to transparently access gLite computing resources via the usual ARC command line tools.
+The gLite gateway functionality integrated into the ARCLIB, thus into the ARC client allow the user to transparently access gLite computing resources via the usual ARC command line tools.
 The gLite modules perform authentication, credential delagation and job submission to the CREAM2 CE.
-The CREAM2 CE may perform the necessary input file download from the different storage element.
+The CREAM2 CE may perform the necessary input file download from the different storage elements.
 The library module gets job information through CREAM2 and sends management commands, like kill a job or undelegate credentials.
-Note, that not all the functionality described on the overview figure is implemented yet, for the current limitations see section \ref{Current limitations}!\par
+Note that not all the functionality described in the overview figure is implemented yet, for the current limitations see section \ref{Current limitations}!\par
 \begin{figure}[ht]
@@
 The gLite computing element (CE) offers several interfaces.
-Some of them are only suitabe for the WMS (glite resource broker) and was not meant to be used by 3rd party clients.
+Some of them are only suitabe for the WMS (gLite resource broker) and was not meant to be used by 3rd party clients.
 Fortunately gLite comes with a webservice based computing element interface, known as the CREAM CE.
 The CREAM has undergone several major revisions.
-We can only hope that some stability will soon arise.
-The current CREAM version has entered the EGEE certification process newertheless it is a full scale deplomeyment on the EGEE worldwide grid has not taken place yet.
-CREAM2 is available on several preproduction and testing grid sites.\par
-Nevertheless the KnowARC project chose the CREAM version 2\cite{cream} as the gLite computing element interface.\par
+The current CREAM version has entered the EGEE certification process nevertheless it is a full scale deplomeyment on the EGEE worldwide grid has not taken place yet.
+CREAM2 is available on several preproduction and testing grid sites, KnowARC has chosen the CREAM version 2\cite{cream} as the gLite computing element interface.\par
 In particular, the gateway library was developed against version glite-ce-cream\_R\_1\_8\_3\_0\footnote{\url{http://grid.pd.infn.it/cream/field.php?n=Main.ReleaseNotes}}.\par
@@
 \item CREAM2 contains numerous improvements compared to the earlier CREAM versions.
-\item Unlike the other gLite CE interfaces it supports direct job status queries 
+\item Unlike the other gLite CE interfaces it supports direct job status queries.
 \item It is an official interface to gLite computing elements and as EGEE evolves it is expected that more and more computing elements will be deployed with CREAM2 interface.
@@
 \end{itemize}
-In the near future CREAM will offer yet another interface which will be based on the BES/JSDL/GLUE OGF standards. This CREAM BES interface will coexists with the CREAM2 too.\par
+In the near future CREAM will offer yet another interface which will be based on the BES/JSDL/GLUE OGF standards. This CREAM BES interface will coexists with CREAM2 too.\par
 \section{External dependencies and installation instruction}
@@
 This VOMS proxy can be generated by using the VOMS client tools.
-It is only the VOMS enabled proxy certificate that is needed here, appart from this no VOMS component needed.\par
-The ARC gLite gateway plugins library and the temporary command line utilities are incorporated into ARC1 releases.
+It is only the VOMS enabled proxy certificate that is needed here, apart from this no VOMS components are needed.\par
+The ARC$\rightarrow$gLite gateway plugins library and the temporary command line utilities are incorporated into ARC1 releases.
 The source code is available from the ARC1 SVN repository\footnote{\url{http://svn.nordugrid.org/repos/nordugrid/arc1/}} and binary packages are provided at the NorduGrid download area.
 The build and installation instructions are given in the README file distributed together with the software.
-A step-by-step guide for Debian binary installation can be found in the appendix \ref{Debian install} \textit{Installation from Debian binary packages}.
+A step-by-step guide for Debian binary installation can be found in appendix \ref{Debian install} \textit{Installation from Debian binary packages}.
 \section{ARCLIB gLite gateway plugins}
@@
 The ARCLIB gLite gateway plugins constist of the wrapper classes, the job description handling facility and the gLite specific CREAMClient base class.
-The wrapper classes coveys the function provided by the CREAMClient to ARCLIB.
+The wrapper classes covey the function provided by the CREAMClient to ARCLIB.
 The job description handling framework makes a general job request representation available for the other classes.
 CREAMClient base class implements the interface towards CREAM2 computing elements.
-This client object has an easy-to-use and intuitive interface. It uses the VOMS proxy certificate to build the secure channel to the server and to sign that of the server's.
+This client object has an easy-to-use and intuitive interface. It uses the VOMS proxy certificate to build the secure channel to the server and also to sign that of the server's.
 The different functions throw \textit{CREAMClientError} exceptions that should be caught in the application.\par
@@
 \subsection{Wrapper classes}
-... mik ezek a wrapper osztályok ... milyen base class, mit csinál, mi a funckiójuk ... viszonylag részletesen ...
+\label{Wrapper classes}
+The gLite gateway library can be integrated into the ARCLIB as a plugin.
+To expose the CREAM2 functionality into the generic ARCLIB target object functions, the following three wrapper classes have been implemented: the JobController, the Submitter and the TargetRetriever.\par
+The JobControllerCREAM wrapper class provides the basic functions related to job management, such as getting job information or killing and cleaning the job on the remote CE.\par
+The SubmitterCREAM instantiates the gLite client library and calls the CREAM2-specific submitter function in there.\par
+Class TargetRetrieverCREAM performs a BDII database query and extracts the possible CREAM2 target objects.\par
+Unfortunately handling CREAM2 target objects do not perfectly fit into the philosophy of ARCLIB, there are several functions, such as proxy credential delegation or CREAM2-specific job information retrieval that behaves differently on an ARC1 and on a CREAM2 target.
+The wrapper classes have to provide solutions for these special cases.
 \subsection{Job description translators}
-... Bárhol meghívható ... objektumot példányosítani ... milyen transzformációk állank rendelkezésre ...
+\label{Job description}
+When addressing interoperability it is of paramount importance to transparently address grid job descriptions written in different job description languages by translating them automatically.
+In the gLite gateway library code this functionality is implemented in class JobDescription.
+This is a generic class that takes a job description as input in any supported formats (currently xRSL, JDL, JSDL), converts and stores it in a JSDL-like internal job description format.
+Then different operations, like getting the description in other formats or getting job-related information, can be performed using the description-independent functions of this class.\par
+JobDescription has tree back-end classes, sometimes are referred to as back-end modules or translator modules, corresponding to the three supported job description languages above.
+The back-end classes are used for parsing and also for generating the job descriptions.
+These classes do not store information and are not intended to be used directly, instead, the generic class uses them.\par
+There is a class called JobDescriptionOrderer which tries to find out the format of the input file by pattern matching; the generic class uses this piece of information to choose the appropriate back-end.
+\begin{figure}[ht]
+\centering{\resizebox{255px}{!}{\includegraphics{JobDescriptionExample-XRSLInput.jpg}}
+\caption{\label{fig:Chart 1}The figure shows how the different classes participating in the job description translation process interrelate in an example about xRSL parsing.}}
+\end{figure}
 \subsection{CREAMClient base class}
 \label{CREAMClient}
-The client object uses the VOMS proxy certificate to build the secure channel to the server and to sign that of the server's.\par
-...
-1 bekezdés arról, hogy SOAP a hed(Hivatkozni!!!) SOAP framework-jét használja 'fully integrated'...
-az előző deliverable-ben a kommunikáció gSOAP-ot használt, ez az egyik nagy előrelépés, hogy ezt lecseréltük...\cite{gsoap}...
-\subsubsection*{CREAMClient(Arc::URL, Arc::MCCConfig)}
+The client object uses the VOMS proxy certificate to build the secure channel to the server and also to sign that of the server's.\par
+To establish the SOAP connection via the secure channel, the library use the native ARC1 HED\footnote{\url{https://www.knowarc.eu/documents/Knowarci\_D1.1-1\_07.pdf}} Message Chain Component (MCC), that provides full integration with the ARC1 core and corresponding services.
+This replaced the former gSOAP-based\cite{gsoap} communication libraries, which enabled the full elimination of the gSOAP dependency.
+\subsubsection*{CREAMClient(Arc::URL, Arc::MCCConfig)\footnote{For class references see: \url{https://www.knowarc.eu/download/D1.2-2\_documentation.pdf}}}
 The constructor of the class performs the necessary initialization work.
@@
 This function sets the previously registered and referred delegation ID on the client.
-The function is very simple: just set a private variable and returns with no value.
+The function is simple: it sets a private variable and returns with no value.
 \subsubsection*{createDelegation(std::string)}
@@
 \subsubsection*{destroyDelegation(std::string)}
-This method has the reverse functionality of the \textbf{createDelegation} function. It sends the \textit{destroy} message of CREAM2 to the remote server.
+This method has the reverse functionality of function \textbf{createDelegation}. It sends the \textit{destroy} message of CREAM2 to the remote server.
 \subsubsection*{submit(std::string)}
 The \textbf{submit} function is the most complex part of the class.
-It translates the job description received as the argument, registers the job with a \textit{JobRegisterRequest} message, uploads the locally stored files, if necessary, then enables the job execution on the server by sending a \textit{JobStartRequest} message.
+It translates the job description received as the argument, registers the job with a \textit{JobRegisterRequest} message, uploads the locally stored files, if necessary, then launches the job execution on the server by sending a \textit{JobStartRequest} message.
 Finally it returns with a \textbf{creamJobInfo} object that contains a \textit{jobId} (job indetifier), a \textit{creamURL} (as the service URL), an \textit{ISB} (reference to the Input Sandbox) and an \textit{OSB} (Output Sandbox) member to describe the registered job.
@@
 Unfortunately the interoperability features offered by the gLite gateway library are not yet integrated into the NorduGrid commands.
-Also the current version of ARCLIB does not offer the full spectrum of language transformations.
-However only a set of temporary command line tools exposing the available features of library are provided.
-These temporary commands make the already existing interoparabilily features usable for ordinary users.
+Also the current version of ARCLIB does not offer the full spectrum of language transformations.\par
+However only a set of temporary command line tools exposing the available features of the library are provided.
+These temporary commands make the already existing interoparabilily features usable for ordinary users.\par
 There are seven different command line tools for using the gLite gateway functions.
 In this section these tools will be presented in detail through examples.
-The job descriptions and services shown are just simple examples, they might be different in real-life usage.\par
+The job descriptions and services shown are just simple examples, their arguments might be different in real-life usage.\par
 Every tool has a short manual page being accessible by using the \textbf{man} on-line manual reader command or some more help can be achieved by using the $<$command$>$ -? command line option.\par
@@
 This delegation ID will be associated to the remote resource.
-If you are not sure wether your favourite delegation ID is occupied or not, then try to delete it before registering it again.
+If you are not sure whether your favourite delegation ID is occupied or not, then try to delete it before registering it again.
 (See Section \ref{gliteundelegate} for further details!) The delegation command has the following syntax:
@@
 Ordinary ldapsearch can be used to find such an endpoint.
-Below an example:\par
+Below comes an example:\par
 \begin{shaded}
@@
 See Section \ref{glitedelegate} how to create one.
-By the way this is the only command which needs the delegation ID.\par The next step is to prepare the necessary input files and the job description!\par
+By the way this is the only command which needs the delegation ID.\par The next step is to prepare the necessary input files and the job description.\par
 The glitesub command syntax is the following (please note that the service URL, job description, info file are mandatory and can only be given in the specified order):
@@
 The info file stores information about the job submitted.
-It should either be a non-existing file otherwise it is overwritten.
+It should be a non-existing file otherwise it will be overwritten.
 This file contains the job ID, the input sandbox URL, the output sandbox URL and the service URL.\par
@@
 \verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/CREAM2 description.jdl job.info#\end{shaded}
-... mit csinál a parancs sikeres és sikertelen futás esetén... \par
+After successful completion the command reports the success and tells the user where the job related data are stored. Otherwise the message written to the output contains the reason of the failure.
 \subsection{glitestat}
@@
 \label{gliteundelegate}
-After the work is complete it is possible to unregister and delete the delegation entry from the remote site by using command \textbf{gliteundelegate}.
-It can also be used when one needs to register a delegation ID but are unsure whether it is already used or not.
-Put the delegation delegation service URL into the argument!\par
+After the job execution is complete it is possible to unregister and delete the delegation entry from the remote site by using command \textbf{gliteundelegate}.
+It can also be used when one needs to register a delegation ID but is unsure whether it is already used or not.
+Put the delegation service URL into the argument!\par
 Usage:
@@
 \end{framed}
-It is of course mandatory to define \textit{SERVICE\_URL} and \textit{JOBID} strings to their real values to reach the proper functionality.
+It is of course mandatory to set \textit{SERVICE\_URL} and \textit{JOBID} strings to their real values to obtain the proper functionality.
 \section{Open issues}
 In order to achieve the full scale transparent interoperability width the gLite computing element transparently offered through the NorduGrid client a couple of issues still need to be resolved.
-Most of these are related to the incomplete integration of the glite gateway library into the ARCLIB framework, even though the gateway library code is ready for production use.
-There are also some conceptual chanlenges due to the incompatible nature of the two systems which require development of some workarounds.
+Most of these are related to the incomplete integration of the gLite gateway library into the ARCLIB framework, even though the it is ready for production use.
+There are also some conceptual challenges due to the incompatible nature of the two systems, which requires workarounds.
 \subsection{Current limitations}
-Full ARC1 client integration would mean that the temporary commands developed on top of the CREAMClient\ref{CREAMClient} presented in the previous section will be obsoleted by the ordinary ARC commands built on top of the ARCLIB.
-To achieve this the Wrapper class integration of the gLite gateway plugins library should finalized and exposed to extended testing.\par
+\label{Current limitations}
+Full ARC1 client integration would mean that the temporary commands developed on top of the CREAMClient presented in the previous section \ref{CREAMClient} will be overriden by the ordinary ARC commands built on top of the ARCLIB.
+To achieve this the Wrapper class integration of the gLite gateway plugins library should be finalized and exposed to extensive testing.\par
 Some of the Wrapper class methods and data structures should be extracted from the information available in the CREAMClient class.
 In particular, the detailed job representation needed for job monitoring is an open area.\par
-The CREAMClient class will be extended to support operations such as SUSPEND available both in the CREAM2 interface and the ARCLIB's jobcontrol component.\par
+The CREAMClient class will be extended to support operations such as \textit{suspend} available both in the CREAM2 interface and the ARCLIB's jobcontrol component.\par
 Due to the incomplete ARCLIB integration currently only direct job submission to a manually selected CE is possible.
-The ARCLIB integration will combine the existing resource discovery (REF ...
-a Wrapperre) and job submission functionality (REF ...
-a Wrapperre)  of the glite gateway plugins with the brokering module of ARCLIB, which is  currently under development.
-The combination of these ARCLIB modules will provide intelligent brokered job submission over glite and nordugrid computing element.\par
-For a fully transparent integration the automatic translation among JSDL/XRSL and JDL should also be provided.
-A user should be able to specify his/her job request in any of these languages and the gateway do the necessary translations.
-The current language translator module currently is not capable generating JDL output.
-Therefore, the user still needs to use JDL if he/she wants to submit jobs to CREAM2.\par
-Further investigation is needed how a CREAM2 service handles data staging and whether the data movement related ARC libraries can be used to manage data staging in the client side or development of special data staging wrapper scripts to be executed on the computing element are needed.
-A set of temporary workaround command line clients had to be developed (the glite* commands [...
-REF]) due to a discovered OpenSSL version conflict.
-The conflict appeared when we tried to integrate the CREAMClient gateway library with the rest of the ARCLIB.
+The ARCLIB integration will combine the existing resource discovery (See section \ref{Wrapper classes}) and job submission functionality of the gLite gateway plugins with the brokering module of ARCLIB, which is  currently under development.
+The combination of these ARCLIB modules will provide intelligent brokered job submission over gLite and NorduGrid computing element.\par
+For a fully transparent integration automatic translation among JSDL/xRSL and JDL should also be provided.
+A user should be able to specify his/her job request in any of these languages and the gateway does the necessary translations.
+The current language translator module currently is not capable of generating JDL output.
+Therefore, the user still needs to use JDL to submit jobs to CREAM2.\par
+Further investigation is needed on how a CREAM2 service handles data staging and whether the data movement related ARC libraries can be used to manage data staging on the client side or development of special data staging wrapper scripts executed on the computing element are needed.
+A set of temporary workaround command line clients have been developed (the glite* commands; see section \ref{Users guide}) due to a discovered OpenSSL version conflict.
+The conflict appeared when we tried to integrate the CREAMClient gateway library to the rest of the ARCLIB.
 A tedious and time consuming debugging process took place.
-The debugging revealed the following:  A VOMS proxy certificate delegation method was implemented in the CREAMClient gateway library.
+It revealed the following:  A VOMS proxy certificate delegation method was implemented in the CREAMClient gateway library.
 During this process it was needed to sign a certificate and the original OpenSSL was used for this purpose.
-In the used communication library taken from the ARC1 framework there is a built-in Message Chain Component (MCC) which already utilizes the Globus Toolkit OpenSSL functions to enable establishing secure communication.
+In the communication library used and taken from the ARC1 framework there is a built-in Message Chain Component (MCC) which already utilizes the Globus Toolkit OpenSSL functions to enable establishing secure communication.
 Linking the CREAMClient library directly against the standard OpenSSL libs caused congesting SSL functions.
-In order to resolve the OpenSSL conflict and to make possible the complete ARCLIB integration  a globus patch has been being worked on.
-Once the OpenSSL conflict is resolved the ARCLIB integration will be completed this way obsoleting the glite* temporary commands.\par
+In order to resolve the OpenSSL conflict and to make possible the complete ARCLIB integration  a Globus patch has been worked out.
+Once the OpenSSL conflict is resolved the ARCLIB integration will be completed this way overriding the temporary glite* commands.\par
 \subsection{Conceptional problems}
 During the development of the gateway library a few conceptual incompatibilities were found.
-These are the typical cases when one of the middlewares offers richer capabilities not available in the other one.
+These are the typical cases when one of the systems offers richer capabilities not available on the other one.
 These conceptual gaps can only be bridged by developing customized case by case workarounds, nevertheless  the interoperation always would result in some information loss in these cases.\par
-Below we collected the most important challenges for which we still have to find some solution:
+Below are the most important challenges for which solution is still sought:
 \subsubsection*{Expressing file staging in job description}
-XRSL  offers reacher capability to describe input and output date of a job than the JDL.
-Speciﬁcally, in xRSL it is possible for input files downloaded from storage servers to have their names changed.
+The xRSL offers richer capability to describe input and output date of a job than JDL.
+Especially, in xRSL it is possible for input files downloaded from storage servers to have their names changed.
 This is not possible in JDL.
-So a user might in xRSL ask for this:
+So in xRSL user might ask for this:
 \begin{shaded}
@@
 This could be solved by a wrapper script that renames files according to the xRSL speciﬁcation.
-This does not solve our problem completely however.
+This does not solve the problem completely however.
 For example if a user wants to concatenate the stdout from two jobs, he would write something like the following in xRSL:
@@
 \end{shaded}
-This cannot be solved just using the renaming method, because both input files would be retrieved before renaming is done and they would therefore overwrite each other.\par
+This cannot be solved just using the renaming method, because both input files would be retrieved before the renaming and they would therefore overwrite each other.\par
 A possible  solution is to recognize jobs of this type and simply barring them from using gLite resources or to let such jobs fail with an error message explaining that the user has to express their job in a different way.
@@
 While the library is capable of translating the RTE request itself it is not capable of translating the name of the RTE.
-No standardized cross Grid nomenclature has been adopted, nor are the authors aware of any such activity.
+No standardized cross-grid nomenclature has been adopted, nor are the authors aware of any such activity.
 A temporary solution would be to have a remotely accessible list of RTEs and their names in each job description language.
@@
 \begin{enumerate}
-\item Put the Debian repository into the APT source files (/etc/apt/sources.list). This repository in Luebeck is maintained as long as the KnowARC procject continues.
+\item Put the Debian repository into the APT source files (/etc/apt/sources.list). This repository in Luebeck is maintained as long as the KnowARC project continues.
 \begin{shaded}\verb#$ echo "deb http://pc02.inb.uni-luebeck.de:8080/~moeller/debian/stable ./" \#\\*
@@
 \begin{shaded}\verb#$ aptitude install nordugrid-arc1-client#\end{shaded}
-\item Create the vomses file:
+\item Create the vomses directory
 \begin{shaded}\verb#$ mkdir -p /opt/glite/etc/#\end{shaded}

### 2008-08-14 doc/glite_gateway/glite_gateway.tex
Layout and biblography corrections

@@
 \begin{tabular}{rl}
-\resizebox*{3cm}{!}{\includegraphics{logo-knowarc.png}}
+\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
+&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
 \end{tabular}
@@
 {\centering \large Mattias Ellert\footnote{mattias.ellert@fysast.uu.se} \large \par}
-{\centering \large Ivan Marton\footnote{martoni@niif.hu} \large \par}
-{\centering \large Peter Stefan\footnote{stefan@niif.hu} \large \par}
+{\centering \large Iv\'{a}n M\'{a}rton\footnote{martoni@niif.hu} \large \par}
+{\centering \large P\'{e}ter Stef\'{a}n\footnote{stefan@niif.hu} \large \par}
 %%%%---- An abstract - if style is article ----
@@
 The latter solution was preferred by ARC1 developers, the gLite gateway library module has been implemented in the ARCLIB and the gateway functionality will be transparently offered via the usual NorduGrid command line tools.\par
-An overview of the preferred soulution is presented in figure \ref{fig:Chart 1}.
+An overview of the preferred soulution is presented in figure \ref{fig:Chart1}.
 The user will be able to express the job requirements in the job description document using any of the xRSL\cite{xrsl}, JDL or JSDL\cite{jsdl} languages.
@@
 \centering{\resizebox{\textwidth}{!}{\includegraphics{D3-2-1-figure.png}}
-\caption{\label{fig:Chart 1}This chart illustrates how the ARC1$\rightarrow$gLite gateway library module based approach works.}}
+\caption{\label{fig:Chart1}This chart illustrates how the ARC1$\rightarrow$gLite gateway library module based approach works.}}
 \end{figure}
@@
 \begin{figure}[ht]
-\centering{\resizebox{255px}{!}{\includegraphics{JobDescriptionExample-XRSLInput.jpg}}
-\caption{\label{fig:Chart 1}The figure shows how the different classes participating in the job description translation process interrelate in an example about xRSL parsing.}}
+\centering{\resizebox{255pt}{!}{\includegraphics{JobDescriptionExample-XRSLInput.jpg}}
+\caption{\label{fig:Chart2}The figure shows how the different classes participating in the job description translation process interrelate in an example about xRSL parsing.}}
 \end{figure}

### 2008-08-14 doc/client/arcdata.tex
Something

@@
 
-\texttt{ngcp} can perform multi-stream transfers if \texttt{threads}
+\texttt{arccp} can perform multi-stream transfers if \texttt{threads}
 URL option is specified and server supports it.
@@
    Since the job ID is in fact a \verb#gsiftp://# URL of the job
-   top directory, you can use \texttt{ngcp} to copy files from the
+   top directory, you can use \texttt{arccp} to copy files from the
    job directory at any time.
@@
 
-Usage examples of \texttt{ngcp} are:
+Usage examples of \texttt{arccp} are:
 
@@
 
-\subsection{ngrm}\label{sec:ngrm}
+\subsection{arcrm}\label{sec:arcrm}
 
-The \texttt{ngrm}\footnote{In ARC $\leq$ 0.5.28 was called
-  \texttt{ngremove}} \index{ngrm}\index{commands:ngrm}
+The \texttt{arcrm}\footnote{In ARC $\leq$ 0.5.28 was called
+  \texttt{arcremove}} \index{arcrm}\index{commands:arcrm}
 command allows users to erase files at any location specified by a
@@
 \begin{framed}
-   A convenient use for \texttt{ngrm} is to erase the files in a data
+   A convenient use for \texttt{ngarc} is to erase the files in a data
    indexing catalog (RC, RLS or such), as it will not only remove the
@@
 
-Here is an \texttt{ngrm} example:
+Here is an \texttt{arcrm} example:
 
@@
 
-\subsection{ngacl}\label{sec:ngacl}
+\subsection{arcacl}\label{sec:arcacl}
 
-\index{ngacl}\index{commands:ngacl}This command retrieves or modifies
+\index{arcacl}\index{commands:arcacl}This command retrieves or modifies
 access control information associated with a stored object if service
@@
 \begin{shaded}
-   \uicommand{ngacl [options] get$|$put $<$URL$>$}
+   \uicommand{arcacl [options] get$|$put $<$URL$>$}
 \end{shaded}
-\versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
+\versions{ARC 0.9}
 \begin{longtable}{llp{8cm}}
@@
 
-\subsection{ngtransfer}\label{sec:ngtransfer}
-\index{ngtransfer}\index{commands:ngtransfer}The \texttt{ngtransfer}
-command\footnote{In ARC $\leq$ 0.5.28, was called \texttt{ngrequest}}
-initiates direct transfer of data between 2 servers (known as
-\textit{third-party transfer}).  
-\hspace*{0.5cm}
-\begin{shaded}
-   \uicommand{ngtransfer [options] $<$destination$>$}
-\end{shaded}
-\versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-\begin{longtable}{llp{8cm}}
-   Options:&&\\
-    \texttt{-s, -source} & \textit{URL} & source file URL\\
-    \texttt{-d, -debug} & \textit{debuglevel} &debug level: 0 = some, 1 = more, 2 = a lot \\
-    \texttt{-v} && print version information\\
-    \texttt{-h} && short help\\
-   Arguments:&&\\
-    \texttt{destination} && destination URL; currently only se:// and
-    (gsi)ftp:// are supported\\
-\end{longtable}
-
-This command initiates file copy from multiple source instances to a
-destination URL. Destination of \texttt{(gsi)ftp} type accepts only
-similar kinds of sources.  Destination can also be URL of an Indexing
-Service. In such a case, real destinations with suitable protocols are
-chosen, when available. Requests are sent to the corresponding
-services/servers to initiate file transfer from one of the sources.
-
-\begin{framed}
-   Absence of \verb#-s# option is treated as a file replication
-   request. In this case, destination must be an Indexing service or
-   an SRM.
-\end{framed}
-
-Following source URL types are supported: \texttt{http},
-\texttt{https}, \texttt{httpg}, \texttt{ftp}, \texttt{gsiftp},
-\texttt{rc}, \texttt{rls}, \texttt{se}, \texttt{srm}, \texttt{fireman}
-(see Section~\ref{sec:urls} for URL details).
-
-Example:
-\begin{verbatim}
-    ngtransfer -s http://www.host1.org/dat1  -s gsiftp://host2.org/dir/dat1 \
-               se://se.host.org/se_service?new_file_lfn
-\end{verbatim}
+\subsection{arctransfer}\label{sec:arctransfer}
+\index{arctransfer}\index{commands:arctransfer}The \texttt{arctransfer}
+This command is not implemented.

### 2008-08-14 doc/client/arcjob.tex
Something

@@
    Options:&&\\
-   \texttt{ -c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
-   \texttt{ -C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
-   \texttt{ -g, -giisurl}&\textit{url}&URL of a central Information System server (GIIS)\\
-   \texttt{ -G, -giislist}&\textit{filename}&list of GIIS URLs\\
-   \texttt{ -e, -xrsl}&\textit{filename}&string describing the job to be submitted\\
-   \texttt{ -f, -file}&\textit{filename}&file describing the job to be submitted\\
-   \texttt{ -o, -joblist}&\textit{filename}&file where the job IDs will be stored\\
-   \texttt{ ~~~~-dryrun}&&add dryrun option to the job description\\
-   \texttt{ ~~~~-dumpxrsl}&&do not submit -- dump transformed job description to stdout\\
-   \texttt{ -t, -timeout}&\textit{time}&timeout for queries (default 40 sec)\\
-   \texttt{ -d, -debug}&\textit{debuglevel}&debug level, from -3 (quiet) to 3 (verbose) - default 0\\
-   \texttt{ -x, -anonymous}&&use anonymous bind for queries (default)\\
-   \texttt{ -X, -gsi}&&use GSI-GSSAPI bind for queries\\
-   \texttt{ -v, -version}&&print version information\\
-   \texttt{ -h, -help}&&print help page\\
+   \texttt{ -c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject a specific site (cluster)\\
+   \texttt{ -i, --indexurl}&\textit{url}&URL of an index server\\
+   \texttt{ -e, --jobdescrstring}&\textit{filename}&string describing the job to be submitted\\
+   \texttt{ -f, --jobdescrfile}&\textit{filename}&file describing the job to be submitted\\
+   \texttt{ -j, --joblist}&\textit{filename}&file where the job IDs will be stored\\
+   \texttt{ -D, --dryrun}&&add dryrun option to the job description\\
+   \texttt{ -x, --dumpdescription}&&do not submit -- dump transformed job description to stdout\\
+   \texttt{ -t, --timeout}&\textit{time}&timeout for queries in seconds (default 20)\\
+   \texttt{ -d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
+   \texttt{ -U, --unknownattr}&&allow unknown attributes in the job description\\
+   \texttt{ -v, --version}&&print version information\\
+   \texttt{ -?, --help}&&print help page\\
    Arguments:&&\\
@@
 \begin{framed}
-Some text can be in frames
+The -c and -i arguments accept ``urls'' of the format GRID:URL,
+
+e.g. for index servers:
+
+ARC0:ldap://grid.tsl.uu.se:2135/mds-vo-name=sweden,O=grid
+
+CREAM:ldap://cream.grid.upjs.sk:2170/o=grid
+
+or clusters:
+
+ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid
 \end{framed}
 
-A simple \textit{"Hello World"} job would look like
+A simple \textit{"Hello World"} job would look like...
 

### 2008-08-14 doc/a-rex/arex_tech_doc.tex
More changes in a-rex document

@@
 \begin{center}
-\textsc{\large The ARC Computational Job Management Component
-- A-Rex}
+\textsc{\large The ARC Computational Job Management Component - A-Rex}
 \par\end{center}{\large \par}
@@
 
-The aim of the   A-Rex is to take care of job pre-
-and post-processing. It provides functionality to stage-in files containing
-input data and program modules from wide range of sources and transfer
-or store output results.
-
-The A-Rex implements Web Service (WS) interface which   
-    provides a way to submit and control
-computational tasks (jobs) to be executed by the A-Rex and undelying
-Local Resource Management System.
+The aim of the A-Rex is to take care of job pre- and post-processing.
+It provides functionality to stage-in files containing input data
+and program modules from wide range of sources and transfer or store
+output results.
+
+The A-Rex implements Web Service (WS) interface which provides a way
+to submit and control computational tasks (jobs) to be executed by
+the A-Rex and undelying Local Resource Management System.
 
@@
 and understanding of internals of the aforementioned tools. For general
-instalaltion and configuration refer to other documents avaialble
-at \url{http://www.nordugrid.org/papers.html}. }
+instalaltion and configuration of whole system refer to other documents
+avaialble at \url{http://www.nordugrid.org/papers.html}. }
 
@@
 
-A job is a set of input files (which may or may not include executables),
-a main executable and a set of output files. The process of gathering
-input files, executing a job, and transferring/storing output files
-is called a \emph{session}.
+A job is a set of input files (which may include executables), a main
+executable and a set of output files. The process of gathering input
+files, executing a job, and transferring/storing output files is called
+a \emph{session}.
 
@@
 (SD). Input files are gathered in the SD. The job is supposed to produce
-new data files  in the SD as well. The A-Rex does not guarantee
-the availability of any other places accessible by the job other than
+new data files in the SD as well. The A-Rex does not guarantee the
+availability of any other places accessible by the job other than
 SD (unless such a place is part of a requested Runtime Environement).
-The SD is also the only place which is controlled by the A-Rex.
-It is  accessible by the user from outside through HTTP(S)  
-protocol. Any file created outside the SD is not controlled by the
-A-Rex. Any exchange of data between client and A-Rex (including
-also program modules) is done through    HTTP(S) protocol
-   . A URL for accessing input/output
-files is       obtained through
-the WS Local Information Description Interface (LIDI) of A-Rex.
-
-Each job gets an identifier (\textit{jobid}). This is a handle which
-identifies the job in the A-Rex and the Information  Interface.
-
-  Jobs are initiated and controlled through   WS interface.
-Complete job descriptions   (JD)   are passed to the
-A-Rex  through       WS in JSDL \cite{jsdl}
-coded description     . Input data files
-and job executables are transferred separately through the same interface,
-as described in the next section.
+The SD is also the only place which is controlled by the A-Rex. It
+is accessible by the user from outside through HTTP(S) protocol. Any
+file created outside the SD is not controlled by the A-Rex. Any exchange
+of data between client and A-Rex (including also program modules)
+is done through HTTP(S) protocol. A URL for accessing input/output
+files is  obtained through the WS Local Information Description Interface
+(LIDI) of A-Rex.
+
+Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
+\cite{jsdl} XML document - which identifies the job in the A-Rex
+and the Information Interface.
+
+Jobs are initiated and controlled through WS interface. Complete job
+descriptions (JD)are passed to the A-Rex through WS in JSDL\cite{jsdl}
+coded description. Input data files and job executables are transferred
+separately through the same interface, as described in the next section.
 
@@
 
-One of the most important tasks   of the A-Rex is to
-take care of processing input and output data (files) of the job.
-Input files are gathered in the SD. There are two ways to put a file
+One of the most important tasks of the A-Rex is to take care of processing
+input and output data (files) of the job. Input files are gathered
+in the SD or associated cache area. There are two ways to put a file
 into the SD:
@@
 \begin{itemize}
-\item Download is initiated by the A-Rex --  This is the case
-for files     defined in the JD (with name and source).
-      The A-Rex alone is responsible to
+\item Download is initiated by the A-Rex -- This is the case for files defined
+in the JD (with name and source). The A-Rex alone is responsible to
 make sure that all required files will be available in the SD.\\
-The supported protocols for sources  at the moment are: GridFTP,
-FTP, HTTP, HTTPS (HTTP over SSLv3) and HTTPg (HTTP over GSI). Also
-some nonstandard sources are supported. Those are described below.
+The supported protocols for sources at the moment are (in case of
+full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
+some less standard sources are supported. Those are described below.
 The A-Rex fully relies on HED framework \cite{hed} for data transfering
@@
 submission of JD, UI can (and should) use that to upload data files
-which are not otherwise accessible by the A-Rex.  Examples of
-such files  are the main executable of the job,  
-the job's input files, etc. These files can (and
-should) also be specified in the JD.
+which are not otherwise accessible by the A-Rex. Examples of such
+files are the main executable of the job, the job's input files, etc.
+These files can (and should) also be specified in the JD.
 \end{itemize}
 \textbf{There is no} other reliable way for a job to obtain input
-data on the CE based on the A-Rex  . Access to
-AFS, NFS, FTP, HTTP and any other remote data transport during execution
-of a job is not guaranteed (at least not by A-Rex).
-
-Jobs should store output files in their SD. Like input files,
-output files  belong into two groups:
+data on the CE based on the A-Rex. Access to AFS, NFS, FTP, HTTP and
+any other remote data transport during execution of a job is not guaranteed
+(at least not by A-Rex).
+
+Jobs should store output files in their SD. Like input files, output
+files belong into two groups:
 
@@
 and optionally registered in some \emph{Indexing Service} like the
-Globus \emph{Replica Location Service } (RLS)  --
-The A-Rex takes care of these files. They have to be specified
-in the JD. If the job fails during any stage of processing, no attept
-is made to transfer those files to their final destination, unless
-the  option \emph{preserve=yes} is specified in their URLs.
+Globus \emph{Replica Location Service} (RLS) -- The A-Rex takes care
+of these files. They have to be specified in the JD. If the job fails
+during any stage of processing, no attept is made to transfer those
+files to their final destination, unless the option \emph{preserve=yes}
+is specified in their URLs.
 \item Files which are supposed to be fetched by the user -- The user has
@@
 \end{itemize}
+All files not specified in the JD are deleted after job execution
+finished. If job execution fails for any reason (if exit code of main
+executable is not 0) all files from furst group are transfered to
+second one.
 
@@
 
-From the point of view of the A-Rex a job passes through various
-states. Picture \ref{job states diagram} presents a diagram of the
-possible states of a job.%
+From the point of view of the A-Rex a job passes through various states.
+Picture \ref{job states diagram} presents a diagram of the possible
+states of a job.%
 \begin{figure}[tbh]
@@
 A user can examine the state of a job by querying the dedicated Local
-Information Description Interface of A-Rex  
- using the UI or any other suitable tool
-
-     or through query method of WS interface.
+Information Description Interface of A-Rex using the UI or any other
+suitable tool or through query method of WS interface.
 
@@
 the state in question will stay in it's current state waiting for
-a free slot. This situation is presented by prepending the prefix
-\textbf{PENDING:} to the current state name  \textbf{}
-in the job's status mark.
+a free slot. This situation is presented by additional state mark
+\textbf{PENDING} to the current state name in the job's status desription.
 
@@
 \begin{itemize}
-\item \textbf{Accepted} - In this state the job has been submitted to
-a CE but is not processed yet. The A-Rex will analyze the JD and
-move to the next stage. If the JD can not be processed the job will
-be canceled and moved to the state \textbf{Finishing}.
+\item \textbf{Accepted} - In this state the job has been submitted to a
+CE but is not processed yet. The A-Rex will analyze the JD and move
+to the next stage. If the JD can not be processed the job will be
+canceled and moved to the state \textbf{Finishing}.
 \item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
 The A-Rex is downloading the files specified in the JD and is waiting
-for files which are supposed to be uploaded by the UI. If
-all files are successfully gathered the job moves to the next state.
-If \textbf{any} file can't be downloaded or it takes the UI too long
-to upload a file,  the job moves to \textbf{Finishing} state. It
-is possible to put a limit on the number of simultaneous \textbf{Preparing}
-jobs. If this limit is exceeded, jobs ready to enter the \textbf{Preparing}
-state      will stay in the \textbf{Accepted}
-state, but prefixed with the PENDING: mark. Exceptions are jobs which
-have no files to be downloaded. These are processed out of limits.
-\item \textbf{Submitting} -      The job is being
-passed for execution to the \emph{Local Resource Management System}
-(LRMS).         The correspoding backends
-for many LRMSs are provided with the default installation.  
-     If the local job submission is successful
-the job moves to the \textbf{InLRMS} state. Otherwise it moves
-to \textbf{Finishing}. It is possible to limit the agregate number
-of jobs in \textbf{Submitting} and  \textbf{InLRMS} states.
-\item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
+for files which are supposed to be uploaded by the UI. If all files
+are successfully gathered the job moves to the next state. If \textbf{any}
+file can't be downloaded or it takes the UI too long to upload a file,
+the job moves to \textbf{Finishing} state. It is possible to put a
+limit on the number of simultaneous \textbf{Preparing} jobs. If this
+limit is exceeded, jobs ready to enter the \textbf{Preparing} state
+ will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
+mark. Exceptions are jobs which have no files to be downloaded. These
+are processed out of limits.
+\item \textbf{Submiting} - The job is being passed for execution to the
+\emph{Local Resource Management System} (LRMS). The correspoding backends
+for many LRMSs are provided with the default installation. If the
+local job submission is successful the job moves to the \textbf{Executing}
+state. Otherwise it moves to \textbf{Finishing}. It is possible to
+limit the agregate number of jobs in \textbf{Submiting} and \textbf{Executing}
+states.
+\item \textbf{Executing} - The job is queued or being executed in the LRMS.
 The A-Rex takes no actions except waiting until the job finishes.
-\item \textbf{Cancelling} - Necessary action to cancel the job in the LRMS
+\item \textbf{Killing} - Necessary action to cancel the job in the LRMS
 is being taken.
@@
 The user can continue to download data files from the SD. The SD is
-kept available for some time (default is 1 week). After that the
-job is moved to the state \textbf{Deleted}. The 'deletion' time can
-be obtained by querying the Informatrion Interface of the A-Rex
-      . If a job was moved to \textbf{Finished}
-because of failure, it may be restarted on request of a client. When
-restarted, a job is moved to the state previous to the one in which
-it failed and is assigned mark PENDING. This is needed in order to
-not break the configuration limits. Exception is a job failed in \textbf{InLRMS}
-state and lacking input files specified in JD. Such a job is treated
-like failed in \textbf{Preparing} state.
+kept available for some time (default is 1 week). After that the job
+is moved to the state \textbf{Deleted}. The 'deletion' time can be
+obtained by querying the Informatrion Interface of the A-Rex. If a
+job was moved to \textbf{Finished} because of failure, it may be restarted
+on request of a client. When restarted, a job is moved to the state
+previous to the one in which it failed and is assigned mark PENDING.
+This is needed in order to not break the configuration limits. Exception
+is a job failed in \textbf{Executing} state and lacking input files
+specified in JD. Such a job is treated like failed in \textbf{Preparing}
+state.
 \item \textbf{Deleted} - The job is moved to this state if the user have
@@
 \end{itemize}
-In  case of a failure special processing is applied to output
-files. All specified output files are treated as \textbf{downloadable
-by the user}. No files will be moved to their destination SE.
+In case of a failure special processing is applied to output files.
+All specified output files are treated as \textbf{downloadable by
+the user}. No files will be moved to their destination SE.
 
@@
 
-In full installation the A-Rex and it's components support the
-following data transfer protocols and corresponding URLs: \emph{ftp,
-gsiftp, http, httpg, https, se, rc} and \emph{rls.} For more information
-please see {}``The Hosting Environment of the Advanced Resource Connector
-middleware      
-  '' document \cite{hed}.
+In full installation the A-Rex and it's components support the following
+data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
+http, https, lfc} and \emph{rls.} For more information please see
+{}``The Hosting Environment of the Advanced Resource Connector middleware''
+document \cite{hed}.
 
@@
 
-\subsection{Internal Files of the A-Rex }
+\subsection{Internal Files of the A-Rex}
 
@@
 about jobs belonging to that user. Multiple users can share the same
-\textit{control directory}. To make it easier to recover in  case
-of failure, the A-Rex stores most information in files rather than
-in memory. All files belonging to the same job have names starting
-with \textbf{job.ID.,} where ID is the job identifier.
+\textit{control directory}. In most common configuration case A-Rex
+serves all users defined by Operating System and stores their control
+file in same directory. To make it easier to recover in case of failure,
+the A-Rex stores most information in files rather than in memory.
+All files belonging to the same job have names starting with \textbf{job.ID.,}
+where ID is the job identifier.
 
@@
 \item \textit{job.ID.status} - current state of the job. This is a plain
-text file containing a single word   representing the current
-state of the job. Possible values are :
-
-\begin{itemize}
-\item ACCEPTED
-\item PREPARING
-\item SUBMITTING
-\item INLRMS
-\item FINISHING
-\item FINISHED
-\item CANCELING
-\item DELETED
+text file containing a single word representing the internal name
+of current state of the job. Possible values and corrsponding external
+job states are :
+
+\begin{itemize}
+\item ACCEPTED/\textbf{Accepted}
+\item PREPARING/\textbf{Preparing}
+\item SUBMITTING/\textbf{Submiting}
+\item INLRMS/\textbf{Executing}
+\item FINISHING/\textbf{Finishing}
+\item FINISHED/\textbf{Finished}
+\item CANCELING/\textbf{Killing}
+\item DELETED/\textbf{Deleted}
 \end{itemize}
@@
 See section \ref{Section: Job Flow} for a description of the various
-states. Additionally each value can be prepended  the prefix
-{}``PENDING:'' (like PENDING:ACCEPTED, see section \ref{Section: Job Flow}).
-This is used to show that a job  is \emph{ready} to be moved
-to the next state but it has to stay in it's current state \emph{only}
-because otherwise some limits set in the configuration would be
-exceeded.
-
-\begin{itemize}
-\item \textit{job.ID.description} - contains the  description of the
-job (JD).
+states. Additionally each value can be prepended the prefix {}``PENDING:''
+(like PENDING:ACCEPTED, see section \ref{Section: Job Flow}). This
+is used to show that a job is \emph{ready} to be moved to the next
+state but it has to stay in it's current state \emph{only} because
+otherwise some limits set in the configuration would be exceeded.
+
+\begin{itemize}
+\item \textit{job.ID.description} - contains the description of the job
+(JD).
 \item \textit{job.ID.local} - information about the job used by the A-Rex.
-It consists of lines of format \textit{{}``name = value''} . Not
+It consists of lines of format \textit{{}``name = value''}. Not
 all of them are always available. The following names are defined:
@@
 in the Generalized Time format of LDAP 
-\item \textit{lifetime} - time period to  preserve the SD after the
-job has finished in seconds
-\item \textit{cleanuptime} -  GMT time when the job should be removed
-from the cluster and it's SD deleted in Generalized Time format
+\item \textit{lifetime} - time period to preserve the SD after the job has
+finished in seconds
+\item \textit{cleanuptime} - GMT time when the job should be removed from
+the cluster and it's SD deleted in Generalized Time format
 \item \textit{notify} - email addresses and flags to send mail to about
 the job specified status changes
-\item \textit{processtime} -  GMT time when to start processing the job
-in Generalized Time format
+\item \textit{processtime} - GMT time when to start processing the job in
+Generalized Time format
 \item \textit{exectime} - GMT time when to start job execution in Generalized
 Time format
-\item \textit{expiretime} -  GMT time when the credentials delegated
-to the job expire in Generalized Time format
+\item \textit{expiretime} - GMT time when the credentials delegated to the
+job expire in Generalized Time format
 \item \textit{rerun} - number of retries left to rerun the job
 \item \textit{jobname} - name of the job as supplied by the user
-\item \textit{lrms} - name of the LRMS backend     to be used
-for local submission
+\item \textit{lrms} - name of the LRMS backend to be used for local submission
 \item \textit{queue} - name of the queue to run the job at
-\item \textit{localid} - job id in LRMS (appears only after the job
- has reached state \textbf{InLRMS})
+\item \textit{localid} - job id in LRMS (appears only after the job has
+reached state \textbf{InLRMS})
 \item \textit{args} - executable name followed by a list of command-line
@@
 ip address:port of the submitting client machine
-
 \item \textit{clientsoftware} - version of software used to submit the job
 \item \textit{sessiondir} - the job's SD
-\item \textit{failedstate} - state in which job failed (available only
-if it is possible to restart the job)
+\item \textit{failedstate} - state in which job failed (available only if
+it is possible to restart the job)
 \item \textit{jobreport} - URL of a user requested \emph{logger service}.
 The GM will also send job records to this service in addition to the
-default logger service configured in the configuration  
-        (This feature does
-not work in current implementation).
+default logger service configured in the configuration (This feature
+does not work in current implementation).
 \end{itemize}
@@
 
-\textit{url} - ordinary URL for gsiftp, ftp, http, https or httpg
-protocols with the addition of '\textbf{replica catalog url}' (RC
-URL) and '\textbf{replica location service url}' (RLS URL).\\
+\textit{url} - ordinary URL for gsiftp, ftp, http, https protocols
+with the addition of '\textbf{replica location service url}' (RLS
+URL).\\
 Each URL can contain additional options.
@@
 
-\item \textit{job.ID.proxy} -delegated X509 credentials .
-\item \textit{job.ID.proxy.tmp} - temporary X509 credentials  with
-different UNIX ownership used by processes run with effective \emph{user
-id} different form job owner's \emph{id}.
+\item \textit{job.ID.proxy} -delegated X509 credentials.
+\item \textit{job.ID.proxy.tmp} - temporary X509 credentials with different
+UNIX ownership used by processes run with effective \emph{user id}
+different form job owner's \emph{id}.
 \end{itemize}
 There are other files with names like job.ID.{*} which are created
-and used by different parts of the A-Rex. Their presence in the
-\textit{control directory} can not be guaranteed and can change depending
-on changes in the A-Rex code.
-
-\subsection{ }
-
-    \textit{ }  \textit{}
-
- \textit{}
-
- \textit{}
-
- \textit{}
+and used by different parts of the A-Rex. Their presence in the \textit{control
+directory} can not be guaranteed and can change depending on changes
+in the A-Rex code.
 
@@
 and thus defines operations to start and stop the functionality of
-the BES service. The A-Rex does not implement remote control of
-service functionality. Hence the BES-Management port-type is not functional.
+the BES service. The A-Rex does not implement remote control of service
+functionality. Hence the BES-Management port-type is not functional.
 The BES-Factory port-type provides operations to submit new jobs (to
@@
 
-A-R introduces a new operation in addition to those provided by
-BES. It does that by defining its own port-type  with the single
-operation \emph{ChangeActivityStatus} (see Appendix D). This operation
-provides a way to request simple transfers between states of jobs
-  and corresponding actions.
+A-Rex introduces a new operation in addition to those provided by
+BES. It does that by defining its own port-type with the single operation
+\emph{ChangeActivityStatus} (see Appendix D). This operation provides
+a way to request simple transfers between states of jobs and corresponding
+actions.
 
@@
 Since BES allows implementators to extend their initial activity states
-with additional sub-states, A-Rex defines a set of sub-states of
-activity processing in addition to those defined by the BES, as listed
-in Table \ref{tab:Job-states-definitions}.
+with additional sub-states, A-Rex defines a set of sub-states of activity
+processing in addition to those defined by the BES, as listed in Table
+\ref{tab:Job-states-definitions}. Their meaning is described in section
+\ref{Section: Job Flow}.
 
@@
 \multicolumn{1}{|c|}{}&
-Submitting&
+Submiting&
 Communication with local batch system is in process\tabularnewline
@@
 Job processing is suspended on client request; this sub-state appears
-in parallel with other sub-states\tabularnewline
+in parallel with other sub-states. This state is reserved for future
+and i snot implemented yet.\tabularnewline
 \hline
@@
 
-       The A-Rex
-also supportd  the Delegation Interface (see Appendix E). This
-is a common purpose interface to be used by ARC services which accepts
-delegated credentials from clients. The Delegation Interface implements
-two operations: initiatialization of credentials delegation (DelegateCredentialsInit)
-and update/renewal of credentials (UpdateCredentials).
+The A-Rex also supportd the Delegation Interface (see Appendix E).
+This is a common purpose interface to be used by ARC services which
+accepts delegated credentials from clients. The Delegation Interface
+implements two operations: initiatialization of credentials delegation
+(DelegateCredentialsInit) and update/renewal of credentials (UpdateCredentials).
 
@@
 response to the previous \emph{DelegateCredentialsInit} operation.
+For more information about delegations and delegation interface refer
+to \cite{wsrf-rp}.
 
@@
 The A-REX implements the Local Information Description Interface (LIDI)
-interface  common for all ARC services. This interface is based
-on OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
+interface common for all ARC services. This interface is based on
+OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
 Information about resources and maintained activities/jobs are represented
@@
 \begin{lyxlist}{00.00.0000}
-\item [{\emph{BESFactory}}] - collection of BES Factory attributes as defined
-in the BES specifications.
-\item [{\emph{Glue2Resource}}] \emph{-} description of a computation resource
-that uses Glue2 schema. This one is going to be specified after the
-Glue2 XML bindings will be available.
-\item [{\emph{Activities}}] \emph{-} list of maintained activities. Each
-activity contains an identifier (\emph{ActivityIdentifier}), BES (\emph{ActivityDocument})
-and Glue2 (\emph{Glue2Job}) descriptions of the activity.
+\item [{\emph{nordugrid}}] - desription of computing resource that uses
+NorudGrid LDAP schema {[}] converted to XML document.
+\item [{\emph{Domains}}] \emph{-} description of a computation resource
+that uses Glue2 schema.
 \end{lyxlist}
@@
 
-If a job requests an input file which is subject  for
-cacheing, it is stored in the cache directory instead of the SD.
-The file is made available to the job by either  soft-linking it
-  in the SD         
-    or copying it to the SD. The latter option
-is more secure and hence advised.
+If a job requests an input file which is subject for cacheing, it
+is stored in the cache directory instead of the SD. The file is made
+available to the job by either soft-linking it in the SD or copying
+it to the SD. The latter option is more secure and hence advised.
 
@@
 system has no more space), it tries to the remove oldest cached files
-which are not being used by any job. This means that \textbf{the
-hard limit of cache size is the space available in the file-system}.
-In case cache gets full and it is impossible to free enough space,
-the download fails and then is retried without using cache.
+which are not being used by any job. This means that \textbf{the hard
+limit of cache size is the space available in the file-system}. In
+case cache gets full and it is impossible to free enough space, the
+download fails and then is retried without using cache.
 
 Before giving access to an already cached file, the GM contacts the
-original file source to check if the user     has
-suficient access rights to the file. Not all protocols  
- make this check possible.
-
-Also, file creation or validity times are checked to make sure the
-cached file is fresh . If it is impossible to obtain creation
-and invalidation times  , the file is invalidated 24 hours
-after it was downloaded.
-
- The GM checks the cache periodically. If  the space used
-by cache exceeds the high water-mark given in the configuration file
-(\emph{softsize}), it tries to remove the oldest unused files until
-cache size drops below the  low water-mark. This sets the
-soft limit of cache size.
+original file source to check if the user has suficient access rights
+to the file. Not all protocols make this check possible. Also, file
+creation or validity times are checked to make sure the cached file
+is fresh. If it is impossible to obtain creation and invalidation
+times, the file is invalidated 24 hours after it was downloaded.
+
+The GM checks the cache periodically. If the space used by cache exceeds
+the high water-mark given in the configuration file (\emph{softsize}),
+it tries to remove the oldest unused files until cache size drops
+below the low water-mark. This sets the soft limit of cache size.
 
@@
 
-The GM consists of several separate executable modules. These
-are:
+The GM consists of several separate executable modules. These are:
 
@@
 \item \textit{uploader} - This module is responsible for delivering output
-files to the specified SEs and registration at  an Indexding
-Service (like RLS) as needed. It processes and updates the \textit{job.ID.output}
+files to the specified SEs and registration at an Indexding Service
+(like RLS) as needed. It processes and updates the \textit{job.ID.output}
 file.
 \item \emph{cache-register} - Utility to register cached data into an Indexing
-Service    . It reads and modifies cache informational
-files \emph{old} and \emph{new} (as described in section \ref{sec:cache}).
-Configuration is read directly from the GM's configuration file (see
-section \ref{SubSection: ConfigFile}). It is run by the GM every
-5 minutes.
+Service. It reads and modifies cache informational files \emph{old}
+and \emph{new} (as described in section \ref{sec:cache}). Configuration
+is read directly from the GM's configuration file (see section \ref{SubSection: ConfigFile}).
+It is run by the GM every 5 minutes.
 \item \emph{frontend-info-collector} - Utility to gather information about
-the frontend  . It puts  collected information into the \emph{job.ID.diag}
+the frontend. It puts collected information into the \emph{job.ID.diag}
 file.
@@
 for job submission to the LRMS.
-\item \textit{cancel-{*}-job} - This script is for canceling  jobs
-which have been already submitted to the LRMS.
+\item \textit{cancel-{*}-job} - This script is for canceling jobs which
+have been already submitted to the LRMS.
 \item \textit{scan-{*}-job} -This shell script is responsible for notifying
-the GM about completion of  jobs. It's implementation for PBS 
-uses server logs to extract information about jobs. If logs are not
-available it uses the less reliable \emph{qstat} command for that.
-Other backends use different techniques.
+the GM about completion of jobs. It's implementation for PBS uses
+server logs to extract information about jobs. If logs are not available
+it uses the less reliable \emph{qstat} command for that. Other backends
+use different techniques.
 \end{itemize}
@@
 \begin{itemize}
-\item \textit{session root directory} -  Tthis is the directory under
-which a user's SDs are created. It's location is configurable
-per UNIX user. Several (or even all) users may share the same session
-root directory.        
-    .\\
+\item \textit{session root directory} - Tthis is the directory under which
+a user's SDs are created. It's location is configurable per UNIX user.
+Several (or even all) users may share the same session root directory..\\
 There are 2 processes which need to have permissions to create new
-files and directories in it  : GM and GFS. \\
-If any of these processes  are run under a dedicated user
-account, that account needs full permissions in the \textit{session
-root directory}.\\
+files and directories in it: GM and GFS. \\
+If any of these processes are run under a dedicated user account,
+that account needs full permissions in the \textit{session root directory}.\\
 If these processes are run under the \emph{root} account, make sure
-\textit{session root directory}  resides on a filesystem which
-does not limit the capabilities of the \emph{root} user  (as does
-for example NFS with \emph{root\_squash} option).\\
+\textit{session root directory} resides on a filesystem which does
+not limit the capabilities of the \emph{root} user (as does for example
+NFS with \emph{root\_squash} option).\\
 If there is a need to run processes under the \emph{root} account
@@
 root directory,} use the \emph{norootpower} command in GM's configuration
-  file. In that case GM and GFS will use the identity of the
-local user to which a Grid identity is mapped to access the \textit{session
-root directory}. Hence those users will need full access there.\\
+file. In that case GM and GFS will use the identity of the local user
+to which a Grid identity is mapped to access the \textit{session root
+directory}. Hence those users will need full access there.\\
 The GM creates SDs with proper ownership and permissions for the local
 identity used to run a job. Some filesystems require users to have
-\emph{execute} permission on the \textit{session root directory}
-      in order to access any file or subdirectory
+\emph{execute} permission on the \textit{session root directory} in
+order to access any file or subdirectory there.\\
+In order for jobs to access their input files, session root directories
+should be shared across cluster nodes. Otherwise, LRMS-specific methods
+must be used to transfer files to execution nodes. For more information
+see section \ref{sub:LRMS}.
+\item \textit{control directory} - In this directoryGM and GFS store an
+information about accepted jobs. Both processes must have full permissions
 there.\\
-         In order
-for jobs to access their input files, session root directories should
-be shared across cluster nodes. Otherwise,    LRMS-specific
-methods must be used to transfer files to execution nodes. For more
-information see section \ref{sub:LRMS}.
-\item \textit{control directory} - In this directory  GM and GFS store
-an information about  accepted jobs. Both    processes
-must have full permissions there.\\
-A subdirectory called \textit{log} is created there. It is used
-to accumulate information about started and finished jobs. This information
+A subdirectory called \textit{log} is created there. It is used to
+accumulate information about started and finished jobs. This information
 is periodically sent to the desired \emph{logger service}(s). For
@@
 \hspace*{0.5cm}\emph{-c} - name od configuration file,\\
-\hspace*{0.5cm}\emph{-C} - remove old information before starting:
-1- remove finished jobs, 2 - remove active jobs too, 3- also remove
-everything that looks like junk.
-
-\subsection{    }
-
-        \textit{}
- \textit{}
-
-\begin{itemize}
-\item \textbf{\textit{}}\textit{}
-
-\item \textbf{\textit{}}\textit{}
-
-\item \textbf{\textit{}}\textit{}
-
-\item \textbf{\textit{}}\textit{}
-
-\item \textbf{\textit{}}\textit{}     \emph{}
-
-\item \textbf{\textit{}}\textit{}      
-\textit{}
-
-  \textit{}
-
-\item \textbf{\textit{}}\textit{ }
-
- \textit{}       
-        \textit{}
-
-\begin{itemize}
-\item \textbf{\textit{}} \textit{}
-
-\item \textbf{\textit{}} \textit{}    
-         \textit{}
-
-\textit{}
-
-\item \textbf{\textit{}} \textit{  }  
-      \textit{}    
-        \textit{}
-\textit{}       
-   \textit{}     
-       \textit{}
-
-\item \textbf{\textit{}} \textit{ } \textit{}
-\textit{}     \emph{}  
-    \emph{}    
- \emph{}          
-     \emph{}
-
-\begin{itemize}
-\item      
-\item       
-\end{itemize}
-\end{itemize}
-\item \textbf{\textit{}}\textit{ }    \textbf{\textit{}}
-       \textit{}
-\item \textbf{\textit{}}\textit{ }
-
-\textit{}       \textbf{\textit{
-}}      \textbf{\textit{}}\textbf{\textit{}}
-
- \textit{}    \textit{}
- \textit{}
-
-\item \textbf{\textit{}}\textit{}
-
-      \textbf{\textit{}}
-
-\end{itemize}
-  \emph{}
-
-          \textbf{\textit{}}
-\textit{ }     \textbf{\textit{}}
-
-\begin{itemize}
-\item \textbf{\textit{}}\textit{}    \textit{}
-
-\item \textbf{\textit{}}\textit{}
-
-     \textit{ } 
-\textit{}
-
-\begin{itemize}
-\item \textit{}
-
-\begin{itemize}
-\item \textbf{\textit{}}\textit{}
-
-\item \textbf{\textit{}}\textit{}     
-     \emph{}
-\item \textbf{\textit{}}\textbf{\textit{}}\textbf{\textit{}}
-
-\end{itemize}
-\item \textit{}
-
-\begin{itemize}
-\item \textbf{\textit{}}\textit{}
-
-\item \textbf{\textit{}}\textit{ }    
-    \textit{}     
-      \textit{}
-
-\begin{itemize}
-\item \textbf{\textit{}}
-
-\item \textbf{\textit{}}       
-\item \textbf{\textit{}}      
-\item \textbf{\textit{}}      
-\end{itemize}
-
-\begin{itemize}
-\item \textbf{\textit{}}    
-\item \textbf{\textit{}}    
-\item \textbf{\textit{}}        
-\item \textbf{\textit{}}
-
-\item \textbf{\textit{}}       
-\item \textbf{\textit{}}       
-\item \textbf{\textit{}} \textit{ }
-          \textit{} 
-    \textit{}
-
-   \textit{}  \textit{}
-
-\item \textbf{\textit{}} \textit{ }
-
-\end{itemize}
-\end{itemize}
-\item \textit{}
-
-      \textbf{\textit{}}
-
-   \textbf{\textit{}}\textit{}  
- \textbf{\textit{}}\textit{}
-
-\begin{lyxlist}{00.00.0000}
-\item [{\emph{}}]      
-\item [{\emph{}}]      
-\item [{\emph{}}]       
-\item [{\emph{}}]     
-\item [{\emph{}}]     
-\item [{\emph{}}]        
-\end{lyxlist}
-     \emph{}
-
-\end{itemize}
-\end{itemize}
-
-\emph{         
-          }\emph{}
-  \emph{}   \emph{}    
-    \emph{}    
-       \emph{}
-         \emph{}   
-  \emph{}     \emph{}
-   \emph{}     
- \emph{}        
-   \emph{}
+\hspace*{0.5cm}\emph{-C} - remove
+
+old information before starting:
+
+1- remove finished jobs, 2 -
+
+remove active jobs too, 3- also remove everything that looks like
+junk.
 
@@
 performed. Pictures a) corresponds to the situation right after all
-input files have been gathered in the session directory and show
-the actions taken right after the job's script starts. Pictures b)
-shows   the situation while the job is running and the actions
-which are taken right after it has finished. Pictures c)  illustrates
-the final situation, when the job's output files are ready to be uploaded
-to an external storage element or be downloaded by the user.
+input files have been gathered in the session directory and show the
+actions taken right after the job's script starts. Pictures b) shows
+the situation while the job is running and the actions which are taken
+right after it has finished. Pictures c) illustrates the final situation,
+when the job's output files are ready to be uploaded to an external
+storage element or be downloaded by the user.
 
@@
 \caption{\label{fig:node scratch, not vis on front}RUNTIME\_LOCAL\_SCRATCH\_DIR
-is set to  the location of the scratch directory
-on the computing node, RUNTIME\_FRONTEND\_SEES\_NODE is undefined.}
+is set to the location of the scratch directory on the computing node,
+RUNTIME\_FRONTEND\_SEES\_NODE is undefined.}
 
@@
 \caption{\label{fig:node scratch, vis on front}Both RUNTIME\_LOCAL\_SCRATCH\_DIR
-and RUNTIME\_FRONTEND\_SEES\_NODE are set to the  location
-of the scratch directory on the computing node and    the
-location where this scratch directory is accessible from the frontend,
-respectively .}
+and RUNTIME\_FRONTEND\_SEES\_NODE are set to the location of the scratch
+directory on the computing node and the location where this scratch
+directory is accessible from the frontend, respectively.}
 
@@
 
-\subsection{ Running as non-root}
+\subsection{Running as non-root}
 
 Bot GM and GFS are primarily designed to be run by the \emph{root}
-UNIX account and serve multiple    global Grid identities
-mapped to several UNIX accounts. Nevertheless it is possible to use
-\emph{non-root} accounts to run these services   
-at the cost of some functionality loss as described below.
-
-There are no drawbacks of running GFS with \emph{gaclplugin}
-or \emph{fileplugin} under a \emph{non-root} account as long as the
-only Grid identity used is that of the user   who runs the services
-and all served files and directories are owned by the server's account.
-
-  In order for GM and GFS with \emph{jobplugin} to
-be able to cooperate, both services must be run either by the same
-non-root account or one of the services must be run under the \emph{root}
-account. This is needed because the two services communicate over
-a local filesystem, hence they must have \emph{full} access to the
-same set of files.
-
-As long as GFS with \emph{jobplugin} is run under a non-root account
- , no mapping from Grid identity to local UNIX account takes
-place. All alowed Grid users are assigned the server's account and
-are then processes by GM using same account. The only way to overcome
-this limitation is to run one GFS per local account with proper access
-control configured.
-
-Because GM has to impersonate a user's local account while
-communicating with the LRMS, it can serve only the account it is
-run under (unless it is run under the \emph{root} account, of course).
-As in the case of GFS, multiple instances of GM may be run, one
-per local account. This solution   has
-some drawbacks. The GM looses possibility to share cached files among
-serviced users. In addition, it is not  possible to control the
-load on a frontend by limiting the number of simultatenuosly running
-\emph{downloader} and \emph{uploader} modules.
-
-One  also has to take into account that the private part of the
-GSI infrastructure (private key of a host at least) has to be duplicated
+UNIX account and serve multiple global Grid identities mapped to several
+UNIX accounts. Nevertheless it is possible to use \emph{non-root}
+accounts to run these services at the cost of some functionality loss
+as described below.
+
+There are no drawbacks of running GFS with \emph{gaclplugin} or \emph{fileplugin}
+under a \emph{non-root} account as long as the only Grid identity
+used is that of the user who runs the services and all served files
+and directories are owned by the server's account.
+
+In order for GM and GFS with \emph{jobplugin} to be able to cooperate,
+both services must be run either by the same non-root account or one
+of the services must be run under the \emph{root} account. This is
+needed because the two services communicate over a local filesystem,
+hence they must have \emph{full} access to the same set of files.
+
+As long as GFS with \emph{jobplugin} is run under a non-root account,
+no mapping from Grid identity to local UNIX account takes place. All
+alowed Grid users are assigned the server's account and are then processes
+by GM using same account. The only way to overcome this limitation
+is to run one GFS per local account with proper access control configured.
+
+Because GM has to impersonate a user's local account while communicating
+with the LRMS, it can serve only the account it is run under (unless
+it is run under the \emph{root} account, of course). As in the case
+of GFS, multiple instances of GM may be run, one per local account.
+This solution has some drawbacks. The GM looses possibility to share
+cached files among serviced users. In addition, it is not possible
+to control the load on a frontend by limiting the number of simultatenuosly
+running \emph{downloader} and \emph{uploader} modules.
+
+One also has to take into account that the private part of the GSI
+infrastructure (private key of a host at least) has to be duplicated
 for every account used to run GFS.
@@
 client will see virtual directories representing the jobs belonging
-to    him/her. Directory names are job identifiers,
-each representing one job      
-  . These directories are directly connected to
-session directories of jobs and contain the  files and subdirectories
-that are visible on the frontend.
+to him/her. Directory names are job identifiers, each representing
+one job. These directories are directly connected to session directories
+of jobs and contain the files and subdirectories that are visible
+on the frontend.
 
 If a job's xRSL has \emph{gmlog} specified, then the job's directory
-also contains a virtual subdirectory  of that name 
- holding files with information about the job as created by
-GM. The    'errors' and 'status' files are especially
-usefull for troubleshooting. 'errors' contains the stderr output of
-the various modules run by GM   during job processing
+also contains a virtual subdirectory of that name holding files with
+information about the job as created by GM. The 'errors' and 'status'
+files are especially usefull for troubleshooting. 'errors' contains
+the stderr output of the various modules run by GM during job processing
 stages (downloader, uploader, job's submission to LRMS). 'status'
-contains one word representing the job's state  .
+contains one word representing the job's state.
 
 Also under the jobplugin's mount point there is an additional directory
-named \char`\"{}new\char`\"{} -- used to submit new jobs -- and
-another directory named ''info''. The latter has subdirectories
-named after job ids  , each of which contains
-files with information about a job. These are the same files 
- that can be accessed in the subdirectory specified through
-\emph{gmlog}.
+named \char`\"{}new\char`\"{} -- used to submit new jobs -- and another
+directory named ''info''. The latter has subdirectories named after
+job ids, each of which contains files with information about a job.
+These are the same files that can be accessed in the subdirectory
+specified through \emph{gmlog}.
 
@@
 
-Each xRSL put into the {}``new'' directory by a gridftp client
-is treated as a new job's description. GFS's jobplugin parses
-the xRSL and returns the client a positive response if there
-were no errors in the request.
+Each xRSL put into the {}``new'' directory by a gridftp client is
+treated as a new job's description. GFS's jobplugin parses the xRSL
+and returns the client a positive response if there were no errors
+in the request.
 
@@
 name appears. If the job's description contains input files which
-should be transferred from the client's machine, the client
-must upload them to that directory under specified names.
-
-   The job identifier reserved by GFS to a new job must
-be somehow communicated back to the client. Within the bounds of the
-FTP protocol, this is achieved in the following way. Prior to uploading
-the xRSL, the client issues a CWD command to change the current directory
-to {}``new''. In turn, the server responds with a redirect to the
-new session directory named after the reserved identifier. The client
-has now found out the job's id, and proceeds with uploading the xRLS.
+should be transferred from the client's machine, the client must upload
+them to that directory under specified names.
+
+The job identifier reserved by GFS to a new job must be somehow communicated
+back to the client. Within the bounds of the FTP protocol, this is
+achieved in the following way. Prior to uploading the xRSL, the client
+issues a CWD command to change the current directory to {}``new''.
+In turn, the server responds with a redirect to the new session directory
+named after the reserved identifier. The client has now found out
+the job's id, and proceeds with uploading the xRLS.
 
@@
 by uploading special xRSL files into directory {}``new''. Such an
-  xRSL may consist of only 2 parameters - \emph{action} for
-action to be performed, and \emph{jobid} to identify the job to be
-affected.  All other parameters are ignored.
+xRSL may consist of only 2 parameters - \emph{action} for action to
+be performed, and \emph{jobid} to identify the job to be affected.
+All other parameters are ignored.
 

### 2008-08-16 doc/a-rex/arex_tech_doc.tex
More changes to A-REX doc.

@@
 \begin{center}
-\textsc{\large The ARC Computational Job Management Component - A-Rex}
+\textsc{\large The ARC Computational Job Management Component - A-REX}
 \par\end{center}{\large \par}
@@
 
-A-Rex is one of ARC middleware componets implementing functions of
+A-REX is one of ARC middleware componets implementing functions of
 so called \emph{Computing Element} (CE).
 
-The aim of the A-Rex is to take care of job pre- and post-processing.
+The aim of the A-REX is to take care of job pre- and post-processing.
 It provides functionality to stage-in files containing input data
@@
 
-The A-Rex implements Web Service (WS) interface which provides a way
+The A-REX implements Web Service (WS) interface which provides a way
 to submit and control computational tasks (jobs) to be executed by
-the A-Rex and undelying Local Resource Management System.
+the A-REX and undelying Local Resource Management System.
 
@@
 (SD). Input files are gathered in the SD. The job is supposed to produce
-new data files in the SD as well. The A-Rex does not guarantee the
+new data files in the SD as well. The A-REX does not guarantee the
 availability of any other places accessible by the job other than
 SD (unless such a place is part of a requested Runtime Environement).
-The SD is also the only place which is controlled by the A-Rex. It
+The SD is also the only place which is controlled by the A-REX. It
 is accessible by the user from outside through HTTP(S) protocol. Any
-file created outside the SD is not controlled by the A-Rex. Any exchange
-of data between client and A-Rex (including also program modules)
+file created outside the SD is not controlled by the A-REX. Any exchange
+of data between client and A-REX (including also program modules)
 is done through HTTP(S) protocol. A URL for accessing input/output
 files is  obtained through the WS Local Information Description Interface
-(LIDI) of A-Rex.
+(LIDI) of A-REX.
 
 Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
-\cite{jsdl} XML document - which identifies the job in the A-Rex
+\cite{jsdl} XML document - which identifies the job in the A-REX
 and the Information Interface.
@@
 Jobs are initiated and controlled through WS interface. Complete job
-descriptions (JD)are passed to the A-Rex through WS in JSDL\cite{jsdl}
+descriptions (JD)are passed to the A-REX through WS in JSDL\cite{jsdl}
 coded description. Input data files and job executables are transferred
@@
 
-One of the most important tasks of the A-Rex is to take care of processing
+One of the most important tasks of the A-REX is to take care of processing
 input and output data (files) of the job. Input files are gathered
@@
 \begin{itemize}
-\item Download is initiated by the A-Rex -- This is the case for files defined
-in the JD (with name and source). The A-Rex alone is responsible to
+\item Download is initiated by the A-REX -- This is the case for files defined
+in the JD (with name and source). The A-REX alone is responsible to
 make sure that all required files will be available in the SD.\\
@@
 some less standard sources are supported. Those are described below.
-The A-Rex fully relies on HED framework \cite{hed} for data transfering
+The A-REX fully relies on HED framework \cite{hed} for data transfering
 capabilities. Hence actual set of supported protocols depends on installed
@@
 submission of JD, UI can (and should) use that to upload data files
-which are not otherwise accessible by the A-Rex. Examples of such
+which are not otherwise accessible by the A-REX. Examples of such
 files are the main executable of the job, the job's input files, etc.
@@
 \textbf{There is no} other reliable way for a job to obtain input
-data on the CE based on the A-Rex. Access to AFS, NFS, FTP, HTTP and
+data on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and
 any other remote data transport during execution of a job is not guaranteed
-(at least not by A-Rex).
+(at least not by A-REX).
 
@@
 and optionally registered in some \emph{Indexing Service} like the
-Globus \emph{Replica Location Service} (RLS) -- The A-Rex takes care
+Globus \emph{Replica Location Service} (RLS) -- The A-REX takes care
 of these files. They have to be specified in the JD. If the job fails
@@
 
-From the point of view of the A-Rex a job passes through various states.
+From the point of view of the A-REX a job passes through various states.
 Picture \ref{job states diagram} presents a diagram of the possible
@@
 A user can examine the state of a job by querying the dedicated Local
-Information Description Interface of A-Rex using the UI or any other
+Information Description Interface of A-REX using the UI or any other
 suitable tool or through query method of WS interface.
@@
 
-Below is the description of all actions taken by the A-Rex at every
+Below is the description of all actions taken by the A-REX at every
 state:
@@
 \item \textbf{Accepted} - In this state the job has been submitted to a
-CE but is not processed yet. The A-Rex will analyze the JD and move
+CE but is not processed yet. The A-REX will analyze the JD and move
 to the next stage. If the JD can not be processed the job will be
@@
 \item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
-The A-Rex is downloading the files specified in the JD and is waiting
+The A-REX is downloading the files specified in the JD and is waiting
 for files which are supposed to be uploaded by the UI. If all files
@@
 \item \textbf{Executing} - The job is queued or being executed in the LRMS.
-The A-Rex takes no actions except waiting until the job finishes.
+The A-REX takes no actions except waiting until the job finishes.
 \item \textbf{Killing} - Necessary action to cancel the job in the LRMS
@@
 in this state.
-\item \textbf{Finished} - No more processing is performed by the A-Rex.
+\item \textbf{Finished} - No more processing is performed by the A-REX.
 The user can continue to download data files from the SD. The SD is
@@
 is moved to the state \textbf{Deleted}. The 'deletion' time can be
-obtained by querying the Informatrion Interface of the A-Rex. If a
+obtained by querying the Informatrion Interface of the A-REX. If a
 job was moved to \textbf{Finished} because of failure, it may be restarted
@@
 
-In full installation the A-Rex and it's components support the following
+In full installation the A-REX and it's components support the following
 data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
@@
 
-\subsection{Internal Files of the A-Rex}
-
-For each local UNIX user listed in the A-Rex configuration - including
+\subsection{Internal Files of the A-REX}
+
+For each local UNIX user listed in the A-REX configuration - including
 a generic one which covers all local user identities - a \textit{control
-directory} exists. In this directory the A-Rex stores information
+directory} exists. In this directory the A-REX stores information
 about jobs belonging to that user. Multiple users can share the same
-\textit{control directory}. In most common configuration case A-Rex
+\textit{control directory}. In most common configuration case A-REX
 serves all users defined by Operating System and stores their control
 file in same directory. To make it easier to recover in case of failure,
-the A-Rex stores most information in files rather than in memory.
+the A-REX stores most information in files rather than in memory.
 All files belonging to the same job have names starting with \textbf{job.ID.,}
@@
 (JD).
-\item \textit{job.ID.local} - information about the job used by the A-Rex.
+\item \textit{job.ID.local} - information about the job used by the A-REX.
 It consists of lines of format \textit{{}``name = value''}. Not
@@
 There are other files with names like job.ID.{*} which are created
-and used by different parts of the A-Rex. Their presence in the \textit{control
+and used by different parts of the A-REX. Their presence in the \textit{control
 directory} can not be guaranteed and can change depending on changes
-in the A-Rex code.
+in the A-REX code.
 
@@
 and thus defines operations to start and stop the functionality of
-the BES service. The A-Rex does not implement remote control of service
+the BES service. The A-REX does not implement remote control of service
 functionality. Hence the BES-Management port-type is not functional.
@@
 create an activity in terms of BES) and to monitor its state. It also
-has an ability to provide information about the service. A-Rex fully
+has an ability to provide information about the service. A-REX fully
 implements the functionality of this port-type.
 
-For job descriptions A-Rex accepts the Job Submission Description
+For job descriptions A-REX accepts the Job Submission Description
 Language (JSDL)\cite{jsdl} documents as defined by the OGF Job Submission
@@
 
-A-Rex introduces a new operation in addition to those provided by
+A-REX introduces a new operation in addition to those provided by
 BES. It does that by defining its own port-type with the single operation
@@
 Since BES allows implementators to extend their initial activity states
-with additional sub-states, A-Rex defines a set of sub-states of activity
+with additional sub-states, A-REX defines a set of sub-states of activity
 processing in addition to those defined by the BES, as listed in Table
@@
 
-The A-Rex also supportd the Delegation Interface (see Appendix E).
+The A-REX also supportd the Delegation Interface (see Appendix E).
 This is a common purpose interface to be used by ARC services which
@@
 \end{itemize}
-Additionally, A-Rex Web Service Interface allows delegation to be
+Additionally, A-REX Web Service Interface allows delegation to be
 performed as part of the \emph{CreateActivity} operation of the BES-Factory
@@
 in a \emph{WS-Resource Properties} informational XML document. The
-document type is defined in the A-Rex WSDL as a \emph{ResourceInformationDocumentType}.
+document type is defined in the A-REX WSDL as a \emph{ResourceInformationDocumentType}.
 It contains the following elements/resources:
@@
 A-REX supports the following elements from the JSDL version 1.0 specification\cite{jsdl}
-including POSIX Applications extension:
+including POSIX Applications extension and JSDL HPC Profile Application
+Extension {[}]:
 
@@
 \item [{\emph{JobName}}] \emph{-} name of the job as assigned by the user.
-\item [{\emph{Executable}\ (POSIX)}] - name of the executable file.
-\item [{\emph{Argument}\ (POSIX)}] - arguments the executable will be
-launched with.
+\item [{\emph{Executable}\ (POSIX,HPC)}] - name of the executable file.
+\item [{\emph{Argument}\ (POSIX,HPC)}] - arguments the executable will
+be launched with.
 \item [{\emph{DataStaging}}]~
@@
 \end{lyxlist}
-\item [{\emph{Input}\ (POSIX)}] - file to be used as standard input for
-the executable.
-\item [{\emph{Output}\ (POSIX)}] - file to be used as standard output
+\item [{\emph{Input}\ (POSIX,HPC)}] - file to be used as standard input
 for the executable.
-\item [{\emph{Error}\ (POSIX)}] - file to be used as standard error for
-the executable.
+\item [{\emph{Output}\ (POSIX,HPC)}] - file to be used as standard output
+for the executable.
+\item [{\emph{Error}\ (POSIX,HPC)}] - file to be used as standard error
+for the executable.
 \item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
@@
 \item [{\emph{IndividualCPUCount}}] - same as TotalCPUCount.
+\item [{\emph{Environment}\ (POSIX)}] - environement variable.
 \end{lyxlist}
@@
 
-The GM can cache input files. Caching is enabled if corresponding
-command is present in configuration file. The GM does not cache files
-marked as executable in job. Caching can also be explicitly turned
-off by user for each file by using \textit{cache=no} option in URL
-(for URL options read {}``Protocols, Uniform Resource Locators (URL)
-and extensions supported in ARC'' \cite{urls}). The disc space occupied
-by cache is controlled by removing unused files. For more information
-look in section \ref{SubSection: ConfigFile}.
+The A-REX can cache input files. Caching is enabled if corresponding
+command is present in configuration file. The A-REX does not cache
+files marked as executable in job. Caching can also be explicitly
+turned off by user for each file by using \textit{cache=no} option
+in URL (for URL options read {}``Protocols, Uniform Resource Locators
+(URL) and extensions supported in ARC'' \cite{urls}). The disc space
+occupied by cache is controlled by removing unused files. For more
+information look in section \ref{SubSection: ConfigFile}.
+
+\emph{NOTE: The caching algorithm used by A-REX and described above
+is currently being replaced by a better one which is no robust and
+provides better scalability.}
 
@@
 
-Before downloading a file the GM tries to determine it's size and
+Before downloading a file the A-REX tries to determine it's size and
 then tries to preallocate space in the cache directory by writing
-a file of the same size. If this fails (possibly because the file
-system has no more space), it tries to the remove oldest cached files
-which are not being used by any job. This means that \textbf{the hard
-limit of cache size is the space available in the file-system}. In
-case cache gets full and it is impossible to free enough space, the
-download fails and then is retried without using cache.
-
-Before giving access to an already cached file, the GM contacts the
-original file source to check if the user has suficient access rights
-to the file. Not all protocols make this check possible. Also, file
-creation or validity times are checked to make sure the cached file
-is fresh. If it is impossible to obtain creation and invalidation
+an empty file of the same size. If this fails (possibly because the
+file system has no more space), it tries to the remove oldest cached
+files which are not being used by any job. This means that \textbf{the
+real hard limit of cache size is the space available in the file-system}.
+In case cache gets full and it is impossible to free enough space,
+the download fails and then is retried without using cache.
+
+Before giving access to an already cached file, the A-REX contacts
+the original file source to check if the user has suficient access
+rights to the file. Not all protocols make this check possible. Also,
+file creation or validity times are checked to make sure the cached
+file is fresh. If it is impossible to obtain creation and invalidation
 times, the file is invalidated 24 hours after it was downloaded.
 
-The GM checks the cache periodically. If the space used by cache exceeds
-the high water-mark given in the configuration file (\emph{softsize}),
+The A-REX checks the cache periodically. If the space used by cache
+exceeds the high water-mark given in the configuration file (\emph{softsize}),
 it tries to remove the oldest unused files until cache size drops
@@
 There are two kinds of caches supported. Files in \textit{private}
-cache are owned by the Unix user to which a grid user is mapped. Those
+cache are owned by the Unix user to which a Grid user is mapped. Those
 files are readable only by that particular Unix user. Another kind
 of cache is \textit{shared}. Files in the shared cache are owned by
-the Unix user who started GM and are readable by everyone.
+the Unix user who started A-REX and are readable by everyone.
 
@@
 
-The GM consists of several separate executable modules. These are:
-
-\begin{itemize}
-\item \textit{grid-manager} - The main module. It is responsible for processing
-jobs, moving them through states and running other modules.
+The A-REX consists of several separate modules. These are:
+
+\begin{itemize}
+\item \textit{libarex.so} - The main module providing main functionality
+and web interface. It is implemented as HTTP and SOAP service inside
+HED. It is responsible for processing jobs, moving them through states
+and running other modules.
 \item \textit{downloader} - This is a module responsible for gathering input
@@
 file.
-\item \emph{gm-kick} - Sends a signal to the GM though a FIFO file to wake
-it up. It's used to increase responsiveness of GM.
+\item \emph{gm-kick} - Sends a signal to the A-REX though a FIFO file to
+wake it up. It's used to increase responsiveness of A-REX.
+\item \emph{cluster+qju.pl} - Collects and generates information about computing
+resource as XML document in NoruGrid and Glue 2 format.
 \end{itemize}
 The following modules are always run under the Unix account to which
-a grid user is mapped.
+a Grid user is mapped.
 
@@
 \item \textit{scan-{*}-job} -This shell script is responsible for notifying
-the GM about completion of jobs. It's implementation for PBS uses
+the A-REX about completion of jobs. It's implementation for PBS uses
 server logs to extract information about jobs. If logs are not available
@@
 \end{itemize}
-GM comes with plugins useable for various authorization purposes (see
-for example the description of \emph{authplugin} command below):
-
-\begin{itemize}
-\item \textit{inputcheck} - checks if all input files specified in job description
-are downloadable.\\
-\hspace*{0.5cm}inputcheck {[}-h] {[}-d debug\_level] RSL\_fle {[}proxy\_file]\\
-\hspace*{0.5cm}- RSL\_file -- file with job description,\\
-\hspace*{0.5cm}- proxy\_file -- credentials proxy.
-\item \textit{lcas} - executes LCAS plugins on credentials and returns 0
-if authorization passed.\\
-\hspace*{0.5cm} lcas credentials description {[}library {[}db {[}directory]]]\\
-\hspace*{0.5cm}- credentials -- path to file with credentials to
-authorize,\\
-\hspace*{0.5cm}- description -- path to file with job description,\\
-\hspace*{0.5cm}- library -- path to LCAS library (full or relative
-to LCAS directory),\\
-\hspace*{0.5cm}- db -- path to LCAS DB file (full or relative to
-LCAS directory),\\
-\hspace*{0.5cm}- directory -- LCAS directory.
-\end{itemize}
 
@@
 
-The GM is installed into a single installation point referred as \$NORDUGRID\_LOCATION
-and the following sub-directories are used:
+The A-REX is installed into a single installation point referred as
+\$ARC\_LOCATION and the following sub-directories are used:
 
 \begin{LyXParagraphLeftIndent}{1cm}
-\$NORDUGRID\_LOCATION/bin -- tools
-
-\$NORDUGRID\_LOCATION/libexec -- program modules used by GM
-
-\$NORDUGRID\_LOCATION/etc -- configuration files, deprecated, central
-configuration file is used by default
-
-\$NORDUGRID\_LOCATION/sbin -- daemons
-
-\$NORDUGRID\_LOCATION/lib -- gridftp server's plugins and API libraries
+\$ARC\_LOCATION/bin -- tools
+
+\$ARC\_LOCATION/libexec -- program modules used by A-REX
+
+/etc -- central configuration file - location used by default
+
+\$ARC\_LOCATION/lib/arc -- service module
 
 \end{LyXParagraphLeftIndent}
-The GM also uses following directories:
-
-\begin{itemize}
-\item \textit{session root directory} - Tthis is the directory under which
+The A-REX also uses following directories:
+
+\begin{itemize}
+\item \textit{session root directory} - Tthis is the directory in which
 a user's SDs are created. It's location is configurable per UNIX user.
-Several (or even all) users may share the same session root directory..\\
-There are 2 processes which need to have permissions to create new
-files and directories in it: GM and GFS. \\
-If any of these processes are run under a dedicated user account,
-that account needs full permissions in the \textit{session root directory}.\\
-If these processes are run under the \emph{root} account, make sure
-\textit{session root directory} resides on a filesystem which does
-not limit the capabilities of the \emph{root} user (as does for example
-NFS with \emph{root\_squash} option).\\
-If there is a need to run processes under the \emph{root} account
-(to be able to run jobs in LRMS under different users' accounts, for
-example) but there is no way to provide a suitable \textit{session
-root directory,} use the \emph{norootpower} command in GM's configuration
-file. In that case GM and GFS will use the identity of the local user
-to which a Grid identity is mapped to access the \textit{session root
-directory}. Hence those users will need full access there.\\
-The GM creates SDs with proper ownership and permissions for the local
-identity used to run a job. Some filesystems require users to have
-\emph{execute} permission on the \textit{session root directory} in
-order to access any file or subdirectory there.\\
+Several (or even all) users may share the same session root directory.\\
+The A-REX need to have permission to create new files and directories
+in the session root directory. If A-REX is run under a dedicated user
+account, that account needs full permissions in the \textit{session
+root directory}.\\
+If A-REX is run under the \emph{root} account, make sure \textit{session
+root directory} resides on a filesystem which does not limit the capabilities
+of the \emph{root} user (as does for example NFS with \emph{root\_squash}
+option).\\
+If there is a need to run A-REX under the \emph{root} account (to
+be able to run jobs in LRMS under different users' accounts, for example)
+but there is no way to provide a suitable \textit{session root directory,}
+use the \emph{norootpower} command in configuration file. In that
+case A-REX will use the identity of the local user to which a Grid
+identity is mapped to access the \textit{session root directory}.
+Hence those users will need full access there.\\
+The A-REX creates SDs with proper ownership and permissions for the
+local identity used to run a job. Some filesystems require users to
+have \emph{execute} permission on the \textit{session root directory}
+in order to access any file or subdirectory there.\\
 In order for jobs to access their input files, session root directories
@@
 see section \ref{sub:LRMS}.
-\item \textit{control directory} - In this directoryGM and GFS store an
-information about accepted jobs. Both processes must have full permissions
-there.\\
+\item \textit{control directory} - In this directory A-REX stores an information
+about accepted jobs. It must have full permissions there.\\
 A subdirectory called \textit{log} is created there. It is used to
@@
 
-\subsection{\label{SubSection: ConfigFile}Configuration of the Grid Manager}
-
-The GM configuration is done through a single configuration file.
-Historically GM supports 2 kinds of configuration files. For old one
-(deprecated) it looks at following places:
-
-\begin{itemize}
-\item \textit{\$NORDUGRID\_LOCATION/etc/grid-manager.conf}
-\item \textit{/etc/grid-manager.conf}
-\end{itemize}
-And for new one in
-
-\begin{itemize}
-\item \textit{/etc/arc.conf}
-\end{itemize}
-The old configuration file consists of empty lines, lines containing
-comment (line starts from \#) or configuration commands. Blank spaces
-in arguments must be escaped using '\textbackslash{}' or arguments
-must be enclosed in '\char`\"{}'. Command line starts from command
-followed by arguments separated from command and between them by spaces.
-
-The new configuration file can also contain empty lines and comments
+\subsection{\label{SubSection: ConfigFile}Configuration of the A-REX}
+
+Due to historical resons configuration of the A-REX is split into
+2 parts. The small part is located inside HED configuration (see appendix
+B) for schema and description of supported elements. It refers to
+file containing main part of configuration. The default location of
+main configuration file is \textit{/etc/arc.conf .}
+
+The configuration file can contain empty lines and comments in lines
 starting from \#. It is separated into sections. Each sections starts
-from string containing
+from string containing pattern
 
@@
 \end{itemize}
-Each section continues till next section of end of file. One configuration
+Each section continues till next section or end of file. One configuration
 file can have commands for multiple services/modules/programs. Each
-service get it's own section named after it. The GM uses section \emph{{[}grid-manager]}.
-Some services can make use of multiple subsections to reflect their
-internal modular structure. Commands in section \emph{{[}common]}
+service get it's own section named after it. The A-REX uses section
+\emph{{[}grid-manager]}. Some services can make use of multiple subsections
+to reflect their internal modular structure. Commands in section \emph{{[}common]}
 apply to all services. Command lines have format
@@
 \end{itemize}
-Names are same as in old configuration file. The \emph{argument string}
-consists of same arguments as in old format. And they must obey same
-rules.
-
-Both files support almost same commands. Following commands are defined
-(examples are given for new format):
-
-\noindent Global commads (those which affect global parameters of
-the GM and affect all serviced users, also described in \cite{config}):
-
-\begin{itemize}
-\item \textbf{\textit{daemon}}\textit{=yes|no} - specifies whether the GM
-should go to background after started. Defaults to \emph{yes}.
-\item \textbf{\textit{logfile}}\textit{={[}path]} - specifies name of file
-for logging debug/informational output. Defaults to \emph{/dev/null}
-for daemon mode and \emph{stderr} for foreground mode.
-\item \textbf{\textit{user}}\textit{={[}uid{[}:gid]]} - specifies user id
-(and optionally group id) to which the GM must switch after reading
-configuration. Defaults to \emph{not switch}.
-\item \textbf{\textit{pidfile}}\textit{={[}path]} - specifies file where
-id if GM process will be stored. Defaults to \emph{not write}.
-\item \textbf{\textit{debug}}\textit{=number} - specifies level of debug
-information. More information is printed for higher levels. Currently
-highest effective number is 3 and lowest 0. Defaults to 2.
-\end{itemize}
-All commands above are generic for every daemon-enabled server in
-ARC NorduGrid toolkit (like GFS and HTTPSD).
+Following commands are defined:
 
@@
 \item \textbf{\textit{jobreport}}\textit{={[}URL ... number]} - specifies
-that GM has to report information about jobs being processed (started,
+that A-REX has to report information about jobs being processed (started,
 finished) to centralized service running at given \textit{URL}. Multiple
 entries and multiple URLs are allowed. \textit{number} specifies how
-long old records have to be kept if failed to be reported. That time
-is in days. Last specified value becomes effective.
+long old records have to be kept if failed to be reported. That value
+is specified in days. Last specified value becomes effective.
 \item \textbf{\textit{securetransfer}}\textit{=yes|no} - specifies whether
 to use encryption while transferring data. Currently works for GridFTP
-only. Default is no. It is overridden by value specified in URL options.
+only. Default is \emph{no}. It is overridden by value specified in
+URL options.
 \item \textbf{\textit{localtransfer}}\textit{=yes|no} - specifies whether
 to pass file downloading/uploading task to computing node. If set
-to yes the GM won't download/upload files. Instead it composes script
-submitted to LRMS in way to make it do that. This requires installation
-of GM and Globus to be accessible from computing nodes and environment
-variables GLOBUS\_LOCATION and NORDUGRID\_LOCATION to be set accordingly.
-Default is no.
+to yes the A-REX won't initiate download/upload files. Instead it
+composes script submitted to LRMS in way to make it do that. This
+requires installation of A-REX and all realted software to be accessible
+from computing nodes and environment variable ARC\_LOCATION to be
+set accordingly. Default is \emph{no}.
 \item \textbf{\textit{maxjobs}}\textit{={[}max\_processed\_jobs {[}max\_running\_jobs]]}
-- specifies maximum number of jobs being processed by the GM at different
-stages:\\
+- specifies maximum number of jobs being processed by the A-REX at
+different stages:\\
 \textit{max\_processed\_jobs} - maximal amount of jobs being processed
-by GM. This does not limit amount of jobs, which can be suNOTE:bmitted
-to cluster\\
+by A-REX. This does not limit amount of jobs, which can be submitted
+to A-REX\\
 \textit{max\_running jobs} - maximal amount of jobs passed to Local
@@
 \item \textbf{\textit{wakeupperiod}}\textit{=time} - specifies how often
-for external changes are performed (like new arrived job, job finished
-in LRMS, etc.). \textit{time} is a minimal time period specified in
-seconds. Default is 3 minutes.
+the A-REX checks for job state changes (like new arrived job, job
+finished in LRMS, etc.). \textit{time} is a minimal time period specified
+in seconds. Default is \emph{3 minutes}.
 \item \textbf{\textit{cacheregistration}}\textit{=yes|no} - enables or disables
 registration of cache data into Indexing Services like RC or RLS.
-The default is \emph{no}. Only files dowmloaded through \emph{meta-url}
+The default is \emph{no}. Only files downloaded through \emph{meta-url}
 are registred. Registration is done to same service used for obtaining
-information about file. For this opeartion credentials of the GM (host
-key and certificate) are used. If required new files storage location
-is registered at Indexing Service with quasi-url \emph{cache://hostname/}
+information about file. For this opeartion credentials of the A-REX
+(host key and certificate) are used. If required new files storage
+location is registered at Indexing Service with quasi-url \emph{cache://hostname/}
 and name \emph{hostname:cache} .
@@
 code is not 0 current operation will fail.
-\item \textbf{\textit{norootpower}}\textit{=yes/no} - if set to yes all
-processes involved in job management will use local identity of a
-user to which Grid identity is mapped in order to access filesystem
+\item \textbf{\textit{norootpower}}\textit{=yes/no} - if set to \emph{yes}
+all processes involved in job management will use local identity of
+a user to which Grid identity is mapped in order to access filesystem
 at path specified in \textbf{\textit{session}} command (see below).
 Sometimes this may involve running temporary external process.
-\item \textbf{\textit{allowsubmit}}\textit{={[}group ...]} - list of authorization
-groups of users allowed to submit new jobs while \char`\"{}allownew=no\char`\"{}
-is active in \emph{jobplugin.so} configuration (see below in section
-\ref{subsec: configfile of the gfs}). Multiple commands are allowed.
 \item \textbf{\textit{speedcontrol}}\textit{=min\_speed min\_time min\_average\_speed
@@
 NOTE: URLs which fit into \textit{copyurl} or \textit{linkurl} are
-treated as more easily accessible than other URLs. That means if GM
+treated as more easily accessible than other URLs. That means if A-REX
 has to choose between few URLs from which should it download input
@@
 is called \textbf{\textit{sessiondir}}.
-\item \textbf{\textit{cache}}\textit{dir=path {[}link\_path]} - specifies
+\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} - specifies
 the directory to store cached data. Empty \emph{path} disables caching.
@@
 is set to '.' files are not soft-linked, but copied to session directory.
-In old configuration file this command is called \textbf{\textit{cache}}.
 \item \textbf{\textit{privatecache}}\textit{=path {[}link\_path]} - same
@@
 \item \textbf{\textit{maxrerun}}\textit{=number} - specifies maximal number
-of times job will be allowed to rerun after it failed in LRMS. Default
-value is 2. This only specifies a upper limit. Actual number is provided
-in job description and defaults to 0.
+of times job will be allowed to rerun after it failed at any stage.
+Default value is \emph{5}. This only specifies a upper limit. Actual
+number is provided in job description and defaults to 0.
 \end{itemize}
@@
 \item \textbf{\textit{control}}\textit{=path username {[}username {[}...]]}
-- This option initiates UNIX user as being serviced by the GM. \textit{path}
-refers to the control directory (see section \ref{section: internals}
+- This option initiates UNIX user as being serviced by the A-REX.
+\textit{path} refers to the control directory (see section \ref{section: internals}
 for the description of control directory). If the path is {*} the
@@
 directory will be used for \textbf{any} user. This option should be
-the last one in the configuration file. In new configuration file
-command \textbf{\textit{controldir}}\textit{=path} is also available.
+the last one in the configuration file. There is also command \textbf{\textit{controldir}}\textit{=path}.
 It uses special username '.' and is always executed last independent
@@
 \end{itemize}
-Following are global commands supported only in new configuration
-file. Most of them are specific to underlying LRMS (PBS in this case)
-and are passed in environment variables if old configuration file
-is used.
+Following are global commands specific to communication with underlying
+LRMS (PBS in this case).
 
@@
 contains \emph{runtimenvironment} scripts.
-\item \textbf{\textit{shared\_filesystem}}\textit{=yes|no} - if compiting
+\item \textbf{\textit{shared\_filesystem}}\textit{=yes|no} - if computing
 nodes have an access to session directory through a shared filesystem
-like NFS. Corresponds to an environement variable RUNTIME\_NODE\_SEES\_FRONTEND.
+like NFS. Corresponds to an environement variable RUNTIME\_NODE\_SEES\_FRONTEND
+(see section \ref{sub:LRMS}).
 \item \textbf{\textit{nodename}}\textit{=command} - command to obtain hostname
@@
 where \textbf{\textit{scratchdir}} can be found.
-\item \textbf{\textit{nodename}}\textit{=command} - command to obtain hostname
-of computing node.
 \end{itemize}
@@
 \end{lyxlist}
-Some configuration parameters can be specified from command line while
-starting the GM:
-
-\emph{grid-manager {[}-h] {[}-C level] {[}-d level] {[}-c path] {[}-F]
-{[}-U uid{[}:gid]] {[}-L path] {[}-P path]}\\
-\hspace*{0.5cm}\emph{-h} - short help,\\
-\hspace*{0.5cm}\emph{-d} - debug level,\\
-\hspace*{0.5cm}\emph{-L} - name log file (overwrites value in configuration
-file),\\
-\hspace*{0.5cm}\emph{-P} - name for file containing process id (overwrites
-value in configuration file),\\
-\hspace*{0.5cm}\emph{-U} - user and gropu id to use for running daemon,\\
-\hspace*{0.5cm}\emph{-F} - do not make process daemon,\\
-\hspace*{0.5cm}\emph{-c} - name od configuration file,\\
-\hspace*{0.5cm}\emph{-C} - remove
-
-old information before starting:
-
-1- remove finished jobs, 2 -
-
-remove active jobs too, 3- also remove everything that looks like
-junk.
 
@@
 
-Authorization is performed at GFS by applying set of rules. Each rule
-takes one line in the \emph{group} section. For information about
-supported rules please read {}``Configuration and authorisation of
-ARC (NorduGrid) Services'' \cite{config}.
+Authorization is performed by generic means provided by HED framework.
+Currently A-REX does not implement any internal authorization tecniques.
 
@@
 The GM comes with support for several LRMS. And this number is slowly
-growing. Features explained below are for \textbf{PBS} backend. This
-support is provided through \textit{submit-pbs-job}, \textit{cancel-pbs-job},
+growing. Features explained below are for \textbf{PBS/Torque} backend.
+This support is provided through \textit{submit-pbs-job}, \textit{cancel-pbs-job},
 \textit{scan-pbs-job} scripts. \textit{submit-pbs-job} creates job's
@@
 
-The GM can run specially prepared \emph{BASH} scripts prior creation
+The A-REX can run specially prepared \emph{BASH} scripts prior creation
 of job's script, before and after executing job's main executable.
@@
 are run through BASH's 'source' command, and hence can manipulate
-with shell variables. With argument '0' scripts are run by the GM
-on frontend. Some environment variables are defined in that case and
+shell variables. With argument '0' scripts are run by the A-REX on
+frontend. Some environment variables are defined in that case and
 can be changed to influence job's execution later:
@@
 
+For publicly available runtime environments please see RTE \url{http://gridrer.csc.fi/}
+.
+
 \section{Installation\label{sec:installation}}
 
-To install GM as part of ARC-enabled site please read {}``NorduGrid
-ARC server installation instructions'' at \url{http://www.nordugrid.org/documents/ng-server-install.html}.
+The A-REX is installed as part of ARC 1 middleware and is available
+in NorduGrid download area at ARC1 Download \url{http://download.nordugrid.org/software/nordugrid-arc1/}
+.There are packages available for various Linux distributions. The
+A-REX comes in nordugrid-arc1-arex package. Source code ready for
+compilation is available too.
 
@@
 
-The GM is mostly written using C++. It was tested and should compile
-on recent enough \emph{Linux} systems using \emph{gcc} compiler and
-\emph{GNU make} (gcc versions 2.95, 2.96, 3.2, 3.4 were tested). You
-will also need \emph{Globus Toolkit$^{TM}$} of version higher than
-2.2 installed \url{http://www-unix.globus.org/toolkit/}.
+If installed from binary packages all dependencies are handled automatically.
+For compilation from source code please read included README file.
 
@@
 it finished).
-
-\subsection{Setup of the GridFTP Server}
-
-For in-depth information about how to properly setup the GM and related
-software please read {}``NorduGrid ARC server installation instructions''
-at \url{http://www.nordugrid.org/documents/ng-server-install.html}.
-Follow that manual to install GM, configure and run it. Additional
-tips are described here.
-
-To make GFS to interoperate with other parts of the ARC only one \emph{jobplugin.so}
-needs to be configured. It is advisable to use the template configuration
-file. You can leave only part which configures \textit{jobplugin.so}
-plugin.
 
@@
 
-\section*{Appendix A. Job control over jobplugin.so}
+\section*{Appendix A. Session directory access through HTTP(S) interface}
 
@@
 
-\section*{Appendix B. Library \textit{libarcdata}}
-
-\emph{libarcdata} is now part of \emph{libngui} library. It's functions
-are declared in a header file \emph{arcdata.h}. They correspond to
-ng{*} utilities meant for data handling - \emph{arcacl}, \emph{arccp},
-\emph{arcls}, \emph{arcrm}, \emph{arctransfer}. It consists of following
-functions:
+\section*{Appendix C. Configuration schema of A-REX}
 
 \begin{lyxcode}
-void~arcacl(const~std::string\&~file\_url,~const~std::string\&~command,~int~timeout~=~0);
-
-void~arcregister~(const~std::string\&~source\_url,~const~std::string\&~destination\_url,~bool~secure~=~false,~bool~passive~=~true,~bool~force\_meta~=~false,~int~timeout~=~0);
-
-void~arccp~(const~std::string\&~source\_url,~const~std::string\&~destination\_url,~bool~secure~=~false,~bool~passive~=~true,~bool~force\_meta~=~false,~int~recursion~=~0,~bool~verbose~=~false,~int~timeout~=~0);
-
-void~arcls(const~std::string\&~dir\_url,~bool~show\_details~=~false,~bool~show\_urls~=~false,~int~recursion~=~0,~int~timeout~=~0);
-
-void~arcrm(const~std::string\&~file\_url,~bool~errcont~=~false,~int~timeout~=~0);
-
-void~arctransfer(const~std::string\&~destination,~std::list<std::string>\&~sources,~int~timeout~=~0);
+<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
+
+<xsd:schema
+
+xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
+
+xmlns=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
+
+xmlns:arc=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
+
+targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
+
+elementFormDefault=\char`\"{}qualified\char`\"{}>
+
+~~<xsd:complexType~name=\char`\"{}endpoint\_Type\char`\"{}>
+
+~~~~<!-{}-
+
+~~~~This~element~defines~URL~of~A-REX~service~as~seen~from~outside.
+
+~~~~-{}->
+
+~~~~<xsd:simpleContent>
+
+~~~~~~<xsd:extension~base=\char`\"{}xsd:string\char`\"{}>~
+
+~~~~~~</xsd:extension>
+
+~~~~</xsd:simpleContent>
+
+~~</xsd:complexType>
+
+~~<xsd:element~name=\char`\"{}endpoint\char`\"{}~type=\char`\"{}endpoint\_Type\char`\"{}/>
+
+~~<xsd:complexType~name=\char`\"{}gmconfig\_Type\char`\"{}>
+
+~~~~<!-{}-
+
+~~~~This~element~defines~path~to~arc0~Grid~Manager~configuartion~file.
+
+~~~~By~default~it~is~/etc/arc.conf.
+
+~~~~-{}->
+
+~~~~<xsd:simpleContent>
+
+~~~~~~<xsd:extension~base=\char`\"{}xsd:string\char`\"{}>
+
+~~~~~~</xsd:extension>
+
+~~~~</xsd:simpleContent>
+
+~~</xsd:complexType>
+
+~~<xsd:element~name=\char`\"{}gmconfig\char`\"{}~type=\char`\"{}gmconfig\_Type\char`\"{}/>
+
+~~<xsd:simpleType~name=\char`\"{}gmrun\_Type\char`\"{}>
+
+~~~~<!-{}-
+
+~~~~This~element~defines~how~grid-manager~part~of~A-Rex~is~run.
+
+~~~~{*}~internal~-~as~a~thread~inside~service~container.
+
+~~~~{*}~none~-~no~grid-manager~is~run.
+
+~~~~{*}~external~-~as~a~separate~executable~(not~supported~anymore).
+
+~~~~Default~is~'internal'.
+
+~~~~-{}->
+
+~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
+
+~~~~~~<xsd:enumeration~value=\char`\"{}internal\char`\"{}/>
+
+~~~~~~<xsd:enumeration~value=\char`\"{}external\char`\"{}/>
+
+~~~~~~<xsd:enumeration~value=\char`\"{}none\char`\"{}/>
+
+~~~~</xsd:restriction>
+
+~~</xsd:simpleType>
+
+~~<xsd:element~name=\char`\"{}gmrun\char`\"{}~type=\char`\"{}gmrun\_Type\char`\"{}/>
+
+~~<xsd:complexType~name=\char`\"{}usermap\_Type\char`\"{}>
+
+~~~~<xsd:sequence>
+
+~~~~~~<xsd:element~name=\char`\"{}defaultLocalName\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~~</xsd:sequence>
+
+~~</xsd:complexType>
+
+~~<xsd:element~name=\char`\"{}usermap\char`\"{}~type=\char`\"{}usermap\_Type\char`\"{}/>
+
+~~<!-{}-~CommonName~attribute~of~bes-factory.~-{}->
+
+~~<xsd:element~name=\char`\"{}commonName\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~<!-{}-~LongDescription~attribute~of~bes-factory.~-{}->
+
+~~<xsd:element~name=\char`\"{}longDescription\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~<!-{}-~Name~of~Local~Resource~Management~System.~-{}->
+
+~~<xsd:element~name=\char`\"{}LRMSName\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~<!-{}-
+
+~~Name~of~Operating~System.
+
+~~The~values~are~based~on~the~OSType~field~of~the~CIM\_OperatingSystem~model:
+
+~~http://www.dmtf.org/standards/cim/cim\_schema\_v29
+
+~~-{}->
+
+~~<xsd:element~name=\char`\"{}OperatingSystem\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+</xsd:schema>
 \end{lyxcode}
-Additionally this library contains C++ classes used by \emph{ng{*}}
-data management utilities. Those are described in {}``ARC::DataMove
-Reference Manual''.
-
-\section*{Appendix C. Error messages of GM}
-
-If job has not finished successfully the GM put one or more lines
-into \textit{job.ID.failed}. Possible valuesinclude those generated
-by the GM itself:
+
+\section*{Appendix C. Error messages of A-REX}
+
+If job has not finished successfully the A-REX puts one or more lines
+into \textit{job.ID.failed}. Possible values include those generated
+by the A-REX itself:
 
@@
 
-\section*{Appendix D. A-Rex WSDL}
+\section*{Appendix D. A-REX WSDL}
 

### 2008-08-17 doc/glite_gateway/glite_gateway.tex
Some minor editing

@@
 \end{titlepage}
+
 \tableofcontents
+
 \newpage
@@
 \label{Introduction}
-The purpose of this document is to describe this gLite\cite{glite} specific plugins of the ARCLIB\cite{arclib} to present the related command line utilities built on top of these libraries.
-The gateway library together with the tools enables seemless interoparability with gLite computing elements.
-It is possible, that relevant sections of this document will be moved into ARC1 userguide and into ARCLIB technical documentation.
-Interoperability with gLite storage element is provided by an other set of ARC library.
-This data library provides native interfaces to GridFTP\cite{gridftp} SRM which is the main bulding blocks of a gLite storage element.
-A storage interopability is out of scope of this library.
+The purpose of this document is to describe gLite~\cite{glite} specific plugins of ARCLIB~\cite{arclib} and to present the related command line utilities built on top of them.
+The gateway library together with these tools enables seemless interoparability with gLite computing elements (CE).
+It is possible, that relevant sections of this document will be moved into the ARC1 user guide and into the ARCLIB technical documentation.
+Interoperability with gLite storage elements (SE) is provided by another subset of ARC library.
+This data library provides native interfaces to GridFTP~\cite{gridftp} and SRM~\cite{srm} which are the main bulding blocks of a gLite storage element.
+Storage interopability is out of scope of this library.
+
 \section{Overview}
 \label{Overview}
-Interoperability has been one of the most important disciplines in recent grid middleware development.
+Interoperability has been one of the most important goals in recent grid middleware development.
 This term is often used in the sense of accessing resources operated by one kind of grid middleware (e.g.
 gLite) from a user interface operated by another kind (e.g.
-ARC1).
-There are two main possibilities to achieve this result:
+ARC).
+There are two main ways to achieve this result:
 \begin{itemize}
-\item either to use a special infrastructure element, the gateway, which performs a full data structure and protocol translation between the two inter-operating grid middleware solutions,
+\item either to use a special infrastructure element, the gateway, which performs a full data structure and protocol translation between the two interoperating grid middleware solutions,
 \item or to explicitly implement the interface of the other side in the client.
 \end{itemize}
-The latter solution was preferred by ARC1 developers, the gLite gateway library module has been implemented in the ARCLIB and the gateway functionality will be transparently offered via the usual NorduGrid command line tools.\par
-An overview of the preferred soulution is presented in figure \ref{fig:Chart1}.
-The user will be able to express the job requirements in the job description document using any of the xRSL\cite{xrsl}, JDL or JSDL\cite{jsdl} languages.
-The ARC client library will perform an automatic transformation.
-The gLite gateway functionality integrated into the ARCLIB, thus into the ARC client allow the user to transparently access gLite computing resources via the usual ARC command line tools.
-The gLite modules perform authentication, credential delagation and job submission to the CREAM2 CE.
-The CREAM2 CE may perform the necessary input file download from the different storage elements.
-The library module gets job information through CREAM2 and sends management commands, like kill a job or undelegate credentials.
-Note that not all the functionality described in the overview figure is implemented yet, for the current limitations see section \ref{Current limitations}!\par
+The latter solution has been chosen by ARC developers, the gLite gateway library module has been implemented in the ARCLIB and the gateway functionality will be transparently offered via the usual ARC command line tools.
+
+An overview of the implemented soulution is presented in Figure~\ref{fig:Chart1}.
+The gLite gateway functionality integrated into the ARCLIB and thus into the ARC client, allows the user to transparently access gLite computing resources via the usual ARC command line tools.
+The user will be able to express the job requirements in the job description document using any of the xRSL~\cite{xrsl}, JDL or JSDL~\cite{jsdl} languages.
+The ARC client library will perform an automatic translation.
+The gLite modules perform authentication, credential delegation and job submission to a CREAM2~\cite{cream} CE.
+CREAM2 CE may perform the necessary input file download from different storage elements.
+The library module gets job information through CREAM2 and sends management commands, like killing a job or undelegating credentials.
+Note that not all the functionality described in the overview figure is implemented yet, for the current limitations see Section~\ref{Current limitations}.
+
 \begin{figure}[ht]
@@
 \end{figure}
+
 \subsection{The supported gLite CE interface: CREAM2}
 The gLite computing element (CE) offers several interfaces.
-Some of them are only suitabe for the WMS (gLite resource broker) and was not meant to be used by 3rd party clients.
-Fortunately gLite comes with a webservice based computing element interface, known as the CREAM CE.
+Some of them are only suitabe for the gLite workload management system and were not meant to be used by 3rd party clients.
+Fortunately, gLite comes with a Web Service based computing element interface, known as the CREAM CE~\cite{cream}.
 The CREAM has undergone several major revisions.
-The current CREAM version has entered the EGEE certification process nevertheless it is a full scale deplomeyment on the EGEE worldwide grid has not taken place yet.
-CREAM2 is available on several preproduction and testing grid sites, KnowARC has chosen the CREAM version 2\cite{cream} as the gLite computing element interface.\par
-In particular, the gateway library was developed against version glite-ce-cream\_R\_1\_8\_3\_0\footnote{\url{http://grid.pd.infn.it/cream/field.php?n=Main.ReleaseNotes}}.\par
+The current CREAM version has recently entered the EGEE certification process, nevertheless its full-scale deplomeyment on the EGEE worldwide grid has not taken place yet.
+CREAM2 is currently available on several preproduction and testing grid sites.
+KnowARC has chosen the CREAM version 2 (CREAM2) as the target gLite computing element interface.
+
+In particular, the gateway library was developed against version glite-ce-cream\_R\_1\_8\_3\_0\footnote{\url{http://grid.pd.infn.it/cream/field.php?n=Main.ReleaseNotes}}.
+
 CREAM2 was chosen for the following reasons:
 \begin{itemize}
-\item CREAM2 has a web service interface that fits the web service based ARC1.
-\item CREAM2 enables direct access to the gLite computing element without the need of going through the WMS.
-\item CREAM2 contains numerous improvements compared to the earlier CREAM versions.
-\item Unlike the other gLite CE interfaces it supports direct job status queries.
+\item CREAM2 has a Web Service interface that fits the Web Service based ARC.
+\item CREAM2 enables direct access to the gLite computing element without the need of going through the gLite workload management system.
+\item CREAM2 contains numerous improvements when compared to the earlier CREAM versions.
+\item Unlike other gLite CE interfaces it supports direct job status queries.
 \item It is an official interface to gLite computing elements and as EGEE evolves it is expected that more and more computing elements will be deployed with CREAM2 interface.
@@
 \end{itemize}
-In the near future CREAM will offer yet another interface which will be based on the BES/JSDL/GLUE OGF standards. This CREAM BES interface will coexists with CREAM2 too.\par
+In the near future CREAM will offer yet another interface which will be based on the BES~\cite{ogsa-bes}, JSDL and GLUE~\cite{glue} OGF standards.
+This CREAM BES interface will coexist with CREAM2 too.
+
 \section{External dependencies and installation instruction}
 \label{External dependencies and installation instruction}
-The ARCLIB gLite plugins have only one dependency.
-GridFTP components of the Globus Toolkit\cite{globus} are needed to enable data staging functionality to those CREAM2 computing elements which offer input and output sandbox over a GridFTP enabled storage.
-Unfortunately the gateway library only works together with a strongly patched Globus verion.
+The ARCLIB gLite plugins have only one dependency: GridFTP components of the Globus Toolkit~\cite{globus} are needed to enable data staging functionality to those CREAM2 computing elements which offer input and output sandbox over a GridFTP enabled storage.
+Unfortunately, the gateway library only works together with a strongly patched Globus verion.
 The NorduGrid patched Globus can be dowloaded from the NorduGrid download area\footnote{\url{http://download.nordugrid.org/software/globus/}}.
 Further information about the NorduGrid Globus patches can be read on the NorduGrid page\footnote{\url{http://wiki.nordugrid.org/index.php/Globus\_Libraries}}.
-It is worth to mention that the user will need a VOMS\cite{voms} proxy certificate in order to use the command line tools against a CREAM2 service.
-This VOMS proxy can be generated by using the VOMS client tools.
-It is only the VOMS enabled proxy certificate that is needed here, apart from this no VOMS components are needed.\par
-The ARC$\rightarrow$gLite gateway plugins library and the temporary command line utilities are incorporated into ARC1 releases.
-The source code is available from the ARC1 SVN repository\footnote{\url{http://svn.nordugrid.org/repos/nordugrid/arc1/}} and binary packages are provided at the NorduGrid download area.
-The build and installation instructions are given in the README file distributed together with the software.
-A step-by-step guide for Debian binary installation can be found in appendix \ref{Debian install} \textit{Installation from Debian binary packages}.
+It is worth mentioning that the user would need a VOMS~\cite{voms} proxy certificate in order to use command line tools against a CREAM2 service.
+This VOMS proxy can be generated by using standard VOMS client tools.
+It is only the VOMS enabled proxy certificate that is needed here, apart from this no VOMS components are needed.
+
+The ARC$\rightarrow$gLite gateway plugins library and the temporary command line utilities are incorporated into ARC releases.
+The source code is available from the ARC1 SVN repository\footnote{\url{http://svn.nordugrid.org/repos/nordugrid/arc1/}} and binary packages are provided in the NorduGrid download area.
+Build and installation instructions are given in the README file distributed together with the software.
+A step-by-step guide for Debian binary installation can be found in Appendix~\ref{Debian install} \textit{Installation from Debian binary packages}.
+
 \section{ARCLIB gLite gateway plugins}
 \label{ARCLIB gLite gateway plugins}
-The ARCLIB gLite gateway plugins constist of the wrapper classes, the job description handling facility and the gLite specific CREAMClient base class.
-The wrapper classes covey the function provided by the CREAMClient to ARCLIB.
+The ARCLIB gLite gateway plugins constist of the wrapper classes, the job description handling facility and the gLite specific \texttt{CREAMClient} base class.
+The wrapper classes convey the function provided by the \texttt{CREAMClient} to ARCLIB.
 The job description handling framework makes a general job request representation available for the other classes.
-CREAMClient base class implements the interface towards CREAM2 computing elements.
-This client object has an easy-to-use and intuitive interface. It uses the VOMS proxy certificate to build the secure channel to the server and also to sign that of the server's.
-The different functions throw \textit{CREAMClientError} exceptions that should be caught in the application.\par
-The following sections present the set of functions provided by the library.\par
+\texttt{CREAMClient} base class implements the interface towards CREAM2 computing elements.
+
+This client object has an easy-to-use and intuitive interface. It uses VOMS proxy certificates to build the secure channel to the server and also to sign that of the server.
+
+Different functions throw \textit{CREAMClientError} exceptions that should be caught in the application.
+
+The following sections present the set of functions provided by the library.
+
 \subsection{Wrapper classes}
 \label{Wrapper classes}
-The gLite gateway library can be integrated into the ARCLIB as a plugin.
-To expose the CREAM2 functionality into the generic ARCLIB target object functions, the following three wrapper classes have been implemented: the JobController, the Submitter and the TargetRetriever.\par
-The JobControllerCREAM wrapper class provides the basic functions related to job management, such as getting job information or killing and cleaning the job on the remote CE.\par
-The SubmitterCREAM instantiates the gLite client library and calls the CREAM2-specific submitter function in there.\par
-Class TargetRetrieverCREAM performs a BDII database query and extracts the possible CREAM2 target objects.\par
-Unfortunately handling CREAM2 target objects do not perfectly fit into the philosophy of ARCLIB, there are several functions, such as proxy credential delegation or CREAM2-specific job information retrieval that behaves differently on an ARC1 and on a CREAM2 target.
+The gLite gateway library can be integrated into ARCLIB as a plugin.
+To expose the CREAM2 functionality into the generic ARCLIB target object functions, the following three wrapper classes have been implemented: the \texttt{JobController}, the \texttt{Submitter} and the \texttt{TargetRetriever}.
+
+The \texttt{JobControllerCREAM} wrapper class provides the basic functions related to job management, such as getting job information or killing and cleaning the job on the remote CE.
+
+The \texttt{SubmitterCREAM} instantiates the gLite client library and calls the CREAM2-specific submitter function in there.
+
+Class \texttt{TargetRetrieverCREAM} performs a gLite BDII database query and extracts possible CREAM2 target objects.
+
+Unfortunately, handling CREAM2 target objects does not perfectly fit into the philosophy of ARCLIB.
+There are several functions, such as proxy credential delegation or CREAM2-specific job information retrieval, that behave differently on an ARC and on a CREAM2 target.
 The wrapper classes have to provide solutions for these special cases.
+
 \subsection{Job description translators}
@@
 When addressing interoperability it is of paramount importance to transparently address grid job descriptions written in different job description languages by translating them automatically.
-In the gLite gateway library code this functionality is implemented in class JobDescription.
-This is a generic class that takes a job description as input in any supported formats (currently xRSL, JDL, JSDL), converts and stores it in a JSDL-like internal job description format.
-Then different operations, like getting the description in other formats or getting job-related information, can be performed using the description-independent functions of this class.\par
-JobDescription has tree back-end classes, sometimes are referred to as back-end modules or translator modules, corresponding to the three supported job description languages above.
+In the gLite gateway library code this functionality is implemented in class \texttt{JobDescription}.
+This is a generic class that takes a job description as input in any supported format (currently xRSL, JDL, JSDL), converts and stores it in a JSDL-like internal job description format.
+Then different operations, like getting the description in other formats or getting job-related information, can be performed using the description-independent functions of this class.
+
+\texttt{JobDescription} has tree back-end classes, sometimes are referred to as back-end modules or translator modules, corresponding to the three supported job description languages above.
 The back-end classes are used for parsing and also for generating the job descriptions.
-These classes do not store information and are not intended to be used directly, instead, the generic class uses them.\par
-There is a class called JobDescriptionOrderer which tries to find out the format of the input file by pattern matching; the generic class uses this piece of information to choose the appropriate back-end.
+These classes do not store information and are not intended to be used directly, instead, the generic class uses them.
+
+There is a class called \texttt{JobDescriptionOrderer} which tries to find out the format of the input file by pattern matching; the generic class uses this piece of information to choose the appropriate back-end.
+
 \begin{figure}[ht]
@@
 \end{figure}
+
 \subsection{CREAMClient base class}
 \label{CREAMClient}
-The client object uses the VOMS proxy certificate to build the secure channel to the server and also to sign that of the server's.\par
-To establish the SOAP connection via the secure channel, the library use the native ARC1 HED\footnote{\url{https://www.knowarc.eu/documents/Knowarci\_D1.1-1\_07.pdf}} Message Chain Component (MCC), that provides full integration with the ARC1 core and corresponding services.
-This replaced the former gSOAP-based\cite{gsoap} communication libraries, which enabled the full elimination of the gSOAP dependency.
-\subsubsection*{CREAMClient(Arc::URL, Arc::MCCConfig)\footnote{For class references see: \url{https://www.knowarc.eu/download/D1.2-2\_documentation.pdf}}}
+The client object uses the VOMS proxy certificate to build the secure channel to the server and also to sign that of the server.
+
+To establish SOAP connection via the secure channel, the library uses the native ARC HED\footnote{\url{https://www.knowarc.eu/documents/Knowarci\_D1.1-1\_07.pdf}} Message Chain Component (MCC), that provides full integration with the ARC core and corresponding services.
+This replaces the former gSOAP-based~\cite{gsoap} communication libraries, thus completely eliminating gSOAP dependency.
+
+\subsubsection*{CREAMClient(Arc::URL, Arc::MCCConfig)\footnote{For class references, see: \url{http://www.knowarc.eu/download/D1.2-2\_documentation.pdf}}}
 The constructor of the class performs the necessary initialization work.
 It receives two arguments: the service URL and the ARC message chain component configuration file to establish the communication channel to the server.
-These pieces of information are indispensable to create your client and to communicate with the remote site.
+These pieces of information are indispensable to create the client and to communicate with the remote site.
+
 \subsubsection*{setDelegationId(std::string)}
@@
 The function is simple: it sets a private variable and returns with no value.
+
 \subsubsection*{createDelegation(std::string)}
-The \textbf{createDelegation} function is about to perform the whole delegation registration process.
+The \texttt{createDelegation} function performs the whole delegation registration process.
 It sends a \textit{getProxyReq} message having the requested delegation ID, signs the received certificate and sends it back in a \textit{putProxy} SOAP message to the server.
 These three steps constitute the delegation registration process.
+
 As almost every funtion, this one has no return value either.
@@
 If there are no exceptions thrown, then the command is considered to have successfully completed.
+
 \subsubsection*{destroyDelegation(std::string)}
-This method has the reverse functionality of function \textbf{createDelegation}. It sends the \textit{destroy} message of CREAM2 to the remote server.
+This method has the reverse functionality of function \texttt{createDelegation}. It sends the \textit{destroy} message of CREAM2 to the remote server.
+
 \subsubsection*{submit(std::string)}
-The \textbf{submit} function is the most complex part of the class.
+The \texttt{submit} function is the most complex part of the class.
 It translates the job description received as the argument, registers the job with a \textit{JobRegisterRequest} message, uploads the locally stored files, if necessary, then launches the job execution on the server by sending a \textit{JobStartRequest} message.
-Finally it returns with a \textbf{creamJobInfo} object that contains a \textit{jobId} (job indetifier), a \textit{creamURL} (as the service URL), an \textit{ISB} (reference to the Input Sandbox) and an \textit{OSB} (Output Sandbox) member to describe the registered job.
-In the \textbf{glitesub} command (See Section \ref{glitesub} for details) these pieces of information are stored in the information file.
+Finally it returns with a \texttt{creamJobInfo} object that contains a \textit{jobId} (job indetifier), a \textit{creamURL} (as the service URL), an \textit{ISB} (reference to the Input Sandbox) and an \textit{OSB} (Output Sandbox) member to describe the registered job.
+In the \texttt{glitesub} command (See Section~\ref{glitesub} for details) these pieces of information are stored in the information file.
+
 \subsubsection*{stat(std::string)}
@@
 The return value is the job status translated to the ARC terminology.
-The possible values are presented in the section of the \textbf{glitestat} command line tool.
-(See Section \ref{glitestat}!)
+The possible values are presented in the section of the \texttt{glitestat} command line tool (see Section~\ref{glitestat}).
+
 \subsubsection*{cancel(std::string)}
-The \textbf{cancel} function sends a \textit{JobCancelRequest} SOAP message to the server and handles the emerging exceptions.
+The \texttt{cancel} function sends a \textit{JobCancelRequest} SOAP message to the server and handles the emerging exceptions.
 It can be used for canceling a remotely registered and possibly running job.
+
 \subsubsection*{purge(std::string)}
-The \textbf{purge} function sends a \textit{JobPurgeRequest} message to server and throws an exception in case of any emerging problems.
+The \texttt{purge} function sends a \textit{JobPurgeRequest} message to the server and throws an exception in case of any emerging problems.
+
 \section{ARC command line tools for gLite CREAM2 CE}
 \label{Users guide}
-The well known ARC commands (ngsub, ngget, See: The NorduGrid/ARC User Guide\cite{userguide}!) will provide seemless access and integration with gLite computing elements.
+The well known ARC commands (\texttt{ngsub}, \texttt{ngget}, see The NorduGrid/ARC User Guide\cite{userguide}!) will provide seemless access and integration with gLite computing elements.
 Full transparency will be achieved through the gLite gateway ARCLIB plugins.
-An ordinary user will use the same set of commands regardless of what type of CE to access.
-The user will be able to specify grid jobs described both in the gLite (JDL) and the NorduGrid (JSDL, xRSL) job description languages.
-The client will perform automatic conversion between the different job description languages.\par
-Unfortunately the interoperability features offered by the gLite gateway library are not yet integrated into the NorduGrid commands.
-Also the current version of ARCLIB does not offer the full spectrum of language transformations.\par
-However only a set of temporary command line tools exposing the available features of the library are provided.
-These temporary commands make the already existing interoparabilily features usable for ordinary users.\par
-There are seven different command line tools for using the gLite gateway functions.
-In this section these tools will be presented in detail through examples.
-The job descriptions and services shown are just simple examples, their arguments might be different in real-life usage.\par
-Every tool has a short manual page being accessible by using the \textbf{man} on-line manual reader command or some more help can be achieved by using the $<$command$>$ -? command line option.\par
+An ordinary user will use the same set of commands regardless of what type of CE is accessed.
+The user will be able to specify grid jobs described both in gLite (JDL) and ARC (JSDL, xRSL) job description languages.
+The client will perform automatic translation between the different job description languages.
+
+Unfortunately, the interoperability features offered by the gLite gateway library are not yet integrated into the ARC commands.
+Also, the current version of ARCLIB does not offer the full spectrum of language transformations yet.
+
+Currently, only a set of temporary command line tools exposing the available features of the library are provided.
+These temporary commands make the already existing interoparabilily features usable for ordinary users.
+
+There are seven different commands for using the gLite gateway functions.
+In this section these commands will be presented in details through examples.
+Job descriptions and services shown are just simple examples, their arguments might be different in real-life usage.
+
+Every tool has a short manual page being accessible by using the \texttt{man} on-line manual reader command; some more help can be achieved by using the $<$command$>$ -? command line option.
+
 \subsection{glitedelegate}
 \label{glitedelegate}
-In order to use gLite resources user must have a VOMS proxy. This proxy can be generated by command \textit{voms-proxy-init}.\par
+In order to use gLite resources user must have a VOMS proxy. This proxy can be generated by command \textit{voms-proxy-init}.
+
 For example:
-\begin{shaded}\verb#$ voms-proxy-init -voms knowarc.eu#\end{shaded}
-Once a VOMS proxy has been generated everything is in place to submit a new job.
+\begin{shaded}\verb#$ voms-proxy-init -voms knowarc.eu#\end{shaded}%$
+Once a VOMS proxy has been generated, everything is in place to submit a new job.
 If one wants submit a job to a gLite server, first has to own a valid delegation on the remote site.
 This can either be an old, but still valid delegation identifier, or can be a new one.
-There is a command line tool to register a delegation on the CE to be used.\par
+There is a command line tool to register a delegation on the CE to be used.
+
 The usage of command \textbf{glitedelegate} is straighforward.
@@
 This delegation ID will be associated to the remote resource.
-If you are not sure whether your favourite delegation ID is occupied or not, then try to delete it before registering it again.
-(See Section \ref{gliteundelegate} for further details!) The delegation command has the following syntax:
-\begin{shaded}\verb#$ glitedelegate <delegation ID> <delegation service URL>#\end{shaded}
+If you are not sure whether your favourite delegation ID is occupied or not, then try to delete it before registering it again (see Section~\ref{gliteundelegate} for further details).
+The delegation command has the following syntax:
+\begin{shaded}\verb#$ glitedelegate <delegation ID> <delegation service URL>#\end{shaded}%$
 For example:
 \begin{shaded}\verb#$ glitedelegate test_delegation \#\\*
-\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}
+\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}%$
+
 \subsection{glitesub}
 \label{glitesub}
-The current command only supports direct job submission that is the user must know service endpoints.
-These kind of endpoints usually stored in LDAP databases.
-Ordinary ldapsearch can be used to find such an endpoint.
-Below comes an example:\par
+The current command only supports direct job submission to a known CE, that is, the user must know service endpoints.
+These kind of endpoints are usually stored in LDAP/BDII databases.
+Ordinary \texttt{ldapsearch} can be used to find such an endpoint.
+Below comes an example:
+
 \begin{shaded}
@@
 \end{shaded}
-The \textbf{glitesub} command implements the actual job submission in three phases: registration of a job, uploading the local files to the input sandbox URL received during the registration and starting the job.\par
-The \textbf{glitesub} command requires a valid delegation.
-See Section \ref{glitedelegate} how to create one.
-By the way this is the only command which needs the delegation ID.\par The next step is to prepare the necessary input files and the job description.\par
-The glitesub command syntax is the following (please note that the service URL, job description, info file are mandatory and can only be given in the specified order):
-\begin{shaded}\verb#$ glitesub -D <delegation ID> <service URL> <job description> <info file>#\end{shaded}
+The \textbf{glitesub} command implements the actual job submission in three phases: registration of a job, uploading the local files to the input sandbox URL received during the registration, and starting the job.
+
+The \textbf{glitesub} command requires a valid delegation. See Section~\ref{glitedelegate} how to create one. By the way, this is the only command which needs the delegation ID.
+
+The next step is to prepare the necessary input files and the job description.
+
+The \textbf{glitesub} command syntax is the following (please note that the service URL, job description, info file are mandatory and can only be given in the specified order):
+\begin{shaded}\verb#$ glitesub -D <delegation ID> <service URL> <job description> <info file>#\end{shaded}%$
 Here the delegation ID is the identification registered previously.
@@
 It should be a non-existing file otherwise it will be overwritten.
-This file contains the job ID, the input sandbox URL, the output sandbox URL and the service URL.\par
-Below I present a JDL description example.
-Note that the \textit{VirtualOrganisation} and \textit{QueueName} attributes should be the same as your virtual organisation name and the corresponding queue name on the server side.\par
+This file contains the job ID, the input sandbox URL, the output sandbox URL and the service URL.
+
+Below is a JDL file example.
+Note that the \textit{VirtualOrganisation} and \textit{QueueName} attributes should be the same as your virtual organisation name and the corresponding queue name on the server side.
 
@@
 \begin{shaded}\verb#$ glitesub -D test_delegation \#\\*
-\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/CREAM2 description.jdl job.info#\end{shaded}
-After successful completion the command reports the success and tells the user where the job related data are stored. Otherwise the message written to the output contains the reason of the failure.
+\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/CREAM2 description.jdl job.info#\end{shaded}%$
+After successful completion the command reports success and tells the user where the job related data are stored. Otherwise the message written to the output contains the reason of the failure.
+
 \subsection{glitestat}
 \label{glitestat}
-The \textbf{glitestat} command contacts to the CREAM2 service on the service URL and queries job status of a job specified by the job ID given in the info file.\par
+The \textbf{glitestat} command contacts the CREAM2 service on the service URL and queries status of a job specified by the job ID given in the info file.
+
 Usage:
-\begin{shaded}\verb#glitestat <info file>#\end{shaded}
+\begin{shaded}\verb#$ glitestat <info file>#\end{shaded}%$
 For example:
-\begin{shaded}\verb#glitestat job.info#\end{shaded}
-The glitesub command implements a job status mapping between gLite and ARC job status semantics. The possible responses and their meanings are the following:
+\begin{shaded}\verb#$ glitestat job.info#\end{shaded}%$
+The \textbf{glitesub} command implements job status mapping between gLite and ARC job status semantics. The possible responses and their meanings are the following:
 \begin{itemize}
@@
 \end{itemize}
+
 \subsection{glitekill}
@@
 The \textbf{glitekill} command can be used for killing a remote job running in the underlying batch system.
-The only necessary argument needed is the information file made by \textbf{glitesub}.
-(See Section \ref{glitesub} for further details!) This command initiates stopping the job in the batch system.
-You can check the effect by using \textbf{glitestat}.
-(See Section \ref{glitestat} for its usage!)\par
+The only necessary argument needed is the information file made by \textbf{glitesub} (see Section~\ref{glitesub} for further details).
+This command initiates stopping the job in the batch system.
+You can check the effect by using \textbf{glitestat} (see Section~\ref{glitestat} for its usage).
+
 Usage:
-\begin{shaded}\verb#glitekill <info file>#\end{shaded}
+\begin{shaded}\verb#$ glitekill <info file>#\end{shaded}%$
 For example:
-\begin{shaded}\verb#glitekill job.info#\end{shaded}
+\begin{shaded}\verb#$ glitekill job.info#\end{shaded}%$
+
 \subsection{gliteclean}
 \label{gliteclean}
-Either after killing a job or because of some remote error occurs garbage files may remain on the server.
-These files can be purged by using command \textbf{gliteclean}.
-If you intentionally kill a job on the computing node, the job related stuff might also remain as garbage in the remote queue.
+Either after killing a job or because some remote error occurs, garbage files may remain on the server.
+These files can be removed by using command \textbf{gliteclean}.
+If you intentionally kill a job on a computing node, the job related stuff might also remain as garbage in the remote queue.
 This command also removes the job from the underlying batch system on the gLite computing element.
 It is important to note that this command removes the local information file as well.
-It might be useful to run this command after killing a job.
-(See Section \ref{glitekill}!)\par
+It might be useful to run this command after killing a job (see Section~\ref{glitekill}).
+
 Usage:
-\begin{shaded}\verb#gliteclean <info file>#\end{shaded}
+\begin{shaded}\verb#$ gliteclean <info file>#\end{shaded}%$
 For example:
-\begin{shaded}\verb#gliteclean job.info#\end{shaded}
+\begin{shaded}\verb#$ gliteclean job.info#\end{shaded}%$
+
 \subsection{gliteundelegate}
 \label{gliteundelegate}
-After the job execution is complete it is possible to unregister and delete the delegation entry from the remote site by using command \textbf{gliteundelegate}.
+After the job execution is completed, it is possible to unregister and delete the delegation entry from the remote site by using command \textbf{gliteundelegate}.
 It can also be used when one needs to register a delegation ID but is unsure whether it is already used or not.
-Put the delegation service URL into the argument!\par
+Put the delegation service URL into the argument!
+
 Usage:
-\begin{shaded}\verb#$ gliteundelegate <delegation ID> <delegation service URL>#\end{shaded}
+\begin{shaded}\verb#$ gliteundelegate <delegation ID> <delegation service URL>#\end{shaded}%$
 For example:
 \begin{shaded}\verb#$ gliteundelegate test_delegation \#\\*
-\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}
+\verb#        https://cream.grid.upjs.sk:8443/ce-cream/services/gridsite-delegation#\end{shaded}%$
+
 \subsection{Skeleton of a temporary example gLite client}
-Finally, here is an example code to present how easy and simple it is to write a new client using the CREAMClient class (See section \ref{CREAMClient}!).
+Finally, here is an example code which shows how easy and simple it is to write a new client using the \texttt{CREAMClient} class (see Section~\ref{CREAMClient}).
 This code sample requests the job status of a previously submitted remote job, and writes it to the standard output.
@@
 \verb?}?
-\end{framed}
+\end{framed}%>>
 It is of course mandatory to set \textit{SERVICE\_URL} and \textit{JOBID} strings to their real values to obtain the proper functionality.
+
 \section{Open issues}
-In order to achieve the full scale transparent interoperability width the gLite computing element transparently offered through the NorduGrid client a couple of issues still need to be resolved.
-Most of these are related to the incomplete integration of the gLite gateway library into the ARCLIB framework, even though the it is ready for production use.
-There are also some conceptual challenges due to the incompatible nature of the two systems, which requires workarounds.
+In order to achieve the full scale transparent interoperability width the gLite computing element offered through the ARC client, a couple of issues still need to be resolved.
+Most of these are related to the incomplete integration of the gLite gateway library into the ARCLIB framework, even though it is ready for production use.
+There are also some conceptual challenges due to the incompatible nature of the two systems, which require workarounds.
+
 \subsection{Current limitations}
 \label{Current limitations}
-Full ARC1 client integration would mean that the temporary commands developed on top of the CREAMClient presented in the previous section \ref{CREAMClient} will be overriden by the ordinary ARC commands built on top of the ARCLIB.
-To achieve this the Wrapper class integration of the gLite gateway plugins library should be finalized and exposed to extensive testing.\par
-Some of the Wrapper class methods and data structures should be extracted from the information available in the CREAMClient class.
-In particular, the detailed job representation needed for job monitoring is an open area.\par
-The CREAMClient class will be extended to support operations such as \textit{suspend} available both in the CREAM2 interface and the ARCLIB's jobcontrol component.\par
-Due to the incomplete ARCLIB integration currently only direct job submission to a manually selected CE is possible.
-The ARCLIB integration will combine the existing resource discovery (See section \ref{Wrapper classes}) and job submission functionality of the gLite gateway plugins with the brokering module of ARCLIB, which is  currently under development.
-The combination of these ARCLIB modules will provide intelligent brokered job submission over gLite and NorduGrid computing element.\par
-For a fully transparent integration automatic translation among JSDL/xRSL and JDL should also be provided.
-A user should be able to specify his/her job request in any of these languages and the gateway does the necessary translations.
-The current language translator module currently is not capable of generating JDL output.
-Therefore, the user still needs to use JDL to submit jobs to CREAM2.\par
-Further investigation is needed on how a CREAM2 service handles data staging and whether the data movement related ARC libraries can be used to manage data staging on the client side or development of special data staging wrapper scripts executed on the computing element are needed.
-A set of temporary workaround command line clients have been developed (the glite* commands; see section \ref{Users guide}) due to a discovered OpenSSL version conflict.
-The conflict appeared when we tried to integrate the CREAMClient gateway library to the rest of the ARCLIB.
-A tedious and time consuming debugging process took place.
-It revealed the following:  A VOMS proxy certificate delegation method was implemented in the CREAMClient gateway library.
+Full ARC client integration would mean that the temporary commands developed on top of the \texttt{CREAMClient} presented in Section~\ref{CREAMClient} will be overridden by the ordinary ARC commands built on top of the ARCLIB.
+To achieve this, the \texttt{Wrapper} class integration of the gLite gateway plugins library should be finalized and exposed to extensive testing.
+
+Some of the \texttt{Wrapper} class methods and data structures should be extracted from the information available in the \texttt{CREAMClient} class.
+In particular, detailed job representation needed for job monitoring is an open area.
+
+The \texttt{CREAMClient} class will be extended to support operations such as \textit{suspend} available both in the CREAM2 interface and the ARCLIB's jobcontrol component.
+
+Due to the incomplete ARCLIB integration, currently only direct job submission to a manually selected CE is possible.
+The ARCLIB integration will combine the existing resource discovery (see Section~\ref{Wrapper classes}) and job submission functionality of the gLite gateway plugins with the brokering module of ARCLIB, which is currently under development.
+The combination of these ARCLIB modules will provide intelligent brokered job submission over gLite and ARC computing element.
+
+For a fully transparent integration, automatic translation among JSDL/xRSL and JDL should also be provided.
+A user should be able to specify her job request in any of these languages and the gateway should perform the necessary translations.
+The existing language translator module currently is not capable of generating JDL output.
+Therefore, the user still needs to use JDL to submit jobs to CREAM2.
+
+Further investigation is needed on how a CREAM2 service handles data staging and whether the data movement related ARC libraries can be used to manage data staging on the client side, or development of special data staging wrapper scripts executed on the computing element is needed.
+
+A set of temporary workaround command line clients have been developed (the glite* commands; see Section~\ref{Users guide}) due to a discovered OpenSSL version conflict.
+The conflict appeared when we tried to integrate the \texttt{CREAMClient} gateway library into the rest of ARCLIB.
+A tedious and time-consuming debugging process took place.
+It revealed the following: a VOMS proxy certificate delegation method was implemented in the \texttt{CREAMClient} gateway library.
 During this process it was needed to sign a certificate and the original OpenSSL was used for this purpose.
 In the communication library used and taken from the ARC1 framework there is a built-in Message Chain Component (MCC) which already utilizes the Globus Toolkit OpenSSL functions to enable establishing secure communication.
-Linking the CREAMClient library directly against the standard OpenSSL libs caused congesting SSL functions.
+Linking the \texttt{CREAMClient} library directly against the standard OpenSSL libs caused congesting SSL functions.
 In order to resolve the OpenSSL conflict and to make possible the complete ARCLIB integration  a Globus patch has been worked out.
-Once the OpenSSL conflict is resolved the ARCLIB integration will be completed this way overriding the temporary glite* commands.\par
-\subsection{Conceptional problems}
+Once the OpenSSL conflict is resolved, the ARCLIB integration will be completed, this way overriding the temporary glite* commands.
+
+\subsection{Conceptual problems}
 During the development of the gateway library a few conceptual incompatibilities were found.
-These are the typical cases when one of the systems offers richer capabilities not available on the other one.
-These conceptual gaps can only be bridged by developing customized case by case workarounds, nevertheless  the interoperation always would result in some information loss in these cases.\par
+These are typical when one of the systems offers richer capabilities not available in the other.
+These conceptual gaps can only be bridged by developing customized case-by-case workarounds.
+Still, the interoperation will always result in some information or functionality loss in these cases.
+
 Below are the most important challenges for which solution is still sought:
+
 \subsubsection*{Expressing file staging in job description}
-The xRSL offers richer capability to describe input and output date of a job than JDL.
+The xRSL offers richer capability to describe input and output data of a job than JDL.
 Especially, in xRSL it is possible for input files downloaded from storage servers to have their names changed.
@@
 \end{shaded}
-That is not possible to express in JDL.
-It only allows the file to be downloaded as datafile1.txt.
+This is not possible to express in JDL.
+It only allows the file to be downloaded as \texttt{datafile1.txt}.
 This could be solved by a wrapper script that renames files according to the xRSL speciﬁcation.
-This does not solve the problem completely however.
-For example if a user wants to concatenate the stdout from two jobs, he would write something like the following in xRSL:
+This however does not solve the problem completely.
+For example, if a user wants to concatenate standard outputs from two jobs, she would write something like the following in xRSL:
 \begin{shaded}
@@
 \end{shaded}
-This cannot be solved just using the renaming method, because both input files would be retrieved before the renaming and they would therefore overwrite each other.\par
+This cannot be solved just using the renaming method, because both input files would be retrieved before the renaming and they would therefore overwrite each other.
+
 A possible  solution is to recognize jobs of this type and simply barring them from using gLite resources or to let such jobs fail with an error message explaining that the user has to express their job in a different way.
+
 \subsubsection*{RunTime Environments related issues}
@@
 The requirement as well as the initialization of such software is expressed through the use of Run-Time Environments (RTEs).
-While the library is capable of translating the RTE request itself it is not capable of translating the name of the RTE.
+While the library is capable of translating the RTE request itself, it is not capable of translating the name of the RTE.
 No standardized cross-grid nomenclature has been adopted, nor are the authors aware of any such activity.
@@
 This could then be used for translating jobs.
+
 \appendix
@@
 \label{Debian install}
-This section contains instructions how to install the gateway libraries and glite* command line tools on Debian Linux.
+This section contains instructions on how to install the gateway libraries and glite* command line tools on Debian Linux.
 \begin{enumerate}
-\item Put the Debian repository into the APT source files (/etc/apt/sources.list). This repository in Luebeck is maintained as long as the KnowARC project continues.
+\item Put the Debian repository into the APT source files (\texttt{/etc/apt/sources.list}). This repository in Luebeck is maintained as long as the KnowARC project continues.
 \begin{shaded}\verb#$ echo "deb http://pc02.inb.uni-luebeck.de:8080/~moeller/debian/stable ./" \#\\*

### 2008-08-17 doc/a-rex/arex_tech_doc.tex
Conversion from GM to A-REX is done in A-REX document. Language and citations fixes to follow.

@@
 \item \textit{jobreport} - URL of a user requested \emph{logger service}.
-The GM will also send job records to this service in addition to the
-default logger service configured in the configuration (This feature
+The A-REX will also send job records to this service in addition to
+the default logger service configured in the configuration (This feature
 does not work in current implementation).
@@
 BES. It does that by defining its own port-type with the single operation
-\emph{ChangeActivityStatus} (see Appendix D). This operation provides
+\emph{ChangeActivityStatus} (see Appendix C). This operation provides
 a way to request simple transfers between states of jobs and corresponding
@@
 
-The A-REX also supportd the Delegation Interface (see Appendix E).
+The A-REX also supportd the Delegation Interface (see Appendix D).
 This is a common purpose interface to be used by ARC services which
@@
 A-REX accepts JSDL documents having the following additional elements
-(see Appendix F):
+(see Appendix E):
 
@@
 and \emph{new} (as described in section \ref{sec:cache}). Configuration
-is read directly from the GM's configuration file (see section \ref{SubSection: ConfigFile}).
-It is run by the GM every 5 minutes.
+is read directly from the A-REX's configuration file (see section
+\ref{SubSection: ConfigFile}). It is run by the A-REX every 5 minutes.
 \item \emph{frontend-info-collector} - Utility to gather information about
@@
 \item [{renew}] - old job, new credentials
-\item [{write}] - write/delete file, create/delete directory (through gridftp)
-\item [{read}] - read file, directory, etc. (through gridftp)
-\item [{extern}] - call external program (grid-manager)
+\item [{write}] - write/delete file, create/delete directory
+\item [{read}] - read file, directory, etc.
+\item [{extern}] - call external program 
 \end{lyxlist}
@@
 Authorization is performed by generic means provided by HED framework.
-Currently A-REX does not implement any internal authorization tecniques.
+Currently A-REX does not implement any internal authorization tecniques
+except those imposed by Access Policy assigned to jobs through AccessControl
+element of assigned JSDL.
 
@@
 
-The GM comes with support for several LRMS. And this number is slowly
-growing. Features explained below are for \textbf{PBS/Torque} backend.
-This support is provided through \textit{submit-pbs-job}, \textit{cancel-pbs-job},
-\textit{scan-pbs-job} scripts. \textit{submit-pbs-job} creates job's
-script and submits it to PBS. Created job's script is responsible
-for moving data between frontend machine and cluster node (if required)
-and execution of actual job. Alternatively it can download input files
-and upload output if \emph{{}``localtransfer no''} is specified
-in the configuration file.
+The A-REX comes with support for several LRMS. And this number is
+slowly growing. Features explained below are for \textbf{PBS/Torque}
+backend. This support is provided through \textit{submit-pbs-job},
+\textit{cancel-pbs-job}, \textit{scan-pbs-job} scripts. \textit{submit-pbs-job}
+creates job's script and submits it to PBS. Created job's script is
+responsible for moving data between frontend machine and cluster node
+(if required) and execution of actual job. Alternatively it can download
+input files and upload output if \emph{{}``localtransfer=no''} is
+specified in the configuration file.
 
 Behavior of submission script is mostly controlled using environment
-variables. Most of them can be specified on frontend in GM's environment
+variables. Most of them can be specified on frontend in A-REX's environment
 and overwritten on cluster's node through PBS configuration. Some
@@
 \textbf{\textit{PBS\_BIN\_PATH}} - path to PBS executables. Like \emph{/usr/local/bin}
-for example. \emph{pbs\_bin\_path} configuration command.
-
-\textbf{\textit{PBS\_LOG\_PATH}} - path to PBS server logs. \emph{pbs\_log\_path}
+for example. \emph{}Corresponds to \emph{pbs\_bin\_path} configuration
+command.
+
+\textbf{\textit{PBS\_LOG\_PATH}} - path to PBS server logs. Corresponds
+to \emph{pbs\_log\_path} configuration command.
+
+\textbf{\textit{TMP\_DIR}} - path to directory to store temporary
+files. Default value is \emph{/tmp}. Corresponds to \emph{tmpdir}
 configuration command.
 
-\textbf{\textit{TMP\_DIR}} - path to directory to store temporary
-files. Default value is \emph{/tmp}. \emph{tmpdir} configuration command.
-
 \textbf{\textit{RUNTIME\_CONFIG\_DIR}} - path where runtime setup
-scripts can be found. \emph{runtimedir} configuration command.
+scripts can be found. Corresponds to \emph{runtimedir} configuration
+command.
 
 \textbf{\textit{GNU\_TIME}} - path to GNU time utility. It is important
-to path to utility compatible with GNU time. If such utility is not
-available, modify \textit{submit-pbs-job} to either reset this variable
-or change usage of available utility. \emph{gnu\_time} configuration
-command.
+to provide path to utility compatible with GNU time. If such utility
+is not available, modify \textit{submit-pbs-job} to either reset this
+variable or change usage of available utility. Corresponds to \emph{gnu\_time}
+configuration command.
 
 \textbf{\textit{NODENAME}} - command to obtain name of cluster's node.
-Default is \emph{/bin/hostname -f}. \emph{nodename} configuration
-command.
+Default is \emph{/bin/hostname -f}. Corresponds to \emph{nodename}
+configuration command.
 
@@
 contain path corresponding to \textit{RUNTIME\_LOCAL\_SCRATCH\_DIR}
-as seen on \textbf{frontend} machine. \emph{shared\_scratch} configuration
-command.
+as seen on \textbf{frontend} machine. Corresponds to \emph{shared\_scratch}
+configuration command.
 
@@
 by the LRMS. Results are moved back after job's execution in a same
-way. \emph{shared\_filesystem} configuration command.
+way. Corresponds to \emph{shared\_filesystem} configuration command.
 
@@
 and RUNTIME\_FRONTEND\_SEES\_NODE and explain how data movement is
-performed. Pictures a) corresponds to the situation right after all
+performed. Pictures a) correspond to the situation right after all
 input files have been gathered in the session directory and show the
-actions taken right after the job's script starts. Pictures b) shows
+actions taken right after the job's script starts. Pictures b) show
 the situation while the job is running and the actions which are taken
-right after it has finished. Pictures c) illustrates the final situation,
+right after it has finished. Pictures c) illustrate the final situation,
 when the job's output files are ready to be uploaded to an external
@@
 directories are visible from computing nodes. The job is executed
-directly in the session directory prepared by GM on the frontend.}
+directly in the session directory prepared by the A-REX on the frontend.}
 \end{figure}
@@
 
-\subsection{Setup of the Grid Manager}
-
-For in-depth information about how to properly setup the GM and related
-software please read {}``NorduGrid ARC server installation instructions''
-at \url{http://www.nordugrid.org/documents/ng-server-install.html}.
-Follow that manual to install GM, configure and run it. Additional
-tips are described here.
-
-The GM is designed to be able to run both as root and as ordinary
-user. You can chose the name of the user by using corresponding command
-in configuration file. It is better run GM as root if You want to
-serve few users.
-
-The GM writes debug information into a file /var/log/grid-manager.log
-by default. . Also file /var/log/gm-jobs.log (default path in configuration
-template, turned off by default) contains information about all started
-and finished jobs, 2 lines per job (1 when job is started and 1 after
-it finished).
+\subsection{Setup of the A-REX}
+
+The A-REX service is a plugable module of the HED. Si first You need
+to setup your HED - see HED documentation.
+
+Then You need to add A-REX to the HED configuration. Add new <Name>
+element inside <Plugins> containing string \emph{arex}. That will
+make HED load libarex plugin library. Then add new <Service> element
+with attribute name=''a-rex''. That will instantiate A-REX service.
+Now to make service accessible extend <Plexer> element with new <next>
+refering to service's id. Write <Service> element carefully - here
+is an example:
+
+\begin{lyxcode}
+~~~~<!-{}-~A-Rex~service~-{}->
+
+~~~~<Service~name=\char`\"{}a-rex\char`\"{}~id=\char`\"{}a-rex\char`\"{}>
+
+~~~~~~<!-{}-~Optional~endpoint~element~is~advised~in~case~of~multiple~IP~adresses~-{}->
+
+~~~~~~<arex:endpoint>https://localhost:60000/arex</arex:endpoint>
+
+~~~~~~<!-{}-~Use~information~generated~by~identity.map~plugin~or~default~provided~below~-{}->
+
+~~~~~~<arex:usermap><arex:defaultLocalName>nobody</arex:defaultLocalName></arex:usermap>
+
+~~~~~~<!-{}-~grid-manager~part~of~a-rex~requires~legacy~configuration~file.
+
+~~~~~~~~~~~Use~arc\_arex.conf~example~or~write~own.~-{}->
+
+~~~~~~<arex:gmconfig>/etc/arc\_arex.conf</arex:gmconfig>
+
+~~~~</Service>
+\end{lyxcode}
+For in-depth information about available elements see Appendix B.
+
+Use a template arc\_arex.conf or write own A-REX configuration file.
+For information about format and available configuration commands
+see Section \ref{SubSection: ConfigFile}.
 
@@
 
-Refer to the description of the \emph{User Interface} part \cite{ui}
-and extensions to RSL \cite{xrsl} for using the GM.
+Please read \emph{User Manual {[}]} for usage instructions.
 
@@
 
-Bot GM and GFS are primarily designed to be run by the \emph{root}
-UNIX account and serve multiple global Grid identities mapped to several
+The A-REX is primarily designed to be run by the \emph{root} UNIX
+account and serve multiple global Grid identities mapped to several
 UNIX accounts. Nevertheless it is possible to use \emph{non-root}
-accounts to run these services at the cost of some functionality loss
+accounts to run that service at the cost of some functionality loss
 as described below.
 
-There are no drawbacks of running GFS with \emph{gaclplugin} or \emph{fileplugin}
-under a \emph{non-root} account as long as the only Grid identity
-used is that of the user who runs the services and all served files
-and directories are owned by the server's account.
-
-In order for GM and GFS with \emph{jobplugin} to be able to cooperate,
-both services must be run either by the same non-root account or one
-of the services must be run under the \emph{root} account. This is
-needed because the two services communicate over a local filesystem,
-hence they must have \emph{full} access to the same set of files.
-
-As long as GFS with \emph{jobplugin} is run under a non-root account,
-no mapping from Grid identity to local UNIX account takes place. All
-alowed Grid users are assigned the server's account and are then processes
-by GM using same account. The only way to overcome this limitation
-is to run one GFS per local account with proper access control configured.
-
-Because GM has to impersonate a user's local account while communicating
-with the LRMS, it can serve only the account it is run under (unless
-it is run under the \emph{root} account, of course). As in the case
-of GFS, multiple instances of GM may be run, one per local account.
-This solution has some drawbacks. The GM looses possibility to share
-cached files among serviced users. In addition, it is not possible
-to control the load on a frontend by limiting the number of simultatenuosly
-running \emph{downloader} and \emph{uploader} modules.
-
-One also has to take into account that the private part of the GSI
-infrastructure (private key of a host at least) has to be duplicated
-for every account used to run GFS.
+There are no drawbacks of running A-REX under a \emph{non-root} account
+as long as the only UNIX identity used is that of the user who runs
+the services and all served files and directories are owned by the
+server's account. Because A-REX has to impersonate a user's local
+account while communicating with the LRMS, it can serve only the account
+it is run under (unless it is run under the \emph{root} account, of
+course).
 
@@
 
-\subsection*{Virtual tree}
-
-Under the mount point of the jobplugin, a user connecting with a gridftp
-client will see virtual directories representing the jobs belonging
-to him/her. Directory names are job identifiers, each representing
-one job. These directories are directly connected to session directories
-of jobs and contain the files and subdirectories that are visible
-on the frontend.
-
-If a job's xRSL has \emph{gmlog} specified, then the job's directory
-also contains a virtual subdirectory of that name holding files with
-information about the job as created by GM. The 'errors' and 'status'
-files are especially usefull for troubleshooting. 'errors' contains
-the stderr output of the various modules run by GM during job processing
-stages (downloader, uploader, job's submission to LRMS). 'status'
-contains one word representing the job's state.
-
-Also under the jobplugin's mount point there is an additional directory
-named \char`\"{}new\char`\"{} -- used to submit new jobs -- and another
-directory named ''info''. The latter has subdirectories named after
-job ids, each of which contains files with information about a job.
-These are the same files that can be accessed in the subdirectory
-specified through \emph{gmlog}.
-
-\subsection*{Submission}
-
-Each xRSL put into the {}``new'' directory by a gridftp client is
-treated as a new job's description. GFS's jobplugin parses the xRSL
-and returns the client a positive response if there were no errors
-in the request.
-
-The new job gets an identifier and a directory with the corresponding
-name appears. If the job's description contains input files which
-should be transferred from the client's machine, the client must upload
-them to that directory under specified names.
-
-The job identifier reserved by GFS to a new job must be somehow communicated
-back to the client. Within the bounds of the FTP protocol, this is
-achieved in the following way. Prior to uploading the xRSL, the client
-issues a CWD command to change the current directory to {}``new''.
-In turn, the server responds with a redirect to the new session directory
-named after the reserved identifier. The client has now found out
-the job's id, and proceeds with uploading the xRLS.
-
-\subsection*{Actions}
-
-Various actions to affect processing of an existing job are requested
-by uploading special xRSL files into directory {}``new''. Such an
-xRSL may consist of only 2 parameters - \emph{action} for action to
-be performed, and \emph{jobid} to identify the job to be affected.
-All other parameters are ignored.
-
-The currently supported actions are:
-
-\begin{lyxlist}{00.00.0000}
-\item [{\emph{cancel}}] to cancel a job
-\item [{\emph{clean}}] to remove a job from computing resource
-\item [{\emph{renew}}] to renew credentials delegated to a job
-\item [{\emph{restart}}] to restart a job after failure at some phases
-\end{lyxlist}
-Alternatively, it is also possible to perform some of these actions
-by using the shortcut FTP operations described below.
-
-\subsubsection*{Cancel}
-
-A job is canceled by performing a DELE (delete file) command on the
-directory representing the job. It can take some time (a few minutes)
-before the job is actually canceled. Nevertheless, the client gets
-a response immediately.
-
-\subsubsection*{Clean}
-
-A job's content is cleaned by performing a RMD (remove directory)
-command on the directory representing the job. If the job is in \char`\"{}FINISHED\char`\"{}
-state it will be cleaned immediately. Otherwise it will be cleaned
-after it reaches state \char`\"{}FINISHED\char`\"{}.
-
-\subsubsection*{Renew}
-
-If a client requests CWD to a job's session directory, credentials
-passed during authentication are compared to the current credentials
-of the job. If the validity time of the new credentials is longer,
-the job's current credentials are replaced with the new ones.
-
-\section*{Appendix C. Configuration schema of A-REX}
+In addition to the BES interface A-REX provides access to the SD through
+pure HTTP(S) interface. This functionality is used for uploading user-stageable
+files during job submission and for staging out result files produced
+by job. It can also be used to monitor job execution by checking content
+of application dependant files in SD.
+
+The BES defines job identifier as WS Addressing {[}] Endpoint Reference
+(EPR) - XML document. The EPR is extendable and the A-REX adds it
+own element JobSessionDir belonging to namespace http://www.nordugrid.org/schemas/a-rex
+as direct child of ReferenceParameters element. This new element contains
+the URL of SD.
+
+Obtained URL shoud be extended with file names relative to SD and
+HTTP methods PUT and GET may be used to upload/download content of
+those files. For directories - including SD itself - GET method is
+supported shich returns HTML encoded non-recursive list of files and
+firectories. The files and subdirectories have their URLs inside HTML
+element <A>.
+
+\section*{Appendix B. Configuration schema of A-REX}
 
@@
 
-\section*{Appendix C. Error messages of A-REX}
+\section*{Appendix C. A-REX WSDL}
+
+\begin{lyxcode}
+<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
+
+<wsdl:definitions~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}
+
+~xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
+
+~xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
+
+~xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
+
+~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
+
+~xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
+
+~xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
+
+~xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
+
+~xmlns:bes-factory=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}
+
+~xmlns:bes-mgmt=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}
+
+~xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
+
+~xmlns:wsrf-rpw=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}
+
+~xmlns:a-rex=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
+
+~~<wsdl:import~namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}~location=\char`\"{}./bes-factory.wsdl\char`\"{}/>
+
+~~<wsdl:import~namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}~location=\char`\"{}./bes-management.wsdl\char`\"{}/>
+
+~~<wsdl:import~namespase=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}~location=\char`\"{}../schemas/delegation.wsdl\char`\"{}/>
+
+~~<wsdl:import~namespase=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}~location=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2.wsdl\char`\"{}/>
+
+~~<wsdl:types>
+
+~~~~<xsd:schema~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
+
+~~~~~~<xsd:import~namespace=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}~schemaLocation=\char`\"{}./ws-addr.xsd\char`\"{}/>
+
+~~~~~~<xsd:simpleType~name=\char`\"{}ActivitySubStateType\char`\"{}>
+
+~~~~~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Accepting\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Accepted\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Preparing\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Prepared\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Submiting\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Executing\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Killing\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Executed\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Finishing\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Finished\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Failed\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Deleted\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Pending\char`\"{}/>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Held\char`\"{}/>
+
+~~~~~~~~</xsd:restriction>
+
+~~~~~~</xsd:simpleType>
+
+~~~~~~<xsd:element~name=\char`\"{}State\char`\"{}~type=\char`\"{}a-rex:ActivitySubStateType\char`\"{}/>
+
+~~~~~~<xsd:complexType~name=\char`\"{}ResourceInformationDocumentType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~~<xsd:element~name=\char`\"{}BESFactory\char`\"{}~type=\char`\"{}bes-factory:FactoryResourceAttributesDocumentType\char`\"{}/>
+
+~~~~~~~~~~<xsd:complexType~name=\char`\"{}Glue2Resource\char`\"{}~minOccurs='0'>
+
+~~~~~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
+
+~~~~~~~~~~~~</xsd:sequence>
+
+~~~~~~~~~~</xsd:complexType>
+
+~~~~~~~~~~<xsd:complexType~name=\char`\"{}Activities\char`\"{}~minOccurs='0'>
+
+~~~~~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~~~~~<xsd:complexType~name=\char`\"{}Activity\char`\"{}~minOccurs='0'~maxOccurs='unbounded'>
+
+~~~~~~~~~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~~~~~~~~~<xsd:element~name=\char`\"{}ActivityIdentifier\char`\"{}~type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
+
+~~~~~~~~~~~~~~~~~~<xsd:element~ref=\char`\"{}bes-factory:ActivityDocument\char`\"{}~minOccurs='0'/>
+
+~~~~~~~~~~~~~~~~~~<xsd:complexType~name=\char`\"{}Glue2Job\char`\"{}~minOccurs='0'>
+
+~~~~~~~~~~~~~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~~~~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
+
+~~~~~~~~~~~~~~~~~~~~</xsd:sequence>
+
+~~~~~~~~~~~~~~~~~~</xsd:complexType>
+
+~~~~~~~~~~~~~~~~</xsd:sequence>
+
+~~~~~~~~~~~~~~</xsd:complexType>
+
+~~~~~~~~~~~~</xsd:sequence>
+
+~~~~~~~~~~</xsd:complexType>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:complexType~name=\char`\"{}ChangeActivityStatusRequestType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}ActivityIdentifier\char`\"{}~type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}OldStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}NewStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}ChangeActivityStatus\char`\"{}~type=\char`\"{}a-rex:ChangeActivityStatusRequestType\char`\"{}/>
+
+~~~~~~<xsd:complexType~name=\char`\"{}ChangeActivityStatusResponseType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}NewStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}~type=\char`\"{}a-rex:ChangeActivityStatusResponseType\char`\"{}/>
+
+~~~~</xsd:schema>
+
+~~</wsdl:types>
+
+~~<wsdl:message~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}~element=\char`\"{}a-rex:ChangeActivityStatus\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:message~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}~element=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:portType~name=\char`\"{}a-rex\char`\"{}>
+
+~~~~<wsdl:operation~name=\char`\"{}ChangeActivityStatus\char`\"{}>
+
+~~~~~~<wsdl:documentation>
+
+~~~~~~~~This~operation~allows~any~simple~status~change~request
+
+~~~~~~~~which~involves~no~additional~parameters.~It~should~be
+
+~~~~~~~~used~to~modify~job/activity~execution~flow:
+
+~~~~~~~~~~-~To~put~job~on~hold
+
+~~~~~~~~~~-~To~rerun~job~in~case~of~failure
+
+~~~~~~~~~~-~To~cancel~job~(same~as~TerminateActivity~of~BESFActory)
+
+~~~~~~~~~~-~To~remove/release~job~-~as~long~as~non-existence~is~a~state
+
+~~~~~~~~~~-~Any~other~status~change~no~supported~by~BES
+
+~~~~~~</wsdl:documentation>
+
+~~~~~~<wsdl:input~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}
+
+~~~~~~~~message=\char`\"{}a-rex:ChangeActivityStatusRequest\char`\"{}/>
+
+~~~~~~<wsdl:output~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
+
+~~~~~~~~message=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}
+
+~~~~~~~~message=\char`\"{}bes-factory:NotAuthorizedFault\char`\"{}
+
+~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}
+
+~~~~~~~~message=\char`\"{}bes-factory:InvalidActivityIdentifierFault\char`\"{}
+
+~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}
+
+~~~~~~~~~message=\char`\"{}bes-factory:CantApplyOperationToCurrentStateFault\char`\"{}
+
+~~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}
+
+~~~~~~~~~message=\char`\"{}bes-factory:OperationWillBeAppliedEventuallyFault\char`\"{}
+
+~~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+
+~~~~</wsdl:operation>
+
+~~</wsdl:portType>
+
+~~<wsdl:binding~name=\char`\"{}a-rex\char`\"{}~type=\char`\"{}a-rex:a-rex\char`\"{}>
+
+~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+
+~~~~<wsdl:operation~name=\char`\"{}ChangeActivityStatus\char`\"{}>
+
+~~~~~~<soap:operation~soapAction=\char`\"{}ChangeActivityStatus\char`\"{}/>
+
+~~~~~~<wsdl:input~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
+
+~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:input>
+
+~~~~~~<wsdl:output~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
+
+~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:output>
+
+~~~~~~<wsdl:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~</wsdl:operation>
+
+~~</wsdl:binding>
+
+~~<wsdl:binding~name=\char`\"{}GetResourcePropertyDocument\char`\"{}~type=\char`\"{}wsrf-rpw:GetResourcePropertyDocument\char`\"{}>
+
+~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+
+~~~~<wsdl:operation~name=\char`\"{}GetResourcePropertyDocument\char`\"{}>
+
+~~~~~~<soap:operation~soapAction=\char`\"{}GetResourcePropertyDocument\char`\"{}/>
+
+~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentRequest\char`\"{}>
+
+~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:input>
+
+~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentResponse\char`\"{}>
+
+~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:output>
+
+~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~</wsdl:operation>
+
+~~</wsdl:binding>
+
+~~<wsdl:binding~name=\char`\"{}GetResourceProperty\char`\"{}~type=\char`\"{}wsrf-rpw:GetResourceProperty\char`\"{}>
+
+~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+
+~~~~<wsdl:operation~name=\char`\"{}GetResourceProperty\char`\"{}>
+
+~~~~~~<soap:operation~soapAction=\char`\"{}GetResourceProperty\char`\"{}/>
+
+~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:GetResourcePropertyRequest\char`\"{}>
+
+~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:input>
+
+~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:GetResourcePropertyResponse\char`\"{}>
+
+~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:output>
+
+~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~</wsdl:operation>
+
+~~</wsdl:binding>
+
+~~<wsdl:binding~name=\char`\"{}QueryResourceProperties\char`\"{}~type=\char`\"{}wsrf:QueryResourceProperties\char`\"{}>
+
+~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+
+~~~~<wsdl:operation~name=\char`\"{}QueryResourceProperties\char`\"{}>
+
+~~~~~~<soap:operation~soapAction=\char`\"{}QueryResourceProperties\char`\"{}/>
+
+~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:QueryResourcePropertiesRequest\char`\"{}>
+
+~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:input>
+
+~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:QueryResourcePropertiesResponse\char`\"{}>
+
+~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:output>
+
+~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}InvalidQueryExpressionFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}InvalidQueryExpressionFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~~~<wsdl:fault~name=\char`\"{}QueryEvaluationErrorFault\char`\"{}>
+
+~~~~~~~~<soap:fault~name=\char`\"{}QueryEvaluationErrorFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
+
+~~~~~~</wsdl:fault>
+
+~~~~</wsdl:operation>
+
+~~</wsdl:binding>
+
+~~<wsdl:service~name=\char`\"{}a-rex\char`\"{}>
+
+~~~~<wsdl:port~name=\char`\"{}delegation\char`\"{}~binding=\char`\"{}deleg:DelegationBinding\char`\"{}>
+
+~~~~</wsdl:port>
+
+~~~~<wsdl:port~name=\char`\"{}bes-factory\char`\"{}~binding=\char`\"{}bes-factory:BESFactoryBinding\char`\"{}>
+
+~~~~</wsdl:port>
+
+~~~~<wsdl:port~name=\char`\"{}bes-mgmt\char`\"{}~binding=\char`\"{}bes-mgmt:BESManagementBinding\char`\"{}>
+
+~~~~</wsdl:port>
+
+~~~~<wsdl:port~name=\char`\"{}GetResourcePropertyDocument\char`\"{}~binding=\char`\"{}a-rex:GetResourcePropertyDocument\char`\"{}>
+
+~~~~</wsdl:port>
+
+~~~~<wsdl:port~name=\char`\"{}GetResourceProperty\char`\"{}~binding=\char`\"{}a-rex:GetResourceProperty\char`\"{}>
+
+~~~~</wsdl:port>
+
+~~~~<wsdl:port~name=\char`\"{}QueryResourceProperties\char`\"{}~binding=\char`\"{}a-rex:QueryResourceProperties\char`\"{}>
+
+~~~~</wsdl:port>
+
+~~~~<wsdl:port~name=\char`\"{}a-rex\char`\"{}~binding=\char`\"{}a-rex:a-rex\char`\"{}>
+
+~~~~</wsdl:port>
+
+~~</wsdl:service>
+
+</wsdl:definitions>
+\end{lyxcode}
+
+\section*{Appendix D. Delegation WSDL}
+
+\begin{lyxcode}
+<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
+
+<wsdl:definitions~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
+
+~xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
+
+~xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
+
+~xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
+
+~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
+
+~xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
+
+~xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
+
+~xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
+
+~xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
+
+~~<wsdl:types>
+
+~~~~<xsd:schema~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
+
+~~~~~~<!-{}-~Common~types~-{}->
+
+~~~~~~<xsd:simpleType~name=\char`\"{}TokenFormatType\char`\"{}>
+
+~~~~~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
+
+~~~~~~~~~~<xsd:enumeration~value=\char`\"{}x509\char`\"{}/>
+
+~~~~~~~~</xsd:restriction>
+
+~~~~~~</xsd:simpleType>
+
+~~~~~~<xsd:complexType~name=\char`\"{}ReferenceType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:complexType~name=\char`\"{}DelegatedTokenType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}Id\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}Value\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}Reference\char`\"{}~type=\char`\"{}deleg:ReferenceType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~~~<xsd:attribute~name=\char`\"{}Format\char`\"{}~type=\char`\"{}deleg:TokenFormatType\char`\"{}~use=\char`\"{}required\char`\"{}/>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}DelegatedToken\char`\"{}~type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
+
+~~~~~~<xsd:complexType~name=\char`\"{}TokenRequestType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}Id\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}Value\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~~~<xsd:attribute~name=\char`\"{}Format\char`\"{}~type=\char`\"{}deleg:TokenFormatType\char`\"{}~use=\char`\"{}required\char`\"{}/>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}TokenRequest\char`\"{}~type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
+
+~~~~~~<!-{}-~Types~for~messages~-{}->
+
+~~~~~~<xsd:complexType~name=\char`\"{}DelegateCredentialsInitRequestType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}DelegateCredentialsInit\char`\"{}~type=\char`\"{}deleg:DelegateCredentialsInitRequestType\char`\"{}/>
+
+~~~~~~<xsd:complexType~name=\char`\"{}DelegateCredentialsInitResponseType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}TokenRequest\char`\"{}~type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}~type=\char`\"{}deleg:DelegateCredentialsInitResponseType\char`\"{}/>
+
+~~~~~~<xsd:complexType~name=\char`\"{}UpdateCredentialsRequestType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}DelegatedToken\char`\"{}~type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}UpdateCredentials\char`\"{}~type=\char`\"{}deleg:UpdateCredentialsRequestType\char`\"{}/>
+
+~~~~~~<xsd:complexType~name=\char`\"{}UpdateCredentialsResponseType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}UpdateCredentialsResponse\char`\"{}~type=\char`\"{}deleg:UpdateCredentialsResponseType\char`\"{}/>
+
+~~~~~~<!-{}-~Faults~-{}->
+
+~~~~~~<xsd:complexType~name=\char`\"{}UnsupportedFaultType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}UnsupportedFault\char`\"{}~type=\char`\"{}deleg:UnsupportedFaultType\char`\"{}/>
+
+~~~~~~<xsd:complexType~name=\char`\"{}ProcessingFaultType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}ProcessingFault\char`\"{}~type=\char`\"{}deleg:ProcessingFaultType\char`\"{}/>
+
+~~~~~~<xsd:complexType~name=\char`\"{}WrongReferenceFaultType\char`\"{}>
+
+~~~~~~~~<xsd:sequence>
+
+~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~~~~~~</xsd:sequence>
+
+~~~~~~</xsd:complexType>
+
+~~~~~~<xsd:element~name=\char`\"{}WrongReferenceFault\char`\"{}~type=\char`\"{}deleg:WrongReferenceFaultType\char`\"{}/>
+
+~~~~</xsd:schema>
+
+~~</wsdl:types>
+
+~~<wsdl:message~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}~element=\char`\"{}deleg:DelegateCredentialsInit\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:message~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}~element=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:message~name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}UpdateCredentialsRequest\char`\"{}~element=\char`\"{}deleg:UpdateCredentials\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:message~name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}UpdateCredentialsResponse\char`\"{}~element=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:message~name=\char`\"{}UnsupportedFault\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:message~name=\char`\"{}ProcessingFault\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:ProcessingFault\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:message~name=\char`\"{}WrongReferenceFault\char`\"{}>
+
+~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
+
+~~</wsdl:message>
+
+~~<wsdl:portType~name=\char`\"{}DelegationPortType\char`\"{}>
+
+~~~<wsdl:operation~name=\char`\"{}DelegateCredentialsInit\char`\"{}>
+
+~~~~~~<wsdl:documentation>
+
+~~~~~~</wsdl:documentation>
+
+~~~~~~<wsdl:input~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:DelegateCredentialsInitRequest\char`\"{}/>
+
+~~~~~~<wsdl:output~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}UnsupportedFault\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}ProcessingFault\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
+
+~~~~</wsdl:operation>
+
+~~~<wsdl:operation~name=\char`\"{}UpdateCredentials\char`\"{}>
+
+~~~~~~<wsdl:documentation>
+
+~~~~~~</wsdl:documentation>
+
+~~~~~~<wsdl:input~name=\char`\"{}UpdateCredentialsRequest\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:UpdateCredentialsRequest\char`\"{}/>
+
+~~~~~~<wsdl:output~name=\char`\"{}UpdateCredentialsResponse\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}UnsupportedFault\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}ProcessingFault\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
+
+~~~~~~<wsdl:fault~name=\char`\"{}WrongReferenceFault\char`\"{}
+
+~~~~~~~~message=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
+
+~~~~</wsdl:operation>
+
+~~</wsdl:portType>
+
+~~<wsdl:binding~name=\char`\"{}DelegationBinding\char`\"{}~type=\char`\"{}deleg:DelegationPortType\char`\"{}>
+
+~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+
+~~~~<wsdl:operation~name=\char`\"{}DelegateCredentialsInit\char`\"{}>
+
+~~~~~~<soap:operation~soapAction=\char`\"{}DelegateCredentialsInit\char`\"{}/>
+
+~~~~~~<wsdl:input~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
+
+~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:input>
+
+~~~~~~<wsdl:output~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
+
+~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:output>
+
+~~~~</wsdl:operation>
+
+~~~~<wsdl:operation~name=\char`\"{}UpdateCredentials\char`\"{}>
+
+~~~~~~<soap:operation~soapAction=\char`\"{}UpdateCredentials\char`\"{}/>
+
+~~~~~~<wsdl:input~name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
+
+~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:input>
+
+~~~~~~<wsdl:output~name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
+
+~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
+
+~~~~~~</wsdl:output>
+
+~~~~</wsdl:operation>
+
+~~</wsdl:binding>
+
+</wsdl:definitions>
+\end{lyxcode}
+
+\section*{Appendix E. ARC extensions for JSDL schema}
+
+\begin{lyxcode}
+<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
+
+<xsd:schema~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
+
+~~~~~~~~~~~~xmlns=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
+
+~~~~~~~~~~~~xmlns:jsdl-arc=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
+
+~~~~~~~~~~~~targetNamespace=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}>
+
+~<xsd:simpleType~name=\char`\"{}GMState\_Type\char`\"{}>
+
+~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
+
+~~~<xsd:enumeration~value=\char`\"{}ACCEPTED\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}PREPARING\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}SUBMIT\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}INLRMS\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}FINISHING\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}FINISHED\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}DELETED\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}CANCELING\char`\"{}/>
+
+~~</xsd:restriction>
+
+~</xsd:simpleType>
+
+~<xsd:complexType~name=\char`\"{}Version\_Type\char`\"{}>
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}UpperExclusive\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}LowerExclusive\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}Exact\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}Exclusive\char`\"{}~type=\char`\"{}xsd:boolean\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:simpleType~name=\char`\"{}SessionType\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}SessionType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
+
+~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
+
+~~~<xsd:enumeration~value=\char`\"{}INTERNAL\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}LIMITED\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}READONLY\char`\"{}/>
+
+~~~<xsd:enumeration~value=\char`\"{}FULL\char`\"{}/>
+
+~~</xsd:restriction>
+
+~</xsd:simpleType>
+
+~<xsd:simpleType~name=\char`\"{}IsExecutable\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:DataStaging\_Type~(default:~false)~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}IsExecutable\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
+
+~~<xsd:restriction~base=\char`\"{}xsd:boolean\char`\"{}/>
+
+~</xsd:simpleType>
+
+~<xsd:simpleType~name=\char`\"{}FileParameters\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:DataStaging\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}IsExecutable\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
+
+~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}/>
+
+~</xsd:simpleType>
+
+~<xsd:simpleType~name=\char`\"{}JoinOutputs\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:JobDescription\_Type~(default:~false)~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}JoinOutputs\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
+
+~~<xsd:restriction~base=\char`\"{}xsd:boolean\char`\"{}/>
+
+~</xsd:simpleType>
+
+~<xsd:simpleType~name=\char`\"{}Reruns\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:JobDescription\_Type~(default:~false)~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}Reruns\char`\"{}~minOccurs=\char`\"{}0\char`\"{}//~-{}->
+
+~~<xsd:restriction~base=\char`\"{}xsd:integer\char`\"{}/>
+
+~</xsd:simpleType>
+
+~<xsd:complexType~name=\char`\"{}RunTimeEnvironment\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}RunTimeEnvironment\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/~-{}->
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}Name\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}Version\char`\"{}~type=\char`\"{}Version\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:complexType~name=\char`\"{}Middleware\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}Middleware\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/~-{}->
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}Name\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}Version\char`\"{}~type=\char`\"{}Version\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:complexType~name=\char`\"{}RemoteLogging\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}RemoteLogging\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}3\char`\"{}/~-{}->
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}URL\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:anyURI\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:complexType~name=\char`\"{}CredentialServer\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}CredentialServer\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}URL\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:anyURI\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:complexType~name=\char`\"{}LocalLogging\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}LocalLogging\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}Directory\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:simpleType~name=\char`\"{}AccessControlType\_Type\char`\"{}>
+
+~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
+
+~~~<xsd:enumeration~value=\char`\"{}GACL\char`\"{}/>
+
+~~</xsd:restriction>
+
+~</xsd:simpleType>
+
+~<xsd:complexType~name=\char`\"{}AccessControl\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}AccessControl\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}OwnerAlwaysAllowed\char`\"{}~type=\char`\"{}xsd:boolean\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}Type\char`\"{}~type=\char`\"{}AccessControlType\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}Content\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:simpleType~name=\char`\"{}NotificationType\_Type\char`\"{}>
+
+~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
+
+~~~<xsd:enumeration~value=\char`\"{}Email\char`\"{}/>
+
+~~</xsd:restriction>
+
+~</xsd:simpleType>
+
+~<xsd:complexType~name=\char`\"{}Notify\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}Notify\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}3\char`\"{}/~-{}->
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}Type\char`\"{}~type=\char`\"{}NotificationType\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}Endpoint\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}State\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}~type=\char`\"{}GMState\_Type\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:simpleType~name=\char`\"{}SessionLifeTime\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}SessionLifeTime\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
+
+~~<xsd:restriction~base=\char`\"{}xsd:long\char`\"{}/>
+
+~</xsd:simpleType>
+
+~<xsd:simpleType~name=\char`\"{}GridTimeLimit\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}GridTimeLimit\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
+
+~~<xsd:restriction~base=\char`\"{}xsd:positiveInteger\char`\"{}/>
+
+~</xsd:simpleType>
+
+~<xsd:complexType~name=\char`\"{}CandidateTarget\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}jsdl-arc:CandidateTarget\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
+
+~~<xsd:sequence>
+
+~~~<xsd:element~name=\char`\"{}HostName\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~~<xsd:element~name=\char`\"{}QueueName\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
+
+~~</xsd:sequence>
+
+~</xsd:complexType>
+
+~<xsd:simpleType~name=\char`\"{}Time\_Type\char`\"{}>
+
+~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
+
+~~<!-{}-~xsd:element~ref=\char`\"{}ProcessingStartTime\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
+
+~~<xsd:restriction~base=\char`\"{}xsd:dateTime\char`\"{}/>
+
+~</xsd:simpleType>
+
+~<!-{}-=========================================================-{}->
+
+~<xsd:element~name=\char`\"{}IsExecutable\char`\"{}~type=\char`\"{}IsExecutable\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}FileParameters\char`\"{}~type=\char`\"{}FileParameters\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}RunTimeEnvironment\char`\"{}~type=\char`\"{}RunTimeEnvironment\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}Middleware\char`\"{}~type=\char`\"{}Middleware\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}RemoteLogging\char`\"{}~type=\char`\"{}RemoteLogging\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}LocalLogging\char`\"{}~type=\char`\"{}LocalLogging\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}AccessControl\char`\"{}~type=\char`\"{}AccessControl\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}Notify\char`\"{}~type=\char`\"{}Notify\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}SessionLifeTime\char`\"{}~type=\char`\"{}SessionLifeTime\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}SessionType\char`\"{}~type=\char`\"{}SessionType\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}JoinOutputs\char`\"{}~type=\char`\"{}JoinOutputs\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}Reruns\char`\"{}~type=\char`\"{}Reruns\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}CredentialServer\char`\"{}~type=\char`\"{}CredentialServer\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}GridTimeLimit\char`\"{}~type=\char`\"{}GridTimeLimit\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}CandidateTarget\char`\"{}~type=\char`\"{}CandidateTarget\_Type\char`\"{}/>
+
+~<xsd:element~name=\char`\"{}ProcessingStartTime\char`\"{}~type=\char`\"{}Time\_Type\char`\"{}/>
+
+</xsd:schema>
+\end{lyxcode}
+
+\section*{Appendix F. Error messages of A-REX}
 
@@
 User requested to cancel the job&
-GM detected external request to cancel this job, most probably issued
-by user\tabularnewline
+A-REX detected external request to cancel this job, most probably
+issued by user\tabularnewline
 \hline
@@
 Plugin at state STATE failed: OUTPUT&
-External plugin specified in GM's configuration returned non-zero
+External plugin specified in A-REX configuration returned non-zero
 exit code. STATE is replcaced by name of state to which job was going
@@
 Failed running plugin at state STATE&
-External plugin specified in GM's configuration could not be executed.\tabularnewline
+External plugin specified in A-REX configuration could not be executed.\tabularnewline
 \hline
@@
 Input file: URL - can't use local cache&
-Problems with GM cache \tabularnewline
+Problems with A-REX cache \tabularnewline
 \hline 
@@
 User file: FILENAME - Expected file. Directory found.&
-Instead of file uploadable by user interface GM found directory with
-same name in a session directory.\tabularnewline
+Instead of file uploadable by user interface A-REX found directory
+with same name in a session directory.\tabularnewline
 \hline 
 User file: FILENAME - Expected ordinary file. Special object found.&
-Instead of file uploadable by user interface GM found special object
+Instead of file uploadable by user interface A-REX found special object
 with same name in a session directory.\tabularnewline
@@
 User file: FILENAME - Delivered file is bigger than specified.&
-The size of file uploadable by user interface is bigger than specified
-in job description.\tabularnewline
+The size of file uploadable by user interface is bigger \tabularnewline
 \hline
 User file: FILENAME - Delivered file is unreadable.&
-GM can't check user uploadable file due to some internal error. Most
-probably due to improperly configured local permissions.\tabularnewline
+A-REX can't check user uploadable file due to some internal error.
+Most probably due to improperly configured local permissions.\tabularnewline
 \hline
 User file: FILENAME - Could not read file to compute checksum.&
-GM can't read user uploadable file due to some internal error. Most
-probably due to improperly configured local permissions.\tabularnewline
+A-REX can't read user uploadable file due to some internal error.
+Most probably due to improperly configured local permissions.\tabularnewline
 \hline
 User file: FILENAME - Timeout waiting&
-GM waited for user uploadable file too long.\tabularnewline
+A-REX waited for user uploadable file too long.\tabularnewline
 \hline
@@
 \begin{lyxcode}
-~
-\end{lyxcode}
-
-\section*{Appendix D. A-REX WSDL}
-
-\begin{lyxcode}
-<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
-
-<wsdl:definitions~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}
-
-~xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
-
-~xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
-
-~xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
-
-~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-
-~xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
-
-~xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
-
-~xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
-
-~xmlns:bes-factory=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}
-
-~xmlns:bes-mgmt=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}
-
-~xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
-
-~xmlns:wsrf-rpw=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}
-
-~xmlns:a-rex=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
-
-~~<wsdl:import~namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}~location=\char`\"{}./bes-factory.wsdl\char`\"{}/>
-
-~~<wsdl:import~namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}~location=\char`\"{}./bes-management.wsdl\char`\"{}/>
-
-~~<wsdl:import~namespase=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}~location=\char`\"{}../schemas/delegation.wsdl\char`\"{}/>
-
-~~<wsdl:import~namespase=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}~location=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2.wsdl\char`\"{}/>
-
-~~<wsdl:types>
-
-~~~~<xsd:schema~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
-
-~~~~~~<xsd:import~namespace=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}~schemaLocation=\char`\"{}./ws-addr.xsd\char`\"{}/>
-
-~~~~~~<xsd:simpleType~name=\char`\"{}ActivitySubStateType\char`\"{}>
-
-~~~~~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Accepting\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Accepted\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Preparing\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Prepared\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Submiting\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Executing\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Killing\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Executed\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Finishing\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Finished\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Failed\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Deleted\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Pending\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Held\char`\"{}/>
-
-~~~~~~~~</xsd:restriction>
-
-~~~~~~</xsd:simpleType>
-
-~~~~~~<xsd:element~name=\char`\"{}State\char`\"{}~type=\char`\"{}a-rex:ActivitySubStateType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ResourceInformationDocumentType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~<xsd:element~name=\char`\"{}BESFactory\char`\"{}~type=\char`\"{}bes-factory:FactoryResourceAttributesDocumentType\char`\"{}/>
-
-~~~~~~~~~~<xsd:complexType~name=\char`\"{}Glue2Resource\char`\"{}~minOccurs='0'>
-
-~~~~~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~~~~~~~~~~</xsd:sequence>
-
-~~~~~~~~~~</xsd:complexType>
-
-~~~~~~~~~~<xsd:complexType~name=\char`\"{}Activities\char`\"{}~minOccurs='0'>
-
-~~~~~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~~~~<xsd:complexType~name=\char`\"{}Activity\char`\"{}~minOccurs='0'~maxOccurs='unbounded'>
-
-~~~~~~~~~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~~~~~~~~<xsd:element~name=\char`\"{}ActivityIdentifier\char`\"{}~type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
-
-~~~~~~~~~~~~~~~~~~<xsd:element~ref=\char`\"{}bes-factory:ActivityDocument\char`\"{}~minOccurs='0'/>
-
-~~~~~~~~~~~~~~~~~~<xsd:complexType~name=\char`\"{}Glue2Job\char`\"{}~minOccurs='0'>
-
-~~~~~~~~~~~~~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~~~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~~~~~~~~~~~~~~~~~~</xsd:sequence>
-
-~~~~~~~~~~~~~~~~~~</xsd:complexType>
-
-~~~~~~~~~~~~~~~~</xsd:sequence>
-
-~~~~~~~~~~~~~~</xsd:complexType>
-
-~~~~~~~~~~~~</xsd:sequence>
-
-~~~~~~~~~~</xsd:complexType>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ChangeActivityStatusRequestType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}ActivityIdentifier\char`\"{}~type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}OldStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}NewStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}ChangeActivityStatus\char`\"{}~type=\char`\"{}a-rex:ChangeActivityStatusRequestType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ChangeActivityStatusResponseType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}NewStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}~type=\char`\"{}a-rex:ChangeActivityStatusResponseType\char`\"{}/>
-
-~~~~</xsd:schema>
-
-~~</wsdl:types>
-
-~~<wsdl:message~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}~element=\char`\"{}a-rex:ChangeActivityStatus\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}~element=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:portType~name=\char`\"{}a-rex\char`\"{}>
-
-~~~~<wsdl:operation~name=\char`\"{}ChangeActivityStatus\char`\"{}>
-
-~~~~~~<wsdl:documentation>
-
-~~~~~~~~This~operation~allows~any~simple~status~change~request
-
-~~~~~~~~which~involves~no~additional~parameters.~It~should~be
-
-~~~~~~~~used~to~modify~job/activity~execution~flow:
-
-~~~~~~~~~~-~To~put~job~on~hold
-
-~~~~~~~~~~-~To~rerun~job~in~case~of~failure
-
-~~~~~~~~~~-~To~cancel~job~(same~as~TerminateActivity~of~BESFActory)
-
-~~~~~~~~~~-~To~remove/release~job~-~as~long~as~non-existence~is~a~state
-
-~~~~~~~~~~-~Any~other~status~change~no~supported~by~BES
-
-~~~~~~</wsdl:documentation>
-
-~~~~~~<wsdl:input~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}
-
-~~~~~~~~message=\char`\"{}a-rex:ChangeActivityStatusRequest\char`\"{}/>
-
-~~~~~~<wsdl:output~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
-
-~~~~~~~~message=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}bes-factory:NotAuthorizedFault\char`\"{}
-
-~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}bes-factory:InvalidActivityIdentifierFault\char`\"{}
-
-~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}
-
-~~~~~~~~~message=\char`\"{}bes-factory:CantApplyOperationToCurrentStateFault\char`\"{}
-
-~~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}
-
-~~~~~~~~~message=\char`\"{}bes-factory:OperationWillBeAppliedEventuallyFault\char`\"{}
-
-~~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:portType>
-
-~~<wsdl:binding~name=\char`\"{}a-rex\char`\"{}~type=\char`\"{}a-rex:a-rex\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}ChangeActivityStatus\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}ChangeActivityStatus\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
-
-~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
-
-~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~~~<wsdl:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-~~<wsdl:binding~name=\char`\"{}GetResourcePropertyDocument\char`\"{}~type=\char`\"{}wsrf-rpw:GetResourcePropertyDocument\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}GetResourcePropertyDocument\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}GetResourcePropertyDocument\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentRequest\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentResponse\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-~~<wsdl:binding~name=\char`\"{}GetResourceProperty\char`\"{}~type=\char`\"{}wsrf-rpw:GetResourceProperty\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}GetResourceProperty\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}GetResourceProperty\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:GetResourcePropertyRequest\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:GetResourcePropertyResponse\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-~~<wsdl:binding~name=\char`\"{}QueryResourceProperties\char`\"{}~type=\char`\"{}wsrf:QueryResourceProperties\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}QueryResourceProperties\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}QueryResourceProperties\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:QueryResourcePropertiesRequest\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:QueryResourcePropertiesResponse\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidQueryExpressionFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}InvalidQueryExpressionFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}QueryEvaluationErrorFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}QueryEvaluationErrorFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-~~<wsdl:service~name=\char`\"{}a-rex\char`\"{}>
-
-~~~~<wsdl:port~name=\char`\"{}delegation\char`\"{}~binding=\char`\"{}deleg:DelegationBinding\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}bes-factory\char`\"{}~binding=\char`\"{}bes-factory:BESFactoryBinding\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}bes-mgmt\char`\"{}~binding=\char`\"{}bes-mgmt:BESManagementBinding\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}GetResourcePropertyDocument\char`\"{}~binding=\char`\"{}a-rex:GetResourcePropertyDocument\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}GetResourceProperty\char`\"{}~binding=\char`\"{}a-rex:GetResourceProperty\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}QueryResourceProperties\char`\"{}~binding=\char`\"{}a-rex:QueryResourceProperties\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}a-rex\char`\"{}~binding=\char`\"{}a-rex:a-rex\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~</wsdl:service>
-
-</wsdl:definitions>
-\end{lyxcode}
-
-\section*{Appendix E. Delegation WSDL}
-
-\begin{lyxcode}
-<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
-
-<wsdl:definitions~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
-
-~xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
-
-~xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
-
-~xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
-
-~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-
-~xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
-
-~xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
-
-~xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
-
-~xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
-
-~~<wsdl:types>
-
-~~~~<xsd:schema~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
-
-~~~~~~<!-{}-~Common~types~-{}->
-
-~~~~~~<xsd:simpleType~name=\char`\"{}TokenFormatType\char`\"{}>
-
-~~~~~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}x509\char`\"{}/>
-
-~~~~~~~~</xsd:restriction>
-
-~~~~~~</xsd:simpleType>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ReferenceType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:complexType~name=\char`\"{}DelegatedTokenType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Id\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Value\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Reference\char`\"{}~type=\char`\"{}deleg:ReferenceType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~~~<xsd:attribute~name=\char`\"{}Format\char`\"{}~type=\char`\"{}deleg:TokenFormatType\char`\"{}~use=\char`\"{}required\char`\"{}/>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}DelegatedToken\char`\"{}~type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}TokenRequestType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Id\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Value\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~~~<xsd:attribute~name=\char`\"{}Format\char`\"{}~type=\char`\"{}deleg:TokenFormatType\char`\"{}~use=\char`\"{}required\char`\"{}/>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}TokenRequest\char`\"{}~type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
-
-~~~~~~<!-{}-~Types~for~messages~-{}->
-
-~~~~~~<xsd:complexType~name=\char`\"{}DelegateCredentialsInitRequestType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}DelegateCredentialsInit\char`\"{}~type=\char`\"{}deleg:DelegateCredentialsInitRequestType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}DelegateCredentialsInitResponseType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}TokenRequest\char`\"{}~type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}~type=\char`\"{}deleg:DelegateCredentialsInitResponseType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}UpdateCredentialsRequestType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}DelegatedToken\char`\"{}~type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}UpdateCredentials\char`\"{}~type=\char`\"{}deleg:UpdateCredentialsRequestType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}UpdateCredentialsResponseType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}UpdateCredentialsResponse\char`\"{}~type=\char`\"{}deleg:UpdateCredentialsResponseType\char`\"{}/>
-
-~~~~~~<!-{}-~Faults~-{}->
-
-~~~~~~<xsd:complexType~name=\char`\"{}UnsupportedFaultType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}UnsupportedFault\char`\"{}~type=\char`\"{}deleg:UnsupportedFaultType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ProcessingFaultType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}ProcessingFault\char`\"{}~type=\char`\"{}deleg:ProcessingFaultType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}WrongReferenceFaultType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}WrongReferenceFault\char`\"{}~type=\char`\"{}deleg:WrongReferenceFaultType\char`\"{}/>
-
-~~~~</xsd:schema>
-
-~~</wsdl:types>
-
-~~<wsdl:message~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}~element=\char`\"{}deleg:DelegateCredentialsInit\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}~element=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}UpdateCredentialsRequest\char`\"{}~element=\char`\"{}deleg:UpdateCredentials\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}UpdateCredentialsResponse\char`\"{}~element=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}UnsupportedFault\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}ProcessingFault\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:ProcessingFault\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}WrongReferenceFault\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:portType~name=\char`\"{}DelegationPortType\char`\"{}>
-
-~~~<wsdl:operation~name=\char`\"{}DelegateCredentialsInit\char`\"{}>
-
-~~~~~~<wsdl:documentation>
-
-~~~~~~</wsdl:documentation>
-
-~~~~~~<wsdl:input~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:DelegateCredentialsInitRequest\char`\"{}/>
-
-~~~~~~<wsdl:output~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}UnsupportedFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ProcessingFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
-
-~~~~</wsdl:operation>
-
-~~~<wsdl:operation~name=\char`\"{}UpdateCredentials\char`\"{}>
-
-~~~~~~<wsdl:documentation>
-
-~~~~~~</wsdl:documentation>
-
-~~~~~~<wsdl:input~name=\char`\"{}UpdateCredentialsRequest\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:UpdateCredentialsRequest\char`\"{}/>
-
-~~~~~~<wsdl:output~name=\char`\"{}UpdateCredentialsResponse\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}UnsupportedFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ProcessingFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}WrongReferenceFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:portType>
-
-~~<wsdl:binding~name=\char`\"{}DelegationBinding\char`\"{}~type=\char`\"{}deleg:DelegationPortType\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}DelegateCredentialsInit\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}DelegateCredentialsInit\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
-
-~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
-
-~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~</wsdl:operation>
-
-~~~~<wsdl:operation~name=\char`\"{}UpdateCredentials\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}UpdateCredentials\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-</wsdl:definitions>
-\end{lyxcode}
-
-\section*{Appendix F. ARC extensions for JSDL schema}
-
-\begin{lyxcode}
-<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
-
-<xsd:schema~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-
-~~~~~~~~~~~~xmlns=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
-
-~~~~~~~~~~~~xmlns:jsdl-arc=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
-
-~~~~~~~~~~~~targetNamespace=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}>
-
-~<xsd:simpleType~name=\char`\"{}GMState\_Type\char`\"{}>
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~<xsd:enumeration~value=\char`\"{}ACCEPTED\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}PREPARING\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}SUBMIT\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}INLRMS\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}FINISHING\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}FINISHED\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}DELETED\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}CANCELING\char`\"{}/>
-
-~~</xsd:restriction>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}Version\_Type\char`\"{}>
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}UpperExclusive\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}LowerExclusive\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Exact\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Exclusive\char`\"{}~type=\char`\"{}xsd:boolean\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}SessionType\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}SessionType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~<xsd:enumeration~value=\char`\"{}INTERNAL\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}LIMITED\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}READONLY\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}FULL\char`\"{}/>
-
-~~</xsd:restriction>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}IsExecutable\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:DataStaging\_Type~(default:~false)~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}IsExecutable\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:boolean\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}FileParameters\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:DataStaging\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}IsExecutable\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}JoinOutputs\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~(default:~false)~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}JoinOutputs\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:boolean\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}Reruns\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~(default:~false)~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}Reruns\char`\"{}~minOccurs=\char`\"{}0\char`\"{}//~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:integer\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}RunTimeEnvironment\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}RunTimeEnvironment\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}Name\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Version\char`\"{}~type=\char`\"{}Version\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:complexType~name=\char`\"{}Middleware\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}Middleware\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}Name\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Version\char`\"{}~type=\char`\"{}Version\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:complexType~name=\char`\"{}RemoteLogging\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}RemoteLogging\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}3\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}URL\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:anyURI\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:complexType~name=\char`\"{}CredentialServer\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}CredentialServer\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}URL\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:anyURI\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:complexType~name=\char`\"{}LocalLogging\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}LocalLogging\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}Directory\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}AccessControlType\_Type\char`\"{}>
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~<xsd:enumeration~value=\char`\"{}GACL\char`\"{}/>
-
-~~</xsd:restriction>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}AccessControl\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}AccessControl\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}OwnerAlwaysAllowed\char`\"{}~type=\char`\"{}xsd:boolean\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Type\char`\"{}~type=\char`\"{}AccessControlType\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Content\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}NotificationType\_Type\char`\"{}>
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~<xsd:enumeration~value=\char`\"{}Email\char`\"{}/>
-
-~~</xsd:restriction>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}Notify\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}Notify\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}3\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}Type\char`\"{}~type=\char`\"{}NotificationType\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Endpoint\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}State\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}~type=\char`\"{}GMState\_Type\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}SessionLifeTime\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}SessionLifeTime\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:long\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}GridTimeLimit\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}GridTimeLimit\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:positiveInteger\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}CandidateTarget\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}jsdl-arc:CandidateTarget\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}HostName\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}QueueName\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}Time\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}ProcessingStartTime\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:dateTime\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<!-{}-=========================================================-{}->
-
-~<xsd:element~name=\char`\"{}IsExecutable\char`\"{}~type=\char`\"{}IsExecutable\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}FileParameters\char`\"{}~type=\char`\"{}FileParameters\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}RunTimeEnvironment\char`\"{}~type=\char`\"{}RunTimeEnvironment\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}Middleware\char`\"{}~type=\char`\"{}Middleware\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}RemoteLogging\char`\"{}~type=\char`\"{}RemoteLogging\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}LocalLogging\char`\"{}~type=\char`\"{}LocalLogging\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}AccessControl\char`\"{}~type=\char`\"{}AccessControl\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}Notify\char`\"{}~type=\char`\"{}Notify\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}SessionLifeTime\char`\"{}~type=\char`\"{}SessionLifeTime\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}SessionType\char`\"{}~type=\char`\"{}SessionType\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}JoinOutputs\char`\"{}~type=\char`\"{}JoinOutputs\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}Reruns\char`\"{}~type=\char`\"{}Reruns\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}CredentialServer\char`\"{}~type=\char`\"{}CredentialServer\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}GridTimeLimit\char`\"{}~type=\char`\"{}GridTimeLimit\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}CandidateTarget\char`\"{}~type=\char`\"{}CandidateTarget\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}ProcessingStartTime\char`\"{}~type=\char`\"{}Time\_Type\char`\"{}/>
-
-</xsd:schema>
 

### 2008-08-17 doc/ws-quick-guide/ws-quick-guide.tex
Easy read - but it does not give instructions on what to do after a "make install" of the ARC-1 sources (or after the installation of any binary package). There was some recent update on the grid monitor, right? This would be nice to be mentioned, so people would know better how to trace their jobs.

@@
 enabling production quality computational and data grids.
-
 Since the first release (May 2002) the middleware has been deployed and
@@
 number of grid projects, like NDGF~\cite{ndgf}, Swegrid~\cite{swegrid} or Swiss
-National Grid Association~\cite{swing} has chosen ARC as their middleware.
+National Grid Association~\cite{swing} have chosen ARC as their middleware.
 
@@
 based on a service container -- the \textit{Hosting Environment Daemon (HED)},
-and different grid capabilities are implemented as Web Services residing in HED.
+and various capabilities of the grid are implemented as distinct Web Services
+that reside in HED.
 
-Currently included components are:
+At the time of writing, this comprises the following components:
 \begin{itemize}
@@
 
-The core part of middleware is written in C/C++. Building the software
-from source or installing a precompiled binary requires different external
-packages, furthermore the client and server packages have different dependencies
-too. Below a list of the explicit requirements is shown:
+The core part of middleware is written in C/C++. To build the software
+from source or installing a precompiled binary requires the prior installation
+of multiple external packages. Client and server packages differ in their
+dependencies to some degree. The following lists the explicit requirements:
 
@@
 \end{itemize}
+
+The developers of ARC have prepared packages for several Linux
+distributions\footnote{The nightly build system current comprises Debian,
+Ubuntu, Fedora, RedHat and OpenSuSE}. As a user of any such operating
+system, the external packages are all readily available.
 
@@
 Please note that you can run similar \texttt{apinfo} request against any ARC1
-service except Echo service. Output should always be XML description of service.
+service except Echo service. Output should always be XML description of the service.
 
@@
 Many information about functionality and configuration of various components
-may be found inside corresponding configuratrion XML schemas.
+may be found inside the corresponding configuration XML schemas.
 
@@
 The open source development of the ARC middleware is coordinated by the
-NorduGrid Collaboration. Currently, the main contributor is the KnowARC
+NorduGrid collaboration. Currently, the main contributor is the KnowARC
 project~\cite{knowarc}, but the Collaboration is open to new members.
@@
 further questions, write to the \texttt{nordugrid-discuss} mailing list (see
-NorduGrid Web site~\cite{nordugrid} for details). Ongoing and completed Grid
-Research projects and student assignments related to the middleware are listed
-on the NorduGrid Web site as well.
+NorduGrid website~\cite{nordugrid} for details). Ongoing and completed Grid
+research projects and student assignments related to the middleware are listed
+on the NorduGrid website as well.
 
@@
 tutorial materials can be fetched from the documentation section of the
-NorduGrid Web site.
+NorduGrid website.
 
-Contact information is kept updated on the NorduGrid Web site.
+Contact information is kept updated on the NorduGrid website.
 

### 2008-08-18 doc/a-rex/arex_tech_doc.tex
A-REX tech doc update. Language fixes from Steffen and spellchecker. Citations if available.

@@
 
-A-REX is one of ARC middleware componets implementing functions of
-so called \emph{Computing Element} (CE).
-
-The aim of the A-REX is to take care of job pre- and post-processing.
-It provides functionality to stage-in files containing input data
-and program modules from wide range of sources and transfer or store
-output results.
-
-The A-REX implements Web Service (WS) interface which provides a way
-to submit and control computational tasks (jobs) to be executed by
-the A-REX and undelying Local Resource Management System.
+A-REX is one of ARC middleware components that implements functions
+of so called \emph{Computing Element} (CE).
+
+It takes care of job pre- and post-processing. This comprises the
+stage-in of files containing input data or program modules from wide
+range of sources and the transfer or storing of output results.
+
+The A-REX implements a Web Service (WS) interface which provides a
+way to submit and control computational tasks (jobs) to be executed
+by the A-REX and he underlying Local Resource Management System.
 
@@
 and understanding of internals of the aforementioned tools. For general
-instalaltion and configuration of whole system refer to other documents
-avaialble at \url{http://www.nordugrid.org/papers.html}. }
+installation and configuration of whole system refer to other documents
+available at \url{http://www.nordugrid.org/papers.html}. }
 
@@
 
-A job is a set of input files (which may include executables), a main
-executable and a set of output files. The process of gathering input
-files, executing a job, and transferring/storing output files is called
-a \emph{session}.
+On the computing element a job is described as a set of input files
+(which may include executables), a main executable and a set of output
+files. The process of gathering input files, executing a job, and
+transferring/storing output files is called a \emph{session}.
 
@@
 availability of any other places accessible by the job other than
-SD (unless such a place is part of a requested Runtime Environement).
+SD (unless such a place is part of a requested Runtime Environment).
 The SD is also the only place which is controlled by the A-REX. It
-is accessible by the user from outside through HTTP(S) protocol. Any
-file created outside the SD is not controlled by the A-REX. Any exchange
-of data between client and A-REX (including also program modules)
-is done through HTTP(S) protocol. A URL for accessing input/output
+is accessible by the user from outside through the HTTP(S) protocol.
+Any file created outside the SD is not controlled by the A-REX. Any
+exchange of data between client and A-REX (including also program
+modules) is performed via HTTP(S). A URL for accessing input/output
 files is  obtained through the WS Local Information Description Interface
@@
 Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
-\cite{jsdl} XML document - which identifies the job in the A-REX
-and the Information Interface.
-
-Jobs are initiated and controlled through WS interface. Complete job
-descriptions (JD)are passed to the A-REX through WS in JSDL\cite{jsdl}
+\cite{ws-addr-soap} XML document - which identifies the job in the
+A-REX and the Information Interface.
+
+Jobs are initiated and controlled through the WS interface. Complete
+job descriptions (JD)are passed to the A-REX through WS in JSDL\cite{jsdl}
 coded description. Input data files and job executables are transferred
@@
 in the JD (with name and source). The A-REX alone is responsible to
-make sure that all required files will be available in the SD.\\
+ensure that all required files will be available in the SD.\\
 The supported protocols for sources at the moment are (in case of
@@
 some less standard sources are supported. Those are described below.
-The A-REX fully relies on HED framework \cite{hed} for data transfering
+The A-REX fully relies on HED framework \cite{hed} for data transferring
 capabilities. Hence actual set of supported protocols depends on installed
@@
 of these files. They have to be specified in the JD. If the job fails
-during any stage of processing, no attept is made to transfer those
+during any stage of processing, no attempt is made to transfer those
 files to their final destination, unless the option \emph{preserve=yes}
@@
 finished. If job execution fails for any reason (if exit code of main
-executable is not 0) all files from furst group are transfered to
+executable is not 0) all files from first group are transferred to
 second one.
@@
 
-Configuration can put limits on the amout of simultaneous jobs in
+Configuration can put limits on the amount of simultaneous jobs in
 some states. If such a limit is reached, a job ready to enter into
@@
 a free slot. This situation is presented by additional state mark
-\textbf{PENDING} to the current state name in the job's status desription.
+\textbf{PENDING} to the current state name in the job's status description.
 
@@
 \item \textbf{Submiting} - The job is being passed for execution to the
-\emph{Local Resource Management System} (LRMS). The correspoding backends
-for many LRMSs are provided with the default installation. If the
-local job submission is successful the job moves to the \textbf{Executing}
+\emph{Local Resource Management System} (LRMS). The corresponding
+backends for many LRMSs are provided with the default installation.
+If the local job submission is successful the job moves to the \textbf{Executing}
 state. Otherwise it moves to \textbf{Finishing}. It is possible to
-limit the agregate number of jobs in \textbf{Submiting} and \textbf{Executing}
+limit the aggregate number of jobs in \textbf{Submiting} and \textbf{Executing}
 states.
@@
 is moved to the state \textbf{Deleted}. The 'deletion' time can be
-obtained by querying the Informatrion Interface of the A-REX. If a
+obtained by querying the Information Interface of the A-REX. If a
 job was moved to \textbf{Finished} because of failure, it may be restarted
@@
 text file containing a single word representing the internal name
-of current state of the job. Possible values and corrsponding external
+of current state of the job. Possible values and corresponding external
 job states are :
@@
 \item \textit{clientname} - name (as provided by the user interface) and
-ip address:port of the submitting client machine
+IP address:port of the submitting client machine
 \item \textit{clientsoftware} - version of software used to submit the job
@@
 
-\textit{url} - ordinary URL for gsiftp, ftp, http, https protocols
-with the addition of '\textbf{replica location service url}' (RLS
-URL).\\
+\emph{URL} represents location from which file can be downloaded.
 Each URL can contain additional options.
@@
 This file is used by the '\textbf{\textit{downloader}}' utility. Files
-with 'url' will be downloaded to the SD and files with 'file description'
-will simply be checked to exist. Each time a new \textbf{valid} file
-appears in the SD it is removed from the list and \textit{job.ID.input}
+with \emph{URL} will be downloaded to the SD and files with 'file
+description' will simply be checked to exist. Each time a new \textbf{valid}
+file appears in the SD it is removed from the list and \textit{job.ID.input}
 is updated. Any external tool can thus track the process of collecting
@@
 This file is used by the '\textbf{\textit{uploader}}' utility. Files
-with \textit{url} will be uploaded to SE and remaining files will
+with \textit{URL} will be uploaded to SE and remaining files will
 be left in the SD. Each time a file is uploaded it is removed from
@@
 during the execution of a job. The WS Interface is built and deployed
-inside Hosting Envirnoment Daemon (HED) infrastructure\cite{hed}.
+inside the Hosting Environment Daemon (HED) infrastructure\cite{hed}.
 
@@
 
-The job submission and control interface is based on document produced
-by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
-The exchange of SOAP messages is performed over the HTTP(S) protocol.
-The BES interface is represented by two port-types - BES-Management
-and BES-Factory. The former is made to control the A-REX service itself
-and thus defines operations to start and stop the functionality of
-the BES service. The A-REX does not implement remote control of service
-functionality. Hence the BES-Management port-type is not functional.
-The BES-Factory port-type provides operations to submit new jobs (to
-create an activity in terms of BES) and to monitor its state. It also
-has an ability to provide information about the service. A-REX fully
-implements the functionality of this port-type.
+The job submission and control interface is based on a document produced
+by the OGF OGSA Basic Execution Services (BES) Working Group\cite{ogsa-bes}.
+
+The exchange of SOAP messages is performed via HTTP(S). The BES interface
+is represented by two port-types - BES-Management and BES-Factory.
+The former is made to control the A-REX service itself and thus defines
+operations to start and stop the functionality of the BES service.
+The A-REX does not implement remote control of service functionality.
+Hence the BES-Management port-type is not functional. The BES-Factory
+port-type provides operations to submit new jobs (to create an activity
+in terms of BES) and to monitor its state. It also has an ability
+to provide information about the service. A-REX fully implements the
+functionality of this port-type.
 
@@
 
-Since BES allows implementators to extend their initial activity states
-with additional sub-states, A-REX defines a set of sub-states of activity
-processing in addition to those defined by the BES, as listed in Table
-\ref{tab:Job-states-definitions}. Their meaning is described in section
-\ref{Section: Job Flow}.
+Since BES allows implementations to extend their initial activity
+states with additional sub-states, A-REX defines a set of sub-states
+of activity processing in addition to those defined by the BES, as
+listed in Table \ref{tab:Job-states-definitions}. Their meaning is
+described in section \ref{Section: Job Flow}.
 
@@
 
-The A-REX also supportd the Delegation Interface (see Appendix D).
+The A-REX also supports the Delegation Interface (see Appendix D).
 This is a common purpose interface to be used by ARC services which
 accepts delegated credentials from clients. The Delegation Interface
-implements two operations: initiatialization of credentials delegation
+implements two operations: initialization of credentials delegation
 (DelegateCredentialsInit) and update/renewal of credentials (UpdateCredentials).
@@
 \begin{lyxlist}{00.00.0000}
-\item [{\emph{nordugrid}}] - desription of computing resource that uses
-NorudGrid LDAP schema {[}] converted to XML document.
+\item [{\emph{nordugrid}}] - description of computing resource that uses
+NorudGrid LDAP schema \cite{is} converted to XML document.
 \item [{\emph{Domains}}] \emph{-} description of a computation resource
@@
 including POSIX Applications extension and JSDL HPC Profile Application
-Extension {[}]:
+Extension \cite{jsdl-hpc}:
 
@@
 \item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
-\item [{\emph{IndividualCPUCount}}] - same as TotalCPUCount.
-\item [{\emph{Environment}\ (POSIX)}] - environement variable.
+\item [{\emph{IndividualCPUCount}}] - same as 6.
+\item [{\emph{Environment}\ (POSIX)}] - environment variable.
 \end{lyxlist}
@@
 Files \textit{list, old , new} and \textit{\#\#\#\#\#\#\#\#.info}
-has to be stored on filesystem which has support for files' locking.
+has to be stored on file system which has support for files' locking.
 
@@
 
-If a job requests an input file which is subject for cacheing, it
-is stored in the cache directory instead of the SD. The file is made
+If a job requests an input file which is subject for caching, it is
+stored in the cache directory instead of the SD. The file is made
 available to the job by either soft-linking it in the SD or copying
@@
 Before giving access to an already cached file, the A-REX contacts
-the original file source to check if the user has suficient access
+the original file source to check if the user has sufficient access
 rights to the file. Not all protocols make this check possible. Also,
@@
 \item \textit{uploader} - This module is responsible for delivering output
-files to the specified SEs and registration at an Indexding Service
+files to the specified SEs and registration at an Indexing Service
 (like RLS) as needed. It processes and updates the \textit{job.ID.output}
@@
 \item \emph{cluster+qju.pl} - Collects and generates information about computing
-resource as XML document in NoruGrid and Glue 2 format.
+resource as XML document in NorduGrid and Glue 2 format.
 \end{itemize}
@@
 \item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
-Curently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
+Currently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
 pseudo-LRMS is supported for testing purposes. This module is responsible
@@
 \begin{itemize}
-\item \textit{session root directory} - Tthis is the directory in which
-a user's SDs are created. It's location is configurable per UNIX user.
+\item \textit{session root directory} - This is the directory in which a
+user's SDs are created. It's location is configurable per UNIX user.
 Several (or even all) users may share the same session root directory.\\
@@
 If A-REX is run under the \emph{root} account, make sure \textit{session
-root directory} resides on a filesystem which does not limit the capabilities
-of the \emph{root} user (as does for example NFS with \emph{root\_squash}
-option).\\
+root directory} resides on a file system which does not limit the
+capabilities of the \emph{root} user (as does for example NFS with
+\emph{root\_squash} option).\\
 If there is a need to run A-REX under the \emph{root} account (to
@@
 The A-REX creates SDs with proper ownership and permissions for the
-local identity used to run a job. Some filesystems require users to
-have \emph{execute} permission on the \textit{session root directory}
+local identity used to run a job. Some file systems require users
+to have \emph{execute} permission on the \textit{session root directory}
 in order to access any file or subdirectory there.\\
@@
 
-Due to historical resons configuration of the A-REX is split into
+Due to historical reasons configuration of the A-REX is split into
 2 parts. The small part is located inside HED configuration (see appendix
@@
 composes script submitted to LRMS in way to make it do that. This
-requires installation of A-REX and all realted software to be accessible
+requires installation of A-REX and all related software to be accessible
 from computing nodes and environment variable ARC\_LOCATION to be
@@
 then jobs can't finish due to big amount of recently submitted jobs.\textit{}\\
-\textit{max\_transfered\_files} - maximal number of files being transfered
+\textit{max\_transfered\_files} - maximal number of files being transferred
 in parallel by every job. Used to decrease load on not so powerful
@@
 The default is \emph{no}. Only files downloaded through \emph{meta-url}
-are registred. Registration is done to same service used for obtaining
-information about file. For this opeartion credentials of the A-REX
+are registered. Registration is done to same service used for obtaining
+information about file. For this operation credentials of the A-REX
 (host key and certificate) are used. If required new files storage
@@
 \textit{pass} - continue execution,\\
-\textit{log} - write information about result into logfile and continue
+\textit{log} - write information about result into log file and continue
 execution,\\
-\textit{fail} - write information about result into logfile and cancel
+\textit{fail} - write information about result into log file and cancel
 job.
@@
 all processes involved in job management will use local identity of
-a user to which Grid identity is mapped in order to access filesystem
+a user to which Grid identity is mapped in order to access file system
 at path specified in \textbf{\textit{session}} command (see below).
@@
 \item \textbf{\textit{speedcontrol}}\textit{=min\_speed min\_time min\_average\_speed
-max\_inactivity} - specifies how long/slow data fransfer is allowed
+max\_inactivity} - specifies how long/slow data transfer is allowed
 to take place. Transfer is canceled if transfer rate (bytes per second)
@@
 or if average rate is lower than \emph{min\_average\_speed}, or no
-data is receved for longer than \textit{max\_inactivity} seconds.
+data is received for longer than \textit{max\_inactivity} seconds.
 \item \textbf{\textit{copyurl}}\textit{=template replacement} - specifies
@@
 are 7days for \emph{ttl} and 30 days for \emph{ttr}.
-\item \textbf{\textit{defaultlrms}}\textit{=default\_lrms\_name default\_queue\_name}
+\item \textbf{\textit{lrms}}\textit{=default\_lrms\_name default\_queue\_name}
 - specifies names for the LRMS and queue. Queue name can also be specified
-in the JD (currently it is not allowed to override used LRMS by using
-JD). In new configuration file this command is called \textbf{\textit{lrms}}.
+in the JD (currently it is not allowed to override LRMS by usingon
+the JD).
 \item \textbf{\textit{session}}\textit{=path} - specifies path to the directory
@@
 values. In that case corresponding positive value means space left
-on filesystem. If \textit{low\_mark} is omitted it becomes equal to
-\textit{high\_mark}. By default this feature is turned off. To turn
-it off explicitly \textit{cachesize} without parameters should be
-specified. If turned off cache will grow up till it fills whole file
-system.
+on file system. If \textit{low\_mark} is omitted it becomes equal
+to \textit{high\_mark}. By default this feature is turned off. To
+turn it off explicitly \textit{cachesize} without parameters should
+be specified. If turned off cache will grow up till it fills whole
+file system.
 \item \textbf{\textit{maxrerun}}\textit{=number} - specifies maximal number
@@
 \item \textbf{\textit{shared\_filesystem}}\textit{=yes|no} - if computing
-nodes have an access to session directory through a shared filesystem
-like NFS. Corresponds to an environement variable RUNTIME\_NODE\_SEES\_FRONTEND
+nodes have an access to session directory through a shared file system
+like NFS. Corresponds to an environment variable RUNTIME\_NODE\_SEES\_FRONTEND
 (see section \ref{sub:LRMS}).
@@
 \item [{\%L}] - default lrms - look command 'defaultlrms'
-\item [{\%W}] - installation path - \$\{NORDUGRID\_LOCATION\}
+\item [{\%W}] - installation path - \$\{ARC\_LOCATION\}
 \item [{\%G}] - globus path - \$\{GLOBUS\_LOCATION\}
@@
 Authorization is performed by generic means provided by HED framework.
-Currently A-REX does not implement any internal authorization tecniques
+Currently A-REX does not implement any internal authorization techniques
 except those imposed by Access Policy assigned to jobs through AccessControl
@@
 \textbf{\textit{RUNTIME\_NODE\_SEES\_FRONTEND}} - if set to {}``\emph{no}''
-means computing node does not share filesystem with frontend. In that
-case content of the SD is moved to computing node by using means provided
-by the LRMS. Results are moved back after job's execution in a same
-way. Corresponds to \emph{shared\_filesystem} configuration command.
+means computing node does not share file system with frontend. In
+that case content of the SD is moved to computing node by using means
+provided by the LRMS. Results are moved back after job's execution
+in a same way. Corresponds to \emph{shared\_filesystem} configuration
+command.
 
@@
 \item joboption\_directory - session directory.
-\item joboption\_args - command to be executed as specified in RSL.
+\item joboption\_args - command to be executed as specified in the JD.
 \item joboption\_env\_\# - array of 'NAME=VALUE' environment variables (\textbf{not}
@@
 \item joboption\_stderr - same for stderr.
-\item joboption\_maxcputime - amout of CPU time requested (minutes).
-\item joboption\_maxmemory - amout of memory requested (megabytes).
+\item joboption\_maxcputime - amount of CPU time requested (minutes).
+\item joboption\_maxmemory - amount of memory requested (megabytes).
 \item joboption\_count - number of processors requested.
@@
 
-The A-REX service is a plugable module of the HED. Si first You need
+The A-REX service is a plugable module of the HED. So first You need
 to setup your HED - see HED documentation.
@@
 element inside <Plugins> containing string \emph{arex}. That will
-make HED load libarex plugin library. Then add new <Service> element
-with attribute name=''a-rex''. That will instantiate A-REX service.
-Now to make service accessible extend <Plexer> element with new <next>
-refering to service's id. Write <Service> element carefully - here
-is an example:
+make HED load libarex plugin library.
+
+Then add new <Service> element with attribute name=''a-rex''. That
+will instantiate A-REX service. Now to make service accessible extend
+<Plexer> element with new <next> referring to an id of the service.
+Write <Service> element carefully - here is an example:
 
@@
 
-Please read \emph{User Manual {[}]} for usage instructions.
+Please read \emph{User Manual {[}??]} for usage instructions.
 
@@
 by job. It can also be used to monitor job execution by checking content
-of application dependant files in SD.
-
-The BES defines job identifier as WS Addressing {[}] Endpoint Reference
-(EPR) - XML document. The EPR is extendable and the A-REX adds it
-own element JobSessionDir belonging to namespace http://www.nordugrid.org/schemas/a-rex
-as direct child of ReferenceParameters element. This new element contains
-the URL of SD.
-
-Obtained URL shoud be extended with file names relative to SD and
+of application dependent files in SD.
+
+The BES defines job identifier as WS Addressing \cite{ws-addr-soap}
+Endpoint Reference (EPR) - XML document. The EPR is extendable and
+the A-REX adds it own element JobSessionDir belonging to namespace
+http://www.nordugrid.org/schemas/a-rex as direct child of ReferenceParameters
+element. This new element contains the URL of SD.
+
+Obtained URL should be extended with file names relative to SD and
 HTTP methods PUT and GET may be used to upload/download content of
 those files. For directories - including SD itself - GET method is
-supported shich returns HTML encoded non-recursive list of files and
-firectories. The files and subdirectories have their URLs inside HTML
+supported which returns HTML encoded non-recursive list of files and
+directories. The files and subdirectories have their URLs inside HTML
 element <A>.

### 2008-08-19 doc/ws-quick-guide/ws-quick-guide.tex
Rename apinfo -> arcinfo

@@
 
-Now you may use the command line utility ``\texttt{apinfo}'' to obtain a service
+Now you may use the command line utility ``\texttt{arcinfo}'' to obtain a service
 description. Make sure user key and certificate and CA certificate(s) are
@@
 \begin{lstlisting}
-./apinfo -d VERBOSE -C $HOME/usercert.pem -K $HOME/userkey.pem
+./arcinfo -d VERBOSE -C $HOME/usercert.pem -K $HOME/userkey.pem
  -A /etc/grid-security/certificates https://localhost:60000/arex 2>error.log
@@
 This should produce XML output describing resource A-REX represents.
-Listing~\ref{lst:apinfo} shows an example of proper output. In case of error see
+Listing~\ref{lst:arcinfo} shows an example of proper output. In case of error see
 \texttt{error.log} file and report problems to ARC development team. Include
@@
 
-\begin{lstlisting}[caption={Example of a good apinfo output},label=lst:apinfo]
+\begin{lstlisting}[caption={Example of a good arcinfo output},label=lst:arcinfo]
 <n:nordugrid xmlns:n="urn:nordugrid" xmlns:M0="urn:Mds"
@@
 
-Please note that you can run similar \texttt{apinfo} request against any ARC1
+Please note that you can run similar \texttt{arcinfo} request against any ARC1
 service except Echo service. Output should always be XML description of the service.

### 2008-08-19 doc/ws-quick-guide/ws-quick-guide.tex
Rename approxy -> arcproxy

@@
 
-ARC comes with utility \texttt{approxy} which generates proxy credentials from
+ARC comes with utility \texttt{arcproxy} which generates proxy credentials from
 certificate/private key pair. It provides only basic functionality and is meant
@@
 then you won't be able to use proxy credentials generated by utilities
-like \texttt{grid-proxy-init}, \texttt{voms-proxy-init} or \texttt{approxy}.
+like \texttt{grid-proxy-init}, \texttt{voms-proxy-init} or \texttt{arcproxy}.
 Because of that all user private keys will have to be kept unencrypted.
@@
 As metioned above, ARC comes with proxy generation utility
--- \texttt{approxy}, installed in \verb|$ARC_LOCATION/bin|. The usage
-of \texttt{approxy} is like:
+-- \texttt{arcproxy}, installed in \verb|$ARC_LOCATION/bin|. The usage
+of \texttt{arcproxy} is like:
 
 \begin{lstlisting}
-$ARC_LOCATION/bin/approxy -P proxy.pem -C cert.pem -K key.pem
+$ARC_LOCATION/bin/arcproxy -P proxy.pem -C cert.pem -K key.pem
   -c validityStart=2008-05-29T10:20:30Z

### 2008-08-19 doc/storage_system/arc1-storage-documentation.tex
Rename arc_storage_cli.py arc_storage_cli

@@
 
-In the first prototype release there is one client tool called \verb!arc_storage_cli.py!, which is written in Python, and only need a basic Python installation to run.
+In the first prototype release there is one client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run.
 It is capable of communicating with a given Bartender service, and uploading and downloading TURLs via HTTP.
@@
 \begin{verbatim}
-$ arc_storage_cli.py
+$ arc_storage_cli
 Usage:
-  arc_storage_cli.py <method> [<arguments>]
+  arc_storage_cli <method> [<arguments>]
 Supported methods: stat, make[Collection], unmake[Collection], list, move,
@@
 \begin{verbatim}
-$ arc_storage_cli.py move
+$ arc_storage_cli move
 Usage: move <sourceLN> <targetLN>
@@
 This is a testfile.
-$ arc_storage_cli.py put testfile /tmp/
+$ arc_storage_cli put testfile /tmp/
 - The size of the file is 20 bytes
@@
 'testfile' (20 bytes) uploaded as '/tmp/testfile'.
-$ arc_storage_cli.py stat /tmp/testfile
+$ arc_storage_cli stat /tmp/testfile
 - ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
@@
     type: file
-$ arc_storage_cli.py get /tmp/testfile newfile
+$ arc_storage_cli get /tmp/testfile newfile
 - ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender

### 2008-08-21 doc/a-rex/arex_tech_doc.tex
Extending BES/A-REX states table to include Deleted and Finished states
Fixes for A-REX document implementing some of Gabor's comments

@@
 \usepackage{array}
+\usepackage{longtable}
 \usepackage{amsmath}
@@
 \usepackage{multirow}
+\usepackage{verbatim}
 %
@@
 exchange of data between client and A-REX (including also program
-modules) is performed via HTTP(S). A URL for accessing input/output
+modules) is performed via HTTP(S). An URL for accessing input/output
 files is  obtained through the WS Local Information Description Interface
@@
 Cancelled&
-Failed&
+Finished&
 There was a failure during execution\tabularnewline
@@
 \multicolumn{1}{l|}{}\tabularnewline
+\hline 
+Finished&
+Finished&
+Job finished successfully\tabularnewline
+\hline 
+Finished&
+Deleted&
+Job finished and was left in A-REX too long\tabularnewline
 \hline 
@@
 \item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
-\item [{\emph{IndividualCPUCount}}] - same as 6.
-\item [{\emph{Environment}\ (POSIX)}] - environment variable.
+\item [{\emph{IndividualCPUCount}}] - same as \emph{TotalCPUCount}.
 \end{lyxlist}
+%Environment(POSIX) - environment variable.
 
@@
 file.
-\item \emph{cache-register} - Utility to register cached data into an Indexing
-Service. It reads and modifies cache informational files \emph{old}
-and \emph{new} (as described in section \ref{sec:cache}). Configuration
-is read directly from the A-REX's configuration file (see section
-\ref{SubSection: ConfigFile}). It is run by the A-REX every 5 minutes.
-\item \emph{frontend-info-collector} - Utility to gather information about
-the frontend. It puts collected information into the \emph{job.ID.diag}
-file.
+\end{itemize}
+%cache-register - Utility to register cached data into an Indexing Service.
+It reads and modifies cache informational files old and new (as described in section ).
+Configuration is read directly from the A-REX's configuration file (see section ).
+It is run by the A-REX every 5 minutes.
+
+%frontend-info-collector - Utility to gather information about the frontend. It puts collected information into the job.ID.diag file.
+
+\begin{itemize}
 \item \emph{gm-kick} - Sends a signal to the A-REX though a FIFO file to
@@
 \end{itemize}
-In addition, there is also an administration utility:
+\begin{comment}In addition, there is also an administration utility:
 
@@
 \end{itemize}
+\end{comment}
 
@@
 place to store cache files containing data itself. This can be useful
-in case of big data storage available only on NSF server which does
+in case of big data storage available only on NFS server which does
 not support file locking. If command or \textit{path} is missing -
@@
 \textbf{\textit{RUNTIME\_FRONTEND\_SEES\_NODE}} - if defined should
-contain path corresponding to \textit{RUNTIME\_LOCAL\_SCRATCH\_DIR}
-as seen on \textbf{frontend} machine. Corresponds to \emph{shared\_scratch}
-configuration command.
+contain path corresponding to \textit{}\\
+\textit{RUNTIME\_LOCAL\_SCRATCH\_DIR} as seen on \textbf{frontend}
+machine. Corresponds to \emph{shared\_scratch} configuration command.
 
@@
 present some possible combinations for RUNTIME\_LOCAL\_SCRATCH\_DIR
-and RUNTIME\_FRONTEND\_SEES\_NODE and explain how data movement is
-performed. Pictures a) correspond to the situation right after all
-input files have been gathered in the session directory and show the
-actions taken right after the job's script starts. Pictures b) show
-the situation while the job is running and the actions which are taken
-right after it has finished. Pictures c) illustrate the final situation,
-when the job's output files are ready to be uploaded to an external
-storage element or be downloaded by the user.
+and \\
+RUNTIME\_FRONTEND\_SEES\_NODE and explain how data movement is performed.
+Pictures a) correspond to the situation right after all input files
+have been gathered in the session directory and show the actions taken
+right after the job's script starts. Pictures b) show the situation
+while the job is running and the actions which are taken right after
+it has finished. Pictures c) illustrate the final situation, when
+the job's output files are ready to be uploaded to an external storage
+element or be downloaded by the user.
 
@@
 \item joboption\_directory - session directory.
-\item joboption\_args - command to be executed as specified in the JD.
+\item joboption\_arg\_\# - command with arguments to be executed as specified
+in the JD (\textbf{not} bash array).
 \item joboption\_env\_\# - array of 'NAME=VALUE' environment variables (\textbf{not}
@@
 \end{itemize}
-For example \emph{joboption\_args} could be changed to wrap main executable.
-Or \emph{joboption\_runtime} could be expanded if current one depends
-on others.
+For example \emph{joboption\_arg\_\#} could be changed to wrap main
+executable. Or \emph{joboption\_runtime} could be expanded if current
+one depends on others.
 
@@
 
-The A-REX service is a plugable module of the HED. So first You need
+The A-REX service is a pluggable module of the HED. So first You need
 to setup your HED - see HED documentation.
@@
 
-\section*{Appendix A. Session directory access through HTTP(S) interface}
+\appendix
+
+\section{Session directory access through HTTP(S) interface}
 
@@
 
-\section*{Appendix B. Configuration schema of A-REX}
+\section{Configuration schema of A-REX}
 
@@
 
-\section*{Appendix C. A-REX WSDL}
+\section{A-REX WSDL}
 
@@
 
-\section*{Appendix D. Delegation WSDL}
+\section{Delegation WSDL}
 
@@
 
-\section*{Appendix E. ARC extensions for JSDL schema}
+\section{ARC extensions for JSDL schema}
 
@@
 
-\section*{Appendix F. Error messages of A-REX}
+\section{Error messages of A-REX}
 
@@
 into \textit{job.ID.failed}. Possible values include those generated
-by the A-REX itself:
-
-\begin{tabular}{|p{2.5in}|p{4in}|}
+by the A-REX itself:\\
+\begin{longtable}{|p{2.5in}|p{4in}|}
 \hline 
@@
 Job submission to LRMS failed&
-Backend executable supposed to pass job to LRMs returned non-zero
+Backend executable supposed to pass job to LRMS returned non-zero
 exit code\tabularnewline
@@
 \hline
-\end{tabular}
+\end{longtable}
 
@@
 Provided by downloader component (URL is replcaced by source of input
-file, FILE by name of file):
-
-\begin{tabular}{|p{2.5in}|p{4in}|}
+file, FILE by name of file):\\
+\begin{longtable}{|p{2.5in}|p{4in}|}
 \hline 
@@
 \hline
-\end{tabular}
+\end{longtable}
 
 \medskip{}
-Provided by uploader component (URL is replcaced by destination of
-output file) :
-
-\begin{tabular}{|p{2.5in}|p{4in}|}
+Provided by uploader component (URL is replaced by destination of
+output file) :\\
+\begin{longtable}{|p{2.5in}|p{4in}|}
 \hline 
@@
 \hline
-\end{tabular}
+\end{longtable}
 
 \medskip{}
-Coming from LRMS (PBS) backend:
-
-\begin{tabular}{|p{2.5in}|p{4in}|}
+Coming from LRMS (PBS) backend:\\
+\begin{longtable}{|p{2.5in}|p{4in}|}
 \hline 

### 2008-08-22 doc/a-rex/arex_tech_doc.tex
Fixing a-rex document to represent bes/internal/reported staates consistently.

@@
 \begin{itemize}
-\item ACCEPTED/\textbf{Accepted}
-\item PREPARING/\textbf{Preparing}
-\item SUBMITTING/\textbf{Submiting}
-\item INLRMS/\textbf{Executing}
-\item FINISHING/\textbf{Finishing}
-\item FINISHED/\textbf{Finished}
-\item CANCELING/\textbf{Killing}
-\item DELETED/\textbf{Deleted}
+\item ACCEPTED
+\item PREPARING
+\item SUBMIT
+\item INLRMS
+\item FINISHING
+\item FINISHED
+\item CANCELING
+\item DELETED
 \end{itemize}
@@
 
-\begin{tabular}{|c|c|>{\raggedright}p{10cm}|}
+\begin{tabular}{|c|c|c|>{\raggedright}p{10cm}|}
 \hline 
@@
 ARC Sub-state&
+A-REX internal state&
 \multicolumn{1}{c|}{Description}\tabularnewline
@@
 Accepting&
-Job is in the process of being submitted\tabularnewline
-\cline{2-2} \cline{3-3} 
+ACCEPTED&
+Job is in the process of being submitted. This state is not recognised
+by the A-REX yet. \emph{Accepted} is first reported state.\tabularnewline
+\cline{2-2} \cline{3-3} \cline{4-4} 
 \multicolumn{1}{|c|}{}&
 Accepted&
+ACCEPTED&
 Job was submitted\tabularnewline
@@
 Preparing&
+PREPARING&
 Stage-in process is going on\tabularnewline
-\cline{2-2} \cline{3-3} 
+\cline{2-2} \cline{3-3} \cline{4-4} 
 \multicolumn{1}{|c|}{}&
 Prepared&
+PREPARING+PENDING&
 Stage-in process has finished\tabularnewline
-\cline{2-2} \cline{3-3} 
+\cline{2-2} \cline{3-3} \cline{4-4} 
 \multicolumn{1}{|c|}{}&
 Submiting&
+SUBMIT&
 Communication with local batch system is in process\tabularnewline
-\cline{2-2} \cline{3-3} 
+\cline{2-2} \cline{3-3} \cline{4-4} 
+&
+\multicolumn{1}{c|}{Queued}&
+\multicolumn{1}{c|}{INLRMS}&
+\multicolumn{1}{p{10cm}|}{Job entered local batch system but is not runnning now. This state
+is not recognised by the A-REX yet. \emph{Executing} is reported instead }\tabularnewline
+\cline{2-2} \cline{3-3} \cline{4-4} 
 \multicolumn{1}{|c|}{}&
 Executing&
+INLRMS&
 Job is being executed in local batch system\tabularnewline
-\cline{2-2} \cline{3-3} 
+\cline{2-2} \cline{3-3} \cline{4-4} 
+\multicolumn{1}{|c|}{}&
+Executed&
+INLRMS,INLRMS+PENDING&
+Job execution in local batch system has finished. The A-REX dos not
+detect job states inside local batch system yet. As result this state
+is reported only if job is \emph{Pending}.\tabularnewline
+\cline{2-2} \cline{3-3} \cline{4-4} 
 \multicolumn{1}{|c|}{}&
 Killing&
+CANCELING&
 Communication with local batch system to terminate execution is in
 process\tabularnewline
-\cline{2-2} \cline{3-3} 
-\multicolumn{1}{|c|}{}&
-Executed&
-Job execution in local batch system has finished\tabularnewline
-\cline{2-2} \cline{3-3} 
+\cline{2-2} \cline{3-3} \cline{4-4} 
 \multicolumn{1}{|c|}{}&
 Finishing&
+FINISHING&
 Stage-out process is going on\tabularnewline
@@
 Cancelled&
-Finished&
-There was a failure during execution\tabularnewline
-\cline{1-1} 
-Failed&
-\multicolumn{1}{c|}{}&
-\multicolumn{1}{l|}{}\tabularnewline
+Killed&
+FINISHED&
+Job was stopped by explicit user request. The A-REX currently does
+not remember this request. \emph{Failed} is reported instead.\tabularnewline
+\hline 
+\multicolumn{1}{|c|}{Failed}&
+\multicolumn{1}{c|}{Failed}&
+FINISHED&
+\multicolumn{1}{l|}{There was a failure during execution}\tabularnewline
 \hline 
@@
 Finished&
+FINISHED&
 Job finished successfully\tabularnewline
@@
 Deleted&
+DELETED&
 Job finished and was left in A-REX too long\tabularnewline
@@
 Pending&
+PENDING&
 Job is prevented from going to the next state due to some internal
@@
 Held&
+&
 Job processing is suspended on client request; this sub-state appears
 in parallel with other sub-states. This state is reserved for future
-and i snot implemented yet.\tabularnewline
+and is not implemented yet.\tabularnewline
 \hline

### 2008-08-22 doc/hed/ARCHED_article.tex
Few more words to HED document

@@
 In the technical design it turned out that the endeavours to provide dynamic loading, portability and a well tested high level memory management could all be greatly assisted by introducing glibmm - the C++ interface to the gnome projects library for memory management and related functionality.
-This enables the daemon process to dynamically load the components it is going to use in a portable way.
+This enables the developers to write code in a way easily portable across various operating systems and architectures.
 
-Most of the configuration
+%Most of the configuration
 
-The HED itself means two things:
+The HED itself means three things:
 \begin{enumerate}
+
 \item the daemon (called \texttt{arched}) which hooks up the system and initialize components the way as it is described by the configuration files.
 This configuration describes the components and their relations to each other.
-In optimal cases these single services runs on any node where ARC1 is deployed and started.
-Without loadable components the daemon itself does nothing at all.
+In optimal cases these single services run on any node where ARC1 is deployed and started.
+Without loadable components the daemon itself does nothing usefull.
+
 \item sometimes using the HED terms to refer to collection of libraries which is used by service or other component developers.
-These libraries implements some common classes which may simplify the life of service and component developes however only few of these classes are mandatory to use to make the components and services loadable and hookable by the daemon.
+These libraries define interfaces and implement some common classes which may simplify the life of service and component developes however only few of these classes are mandatory to use to make the components and services loadable and hookable by the daemon.
+
+\item the collection of components implementing minimal set of protocols needed for implementing so called Web Services.
+
+Unless otherwise stated the term HED will be used through this document to refer to second option - framework of C++ classes.
+
 \end{enumerate}
@@
 
-In the HED all data channels to the outside world are set up by chains of small processing units called Message Chain Components (MCCs).
-These work on units called messages which represent data going in to or out of the HED.
-The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general attributes and security attributes where information relevant to each protocol is accumulated as the message progresses.
+In the HED data channels to the outside world may be set up by chains of small processing units called Message Chain Components (MCCs).
+These work on units called Messages which represent data going in to or out of the HED.
+The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general Attributes and Security Attributes where information relevant to each protocol is accumulated as the message progresses.
 Each MCC typically implements one level in the Internet Protocol suite by transforming a message to an input suitable to propagate to the next component and then performs the corresponding transformation of the response on the way back.
@@
 
-Each MCC can also be configured to have loadable modules called security handlers attached to it in order to enforce security policies such as authentication and authorization or to assist such activities by gathering specialized security information.
+\paragraph{Payload}
+
+Main content of the information is transfered using the Payload part of the Message.
+There are no limitations on functionality of Payload object except that it must be inherited from MessagePayload class.
+Despite being flexible such approach would be useless.
+This is why HED defines three Payload interfaces and their simplest implementations.
+All MCCs which are distributed with the HED use, implement and extend those interfaces.
+Those include:
+
+\begin{enumerate}
+
+\item PayloadRawInterface and it's implementation PayloadRaw.
+This interface represents set of catenated in-memory chunks.
+It's meant to be used for information available as whole.
+And also for prepending and appending information without actually moving and copying data chunks in memory.
+
+\item PayloadStreamInterface and it's implementation PayloadStream.
+It covers case of sequentially accessible information.
+The main purpose of that Payload is to serve protocols which define continous data stream like TCP.
+
+\item PayloadSOAP represents parsed SOAP message. It's introduced to cover need for writing SOAP based Web Services in unified way.
+
+\end{enumerate}
+
+\paragraph{Security Handler}
+
+Each MCC can also be configured to have loadable modules called Security Handlers attached to it in order to enforce security policies such as authentication and authorization or to assist such activities by gathering specialized security related information into Security Attributes.
+There is no strict distinction of capabilities between Security Handlers and MCCs.
+Both can and do populate Security Attributes.
+The distinction is more of logical nature.
+It also makes possible to have Security Handlers dealing with similar kind of information and capable of acting on different protocol levels.
+
+\paragraph{Configuration}
 

### 2008-08-24 doc/hed/ARCHED_article.tex
A bit about attributes and secattr in HED document

@@
 As a result essentials like SOAP and HTTP are implemented inside HED, while external software is used whenever that is found to be appropriate - as in the case of TLS, (Grid)FTP, LDAP and some other cases.
-?That does not exclude possibility to have plugins using entirely external solutions either developed or accepted from third parties.?
+That does not exclude possibility to have plugins using entirely external solutions either developed or accepted from third parties.?
 
@@
 
-%As the data is passed through the individual MCCs, they each populate structures with both general attributes and special security attributes that are available at that particular protocol level.
+Each MCC has an entry method which is called with Message being processed.
+It then processes Message by modifying it or creating new Message.
+Then MCC calls entry method of next MCC in the chain.
+For information how mwssages are handled and about memory management policies please see API description of MCC class in [].
+
+As the data is passed through the individual MCCs, they each populate structures with both general attributes and special security attributes that are available at that particular protocol level.
 
@@
 
+Each MCC developed inside and outside the HED must be accompanied with description of the Payloads it supports on input and those generated on output.
+Those types should be taken into account while creating chains of the MCCs.
+There are no Payload type checks done during the chain configuration phase.
+Hence Payload incompatibility problems will be detected only during runtime.
+
+\paragraph{Attributes}
+
+The Message object may contain general purpose key and value pairs called Attributes.
+Keys and values are simple stings.
+Each key may have multiple corresponding values.
+All pairs are handled by MessageAttributes class.
+Codewise there are no limitations put on content and purpose of Attributes.
+
+By convention keys are composed of two parts: name of MCC/Protocol at which Attribute was generated or must be consumed and name of Attribute itself separated by column.
+For example the Attribute with key HTTP:METHOD holds HTTP protocol method like GET, PUT, HEAD, etc.
+It is either generated MCC implementing HTTP protocol or is filled by other code and is used by HTTP MCC to generate proper HTTP header.
+
+Each MCC developed inside and outside the HED must be have generated and consumed Attributes described in accompanying documentation.
+
+\paragraph{Security Attributes}
+
+Here only basic information about security related objects is presented. For more detailed information please see "Security Framework of ARC1" [].
+
+Security Attributes are storing various aspects of Message useful for authorization decisions to be made.
+Those normally include operation being requested, target of operation and identity of subject making request.
+They can also contain authorization policies.
+Actually nothing stops from storing an other type of information but there is no convention developed for that.
+
+The Secuirity Attributes are stored as key and value pairs.
+Each key may have only single value attached.
+Keys are simple strings.
+By convention each MCC or Security Handler (see below) produce Security Attribute with name corresponding to protocol name.
+For example Security Attribute stored under name TLS holds information collected at Transport Level Security layer.
+
+The value if Secuirity Attribute is an object of class inherited from SecAttr.
+The HED implements class SecAttr which serves as definition of interface for all Security Attributes.
+It defines way collected information may be turned into useful format.
+For that each Security Attribute value implements method Export for converting internal information into one of supported formats.
+Currently only implemented is ARC Authorization Request/Policy (see below).
+Please see API description of SecAttr and MessageAuth classes in [] for more information.
+
+Each MCC developed inside and outside the HED must come with explanation of generated and consumed Security Attributes.
+Dedicated components for dealing with Security Attribute - Security Handlers are described below.
+
 \paragraph{Security Handler}
@@
 It also makes possible to have Security Handlers dealing with similar kind of information and capable of acting on different protocol levels.
+
+\paragraph{Client and Server MCCs}
 

### 2008-08-28 doc/client/client_technical.tex
started working on technical documentation

\documentclass{book}
%\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-XXXXXXX-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large Paper title}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Paper subtitle}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Paper author\footnote{authors@address} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage
\chapter{Preface}
\label{sec:intro}

This document describes from a technical viewpoint a plugin based client library for the new 
Web Service (WS) based Advanced Resource Connector (ARC) middleware. The library is capable of 
handling resource discovery, job submission and job control and interoperates with several 
different grid flavours. The support for different grid flavours is enabled through the modular 
design using plugins specialized for each supported middleware, and presently supported are ARC0 
(ARC classic), ARC1 and the gLite middlewares. Future extensions to support additional middlewares 
involve plugin compilation only i.e. no recompilation of main libraries or client.

Using the library a set of command line tools have been built which puts the library's functionality 
at the fingertips of users. While this documentation will illustrate how such command line tool can be 
built, the main documention of the command line tools is given in the client user manual [ref].

Finally the library has been designed with care for later usage within a graphical user interface. This 
particularly applies to the job management part.

\chapter{Resource Discovery and Information Retrieval}
\label{sec:TargetDiscovery}
Almost every day a new grid cluster is purchased and made available around the world (either to the 
general public or restricted to authorized users). More available clusters create a need for better and 
faster resource discovery and information retrieval, making this task an utmost important component of the 
new client library.

The new client library's resource discovery and information retrieval component consists of three classes, one 
of which a base class for further grid middleware specific specialization (plugin):

\begin{itemize}

\item{{\bf TargetGenerator} is responsible for loading TargetRetriever plugins in accordance with the input URLs. 
e.g. if an URL pointing to an ARC1 resource is given the TargetRetrieverARC1 is loaded. If no URLs are given, the 
TargetGenerator makes use of the user/system defaults. When queried for resources (``targets'') the TargetGenerator 
passes the request to the loaded TargetRetrievers running them as individual threads for improved performance. The 
TargetGenerator keeps records of by TargetRetriever found services (index of computing) in order to avoid multiple 
queries to the same resource. Accepted targets are registered in the FoundTargets array of the TargetGenerator.}

\item{{\bf TargetRetriever} is base class for TargetRetriever grid middleware specializations. Contains the pure 
virtual method GetTargets which is to be implemented by the specialized class. While it is not mandatory it is recommended 
that the specialized class divides this function into two components: QueryIndex and InterrogateTarget. The former is 
responsible for querying index servers, and if the result yields an addresss to a different index server then the 
specialized TargetRetriever should call itself recursively with the found url if the TargetGenerator allows. Discovered 
computing resources are sent to the InterrogateTarget function (or if the TargetRetriever was initialized with a computing 
service url then it should jump directly to this point) which collects the detailed information about the resource.}

\item{{\bf ExecutionTarget} is the class representation of a computing resource (queue) capable of running a grid job. It
serves as input to the broker which is foreseen to be able to select between different ExecutionTargets from different 
grid flavours without apriori knowing their difference. The ExecutionTarget class mimics the glue2 information model with 
a flattened structure, and

Consequently a mapping a is needed

Mapping should be 1 queue equals 1 ExectutionTarget ()}

\end{itemize}

%\begin{figure}[ht]
%\centering{{\scalebox{0.9}{\includegraphics{ng-logo.png}}}
%\caption{\label{fig:myfigure1}The figure shows a logo.} }
%\end{figure}

\chapter{Job Management}
\label{sec:JobManagement}

\chapter{Data Management}
\label{sec:DataManagement}

\chapter{Specialized Classes (Grid Flavour Plugins)}
\label{sec:plugins}
\section{ARC0 Plugins}
\section{ARC1 Plugins}
\section{gLite Plugins}

\chapter{Building Command Line Interfaces}
\label{sec:cli}

\bibliography{grid}
\end{document}

### 2008-08-29 doc/hed/ARCHED_article.tex
Converting HED document into NorduGrid Tech. Doc. according to template. Adding more text about MCCs.

@@
-\documentclass[a4paper,10pt]{article}
-
+\documentclass{book}
+%\documentclass{article}                            %for shorter notes
+\usepackage{graphicx}                              %for PNG images (pdflatex)
+%\usepackage{graphics}                              %for EPS images (latex)
+\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
+\usepackage{color}                                 %for defining custom colors
+\usepackage{framed}                                %for shaded and framed paragraphs
+\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
+\usepackage{geometry}                              %for defining page size
+\usepackage{longtable}                             %for breaking tables
 \usepackage{ifthen}
@@
 \setboolean{verbtechdesc}{true}
+%
+\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
+\hypersetup{
+  pdfauthor = {J. J\"{o}nemo, et al},
+  pdftitle = {The Hosting Environment of the Advanced Resource Connector middleware},
+  pdfsubject = {Paper subject},
+  pdfkeywords = {HED,ARC},
+  pdfcreator = {PDFLaTeX with hyperref package},
+  pdfproducer = {PDFLaTeX}
+}
+%
+\bibliographystyle{IEEEtran}                       %a nice bibliography style
+%
+\def\efill{\hfill\nopagebreak}%
+\hyphenation{Nordu-Grid}
+\setlength{\parindent}{0cm}
+\setlength{\FrameRule}{1pt}
+\setlength{\FrameSep}{8pt}
+\addtolength{\parskip}{5pt}
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+\renewcommand{\arraystretch}{1.3}
+\newcommand{\dothis}{\colorbox{shadecolor}}
+\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
+\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
+\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
+\definecolor{shadecolor}{rgb}{1,1,0.6}
+\definecolor{salmon}{rgb}{1,0.9,1}
+\definecolor{bordeaux}{rgb}{0.75,0.,0.}
+\definecolor{cyan}{rgb}{0,1,1}
+%
+%----- DON'T CHANGE HEADER MATTER
+\begin{document}
+\def\today{\number\day/\number\month/\number\year}
 
-%opening
-\title{The Hosting Environment of the Advanced Resource Connector middleware}
+\begin{titlepage}
 
-\author{J. J\"{o}nemo, et al}
+\begin{tabular}{rl}
+\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
+&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
+\end{tabular}
 
-\begin{document}
+\hrulefill
 
-\maketitle
+%-------- Change this to NORDUGRID-XXXXXXX-NN
 
-\begin{abstract}
+{\raggedleft NORDUGRID-XXXXXXX-NN\par}
 
-\end{abstract}
+{\raggedleft \today\par}
 
-\section{Introduction}
+\vspace*{2cm}
+
+%%%%---- The title ----
+{\centering \textsc{\Large The Hosting Environment of the Advanced Resource Connector middleware}\Large \par}
+\vspace*{0.5cm}
+
+%%%%---- A subtitle, if necessary ----
+%{\centering \textit{\large Paper subtitle}\large \par}
+
+\vspace*{1.5cm}
+%%%%---- A list of authors ----
+    {\centering \large J. J\"{o}nemo\footnote{authors@address}, et al \large \par}
+
+%%%%---- An abstract - if style is article ----
+%\begin{abstract}
+%The abstract
+%\end{abstract}
+\end{titlepage}
+
+\tableofcontents                          %Comment if use article style
+\newpage
+%\chapter{Preface}
+%\section{Introduction}                    %Use Sections for articles
+%\label{sec:intro}
+
+\chapter{Introduction}
 
@@
 
-\section{Architecture}
+\chapter{Architecture}
 
-\subsection{Requirements}
+\section{Requirements}
 % Perhaps this can be an introductory part of the section without any subsection, we'll see
@@
 
-%\subsection{Functional design}
+%\section{Functional design}
 
@@
 
-\subsection{Technical design}
+\section{Technical design}
 
@@
 
-\subsubsection{MCC}
+\subsection{MCC}
 
 In the HED data channels to the outside world may be set up by chains of small processing units called Message Chain Components (MCCs).
-These work on units called Messages which represent data going in to or out of the HED.
+The chain is an ordered list of MCCs and their interconection can be described in the configuration file.
+The MCCs work on units called Messages which represent data going in to or out of the HED.
 The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general Attributes and Security Attributes where information relevant to each protocol is accumulated as the message progresses.
@@
 
-Each MCC has an entry method which is called with Message being processed.
+Each MCC has an entry method \texttt{process} which is called with Message being processed.
 It then processes Message by modifying it or creating new Message.
 Then MCC calls entry method of next MCC in the chain.
-For information how mwssages are handled and about memory management policies please see API description of MCC class in [].
+For information how messages are handled and about memory management policies please see API description of MCC class in [].
+
+The developer who writes an MCC is free to choose any 3rd party library and component to implement the functionalty of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by \texttt{arcloader} and \texttt{arcmessage} libraries of ARC1.
+
+The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs.
+Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component.
 
@@
 
-\paragraph{Payload}
+In general every MCC have optimal and natural places in certain chains and this place cannot always be modified.
+For example on the server side the TCP MCC must be the first MCC in any chain where it is used and the TLS MCC should be right after the TCP MCC.
+
+\subsubsection{Server and Client Side MCCs}
+
+Most of the MCCs has a client and a server version because the behavior of an MCC should be different depending on whether it is sitting on the server or client side.
+The typical scenario here is illustrated by the TCP MCC which should listen and wait for incomming messages on a socket on the server side but call \texttt{connect()} on the client side.
+
+Server and Client side MCCs are separate elements although definitely sharing some code and normally provided inside same plugin module.
+
+\subsubsection{Payload}
 
@@
 
-\paragraph{Attributes}
+\subsubsection{Attributes}
 
@@
 For example the Attribute with key HTTP:METHOD holds HTTP protocol method like GET, PUT, HEAD, etc.
-It is either generated MCC implementing HTTP protocol or is filled by other code and is used by HTTP MCC to generate proper HTTP header.
+It is either generated by MCC implementing HTTP protocol or is filled by other code and is used by HTTP MCC to generate proper HTTP header.
 
-Each MCC developed inside and outside the HED must be have generated and consumed Attributes described in accompanying documentation.
+Each MCC developed inside and outside the HED must have generated and consumed Attributes described in accompanying documentation.
 
-\paragraph{Security Attributes}
+\subsubsection{Security Attributes}
 
@@
 
-\paragraph{Security Handler}
+\subsubsection{Security Handlers}
 
@@
 
-\paragraph{Client and Server MCCs}
+The Security Handlers are arranged into named queues.
+Elements in every queue are executed sequentially with Message as only argument.
+Different queues are executed at  different times.
+Which queue is executed at which case depends on MCC.
+Most MCCs implement two queues named "incoming" and "outgoing".
+The queue "incoming" is executed for Messages moving through the chain towards hosted application and passes as argument the Message with Payload of type corresponding to MCC type i.e.
+HTTP MCC passes Payload with parsed HTTP message.
+The "outgoing" queue is executed for Messages travelling to outside HED.
 
-\paragraph{Configuration}
+\subsection{Services and Clients}
 
-The chain formulation using MCCs is up to the administrator.
-The chain is an ordered list of MCCs and their interconection can be described in the configuration file.
-Based on the configuration information the Loader components formulate the actual chain in the memory of the daemon.
-The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs.
-Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component.
+The services are dynamically loaded on start up just like the MCCs.
+They almost identic to MCCs with an exception that they constitute the last link in the Message Chain.
+They are attached to the Chain in the same manner as other MCCs.
+But differently from MCC the \texttt{process} method of the Service does not pass Message to other components of the Chain.
+Instead they have to process incoming Messages and produce outgoing ones.
 
-Most of the MCCs has a client and a server version because the behavior of an MCC should be different depending on whether it is sitting on the server or client side.
-The typical scenario here is illustrated by the TCP MCC which should listen and wait for incomming messages on a socket on the server side but call \texttt{connect} on the client side.
+The clients are not represented by any specific component.
+The client code sees the Chain as single object with named entry points.
+Those entries are used by clients to insert request Message and get result Message on output.
+For simplifying task of writing clients there is library \texttt{arcclient} provided which wraps task of creating Chains and Messages for widely used SOAP, HTTP, TLS, TCP communications.
 
-In general every MCC have optimal and natural places in certain chains and this place cannot always be modified.
-For example on the server side the TCP MCC must be the first MCC in any chain where it is used and the TLS MCC should be right after the TCP MCC.
+\subsection{Plexer}
 
-The developer who writes an MCC is free to choose any 3rd party library and component to implement the functionalty of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by libarcloader and libarcmessage libraries of ARC1.
+In general case multiple Services living in the HED so the incomming Message should route to the proper Service.
+The Plexer MCC does this job.
+It takes the the ENDPOINT attributes of the message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forward the message to the all matching service.
+It acts as a dispacher.
+The Plexer is also special in a sence that it is not a plugin but part of the \texttt{arcloader} library.
 
-\paragraph{TCP MCC}
+This Plexer provides only basic functionality and is capable of doing only simple routing.
+But because each MCC has multiple routing capabilities it is possible to provide pluggable MCC implementing more sophisticated and/or more specific routing algorithms.
+
+\subsection{Error handling}
+
+\texttt{Relevant classes: Arc::MCC\_Status}
+
+For reporting errors each \texttt{process} method returns instance of MCC\_Status class. That object carries predefined set of common error codes.
+
+This way for error reporting is mostly meant to be used for reporting problems related to code execution.
+For errors caused by processing corresponding protocol MCC should generate proper response Message which carries error description.
+Only if protocol does not provide error handling MCC\_Status should be used.
+It is also advisable to convert MCC\_Status error obtained from next MCC in the Chain into protocol psecific error Message if possible.
+
+\subsection{Instantiation of the Chain}
+
+\texttt{Relevant classes: Arc::Loader}
+
+Chain instantiation is handled by Loader component.
+It takes XML configuration on input and then handles tasks of loading plugin libraries, identifying plugins in them, creating and linking objects of corresponding classes.
+
+Each Loader object may create multiple non-intersecting chains and there may be multiple Loader objects in same executable.
+For each component mentioned in configuration Loader creates single object of corresponding class.
+On Loader destruction all handled components are destroyed too.
+
+\subsection{Sessions and Contexts}
+
+The HED defines few lifetimes for operations happening inside the Chain and components which can be associated with them:
+
+%The HED provides few context containers associated with the Chain and the Message. Those are:
+
+\begin{enumerate}
+
+\item The Message lifetime - lasts as long as incoming and outgoing Messages are passing through the Chain forth and back.
+The Message it self in this case is used as container for associated components.
+
+\item The Session lifetime - is defined by existence of some logical connection between Messages being processed.
+Corresponding container MessageContext is normally created by first MCC in a chain.
+Actual lifetime of that container is MCC specific.
+For TCP MCC it corresponds to TCP connection.
+The MessageContext holds objects inherited from MessageContextElement class.
+When Session ends all associated objects are destroyed.
+
+%\item The Session Context. It holds objects inherited from MessageContextElement class and is persistent withion
+
+\end{enumerate}
+
+\subsection{Alternative implementation languages}
+
+In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done 
+%Not much aboút the actual services, more about what they need to implement and what
+
+The chain formulation using MCCs is up to the administrator.  Based on the configuration information the Loader components formulate the actual chain in the memory of the daemon.
+
+\chapter{Implemented elements}
+
+\section{Implemented MCCs}
+
+\subsection{TCP MCC}
 
@@
 
-\paragraph{TLS MCC}
+\subsubsection{TLS MCC}
 The server TLS MCC enables ...
@@
 
-\paragraph{HTTP MCC}
+\subsection{HTTP MCC}
 
-\paragraph{SOAP MCC}
+\subsection{SOAP MCC}
 
-\paragraph{Plexer}
+\section{Implemented Security Handlers}
 
-In general case multiple service living in the HED so the incomming message should route to the proper service.
-The Plexer MCC does this job.
-It takes the the ENDPOINT attributes of the message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forward the message to the all matching service.
-It acts as a dispacher.
-The Plexer is special in a sence that it is not a plugin but part of the libarcloader library.
+\section{Implemented PDCs}
 
-\subsubsection{Services}
-
-The services are dynamically loaded on start up just like the MCCs.
-They can even be regarded as special cases of MCCs in that they constitute the last link in the so called Message Chain and are attached in much the same manner as other MCCs.
-After being loaded, instanciated and configured by the HED, services are provided with a range of...
-
-The services are typically expected to receive parsed SOAP messages which they service - possibly by invoking or communicating with external processes.
-
-\subsubsection{PDC}
-
-\subsubsection{Security Handlers}
-
-\subsubsection{Alternative implementation languages}
-
-In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done 
-%Not much aboút the actual services, more about what they need to implement and what 
-\section{Conclusion}
+\chapter{Conclusion}
 

### 2008-08-30 doc/hed/ARCHED_article.tex
More of HED documentation.

@@
 Statements like "you are doing crap, I'll better use Axis" are understood but not welcome and are usually reacted to adequately.
+
+\chapter{How to read}
+
+This document does not (yet?) include in depth description of C++ classes which constitute the HED.
+Instead most section contain notes entitled "Relevant classes".
+Technical description of those classes can be found in automatically generated "Hosting Environment (Daemon) Reference Manual" document [].
+
+For examples please see source code in repository []. Many components and libraries are accompanied with test and example applications.
 
@@
 
+\texttt{Relevant classes: Arc::Plexer }
+
 In general case multiple Services living in the HED so the incomming Message should route to the proper Service.
@@
 
-The HED defines few lifetimes for operations happening inside the Chain and components which can be associated with them:
+\texttt{Relevant classes: Arc::MessageContext, Arc::MessageContextElement, Arc::MessageAuthContext, Arc::ChianContext }
+
+The HED defines three lifetimes for operations happening inside the Chain and components which can be associated with them:
 
@@
 \item The Session lifetime - is defined by existence of some logical connection between Messages being processed.
-Corresponding container MessageContext is normally created by first MCC in a chain.
+Corresponding container MessageContext is normally created by first MCC in a chain and attached to the Message.
 Actual lifetime of that container is MCC specific.
@@
 The MessageContext holds objects inherited from MessageContextElement class.
-When Session ends all associated objects are destroyed.
+The Security Attributes can also be stored in dedicated container - MessageAuthContext - with Session lifetime.
+At end of  the Session all associated objects are destroyed.
 
-%\item The Session Context. It holds objects inherited from MessageContextElement class and is persistent withion
+\item The Chain lifetime - is time period while componnets making the Chain exist.
+This lifetime is represented by ChainContext class.
+Differently from other context objects this one doe not allow free manipulation of contained objects.
+Instead it provides an interface to some internal structures of Loader object.
+Those include factories and lists of object of all types created by particular instance of the Loader.
 
@@
 
+\subsection{DMC}
+
+\texttt{Relevant classes: Arc::DataPoint, Arc::DataMover and related classes.}
+
+The HED defines an interface for pluggable components implementing higher-level information transfer and query.
+Those are Data Management Components (DMC).
+Each DMC is inherited from Datapoint class and provides subset of methods for performing following operations on data endpoint:
+
+\begin{enumerate}
+
+\item Read data from specified endpoint into DataBuffer class object
+
+\item Write data into specified endpoint from DataBuffer class object
+
+\item List subcontent of endpoint (i.e. list files in directory)
+
+\item Register and unregister presence of data
+
+\end{enumerate}
+
+The DMC may be implemented using third party software like it is currently done for (Grid)FTP DMC. But implementation may use Message Chains too like in case of HTTP MCC.
+
+Along with ordinary endpoints defining location of data directly - like HTTP, FTP, LDAP - the DMC can be used with indirect/indexing endpoints.
+Those are endpoints which define only location of meta-data associated with actual data or an interface/service providing functionality of managing/requesting data.
+
+% TODO a lot
+
 \subsection{Alternative implementation languages}
 
-In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done 
-%Not much aboút the actual services, more about what they need to implement and what
+In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done in Python language.
 
-The chain formulation using MCCs is up to the administrator.  Based on the configuration information the Loader components formulate the actual chain in the memory of the daemon.
+% TODO it
 
 \chapter{Implemented elements}
+
+This chapter describes components which are implemented alongside with the HED infractructure itself.
+Although strictly not belonging to infrastructure this minimal set of components is necessary to make inrastructure useful.
 
@@
 
-The server TCP MCC in the HED is special in that it produces messages by listening on a socket rather than passing on messages from other MCCs.
-As such it spawns new threads to handle the message and its response throughout the message chain.
-One could envision other parallel MCCs having these properties but producing messages from other sources such as e.g.
+The server TCP MCC in the HED is special in that it produces messages by listening on a network socket rather than passing on messages from other MCCs.
+As such it spawns new thread for every new connection to handle Messages and their responses throughout the Message Chain.
+One could envision other MCCs having these properties but producing messages from other sources such as e.g.
 unix sockets.
@@
 
-\section{Implemented PDCs}
+\section{Implemented DMCs}
+
+\subsection{File DMC}
+
+\subsection{GridFTP/FTP DMC}
+
+\subsection{HTTP DMC}
+
+\subsection{LDAP DMC}
+
+\subsection{LFC DMC}
+
+\subsection{RLS DMC}
 

### 2008-08-31 doc/hed/ARCHED_article.tex
One more update of HED document

@@
 
-%-------- Change this to NORDUGRID-XXXXXXX-NN
-
-{\raggedleft NORDUGRID-XXXXXXX-NN\par}
+{\raggedleft NORDUGRID-TECH-19\par}
 
@@
 Instead most section contain notes entitled "Relevant classes".
-Technical description of those classes can be found in automatically generated "Hosting Environment (Daemon) Reference Manual" document [].
+Technical description of those classes can be found in automatically generated "Hosting Environment (Daemon) Reference Manual" document \cite{hed-api}.
 
-For examples please see source code in repository []. Many components and libraries are accompanied with test and example applications.
+For examples please see source code in repository \cite{svn-arc1}. Many components and libraries are accompanied with test and example applications.
 
@@
 Then MCC calls entry method of next MCC in the chain.
-For information how messages are handled and about memory management policies please see API description of MCC class in [].
+For information how messages are handled and about memory management policies please see API description of MCC class in \cite{hed-api}.
 
@@
 
-Here only basic information about security related objects is presented. For more detailed information please see "Security Framework of ARC1" [].
+\texttt{Relevant classes: Arc::SecAttr}
+
+Here only basic information about security related objects is presented. For more detailed information please see "Security Framework of ARC1" \cite{sec-arc1}.
 
@@
 Currently only implemented is ARC Authorization Request/Policy (see below).
-Please see API description of SecAttr and MessageAuth classes in [] for more information.
+Please see API description of SecAttr and MessageAuth classes in \cite{hed-api} for more information.
 
@@
 \subsubsection{Security Handlers}
+
+\texttt{Relevant classes: ArcSec::SecHander}
 
@@
 
-\texttt{Relevant classes: Arc::DataPoint, Arc::DataMover and related classes.}
+\texttt{Relevant classes: Arc::DataPoint, Arc::DataMover, Arc::DataBuffer and related classes.}
 
@@
 Those are endpoints which define only location of meta-data associated with actual data or an interface/service providing functionality of managing/requesting data.
+For more information about indexing endpoints see description of supported URLs in \cite{urls}.
 
-% TODO a lot
+More in depth techical information about the DMC can be found in \cite{dmc}.
 
@@
 
+The server and client side MCCs are called \texttt{tcp.service} and \texttt{tcp.client} correspondingly. They are located in the shared library called \texttt{mcctcp}.
+
 \subsubsection{TLS MCC}
@@
 
+\bibliography{grid}
+
 \end{document}

### 2008-09-01 doc/client/client_technical.tex
added text and illustration

@@
 \begin{itemize}
-
 \item{{\bf TargetGenerator} is responsible for loading TargetRetriever plugins in accordance with the input URLs. 
-e.g. if an URL pointing to an ARC1 resource is given the TargetRetrieverARC1 is loaded. If no URLs are given, the 
-TargetGenerator makes use of the user/system defaults. When queried for resources (``targets'') the TargetGenerator 
-passes the request to the loaded TargetRetrievers running them as individual threads for improved performance. The 
-TargetGenerator keeps records of by TargetRetriever found services (index of computing) in order to avoid multiple 
-queries to the same resource. Accepted targets are registered in the FoundTargets array of the TargetGenerator.}
+e.g. if an URL pointing to an ARC1 resource is given the TargetRetrieverARC1 is loaded. When queried for resources 
+(``targets'') the TargetGenerator passes the request to the loaded TargetRetrievers running them as individual threads 
+for improved performance. The TargetGenerator keeps records of by TargetRetriever found servers and services 
+(index or computing) in order to avoid multiple queries. Accepted targets are stored in the FoundTargets array kept by 
+the TargetGenerator.}
 
@@
 responsible for querying index servers, and if the result yields an addresss to a different index server then the 
-specialized TargetRetriever should call itself recursively with the found url if the TargetGenerator allows. Discovered 
-computing resources are sent to the InterrogateTarget function (or if the TargetRetriever was initialized with a computing 
-service url then it should jump directly to this point) which collects the detailed information about the resource.}
+specialized TargetRetriever should call itself recursively with the found url. Discovered computing services are sent 
+to the InterrogateTarget function (or if the TargetRetriever was initialized with a computing service url then it should 
+jump directly to this point) which collects the detailed information about the service.}
 
@@
 serves as input to the broker which is foreseen to be able to select between different ExecutionTargets from different 
-grid flavours without apriori knowing their difference. The ExecutionTarget class mimics the glue2 information model with 
-a flattened structure, and
+grid flavours without apriori knowing their difference. The ExecutionTarget class mimics the glue2 information model (with 
+a flattened structure), and thus a mapping between attributes from other information systems into the glue2 format is needed. 
+Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares. All information can be printed by 
+invoking the ExecutionTarget's ``Stat'' command which prints all variables (if they have been filled during the information 
+retrieval process).
 
-Consequently a mapping a is needed
-
-Mapping should be 1 queue equals 1 ExectutionTarget ()}
-
+Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
+selected ExecutionTarget, the ExecutionTarget is capable of returning a submitter (see next section) suitable submitting a 
+job to the resource which is represents.}
 \end{itemize}
 
-%\begin{figure}[ht]
-%\centering{{\scalebox{0.9}{\includegraphics{ng-logo.png}}}
-%\caption{\label{fig:myfigure1}The figure shows a logo.} }
-%\end{figure}
+Figure \ref{fig:ResDisc} depicts how the classes work together in a command chain to discover all resources registered with 
+a certain information server. Below a description of each step is given:
 
-\chapter{Job Management}
+\begin{figure}[ht]
+\centering{{\scalebox{0.75}{\includegraphics{TargetDiscovery.png}}}
+\caption{\label{fig:ResDisc}Diagram depicting the process of resource discovery.} }
+\end{figure}
+
+\begin{enumerate}
+\item{The TargetGenerator takes three URL arrays as input. First array contains individually selected clusters, while the 
+second array contains individually selected index servers (1a). The last array contains a list of rejected clusters. This 
+list is parsed through alias resolve before being inserted as filter to the storage of found services (1b).}
+\item{The complete list of URLs pointing to services and index servers.}
+\item{For each URL a TargetRetriever plugin is loaded using the HED ? loader. The TargetRetriever is initialized 
+with the URL and the information about wether the URL points to a service or an index server.}
+\item{An external call is received calling for targets to be prepared. Call is propagated to all loaded TargetRetrievers.}
+\item{The call for targets is processed by each TargetRetriever in parallel. If the TargetRetriever is initialized with an 
+index server URL it first registers at the Index Servers store kept by the TargetGenerator. If allowed to register, data 
+is read out from the server and processed.}
+\item{The TargetRetriver finds a registered service and contacts the Found Services store kept by the TargetGenerator to check 
+if the service in question has been found be another TargetRetriever. (Services often registers at more than one index server, 
+thus different TargetRetrievers may discover the same service.)}
+\item{If registered in the Found Services store the TargetRetriever progresses to interrogate the service.}
+\item{Found and processed ExecutionTargets are stored in the Found Targets store kept by the TargetGenerator for later usage 
+(e.g. status printing or job submission).}
+\item{Index servers often registers with other index servers creating a nested hierarchical information structure. For this 
+reasons a TargetRetriever interpreting the read out result from an index server may encounter other index servers. In that 
+event the TargetRetriever calls itself recursively creating a new thread for the new URL. This ensures optimum speed and 
+performance for resource discovery.}
+\end{enumerate}
+
+The example above outlines how the TargetRetriever works when initialized with an URL pointing to and index server. However, 
+as indicated it also accepts service URLs in which case the TargetRetriever will try to register the service and if allowed 
+read all its details through the InterrogateTarget method.
+
+\chapter{Job Submission and Management}
 \label{sec:JobManagement}
+\section{Submission}
+Almost every day a new grid cluster is purchased and made available around the world (either to the 
+general public or restricted to authorized users). More available clusters create a need for better and 
+faster resource discovery and information retrieval, making this task an utmost important component of the 
+new client library.
+
+\begin{itemize}
+\item{{\bf Submitter} is base class}
+
+\item{{\bf JobDescription} is }
+
+\item{{\bf Broker} is }
+\end{itemize}
+
+\section{Management}
+Almost every day a new grid cluster is purchased and made available around the world (either to the 
+general public or restricted to authorized users). More available clusters create a need for better and 
+faster resource discovery and information retrieval, making this task an utmost important component of the 
+new client library.
+
+\begin{itemize}
+\item{{\bf JobSupervisor} is base class}
+
+\item{{\bf JobController} is }
+
+\item{{\bf Job} is }
+\end{itemize}
 

### 2008-09-02 doc/client/client_technical.tex
added text about job submission

@@
 Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
-selected ExecutionTarget, the ExecutionTarget is capable of returning a submitter (see next section) suitable submitting a 
+selected ExecutionTarget, the ExecutionTarget is capable of returning a submitter (see next section) capable of submitting a 
 job to the resource which is represents.}
@@
 \section{Submission}
-Almost every day a new grid cluster is purchased and made available around the world (either to the 
-general public or restricted to authorized users). More available clusters create a need for better and 
-faster resource discovery and information retrieval, making this task an utmost important component of the 
-new client library.
+Job submission starts with the resource discovery and target preparation as outlined in the section above. Not before a list 
+of possible targets (which allows the user) is available is the job description read in order to enable bulk job submission of 
+widely different jobs without having to reperform the resource discovery. The job submission makes use of the following classes:
 
 \begin{itemize}
-\item{{\bf Submitter} is base class}
+\item{{\bf Broker} is not yet implemented and thus not described here.}
 
-\item{{\bf JobDescription} is }
+\item{{\bf JobDescription} is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
+JSDL job description languages and to some extent also translate between them. Validation (syntax and semantical) is performed 
+automatically when the source is set. Features easy retrieval of list of local input files which should be uploaded to the 
+computing service.}
 
-\item{{\bf Broker} is }
+\item{{\bf Submitter} is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
+uploads (by the job needed) local input files. Returns an XML node containing all needed information about the job 
+for later job management. The Submitter is returned by the ExecutionTarget selected for job execution and thus the 
+ExecutionTarget populates (through XML config element) the Submitter with information about submission endpoint (URL) 
+and job description languages ``spoken'' by the target.}
+
 \end{itemize}
 
+\begin{figure}[ht]
+\centering{{\scalebox{0.75}{\includegraphics{JobSubmission.png}}}
+\caption{\label{fig:JobSub}Diagram depicting the process of submitting a job to a computing service, including finding the 
+best possible resource for the job.} }
+\end{figure}
+
+Figure \ref{fig:JobSub} shows a job submission sequence and below a description of each step is given:
+
+\begin{enumerate}
+\item{The TargetGenerator has prepared a list of targets. Depending on the URLs provided to the TargetGenerator the list of 
+found targets may be empty or contain several targets. Targets may even represent more than one grid flavour. The list of 
+found targets are taken as input to the Broker.}
+\item{In order to choose between services the Broker needs to know the specific job requirements, thus the JobDescription is 
+needed as input to the brokering process.}
+\item{The broker outputs a ordered list of ExecutionTargets according to the provided JobDescription.}
+\item{As explained in Section~\ref{sec:TargetDiscovery} each ExectionTarget has a method to return a specialized Submitter 
+capable of submitting jobs to the service it represents. Thus the best suitable ExecutionTarget for the job is asked to return 
+a Submitter for the submission.}
+\item{The Submitter takes the JobDescription as input and uploads it to the computing service.}
+\item{The Submitter identifies local input files from the JobDescription and uploads them to the computing service.}
+\end{enumerate}
+
 \section{Management}
-Almost every day a new grid cluster is purchased and made available around the world (either to the 
-general public or restricted to authorized users). More available clusters create a need for better and 
-faster resource discovery and information retrieval, making this task an utmost important component of the 
-new client library.
+Information about active jobs is stored in a local xml file which is used for later job management. As this file may contain 
+information about jobs running on completely different grid flavours, job management should be handled using plugins similar to 
+resource discovery and job submission. The following classes are used:
 
 \begin{itemize}
-\item{{\bf JobSupervisor} is base class}
+\item{{\bf JobSupervisor} is responsible for loading the appropriate JobController for managing a job running on a specific 
+grid flavour. In order to match a job identification string (jobid) with a grid flavour a lookup in the local xml file is 
+performed. The JobSupervisor main functionality is to return a list of JobControllers which are capable of managing all jobs 
+addressed by a certain command.}
 
-\item{{\bf JobController} is }
+\item{{\bf JobController} is both base class for grid specific specializations, but also the implementor of all public functionality 
+offered by the JobControllers. In other words all virtual functions of the JobController are private. Upon initialization the 
+JobController reads into and fills a local store of jobs called JobsStore.
+The public functions of the 
+JobController offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations for issueing the command. 
+This ensures consistency in how the commands are handled }
 

### 2008-09-02 doc/client/client_technical.tex
new structure

@@
 This document describes from a technical viewpoint a plugin based client library for the new 
-Web Service (WS) based Advanced Resource Connector (ARC) middleware. The library is capable of 
-handling resource discovery, job submission and job control and interoperates with several 
-different grid flavours. The support for different grid flavours is enabled through the modular 
-design using plugins specialized for each supported middleware, and presently supported are ARC0 
-(ARC classic), ARC1 and the gLite middlewares. Future extensions to support additional middlewares 
-involve plugin compilation only i.e. no recompilation of main libraries or client.
+Web Service (WS) based Advanced Resource Connector (ARC) middleware. The library consists of 
+a set of C++ classes for
+
+\begin{itemize}
+\item{handling proxy, user and host certificates,}
+\item{performing resource discovery and information retrieval,}
+\item{job submission and management and}
+\item{data handling.}
+\end{itemize}
+
+All capabilities are enabled for three different grid flavours (Production ARC, ARC1 and gLite) 
+through the modular design using plugins specialized for each supported middleware. Future 
+extensions to support additional middlewares involve plugin compilation only i.e. no recompilation 
+of main libraries or client.
 
@@
 
-Finally the library has been designed with care for later usage within a graphical user interface. This 
-particularly applies to the job management part.
+In the following we will give a functionality overview in Section~\ref{sec:FuncOver} while all technical 
+details will be given in Section~\ref{sec:Implementation}. Section~\ref{sec:cli} will through examples 
+show how command line interfaces can be built upon the library.
 
-\chapter{Resource Discovery and Information Retrieval}
+\chapter{Functionality Overview}
+\label{sec:FuncOver}
+The new ARClib makes extensive use of plugins for command handling. These plugins are handled by a set 
+of higher level classes which thus are the ones to offer the plugin functionality to external calls. In 
+this section an overview of the library's main functionality is given which also introduces the most 
+important higher level classes.
+\section{Resource Discovery and Information Retrieval}
 \label{sec:TargetDiscovery}
-Almost every day a new grid cluster is purchased and made available around the world (either to the 
-general public or restricted to authorized users). More available clusters create a need for better and 
-faster resource discovery and information retrieval, making this task an utmost important component of the 
-new client library.
+With the increasing number of grid clusters around the world, a reliable and fast resource discovery and 
+information retrieval capability is of crucial importance for a user interface. The new ARClib resource 
+discovery and information retrieval component consists of three classes; The TargetGenerator, the 
+TargetRetriever and the ExecutionTarget. Of these the TargetRetriever is a base class for further grid 
+middleware specific specialization (plugin).
 
-The new client library's resource discovery and information retrieval component consists of three classes, one 
-of which a base class for further grid middleware specific specialization (plugin):
-
-\begin{itemize}
-\item{{\bf TargetGenerator} is responsible for loading TargetRetriever plugins in accordance with the input URLs. 
-e.g. if an URL pointing to an ARC1 resource is given the TargetRetrieverARC1 is loaded. When queried for resources 
-(``targets'') the TargetGenerator passes the request to the loaded TargetRetrievers running them as individual threads 
-for improved performance. The TargetGenerator keeps records of by TargetRetriever found servers and services 
-(index or computing) in order to avoid multiple queries. Accepted targets are stored in the FoundTargets array kept by 
-the TargetGenerator.}
-
-\item{{\bf TargetRetriever} is base class for TargetRetriever grid middleware specializations. Contains the pure 
-virtual method GetTargets which is to be implemented by the specialized class. While it is not mandatory it is recommended 
-that the specialized class divides this function into two components: QueryIndex and InterrogateTarget. The former is 
-responsible for querying index servers, and if the result yields an addresss to a different index server then the 
-specialized TargetRetriever should call itself recursively with the found url. Discovered computing services are sent 
-to the InterrogateTarget function (or if the TargetRetriever was initialized with a computing service url then it should 
-jump directly to this point) which collects the detailed information about the service.}
-
-\item{{\bf ExecutionTarget} is the class representation of a computing resource (queue) capable of running a grid job. It
-serves as input to the broker which is foreseen to be able to select between different ExecutionTargets from different 
-grid flavours without apriori knowing their difference. The ExecutionTarget class mimics the glue2 information model (with 
-a flattened structure), and thus a mapping between attributes from other information systems into the glue2 format is needed. 
-Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares. All information can be printed by 
-invoking the ExecutionTarget's ``Stat'' command which prints all variables (if they have been filled during the information 
-retrieval process).
-
-Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
-selected ExecutionTarget, the ExecutionTarget is capable of returning a submitter (see next section) capable of submitting a 
-job to the resource which is represents.}
-\end{itemize}
-
-Figure \ref{fig:ResDisc} depicts how the classes work together in a command chain to discover all resources registered with 
-a certain information server. Below a description of each step is given:
+Figure \ref{fig:ResDisc} depicts how the classes work together in a command chain to discover all resources 
+registered with a certain information server. Below a description of each step is given:
 
@@
 \centering{{\scalebox{0.75}{\includegraphics{TargetDiscovery.png}}}
-\caption{\label{fig:ResDisc}Diagram depicting the process of resource discovery.} }
+\caption{\label{fig:ResDisc}Diagram depicting the process of resource discovery and information retrieval.} }
 \end{figure}
@@
 
-\chapter{Job Submission and Management}
-\label{sec:JobManagement}
-\section{Submission}
+\section{Job Submission}
+\label{sec:JobSubmission}
 Job submission starts with the resource discovery and target preparation as outlined in the section above. Not before a list 
@@
 widely different jobs without having to reperform the resource discovery. The job submission makes use of the following classes:
-
-\begin{itemize}
-\item{{\bf Broker} is not yet implemented and thus not described here.}
-
-\item{{\bf JobDescription} is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
-JSDL job description languages and to some extent also translate between them. Validation (syntax and semantical) is performed 
-automatically when the source is set. Features easy retrieval of list of local input files which should be uploaded to the 
-computing service.}
-
-\item{{\bf Submitter} is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
-uploads (by the job needed) local input files. Returns an XML node containing all needed information about the job 
-for later job management. The Submitter is returned by the ExecutionTarget selected for job execution and thus the 
-ExecutionTarget populates (through XML config element) the Submitter with information about submission endpoint (URL) 
-and job description languages ``spoken'' by the target.}
-
-\end{itemize}
 
@@
 
-\section{Management}
+\section{Job Management}
 Information about active jobs is stored in a local xml file which is used for later job management. As this file may contain 
@@
 
-\begin{itemize}
-\item{{\bf JobSupervisor} is responsible for loading the appropriate JobController for managing a job running on a specific 
+\section{Data Handling}
+\label{sec:DataHandling}
+
+\chapter{Implementation}
+\label{sec:Implementation}
+Clever text here.
+\section{Generic Classes}
+\subsection{TargetGenerator} The TargetGenerator is responsible for loading TargetRetriever plugins in accordance with the 
+input URLs. e.g. if an URL pointing to an ARC1 resource is given the TargetRetrieverARC1 is loaded. When queried for resources 
+(``targets'') the TargetGenerator passes the request to the loaded TargetRetrievers running them as individual threads 
+for improved performance. The TargetGenerator keeps records of by TargetRetriever found servers and services 
+(index or computing) in order to avoid multiple queries. Accepted targets are stored in the FoundTargets array kept by 
+the TargetGenerator.
+
+\subsection{TargetRetriever} The TargetRetriever is base class for TargetRetriever grid middleware specializations. Contains the pure 
+virtual method GetTargets which is to be implemented by the specialized class. While it is not mandatory it is recommended 
+that the specialized class divides this function into two components: QueryIndex and InterrogateTarget. The former is 
+responsible for querying index servers, and if the result yields an addresss to a different index server then the 
+specialized TargetRetriever should call itself recursively with the found url. Discovered computing services are sent 
+to the InterrogateTarget function (or if the TargetRetriever was initialized with a computing service url then it should 
+jump directly to this point) which collects the detailed information about the service.
+
+\subsection{ExecutionTarget} The ExecutionTarget is the class representation of a computing resource (queue) capable of running a grid job. It
+serves as input to the broker which is foreseen to be able to select between different ExecutionTargets from different 
+grid flavours without apriori knowing their difference. The ExecutionTarget class mimics the glue2 information model (with 
+a flattened structure), and thus a mapping between attributes from other information systems into the glue2 format is needed. 
+Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares. All information can be printed by 
+invoking the ExecutionTarget's ``Stat'' command which prints all variables (if they have been filled during the information 
+retrieval process).
+
+Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
+selected ExecutionTarget, the ExecutionTarget is capable of returning a submitter (see next section) capable of submitting a 
+job to the resource which is represents.
+
+\subsection{Broker} is not yet implemented and thus not described here.
+
+\subsection{JobDescription} The JobDescription is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
+JSDL job description languages and to some extent also translate between them. Validation (syntax and semantical) is performed 
+automatically when the source is set. Features easy retrieval of list of local input files which should be uploaded to the 
+computing service.
+
+\subsection{Submitter} Submitter is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
+uploads (by the job needed) local input files. Returns an XML node containing all needed information about the job 
+for later job management. The Submitter is returned by the ExecutionTarget selected for job execution and thus the 
+ExecutionTarget populates (through XML config element) the Submitter with information about submission endpoint (URL) 
+and job description languages ``spoken'' by the target.
+
+\subsection{JobSupervisor} is responsible for loading the appropriate JobController for managing a job running on a specific 
 grid flavour. In order to match a job identification string (jobid) with a grid flavour a lookup in the local xml file is 
 performed. The JobSupervisor main functionality is to return a list of JobControllers which are capable of managing all jobs 
-addressed by a certain command.}
+addressed by a certain command.
 
-\item{{\bf JobController} is both base class for grid specific specializations, but also the implementor of all public functionality 
+\subsection{JobController} is both base class for grid specific specializations, but also the implementor of all public functionality 
 offered by the JobControllers. In other words all virtual functions of the JobController are private. Upon initialization the 
-JobController reads into and fills a local store of jobs called JobsStore.
-The public functions of the 
-JobController offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations for issueing the command. 
-This ensures consistency in how the commands are handled }
+JobController converts the information stored in the local XML to Jobs which are stored in a JobControler internal store called JobsStore. 
+The public functions of the JobController offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations 
+for issueing the command. This ensures consistency in how the commands are handled between the different grid flavours. Jobs that are 
+permanently removed following a command are removed from both the JobStore and local xml file. Thus it is foreseen that the JobControllers 
+may be used in a future graphical user interface where it becomes a persistent object continously monitoring jobs.
 
-\item{{\bf Job} is }
-\end{itemize}
+\subsection{Job} is generic job class for storing all job related information. Attributes are derived from the glue2 information model 
+and thus a mapping is needed for non glue2 compliant grid middlewares. Appendix XX shows the present mapping schema.
 
-\chapter{Data Management}
-\label{sec:DataManagement}
+\subsection{UserConfig}
 
-\chapter{Specialized Classes (Grid Flavour Plugins)}
+\section{Specialized Classes (Grid Flavour Plugins)}
 \label{sec:plugins}
-\section{ARC0 Plugins}
-\section{ARC1 Plugins}
-\section{gLite Plugins}
+\subsection{ARC0 Plugins}
+\subsection{ARC1 Plugins}
+\subsection{gLite Plugins}
+
+\section{Support Classes}
+\subsection{URL}
+\subsection{Logger}
 

### 2008-09-02 doc/client/client_technical.tex
even more text

@@
 \newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
+\newcommand{\TargetGenerator}{\texttt{TargetGenerator}}
+\newcommand{\TargetRetriever}{\texttt{TargetRetriever}}
+\newcommand{\ExecutionTarget}{\texttt{ExecutionTarget}}
+\newcommand{\Broker}{\texttt{Broker}}
+\newcommand{\Submitter}{\texttt{Submitter}}
+\newcommand{\JobDescription}{\texttt{JobDescription}}
+\newcommand{\JobSupervisor}{\texttt{JobSupervisor}}
+\newcommand{\JobController}{\texttt{JobController}}
+\newcommand{\Job}{\texttt{Job}}
 \definecolor{shadecolor}{rgb}{1,1,0.6}
@@
 information retrieval capability is of crucial importance for a user interface. The new ARClib resource 
-discovery and information retrieval component consists of three classes; The TargetGenerator, the 
-TargetRetriever and the ExecutionTarget. Of these the TargetRetriever is a base class for further grid 
+discovery and information retrieval component consists of three classes; The {\TargetGenerator}, the 
+{\TargetRetriever} and the {\ExecutionTarget}. Of these the {\TargetRetriever} is a base class for further grid 
 middleware specific specialization (plugin).
@@
 \begin{enumerate}
-\item{The TargetGenerator takes three URL arrays as input. First array contains individually selected clusters, while the 
-second array contains individually selected index servers (1a). The last array contains a list of rejected clusters. This 
+\item{The {\TargetGenerator} takes three URL arrays as input. First array contains individually selected computing services, while the 
+second array contains individually selected index servers (1a). The last array contains a list of rejected services. This 
 list is parsed through alias resolve before being inserted as filter to the storage of found services (1b).}
-\item{The complete list of URLs pointing to services and index servers.}
-\item{For each URL a TargetRetriever plugin is loaded using the HED ? loader. The TargetRetriever is initialized 
-with the URL and the information about wether the URL points to a service or an index server.}
-\item{An external call is received calling for targets to be prepared. Call is propagated to all loaded TargetRetrievers.}
-\item{The call for targets is processed by each TargetRetriever in parallel. If the TargetRetriever is initialized with an 
-index server URL it first registers at the Index Servers store kept by the TargetGenerator. If allowed to register, data 
-is read out from the server and processed.}
-\item{The TargetRetriver finds a registered service and contacts the Found Services store kept by the TargetGenerator to check 
-if the service in question has been found be another TargetRetriever. (Services often registers at more than one index server, 
-thus different TargetRetrievers may discover the same service.)}
-\item{If registered in the Found Services store the TargetRetriever progresses to interrogate the service.}
-\item{Found and processed ExecutionTargets are stored in the Found Targets store kept by the TargetGenerator for later usage 
+\item{The complete list of URLs pointing to computing services and index servers.}
+\item{For each URL a {\TargetRetriever} plugin is loaded using the ARC loader. The {\TargetRetriever} is initialized 
+with the URL and the information about this URL pointing to a computing service or an index server.}
+\item{An external call is received calling for targets to be prepared. Call is propagated to all loaded {\TargetRetriever}s.}
+\item{The call for targets is processed by each {\TargetRetriever} in parallel. If the {\TargetRetriever} is initialized with an 
+index server URL it first registers at the Index Servers store kept by the {\TargetGenerator}. If allowed to register, the index 
+server is queried and the query result processed.}
+\item{The TargetRetriver finds a registered service and contacts the Found Services store kept by the {\TargetGenerator} to check 
+if the service in question has been found be another {\TargetRetriever}. (Services often registers at more than one index server, 
+thus different {\TargetRetriever}s may discover the same service.)}
+\item{If the {\TargetGenerator} allows registration in the Found Services store, the {\TargetRetriever} progresses to interrogate the service.}
+\item{Found and processed {\ExecutionTarget}s are stored in the Found Targets store kept by the {\TargetGenerator} for later usage 
 (e.g. status printing or job submission).}
 \item{Index servers often registers with other index servers creating a nested hierarchical information structure. For this 
-reasons a TargetRetriever interpreting the read out result from an index server may encounter other index servers. In that 
-event the TargetRetriever calls itself recursively creating a new thread for the new URL. This ensures optimum speed and 
+reason a {\TargetRetriever} processing the query result from an index server may encounter other index servers. In such
+events the {\TargetRetriever} will call itself recursively creating a new thread for the new URL. This ensures optimum speed and 
 performance for resource discovery.}
@@
 
-The example above outlines how the TargetRetriever works when initialized with an URL pointing to and index server. However, 
-as indicated it also accepts service URLs in which case the TargetRetriever will try to register the service and if allowed 
-read all its details through the InterrogateTarget method.
+The example above outlines how the {\TargetRetriever} works when initialized with an URL pointing to and index server. However, 
+as indicated it also accepts initialization with a service URLs, in which case the {\TargetRetriever} will try to register the 
+service in the {\TargetGenerator} Found Services store. If allowed the {\TargetRetriever} will progress with the target interrogation 
+method.
 
@@
 \label{sec:JobSubmission}
-Job submission starts with the resource discovery and target preparation as outlined in the section above. Not before a list 
-of possible targets (which allows the user) is available is the job description read in order to enable bulk job submission of 
-widely different jobs without having to reperform the resource discovery. The job submission makes use of the following classes:
+Job submission starts with the resource discovery and target preparation as outlined in the Section~\ref{sec:TargetDiscovery}. 
+Not until a list of possible targets (which allows the user) is available is the job description read in order to enable bulk 
+job submission of widely different jobs without having to reperform the resource discovery. In addition to the classes mentioned 
+above the job submission makes use of the {\Broker}, {\JobDescription} and {\Submitter} classes. At the time of writing the {\Broker} is 
+not implemented, but it's functionality and interface is well known, and thus we will include it in the overview given here.
+The {\Submitter} is base class for further grid middleware specific specialization (plugin) similarly to the {\TargetRetriever}.
 
@@
 \centering{{\scalebox{0.75}{\includegraphics{JobSubmission.png}}}
-\caption{\label{fig:JobSub}Diagram depicting the process of submitting a job to a computing service, including finding the 
-best possible resource for the job.} }
+\caption{\label{fig:JobSub}Diagram depicting the process of submitting a job to a computing service.}}
 \end{figure}
@@
 \begin{enumerate}
-\item{The TargetGenerator has prepared a list of targets. Depending on the URLs provided to the TargetGenerator the list of 
-found targets may be empty or contain several targets. Targets may even represent more than one grid flavour. The list of 
-found targets are taken as input to the Broker.}
-\item{In order to choose between services the Broker needs to know the specific job requirements, thus the JobDescription is 
-needed as input to the brokering process.}
-\item{The broker outputs a ordered list of ExecutionTargets according to the provided JobDescription.}
-\item{As explained in Section~\ref{sec:TargetDiscovery} each ExectionTarget has a method to return a specialized Submitter 
-capable of submitting jobs to the service it represents. Thus the best suitable ExecutionTarget for the job is asked to return 
-a Submitter for the submission.}
-\item{The Submitter takes the JobDescription as input and uploads it to the computing service.}
-\item{The Submitter identifies local input files from the JobDescription and uploads them to the computing service.}
+\item{The {\TargetGenerator} has prepared a list of {\ExecutionTarget}s. Depending on the URLs provided to the {\TargetGenerator} the list of 
+found {\ExecutionTarget}s may be empty or contain several targets. Targets may even represent more than one grid flavour. The list of 
+found targets are given as input to the {\Broker}.}
+\item{In order to rank the found services ({\ExecutionTarget}s) the {\Broker} needs detailed knowledge about the job requirements, thus the 
+{\JobDescription} is passed as input to the brokering process.}
+\item{The {\Broker} outputs a ordered list of {\ExecutionTarget}s according to the provided {\JobDescription}.}
+\item{Each ExectionTarget has a method to return a specialized {\Submitter} which is capable of submitting jobs to the service it 
+represents. The best suitable {\ExecutionTarget} for the job is asked to return a {\Submitter} for job submission.}
+\item{The {\Submitter} takes the {\JobDescription} as input and uploads it to the computing service.}
+\item{The {\Submitter} identifies local input files from the {\JobDescription} and uploads them to the computing service.}
 \end{enumerate}
@@
 \section{Generic Classes}
-\subsection{TargetGenerator} The TargetGenerator is responsible for loading TargetRetriever plugins in accordance with the 
-input URLs. e.g. if an URL pointing to an ARC1 resource is given the TargetRetrieverARC1 is loaded. When queried for resources 
-(``targets'') the TargetGenerator passes the request to the loaded TargetRetrievers running them as individual threads 
-for improved performance. The TargetGenerator keeps records of by TargetRetriever found servers and services 
-(index or computing) in order to avoid multiple queries. Accepted targets are stored in the FoundTargets array kept by 
-the TargetGenerator.
+\subsection{{\TargetGenerator}} The {\TargetGenerator} class is the main class for resource discovery and information retrieval. It loads 
+{\TargetRetriever} plugins in accordance with the input URLs using the ARC Loader (e.g. if an URL pointing to an ARC1 resource is given 
+the TargetRetrieverARC1 is loaded).
 
-\subsection{TargetRetriever} The TargetRetriever is base class for TargetRetriever grid middleware specializations. Contains the pure 
+To perform a resource discovery, first construct a {\TargetGenerator} object using the {\TargetGenerator} constructor,
+
+\begin{shaded}
+\begin{verbatim}
+TargetGenerator(const std::list<std::string>& clusterselect,
+                const std::list<std::string>& clusterreject,
+                const std::list<std::string>& indexurls);
+\end{verbatim}
+\end{shaded}
+
+where \texttt{clusterselect}, \texttt{clusterreject} and \texttt{indexurls} are lists of strings where each string is either 
+an URL or a regular string defined as an alias in the user (or system) configuration. Without any further explanation it is 
+hard to decode from an URL if it points to a index server or computing service of a certain grid flavour. Thus in order for 
+the {\TargetGenerator} to know which {\TargetRetriever} should be loaded URLs have to be given in a format where the grid 
+flavour preceeds the actual URL, e.g:
+
+\begin{shaded}
+\begin{small}
+\begin{verbatim}
+ARC0:ldap://grid.tsl.uu.se:2135/mds-vo-name=sweden,O=grid
+ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid
+\end{verbatim}
+\end{small}
+\end{shaded}
+
+where the former is the URL to an index server while the latter is an URL to a computing service.
+
+To prepare a list of {\ExecutionTarget}s use the {\TargetGenerator} object and invoke its method
+
+\begin{shaded}
+\begin{verbatim}
+GetTargets(int targetType, int detailLevel);
+\end{verbatim}
+\end{shaded}
+
+The {\TargetGenerator} will pass this request to the loaded {\TargetRetriever}s running them as individual threads 
+for improved performance. The {\TargetGenerator} keeps records of by {\TargetRetriever} found index servers and computing 
+services in order to avoid multiple identical queries. Accepted {\ExecutionTarget}s are stored in the FoundTargets 
+array kept by the {\TargetGenerator}. See also Section~\ref{sec:TargetDiscovery} for a schematic drawing of the 
+resource discovery process.
+
+Information about the found {\ExecutionTarget}s can be printed by
+
+\begin{shaded}
+\begin{verbatim}
+PrintTargetInfo(bool longlist) const;
+\end{verbatim}
+\end{shaded}
+
+\subsection{{\TargetRetriever}} The {\TargetRetriever} is base class for {\TargetRetriever} grid middleware specializations. Contains the pure 
 virtual method GetTargets which is to be implemented by the specialized class. While it is not mandatory it is recommended 
@@
 responsible for querying index servers, and if the result yields an addresss to a different index server then the 
-specialized TargetRetriever should call itself recursively with the found url. Discovered computing services are sent 
-to the InterrogateTarget function (or if the TargetRetriever was initialized with a computing service url then it should 
+specialized {\TargetRetriever} should call itself recursively with the found url. Discovered computing services are sent 
+to the InterrogateTarget function (or if the {\TargetRetriever} was initialized with a computing service url then it should 
 jump directly to this point) which collects the detailed information about the service.
 
-\subsection{ExecutionTarget} The ExecutionTarget is the class representation of a computing resource (queue) capable of running a grid job. It
-serves as input to the broker which is foreseen to be able to select between different ExecutionTargets from different 
-grid flavours without apriori knowing their difference. The ExecutionTarget class mimics the glue2 information model (with 
+\subsection{{\ExecutionTarget}} The {\ExecutionTarget} is the class representation of a computing resource (queue) capable of running a grid job. It
+serves as input to the broker which is foreseen to be able to select between different {\ExecutionTarget}s from different 
+grid flavours without apriori knowing their difference. The {\ExecutionTarget} class mimics the glue2 information model (with 
 a flattened structure), and thus a mapping between attributes from other information systems into the glue2 format is needed. 
 Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares. All information can be printed by 
-invoking the ExecutionTarget's ``Stat'' command which prints all variables (if they have been filled during the information 
+invoking the {\ExecutionTarget}'s ``Stat'' command which prints all variables (if they have been filled during the information 
 retrieval process).
@@
 Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
-selected ExecutionTarget, the ExecutionTarget is capable of returning a submitter (see next section) capable of submitting a 
+selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of returning a submitter (see next section) capable of submitting a 
 job to the resource which is represents.
 
-\subsection{Broker} is not yet implemented and thus not described here.
+\subsection{{\Broker}} is not yet implemented and thus not described here.
 
-\subsection{JobDescription} The JobDescription is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
+\subsection{{\JobDescription}} The {\JobDescription} is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
 JSDL job description languages and to some extent also translate between them. Validation (syntax and semantical) is performed 
@@
 
-\subsection{Submitter} Submitter is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
+\subsection{{\Submitter}} {\Submitter} is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
 uploads (by the job needed) local input files. Returns an XML node containing all needed information about the job 
-for later job management. The Submitter is returned by the ExecutionTarget selected for job execution and thus the 
-ExecutionTarget populates (through XML config element) the Submitter with information about submission endpoint (URL) 
+for later job management. The {\Submitter} is returned by the {\ExecutionTarget} selected for job execution and thus the 
+{\ExecutionTarget} populates (through XML config element) the {\Submitter} with information about submission endpoint (URL) 
 and job description languages ``spoken'' by the target.
 
-\subsection{JobSupervisor} is responsible for loading the appropriate JobController for managing a job running on a specific 
+\subsection{{\JobSupervisor}} is responsible for loading the appropriate {\JobController} for managing a job running on a specific 
 grid flavour. In order to match a job identification string (jobid) with a grid flavour a lookup in the local xml file is 
-performed. The JobSupervisor main functionality is to return a list of JobControllers which are capable of managing all jobs 
+performed. The {\JobSupervisor} main functionality is to return a list of {\JobController}s which are capable of managing all jobs 
 addressed by a certain command.
 
-\subsection{JobController} is both base class for grid specific specializations, but also the implementor of all public functionality 
-offered by the JobControllers. In other words all virtual functions of the JobController are private. Upon initialization the 
-JobController converts the information stored in the local XML to Jobs which are stored in a JobControler internal store called JobsStore. 
-The public functions of the JobController offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations 
+\subsection{{\JobController}} is both base class for grid specific specializations, but also the implementor of all public functionality 
+offered by the {\JobController}s. In other words all virtual functions of the {\JobController} are private. Upon initialization the 
+{\JobController} converts the information stored in the local XML to Jobs which are stored in a JobControler internal store called JobsStore. 
+The public functions of the {\JobController} offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations 
 for issueing the command. This ensures consistency in how the commands are handled between the different grid flavours. Jobs that are 
-permanently removed following a command are removed from both the JobStore and local xml file. Thus it is foreseen that the JobControllers 
+permanently removed following a command are removed from both the JobStore and local xml file. Thus it is foreseen that the {\JobController}s 
 may be used in a future graphical user interface where it becomes a persistent object continously monitoring jobs.
 
-\subsection{Job} is generic job class for storing all job related information. Attributes are derived from the glue2 information model 
+\subsection{{\Job}} is generic job class for storing all job related information. Attributes are derived from the glue2 information model 
 and thus a mapping is needed for non glue2 compliant grid middlewares. Appendix XX shows the present mapping schema.

### 2008-09-02 doc/storage_system/arc1-storage-documentation.tex
renaming httpd --> hopi

@@
 The Shepherd could communicate with the file transfer services via backend modules.
-Currently there are two kinds of backend modules, one for the \emph{byteio} service (which is a simple implementation of a subset of the ByteIO interface) and one for the \emph{httpd} service (which is a simple HED-based HTTP server).
+Currently there are two kinds of backend modules, one for the \emph{byteio} service (which is a simple implementation of a subset of the ByteIO interface) and one for the \emph{hopi} service (which is a simple HED-based HTTP server).
 
@@
 
-Let's see the scenario for the httpd service which should be in a special `slave' mode for this kind of operation: if a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), then the Shepherd service creates a hardlink into the transfer directory (e.g.~\verb!./transfer/abc!) and sets this file read-only.
-If the httpd service is configured that way that it handles the HTTP path \verb!/prb!
+Let's see the scenario for the hopi service which should be in a special `slave' mode for this kind of operation: if a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), then the Shepherd service creates a hardlink into the transfer directory (e.g.~\verb!./transfer/abc!) and sets this file read-only.
+If the hopi service is configured that way that it handles the HTTP path \verb!/prb!
 and it is serving files from the directory \verb!./transfer!
@@
 Then the client \verb!GET!s this URL and gets the file.
-The httpd service removes (unlinks) this file immediately after the \verb!GET!
+The hopi service removes (unlinks) this file immediately after the \verb!GET!
 request arrived, which makes this \verb!http://localhost:60000/prb/abc!
@@
 to this URL.
-When the client \verb!PUT!s the file there, the httpd service immediately removes the uploaded file from the transfer directory, but because it has a hardlink in the store directory, the file is stored there as \verb!./store/file2!.
-The backend module for the httpd service periodically checks whether a new file has two or just one hard links.
+When the client \verb!PUT!s the file there, the hopi service immediately removes the uploaded file from the transfer directory, but because it has a hardlink in the store directory, the file is stored there as \verb!./store/file2!.
+The backend module for the hopi service periodically checks whether a new file has two or just one hard links.
 If it has only one that means that a file is uploaded, so it could notify the Shepherd that the file is arrived.
@@
 - done in 0.08 seconds.
-- Got transfer URL: http://localhost:60000/httpd/d15900f5-34ee-4bba-bb10-73d60d1c0d75
+- Got transfer URL: http://localhost:60000/hopi/d15900f5-34ee-4bba-bb10-73d60d1c0d75
 - Uploading from 'testfile'
-    to 'http://localhost:60000/httpd/d15900f5-34ee-4bba-bb10-73d60d1c0d75' with http...
+    to 'http://localhost:60000/hopi/d15900f5-34ee-4bba-bb10-73d60d1c0d75' with http...
 Uploading 20 bytes... data sent, waiting... done.
@@
 - done in 0.05 seconds.
-- Got transfer URL: http://localhost:60000/httpd/dab911d0-110f-468e-b0c3-627af6e3af31
-- Downloading from 'http://localhost:60000/httpd/dab911d0-110f-468e-b0c3-627af6e3af31'
+- Got transfer URL: http://localhost:60000/hopi/dab911d0-110f-468e-b0c3-627af6e3af31
+- Downloading from 'http://localhost:60000/hopi/dab911d0-110f-468e-b0c3-627af6e3af31'
     to 'newfile' with http...

### 2008-09-02 src/services/hopi/hopi.cpp

@@ -14,37 +14,37 @@
 #include <arc/message/MessageAttributes.h>
 
 
-#include "httpd.h"
+#include "hopi.h"
 #include "PayloadFile.h"
 
-namespace HTTPD {
+namespace Hopi {
 
 static Arc::Service *get_service(Arc::Config *cfg, Arc::ChainContext*)
 {
-    return new HTTPD(cfg);
+    return new Hopi(cfg);
 }
 
-HTTPD::HTTPD(Arc::Config *cfg):Service(cfg), logger(Arc::Logger::rootLogger, "HTTPD")
+Hopi::Hopi(Arc::Config *cfg):Service(cfg), logger(Arc::Logger::rootLogger, "Hopi")
 {
-    logger.msg(Arc::INFO, "HTTPD Initialized"); 
+    logger.msg(Arc::INFO, "Hopi Initialized"); 
     doc_root = (std::string)((*cfg)["DocumentRoot"]);
     if (doc_root.empty() == 1) {
         doc_root = "./";
     }
-    logger.msg(Arc::INFO, "HTTPD DocumentRoot is " + doc_root);
+    logger.msg(Arc::INFO, "Hopi DocumentRoot is " + doc_root);
     slave_mode = (std::string)((*cfg)["SlaveMode"]);
     if (slave_mode.empty() == 1) {
         slave_mode = "0";
     }  
-    if (slave_mode == "1") logger.msg(Arc::INFO, "HTTPD SlaveMode is on!");
+    if (slave_mode == "1") logger.msg(Arc::INFO, "Hopi SlaveMode is on!");
 }
 
-HTTPD::~HTTPD(void)
+Hopi::~Hopi(void)
 {
-    logger.msg(Arc::INFO, "HTTPD shutdown");
+    logger.msg(Arc::INFO, "Hopi shutdown");
 }
 
-Arc::PayloadRawInterface *HTTPD::Get(const std::string &path, const std::string &base_url)
+Arc::PayloadRawInterface *Hopi::Get(const std::string &path, const std::string &base_url)
 {
     // XXX eliminate relativ paths first
     std::string full_path = Glib::build_filename(doc_root, path);
@@ -75,13 +75,13 @@ Arc::PayloadRawInterface *HTTPD::Get(const std::string &path, const std::string 
     return NULL;
 }
 
-Arc::MCC_Status HTTPD::Put(const std::string &path, Arc::PayloadRawInterface &buf)
[…]

### 2008-09-02 src/services/hopi/hopi.h

@@ -1,14 +1,14 @@
-#ifndef __ARC_HTTPD_H__
-#define __ARC_HTTPD_H__
+#ifndef __ARC_Hopi_H__
+#define __ARC_Hopi_H__
 
 #include <arc/message/Service.h>
 #include <arc/message/PayloadRaw.h>
 #include <arc/Logger.h>
 #include <string>
 
-namespace HTTPD {
+namespace Hopi {
 
-class HTTPD: public Arc::Service {
+class Hopi: public Arc::Service {
     protected:
         Arc::Logger logger;
         std::string doc_root;
@@ -16,11 +16,11 @@ class HTTPD: public Arc::Service {
         Arc::PayloadRawInterface* Get(const std::string &path, const std::string &base_url);
         Arc::MCC_Status Put(const std::string &path, Arc::PayloadRawInterface &buf);
     public:
-        HTTPD(Arc::Config *cfg);
-        virtual ~HTTPD(void);
+        Hopi(Arc::Config *cfg);
+        virtual ~Hopi(void);
         virtual Arc::MCC_Status process(Arc::Message &inmsg, Arc::Message &outmsg);
 };
 
-} // namespace HTTPD
+} // namespace Hopi
 
-#endif // __ARC_HTTPD_H__
+#endif // __ARC_Hopi_H__

### 2008-09-02 doc/storage_system/arc1-storage-documentation.tex
hopi -> Hopi (uppercase)

@@
 The Shepherd could communicate with the file transfer services via backend modules.
-Currently there are two kinds of backend modules, one for the \emph{byteio} service (which is a simple implementation of a subset of the ByteIO interface) and one for the \emph{hopi} service (which is a simple HED-based HTTP server).
+Currently there are two kinds of backend modules, one for the \emph{byteio} service (which is a simple implementation of a subset of the ByteIO interface) and one for the \emph{Hopi} service (which is a simple HED-based HTTP server).
 
@@
 
-Let's see the scenario for the hopi service which should be in a special `slave' mode for this kind of operation: if a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), then the Shepherd service creates a hardlink into the transfer directory (e.g.~\verb!./transfer/abc!) and sets this file read-only.
-If the hopi service is configured that way that it handles the HTTP path \verb!/prb!
+Let's see the scenario for the Hopi service which should be in a special `slave' mode for this kind of operation: if a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), then the Shepherd service creates a hardlink into the transfer directory (e.g.~\verb!./transfer/abc!) and sets this file read-only.
+If the Hopi service is configured that way that it handles the HTTP path \verb!/prb!
 and it is serving files from the directory \verb!./transfer!
@@
 Then the client \verb!GET!s this URL and gets the file.
-The hopi service removes (unlinks) this file immediately after the \verb!GET!
+The Hopi service removes (unlinks) this file immediately after the \verb!GET!
 request arrived, which makes this \verb!http://localhost:60000/prb/abc!
@@
 to this URL.
-When the client \verb!PUT!s the file there, the hopi service immediately removes the uploaded file from the transfer directory, but because it has a hardlink in the store directory, the file is stored there as \verb!./store/file2!.
-The backend module for the hopi service periodically checks whether a new file has two or just one hard links.
+When the client \verb!PUT!s the file there, the Hopi service immediately removes the uploaded file from the transfer directory, but because it has a hardlink in the store directory, the file is stored there as \verb!./store/file2!.
+The backend module for the Hopi service periodically checks whether a new file has two or just one hard links.
 If it has only one that means that a file is uploaded, so it could notify the Shepherd that the file is arrived.

### 2008-09-03 src/services/a-rex/infoproviders/CEinfo.pl
rename cluster+qju.pl to CEinfo.pl

@@ -1,0 +1,317 @@
+#!/usr/bin/perl -w
+
+package ClusterInfo;
+
+use File::Basename;
+use lib dirname($0);
+
+use Getopt::Long;
+use XML::Simple;
+use Data::Dumper;
+
+########################################################
+# Driver for information collection
+# Reads old style arc.config and prints out XML
+########################################################
+
+use base "InfoCollector";
+
+use HostInfo;
+use GMJobsInfo;
+use LRMSInfo;
+
+use ARC0ClusterInfo;
+use ARC1ClusterInfo;
+
+use ConfigParser;
+use LogUtils; 
+
+use strict;
+
+our $log = LogUtils->getLogger(__PACKAGE__);
+
+
+sub main {
+
+    LogUtils->getLogger()->level($LogUtils::INFO);
+
+    ########################################################
+    # Config defaults
+    ########################################################
+    
+    my %config = (
+                   ttl            =>  600,
+                   gm_mount_point => "/jobs",
+                   gm_port        => 2811,
+                   homogeneity    => "TRUE",
+                   providerlog    => "/var/log/grid/infoprovider.log",
+                   defaultttl     => "604800",
+                   x509_user_cert => "/etc/grid-security/hostcert.pem",
+                   x509_cert_dir  => "/etc/grid-security/certificates/",
+                   ng_location    => $ENV{ARC_LOCATION} ||= "/usr/local",
+                   gridmap        => "/etc/grid-security/grid-mapfile",
+                   processes      => [qw(arched)]
+    
+    );
+
+
+    ########################################################
+    # Parse command line options
+    ########################################################
+    
+    my $configfile;
+    my $print_help;
+    GetOptions("config:s" => \$configfile,
+               "help|h"   => \$print_help ); 
+        
+    if ($print_help) { 
+        print "Usage: $0 <options>
+        --config   - location of arc.conf
+        --help     - this help\n";
+        exit 1;
+    }
+    
+    unless ( $configfile ) {
+        $log->error("a command line argument is missing, see --help ") and die;
+    }
+    
+    ##################################################
+    # Read ARC configuration
[…]

### 2008-09-03 doc/client/client_technical.tex
more text

@@
 \newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
+\newcommand{\ACC}{\texttt{ACC}}
 \newcommand{\TargetGenerator}{\texttt{TargetGenerator}}
@@
 \newcommand{\Job}{\texttt{Job}}
+\newcommand{\UserConfig}{\texttt{UserConfig}}
+\newcommand{\ClientInterface}{\texttt{ClientInterface}}
 \definecolor{shadecolor}{rgb}{1,1,0.6}
@@
 \section{Generic Classes}
+\subsection{{\ACC}}
+
 \subsection{{\TargetGenerator}} The {\TargetGenerator} class is the main class for resource discovery and information retrieval. It loads 
@@
 
-\subsection{{\TargetRetriever}} The {\TargetRetriever} is base class for {\TargetRetriever} grid middleware specializations. Contains the pure 
-virtual method GetTargets which is to be implemented by the specialized class. While it is not mandatory it is recommended 
-that the specialized class divides this function into two components: QueryIndex and InterrogateTarget. The former is 
-responsible for querying index servers, and if the result yields an addresss to a different index server then the 
-specialized {\TargetRetriever} should call itself recursively with the found url. Discovered computing services are sent 
-to the InterrogateTarget function (or if the {\TargetRetriever} was initialized with a computing service url then it should 
-jump directly to this point) which collects the detailed information about the service.
+\subsection{{\TargetRetriever}} The {\TargetRetriever} is base class for {\TargetRetriever} grid middleware specializations and 
+inherits from the \texttt{ACC} base class in order to be loadable by the ARC \texttt{Loader}. It is designed to work on conjunction 
+with the {\TargetGenerator} contains and contains the pure virtual method
 
-\subsection{{\ExecutionTarget}} The {\ExecutionTarget} is the class representation of a computing resource (queue) capable of running a grid job. It
-serves as input to the broker which is foreseen to be able to select between different {\ExecutionTarget}s from different 
-grid flavours without apriori knowing their difference. The {\ExecutionTarget} class mimics the glue2 information model (with 
-a flattened structure), and thus a mapping between attributes from other information systems into the glue2 format is needed. 
-Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares. All information can be printed by 
-invoking the {\ExecutionTarget}'s ``Stat'' command which prints all variables (if they have been filled during the information 
-retrieval process).
+\begin{shaded}
+\begin{verbatim}
+virtual void GetTargets(TargetGenerator& mom, int targetType,
+                        int detailLevel) = 0;
+\end{verbatim}
+\end{shaded}
 
-Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
-selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of returning a submitter (see next section) capable of submitting a 
-job to the resource which is represents.
+which is to be implemented by the specialized class. While it is not mandatory it is recommended that the specialized class 
+divides this method into two components: QueryIndex and InterrogateTarget. The former is handles index server queries, and the 
+latter the computing service queries and {\ExecutionTarget} preparation.
 
-\subsection{{\Broker}} is not yet implemented and thus not described here.
+If an index server query yields an url to a different index server then the one queried, then the {\TargetRetriever} should 
+call itself recursively creating a new {\TargetRetriever} for the discovered index server.
 
-\subsection{{\JobDescription}} The {\JobDescription} is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
-JSDL job description languages and to some extent also translate between them. Validation (syntax and semantical) is performed 
-automatically when the source is set. Features easy retrieval of list of local input files which should be uploaded to the 
-computing service.
+\subsection{{\ExecutionTarget}} The {\ExecutionTarget} is the class representation of a computing resource (queue) capable of 
+executing a grid job. It serves as input to the {\Broker} which is foreseen to be able to select between different 
+{\ExecutionTarget}s from different grid flavours without apriori knowing their difference. The {\ExecutionTarget} class 
+mimics the glue2 information model (with a flattened structure), and thus a mapping between attributes from other information 
+systems into the glue2 format is needed.  Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares.
 
-\subsection{{\Submitter}} {\Submitter} is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
-uploads (by the job needed) local input files. Returns an XML node containing all needed information about the job 
-for later job management. The {\Submitter} is returned by the {\ExecutionTarget} selected for job execution and thus the 
-{\ExecutionTarget} populates (through XML config element) the {\Submitter} with information about submission endpoint (URL) 
-and job description languages ``spoken'' by the target.
+All attributes of the {\ExecutionTarget} can be printed by the method
 
-\subsection{{\JobSupervisor}} is responsible for loading the appropriate {\JobController} for managing a job running on a specific 
-grid flavour. In order to match a job identification string (jobid) with a grid flavour a lookup in the local xml file is 
-performed. The {\JobSupervisor} main functionality is to return a list of {\JobController}s which are capable of managing all jobs 
-addressed by a certain command.
+\begin{shaded}
+\begin{verbatim}                                                                                                                            
+void Print(bool longlist) const;
+\end{verbatim}
+\end{shaded}
 
-\subsection{{\JobController}} is both base class for grid specific specializations, but also the implementor of all public functionality 
-offered by the {\JobController}s. In other words all virtual functions of the {\JobController} are private. Upon initialization the 
-{\JobController} converts the information stored in the local XML to Jobs which are stored in a JobControler internal store called JobsStore. 
-The public functions of the {\JobController} offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations 
-for issueing the command. This ensures consistency in how the commands are handled between the different grid flavours. Jobs that are 
-permanently removed following a command are removed from both the JobStore and local xml file. Thus it is foreseen that the {\JobController}s 
-may be used in a future graphical user interface where it becomes a persistent object continously monitoring jobs.
+Following a broker decision jobs are to be submitted. Since all information about the selected computing service resides within the 
+selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of returning a {\Submitter} capable of submitting a 
+job to the service which is represents.
 
-\subsection{{\Job}} is generic job class for storing all job related information. Attributes are derived from the glue2 information model 
-and thus a mapping is needed for non glue2 compliant grid middlewares. Appendix XX shows the present mapping schema.
+\begin{shaded}
+\begin{verbatim}
+Submitter *GetSubmitter() const;
+\end{verbatim}
+\end{shaded}
 
-\subsection{UserConfig}
+\subsection{{\Broker}} The {\Broker} is not yet implemented and thus not described here.
+
+\subsection{{\JobDescription}} When addressing interoperability it is of paramount importance to transparently address grid job 
+descriptions written in different job description languages by translating them automatically. In the ARClib library this functionality 
+is implemented in the {\JobDescription} class. This is a generic class that takes a job description (string )as input
+
+\begin{shaded}
+\begin{verbatim}
+setSource( const std::string source ) throw(JobDescriptionError);
+\end{verbatim}
+\end{shaded}
+
+in any supported formats (currently xRSL, JDL, JSDL), converts and stores it in a JSDL-like internal job description format. The identification 
+of the source's description format is internally handled by a \texttt{JobDescriptionOrderer} class which identifies the format by pattern
+matching.
+
+Different operations, like getting the job description in other formats or getting job-related information, can be performed using the 
+description-independent functions of this class, e.g.:
+
+\begin{shaded}
+\begin{verbatim}
+getProduct(std::string& product, std::string format="JSDL") throw(JobDescriptionError);
+\end{verbatim}
+\end{shaded}
+
+The {\JobDescription} class has tree back-end classes, sometimes referred to as back-end or translator modules, corresponding to the three 
+supported job description languages. The {\JobDescription} class chooses the appropriate back-end module according to the pattern match 
+performed by the \texttt{JobDescriptionOrderer}, and uses the back-end module for parsing and generating the job descriptions.
+
+\subsection{{\Submitter}} {\Submitter} is base class for grid specific specializations (plugin). It submits job(s) to the service 
+it represents and uploads (by the job needed) local input files.
+
+\begin{shaded}
+\begin{verbatim}
+virtual bool Submit(JobDescription& jobdesc, XMLNode& info) = 0;
+\end{verbatim}
+\end{shaded}
+
+The \texttt{Submit} method fills the XML node \texttt{info} with all needed information about the job for later job management. 
+The {\Submitter} is returned by the {\ExecutionTarget} selected for job execution and thus the {\ExecutionTarget} populates 
+(through XML config element) the {\Submitter} with information about submission endpoint (URL) and job description languages 
+``spoken'' by the target.
+
+\subsection{{\JobSupervisor}} The {\JobSupervisor} is responsible for loading the appropriate {\JobController}s for managing 
+jobs running on a certain grid flavour. Job manipulation can be performed either on individual jobs or on groups of jobs (e.g. 
+all jobs running on certain cluster or all jobs with job state ``FINISHED''), and in order to translate the by user given 
+information into a set of loadable {\JobController}s the {\JobSupervisor} makes extensive use of the local XML file housing 
+information about all active jobs. Thus the {\JobSupervisor} constructor becomes
+
+\begin{shaded}
+\begin{verbatim}
+JobSupervisor(const std::list<std::string>& jobs,
+              const std::list<std::string>& clusterselect,
+              const std::list<std::string>& clusterreject,
+              const std::string joblist);
+\end{verbatim}
+\end{shaded}
+
+where \texttt{jobs} is a list of job identification strings (jobids), \texttt{clusterselect} is a list of selected computing services, 
+\texttt{clusterreject} is a list of rejected computing services and \texttt{joblist} is a string pointing to the local XML file 
+with job information. Syntax for \texttt{clusterselect} and \texttt{clusterreject} is identical to the one used by the {\TargetGenerator}
+for for internal consistency.
+
+Although being loaded by the {\JobSupervisor} the {\JobController} objects truly resides within the ARC Loader which is a member 
+of the {\JobSupervisor} class. In order to get handles on the {\JobController}s the inline method
+
+\begin{shaded}
+\begin{verbatim}
+std::list<Arc::JobController*> GetJobControllers(){return JobControllers;}
+\end{verbatim}
+\end{shaded}
+
+returns a list of pointers to the loaded {\JobController}s.
+
+\subsection{{\JobController}} The {\JobController} is both base class for grid specific specializations, but also the implementor of all 
+public functionality offered by the {\JobController}s. In other words all virtual functions of the {\JobController} are private. The 
+initialization of a (specialized) {\JobController} object takes two steps. First the constructor has to invoked
+
+\begin{shaded}
+\begin{verbatim}
+JobController(Arc::Config *cfg, std::string flavour);
+\end{verbatim}
+\end{shaded}
+
+This is usually done by the ARC Loader which sees to that the {\JobController} receives information about its flavour (grid) and 
+the local XML file containing information about all active jobs (flavour independent). Next step is the filling of the {\JobController}'s
+job pool \texttt{JobStore} which is the pool of jobs that the {\JobController} can manage.
+
+\begin{shaded}
+\begin{verbatim}
+void FillJobStore(std::list<std::string> jobs,
+                  std::list<std::string> clusterselect,
+                  std::list<std::string> cluterreject);
+\end{verbatim}
+\end{shaded}
+
+Here \texttt{jobs}, \texttt{clusterselect} and \texttt{clusterreject} have the same interpretation as in the {\JobSupervisor} constructor. 
+However, while the latter only had to identify which {\JobController}s should be loaded, the {\JobController} itself has to decode the 
+supplied lists into a list of jobs with non-overlapping entries which will be filled into the JobStore. Due to possible conflicts 
+within the supplied lists, rules have to be set up for dealing with contradictions:
+
+\begin{enumerate}
+\item{If the \texttt{jobs} list has entries, fill \texttt{JobStore} with the by user requested jobs.}
+\item{If the \texttt{clusterselect} list has entries, fill \texttt{JobStore} with the jobs running on the selected clusters.}
+\item{If clusters are rejected and \texttt{JobStore} has entries, remove from \texttt{JobStore} all jobs running on rejected clusters.}
+\item{If \texttt{jobs} and \texttt{clusterselect} are both empty lists, fill \texttt{JobStore} with all jobs except those 
+running on possible rejected clusters.}
+\end{enumerate}
+
+The steps above completes the initialization of the {\JobController} which is now ready for handling jobs. The public functions of the 
+{\JobController} offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations for issueing the command. 
+Here examplified by the \texttt{Stat} command:
+
+\begin{shaded}
+\begin{small}
+\begin{verbatim}
+void JobController::Stat(const std::list<std::string>& jobs,
+                    const std::list<std::string>& clusterselect,
+                    const std::list<std::string>& clusterreject,
+                    const std::list<std::string>& status,
+                    const bool longlist,
+                    const int timeout){
+
+  GetJobInformation();
+
+  //Add here functionality for identifying jobs according to clusterselect, clusterreject & status
+
+  for(std::list<Arc::Job>::iterator jobiter = JobStore.begin();jobiter!= JobStore.end();jobiter++){
+    if(jobiter->State.empty()){
+      std::cout<<Arc::IString("Job state information not found:%s",jobiter->JobID.str())<<std::endl;
+      Time now;
+      if(now - jobiter->LocalSubmissionTime < 90){
+        std::cout<<Arc::IString("This job was very recently "
+                                "submitted and might not yet" 
+                                "have reached the information-system")<<std::endl;
+      }
+      continue;
+    }
+    jobiter->Print(longlist);
+  }
+}
+\end{verbatim}
+\end{small}
+\end{shaded}
+
+The \texttt{Stat} command prints the job information to screen and in order to do so the {\JobController} has to query local information
+server for the latest status. Due to different protocols used for different grid flavours (e.g. ldap for production ARC), the 
+\texttt{GetJobInformation} has to be grid flavour specific and is only declared as a private virtual method within the {\JobController}
+base class. For details about the flavour specific implementations see Section~\ref{sec:plugins}.
+
+It should be noted that the \texttt{Stat} example above appears to have several unused variables (arguments). This is an artifact 
+caused by the ongoing development in which the {\JobController} is being prepared for usage in a future graphical user interface (GUI).
+In a GUI it would be desirable to make the {\JobController} a persistent object which exists throughout the lifetime of the GUI and 
+which is capable of monitoring and managing all jobs running on a certain grid flavour. In order to do so the \texttt{JobStore} should 
+become a pool of all jobs, and the public functions should take arguments which identifies which jobs from the pool should be subject 
+to the command. In the present implementation commands are performed on all jobs in \texttt{JobStore}.
+
+\subsection{{\Job}} The {\Job} is a generic job class for storing all job related information. Attributes are derived from the 
+glue2 information model and thus a mapping is needed for non glue2 compliant grid middlewares. Appendix XX shows the present 
+mapping schema.
+
+All attributes of the {\Job} can be printed by the method
+
+\begin{shaded}
+\begin{verbatim}                                                                                                                            
+void Print(bool longlist) const;
+\end{verbatim}
+\end{shaded}
+
+\subsection{{\UserConfig}} The {\UserConfig} class handles the client setup i.e. proxy, certificate and key location, user and system 
+configuration and local joblist location. Upon initialization (constructor) the {\UserConfig} locates the user files
+
+\begin{shaded}
+\begin{verbatim}                                                                                                                            
+$HOME/.arc/client.xml
+$HOME/.arc/jobs.xml
+\end{verbatim}
+\end{shaded}
+
+and if either of them is non-existing a default (empty) one is created.
+
+The {\UserConfig} has two main public methods
+
+\begin{shaded}
+\begin{verbatim}
+const std::string& JobsFile();
+const XMLNode ConfTree();
+\end{verbatim}
+\end{shaded}
+
+where \texttt{JobsFile()} returns the string pointing to the jobs.xml file while \texttt{ConfTree()}
+returns a configuration XML object which is the merge between the user and system configurations.
+In order to do this the method has to locate the system configuration and resolve possible conflicts 
+with the user configuration. This proceeds through the following chain of actions:
+
+\begin{enumerate}
+\item{Try reading system configuration from \texttt{<ARC Install Location>/etc/arcclient.xml}}
+\item{If the previous step failed try reading system configuration from \texttt{/etc/arcclient.xml}}
+\item{Merge system and user configuration by adding all system configuration not already listed 
+in the user configuration to the latter.}
+\item{Check if \texttt{\$X509\_USER\_PROXY} is set, if set add this as \texttt{ProxyPath} to configuration}
+\end{enumerate}
+
+\subsection{{\ClientInterface}}
 
@@
 \subsection{Logger}
+\subsection{XMLNode}
 

### 2008-09-04 doc/client/client_technical.tex
more text

@@
 All capabilities are enabled for three different grid flavours (Production ARC, ARC1 and gLite) 
-through the modular design using plugins specialized for each supported middleware. Future 
+through a modular design using plugins specialized for each supported middleware. Future 
 extensions to support additional middlewares involve plugin compilation only i.e. no recompilation 
@@
 \centering{{\scalebox{0.75}{\includegraphics{TargetDiscovery.png}}}
-\caption{\label{fig:ResDisc}Diagram depicting the process of resource discovery and information retrieval.} }
+\caption{\label{fig:ResDisc}Diagram depicting the resource discovery and information retrieval process} }
 \end{figure}
@@
 \centering{{\scalebox{0.75}{\includegraphics{JobSubmission.png}}}
-\caption{\label{fig:JobSub}Diagram depicting the process of submitting a job to a computing service.}}
+\caption{\label{fig:JobSub}Diagram depicting the submission of a job to a computing service.}}
 \end{figure}
@@
 \section{Job Management}
-Information about active jobs is stored in a local xml file which is used for later job management. As this file may contain 
-information about jobs running on completely different grid flavours, job management should be handled using plugins similar to 
-resource discovery and job submission. The following classes are used:
+Once a job is submitted job related information (job identification string, cluster etc) is stored in a local xml file which hosts 
+similar information for all active jobs. This file may contain jobs runnng on completely different grid flavour, and thus job 
+management should be handled using plugins similar to resource discovery and job submission. The job managing plugin is called 
+the {\JobController} and it is supported by the {\JobSupervisor} and {\Job} classes.
+
+\begin{figure}[ht]
+\centering{{\scalebox{0.75}{\includegraphics{JobManagement.png}}}
+\caption{\label{fig:JobMan}Diagram depicting how job controlling plugins, {\JobController}s, are loaded and initialized.}}
+\end{figure}
+
+Figure~\ref{fig:JobMan} shows how the three different classes work together and below a step by step description is given:
+
+\begin{enumerate}
+\item{The {\JobSupervisor} takes three arguments Jobs (job identification strings), ClusterSelect and ClusterReject.}
+\item{A job identification string does not uniguely define which grid flavour runs a certain job. Thus this information is stored 
+in joblist.xml upon submission by the {\Submitter} and joblist.xml is extensively used by the {\JobSupervisor} to identify the 
+{\JobController} flavours which are to be loaded.}
+\item{Suitable {\JobController}s are loaded}
+\item{Jobs, ClusterSelect and ClusterReject are passed to the {\JobController} which uses the information to 
+fill its internal JobStore.}
+\item{Residing within the {\JobSupervisor} the {\JobController}s are now accesible for external calls (i.e. job handling).}
+\end{enumerate}
 
@@
 \item{Check if \texttt{\$X509\_USER\_PROXY} is set, if set add this as \texttt{ProxyPath} to configuration}
+\item{If \texttt{\$X509\_USER\_PROXY} is not set and configuration tree does not already contain 
+a \texttt{ProxyPath} set \texttt{ProxyPath} to \texttt{/tmp/x509up\_u + userid}.}
+\item{Check if \texttt{\$X509\_USER\_CERT} is set, if set add this as \texttt{CertificatePath} to configuration}
+\item{If \texttt{\$X509\_USER\_CERT} is not set and configuration tree does not already contain 
+a \texttt{CertificatePath} set \texttt{CertificatePath} to \texttt{/etc/grid-security/hostcert.pem} 
+if user is superuser otherwise \texttt{\$HOME/.globus/usercert.pem}.}
+\item{Check if \texttt{\$X509\_USER\_KEY} is set, if set add this as \texttt{KeyPath} to configuration}
+\item{If \texttt{\$X509\_USER\_KEY} is not set and configuration tree does not already contain 
+a \texttt{KeyPath} set \texttt{KeyPath} to \texttt{/etc/grid-security/hostkey.pem} 
+if user is superuser otherwise \texttt{\$HOME/.globus/userkey.pem}.}
+\item{Check if \texttt{\$X509\_CERT\_DIR} is set, if set add this as \texttt{CACertificatesDir} to configuration}
+\item{If \texttt{\$X509\_CERT\_DIR} is not set and configuration tree does not already contain 
+a \texttt{CACertificatesDir} set \texttt{CACertificatesDir} to \texttt{/etc/grid-security/certificates}.}
 \end{enumerate}

### 2008-09-08 doc/client/client_technical.tex
Client plugin documentation

@@
 at the fingertips of users. While this documentation will illustrate how such command line tool can be 
-built, the main documention of the command line tools is given in the client user manual [ref].
+built, the main documentation of the command line tools is given in the client user manual [ref].
 
@@
 Once a job is submitted job related information (job identification string, cluster etc) is stored in a local xml file which hosts 
-similar information for all active jobs. This file may contain jobs runnng on completely different grid flavour, and thus job 
+similar information for all active jobs. This file may contain jobs running on completely different grid flavour, and thus job 
 management should be handled using plugins similar to resource discovery and job submission. The job managing plugin is called 
@@
 \item{The {\JobSupervisor} takes three arguments Jobs (job identification strings), ClusterSelect and ClusterReject.}
-\item{A job identification string does not uniguely define which grid flavour runs a certain job. Thus this information is stored 
+\item{A job identification string does not uniquely define which grid flavour runs a certain job. Thus this information is stored 
 in joblist.xml upon submission by the {\Submitter} and joblist.xml is extensively used by the {\JobSupervisor} to identify the 
@@
 fill its internal JobStore.}
-\item{Residing within the {\JobSupervisor} the {\JobController}s are now accesible for external calls (i.e. job handling).}
+\item{Residing within the {\JobSupervisor} the {\JobController}s are now accessible for external calls (i.e. job handling).}
 \end{enumerate}
@@
 the {\TargetGenerator} to know which {\TargetRetriever} should be loaded URLs have to be given in a format where the grid 
-flavour preceeds the actual URL, e.g:
+flavour precedes the actual URL, e.g:
 
@@
 The steps above completes the initialization of the {\JobController} which is now ready for handling jobs. The public functions of the 
-{\JobController} offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations for issueing the command. 
-Here examplified by the \texttt{Stat} command:
+{\JobController} offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations for issuing the command. 
+Here exemplified by the \texttt{Stat} command:
 
@@
 \subsection{ARC0 Plugins}
+
+The ARC0 plugins enables support for the interfaces used by computing
+elements running ARC version 0.x.
+
+The ARC 0.x local information system uses the globus GRIS with a
+propriety ARC schema. The information index server used is the globus
+GIIS. Both these servers are using the LDAP protocol. The
+specialization of the {\TargetRetriever} class for ARC0 is implemented
+using the LDAP DMC.
+
+Jobs running on an ARC 0.x computing element are handled by the ARC
+grid-manager. Job submission and job control are done using the
+gridftp protocol. The specializations of the {\Submitter} and
+{\JobController} classes use the globus ftp control library.
+
+Stage-in and stage-out of input and output files are also done using
+the gridftp protocol. This means that proper functionality of the ARC0
+plugins requires the gridftp DMC.
+
 \subsection{ARC1 Plugins}
+
+The computing element in ARC version 1.x is the A-Rex service running
+in a HED container.
+
+A-Rex implements the BES standard interface. Since this is a SOAP
+based interface the specializations of the {\TargetRetriever},
+{\Submitter} and {\JobController} classes make use of an MCC chain
+ending with the SOAP client MCC.
+
+The A-Rex service uses the https protocol put and get methods for
+stage-in and stage-out of input and output files. Therefore, the ARC1
+plugins requires the http DMC.
+
 \subsection{gLite Plugins}
+
+The CREAM plugin implements support for the CREAM interface to the
+gLite computing element.
+
+CREAM resources are registered in top level and site BDIIs. The CREAM
+specialization of the {\TargetRetriever} therefore makes use of the
+LDAP DMC.
+
+The CREAM interface is a propriety SOAP based interface. The CREAM
+specializations of the {\Submitter} and {\JobController} classes
+therefore use an MCC chain ending with the SOAP client MCC.
+
+Stage-in and stage-out of input and output files are done using the
+gridftp protocol. The gridftp DMC is therefore required.
 

### 2008-09-08 doc/client/client_technical.tex
more text

@@
 \hypersetup{
-  pdfauthor = {Author Name},
+  pdfauthor = {Paper author},
   pdftitle = {Paper title},
@@
 \newcommand{\ClientInterface}{\texttt{ClientInterface}}
+\newcommand{\URL}{\texttt{URL}}
+\newcommand{\XML}{\texttt{XML}}
+\newcommand{\Loader}{\texttt{Loader}}
+\newcommand{\ACCConfig}{\texttt{ACCConfig}}
+\newcommand{\Logger}{\texttt{Logger}}
+\newcommand{\XMLNode}{\texttt{XMLNode}}
 \definecolor{shadecolor}{rgb}{1,1,0.6}
@@
 
-{\raggedleft NORDUGRID-XXXXXXX-NN\par}
+{\raggedleft NORDUGRID-TECH-20\par}
 
@@
 %%%%---- The title ----
-{\centering \textsc{\Large Paper title}\Large \par}
+{\centering \textsc{\Large ARC1lib, A Client Library for ARC1}\Large \par}
 \vspace*{0.5cm}
@@
 %%%%---- A list of authors ----
-    {\centering \large Paper author\footnote{authors@address} \large \par}
-
+{\centering \large Mattias Ellert\footnote{mattias.ellert@fysast.uu.se} \large \par}
+{\centering \large Iv\'{a}n M\'{a}rton\footnote{martoni@niif.hu} \large \par}
+{\centering \large Bjarte Mohn\footnote{bjarte.mohn@fysast.uu.se} \large \par}
+{\centering \large P\'{e}ter Stef\'{a}n\footnote{stefan@niif.hu} \large \par}
 %%%%---- An abstract - if style is article ----
@@
 This document describes from a technical viewpoint a plugin based client library for the new 
-Web Service (WS) based Advanced Resource Connector (ARC) middleware. The library consists of 
+Web Service (WS) based Advanced Resource Connector~\cite{arc} ARC middleware. The library consists of 
 a set of C++ classes for
@@
 
-All capabilities are enabled for three different grid flavours (Production ARC, ARC1 and gLite) 
+All capabilities are enabled for three different grid flavours (Production ARC, ARC1 and gLite~\cite{glite}) 
 through a modular design using plugins specialized for each supported middleware. Future 
@@
 at the fingertips of users. While this documentation will illustrate how such command line tool can be 
-built, the main documentation of the command line tools is given in the client user manual [ref].
+built, the main documentation of the command line tools is given in the client user manual~\cite{ui}.
 
@@
 \label{sec:FuncOver}
-The new ARClib makes extensive use of plugins for command handling. These plugins are handled by a set 
+The new ARC1lib makes extensive use of plugins for command handling. These plugins are handled by a set 
 of higher level classes which thus are the ones to offer the plugin functionality to external calls. In 
@@
 With the increasing number of grid clusters around the world, a reliable and fast resource discovery and 
-information retrieval capability is of crucial importance for a user interface. The new ARClib resource 
+information retrieval capability is of crucial importance for a user interface. The new ARC1lib resource 
 discovery and information retrieval component consists of three classes; The {\TargetGenerator}, the 
@@
 \begin{enumerate}
-\item{The {\TargetGenerator} takes three URL arrays as input. First array contains individually selected computing services, while the 
+\item{The {\TargetGenerator} takes three {\URL} arrays as input. First array contains individually selected computing services, while the 
 second array contains individually selected index servers (1a). The last array contains a list of rejected services. This 
 list is parsed through alias resolve before being inserted as filter to the storage of found services (1b).}
-\item{The complete list of URLs pointing to computing services and index servers.}
-\item{For each URL a {\TargetRetriever} plugin is loaded using the ARC loader. The {\TargetRetriever} is initialized 
-with the URL and the information about this URL pointing to a computing service or an index server.}
+\item{The complete list of {\URL}s pointing to computing services and index servers.}
+\item{For each {\URL} a {\TargetRetriever} plugin is loaded using the ARC loader. The {\TargetRetriever} is initialized 
+with the {\URL} and the information about this {\URL} pointing to a computing service or an index server.}
 \item{An external call is received calling for targets to be prepared. Call is propagated to all loaded {\TargetRetriever}s.}
 \item{The call for targets is processed by each {\TargetRetriever} in parallel. If the {\TargetRetriever} is initialized with an 
-index server URL it first registers at the Index Servers store kept by the {\TargetGenerator}. If allowed to register, the index 
+index server {\URL} it first registers at the Index Servers store kept by the {\TargetGenerator}. If allowed to register, the index 
 server is queried and the query result processed.}
@@
 reason a {\TargetRetriever} processing the query result from an index server may encounter other index servers. In such
-events the {\TargetRetriever} will call itself recursively creating a new thread for the new URL. This ensures optimum speed and 
+events the {\TargetRetriever} will call itself recursively creating a new thread for the new {\URL}. This ensures optimum speed and 
 performance for resource discovery.}
@@
 
-The example above outlines how the {\TargetRetriever} works when initialized with an URL pointing to and index server. However, 
-as indicated it also accepts initialization with a service URLs, in which case the {\TargetRetriever} will try to register the 
+The example above outlines how the {\TargetRetriever} works when initialized with an {\URL} pointing to and index server. However, 
+as indicated it also accepts initialization with a service {\URL}s, in which case the {\TargetRetriever} will try to register the 
 service in the {\TargetGenerator} Found Services store. If allowed the {\TargetRetriever} will progress with the target interrogation 
@@
 \begin{enumerate}
-\item{The {\TargetGenerator} has prepared a list of {\ExecutionTarget}s. Depending on the URLs provided to the {\TargetGenerator} the list of 
+\item{The {\TargetGenerator} has prepared a list of {\ExecutionTarget}s. Depending on the {\URL}s provided to the {\TargetGenerator} the list of 
 found {\ExecutionTarget}s may be empty or contain several targets. Targets may even represent more than one grid flavour. The list of 
@@
 
-\section{Data Handling}
-\label{sec:DataHandling}
-
 \chapter{Implementation}
@@
 \subsection{{\ACC}}
+The Arc Client Component ({\ACC}) class is a completely empty base class needed for the {\Loader} 
+in order to create loadable classes.
+
+\begin{shaded}
+\begin{verbatim}
+class ACC {
+protected:
+  ACC() {}
+public:
+  virtual ~ACC() {}
+};
+\end{verbatim}
+\end{shaded}
 
 \subsection{{\TargetGenerator}} The {\TargetGenerator} class is the main class for resource discovery and information retrieval. It loads 
-{\TargetRetriever} plugins in accordance with the input URLs using the ARC Loader (e.g. if an URL pointing to an ARC1 resource is given 
-the TargetRetrieverARC1 is loaded).
+{\TargetRetriever} plugins in accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed} (e.g. if an {\URL} pointing to an ARC1 
+resource is given the TargetRetrieverARC1 is loaded).
 
@@
 where \texttt{clusterselect}, \texttt{clusterreject} and \texttt{indexurls} are lists of strings where each string is either 
-an URL or a regular string defined as an alias in the user (or system) configuration. Without any further explanation it is 
-hard to decode from an URL if it points to a index server or computing service of a certain grid flavour. Thus in order for 
-the {\TargetGenerator} to know which {\TargetRetriever} should be loaded URLs have to be given in a format where the grid 
-flavour precedes the actual URL, e.g:
+an {\URL} or a regular string defined as an alias in the user (or system) configuration. Without any further explanation it is 
+hard to decode from an {\URL} if it points to a index server or computing service of a certain grid flavour. Thus in order for 
+the {\TargetGenerator} to know which {\TargetRetriever} should be loaded {\URL}s have to be given in a format where the grid 
+flavour precedes the actual {\URL}, e.g:
 
@@
 
-where the former is the URL to an index server while the latter is an URL to a computing service.
+where the former is the {\URL} to an index server while the latter is an {\URL} to a computing service.
 
@@
 \subsection{{\JobDescription}} When addressing interoperability it is of paramount importance to transparently address grid job 
-descriptions written in different job description languages by translating them automatically. In the ARClib library this functionality 
-is implemented in the {\JobDescription} class. This is a generic class that takes a job description (string )as input
+descriptions written in different job description languages by translating them automatically. In the ARC1lib library this functionality 
+is implemented in the {\JobDescription} class. This is a generic class that takes a job description (string) as input
 
@@
 
-The \texttt{Submit} method fills the XML node \texttt{info} with all needed information about the job for later job management. 
+The \texttt{Submit} method fills the {\XML} node \texttt{info} with all needed information about the job for later job management. 
 The {\Submitter} is returned by the {\ExecutionTarget} selected for job execution and thus the {\ExecutionTarget} populates 
-(through XML config element) the {\Submitter} with information about submission endpoint (URL) and job description languages 
+(through {\XML} config element) the {\Submitter} with information about submission endpoint ({\URL}) and job description languages 
 ``spoken'' by the target.
@@
 all jobs running on certain cluster or all jobs with job state ``FINISHED''), and in order to translate the by user given 
-information into a set of loadable {\JobController}s the {\JobSupervisor} makes extensive use of the local XML file housing 
+information into a set of loadable {\JobController}s the {\JobSupervisor} makes extensive use of the local {\XML} file housing 
 information about all active jobs. Thus the {\JobSupervisor} constructor becomes
@@
 where \texttt{jobs} is a list of job identification strings (jobids), \texttt{clusterselect} is a list of selected computing services, 
-\texttt{clusterreject} is a list of rejected computing services and \texttt{joblist} is a string pointing to the local XML file 
+\texttt{clusterreject} is a list of rejected computing services and \texttt{joblist} is a string pointing to the local {\XML} file 
 with job information. Syntax for \texttt{clusterselect} and \texttt{clusterreject} is identical to the one used by the {\TargetGenerator}
@@
 
-Although being loaded by the {\JobSupervisor} the {\JobController} objects truly resides within the ARC Loader which is a member 
+Although being loaded by the {\JobSupervisor} the {\JobController} objects truly resides within the ARC {\Loader} which is a member 
 of the {\JobSupervisor} class. In order to get handles on the {\JobController}s the inline method
@@
 
-This is usually done by the ARC Loader which sees to that the {\JobController} receives information about its flavour (grid) and 
-the local XML file containing information about all active jobs (flavour independent). Next step is the filling of the {\JobController}'s
+This is usually done by the ARC {\Loader} which sees to that the {\JobController} receives information about its flavour (grid) and 
+the local {\XML} file containing information about all active jobs (flavour independent). Next step is the filling of the {\JobController}'s
 job pool \texttt{JobStore} which is the pool of jobs that the {\JobController} can manage.
@@
 where \texttt{JobsFile()} returns the string pointing to the jobs.xml file while \texttt{ConfTree()}
-returns a configuration XML object which is the merge between the user and system configurations.
+returns a configuration {\XML} object which is the merge between the user and system configurations.
 In order to do this the method has to locate the system configuration and resolve possible conflicts 
@@
 a \texttt{CertificatePath} set \texttt{CertificatePath} to \texttt{/etc/grid-security/hostcert.pem} 
-if user is superuser otherwise \texttt{\$HOME/.globus/usercert.pem}.}
+if user is superuser otherwise 
+\texttt{\$HOME/.globus/usercert.pem}.}
 \item{Check if \texttt{\$X509\_USER\_KEY} is set, if set add this as \texttt{KeyPath} to configuration}
@@
 
+\subsection{{\ACCConfig}}
+Locating Arc Client Components (plugins) is handled by the {\ACCConfig} class. It inherits from \texttt{BaseConfig} 
+and implements only one method
+
+\begin{shaded}
+\begin{verbatim}                                                                                                                            
+virtual XMLNode MakeConfig(XMLNode cfg) const;
+\end{verbatim}
+\end{shaded}
+
+The \texttt{MakeConfig} method searches plugin paths for all libraries named \texttt{libacc*}. Matching 
+libraries are added as plugins to the configuration object \texttt{cfg}.
+
 \subsection{{\ClientInterface}}
+\subsubsection{ClientTCP}
+\subsubsection{ClientHTTP}
+\subsubsection{ClientSOAP}
 
@@
 \subsection{ARC0 Plugins}
-
 The ARC0 plugins enables support for the interfaces used by computing
@@
 
-The ARC 0.x local information system uses the globus GRIS with a
-propriety ARC schema. The information index server used is the globus
-GIIS. Both these servers are using the LDAP protocol. The
+The ARC 0.x local information system uses the {\GT}~\cite{globus} GRIS with a
+propriety ARC schema. The information index server used is the {\GT}
+GIIS. Both these servers are using the LDAP~\cite{ldap} protocol. The
 specialization of the {\TargetRetriever} class for ARC0 is implemented
-using the LDAP DMC.
+using the ARC LDAP Data Management Component (DMC) (see \cite{hed} for 
+technical details).
 
 Jobs running on an ARC 0.x computing element are handled by the ARC
-grid-manager. Job submission and job control are done using the
-gridftp protocol. The specializations of the {\Submitter} and
+grid-manager~\cite{gm}. Job submission and job control are done using the
+gridftp~\cite{gridftp} protocol. The specializations of the {\Submitter} and
 {\JobController} classes use the globus ftp control library.
@@
 Stage-in and stage-out of input and output files are also done using
-the gridftp protocol. This means that proper functionality of the ARC0
+the gridftp~\cite{gridftp} protocol. This means that proper functionality of the ARC0
 plugins requires the gridftp DMC.
@@
 \subsection{ARC1 Plugins}
+The computing element in ARC version 1.x is the A-Rex~\cite{arex} service running
+in a HED~\cite{hed} container.
 
-The computing element in ARC version 1.x is the A-Rex service running
-in a HED container.
-
-A-Rex implements the BES standard interface. Since this is a SOAP
+A-Rex implements the BES~\cite{ogsa-bes} standard interface. Since this is a SOAP~\cite{soap}
 based interface the specializations of the {\TargetRetriever},
-{\Submitter} and {\JobController} classes make use of an MCC chain
-ending with the SOAP client MCC.
+{\Submitter} and {\JobController} classes make use of a chain of ARC Message Chain Components 
+(MCC~\cite{hed}) ending with the SOAP client MCC.
 
@@
 \subsection{gLite Plugins}
+The gLite computing element offers several interfaces, one of them being the Web Service based 
+computing element interface known as the CREAM CE~\cite{cream}. The current revision of this 
+interface (CREAM version 2) has been chosen for implementation within ARC1lib for the following 
+reasons:
 
-The CREAM plugin implements support for the CREAM interface to the
-gLite computing element.
+\begin{itemize}
+\item CREAM2 has a Web Service interface that fits the Web Service based ARC.
+\item CREAM2 enables direct access to the gLite computing element without having to go via the gLite workload management system.
+\item CREAM2 contains numerous improvements when compared to the earlier CREAM versions.
+\item CREAM2 supports direct job status queries.
+\item CREAM2 offers a convenient way of handling input and output files through accessing the input and output sandbox via GridFTP.
+\end{itemize}
 
-CREAM resources are registered in top level and site BDIIs. The CREAM
+gLite resources are registered in top level and site BDIIs. The CREAM
 specialization of the {\TargetRetriever} therefore makes use of the
-LDAP DMC.
+LDAP DMC similarly to the ARC0 plugins.
 
@@
 specializations of the {\Submitter} and {\JobController} classes
-therefore use an MCC chain ending with the SOAP client MCC.
+therefore use an MCC chain ending with the SOAP client MCC the same 
+way the ARC1 plugin does.
 
@@
 \section{Support Classes}
-\subsection{URL}
+\subsection{{\URL}}
 \subsection{Logger}

### 2008-09-08 doc/client/client_technical.tex
Getting there

@@
 \label{sec:Implementation}
-Clever text here.
+In this section an overview of all important classes in the new ARC1lib is given. The overview is subdivided in two 
+parts where first all the generic classes are presented, Section~\ref{sec:genclass}, before the grid flavour specializations 
+are presented in Section~\ref{sec:plugins}. Classes are described both in words and by code.
+
 \section{Generic Classes}
+\label{sec:genclass}
 \subsection{{\ACC}}
@@
 
-in any supported formats (currently xRSL, JDL, JSDL), converts and stores it in a JSDL-like internal job description format. The identification 
+in any supported formats (currently xRSL~\cite{xrsl}, JDL~\cite{jdl}, JSDL~\cite{jsdl}), converts and stores it in a JSDL-like internal job description format. The identification 
 of the source's description format is internally handled by a \texttt{JobDescriptionOrderer} class which identifies the format by pattern
@@
 a \texttt{CertificatePath} set \texttt{CertificatePath} to \texttt{/etc/grid-security/hostcert.pem} 
-if user is superuser otherwise 
+if user is superuser otherwise \newline
 \texttt{\$HOME/.globus/usercert.pem}.}
@@
 
-\section{Support Classes}
-\subsection{{\URL}}
-\subsection{Logger}
-\subsection{XMLNode}
-
 \chapter{Building Command Line Interfaces}
 \label{sec:cli}
+Given all components listed above it possible to write versatile command line interfaces (cli) for 
+grid job submission and management. ARC1lib offers the following native commands:
+
+\begin{enumerate}
+\item{\texttt{arcstat} - for computing resource or grid job status queries.}
+\item{\texttt{arcsub} - for grid job submission}
+\item{\texttt{arcget} - for downloading output of finished, cancelled or failed grid jobs.}
+\item{\texttt{arckill} - for terminating grid jobs.}
+\item{\texttt{arcclean} - for removing a grid job's session directory including all contents}
+\item{\texttt{arccat} - for performing the \texttt{cat} command to a running grid job's std out or std error file.}
+\end{enumerate}
+
+Each of the commands above are encoded within one C++ file with the following structure:
+
+\begin{shaded}
+\begin{verbatim}
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <arc/ArcLocation.h>
+#include <arc/IString.h>
+#include <arc/Logger.h>
+#include <arc/OptionParser.h>
+#include <arc/client/UserConfig.h>
+#include <arc/IString.h>
+
+void MyCommand(MyArgument1,
+               MyArgument2){
+
+  //Implementation of MyCommand
+
+}
+
+int main(int argc, char **argv){
+
+  setlocale(LC_ALL, "");
+
+  Arc::LogStream logcerr(std::cerr);
+  Arc::Logger::getRootLogger().addDestination(logcerr);
+  Arc::Logger::getRootLogger().setThreshold(Arc::WARNING);
+
+  Arc::ArcLocation::Init(argv[0]);
+
+  Arc::OptionParser options(istring("argument"), "", istring(
+                              "Clarifying text here));
+
+  //add command options
+  bool all = false;
+  options.AddOption('a', "all",
+		    istring("all jobs"),
+		    all);
+
+  std::list<std::string> params = options.Parse(argc, argv);
+
+  //call command
+  MyCommand(all, params);
+
+  return 0;
+
+}
+\end{verbatim}
+\end{shaded}
+
+The structure above has two separate blocks of code where the first block is the MyCommand method which is be 
+called by the main program (second block). Certainly this is not the only way of building (structuring) a 
+command line interface, but it has the benefit of a clear structure in addition to that the method implemented 
+can easily be reused in a different location without code duplication.
 

### 2008-09-08 doc/hed/ARCHED_article.tex
More text into HED document

@@
 
+\texttt{Relevant classes: Arc::MCC, Arc::Loader}
+
 In the HED data channels to the outside world may be set up by chains of small processing units called Message Chain Components (MCCs).
@@
 
-Each MCC has an entry method \texttt{process} which is called with Message being processed.
+Each MCC has an entry method \texttt{process()} which is called with Message being processed.
 It then processes Message by modifying it or creating new Message.
@@
 Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component.
+For that purpose <next> elements in MCC configuraton may have optional \texttt{id} attribute which allows to assign labels to all chain links to next MCCs in the chain.
+Supported labels are MCC dependant.
+By default simple MCCs support only one unnamed link.
+The MCCs with routing capabilities must have all supported labels documented.
 
@@
 \subsubsection{Payload}
+
+\texttt{Relevant classes: Arc::Payload, Arc::PayloadRawInterface, Arc::PayloadRaw, Arc::PayloadStreamInterface, Arc::PayloadStream, Arc::PayloadSOAP}
 
@@
 
+\texttt{Relevant classes: Arc::MessageAttributes, Arc::Message}
+
 The Message object may contain general purpose key and value pairs called Attributes.
@@
 
-\texttt{Relevant classes: Arc::SecAttr}
+\texttt{Relevant classes: Arc::SecAttr, Arc::MessageAuth, Arc::MessageAuthContext, Arc::Message}
 
@@
 
-\texttt{Relevant classes: ArcSec::SecHander}
+\texttt{Relevant classes: ArcSec::SecHander, Arc::MCC}
 
@@
 They are attached to the Chain in the same manner as other MCCs.
-But differently from MCC the \texttt{process} method of the Service does not pass Message to other components of the Chain.
+But differently from MCC the \texttt{process()} method of the Service does not pass Message to other components of the Chain.
 Instead they have to process incoming Messages and produce outgoing ones.
@@
 
-For reporting errors each \texttt{process} method returns instance of MCC\_Status class. That object carries predefined set of common error codes.
+For reporting errors each \texttt{process()} method returns instance of MCC\_Status class. That object carries predefined set of common error codes.
 
@@
 
-\texttt{Relevant classes: Arc::Loader}
+\texttt{Relevant classes: Arc::Loader, Arc::Config, Arc::LoaderFactory, Arc::MCCFactory, Arc::ServiceFactory, Arc::DMCFactory, Arc::SecHandlerFactory, Arc::PDPFactory, Arc::ACCFactory}
 
@@
 In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done in Python language.
+Currently only available language bindings are Python and Java.
+And currently it is possible to write only SOAP Service modules in Python and Java due to multiple inheritance limitation.
 
-% TODO it
+% TODO more of it
 
@@
 
-The server TCP MCC in the HED is special in that it produces messages by listening on a network socket rather than passing on messages from other MCCs.
+\texttt{Plugin names: tcp.service, tcp.client}
+
+\texttt{Library name: \{lib\}mcctcp}
+
+\texttt{SEcuruty handler queues: incoming, outgoing}
+
+% TODO: attributes and security attributes
+
+The server side TCP MCC in the HED is special in that it produces messages by listening on a network socket rather than passing on messages from other MCCs.
 As such it spawns new thread for every new connection to handle Messages and their responses throughout the Message Chain.
@@
 It is optional and should if present contain the single digit 4 or 6.
+
+The server side TCP MCC generates \texttt{PayloadStreamInterface} payload in the Message passed to next MCC whihc can be used for communicationg through open TCP channel.
+Currently it ignores any payload attached to the returned Message.
 }{}
 
-The server and client side MCCs are called \texttt{tcp.service} and \texttt{tcp.client} correspondingly. They are located in the shared library called \texttt{mcctcp}.
+The client side TCP MCC performs TCP connection to host and port specified in <Host> and <Port> elements inside <Connect> element of the MCC configuration.
+Then all incoming Messages of \texttt{process()} method are transfered ober TCP connection.
+Accepted Payload type of incoming Message is \texttt{PayloadRawInterface}.
+Returned Payload is of \texttt{PayloadStreamInterface} type.
+It represents established TCP conenction and \texttt{may} be used by previous MCCs in chain for direct communication.
+It is still prefered to call \texttt{process()} method instead.
 
-\subsubsection{TLS MCC}
-The server TLS MCC enables ...
+\paragraph{Configuration schema}
 
-\ifthenelse{\boolean{brieftechdesc}}{
-This MCC needs to be configured with a host certificate as well as trusted CA certificates.
-}{}
+\begin{verbatim}
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
+  xmlns:arc="http://www.nordugrid.org/schemas/ArcConfig/2007"
+  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
+  elementFormDefault="qualified">
 
-\ifthenelse{\boolean{verbtechdesc}}{
-The
-}{}
+    <xsd:simpleType name="Version_Type">
+        <!-- This element defines TCP/IP protocol version.  -->
+        <xsd:restriction base="xsd:string">
+            <xsd:enumeration value="4"/>
+            <xsd:enumeration value="6"/>
+        </xsd:restriction>
+    </xsd:simpleType>
+    <xsd:element name="Version" type="Version_Type"/>
+
+    <xsd:complexType name="Listen_Type">
+        <!--
+            This element defines listening TCP socket. If interface is missing socket
+            is bound to all local interfaces (not supported). There may be multiple Listen elements.
+        -->
+        <xsd:sequence>
+            <xsd:element name="Interface" type="xsd:string" minOccurs="0" maxOccurs="1"/>
+            <xsd:element name="Port" type="xsd:int" minOccurs="1" maxOccurs="1"/>
+            <xsd:element name="Version" type="Version_Type" minOccurs="0" maxOccurs="1"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="Listen" type="Listen_Type"/>
+    <xsd:complexType name="Connect_Type">
+        <!--
+            This element defines TCP connection to be established to specified Host at specified Port.
+            If LocalPort is defined TCP socket will be bound to this port number (not supported).
+        -->
+        <xsd:sequence>
+            <xsd:element name="Host" type="xsd:string" minOccurs="1" maxOccurs="1"/>
+            <xsd:element name="Port" type="xsd:int" minOccurs="1" maxOccurs="1"/>
+            <xsd:element name="LocalPort" type="xsd:int" minOccurs="0" maxOccurs="1"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="Connect" type="Connect_Type"/>
+</xsd:schema>
+\end{verbatim}
+
+\subsection{TLS MCC}
+
+\texttt{Plugin names: tls.service, tls.client}
+
+\texttt{Library name: \{lib\}mcctls}
+
+\texttt{Securuty handler queues: incoming, outgoing}
+
+% TODO: attributes and security attributes
+
+The server and client TLS MCCs provide transport level security (TLS) over any stream channel. Curently they interoperate well with TCP MCCs.
+
+The server side MCC accepts payload of type PayloadStreamInterface.
+It then creates own instance of object inherited from PayloadStreamInterface bound to initial payload and passes it to next MCC.
+This object is maintained inside Message Context under name \texttt{tls.service} and is destroyed when Context becomes inactive.
+Currenlty this MCC does not expect any payload to be returned from rest of the chain and passes no payload to previous MCC.
+
+The client side MCC behaves in similar way.
+It also establishes PayloadStreamInterface type object linked to same type of payload of next MCC.
+To obtain that last payload it makes a first call to next MCC with payload of type PayloadRawInterface and then uses returned payload - which is expected to be of PayloadStreamInterface type - to create own payload object with streaming capabilities and returns it back to previous MCC for further usage.
+
+Both client ans server side MCCs are implemented using OpenSSL toolkit \cite{openssl}, use X.509 infrastructure \cite{x509} for establishing secure connection and may be configured to get private key, certfificate or proxy credentials from files esiding at local file system.
+It also possible to specify location of Certification Authority cerfificate or to use all certificates located in specified directory.
+for more information see configuration schema with comments below.
+
+\paragraph{Configuration schema}
+
+\begin{verbatim}
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns="http://www.nordugrid.org/schemas/ArcMCCTLS/2007"
+  xmlns:arc="http://www.nordugrid.org/schemas/ArcConfig/2007"
+  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCTLS/2007"
+  elementFormDefault="qualified">
+    <xsd:complexType name="CACertificatesDir_Type">
+      <xsd:simpleContent>
+        <xsd:extension base="xsd:string">
+          <xsd:attribute name="PolicyGlobus" type="xsd:boolean" use="optional" default="false"/>
+        </xsd:extension>
+      </xsd:simpleContent>
+    </xsd:complexType>
+    <!-- Location of private key.
+      Default is /etc/grid-security/hostkey.pem for service
+      and none for client. -->
+    <xsd:element name="KeyPath" type="xsd:string"/>
+    <!-- Content of private key - not supported -->
+    <xsd:element name="Key" type="xsd:string"/>
+    <!-- Location of public certificate.
+      Default is /etc/grid-security/hostcert.pem for service
+      and none for client. -->
+    <xsd:element name="CertificatePath" type="xsd:string"/>
+    <!-- Content of public certificate - not supported -->
+    <xsd:element name="Certificate" type="xsd:string"/>
+    <!-- Location of proxy credentials - includes certificates, key and
+      chain of involved certificates. Overwrites elements Key, KeyPath,
+      Certificate and CertificatePath.
+      Default is none for client and none for service. -->
+    <xsd:element name="ProxyPath" type="xsd:string"/>
+    <!-- Content of proxy credentials - not supported -->
+    <xsd:element name="Proxy" type="xsd:string"/>
+    <!-- Location of certificate of CA. Default is none. -->
+    <xsd:element name="CACertificatePath" type="xsd:string"/>
+    <!-- Content of certificate of CA - not supported -->
+    <xsd:element name="CACertificate" type="xsd:string"/>
+    <!-- Directory containing certificates of accepted CAs.
+      Default is /etc/grid-security/ . -->
+    <xsd:element name="CACertificatesDir" type="xsd:string"/>
+</xsd:schema>
+\end{verbatim}
 
@@
 
+\texttt{Plugin names: http.service, http.client}
+
+\texttt{Library name: \{lib\}mcchttp}
+
+\texttt{Securuty handler queues: incoming, outgoing}
+
+% TODO: attributes and security attributes
+
+The server side HTTP MCC accepts messages with \texttt{PayloadStreamInterface} payload and parses HTTP related information from it.
+Information from the HTTP header is added to the Message Attributes.
+The body of HTTP message is passed to next MCC as \texttt{PayloadRawInterface} payload.
+In response this MCC expects also the Message with \texttt{PayloadRawInterface}.
+It is then prepended with HTTP reponse header and pushes it into initially providded stream channel.
+for output it returns empty \texttt{PayloadRawInterface} payload.
+
+This MCC routes results to multiple next MCCs in the chain.
+For that it accepts only labeled <next> elements in cinfoguration.
+Label names are those of HTTP methods (uppercase).
+HTTP messages will be routed to their destinations according HTTP method requested.
+
+The client side HTTP MCC will accept \texttt{PayloadRawInterface} payload as HTTP body and after prepending it with HTTP information passes to next MCC also as \texttt{PayloadRawInterface}.
+It accepts \texttt{PayloadStreamInterface} in response and after processing passes \texttt{PayloadRawInterface} back trhough the chain.
+
+\paragraph{Configuration schema}
+
+\begin{verbatim}
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
+  xmlns:arc="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
+  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
+  elementFormDefault="qualified">
+
+    <!--
+        These elements define endpoint and HTTP method for client HTTP MCC.
+    -->
+    <xsd:element name="Endpoint" type="xsd:string"/>
+    <xsd:element name="Method" type="xsd:string"/>
+</xsd:schema>
+\end{verbatim}
+
 \subsection{SOAP MCC}
 
+\texttt{Plugin names: soap.service, soap.client}
+
+\texttt{Library name: \{lib\}mccsoap}
+
+\texttt{Securuty handler queues: incoming, outgoing}
+
+% TODO: attributes and security attributes
+
+These MCCs convert payloads between dat chunks presented by PayloadRawInterface type object into dedicated PayloadSOAP payloads and vice versa. Currently it has no specific configuration parameters.
+
 \section{Implemented Security Handlers}
+
+\subsection{Simple List Auth}
+
+\texttt{Plugin name: arc.authz}
+
+\texttt{Library name: \{lib\}arcpdc}
+
+\subsection{Username Token}
+
+\texttt{Plugin name: usernametoken.handler}
+
+\texttt{Library name: \{lib\}arcpdc}
+
+\subsection{X.509 Token}
+
+\texttt{Plugin name: x509token.handler}
+
+\texttt{Library name: \{lib\}arcpdc}
 

### 2008-09-09 src/clients/echo/arcecho.cpp
Rename apecho -> arcecho

@@ -1,4 +1,4 @@
-// apecho.cpp
+// arcecho.cpp
 
 #include <iostream>
 #include <fstream>
@@ -11,7 +11,7 @@
   "ap" means "Arc Prototype".
   
   Usage:
-  apecho <message>
+  arcecho <message>
 
   Arguments:
   <message> A message that the echo service shall return back again.

### 2008-09-09 doc/client/arc-ws-clients.tex
Rename ap clients to arc clients and remove asstat??

@@
 \section{Job Submission}
-\label{sec:apsub}
+\label{sec:arcsub}
 
-The \verb#apsub# command is used to submit jobs to an A-REX service.
+The \verb#arcsub# command is used to submit jobs to an A-REX service.
 
 \begin{shaded}
-\paragraph{Usage} \verb:apsub <JSDL-file> <JobID-file>:
+\paragraph{Usage} \verb:arcsub <JSDL-file> <JobID-file>:
 \end{shaded}
@@
 \section{Job Status Query}
-\label{sec:apstat}
+\label{sec:arcstat}
 
-The \verb#apstat# command is used to query the status of a job that
+The \verb#arcstat# command is used to query the status of a job that
 has been submitted to an A-REX service.
@@
 \begin{shaded}
-\paragraph{Usage} \verb:apstat <JobID-file>:
+\paragraph{Usage} \verb:arcstat <JobID-file>:
 \end{shaded}
@@
 \item[\texttt{<JobID-file>}] The name of a file in which the
-Job ID is stored. This file was created by \verb:apsub: when the job
+Job ID is stored. This file was created by \verb:arcsub: when the job
 was submitted.
@@
 \section{Job Termination}
-\label{sec:apkill}
+\label{sec:arckill}
 
-The \verb#apkill# command is used to terminate a job that has been
+The \verb#arckill# command is used to terminate a job that has been
 submitted to an A-REX service.
@@
 \begin{shaded}
-\paragraph{Usage} \verb:apkill <JobID-file>:
+\paragraph{Usage} \verb:arckill <JobID-file>:
 \end{shaded}

### 2008-09-09 doc/ws-quick-guide/ws-quick-guide.tex
Rename apecho -> arcecho
Rename ap clients to arc clients and remove asstat??

@@
 \begin{lstlisting}
-$ARC_LOCATION/bin/apecho <message>
+$ARC_LOCATION/bin/arcecho <message>
 \end{lstlisting}
@@
 \verb|jsdl_stage.xml|. To submit job to A-REX service one may use
-\texttt{apsub} command:
+\texttt{arcsub} command:
 
 \begin{lstlisting}
-$ARC_LOCATION/bin/apsub https://localhost:60000/arex
+$ARC_LOCATION/bin/arcsub https://localhost:60000/arex
  /usr/local/share/doc/arc/jsdl_simple.xml id.xml
@@
 ``Submitted the job!''. Obtained job identifier is an XML document and is stored
-in id.xml file. It may then be used to query job state with \texttt{apstat}
+in id.xml file. It may then be used to query job state with \texttt{arcstat}
 utility:
@@
 \begin{lstlisting}
-$ARC_LOCATION/bin/apstat id.xml 2>/dev/null
+$ARC_LOCATION/bin/arcstat id.xml 2>/dev/null
 Job status: Running/Submiting
 
-$ARC_LOCATION/bin/apstat id.xml 2>/dev/null
+$ARC_LOCATION/bin/arcstat id.xml 2>/dev/null
 Job status: Running/Finishing
 
-$ARC_LOCATION/bin/apstat id.xml 2>/dev/null
+$ARC_LOCATION/bin/arcstat id.xml 2>/dev/null
 Job status: Finished/Finished
@@
 
-Full set of A-REX client tools consists of \texttt{apsub}, \texttt{apstat},
-\texttt{apkill}, \texttt{asstat}, \texttt{apget} and \texttt{apclean} commands.
+Full set of A-REX client tools consists of \texttt{arcsub}, \texttt{arcstat},
+\texttt{arckill}, \texttt{arcget} and \texttt{arcclean} commands.
 For more information please see man pages ot those utilities.

### 2008-09-10 doc/a-rex/arex_tech_doc.tex
rename cluster+qju.pl to CEinfo.pl
Implementing comments from Gabor. Commentig LRMS specific parts.

@@
 
-\begin{tabular}{|c|c|c|>{\raggedright}p{10cm}|}
+\begin{tabular}{|>{\centering}p{0.15\columnwidth}|>{\centering}p{0.15\columnwidth}|c|>{\raggedright}p{0.4\columnwidth}|}
 \hline 
@@
 A-REX internal state&
-\multicolumn{1}{c|}{Description}\tabularnewline
+Description\tabularnewline
 \hline
@@
 Communication with local batch system is in process\tabularnewline
-\cline{2-2} \cline{3-3} \cline{4-4} 
+\hline 
 &
-\multicolumn{1}{c|}{Queued}&
-\multicolumn{1}{c|}{INLRMS}&
-\multicolumn{1}{p{10cm}|}{Job entered local batch system but is not runnning now. This state
-is not recognised by the A-REX yet. \emph{Executing} is reported instead }\tabularnewline
+Queued&
+INLRMS&
+Job entered local batch system but is not runnning now. This state
+is not recognised by the A-REX yet. \emph{Executing} is reported instead \tabularnewline
 \cline{2-2} \cline{3-3} \cline{4-4} 
@@
 Executed&
-INLRMS,INLRMS+PENDING&
+INLRMS, INLRMS+PENDING&
 Job execution in local batch system has finished. The A-REX dos not
@@
 \hline 
-\multicolumn{1}{|c|}{Failed}&
-\multicolumn{1}{c|}{Failed}&
+Failed&
+Failed&
 FINISHED&
@@
 wake it up. It's used to increase responsiveness of A-REX.
-\item \emph{cluster+qju.pl} - Collects and generates information about computing
+\item \emph{CEinfo.pl} - Collects and generates information about computing
 resource as XML document in NorduGrid and Glue 2 format.
@@
 \item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
-Currently supported LRMS are PBS/Torque, Condor and SGE. Also \emph{fork}
-pseudo-LRMS is supported for testing purposes. This module is responsible
-for job submission to the LRMS.
+Currently supported LRMS are PBS/Torque, Condor, LoadLeveler, LSF
+and SGE. Also \emph{fork} pseudo-LRMS is supported for testing purposes.
+This module is responsible for job submission to the LRMS.
 \item \textit{cancel-{*}-job} - This script is for canceling jobs which
@@
 should be shared across cluster nodes. Otherwise, LRMS-specific methods
-must be used to transfer files to execution nodes. For more information
-see section \ref{sub:LRMS}.
+must be used to transfer files to execution nodes. 
+\end{itemize}
+\begin{comment} (LRMS\_COMMENT) For more information see section
+\ref{sub:LRMS}.\end{comment}
+
+\begin{itemize}
 \item \textit{control directory} - In this directory A-REX stores an information
@@
 \end{itemize}
+\begin{comment}
+
 Following are global commands specific to communication with underlying
@@
 with PBS server's log files.
+\end{itemize}
+\end{comment}
+
+\begin{itemize}
 \item \textbf{\textit{gnu\_time}}\textit{=path} - path to \emph{time} utility.
@@
 nodes have an access to session directory through a shared file system
-like NFS. Corresponds to an environment variable RUNTIME\_NODE\_SEES\_FRONTEND
-(see section \ref{sub:LRMS}).
+like NFS. \begin{comment} Corresponds to an environment variable
+RUNTIME\_NODE\_SEES\_FRONTEND (see section \ref{sub:LRMS}). \end{comment}
 \item \textbf{\textit{nodename}}\textit{=command} - command to obtain hostname
@@
 
+\begin{comment}(LRMS\_COMMENT)
+
 \subsection{\label{sub:LRMS}LRMS support}
@@
 \end{figure}
+
+\end{comment}
 
@@
 
-Please read \emph{User Manual {[}??]} for usage instructions.
+Please read \emph{User Guide} \cite{userguide1} for usage instructions.
 
@@
 \hline
-&
-\tabularnewline
-\hline
-&
-\tabularnewline
-\hline
 \end{longtable}
+
+\begin{comment}(LRMS\_COMMENT)
 
@@
 
-\begin{lyxcode}
-
-\end{lyxcode}
+\end{comment}
+
 \bibliography{grid}

### 2008-09-11 doc/hed/ARCHED_article.tex
More for HED document. MCCs are done except secattr - those should go to separate doc. Minimal description of DMCs is added. Reference/description of common purpose functionality is missing yet.

@@
 
-\texttt{SEcuruty handler queues: incoming, outgoing}
+\texttt{Security handler queues: incoming, outgoing}
+
+\texttt{Message attrbiutes: TCP:HOST, TCP:PORT, TCP:REMOTEHOST, TCP:REMOTEPORT, TCP:ENDPOINT, ENDPOINT}
 
@@
 \ifthenelse{\boolean{verbtechdesc}}{
-This MCC can be configuered with one or more \texttt{<tcp:Listen>} elements which in turn contain the elements \texttt{<tcp:Port>}, \texttt{<tcp:Interface>} and \texttt{<tcp:Version>}.
+This MCC can be configured with one or more \texttt{<tcp:Listen>} elements which in turn contain the elements \texttt{<tcp:Port>}, \texttt{<tcp:Interface>} and \texttt{<tcp:Version>}.
 The \texttt{<tcp:Port>} element is mandatory and should contain an integer corresponding to the TCP port to listen to.
@@
 Currently it ignores any payload attached to the returned Message.
+That MCC also fills following Message Attrbutes in incoming message while passing to next MCC:
+\begin{enumerate}
+
+\item TCP:HOST - IP address of local interface which was used to establish TCP connection
+
+\item TCP:PORT - local TCP port which was used for connection
+
+\item TCP:REMOTEHOST - IP address of contacting client
+
+\item TCP:REMOTEPORT - TCP port of contacting client
+
+\item TCP:ENDPOINT - URL-like combination of ://(TCP:HOST):(TCP:PORT)
+
+\item ENDPOINT - same as TCP:ENDPOINT
+
+\end{enumerate}
 }{}
@@
 
+\texttt{Message attrbiutes: TLS:PEERDN, TLS:IDENTITYDN}
+
 % TODO: attributes and security attributes
@@
 Currenlty this MCC does not expect any payload to be returned from rest of the chain and passes no payload to previous MCC.
+
+The server MCC fills Message Attributes TLS:PEERDN and TLS:IDENTITYDN representing subjects of last cretificate in client's certificate chain used for establishing secure connection and subject of last certificate which is not a proxy certificate correspondingly.
+Use of thos attributes is deprecated.
+It is advised to use Security Attributes instead.
 
@@
 
+\texttt{Message attrbiutes: HTTP:METHOD, HTTP:CODE, HTTP:REASON, HTTP:RANGESTART, HTTP:RANGEEND, HTTP:ENDPOINT, HTTP:*, ENDPOINT}
+
 % TODO: attributes and security attributes
@@
 
-This MCC routes results to multiple next MCCs in the chain.
-For that it accepts only labeled <next> elements in cinfoguration.
+That MCC routes results to multiple next MCCs in the chain.
+For that it accepts only labeled <next> elements in confoguration.
 Label names are those of HTTP methods (uppercase).
@@
 
+That MCC also fills following Message Attrbutes in incoming message while passing to next MCC:
+\begin{enumerate}
+
+\item HTTP:METHOD - HTTP method as defined in HTTP request header.
+
+\item HTTP:RANGESTART - Range request start offset
+
+\item HTTP:RANGEEND - Range request end offset
+
+\item HTTP:ENDPOINT - URL or path as specified in HTTP request header
+
+\item HTTP:* - Here * stands for any name. All HTTP options from HTTP request header are converted into Message Attributes named HTTP:\{option name\}
+
+\item ENDPOINT - same as HTTP:ENDPOINT
+
+\end{enumerate}
+
+For outgoing message server MCC converts all HTTP:* Message Attriutes into corresponding HTTP response header attribute.
+
 The client side HTTP MCC will accept \texttt{PayloadRawInterface} payload as HTTP body and after prepending it with HTTP information passes to next MCC also as \texttt{PayloadRawInterface}.
 It accepts \texttt{PayloadStreamInterface} in response and after processing passes \texttt{PayloadRawInterface} back trhough the chain.
+
+The client MCC also convert all HTTP:* Message Attributes into corresponding HTTP header options while creating HTTP request.
+If define it uses HTTP:METHOD and HTTP:ENDPOINT as method and URL/path of HTTP request.
+If not specified those defined in configuration are used.
+For HTTP response it sets HTTP:CODE and HTTP:REASON to response code and reason correspondingly.
+It also performs conversion from header options into HTTP:* Message Attributes.
 
@@
 
+\texttt{Message attrbiutes: SOAP:ENDPOINT, ENDPOINT}
+
 % TODO: attributes and security attributes
 
-These MCCs convert payloads between dat chunks presented by PayloadRawInterface type object into dedicated PayloadSOAP payloads and vice versa. Currently it has no specific configuration parameters.
+These MCCs convert payloads between data chunks presented by PayloadRawInterface type object into dedicated PayloadSOAP payloads and vice versa. Currently it has no specific configuration parameters.
+
+The server side MCC sets Message Attributes SOAP:ENDPOINT and ENDPOINT to URL present in 'texttt{To} element of Web Service Addressing information stored in SOAP header.
+For proper SOAP over HTTP binding it also sets \texttt{HTTP:Content-Type} Message Attribute to value \texttt{application/soap+xml} or \texttt{text/xml} for SOAP version 1.2 and 1.1 respectively for response message.
+Also in case of SOAP fault response \texttt{HTTP:CODE} is set to \texttt{500} and \texttt{HTTP:REASON} to \texttt{SOAP FAULT}.
+
+The client side MCC for outgoing message accepts \texttt{SOAP:ACTION} Message Attribute and uses it to fill \texttt{HTTP:Content-Type} or \texttt{HTTP:SOAPAction} depending on SOAP version.
 
@@
 
+\texttt{Protocol name: file}
+
+This DMC implements access to local or remote mounted file system. It allows reading and writing content of file, listing content of directories and checking presence of the object.
+
+Name of pligin is \texttt{file} and it is located in \texttt{\{lib\}dmcfile} loadable module.
+
 \subsection{GridFTP/FTP DMC}
+
+\texttt{Protocol name: ftp, gsiftp}
+
+This DMC implements access to a data server using FTP or GridFTP protocol. It allows reading and writing content of stored file, listing content of directories and checking presence of the object.
+
+Name of pligin is \texttt{gridftp} and it is located in \texttt{\{lib\}dmcgridftp} loadable module.
 
@@
 
+\texttt{Protocol name: http, https, httpg (not implemented yet)}
+
+This DMC implements access to a data server using HTTP over TCP or over TLS protocol.
+It uses GET for reading and PUT for storing files.
+Data is always transfered in chunks due to current limitation of HTTP MCC.
+Listing is also supported.
+For that DMC extracts references from \texttt{A} tags of obtained HTLM content.
+If retrieved content is not HTML or if content of it's \texttt{TITLE} tag does not start from ""Index of /"" then only information about resquested URL itself is listed.
+
+Name of pligin is \texttt{http} and it is located in \texttt{\{lib\}dmchttp} loadable module.
+
 \subsection{LDAP DMC}
+
+\texttt{Protocol name: ldap}
+
+The LDAP DMC implements access to a data accessible through LDAP protocol.
+Currently it can only retrieve content of LDAP tree.
+Retrieved content is converted into XML.
+The DMC supports following kind of URL
+
+\texttt{ldap://host[:port]/base[?[attributes][?[scope][?filter]]]}
+
+\begin{itemize}
+\item \texttt{attributes} is coma separted list of attributes which has to be retrieved,
+\item \texttt{scope} is either \texttt{base}, \texttt{one} or \texttt{sub} as defined by LDAP,
+\item \texttt{filter} is LDAP filter.
+\end{itemize}
+
+Name of pligin is \texttt{ldap} and it is located in \texttt{\{lib\}dmcldap} loadable module.
 
@@
 
+\texttt{Protocol name: lfc}
+
+The LFC DMC provides an access to LCG File Catalog service and supports file listing, URL resolution and (un)registration of file locations.
+It uses LFC protocol implementation provided by gLite middleware.
+
+Name of pligin is \texttt{lfc} and it is located in \texttt{\{lib\}dmclfc} loadable module.
+
 \subsection{RLS DMC}
 
-\chapter{Conclusion}
+\texttt{Protocol name: rls}
+
+The RLS DMC provides an access to Relica Location Service and supports file listing, URL resolution and (un)registration of file locations.
+It uses RLS protocol implementation provided by Globus Toolkit.
+
+Name of pligin is \texttt{rls} and it is located in \texttt{\{lib\}dmcrls} loadable module.
+
+%\chapter{Conclusion}
 

### 2008-09-12 doc/hed/ARCHED_article.tex
Mostly done HED document.

@@
 
+The configuration of Loader is made of following elements (see also configuration schema at http://svn.nordugrid.org/trac/nordugrid/export/10146/arc1/trunk/src/hed/libs/loader/mcc.xsd):
+
+\begin{itemize}
+
+\item \texttt{Chain} is base element. It has no special meaning for configuration of chain and is only sed to group other elements in a logical way. Chains may be nested.
+
+\item \texttt{Component} represents the MCC.
+It's attribute \texttt{name} defines name of MCC which has to be instantiated.
+This MCC is being looked among those contained in modules loaded by \texttt{ModuleManager} element.
+If not found then the Loader tries to find module with name corresponding to name of the MCC.
+Each MCC also must have unique identifier specified by \texttt{id} attribute.
+Subelement \texttt{next} specifies link to next MCC in the Chain.
+It uses uinique identifier specified by \texttt{id}.
+The \texttt{next} element may have an optional text which is used to distinguish different Message propagation path in MCC specific way.
+The \texttt{SecurityHandler} subelement defines the Security Handler and queue ato which it has to be attached.
+Also attribute \texttt{entry} my be used to create an entry point to the Chain.
+
+\item \texttt{Plexer} element instantiates Plexer component. It's configuration is similar to one of \texttt{Component} element.
+
+\item \texttt{Service} element defines plugin implementing final Component in the Chain - service. Form Loader point of view it's configuration is identical to that of the \texttt{Component}.
+
+\item \texttt{ModuleManager} defines parameters needed to find loadable modules containiong plugins.
+Currently only subelement \texttt{path} is supported whihc defines path on file system where loadable modules may be found.
+
+\item \texttt{Plugins} specifies name of the loadable module.
+
+\end{itemize}
+
 \subsection{Sessions and Contexts}
@@
 
-\item Register and unregister presence of data
+\item Register and unregister presence of data - for indexing endpoints
+
+\item Resolve final or lower level location of data from stored metadata - for indexing endpoints
 
@@
 More in depth techical information about the DMC can be found in \cite{dmc}.
+
+\subsection{Generic purpose components}
+
+The HED also includes vaste amount of common purpose components:
+
+\begin{itemize}
+
+\item XMLNode is class for managing parsed XML structures. It provides minimal functionality for operationg on XML elements, attributes and namespaces.
+
+\item Thread management functions provide few mostly used functions for thread creation.
+
+\item URL class gives access to various parts of URL. It has support for many specific URLs built-in.
+
+\item String utilities for easy conversion between various types to strings and back.
+
+\item Run class for starting, communicatiing and monitoring external processes.
+
+\item Logger class provides multilevel controllable logging functionality.
+
+\item Time and Period classes alows parsing and generating textual description of times, dates and time periods in various formats.
+
+\item RegularExpression class providing C++ wrapper for regexp related functions.
+
+\item Config class makes it possible to manipulate configuration file. And related classes give an access to specific parts of configuration.
+
+\item Counter and IntraProcessCounter provide a way to count abstract resources.
+
+\end{itemize}
+
+\subsection{Web Service related components}
+
+There is also a set of components implementing various Web Service related functionality. Those include:
+
+\begin{itemize}
+
+\item WSAEnpointReference and WSAHeader manipulate WS-Addressing \cite{ws-addr-soap} information in XML element and SOAP header correspondingly.
+
+\item SAMLToken, UsernameToken and X509Token are for consuming and generating token of same name according to various profiles of WS Security specifications \cite{ws-security}.
+
+\item WSRF and related and inherited classes offer a way to generate and anaylyze various Web Services Resource specification \cite{wsrf} related elements.
+
+\item DelegationConsumerSOAP, DelegationProviderSOAP and DelegationContainerSOAP implement web service interface which enables client to delegate X509 credentials to service.
+
+\item InformationInterface and related classes may be used to implement information interface service and client part in a way common for services built on top of HED.
+Also InfoCacheInterface extends it with caching functionality and InfoRegister provides service registration ability.
+
+\end{itemize}
+
+\subsection{Daemon}
+
+All componets of the Hosting Environment are compiled into loadable libraries and may be used in various executables.
+But there is also a dedicated executable provided - \texttt{arched}.
+It accepts configuration file containing XML document and passes it to Loader component.
+Then the Loader component takes care of loading all modules and instantiating all Message Chains.
+The \texttt{arched} also initializes Logger component, configures it and directs it's output to specified or to screen.
+The \texttt{arched} normally runs as background process but can be run in foreground as well.
+For more information about \texttt{arched} capabilities please read it's manual page.
 

### 2008-09-15 doc/hed/ARCHED_article.tex
Spelling fixes for HED document

@@
 External dependencies needed to be kept to a minimum while also taking into consideration their ubiquity or relative rarity as well as license related concerns.
-Software of this level of complexity must of course depend on many external libraries and components but each such dependenscy has been introduced only after due consideration.
+Software of this level of complexity must of course depend on many external libraries and components but each such dependency has been introduced only after due consideration.
 
@@
 %How was this implemented, dynamic loading, mcc...
-In the technical design it turned out that the endeavours to provide dynamic loading, portability and a well tested high level memory management could all be greatly assisted by introducing glibmm - the C++ interface to the gnome projects library for memory management and related functionality.
+In the technical design it turned out that the endeavors to provide dynamic loading, portability and a well tested high level memory management could all be greatly assisted by introducing glibmm - the C++ interface to the gnome projects library for memory management and related functionality.
 This enables the developers to write code in a way easily portable across various operating systems and architectures.
@@
 \item sometimes using the HED terms to refer to collection of libraries which is used by service or other component developers.
-These libraries define interfaces and implement some common classes which may simplify the life of service and component developes however only few of these classes are mandatory to use to make the components and services loadable and hookable by the daemon.
+These libraries define interfaces and implement some common classes which may simplify the life of service and component developers however only few of these classes are mandatory to use to make the components and services loadable and hookable by the daemon.
 
@@
 In the HED data channels to the outside world may be set up by chains of small processing units called Message Chain Components (MCCs).
-The chain is an ordered list of MCCs and their interconection can be described in the configuration file.
+The chain is an ordered list of MCCs and their interconnection can be described in the configuration file.
 The MCCs work on units called Messages which represent data going in to or out of the HED.
@@
 
-The developer who writes an MCC is free to choose any 3rd party library and component to implement the functionalty of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by \texttt{arcloader} and \texttt{arcmessage} libraries of ARC1.
+The developer who writes an MCC is free to choose any 3rd party library and component to implement the functionality of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by \texttt{arcloader} and \texttt{arcmessage} libraries of ARC1.
 
 The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs.
-Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component.
-For that purpose <next> elements in MCC configuraton may have optional \texttt{id} attribute which allows to assign labels to all chain links to next MCCs in the chain.
+Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTP service component.
+For that purpose <next> elements in MCC configuration may have optional \texttt{id} attribute which allows to assign labels to all chain links to next MCCs in the chain.
 Supported labels are MCC dependant.
@@
 Most of the MCCs has a client and a server version because the behavior of an MCC should be different depending on whether it is sitting on the server or client side.
-The typical scenario here is illustrated by the TCP MCC which should listen and wait for incomming messages on a socket on the server side but call \texttt{connect()} on the client side.
+The typical scenario here is illustrated by the TCP MCC which should listen and wait for incoming messages on a socket on the server side but call \texttt{connect()} on the client side.
 
@@
 
-Main content of the information is transfered using the Payload part of the Message.
+Main content of the information is transferred using the Payload part of the Message.
 There are no limitations on functionality of Payload object except that it must be inherited from MessagePayload class.
@@
 It covers case of sequentially accessible information.
-The main purpose of that Payload is to serve protocols which define continous data stream like TCP.
+The main purpose of that Payload is to serve protocols which define continuous data stream like TCP.
 
@@
 
-The Secuirity Attributes are stored as key and value pairs.
+The Security Attributes are stored as key and value pairs.
 Each key may have only single value attached.
@@
 
-The value if Secuirity Attribute is an object of class inherited from SecAttr.
+The value if Security Attribute is an object of class inherited from SecAttr.
 The HED implements class SecAttr which serves as definition of interface for all Security Attributes.
@@
 The services are dynamically loaded on start up just like the MCCs.
-They almost identic to MCCs with an exception that they constitute the last link in the Message Chain.
+They are almost identical to MCCs with an exception that they constitute the last link in the Message Chain.
 They are attached to the Chain in the same manner as other MCCs.
@@
 
-In general case multiple Services living in the HED so the incomming Message should route to the proper Service.
+In general case multiple Services living in the HED so the incoming Message should route to the proper Service.
 The Plexer MCC does this job.
 It takes the the ENDPOINT attributes of the message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forward the message to the all matching service.
-It acts as a dispacher.
-The Plexer is also special in a sence that it is not a plugin but part of the \texttt{arcloader} library.
+It acts as a dispatcher.
+The Plexer is also special in a sense that it is not a plugin but part of the \texttt{arcloader} library.
 
@@
 Only if protocol does not provide error handling MCC\_Status should be used.
-It is also advisable to convert MCC\_Status error obtained from next MCC in the Chain into protocol psecific error Message if possible.
+It is also advisable to convert MCC\_Status error obtained from next MCC in the Chain into protocol specific error Message if possible.
 
@@
 
-\item \texttt{Chain} is base element. It has no special meaning for configuration of chain and is only sed to group other elements in a logical way. Chains may be nested.
+\item \texttt{Chain} is base element. It has no special meaning for configuration of chain and is only used to group other elements in a logical way. Chains may be nested.
 
@@
 Subelement \texttt{next} specifies link to next MCC in the Chain.
-It uses uinique identifier specified by \texttt{id}.
+It uses unique identifier specified by \texttt{id}.
 The \texttt{next} element may have an optional text which is used to distinguish different Message propagation path in MCC specific way.
-The \texttt{SecurityHandler} subelement defines the Security Handler and queue ato which it has to be attached.
+The \texttt{SecurityHandler} subelement defines the Security Handler and queue to which it has to be attached.
 Also attribute \texttt{entry} my be used to create an entry point to the Chain.
@@
 
-\item \texttt{ModuleManager} defines parameters needed to find loadable modules containiong plugins.
-Currently only subelement \texttt{path} is supported whihc defines path on file system where loadable modules may be found.
+\item \texttt{ModuleManager} defines parameters needed to find loadable modules containing plugins.
+Currently only subelement \texttt{path} is supported which defines path on file system where loadable modules may be found.
 
@@
 
-\texttt{Relevant classes: Arc::MessageContext, Arc::MessageContextElement, Arc::MessageAuthContext, Arc::ChianContext }
+\texttt{Relevant classes: Arc::MessageContext, Arc::MessageContextElement, Arc::MessageAuthContext, Arc::ChainContext }
 
@@
 
-\item The Chain lifetime - is time period while componnets making the Chain exist.
+\item The Chain lifetime - is time period while components making the Chain exist.
 This lifetime is represented by ChainContext class.
@@
 
-More in depth techical information about the DMC can be found in \cite{dmc}.
+More in depth technical information about the DMC can be found in \cite{dmc}.
 
@@
 
-The HED also includes vaste amount of common purpose components:
+The HED also includes vast amount of common purpose components:
 
@@
 
-\item XMLNode is class for managing parsed XML structures. It provides minimal functionality for operationg on XML elements, attributes and namespaces.
+\item XMLNode is class for managing parsed XML structures. It provides minimal functionality for operating on XML elements, attributes and namespaces.
 
@@
 
-\item Run class for starting, communicatiing and monitoring external processes.
+\item Run class for starting, communicating and monitoring external processes.
 
@@
 
-\item Time and Period classes alows parsing and generating textual description of times, dates and time periods in various formats.
+\item Time and Period classes allows parsing and generating textual description of times, dates and time periods in various formats.
 
@@
 
-\item WSRF and related and inherited classes offer a way to generate and anaylyze various Web Services Resource specification \cite{wsrf} related elements.
+\item WSRF and related and inherited classes offer a way to generate and analyze various Web Services Resource specification \cite{wsrf} related elements.
 
@@
 
-In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done in Python language.
+In order to facilitate the development of services, API bindings for languages other than C++ are provided and some service development has already been done in Python language.
 Currently only available language bindings are Python and Java.
@@
 
-This chapter describes components which are implemented alongside with the HED infractructure itself.
-Although strictly not belonging to infrastructure this minimal set of components is necessary to make inrastructure useful.
+This chapter describes components which are implemented alongside with the HED infrastructure itself.
+Although strictly not belonging to infrastructure this minimal set of components is necessary to make infrastructure useful.
 
@@
 
-\texttt{Message attrbiutes: TCP:HOST, TCP:PORT, TCP:REMOTEHOST, TCP:REMOTEPORT, TCP:ENDPOINT, ENDPOINT}
+\texttt{Message attributes: TCP:HOST, TCP:PORT, TCP:REMOTEHOST, TCP:REMOTEPORT, TCP:ENDPOINT, ENDPOINT}
 
@@
 One could envision other MCCs having these properties but producing messages from other sources such as e.g.
-unix sockets.
+UNIX sockets.
 
@@
 
-The server side TCP MCC generates \texttt{PayloadStreamInterface} payload in the Message passed to next MCC whihc can be used for communicationg through open TCP channel.
+The server side TCP MCC generates \texttt{PayloadStreamInterface} payload in the Message passed to next MCC which can be used for communicating through open TCP channel.
 Currently it ignores any payload attached to the returned Message.
-That MCC also fills following Message Attrbutes in incoming message while passing to next MCC:
+That MCC also fills following Message Attributes in incoming message while passing to next MCC:
 \begin{enumerate}
@@
 The client side TCP MCC performs TCP connection to host and port specified in <Host> and <Port> elements inside <Connect> element of the MCC configuration.
-Then all incoming Messages of \texttt{process()} method are transfered ober TCP connection.
+Then all incoming Messages of \texttt{process()} method are transferred over TCP connection.
 Accepted Payload type of incoming Message is \texttt{PayloadRawInterface}.
 Returned Payload is of \texttt{PayloadStreamInterface} type.
-It represents established TCP conenction and \texttt{may} be used by previous MCCs in chain for direct communication.
-It is still prefered to call \texttt{process()} method instead.
+It represents established TCP connection and \texttt{may} be used by previous MCCs in chain for direct communication.
+It is still preferred to call \texttt{process()} method instead.
 
@@
 
-\texttt{Securuty handler queues: incoming, outgoing}
+\texttt{Security handler queues: incoming, outgoing}
 
-\texttt{Message attrbiutes: TLS:PEERDN, TLS:IDENTITYDN}
+\texttt{Message attributes: TLS:PEERDN, TLS:IDENTITYDN}
 
@@
 
-The server and client TLS MCCs provide transport level security (TLS) over any stream channel. Curently they interoperate well with TCP MCCs.
+The server and client TLS MCCs provide transport level security (TLS) over any stream channel. Currently they interoperate well with TCP MCCs.
 
@@
 This object is maintained inside Message Context under name \texttt{tls.service} and is destroyed when Context becomes inactive.
-Currenlty this MCC does not expect any payload to be returned from rest of the chain and passes no payload to previous MCC.
+Currently this MCC does not expect any payload to be returned from rest of the chain and passes no payload to previous MCC.
 
-The server MCC fills Message Attributes TLS:PEERDN and TLS:IDENTITYDN representing subjects of last cretificate in client's certificate chain used for establishing secure connection and subject of last certificate which is not a proxy certificate correspondingly.
-Use of thos attributes is deprecated.
+The server MCC fills Message Attributes TLS:PEERDN and TLS:IDENTITYDN representing subjects of last certificate in client's certificate chain used for establishing secure connection and subject of last certificate which is not a proxy certificate correspondingly.
+Use of those attributes is deprecated.
 It is advised to use Security Attributes instead.
@@
 
-Both client ans server side MCCs are implemented using OpenSSL toolkit \cite{openssl}, use X.509 infrastructure \cite{x509} for establishing secure connection and may be configured to get private key, certfificate or proxy credentials from files esiding at local file system.
-It also possible to specify location of Certification Authority cerfificate or to use all certificates located in specified directory.
+Both client ans server side MCCs are implemented using OpenSSL toolkit \cite{openssl}, use X.509 infrastructure \cite{x509} for establishing secure connection and may be configured to get private key, certificate or proxy credentials from files residing at local file system.
+It also possible to specify location of Certification Authority certificate or to use all certificates located in specified directory.
 for more information see configuration schema with comments below.
@@
 
-\texttt{Securuty handler queues: incoming, outgoing}
+\texttt{Security handler queues: incoming, outgoing}
 
-\texttt{Message attrbiutes: HTTP:METHOD, HTTP:CODE, HTTP:REASON, HTTP:RANGESTART, HTTP:RANGEEND, HTTP:ENDPOINT, HTTP:*, ENDPOINT}
+\texttt{Message attributes: HTTP:METHOD, HTTP:CODE, HTTP:REASON, HTTP:RANGESTART, HTTP:RANGEEND, HTTP:ENDPOINT, HTTP:*, ENDPOINT}
 
@@
 In response this MCC expects also the Message with \texttt{PayloadRawInterface}.
-It is then prepended with HTTP reponse header and pushes it into initially providded stream channel.
+It is then prepended with HTTP response header and pushes it into initially provided stream channel.
 for output it returns empty \texttt{PayloadRawInterface} payload.
@@
 That MCC routes results to multiple next MCCs in the chain.
-For that it accepts only labeled <next> elements in confoguration.
+For that it accepts only labeled <next> elements in configuration.
 Label names are those of HTTP methods (uppercase).
@@
 
-That MCC also fills following Message Attrbutes in incoming message while passing to next MCC:
+That MCC also fills following Message Attributes in incoming message while passing to next MCC:
 \begin{enumerate}
@@
 
-For outgoing message server MCC converts all HTTP:* Message Attriutes into corresponding HTTP response header attribute.
+For outgoing message server MCC converts all HTTP:* Message Attributes into corresponding HTTP response header attribute.
 
 The client side HTTP MCC will accept \texttt{PayloadRawInterface} payload as HTTP body and after prepending it with HTTP information passes to next MCC also as \texttt{PayloadRawInterface}.
-It accepts \texttt{PayloadStreamInterface} in response and after processing passes \texttt{PayloadRawInterface} back trhough the chain.
+It accepts \texttt{PayloadStreamInterface} in response and after processing passes \texttt{PayloadRawInterface} back through the chain.
 
@@
 
-\texttt{Securuty handler queues: incoming, outgoing}
+\texttt{Security handler queues: incoming, outgoing}
 
-\texttt{Message attrbiutes: SOAP:ENDPOINT, ENDPOINT}
+\texttt{Message attributes: SOAP:ENDPOINT, ENDPOINT}
 
@@
 
-\subsection{Simple List Auth}
+For more information about security related capabilities of the HED please see dedicated document at \cite{sec-arc1}.
 
-\texttt{Plugin name: arc.authz}
+%\subsection{Simple List Auth}
+%
+%\texttt{Plugin name: arc.authz}
+%
+%\texttt{Library name: \{lib\}arcpdc}
 
-\texttt{Library name: \{lib\}arcpdc}
+%\subsection{Username Token}
+%
+%\texttt{Plugin name: usernametoken.handler}
+%
+%\texttt{Library name: \{lib\}arcpdc}
 
-\subsection{Username Token}
-
-\texttt{Plugin name: usernametoken.handler}
-
-\texttt{Library name: \{lib\}arcpdc}
-
-\subsection{X.509 Token}
-
-\texttt{Plugin name: x509token.handler}
-
-\texttt{Library name: \{lib\}arcpdc}
+%\subsection{X.509 Token}
+%
+%\texttt{Plugin name: x509token.handler}
+%
+%\texttt{Library name: \{lib\}arcpdc}
 
@@
 
-Name of pligin is \texttt{file} and it is located in \texttt{\{lib\}dmcfile} loadable module.
+Name of plugin is \texttt{file} and it is located in \texttt{\{lib\}dmcfile} loadable module.
 
@@
 
-Name of pligin is \texttt{gridftp} and it is located in \texttt{\{lib\}dmcgridftp} loadable module.
+Name of plugin is \texttt{gridftp} and it is located in \texttt{\{lib\}dmcgridftp} loadable module.
 
@@
 It uses GET for reading and PUT for storing files.
-Data is always transfered in chunks due to current limitation of HTTP MCC.
+Data is always transferred in chunks due to current limitation of HTTP MCC.
 Listing is also supported.
-For that DMC extracts references from \texttt{A} tags of obtained HTLM content.
+For that DMC extracts references from \texttt{A} tags of obtained HTML content.
 If retrieved content is not HTML or if content of it's \texttt{TITLE} tag does not start from ""Index of /"" then only information about resquested URL itself is listed.
 
-Name of pligin is \texttt{http} and it is located in \texttt{\{lib\}dmchttp} loadable module.
+Name of plugin is \texttt{http} and it is located in \texttt{\{lib\}dmchttp} loadable module.
 
@@
 \begin{itemize}
-\item \texttt{attributes} is coma separted list of attributes which has to be retrieved,
+\item \texttt{attributes} is coma separated list of attributes which has to be retrieved,
 \item \texttt{scope} is either \texttt{base}, \texttt{one} or \texttt{sub} as defined by LDAP,
@@
 
-Name of pligin is \texttt{ldap} and it is located in \texttt{\{lib\}dmcldap} loadable module.
+Name of plugin is \texttt{ldap} and it is located in \texttt{\{lib\}dmcldap} loadable module.
 
@@
 
-Name of pligin is \texttt{lfc} and it is located in \texttt{\{lib\}dmclfc} loadable module.
+Name of plugin is \texttt{lfc} and it is located in \texttt{\{lib\}dmclfc} loadable module.
 
@@
 
-The RLS DMC provides an access to Relica Location Service and supports file listing, URL resolution and (un)registration of file locations.
+The RLS DMC provides an access to Replica Location Service and supports file listing, URL resolution and (un)registration of file locations.
 It uses RLS protocol implementation provided by Globus Toolkit.
 
-Name of pligin is \texttt{rls} and it is located in \texttt{\{lib\}dmcrls} loadable module.
+Name of plugin is \texttt{rls} and it is located in \texttt{\{lib\}dmcrls} loadable module.
 

### 2008-09-17 doc/client/client_technical.tex
Client documentation update

--- 2008-09-08 doc/client/client_technical.tex
@@
 \newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
+\newcommand{\arclib}{libarcclient}
 \newcommand{\ACC}{\texttt{ACC}}
+\newcommand{\ACCConfig}{\texttt{ACCConfig}}
+\newcommand{\Broker}{\texttt{Broker}}
+\newcommand{\ClientInterface}{\texttt{ClientInterface}}
+\newcommand{\ClientTCP}{\texttt{ClientTCP}}
+\newcommand{\ClientHTTP}{\texttt{ClientHTTP}}
+\newcommand{\ClientSOAP}{\texttt{ClientSOAP}}
+\newcommand{\ExecutionTarget}{\texttt{ExecutionTarget}}
+\newcommand{\Job}{\texttt{Job}}
+\newcommand{\JobController}{\texttt{JobController}}
+\newcommand{\JobDescription}{\texttt{JobDescription}}
+\newcommand{\JobSupervisor}{\texttt{JobSupervisor}}
+\newcommand{\Loader}{\texttt{Loader}}
+\newcommand{\Logger}{\texttt{Logger}}
 \newcommand{\TargetGenerator}{\texttt{TargetGenerator}}
 \newcommand{\TargetRetriever}{\texttt{TargetRetriever}}
-\newcommand{\ExecutionTarget}{\texttt{ExecutionTarget}}
-\newcommand{\Broker}{\texttt{Broker}}
 \newcommand{\Submitter}{\texttt{Submitter}}
-\newcommand{\JobDescription}{\texttt{JobDescription}}
-\newcommand{\JobSupervisor}{\texttt{JobSupervisor}}
-\newcommand{\JobController}{\texttt{JobController}}
-\newcommand{\Job}{\texttt{Job}}
+\newcommand{\URL}{\texttt{URL}}
 \newcommand{\UserConfig}{\texttt{UserConfig}}
-\newcommand{\ClientInterface}{\texttt{ClientInterface}}
-\newcommand{\URL}{\texttt{URL}}
 \newcommand{\XML}{\texttt{XML}}
-\newcommand{\Loader}{\texttt{Loader}}
-\newcommand{\ACCConfig}{\texttt{ACCConfig}}
-\newcommand{\Logger}{\texttt{Logger}}
 \newcommand{\XMLNode}{\texttt{XMLNode}}
@@
 %%%%---- The title ----
-{\centering \textsc{\Large ARC1lib, A Client Library for ARC1}\Large \par}
+{\centering \textsc{\Large {\arclib}}\Large \par}
 \vspace*{0.5cm}
@@
 %%%%---- A subtitle, if necessary ----
-{\centering \textit{\large Paper subtitle}\large \par}
+{\centering \textit{\large A Client Library for ARC}\large \par}
 
@@
 
-This document describes from a technical viewpoint a plugin based client library for the new 
-Web Service (WS) based Advanced Resource Connector~\cite{arc} ARC middleware. The library consists of 
-a set of C++ classes for
+This document describes from a technical viewpoint a plugin based
+client library for the new Web Service (WS) based Advanced Resource
+Connector~\cite{arc} (ARC) middleware. The library consists of a set
+of C++ classes for
 
@@
 
-All capabilities are enabled for three different grid flavours (Production ARC, ARC1 and gLite~\cite{glite}) 
-through a modular design using plugins specialized for each supported middleware. Future 
-extensions to support additional middlewares involve plugin compilation only i.e. no recompilation 
-of main libraries or client.
+All capabilities are enabled for three different grid flavours
+(Production ARC, ARC1 and gLite~\cite{glite}) through a modular design
+using plugins specialized for each supported middleware. Future
+extensions to support additional middlewares involve plugin
+compilation only i.e.\ no recompilation of main libraries or clients
+is necessary.
 
-Using the library a set of command line tools have been built which puts the library's functionality 
-at the fingertips of users. While this documentation will illustrate how such command line tool can be 
-built, the main documentation of the command line tools is given in the client user manual~\cite{ui}.
+Using the library, a set of command line tools have been built which
+puts the library's functionality at the fingertips of users. While
+this documentation will illustrate how such command line tools can be
+built, the main documentation of the command line tools is given in
+the client user manual~\cite{ui}.
 
-In the following we will give a functionality overview in Section~\ref{sec:FuncOver} while all technical 
-details will be given in Section~\ref{sec:Implementation}. Section~\ref{sec:cli} will through examples 
-show how command line interfaces can be built upon the library.
+In the following we will give a functionality overview in
+Section~\ref{sec:FuncOver} while all technical details will be given
+in Section~\ref{sec:Implementation}. Section~\ref{sec:cli} will
+through examples show how command line interfaces can be built upon
+the library.
 
@@
 \label{sec:FuncOver}
-The new ARC1lib makes extensive use of plugins for command handling. These plugins are handled by a set 
-of higher level classes which thus are the ones to offer the plugin functionality to external calls. In 
-this section an overview of the library's main functionality is given which also introduces the most 
-important higher level classes.
+
+The new {\arclib} makes extensive use of plugins for command
+handling. These plugins are handled by a set of higher level classes
+which thus are the ones to offer the plugin functionality to external
+calls. In this section an overview of the library's main functionality
+is given which also introduces the most important higher level
+classes.
+
 \section{Resource Discovery and Information Retrieval}
 \label{sec:TargetDiscovery}
-With the increasing number of grid clusters around the world, a reliable and fast resource discovery and 
-information retrieval capability is of crucial importance for a user interface. The new ARC1lib resource 
-discovery and information retrieval component consists of three classes; The {\TargetGenerator}, the 
-{\TargetRetriever} and the {\ExecutionTarget}. Of these the {\TargetRetriever} is a base class for further grid 
-middleware specific specialization (plugin).
 
-Figure \ref{fig:ResDisc} depicts how the classes work together in a command chain to discover all resources 
-registered with a certain information server. Below a description of each step is given:
+With the increasing number of grid clusters around the world, a
+reliable and fast resource discovery and information retrieval
+capability is of crucial importance for a user interface. The new
+{\arclib} resource discovery and information retrieval component
+consists of three classes; the {\TargetGenerator}, the
+{\TargetRetriever} and the {\ExecutionTarget}. Of these the
+{\TargetRetriever} is a base class for further grid middleware
+specific specialization (plugin).
+
+Figure \ref{fig:ResDisc} depicts how the classes work together in a
+command chain to discover all resources registered with a certain
+information server. Below a description of each step is given:
 
@@
 \centering{{\scalebox{0.75}{\includegraphics{TargetDiscovery.png}}}
-\caption{\label{fig:ResDisc}Diagram depicting the resource discovery and information retrieval process} }
+\caption{\label{fig:ResDisc}Diagram depicting the resource discovery
+  and information retrieval process}}
 \end{figure}
@@
 \begin{enumerate}
-\item{The {\TargetGenerator} takes three {\URL} arrays as input. First array contains individually selected computing services, while the 
-second array contains individually selected index servers (1a). The last array contains a list of rejected services. This 
-list is parsed through alias resolve before being inserted as filter to the storage of found services (1b).}
-\item{The complete list of {\URL}s pointing to computing services and index servers.}
-\item{For each {\URL} a {\TargetRetriever} plugin is loaded using the ARC loader. The {\TargetRetriever} is initialized 
-with the {\URL} and the information about this {\URL} pointing to a computing service or an index server.}
-\item{An external call is received calling for targets to be prepared. Call is propagated to all loaded {\TargetRetriever}s.}
-\item{The call for targets is processed by each {\TargetRetriever} in parallel. If the {\TargetRetriever} is initialized with an 
-index server {\URL} it first registers at the Index Servers store kept by the {\TargetGenerator}. If allowed to register, the index 
-server is queried and the query result processed.}
-\item{The TargetRetriver finds a registered service and contacts the Found Services store kept by the {\TargetGenerator} to check 
-if the service in question has been found be another {\TargetRetriever}. (Services often registers at more than one index server, 
-thus different {\TargetRetriever}s may discover the same service.)}
-\item{If the {\TargetGenerator} allows registration in the Found Services store, the {\TargetRetriever} progresses to interrogate the service.}
-\item{Found and processed {\ExecutionTarget}s are stored in the Found Targets store kept by the {\TargetGenerator} for later usage 
-(e.g. status printing or job submission).}
-\item{Index servers often registers with other index servers creating a nested hierarchical information structure. For this 
-reason a {\TargetRetriever} processing the query result from an index server may encounter other index servers. In such
-events the {\TargetRetriever} will call itself recursively creating a new thread for the new {\URL}. This ensures optimum speed and 
-performance for resource discovery.}
+\item{The {\TargetGenerator} takes three arguments as input. The first
+  argument is a reference to a {\UserConfig} object containing a
+  representing the contents of the user's configuration file. The
+  second and third arguments contain lists of strings. The first list
+  contains individually selected and rejected computing services,
+  while the second list contains individually selected and rejected
+  index servers. Rejected servers and services are identified by that
+  its name is prefixed by a minus sign in the lists. The name of the
+  servers and services should be given either in the form of an alias
+  defined in the {\UserConfig} object or as the name of its grid
+  flavour followed by a colon and the URL of its information contact
+  endpoint.}
+\item{These lists are parsed through alias resolution before being
+  used to initialize the complete list of selected and rejected
+  {\URL}s pointing to computing services and index servers.}
+\item{An external call is received calling for targets to be
+  prepared.}
+\item{For each selected index server and computing service a
+  {\TargetRetriever} plugin for the server's or service's grid flavour
+  is loaded using the ARC loader. The {\TargetRetriever} is
+  initialized with its {\URL} and the information about whether it
+  represents a computing service or an index server. The call for
+  targets is processed by each {\TargetRetriever} in parallel.}
+\item{A {\TargetRetriever} representing an index server first tries to
+  register at the index server store kept by the {\TargetGenerator}.
+  If allowed to register, the index server is queried and the query
+  result processed. The {\TargetGenerator} will not allow
+  registrations from index servers present in its list of rejected
+  index servers or from servers that have already registered
+  once. Index servers often registers at more than one index server,
+  thus different {\TargetRetriever} may discover the same server.)}
+\item{If while processing the query result the {\TargetRetriever}
+  finds an other registered index server or a registered computing
+  service it creates an a new {\TargetRetriever} for the found server
+  or service and forwards the call for targets to the new
+  {\TargetRetriever}.}
+\item{A {\TargetGenerator} representing a computing service first
+  tries to register at the service store kept by the
+  {\TargetGenerator}. If allowed to register, the computing server is
+  queried and the query result processed. The {\TargetGenerator} will
+  not allow registrations from computing services present in its list
+  of rejected computing services or from service that have already
+  registered once. Computing and services often registers at more than
+  one index server, thus different {\TargetRetriever} may discover the
+  same service.)}
+\item{When processing the query result the {\TargetRetriever} will
+  create an {\ExecutionTarget} for each queue found on the computing
+  service and collect all possible information about them. It will
+  then store the {\ExecutionTarget} in the found targets store kept
+  by the {\TargetGenerator} for later usage (e.g. status printing or
+  job submission).}
 \end{enumerate}
-
-The example above outlines how the {\TargetRetriever} works when initialized with an {\URL} pointing to and index server. However, 
-as indicated it also accepts initialization with a service {\URL}s, in which case the {\TargetRetriever} will try to register the 
-service in the {\TargetGenerator} Found Services store. If allowed the {\TargetRetriever} will progress with the target interrogation 
-method.
 
@@
 \label{sec:JobSubmission}
-Job submission starts with the resource discovery and target preparation as outlined in the Section~\ref{sec:TargetDiscovery}. 
-Not until a list of possible targets (which allows the user) is available is the job description read in order to enable bulk 
-job submission of widely different jobs without having to reperform the resource discovery. In addition to the classes mentioned 
-above the job submission makes use of the {\Broker}, {\JobDescription} and {\Submitter} classes. At the time of writing the {\Broker} is 
-not implemented, but it's functionality and interface is well known, and thus we will include it in the overview given here.
-The {\Submitter} is base class for further grid middleware specific specialization (plugin) similarly to the {\TargetRetriever}.
+
+Job submission starts with the resource discovery and target
+preparation as outlined in the Section~\ref{sec:TargetDiscovery}.  Not
+until a list of possible targets (which allows the user) is available
+is the job description read in order to enable bulk job submission of
+widely different jobs without having to reperform the resource
+discovery. In addition to the classes mentioned above the job
+submission makes use of the {\Broker}, {\JobDescription} and
+{\Submitter} classes. At the time of writing the {\Broker} is not
+implemented, but it's functionality and interface is well known, and
+thus we will include it in the overview given here.  The {\Submitter}
+is base class for further grid middleware specific specialization
+(plugin) similarly to the {\TargetRetriever}.
 
@@
 \centering{{\scalebox{0.75}{\includegraphics{JobSubmission.png}}}
-\caption{\label{fig:JobSub}Diagram depicting the submission of a job to a computing service.}}
+\caption{\label{fig:JobSub}Diagram depicting the submission of a job
+  to a computing service.}}
 \end{figure}
 
-Figure \ref{fig:JobSub} shows a job submission sequence and below a description of each step is given:
+Figure \ref{fig:JobSub} shows a job submission sequence and below a
+description of each step is given:
 
 \begin{enumerate}
-\item{The {\TargetGenerator} has prepared a list of {\ExecutionTarget}s. Depending on the {\URL}s provided to the {\TargetGenerator} the list of 
-found {\ExecutionTarget}s may be empty or contain several targets. Targets may even represent more than one grid flavour. The list of 
-found targets are given as input to the {\Broker}.}
-\item{In order to rank the found services ({\ExecutionTarget}s) the {\Broker} needs detailed knowledge about the job requirements, thus the 
-{\JobDescription} is passed as input to the brokering process.}
-\item{The {\Broker} outputs a ordered list of {\ExecutionTarget}s according to the provided {\JobDescription}.}
-\item{Each ExectionTarget has a method to return a specialized {\Submitter} which is capable of submitting jobs to the service it 
-represents. The best suitable {\ExecutionTarget} for the job is asked to return a {\Submitter} for job submission.}
-\item{The {\Submitter} takes the {\JobDescription} as input and uploads it to the computing service.}
-\item{The {\Submitter} identifies local input files from the {\JobDescription} and uploads them to the computing service.}
+\item{The {\TargetGenerator} has prepared a list of
+  {\ExecutionTarget}s. Depending on the {\URL}s provided to the
+  {\TargetGenerator} the list of found {\ExecutionTarget}s may be
+  empty or contain several targets. Targets may even represent more
+  than one grid flavour. The list of found targets are given as input
+  to the {\Broker}.}
+\item{In order to rank the found services ({\ExecutionTarget}s) the
+  {\Broker} needs detailed knowledge about the job requirements, thus
+  the {\JobDescription} is passed as input to the brokering process.}
+\item{The {\Broker} outputs a ordered list of {\ExecutionTarget}s
+  according to the provided {\JobDescription}.}
+\item{Each ExectionTarget has a method to return a specialized
+  {\Submitter} which is capable of submitting jobs to the service it
+  represents. The best suitable {\ExecutionTarget} for the job is
+  asked to return a {\Submitter} for job submission.}
+\item{The {\Submitter} takes the {\JobDescription} as input and
+  uploads it to the computing service.}
+\item{The {\Submitter} identifies local input files from the
+  {\JobDescription} and uploads them to the computing service.}
 \end{enumerate}
@@
 \section{Job Management}
-Once a job is submitted job related information (job identification string, cluster etc) is stored in a local xml file which hosts 
-similar information for all active jobs. This file may contain jobs running on completely different grid flavour, and thus job 
-management should be handled using plugins similar to resource discovery and job submission. The job managing plugin is called 
-the {\JobController} and it is supported by the {\JobSupervisor} and {\Job} classes.
+
+Once a job is submitted job related information (job identification
+string, cluster etc) is stored in a local xml file which hosts similar
+information for all active jobs. This file may contain jobs running on
+completely different grid flavours, and thus job management should be
+handled using plugins similar to resource discovery and job
+submission. The job managing plugin is called the {\JobController} and
+it is supported by the {\JobSupervisor} and {\Job} classes.
 
@@
 \centering{{\scalebox{0.75}{\includegraphics{JobManagement.png}}}
-\caption{\label{fig:JobMan}Diagram depicting how job controlling plugins, {\JobController}s, are loaded and initialized.}}
+\caption{\label{fig:JobMan}Diagram depicting how job controlling
+  plugins, {\JobController}s, are loaded and initialized.}}
 \end{figure}
 
-Figure~\ref{fig:JobMan} shows how the three different classes work together and below a step by step description is given:
+Figure~\ref{fig:JobMan} shows how the three different classes work
+together and below a step by step description is given:
 
 \begin{enumerate}
-\item{The {\JobSupervisor} takes three arguments Jobs (job identification strings), ClusterSelect and ClusterReject.}
-\item{A job identification string does not uniquely define which grid flavour runs a certain job. Thus this information is stored 
-in joblist.xml upon submission by the {\Submitter} and joblist.xml is extensively used by the {\JobSupervisor} to identify the 
-{\JobController} flavours which are to be loaded.}
+\item{The {\JobSupervisor} takes four arguments as input. The first
+  argument is a reference to a {\UserConfig} object containing a
+  representing the contents of the user's configuration file. The
+  second is a list of strings containing job identifiers and job
+  names, the third is a list of strings of clusters to select or
+  reject (in the same format as described for the {\TargetGenerator}
+  above). The last argument is the name of the file containing the
+  local information about active jobs, hereafter called the joblist
+  file.}
+\item{A job identifier does not uniquely define which grid flavour
+  runs a certain job. Thus this information is stored in the joblist
+  file upon submission by the {\Submitter} and the joblist file is
+  extensively used by the {\JobSupervisor} to identify the
+  {\JobController} flavours which are to be loaded. The information in
+  the joblist file is also used to lookup the job identifier for jobs
+  specified using job names. Alias resolution for the selected and
+  rejected clusters are performed using the information in the
+  {\UserConfig} object.}
 \item{Suitable {\JobController}s are loaded}
-\item{Jobs, ClusterSelect and ClusterReject are passed to the {\JobController} which uses the information to 
+\item{The list of job identifiers and selected and rejected clusters
+  are passed to each {\JobController} which uses the information to
   fill its internal JobStore.}
-\item{Residing within the {\JobSupervisor} the {\JobController}s are now accessible for external calls (i.e. job handling).}
+\item{Residing within the {\JobSupervisor} the {\JobController}s are
+  now accessible for external calls (i.e. job handling).}
 \end{enumerate}
@@
 \label{sec:Implementation}
-Clever text here.
+
+In this section an overview of all important classes in the new
+{\arclib} is given. The overview is subdivided in two parts where
+first all the generic classes are presented,
+Section~\ref{sec:genclass}, before the grid flavour specializations
+are presented in Section~\ref{sec:plugins}. Classes are described both
+in words and by code.
+
 \section{Generic Classes}
+\label{sec:genclass}
+
 \subsection{{\ACC}}
-The Arc Client Component ({\ACC}) class is a completely empty base class needed for the {\Loader} 
-in order to create loadable classes.
+
+The Arc Client Component ({\ACC}) class is a base class needed for the
+{\Loader} in order to create loadable classes. It stored information
+about which flavour it supports and security related information
+needed by all ACC specializations.
 
@@
 protected:
-  ACC() {}
+  ACC(Config *cfg, const std::string& flavour);
 public:
-  virtual ~ACC() {}
+  virtual ~ACC();
+  const std::string& Flavour();
+protected:
+  std::string flavour;
+  std::string proxyPath;
+  std::string certificatePath;
+  std::string keyPath;
+  std::string caCertificatesDir;
 };
@@
 
-\subsection{{\TargetGenerator}} The {\TargetGenerator} class is the main class for resource discovery and information retrieval. It loads 
-{\TargetRetriever} plugins in accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed} (e.g. if an {\URL} pointing to an ARC1 
-resource is given the TargetRetrieverARC1 is loaded).
+\subsection{{\TargetGenerator}}
 
-To perform a resource discovery, first construct a {\TargetGenerator} object using the {\TargetGenerator} constructor,
+ The {\TargetGenerator} class is the main class for resource discovery
+ and information retrieval. It loads {\TargetRetriever} plugins in
+ accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed}
+ (e.g. if an {\URL} pointing to an ARC1 resource is given the
+ TargetRetrieverARC1 is loaded).
+
+To perform a resource discovery, first construct a {\TargetGenerator}
+object using the {\TargetGenerator} constructor,
 
@@
 \begin{verbatim}
-TargetGenerator(const std::list<std::string>& clusterselect,
-                const std::list<std::string>& clusterreject,
+    TargetGenerator(const UserConfig& usercfg,
+		    const std::list<std::string>& clusters,
 		    const std::list<std::string>& indexurls);
@@
 
-where \texttt{clusterselect}, \texttt{clusterreject} and \texttt{indexurls} are lists of strings where each string is either 
-an {\URL} or a regular string defined as an alias in the user (or system) configuration. Without any further explanation it is 
-hard to decode from an {\URL} if it points to a index server or computing service of a certain grid flavour. Thus in order for 
-the {\TargetGenerator} to know which {\TargetRetriever} should be loaded {\URL}s have to be given in a format where the grid 
-flavour precedes the actual {\URL}, e.g:
+where \texttt{clusters}, \texttt{indexurls} are lists of strings where
+each string is either the name of a grid flavour followed by a colon
+and the URL of an information contact endpoint for a computing service
+or index server, respectively, of that grid flavour or an alias
+defined in the {\UserConfig} object passed as the first
+argument, e.g:
 
@@
 \begin{verbatim}
-ARC0:ldap://grid.tsl.uu.se:2135/mds-vo-name=sweden,O=grid
+ARC0:ldap://grid.tsl.uu.se:2135/Mds-Vo-name=Sweden,o=grid
 ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid
@@
 
-where the former is the {\URL} to an index server while the latter is an {\URL} to a computing service.
+where the former is the URL to an index server while the latter is an
+URL to a computing service.
 
-To prepare a list of {\ExecutionTarget}s use the {\TargetGenerator} object and invoke its method
+To prepare a list of {\ExecutionTarget}s use the {\TargetGenerator}
+object and invoke its method
 
@@
 
-The {\TargetGenerator} will pass this request to the loaded {\TargetRetriever}s running them as individual threads 
-for improved performance. The {\TargetGenerator} keeps records of by {\TargetRetriever} found index servers and computing 
-services in order to avoid multiple identical queries. Accepted {\ExecutionTarget}s are stored in the FoundTargets 
-array kept by the {\TargetGenerator}. See also Section~\ref{sec:TargetDiscovery} for a schematic drawing of the 
+The {\TargetGenerator} will pass this request to the loaded
+{\TargetRetriever}s running them as individual threads for improved
+performance. The {\TargetGenerator} keeps records of index servers and
+computing services found by the {\TargetRetriever}s in order to avoid
+multiple identical queries. Accepted {\ExecutionTarget}s are stored in
+the FoundTargets array kept by the {\TargetGenerator}. See also
+Section~\ref{sec:TargetDiscovery} for a schematic drawing of the
 resource discovery process.
@@
 
-\subsection{{\TargetRetriever}} The {\TargetRetriever} is base class for {\TargetRetriever} grid middleware specializations and 
-inherits from the \texttt{ACC} base class in order to be loadable by the ARC \texttt{Loader}. It is designed to work on conjunction 
-with the {\TargetGenerator} contains and contains the pure virtual method
+\subsection{{\TargetRetriever}}
+
+The {\TargetRetriever} is base class for {\TargetRetriever} grid
+middleware specializations and inherits from the \texttt{ACC} base
+class in order to be loadable by the ARC \texttt{Loader}. It is
+designed to work in conjunction with the {\TargetGenerator} and
+contains the pure virtual method
 
@@
 
-which is to be implemented by the specialized class. While it is not mandatory it is recommended that the specialized class 
-divides this method into two components: QueryIndex and InterrogateTarget. The former is handles index server queries, and the 
-latter the computing service queries and {\ExecutionTarget} preparation.
+which is to be implemented by the specialized class. While it is not
+mandatory it is recommended that the specialized class divides this
+method into two components: QueryIndex and InterrogateTarget. The
+former handles index server queries, and the latter the computing
+service queries and {\ExecutionTarget} preparation.
 
-If an index server query yields an url to a different index server then the one queried, then the {\TargetRetriever} should 
-call itself recursively creating a new {\TargetRetriever} for the discovered index server.
+If an index server query yields an url to a different index server
+than the one queried, then the {\TargetRetriever} should call itself
+recursively creating a new {\TargetRetriever} for the discovered index
+server.
 
-\subsection{{\ExecutionTarget}} The {\ExecutionTarget} is the class representation of a computing resource (queue) capable of 
-executing a grid job. It serves as input to the {\Broker} which is foreseen to be able to select between different 
-{\ExecutionTarget}s from different grid flavours without apriori knowing their difference. The {\ExecutionTarget} class 
-mimics the glue2 information model (with a flattened structure), and thus a mapping between attributes from other information 
-systems into the glue2 format is needed.  Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares.
+\subsection{{\ExecutionTarget}}
+
+The {\ExecutionTarget} is the class representation of a computing
+resource (queue) capable of executing a grid job. It serves as input
+to the {\Broker} which is foreseen to be able to select between
+different {\ExecutionTarget}s from different grid flavours without a
+priori knowing their difference. The {\ExecutionTarget} class mimics
+the glue2 information model (with a flattened structure), and thus a
+mapping between attributes from other information systems into the
+glue2 format is needed. Appendix XX shows the current mapping for the
+production ARC, ARC1 and gLite middlewares.
 
@@
 
-Following a broker decision jobs are to be submitted. Since all information about the selected computing service resides within the 
-selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of returning a {\Submitter} capable of submitting a 
-job to the service which is represents.
+Following a broker decision jobs are to be submitted. Since all
+information about the selected computing service resides within the
+selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of
+returning a {\Submitter} capable of submitting a job to the service
+which is represents.
 
@@
 \begin{verbatim}
-Submitter *GetSubmitter() const;
+Submitter *GetSubmitter(const UserConfig& ucfg) const;
 \end{verbatim}
@@
 
-\subsection{{\Broker}} The {\Broker} is not yet implemented and thus not described here.
+The {\UserConfig} object is used to configure the security related
+information needed by the {\Submitter}.
 
-\subsection{{\JobDescription}} When addressing interoperability it is of paramount importance to transparently address grid job 
-descriptions written in different job description languages by translating them automatically. In the ARC1lib library this functionality 
-is implemented in the {\JobDescription} class. This is a generic class that takes a job description (string) as input
+\subsection{{\Broker}}
+
+The {\Broker} is not yet implemented and thus not described here.
+
+\subsection{{\JobDescription}}
+
+When addressing interoperability it is of paramount importance to
+transparently address grid job descriptions written in different job
+description languages by translating them automatically. In the
+{\arclib} library this functionality is implemented in the
+{\JobDescription} class. This is a generic class that takes a job
+description (string) as input
 
@@
 \begin{verbatim}
-setSource( const std::string source ) throw(JobDescriptionError);
+void setSource(const std::string source);
 \end{verbatim}
@@
 
-in any supported formats (currently xRSL, JDL, JSDL), converts and stores it in a JSDL-like internal job description format. The identification 
-of the source's description format is internally handled by a \texttt{JobDescriptionOrderer} class which identifies the format by pattern
-matching.
+in any supported formats (currently xRSL~\cite{xrsl}, JDL~\cite{jdl},
+JSDL~\cite{jsdl}), converts and stores it in a JSDL-like internal job
+description format. The identification of the source's description
+format is internally handled by a \texttt{JobDescriptionOrderer} class
+which identifies the format by pattern matching.
 
-Different operations, like getting the job description in other formats or getting job-related information, can be performed using the 
+Different operations, like getting the job description in other
+formats or getting job-related information, can be performed using the
 description-independent functions of this class, e.g.:
@@
 \begin{verbatim}
-getProduct(std::string& product, std::string format="JSDL") throw(JobDescriptionError);
+void getProduct(std::string& product, std::string format = "JSDL")
 \end{verbatim}
@@
 
-The {\JobDescription} class has tree back-end classes, sometimes referred to as back-end or translator modules, corresponding to the three 
-supported job description languages. The {\JobDescription} class chooses the appropriate back-end module according to the pattern match 
-performed by the \texttt{JobDescriptionOrderer}, and uses the back-end module for parsing and generating the job descriptions.
+The {\JobDescription} class has tree back-end classes, sometimes
+referred to as back-end or translator modules, corresponding to the
+three supported job description languages. The {\JobDescription} class
+chooses the appropriate back-end module according to the pattern match
+performed by the \texttt{JobDescriptionOrderer}, and uses the back-end
+module for parsing and generating the job descriptions.
 
-\subsection{{\Submitter}} {\Submitter} is base class for grid specific specializations (plugin). It submits job(s) to the service 
-it represents and uploads (by the job needed) local input files.
+\subsection{{\Submitter}}
+
+{\Submitter} is base class for grid specific specializations (plugin).
+It submits job(s) to the service it represents and uploads (by the job
+needed) local input files.
 
@@
 
-The \texttt{Submit} method fills the {\XML} node \texttt{info} with all needed information about the job for later job management. 
-The {\Submitter} is returned by the {\ExecutionTarget} selected for job execution and thus the {\ExecutionTarget} populates 
-(through {\XML} config element) the {\Submitter} with information about submission endpoint ({\URL}) and job description languages 
-``spoken'' by the target.
+The \texttt{Submit} method fills the {\XMLNode} \texttt{info} with all
+needed information about the job for later job management. The
+{\Submitter} is returned by the {\ExecutionTarget} selected for job
+execution and thus the {\ExecutionTarget} populates (through its
+{\XMLNode} config element) the {\Submitter} with information about
+submission endpoint ({\URL}) and job description languages understood
+by the target.
 
-\subsection{{\JobSupervisor}} The {\JobSupervisor} is responsible for loading the appropriate {\JobController}s for managing 
-jobs running on a certain grid flavour. Job manipulation can be performed either on individual jobs or on groups of jobs (e.g. 
-all jobs running on certain cluster or all jobs with job state ``FINISHED''), and in order to translate the by user given 
-information into a set of loadable {\JobController}s the {\JobSupervisor} makes extensive use of the local {\XML} file housing 
-information about all active jobs. Thus the {\JobSupervisor} constructor becomes
+\subsection{{\JobSupervisor}}
+
+The {\JobSupervisor} is responsible for loading the appropriate
+{\JobController}s for managing jobs running on a certain grid
+flavour. Job manipulation can be performed either on individual jobs
+or on groups of jobs (e.g.\ all jobs running on certain cluster or all
+jobs with job state ``FINISHED''), and in order to translate the
+information given by the user into a set of loadable {\JobController}s
+the {\JobSupervisor} makes extensive use of the local joblist file
+housing information about all active jobs. Thus the {\JobSupervisor}
+constructor becomes
 
@@
 \begin{verbatim}
-JobSupervisor(const std::list<std::string>& jobs,
-              const std::list<std::string>& clusterselect,
-              const std::list<std::string>& clusterreject,
-              const std::string joblist);
+JobSupervisor(const UserConfig& usercfg,
+              const std::list<std::string>& jobs,
+              const std::list<std::string>& clusters,
+              const std::string& joblist);
 \end{verbatim}
@@
 
-where \texttt{jobs} is a list of job identification strings (jobids), \texttt{clusterselect} is a list of selected computing services, 
-\texttt{clusterreject} is a list of rejected computing services and \texttt{joblist} is a string pointing to the local {\XML} file 
-with job information. Syntax for \texttt{clusterselect} and \texttt{clusterreject} is identical to the one used by the {\TargetGenerator}
-for for internal consistency.
+where \texttt{jobs} is a list of job identifiers and job names,
+\texttt{clusters} is a list of selected and rejected computing
+services in the same format as described above for the
+{\TargetGenerator}, \texttt{joblist} is a string containing the name
+of the joblist file.
 
-Although being loaded by the {\JobSupervisor} the {\JobController} objects truly resides within the ARC {\Loader} which is a member 
-of the {\JobSupervisor} class. In order to get handles on the {\JobController}s the inline method
+Although being loaded by the {\JobSupervisor} the {\JobController}
+objects truly resides within the ARC {\Loader} which is a member of
+the {\JobSupervisor} class. In order to get handles on the
+{\JobController}s the inline method
 
@@
 \begin{verbatim}
-std::list<Arc::JobController*> GetJobControllers(){return JobControllers;}
+const std::list<Arc::JobController*>& GetJobControllers();
 \end{verbatim}
@@
 
-\subsection{{\JobController}} The {\JobController} is both base class for grid specific specializations, but also the implementor of all 
-public functionality offered by the {\JobController}s. In other words all virtual functions of the {\JobController} are private. The 
-initialization of a (specialized) {\JobController} object takes two steps. First the constructor has to invoked
+\subsection{{\JobController}}
+
+The {\JobController} is a base class for grid specific
+specializations, but also the implementor of all public functionality
+offered by the {\JobController}s. In other words all virtual functions
+of the {\JobController} are private. The initialization of a
+(specialized) {\JobController} object takes two steps. First the
+{\JobController} specialization for the required grid flavour must be
+loaded by the ARC {\Loader}, which sees to that the {\JobController}
+receives information about its flavour (grid) and the local joblist
+file containing information about all active jobs (flavour
+independent). Next step is the filling of the {\JobController}'s job
+pool \texttt{JobStore} which is the pool of jobs that the
+{\JobController} can manage.
 
@@
 \begin{verbatim}
-JobController(Arc::Config *cfg, std::string flavour);
+void FillJobStore(const std::list<URL>& jobids,
+                  const std::list<URL>& clusterselect,
+                  const std::list<URL>& cluterreject);
 \end{verbatim}
@@
 
-This is usually done by the ARC {\Loader} which sees to that the {\JobController} receives information about its flavour (grid) and 
-the local {\XML} file containing information about all active jobs (flavour independent). Next step is the filling of the {\JobController}'s
-job pool \texttt{JobStore} which is the pool of jobs that the {\JobController} can manage.
-
-\begin{shaded}
-\begin{verbatim}
-void FillJobStore(std::list<std::string> jobs,
-                  std::list<std::string> clusterselect,
-                  std::list<std::string> cluterreject);
-\end{verbatim}
-\end{shaded}
-
-Here \texttt{jobs}, \texttt{clusterselect} and \texttt{clusterreject} have the same interpretation as in the {\JobSupervisor} constructor. 
-However, while the latter only had to identify which {\JobController}s should be loaded, the {\JobController} itself has to decode the 
-supplied lists into a list of jobs with non-overlapping entries which will be filled into the JobStore. Due to possible conflicts 
-within the supplied lists, rules have to be set up for dealing with contradictions:
+Here \texttt{jobids}, \texttt{clusterselect} and
+\texttt{clusterreject} have been resolved for job names and aliases by
+the {\JobSupervisor}, and no further resolution is needed. The
+following rules are observed when filling the \texttt{JobStore}:
 
 \begin{enumerate}
-\item{If the \texttt{jobs} list has entries, fill \texttt{JobStore} with the by user requested jobs.}
-\item{If the \texttt{clusterselect} list has entries, fill \texttt{JobStore} with the jobs running on the selected clusters.}
-\item{If clusters are rejected and \texttt{JobStore} has entries, remove from \texttt{JobStore} all jobs running on rejected clusters.}
-\item{If \texttt{jobs} and \texttt{clusterselect} are both empty lists, fill \texttt{JobStore} with all jobs except those 
-running on possible rejected clusters.}
+\item{If the \texttt{jobids} list has entries, fill \texttt{JobStore}
+  with the by user requested jobs.}
+\item{If the \texttt{clusterselect} list has entries, fill
+  \texttt{JobStore} with the jobs running on the selected clusters.}
+\item{If clusters are rejected and \texttt{JobStore} has entries,
+  remove from \texttt{JobStore} all jobs running on rejected
+  clusters.}
+\item{If \texttt{jobs} and \texttt{clusterselect} are both empty
+  lists, fill \texttt{JobStore} with all jobs except those running on
+  possible rejected clusters.}
 \end{enumerate}
 
-The steps above completes the initialization of the {\JobController} which is now ready for handling jobs. The public functions of the 
-{\JobController} offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations for issuing the command. 
-Here exemplified by the \texttt{Stat} command:
+The steps above completes the initialization of the {\JobController}
+which is now ready for handling jobs. The public functions of the
+{\JobController} offer to get (download), clean, cancel, etc one or
+more jobs and uses the private specializations for issuing the
+command. Here exemplified by the \texttt{Stat} command:
 
@@
 \begin{verbatim}
-void JobController::Stat(const std::list<std::string>& jobs,
-                    const std::list<std::string>& clusterselect,
-                    const std::list<std::string>& clusterreject,
-                    const std::list<std::string>& status,
+bool JobController::Stat(const std::list<std::string>& status,
                          const bool longlist,
-                    const int timeout){
+                         const int timeout) {
 
@@
 
-  //Add here functionality for identifying jobs according to clusterselect, clusterreject & status
-
-  for(std::list<Arc::Job>::iterator jobiter = JobStore.begin();jobiter!= JobStore.end();jobiter++){
-    if(jobiter->State.empty()){
-      std::cout<<Arc::IString("Job state information not found:%s",jobiter->JobID.str())<<std::endl;
+  for (std::list<Job>::iterator it = jobstore.begin();
+       it != jobstore.end(); it++) {
+    if (it->State.empty()) {
+      logger.msg(WARNING, "Job state information not found: %s",
+                 it->JobID.str());
       Time now;
-      if(now - jobiter->LocalSubmissionTime < 90){
-        std::cout<<Arc::IString("This job was very recently "
+      if (now - it->LocalSubmissionTime < 90)
+        logger.msg(WARNING, "This job was very recently "
                    "submitted and might not yet"
-                                "have reached the information-system")<<std::endl;
-      }
+                   "have reached the information-system");
       continue;
     }
-    jobiter->Print(longlist);
+    it->Print(longlist);
   }
+  return true;
 }
@@
 
-The \texttt{Stat} command prints the job information to screen and in order to do so the {\JobController} has to query local information
-server for the latest status. Due to different protocols used for different grid flavours (e.g. ldap for production ARC), the 
-\texttt{GetJobInformation} has to be grid flavour specific and is only declared as a private virtual method within the {\JobController}
-base class. For details about the flavour specific implementations see Section~\ref{sec:plugins}.
+The \texttt{Stat} command prints the job information to screen and in
+order to do so the {\JobController} has to query local information
+server for the latest status. Due to different protocols used for
+different grid flavours (e.g. ldap for production ARC), the
+\texttt{GetJobInformation} has to be grid flavour specific and is only
+declared as a private virtual method within the {\JobController} base
+class. For details about the flavour specific implementations see
+Section~\ref{sec:plugins}.
 
-It should be noted that the \texttt{Stat} example above appears to have several unused variables (arguments). This is an artifact 
-caused by the ongoing development in which the {\JobController} is being prepared for usage in a future graphical user interface (GUI).
-In a GUI it would be desirable to make the {\JobController} a persistent object which exists throughout the lifetime of the GUI and 
-which is capable of monitoring and managing all jobs running on a certain grid flavour. In order to do so the \texttt{JobStore} should 
-become a pool of all jobs, and the public functions should take arguments which identifies which jobs from the pool should be subject 
-to the command. In the present implementation commands are performed on all jobs in \texttt{JobStore}.
+\subsection{{\Job}} The
 
-\subsection{{\Job}} The {\Job} is a generic job class for storing all job related information. Attributes are derived from the 
-glue2 information model and thus a mapping is needed for non glue2 compliant grid middlewares. Appendix XX shows the present 
-mapping schema.
+{\Job} is a generic job class for storing all job related information.
+Attributes are derived from the glue2 information model and thus a
+mapping is needed for non glue2 compliant grid middlewares. Appendix
+XX shows the present mapping schema.
 
@@
 
-\subsection{{\UserConfig}} The {\UserConfig} class handles the client setup i.e. proxy, certificate and key location, user and system 
-configuration and local joblist location. Upon initialization (constructor) the {\UserConfig} locates the user files
+\subsection{{\UserConfig}}
+
+The {\UserConfig} class handles the client setup i.e. proxy,
+certificate and key location, user and system configuration and local
+joblist location. Upon initialization (constructor) the {\UserConfig}
+locates the user files
 
@@
 
-The {\UserConfig} has two main public methods
+The {\UserConfig} has three main public methods
 
@@
 \begin{verbatim}
-const std::string& JobsFile();
-const XMLNode ConfTree();
+const std::string& JobListFile() const;
+const XMLNode& ConfTree() const;
+bool ApplySecurity(XMLNode& ccfg) const;
 \end{verbatim}
@@
 
-where \texttt{JobsFile()} returns the string pointing to the jobs.xml file while \texttt{ConfTree()}
-returns a configuration {\XML} object which is the merge between the user and system configurations.
-In order to do this the method has to locate the system configuration and resolve possible conflicts 
-with the user configuration. This proceeds through the following chain of actions:
+where \texttt{JobListFile()} returns the string pointing to the
+jobs.xml file while \texttt{ConfTree()} returns a configuration
+{\XMLNode} object which is the merge between the user and system
+configurations. In order to do this the method has to locate the
+system configuration and resolve possible conflicts with the user
+configuration. This proceeds through the following chain of actions:
 
 \begin{enumerate}
-\item{Try reading system configuration from \texttt{<ARC Install Location>/etc/arcclient.xml}}
-\item{If the previous step failed try reading system configuration from \texttt{/etc/arcclient.xml}}
-\item{Merge system and user configuration by adding all system configuration not already listed 
-in the user configuration to the latter.}
-\item{Check if \texttt{\$X509\_USER\_PROXY} is set, if set add this as \texttt{ProxyPath} to configuration}
-\item{If \texttt{\$X509\_USER\_PROXY} is not set and configuration tree does not already contain 
-a \texttt{ProxyPath} set \texttt{ProxyPath} to \texttt{/tmp/x509up\_u + userid}.}
-\item{Check if \texttt{\$X509\_USER\_CERT} is set, if set add this as \texttt{CertificatePath} to configuration}
-\item{If \texttt{\$X509\_USER\_CERT} is not set and configuration tree does not already contain 
-a \texttt{CertificatePath} set \texttt{CertificatePath} to \texttt{/etc/grid-security/hostcert.pem} 
-if user is superuser otherwise 
-\texttt{\$HOME/.globus/usercert.pem}.}
-\item{Check if \texttt{\$X509\_USER\_KEY} is set, if set add this as \texttt{KeyPath} to configuration}
-\item{If \texttt{\$X509\_USER\_KEY} is not set and configuration tree does not already contain 
-a \texttt{KeyPath} set \texttt{KeyPath} to \texttt{/etc/grid-security/hostkey.pem} 
-if user is superuser otherwise \texttt{\$HOME/.globus/userkey.pem}.}
-\item{Check if \texttt{\$X509\_CERT\_DIR} is set, if set add this as \texttt{CACertificatesDir} to configuration}
-\item{If \texttt{\$X509\_CERT\_DIR} is not set and configuration tree does not already contain 
-a \texttt{CACertificatesDir} set \texttt{CACertificatesDir} to \texttt{/etc/grid-security/certificates}.}
+\item{Try reading system configuration from \texttt{<ARC Install
+    Location>/etc/arcclient.xml}}
+\item{If the previous step failed try reading system configuration
+  from \texttt{/etc/arcclient.xml}}
+\item{Merge system and user configuration by adding all system
+  configuration not already listed in the user configuration to the
+  latter.}
 \end{enumerate}
 
+The \texttt{ApplySecurity(XMLNode\& ccfg)} adds security tags to the
+ccfg {\XMLNode} as follows:
+
+\begin{itemize}
+\item{If the \texttt{\$X509\_USER\_PROXY} environment variable is set,
+  use its value to define a \texttt{ProxyPath} tag in \texttt{cfg}. If
+  the file does not exist or can't be read exit with an error.}
+\item{Otherwise, if the \texttt{\$X509\_USER\_CERT} environment
+  variable is set, use its value and the value of the
+  \texttt{\$X509\_USER\_KEY} environment variable to define
+  \texttt{CertificatePath} and \texttt{KeyPath} tags in \texttt{cfg}.
+  If \texttt{\$X509\_USER\_KEY} is not set or either file can't be
+  read exit with an error.}
+\item{Otherwise, if the merged user configuration tree (see
+  \texttt{ConfTree} containns a \texttt{ProxyPath} tag, copy it to
+  \texttt{cfg}.  If the file does not exist or can't be read exit with
+  an error.}
+\item{Otherwise, if the merged user configuration tree containns a
+  \texttt{CertificatePath} tag, copy it to \texttt{cfg} along with the
+  accompanying \texttt{KeyPath} tag. If the \texttt{KeyPath} tag is
+  undefiend or either filefile does not exist or can't be read exit
+  with an error.}
+\item{Otherwise, if the file \texttt{/tmp/x509up\_u + userid} exists
+  add a \texttt{ProxyPath} tag in \texttt{cfg} pointing to it.}
+\item{Otherwise, add \texttt{CertificatePath} and \texttt{KeyPath}
+  tags to \texttt{cfg} pointing to
+  \texttt{\$HOME/.globus/usercert.pem} and
+  \texttt{\$HOME/.globus/userkey.pem}, respectively. If either file
+  does not exist exit with error.}
+\item{If the \texttt{\$X509\_CERT\_DIR} environment variable is set,
+  use its value to define a \texttt{CACertificatesDir} tag in
+  \texttt{cfg}. If the directory does not exist, exit with an error.}
+\item{Otherwise, if the user ID is not zero and the directory
+  \texttt{\$HOME/.globus/certificates} exists add a
+  \texttt{CACertificatesDir} tag in \texttt{cfg} pointing to it.}
+\item{Otherwise, add a \texttt{CACertificatesDir} tag in \texttt{cfg}
+  pointing to \texttt{/etc/grid-security/certificates}. If the
+  directory does not exist, exit with an error.}
+\end{itemize}
+
 \subsection{{\ACCConfig}}
-Locating Arc Client Components (plugins) is handled by the {\ACCConfig} class. It inherits from \texttt{BaseConfig} 
-and implements only one method
+
+Locating Arc Client Components (plugins) is handled by the
+{\ACCConfig} class. It inherits from \texttt{BaseConfig} and
+implements only one method
 
@@
 
-The \texttt{MakeConfig} method searches plugin paths for all libraries named \texttt{libacc*}. Matching 
-libraries are added as plugins to the configuration object \texttt{cfg}.
+The \texttt{MakeConfig} method searches plugin paths for all libraries
+named \texttt{libacc*}. Matching libraries are added as plugins to the
+configuration object \texttt{cfg}.
 
 \subsection{{\ClientInterface}}
-\subsubsection{ClientTCP}
-\subsubsection{ClientHTTP}
-\subsubsection{ClientSOAP}
+
+The {\ClientInterface} class is a utility base class used for
+configuring a client side Message Chain Component (MCC) chain and
+loading it into memory. It has several specializations of increasing
+complexity of the MCC chains.
+
+\subsubsection{{\ClientTCP}}
+
+The {\ClientTCP} class is a specialization of the {\ClientInterface}
+which sets up a client MCC chain for TCP communication, and optionally
+with a TLS layer on top.
+
+\subsubsection{{\ClientHTTP}}
+
+The {\ClientHTTP} class inherits from the {\ClientTCP} class and adds
+an HTTP MCC to the chain.
+
+\subsubsection{{\ClientSOAP}}
+
+The {\ClientSOAP} class inherits from the {\ClientHTTP} class and adds
+a SOAP MCC to the chain. Specializations of the {\TargetRetriever},
+{\Submitter} and {\JobController} classes that communicate with SOAP
+based services make use of this class.
 
@@
 \label{sec:plugins}
+
 \subsection{ARC0 Plugins}
+
 The ARC0 plugins enables support for the interfaces used by computing
@@
 
-The ARC 0.x local information system uses the {\GT}~\cite{globus} GRIS with a
-propriety ARC schema. The information index server used is the {\GT}
-GIIS. Both these servers are using the LDAP~\cite{ldap} protocol. The
-specialization of the {\TargetRetriever} class for ARC0 is implemented
-using the ARC LDAP Data Management Component (DMC) (see \cite{hed} for 
-technical details).
+The ARC 0.x local information system uses the {\GT}~\cite{globus} GRIS
+with a propriety ARC schema. The information index server used is the
+{\GT} GIIS. Both these servers are using the LDAP~\cite{ldap}
+protocol. The specialization of the {\TargetRetriever} class for ARC0
+is implemented using the ARC LDAP Data Management Component (DMC) (see
+\cite{hed} for technical details).
 
 Jobs running on an ARC 0.x computing element are handled by the ARC
-grid-manager~\cite{gm}. Job submission and job control are done using the
-gridftp~\cite{gridftp} protocol. The specializations of the {\Submitter} and
-{\JobController} classes use the globus ftp control library.
+grid-manager~\cite{gm}. Job submission and job control are done using
+the gridftp~\cite{gridftp} protocol. The specializations of the
+{\Submitter} and {\JobController} classes use the globus ftp control
+library.
 
 Stage-in and stage-out of input and output files are also done using
-the gridftp~\cite{gridftp} protocol. This means that proper functionality of the ARC0
-plugins requires the gridftp DMC.
+the gridftp~\cite{gridftp} protocol. This means that proper
+functionality of the ARC0 plugins requires the gridftp DMC.
 
 \subsection{ARC1 Plugins}
-The computing element in ARC version 1.x is the A-Rex~\cite{arex} service running
-in a HED~\cite{hed} container.
 
-A-Rex implements the BES~\cite{ogsa-bes} standard interface. Since this is a SOAP~\cite{soap}
-based interface the specializations of the {\TargetRetriever},
-{\Submitter} and {\JobController} classes make use of a chain of ARC Message Chain Components 
-(MCC~\cite{hed}) ending with the SOAP client MCC.
+The computing element in ARC version 1.x is the A-Rex~\cite{arex}
+service running in a HED~\cite{hed} container.
+
+A-Rex implements the BES~\cite{ogsa-bes} standard interface. Since
+this is a SOAP~\cite{soap} based interface the specializations of the
+{\TargetRetriever}, {\Submitter} and {\JobController} classes make use
+of a chain of ARC Message Chain Components (MCC~\cite{hed}) ending
+with the SOAP client MCC.
 
@@
 \subsection{gLite Plugins}
-The gLite computing element offers several interfaces, one of them being the Web Service based 
-computing element interface known as the CREAM CE~\cite{cream}. The current revision of this 
-interface (CREAM version 2) has been chosen for implementation within ARC1lib for the following 
-reasons:
+
+The gLite computing element offers several interfaces, one of them
+being the Web Service based computing element interface known as the
+CREAM CE~\cite{cream}. The current revision of this interface (CREAM
+version 2) has been chosen for implementation within {\arclib} for the
+following reasons:
 
 \begin{itemize}
-\item CREAM2 has a Web Service interface that fits the Web Service based ARC.
-\item CREAM2 enables direct access to the gLite computing element without having to go via the gLite workload management system.
-\item CREAM2 contains numerous improvements when compared to the earlier CREAM versions.
+\item CREAM2 has a Web Service interface that fits the Web Service
+  based ARC.
+\item CREAM2 enables direct access to the gLite computing element
+  without having to go via the gLite workload management system.
+\item CREAM2 contains numerous improvements when compared to the
+  earlier CREAM versions.
 \item CREAM2 supports direct job status queries.
-\item CREAM2 offers a convenient way of handling input and output files through accessing the input and output sandbox via GridFTP.
+\item CREAM2 offers a convenient way of handling input and output
+  files through accessing the input and output sandbox via GridFTP.
 \end{itemize}
@@
 
-\section{Support Classes}
-\subsection{{\URL}}
-\subsection{Logger}
-\subsection{XMLNode}
-
 \chapter{Building Command Line Interfaces}
@@
 
+Given all components listed above it possible to write versatile
+command line interfaces (cli) for grid job submission and
+management. {\arclib} offers the following native commands:
+
+\begin{enumerate}
+\item{\texttt{arcstat} - for computing resource or grid job status
+  queries.}
+\item{\texttt{arcsub} - for grid job submission}
+\item{\texttt{arcget} - for downloading output of finished, cancelled
+  or failed grid jobs.}
+\item{\texttt{arckill} - for terminating grid jobs.}
+\item{\texttt{arcclean} - for removing a grid job's session directory
+  including all contents}
+\item{\texttt{arccat} - for performing the \texttt{cat} command to a
+  running grid job's std out or std error file.}
+\end{enumerate}
+
+Each of the commands above are encoded within one C++ file with the
+following structure, here exemplified with the arcget command:
+
+\begin{shaded}
+\begin{verbatim}
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <iostream>
+#include <list>
+#include <string>
+
+#include <arc/ArcLocation.h>
+#include <arc/IString.h>
+#include <arc/Logger.h>
+#include <arc/OptionParser.h>
+#include <arc/client/JobController.h>
+#include <arc/client/JobSupervisor.h>
+#include <arc/client/UserConfig.h>
+
+int main(int argc, char **argv) {
+
+  setlocale(LC_ALL, "");
+
+  Arc::Logger logger(Arc::Logger::getRootLogger(), "arcget");
+  Arc::LogStream logcerr(std::cerr);
+  Arc::Logger::getRootLogger().addDestination(logcerr);
+  Arc::Logger::getRootLogger().setThreshold(Arc::WARNING);
+
+  Arc::ArcLocation::Init(argv[0]);
+
+  Arc::OptionParser options(istring("[job ...]"),
+			    istring("The arcget command is used for "
+				    "retrieving the results from a job."),
+			    istring("Argument to -c has the format "
+				    "Flavour:URL e.g.\n"
+				    "ARC0:ldap://grid.tsl.uu.se:2135/"
+				    "nordugrid-cluster-name=grid.tsl.uu.se,"
+				    "Mds-Vo-name=local,o=grid"));
+
+  bool all = false;
+  options.AddOption('a', "all",
+		    istring("all jobs"),
+		    all);
+
+  // Removed most of the option definition from this write-up to
+  // save space. See the real source file for the complete list.
+  // ...
+
+  bool version = false;
+  options.AddOption('v', "version", istring("print version information"),
+		    version);
+
+  std::list<std::string> jobs = options.Parse(argc, argv);
+
+  if (!debug.empty())
+    Arc::Logger::getRootLogger().setThreshold(Arc::string_to_level(debug));
+
+  Arc::UserConfig usercfg(conffile);
+  if (!usercfg) {
+    logger.msg(Arc::ERROR, "Failed configuration initialization");
+    return 1;
+  }
+
+  if (debug.empty() && usercfg.ConfTree()["Debug"]) {
+    debug = (std::string)usercfg.ConfTree()["Debug"];
+    Arc::Logger::getRootLogger().setThreshold(Arc::string_to_level(debug));
+  }
+
+  if (version) {
+    std::cout << Arc::IString("%s version %s", "arcget", VERSION)
+	      << std::endl;
+    return 0;
+  }
+
+  if (jobs.empty() && joblist.empty() && !all) {
+    logger.msg(Arc::ERROR, "No jobs given");
+    return 1;
+  }
+
+  if (joblist.empty())
+    joblist = usercfg.JobListFile();
+
+  Arc::JobSupervisor jobmaster(usercfg, jobs, clusters, joblist);
+
+  std::list<Arc::JobController*> jobcont = jobmaster.GetJobControllers();
+
+  if (jobcont.empty()) {
+    logger.msg(Arc::ERROR, "No job controllers loaded");
+    return 1;
+  }
+
+  int retval = 0;
+  for (std::list<Arc::JobController*>::iterator it = jobcont.begin();
+       it != jobcont.end(); it++)
+    if (!(*it)->Get(status, downloaddir, keep, timeout))
+      retval = 1;
+
+  return retval;
+}
+\end{verbatim}
+\end{shaded}
+
 \bibliography{grid}
+
 \end{document}

### 2008-09-18 doc/client/client_technical.tex
Info sys mapping documentation

@@
 {\centering \large Bjarte Mohn\footnote{bjarte.mohn@fysast.uu.se} \large \par}
-{\centering \large P\'{e}ter Stef\'{a}n\footnote{stefan@niif.hu} \large \par}
+{\centering \large G\'{a}bor R\H{o}czei\footnote{roczei@niif.hu} \large \par}
 %%%%---- An abstract - if style is article ----
@@
 priori knowing their difference. The {\ExecutionTarget} class mimics
-the glue2 information model (with a flattened structure), and thus a
+the Glue2 information model (with a flattened structure), and thus a
 mapping between attributes from other information systems into the
-glue2 format is needed. Appendix XX shows the current mapping for the
-production ARC, ARC1 and gLite middlewares.
+Glue2 format is needed. Appendix \ref{app:ExTarget} shows the current
+mapping for the production ARC, ARC1 and gLite middlewares.
 
@@
 {\Job} is a generic job class for storing all job related information.
-Attributes are derived from the glue2 information model and thus a
-mapping is needed for non glue2 compliant grid middlewares. Appendix
-XX shows the present mapping schema.
+Attributes are derived from the Glue2 information model and thus a
+mapping is needed for non Glue2 compliant grid middlewares. Appendix
+\ref{app:job} shows the present mapping schema.
 
@@
 
+\appendix
+
+\chapter{{\ExecutionTarget}}
+\label{app:ExTarget}
+
+\section{Domain and Location attributes}
+
+\subsection*{std::string DomainName}
+
+ARC0: nordugrid-cluster-name \\
+CREAM: GlueSiteName
+
+\subsection*{std::string Owner}
+
+ARC0: nordugrid-cluster-owner
+
+\subsection*{std::string Address}
+
+\subsection*{std::string Place}
+
+CREAM: GlueSiteLocation
+
+\subsection*{std::string PostCode}
+
+ARC0: nordugrid-cluster-location
+
+\subsection*{float Latitude}
+
+CREAM: GlueSiteLatitude
+
+\subsection*{float Longitude}
+
+CREAM: GlueSiteLongitude
+
+\section{ComputingService and ComputingEndpoint attributes}
+
+\subsection*{std::string CEID}
+
+\subsection*{std::string CEName}
+
+\subsection*{std::string Capability}
+
+\subsection*{std::string Type}
+
+\subsection*{std::string QualityLevel}
+
+\subsection*{URL url}
+
+ARC0: nordugrid-cluster-contactstring \\
+CREAM: GlueCEInfoContactString
+
+\subsection*{std::string Technology}
+
+\subsection*{std::string Interface}
+
+ARC0: ``GridFTP''
+
+\subsection*{std::string InterfaceExtension}
+
+\subsection*{std::string SupportedProfile}
+
+\subsection*{std::string Implementor}
+
+ARC0: ``NorduGrid''
+
+\subsection*{std::string ImplementationName}
+
+ARC0: ``ARC0'' \\
+CREAM: GlueCEImplementationName
+
+\subsection*{std::string ImplementationVersion}
+
+ARC0: nordugrid-cluster-middleware \\
+CREAM: GlueCEImplementaionVersion
+
+\subsection*{std::string HealthState}
+
+ARC0: nordugrid-queue-status
+
+\subsection*{std::string ServingState}
+
+CREAM: GlueCEStateStatus (GlueVOView $||$ GlueCE)
+
+\subsection*{std::string IssuerCA}
+
+ARC0: nordugrid-cluster-issuerca (nordugrid-cluster-issuerca-hash)
+
+\subsection*{std::list<std::string> TrustedCA}
+
+ARC0: nordugrid-cluster-trustedca
+
+\subsection*{Time DowntimeStarts}
+
+\subsection*{Time DowntimeEnds}
+
+\subsection*{std::string Staging}
+
+ARC0: nordugrid-cluster-nodeaccess
+
+\subsection*{std::string Jobdescription}
+
+\section{ComputingService and ComputingShare load attributes}
+
+\subsection*{int TotalJobs}
+
+ARC0: nordugrid-cluster-totaljobs \\
+CREAM: GlueCEStateTotalJobs (GlueVOView $||$ GlueCE)
+
+\subsection*{int RunningJobs}
+
+ARC0: nordugrid-queue-running \\
+CREAM: GlueCEStateRunningJobs (GlueVOView $||$ GlueCE)
+
+\subsection*{int WaitingJobs}
+
+ARC0: nordugrid-queue-queued $||$ nordugrid-cluster-queuedjobs \\
+CREAM: GlueCEStateWaitingJobs (GlueVOView $||$ GlueCE)
+
+\subsection*{int StagingJobs}
+
+\subsection*{int SuspendedJobs}
+
+\subsection*{int PreLRMSWaitingJobs}
+
+ARC0: nordugrid-queue-prelrmsqueued
+
+\subsection*{int LocalRunningJobs}
+
+ARC0: nordugrid-queue-running $-$ nordugrid-queue-gridrunning
+
+\subsection*{int LocalWaitingJobs}
+
+ARC0: nordugrid-queue-queued $-$ nordugrid-queue-gridqueue
+
+\subsection*{int FreeSlots}
+
+CREAM: GlueCEStateFreeJobSlots $||$ GlueCEStateFreeCPUs (GlueVOView $||$ GlueCE)
+
+\subsection*{std::string FreeSlotsWithDuration}
+
+\subsection*{int UsedSlots}
+
+ARC0: nordugrid-queue-usedcpus
+
+\subsection*{int RequestedSlots}
+
+\section{ComputingShare attributes}
+
+\subsection*{std::string MappingQueue}
+
+ARC0: nordugrid-queue-name
+
+\subsection*{Period MaxWallTime}
+
+ARC0: nordugrid-queue-maxwalltime \\
+CREAM: GlueCEPolicyMaxWallClockTime (GlueVOView $||$ GlueCE)
+
+\subsection*{Period MaxTotalWallTime}
+
+\subsection*{Period MinWallTime}
+
+ARC0: nordugrid-queue-minwalltime
+
+\subsection*{Period DefaultWallTime}
+
+ARC0: nordugrid-queue-defaultwalltime
+
+\subsection*{Period MaxCPUTime}
+
+ARC0: nordugrid-queue-maxcputime \\
+CREAM: GlueCEPolicyMaxCPUTime (GlueVOView $||$ GlueCE)
+
+\subsection*{Period MaxTotalCPUTime}
+
+\subsection*{Period MinCPUTime}
+
+ARC0: nordugrid-queue-defaultcputime
+
+\subsection*{Period DefaultCPUTime}
+
+ARC0: nordugrid-queue-defaultcputime
+
+\subsection*{int MaxTotalJobs}
+
+CREAM: GlueCEPolicyMaxTotalJobs (GlueVOView $||$ GlueCE)
+
+\subsection*{int MaxRunningJobs}
+
+ARC0: nordugrid-queue-maxrunning \\
+CREAM: GlueCEPolicyMaxRunningJobs (GlueVOView $||$ GlueCE)
+
+\subsection*{int MaxWaitingJobs}
+
+ARC0: nordugrid-queue-maxqueable \\
+CREAM: GlueCEPolicyMaxWaitingJobs (GlueVOView $||$ GlueCE)
+
+\subsection*{int NodeMemory}
+
+ARC0: nordugrid-queue-nodememory $||$ nordugrid-cluster-nodememory \\
+CREAM: GlueHostMainMemoryRAMSize
+
+\subsection*{int MaxPreLRMSWaitingJobs}
+
+\subsection*{int MaxUserRunningJobs}
+
+ARC0: nordugrid-queue-maxuserrun \\
+CREAM: GlueCEPolicyAssignedJobSlots (GlueVOView $||$ GlueCE) % ??
+
+\subsection*{int MaxSlotsPerJob}
+
+CREAM: GlueCEPolicyMaxSlotsPerJob (GlueVOView $||$ GlueCE)
+
+\subsection*{int MaxStageInStreams}
+\subsection*{int MaxStageOutStreams}
+\subsection*{std::string SchedulingPolicy}
+
+ARC0: nordugrid-queue-schedulingpolicy
+
+\subsection*{int MaxMemory}
+
+ARC0: nordugrid-queue-nodememory $||$ nordugrid-cluster-nodememory \\
+CREAM: GlueHostMainMemoryVirtualSize
+
+\subsection*{int MaxDiskSpace}
+\subsection*{URL DefaultStorageService}
+
+ARC0: nordugrid-cluster-localse \\
+CREAM: GlueCEInfoDefaultSE (GlueVOView $||$ GlueCE)
+
+\subsection*{bool Preemption}
+
+CREAM: GlueCEPolicyPreemption (GlueVOView $||$ GlueCE)
+
+\subsection*{Period EstimatedAverageWaitingTime}
+
+CREAM: GlueCEStateEstimatedResponseTime (GlueVOView $||$ GlueCE)
+
+\subsection*{Period EstimatedWorstWaitingTime}
+
+CREAM: GlueCEStateWorstResponseTime (GlueVOView $||$ GlueCE)
+
+\subsection*{std::string ReservationPolicy}
+
+\section{ComputingManager attributes}
+
+\subsection*{std::string ManagerType}
+
+ARC0: nordugrid-cluster-lrms-type (nordugrid-cluster-lrms-version)
+
+\subsection*{bool Reservation}
+\subsection*{bool BulkSubmission}
+\subsection*{int TotalPhysicalCPUs}
+
+ARC0: nordugrid-queue-totalcpus $||$ nordugrid-cluster-totalcpus
+
+\subsection*{int TotalLogicalCPUs}
+
+ARC0: nordugrid-queue-totalcpus $||$ nordugrid-cluster-totalcpus
+
+\subsection*{int TotalSlots}
+
+ARC0: nordugrid-queue-totalcpus $||$ nordugrid-cluster-totalcpus
+
+\subsection*{bool Homogeneity}
+
+ARC0: nordugrid-queue-homogeneity $||$ nordugrid-cluster-homogeneity
+
+\subsection*{std::string NetworkInfo}
+
+\subsection*{bool WorkingAreaShared}
+
+\subsection*{int WorkingAreaFree}
+
+ARC0: nordugrid-cluster-sessiondir-free
+
+\subsection*{Period WorkingAreaLifeTime}
+
+ARC0: nordugrid-cluster-sessiondir-lifetime
+
+\subsection*{int CacheFree}
+
+ARC0: nordugrid-cluster-cache-free
+
+\section{ExecutionEnvironment attributes}
+
+\subsection*{std::string Platform}
+
+ARC0: nordugrid-queue-architecture $||$ nordugrid-cluster-architecture
+
+\subsection*{bool VirtualMachine}
+
+\subsection*{std::string CPUVendor}
+
+ARC0: nordugrid-queue-nodecpu $||$ nordugrid-cluster-nodecpu
+
+\subsection*{std::string CPUModel}
+
+ARC0: nordugrid-queue-nodecpu $||$ nordugrid-cluster-nodecpu
+
+\subsection*{std::string CPUVersion}
+
+ARC0: nordugrid-queue-nodecpu $||$ nordugrid-cluster-nodecpu
+
+\subsection*{int CPUClockSpeed}
+
+\subsection*{int MainMemorySize}
+
+\subsection*{std::string OSFamily}
+
+ARC0: nordugrid-queue-opsys $||$ nordugrid-cluster-opsys
+
+\subsection*{std::string OSName}
+
+ARC0: nordugrid-queue-opsys $||$ nordugrid-cluster-opsys
+
+\subsection*{std::string OSVersion}
+
+\subsection*{bool ConnectivityIn}
+
+\subsection*{bool ConnectivityOut}
+
+\subsection*{std::list<Benchmark> Benchmarks}
+
+ARC0: nordugrid-queue-benchmark $||$ nordugrid-cluster-benchmark
+
+\section{ApplicationEnvironment}
+
+\subsection*{std::list<ApplicationEnvironment> ApplicationEnvironments}
+
+ARC0: nordugrid-cluster-runtimeenvironment \\
+CREAM: GlueHostApplicationSoftwareRunTimeEnvironment
+
+\chapter{{\Job}}
+\label{app:job}
+
+\section{Information stored in the job list file}
+
+\subsection*{std::string Flavour}
+
+\subsection*{URL JobID}
+
+\subsection*{URL Cluster}
+
+\subsection*{URL SubmissionEndpoint}
+
+\subsection*{URL InfoEndpoint}
+
+\subsection*{URL ISB}
+
+\subsection*{URL OSB}
+
+\subsection*{URL AuxURL}
+
+\subsection*{std::string AuxInfo}
+
+\section{Information retrieved from the informtaion system}
+
+\subsection*{std::string Name}
+
+ARC0: nordugrid-job-jobname
+
+\subsection*{std::string Type}
+
+\subsection*{URL IDFromEndpoint}
+
+\subsection*{std::string LocalIdFromManager}
+
+\subsection*{std::string JobDescription}
+
+\subsection*{std::string State}
+
+ARC0: nordugrid-job-status
+
+\subsection*{std::string RestartState}
+
+ARC0: nordugrid-job-rerunable
+
+\subsection*{int ExitCode}
+
+ARC0: nordugrid-job-exitcode
+
+\subsection*{std::string ComputingManagerExitCode}
+
+\subsection*{std::list<std::string> Error}
+
+ARC0: nordugrid-job-errors
+
+\subsection*{int WaitingPosition}
+
+ARC0: nordugrid-job-queuerank
+
+\subsection*{std::string UserDomain}
+
+\subsection*{std::string Owner}
+
+ARC0: nordugrid-job-globalowner
+
+\subsection*{std::string LocalOwner}
+
+\subsection*{Period RequestedWallTime}
+
+ARC0: nordugrid-job-reqwalltime
+
+\subsection*{Period RequestedTotalCPUTime}
+
+ARC0: nordugrid-job-reqcputime
+
+\subsection*{int RequestedMainMemory}
+
+\subsection*{int RequestedSlots}
+
+\subsection*{std::string StdIn}
+
+ARC0: nordugrid-job-stdin
+
+\subsection*{std::string StdOut}
+
+ARC0: nordugrid-job-stdout
+
+\subsection*{std::string StdErr}
+
+ARC0: nordugrid-job-stderr
+
+\subsection*{std::string LogDir}
+
+ARC0: nordugrid-job-gmlog
+
+\subsection*{std::list<std::string> ExecutionNode}
+
+Arc0: nordugrid-job-executionnodes
+
+\subsection*{std::string ExecutionCE}
+
+ARC0: nordugrid-job-execcluster
+
+\subsection*{std::string Queue}
+
+ARC0: nordugrid-job-execqueue
+
+\subsection*{Period UsedWallTime}
+
+ARC0: nordugrid-job-usedwalltime
+
+\subsection*{Period UsedTotalCPUTime}
+
+ARC0: nordugrid-job-usedcputime
+
+\subsection*{int UsedMainMemory}
+
+ARC0: nordugrid-job-usedmem
+
+\subsection*{std::list<std::string> UsedApplicationEnvironment}
+
+ARC0: nordugrid-job-runtimeenvironment
+
+\subsection*{int UsedSlots}
+
+ARC0: nordugrid-job-cpucount
+
+\subsection*{Time LocalSubmissionTime}
+
+\subsection*{Time SubmissionTime}
+
+ARC0: nordugrid-job-submissiontime
+
+\subsection*{Time ComputingManagerSubmissionTime}
+
+\subsection*{Time StartTime}
+
+\subsection*{Time ComputingManagerEndTime}
+
+\subsection*{Time EndTime}
+
+ARC0: nordugrid-job-completiontime
+
+\subsection*{Time WorkingAreaEraseTime}
+
+ARC0: nordugrid-job-sessiondirerasetime
+
+\subsection*{Time ProxyExpirationTime}
+
+ARC0: nordugrid-job-proxyexpirationtime
+
+\subsection*{std::string SubmissionHost}
+
+ARC0: nordugrid-job-submissionui
+
+\subsection*{std::string SubmissionClientName}
+
+ARC0: nordugrid-job-clientsoftware
+
+\subsection*{Time CreationTime}
+
+ARC0: Mds-validfrom
+
+\subsection*{Period Validity}
+
+ARC0: Mds-validto $-$ Mds-validfrom
+
+\subsection*{std::string OtherMessages}
+
+ARC0: nordugrid-job-comment
+
+\section{Associations}
+
+\subsection*{URL JobManagementEndpoint}
+
+\subsection*{URL DataStagingEndpoint}
+
+\section{ExecutionEnvironment (condensed)}
+
+\subsection*{bool VirtualMachine}
+
+\subsection*{std::string UsedCPUType}
+
+\subsection*{std::string UsedOSFamily}
+
+\subsection*{std::string UsedPlatform}
+
 \bibliography{grid}

### 2008-09-18 doc/client/client_technical.tex
JobDescription details

@@
 
+The parsing is possible from any language but by the other hand the
+output generation is not so widespread supported. There is only the
+JSDL and JDL output generation implemented yet and there are limitations
+too.
+
+The conversion means mainly a syntactial transformation and of course
+can do nothing in that cases where is not enough data to assemble the
+given attribute or can loose information where the received attribute
+has no equivalent pair in the output language.
+
+By the JSDL output generation the {\JobDescription} class is using the
+core JSDL's capabilities raised with the JSDL-POSIX and JSDL-ARC extensions.
+The loss of information is minimal in case of generating such an output.
+
+By the JDL generation the last version of JDL specification was used 
+(v0.9\footnote{\url{https://edms.cern.ch/file/590869/1/EGEE-JRA1-TEC-590869-JDL-Attributes-v0-9.pdf}})
+and only the not lapsing attributes are implemented (the deprecated
+attributes are excluded).
+
+Actually there is no XRSL generation implemented yet.
+
 \subsection{{\Submitter}}

### 2008-09-18 doc/client/client_technical.tex
Job description class documentation update

--- 2008-09-18 doc/client/client_technical.tex
@@
 
-The {\JobDescription} class has tree back-end classes, sometimes
+The {\JobDescription} class has three back-end classes, sometimes
 referred to as back-end or translator modules, corresponding to the
@@
 module for parsing and generating the job descriptions.
+
+Parsing is currently possible from any of the languages but output
+generation is only partially supported. Only JSDL and JDL output
+generation is implemented yet and even for these there are some
+limitations.
+
+The conversion mainly means a syntactial transformation and the
+{\JobDescription} class can of course do nothing in those cases where
+not enough data to assemble a given attribute is available or when
+information can be lost because the received attribute has no
+equivalent in the output language.
+
+For the JSDL output generation the {\JobDescription} class is using
+the core JSDL's capabilities amended with the JSDL-POSIX and JSDL-ARC
+extensions. The loss of information is minimal in case of generating
+such an output.
+
+For the JDL generation the latest version of JDL specification was
+used (v0.9) \cite{jdl}. Deprecated and backward compatibility
+attributes are not implemented.
+
+No XRSL output generation is implemented yet.
 

### 2008-09-18 doc/client/client_technical.tex
Fix some PDF stuff

@@
 \hypersetup{
-  pdfauthor = {Paper author},
-  pdftitle = {Paper title},
-  pdfsubject = {Paper subject},
-  pdfkeywords = {Paper,keyword,comma-separated},
+  pdfauthor = {Mattias Ellert, Bjarte Mohn, Ivan Marton, Gabor Roczei},
+  pdftitle = {libarcclient},
+  pdfsubject = {A Client Library for ARC},
+  pdfkeywords = {ARC,client,library},
   pdfcreator = {PDFLaTeX with hyperref package},

### 2008-09-22 doc/a-rex/arex_tech_doc.tex
Adding referencce to backends documentation - better than nothing
Implementing A-REX doc changes from Steffen.

@@
-%% LyX 1.4.5.1 created this file.  For more info, see http://www.lyx.org/.
+%% LyX 1.4.5 created this file.  For more info, see http://www.lyx.org/.
 %% Do not edit unless you really know what you are doing.
@@
 A-REX is one of ARC middleware components that implements functions
-of so called \emph{Computing Element} (CE).
+of the so called \emph{Computing Element} (CE). Here Computing Element
+is a service accepting requests containing description of generic
+computational job and executing it in underlying local batch system.
 
@@
 
+\subsection{\label{sub:LRMS}LRMS support}
+
+For information about supported LRMSes and their specific features
+and configuration options please read dedicated documentation at \cite{backends-arc1}.
+
 \begin{comment}(LRMS\_COMMENT)
-
-\subsection{\label{sub:LRMS}LRMS support}
 

### 2008-09-24 doc/a-rex/arex_tech_doc.tex
Some readability improvements, standard appearance

--- 2008-09-10 doc/a-rex/arex_tech_doc.tex
@@
-%% LyX 1.4.5.1 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[twoside,english]{article}
-\usepackage{pslatex}
-\usepackage[T1]{fontenc}
-\usepackage[latin1]{inputenc}
-\usepackage{geometry}
-\geometry{verbose,a4paper,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
-\usepackage{array}
-\usepackage{longtable}
-\usepackage{amsmath}
-\usepackage{setspace}
-\onehalfspacing
-\usepackage{amssymb}
-\IfFileExists{url.sty}{\usepackage{url}}
-                      {\newcommand{\url}{\texttt}}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\newenvironment{LyXParagraphLeftIndent}[1]%
-{
-  \begin{list}{}{%
-    \setlength{\topsep}{0pt}%
-    \addtolength{\leftmargin}{#1}
-    \setlength{\parsep}{0pt plus 1pt}%
-  }
-  \item[]
-}
-{\end{list}}
-%% Because html converters don't know tabularnewline
-\providecommand{\tabularnewline}{\\}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
-\newenvironment{lyxlist}[1]
-{\begin{list}{}
-{\settowidth{\labelwidth}{#1}
- \setlength{\leftmargin}{\labelwidth}
- \addtolength{\leftmargin}{\labelsep}
- \renewcommand{\makelabel}[1]{##1\hfil}}}
-{\end{list}}
-\newenvironment{lyxcode}
-{\begin{list}{}{
-\setlength{\rightmargin}{\leftmargin}
-\setlength{\listparindent}{0pt}% needed for AMS classes
-\raggedright
-\setlength{\itemsep}{0pt}
-\setlength{\parsep}{0pt}
-\normalfont\ttfamily}%
- \item[]}
-{\end{list}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-\usepackage{graphicx}                              %for PNG images (pdflatex)
-%\usepackage{graphics}                              %for EPS images (latex)
-\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
+\documentclass{article}                            %for shorter notes
+\usepackage{graphics}                              %for EPS images (latex)
 \usepackage{color}                                 %for defining custom colors
@@
 \usepackage{longtable}                             %for breaking tables
+\usepackage{listings}                              %for code listings
+\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
+\lstloadlanguages{bash,XML}
 \usepackage{multirow}
 \usepackage{verbatim}
+\usepackage{array}
 %
@@
   pdfauthor = {A.Konstantinov},
-  pdftitle = {The NorduGrid Grid Manager and GridFTP Server.},
+  pdftitle = {The ARC Computational Job Management Component -- A-REX},
   pdfsubject = {},
@@
 \definecolor{cyan}{rgb}{0,1,1}
+\definecolor{lightblue}{cmyk}{0.05,0,0,0.05}
+\lstset{
+frame=single,
+backgroundcolor=\color{shadecolor},
+fillcolor=\color{shadecolor},
+rulecolor=\color{lightblue},
+basicstyle=\ttfamily,
+xleftmargin=0.05\linewidth,
+aboveskip=2\medskipamount
+}
 %
-
-\usepackage{babel}
-\makeatother
+%
+%----- DON'T CHANGE HEADER MATTER
 \begin{document}
@@
 \begin{tabular}{rl}
-
-\resizebox*{3cm}{!}{\includegraphics{ng-logo}}
-
+\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
 &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
-
 \end{tabular}
@@
 
+%-------- Change this to NORDUGRID-XXXXXXX-NN
+
 {\raggedleft NORDUGRID-TECH-14\par}
@@
 
-\begin{center}
-\textsc{\large The ARC Computational Job Management Component - A-REX}
-\par\end{center}{\large \par}
-
+%%%%---- The title ----
+{\centering \textsc{\Large ARC Computational Job Management Component -- A-REX}\Large \par}
 \vspace*{0.5cm}
 
-\begin{center}
-\textit{\large Description and Administrator's Manual}
-\par\end{center}{\large \par}
+%%%%---- A subtitle, if necessary ----
+{\centering \textit{\large Description and Administrator's Manual}\large \par}
 
 \vspace*{1.5cm}
-
-{\centering \large A.Konstantinov\footnote{aleks@fys.uio.no} \large \par}
-
+%%%%---- A list of authors ----
+    {\centering \large A.~Konstantinov\footnote{aleks@fys.uio.no} \large \par}
+
+%%%%---- An abstract - if style is article ----
+%\begin{abstract}
+%The abstract
+%\end{abstract}
 \end{titlepage}
 
-\tableofcontents{}
-
+%\tableofcontents                          %Comment if use article style
 \newpage
-
 \section{Introduction\label{sec:intro}}
@@
 A-REX is one of ARC middleware components that implements functions
-of so called \emph{Computing Element} (CE).
-
-It takes care of job pre- and post-processing. This comprises the
+of the so-called \emph{Computing Element} (CE). Here Computing Element
+is a service accepting requests containing description of generic
+computational job and executing it in the underlying local batch system.
+
+It takes care of job pre- and post-processing. This comprises
 stage-in of files containing input data or program modules from wide
-range of sources and the transfer or storing of output results.
-
-The A-REX implements a Web Service (WS) interface which provides a
+range of sources and transfer or storing of output results.
+
+A-REX implements a Web Service (WS) interface which provides a
 way to submit and control computational tasks (jobs) to be executed
-by the A-REX and he underlying Local Resource Management System.
-
-\textbf{You should use this document for advanced configuration purposes
+by A-REX and the underlying Local Resource Management System.
+
+\begin{framed}You should use this document for advanced configuration purposes
 and understanding of internals of the aforementioned tools. For general
 installation and configuration of whole system refer to other documents
-available at \url{http://www.nordugrid.org/papers.html}. }
+available at \url{http://www.nordugrid.org/papers.html}.
+\end{framed}
 
@@
 
-Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
-\cite{ws-addr-soap} XML document - which identifies the job in the
-A-REX and the Information Interface.
+Each job gets an identifier (\textit{jobid}). This is a handle -- WS-Addressing~\cite{ws-addr-soap} XML document -- which identifies the job in the
+A-REX and in the Information Interface.
 
 Jobs are initiated and controlled through the WS interface. Complete
-job descriptions (JD)are passed to the A-REX through WS in JSDL\cite{jsdl}
+job descriptions (JD) are passed to the A-REX through WS in JSDL~\cite{jsdl}
 coded description. Input data files and job executables are transferred
-separately through the same interface, as described in the next section.
+separately through the same interface, as described in Section~\ref{sec:input/output data}.
 
@@
 
-One of the most important tasks of the A-REX is to take care of processing
+One of the most important tasks of the A-REX is to take care of processing of the
 input and output data (files) of the job. Input files are gathered
-in the SD or associated cache area. There are two ways to put a file
+in the SD or in the associated cache area. There are two ways to put a file
 into the SD:
@@
 some less standard sources are supported. Those are described below.
-The A-REX fully relies on HED framework \cite{hed} for data transferring
+The A-REX fully relies on HED framework~\cite{hed} for data transferring
 capabilities. Hence actual set of supported protocols depends on installed
@@
 
-\section{\label{Section: Job Flow}Job flow}
+\section{Job flow\label{Section:Job Flow}}
 
 From the point of view of the A-REX a job passes through various states.
-Picture \ref{job states diagram} presents a diagram of the possible
+Figure~\ref{job states diagram} presents a diagram of the possible
 states of a job.%
-\begin{figure}[tbh]
-\begin{centering}
-\includegraphics{pic1}
-\par\end{centering}
-
+
+\begin{figure}[th]
+\centering \includegraphics{pic1.pdf}
 \caption{\label{job states diagram}Job states}
 \end{figure}
+
 A user can examine the state of a job by querying the dedicated Local
@@
 \begin{itemize}
-\item \textbf{Accepted} - In this state the job has been submitted to a
+\item \textbf{Accepted} -- In this state the job has been submitted to a
 CE but is not processed yet. The A-REX will analyze the JD and move
@@
 canceled and moved to the state \textbf{Finishing}.
-\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
+\item \textbf{Preparing} -- The input data is being gathered in the SD (stage-in).
 The A-REX is downloading the files specified in the JD and is waiting
@@
 are processed out of limits.
-\item \textbf{Submiting} - The job is being passed for execution to the
+\item \textbf{Submiting} -- The job is being passed for execution to the
 \emph{Local Resource Management System} (LRMS). The corresponding
@@
 states.
-\item \textbf{Executing} - The job is queued or being executed in the LRMS.
+\item \textbf{Executing} -- The job is queued or being executed in the LRMS.
 The A-REX takes no actions except waiting until the job finishes.
-\item \textbf{Killing} - Necessary action to cancel the job in the LRMS
+\item \textbf{Killing} -- Necessary action to cancel the job in the LRMS
 is being taken.
-\item \textbf{Finishing} - The output data is being processed (stage-out).
+\item \textbf{Finishing} -- The output data is being processed (stage-out).
 Specified data files are moved to the specified SEs and are optionally
@@
 in this state.
-\item \textbf{Finished} - No more processing is performed by the A-REX.
+\item \textbf{Finished} -- No more processing is performed by the A-REX.
 The user can continue to download data files from the SD. The SD is
@@
 state.
-\item \textbf{Deleted} - The job is moved to this state if the user have
+\item \textbf{Deleted} -- The job is moved to this state if the user have
 not requested job to be cleaned before the SD's lifetime expires.
@@
 
-\section{URLs}
+\section{URLs\label{sec:urls}}
 
@@
 {}``The Hosting Environment of the Advanced Resource Connector middleware''
-document \cite{hed}.
-
-\section{\label{section: internals}Internals}
+document~\cite{hed}.
+
+\section{Internals\label{section:internals}}
 
@@
 
-For each local UNIX user listed in the A-REX configuration - including
-a generic one which covers all local user identities - a \textit{control
+For each local UNIX user listed in the A-REX configuration -- including
+a generic one which covers all local user identities -- a \textit{control
 directory} exists. In this directory the A-REX stores information
@@
 \begin{itemize}
-\item \textit{job.ID.status} - current state of the job. This is a plain
+\item \textit{job.ID.status} -- current state of the job. This is a plain
 text file containing a single word representing the internal name
 of current state of the job. Possible values and corresponding external
-job states are :
+job states are:
 
@@
 \end{itemize}
-See section \ref{Section: Job Flow} for a description of the various
+
+See Section~\ref{Section:Job Flow} for a description of the various
 states. Additionally each value can be prepended the prefix {}``PENDING:''
-(like PENDING:ACCEPTED, see section \ref{Section: Job Flow}). This
+(like PENDING:ACCEPTED, see Section~\ref{Section:Job Flow}). This
 is used to show that a job is \emph{ready} to be moved to the next
@@
 \begin{itemize}
-\item \textit{job.ID.description} - contains the description of the job
+\item \textit{job.ID.description} -- contains the description of the job
 (JD).
-\item \textit{job.ID.local} - information about the job used by the A-REX.
+\item \textit{job.ID.local} -- information about the job used by the A-REX.
 It consists of lines of format \textit{{}``name = value''}. Not
@@
 \begin{itemize}
-\item \textit{subject} - user certificate's subject, also known as the distinguished
+\item \textit{subject} -- user certificate's subject, also known as the distinguished
 name (DN)
-\item \textit{starttime} - GMT time when the job was accepted represented
+\item \textit{starttime} -- GMT time when the job was accepted represented
 in the Generalized Time format of LDAP 
-\item \textit{lifetime} - time period to preserve the SD after the job has
+\item \textit{lifetime} -- time period to preserve the SD after the job has
 finished in seconds
-\item \textit{cleanuptime} - GMT time when the job should be removed from
+\item \textit{cleanuptime} -- GMT time when the job should be removed from
 the cluster and it's SD deleted in Generalized Time format
-\item \textit{notify} - email addresses and flags to send mail to about
+\item \textit{notify} -- email addresses and flags to send mail to about
 the job specified status changes
-\item \textit{processtime} - GMT time when to start processing the job in
+\item \textit{processtime} -- GMT time when to start processing the job in
 Generalized Time format
-\item \textit{exectime} - GMT time when to start job execution in Generalized
+\item \textit{exectime} -- GMT time when to start job execution in Generalized
 Time format
-\item \textit{expiretime} - GMT time when the credentials delegated to the
+\item \textit{expiretime} -- GMT time when the credentials delegated to the
 job expire in Generalized Time format
-\item \textit{rerun} - number of retries left to rerun the job
-\item \textit{jobname} - name of the job as supplied by the user
-\item \textit{lrms} - name of the LRMS backend to be used for local submission
-\item \textit{queue} - name of the queue to run the job at
-\item \textit{localid} - job id in LRMS (appears only after the job has
+\item \textit{rerun} -- number of retries left to rerun the job
+\item \textit{jobname} -- name of the job as supplied by the user
+\item \textit{lrms} -- name of the LRMS backend to be used for local submission
+\item \textit{queue} -- name of the queue to run the job at
+\item \textit{localid} -- job id in LRMS (appears only after the job has
 reached state \textbf{InLRMS})
-\item \textit{args} - executable name followed by a list of command-line
+\item \textit{args} -- executable name followed by a list of command-line
 arguments
-\item \textit{downloads} - number of files to download into the SD before
+\item \textit{downloads} -- number of files to download into the SD before
 execution
-\item \textit{uploads} - number of files to upload from the SD after execution
-\item \textit{gmlog} - directory name which holds files containing information
+\item \textit{uploads} -- number of files to upload from the SD after execution
+\item \textit{gmlog} -- directory name which holds files containing information
 about the job when accessed through GridFTP interface
-\item \textit{clientname} - name (as provided by the user interface) and
+\item \textit{clientname} -- name (as provided by the user interface) and
 IP address:port of the submitting client machine
-\item \textit{clientsoftware} - version of software used to submit the job
-\item \textit{sessiondir} - the job's SD
-\item \textit{failedstate} - state in which job failed (available only if
+\item \textit{clientsoftware} -- version of software used to submit the job
+\item \textit{sessiondir} -- the job's SD
+\item \textit{failedstate} -- state in which job failed (available only if
 it is possible to restart the job)
-\item \textit{jobreport} - URL of a user requested \emph{logger service}.
+\item \textit{jobreport} -- URL of a user requested \emph{logger service}.
 The A-REX will also send job records to this service in addition to
@@
 \end{itemize}
+
 This file is filled partially during job submission and fully when
@@
 \begin{itemize}
-\item \textit{job.ID.input} - list of input files. Each line contains 2
+\item \textit{job.ID.input} -- list of input files. Each line contains 2
 values separated by a space. First value contains name of the file
@@
 
-\textit{file description} - {[}size]{[}.checksum].
-
-\hspace*{1cm}\textit{size} - size of the file in bytes.
-
-\hspace*{1cm}\textit{checksum} - checksum of the file identical to
+\textit{file description} -- {[}size]{[}.checksum].
+
+\hspace*{1cm}\textit{size} -- size of the file in bytes.
+
+\hspace*{1cm}\textit{checksum} -- checksum of the file identical to
 the one produced by \textbf{\textit{cksum}} (1).
@@
 exist.
-
-\end{itemize}
+\end{itemize}
+
 This file is used by the '\textbf{\textit{downloader}}' utility. Files
@@
 \begin{itemize}
-\item \textit{job.ID.output} - list of output files. Each line contains
+\item \textit{job.ID.output} -- list of output files. Each line contains
 1 or 2 values separated by a space. First value is the name of the
@@
 \end{itemize}
+
 This file is used by the '\textbf{\textit{uploader}}' utility. Files
@@
 \begin{itemize}
-\item \textit{job.ID.failed} - the existence of this file marks the failure
+\item \textit{job.ID.failed} -- the existence of this file marks the failure
 of the job. It can also contain one or more lines of text describing
@@
 from zero of the job itself.
-\item \textit{job.ID.errors} - this file contains the output produced by
+\item \textit{job.ID.errors} -- this file contains the output produced by
 external utilities like \textbf{\textit{downloader}}, \textbf{\textit{uploader}},
@@
 content of that file while asking for help.
-\item \textit{job.ID.diag} - information about resources used during execution
+\item \textit{job.ID.diag} -- information about resources used during execution
 of job and other information suitable for diagnostics and statistics.
@@
 names are at least defined:
-
-\begin{itemize}
-\item \textit{nodename} - name of computing node which was used to execute
+\begin{itemize}
+\item \textit{nodename} -- name of computing node which was used to execute
 job,
-\item \textit{runtimeenvironments} - used runtime environments separated
+\item \textit{runtimeenvironments} -- used runtime environments separated
 by ';',
-\item \textit{exitcode} - numerical exit code of job,
-\item \textit{frontend\_distribution} - name and version of operating system
+\item \textit{exitcode} -- numerical exit code of job,
+\item \textit{frontend\_distribution} -- name and version of operating system
 distribution on frontend computer,
-\item \textit{frontend\_system} - name of operating on frontend computer,
-\item \textit{frontend\_subject} - subject (DN) of certificate representing
+\item \textit{frontend\_system} -- name of operating on frontend computer,
+\item \textit{frontend\_subject} -- subject (DN) of certificate representing
 frontend computer,
-\item \textit{frontend\_ca} - subject (DN) of issuer of certificate representing
+\item \textit{frontend\_ca} -- subject (DN) of issuer of certificate representing
 frontend computer,
@@
 in their output. Hence some lines can have broken format.
-
 \item \textit{job.ID.proxy} -delegated X509 credentials.
-\item \textit{job.ID.proxy.tmp} - temporary X509 credentials with different
+\item \textit{job.ID.proxy.tmp} -- temporary X509 credentials with different
 UNIX ownership used by processes run with effective \emph{user id}
@@
 \end{itemize}
+
 There are other files with names like job.ID.{*} which are created
@@
 during the execution of a job. The WS Interface is built and deployed
-inside the Hosting Environment Daemon (HED) infrastructure\cite{hed}.
+inside the Hosting Environment Daemon (HED) infrastructure~\cite{hed}.
 
@@
 The job submission and control interface is based on a document produced
-by the OGF OGSA Basic Execution Services (BES) Working Group\cite{ogsa-bes}.
+by the OGF OGSA Basic Execution Services (BES) Working Group~\cite{ogsa-bes}.
 
 The exchange of SOAP messages is performed via HTTP(S). The BES interface
-is represented by two port-types - BES-Management and BES-Factory.
+is represented by two port-types -- BES-Management and BES-Factory.
 The former is made to control the A-REX service itself and thus defines
@@
 For job descriptions A-REX accepts the Job Submission Description
-Language (JSDL)\cite{jsdl} documents as defined by the OGF Job Submission
+Language (JSDL)~\cite{jsdl} documents as defined by the OGF Job Submission
 Description Language Working Group. Supported elements and extensions
@@
 BES. It does that by defining its own port-type with the single operation
-\emph{ChangeActivityStatus} (see Appendix C). This operation provides
+\emph{ChangeActivityStatus} (see Appendix~\ref{annex:arex-wsdl}). This operation provides
 a way to request simple transfers between states of jobs and corresponding
@@
 \item \emph{ChangeActivityStatus}
-
 \begin{itemize}
 \item Input
-
 \begin{itemize}
@@
 \item Output
-
 \begin{itemize}
@@
 \item Fault(s)
-
 \begin{itemize}
@@
 \end{itemize}
+
 On result of this command, the job should be put into the requested
@@
 of activity processing in addition to those defined by the BES, as
-listed in Table \ref{tab:Job-states-definitions}. Their meaning is
-described in section \ref{Section: Job Flow}.
+listed in Table~\ref{tab:Job-states-definitions}. Their meaning is
+described in Section~\ref{Section:Job Flow}.
 
@@
 
-\begin{tabular}{|>{\centering}p{0.15\columnwidth}|>{\centering}p{0.15\columnwidth}|c|>{\raggedright}p{0.4\columnwidth}|}
+\begin{tabular}{|>{\centering}p{0.15\columnwidth}|>{\centering}p{0.15\columnwidth}|>{\centering}p{0.2\columnwidth}|p{0.4\columnwidth}|}
 \hline 
@@
 \hline 
-\renewcommand{\multirowsetup}{\centering}\multirow{2}{2cm}{Pending}&
-Accepting&
-ACCEPTED&
+\multirow{2}{*}{Pending}& Accepting& ACCEPTED &
 Job is in the process of being submitted. This state is not recognised
-by the A-REX yet. \emph{Accepted} is first reported state.\tabularnewline
-\cline{2-2} \cline{3-3} \cline{4-4} 
-\multicolumn{1}{|c|}{}&
-Accepted&
-ACCEPTED&
-Job was submitted\tabularnewline
-\hline 
-\renewcommand{\multirowsetup}{\centering}\multirow{7}{2cm}{Running}&
-Preparing&
-PREPARING&
-Stage-in process is going on\tabularnewline
-\cline{2-2} \cline{3-3} \cline{4-4} 
-\multicolumn{1}{|c|}{}&
-Prepared&
-PREPARING+PENDING&
-Stage-in process has finished\tabularnewline
-\cline{2-2} \cline{3-3} \cline{4-4} 
-\multicolumn{1}{|c|}{}&
-Submiting&
-SUBMIT&
-Communication with local batch system is in process\tabularnewline
-\hline 
-&
-Queued&
-INLRMS&
+by the A-REX yet. \emph{Accepted} is first reported state\\ \cline{2-4}
+&Accepted& ACCEPTED& Job was submitted\\ \hline 
+\multirow{8}{*}{Running}& Preparing& PREPARING& Stage-in process is going on\\ \cline{2-4} 
+&Prepared& PREPARING + PENDING& Stage-in process has finished\\ \cline{2-4} 
+&Submiting& SUBMIT& Communication with local batch system is in process\\ \cline{2-4} 
+&Queued& INLRMS&
 Job entered local batch system but is not runnning now. This state
-is not recognised by the A-REX yet. \emph{Executing} is reported instead \tabularnewline
-\cline{2-2} \cline{3-3} \cline{4-4} 
-\multicolumn{1}{|c|}{}&
-Executing&
-INLRMS&
-Job is being executed in local batch system\tabularnewline
-\cline{2-2} \cline{3-3} \cline{4-4} 
-\multicolumn{1}{|c|}{}&
-Executed&
-INLRMS, INLRMS+PENDING&
+is not recognised by the A-REX yet. \emph{Executing} is reported instead\\ \cline{2-4} 
+&Executing& INLRMS& Job is being executed in local batch system\\ \cline{2-4} 
+&Executed& INLRMS, INLRMS + PENDING&
 Job execution in local batch system has finished. The A-REX dos not
 detect job states inside local batch system yet. As result this state
-is reported only if job is \emph{Pending}.\tabularnewline
-\cline{2-2} \cline{3-3} \cline{4-4} 
-\multicolumn{1}{|c|}{}&
-Killing&
-CANCELING&
-Communication with local batch system to terminate execution is in
-process\tabularnewline
-\cline{2-2} \cline{3-3} \cline{4-4} 
-\multicolumn{1}{|c|}{}&
-Finishing&
-FINISHING&
-Stage-out process is going on\tabularnewline
-\hline 
-Cancelled&
-Killed&
-FINISHED&
+is reported only if job is \emph{Pending}.\\ \cline{2-4} 
+&Killing& CANCELING& Communication with local batch system to terminate execution is in process\\ \cline{2-4} 
+&Finishing& FINISHING& Stage-out process is going on\\ \hline 
+Cancelled& Killed& FINISHED&
 Job was stopped by explicit user request. The A-REX currently does
-not remember this request. \emph{Failed} is reported instead.\tabularnewline
-\hline 
-Failed&
-Failed&
-FINISHED&
-\multicolumn{1}{l|}{There was a failure during execution}\tabularnewline
-\hline 
-Finished&
-Finished&
-FINISHED&
-Job finished successfully\tabularnewline
-\hline 
-Finished&
-Deleted&
-DELETED&
-Job finished and was left in A-REX too long\tabularnewline
-\hline 
-All&
-Pending&
-PENDING&
+not remember this request. \emph{Failed} is reported instead.\\ \hline 
+Failed& Failed& FINISHED& There was a failure during execution\\ \hline 
+Finished& Finished& FINISHED& Job finished successfully\\ \hline 
+Finished& Deleted& DELETED& Job finished and was left in A-REX too long\\ \hline 
+All& Pending& PENDING&
 Job is prevented from going to the next state due to some internal
-limits; this sub-state appears in parallel with other sub-states\tabularnewline
-\hline 
-All&
-Held&
-&
+limits; this sub-state appears in parallel with other sub-states\\ \hline 
+All& Held& &
 Job processing is suspended on client request; this sub-state appears
 in parallel with other sub-states. This state is reserved for future
-and is not implemented yet.\tabularnewline
-\hline
+and is not implemented yet.\\ \hline
 \end{tabular}
@@
 
-The A-REX also supports the Delegation Interface (see Appendix D).
+The A-REX also supports the Delegation Interface (see Appendix~\ref{annex:delegation-wsdl}).
 This is a common purpose interface to be used by ARC services which
@@
 \begin{itemize}
-\item \emph{DelegateCredentialsInit} operation - this operation performs
+\item \emph{DelegateCredentialsInit} operation -- this operation performs
 the first half of the credentials delegation sequence.
-
 \begin{itemize}
 \item Input
-
 \begin{itemize}
@@
 \item Output(s)
-
 \begin{itemize}
@@
 \end{itemize}
-\item \emph{UpdateCredentials} operation - this operation makes it possible
+\item \emph{UpdateCredentials} operation -- this operation makes it possible
 to update the content of delegated credentials (like in the case of
 credentials being renewed) unrelated to other operations of the service.
-
 \begin{itemize}
 \item Input
-
 \begin{itemize}
@@
 \end{itemize}
+
 Additionally, A-REX Web Service Interface allows delegation to be
@@
 For more information about delegations and delegation interface refer
-to \cite{wsrf-rp}.
+to~\cite{wsrf-rp}.
 
@@
 interface common for all ARC services. This interface is based on
-OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
+OASIS Web Services Resource Properties specification~\cite{wsrf-rp}.
 Information about resources and maintained activities/jobs are represented
@@
 
-\begin{lyxlist}{00.00.0000}
-\item [{\emph{nordugrid}}] - description of computing resource that uses
-NorudGrid LDAP schema \cite{is} converted to XML document.
-\item [{\emph{Domains}}] \emph{-} description of a computation resource
+\begin{list}{--}{\setlength{\labelwidth}{0.5cm}\setlength{\rightmargin}{\leftmargin}}
+\item [{\emph{nordugrid}}] -- description of computing resource that uses
+NorudGrid LDAP schema~\cite{is} converted to XML document.
+\item [{\emph{Domains}}] -- description of a computation resource
 that uses Glue2 schema.
-\end{lyxlist}
+\end{list}
 All information can be accessed either through requests on particular
@@
 
-A-REX supports the following elements from the JSDL version 1.0 specification\cite{jsdl}
+A-REX supports the following elements from the JSDL version 1.0 specification~\cite{jsdl}
 including POSIX Applications extension and JSDL HPC Profile Application
-Extension \cite{jsdl-hpc}:
-
-\begin{lyxlist}{00.00.0000}
-\item [{\emph{JobName}}] \emph{-} name of the job as assigned by the user.
-\item [{\emph{Executable}\ (POSIX,HPC)}] - name of the executable file.
-\item [{\emph{Argument}\ (POSIX,HPC)}] - arguments the executable will
+Extension~\cite{jsdl-hpc}:
+
+\begin{list}{--}{\setlength{\labelwidth}{0.5cm}\setlength{\rightmargin}{\leftmargin}}
+\item [{\emph{JobName}}] -- name of the job as assigned by the user.
+\item [{\emph{Executable}\ (POSIX,HPC)}] -- name of the executable file.
+\item [{\emph{Argument}\ (POSIX,HPC)}] -- arguments the executable will
 be launched with.
 \item [{\emph{DataStaging}}]~
-
-\begin{lyxlist}{00.00.0000}
-\item [{Filename}] - name of the data file on the executing node.
-\item [{\emph{Source}}] - source where the file will be taken from before
+\begin{list}{--}{\setlength{\labelwidth}{0.5cm}\setlength{\rightmargin}{\leftmargin}}
+\item [{Filename}] -- name of the data file on the executing node.
+\item [{\emph{Source}}] -- source where the file will be taken from before
 execution.
-\item [{\emph{Target}}] \emph{-} destination the file will be delivered
+\item [{\emph{Target}}] -- destination the file will be delivered
 to after execution.
-\end{lyxlist}
-\item [{\emph{Input}\ (POSIX,HPC)}] - file to be used as standard input
+\end{list}
+\item [{\emph{Input}\ (POSIX,HPC)}] -- file to be used as standard input
 for the executable.
-\item [{\emph{Output}\ (POSIX,HPC)}] - file to be used as standard output
+\item [{\emph{Output}\ (POSIX,HPC)}] -- file to be used as standard output
 for the executable.
-\item [{\emph{Error}\ (POSIX,HPC)}] - file to be used as standard error
+\item [{\emph{Error}\ (POSIX,HPC)}] -- file to be used as standard error
 for the executable.
-\item [{\emph{MemoryLimit}\ (POSIX)}] - amount of physical memory needed
+\item [{\emph{MemoryLimit}\ (POSIX)}] -- amount of physical memory needed
 for execution.
-\item [{\emph{TotalPhysicalMemory}}] - same as MemoryLimit.
-\item [{\emph{IndividualPhysicalMemory}}] - same as MemoryLimit.
-\item [{\emph{CPUTimeLimit}\ (POSIX)}] - maximal amount of CPU time needed
+\item [{\emph{TotalPhysicalMemory}}] -- same as MemoryLimit.
+\item [{\emph{IndividualPhysicalMemory}}] -- same as MemoryLimit.
+\item [{\emph{CPUTimeLimit}\ (POSIX)}] -- maximal amount of CPU time needed
 for execution.
-\item [{\emph{TotalCPUTime}}] - same as CPUTimeLimit.
-\item [{\emph{IndividualCPUTime}}] - same as CPUTimeLimit.
-\item [{\emph{WallTimeLimit}\ (POSIX)}] - amount of clock time needed
+\item [{\emph{TotalCPUTime}}] -- same as CPUTimeLimit.
+\item [{\emph{IndividualCPUTime}}] -- same as CPUTimeLimit.
+\item [{\emph{WallTimeLimit}\ (POSIX)}] -- amount of clock time needed
 for execution.
-\item [{\emph{TotalCPUCount}}] \emph{-} number of CPUs needed for execution.
-\item [{\emph{IndividualCPUCount}}] - same as \emph{TotalCPUCount}.
-\end{lyxlist}
-%Environment(POSIX) - environment variable.
+\item [{\emph{TotalCPUCount}}] -- number of CPUs needed for execution.
+\item [{\emph{IndividualCPUCount}}] -- same as \emph{TotalCPUCount}.
+\end{list}
+%Environment(POSIX) -- environment variable.
 
@@
 A-REX accepts JSDL documents having the following additional elements
-(see Appendix E):
-
-\begin{lyxlist}{00.00.0000}
-\item [{\emph{IsExecutable}}] \emph{-} marks file to become executable
+(see Appendix~\ref{annex:jsdl-extension}):
+
+\begin{list}{--}{\setlength{\labelwidth}{0.5cm}\setlength{\rightmargin}{\leftmargin}}
+\item [{\emph{IsExecutable}}] -- marks file to become executable
 after being delivered to the computing resource.
-\item [{\emph{RunTimeEnvironment}}] - specifies the name of the Runtime
+\item [{\emph{RunTimeEnvironment}}] -- specifies the name of the Runtime
 Environment needed for job execution.
-\item [{\emph{Middleware}}] - request for specific middleware on the computing
+\item [{\emph{Middleware}}] -- request for specific middleware on the computing
 resource frontend.
-\item [{\emph{RemoteLogging}}] \emph{-} destination for the usage record
+\item [{\emph{RemoteLogging}}] -- destination for the usage record
 report of the executed job.
-\item [{\emph{LocalLogging}}] - name for the virtual directory available
+\item [{\emph{LocalLogging}}] -- name for the virtual directory available
 through job interface and containing various debug information about
 job execution.
-\item [{\emph{AccessControl}}] - ACL expression which describes the identities
+\item [{\emph{AccessControl}}] -- ACL expression which describes the identities
 of those clients who are allowed to perform operations on this job.
-\item [{\emph{Notify}}] - Email destination for notification of job state
+\item [{\emph{Notify}}] -- Email destination for notification of job state
 changes.
-\item [{\emph{SessionLifeTime}}] - duration for the directory containing
+\item [{\emph{SessionLifeTime}}] -- duration for the directory containing
 job-related files to exist after the job finished executing.
-\item [{\emph{JoinOutputs}}] - specifies if standard output and standard
+\item [{\emph{JoinOutputs}}] -- specifies if standard output and standard
 error channels must be merged.
-\item [{\emph{Reruns}}] \emph{-} defines how many times a job is allowed
+\item [{\emph{Reruns}}] -- defines how many times a job is allowed
 to rerun in case of failure.
-\item [{\emph{CredentialServer}}] \emph{-} URL of MyProxy service which
+\item [{\emph{CredentialServer}}] -- URL of MyProxy service which
 may be used for renewing the expired delegated job credentials.
-\item [{\emph{CandidateTarget}}] \emph{-} specifies host name and queue
+\item [{\emph{CandidateTarget}}] -- specifies host name and queue
 of a computing resource.
-\end{lyxlist}
+\end{list}
 
@@
 in URL (for URL options read {}``Protocols, Uniform Resource Locators
-(URL) and extensions supported in ARC'' \cite{urls}). The disc space
+(URL) and extensions supported in ARC''~\cite{urls}). The disc space
 occupied by cache is controlled by removing unused files. For more
-information look in section \ref{SubSection: ConfigFile}.
-
-\emph{NOTE: The caching algorithm used by A-REX and described above
+information look in Section~\ref{SubSection:ConfigFile}.
+
+\begin{framed}
+NOTE: The caching algorithm used by A-REX and described above
 is currently being replaced by a better one which is no robust and
-provides better scalability.}
+provides better scalability.
+\end{framed}
 
@@
 \begin{itemize}
-\item \textit{list} - stores names of the files (8 digit numbers) and corresponding
+\item \textit{list} -- stores names of the files (8 digit numbers) and corresponding
 URLs delimited by blank space. Each pair is delimited by some amount
@@
 stored if available
-\item \textit{old} - stores URLs which have been removed from cache. Records
+\item \textit{old} -- stores URLs which have been removed from cache. Records
 are delimited by some amount of \textbackslash{}0 codes and are meant
 to be removed by some external routine.
-\item \textit{new} - stores URLs which have been added to cache. Records
+\item \textit{new} -- stores URLs which have been added to cache. Records
 are delimited by some amount of \textbackslash{}0 codes and are removed
@@
 it is removed from \emph{new}.
-\item \textit{statistics} - consists of strings containing \emph{name=value}
+\item \textit{statistics} -- consists of strings containing \emph{name=value}
 pairs. Following names are defined:
@@
 \begin{itemize}
-\item \textit{hardsize} -size of file system for storing cached data
-\item \textit{hardfree} - amount of disc space available on that file system
-\item \textit{softsize} - if cache exceeds this size files are started being
+\item \textit{hardsize} -- size of file system for storing cached data
+\item \textit{hardfree} -- amount of disc space available on that file system
+\item \textit{softsize} -- if cache exceeds this size files are started being
 removed 
-\item \textit{softfree} - space left till softsize (can be negative)
-\item \textit{claimed} - space used by files claimed by running jobs
-\item \textit{unclaimed} - space used by files not being currently used
+\item \textit{softfree} -- space left till softsize (can be negative)
+\item \textit{claimed} -- space used by files claimed by running jobs
+\item \textit{unclaimed} -- space used by files not being currently used
 by any job
 \end{itemize}
-\item \textit{\#\#\#\#\#\#\#\#.info} - stores state of file (\textit{\#\#\#\#\#\#\#\#}
+\item \textit{\#\#\#\#\#\#\#\#.info} -- stores state of file (\textit{\#\#\#\#\#\#\#\#}
 stands for 8 digits). State is represented by one character:\\
-\textit{c} - just created, content is empty.\\
-\textit{f} - failed to download (treated same as 'c').\\
-\textit{r} - ready to be used, content is valid.\\
-\textit{d} - being downloaded. 'd' is followed by identifier of application/job
+\textit{c} -- just created, content is empty.\\
+\textit{f} -- failed to download (treated same as 'c').\\
+\textit{r} -- ready to be used, content is valid.\\
+\textit{d} -- being downloaded. 'd' is followed by identifier of application/job
 downloading that file. During content's download this file has write
 lock set.
-\item \textit{\#\#\#\#\#\#\#\#.claim} - stores list of identifiers of applications/jobs
+\item \textit{\#\#\#\#\#\#\#\#.claim} -- stores list of identifiers of applications/jobs
 using this file. Identifiers are stored one per line.
-\item \textit{\#\#\#\#\#\#\#\#} - files storing content of corresponding
+\item \textit{\#\#\#\#\#\#\#\#} -- files storing content of corresponding
 URL. These can be stored in separate directory.
 \end{itemize}
-Files \textit{list, old , new} and \textit{\#\#\#\#\#\#\#\#.info}
+
+Files \textit{list, old, new} and \textit{\#\#\#\#\#\#\#\#.info}
 has to be stored on file system which has support for files' locking.
@@
 \begin{itemize}
-\item \textit{libarex.so} - The main module providing main functionality
+\item \textit{libarex.so} -- The main module providing main functionality
 and web interface. It is implemented as HTTP and SOAP service inside
@@
 and running other modules.
-\item \textit{downloader} - This is a module responsible for gathering input
+\item \textit{downloader} -- This is a module responsible for gathering input
 files in the SD. It processes the \textit{job.ID.input} file and updates
 it.
-\item \textit{uploader} - This module is responsible for delivering output
+\item \textit{uploader} -- This module is responsible for delivering output
 files to the specified SEs and registration at an Indexing Service
@@
 \end{itemize}
-%cache-register - Utility to register cached data into an Indexing Service.
-It reads and modifies cache informational files old and new (as described in section ).
-Configuration is read directly from the A-REX's configuration file (see section ).
+
+%cache-register -- Utility to register cached data into an Indexing Service.
+It reads and modifies cache informational files old and new (as described in Section ).
+Configuration is read directly from the A-REX's configuration file (see Section ).
 It is run by the A-REX every 5 minutes.
 
-%frontend-info-collector - Utility to gather information about the frontend. It puts collected information into the job.ID.diag file.
-
-\begin{itemize}
-\item \emph{gm-kick} - Sends a signal to the A-REX though a FIFO file to
+%frontend-info-collector -- Utility to gather information about the frontend. It puts collected information into the job.ID.diag file.
+
+\begin{itemize}
+\item \emph{gm-kick} -- Sends a signal to the A-REX though a FIFO file to
 wake it up. It's used to increase responsiveness of A-REX.
-\item \emph{CEinfo.pl} - Collects and generates information about computing
+\item \emph{CEinfo.pl} -- Collects and generates information about computing
 resource as XML document in NorduGrid and Glue 2 format.
 \end{itemize}
+
 The following modules are always run under the Unix account to which
@@
 \begin{itemize}
-\item \textit{smtp-send.sh} and \textit{smtp-send} - These are the modules
+\item \textit{smtp-send.sh} and \textit{smtp-send} -- These are the modules
 responsible for sending e-mail notifications to the user. The format
@@
 script \textit{smtp-send.sh}. 
-\item \textit{submit-{*}-job} - Here {*} stands for the name of the LRMS.
+\item \textit{submit-{*}-job} -- Here {*} stands for the name of the LRMS.
 Currently supported LRMS are PBS/Torque, Condor, LoadLeveler, LSF
@@
 This module is responsible for job submission to the LRMS.
-\item \textit{cancel-{*}-job} - This script is for canceling jobs which
+\item \textit{cancel-{*}-job} -- This script is for canceling jobs which
 have been already submitted to the LRMS.
@@
 \end{itemize}
+
 \begin{comment}In addition, there is also an administration utility:
@@
 \begin{itemize}
-\item \textit{gm-jobs} - prints a list of jobs available on the cluster
+\item \textit{gm-jobs} -- prints a list of jobs available on the cluster
 and the number of jobs in each state.\\
@@
 
-\begin{LyXParagraphLeftIndent}{1cm}
-\$ARC\_LOCATION/bin -- tools
-
-\$ARC\_LOCATION/libexec -- program modules used by A-REX
-
-/etc -- central configuration file - location used by default
-
-\$ARC\_LOCATION/lib/arc -- service module
-
-\end{LyXParagraphLeftIndent}
+\begin{itemize}
+\item[] \$ARC\_LOCATION/bin -- tools
+\item[] \$ARC\_LOCATION/libexec -- program modules used by A-REX
+\item[] /etc -- central configuration file -- location used by default
+\item[] \$ARC\_LOCATION/lib/arc -- service module
+\end{itemize}
+
 The A-REX also uses following directories:
@@
 \begin{itemize}
-\item \textit{session root directory} - This is the directory in which a
+\item \textit{session root directory} -- This is the directory in which a
 user's SDs are created. It's location is configurable per UNIX user.
@@
 \end{itemize}
-\begin{comment} (LRMS\_COMMENT) For more information see section
-\ref{sub:LRMS}.\end{comment}
-
-\begin{itemize}
-\item \textit{control directory} - In this directory A-REX stores an information
+
+\begin{comment} (LRMS\_COMMENT) For more information see Section~\ref{sub:LRMS}.\end{comment}
+
+\begin{itemize}
+\item \textit{control directory} -- In this directory A-REX stores an information
 about accepted jobs. It must have full permissions there.\\
@@
 
-\subsection{\label{SubSection: ConfigFile}Configuration of the A-REX}
+\subsection{Configuration of the A-REX\label{SubSection:ConfigFile}}
 
 Due to historical reasons configuration of the A-REX is split into
-2 parts. The small part is located inside HED configuration (see appendix
-B) for schema and description of supported elements. It refers to
+2 parts. The small part is located inside HED configuration (see Appendix~\ref{annex:arex-conf}) for schema and description of supported elements. It refers to
 file containing main part of configuration. The default location of
-main configuration file is \textit{/etc/arc.conf .}
+main configuration file is \textit{/etc/arc.conf}.
 
 The configuration file can contain empty lines and comments in lines
-starting from \#. It is separated into sections. Each sections starts
+starting with \#. It is separated into sections. Each section starts
 from string containing pattern
 
-\begin{itemize}
-\item \emph{{[}section name/subsection name/subsubsection name]}. 
-\end{itemize}
+\begin{shaded}
+\verb|[section name/subsection name/subsubsection name]|
+\end{shaded}
+
 Each section continues till next section or end of file. One configuration
@@
 
-\begin{itemize}
-\item \emph{name=''arguments string''.}
-\end{itemize}
+\begin{shaded}
+\verb|name=''arguments string''|
+\end{shaded}
 Following commands are defined:
@@
 \begin{itemize}
-\item \textbf{\textit{joblog}}\textit{={[}path]} - specifies where to store
+\item \textbf{\textit{joblog}}\textit{={[}path]} -- specifies where to store
 log file containing information about started and finished jobs.
-\item \textbf{\textit{jobreport}}\textit{={[}URL ... number]} - specifies
+\item \textbf{\textit{jobreport}}\textit{={[}URL ... number]} -- specifies
 that A-REX has to report information about jobs being processed (started,
@@
 is specified in days. Last specified value becomes effective.
-\item \textbf{\textit{securetransfer}}\textit{=yes|no} - specifies whether
+\item \textbf{\textit{securetransfer}}\textit{=yes|no} -- specifies whether
 to use encryption while transferring data. Currently works for GridFTP
@@
 URL options.
-\item \textbf{\textit{localtransfer}}\textit{=yes|no} - specifies whether
+\item \textbf{\textit{localtransfer}}\textit{=yes|no} -- specifies whether
 to pass file downloading/uploading task to computing node. If set
@@
 different stages:\\
-\textit{max\_processed\_jobs} - maximal amount of jobs being processed
+\textit{max\_processed\_jobs} -- maximal amount of jobs being processed
 by A-REX. This does not limit amount of jobs, which can be submitted
 to A-REX\\
-\textit{max\_running jobs} - maximal amount of jobs passed to Local
+\textit{max\_running jobs} -- maximal amount of jobs passed to Local
 Resource Management System\\
@@
 \item \textbf{\textit{maxlod}}\textit{={[}max\_frontend\_jobs {[}emergency\_frontend\_jobs
-{[}max\_transferred\_files]]]} - specifies maximum load caused by
+{[}max\_transferred\_files]]]} -- specifies maximum load caused by
 jobs being processed on frontend:\\
-\textit{max\_frontend\_jobs} - maximal amount of jobs heavily using
+\textit{max\_frontend\_jobs} -- maximal amount of jobs heavily using
 resources of frontend (applied before moving job to PREPARING and
 FINISHING states)\textit{}\\
-\textit{emergency\_frontend\_jobs} - if limit of \textit{max\_frontend\_jobs}
+\textit{emergency\_frontend\_jobs} -- if limit of \textit{max\_frontend\_jobs}
 is used only by PREPARING or by FINISHING jobs aforementioned number
-of jobs can be moved to another state .This is used to avoid case
+of jobs can be moved to another state. This is used to avoid case
 then jobs can't finish due to big amount of recently submitted jobs.\textit{}\\
-\textit{max\_transfered\_files} - maximal number of files being transferred
+\textit{max\_transfered\_files} -- maximal number of files being transferred
 in parallel by every job. Used to decrease load on not so powerful
@@
 Missing value or -1 means no limit.
-\item \textbf{\textit{wakeupperiod}}\textit{=time} - specifies how often
+\item \textbf{\textit{wakeupperiod}}\textit{=time} -- specifies how often
 the A-REX checks for job state changes (like new arrived job, job
@@
 in seconds. Default is \emph{3 minutes}.
-\item \textbf{\textit{cacheregistration}}\textit{=yes|no} - enables or disables
+\item \textbf{\textit{cacheregistration}}\textit{=yes|no} -- enables or disables
 registration of cache data into Indexing Services like RC or RLS.
@@
 location is registered at Indexing Service with quasi-url \emph{cache://hostname/}
-and name \emph{hostname:cache} .
-\item \textbf{\textit{authplugin}}\textit{=state options plugin} - specifies
+and name \emph{hostname:cache}.
+\item \textbf{\textit{authplugin}}\textit{=state options plugin} -- specifies
 \emph{plugin} (external executable) to be run every time job is going
@@
 pairs separated by a comma. Following \textit{name}s are supported:\\
-\textit{timeout} - specifies how long in seconds execution of the
+\textit{timeout} -- specifies how long in seconds execution of the
 plugin allowed to last (mandatory, {}``\textit{timeout=}{}`` can
 be skipped for backward compatibility).\\
-\textit{onsuccess}, \textit{onfailure} and \textit{ontimeout} - defines
+\textit{onsuccess}, \textit{onfailure} and \textit{ontimeout} -- defines
 action taken in each case (\textit{onsuccess} happens if exit code
 is 0). Possible actions are:\\
-\textit{pass} - continue execution,\\
-\textit{log} - write information about result into log file and continue
+\textit{pass} -- continue execution,\\
+\textit{log} -- write information about result into log file and continue
 execution,\\
-\textit{fail} - write information about result into log file and cancel
+\textit{fail} -- write information about result into log file and cancel
 job.
-\item \textbf{\textit{localcred}}\textit{=timeout plugin} - specifies \emph{plugin}
+\item \textbf{\textit{localcred}}\textit{=timeout plugin} -- specifies \emph{plugin}
 (external executable or function in shared library) to be run every
@@
 code is not 0 current operation will fail.
-\item \textbf{\textit{norootpower}}\textit{=yes/no} - if set to \emph{yes}
+\item \textbf{\textit{norootpower}}\textit{=yes/no} -- if set to \emph{yes}
 all processes involved in job management will use local identity of
@@
 \item \textbf{\textit{speedcontrol}}\textit{=min\_speed min\_time min\_average\_speed
-max\_inactivity} - specifies how long/slow data transfer is allowed
+max\_inactivity} -- specifies how long/slow data transfer is allowed
 to take place. Transfer is canceled if transfer rate (bytes per second)
@@
 data is received for longer than \textit{max\_inactivity} seconds.
-\item \textbf{\textit{copyurl}}\textit{=template replacement} - specifies
+\item \textbf{\textit{copyurl}}\textit{=template replacement} -- specifies
 that URLs, starting from template should be accessed in a different
@@
 \item \textbf{\textit{linkurl}}\textit{=template replacement {[}node\_path]}
-- mostly identical to \textit{copyurl} but file won't be copied. Instead
+-- mostly identical to \textit{copyurl} but file won't be copied. Instead
 soft-link will be created. \textit{replacement} specifies the way
@@
 computing nodes, and will be used for soft-link creation. If \textit{node\_path}
-is missing - \textit{local\_path} will be used instead. Both \textit{node\_path}
+is missing, \textit{local\_path} will be used instead. Both \textit{node\_path}
 and \textit{replacement} should not be URLs.
-
+\begin{framed}
 NOTE: URLs which fit into \textit{copyurl} or \textit{linkurl} are
@@
 file, these will be tried first.
-
-\end{itemize}
-\noindent Per UNIX user commands:
-
-\begin{itemize}
-\item \textbf{\textit{mail}}\textit{=e-mail\_address} - specifies an email
+\end{framed}
+\end{itemize}
+
+Per UNIX user commands:
+
+\begin{itemize}
+\item \textbf{\textit{mail}}\textit{=e-mail\_address} -- specifies an email
 address \textbf{from} which the notification mails are sent.
-\item \textbf{\textit{defaultttl}}\textit{=ttl {[}ttr]} - specifies the
+\item \textbf{\textit{defaultttl}}\textit{=ttl {[}ttr]} -- specifies the
 time in seconds for the SD to be available after job finished (\emph{ttl})
@@
 the JD).
-\item \textbf{\textit{session}}\textit{=path} - specifies path to the directory
+\item \textbf{\textit{session}}\textit{=path} -- specifies path to the directory
 in which the SD is created. If the path is {*} the default one is
-used - \textit{\$HOME/.jobs} . In new configuration file this command
+used -- \textit{\$HOME/.jobs}. In new configuration file this command
 is called \textbf{\textit{sessiondir}}.
-\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} - specifies
+\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} -- specifies
 the directory to store cached data. Empty \emph{path} disables caching.
@@
 is set to '.' files are not soft-linked, but copied to session directory.
-\item \textbf{\textit{privatecache}}\textit{=path {[}link\_path]} - same
+\item \textbf{\textit{privatecache}}\textit{=path {[}link\_path]} -- same
 as \textit{cache} command, but cache belongs (owned) to user. For
 shared caches use 'cache'.
-\item \textbf{\textit{cachedata}}\textit{=path} - allows to specify separate
+\item \textbf{\textit{cachedata}}\textit{=path} -- allows to specify separate
 place to store cache files containing data itself. This can be useful
@@
 command, together with control files. 
-\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} - specifies
+\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} -- specifies
 high and low water-mark for space used by cache. Values are specified
@@
 file system.
-\item \textbf{\textit{maxrerun}}\textit{=number} - specifies maximal number
+\item \textbf{\textit{maxrerun}}\textit{=number} -- specifies maximal number
 of times job will be allowed to rerun after it failed at any stage.
@@
 \end{itemize}
+
 All per-user commands should be put before \textit{control} command
@@
 - This option initiates UNIX user as being serviced by the A-REX.
-\textit{path} refers to the control directory (see section \ref{section: internals}
+\textit{path} refers to the control directory (see Section~\ref{section:internals}
 for the description of control directory). If the path is {*} the
-default one is used - \$HOME/.jobstatus . \textit{username} stands
+default one is used -- \verb|$HOME/.jobstatus|. \textit{username} stands
 for UNIX name of the local user. Multiple names can be specified.
-If the name is {*} it is substituted by all names found in file /etc/grid-security/grid-mapfile
-(for the format of this file one should study the Globus project \cite{globus}).\\
+If the name is {*} it is substituted by all names found in file \texttt{/etc/grid-security/grid-mapfile}
+(for the format of this file one should study the Globus project~\cite{globus}).\\
 Also the special name '.'(dot) can be used. Corresponding control
@@
 \item \textbf{\textit{helper}}\textit{=username command {[}argument {[}argument
-{[}...]]]} - associates external program with the local UNIX user.
+{[}...]]]} -- associates external program with the local UNIX user.
 This program will be kept running under account of the specified user.
 \textit{username} stands for the name of the user. Special names can
-be used: '{*}' - all names from /etc/grid-security/grid-mapfile, '.'
+be used: '{*}' -- all names from /etc/grid-security/grid-mapfile, '.'
 - root user. The user should be already configured with \textit{control}
@@
 \end{itemize}
+
 \begin{comment}
@@
 \begin{itemize}
-\item \textbf{\textit{pbs\_bin\_path}}\textit{=path} - path to directory
+\item \textbf{\textit{pbs\_bin\_path}}\textit{=path} -- path to directory
 which contains PBS commands.
-\item \textbf{\textit{pbs\_log\_path}}\textit{=path} - path to directory
+\item \textbf{\textit{pbs\_log\_path}}\textit{=path} -- path to directory
 with PBS server's log files.
@@
 \begin{itemize}
-\item \textbf{\textit{gnu\_time}}\textit{=path} - path to \emph{time} utility.
-\item \textbf{\textit{tmpdir}}\textit{=path} - path to directory for temporary
+\item \textbf{\textit{gnu\_time}}\textit{=path} -- path to \emph{time} utility.
+\item \textbf{\textit{tmpdir}}\textit{=path} -- path to directory for temporary
 files.
-\item \textbf{\textit{runtimedir}}\textit{=path} - path to directory which
+\item \textbf{\textit{runtimedir}}\textit{=path} -- path to directory which
 contains \emph{runtimenvironment} scripts.
-\item \textbf{\textit{shared\_filesystem}}\textit{=yes|no} - if computing
+\item \textbf{\textit{shared\_filesystem}}\textit{=yes|no} -- if computing
 nodes have an access to session directory through a shared file system
 like NFS. \begin{comment} Corresponds to an environment variable
-RUNTIME\_NODE\_SEES\_FRONTEND (see section \ref{sub:LRMS}). \end{comment}
-\item \textbf{\textit{nodename}}\textit{=command} - command to obtain hostname
+RUNTIME\_NODE\_SEES\_FRONTEND (see Section~\ref{sub:LRMS}). \end{comment}
+\item \textbf{\textit{nodename}}\textit{=command} -- command to obtain hostname
 of computing node.
-\item \textbf{\textit{scratchdir}}\textit{=path} - path on computing node
+\item \textbf{\textit{scratchdir}}\textit{=path} -- path on computing node
 where to move session directory before execution.
-\item \textbf{\textit{shared\_scratch}}\textit{=path} - path on frontend
+\item \textbf{\textit{shared\_scratch}}\textit{=path} -- path on frontend
 where \textbf{\textit{scratchdir}} can be found.
 \end{itemize}
+
 In the command arguments (paths, executables, ...) following substitutions
@@
 
-\begin{lyxlist}{00.00.0000}
-\item [{\%R}] - session root - see command \emph{session}
-\item [{\%C}] - control dir - see command \emph{control}
-\item [{\%U}] - username
-\item [{\%u}] - userid - numerical
-\item [{\%g}] - groupid - numerical
-\item [{\%H}] - home dir - home specified in /etc/passwd
-\item [{\%Q}] - default queue - look command 'defaultlrms'
-\item [{\%L}] - default lrms - look command 'defaultlrms'
-\item [{\%W}] - installation path - \$\{ARC\_LOCATION\}
-\item [{\%G}] - globus path - \$\{GLOBUS\_LOCATION\}
-\item [{\%c}] - list of all control directories
-\item [{\%I}] - job's ID (for plugins only, substituted in runtime)
-\item [{\%S}] - job's state (for \emph{authplugin} plugins only, substituted
+\begin{itemize}
+\item [{\%R}] -- session root -- see command \emph{session}
+\item [{\%C}] -- control dir -- see command \emph{control}
+\item [{\%U}] -- username
+\item [{\%u}] -- userid -- numerical
+\item [{\%g}] -- groupid -- numerical
+\item [{\%H}] -- home dir -- home specified in /etc/passwd
+\item [{\%Q}] -- default queue -- look command 'defaultlrms'
+\item [{\%L}] -- default lrms -- look command 'defaultlrms'
+\item [{\%W}] -- installation path -- \$\{ARC\_LOCATION\}
+\item [{\%G}] -- globus path -- \$\{GLOBUS\_LOCATION\}
+\item [{\%c}] -- list of all control directories
+\item [{\%I}] -- job's ID (for plugins only, substituted in runtime)
+\item [{\%S}] -- job's state (for \emph{authplugin} plugins only, substituted
 in runtime)
-\item [{\%O}] - reason (for \emph{localcred} plugins only, substituted
-in runtime).\\
-Possible reasons are:
-
-\begin{lyxlist}{00.00.0000}
-\item [{new}] - new job, new credentials
-\item [{renew}] - old job, new credentials
-\item [{write}] - write/delete file, create/delete directory
-\item [{read}] - read file, directory, etc.
-\item [{extern}] - call external program 
-\end{lyxlist}
-\end{lyxlist}
+\item [{\%O}] -- reason (for \emph{localcred} plugins only, substituted
+in runtime). Possible reasons are:
+\begin{list}{--}{\setlength{\labelwidth}{0.5cm}\setlength{\rightmargin}{\leftmargin}}
+\item [{new}] -- new job, new credentials
+\item [{renew}] -- old job, new credentials
+\item [{write}] -- write/delete file, create/delete directory
+\item [{read}] -- read file, directory, etc.
+\item [{extern}] -- call external program 
+\end{list}
+\end{itemize}
 
@@
 
+\subsection{LRMS support\label{sub:LRMS}}
+
+For information about supported LRMSes and their specific features
+and configuration options please read dedicated documentation~\cite{backends-arc1}.
+
 \begin{comment}(LRMS\_COMMENT)
-
-\subsection{\label{sub:LRMS}LRMS support}
 
@@
 
-\textbf{\textit{PBS\_BIN\_PATH}} - path to PBS executables. Like \emph{/usr/local/bin}
+\textbf{\textit{PBS\_BIN\_PATH}} -- path to PBS executables. Like \emph{/usr/local/bin}
 for example. \emph{}Corresponds to \emph{pbs\_bin\_path} configuration
@@
 
-\textbf{\textit{PBS\_LOG\_PATH}} - path to PBS server logs. Corresponds
+\textbf{\textit{PBS\_LOG\_PATH}} -- path to PBS server logs. Corresponds
 to \emph{pbs\_log\_path} configuration command.
 
-\textbf{\textit{TMP\_DIR}} - path to directory to store temporary
+\textbf{\textit{TMP\_DIR}} -- path to directory to store temporary
 files. Default value is \emph{/tmp}. Corresponds to \emph{tmpdir}
@@
 
-\textbf{\textit{RUNTIME\_CONFIG\_DIR}} - path where runtime setup
+\textbf{\textit{RUNTIME\_CONFIG\_DIR}} -- path where runtime setup
 scripts can be found. Corresponds to \emph{runtimedir} configuration
@@
 
-\textbf{\textit{GNU\_TIME}} - path to GNU time utility. It is important
+\textbf{\textit{GNU\_TIME}} -- path to GNU time utility. It is important
 to provide path to utility compatible with GNU time. If such utility
@@
 
-\textbf{\textit{NODENAME}} - command to obtain name of cluster's node.
+\textbf{\textit{NODENAME}} -- command to obtain name of cluster's node.
 Default is \emph{/bin/hostname -f}. Corresponds to \emph{nodename}
@@
 
-\textbf{\textit{RUNTIME\_LOCAL\_SCRATCH\_DIR}} - if defined should
+\textbf{\textit{RUNTIME\_LOCAL\_SCRATCH\_DIR}} -- if defined should
 contain path to the directory on computing node, which can be used
@@
 
-\textbf{\textit{RUNTIME\_FRONTEND\_SEES\_NODE}} - if defined should
+\textbf{\textit{RUNTIME\_FRONTEND\_SEES\_NODE}} -- if defined should
 contain path corresponding to \textit{}\\
@@
 
-\textbf{\textit{RUNTIME\_NODE\_SEES\_FRONTEND}} - if set to {}``\emph{no}''
+\textbf{\textit{RUNTIME\_NODE\_SEES\_FRONTEND}} -- if set to {}``\emph{no}''
 means computing node does not share file system with frontend. In
@@
 
-Figures \ref{fig:no node scratch},\ref{fig:node scratch, not vis on front},\ref{fig:node scratch, vis on front}
+Figures~\ref{fig:no-node-scratch},~\ref{fig:node-scratch-not-vis-on-front},~\ref{fig:node-scratch-vis-on-front}
 present some possible combinations for RUNTIME\_LOCAL\_SCRATCH\_DIR
@@
 %
-\begin{figure}[tbh]
+\begin{figure}[ht]
 \begin{centering}
-\includegraphics{pic2}
-\par\end{centering}
-
-\caption{\label{fig:no node scratch}Both RUNTIME\_LOCAL\_SCRATCH\_DIR and
+\includegraphics{pic2.pdf}
+\end{centering}
+\caption{\label{fig:no-node-scratch}Both RUNTIME\_LOCAL\_SCRATCH\_DIR and
 RUNTIME\_FRONTEND\_SEES\_NODE are undefined. It is assumed that session
@@
 %
-\begin{figure}[htbp]
+\begin{figure}[ht]
 \begin{centering}
-\includegraphics{pic3}
-\par\end{centering}
-
-\caption{\label{fig:node scratch, not vis on front}RUNTIME\_LOCAL\_SCRATCH\_DIR
+\includegraphics{pic3.pdf}
+\end{centering}
+\caption{\label{fig:node-scratch-not-vis-on-front}RUNTIME\_LOCAL\_SCRATCH\_DIR
 is set to the location of the scratch directory on the computing node,
@@
 
-\begin{lyxlist}{00.00.0000}
-\item [{a)}] After the job script starts all input files are moved to 'scratch
+\begin{itemize}
+\item [{a)}] After the job script starts all the input files are moved to the 'scratch
 directory' on the computing node.
-\item [{b)}] The job runs in a separate directory in 'scratch directory'.
+\item [{b)}] The job runs in a separate directory in the 'scratch directory'.
 Only the files representing the job's \textit{stdout} and \textit{stderr}
 are placed in the original 'session directory' and soft-linked in
-'scratch'. After execution all files from 'scratch' are moved back
+'scratch'. After execution all the files from 'scratch' are moved back
 to the original 'session directory'.
-\item [{c)}] All output files are in 'session directory' and are ready
+\item [{c)}] All output files are in the 'session directory' and are ready
 to be uploaded/downloaded.
-\end{lyxlist}
+\end{itemize}
 
@@
 %
-\begin{figure}[htbp]
+\begin{figure}[ht]
 \begin{centering}
-\includegraphics{pic4}
-\par\end{centering}
-
-\caption{\label{fig:node scratch, vis on front}Both RUNTIME\_LOCAL\_SCRATCH\_DIR
+\includegraphics{pic4.pdf}
+\end{centering}
+\caption{\label{fig:node-scratch-vis-on-front}Both RUNTIME\_LOCAL\_SCRATCH\_DIR
 and RUNTIME\_FRONTEND\_SEES\_NODE are set to the location of the scratch
 directory on the computing node and the location where this scratch
-directory is accessible from the frontend, respectively.}
-
-\begin{lyxlist}{00.00.0000}
-\item [{a)}] After the job script starts all input files are moved to 'scratch
+directory is accessible from the frontend, respectively}.
+
+\begin{itemize}
+\item [{a)}] After the job script starts all input files are moved to the 'scratch
 directory' on the computing node. The original 'session directory'
@@
 from 'scratch' are moved back to the original 'session directory'.
-\item [{c)}] All output files are in 'session directory' and are ready
+\item [{c)}] All output files are in the 'session directory' and are ready
 to be uploaded/downloaded. 
-\end{lyxlist}
+\end{itemize}
 
@@
 
-The A-REX can run specially prepared \emph{BASH} scripts prior creation
-of job's script, before and after executing job's main executable.
-Those scripts are requested by user through \emph{runtimeenvironment}
-attribute in RSL and are run with only argument set equal to '0',
-'1' or '2' during creation of job's script, before execution of main
-executable and after main executable finished accordingly. They all
+The A-REX can run specially prepared \emph{BASH} scripts prior to creation
+of the job's script, before and after executing job's main executable.
+Those scripts are requested by the user through the \emph{runtimeenvironment}
+attribute in JSDL and are run with the only argument set either equal to '0',
+'1' or '2' during creation of the job's script, before execution of the main
+executable and after main the executable is finished, respectively. They all
 are run through BASH's 'source' command, and hence can manipulate
-shell variables. With argument '0' scripts are run by the A-REX on
+shell variables. With argument '0' scripts are run by the A-REX on the
 frontend. Some environment variables are defined in that case and
@@
 \begin{itemize}
-\item joboption\_directory - session directory.
-\item joboption\_arg\_\# - command with arguments to be executed as specified
+\item joboption\_directory -- session directory.
+\item joboption\_arg\_\# -- command with arguments to be executed as specified
 in the JD (\textbf{not} bash array).
-\item joboption\_env\_\# - array of 'NAME=VALUE' environment variables (\textbf{not}
+\item joboption\_env\_\# -- array of 'NAME=VALUE' environment variables (\textbf{not}
 bash array).
-\item joboption\_runtime\_\# - array of requested \emph{runtimeenvironment}
+\item joboption\_runtime\_\# -- array of requested \emph{runtimeenvironment}
 names (\textbf{not} bash array).
-\item joboption\_num - \emph{runtimeenvironment} currently beeing processed
+\item joboption\_num -- \emph{runtimeenvironment} currently beeing processed
 (number starting from 0).
-\item joboption\_stdin - name of file to be attached to stdin handle.
-\item joboption\_stdout - same for stdout.
-\item joboption\_stderr - same for stderr.
-\item joboption\_maxcputime - amount of CPU time requested (minutes).
-\item joboption\_maxmemory - amount of memory requested (megabytes).
-\item joboption\_count - number of processors requested.
-\item joboption\_lrms - LRMS to be used to run job.
-\item joboption\_queue - name of a queue of LRMS to put job into.
-\item joboption\_nodeproperty\_\# - array of properties of computing nodes
+\item joboption\_stdin -- name of file to be attached to stdin handle.
+\item joboption\_stdout -- same for stdout.
+\item joboption\_stderr -- same for stderr.
+\item joboption\_maxcputime -- amount of CPU time requested (minutes).
+\item joboption\_maxmemory -- amount of memory requested (megabytes).
+\item joboption\_count -- number of processors requested.
+\item joboption\_lrms -- LRMS to be used to run job.
+\item joboption\_queue -- name of a queue of LRMS to put job into.
+\item joboption\_nodeproperty\_\# -- array of properties of computing nodes
 (LRMS specific, \textbf{not} bash array).
-\item joboption\_jobname - name of the job as given by user.
-\item joboption\_rsl - whole RSL for very clever submission scripts.
-\item joboption\_rsl\_\emph{name} - RSL attributes and values (like joboption\_rsl\_executable=''/bin/echo'')
-\end{itemize}
-For example \emph{joboption\_arg\_\#} could be changed to wrap main
+\item joboption\_jobname -- name of the job as given by user.
+\item joboption\_rsl -- whole RSL for very clever submission scripts.
+\item joboption\_rsl\_\emph{name} -- RSL attributes and values (like joboption\_rsl\_executable=''/bin/echo'')
+\end{itemize}
+
+For example \emph{joboption\_arg\_\#} could be changed to wrap the main
 executable. Or \emph{joboption\_runtime} could be expanded if current
@@
 
-With argument '1' scripts are run just before main executable is run.
-They are executed on computing node. Such script can prepare environment
+With argument '1' scripts are run just before the main executable is run.
+They are executed on the computing node. Such a script can prepare environment
 for some third-party software package. A current directory in that
-case is one which would be used for execution of job. Variable HOME
+case is the one which would be used for execution of the job. Variable \verb|$HOME|
 also points to that directory.
@@
 
-For publicly available runtime environments please see RTE \url{http://gridrer.csc.fi/}
-.
+For publicly available runtime environments please see the RTE repository at \url{http://gridrer.csc.fi/}.
 
@@
 
-The A-REX is installed as part of ARC 1 middleware and is available
-in NorduGrid download area at ARC1 Download \url{http://download.nordugrid.org/software/nordugrid-arc1/}
-.There are packages available for various Linux distributions. The
-A-REX comes in nordugrid-arc1-arex package. Source code ready for
+The A-REX is installed as an ARC1 component of ARC middleware and is available
+in NorduGrid download area at ARC1 Download \url{http://download.nordugrid.org/software/nordugrid-arc1/}.There are packages available for various Linux distributions. The
+A-REX comes in the \texttt{nordugrid-arc1-arex} package. Source code ready for
 compilation is available too.
@@
 
-If installed from binary packages all dependencies are handled automatically.
-For compilation from source code please read included README file.
+When installed from binary packages, all the dependencies are handled automatically.
+For compilation from source code please read included README files.
 
@@
 The A-REX service is a pluggable module of the HED. So first You need
-to setup your HED - see HED documentation.
-
-Then You need to add A-REX to the HED configuration. Add new <Name>
-element inside <Plugins> containing string \emph{arex}. That will
+to setup your HED -- see HED documentation.
+
+Then You need to add A-REX to the HED configuration. Add new \verb|<Name>|
+element inside \verb|<Plugins>| containing string \emph{arex}. That will
 make HED load libarex plugin library.
 
-Then add new <Service> element with attribute name=''a-rex''. That
+Then add new \verb|<Service>| element with attribute \verb|name=``a-rex''|. That
 will instantiate A-REX service. Now to make service accessible extend
-<Plexer> element with new <next> referring to an id of the service.
-Write <Service> element carefully - here is an example:
-
-\begin{lyxcode}
-~~~~<!-{}-~A-Rex~service~-{}->
-
-~~~~<Service~name=\char`\"{}a-rex\char`\"{}~id=\char`\"{}a-rex\char`\"{}>
-
-~~~~~~<!-{}-~Optional~endpoint~element~is~advised~in~case~of~multiple~IP~adresses~-{}->
-
-~~~~~~<arex:endpoint>https://localhost:60000/arex</arex:endpoint>
-
-~~~~~~<!-{}-~Use~information~generated~by~identity.map~plugin~or~default~provided~below~-{}->
-
-~~~~~~<arex:usermap><arex:defaultLocalName>nobody</arex:defaultLocalName></arex:usermap>
-
-~~~~~~<!-{}-~grid-manager~part~of~a-rex~requires~legacy~configuration~file.
-
-~~~~~~~~~~~Use~arc\_arex.conf~example~or~write~own.~-{}->
-
-~~~~~~<arex:gmconfig>/etc/arc\_arex.conf</arex:gmconfig>
-
-~~~~</Service>
-\end{lyxcode}
-For in-depth information about available elements see Appendix B.
+\verb|<Plexer>| element with new \verb|<next>| referring to an id of the service.
+Write \verb|<Service>| element carefully -- here is an example:
+
+\begin{shaded}
+\begin{verbatim}
+    <!-{}- A-Rex service -{}->
+    <Service name=\char`\"{}a-rex\char`\"{} id=\char`\"{}a-rex\char`\"{}>
+      <!-{}- Optional endpoint element is advised in case of multiple IP adresses -{}->
+      <arex:endpoint>https://localhost:60000/arex</arex:endpoint>
+      <!-{}- Use information generated by identity.map plugin or default provided below -{}->
+      <arex:usermap><arex:defaultLocalName>nobody</arex:defaultLocalName></arex:usermap>
+      <!-{}- grid-manager part of a-rex requires legacy configuration file.
+           Use arc\_arex.conf example or write own. -{}->
+      <arex:gmconfig>/etc/arc\_arex.conf</arex:gmconfig>
+    </Service>
+\end{verbatim}
+\end{shaded}
+
+For in-depth information about available elements see Appendix~\ref{annex:arex-conf}.
 
@@
 For information about format and available configuration commands
-see Section \ref{SubSection: ConfigFile}.
+see Section~\ref{SubSection:ConfigFile}.
 
@@
 
-Please read \emph{User Guide} \cite{userguide1} for usage instructions.
+Please read \emph{User Guide}~\cite{userguide1} for usage instructions.
 
@@
 
-\section{Session directory access through HTTP(S) interface}
+\section{Session directory access through HTTP(S) interface\label{annex:a}}
 
@@
 
-The BES defines job identifier as WS Addressing \cite{ws-addr-soap}
-Endpoint Reference (EPR) - XML document. The EPR is extendable and
-the A-REX adds it own element JobSessionDir belonging to namespace
-http://www.nordugrid.org/schemas/a-rex as direct child of ReferenceParameters
+The BES defines job identifier as WS Addressing~\cite{ws-addr-soap}
+Endpoint Reference (EPR) -- XML document. The EPR is extendable and 
+the A-REX adds it own element \texttt{JobSessionDir} belonging to the namespace
+\texttt{http://www.nordugrid.org/schemas/a-rex} as a direct child of ReferenceParameters
 element. This new element contains the URL of SD.
@@
 HTTP methods PUT and GET may be used to upload/download content of
-those files. For directories - including SD itself - GET method is
+those files. For directories -- including SD itself -- GET method is
 supported which returns HTML encoded non-recursive list of files and
 directories. The files and subdirectories have their URLs inside HTML
-element <A>.
-
-\section{Configuration schema of A-REX}
-
-\begin{lyxcode}
-<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
-
+element \verb|<A>|.
+
+\section{Configuration schema of A-REX\label{annex:arex-conf}}
+
+\begin{footnotesize}
+\begin{verbatim}
+<?xml version=\char`\"{}1.0\char`\"{} encoding=\char`\"{}UTF-8\char`\"{}?>
 <xsd:schema
-
 xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-
 xmlns=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
-
 xmlns:arc=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
-
 targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
-
 elementFormDefault=\char`\"{}qualified\char`\"{}>
-
-~~<xsd:complexType~name=\char`\"{}endpoint\_Type\char`\"{}>
-
-~~~~<!-{}-
-
-~~~~This~element~defines~URL~of~A-REX~service~as~seen~from~outside.
-
-~~~~-{}->
-
-~~~~<xsd:simpleContent>
-
-~~~~~~<xsd:extension~base=\char`\"{}xsd:string\char`\"{}>~
-
-~~~~~~</xsd:extension>
-
-~~~~</xsd:simpleContent>
-
-~~</xsd:complexType>
-
-~~<xsd:element~name=\char`\"{}endpoint\char`\"{}~type=\char`\"{}endpoint\_Type\char`\"{}/>
-
-~~<xsd:complexType~name=\char`\"{}gmconfig\_Type\char`\"{}>
-
-~~~~<!-{}-
-
-~~~~This~element~defines~path~to~arc0~Grid~Manager~configuartion~file.
-
-~~~~By~default~it~is~/etc/arc.conf.
-
-~~~~-{}->
-
-~~~~<xsd:simpleContent>
-
-~~~~~~<xsd:extension~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~~~~</xsd:extension>
-
-~~~~</xsd:simpleContent>
-
-~~</xsd:complexType>
-
-~~<xsd:element~name=\char`\"{}gmconfig\char`\"{}~type=\char`\"{}gmconfig\_Type\char`\"{}/>
-
-~~<xsd:simpleType~name=\char`\"{}gmrun\_Type\char`\"{}>
-
-~~~~<!-{}-
-
-~~~~This~element~defines~how~grid-manager~part~of~A-Rex~is~run.
-
-~~~~{*}~internal~-~as~a~thread~inside~service~container.
-
-~~~~{*}~none~-~no~grid-manager~is~run.
-
-~~~~{*}~external~-~as~a~separate~executable~(not~supported~anymore).
-
-~~~~Default~is~'internal'.
-
-~~~~-{}->
-
-~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~~~~<xsd:enumeration~value=\char`\"{}internal\char`\"{}/>
-
-~~~~~~<xsd:enumeration~value=\char`\"{}external\char`\"{}/>
-
-~~~~~~<xsd:enumeration~value=\char`\"{}none\char`\"{}/>
-
-~~~~</xsd:restriction>
-
-~~</xsd:simpleType>
-
-~~<xsd:element~name=\char`\"{}gmrun\char`\"{}~type=\char`\"{}gmrun\_Type\char`\"{}/>
-
-~~<xsd:complexType~name=\char`\"{}usermap\_Type\char`\"{}>
-
-~~~~<xsd:sequence>
-
-~~~~~~<xsd:element~name=\char`\"{}defaultLocalName\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~</xsd:sequence>
-
-~~</xsd:complexType>
-
-~~<xsd:element~name=\char`\"{}usermap\char`\"{}~type=\char`\"{}usermap\_Type\char`\"{}/>
-
-~~<!-{}-~CommonName~attribute~of~bes-factory.~-{}->
-
-~~<xsd:element~name=\char`\"{}commonName\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~<!-{}-~LongDescription~attribute~of~bes-factory.~-{}->
-
-~~<xsd:element~name=\char`\"{}longDescription\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~<!-{}-~Name~of~Local~Resource~Management~System.~-{}->
-
-~~<xsd:element~name=\char`\"{}LRMSName\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~<!-{}-
-
-~~Name~of~Operating~System.
-
-~~The~values~are~based~on~the~OSType~field~of~the~CIM\_OperatingSystem~model:
-
-~~http://www.dmtf.org/standards/cim/cim\_schema\_v29
-
-~~-{}->
-
-~~<xsd:element~name=\char`\"{}OperatingSystem\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
+  <xsd:complexType name=\char`\"{}endpoint\_Type\char`\"{}>
+    <!-{}-
+    This element defines URL of A-REX service as seen from outside.
+    -{}->
+    <xsd:simpleContent>
+      <xsd:extension base=\char`\"{}xsd:string\char`\"{}> 
+      </xsd:extension>
+    </xsd:simpleContent>
+  </xsd:complexType>
+  <xsd:element name=\char`\"{}endpoint\char`\"{} type=\char`\"{}endpoint\_Type\char`\"{}/>
+  <xsd:complexType name=\char`\"{}gmconfig\_Type\char`\"{}>
+    <!-{}-
+    This element defines path to arc0 Grid Manager configuartion file.
+    By default it is /etc/arc.conf.
+    -{}->
+    <xsd:simpleContent>
+      <xsd:extension base=\char`\"{}xsd:string\char`\"{}>
+      </xsd:extension>
+    </xsd:simpleContent>
+  </xsd:complexType>
+  <xsd:element name=\char`\"{}gmconfig\char`\"{} type=\char`\"{}gmconfig\_Type\char`\"{}/>
+  <xsd:simpleType name=\char`\"{}gmrun\_Type\char`\"{}>
+    <!-{}-
+    This element defines how grid-manager part of A-Rex is run.
+    {*} internal - as a thread inside service container.
+    {*} none - no grid-manager is run.
+    {*} external - as a separate executable (not supported anymore).
+    Default is 'internal'.
+    -{}->
+    <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
+      <xsd:enumeration value=\char`\"{}internal\char`\"{}/>
+      <xsd:enumeration value=\char`\"{}external\char`\"{}/>
+      <xsd:enumeration value=\char`\"{}none\char`\"{}/>
+    </xsd:restriction>
+  </xsd:simpleType>
+  <xsd:element name=\char`\"{}gmrun\char`\"{} type=\char`\"{}gmrun\_Type\char`\"{}/>
+  <xsd:complexType name=\char`\"{}usermap\_Type\char`\"{}>
+    <xsd:sequence>
+      <xsd:element name=\char`\"{}defaultLocalName\char`\"{} type=\char`\"{}xsd:string\char`\"{}
+                   minOccurs=\char`\"{}0\char`\"{}/>
+    </xsd:sequence>
+  </xsd:complexType>
+  <xsd:element name=\char`\"{}usermap\char`\"{} type=\char`\"{}usermap\_Type\char`\"{}/>
+  <!-{}- CommonName attribute of bes-factory. -{}->
+  <xsd:element name=\char`\"{}commonName\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+  <!-{}- LongDescription attribute of bes-factory. -{}->
+  <xsd:element name=\char`\"{}longDescription\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+  <!-{}- Name of Local Resource Management System. -{}->
+  <xsd:element name=\char`\"{}LRMSName\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+  <!-{}-
+  Name of Operating System.
+  The values are based on the OSType field of the CIM\_OperatingSystem model:
+  http://www.dmtf.org/standards/cim/cim\_schema\_v29
+  -{}->
+  <xsd:element name=\char`\"{}OperatingSystem\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
 </xsd:schema>
-\end{lyxcode}
-
-\section{A-REX WSDL}
-
-\begin{lyxcode}
-<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
-
-<wsdl:definitions~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}
-
-~xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
-
-~xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
-
-~xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
-
-~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-
-~xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
-
-~xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
-
-~xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
-
-~xmlns:bes-factory=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}
-
-~xmlns:bes-mgmt=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}
-
-~xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
-
-~xmlns:wsrf-rpw=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}
-
-~xmlns:a-rex=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
-
-~~<wsdl:import~namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}~location=\char`\"{}./bes-factory.wsdl\char`\"{}/>
-
-~~<wsdl:import~namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}~location=\char`\"{}./bes-management.wsdl\char`\"{}/>
-
-~~<wsdl:import~namespase=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}~location=\char`\"{}../schemas/delegation.wsdl\char`\"{}/>
-
-~~<wsdl:import~namespase=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}~location=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2.wsdl\char`\"{}/>
-
-~~<wsdl:types>
-
-~~~~<xsd:schema~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
-
-~~~~~~<xsd:import~namespace=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}~schemaLocation=\char`\"{}./ws-addr.xsd\char`\"{}/>
-
-~~~~~~<xsd:simpleType~name=\char`\"{}ActivitySubStateType\char`\"{}>
-
-~~~~~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Accepting\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Accepted\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Preparing\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Prepared\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Submiting\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Executing\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Killing\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Executed\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Finishing\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Finished\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Failed\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Deleted\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Pending\char`\"{}/>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}Held\char`\"{}/>
-
-~~~~~~~~</xsd:restriction>
-
-~~~~~~</xsd:simpleType>
-
-~~~~~~<xsd:element~name=\char`\"{}State\char`\"{}~type=\char`\"{}a-rex:ActivitySubStateType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ResourceInformationDocumentType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~<xsd:element~name=\char`\"{}BESFactory\char`\"{}~type=\char`\"{}bes-factory:FactoryResourceAttributesDocumentType\char`\"{}/>
-
-~~~~~~~~~~<xsd:complexType~name=\char`\"{}Glue2Resource\char`\"{}~minOccurs='0'>
-
-~~~~~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~~~~~~~~~~</xsd:sequence>
-
-~~~~~~~~~~</xsd:complexType>
-
-~~~~~~~~~~<xsd:complexType~name=\char`\"{}Activities\char`\"{}~minOccurs='0'>
-
-~~~~~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~~~~<xsd:complexType~name=\char`\"{}Activity\char`\"{}~minOccurs='0'~maxOccurs='unbounded'>
-
-~~~~~~~~~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~~~~~~~~<xsd:element~name=\char`\"{}ActivityIdentifier\char`\"{}~type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
-
-~~~~~~~~~~~~~~~~~~<xsd:element~ref=\char`\"{}bes-factory:ActivityDocument\char`\"{}~minOccurs='0'/>
-
-~~~~~~~~~~~~~~~~~~<xsd:complexType~name=\char`\"{}Glue2Job\char`\"{}~minOccurs='0'>
-
-~~~~~~~~~~~~~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~~~~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~~~~~~~~~~~~~~~~~~</xsd:sequence>
-
-~~~~~~~~~~~~~~~~~~</xsd:complexType>
-
-~~~~~~~~~~~~~~~~</xsd:sequence>
-
-~~~~~~~~~~~~~~</xsd:complexType>
-
-~~~~~~~~~~~~</xsd:sequence>
-
-~~~~~~~~~~</xsd:complexType>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ChangeActivityStatusRequestType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}ActivityIdentifier\char`\"{}~type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}OldStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}NewStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}ChangeActivityStatus\char`\"{}~type=\char`\"{}a-rex:ChangeActivityStatusRequestType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ChangeActivityStatusResponseType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}NewStatus\char`\"{}~type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}~type=\char`\"{}a-rex:ChangeActivityStatusResponseType\char`\"{}/>
-
-~~~~</xsd:schema>
-
-~~</wsdl:types>
-
-~~<wsdl:message~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}~element=\char`\"{}a-rex:ChangeActivityStatus\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}~element=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:portType~name=\char`\"{}a-rex\char`\"{}>
-
-~~~~<wsdl:operation~name=\char`\"{}ChangeActivityStatus\char`\"{}>
-
-~~~~~~<wsdl:documentation>
-
-~~~~~~~~This~operation~allows~any~simple~status~change~request
-
-~~~~~~~~which~involves~no~additional~parameters.~It~should~be
-
-~~~~~~~~used~to~modify~job/activity~execution~flow:
-
-~~~~~~~~~~-~To~put~job~on~hold
-
-~~~~~~~~~~-~To~rerun~job~in~case~of~failure
-
-~~~~~~~~~~-~To~cancel~job~(same~as~TerminateActivity~of~BESFActory)
-
-~~~~~~~~~~-~To~remove/release~job~-~as~long~as~non-existence~is~a~state
-
-~~~~~~~~~~-~Any~other~status~change~no~supported~by~BES
-
-~~~~~~</wsdl:documentation>
-
-~~~~~~<wsdl:input~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}
-
-~~~~~~~~message=\char`\"{}a-rex:ChangeActivityStatusRequest\char`\"{}/>
-
-~~~~~~<wsdl:output~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
-
-~~~~~~~~message=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}bes-factory:NotAuthorizedFault\char`\"{}
-
-~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}bes-factory:InvalidActivityIdentifierFault\char`\"{}
-
-~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}
-
-~~~~~~~~~message=\char`\"{}bes-factory:CantApplyOperationToCurrentStateFault\char`\"{}
-
-~~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}
-
-~~~~~~~~~message=\char`\"{}bes-factory:OperationWillBeAppliedEventuallyFault\char`\"{}
-
-~~~~~~~~~wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:portType>
-
-~~<wsdl:binding~name=\char`\"{}a-rex\char`\"{}~type=\char`\"{}a-rex:a-rex\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}ChangeActivityStatus\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}ChangeActivityStatus\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
-
-~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
-
-~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~~~<wsdl:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}NotAuthorizedFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-~~<wsdl:binding~name=\char`\"{}GetResourcePropertyDocument\char`\"{}~type=\char`\"{}wsrf-rpw:GetResourcePropertyDocument\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}GetResourcePropertyDocument\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}GetResourcePropertyDocument\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentRequest\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentResponse\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-~~<wsdl:binding~name=\char`\"{}GetResourceProperty\char`\"{}~type=\char`\"{}wsrf-rpw:GetResourceProperty\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}GetResourceProperty\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}GetResourceProperty\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:GetResourcePropertyRequest\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:GetResourcePropertyResponse\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-~~<wsdl:binding~name=\char`\"{}QueryResourceProperties\char`\"{}~type=\char`\"{}wsrf:QueryResourceProperties\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}QueryResourceProperties\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}QueryResourceProperties\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}wsrf-rpw:QueryResourcePropertiesRequest\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}wsrf-rpw:QueryResourcePropertiesResponse\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnknownFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}ResourceUnavailabbleFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}InvalidQueryExpressionFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}InvalidQueryExpressionFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~~~<wsdl:fault~name=\char`\"{}QueryEvaluationErrorFault\char`\"{}>
-
-~~~~~~~~<soap:fault~name=\char`\"{}QueryEvaluationErrorFault\char`\"{}~use=\char`\"{}literal\char`\"{}~/>
-
-~~~~~~</wsdl:fault>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
-~~<wsdl:service~name=\char`\"{}a-rex\char`\"{}>
-
-~~~~<wsdl:port~name=\char`\"{}delegation\char`\"{}~binding=\char`\"{}deleg:DelegationBinding\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}bes-factory\char`\"{}~binding=\char`\"{}bes-factory:BESFactoryBinding\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}bes-mgmt\char`\"{}~binding=\char`\"{}bes-mgmt:BESManagementBinding\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}GetResourcePropertyDocument\char`\"{}~binding=\char`\"{}a-rex:GetResourcePropertyDocument\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}GetResourceProperty\char`\"{}~binding=\char`\"{}a-rex:GetResourceProperty\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}QueryResourceProperties\char`\"{}~binding=\char`\"{}a-rex:QueryResourceProperties\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~~~<wsdl:port~name=\char`\"{}a-rex\char`\"{}~binding=\char`\"{}a-rex:a-rex\char`\"{}>
-
-~~~~</wsdl:port>
-
-~~</wsdl:service>
-
+\end{verbatim}
+\end{footnotesize}
+
+\section{A-REX WSDL\label{annex:arex-wsdl}}
+
+\begin{footnotesize}\begin{verbatim}
+<?xml version=\char`\"{}1.0\char`\"{} encoding=\char`\"{}UTF-8\char`\"{}?>
+<wsdl:definitions targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}
+ xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
+ xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
+ xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
+ xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
+ xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
+ xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
+ xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
+ xmlns:bes-factory=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}
+ xmlns:bes-mgmt=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}
+ xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
+ xmlns:wsrf-rpw=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}
+ xmlns:a-rex=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
+  <wsdl:import namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}
+               location=\char`\"{}./bes-factory.wsdl\char`\"{}/>
+  <wsdl:import namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}
+               location=\char`\"{}./bes-management.wsdl\char`\"{}/>
+  <wsdl:import namespase=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
+               location=\char`\"{}../schemas/delegation.wsdl\char`\"{}/>
+  <wsdl:import namespase=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}
+               location=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2.wsdl\char`\"{}/>
+  <wsdl:types>
+    <xsd:schema targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
+      <xsd:import namespace=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
+                  schemaLocation=\char`\"{}./ws-addr.xsd\char`\"{}/>
+      <xsd:simpleType name=\char`\"{}ActivitySubStateType\char`\"{}>
+        <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
+          <xsd:enumeration value=\char`\"{}Accepting\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Accepted\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Preparing\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Prepared\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Submiting\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Executing\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Killing\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Executed\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Finishing\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Finished\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Failed\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Deleted\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Pending\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Held\char`\"{}/>
+        </xsd:restriction>
+      </xsd:simpleType>
+      <xsd:element name=\char`\"{}State\char`\"{} type=\char`\"{}a-rex:ActivitySubStateType\char`\"{}/>
+      <xsd:complexType name=\char`\"{}ResourceInformationDocumentType\char`\"{}>
+        <xsd:sequence>
+           <xsd:element name=\char`\"{}BESFactory\char`\"{}
+                        type=\char`\"{}bes-factory:FactoryResourceAttributesDocumentType\char`\"{}/>
+          <xsd:complexType name=\char`\"{}Glue2Resource\char`\"{} minOccurs='0'>
+            <xsd:sequence>
+              <xsd:any namespace=\char`\"{}\#\#other\char`\"{} processContents=\char`\"{}lax\char`\"{}
+                       minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>
+            </xsd:sequence>
+          </xsd:complexType>
+          <xsd:complexType name=\char`\"{}Activities\char`\"{} minOccurs='0'>
+            <xsd:sequence>
+              <xsd:complexType name=\char`\"{}Activity\char`\"{} minOccurs='0' maxOccurs='unbounded'>
+                <xsd:sequence>
+                  <xsd:element name=\char`\"{}ActivityIdentifier\char`\"{}
+                               type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
+                  <xsd:element ref=\char`\"{}bes-factory:ActivityDocument\char`\"{} minOccurs='0'/>
+                  <xsd:complexType name=\char`\"{}Glue2Job\char`\"{} minOccurs='0'>
+                    <xsd:sequence>
+                      <xsd:any namespace=\char`\"{}\#\#other\char`\"{} processContents=\char`\"{}lax\char`\"{}
+                               minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>
+                    </xsd:sequence>
+                  </xsd:complexType>
+                </xsd:sequence>
+              </xsd:complexType>
+            </xsd:sequence>
+          </xsd:complexType>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:complexType name=\char`\"{}ChangeActivityStatusRequestType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}ActivityIdentifier\char`\"{} type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
+          <xsd:element name=\char`\"{}OldStatus\char`\"{} type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}
+                       minOccurs=\char`\"{}0\char`\"{}/>
+          <xsd:element name=\char`\"{}NewStatus\char`\"{} type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}ChangeActivityStatus\char`\"{}
+                   type=\char`\"{}a-rex:ChangeActivityStatusRequestType\char`\"{}/>
+      <xsd:complexType name=\char`\"{}ChangeActivityStatusResponseType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}NewStatus\char`\"{} type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
+                   type=\char`\"{}a-rex:ChangeActivityStatusResponseType\char`\"{}/>
+    </xsd:schema>
+  </wsdl:types>
+  <wsdl:message name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
+    <wsdl:part name=\char`\"{}ChangeActivityStatusRequest\char`\"{}
+               element=\char`\"{}a-rex:ChangeActivityStatus\char`\"{}/>
+  </wsdl:message>
+  <wsdl:message name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
+    <wsdl:part name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
+               element=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
+  </wsdl:message>
+  <wsdl:portType name=\char`\"{}a-rex\char`\"{}>
+    <wsdl:operation name=\char`\"{}ChangeActivityStatus\char`\"{}>
+      <wsdl:documentation>
+        This operation allows any simple status change request
+        which involves no additional parameters. It should be
+        used to modify job/activity execution flow:
+          - To put job on hold
+          - To rerun job in case of failure
+          - To cancel job (same as TerminateActivity of BESFActory)
+          - To remove/release job - as long as non-existence is a state
+          - Any other status change no supported by BES
+      </wsdl:documentation>
+      <wsdl:input name=\char`\"{}ChangeActivityStatusRequest\char`\"{}
+        message=\char`\"{}a-rex:ChangeActivityStatusRequest\char`\"{}/>
+      <wsdl:output name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
+        message=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
+      <wsdl:fault name=\char`\"{}NotAuthorizedFault\char`\"{}
+        message=\char`\"{}bes-factory:NotAuthorizedFault\char`\"{}
+        wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}
+        message=\char`\"{}bes-factory:InvalidActivityIdentifierFault\char`\"{}
+        wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}
+         message=\char`\"{}bes-factory:CantApplyOperationToCurrentStateFault\char`\"{}
+         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}
+         message=\char`\"{}bes-factory:OperationWillBeAppliedEventuallyFault\char`\"{}
+         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+    </wsdl:operation>
+  </wsdl:portType>
+  <wsdl:binding name=\char`\"{}a-rex\char`\"{} type=\char`\"{}a-rex:a-rex\char`\"{}>
+    <soap:binding style=\char`\"{}document\char`\"{}
+                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+    <wsdl:operation name=\char`\"{}ChangeActivityStatus\char`\"{}>
+      <soap:operation soapAction=\char`\"{}ChangeActivityStatus\char`\"{}/>
+      <wsdl:input name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
+         <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:input>
+      <wsdl:output name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
+         <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:output>
+      <wsdl:fault name=\char`\"{}NotAuthorizedFault\char`\"{}>
+        <soap:fault name=\char`\"{}NotAuthorizedFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}>
+        <soap:fault name=\char`\"{}InvalidActivityIdentifierFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}>
+        <soap:fault name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}>
+        <soap:fault name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+    </wsdl:operation>
+  </wsdl:binding>
+  <wsdl:binding name=\char`\"{}GetResourcePropertyDocument\char`\"{}
+                type=\char`\"{}wsrf-rpw:GetResourcePropertyDocument\char`\"{}>
+    <soap:binding style=\char`\"{}document\char`\"{}
+                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+    <wsdl:operation name=\char`\"{}GetResourcePropertyDocument\char`\"{}>
+      <soap:operation soapAction=\char`\"{}GetResourcePropertyDocument\char`\"{}/>
+      <wsdl:input name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentRequest\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:input>
+      <wsdl:output name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentResponse\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:output>
+      <wsdl:fault name=\char`\"{}ResourceUnknownFault\char`\"{}>
+        <soap:fault name=\char`\"{}ResourceUnknownFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}ResourceUnavailableFault\char`\"{}>
+        <soap:fault name=\char`\"{}ResourceUnavailabbleFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+    </wsdl:operation>
+  </wsdl:binding>
+  <wsdl:binding name=\char`\"{}GetResourceProperty\char`\"{} type=\char`\"{}wsrf-rpw:GetResourceProperty\char`\"{}>
+    <soap:binding style=\char`\"{}document\char`\"{}
+                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+    <wsdl:operation name=\char`\"{}GetResourceProperty\char`\"{}>
+      <soap:operation soapAction=\char`\"{}GetResourceProperty\char`\"{}/>
+      <wsdl:input name=\char`\"{}wsrf-rpw:GetResourcePropertyRequest\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:input>
+      <wsdl:output name=\char`\"{}wsrf-rpw:GetResourcePropertyResponse\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:output>
+      <wsdl:fault name=\char`\"{}ResourceUnknownFault\char`\"{}>
+        <soap:fault name=\char`\"{}ResourceUnknownFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}ResourceUnavailableFault\char`\"{}>
+        <soap:fault name=\char`\"{}ResourceUnavailabbleFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
+        <soap:fault name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+    </wsdl:operation>
+  </wsdl:binding>
+  <wsdl:binding name=\char`\"{}QueryResourceProperties\char`\"{} type=\char`\"{}wsrf:QueryResourceProperties\char`\"{}>
+    <soap:binding style=\char`\"{}document\char`\"{}
+                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+    <wsdl:operation name=\char`\"{}QueryResourceProperties\char`\"{}>
+      <soap:operation soapAction=\char`\"{}QueryResourceProperties\char`\"{}/>
+      <wsdl:input name=\char`\"{}wsrf-rpw:QueryResourcePropertiesRequest\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:input>
+      <wsdl:output name=\char`\"{}wsrf-rpw:QueryResourcePropertiesResponse\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:output>
+      <wsdl:fault name=\char`\"{}ResourceUnknownFault\char`\"{}>
+        <soap:fault name=\char`\"{}ResourceUnknownFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}ResourceUnavailableFault\char`\"{}>
+        <soap:fault name=\char`\"{}ResourceUnavailabbleFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
+        <soap:fault name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}>
+        <soap:fault name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}InvalidQueryExpressionFault\char`\"{}>
+        <soap:fault name=\char`\"{}InvalidQueryExpressionFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}QueryEvaluationErrorFault\char`\"{}>
+        <soap:fault name=\char`\"{}QueryEvaluationErrorFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+    </wsdl:operation>
+  </wsdl:binding>
+  <wsdl:service name=\char`\"{}a-rex\char`\"{}>
+    <wsdl:port name=\char`\"{}delegation\char`\"{} binding=\char`\"{}deleg:DelegationBinding\char`\"{}>
+    </wsdl:port>
+    <wsdl:port name=\char`\"{}bes-factory\char`\"{} binding=\char`\"{}bes-factory:BESFactoryBinding\char`\"{}>
+    </wsdl:port>
+    <wsdl:port name=\char`\"{}bes-mgmt\char`\"{} binding=\char`\"{}bes-mgmt:BESManagementBinding\char`\"{}>
+    </wsdl:port>
+    <wsdl:port name=\char`\"{}GetResourcePropertyDocument\char`\"{}
+               binding=\char`\"{}a-rex:GetResourcePropertyDocument\char`\"{}>
+    </wsdl:port>
+    <wsdl:port name=\char`\"{}GetResourceProperty\char`\"{} binding=\char`\"{}a-rex:GetResourceProperty\char`\"{}>
+    </wsdl:port>
+    <wsdl:port name=\char`\"{}QueryResourceProperties\char`\"{}
+               binding=\char`\"{}a-rex:QueryResourceProperties\char`\"{}>
+    </wsdl:port>
+    <wsdl:port name=\char`\"{}a-rex\char`\"{} binding=\char`\"{}a-rex:a-rex\char`\"{}>
+    </wsdl:port>
+  </wsdl:service>
 </wsdl:definitions>
-\end{lyxcode}
-
-\section{Delegation WSDL}
-
-\begin{lyxcode}
-<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
-
-<wsdl:definitions~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
-
-~xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
-
-~xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
-
-~xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
-
-~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-
-~xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
-
-~xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
-
-~xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
-
-~xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
-
-~~<wsdl:types>
-
-~~~~<xsd:schema~targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
-
-~~~~~~<!-{}-~Common~types~-{}->
-
-~~~~~~<xsd:simpleType~name=\char`\"{}TokenFormatType\char`\"{}>
-
-~~~~~~~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~~~~~~~~<xsd:enumeration~value=\char`\"{}x509\char`\"{}/>
-
-~~~~~~~~</xsd:restriction>
-
-~~~~~~</xsd:simpleType>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ReferenceType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:any~namespace=\char`\"{}\#\#other\char`\"{}~processContents=\char`\"{}lax\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:complexType~name=\char`\"{}DelegatedTokenType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Id\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Value\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Reference\char`\"{}~type=\char`\"{}deleg:ReferenceType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~~~<xsd:attribute~name=\char`\"{}Format\char`\"{}~type=\char`\"{}deleg:TokenFormatType\char`\"{}~use=\char`\"{}required\char`\"{}/>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}DelegatedToken\char`\"{}~type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}TokenRequestType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Id\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Value\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~~~<xsd:attribute~name=\char`\"{}Format\char`\"{}~type=\char`\"{}deleg:TokenFormatType\char`\"{}~use=\char`\"{}required\char`\"{}/>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}TokenRequest\char`\"{}~type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
-
-~~~~~~<!-{}-~Types~for~messages~-{}->
-
-~~~~~~<xsd:complexType~name=\char`\"{}DelegateCredentialsInitRequestType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}DelegateCredentialsInit\char`\"{}~type=\char`\"{}deleg:DelegateCredentialsInitRequestType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}DelegateCredentialsInitResponseType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}TokenRequest\char`\"{}~type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}~type=\char`\"{}deleg:DelegateCredentialsInitResponseType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}UpdateCredentialsRequestType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}DelegatedToken\char`\"{}~type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}UpdateCredentials\char`\"{}~type=\char`\"{}deleg:UpdateCredentialsRequestType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}UpdateCredentialsResponseType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}UpdateCredentialsResponse\char`\"{}~type=\char`\"{}deleg:UpdateCredentialsResponseType\char`\"{}/>
-
-~~~~~~<!-{}-~Faults~-{}->
-
-~~~~~~<xsd:complexType~name=\char`\"{}UnsupportedFaultType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}UnsupportedFault\char`\"{}~type=\char`\"{}deleg:UnsupportedFaultType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}ProcessingFaultType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}ProcessingFault\char`\"{}~type=\char`\"{}deleg:ProcessingFaultType\char`\"{}/>
-
-~~~~~~<xsd:complexType~name=\char`\"{}WrongReferenceFaultType\char`\"{}>
-
-~~~~~~~~<xsd:sequence>
-
-~~~~~~~~~~<xsd:element~name=\char`\"{}Description\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~~~~~~</xsd:sequence>
-
-~~~~~~</xsd:complexType>
-
-~~~~~~<xsd:element~name=\char`\"{}WrongReferenceFault\char`\"{}~type=\char`\"{}deleg:WrongReferenceFaultType\char`\"{}/>
-
-~~~~</xsd:schema>
-
-~~</wsdl:types>
-
-~~<wsdl:message~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}~element=\char`\"{}deleg:DelegateCredentialsInit\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}~element=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}UpdateCredentialsRequest\char`\"{}~element=\char`\"{}deleg:UpdateCredentials\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}UpdateCredentialsResponse\char`\"{}~element=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}UnsupportedFault\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}ProcessingFault\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:ProcessingFault\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:message~name=\char`\"{}WrongReferenceFault\char`\"{}>
-
-~~~~<wsdl:part~name=\char`\"{}Detail\char`\"{}~element=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
-
-~~</wsdl:message>
-
-~~<wsdl:portType~name=\char`\"{}DelegationPortType\char`\"{}>
-
-~~~<wsdl:operation~name=\char`\"{}DelegateCredentialsInit\char`\"{}>
-
-~~~~~~<wsdl:documentation>
-
-~~~~~~</wsdl:documentation>
-
-~~~~~~<wsdl:input~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:DelegateCredentialsInitRequest\char`\"{}/>
-
-~~~~~~<wsdl:output~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}UnsupportedFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ProcessingFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
-
-~~~~</wsdl:operation>
-
-~~~<wsdl:operation~name=\char`\"{}UpdateCredentials\char`\"{}>
-
-~~~~~~<wsdl:documentation>
-
-~~~~~~</wsdl:documentation>
-
-~~~~~~<wsdl:input~name=\char`\"{}UpdateCredentialsRequest\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:UpdateCredentialsRequest\char`\"{}/>
-
-~~~~~~<wsdl:output~name=\char`\"{}UpdateCredentialsResponse\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}UnsupportedFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}ProcessingFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
-
-~~~~~~<wsdl:fault~name=\char`\"{}WrongReferenceFault\char`\"{}
-
-~~~~~~~~message=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:portType>
-
-~~<wsdl:binding~name=\char`\"{}DelegationBinding\char`\"{}~type=\char`\"{}deleg:DelegationPortType\char`\"{}>
-
-~~~~<soap:binding~style=\char`\"{}document\char`\"{}~transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-
-~~~~<wsdl:operation~name=\char`\"{}DelegateCredentialsInit\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}DelegateCredentialsInit\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
-
-~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
-
-~~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~</wsdl:operation>
-
-~~~~<wsdl:operation~name=\char`\"{}UpdateCredentials\char`\"{}>
-
-~~~~~~<soap:operation~soapAction=\char`\"{}UpdateCredentials\char`\"{}/>
-
-~~~~~~<wsdl:input~name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:input>
-
-~~~~~~<wsdl:output~name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
-
-~~~~~~~~<soap:body~use=\char`\"{}literal\char`\"{}/>
-
-~~~~~~</wsdl:output>
-
-~~~~</wsdl:operation>
-
-~~</wsdl:binding>
-
+\end{verbatim}
+\end{footnotesize}
+
+\section{Delegation WSDL\label{annex:delegation-wsdl}}
+
+\begin{footnotesize}\begin{verbatim}
+<?xml version=\char`\"{}1.0\char`\"{} encoding=\char`\"{}UTF-8\char`\"{}?>
+<wsdl:definitions targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
+ xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
+ xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
+ xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
+ xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
+ xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
+ xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
+ xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
+ xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
+  <wsdl:types>
+    <xsd:schema targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
+      <!-{}- Common types -{}->
+      <xsd:simpleType name=\char`\"{}TokenFormatType\char`\"{}>
+        <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
+          <xsd:enumeration value=\char`\"{}x509\char`\"{}/>
+        </xsd:restriction>
+      </xsd:simpleType>
+      <xsd:complexType name=\char`\"{}ReferenceType\char`\"{}>
+        <xsd:sequence>
+          <xsd:any namespace=\char`\"{}\#\#other\char`\"{} processContents=\char`\"{}lax\char`\"{}
+                   minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:complexType name=\char`\"{}DelegatedTokenType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}Id\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+          <xsd:element name=\char`\"{}Value\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+          <xsd:element name=\char`\"{}Reference\char`\"{} type=\char`\"{}deleg:ReferenceType\char`\"{}
+                       minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>
+        </xsd:sequence>
+        <xsd:attribute name=\char`\"{}Format\char`\"{} type=\char`\"{}deleg:TokenFormatType\char`\"{}
+                       use=\char`\"{}required\char`\"{}/>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}DelegatedToken\char`\"{} type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
+      <xsd:complexType name=\char`\"{}TokenRequestType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}Id\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+          <xsd:element name=\char`\"{}Value\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+        </xsd:sequence>
+        <xsd:attribute name=\char`\"{}Format\char`\"{} type=\char`\"{}deleg:TokenFormatType\char`\"{}
+                       use=\char`\"{}required\char`\"{}/>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}TokenRequest\char`\"{} type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
+      <!-{}- Types for messages -{}->
+      <xsd:complexType name=\char`\"{}DelegateCredentialsInitRequestType\char`\"{}>
+        <xsd:sequence>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}DelegateCredentialsInit\char`\"{}
+                   type=\char`\"{}deleg:DelegateCredentialsInitRequestType\char`\"{}/>
+      <xsd:complexType name=\char`\"{}DelegateCredentialsInitResponseType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}TokenRequest\char`\"{} type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
+                   type=\char`\"{}deleg:DelegateCredentialsInitResponseType\char`\"{}/>
+      <xsd:complexType name=\char`\"{}UpdateCredentialsRequestType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}DelegatedToken\char`\"{} type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}UpdateCredentials\char`\"{}
+                   type=\char`\"{}deleg:UpdateCredentialsRequestType\char`\"{}/>
+      <xsd:complexType name=\char`\"{}UpdateCredentialsResponseType\char`\"{}>
+        <xsd:sequence>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}UpdateCredentialsResponse\char`\"{}
+                   type=\char`\"{}deleg:UpdateCredentialsResponseType\char`\"{}/>
+      <!-{}- Faults -{}->
+      <xsd:complexType name=\char`\"{}UnsupportedFaultType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}Description\char`\"{} type=\char`\"{}xsd:string\char`\"{}
+                       minOccurs=\char`\"{}0\char`\"{}/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}UnsupportedFault\char`\"{} type=\char`\"{}deleg:UnsupportedFaultType\char`\"{}/>
+      <xsd:complexType name=\char`\"{}ProcessingFaultType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}Description\char`\"{} type=\char`\"{}xsd:string\char`\"{}
+                       minOccurs=\char`\"{}0\char`\"{}/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}ProcessingFault\char`\"{} type=\char`\"{}deleg:ProcessingFaultType\char`\"{}/>
+      <xsd:complexType name=\char`\"{}WrongReferenceFaultType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}Description\char`\"{} type=\char`\"{}xsd:string\char`\"{}
+                       minOccurs=\char`\"{}0\char`\"{}/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name=\char`\"{}WrongReferenceFault\char`\"{} type=\char`\"{}deleg:WrongReferenceFaultType\char`\"{}/>
+    </xsd:schema>
+  </wsdl:types>
+  <wsdl:message name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
+    <wsdl:part name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}
+               element=\char`\"{}deleg:DelegateCredentialsInit\char`\"{}/>
+  </wsdl:message>
+  <wsdl:message name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
+    <wsdl:part name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
+               element=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
+  </wsdl:message>
+  <wsdl:message name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
+    <wsdl:part name=\char`\"{}UpdateCredentialsRequest\char`\"{} element=\char`\"{}deleg:UpdateCredentials\char`\"{}/>
+  </wsdl:message>
+  <wsdl:message name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
+    <wsdl:part name=\char`\"{}UpdateCredentialsResponse\char`\"{}
+               element=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
+  </wsdl:message>
+  <wsdl:message name=\char`\"{}UnsupportedFault\char`\"{}>
+    <wsdl:part name=\char`\"{}Detail\char`\"{} element=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
+  </wsdl:message>
+  <wsdl:message name=\char`\"{}ProcessingFault\char`\"{}>
+    <wsdl:part name=\char`\"{}Detail\char`\"{} element=\char`\"{}deleg:ProcessingFault\char`\"{}/>
+  </wsdl:message>
+  <wsdl:message name=\char`\"{}WrongReferenceFault\char`\"{}>
+    <wsdl:part name=\char`\"{}Detail\char`\"{} element=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
+  </wsdl:message>
+  <wsdl:portType name=\char`\"{}DelegationPortType\char`\"{}>
+   <wsdl:operation name=\char`\"{}DelegateCredentialsInit\char`\"{}>
+      <wsdl:documentation>
+      </wsdl:documentation>
+      <wsdl:input name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}
+        message=\char`\"{}deleg:DelegateCredentialsInitRequest\char`\"{}/>
+      <wsdl:output name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
+        message=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
+      <wsdl:fault name=\char`\"{}UnsupportedFault\char`\"{}
+        message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}ProcessingFault\char`\"{}
+        message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
+    </wsdl:operation>
+   <wsdl:operation name=\char`\"{}UpdateCredentials\char`\"{}>
+      <wsdl:documentation>
+      </wsdl:documentation>
+      <wsdl:input name=\char`\"{}UpdateCredentialsRequest\char`\"{}
+        message=\char`\"{}deleg:UpdateCredentialsRequest\char`\"{}/>
+      <wsdl:output name=\char`\"{}UpdateCredentialsResponse\char`\"{}
+        message=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
+      <wsdl:fault name=\char`\"{}UnsupportedFault\char`\"{}
+        message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}ProcessingFault\char`\"{}
+        message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}WrongReferenceFault\char`\"{}
+        message=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
+    </wsdl:operation>
+  </wsdl:portType>
+  <wsdl:binding name=\char`\"{}DelegationBinding\char`\"{} type=\char`\"{}deleg:DelegationPortType\char`\"{}>
+    <soap:binding style=\char`\"{}document\char`\"{}
+                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
+    <wsdl:operation name=\char`\"{}DelegateCredentialsInit\char`\"{}>
+      <soap:operation soapAction=\char`\"{}DelegateCredentialsInit\char`\"{}/>
+      <wsdl:input name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
+         <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:input>
+      <wsdl:output name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
+         <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:output>
+    </wsdl:operation>
+    <wsdl:operation name=\char`\"{}UpdateCredentials\char`\"{}>
+      <soap:operation soapAction=\char`\"{}UpdateCredentials\char`\"{}/>
+      <wsdl:input name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:input>
+      <wsdl:output name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{}/>
+      </wsdl:output>
+    </wsdl:operation>
+  </wsdl:binding>
 </wsdl:definitions>
-\end{lyxcode}
-
-\section{ARC extensions for JSDL schema}
-
-\begin{lyxcode}
-<?xml~version=\char`\"{}1.0\char`\"{}~encoding=\char`\"{}UTF-8\char`\"{}?>
-
-<xsd:schema~xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-
-~~~~~~~~~~~~xmlns=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
-
-~~~~~~~~~~~~xmlns:jsdl-arc=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
-
-~~~~~~~~~~~~targetNamespace=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}>
-
-~<xsd:simpleType~name=\char`\"{}GMState\_Type\char`\"{}>
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~<xsd:enumeration~value=\char`\"{}ACCEPTED\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}PREPARING\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}SUBMIT\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}INLRMS\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}FINISHING\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}FINISHED\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}DELETED\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}CANCELING\char`\"{}/>
-
-~~</xsd:restriction>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}Version\_Type\char`\"{}>
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}UpperExclusive\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}LowerExclusive\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Exact\char`\"{}~type=\char`\"{}xsd:string\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Exclusive\char`\"{}~type=\char`\"{}xsd:boolean\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}SessionType\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}SessionType\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~<xsd:enumeration~value=\char`\"{}INTERNAL\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}LIMITED\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}READONLY\char`\"{}/>
-
-~~~<xsd:enumeration~value=\char`\"{}FULL\char`\"{}/>
-
-~~</xsd:restriction>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}IsExecutable\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:DataStaging\_Type~(default:~false)~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}IsExecutable\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:boolean\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}FileParameters\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:DataStaging\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}IsExecutable\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}JoinOutputs\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~(default:~false)~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}JoinOutputs\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:boolean\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}Reruns\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~(default:~false)~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}Reruns\char`\"{}~minOccurs=\char`\"{}0\char`\"{}//~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:integer\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}RunTimeEnvironment\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}RunTimeEnvironment\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}Name\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Version\char`\"{}~type=\char`\"{}Version\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:complexType~name=\char`\"{}Middleware\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}Middleware\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}Name\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Version\char`\"{}~type=\char`\"{}Version\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:complexType~name=\char`\"{}RemoteLogging\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}RemoteLogging\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}3\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}URL\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:anyURI\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:complexType~name=\char`\"{}CredentialServer\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}CredentialServer\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}URL\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:anyURI\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:complexType~name=\char`\"{}LocalLogging\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}LocalLogging\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}Directory\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}AccessControlType\_Type\char`\"{}>
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~<xsd:enumeration~value=\char`\"{}GACL\char`\"{}/>
-
-~~</xsd:restriction>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}AccessControl\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}AccessControl\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}OwnerAlwaysAllowed\char`\"{}~type=\char`\"{}xsd:boolean\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Type\char`\"{}~type=\char`\"{}AccessControlType\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Content\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}NotificationType\_Type\char`\"{}>
-
-~~<xsd:restriction~base=\char`\"{}xsd:string\char`\"{}>
-
-~~~<xsd:enumeration~value=\char`\"{}Email\char`\"{}/>
-
-~~</xsd:restriction>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}Notify\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}Notify\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}3\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}Type\char`\"{}~type=\char`\"{}NotificationType\_Type\char`\"{}~minOccurs=\char`\"{}0\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}Endpoint\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}State\char`\"{}~minOccurs=\char`\"{}1\char`\"{}~maxOccurs=\char`\"{}unbounded\char`\"{}~type=\char`\"{}GMState\_Type\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}SessionLifeTime\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}SessionLifeTime\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:long\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:simpleType~name=\char`\"{}GridTimeLimit\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}GridTimeLimit\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:positiveInteger\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<xsd:complexType~name=\char`\"{}CandidateTarget\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:Resources\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}jsdl-arc:CandidateTarget\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:sequence>
-
-~~~<xsd:element~name=\char`\"{}HostName\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~~<xsd:element~name=\char`\"{}QueueName\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~type=\char`\"{}xsd:string\char`\"{}/>
-
-~~</xsd:sequence>
-
-~</xsd:complexType>
-
-~<xsd:simpleType~name=\char`\"{}Time\_Type\char`\"{}>
-
-~~<xsd:documentation>~For~jsdl:JobDescription\_Type~</xsd:documentation>
-
-~~<!-{}-~xsd:element~ref=\char`\"{}ProcessingStartTime\char`\"{}~minOccurs=\char`\"{}0\char`\"{}~maxOccurs=\char`\"{}1\char`\"{}/~-{}->
-
-~~<xsd:restriction~base=\char`\"{}xsd:dateTime\char`\"{}/>
-
-~</xsd:simpleType>
-
-~<!-{}-=========================================================-{}->
-
-~<xsd:element~name=\char`\"{}IsExecutable\char`\"{}~type=\char`\"{}IsExecutable\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}FileParameters\char`\"{}~type=\char`\"{}FileParameters\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}RunTimeEnvironment\char`\"{}~type=\char`\"{}RunTimeEnvironment\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}Middleware\char`\"{}~type=\char`\"{}Middleware\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}RemoteLogging\char`\"{}~type=\char`\"{}RemoteLogging\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}LocalLogging\char`\"{}~type=\char`\"{}LocalLogging\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}AccessControl\char`\"{}~type=\char`\"{}AccessControl\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}Notify\char`\"{}~type=\char`\"{}Notify\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}SessionLifeTime\char`\"{}~type=\char`\"{}SessionLifeTime\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}SessionType\char`\"{}~type=\char`\"{}SessionType\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}JoinOutputs\char`\"{}~type=\char`\"{}JoinOutputs\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}Reruns\char`\"{}~type=\char`\"{}Reruns\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}CredentialServer\char`\"{}~type=\char`\"{}CredentialServer\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}GridTimeLimit\char`\"{}~type=\char`\"{}GridTimeLimit\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}CandidateTarget\char`\"{}~type=\char`\"{}CandidateTarget\_Type\char`\"{}/>
-
-~<xsd:element~name=\char`\"{}ProcessingStartTime\char`\"{}~type=\char`\"{}Time\_Type\char`\"{}/>
-
+\end{verbatim}
+\end{footnotesize}
+
+\section{ARC extensions for JSDL schema\label{annex:jsdl-extension}}
+
+\begin{footnotesize}\begin{verbatim}
+<?xml version=\char`\"{}1.0\char`\"{} encoding=\char`\"{}UTF-8\char`\"{}?>
+<xsd:schema xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
+            xmlns=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
+            xmlns:jsdl-arc=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
+            targetNamespace=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}>
+ <xsd:simpleType name=\char`\"{}GMState\_Type\char`\"{}>
+  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
+   <xsd:enumeration value=\char`\"{}ACCEPTED\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}PREPARING\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}SUBMIT\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}INLRMS\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}FINISHING\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}FINISHED\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}DELETED\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}CANCELING\char`\"{}/>
+  </xsd:restriction>
+ </xsd:simpleType>
+ <xsd:complexType name=\char`\"{}Version\_Type\char`\"{}>
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}UpperExclusive\char`\"{} type=\char`\"{}xsd:string\char`\"{}
+                minOccurs=\char`\"{}0\char`\"{}/>
+   <xsd:element name=\char`\"{}LowerExclusive\char`\"{} type=\char`\"{}xsd:string\char`\"{}
+                minOccurs=\char`\"{}0\char`\"{}/>
+   <xsd:element name=\char`\"{}Exact\char`\"{} type=\char`\"{}xsd:string\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                maxOccurs=\char`\"{}unbounded\char`\"{}/>
+   <xsd:element name=\char`\"{}Exclusive\char`\"{} type=\char`\"{}xsd:boolean\char`\"{}
+                minOccurs=\char`\"{}0\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name=\char`\"{}SessionType\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}SessionType\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
+  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
+   <xsd:enumeration value=\char`\"{}INTERNAL\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}LIMITED\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}READONLY\char`\"{}/>
+   <xsd:enumeration value=\char`\"{}FULL\char`\"{}/>
+  </xsd:restriction>
+ </xsd:simpleType>
+ <xsd:simpleType name=\char`\"{}IsExecutable\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:DataStaging\_Type (default: false) </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}IsExecutable\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
+  <xsd:restriction base=\char`\"{}xsd:boolean\char`\"{}/>
+ </xsd:simpleType>
+ <xsd:simpleType name=\char`\"{}FileParameters\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:DataStaging\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}IsExecutable\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
+  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}/>
+ </xsd:simpleType>
+ <xsd:simpleType name=\char`\"{}JoinOutputs\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:JobDescription\_Type (default: false) </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}JoinOutputs\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
+  <xsd:restriction base=\char`\"{}xsd:boolean\char`\"{}/>
+ </xsd:simpleType>
+ <xsd:simpleType name=\char`\"{}Reruns\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:JobDescription\_Type (default: false) </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}Reruns\char`\"{} minOccurs=\char`\"{}0\char`\"{}// -{}->
+  <xsd:restriction base=\char`\"{}xsd:integer\char`\"{}/>
+ </xsd:simpleType>
+ <xsd:complexType name=\char`\"{}RunTimeEnvironment\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}RunTimeEnvironment\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                     maxOccurs=\char`\"{}unbounded\char`\"{}/ -{}->
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}Name\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+   <xsd:element name=\char`\"{}Version\char`\"{} type=\char`\"{}Version\_Type\char`\"{}
+                minOccurs=\char`\"{}0\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:complexType name=\char`\"{}Middleware\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}Middleware\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                     maxOccurs=\char`\"{}unbounded\char`\"{}/ -{}->
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}Name\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+   <xsd:element name=\char`\"{}Version\char`\"{} type=\char`\"{}Version\_Type\char`\"{}
+                minOccurs=\char`\"{}0\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:complexType name=\char`\"{}RemoteLogging\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}RemoteLogging\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                     maxOccurs=\char`\"{}3\char`\"{}/ -{}->
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}URL\char`\"{} minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}1\char`\"{}
+                type=\char`\"{}xsd:anyURI\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:complexType name=\char`\"{}CredentialServer\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}CredentialServer\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}URL\char`\"{} minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}1\char`\"{}
+                type=\char`\"{}xsd:anyURI\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:complexType name=\char`\"{}LocalLogging\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}LocalLogging\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}Directory\char`\"{} minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}1\char`\"{}
+                type=\char`\"{}xsd:string\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name=\char`\"{}AccessControlType\_Type\char`\"{}>
+  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
+   <xsd:enumeration value=\char`\"{}GACL\char`\"{}/>
+  </xsd:restriction>
+ </xsd:simpleType>
+ <xsd:complexType name=\char`\"{}AccessControl\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}AccessControl\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}OwnerAlwaysAllowed\char`\"{} type=\char`\"{}xsd:boolean\char`\"{}
+                minOccurs=\char`\"{}0\char`\"{}/>
+   <xsd:element name=\char`\"{}Type\char`\"{} type=\char`\"{}AccessControlType\_Type\char`\"{}
+                minOccurs=\char`\"{}0\char`\"{}/>
+   <xsd:element name=\char`\"{}Content\char`\"{} minOccurs=\char`\"{}0\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name=\char`\"{}NotificationType\_Type\char`\"{}>
+  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
+   <xsd:enumeration value=\char`\"{}Email\char`\"{}/>
+  </xsd:restriction>
+ </xsd:simpleType>
+ <xsd:complexType name=\char`\"{}Notify\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}Notify\char`\"{} minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}3\char`\"{}/
+    -{}->
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}Type\char`\"{} type=\char`\"{}NotificationType\_Type\char`\"{}
+                minOccurs=\char`\"{}0\char`\"{}/>
+   <xsd:element name=\char`\"{}Endpoint\char`\"{} minOccurs=\char`\"{}0\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+   <xsd:element name=\char`\"{}State\char`\"{} minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}
+                type=\char`\"{}GMState\_Type\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name=\char`\"{}SessionLifeTime\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}SessionLifeTime\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
+  <xsd:restriction base=\char`\"{}xsd:long\char`\"{}/>
+ </xsd:simpleType>
+ <xsd:simpleType name=\char`\"{}GridTimeLimit\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}GridTimeLimit\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
+  <xsd:restriction base=\char`\"{}xsd:positiveInteger\char`\"{}/>
+ </xsd:simpleType>
+ <xsd:complexType name=\char`\"{}CandidateTarget\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}jsdl-arc:CandidateTarget\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
+  <xsd:sequence>
+   <xsd:element name=\char`\"{}HostName\char`\"{} minOccurs=\char`\"{}0\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+   <xsd:element name=\char`\"{}QueueName\char`\"{} minOccurs=\char`\"{}0\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name=\char`\"{}Time\_Type\char`\"{}>
+  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
+  <!-{}- xsd:element ref=\char`\"{}ProcessingStartTime\char`\"{} minOccurs=\char`\"{}0\char`\"{}
+                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
+  <xsd:restriction base=\char`\"{}xsd:dateTime\char`\"{}/>
+ </xsd:simpleType>
+ <!-{}-=========================================================-{}->
+ <xsd:element name=\char`\"{}IsExecutable\char`\"{} type=\char`\"{}IsExecutable\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}FileParameters\char`\"{} type=\char`\"{}FileParameters\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}RunTimeEnvironment\char`\"{} type=\char`\"{}RunTimeEnvironment\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}Middleware\char`\"{} type=\char`\"{}Middleware\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}RemoteLogging\char`\"{} type=\char`\"{}RemoteLogging\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}LocalLogging\char`\"{} type=\char`\"{}LocalLogging\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}AccessControl\char`\"{} type=\char`\"{}AccessControl\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}Notify\char`\"{} type=\char`\"{}Notify\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}SessionLifeTime\char`\"{} type=\char`\"{}SessionLifeTime\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}SessionType\char`\"{} type=\char`\"{}SessionType\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}JoinOutputs\char`\"{} type=\char`\"{}JoinOutputs\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}Reruns\char`\"{} type=\char`\"{}Reruns\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}CredentialServer\char`\"{} type=\char`\"{}CredentialServer\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}GridTimeLimit\char`\"{} type=\char`\"{}GridTimeLimit\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}CandidateTarget\char`\"{} type=\char`\"{}CandidateTarget\_Type\char`\"{}/>
+ <xsd:element name=\char`\"{}ProcessingStartTime\char`\"{} type=\char`\"{}Time\_Type\char`\"{}/>
 </xsd:schema>
-\end{lyxcode}
-
-\section{Error messages of A-REX}
+\end{verbatim}
+\end{footnotesize}
+
+\section{Error messages of A-REX\label{annex:arex-errors}}
 
@@
 by the A-REX itself:\\
-\begin{longtable}{|p{2.5in}|p{4in}|}
+\begin{longtable}{|p{5cm}|p{10cm}|}
 \hline 
@@
 \hline 
-Failed in files upload due to expired credentials - try to renew&
+Failed in files upload due to expired credentials -- try to renew&
 Failed to upload some or all output files most probably due to expired
@@
 \hline 
-Failed in files download due to expired credentials - try to renew&
+Failed in files download due to expired credentials -- try to renew&
 Failed to download some or all input files most probably due to expired
@@
 file, FILE by name of file):\\
-\begin{longtable}{|p{2.5in}|p{4in}|}
+\begin{longtable}{|p{5cm}|p{10cm}|}
 \hline 
@@
 \hline 
-Input file: URL - unknown error&
+Input file: URL -- unknown error&
 Generic error \tabularnewline
 \hline 
-Input file: URL - unexpected error&
+Input file: URL -- unexpected error&
 Generic error \tabularnewline
 \hline 
-Input file: URL - bad source URL&
+Input file: URL -- bad source URL&
 Source URL is either malformed or not supported\tabularnewline
 \hline 
-Input file: URL - bad destination URL&
+Input file: URL -- bad destination URL&
 Shouldn't happen\tabularnewline
 \hline 
-Input file: URL - failed to resolve source locations&
+Input file: URL -- failed to resolve source locations&
 File either not registred or other problems related to Data Indexing
@@
 \hline 
-Input file: URL - failed to resolve destination locations&
+Input file: URL -- failed to resolve destination locations&
 Shouldn't happen\tabularnewline
 \hline 
-Input file: URL - failed to register new destination file&
+Input file: URL -- failed to register new destination file&
 Shouldn't happen\tabularnewline
 \hline 
-Input file: URL - can't start reading from source&
+Input file: URL -- can't start reading from source&
 Problems related to accessing instance of file at Data Storing service.\tabularnewline
 \hline 
-Input file: URL - can't read from source&
+Input file: URL -- can't read from source&
 -//-\tabularnewline
 \hline 
-Input file: URL - can't start writing to destination&
+Input file: URL -- can't start writing to destination&
 Access problems in a session directory\tabularnewline
 \hline 
-Input file: URL - can't write to destination&
+Input file: URL -- can't write to destination&
 -//-\tabularnewline
 \hline 
-Input file: URL - data transfer was too slow&
+Input file: URL -- data transfer was too slow&
 Timeouted while trying to download file\tabularnewline
 \hline 
-Input file: URL - failed while closing connection to source&
+Input file: URL -- failed while closing connection to source&
 Shouldn't happen\tabularnewline
 \hline 
-Input file: URL - failed while closing connection to destination&
+Input file: URL -- failed while closing connection to destination&
 Shouldn't happen\tabularnewline
 \hline 
-Input file: URL - failed to register new location&
+Input file: URL -- failed to register new location&
 Shouldn't happen\tabularnewline
 \hline 
-Input file: URL - can't use local cache&
+Input file: URL -- can't use local cache&
 Problems with A-REX cache \tabularnewline
 \hline 
-Input file: URL - system error&
+Input file: URL -- system error&
 Operating System returned error code where unexpected\tabularnewline
 \hline 
-Input file: URL - delegated credentials expired&
+Input file: URL -- delegated credentials expired&
 Access to source requires credententials and they are either outdated
@@
 \hline 
-User file: FILENAME - Bad information about file: checksum can't be
+User file: FILENAME -- Bad information about file: checksum can't be
 parsed.&
@@
 \hline 
-User file: FILENAME - Bad information about file: size can't be parsed.&
+User file: FILENAME -- Bad information about file: size can't be parsed.&
 In job description there is a size provided for file uploadable by
@@
 \hline 
-User file: FILENAME - Expected file. Directory found.&
+User file: FILENAME -- Expected file. Directory found.&
 Instead of file uploadable by user interface A-REX found directory
@@
 \hline 
-User file: FILENAME - Expected ordinary file. Special object found.&
+User file: FILENAME -- Expected ordinary file. Special object found.&
 Instead of file uploadable by user interface A-REX found special object
@@
 \hline 
-User file: FILENAME - Delivered file is bigger than specified.&
+User file: FILENAME -- Delivered file is bigger than specified.&
 The size of file uploadable by user interface is bigger \tabularnewline
 \hline
-User file: FILENAME - Delivered file is unreadable.&
+User file: FILENAME -- Delivered file is unreadable.&
 A-REX can't check user uploadable file due to some internal error.
@@
 \hline
-User file: FILENAME - Could not read file to compute checksum.&
+User file: FILENAME -- Could not read file to compute checksum.&
 A-REX can't read user uploadable file due to some internal error.
@@
 \hline
-User file: FILENAME - Timeout waiting&
+User file: FILENAME -- Timeout waiting&
 A-REX waited for user uploadable file too long.\tabularnewline
@@
 
-\medskip{}
 Provided by uploader component (URL is replaced by destination of
-output file) :\\
-\begin{longtable}{|p{2.5in}|p{4in}|}
+output file) :
+
+\begin{longtable}{|p{5cm}|p{10cm}|}
 \hline 
@@
 \hline 
-Output file: URL - unknown error&
+Output file: URL -- unknown error&
 Generic error \tabularnewline
 \hline 
-Output file: URL - unexpected error&
+Output file: URL -- unexpected error&
 Generic error \tabularnewline
@@
 \hline 
-Output file: URL - bad source URL&
+Output file: URL -- bad source URL&
 Shouldn't happen\tabularnewline
 \hline 
-Output file: URL - bad destination URL&
+Output file: URL -- bad destination URL&
 Destination URL is either malformed or not supported\tabularnewline
 \hline 
-Output file: URL - failed to resolve source locations&
+Output file: URL -- failed to resolve source locations&
 Shouldn't happen\tabularnewline
 \hline 
-Output file: URL - failed to resolve destination locations&
+Output file: URL -- failed to resolve destination locations&
 Problems related to Data Indexing service.\tabularnewline
 \hline 
-Output file: URL - failed to register new destination file&
+Output file: URL -- failed to register new destination file&
 -//-\tabularnewline
 \hline 
-Output file: URL - can't start reading from source&
+Output file: URL -- can't start reading from source&
 User request to store output file, but there is no such file or there
@@
 \hline 
-Output file: URL - can't start writing to destination&
+Output file: URL -- can't start writing to destination&
 Problems with Data Storing services\tabularnewline
 \hline 
-Output file: URL - can't read from source&
+Output file: URL -- can't read from source&
 Problems accessing session directory\tabularnewline
 \hline 
-Output file: URL - can't write to destination&
+Output file: URL -- can't write to destination&
 Problems with Data Storing services\tabularnewline
 \hline 
-Output file: URL - data transfer was too slow&
+Output file: URL -- data transfer was too slow&
 Timeout during transfer\tabularnewline
 \hline 
-Output file: URL - failed while closing connection to source&
+Output file: URL -- failed while closing connection to source&
 Shouldn't happen\tabularnewline
 \hline 
-Output file: URL - failed while closing connection to destination&
+Output file: URL -- failed while closing connection to destination&
 Shouldn't happen\tabularnewline
 \hline 
-Output file: URL - failed to register new location&
+Output file: URL -- failed to register new location&
 Problems related to Data Indexing service.\tabularnewline
 \hline 
-Output file: URL - can't use local cache&
+Output file: URL -- can't use local cache&
 Shouldn't happen\tabularnewline
 \hline 
-Output file: URL - system error&
+Output file: URL -- system error&
 Operating System returned error code where unexpected\tabularnewline
 \hline 
-Output file: URL - delegated credentials expired&
+Output file: URL -- delegated credentials expired&
 Access to destination requires credententials and they are either

### 2008-09-25 doc/client/client_technical.tex
Documentation update from review comments

@@
   argument is a reference to a {\UserConfig} object containing a
-  representing the contents of the user's configuration file. The
+  representation of the contents of the user's configuration file. The
   second and third arguments contain lists of strings. The first list
@@
   index servers or from servers that have already registered
-  once. Index servers often registers at more than one index server,
-  thus different {\TargetRetriever} may discover the same server.)}
+  once. Index servers often register at more than one index server,
+  thus different {\TargetRetriever}s may discover the same server.}
 \item{If while processing the query result the {\TargetRetriever}
   finds an other registered index server or a registered computing
-  service it creates an a new {\TargetRetriever} for the found server
-  or service and forwards the call for targets to the new
+  service it creates a new {\TargetRetriever} for the found server or
+  service and forwards the call for targets to the new
   {\TargetRetriever}.}
@@
   of rejected computing services or from service that have already
-  registered once. Computing and services often registers at more than
-  one index server, thus different {\TargetRetriever} may discover the
-  same service.)}
+  registered once. Computing services often register at more than one
+  index server, thus different {\TargetRetriever}s may discover the
+  same service.}
 \item{When processing the query result the {\TargetRetriever} will
@@
   then store the {\ExecutionTarget} in the found targets store kept
-  by the {\TargetGenerator} for later usage (e.g. status printing or
+  by the {\TargetGenerator} for later usage (e.g.\ status printing or
   job submission).}
@@
   according to the provided {\JobDescription}.}
-\item{Each ExectionTarget has a method to return a specialized
+\item{Each {\ExecutionTarget} has a method to return a specialized
   {\Submitter} which is capable of submitting jobs to the service it
@@
 Once a job is submitted job related information (job identification
-string, cluster etc) is stored in a local xml file which hosts similar
+string, cluster etc) is stored in a local XML file which hosts similar
 information for all active jobs. This file may contain jobs running on
@@
   argument is a reference to a {\UserConfig} object containing a
-  representing the contents of the user's configuration file. The
+  representation of the contents of the user's configuration file. The
   second is a list of strings containing job identifiers and job
@@
   {\JobController} flavours which are to be loaded. The information in
-  the joblist file is also used to lookup the job identifier for jobs
+  the joblist file is also used to look up the job identifier for jobs
   specified using job names. Alias resolution for the selected and
@@
   are passed to each {\JobController} which uses the information to
-  fill its internal JobStore.}
+  fill its internal \texttt{JobStore}.}
 \item{Residing within the {\JobSupervisor} the {\JobController}s are
@@
 The Arc Client Component ({\ACC}) class is a base class needed for the
-{\Loader} in order to create loadable classes. It stored information
+{\Loader} in order to create loadable classes. It stores information
 about which flavour it supports and security related information
@@
 accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed}
- (e.g. if an {\URL} pointing to an ARC1 resource is given the
+(e.g.\ if an {\URL} pointing to an ARC1 resource is given the
 TargetRetrieverARC1 is loaded).
@@
 
-where \texttt{clusters}, \texttt{indexurls} are lists of strings where
-each string is either the name of a grid flavour followed by a colon
-and the URL of an information contact endpoint for a computing service
-or index server, respectively, of that grid flavour or an alias
-defined in the {\UserConfig} object passed as the first
-argument, e.g:
+where \texttt{clusters} and \texttt{indexurls} are lists of strings
+where each string is either the name of a grid flavour followed by a
+colon and the URL of an information contact endpoint for a computing
+service or index server, respectively, of that grid flavour or an
+alias defined in the {\UserConfig} object passed as the first
+argument, e.g.:
 
@@
 where the former is the URL to an index server while the latter is an
-URL to a computing service.
+URL to a computing service. If a string in \texttt{clusters} or
+\texttt{indexurls} is prefixed by a minus sign, the corresponding
+computing service or index server is rejected and excluded during
+resource discovery.
 
@@
 The {\TargetRetriever} is base class for {\TargetRetriever} grid
-middleware specializations and inherits from the \texttt{ACC} base
-class in order to be loadable by the ARC \texttt{Loader}. It is
-designed to work in conjunction with the {\TargetGenerator} and
-contains the pure virtual method
+middleware specializations and inherits from the {\ACC} base class in
+order to be loadable by the ARC \texttt{Loader}. It is designed to
+work in conjunction with the {\TargetGenerator} and contains the pure
+virtual method
 
@@
 
-If an index server query yields an url to a different index server
+If an index server query yields a URL to a different index server
 than the one queried, then the {\TargetRetriever} should call itself
@@
 selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of
-returning a {\Submitter} capable of submitting a job to the service
-which is represents.
+returning a {\Submitter} capable of submitting a job to the service it
+represents.
 
@@
 
-in any supported formats (currently xRSL~\cite{xrsl}, JDL~\cite{jdl},
+in any supported formats (currently XRSL~\cite{xrsl}, JDL~\cite{jdl},
 JSDL~\cite{jsdl}), converts and stores it in a JSDL-like internal job
@@
 
-The conversion mainly means a syntactial transformation and the
+The conversion mainly means a syntactical transformation and the
 {\JobDescription} class can of course do nothing in those cases where
@@
 
-No XRSL output generation is implemented yet.
+Support for XRSL output generation is currently being developed.
 
@@
 The {\JobController} is a base class for grid specific
-specializations, but also the implementor of all public functionality
+specializations, but also the implementer of all public functionality
 offered by the {\JobController}s. In other words all virtual functions
@@
 server for the latest status. Due to different protocols used for
-different grid flavours (e.g. ldap for production ARC), the
+different grid flavours (e.g.\ ldap for production ARC), the
 \texttt{GetJobInformation} has to be grid flavour specific and is only
@@
 
-\subsection{{\Job}} The
-
-{\Job} is a generic job class for storing all job related information.
-Attributes are derived from the Glue2 information model and thus a
-mapping is needed for non Glue2 compliant grid middlewares. Appendix
-\ref{app:job} shows the present mapping schema.
+\subsection{{\Job}}
+
+The {\Job} is a generic job class for storing all job related
+information. Attributes are derived from the Glue2 information model
+and thus a mapping is needed for non Glue2 compliant grid middlewares.
+Appendix \ref{app:job} shows the present mapping schema.
 
@@
   \texttt{CertificatePath} and \texttt{KeyPath} tags in \texttt{cfg}.
-  If \texttt{\$X509\_USER\_KEY} is not set or either file can't be
-  read exit with an error.}
+  If \texttt{\$X509\_USER\_KEY} is not set or either file does not
+  exist or can't be read exit with an error.}
 \item{Otherwise, if the merged user configuration tree (see
-  \texttt{ConfTree} containns a \texttt{ProxyPath} tag, copy it to
+  \texttt{ConfTree} contains a \texttt{ProxyPath} tag, copy it to
   \texttt{cfg}. If the file does not exist or can't be read exit with
   an error.}
-\item{Otherwise, if the merged user configuration tree containns a
+\item{Otherwise, if the merged user configuration tree contains a
   \texttt{CertificatePath} tag, copy it to \texttt{cfg} along with the
   accompanying \texttt{KeyPath} tag. If the \texttt{KeyPath} tag is
-  undefiend or either filefile does not exist or can't be read exit
-  with an error.}
+  undefined or either file does not exist or can't be read exit with
+  an error.}
 \item{Otherwise, if the file \texttt{/tmp/x509up\_u + userid} exists
@@
 
-Given all components listed above it possible to write versatile
+Given all components listed above it is possible to write versatile
 command line interfaces (cli) for grid job submission and

### 2008-09-25 doc/client/client_technical.tex
Fix description to match the figure

@@
   {\URL}s pointing to computing services and index servers.}
-\item{An external call is received calling for targets to be
-  prepared.}
 \item{For each selected index server and computing service a
@@
   initialized with its {\URL} and the information about whether it
-  represents a computing service or an index server. The call for
-  targets is processed by each {\TargetRetriever} in parallel.}
+  represents a computing service or an index server.}
+\item{An external call is received calling for targets to be
+  prepared. The call for targets is processed by each
+  {\TargetRetriever} in parallel.}
 \item{A {\TargetRetriever} representing an index server first tries to

### 2008-09-26 doc/hed/ARCHED_article.tex
Implementing Peter's comments for HED tech. doc. and few more enhancements.

@@
 \hypersetup{
-  pdfauthor = {J. J\"{o}nemo, et al},
+  pdfauthor = {J. J\"{o}nemo, Aleksandr Konstantinov},
   pdftitle = {The Hosting Environment of the Advanced Resource Connector middleware},
@@
 %%%%---- A list of authors ----
-    {\centering \large J. J\"{o}nemo\footnote{authors@address}, et al \large \par}
+    {\centering \large J. J\"{o}nemo\footnote{johan.jonemo@hep.lu.se}, Aleksandr Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no} \large \par}
 
@@
 
+\sloppy
+
 \chapter{Introduction}
 
-The Hosting Environment Daemon (HED) is the container of all the functional components of the new generation of the Advanced Resource Connector (ARC) middleware on the server side.
+The Hosting Environment Daemon (HED) is the container of all the functional components of the next generation of the Advanced Resource Connector (ARC) middleware on the server side.
 It is the central part in a new very lightweight incarnation of ARC that is aimed at - but not limited to - providing Web Service.
@@
 The whole design of the HED is built around the idea of flexibility and modularity.
-Inside HED the developer or deployer is supposed to use only as much as needed.
-This is why the HED mostly consists of pluggable modules with some glue between them.
+Inside HED the developer or deployer is supposed to use the minimum amount of components and external dependencies only.
+This is why the HED mostly consists of pluggable modules with some glue among them.
 
-Because in it's current state it mostly provides modules for building SOAP based Web Services, it is easy to think that HED is just another Web Services development framework like Axis, gSOAP, XFire or any other of numerous implementations.
-But instead the idea of HED is to provide framework for gluing functionalities and not a re-implementation of various standards.
-Effectively that means if Apache 2 web server is considered by developers as necessary for serving as frontend to services there could be plugin written which puts Apache 2 into a chain of other plugins of the HED.
+Because in its current state it mostly provides modules for building SOAP based Web Services, it is easy to think that HED is just another Web Services development framework like Axis, gSOAP, XFire or any other out of the numerous implementations.
+Instead, the idea of HED is to provide framework for gluing functionalities and not a re-implementation of various standards.
+Effectively that means if Apache 2 web server is considered by developers as necessary for serving as frontend to services there could be plugin written which places Apache 2 into a chain of other plugins of the HED.
 
 In the current implementation there are no Apache or Axis plugins.
-That is because the developers of HED were very much concerned about making the solution lightweight and needed an implementation of the supported protocols that was both simple and lightweight.
-As a result essentials like SOAP and HTTP are implemented inside HED, while external software is used whenever that is found to be appropriate - as in the case of TLS, (Grid)FTP, LDAP and some other cases.
-That does not exclude possibility to have plugins using entirely external solutions either developed or accepted from third parties.?
+That is because the developers of HED were very much concerned about making the solution lightweight and needed an implementation of those supported protocols that was both simple and lightweight.
+As a result essentials like SOAP and HTTP are implemented inside HED, while external software is used whenever was appropriate - as in the case of TLS, (Grid)FTP, LDAP and some other cases.
+That does not exclude possibility to have plugins using entirely external solutions either developed or accepted from third parties.
 
@@
 The situation will hopefully improve with time.
-We would be grateful for any suggestions how to improve architecture and code of the HED.
-Statements like "you are doing crap, I'll better use Axis" are understood but not welcome and are usually reacted to adequately.
+We would be grateful for any suggestions how to improve the architecture and the code of the HED.
+Statements like "you are doing crap, I shall use Axis instead" are understood but not welcome and are usually reacted adequately.
 
@@
 This document does not (yet?) include in depth description of C++ classes which constitute the HED.
-Instead most section contain notes entitled "Relevant classes".
+Instead most sections contain notes entitled "Relevant classes".
 Technical description of those classes can be found in automatically generated "Hosting Environment (Daemon) Reference Manual" document \cite{hed-api}.
 
-For examples please see source code in repository \cite{svn-arc1}. Many components and libraries are accompanied with test and example applications.
+For examples please see the source code in the repository \cite{svn-arc1}. Many components and libraries are accompanied with test and example applications.
 
@@
 
-The implementation language needed to be object oriented, efficient and provide easy access to system functionality.
+The implementation language had to be object oriented, efficient and capable of providing easy access to system functionality.
 This eventually lead to the adoption of C++.
-but languages such as Java and Python were also considered at an early stage.
+But languages such as Java and Python were also considered in the early stage.
 
 External dependencies needed to be kept to a minimum while also taking into consideration their ubiquity or relative rarity as well as license related concerns.
-Software of this level of complexity must of course depend on many external libraries and components but each such dependency has been introduced only after due consideration.
+Software of this level of complexity must depend on many external libraries and components but each such dependency has been introduced only after thorough consideration.
 
-Conservation of resources was an important goal.
-The present design enables many services sharing both the same process and the same network ports or even port while at the same time exhibiting a remarkably low memory footprint.
+Using resources in efficient way was an important goal.
+The present design enables many services sharing both the same process and the same network port(s) while exhibiting a remarkably low memory footprint.
 
@@
 %How was this implemented, dynamic loading, mcc...
-In the technical design it turned out that the endeavors to provide dynamic loading, portability and a well tested high level memory management could all be greatly assisted by introducing glibmm - the C++ interface to the gnome projects library for memory management and related functionality.
-This enables the developers to write code in a way easily portable across various operating systems and architectures.
+In the technical design it turned out that providing dynamic loading, portability and a well tested high level memory management could all be greatly assisted by introducing glibmm \cite{gtkmm} - the C++ interface made for the GNOME \cite{gnome} projects library for memory management and related functionality.
+This enables the developers to write code easily portable across various operating systems and architectures.
 
@@
 
-The HED itself means three things:
+The term HED means three things:
 \begin{enumerate}
 
-\item the daemon (called \texttt{arched}) which hooks up the system and initialize components the way as it is described by the configuration files.
+\item the daemon (called \texttt{arched}) which hooks up the system and initialize components the way as it is described in the configuration files.
 This configuration describes the components and their relations to each other.
@@
 
-Unless otherwise stated the term HED will be used through this document to refer to second option - framework of C++ classes.
+Unless otherwise stated the term HED will be used through this document to refer to the second option - framework of C++ classes.
 
@@
 The MCCs work on units called Messages which represent data going in to or out of the HED.
-The message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general Attributes and Security Attributes where information relevant to each protocol is accumulated as the message progresses.
-Each MCC typically implements one level in the Internet Protocol suite by transforming a message to an input suitable to propagate to the next component and then performs the corresponding transformation of the response on the way back.
+The Message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general Attributes and Security Attributes where information relevant to each protocol is accumulated as the Message progresses.
+Each MCC typically implements one level in the Internet Protocol suite by transforming the Message to an input suitable to propagate to the next component and then performs the corresponding transformation of the response on the way back.
 The components are all dynamically loaded to provide maximum flexibility and extensibility.
@@
 It then processes Message by modifying it or creating new Message.
-Then MCC calls entry method of next MCC in the chain.
-For information how messages are handled and about memory management policies please see API description of MCC class in \cite{hed-api}.
+Then MCC calls entry method of the next MCC in the chain.
+For information how Messages are handled and about memory management policies please see API description of MCC class in \cite{hed-api}.
 
@@
 The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs.
-Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTP service component.
-For that purpose <next> elements in MCC configuration may have optional \texttt{id} attribute which allows to assign labels to all chain links to next MCCs in the chain.
-Supported labels are MCC dependant.
+Typical scenario is that the HTTP MCC at the server side routes the HTTP Messages with POST HTTP operation to a SOAP MCC but the Messages with GET operation to for example a simple HTTP service component.
+For that purpose <next> elements in MCC configuration may have optional \texttt{id} attribute which allows to assign labels to all chain links to the next MCCs in the chain.
+Supported labels are MCC dependent.
 By default simple MCCs support only one unnamed link.
@@
 
-In general every MCC have optimal and natural places in certain chains and this place cannot always be modified.
+In general every MCC has optimal and natural places in certain chains and this place cannot always be modified.
 For example on the server side the TCP MCC must be the first MCC in any chain where it is used and the TLS MCC should be right after the TCP MCC.
@@
 
-Server and Client side MCCs are separate elements although definitely sharing some code and normally provided inside same plugin module.
+Server and Client side MCCs are separate elements although definitely sharing some code and normally provided inside the same plugin module.
 
@@
 
-\item PayloadRawInterface and it's implementation PayloadRaw.
+\item PayloadRawInterface and its implementation PayloadRaw.
 This interface represents set of catenated in-memory chunks.
@@
 
-\item PayloadStreamInterface and it's implementation PayloadStream.
+\item PayloadStreamInterface and its implementation PayloadStream.
 It covers case of sequentially accessible information.
@@
 The Message object may contain general purpose key and value pairs called Attributes.
-Keys and values are simple stings.
+Keys and values are simple strings.
 Each key may have multiple corresponding values.
 All pairs are handled by MessageAttributes class.
-Codewise there are no limitations put on content and purpose of Attributes.
+Codewise there are no limitations posed on content and purpose of Attributes.
 
@@
 Security Attributes are storing various aspects of Message useful for authorization decisions to be made.
-Those normally include operation being requested, target of operation and identity of subject making request.
+Those normally include the operation being requested, target of operation and identity of subject making request.
 They can also contain authorization policies.
@@
 
-The value if Security Attribute is an object of class inherited from SecAttr.
+The value of Security Attribute is an object of class inherited from SecAttr.
 The HED implements class SecAttr which serves as definition of interface for all Security Attributes.
-It defines way collected information may be turned into useful format.
-For that each Security Attribute value implements method Export for converting internal information into one of supported formats.
+It defines the way how collected information may be turned into format suitable for making authorization decisions.
+For that each Security Attribute value implements method Export for converting internal information into one of the supported formats.
 Currently only implemented is ARC Authorization Request/Policy (see below).
@@
 The Security Handlers are arranged into named queues.
-Elements in every queue are executed sequentially with Message as only argument.
+Elements in every queue are executed sequentially with the Message as the only argument.
 Different queues are executed at  different times.
-Which queue is executed at which case depends on MCC.
+Which queue is executed in which case it depends on the MCC.
 Most MCCs implement two queues named "incoming" and "outgoing".
-The queue "incoming" is executed for Messages moving through the chain towards hosted application and passes as argument the Message with Payload of type corresponding to MCC type i.e.
+The queue "incoming" is executed for Messages moving through the chain towards hosted application and passes the Message as argument with Payload of type corresponding to MCC type i.e.
 HTTP MCC passes Payload with parsed HTTP message.
-The "outgoing" queue is executed for Messages travelling to outside HED.
+The "outgoing" queue is executed for Messages travelling to outside the HED.
 
@@
 Those entries are used by clients to insert request Message and get result Message on output.
-For simplifying task of writing clients there is library \texttt{arcclient} provided which wraps task of creating Chains and Messages for widely used SOAP, HTTP, TLS, TCP communications.
+For simplifying the task of writing clients there is a library \texttt{arcclient} provided which wraps the task of creating Chains and Messages for the widely used SOAP, HTTP, TLS, TCP communications.
 
@@
 
-In general case multiple Services living in the HED so the incoming Message should route to the proper Service.
+In general case multiple Services are living in the HED so the incoming Message should be routed to the proper Service.
 The Plexer MCC does this job.
-It takes the the ENDPOINT attributes of the message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forward the message to the all matching service.
+It takes the ENDPOINT attribute of the Message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forwards the Message to the matching service.
 It acts as a dispatcher.
@@
 
-This Plexer provides only basic functionality and is capable of doing only simple routing.
+The Plexer provides only basic functionality and is capable of doing only simple routing.
 But because each MCC has multiple routing capabilities it is possible to provide pluggable MCC implementing more sophisticated and/or more specific routing algorithms.
@@
 This way for error reporting is mostly meant to be used for reporting problems related to code execution.
-For errors caused by processing corresponding protocol MCC should generate proper response Message which carries error description.
+For errors caused by processing the corresponding protocol MCC should generate proper response Message which carries error description.
 Only if protocol does not provide error handling MCC\_Status should be used.
-It is also advisable to convert MCC\_Status error obtained from next MCC in the Chain into protocol specific error Message if possible.
+It is also advised to convert MCC\_Status error obtained from the next MCC in the Chain into protocol specific error Message if possible.
 
@@
 Each Loader object may create multiple non-intersecting chains and there may be multiple Loader objects in same executable.
-For each component mentioned in configuration Loader creates single object of corresponding class.
+For each component mentioned in the configuration the Loader creates single object of the corresponding class.
 On Loader destruction all handled components are destroyed too.
 
-The configuration of Loader is made of following elements (see also configuration schema at http://svn.nordugrid.org/trac/nordugrid/export/10146/arc1/trunk/src/hed/libs/loader/mcc.xsd):
+The configuration of Loader is made up of the following elements (see also configuration schema at http://svn.nordugrid.org/trac/nordugrid/export/10146/arc1/trunk/src/hed/libs/loader/mcc.xsd):
 
@@
 \item \texttt{Component} represents the MCC.
-It's attribute \texttt{name} defines name of MCC which has to be instantiated.
-This MCC is being looked among those contained in modules loaded by \texttt{ModuleManager} element.
-If not found then the Loader tries to find module with name corresponding to name of the MCC.
+Its attribute \texttt{name} defines the name of the MCC which has to be instantiated.
+This MCC is being looked up among those contained in modules loaded by \texttt{ModuleManager} element.
+If not found then the Loader tries to find a module with name corresponding to name of the MCC.
 Each MCC also must have unique identifier specified by \texttt{id} attribute.
-Subelement \texttt{next} specifies link to next MCC in the Chain.
+Subelement \texttt{next} specifies link to the next MCC in the Chain.
 It uses unique identifier specified by \texttt{id}.
 The \texttt{next} element may have an optional text which is used to distinguish different Message propagation path in MCC specific way.
-The \texttt{SecurityHandler} subelement defines the Security Handler and queue to which it has to be attached.
-Also attribute \texttt{entry} my be used to create an entry point to the Chain.
+The \texttt{SecurityHandler} subelement defines the Security Handler and the queue to which it has to be attached.
+Also attribute \texttt{entry} may be used to create an entry point to the Chain.
 
-\item \texttt{Plexer} element instantiates Plexer component. It's configuration is similar to one of \texttt{Component} element.
+\item \texttt{Plexer} element instantiates Plexer component. Its configuration is similar to one of the \texttt{Component} elements.
 
-\item \texttt{Service} element defines plugin implementing final Component in the Chain - service. Form Loader point of view it's configuration is identical to that of the \texttt{Component}.
+\item \texttt{Service} element defines plugin implementing final Component in the Chain - service. From Loader point of view its configuration is identical to that of the \texttt{Component}.
 
@@
 \item The Message lifetime - lasts as long as incoming and outgoing Messages are passing through the Chain forth and back.
-The Message it self in this case is used as container for associated components.
+The Message itself in this case is used as container for associated components.
 
 \item The Session lifetime - is defined by existence of some logical connection between Messages being processed.
-Corresponding container MessageContext is normally created by first MCC in a chain and attached to the Message.
+Corresponding container MessageContext is normally created by the first MCC in a chain and attached to the Message.
 Actual lifetime of that container is MCC specific.
 For TCP MCC it corresponds to TCP connection.
-The MessageContext holds objects inherited from MessageContextElement class.
+The MessageContext holds objects inherited from the MessageContextElement class.
 The Security Attributes can also be stored in dedicated container - MessageAuthContext - with Session lifetime.
-At end of  the Session all associated objects are destroyed.
+At the end of  the Session all associated objects are destroyed.
 
-\item The Chain lifetime - is time period while components making the Chain exist.
+\item The Chain lifetime - is the time period while component that made the Chain exists.
 This lifetime is represented by ChainContext class.
-Differently from other context objects this one doe not allow free manipulation of contained objects.
-Instead it provides an interface to some internal structures of Loader object.
-Those include factories and lists of object of all types created by particular instance of the Loader.
+Differently from other context objects this one does not allow free manipulation of contained objects.
+Instead, it provides an interface to some internal structures of the Loader object.
+Those include factories and lists of objects of all types created by particular instance of the Loader.
 
@@
 Those are Data Management Components (DMC).
-Each DMC is inherited from Datapoint class and provides subset of methods for performing following operations on data endpoint:
+Each DMC is inherited from Datapoint class and provides subset of methods for performing the following operations on data endpoint:
 
@@
 
-Along with ordinary endpoints defining location of data directly - like HTTP, FTP, LDAP - the DMC can be used with indirect/indexing endpoints.
-Those are endpoints which define only location of meta-data associated with actual data or an interface/service providing functionality of managing/requesting data.
+Along with ordinary endpoints defining the location of data directly - like HTTP, FTP, LDAP - the DMC can be used with indirect/indexing endpoints.
+Those are endpoints which define only the location of meta-data associated with actual data or an interface/service providing functionality of managing/requesting data.
 For more information about indexing endpoints see description of supported URLs in \cite{urls}.
@@
 But there is also a dedicated executable provided - \texttt{arched}.
-It accepts configuration file containing XML document and passes it to Loader component.
+It accepts configuration file containing XML document and passes it to the Loader component.
 Then the Loader component takes care of loading all modules and instantiating all Message Chains.
-The \texttt{arched} also initializes Logger component, configures it and directs it's output to specified or to screen.
+The \texttt{arched} also initializes the Logger component, configures it and directs its output to where it is specified or to the standard output.
 The \texttt{arched} normally runs as background process but can be run in foreground as well.
-For more information about \texttt{arched} capabilities please read it's manual page.
+For more information about \texttt{arched} capabilities please read its manual page.
 
@@
 In order to facilitate the development of services, API bindings for languages other than C++ are provided and some service development has already been done in Python language.
-Currently only available language bindings are Python and Java.
-And currently it is possible to write only SOAP Service modules in Python and Java due to multiple inheritance limitation.
+Currently the only available language bindings are Python and Java.
+Currently it is possible to write only SOAP Service modules in Python and Java due to multiple inheritance limitation.
 
@@
 
-This chapter describes components which are implemented alongside with the HED infrastructure itself.
-Although strictly not belonging to infrastructure this minimal set of components is necessary to make infrastructure useful.
+This chapter describes components which are implemented alongside with the HED infrastructure.
+Although strictly not belonging to the infrastructure this minimal set of components is necessary to make the infrastructure usable.
 
@@
 
-The server side TCP MCC in the HED is special in that it produces messages by listening on a network socket rather than passing on messages from other MCCs.
+The server side TCP MCC in the HED is special in that it produces Messages by listening on a network socket rather than processing Messages from other MCCs.
 As such it spawns new thread for every new connection to handle Messages and their responses throughout the Message Chain.
-One could envision other MCCs having these properties but producing messages from other sources such as e.g.
+One could envision other MCCs having these properties but producing Messages from other sources such as e.g.
 UNIX sockets.
@@
 
-The server side TCP MCC generates \texttt{PayloadStreamInterface} payload in the Message passed to next MCC which can be used for communicating through open TCP channel.
+The server side TCP MCC generates \texttt{PayloadStreamInterface} payload in the Message passed to the next MCC which can be used for communicating through open TCP channel.
 Currently it ignores any payload attached to the returned Message.
-That MCC also fills following Message Attributes in incoming message while passing to next MCC:
+That MCC also fills following Message Attributes in the incoming Message while passing to the next MCC:
 \begin{enumerate}
@@
 The server side MCC accepts payload of type PayloadStreamInterface.
-It then creates own instance of object inherited from PayloadStreamInterface bound to initial payload and passes it to next MCC.
+It then creates own instance of object inherited from PayloadStreamInterface bound to initial payload and passes it to the next MCC.
 This object is maintained inside Message Context under name \texttt{tls.service} and is destroyed when Context becomes inactive.
-Currently this MCC does not expect any payload to be returned from rest of the chain and passes no payload to previous MCC.
+Currently this MCC does not expect any payload to be returned from the rest of the chain and passes no payload to previous MCC.
 
-The server MCC fills Message Attributes TLS:PEERDN and TLS:IDENTITYDN representing subjects of last certificate in client's certificate chain used for establishing secure connection and subject of last certificate which is not a proxy certificate correspondingly.
+The server MCC fills Message Attributes TLS:PEERDN and TLS:IDENTITYDN representing subjects of last certificate in client's certificate chain used for establishing secure connection and the subject of last certificate which is not a proxy certificate correspondingly.
 Use of those attributes is deprecated.
@@
 The client side MCC behaves in similar way.
-It also establishes PayloadStreamInterface type object linked to same type of payload of next MCC.
-To obtain that last payload it makes a first call to next MCC with payload of type PayloadRawInterface and then uses returned payload - which is expected to be of PayloadStreamInterface type - to create own payload object with streaming capabilities and returns it back to previous MCC for further usage.
+It also establishes PayloadStreamInterface type object linked to same type of payload of the next MCC.
+To obtain that last payload it makes a first call to the next MCC with payload of type PayloadRawInterface and then uses the returned payload - which is expected to be of PayloadStreamInterface type - to create its own payload object with streaming capabilities and returns it to previous MCC for further usage.
 
-Both client ans server side MCCs are implemented using OpenSSL toolkit \cite{openssl}, use X.509 infrastructure \cite{x509} for establishing secure connection and may be configured to get private key, certificate or proxy credentials from files residing at local file system.
-It also possible to specify location of Certification Authority certificate or to use all certificates located in specified directory.
+Both client and server side MCCs are implemented using OpenSSL toolkit \cite{openssl} and use X.509 infrastructure \cite{x509} for establishing secure connection and may be configured to get private key, certificate or proxy credentials from files residing at local file systems.
+It is also possible to specify the location of Certification Authority certificate or to use all certificates located in specified directory.
 for more information see configuration schema with comments below.
@@
 
-The server side HTTP MCC accepts messages with \texttt{PayloadStreamInterface} payload and parses HTTP related information from it.
+The server side HTTP MCC accepts Messages with \texttt{PayloadStreamInterface} payload and parses HTTP related information from it.
 Information from the HTTP header is added to the Message Attributes.
-The body of HTTP message is passed to next MCC as \texttt{PayloadRawInterface} payload.
+The body of HTTP message is passed to the next MCC as \texttt{PayloadRawInterface} payload.
 In response this MCC expects also the Message with \texttt{PayloadRawInterface}.
 It is then prepended with HTTP response header and pushes it into initially provided stream channel.
-for output it returns empty \texttt{PayloadRawInterface} payload.
+As an output it returns empty \texttt{PayloadRawInterface} payload.
 
 That MCC routes results to multiple next MCCs in the chain.
-For that it accepts only labeled <next> elements in configuration.
+For that it accepts only labeled <next> elements in the configuration.
 Label names are those of HTTP methods (uppercase).
-HTTP messages will be routed to their destinations according HTTP method requested.
+HTTP Messages will be routed to their destinations according HTTP method requested.
 
-That MCC also fills following Message Attributes in incoming message while passing to next MCC:
+That MCC also fills following Message Attributes in the incoming Message while passing to the next MCC:
 \begin{enumerate}
@@
 
-For outgoing message server MCC converts all HTTP:* Message Attributes into corresponding HTTP response header attribute.
+For outgoing Message server MCC converts all HTTP:* Message Attributes into corresponding HTTP response header attribute.
 
-The client side HTTP MCC will accept \texttt{PayloadRawInterface} payload as HTTP body and after prepending it with HTTP information passes to next MCC also as \texttt{PayloadRawInterface}.
+The client side HTTP MCC will accept \texttt{PayloadRawInterface} payload as HTTP body and after prepending it with HTTP information passes to the next MCC also as \texttt{PayloadRawInterface}.
 It accepts \texttt{PayloadStreamInterface} in response and after processing passes \texttt{PayloadRawInterface} back through the chain.
 
-The client MCC also convert all HTTP:* Message Attributes into corresponding HTTP header options while creating HTTP request.
-If define it uses HTTP:METHOD and HTTP:ENDPOINT as method and URL/path of HTTP request.
-If not specified those defined in configuration are used.
+The client MCC also converts all HTTP:* Message Attributes into corresponding HTTP header options while creating HTTP request.
+If defined it uses HTTP:METHOD and HTTP:ENDPOINT as method and URL/path of HTTP request.
+If not specified parameters defined in the configuration are used.
 For HTTP response it sets HTTP:CODE and HTTP:REASON to response code and reason correspondingly.
@@
 
-The server side MCC sets Message Attributes SOAP:ENDPOINT and ENDPOINT to URL present in 'texttt{To} element of Web Service Addressing information stored in SOAP header.
-For proper SOAP over HTTP binding it also sets \texttt{HTTP:Content-Type} Message Attribute to value \texttt{application/soap+xml} or \texttt{text/xml} for SOAP version 1.2 and 1.1 respectively for response message.
+The server side MCC sets Message Attributes SOAP:ENDPOINT and ENDPOINT to URL present in \texttt{To} element of Web Service Addressing information stored in SOAP header.
+For proper SOAP over HTTP binding it also sets \texttt{HTTP:Content-Type} Message Attribute to value \texttt{application/soap+xml} or \texttt{text/xml} for SOAP version 1.2 and 1.1 respectively for response Message.
 Also in case of SOAP fault response \texttt{HTTP:CODE} is set to \texttt{500} and \texttt{HTTP:REASON} to \texttt{SOAP FAULT}.
 
-The client side MCC for outgoing message accepts \texttt{SOAP:ACTION} Message Attribute and uses it to fill \texttt{HTTP:Content-Type} or \texttt{HTTP:SOAPAction} depending on SOAP version.
+The client side MCC for outgoing Message accepts \texttt{SOAP:ACTION} Message Attribute and uses it to fill \texttt{HTTP:Content-Type} or \texttt{HTTP:SOAPAction} depending on SOAP version.
 
@@
 
-This DMC implements access to local or remote mounted file system. It allows reading and writing content of file, listing content of directories and checking presence of the object.
+This DMC implements access to local or remote mounted file systems. It allows reading and writing the content of file, listing the content of directories and checking the presence of the object.
 
@@
 
-This DMC implements access to a data server using FTP or GridFTP protocol. It allows reading and writing content of stored file, listing content of directories and checking presence of the object.
+This DMC implements access to a data server using FTP or GridFTP protocol.
+It allows reading and writing the content of stored file, listing the content of directories and checking the presence of the object.
 
@@
 Listing is also supported.
-For that DMC extracts references from \texttt{A} tags of obtained HTML content.
-If retrieved content is not HTML or if content of it's \texttt{TITLE} tag does not start from ""Index of /"" then only information about resquested URL itself is listed.
+For that DMC extracts references from the \texttt{A} tags of the obtained HTML content.
+If retrieved content is not HTML or if content of its \texttt{TITLE} tag does not start from ""Index of /"" then only information about resquested URL itself is listed.
 
@@
 Retrieved content is converted into XML.
-The DMC supports following kind of URL
+The DMC supports following kind of URL:
 
@@
 \begin{itemize}
-\item \texttt{attributes} is coma separated list of attributes which has to be retrieved,
+\item \texttt{attributes} is comma separated list of attributes which has to be retrieved,
 \item \texttt{scope} is either \texttt{base}, \texttt{one} or \texttt{sub} as defined by LDAP,
@@
 
+\chapter{Future work}
+
+Currently the interfaces in the HED have matured enough and stay quite stable. Most changes being introduced in the field of Security infrastructure.
+
+Most of the work has shifted to the development of higher level libraries and services based on the HED. But the HED also continiously being expanded by adding new plugins.
+
+There are also significant pieces of functionality not addressed in the HED yet like user-friendly configuration, efficient inter-service internal Message routing, etc.
+
 %\chapter{Conclusion}

### 2008-09-26 doc/hed/ARCHED_article.tex
Converting list of attributes to bullets and adding full stop after each.

@@
 That MCC also fills following Message Attributes in the incoming Message while passing to the next MCC:
-\begin{enumerate}
+\begin{itemize}
 
@@
 
-\item HTTP:RANGESTART - Range request start offset
+\item HTTP:RANGESTART - Range request start offset.
 
-\item HTTP:RANGEEND - Range request end offset
+\item HTTP:RANGEEND - Range request end offset.
 
-\item HTTP:ENDPOINT - URL or path as specified in HTTP request header
+\item HTTP:ENDPOINT - URL or path as specified in HTTP request header.
 
-\item HTTP:* - Here * stands for any name. All HTTP options from HTTP request header are converted into Message Attributes named HTTP:\{option name\}
+\item HTTP:* - Here * stands for any name. All HTTP options from HTTP request header are converted into Message Attributes named HTTP:\{option name\}.
 
-\item ENDPOINT - same as HTTP:ENDPOINT
+\item ENDPOINT - same as HTTP:ENDPOINT.
 
-\end{enumerate}
+\end{itemize}
 

### 2008-09-27 doc/client/arc-ws-clients.tex
Port number typo

@@
 that have been loaded into a communication chain. In this case, a TCP
-MCC that connects to port 6000 on the local host is connected to a TLS
+MCC that connects to port 60000 on the local host is connected to a TLS
 MCC, which is connected to a HTTP that uses the POST method, and

### 2008-10-01 doc/hed/ARCHED_article.tex
Removed "crap" and fixed bibliography

@@
 \hypersetup{
-  pdfauthor = {J. J\"{o}nemo, Aleksandr Konstantinov},
+  pdfauthor = {J. J\"{o}nemo, A. Konstantinov},
   pdftitle = {The Hosting Environment of the Advanced Resource Connector middleware},
@@
 %%%%---- A list of authors ----
-    {\centering \large J. J\"{o}nemo\footnote{johan.jonemo@hep.lu.se}, Aleksandr Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no} \large \par}
+    {\centering \large J. J\"{o}nemo\footnote{johan.jonemo@hep.lu.se}, A. Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no} \large \par}
 
@@
 We would be grateful for any suggestions how to improve the architecture and the code of the HED.
-Statements like "you are doing crap, I shall use Axis instead" are understood but not welcome and are usually reacted adequately.
+%Statements like "you are doing crap, I shall use Axis instead" are understood but not welcome and are usually reacted adequately.
 

### 2008-10-02 doc/hed/ARCHED_article.tex
Added authors, fixed "quite stable" as Bjarte suggested

@@
 \hypersetup{
-  pdfauthor = {J. J\"{o}nemo, A. Konstantinov},
+  pdfauthor = {M. Ellert, J. J\"{o}nemo, A. Konstantinov, M. Norden, W. Qiang, F. Szalai},
   pdftitle = {The Hosting Environment of the Advanced Resource Connector middleware},
@@
 %%%%---- A list of authors ----
-    {\centering \large J. J\"{o}nemo\footnote{johan.jonemo@hep.lu.se}, A. Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no} \large \par}
+    {\centering \large M. Ellert, J. J\"{o}nemo\footnote{johan.jonemo@hep.lu.se}, A. Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no}, M. Nord\'{e}n, W. Qiang, F. Szalai \large \par}
 
@@
 
-Currently the interfaces in the HED have matured enough and stay quite stable. Most changes being introduced in the field of Security infrastructure.
+Interfaces in HED are mature and stable with the exception of the security infrastructure which still needs improvements.
 

### 2008-10-03 doc/hed/ARCHED_article.tex
Modified changes requested by Bjarte implemented.

@@
 \hypersetup{
-  pdfauthor = {M. Ellert, J. J\"{o}nemo, A. Konstantinov, M. Norden, W. Qiang, F. Szalai},
+  pdfauthor = {D. Cameron, M. Ellert, J. J\"{o}nemo, A. Konstantinov, I. Marton, B. Mohn, J. K. Nilsen, M. Norden, W. Qiang, G. Roczei, O. Smirnova, F. Szalai, A. Waananen},
   pdftitle = {The Hosting Environment of the Advanced Resource Connector middleware},
@@
 %%%%---- A list of authors ----
-    {\centering \large M. Ellert, J. J\"{o}nemo\footnote{johan.jonemo@hep.lu.se}, A. Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no}, M. Nord\'{e}n, W. Qiang, F. Szalai \large \par}
+    {\centering \large D.
+Cameron, M.
+Ellert, J.
+J\"{o}nemo, A.
+Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no}, I.
+Marton, B.
+Mohn, J.
+K.
+Nilsen, M.
+N\'{o}rden, W.
+Qiang, G.
+Roczei, O.
+Smirnova, F.
+Szalai, A.
+W\"{a}\"{a}n\"{a}nen \large \par}
 
@@
 But languages such as Java and Python were also considered in the early stage.
+Our investigation showed that there are Open Source tools which allow C++ modules to be relatively easy interfaced from modules written in other considered programming languages.
+C++ modules also work in opposite way being capable to use modules written in various other languages.
 
@@
 
-Interfaces in HED are mature and stable with the exception of the security infrastructure which still needs improvements.
+Interfaces in HED are mature and stable with only the security infrastructure still being extended.
 

### 2008-10-08 doc/a-rex/arex_tech_doc.tex
update for new cache

@@
 
-The A-REX can cache input files. Caching is enabled if corresponding
-command is present in configuration file. The A-REX does not cache
-files marked as executable in job. Caching can also be explicitly
-turned off by user for each file by using \textit{cache=no} option
-in URL (for URL options read {}``Protocols, Uniform Resource Locators
-(URL) and extensions supported in ARC''~\cite{urls}). The disc space
-occupied by cache is controlled by removing unused files. For more
-information look in Section~\ref{SubSection:ConfigFile}.
-
-\begin{framed}
-NOTE: The caching algorithm used by A-REX and described above
-is currently being replaced by a better one which is no robust and
-provides better scalability.
-\end{framed}
+The A-REX can cache input files. Caching is enabled if one or more cache
+directories are specified in the configuration file. The A-REX does not
+cache files marked as executable in a job. Caching can also be explicitly
+turned off by the user for each file by using the\textit{ cache=no}
+option in the URL (for URL options read {}``Protocols, Uniform Resource
+Locators (URL) and extensions supported in ARC'' \cite{urls}). The
+disk space occupied by the cache is controlled by removing old files.
+For more information look in section \ref{SubSection:ConfigFile}.
 
@@
 
-Cache directory contains plain files. Those are
-
-\begin{itemize}
-\item \textit{list} -- stores names of the files (8 digit numbers) and corresponding
-URLs delimited by blank space. Each pair is delimited by some amount
-of \textbackslash{}0 codes. Also creation and expiration times are
-stored if available
-\item \textit{old} -- stores URLs which have been removed from cache. Records
-are delimited by some amount of \textbackslash{}0 codes and are meant
-to be removed by some external routine.
-\item \textit{new} -- stores URLs which have been added to cache. Records
-are delimited by some amount of \textbackslash{}0 codes and are removed
-when corresponding files are removed from cache. They can also be
-handled by some external routines. Every time record is added to \emph{old}
-it is removed from \emph{new}.
-\item \textit{statistics} -- consists of strings containing \emph{name=value}
-pairs. Following names are defined:
-
-\begin{itemize}
-\item \textit{hardsize} -- size of file system for storing cached data
-\item \textit{hardfree} -- amount of disc space available on that file system
-\item \textit{softsize} -- if cache exceeds this size files are started being
-removed 
-\item \textit{softfree} -- space left till softsize (can be negative)
-\item \textit{claimed} -- space used by files claimed by running jobs
-\item \textit{unclaimed} -- space used by files not being currently used
-by any job
-\end{itemize}
-\item \textit{\#\#\#\#\#\#\#\#.info} -- stores state of file (\textit{\#\#\#\#\#\#\#\#}
-stands for 8 digits). State is represented by one character:\\
-\textit{c} -- just created, content is empty.\\
-\textit{f} -- failed to download (treated same as 'c').\\
-\textit{r} -- ready to be used, content is valid.\\
-\textit{d} -- being downloaded. 'd' is followed by identifier of application/job
-downloading that file. During content's download this file has write
-lock set.
-\item \textit{\#\#\#\#\#\#\#\#.claim} -- stores list of identifiers of applications/jobs
-using this file. Identifiers are stored one per line.
-\item \textit{\#\#\#\#\#\#\#\#} -- files storing content of corresponding
-URL. These can be stored in separate directory.
-\end{itemize}
-
-Files \textit{list, old, new} and \textit{\#\#\#\#\#\#\#\#.info}
-has to be stored on file system which has support for files' locking.
+Cached files are stored in sub-direcories under the \emph{data} directory
+in each main cache directory. Filenames are constructed from an SHA-1
+hash of the URL of the file, and split into subdirectories based on
+the two initial characters of the hash. This enables the cached files
+to be evenly split over a number of subdirectories. If
+more than one cache directory is used the initial letter of the hash
+also determines which cache is to be used. The algorithm is simply
+to use the cache directory at index \emph{i} in the list of directories,
+where \emph{i} is found from the initial letter of the hash mod the
+number of caches. This limits the maximum number of cache directories
+to 16, as SHA-1 hashes use the character set {[}\emph{0-9,a-f}]. In
+the extremely unlikely event of a collision between two URLs having
+the same SHA-1 hash, caching will not be used for the second file.
+
+There is no indexing system for the cache, and any cache filename
+can be easily determined from a URL by using the same hashing algorithm
+as the A-REX, e.g. the standard command line tool \emph{sha1sum. }The
+cache directory used by a file can be determined from the initial
+letter of the hash and the order of cache directories in the configuration
+file. Some associated metadata (the corresponding URL and an expiry
+time, if available) are stored in a file with the same name as the
+cache file, with a \emph{.meta} suffix.
+
+For example, with a cache directory \emph{/cache}, the file
+
+\begin{center}
+\emph{lfc://atlaslfc.nordugrid.org//grid/atlas/file1 }\\
+is mapped to\\
+\emph{/cache/data/78/f607405ab1df6b647fac7aa97dfb6089c19fb3}
+\par\end{center}
+
+and the file \emph{/cache/data/78/f607405ab1df6b647fac7aa97dfb6089c19fb3.meta
+}contains the original URL and an expiry time if one is available.
+
+At the start of a file download, the cache file is locked, so that
+it cannot be deleted and so that another download process cannot write
+the same file simultaneously. This done by creating a file with the
+same name as the cache filename but with a \emph{.lock} suffix. This
+file contains the process ID of the process and the hostname of the
+host holding the lock. If this file is present, another process cannot
+do anything with the cache file and must wait until the cache file
+is unlocked (i.e. the \emph{.lock} file no longer exists). The lock
+has a timeout of one day, so that stale locks left behind by a download
+process exiting abnormally will eventually be cleaned up. Also, if
+the process corresponding to the process ID stored inside the lock
+is no longer running on the host specified in the lock, it is safe
+to assume that the lock file can be deleted.
 
@@
 
-If a job requests an input file which is subject for caching, it is
-stored in the cache directory instead of the SD. The file is made
-available to the job by either soft-linking it in the SD or copying
-it to the SD. The latter option is more secure and hence advised.
-
-Before downloading a file the A-REX tries to determine it's size and
-then tries to preallocate space in the cache directory by writing
-an empty file of the same size. If this fails (possibly because the
-file system has no more space), it tries to the remove oldest cached
-files which are not being used by any job. This means that \textbf{the
-real hard limit of cache size is the space available in the file-system}.
-In case cache gets full and it is impossible to free enough space,
-the download fails and then is retried without using cache.
-
-Before giving access to an already cached file, the A-REX contacts
-the original file source to check if the user has sufficient access
-rights to the file. Not all protocols make this check possible. Also,
-file creation or validity times are checked to make sure the cached
-file is fresh. If it is impossible to obtain creation and invalidation
-times, the file is invalidated 24 hours after it was downloaded.
-
-The A-REX checks the cache periodically. If the space used by cache
-exceeds the high water-mark given in the configuration file (\emph{softsize}),
-it tries to remove the oldest unused files until cache size drops
-below the low water-mark. This sets the soft limit of cache size.
-
-There are two kinds of caches supported. Files in \textit{private}
-cache are owned by the Unix user to which a Grid user is mapped. Those
-files are readable only by that particular Unix user. Another kind
-of cache is \textit{shared}. Files in the shared cache are owned by
-the Unix user who started A-REX and are readable by everyone.
+If a job requests an input file which can be cached or is allowed to
+be cached, it is stored in the selected cache directory, and depending
+on the configuration, either the file is copied to the SD or a hard
+link is created in a per-job directory and a soft link is created in
+the SD to there. The per-job directories are in the \emph{joblinks}
+subdirectory of the main cache directory. The former option is advised
+if the cache is on a file system which will suffer poor performance
+from a large number of jobs reading files on it, or the file system
+containing the cache is not accessible from worker nodes. The latter
+option is the default option. The per-job directory is only readable
+by the local user running the job, and the cache directory is readable
+only by the A-REX user (usually root). This means that the local user
+cannot access any other users' cache files. It also means that cache
+files can be removed without needing to know whether they are in use
+by a currently running job.
+
+If the file system containing the cache is full and it is impossible
+to free any space, the download fails and is retried without using
+cacheing.
+
+Before giving access to a file already in the cache, the A-REX contacts
+the initial file source to check if the user has read permission on
+the file. Also file creation or validity times from the original source
+are checked to make sure the cached file is fresh enough. If it is
+impossible to obtain creation and invalidation times for the file,
+it is invalidated 24 hours after download.
+
+The A-REX checks the cache periodically. If the used space on the file
+system containing the cache exceeds the high water-mark given in the
+configuration file it tries to remove the least-recently accessed
+files to reduce size to the low water-mark.
+
+\subsection{Administration tools}
+
+The following tools (installed in \emph{\$ARC\_LOCATION/libexec})
+exist to help with administration of the cache:
+
+\begin{itemize}
+\item \emph{cache-clean} - This tool is used periodically (every 2 minutes)
+by the A-REX to keep the size of each cache within the configurated limits.
+It removes files from the cache if the total size of the cache is
+greater than the configured limit. It will attempt to remove files
+which are not locked in order of access time, starting with the earliest,
+until the size is lower than the configured lower limit. If the lower
+limit cannot be reached (because too many files are locked, or other
+files outside the cache are taking up space on the file system), the
+tool will exit before the lower limit is reached.\\
+\emph{cache-clean -h} gives a list of options. The most useful option
+for administrators is \emph{-s}, which does not delete anything, but
+gives summary information on the files in the cache, including information
+on the ages of the files in the cache.\\
+It is not recommended to run \emph{cache-clean }manually to clean
+up the cache, unless it is desired to temporarily clean up the cache
+with different size limits to those specified in the configuration.
+\item \emph{cache-list} - This tool is used to list all files present in
+each cache. It simply reads through all the \emph{.meta }files and
+prints to stdout a list of all URLs stored in each cache and their
+corresponding cache filename, one per line.
+\end{itemize}
 
@@
 in seconds. Default is \emph{3 minutes}.
-\item \textbf{\textit{cacheregistration}}\textit{=yes|no} -- enables or disables
-registration of cache data into Indexing Services like RC or RLS.
-The default is \emph{no}. Only files downloaded through \emph{meta-url}
-are registered. Registration is done to same service used for obtaining
-information about file. For this operation credentials of the A-REX
-(host key and certificate) are used. If required new files storage
-location is registered at Indexing Service with quasi-url \emph{cache://hostname/}
-and name \emph{hostname:cache}.
 \item \textbf{\textit{authplugin}}\textit{=state options plugin} -- specifies
@@
 is called \textbf{\textit{sessiondir}}.
-\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} -- specifies
-the directory to store cached data. Empty \emph{path} disables caching.
-Default is not to cache data. Optional \textit{link\_path} specifies
-the path at which cache is accessible at computing nodes. If \textit{link\_path}
-is set to '.' files are not soft-linked, but copied to session directory.
-\item \textbf{\textit{privatecache}}\textit{=path {[}link\_path]} -- same
-as \textit{cache} command, but cache belongs (owned) to user. For
-shared caches use 'cache'.
-\item \textbf{\textit{cachedata}}\textit{=path} -- allows to specify separate
-place to store cache files containing data itself. This can be useful
-in case of big data storage available only on NFS server which does
-not support file locking. If command or \textit{path} is missing -
-default is to store data at place specified in \textit{cache} or \textit{privatecache}
-command, together with control files. 
-\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} -- specifies
-high and low water-mark for space used by cache. Values are specified
-in bytes. Both \textit{high\_mark} and \textit{low\_mark} can be negative
-values. In that case corresponding positive value means space left
-on file system. If \textit{low\_mark} is omitted it becomes equal
-to \textit{high\_mark}. By default this feature is turned off. To
-turn it off explicitly \textit{cachesize} without parameters should
-be specified. If turned off cache will grow up till it fills whole
-file system.
+\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} - specifies
+a directory to store cached data (see \ref{sec:cache}).\emph{ }Multiple
+cache directories may be specified by specifying multiple \emph{cachedir}
+commands. Cached data will be distributed evenly over the caches.
+Specifying no \emph{cachedir }command or commands with an empty path
+disables caching. Optional \textit{link\_path} specifies the path
+at which \emph{path} is accessible on computing nodes, if it is different
+from the path on the A-REX host. If \textit{link\_path} is set to '.'
+files are not soft-linked, nor are per-job links created, but files
+are copied to the session directory. In old configuration file this
+command is called \textbf{\textit{cache}}.
+\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} - specifies
+high and low watermarks for space used by each cache, as a percentage
+of the space on the file system on which the cache directory is located.
+When the max is exceeded, files will be deleted to bring the used
+space down to the min level. It is a good idea to have each cache
+on its own separate file system. To turn off this feature, \char`\"{}cachesize\char`\"{}
+without parameters can be specified. These cache settings apply to
+all caches specified by \emph{cachedir} commands.
 \item \textbf{\textit{maxrerun}}\textit{=number} -- specifies maximal number

### 2008-10-25 doc/sec/SecurityFrameworkofARC1.tex
Add x509token description into technical documentation

% This file was converted to LaTeX by Writer2LaTeX ver. 0.5
% see http://www.hj-gym.dk/~hj/writer2latex for more info
\documentclass{article}
\usepackage[ascii]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[spanish,english]{babel}
\usepackage{amsmath,amssymb,amsfonts,textcomp}
\usepackage{color}
\usepackage{array}
\usepackage{supertabular}
\usepackage{hhline}
\usepackage{hyperref}
\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=name, pdfsubject=, pdfkeywords=}
\usepackage[pdftex]{graphicx}
% Text styles
\newcommand\textstyleInternetlink[1]{\textcolor[rgb]{0.0,0.0,0.5019608}{#1}}
% Outline numbering
\setcounter{secnumdepth}{4}
\renewcommand\thesection{\arabic{section}.}
\renewcommand\thesubsection{\arabic{subsection}.}
\renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection}.}
\renewcommand\theparagraph{\arabic{subsection}.\arabic{subsubsection}.\arabic{paragraph}.}
\makeatletter
\newcommand\arraybslash{\let\\\@arraycr}
\makeatother
% List styles
\newcounter{saveenum}
\newcommand\liststyleWWviiiNumii{%
\renewcommand\labelitemi{[F0B7?]}
\renewcommand\labelitemii{[F0B7?]}
\renewcommand\labelitemiii{[F0B7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumxiii{%
\renewcommand\labelitemi{[F0B7?]}
\renewcommand\labelitemii{[F0B7?]}
\renewcommand\labelitemiii{[F0B7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumxxi{%
\renewcommand\labelitemi{[F0B7?]}
\renewcommand\labelitemii{o}
\renewcommand\labelitemiii{[F0A7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumxvii{%
\renewcommand\labelitemi{o}
\renewcommand\labelitemii{o}
\renewcommand\labelitemiii{[F0A7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumxix{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\labelitemi{o}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
}
\newcommand\liststyleWWviiiNumxviii{%
\renewcommand\labelitemi{o}
\renewcommand\labelitemii{o}
\renewcommand\labelitemiii{[F0A7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumxx{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\labelitemi{o}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
}
\newcommand\liststyleWWviiiNumxxii{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\labelitemi{o}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
}
\newcommand\liststyleWWviiiNumxxiii{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\labelitemi{{}-}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
}
\newcommand\liststyleWWviiiNumiii{%
\renewcommand\theenumi{\alph{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi)}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumiv{%
\renewcommand\theenumi{\alph{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi)}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumv{%
\renewcommand\theenumi{\alph{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi)}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumvi{%
\renewcommand\theenumi{\alph{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi)}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumvii{%
\renewcommand\theenumi{\alph{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi)}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumviii{%
\renewcommand\labelitemi{[F0B7?]}
\renewcommand\labelitemii{[F081?]}
\renewcommand\labelitemiii{${\blacksquare}$}
\renewcommand\labelitemiv{[F06C?]}
}
\newcommand\liststyleWWviiiNumix{%
\renewcommand\labelitemi{[F0B7?]}
\renewcommand\labelitemii{{}-}
\renewcommand\labelitemiii{${\blacksquare}$}
\renewcommand\labelitemiv{[F06C?]}
}
\newcommand\liststyleWWviiiNumx{%
\renewcommand\labelitemi{[F0B7?]}
\renewcommand\labelitemii{[F081?]}
\renewcommand\labelitemiii{${\blacksquare}$}
\renewcommand\labelitemiv{[F06C?]}
}
\newcommand\liststyleWWviiiNumxiv{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumxi{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumxii{%
\renewcommand\labelitemi{[F0B7?]}
\renewcommand\labelitemii{[F0B7?]}
\renewcommand\labelitemiii{[F0B7?]}
\renewcommand\labelitemiv{[F0B7?]}
}
\newcommand\liststyleWWviiiNumxv{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumxvi{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleWWviiiNumxxiv{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
% Page layout (geometry)
\setlength\paperwidth{8.2673in}
\setlength\paperheight{11.6925in}
\setlength\voffset{-1in}
\setlength\hoffset{-1in}
\setlength\topmargin{1in}
\setlength\oddsidemargin{1.25in}
\setlength\textheight{9.0315in}
\setlength\textwidth{5.7672997in}
\setlength\footskip{0.661in}
\setlength\headheight{0cm}
\setlength\headsep{0cm}
% Footnote rule
\setlength{\skip\footins}{0.0469in}
\renewcommand\footnoterule{\vspace*{-0.0071in}\setlength\leftskip{0pt}\setlength\rightskip{0pt plus 1fil}\noindent\textcolor{black}{\rule{0.25\columnwidth}{0.0071in}}\vspace*{0.0398in}}
% Pages styles
\makeatletter
\newcommand\ps@Convertiv{
  \renewcommand\@oddhead{}
  \renewcommand\@evenhead{}
  \renewcommand\@oddfoot{}
  \renewcommand\@evenfoot{\@oddfoot}
  \renewcommand\thepage{\arabic{page}}
}
\newcommand\ps@Standard{
  \renewcommand\@oddhead{}
  \renewcommand\@evenhead{}
  \renewcommand\@oddfoot{\thepage{}}
  \renewcommand\@evenfoot{\@oddfoot}
  \renewcommand\thepage{\arabic{page}}
}
\makeatother
\pagestyle{Standard}
\setlength\tabcolsep{1mm}
\renewcommand\arraystretch{1.3}
\newcounter{Bibitem}
\renewcommand\theBibitem{\arabic{Bibitem}}
\newcounter{Figure}
\renewcommand\theFigure{\arabic{Figure}}
\title{}
\begin{document}
\begin{flushleft}
\tablehead{}\begin{supertabular}{m{1.3809599in}m{2.0976598in}}
\begin{center}
\includegraphics[width=1.4228in,height=0.9098in]{SecurityFrameworkofARC1-img1.jpg}
\end{center}
 &
\selectlanguage{english}\scshape\color{black} \newline
\newline
NORDUGRID\\
\end{supertabular}
\end{flushleft}

\bigskip

\bigskip

\begin{flushleft}
\tablehead{}\begin{supertabular}{m{0.83445984in}m{4.78586in}m{0.83655983in}}
\multicolumn{3}{m{6.61436in}}{\raggedleft
\selectlanguage{english}\color{black} NORDUGRID-TECH-16\newline
6/8/08}\\
\multicolumn{3}{m{6.61436in}}{\centering
{\selectlanguage{english}\bfseries\scshape\color{black} Security
Framework of ARC1}\par

~
}\\
\multicolumn{3}{m{6.61436in}}{\centering
\selectlanguage{english}\color{black} W.Qiang, A.Konstantinov}\\
\multicolumn{3}{m{6.61436in}}{\centering
\selectlanguage{english}\itshape\color{black} Abstract}\\
~
 &
\selectlanguage{english}\color{black} This document is about security
design concerns and ideas, as well as security framework implementation
in the ARC1 middleware. &
~
\\
\end{supertabular}
\end{flushleft}
\clearpage{\centering\selectlanguage{english}\bfseries\scshape\color{black}
Table of Contents
\par}

\setcounter{tocdepth}{9}
\renewcommand\contentsname{}
\tableofcontents

\bigskip

\bigskip
\clearpage\clearpage\setcounter{page}{1}\pagestyle{Convertiv}

\bigskip

\subsection{Introduction}
{\selectlanguage{english}\color{black}
The security framework of the ARC1 includes two parts of capabilities:
security capability embedded in hosting environment, and security
capability implemented as plug-ins with well-defined interfaces which
can be accessed by hosting environment and applications. The following
design concerns were employed when designing:}

\liststyleWWviiiNumii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Interoperability and standardization}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
In consistent with the main design concern of the ARC1, interoperability
and standardization is considered in security framework. For example,
in terms of authentication, PKI infrastructure and proxy certificate
(RFC3820 [1]) is used as most of the other grid middle-wares do. Since
supporting of standardization is a way for implementing
interoperability, some standard specifications have been implemented as
prototype and tested, such as SAML specification.}

\liststyleWWviiiNumii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Modularity and extensibility }
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
Besides the security functionality which is embedded in hosting
environment, the other security functionality is implemented as
plug-ins which has well-defined interfaces, and is configurable and
dynamically loadable. Since the interoperation interface between
security plug-in and hosting environment or applications is predefined,
it is easy to extend the security functionality in order to support
some other security capability by implementing the interface.}

\liststyleWWviiiNumii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Backward compatibility }
\end{itemize}
{\selectlanguage{english}\color{black}
The GSI (Grid Security Infrastructure) based mechanism has been a
de-facto solution for grid security. The design of security framework
should be compatible to it.}

\bigskip

\subsection{Security architecture in HED. SecHandler and PDP}
\subsubsection{Structure of SecHandler and PDP}
{\selectlanguage{english}\upshape\color{black}
In the implementation of the ARC1, there is a Service Container -- the
Hosting Environment Daemon (HED) (D1.2-2, [2]) which provides a hosting
place for various services in application level, as well as a flexible
and efficient communication mechanism.}

{\selectlanguage{english}\upshape\color{black}
HED contains a framework for implementing and enforcing authentication
and authorization. Each Message Chain Component (MCC) or service has a
common interface for implementing various authentication and
authorization functionality. This functionality is implemented by using
pluggable components (plug-ins) called SecHandler. The SecHandler
components are C++ classes and provide method for processing messages
traveling through Message Chains of the HED. Each MCC or Service
\ usually implement two queues of SecHandlers -- one for incoming
messages and one for outgoing called
{\textquotedblleft}incoming{\textquotedblright} and
{\textquotedblleft}outgoing{\textquotedblright} respectively. It is
possible for MCC or Service to implement other set of queues. Please
check documentation of particular component for that particular
information. All SecHandler components attached to the queue are
executed sequentially. If any of them fails, message processing fails
as well.}

{\selectlanguage{english}\upshape\color{black}
Each SecHandler is configured inside same configuration file used for
configuring whole chain of MCCs. Some of implemented SecHandler
components also make use of pluggable and configurable sub-modules
which specifically handle various security functionalities, such as
authorization, authentication, etc. The currently implemented
sub-modules used by some SecHandlers are Policy Decision Point (PDP)
components such as Arc PDP which can process ARC specific Request and
Policy documents. Figure 1 gives the structure of a MCC/Service, and
the message sequence inside it. And Figure 2 shows the configuration of
SecHandler components for an example
{\textquotedblleft}Echo{\textquotedblright} service.}

\bigskip

\includegraphics[width=5.7709in,height=4.5209in]{SecurityFrameworkofARC1-img2.jpg}

{\selectlanguage{english}\itshape\color{black}
Figure \stepcounter{Figure}{\theFigure}. There are usually two chains of
SecHandlers inside the MCC or service. Each SecHandler will parse the
Security Attributes which are generated by the upstream MCC/services or
probably upstream SecHandlers in the same or other MCC/Service, and do
message processing or authenticate or authorize the incoming/outgoing
message based on the collected information. The SecHandler can also
change the payload and attributes of Messsage itself. For example, the
Username-Token SecHandler will insert the WSS Username Token [3] into
header part of SOAP message. The PDPs are called by the SecHandlers and
are supposed to make authorization decision. Here the two local PDP and
one remote PDP service is just for demonstration, and one or any number
of PDPs can be configured under each SecHandler.}

\bigskip

[Warning: Draw object ignored]

{\centering\selectlanguage{english}\itshape\color{black}
Figure \stepcounter{Figure}{\theFigure}. Example Echo service is
configured to use two SecHandlers, both responsible for authorization.
First SecHandler uses the identity of client extracted from the
incoming message to map it into local identity like local Linux
username. In this case all clients are mapped to local account
{\textquotedblleft}test{\textquotedblright}. The second one uses two
PDPs: one will compose ARC specific authorization request based on the
Security Attributes collected from
{\textquotedblleft}incoming{\textquotedblright} message and evaluate it
against the ARC specific authorization policy which is defined in local
file {\textquotedblleft}policy.xml{\textquotedblright}; the other will
compare the X509 identity of client extracted from the incoming message
against list of identities stored locally.
\par}

\subsubsection{Interface of SecHandler}
{\selectlanguage{english}\upshape\color{black}
When one component (MCC or service) is loaded according to the
configuration information, the SecHandler under the component and the
plug-ins like PDP which are attached to the SecHandler will be loaded
as well. }

{\selectlanguage{english}\upshape\color{black}
There is one simple interface (see Figure 3) defined in class
SecHandler, which will be called by the containing MCC/Service once
there is message (incoming or outgoing) need to be processed.}

{\centering\selectlanguage{english}\itshape\color{black}
Figure \stepcounter{Figure}{\theFigure}. class SecHandler is an abstract
class which includes a general interface called Handle which uses
Message object as argument. Any security handler implementation should
inherit class SecHandler and implement the interface according to the
actual functionality. The interface only return simple Boolean value,
and any useful information generated during the calling of this
interface should be put into the security attribute of the message, or
put into the payload itself. 
\par}

\begin{center}
\begin{minipage}{3.3756in}
{\selectlanguage{english}\color{black}
\ \ class SecHandler \{}

{\selectlanguage{english}\upshape\color{black}
\ \ \ public:}

{\selectlanguage{english}\upshape\color{black}
\ \ \ \ SecHandler(Arc::Config*) \{\};}

{\selectlanguage{english}\upshape\color{black}
\ \ \ \ virtual \~{}SecHandler() \{\};}

{\selectlanguage{english}\upshape\color{black}
\ \ \ \ virtual bool \textit{Handle}(Arc::Message *msg) = 0;}

{\selectlanguage{english}\color{black}
\ \ \};}
\end{minipage}
\end{center}
{\selectlanguage{english}\upshape\color{black}
Currently, the ARC1 comes with the following four security handler
implemented: }

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{arc.authz} -- Authorization SecHandler}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The \textit{arc.authz} is responsible for calling the interface of
policy decision point and getting back the authorization result, and
then making decision according to this authorization result. There is
one simple interface (see Figure 4) defined in PDP, which will be
called by \textit{arc.authz} if configured inside once there is message
(incoming or outgoing) need to be processed.}

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{identity.map} -- Identity Mapping SecHandler}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The \textit{identity.map} is a specific authorization oriented security
handler. It will map the global identity in the message into local
identity like system username based on the result returned by Policy
Decision Point components.}

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{delegation.collector} -- Delegation SecHandler}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The \textit{delegation.collector} is responsible for collecting the
delegation policy information from the remote proxy credential (proxy
certificate is compatible to RFC3820) inside the message, and putting
this policy into message{\textquoteright}s security attribute for the
usage of other components, such as \ \textit{delegation.pdp}.}

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{usernametoken.handler} -- UseranemToken SecHandler}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The task of the \ \textit{usernametoken.handler} is to generate the
WS-Security Username-Token and add it into header of SOAP message which
is the payload of outgoing message. It can also extract the WS-Security
Username-Token from the header of SOAP message which is the payload of
incoming message.}

\subsubsection[Interface of PDP]{Interface of PDP}
\label{bkm:Ref204009855}{\selectlanguage{english}\upshape\color{black}
Figure 4 shows the definition of abstract class PDP. The implementation
could be some function which implements the interface by composing the
policy evaluation request, evaluating this request against some policy,
and returning the evaluation result, or just by composing the policy
evaluation request, invoking some remote policy decision web service
and getting back the evaluation result.}

{\centering\selectlanguage{english}\itshape\color{black}
Figure \stepcounter{Figure}{\theFigure}. class PDP is an abstract class
which includes a general interface called isPermitted which uses
Message object as argument. Any policy decision point implementation
should inherit class PDP and implement the interface according to the
actual functionality. The interface only return simple Boolean value,
and any useful information generated during the calling of this
interface should be put into the security attribute of the message, or
put into the payload itself.
\par}

\begin{center}
\begin{minipage}{3.8756in}
{\selectlanguage{english}\color{black}
\ \ class PDP \{}

{\selectlanguage{english}\upshape\color{black}
\ \ \ public:}

{\selectlanguage{english}\upshape\color{black}
\ \ \ \ PDP(Arc::Config* cfg) \{ \};}

{\selectlanguage{english}\upshape\color{black}
\ \ \ \ virtual \~{}PDP() \{\};}

{\selectlanguage{english}\upshape\color{black}
\ \ \ \ virtual bool \textit{isPermitted}(Arc::Message *msg) = 0;}

{\selectlanguage{english}\color{black}
\ \ \};}
\end{minipage}
\end{center}
{\selectlanguage{english}\upshape\color{black}
Currently, the ARC1 comes with the following four policy decision point
implementation: }

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{arc.pdp }{}-- Arc PDP}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The Arc PDP will organize the security attributes into the ARC specific
authorization request, call the policy evaluator to evaluate the
request against the policy (which is in ARC specific format)
repository, and get back the evaluation result. See paragraph 3 for
detail information about request schema and policy schema.}

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{delegtion.pdp} -- Delegation PDP}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The Delegation PDP is basically similar to Arc PDP, except it uses the
delegation policy parsed from remote proxy credential by
\textit{delegation.collector}, and evaluates the request against
delegation policy. See section 6for the design idea and use case of
delegation policy in fine-grained identity delegation.}

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{simplelist.pdp} -- Simplelist PDP}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The Simplelist PDP is a simplest implementation of policy decision
point. It will match the identity extracted from the remote credential
(or proxy credential) with local list of permitted identities.}

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{pdpservice.invoker} -- PDP Service Invoker}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The PDP Service Invoker is a client which can be used to invoke the PDP
Service which implements the same functionality as Arc PDP, except that
the evaluation request and response are carried by SOAP message. The
benefit of implementing PDP Service and PDP Service Invoker is that the
policy evaluation engine can be accessed remotely and maintained
centrally.}

\subsection[Policy Evaluation Engine]{Policy Evaluation Engine}
\subsubsection{Design of policy evaluation engine}
{\selectlanguage{english}\upshape\color{black}
The ARC1 defines specific evaluation request and policy schema. Based on
the schema definition, one policy evaluation engine is implemented. The
design principal of policy evaluation engine is generality by which the
implementation of the policy evaluation engine can be easily extended
to adopt some other policy schema, such as XACML policy schema.}

{\selectlanguage{english}\upshape\color{black}
Figure 5 shows the UML class diagram about the policy evaluation engine.
It shows all classes and relations simultaneously for getting the
overall picture. }

{\selectlanguage{english}\upshape\color{black}
The \textit{Evaluator} class is the key class for policy evaluation. It
accepts request evaluates it against loaded policy and returns
evaluation response. }

{\selectlanguage{english}\upshape\color{black}
Three abstract factories - \textit{FnFactory}, \textit{AlgFactory},
\textit{AttributeFactory} - are responsible for creating the
\textit{Function}, \textit{CombiningAlg} and \textit{AttributeValue}
objects correspondingly. The classes inherited from
\textit{CombiningAlg} class take care of implementing various combining
algorithms which define relations between
{\textless}Rule/{\textgreater} elements in policy. The
\textit{AttributeValue} type of classes are used for processing
different types of {\textless}Attribute/{\textgreater} and similar
elements. The \textit{Function} classes take care of comparing
{\textless}Attribute/{\textgreater} elements of request and policy. }

{\selectlanguage{english}\upshape\color{black}
The \textit{Policy} class parses {\textless}Policy/{\textgreater} or
{\textless}Rule/{\textgreater} elements and creates
\ \textit{CombingAlg} objects according to the
{\textless}RuleCombiningAlg/{\textgreater} attribute of
{\textless}Policy/{\textgreater}, \textit{Function} objects according
to the {\textless}Function/{\textgreater} attribute of
{\textless}Attribute/{\textgreater} and \textit{AttributeValue} objects
according to the {\textless}Type/{\textgreater} attribute of
{\textless}Attribute/{\textgreater}. Those objects will be used when
evaluating the request.}

{\selectlanguage{english}\upshape\color{black}
The \textit{Request} class is responsible for parsing
{\textless}Request/{\textgreater} element and creates corresponding
\textit{AttributeValue} objects according to the
{\textless}Type/{\textgreater} attribute of
{\textless}Attribute/{\textgreater}. When evaluating, each
\textit{AttributeValue} in request will be evaluated against
corresponding \textit{AttributeValue} in the policy by using relevant
\textit{Function}.}

{\selectlanguage{english}\upshape\color{black}
Due to extensible architecture of code it is relatively easy to add
support for new types of \ \textit{AttributeValue}, \textit{Function
and CombingAlg} objects in this way supporting various types of XML
based policy languages.}

\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img3.png}

\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img4.png}

\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img5.png}

\includegraphics[width=5.7043in,height=5.6354in]{SecurityFrameworkofARC1-img6.png}

{\centering\selectlanguage{english}\itshape\color{black}
Figure \stepcounter{Figure}{\theFigure}. The UML class diagram of the
classes inside policy evaluation engine. 
\par}

\subsubsection{Schemas for policy evaluation engine}
{\selectlanguage{english}\upshape\color{black}
The schema for ARC Policy is available at}

{\selectlanguage{english}\upshape\color{black}
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd}
.}

{\selectlanguage{english}\color{black}
The hierarchy tree of ARC Policy is show below (numbers show
multiplicity of elements)}

{\selectlanguage{english}\itshape\color{black}
Policy (1)}

{\selectlanguage{english}\itshape\color{black}
 Rule (1-)}

{\selectlanguage{english}\itshape\color{black}
  Subjects (1)}

{\selectlanguage{english}\itshape\color{black}
   Subject (1-)}

{\selectlanguage{english}\itshape\color{black}
    Attribute (1-)}

{\selectlanguage{english}\itshape\color{black}
   Resources (0-1)}

{\selectlanguage{english}\itshape\color{black}
   Resource (1-)}

{\selectlanguage{english}\itshape\color{black}
  Actions (0-1)}

{\selectlanguage{english}\itshape\color{black}
   Action (1-)}

{\selectlanguage{english}\itshape\color{black}
  Conditions (0-1)}

{\selectlanguage{english}\itshape\color{black}
   Condition (1-)}

{\selectlanguage{english}\itshape\color{black}
    Attribute (1-)}

{\selectlanguage{english}\upshape\color{black}
The schema for ARC Request is available at \ }

{\selectlanguage{english}\upshape\color{black}
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd}
.}

{\selectlanguage{english}\upshape\color{black}
The hierarchy tree of ARC Request is show below (numbers show
multiplicity of elements)}

{\selectlanguage{english}\itshape\color{black}
Request (1)}

{\selectlanguage{english}\itshape\color{black}
 RequestItem (1-)}

{\selectlanguage{english}\itshape\color{black}
  Subject (1-)}

{\selectlanguage{english}\itshape\color{black}
   SubjectAttribute (1-)}

{\selectlanguage{english}\itshape\color{black}
  Resource (0-)}

{\selectlanguage{english}\itshape\color{black}
  Action (0-)}

{\selectlanguage{english}\itshape\color{black}
  Context (0-)}

{\selectlanguage{english}\itshape\color{black}
   ContextAttribute (1-)}

{\selectlanguage{english}\upshape\color{black}
The schema for ARC Response is available at}

{\selectlanguage{english}\upshape\color{black}
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Response.xsd}
.}

{\selectlanguage{english}\upshape\color{black}
The ARC Response is not used directly in code. It is in use by PDP
Service which provides remote evaluation of policies. }

\subsubsection{Basic Elements of Policy}
{\selectlanguage{english}\upshape\color{black}
There are 2 basic objects - {\textquotedbl}policy{\textquotedbl} and
{\textquotedbl}request{\textquotedbl}. There is 1 main actor -
Evaluator. Curretly there are two types of elements in policy: Policy
and Rule. Policy element is made of Rule elements.}

{\color{black}
Evaluator matches request to policy and produces one of 4 following
results:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{PERMIT} - policy explicitely permits activity specified in
request because request matches some part of policy and corresponding
effect specified in policy is PERMIT. }
\end{itemize}
{\bfseries\itshape\color{black}
Example: }

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Rule: }}\textit{PERMIT
person ALICE to PLAY in place called WONDERLAND}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Request: }}\textit{person
ALICE wants to PLAY in place called WONDERLAND}}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{DENY} - policy explicitely denies activity specified in Request
because Request matches some part of policy and corresponding effect
specified in policy is DENY.}
\end{itemize}
{\bfseries\itshape\color{black}
Example:}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Rule: DENY }}\textit{fruit
APPLE to GROW on PEACH tree}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Request}}\textit{: fruit
APPLE to be GROWN on PEACH tree}}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{INDETERMINATE} - request has some part which does not correspond
to policy.}
\end{itemize}
{\bfseries\itshape\color{black}
Example: }

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Rule}}\textit{: DENY fruit APPLE to
GROW on PEACH tree}}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit APPLE to
be GROWN on WHEAT ground}}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: flower SUNFLOWER
to be grown on PEACH tree}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{Explaination}}: Here, it is not possible to obtain any
matching result - neither positive (DENY or PERMIT) nor negative
(NOT\_APPLICABLE, see below)}

{\color{black}
In the request, the {\textquotedbl}ground{\textquotedbl} is completely
uncomparable to the {\textquotedbl}tree{\textquotedbl} in policy. One
can compare {\textquotedbl}PEACH tree{\textquotedbl} and
{\textquotedbl}APPLE tree{\textquotedbl} because they are both
{\textquotedbl}tree{\textquotedbl}; But it is impossible to compare
{\textquotedbl}PEACH tree{\textquotedbl} and {\textquotedbl}WHEAT
ground{\textquotedbl} becaue they of different kind (Policy is about
tree and Request is about ground).}

{\color{black}
In a similar way one can{\textquotesingle}t compare {\textquotedbl}fruit
APPLE{\textquotedbl} and {\textquotedbl}flower SUNFLOWER{\textquotedbl}
(here policy is about fruits and Request is about flower).}

{\color{black}
Any other situation which makes it impossibile to compare two attributes
will also cause {\textquotedbl}INDETERMINATE{\textquotedbl}.}

\bigskip

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{NOT\_APPLICABLE} - all parts of the Request have corresponding
parts in the Policy, but some value of those parts are not the same.
Hence request does not match policy.}
\end{itemize}
{\bfseries\itshape\color{black}
Example:}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Rule}}\textit{: DENY fruit APPLE to
GROW on PEACH tree}}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit APPLE to
be GROWN on APPLE tree }}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit ORANGE to
be GROWN on PEACH tree }}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit ORANGE to
be GROWN on APPLE tree }}

{\color{black}
Explanation: for each part of the Request evaluator can find relevant
part in the Policy - both Policy and Request are about fruit and tree.
But the values do not match.}

{\color{black}
If it is required to reduce evaluation results to boolean value PERMIT
maps to TRUE and rest of results to FALSE.}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{Note}}\textit{: It would be useful to make it possible
to specify secondary effect which would become }active in case Request
is NOT\_APPLICABLE. For example:}

{\itshape\color{black}
DENY fruit APPLE to GROW on PEACH tree otherwise PERMIT}

{\color{black}
But one should be careful because example above would allow fruit PLUMS
to grow on APPLE trees :)}

{\color{black}
This kind of requirement can be supported by using the algorithm between
policies. \ For example, in case of above scenario, we can use some
algorithm like {\textquotedbl}Permit-if-notapplicable{\textquotedbl}.
See below the {\textquotedbl}Policy matching{\textquotedbl} part for
more explaination. \ }

\subsubsection{Policy MAtching}
{\selectlanguage{english}\upshape\color{black}
Policy is made of Rule elements. Request is evaluated against each Rule.
Each evaluation produces same results as policy evaluation described
above. The results from all Rules are then combined in order to produce
final result for whole policy. Results Combining Algorithm is specified
in Policy. There are 26 algorithms currently:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{\textit{Deny-Overrides}} - this is default if no algorithm
specified.}

\begin{itemize}
\item {\color{black}
If there is at least one DENY in results final result is DENY.}
\item {\color{black}
Otherwise if there is at least one PERMIT, the final result is PERMIT.}
\item {\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
\textbf{NOTE}\textbf{: CHECK if algorithm description is correct}}

{\color{black}
Special case is Policy wih no rules. Probably such policy shoudl be
treated as always producing DENY.}

{\bfseries\color{black}
Discussion}

{\itshape\color{black}
If there is no Rule under Policy, it means no restriction from this
Policy, what exactly shoud the Policy give? DENY or NOT\_APPLICABLE?}

{\itshape\color{black}
It seems to be more logical to produce INDETERMINATE because there is no
rule with elements which could be compared to request.}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\bfseries\itshape\color{black}
Permit-Overrides.}

\begin{itemize}
\item {\color{black}
If there is at least one PERMIT in results final result is PERMIT.}
\item {\color{black}
Otherwise if there is at least one DENY the final result is DENY.}
\item {\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\end{itemize}
{\color{black}
Special case is Policy wih no rules. Probably such policy should be
treated as always producing DENY.}

{\bfseries\color{black}
NOTE: CHECK if algorithm description is correct}

{\bfseries\color{black}
Discussion}

{\itshape\color{black}
Question? The same as above.}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\bfseries\itshape\color{black}
Ordered algorithms.}
\end{itemize}
{\selectlanguage{english}\color{black}
These specify priorities for all four possible results. Their names look
like Result1-Result2-Result3-Result4 with Result\# naming result types,
for example Permit-Deny-NotApplicable-Indeterminate. The results are
combined in following way:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item \begin{itemize}
\item {\color{black}
If there is at least one result of Result1 type then final result is
Result1.}
\item {\color{black}
Otherwise if there is at least one result of Result2 type then final
result is Result2.}
\item {\color{black}
Otherwise if there is at least one result of Result3 type then final
result is Result3.}
\item {\color{black}
Otherwise final result is Result4.}
\end{itemize}
\end{itemize}
{\color{black}
There are 24 possible combinations of those algorithms.}

{\selectlanguage{english}\upshape\color{black}
\textbf{Note:} It would be useful to have more combining algorithms. For
example }

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Permit-if-notapplicable - the use case could be {\textquotedbl}DENY
fruit APPLE to GROW on PEACH tree otherwise PERMIT{\textquotedbl}. In
this case there is only one Rule under Policy, and this Rule is with
{\textquotedbl}Deny{\textquotedbl} effect. }

\begin{itemize}
\item {\color{black}
If this Rule gives DENY in results, final result is DENY.}
\item {\color{black}
Otherwise if this Rule gives NOT\_APPLICABLE, final result is PERMIT.}
\item {\color{black}
Otherwise final result is INDETERMINATE. }
\end{itemize}
\item {\selectlanguage{english}\color{black}
{\textquotedblleft}Permit-if-allPermit{\textquotedbl} - Permit if all
the Rules gives Permit, this algorithm is useful in case if we are
collecting different policies from a few sources, and we want the
request to satisfy all of them.}

\begin{itemize}
\item {\color{black}
If all of the Rule give PERMIT, the final result is PERMIT.}
\item {\color{black}
Otherwise if there is at least one DENY the final result is DENY.}
\item {\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\item {\color{black}
OnlyOneApplicable}

\begin{itemize}
\item {\color{black}
If there is one gives INDETERMINATE, final result INDETERMINATE is given
immediately.}
\item {\color{black}
Otherwise if there is exactly only one gives applicable result (DENY or
PERMIT), final result is as this result.}
\item {\color{black}
Otherwise if there is more than one gives applicable result, final
result is INDTERMINATE.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is NOT\_APPLICABLE.}
\end{itemize}
\end{itemize}
{\color{black}
This algorithm makes sure that only one Rule is selected when making
decision.}

\liststyleWWviiiNumxvii
\begin{itemize}
\item {\color{black}
FirstApplicable}

\begin{itemize}
\item {\color{black}
If there is one give DENY, PERMIT or INDETERMINATE result, final result
is given immediately as this result.}
\item {\color{black}
Otherwise final result is NOT\_APPLICABLE.}
\end{itemize}
\end{itemize}
\subsubsection{Request Structure}
{\selectlanguage{english}\upshape\color{black}
Request is made of RequestItem elements. Each RequestItem is evaluated
against Policy Rule and for each evaluation separate result is
generated as described above.RequestItem is made of 4 elements:}

{\color{black}
Subject - represents entity requesting specified action}

\liststyleWWviiiNumxix
\begin{itemize}
\item {\color{black}
Resource - destination/object of the action}
\item {\color{black}
Action - specifies what has to be done on resource}
\item {\color{black}
Context - for additional information which does not fit anywhere else,
like the current time.}
\end{itemize}
{\color{black}
Effectively RequestItem may have only one Subject, one Resource, one
Action and one Context. If there are more than one element of any kind
of sub-elelemt, then in the evaluator this RequestItem is split into
several items containing all possible permutations and results are
obtained for every item separately. How results are combined will be
explained later.}

{\color{black}
Additionally Subject could contain sub-elements SubjectAttribute. Those
are meant to represent different kinds of requesters{\textquotesingle}
identities. Example:}

\liststyleWWviiiNumxviii
\begin{itemize}
\item {\color{black}
Subject}

\begin{itemize}
\item {\color{black}
SubjectAttribute: name is ALICE}
\item {\color{black}
SubjectAttribute: age is YOUNG}
\item {\color{black}
SubjectAttribute: gender is GIRL}
\end{itemize}
\end{itemize}
{\color{black}
Context could also be made of ContextAttribute elements in the same way
as Subject.}

{\color{black}
The following is an example of the Request:}

{\itshape\color{black}
{\textless}Request
xmlns={\textquotedbl}http://www.nordugrid.org/schemas/request-arc{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Subject
AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Action
AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Resource
AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Context
AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}

{\itshape\color{black}
{\textless}/Request{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
While evaluating this RequestItem will be split into two RequestItems:}

{\itshape\color{black}
{\textless}Request
xmlns={\textquotedbl}http://www.nordugrid.org/schemas/request-arc{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Action
AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Resource
AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Context
AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Subject
AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Action
AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Resource
AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Context
AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}

{\itshape\color{black}
{\textless}/Request{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
The following means this Subject possesses both of these Attributes.}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\color{black}
However, the following means two Subject each of which possesses one
Attribute.}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Subject
AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Subject
AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Subject{\textgreater}}}

{\color{black}
The {\textquotedbl}Type{\textquotedbl} xml-attribute is for
distinguishing how to process the xml-node value, which is critical
when evaluate two value from request side and policy side because
different type requires different evaluating/comparing approach. The
default {\textquotedbl}Type{\textquotedbl} is
{\textquotedbl}string{\textquotedbl}, in this case (also with the
{\textquotedbl}Function{\textquotedbl} xml-attribute on the policy side
is {\textquotedbl}equal{\textquotedbl}, which will be explained later),
each letters of these two values will be compared one by one when
evaluating them.}

{\color{black}
The {\textquotedbl}AttributeId{\textquotedbl} xml-attribute is for
evaluator to find the Attribute with AttributeId from the request side
which corresponds to the Attribute with the same AttributeId on the
policy side. Only if two Attributes{\textquotesingle} AttributeId are
equal, the evaluator will then compare the value.}

{\color{black}
Each RequestItem will be sequencialy and independently evaluated against
policy/policies. So for one Request (including few RequestItems), some
RequestItem could get positive evaluation result (PERMIT) from policy
engine, others could get negative evaluation result (DENY,
NOT\_APPLICABLE, INDETERMINATE).}

{\color{black}
It is up to policy decision point to make final decision according to
the evaluation results returned by evaluator, and the evaluator itself
can not give this kind of final decision.}

{\selectlanguage{english}\upshape\color{black}
\textbf{NOTE: This probably should be changed because evaluator is fed
with complete policy and complete request. Hence it is illogical that
it retu}\textbf{rns multiple decisions.}}

{\color{black}
Basically the policy decision point will feed policy engine with
request, get back evaluation results, }

{\color{black}
and make final decision.}

\bigskip

\subsubsection{Rule Composition and Matching}
{\selectlanguage{english}\upshape\color{black}
Policy rule is made of 4 elements - Subjects, Resources, Actions,
Conditions (See the following example). Those are only used to group
multiple elements Subject, Resource, Action, Condition. For instance,
you can merge two Rules with the same Resources, Actions, Conditions,
and the same {\textquotedbl}Effect{\textquotedbl} but different
Subjects into one Rule.}

{\selectlanguage{english}\upshape\color{black}
\textbf{NOTE: That is strange. Why do we need }\textbf{Rule at all?}}

{\selectlanguage{english}\upshape\color{black}
There is no logical relationship between Subject, which means you can
split one Rule into two Subject (under Subjects) into two Rule (each of
which has one Subject (under Subjects)).From now only later ones
(Subjects with only one Subject as sub-element, and the same for
others) are described. Their meaning is same as in request with
Condition corresponding to Context. Subject and Condition elements are
also made of Attributes. All elements may be present more than one
time. During procedure of matching each element in RequestItem is
matched against all elements of same kind in Policy - Subject is
matched to Subject, Resource to Resource, etc. For every combination 3
possible results are produced:}

\liststyleWWviiiNumxx
\begin{itemize}
\item {\color{black}
MATCHED - element from RequestItem matched element in Policy Rule.
Example:}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{RequestItem Resource}}\textit{: place called
WONDERLAND}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{PolicyItem Resource}}\textit{: place called WONDERLAND}}

\liststyleWWviiiNumxx
\begin{itemize}
\item {\color{black}
NOT MATCHED - element from RequestItem did not match element in Policy
Rule. Example:}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{RequestItem Resource}}\textit{: place called
WONDERLAND}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{PolicyItem Resource}}\textit{: place called PLAYGROUND}}

\liststyleWWviiiNumxx
\begin{itemize}
\item {\color{black}
INDETERMINATE- element from RequestItem could not be compared to element
in Policy \ Rule because they are of incompatible ids/belong to
different namespaces. Example:}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{RequestItem Resource}}\textit{: place called WONDERLAND
(with namespace {\textquotedbl}place{\textquotedbl})}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{PolicyItem Resource}}\textit{: LEMON tree (with
namespace {\textquotedbl}tree{\textquotedbl})}}

{\color{black}
The produced results then combined to produce final 4 types of results
in following way:}

\liststyleWWviiiNumxx
\begin{itemize}
\item {\color{black}
If for every element in RequestItem there is at least one MATCHED result
then result for this Policy Rule is as specified in the corresponding
Effect (Deny or Permit).}
\item {\color{black}
Otherwise if for every element in RequestItem there is at least one gets
INDETERMINATE result then result for Policy Rule is INDETERMINATE.}
\item {\color{black}
Otherwise result is NOT\_APPLICABLE.}
\end{itemize}
{\color{black}
Special case is then RequestItem does not have the element(s) of some
kind (Subject, Action, Resource or Context/Condition). If there are
elements of corresponding kind in the Policy Rule then such situation
should be considered as INDETERMINATE.}

{\color{black}
The following is an example of the Policy:}

{\itshape\color{black}
{\textless}Policy
xmlns={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc{\textquotedbl}
CombiningAlg={\textquotedbl}Permit-Overrides{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}Rule
Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ }\textit{{\textless}Subjects{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject
AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ }\textit{{\textless}/Subjects{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ }\textit{{\textless}Actions{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Action
AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Action
AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Delete{\textless}/Action{\textgreater}
\ \ }}

{\selectlanguage{english}\color{black}
\textit{\ \ }\textit{{\textless}/Actions{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ }\textit{{\textless}Resources{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Resource
AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textquotedbl}{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ }\textit{{\textless}/Resources{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ }\textit{{\textless}Conditions{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Condition
AttributeId={\textquotedbl}urn:knowarc:period{\textquotedbl}
Type={\textquotedbl}period{\textquotedbl}
Function={\textquotedbl}Inrange{\textquotedbl}{\textgreater}2008-09-10T20:30:20/P1Y1M{\textless}/Condition{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ }\textit{{\textless}/Conditions{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}/Rule{\textgreater}}}

{\itshape\color{black}
{\textless}/Policy{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
For the Subject which includes two Attributes in this example:}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
These two attributes mean the Rule requires the request should possess
both of these two attributes.}

{\color{black}
However, if You put these above two Attribute into two Subject
elements:}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject
AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject
AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
Then it means the Rule requires the request to possess at least one of
these two attributes.}

{\color{black}
For the xml-attribute {\textquotedbl}Type{\textquotedbl} and
{\textquotedbl}AttributeId{\textquotedbl}, the explaination for Request
example also applies here.}

{\color{black}
The {\textquotedbl}Function{\textquotedbl} xml-attribute is for
distinguishing different comparison algorithm when comparing these two
xml-node value. If Function is absent,
{\textquotedbl}equal{\textquotedbl} will be used as default.}

\subsubsection{Rule Elements Matching}
{\color{black}
For elements without attributes those elements have:}

\liststyleWWviiiNumxxii
\begin{itemize}
\item {\color{black}
Kind specified by AttributeId XML attribute. There is no default.}
\item {\color{black}
Matching algorithm specified by Id XML attribute. By default
string-equal matching is used.}
\item {\color{black}
Content}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
Example: \textit{LEMON tree}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{Kind}}\textit{: tree}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{Matching algorithm}}\textit{: default}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{Content}}\textit{: LEMON}}

{\color{black}
Matching procedure consists of following steps:}

\liststyleWWviiiNumxxiii
\begin{itemize}
\item {\color{black}
Kinds are compared using simple string equal matching. If those do not
match then result is INDETERMINATE.}
\item {\color{black}
Matching algorithm is used to compare content of elements. Result is
either MATCH or NO\_MATCH according to matching algorithm.}
\end{itemize}
{\color{black}
Each element on the RequestItem must satisfy corresponding element in
Rule.In detail, for Subjects element under Rule, if there is at least
one Subject (with one Attribute or a few Attribute) which is matched by
a Subject on this RequestItem, we say this Subjects is matched by the
RequestItem; and also the same for the other elements (Actions,
Resources, Conditions).}

{\selectlanguage{english}\upshape\color{black}
For elements with multiple Attribute sub-elements the way to judging
whether elements match is if and only if all of the Attribute under the
Rule have matching Attributes at RequestItem side.}

{\selectlanguage{english}\upshape\color{black}
Example of the Subject with three Attributes:}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{Subject}}\textit{:}}

{\selectlanguage{english}\color{black}
\textbf{\textit{SubjectAttribute}}\textit{: name is ALICE}}

{\selectlanguage{english}\color{black}
\textbf{\textit{SubjectAttribute}}\textit{: age is YOUNG}}

{\selectlanguage{english}\color{black}
\textbf{\textit{SubjectAttribute}}\textit{: gender is GIRL}}

{\selectlanguage{english}\upshape\color{black}
In XML that is:}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}gender{\textgreater}GIRL{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
That requires the Subject in the RequestItem to possess at least these
three Attributes.}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}age{\textquotedbl}{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}gender{\textquotedbl}{\textgreater}GIRL{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other
Attribute-{}-{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\itshape\color{black}
{\textless}/RequestItem{\textgreater}}

{\color{black}
The above example shows that the Subject in the RequestItem
{\textquotedbl}MATCH{\textquotedbl} one Subject on the Rule side.}

{\color{black}
If the Subject in the RequestItem is like this:}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}from{\textquotedbl}{\textgreater}OSLO{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other Attribute, buts
not a {\textquotedbl}gender{\textquotedbl}-{}-{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
Then evaluator will produce INDETERMINATE as the match-making result of
this Subject.}

{\color{black}
If the Subject in the RequestItem is like this:}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Bob{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
AttributeId={\textquotedbl}gender{\textquotedbl}{\textgreater}BOY{\textless}/Attribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other
Attribute-{}-{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}

{\color{black}
Then evaluator will give NO\_MATCH as the match-making result of this
Subject.}

{\selectlanguage{english}\upshape\color{black}
Finally if and only if all of the elemens (Subjects, Actions, Resources,
Conditions) which are not empty under the Rule have been matched (gets
MATCH) to the RequestItem, then the whole Rule is considered to be
matched (produces MATCH result). MATCH is then mapped to final
evaluation result depending on the specified Effect. If Effect is set
to Deny then DENY decision will be produced for this Rule; if Effect is
Permit then PERMIT.}

{\color{black}
Otherwise if any of the element (Subjects, Actions, Resources,
Conditions) of RequestItem got INDETERMINATE decision then the
INDETERMINATE decision will be made for this Rule.}

{\color{black}
Otherwise the NOT\_APPLICABLE decision will be made for this Rule. In
other words that means at least one of the elements of this Rule got
NO\_MATCH and the other elements got MATCH.}

\subsubsection{Interface for using the policy evaluation engine}
{\selectlanguage{english}\upshape\color{black}
For making usage of policy evaluation engine more convenient basic
\textit{Evaluator} class is complemented by additional interfaces.
Below are examples of steps needed to carry out policy evaluation and
corresponding helper interfaces.}

\liststyleWWviiiNumiii
\begin{enumerate}
\item {\selectlanguage{english}\color{black}
Create the policy evaluation object:}
\end{enumerate}
{\selectlanguage{english}\ttfamily\color{black}
// Create object which provides an interface }

{\selectlanguage{english}\ttfamily\color{black}
// for loading other objects}

{\selectlanguage{english}\ttfamily\color{black}
ArcSec::EvaluatorLoader eval\_loader; }

{\selectlanguage{english}\ttfamily\color{black}
//Load the Evaluator }

{\selectlanguage{english}\ttfamily\color{black}
ArcSec::Evaluator* eval = NULL;}

{\selectlanguage{english}\ttfamily\color{black}
// Define name of policy evaluator.}

{\selectlanguage{english}\ttfamily\color{black}
// This one is for evaluation ARC policies}

{\selectlanguage{english}\ttfamily\color{black}
std::string evaluator = {\textquotedbl}arc.evaluator{\textquotedbl}; }

{\selectlanguage{english}\ttfamily\color{black}
eval = eval\_loader.getEvaluator(evaluator);}

\liststyleWWviiiNumiv
\begin{enumerate}
\item {\selectlanguage{english}\color{black}
Create the policy object:}
\end{enumerate}
{\selectlanguage{english}\ttfamily\color{black}
ArcSec::Policy* policy = NULL; }

{\selectlanguage{english}\ttfamily\color{black}
// Define type of policy -- ARC policy in this case}

{\selectlanguage{english}\ttfamily\color{black}
std::string policyclassname = {\textquotedbl}arc.policy{\textquotedbl};
}

{\selectlanguage{english}\ttfamily\color{black}
// Define source from which policy to be taken}

{\selectlanguage{english}\ttfamily\color{black}
ArcSec::SourceFile
policy\_source({\textquotedbl}Policy\_Example.xml{\textquotedbl}); }

{\selectlanguage{english}\ttfamily\color{black}
// Load and parse policy}

{\selectlanguage{english}\ttfamily\color{black}
policy = eval\_loader.getPolicy(policyclassname, policy\_source); }

\liststyleWWviiiNumv
\begin{enumerate}
\item {\selectlanguage{english}\color{black}
Create the request:}
\end{enumerate}
{\selectlanguage{english}\ttfamily\color{black}
ArcSec::Request* request = NULL;}

{\selectlanguage{english}\ttfamily\color{black}
// Define type of request -- ARC request in this case}

{\selectlanguage{english}\ttfamily\color{black}
std::string requestclassname =
{\textquotedbl}arc.request{\textquotedbl};}

{\selectlanguage{english}\ttfamily\color{black}
// Define source from which request to be taken}

{\selectlanguage{english}\ttfamily\color{black}
ArcSec::SourceFile
request\_source({\textquotedbl}Request.xml{\textquotedbl});}

{\selectlanguage{english}\ttfamily\color{black}
// Load and parse request}

{\selectlanguage{english}\ttfamily\color{black}
request = eval\_loader.getRequest(requestclassname, request\_source);}

\liststyleWWviiiNumvi
\begin{enumerate}
\item {\selectlanguage{english}\color{black}
Add the policy into \textit{Evaluator} object:}
\end{enumerate}
{\selectlanguage{english}\ttfamily\color{black}
eval-{\textgreater}addPolicy(policy);}

\liststyleWWviiiNumvii
\begin{enumerate}
\item {\selectlanguage{english}\color{black}
Evaluate the request object:}
\end{enumerate}
{\selectlanguage{english}\ttfamily\color{black}
ArcSec::Response *resp = NULL;}

{\selectlanguage{english}\ttfamily\color{black}
resp = eval-{\textgreater}evaluate(request);}

{\selectlanguage{english}\color{black}
The steps d) and e) can also be replaced by:}

{\selectlanguage{english}\ttfamily\color{black}
resp = eval-{\textgreater}evaluate(request, policy);}

\bigskip

{\selectlanguage{english}\upshape\color{black}
The \textit{Evalutor::evaluate()} method can also be feed up with both
\textit{Policy}/\textit{Request} objects and their sources in any
combination. See example code at
http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/testinterface.cpp
for more details about usage of the interface.}

{\selectlanguage{english}\upshape\color{black}
The description of mentioned classes and their methods are avaialble in
\ API document at
\ \url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/KnowARC-API.pdf?format=raw}
.}

\subsection{Policy Decision Service}
{\selectlanguage{english}\upshape\color{black}
Policy decision service is a service implementation which contains the
functionality of ArcPDP. It will accept the soap request containing
policy decision request and return soap response containing policy
decision response.}

{\selectlanguage{english}\upshape\color{black}
The WSDL description of policy decision service is available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl}
. It{\textquotesingle}s configuration is presented in Section 7.5. }

\subsection{Security Attributes}
\subsubsection[Infrastructure]{Infrastructure}
\label{bkm:Ref204009818}{\selectlanguage{english}\upshape\color{black}
Security Attributes represent security related information inside HED
framework and store information representing various aspects needed to
perform authorization decison - identity of client, requested action,
targeted resource, constraint policies. }

{\selectlanguage{english}\upshape\color{black}
Each kind of Security Attribute is represented by own class inherited
from parent SecAttr class
{\textless}arc/message/SecAttr.h{\textgreater}. Each Security Attribute
stores it{\textquotesingle}s information in internal format and is
capable to export it to one of predefined formats using Export()
method. Currently only supported format is ARC Policy/Request XML
document described in Section 7.1and 7.2}

{\selectlanguage{english}\upshape\color{black}
Collectors of Security Attributes instantiate corresponding classes and
link them to Secuirity Attributes containers - MessageAuth
{\textless}arc/message/MessageAuth.h{\textgreater} and
MessageAuthContext {\textless}arc/message/Message.h{\textgreater}
storing collected attributes per request and per session
correspondingly. Each attribute is assigned a name. Current
implementations of Security Attributes Collectors are either integrated
into existing MCCs or implemented as separate SecHandler plugins. See
Section 5.2for available Collectors and corresponding Security
Attributes. }

{\selectlanguage{english}\upshape\color{black}
Note for service developers: Services may implement own authorization
algorithms. But they may use Security Atributes as well by providing
instances of classes inherited from SecAttr and running them through
either configured or hardcoded processors/PDPs.}

{\selectlanguage{english}\upshape\color{black}
Processors of Security Attributes are implemented as Policy Decision
Point components. Currently there are 2 PDP components available:}

\liststyleWWviiiNumviii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Arc PDP makes use of Security Attributes containing identities of
client, resource and requested action. It evaluates either all or
selected set of attributes against specified Policy documents thus
making it possible to enforce policies defined/selected by service
providers. }
\item {\selectlanguage{english}\color{black}
Delegation PDP is described below in Section 6.3}
\end{itemize}
\subsubsection[Available collectors]{Available collectors}
\label{bkm:Ref204009636}{\selectlanguage{english}\upshape\color{black}
Here Security Attribute collectors distributed as part of the ARC1 are
described except those used for Delegation Restrictions. Those are
described in Section 6.2}

\paragraph{TCP}
{\selectlanguage{english}\upshape\color{black}
Information is collected inside TCP MCC. The Security Attribute is
stored under name {\textquotesingle}TCP{\textquotesingle} and exports
ARC Request with following attributes:}

\begin{flushleft}
\bottomcaption{Security Attributes collected at TCP MCC}
\tablehead{\hline
\centering \selectlanguage{english}\bfseries\itshape\color{black}
Element &
\centering \selectlanguage{english}\bfseries\itshape\color{black}
AttributeId &
\centering\arraybslash
\selectlanguage{english}\bfseries\itshape\color{black}
Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
\selectlanguage{english}\color{black} Resource &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/localendpoint &
\selectlanguage{english}\color{black} service\_ip[:service\_port]
\\\hline
\selectlanguage{english}\color{black} SubjectAttribute &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/remoteendpoint &
\selectlanguage{english}\color{black} client\_ip[:client\_port] \\\hline
\end{supertabular}
\end{flushleft}

\bigskip

\paragraph{TLS}
{\selectlanguage{english}\upshape\color{black}
Information is collected inside TLS MCC. Generated Security Attribute
class is stored under name {\textquotesingle}TLS{\textquotesingle} and
exports ARC Request with following attributes:}

\begin{flushleft}
\bottomcaption{Security Attributes collected at TLS MCC}
\tablehead{\hline
\centering \selectlanguage{english}\bfseries\itshape\color{black}
Element &
\centering \selectlanguage{english}\bfseries\itshape\color{black}
AttributeId &
\centering\arraybslash
\selectlanguage{english}\bfseries\itshape\color{black}
Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
\selectlanguage{english}\color{black} SubjectAttribute &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/tls/ca &
\selectlanguage{english}\color{black} signer of first certificate in
client{\textquotesingle}s chain\\\hline
\selectlanguage{english}\color{black} SubjectAttribute &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/tls/chain &
\selectlanguage{english}\color{black} Subject of certificate in
client{\textquotesingle}s chain - multiple items\\\hline
\selectlanguage{english}\color{black} SubjectAttribute &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/tls/subject &
\selectlanguage{english}\color{black} Subject of last certificate in
client{\textquotesingle}s chain\\\hline
\selectlanguage{english}\color{black} SubjectAttribute &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/tls/identity &
\selectlanguage{english}\color{black} Subject of last non-proxy
certificate in client{\textquotesingle}s chain\\\hline
\end{supertabular}
\end{flushleft}
\paragraph{HTTP}
{\selectlanguage{english}\upshape\color{black}
Information is collected inside HTTP MCC. The Security Attribute is
stored under name {\textquotesingle}HTTP{\textquotesingle} and exports
ARC Request with following attributes:}

\begin{flushleft}
\bottomcaption{Security Attributes collected at HTTP MCC}
\tablehead{\hline
\centering \selectlanguage{english}\bfseries\itshape\color{black}
Element &
\centering \selectlanguage{english}\bfseries\itshape\color{black}
AttributeId &
\centering\arraybslash
\selectlanguage{english}\bfseries\itshape\color{black}
Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
\selectlanguage{english}\color{black} Resource &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/http/path &
\selectlanguage{english}\color{black} HTTP path without host and port
part\\\hline
\selectlanguage{english}\color{black} Action &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/http/method &
\selectlanguage{english}\color{black} HTTP method\\\hline
\end{supertabular}
\end{flushleft}
\paragraph{SOAP}
{\selectlanguage{english}\upshape\color{black}
Information is collected inside SOAP MCC. Security Attribute is stored
under name {\textquotesingle}SOAP{\textquotesingle} and exports ARC
Request with following attributes:}

\begin{flushleft}
\bottomcaption{Security Attributes collected at SOAP MCC}
\tablehead{\hline
\centering \selectlanguage{english}\bfseries\itshape\color{black}
Element &
\centering \selectlanguage{english}\bfseries\itshape\color{black}
AttributeId &
\centering\arraybslash
\selectlanguage{english}\bfseries\itshape\color{black}
Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
\selectlanguage{english}\color{black} Resource &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint &
\selectlanguage{english}\color{black} To element of WS-Addressing
structure\\\hline
\selectlanguage{english}\color{black} Action &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/soap/operation &
\selectlanguage{english}\color{black} SOAP top level element name
without namespace prefix\\\hline
\selectlanguage{english}\color{black} Context &
\selectlanguage{english}\color{black}
http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace &
\selectlanguage{english}\color{black} Namespace of SOAP top level
element\\\hline
\end{supertabular}
\end{flushleft}
\subsection[Delegation Restrictions]{Delegation Restrictions}
\label{bkm:Ref204009393}\subsubsection[Delegation
Architecture]{Delegation Architecture}
\label{bkm:Ref204009914}{\selectlanguage{english}\upshape\color{black}
In current implementation delegation is achieved through Identity
Delegation implemented using X509 Proxy Certificates as defined in RFC
3820. Client wishing to allow service to act on it{\textquotesingle}s
behalf provides Proxy Certificate to the service using Web Service
based Delegation interface described in Section 6.4}

{\selectlanguage{english}\upshape\color{black}
For limiting the scope of delegated credentials along with usually used
time constraints it is possible to attach Policy document to Proxy
Certificate. According to RFC 3820 Policy is stored in ProxyPolicy
extension. In order not to introduce new type of object Policy is
assigned id-ppl-anyLanguage identifier. RFC 3820 allows any octet
string associated with such object. We are using textual representation
of ARC Policy XML document.}

{\selectlanguage{english}\upshape\color{black}
Each deployment implementing Delegation Restrictions must use dedicated
Security Handler plugin (see section 5.1) to collect all Policy
documents from Proxy Certificates used for establishing secure
connection. Then those documents must be processed by dedicated Policy
Decision Point plugin (see section 2.3) to make a final decision based
on collected Policies and various information about
client{\textquotesingle}s identity and requested operation. Service or
MCC chain supporting Delegation Restrictions must accept negative
decision of this PDP as final and do not override it with any other
decision based on other policies.}

\subsubsection[Delegation Collector]{Delegation Collector}
\label{bkm:Ref204009734}{\selectlanguage{english}\upshape\color{black}
This Security Attribute is collected by dedicated Security Handler
plugin named {\textquotedbl}delegation.collector{\textquotedbl}
avaialble as part of the ARC1 distribution. It extracts policy document
stored inside X509 certificate proxy extension as defined in RFC3820
and described in Section 6.1All proxy certificates in a chain provided
by client are examined and all available policies are extracted. }

{\selectlanguage{english}\upshape\color{black}
Extracted content is converted into XML document. Then document is
checked to be of ARC Policy kind. If policy is not recognized as ARC
Policy procedure fails and that causes failure of communication.}

{\selectlanguage{english}\upshape\color{black}
Proxy certificates with id-ppl-inheritAll [5. RFC3820.
http://www.faqs.org/rfcs/rfc3820.html] property are passed through and
no policy document is generated for them. Proxies with other type of
policies including id-ppl-independent are not accepted and generate
immediate failure.}

\subsubsection[Delegation PDP]{Delegation PDP}
\label{bkm:Ref204009695}{\selectlanguage{english}\upshape\color{black}
The Delegation PDP is similar to the Arc PDP described above except that
it takes it{\textquotesingle}s Policy documents directly from Security
Attributes. Differently from Arc PDP it is meant to be used for
enforcing policies defined by client.}

\subsubsection[Delegation interface]{Delegation interface}
\label{bkm:Ref204009766}{\selectlanguage{english}\upshape\color{black}
Delegation interface in the ARC1 is implemented using Web Service
approach. Each ARC1 service wishing to accept delegated credentials
implements this interface. Here is how delegation procedure works:}

\liststyleWWviiiNumix
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Step 1}
\item {\selectlanguage{english}\color{black}
Client contacts service requesting operation DelegateCredentialsInit.
This operation has no arguments.}
\item {\selectlanguage{english}\color{black}
Service responds with DelegateCredentialsInitResponse message with
element TokenRequest. That element contains credentials request
generated by service in Value. Type of request is defined by attribute
Format. Currently only supported format is x509. Along with Value
service provides identifier Id which is used in second step.}
\end{itemize}
\liststyleWWviiiNumx
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Step 2}
\item {\selectlanguage{english}\color{black}
Client requests UpdateCredentials operation with DelegatedToken
argument. This element contains Value with serialized delegated
credentials and Id which links it to first step. Delegated token
element may also contain multiple Reference elements. Reference refers
to the object which these credentials should be applied to in a way
specific to the service. The DelegatedToken element may also be used
for delegating credentials when Step 2 is combined with other
operations on service in service specific way.}
\item {\selectlanguage{english}\color{black}
Service responds with empty UpdateCredentialsResponse message.}
\end{itemize}
\subsection[Schemas, descriptions and examples]{Schemas, descriptions
and examples}
\label{bkm:Ref204010074}\subsubsection[Authorization
Policy]{Authorization Policy}
\label{bkm:Ref204009564}{\selectlanguage{english}\upshape\color{black}
XML schema with comments available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd}
.}

\subsubsection[Authorization Request]{Authorization Request}
\label{bkm:Ref204009595}{\selectlanguage{english}\upshape\color{black}
XML schema with comments available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd}{\textstyleInternetlink{
.}}}

\subsubsection{Authorization Response}
{\selectlanguage{english}\upshape\color{black}
XML schema with comments available at
\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Response.xsd}}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd}{\textstyleInternetlink{
.}}}

\subsubsection[Interface of policy decision service]{Interface of policy
decision service}
{\selectlanguage{english}\upshape\color{black}
WSDL with comments available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl}
.}

\subsubsection{Configuration of PDP service}
{\selectlanguage{english}\upshape\color{black}
XML schema with comments available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.xsd}
.}

{\selectlanguage{english}\upshape\color{black}
Below is an example configuration of PDP service which can evaluate ARC
Request against ARC Policy stored in local file.}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}Service name={\textquotedbl}pdp.service{\textquotedbl}
id={\textquotedbl}pdp\_service{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}!-{}-The element {\textless}Evaluator/{\textgreater},
{\textless}Policy/{\textgreater} and {\textless}Request/{\textgreater}
configuration}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ are supposed to be used to load object; element
{\textless}PolicyStore/{\textgreater} is}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ supposed to be used to get the location of
policy-{}-{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ {\textless}pdp:PDPConfig{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}pdp:PolicyStore{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}Location
Type={\textquotedbl}file{\textquotedbl}{\textgreater}Policy\_Example.xml{\textless}/Location{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}- other policy
location-{}-{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}/pdp:PolicyStore{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}pdp:Evaluator
name={\textquotedbl}arc.evaluator{\textquotedbl} /{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}pdp:Policy
name={\textquotedbl}arc.policy{\textquotedbl} /{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}pdp:Request
name={\textquotedbl}arc.request{\textquotedbl} /{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ {\textless}/pdp:PDPConfig{\textgreater}}}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}/Service{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
See Section 7.7for the explanation of ARC Policy.}

\subsubsection{SimpleList PDP configuration and Policy Example}
{\selectlanguage{english}\upshape\color{black}
XML schema with comments available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/simplelistpdp/SimpleListPDP.xsd}
.}

{\selectlanguage{english}\upshape\color{black}
Below is an example configuration of SimpleList PDP inside
{\textquotedblleft}echo{\textquotedblright} service.}

{\selectlanguage{spanish}\ttfamily\color{black}
{\textless}Service name={\textquotedbl}echo{\textquotedbl}
id={\textquotedbl}echo{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}SecHandler
name={\textquotedbl}arc.authz{\textquotedbl}
id={\textquotedbl}authz{\textquotedbl}
event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}PDP
name={\textquotedbl}}\texttt{simplelist}\texttt{.pdp{\textquotedbl}}\texttt{
location=}\texttt{{\textquotedbl}}\texttt{simplelist}\texttt{{\textquotedbl}}\texttt{/}\texttt{{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ }\foreignlanguage{spanish}{\texttt{{\textless}/SecHandler{\textgreater}}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
{\textless}/echo:prefix{\textgreater}}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}echo:suffix{\textgreater}
]{\textless}/echo:suffix{\textgreater}}}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}/Service{\textgreater}}

{\selectlanguage{english}\color{black}
The attribute {\textquotedblleft}name{\textquotedblright} of
{\textless}PDP/{\textgreater} is critical for loading the object.
Specifically, the name
{\textquotedblleft}simplelist.pdp{\textquotedblright} is for loading
the SimpleList PDP object.}

{\selectlanguage{english}\upshape\color{black}
The policy file {\textquotedblleft}simplelist{\textquotedblright} is a
local file which contains the list of X509 subjects of authorized
entities. It the peer certificate is proxy certificate, the identity in
this list should only include the original DN of
users{\textquotesingle}s certificate.}

{\selectlanguage{english}\upshape\color{black}
For example content of \textit{simplelist} file may look like this:}

{\selectlanguage{spanish}\color{black}
/C=NO/O=UiO/CN=test1}

{\selectlanguage{spanish}\color{black}
/C=NO/O=UiO/CN=test2}

\subsubsection[Arc PDP configuration and Policy Example]{Arc PDP
configuration and Policy Example}
\label{bkm:Ref204009974}\label{bkm:RefSecarcpdpconfigandpolicy}{\selectlanguage{english}\upshape\color{black}
XML schema with comments available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/ArcPDP.xsd}
.}

{\selectlanguage{english}\upshape\color{black}
Below is an example of configuration of Arc PDP inside
{\textquotedblleft}echo{\textquotedblright} service.}

{\selectlanguage{spanish}\ttfamily\color{black}
{\textless}Service name={\textquotedbl}echo{\textquotedbl}
id={\textquotedbl}echo{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}SecHandler
name={\textquotedbl}arc.authz{\textquotedbl}
id={\textquotedbl}authz{\textquotedbl}
event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}PDP
name={\textquotedbl}arc.pdp{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ {\textless}PolicyStore{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ {\textless}Location
type={\textquotedbl}file{\textquotedbl}{\textgreater}Policy\_Example.xml{\textless}/Location{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ {\textless}!-{}-other policy
location-{}-{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ {\textless}/PolicyStore{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}/PDP{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ }\foreignlanguage{spanish}{\texttt{{\textless}/SecHandler{\textgreater}}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
{\textless}/echo:prefix{\textgreater}}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}echo:suffix{\textgreater}
]{\textless}/echo:suffix{\textgreater}}}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}/Service{\textgreater}}

{\selectlanguage{english}\color{black}
The name {\textquotedblleft}arc.pdp{\textquotedblright} is for loading
the ArcPDP object.}

{\selectlanguage{english}\color{black}
There could be a few policy files under
{\textless}PolicyStore/{\textgreater}. The request will be checked
against all of the policies.}

{\selectlanguage{english}\upshape\color{black}
There is an example policy for echo service below. See Section 7.1for
the policy schema. The example policy is \ made of following elements:
}

{\selectlanguage{english}\color{black}
1. Line 14 \ defines resource being protected. In this it is everything
located under HTTP path
\ {\textquotedblleft}/Echo{\textquotedblright}.}

{\selectlanguage{english}\color{black}
2. Lines 17 and 18 define allowed HTTP operations to be
{\textquotedblleft}POST{\textquotedblright} and
{\textquotedblleft}GET{\textquotedblright}. Line 19 also defines SOAP
operation {\textquotedblleft}echo{\textquotedblright} to be applied to
service at path defined above.}

{\selectlanguage{english}\color{black}
3. Lines 10 and 9 require the requester to present X509 certificate with
specified identity and signed by specified Certification Authority.}

{\selectlanguage{english}\color{black}
4. No {\textless}Conditions/{\textgreater} defined.}

{\selectlanguage{english}\color{black}
5. Line 3 defines that if and only if all of the above constraints have
been satisfied by requester, the {\textless}Rule/{\textgreater}
evaluates to Permit decision.}

{\selectlanguage{english}\color{black}
The Secuirity Attributes used by Arc PDP are collected by different
MCCs. It is possible for service to collect some application-specific
attributes by implementing class inherited from SecAtt. And that should
be the task of application developer.}

{\selectlanguage{english}\upshape\color{black}
Administrator of service can configure Authorization SecHandler -
arc.authz - for each MCC and Service and define reasonable and
meaningful policy. While defining policy the administrator must take
into account that the attributes defined in the policy should be
already collected by previous components in a chain. For instance,
policy with AttributeId
{\textquotedblleft}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedblright}
should not be configured inside SecHandler attached to MCCTLS. }

\liststyleWWviiiNumxiv
\begin{enumerate}
\item {\selectlanguage{english}\ttfamily\color{black}
{\textless}?xml version={\textquotedbl}1.0{\textquotedbl}
encoding={\textquotedbl}UTF-8{\textquotedbl}?{\textgreater}}
\item {\selectlanguage{english}\ttfamily\color{black}
{\textless}Policy
xmlns={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc{\textquotedbl}
PolicyId={\textquotedbl}sm-example:arcpdppolicy{\textquotedbl}
CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ {\textless}Rule
Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ {\textless}Description{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ Example policy for echo service}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ {\textless}/Description{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Subjects{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Subject{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/ca{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=CA{\textless}/Attribute{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/identity{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=test{\textless}/Attribute{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Subject{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}/Subjects{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Resources{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Resource
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}/Echo{\textless}/Resource{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}/Resources{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}POST{\textless}/Action{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}GET{\textless}/Action{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/operation{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}echo{\textless}/Action{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Conditions/{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
\item {\selectlanguage{english}\ttfamily\color{black}
{\textless}/Policy{\textgreater}}
\end{enumerate}
\subsubsection{PDP Service Invoker configuration}
{\selectlanguage{english}\upshape\color{black}
Configuration XML schema with comments is available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/pdpserviceinvoker/ArcPDPServiceInvoker.xsd}
.}

{\selectlanguage{english}\upshape\color{black}
Below is an example of configuration of PDP Service Invoker inside
{\textquotedblleft}echo{\textquotedblright} service.}

{\selectlanguage{spanish}\ttfamily\color{black}
{\textless}Service name={\textquotedbl}echo{\textquotedbl}
id={\textquotedbl}echo{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}SecHandler
name={\textquotedbl}arc.authz{\textquotedbl}
id={\textquotedbl}authz{\textquotedbl}
event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}!-{}-Remote pdp service
invoking-{}-{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ {\textless}PDP
name={\textquotedbl}pdpservice.invoker{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ {\textless}ServiceEndpoint{\textgreater}https://127.0.0.1:60001/pdp.service{\textless}/ServiceEndpoint{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ {\textless}KeyPath{\textgreater}./key.pem{\textless}/KeyPath{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ {\textless}CertificatePath{\textgreater}./cert.pem{\textless}/CertificatePath{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ {\textless}CACertificatePath{\textgreater}./ca.pem{\textless}/CACertificatePath{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ {\textless}/PDP{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ }\foreignlanguage{spanish}{\texttt{{\textless}/SecHandler{\textgreater}}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}next
id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
{\textless}/echo:prefix{\textgreater}}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}echo:suffix{\textgreater}
]{\textless}/echo:suffix{\textgreater}}}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}/Service{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
The name {\textquotedblleft}pdpservice.invoker{\textquotedblright}
defines the PDP Service Invoker object.}

{\selectlanguage{english}\color{black}
The PDP Service Invoker is a client of PDP Service. The configuration
options include endpoint of service and credentials to be used for
establishing secure connection.}

\bigskip

\subsubsection{Delegation PDP configuration}
{\selectlanguage{english}\upshape\color{black}
Configuration XML schema with comments available at
\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/delegationsh/DelegationPDP.xsd}
.}

{\selectlanguage{english}\upshape\color{black}
Below is an example of configuration of Delegation PDP inside
{\textquotedblleft}echo{\textquotedblright} service.}

{\selectlanguage{spanish}\ttfamily\color{black}
{\textless}Service name={\textquotedbl}echo{\textquotedbl}
id={\textquotedbl}echo{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}SecHandler
name={\textquotedbl}arc.authz{\textquotedbl}
id={\textquotedbl}authz{\textquotedbl}
event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}PDP
name={\textquotedbl}delegation.pdp{\textquotedbl}/{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ {\textless}/SecHandler{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ {\textless}next
id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ }\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
{\textless}/echo:prefix{\textgreater}}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:suffix{\textgreater}
]{\textless}/echo:suffix{\textgreater}}}}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}/Service{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
For Delegation PDP, no specific configuration is needed. We only need to
switch it on by adding {\textless}PDP
name={\textquotedbl}delegation.pdp{\textquotedbl}/{\textgreater} under
{\textless}SecHandler/{\textgreater}}

\subsubsection[Delegation SecHandler Configuration]{Delegation
SecHandler Configuration}
\label{bkm:RefSecdelegsechandleconfig}{\selectlanguage{english}\upshape\color{black}
Below is an example of configuration of Delegation SecHandler inside TLS
MCC component.}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Component
name={\textquotedbl}tls.service{\textquotedbl}
id={\textquotedbl}tls{\textquotedbl}{\textgreater} {\textless}next
id={\textquotedbl}http{\textquotedbl}/{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:KeyPath{\textgreater}./key.pem{\textless}/tls:KeyPath{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CertificatePath{\textgreater}./cert.pem{\textless}/tls:CertificatePath{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CACertificatePath{\textgreater}./ca.pem{\textless}/tls:CACertificatePath{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}-delegation.collector
must be inside tls MCC-{}-{\textgreater}}}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}SecHandler
name={\textquotedbl}delegation.collector{\textquotedbl} }

{\selectlanguage{english}\ttfamily\color{black}
id={\textquotedbl}delegation{\textquotedbl}
event={\textquotedbl}incoming{\textquotedbl}{\textgreater}{\textless}/SecHandler{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
Current implementation of Delegation\textbf{ }SecHandler must be
attached to TLS MCC. }

\subsubsection[UsernameToken SecHandler configuration]{UsernameToken
SecHandler configuration}
\label{bkm:RefSecuserametokenconfig}{\selectlanguage{english}\upshape\color{black}
Configuration XML schema with comments available at
\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTo}}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{kenSH.xsd}}
.}

{\selectlanguage{english}\upshape\color{black}
Below is an example of configuration of UsernameToken SecHandler inside
MCCSOAP component.}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}Component
name={\textquotedbl}soap.service{\textquotedbl}
id={\textquotedbl}soap{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}next
id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
name={\textquotedbl}usernametoken.handler{\textquotedbl}
id={\textquotedbl}usernametoken{\textquotedbl}
event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}Process{\textgreater}extract{\textless}/Process{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}PasswordSource{\textgreater}password.txt{\textless}/PasswordSource{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}/SecHandler{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}

{\selectlanguage{english}\color{black}
UsernameToken SecHandler must be configured under SOAP MCC.}

\subsubsection[X509Token SecHandler configuration]{X509Token SecHandler
configuration}
{\selectlanguage{english}\upshape\color{black}
Configuration XML schema with comments available at
\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/x509tokensh/X509TokenSH.xsd}}.}

{\selectlanguage{english}\upshape\color{black}
Below is an example of configuration of X509Token SecHandler inside
MCCSOAP component.}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}Component
name={\textquotedbl}soap.service{\textquotedbl}
id={\textquotedbl}soap{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}next
id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
name={\textquotedbl}x509token.handler{\textquotedbl}
id={\textquotedbl}x509token{\textquotedbl}
event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}

\bigskip

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}Process{\textgreater}extract{\textless}/Process{\textgreater}}}

\bigskip

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}CACertificatePath{\textgreater}ca.pem{\textless}/CACertificatePath{\textgreater}}}

\bigskip

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/SecHandler{\textgreater}}}

\bigskip

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}

{\selectlanguage{english}\color{black}
X509Token SecHandler must be configured under SOAP MCC.}

\bigskip

\subsection{User Manual}
{\selectlanguage{english}\color{black}
This section describes how to configure and use SecHandler and PDP
elements included in the ARC1 and provides few examples of the ARC
Policy documents. The target readers are those users who will use the
ARC1 middleware. Currently this section is very short on details. It is
going to be continuously extended. Especially taking user feedback into
account.}

\subsubsection[Authorization SecHandler and PDPs]{Authorization
SecHandler and PDPs}
{\selectlanguage{english}\color{black}
There is a specific Authorization SecHandler (arc.authz) which is
implemented for calling the Policy Decision Points (PDP) and serves as
their container. }

{\selectlanguage{english}\color{black}
Usually the Authorization SecHandler and included PDPs are used on the
service side of communication channel. Although it is also possible to
use them on the client side. All possibilities are achieved by
modifying the configuration file (hereafter mentioned as service.xml)
and possibly providing the authorization policy in a separate file.}

{\selectlanguage{english}\color{black}
Here the {\textquotedblleft}echo{\textquotedblright} test service is
used to explaining the usage, but the explanation applies to other
services as well.}

{\selectlanguage{english}\color{black}
The procedure for configuring Authorization SecHandler in service.xml is
following:}

\liststyleWWviiiNumxi
\begin{enumerate}
\item {\selectlanguage{english}\color{black}
Add the Authorization SecHandler as child element
{\textless}SecHandler/{\textgreater} of
{\textless}Service/{\textgreater} element. }
\end{enumerate}
{\selectlanguage{english}\color{black}
The {\textquotedblleft}name{\textquotedblright} and
{\textquotedblleft}event{\textquotedblright} attribute of
{\textless}SecHandler/{\textgreater} element are both important. The
{\textquotedblleft}name{\textquotedblright} attribute is used for
distinguishing betwenn loaded SecHandler objects. The
{\textquotedblleft}event{\textquotedblright} attribute defines for
which message authorization would be enforced. Usually and reasonably
it is done for {\textquotedblleft}incoming{\textquotedblright}
messages. But some services and other Message Chain components may
define other internal types of messages. For possible values please
refer to documentation of particular Service or MCC. In our particular
case {\textquotedblleft}echo{\textquotedblright} service only supports
{\textquotedblleft}incoming{\textquotedblright} messages for this
purpose.}

\liststyleWWviiiNumxi
\setcounter{saveenum}{\value{enumi}}
\begin{enumerate}
\setcounter{enumi}{\value{saveenum}}
\item {\selectlanguage{english}\color{black}
Add the PDP configuration as \ child element
{\textless}PDP/{\textgreater} under
{\textless}SecHandler/{\textgreater}. Currently there are four usable
PDPs distributed as part of the ARC1 middleware:}
\end{enumerate}
\liststyleWWviiiNumxii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
simplelist.pdp -- compares Subject of user{\textquotesingle}s X509
certificate to those stored in a file.}
\item {\selectlanguage{english}\color{black}
arc.pdp -- compares authorization related information parsed from
message at various processing steps \ to Policy document specified in
configuration of this PDP.}
\item {\selectlanguage{english}\color{black}
pdpservice.invoker - composes the ARC Request, puts request into SOAP
message, and invokes the remote PDP service to get the response SOAP
which includes authorization decision. The PDP service functionality is
similar to arc.pdp.}
\item {\selectlanguage{english}\color{black}
delegation.pdp -- compares authorization related information parsed from
message at various processing steps and Policy document embedded in
proxy certificate used by remote side.}
\end{itemize}
{\selectlanguage{english}\color{black}
Default behavior of Authorization SecHandler is to execute all
configured PDPs sequentially till eithe r one of them fails or all
produced positive results. This behavior may be modified by attribute
{\textquotedblleft}action{\textquotedblright} of
{\textless}PDP/{\textgreater} element.}

{\selectlanguage{english}\upshape\color{black}
The description of PDP configuration and ARC Policy example are
available in Section 7}

\subsubsection[Delegation SecHandler, Delegation PDP and Proxy
Certificate Generation]{Delegation SecHandler, Delegation PDP and Proxy
Certificate Generation}
{\selectlanguage{english}\upshape\color{black}
Delegation SecHandler and Delegation PDP in their current state provide
an infrastructure for limiting capabilities of delegated credentials.
Their collect and process policies attached to X509 Proxy Certificates
respectively. Hence to have delegation restriction working both must be
enabled in configuration of service. Configuration of Delegation
SecHandler is described in section 7.10. }

{\selectlanguage{english}\color{black}
The possible location for Delegation PDP is inside Authorization
SecHandler (arc.authz). Depending on how fine grained policy of
delegated credentials is supposed to be corresponding Authrorization
SecHandler may be attached to different MCCs or directly to Service
component. However, the precondition for using Delegation PDP is that
there must be Delegation SecHandler instantiated earlier in chain.}

{\selectlanguage{english}\color{black}
On the client side, command line utility
{\textquotedblleft}approxy{\textquotedblright} utility can be used to
generate Proxy Certificate with Delegation Policy embedded.}

{\selectlanguage{english}\color{black}
Normally approxy appears in \$ARC\_LOCATION/bin. The usage of approxy is
like:}

{\selectlanguage{english}\color{black}
approxy -P proxy.pem -C cert.pem -K key.pem -c constraints}

{\selectlanguage{english}\color{black}
By using argument {\textquotedbl}-c{\textquotedbl}, some constraints can
be specified for proxy certificate. Currently, the life time can be
limited by using {\textquotedbl}-c validityStart=...{\textquotedbl} and
{\textquotedbl}-c validityEnd=...{\textquotedbl}, {\textquotedbl}-c
validityStart=...{\textquotedbl} and {\textquotedbl}-c
validityPeriod=...{\textquotedbl}. Like for example }

{\selectlanguage{english}\color{black}
{}-c validityStart=2008-05-29T10:20:30Z}

{\selectlanguage{english}\color{black}
{}-c validityEnd=2008-06-29T10:20:30Z}

{\selectlanguage{english}\color{black}
The Delegation Policy can be specified by using {\textquotedbl}-c
proxyPolicyFile=...{\textquotedbl} or {\textquotedbl}-c
proxyPolicy=...{\textquotedbl}. Like }

{\selectlanguage{english}\color{black}
{}-c proxyPolicyFile=delegation\_policy.xml}

{\selectlanguage{english}\upshape\color{black}
The Delegation Policy is the same as the ARC Policy explained in section
7.7. Simple example below renders delegated credentials usable only for
contacting service attached to HTTP communication channel under path
/arex (line 6) \ and allows HTTP operation POST (line 9) on it. }

\liststyleWWviiiNumxv
\begin{enumerate}
\item {\selectlanguage{english}\ttfamily\color{black}
{\textless}?xml version={\textquotedbl}1.0{\textquotedbl}
encoding={\textquotedbl}UTF-8{\textquotedbl}?{\textgreater}}
\item {\selectlanguage{english}\color{black}
\texttt{{\textless}Policy
xmlns={\textquotedbl}}\url{http://www.nordugrid.org/schemas/policy-arc}\texttt{{\textquotedbl}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ PolicyId={\textquotedbl}sm-example:policy1{\textquotedbl}
CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ {\textless}Rule RuleId={\textquotedbl}rule1{\textquotedbl}
Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Resources{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Resource
Type={\textquotedbl}string{\textquotedbl}
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedbl}{\textgreater}/arex{\textless}/Resource{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}/Resources{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
Type={\textquotedbl}string{\textquotedbl}
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}{\textgreater}POST{\textless}/Action{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
\item {\selectlanguage{english}\ttfamily\color{black}
{\textless}/Policy{\textgreater}}
\end{enumerate}
{\selectlanguage{english}\color{black}
Another example of delegation policy is presented below. This policy
restricts usage of delegated credentials to SOAP operation
CreateActivity (line 5) of Basic Execution Service (BES) [4] namespace
(line 9). Such policy could be embedded into credentials delegated to
high level Brokering service performing Grid job submission to low
level BES on behalf of user.}

\liststyleWWviiiNumxvi
\begin{enumerate}
\item {\selectlanguage{english}\ttfamily\color{black}
{\textless}?xml version={\textquotedbl}1.0{\textquotedbl}
encoding={\textquotedbl}UTF-8{\textquotedbl}?{\textgreater}}
\item {\selectlanguage{english}\ttfamily\color{black}
{\textless}Policy
xmlns={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc{\textquotedbl}
PolicyId={\textquotedbl}sm-example:policy1{\textquotedbl}
CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ {\textless}Rule RuleId={\textquotedbl}rule1{\textquotedbl}
Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
Type={\textquotedbl}string{\textquotedbl}
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/operation{\textquotedbl}{\textgreater}CreateActivity{\textless}/Action{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Conditions{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ {\textless}Condition{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ {\textless}Attribute
Type={\textquotedbl}string{\textquotedbl}
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace{\textquotedbl}{\textgreater}http://schemas.ggf.org/bes/2006/08/bes-factory{\textless}/Attribute{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ {\textless}/Condition{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}/Conditions{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
\item {\selectlanguage{english}\ttfamily\color{black}
{\textless}/Policy{\textgreater}}
\end{enumerate}
\subsubsection{VOMS Proxy Certificate}
{\selectlanguage{english}\color{black}
The commonly used voms proxy certificate can be used in ARC1 for
authentication as normal proxy certificate, and attribute acquiring in
order to make authorization decision based on attributes.}

{\selectlanguage{english}\color{black}
1.How to create voms proxy certificate}

{\selectlanguage{english}\color{black}
Currently the proxy creation utility in ARC1---arcproxy can not be used
for creating voms proxy certificate because the GSI-based communication
requirement from voms server side has not been supported in ARC1 (ARC1
uses standard SSL/TLS communication). So the way to create voms proxy
certificate is still to use the
{\textquotedblleft}voms-proxy-init{\textquotedblright} utility.}

{\selectlanguage{english}\color{black}
2.How to use voms proxy certificate in ARC1}

{\selectlanguage{english}\color{black}
The attribute certificate (AC, which is created by voms server and then
embedded in voms proxy certificate \ as one of the certificate
extension by {\textquotedblleft}voms-proxy-init{\textquotedblright}
client) \ will be parsed by TLS handling plugin (called MCCTLS) of
ARC1, and saved in the message context in the format
{\textquotedblleft}grantor=knowarc://testvoms.knowarc.eu:50000/knowarc:role=guest{\textquotedblright}.}

{\selectlanguage{english}\color{black}
Service administrator needs to configure the trusted DN chain for
verifying the AC under tls MCC{\textquotesingle}s configurationwith the
following format. You should specify one node
({\textless}tls:VOMSCertTrustDNChain/{\textgreater}) for each vo which
you trust. The reason why using trusted DN chain is for service to
restrict which vo server is trusted. }

{\selectlanguage{english}\color{black}
In the following table, the first
{\textless}tls:VOMSCertTrustDNChain{\textgreater} in the following
table defines two items: the first one is the DN of voms server, the
second line is the DN of the corresponding CA. The second
{\textless}tls:VOMSCertTrustDNChain{\textgreater} defines regular
expression for matching, the DN of voms server and DN of the
corresponding CA should both match it. The third one defines the
external file which includes the
{\textless}tls:VOMSCertTrustDNChain{\textgreater}.}

{\selectlanguage{english}\color{black}
\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChain{\textgreater}}

\bigskip

{\selectlanguage{english}\color{black}
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDN{\textgreater}/O=Grid/O=NorduGrid/OU=fys.uio.no/CN=ABCDE{\textless}/tls:VOMSCertTrustDN{\textgreater}}

\bigskip

{\selectlanguage{english}\color{black}
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDN{\textgreater}/O=Grid/O=NorduGrid/CN=NorduGrid
Certification Authority{\textless}/tls:VOMSCertTrustDN{\textgreater}}

\bigskip

{\selectlanguage{english}\color{black}
\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/tls:VOMSCertTrustDNChain{\textgreater}}

\bigskip

{\selectlanguage{english}\color{black}
\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChain{\textgreater}}

\bigskip

{\selectlanguage{english}\color{black}
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustRegex{\textgreater}\^{}/O=Grid/O=NorduGrid{\textless}/tls:VOMSCertTrustRegex{\textgreater}}

\bigskip

{\selectlanguage{english}\color{black}
\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/tls:VOMSCertTrustDNChain{\textgreater}}

\bigskip

{\selectlanguage{english}\color{black}
\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChainsLocation{\textgreater}./pathto/external/file{\textless}/tls:VOMSCertTrustDNChainsLocation{\textgreater}}

\bigskip

{\selectlanguage{english}\color{black}
Service administrator can then use the attribute into access control
policy:}

\liststyleWWviiiNumxxiv
\begin{enumerate}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Subjects{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}Subject{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}grantor=knowarc://testvoms.knowarc.eu:50000/knowarc:role=guest{\textless}/Attribute{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}Attribute
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/ca{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=CA{\textless}/Attribute{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/identity{\textquotedbl}
Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=test{\textless}/Attribute{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Subject{\textgreater}}}
\item {\selectlanguage{english}\color{black}
\texttt{\ \ \ \ \ \ {\textless}/Subjects{\textgreater}}}
\end{enumerate}

\bigskip

\subsubsection[UsernameToken SecHandler]{UsernameToken SecHandler}
{\selectlanguage{english}\color{black}
The UsernameToken SecHandler is meant for processing - generating and
extracting - WS-Security [5] UsernameToken from SOAP header. Hence it
must be attached to SOAP MCC of service or/and client communication
channel. }

{\selectlanguage{english}\upshape\color{black}
On the service side, the functionality of extracting UsernameToken may
be configured as described in section 7.11. }

{\selectlanguage{english}\color{black}
On the client side, the UsernameToken SecHandler may be configured
either by using client specific methods (for example see
test\_clientinterface.cpp src/tests/echo directory of source tree) or
through generic client configuration file as shown in example below
(also can see share/doc/arc/echo\_client.xml.example under the
installation directory). This example will generate token with username
{\textquotedblleft}user{\textquotedblright} and password
{\textquotedblleft}pass{\textquotedblright} inside any SOAP message
sent by client tools of the ARC1.}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}ArcConfig{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ {\textless}Plugins
overlay={\textquotedbl}add{\textquotedbl}{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ {\textless}Chain{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ {\textless}Component
name={\textquotedbl}soap.client{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ {\textless}SecHandler
name={\textquotedbl}usernametoken.handler{\textquotedbl}
event={\textquotedbl}outgoing{\textquotedbl}
overlay={\textquotedbl}add{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}Process{\textgreater}generate{\textless}/Process{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}Username{\textgreater}user{\textless}/Username{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}Password{\textgreater}pass{\textless}/Password{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}PasswordEncoding{\textgreater}digest{\textless}/PasswordEncoding{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ {\textless}/SecHandler{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ {\textless}/Component{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ {\textless}/Chain{\textgreater}}}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}/ArcConfig{\textgreater}}

\subsubsection[X509Token SecHandler]{X509Token SecHandler}
{\selectlanguage{english}\color{black}
The X509Token SecHandler is meant for processing - generating and
extracting - WS-Security [5] X509Token from SOAP header. Hence it must
be attached to SOAP MCC of service or/and client communication channel.
}

{\selectlanguage{english}\color{black}
On the service side, the functionality of extracting X509Token may be
configured as described in section 7.12. }

{\selectlanguage{english}\color{black}
On the client side, the X509Token SecHandler may be configured either by
using client specific methods (for example see
test\_clientinterface.cpp src/tests/echo directory of source tree) or
through generic client configuration file as shown in example below.
This example will generate x509 token (by using the specified
certificate and key file) inside any SOAP message sent by client tools
of the ARC1.}

{\selectlanguage{english}\ttfamily\color{black}
{\textless}ArcConfig{\textgreater}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ {\textless}Plugins
overlay={\textquotedbl}add{\textquotedbl}{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ {\textless}Chain{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ {\textless}Component
name={\textquotedbl}soap.client{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ {\textless}SecHandler
name={\textquotedbl}x509token.handler{\textquotedbl}
event={\textquotedbl}outgoing{\textquotedbl}
overlay={\textquotedbl}add{\textquotedbl}{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}Process{\textgreater}generate{\textless}/Process{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}KeyPath{\textgreater}/path/to/key.pem{\textless}/KeyPath{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ \ {\textless}CertificatePath{\textgreater}/path/to/cert.pem{\textless}/CertificatePath{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ \ {\textless}/SecHandler{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ \ {\textless}/Component{\textgreater}}}

{\selectlanguage{english}\upshape\color{black}
\texttt{\ {\textless}/Chain{\textgreater}{\textless}/ArcConfig{\textgreater}}}

\clearpage{\centering\selectlanguage{english}\bfseries\scshape\color{black}
References
\par}

{\selectlanguage{english}\color{black}
\stepcounter{Bibitem}{\theBibitem}. RFC3820- Internet X.509 Public Key
Infrastracture (PKI) Proxy, \url{http://rfc.net/rfc3820.html}}

{\selectlanguage{english}\color{black}
\stepcounter{Bibitem}{\theBibitem}. D1.2-2 The ARC container (first
prototype),
\url{http://www.knowarc.eu/documents/Knowarc_D1.2-2_07.pdf}}

{\selectlanguage{english}\color{black}
\textstyleInternetlink{\textcolor{black}{\stepcounter{Bibitem}{\theBibitem}}}\textstyleInternetlink{\textcolor{black}{.}}\textstyleInternetlink{\foreignlanguage{english}{\textcolor{black}{Web
Service Security Username Token Profile 1.1.
}}}\url{http://www.oasis-open.org/committees/download.php/16782/wss-v1.1-spec-os-UsernameTokenProfile.pdf}}

{\selectlanguage{english}\color{black}
\stepcounter{Bibitem}{\theBibitem}. OGSA{\textregistered} Basic
Execution Service Version 1.0,
\ \url{http://www.ogf.org/documents/GFD.108.pdf} }

{\selectlanguage{english}\color{black}
\textstyleInternetlink{\textcolor{black}{\stepcounter{Bibitem}{\theBibitem}}}\textstyleInternetlink{\textcolor{black}{.
OASIS Web Services Security (WSS) TC.
}}\url{http://www.oasis-open.org/committees/tc_home.php?wg_abbrev=wss}}
\end{document}

### 2008-11-10 doc/a-rex/arex_tech_doc.tex
submiting → submitting

@@
 are processed out of limits.
-\item \textbf{Submiting} -- The job is being passed for execution to the
+\item \textbf{Submitting} -- The job is being passed for execution to the
 \emph{Local Resource Management System} (LRMS). The corresponding
@@
 state. Otherwise it moves to \textbf{Finishing}. It is possible to
-limit the aggregate number of jobs in \textbf{Submiting} and \textbf{Executing}
+limit the aggregate number of jobs in \textbf{Submitting} and \textbf{Executing}
 states.
@@
 &Prepared& PREPARING + PENDING& Stage-in process has finished\\ \cline{2-4} 
-&Submiting& SUBMIT& Communication with local batch system is in process\\ \cline{2-4} 
+&Submitting& SUBMIT& Communication with local batch system is in process\\ \cline{2-4} 
 &Queued& INLRMS&
@@
           <xsd:enumeration value=\char`\"{}Prepared\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Submiting\char`\"{}/>
+          <xsd:enumeration value=\char`\"{}Submitting\char`\"{}/>
           <xsd:enumeration value=\char`\"{}Executing\char`\"{}/>

### 2008-11-10 src/services/a-rex/grid-manager/jobs/job.h

@@ -11,15 +11,15 @@ class JobLocalDescription;
 class JobUser;
 
 typedef enum {
-  JOB_STATE_ACCEPTED  = 0,
-  JOB_STATE_PREPARING = 1,
-  JOB_STATE_SUBMITING = 2,
-  JOB_STATE_INLRMS    = 3,
-  JOB_STATE_FINISHING = 4,
-  JOB_STATE_FINISHED  = 5,
-  JOB_STATE_DELETED   = 6,
-  JOB_STATE_CANCELING = 7,
-  JOB_STATE_UNDEFINED = 8
+  JOB_STATE_ACCEPTED   = 0,
+  JOB_STATE_PREPARING  = 1,
+  JOB_STATE_SUBMITTING = 2,
+  JOB_STATE_INLRMS     = 3,
+  JOB_STATE_FINISHING  = 4,
+  JOB_STATE_FINISHED   = 5,
+  JOB_STATE_DELETED    = 6,
+  JOB_STATE_CANCELING  = 7,
+  JOB_STATE_UNDEFINED  = 8
 } job_state_t;
 
 typedef struct {
@@ -36,7 +36,7 @@ extern job_state_rec_t states_all[JOB_STATE_UNDEFINED+1];
 #define JOB_NUM_ACCEPTED (\
   jobs_num[JOB_STATE_ACCEPTED] + \
   jobs_num[JOB_STATE_PREPARING] + \
-  jobs_num[JOB_STATE_SUBMITING] + \
+  jobs_num[JOB_STATE_SUBMITTING] + \
   jobs_num[JOB_STATE_INLRMS] + \
   jobs_num[JOB_STATE_FINISHING] \
 )
@@ -57,7 +57,7 @@ extern job_state_rec_t states_all[JOB_STATE_UNDEFINED+1];
 
 /* states doing their job mostly on computing nodes */
 #define JOB_NUM_RUNNING (\
-  jobs_num[JOB_STATE_SUBMITING] + \
+  jobs_num[JOB_STATE_SUBMITTING] + \
   jobs_num[JOB_STATE_INLRMS] \
 )
 

### 2008-11-10 src/services/a-rex/grid-manager/jobs/states.cpp

@@ -194,7 +194,7 @@ bool JobsList::DestroyJob(JobsList::iterator &i,bool finished,bool active) {
   };
   logger.msg(Arc::INFO,"%s: This job may be still running - canceling",i->job_id);
   bool state_changed = false;
-  if(!state_submiting(i,state_changed,true)) {
+  if(!state_submitting(i,state_changed,true)) {
     logger.msg(Arc::WARNING,"%s: Cancelation failed (probably job finished) - cleaning anyway",i->job_id);
     job_clean_final(*i,*user);
     if(i->local) { delete i->local; }; i=jobs.erase(i);
@@ -246,7 +246,7 @@ bool JobsList::GetLocalDescription(const JobsList::iterator &i) {
   return true;
 }
 
-bool JobsList::state_submiting(const JobsList::iterator &i,bool &state_changed,bool cancel) {
+bool JobsList::state_submitting(const JobsList::iterator &i,bool &state_changed,bool cancel) {
   if(i->child == NULL) {
     /* no child was running yet, or recovering from fault */
     /* write grami file for globus-script-X-submit */
@@ -317,7 +317,7 @@ bool JobsList::state_submiting(const JobsList::iterator &i,bool &state_changed,b
     if(i->child->Result() != 0) { 
       if(!cancel) {
         logger.msg(Arc::ERROR,"%s: Job submission to LRMS failed.",i->job_id);
-        JobFailStateRemember(i,JOB_STATE_SUBMITING);
+        JobFailStateRemember(i,JOB_STATE_SUBMITTING);
       } else {
         logger.msg(Arc::ERROR,"%s: Failed to cancel running job.",i->job_id);
       };
@@ -332,7 +332,7 @@ bool JobsList::state_submiting(const JobsList::iterator &i,bool &state_changed,b
       if(local_id.length() == 0) {
         logger.msg(Arc::ERROR,"%s: Failed obtaining lrms id.",i->job_id);
         i->AddFailure("Failed extracting LRMS ID due to some internal error");
-        JobFailStateRemember(i,JOB_STATE_SUBMITING);
+        JobFailStateRemember(i,JOB_STATE_SUBMITTING);
         return false;
       };
       /* put id into local information file */
@@ -769,7 +769,7 @@ void JobsList::ActJobPreparing(JobsList::iterator &i,bool /*hard_job*/,
         if(i->job_pending || state_loading(i,state_changed,false)) {
[…]

### 2008-11-17 doc/developer/python/ARC1-Python-Developers-Guide.tex
Python Dev. Guide Draft

\documentclass{article}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%

%---CUSTOM FLOAT---
\usepackage{float}
\floatstyle{plain}
\newfloat{illustration}{htp}{loi}
\floatname{illustration}{Illustration}

%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
%\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://download.nordugrid.org/}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{fig/logo-knowarc.png}}
\end{tabular}

\hrulefill

%-------- Change this to Knowarc_D<w>.<d>-<n>_<yy>

{\raggedleft Knowarc\_D0.0-0\_00\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large ARC1 Python Web Services Developer's Guide }\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large Paper subtitle}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Tam\'as Kazinczy\footnote{kazy@niif.hu} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

\tableofcontents
\newpage
\listoffigures
\newpage

\section{Preface}

\subsection{Purpose of this document}

This document has two main objectives. On one hand it guides you through the
creation of a new Python-based Web Service in ARC1\footnote[1]{See
  \url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/README}
  and \url{http://www.knowarc.eu/}}. On the other hand it describes the steps
needed to test the newly-created ARC1 Service.

The document has the following structure:

\begin{itemize}

  \item{In the first section we provide a brief description about the
    intended audience and the service to be implemented.}

  \item{In the next section Service Design will be discussed.}

  \item{In section Implementing the Service we provide an overview of the
    Python implementation.}

  \item{In section Putting GreetingService to work we describe the steps
    needed to make the Python service work.}

  \item{In section Testing the Service we introduce a tool to test the
    service with and the steps of testing.}

\end{itemize}

\subsection{Intended audience}

ARC1 Python Web Services Developer's Guide is intended for developers, who:

\begin{itemize}

  \item{have some experience with Python;}
  \item{want to create Python-based services in ARC1;}
  \item{are familiar with XML\footnote{See
      \url{http://www.w3.org/XML}}, XSD\footnote{See
      \url{http://www.w3.org/XML/Schema}}, SOAP\footnote{See
      \url{http://www.w3.org/TR/soap/}} and WSDL\footnote{See
      \url{http://www.w3.org/TR/wsdl} and
      \url{http://www.w3.org/TR/wsdl20}}.}

\end{itemize}

\subsection{About the service to be implemented}

The service (GreetingService) to be implemented is quite
simple. Users send their names to the service in a message and,
as an answer, they get a
personal greeting that tells them the local time on the server. So
if the name of the user is Anonymous, the answer would be something
like: 'Hello, Anonymous! Local time is: [Mon Aug 11 14:17:53 2008]'.

\newpage

\section{Service Design}

For the sake of clarity and the simplicity of testing, we use a
top-down approach in the service design. First, we create an XML
Schema Definition (XSD), then use this XSD to construct the WSDL of
the service.

\subsection{Designing the XML Schema Definition (XSD)}

The GreetingService is quite simple: it has a single input and a single output.
As an input, it takes a full personal name; as an output, it sends a
greeting message in the following form: “Hello, XY! Local time is:
[SERVER-DATE-AND-TIME]”.

In designing the service first, we provide a targetNamespace for the schema. 
Let it be, for example \linebreak
\verb#http://example.com/GreetingService/types#. We also define a 
namespace prefix \verb#tns# for this namespace so that we can refer to 
the types defined in this schema.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<schema#\\
\verb#  targetNamespace="http://example.com/GreetingService/types"#\\
\verb#  xmlns="http://www.w3.org/2001/XMLSchema"#\\
\verb#  xmlns:xsd="http://www.w3.org/2001/XMLSchema"#\\
\verb#  xmlns:tns="http://example.com/GreetingService/types"#\\
\verb#  elementFormDefault="qualified">#\\
\hline
\end{tabular}
\end{center}
\caption{Definition of target namespace}
\end{illustration}

Next, we create simple types for the input (name) and output (greeting).

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<simpleType name="name">#\\
\verb#  <restriction base="xsd:string"/>#\\
\verb#</simpleType>#\\

\verb#<simpleType name="greeting">#\\
\verb#  <restriction base="xsd:string"/>#\\
\verb#</simpleType>#\\
\hline
\end{tabular}
\end{center}
\caption{Simple types for input (name) and output (greeting)}
\end{illustration}

GreetingService is of a request-response type, so we create types for
the request and response as well.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<complexType name="GreetingServiceRequestType">#\\ 
\verb#  <sequence>#\\ 
\verb#    <element name="yourName" type="tns:name" minOccurs="1" maxOccurs="1"/>#\\ 
\verb#  </sequence>#\\ 
\verb#</complexType>#\\
\hline
\end{tabular}
\end{center}
\caption{Complex type to represent request}
\end{illustration}

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<complexType name="GreetingServiceResponseType">#\\ 
\verb#  <sequence>#\\ 
\verb#    <element name="yourGreeting" type="tns:greeting" minOccurs="1" maxOccurs="1"/>#\\ 
\verb#  </sequence>#\\ 
\verb#</complexType>#\\
\hline
\end{tabular}
\end{center}
\caption{Complex type to represent response}
\end{illustration}

\pagebreak

Here we have used the previously created simple types (referred to as
\verb#tns:name# and \verb#tns:greeting#) to define the elements
\verb#yourName# and \verb#yourGreeting# of complex types that will
represent the request and response.

All we need to do now is to create the request and response elements
which are of complex types mentioned before.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<element name="GreetingServiceRequest" type="tns:GreetingServiceRequestType"/>#\\ 
\verb#<element name="GreetingServiceResponse" type="tns:GreetingServiceResponseType"/>#\\ 
\hline
\end{tabular}
\end{center}
\caption{Request and response element}
\end{illustration}

The complete XSD can be found in Appendix \ref{app:xsd} of this guide.

%\clearpage

\subsection{Designing the Service WSDL}

Construction of the service WSDL is relatively straightforward.

We use those elements defined in our XML schema so we have to
provide a namespace prefix for the namespace
\verb#http://example.com/GreetingService/types# to be able to refer to
them. We also define the target namespace
(\verb#http://example.com/GreetingService#) as well as a prefix for
our service definition.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<definitions#\\ 
\verb#  targetNamespace="http://example.com/GreetingService"#\\ 
\verb#  xmlns="http://schemas.xmlsoap.org/wsdl/"#\\ 
\verb#  xmlns:xsd="http://www.w3.org/2001/XMLSchema"#\\ 
\verb#  xmlns:soap="http://schemas.xmlsoap.org/wsdl/soap/"#\\ 
\verb#  xmlns:wsdl="http://schemas.xmlsoap.org/wsdl/"#\\ 
\verb#  xmlns:gs="http://example.com/GreetingService"#\\ 
\verb#  xmlns:gst="http://example.com/GreetingService/types">#\\
\hline
\end{tabular}
\end{center}
\caption{Namespaces}
\end{illustration}

\pagebreak

We also need to import our XML schema in the
types section. It is worth noting that (according to the WS-I
Basic Profile Version 1.1): ``To import XML Schema Definitions, a
DESCRIPTION MUST use the XML Schema \texttt{import} statement.''

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<wsdl:types>#\\ 
\verb#  <xsd:schema xmlns:xsd="http://www.w3.org/2001/XMLSchema">#\\
\verb#    <xsd:import namespace="http://example.com/GreetingService/types"#\\
\verb#                schemaLocation="GreetingService.xsd"/>#\\ 
\verb#  </xsd:schema>#\\ 
\verb#</wsdl:types>#\\
\hline
\end{tabular}
\end{center}
\caption{Types}
\end{illustration}

Then, we create the messages to be used by the service. These messages
will use the request and response elements that we created earlier in the
schema.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<wsdl:message name="GreetingServiceRequestMessage">#\\ 
\verb#  <wsdl:part name="payload" element="gst:GreetingServiceRequest"/>#\\ 
\verb#</wsdl:message>#\\

\verb#<wsdl:message name="GreetingServiceResponseMessage">#\\ 
\verb#  <wsdl:part name="payload" element="gst:GreetingServiceResponse"/>#\\ 
\verb#</wsdl:message>#\\
\hline
\end{tabular}
\end{center}
\caption{Definition of WSDL messages}
\end{illustration}

The next step is to create a portType. Here we define the operation (greet)
along with its input and output messages.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<wsdl:portType name="GreetingServicePT">#\\ 
\verb#  <wsdl:operation name="greet">#\\ 
\verb#    <wsdl:input name="GreetingServiceInput"#\\ 
\verb#                message="gs:GreetingServiceRequestMessage"/>#\\ 
\verb#    <wsdl:output name="GreetingServiceOutput"#\\ 
\verb#                 message="gs:GreetingServiceResponseMessage"/>#\\ 
\verb#  </wsdl:operation>#\\ 
\verb#</wsdl:portType>#\\ 
\hline
\end{tabular}
\end{center}
\caption{Definition of WSDL portType}
\end{illustration}

\pagebreak

Then we define the binding where we use the portType created previously. 
The key point here is that we use \texttt{document/literal}
style SOAP HTTP binding.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<wsdl:binding name="GreetingServiceBinding" type="gs:GreetingServicePT">#\\ 
\verb#  <soap:binding style="document" transport="http://schemas.xmlsoap.org/soap/http"/>#\\ 
\verb#  <wsdl:operation name="greet">#\\ 
\verb#    <soap:operation soapAction="greet"/>#\\ 
\verb#    <wsdl:input name="GreetingServiceInput">#\\ 
\verb#      <soap:body use="literal"/>#\\ 
\verb#    </wsdl:input>#\\ 
\verb#    <wsdl:output name="GreetingServiceOutput">#\\ 
\verb#      <soap:body use="literal"/>#\\
\verb#    </wsdl:output>#\\ 
\verb#  </wsdl:operation>#\\ 
\verb#</wsdl:binding>#\\
\hline
\end{tabular}
\end{center}
\caption{Definition of WSDL binding}
\end{illustration}

As the last step of the design, we define the service with a port that 
uses this binding. For the sake of simplicity, we should provide an 
endpoint address where our service will listen.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<wsdl:service name="GreetingService">#\\
\verb#  <wsdl:port name="GreetingServicePort" binding="gs:GreetingServiceBinding">#\\
\verb#    <soap:address location="http://arctest.ki.iif.hu:50000/GreetingService"/>#\\
\verb#  </wsdl:port>#\\
\verb#</wsdl:service>#\\
\hline
\end{tabular}
\end{center}
\caption{Definition of WSDL service}
\end{illustration}

The complete WSDL can be found in Appendix \ref{app:wsdl} of this guide.

\clearpage
\newpage

\section{Implementing the Service}

\subsection{Before jumping in}

Though configuration is discussed later in this document, some points
need prior explanation.

For demonstration purposes, fixed parts of the greeting of
GreetingService are stored in ARC HED's\footnote[1]{HED: Hosting
  Environment Daemon} XML configuration file. In the implementation we
will get these parts from the XML right after the instantiation of the
Python class (in the \verb#__init__()# method).

\subsection{Overview of the Python implementation}

As it was pointed out before, we get fixed parts of the greeting in the
\verb#__init__()# method. This also includes resolving the namespace
prefix of the service namespace \verb#http://example.com/GreetingService#.

In ARC1, Python-based services use a wrapper to access the core service
classes and their methods.
This means that upon starting ARC HED, the wrapper loads the Python based
services as modules. The nature of this wrapper indicates that every
such (Python based) service is required to provide a process()
method. In this method, we get the payload from the incoming SOAP
message, get the name of the consumer, get the local time, then construct
the greeting message, assemble the outgoing SOAP message and finally
return.

The source code including comments can be found in Appendix \ref{app:pysrc}
of this guide.

\newpage

\section{Putting GreetingService to work}

\subsection{Creating \texttt{\_\_init\_\_.py}}

Let us assume that our Python implementation (greetingservice.py)
resides in a directory called GreetingService\footnote[1]{
This directory is required to be under ARC1 home directory ($\$ARC1\_HOME$)}. 
Without \verb#__init__.py# in it, this is only a simple directory; it is
\verb#__init__.py# what makes GreetingService a Python package, so we
need to have one, even if it is an empty file.

\subsection{Creating \texttt{Makefile.am}}

We need \verb#Makefile.am# to generate the \verb#Makefile#. It is
rather simple in our example, consisting of only three lines:

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#pythondir= $(PYTHON_SITE_PACKAGES)/GreetingService#\\

\verb#python_DATA = __init__.py greetingservice.py#\\

\verb#EXTRA_DISTS = $(python_DATA)#\\
\hline
\end{tabular}
\end{center}
\caption{Contents of \texttt{Makefile.am}}
\end{illustration}

The first line tells where GreetingService will be installed. Here
\verb#$(PYTHON_SITE_PACKAGES)# refers to something like
\verb#/usr/local/lib/python2.5/site-packages/#.

Second line tells what the Python sources are (here you see the Python
source files separated by spaces) and the third one that we will use
these files in our service.

\subsection{Generating the \texttt{Makefile}}

Open a terminal, go to ARC1 home directory (\verb#$ARC1_HOME#), then
edit \verb#./configure.ac# (you may need root privileges to do this):

\begin{enumerate}

\item search for the text \verb#AC_CONFIG_FILES#;

\item put an extra path there for the \verb#GreetingService#
  \verb#Makefile# (this is required to be the relative path from
  \verb#$ARC1_HOME#, so if the \verb#GreetingService# directory was in
  \verb#$ARC1_HOME/src/services# then it should be
  \verb#src/services/GreetingService/Makefile#).

\end{enumerate}

After saving configure.ac run `\verb#./autogen.sh#` and
`\verb#./configure#` in this order as in case of ordinary ARC1 configuration (doing this needs root privileges).

\clearpage

\subsection{Configuring the ARC HED}

Edit ARC1's XML configuration file used when starting ARC HED.

We define the \verb#http://example.com/GreetingService# namespace in
the \verb#<ArcConfig># node (see below), as we also use this
configuration file to store fixed parts of the greeting of
\verb#GreetingService#.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<ArcConfig#\\ 
\verb#  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"#\\
\verb#  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"#\\
\verb#  xmlns:arex="http://www.nordugrid.org/schemas/a-rex/Config"#\\
\verb#  xmlns:gsrv="http://example.com/GreetingService">#\\
\hline
\end{tabular}
\end{center}
\caption{Definition of service namespace in ArcConfig node}
\end{illustration}

We have to create a new \verb#<Service># node in \verb#<Chain># to
represent our service.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<Service name="pythonservice" id="greetingservice">#\\ 
\verb#  <ClassName>GreetingService.greetingservice.GreetingClass</ClassName>#\\ 
\verb#  <gsrv:part1>Hello, </gsrv:part1>#\\ 
\verb#  <gsrv:part2>! Local time is: [</gsrv:part2>#\\ 
\verb#  <gsrv:part3>]</gsrv:part3>#\\ 
\verb#</Service>#\\
\hline
\end{tabular}
\end{center}
\caption{Definition of service in ARC HED}
\end{illustration}

As mentioned before, Python-based services are handled through a
wrapper. This wrapper is called \linebreak{}\verb#pythonservice#, that is where the
name attribute of \verb#<Service># node comes from.

ClassName consists of three parts, separated by periods. The first
part refers to the package \linebreak{}(\verb#GreetingService# here), the second
part refers to the Python module (\verb#greetingservice# here -- so
the Python source file is \verb#greetingservice.py#) while the third
one is the name of the Python Class \linebreak{}(\verb#GreetingClass#).

The next three nodes (\verb#<gsrv:part1>#, \verb#<gsrv:part2># and
\verb#<gsrv:part3>#) are the fixed parts of the greeting.

All we have to do now is to check the \verb#<Component># nodes (so we
have TCP, HTTP, SOAP and POST present in the HTTP component), then
create a new node for \verb#GreetingService# in \verb#<Plexer>#.

\begin{illustration}
\begin{center}
\begin{tabular}{|l|}
\hline
\verb#<Plexer name="plexer.service" id="plexer">#\\
\verb#  <next id="a-rex">^/arex$</next>#\\
\verb#  <next id="count">^/count$</next>#\\
\verb#  <next id="greetingservice">^/GreetingService$</next>#\\
\verb#</Plexer>#\\
\hline
\end{tabular}
\end{center}
\caption{Configuration of Plexer}
\end{illustration}

The new node tells that the Service, whose ID is
\verb#greetingservice#, will be made accessible at the relative URL
\verb#/GreetingService#.

\subsection{Installation}

Installation of the service is also straightforward. Open a terminal, go to
the directory of the service (for example
\verb#$ARC1_HOME/src/services/GreetingService#) and execute the
command: `\verb#make install#` (doing this may need root privileges).
After the execution of the command ARC HED needs to be restarted.

\newpage

\section{Testing the Service}

The top-down design approach helps us to test
\verb#GreetingService#. There are several Open Source tools that can help
to test the service without having to write a client for it.

Basically, when considering which tool to use, we look for the
following features:

\begin{itemize}
  \item{capability of sending SOAP messages and receiving answers;}
  \item{capability of checking XML schema compliance;}
  \item{capability of generating message from WSDL;}
  \item{possibility of using custom assertions;}
  \item{possibility of using parameters.}
\end{itemize}

Our choice for testing the service had been soapUI. It is a widely used open
source tool for Web Service Testing that has all the features (even
more) we need.

\subsection{Creating the Test Project}
\label{sec:createtp}

After starting soapUI, we can create a new WSDL project by pressing
Ctrl+N or selecting \textit{New WSDL Project} from the \textit{File}
menu. [See \emph{Figure \ref{fig:newwsdl}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img2_resize.jpg}
\caption{New WSDL Project}
\label{fig:newwsdl}
\end{center}
\end{figure}

Fill in the \textit{Project Name} (it could be any of your choice) and
browse for the WSDL that we created before. You should leave
\textit{Create Requests} checked. This means that requests are
automatically generated from the WSDL. You may also check the
\textit{Create Project File} checkbox as it will allow you to save
your project.

\subsection{Adding the Test Request to a Test Case}
\label{sec:addtr2tc}

Now you should see the project in the \textit{Navigator
  view}. Navigate to \textit{Request 1} in the tree and open it (press
Enter or select \textit{Show Request Editor} from the context
menu). [See \emph{Figure \ref{fig:showreqed}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img3_resize.jpg}
\caption{Show Request Editor}
\label{fig:showreqed}
\end{center}
\end{figure}

Next we add this request to a Test Case. Click on the second icon from
the left in the Request Editor (tooltip says: Adds this request to a
TestCase). [See \emph{Figure \ref{fig:addreq2tc}}]

\begin{figure}
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img4_resize.jpg}
\caption{Adding Request to a Test Case 1}
\label{fig:addreq2tc}
\end{center}
\end{figure}

As we do not have a Test Suite yet, we need to create a Test Suite as
well. Choose a name for your Test Suite and Test Case (e.g.:
GreetingServiceTS and GreetingServiceTC1).

Now, a window appears. Here you can specify the name of the request
(e.g.: greetRequest1) and add predefined assertions. For example, you
may opt in the \textit{Add Schema Assertion} checkbox to validate
messages against our XML schema. [See \emph{Figure \ref{fig:addreq2tc2}}]

\begin{figure}
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img5.jpg}
\caption{Adding Request to a Test Case 2}
\label{fig:addreq2tc2}
\end{center}
\end{figure}

You should leave \textit{Show TestCase Editor} checked as we will
edit the Test Case in the following step.

\subsection{Editing the Test Case}
\label{sec:edittc}

After the Test Case Editor appears, click on greetRequest and select
\textit{Insert Step} $\to$ \textit{Properties} in the context menu
(tooltip says: Defines / Loads global TestCase properties). [See 
\emph{Figure \ref{fig:insprop}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img6_resize.jpg}
\caption{Insert Properties}
\label{fig:insprop}
\end{center}
\end{figure}

Fill in the name of the step (e.g.: greetProperties). The properties
window shows up. It has a table with two columns (Name and
Value). Here you can add a property (e.g.: inputName) by clicking on
the first icon from the left (tooltip says: Adds a property to the
property list) that could be used in the request and in assertions as
well. [See \emph{Figure \ref{fig:addprop1}} and \emph{Figure \ref{fig:addprop2}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img7_resize.jpg}
\caption{Adding a Property 1}
\label{fig:addprop1}
\end{center}
\end{figure}

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img8.jpg}
\caption{Adding a Property 2}
\label{fig:addprop2}
\end{center}
\end{figure}

Close the properties window after adding the property (and providing a
value for it). [See \emph{Figure \ref{fig:addprop3}}]

\begin{figure}
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img9_resize.jpg}
\caption{Adding a Property 3}
\label{fig:addprop3}
\end{center}
\end{figure}

Now double click on greetRequest1; this opens the Request Editor. In
\verb#<yourName>#, write \verb#${inputName}# instead of the default value
(a question mark).

\subsection{Assertions}
\label{sec:assertions}

Still in the Request Editor, click on the second icon from the left
(tooltip says: Adds an assertion to this item). [See \emph{Figure \ref{fig:assert1}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img10_resize.jpg}
\caption{Adding an Assertion 1}
\label{fig:assert1}
\end{center}
\end{figure}

Next, you have to choose the type of assertion. For now, this should
be \textit{Contains}. [See \emph{Figure \ref{fig:assert2}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img11.jpg}
\caption{Adding an Assertion 2}
\label{fig:assert2}
\end{center}
\end{figure}

In the next step, \textit{Content} should be
``\verb#Hello, ${inputName}!#'', so if we were to provide
``Anonymous'' as the value of \verb#inputName#, we could require the
answer to contain ``Hello, Anonymous!''. [See \emph{Figure \ref{fig:assert3}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img12.jpg}
\caption{Adding an Assertion 3}
\label{fig:assert3}
\end{center}
\end{figure}

\subsection{Running the Test Case}
\label{sec:runtc}

Close the Request Editor. In the Test Case Editor, click on the first
icon from the left (tooltip says: Runs this testcase). [See 
\emph{Figure \ref{fig:runtc}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img13_resize.jpg}
\caption{Running the Test Case}
\label{fig:runtc}
\end{center}
\end{figure}

\subsection{Evaluating Results}
\label{sec:evalresults}

If nothing went wrong, you would see the text FINISHED on a green
stripe. Details of the results are shown in the \textit{TestCase Log}.
[See \emph{Figure \ref{fig:tcresult}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img14_resize.jpg}
\caption{Test Case Results}
\label{fig:tcresult}
\end{center}
\end{figure}

Double clicking greetRequest1 allows you see the request and response
message and the results of assertions as well. [See \emph{Figure \ref{fig:respmsg}} 
and \emph{Figure \ref{fig:evalres}}]

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img15_resize.jpg}
\caption{Response Message}
\label{fig:respmsg}
\end{center}
\end{figure}

\begin{figure}[!hbp]
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img16_resize.jpg}
\caption{Assertions}
\label{fig:evalres}
\end{center}
\end{figure}

Although the value of the inputName property is not shown in the request
message, if you open the \textit{http log}, you will see that the
correct value has been used while creating the outgoing
request. [See \emph{Figure \ref{fig:httplog}}]

\begin{figure}
\begin{center}
\includegraphics{fig/ARC1PythonDGDraft-img17_resize.jpg}
\caption{HTTP Log}
\label{fig:httplog}
\end{center}
\end{figure}

\clearpage
\newpage

\appendix

\section{Appendices}

\subsection{GreetingService.xsd}
\label{app:xsd}

\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?> 
<schema 
  targetNamespace="http://example.com/GreetingService/types"
  xmlns="http://www.w3.org/2001/XMLSchema"
  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
  xmlns:tns="http://example.com/GreetingService/types"
  elementFormDefault="qualified">

  <simpleType name="name">
    <restriction base="xsd:string"/>
  </simpleType>

  <simpleType name="greeting">
    <restriction base="xsd:string"/>
  </simpleType>

  <complexType name="GreetingServiceRequestType">
    <sequence>
        <element name="yourName" type="tns:name" 
                 minOccurs="1" maxOccurs="1"/>
    </sequence>
  </complexType>

  <complexType name="GreetingServiceResponseType">
    <sequence>
      <element name="yourGreeting" type="tns:greeting" 
               minOccurs="1" maxOccurs="1"/>
    </sequence>
  </complexType>

  <element name="GreetingServiceRequest" 
           type="tns:GreetingServiceRequestType"/>
  <element name="GreetingServiceResponse" 
           type="tns:GreetingServiceResponseType"/>
</schema>
\end{verbatim}

\subsection{GreetingService.wsdl}
\label{app:wsdl}

\begin{verbatim}
<?xml version="1.0"
 encoding="UTF-8"?>
<definitions
  targetNamespace="http://example.com/GreetingService"
  xmlns="http://schemas.xmlsoap.org/wsdl/"
  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
  xmlns:soap="http://schemas.xmlsoap.org/wsdl/soap/"
  xmlns:wsdl="http://schemas.xmlsoap.org/wsdl/"
  xmlns:gs="http://example.com/GreetingService"
  xmlns:gst="http://example.com/GreetingService/types">

  <wsdl:types> 
    <xsd:schema xmlns:xsd="http://www.w3.org/2001/XMLSchema"> 
      <xsd:import namespace="http://example.com/GreetingService/types" 
                  schemaLocation="GreetingService.xsd"/> 
    </xsd:schema> 
  </wsdl:types>

  <wsdl:message name="GreetingServiceRequestMessage"> 
    <wsdl:part name="payload" element="gst:GreetingServiceRequest"/> 
  </wsdl:message>

  <wsdl:message name="GreetingServiceResponseMessage"> 
    <wsdl:part name="payload" element="gst:GreetingServiceResponse"/> 
  </wsdl:message>

  <wsdl:portType name="GreetingServicePT"> 
    <wsdl:operation name="greet"> 
      <wsdl:input name="GreetingServiceInput"
                  message="gs:GreetingServiceRequestMessage"/> 
      <wsdl:output name="GreetingServiceOutput" 
                   message="gs:GreetingServiceResponseMessage"/> 
    </wsdl:operation> 
  </wsdl:portType>

  <wsdl:binding name="GreetingServiceBinding" type="gs:GreetingServicePT"> 
    <soap:binding style="document" 
                  transport="http://schemas.xmlsoap.org/soap/http"/> 
    <wsdl:operation name="greet"> 
      <soap:operation soapAction="greet"/> 
      <wsdl:input name="GreetingServiceInput"> 
        <soap:body use="literal"/> 
      </wsdl:input> 
      <wsdl:output name="GreetingServiceOutput"> 
        <soap:body use="literal"/> 
      </wsdl:output> 
    </wsdl:operation> 
  </wsdl:binding>

  <wsdl:service name="GreetingService"> 
    <wsdl:port name="GreetingServicePort" 
               binding="gs:GreetingServiceBinding"> 
      <soap:address location="http://arctest.ki.iif.hu:50000/GreetingService"/> 
    </wsdl:port> 
  </wsdl:service>

</definitions>
\end{verbatim}

\subsection{greetingservice.py}
\label{app:pysrc}

\begin{verbatim}
import arc
import time

class GreetingClass:

    def __init__(self, cfg):
        # we want to use members of the 'http://example.com/GreetingService' namespace
        # so we need to know which namespace prefix was assigned to it
        # we get the namespace prefix into 'self.gs_ns'
        self.gs_ns = cfg.NamespacePrefix('http://example.com/GreetingService')

        # get part1 of greeting from the config XML
        self.part1 = str(cfg.Get(self.gs_ns + ':part1'))

        # get part2 of greeting from the config XML
        self.part2 = str(cfg.Get(self.gs_ns + ':part2'))

        # get part3 of greeting from the config XML
        self.part3 = str(cfg.Get(self.gs_ns + ':part3'))

    def process(self, inmsg, outmsg):

        # get the payload from the message
        inpayload = inmsg.Payload()

        # we want to use members of the 'http://example.com/GreetingService/types'
        # namespace so we need to know which namespace prefix was assigned to it
        # we get the namespace prefix into a variable named 'gst_ns'
        gst_ns = inpayload.NamespacePrefix('http://example.com/GreetingService/types')

        # then we can get 'gst_ns:GreetingServiceRequest/gst_ns:yourName'
        GreetingServiceRequest = inpayload.Get(gst_ns + ':GreetingServiceRequest')

        name = str(GreetingServiceRequest.Get(gst_ns + ':yourName'))

        localtime = time.localtime()

        strlocaltime = time.asctime(localtime)

        # forge the greeting message
        greet = self.part1 + name + self.part2 + strlocaltime + self.part3

        # create an answer payload
        outpayload = arc.PayloadSOAP(\
                        arc.NS({'gst':'http://example.com/GreetingService/types'}))

        # here we defined that 'gst' prefix will be the namespace prefix of
        # 'http://example.com/GreetingService/types' and
        # we create a node at '/gst:GreetingServiceResponse/gst:yourGreeting' 
        # and put the string in it
        outpayload.NewChild('gst:GreetingServiceResponse').\
                        NewChild('gst:yourGreeting').Set(greet)

        # put the payload into the outgoing message
        outmsg.Payload(outpayload)

        # return with STATUS_OK
        return arc.MCC_Status(arc.STATUS_OK)
\end{verbatim}

\end{document}

### 2008-11-25 doc/sec/ManualofSecurityFramework.tex
Put the user manual away from security technical documentation

--- 2008-10-25 doc/sec/SecurityFrameworkofARC1.tex
@@
 \usepackage[T1]{fontenc}
-\usepackage[spanish,english]{babel}
+\usepackage[english]{babel}
 \usepackage{amsmath,amssymb,amsfonts,textcomp}
@@
 \usepackage{hyperref}
-\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=name, pdfsubject=, pdfkeywords=}
+\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=qiangwz, pdfsubject=, pdfkeywords=}
 \usepackage[pdftex]{graphicx}
@@
 % Outline numbering
-\setcounter{secnumdepth}{4}
+\setcounter{secnumdepth}{3}
 \renewcommand\thesection{\arabic{section}.}
@@
 \renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection}.}
-\renewcommand\theparagraph{\arabic{subsection}.\arabic{subsubsection}.\arabic{paragraph}.}
 \makeatletter
@@
 \newcounter{saveenum}
-\newcommand\liststyleWWviiiNumii{%
-\renewcommand\labelitemi{[F0B7?]}
-\renewcommand\labelitemii{[F0B7?]}
-\renewcommand\labelitemiii{[F0B7?]}
-\renewcommand\labelitemiv{[F0B7?]}
-}
-\newcommand\liststyleWWviiiNumxiii{%
-\renewcommand\labelitemi{[F0B7?]}
-\renewcommand\labelitemii{[F0B7?]}
-\renewcommand\labelitemiii{[F0B7?]}
-\renewcommand\labelitemiv{[F0B7?]}
-}
-\newcommand\liststyleWWviiiNumxxi{%
-\renewcommand\labelitemi{[F0B7?]}
-\renewcommand\labelitemii{o}
-\renewcommand\labelitemiii{[F0A7?]}
-\renewcommand\labelitemiv{[F0B7?]}
-}
-\newcommand\liststyleWWviiiNumxvii{%
-\renewcommand\labelitemi{o}
-\renewcommand\labelitemii{o}
-\renewcommand\labelitemiii{[F0A7?]}
-\renewcommand\labelitemiv{[F0B7?]}
-}
-\newcommand\liststyleWWviiiNumxix{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\labelitemi{o}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-}
-\newcommand\liststyleWWviiiNumxviii{%
-\renewcommand\labelitemi{o}
-\renewcommand\labelitemii{o}
-\renewcommand\labelitemiii{[F0A7?]}
-\renewcommand\labelitemiv{[F0B7?]}
-}
-\newcommand\liststyleWWviiiNumxx{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\labelitemi{o}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-}
-\newcommand\liststyleWWviiiNumxxii{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\labelitemi{o}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-}
-\newcommand\liststyleWWviiiNumxxiii{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\labelitemi{{}-}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-}
-\newcommand\liststyleWWviiiNumiii{%
-\renewcommand\theenumi{\alph{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi)}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
-\newcommand\liststyleWWviiiNumiv{%
-\renewcommand\theenumi{\alph{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi)}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
-\newcommand\liststyleWWviiiNumv{%
-\renewcommand\theenumi{\alph{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi)}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
-\newcommand\liststyleWWviiiNumvi{%
-\renewcommand\theenumi{\alph{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi)}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
-\newcommand\liststyleWWviiiNumvii{%
-\renewcommand\theenumi{\alph{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi)}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
-\newcommand\liststyleWWviiiNumviii{%
-\renewcommand\labelitemi{[F0B7?]}
-\renewcommand\labelitemii{[F081?]}
-\renewcommand\labelitemiii{${\blacksquare}$}
-\renewcommand\labelitemiv{[F06C?]}
-}
-\newcommand\liststyleWWviiiNumix{%
-\renewcommand\labelitemi{[F0B7?]}
-\renewcommand\labelitemii{{}-}
-\renewcommand\labelitemiii{${\blacksquare}$}
-\renewcommand\labelitemiv{[F06C?]}
-}
-\newcommand\liststyleWWviiiNumx{%
-\renewcommand\labelitemi{[F0B7?]}
-\renewcommand\labelitemii{[F081?]}
-\renewcommand\labelitemiii{${\blacksquare}$}
-\renewcommand\labelitemiv{[F06C?]}
-}
-\newcommand\liststyleWWviiiNumxiv{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
 \newcommand\liststyleWWviiiNumxi{%
@@
 \makeatletter
-\newcommand\ps@Convertiv{
+\newcommand\ps@Convertv{
   \renewcommand\@oddhead{}
@@
 \renewcommand\theBibitem{\arabic{Bibitem}}
-\newcounter{Figure}
-\renewcommand\theFigure{\arabic{Figure}}
 \title{}
@@
 \begin{center}
-\includegraphics[width=1.4228in,height=0.9098in]{SecurityFrameworkofARC1-img1.jpg}
+\includegraphics[width=1.422in,height=0.9091in]{ManualofSecurityFramework-img1.jpg}
 \end{center}
@@
 \multicolumn{3}{m{6.61436in}}{\raggedleft
-\selectlanguage{english}\color{black} NORDUGRID-TECH-16\newline
-6/8/08}\\
+\selectlanguage{english}\color{black} NORDUGRID-TECH-17\newline
+15/11/08}\\
 \multicolumn{3}{m{6.61436in}}{\centering
-{\selectlanguage{english}\bfseries\scshape\color{black} Security
-Framework of ARC1}\par
+{\selectlanguage{english}\bfseries\scshape\color{black} Manual of
+Security Framework}\par
 
@@
  &
-\selectlanguage{english}\color{black} This document is about security
-design concerns and ideas, as well as security framework implementation
-in the ARC1 middleware. &
+\selectlanguage{english}\color{black} This document is about user manual
+of security framework, as well as development manual and administration
+manual. &
 ~
@@
 \bigskip
-\clearpage\clearpage\setcounter{page}{1}\pagestyle{Convertiv}
-
-\bigskip
-
-\subsection{Introduction}
-{\selectlanguage{english}\color{black}
-The security framework of the ARC1 includes two parts of capabilities:
-security capability embedded in hosting environment, and security
-capability implemented as plug-ins with well-defined interfaces which
-can be accessed by hosting environment and applications. The following
-design concerns were employed when designing:}
-
-\liststyleWWviiiNumii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-Interoperability and standardization}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-In consistent with the main design concern of the ARC1, interoperability
-and standardization is considered in security framework. For example,
-in terms of authentication, PKI infrastructure and proxy certificate
-(RFC3820 [1]) is used as most of the other grid middle-wares do. Since
-supporting of standardization is a way for implementing
-interoperability, some standard specifications have been implemented as
-prototype and tested, such as SAML specification.}
-
-\liststyleWWviiiNumii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-Modularity and extensibility }
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-Besides the security functionality which is embedded in hosting
-environment, the other security functionality is implemented as
-plug-ins which has well-defined interfaces, and is configurable and
-dynamically loadable. Since the interoperation interface between
-security plug-in and hosting environment or applications is predefined,
-it is easy to extend the security functionality in order to support
-some other security capability by implementing the interface.}
-
-\liststyleWWviiiNumii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-Backward compatibility }
-\end{itemize}
-{\selectlanguage{english}\color{black}
-The GSI (Grid Security Infrastructure) based mechanism has been a
-de-facto solution for grid security. The design of security framework
-should be compatible to it.}
-
-\bigskip
-
-\subsection{Security architecture in HED. SecHandler and PDP}
-\subsubsection{Structure of SecHandler and PDP}
-{\selectlanguage{english}\upshape\color{black}
-In the implementation of the ARC1, there is a Service Container -- the
-Hosting Environment Daemon (HED) (D1.2-2, [2]) which provides a hosting
-place for various services in application level, as well as a flexible
-and efficient communication mechanism.}
-
-{\selectlanguage{english}\upshape\color{black}
-HED contains a framework for implementing and enforcing authentication
-and authorization. Each Message Chain Component (MCC) or service has a
-common interface for implementing various authentication and
-authorization functionality. This functionality is implemented by using
-pluggable components (plug-ins) called SecHandler. The SecHandler
-components are C++ classes and provide method for processing messages
-traveling through Message Chains of the HED. Each MCC or Service
-\ usually implement two queues of SecHandlers -- one for incoming
-messages and one for outgoing called
-{\textquotedblleft}incoming{\textquotedblright} and
-{\textquotedblleft}outgoing{\textquotedblright} respectively. It is
-possible for MCC or Service to implement other set of queues. Please
-check documentation of particular component for that particular
-information. All SecHandler components attached to the queue are
-executed sequentially. If any of them fails, message processing fails
-as well.}
-
-{\selectlanguage{english}\upshape\color{black}
-Each SecHandler is configured inside same configuration file used for
-configuring whole chain of MCCs. Some of implemented SecHandler
-components also make use of pluggable and configurable sub-modules
-which specifically handle various security functionalities, such as
-authorization, authentication, etc. The currently implemented
-sub-modules used by some SecHandlers are Policy Decision Point (PDP)
-components such as Arc PDP which can process ARC specific Request and
-Policy documents. Figure 1 gives the structure of a MCC/Service, and
-the message sequence inside it. And Figure 2 shows the configuration of
-SecHandler components for an example
-{\textquotedblleft}Echo{\textquotedblright} service.}
-
-\bigskip
-
-\includegraphics[width=5.7709in,height=4.5209in]{SecurityFrameworkofARC1-img2.jpg}
-
-{\selectlanguage{english}\itshape\color{black}
-Figure \stepcounter{Figure}{\theFigure}. There are usually two chains of
-SecHandlers inside the MCC or service. Each SecHandler will parse the
-Security Attributes which are generated by the upstream MCC/services or
-probably upstream SecHandlers in the same or other MCC/Service, and do
-message processing or authenticate or authorize the incoming/outgoing
-message based on the collected information. The SecHandler can also
-change the payload and attributes of Messsage itself. For example, the
-Username-Token SecHandler will insert the WSS Username Token [3] into
-header part of SOAP message. The PDPs are called by the SecHandlers and
-are supposed to make authorization decision. Here the two local PDP and
-one remote PDP service is just for demonstration, and one or any number
-of PDPs can be configured under each SecHandler.}
-
-\bigskip
-
-[Warning: Draw object ignored]
-
-{\centering\selectlanguage{english}\itshape\color{black}
-Figure \stepcounter{Figure}{\theFigure}. Example Echo service is
-configured to use two SecHandlers, both responsible for authorization.
-First SecHandler uses the identity of client extracted from the
-incoming message to map it into local identity like local Linux
-username. In this case all clients are mapped to local account
-{\textquotedblleft}test{\textquotedblright}. The second one uses two
-PDPs: one will compose ARC specific authorization request based on the
-Security Attributes collected from
-{\textquotedblleft}incoming{\textquotedblright} message and evaluate it
-against the ARC specific authorization policy which is defined in local
-file {\textquotedblleft}policy.xml{\textquotedblright}; the other will
-compare the X509 identity of client extracted from the incoming message
-against list of identities stored locally.
-\par}
-
-\subsubsection{Interface of SecHandler}
-{\selectlanguage{english}\upshape\color{black}
-When one component (MCC or service) is loaded according to the
-configuration information, the SecHandler under the component and the
-plug-ins like PDP which are attached to the SecHandler will be loaded
-as well. }
-
-{\selectlanguage{english}\upshape\color{black}
-There is one simple interface (see Figure 3) defined in class
-SecHandler, which will be called by the containing MCC/Service once
-there is message (incoming or outgoing) need to be processed.}
-
-{\centering\selectlanguage{english}\itshape\color{black}
-Figure \stepcounter{Figure}{\theFigure}. class SecHandler is an abstract
-class which includes a general interface called Handle which uses
-Message object as argument. Any security handler implementation should
-inherit class SecHandler and implement the interface according to the
-actual functionality. The interface only return simple Boolean value,
-and any useful information generated during the calling of this
-interface should be put into the security attribute of the message, or
-put into the payload itself. 
-\par}
-
-\begin{center}
-\begin{minipage}{3.3756in}
-{\selectlanguage{english}\color{black}
-\ \ class SecHandler \{}
-
-{\selectlanguage{english}\upshape\color{black}
-\ \ \ public:}
-
-{\selectlanguage{english}\upshape\color{black}
-\ \ \ \ SecHandler(Arc::Config*) \{\};}
-
-{\selectlanguage{english}\upshape\color{black}
-\ \ \ \ virtual \~{}SecHandler() \{\};}
-
-{\selectlanguage{english}\upshape\color{black}
-\ \ \ \ virtual bool \textit{Handle}(Arc::Message *msg) = 0;}
-
-{\selectlanguage{english}\color{black}
-\ \ \};}
-\end{minipage}
-\end{center}
-{\selectlanguage{english}\upshape\color{black}
-Currently, the ARC1 comes with the following four security handler
-implemented: }
-
-\liststyleWWviiiNumxiii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textit{arc.authz} -- Authorization SecHandler}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-The \textit{arc.authz} is responsible for calling the interface of
-policy decision point and getting back the authorization result, and
-then making decision according to this authorization result. There is
-one simple interface (see Figure 4) defined in PDP, which will be
-called by \textit{arc.authz} if configured inside once there is message
-(incoming or outgoing) need to be processed.}
-
-\liststyleWWviiiNumxiii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textit{identity.map} -- Identity Mapping SecHandler}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-The \textit{identity.map} is a specific authorization oriented security
-handler. It will map the global identity in the message into local
-identity like system username based on the result returned by Policy
-Decision Point components.}
-
-\liststyleWWviiiNumxiii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textit{delegation.collector} -- Delegation SecHandler}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-The \textit{delegation.collector} is responsible for collecting the
-delegation policy information from the remote proxy credential (proxy
-certificate is compatible to RFC3820) inside the message, and putting
-this policy into message{\textquoteright}s security attribute for the
-usage of other components, such as \ \textit{delegation.pdp}.}
-
-\liststyleWWviiiNumxiii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textit{usernametoken.handler} -- UseranemToken SecHandler}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-The task of the \ \textit{usernametoken.handler} is to generate the
-WS-Security Username-Token and add it into header of SOAP message which
-is the payload of outgoing message. It can also extract the WS-Security
-Username-Token from the header of SOAP message which is the payload of
-incoming message.}
-
-\subsubsection[Interface of PDP]{Interface of PDP}
-\label{bkm:Ref204009855}{\selectlanguage{english}\upshape\color{black}
-Figure 4 shows the definition of abstract class PDP. The implementation
-could be some function which implements the interface by composing the
-policy evaluation request, evaluating this request against some policy,
-and returning the evaluation result, or just by composing the policy
-evaluation request, invoking some remote policy decision web service
-and getting back the evaluation result.}
-
-{\centering\selectlanguage{english}\itshape\color{black}
-Figure \stepcounter{Figure}{\theFigure}. class PDP is an abstract class
-which includes a general interface called isPermitted which uses
-Message object as argument. Any policy decision point implementation
-should inherit class PDP and implement the interface according to the
-actual functionality. The interface only return simple Boolean value,
-and any useful information generated during the calling of this
-interface should be put into the security attribute of the message, or
-put into the payload itself.
-\par}
-
-\begin{center}
-\begin{minipage}{3.8756in}
-{\selectlanguage{english}\color{black}
-\ \ class PDP \{}
-
-{\selectlanguage{english}\upshape\color{black}
-\ \ \ public:}
-
-{\selectlanguage{english}\upshape\color{black}
-\ \ \ \ PDP(Arc::Config* cfg) \{ \};}
-
-{\selectlanguage{english}\upshape\color{black}
-\ \ \ \ virtual \~{}PDP() \{\};}
-
-{\selectlanguage{english}\upshape\color{black}
-\ \ \ \ virtual bool \textit{isPermitted}(Arc::Message *msg) = 0;}
-
-{\selectlanguage{english}\color{black}
-\ \ \};}
-\end{minipage}
-\end{center}
-{\selectlanguage{english}\upshape\color{black}
-Currently, the ARC1 comes with the following four policy decision point
-implementation: }
-
-\liststyleWWviiiNumxiii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textit{arc.pdp }{}-- Arc PDP}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-The Arc PDP will organize the security attributes into the ARC specific
-authorization request, call the policy evaluator to evaluate the
-request against the policy (which is in ARC specific format)
-repository, and get back the evaluation result. See paragraph 3 for
-detail information about request schema and policy schema.}
-
-\liststyleWWviiiNumxiii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textit{delegtion.pdp} -- Delegation PDP}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-The Delegation PDP is basically similar to Arc PDP, except it uses the
-delegation policy parsed from remote proxy credential by
-\textit{delegation.collector}, and evaluates the request against
-delegation policy. See section 6for the design idea and use case of
-delegation policy in fine-grained identity delegation.}
-
-\liststyleWWviiiNumxiii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textit{simplelist.pdp} -- Simplelist PDP}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-The Simplelist PDP is a simplest implementation of policy decision
-point. It will match the identity extracted from the remote credential
-(or proxy credential) with local list of permitted identities.}
-
-\liststyleWWviiiNumxiii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textit{pdpservice.invoker} -- PDP Service Invoker}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-The PDP Service Invoker is a client which can be used to invoke the PDP
-Service which implements the same functionality as Arc PDP, except that
-the evaluation request and response are carried by SOAP message. The
-benefit of implementing PDP Service and PDP Service Invoker is that the
-policy evaluation engine can be accessed remotely and maintained
-centrally.}
-
-\subsection[Policy Evaluation Engine]{Policy Evaluation Engine}
-\subsubsection{Design of policy evaluation engine}
-{\selectlanguage{english}\upshape\color{black}
-The ARC1 defines specific evaluation request and policy schema. Based on
-the schema definition, one policy evaluation engine is implemented. The
-design principal of policy evaluation engine is generality by which the
-implementation of the policy evaluation engine can be easily extended
-to adopt some other policy schema, such as XACML policy schema.}
-
-{\selectlanguage{english}\upshape\color{black}
-Figure 5 shows the UML class diagram about the policy evaluation engine.
-It shows all classes and relations simultaneously for getting the
-overall picture. }
-
-{\selectlanguage{english}\upshape\color{black}
-The \textit{Evaluator} class is the key class for policy evaluation. It
-accepts request evaluates it against loaded policy and returns
-evaluation response. }
-
-{\selectlanguage{english}\upshape\color{black}
-Three abstract factories - \textit{FnFactory}, \textit{AlgFactory},
-\textit{AttributeFactory} - are responsible for creating the
-\textit{Function}, \textit{CombiningAlg} and \textit{AttributeValue}
-objects correspondingly. The classes inherited from
-\textit{CombiningAlg} class take care of implementing various combining
-algorithms which define relations between
-{\textless}Rule/{\textgreater} elements in policy. The
-\textit{AttributeValue} type of classes are used for processing
-different types of {\textless}Attribute/{\textgreater} and similar
-elements. The \textit{Function} classes take care of comparing
-{\textless}Attribute/{\textgreater} elements of request and policy. }
-
-{\selectlanguage{english}\upshape\color{black}
-The \textit{Policy} class parses {\textless}Policy/{\textgreater} or
-{\textless}Rule/{\textgreater} elements and creates
-\ \textit{CombingAlg} objects according to the
-{\textless}RuleCombiningAlg/{\textgreater} attribute of
-{\textless}Policy/{\textgreater}, \textit{Function} objects according
-to the {\textless}Function/{\textgreater} attribute of
-{\textless}Attribute/{\textgreater} and \textit{AttributeValue} objects
-according to the {\textless}Type/{\textgreater} attribute of
-{\textless}Attribute/{\textgreater}. Those objects will be used when
-evaluating the request.}
-
-{\selectlanguage{english}\upshape\color{black}
-The \textit{Request} class is responsible for parsing
-{\textless}Request/{\textgreater} element and creates corresponding
-\textit{AttributeValue} objects according to the
-{\textless}Type/{\textgreater} attribute of
-{\textless}Attribute/{\textgreater}. When evaluating, each
-\textit{AttributeValue} in request will be evaluated against
-corresponding \textit{AttributeValue} in the policy by using relevant
-\textit{Function}.}
-
-{\selectlanguage{english}\upshape\color{black}
-Due to extensible architecture of code it is relatively easy to add
-support for new types of \ \textit{AttributeValue}, \textit{Function
-and CombingAlg} objects in this way supporting various types of XML
-based policy languages.}
-
-\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img3.png}
-
-\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img4.png}
-
-\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img5.png}
-
-\includegraphics[width=5.7043in,height=5.6354in]{SecurityFrameworkofARC1-img6.png}
-
-{\centering\selectlanguage{english}\itshape\color{black}
-Figure \stepcounter{Figure}{\theFigure}. The UML class diagram of the
-classes inside policy evaluation engine. 
-\par}
-
-\subsubsection{Schemas for policy evaluation engine}
-{\selectlanguage{english}\upshape\color{black}
-The schema for ARC Policy is available at}
-
-{\selectlanguage{english}\upshape\color{black}
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd}
-.}
-
-{\selectlanguage{english}\color{black}
-The hierarchy tree of ARC Policy is show below (numbers show
-multiplicity of elements)}
-
-{\selectlanguage{english}\itshape\color{black}
-Policy (1)}
-
-{\selectlanguage{english}\itshape\color{black}
- Rule (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Subjects (1)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Subject (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-    Attribute (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Resources (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Resource (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Actions (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Action (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Conditions (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Condition (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-    Attribute (1-)}
-
-{\selectlanguage{english}\upshape\color{black}
-The schema for ARC Request is available at \ }
-
-{\selectlanguage{english}\upshape\color{black}
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd}
-.}
-
-{\selectlanguage{english}\upshape\color{black}
-The hierarchy tree of ARC Request is show below (numbers show
-multiplicity of elements)}
-
-{\selectlanguage{english}\itshape\color{black}
-Request (1)}
-
-{\selectlanguage{english}\itshape\color{black}
- RequestItem (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Subject (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-   SubjectAttribute (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Resource (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Action (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Context (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-   ContextAttribute (1-)}
-
-{\selectlanguage{english}\upshape\color{black}
-The schema for ARC Response is available at}
-
-{\selectlanguage{english}\upshape\color{black}
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Response.xsd}
-.}
-
-{\selectlanguage{english}\upshape\color{black}
-The ARC Response is not used directly in code. It is in use by PDP
-Service which provides remote evaluation of policies. }
-
-\subsubsection{Basic Elements of Policy}
-{\selectlanguage{english}\upshape\color{black}
-There are 2 basic objects - {\textquotedbl}policy{\textquotedbl} and
-{\textquotedbl}request{\textquotedbl}. There is 1 main actor -
-Evaluator. Curretly there are two types of elements in policy: Policy
-and Rule. Policy element is made of Rule elements.}
-
-{\color{black}
-Evaluator matches request to policy and produces one of 4 following
-results:}
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textbf{PERMIT} - policy explicitely permits activity specified in
-request because request matches some part of policy and corresponding
-effect specified in policy is PERMIT. }
-\end{itemize}
-{\bfseries\itshape\color{black}
-Example: }
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Rule: }}\textit{PERMIT
-person ALICE to PLAY in place called WONDERLAND}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Request: }}\textit{person
-ALICE wants to PLAY in place called WONDERLAND}}
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textbf{DENY} - policy explicitely denies activity specified in Request
-because Request matches some part of policy and corresponding effect
-specified in policy is DENY.}
-\end{itemize}
-{\bfseries\itshape\color{black}
-Example:}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Rule: DENY }}\textit{fruit
-APPLE to GROW on PEACH tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Request}}\textit{: fruit
-APPLE to be GROWN on PEACH tree}}
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textbf{INDETERMINATE} - request has some part which does not correspond
-to policy.}
-\end{itemize}
-{\bfseries\itshape\color{black}
-Example: }
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Rule}}\textit{: DENY fruit APPLE to
-GROW on PEACH tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit APPLE to
-be GROWN on WHEAT ground}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: flower SUNFLOWER
-to be grown on PEACH tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Explaination}}: Here, it is not possible to obtain any
-matching result - neither positive (DENY or PERMIT) nor negative
-(NOT\_APPLICABLE, see below)}
-
-{\color{black}
-In the request, the {\textquotedbl}ground{\textquotedbl} is completely
-uncomparable to the {\textquotedbl}tree{\textquotedbl} in policy. One
-can compare {\textquotedbl}PEACH tree{\textquotedbl} and
-{\textquotedbl}APPLE tree{\textquotedbl} because they are both
-{\textquotedbl}tree{\textquotedbl}; But it is impossible to compare
-{\textquotedbl}PEACH tree{\textquotedbl} and {\textquotedbl}WHEAT
-ground{\textquotedbl} becaue they of different kind (Policy is about
-tree and Request is about ground).}
-
-{\color{black}
-In a similar way one can{\textquotesingle}t compare {\textquotedbl}fruit
-APPLE{\textquotedbl} and {\textquotedbl}flower SUNFLOWER{\textquotedbl}
-(here policy is about fruits and Request is about flower).}
-
-{\color{black}
-Any other situation which makes it impossibile to compare two attributes
-will also cause {\textquotedbl}INDETERMINATE{\textquotedbl}.}
-
-\bigskip
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textbf{NOT\_APPLICABLE} - all parts of the Request have corresponding
-parts in the Policy, but some value of those parts are not the same.
-Hence request does not match policy.}
-\end{itemize}
-{\bfseries\itshape\color{black}
-Example:}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Rule}}\textit{: DENY fruit APPLE to
-GROW on PEACH tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit APPLE to
-be GROWN on APPLE tree }}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit ORANGE to
-be GROWN on PEACH tree }}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit ORANGE to
-be GROWN on APPLE tree }}
-
-{\color{black}
-Explanation: for each part of the Request evaluator can find relevant
-part in the Policy - both Policy and Request are about fruit and tree.
-But the values do not match.}
-
-{\color{black}
-If it is required to reduce evaluation results to boolean value PERMIT
-maps to TRUE and rest of results to FALSE.}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Note}}\textit{: It would be useful to make it possible
-to specify secondary effect which would become }active in case Request
-is NOT\_APPLICABLE. For example:}
-
-{\itshape\color{black}
-DENY fruit APPLE to GROW on PEACH tree otherwise PERMIT}
-
-{\color{black}
-But one should be careful because example above would allow fruit PLUMS
-to grow on APPLE trees :)}
-
-{\color{black}
-This kind of requirement can be supported by using the algorithm between
-policies. \ For example, in case of above scenario, we can use some
-algorithm like {\textquotedbl}Permit-if-notapplicable{\textquotedbl}.
-See below the {\textquotedbl}Policy matching{\textquotedbl} part for
-more explaination. \ }
-
-\subsubsection{Policy MAtching}
-{\selectlanguage{english}\upshape\color{black}
-Policy is made of Rule elements. Request is evaluated against each Rule.
-Each evaluation produces same results as policy evaluation described
-above. The results from all Rules are then combined in order to produce
-final result for whole policy. Results Combining Algorithm is specified
-in Policy. There are 26 algorithms currently:}
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-\textbf{\textit{Deny-Overrides}} - this is default if no algorithm
-specified.}
-
-\begin{itemize}
-\item {\color{black}
-If there is at least one DENY in results final result is DENY.}
-\item {\color{black}
-Otherwise if there is at least one PERMIT, the final result is PERMIT.}
-\item {\color{black}
-Otherwise if there is at least one NOT\_APPLICABLE final result is
-NOT\_APPLICABLE.}
-\item {\color{black}
-Otherwise final result is INDETERMINATE.}
-\end{itemize}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-\textbf{NOTE}\textbf{: CHECK if algorithm description is correct}}
-
-{\color{black}
-Special case is Policy wih no rules. Probably such policy shoudl be
-treated as always producing DENY.}
-
-{\bfseries\color{black}
-Discussion}
-
-{\itshape\color{black}
-If there is no Rule under Policy, it means no restriction from this
-Policy, what exactly shoud the Policy give? DENY or NOT\_APPLICABLE?}
-
-{\itshape\color{black}
-It seems to be more logical to produce INDETERMINATE because there is no
-rule with elements which could be compared to request.}
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item {\bfseries\itshape\color{black}
-Permit-Overrides.}
-
-\begin{itemize}
-\item {\color{black}
-If there is at least one PERMIT in results final result is PERMIT.}
-\item {\color{black}
-Otherwise if there is at least one DENY the final result is DENY.}
-\item {\color{black}
-Otherwise if there is at least one NOT\_APPLICABLE final result is
-NOT\_APPLICABLE.}
-\item {\color{black}
-Otherwise final result is INDETERMINATE.}
-\end{itemize}
-\end{itemize}
-{\color{black}
-Special case is Policy wih no rules. Probably such policy should be
-treated as always producing DENY.}
-
-{\bfseries\color{black}
-NOTE: CHECK if algorithm description is correct}
-
-{\bfseries\color{black}
-Discussion}
-
-{\itshape\color{black}
-Question? The same as above.}
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item {\bfseries\itshape\color{black}
-Ordered algorithms.}
-\end{itemize}
-{\selectlanguage{english}\color{black}
-These specify priorities for all four possible results. Their names look
-like Result1-Result2-Result3-Result4 with Result\# naming result types,
-for example Permit-Deny-NotApplicable-Indeterminate. The results are
-combined in following way:}
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item \begin{itemize}
-\item {\color{black}
-If there is at least one result of Result1 type then final result is
-Result1.}
-\item {\color{black}
-Otherwise if there is at least one result of Result2 type then final
-result is Result2.}
-\item {\color{black}
-Otherwise if there is at least one result of Result3 type then final
-result is Result3.}
-\item {\color{black}
-Otherwise final result is Result4.}
-\end{itemize}
-\end{itemize}
-{\color{black}
-There are 24 possible combinations of those algorithms.}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{Note:} It would be useful to have more combining algorithms. For
-example }
-
-\liststyleWWviiiNumxxi
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-Permit-if-notapplicable - the use case could be {\textquotedbl}DENY
-fruit APPLE to GROW on PEACH tree otherwise PERMIT{\textquotedbl}. In
-this case there is only one Rule under Policy, and this Rule is with
-{\textquotedbl}Deny{\textquotedbl} effect. }
-
-\begin{itemize}
-\item {\color{black}
-If this Rule gives DENY in results, final result is DENY.}
-\item {\color{black}
-Otherwise if this Rule gives NOT\_APPLICABLE, final result is PERMIT.}
-\item {\color{black}
-Otherwise final result is INDETERMINATE. }
-\end{itemize}
-\item {\selectlanguage{english}\color{black}
-{\textquotedblleft}Permit-if-allPermit{\textquotedbl} - Permit if all
-the Rules gives Permit, this algorithm is useful in case if we are
-collecting different policies from a few sources, and we want the
-request to satisfy all of them.}
-
-\begin{itemize}
-\item {\color{black}
-If all of the Rule give PERMIT, the final result is PERMIT.}
-\item {\color{black}
-Otherwise if there is at least one DENY the final result is DENY.}
-\item {\color{black}
-Otherwise if there is at least one NOT\_APPLICABLE final result is
-NOT\_APPLICABLE.}
-\item {\color{black}
-Otherwise final result is INDETERMINATE.}
-\end{itemize}
-\item {\color{black}
-OnlyOneApplicable}
-
-\begin{itemize}
-\item {\color{black}
-If there is one gives INDETERMINATE, final result INDETERMINATE is given
-immediately.}
-\item {\color{black}
-Otherwise if there is exactly only one gives applicable result (DENY or
-PERMIT), final result is as this result.}
-\item {\color{black}
-Otherwise if there is more than one gives applicable result, final
-result is INDTERMINATE.}
-\item {\selectlanguage{english}\color{black}
-Otherwise final result is NOT\_APPLICABLE.}
-\end{itemize}
-\end{itemize}
-{\color{black}
-This algorithm makes sure that only one Rule is selected when making
-decision.}
-
-\liststyleWWviiiNumxvii
-\begin{itemize}
-\item {\color{black}
-FirstApplicable}
-
-\begin{itemize}
-\item {\color{black}
-If there is one give DENY, PERMIT or INDETERMINATE result, final result
-is given immediately as this result.}
-\item {\color{black}
-Otherwise final result is NOT\_APPLICABLE.}
-\end{itemize}
-\end{itemize}
-\subsubsection{Request Structure}
-{\selectlanguage{english}\upshape\color{black}
-Request is made of RequestItem elements. Each RequestItem is evaluated
-against Policy Rule and for each evaluation separate result is
-generated as described above.RequestItem is made of 4 elements:}
-
-{\color{black}
-Subject - represents entity requesting specified action}
-
-\liststyleWWviiiNumxix
-\begin{itemize}
-\item {\color{black}
-Resource - destination/object of the action}
-\item {\color{black}
-Action - specifies what has to be done on resource}
-\item {\color{black}
-Context - for additional information which does not fit anywhere else,
-like the current time.}
-\end{itemize}
-{\color{black}
-Effectively RequestItem may have only one Subject, one Resource, one
-Action and one Context. If there are more than one element of any kind
-of sub-elelemt, then in the evaluator this RequestItem is split into
-several items containing all possible permutations and results are
-obtained for every item separately. How results are combined will be
-explained later.}
-
-{\color{black}
-Additionally Subject could contain sub-elements SubjectAttribute. Those
-are meant to represent different kinds of requesters{\textquotesingle}
-identities. Example:}
-
-\liststyleWWviiiNumxviii
-\begin{itemize}
-\item {\color{black}
-Subject}
-
-\begin{itemize}
-\item {\color{black}
-SubjectAttribute: name is ALICE}
-\item {\color{black}
-SubjectAttribute: age is YOUNG}
-\item {\color{black}
-SubjectAttribute: gender is GIRL}
-\end{itemize}
-\end{itemize}
-{\color{black}
-Context could also be made of ContextAttribute elements in the same way
-as Subject.}
-
-{\color{black}
-The following is an example of the Request:}
-
-{\itshape\color{black}
-{\textless}Request
-xmlns={\textquotedbl}http://www.nordugrid.org/schemas/request-arc{\textquotedbl}{\textgreater}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Resource
-AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Context
-AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
-Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}
-
-{\itshape\color{black}
-{\textless}/Request{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-While evaluating this RequestItem will be split into two RequestItems:}
-
-{\itshape\color{black}
-{\textless}Request
-xmlns={\textquotedbl}http://www.nordugrid.org/schemas/request-arc{\textquotedbl}{\textgreater}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Resource
-AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Context
-AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
-Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Resource
-AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Context
-AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
-Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}
-
-{\itshape\color{black}
-{\textless}/Request{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-The following means this Subject possesses both of these Attributes.}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\color{black}
-However, the following means two Subject each of which possesses one
-Attribute.}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Subject{\textgreater}}}
-
-{\color{black}
-The {\textquotedbl}Type{\textquotedbl} xml-attribute is for
-distinguishing how to process the xml-node value, which is critical
-when evaluate two value from request side and policy side because
-different type requires different evaluating/comparing approach. The
-default {\textquotedbl}Type{\textquotedbl} is
-{\textquotedbl}string{\textquotedbl}, in this case (also with the
-{\textquotedbl}Function{\textquotedbl} xml-attribute on the policy side
-is {\textquotedbl}equal{\textquotedbl}, which will be explained later),
-each letters of these two values will be compared one by one when
-evaluating them.}
-
-{\color{black}
-The {\textquotedbl}AttributeId{\textquotedbl} xml-attribute is for
-evaluator to find the Attribute with AttributeId from the request side
-which corresponds to the Attribute with the same AttributeId on the
-policy side. Only if two Attributes{\textquotesingle} AttributeId are
-equal, the evaluator will then compare the value.}
-
-{\color{black}
-Each RequestItem will be sequencialy and independently evaluated against
-policy/policies. So for one Request (including few RequestItems), some
-RequestItem could get positive evaluation result (PERMIT) from policy
-engine, others could get negative evaluation result (DENY,
-NOT\_APPLICABLE, INDETERMINATE).}
-
-{\color{black}
-It is up to policy decision point to make final decision according to
-the evaluation results returned by evaluator, and the evaluator itself
-can not give this kind of final decision.}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{NOTE: This probably should be changed because evaluator is fed
-with complete policy and complete request. Hence it is illogical that
-it retu}\textbf{rns multiple decisions.}}
-
-{\color{black}
-Basically the policy decision point will feed policy engine with
-request, get back evaluation results, }
-
-{\color{black}
-and make final decision.}
-
-\bigskip
-
-\subsubsection{Rule Composition and Matching}
-{\selectlanguage{english}\upshape\color{black}
-Policy rule is made of 4 elements - Subjects, Resources, Actions,
-Conditions (See the following example). Those are only used to group
-multiple elements Subject, Resource, Action, Condition. For instance,
-you can merge two Rules with the same Resources, Actions, Conditions,
-and the same {\textquotedbl}Effect{\textquotedbl} but different
-Subjects into one Rule.}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{NOTE: That is strange. Why do we need }\textbf{Rule at all?}}
-
-{\selectlanguage{english}\upshape\color{black}
-There is no logical relationship between Subject, which means you can
-split one Rule into two Subject (under Subjects) into two Rule (each of
-which has one Subject (under Subjects)).From now only later ones
-(Subjects with only one Subject as sub-element, and the same for
-others) are described. Their meaning is same as in request with
-Condition corresponding to Context. Subject and Condition elements are
-also made of Attributes. All elements may be present more than one
-time. During procedure of matching each element in RequestItem is
-matched against all elements of same kind in Policy - Subject is
-matched to Subject, Resource to Resource, etc. For every combination 3
-possible results are produced:}
-
-\liststyleWWviiiNumxx
-\begin{itemize}
-\item {\color{black}
-MATCHED - element from RequestItem matched element in Policy Rule.
-Example:}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{RequestItem Resource}}\textit{: place called
-WONDERLAND}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{PolicyItem Resource}}\textit{: place called WONDERLAND}}
-
-\liststyleWWviiiNumxx
-\begin{itemize}
-\item {\color{black}
-NOT MATCHED - element from RequestItem did not match element in Policy
-Rule. Example:}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{RequestItem Resource}}\textit{: place called
-WONDERLAND}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{PolicyItem Resource}}\textit{: place called PLAYGROUND}}
-
-\liststyleWWviiiNumxx
-\begin{itemize}
-\item {\color{black}
-INDETERMINATE- element from RequestItem could not be compared to element
-in Policy \ Rule because they are of incompatible ids/belong to
-different namespaces. Example:}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{RequestItem Resource}}\textit{: place called WONDERLAND
-(with namespace {\textquotedbl}place{\textquotedbl})}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{PolicyItem Resource}}\textit{: LEMON tree (with
-namespace {\textquotedbl}tree{\textquotedbl})}}
-
-{\color{black}
-The produced results then combined to produce final 4 types of results
-in following way:}
-
-\liststyleWWviiiNumxx
-\begin{itemize}
-\item {\color{black}
-If for every element in RequestItem there is at least one MATCHED result
-then result for this Policy Rule is as specified in the corresponding
-Effect (Deny or Permit).}
-\item {\color{black}
-Otherwise if for every element in RequestItem there is at least one gets
-INDETERMINATE result then result for Policy Rule is INDETERMINATE.}
-\item {\color{black}
-Otherwise result is NOT\_APPLICABLE.}
-\end{itemize}
-{\color{black}
-Special case is then RequestItem does not have the element(s) of some
-kind (Subject, Action, Resource or Context/Condition). If there are
-elements of corresponding kind in the Policy Rule then such situation
-should be considered as INDETERMINATE.}
-
-{\color{black}
-The following is an example of the Policy:}
-
-{\itshape\color{black}
-{\textless}Policy
-xmlns={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc{\textquotedbl}
-CombiningAlg={\textquotedbl}Permit-Overrides{\textquotedbl}{\textgreater}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}Rule
-Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}Subjects{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}/Subjects{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}Actions{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Delete{\textless}/Action{\textgreater}
-\ \ }}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}/Actions{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}Resources{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Resource
-AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textquotedbl}{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}/Resources{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}Conditions{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Condition
-AttributeId={\textquotedbl}urn:knowarc:period{\textquotedbl}
-Type={\textquotedbl}period{\textquotedbl}
-Function={\textquotedbl}Inrange{\textquotedbl}{\textgreater}2008-09-10T20:30:20/P1Y1M{\textless}/Condition{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}/Conditions{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}/Rule{\textgreater}}}
-
-{\itshape\color{black}
-{\textless}/Policy{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-For the Subject which includes two Attributes in this example:}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-These two attributes mean the Rule requires the request should possess
-both of these two attributes.}
-
-{\color{black}
-However, if You put these above two Attribute into two Subject
-elements:}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-Then it means the Rule requires the request to possess at least one of
-these two attributes.}
-
-{\color{black}
-For the xml-attribute {\textquotedbl}Type{\textquotedbl} and
-{\textquotedbl}AttributeId{\textquotedbl}, the explaination for Request
-example also applies here.}
-
-{\color{black}
-The {\textquotedbl}Function{\textquotedbl} xml-attribute is for
-distinguishing different comparison algorithm when comparing these two
-xml-node value. If Function is absent,
-{\textquotedbl}equal{\textquotedbl} will be used as default.}
-
-\subsubsection{Rule Elements Matching}
-{\color{black}
-For elements without attributes those elements have:}
-
-\liststyleWWviiiNumxxii
-\begin{itemize}
-\item {\color{black}
-Kind specified by AttributeId XML attribute. There is no default.}
-\item {\color{black}
-Matching algorithm specified by Id XML attribute. By default
-string-equal matching is used.}
-\item {\color{black}
-Content}
-\end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-Example: \textit{LEMON tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Kind}}\textit{: tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Matching algorithm}}\textit{: default}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Content}}\textit{: LEMON}}
-
-{\color{black}
-Matching procedure consists of following steps:}
-
-\liststyleWWviiiNumxxiii
-\begin{itemize}
-\item {\color{black}
-Kinds are compared using simple string equal matching. If those do not
-match then result is INDETERMINATE.}
-\item {\color{black}
-Matching algorithm is used to compare content of elements. Result is
-either MATCH or NO\_MATCH according to matching algorithm.}
-\end{itemize}
-{\color{black}
-Each element on the RequestItem must satisfy corresponding element in
-Rule.In detail, for Subjects element under Rule, if there is at least
-one Subject (with one Attribute or a few Attribute) which is matched by
-a Subject on this RequestItem, we say this Subjects is matched by the
-RequestItem; and also the same for the other elements (Actions,
-Resources, Conditions).}
-
-{\selectlanguage{english}\upshape\color{black}
-For elements with multiple Attribute sub-elements the way to judging
-whether elements match is if and only if all of the Attribute under the
-Rule have matching Attributes at RequestItem side.}
-
-{\selectlanguage{english}\upshape\color{black}
-Example of the Subject with three Attributes:}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Subject}}\textit{:}}
-
-{\selectlanguage{english}\color{black}
-\textbf{\textit{SubjectAttribute}}\textit{: name is ALICE}}
-
-{\selectlanguage{english}\color{black}
-\textbf{\textit{SubjectAttribute}}\textit{: age is YOUNG}}
-
-{\selectlanguage{english}\color{black}
-\textbf{\textit{SubjectAttribute}}\textit{: gender is GIRL}}
-
-{\selectlanguage{english}\upshape\color{black}
-In XML that is:}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}gender{\textgreater}GIRL{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-That requires the Subject in the RequestItem to possess at least these
-three Attributes.}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}age{\textquotedbl}{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}gender{\textquotedbl}{\textgreater}GIRL{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other
-Attribute-{}-{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\itshape\color{black}
-{\textless}/RequestItem{\textgreater}}
-
-{\color{black}
-The above example shows that the Subject in the RequestItem
-{\textquotedbl}MATCH{\textquotedbl} one Subject on the Rule side.}
-
-{\color{black}
-If the Subject in the RequestItem is like this:}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}from{\textquotedbl}{\textgreater}OSLO{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other Attribute, buts
-not a {\textquotedbl}gender{\textquotedbl}-{}-{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-Then evaluator will produce INDETERMINATE as the match-making result of
-this Subject.}
-
-{\color{black}
-If the Subject in the RequestItem is like this:}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Bob{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}gender{\textquotedbl}{\textgreater}BOY{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other
-Attribute-{}-{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\color{black}
-Then evaluator will give NO\_MATCH as the match-making result of this
-Subject.}
-
-{\selectlanguage{english}\upshape\color{black}
-Finally if and only if all of the elemens (Subjects, Actions, Resources,
-Conditions) which are not empty under the Rule have been matched (gets
-MATCH) to the RequestItem, then the whole Rule is considered to be
-matched (produces MATCH result). MATCH is then mapped to final
-evaluation result depending on the specified Effect. If Effect is set
-to Deny then DENY decision will be produced for this Rule; if Effect is
-Permit then PERMIT.}
-
-{\color{black}
-Otherwise if any of the element (Subjects, Actions, Resources,
-Conditions) of RequestItem got INDETERMINATE decision then the
-INDETERMINATE decision will be made for this Rule.}
-
-{\color{black}
-Otherwise the NOT\_APPLICABLE decision will be made for this Rule. In
-other words that means at least one of the elements of this Rule got
-NO\_MATCH and the other elements got MATCH.}
-
-\subsubsection{Interface for using the policy evaluation engine}
-{\selectlanguage{english}\upshape\color{black}
-For making usage of policy evaluation engine more convenient basic
-\textit{Evaluator} class is complemented by additional interfaces.
-Below are examples of steps needed to carry out policy evaluation and
-corresponding helper interfaces.}
-
-\liststyleWWviiiNumiii
-\begin{enumerate}
-\item {\selectlanguage{english}\color{black}
-Create the policy evaluation object:}
-\end{enumerate}
-{\selectlanguage{english}\ttfamily\color{black}
-// Create object which provides an interface }
-
-{\selectlanguage{english}\ttfamily\color{black}
-// for loading other objects}
-
-{\selectlanguage{english}\ttfamily\color{black}
-ArcSec::EvaluatorLoader eval\_loader; }
-
-{\selectlanguage{english}\ttfamily\color{black}
-//Load the Evaluator }
-
-{\selectlanguage{english}\ttfamily\color{black}
-ArcSec::Evaluator* eval = NULL;}
-
-{\selectlanguage{english}\ttfamily\color{black}
-// Define name of policy evaluator.}
-
-{\selectlanguage{english}\ttfamily\color{black}
-// This one is for evaluation ARC policies}
-
-{\selectlanguage{english}\ttfamily\color{black}
-std::string evaluator = {\textquotedbl}arc.evaluator{\textquotedbl}; }
-
-{\selectlanguage{english}\ttfamily\color{black}
-eval = eval\_loader.getEvaluator(evaluator);}
-
-\liststyleWWviiiNumiv
-\begin{enumerate}
-\item {\selectlanguage{english}\color{black}
-Create the policy object:}
-\end{enumerate}
-{\selectlanguage{english}\ttfamily\color{black}
-ArcSec::Policy* policy = NULL; }
-
-{\selectlanguage{english}\ttfamily\color{black}
-// Define type of policy -- ARC policy in this case}
-
-{\selectlanguage{english}\ttfamily\color{black}
-std::string policyclassname = {\textquotedbl}arc.policy{\textquotedbl};
-}
-
-{\selectlanguage{english}\ttfamily\color{black}
-// Define source from which policy to be taken}
-
-{\selectlanguage{english}\ttfamily\color{black}
-ArcSec::SourceFile
-policy\_source({\textquotedbl}Policy\_Example.xml{\textquotedbl}); }
-
-{\selectlanguage{english}\ttfamily\color{black}
-// Load and parse policy}
-
-{\selectlanguage{english}\ttfamily\color{black}
-policy = eval\_loader.getPolicy(policyclassname, policy\_source); }
-
-\liststyleWWviiiNumv
-\begin{enumerate}
-\item {\selectlanguage{english}\color{black}
-Create the request:}
-\end{enumerate}
-{\selectlanguage{english}\ttfamily\color{black}
-ArcSec::Request* request = NULL;}
-
-{\selectlanguage{english}\ttfamily\color{black}
-// Define type of request -- ARC request in this case}
-
-{\selectlanguage{english}\ttfamily\color{black}
-std::string requestclassname =
-{\textquotedbl}arc.request{\textquotedbl};}
-
-{\selectlanguage{english}\ttfamily\color{black}
-// Define source from which request to be taken}
-
-{\selectlanguage{english}\ttfamily\color{black}
-ArcSec::SourceFile
-request\_source({\textquotedbl}Request.xml{\textquotedbl});}
-
-{\selectlanguage{english}\ttfamily\color{black}
-// Load and parse request}
-
-{\selectlanguage{english}\ttfamily\color{black}
-request = eval\_loader.getRequest(requestclassname, request\_source);}
-
-\liststyleWWviiiNumvi
-\begin{enumerate}
-\item {\selectlanguage{english}\color{black}
-Add the policy into \textit{Evaluator} object:}
-\end{enumerate}
-{\selectlanguage{english}\ttfamily\color{black}
-eval-{\textgreater}addPolicy(policy);}
-
-\liststyleWWviiiNumvii
-\begin{enumerate}
-\item {\selectlanguage{english}\color{black}
-Evaluate the request object:}
-\end{enumerate}
-{\selectlanguage{english}\ttfamily\color{black}
-ArcSec::Response *resp = NULL;}
-
-{\selectlanguage{english}\ttfamily\color{black}
-resp = eval-{\textgreater}evaluate(request);}
-
-{\selectlanguage{english}\color{black}
-The steps d) and e) can also be replaced by:}
-
-{\selectlanguage{english}\ttfamily\color{black}
-resp = eval-{\textgreater}evaluate(request, policy);}
-
-\bigskip
-
-{\selectlanguage{english}\upshape\color{black}
-The \textit{Evalutor::evaluate()} method can also be feed up with both
-\textit{Policy}/\textit{Request} objects and their sources in any
-combination. See example code at
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/testinterface.cpp
-for more details about usage of the interface.}
-
-{\selectlanguage{english}\upshape\color{black}
-The description of mentioned classes and their methods are avaialble in
-\ API document at
-\ \url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/KnowARC-API.pdf?format=raw}
-.}
-
-\subsection{Policy Decision Service}
-{\selectlanguage{english}\upshape\color{black}
-Policy decision service is a service implementation which contains the
-functionality of ArcPDP. It will accept the soap request containing
-policy decision request and return soap response containing policy
-decision response.}
-
-{\selectlanguage{english}\upshape\color{black}
-The WSDL description of policy decision service is available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl}
-. It{\textquotesingle}s configuration is presented in Section 7.5. }
-
-\subsection{Security Attributes}
-\subsubsection[Infrastructure]{Infrastructure}
-\label{bkm:Ref204009818}{\selectlanguage{english}\upshape\color{black}
-Security Attributes represent security related information inside HED
-framework and store information representing various aspects needed to
-perform authorization decison - identity of client, requested action,
-targeted resource, constraint policies. }
-
-{\selectlanguage{english}\upshape\color{black}
-Each kind of Security Attribute is represented by own class inherited
-from parent SecAttr class
-{\textless}arc/message/SecAttr.h{\textgreater}. Each Security Attribute
-stores it{\textquotesingle}s information in internal format and is
-capable to export it to one of predefined formats using Export()
-method. Currently only supported format is ARC Policy/Request XML
-document described in Section 7.1and 7.2}
-
-{\selectlanguage{english}\upshape\color{black}
-Collectors of Security Attributes instantiate corresponding classes and
-link them to Secuirity Attributes containers - MessageAuth
-{\textless}arc/message/MessageAuth.h{\textgreater} and
-MessageAuthContext {\textless}arc/message/Message.h{\textgreater}
-storing collected attributes per request and per session
-correspondingly. Each attribute is assigned a name. Current
-implementations of Security Attributes Collectors are either integrated
-into existing MCCs or implemented as separate SecHandler plugins. See
-Section 5.2for available Collectors and corresponding Security
-Attributes. }
-
-{\selectlanguage{english}\upshape\color{black}
-Note for service developers: Services may implement own authorization
-algorithms. But they may use Security Atributes as well by providing
-instances of classes inherited from SecAttr and running them through
-either configured or hardcoded processors/PDPs.}
-
-{\selectlanguage{english}\upshape\color{black}
-Processors of Security Attributes are implemented as Policy Decision
-Point components. Currently there are 2 PDP components available:}
-
-\liststyleWWviiiNumviii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-Arc PDP makes use of Security Attributes containing identities of
-client, resource and requested action. It evaluates either all or
-selected set of attributes against specified Policy documents thus
-making it possible to enforce policies defined/selected by service
-providers. }
-\item {\selectlanguage{english}\color{black}
-Delegation PDP is described below in Section 6.3}
-\end{itemize}
-\subsubsection[Available collectors]{Available collectors}
-\label{bkm:Ref204009636}{\selectlanguage{english}\upshape\color{black}
-Here Security Attribute collectors distributed as part of the ARC1 are
-described except those used for Delegation Restrictions. Those are
-described in Section 6.2}
-
-\paragraph{TCP}
-{\selectlanguage{english}\upshape\color{black}
-Information is collected inside TCP MCC. The Security Attribute is
-stored under name {\textquotesingle}TCP{\textquotesingle} and exports
-ARC Request with following attributes:}
-
-\begin{flushleft}
-\bottomcaption{Security Attributes collected at TCP MCC}
-\tablehead{\hline
-\centering \selectlanguage{english}\bfseries\itshape\color{black}
-Element &
-\centering \selectlanguage{english}\bfseries\itshape\color{black}
-AttributeId &
-\centering\arraybslash
-\selectlanguage{english}\bfseries\itshape\color{black}
-Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
-\selectlanguage{english}\color{black} Resource &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/localendpoint &
-\selectlanguage{english}\color{black} service\_ip[:service\_port]
-\\\hline
-\selectlanguage{english}\color{black} SubjectAttribute &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/remoteendpoint &
-\selectlanguage{english}\color{black} client\_ip[:client\_port] \\\hline
-\end{supertabular}
-\end{flushleft}
-
-\bigskip
-
-\paragraph{TLS}
-{\selectlanguage{english}\upshape\color{black}
-Information is collected inside TLS MCC. Generated Security Attribute
-class is stored under name {\textquotesingle}TLS{\textquotesingle} and
-exports ARC Request with following attributes:}
-
-\begin{flushleft}
-\bottomcaption{Security Attributes collected at TLS MCC}
-\tablehead{\hline
-\centering \selectlanguage{english}\bfseries\itshape\color{black}
-Element &
-\centering \selectlanguage{english}\bfseries\itshape\color{black}
-AttributeId &
-\centering\arraybslash
-\selectlanguage{english}\bfseries\itshape\color{black}
-Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
-\selectlanguage{english}\color{black} SubjectAttribute &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/tls/ca &
-\selectlanguage{english}\color{black} signer of first certificate in
-client{\textquotesingle}s chain\\\hline
-\selectlanguage{english}\color{black} SubjectAttribute &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/tls/chain &
-\selectlanguage{english}\color{black} Subject of certificate in
-client{\textquotesingle}s chain - multiple items\\\hline
-\selectlanguage{english}\color{black} SubjectAttribute &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/tls/subject &
-\selectlanguage{english}\color{black} Subject of last certificate in
-client{\textquotesingle}s chain\\\hline
-\selectlanguage{english}\color{black} SubjectAttribute &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/tls/identity &
-\selectlanguage{english}\color{black} Subject of last non-proxy
-certificate in client{\textquotesingle}s chain\\\hline
-\end{supertabular}
-\end{flushleft}
-\paragraph{HTTP}
-{\selectlanguage{english}\upshape\color{black}
-Information is collected inside HTTP MCC. The Security Attribute is
-stored under name {\textquotesingle}HTTP{\textquotesingle} and exports
-ARC Request with following attributes:}
-
-\begin{flushleft}
-\bottomcaption{Security Attributes collected at HTTP MCC}
-\tablehead{\hline
-\centering \selectlanguage{english}\bfseries\itshape\color{black}
-Element &
-\centering \selectlanguage{english}\bfseries\itshape\color{black}
-AttributeId &
-\centering\arraybslash
-\selectlanguage{english}\bfseries\itshape\color{black}
-Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
-\selectlanguage{english}\color{black} Resource &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/http/path &
-\selectlanguage{english}\color{black} HTTP path without host and port
-part\\\hline
-\selectlanguage{english}\color{black} Action &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/http/method &
-\selectlanguage{english}\color{black} HTTP method\\\hline
-\end{supertabular}
-\end{flushleft}
-\paragraph{SOAP}
-{\selectlanguage{english}\upshape\color{black}
-Information is collected inside SOAP MCC. Security Attribute is stored
-under name {\textquotesingle}SOAP{\textquotesingle} and exports ARC
-Request with following attributes:}
-
-\begin{flushleft}
-\bottomcaption{Security Attributes collected at SOAP MCC}
-\tablehead{\hline
-\centering \selectlanguage{english}\bfseries\itshape\color{black}
-Element &
-\centering \selectlanguage{english}\bfseries\itshape\color{black}
-AttributeId &
-\centering\arraybslash
-\selectlanguage{english}\bfseries\itshape\color{black}
-Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
-\selectlanguage{english}\color{black} Resource &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint &
-\selectlanguage{english}\color{black} To element of WS-Addressing
-structure\\\hline
-\selectlanguage{english}\color{black} Action &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/soap/operation &
-\selectlanguage{english}\color{black} SOAP top level element name
-without namespace prefix\\\hline
-\selectlanguage{english}\color{black} Context &
-\selectlanguage{english}\color{black}
-http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace &
-\selectlanguage{english}\color{black} Namespace of SOAP top level
-element\\\hline
-\end{supertabular}
-\end{flushleft}
-\subsection[Delegation Restrictions]{Delegation Restrictions}
-\label{bkm:Ref204009393}\subsubsection[Delegation
-Architecture]{Delegation Architecture}
-\label{bkm:Ref204009914}{\selectlanguage{english}\upshape\color{black}
-In current implementation delegation is achieved through Identity
-Delegation implemented using X509 Proxy Certificates as defined in RFC
-3820. Client wishing to allow service to act on it{\textquotesingle}s
-behalf provides Proxy Certificate to the service using Web Service
-based Delegation interface described in Section 6.4}
-
-{\selectlanguage{english}\upshape\color{black}
-For limiting the scope of delegated credentials along with usually used
-time constraints it is possible to attach Policy document to Proxy
-Certificate. According to RFC 3820 Policy is stored in ProxyPolicy
-extension. In order not to introduce new type of object Policy is
-assigned id-ppl-anyLanguage identifier. RFC 3820 allows any octet
-string associated with such object. We are using textual representation
-of ARC Policy XML document.}
-
-{\selectlanguage{english}\upshape\color{black}
-Each deployment implementing Delegation Restrictions must use dedicated
-Security Handler plugin (see section 5.1) to collect all Policy
-documents from Proxy Certificates used for establishing secure
-connection. Then those documents must be processed by dedicated Policy
-Decision Point plugin (see section 2.3) to make a final decision based
-on collected Policies and various information about
-client{\textquotesingle}s identity and requested operation. Service or
-MCC chain supporting Delegation Restrictions must accept negative
-decision of this PDP as final and do not override it with any other
-decision based on other policies.}
-
-\subsubsection[Delegation Collector]{Delegation Collector}
-\label{bkm:Ref204009734}{\selectlanguage{english}\upshape\color{black}
-This Security Attribute is collected by dedicated Security Handler
-plugin named {\textquotedbl}delegation.collector{\textquotedbl}
-avaialble as part of the ARC1 distribution. It extracts policy document
-stored inside X509 certificate proxy extension as defined in RFC3820
-and described in Section 6.1All proxy certificates in a chain provided
-by client are examined and all available policies are extracted. }
-
-{\selectlanguage{english}\upshape\color{black}
-Extracted content is converted into XML document. Then document is
-checked to be of ARC Policy kind. If policy is not recognized as ARC
-Policy procedure fails and that causes failure of communication.}
-
-{\selectlanguage{english}\upshape\color{black}
-Proxy certificates with id-ppl-inheritAll [5. RFC3820.
-http://www.faqs.org/rfcs/rfc3820.html] property are passed through and
-no policy document is generated for them. Proxies with other type of
-policies including id-ppl-independent are not accepted and generate
-immediate failure.}
-
-\subsubsection[Delegation PDP]{Delegation PDP}
-\label{bkm:Ref204009695}{\selectlanguage{english}\upshape\color{black}
-The Delegation PDP is similar to the Arc PDP described above except that
-it takes it{\textquotesingle}s Policy documents directly from Security
-Attributes. Differently from Arc PDP it is meant to be used for
-enforcing policies defined by client.}
-
-\subsubsection[Delegation interface]{Delegation interface}
-\label{bkm:Ref204009766}{\selectlanguage{english}\upshape\color{black}
-Delegation interface in the ARC1 is implemented using Web Service
-approach. Each ARC1 service wishing to accept delegated credentials
-implements this interface. Here is how delegation procedure works:}
-
-\liststyleWWviiiNumix
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-Step 1}
-\item {\selectlanguage{english}\color{black}
-Client contacts service requesting operation DelegateCredentialsInit.
-This operation has no arguments.}
-\item {\selectlanguage{english}\color{black}
-Service responds with DelegateCredentialsInitResponse message with
-element TokenRequest. That element contains credentials request
-generated by service in Value. Type of request is defined by attribute
-Format. Currently only supported format is x509. Along with Value
-service provides identifier Id which is used in second step.}
-\end{itemize}
-\liststyleWWviiiNumx
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-Step 2}
-\item {\selectlanguage{english}\color{black}
-Client requests UpdateCredentials operation with DelegatedToken
-argument. This element contains Value with serialized delegated
-credentials and Id which links it to first step. Delegated token
-element may also contain multiple Reference elements. Reference refers
-to the object which these credentials should be applied to in a way
-specific to the service. The DelegatedToken element may also be used
-for delegating credentials when Step 2 is combined with other
-operations on service in service specific way.}
-\item {\selectlanguage{english}\color{black}
-Service responds with empty UpdateCredentialsResponse message.}
-\end{itemize}
-\subsection[Schemas, descriptions and examples]{Schemas, descriptions
-and examples}
-\label{bkm:Ref204010074}\subsubsection[Authorization
-Policy]{Authorization Policy}
-\label{bkm:Ref204009564}{\selectlanguage{english}\upshape\color{black}
-XML schema with comments available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd}
-.}
-
-\subsubsection[Authorization Request]{Authorization Request}
-\label{bkm:Ref204009595}{\selectlanguage{english}\upshape\color{black}
-XML schema with comments available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd}{\textstyleInternetlink{
-.}}}
-
-\subsubsection{Authorization Response}
-{\selectlanguage{english}\upshape\color{black}
-XML schema with comments available at
-\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Response.xsd}}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd}{\textstyleInternetlink{
-.}}}
-
-\subsubsection[Interface of policy decision service]{Interface of policy
-decision service}
-{\selectlanguage{english}\upshape\color{black}
-WSDL with comments available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl}
-.}
-
-\subsubsection{Configuration of PDP service}
-{\selectlanguage{english}\upshape\color{black}
-XML schema with comments available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.xsd}
-.}
-
-{\selectlanguage{english}\upshape\color{black}
-Below is an example configuration of PDP service which can evaluate ARC
-Request against ARC Policy stored in local file.}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}Service name={\textquotedbl}pdp.service{\textquotedbl}
-id={\textquotedbl}pdp\_service{\textquotedbl}{\textgreater}}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}!-{}-The element {\textless}Evaluator/{\textgreater},
-{\textless}Policy/{\textgreater} and {\textless}Request/{\textgreater}
-configuration}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ are supposed to be used to load object; element
-{\textless}PolicyStore/{\textgreater} is}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ supposed to be used to get the location of
-policy-{}-{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}pdp:PDPConfig{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}pdp:PolicyStore{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}Location
-Type={\textquotedbl}file{\textquotedbl}{\textgreater}Policy\_Example.xml{\textless}/Location{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}- other policy
-location-{}-{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}/pdp:PolicyStore{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}pdp:Evaluator
-name={\textquotedbl}arc.evaluator{\textquotedbl} /{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}pdp:Policy
-name={\textquotedbl}arc.policy{\textquotedbl} /{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}pdp:Request
-name={\textquotedbl}arc.request{\textquotedbl} /{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}/pdp:PDPConfig{\textgreater}}}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}/Service{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-See Section 7.7for the explanation of ARC Policy.}
-
-\subsubsection{SimpleList PDP configuration and Policy Example}
-{\selectlanguage{english}\upshape\color{black}
-XML schema with comments available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/simplelistpdp/SimpleListPDP.xsd}
-.}
-
-{\selectlanguage{english}\upshape\color{black}
-Below is an example configuration of SimpleList PDP inside
-{\textquotedblleft}echo{\textquotedblright} service.}
-
-{\selectlanguage{spanish}\ttfamily\color{black}
-{\textless}Service name={\textquotedbl}echo{\textquotedbl}
-id={\textquotedbl}echo{\textquotedbl}{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}SecHandler
-name={\textquotedbl}arc.authz{\textquotedbl}
-id={\textquotedbl}authz{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}PDP
-name={\textquotedbl}}\texttt{simplelist}\texttt{.pdp{\textquotedbl}}\texttt{
-location=}\texttt{{\textquotedbl}}\texttt{simplelist}\texttt{{\textquotedbl}}\texttt{/}\texttt{{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ }\foreignlanguage{spanish}{\texttt{{\textless}/SecHandler{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
-{\textless}/echo:prefix{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}echo:suffix{\textgreater}
-]{\textless}/echo:suffix{\textgreater}}}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}/Service{\textgreater}}
-
-{\selectlanguage{english}\color{black}
-The attribute {\textquotedblleft}name{\textquotedblright} of
-{\textless}PDP/{\textgreater} is critical for loading the object.
-Specifically, the name
-{\textquotedblleft}simplelist.pdp{\textquotedblright} is for loading
-the SimpleList PDP object.}
-
-{\selectlanguage{english}\upshape\color{black}
-The policy file {\textquotedblleft}simplelist{\textquotedblright} is a
-local file which contains the list of X509 subjects of authorized
-entities. It the peer certificate is proxy certificate, the identity in
-this list should only include the original DN of
-users{\textquotesingle}s certificate.}
-
-{\selectlanguage{english}\upshape\color{black}
-For example content of \textit{simplelist} file may look like this:}
-
-{\selectlanguage{spanish}\color{black}
-/C=NO/O=UiO/CN=test1}
-
-{\selectlanguage{spanish}\color{black}
-/C=NO/O=UiO/CN=test2}
-
-\subsubsection[Arc PDP configuration and Policy Example]{Arc PDP
-configuration and Policy Example}
-\label{bkm:Ref204009974}\label{bkm:RefSecarcpdpconfigandpolicy}{\selectlanguage{english}\upshape\color{black}
-XML schema with comments available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/ArcPDP.xsd}
-.}
-
-{\selectlanguage{english}\upshape\color{black}
-Below is an example of configuration of Arc PDP inside
-{\textquotedblleft}echo{\textquotedblright} service.}
-
-{\selectlanguage{spanish}\ttfamily\color{black}
-{\textless}Service name={\textquotedbl}echo{\textquotedbl}
-id={\textquotedbl}echo{\textquotedbl}{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}SecHandler
-name={\textquotedbl}arc.authz{\textquotedbl}
-id={\textquotedbl}authz{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}PDP
-name={\textquotedbl}arc.pdp{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ {\textless}PolicyStore{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}Location
-type={\textquotedbl}file{\textquotedbl}{\textgreater}Policy\_Example.xml{\textless}/Location{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}!-{}-other policy
-location-{}-{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/PolicyStore{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}/PDP{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ }\foreignlanguage{spanish}{\texttt{{\textless}/SecHandler{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
-{\textless}/echo:prefix{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}echo:suffix{\textgreater}
-]{\textless}/echo:suffix{\textgreater}}}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}/Service{\textgreater}}
-
-{\selectlanguage{english}\color{black}
-The name {\textquotedblleft}arc.pdp{\textquotedblright} is for loading
-the ArcPDP object.}
-
-{\selectlanguage{english}\color{black}
-There could be a few policy files under
-{\textless}PolicyStore/{\textgreater}. The request will be checked
-against all of the policies.}
-
-{\selectlanguage{english}\upshape\color{black}
-There is an example policy for echo service below. See Section 7.1for
-the policy schema. The example policy is \ made of following elements:
-}
-
-{\selectlanguage{english}\color{black}
-1. Line 14 \ defines resource being protected. In this it is everything
-located under HTTP path
-\ {\textquotedblleft}/Echo{\textquotedblright}.}
-
-{\selectlanguage{english}\color{black}
-2. Lines 17 and 18 define allowed HTTP operations to be
-{\textquotedblleft}POST{\textquotedblright} and
-{\textquotedblleft}GET{\textquotedblright}. Line 19 also defines SOAP
-operation {\textquotedblleft}echo{\textquotedblright} to be applied to
-service at path defined above.}
-
-{\selectlanguage{english}\color{black}
-3. Lines 10 and 9 require the requester to present X509 certificate with
-specified identity and signed by specified Certification Authority.}
-
-{\selectlanguage{english}\color{black}
-4. No {\textless}Conditions/{\textgreater} defined.}
-
-{\selectlanguage{english}\color{black}
-5. Line 3 defines that if and only if all of the above constraints have
-been satisfied by requester, the {\textless}Rule/{\textgreater}
-evaluates to Permit decision.}
-
-{\selectlanguage{english}\color{black}
-The Secuirity Attributes used by Arc PDP are collected by different
-MCCs. It is possible for service to collect some application-specific
-attributes by implementing class inherited from SecAtt. And that should
-be the task of application developer.}
-
-{\selectlanguage{english}\upshape\color{black}
-Administrator of service can configure Authorization SecHandler -
-arc.authz - for each MCC and Service and define reasonable and
-meaningful policy. While defining policy the administrator must take
-into account that the attributes defined in the policy should be
-already collected by previous components in a chain. For instance,
-policy with AttributeId
-{\textquotedblleft}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedblright}
-should not be configured inside SecHandler attached to MCCTLS. }
-
-\liststyleWWviiiNumxiv
-\begin{enumerate}
-\item {\selectlanguage{english}\ttfamily\color{black}
-{\textless}?xml version={\textquotedbl}1.0{\textquotedbl}
-encoding={\textquotedbl}UTF-8{\textquotedbl}?{\textgreater}}
-\item {\selectlanguage{english}\ttfamily\color{black}
-{\textless}Policy
-xmlns={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc{\textquotedbl}
-PolicyId={\textquotedbl}sm-example:arcpdppolicy{\textquotedbl}
-CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}Rule
-Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ {\textless}Description{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ Example policy for echo service}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ {\textless}/Description{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Subjects{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Subject{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/ca{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=CA{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/identity{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=test{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Subject{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Subjects{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Resources{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Resource
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/Echo{\textless}/Resource{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Resources{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}POST{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}GET{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/operation{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}echo{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Conditions/{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
-\item {\selectlanguage{english}\ttfamily\color{black}
-{\textless}/Policy{\textgreater}}
-\end{enumerate}
-\subsubsection{PDP Service Invoker configuration}
-{\selectlanguage{english}\upshape\color{black}
-Configuration XML schema with comments is available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/pdpserviceinvoker/ArcPDPServiceInvoker.xsd}
-.}
-
-{\selectlanguage{english}\upshape\color{black}
-Below is an example of configuration of PDP Service Invoker inside
-{\textquotedblleft}echo{\textquotedblright} service.}
-
-{\selectlanguage{spanish}\ttfamily\color{black}
-{\textless}Service name={\textquotedbl}echo{\textquotedbl}
-id={\textquotedbl}echo{\textquotedbl}{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}SecHandler
-name={\textquotedbl}arc.authz{\textquotedbl}
-id={\textquotedbl}authz{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}!-{}-Remote pdp service
-invoking-{}-{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ {\textless}PDP
-name={\textquotedbl}pdpservice.invoker{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}ServiceEndpoint{\textgreater}https://127.0.0.1:60001/pdp.service{\textless}/ServiceEndpoint{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}KeyPath{\textgreater}./key.pem{\textless}/KeyPath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}CertificatePath{\textgreater}./cert.pem{\textless}/CertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}CACertificatePath{\textgreater}./ca.pem{\textless}/CACertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/PDP{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ }\foreignlanguage{spanish}{\texttt{{\textless}/SecHandler{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}next
-id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
-{\textless}/echo:prefix{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}echo:suffix{\textgreater}
-]{\textless}/echo:suffix{\textgreater}}}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}/Service{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-The name {\textquotedblleft}pdpservice.invoker{\textquotedblright}
-defines the PDP Service Invoker object.}
-
-{\selectlanguage{english}\color{black}
-The PDP Service Invoker is a client of PDP Service. The configuration
-options include endpoint of service and credentials to be used for
-establishing secure connection.}
-
-\bigskip
-
-\subsubsection{Delegation PDP configuration}
-{\selectlanguage{english}\upshape\color{black}
-Configuration XML schema with comments available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/delegationsh/DelegationPDP.xsd}
-.}
-
-{\selectlanguage{english}\upshape\color{black}
-Below is an example of configuration of Delegation PDP inside
-{\textquotedblleft}echo{\textquotedblright} service.}
-
-{\selectlanguage{spanish}\ttfamily\color{black}
-{\textless}Service name={\textquotedbl}echo{\textquotedbl}
-id={\textquotedbl}echo{\textquotedbl}{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\texttt{{\textless}SecHandler
-name={\textquotedbl}arc.authz{\textquotedbl}
-id={\textquotedbl}authz{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}PDP
-name={\textquotedbl}delegation.pdp{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}/SecHandler{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}next
-id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ }\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
-{\textless}/echo:prefix{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:suffix{\textgreater}
-]{\textless}/echo:suffix{\textgreater}}}}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}/Service{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-For Delegation PDP, no specific configuration is needed. We only need to
-switch it on by adding {\textless}PDP
-name={\textquotedbl}delegation.pdp{\textquotedbl}/{\textgreater} under
-{\textless}SecHandler/{\textgreater}}
-
-\subsubsection[Delegation SecHandler Configuration]{Delegation
-SecHandler Configuration}
-\label{bkm:RefSecdelegsechandleconfig}{\selectlanguage{english}\upshape\color{black}
-Below is an example of configuration of Delegation SecHandler inside TLS
-MCC component.}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Component
-name={\textquotedbl}tls.service{\textquotedbl}
-id={\textquotedbl}tls{\textquotedbl}{\textgreater} {\textless}next
-id={\textquotedbl}http{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:KeyPath{\textgreater}./key.pem{\textless}/tls:KeyPath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CertificatePath{\textgreater}./cert.pem{\textless}/tls:CertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CACertificatePath{\textgreater}./ca.pem{\textless}/tls:CACertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}-delegation.collector
-must be inside tls MCC-{}-{\textgreater}}}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}SecHandler
-name={\textquotedbl}delegation.collector{\textquotedbl} }
-
-{\selectlanguage{english}\ttfamily\color{black}
-id={\textquotedbl}delegation{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}{\textless}/SecHandler{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-Current implementation of Delegation\textbf{ }SecHandler must be
-attached to TLS MCC. }
-
-\subsubsection[UsernameToken SecHandler configuration]{UsernameToken
-SecHandler configuration}
-\label{bkm:RefSecuserametokenconfig}{\selectlanguage{english}\upshape\color{black}
-Configuration XML schema with comments available at
-\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTo}}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{kenSH.xsd}}
-.}
-
-{\selectlanguage{english}\upshape\color{black}
-Below is an example of configuration of UsernameToken SecHandler inside
-MCCSOAP component.}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}Component
-name={\textquotedbl}soap.service{\textquotedbl}
-id={\textquotedbl}soap{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}next
-id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
-name={\textquotedbl}usernametoken.handler{\textquotedbl}
-id={\textquotedbl}usernametoken{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}Process{\textgreater}extract{\textless}/Process{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}PasswordSource{\textgreater}password.txt{\textless}/PasswordSource{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}/SecHandler{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-UsernameToken SecHandler must be configured under SOAP MCC.}
-
-\subsubsection[X509Token SecHandler configuration]{X509Token SecHandler
-configuration}
-{\selectlanguage{english}\upshape\color{black}
-Configuration XML schema with comments available at
-\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/x509tokensh/X509TokenSH.xsd}}.}
-
-{\selectlanguage{english}\upshape\color{black}
-Below is an example of configuration of X509Token SecHandler inside
-MCCSOAP component.}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}Component
-name={\textquotedbl}soap.service{\textquotedbl}
-id={\textquotedbl}soap{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}next
-id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
-name={\textquotedbl}x509token.handler{\textquotedbl}
-id={\textquotedbl}x509token{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}
-
-\bigskip
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}Process{\textgreater}extract{\textless}/Process{\textgreater}}}
-
-\bigskip
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}CACertificatePath{\textgreater}ca.pem{\textless}/CACertificatePath{\textgreater}}}
-
-\bigskip
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/SecHandler{\textgreater}}}
-
-\bigskip
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-X509Token SecHandler must be configured under SOAP MCC.}
-
-\bigskip
-
-\subsection{User Manual}
+
+\bigskip
+\clearpage\clearpage\setcounter{page}{1}\pagestyle{Convertv}
+
+\bigskip
+
+\subsection{1. User Manual}
 {\selectlanguage{english}\color{black}
@@
 The description of PDP configuration and ARC Policy example are
-available in Section 7}
+available in Section }
 
@@
 enabled in configuration of service. Configuration of Delegation
-SecHandler is described in section 7.10. }
+SecHandler is described in section . }
 
@@
 The Delegation Policy is the same as the ARC Policy explained in section
-7.7. Simple example below renders delegated credentials usable only for
+. Simple example below renders delegated credentials usable only for
 contacting service attached to HTTP communication channel under path
@@
 xmlns={\textquotedbl}}\url{http://www.nordugrid.org/schemas/policy-arc}\texttt{{\textquotedbl}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ PolicyId={\textquotedbl}sm-example:policy1{\textquotedbl}
-CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}Rule RuleId={\textquotedbl}rule1{\textquotedbl}
-Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Resources{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Resource
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ PolicyId={\textquotedbl}sm-example:policy1{\textquotedbl}
+CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}Rule RuleId={\textquotedbl}rule1{\textquotedbl}
+Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Resources{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Resource
 Type={\textquotedbl}string{\textquotedbl}
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedbl}{\textgreater}/arex{\textless}/Resource{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Resources{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
+AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedbl}{\textgreater}/arex{\textless}/Resource{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/Resources{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Actions{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Action
 Type={\textquotedbl}string{\textquotedbl}
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}{\textgreater}POST{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
+AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}{\textgreater}POST{\textless}/Action{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/Actions{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}/Rule{\textgreater}}
 \item {\selectlanguage{english}\ttfamily\color{black}
@@
 CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}Rule RuleId={\textquotedbl}rule1{\textquotedbl}
-Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}Rule RuleId={\textquotedbl}rule1{\textquotedbl}
+Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Actions{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Action
 Type={\textquotedbl}string{\textquotedbl}
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/operation{\textquotedbl}{\textgreater}CreateActivity{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Conditions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}Condition{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/operation{\textquotedbl}{\textgreater}CreateActivity{\textless}/Action{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/Actions{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Conditions{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Condition{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ {\textless}Attribute
 Type={\textquotedbl}string{\textquotedbl}
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace{\textquotedbl}{\textgreater}http://schemas.ggf.org/bes/2006/08/bes-factory{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}/Condition{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Conditions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
+AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace{\textquotedbl}{\textgreater}http://schemas.ggf.org/bes/2006/08/bes-factory{\textless}/Attribute{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Condition{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/Conditions{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}/Rule{\textgreater}}
 \item {\selectlanguage{english}\ttfamily\color{black}
@@
 \begin{enumerate}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Subjects{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Subject{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Subjects{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Subject{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}grantor=knowarc://testvoms.knowarc.eu:50000/knowarc:role=guest{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Attribute
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}grantor=knowarc://testvoms.knowarc.eu:50000/knowarc:role=guest{\textless}/Attribute{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Attribute
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/ca{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=CA{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=CA{\textless}/Attribute{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/identity{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=test{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Subject{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Subjects{\textgreater}}}
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=test{\textless}/Attribute{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}/Subject{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/Subjects{\textgreater}}
 \end{enumerate}
@@
 On the service side, the functionality of extracting UsernameToken may
-be configured as described in section 7.11. }
+be configured as described in section . }
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}Plugins
-overlay={\textquotedbl}add{\textquotedbl}{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}Chain{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}Component
-name={\textquotedbl}soap.client{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}SecHandler
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Plugins
+overlay={\textquotedbl}add{\textquotedbl}{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Chain{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ {\textless}Component
+name={\textquotedbl}soap.client{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}SecHandler
 name={\textquotedbl}usernametoken.handler{\textquotedbl}
 event={\textquotedbl}outgoing{\textquotedbl}
-overlay={\textquotedbl}add{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}Process{\textgreater}generate{\textless}/Process{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}Username{\textgreater}user{\textless}/Username{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}Password{\textgreater}pass{\textless}/Password{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}PasswordEncoding{\textgreater}digest{\textless}/PasswordEncoding{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}/SecHandler{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}/Component{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}/Chain{\textgreater}}}
+overlay={\textquotedbl}add{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Process{\textgreater}generate{\textless}/Process{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Username{\textgreater}user{\textless}/Username{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Password{\textgreater}pass{\textless}/Password{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}PasswordEncoding{\textgreater}digest{\textless}/PasswordEncoding{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}/SecHandler{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}/Chain{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}Plugins
-overlay={\textquotedbl}add{\textquotedbl}{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}Chain{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}Component
-name={\textquotedbl}soap.client{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}SecHandler
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Plugins
+overlay={\textquotedbl}add{\textquotedbl}{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Chain{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ {\textless}Component
+name={\textquotedbl}soap.client{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}SecHandler
 name={\textquotedbl}x509token.handler{\textquotedbl}
 event={\textquotedbl}outgoing{\textquotedbl}
-overlay={\textquotedbl}add{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}Process{\textgreater}generate{\textless}/Process{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}KeyPath{\textgreater}/path/to/key.pem{\textless}/KeyPath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}CertificatePath{\textgreater}/path/to/cert.pem{\textless}/CertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}/SecHandler{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}/Component{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}/Chain{\textgreater}{\textless}/ArcConfig{\textgreater}}}
+overlay={\textquotedbl}add{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Process{\textgreater}generate{\textless}/Process{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}KeyPath{\textgreater}/path/to/key.pem{\textless}/KeyPath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}CertificatePath{\textgreater}/path/to/cert.pem{\textless}/CertificatePath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}/SecHandler{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}/Chain{\textgreater}{\textless}/ArcConfig{\textgreater}}
 

### 2008-11-25 doc/sec/SecurityFrameworkofARC1.tex
Put the user manual away from security technical documentation

@@
 % List styles
-\newcounter{saveenum}
 \newcommand\liststyleWWviiiNumii{%
@@
 }
-\newcommand\liststyleWWviiiNumxi{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
-\newcommand\liststyleWWviiiNumxii{%
-\renewcommand\labelitemi{[F0B7?]}
-\renewcommand\labelitemii{[F0B7?]}
-\renewcommand\labelitemiii{[F0B7?]}
-\renewcommand\labelitemiv{[F0B7?]}
-}
-\newcommand\liststyleWWviiiNumxv{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
-\newcommand\liststyleWWviiiNumxvi{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
-\newcommand\liststyleWWviiiNumxxiv{%
-\renewcommand\theenumi{\arabic{enumi}}
-\renewcommand\theenumii{\arabic{enumii}}
-\renewcommand\theenumiii{\arabic{enumiii}}
-\renewcommand\theenumiv{\arabic{enumiv}}
-\renewcommand\labelenumi{\theenumi.}
-\renewcommand\labelenumii{\theenumii.}
-\renewcommand\labelenumiii{\theenumiii.}
-\renewcommand\labelenumiv{\theenumiv.}
-}
 % Page layout (geometry)
@@
 \begin{center}
-\includegraphics[width=1.4228in,height=0.9098in]{SecurityFrameworkofARC1-img1.jpg}
+\includegraphics[width=1.422in,height=0.9091in]{SecurityFrameworkofARC1-img1.jpg}
 \end{center}
@@
 \bigskip
-\clearpage\clearpage\setcounter{page}{1}\pagestyle{Convertiv}
+
+\bigskip
+\clearpage\clearpage\setcounter{page}{1}\pagestyle{Convertv}
 
@@
 \begin{center}
-\begin{minipage}{3.3756in}
+\begin{minipage}{3.3752in}
 {\selectlanguage{english}\color{black}
@@
 \begin{center}
-\begin{minipage}{3.8756in}
+\begin{minipage}{3.8752in}
 {\selectlanguage{english}\color{black}
@@
 
-\includegraphics[width=5.7043in,height=5.6354in]{SecurityFrameworkofARC1-img6.png}
+\includegraphics[width=5.9709in,height=6.3957in]{SecurityFrameworkofARC1-img6.png}
 
@@
 \selectlanguage{english}\bfseries\itshape\color{black}
-Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
+Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7698599in}|}
 \selectlanguage{english}\color{black} Resource &
@@
 \selectlanguage{english}\bfseries\itshape\color{black}
-Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7684599in}|}
+Content\\\hline}\begin{supertabular}{|m{1.2330599in}|m{2.5476599in}|m{1.7698599in}|}
 \selectlanguage{english}\color{black} SubjectAttribute &
@@
 and examples}
-\label{bkm:Ref204010074}\subsubsection[Authorization
-Policy]{Authorization Policy}
+\subsubsection[Authorization Policy]{Authorization Policy}
 \label{bkm:Ref204009564}{\selectlanguage{english}\upshape\color{black}
@@
 XML schema with comments available at
-\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/simplelistpdp/SimpleListPDP.xsd}
+\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/simplelistpdp/SimpleListPDP.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/simplelistpdp/SimpleListPDP.}}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/simplelistpdp/SimpleListPDP.xsd}{\textstyleInternetlink{xsd}}
 .}
@@
 configuration and Policy Example}
-\label{bkm:Ref204009974}\label{bkm:RefSecarcpdpconfigandpolicy}{\selectlanguage{english}\upshape\color{black}
+\label{bkm:Ref204009974}{\selectlanguage{english}\upshape\color{black}
 XML schema with comments available at
@@
 SecHandler Configuration}
-\label{bkm:RefSecdelegsechandleconfig}{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\upshape\color{black}
 Below is an example of configuration of Delegation SecHandler inside TLS
@@
 SecHandler configuration}
-\label{bkm:RefSecuserametokenconfig}{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\upshape\color{black}
 Configuration XML schema with comments available at
-\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTo}}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{kenSH.xsd}}
+\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}
 .}
@@
 
-{\selectlanguage{english}\color{black}
+{\selectlanguage{english}\upshape\color{black}
 X509Token SecHandler must be configured under SOAP MCC.}
@@
 \bigskip
-
-\subsection{User Manual}
-{\selectlanguage{english}\color{black}
-This section describes how to configure and use SecHandler and PDP
-elements included in the ARC1 and provides few examples of the ARC
-Policy documents. The target readers are those users who will use the
-ARC1 middleware. Currently this section is very short on details. It is
-going to be continuously extended. Especially taking user feedback into
-account.}
-
-\subsubsection[Authorization SecHandler and PDPs]{Authorization
-SecHandler and PDPs}
-{\selectlanguage{english}\color{black}
-There is a specific Authorization SecHandler (arc.authz) which is
-implemented for calling the Policy Decision Points (PDP) and serves as
-their container. }
-
-{\selectlanguage{english}\color{black}
-Usually the Authorization SecHandler and included PDPs are used on the
-service side of communication channel. Although it is also possible to
-use them on the client side. All possibilities are achieved by
-modifying the configuration file (hereafter mentioned as service.xml)
-and possibly providing the authorization policy in a separate file.}
-
-{\selectlanguage{english}\color{black}
-Here the {\textquotedblleft}echo{\textquotedblright} test service is
-used to explaining the usage, but the explanation applies to other
-services as well.}
-
-{\selectlanguage{english}\color{black}
-The procedure for configuring Authorization SecHandler in service.xml is
-following:}
-
-\liststyleWWviiiNumxi
-\begin{enumerate}
-\item {\selectlanguage{english}\color{black}
-Add the Authorization SecHandler as child element
-{\textless}SecHandler/{\textgreater} of
-{\textless}Service/{\textgreater} element. }
-\end{enumerate}
-{\selectlanguage{english}\color{black}
-The {\textquotedblleft}name{\textquotedblright} and
-{\textquotedblleft}event{\textquotedblright} attribute of
-{\textless}SecHandler/{\textgreater} element are both important. The
-{\textquotedblleft}name{\textquotedblright} attribute is used for
-distinguishing betwenn loaded SecHandler objects. The
-{\textquotedblleft}event{\textquotedblright} attribute defines for
-which message authorization would be enforced. Usually and reasonably
-it is done for {\textquotedblleft}incoming{\textquotedblright}
-messages. But some services and other Message Chain components may
-define other internal types of messages. For possible values please
-refer to documentation of particular Service or MCC. In our particular
-case {\textquotedblleft}echo{\textquotedblright} service only supports
-{\textquotedblleft}incoming{\textquotedblright} messages for this
-purpose.}
-
-\liststyleWWviiiNumxi
-\setcounter{saveenum}{\value{enumi}}
-\begin{enumerate}
-\setcounter{enumi}{\value{saveenum}}
-\item {\selectlanguage{english}\color{black}
-Add the PDP configuration as \ child element
-{\textless}PDP/{\textgreater} under
-{\textless}SecHandler/{\textgreater}. Currently there are four usable
-PDPs distributed as part of the ARC1 middleware:}
-\end{enumerate}
-\liststyleWWviiiNumxii
-\begin{itemize}
-\item {\selectlanguage{english}\color{black}
-simplelist.pdp -- compares Subject of user{\textquotesingle}s X509
-certificate to those stored in a file.}
-\item {\selectlanguage{english}\color{black}
-arc.pdp -- compares authorization related information parsed from
-message at various processing steps \ to Policy document specified in
-configuration of this PDP.}
-\item {\selectlanguage{english}\color{black}
-pdpservice.invoker - composes the ARC Request, puts request into SOAP
-message, and invokes the remote PDP service to get the response SOAP
-which includes authorization decision. The PDP service functionality is
-similar to arc.pdp.}
-\item {\selectlanguage{english}\color{black}
-delegation.pdp -- compares authorization related information parsed from
-message at various processing steps and Policy document embedded in
-proxy certificate used by remote side.}
-\end{itemize}
-{\selectlanguage{english}\color{black}
-Default behavior of Authorization SecHandler is to execute all
-configured PDPs sequentially till eithe r one of them fails or all
-produced positive results. This behavior may be modified by attribute
-{\textquotedblleft}action{\textquotedblright} of
-{\textless}PDP/{\textgreater} element.}
-
-{\selectlanguage{english}\upshape\color{black}
-The description of PDP configuration and ARC Policy example are
-available in Section 7}
-
-\subsubsection[Delegation SecHandler, Delegation PDP and Proxy
-Certificate Generation]{Delegation SecHandler, Delegation PDP and Proxy
-Certificate Generation}
-{\selectlanguage{english}\upshape\color{black}
-Delegation SecHandler and Delegation PDP in their current state provide
-an infrastructure for limiting capabilities of delegated credentials.
-Their collect and process policies attached to X509 Proxy Certificates
-respectively. Hence to have delegation restriction working both must be
-enabled in configuration of service. Configuration of Delegation
-SecHandler is described in section 7.10. }
-
-{\selectlanguage{english}\color{black}
-The possible location for Delegation PDP is inside Authorization
-SecHandler (arc.authz). Depending on how fine grained policy of
-delegated credentials is supposed to be corresponding Authrorization
-SecHandler may be attached to different MCCs or directly to Service
-component. However, the precondition for using Delegation PDP is that
-there must be Delegation SecHandler instantiated earlier in chain.}
-
-{\selectlanguage{english}\color{black}
-On the client side, command line utility
-{\textquotedblleft}approxy{\textquotedblright} utility can be used to
-generate Proxy Certificate with Delegation Policy embedded.}
-
-{\selectlanguage{english}\color{black}
-Normally approxy appears in \$ARC\_LOCATION/bin. The usage of approxy is
-like:}
-
-{\selectlanguage{english}\color{black}
-approxy -P proxy.pem -C cert.pem -K key.pem -c constraints}
-
-{\selectlanguage{english}\color{black}
-By using argument {\textquotedbl}-c{\textquotedbl}, some constraints can
-be specified for proxy certificate. Currently, the life time can be
-limited by using {\textquotedbl}-c validityStart=...{\textquotedbl} and
-{\textquotedbl}-c validityEnd=...{\textquotedbl}, {\textquotedbl}-c
-validityStart=...{\textquotedbl} and {\textquotedbl}-c
-validityPeriod=...{\textquotedbl}. Like for example }
-
-{\selectlanguage{english}\color{black}
-{}-c validityStart=2008-05-29T10:20:30Z}
-
-{\selectlanguage{english}\color{black}
-{}-c validityEnd=2008-06-29T10:20:30Z}
-
-{\selectlanguage{english}\color{black}
-The Delegation Policy can be specified by using {\textquotedbl}-c
-proxyPolicyFile=...{\textquotedbl} or {\textquotedbl}-c
-proxyPolicy=...{\textquotedbl}. Like }
-
-{\selectlanguage{english}\color{black}
-{}-c proxyPolicyFile=delegation\_policy.xml}
-
-{\selectlanguage{english}\upshape\color{black}
-The Delegation Policy is the same as the ARC Policy explained in section
-7.7. Simple example below renders delegated credentials usable only for
-contacting service attached to HTTP communication channel under path
-/arex (line 6) \ and allows HTTP operation POST (line 9) on it. }
-
-\liststyleWWviiiNumxv
-\begin{enumerate}
-\item {\selectlanguage{english}\ttfamily\color{black}
-{\textless}?xml version={\textquotedbl}1.0{\textquotedbl}
-encoding={\textquotedbl}UTF-8{\textquotedbl}?{\textgreater}}
-\item {\selectlanguage{english}\color{black}
-\texttt{{\textless}Policy
-xmlns={\textquotedbl}}\url{http://www.nordugrid.org/schemas/policy-arc}\texttt{{\textquotedbl}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ PolicyId={\textquotedbl}sm-example:policy1{\textquotedbl}
-CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}Rule RuleId={\textquotedbl}rule1{\textquotedbl}
-Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Resources{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Resource
-Type={\textquotedbl}string{\textquotedbl}
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedbl}{\textgreater}/arex{\textless}/Resource{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Resources{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
-Type={\textquotedbl}string{\textquotedbl}
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}{\textgreater}POST{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
-\item {\selectlanguage{english}\ttfamily\color{black}
-{\textless}/Policy{\textgreater}}
-\end{enumerate}
-{\selectlanguage{english}\color{black}
-Another example of delegation policy is presented below. This policy
-restricts usage of delegated credentials to SOAP operation
-CreateActivity (line 5) of Basic Execution Service (BES) [4] namespace
-(line 9). Such policy could be embedded into credentials delegated to
-high level Brokering service performing Grid job submission to low
-level BES on behalf of user.}
-
-\liststyleWWviiiNumxvi
-\begin{enumerate}
-\item {\selectlanguage{english}\ttfamily\color{black}
-{\textless}?xml version={\textquotedbl}1.0{\textquotedbl}
-encoding={\textquotedbl}UTF-8{\textquotedbl}?{\textgreater}}
-\item {\selectlanguage{english}\ttfamily\color{black}
-{\textless}Policy
-xmlns={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc{\textquotedbl}
-PolicyId={\textquotedbl}sm-example:policy1{\textquotedbl}
-CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}Rule RuleId={\textquotedbl}rule1{\textquotedbl}
-Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
-Type={\textquotedbl}string{\textquotedbl}
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/operation{\textquotedbl}{\textgreater}CreateActivity{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Conditions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}Condition{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ {\textless}Attribute
-Type={\textquotedbl}string{\textquotedbl}
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace{\textquotedbl}{\textgreater}http://schemas.ggf.org/bes/2006/08/bes-factory{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}/Condition{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Conditions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
-\item {\selectlanguage{english}\ttfamily\color{black}
-{\textless}/Policy{\textgreater}}
-\end{enumerate}
-\subsubsection{VOMS Proxy Certificate}
-{\selectlanguage{english}\color{black}
-The commonly used voms proxy certificate can be used in ARC1 for
-authentication as normal proxy certificate, and attribute acquiring in
-order to make authorization decision based on attributes.}
-
-{\selectlanguage{english}\color{black}
-1.How to create voms proxy certificate}
-
-{\selectlanguage{english}\color{black}
-Currently the proxy creation utility in ARC1---arcproxy can not be used
-for creating voms proxy certificate because the GSI-based communication
-requirement from voms server side has not been supported in ARC1 (ARC1
-uses standard SSL/TLS communication). So the way to create voms proxy
-certificate is still to use the
-{\textquotedblleft}voms-proxy-init{\textquotedblright} utility.}
-
-{\selectlanguage{english}\color{black}
-2.How to use voms proxy certificate in ARC1}
-
-{\selectlanguage{english}\color{black}
-The attribute certificate (AC, which is created by voms server and then
-embedded in voms proxy certificate \ as one of the certificate
-extension by {\textquotedblleft}voms-proxy-init{\textquotedblright}
-client) \ will be parsed by TLS handling plugin (called MCCTLS) of
-ARC1, and saved in the message context in the format
-{\textquotedblleft}grantor=knowarc://testvoms.knowarc.eu:50000/knowarc:role=guest{\textquotedblright}.}
-
-{\selectlanguage{english}\color{black}
-Service administrator needs to configure the trusted DN chain for
-verifying the AC under tls MCC{\textquotesingle}s configurationwith the
-following format. You should specify one node
-({\textless}tls:VOMSCertTrustDNChain/{\textgreater}) for each vo which
-you trust. The reason why using trusted DN chain is for service to
-restrict which vo server is trusted. }
-
-{\selectlanguage{english}\color{black}
-In the following table, the first
-{\textless}tls:VOMSCertTrustDNChain{\textgreater} in the following
-table defines two items: the first one is the DN of voms server, the
-second line is the DN of the corresponding CA. The second
-{\textless}tls:VOMSCertTrustDNChain{\textgreater} defines regular
-expression for matching, the DN of voms server and DN of the
-corresponding CA should both match it. The third one defines the
-external file which includes the
-{\textless}tls:VOMSCertTrustDNChain{\textgreater}.}
-
-{\selectlanguage{english}\color{black}
-\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChain{\textgreater}}
-
-\bigskip
-
-{\selectlanguage{english}\color{black}
-\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDN{\textgreater}/O=Grid/O=NorduGrid/OU=fys.uio.no/CN=ABCDE{\textless}/tls:VOMSCertTrustDN{\textgreater}}
-
-\bigskip
-
-{\selectlanguage{english}\color{black}
-\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDN{\textgreater}/O=Grid/O=NorduGrid/CN=NorduGrid
-Certification Authority{\textless}/tls:VOMSCertTrustDN{\textgreater}}
-
-\bigskip
-
-{\selectlanguage{english}\color{black}
-\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/tls:VOMSCertTrustDNChain{\textgreater}}
-
-\bigskip
-
-{\selectlanguage{english}\color{black}
-\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChain{\textgreater}}
-
-\bigskip
-
-{\selectlanguage{english}\color{black}
-\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustRegex{\textgreater}\^{}/O=Grid/O=NorduGrid{\textless}/tls:VOMSCertTrustRegex{\textgreater}}
-
-\bigskip
-
-{\selectlanguage{english}\color{black}
-\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/tls:VOMSCertTrustDNChain{\textgreater}}
-
-\bigskip
-
-{\selectlanguage{english}\color{black}
-\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChainsLocation{\textgreater}./pathto/external/file{\textless}/tls:VOMSCertTrustDNChainsLocation{\textgreater}}
-
-\bigskip
-
-{\selectlanguage{english}\color{black}
-Service administrator can then use the attribute into access control
-policy:}
-
-\liststyleWWviiiNumxxiv
-\begin{enumerate}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Subjects{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Subject{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}grantor=knowarc://testvoms.knowarc.eu:50000/knowarc:role=guest{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Attribute
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/ca{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=CA{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
-AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/identity{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=test{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Subject{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Subjects{\textgreater}}}
-\end{enumerate}
-
-\bigskip
-
-\subsubsection[UsernameToken SecHandler]{UsernameToken SecHandler}
-{\selectlanguage{english}\color{black}
-The UsernameToken SecHandler is meant for processing - generating and
-extracting - WS-Security [5] UsernameToken from SOAP header. Hence it
-must be attached to SOAP MCC of service or/and client communication
-channel. }
-
-{\selectlanguage{english}\upshape\color{black}
-On the service side, the functionality of extracting UsernameToken may
-be configured as described in section 7.11. }
-
-{\selectlanguage{english}\color{black}
-On the client side, the UsernameToken SecHandler may be configured
-either by using client specific methods (for example see
-test\_clientinterface.cpp src/tests/echo directory of source tree) or
-through generic client configuration file as shown in example below
-(also can see share/doc/arc/echo\_client.xml.example under the
-installation directory). This example will generate token with username
-{\textquotedblleft}user{\textquotedblright} and password
-{\textquotedblleft}pass{\textquotedblright} inside any SOAP message
-sent by client tools of the ARC1.}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}ArcConfig{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}Plugins
-overlay={\textquotedbl}add{\textquotedbl}{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}Chain{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}Component
-name={\textquotedbl}soap.client{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}SecHandler
-name={\textquotedbl}usernametoken.handler{\textquotedbl}
-event={\textquotedbl}outgoing{\textquotedbl}
-overlay={\textquotedbl}add{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}Process{\textgreater}generate{\textless}/Process{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}Username{\textgreater}user{\textless}/Username{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}Password{\textgreater}pass{\textless}/Password{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}PasswordEncoding{\textgreater}digest{\textless}/PasswordEncoding{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}/SecHandler{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}/Component{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}/Chain{\textgreater}}}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}/ArcConfig{\textgreater}}
-
-\subsubsection[X509Token SecHandler]{X509Token SecHandler}
-{\selectlanguage{english}\color{black}
-The X509Token SecHandler is meant for processing - generating and
-extracting - WS-Security [5] X509Token from SOAP header. Hence it must
-be attached to SOAP MCC of service or/and client communication channel.
-}
-
-{\selectlanguage{english}\color{black}
-On the service side, the functionality of extracting X509Token may be
-configured as described in section 7.12. }
-
-{\selectlanguage{english}\color{black}
-On the client side, the X509Token SecHandler may be configured either by
-using client specific methods (for example see
-test\_clientinterface.cpp src/tests/echo directory of source tree) or
-through generic client configuration file as shown in example below.
-This example will generate x509 token (by using the specified
-certificate and key file) inside any SOAP message sent by client tools
-of the ARC1.}
-
-{\selectlanguage{english}\ttfamily\color{black}
-{\textless}ArcConfig{\textgreater}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}Plugins
-overlay={\textquotedbl}add{\textquotedbl}{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}Chain{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}Component
-name={\textquotedbl}soap.client{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}SecHandler
-name={\textquotedbl}x509token.handler{\textquotedbl}
-event={\textquotedbl}outgoing{\textquotedbl}
-overlay={\textquotedbl}add{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}Process{\textgreater}generate{\textless}/Process{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}KeyPath{\textgreater}/path/to/key.pem{\textless}/KeyPath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}CertificatePath{\textgreater}/path/to/cert.pem{\textless}/CertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}/SecHandler{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}/Component{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ {\textless}/Chain{\textgreater}{\textless}/ArcConfig{\textgreater}}}
 

### 2008-11-25 doc/sec/ManualofSecurityFramework.tex
Put the slcs service/client into user manual

@@
 
+\subsubsection[SLCS service and client]{SLCS service and client}
+{\selectlanguage{english}\color{black}
+Short-lived credential service (SLCS) is for signing short-lived x509
+credential for user based user{\textquotesingle}s username/password
+credential. Then the user can use the short-lived x509 credential to
+access grid services/resources where x509 credential is required by
+default. \ SLCS service should depend on the SP (Service provider)
+service which is one of the participants of SAML2 SSO profile (SAML2
+SSO profile is used for authenticating based on username/password
+credential and getting SAML authentication assertion; and SAML
+authentication assertion is then used as basis for signing short-lived
+x509 credential).}
+
+{\selectlanguage{english}\color{black}
+a. \textbf{SLCS Service} }
+
+{\selectlanguage{english}\color{black}
+On the SLCS service side, a typical configuration should be like the
+following.}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}?xml
+version={\textquotedbl}1.0{\textquotedbl}?{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}ArcConfig}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns={\textquotedbl}http://www.nordugrid.org/schemas/ArcConfig/2007{\textquotedbl}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns:tcp={\textquotedbl}http://www.nordugrid.org/schemas/ArcMCCTCP/2007{\textquotedbl}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns:tls={\textquotedbl}http://www.nordugrid.org/schemas/ArcMCCTLS/2007{\textquotedbl}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns:arcpdp={\textquotedbl}http://www.nordugrid.org/schemas/ArcPDP{\textquotedbl}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns:slcs={\textquotedbl}http://www.nordugrid.org/schemas/ArcConfig/2007/slcs{\textquotedbl}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+ \ \ \ {\textless}ModuleManager{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}.libs/{\textless}/Path{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/mcc/http/.libs/{\textless}/Path{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/mcc/tls/.libs/{\textless}/Path{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/mcc/soap/.libs/{\textless}/Path{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/mcc/tcp/.libs/{\textless}/Path{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/pdc/.libs/{\textless}/Path{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../services/saml/.libs/{\textless}/Path{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../services/slcs/.libs/{\textless}/Path{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}/ModuleManager{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}mcctcp{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}mcctls{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}mcchttp{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}mccsoap{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}saml2sp{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}slcs{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Chain{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}tcp.service{\textquotedbl}
+id={\textquotedbl}tcp{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}tls{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tcp:Listen{\textgreater}{\textless}tcp:Port{\textgreater}60000{\textless}/tcp:Port{\textgreater}{\textless}/tcp:Listen{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}tls.service{\textquotedbl}
+id={\textquotedbl}tls{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}http{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:KeyPath{\textgreater}./testkey-nopass.pem{\textless}/tls:KeyPath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CertificatePath{\textgreater}./testcert.pem{\textless}/tls:CertificatePath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}-tls:CACertificatePath{\textgreater}./cacert.pem{\textless}/tls:CACertificatePath-{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:ClientAuthn{\textgreater}false{\textless}/tls:ClientAuthn{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}http.service{\textquotedbl}
+id={\textquotedbl}http{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}plexer{\textquotedbl}{\textgreater}POST{\textless}/next{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Plexer
+name={\textquotedbl}plexer.service{\textquotedbl}
+id={\textquotedbl}plexer{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}samlsp{\textquotedbl}{\textgreater}/saml2sp{\textless}/next{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}soap{\textquotedbl}{\textgreater}/slcs{\textless}/next{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Plexer{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}soap.service{\textquotedbl}
+id={\textquotedbl}soap{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}slcs{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Service
+name={\textquotesingle}saml.sp{\textquotesingle}
+id={\textquotesingle}samlsp{\textquotesingle}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}MetaDataLocation{\textgreater}./test\_metadata.xml{\textless}/MetaDataLocation{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}ServiceProviderName{\textgreater}https://squark.uio.no/shibboleth-sp{\textless}/ServiceProviderName{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Service{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Service
+name={\textquotedbl}slcs.service{\textquotedbl}
+id={\textquotedbl}slcs{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}slcs{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CACertificate{\textgreater}./CAcert.pem{\textless}/slcs:CACertificate{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CAKey{\textgreater}./CAkey.pem{\textless}/slcs:CAKey{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CASerial{\textgreater}./CAserial{\textless}/slcs:CASerial{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Service{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}/Chain{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}/ArcConfig{\textgreater}}
+
+{\selectlanguage{english}\color{black}
+SLCS service and SP service should together be configured by using the
+Plexer. SP service is directly based on http and SLCS is directly based
+on soap. }
+
+{\selectlanguage{english}\color{black}
+SLCS service should specifically include the CA credential (certificate
+and key file) and the serial number file (which includes the serial
+number of each signed certificate).}
+
+{\selectlanguage{english}\color{black}
+\textbf{Note}: When deploying the SLCS service, service administrator
+should deploy a dedicated IdP (Identity Provider) which can be assigned
+as authentication/attribute authority, or the users should have already
+had his own IdP. And the IdP information (authentication URL and
+attribute authority URL) should have already been included into the
+metadata of above configuration (e.g. test\_metadata.xml). On the other
+hand, the SP (SP service) information (assertion consuming URL) should
+also have been included into the metadata of the IdP.}
+
+{\selectlanguage{english}\color{black}
+After the whole process, user only needs to authentication with his
+existing IdP and can get back the x509 credential.}
+
+{\selectlanguage{english}\color{black}
+Shibboleth IdP (http://shibboleth.internet2.edu) is used for the current
+solution of IdP, since it is widely deployed for other AAI
+(Authentication and Authority Infrastructure). There is a test IdP
+deployed in
+\href{https://squark.uio.no:8443/}{https://squark.uio.no:8443}
+(idpname: https://squark.uio.no/idp/shibboleth), you can use the SP
+from \url{https://sp.testshib.org/} to test the validity of the test
+IdP. And also you can installed your own IdP which is supposed to
+authenticates the users from your own organization.}
+
+{\selectlanguage{english}\color{black}
+b. \textbf{SLCS client}}
+
+{\selectlanguage{english}\color{black}
+On the SLCS client side, there is a client utility called
+{\textquotedblleft}arcslcs{\textquotedblright}. \ The command option
+for arcslcs is as following:}
+
+{\selectlanguage{english}\color{black}
+Application Options:}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-S, -{}-url=url \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ URL of SLCS service}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-I, -{}-idp=string \ \ \ \ \ \ \ \ \ \ \ \ IdP name}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-U, -{}-user=string \ \ \ \ \ \ \ \ \ \ \ User account to IdP}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-P, -{}-password=string \ \ \ \ \ \ \ Password for user account to
+IdP}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-Z, -{}-keysize=number \ \ \ \ \ \ \ \ Key size of the private key
+(512, 1024, 2048)}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-K, -{}-keypass=passphrase \ \ \ \ Private key passphrase}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-L, -{}-lifetime=period \ \ \ \ \ \ \ Lifetime of the certificate,
+start with current time, hour as unit}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-D, -{}-storedir=directory \ \ \ \ Store directory for key and
+signed certificate}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-c, -{}-conffile=filename \ \ \ \ \ configuration file (default
+\~{}/.arc/client.xml)}
+
+{\selectlanguage{english}\color{black}
+An example is:}
+
+{\selectlanguage{english}\color{black}
+./arcslcs -S https://127.0.0.1:60000/slcs -I
+https://squark.uio.no/idp/shibboleth -U root -P aa1122 -D
+/home/wzqiang/arc-0.9/src/clients/credentials -c client.xml}
+
+{\selectlanguage{english}\color{black}
+Note user should input the {\textquotedblleft}IdP
+name{\textquotedblright} is the corresponding Identity Provider (one of
+the participants of SAML2 SSO profile) name to which the user would
+authentication against by using its username/password credential. And
+the name is stored inside the metadata on both SP service and IdP
+provider, and it is used by SP service to get the authentication URL
+for username/password based authentication.}
+
+{\selectlanguage{english}\color{black}
+And the {\textquotedblleft}user{\textquotedblright} and
+{\textquotedblleft}password{\textquotedblright} is the credential which
+will be used to authentication against IdP.}
+
+{\selectlanguage{english}\color{black}
+The {\textquotedblleft}conffile{\textquotedblright} can be another
+option for all of the above options, see the following as an example:}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}ArcConfig{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}!-{}- change the paths below to the location of your user
+certs -{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}!-{}- here only trusted certificates are set. Because this
+arcslcs is}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ not supposed to have x509 credential before the samlsso profile has
+been}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ (by using client{\textquotesingle}s username/password credential)
+passed and the SLCS}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ service succeeded to repond this client with the signed x509
+credential.}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ On the SLCS service{\textquotesingle}s configuration: ClientAuthn
+should be off.-{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}!-{}-KeyPath{\textgreater}./testkey-nopass.pem{\textless}/KeyPath-{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}!-{}-CertificatePath{\textgreater}./testcert.pem{\textless}/CertificatePath-{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}CACertificatesDir{\textgreater}./certificates{\textless}/CACertificatesDir{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}SLCSURL{\textgreater}https://127.0.0.1:60000/slcs{\textless}/SLCSURL{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}IdPname{\textgreater}https://squark.uio.no/idp/shibboleth{\textless}/IdPname{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Username{\textgreater}root{\textless}/Username{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Password{\textgreater}aa1122{\textless}/Password{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Keysize{\textgreater}1024{\textless}/Keysize{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Keypass{\textgreater}123456{\textless}/Keypass{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}CertLifetime{\textgreater}24{\textless}/CertLifetime{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}StoreDir{\textgreater}./{\textless}/StoreDir{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Debug{\textgreater}INFO{\textless}/Debug{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}/ArcConfig{\textgreater}}
+
+\bigskip
+
+\bigskip
+
+\bigskip
+
 \clearpage{\centering\selectlanguage{english}\bfseries\scshape\color{black}

### 2008-11-27 doc/sec/ManualofSecurityFramework.tex
Description about saml sp service in manual

--- 2008-11-25 doc/sec/ManualofSecurityFramework.tex
@@
 
+\subsubsection{SP Service and SAML2SSO Service Provider Handler}
+{\selectlanguage{english}\color{black}
+Service Provider servicer (SP service) implements the functionality of
+Service Provider in SAML2 Web SSO profile. }
+
+{\selectlanguage{english}\color{black}
+Normally SP service will not be deployed independently, instead, it
+should be deployed together with other services. The SLCS service is
+the typical deployment about SP service. On the client side, client
+developer should use the ClientSAML2Interface instead of
+ClientInterface to call the client functionality. The SLCS client
+(arcslcs) is the typical usage of \ ClientSAML2Interface.}
+
+{\selectlanguage{english}\color{black}
+Once the SP Service has cooperated with user agent and Identity Provider
+(external) and succeeded to accomplish the SAML2 SSO profile, \ the SP
+service will get the saml authentication assertion which asserts that
+the authentication has succeeded, and then SP service will store this
+assertion into session context.}
+
+{\selectlanguage{english}\color{black}
+SAML2 service provider handler (saml2ssoserviceprovider.handler) is the
+security handler which will understand the authentication assertion and
+make authorization decision according to this assertion.}
+
+{\selectlanguage{english}\color{black}
+The example configuration about using SP service and SAML2SSO service
+provider handler can be seen in the configuration of SLCS service. }
+
+{\selectlanguage{english}\color{black}
+SP service is not only supposed to work together with SLCS service. But
+if it is used to work together with other services, the client should
+be developed on \ ClientSAML2Interface instead of ClientInterface.}
+
+\subsubsection{SLCS service and client}
+{\selectlanguage{english}\color{black}
+Short-lived credential service (SLCS) is for signing short-lived x509
+credential for user based user{\textquotesingle}s username/password
+credential. Then the user can use the short-lived x509 credential to
+access grid services/resources where x509 credential is required by
+default. \ SLCS service should depend on the SP (Service provider)
+service which is one of the participants of SAML2 SSO profile (SAML2
+SSO profile is used for authenticating based on username/password
+credential and getting SAML authentication assertion; and SAML
+authentication assertion is then used as basis for signing short-lived
+x509 credential).}
+
+{\selectlanguage{english}\color{black}
+a. \textbf{SLCS Service} }
+
+{\selectlanguage{english}\color{black}
+On the SLCS service side, a typical configuration should be like the
+following.}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}?xml version={\textquotedbl}1.0{\textquotedbl}?{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}ArcConfig }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns={\textquotedbl}http://www.nordugrid.org/schemas/ArcConfig/2007{\textquotedbl}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns:tcp={\textquotedbl}http://www.nordugrid.org/schemas/ArcMCCTCP/2007{\textquotedbl}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns:tls={\textquotedbl}http://www.nordugrid.org/schemas/ArcMCCTLS/2007{\textquotedbl}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns:arcpdp={\textquotedbl}http://www.nordugrid.org/schemas/ArcPDP{\textquotedbl}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ xmlns:slcs={\textquotedbl}http://www.nordugrid.org/schemas/ArcConfig/2007/slcs{\textquotedbl}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}ModuleManager{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}.libs/{\textless}/Path{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/mcc/http/.libs/{\textless}/Path{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/mcc/tls/.libs/{\textless}/Path{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/mcc/soap/.libs/{\textless}/Path{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/mcc/tcp/.libs/{\textless}/Path{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../hed/pdc/.libs/{\textless}/Path{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../services/saml/.libs/{\textless}/Path{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Path{\textgreater}../../services/slcs/.libs/{\textless}/Path{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}/ModuleManager{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}mcctcp{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}mcctls{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}mcchttp{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}mccsoap{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}arcpdc{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}saml2sp{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Plugins{\textgreater}{\textless}Name{\textgreater}slcs{\textless}/Name{\textgreater}{\textless}/Plugins{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}Chain{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}tcp.service{\textquotedbl}
+id={\textquotedbl}tcp{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}tls{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tcp:Listen{\textgreater}{\textless}tcp:Port{\textgreater}60000{\textless}/tcp:Port{\textgreater}{\textless}/tcp:Listen{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}tls.service{\textquotedbl}
+id={\textquotedbl}tls{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}http{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:KeyPath{\textgreater}./testkey-nopass.pem{\textless}/tls:KeyPath{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CertificatePath{\textgreater}./testcert.pem{\textless}/tls:CertificatePath{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}-tls:CACertificatePath{\textgreater}./cacert.pem{\textless}/tls:CACertificatePath-{}-{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:ClientAuthn{\textgreater}false{\textless}/tls:ClientAuthn{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}http.service{\textquotedbl}
+id={\textquotedbl}http{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}plexer{\textquotedbl}{\textgreater}POST{\textless}/next{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Plexer
+name={\textquotedbl}plexer.service{\textquotedbl}
+id={\textquotedbl}plexer{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}samlsp{\textquotedbl}{\textgreater}/saml2sp{\textless}/next{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}soap{\textquotedbl}{\textgreater}/slcs{\textless}/next{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Plexer{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}soap.service{\textquotedbl}
+id={\textquotedbl}soap{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}slcs{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}SecHandler
+name={\textquotedbl}\textcolor[rgb]{0.8627451,0.13725491,0.0}{saml2ssoserviceprovider.handler}{\textquotedbl}
+id={\textquotedbl}saml2ssosp{\textquotedbl}
+event={\textquotedbl}incoming{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}/Component{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Service
+name={\textquotesingle}\textcolor[rgb]{0.77254903,0.0,0.043137256}{saml.sp}{\textquotesingle}
+id={\textquotesingle}samlsp{\textquotesingle}{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}MetaDataLocation{\textgreater}./test\_metadata.xml{\textless}/MetaDataLocation{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}ServiceProviderName{\textgreater}https://squark.uio.no/shibboleth-sp{\textless}/ServiceProviderName{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Service{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Service
+name={\textquotedbl}\textcolor[rgb]{0.8627451,0.13725491,0.0}{slcs.service}{\textquotedbl}
+id={\textquotedbl}slcs{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}slcs{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CACertificate{\textgreater}./CAcert.pem{\textless}/slcs:CACertificate{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CAKey{\textgreater}./CAkey.pem{\textless}/slcs:CAKey{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CASerial{\textgreater}./CAserial{\textless}/slcs:CASerial{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Service{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}/Chain{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}/ArcConfig{\textgreater} }
+
+{\selectlanguage{english}\color{black}
+SLCS service and SP service should together be configured by using the
+Plexer. SP service is directly based on http and SLCS is directly based
+on soap. }
+
+{\selectlanguage{english}\color{black}
+SLCS service should specifically include the CA credential (certificate
+and key file) and the serial number file (which includes the serial
+number of each signed certificate).}
+
+{\selectlanguage{english}\color{black}
+\textbf{Note}: When deploying the SLCS service, service administrator
+should deploy a dedicated IdP (Identity Provider) which can be assigned
+as authentication/attribute authority, or the users should have already
+had his own IdP. And the IdP information (authentication URL and
+attribute authority URL) should have already been included into the
+metadata of above configuration (e.g. test\_metadata.xml). On the other
+hand, the SP (SP service) information (assertion consuming URL) should
+also have been included into the metadata of the IdP.}
+
+{\selectlanguage{english}\color{black}
+After the whole process, user only needs to authentication with his
+existing IdP and can get back the x509 credential.}
+
+{\selectlanguage{english}\color{black}
+Shibboleth IdP (http://shibboleth.internet2.edu) is used for the current
+solution of IdP, since it is widely deployed for other AAI
+(Authentication and Authority Infrastructure). There is a test IdP
+deployed in
+\href{https://squark.uio.no:8443/}{https://squark.uio.no:8443}
+(idpname: https://squark.uio.no/idp/shibboleth), you can use the SP
+from \url{https://sp.testshib.org/} to test the validity of the test
+IdP. And also you can installed your own IdP which is supposed to
+authenticates the users from your own organization.}
+
+{\selectlanguage{english}\color{black}
+b. \textbf{SLCS client}}
+
+{\selectlanguage{english}\color{black}
+On the SLCS client side, there is a client utility called
+{\textquotedblleft}arcslcs{\textquotedblright}. \ The command option
+for arcslcs is as following:}
+
+{\selectlanguage{english}\color{black}
+Application Options: }
+
+{\selectlanguage{english}\color{black}
+\ \ {}-S, -{}-url=url \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ URL of SLCS service
+}
+
+{\selectlanguage{english}\color{black}
+\ \ {}-I, -{}-idp=string \ \ \ \ \ \ \ \ \ \ \ \ IdP name }
+
+{\selectlanguage{english}\color{black}
+\ \ {}-U, -{}-user=string \ \ \ \ \ \ \ \ \ \ \ User account to IdP }
+
+{\selectlanguage{english}\color{black}
+\ \ {}-P, -{}-password=string \ \ \ \ \ \ \ Password for user account to
+IdP }
+
+{\selectlanguage{english}\color{black}
+\ \ {}-Z, -{}-keysize=number \ \ \ \ \ \ \ \ Key size of the private key
+(512, 1024, 2048) }
+
+{\selectlanguage{english}\color{black}
+\ \ {}-K, -{}-keypass=passphrase \ \ \ \ Private key passphrase }
+
+{\selectlanguage{english}\color{black}
+\ \ {}-L, -{}-lifetime=period \ \ \ \ \ \ \ Lifetime of the certificate,
+start with current time, hour as unit }
+
+{\selectlanguage{english}\color{black}
+\ \ {}-D, -{}-storedir=directory \ \ \ \ Store directory for key and
+signed certificate }
+
+{\selectlanguage{english}\color{black}
+\ \ {}-c, -{}-conffile=filename \ \ \ \ \ configuration file (default
+\~{}/.arc/client.xml) }
+
+{\selectlanguage{english}\color{black}
+An example is:}
+
+{\selectlanguage{english}\color{black}
+./arcslcs -S https://127.0.0.1:60000/slcs -I
+https://squark.uio.no/idp/shibboleth -U root -P aa1122 -D
+/home/wzqiang/arc-0.9/src/clients/credentials -c client.xml }
+
+{\selectlanguage{english}\color{black}
+Note user should input the {\textquotedblleft}IdP
+name{\textquotedblright} is the corresponding Identity Provider (one of
+the participants of SAML2 SSO profile) name to which the user would
+authentication against by using its username/password credential. And
+the name is stored inside the metadata on both SP service and IdP
+provider, and it is used by SP service to get the authentication URL
+for username/password based authentication.}
+
+{\selectlanguage{english}\color{black}
+And the {\textquotedblleft}user{\textquotedblright} and
+{\textquotedblleft}password{\textquotedblright} is the credential which
+will be used to authentication against IdP.}
+
+{\selectlanguage{english}\color{black}
+The {\textquotedblleft}conffile{\textquotedblright} can be another
+option for all of the above options, see the following as an example:}
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}ArcConfig{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}!-{}- change the paths below to the location of your user
+certs -{}-{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}!-{}- here only trusted certificates are set. Because this
+arcslcs is }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ not supposed to have x509 credential before the samlsso profile has
+been }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ (by using client{\textquotesingle}s username/password credential)
+passed and the SLCS }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ service succeeded to repond this client with the signed x509
+credential. }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ On the SLCS service{\textquotesingle}s configuration: ClientAuthn
+should be off.-{}-{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}!-{}-KeyPath{\textgreater}./testkey-nopass.pem{\textless}/KeyPath-{}-{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}!-{}-CertificatePath{\textgreater}./testcert.pem{\textless}/CertificatePath-{}-{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}CACertificatesDir{\textgreater}./certificates{\textless}/CACertificatesDir{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}SLCSURL{\textgreater}https://127.0.0.1:60000/slcs{\textless}/SLCSURL{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}IdPname{\textgreater}https://squark.uio.no/idp/shibboleth{\textless}/IdPname{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Username{\textgreater}root{\textless}/Username{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Password{\textgreater}aa1122{\textless}/Password{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Keysize{\textgreater}1024{\textless}/Keysize{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Keypass{\textgreater}123456{\textless}/Keypass{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}CertLifetime{\textgreater}24{\textless}/CertLifetime{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}StoreDir{\textgreater}./{\textless}/StoreDir{\textgreater}
+}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ {\textless}Debug{\textgreater}INFO{\textless}/Debug{\textgreater} }
+
+{\selectlanguage{english}\ttfamily\color{black}
+{\textless}/ArcConfig{\textgreater} }
+
+\bigskip
+
+\bigskip
+
+\bigskip
+
 \clearpage{\centering\selectlanguage{english}\bfseries\scshape\color{black}

### 2008-12-01 doc/sec/Sec_paper.tex
start sec paper

% This file was converted to LaTeX by Writer2LaTeX ver. 0.5
% see http://www.hj-gym.dk/~hj/writer2latex for more info
\documentclass{article}
\usepackage[ascii]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsfonts,textcomp}
\usepackage{color}
\usepackage{array}
\usepackage{hhline}
\usepackage{hyperref}
\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=qiangwz, pdfsubject=, pdfkeywords=}
\usepackage[pdftex]{graphicx}
% List styles
\newcommand\liststyleLi{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
% Page layout (geometry)
\setlength\paperwidth{8.5in}
\setlength\paperheight{11in}
\setlength\voffset{-1in}
\setlength\hoffset{-1in}
\setlength\topmargin{0.7874in}
\setlength\oddsidemargin{0.7874in}
\setlength\textheight{9.4251995in}
\setlength\textwidth{6.9251995in}
\setlength\footskip{0.0cm}
\setlength\headheight{0cm}
\setlength\headsep{0cm}
% Footnote rule
\setlength{\skip\footins}{0.0469in}
\renewcommand\footnoterule{\vspace*{-0.0071in}\setlength\leftskip{0pt}\setlength\rightskip{0pt plus 1fil}\noindent\textcolor{black}{\rule{0.25\columnwidth}{0.0071in}}\vspace*{0.0398in}}
% Pages styles
\makeatletter
\newcommand\ps@Standard{
  \renewcommand\@oddhead{}
  \renewcommand\@evenhead{}
  \renewcommand\@oddfoot{}
  \renewcommand\@evenfoot{}
  \renewcommand\thepage{\arabic{page}}
}
\makeatother
\pagestyle{Standard}
\title{}
\begin{document}
{\centering
Towards cross-domain authentication and standardized attribute-based
authorization for ARC grid middleware
\par}

\bigskip

Abstract-{}-

\bigskip

\bigskip

\bigskip

\liststyleLi
\begin{enumerate}
\item {\centering
Introduction
\par}

\bigskip
\end{enumerate}

\bigskip

\bigskip

{\centering
2. ARC grid middleware 
\par}

ARC(Advanced Resource Connector) is an open source grid middleware
solution released under GPL licence. ARC middleware initially aimed at
developing a grid middleware which provides characteristics such as
self-organized, fault-tolerant, non-intrusive, easy-management [fgcs
ARC paper]. \ Now \ the classic version of ARC provides grid services
such as grid job submission and management, resource characterization,
resource aggregation and discovery, basic data management, integration
of grid security solution, etc. Classic ARC has been deployed and used
in production environment, and been one of the mainly deployed grid
middlewares in Europe.

The new generation of ARC is funded by KnowARC project, and based on the
functionality and capabilities of the classic ARC middleware, it aims
at implementing a service-oriented middleware which will provide higher
levels of resource and user abstracting through well-defined web
service interface [Design document] in order to provide
interoperability with other service-based grid middlewares, as well as
other Web Service compatible applications.

As the key part of the implementation of new ARC middleware, there is a
lightweight web service container called HED (Hosting Environment
Daemon) which provides a host place for various services in application
level, as well as a host place for a bunch of components to support
flexible, interoperatible, and efficient communication mechanism. The
whole design of the HED is built around the idea of flexibility and
modularity, which means for the developer, he can easily concentrate on
the application level Web Service implementation by only using the core
minimum amount of components if he only would do application level
development, or he can easily concentrate on the middleware level
implementation such as supporting another communication protocol or
implementing authentication mechanism by using the core minimum amount
of components and external dependencies; also for the deployer, he can
easily configure and deploy the middleware and application for
different kinds of requirements without being bothered to know much
about the implementation.

{\centering
Figure 1. The architecture of Host Environment Daemon
\par}

\begin{center}
\includegraphics[width=4.0673in,height=3.1047in]{Secpaper-img1.png}
\end{center}
The architecture of the Host Environment Daemon is illustrated as Figure
1. In general, there are a few \ components called MCC (Message Chain
Components) which are in charge of processing different messages for
different protocol levels. For instance, as shown in the example
message flow, HTTP MCC will process secure socket stream from TLS MCC
to generate output HTTP message to SOAP MCC, and also process SOAP
documentation from SOAP MCC to generate HTTP message to TLS MCC.

Two examples of MCC components configuration in Figure 1 shows MCC are
configurable: real line shows SOAP over HTTPS (HTTP over TLS), while
dotted line shows SOAP directly over TCP. Service administrator can
configure the MCCs according to his requirement, such as the
interoperability requirement with the other part. For instance, the
later configuration is compatible to WSE (Web Services Enhancement for
.NET) {\textquotesingle}s SOAP message mechanism (see
WSE{\textquotesingle}s \textit{SoapSender} and \textit{SoapReceiver}).
Another configuration could be SOAP over HTTPG (HTTP over GSI) which
can be user to interoperate with some service based on HTTPG such as
SRM (Storage Resource Manager) [ref] service. The easily configurable
characteristics shows the flexibility of HED in terms of protocol
supporting.

Because in the current state HED mostly provides modules for building
SOAP based Web Services, it is easy to think that HED is just another
Web Services development framework like Axis, gSOAP, XFire or any other
out of the numerous implementations. Instead, the idea of HED is to
provide framework for gluing functionalities and not a
re-implementation of various standards. Effectively that means if
Apache httpd web server is considered by developers as necessary for
serving as front-end to services there could be plugin code written
which places Apache2 into a chain of other plugins of the HED, the same
for the Apache Axis SOAP implementation, as shown in Figure 1.

Even though the external solutions like Axis and Apache httpd can be
used for protocol supporting, \ two protocols are completely developed
in HED to support HTTP and SOAP in order to make the solution
lightweight and needed an implementation of those supported protocols
that is both simple and lightweight. However external solutions are
used whenever is appropriate - as in the case of TLS (openssl is used),
GridFTP (GridFTP API is used), LDAP and some other cases, this
principle applies to Axis and Apache httpd if we used the API from
those solution to implement the plugins for supporting SOAP and HTTP
protocol.

\bigskip

\bigskip

\bigskip

3. Cross-domain authentication and Single Sign-On

\bigskip

\ \ Community based authentication: Utilizing Shibboleth IdP

\bigskip

\ \ Short-lived certificate service

\bigskip

\ \ Certificate delegation service

\bigskip

\ \ From transport level security to message level security: X.509 Token
and SAML Token

\bigskip

\ \ Delegation between WS-Security tokens

\bigskip

\ \ \ \ \ \ \ \ \ trust relationship

\bigskip

4. Standardized Attribute-based authorization

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip
\end{document}

### 2008-12-01 doc/sec/Sec_paper.tex
Abstract about sec paper

@@
 
-Abstract-{}-
+{\centering
+Abstract
+\par}
 
-\bigskip
-
-\bigskip
+\ In order to make the access to Grids as simple as possible, security
+including authentication and authorization is one of the most important
+challenges in production Grids, especially when the Grid applications
+are across multiple administration domains as well as across
+heterogeneous Grid middlewares, such as some widely scaled eScience
+applications which needs to coordinate resources sharing among
+collection of autonomous institutions with different Grid middlewares
+running on these resources. In this paper, we describe the security
+implementation and consideration in new version of Advanced Resource
+Connector (ARC) middleware, where the above issues have been exploited.
+The main goal of ARC implementation in terms of security is to let the
+middleware be capable to \ interoperate with other Grid middlewares by
+leveraging standard specifications. The key aspect of the work we have
+done is to enhance the current proxy certificate based delegation and
+authentication by utilizing the Security Assertion Markup Languages
+(\textit{SAML}) single sign-on (SSO) profile, WS-Security in order to
+achieve cross-domain authentication, as well as to enhance the
+Attribute Certificate (AC) based authorization by utilizing the SAML
+\ attribute query profile in order to achieve standardized
+attribute-based authorization. \
 
@@
 \begin{enumerate}
-\item {\centering
-Introduction
+\item[] {\centering
+1. Introduction
 \par}
+\end{enumerate}
 
 \bigskip
-\end{enumerate}
 
@@
 
+HED contains a framework for implementing and enforcing authentication
+and authorization. Each Message Chain Component (MCC) or service has a
+common interface for implementing various authentication and
+authorization functionality. This functionality is implemented by using
+pluggable and configurable components (plug-ins) called
+\textit{SecHandler}. The \textit{SecHandler} components are C++ classes
+and provide method for processing messages traveling through Message
+Chains of the HED. Each MCC or Service \ usually implement two queues
+of SecHandlers -- one for incoming messages and one for outgoing called
+{\textquotedblleft}incoming{\textquotedblright} and
+{\textquotedblleft}outgoing{\textquotedblright} respectively. All
+\textit{SecHandler} components attached to the queue are executed
+sequentially. If any of them fails, message processing fails as well.
+In figure 1, the {\textquotedblleft}AuthZ{\textquotedblright} and
+{\textquotedblleft}AuthN{\textquotedblright} sub-module inside MCCs and
+Services are examples of \textit{SecHandler}.
+
+\bigskip
+
+{\centering
+3. Cross-domain authentication and Single Sign-On
+\par}
+
 \bigskip
@@
 
-3. Cross-domain authentication and Single Sign-On
+\ \ 3.1 Community based authentication: Utilizing Shibboleth IdP
 
@@
 
-\ \ Community based authentication: Utilizing Shibboleth IdP
+\ \ 3.2 Short-lived certificate service
 
@@
 
-\ \ Short-lived certificate service
+\ \ 3.3 Certificate delegation service
 
@@
 
-\ \ Certificate delegation service
+\ \ 3.4 From transport level security to message level security: X.509
+Token and SAML Token
 
@@
 
-\ \ From transport level security to message level security: X.509 Token
-and SAML Token
-
 \bigskip
 
-\ \ Delegation between WS-Security tokens
+\ \ 3.5 Delegation between WS-Security tokens
 
@@
 
+{\centering
 4. Standardized Attribute-based authorization 
+\par}
+
+\bigskip
+
+4.1 SAML Attribute Query profile
+
+\bigskip
+
+4.2 Client to VOMS Attribute authority Service
+
+\bigskip
+
+4.3 Attribute parsing and Policy decision point
 
@@
 
+{\centering
+5. Related work
+\par}
+
+VOMS AA Service
+
+Shibboleth
+
+GridShib
+
+gLite Delegation Service?
+
+SWITCH SLCS Service
+
+\bigskip
+
+{\centering
+6. Conclusion and Future work
+\par}
+
 \bigskip
@@
 
+Reference:
+
 \bigskip

### 2008-12-03 doc/storage_system/ARCStorage_src/ARCStorage.tex
ARC Storage paper, source

%\documentclass[10pts, twocolumn]{article}
%\documentclass[twocolumn]{IEEEtran}
\documentclass[final]{ieee}
\usepackage { graphicx}
\usepackage{wrapfig}

\hyphenation{librar-ian librar-ians shep-herd bar-tender gate-way name-space}

\begin{document}
\pagestyle{headings} 
\title{Service Oriented ARC Storage; \\An Overview and Design}
%% \author{  Jon K. Nilsen$^{2,3}$ \and Salman Toor$^{1}$
%%   \and Zsombor Nagy$^{5}$ \and Bjarte Mohn$^{4}$  \\ $^{1}$Scientific
%%   Computing Division, IT Department, Uppsala University, Sweden
%%   \\ $^{2}$Center for Information Technology, University of Oslo, Norway
%%   \\ $^{3}$Department of Physics,  University of Oslo, Norway
%%   \\  $^{4}$Department of Physics and Astronomy, Uppsala University,
%%   Sweden \\ $^{5}$National Information Infrastructure Development
%%   Institute, Hungary}
\author{  Jon K. Nilsen$^{\dagger,\ddagger}$ \and Salman Toor$^{*}$
  \and Zsombor Nagy$^{\mathparagraph}$ \and Bjarte Mohn$^{\mathsection}$  \\ $^{*}$Scientific
  Computing Division, IT Department, Uppsala University, Sweden
  \\ $^{\dagger}$Center for Information Technology, University of Oslo, Norway
  \\ $^{\ddagger}$Department of Physics,  University of Oslo, Norway
  \\  $^{\mathsection}$Department of Physics and Astronomy, Uppsala University,
  Sweden \\ $^{\mathparagraph}$National Information Infrastructure Development
  Institute, Hungary\\ \\  j.k.nilsen@usit.uio.no \hspace{5 mm}
  salman.toor@it.uu.se \hspace{5 mm} zsombor@niif.hu \\  \hspace{7 mm}
  bjarte.mohn@fysast.uu.se}

  % \\ \small{salman.toor \and bjarte.mohn  }}

\maketitle             
%% \begin{abstract}
%% The KnowARC project will deliver the next generation Advanced Resource
%% Connector Fall  2009. A major component of this
%% service oriented Grid middleware will be its storage system, the
%% ARC Storage, of which a
%% prototype is due Spring 2009. We will in this paper present an
%% overview of the storage system, its design and architecture. We will
%% also present some first, proof-of-concept test results, deploying the
%% the storage with a geographically distribution among three different countries.
%% ~\\~\\~
%% \end{abstract}
\begin{abstract}
There is an ever increasing need to utilize geographically distributed
hardware resources, both in terms of CPU and in terms of storage. The
service oriented architecture provides a natural framework for
managing these resources. The next generation Advanced Resource
Connector (ARC) is a service oriented Grid solution that will provide
the middleware to represent distributed resources in one simple
framework. In this paper, we will present an overview of the ARC
storage system, itself a set of services providing a self-healing,
flexible Grid storage solution. We will also present some first
proof-of-concept test results, deploying the storage system
distributed between three different countries. 
~\\~\\~
\end{abstract}
%-----------------------------------------------------------
\section{Introduction}
\label{Introduction}

The challenge of building a reliable, self-healing, fault-tolerant,
consistent data management system in a web scale is an interesting
task. Making the system work in a heterogeneous, distributed
environment like the Grid is even more interesting. An increasing
amount of applications demands not only increased CPU power, but also
vast amounts of storage space. The required storage space is not only
restricted to the duration in which the application runs; the data
should often be available for years afterwards, in certain cases even
for decades. Nowadays, we can easily find single Grid jobs, producing
gigabytes or even terabytes of data, ramping up the requirements of
storage systems to the petabyte scale and beyond. Taking into account
that hardware is inherently unstable\footnote{E.g., if one hard drive
  has an uptime of 10,000 hours, one of 1,000 hard drives will fail
  after 10 hours.}, the need for replicating, self-healing storage is evident. To
make the storage system useable to Grid users, the system must provide
reliable and secure file transfer protocols, a cataloging system and
secure storage of the data. Several projects and designs have emerged
to address such challenges \cite{Hoschek00datamanagement,DengWang}.

In the advent of the next generation Advanced Resource Connector (ARC)
Grid middleware \cite{NorduGridsite} (due during Fall 2009), we present 
the ARC storage system \cite{ARCStoragedesigndoc}. This distributed
storage system is designed to provide an easy
to use, flexible and scalable system that can offer native storage and
at the same time provide access to third-party solutions like dCache by
using the same, uniform interface. Being part of ARC, the
storage system is based on a service
oriented architecture, in which each major component of the
system runs as a separate service within the ARC Hosting Environment
Daemon (HED) \cite{HEDdesigndoc}. The HED service container gives the
capability of flexible replacement of the components as well as the
possibility to introduce modifications in the future. %% The HED service container gives
%% the capability of flexible replacement of the components as well as
%% the future modification/enhancement of the system.

This paper is organized as follows. After describing
related work in Section \ref{Related Work}, we give the bird's eye
view of the next generation Advanced Resource Connector (ARC) in
Section \ref{The Advance Resource Connector}. An overview of the
ARC Storage is given in Section \ref{The ARC Storage System}, while
the architecture of the storage system is elaborated in Section
\ref{Architecture of the ARC Storage}. In Section \ref{Testing and
  Discussion} we give some early, proof-of-concept results, before 
concluding in Section \ref{Conclusion and Future Work}.

\section{Related Work}
\label{Related Work}
\textbf{dCache:} dCache is a storage system \cite{dCache} \cite{DSSWithdCache} which combines heterogeneous storage
elements in the order of several hundreds of terabytes in a single
namespace. It additionally supports standard and native
protocols like gridftp, srm, dcap and gsidcap. dCache is
a joint effort between Desy  \cite{dCachesite}, Fermi National Laboratory \cite{FermiLab}, the  Nordic Data
Grid Facility (NDGF) \cite{NDGF}  and several other collaborators. %% The services or
%% components of dCache are known as cells. The standard deployment of
%% dCache consist of  storage pools; used to store file, optionally be
%% connected with HSM\footnote{Hierarchical Storage Manager, used for the
%%   tape storage}. Namespace manager; used PNFS \footnote{Perfectly
%%   Normal FileSystem; A network file system, designed in Desy} as a
%% filesystem  for storing matadata. And the number of protocol
%% doors. One of the biggest deployment of dCache is the distributed
%% infrastructure of  NDGF.\\ 
dCache has proven to be a very stable and scalable solution. However,
it has a relatively high threshold for understanding and
deployment. The ARC Storage, being a light-weight and flexible
storage solution, aims more towards new user groups less familiar with
Grid solutions. 
\newline
\textbf{BigTable:} Bigtable is a distributed storage system managing structured data in
the petabyte scale \cite{Bigtable}. It is currently used within
Google, in projects like web indexing, Google Earth and Google
Financing. %% Bigtable is mainly built on top of three pieces of Google's
%% infrastructure: The Google Filesystem \cite{GFS} a distributed,
%% replicating file system, the Google SSTable, providing
%% a persistent, ordered immutable map and Chubby \cite{Chubby}, a highly
%% available and persistent distributed lock service. The system uses a
%% three-level B$^+$-tree to store \textit{tablets}. The location of the
%% first level is stored in Chubby, making Chubby a vitaly important part
%% of the design. To ensure a persistent solution, Chubby takes advantage
%% of the Paxos algorithm \cite{Paxos,PaxosLive}.\\
Bigtable has several interesting features, one of which is the
distributed lock service, Chubby \cite{Chubby}. Chubby is a high-availability, distributed
locking service sitting on top of the B$^+$-tree architecture of
Bigtable. Chubby has several features similar to the distributed A-Hash service,
among which the use of the Paxos algorithm \cite{Paxos,PaxosLive}, is
the most striking. A major caveat for the Grid community is that
Bigtable is neither free, open-source, nor available to the public.
\newline
\textbf{Storage Resource Broker:}  Based on the client-server model,
The Storage Resource Broker (SRB) \cite{earlySRB,SRB} provides a flexible data grid management
system. It allows  uniform access for connecting to heterogeneous
storage resources over a wide area network. %% SRB also manages the
%% namespaces used to identify files, storage resources, users, metadata,
%% and access controls independently of the physical storage systems. It
%% uses centralized matadata catalog (MCat). Also supports several file
%% systems of UNIX, Windows, Linux, Mac OS X, HPSS and large scale
%% storage systems like Castor and dCache.\\
Its functionality, with a uniform namespace for several Data Grid
Managers and file systems, is quite similar to the functionality
offered by our Gateway service. However, being built as a middleware on top of
other major storage solutions, SRB does not offer its own storage solution.
\newline
\textbf{Scalla:} Scalla is a widely used software suite consisting of an
xrootd server for data access, and an olbd
server for building scalable xrootd clusters \cite{Scalla}. %% The xrootd
%% server consists of several plug-in based components, of which the most
%% important are; the xrd component, responsible for network, threading,
%% data buffer and protocol management; the xroot protocol component,
%% providing the specialized and increasingly popular xroot protocol; and
%% components for logical and physical file systems. The olbd server is
%% used for setting up storage clusters in a B-64 tree with a
%% redirector on top appointing one supervisor for every 64 data servers
%% in the lower layers of the tree.
Originally developed for use with the physics analysis tool root,
xrootd offers data access both through the specialized xroot protocol
and through other third-party protocols. The combination of the xrootd
and olbd components offers a cluster storage designed for low latency
high bandwidth environments. In contrast, the ARC Storage is optimized
for high latency, more suitable for the Grid environment.

%% {\large Comparison:}

%% dCache is a very scalable solution, but it has a relatively high
%% threshold (more under standing is required) for learning and
%% deployment.  Proposed solution is more light-weight, flexible and
%% hence more suitable for the new users.

%% Xrootd with OBLD is a cluster storage designed for low latency high
%% bandwidth. Our solution is optimized for high latency more suitable
%% for Grid environment.

%% Big table can handle extreme amount  of data and vast number of
%% users. How ever it is still an in-house solution.

%% The functionality of SRB, where there is a uniform namespace for
%% several DGMS and file systems, is very similar to our gateway service.

\section{The Advance Resource Connector}
\label{The Advance Resource Connector}

The next generation Advanced Resource Connector (ARC) Grid middleware \cite{NorduGridsite} \cite{KnowARCsite} consists of a set of pluggable components. These components are the
fundamental building blocks of the ARC services and clients. ARC
services are running inside a container called the Hosting Environment
Daemon (HED). There are four kinds of pluggable components having well
defined tasks: Data Management Components are used to transfer the
data using various protocols, Message Chain Components are
responsible for the communication within clients and services as well as
between the clients and the services, ARC Client Components are
plug-ins used by the clients to connect to different Grid flavors and
Policy Decision Components are responsible for the security model
within the system.

To deliver the non-trivial quality of services required by the Grid,
there are a number of services running inside the HED. For example, Grid
job execution and management is handled by the A-REX service \cite{AREXdesigndoc}, policy
decision is the task for the Charon service, the ISIS service is
responsible for the information indexing, batch jobs submission is
handled by the Sched service etc. In the later sections our discussion
will focus on the architecture and the design of another major set
of services, i.e., the ARC storage.

\section{The ARC Storage System}
\label{The ARC Storage System}

The ARC Storage consists of a set of SOAP based services residing
within HED. Together the services provide a self-healing, reliable, robust,
scalable, resilient and consistent data storage system. Data is
managed in a hierarchical global namespace with files and
subcollections grouped into collections\footnote{A concept very
  similar to files and directories in most common file systems.}.  A
dedicated root collection serves as a reference point for accessing the namespace. The hierarchy can then be
referenced using Logical Names. The global namespace is accessed
in the same manner as in local filesystems.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.8\columnwidth]{arc1-storage-services-with-gateway.pdf}
  \end{center}
  \caption{Schematic of the ARC Storage architecture. The figure shows
    the main services of ARC Storage and how they communicate. See
    text for further details.}
  \label{fig:arc1_storage_service}
\end{figure}

Being a service oriented architecture based system, the ARC
Storage consists of a set of services as shown in figure
\ref{fig:arc1_storage_service}. The services are as follows: The Bartender (B),
providing the high-level interface to the user; the Librarian (L),
handling the entire storage namespace, using the A-Hash (A-H) as a metadatabase; the
Gateway (G) providing means to access third-party storage systems; and
the Shepherd (S), the frontend for the physical storage node. 
See Section \ref{Architecture of the ARC Storage} for a detailed
discussion of the different services.   
The services communicates with each other through the
Message Chain Components in HED. The communication is
depicted by straight lines in figure \ref{fig:arc1_storage_service}.

The system supports file transfer through several transfer protocols,
with client side tools hiding technical
details, such as specifying protocols, port numbers
and so forth. To enable for fault-tolerance, the system implements
automatic file replication, where
the replicas of a file\footnote{In this paper, a file
denotes an entry in the global namespace, while replica denotes the
physical file stored on a storage node.} are always stored on separate
storage nodes\footnote{A
  storage node is a server with a Shepherd, a Shepherd backend
and some storage service.}. To ensure a resilient, self-healing system, the Shepherd
regularly sends heartbeats to one of the Librarians. If the Shepherd fails
to send a heartbeat one of the Librarians will automatically initiate
re-replication of the replicas between the other Shepherds.

In the default implementation the services will provide a
full-featured and consistent data storage system using files as an
atomic unit. The scope of the ARC data management system in the foreseen
future is restricted to provide support for file-based data services.
%% there are currently no plans for intrinsic support for database
%% access.

The ARC Storage supports third-party storage services in two
ways. Using the Gateway service, files already stored in some
third-party storage can be accessed using the client tools provided by
the ARC Storage. To make use of third-party services for storing new
files, with the replication and fault-tolerance offered by the ARC
Storage, third-party storage elements can also be used as backends for the
Shepherd service.

External grid middleware components can
access the ARC storage system by adapting and using ARC data service
interfaces directly. ARC will provide interface
components that communicate via standard protocols like SRM, which
will give a single access point to the whole system.

\section{Architecture of the ARC Storage}
\label{Architecture of the ARC Storage}

In a service oriented architecture
the role of each service is well defined. Available
objects\footnote{Files, collections, mount points, etc.} in the
system are identified by unique global IDs. These IDs are
categorized in three different categories according to the object type.

\begin{itemize} 
\item Each  file and collection have a unique ID called GUID.
\item Services are uniquely identified by using a serviceID. 
\item The Shepherds in the system identify their files using a referenceID.
\end{itemize}

Each object in the ARC Storage has a globally
unique ID. A collection contains files and other collections,
and each of these entries has a name unique within the collection very
much like entries in a standard directory on a local filesystem. Besides
files and collections the ARC Storage has a third type of entries
called mount-points which are references to the third-party storages
within the global namespace.

Replicas in a distributed storage system can have different states;
they can be broken, deleted, partially uploaded, etc. In the
ARC Storage, all replicas has assigned a state, some of which are
`\textbf{alive}' (if the replica passed the checksum test, and the
Shepherd reports that the storage node is healthy), `\textbf{invalid}'
(if the replica has wrong checksum, or the Shepherd claims it has no
such replica), `\textbf{creating}' (if the replica is in
the state of uploading).

The details of the
architecture are divided into two parts: First we will discuss the
details of the core components, second we will discuss some of the
important features provided by the system.

\subsection{Core Components} 
\subsubsection{Bartender}
\label{Bartender}
The Bartender service provides a high-level interface for the storage
system. Clients connect to the Bartender to create and remove
files, collections and mount-points using their Logical Names. The
Bartender communicates 
with the Librarian and Shepherd services to accomplish the clients
requests. However, the actual file data does not go through the
Bartender, instead file
transfers are directly performed between the storage nodes and the
clients. There could be any number of independent Bartender services
running in the system, providing high-availability and
load-balancing. 
\subsubsection{Librarian}
\label{librarian}
The Librarian manages the hierarchy and metadata of
files, collections and mount points, as well as the health information of the Shepherd
services. In addition, the Librarian handles
the information about registered Shepherd services. The Librarian receives
heartbeat messages from the Shepherds and change replica states automatically
if needed. The Librarian uses the A-Hash for consistent storing. This makes the
Librarian a stateless service, thus enabling the system to have any
number of independent
Librarian services, again providing high-availability in the
system.

\subsubsection{Shepherd}
\label{Shepherd}

The Shepherd services run as front-ends on storage nodes. A Shepherd
service reports to a Librarian about the 
node's health state in terms of replicas. While the Bartender initiates
file transfers, the actual transfers goes directly between the Shepherd
and the clients.

%% The file-names used by a Shepherd is completely independent from the
%% hierarchy of collections or Logical Names.
When a new replica upload
is initiated, the Shepherd generates a referenceID which refers to
the replica within
that Shepherd. Each Shepherd has a unique serviceID, so with these two
IDs the replica can be unambiguously referenced. This is called a Location
of the replica. %% The namespace of these Locations is independent from
%% the namespace of GUIDs and Logical Names.

\subsubsection{A-Hash}
\label{Hash Algorithm}

A-Hash is a hash table for consistently storing data in property-value
pairs. All metadata about files, collections, mount point, Shepherd's
health status, and so forth, is stored in the A-Hash. As the A-Hash is
the service storing the entire state of the storage system, its
absolutely crucial for the ARC Storage that the A-Hash is
consistent. The distribution and replication of this service is
therefore both challenging and needed.

\begin{figure}
\centering
\includegraphics[width=1.0\columnwidth]{arc1-storage-downloading.pdf}
\caption{Schematic showing the scenario of file download from the ARC Storage.}
\label{arc1-storage-downloading}
\end{figure}

\subsection{Features}
\subsubsection{Heartbeat monitoring}
\label{Heartbeat and Replication}
In the proposed architecture, each Shepherd periodically send
heartbeats to a Librarian with information about replicas
whose state changed since the last heartbeat, and the corresponding
GUIDs. These heartbeats are then stored in the A-Hash, making them
visible to all the Librarians in the system. If any of the Librarians
notices that a Shepherd is late with its heartbeat, it will invalidate
all the replicas in that Shepherd.

\subsubsection{Replication}
\label{Replication}
Shepherds periodically ask the Librarian if the file of the replica stored on
its storage node have enough replicas. If the file does not have
enough replicas, the Shepherd informs the Bartender and the Bartender initiates
a put request that returns a Transfer URL to the Shepherd. The
Shepherd finally uploads the new replica. The Shepherds which gets the new replica notifies the
Librarian that the replica is alive. The Librarian then add this to
the corresponding file.

\subsubsection{Deletion}
\label{Deletion}
In a replicating storage system, there are several possible solutions
for file deletion, since both the replicas and the metadata need to be
removed. In our solution, we use the process of \textit{lazy
  deletion}, \cite{LazyDeletion}. When a client requests that a file
should be deleted (and the user has the proper permissions) the
Bartender instructs the Librarian to remove the file's GUID from the
A-Hash. When the Shepherd, periodically checking all its replicas,
discovers that a replica has no file (and hence, no Location), it will
automatically delete the replica.

\subsubsection{Fault tolerance}
\label{Fault tolerance}

Due to the unpredictable nature of the Grid environment, it is essential to
have some degree of  automatic recovery system in case of unexpected
failures. Fault tolerant behavior is required both at the level of
metadata and on the level of physical storage. While the work on fault-tolerant metadata is still in
progress, the following two
scenarios will explain the currently available recovery mechanism for
physical storage: 
\begin{itemize}
\item In the case of a file having invalid checksum, the Shepherd
immediately informs the Librarian and the Librarian changes the state of the given replica to
`\textbf{invalid}'. To recover its replica, the Shepherd contacts a
Bartender and asks for another replica of the file. The Bartender chooses a valid
replica, initiate file transfer from a Shepherd having the replica,
and returns the TURL to the Shepherd with the invalid replica. When the
Shepherd has recieved the replica and compared the checksum, it notifies the
Librarian that the replica is alive again.  
\item In the case of a Shepherd becoming offline, a Librarian will
  as mentioned notice the lack of heartbeats and invalidate all the
  replicas, initiating new replication for all the files stored in
  this storage node. However, if the Shepherd
  again comes online, there will evidently be more replicas than
  needed. The first Shepherd to notice this will set its replica's
  state to `\textbf{thirdwheel}', i.e., obsolete. At the next occasion,
  the Shepherd will remove the replica, if and only if it has the only
  `\textbf{thirdwheel}' replica of this file. If there are more
  replicas with this state, all replicas will be set back to
  `\textbf{alive}' and the process is repeated. This scenario will be discussed further in Section
  \ref{Testing and Discussion}.
\end{itemize}
\subsubsection{Client tools }
\label{Client extension using FUSE module}

Being the only part a user will (and should) see from a storage
system, the client tools is an important part of the ARC
Storage. Currently ARC supports two ways of accessing the
storage solution:
\begin{itemize}
\item The \textbf{Command-line Interface} (CLI) provides
  access to the storage through the methods \texttt{stat}, \texttt{makeCollection},
  \texttt{unmakeCollection}, \texttt{putFile}, \texttt{getFile}, \texttt{delFile},
  \texttt{list} and \texttt{move}. Methods for modifying
  access and ownership will be available in the near future. The CLI
  does assume a relatively high level of computer experience
  from the user. However, the CLI, being a stand-alone tool, can be
  used to access ARC Storage from any computer\footnote{Also including
  Microsoft Windows PCs.} that has network access and
  a Python installation.
\item The \textbf{FUSE module} provides a high-level access to the
  ARC Storage. Filesystem in Userspace (FUSE) provides a simple
  library and a kernel-userspace interface. Using FUSE and the ARC
  Storage Python interface, the FUSE module allows users to mount the
  ARC Storage namespace into the local namespace. This way, the user
  can use his/her favorite file manager to access his/her files and
  collections. The FUSE module provides most of the features provided
  by the CLI, with the exception of modifying some non-posix metadata.
\end{itemize}
Its worth mentioning that the client tools access the storage system
through the Bartender only. Currently upload and download is realized
through HTTP(S), but there are plans to add support for other
protocols, such as SRM and GridFTP.

\subsubsection{Gateways }
\label{Gateways}

Gateways are used to communicate with the external storage
managers. While designing this service, care was taken to: 
\begin{itemize}
\item Retain the transparency of the global namespace while using the external storage systems.
\item  Develop a protocol oriented  service i.e., all the external
  storage managers which support a certain protocol should be handled
  using the corresponding gateway service.  
\end{itemize}
This approach provides the flexibility of  avoiding multiple Gateway
services for different storage managers. Currently the available
Gateway service is based on the gridftp protocol.

When the user requests comes, the Librarian provides the metadata related
with the request to the Bartender. Requests can be related to
files, collections or external mount-points. In case of creating
external mount-points, the Bartender contacts the Librarian to store
the mount-point for later use. In the case where a request is related
to the downloading of files from the external store, the Gateway service first checks the status of the file
and then send the Transfer URL (TURL) to the client via the
Bartender. Using this TURL client can directly get the file from
external store. %% For the experiments we use dCache with gridftp
%% protocol as an external storage system.

\subsection{Security Model }
\label{Security Model }

As is the case for all openly accessible web services, the security
model is of crucial importance for the ARC Storage. The security
architecture of the storage can be split into three parts; the
inter-service authorization; the transfer-level authorization; and the
high-level authorization.
\begin{itemize}
\item The \textbf{inter-service authorization} takes care of the
  internal communication between services. There is a lot of
  communication between the services of the ARC Storage. The
  Bartenders send requests to the Librarians and the Shepherds, the
  Shepherds communicate with the Librarians, the Librarians talk with
  the A-Hash. If any of these services get compromised or a new
  rogue service gets inserted in the system, the security of the
  entire system is compromised. To enable trust between the services,
  they need to know each other's Distinguished Names (DNs). This way,
  a rogue service would need to obtain a certificate with that exact
  DN from some trusted Certificate Authority (CA).
\item The \textbf{transfer-level authorization} handles the
  authorization in the cases of uploading and downloading files. When
  a transfer is requested the Shepherd will provide a one-time
  Transfer URL (TURL) which the client can connect to. In the most
  current architecture, this TURL is world-accessible. This may not seem
  very secure at first. However, provided that the TURL has a very long,
  unguessable name, that it is transfered to the user in a secure way
  and that it can only be accessed once before it is deleted, the
  chance of compromitation is very low.
\item The \textbf{high-level authorization} considers the access
  policies for the files and collections in the system. These policies
  are stored in A-Hash, in the metadata of the corresponding file or
  collection, offering a fine-grained security in the system.
\end{itemize}
The communication with and within the storage system is realized
through HTTPS with regular X.509 authentication.

\section{Testing and Discussion}
\label{Testing and Discussion}

Even though the ARC Storage is in a pre-prototype state, it is already
possible to deploy and use the system for testing purposes. To
properly run a proof-of-concept test, the resources need to be
geographically distributed. In our test scenarios we utilized
resources in three different countries.

In our test deployment, we used two nodes from Uppsala
Multidisciplinary Center for Advanced Computational Science (UPPMAX), Sweden, 
%each with a 3.00GHz Dual Core Intel(R) Pentium(R) 4 CPU and 4 GB RAM, 
three nodes from the Center for Information Technology (USIT) at the
University of Oslo, Norway, 
%each with two 3.40GHz Dual core Intel(R) Xeon(TM) CPUs with 2GB RAM
and one node from National Information Infrastructure Development Institute (NIIF)
Hungary. %with a 2.00GHz Dual Core Intel(R) Pentium(R) 4 CPU with 4 GB RAM.

The services were distributed as shown in figure \ref{fig:map}:
\begin{itemize}
\item An A-Hash runs at UPPMAX.
\item A Bartender runs at USIT.
\item A Librarian runs on UPPMAX.
\item In total we are running five Shepherds: Three at USIT, having
  100GB storage space each, one at UPPMAX, with 20GB, and one at NIIF, providing 16GB of storage space.
 \end{itemize}

\begin{figure}
\centering
\includegraphics[width=0.7\columnwidth]{map}
\caption{The map shows the geographical distribution of the services in the test setup.}
\label{fig:map}
\end{figure}

The following test where considered:

\begin{itemize}
\item Test 1: The distribution after uploading 10 large size (1 GB)
  files with one replica each.  
\item Test 2: The distribution after uploading 100 small size (1 kB)
  files with one replica each, and with simultanious uploading
\item Test 3: Distribution after repeating Tests 1 and 2 with three replicas.  
\item Test 4: System behavior while some of the shepherd becomes offline.   
\end{itemize}

All the tests were performed and worked as expected. However, some of
the tests deserve extra attention.

Figure \ref{fig:FileDistribution} illustrate two different scenarios,
corresponding to Test 2 and Test 3: The orange bars show the distribution after uploading 100 small files
simultaneously, without replication. This gives an indication of how
the system balances the load of many clients uploading at the same
time. The green bars show distribution with the same kind of
uploading, but with the clients requesting three replicas each. The
difference here is that the system simultaneously has to handle both
replication and the clients requesting uploads. When the Bartender chooses
a location for which to put a replica, it generates a list of
Shepherds that (a) do not have a replica of the file, and (b) is not
already in the process of uploading the replica. Next, it draws a
Shepherd from the list at random, using a uniform random number
generator. When uploading without asking for replicas we would
therefore expect a relatively flat distribution of the files, as can
be seen in figure \ref{fig:FileDistribution}.

\begin{figure}
\centering
\includegraphics[width=0.8\columnwidth]{FileDistribution}
\caption{The distribution of replicas over the geographically distributed
  storage nodes. Green bars shows the distribution in the case where
  we upload 100 files requesting 3 replicas where as orange bars show
  the distribution after uploading 100 files without replication. }
\label{fig:FileDistribution}
\end{figure}

\begin{table}[ht]
\centering
%% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
\begin{tabular}[width = \columnwidth]{llllll}
\hline \hline
&\small{UPPMAX}&\small{USIT-1}&\small{USIT-2}&\small{USIT-3}&\small{NIIF}\\
\hline
           \small{Load(MB)}&6683&7638&7650&4773&2289\\
\hline
\end{tabular}
\caption{Overall load distribution on the storage nodes after Tests 1,
2 and 3 were finished}
\label{table:load distribution}  
\end{table}

Table \ref{table:load distribution} shows the disk usage after Tests
1, 2 and 3. Worth noticing here is that the bandwidth NIIF was
significantly lower than between UPPMAX and USIT. If the bandwith is
saturated and the storage node is busy with, e.g., checksumming a
large file, the heartbeat from this Shepherd may be delayed, causing the
Bartender not to choose this Shepherd for the next replica. This may explain the
relatively low storage load on the  NIIF storage node.

%% \begin{table}[ht]
%% \centering
%% %% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
%% \begin{tabular}[width = \columnwidth]{llllll}
%% \hline \hline
%% &\small{UPPMAX}&\small{USIT-1}&\small{USIT-2}&\small{USIT-3}&\small{NIIF}\\
%% \hline
%%            \small{Before(MB)}&4888&7820&6843&7820&3320\\
%%            \small{During(MB)}&8798&9776& - &9775&2344 \\
%%            \small{After(MB)}&8798&6843&4888&8798&1367\\
%% \hline
%% \end{tabular}
%% \caption{Storage load before, during and after a Shepherd outage.}
%% \label{table:self healing}  
%% \end{table}

\begin{table}[ht]
\centering
%% \begin{tabular}[width = \columnwidth]{p{11mm} p{12mm} p{11mm} p{11mm} p{11mm} p{5mm}}
\begin{tabular}[width = \columnwidth]{llll}
\hline \hline
&\small{Before (MB)}&\small{During (MB)}&\small{After (MB)}\\
\hline
\small{UPPMAX}&4888&8799&8798\\
\small{USIT-1}&7820&9778&6843\\
\small{USIT-2}&6843&-&4888\\
\small{USIT-3}&7820&9774&8798\\
\small{NIIF}&3320&2344&1367\\
\hline
\small{Sum}&30691&30693&30694\\
\hline
\end{tabular}
\caption{Storage load before, during and after a Shepherd outage.}
\label{table:self healing}  
\end{table}

A significant feature of the ARC Storage is its automatic self
healing. Test 4 addresses this feature by studying the effect of
taking out one Shepherd from the system, and later reinsert it. Table
\ref{table:self healing} shows the storage distribution in three
states: Before interrupting a Shepherd, after the USIT-2
Shepherd is interrupted and redistribution of replicas is finished; and
alive, when the Shepherd is restarted and the system again has stabilized. We
can see that all files are properly distributed between the remaining
Shepherds when one of them is disrupted. Interestingly, the storage load
automatically evens out when the Shepherd comes back online. When the
system discovers that there are more replicas than needed, the first
Shepherd noticing will set mark its replica as obsolete. Since the failing Shepherd didn't lose any replicas while
being down, redistribution of replicas is just a matter of deleting
obsolete replicas. Interestingly, the NIIF node actually got fewer
files when USIT-2 went offline. If two Shepherds simultaniously
starts uploading replicating a missing replica, there will be too many
replicas. Here, a big replica on the NIIF node has randomly been chosen as
obsolete. However, the total storage usage remains the same in all
three cases.

\section{Conclusion and Future Work}
\label{Conclusion and Future Work}

The proposed system is still in an early phase of development, but our test
results depict that the architecture is strong enough to handle
the challenges for distributed large scale storage. Much
effort is required to make the system production
ready. However, we believe that continuing in the same direction
will enable us to provide a persistent  and flexible storage system
which can fulfill the needs of the scientific community.

Some key areas need special attention and effort to make the proposed
storage system even more stable, reliable and consistent:
\begin{itemize}
\item \textbf{Security:} This is still under development. The design is more
  or less ready, but it still needs to be implemented and properly
  tested.
\item \textbf{The A-Hash}: This is currently centralized. To avoid a singel
  point of failure in the system, and to improve the system
  performance, the A-Hash needs to be distributed.
\item \textbf{Performance optimization:} To make a storage system
  ready for production, one needs to discover and improve on
  possible bottlenecks within the system. As soon as the A-Hash is
  distributed, other possible bottlenecks, such as load-balancing
  and self-healing mechanisms, will be investigated.
\item \textbf{Protocols:} To ease the interoperability with third-party
  storage solutions and clients, the system needs to support storage
  protocols such as SRM and GridFTP. In addition the system will come
  with its own ARC protocol.
\end{itemize}

Even though we have more work in front of us, we have shown that the
ARC Storage already is in a state where initial real-life tests can be
done. We have described a simple, yet strong architecture which we
believe will benefit communities in need of a lightweight,
yet distributed storage solution.

\section{Acknowledgements}
\label{Acknowledgements}

We wish to thank Mattias Ellert for helpful discussions and guidance
on the ARC middleware. In addition, we like to thank UPPMAX, NIIF and USIT for
providing resources for running the storage tests.

%% In the near future proposed solution will be the part of the ARC
%% middleware production line. We have identified some key areas in which
%% it is essential to put more efforts, to make the system more stable,
%% reliable and consistent. For example: A-Hash is currently centralized
%% and it need to be distributed to avoid single point of failure,
%% require to support more standard protocols like SRM, Better load
%% balancing system for Bartender, more improvements in the self healing
%% mechanism. The work in all these areas is in progress. Hopefully, we
%% will come-up with a reasonably good solutions (One or two last
%% lines)....  
%--------------------------------------------------------
%-----------------------------------------------------------
%% \begin{thebibliography}{5}

%% International Conference on Computing in High Energy and Nuclear Physics (CHEPê07) IOP Publishing 
%% Journal of Physics: Conference Series 119 (2008) 062014 doi:10.1088/1742-6596/119/6/062014 
%% \newline
%% A Simple Mass Storage System for the SRB Data Grid, Michael Wan, Arcot Rajasekar, Reagan Moore, Phil Andrew, 20th IEEE/ 11th NASA Goddard Conference on Mass Storage Systems \& Technologies (MSST2003) San Diego, California, April 7-10, 2003.
%% \newline
%% Advanced Resource Connector middleware for lightweight computational Grids, ELLERT M. ; GRONAGER M. ; KONSTANTINOV A. ; KONYA B. ; LINDEMANN J. ; LIVENSON I. ; NIELSEN J. L. ; NIINIM?KI M. ; SMIRNOVA O. ; WAANANEN A. ;FGCS. Future generations computer systems   ISSN 0167-739X   CODEN FGCSEV

%% \end{thebibliography}

\bibliography{arcstorage_refs}
\bibliographystyle{plain}

\end{document}

### 2008-12-03 doc/sec/Sec_paper.tex
Some more sections for the sec doc

@@
 \newcommand\liststyleLi{%
+\renewcommand\theenumi{\arabic{enumi}}
+\renewcommand\theenumii{\arabic{enumii}}
+\renewcommand\theenumiii{\arabic{enumiii}}
+\renewcommand\theenumiv{\arabic{enumiv}}
+\renewcommand\labelenumi{\theenumi.}
+\renewcommand\labelenumii{\theenumii.}
+\renewcommand\labelenumiii{\theenumiii.}
+\renewcommand\labelenumiv{\theenumiv.}
+}
+\newcommand\liststyleLii{%
+\renewcommand\labelitemi{${\bullet}$}
+\renewcommand\labelitemii{${\circ}$}
+\renewcommand\labelitemiii{${\blacksquare}$}
+\renewcommand\labelitemiv{${\bullet}$}
+}
+\newcommand\liststyleLiii{%
+\renewcommand\labelitemi{${\bullet}$}
+\renewcommand\labelitemii{${\circ}$}
+\renewcommand\labelitemiii{${\blacksquare}$}
+\renewcommand\labelitemiv{${\bullet}$}
+}
+\newcommand\liststyleLv{%
 \renewcommand\theenumi{\arabic{enumi}}
@@
 \end{enumerate}
+Grids provides the technology for large scaled, cross domains
+collaboration where the users work on resources and sharing data which
+are distributed across domains. This kind of collaboration is called
+virtual organization (VO)[ref1]. The resources that contribute to the
+VO could come from independent administrative domains, and even require
+different security approaches for control the access to themselves. The
+same heterogeneous situation even critically applies to users who need
+to access the VO, because users holding one kind of credential could
+need to travel through multiple resources with different security
+approaches. To solve the above issues, the Grid middleware should
+provide the following ability: The user which participates the VO
+should be able to access data and resources without being required to
+constantly do authentication at each resource or data; The VO should be
+able to easily manage the privileges or memberships to the users, and
+VO together with its resources should be able to easily manage the
+authorization, given the condition that the resources are actually
+owned by different administrative domains.
+
+In grid community, Grid Security Infrastructure (GSI)[ref2] is the
+\textit{de facto} standard about authentication and \ transport level
+communication, which is builds on Public Key Infrastructure (PKI), an
+architecture based on X.509 public key certificate. GSI implements some
+enhancement (such as certificate delegation) based on standard SSL/TLS
+protocol. Mutual authentication is required by GSI, and is the default
+configuration for GSI based grid deployment; and X.509 certificates is
+required for both of the client and service sides in order to achieve
+mutual authentication. X.509 certificates is issued by trusted third
+party called certificate authorities (CA), and then CAs constitute
+trust federation and guarantee two different X.509 certificates from
+different CAs can accomplish authentication to each other. So if a user
+would access grid system, he/she should own a X.509 certificate which
+is issued by a CA that is trusted by other{\textquotesingle}s entity in
+the GSI based grid system.
+
+In terms of the virtual organization issues mentioned earlier, GSI and
+some other related solutions like virtual organization management
+service (VOMS) [ref3] has partly solved the them in some sense. The
+identity delegation and X.509 proxy certificate initiated by GSI can
+achieve single sign-on, so that the need to re-enter the
+user{\textquotesingle}s pass phrase for authentication can be avoided
+by creating the proxy certificate, even if some local or remote
+entities access access resources on behalf of this user. \ As the most
+successful grid authorization model, VOMS adopts X.509 Attribute
+Certificate[ref4]. Attribute certificate is signed by VOMS server and
+then binded to proxy certificates by client, \ and attributes
+certificate includes the VO memberships which is associated with the
+proxy certificate{\textquotesingle}s identity. Essentially the
+resources can make access control decision based on parsing the VO
+membership attributes from the homogeneous attribute certificate.
+
+\ \ \ \ The VO is essentially supposed to be able to provide resource
+sharing capability with resources running under heterogeneous
+systems/middlewares. This goal can be achieved using Web Service
+Architecture (WSA). Therefore, the Web Service technology has been
+converged into Grid computing middlewares. The Web Service technology
+has been adopted by Globus toolkit, gLite, as well as the ARC.
+Actually, utilizing Web Service technology and providing seamless
+interoperability with other middlewares is one of the main goals of the
+new version of ARC middleware.
+
+Unfortunately, the existing grid security solutions such as GSI and VOMS
+do not apply to Web Service \ based application very well due to the
+following reasons:
+
+\liststyleLii
+\begin{itemize}
+\item Firstly, the common implementation for Web Service does not
+necessarily recognize and then properly verify and parse the
+certificate, since comparing with usual X.509 certificate, proxy
+certificate needs different treatment such as proxy path verification. 
+\item Secondly, the X.509 Attribute Certificate has not been adopted by
+any Web Service implementation at all. 
+\item Thirdly, what is even worse is that the GSS API based confidential
+communication has also not been adopted by Web Service implementation.
+\item Finally, identity delegation is a good solution for user or
+user{\textquotesingle}s representative (on behalf of the user) to
+move/travel from one resource to another resource. But since the
+identity delegation has been completely coupled with the GSI
+confidential communication process, it is not possible for the existing
+identity delegation solution to be used by Web Service applications.
+\end{itemize}
+There is some effort to add GSI support to adapt to the strict security
+requirement including GSI from grid middlewares, such as GSI plug-in
+for gSOAP[ref6]. However, this kind of solution
+doesn{\textquotesingle}t help the existing grid middlewares
+interoperate with other common Web Service implementation. Instead, Web
+Service implementations commonly use the WS-Security, Security
+Assertion Markup Languages (SAML), and standard SSL/TLS as well as
+standard X.509 certificate when considering security. \
+
+\ \ \ \ Moreover, the current grid security solutions require each user
+to possess a X.509 certificate. But the process to get a certificate
+may take quite long time, since the requester needs to be checked and
+approved by the Registration Authority (RA) and then the Certificate
+Authority (CA) can issue the certificate according to the approval from
+RA. Meanwhile the process to use the certificate is also not so easy
+for the wider less-IT focused research community to deal with, since
+the user need to generate the proxy certificate by using the grid
+specific command lines, such as \textit{grid-proxy-init}, and
+\textit{voms-proxy-init}. On the other hand, users could most normally
+have some local community/institutional credentials such as
+username/passwords to which users are more familiar. To enable users
+use their own community credential instead of the X.509 certificate to
+access grid is a promising solution to make the grid be more easily
+accessible.
+
+\ \ \ \ The above analysis shows that in order to achieve
+interoperability and accessibility for user to access grid system, we
+need to address the mismatch between the different security solutions
+from different middlewares while keeping the changes as minimum as
+possible and the benefits as much as possible; also need to address the
+mismatch between user{\textquotesingle}s normal experience and
+middleware requirement in order to ensure the access as simple as
+possible.
+
+\ \ The approach in the new version of ARC is to utilize the existing
+Web Service standards, such as SAML and WS-Security, to achieve the
+interoperability and accessibility without breaking existing grid
+specific security protocols.
+
+\ The rest of this paper is organized as follows: Section 2 shows the
+ARC grid middleware, especially the architecture if the new version of
+ARC. Section 3 describes the solution about cross-domain authentication
+and single sign-on is presented, including a few solutions implemented
+for bridging the mismatch in terms of authentication. \ Section 4
+describes the implementation about standardized attribute querying for
+authorization. Section 5 discusses the related research, and Section 6
+concludes.
 
@@
 \bigskip
+
+\bigskip
+
+Based on OASIS WS-Trust 1.3 standard
+
+\bigskip
+
+\bigskip
+
+Allows us to bridge security domains in a scalable manner
+
+Allows the use of multiple token formats
+
+pick which is best/supported for your product
+
+Privacy can be protected across applications because:
+
+\ Each application can get a different, opaque, ID
+
+\ Each application receives only the attributes it needs
+
+\ \ \ \ Attributes may be retrieved from places like VOMS, GUMS, etc.
+
+\ \ \ \ When one application requests a token for use with another that
+token \ may be encrypted so that only the receiving app may view its
+data
+
+The STS serves as a point where policy decisions, such as \ whether App1
+can talk to App2, can be made.
+
+Removes the need for proprietary token related protocols
+
+\ \ Supports a superset of current operations and works for more than
+\ one token format
+
+\bigskip
+
+\bigskip
+
+While we started our work on the STS in the grid space it \ became clear
+it had other uses:
+
+\ \ A portal wants to connect to a SOAP service to get information
+\ about you but needs a SAML assertion to do it
+
+\ \ A user wants to connect to an AFS server in a different domain
+\ using the credentials from their domain
+
+\ \ A user has a SAML assertion but needs a Kerberos ticket to
+\ communicate with SharePoint
+
+There are a number of token formats out there. \ Chances \ are high,
+amongst all the software you have, more than one \ format is supported.
+
+How will you prepare for the next greatest format to hit the \ street?
+
+\bigskip
+
+\bigskip
+
+\bigskip
+
+\bigskip
+
+In ARC1, besides that GSI is supported for talking with external grid
+services which are based on GSI, the standard TLS/SSL is also
+supported.
 
@@
 
-\bigskip
-
 HED contains a framework for implementing and enforcing authentication
@@
 
+As explained above, HED is supposed to act as a service holder, but on
+the client side, from implementation perspective, there is similar
+architecture implemented for processing messages from different
+protocols and handling security functionality, also there is a specific
+application programming interface (API) for developers to easily write
+Web Service client code to service.
+
 \bigskip
@@
 
+For cross domain authentication, our work comprises the following five
+major aspects:
+
+\liststyleLiii
+\begin{itemize}
+\item \textit{Integration of community authentication} -{}-{}- to
+utilize the standardized community authentication from Shibboleth
+Identity Provider software when users authenticate against Grid
+services.
+\item \textit{Short lived credential service} \ {}-{}-{}- to issue the
+short lived X.509 certificate for user by using the community
+authentication, in order to interoperate with services that depends on
+PKI based mutual authentication.
+\end{itemize}
+
+\bigskip
+
+\ \ 3.1 Community based authentication for Grid by utilizing Shibboleth
+Identity Provider
+
+AAI (Authentication and authorization Infrastructure) is a solution for
+the authentication and authorization for inter-organization resource
+sharing, such as electronic resource sharing between libraries, etc.
+AAI implicitly applies to community/institutional/organization based
+authentication where users are from different home communities but need
+to get resources from other communities by using some federation
+mechanism. Unlike the X.509 based authentication solution in current
+grid systems, AAI does not require user to provide X.509 certificate,
+instead, it can support different types of authentication, such as
+username/password based authentication, IP address based
+authentication, etc. There are a few implementations about AAI, among
+which Shibboleth [ref] is one implementation which has been widely
+deployed.
+
+Shibboleth provides cross-domain single sign-on and attribute-based
+authorization while preserving user privacy. Shibboleth is based on the
+OASIS Security Assertion Markup Language (SAML), especially the \ new
+version of Shibboleth supports SAML 2.0 specification. For
+authentication, the main SAML profile that Shibboleth implements is the
+SAML2.0 web browser SSO profile, which \ define two functional
+components, an Identity Provider and a Service Provider. The Identity
+Provider (IdP) is responsible for creating, maintaining, and managing
+user identity, while the Service Provider (SP) is responsible for
+controlling access to services and resources by using the SAML
+assertion produced and issued by IdP upon request. In order to discover
+which home community does a user from, Shibboleth specifies an optional
+third component called {\textquotedblleft}Where Are You
+From?{\textquotedblright} (WAYF) service to aid in the process of IdP
+discovery, and this IdP discovery process is also standardized and
+\ defined in SAML 2.0 specification and called as
+{\textquotedblleft}Identity Profile Discovery
+Profile{\textquotedblright}.
+
+We utilized the SAML2.0 web browser SSO profile for the authentication
+in ARC middleware. But since the SSO profile is initially supposed to
+protect web applications and provide authentication for web users, we
+implemented some external code on the client and service side to
+integrate SSO profile. On the client side, apart from the WS-Client API
+for writing Web Service client, we implemented the user agent
+functionality of web browser in order to mimic the behavior of web
+browser{\textquotesingle}s user agent, such as http redirection, cookie
+processing, etc. In fact, the implementation of user agent is also
+based on the client interface of ARC, specifically, the https client
+interface, since the client interface of ARC can support different
+protocols which are incarnated by different message chain component
+(MCC). Hence client developers who would use SAML2.0 SSO profile should
+call the user agent interface and then the WS-Client interface. On the
+service side, we implemented the Service Provider functionality (based
+on the HTTP MCC configured together with TLS MCC) which is called SP
+Service. For Identity Provider, the Shibboleth IdP implementation is
+used.
+
+Figure 2 shows the process of SAML2.0 SSO integrated in ARC client and
+service.
+
+\begin{center}
+\includegraphics[width=3.398in,height=2.1555in]{Secpaper-img2.png}
+\end{center}
+
+\bigskip
+
+{\centering
+Figure 2. SAML2.0 SSO profile in ARC
+\par}
+
+The steps shows in Figure 2 are described as follows:
+
+\liststyleLv
+\begin{enumerate}
+\item The client use the user agent interface to launches a http request
+including the IdP name (to which the user belongs) to the service side.
+The host-name and port of destination URL is the same as that of the
+real service, the last part of the destination URL is
+{\textquotedblleft}saml2sp{\textquotedblright} which is specific for
+pointing to the Service Provider. Note that we use Identity Provider
+(IdP) name here to simplify the IdP discovery process, because we
+suppose that user who would access the service should better know where
+is he from initially.
+\item The SP Service (Service Provider) searches the metadata (we use
+the same metadata with the same format defined in Shibboleth) and gets
+the location of the single sign-on service (hosted in IdP) and also the
+location of assertion consuming service (hosted in this SP itself) in
+order to compose the SAML {\textless}AuthnRequest{\textgreater}
+message. Then SP Service issues this
+{\textless}AuthnRequest{\textgreater} message by using its own X.509
+certificate (Note in the SAML SSO profile, certificate is still needed
+by IdP and SP) and sends back to user agent.
+\item User agent sends the {\textless}AuthnRequest{\textgreater} message
+to Identity Provider.
+\item Identity Provider requires an act of authentication. The
+authentication mechanism is outside of the SAML2.0 SSO profile. The
+shibboleth IdP implementation will choose some login handler for
+authentication. The current user agent implementation is compatible to
+the Username/Password login handler of Shibboleth IdP. Through the http
+protocol, user agent will feed IdP with the username/password which has
+been given by the caller of user agent interface.
+\item Once the authentication has been succeeded, the IdP issues a SAML
+response including an encrypted (encrypted by destination
+SP{\textquotesingle}s public key) SAML assertion, and then this SAML
+response will be delivered by the user agent to the Service Provider. 
+\item The SP Service verifies and checks the SAML response, and decrypts
+and stores the SAML assertion into session/connection context. The SAML
+assertion actually includes the {\textless}AuthnStatement{\textgreater}
+and {\textless}AttributeStatement{\textgreater}.
+\item The WS-Client launches the real Grid/Web Service request via the
+same connection as the one which is used by user agent to contacts SP
+Service.
+\item The Grid/Web Service checks the
+{\textless}AuthnStatement{\textgreater} from the session context to see
+if the session is still valid through the \textit{SecHandler} called
+{\textquotedblleft}SAML SecHandler{\textquotedblright}. And if valid,
+service handles the real service processing and returns the response to
+WS client. Note that service requires WS client is from the same
+connection as the one on which user agent contact SP service in order
+to guarantee that the validity of SSO profile result effects the real
+client/service interaction.
+\end{enumerate}
+The SP service and real functional service(s) are hosted by the same
+container, and they use the same X.509 credential. The client side
+authentication is switched off, so that client doesn{\textquotesingle}t
+need to use any X.509 credential. Only the trusted certificates (CA
+certificate for SP and IdP) need to be configured for client side so
+that SP and IdP can authenticate themselves to the client. As required
+by SAML2.0 profile, the SP and IdP should have trust relationship to
+each other.
+
+Since the Shibboleth implementation of SAML is standard-compliant and
+widely deployed, the solution implemented in ARC can easily
+interoperate with other SAML implementations with minimum changes, and
+more importantly, this solution can succeed to utilize the widely
+deployed SAML implementation for authentication in grid systems by
+avoiding the usage of X.509 certificate.
+
+Moreover, even though the implementation is based on ARC middlerware,
+the idea can be adopted by other grid middlewares if they can only
+require server SSL authentication instead of mutual SSL authentication.
+
+\bigskip
+
+\ \ 3.2 Short lived credential service
+
+However, the \ \ Based on the solution described in Section 3.1,
+
+\bigskip
+
+\begin{center}
+\includegraphics[width=4.2134in,height=3.8972in]{Secpaper-img3.png}
+\end{center}
+{\centering
+Figure 3. Short lived credential service
+\par}
+
+\bigskip
+
+how to compose the DN
+
+\bigskip
+
+command line
+
 \bigskip
@@
 
-\ \ 3.1 Community based authentication: Utilizing Shibboleth IdP
-
-\bigskip
-
-\ \ 3.2 Short-lived certificate service
-
 \bigskip
@@
 \ \ 3.3 Certificate delegation service
+
+\bigskip
 
@@
 \bigskip
-
-\ \ \ \ \ \ \ \ \ trust relationship
 
@@
 
+the CCGrid 06 paper: Shibboleth-Protected privilege management
+infrastructure
+
+\bigskip
+
 \bigskip
@@
 
+Foster I, Kesselman C, Tuecke S. The anatomy of the Grid: Enabling
+scalable virtual organizations. International Journal of Supercomputer
+Applications 2001; 15(3):200--222.
+
+\bigskip
+
+The Shibboleth Project.
+\href{http://shibboleth.internet2.edu/}{Http://shibboleth.internet2.edu/}
+
 \bigskip

### 2008-12-03 doc/infosys/TechnicalHandbook/ISIS - TechnicalHandbook.tex
ISIS Technical handbook - half ready

%\documentclass{book}
\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://download.nordugrid.org/}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{logo-knowarc.png}}
\end{tabular}

\hrulefill

%-------- Change this to Knowarc_D<w>.<d>-<n>_<yy>

%{\raggedleft Knowarc\_D0.0-0\_00\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large ISIS Information Indexing Service}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Technical handbook}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
{\centering \large Ivan Marton\footnote{martoni@niif.hu} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}
\tableofcontents                          %Comment if use article style
\newpage
%\chapter{Preface}
\section{Introduction}
\label{Introduction}

\section{Components}
\label{Components}
\subsection{Registrant}
\label{Registrant}
The Registrant is active module of the HED (Hosting Environment Daemon) which is bound to a set of ISIS's. In practice, the configuration part of the Registrant contains exactly one ISIS to bind, and the Registrant will collect the necessary information about the other ISIS's belonging to the same network.\par
If the administrator wants to use more then one network for registering his services then more then one Registrant have to be instanced for this purpose. In this case, the default Registrant will be used for registering every services that isn't configured in a different way. \par
The registration can be done once or periodically. It based on the configuration of the Registrant as default behavior but can be overwritten for every service separately.\par
The Registrant is doing the message aggregation of every service belongs to it if it's possible. The algorithm of the Registrant is following:
\begin{framed}
  Registrant - pseudo algorithm\\
  \\
  \verb#// Initialize phase#\\
  \verb# Read the configuration and store the information about the services in a list#\\
  \verb# do { // In a different Thread#\\
  \verb# wake_up_time = now();#\\
  \verb# messages = null;#\\
  \verb#  if ( 0 < count(service where service.next_run <= wake_up_time)) {#\\
  \verb#    foreach( service where service.next_run <= wake_up_time) {#\\
  \verb#      messages.add(service.RegistrationCollector);#\\
  \verb#      service.next_run = wake_up_time + service.period;#\\
  \verb#    }#\\
  \verb#    if (0 < count(messages)) {#\\
  \verb#      sent_message = assemble message with headers(messages);#\\
  \verb#      send(sent_message);#\\
  \verb#    }#\\
  \verb#  } else {#\\
  \verb#    sleep(min(service.next_run) - now()); #\\
  \verb#  }#\\
  \verb#} while(true)#\\
\end{framed}
There is a minimum value of the period defined. If the settings is less then 2 minutes then it will be replaced with this value.
Before sending the information received from the services (\textbf{Service Advertisement}, see Section \ref{Service Advertisement}), the Registrant extend them some additional data (\textbf{Service Advertisement Metadata}). This part of the message contains the following information:
\begin{itemize}
  \item GenTime: (Generation Time) The actual timestamp (called wake\_up\_time in the pseudo code)
  \item Expiration: (The time of Service Advertisement's expiration.)  Generation Time + Renewal period length.
  \item Source: The Endpoint of the Service
  \item Status: Status of the Service Advertisement.
\end{itemize}
A \textbf{Registration Entry} contains a \textbf{Service Advertisement} and a corresponding \textbf{Service Advertisement Metadata}.\par
The \textbf{Registration Message} is the one that will be sent to an ISIS service. This message is consists of more \textbf{Registration Entry} and at most \textbf{Registration Header}.  This embedded structure can be shown on Figure \ref{fig:RegistrationMessage}.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{RegistrationMessage.png}}}
\caption{\label{fig:RegistrationMessage}Embedded structure of Registration Message} }
\end{figure}
\subsection{Services}
\label{Services}
The service developers have to ensure that the services are able to provide the necessary information about themselves. It can be easily done by implementing the proper interface of the Arc::Service class. It means to fill the RegistrationCollector function with the necessary code parts to assemble the always up-to-date status information about the service and anything else wants to be advertised. This information package is called \textbf{Service Advertisement}.\par
\label{Service Advertisement}
The \textbf{Service Advertisement} can contains any information wants to advertised but the mandatory elements are the following attributes these one have to be present:
\begin{itemize}
  \item Service ID: A globally unique identifier.
  \item Service Type: The Glue2 type of service.
  \item Endpoint URL: The URL where the service can be achieved.
\end{itemize}
\subsection{ISIS Service}
\label{ISIS Service}
The ISISs perform two completely different function. On the one hand they are working as ordinary HED services, and on the other hand maintain a peer-to-peer network between each other.
\subsubsection{Service functions}
\label{ISIS Service functions}
Every ISIS service has two different interfaces:
\begin{itemize}
  \item WSRF interface
  \item service specific interface
\end{itemize}
The WSRF interface returns the service specific information expect of the stored database. (The Property Document doesn't contains the Activities that are the Registration Entries in this case.) The service specific interface is for retrieve these data. This interface has 5 operation defined:
\begin{itemize}
  \item Register\par
  The register operation is accessed by the Registrant. See Section \ref{Registration} for further details.
  \item RemoveRegistration\par
  This operation is used for requesting the removal of a list of zero or more Registration Entries stored in the Information System. The attribute of this operation is a list of Service IDs. Every entry belonging to any of the given Service ID will be removed.
  \item GetRegistrationStatuses\par
  This operation is used for requesting the current status(es) of zero or more Registration Entries previously pushed into the system by the Register operation. The operation receives a list of Service ID as its input and returns a list of <ServiceID, Status> pairs.
  \item GetISISList\par
  The operation is used for obtaining a list of known ISIS instances from any particular ISIS. Clients can then use the obtained list to run direct queries against the ISIS instances.  The operation is provided for fault tolerance and for providing optional performance boost. This operation returns the known peer-to-peer neighbors. If you are interested in every known ISIS then they can be queried with the Query operation because they are ordinary services registered into the Information System.
  \item Query
  The operation is used for request any kind of queries related to the Indexing Database. For providing as much flexibility as possible any XQuery searches can be executed. This operation is accessed by clients. For an example and further details see Section \ref{Query data}. 
\end{itemize}

\subsubsection{Peer-to-Peer functions}
\label{ISIS Peer-to-Peer functions}
\subsection{Client}
\label{Client}
The client part of the information service is for querying the data stored in the database of the ISIS services. This is implemented as a part of the ARCLib and using the ISIS service's interface for the various data queries. For further details see Section \ref{Query data}.  
\section{Operations}
\label{Operations}
The different operations are performed by different Information system components. These cases are described below in the dedicated subsections. 
\subsection{Registration}
\label{Registration}
The registration operation is done by the Registrants and described in Section \ref{Registrant}. An example layout can be shown on Figure \ref{fig:RegistrationMessage}. In this configuration there is two different Registrant configured in one HED for three services. Let the \textit{Registrant A} the default Registrant and the services configured in the following way:\par
\begin{itemize}
  \item Service 1: There is no Registrant configured so the default one will register it.
  \item Service 2: The \textit{Registrant A} is configured explicitly.
  \item Service 3: The \textit{Registrant B} is configured explicitly.
\end{itemize}
In the sketched case the Service 1 and 2 are handled equivalently by \textit{Registrant A} and Service 3 by \textit{Registrant B}.\par
In the first step the information collection from the first handled services is done parallel (\textit{A1}-\textit{A2}, \textit{B1}-\textit{B2}). The retrieved data are stored and other services are queried if there is any (\textit{A3}, \textit{A4}). Finally, the aggregated information is sent to the corresponding ISISs independently from each other (\textit{A5}, \textit{B5}).\par
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{Registration.png}}}
\caption{\label{fig:RegistrationMessage}Overview of the registration operation} }
\end{figure}
This registration operation is done once during the start-up phase and cyclic later by right of the (per service) configured periods. This periodical operation follows the algorithm described in Section \ref{Registrant}, and do every possible message aggregation.
\subsection{Query data}
\label{Query data}

%Advanced Resource Connector (ARC) was invented by
%NorduGrid~\cite{nordugrid}. This \textit{middleware}\index{middleware}
%uses \GT~4 by Globus.

%In expressions, the following operands are allowed:
%\begin{shaded}
%  \verb#=   !=   >   <   >=   <=#
%\end{shaded}

%\begin{framed}
%  Examples of URLs are:\\
%  \\
%  \verb#http://grid.domain.org/dir/script.sh#\\
%  \verb#gsiftp://grid.domain.org:2811;threads=10/dir/input_12378.dat#\\
%  \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
%  \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#
%  \verb#file:///home/auser/griddir/steer.cra#\\
%\end{framed}

%\subsection{Subsection}
%\label{sec:subsection}

%\begin{figure}[ht]
%\centering{{\scalebox{0.9}{\includegraphics{logo-knowarc.png}}}
%\caption{\label{fig:myfigure1}The figure shows a logo.} }
%\end{figure}

\bibliography{grid}
\end{document}

### 2008-12-05 doc/sec/Sec_paper.tex
More sections of the sec doc

@@
 \renewcommand\labelitemiv{${\bullet}$}
+}
+\newcommand\liststyleLiv{%
+\renewcommand\theenumi{\arabic{enumi}}
+\renewcommand\theenumii{\arabic{enumii}}
+\renewcommand\theenumiii{\arabic{enumiii}}
+\renewcommand\theenumiv{\arabic{enumiv}}
+\renewcommand\labelenumi{\theenumi.}
+\renewcommand\labelenumii{\theenumii.}
+\renewcommand\labelenumiii{\theenumiii.}
+\renewcommand\labelenumiv{\theenumiv.}
 }
@@
 {\centering
-Towards cross-domain authentication and standardized attribute-based
-authorization for ARC grid middleware
+Towards cross-middleware authentication and single sign-on for ARC grid
+middleware
 \par}
@@
 \ In order to make the access to Grids as simple as possible, security
-including authentication and authorization is one of the most important
-challenges in production Grids, especially when the Grid applications
-are across multiple administration domains as well as across
-heterogeneous Grid middlewares, such as some widely scaled eScience
-applications which needs to coordinate resources sharing among
-collection of autonomous institutions with different Grid middlewares
-running on these resources. In this paper, we describe the security
-implementation and consideration in new version of Advanced Resource
-Connector (ARC) middleware, where the above issues have been exploited.
-The main goal of ARC implementation in terms of security is to let the
-middleware be capable to \ interoperate with other Grid middlewares by
-leveraging standard specifications. The key aspect of the work we have
-done is to enhance the current proxy certificate based delegation and
-authentication by utilizing the Security Assertion Markup Languages
-(\textit{SAML}) single sign-on (SSO) profile, WS-Security in order to
-achieve cross-domain authentication, as well as to enhance the
-Attribute Certificate (AC) based authorization by utilizing the SAML
-\ attribute query profile in order to achieve standardized
-attribute-based authorization. \
+is one of the most important challenges in production Grids, especially
+when the Grid applications are across multiple administration domains
+as well as across heterogeneous Grid middlewares, such as some widely
+scaled eScience applications which needs to coordinate resources
+sharing among collection of autonomous institutions with different Grid
+middlewares running on these resources. In this paper, we describe the
+security implementation and consideration in new version of Advanced
+Resource Connector (ARC) middleware, where the above issues have been
+exploited. The main goal of ARC implementation in terms of security is
+to let the middleware be capable to \ interoperate with other Grid
+middlewares by leveraging standard specifications. The key aspect of
+the work we have done is to enhance the current proxy certificate based
+delegation and authentication by utilizing the Security Assertion
+Markup Languages (\textit{SAML}) single sign-on (SSO) profile,
+WS-Security in order to achieve cross-domain authentication. \
 
@@
 and single sign-on is presented, including a few solutions implemented
-for bridging the mismatch in terms of authentication. \ Section 4
-describes the implementation about standardized attribute querying for
-authorization. Section 5 discusses the related research, and Section 6
-concludes.
-
-\bigskip
-
-\bigskip
-
-\bigskip
-
-Based on OASIS WS-Trust 1.3 standard
-
-\bigskip
-
-\bigskip
-
-Allows us to bridge security domains in a scalable manner
-
-Allows the use of multiple token formats
-
-pick which is best/supported for your product
-
-Privacy can be protected across applications because:
-
-\ Each application can get a different, opaque, ID
-
-\ Each application receives only the attributes it needs
-
-\ \ \ \ Attributes may be retrieved from places like VOMS, GUMS, etc.
-
-\ \ \ \ When one application requests a token for use with another that
-token \ may be encrypted so that only the receiving app may view its
-data
-
-The STS serves as a point where policy decisions, such as \ whether App1
-can talk to App2, can be made.
-
-Removes the need for proprietary token related protocols
-
-\ \ Supports a superset of current operations and works for more than
-\ one token format
-
-\bigskip
-
-\bigskip
-
-While we started our work on the STS in the grid space it \ became clear
-it had other uses:
-
-\ \ A portal wants to connect to a SOAP service to get information
-\ about you but needs a SAML assertion to do it
-
-\ \ A user wants to connect to an AFS server in a different domain
-\ using the credentials from their domain
-
-\ \ A user has a SAML assertion but needs a Kerberos ticket to
-\ communicate with SharePoint
-
-There are a number of token formats out there. \ Chances \ are high,
-amongst all the software you have, more than one \ format is supported.
-
-How will you prepare for the next greatest format to hit the \ street?
-
-\bigskip
-
-\bigskip
-
-\bigskip
-
-\bigskip
-
-In ARC1, besides that GSI is supported for talking with external grid
-services which are based on GSI, the standard TLS/SSL is also
-supported.
+for bridging the mismatch in terms of authentication. Section 4
+discusses the related research, and Section 5 concludes.
 
@@
 Another configuration could be SOAP over HTTPG (HTTP over GSI) which
-can be user to interoperate with some service based on HTTPG such as
+can be used to interoperate with some service based on HTTPG such as
 SRM (Storage Resource Manager) [ref] service. The easily configurable
@@
 PKI based mutual authentication.
+\item \textit{Credential delegation service} -{}-{}- to process the
+X.509 credential delegation as a normal Web Service, in order to
+provide single sign-on characteristic.
+\item \textit{Message level authentication} -{}-{}- to provide message
+level authentication by implementing WS-Security.
+\item 
+\bigskip
 \end{itemize}
@@
 \begin{center}
-\includegraphics[width=3.398in,height=2.1555in]{Secpaper-img2.png}
+\includegraphics[width=3.4063in,height=2.1417in]{Secpaper-img2.png}
 \end{center}
+
+\bigskip
 
@@
 
-\liststyleLv
+\liststyleLiv
 \begin{enumerate}
@@
 
+One benefit of SAML2.0 SSO profile that is worth mentioning is that: the
+Identity Provider could cache the authentication result through session
+management once the user agent has succeeded to authenticate, then in a
+short period this authentication result is valid so that user agent
+don{\textquotesingle}t need to feed IdP with user{\textquotesingle}s
+username and password the next time (if the time is not out of the
+scope of valid period) it authenticates against IdP. So user (or the
+client on behalf of this user) can travel across multiple security
+domains with only inputing his name and password once, which is the
+characteristic of single sign-on.
+
 Since the Shibboleth implementation of SAML is standard-compliant and
@@
 
-However, the \ \ Based on the solution described in Section 3.1,
+However, most of the widely used grid middlewares are based on GSI while
+GSI requires mutual authentication. Also for Web Service based grid
+solution, we can not prevent service side from requiring client X.509
+certificate. \ Based on the solution described in Section 3.1, we
+implemented a short lived credential service (SLCS) by which user can
+get a short-lived X.509 certificate without being bothered to contact
+any registration authority (RA) or certificate authority (CA).
+
+The SLCS service is actually a specific Web Service (standard web
+service implemented by using ARC service interface) in Figure 2, and
+the SLCS client is a specific ARC client which includes the user agent
+and WS client. The whole process from step 1 to step 8 is the same as
+in Figure 2.
 
@@
 \begin{center}
-\includegraphics[width=4.2134in,height=3.8972in]{Secpaper-img3.png}
+\includegraphics[width=3.7689in,height=3.2055in]{Secpaper-img3.png}
 \end{center}
@@
 
-\bigskip
+\ \ \ SLCS client generates a X.509 certificate request, launches a Web
+Service request which includes the certificate request; SLCS service
+then gets the certificate request, composes a distinguished name (DN),
+issues a certificate (short lived, 12 hours by default) with the SAML
+attribute (from the SAML2.0 SSO profile) as the X.509 certificate
+extension, and puts the certificate in to the Web Service response;
+SLCS client get the response and stores the X.509 certificate into
+local repository.
 
-how to compose the DN
+The command line interface for the SLCS client is like this:
+
+\$ ./arcslcs -S https://127.0.0.1:60000/slcs -I
+https://idp.testshib.org/idp/shibboleth -U myself -P myself
+
+\ \ \ Since the lifetime of the short lived credential is normally
+short, it is not a must to protect the private key by pass-phrase. As
+illustrated in step a and step b in Figure 3, if the private key is not
+protected, through Web Service client, the user can use the X.509
+certificate to access Grid Service or Web Service from any kind of
+middleware. If the private key is protected, the can use the X.509
+certificate to generate a proxy certificate (by using command line
+interface utility such as \textit{grid-proxy-init},
+\textit{voms-proxy-init}, or \textit{arcproxy}), and the use the proxy
+certificate to access Grid Service or Web Service.
+
+\ \ \ It is worth mentioning that since ARC middleware can support GSI
+communication by configuring the GSI MCC, \ together with the X.509
+certificate, the Web Service client developed by ARC client interface
+can interoperate with Grid Service that requires GSI communication.
+
+\ \ \ It might be noticed that how to compose the distinguished name
+(DN) for the certificate is critical issue for SLCS service. Since the
+Shibboleth Identity Provider uses the eduPerson schema [ref] for the
+definition of {\textless}saml:Attribute{\textgreater} in
+{\textless}saml:AttributeStatement{\textgreater}, we pick the
+relatively distinguishable attribute
+{\textquotedblleft}eduPersonPrincipalName{\textquotedblright} for the
+DN. A typical \ eduPersonPrincipalName \ value could be
+\href{mailto:alice@example.org}{alice@example.org}, then the DN is
+{\textquotedblleft}/O=knowarc/OU=example.org/CN=alice{\textquotedblright}.
+
+The obvious benefit of SLCS service is that: If a user passes the
+authentication to his home Identity Provider, he can get the X.509
+credential at anywhere only by running the SLCS client command together
+with inputing his username and password to this home IdP, and then
+access the grid system conveniently; meanwhile, thanks to the single
+sign-on characteristic discussed in Section 3.1, this user
+\ doesn{\textquotesingle}t not need to input his username and password
+in a valid period after the first time he succeeds to authenticate
+against his home IdP by running the SLCS client command on the same
+node, even if this SLCS client command is supposed to run against a few
+SLCS services to get a few SLCS credentials.
 
@@
 
-command line
+\ \ 3.3 X.509 Credential delegation service
 
-\bigskip
+The delegation of credentials proposed in Grid Security Infrastructure
+(GSI) is a good solution for the supporting of single sign-on for
+users, except the coupling of the GSI communication which is not so
+widely accepted in Web Service applications. In the design of the new
+version ARC middleware, we keep the credential delegation idea, while
+utilizing the SOAP communication (and normal TCP or TLS for transport
+level communication) instead of the GSI communication for the
+credential delegation process.
 
-\bigskip
+As shown in Figure 4, there is a specific WS client and Web Service as
+well for processing delegation: delegation client, and delegation
+service. The delegation service has two main wsdl operations, one for
+processing the delegation initiation; the other for processing/storing
+the signed proxy certificate. The delegation process is detailed in
+Figure 5.
 
-\bigskip
+For the service invocation process, the WS client functionality (inside
+each service or the initial client) which is included to access the
+target service (step a1 or a2 in Figure 4), should use the
+user{\textquotesingle}s X.509 certificate or proxy certificate (on
+behalf of the user) for authentication and secure communication. For
+the delegation process, the principle is the same.
 
-\ \ 3.3 Certificate delegation service
+Since each WS client accesses the target service (delegation service or
+other functional service) on behalf of the user, the trust relationship
+between the user{\textquotesingle}s certificates (X.509 certificate and
+proxy certificate) and services is required, while the trust
+relationship between service and service is not required.
 
-\bigskip
-
-\bigskip
-
-\ \ 3.4 From transport level security to message level security: X.509
-Token and SAML Token
-
-\bigskip
-
-\bigskip
-
-\ \ 3.5 Delegation between WS-Security tokens
+\begin{center}
+\includegraphics[width=3.802in,height=3.0689in]{Secpaper-img4.png}
+\end{center}
 
@@
 {\centering
-4. Standardized Attribute-based authorization 
+Figure 4. Credential delegation service
 \par}
@@
 
-4.1 SAML Attribute Query profile
+\begin{center}
+\includegraphics[width=2.5319in,height=1.8571in]{Secpaper-img5.png}
+\end{center}
+{\centering
+Figure 5. Credential delegation process
+\par}
 
@@
 
-4.2 Client to VOMS Attribute authority Service
+\ \ 3.4 Message level authentication
+
+Transport level security (TLS or SSL) is not sufficient for securing Web
+Services, because the secure channel created by SSL/TLS make it
+impossible to differentially protect the SOAP message, e.g. \ to
+encrypt and/or sign only particular components of the SOAP message,
+which is relevant when non-sensitive portions of the message need to be
+accessed or changed by intermediate actors. Additionally, SSL/TLS is
+incapable of providing end-to-end protection for SOAP message which
+might flow through multiple intermediate actors, because it only allows
+each hop to be protected with the resultant security gaps at
+intermediate actors. Web Services Security (WS-Security) specification
+(ref) provides means for applying security to Web Service.
+
+The WS-Security related specifications are implemented in ARC as
+\textit{SecHandler}. There are three kinds of WS-Security version 1.1
+specifications are completely or partly implemented: Username Token
+profile, X.509 Token profile, and SAML Token profile.
+
+For those WS clients on behalf of a user, at any step of service
+invocation, the WS client can augment the SOAP message with WS-Security
+token by simply configuring the WS-Security \textit{SecHandler} into
+container{\textquotesingle}s configuration, if the relying party
+(target service) requires.
+
+If the WS client owns an X.509 certificate or proxy certificate, the
+process of generating an X.509 Token for a SOAP message is: the client
+puts the content of certificate into X.509 token as
+{\textless}wsse:BinarySecurityToken{\textgreater}, and uses the private
+key (corresponding to the certificate) to create XML signature for the
+SOAP message body.
+
+However the process of generating a SAML Token is more complicated. In
+SAML Token profile, \textit{ConfirmationMethod} is the verifying method
+for establishing proof-of-possession for the subject. Two methods are
+proposed in SAML Token profile: \textit{holder-of-key},
+\textit{sender-vouches}. \ We rely on the \textit{holder-of-key
+}method. For the \textit{hold-of-key} subject confirmation method,
+there are three interaction parts: the attesting entity, the relying
+party and the issuing authority. Firstly the attesting entity
+authenticates to issuing authority by using some authentication scheme
+such as X.509 Token profile. So then issuing authority is able to make
+a definitive statement (sign a SAML assertion) about an act of
+authentication that has already taken place. The attesting entity gets
+the SAML assertion and then signs the SOAP message together with the
+assertion by using its private key (the relevant certificate has been
+authenticated by issuing authority, and its relevant public key has
+been put into \textit{SubjectConfirmation}
+
+element under SAML assertion by issuing authority. Only the actual owner
+of the SAML assertion can do this, as only the subject possesses the
+private key paired with the public key in the assertion. This
+establishes an irrefutable connection between the author of the SOAP
+message and the assertion describing an authentication event.)
+
+\ The relying party is supposed to trust the issuing authority. When it
+receives a message from the asserting entity, it will check the SAML
+assertion based on its predetermined trust relationship with the SAML
+issuing authority, and check the signature of the soap message based on
+the public key in the SAML assertion without directly trust
+relationship with attesting entity (subject owner).
 
@@
 
-4.3 Attribute parsing and Policy decision point
+\ \ 3.5 SAML Token delegation service
+
+\ There is also delegation requirement on the message level, some
+entities (client or user) often need to temporarily delegate itself to
+another entity to perform action on their behalf. We propose a
+delegation mechanism for SAML Token.
+
+The delegation process is similar to the process for delegation proxy
+certificate, with the following differences:
+
+\liststyleLv
+\begin{enumerate}
+\item The communication between delegation client and service is based
+on mutual authentication between proxy certificate and service
+certificate, instead, it is based on mutual authentication between
+service certificate and service certificate.
+\item The delegation service will not generate any certificate request,
+instead, it picks the public key from its own service certificate, and
+sends back to delegation client.
+\end{enumerate}
+Figure 6 shows the token change due to delegation. On the left side, B
+is the identity of the attesting entity{\textquotesingle}s X509 token
+mentioned above, while A is the identity of issuing authority, and the
+SOAP message is signed by B{\textquotesingle}s private key. On the
+right side, C is the identity of the delegation service, ans now the
+SOAP message is signed by C{\textquotesingle}s private key.
+
+\bigskip
+
+\begin{center}
+\includegraphics[width=4.2571in,height=2.152in]{Secpaper-img6.png}
+\end{center}
+{\centering
+\ \ \ \ \ \ \ \ \ Figure 6. Token before and after delegation
+\par}
+
+\bigskip
 
@@
 {\centering
-5. Related work
+4. Related work
 \par}
-
-VOMS AA Service
 
@@
 
+Delegation service : gridsite
+
 \bigskip
@@
 {\centering
-6. Conclusion and Future work
+5. Conclusion and Future work
 \par}
@@
 \bigskip
+
+eduPerson and eduOrg Object shema.
+\url{http://middleware.internet2.edu/eduperson/}
+
+\bigskip
 \end{document}

### 2008-12-05 doc/sec/Sec_paper.tex
little changes in sec doc

@@
 level authentication by implementing WS-Security.
-\item 
-\bigskip
+\item \textit{SAML token delegation service} -{}-{}- to process the WS
+Security SAML token in order to provide single sign-on for SOAP message
+level authentication
 \end{itemize}
@@
 \end{enumerate}
-Figure 6 shows the token change due to delegation. On the left side, B
-is the identity of the attesting entity{\textquotesingle}s X509 token
-mentioned above, while A is the identity of issuing authority, and the
-SOAP message is signed by B{\textquotesingle}s private key. On the
-right side, C is the identity of the delegation service, ans now the
-SOAP message is signed by C{\textquotesingle}s private key.
+Figure 6 shows the SOAP message header is changed due to delegation. On
+the left side, B is the identity of the attesting
+entity{\textquotesingle}s X509 token mentioned above, while A is the
+identity of issuing authority, and the SOAP message is signed by
+B{\textquotesingle}s private key. On the right side, C is the identity
+of the delegation service (also C is the identity of the target service
+\ to which the B will access), and now the SOAP message is signed by
+C{\textquotesingle}s private key.
 
@@
 {\centering
-\ \ \ \ \ \ \ \ \ Figure 6. Token before and after delegation
+\ \ \ \ \ \ \ \ \ Figure 6. SOAP message header with SAML delegation
+token inside
 \par}

### 2008-12-05 doc/sec/Sec_paper.tex
Some content improvement, and reference for Sec doc

@@
 \usepackage[pdftex]{graphicx}
+% Text styles
+\newcommand\textstyleStrongEmphasis[1]{\textbf{#1}}
+\newcommand\textstyleCitation[1]{\textit{#1}}
 % List styles
@@
 
-\ In order to make the access to Grids as simple as possible, security
-is one of the most important challenges in production Grids, especially
-when the Grid applications are across multiple administration domains
-as well as across heterogeneous Grid middlewares, such as some widely
-scaled eScience applications which needs to coordinate resources
-sharing among collection of autonomous institutions with different Grid
-middlewares running on these resources. In this paper, we describe the
-security implementation and consideration in new version of Advanced
-Resource Connector (ARC) middleware, where the above issues have been
-exploited. The main goal of ARC implementation in terms of security is
-to let the middleware be capable to \ interoperate with other Grid
-middlewares by leveraging standard specifications. The key aspect of
-the work we have done is to enhance the current proxy certificate based
-delegation and authentication by utilizing the Security Assertion
-Markup Languages (\textit{SAML}) single sign-on (SSO) profile,
-WS-Security in order to achieve cross-domain authentication. \
+\ \ \ \ In order to make the access to Grids as simple as possible,
+security is one of the most important challenges in production Grids,
+especially when the Grid applications are across multiple
+administration domains as well as across heterogeneous Grid
+middlewares, such as some wide scale eScience applications which needs
+to coordinate resources sharing among collection of autonomous
+institutions with different Grid middlewares running on these
+resources. In this paper, we describe the security implementation and
+consideration in new version of Advanced Resource Connector (ARC)
+middleware, where the heterogeneity issue have been exploited. The main
+goal of ARC implementation in terms of security is to let the
+middleware be capable of interoperating with other Grid middlewares by
+leveraging standard specifications. The key aspect of the work we have
+done is to enhance the current proxy certificate based authentication
+and single sign-on by utilizing and enhancing the standardized Web
+Service specifications such as Security Assertion Markup Languages
+(\textit{SAML}) single sign-on (SSO) profile, Web Services Security in
+order to achieve cross-middleware authentication and single sign-on.
 
@@
 \end{enumerate}
-Grids provides the technology for large scaled, cross domains
+\ \ \ \ Grids provides the technology for wide-scale, cross domains
 collaboration where the users work on resources and sharing data which
 are distributed across domains. This kind of collaboration is called
-virtual organization (VO)[ref1]. The resources that contribute to the
-VO could come from independent administrative domains, and even require
-different security approaches for control the access to themselves. The
-same heterogeneous situation even critically applies to users who need
-to access the VO, because users holding one kind of credential could
-need to travel through multiple resources with different security
-approaches. To solve the above issues, the Grid middleware should
+virtual organization (VO)[1]. The resources that contribute to the VO
+could come from independent administrative domains, and even require
+different security approaches for control the access to themselves
+because different middlerwares are used as platform. The same
+heterogeneous situation even critically applies to users who need to
+access the VO, because users holding one kind of credential could need
+to travel through multiple resources with different security
+approaches. To solve the above issues, the Grid middlewares should
 provide the following ability: The user which participates the VO
@@
 VO together with its resources should be able to easily manage the
-authorization, given the condition that the resources are actually
-owned by different administrative domains.
+authorization, given the condition that the resources are owned by
+different administrative domains.
 
-In grid community, Grid Security Infrastructure (GSI)[ref2] is the
+\ \ \ \ In grid community, Grid Security Infrastructure (GSI)[2] is the
 \textit{de facto} standard about authentication and \ transport level
-communication, which is builds on Public Key Infrastructure (PKI), an
+communication, which builds on Public Key Infrastructure (PKI), an
 architecture based on X.509 public key certificate. GSI implements some
@@
 trust federation and guarantee two different X.509 certificates from
-different CAs can accomplish authentication to each other. So if a user
-would access grid system, he/she should own a X.509 certificate which
-is issued by a CA that is trusted by other{\textquotesingle}s entity in
-the GSI based grid system.
+different CAs can accomplish authentication to each other. So in the
+GSI based grid system, if a user would access grid system, he/she
+should own a X.509 certificate which is issued by a CA that is trusted
+by other entity.
 
-In terms of the virtual organization issues mentioned earlier, GSI and
-some other related solutions like virtual organization management
-service (VOMS) [ref3] has partly solved the them in some sense. The
-identity delegation and X.509 proxy certificate initiated by GSI can
-achieve single sign-on, so that the need to re-enter the
+\ \ \ \ In terms of the virtual organization issues mentioned earlier,
+GSI and some other related solutions like virtual organization
+management service (VOMS) [3] have partly solved the them in some
+sense. The identity delegation and X.509 proxy certificate initiated by
+GSI can achieve single sign-on, so that the need to re-enter the
 user{\textquotesingle}s pass phrase for authentication can be avoided
-by creating the proxy certificate, even if some local or remote
-entities access access resources on behalf of this user. \ As the most
-successful grid authorization model, VOMS adopts X.509 Attribute
-Certificate[ref4]. Attribute certificate is signed by VOMS server and
-then binded to proxy certificates by client, \ and attributes
-certificate includes the VO memberships which is associated with the
-proxy certificate{\textquotesingle}s identity. Essentially the
-resources can make access control decision based on parsing the VO
-membership attributes from the homogeneous attribute certificate.
+by creating the proxy certificate. \ As the most successful grid
+authorization model, VOMS adopts Attribute Certificate[4]. Attribute
+certificate is signed by VOMS server and then binded to proxy
+certificates by client, \ and attributes certificate includes the VO
+memberships which is associated with the proxy
+certificate{\textquotesingle}s identity. Eventually the resources can
+make access control decision based on parsing the VO membership
+attributes from the homogeneous attribute certificate.
 
@@
 converged into Grid computing middlewares. The Web Service technology
-has been adopted by Globus toolkit, gLite, as well as the ARC.
-Actually, utilizing Web Service technology and providing seamless
-interoperability with other middlewares is one of the main goals of the
-new version of ARC middleware.
+has been adopted by Globus toolkit[5], gLite[6], as well as \ Advanced
+Resource Connector(ARC) [7]. In fact, utilizing Web Service technology
+and providing seamless interoperability with other middlewares is one
+of the main goals of the new version of ARC middleware.
 
-Unfortunately, the existing grid security solutions such as GSI and VOMS
-do not apply to Web Service \ based application very well due to the
-following reasons:
+\ \ \ Unfortunately, the existing grid security solutions such as GSI
+and VOMS do not apply to Web Service \ based application very well due
+to the following reasons:
 
@@
 \begin{itemize}
-\item Firstly, the common implementation for Web Service does not
-necessarily recognize and then properly verify and parse the
-certificate, since comparing with usual X.509 certificate, proxy
-certificate needs different treatment such as proxy path verification. 
-\item Secondly, the X.509 Attribute Certificate has not been adopted by
-any Web Service implementation at all. 
-\item Thirdly, what is even worse is that the GSS API based confidential
-communication has also not been adopted by Web Service implementation.
+\item Firstly, \ the GSS API based confidential communication has also
+not been adopted by Web Service implementation.
+\item Secondly, \ Attribute Certificate has not been adopted by any Web
+Service implementation at all.
+\item Thirdly, Web Service implementations commonly use the Web Services
+Security(WS-Security)[8], Security Assertion Markup Languages
+(SAML)[9], and standard SSL/TLS as well as standard X.509 certificate
+when considering security, which make them impossible to interoperate
+with the GSI based grid middlewares.
 \item Finally, identity delegation is a good solution for user or
-user{\textquotesingle}s representative (on behalf of the user) to
+user{\textquotesingle}s delegatee (on behalf of the user) to
 move/travel from one resource to another resource. But since the
@@
 \end{itemize}
-There is some effort to add GSI support to adapt to the strict security
-requirement including GSI from grid middlewares, such as GSI plug-in
-for gSOAP[ref6]. However, this kind of solution
-doesn{\textquotesingle}t help the existing grid middlewares
-interoperate with other common Web Service implementation. Instead, Web
-Service implementations commonly use the WS-Security, Security
-Assertion Markup Languages (SAML), and standard SSL/TLS as well as
-standard X.509 certificate when considering security. \
-
 \ \ \ \ Moreover, the current grid security solutions require each user
@@
 have some local community/institutional credentials such as
-username/passwords to which users are more familiar. To enable users
+username/passwords to which users are more familiar. To enable users to
 use their own community credential instead of the X.509 certificate to
@@
 \ \ \ \ The above analysis shows that in order to achieve
-interoperability and accessibility for user to access grid system, we
-need to address the mismatch between the different security solutions
-from different middlewares while keeping the changes as minimum as
-possible and the benefits as much as possible; also need to address the
-mismatch between user{\textquotesingle}s normal experience and
-middleware requirement in order to ensure the access as simple as
-possible.
+interoperability and accessibility for user to access grid system which
+is established on different middlewares, we need to address the
+mismatch between the different security solutions from these
+middlewares in order to ensure the interoperability as wide as
+possible; also need to address the mismatch between
+user{\textquotesingle}s normal experience and
+middleware{\textquotesingle}s requirement in order to ensure the access
+as simple as possible.
 
-\ \ The approach in the new version of ARC is to utilize the existing
-Web Service standards, such as SAML and WS-Security, to achieve the
-interoperability and accessibility without breaking existing grid
-specific security protocols.
+\ \ \ \ The approach in the new version of ARC is to utilize the
+existing Web Service standards, such as SAML and WS-Security, to
+achieve the interoperability and accessibility without breaking
+existing grid specific security protocols.
 
-\ The rest of this paper is organized as follows: Section 2 shows the
-ARC grid middleware, especially the architecture if the new version of
-ARC. Section 3 describes the solution about cross-domain authentication
-and single sign-on is presented, including a few solutions implemented
-for bridging the mismatch in terms of authentication. Section 4
-discusses the related research, and Section 5 concludes.
+\ \ \ \ The rest of this paper is organized as follows: Section 2 shows
+the ARC grid middleware, especially the architecture of the new version
+of ARC. Section 3 describes the consideration and implementation about
+cross-middleware authentication and single sign-on is presented,
+including a few solutions implemented for bridging the mismatch in
+terms of authentication and single sign-on. Section 4 discusses the
+related research, and Section 5 concludes.
 
@@
 
-ARC(Advanced Resource Connector) is an open source grid middleware
-solution released under GPL licence. ARC middleware initially aimed at
-developing a grid middleware which provides characteristics such as
-self-organized, fault-tolerant, non-intrusive, easy-management [fgcs
-ARC paper]. \ Now \ the classic version of ARC provides grid services
-such as grid job submission and management, resource characterization,
-resource aggregation and discovery, basic data management, integration
-of grid security solution, etc. Classic ARC has been deployed and used
-in production environment, and been one of the mainly deployed grid
-middlewares in Europe.
+\ \ \ \ ARC(Advanced Resource Connector) is an open source grid
+middleware solution released under Apache license. ARC middleware
+initially aimed at developing a grid middleware which provides
+characteristics such as self-organized, fault-tolerant, non-intrusive,
+easy-management [10]. \ Now \ the classic version of ARC provides grid
+services such as grid job submission and management, resource
+characterization, resource aggregation and discovery, basic data
+management, integration of grid security solution, etc. Classic ARC has
+been deployed and used in production environment, and been one of the
+mainly deployed grid middlewares in Europe.
 
-The new generation of ARC is funded by KnowARC project, and based on the
-functionality and capabilities of the classic ARC middleware, it aims
-at implementing a service-oriented middleware which will provide higher
-levels of resource and user abstracting through well-defined web
-service interface [Design document] in order to provide
-interoperability with other service-based grid middlewares, as well as
-other Web Service compatible applications.
+\ \ \ \ The new generation of ARC is based on the functionality and
+capabilities of the classic ARC middleware, and it aims at implementing
+a service-oriented middleware which will provide higher levels of
+resource and user abstracting through well-defined web service
+interface [11] in order to provide interoperability with other
+service-based grid middlewares, as well as other Web Service compatible
+applications.
 
-As the key part of the implementation of new ARC middleware, there is a
-lightweight web service container called HED (Hosting Environment
-Daemon) which provides a host place for various services in application
-level, as well as a host place for a bunch of components to support
-flexible, interoperatible, and efficient communication mechanism. The
-whole design of the HED is built around the idea of flexibility and
-modularity, which means for the developer, he can easily concentrate on
-the application level Web Service implementation by only using the core
-minimum amount of components if he only would do application level
-development, or he can easily concentrate on the middleware level
-implementation such as supporting another communication protocol or
-implementing authentication mechanism by using the core minimum amount
-of components and external dependencies; also for the deployer, he can
-easily configure and deploy the middleware and application for
-different kinds of requirements without being bothered to know much
-about the implementation.
+\ \ \ \ As the key part of the implementation of new ARC middleware,
+there is a lightweight Web Service container called Hosting Environment
+Daemon (HED) which provides a host place for various services in
+application level, as well as a host place for a bunch of modules to
+support flexible, interoperatible, and efficient communication
+mechanism for building SOAP based Web Services. The whole design of the
+HED is built around the idea of flexibility and modularity, which means
+for the developer, he can easily concentrate on the application level
+Web Service implementation by only using the core minimum amount of
+components if he only would do application level development, or he can
+easily concentrate on the middleware level implementation such as
+supporting another communication protocol or implementing
+authentication mechanism by using the core minimum amount of components
+and external dependencies; also for the deployer, he can easily
+configure and deploy the middleware and application for different kinds
+of requirements without being bothered to know much about the
+implementation.
 
@@
 \end{center}
-The architecture of the Host Environment Daemon is illustrated as Figure
-1. In general, there are a few \ components called MCC (Message Chain
-Components) which are in charge of processing different messages for
+\ \ \ \ The architecture of the HED is illustrated as Figure 1. In
+general, there are a few \ components called Message Chain Component
+(MCC) which are in charge of processing different messages for
 different protocol levels. For instance, as shown in the example
@@
 
-Two examples of MCC components configuration in Figure 1 shows MCC are
-configurable: real line shows SOAP over HTTPS (HTTP over TLS), while
-dotted line shows SOAP directly over TCP. Service administrator can
-configure the MCCs according to his requirement, such as the
-interoperability requirement with the other part. For instance, the
-later configuration is compatible to WSE (Web Services Enhancement for
-.NET) {\textquotesingle}s SOAP message mechanism (see
-WSE{\textquotesingle}s \textit{SoapSender} and \textit{SoapReceiver}).
-Another configuration could be SOAP over HTTPG (HTTP over GSI) which
-can be used to interoperate with some service based on HTTPG such as
-SRM (Storage Resource Manager) [ref] service. The easily configurable
-characteristics shows the flexibility of HED in terms of protocol
-supporting.
+\ \ \ \ Two examples of MCC components configuration in Figure 1 shows
+MCC are flexibly configurable: real line shows SOAP over HTTPS (HTTP
+over TLS), while dotted line shows SOAP directly over TCP. Service
+administrator can configure the MCCs according to the interoperability
+requirement with the other part. For instance, the later example
+configuration is compatible to WSE (Web Services Enhancement for .NET)
+{\textquotesingle}s SOAP message mechanism (see WSE{\textquotesingle}s
+\textit{SoapSender} and \textit{SoapReceiver}). Another configuration
+could be SOAP over HTTPG (HTTP over GSI) which can be used to
+interoperate with some service based on HTTPG such as Storage Resource
+Manager (SRM)[12] \ service. The easily configurable characteristics
+shows the flexibility of HED in terms of protocol supporting.
 
-Because in the current state HED mostly provides modules for building
-SOAP based Web Services, it is easy to think that HED is just another
-Web Services development framework like Axis, gSOAP, XFire or any other
-out of the numerous implementations. Instead, the idea of HED is to
-provide framework for gluing functionalities and not a
-re-implementation of various standards. Effectively that means if
-Apache httpd web server is considered by developers as necessary for
-serving as front-end to services there could be plugin code written
-which places Apache2 into a chain of other plugins of the HED, the same
-for the Apache Axis SOAP implementation, as shown in Figure 1.
-
-Even though the external solutions like Axis and Apache httpd can be
-used for protocol supporting, \ two protocols are completely developed
-in HED to support HTTP and SOAP in order to make the solution
-lightweight and needed an implementation of those supported protocols
-that is both simple and lightweight. However external solutions are
-used whenever is appropriate - as in the case of TLS (openssl is used),
-GridFTP (GridFTP API is used), LDAP and some other cases, this
-principle applies to Axis and Apache httpd if we used the API from
-those solution to implement the plugins for supporting SOAP and HTTP
-protocol.
-
-HED contains a framework for implementing and enforcing authentication
-and authorization. Each Message Chain Component (MCC) or service has a
-common interface for implementing various authentication and
-authorization functionality. This functionality is implemented by using
-pluggable and configurable components (plug-ins) called
-\textit{SecHandler}. The \textit{SecHandler} components are C++ classes
-and provide method for processing messages traveling through Message
-Chains of the HED. Each MCC or Service \ usually implement two queues
-of SecHandlers -- one for incoming messages and one for outgoing called
-{\textquotedblleft}incoming{\textquotedblright} and
-{\textquotedblleft}outgoing{\textquotedblright} respectively. All
-\textit{SecHandler} components attached to the queue are executed
-sequentially. If any of them fails, message processing fails as well.
-In figure 1, the {\textquotedblleft}AuthZ{\textquotedblright} and
+\ \ \ \ HED contains a framework for implementing and enforcing
+authentication and authorization. Each MCC or Service has a common
+interface for implementing various authentication and authorization
+functionality. Each functionality can be implemented as a pluggable and
+configurable component (plug-ins) called \textit{SecHandler} which is
+C++ class and provides method for processing message that travels
+through MCC or Service. Each MCC or Service is usually configured with
+two queues of \textit{SecHandler} -- one for incoming message and one
+for outgoing message respectively. All \textit{SecHandler} components
+attached to the queue are executed sequentially. If any of them fails,
+message processing fails as well. In Figure 1, the
+{\textquotedblleft}AuthZ{\textquotedblright} and
 {\textquotedblleft}AuthN{\textquotedblright} sub-module inside MCCs and
@@
 
-As explained above, HED is supposed to act as a service holder, but on
-the client side, from implementation perspective, there is similar
-architecture implemented for processing messages from different
-protocols and handling security functionality, also there is a specific
-application programming interface (API) for developers to easily write
-Web Service client code to service.
+As explained above, HED is supposed to act as a Web/Grid Service
+container, but on the client side, from implementation perspective,
+there is similar architecture implemented for processing messages from
+different protocols and handling security functionality, also there is
+a specific application programming interface (API) for developers to
+easily write Web Service client code.
 
@@
 {\centering
-3. Cross-domain authentication and Single Sign-On
+3. Cross-middleware authentication and Single Sign-On
 \par}
 
-For cross domain authentication, our work comprises the following five
-major aspects:
+\ \ \ \ We exploited the authentication and single sign-on issue in the
+scenario that user or user{\textquotesingle}s delegatee (act on behalf
+of this user) needs to access across different middlewares. Our work
+comprises the following five major aspects:
 
@@
 \item \textit{Integration of community authentication} -{}-{}- to
-utilize the standardized community authentication from Shibboleth
-Identity Provider software when users authenticate against Grid
-services.
+utilize the standard-compatible community authentication mechanism from
+Shibboleth [13] Identity Provider software when users authenticate
+against Grid services.
 \item \textit{Short lived credential service} \ {}-{}-{}- to issue the
@@
 Security SAML token in order to provide single sign-on for SOAP message
-level authentication
+level authentication.
 \end{itemize}
-
-\bigskip
-
-\ \ 3.1 Community based authentication for Grid by utilizing Shibboleth
+3.1 Community based authentication for Grid by utilizing Shibboleth
 Identity Provider
 
-AAI (Authentication and authorization Infrastructure) is a solution for
-the authentication and authorization for inter-organization resource
-sharing, such as electronic resource sharing between libraries, etc.
-AAI implicitly applies to community/institutional/organization based
-authentication where users are from different home communities but need
-to get resources from other communities by using some federation
-mechanism. Unlike the X.509 based authentication solution in current
-grid systems, AAI does not require user to provide X.509 certificate,
-instead, it can support different types of authentication, such as
-username/password based authentication, IP address based
+\ \ \ \ AAI (Authentication and authorization Infrastructure) is a
+solution for the authentication and authorization for
+inter-organization resource sharing, such as electronic resource
+sharing between libraries, etc. AAI implicitly applies to community or
+institutional based authentication where users are from different home
+communities but need to get resources from other communities by using
+some federation mechanism. Unlike the X.509 based authentication
+solution in current grid systems, AAI does not require user to provide
+X.509 certificate, instead, it can support different types of
+authentication, such as username/password authentication, IP address
 authentication, etc. There are a few implementations about AAI, among
-which Shibboleth [ref] is one implementation which has been widely
-deployed.
+which Shibboleth is one implementation which has been widely deployed.
 
-Shibboleth provides cross-domain single sign-on and attribute-based
-authorization while preserving user privacy. Shibboleth is based on the
-OASIS Security Assertion Markup Language (SAML), especially the \ new
-version of Shibboleth supports SAML 2.0 specification. For
-authentication, the main SAML profile that Shibboleth implements is the
-SAML2.0 web browser SSO profile, which \ define two functional
-components, an Identity Provider and a Service Provider. The Identity
-Provider (IdP) is responsible for creating, maintaining, and managing
-user identity, while the Service Provider (SP) is responsible for
-controlling access to services and resources by using the SAML
-assertion produced and issued by IdP upon request. In order to discover
-which home community does a user from, Shibboleth specifies an optional
-third component called {\textquotedblleft}Where Are You
-From?{\textquotedblright} (WAYF) service to aid in the process of IdP
-discovery, and this IdP discovery process is also standardized and
-\ defined in SAML 2.0 specification and called as
+\ \ \ \ Shibboleth provides cross-domain single sign-on and
+attribute-based authorization while preserving user privacy. It is
+based on the OASIS Security Assertion Markup Language (SAML),
+especially the \ new version of Shibboleth supports SAML 2.0
+specification. For authentication, the main SAML profile that
+Shibboleth implements is the SAML2.0 web browser SSO profile, which
+\ define two functional components, an Identity Provider and a Service
+Provider. The Identity Provider (IdP) is responsible for creating,
+maintaining, and managing user identity, while the Service Provider
+(SP) is responsible for controlling access to services and resources by
+using the SAML assertion produced and issued by IdP upon request. In
+order to discover which home community does a user from, Shibboleth
+specifies an optional third component called {\textquotedblleft}Where
+Are You From?{\textquotedblright} (WAYF) service to aid in the process
+of IdP discovery, and this IdP discovery process is also standardized
+and \ defined in SAML 2.0 specification and called as
 {\textquotedblleft}Identity Profile Discovery
@@
 
-We utilized the SAML2.0 web browser SSO profile for the authentication
-in ARC middleware. But since the SSO profile is initially supposed to
-protect web applications and provide authentication for web users, we
-implemented some external code on the client and service side to
-integrate SSO profile. On the client side, apart from the WS-Client API
-for writing Web Service client, we implemented the user agent
-functionality of web browser in order to mimic the behavior of web
-browser{\textquotesingle}s user agent, such as http redirection, cookie
-processing, etc. In fact, the implementation of user agent is also
-based on the client interface of ARC, specifically, the https client
-interface, since the client interface of ARC can support different
-protocols which are incarnated by different message chain component
-(MCC). Hence client developers who would use SAML2.0 SSO profile should
-call the user agent interface and then the WS-Client interface. On the
-service side, we implemented the Service Provider functionality (based
-on the HTTP MCC configured together with TLS MCC) which is called SP
-Service. For Identity Provider, the Shibboleth IdP implementation is
-used.
+\ \ \ \ We utilized the SAML2.0 web browser SSO profile for the
+authentication in ARC middleware. But since the SSO profile is
+initially supposed to protect web applications and provide
+authentication for web users, we implemented some external code on the
+client and service side to integrate SSO profile. On the client side,
+apart from the client interface for writing Web Service client, we
+implemented the user agent functionality of web browser in order to
+mimic its behavior, such as http redirection, cookie processing (In
+fact, the implementation of user agent is also based on the client
+interface of ARC, specifically, the https client interface, since the
+client interface of ARC can support different protocols which are
+incarnated by different MCC). Hence client developers who would use
+SAML2.0 SSO profile should call the user agent interface and then the
+Web Service client interface. On the service side, we implemented the
+Service Provider functionality (based on the HTTP MCC configured
+together with TLS MCC) which is called SP Service. For Identity
+Provider, the Shibboleth IdP implementation is used.
 
-Figure 2 shows the process of SAML2.0 SSO integrated in ARC client and
-service.
+\ \ \ \ Figure 2 shows the process of SAML2.0 SSO integrated in ARC
+client and service.
 
@@
 \end{center}
-
-\bigskip
 
@@
 \begin{enumerate}
-\item The client use the user agent interface to launches a http request
-including the IdP name (to which the user belongs) to the service side.
-The host-name and port of destination URL is the same as that of the
-real service, the last part of the destination URL is
+\item The client uses the user agent interface to launches a http
+request including the IdP name (to which the user belongs) to the
+service side. The endpoint of the SP service is the same as that of the
+other target services, except the last part of the endpoint is
 {\textquotedblleft}saml2sp{\textquotedblright} which is specific for
-pointing to the Service Provider. Note that we use Identity Provider
-(IdP) name here to simplify the IdP discovery process, because we
-suppose that user who would access the service should better know where
-is he from initially.
+pointing to the SP Service. Note that we use Identity Provider (IdP)
+name here to simplify the IdP discovery process in order to avoid the
+IdP discovery process, because we suppose that user who would access
+the target services should better know where is he from initially.
 \item The SP Service (Service Provider) searches the metadata (we use
-the same metadata with the same format defined in Shibboleth) and gets
-the location of the single sign-on service (hosted in IdP) and also the
-location of assertion consuming service (hosted in this SP itself) in
-order to compose the SAML {\textless}AuthnRequest{\textgreater}
+the same metadata format defined in Shibboleth) and gets the location
+of the single sign-on service (hosted in IdP) and also the location of
+assertion consuming service (hosted in this SP itself) in order to
+compose the SAML {\textless}\textit{samlp:AuthnRequest}{\textgreater}
 message. Then SP Service issues this
-{\textless}AuthnRequest{\textgreater} message by using its own X.509
-certificate (Note in the SAML SSO profile, certificate is still needed
-by IdP and SP) and sends back to user agent.
-\item User agent sends the {\textless}AuthnRequest{\textgreater} message
-to Identity Provider.
+{\textless}\textit{samlp:AuthnRequest}{\textgreater} message by using
+its own X.509 certificate (Note in the SAML SSO profile, certificate is
+still needed by IdP and SP) and sends back to user agent.
+\item User agent sends the
+{\textless}\textit{samlp:AuthnRequest}{\textgreater} message to
+Identity Provider.
 \item Identity Provider requires an act of authentication. The
-authentication mechanism is outside of the SAML2.0 SSO profile. The
-shibboleth IdP implementation will choose some login handler for
+authentication mechanism is outside of the SAML2.0 SSO profile.
+Shibboleth IdP implementation chooses some login handler for
 authentication. The current user agent implementation is compatible to
-the Username/Password login handler of Shibboleth IdP. Through the http
-protocol, user agent will feed IdP with the username/password which has
-been given by the caller of user agent interface.
+\textit{Username/Password} login handler of Shibboleth IdP. Through the
+http protocol, user agent will feed IdP with the username/password
+which has been given by the caller of user agent interface.
 \item Once the authentication has been succeeded, the IdP issues a SAML
@@
 and stores the SAML assertion into session/connection context. The SAML
-assertion actually includes the {\textless}AuthnStatement{\textgreater}
-and {\textless}AttributeStatement{\textgreater}.
-\item The WS-Client launches the real Grid/Web Service request via the
-same connection as the one which is used by user agent to contacts SP
+assertion includes the
+{\textless}\textit{saml:AuthnStatement}{\textgreater} and
+{\textless}\textit{saml:AttributeStatement}{\textgreater}.
+\item The WS client launches the Grid/Web Service request via the same
+connection as the one which is used by user agent to contact SP
 Service.
 \item The Grid/Web Service checks the
-{\textless}AuthnStatement{\textgreater} from the session context to see
-if the session is still valid through the \textit{SecHandler} called
-{\textquotedblleft}SAML SecHandler{\textquotedblright}. And if valid,
-service handles the real service processing and returns the response to
-WS client. Note that service requires WS client is from the same
-connection as the one on which user agent contact SP service in order
-to guarantee that the validity of SSO profile result effects the real
-client/service interaction.
+{\textless}\textit{saml:AuthnStatement}{\textgreater} from the session
+context to see if the session is still valid through the
+\textit{SecHandler} called {\textquotedblleft}\textit{SAML
+SecHandler}{\textquotedblright}. If valid, service handles the service
+processing and returns the response to WS client. Note that service
+requires that WS client is from the same connection as the one on which
+user agent contact SP service, in order to guarantee that the validity
+of SSO profile result effects the WS client/Web Service interaction.
 \end{enumerate}
-The SP service and real functional service(s) are hosted by the same
-container, and they use the same X.509 credential. The client side
+\ \ \ \ The SP service and functional service(s) are hosted by the same
+container, and they use the same X.509 credential. The client
 authentication is switched off, so that client doesn{\textquotesingle}t
 need to use any X.509 credential. Only the trusted certificates (CA
-certificate for SP and IdP) need to be configured for client side so
-that SP and IdP can authenticate themselves to the client. As required
-by SAML2.0 profile, the SP and IdP should have trust relationship to
-each other.
+certificates for both SP and IdP) need to be configured for client side
+so that SP and IdP can authenticate themselves to the client. As
+required by SAML2.0 profile, the SP and IdP should have trust
+relationship to each other.
 
-One benefit of SAML2.0 SSO profile that is worth mentioning is that: the
-Identity Provider could cache the authentication result through session
-management once the user agent has succeeded to authenticate, then in a
-short period this authentication result is valid so that user agent
-don{\textquotesingle}t need to feed IdP with user{\textquotesingle}s
-username and password the next time (if the time is not out of the
-scope of valid period) it authenticates against IdP. So user (or the
-client on behalf of this user) can travel across multiple security
-domains with only inputing his name and password once, which is the
-characteristic of single sign-on.
+\ \ \ \ One benefit of SAML2.0 SSO profile that is worth mentioning is:
+the Identity Provider could cache the authentication result through
+session management once the user agent has succeeded to authenticate,
+then in a short period this authentication result is valid so that user
+agent doesn{\textquotesingle}t need to feed IdP with
+user{\textquotesingle}s username and password the next time (if this
+point of time is not out of the scope of valid period) it authenticates
+against IdP. So user (or the client on behalf of this user) can travel
+across multiple security domains with only inputing his name and
+password once, which is the characteristic of single sign-on.
 
-Since the Shibboleth implementation of SAML is standard-compliant and
-widely deployed, the solution implemented in ARC can easily
-interoperate with other SAML implementations with minimum changes, and
-more importantly, this solution can succeed to utilize the widely
-deployed SAML implementation for authentication in grid systems by
-avoiding the usage of X.509 certificate.
+\ \ \ \ Since the Shibboleth implementation of SAML is
+standard-compliant and widely deployed, the solution implemented in ARC
+can easily interoperate with other SAML implementations with minimum
+change, and more importantly, this solution can succeed to utilize the
+widely deployed SAML implementation for authentication in grid systems
+by avoiding the usage of X.509 certificate.
 
 Moreover, even though the implementation is based on ARC middlerware,
-the idea can be adopted by other grid middlewares if they can only
-require server SSL authentication instead of mutual SSL authentication.
+the idea can be adopted by other grid middlewares if they only require
+server authentication instead of mutual authentication.
 
-\bigskip
+3.2 Short-lived credential service
 
-\ \ 3.2 Short lived credential service
+\ \ \ \ However, most of the widely used grid middlewares are based on
+GSI while GSI requires mutual authentication. Also for Web Service
+based grid solution, we can not prevent service side from requiring
+client X.509 certificate. \ Based on the solution described in Section
+3.1, we implemented a short lived credential service (SLCS) by which
+user can get a short-lived X.509 certificate without being bothered to
+contact any registration authority (RA) or certificate authority (CA).
 
-However, most of the widely used grid middlewares are based on GSI while
-GSI requires mutual authentication. Also for Web Service based grid
-solution, we can not prevent service side from requiring client X.509
-certificate. \ Based on the solution described in Section 3.1, we
-implemented a short lived credential service (SLCS) by which user can
-get a short-lived X.509 certificate without being bothered to contact
-any registration authority (RA) or certificate authority (CA).
-
-The SLCS service is actually a specific Web Service (standard web
-service implemented by using ARC service interface) in Figure 2, and
-the SLCS client is a specific ARC client which includes the user agent
-and WS client. The whole process from step 1 to step 8 is the same as
-in Figure 2.
+The SLCS service is also a Web Service (standard web service implemented
+by using ARC service interface), and the SLCS client is a specific
+command-line interface(CLI) which includes the user agent and WS
+client. The whole process of SLCS invocation is showed in Figure3 (from
+step 1 to step 8), which is the same as in Figure 2, except that step7
+and step8 is incarnated for SLCS certificate request and response.
 
@@
 
-\ \ \ SLCS client generates a X.509 certificate request, launches a Web
-Service request which includes the certificate request; SLCS service
-then gets the certificate request, composes a distinguished name (DN),
-issues a certificate (short lived, 12 hours by default) with the SAML
-attribute (from the SAML2.0 SSO profile) as the X.509 certificate
-extension, and puts the certificate in to the Web Service response;
-SLCS client get the response and stores the X.509 certificate into
-local repository.
+\ \ \ \ SLCS client generates a X.509 certificate request, launches a
+Web Service request which includes the certificate request; SLCS
+service then gets the certificate request, composes a distinguished
+name (DN), issues a certificate (short lived, 12 hours by default) with
+the SAML attribute (from the SAML2.0 SSO profile) as the X.509
+certificate extension, and puts the certificate in to the Web Service
+response; SLCS client get the response and stores the X.509 certificate
+into local repository.
 
-The command line interface for the SLCS client is like this:
+The CLI for the SLCS client is like this:
 
@@
 
-\ \ \ Since the lifetime of the short lived credential is normally
+\ \ \ \ Since the lifetime of the short lived credential is normally
 short, it is not a must to protect the private key by pass-phrase. As
@@
 middleware. If the private key is protected, the can use the X.509
-certificate to generate a proxy certificate (by using command line
+certificate to generate a proxy certificate (by using command-line
 interface utility such as \textit{grid-proxy-init},
-\textit{voms-proxy-init}, or \textit{arcproxy}), and the use the proxy
+\textit{voms-proxy-init}, or \textit{arcproxy}), and then use the proxy
 certificate to access Grid Service or Web Service.
 
-\ \ \ It is worth mentioning that since ARC middleware can support GSI
+\ \ \ \ It is worth mentioning that since ARC middleware can support GSI
 communication by configuring the GSI MCC, \ together with the X.509
-certificate, the Web Service client developed by ARC client interface
-can interoperate with Grid Service that requires GSI communication.
+certificate, the Web Service client developed by ARC Web Service client
+interface can interoperate with Grid Service that requires GSI
+communication.
 
-\ \ \ It might be noticed that how to compose the distinguished name
-(DN) for the certificate is critical issue for SLCS service. Since the
-Shibboleth Identity Provider uses the eduPerson schema [ref] for the
-definition of {\textless}saml:Attribute{\textgreater} in
-{\textless}saml:AttributeStatement{\textgreater}, we pick the
+\ \ \ \ It might be noticed that how to compose the distinguished name
+(DN) for the certificate is a critical issue for SLCS service. Since
+the Shibboleth Identity Provider uses the eduPerson schema [14] for the
+definition of {\textless}\textit{saml:Attribute}{\textgreater} in
+{\textless}\textit{saml:AttributeStatement}{\textgreater}, we pick the
 relatively distinguishable attribute
-{\textquotedblleft}eduPersonPrincipalName{\textquotedblright} for the
-DN. A typical \ eduPersonPrincipalName \ value could be
-\href{mailto:alice@example.org}{alice@example.org}, then the DN is
-{\textquotedblleft}/O=knowarc/OU=example.org/CN=alice{\textquotedblright}.
+{\textquotedblleft}\textit{eduPersonPrincipalName}{\textquotedblright}
+for the DN. A typical \ \textit{eduPersonPrincipalName} \ value could
+be \href{mailto:alice@example.org}{alice@example.org}, then the DN is
+{\textquotedblleft}\textit{/O=knowarc/OU=example.org/CN=alice}{\textquotedblright}.
 
-The obvious benefit of SLCS service is that: If a user passes the
-authentication to his home Identity Provider, he can get the X.509
+\ \ \ \ The obvious benefit of SLCS service is that: If a user passes
+the authentication to his home Identity Provider, he can get the X.509
 credential at anywhere only by running the SLCS client command together
@@
 access the grid system conveniently; meanwhile, thanks to the single
-sign-on characteristic discussed in Section 3.1, this user
-\ doesn{\textquotesingle}t not need to input his username and password
-in a valid period after the first time he succeeds to authenticate
-against his home IdP by running the SLCS client command on the same
-node, even if this SLCS client command is supposed to run against a few
-SLCS services to get a few SLCS credentials.
+sign-on characteristic of SAML2.0 SSO profile which is discussed in
+Section 3.1, this user doesn{\textquotesingle}t not need to input his
+username and password in a valid period after the first time he
+succeeds to authenticate against his home IdP by running the SLCS
+client command on the same node, even if this SLCS client command is
+supposed to run against a few SLCS services to get a few SLCS
+credentials.
 
@@
 
-\ \ 3.3 X.509 Credential delegation service
+3.3 X.509 Credential delegation service
 
-The delegation of credentials proposed in Grid Security Infrastructure
-(GSI) is a good solution for the supporting of single sign-on for
-users, except the coupling of the GSI communication which is not so
-widely accepted in Web Service applications. In the design of the new
-version ARC middleware, we keep the credential delegation idea, while
-utilizing the SOAP communication (and normal TCP or TLS for transport
+\ \ \ \ The delegation of credentials proposed in Grid Security
+Infrastructure (GSI) is a good solution for the supporting of single
+sign-on for users, except the coupling of the GSI communication which
+is not so widely accepted in Web Service applications. In the design of
+the new version ARC middleware, we keep the credential delegation idea,
+while utilizing the SOAP communication (and TCP or TLS for transport
 level communication) instead of the GSI communication for the
@@
 
-As shown in Figure 4, there is a specific WS client and Web Service as
-well for processing delegation: delegation client, and delegation
-service. The delegation service has two main wsdl operations, one for
-processing the delegation initiation; the other for processing/storing
-the signed proxy certificate. The delegation process is detailed in
-Figure 5.
+\ \ \ \ As shown in Figure 4, there is a specific WS client and Web
+Service for processing delegation: delegation client, and delegation
+service. The delegation service has two main \textit{WSDL}(Web Service
+Description Language) operations, one for processing the delegation
+initiation; the other for storing the signed proxy certificate. The
+delegation process is detailed in Figure 5.
 
-For the service invocation process, the WS client functionality (inside
-each service or the initial client) which is included to access the
-target service (step a1 or a2 in Figure 4), should use the
+\ \ \ \ For the service invocation process, the WS client functionality
+(inside each service or the initial client) which is included to access
+the target service (step a1 or a2 in Figure 4), should use the
 user{\textquotesingle}s X.509 certificate or proxy certificate (on
@@
 
-Since each WS client accesses the target service (delegation service or
-other functional service) on behalf of the user, the trust relationship
-between the user{\textquotesingle}s certificates (X.509 certificate and
-proxy certificate) and services is required, while the trust
-relationship between service and service is not required.
+\ \ \ \ Since each WS client accesses the target service (delegation
+service or other functional service) on behalf of the user, the trust
+relationship between the user{\textquotesingle}s certificates (X.509
+certificate and proxy certificate) and services is required, while the
+trust relationship between service and service is not required.
 
@@
 \begin{center}
-\includegraphics[width=2.5319in,height=1.8571in]{Secpaper-img5.png}
+\includegraphics[width=2.3028in,height=1.6465in]{Secpaper-img5.png}
 \end{center}
@@
 
-\bigskip
+3.4 Message level authentication
 
-\ \ 3.4 Message level authentication
+\ \ \ \ Transport level security (TLS or SSL) is not sufficient for
+securing Web Services, because the secure channel created by SSL/TLS
+make it impossible to differentially protect the SOAP message, e.g.
+\ to encrypt and/or sign only particular components of the SOAP
+message, which is relevant when non-sensitive portions of the message
+need to be accessed or changed by intermediate actors. Additionally,
+SSL/TLS is incapable of providing end-to-end protection for SOAP
+message which might flow through multiple intermediate actors, because
+it only allows each hop to be protected with the resultant security
+gaps at intermediate actors. Web Services Security (WS-Security)
+specification (ref) provides means for applying security to Web
+Service.
 
-Transport level security (TLS or SSL) is not sufficient for securing Web
-Services, because the secure channel created by SSL/TLS make it
-impossible to differentially protect the SOAP message, e.g. \ to
-encrypt and/or sign only particular components of the SOAP message,
-which is relevant when non-sensitive portions of the message need to be
-accessed or changed by intermediate actors. Additionally, SSL/TLS is
-incapable of providing end-to-end protection for SOAP message which
-might flow through multiple intermediate actors, because it only allows
-each hop to be protected with the resultant security gaps at
-intermediate actors. Web Services Security (WS-Security) specification
-(ref) provides means for applying security to Web Service.
-
-The WS-Security related specifications are implemented in ARC as
+\ \ \ \ The WS-Security related specifications are implemented in ARC as
 \textit{SecHandler}. There are three kinds of WS-Security version 1.1
@@
 
-For those WS clients on behalf of a user, at any step of service
+\ \ \ \ For those WS clients on behalf of a user, at any step of service
 invocation, the WS client can augment the SOAP message with WS-Security
@@
 
-If the WS client owns an X.509 certificate or proxy certificate, the
-process of generating an X.509 Token for a SOAP message is: the client
-puts the content of certificate into X.509 token as
+\ \ \ \ If the WS client owns an X.509 certificate or proxy certificate,
+the process of generating an X.509 Token for a SOAP message is: the
+client puts the content of certificate into X.509 token as
 {\textless}wsse:BinarySecurityToken{\textgreater}, and uses the private
@@
 
-However the process of generating a SAML Token is more complicated. In
-SAML Token profile, \textit{ConfirmationMethod} is the verifying method
-for establishing proof-of-possession for the subject. Two methods are
-proposed in SAML Token profile: \textit{holder-of-key},
+\ \ \ \ However the process of generating a SAML Token is more
+complicated. In SAML Token profile, \textit{ConfirmationMethod} is the
+verifying method for establishing proof-of-possession for the subject.
+Two methods are proposed in SAML Token profile: \textit{holder-of-key},
 \textit{sender-vouches}. \ We rely on the \textit{holder-of-key
@@
 
-\ The relying party is supposed to trust the issuing authority. When it
-receives a message from the asserting entity, it will check the SAML
-assertion based on its predetermined trust relationship with the SAML
-issuing authority, and check the signature of the soap message based on
-the public key in the SAML assertion without directly trust
+\ \ \ \ The relying party is supposed to trust the issuing authority.
+When it receives a message from the asserting entity, it will check the
+SAML assertion based on its predetermined trust relationship with the
+SAML issuing authority, and check the signature of the soap message
+based on the public key in the SAML assertion without directly trust
 relationship with attesting entity (subject owner).
+
+\ \ \ In terms of verification, since the of proxy certificates has been
+introduced in IETF (RFC 3820), \ and OpenSSL has also implemented
+RFC3820, it could not be difficult to use the proxy certificate for
+X.509 Token profile.
 
@@
 
-\ \ 3.5 SAML Token delegation service
+3.5 SAML Token delegation service
 
-\ There is also delegation requirement on the message level, some
+\ \ \ \ There is also delegation requirement on the message level, some
 entities (client or user) often need to temporarily delegate itself to
@@
 
-The delegation process is similar to the process for delegation proxy
-certificate, with the following differences:
+\ \ \ \ The delegation process is similar to the process for delegation
+proxy certificate, with the following differences:
 
@@
 \end{enumerate}
-Figure 6 shows the SOAP message header is changed due to delegation. On
-the left side, B is the identity of the attesting
+\ \ \ \ Figure 6 shows the SOAP message header is changed due to
+delegation. On the left side, B is the identity of the attesting
 entity{\textquotesingle}s X509 token mentioned above, while A is the
@@
 
-\bigskip
+3.6 Discussion
 
@@
 \par}
+
+\bigskip
+
+\ \ There is some effort to add GSI support to adapt to the strict
+security requirement including GSI from grid middlewares, such as GSI
+plug-in for gSOAP[15]. The GSI MCC implemented in ARC can also achieve
+the same functionality as gSOAP. Moreover, the ARC solution is flexible
+than the gSOAP solution, because both the GSI MCC of service and client
+sides are pluggable, which means service and client developer of ARC do
+not need to modify the service or client implementation in order to add
+or remove the GSI functionality.
+
+\bigskip
 
@@
 
-Foster I, Kesselman C, Tuecke S. The anatomy of the Grid: Enabling
+[1] Foster I, Kesselman C, Tuecke S. The anatomy of the Grid: Enabling
 scalable virtual organizations. International Journal of Supercomputer
@@
 
-\bigskip
+\textstyleStrongEmphasis{\textmd{[2]}}\textstyleStrongEmphasis{\textmd{Foster,
+I., Kesselman, C., Tsudik, G. and }}Tuecke, S. A Security Architecture
+for Computational Grids. ACM Conference on Computers and Security,
+1998, 83-91.
 
-The Shibboleth Project.
+[3]R. Alfieri, R. Cecchini, V. Ciaschini, L.
+dell{\textquoteright}Agnello, A. Frohner, K. Lorentey, and F. Spataro.
+From gridmap- file to voms: managing authorization in a grid
+environment. Future Generation Comp. Syst., 21(4):549--558, 2005.
+
+[4] RFC 3821- An Internet Attribute Certificate Profile for
+Authorization. \url{http://www.faqs.org/rfcs/rfc3281.html}
+
+[5] Globus Toolkit. http://www.globus.org/toolkit/
+
+[6]gLite: lightweight middleware for grid computing.
+\href{http://glite.web.cern.ch/}{Http://}\href{http://glite.web.cern.ch/}{\textstyleCitation{\textup{glite.web.cern.ch}}}
+
+\textstyleCitation{\textup{[7]Advanced Resource Connector.
+}}\url{http://www.nordugrid.org/middleware/}
+
+\textstyleCitation{\textup{[8]Web Services Security(WS-Security).
+www.oasis-open.org/committees/wss/}}
+
+\textstyleCitation{\textup{[9]OASIS Security Assertion Markup Languages
+(SAML). www.oasis-open.org/committees/security/}}
+
+\textstyleCitation{\textup{[10]}}\textstyleCitation{\textup{M. Ellert,
+M. Gr{\o}nager, A. Konstantinov, B. Konya, }}J. Lindemann, I. Livenson,
+J. Langgaard Nielsen, M. Niinimaki, O. Smirnova, and A. Waananen.
+Advanced resource connector middleware for lightweight computational
+grids. Future Generation computer systems, 23(2):219--240, 2007.
+
+\textstyleCitation{\textup{[11] Design document of new version ARC.
+}}\url{https://www.knowarc.eu/documents/Knowarc_D1.1-1_07.pdf}
+
+\textstyleCitation{\textup{[12]A. Shoshani, A. Sim, and J. Gu,
+}}\textstyleCitation{Storage Resource Managers: Essential Components
+for the Grid}\textstyleCitation{\textup{, Grid Resource Management:
+State of the Art and Future Trends, Kluwer Publishing, 2003.}}
+
+[13]The Shibboleth Project.
 \href{http://shibboleth.internet2.edu/}{Http://shibboleth.internet2.edu/}
+
+[14]eduPerson and eduOrg Object shema.
+\url{http://middleware.internet2.edu/eduperson/}
+
+\textstyleCitation{\textup{[15]}}\textstyleCitation{\textup{Giovanni
+Aloisio, Massimo Cafaro, Italo Epicoco, Daniele Lezzi, and Robert van
+Engelen, }}\textstyleCitation{The GSI plug-in for gSOAP: Enhanced
+Security, Performance, and Reliability}\textstyleCitation{\textup{, in
+the ITCC conference 2005, IEEE Press, Volume I, pages 304-309.}}
 
@@
 
-eduPerson and eduOrg Object shema.
-\url{http://middleware.internet2.edu/eduperson/}
+\bigskip
 

### 2008-12-06 doc/sec/Sec_paper.tex
changes including Aleks's comments, and other improvement

@@
 \newcommand\liststyleLiv{%
+\renewcommand\labelitemi{${\bullet}$}
+\renewcommand\labelitemii{${\circ}$}
+\renewcommand\labelitemiii{${\blacksquare}$}
+\renewcommand\labelitemiv{${\bullet}$}
+}
+\newcommand\liststyleLv{%
 \renewcommand\theenumi{\arabic{enumi}}
@@
 }
-\newcommand\liststyleLv{%
+\newcommand\liststyleLvi{%
 \renewcommand\theenumi{\arabic{enumi}}
@@
 
-\ \ \ \ In order to make the access to Grids as simple as possible,
-security is one of the most important challenges in production Grids,
-especially when the Grid applications are across multiple
-administration domains as well as across heterogeneous Grid
-middlewares, such as some wide scale eScience applications which needs
+\ \ \ \ In task of making the access to Grids as simple as possible,
+security is one of the most important challenges in production Grid
+infrastructures, especially when the Grid applications span across
+multiple administration domains as well as across heterogeneous Grid
+middlewares, such as some wide scale eScience applications which need
 to coordinate resources sharing among collection of autonomous
@@
 \end{enumerate}
-\ \ \ \ Grids provides the technology for wide-scale, cross domains
+\ \ \ \ Grid provides the technology for wide-scale, cross domains
 collaboration where the users work on resources and sharing data which
@@
 because different middlerwares are used as platform. The same
-heterogeneous situation even critically applies to users who need to
-access the VO, because users holding one kind of credential could need
-to travel through multiple resources with different security
-approaches. To solve the above issues, the Grid middlewares should
-provide the following ability: The user which participates the VO
-should be able to access data and resources without being required to
-constantly do authentication at each resource or data; The VO should be
-able to easily manage the privileges or memberships to the users, and
-VO together with its resources should be able to easily manage the
-authorization, given the condition that the resources are owned by
+heterogeneous situation applies to users who need to access the VO,
+because users holding one kind of credential could need to travel
+through multiple resources with different security approaches. To solve
+the above issues, the Grid middlewares should provide the following
+ability:
+
+\liststyleLii
+\begin{itemize}
+\item The user which participates the VO should be able to access data
+and resources without being required to constantly do authentication at
+each resource or data; 
+\item The VO should be able to easily manage the privileges or
+memberships of the users; 
+\item The VO together with its resources should be able to easily manage
+the authorization, given the condition that the resources are owned by
 different administrative domains.
-
+\end{itemize}
 \ \ \ \ In grid community, Grid Security Infrastructure (GSI)[2] is the
-\textit{de facto} standard about authentication and \ transport level
+\textit{de facto} for authentication and \ transport level
 communication, which builds on Public Key Infrastructure (PKI), an
@@
 protocol. Mutual authentication is required by GSI, and is the default
-configuration for GSI based grid deployment; and X.509 certificates is
-required for both of the client and service sides in order to achieve
-mutual authentication. X.509 certificates is issued by trusted third
-party called certificate authorities (CA), and then CAs constitute
-trust federation and guarantee two different X.509 certificates from
-different CAs can accomplish authentication to each other. So in the
-GSI based grid system, if a user would access grid system, he/she
-should own a X.509 certificate which is issued by a CA that is trusted
-by other entity.
+configuration for GSI based grid deployment. The X.509 certificates are
+required for both the client and service sides in order to achieve
+mutual authentication. The X.509 certificates are issued by trusted
+third parties called certificate authorities (CA), and then CAs
+constitute trust federation and guarantee two different X.509
+certificates from different CAs can accomplish authentication to each
+other. So in the GSI based Grid system, if a user would access Grid
+system, he/she should own a X.509 certificate which is issued by a CA
+that is trusted by other entity.
 
 \ \ \ \ In terms of the virtual organization issues mentioned earlier,
-GSI and some other related solutions like virtual organization
-management service (VOMS) [3] have partly solved the them in some
-sense. The identity delegation and X.509 proxy certificate initiated by
-GSI can achieve single sign-on, so that the need to re-enter the
+GSI and some other related solutions like Virtual Organization
+Management Service (VOMS) [3] have partly solved them in some sense.
+The identity delegation and X.509 proxy certificate initiated by GSI
+can achieve single sign-on, so that the need to re-enter the
 user{\textquotesingle}s pass phrase for authentication can be avoided
-by creating the proxy certificate. \ As the most successful grid
-authorization model, VOMS adopts Attribute Certificate[4]. Attribute
-certificate is signed by VOMS server and then binded to proxy
+by creating the proxy certificate. \ As the most successful Grid
+authorization model, VOMS adopts Attribute Certificate[4] approach.
+Attribute Certificate is signed by VOMS server and then binded to proxy
 certificates by client, \ and attributes certificate includes the VO
@@
 \ \ \ Unfortunately, the existing grid security solutions such as GSI
-and VOMS do not apply to Web Service \ based application very well due
+and VOMS do not apply to Web Service \ based applications very well due
 to the following reasons:
 
-\liststyleLii
+\liststyleLiii
 \begin{itemize}
-\item Firstly, \ the GSS API based confidential communication has also
-not been adopted by Web Service implementation.
-\item Secondly, \ Attribute Certificate has not been adopted by any Web
-Service implementation at all.
-\item Thirdly, Web Service implementations commonly use the Web Services
+\item The GSS API based confidential communication has also not been
+adopted by Web Service implementation.
+\item The Attribute Certificate has not been adopted by any Web Service
+implementation at all.
+\item Web Service implementations commonly use the Web Services
 Security(WS-Security)[8], Security Assertion Markup Languages
@@
 \end{itemize}
-\ \ \ \ Moreover, the current grid security solutions require each user
-to possess a X.509 certificate. But the process to get a certificate
-may take quite long time, since the requester needs to be checked and
+\ \ \ \ Moreover, the grid security solutions require each user to
+possess a X.509 certificate. But the process to get a certificate may
+take quite long time, since the requester needs to be checked and
 approved by the Registration Authority (RA) and then the Certificate
@@
 specific command lines, such as \textit{grid-proxy-init}, and
-\textit{voms-proxy-init}. On the other hand, users could most normally
-have some local community/institutional credentials such as
-username/passwords to which users are more familiar. To enable users to
+\textit{voms-proxy-init}. On the other hand, users could often have
+some local community/institutional credentials such as
+username/passwords to which users are more familiar. Enabling users to
 use their own community credential instead of the X.509 certificate to
-access grid is a promising solution to make the grid be more easily
+access grid is a promising solution to make the Grid more easily
 accessible.
@@
 \ \ \ \ The above analysis shows that in order to achieve
-interoperability and accessibility for user to access grid system which
+interoperability and accessibility for user to access Grid system which
 is established on different middlewares, we need to address the
@@
 middlewares in order to ensure the interoperability as wide as
-possible; also need to address the mismatch between
-user{\textquotesingle}s normal experience and
-middleware{\textquotesingle}s requirement in order to ensure the access
-as simple as possible.
+possible; it is also necessary to address the mismatch between
+users{\textquotesingle} usual experience and
+middleware{\textquotesingle}s requirements in order to ensure the
+access as simple as possible.
 
@@
 achieve the interoperability and accessibility without breaking
-existing grid specific security protocols.
+existing Grid specific security protocols.
 
-\ \ \ \ The rest of this paper is organized as follows: Section 2 shows
-the ARC grid middleware, especially the architecture of the new version
-of ARC. Section 3 describes the consideration and implementation about
-cross-middleware authentication and single sign-on is presented,
-including a few solutions implemented for bridging the mismatch in
-terms of authentication and single sign-on. Section 4 discusses the
-related research, and Section 5 concludes.
+\ \ \ \ The rest of this paper is organized as follows: Section 2
+presents the ARC Grid middleware, especially the architecture of the
+new version of ARC. Section 3 describes the solution for
+cross-middleware authentication and single sign-on, including
+implementations for bridging the mismatch in terms of authentication
+and single sign-on. Section 4 discusses the related research, and
+Section 5 presents conclusion.
 
@@
 {\centering
-2. ARC grid middleware 
+2. ARC Grid Middleware 
 \par}
 
-\ \ \ \ ARC(Advanced Resource Connector) is an open source grid
+\ \ \ \ ARC(Advanced Resource Connector) is an open source Grid
 middleware solution released under Apache license. ARC middleware
-initially aimed at developing a grid middleware which provides
-characteristics such as self-organized, fault-tolerant, non-intrusive,
-easy-management [10]. \ Now \ the classic version of ARC provides grid
-services such as grid job submission and management, resource
-characterization, resource aggregation and discovery, basic data
-management, integration of grid security solution, etc. Classic ARC has
-been deployed and used in production environment, and been one of the
-mainly deployed grid middlewares in Europe.
+initially aimed at developing self-organized, fault-tolerant,
+non-intrusive, easy-manageable Grid middleware [10]. \ Now \ the
+classic version of ARC provides Grid services for job submission and
+management, resource characterization, resource aggregation and
+discovery, basic data management, integration of grid security
+solution, etc. Classic ARC has been deployed and used in production
+environment, and been one of the widely deployed Grid middlewares in
+Europe.
 
-\ \ \ \ The new generation of ARC is based on the functionality and
-capabilities of the classic ARC middleware, and it aims at implementing
-a service-oriented middleware which will provide higher levels of
-resource and user abstracting through well-defined web service
-interface [11] in order to provide interoperability with other
-service-based grid middlewares, as well as other Web Service compatible
-applications.
+\ \ \ \ The new generation of ARC is developed by KnowARC project [11],
+based on the functionality and capabilities of the classic ARC
+middleware. It aims at implementing a Web Service oriented middleware
+which will provide higher levels of resource and user abstraction
+through well-defined web service interface [12] in order to provide
+interoperability with other service-based Grid middlewares, as well as
+other Web Service compatible applications.
 
@@
 there is a lightweight Web Service container called Hosting Environment
-Daemon (HED) which provides a host place for various services in
-application level, as well as a host place for a bunch of modules to
-support flexible, interoperatible, and efficient communication
-mechanism for building SOAP based Web Services. The whole design of the
-HED is built around the idea of flexibility and modularity, which means
-for the developer, he can easily concentrate on the application level
-Web Service implementation by only using the core minimum amount of
-components if he only would do application level development, or he can
-easily concentrate on the middleware level implementation such as
-supporting another communication protocol or implementing
+Daemon (HED) which provides a hosting for various services at
+application level, as well as a bunch of modules to support flexible,
+interoperatible, and efficient communication mechanism for building
+SOAP based Web Services. The whole design of the HED is built around
+the idea of flexibility and modularity, which means for the developer,
+he can easily concentrate on the application level Web Service
+implementation by only using the core minimum amount of components if
+he only would do application level development, or he can work on the
+middleware level implementing another communication protocol or
 authentication mechanism by using the core minimum amount of components
@@
 
+\begin{center}
+\includegraphics[width=3.6362in,height=3.3339in]{Secpaper-img1.png}
+\end{center}
 {\centering
-Figure 1. The architecture of Host Environment Daemon
+Figure 1. The example of Host Environment Daemon deployed with A-REX and
+File services
 \par}
 
-\begin{center}
-\includegraphics[width=4.0673in,height=3.1047in]{Secpaper-img1.png}
-\end{center}
-\ \ \ \ The architecture of the HED is illustrated as Figure 1. In
-general, there are a few \ components called Message Chain Component
-(MCC) which are in charge of processing different messages for
-different protocol levels. For instance, as shown in the example
-message flow, HTTP MCC will process secure socket stream from TLS MCC
-to generate output HTTP message to SOAP MCC, and also process SOAP
-documentation from SOAP MCC to generate HTTP message to TLS MCC.
+\ \ \ \ The architecture of the HED is illustrated by Figure 1. In
+general, there are few \ components called Message Chain Component
+(MCC) which are in charge of implementing different protocol levels.
+For instance, as shown in the example message flow, HTTP MCC will
+process stream from TLS MCC to \ parse HTTP message and pass its body
+to SOAP MCC, and also process SOAP response from SOAP MCC to generate
+HTTP message for TLS MCC.
 
-\ \ \ \ Two examples of MCC components configuration in Figure 1 shows
-MCC are flexibly configurable: real line shows SOAP over HTTPS (HTTP
-over TLS), while dotted line shows SOAP directly over TCP. Service
-administrator can configure the MCCs according to the interoperability
-requirement with the other part. For instance, the later example
-configuration is compatible to WSE (Web Services Enhancement for .NET)
-{\textquotesingle}s SOAP message mechanism (see WSE{\textquotesingle}s
-\textit{SoapSender} and \textit{SoapReceiver}). Another configuration
-could be SOAP over HTTPG (HTTP over GSI) which can be used to
-interoperate with some service based on HTTPG such as Storage Resource
-Manager (SRM)[12] \ service. The easily configurable characteristics
-shows the flexibility of HED in terms of protocol supporting.
+\ \ \ \ Dotted lines at Figure 1 show alternative path information can
+propagate among MCCs. Service administrator can configure the MCCs
+according to the interoperability requirements with the other part. For
+instance, the configuration marked with dashed lines is compatible to
+WSE (Web Services Enhancement for .NET) {\textquotesingle}s SOAP
+message mechanism (see WSE{\textquotesingle}s \textit{SoapSender} and
+\textit{SoapReceiver}). Another configuration could be SOAP over HTTPG
+(HTTP over GSI) which is needed to interoperate with services like
+Storage Resource Manager (SRM)[13] \ service. This shows the
+flexibility of HED in terms of protocols support.
 
@@
 {\textquotedblleft}AuthZ{\textquotedblright} and
-{\textquotedblleft}AuthN{\textquotedblright} sub-module inside MCCs and
-Services are examples of \textit{SecHandler}.
+{\textquotedblleft}AuthN{\textquotedblright} sub-modules inside MCCs
+and Services are examples of \textit{SecHandler}.
 
@@
 
-\liststyleLiii
+\liststyleLiv
 \begin{itemize}
@@
 utilize the standard-compatible community authentication mechanism from
-Shibboleth [13] Identity Provider software when users authenticate
+Shibboleth [14] Identity Provider software when users authenticate
 against Grid services.
 \item \textit{Short lived credential service} \ {}-{}-{}- to issue the
-short lived X.509 certificate for user by using the community
-authentication, in order to interoperate with services that depends on
+short lived X.509 certificates for users by using the community
+authentication, in order to interoperate with services that depend on
 PKI based mutual authentication.
@@
 
-\liststyleLiv
+\liststyleLv
 \begin{enumerate}
@@
 
+\bigskip
+
 3.2 Short-lived credential service
@@
 interface utility such as \textit{grid-proxy-init},
-\textit{voms-proxy-init}, or \textit{arcproxy}), and then use the proxy
-certificate to access Grid Service or Web Service.
+\textit{voms-}\textit{proxy-init}, or \textit{arcproxy}), and then use
+the proxy certificate to access Grid Service or Web Service.
 
@@
 (DN) for the certificate is a critical issue for SLCS service. Since
-the Shibboleth Identity Provider uses the eduPerson schema [14] for the
+the Shibboleth Identity Provider uses the eduPerson schema [15] for the
 definition of {\textless}\textit{saml:Attribute}{\textgreater} in
@@
 Service for processing delegation: delegation client, and delegation
-service. The delegation service has two main \textit{WSDL}(Web Service
-Description Language) operations, one for processing the delegation
-initiation; the other for storing the signed proxy certificate. The
-delegation process is detailed in Figure 5.
+service. The \textit{WSDL}(Web Service Description Language) of
+delegation service has two main operations, one for processing the
+delegation initiation, and \ the other for storing the signed proxy
+certificate. The delegation process is detailed in Figure 5.
 
-\ \ \ \ For the service invocation process, the WS client functionality
-(inside each service or the initial client) which is included to access
-the target service (step a1 or a2 in Figure 4), should use the
-user{\textquotesingle}s X.509 certificate or proxy certificate (on
-behalf of the user) for authentication and secure communication. For
-the delegation process, the principle is the same.
+\ \ \ \ For the delegation process, the delegation WS client
+functionality (inside each delegation service or the initial client)
+which is included to access the target delegation service (d1 or d2 in
+Figure 4), should use the user{\textquotesingle}s X.509 credential or
+delegated credential (on behalf of the user) for authentication and
+secure communication, as well as delegating one more level of delegated
+credential. For other general \ service invocation process, the WS
+client functionality (inside each service or the initial client) which
+is included to access the target service (step a1 or a2 in Figure 4),
+should also use the user{\textquotesingle}s X.509 credential or
+delegated credential (on behalf of the user) for authentication and
+secure communication.
 
-\ \ \ \ Since each WS client accesses the target service (delegation
-service or other functional service) on behalf of the user, the trust
+\ \ \ \ \ Since each WS client accesses the target service (delegation
+service or other general services) on behalf of the user, the trust
 relationship between the user{\textquotesingle}s certificates (X.509
 certificate and proxy certificate) and services is required, while the
-trust relationship between service and service is not required.
+trust relationship between service{\textquotesingle}s certificate and
+another service{\textquotesingle}s certificate is not required.
+
+\ \ \ \ What should be mentioned is that the delegation service is also
+a normal Web Service. Therefore, \ delegation service implemented in
+ARC can be accessed by any standard-compliant WS client that intends to
+delegate credential, also delegation client implemented in ARC can
+access any other standard-compliant Web Service which accepts
+delegation request.
 
@@
 {\centering
-Figure 4. Credential delegation service
+Figure 4. Credential delegation service and client
 \par}
@@
 
+\bigskip
+
 \begin{center}
-\includegraphics[width=2.3028in,height=1.6465in]{Secpaper-img5.png}
+\includegraphics[width=3.0638in,height=1.7827in]{Secpaper-img5.png}
 \end{center}
@@
 
+\bigskip
+
 3.4 Message level authentication
@@
 \ \ \ \ Transport level security (TLS or SSL) is not sufficient for
-securing Web Services, because the secure channel created by SSL/TLS
-make it impossible to differentially protect the SOAP message, e.g.
-\ to encrypt and/or sign only particular components of the SOAP
-message, which is relevant when non-sensitive portions of the message
-need to be accessed or changed by intermediate actors. Additionally,
-SSL/TLS is incapable of providing end-to-end protection for SOAP
-message which might flow through multiple intermediate actors, because
-it only allows each hop to be protected with the resultant security
-gaps at intermediate actors. Web Services Security (WS-Security)
-specification (ref) provides means for applying security to Web
-Service.
+securing Web Service, because the secure channel created by SSL/TLS
+make it impossible to differentially protect the SOAP message, e.g. to
+encrypt or sign only particular components of the SOAP message, which
+is relevant when non-sensitive portions of the message need to be
+accessed or changed by intermediate actors. Additionally, SSL/TLS is
+incapable of providing end-to-end protection for SOAP message which
+might flow through multiple intermediate actors, because it only allows
+each hop to be protected with the resultant security gaps at
+intermediate actors. Web Services Security (WS-Security)[8]
+specification provides means for applying security to Web Service.
 
-\ \ \ \ The WS-Security related specifications are implemented in ARC as
+\ \ \ \ WS-Security related specifications are implemented in ARC as
 \textit{SecHandler}. There are three kinds of WS-Security version 1.1
-specifications are completely or partly implemented: Username Token
-profile, X.509 Token profile, and SAML Token profile.
+specifications which are completely or partly implemented: Username
+Token profile, X.509 Token profile, and SAML Token profile.
 
@@
 \ \ \ \ If the WS client owns an X.509 certificate or proxy certificate,
-the process of generating an X.509 Token for a SOAP message is: the
+the process of generating an X.509 Token for a SOAP message is: this
 client puts the content of certificate into X.509 token as
-{\textless}wsse:BinarySecurityToken{\textgreater}, and uses the private
-key (corresponding to the certificate) to create XML signature for the
-SOAP message body.
+{\textless}\textit{wsse:BinarySecurityToken}{\textgreater}, and uses
+the private key (corresponding to the certificate) to create XML
+signature for the SOAP message body.
 
 \ \ \ \ However the process of generating a SAML Token is more
-complicated. In SAML Token profile, \textit{ConfirmationMethod} is the
-verifying method for establishing proof-of-possession for the subject.
-Two methods are proposed in SAML Token profile: \textit{holder-of-key},
-\textit{sender-vouches}. \ We rely on the \textit{holder-of-key
-}method. For the \textit{hold-of-key} subject confirmation method,
-there are three interaction parts: the attesting entity, the relying
-party and the issuing authority. Firstly the attesting entity
-authenticates to issuing authority by using some authentication scheme
-such as X.509 Token profile. So then issuing authority is able to make
-a definitive statement (sign a SAML assertion) about an act of
-authentication that has already taken place. The attesting entity gets
-the SAML assertion and then signs the SOAP message together with the
-assertion by using its private key (the relevant certificate has been
-authenticated by issuing authority, and its relevant public key has
-been put into \textit{SubjectConfirmation}
-
-element under SAML assertion by issuing authority. Only the actual owner
-of the SAML assertion can do this, as only the subject possesses the
-private key paired with the public key in the assertion. This
-establishes an irrefutable connection between the author of the SOAP
-message and the assertion describing an authentication event.)
+complicated. In WS-Security SAML Token profile,
+\textit{ConfirmationMethod} is the verifying method for establishing
+proof-of-possession for the subject. Two methods are proposed in SAML
+Token profile: \textit{holder-of-key}, \textit{sender-vouches}. We rely
+on the \textit{holder-of-key }method. For the \textit{hold-of-key}
+subject confirmation method, there are three interaction parts: the
+\textit{attesting entity}, the \textit{relying party} and the
+\textit{issuing authority}. Firstly the attesting entity authenticates
+to issuing authority by using some authentication scheme such as X.509
+Token profile, or SSL/TLS mutual authentication. So then issuing
+authority is able to make a definitive statement (sign a SAML
+assertion) about an act of authentication that has already taken place.
+The attesting entity gets the SAML assertion and then signs the SOAP
+message together with the assertion by using its private key (the
+relevant certificate has been authenticated by issuing authority, and
+its relevant public key has been put into \textit{SubjectConfirmation}
+element under SAML assertion by issuing authority). Only the exact
+entity that possesses the private key which is paired with the public
+key in the SAML assertion can sign the SOAP message, which establishes
+an irrefutable connection between the author of the SOAP message and
+the assertion describing an authentication event.
 
@@
 based on the public key in the SAML assertion without directly trust
-relationship with attesting entity (subject owner).
+relationship with attesting entity (holder of this public key).
 
-\ \ \ In terms of verification, since the of proxy certificates has been
-introduced in IETF (RFC 3820), \ and OpenSSL has also implemented
-RFC3820, it could not be difficult to use the proxy certificate for
-X.509 Token profile.
+\ \ \ \ In terms of verification, since the proxy certificates has been
+introduced in IETF (RFC 3820)[16], \ and OpenSSL has also implemented
+RFC3820, it is not difficult to use the proxy certificate for \ X.509
+Token profile to issuing X.509 token.
 
@@
 
-\ \ \ \ There is also delegation requirement on the message level, some
-entities (client or user) often need to temporarily delegate itself to
-another entity to perform action on their behalf. We propose a
-delegation mechanism for SAML Token.
+\ \ \ \ There is also delegation requirement on the message level: some
+entities often need to temporarily delegate itself to another entity to
+perform action on their behalf. We propose a delegation mechanism for
+SAML Token. Figure 6 shows the participant parties of SAML Token
+delegation.
 
-\ \ \ \ The delegation process is similar to the process for delegation
-proxy certificate, with the following differences:
+\ \ \ \ The delegation process is similar to the process in X.509
+credential delegation service, with the following differences:
 
-\liststyleLv
+\liststyleLvi
 \begin{enumerate}
-\item The communication between delegation client and service is based
-on mutual authentication between proxy certificate and service
-certificate, instead, it is based on mutual authentication between
-service certificate and service certificate.
-\item The delegation service will not generate any certificate request,
-instead, it picks the public key from its own service certificate, and
-sends back to delegation client.
+\item Transport level communication between delegation client and
+service is not based on mutual authentication between proxy certificate
+and service certificate, instead, it is based on mutual authentication
+between service certificate and service certificate.
+\item Delegation service will not generate any certificate request,
+instead, the delegation client will pick the public key from peer
+certificate after successful SSL/TLS authentication. So \ step2 of the
+delegation protocol will only be some confirmation information which
+lets delegation client to proceed. 
+\item For those multiple intermediate services through which a SOAP
+message flows, delegation service does not need to appear, instead, it
+only need to appear on the end of the SOAP message flow.
+\item SAML Token might include one or a few
+{\textless}\textit{saml:Attribute}{\textgreater} which carries the
+attribute information related to the subject, instead of the Attribute
+Certificate which is used in VOMS proxy certificate for carry the
+attribute information.
 \end{enumerate}
-\ \ \ \ Figure 6 shows the SOAP message header is changed due to
-delegation. On the left side, B is the identity of the attesting
-entity{\textquotesingle}s X509 token mentioned above, while A is the
-identity of issuing authority, and the SOAP message is signed by
-B{\textquotesingle}s private key. On the right side, C is the identity
-of the delegation service (also C is the identity of the target service
-\ to which the B will access), and now the SOAP message is signed by
-C{\textquotesingle}s private key.
+\ \ \ \ Figure 7 shows the SAML Token (SAML Assertion) inside SOAP
+message is changed due to delegation (we suppose X.509 token is used
+for attesting entity to authenticate against issuing authority). The
+SOAP message on the left side is the one which B sends to relying party
+(intermediate service), B is the identity of the attesting
+entity{\textquotesingle}s X.509 token, while A is the identity of
+issuing authority, and the SOAP message is signed by
+B{\textquotesingle}s private key which is corresponding to the X.509
+certificate that is used to create the X.509 token. On the right side,
+C is the identity of the delegation service (also C is the identity of
+the target service to which B will access), and now the SOAP message is
+signed by C{\textquotesingle}s private key which is corresponding to
+the target service{\textquotesingle}s certificate.
 
@@
 \begin{center}
-\includegraphics[width=4.2571in,height=2.152in]{Secpaper-img6.png}
+\includegraphics[width=4.1866in,height=2.7874in]{Secpaper-img6.png}
+\end{center}
+\begin{center}
+\includegraphics[width=6.9252in,height=0.0161in]{Secpaper-img7.png}
+\end{center}
+\begin{center}
+\includegraphics[width=6.9252in,height=0.0161in]{Secpaper-img8.png}
 \end{center}
 {\centering
-\ \ \ \ \ \ \ \ \ Figure 6. SOAP message header with SAML delegation
+Figure 6. SAML Token delegation service
+\par}
+
+\bigskip
+
+\bigskip
+
+\begin{center}
+\includegraphics[width=4.2957in,height=2.2575in]{Secpaper-img9.png}
+\end{center}
+
+\bigskip
+
+{\centering
+\ \ \ \ \ \ \ \ \ Figure 7. SOAP message header with SAML delegation
 token inside
 \par}
+
+\bigskip
 
@@
 
-\ \ There is some effort to add GSI support to adapt to the strict
-security requirement including GSI from grid middlewares, such as GSI
-plug-in for gSOAP[15]. The GSI MCC implemented in ARC can also achieve
-the same functionality as gSOAP. Moreover, the ARC solution is flexible
-than the gSOAP solution, because both the GSI MCC of service and client
-sides are pluggable, which means service and client developer of ARC do
-not need to modify the service or client implementation in order to add
-or remove the GSI functionality.
+There is some effort to add GSI support to adapt to the strict security
+requirement including GSI from grid middlewares, such as GSI plug-in
+for gSOAP[17]. The GSI MCC implemented in ARC can also achieve the same
+functionality as gSOAP. Moreover, the ARC solution is flexible than the
+gSOAP solution, because both the GSI MCC of service and client sides
+are pluggable, which means service and client developer of ARC do not
+need to modify the service or client implementation in order to add or
+remove the GSI functionality.
 
@@
 
-\textstyleCitation{\textup{[11] Design document of new version ARC.
+[11]KnowARC project. \ https://www.knowarc.eu/
+
+\textstyleCitation{\textup{[12] Design document of new version ARC.
 }}\url{https://www.knowarc.eu/documents/Knowarc_D1.1-1_07.pdf}
 
-\textstyleCitation{\textup{[12]A. Shoshani, A. Sim, and J. Gu,
+\textstyleCitation{\textup{[13]A. Shoshani, A. Sim, and J. Gu,
 }}\textstyleCitation{Storage Resource Managers: Essential Components
@@
 
-[13]The Shibboleth Project.
+[14]The Shibboleth Project.
 \href{http://shibboleth.internet2.edu/}{Http://shibboleth.internet2.edu/}
 
-[14]eduPerson and eduOrg Object shema.
+[15]eduPerson and eduOrg Object shema.
 \url{http://middleware.internet2.edu/eduperson/}
 
-\textstyleCitation{\textup{[15]}}\textstyleCitation{\textup{Giovanni
+\textstyleCitation{\textup{[16]RFC 3820 - \ Internet X.509 Public Key
+Infrastructure (PKI)}} Proxy Certificate Profile.
+\url{http://www.ietf.org/rfc/rfc3820.txt}
+
+\textstyleCitation{\textup{[17]}}\textstyleCitation{\textup{Giovanni
 Aloisio, Massimo Cafaro, Italo Epicoco, Daniele Lezzi, and Robert van
@@
 \bigskip
+
+\bigskip
+
+\bigskip
 \end{document}

### 2008-12-06 doc/sec/Sec_paper.tex
related work and reference

@@
 \begin{center}
-\includegraphics[width=3.6362in,height=3.3339in]{Secpaper-img1.png}
+\includegraphics[width=3.7465in,height=3.3764in]{Secpaper-img1.png}
 \end{center}
+
+\bigskip
+
 {\centering
@@
 
-\bigskip
-
-3.6 Discussion
-
-\bigskip
-
-\bigskip
-
 {\centering
@@
 
-\bigskip
+Similar work that integrates Shibboleth and Grid middleware does exist.
+GridShib[17][18] uses Shibboleth to get entity{\textquotesingle}s
+attribute for the authorization of services developed in Globus Tookit.
+Unlike the work of this paper, GridShib does not use
+Shibboleth{\textquotesingle}s native login handlers for authentication
+to Shibboleth IdP, instead, a new login handler specific for X.509
+based authentication is developed for Shibboleth IdP, and a credential
+service is developed for acting as an online-CA, authenticating the
+user, and issuing short-lived X.509 credential.
 
-There is some effort to add GSI support to adapt to the strict security
-requirement including GSI from grid middlewares, such as GSI plug-in
-for gSOAP[17]. The GSI MCC implemented in ARC can also achieve the same
-functionality as gSOAP. Moreover, the ARC solution is flexible than the
-gSOAP solution, because both the GSI MCC of service and client sides
-are pluggable, which means service and client developer of ARC do not
-need to modify the service or client implementation in order to add or
-remove the GSI functionality.
+SWITCH short-lived credential service (SLCS) [19] provides a SLCS
+service and client. The benefit of our work comparing to the
+SWTCH{\textquotesingle}s solution is the SLCS service and client is
+based on Web Service, so that it can be implemented in any Web Service
+container while keeping interoperability.
 
-\bigskip
+In terms of delegation, apart from the credential delegation mechanism
+from GSI [2][20], [21] proposes a delegation protocol based on the
+WS-Trust [22] in order to provide a standard and interoperable protocol
+for delegation in Grids. WS-Trust will also be adopted by ARC
+middleware to express the specifications required to define delegation
+protocol in a standard way. Gridsite implements a X.509 credential
+delegation solution [23] based on Web Service, with which ARC
+delegation client can easily interoperate.
 
-Shibboleth
-
-GridShib
-
-gLite Delegation Service?
-
-SWITCH SLCS Service
-
-the CCGrid 06 paper: Shibboleth-Protected privilege management
-infrastructure
-
-Delegation service : gridsite
-
-\bigskip
+GSI plug-in for gSOAP[24] adds GSI support for gSOAP Web Service tookit.
+The GSI MCC implemented in ARC can also achieve the same functionality
+as this plug-in. Moreover, the ARC solution is more flexible than it,
+because GSI MCC of both service and client sides are pluggable, which
+means service and client developer of ARC do not need to modify the
+service or client implementation in order to add or replace the GSI
+functionality.
 
@@
 
-\textstyleCitation{\textup{[8]Web Services Security(WS-Security).
+\textstyleCitation{\textup{[8]OASIS Web Services Security.
 www.oasis-open.org/committees/wss/}}
@@
 
-\textstyleCitation{\textup{[12] Design document of new version ARC.
+\textstyleCitation{\textup{[12]Design document of new version ARC.
 }}\url{https://www.knowarc.eu/documents/Knowarc_D1.1-1_07.pdf}
@@
 
-\textstyleCitation{\textup{[17]}}\textstyleCitation{\textup{Giovanni
+[17]V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes,
+Anonymity, and Access: Shibboleth and Globus Integration to Facilitate
+Grid Collaboration. 4th Annual PKI R\&D Workshop, 2005.
+
+[18]T. Barton, J. Basney, T. Freeman, T. Scavo, F. Siebenlist, V. Welch,
+R. Ananthakrishnan, B. Baker, and K. Keahey. Identity Federation and
+Attribute-based Authorization through the Globus Toolkit, Shibboleth,
+Gridshib, and MyProxy. 5th Annual PKI R\&D Workshop, 2006.
+
+[19]SWITCH Short Lived Credential Service.
+\url{http://www.switch.ch/grid/slcs/}
+
+[20]V. Welch, I. Foster, C. Kesselman, O. Mulmo, L. Pearlman, S. Teucke,
+J. Gawor, S. Meder and F. Siebenlist. X.509 proxy certificate for
+dynamic delegation, Proceeding of the 3rd Annual PKI R\&D Workshop,
+2004.
+
+[21]M. Ahsant, J. Basney, and O. Mulmo. Grid Delegation Protocol, UK
+Workshop on Grid Security Experiences, Oxford. 2004.
+
+[22]OASIS WS-Trust specification.
+\url{http://docs.oasis-open.org/ws-sx/ws-trust/200512}
+
+[23]gridsite elegation service.
+http://www.gridsite.org/wiki/Delegation\_protocol
+
+\textstyleCitation{\textup{[24]}}\textstyleCitation{\textup{Giovanni
 Aloisio, Massimo Cafaro, Italo Epicoco, Daniele Lezzi, and Robert van
-Engelen, }}\textstyleCitation{The GSI plug-in for gSOAP: Enhanced
-Security, Performance, and Reliability}\textstyleCitation{\textup{, in
-the ITCC conference 2005, IEEE Press, Volume I, pages 304-309.}}
+Engelen, The GSI plug-in for gSOAP: Enhanced Security, Performance, and
+Reliability, in the ITCC conference 2005, IEEE Press, Volume I, pages
+304-309.}}
 

### 2008-12-06 doc/sec/Sec_paper.tex
conclusion

@@
 and single sign-on. Section 4 discusses the related research, and
-Section 5 presents conclusion.
+Section 5 presents conclusion and future work.
 
@@
 
-\begin{center}
-\includegraphics[width=3.4063in,height=2.1417in]{Secpaper-img2.png}
-\end{center}
-
 \bigskip
 
+\begin{center}
+\includegraphics[width=3.2256in,height=1.8472in]{Secpaper-img2.png}
+\end{center}
 {\centering
@@
 \end{enumerate}
-\ \ \ \ The SP service and functional service(s) are hosted by the same
-container, and they use the same X.509 credential. The client
+\ \ \ \ The SP service and other functional service(s) are hosted by the
+same container, and they use the same X.509 credential. The client
 authentication is switched off, so that client doesn{\textquotesingle}t
@@
 
+\ \ \ It is worth mentioning is: since the SLCS X.509 credential or
+delegated credential can be used to issue X.509 token, \ the X.509
+credential delegation and SAML Token delegation are smoothly bridged,
+which benefits the single sign-on.
+
 \bigskip
@@
 \begin{center}
-\includegraphics[width=4.1866in,height=2.7874in]{Secpaper-img6.png}
+\includegraphics[width=4.0016in,height=2.639in]{Secpaper-img6.png}
 \end{center}
@@
 
-\begin{center}
-\includegraphics[width=4.2957in,height=2.2575in]{Secpaper-img9.png}
-\end{center}
-
 \bigskip
 
+\begin{center}
+\includegraphics[width=3.8374in,height=1.9063in]{Secpaper-img9.png}
+\end{center}
 {\centering
@@
 service is developed for acting as an online-CA, authenticating the
-user, and issuing short-lived X.509 credential.
+user, and issuing short-lived X.509 credential. Moreover, this paper
+provides a single sign-on solution for user authenticating to Grid
+services by directly using SAML2.0 SSO profile (section 3.1).
 
@@
 
-\bigskip
+\ \ \ \ Within this paper we presented the solution inside ARC Grid
+middleware towards the cross-middleware authentication and single
+sing-on. The goal of our work is to develop an interoperatable and
+accessible single sign-on solution to be used for VO scenario in
+production Grids, where users can easily authenticate against the Grids
+with any of his credentials (not only X.509 credential) from anywhere,
+and travel inside Grid middleware boundaries and even across
+heterogeneous Grids middleware boundaries, \ without re-authentication
+in a reasonable period. In order to do so, we implemented some standard
+based Web Services (including SLCS service, X.509 credential delegation
+service, and SAML Token delegation service), as well as the user agent
+and Service Provider in context of SAML2 SSO profile.
+
+\ \ \ It must be noted that due to standard specifications utilized
+(SAML and WS-Security) which are independent from any particular
+middleware, though the WS clients and Web/Grid Services presented in
+this paper are mostly developed on ARC middleware, they can
+interoperate with WS clients and Web/Grid Services developed on any
+other middlware in a standardized way. ARC also supports
+interoperability with Grid middlewares that require GSI based
+communication.
+
+\ \ \ \ Since authorization is also a key problem for VO, as the future
+work, we will implement a solution for cross-middleware authorization
+by using the standard specification such as SAML.
 

### 2008-12-11 doc/sec/SecurityFrameworkofARC1.tex
Put the added functionality into Sec technical documentation

@@
 }
+\newcommand\liststyleLi{%
+\renewcommand\theenumi{\arabic{enumi}}
+\renewcommand\theenumii{\arabic{enumii}}
+\renewcommand\theenumiii{\arabic{enumiii}}
+\renewcommand\theenumiv{\arabic{enumiv}}
+\renewcommand\labelenumi{\theenumi.}
+\renewcommand\labelenumii{\theenumii.}
+\renewcommand\labelenumiii{\theenumiii.}
+\renewcommand\labelenumiv{\theenumiv.}
+}
 % Page layout (geometry)
@@
 
-\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img3.png}
-
-\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img4.png}
-
-\includegraphics[width=0in,height=0.0161in]{SecurityFrameworkofARC1-img5.png}
+\includegraphics[width=0.0161in,height=0.0161in]{SecurityFrameworkofARC1-img3.png}
+
+\includegraphics[width=0.0161in,height=0.0161in]{SecurityFrameworkofARC1-img4.png}
+
+\includegraphics[width=0.0161in,height=0.0161in]{SecurityFrameworkofARC1-img5.png}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ are supposed to be used to load object; element
-{\textless}PolicyStore/{\textgreater} is}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ supposed to be used to get the location of
-policy-{}-{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}pdp:PDPConfig{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}pdp:PolicyStore{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}Location
-Type={\textquotedbl}file{\textquotedbl}{\textgreater}Policy\_Example.xml{\textless}/Location{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}- other policy
-location-{}-{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}/pdp:PolicyStore{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}pdp:Evaluator
-name={\textquotedbl}arc.evaluator{\textquotedbl} /{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}pdp:Policy
-name={\textquotedbl}arc.policy{\textquotedbl} /{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}pdp:Request
-name={\textquotedbl}arc.request{\textquotedbl} /{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ {\textless}/pdp:PDPConfig{\textgreater}}}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ are supposed to be used to load object; element
+{\textless}PolicyStore/{\textgreater} is}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ supposed to be used to get the location of
+policy-{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}pdp:PDPConfig{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}pdp:PolicyStore{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ {\textless}Location
+Type={\textquotedbl}file{\textquotedbl}{\textgreater}Policy\_Example.xml{\textless}/Location{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}- other policy
+location-{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}/pdp:PolicyStore{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}pdp:Evaluator
+name={\textquotedbl}arc.evaluator{\textquotedbl} /{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}pdp:Policy
+name={\textquotedbl}arc.policy{\textquotedbl} /{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}pdp:Request
+name={\textquotedbl}arc.request{\textquotedbl} /{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}/pdp:PDPConfig{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}PDP
-name={\textquotedbl}arc.pdp{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ {\textless}PolicyStore{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}Location
-type={\textquotedbl}file{\textquotedbl}{\textgreater}Policy\_Example.xml{\textless}/Location{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}!-{}-other policy
-location-{}-{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/PolicyStore{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}/PDP{\textgreater}}}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}PDP
+name={\textquotedbl}arc.pdp{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}PolicyStore{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}Location
+type={\textquotedbl}file{\textquotedbl}{\textgreater}Policy\_Example.xml{\textless}/Location{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}!-{}-other policy location-{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/PolicyStore{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}/PDP{\textgreater}}
 
@@
 CombiningAlg={\textquotedbl}Deny-Overrides{\textquotedbl}{\textgreater}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}Rule
-Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ {\textless}Description{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ Example policy for echo service}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ {\textless}/Description{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Subjects{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Subject{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}Rule
+Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ {\textless}Description{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ Example policy for echo service}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ {\textless}/Description{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Subjects{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Subject{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/ca{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=CA{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=CA{\textless}/Attribute{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}Attribute
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/tls/identity{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=test{\textless}/Attribute{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Subject{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Subjects{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Resources{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Resource
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}/C=NO/ST=Oslo/O=UiO/CN=test{\textless}/Attribute{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}/Subject{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/Subjects{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Resources{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Resource
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/path{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}/Echo{\textless}/Resource{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Resources{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}/Echo{\textless}/Resource{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/Resources{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Actions{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Action
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}POST{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}POST{\textless}/Action{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Action
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/http/method{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}GET{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Action
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}GET{\textless}/Action{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Action
 AttributeId={\textquotedbl}http://www.nordugrid.org/schemas/policy-arc/types/soap/operation{\textquotedbl}
-Type={\textquotedbl}string{\textquotedbl}{\textgreater}echo{\textless}/Action{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/Actions{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ \ \ \ {\textless}Conditions/{\textgreater}}}
-\item {\selectlanguage{english}\color{black}
-\texttt{\ \ \ {\textless}/Rule{\textgreater}}}
+Type={\textquotedbl}string{\textquotedbl}{\textgreater}echo{\textless}/Action{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/Actions{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}Conditions/{\textgreater}}
+\item {\selectlanguage{english}\ttfamily\color{black}
+\ \ \ {\textless}/Rule{\textgreater}}
 \item {\selectlanguage{english}\ttfamily\color{black}
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}!-{}-Remote pdp service
-invoking-{}-{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ {\textless}PDP
-name={\textquotedbl}pdpservice.invoker{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}ServiceEndpoint{\textgreater}https://127.0.0.1:60001/pdp.service{\textless}/ServiceEndpoint{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}KeyPath{\textgreater}./key.pem{\textless}/KeyPath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}CertificatePath{\textgreater}./cert.pem{\textless}/CertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ {\textless}CACertificatePath{\textgreater}./ca.pem{\textless}/CACertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ {\textless}/PDP{\textgreater}}}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}!-{}-Remote pdp service invoking-{}-{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}PDP
+name={\textquotedbl}pdpservice.invoker{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}ServiceEndpoint{\textgreater}https://127.0.0.1:60001/pdp.service{\textless}/ServiceEndpoint{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}KeyPath{\textgreater}./key.pem{\textless}/KeyPath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}CertificatePath{\textgreater}./cert.pem{\textless}/CertificatePath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ {\textless}CACertificatePath{\textgreater}./ca.pem{\textless}/CACertificatePath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ {\textless}/PDP{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ {\textless}PDP
-name={\textquotedbl}delegation.pdp{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}/SecHandler{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ {\textless}next
-id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ {\textless}PDP
+name={\textquotedbl}delegation.pdp{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ {\textless}/SecHandler{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ {\textless}next
+id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}Component
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}Component
 name={\textquotedbl}tls.service{\textquotedbl}
 id={\textquotedbl}tls{\textquotedbl}{\textgreater} {\textless}next
-id={\textquotedbl}http{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:KeyPath{\textgreater}./key.pem{\textless}/tls:KeyPath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CertificatePath{\textgreater}./cert.pem{\textless}/tls:CertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CACertificatePath{\textgreater}./ca.pem{\textless}/tls:CACertificatePath{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}-delegation.collector
-must be inside tls MCC-{}-{\textgreater}}}
+id={\textquotedbl}http{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:KeyPath{\textgreater}./key.pem{\textless}/tls:KeyPath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CertificatePath{\textgreater}./cert.pem{\textless}/tls:CertificatePath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CACertificatePath{\textgreater}./ca.pem{\textless}/tls:CACertificatePath{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}-delegation.collector must be
+inside tls MCC-{}-{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
 
@@
 
+\subsection[Web Service Security Support]{\color{black} Web Service
+Security Support}
 \subsubsection[UsernameToken SecHandler configuration]{UsernameToken
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}Component
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}Component
 name={\textquotedbl}soap.service{\textquotedbl}
-id={\textquotedbl}soap{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}next
-id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
+id={\textquotedbl}soap{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
 name={\textquotedbl}usernametoken.handler{\textquotedbl}
 id={\textquotedbl}usernametoken{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}Process{\textgreater}extract{\textless}/Process{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}PasswordSource{\textgreater}password.txt{\textless}/PasswordSource{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}/SecHandler{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}
+event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}Process{\textgreater}extract{\textless}/Process{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}PasswordSource{\textgreater}password.txt{\textless}/PasswordSource{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ {\textless}/SecHandler{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}Component
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}Component
 name={\textquotedbl}soap.service{\textquotedbl}
-id={\textquotedbl}soap{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ {\textless}next
-id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
+id={\textquotedbl}soap{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
 name={\textquotedbl}x509token.handler{\textquotedbl}
 id={\textquotedbl}x509token{\textquotedbl}
-event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}}
+event={\textquotedbl}incoming{\textquotedbl}{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}Process{\textgreater}extract{\textless}/Process{\textgreater}}}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}Process{\textgreater}extract{\textless}/Process{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}CACertificatePath{\textgreater}ca.pem{\textless}/CACertificatePath{\textgreater}}}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}CACertificatePath{\textgreater}ca.pem{\textless}/CACertificatePath{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/SecHandler{\textgreater}}}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/SecHandler{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\texttt{\ \ \ \ \ \ \ {\textless}/Component{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\ttfamily\color{black}
+\ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\color{black}
 X509Token SecHandler must be configured under SOAP MCC.}
+
+\bigskip
+
+\subsection[Using Shibboleth IdP for Authentication and Attribute-based
+Authorization]{Using \foreignlanguage{english}{Shibboleth} IdP for
+Authentication and Attribute-based Authorization}
+{\selectlanguage{english}\color{black}
+\ \ \ \ In grid community, GSI (grid security infrastructure) is the
+de-facto standard about transport level communication for the legacy
+grid solution, which implements some enhancement (such as delegation)
+based on standard SSL3.0/TLS1.0. In ARC1, besides that GSI is supported
+for talking with external grid services which are based on GSI, the
+standard TLS/SSL is also supported.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ No matter standard SSL/TLS or GSI is used, SSL/TLS based mutual
+authentication applies for both of them, and is the default
+configuration for grid deployment; and X.509 certificate is required
+for both of the client and service sides. X.509 certificates is issued
+by certificate authorities (CA), and then CAs constitute trust
+federation and guarantee two different X.509 certificates from
+different CAs can accomplish authentication to each other. So if a user
+would access grid system, he/she should own a X.509 certificate which
+is issued by a CA that is trusted by other{\textquotesingle}s entity in
+the grid system.}
+
+{\selectlanguage{english}\upshape\color{black}
+\ AAI (Authentication and Authorization Infrastructure) is a solution
+for the authentication and authorization in inter-organization resource
+sharing, such as electronic resource sharing between libraries, etc.
+AAI implicitly applies to community or institutional based
+authentication where users from different home communities need to get
+resources from other communities by using some federation mechanism.
+Unlike the X.509 based authentication solution in current Grid systems,
+AAI does not require users to provide X.509 certificate, instead, it
+can support different types of authentication, such as
+username/password authentication, IP address authentication, etc. There
+are several implementations of AAI, among which Shibboleth is one
+implementation which has been widely deployed.}
+
+{\selectlanguage{english}\upshape\color{black}
+\ \ \ \ Shibboleth provides cross-domain single sign-on and
+attribute-based authorization while preserving user privacy. It is
+based on the OASIS Security Assertion Markup Language (SAML),
+specifically, the new version of Shibboleth supports SAML 2.0
+specification. For authentication, the main SAML profile that
+Shibboleth implements is the SAML2.0 web browser SSO profile, which
+defines two functional components, an Identity Provider and a Service
+Provider. The Identity Provider (IdP) is responsible for creating,
+maintaining, and managing user identity, while the Service Provider
+(SP) is responsible for controlling access to services and resources by
+using the SAML assertion produced and issued by IdP upon request. In
+order to discover which home community does a user come from,
+Shibboleth specifies an optional third component called
+{\textquotedblleft}Where Are You From?{\textquotedblright} (WAYF)
+service to \ aid in the process of IdP discovery, and this IdP
+discovery process is also standardized and defined in SAML 2.0
+specification and called as {\textquotedblleft}Identity Profile
+Discovery Profile{\textquotedblright}.}
+
+{\selectlanguage{english}\upshape\color{black}
+\ \ \ \ The SAML2.0 web browser SSO profile is utilized for the
+authentication in ARC middleware. But since the SSO profile is
+primarily supposed to protect Web applications and provide
+authentication for Web users, some external code on the client and
+service side is implemented to integrate the SSO profile. On the client
+side, apart from the client interface for writing Web Service client,
+the user agent functionality of the Web browser is implemented in order
+to mimic its behavior, such as HTTP redirection and cookie processing.
+In fact, implementation of the user agent is also based on the client
+interface of ARC, specifically, the HTTPs client interface, since the
+client interface of ARC can support different protocols which are
+incarnated by different MCCs. The client developers who would use
+SAML2.0 SSO profile should call the user agent interface and then the
+Web Service client interface. On the service side, the Service Provider
+functionality (based on the HTTP MCC configured together with TLS MCC)
+is implemented, which is called SP Service. For Identity Provider, the
+Shibboleth IdP implementation is used. Figure 6 shows the process of
+SAML2.0 SSO integrated in ARC client and service.}
+
+\begin{center}
+\includegraphics[width=3.2256in,height=1.8472in]{SecurityFrameworkofARC1-img7.png}
+\end{center}
+
+\bigskip
+
+{\centering\selectlanguage{english}\itshape\color{black}
+Figure \stepcounter{Figure}{\theFigure}.\textbf{ }SAML2.0 SSO profile in
+ARC
+\par}
+
+{\selectlanguage{english}\color{black}
+\ The steps shows in Figure 6 are described as follows:}
+
+\liststyleLi
+\begin{enumerate}
+\item {\selectlanguage{english}\color{black}
+The client uses the user agent interface to launch a HTTP request
+including the IdP name (to which the user belongs) to the service side.
+The endpoint of the SP (Service Provider) service is the same as that
+of the other target services, except the last part of the endpoint is
+{\textquotedblleft}saml2sp{\textquotedblright} which is specific for
+pointing to the SP Service. Note that we use Identity Provider (IdP)
+name here to simplify the IdP discovery process in order to avoid the
+IdP discovery process, because we suppose that the user who would
+access the target services should better know where is he from
+initially.}
+\item {\selectlanguage{english}\color{black}
+The SP Service searches the metadata (we use the same metadata format as
+defined in Shibboleth) and gets the location of the single sign-on
+service (hosted in IdP) and also the location of assertion consuming
+service (hosted in this SP itself) in order to compose the SAML
+{\textless}samlp:AuthnRequest{\textgreater} message. Then SP Service
+issues this {\textless}samlp:AuthnRequest{\textgreater} message by
+using its own X.509 certificate (note that in the SAML SSO profile,
+X.509 certificates are still needed for IdP and SP) and sends back to
+user agent.}
+\item {\selectlanguage{english}\color{black}
+User agent sends the {\textless}samlp:AuthnRequest{\textgreater} message
+to the Identity Provider.}
+\item {\selectlanguage{english}\color{black}
+Identity Provider requires an act of authentication. The authentication
+mechanism is outside of the SAML2.0 SSO profile. Shibboleth IdP
+implementation chooses some login handlers for authentication. The
+current user agent implementation is compatible with the
+Username/Password login handler of Shibboleth IdP. Through the HTTP
+protocol, the user agent will feed IdP with the username/password which
+has been given by the caller of user agent interface.}
+\item {\selectlanguage{english}\color{black}
+Once the authentication has been succeeded, the IdP issues a SAML
+response including an encrypted (encrypted by destination
+SP{\textquoteright}s public key) SAML assertion, and then this SAML
+response will be delivered by the user agent to the Service Provider.}
+\item {\selectlanguage{english}\color{black}
+The SP Service verifies and checks the SAML response, decrypts and
+stores the SAML assertion into session/connection context. The SAML
+assertion includes the {\textless}saml:AuthnStatement{\textgreater} and
+{\textless}saml:AttributeStatement{\textgreater}.}
+\item {\selectlanguage{english}\color{black}
+The WS client launches the Grid/Web Service request via the same
+connection as the one which is used by user agent to contact SP
+Service.}
+\item {\selectlanguage{english}\color{black}
+The Grid/Web Service checks the
+{\textless}saml:AuthnStatement{\textgreater} from the session context
+to see if the session is still valid through the SecHandler called
+{\textquotedblleft}SAML SecHandler{\textquotedblright}. If valid,
+service handles the service processing and returns the response to WS
+client. Note that service requires that WS client is from the same
+connection as the one on which user agent contact SP service, in order
+to guarantee that the validity of SSO profile result effects the WS
+client/Web Service interaction.}
+\end{enumerate}
+{\selectlanguage{english}\upshape\color{black}
+\ \ \ \ The SP service and other functional service(s) are hosted by the
+same container, and they use the same X.509 credential. The client
+authentication is switched off, so that client doesn{\textquoteright}t
+need to use any X.509 credential. Only the trusted certificates (CA
+certificates for both SP and IdP) need to be configured for the client
+side, so that SP and IdP can authenticate themselves to the client. As
+required by the SAML2.0 profile, the SP and IdP should have trust
+relationship to each other.}
+
+{\selectlanguage{english}\upshape\color{black}
+\ \ \ \ One benefit of the SAML2.0 SSO profile that is worth mentioning
+is: the Identity Provider could cache the authentication result through
+session management once the user agent has succeeded to authenticate;
+then for a short period this authentication result is valid so that the
+user agent doesn{\textquoteright}t need to feed IdP with
+user{\textquoteright}s username and password the next time (if this
+point of time is not out of the scope of valid period) it authenticates
+against IdP. So user (or the client on behalf of this user) can travel
+across multiple security \ domains with only providing his name and
+password once, which is the characteristic of single sign-on.}
+
+{\selectlanguage{english}\upshape\color{black}
+\ \ \ \ Since the Shibboleth implementation of SAML is
+standard-compliant and widely deployed, the solution implemented in ARC
+can easily interoperate with other SAML implementations with minimum
+change, and more importantly, this solution can succeed to utilize the
+widely deployed SAML implementation for authentication in Grid systems
+by avoiding the usage of X.509 certificate.}
+
+{\selectlanguage{english}\upshape\color{black}
+\ \ \ \ Moreover, even though the implementation is based on the ARC
+middleware, the idea can be adopted by other Grid middlewares if they
+only require server authentication instead of mutual authentication.}
+
+\bigskip
+
+\subsection[\ Short-Lived Credential
+Service]{\foreignlanguage{english}{\ }Short-Lived Credential Service}
+{\selectlanguage{english}\color{black}
+\ \ \ \ However, most of the widely used Grid middlewares are based on
+GSI while GSI requires mutual authentication. Also for Web Service
+based Grid solution, we cannot prevent service side from requiring
+client X.509 certificate. Based on the solution described in Section 9,
+a short lived credential service (SLCS) is implemented by which user
+can get a short-lived X.509 certificate without being bothered to
+contact any registration authority (RA) or certificate authority (CA).}
+
+{\selectlanguage{english}\upshape\color{black}
+\ \ \ \ The SLCS service is also a Web Service (standard Web Service
+implemented by using ARC service interface), and the SLCS client is a
+specific command-line interface (CLI) which includes the user agent and
+WS client. The whole process of SLCS invocation is showed in Figure 7
+(from step 1 to step 8), which is the same as in Figure 6, except that
+step 7 and step 8 are invoked for the SLCS certificate request and
+response.}
+
+\begin{center}
+\includegraphics[width=3.7689in,height=3.2055in]{SecurityFrameworkofARC1-img8.png}
+\end{center}
+{\centering\selectlanguage{english}\itshape\color{black}
+Figure \stepcounter{Figure}{\theFigure}.Short lived credential service
+\par}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SLCS client generates \ a X.509 certificate \ request, launches
+a Web Service request which includes the certificate request; SLCS
+service then gets the certificate request, composes a distinguished
+name (DN), issues a certificate (short lived, 12 hours by default) with
+the SAML attribute (from the SAML2.0 SSO profile) as the X.509
+certificate extension, and puts the certificate in to the Web Service
+response; SLCS client get the response and stores the X.509 certificate
+into local repository.}
+
+{\selectlanguage{english}\color{black}
+The CLI for the SLCS client is like this:}
+
+{\selectlanguage{english}\ttfamily\color{black}
+\$ ./arcslcs -S https://127.0.0.1:60000/slcs -I
+https://idp.testshib.org/idp/shibboleth -U myself -P myself}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ Since the lifetime of the short lived credential is normally
+short, it is not a must to protect the private key by a pass phrase. As
+illustrated in steps (a) and (b) in Figure 7, if the private key is not
+protected through the Web Service client, the user can use the X.509
+certificate to access Grid Service or Web Service from any kind of
+middleware. If the private key is protected, she can use the X.509
+certificate to generate a proxy certificate (by using a command-line
+interface utility such as grid-proxy-init , voms-proxy-init, or
+arcproxy ), and then use the proxy certificate to access a Grid/Web
+Service.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ It is worth mentioning that since the ARC middleware can support
+GSI communication by configuring the GSI MCC, together with the X.509
+certificate, the Web Service client developed with the ARC Web Service
+client interface can interoperate with Grid Service that requires GSI
+communication.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ It should be noticed that the process of composing the
+distinguished name (DN) for the certificate is a critical issue for the
+SLCS service. Since the Shibboleth Identity Provider uses the eduPerson
+schema for the definition of {\textless}saml:Attribute{\textgreater} in
+{\textless}saml:AttributeStatement{\textgreater}, we pick the
+relatively distinguishable attribute
+{\textquotedblleft}eduPersonPrincipalName{\textquotedblright} for the
+DN. A typical eduPersonPrincipalName value could be alice@example.org,
+then the DN is
+{\textquotedblleft}/O=knowarc/OU=example.org/CN=alice{\textquotedblright}.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ The obvious benefit of the SLCS service is that: If a user
+passes the authentication to her home Identity Provider, she can get
+the X.509 credential anywhere simply by running the SLCS client command
+together with providing her username and password to this home IdP, and
+then access the Grid system conveniently. Thanks to the single sign-on
+characteristic of SAML2.0 SSO profile, this user
+doesn{\textquoteright}t not need to input her username and password in
+a valid period after the first time she succeeds to authenticate
+against her home IdP by running the SLCS client command on the same
+node, even if this SLCS client command is supposed to run against a few
+SLCS services to get a few SLCS credentials.}
+
+\bigskip
+
+\subsection[\ X.509 Credential Delegation
+Service]{\selectlanguage{english}\scshape\color{black} \ X.509
+Credential Delegation Service}
 

### 2008-12-11 doc/client/client_technical.tex
Replaced sentences containing the word propriety. Probably the intention was that they should say propietary, but since the things being described weren't proprietary other words were used instead.
Starting broker latex documentation

@@
 submission makes use of the {\Broker}, {\JobDescription} and
-{\Submitter} classes. At the time of writing the {\Broker} is not
-implemented, but it's functionality and interface is well known, and
-thus we will include it in the overview given here. The {\Submitter}
-is base class for further grid middleware specific specialization
-(plugin) similarly to the {\TargetRetriever}.
+{\Submitter} classes.  The {\Submitter} is base class for further grid
+middleware specific specialization (plugin) similarly to the {\TargetRetriever}.
 
@@
 
-The {\Broker} is not yet implemented and thus not described here.
+There are two usable broker type at the current state: \textbf{FastQueueBroker} and \textbf{RandomBroker}.
+These classes are inheritance of the "super \textbf{Broker}" class.
+This parent class has a prefiltering part and its task is to create a mapping between the \textbf{JobDescription} and the \textbf{ExecutionTarget}.
+The mapping is working but it will be change at the near future because the \textbf{JobDescription} class will be  recreate with \textbf{GIN Profile JSDL}.
+The mapping result is the \textbf{PossibleTargets} vector which content the "good" \textbf{ExecutionTargets}.
+The descendant broker classes task is to sort the \textbf{PossibleTargets} vector with its name type, for example: \textbf{RandomBroker} sort type is \textbf{Random}.
+The broker separated into a module therefore it is loadable.
 
@@
 The ARC 0.x local information system uses the {\GT}~\cite{globus} GRIS
-with a propriety ARC schema. The information index server used is the
+with a purpose made ARC schema. The information index server used is the
 {\GT} GIIS. Both these servers are using the LDAP~\cite{ldap}
@@
 
-The CREAM interface is a propriety SOAP based interface. The CREAM
+CREAM has its own SOAP based interface. The CREAM
 specializations of the {\Submitter} and {\JobController} classes

### 2008-12-15 doc/sec/SecurityFrameworkofARC1.tex
Description for slcs configuration into Security techinical document

@@
 
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 \ \ \ \ The SLCS service is also a Web Service (standard Web Service
@@
 \par}
+
+\bigskip
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SLCS service is supposed to run together with SP service for
+achieving the SAML2 SSO profile. A typical configuration for SLCS
+service is as follows:}
+
+\begin{center}
+\tablehead{}\begin{supertabular}{|m{5.6886597in}|}
+\hline
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Chain{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}tcp.service{\textquotedbl}
+id={\textquotedbl}tcp{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}tls{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tcp:Listen{\textgreater}{\textless}tcp:Port{\textgreater}60000{\textless}/tcp:Port{\textgreater}{\textless}/tcp:Listen{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}tls.service{\textquotedbl}
+id={\textquotedbl}tls{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}http{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:KeyPath{\textgreater}./testkey-nopass.pem{\textless}/tls:KeyPath{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CertificatePath{\textgreater}./testcert.pem{\textless}/tls:CertificatePath{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}-tls:CACertificatePath{\textgreater}./cacert.pem{\textless}/tls:CACertificatePath-{}-{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:ClientAuthn{\textgreater}false{\textless}/tls:ClientAuthn{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}http.service{\textquotedbl}
+id={\textquotedbl}http{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}plexer{\textquotedbl}{\textgreater}POST{\textless}/next{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Plexer
+name={\textquotedbl}plexer.service{\textquotedbl}
+id={\textquotedbl}plexer{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}samlsp{\textquotedbl}{\textgreater}/saml2sp{\textless}/next{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}soap{\textquotedbl}{\textgreater}/slcs{\textless}/next{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Plexer{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}soap.service{\textquotedbl}
+id={\textquotedbl}soap{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}slcs{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
+name={\textquotedbl}saml2ssoserviceprovider.handler{\textquotedbl}
+id={\textquotedbl}saml2ssosp{\textquotedbl}
+event={\textquotedbl}incoming{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Service
+name={\textquotesingle}saml.sp{\textquotesingle}
+id={\textquotesingle}samlsp{\textquotesingle}{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}MetaDataLocation{\textgreater}./test\_metadata.xml{\textless}/MetaDataLocation{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}ServiceProviderName{\textgreater}https://squark.uio.no/shibboleth-sp{\textless}/ServiceProviderName{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Service{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Service
+name={\textquotedbl}slcs.service{\textquotedbl}
+id={\textquotedbl}slcs{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}slcs{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CACertificate{\textgreater}./CAcert.pem{\textless}/slcs:CACertificate{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CAKey{\textgreater}./CAkey.pem{\textless}/slcs:CAKey{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CASerial{\textgreater}./CAserial{\textless}/slcs:CASerial{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Service{\textgreater}}
+
+\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Chain{\textgreater}\\\hline
+\end{supertabular}
+\end{center}
+{\selectlanguage{english}\color{black}
+\ \ \ \ As shown in the above table, a plexer dispatches the message
+flow (outgoing from \textit{http.service}) into two SP service
+(\textit{saml.sp}) and SOAP service (\textit{soap.service}). Since the
+client authentication is switched off, it is not necessary to configure
+{\textless}\textit{CACertificatePath}{\textgreater} or
+{\textless}\textit{CACertificateDir}{\textgreater} for TLS MCC.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SP service needs to be configured with
+{\textless}\textit{MetaDataLocation}{\textgreater} and
+{\textless}\textit{ServiceProviderName}{\textgreater}. }
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SLCS service needs to be configured with
+{\textless}\textit{CACertificate}{\textgreater}
+{\textless}\textit{CAKey}{\textgreater} and
+{\textless}\textit{Caserial}{\textgreater}.}
 
@@
 
+\bigskip
+
 \subsection[\ X.509 Credential Delegation

### 2008-12-15 doc/infosys/TechnicalHandbook/ISIS - TechnicalHandbook.tex
Registrant configuration schema

@@
 \hypersetup{
-  pdfauthor = {Author Name},
-  pdftitle = {Paper title},
-  pdfsubject = {Paper subject},
+  pdfauthor = {Ivan Marton},
+  pdftitle = {ISIS Technical handbook},
+  pdfsubject = {ISIS Technical handbook},
   pdfkeywords = {Paper,keyword,comma-separated},
@@
   \verb#// Initialize phase#\\
-  \verb# Read the configuration and store the information about the services in a list#\\
-  \verb# do { // In a different Thread#\\
+  \verb#Read the configuration and store the information about the services in a list#\\
+  \verb#do { // Cyclic phase in a different Thread#\\
   \verb# wake_up_time = now();#\\
@@
 
+\subsection{Configuration}
+\label{Configuration}
+The configuration of the different modules can be done through the ARC Configuration file. The format of these entries is shown by the relevant XSD Schema files.
+
+\begin{framed}
+Registrant.xsd\\
+\\
+\verb#<?xml version="1.0" encoding="UTF-8"?>#\\
+\verb#<xsd:schema#\\
+\verb#  xmlns:xsd="http://www.w3.org/2001/XMLSchema"#\\
+\verb#  xmlns:registrant="http://www.nordugrid.org/schemas/RegistrantConfig/2008"#\\
+\verb#  xmlns:arcConfig="http://www.nordugrid.org/schemas/ArcConfig/2007"#\\
+\verb#  targetNamespace="http://www.nordugrid.org/schemas/RegistrantConfig/2008"#\\
+\verb#  elementFormDefault="qualified">#\\
+\verb# #\\
+\verb#    <xsd:complexType name="Registrant_Type">#\\
+\verb#        <!-- This element defines configuration of Information #\\
+\verb#            Registration interface. -->#\\
+\verb#        <xsd:sequence>#\\
+\verb#            <!-- Period specifies how often registration to be done in seconds.#\\
+\verb#                The missing or 0 value means a non period registration. -->#\\
+\verb#            <xsd:element name="Period" type="xsd:nonNegativeInteger"/>#\\
+\verb#            <!-- URL specifies contact endpoint of a bootstrap Information#\\
+\verb#                Registration service. Further ISIS's adresses will be queried#\\
+\verb#                from this service. -->#\\
+\verb#            <xsd:element name="ISIS_URL" type="xsd:string" />#\\
+\verb#            <!-- Optional KeyPath, CertificatePath, ProxyPath and CACertificatesDir#\\
+\verb#                are paths to files storing X509 credentials used for establishing#\\
+\verb#                connections. -->#\\
+\verb#            <xsd:element name="KeyPath" type="xsd:string" minOccurs="0" />#\\
+\verb#            <xsd:element name="CertificatePath" type="xsd:string" minOccurs="0" />#\\
+\verb#            <xsd:element name="ProxyPath" type="xsd:string" minOccurs="0" />#\\
+\verb#            <xsd:element name="CACertificatesDir" type="xsd:string" minOccurs="0" />#\\
+\verb#        </xsd:sequence>#\\
+\verb#    </xsd:complexType>#\\
+\verb#    <xsd:element name="Registrant" type="registrant:Registrant_Type"/>#\\
+\verb# #\\
+\verb#    <!-- Optional element for Service element overriding the default Period length. -->#\\
+\verb#    <xsd:element name="Period" type="xsd:nonNegativeInteger"/>#\\
+\verb# #\\
+\verb#</xsd:schema>#\\
+\end{framed}
+
 %Advanced Resource Connector (ARC) was invented by

### 2008-12-15 doc/sec/ManualofSecurityFramework.tex
some changes in documentation

@@
 \usepackage[pdftex]{graphicx}
-% Text styles
-\newcommand\textstyleInternetlink[1]{\textcolor[rgb]{0.0,0.0,0.5019608}{#1}}
 % Outline numbering
@@
 \renewcommand\arraystretch{1.3}
-\newcounter{Bibitem}
-\renewcommand\theBibitem{\arabic{Bibitem}}
 \title{}
@@
 
+\bigskip
+
 \subsubsection{SP Service and SAML2SSO Service Provider Handler}
 {\selectlanguage{english}\color{black}
-Service Provider servicer (SP service) implements the functionality of
-Service Provider in SAML2 Web SSO profile. }
-
-{\selectlanguage{english}\color{black}
-Normally SP service will not be deployed independently, instead, it
-should be deployed together with other services. The SLCS service is
+\ \ \ \ Service Provider servicer (SP service) implements the
+functionality of Service Provider in SAML2 Web SSO profile.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ Normally SP service will not be deployed independently, instead,
+it should be deployed together with other services. The SLCS service is
 the typical deployment about SP service. On the client side, client
@@
 {\selectlanguage{english}\color{black}
-Once the SP Service has cooperated with user agent and Identity Provider
-(external) and succeeded to accomplish the SAML2 SSO profile, \ the SP
-service will get the saml authentication assertion which asserts that
-the authentication has succeeded, and then SP service will store this
-assertion into session context.}
-
-{\selectlanguage{english}\color{black}
-SAML2 service provider handler (saml2ssoserviceprovider.handler) is the
-security handler which will understand the authentication assertion and
-make authorization decision according to this assertion.}
-
-{\selectlanguage{english}\color{black}
-The example configuration about using SP service and SAML2SSO service
-provider handler can be seen in the configuration of SLCS service. }
-
-{\selectlanguage{english}\color{black}
-SP service is not only supposed to work together with SLCS service. But
-if it is used to work together with other services, the client should
-be developed on \ ClientSAML2Interface instead of ClientInterface.}
+\ \ \ \ Once the SP Service has cooperated with user agent and Identity
+Provider (external) and succeeded to accomplish the SAML2 SSO profile,
+\ the SP service will get the saml authentication assertion which
+asserts that the authentication has succeeded, and then SP service will
+store this assertion into session context.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SAML2 assertion consumer handler
+(saml2ssoassertionconsumer.handler) is the security handler which will
+understand the authentication assertion and attribute assertion from
+IdP, and make authorization decision according to the attribute values
+inside these assertions. Currently the SAML2 assertion consumer handler
+is an empty security handler and does not effect the services.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ The example configuration about using SP service and SAML2SSO
+assertion consumer handler can be seen in the configuration of SLCS
+service. }
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SP service is not only supposed to work together with SLCS
+service. If it is used to work together with other services, the client
+should be developed on ClientSAML2Interface instead of ClientInterface,
+which is the same as that in SLCS client (\textit{arcslcs}).}
+
+\bigskip
 
@@
 {\selectlanguage{english}\color{black}
-Short-lived credential service (SLCS) is for signing short-lived x509
-credential for user based user{\textquotesingle}s username/password
-credential. Then the user can use the short-lived x509 credential to
-access grid services/resources where x509 credential is required by
-default. \ SLCS service should depend on the SP (Service provider)
-service which is one of the participants of SAML2 SSO profile (SAML2
-SSO profile is used for authenticating based on username/password
-credential and getting SAML authentication assertion; and SAML
-authentication assertion is then used as basis for signing short-lived
-x509 credential).}
+\ \ \ \ Short-lived credential service (SLCS) is for signing short-lived
+x509 credential for user based user{\textquotesingle}s
+username/password credential. Then the user can use the short-lived
+x509 credential to access grid services/resources where x509 credential
+is required by default. \ SLCS service should depend on the SP (Service
+provider) service which is one of the participants of SAML2 SSO profile
+(SAML2 SSO profile is used for authenticating based on
+username/password credential and getting SAML authentication assertion;
+and SAML authentication assertion is then used as basis for signing
+short-lived x509 credential).}
 
@@
 \ \ \ \ {\textless}SecHandler
-name={\textquotedbl}\textcolor[rgb]{0.8627451,0.13725491,0.0}{saml2ssoserviceprovider.handler}{\textquotedbl}
+name={\textquotedbl}\textcolor[rgb]{0.8627451,0.13725491,0.0}{saml2ssoassertionconsumer.handler}{\textquotedbl}
 id={\textquotedbl}saml2ssosp{\textquotedbl}
@@
 {\selectlanguage{english}\color{black}
-SLCS service and SP service should together be configured by using the
-Plexer. SP service is directly based on http and SLCS is directly based
-on soap. }
-
-{\selectlanguage{english}\color{black}
-SLCS service should specifically include the CA credential (certificate
-and key file) and the serial number file (which includes the serial
-number of each signed certificate).}
+\ \ \ \ SLCS service and SP service should together be configured by
+using the Plexer. SP service is directly based on http and SLCS is
+directly based on soap. }
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SLCS service should specifically include the CA credential
+(certificate and key file) and the serial number file (which includes
+the serial number of each signed certificate). Therefore, the service
+administrator should firstly create a CA credential (note for
+interoperability purpose in the production grid deployment, the CA
+credential should be trusted by others). }
 
@@
 \ \ {}-D, -{}-storedir=directory \ \ \ \ Store directory for key and
-signed certificate }
+signed certificate}
 
@@
 {\selectlanguage{english}\color{black}
-Note user should input the {\textquotedblleft}IdP
+\ \ \ \ Note user should input the {\textquotedblleft}IdP
 name{\textquotedblright} is the corresponding Identity Provider (one of
 the participants of SAML2 SSO profile) name to which the user would
-authentication against by using its username/password credential. And
-the name is stored inside the metadata on both SP service and IdP
-provider, and it is used by SP service to get the authentication URL
-for username/password based authentication.}
-
-{\selectlanguage{english}\color{black}
-And the {\textquotedblleft}user{\textquotedblright} and
+authenticate against by using its username/password credential. And the
+name is stored inside the metadata on both SP service and IdP provider,
+and it is used by SP service to get the authentication URL for
+username/password based authentication.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ And the {\textquotedblleft}user{\textquotedblright} and
 {\textquotedblleft}password{\textquotedblright} is the credential which
@@
 
-\bigskip
-
-\bigskip
-
-\bigskip
-
-\clearpage{\centering\selectlanguage{english}\bfseries\scshape\color{black}
-References
-\par}
-
-{\selectlanguage{english}\color{black}
-\stepcounter{Bibitem}{\theBibitem}. RFC3820- Internet X.509 Public Key
-Infrastracture (PKI) Proxy, \url{http://rfc.net/rfc3820.html}}
-
-{\selectlanguage{english}\color{black}
-\stepcounter{Bibitem}{\theBibitem}. D1.2-2 The ARC container (first
-prototype),
-\url{http://www.knowarc.eu/documents/Knowarc_D1.2-2_07.pdf}}
-
-{\selectlanguage{english}\color{black}
-\textstyleInternetlink{\textcolor{black}{\stepcounter{Bibitem}{\theBibitem}}}\textstyleInternetlink{\textcolor{black}{.}}\textstyleInternetlink{\foreignlanguage{english}{\textcolor{black}{Web
-Service Security Username Token Profile 1.1.
-}}}\url{http://www.oasis-open.org/committees/download.php/16782/wss-v1.1-spec-os-UsernameTokenProfile.pdf}}
-
-{\selectlanguage{english}\color{black}
-\stepcounter{Bibitem}{\theBibitem}. OGSA{\textregistered} Basic
-Execution Service Version 1.0,
-\ \url{http://www.ogf.org/documents/GFD.108.pdf} }
-
-{\selectlanguage{english}\color{black}
-\textstyleInternetlink{\textcolor{black}{\stepcounter{Bibitem}{\theBibitem}}}\textstyleInternetlink{\textcolor{black}{.
-OASIS Web Services Security (WSS) TC.
-}}\url{http://www.oasis-open.org/committees/tc_home.php?wg_abbrev=wss}}
+{\selectlanguage{english}\color{black}
+\ \ \ \ There is a temporary Identity Provider (IdP) deployed on
+\textit{squark.uio.no} for test with the following test username and
+password: staff, researcher, librarian, binduser; with the same
+password {\textquotedbl}123456{\textquotedbl}}
+
+{\selectlanguage{english}\color{black}
+./arcslcs -S https://127.0.0.1:60000/slcs -I
+https://squark.uio.no/idp/shibboleth -U staff -P 123456 -D
+/home/wzqiang/arc-0.9/src/clients/credentials -c client.xml}
+
+{\selectlanguage{english}\upshape\color{black}
+\ \ \ \ The short-lived credential issued by SLCS service will include
+the SAML assertion as the extension of X.509 certificate as a proof of
+passing SAML2 SSO profile.}
 \end{document}

### 2008-12-15 doc/sec/SecurityFrameworkofARC1.tex
some changes in documentation

@@
 \ \ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
-name={\textquotedbl}saml2ssoserviceprovider.handler{\textquotedbl}
+name={\textquotedbl}saml2ssoassertionconsumer.handler{\textquotedbl}
 id={\textquotedbl}saml2ssosp{\textquotedbl}

### 2008-12-23 doc/sec/SecurityFrameworkofARC1.tex
Adding VOMS attributes description.

--- 2008-12-11 doc/sec/SecurityFrameworkofARC1.tex
@@
-% This file was converted to LaTeX by Writer2LaTeX ver. 0.5
+% This file was converted to LaTeX by Writer2LaTeX ver. 0.9.3
 % see http://www.hj-gym.dk/~hj/writer2latex for more info
-\documentclass{article}
+\documentclass[a4paper]{article}
 \usepackage[ascii]{inputenc}
@@
 \usepackage[spanish,english]{babel}
-\usepackage{amsmath,amssymb,amsfonts,textcomp}
+\usepackage{amsmath}
+\usepackage{amssymb,amsfonts,textcomp}
 \usepackage{color}
@@
 \usepackage{hyperref}
-\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=qiangwz, pdfsubject=, pdfkeywords=}
+\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=Aleksandr Konstantinov, pdfsubject=, pdfkeywords=}
 \usepackage[pdftex]{graphicx}
@@
 \setcounter{secnumdepth}{4}
-\renewcommand\thesection{\arabic{section}.}
-\renewcommand\thesubsection{\arabic{subsection}.}
-\renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection}.}
-\renewcommand\theparagraph{\arabic{subsection}.\arabic{subsubsection}.\arabic{paragraph}.}
+\renewcommand\thesection{\arabic{section}}
+\renewcommand\thesubsection{\arabic{subsection}}
+\renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection}}
+\renewcommand\theparagraph{\arabic{subsection}.\arabic{subsubsection}.\arabic{paragraph}}
 \makeatletter
@@
 }
-\newcommand\liststyleLi{%
+\newcommand\liststyleLii{%
 \renewcommand\theenumi{\arabic{enumi}}
@@
 % Page layout (geometry)
-\setlength\paperwidth{8.2673in}
-\setlength\paperheight{11.6925in}
 \setlength\voffset{-1in}
@@
 \title{}
+\author{Aleksandr Konstantinov}
+\date{2008-12-23}
 \begin{document}
@@
 \bigskip
-\clearpage\clearpage\setcounter{page}{1}\pagestyle{Convertv}
+\clearpage\clearpage\pagestyle{Convertv}
 
@@
 \textit{delegation.collector}, and evaluates the request against
-delegation policy. See section 6for the design idea and use case of
-delegation policy in fine-grained identity delegation.}
+delegation policy. See section \ref{bkm:Ref204009393}for the design
+idea and use case of delegation policy in fine-grained identity
+delegation.}
 
@@
 {\selectlanguage{english}\itshape\color{black}
- Rule (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Subjects (1)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Subject (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-    Attribute (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Resources (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Resource (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Actions (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Action (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Conditions (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-   Condition (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-    Attribute (1-)}
+\ \ Rule (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ Subjects (1)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ Subject (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ Attribute (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ Resources (0-1)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ Resource (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ Actions (0-1)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ Action (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ Conditions (0-1)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ Condition (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ Attribute (1-)}
 
@@
 {\selectlanguage{english}\itshape\color{black}
- RequestItem (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Subject (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-   SubjectAttribute (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Resource (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Action (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-  Context (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-   ContextAttribute (1-)}
+\ \ RequestItem (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ Subject (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ SubjectAttribute (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ Resource (0-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ Action (0-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ Context (0-)}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ ContextAttribute (1-)}
 
@@
 \subsubsection{Basic Elements of Policy}
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 There are 2 basic objects - {\textquotedbl}policy{\textquotedbl} and
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 Evaluator matches request to policy and produces one of 4 following
@@
 \item {\selectlanguage{english}\color{black}
-\textbf{PERMIT} - policy explicitely permits activity specified in
-request because request matches some part of policy and corresponding
-effect specified in policy is PERMIT. }
-\end{itemize}
-{\bfseries\itshape\color{black}
+\foreignlanguage{english}{\textbf{PERMIT}}\foreignlanguage{english}{ -
+policy explicitely permits activity specified in request because
+request matches some part of policy and corresponding effect specified
+in policy is PERMIT. }}
+\end{itemize}
+{\selectlanguage{english}\bfseries\itshape\color{black}
 Example: }
@@
 {\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Rule: }}\textit{PERMIT
-person ALICE to PLAY in place called WONDERLAND}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Request: }}\textit{person
-ALICE wants to PLAY in place called WONDERLAND}}
+\foreignlanguage{english}{\textbf{\textit{\ \ \ \ \ \ Rule:
+}}}\foreignlanguage{english}{\textit{PERMIT person ALICE to PLAY in
+place called WONDERLAND}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{\ \ \ \ \ \ Request:
+}}}\foreignlanguage{english}{\textit{person ALICE wants to PLAY in
+place called WONDERLAND}}}
 
@@
 \item {\selectlanguage{english}\color{black}
-\textbf{DENY} - policy explicitely denies activity specified in Request
-because Request matches some part of policy and corresponding effect
-specified in policy is DENY.}
-\end{itemize}
-{\bfseries\itshape\color{black}
+\foreignlanguage{english}{\textbf{DENY}}\foreignlanguage{english}{ -
+policy explicitely denies activity specified in Request because Request
+matches some part of policy and corresponding effect specified in
+policy is DENY.}}
+\end{itemize}
+{\selectlanguage{english}\bfseries\itshape\color{black}
 Example:}
@@
 {\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Rule: DENY }}\textit{fruit
-APPLE to GROW on PEACH tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Request}}\textit{: fruit
-APPLE to be GROWN on PEACH tree}}
+\foreignlanguage{english}{\textbf{\textit{\ \ \ \ \ \ Rule: DENY
+}}}\foreignlanguage{english}{\textit{fruit APPLE to GROW on PEACH
+tree}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{\ \ \ \ \ \ Request}}}\foreignlanguage{english}{\textit{:
+fruit APPLE to be GROWN on PEACH tree}}}
 
@@
 \item {\selectlanguage{english}\color{black}
-\textbf{INDETERMINATE} - request has some part which does not correspond
-to policy.}
-\end{itemize}
-{\bfseries\itshape\color{black}
+\foreignlanguage{english}{\textbf{INDETERMINATE}}\foreignlanguage{english}{
+- request has some part which does not correspond to policy.}}
+\end{itemize}
+{\selectlanguage{english}\bfseries\itshape\color{black}
 Example: }
@@
 {\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Rule}}\textit{: DENY fruit APPLE to
-GROW on PEACH tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit APPLE to
-be GROWN on WHEAT ground}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: flower SUNFLOWER
-to be grown on PEACH tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Explaination}}: Here, it is not possible to obtain any
-matching result - neither positive (DENY or PERMIT) nor negative
-(NOT\_APPLICABLE, see below)}
-
-{\color{black}
+\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Rule}}}\foreignlanguage{english}{\textit{:
+DENY fruit APPLE to GROW on PEACH tree}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
+fruit APPLE to be GROWN on WHEAT ground}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
+flower SUNFLOWER to be grown on PEACH tree}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{Explaination}}}\foreignlanguage{english}{:
+Here, it is not possible to obtain any matching result - neither
+positive (DENY or PERMIT) nor negative (NOT\_APPLICABLE, see below)}}
+
+{\selectlanguage{english}\color{black}
 In the request, the {\textquotedbl}ground{\textquotedbl} is completely
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 In a similar way one can{\textquotesingle}t compare {\textquotedbl}fruit
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 Any other situation which makes it impossibile to compare two attributes
@@
 \item {\selectlanguage{english}\color{black}
-\textbf{NOT\_APPLICABLE} - all parts of the Request have corresponding
-parts in the Policy, but some value of those parts are not the same.
-Hence request does not match policy.}
-\end{itemize}
-{\bfseries\itshape\color{black}
+\foreignlanguage{english}{\textbf{NOT\_APPLICABLE}}\foreignlanguage{english}{
+- all parts of the Request have corresponding parts in the Policy, but
+some value of those parts are not the same. Hence request does not
+match policy.}}
+\end{itemize}
+{\selectlanguage{english}\bfseries\itshape\color{black}
 Example:}
@@
 {\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Rule}}\textit{: DENY fruit APPLE to
-GROW on PEACH tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit APPLE to
-be GROWN on APPLE tree }}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit ORANGE to
-be GROWN on PEACH tree }}
-
-{\selectlanguage{english}\upshape\color{black}
-\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit ORANGE to
-be GROWN on APPLE tree }}
-
-{\color{black}
+\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Rule}}}\foreignlanguage{english}{\textit{:
+DENY fruit APPLE to GROW on PEACH tree}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
+fruit APPLE to be GROWN on APPLE tree }}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
+fruit ORANGE to be GROWN on PEACH tree }}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
+fruit ORANGE to be GROWN on APPLE tree }}}
+
+{\selectlanguage{english}\color{black}
 Explanation: for each part of the Request evaluator can find relevant
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 If it is required to reduce evaluation results to boolean value PERMIT
@@
 {\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Note}}\textit{: It would be useful to make it possible
-to specify secondary effect which would become }active in case Request
-is NOT\_APPLICABLE. For example:}
-
-{\itshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{Note}}}\foreignlanguage{english}{\textit{:
+It would be useful to make it possible to specify secondary effect
+which would become }}\foreignlanguage{english}{active in case Request
+is NOT\_APPLICABLE. For example:}}
+
+{\selectlanguage{english}\itshape\color{black}
 DENY fruit APPLE to GROW on PEACH tree otherwise PERMIT}
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 But one should be careful because example above would allow fruit PLUMS
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 This kind of requirement can be supported by using the algorithm between
@@
 \subsubsection{Policy MAtching}
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 Policy is made of Rule elements. Request is evaluated against each Rule.
@@
 \item {\selectlanguage{english}\color{black}
-\textbf{\textit{Deny-Overrides}} - this is default if no algorithm
-specified.}
-
-\begin{itemize}
-\item {\color{black}
+\foreignlanguage{english}{\textbf{\textit{Deny-Overrides}}}\foreignlanguage{english}{
+- this is default if no algorithm specified.}}
+
+\begin{itemize}
+\item {\selectlanguage{english}\color{black}
 If there is at least one DENY in results final result is DENY.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is at least one PERMIT, the final result is PERMIT.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is at least one NOT\_APPLICABLE final result is
 NOT\_APPLICABLE.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise final result is INDETERMINATE.}
@@
 \end{itemize}
-{\selectlanguage{english}\upshape\color{black}
-\textbf{NOTE}\textbf{: CHECK if algorithm description is correct}}
-
-{\color{black}
+{\selectlanguage{english}\bfseries\color{black}
+NOTE: CHECK if algorithm description is correct}
+
+{\selectlanguage{english}\color{black}
 Special case is Policy wih no rules. Probably such policy shoudl be
@@
 
-{\bfseries\color{black}
+{\selectlanguage{english}\bfseries\color{black}
 Discussion}
 
-{\itshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
 If there is no Rule under Policy, it means no restriction from this
@@
 
-{\itshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
 It seems to be more logical to produce INDETERMINATE because there is no
@@
 \begin{itemize}
-\item {\bfseries\itshape\color{black}
+\item {\selectlanguage{english}\bfseries\itshape\color{black}
 Permit-Overrides.}
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 If there is at least one PERMIT in results final result is PERMIT.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is at least one DENY the final result is DENY.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is at least one NOT\_APPLICABLE final result is
 NOT\_APPLICABLE.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise final result is INDETERMINATE.}
@@
 \end{itemize}
-{\color{black}
+{\selectlanguage{english}\color{black}
 Special case is Policy wih no rules. Probably such policy should be
@@
 
-{\bfseries\color{black}
+{\selectlanguage{english}\bfseries\color{black}
 NOTE: CHECK if algorithm description is correct}
 
-{\bfseries\color{black}
+{\selectlanguage{english}\bfseries\color{black}
 Discussion}
 
-{\itshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
 Question? The same as above.}
@@
 \begin{itemize}
-\item {\bfseries\itshape\color{black}
+\item {\selectlanguage{english}\bfseries\itshape\color{black}
 Ordered algorithms.}
@@
 \item \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 If there is at least one result of Result1 type then final result is
 Result1.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is at least one result of Result2 type then final
 result is Result2.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is at least one result of Result3 type then final
 result is Result3.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise final result is Result4.}
@@
 \end{itemize}
-{\color{black}
+{\selectlanguage{english}\color{black}
 There are 24 possible combinations of those algorithms.}
@@
 {\selectlanguage{english}\upshape\color{black}
-\textbf{Note:} It would be useful to have more combining algorithms. For
-example }
+\foreignlanguage{english}{\textbf{Note:}}\foreignlanguage{english}{ It
+would be useful to have more combining algorithms. For example }}
 
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 If this Rule gives DENY in results, final result is DENY.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if this Rule gives NOT\_APPLICABLE, final result is PERMIT.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise final result is INDETERMINATE. }
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 If all of the Rule give PERMIT, the final result is PERMIT.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is at least one DENY the final result is DENY.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is at least one NOT\_APPLICABLE final result is
 NOT\_APPLICABLE.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise final result is INDETERMINATE.}
 \end{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 OnlyOneApplicable}
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 If there is one gives INDETERMINATE, final result INDETERMINATE is given
 immediately.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is exactly only one gives applicable result (DENY or
 PERMIT), final result is as this result.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if there is more than one gives applicable result, final
@@
 \end{itemize}
-{\color{black}
+{\selectlanguage{english}\color{black}
 This algorithm makes sure that only one Rule is selected when making
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 FirstApplicable}
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 If there is one give DENY, PERMIT or INDETERMINATE result, final result
 is given immediately as this result.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise final result is NOT\_APPLICABLE.}
@@
 \subsubsection{Request Structure}
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 Request is made of RequestItem elements. Each RequestItem is evaluated
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 Subject - represents entity requesting specified action}
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Resource - destination/object of the action}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Action - specifies what has to be done on resource}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Context - for additional information which does not fit anywhere else,
@@
 \end{itemize}
-{\color{black}
+{\selectlanguage{english}\color{black}
 Effectively RequestItem may have only one Subject, one Resource, one
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 Additionally Subject could contain sub-elements SubjectAttribute. Those
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Subject}
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 SubjectAttribute: name is ALICE}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 SubjectAttribute: age is YOUNG}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 SubjectAttribute: gender is GIRL}
@@
 \end{itemize}
-{\color{black}
+{\selectlanguage{english}\color{black}
 Context could also be made of ContextAttribute elements in the same way
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 The following is an example of the Request:}
 
-{\itshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
 {\textless}Request
@@
 
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Resource
-AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Context
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}RequestItem{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}SubjectAttribute
+AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}SubjectAttribute
+AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Subject
+AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Action
+AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Resource
+AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Context
 AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
-Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}
-
-{\itshape\color{black}
+Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}/RequestItem{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
 {\textless}/Request{\textgreater}}
 
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 While evaluating this RequestItem will be split into two RequestItems:}
 
-{\itshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
 {\textless}Request
@@
 
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Resource
-AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Context
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}RequestItem{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}SubjectAttribute
+AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}SubjectAttribute
+AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Action
+AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Resource
+AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Context
 AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
-Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Resource
-AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Context
+Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}/RequestItem{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}RequestItem{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Subject
+AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Action
+AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Resource
+AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Context
 AttributeId={\textquotedbl}urn:knowarc:time{\textquotedbl}
-Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}/RequestItem{\textgreater}}}
-
-{\itshape\color{black}
+Type={\textquotedbl}time{\textquotedbl}{\textgreater}2008-09-15T20:30:20{\textless}/Context{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}/RequestItem{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
 {\textless}/Request{\textgreater}}
 
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 The following means this Subject possesses both of these Attributes.}
 
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\color{black}
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}SubjectAttribute
+AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}SubjectAttribute
+AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\color{black}
 However, the following means two Subject each of which possesses one
@@
 
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Subject{\textgreater}}}
-
-{\color{black}
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Subject
+AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Subject
+AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\color{black}
 The {\textquotedbl}Type{\textquotedbl} xml-attribute is for
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 The {\textquotedbl}AttributeId{\textquotedbl} xml-attribute is for
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 Each RequestItem will be sequencialy and independently evaluated against
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 It is up to policy decision point to make final decision according to
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\textbf{NOTE: This probably should be changed because evaluator is fed
-with complete policy and complete request. Hence it is illogical that
-it retu}\textbf{rns multiple decisions.}}
-
-{\color{black}
+{\selectlanguage{english}\bfseries\color{black}
+NOTE: This probably should be changed because evaluator is fed with
+complete policy and complete request. Hence it is illogical that it
+returns multiple decisions.}
+
+{\selectlanguage{english}\color{black}
 Basically the policy decision point will feed policy engine with
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 and make final decision.}
@@
 \subsubsection{Rule Composition and Matching}
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 Policy rule is made of 4 elements - Subjects, Resources, Actions,
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\textbf{NOTE: That is strange. Why do we need }\textbf{Rule at all?}}
-
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\bfseries\color{black}
+NOTE: That is strange. Why do we need Rule at all?}
+
+{\selectlanguage{english}\color{black}
 There is no logical relationship between Subject, which means you can
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 MATCHED - element from RequestItem matched element in Policy Rule.
@@
 {\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{RequestItem Resource}}\textit{: place called
-WONDERLAND}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{PolicyItem Resource}}\textit{: place called WONDERLAND}}
+\foreignlanguage{english}{\textbf{\textit{RequestItem
+Resource}}}\foreignlanguage{english}{\textit{: place called
+WONDERLAND}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{PolicyItem
+Resource}}}\foreignlanguage{english}{\textit{: place called
+WONDERLAND}}}
 
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 NOT MATCHED - element from RequestItem did not match element in Policy
@@
 {\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{RequestItem Resource}}\textit{: place called
-WONDERLAND}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{PolicyItem Resource}}\textit{: place called PLAYGROUND}}
+\foreignlanguage{english}{\textbf{\textit{RequestItem
+Resource}}}\foreignlanguage{english}{\textit{: place called
+WONDERLAND}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{PolicyItem
+Resource}}}\foreignlanguage{english}{\textit{: place called
+PLAYGROUND}}}
 
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 INDETERMINATE- element from RequestItem could not be compared to element
@@
 {\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{RequestItem Resource}}\textit{: place called WONDERLAND
-(with namespace {\textquotedbl}place{\textquotedbl})}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{PolicyItem Resource}}\textit{: LEMON tree (with
-namespace {\textquotedbl}tree{\textquotedbl})}}
-
-{\color{black}
+\foreignlanguage{english}{\textbf{\textit{RequestItem
+Resource}}}\foreignlanguage{english}{\textit{: place called WONDERLAND
+(with namespace {\textquotedbl}place{\textquotedbl})}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{PolicyItem
+Resource}}}\foreignlanguage{english}{\textit{: LEMON tree (with
+namespace {\textquotedbl}tree{\textquotedbl})}}}
+
+{\selectlanguage{english}\color{black}
 The produced results then combined to produce final 4 types of results
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 If for every element in RequestItem there is at least one MATCHED result
@@
 Effect (Deny or Permit).}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise if for every element in RequestItem there is at least one gets
 INDETERMINATE result then result for Policy Rule is INDETERMINATE.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Otherwise result is NOT\_APPLICABLE.}
 \end{itemize}
-{\color{black}
+{\selectlanguage{english}\color{black}
 Special case is then RequestItem does not have the element(s) of some
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 The following is an example of the Policy:}
 
-{\itshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
 {\textless}Policy
@@
 
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}Rule
-Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}Subjects{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}/Subjects{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}Actions{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Action
-AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Action
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}Rule
+Effect={\textquotedbl}Permit{\textquotedbl}{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ {\textless}Subjects{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject
+AttributeId={\textquotedbl}urn:knowarc:shibboleth:attribute{\textquotedbl}{\textgreater}member{\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ {\textless}/Subjects{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ {\textless}Actions{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Action
+AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Read{\textless}/Action{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Action
 AttributeId={\textquotedbl}urn:knowarc:fileoperation{\textquotedbl}{\textgreater}Delete{\textless}/Action{\textgreater}
-\ \ }}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}/Actions{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}Resources{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Resource
-AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textquotedbl}{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}/Resources{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}Conditions{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ }\textit{{\textless}Condition
+\ \ }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ {\textless}/Actions{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ {\textless}Resources{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Resource
+AttributeId={\textquotedbl}urn:knowarc:fileidentity{\textquotedbl}{\textgreater}file:///home/test{\textless}/Resource{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ {\textless}/Resources{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ {\textless}Conditions{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ {\textless}Condition
 AttributeId={\textquotedbl}urn:knowarc:period{\textquotedbl}
 Type={\textquotedbl}period{\textquotedbl}
-Function={\textquotedbl}Inrange{\textquotedbl}{\textgreater}2008-09-10T20:30:20/P1Y1M{\textless}/Condition{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ }\textit{{\textless}/Conditions{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}/Rule{\textgreater}}}
-
-{\itshape\color{black}
+Function={\textquotedbl}Inrange{\textquotedbl}{\textgreater}2008-09-10T20:30:20/P1Y1M{\textless}/Condition{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ {\textless}/Conditions{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}/Rule{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
 {\textless}/Policy{\textgreater}}
 
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 For the Subject which includes two Attributes in this example:}
 
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\color{black}
 These two attributes mean the Rule requires the request should possess
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 However, if You put these above two Attribute into two Subject
@@
 
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject
-AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject
+AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject
+AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\color{black}
 Then it means the Rule requires the request to possess at least one of
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 For the xml-attribute {\textquotedbl}Type{\textquotedbl} and
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 The {\textquotedbl}Function{\textquotedbl} xml-attribute is for
@@
 \subsubsection{Rule Elements Matching}
-{\color{black}
+{\selectlanguage{english}\color{black}
 For elements without attributes those elements have:}
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Kind specified by AttributeId XML attribute. There is no default.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Matching algorithm specified by Id XML attribute. By default
 string-equal matching is used.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Content}
@@
 {\selectlanguage{english}\upshape\color{black}
-Example: \textit{LEMON tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Kind}}\textit{: tree}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Matching algorithm}}\textit{: default}}
-
-{\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Content}}\textit{: LEMON}}
-
-{\color{black}
+\foreignlanguage{english}{Example:
+}\foreignlanguage{english}{\textit{LEMON tree}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{Kind}}}\foreignlanguage{english}{\textit{:
+tree}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{Matching
+algorithm}}}\foreignlanguage{english}{\textit{: default}}}
+
+{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{Content}}}\foreignlanguage{english}{\textit{:
+LEMON}}}
+
+{\selectlanguage{english}\color{black}
 Matching procedure consists of following steps:}
@@
 \begin{itemize}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Kinds are compared using simple string equal matching. If those do not
 match then result is INDETERMINATE.}
-\item {\color{black}
+\item {\selectlanguage{english}\color{black}
 Matching algorithm is used to compare content of elements. Result is
@@
 \end{itemize}
-{\color{black}
+{\selectlanguage{english}\color{black}
 Each element on the RequestItem must satisfy corresponding element in
@@
 
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 For elements with multiple Attribute sub-elements the way to judging
@@
 
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 Example of the Subject with three Attributes:}
@@
 {\selectlanguage{english}\upshape\color{black}
-\textbf{\textit{Subject}}\textit{:}}
-
-{\selectlanguage{english}\color{black}
-\textbf{\textit{SubjectAttribute}}\textit{: name is ALICE}}
-
-{\selectlanguage{english}\color{black}
-\textbf{\textit{SubjectAttribute}}\textit{: age is YOUNG}}
-
-{\selectlanguage{english}\color{black}
-\textbf{\textit{SubjectAttribute}}\textit{: gender is GIRL}}
-
-{\selectlanguage{english}\upshape\color{black}
+\foreignlanguage{english}{\textbf{\textit{Subject}}}\foreignlanguage{english}{\textit{:}}}
+
+{\selectlanguage{english}\color{black}
+\foreignlanguage{english}{\textbf{\textit{SubjectAttribute}}}\foreignlanguage{english}{\textit{:
+name is ALICE}}}
+
+{\selectlanguage{english}\color{black}
+\foreignlanguage{english}{\textbf{\textit{SubjectAttribute}}}\foreignlanguage{english}{\textit{:
+age is YOUNG}}}
+
+{\selectlanguage{english}\color{black}
+\foreignlanguage{english}{\textbf{\textit{SubjectAttribute}}}\foreignlanguage{english}{\textit{:
+gender is GIRL}}}
+
+{\selectlanguage{english}\color{black}
 In XML that is:}
 
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}gender{\textgreater}GIRL{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}gender{\textgreater}GIRL{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\color{black}
 That requires the Subject in the RequestItem to possess at least these
@@
 
-{\selectlanguage{english}\color{black}
-\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}age{\textquotedbl}{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}gender{\textquotedbl}{\textgreater}GIRL{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other
-Attribute-{}-{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\itshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
+\ {\textless}RequestItem{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}age{\textquotedbl}{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}gender{\textquotedbl}{\textgreater}GIRL{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}!-{}-Some other Attribute-{}-{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
 {\textless}/RequestItem{\textgreater}}
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 The above example shows that the Subject in the RequestItem
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 If the Subject in the RequestItem is like this:}
 
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}from{\textquotedbl}{\textgreater}OSLO{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other Attribute, buts
-not a {\textquotedbl}gender{\textquotedbl}-{}-{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Alice{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}from{\textquotedbl}{\textgreater}OSLO{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}!-{}-Some other Attribute, buts not a
+{\textquotedbl}gender{\textquotedbl}-{}-{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\color{black}
 Then evaluator will produce INDETERMINATE as the match-making result of
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 If the Subject in the RequestItem is like this:}
 
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}Subject{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Bob{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}Attribute
-AttributeId={\textquotedbl}gender{\textquotedbl}{\textgreater}BOY{\textless}/Attribute{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ \ \ }\textit{{\textless}!-{}-Some other
-Attribute-{}-{\textgreater}}}
-
-{\selectlanguage{english}\color{black}
-\textit{\ \ \ \ }\textit{{\textless}/Subject{\textgreater}}}
-
-{\color{black}
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}Subject{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}name{\textquotedbl}{\textgreater}Bob{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}age{\textgreater}YOUNG{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}Attribute
+AttributeId={\textquotedbl}gender{\textquotedbl}{\textgreater}BOY{\textless}/Attribute{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ {\textless}!-{}-Some other Attribute-{}-{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Subject{\textgreater}}
+
+{\selectlanguage{english}\color{black}
 Then evaluator will give NO\_MATCH as the match-making result of this
@@
 
-{\selectlanguage{english}\upshape\color{black}
+{\selectlanguage{english}\color{black}
 Finally if and only if all of the elemens (Subjects, Actions, Resources,
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 Otherwise if any of the element (Subjects, Actions, Resources,
@@
 
-{\color{black}
+{\selectlanguage{english}\color{black}
 Otherwise the NOT\_APPLICABLE decision will be made for this Rule. In
@@
 method. Currently only supported format is ARC Policy/Request XML
-document described in Section 7.1and 7.2}
+document described in Section \ref{bkm:Ref204009564}and
+\ref{bkm:Ref204009595}}
 
@@
 into existing MCCs or implemented as separate SecHandler plugins. See
-Section 5.2for available Collectors and corresponding Security
-Attributes. }
+Section \ref{bkm:Ref204009636}for available Collectors and
+corresponding Security Attributes. }
 
@@
 \item {\selectlanguage{english}\color{black}
-Delegation PDP is described below in Section 6.3}
+Delegation PDP is described below in Section \ref{bkm:Ref204009695}}
 \end{itemize}
@@
 described except those used for Delegation Restrictions. Those are
-described in Section 6.2}
+described in Section \ref{bkm:Ref204009734}}
 
@@
 certificate in client{\textquotesingle}s chain\\\hline
+\selectlanguage{english}\color{black} SubjectAttribute &
+\selectlanguage{english}\color{black}
+\ http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute &
+\selectlanguage{english}\color{black} VOMS attributes extracted from
+whole client{\textquotesingle}s chain of certificates\\\hline
 \end{supertabular}
 \end{flushleft}
+{\selectlanguage{english}\upshape\color{black}
+The VOMS attributes are presented in format similar to VOMS FQAN with
+slight modifications. Differently from FQAN all values are prepended
+with their names like VO and Group. Missing elements are reported as
+having string value NULL. Example of VOMS attribute looks like }
+
+{\selectlanguage{english}\bfseries\color{black}
+/VO=knowarc.eu/Group=testers/Role=admin/Capability=NULL}
+
+{\selectlanguage{english}\color{black}
+Each set of attributes is acompanied by identifier of service which
+provided those attributes. It is made of \textit{voname} element with
+name of VO followed by optional element \textit{hostname} with hostname
+and port of service. Here is an example}
+
+{\selectlanguage{english}\bfseries\color{black}
+/voname=knowarc.eu/hostname=arthur.hep.lu.se:15001}
+
+{\selectlanguage{english}\color{black}
+If VOMS extensions contain user definable attributes those are presented
+together with the information of their grantor. They consist of
+\textit{voname} and \textit{hostname} elements presented above (if
+\textit{hostname} is missing it is assigned string value NULL) followed
+by user attribute. It{\textquotesingle}s pattern is
+\textit{qualifier:name=value}. Here qualifier acts as namespace and is
+usually same as VO name. Below is an example of such attribute}
+
+{\selectlanguage{english}\bfseries\color{black}
+/voname=knowarc.eu/hostname=arthur.hep.lu.se:15001/knowarc.eu:UniqueKnowarcAttribute=False}
+
 \paragraph{HTTP}
@@
 behalf provides Proxy Certificate to the service using Web Service
-based Delegation interface described in Section 6.4}
+based Delegation interface described in Section \ref{bkm:Ref204009766}}
 
@@
 Each deployment implementing Delegation Restrictions must use dedicated
-Security Handler plugin (see section 5.1) to collect all Policy
-documents from Proxy Certificates used for establishing secure
-connection. Then those documents must be processed by dedicated Policy
-Decision Point plugin (see section 2.3) to make a final decision based
-on collected Policies and various information about
-client{\textquotesingle}s identity and requested operation. Service or
-MCC chain supporting Delegation Restrictions must accept negative
-decision of this PDP as final and do not override it with any other
-decision based on other policies.}
+Security Handler plugin (see section \ref{bkm:Ref204009818}) to collect
+all Policy documents from Proxy Certificates used for establishing
+secure connection. Then those documents must be processed by dedicated
+Policy Decision Point plugin (see section \ref{bkm:Ref204009855}) to
+make a final decision based on collected Policies and various
+information about client{\textquotesingle}s identity and requested
+operation. Service or MCC chain supporting Delegation Restrictions must
+accept negative decision of this PDP as final and do not override it
+with any other decision based on other policies.}
 
@@
 stored inside X509 certificate proxy extension as defined in RFC3820
-and described in Section 6.1All proxy certificates in a chain provided
-by client are examined and all available policies are extracted. }
+and described in Section \ref{bkm:Ref204009914}All proxy certificates
+in a chain provided by client are examined and all available policies
+are extracted. }
 
@@
 {\selectlanguage{english}\upshape\color{black}
-See Section 7.7for the explanation of ARC Policy.}
+See Section \ref{bkm:Ref204009974}for the explanation of ARC Policy.}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
-{\textless}/echo:prefix{\textgreater}}}}
+{\selectlanguage{spanish}\ttfamily\color{black}
+\ \ {\textless}echo:prefix{\textgreater}[
+{\textless}/echo:prefix{\textgreater}}
 
@@
 {\selectlanguage{english}\upshape\color{black}
-There is an example policy for echo service below. See Section 7.1for
-the policy schema. The example policy is \ made of following elements:
-}
+There is an example policy for echo service below. See Section
+\ref{bkm:Ref204009564}for the policy schema. The example policy is
+\ made of following elements: }
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}next
-id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}}}
-
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:prefix{\textgreater}[
-{\textless}/echo:prefix{\textgreater}}}}
+{\selectlanguage{spanish}\ttfamily\color{black}
+\ \ {\textless}next
+id={\textquotedbl}echo{\textquotedbl}/{\textgreater}}
+
+{\selectlanguage{spanish}\ttfamily\color{black}
+\ \ {\textless}echo:prefix{\textgreater}[
+{\textless}/echo:prefix{\textgreater}}
 
@@
 
-{\selectlanguage{english}\upshape\color{black}
-\foreignlanguage{spanish}{\texttt{\ \ }}\foreignlanguage{spanish}{\texttt{{\textless}echo:suffix{\textgreater}
-]{\textless}/echo:suffix{\textgreater}}}}
+{\selectlanguage{spanish}\ttfamily\color{black}
+\ \ {\textless}echo:suffix{\textgreater}
+]{\textless}/echo:suffix{\textgreater}}
 
@@
 
-\subsection[Using Shibboleth IdP for Authentication and Attribute-based
-Authorization]{Using \foreignlanguage{english}{Shibboleth} IdP for
-Authentication and Attribute-based Authorization}
+\subsection[Using Shibboleth IdP for Authentication and
+Attribute{}-based Authorization]{Using
+\foreignlanguage{english}{Shibboleth} IdP for Authentication and
+Attribute-based Authorization}
 {\selectlanguage{english}\color{black}
@@
 
-\liststyleLi
+\liststyleLii
 \begin{enumerate}
@@
 
-\subsection[\ Short-Lived Credential
+\subsection[\ Short{}-Lived Credential
 Service]{\foreignlanguage{english}{\ }Short-Lived Credential Service}
@@
 \par}
+
+\bigskip
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SLCS service is supposed to run together with SP service for
+achieving the SAML2 SSO profile. A typical configuration for SLCS
+service is as follows:}
+
+\begin{center}
+\tablehead{}\begin{supertabular}{|m{5.6886597in}|}
+\hline
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Chain{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}tcp.service{\textquotedbl}
+id={\textquotedbl}tcp{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}tls{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tcp:Listen{\textgreater}{\textless}tcp:Port{\textgreater}60000{\textless}/tcp:Port{\textgreater}{\textless}/tcp:Listen{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}tls.service{\textquotedbl}
+id={\textquotedbl}tls{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}http{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:KeyPath{\textgreater}./testkey-nopass.pem{\textless}/tls:KeyPath{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:CertificatePath{\textgreater}./testcert.pem{\textless}/tls:CertificatePath{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}!-{}-tls:CACertificatePath{\textgreater}./cacert.pem{\textless}/tls:CACertificatePath-{}-{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:ClientAuthn{\textgreater}false{\textless}/tls:ClientAuthn{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}http.service{\textquotedbl}
+id={\textquotedbl}http{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}plexer{\textquotedbl}{\textgreater}POST{\textless}/next{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Plexer
+name={\textquotedbl}plexer.service{\textquotedbl}
+id={\textquotedbl}plexer{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}samlsp{\textquotedbl}{\textgreater}/saml2sp{\textless}/next{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}soap{\textquotedbl}{\textgreater}/slcs{\textless}/next{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Plexer{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Component
+name={\textquotedbl}soap.service{\textquotedbl}
+id={\textquotedbl}soap{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}slcs{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}SecHandler
+name={\textquotedbl}saml2ssoassertionconsumer.handler{\textquotedbl}
+id={\textquotedbl}saml2ssosp{\textquotedbl}
+event={\textquotedbl}incoming{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Component{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Service
+name={\textquotesingle}saml.sp{\textquotesingle}
+id={\textquotesingle}samlsp{\textquotesingle}{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}MetaDataLocation{\textgreater}./test\_metadata.xml{\textless}/MetaDataLocation{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}ServiceProviderName{\textgreater}https://squark.uio.no/shibboleth-sp{\textless}/ServiceProviderName{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Service{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}Service
+name={\textquotedbl}slcs.service{\textquotedbl}
+id={\textquotedbl}slcs{\textquotedbl}{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}next
+id={\textquotedbl}slcs{\textquotedbl}/{\textgreater} }
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CACertificate{\textgreater}./CAcert.pem{\textless}/slcs:CACertificate{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CAKey{\textgreater}./CAkey.pem{\textless}/slcs:CAKey{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}slcs:CASerial{\textgreater}./CAserial{\textless}/slcs:CASerial{\textgreater}
+}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ {\textless}/Service{\textgreater} }
+
+\selectlanguage{english}\itshape\color{black}
+\ \ \ \ {\textless}/Chain{\textgreater} \\\hline
+\end{supertabular}
+\end{center}
+{\selectlanguage{english}\color{black}
+\ \ \ \ As shown in the above table, a plexer dispatches the message
+flow (outgoing from \textit{http.service}) into two SP service
+(\textit{saml.sp}) and SOAP service (\textit{soap.service}). Since the
+client authentication is switched off, it is not necessary to configure
+{\textless}\textit{CACertificatePath}{\textgreater} or
+{\textless}\textit{CACertificateDir}{\textgreater} for TLS MCC.}
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SP service needs to be configured with
+{\textless}\textit{MetaDataLocation}{\textgreater} and
+{\textless}\textit{ServiceProviderName}{\textgreater}. }
+
+{\selectlanguage{english}\color{black}
+\ \ \ \ SLCS service needs to be configured with
+{\textless}\textit{CACertificate}{\textgreater}
+{\textless}\textit{CAKey}{\textgreater} and
+{\textless}\textit{Caserial}{\textgreater}.}
 
@@
 {\selectlanguage{english}\color{black}
-\textstyleInternetlink{\textcolor{black}{\stepcounter{Bibitem}{\theBibitem}}}\textstyleInternetlink{\textcolor{black}{.}}\textstyleInternetlink{\foreignlanguage{english}{\textcolor{black}{Web
-Service Security Username Token Profile 1.1.
-}}}\url{http://www.oasis-open.org/committees/download.php/16782/wss-v1.1-spec-os-UsernameTokenProfile.pdf}}
+\stepcounter{Bibitem}{\theBibitem}.\foreignlanguage{english}{Web Service
+Security Username Token Profile 1.1.
+}\url{http://www.oasis-open.org/committees/download.php/16782/wss-v1.1-spec-os-UsernameTokenProfile.pdf}}
 
@@
 {\selectlanguage{english}\color{black}
-\textstyleInternetlink{\textcolor{black}{\stepcounter{Bibitem}{\theBibitem}}}\textstyleInternetlink{\textcolor{black}{.
-OASIS Web Services Security (WSS) TC.
-}}\url{http://www.oasis-open.org/committees/tc_home.php?wg_abbrev=wss}}
+\stepcounter{Bibitem}{\theBibitem}. OASIS Web Services Security (WSS)
+TC.
+\url{http://www.oasis-open.org/committees/tc_home.php?wg_abbrev=wss}}
 \end{document}

### 2009-01-19 doc/sec/SecurityFrameworkofARC1.tex
Enhancing description of delegation interface, adding diagrams.

@@
 \author{Aleksandr Konstantinov}
-\date{2008-12-23}
+\date{2009-01-19}
 \begin{document}
@@
 \end{flushleft}
-\subsection[Delegation Restrictions]{Delegation Restrictions}
+\subsection[Delegation]{Delegation}
 \label{bkm:Ref204009393}\subsubsection[Delegation
@@
 Delegation interface in the ARC1 is implemented using Web Service
-approach. Each ARC1 service wishing to accept delegated credentials
-implements this interface. Here is how delegation procedure works:}
+approach. Each ARC1 service wishing to act on behalf of client identity
+implements this interface in order to accept delegated credentials.
+Here is how delegation procedure works (also shown in figures
+\ref{seq:refFigure7} and \ref{seq:refFigure8}) :}
 
@@
 \end{itemize}
+{\selectlanguage{english}\color{black}
+Optionally step 2 can be skipped and the DelegatedToken element provided
+to Service as additional payload of other service specific message. }
+
+{\selectlanguage{english}\color{black}
+The Id element obtained in the step 1 can be reused multiple times with
+different content of the Value element.}
+
+{\selectlanguage{english}\color{black}
+WSDL of portType implementing delegation functionality can be found at
+\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/libs/delegation/delegation.wsdl}
+.}
+
+{\centering [Warning: Draw object ignored]\par}
+
+{\centering\selectlanguage{english}\itshape\color{black}
+Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure7}}. The
+flow diaram of delegation procedure with multiple second step and
+session expiration. 
+\par}
+
+{\centering [Warning: Draw object ignored]\par}
+
+{\centering\selectlanguage{english}\itshape\color{black}
+Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure8}}. The
+flow diagram of delegation procedure with certificate transfered as
+payload of service specific message. 
+\par}
+
 \subsection[Schemas, descriptions and examples]{Schemas, descriptions
@@
 
-\begin{center}
+\bigskip
+
+{\centering 
 \includegraphics[width=3.2256in,height=1.8472in]{SecurityFrameworkofARC1-img7.png}
-\end{center}
-
-\bigskip
+\par}
 
 {\centering\selectlanguage{english}\itshape\color{black}
-Figure \stepcounter{Figure}{\theFigure}.\textbf{ }SAML2.0 SSO profile in
-ARC
+Figure
+{\refstepcounter{Figure}\theFigure\label{seq:refFigure5}}.\textbf{
+}SAML2.0 SSO profile in ARC
 \par}
@@
 {\selectlanguage{english}\color{black}
-\ The steps shows in Figure 6 are described as follows:}
+\ The steps shows in Figure \ref{seq:refFigure5} are described as
+follows:}
 
@@
 specific command-line interface (CLI) which includes the user agent and
-WS client. The whole process of SLCS invocation is showed in Figure 7
-(from step 1 to step 8), which is the same as in Figure 6, except that
-step 7 and step 8 are invoked for the SLCS certificate request and
-response.}
+WS client. The whole process of SLCS invocation is showed in Figure
+\ref{seq:refFigure6} (from step 1 to step 8), which is the same as in
+Figure \ref{seq:refFigure5}, except that step 7 and step 8 are invoked
+for the SLCS certificate request and response.}
 
@@
 {\centering\selectlanguage{english}\itshape\color{black}
-Figure \stepcounter{Figure}{\theFigure}.Short lived credential service
+Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure6}}.Short
+lived credential service
 \par}

### 2009-01-30 doc/sec/papers/CCGRID09/Sec_paper.tex
Moving papers to dedicated places

--- 2008-12-06 doc/sec/Sec_paper.tex
@@
 
+{\centering
+Weizhong Qiang 1, Aleksandr Konstantinov 1,2 
+\par}
+
+{\centering
+1University of Oslo, Department of Physics, P.O. Box 1048, Blindern,
+0316 Oslo, Norway
+\par}
+
+{\centering
+2Vilnius University, Institute of Material Science and Applied Research,
+Sauletekio al. 9, Vilnius 2040, Lithuania 
+\par}
+
 \bigskip
@@
 
-\bigskip
+Keyword: single sign-on, delegation, virtual organization, ARC
 
@@
 \end{enumerate}
-\ \ \ \ Grid provides the technology for wide-scale, cross domains
+\ \ \ \ Grid provides the technology for wide-scale, cross-domain
 collaboration where the users work on resources and sharing data which
@@
 system, he/she should own a X.509 certificate which is issued by a CA
-that is trusted by other entity.
+that is trusted by other entities.
 
@@
 make access control decision based on parsing the VO membership
-attributes from the homogeneous attribute certificate.
+attributes from the Attribute Certificate.
 
@@
 \begin{itemize}
-\item The GSS API based confidential communication has also not been
-adopted by Web Service implementation.
+\item The GSS API based confidential communication has not been adopted
+by Web Service implementations.
 \item The Attribute Certificate has not been adopted by any Web Service
@@
 when considering security, which make them impossible to interoperate
-with the GSI based grid middlewares.
+with the GSI based Grid middlewares.
 \item Finally, identity delegation is a good solution for user or
@@
 use their own community credential instead of the X.509 certificate to
-access grid is a promising solution to make the Grid more easily
+access Grids is a promising solution to make the Grids more easily
 accessible.
@@
 
-\bigskip
-
 {\centering
@@
 initially aimed at developing self-organized, fault-tolerant,
-non-intrusive, easy-manageable Grid middleware [10]. \ Now \ the
-classic version of ARC provides Grid services for job submission and
+non-intrusive, easy-manageable Grid middleware [10]. Now the classic
+version of ARC provides Grid services for job submission and
 management, resource characterization, resource aggregation and
@@
 based on the functionality and capabilities of the classic ARC
-middleware. It aims at implementing a Web Service oriented middleware
-which will provide higher levels of resource and user abstraction
-through well-defined web service interface [12] in order to provide
-interoperability with other service-based Grid middlewares, as well as
-other Web Service compatible applications.
+middleware. It aims at implementing a Web Service oriented Grid
+middleware which will provide higher levels of resource and user
+abstraction through well-defined Web Service interface [12] in order to
+provide interoperability with other service-based Grid middlewares, as
+well as other Web Service compatible applications.
 
@@
 \begin{center}
-\includegraphics[width=3.7465in,height=3.3764in]{Secpaper-img1.png}
+\includegraphics[width=3.528in,height=3.098in]{Secpaper-img1.png}
 \end{center}
@@
 message processing fails as well. In Figure 1, the
-{\textquotedblleft}AuthZ{\textquotedblright} and
-{\textquotedblleft}AuthN{\textquotedblright} sub-modules inside MCCs
-and Services are examples of \textit{SecHandler}.
-
-As explained above, HED is supposed to act as a Web/Grid Service
+{\textquotedblleft}\textit{AuthZ}{\textquotedblright} and
+{\textquotedblleft}\textit{AuthN}{\textquotedblright} sub-modules
+inside MCCs and Services are examples of \textit{SecHandler}.
+
+\ \ \ \ As explained above, HED is supposed to act as a Web/Grid Service
 container, but on the client side, from implementation perspective,
@@
 X.509 credential delegation as a normal Web Service, in order to
-provide single sign-on characteristic.
+provide single sign-on ability.
 \item \textit{Message level authentication} -{}-{}- to provide message
 level authentication by implementing WS-Security.
-\item \textit{SAML token delegation service} -{}-{}- to process the WS
-Security SAML token in order to provide single sign-on for SOAP message
-level authentication.
+\item \textit{SAML token delegation service} -{}-{}- to process the
+delegation of WS Security SAML token in order to provide single sign-on
+for SOAP message level authentication.
 \end{itemize}
@@
 using the SAML assertion produced and issued by IdP upon request. In
-order to discover which home community does a user from, Shibboleth
-specifies an optional third component called {\textquotedblleft}Where
-Are You From?{\textquotedblright} (WAYF) service to aid in the process
-of IdP discovery, and this IdP discovery process is also standardized
-and \ defined in SAML 2.0 specification and called as
-{\textquotedblleft}Identity Profile Discovery
+order to discover which home community does a user come from,
+Shibboleth specifies an optional third component called
+{\textquotedblleft}Where Are You From?{\textquotedblright} (WAYF)
+service to aid in the process of IdP discovery, and this IdP discovery
+process is also standardized and \ defined in SAML 2.0 specification
+and called as {\textquotedblleft}Identity Profile Discovery
 Profile{\textquotedblright}.
@@
 {\textless}\textit{samlp:AuthnRequest}{\textgreater} message by using
-its own X.509 certificate (Note in the SAML SSO profile, certificate is
-still needed by IdP and SP) and sends back to user agent.
+its own X.509 certificate (Note in the SAML SSO profile, X.509
+certificates are still needed for IdP and SP) and sends back to user
+agent.
 \item User agent sends the
@@
 authentication mechanism is outside of the SAML2.0 SSO profile.
-Shibboleth IdP implementation chooses some login handler for
+Shibboleth IdP implementation chooses some login handlers for
 authentication. The current user agent implementation is compatible to
@@
 response will be delivered by the user agent to the Service Provider. 
-\item The SP Service verifies and checks the SAML response, and decrypts
-and stores the SAML assertion into session/connection context. The SAML
+\item The SP Service verifies and checks the SAML response, decrypts and
+stores the SAML assertion into session/connection context. The SAML
 assertion includes the
@@
 Moreover, even though the implementation is based on ARC middlerware,
-the idea can be adopted by other grid middlewares if they only require
+the idea can be adopted by other Grid middlewares if they only require
 server authentication instead of mutual authentication.
@@
 \ \ \ \ Since the lifetime of the short lived credential is normally
-short, it is not a must to protect the private key by pass-phrase. As
+short, it is not a must to protect the private key by pass phrase. As
 illustrated in step a and step b in Figure 3, if the private key is not
@@
 certificate to access Grid Service or Web Service from any kind of
-middleware. If the private key is protected, the can use the X.509
+middleware. If the private key is protected, he can use the X.509
 certificate to generate a proxy certificate (by using command-line
 interface utility such as \textit{grid-proxy-init},
-\textit{voms-}\textit{proxy-init}, or \textit{arcproxy}), and then use
-the proxy certificate to access Grid Service or Web Service.
+\textit{voms-}\textit{proxy-init}), and then use the proxy certificate
+to access Grid/Web Service.
 
@@
 access the grid system conveniently; meanwhile, thanks to the single
-sign-on characteristic of SAML2.0 SSO profile which is discussed in
-Section 3.1, this user doesn{\textquotesingle}t not need to input his
-username and password in a valid period after the first time he
-succeeds to authenticate against his home IdP by running the SLCS
-client command on the same node, even if this SLCS client command is
-supposed to run against a few SLCS services to get a few SLCS
-credentials.
+sign-on characteristic of SAML2.0 SSO profile, this user
+doesn{\textquotesingle}t not need to input his username and password in
+a valid period after the first time he succeeds to authenticate against
+his home IdP by running the SLCS client command on the same node, even
+if this SLCS client command is supposed to run against a few SLCS
+services to get a few SLCS credentials.
 
@@
 the new version ARC middleware, we keep the credential delegation idea,
-while utilizing the SOAP communication (and TCP or TLS for transport
-level communication) instead of the GSI communication for the
+while utilizing the SOAP communication (together with TCP or TLS for
+transport level communication) instead of the GSI communication for the
 credential delegation process.
@@
 delegation initiation, and \ the other for storing the signed proxy
-certificate. The delegation process is detailed in Figure 5.
+certificate. The delegation protocol is detailed in Figure 5.
 
@@
 should also use the user{\textquotesingle}s X.509 credential or
-delegated credential (on behalf of the user) for authentication and
-secure communication.
+delegated credential for authentication and secure communication.
 
@@
 {\centering
-Figure 5. Credential delegation process
+Figure 5. Credential delegation protocol
 \par}
@@
 
-\ \ \ \ WS-Security related specifications are implemented in ARC as
-\textit{SecHandler}. There are three kinds of WS-Security version 1.1
-specifications which are completely or partly implemented: Username
-Token profile, X.509 Token profile, and SAML Token profile.
+\ \ \ \ WS-Security related specifications are implemented as
+\textit{SecHandler} in ARC. There are three kinds of WS-Security
+version 1.1 specifications which are completely or partly implemented:
+Username Token profile, X.509 Token profile, and SAML Token profile.
 
@@
 instead, the delegation client will pick the public key from peer
-certificate after successful SSL/TLS authentication. So \ step2 of the
-delegation protocol will only be some confirmation information which
-lets delegation client to proceed. 
+certificate after successful SSL/TLS authentication. So step2 of the
+delegation protocol is only be some confirmation information which
+informs delegation client to proceed. 
 \item For those multiple intermediate services through which a SOAP
@@
 SOAP message on the left side is the one which B sends to relying party
-(intermediate service), B is the identity of the attesting
+(intermediate service). B is the identity of the attesting
 entity{\textquotesingle}s X.509 token, while A is the identity of
@@
 
-\ \ \ It is worth mentioning is: since the SLCS X.509 credential or
+\ \ \ What is worth mentioning is: since the SLCS X.509 credential or
 delegated credential can be used to issue X.509 token, \ the X.509
@@
 {\centering
-Figure 6. SAML Token delegation service
+Figure 6. SAML Token delegation service and client
 \par}
@@
 
-Similar work that integrates Shibboleth and Grid middleware does exist.
-GridShib[17][18] uses Shibboleth to get entity{\textquotesingle}s
-attribute for the authorization of services developed in Globus Tookit.
-Unlike the work of this paper, GridShib does not use
-Shibboleth{\textquotesingle}s native login handlers for authentication
-to Shibboleth IdP, instead, a new login handler specific for X.509
-based authentication is developed for Shibboleth IdP, and a credential
-service is developed for acting as an online-CA, authenticating the
-user, and issuing short-lived X.509 credential. Moreover, this paper
-provides a single sign-on solution for user authenticating to Grid
-services by directly using SAML2.0 SSO profile (section 3.1).
-
-SWITCH short-lived credential service (SLCS) [19] provides a SLCS
-service and client. The benefit of our work comparing to the
+\ \ \ \ Similar work that integrates Shibboleth and Grid middleware does
+exist. GridShib[17][18] uses Shibboleth to get
+entity{\textquotesingle}s attribute for the authorization of services
+developed in Globus Tookit. Unlike the work of this paper, GridShib
+does not use Shibboleth{\textquotesingle}s native login handlers for
+authentication to Shibboleth IdP, instead, a new login handler specific
+for X.509 based authentication is developed for Shibboleth IdP, and a
+credential service is developed for acting as an online-CA,
+authenticating the user, and issuing short-lived X.509 credential.
+Moreover, this paper provides a single sign-on solution for user with
+community credential authenticating to Grid services by directly using
+SAML2.0 SSO profile.
+
+\ \ \ \ SWITCH short-lived credential service (SLCS) [19] provides a
+SLCS service and client. The benefit of our work comparing to the
 SWTCH{\textquotesingle}s solution is the SLCS service and client is
@@
 
-In terms of delegation, apart from the credential delegation mechanism
-from GSI [2][20], [21] proposes a delegation protocol based on the
-WS-Trust [22] in order to provide a standard and interoperable protocol
-for delegation in Grids. WS-Trust will also be adopted by ARC
+\ \ \ \ In terms of delegation, apart from the credential delegation
+mechanism from GSI [2][20], [21] proposes a delegation protocol based
+on the WS-Trust [22] in order to provide a standard and interoperable
+protocol for delegation in Grids. WS-Trust will also be adopted by ARC
 middleware to express the specifications required to define delegation
-protocol in a standard way. Gridsite implements a X.509 credential
-delegation solution [23] based on Web Service, with which ARC
-delegation client can easily interoperate.
-
-GSI plug-in for gSOAP[24] adds GSI support for gSOAP Web Service tookit.
-The GSI MCC implemented in ARC can also achieve the same functionality
-as this plug-in. Moreover, the ARC solution is more flexible than it,
-because GSI MCC of both service and client sides are pluggable, which
-means service and client developer of ARC do not need to modify the
-service or client implementation in order to add or replace the GSI
-functionality.
+protocol in a standard way. Gridsite project implements a X.509
+credential delegation solution [23] based on Web Service, with which
+ARC delegation client can easily interoperate.
+
+\ \ \ \ GSI plug-in for gSOAP[24] adds GSI support for gSOAP Web Service
+tookit. The GSI MCC implemented in ARC can also achieve the same
+functionality as this plug-in. Moreover, the ARC solution is more
+flexible than it, because GSI MCC of both service and client sides are
+pluggable, which means service and client developer of ARC do not need
+to modify the service or client implementation in order to add or
+replace the GSI functionality.
 
@@
 
-\ \ \ It must be noted that due to standard specifications utilized
+\ \ \ \ It must be noted that due to standard specifications utilized
 (SAML and WS-Security) which are independent from any particular
@@
 
-\ \ \ \ Since authorization is also a key problem for VO, as the future
-work, we will implement a solution for cross-middleware authorization
-by using the standard specification such as SAML.
+\ \ \ \ Though only authentication issue is discussed in this paper, as
+the future work, we will implement a solution for cross-middleware
+attribute-based authorization by using the standard specification such
+as SAML.
+
+{\centering
+6. Acknowledgment
+\par}
+
+We would thank all of the ARC middleware developers for their
+intelligent work. The work has been supported by the European
+Commission through the KnowARC project (contract Nr. 032691).
 

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
tex source for ISC; format should be two-column

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Towards cross-middleware authentication and single sign-on for ARC Grid middleware
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Weizhong Qiang         \and
        Aleksandr Konstantinov %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
When pursuing the task of making access to Grids as simple as possible, security is one of the most important challenges in production Grid infrastructures, especially when the Grid applications span multiple administrative domains as well as heterogeneous Grid middlewares. A typical example is wide scale e-Science applications which need to coordinate resources shared among a number of independent institutions with different Grid middlewares deployed on these resources. In this paper, we describe security implementation and considerations used in the upcoming version of the Advanced Resource Connector (ARC) middleware, where the heterogeneity issue has been addressed. The main goal of ARC implementation in terms of security is to let the middleware be capable of interoperating with other Grid middlewares by leveraging on standard specifications. The key aspect of the work is to enhance the current proxy certificate based authentication and single sign-on by utilizing and enhancing the standardized Web Service specifications such as Security Assertion Markup Languages (SAML), single sign-on (SSO) profile and Web Services Security in order to achieve cross-middleware authentication and single sign-on.
\keywords{single sign-on \and delegation \and virtual organization \and ARC middleware }
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction} 
\label{sec:intro}
Grid provides the technology for wide-scale, cross-domain collaboration where users work with resources and share data which are distributed across administrative domains. This kind of collaboration is called Virtual Organization (VO)\cite{IFoster01}. The resources that contribute to the VO can come from independent administrative domains, and even require different security approaches to access control due to usage of different underlying middlewares. The same heterogeneous situation applies to users who need to access the VO resources, because users holding one kind of a credential may need to utilize multiple resources with different security approaches. To solve the above issues, Grid middlewares should provide the following abilities:
The user which participates a VO should be able to access data and resources without being required to constantly do authentication at each resource or service;
The VO should be able to easily manage the privileges and memberships of the users;
The VO together with its resources should be able to easily manage the authorization, given the condition that the resources are owned by different administrative domains.

In the current Grid community, the Grid Security Infrastructure (GSI)\cite{IFoster98} is the de facto standard for authentication and transport level security. It builds on the Public Key Infrastructure (PKI), an architecture based on the X.509 public key certificate. GSI implements some enhancement (such as certificate delegation) based on the standard SSL/TLS protocol. Mutual authentication is required by GSI, and is the default configuration for GSI based Grid deployment. The X.509 certificates are required for both the client and service sides in order to achieve mutual authentication. The X.509 certificates are issued by trusted third parties called certificate authorities (CA), and the CAs constitute trust federation which guarantees that two different X.509 certificates from different CAs can accomplish authentication against each other. Thus in the GSI based Grid system, if a user would like to access a Grid system, he/she should own a X.509 certificate which is issued by a CA that is trusted by other entities.

In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS)\cite{AlfieriR05} have partly solved them, to a different extent. The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided by creating the proxy certificate.  As the most successful Grid authorization model, VOMS adopts the Attribute Certificate\cite{RFC3821link} approach. The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.Attributes include the VO membership data which is associated with the proxy certificate’s identity. Eventually the resources can make access control decisions based on parsing the VO membership attributes from the Attribute Certificate.

The VO is essentially supposed to be able to provide resource sharing capability with resources running under heterogeneous systems/middlewares. This goal can be achieved using Web Service Architecture (WSA). This way, the Web Service technology has been utilized in Grid computing middlewares. The Web Service technology has been adopted by the Globus Toolkit\cite{GTlink}, gLite\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)\cite{ARClink}. In fact, utilizing Web Service technology and providing seamless interoperability with other middlewares is one of the main goals of the new version of ARC middleware.

Unfortunately, the existing Grid security solutions such as GSI and VOMS do not apply to Web Service based applications very well due to the following reasons:
The GSS API based confidential communication has not been adopted by Web Service implementations.
The Attribute Certificate has not been adopted by any Web Service implementation at all.
Web Service implementations commonly use the Web Services Security (WS-Security)\cite{WSSeclink}, Security Assertion Markup Languages (SAML)\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
Finally, identity delegation is a good solution for user or user’s representative (acting on behalf of the user) to move/travel from one resource to another. But since the identity delegation has been completely coupled with the GSI confidential communication process, it is not possible for the existing identity delegation solution to be used by Web Service applications.

Moreover, the Grid security solutions require each user to possess a X.509 certificate. However, the process to get a certificate may take quite long time, since the requester needs to be checked and approved by the Registration Authority (RA) and only then the Certificate Authority (CA) can issue the certificate according to the approval from the RA. Meanwhile, the process to use the certificate is also not so easy for the non-IT advanced research community to deal with, since the user has to generate proxy certificates by using Grid-specific command lines, such as grid-proxy-init or voms-proxy-init. On the other hand, users could often have some more familiar local community or institutional credentials such as username/password. Enabling users to use their own community credentials instead of the X.509 certificate to access Grids is a promising solution to make the Grids more easily accessible.

The above analysis shows that in order to achieve interoperability and transparency for user to access Grid system which is established on different middlewares, we need to address the mismatch between the different security solutions from these middlewares in order to ensure the interoperability as wide as possible; it is also necessary to address the mismatch between users’ daily experience and middleware’s requirements in order to ensure the access as simple as possible.

The approach in the new version of ARC is to utilize the existing Web Service standards, such as SAML and WS-Security, to achieve the interoperability and accessibility without breaking existing Grid specific security protocols.

The rest of this paper is organized as follows: Section II presents the ARC Grid middleware, especially the architecture of the new version of ARC. Section III describes the solution for cross-middleware authentication and single sign-on, including implementations for bridging the mismatch in terms of authentication and single sign-on. Section IV discusses the related research, and Section V presents conclusions and future work outlook.

\section{ARC Grid middleware}
\label{sec:arcmiddleware}
ARC (Advanced Resource Connector) is an open source Grid middleware solution released under Apache license. ARC middleware aims at developing self-organized, fault-tolerant, non-intrusive, easy-manageable Grid middleware\cite{MEllert07}. The current production version of ARC provides Grid services for job submission and management, resource characterization, resource aggregation and discovery, basic data management, integration of Grid security solution, and so on. This release has been deployed and used in production environment, and is one of the widely deployed Grid middlewares in Europe.

The next generation of ARC components is developed by the KnowARC project\cite{KnowARClink}, based on the functionality and capabilities of the classic ARC middleware. It aims at implementing a Web Service oriented Grid middleware which will provide higher levels of resource and user abstraction through well-defined Web Service interface\cite{KnowARCDesignlink} in order to provide interoperability with other service-based Grid middlewares, as well as other Web Service compatible applications.

As the key part of the implementation of new ARC middleware, there is a lightweight Web Service container called Hosting Environment Daemon (HED) which provides hosting for various services at application level, as well as a number of modules to support flexible, interoperable, and efficient communication mechanism for building SOAP based Web Services. The whole design of the HED is built around the idea of flexibility and modularity, such that the application service developers can simply concentrate on the application level Web Service implementation by only using the core minimum amount of components. It also simplifies work on the middleware level, e.g. making possible to implement another communication protocol or authentication mechanism by using the core minimum amount of components and external dependencies. A system administrator can easily configure and deploy the middleware and application for different kinds of requirements without having to know much about the implementation.

The architecture of the HED is illustrated in Figure \ref{fig:HED}. In general, there are few components called Message Chain Component (MCC) which are in charge of implementing different protocol levels. For instance, as shown in the example message flow, the HTTP MCC will process stream from the TLS MCC to parse the HTTP message and pass its body to the SOAP MCC, and also process the SOAP response from the SOAP MCC to generate the HTTP message for the TLS MCC.

Dotted line in Figure \ref{fig:HED} shows an alternative path for the information to propagate among MCCs. A service administrator can configure the MCCs according to the interoperability requirements with the other part. For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE’s SoapSender and SoapReceiver)\cite{WSElink}. Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)\cite{AShoshani03} service. This shows the flexibility of HED in terms of protocols support.

\begin{figure}
\includegraphics[width=0.75\textwidth]{HED.eps}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}
HED contains a framework for implementing and enforcing authentication and authorization. Each MCC or Service has a common interface for implementing various authentication and authorization functionality. Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service. Each MCC or Service is usually configured with two queues of SecHandler – one for the incoming message and one for the outgoing message respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well. In Figure \ref{fig:HED}, the “AuthZ” and “AuthN” sub-modules inside MCCs and Services are examples of SecHandler.

As explained above, HED is supposed to act as a Web/Grid Service container, but on the client side, from implementation perspective, there is a similar architecture implemented for processing messages from different protocols and handling security functionality; also there is a specific application programming interface (API) for developers to easily write Web Service client code.

\section{Cross-Middleware Authentication and Single Sign-On}
\label{sec:siglesignon}
We investigated the authentication and single sign-on issue in the scenario when a user or user’s representative (acting on behalf of this user) needs to access resources across different middlewares. Our work focuses on the following five major aspects:
Integration of community authentication – to utilize the standard-compatible community authentication mechanism from the Shibboleth\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
Short lived credentials service – to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
Credential delegation service – to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
Message level authentication – to provide message level authentication by implementing WS-Security.
SAML token based delegation service – to process the delegation of an WS Security SAML token in order to provide single sign-on for SOAP message level authentication.

\subsection{Community based authentication for Grid by utilizing Shibboleth Identity Provider}
\label{sec:communityauthn}
AAI (Authentication and Authorization Infrastructure) is a solution for the authentication and authorization in cross-domain resource sharing, such as electronic resource sharing between digital libraries, etc. AAI implicitly applies to the community or institutional based authentication where users from the various communities need to get resources from other communities by using some federation mechanism. Unlike the X.509 based authentication solution in some Grid systems, the AAI does not require users to provide the X.509 certificate, instead, it supports different types of authentication, such as username/password authentication, IP address authentication, etc. There are several implementations of the AAI, among which Shibboleth is the one which has been widely deployed.

The Shibboleth provides cross-domain single sign-on and attribute-based authorization while preserving user privacy. It is based on the OASIS Security Assertion Markup Language (SAML), specifically, the new version of Shibboleth supports SAML 2.0 specification. For the authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines two functional components, an Identity Provider and a Service Provider. The Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity, while the Service Provider (SP) is responsible for controlling access to services and resources by analyzing the SAML assertions produced and issued by the IdP upon request. In order to discover which home community does a user come from, Shibboleth specifies an optional third component called “Where Are You From?” (WAYF) service to aid in the process of IdP discovery. This IdP discovery process is also standardized and defined in SAML 2.0 specification as “Identity Provider Discovery Profile”.

We also utilized the SAML2.0 web browser SSO profile for the authentication in ARC middleware. On the client side the user agent functionality of the Web browser was implemented in order to mimic its behavior, such as the HTTP redirection and the Cookie processing. In fact, implementation of the user agent is also based on the client interface of ARC, specifically, the HTTPS client interface, since the client interface of ARC can support different protocols which are incarnated by different MCCs. On the service side, we implemented the Service Provider functionality (based on the HTTP MCC chained to TLS MCC). For Identity Provider, the Shibboleth IdP implementation is used. Figure \ref{fig:SAML2SSO} shows the process of SAML2.0 SSO integrated in ARC client and service.

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAML2SSO.eps}
\caption{SAML2.0 SSO profile in ARC}
\label{fig:SAML2SSO}
\end{figure}
Here is the explanation of steps shown in Figure \ref{fig:SAML2SSO}:
1.The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side. The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is “saml2sp” which is specific for pointing to the SP Service. Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
2.The SP Service searches the metadata (we use the same metadata format as defined in the Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself). Then SP Service uses obtained information to compose and issue the SAML authentication request also by utilizing own X.509 certificate (note that in the SAML2.0 SSO profile, X.509 certificates are still needed for IdP and SP) and sends the response back to the user agent.
3.User agent sends the SAML authentication request to the Identity Provider.
4.Identity Provider requires an act of authentication. The authentication mechanism is outside of the SAML2.0 SSO profile. The Shibboleth IdP implementation uses login handlers for authentication. The current user agent implementation is compatible with the Username/Password login handler of Shibboleth IdP. Through the HTTP protocol, the user agent will feed IdP with the username/password which has been given by the caller of user agent interface.
5.Once the authentication has been succeeded, the IdP issues a SAML response including SAML assertion encrypted by destination SP’s public key, and then this SAML response will be delivered by the user agent to the Service Provider.
6.The SP Service verifies and checks the SAML response, decrypts and stores the SAML assertion into session/connection context. That assertion includes the AuthnStatement and AttributeStatement elements.
7.The WS client launches the Grid/Web Service request via the same connection as the one which is used by user agent to contact SP Service.
8.The Grid/Web Service checks the AuthnStatement stored in the session context to see if the session is still valid. If valid, service handles the  request and returns the response to the WS client. Note that in the current implementation the service requires WS client to reuse same TCP connection as the one used by user agent is step 5, in order to guarantee  validity of the SSO result.

The SP service and other functional service(s) are hosted by the same container, and they use the same X.509 credential for the transport layer security. The client authentication may be switched off, so that client doesn’t need to use any X.509 credential. Only the trusted CAs of SP and IdP services need to be configured on the client side to ensure SP and IdP can authentication.

The benefit of the SAML2.0 SSO profile worth mentioning is that the Identity Provider can cache the authentication result through session management once the user agent has succeeded to authenticate. Then for a short period this authentication result is valid so that the user agent doesn’t need to feed IdP with user’s username and password the next time it authenticates against IdP. So the user (or the client on behalf of that user) can travel across multiple security domains with only providing his name and password once, which is the characteristic of single sign-on.

Since the Shibboleth implementation of SAML is standard-compliant and widely deployed, the solution implemented in ARC can interoperate with other SAML implementations with minimum change, and more importantly, this solution can utilize the widely deployed SAML based authentication implementation in Grid systems avoiding the use of X.509 certificate on the client side.

\subsection{Short-Lived Credential Service}
\label{sec:slcs}
However, many widely used Grid middlewares are based on GSI while GSI requires mutual X.509 authentication. Also for Web Service based Grid solutions, we cannot prevent service side from requiring client's X.509 certificate.  Based on the solution described in previous section, we implemented a Short Lived Credential Service (SLCS) through which user can get a short-lived X.509 certificate without being bothered to contact any registration authority (RA) or certificate authority (CA).

\begin{figure}
\includegraphics[width=0.75\textwidth]{SLCS.eps}
\caption{Short lived credential service}
\label{fig:SLCS}
\end{figure}
The SLCS service is also a Web Service implemented using ARC service interface. The SLCS client is implemented as command-line interface (CLI) utility which implements the user agent and WS client functionalities. The whole process of SLCS invocation is shown in Figure \ref{fig:SLCS} (steps 1‑8). The process resembles one shown in in Figure \ref{fig:SAML2SSO}, except that steps 7 and 8 use SLCS certificate request and response.

The SLCS client sends request to the SLCS service with the generated X.509 certificate request embedded. Upon reception of the request the SLCS service  composes a distinguished name (DN), andissues a certificate (short lived, 12 hours by default) with the SAML attribute (from the SAML2.0 SSO profile) as the X.509 certificate extension and puts the certificate in to the Web Service response. The SLCS client gets the response and stores the X.509 certificate into local repository along with corresponding private key.

Since the lifetime of the short lived credential is usually short, it is not a must to protect the private key by a pass phrase. It is considered enough to use protection  provided by underlying file system. If the private key is not protected by passphrase then it can be used direcctly for steps (a) and (b) in Figure \ref{fig:SLCS} to access Grid Service or Web Service. If the private key is protected, the user can generate temporary  X.509 proxy by using a command-line interface utility such as grid-proxy-init, voms-proxy-init, etc. Then this proxy can by used  to access the Grid/Web Service.

It should be noticed that the process of composing the distinguished name (DN) for the certificate is a critical issue for the SLCS service. Since the Shibboleth Identity Provider uses the eduPerson schema\cite{eduSchemalink} for the composing the SAML Attribute we pick the relatively distinguishable attribute “eduPersonPrincipalName” for the DN. A typical example of the eduPersonPrincipalName  could be alice@example.org. In such case composed DN is for example “/O=knowarc/OU=example.org/CN=alice”.

The obvious benefit of the SLCS service is that the user can get the X.509 credentials anywhere simply by running the SLCS client command and providing her username and password, and then access the X.509 based Grid infrastructure conveniently. Thanks to the single sign-on capability of SAML2.0 SSO profile, the user doesn’t not need to provide her username and password during valid period after the first time she succeeds to authenticate against her home IdP.

\subsection{X.509 Credential Delegation Service}
\label{sec:creddeleg}
The delegation of credentials proposed in the Grid Security Infrastructure (GSI) is a good solution for supporting single sign-on for users, except that the coupling of the GSI communication is not so widely accepted in Web Service applications. In the design of the new ARC middleware components, we keep the credential delegation idea, while utilizing the SOAP communication (together with TCP or TLS for transport level communication) instead of the GSI communication for the credential delegation process.

As shown in Figure \ref{fig:Delegation}, there is a specific WS client and Web Service for processing delegation: delegation client, and delegation service. The WSDL (Web Service Description Language) of the delegation service has two main operations, one for processing the delegation initiation, and the other for storing the signed proxy certificate. The delegation protocol is detailed in Figure \ref{fig:Delegation_protocol}.

\begin{figure}
\includegraphics[width=0.75\textwidth]{Delegation.eps}
\caption{Credential delegation service and client}
\label{fig:Delegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\textwidth]{Delegation_protocol.eps}
\caption{Credential delegation protocol}
\label{fig:Delegation_protocol}
\end{figure}
For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure \ref{fig:Delegation}), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential. For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure \ref{fig:Delegation}), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.

Since each WS client accesses the target service (delegation service or other general services) on behalf of the user, the trust relationship between the user’s certificates (X.509 certificate and proxy certificate) and services is required, while the trust relationship between service’s certificate and another service’s certificate is not required.

What should be mentioned is that the delegation service is also a normal Web Service. Therefore, delegation service implemented in ARC can be accessed by any standard-compliant WS client that intends to delegate credential; also delegation client implemented in ARC can access any other standard-compliant Web Service which accepts delegation request.

\subsection{Message Level Authentication}
\label{sec:msgauthn}
Transport level security (TLS or SSL) is not sufficient for securing Web Service, because the secure channel created by SSL/TLS makes it impossible to differentially protect the SOAP message, e.g. to encrypt or sign only particular components of the SOAP message, which is relevant when non-sensitive portions of the message need to be accessed or changed by intermediate actors. Additionally, SSL/TLS is incapable of providing end-to-end protection for SOAP message which might flow through multiple intermediate actors, because it only allows each hop to be protected with the resultant security gaps at intermediate actors. Web Services Security (WS-Security)\cite{WSSeclink} specification provides means for applying security to Web Service.

WS-Security related specifications are implemented as SecHandler in ARC. There are three kinds of WS-Security version 1.1 specifications which are completely or partly implemented: Username Token profile, X.509 Token profile, and SAML Token profile.

For those WS clients on behalf of a user, at any step of service invocation, the WS client can augment the SOAP message with WS-Security token by simply configuring the WS-Security SecHandler into container’s configuration, if the relying party (target service) requires.

If the WS client owns an X.509 certificate or proxy certificate, the process of generating an X.509 Token for a SOAP message is: this client puts the content of certificate into X.509 token as <wsse:BinarySecurityToken>, and uses the private key (corresponding to the certificate) to create XML signature for the SOAP message body.

However the process of generating a SAML Token is more complicated. In the WS-Security SAML Token profile, ConfirmationMethod is the verifying method for establishing proof-of-possession for the subject. Two methods are proposed in the SAML Token profile: holder-of-key and sender-vouches. We rely on the holder-of-key method. For the holder-of-key subject confirmation method, there are three interaction parts: the attesting entity, the relying party and the issuing authority. Firstly, the attesting entity authenticates to issuing authority by using some authentication scheme such as X.509 Token profile, or SSL/TLS mutual authentication. Then the issuing authority is able to make a definitive statement (sign a SAML assertion) about an act of authentication that has already taken place. The attesting entity gets the SAML assertion and then signs the SOAP message together with the assertion by using its private key (the relevant certificate has been authenticated by issuing authority, and its relevant public key has been put into the SubjectConfirmation element under SAML assertion by issuing authority). Only the exact entity that possesses the private key which is paired with the public key in the SAML assertion can sign the SOAP message, which establishes an irrefutable connection between the author of the SOAP message and the assertion describing an authentication event.

The relying party is supposed to trust the issuing authority. When it receives a message from the asserting entity, it will check the SAML assertion based on its predetermined trust relationship with the SAML issuing authority, and check the signature of the SOAP message based on the public key in the SAML assertion without directly trust relationship with attesting entity (holder of this public key).
In terms of verification, since the proxy certificates have been introduced in IETF (RFC 3820)\cite{RFC3820link}, and OpenSSL has also implemented RFC3820, it is not difficult to use the proxy certificate for the X.509 Token profile to issue an X.509 token.

\subsection{SAML Token Delegation Service}
\label{sec:samldeleg}
There is also a delegation requirement on the message level: some entities often need to temporarily delegate themselves to another entity in order to perform actions on their behalf. We propose a delegation mechanism for the SAML Token. Figure \ref{fig:SAMLDelegation} shows the parties of the SAML Token delegation.

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAMLDelegation.eps}
\caption{SAML Token delegation service and client}
\label{fig:SAMLDelegation}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\textwidth]{SAMLDelegationMessage.eps}
\caption{SOAP message header with SAML delegation token inside}
\label{fig:SAMLDelegationMessage}
\end{figure}
The delegation process is similar to the process in the X.509 credential delegation service, with the following differences:
1.Transport level communication between delegation client and service is not based on mutual authentication between the proxy certificate and the service certificate; instead, it is based on mutual authentication between the service certificates.
2.Delegation service will not generate any certificate request; instead, the delegation client will pick the public key from the peer certificate after a successful SSL/TLS authentication. So step 2 of the delegation protocol is just some confirmation information which informs delegation client to proceed.
3.For those multiple intermediate services through which a SOAP message flows, delegation service does not need to appear, instead, it only needs to appear on the end of the SOAP message flow.
4.SAML Token might include one or a few <saml:Attribute> attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.

Figure \ref{fig:SAMLDelegationMessage} shows the SAML Token (SAML Assertion) inside the SOAP message is changed due to delegation (we suppose X.509 token is used for attesting entity to authenticate against issuing authority). The SOAP message on the left side is the one which B sends to the relying party (intermediate service). B is the identity of the attesting entity’s X.509 token, while A is the identity of issuing authority, and the SOAP message is signed by B’s private key which is corresponding to the X.509 certificate that is used to create the X.509 token. On the right side, C is the identity of the delegation service (also C is the identity of the target service to which B will access), and now the SOAP message is signed by C’s private key which is corresponding to the target service’s certificate.

What is worth mentioning is: since the SLCS X.509 credential or the delegated credential can be used to issue an X.509 token,  the X.509 credential delegation and SAML Token delegation are smoothly bridged, which benefits the single sign-on.

\section{Related Work}
\label{sec:relatedwork}
Similar work that integrates Shibboleth and Grid middleware does exist. GridShib\cite{VWelch05},\cite{TBarton06} uses Shibboleth to get entity’s attribute for the authorization of services developed in Globus Tookit. Unlike the work described in this paper, the GridShib does not use Shibboleth’s native login handlers for authentication to Shibboleth IdP, instead, a new login handler specific for X.509 based authentication is developed for the Shibboleth IdP, and a credential service is developed for acting as an online-CA, authenticating the user, and issuing short-lived X.509 credential. Moreover, this paper provides a single sign-on solution for the user with community credentials authenticating to Grid services by directly using SAML2.0 SSO profile.

SWITCH (The Swiss Education and Research Network) short-lived credential service (SLCS)\cite{switchslcslink} provides a SLCS service and a client. The advantage of our work comparing to the SWTCH’s solution is that the SLCS service and client are based on Web Services, so that they can be implemented in any Web Service container while keeping interoperability.

In terms of delegation, apart from the credential delegation mechanism of GSI\cite{IFoster98},{VWelch04}, several authors\cite{MAhsant04} propose a delegation protocol based on the WS-Trust\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way. Gridsite project implements a X.509 credential delegation solution\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.

GSI plug-in for gSOAP\cite{GAloisio05} adds GSI support for gSOAP Web Service toolkit. The GSI MCC implemented in ARC provides samilar functionality. Moreover, the ARC solution is more flexible because GSI MCC of both service and client sides is pluggable, which means service and client developers of the ARC do not need to modify the service or client implementation in order to add or replace the GSI functionality.

\section{Conclusion And Future Work}
\label{sec:conclusion}
In this paper we presented the solution built into the ARC Grid middleware, aiming towards the cross-middleware authentication and single sign-on. The goal of our work is to develop an interoperable and accessible single sign-on solution to be used for a VO scenario in production Grids, where users can easily authenticate against the Grids with any of their credentials (not only X.509) from anywhere, and move inside Grid middleware infrastructure and even across heterogeneous Grids  boundaries, without re-authentication within a reasonable period. In order to do so, we implemented some standards based Web Services (including the SLCS service, the X.509 credential delegation service, and the SAML Token delegation service), as well as the user agent and the Service Provider in the context of the SAML2 SSO profile.

It must be noted that due to the standard specifications utilized (SAML and WS-Security) which are independent from any particular middleware, the WS clients and Web/Grid Services developed within the ARC middleware can interoperate with WS clients and Web/Grid Services developed as part of some other middlewares in a standardized way. ARC also supports interoperability with Grid middlewares that require GSI based communication.

Though only authentication issue is discussed in this paper, as the future work, we will implement a solution for cross-middleware attribute-based authorization by using the standard specifications such as SAML.

% Text with citations \cite{RefB} and \cite{RefJ}.
%\subsection{Subsection title}
%\label{sec:2}
%as required. Don't forget to give each section
%and subsection a unique label (see Sect.~\ref{sec:1}).
%\paragraph{Paragraph headings} Use paragraph headings as needed.

% For one-column wide figures use
\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:1}       % Give a unique label
\end{figure}
%
% For two-column wide figures use
\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[width=0.75\textwidth]{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:2}       % Give a unique label
\end{figure*}
%
% For tables use
\begin{table}
% table caption is above the table
\caption{Please write your table caption here}
\label{tab:1}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lll}
\hline\noalign{\smallskip}
first & second & third  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
number & number & number \\
number & number & number \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

\begin{acknowledgements}
We would thank all of the ARC middleware developers for their work. We wish to thank Oxana Smirnova from Lund University for vital comments and proof-reading.
\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
\begin{thebibliography}{}
%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
%\bibitem{RefJ}
% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
% etc

\bibitem{IFoster01}
Foster I, Kesselman C, Tuecke S. The anatomy of the Grid: Enabling scalable virtual organizations, International Journal of Supercomputer Applications, 15(3), 200–222 (2001)

\bibitem{IFoster98}
Foster, I., Kesselman, C., Tsudik, G. and Tuecke, S. A Security Architecture for Computational Grids, ACM Conference on Computers and Security, 1998, 83-91.

\bibitem{AlfieriR05}
R. Alfieri, R. Cecchini, V. Ciaschini, L. dell’Agnello, A. Frohner, K. Lorentey, and F. Spataro. From gridmap-file to voms: managing authorization in a Grid environment. Future Generation Comp. Syst., 21(4), 549–558 (2005)

\bibitem{RFC3821link}
RFC 3821- An Internet Attribute Certificate Profile for Authorization. http://www.faqs.org/rfcs/rfc3281.html

\bibitem{GTlink}
Globus Toolkit. http://www.globus.org/toolkit/

\bibitem{gLitelink}
gLite: lightweight middleware for Grid computing. http://glite.web.cern.ch

\bibitem{ARClink}
Advanced Resource Connector. http://www.nordugrid.org/middleware/

\bibitem{WSSeclink}
OASIS Web Services Security. www.oasis-open.org/committees/wss/

\bibitem{SAMLlink}
OASIS Security Assertion Markup Languages (SAML). www.oasis-open.org/committees/security/

\bibitem{MEllert07}
M. Ellert et al. Advanced resource connector middleware for lightweight computational grids, Future Generation computer systems, 23(2), 219–240 (2007)

\bibitem{KnowARClink}
KnowARC project.  https://www.knowarc.eu/

\bibitem{KnowARCDesignlink}
Design document of new version ARC. https://www.knowarc.eu/documents/Knowarc_D1.1‑1_07.pdf

\bibitem{WSElink}
Web Services Enhancements 2.0 Service Pack 2. http://msdn.microsoft.com/en-us/library/

\bibitem{AShoshani03}
A. Shoshani, A. Sim, and J. Gu, Storage Resource Managers: Essential Components for the Grid, Grid Resource Management: State of the Art and Future Trends, Kluwer Publishing, 2003.

\bibitem{Shiblink}
The Shibboleth Project. http://shibboleth.internet2.edu/

\bibitem{eduSchemalink}
eduPerson and eduOrg Object shema. http://middleware.internet2.edu/eduperson/

\bibitem{RFC3820link}
Internet X.509 Public Key Infrastructure (PKI) Proxy Certificate Profile. ttp://www.ietf.org/rfc/rfc3820.txt

\bibitem{VWelch05}
V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. 4th Annual PKI R&D Workshop, 2005.

\bibitem{TBarton06}
T. Barton et al. Identity Federation and Attribute-based Authorization through the Globus Toolkit, Shibboleth, Gridshib, and MyProxy. 5th Annual PKI R&D Workshop, 2006.

\bibitem{switchslcslink}
SWITCH Short Lived Credential Service. http://www.switch.ch/Grid/slcs/

\bibitem{VWelch04}
V. Welch et al. X.509 proxy certificate for dynamic delegation, Proceeding of the 3rd Annual KI R&D Workshop, 2004.

\bibitem{MAhsant04}
M. Ahsant, J. Basney, and O. Mulmo. Grid Delegation Protocol, UK Workshop on Grid Security Experiences, Oxford. 2004.

\bibitem{WSTrustlink}
OASIS WS-Trust specification. http://docs.oasis-open.org/ws-sx/ws-trust/200512

\bibitem{GridSitelink}
Gridsite delegation service. http://www.gridsite.org/wiki/Delegation_protocol

\bibitem{GAloisio05}
Giovanni Aloisio, Massimo Cafaro, Italo Epicoco, Daniele Lezzi, and Robert van Engelen, The GSI plug-in for gSOAP: Enhanced Security, Performance, and Reliability, ITCC conference 2005, IEEE Press, Volume I, pages 304-309.

\end{thebibliography}

\end{document}
% end of file template.tex

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
Making pdflatex working

@@
 % your LaTeX will extract the file if required
-\begin{filecontents*}{example.eps}
-%!PS-Adobe-3.0 EPSF-3.0
-%%BoundingBox: 19 19 221 221
-%%CreationDate: Mon Sep 29 1997
-%%Creator: programmed by hand (JK)
-%%EndComments
-gsave
-newpath
-  20 20 moveto
-  20 220 lineto
-  220 220 lineto
-  220 20 lineto
-closepath
-2 setlinewidth
-gsave
-  .4 setgray fill
-grestore
-stroke
-grestore
-\end{filecontents*}
 %
@@
 Grid provides the technology for wide-scale, cross-domain collaboration where users work with resources and share data which are distributed across administrative domains.
-This kind of collaboration is called Virtual Organization (VO)\cite{IFoster01}.
+This kind of collaboration is called Virtual Organization (VO)~\cite{IFoster01}.
 The resources that contribute to the VO can come from independent administrative domains, and even require different security approaches to access control due to usage of different underlying middlewares.
@@
 
-In the current Grid community, the Grid Security Infrastructure (GSI)\cite{IFoster98} is the de facto standard for authentication and transport level security.
+In the current Grid community, the Grid Security Infrastructure (GSI)~\cite{IFoster98} is the de facto standard for authentication and transport level security.
 It builds on the Public Key Infrastructure (PKI), an architecture based on the X.509 public key certificate.
@@
 
-In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS)\cite{AlfieriR05} have partly solved them, to a different extent.
+In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS)~\cite{AlfieriR05} have partly solved them, to a different extent.
 The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided by creating the proxy certificate.
-As the most successful Grid authorization model, VOMS adopts the Attribute Certificate\cite{RFC3821link} approach.
+As the most successful Grid authorization model, VOMS adopts the Attribute Certificate~\cite{RFC3821link} approach.
 The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.Attributes include the VO membership data which is associated with the proxy certificate’s identity.
@@
 This way, the Web Service technology has been utilized in Grid computing middlewares.
-The Web Service technology has been adopted by the Globus Toolkit\cite{GTlink}, gLite\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)\cite{ARClink}.
+The Web Service technology has been adopted by the Globus Toolkit~\cite{GTlink}, gLite~\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)~\cite{ARClink}.
 In fact, utilizing Web Service technology and providing seamless interoperability with other middlewares is one of the main goals of the new version of ARC middleware.
@@
 The Attribute Certificate has not been adopted by any Web Service implementation at all.
-Web Service implementations commonly use the Web Services Security (WS-Security)\cite{WSSeclink}, Security Assertion Markup Languages (SAML)\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
+Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
 Finally, identity delegation is a good solution for user or user’s representative (acting on behalf of the user) to move/travel from one resource to another.
@@
 ARC (Advanced Resource Connector) is an open source Grid middleware solution released under Apache license.
-ARC middleware aims at developing self-organized, fault-tolerant, non-intrusive, easy-manageable Grid middleware\cite{MEllert07}.
+ARC middleware aims at developing self-organized, fault-tolerant, non-intrusive, easy-manageable Grid middleware~\cite{MEllert07}.
 The current production version of ARC provides Grid services for job submission and management, resource characterization, resource aggregation and discovery, basic data management, integration of Grid security solution, and so on.
@@
 
-The next generation of ARC components is developed by the KnowARC project\cite{KnowARClink}, based on the functionality and capabilities of the classic ARC middleware.
-It aims at implementing a Web Service oriented Grid middleware which will provide higher levels of resource and user abstraction through well-defined Web Service interface\cite{KnowARCDesignlink} in order to provide interoperability with other service-based Grid middlewares, as well as other Web Service compatible applications.
+The next generation of ARC components is developed by the KnowARC project~\cite{KnowARClink}, based on the functionality and capabilities of the classic ARC middleware.
+It aims at implementing a Web Service oriented Grid middleware which will provide higher levels of resource and user abstraction through well-defined Web Service interface~\cite{KnowARCDesignlink} in order to provide interoperability with other service-based Grid middlewares, as well as other Web Service compatible applications.
 
@@
 A service administrator can configure the MCCs according to the interoperability requirements with the other part.
-For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE’s SoapSender and SoapReceiver)\cite{WSElink}.
-Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)\cite{AShoshani03} service.
+For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE’s SoapSender and SoapReceiver)~\cite{WSElink}.
+Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)~\cite{AShoshani03} service.
 This shows the flexibility of HED in terms of protocols support.
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{HED.eps}
+\includegraphics[width=0.75\textwidth]{HED.png}
 \caption{The example of Host Environment Daemon deployed with A-REX and File services}
@@
 Our work focuses on the following five major aspects:
-Integration of community authentication – to utilize the standard-compatible community authentication mechanism from the Shibboleth\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
+Integration of community authentication – to utilize the standard-compatible community authentication mechanism from the Shibboleth~\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
 Short lived credentials service – to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{SAML2SSO.eps}
+\includegraphics[width=0.75\textwidth]{SAML2SSO.png}
 \caption{SAML2.0 SSO profile in ARC}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{SLCS.eps}
+\includegraphics[width=0.75\textwidth]{SLCS.png}
 \caption{Short lived credential service}
@@
 It should be noticed that the process of composing the distinguished name (DN) for the certificate is a critical issue for the SLCS service.
-Since the Shibboleth Identity Provider uses the eduPerson schema\cite{eduSchemalink} for the composing the SAML Attribute we pick the relatively distinguishable attribute “eduPersonPrincipalName” for the DN.
+Since the Shibboleth Identity Provider uses the eduPerson schema~\cite{eduSchemalink} for the composing the SAML Attribute we pick the relatively distinguishable attribute “eduPersonPrincipalName” for the DN.
 A typical example of the eduPersonPrincipalName  could be alice@example.org.
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{Delegation.eps}
+\includegraphics[width=0.75\textwidth]{Delegation.png}
 \caption{Credential delegation service and client}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{Delegation_protocol.eps}
+\includegraphics[width=0.75\textwidth]{Delegation_protocol.png}
 \caption{Credential delegation protocol}
@@
 Additionally, SSL/TLS is incapable of providing end-to-end protection for SOAP message which might flow through multiple intermediate actors, because it only allows each hop to be protected with the resultant security gaps at intermediate actors.
-Web Services Security (WS-Security)\cite{WSSeclink} specification provides means for applying security to Web Service.
+Web Services Security (WS-Security)~\cite{WSSeclink} specification provides means for applying security to Web Service.
 
@@
 When it receives a message from the asserting entity, it will check the SAML assertion based on its predetermined trust relationship with the SAML issuing authority, and check the signature of the SOAP message based on the public key in the SAML assertion without directly trust relationship with attesting entity (holder of this public key).
-In terms of verification, since the proxy certificates have been introduced in IETF (RFC 3820)\cite{RFC3820link}, and OpenSSL has also implemented RFC3820, it is not difficult to use the proxy certificate for the X.509 Token profile to issue an X.509 token.
+In terms of verification, since the proxy certificates have been introduced in IETF (RFC 3820)~\cite{RFC3820link}, and OpenSSL has also implemented RFC3820, it is not difficult to use the proxy certificate for the X.509 Token profile to issue an X.509 token.
 
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{SAMLDelegation.eps}
+\includegraphics[width=0.75\textwidth]{SAMLDelegation.png}
 \caption{SAML Token delegation service and client}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{SAMLDelegationMessage.eps}
+\includegraphics[width=0.75\textwidth]{SAMLDelegationMessage.png}
 \caption{SOAP message header with SAML delegation token inside}
@@
 Similar work that integrates Shibboleth and Grid middleware does exist.
-GridShib\cite{VWelch05},\cite{TBarton06} uses Shibboleth to get entity’s attribute for the authorization of services developed in Globus Tookit.
+GridShib~\cite{VWelch05},\cite{TBarton06} uses Shibboleth to get entity’s attribute for the authorization of services developed in Globus Tookit.
 Unlike the work described in this paper, the GridShib does not use Shibboleth’s native login handlers for authentication to Shibboleth IdP, instead, a new login handler specific for X.509 based authentication is developed for the Shibboleth IdP, and a credential service is developed for acting as an online-CA, authenticating the user, and issuing short-lived X.509 credential.
@@
 
-SWITCH (The Swiss Education and Research Network) short-lived credential service (SLCS)\cite{switchslcslink} provides a SLCS service and a client.
+SWITCH (The Swiss Education and Research Network) short-lived credential service (SLCS)~\cite{switchslcslink} provides a SLCS service and a client.
 The advantage of our work comparing to the SWTCH’s solution is that the SLCS service and client are based on Web Services, so that they can be implemented in any Web Service container while keeping interoperability.
 
-In terms of delegation, apart from the credential delegation mechanism of GSI\cite{IFoster98},{VWelch04}, several authors\cite{MAhsant04} propose a delegation protocol based on the WS-Trust\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids.
+In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98},{VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids.
 WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way.
-Gridsite project implements a X.509 credential delegation solution\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.
+Gridsite project implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.
 
-GSI plug-in for gSOAP\cite{GAloisio05} adds GSI support for gSOAP Web Service toolkit.
+GSI plug-in for gSOAP~\cite{GAloisio05} adds GSI support for gSOAP Web Service toolkit.
 The GSI MCC implemented in ARC provides samilar functionality.
@@
 
-% For one-column wide figures use
-\begin{figure}
-% Use the relevant command to insert your figure file.
-% For example, with the graphicx package use
-  \includegraphics{example.eps}
-% figure caption is below the figure
-\caption{Please write your figure caption here}
-\label{fig:1}       % Give a unique label
-\end{figure}
+%% For one-column wide figures use
+%\begin{figure}
+%% Use the relevant command to insert your figure file.
+%% For example, with the graphicx package use
+%  \includegraphics{example.eps}
+%% figure caption is below the figure
+%\caption{Please write your figure caption here}
+%\label{fig:1}       % Give a unique label
+%\end{figure}
+%%
+%% For two-column wide figures use
+%\begin{figure*}
+%% Use the relevant command to insert your figure file.
+%% For example, with the graphicx package use
+%  \includegraphics[width=0.75\textwidth]{example.eps}
+%% figure caption is below the figure
+%\caption{Please write your figure caption here}
+%\label{fig:2}       % Give a unique label
+%\end{figure*}
 %
-% For two-column wide figures use
-\begin{figure*}
-% Use the relevant command to insert your figure file.
-% For example, with the graphicx package use
-  \includegraphics[width=0.75\textwidth]{example.eps}
-% figure caption is below the figure
-\caption{Please write your figure caption here}
-\label{fig:2}       % Give a unique label
-\end{figure*}
-%
-% For tables use
-\begin{table}
-% table caption is above the table
-\caption{Please write your table caption here}
-\label{tab:1}       % Give a unique label
-% For LaTeX tables use
-\begin{tabular}{lll}
-\hline\noalign{\smallskip}
-first & second & third  \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
-number & number & number \\
-number & number & number \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
+%% For tables use
+%\begin{table}
+%% table caption is above the table
+%\caption{Please write your table caption here}
+%\label{tab:1}       % Give a unique label
+%% For LaTeX tables use
+%\begin{tabular}{lll}
+%\hline\noalign{\smallskip}
+%first & second & third  \\
+%\noalign{\smallskip}\hline\noalign{\smallskip}
+%number & number & number \\
+%number & number & number \\
+%\noalign{\smallskip}\hline
+%\end{tabular}
+%\end{table}
 
@@
 \bibitem{KnowARCDesignlink}
-Design document of new version ARC. https://www.knowarc.eu/documents/Knowarc_D1.1‑1_07.pdf
+Design document of new version ARC. https://www.knowarc.eu/documents/Knowarc\_D1.1-1\_07.pdf
 
@@
 \bibitem{VWelch05}
-V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. 4th Annual PKI R&D Workshop, 2005.
+V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. 4th Annual PKI R\&D Workshop, 2005.
 
 \bibitem{TBarton06}
-T. Barton et al. Identity Federation and Attribute-based Authorization through the Globus Toolkit, Shibboleth, Gridshib, and MyProxy. 5th Annual PKI R&D Workshop, 2006.
+T. Barton et al. Identity Federation and Attribute-based Authorization through the Globus Toolkit, Shibboleth, Gridshib, and MyProxy. 5th Annual PKI R\&D Workshop, 2006.
 
@@
 \bibitem{VWelch04}
-V. Welch et al. X.509 proxy certificate for dynamic delegation, Proceeding of the 3rd Annual KI R&D Workshop, 2004.
+V. Welch et al. X.509 proxy certificate for dynamic delegation, Proceeding of the 3rd Annual KI R\&D Workshop, 2004.
 
@@
 \bibitem{GridSitelink}
-Gridsite delegation service. http://www.gridsite.org/wiki/Delegation_protocol
+Gridsite delegation service. http://www.gridsite.org/wiki/Delegation\_protocol
 

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
Some minor changes in tex file

@@
 }
-\subtitle{Do you have a subtitle?\\ If so, write it here}
+%\subtitle{Do you have a subtitle?\\ If so, write it here}
 
@@
 A service administrator can configure the MCCs according to the interoperability requirements with the other part.
-For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE’s SoapSender and SoapReceiver)~\cite{WSElink}.
+For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE's SoapSender and SoapReceiver)~\cite{WSElink}.
 Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)~\cite{AShoshani03} service.
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{HED.png}
+\includegraphics[width=0.9\columnwidth]{HED.png}
 \caption{The example of Host Environment Daemon deployed with A-REX and File services}
@@
 If any of them fails, message processing fails as well.
-In Figure \ref{fig:HED}, the “AuthZ” and “AuthN” sub-modules inside MCCs and Services are examples of SecHandler.
+In Figure \ref{fig:HED}, the "AuthZ" and "AuthN" sub-modules inside MCCs and Services are examples of SecHandler.
 
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{SAML2SSO.png}
+\includegraphics[width=0.9\columnwidth]{SAML2SSO.png}
 \caption{SAML2.0 SSO profile in ARC}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{SLCS.png}
+\includegraphics[width=0.9\columnwidth]{SLCS.png}
 \caption{Short lived credential service}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{Delegation.png}
+\includegraphics[width=0.9\columnwidth]{Delegation.png}
 \caption{Credential delegation service and client}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{Delegation_protocol.png}
+\includegraphics[width=0.9\columnwidth]{Delegation_protocol.png}
 \caption{Credential delegation protocol}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{SAMLDelegation.png}
+\includegraphics[width=0.9\columnwidth]{SAMLDelegation.png}
 \caption{SAML Token delegation service and client}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{SAMLDelegationMessage.png}
+\includegraphics[width=0.9\columnwidth]{SAMLDelegationMessage.png}
 \caption{SOAP message header with SAML delegation token inside}
@@
 
-In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98},{VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids.
+In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98},~\cite{VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids.
 WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way.

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
More polishing

@@
 %\subtitle{Do you have a subtitle?\\ If so, write it here}
-
-%\titlerunning{Short form of title}        % if too long for running head
-
+\titlerunning{Authentication and single sign-on in ARC}        % if too long for running head
 \author{Weizhong Qiang         \and
@@
 %\authorrunning{Short form of author list} % if too long for running head
-
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}           %  \\
+\institute{W. Qiang \at
+              Department of Physics \\
+              University of Oslo, Norway\\
+              \email{weizhong.qiang@fys.uio.no}           %  \\
 %             \emph{Present address:} of F. Author  %  if needed
            \and
-           S. Author \at
-              second address
+           A. Konstantinov \at
+              \email{aleksandr.konstantinov@fys.uio.no}
 }
-
 \date{Received: date / Accepted: date}
 % The correct dates will be entered by the editor
-
 \maketitle
-
 \begin{abstract}
@@
 To solve the above issues, Grid middlewares should provide the following abilities:
-The user which participates a VO should be able to access data and resources without being required to constantly do authentication at each resource or service;
-The VO should be able to easily manage the privileges and memberships of the users;
-The VO together with its resources should be able to easily manage the authorization, given the condition that the resources are owned by different administrative domains.
-
+\begin{itemize}
+\item The user which participates a VO should be able to access data and resources without being required to constantly do authentication at each resource or service;
+\item The VO should be able to easily manage the privileges and memberships of the users;
+\item The VO together with its resources should be able to easily manage the authorization, given the condition that the resources are owned by different administrative domains.
+\end{itemize}
 In the current Grid community, the Grid Security Infrastructure (GSI)~\cite{IFoster98} is the de facto standard for authentication and transport level security.
@@
 Unfortunately, the existing Grid security solutions such as GSI and VOMS do not apply to Web Service based applications very well due to the following reasons:
-The GSS API based confidential communication has not been adopted by Web Service implementations.
-The Attribute Certificate has not been adopted by any Web Service implementation at all.
-Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
-Finally, identity delegation is a good solution for user or user’s representative (acting on behalf of the user) to move/travel from one resource to another.
+\begin{itemize}
+\item The GSS API based confidential communication has not been adopted by Web Service implementations.
+\item The Attribute Certificate has not been adopted by any aspects Web Service implementation at all.
+\item Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
+\item Finally, identity delegation is a good solution for user or user’s representative (acting on behalf of the user) to move/travel from one resource to another.
 But since the identity delegation has been completely coupled with the GSI confidential communication process, it is not possible for the existing identity delegation solution to be used by Web Service applications.
+\end{itemize}
 
@@
 Our work focuses on the following five major aspects:
-Integration of community authentication – to utilize the standard-compatible community authentication mechanism from the Shibboleth~\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
-Short lived credentials service – to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
-Credential delegation service – to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
-Message level authentication – to provide message level authentication by implementing WS-Security.
-SAML token based delegation service – to process the delegation of an WS Security SAML token in order to provide single sign-on for SOAP message level authentication.
+\begin{itemize}
+	\item Integration of community authentication -- to utilize the standard-compatible community authentication mechanism from the Shibboleth~\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
+	\item Short lived credentials service – to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
+	\item Credential delegation service – to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
+	\item Message level authentication – to provide message level authentication by implementing WS-Security.
+	\item SAML token based delegation service – to process the delegation of an WS Security SAML token in order to provide single sign-on for SOAP message level authentication.
+\end{itemize}
 
@@
 Here is the explanation of steps shown in Figure \ref{fig:SAML2SSO}:
-1.The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side.
+\begin{enumerate}
+\item The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side.
 The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is “saml2sp” which is specific for pointing to the SP Service.
 Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
-2.The SP Service searches the metadata (we use the same metadata format as defined in the Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself).
+\item The SP Service searches the metadata (we use the same metadata format as defined in the Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself).
 Then SP Service uses obtained information to compose and issue the SAML authentication request also by utilizing own X.509 certificate (note that in the SAML2.0 SSO profile, X.509 certificates are still needed for IdP and SP) and sends the response back to the user agent.
-3.User agent sends the SAML authentication request to the Identity Provider.
-4.Identity Provider requires an act of authentication.
+\item User agent sends the SAML authentication request to the Identity Provider.
+\item Identity Provider requires an act of authentication.
 The authentication mechanism is outside of the SAML2.0 SSO profile.
@@
 Through the HTTP protocol, the user agent will feed IdP with the username/password which has been given by the caller of user agent interface.
-5.Once the authentication has been succeeded, the IdP issues a SAML response including SAML assertion encrypted by destination SP’s public key, and then this SAML response will be delivered by the user agent to the Service Provider.
-6.The SP Service verifies and checks the SAML response, decrypts and stores the SAML assertion into session/connection context.
+\item Once the authentication has been succeeded, the IdP issues a SAML response including SAML assertion encrypted by destination SP’s public key, and then this SAML response will be delivered by the user agent to the Service Provider.
+\item The SP Service verifies and checks the SAML response, decrypts and stores the SAML assertion into session/connection context.
 That assertion includes the AuthnStatement and AttributeStatement elements.
-7.The WS client launches the Grid/Web Service request via the same connection as the one which is used by user agent to contact SP Service.
-8.The Grid/Web Service checks the AuthnStatement stored in the session context to see if the session is still valid.
+\item The WS client launches the Grid/Web Service request via the same connection as the one which is used by user agent to contact SP Service.
+\item The Grid/Web Service checks the AuthnStatement stored in the session context to see if the session is still valid.
 If valid, service handles the  request and returns the response to the WS client.
 Note that in the current implementation the service requires WS client to reuse same TCP connection as the one used by user agent is step 5, in order to guarantee  validity of the SSO result.
-
+\end{enumerate}
 The SP service and other functional service(s) are hosted by the same container, and they use the same X.509 credential for the transport layer security.
@@
 The delegation process is similar to the process in the X.509 credential delegation service, with the following differences:
-1.Transport level communication between delegation client and service is not based on mutual authentication between the proxy certificate and the service certificate; instead, it is based on mutual authentication between the service certificates.
-2.Delegation service will not generate any certificate request; instead, the delegation client will pick the public key from the peer certificate after a successful SSL/TLS authentication.
+\begin{enumerate}
+	\item Transport level communication between delegation client and service is not based on mutual authentication between the proxy certificate and the service certificate; instead, it is based on mutual authentication between the service certificates.
+	\item Delegation service will not generate any certificate request; instead, the delegation client will pick the public key from the peer certificate after a successful SSL/TLS authentication.
 So step 2 of the delegation protocol is just some confirmation information which informs delegation client to proceed.
-3.For those multiple intermediate services through which a SOAP message flows, delegation service does not need to appear, instead, it only needs to appear on the end of the SOAP message flow.
-4.SAML Token might include one or a few <saml:Attribute> attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.
-
+	\item For those multiple intermediate services through which a SOAP message flows, delegation service does not need to appear, instead, it only needs to appear on the end of the SOAP message flow.
+	\item SAML Token might include one or a few <saml:Attribute> attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.
+\end{enumerate}
 Figure \ref{fig:SAMLDelegationMessage} shows the SAML Token (SAML Assertion) inside the SOAP message is changed due to delegation (we suppose X.509 token is used for attesting entity to authenticate against issuing authority).
@@
 Similar work that integrates Shibboleth and Grid middleware does exist.
-GridShib~\cite{VWelch05},\cite{TBarton06} uses Shibboleth to get entity’s attribute for the authorization of services developed in Globus Tookit.
-Unlike the work described in this paper, the GridShib does not use Shibboleth’s native login handlers for authentication to Shibboleth IdP, instead, a new login handler specific for X.509 based authentication is developed for the Shibboleth IdP, and a credential service is developed for acting as an online-CA, authenticating the user, and issuing short-lived X.509 credential.
+GridShib~\cite{VWelch05,TBarton06} uses Shibboleth to get entity's attribute for the authorization of services developed in Globus Tookit.
+Unlike the work described in this paper, the GridShib does not use Shibboleth's native login handlers for authentication to Shibboleth IdP, instead, a new login handler specific for X.509 based authentication is developed for the Shibboleth IdP, and a credential service is developed for acting as an online-CA, authenticating the user, and issuing short-lived X.509 credential.
 Moreover, this paper provides a single sign-on solution for the user with community credentials authenticating to Grid services by directly using SAML2.0 SSO profile.
@@
 SWITCH (The Swiss Education and Research Network) short-lived credential service (SLCS)~\cite{switchslcslink} provides a SLCS service and a client.
-The advantage of our work comparing to the SWTCH’s solution is that the SLCS service and client are based on Web Services, so that they can be implemented in any Web Service container while keeping interoperability.
+The advantage of our work comparing to the SWTCH's solution is that the SLCS service and client are based on Web Services, so that they can be implemented in any Web Service container while keeping interoperability.
 
-In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98},~\cite{VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids.
+In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98,VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids.
 WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way.
@@
 \bibitem{IFoster01}
-Foster I, Kesselman C, Tuecke S. The anatomy of the Grid: Enabling scalable virtual organizations, International Journal of Supercomputer Applications, 15(3), 200–222 (2001)
+Foster I, Kesselman C, Tuecke S. The anatomy of the Grid: Enabling scalable virtual organizations, International Journal of Supercomputer Applications, 15(3), 200-222 (2001)
 
@@
 Future Generation Comp.
-Syst., 21(4), 549–558 (2005)
+Syst., 21(4), 549-558 (2005)
 
@@
 \bibitem{MEllert07}
-M. Ellert et al. Advanced resource connector middleware for lightweight computational grids, Future Generation computer systems, 23(2), 219–240 (2007)
+M. Ellert et al. Advanced resource connector middleware for lightweight computational grids, Future Generation computer systems, 23(2), 219-240 (2007)
 
@@
 \bibitem{GAloisio05}
-Giovanni Aloisio, Massimo Cafaro, Italo Epicoco, Daniele Lezzi, and Robert van Engelen, The GSI plug-in for gSOAP: Enhanced Security, Performance, and Reliability, ITCC conference 2005, IEEE Press, Volume I, pages 304-309.
+Giovanni Aloisio, Massimo Cafaro, Italo Epicoco, Daniele Lezzi, and Robert van Engelen, The GSI plug-in for gSOAP: Enhanced Security, Performance, and Reliability, ITCC conference 2005, IEEE Press, Volume I, pp.
+304-309.
 

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
more typesetting

@@
 Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service.
-Each MCC or Service is usually configured with two queues of SecHandler – one for the incoming message and one for the outgoing message respectively.
+Each MCC or Service is usually configured with two queues of SecHandler -- one for the incoming message and one for the outgoing message respectively.
 All SecHandler components attached to the queue are executed sequentially.
@@
 \label{sec:siglesignon}
-We investigated the authentication and single sign-on issue in the scenario when a user or user’s representative (acting on behalf of this user) needs to access resources across different middlewares.
+We investigated the authentication and single sign-on issue in the scenario when a user or user's representative (acting on behalf of this user) needs to access resources across different middlewares.
 Our work focuses on the following five major aspects:
@@
 	\item Integration of community authentication -- to utilize the standard-compatible community authentication mechanism from the Shibboleth~\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
-	\item Short lived credentials service – to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
-	\item Credential delegation service – to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
-	\item Message level authentication – to provide message level authentication by implementing WS-Security.
-	\item SAML token based delegation service – to process the delegation of an WS Security SAML token in order to provide single sign-on for SOAP message level authentication.
+	\item Short lived credentials service -- to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
+	\item Credential delegation service -- to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
+	\item Message level authentication -- to provide message level authentication by implementing WS-Security.
+	\item SAML token based delegation service --- to process the delegation of an WS Security SAML token in order to provide single sign-on for SOAP message level authentication.
 \end{itemize}
@@
 The Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity, while the Service Provider (SP) is responsible for controlling access to services and resources by analyzing the SAML assertions produced and issued by the IdP upon request.
-In order to discover which home community does a user come from, Shibboleth specifies an optional third component called “Where Are You From?” (WAYF) service to aid in the process of IdP discovery.
+In order to discover which home community does a user come from, Shibboleth specifies an optional third component called ``Where Are You From?'' (WAYF) service to aid in the process of IdP discovery.
 This IdP discovery process is also standardized and defined in SAML 2.0 specification as “Identity Provider Discovery Profile”.
@@
 \item The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side.
-The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is “saml2sp” which is specific for pointing to the SP Service.
+The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is ``saml2sp'' which is specific for pointing to the SP Service.
 Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
@@
 The SP service and other functional service(s) are hosted by the same container, and they use the same X.509 credential for the transport layer security.
-The client authentication may be switched off, so that client doesn’t need to use any X.509 credential.
+The client authentication may be switched off, so that client doesn't need to use any X.509 credential.
 Only the trusted CAs of SP and IdP services need to be configured on the client side to ensure SP and IdP can authentication.
@@
 The benefit of the SAML2.0 SSO profile worth mentioning is that the Identity Provider can cache the authentication result through session management once the user agent has succeeded to authenticate.
-Then for a short period this authentication result is valid so that the user agent doesn’t need to feed IdP with user’s username and password the next time it authenticates against IdP.
+Then for a short period this authentication result is valid so that the user agent doesn't need to feed IdP with user's username and password the next time it authenticates against IdP.
 So the user (or the client on behalf of that user) can travel across multiple security domains with only providing his name and password once, which is the characteristic of single sign-on.
@@
 The SLCS client is implemented as command-line interface (CLI) utility which implements the user agent and WS client functionalities.
-The whole process of SLCS invocation is shown in Figure \ref{fig:SLCS} (steps 1‑8).
+The whole process of SLCS invocation is shown in Figure \ref{fig:SLCS} (steps 1--8).
 The process resembles one shown in in Figure \ref{fig:SAML2SSO}, except that steps 7 and 8 use SLCS certificate request and response.
@@
 It should be noticed that the process of composing the distinguished name (DN) for the certificate is a critical issue for the SLCS service.
-Since the Shibboleth Identity Provider uses the eduPerson schema~\cite{eduSchemalink} for the composing the SAML Attribute we pick the relatively distinguishable attribute “eduPersonPrincipalName” for the DN.
+Since the Shibboleth Identity Provider uses the eduPerson schema~\cite{eduSchemalink} for the composing the SAML Attribute we pick the relatively distinguishable attribute ``eduPersonPrincipalName'' for the DN.
 A typical example of the eduPersonPrincipalName  could be alice@example.org.
-In such case composed DN is for example “/O=knowarc/OU=example.org/CN=alice”.
+In such case composed DN is for example\\``/O=knowarc/OU=example.org/CN=alice''.
 
 The obvious benefit of the SLCS service is that the user can get the X.509 credentials anywhere simply by running the SLCS client command and providing her username and password, and then access the X.509 based Grid infrastructure conveniently.
-Thanks to the single sign-on capability of SAML2.0 SSO profile, the user doesn’t not need to provide her username and password during valid period after the first time she succeeds to authenticate against her home IdP.
+Thanks to the single sign-on capability of SAML2.0 SSO profile, the user doesn't not need to provide her username and password during valid period after the first time she succeeds to authenticate against her home IdP.
 
@@
 \end{figure}
-For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure \ref{fig:Delegation}), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential.
-For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure \ref{fig:Delegation}), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.
+For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure \ref{fig:Delegation}), should use the user's X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential.
+For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure \ref{fig:Delegation}), should also use the user's X.509 credential or a delegated credential for authentication and secure communication.
 
-Since each WS client accesses the target service (delegation service or other general services) on behalf of the user, the trust relationship between the user’s certificates (X.509 certificate and proxy certificate) and services is required, while the trust relationship between service’s certificate and another service’s certificate is not required.
+Since each WS client accesses the target service (delegation service or other general services) on behalf of the user, the trust relationship between the user's certificates (X.509 certificate and proxy certificate) and services is required, while the trust relationship between service's certificate and another service's certificate is not required.
 
@@
 
-If the WS client owns an X.509 certificate or proxy certificate, the process of generating an X.509 Token for a SOAP message is: this client puts the content of certificate into X.509 token as <wsse:BinarySecurityToken>, and uses the private key (corresponding to the certificate) to create XML signature for the SOAP message body.
+If the WS client owns an X.509 certificate or proxy certificate, the process of generating an X.509 Token for a SOAP message is: this client puts the content of certificate into X.509 token as $<$wsse:BinarySecurityToken$>$, and uses the private key (corresponding to the certificate) to create XML signature for the SOAP message body.
 
@@
 	\item For those multiple intermediate services through which a SOAP message flows, delegation service does not need to appear, instead, it only needs to appear on the end of the SOAP message flow.
-	\item SAML Token might include one or a few <saml:Attribute> attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.
+	\item SAML Token might include one or a few $<$saml:Attribute$>$ attributes which carry the attribute information related to the subject, instead of the Attribute Certificate which is used in the VOMS proxy certificate for carrying the attribute information.
 \end{enumerate}
@@
 The SOAP message on the left side is the one which B sends to the relying party (intermediate service).
-B is the identity of the attesting entity’s X.509 token, while A is the identity of issuing authority, and the SOAP message is signed by B’s private key which is corresponding to the X.509 certificate that is used to create the X.509 token.
-On the right side, C is the identity of the delegation service (also C is the identity of the target service to which B will access), and now the SOAP message is signed by C’s private key which is corresponding to the target service’s certificate.
+B is the identity of the attesting entity's X.509 token, while A is the identity of issuing authority, and the SOAP message is signed by B's private key which is corresponding to the X.509 certificate that is used to create the X.509 token.
+On the right side, C is the identity of the delegation service (also C is the identity of the target service to which B will access), and now the SOAP message is signed by C's private key which is corresponding to the target service's certificate.
 

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
More minor correction

@@
            A. Konstantinov \at
+              Department of Physics \\
+              University of Oslo, Norway\\
               \email{aleksandr.konstantinov@fys.uio.no}
@@
 
-In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS)~\cite{AlfieriR05} have partly solved them, to a different extent.
+In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS) ~\cite{AlfieriR05} have partly solved them, to a different extent.
 The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided by creating the proxy certificate.
@@
 \item Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
-\item Finally, identity delegation is a good solution for user or user’s representative (acting on behalf of the user) to move/travel from one resource to another.
+\item Finally, identity delegation is a good solution for user or user's representative (acting on behalf of the user) to move/travel from one resource to another.
 But since the identity delegation has been completely coupled with the GSI confidential communication process, it is not possible for the existing identity delegation solution to be used by Web Service applications.
@@
 
-The above analysis shows that in order to achieve interoperability and transparency for user to access Grid system which is established on different middlewares, we need to address the mismatch between the different security solutions from these middlewares in order to ensure the interoperability as wide as possible; it is also necessary to address the mismatch between users’ daily experience and middleware’s requirements in order to ensure the access as simple as possible.
+The above analysis shows that in order to achieve interoperability and transparency for user to access Grid system which is established on different middlewares, we need to address the mismatch between the different security solutions from these middlewares in order to ensure the interoperability as wide as possible; it is also necessary to address the mismatch between users' daily experience and middleware's requirements in order to ensure the access as simple as possible.
 
@@
 
-The rest of this paper is organized as follows: Section II presents the ARC Grid middleware, especially the architecture of the new version of ARC.
-Section III describes the solution for cross-middleware authentication and single sign-on, including implementations for bridging the mismatch in terms of authentication and single sign-on.
-Section IV discusses the related research, and Section V presents conclusions and future work outlook.
+The rest of this paper is organized as follows: Section 2 presents the ARC Grid middleware, especially the architecture of the new version of ARC.
+Section 3 describes the solution for cross-middleware authentication and single sign-on, including implementations for bridging the mismatch in terms of authentication and single sign-on.
+Section 4 discusses the related research, and Section 5 presents conclusions and future work outlook.
 
@@
 If any of them fails, message processing fails as well.
-In Figure \ref{fig:HED}, the "AuthZ" and "AuthN" sub-modules inside MCCs and Services are examples of SecHandler.
+In Figure \ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules inside MCCs and Services are examples of SecHandler.
 
@@
 In order to discover which home community does a user come from, Shibboleth specifies an optional third component called ``Where Are You From?'' (WAYF) service to aid in the process of IdP discovery.
-This IdP discovery process is also standardized and defined in SAML 2.0 specification as “Identity Provider Discovery Profile”.
+This IdP discovery process is also standardized and defined in SAML 2.0 specification as ``Identity Provider Discovery Profile''.
 
@@
 \label{sec:conclusion}
-In this paper we presented the solution built into the ARC Grid middleware, aiming towards the cross-middleware authentication and single sign-on.
+In this paper we presented the solution built into the ARC Grid middleware, aiming towards the cross-middle-ware authentication and single sign-on.
 The goal of our work is to develop an interoperable and accessible single sign-on solution to be used for a VO scenario in production Grids, where users can easily authenticate against the Grids with any of their credentials (not only X.509) from anywhere, and move inside Grid middleware infrastructure and even across heterogeneous Grids  boundaries, without re-authentication within a reasonable period.

### 2009-02-01 doc/sec/papers/ISC09/ISC09.tex
More minor correction

@@
 \item Identity Provider requires an act of authentication.
-The authentication mechanism is outside of the SAML2.0 SSO profile.
+The authentication mechanism is outside of the SAML 2.0 SSO profile.
 The Shibboleth IdP implementation uses login handlers for authentication.
@@
 \bibitem{IFoster01}
-Foster I, Kesselman C, Tuecke S. The anatomy of the Grid: Enabling scalable virtual organizations, International Journal of Supercomputer Applications, 15(3), 200-222 (2001)
+I. Foster, C. Kesselman, S. Tuecke. The anatomy of the Grid: Enabling scalable virtual organizations, International Journal of Supercomputer Applications, 15(3), 200-222 (2001)
 
 \bibitem{IFoster98}
-Foster, I., Kesselman, C., Tsudik, G. and Tuecke, S. A Security Architecture for Computational Grids, ACM Conference on Computers and Security, 1998, 83-91.
+I. Foster, C. Kesselman, G. Tsudik, and S. Tuecke, A Security Architecture for Computational Grids, ACM Conference on Computers and Security, 1998, 83-91.
 
@@
 \bibitem{GAloisio05}
-Giovanni Aloisio, Massimo Cafaro, Italo Epicoco, Daniele Lezzi, and Robert van Engelen, The GSI plug-in for gSOAP: Enhanced Security, Performance, and Reliability, ITCC conference 2005, IEEE Press, Volume I, pp.
+G.
+Aloisio, M.
+Cafaro, I.
+Epicoco, D.
+Lezzi, and R.
+van Engelen, The GSI plug-in for gSOAP: Enhanced Security, Performance, and Reliability, International Conference on Information Technology: Coding and Computing (ITCC'05), IEEE Press, Volume I, pp.
 304-309.

### 2009-02-04 doc/client/resubmit_and_migrate_plans.tex
implementations plans for resubmission and migration

\documentclass[12pt,oneside,a4paper,english]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{graphicx,epic,eepic}
\usepackage{amssymb}
\usepackage{amsmath}

\title{Thoughts on resubmission within ARC1}
\author{Peter Rosendahl}

\begin{document}
%\maketitle
\section*{Implementation plan for the resubmission and migration functionality in ARC1}
Before discussing the implementation of the resubmission and migration functionality in ARC1 we first make a distinction between migrating, resubmitting and resuming a job.

{\bf Resubmission} of a job means that the exact same job is redone from scratch, i.e. no changes in the job description or input files. If the job description or input files has changed the job no longer remains the same job and the user should submit a new job instead of resubmitting the previous job.

{\bf Resuming} a job means restarting the job from its last resumable state. The last resumable state of a job is reported by the cluster running the job, thus this action will be dependent on the cluster's middleware. We will only talk about resuming a job if the job has already started running otherwise there will be nothing to resume.

{\bf Migrating} a job is defined as moving a job from a cluster(queue) to a new cluster(queue) where it will then be executed.

\subsection*{Solutions}
In order to solve the scenarios mentioned above, we suggest to implement functionality on both server and client side to provide the commands \texttt{arcresume}, \texttt{arcmigrate} and \texttt{arcresub}.

\subsection*{\texttt{arcresume}}
The {\texttt arcresume} command will restart the job from its last resumable state (as reported by the cluster) on the same cluster (computing element) as the job initially was running on. A typical use of this functionality could be if a job has run successfully, but fails in the uploading of output files, e.g. due to the expiration of the user proxy. The user can resolve this situation by uploading a new proxy to the cluster and invoking the {\texttt arcresume} command which resumes the job and the file upload. Since the job will resume at the same cluster (computing element) the {\texttt arcresume} command should be possible to implement for all supported middlewares if the middleware itself supports the resume action. {\bf This command is neither a migration nor a resubmission of the job, and thus not foreseen to be a part of the D2.2-1 deliverable, but will be implemented as a part of Task 1.2.}

\subsection*{\texttt{arcmigrate}}
The \texttt{arcmigrate} command should enable users to move jobs from one cluster to another. In this first implementation the command will only support the migration of jobs which are queued in the batch system, i.e. there will be no support for check-pointing etc allowing for migration of running jobs. Nevertheless migration of queueing jobs is the foreseen most frequent use case, in which users would like to migrate queueing jobs from one cluster (computing element) to another cluster (computing element) which meanwhile have been freed or made available in other ways.

Upon migration of the job, both the job description and the local input files will be copied directly from the cluster where the job is queueing to the new cluster where the job will be executed\footnote{Input files from storage elements will be copied from the storage element onto the new cluster}. No files or job description will pass via the user's host, thus enabling the user to migrate a job even if the local input files has been changed or deleted. Additionally this will enable the user to perform job migration from a host different from the submission host, e.g. migrating jobs from an internet cafe.

{\bf Keep in mind:} Upon job migration the initial job is killed and hence the job receives a new job-id. The old job-id and accompanying job information (endpoint etc) thus have to be updated in the user's local XML file which holds the list of all active jobs.

\subsubsection*{Technical imlplementation:}

\subsection*{\texttt{arcresub}}
The resubmission command will redo an entire job from scratch by first trying to kill the existing job (if possible) and then resubmit the job. The command could e.g. be used if the contact to the cluster running a job is lost, and the user would like to resubmit the job to a different cluster. In this particular use case no information about the job can be retrieved from the cluster, thus everything constituting the job (job description and input files) will have to be retrieved from elsewhere. This is solved by implementing a local sandbox which upon job submission automatically stores the job description and checksums of the local input files. The local sandbox ensures that the user is able to resubmit the job to a new cluster even if the cluster running the job goes down, and that the resubmitted job contains the exact same input files as the original job (reliable resubmission)\footnote{The reliable resubmission only governs the local input files. Input files taken from storage elements are not under this control}.

The feature of storing job descriptions and checksums in a local sandbox could in the future be expanded to storing these in a ``grid sandbox'', e.g. on a grid storage element. The ``grid sandbox'' could act as a remote joblist which would give an alternative way for the \texttt{arcsync} command to syncronise the local joblist. If the grid sandbox even stores local input files upon submission, it could additionally be used to allow the user to resubmit the job from a different host then the original submission host.

{\bf Comparison with ngresub:} The existing \texttt{ngresub} command relies on retrieving the job description from the cluster before resubmitting the job. This prevents the command from being used in the most likely usage scenario in which the original cluster is down. Additionally the \texttt{ngresub} command does not check that local input files are the same, i.e. it can not guarantee that the new job is exactly the same as the old one.

Both the issues above of the \texttt{ngresub} command are resolved with the proposed new implementation (\texttt{arcresub}). The proposed new solution also resolves the situation with multiple middlewares supported by the client library, and enables resubmission to middlewares different from the middleware originally running the job. All this is achieved through the usage of a local sandbox, which on the other side prevents the user from doing resubmission from a different host then the submission host since everything is stored locally. This could in the future be solved by adopting the above mentioned model of a ``grid sandbox'' but local input files are still likely to be a cause of concern.

\subsubsection*{Technical imlplementation:}

% Comments from Balazs:
%relate new job to the old one
%clarify what it means to be started
\end{document}

### 2009-02-05 doc/client/client_technical.tex
Base broker documentation, the specialised brokers docs are missing

--- 2008-09-25 doc/client/client_technical.tex
@@
 
-The {\Broker} is not yet implemented and thus not described here.
+The Broker inherits from the {\ACC} base class in order to be loadable by the ARC \texttt{Loader}.
+It is the base class of the specialized Brokers.
+If you are using the arcsub command then you can choose a broker type with ``-b`` parameter.
+There is a solution for add parameters to these descendants Brokers.
+Here is an example:
+
+\begin{shaded}
+\begin{verbatim}
+arcsub -b RandomBroker:a,b,c -c ARC1:https://knowarc1.grid.niif.hu:60000/arex job.jsdl
+\end{verbatim}
+\end{shaded}
+
+The Broker's constructor set the PreFilteringDone and the TargetSortingDone values to false because we need to call at first the PreFilterTargets and second the GetBestTarget which trigger the specialised Broker sorting method.
+
+The base Broker operations are the followings:
+
+\begin{enumerate}
+\item{The first step is the PreFilterTargets, this implements the matchmaking: the JobDescription and the TargetGenerator's ExecutionTargets will be compared inside this method.
+JobDescription content the job's JobInnerRepresentation, and the ExecutionTarget content the candidate cluster's queue information.
+There are main elements in the JobInnerRepresentation which are used for matchmaking, these are documented in this Appendix: ...
+(every matchmaking elements and their comparison logic are there) The macthmaking working on this way: it takes every JobInnerrepresentation elements which are relevant and in an ordered way check them.
+If a matchmaking element is not presented in the JobInnerRepresentation then it will be ignored, but if it is there than the ExecutionTarget should content its pair.} For example SessionLifeTime is in the JobInnerRepresentaion and the WorkingAreaLifeTime is in the ExecutionTarget (these are matchmaking pair).
+If the WorkingAreaLifeTime is not defined then the Broker cannot check the times therefore it will ignore this cluster, but if the WorkingAreaLifeTime exist and the comparison between this elements are true then this check will be passed other way the cluster will be ignored.
+We are using this rule for every elements checking.
+There is only one exception and this is the ProcessingStartTime because if the DownTime is not defined in the ExecutionTarget than we can trust by it that the job can anytime start and it will be not killed with shutdown.
+This rule checking runs for every ExecutionTargets and if every matchmaking checks are ok for a target then this ExecutionTarget will be added to the PossibleTargets vector.
+\item{If the matchmaking finished then  the PreFilteringDone variable will be true and the TargetSortingDone will be false. }
+\item{The next step is the GetBestTarget method calling, this will give an ExecutionTarget which is good for our job.
+If the Possibletarget vector empty then it will give nothing.
+The first time the GetBestTarget will trigger the specialid Broker type sorting with this method calling:  SortTargets.
+This SortTargets method is virtual method, therefore every specialised Broker should implement it.}
+\item{If the SortTargets finished then the TargetSortingDone will be true and the GetBestTarget will return the first element of the PossibleTargetVector and the current iterator will be incremented with one.}
+\item{If the submision failed to this cluster then we can call again the GetBestTarget and we will get an other ExecutionTarget till it will not consume.
+If  the targets run out from the PossibleTargetVector then the GetBestTarget's bool parameter will be true, its name is EndOfList''}
+\end{enumerate}
+
+\begin{shaded}
+\begin{verbatim}
+class Broker : public ACC {
+public:
+    ExecutionTarget& GetBestTarget(bool &EndOfList);
+    void PreFilterTargets(Arc::TargetGenerator& targen, Arc::JobDescription jd);    
+protected:
+    Broker(Config *cfg);
+    virtual ~Broker();
+    virtual void SortTargets() = 0;
+    std::vector<Arc::ExecutionTarget> PossibleTargets;
+    bool TargetSortingDone;
+    Arc::JobInnerRepresentation jir;
+    static Logger logger;
+private:
+    std::vector<Arc::ExecutionTarget>::iterator current;
+    bool PreFilteringDone;
+};
+\end{verbatim}
+\end{shaded}
+
+The Broker class content  two useful data elements: logger and JobInnerRepresentation.
+These can be helpful when somebody create a specialised Broker and he would like to use the logger and need some information about the job.
 
@@
 
-\section{Specialized Classes (Grid Flavour Plugins)}
+\section{Specialized Classes (Grid Flavour and Broker plugins)}
 \label{sec:plugins}
@@
 gridftp protocol. The gridftp DMC is therefore required.
+
+\subsection{Broker plugins}
+
+\subsubsection{RandomBroker}
+
+The sorting method is random.
+
+TODO: finish
+
+\subsubsection{FastestCPUBroker}
+
+The sorting method based on the fastest cpu, we are using for this the specint2006 benchmark solution.
+
+TODO: finish
+
+\subsubsection{FastestQueueBroker}
+
+The sorting method based on the shortest queue, where the waiting jobs number is low.
+
+TODO: finish
+
+\subsubsection{DataBroker}
+
+The sorting method based on the A-REX file cache. I will describe how does it working.
+
+TODO: finish
+
+\subsubsection{PythonBroker}
+
+TODO: finish
 

### 2009-02-05 doc/client/client_technical.tex
Technical documentation update and mapping correction.

@@
 {\arclib} library this functionality is implemented in the
-{\JobDescription} class. This is a generic class that takes a job
+{\JobDescription} class. The inner representation (\texttt{JobInnerRepresentation}
+class) is like to the GIN-JSDL and all elements are correspond to
+variables. This variables are simple variables or complex data structures,
+and private on the {\JobDescription} class.
+
+This is a generic class that takes a job
 description (string) as input
@@
 \begin{verbatim}
-void setSource(const std::string source);
+bool setSource(const std::string source);
 \end{verbatim}
@@
 in any supported formats (currently XRSL~\cite{xrsl}, JDL~\cite{jdl},
-JSDL~\cite{jsdl}), converts and stores it in a JSDL-like internal job
+JSDL~\cite{jsdl}), converts and stores it in a GIN-JSDL-like internal job
 description format. The identification of the source's description
 format is internally handled by a \texttt{JobDescriptionOrderer} class
-which identifies the format by pattern matching.
+which identifies the format by pattern matching. If the source string is
+empty or the \texttt{JobDescriptionOrderer} class can not be recognize
+to any formats, then it is not store nothing and return with false, else
+convert and store it.
 
@@
 \begin{verbatim}
-void getProduct(std::string& product, std::string format = "JSDL")
+bool getProduct(std::string& product, std::string format = "POSIXJSDL")
 \end{verbatim}
@@
 module for parsing and generating the job descriptions.
-
-Parsing is currently possible from any of the languages but output
-generation is only partially supported. Only JSDL and JDL output
-generation is implemented yet and even for these there are some
-limitations.
+If the inner representation is empty or the source job descrition is not
+valid, then the output is empty and return with false.
+When the source format is equal with the output's format and it is not XRSL,
+than set the product string with the source string. The XRSL output is
+always generation from the inner representation.
+
+Parsing is currently possible from any of the languages. JSDL, XRSL and
+JDL output generation is implemented yet and even for these there are
+some limitations.
 
@@
 attributes are not implemented.
-
-Support for XRSL output generation is currently being developed.
 

### 2009-02-05 doc/client/resubmit_and_migrate_plans.tex
Additions to the technical implementation of the arcmigrate command

@@
-\documentclass[12pt,oneside,a4paper,english]{article}
+\documentclass[11pt,oneside,a4paper,english]{article}
 \usepackage[english]{babel}
@@
 \usepackage{amsmath}
+\usepackage{a4wide}
+\usepackage{xspace}
 
-\title{Thoughts on resubmission within ARC1}
-\author{Peter Rosendahl}
+\newcommand{\arcmigrate}{\texttt{arcmigrate}\xspace}
 
@@
 \subsection*{Solutions}
-In order to solve the scenarios mentioned above, we suggest to implement functionality on both server and client side to provide the commands \texttt{arcresume}, \texttt{arcmigrate} and \texttt{arcresub}.
+In order to solve the scenarios mentioned above, we suggest to implement functionality on both server and client side to provide the commands \texttt{arcresume}, \arcmigrate and \texttt{arcresub}.
 
@@
 
-\subsection*{\texttt{arcmigrate}}
-The \texttt{arcmigrate} command should enable users to move jobs from one cluster to another.
+\subsection*{\arcmigrate}
+The \arcmigrate command should enable users to move jobs from one cluster to another.
 In this first implementation the command will only support the migration of jobs which are queued in the batch system, i.e.
@@
 
+%% Edited by Martin Skou
 \subsubsection*{Technical imlplementation:}
+In a first implementation the \arcmigrate will only be able to migrate a queued job from an ARC1 cluster (old-cluster) to an ARC1 cluster (new-cluster).
+In its simplest invocation the command should take one argument specified by the user ie.
+the job id of a queued job on old-cluster.
+In case the job is not queued an error should be returned and the command should exit.
+
+From the job id a migration specific JSDL file (migration-JSDL) is generated and used communicate with a possible new-cluster (candidate-cluster).
+For the migration-JSDL file extending JSDL was required.
+
+Now clusters are contacted in the usual way and if a connection to a candidate cluster is accepted migration-JSDL is send.
+
+From the received migration-JSDL, candidate-cluster fetches the JSDL file of the job (original-JSDL) which is located on old-cluster.
+Candidate-cluster then examine through original-JSDL if the job are able to run there.
+If positive, input files specified in original-JSDL should be fetched.
+If not candidate-cluster returns an error and the next cluster should be tried.
+
+Input files can be divided in two categories: local files and storage element files.
+As default candidate-cluster should fetch local files from old-cluster while storage element files should be fetched from the storage element as this in general should be faster.
+If candidate-cluster succeeds in fetching the files the job is started at candidate-cluster now termed new-cluster.
+
+Now the job on old-cluster should be killed.
+Two possible solutions exists.
+First solution is to let the client send a kill signal to old-cluster.
+And in the second solution the signal will be send from new-cluster.
+If the client should send the kill signal, new-cluster should return the new job-id after starting the job.
+While if new-cluster should send the signal, the job-id will be returned after sending the signal.
+In both cases, if the job cannot be killed a warning will be given.
 
@@
 \subsubsection*{Technical imlplementation:}
-
-% Comments from Balazs:
-%relate new job to the old one
-%clarify what it means to be started
 \end{document}

### 2009-02-05 doc/client/client_technical.tex
Specialised brokers documentation

@@
 JobDescription content the job's JobInnerRepresentation, and the ExecutionTarget content the candidate cluster's queue information.
-There are main elements in the JobInnerRepresentation which are used for matchmaking, these are documented in this Appendix: ...
-(every matchmaking elements and their comparison logic are there) The macthmaking working on this way: it takes every JobInnerrepresentation elements which are relevant and in an ordered way check them.
+There are main elements in the JobInnerRepresentation which are used for matchmaking, these are documented in this Appendix: \ref{sec:broker-mapping}, every matchmaking elements and their comparison logic are there.
+The macthmaking working on this way: it takes every JobInnerrepresentation elements which are relevant and in an ordered way check them.
 If a matchmaking element is not presented in the JobInnerRepresentation then it will be ignored, but if it is there than the ExecutionTarget should content its pair.} For example SessionLifeTime is in the JobInnerRepresentaion and the WorkingAreaLifeTime is in the ExecutionTarget (these are matchmaking pair).
@@
 
-The sorting method is random.
-
-TODO: finish
+The ExecutionTarget sorting based on randomization, the only thing is what happen is that the PossibTargets vector will be shuffled.
 
@@
 
-The sorting method based on the fastest cpu, we are using for this the specint2006 benchmark solution.
-
-TODO: finish
+The sorting method based on the fastest CPU, we are using for this the CINT2006 (Integer Component of SPEC CPU2006) benchmark solution, here are more information about this benchmark type: http://www.spec.org/cpu2006/CINT2006/
+
+These is the sorting algorithm for this Broker:
+
+\begin{enumerate}
+\item{There is a cluster filtering at the begining of the sorting. If the cluster is not publish the specint2006 benchmark information then it will be ignored.}
+\item{That cluster will be the first in the PossibleTargets vector which has the highest specint2006 value.}
+\end{enumerate}
 
@@
 
-The sorting method based on the shortest queue, where the waiting jobs number is low.
-
-TODO: finish
+The sorting method based on the shortest queue, where the waiting jobs number is the lowest.
+
+These is the sorting algorithm for this Broker:
+
+\begin{enumerate}
+\item{That clusters will be removed which are not content the target sorting information: WaitingJobs, TotalSlots and FreeSlots.}
+\item{Sort the targets according to the number of waiting jobs (in percent of the cluster size).
+That cluster will be better where the WaitingJobs number division with the TotalSlots give lower number.
+This solution will result that the scale queue will be independent from the cluster size.}
+\item{If several clusters(queues) have free slots (CPUs) do basic load balancing}
+\end{enumerate}
 
@@
 
-The sorting method based on the A-REX file cache. I will describe how does it working.
-
-TODO: finish
+The main idea was that the jobs should submit to that cluster where the data is.
+The sorting method based on the A-REX file cache checking.
+There is a CacheCheck interface inside the A-REX which can be use to query that files present or not in the cache directory.
+There is a limitation at the current state because the CacheCheck interface can use only this cache directory format:
+
+\begin{shaded}
+\begin{verbatim}
+cachedir="/tmp/cache"
+\end{verbatim}
+\end{shaded}
+
+If the cachedir parameter content {\%}U or other substitute element then the CacheCheck will not working.
+
+These is the sorting algorithm for this Broker:
+
+\begin{enumerate}
+\item{Only the A-REX service has CacheCheck method therefore the other clusters will be ignored}
+\item{The files URL collected from the JobInnerRepresentation }
+\item{Every ExecutionTargets which are A-REXs will be called its  CacheCheck method, there is only one query per service, with this it will be not increment its load too.
+The Broker will summarize the file sizes.
+The CacheMappingTable is an array where the A-REX's URL is the key and the summarized file size is the value.
+When there is some problem to query file size from the A-REX then the summarized size will be zero.}
+\item{The last step is to sort the PossibleTargets vector in descending way  (CacheMappingTable used for this), which cluster content the most data that will be at the front of the vector.}
+\end{enumerate}
+
+Example for CacheCheck request what can be sent to an A-REX service:
+
+\begin{shaded}
+\begin{verbatim}
+<CacheCheck>
+   <TheseFilesNeedToCheck>
+       <FileURL>http://knowarc1.grid.niif.hu/storage/Makefile</FileURL>
+       <FileURL>ftp://download.nordugrid.org/test/README</FileURL>
+   </TheseFilesNeedToCheck>
+</CacheCheck>
+\end{verbatim}
+\end{shaded}
+
+Example CacheCheck response of the A-REX service:
+
+\begin{shaded}
+\begin{verbatim}
+<CacheCheckResponse>
+   <CacheCheckResult>
+   <Result>
+         <FileURL>http://knowarc1.grid.niif.hu/storage/Makefile</FileURL>
+         <ExistInTheCache>true</ExistInTheCache>
+         <FileSize>30</FileSize>
+   </Result>
+   <Result>
+         <FileURL>http://knowarc1.grid.niif.hu/storage/Makefile</FileURL>
+         <ExistInTheCache>true</ExistInTheCache>
+         <FileSize>30</FileSize>
+   </Result>
+  </CacheCheckResult>
+</CacheCheckResponse>
+\end{verbatim}
+\end{shaded}
 
 \subsubsection{PythonBroker}
+
+Mattias
 
@@
 
+\chapter{{\Broker}}
+\label{app:broker}
+
+\section{Broker mapping}
+\label{sec:broker-mapping}
+
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/client/broker-mapping.ods
+
+\section{JobInnerrepresentation}
+\label{sec:jobinnerrepresentation}
+
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/client/JobInnerRepresentation.ods
+
 \bibliography{grid}

### 2009-02-09 doc/client/client_technical.tex
Python broker documentation

@@
 
-Mattias
-
-TODO: finish
+This broker allows users to write their customized broker in
+python. To use this broker the user should write a python class. The
+class should contain:
+
+\begin{itemize}
+
+\item{an \_\_init\_\_ method that takes a Config object as input, and}
+\item{a SortTargets method that takes a python list of ExecutionTarget
+  objects and a JobInnerRepresentation object as input.}
+
+\end{itemize}
+
+The Config, ExecutionTarget and JobInnerRepresentation classes are
+available in the swig generated arc python module.
+
+To invoke the python broker, the name of the python module defining
+the broker class and the name of the broker class must be given. If
+e.g. the broker class MyBroker is defined in the python module
+SampleBroker the command line option to arcsub to use this broker is:
+
+\begin{shaded}
+\begin{verbatim}
+-b PythonBroker:SampleBroker.MyBroker
+\end{verbatim}
+\end{shaded}
+
+Additional arguments to the python broker can be added by appending
+them after an additional colon after the python class name:
+
+\begin{shaded}
+\begin{verbatim}
+-b PythonBroker:SampleBroker.MyBroker:args
+\end{verbatim}
+\end{shaded}
+
+Extracting these additional arguments should be done in the python
+broker class's \_\_init\_\_ method.
+
+For a complete example of a simple python broker see the
+SampleBroker.py file that comes with your arc python installation.
 

### 2009-02-10 doc/client/client_technical.tex
grammatical correction

@@
 {\JobDescription} class. The inner representation (\texttt{JobInnerRepresentation}
-class) is like to the GIN-JSDL and all elements are correspond to
-variables. This variables are simple variables or complex data structures,
-and private on the {\JobDescription} class.
+class) is similar to the GIN-JSDL and all elements correspond to
+variables. These variables are either simple variables or complex data structures,
+and are private in the {\JobDescription} class.
 
@@
 JSDL~\cite{jsdl}), converts and stores it in a GIN-JSDL-like internal job
-description format. The identification of the source's description
+description format. The identification of the source description
 format is internally handled by a \texttt{JobDescriptionOrderer} class
 which identifies the format by pattern matching. If the source string is
-empty or the \texttt{JobDescriptionOrderer} class can not be recognize
-to any formats, then it is not store nothing and return with false, else
-convert and store it.
+empty or the \texttt{JobDescriptionOrderer} class can not be recognized
+to any formats, then it does not store anything and returns with false, otherwise
+it converts and stores it.
 
 Different operations, like getting the job description in other
-formats or getting job-related information, can be performed using the
+formats or getting job-related information, can be performed by using the
 description-independent functions of this class, e.g.:
@@
 three supported job description languages. The {\JobDescription} class
-chooses the appropriate back-end module according to the pattern match
+chooses the appropriate back-end module according to the pattern matching
 performed by the \texttt{JobDescriptionOrderer}, and uses the back-end
@@
 If the inner representation is empty or the source job descrition is not
-valid, then the output is empty and return with false.
-When the source format is equal with the output's format and it is not XRSL,
-than set the product string with the source string. The XRSL output is
-always generation from the inner representation.
+valid, then the output is empty too and returns with false.
+When the source format is the same as the output format and it is not XRSL,
+than we set the product string to the source string. The XRSL output is
+always generated from the inner representation.
 
 Parsing is currently possible from any of the languages. JSDL, XRSL and
-JDL output generation is implemented yet and even for these there are
+JDL output generation is implemented already but even for these there are
 some limitations.
@@
 The conversion mainly means a syntactical transformation and the
-{\JobDescription} class can of course do nothing in those cases where
-not enough data to assemble a given attribute is available or when
-information can be lost because the received attribute has no
+{\JobDescription} class can do nothing in those cases where
+there are not enough data is available to assemble a given attribute or
+when information can be lost because the received attribute has no
 equivalent in the output language.
 
-For the JSDL output generation the {\JobDescription} class is using
+For the JSDL output generation the {\JobDescription} class uses
 the core JSDL's capabilities amended with the JSDL-POSIX and JSDL-ARC

### 2009-02-10 doc/client/client_technical.tex
Broker text correction

@@
 It is the base class of the specialized Brokers.
-If you are using the arcsub command then you can choose a broker type with ``-b`` parameter.
-There is a solution for add parameters to these descendants Brokers.
+If you are using command arcsub then you can choose a broker type with with ``-b`` parameter.
+There is a solution for adding parameters to these descendants Brokers.
 Here is an example:
@@
 
-The Broker's constructor set the PreFilteringDone and the TargetSortingDone values to false because we need to call at first the PreFilterTargets and second the GetBestTarget which trigger the specialised Broker sorting method.
+The Broker's constructor sets the PreFilteringDone and the TargetSortingDone values to false because we need to call the PreFilterTargets first then the second the GetBestTarget which trigger the specialised Broker sorting method.
 
@@
 \item{The first step is the PreFilterTargets, this implements the matchmaking: the JobDescription and the TargetGenerator's ExecutionTargets will be compared inside this method.
-JobDescription content the job's JobInnerRepresentation, and the ExecutionTarget content the candidate cluster's queue information.
-There are main elements in the JobInnerRepresentation which are used for matchmaking, these are documented in this Appendix: \ref{sec:broker-mapping}, every matchmaking elements and their comparison logic are there.
-The macthmaking working on this way: it takes every JobInnerrepresentation elements which are relevant and in an ordered way check them.
-If a matchmaking element is not presented in the JobInnerRepresentation then it will be ignored, but if it is there than the ExecutionTarget should content its pair.} For example SessionLifeTime is in the JobInnerRepresentaion and the WorkingAreaLifeTime is in the ExecutionTarget (these are matchmaking pair).
-If the WorkingAreaLifeTime is not defined then the Broker cannot check the times therefore it will ignore this cluster, but if the WorkingAreaLifeTime exist and the comparison between this elements are true then this check will be passed other way the cluster will be ignored.
-We are using this rule for every elements checking.
-There is only one exception and this is the ProcessingStartTime because if the DownTime is not defined in the ExecutionTarget than we can trust by it that the job can anytime start and it will be not killed with shutdown.
-This rule checking runs for every ExecutionTargets and if every matchmaking checks are ok for a target then this ExecutionTarget will be added to the PossibleTargets vector.
-\item{If the matchmaking finished then  the PreFilteringDone variable will be true and the TargetSortingDone will be false. }
-\item{The next step is the GetBestTarget method calling, this will give an ExecutionTarget which is good for our job.
-If the Possibletarget vector empty then it will give nothing.
-The first time the GetBestTarget will trigger the specialid Broker type sorting with this method calling:  SortTargets.
-This SortTargets method is virtual method, therefore every specialised Broker should implement it.}
-\item{If the SortTargets finished then the TargetSortingDone will be true and the GetBestTarget will return the first element of the PossibleTargetVector and the current iterator will be incremented with one.}
-\item{If the submision failed to this cluster then we can call again the GetBestTarget and we will get an other ExecutionTarget till it will not consume.
-If  the targets run out from the PossibleTargetVector then the GetBestTarget's bool parameter will be true, its name is EndOfList''}
+JobDescription encapsulate the job's JobInnerRepresentation, and the ExecutionTarget involves the candidate cluster's queue information.
+There are main elements in the JobInnerRepresentation which are used for matchmaking, these are documented in  Appendix: \ref{sec:broker-mapping}, all matchmaking elements and their comparison logic are there.
+The macthmaking works the following way: it takes every JobInnerrepresentation elements which are relevant and  check them in an ordered way.
+If a matchmaking element is not presented in the JobInnerRepresentation then it will be ignored, but if it is there than the ExecutionTarget should compare.
+For example SessionLifeTime is in the JobInnerRepresentaion and the WorkingAreaLifeTime is in the ExecutionTarget (these are matchmaking pair).
+If the WorkingAreaLifeTime is not defined then the Broker cannot check the times therefore it will ignore this cluster, but if the WorkingAreaLifeTime exist and the comparison between this elements are true then this check will be passed otherwise the cluster will be ignored.
+We are using this rule for checking every elements.
+There is only one exception that is the ProcessingStartTime.
+If the DownTime is not defined in the ExecutionTarget than we can believe it that the job can start anytime and it will not be killed by shutdown.
+This rule checking runs for every ExecutionTargets and if every matchmaking checks are successfull for a target then this ExecutionTarget will be added to the PossibleTargets vector.}
+\item{When the matchmaking has finished then  the PreFilteringDone variable will be set to true and the TargetSortingDone will be set to false. }
+\item{The next step is calling the GetBestTarget method, which will give an ExecutionTarget candidate.
+If the PossibleTargets vector is empty then it will give nothing.
+The GetBestTarget will call the specialid Broker type sorting  method.
+This SortTargets method is a virtual method, therefore every specialised Broker should implement it.}
+\item{When the SortTargets has finished then the TargetSortingDone will be true and the GetBestTarget will return the first element of the PossibleTargetVector and the current iterator will be incremented by one.}
+\item{If the submision fails to this cluster then we can call the GetBestTarget again and we will get an other ExecutionTarget till it will not consumes.
+If  the targets run out from the PossibleTargetVector then the GetBestTarget's boolean parameter (EndOfList) will be true}
 \end{enumerate}
@@
 
-The Broker class content  two useful data elements: logger and JobInnerRepresentation.
-These can be helpful when somebody create a specialised Broker and he would like to use the logger and need some information about the job.
+The Broker class encasulated two useful data elements: logger and JobInnerRepresentation.
+These can be helpful when somebody creates a specialised Broker and wants to use the logger and needs some information about the job.
 
@@
 
-The ExecutionTarget sorting based on randomization, the only thing is what happen is that the PossibTargets vector will be shuffled.
+The ExecutionTarget sorting is based on randomization, the PossibleTargets vector will be shuffled.
 
@@
 
-The sorting method based on the fastest CPU, we are using for this the CINT2006 (Integer Component of SPEC CPU2006) benchmark solution, here are more information about this benchmark type: http://www.spec.org/cpu2006/CINT2006/
-
-These is the sorting algorithm for this Broker:
+The sorting method is based on the fastest CPU, we use the CINT2000 (Integer Component of SPEC CPU2000) benchmark for this purpose , here are more information about this benchmark type: http://www.spec.org/cpu2000/CINT2000/
+
+The sorting algorithm for this Broker is as follows:
 
 \begin{enumerate}
-\item{There is a cluster filtering at the begining of the sorting. If the cluster is not publish the specint2006 benchmark information then it will be ignored.}
-\item{That cluster will be the first in the PossibleTargets vector which has the highest specint2006 value.}
+\item{There is a cluster filtering at the begining of the sorting. If the cluster is not publish the specint2000 benchmark information then it will be ignored.}
+\item{The cluster having the highest specint2000 value will be the first in the PossibleTargets vector which has }
 \end{enumerate}
@@
 
-The sorting method based on the shortest queue, where the waiting jobs number is the lowest.
-
-These is the sorting algorithm for this Broker:
+The sorting method is based on the shortest queue, where the number waiting jobs  of  is the lowest.
+
+The sorting algorithm for this Broker is as follows:
 
 \begin{enumerate}
-\item{That clusters will be removed which are not content the target sorting information: WaitingJobs, TotalSlots and FreeSlots.}
+\item{The clusters which do not involve the target sorting information will be removed  (WaitingJobs, TotalSlots and FreeSlots).}
 \item{Sort the targets according to the number of waiting jobs (in percent of the cluster size).
-That cluster will be better where the WaitingJobs number division with the TotalSlots give lower number.
+The cluster will be better where the WaitingJobs number divided by the TotalSlots gives the lower number.
 This solution will result that the scale queue will be independent from the cluster size.}
@@
 The main idea was that the jobs should submit to that cluster where the data is.
-The sorting method based on the A-REX file cache checking.
-There is a CacheCheck interface inside the A-REX which can be use to query that files present or not in the cache directory.
-There is a limitation at the current state because the CacheCheck interface can use only this cache directory format:
+The sorting method is based on the A-REX file cache checking.
+There is a CacheCheck interface inside the A-REX which can be used to query whether files in question are present or not in the cache directory.
+There is a limitation at the current state because the CacheCheck interface can only use this cache directory format:
 
@@
 
-If the cachedir parameter content {\%}U or other substitute element then the CacheCheck will not working.
-
-These is the sorting algorithm for this Broker:
+If the cachedir parameter involves {\%}U or other substitute element then the CacheCheck will not work.
+
+The sorting algorithm for this Broker is as follows:
 
@@
 \item{Only the A-REX service has CacheCheck method therefore the other clusters will be ignored}
-\item{The files URL collected from the JobInnerRepresentation }
-\item{Every ExecutionTargets which are A-REXs will be called its  CacheCheck method, there is only one query per service, with this it will be not increment its load too.
+\item{The URL files are collected from the JobInnerRepresentation }
+\item{All of the ExecutionTargets which are A-REXs will be called by their  CacheCheck method.
+Single query used for achieving all the necessary information.
 The Broker will summarize the file sizes.
@@
 When there is some problem to query file size from the A-REX then the summarized size will be zero.}
-\item{The last step is to sort the PossibleTargets vector in descending way  (CacheMappingTable used for this), which cluster content the most data that will be at the front of the vector.}
+\item{The last step is to sort the PossibleTargets vector in descending order  (CacheMappingTable used for this), which cluster having the most data that will be at the front of the vector.}
 \end{enumerate}

### 2009-02-10 doc/client/client_technical.tex
Small fix

@@
 
-The Broker's constructor sets the PreFilteringDone and the TargetSortingDone values to false because we need to call the PreFilterTargets first then the second the GetBestTarget which trigger the specialised Broker sorting method.
+The Broker's constructor sets the PreFilteringDone and the TargetSortingDone values to false because we need to call the PreFilterTargets first then the second the GetBestTarget which triggers the specialised Broker sorting method.
 
@@
 
-The sorting method is based on the fastest CPU, we use the CINT2000 (Integer Component of SPEC CPU2000) benchmark for this purpose , here are more information about this benchmark type: http://www.spec.org/cpu2000/CINT2000/
+The sorting method is based on the fastest CPU, we use the CINT2000 (Integer Component of SPEC CPU2000) benchmark for this purpose , here are more information about this benchmark type:
+
+http://www.spec.org/cpu2000/CINT2000/
 

### 2009-02-13 doc/client/client_technical.tex
spell checking, format modifications etc

@@
 \newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
-\newcommand{\arclib}{libarcclient}
+\newcommand{\libarcclient}{libarcclient}
 \newcommand{\ACC}{\texttt{ACC}}
 \newcommand{\ACCConfig}{\texttt{ACCConfig}}
+\newcommand{\Config}{\texttt{Config}}
 \newcommand{\Broker}{\texttt{Broker}}
+\newcommand{\RandomBroker}{\texttt{RandomBroker}}
+\newcommand{\FastestCPUBroker}{\texttt{FastestCPUBroker}}
+\newcommand{\FastestQueueBroker}{\texttt{FastestQueueBroker}}
+\newcommand{\PythonBroker}{\texttt{PythonBroker}}
+\newcommand{\DataBroker}{\texttt{DataBroker}}
 \newcommand{\ClientInterface}{\texttt{ClientInterface}}
@@
 %%%%---- The title ----
-{\centering \textsc{\Large {\arclib}}\Large \par}
+{\centering \textsc{\Large {\libarcclient}}\Large \par}
 \vspace*{0.5cm}
@@
 \item{performing resource discovery and information retrieval,}
+\item{filtering and brokering of found resources,}
 \item{job submission and management and}
@@
 
-The new {\arclib} makes extensive use of plugins for command
+The new {\libarcclient} makes extensive use of plugins for command
 handling. These plugins are handled by a set of higher level classes
@@
 capability is of crucial importance for a user interface. The new
-{\arclib} resource discovery and information retrieval component
+{\libarcclient} resource discovery and information retrieval component
 consists of three classes; the {\TargetGenerator}, the
@@
   the {\JobDescription} is passed as input to the brokering process.}
-\item{The {\Broker} outputs a ordered list of {\ExecutionTarget}s
-  according to the provided {\JobDescription}.}
+\item{The {\Broker} filters and ranks the {\ExecutionTarget}s according 
+  to the ranking method chosen by the user.}
 \item{Each {\ExecutionTarget} has a method to return a specialized
@@
 Once a job is submitted job related information (job identification
-string, cluster etc) is stored in a local XML file which hosts similar
+string, cluster etc) is stored in a local XML file which stores this 
 information for all active jobs. This file may contain jobs running on
@@
   the joblist file is also used to look up the job identifier for jobs
-  specified using job names. Alias resolution for the selected and
+  specified using job names. Alias resolving for the selected and
   rejected clusters are performed using the information in the
@@
 In this section an overview of all important classes in the new
-{\arclib} is given. The overview is subdivided in two parts where
+{\libarcclient} is given. The overview is subdivided in two parts where
 first all the generic classes are presented,
@@
 resource (queue) capable of executing a grid job. It serves as input
-to the {\Broker} which is foreseen to be able to select between
-different {\ExecutionTarget}s from different grid flavours without a
+to the {\Broker} which is able to filter and rank different 
+{\ExecutionTarget}s from different grid flavours without a
 priori knowing their difference. The {\ExecutionTarget} class mimics
@@
 
-Following a broker decision jobs are to be submitted. Since all
+Following a broker ranking jobs are to be submitted. Since all
 information about the selected computing service resides within the
@@
 
-The Broker inherits from the {\ACC} base class in order to be loadable by the ARC \texttt{Loader}.
-It is the base class of the specialized Brokers.
-If you are using command arcsub then you can choose a broker type with with ``-b`` parameter.
-There is a solution for adding parameters to these descendants Brokers.
-Here is an example:
-
-\begin{shaded}
-\begin{verbatim}
-arcsub -b RandomBroker:a,b,c -c ARC1:https://knowarc1.grid.niif.hu:60000/arex job.jsdl
-\end{verbatim}
-\end{shaded}
-
-The Broker's constructor sets the PreFilteringDone and the TargetSortingDone values to false because we need to call the PreFilterTargets first then the second the GetBestTarget which triggers the specialised Broker sorting method.
-
-The base Broker operations are the followings:
-
-\begin{enumerate}
-\item{The first step is the PreFilterTargets, this implements the matchmaking: the JobDescription and the TargetGenerator's ExecutionTargets will be compared inside this method.
-JobDescription encapsulate the job's JobInnerRepresentation, and the ExecutionTarget involves the candidate cluster's queue information.
-There are main elements in the JobInnerRepresentation which are used for matchmaking, these are documented in  Appendix: \ref{sec:broker-mapping}, all matchmaking elements and their comparison logic are there.
-The macthmaking works the following way: it takes every JobInnerrepresentation elements which are relevant and  check them in an ordered way.
-If a matchmaking element is not presented in the JobInnerRepresentation then it will be ignored, but if it is there than the ExecutionTarget should compare.
-For example SessionLifeTime is in the JobInnerRepresentaion and the WorkingAreaLifeTime is in the ExecutionTarget (these are matchmaking pair).
-If the WorkingAreaLifeTime is not defined then the Broker cannot check the times therefore it will ignore this cluster, but if the WorkingAreaLifeTime exist and the comparison between this elements are true then this check will be passed otherwise the cluster will be ignored.
-We are using this rule for checking every elements.
-There is only one exception that is the ProcessingStartTime.
-If the DownTime is not defined in the ExecutionTarget than we can believe it that the job can start anytime and it will not be killed by shutdown.
-This rule checking runs for every ExecutionTargets and if every matchmaking checks are successfull for a target then this ExecutionTarget will be added to the PossibleTargets vector.}
-\item{When the matchmaking has finished then  the PreFilteringDone variable will be set to true and the TargetSortingDone will be set to false. }
-\item{The next step is calling the GetBestTarget method, which will give an ExecutionTarget candidate.
-If the PossibleTargets vector is empty then it will give nothing.
-The GetBestTarget will call the specialid Broker type sorting  method.
-This SortTargets method is a virtual method, therefore every specialised Broker should implement it.}
-\item{When the SortTargets has finished then the TargetSortingDone will be true and the GetBestTarget will return the first element of the PossibleTargetVector and the current iterator will be incremented by one.}
-\item{If the submision fails to this cluster then we can call the GetBestTarget again and we will get an other ExecutionTarget till it will not consumes.
-If  the targets run out from the PossibleTargetVector then the GetBestTarget's boolean parameter (EndOfList) will be true}
-\end{enumerate}
-
-\begin{shaded}
-\begin{verbatim}
-class Broker : public ACC {
-public:
+The {\Broker} inherits from the {\ACC} base class in order to be loadable by the ARC \texttt{Loader}. It is the base class of the 
+specialized {\Broker} and it implements the method for reducing a list of resources found by resource discovery (the list of 
+{\ExecutionTarget}s residing witin the {\TargetGenerator}) to a list of resources capable of running a certain job:
+
+\begin{shaded}
+\begin{verbatim}
+void PreFilterTargets(Arc::TargetGenerator& targen, Arc::JobDescription jd);
+\end{verbatim}
+\end{shaded}
+
+The {\texttt PreFilterTargets} method compares every hardware and software requirement given in the job description against the 
+computing resource (cluster) specifications stored in the {\ExecutionTarget}. If the{\ExecutionTarget} doesn't fulfill 
+the requirements imposed by the job description, or it is impossible to carry out the matchmaking due to incomplete information 
+published by the computing resource, the {\ExecutionTarget} will be removed from the list of possible targets. A detailed overview 
+of the matchmaking between {\JobDescription} and {\ExecutionTarget} attributes is given in Appendix~\ref{sec:broker-mapping}.
+
+Once prefiltered the remaining {\ExecutionTarget}s should be ranked in order to return the ``best'' {\ExecutionTarget} for 
+job submission. Different ranking methods are implemented by the specialized brokers, but for usability and consistency reasons 
+these methods are encapsulated by the {\Broker} base class method
+
+\begin{shaded}
+\begin{verbatim}
 ExecutionTarget& GetBestTarget(bool &EndOfList);
-    void PreFilterTargets(Arc::TargetGenerator& targen, Arc::JobDescription jd);    
-protected:
-    Broker(Config *cfg);
-    virtual ~Broker();
-    virtual void SortTargets() = 0;
-    std::vector<Arc::ExecutionTarget> PossibleTargets;
-    bool TargetSortingDone;
-    Arc::JobInnerRepresentation jir;
-    static Logger logger;
-private:
-    std::vector<Arc::ExecutionTarget>::iterator current;
-    bool PreFilteringDone;
-};
-\end{verbatim}
-\end{shaded}
-
-The Broker class encasulated two useful data elements: logger and JobInnerRepresentation.
-These can be helpful when somebody creates a specialised Broker and wants to use the logger and needs some information about the job.
+\end{verbatim}
+\end{shaded}
+
+which invokes the {\texttt SortTargets} method implemented by the specialized broker (see Section~\ref{sec:brokerplugins}) 
+and returns the best target. The {\texttt GetBestTarget} method is ``incremented'' at each call, thus upon a second call 
+the second best {\ExecutionTarget} will be returned.
 
@@
 description languages by translating them automatically. In the
-{\arclib} library this functionality is implemented in the
-{\JobDescription} class. The inner representation (\texttt{JobInnerRepresentation}
-class) is similar to the GIN-JSDL and all elements correspond to
-variables. These variables are either simple variables or complex data structures,
-and are private in the {\JobDescription} class.
-
-This is a generic class that takes a job
-description (string) as input
+{\libarcclient} library this functionality is implemented in the
+{\JobDescription} class which parses job descriptions given in the 
+XRSL~\cite{xrsl}, JDL~\cite{jdl} or JSDL~\cite{jsdl} formats into an 
+internal representation based upon the upcoming PGI-JSDL proposal.
 
@@
 
-in any supported formats (currently XRSL~\cite{xrsl}, JDL~\cite{jdl},
-JSDL~\cite{jsdl}), converts and stores it in a GIN-JSDL-like internal job
-description format. The identification of the source description
-format is internally handled by a \texttt{JobDescriptionOrderer} class
-which identifies the format by pattern matching. If the source string is
-empty or the \texttt{JobDescriptionOrderer} class can not be recognized
-to any formats, then it does not store anything and returns with false, otherwise
-it converts and stores it.
-
-Different operations, like getting the job description in other
-formats or getting job-related information, can be performed by using the
-description-independent functions of this class, e.g.:
-
-\begin{shaded}
-\begin{verbatim}
-bool getProduct(std::string& product, std::string format = "POSIXJSDL")
-\end{verbatim}
-\end{shaded}
+The identification of the source description format is internally handled by a 
+\texttt{JobDescriptionOrderer} class which identifies the format by pattern matching.
+Once the format is determined from the elements are mapped from the original format (XRSL, JDL or JSDL) 
+into the internal representation (the \texttt{JobInnerRepresentation} class). Mapping details 
+are given in Section~\ref{sec:jobinnerrepresentation}.
 
@@
 performed by the \texttt{JobDescriptionOrderer}, and uses the back-end
-module for parsing and generating the job descriptions.
+module for parsing and translation (if needed) of the job descriptions.
+
+Once the job description language required by the {\ExecutionTarget} chosen 
+for job submission is known the parsed and (if needed) translated job description 
+can be retrieved through the {\texttt GetProduct} funcion as shown below:
+
+\begin{shaded}
+\begin{verbatim}
+bool getProduct(std::string& product, std::string format = "POSIXJSDL")
+\end{verbatim}
+\end{shaded}
+
 If the inner representation is empty or the source job descrition is not
-valid, then the output is empty too and returns with false.
-When the source format is the same as the output format and it is not XRSL,
-than we set the product string to the source string. The XRSL output is
-always generated from the inner representation.
-
-Parsing is currently possible from any of the languages. JSDL, XRSL and
-JDL output generation is implemented already but even for these there are
-some limitations.
-
-The conversion mainly means a syntactical transformation and the
-{\JobDescription} class can do nothing in those cases where
-there are not enough data is available to assemble a given attribute or
-when information can be lost because the received attribute has no
-equivalent in the output language.
+valid, then the {\texttt GetProduct} output is empty too and the method 
+returns with false. When the product format is the same as the source format 
+and not XRSL, than the product string is set to the source string. The XRSL 
+output is always generated from the inner representation, as it includes 
+automatic job description modifications similar to those of the production 
+ARC client. If the product format differs from the source format the 
+{\texttt GetProduct} method calls upon the \texttt{JobDescriptionOrderer} 
+class to carry out the translation.
+
+The translation of a job description basically means a syntactical transformation 
+of the job description from one language to another. Thus the {\JobDescription}
+class can do nothing in those cases where there is not enough data is available 
+to assemble a given attribute or when information can be lost because the received 
+attribute has no equivalent in the output language. In these events the {\texttt GetProduct} 
+method returns with a false and an empty product stringt. For limitations in the translation
+see Section~\ref{sec:jobinnerrepresentation}.
 
@@
 and thus a mapping is needed for non Glue2 compliant grid middlewares.
-Appendix \ref{app:job} shows the present mapping schema.
+Appendix~\ref{app:job} shows the present mapping schema.
 
@@
 CREAM CE~\cite{cream}. The current revision of this interface (CREAM
-version 2) has been chosen for implementation within {\arclib} for the
+version 2) has been chosen for implementation within {\libarcclient} for the
 following reasons:
@@
 \subsection{Broker plugins}
-
-\subsubsection{RandomBroker}
-
-The ExecutionTarget sorting is based on randomization, the PossibleTargets vector will be shuffled.
-
-\subsubsection{FastestCPUBroker}
-
-The sorting method is based on the fastest CPU, we use the CINT2000 (Integer Component of SPEC CPU2000) benchmark for this purpose , here are more information about this benchmark type:
-
-http://www.spec.org/cpu2000/CINT2000/
-
-The sorting algorithm for this Broker is as follows:
+\label{sec:brokerplugins}
+
+\subsubsection{\RandomBroker}
+
+The {\RandomBroker} ranks the {\ExecutionTarget}s randomly.
+
+\subsubsection{\FastestCPUBroker}
+
+The {\FastestCPUBroker} ranks the {\ExecutionTarget}s according to their CPU performance. The most 
+reliable comparison is achieved through the usage of a benchmark scenario, and the {\FastestCPUBroker}
+uses the CINT2000 (Integer Component of SPEC CPU2000)\footnote{http://www.spec.org/cpu2000/CINT2000/} 
+benchmark for this purpose.
+
+The {\texttt SortTargets} method has two steps:
 
 \begin{enumerate}
-\item{There is a cluster filtering at the begining of the sorting. If the cluster is not publish the specint2000 benchmark information then it will be ignored.}
-\item{The cluster having the highest specint2000 value will be the first in the PossibleTargets vector which has }
+\item{{\ExecutionTarget}s not publishing the specint2000 benchmark information are removed from the list 
+  of possible targets as their performance is unknown.}
+\item{The {\ExecutionTarget}s are ranked such that the {\ExecutionTarget} with highest specint2000 value 
+  receives top ranking}
 \end{enumerate}
 
-\subsubsection{FastestQueueBroker}
-
-The sorting method is based on the shortest queue, where the number waiting jobs  of  is the lowest.
-
-The sorting algorithm for this Broker is as follows:
+\subsubsection{\FastestQueueBroker}
+
+The {\FastestQueueBroker} ranks the {\ExecutionTarget}s according to their queue length measured in percentage 
+of the {\ExecutionTarget}'s size (i.e. the queue length divided by the number of total slots/CPUs). If more 
+than one {\ExecutionTarget} has zero queue, the {\FastestQueueBroker} makes use of a basic load balancing method 
+to rearrange the zero queue {\ExecutionTarget}s depending on their number of free slots/CPUs.
+
+The {\texttt SortTargets} method has three steps:
 
 \begin{enumerate}
-\item{The clusters which do not involve the target sorting information will be removed  (WaitingJobs, TotalSlots and FreeSlots).}
-\item{Sort the targets according to the number of waiting jobs (in percent of the cluster size).
-The cluster will be better where the WaitingJobs number divided by the TotalSlots gives the lower number.
-This solution will result that the scale queue will be independent from the cluster size.}
-\item{If several clusters(queues) have free slots (CPUs) do basic load balancing}
+\item{{\ExecutionTarget}s not publishing WaitingJobs, TotalSlots and FreeSlots are removed from the list of possible targets 
+as it is impossible to determine their queue length and perform load balancing.}
+\item{The {\ExecutionTarget}s are ranked according to their queue length relative to the number of slots/CPUs. Zero queue 
+{\ExecutionTarget}s is ranked top.}
+\item{If more than one {\ExecutionTarget} has zero queue the targets are randomly reshuffled based upon how many free slots 
+they have. E.g. if two {\ExecutionTarget}s have zero queue and they have 90 and 10 free slots respectively, then the first
+{\ExecutionTarget} has a 90\% propability of becoming the top ranked cluster.}
 \end{enumerate}
 
-\subsubsection{DataBroker}
-
-The main idea was that the jobs should submit to that cluster where the data is.
-The sorting method is based on the A-REX file cache checking.
-There is a CacheCheck interface inside the A-REX which can be used to query whether files in question are present or not in the cache directory.
-There is a limitation at the current state because the CacheCheck interface can only use this cache directory format:
+\subsubsection{\DataBroker}
+
+The {\DataBroker} ranks the {\ExecutionTarget}s according to how many megabytes of the requested input files that already 
+stored in the cache of the computing resource the {\ExecutionTarget} represents. The broker is motivated by that jobs should 
+be submitted to {\ExecutionTarget}s where the data already is, thus reducing the network load on both the computing resource 
+and client side. The ranking method is based upon the A-REX\footnote{http://www.knowarc.eu/download/D1.2-3\_documentation.pdf} 
+interface {\texttt CacheCheck} for querying for the presence of the file in the cache directory. This interface has a limitation 
+in the current implementation as it can only use cache directory of format
 
@@
 
-If the cachedir parameter involves {\%}U or other substitute element then the CacheCheck will not work.
-
-The sorting algorithm for this Broker is as follows:
+If the cachedir parameter involves {\%}U or other substitute element then the {\texttt CacheCheck} will not work.
+
+The {\texttt SortTargets} method has four steps:
 
 \begin{enumerate}
-\item{Only the A-REX service has CacheCheck method therefore the other clusters will be ignored}
-\item{The URL files are collected from the JobInnerRepresentation }
-\item{All of the ExecutionTargets which are A-REXs will be called by their  CacheCheck method.
-Single query used for achieving all the necessary information.
-The Broker will summarize the file sizes.
-The CacheMappingTable is an array where the A-REX's URL is the key and the summarized file size is the value.
-When there is some problem to query file size from the A-REX then the summarized size will be zero.}
-\item{The last step is to sort the PossibleTargets vector in descending order  (CacheMappingTable used for this), which cluster having the most data that will be at the front of the vector.}
+\item{Only the A-REX service has {\texttt CacheCheck} method, thus {\ExecutionTarget}s not running A-REX are removed.
+therefore the other clusters will be ignored.}
+\item{Information about input files is collected from the JobInnerRepresentation.}
+\item{Each {\ExecutionTarget} is queried (through the {\texttt CacheCheck} method) for the existence of input files. 
+A single query is used for achieving all the necessary information and file sizes are summarized. If there is problems 
+determining file sizes, then the summarized size will be zero.}
+\item{The possible {\ExecutionTarget}s are ranked in a descending order according to the amount of input data they have in cache.}
 \end{enumerate}
 
-Example for CacheCheck request what can be sent to an A-REX service:
+Example of a CacheCheck request that can be sent to an A-REX service:
 
@@
 
-Example CacheCheck response of the A-REX service:
+Example CacheCheck response from the A-REX service:
 
@@
 
-\subsubsection{PythonBroker}
-
-This broker allows users to write their customized broker in
-python. To use this broker the user should write a python class. The
-class should contain:
+\subsubsection{\PythonBroker}
+
+This {\PythonBroker} allows users to write their customized broker in
+python. To use this broker the user should write a python class which 
+should contain:
 
@@
 
-The Config, ExecutionTarget and JobInnerRepresentation classes are
+The {\Config}, {\ExecutionTarget} and {\texttt JobInnerRepresentation} classes are
 available in the swig generated arc python module.
 
-To invoke the python broker, the name of the python module defining
+To invoke the {\PythonBroker}, the name of the python module defining
 the broker class and the name of the broker class must be given. If
@@
 command line interfaces (cli) for grid job submission and
-management. {\arclib} offers the following native commands:
+management. {\libarcclient} offers the following native commands:
 
@@
 
-\section{Domain and Location attributes}
-
-\subsection*{std::string DomainName}
-
-ARC0: nordugrid-cluster-name \\
-CREAM: GlueSiteName
-
-\subsection*{std::string Owner}
-
-ARC0: nordugrid-cluster-owner
-
-\subsection*{std::string Address}
-
-\subsection*{std::string Place}
-
-CREAM: GlueSiteLocation
-
-\subsection*{std::string PostCode}
-
-ARC0: nordugrid-cluster-location
-
-\subsection*{float Latitude}
-
-CREAM: GlueSiteLatitude
-
-\subsection*{float Longitude}
-
-CREAM: GlueSiteLongitude
-
-\section{ComputingService and ComputingEndpoint attributes}
-
-\subsection*{std::string CEID}
-
-\subsection*{std::string CEName}
-
-\subsection*{std::string Capability}
-
-\subsection*{std::string Type}
-
-\subsection*{std::string QualityLevel}
-
-\subsection*{URL url}
-
-ARC0: nordugrid-cluster-contactstring \\
-CREAM: GlueCEInfoContactString
-
-\subsection*{std::string Technology}
-
-\subsection*{std::string Interface}
-
-ARC0: ``GridFTP''
-
-\subsection*{std::string InterfaceExtension}
-
-\subsection*{std::string SupportedProfile}
-
-\subsection*{std::string Implementor}
-
-ARC0: ``NorduGrid''
-
-\subsection*{std::string ImplementationName}
-
-ARC0: ``ARC0'' \\
-CREAM: GlueCEImplementationName
-
-\subsection*{std::string ImplementationVersion}
-
-ARC0: nordugrid-cluster-middleware \\
-CREAM: GlueCEImplementaionVersion
-
-\subsection*{std::string HealthState}
-
-ARC0: nordugrid-queue-status
-
-\subsection*{std::string ServingState}
-
-CREAM: GlueCEStateStatus (GlueVOView $||$ GlueCE)
-
-\subsection*{std::string IssuerCA}
-
-ARC0: nordugrid-cluster-issuerca (nordugrid-cluster-issuerca-hash)
-
-\subsection*{std::list<std::string> TrustedCA}
-
-ARC0: nordugrid-cluster-trustedca
-
-\subsection*{Time DowntimeStarts}
-
-\subsection*{Time DowntimeEnds}
-
-\subsection*{std::string Staging}
-
-ARC0: nordugrid-cluster-nodeaccess
-
-\subsection*{std::string Jobdescription}
-
-\section{ComputingService and ComputingShare load attributes}
-
-\subsection*{int TotalJobs}
-
-ARC0: nordugrid-cluster-totaljobs \\
-CREAM: GlueCEStateTotalJobs (GlueVOView $||$ GlueCE)
-
-\subsection*{int RunningJobs}
-
-ARC0: nordugrid-queue-running \\
-CREAM: GlueCEStateRunningJobs (GlueVOView $||$ GlueCE)
-
-\subsection*{int WaitingJobs}
-
-ARC0: nordugrid-queue-queued $||$ nordugrid-cluster-queuedjobs \\
-CREAM: GlueCEStateWaitingJobs (GlueVOView $||$ GlueCE)
-
-\subsection*{int StagingJobs}
-
-\subsection*{int SuspendedJobs}
-
-\subsection*{int PreLRMSWaitingJobs}
-
-ARC0: nordugrid-queue-prelrmsqueued
-
-\subsection*{int LocalRunningJobs}
-
-ARC0: nordugrid-queue-running $-$ nordugrid-queue-gridrunning
-
-\subsection*{int LocalWaitingJobs}
-
-ARC0: nordugrid-queue-queued $-$ nordugrid-queue-gridqueue
-
-\subsection*{int FreeSlots}
-
-CREAM: GlueCEStateFreeJobSlots $||$ GlueCEStateFreeCPUs (GlueVOView $||$ GlueCE)
-
-\subsection*{std::string FreeSlotsWithDuration}
-
-\subsection*{int UsedSlots}
-
-ARC0: nordugrid-queue-usedcpus
-
-\subsection*{int RequestedSlots}
-
-\section{ComputingShare attributes}
-
-\subsection*{std::string MappingQueue}
-
-ARC0: nordugrid-queue-name
-
-\subsection*{Period MaxWallTime}
-
-ARC0: nordugrid-queue-maxwalltime \\
-CREAM: GlueCEPolicyMaxWallClockTime (GlueVOView $||$ GlueCE)
-
-\subsection*{Period MaxTotalWallTime}
-
-\subsection*{Period MinWallTime}
-
-ARC0: nordugrid-queue-minwalltime
-
-\subsection*{Period DefaultWallTime}
-
-ARC0: nordugrid-queue-defaultwalltime
-
-\subsection*{Period MaxCPUTime}
-
-ARC0: nordugrid-queue-maxcputime \\
-CREAM: GlueCEPolicyMaxCPUTime (GlueVOView $||$ GlueCE)
-
-\subsection*{Period MaxTotalCPUTime}
-
-\subsection*{Period MinCPUTime}
-
-ARC0: nordugrid-queue-defaultcputime
-
-\subsection*{Period DefaultCPUTime}
-
-ARC0: nordugrid-queue-defaultcputime
-
-\subsection*{int MaxTotalJobs}
-
-CREAM: GlueCEPolicyMaxTotalJobs (GlueVOView $||$ GlueCE)
-
-\subsection*{int MaxRunningJobs}
-
-ARC0: nordugrid-queue-maxrunning \\
-CREAM: GlueCEPolicyMaxRunningJobs (GlueVOView $||$ GlueCE)
-
-\subsection*{int MaxWaitingJobs}
-
-ARC0: nordugrid-queue-maxqueable \\
-CREAM: GlueCEPolicyMaxWaitingJobs (GlueVOView $||$ GlueCE)
-
-\subsection*{int NodeMemory}
-
-ARC0: nordugrid-queue-nodememory $||$ nordugrid-cluster-nodememory \\
-CREAM: GlueHostMainMemoryRAMSize
-
-\subsection*{int MaxPreLRMSWaitingJobs}
-
-\subsection*{int MaxUserRunningJobs}
-
-ARC0: nordugrid-queue-maxuserrun \\
-CREAM: GlueCEPolicyAssignedJobSlots (GlueVOView $||$ GlueCE) % ??
-
-\subsection*{int MaxSlotsPerJob}
-
-CREAM: GlueCEPolicyMaxSlotsPerJob (GlueVOView $||$ GlueCE)
-
-\subsection*{int MaxStageInStreams}
-\subsection*{int MaxStageOutStreams}
-\subsection*{std::string SchedulingPolicy}
-
-ARC0: nordugrid-queue-schedulingpolicy
-
-\subsection*{int MaxMemory}
-
-ARC0: nordugrid-queue-nodememory $||$ nordugrid-cluster-nodememory \\
-CREAM: GlueHostMainMemoryVirtualSize
-
-\subsection*{int MaxDiskSpace}
-\subsection*{URL DefaultStorageService}
-
-ARC0: nordugrid-cluster-localse \\
-CREAM: GlueCEInfoDefaultSE (GlueVOView $||$ GlueCE)
-
-\subsection*{bool Preemption}
-
-CREAM: GlueCEPolicyPreemption (GlueVOView $||$ GlueCE)
-
-\subsection*{Period EstimatedAverageWaitingTime}
-
-CREAM: GlueCEStateEstimatedResponseTime (GlueVOView $||$ GlueCE)
-
-\subsection*{Period EstimatedWorstWaitingTime}
-
-CREAM: GlueCEStateWorstResponseTime (GlueVOView $||$ GlueCE)
-
-\subsection*{std::string ReservationPolicy}
-
-\section{ComputingManager attributes}
-
-\subsection*{std::string ManagerType}
-
-ARC0: nordugrid-cluster-lrms-type (nordugrid-cluster-lrms-version)
-
-\subsection*{bool Reservation}
-\subsection*{bool BulkSubmission}
-\subsection*{int TotalPhysicalCPUs}
-
-ARC0: nordugrid-queue-totalcpus $||$ nordugrid-cluster-totalcpus
-
-\subsection*{int TotalLogicalCPUs}
-
-ARC0: nordugrid-queue-totalcpus $||$ nordugrid-cluster-totalcpus
-
-\subsection*{int TotalSlots}
-
-ARC0: nordugrid-queue-totalcpus $||$ nordugrid-cluster-totalcpus
-
-\subsection*{bool Homogeneity}
-
-ARC0: nordugrid-queue-homogeneity $||$ nordugrid-cluster-homogeneity
-
-\subsection*{std::string NetworkInfo}
-
-\subsection*{bool WorkingAreaShared}
-
-\subsection*{int WorkingAreaFree}
-
-ARC0: nordugrid-cluster-sessiondir-free
-
-\subsection*{Period WorkingAreaLifeTime}
-
-ARC0: nordugrid-cluster-sessiondir-lifetime
-
-\subsection*{int CacheFree}
-
-ARC0: nordugrid-cluster-cache-free
-
-\section{ExecutionEnvironment attributes}
-
-\subsection*{std::string Platform}
-
-ARC0: nordugrid-queue-architecture $||$ nordugrid-cluster-architecture
-
-\subsection*{bool VirtualMachine}
-
-\subsection*{std::string CPUVendor}
-
-ARC0: nordugrid-queue-nodecpu $||$ nordugrid-cluster-nodecpu
-
-\subsection*{std::string CPUModel}
-
-ARC0: nordugrid-queue-nodecpu $||$ nordugrid-cluster-nodecpu
-
-\subsection*{std::string CPUVersion}
-
-ARC0: nordugrid-queue-nodecpu $||$ nordugrid-cluster-nodecpu
-
-\subsection*{int CPUClockSpeed}
-
-\subsection*{int MainMemorySize}
-
-\subsection*{std::string OSFamily}
-
-ARC0: nordugrid-queue-opsys $||$ nordugrid-cluster-opsys
-
-\subsection*{std::string OSName}
-
-ARC0: nordugrid-queue-opsys $||$ nordugrid-cluster-opsys
-
-\subsection*{std::string OSVersion}
-
-\subsection*{bool ConnectivityIn}
-
-\subsection*{bool ConnectivityOut}
-
-\subsection*{std::list<Benchmark> Benchmarks}
-
-ARC0: nordugrid-queue-benchmark $||$ nordugrid-cluster-benchmark
-
-\section{ApplicationEnvironment}
-
-\subsection*{std::list<ApplicationEnvironment> ApplicationEnvironments}
-
-ARC0: nordugrid-cluster-runtimeenvironment \\
-CREAM: GlueHostApplicationSoftwareRunTimeEnvironment
-
-\chapter{{\Job}}
-\label{app:job}
-
-\section{Information stored in the job list file}
-
-\subsection*{std::string Flavour}
-
-\subsection*{URL JobID}
-
-\subsection*{URL Cluster}
-
-\subsection*{URL SubmissionEndpoint}
-
-\subsection*{URL InfoEndpoint}
-
-\subsection*{URL ISB}
-
-\subsection*{URL OSB}
-
-\subsection*{URL AuxURL}
-
-\subsection*{std::string AuxInfo}
-
-\section{Information retrieved from the informtaion system}
-
-\subsection*{std::string Name}
-
-ARC0: nordugrid-job-jobname
-
-\subsection*{std::string Type}
-
-\subsection*{URL IDFromEndpoint}
-
-\subsection*{std::string LocalIdFromManager}
-
-\subsection*{std::string JobDescription}
-
-\subsection*{std::string State}
-
-ARC0: nordugrid-job-status
-
-\subsection*{std::string RestartState}
-
-ARC0: nordugrid-job-rerunable
-
-\subsection*{int ExitCode}
-
-ARC0: nordugrid-job-exitcode
-
-\subsection*{std::string ComputingManagerExitCode}
-
-\subsection*{std::list<std::string> Error}
-
-ARC0: nordugrid-job-errors
-
-\subsection*{int WaitingPosition}
-
-ARC0: nordugrid-job-queuerank
-
-\subsection*{std::string UserDomain}
-
-\subsection*{std::string Owner}
-
-ARC0: nordugrid-job-globalowner
-
-\subsection*{std::string LocalOwner}
-
-\subsection*{Period RequestedWallTime}
-
-ARC0: nordugrid-job-reqwalltime
-
-\subsection*{Period RequestedTotalCPUTime}
-
-ARC0: nordugrid-job-reqcputime
-
-\subsection*{int RequestedMainMemory}
-
-\subsection*{int RequestedSlots}
-
-\subsection*{std::string StdIn}
-
-ARC0: nordugrid-job-stdin
-
-\subsection*{std::string StdOut}
-
-ARC0: nordugrid-job-stdout
-
-\subsection*{std::string StdErr}
-
-ARC0: nordugrid-job-stderr
-
-\subsection*{std::string LogDir}
-
-ARC0: nordugrid-job-gmlog
-
-\subsection*{std::list<std::string> ExecutionNode}
-
-Arc0: nordugrid-job-executionnodes
-
-\subsection*{std::string ExecutionCE}
-
-ARC0: nordugrid-job-execcluster
-
-\subsection*{std::string Queue}
-
-ARC0: nordugrid-job-execqueue
-
-\subsection*{Period UsedWallTime}
-
-ARC0: nordugrid-job-usedwalltime
-
-\subsection*{Period UsedTotalCPUTime}
-
-ARC0: nordugrid-job-usedcputime
-
-\subsection*{int UsedMainMemory}
-
-ARC0: nordugrid-job-usedmem
-
-\subsection*{std::list<std::string> UsedApplicationEnvironment}
-
-ARC0: nordugrid-job-runtimeenvironment
-
-\subsection*{int UsedSlots}
-
-ARC0: nordugrid-job-cpucount
-
-\subsection*{Time LocalSubmissionTime}
-
-\subsection*{Time SubmissionTime}
-
-ARC0: nordugrid-job-submissiontime
-
-\subsection*{Time ComputingManagerSubmissionTime}
-
-\subsection*{Time StartTime}
-
-\subsection*{Time ComputingManagerEndTime}
-
-\subsection*{Time EndTime}
-
-ARC0: nordugrid-job-completiontime
-
-\subsection*{Time WorkingAreaEraseTime}
-
-ARC0: nordugrid-job-sessiondirerasetime
-
-\subsection*{Time ProxyExpirationTime}
-
-ARC0: nordugrid-job-proxyexpirationtime
-
-\subsection*{std::string SubmissionHost}
-
-ARC0: nordugrid-job-submissionui
-
-\subsection*{std::string SubmissionClientName}
-
-ARC0: nordugrid-job-clientsoftware
-
-\subsection*{Time CreationTime}
-
-ARC0: Mds-validfrom
-
-\subsection*{Period Validity}
-
-ARC0: Mds-validto $-$ Mds-validfrom
-
-\subsection*{std::string OtherMessages}
-
-ARC0: nordugrid-job-comment
-
-\section{Associations}
-
-\subsection*{URL JobManagementEndpoint}
-
-\subsection*{URL DataStagingEndpoint}
-
-\section{ExecutionEnvironment (condensed)}
-
-\subsection*{bool VirtualMachine}
-
-\subsection*{std::string UsedCPUType}
-
-\subsection*{std::string UsedOSFamily}
-
-\subsection*{std::string UsedPlatform}
-
-\chapter{{\Broker}}
-\label{app:broker}
-
-\section{Broker mapping}
-\label{sec:broker-mapping}
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/client/ExecutionTargetMapping.html
+
+\chapter{Broker mapping}
+\label{app:broker-mapping}
 
@@
 
-\section{JobInnerrepresentation}
-\label{sec:jobinnerrepresentation}
+\chapter{Job Inner Representation}
+\label{app:jobinnerrepresentation}
 

### 2009-02-13 doc/client/client_technical.tex
removed author names

@@
 \hypersetup{
-  pdfauthor = {Mattias Ellert, Bjarte Mohn, Ivan Marton, Gabor Roczei},
+  pdfauthor = {KnorARC collaboration},
   pdftitle = {libarcclient},
@@
 %%%%---- A list of authors ----
-{\centering \large Mattias Ellert\footnote{mattias.ellert@fysast.uu.se} \large \par}
-{\centering \large Iv\'{a}n M\'{a}rton\footnote{martoni@niif.hu} \large \par}
-{\centering \large Bjarte Mohn\footnote{bjarte.mohn@fysast.uu.se} \large \par}
-{\centering \large G\'{a}bor R\H{o}czei\footnote{roczei@niif.hu} \large \par}
+{\centering \large KnowARC \large \par}
 %%%%---- An abstract - if style is article ----

### 2009-02-19 doc/storage_system/arc-storage-documentation.tex
adding the skeleton of the User's manual renaming "arc1-storage" to "arc-storage"

--- 2008-09-02 doc/storage_system/arc1-storage-documentation.tex
@@
   pdfauthor = {Zsombor Nagy},
-  pdftitle = {Documentation of the ARC1 storage system},
+  pdftitle = {Documentation of the ARC storage system},
   pdfsubject = {Paper subject},
@@
 %%%%---- The title ----
-{\centering \textsc{\Large Documentation of the ARC1 storage system}\Large \par}
+{\centering \textsc{\Large Documentation of the ARC storage system}\Large \par}
 \vspace*{0.5cm}
@@
 
-The ARC1 storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.
+The ARC storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.
 The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems), and a collection can contain sub-collections and sub-sub-collections in any depth.
@@
 A storage node here is a network-accessible computer having storage space to share, and one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.).
-For each storage node one of the services of the ARC1 storage system is needed to manage it and to integrate it into the system.
-And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system.
+For each storage node one of the services of the ARC storage system is needed to manage it and to integrate it into the system.
+And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC storage system.
 The main services of the storage system are the following (see Figure~\ref{fig:services}):
@@
 \begin{figure}[ht]
-\centering{{\scalebox{0.9}{\includegraphics{arc1-storage-services.pdf}}}
-\caption{\label{fig:services}The components of the ARC1 storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
+\centering{{\scalebox{0.9}{\includegraphics{arc-storage-services.pdf}}}
+\caption{\label{fig:services}The components of the ARC storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
 \end{figure}
@@
 \begin{figure}[ht]
-\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-namespace.pdf}}}
+\centering{{\scalebox{0.7}{\includegraphics{arc-storage-namespace.pdf}}}
 \caption{\label{fig:namespace}Example of the hierarchy of the global namespace} }
@@
 
-The current version of the ARC1 storage system has no implemented solution for security issues.
+The current version of the ARC storage system has no implemented solution for security issues.
 Designing and implementing a powerful, flexible, fine-grained solution for authentication and authorization is very important.
@@
 \begin{figure}[ht]
-\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-downloading.pdf}}}
+\centering{{\scalebox{0.7}{\includegraphics{arc-storage-downloading.pdf}}}
 \caption{\label{fig:downloading}Downloading a file} }
@@
 \begin{figure}[ht]
-\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-uploading.pdf}}}
+\centering{{\scalebox{0.7}{\includegraphics{arc-storage-uploading.pdf}}}
 \caption{\label{fig:uploading}Uploading a file} }
@@
 \begin{figure}[ht]
-\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-removing.pdf}}}
+\centering{{\scalebox{0.7}{\includegraphics{arc-storage-removing.pdf}}}
 \caption{\label{fig:removing}Removing a file} }
@@
 
-The Bartender provides an easy to use interface of the ARC1 storage system to the users.
+The Bartender provides an easy to use interface of the ARC storage system to the users.
 You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}.
@@
 
-There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, \url{http://fuse.sourceforge.net/}} module and Windows Shell Extensions\footnote{\url{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx}} to be able to mount the ARC1 storage namespace into the local filesystem namespace, and use it with the commands of the operating system.
+There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, \url{http://fuse.sourceforge.net/}} module and Windows Shell Extensions\footnote{\url{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx}} to be able to mount the ARC storage namespace into the local filesystem namespace, and use it with the commands of the operating system.
 
@@
 
-To integrate existing files on a third-party storage system to our namespace thus make them accessible through the interface of the ARC1 storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
-It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC1 storage (Section~\ref{sub:data_model}).
+To integrate existing files on a third-party storage system to our namespace thus make them accessible through the interface of the ARC storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
+It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC storage (Section~\ref{sub:data_model}).
 The files here would be referenced by a path which is local in the namespace of the third-party storage.

### 2009-02-19 doc/storage_system/arc-storage-manual.tex
adding the skeleton of the User's manual renaming "arc1-storage" to "arc-storage"

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {User's manual of the ARC storage system},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-XXXXXXX-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large User's manual of the ARC storage system}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large First prototype status and plans}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
    {\centering \large Jon Nilsen\footnote{j.k.nilsen@usit.uio.no} \large \par}
    {\centering \large Salman Zubair Toor \footnote{salman.toor@it.uu.se} \large \par}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}

\chapter{Quick start guide} % (fold)
\label{cha:quick_start_guide}

\section{Checking the code out from the subversion} % (fold)
\label{sec:checking_the_code_out_from_the_subversion}

% section checking_the_code_out_from_the_subversion (end)

\section{Compiling and installing} % (fold)
\label{sec:compiling_and_installing}

% section compiling_and_installing (end)

\section{Modifying the configuration template} % (fold)
\label{sec:modifying_the_configuration_template}

% section modifying_the_configuration_template (end)

\section{Running the daemon} % (fold)
\label{sec:running_the_daemon}

% section running_the_daemon (end)

\section{Using the CLI tool} % (fold)
\label{sec:using_the_cli_tool}

% section using_the_cli_tool (end)

% chapter quick_start_guide (end)

\chapter{Installing the system} % (fold)
\label{cha:installing_the_system}

\section{Getting the source code} % (fold)
\label{sec:getting_the_source_code}

% section getting_the_source_code (end)

\section{Configuration} % (fold)
\label{sec:configuration}

\subsection{HED service config} % (fold)
\label{sub:hed_service_config}

% subsection hed_service_config (end)

\subsection{Centralized A-Hash} % (fold)
\label{sub:centralized_a_hash}

% subsection centralized_a_hash (end)

\subsection{Distributed A-Hash} % (fold)
\label{sub:distributed_a_hash}

% subsection distributed_a_hash (end)

\subsection{Librarian} % (fold)
\label{sub:librarian}

% subsection librarian (end)

\subsection{Shepherd} % (fold)
\label{sub:shepherd}

% subsection shepherd (end)

\subsection{Hopi} % (fold)
\label{sub:hopi}

% subsection hopi (end)

\subsection{Bartender} % (fold)
\label{sub:bartender}

% subsection bartender (end)

\subsection{Configuring inter-service trust} % (fold)
\label{sub:configuring_inter_service_trust}

% subsection configuring_inter_service_trust (end)

\subsection{Virtual organizations} % (fold)
\label{sub:virtual_organizations}

% subsection virtual_organizations (end)

% section configuration (end)

\section{Deployment scenarios} % (fold)
\label{sec:deployment_scenarios}

% section deployment_scenarios (end)

\section{Single node} % (fold)
\label{sec:single_node}

% section single_node (end)

\section{Centralized management} % (fold)
\label{sec:centralized_management}

% section centralized_management (end)

\section{Simple distribution} % (fold)
\label{sec:simple_distribution}

% section simple_distribution (end)

\section{One service per node} % (fold)
\label{sec:one_service_per_node}

% section one_service_per_node (end)

% chapter installing_the_system (end)

\chapter{Using the system} % (fold)
\label{cha:using_the_system}

\section{Prototype command line tool} % (fold)
\label{sec:prototype_command_line_tool}

\subsection{Configuring the command line tool} % (fold)
\label{sub:configuring_the_command_line_tool}

% subsection configuring_the_command_line_tool (end)

\subsection{Creating the root collection} % (fold)
\label{sub:creating_the_root_collection}

% subsection creating_the_root_collection (end)

\subsection{Uploading files} % (fold)
\label{sub:uploading_files}

% subsection uploading_files (end)

\subsection{Creating collections} % (fold)
\label{sub:creating_collections}

% subsection creating_collections (end)

\subsection{Moving files} % (fold)
\label{sub:moving_files}

% subsection moving_files (end)

\subsection{Creating hardlinks} % (fold)
\label{sub:creating_hardlinks}

% subsection creating_hardlinks (end)

\subsection{Removing files} % (fold)
\label{sub:removing_files}

% subsection removing_files (end)

\subsection{Removing collections} % (fold)
\label{sub:removing_collections}

% subsection removing_collections (end)

\subsection{Close collections} % (fold)
\label{sub:close_collections}

% subsection close_collections (end)

\subsection{Modify access policies} % (fold)
\label{sub:modify_access_policies}

% subsection modify_access_policies (end)

\subsection{Working with mount points} % (fold)
\label{sub:working_with_mount_points}

% subsection working_with_mount_points (end)

% section prototype_command_line_tool (end)

\section{Accessing the system with FUSE} % (fold)
\label{sec:accessing_the_system_with_fuse}

\subsection{Configuring the FUSE module} % (fold)
\label{sub:configuring_the_fuse_module}

% subsection configuring_the_fuse_module (end)

\subsection{Mounting the storage system} % (fold)
\label{sub:mounting_the_storage_system}

% subsection mounting_the_storage_system (end)

\subsection{Supported operations} % (fold)
\label{sub:supported_operations}

% subsection supported_operations (end)

\subsection{Limitations} % (fold)
\label{sub:limitations}

% subsection limitations (end)

% section accessing_the_system_with_fuse (end)

% chapter using_the_system (end)

% \bibliography{grid}
\end{document}

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Tutorial on programming web services with ARC-1. This some interim state, uploaded to help our Luebeck-internal review as a start. An announcement to the list will follow next week.

\chapter{The echo webservice}

\section{Webservice definition}
wsdl

In case of webservices, which are supported by the KnowArc architecture and used in the present tutorial, the interface of the requests and responses may be definied by a \textit{wsdl} file (WebService Definition Language).

\lstsetCPP
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/clients/echoclient/echoclient.cpp}

\lstsetCPP

\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/echoservice/echoservice.cpp}

\lstsetARCHEDXML
\begin{minipage}[t]{\textwidth}
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/echoservice/arched_echoservice.xml}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ rm -f /var/log/arched.log
$ arched -c arched_echoservice.xml  && echo jo ||echo n
$ tail -n100 -f /var/log/arched.log
$ killall arched
\end{lstlisting}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ ./echoclient
Tue Feb 17 17:04:08 2009
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
	<soap-env:Body>
		<echoRequest:echo>
			<echo:say>Der_Berg_ruft</echo:say>
		</echoRequest:echo>
	</soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <echo:echoResponse>
      <echo:hear>[ Der_Berg_ruft ]</echo:hear>
    </echo:echoResponse>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

Folgender XML Aufruf und Antwort soll \textit{automatisch} generiert werden:

%[caption={[]WegDamit},language=XML,basicstyle=\scriptsize,breaklines=true,label=lst:request]

Zertifikate
Zustände (Nutzer wiedererkennen, arbeit aufnehmen)

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Tutorial on programming web services with ARC-1. This some interim state, uploaded to help our Luebeck-internal review as a start. An announcement to the list will follow next week.

\chapter{Introduction}

This guide gives a first inside in how to write a standalone webservice for KnowArc including a suited client.
The practically oriented examples will be followed by explanations of the main architecture concepts.
For being able to comprehense basic knowledge of the programming language C++, the file format XML and the shell is requried.
Furthermore an installation of KnowArc 1 is needed.\\

The KnowArc server provides access to services which are used by clients. The services may cover computational tasks as well 
as access to a database. % job submission
Servers provide access to a certain service and are waiting for a client which demands it. The task of servers can be subdivided into:
\begin{itemize}
 \item Wait for client request
 \item Receive the request
 \item Perform the desired service
 \item Create a response for the client
 \item Send response to the client
 \item Wait for next client request
\end{itemize}
While the server waits passively for a request, the client acts actively:
\begin{itemize}
 \item Create a request
 \item Transmit the request to the server which provides the desired service.
 \item Receive the response from the server.
 \end{itemize}
\forcelinebreak

In general the implementation of a client is easier to be done compared with the server. 
KnowArc is designed as a middleware which encapsulates typical challenges of server-client infrastructures (security, exception handling, extensibility) and abstract from underlying architecture (heterogenous computer, computer location, protocols).
The core of KnowArc is the HED (Hosting Environment Daemon). HED is model consisting of a variable amount of heavily configurable layers. 
%
In the present case the endpoint of the HED will be a SOAP based Web Service. These Web Services exchange messages enveloped in the well known XML format. KnowArc abstracts from this format but nevertheless it is useful to be aware of the underlying message transport mechanism.\\
%which implicate several advantages.
%The XML file format is not restricted to a certain set of tags but can be expanded in such a way as to fit almost any kind of service. It is not a proprietary format and as a result many libraries in almost any language are capable in creating that kind of format.Likewise, XML is also suitable for the usage between heterogeneous platforms.However, while the client can be implemented very straight, the service has to fit a certain layout.In the present guide four kinds of webservices will be introduced: a simple time service, an echo service, a secure echo service and a service with a persistence state .\\

The source code of the examples given in this tutorial are available in the directory \textit{src}.
\task{All examples can be found in the source tree: URL}

\section{Hosting Environment Daemon}

The HED which was already mentioned in the section above is an essential part of the ARC middleware. It provides hosting of various services at application level, as well as a number of modules~\cite{QIANG_2005}. Figure~\ref{fig:HED_internal} illustrates a typical setup of a server sided HED. The HED consists of a few compononents called Message Chain Components (MCC) which are ordered within a layered structure and several \textcolor{urgent}{modules} which shall aid the programmer to simplify the development of the Web Services (Config, Loader, Logging, XMLNode).
%
\begin{figure}
	\centering
	\includegraphics[width=10cm]{tex_introduction/HED.pdf}
	\mycaption{Example of HED structure with an Arex service and a file service.\task{ Replace image by an easier one!} }{\task{Write more stuff}}
\label{fig:HED_internal}
\end{figure}
The MCCs are passing the incoming and outgoing messages to the upper or the respective lower layer.
% AUFZÄHLUNG???
The lowest layer has to be a transport protocol like TCP which enables the transmission between the two applications, the client and the server. (The Internet Protocol (IP) beneath the TCP layer enables a transmission between two computers. IP is needed but not part of the HED.) TCP is well suited for the application because it provides a reliable and ordered delivery of a byte stream which is important to ensure security and a requirement for the usage of HTTP.
The TCP layer is followed by the Hypertext Transfer Protocol (HTTP). 
In case a security layer is desired, the TLS (Transport Layer Security formerly known as SSL, Secure Sockets Layer) can be inserted between the two layers TCP and HTTP.
The HTTP provides the client server architecture. It is stateless but offers several extensions for requests, header information and status codes. Furthermore HTTP enables the usage of Uniform Resource Locators (URL) which is used by KnowArc to multiplex between different services. The multiplexing layer needs to be defined within the HED and is called Plexer. Depending on the path of the URL the Plexer passes messages to a defined SOAP service which again passes the message to the Web Service.\\

The structure of the HED can be configured freely by modifying the HED configuration file. A first impression how the HED can be configured shall be given in the following example which corresponds to the HED shown in figure~\ref{fig:HED_internal}. It uses the internal ARC Echo service.

\subsection{The Arcecho service}
\task{Maybe the word tag needs to be replaced by the word element...}
\task{Where is the xsd file for HED configuration files??}
The first example shall give a basic understanding in how to configure the HED. The task is to setup the ARC intern Echo Web Service which has to be done by invoking the \textit{arched} daemon with a fitting HED configuration file. Such a fitting HED configuration file is shown in Listing~\ref{lst:arched_arcecho_xml}. It is written in XML and its structure is based on the \textcolor{urgent}{??} schema.
\lstsetARCHEDXML\\
\begin{program}
\centering
\lstinputlisting
	[
	label=lst:arched_arcecho_xml,
	caption={[HED configuration. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/arched_arcecho.xml}
\end{program}
The first line of the XML file contains the XML declartion. Several attributes can be defined here but at least the XML version should be specified. \task{What is this ArcConfig for??} % configuration namespaces
The \textit{Server} tag, to be found at line~\ref{lst_code:arched_arcecho_Server}, provides basic settings for the daemon such as location of the Pid-file or the Log-file.
The \textit{ModuleManager} holds the path to the plugin libraries. Several paths may be defined here. Due to the fact that the first library fitting to the plugin name is loaded, the order is relevant. The path should be at least assigned to the ARC installation directory. % for the standard MCC are defined here
The next set of tags are holding the name of the plugins to be loaded (line~\ref{lst_code:arched_arcecho_Plugins}).
The names have to correspond to the name of the dynamic libraries within the path defined by the \textit{ModuleManager}.
Finally the chain of the components are declared initiate with the tag \textit{Chain}.
Each component inside the chain is represented by a tag called \textit{Component} which has at least a name attribute.
Its important that a plugin has been loaded which provides the functionality of it.
Depending on the component certain additional tags may be defined inside of it.
\task{Internet Protocol is required.}
The first component of the chain is the \textit{tcp.service}. As to be seen the port 60000 gets assigned for the server to listen to.
The stream which is received by this component is passed to the component defined by the \textit{next} tag.
In that case the will be passed to the component with the \textit{id} atrribute named \textit{http}. According to this manner the message will be passed to higher layers until it reaches the \textit{Plexer}  Depending on the path of the URL which was demanded by the HTTP the Plexer multiplexes the request to the right Web Service. The path declared within the \textit{next} tag is a regular expression and leads to the service with the id named \textit{echo}. This service finally will process the message and create a new message which will be returned all the way back to the TCP layer. The Echo Web Service itself is provided by the ARC package. Additional fix parameters can be passed by assigning tags which have to be known to the Web Service.
The tags \textit{echo:prefix} and \textit{echo:suffix} seen in this example are defining the characters which will enclose the echo  returned string.
\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}\\

To create the daemon corresponding to the defined HED configuration file ARC-1 provides the program \textit{arched}.
The manner of invokation can been seen in Listing~\ref{lst:arcecho_arched_invokation}.
\lstsetKSH
\begin{program}
\begin{lstlisting}[
label=lst:arcecho_arched_invokation,
caption={[Invokation of the Arched Daemon.]
         \textbf{Invokation of the Arched Daemon.\textcolor{white}{hmf}}}]

$ arched -c arcecho_no_ssl.xml  && echo Daemon started || echo Daemon start failed
 Daemon started
$
\end{lstlisting}
\end{program}
% $ rm -f /var/log/arched.log
% $ tail -n100 -f /var/log/arched.log
%...
%$ killall arched
%
%
The command is followed by the parameter \textit{c} which assignes, that the next string contains the full path of configuration file.
Once the server is running, the client can be used as described in Listing~\ref{lst:arcecho_client_invokation}.
\lstsetKSH
\begin{program}
\begin{lstlisting}[
label=lst:arcecho_client_invokation,
caption={[Usage of the Arc echo client.]
         \textbf{Usage of the Arc echo client.\textcolor{white}{hmf}}}]
$ arcecho http://localhost:60000/Echo text
[ text ]
\end{lstlisting}
\end{program}
In this example the server and the client are running on the same computer. For debugging it is advisable to check the Log-file which was assigned inside the HED configuration file.\\

In the following chapters own Web Services will be developed. Starting with the simple Time Web Service.

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
Tutorial on programming web services with ARC-1. This some interim state, uploaded to help our Luebeck-internal review as a start. An announcement to the list will follow next week.

\chapter{SEC echoservice}

\section{SecureHandler}
And what are PDP's?

Currently, the ARC1 comes in severalfour policy decision point implementations:
\begin{itemize}
 \item arc.pdp – Arc PDP\\
The Arc PDP will organize the security attributes into the ARC specific authorization request, call the
policy evaluator to evaluate the request against the policy (which is in ARC specific format) repository,
and get back the evaluation result. See paragraph 3 for detail information about request schema and policy
schema.
\item delegtion.pdp – Delegation PDP\\
The Delegation PDP is basically similar to Arc PDP, except it uses the delegation policy parsed from
remote proxy credential by delegation.collector, and evaluates the request against delegation policy. See
section 6for the design idea and use case of delegation policy in fine-grained identity delegation.
\item simplelist.pdp – Simplelist PDP\\
The Simplelist PDP is a simplest implementation of policy decision point. It will match the identity
extracted from the remote credential (or proxy credential) with local list of permitted identities.
\item pdpservice.invoker – PDP Service Invoker\\
The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the
same functionality as Arc PDP, except that the evaluation request and response are carried by SOAP
message. The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation
engine can be accessed remotely and maintained centrally.
 \item allow.pdp\\
 Does something unknown!
\end{itemize}

Following Flavors:
\subsection{arc.authz – Authorization SecHandler}
The arc.authz is responsible for calling the interface of policy decision point and getting back the
authorization result, and then making decision according to this authorization result. There is one simple
interface (see Figure 4) defined in PDP, which will be called by arc.authz if configured inside once there
is message (incoming or outgoing) need to be processed.

\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:secHandler_arc.authz,
language=XML,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
...
<Service name="sececho" id="sececho">
	<next id="sececho"/>
	<echo:prefix>[ </echo:prefix>
	<echo:suffix> ]</echo:suffix>
	<SecHandler name="arc.authz" id="pdps" event="incoming">
		<PDP name="simplelist.pdp" location="/etc/grid-security/grid-mapfile"/>
	</SecHandler>
</SecHandler>
...
\end{lstlisting}
\end{minipage}

\begin{lstlisting}[
label=lst:invokation,
language=ksh,
frame=single,
numbers=none,
backgroundcolor=\color{lightgray},
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ tail -n5 /etc/grid-security/grid-mapfile
"/C=Au/O=Science/OU=ARC/CN=Baldur Meier" griduser
"/C=UK/O=MoreScience/OU=ARC/CN=Thor Schulz" griduser
"/C=Au/O=Organelle/OU=ARC/CN=Loki Svensson" griduser
"/C=SK/O=Grid.Org/OU=ARC/CN=Iduna Apple" griduser
"/C=RU/O=MegaGrid/O=ARC/CN=Fenrir Wodaneater" griduser
\end{lstlisting}
Add your client certificate into that list!

Latest: Policy.xsd
http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Policy.xsd

\subsection{identity.map – Identity Mapping SecHandler}

The identity.map is a specific authorization oriented security handler. It will map the global identity in
the message into local identity like system username based on the result returned by Policy Decision Point
components.

\subsection{delegation.collector – Delegation SecHandler}

The delegation.collector is responsible for collecting the delegation policy information from the remote
proxy credential (proxy certificate is compatible to RFC3820) inside the message, and putting this policy
into message’s security attribute for the usage of other components, such as delegation.pdp.

\subsection{ usernametoken.handler – UsernameToken SecHandler}

The task of the usernametoken.handler is to generate the WS-Security Username-Token and add it into
header of SOAP message which is the payload of outgoing message. It can also extract the WS-Security
Username-Token from the header of SOAP message which is the payload of incoming message.

\lstsetCPP
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/clients/secechoclient/secechoclient.cpp}

\lstsetCPP

\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/secechoservice/secechoservice.cpp}

\lstsetARCHEDXML
\begin{minipage}[t]{\textwidth}
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/secechoservice/arched_sec_echoservice.xml}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ rm -f /var/log/arched.log
$ arched -c arched_echoservice.xml  && echo jo ||echo n
$ tail -n100 -f /var/log/arched.log
$ killall arched
\end{lstlisting}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ ./echoclient
Tue Feb 17 17:04:08 2009
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
	<soap-env:Body>
		<echoRequest:echo>
			<echo:say>Der_Berg_ruft</echo:say>
		</echoRequest:echo>
	</soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <echo:echoResponse>
      <echo:hear>[ Der_Berg_ruft ]</echo:hear>
    </echo:echoResponse>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Tutorial on programming web services with ARC-1. This some interim state, uploaded to help our Luebeck-internal review as a start. An announcement to the list will follow next week.

\chapter{The time webservice}

The first implementation of a Web Service which will be presented in this tutorial realises a simple time service.
In case the service receives a request, it will respond with a string containing the systemtime.
For simplicity the service will not parse the request but always return the current time.
In the following subsections the C++ implementation first the service and than the client will be developed.

\section{Service}

In Arc-1 the services are implementated as plugins such that they can easily be included by modifiying the HED configuration file as it was to be seen in Listing~\ref{lst:arched_arcecho_xml}.
The Listing~\ref{lst:time_arched_cpp}.shows the implementation of the time service plugin.
The proper class definition is defined in the corresponding header file which isn't shown here.
The key element for each plugin is the struct PluginDescriptor named PLUGINS\_TABLE\_NAME, see line\ref{lst_code:time_cpp_ptn}. 
%The struct defines basic entities of the service and provides a pointer to the method which is able to create an instance of the time service plugin. 
Within the struct a unique plugin name, the type of the plugin, its version and the pointer to the function which returns the time service instance has to be defined.
The function which returns the pointer to an instance of the service needs to have the PluginArgument as a parameter. 
The actual Time Web Service is implemented within the namespace ArcService. The constructor can be found in line~\ref{lst_code:time_cpp_constructor}, the deconstructor in line~\ref{lst_code:time_cpp_deconstructor} and the method required by the inherited class \textit{Service} file in line~\ref{lst_code:time_cpp_process}. For the Time service is very simple, only the function process needs to be implemented. It gets the incoming and the outgoing message as references and returns a class representing the status the service achieved. 
\lstsetCPP
\lstinputlisting
	[
	label=lst:time_arched_cpp,
	caption={[C++ implementation of the time service. Filename: timeservice.cpp]
	\textbf{C++ implementation of the time service. Filename: timeservice.cpp\textcolor{white}{hmf}}}
	]
{../src/services/timeservice/timeservice.cpp}
The service first determines the current time and than creates a SOAP message using the object \textit{PayloadSOAP} on line~\ref{lst_code:time_cpp_process_payloadSOAP}. The class \textit{PayloadSOAP} is able to create a XML message which is conform to the SOAP protocol (The apperance of the generated request and response will be discussed later). I case of the presented time service the payload consists out of two elements nested into one another. The element \textit{text} holds the element child which again holds the string containing the current time.
Within the line~\ref{lst_code:time_cpp_process_message_payload} the payload is passed to the message. 
The message processing is now completed and the function returns its success status on line~\ref{lst_code:time_cpp_return}.\\

After the service has been implemented, it needs to be loaded into the HED. A suitable HED configuration file is to be seen in Listing~\ref{lst:time_arched_xml}.

\lstsetARCHEDXML
\begin{minipage}[t]{\textwidth}
\lstinputlisting
	[
	label=lst:time_arched_xml,
	caption={[HED configuration for the time service. Filename: arched\_timeservice.xml]
	\textbf{HED configuration for the time service. Filename: arched\_timeservice.xml\textcolor{white}{hmf}}}
	]
{../src/services/timeservice/arched_timeservice.xml}
\end{minipage}

\section{Client}

\task{Hardcoded URL!!!}

\lstsetCPP
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/clients/timeclient/timeclient.cpp}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ arched -c arched_timeservice.xml  && echo jo ||echo n
\end{lstlisting}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ ./timeclient
Wed Feb 18 11:20:30 2009
\end{lstlisting}
\end{minipage}

The data is enveloped by an element called \textit{Envelope} .. See german wiki

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:time="urn:time" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <time>
      <timeRequest></timeRequest>
    </time>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <time>
      <timeResponse>Wed Feb 18 11:20:30 2009</timeResponse>
    </time>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_titlepage/titlepage.tex
Tutorial on programming web services with ARC-1. This some interim state, uploaded to help our Luebeck-internal review as a start. An announcement to the list will follow next week.

--- 2008-10-01 doc/hed/ARCHED_article.tex
@@
-\documentclass{book}
-%\documentclass{article}                            %for shorter notes
-\usepackage{graphicx}                              %for PNG images (pdflatex)
-%\usepackage{graphics}                              %for EPS images (latex)
-\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
-\usepackage{color}                                 %for defining custom colors
-\usepackage{framed}                                %for shaded and framed paragraphs
-\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
-\usepackage{geometry}                              %for defining page size
-\usepackage{longtable}                             %for breaking tables
-\usepackage{ifthen}
-\newboolean{brieftechdesc}
-\newboolean{verbtechdesc}
-\setboolean{brieftechdesc}{false}
-\setboolean{verbtechdesc}{true}
-%
-\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
-\hypersetup{
-  pdfauthor = {J. J\"{o}nemo, A. Konstantinov},
-  pdftitle = {The Hosting Environment of the Advanced Resource Connector middleware},
-  pdfsubject = {Paper subject},
-  pdfkeywords = {HED,ARC},
-  pdfcreator = {PDFLaTeX with hyperref package},
-  pdfproducer = {PDFLaTeX}
-}
-%
-\bibliographystyle{IEEEtran}                       %a nice bibliography style
-%
-\def\efill{\hfill\nopagebreak}%
-\hyphenation{Nordu-Grid}
-\setlength{\parindent}{0cm}
-\setlength{\FrameRule}{1pt}
-\setlength{\FrameSep}{8pt}
-\addtolength{\parskip}{5pt}
-\renewcommand{\thefootnote}{\fnsymbol{footnote}}
-\renewcommand{\arraystretch}{1.3}
-\newcommand{\dothis}{\colorbox{shadecolor}}
-\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
-\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
-\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
-\definecolor{shadecolor}{rgb}{1,1,0.6}
-\definecolor{salmon}{rgb}{1,0.9,1}
-\definecolor{bordeaux}{rgb}{0.75,0.,0.}
-\definecolor{cyan}{rgb}{0,1,1}
-%
-%----- DON'T CHANGE HEADER MATTER
-\begin{document}
 \def\today{\number\day/\number\month/\number\year}
@@
 \begin{tabular}{rl}
-\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
+\resizebox*{3cm}{!}{\includegraphics{images/ng-logo.png}}
 &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
@@
 %%%%---- The title ----
-{\centering \textsc{\Large The Hosting Environment of the Advanced Resource Connector middleware}\Large \par}
+{\centering \textsc{\Large Webservice programming Tutorial}
+\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{I am currently working on this document, but I appreciate all comments and suggestions.
+\\e-mail: glodek@inb.uni-luebeck.de} % DELETE ONLY THAT LINE!!!
+\Large \par}
 \vspace*{0.5cm}
@@
 %%%%---- A list of authors ----
-    {\centering \large J. J\"{o}nemo\footnote{johan.jonemo@hep.lu.se}, A. Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no} \large \par}
+    {\centering \large Michael Glodek, Steffen M\"oller, ... \large \par}
 
@@
 \sloppy
-
-\chapter{Introduction}
-
-The Hosting Environment Daemon (HED) is the container of all the functional components of the next generation of the Advanced Resource Connector (ARC) middleware on the server side.
-It is the central part in a new very lightweight incarnation of ARC that is aimed at - but not limited to - providing Web Service.
-
-The whole design of the HED is built around the idea of flexibility and modularity.
-Inside HED the developer or deployer is supposed to use the minimum amount of components and external dependencies only.
-This is why the HED mostly consists of pluggable modules with some glue among them.
-
-Because in its current state it mostly provides modules for building SOAP based Web Services, it is easy to think that HED is just another Web Services development framework like Axis, gSOAP, XFire or any other out of the numerous implementations.
-Instead, the idea of HED is to provide framework for gluing functionalities and not a re-implementation of various standards.
-Effectively that means if Apache 2 web server is considered by developers as necessary for serving as frontend to services there could be plugin written which places Apache 2 into a chain of other plugins of the HED.
-
-In the current implementation there are no Apache or Axis plugins.
-That is because the developers of HED were very much concerned about making the solution lightweight and needed an implementation of those supported protocols that was both simple and lightweight.
-As a result essentials like SOAP and HTTP are implemented inside HED, while external software is used whenever was appropriate - as in the case of TLS, (Grid)FTP, LDAP and some other cases.
-That does not exclude possibility to have plugins using entirely external solutions either developed or accepted from third parties.
-
-The HED is a relatively young framework and there are quite a few rough edges and non-flexible solutions.
-The situation will hopefully improve with time.
-We would be grateful for any suggestions how to improve the architecture and the code of the HED.
-%Statements like "you are doing crap, I shall use Axis instead" are understood but not welcome and are usually reacted adequately.
-
-\chapter{How to read}
-
-This document does not (yet?) include in depth description of C++ classes which constitute the HED.
-Instead most sections contain notes entitled "Relevant classes".
-Technical description of those classes can be found in automatically generated "Hosting Environment (Daemon) Reference Manual" document \cite{hed-api}.
-
-For examples please see the source code in the repository \cite{svn-arc1}. Many components and libraries are accompanied with test and example applications.
-
-\chapter{Architecture}
-
-\section{Requirements}
-% Perhaps this can be an introductory part of the section without any subsection, we'll see
-In the design of the HED, several goals and requirements were considered. These were weighed against each other and the factual context.
-
-The implementation language had to be object oriented, efficient and capable of providing easy access to system functionality.
-This eventually lead to the adoption of C++.
-But languages such as Java and Python were also considered in the early stage.
-
-External dependencies needed to be kept to a minimum while also taking into consideration their ubiquity or relative rarity as well as license related concerns.
-Software of this level of complexity must depend on many external libraries and components but each such dependency has been introduced only after thorough consideration.
-
-Using resources in efficient way was an important goal.
-The present design enables many services sharing both the same process and the same network port(s) while exhibiting a remarkably low memory footprint.
-
-%\section{Functional design}
-
-%firewalls, modularity, portability, flexibility
-
-\section{Technical design}
-
-%How was this implemented, dynamic loading, mcc...
-In the technical design it turned out that providing dynamic loading, portability and a well tested high level memory management could all be greatly assisted by introducing glibmm \cite{gtkmm} - the C++ interface made for the GNOME \cite{gnome} projects library for memory management and related functionality.
-This enables the developers to write code easily portable across various operating systems and architectures.
-
-%Most of the configuration
-
-The term HED means three things:
-\begin{enumerate}
-
-\item the daemon (called \texttt{arched}) which hooks up the system and initialize components the way as it is described in the configuration files.
-This configuration describes the components and their relations to each other.
-In optimal cases these single services run on any node where ARC1 is deployed and started.
-Without loadable components the daemon itself does nothing usefull.
-
-\item sometimes using the HED terms to refer to collection of libraries which is used by service or other component developers.
-These libraries define interfaces and implement some common classes which may simplify the life of service and component developers however only few of these classes are mandatory to use to make the components and services loadable and hookable by the daemon.
-
-\item the collection of components implementing minimal set of protocols needed for implementing so called Web Services.
-
-Unless otherwise stated the term HED will be used through this document to refer to the second option - framework of C++ classes.
-
-\end{enumerate}
-
-\subsection{MCC}
-
-\texttt{Relevant classes: Arc::MCC, Arc::Loader}
-
-In the HED data channels to the outside world may be set up by chains of small processing units called Message Chain Components (MCCs).
-The chain is an ordered list of MCCs and their interconnection can be described in the configuration file.
-The MCCs work on units called Messages which represent data going in to or out of the HED.
-The Message consists of the so called Payload which is its main content structured in a way relevant to the protocol of the corresponding MCC, and auxiliary structures such as general Attributes and Security Attributes where information relevant to each protocol is accumulated as the Message progresses.
-Each MCC typically implements one level in the Internet Protocol suite by transforming the Message to an input suitable to propagate to the next component and then performs the corresponding transformation of the response on the way back.
-The components are all dynamically loaded to provide maximum flexibility and extensibility.
-Each instance of these MCC's can be individually configured.
-
-Each MCC has an entry method \texttt{process()} which is called with Message being processed.
-It then processes Message by modifying it or creating new Message.
-Then MCC calls entry method of the next MCC in the chain.
-For information how Messages are handled and about memory management policies please see API description of MCC class in \cite{hed-api}.
-
-The developer who writes an MCC is free to choose any 3rd party library and component to implement the functionality of the MCC but at least currently the MCCs should be written in the same language as the HED was written (C++) and should use the \texttt{MCC} interface class and \texttt{Message} class provided by \texttt{arcloader} and \texttt{arcmessage} libraries of ARC1.
-
-The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs.
-Typical scenario is that the HTTP MCC at the server side routes the HTTP Messages with POST HTTP operation to a SOAP MCC but the Messages with GET operation to for example a simple HTTP service component.
-For that purpose <next> elements in MCC configuration may have optional \texttt{id} attribute which allows to assign labels to all chain links to the next MCCs in the chain.
-Supported labels are MCC dependent.
-By default simple MCCs support only one unnamed link.
-The MCCs with routing capabilities must have all supported labels documented.
-
-As the data is passed through the individual MCCs, they each populate structures with both general attributes and special security attributes that are available at that particular protocol level.
-
-In general every MCC has optimal and natural places in certain chains and this place cannot always be modified.
-For example on the server side the TCP MCC must be the first MCC in any chain where it is used and the TLS MCC should be right after the TCP MCC.
-
-\subsubsection{Server and Client Side MCCs}
-
-Most of the MCCs has a client and a server version because the behavior of an MCC should be different depending on whether it is sitting on the server or client side.
-The typical scenario here is illustrated by the TCP MCC which should listen and wait for incoming messages on a socket on the server side but call \texttt{connect()} on the client side.
-
-Server and Client side MCCs are separate elements although definitely sharing some code and normally provided inside the same plugin module.
-
-\subsubsection{Payload}
-
-\texttt{Relevant classes: Arc::Payload, Arc::PayloadRawInterface, Arc::PayloadRaw, Arc::PayloadStreamInterface, Arc::PayloadStream, Arc::PayloadSOAP}
-
-Main content of the information is transferred using the Payload part of the Message.
-There are no limitations on functionality of Payload object except that it must be inherited from MessagePayload class.
-Despite being flexible such approach would be useless.
-This is why HED defines three Payload interfaces and their simplest implementations.
-All MCCs which are distributed with the HED use, implement and extend those interfaces.
-Those include:
-
-\begin{enumerate}
-
-\item PayloadRawInterface and its implementation PayloadRaw.
-This interface represents set of catenated in-memory chunks.
-It's meant to be used for information available as whole.
-And also for prepending and appending information without actually moving and copying data chunks in memory.
-
-\item PayloadStreamInterface and its implementation PayloadStream.
-It covers case of sequentially accessible information.
-The main purpose of that Payload is to serve protocols which define continuous data stream like TCP.
-
-\item PayloadSOAP represents parsed SOAP message. It's introduced to cover need for writing SOAP based Web Services in unified way.
-
-\end{enumerate}
-
-Each MCC developed inside and outside the HED must be accompanied with description of the Payloads it supports on input and those generated on output.
-Those types should be taken into account while creating chains of the MCCs.
-There are no Payload type checks done during the chain configuration phase.
-Hence Payload incompatibility problems will be detected only during runtime.
-
-\subsubsection{Attributes}
-
-\texttt{Relevant classes: Arc::MessageAttributes, Arc::Message}
-
-The Message object may contain general purpose key and value pairs called Attributes.
-Keys and values are simple strings.
-Each key may have multiple corresponding values.
-All pairs are handled by MessageAttributes class.
-Codewise there are no limitations posed on content and purpose of Attributes.
-
-By convention keys are composed of two parts: name of MCC/Protocol at which Attribute was generated or must be consumed and name of Attribute itself separated by column.
-For example the Attribute with key HTTP:METHOD holds HTTP protocol method like GET, PUT, HEAD, etc.
-It is either generated by MCC implementing HTTP protocol or is filled by other code and is used by HTTP MCC to generate proper HTTP header.
-
-Each MCC developed inside and outside the HED must have generated and consumed Attributes described in accompanying documentation.
-
-\subsubsection{Security Attributes}
-
-\texttt{Relevant classes: Arc::SecAttr, Arc::MessageAuth, Arc::MessageAuthContext, Arc::Message}
-
-Here only basic information about security related objects is presented. For more detailed information please see "Security Framework of ARC1" \cite{sec-arc1}.
-
-Security Attributes are storing various aspects of Message useful for authorization decisions to be made.
-Those normally include the operation being requested, target of operation and identity of subject making request.
-They can also contain authorization policies.
-Actually nothing stops from storing an other type of information but there is no convention developed for that.
-
-The Security Attributes are stored as key and value pairs.
-Each key may have only single value attached.
-Keys are simple strings.
-By convention each MCC or Security Handler (see below) produce Security Attribute with name corresponding to protocol name.
-For example Security Attribute stored under name TLS holds information collected at Transport Level Security layer.
-
-The value of Security Attribute is an object of class inherited from SecAttr.
-The HED implements class SecAttr which serves as definition of interface for all Security Attributes.
-It defines the way how collected information may be turned into format suitable for making authorization decisions.
-For that each Security Attribute value implements method Export for converting internal information into one of the supported formats.
-Currently only implemented is ARC Authorization Request/Policy (see below).
-Please see API description of SecAttr and MessageAuth classes in \cite{hed-api} for more information.
-
-Each MCC developed inside and outside the HED must come with explanation of generated and consumed Security Attributes.
-Dedicated components for dealing with Security Attribute - Security Handlers are described below.
-
-\subsubsection{Security Handlers}
-
-\texttt{Relevant classes: ArcSec::SecHander, Arc::MCC}
-
-Each MCC can also be configured to have loadable modules called Security Handlers attached to it in order to enforce security policies such as authentication and authorization or to assist such activities by gathering specialized security related information into Security Attributes.
-There is no strict distinction of capabilities between Security Handlers and MCCs.
-Both can and do populate Security Attributes.
-The distinction is more of logical nature.
-It also makes possible to have Security Handlers dealing with similar kind of information and capable of acting on different protocol levels.
-
-The Security Handlers are arranged into named queues.
-Elements in every queue are executed sequentially with the Message as the only argument.
-Different queues are executed at  different times.
-Which queue is executed in which case it depends on the MCC.
-Most MCCs implement two queues named "incoming" and "outgoing".
-The queue "incoming" is executed for Messages moving through the chain towards hosted application and passes the Message as argument with Payload of type corresponding to MCC type i.e.
-HTTP MCC passes Payload with parsed HTTP message.
-The "outgoing" queue is executed for Messages travelling to outside the HED.
-
-\subsection{Services and Clients}
-
-The services are dynamically loaded on start up just like the MCCs.
-They are almost identical to MCCs with an exception that they constitute the last link in the Message Chain.
-They are attached to the Chain in the same manner as other MCCs.
-But differently from MCC the \texttt{process()} method of the Service does not pass Message to other components of the Chain.
-Instead they have to process incoming Messages and produce outgoing ones.
-
-The clients are not represented by any specific component.
-The client code sees the Chain as single object with named entry points.
-Those entries are used by clients to insert request Message and get result Message on output.
-For simplifying the task of writing clients there is a library \texttt{arcclient} provided which wraps the task of creating Chains and Messages for the widely used SOAP, HTTP, TLS, TCP communications.
-
-\subsection{Plexer}
-
-\texttt{Relevant classes: Arc::Plexer }
-
-In general case multiple Services are living in the HED so the incoming Message should be routed to the proper Service.
-The Plexer MCC does this job.
-It takes the ENDPOINT attribute of the Message collected by other MCCs compares this attribute to regular expression defined in the configuration file and forwards the Message to the matching service.
-It acts as a dispatcher.
-The Plexer is also special in a sense that it is not a plugin but part of the \texttt{arcloader} library.
-
-The Plexer provides only basic functionality and is capable of doing only simple routing.
-But because each MCC has multiple routing capabilities it is possible to provide pluggable MCC implementing more sophisticated and/or more specific routing algorithms.
-
-\subsection{Error handling}
-
-\texttt{Relevant classes: Arc::MCC\_Status}
-
-For reporting errors each \texttt{process()} method returns instance of MCC\_Status class. That object carries predefined set of common error codes.
-
-This way for error reporting is mostly meant to be used for reporting problems related to code execution.
-For errors caused by processing the corresponding protocol MCC should generate proper response Message which carries error description.
-Only if protocol does not provide error handling MCC\_Status should be used.
-It is also advised to convert MCC\_Status error obtained from the next MCC in the Chain into protocol specific error Message if possible.
-
-\subsection{Instantiation of the Chain}
-
-\texttt{Relevant classes: Arc::Loader, Arc::Config, Arc::LoaderFactory, Arc::MCCFactory, Arc::ServiceFactory, Arc::DMCFactory, Arc::SecHandlerFactory, Arc::PDPFactory, Arc::ACCFactory}
-
-Chain instantiation is handled by Loader component.
-It takes XML configuration on input and then handles tasks of loading plugin libraries, identifying plugins in them, creating and linking objects of corresponding classes.
-
-Each Loader object may create multiple non-intersecting chains and there may be multiple Loader objects in same executable.
-For each component mentioned in the configuration the Loader creates single object of the corresponding class.
-On Loader destruction all handled components are destroyed too.
-
-The configuration of Loader is made up of the following elements (see also configuration schema at http://svn.nordugrid.org/trac/nordugrid/export/10146/arc1/trunk/src/hed/libs/loader/mcc.xsd):
-
-\begin{itemize}
-
-\item \texttt{Chain} is base element. It has no special meaning for configuration of chain and is only used to group other elements in a logical way. Chains may be nested.
-
-\item \texttt{Component} represents the MCC.
-Its attribute \texttt{name} defines the name of the MCC which has to be instantiated.
-This MCC is being looked up among those contained in modules loaded by \texttt{ModuleManager} element.
-If not found then the Loader tries to find a module with name corresponding to name of the MCC.
-Each MCC also must have unique identifier specified by \texttt{id} attribute.
-Subelement \texttt{next} specifies link to the next MCC in the Chain.
-It uses unique identifier specified by \texttt{id}.
-The \texttt{next} element may have an optional text which is used to distinguish different Message propagation path in MCC specific way.
-The \texttt{SecurityHandler} subelement defines the Security Handler and the queue to which it has to be attached.
-Also attribute \texttt{entry} may be used to create an entry point to the Chain.
-
-\item \texttt{Plexer} element instantiates Plexer component. Its configuration is similar to one of the \texttt{Component} elements.
-
-\item \texttt{Service} element defines plugin implementing final Component in the Chain - service. From Loader point of view its configuration is identical to that of the \texttt{Component}.
-
-\item \texttt{ModuleManager} defines parameters needed to find loadable modules containing plugins.
-Currently only subelement \texttt{path} is supported which defines path on file system where loadable modules may be found.
-
-\item \texttt{Plugins} specifies name of the loadable module.
-
-\end{itemize}
-
-\subsection{Sessions and Contexts}
-
-\texttt{Relevant classes: Arc::MessageContext, Arc::MessageContextElement, Arc::MessageAuthContext, Arc::ChainContext }
-
-The HED defines three lifetimes for operations happening inside the Chain and components which can be associated with them:
-
-%The HED provides few context containers associated with the Chain and the Message. Those are:
-
-\begin{enumerate}
-
-\item The Message lifetime - lasts as long as incoming and outgoing Messages are passing through the Chain forth and back.
-The Message itself in this case is used as container for associated components.
-
-\item The Session lifetime - is defined by existence of some logical connection between Messages being processed.
-Corresponding container MessageContext is normally created by the first MCC in a chain and attached to the Message.
-Actual lifetime of that container is MCC specific.
-For TCP MCC it corresponds to TCP connection.
-The MessageContext holds objects inherited from the MessageContextElement class.
-The Security Attributes can also be stored in dedicated container - MessageAuthContext - with Session lifetime.
-At the end of  the Session all associated objects are destroyed.
-
-\item The Chain lifetime - is the time period while component that made the Chain exists.
-This lifetime is represented by ChainContext class.
-Differently from other context objects this one does not allow free manipulation of contained objects.
-Instead, it provides an interface to some internal structures of the Loader object.
-Those include factories and lists of objects of all types created by particular instance of the Loader.
-
-\end{enumerate}
-
-\subsection{DMC}
-
-\texttt{Relevant classes: Arc::DataPoint, Arc::DataMover, Arc::DataBuffer and related classes.}
-
-The HED defines an interface for pluggable components implementing higher-level information transfer and query.
-Those are Data Management Components (DMC).
-Each DMC is inherited from Datapoint class and provides subset of methods for performing the following operations on data endpoint:
-
-\begin{enumerate}
-
-\item Read data from specified endpoint into DataBuffer class object
-
-\item Write data into specified endpoint from DataBuffer class object
-
-\item List subcontent of endpoint (i.e. list files in directory)
-
-\item Register and unregister presence of data - for indexing endpoints
-
-\item Resolve final or lower level location of data from stored metadata - for indexing endpoints
-
-\end{enumerate}
-
-The DMC may be implemented using third party software like it is currently done for (Grid)FTP DMC. But implementation may use Message Chains too like in case of HTTP MCC.
-
-Along with ordinary endpoints defining the location of data directly - like HTTP, FTP, LDAP - the DMC can be used with indirect/indexing endpoints.
-Those are endpoints which define only the location of meta-data associated with actual data or an interface/service providing functionality of managing/requesting data.
-For more information about indexing endpoints see description of supported URLs in \cite{urls}.
-
-More in depth technical information about the DMC can be found in \cite{dmc}.
-
-\subsection{Generic purpose components}
-
-The HED also includes vast amount of common purpose components:
-
-\begin{itemize}
-
-\item XMLNode is class for managing parsed XML structures. It provides minimal functionality for operating on XML elements, attributes and namespaces.
-
-\item Thread management functions provide few mostly used functions for thread creation.
-
-\item URL class gives access to various parts of URL. It has support for many specific URLs built-in.
-
-\item String utilities for easy conversion between various types to strings and back.
-
-\item Run class for starting, communicating and monitoring external processes.
-
-\item Logger class provides multilevel controllable logging functionality.
-
-\item Time and Period classes allows parsing and generating textual description of times, dates and time periods in various formats.
-
-\item RegularExpression class providing C++ wrapper for regexp related functions.
-
-\item Config class makes it possible to manipulate configuration file. And related classes give an access to specific parts of configuration.
-
-\item Counter and IntraProcessCounter provide a way to count abstract resources.
-
-\end{itemize}
-
-\subsection{Web Service related components}
-
-There is also a set of components implementing various Web Service related functionality. Those include:
-
-\begin{itemize}
-
-\item WSAEnpointReference and WSAHeader manipulate WS-Addressing \cite{ws-addr-soap} information in XML element and SOAP header correspondingly.
-
-\item SAMLToken, UsernameToken and X509Token are for consuming and generating token of same name according to various profiles of WS Security specifications \cite{ws-security}.
-
-\item WSRF and related and inherited classes offer a way to generate and analyze various Web Services Resource specification \cite{wsrf} related elements.
-
-\item DelegationConsumerSOAP, DelegationProviderSOAP and DelegationContainerSOAP implement web service interface which enables client to delegate X509 credentials to service.
-
-\item InformationInterface and related classes may be used to implement information interface service and client part in a way common for services built on top of HED.
-Also InfoCacheInterface extends it with caching functionality and InfoRegister provides service registration ability.
-
-\end{itemize}
-
-\subsection{Daemon}
-
-All componets of the Hosting Environment are compiled into loadable libraries and may be used in various executables.
-But there is also a dedicated executable provided - \texttt{arched}.
-It accepts configuration file containing XML document and passes it to the Loader component.
-Then the Loader component takes care of loading all modules and instantiating all Message Chains.
-The \texttt{arched} also initializes the Logger component, configures it and directs its output to where it is specified or to the standard output.
-The \texttt{arched} normally runs as background process but can be run in foreground as well.
-For more information about \texttt{arched} capabilities please read its manual page.
-
-\subsection{Alternative implementation languages}
-
-In order to facilitate the development of services, API bindings for languages other than C++ are provided and some service development has already been done in Python language.
-Currently the only available language bindings are Python and Java.
-Currently it is possible to write only SOAP Service modules in Python and Java due to multiple inheritance limitation.
-
-% TODO more of it
-
-\chapter{Implemented elements}
-
-This chapter describes components which are implemented alongside with the HED infrastructure.
-Although strictly not belonging to the infrastructure this minimal set of components is necessary to make the infrastructure usable.
-
-\section{Implemented MCCs}
-
-\subsection{TCP MCC}
-
-\texttt{Plugin names: tcp.service, tcp.client}
-
-\texttt{Library name: \{lib\}mcctcp}
-
-\texttt{Security handler queues: incoming, outgoing}
-
-\texttt{Message attributes: TCP:HOST, TCP:PORT, TCP:REMOTEHOST, TCP:REMOTEPORT, TCP:ENDPOINT, ENDPOINT}
-
-% TODO: attributes and security attributes
-
-The server side TCP MCC in the HED is special in that it produces Messages by listening on a network socket rather than processing Messages from other MCCs.
-As such it spawns new thread for every new connection to handle Messages and their responses throughout the Message Chain.
-One could envision other MCCs having these properties but producing Messages from other sources such as e.g.
-UNIX sockets.
-
-\ifthenelse{\boolean{brieftechdesc}}{
-This MCC can be configured as to what port to listen to.
-}{}
-
-\ifthenelse{\boolean{verbtechdesc}}{
-This MCC can be configured with one or more \texttt{<tcp:Listen>} elements which in turn contain the elements \texttt{<tcp:Port>}, \texttt{<tcp:Interface>} and \texttt{<tcp:Version>}.
-The \texttt{<tcp:Port>} element is mandatory and should contain an integer corresponding to the TCP port to listen to.
-The \texttt{<tcp:Interface>} element is optional and is meant to identify the network interface to bind to.
-It is currently not used.
-The \texttt{<tcp:Version>} element is used to specify IP version.
-It is optional and should if present contain the single digit 4 or 6.
-
-The server side TCP MCC generates \texttt{PayloadStreamInterface} payload in the Message passed to the next MCC which can be used for communicating through open TCP channel.
-Currently it ignores any payload attached to the returned Message.
-That MCC also fills following Message Attributes in the incoming Message while passing to the next MCC:
-\begin{enumerate}
-
-\item TCP:HOST - IP address of local interface which was used to establish TCP connection
-
-\item TCP:PORT - local TCP port which was used for connection
-
-\item TCP:REMOTEHOST - IP address of contacting client
-
-\item TCP:REMOTEPORT - TCP port of contacting client
-
-\item TCP:ENDPOINT - URL-like combination of ://(TCP:HOST):(TCP:PORT)
-
-\item ENDPOINT - same as TCP:ENDPOINT
-
-\end{enumerate}
-}{}
-
-The client side TCP MCC performs TCP connection to host and port specified in <Host> and <Port> elements inside <Connect> element of the MCC configuration.
-Then all incoming Messages of \texttt{process()} method are transferred over TCP connection.
-Accepted Payload type of incoming Message is \texttt{PayloadRawInterface}.
-Returned Payload is of \texttt{PayloadStreamInterface} type.
-It represents established TCP connection and \texttt{may} be used by previous MCCs in chain for direct communication.
-It is still preferred to call \texttt{process()} method instead.
-
-\paragraph{Configuration schema}
-
-\begin{verbatim}
-<?xml version="1.0" encoding="UTF-8"?>
-<xsd:schema
-  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
-  xmlns="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
-  xmlns:arc="http://www.nordugrid.org/schemas/ArcConfig/2007"
-  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
-  elementFormDefault="qualified">
-
-    <xsd:simpleType name="Version_Type">
-        <!-- This element defines TCP/IP protocol version.  -->
-        <xsd:restriction base="xsd:string">
-            <xsd:enumeration value="4"/>
-            <xsd:enumeration value="6"/>
-        </xsd:restriction>
-    </xsd:simpleType>
-    <xsd:element name="Version" type="Version_Type"/>
-
-    <xsd:complexType name="Listen_Type">
-        <!--
-            This element defines listening TCP socket. If interface is missing socket
-            is bound to all local interfaces (not supported). There may be multiple Listen elements.
-        -->
-        <xsd:sequence>
-            <xsd:element name="Interface" type="xsd:string" minOccurs="0" maxOccurs="1"/>
-            <xsd:element name="Port" type="xsd:int" minOccurs="1" maxOccurs="1"/>
-            <xsd:element name="Version" type="Version_Type" minOccurs="0" maxOccurs="1"/>
-        </xsd:sequence>
-    </xsd:complexType>
-    <xsd:element name="Listen" type="Listen_Type"/>
-    <xsd:complexType name="Connect_Type">
-        <!--
-            This element defines TCP connection to be established to specified Host at specified Port.
-            If LocalPort is defined TCP socket will be bound to this port number (not supported).
-        -->
-        <xsd:sequence>
-            <xsd:element name="Host" type="xsd:string" minOccurs="1" maxOccurs="1"/>
-            <xsd:element name="Port" type="xsd:int" minOccurs="1" maxOccurs="1"/>
-            <xsd:element name="LocalPort" type="xsd:int" minOccurs="0" maxOccurs="1"/>
-        </xsd:sequence>
-    </xsd:complexType>
-    <xsd:element name="Connect" type="Connect_Type"/>
-</xsd:schema>
-\end{verbatim}
-
-\subsection{TLS MCC}
-
-\texttt{Plugin names: tls.service, tls.client}
-
-\texttt{Library name: \{lib\}mcctls}
-
-\texttt{Security handler queues: incoming, outgoing}
-
-\texttt{Message attributes: TLS:PEERDN, TLS:IDENTITYDN}
-
-% TODO: attributes and security attributes
-
-The server and client TLS MCCs provide transport level security (TLS) over any stream channel. Currently they interoperate well with TCP MCCs.
-
-The server side MCC accepts payload of type PayloadStreamInterface.
-It then creates own instance of object inherited from PayloadStreamInterface bound to initial payload and passes it to the next MCC.
-This object is maintained inside Message Context under name \texttt{tls.service} and is destroyed when Context becomes inactive.
-Currently this MCC does not expect any payload to be returned from the rest of the chain and passes no payload to previous MCC.
-
-The server MCC fills Message Attributes TLS:PEERDN and TLS:IDENTITYDN representing subjects of last certificate in client's certificate chain used for establishing secure connection and the subject of last certificate which is not a proxy certificate correspondingly.
-Use of those attributes is deprecated.
-It is advised to use Security Attributes instead.
-
-The client side MCC behaves in similar way.
-It also establishes PayloadStreamInterface type object linked to same type of payload of the next MCC.
-To obtain that last payload it makes a first call to the next MCC with payload of type PayloadRawInterface and then uses the returned payload - which is expected to be of PayloadStreamInterface type - to create its own payload object with streaming capabilities and returns it to previous MCC for further usage.
-
-Both client and server side MCCs are implemented using OpenSSL toolkit \cite{openssl} and use X.509 infrastructure \cite{x509} for establishing secure connection and may be configured to get private key, certificate or proxy credentials from files residing at local file systems.
-It is also possible to specify the location of Certification Authority certificate or to use all certificates located in specified directory.
-for more information see configuration schema with comments below.
-
-\paragraph{Configuration schema}
-
-\begin{verbatim}
-<?xml version="1.0" encoding="UTF-8"?>
-<xsd:schema
-  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
-  xmlns="http://www.nordugrid.org/schemas/ArcMCCTLS/2007"
-  xmlns:arc="http://www.nordugrid.org/schemas/ArcConfig/2007"
-  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCTLS/2007"
-  elementFormDefault="qualified">
-    <xsd:complexType name="CACertificatesDir_Type">
-      <xsd:simpleContent>
-        <xsd:extension base="xsd:string">
-          <xsd:attribute name="PolicyGlobus" type="xsd:boolean" use="optional" default="false"/>
-        </xsd:extension>
-      </xsd:simpleContent>
-    </xsd:complexType>
-    <!-- Location of private key.
-      Default is /etc/grid-security/hostkey.pem for service
-      and none for client. -->
-    <xsd:element name="KeyPath" type="xsd:string"/>
-    <!-- Content of private key - not supported -->
-    <xsd:element name="Key" type="xsd:string"/>
-    <!-- Location of public certificate.
-      Default is /etc/grid-security/hostcert.pem for service
-      and none for client. -->
-    <xsd:element name="CertificatePath" type="xsd:string"/>
-    <!-- Content of public certificate - not supported -->
-    <xsd:element name="Certificate" type="xsd:string"/>
-    <!-- Location of proxy credentials - includes certificates, key and
-      chain of involved certificates. Overwrites elements Key, KeyPath,
-      Certificate and CertificatePath.
-      Default is none for client and none for service. -->
-    <xsd:element name="ProxyPath" type="xsd:string"/>
-    <!-- Content of proxy credentials - not supported -->
-    <xsd:element name="Proxy" type="xsd:string"/>
-    <!-- Location of certificate of CA. Default is none. -->
-    <xsd:element name="CACertificatePath" type="xsd:string"/>
-    <!-- Content of certificate of CA - not supported -->
-    <xsd:element name="CACertificate" type="xsd:string"/>
-    <!-- Directory containing certificates of accepted CAs.
-      Default is /etc/grid-security/ . -->
-    <xsd:element name="CACertificatesDir" type="xsd:string"/>
-</xsd:schema>
-\end{verbatim}
-
-\subsection{HTTP MCC}
-
-\texttt{Plugin names: http.service, http.client}
-
-\texttt{Library name: \{lib\}mcchttp}
-
-\texttt{Security handler queues: incoming, outgoing}
-
-\texttt{Message attributes: HTTP:METHOD, HTTP:CODE, HTTP:REASON, HTTP:RANGESTART, HTTP:RANGEEND, HTTP:ENDPOINT, HTTP:*, ENDPOINT}
-
-% TODO: attributes and security attributes
-
-The server side HTTP MCC accepts Messages with \texttt{PayloadStreamInterface} payload and parses HTTP related information from it.
-Information from the HTTP header is added to the Message Attributes.
-The body of HTTP message is passed to the next MCC as \texttt{PayloadRawInterface} payload.
-In response this MCC expects also the Message with \texttt{PayloadRawInterface}.
-It is then prepended with HTTP response header and pushes it into initially provided stream channel.
-As an output it returns empty \texttt{PayloadRawInterface} payload.
-
-That MCC routes results to multiple next MCCs in the chain.
-For that it accepts only labeled <next> elements in the configuration.
-Label names are those of HTTP methods (uppercase).
-HTTP Messages will be routed to their destinations according HTTP method requested.
-
-That MCC also fills following Message Attributes in the incoming Message while passing to the next MCC:
-\begin{itemize}
-
-\item HTTP:METHOD - HTTP method as defined in HTTP request header.
-
-\item HTTP:RANGESTART - Range request start offset.
-
-\item HTTP:RANGEEND - Range request end offset.
-
-\item HTTP:ENDPOINT - URL or path as specified in HTTP request header.
-
-\item HTTP:* - Here * stands for any name. All HTTP options from HTTP request header are converted into Message Attributes named HTTP:\{option name\}.
-
-\item ENDPOINT - same as HTTP:ENDPOINT.
-
-\end{itemize}
-
-For outgoing Message server MCC converts all HTTP:* Message Attributes into corresponding HTTP response header attribute.
-
-The client side HTTP MCC will accept \texttt{PayloadRawInterface} payload as HTTP body and after prepending it with HTTP information passes to the next MCC also as \texttt{PayloadRawInterface}.
-It accepts \texttt{PayloadStreamInterface} in response and after processing passes \texttt{PayloadRawInterface} back through the chain.
-
-The client MCC also converts all HTTP:* Message Attributes into corresponding HTTP header options while creating HTTP request.
-If defined it uses HTTP:METHOD and HTTP:ENDPOINT as method and URL/path of HTTP request.
-If not specified parameters defined in the configuration are used.
-For HTTP response it sets HTTP:CODE and HTTP:REASON to response code and reason correspondingly.
-It also performs conversion from header options into HTTP:* Message Attributes.
-
-\paragraph{Configuration schema}
-
-\begin{verbatim}
-<?xml version="1.0" encoding="UTF-8"?>
-<xsd:schema
-  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
-  xmlns="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
-  xmlns:arc="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
-  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
-  elementFormDefault="qualified">
-
-    <!--
-        These elements define endpoint and HTTP method for client HTTP MCC.
-    -->
-    <xsd:element name="Endpoint" type="xsd:string"/>
-    <xsd:element name="Method" type="xsd:string"/>
-</xsd:schema>
-\end{verbatim}
-
-\subsection{SOAP MCC}
-
-\texttt{Plugin names: soap.service, soap.client}
-
-\texttt{Library name: \{lib\}mccsoap}
-
-\texttt{Security handler queues: incoming, outgoing}
-
-\texttt{Message attributes: SOAP:ENDPOINT, ENDPOINT}
-
-% TODO: attributes and security attributes
-
-These MCCs convert payloads between data chunks presented by PayloadRawInterface type object into dedicated PayloadSOAP payloads and vice versa. Currently it has no specific configuration parameters.
-
-The server side MCC sets Message Attributes SOAP:ENDPOINT and ENDPOINT to URL present in \texttt{To} element of Web Service Addressing information stored in SOAP header.
-For proper SOAP over HTTP binding it also sets \texttt{HTTP:Content-Type} Message Attribute to value \texttt{application/soap+xml} or \texttt{text/xml} for SOAP version 1.2 and 1.1 respectively for response Message.
-Also in case of SOAP fault response \texttt{HTTP:CODE} is set to \texttt{500} and \texttt{HTTP:REASON} to \texttt{SOAP FAULT}.
-
-The client side MCC for outgoing Message accepts \texttt{SOAP:ACTION} Message Attribute and uses it to fill \texttt{HTTP:Content-Type} or \texttt{HTTP:SOAPAction} depending on SOAP version.
-
-\section{Implemented Security Handlers}
-
-For more information about security related capabilities of the HED please see dedicated document at \cite{sec-arc1}.
-
-%\subsection{Simple List Auth}
-%
-%\texttt{Plugin name: arc.authz}
-%
-%\texttt{Library name: \{lib\}arcpdc}
-
-%\subsection{Username Token}
-%
-%\texttt{Plugin name: usernametoken.handler}
-%
-%\texttt{Library name: \{lib\}arcpdc}
-
-%\subsection{X.509 Token}
-%
-%\texttt{Plugin name: x509token.handler}
-%
-%\texttt{Library name: \{lib\}arcpdc}
-
-\section{Implemented DMCs}
-
-\subsection{File DMC}
-
-\texttt{Protocol name: file}
-
-This DMC implements access to local or remote mounted file systems. It allows reading and writing the content of file, listing the content of directories and checking the presence of the object.
-
-Name of plugin is \texttt{file} and it is located in \texttt{\{lib\}dmcfile} loadable module.
-
-\subsection{GridFTP/FTP DMC}
-
-\texttt{Protocol name: ftp, gsiftp}
-
-This DMC implements access to a data server using FTP or GridFTP protocol.
-It allows reading and writing the content of stored file, listing the content of directories and checking the presence of the object.
-
-Name of plugin is \texttt{gridftp} and it is located in \texttt{\{lib\}dmcgridftp} loadable module.
-
-\subsection{HTTP DMC}
-
-\texttt{Protocol name: http, https, httpg (not implemented yet)}
-
-This DMC implements access to a data server using HTTP over TCP or over TLS protocol.
-It uses GET for reading and PUT for storing files.
-Data is always transferred in chunks due to current limitation of HTTP MCC.
-Listing is also supported.
-For that DMC extracts references from the \texttt{A} tags of the obtained HTML content.
-If retrieved content is not HTML or if content of its \texttt{TITLE} tag does not start from ""Index of /"" then only information about resquested URL itself is listed.
-
-Name of plugin is \texttt{http} and it is located in \texttt{\{lib\}dmchttp} loadable module.
-
-\subsection{LDAP DMC}
-
-\texttt{Protocol name: ldap}
-
-The LDAP DMC implements access to a data accessible through LDAP protocol.
-Currently it can only retrieve content of LDAP tree.
-Retrieved content is converted into XML.
-The DMC supports following kind of URL:
-
-\texttt{ldap://host[:port]/base[?[attributes][?[scope][?filter]]]}
-
-\begin{itemize}
-\item \texttt{attributes} is comma separated list of attributes which has to be retrieved,
-\item \texttt{scope} is either \texttt{base}, \texttt{one} or \texttt{sub} as defined by LDAP,
-\item \texttt{filter} is LDAP filter.
-\end{itemize}
-
-Name of plugin is \texttt{ldap} and it is located in \texttt{\{lib\}dmcldap} loadable module.
-
-\subsection{LFC DMC}
-
-\texttt{Protocol name: lfc}
-
-The LFC DMC provides an access to LCG File Catalog service and supports file listing, URL resolution and (un)registration of file locations.
-It uses LFC protocol implementation provided by gLite middleware.
-
-Name of plugin is \texttt{lfc} and it is located in \texttt{\{lib\}dmclfc} loadable module.
-
-\subsection{RLS DMC}
-
-\texttt{Protocol name: rls}
-
-The RLS DMC provides an access to Replica Location Service and supports file listing, URL resolution and (un)registration of file locations.
-It uses RLS protocol implementation provided by Globus Toolkit.
-
-Name of plugin is \texttt{rls} and it is located in \texttt{\{lib\}dmcrls} loadable module.
-
-\chapter{Future work}
-
-Currently the interfaces in the HED have matured enough and stay quite stable. Most changes being introduced in the field of Security infrastructure.
-
-Most of the work has shifted to the development of higher level libraries and services based on the HED. But the HED also continiously being expanded by adding new plugins.
-
-There are also significant pieces of functionality not addressed in the HED yet like user-friendly configuration, efficient inter-service internal Message routing, etc.
-
-%\chapter{Conclusion}
-
-\bibliography{grid}
-
-\end{document}

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Tutorial on programming web services with ARC-1. This some interim state, uploaded to help our Luebeck-internal review as a start. An announcement to the list will follow next week.

\chapter{TLS echoservice}

TLS provides security and data integrity

Diffrent kind of security is intended;
\begin{enumerate}
 \item Integrity (data is not manipulated along the transmission)
 \item Confidentiality 
 \item Authenticity (Server and client sided)
 \item Authorization (Server and client sided)
\end{enumerate}

The first two goals can be reached by encrypting the channel/connection between client and server. An approach 
%sich bewähren / behaupten
which stand the test of time is to introduce a TLS (Transport Layer Security) also known as SSL (Secure Sockets Layer).
Authenticity, Confidentiality, Integrity
The client wants to reassure himself to send his vulnerable data only to the server who can prove its correct identity.
Furthermore a prove of authenticity is needed. % Sich versichern, dass das Gegenüber der ist, für welchen er sich ausgibt
It .... does fancy stuff ...
%
%
Authorization:  The server on the other side takes an interest in limitate its ressources to a close circle of users. Above all no foreign clients shall be able to use the services.\\

One can add the TLS in client and server by simply extending the arched configuration file and a small modification of the client.

clientCA   - Authority guarantee for identity
clientCERT - Certificate given
clientKEY  - Secret key of the client to read messages, encrypted by the public key in the certificate!

serverCA   - Authority guarantee for cert. CA is known by the server. If not, the encrypted cert can be read be its CA.
serverCERT - Certificate given by the CA. Contains a snippet which can be read by the Public key of the CA.
serverKEY  - Secret key of the client to read messages, encrypted by the public key in the certificate!

Server has got have:
Its  serverCERT serverKEY and some kind of root CA knowing the client
Client has got to have
Its  clientCERT clientKEY and some kind of root CA knowing the server

WARNING: If the programm fails, check if one of your certifacte is expired. If so, create new one using the script files in the directory 'certFactory'

For HED is build up to  % entsrprechend einer Mod Structur
be modular, the security can easily be extend to our given example.
Each Message Chain Component (MCC) or service has a common interface for implementing various pluggable components (plug-ins) called SecHandler. The SecHandler components provide a method for processing messages traveling through Message Chains of the HED. 
Each MCC or Service usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called “incoming” and “outgoing” respectively. All SecHandler components attached to the queue are executed sequentially. 
If any of them fails, message processing fails as well. 
Each SecHandler is configured inside the arched configuration file used for configuring whole chain of MCCs 
Some of the currently implemented SecHandler components make use of pluggable and configurable sub-modules called Policy Decision Point (PDP). PDP can process ARC specific Request and Policy documents which are as well written in XML format~\cite{QIANG_2008}..

\lstsetCPP
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/clients/tlsechoclient/tlsechoclient.cpp}

\lstsetCPP

\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/tlsechoservice/tlsechoservice.cpp}

\lstsetARCHEDXML
\begin{minipage}[t]{\textwidth}
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/tlsechoservice/arched_tls_echoservice.xml}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ rm -f /var/log/arched.log
$ arched -c arched_echoservice.xml  && echo jo ||echo n
$ tail -n100 -f /var/log/arched.log
$ killall arched
\end{lstlisting}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ ./echoclient
Tue Feb 17 17:04:08 2009
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
	<soap-env:Body>
		<echoRequest:echo>
			<echo:say>Der_Berg_ruft</echo:say>
		</echoRequest:echo>
	</soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <echo:echoResponse>
      <echo:hear>[ Der_Berg_ruft ]</echo:hear>
    </echo:echoResponse>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

### 2009-02-19 doc/ws-programming-tutorial/src/clients/timeclient/timeclient.cpp

@@ -1,0 +1,83 @@
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
+#include <arc/ArcLocation.h>
+#include <arc/message/MCC.h>
+#include <arc/client/ClientInterface.h>
+#include <arc/client/UserConfig.h>
+
+#include <iostream>
+#include <string>
+
+using namespace Arc;
+
+int main(int argc, char** argv) {
+
+	// Initiate the Logger and set it to the standard error stream
+	Arc::Logger logger(Arc::Logger::getRootLogger(), "arcecho");
+	Arc::LogStream logcerr(std::cerr);
+	Arc::Logger::getRootLogger().addDestination(logcerr);
+	Arc::Logger::rootLogger.setThreshold(Arc::WARNING);
+
+	// Set the ARC installation directory
+	Arc::ArcLocation::Init("/usr/lib/arc"); 
+	setlocale(LC_ALL, "");  
+
+	// URL to the endpoint
+	std::string urlstring("http://localhost:60000/time");
+	printf("URL: %s\n",urlstring.c_str());
+
+
+	// Load user configuration (default ~/.arc/client.xml)
+	std::string conffile; 
+	Arc::UserConfig usercfg(conffile);
+	if (!usercfg) {
+		printf("Failed configuration initialization\n");
+		return 1;
+	}
+
+	// Basic MCC configuration
+	Arc::MCCConfig cfg;
+	Arc::URL service(urlstring);
+
+	// Creates a typical SOAP client based on the 
+	// MCCConfig and the service URL
+	Arc::ClientSOAP client(cfg, service);
+
+	// Defining the namespace and create a request
+	Arc::NS ns("time", "urn:time"); 
+	Arc::PayloadSOAP request(ns);
+  	request.NewChild("time").NewChild("timeRequest") = "";
+
+	//std::string xml;
+	//request.GetXML(xml, true);
+	//printf("Request message:\n%s\n\n\n", xml.c_str());
+
+	// Process request and get response
+	Arc::PayloadSOAP *response = NULL;
+	Arc::MCC_Status status = client.process(&request, &response);
+
+	// Test for possible errors
+	if (!status) {
+		printf("Error %s",((std::string)status).c_str());
+		if (response)
+			delete response;
+		return 1;
+	}
+	if (!response) {
+		printf("No SOAP response");
+		return 1;
+	}
+
+	//response->GetXML(xml, true);
[…]

### 2009-02-19 doc/ws-programming-tutorial/src/services/timeservice/timeservice.cpp

@@ -1,0 +1,104 @@
+#include <arc/loader/Plugin.h>
+#include <arc/message/PayloadSOAP.h>
+
+#include <stdio.h>
+#include <time.h>
+
+#include "timeservice.h"
+
+/**
+ * Within the following variable PLUGINS_TABLE_NAME a pointer to a
+ * method with a certain signature has to be stored.
+ *
+ * This method has to be capable to process the argument PluginArgument 
+ * and must return a pointer to a plugin, which is actually the timeservice.
+ * 
+ * Therefore the method initializes the time service and returns it.
+ */
+static Arc::Plugin* get_service(Arc::PluginArgument* arg) 
+{
+	// The dynamic cast can handle NULL
+	Arc::ServicePluginArgument* mccarg = dynamic_cast<Arc::ServicePluginArgument*>(arg);
+	if(!mccarg) return NULL;
+
+	return new ArcService::TimeService((Arc::Config*)(*mccarg));
+}
+
+
+/**
+ * This PLUGINS_TABLE_NAME is defining basic entities of the implemented .
+ * service. It is used to get the correct entry point to the plugin.
+ */
+Arc::PluginDescriptor PLUGINS_TABLE_NAME[] = {//(@*\label{lst_code:time_cpp_ptn}*@)
+	{
+		"time",					/* Unique name of plugin in scope of its kind */
+		"HED:SERVICE",			/* Type/kind of plugin */
+		1,						/* Version of plugin (0 if not applicable) */
+		&get_service 			/* Pointer to constructor function */
+	},
+	{ NULL, NULL, 0, NULL }		/* The array is terminated by element */
+								/* with all components set to NULL*/
+};
+
+
+using namespace Arc;
+
+namespace ArcService
+{
+
+	/**
+	* Constructor. Nothing to be done here except calling the 
+	* super constructor and initializing the logger.
+	*/
+	TimeService::TimeService(Arc::Config *cfg) //(@*\label{lst_code:time_cpp_constructor}*@) 
+			: Service(cfg),logger(Logger::rootLogger, "Time")
+	{ 
+	}
+
+	/**
+	* Deconstructor. Nothing to be done here.
+	*/
+	TimeService::~TimeService(void)//(@*\label{lst_code:time_deconstructor_ptn}*@)
+	{
+	}
+
+	/**
+	* Implementation of the virtual method defined in MCCInterface.h (to be found in MCC.h). 
+	*
[…]

### 2009-02-19 doc/ws-programming-tutorial/src/services/timeservice/timeservice.h

@@ -1,0 +1,64 @@
+/**      
+ * Implementation of a simple time service
+ *
+ * The time service will always reply with a message which contains
+ * the current time of the service.
+ *
+ */
+
+#ifndef __ARCSERVICE_TIME_H__
+#define __ARCSERVICE_TIME_H__
+
+#if HAVE_CONFIG_H
+	#include <config.h>
+#endif
+
+#include <arc/message/Service.h>
+#include <arc/Logger.h>
+
+namespace ArcService
+{
+
+class TimeService: public Arc::Service
+{
+
+	protected:
+
+		/**
+		*  Arc-intern logger. Generates output into the file specified 
+		*  in the arched configuration file used to invoke arched services.
+		*/
+		Arc::Logger logger;
+
+		/**
+		* Class which specifies a XML namespace i.e. "echo".
+		* Needed to extract the content out of the incoming message
+		*/
+		Arc::NS ns_;
+
+	public:
+
+		/**
+		* Constructor getting the Arc-configuration.
+		*/
+		TimeService(Arc::Config *cfg);
+
+		/**
+		* Destructor.
+		*/
+		virtual ~TimeService(void);
+
+		/**
+		* Implementation of the virtual method defined in MCCInterface
+		* (to be found in MCC.h). 
+		* @param inmsg incoming message
+		* @param inmsg outgoing message
+		* @return Status of the result achieved
+		*/
+		virtual Arc::MCC_Status process(Arc::Message& inmsg, Arc::Message& outmsg);
+
+}; 
+
+} //namespace ArcService
+
+#endif 

### 2009-02-19 doc/ws-programming-tutorial/src/services/tlsechoservice/certFactory/createMasterCA.sh

@@ -1,0 +1,77 @@
+#!/bin/sh
+#
+#     Script to create a new self-signed certifacte authority. 
+#
+# BASED ON
+#    http://www.dylanbeattie.net/docs/openssl_iis_ssl_howto.html
+
+echo ' This script will create a new self-signed certificate authority.'
+echo ' Following files will be created (or overwritten):'
+echo ' - database'
+echo ' - serial'
+echo ' - slaves/'
+echo ' - requests/'
+echo ' - keys/ca,key'
+echo ' - certs/ca.cer'
+echo ' - certs/ca.pem'
+echo ''
+echo 'WARNING: The old CA and all child certificate will be replaced completly !'
+echo ''
+echo    '         Press Strg+C to abort!'
+echo ''
+read -p "         Press a key to proceed"
+
+echo ''
+echo ''
+echo ' Removing directories...'
+echo '   keys/'
+rm -f -r keys
+
+echo '   certs/'
+rm -f -r certs
+
+echo '   slaves/'
+rm -f -r slaves
+
+echo '   requests/'
+rm -f -r requests
+
+
+echo ' Remove files...'
+echo '   database.*'
+rm -f database.*
+
+echo '   serial.*'
+rm -f serial.*
+
+
+echo ' Create new directories...'
+echo '   keys/'
+mkdir keys
+echo '   certs/'
+mkdir certs
+
+
+echo ' Create files...'
+echo '   database.txt'
+touch database.txt
+
+echo '   serial.txt'
+echo '01' > serial.txt
+
+
+
+echo ' Create private and public key for CA (keys/ca.key).'
+openssl genrsa -des3 -out keys/ca.key 1024
+
+echo ' Create certificate for the CA (certs/ca.cer).'
+openssl req -config openssl.conf -new -x509 -days 1001 -key keys/ca.key -out certs/ca.cer
+
+
+#openssl x509 -in certs/ca.cer -out certs/cax509.cer
+echo ' Convert certificate into PEM format (certs/ca.pem).'
+openssl x509 -in certs/ca.cer -out certs/ca.pem -outform PEM
+
+
+
+

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Pushed the state of the chapter time service a little bit further - added a comment in introduction - modified some source code to have LaTeX labels

@@
 According to this manner the message will be passed to higher layers until it reaches the \textit{Plexer}  Depending on the path of the URL which was demanded by the HTTP the Plexer multiplexes the request to the right Web Service.
-The path declared within the \textit{next} tag is a regular expression and leads to the service with the id named \textit{echo}.
+The path is declared within the \textit{next} tag is a regular expression and leads to the service with the id named \textit{echo}.
+Which means that the service will be now reachable under URL \textit{http://localhost:60000/echo}.
 This service finally will process the message and create a new message which will be returned all the way back to the TCP layer.

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Pushed the state of the chapter time service a little bit further - added a comment in introduction - modified some source code to have LaTeX labels

@@
 	caption={[C++ implementation of the time service. Filename: timeservice.cpp]
-	\textbf{C++ implementation of the time service. Filename: timeservice.cpp\textcolor{white}{hmf}}}
+	\textbf{C++ implementation of the time service. Filename: timeservice.cpp}}
 	]
@@
 Within the line~\ref{lst_code:time_cpp_process_message_payload} the payload is passed to the message. 
-The message processing is now completed and the function returns its success status on line~\ref{lst_code:time_cpp_return}.\\
+The message processing is now completed and the function returns its success status on line~\ref{lst_code:time_cpp_return}.
+The source code has to be compiled to a dynamic library named \textit{libtimeservice.so}.\\
 
-After the service has been implemented, it needs to be loaded into the HED. A suitable HED configuration file is to be seen in Listing~\ref{lst:time_arched_xml}.
-
+After the service has been implemented and compiled, it needs to be loaded into the HED. A suitable HED configuration file is to be seen in Listing~\ref{lst:time_arched_xml}. 
 \lstsetARCHEDXML
-\begin{minipage}[t]{\textwidth}
+\begin{program}
 \lstinputlisting
@@
 	caption={[HED configuration for the time service. Filename: arched\_timeservice.xml]
-	\textbf{HED configuration for the time service. Filename: arched\_timeservice.xml\textcolor{white}{hmf}}}
+	\textbf{HED configuration for the time service. Filename: arched\_timeservice.xml}}
 	]
 {../src/services/timeservice/arched_timeservice.xml}
-\end{minipage}
+\end{program}
+A new path element has been introduced into the element \textit{ModuleMananger}, line~\ref{lst_code:arched_time_moduleManager}. It assigns the location to the compliled dynamic library. 
+On line~\ref{lst_code:arched_time_plugin} the library is explict mentioned to be loaded.
+Another important change has been done within the \textit{Plexer} element in line~\ref{lst_code:arched_time_plexer_next}.
+A rule has been introduced which redirects the request to the Time Web Service, if the path of URL is \textit{time}.\\
+
+\lstsetKSH
+\begin{program}
+\begin{lstlisting}[
+label=lst:invokation_arched_timeservice,
+caption={[Transformation in eine uniforme konzentrische Verteilung.]
+         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+$ arched -c arched_timeservice.xml && echo Daemon started || echo Daemon start failed
+\end{lstlisting}
+\end{program}
 
@@
 	[
-	label=lst:arcecho_arched_xml,
+	label=lst:arcecho_time_client_cpp,
 	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
@@
 {../src/clients/timeclient/timeclient.cpp}
-
-\lstsetKSH
-\begin{minipage}[t]{\textwidth}
-\begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ arched -c arched_timeservice.xml  && echo jo ||echo n
-\end{lstlisting}
-\end{minipage}
 
@@
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source,
+label=lst:arched_time_request,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source,
+label=lst:arched_time_response,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Wrote more text for the time client

@@
 	\centering
-	\includegraphics[width=10cm]{tex_introduction/HED.pdf}
-	\mycaption{Example of HED structure with an Arex service and a file service.\task{ Replace image by an easier one!} }{\task{Write more stuff}}
+	\includegraphics[width=14cm]{tex_introduction/HED.pdf}
+	\mycaption{Example of HED structure with an Arex service and a file service.\task{ Replace image by an easier one!} }{\task{Write more stuff}\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}}
 \label{fig:HED_internal}

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Wrote more text for the time client

@@
 
+To run the service again the arched command shown in~\ref{lst:arched_timeservice_ksh} has to be used.
 \lstsetKSH
@@
 \begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
+label=lst:arched_timeservice_ksh,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 $ arched -c arched_timeservice.xml && echo Daemon started || echo Daemon start failed
+Daemon started
 \end{lstlisting}
@@
 
-\task{Hardcoded URL!!!}
+\task{\#ifdef HAVE\_CONFIG\_H\#include <config.h>\#endif - habe ich entfernt... macht keinen Sinn, da keine weiteren Macros im Source Code enthalten sind... oder?}
+In the next step a simple client needs to be implemented whose task is to interprete the returned message of the client.
+The client needs to have the same protocol stack as the daemon but in contrast to it, it does not need the ability to be loaded as a plugin.
+As one will see, the creation of this protocol stack on client side is very simple.
+The source code of the client consists primarily of the main function and is to be seen in Listing~\ref{lst:time_client_cpp}.
+In line~\ref{lst_code:time_client_cpp_logger} the logger is initialized and redirected to the standard error stream.
+For the components of the protocol stack are likewise loaded dynamically, one need to specify the installation directory in line~\ref{lst_code:time_client_cpp_init} too.
+\task{remark in \textbf{introduction}, what the plexer is for. In our case: to have several services running on one port!!!!!!!!}
+The URL of the service is defined in line~\ref{lst_code:time_client_cpp_url}.
+It fits to the HED configuration file which was shown in Listing~\ref{lst:time_arched_xml}.
+The daemon is listing on port 60000 and is running on the localhost.
+The service itself uses HTTP and can be reached by using the path \textit{time}.
+Line~\ref{lst_code:time_client_cpp_config} loads the standard config file which is usually almost empty.
+As a result, no additional security procedures will later be linked into the client sided stack.
+The protocol stack is created in line~\ref{lst_code:time_client_cpp_client}.
+Thanks to the URL the information for the configuration can be completed.
+The constructor of the class \textit{ClientSOAP} creates a SOAP client which uses HTTP, addresses the port 60000 on localhost.
+If i.e.
+HTTPS would have been used, the stack would have been extended by a TLS layer.
+Now the client stack is ready, one can create a request to the service.
+In line~\ref{lst_code:time_client_cpp_message_reqeust} a message which corresponds to the Listing~\ref{lst:time_client_request} is created.
+The message gets processed by the command in line~\ref{lst_code:time_client_cpp_message_response}.
+If everything was fine, the message will finally be send to the standard out in line~\ref{lst_code:time_client_cpp_message_answer}.
 
@@
 	[
-	label=lst:arcecho_time_client_cpp,
+	label=lst:time_client_cpp,
 	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
@@
 {../src/clients/timeclient/timeclient.cpp}
-
-\lstsetKSH
-\begin{minipage}[t]{\textwidth}
-\begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ ./timeclient
-Wed Feb 18 11:20:30 2009
-\end{lstlisting}
-\end{minipage}
 
@@
 \begin{lstlisting}[
-label=lst:arched_time_request,
+label=lst:time_client_request,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \begin{lstlisting}[
-label=lst:arched_time_response,
+label=lst:time_client_response,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{minipage}
+
+\lstsetKSH
+\begin{minipage}[t]{\textwidth}
+\begin{lstlisting}[
+label=st:time_client_ksh,
+caption={[Transformation in eine uniforme konzentrische Verteilung.]
+         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+$ ./timeclient
+Wed Feb 18 11:20:30 2009
+\end{lstlisting}
+\end{minipage}

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Added some images for the TLS service

@@
 
+\begin{figure}[htb]
+	\centering%epstopdf async.eps
+	\subfloat[Texttext\label{fig:cgd}]
+		{\includegraphics[width=13cm]{tex_tls_echoservice/async.pdf}}\\
+	\subfloat[Texttext\label{fig:cgd}]
+		{\includegraphics[width=13cm]{tex_tls_echoservice/async2.pdf}}
+
+	\mycaption{specific}{general}
+\label{fig:HED_internal0}
+\end{figure}
+
+\begin{figure}[htb]
+	\centering%epstopdf certificates.eps
+ 	\includegraphics[width=13cm]{tex_tls_echoservice/certificates.pdf}
+	\mycaption{specific}{general}
+\label{fig:HED_internal2}
+\end{figure}
+
 clientCA   - Authority guarantee for identity
@@
 
+\begin{figure}[htb]
+	\centering%epstopdf verification.eps 
+	\includegraphics[width=13cm]{tex_tls_echoservice/verification.pdf}
+	\mycaption{specific}{general}
+\label{fig:HED_internal3}
+\end{figure}
 Server has got have:
@@
 
+\begin{program}
 \lstsetCPP
@@
 {../src/clients/tlsechoclient/tlsechoclient.cpp}
+\end{program}
 
+\begin{program}
 \lstsetCPP
-
 \lstinputlisting
@@
 {../src/services/tlsechoservice/tlsechoservice.cpp}
+\end{program}
 
 \lstsetARCHEDXML
-\begin{minipage}[t]{\textwidth}
+\begin{program}
 \lstinputlisting
@@
 {../src/services/tlsechoservice/arched_tls_echoservice.xml}
-\end{minipage}
+\end{program}
 
 \lstsetKSH
-\begin{minipage}[t]{\textwidth}
+\begin{program}
 \begin{lstlisting}[
@@
 \end{lstlisting}
-\end{minipage}
+\end{program}
 
 \lstsetKSH
-\begin{minipage}[t]{\textwidth}
+\begin{program}
 \begin{lstlisting}[
@@
 \end{lstlisting}
-\end{minipage}
+\end{program}
 
 \lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
+\begin{program}
 \begin{lstlisting}[
@@
 \end{lstlisting}
-\end{minipage}
+\end{program}

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Various changes to first section.

@@
 
-This guide gives a first inside in how to write a standalone webservice for KnowArc including a suited client.
-The practically oriented examples will be followed by explanations of the main architecture concepts.
-For being able to comprehense basic knowledge of the programming language C++, the file format XML and the shell is requried.
-Furthermore an installation of KnowArc 1 is needed.\\
+This document gently introduces to preparation of standalone webservices and cognate clients with the Advance Resource Connector (ARC).
+The reader is guided through a series of the practically oriented examples.
+Each are accompanied by explanations of the main concepts of the software architecture.
+No particular skills are required to follow the presented steps.
+It is however beneficial to have some basic comprehension of the programming language C++, the file format XML and any regular UNIX shell.
+To start, the user shall have an installation of ARC-1, as it is performed with current RPM or .deb packages\footnote{see http://wiki.nordugrid.org for explicit instructions for your setup}.
 
-The KnowArc server provides access to services which are used by clients. The services may cover computational tasks as well 
-as access to a database. % job submission
-Servers provide access to a certain service and are waiting for a client which demands it. The task of servers can be subdivided into:
+The ARC server provides access to services.
+These are executed on the same machine that the server is installed on.
+Some services, like the a-rex service, will delegate computational efforts to other machines
+via additional software like a local queueing system.
+But the a-rex service itself does not migrate away to other machines - at least this
+has not been implemented yet.
+Services may comprise computational tasks (e.g., some secret and strongly
+patented algorithm), allow arbitrary computations (like ARC's a-rex or
+Amazon's clouds) or provide access to other resources like a particular
+database or to arbitrary disk space (like ARC's hopi or bartender).
+
+The users of computational services are software tools, too.
+They are referred to as clients.
+It is those clients that perform the interactions with a human user.
+However, when a service requests a performance of another service at the same or another server, then this first service is in the role of a client, too.
+
+Servers provide access to a single service or a set of services.
+For ARC, there is one service, named Hosting Environment Demon (hed), more special than other services: hed organises the assignment of other services on the same server to the network.
+But this does not break the concept, it is only a commodity.
+If services are not busy reacting to a client's request, then they are waiting.
+The task of servers can be subdivided into:
 \begin{itemize}
@@
 
-In general the implementation of a client is easier to be done compared with the server. 
-KnowArc is designed as a middleware which encapsulates typical challenges of server-client infrastructures (security, exception handling, extensibility) and abstract from underlying architecture (heterogenous computer, computer location, protocols).
-The core of KnowArc is the HED (Hosting Environment Daemon). HED is model consisting of a variable amount of heavily configurable layers. 
+In general, the implementation of a client is easier that the implementation of the server. 
+ARC is designed as a middleware which encapsulates typical challenges of server-client infrastructures (security, exception handling, extensibility) and abstracts from underlying architecture (heterogeneity of computers, computer location, protocols).
+The core of KnowArc is the prior mentioned HED (Hosting Environment Daemon).
+To the administrator HED hed is mostly visible as the single binary that is started and by the configure file that describes the other services that shall be prepared by HED.
+Conceptionally, the HED determines how services shall be organised both as a principle, and for any given ARC-run server, HED determines the very explicit presence or absence of a service at a particular address.
+Its many configurable layers are described later in this tutorial.
 %
-In the present case the endpoint of the HED will be a SOAP based Web Service.
-These Web Services exchange messages enveloped in the well known XML format.
-KnowArc abstracts from this format but nevertheless it is useful to be aware of the underlying message transport mechanism.\\
+At the time of writing, the endpoint of the HED will be a SOAP based webservice.
+SOAP is a long established XML-based standard for web service communication.
+ARC abstracts from this format, but nevertheless it is useful to be aware of
+the underlying message transport mechanism.
+
 %which implicate several advantages.
@@
 
-The source code of the examples given in this tutorial are available in the directory \textit{src}.
-\task{All examples can be found in the source tree: URL}
+The source code of the examples given in this tutorial are available in the subdirectory \textit{src}\footnote{If this was not shipped with the document you are reading, then please inspect http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/ws-programming-tutorial}.
 
@@
 
-The HED which was already mentioned in the section above is an essential part of the ARC middleware.
+The HED which was already mentioned in the section above as an essential part of the ARC middleware.
 It provides hosting of various services at application level, as well as a number of modules~\cite{QIANG_2005}.

### 2009-02-19 doc/ws-programming-tutorial/doc/tex_titlepage/titlepage.tex
Various changes to first section.

@@
 {\centering \textsc{\Large Webservice programming Tutorial}
-\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{I am currently working on this document, but I appreciate all comments and suggestions.
-\\e-mail: glodek@inb.uni-luebeck.de} % DELETE ONLY THAT LINE!!!
+\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under continous development, but don't hesitate to send comments and suggestions to glodek@inb.uni-luebeck.de, I is much appreciated.} % DELETE ONLY THAT LINE!!!
 \Large \par}
@@
 %%%%---- A list of authors ----
-    {\centering \large Michael Glodek, Steffen M\"oller, ... \large \par}
+    {\centering \large Michael Glodek, Steffen M\"oller \par}
 

### 2009-02-23 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
I wrote more in for the time service and try to push the other services further

@@
 Servers provide access to a single service or a set of services.
-For ARC, there is one service, named Hosting Environment Demon (hed), more special than other services: hed organises the assignment of other services on the same server to the network.
+For ARC, there is one service, named Hosting Environment Demon (HED), more special than other services: HED organises the assignment of other services on the same server to the network.
 But this does not break the concept, it is only a commodity.
@@
 
-In general, the implementation of a client is easier that the implementation of the server. 
+In general, the implementation of a client is easier \task{that} the implementation of the server. 
 ARC is designed as a middleware which encapsulates typical challenges of server-client infrastructures (security, exception handling, extensibility) and abstracts from underlying architecture (heterogeneity of computers, computer location, protocols).
-The core of KnowArc is the prior mentioned HED (Hosting Environment Daemon).
-To the administrator HED hed is mostly visible as the single binary that is started and by the configure file that describes the other services that shall be prepared by HED.
+The core of Arc is the prior mentioned HED (Hosting Environment Daemon).
+To the administrator HED is mostly visible as the single binary that is started and configured by a file that describes services that shall be prepared by HED.
 Conceptionally, the HED determines how services shall be organised both as a principle, and for any given ARC-run server, HED determines the very explicit presence or absence of a service at a particular address.
@@
 
-The source code of the examples given in this tutorial are available in the subdirectory \textit{src}\footnote{If this was not shipped with the document you are reading, then please inspect http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/ws-programming-tutorial}.
+The source code of the examples given in this tutorial are available in the subdirectory \textit{src}\footnote{If this was not shipped with the document you are reading, then please inspect\\ http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/ws-programming-tutorial}.
 
@@
 
-The HED which was already mentioned in the section above as an essential part of the ARC middleware.
-It provides hosting of various services at application level, as well as a number of modules~\cite{QIANG_2005}.
+The HED which was already mentioned in the section above as an essential part of the ARC middleware provides hosting of various services at application level, as well as a number of modules.
 Figure~\ref{fig:HED_internal} illustrates a typical setup of a server sided HED.
@@
 %
-\begin{figure}
+\begin{figure}[htb]
 	\centering
@@
 It is stateless but offers several extensions for requests, header information and status codes.
-Furthermore HTTP enables the usage of Uniform Resource Locators (URL) which is used by KnowArc to multiplex between different services.
-The multiplexing layer needs to be defined within the HED and is called Plexer.
+Furthermore HTTP enables the usage of Uniform Resource Locators (URL) which is used by Arc to multiplex between different services.
+The multiplexing layer can also be defined within the HED and is called Plexer.
 Depending on the path of the URL the Plexer passes messages to a defined SOAP service which again passes the message to the Web Service.\\
@@
 A first impression how the HED can be configured shall be given in the following example which corresponds to the HED shown in figure~\ref{fig:HED_internal}.
-It uses the internal ARC Echo service.
+It uses the internal ARC echo service~\cite{QIANG_2005}.
 
-\subsection{The Arcecho service}
-\task{Maybe the word tag needs to be replaced by the word element...}
+\subsection{The arc echo service}
+
 \task{Where is the xsd file for HED configuration files??}
@@
 \lstsetARCHEDXML\\
-\begin{program}
-\centering
+
 \lstinputlisting
 	[
-	label=lst:arched_arcecho_xml,
+	label=lst:arched_arcecho_xml,float,
 	caption={[HED configuration. Filename: arcecho\_no\_ssl.xml]
@@
 {../src/services/arched_arcecho.xml}
-\end{program}
 The first line of the XML file contains the XML declartion.
@@
 \task{What is this ArcConfig for??} % configuration namespaces
-The \textit{Server} tag, to be found at line~\ref{lst_code:arched_arcecho_Server}, provides basic settings for the daemon such as location of the Pid-file or the Log-file.
+The \textit{Server} element, to be found at line~\ref{lst_code:arched_arcecho_Server}, provides basic settings for the daemon such as location of the Pid-file or the Log-file.
 The \textit{ModuleManager} holds the path to the plugin libraries.
@@
 % for the standard MCC are defined here
-The next set of tags are holding the name of the plugins to be loaded (line~\ref{lst_code:arched_arcecho_Plugins}).
-The names have to correspond to the name of the dynamic libraries within the path defined by the \textit{ModuleManager}.
-Finally the chain of the components are declared initiate with the tag \textit{Chain}.
-Each component inside the chain is represented by a tag called \textit{Component} which has at least a name attribute.
+The next set of elements are holding the name of the plugins to be loaded (line~\ref{lst_code:arched_arcecho_Plugins}).
+The names have to correspond to the name of the dynamic libraries within the path defined in the \textit{ModuleManager}.
+Finally the chain of the components are declared within the element \textit{Chain}.
+Each component inside the chain is represented by an element called \textit{Component} which has at least a name attribute.
 Its important that a plugin has been loaded which provides the functionality of it.
-Depending on the component certain additional tags may be defined inside of it.
+Depending on the component certain additional elements may be defined inside of it.
 \task{Internet Protocol is required.}
-The first component of the chain is the \textit{tcp.service}. As to be seen the port 60000 gets assigned for the server to listen to.
+The first component of the chain is the \textit{tcp.service}. As to be seen the port 60000 is assigned to the server to listen to.
 The stream which is received by this component is passed to the component defined by the \textit{next} tag.
-In that case the will be passed to the component with the \textit{id} atrribute named \textit{http}.
+In the present case it will be passed to the component with the \textit{id} atrribute named \textit{http}.
 According to this manner the message will be passed to higher layers until it reaches the \textit{Plexer}  Depending on the path of the URL which was demanded by the HTTP the Plexer multiplexes the request to the right Web Service.
-The path is declared within the \textit{next} tag is a regular expression and leads to the service with the id named \textit{echo}.
-Which means that the service will be now reachable under URL \textit{http://localhost:60000/echo}.
-This service finally will process the message and create a new message which will be returned all the way back to the TCP layer.
+The path, declared within the \textit{next} element, corresponds to a regular expression and leads to the service with the id named \textit{echo}.
+Hence, the service will now be reachable under URL \textit{http://localhost:60000/echo}.
+It will finally process the message and create a new message which will be returned all the way back to the TCP component.
 The Echo Web Service itself is provided by the ARC package.
-Additional fix parameters can be passed by assigning tags which have to be known to the Web Service.
-The tags \textit{echo:prefix} and \textit{echo:suffix} seen in this example are defining the characters which will enclose the echo  returned string.
+Additional fix parameters are passed by the elements \textit{prefix} and \textit{suffix}.
+They set the characters which will enclose the echoed string.
+%returned
 \task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}\\
 
-To create the daemon corresponding to the defined HED configuration file ARC-1 provides the program \textit{arched}.
-The manner of invokation can been seen in Listing~\ref{lst:arcecho_arched_invokation}.
+To create the daemon corresponding to the defined HED configuration file ARC provides the program called \textit{arched}.
+The manner of invokation is to been seen in Listing~\ref{lst:arcecho_arched_invokation}.
 \lstsetKSH
-\begin{program}
 \begin{lstlisting}[
-label=lst:arcecho_arched_invokation,
+label=lst:arcecho_arched_invokation,float,
 caption={[Invokation of the Arched Daemon.]
@@
 \end{lstlisting}
-\end{program}
+
 % $ rm -f /var/log/arched.log
@@
 %
-The command is followed by the parameter \textit{c} which assignes, that the next string contains the full path of configuration file.
+The name of the program is followed by the parameter \textit{c} which assignes, that the next string contains the full path of configuration file.
 Once the server is running, the client can be used as described in Listing~\ref{lst:arcecho_client_invokation}.
 \lstsetKSH
-\begin{program}
 \begin{lstlisting}[
-label=lst:arcecho_client_invokation,
+label=lst:arcecho_client_invokation,float,
 caption={[Usage of the Arc echo client.]
@@
 \end{lstlisting}
-\end{program}
-In this example the server and the client are running on the same computer. For debugging it is advisable to check the Log-file which was assigned inside the HED configuration file.\\
 
-In the following chapters own Web Services will be developed. Starting with the simple Time Web Service.
+In this example the server and the client are running on the same computer so that the hostname \textit{localhost} can be utilized. 
+For debugging it is advisable to check the Log-file which was assigned inside the HED configuration file.\\
+
+In the following chapters own Web Services will be developed. The first one will be the simple Time Web Service.

### 2009-02-23 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
I wrote more in for the time service and try to push the other services further

 
\chapter{Persistent Webservice}

export ARC\_BARTENDER\_URL

### 2009-02-23 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
I wrote more in for the time service and try to push the other services further

@@
 
+POLICY STUFF:
+copy the request out of the LOGGEr and use that to write your policy.xml
+%/C=DE/ST=Germany/L=L\xFCbeck City/O=Instite of Neuro- and Bioinformatics/OU=Mitarbeiter
+
+Deny-Overrides - this is default if no algorithm specified.
+...
+Administrator of service can configure Authorization SecHandler - arc.authz - for each MCC and Service
+and define reasonable and meaningful policy. While defining policy the administrator must take into
+account that the attributes defined in the policy should be already collected by previous components in a
+chain. For instance, policy with AttributeId “http://www.nordugrid.org/schemas/policy-
+arc/types/http/path” should not be configured inside SecHandler attached to MCCTLS.
+
 \lstsetCPP

### 2009-02-23 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
I wrote more in for the time service and try to push the other services further

@@
 
-The first implementation of a Web Service which will be presented in this tutorial realises a simple time service.
-In case the service receives a request, it will respond with a string containing the systemtime.
+The first implementation of a Web Service will be a simple time service. 
+In case that the service receives a request of a client, it will respond with a string containing the system time.
 For simplicity the service will not parse the request but always return the current time.
-In the following subsections the C++ implementation first the service and than the client will be developed.
+In the following subsections the implementation of the service and the client will be presented.
 
@@
 
-In Arc-1 the services are implementated as plugins such that they can easily be included by modifiying the HED configuration file as it was to be seen in Listing~\ref{lst:arched_arcecho_xml}.
-The Listing~\ref{lst:time_arched_cpp}.shows the implementation of the time service plugin.
-The proper class definition is defined in the corresponding header file which isn't shown here.
-The key element for each plugin is the struct PluginDescriptor named PLUGINS\_TABLE\_NAME, see line\ref{lst_code:time_cpp_ptn}. 
+In Arc the services are implementated as plugins such that they can easily be included or excluded by modifiying the HED configuration file as it was to be seen in Listing~\ref{lst:arched_arcecho_xml}.
+Listing~\ref{lst:time_arched_cpp} contains the C++ implementation of the time service plugin.
+The proper class definition of the object \textit{TimeService} is defined in the corresponding header file which is not shown here.
+It inherits from an object called \textit{Service} defined in the Arc library.
+The key element for each plugin is the struct \textit{PluginDescriptor} named \textit{PLUGINS\_TABLE\_NAME} to be seen in line~\ref{lst_code:time_cpp_ptn}. 
 %The struct defines basic entities of the service and provides a pointer to the method which is able to create an instance of the time service plugin.
-Within the struct a unique plugin name, the type of the plugin, its version and the pointer to the function which returns the time service instance has to be defined.
-The function which returns the pointer to an instance of the service needs to have the PluginArgument as a parameter. 
-The actual Time Web Service is implemented within the namespace ArcService.
-The constructor can be found in line~\ref{lst_code:time_cpp_constructor}, the deconstructor in line~\ref{lst_code:time_cpp_deconstructor} and the method required by the inherited class \textit{Service} file in line~\ref{lst_code:time_cpp_process}.
-For the Time service is very simple, only the function process needs to be implemented.
-It gets the incoming and the outgoing message as references and returns a class representing the status the service achieved.
-\lstsetCPP
-\lstinputlisting
-	[
+Within the struct a unique plugin name, the type of the plugin, its version and a pointer to a function which returns the instance of the time service has to be defined.
+The function which returns the pointer of the instance is desclared in line~\ref{lst_code:time_cpp_get_service}. 
+The interface demands an object \textit{PluginArgument} as a parameter but because we are writing a plugin which inherits from \textit{Service}, we are able to expect the \textit{PluginArgument} to be a \textit{ServicePluginArgument}.
+%	In the first line of the service the object is downcasted with the operator dynamic_cast
+The \textit{ServicePluginArgument} provides access to the HED configuration and the message chain of the service.
+%       operator Config* (void) { return config_; };
+%       operator ChainContext*   			an interface to manipulate chains stored in Loader
+%
+%
+%
+%
+The time service uses the namespace \textit{ArcService}. The constructor can be found in line~\ref{lst_code:time_cpp_constructor} and the deconstructor in line~\ref{lst_code:time_cpp_deconstructor}.%
+\lstsetCPP%
+\lstinputlisting[
 	label=lst:time_arched_cpp,
@@
 	\textbf{C++ implementation of the time service. Filename: timeservice.cpp}}
-	]
-{../src/services/timeservice/timeservice.cpp}
-The service first determines the current time and than creates a SOAP message using the object \textit{PayloadSOAP} on line~\ref{lst_code:time_cpp_process_payloadSOAP}.
-The class \textit{PayloadSOAP} is able to create a XML message which is conform to the SOAP protocol (The apperance of the generated request and response will be discussed later).
-I case of the presented time service the payload consists out of two elements nested into one another.
-The element \textit{text} holds the element child which again holds the string containing the current time.
-Within the line~\ref{lst_code:time_cpp_process_message_payload} the payload is passed to the message. 
-The message processing is now completed and the function returns its success status on line~\ref{lst_code:time_cpp_return}.
-The source code has to be compiled to a dynamic library named \textit{libtimeservice.so}.\\
+	]%
+{../src/services/timeservice/timeservice.cpp}%
+In the context of the simple service they don't contribute with any functionality.
+The method \textit{process} which is required by the inherited class \textit{Service} is defined in line~\ref{lst_code:time_cpp_process} and realises the desired service. 
+The parameters \textit{inmsg} and \textit{outmsg} are references to the incoming and outgoing messages.
+Furthermore the method \textit{process} returns the object \textit{MCCStatus} which is representing the result of the service achieved.
+Within the method \textit{process}, the service first determines the current time and than creates a SOAP payload using the object \textit{PayloadSOAP} on line~\ref{lst_code:time_cpp_process_payloadSOAP}.
+The object \textit{PayloadSOAP} provides enough functionality to create a XML message which is conform to the SOAP protocol 
+(The apperance of the generated request and response will be discussed later on page~\pageref{lst:time_client_request}). 
+%
+%
+In the current case the payload consists of two elements nested into one another.
+The element \textit{time} holds the element \textit{timeResponse} which again holds the string containing the current time.
+Within the line~\ref{lst_code:time_cpp_process_message_payload} the payload is stored in the message. 
+In case nothing unexpected happened during the message is processed, the function returns with an \textit{MCCStatus} set to \textit{OK} as to be seen in line~\ref{lst_code:time_cpp_return}.\\
 
-After the service has been implemented and compiled, it needs to be loaded into the HED. A suitable HED configuration file is to be seen in Listing~\ref{lst:time_arched_xml}. 
+%The service has to be compiled into a dynamic library named \textit{libtimeservice.so}, in order
+%To run the service, one has to compile the source code into a dynamic library named \textit{libtimeservice.so}.\\
+%After the service is shipped into the library \textit{libtimeservice.so} it needs to be copied into the directory \textit{/tmp/arc/tutorial/lib} so it can be found by the \textit{ModuleManager} of the HED.
+
+In order to run the service, it has to be compiled into a dynamic library named \textit{libtimeservice.so}.
+The library has be to be copied into the directory \textit{/tmp/arc/tutorial/lib} such that it can be found by the \textit{ModuleManager} of the HED.
+A suitable HED configuration file which engages the time service is to be seen in Listing~\ref{lst:time_arched_xml}.A new path element has been introduced into the element \textit{ModuleMananger} in line~\ref{lst_code:arched_time_moduleManager}.
+It assigns the location of the created dynamic library.
+On line~\ref{lst_code:arched_time_plugin} the library is explict mentioned to be loaded as a plugin.
+Another important change has been done within the \textit{Plexer} element in line~\ref{lst_code:arched_time_plexer_next}.
+A rule has been introduced which redirects the request to the time service, if the path of URL is \textit{time}.\\
 \lstsetARCHEDXML
-\begin{program}
 \lstinputlisting
 	[
-	label=lst:time_arched_xml,
+	label=lst:time_arched_xml,float=htb,
 	caption={[HED configuration for the time service. Filename: arched\_timeservice.xml]
@@
 {../src/services/timeservice/arched_timeservice.xml}
-\end{program}
-A new path element has been introduced into the element \textit{ModuleMananger}, line~\ref{lst_code:arched_time_moduleManager}. It assigns the location to the compliled dynamic library. 
-On line~\ref{lst_code:arched_time_plugin} the library is explict mentioned to be loaded.
-Another important change has been done within the \textit{Plexer} element in line~\ref{lst_code:arched_time_plexer_next}.
-A rule has been introduced which redirects the request to the Time Web Service, if the path of URL is \textit{time}.\\
 
-To run the service again the arched command shown in~\ref{lst:arched_timeservice_ksh} has to be used.
+To start the service the \textit{arched} command which is shown in Listing~\ref{lst:arched_timeservice_ksh} has to be used.
+It is recommended to examine the log file \textit{/var/log/arched.log} if the HED configuration file has been modified in order to  assert the start was successful
+%To check the start of the HED it is recommended to examine the log file \textit{/var/log/arched.log}.\\
 \lstsetKSH
-\begin{program}
 \begin{lstlisting}[
-label=lst:arched_timeservice_ksh,
+label=lst:arched_timeservice_ksh,float=htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{lstlisting}
-\end{program}
 
@@
 
-\task{\#ifdef HAVE\_CONFIG\_H\#include <config.h>\#endif - habe ich entfernt... macht keinen Sinn, da keine weiteren Macros im Source Code enthalten sind... oder?}
+%\task{\#ifdef HAVE\_CONFIG\_H\#include <config.h>\#endif - habe ich entfernt... macht keinen Sinn, da keine weiteren Macros im Source Code enthalten sind... oder?}
 In the next step a simple client needs to be implemented whose task is to interprete the returned message of the client.
@@
 
+Once the HED is running, the service can
+
 The data is enveloped by an element called \textit{Envelope} .. See german wiki
@@
 \lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:time_client_request,
+label=lst:time_client_request,float,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{lstlisting}
-\end{minipage}
 
 \lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:time_client_response,
+label=lst:time_client_response,float,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{lstlisting}
-\end{minipage}
 

### 2009-02-23 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
I wrote more in for the time service and try to push the other services further

@@
 
+4 types:
+                <KeyPath>./clientKey.pem</KeyPath>
+                <CertificatePath>./clientCert.pem</CertificatePath>
+                <CACertificatePath>./serviceCA.pem</CACertificatePath>
+
 clientCA   - Authority guarantee for identity
@@
 
-\begin{program}
 \lstsetCPP
@@
 	[
-	label=lst:arcecho_arched_xml,
+	label=lst:arcecho_arched_xml,float,
 	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
@@
 {../src/clients/tlsechoclient/tlsechoclient.cpp}
-\end{program}
 
-\begin{program}
 \lstsetCPP
@@
 {../src/services/tlsechoservice/tlsechoservice.cpp}
-\end{program}
 
 \lstsetARCHEDXML
-\begin{program}
 \lstinputlisting
 	[
-	label=lst:arcecho_arched_xml,
+	label=lst:arcecho_arched_xml,float,
 	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
@@
 {../src/services/tlsechoservice/arched_tls_echoservice.xml}
-\end{program}
 
 \lstsetKSH
-\begin{program}
 \begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
+label=lst:invokation_arched_timeservice,float,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{lstlisting}
-\end{program}
 
 \lstsetKSH
-\begin{program}
 \begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
+label=lst:invokation_arched_timeservice,float,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{lstlisting}
-\end{program}
 
 \lstsetJUSTXML
-\begin{program}
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source,
+label=lst:timeservice_cpp_source,float,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{lstlisting}
-\end{program}

### 2009-02-23 doc/ws-programming-tutorial/src/clients/secechoclient/secechoclient.cpp

@@ -51,9 +51,9 @@ int main(int argc, char** argv) {
 		std::cout << "File " << argv[1] << " not found." << std::endl;
 		return -1;
 	}
-printf("sdfsdf");
+
 	std::string message(argv[2]);
-printf("sdfsdf");
+
 
 	// Initiate logger to get the output of ARC classes
 	Arc::Logger logger(Arc::Logger::getRootLogger(), "arcecho");
@@ -66,11 +66,10 @@ printf("sdfsdf");
 	std::string arclib("/usr/lib/arc");
 	Arc::ArcLocation::Init(arclib);  
 
-printf("sdfsdf");
 	// Loading client.xml file which is the counterpart of the arc 
 	// configuration file
 	Arc::XMLNode clientXml(xmlstring);
-printf("sdfsdf");
+
 	// Wrap the xml for hed client configuration
 	Arc::Config clientConfig(clientXml);
 	if(!clientConfig) {
@@ -78,7 +77,7 @@ printf("sdfsdf");
 	  return -1;
 	};
 
-printf("sdfsdf");
+
 	// Take XML configuration and perform configuration part
 	Arc::MCCLoader loader(clientConfig);
 	logger.msg(Arc::INFO, "Client side MCCs are loaded");
@@ -107,7 +106,9 @@ printf("sdfsdf");
 	reqmsg.Context(&context);
 	repmsg.Context(&context);  
 
-	request.NewChild("sececho:secechoRequest").NewChild("sececho:say") = message;
+	XMLNode sayNode = request.NewChild("secechoRequest").NewChild("say") = message;
+	sayNode.NewAttribute("operation") = "reverse";
+
 	reqmsg.Payload(&request);
 
 	request.GetXML(xml, true);
@@ -135,7 +136,7 @@ printf("sdfsdf");
 		return -1;
 	};
 
-	response->GetXML(xml);
+	response->GetXML(xml,true);
 	printf("Response:\n\n%s\n\n\n", xml.c_str());
 
 	std::string answer = (std::string)((*response)["secechoResponse"]["hear"]);//

### 2009-02-23 doc/ws-programming-tutorial/src/services/secechoservice/secechoservice.h

@@ -53,6 +53,9 @@ class SecEchoService: public Arc::Service
 		*/
 		Arc::MCC_Status make_fault(Arc::Message& outmsg);
 
+		#define ECHO_TYPE_ORDINARY "ordinary"
+		#define ECHO_TYPE_REVERSE  "reverse"
+
 	public:
 
         	/**

### 2009-02-23 doc/storage_system/arc-storage-manual.tex
arc storage manual skeleton

@@
 %%%%---- The title ----
-{\centering \textsc{\Large User's manual of the ARC storage system}\Large \par}
+{\centering \textsc{\Large Administrator's and user's manual of the ARC storage system}\Large \par}
 \vspace*{0.5cm}
@@
 \tableofcontents                          %Comment if use article style
+\begin{verbatim}
+    Administrator's and user's manual
+    ---------------------------------
+
+    Administrator's manual
+    - Quick start guide
+        - Checking the code out from the subversion
+        - Compiling and installing
+        - Modifying the configuration template
+        - Running the daemon
+        - Using the CLI tool
+    - Installation:
+        - Getting the source code from subversion
+        - Compilation details
+        - Installation
+    - Configuration:
+        - HED service config
+        - Centralized A-Hash
+        - Distributed A-Hash
+        - Librarian
+        - Shepherd
+        - Hopi
+        - Bartender
+        - Configuring inter-service trust: local listing of DNs, or using an A-Hash to store them
+        - Virtual organizations: how to configure HED to trust a VO
+    - Deployment scenarios:
+        - Single node: putting every service on one single node (including a storage node with Hopi)
+        - Centralized management: using seperate multiple storage nodes with Hopi, but leaving a central A-Hash, a Librarian and a Bartender on the central node
+        - Simple distribution: split the central node to three nodes, each of them has an A-Hash, a Librarian and a Bartender, and we have separate storage nodes
+        - One service per node: we have seperate nodes for all the services, any number of A-Hash nodes, any number of Librarian nodes, Bartender nodes, storage nodes
+
+    User's manual
+    - Overview: what do you need to access a running ARC storage system
+    - Prototype command line tool
+        - Configuring the command line tool
+        - Creating the root collection
+        - Uploading files
+        - Creating collections
+        - Moving files
+        - Creating hardlinks
+        - Removing files
+        - Removing collections
+        - Close collections
+        - Modify access policies
+        - Working with mount points
+    - Grid jobs
+        - Configuring the ARC storage DMC
+        - Specifying files in job descriptions
+    - Accessing the system with FUSE
+        - Configuring the FUSE module
+        - Mounting the storage system
+        - Supported operations
+        - Limitations
+
+\end{verbatim}
+
 \newpage
@@
 
-\chapter{Quick start guide} % (fold)
-\label{cha:quick_start_guide}
+\chapter{Administrator's manual} % (fold)
+\label{cha:administrator_s_manual}
 
-\section{Checking the code out from the subversion} % (fold)
-\label{sec:checking_the_code_out_from_the_subversion}
+\section{Quick start guide} % (fold)
+\label{sec:quick_start_guide}
 
-% section checking_the_code_out_from_the_subversion (end)
+\begin{verbatim}
+http://svn.nordugrid.org/
+svn co http://svn.nordugrid.org/repos/nordugrid/arc1/trunk arc1   
+README file for dependencies
+$ ./autogen.sh
+$ ./configure --disable-java --disable-a-rex-service --disable-isi-service --disable-charon-service --disable-compiler-service --disable-paul-service --disable-sched-service
+--prefix
 
-\section{Compiling and installing} % (fold)
-\label{sec:compiling_and_installing}
+Unit testing:       yes
+Java binding:       no
+Python binding:     yes (2.5)
 
-% section compiling_and_installing (end)
+Available third-party features:
 
-\section{Modifying the configuration template} % (fold)
-\label{sec:modifying_the_configuration_template}
+RLS:                yes
+GridFTP:            yes
+LFC:                no
+RSL:                yes
+SAML:               yes
+MYSQL CLIENT LIB:   no
+gSOAP:              no
 
-% section modifying_the_configuration_template (end)
+Included components:
+A-Rex service:      no
+ISI service:        no
+CHARON service:     no
+HOPI service:       yes
+SCHED service:      no
+STORAGE service:    yes
+PAUL service:       no
+SRM client (DMC):   no
+GSI channel (MCC):  yes
 
-\section{Running the daemon} % (fold)
-\label{sec:running_the_daemon}
+$ make
 
-% section running_the_daemon (end)
+$ sudo make install
 
-\section{Using the CLI tool} % (fold)
-\label{sec:using_the_cli_tool}
+\end{verbatim}
 
-% section using_the_cli_tool (end)
+% section quick_start_guide (end)
 
-% chapter quick_start_guide (end)
+\section{Installation} % (fold)
+\label{sec:installation}
 
-\chapter{Installing the system} % (fold)
-\label{cha:installing_the_system}
-
-\section{Getting the source code} % (fold)
-\label{sec:getting_the_source_code}
-
-% section getting_the_source_code (end)
+% section installation (end)
 
@@
 \label{sec:configuration}
-
-\subsection{HED service config} % (fold)
-\label{sub:hed_service_config}
-
-% subsection hed_service_config (end)
-
-\subsection{Centralized A-Hash} % (fold)
-\label{sub:centralized_a_hash}
-
-% subsection centralized_a_hash (end)
-
-\subsection{Distributed A-Hash} % (fold)
-\label{sub:distributed_a_hash}
-
-% subsection distributed_a_hash (end)
-
-\subsection{Librarian} % (fold)
-\label{sub:librarian}
-
-% subsection librarian (end)
-
-\subsection{Shepherd} % (fold)
-\label{sub:shepherd}
-
-% subsection shepherd (end)
-
-\subsection{Hopi} % (fold)
-\label{sub:hopi}
-
-% subsection hopi (end)
-
-\subsection{Bartender} % (fold)
-\label{sub:bartender}
-
-% subsection bartender (end)
-
-\subsection{Configuring inter-service trust} % (fold)
-\label{sub:configuring_inter_service_trust}
-
-% subsection configuring_inter_service_trust (end)
-
-\subsection{Virtual organizations} % (fold)
-\label{sub:virtual_organizations}
-
-% subsection virtual_organizations (end)
 
@@
 
-\section{Single node} % (fold)
-\label{sec:single_node}
+% chapter administrator_s_manual (end)
 
-% section single_node (end)
+\chapter{User's manual} % (fold)
+\label{cha:user_s_manual}
 
-\section{Centralized management} % (fold)
-\label{sec:centralized_management}
+\section{Overview} % (fold)
+\label{sec:overview}
 
-% section centralized_management (end)
-
-\section{Simple distribution} % (fold)
-\label{sec:simple_distribution}
-
-% section simple_distribution (end)
-
-\section{One service per node} % (fold)
-\label{sec:one_service_per_node}
-
-% section one_service_per_node (end)
-
-% chapter installing_the_system (end)
-
-\chapter{Using the system} % (fold)
-\label{cha:using_the_system}
+% section overview (end)
 
@@
 
-\subsection{Configuring the command line tool} % (fold)
-\label{sub:configuring_the_command_line_tool}
+% section prototype_command_line_tool (end)
 
-% subsection configuring_the_command_line_tool (end)
+\section{Grid jobs} % (fold)
+\label{sec:grid_jobs}
 
-\subsection{Creating the root collection} % (fold)
-\label{sub:creating_the_root_collection}
-
-% subsection creating_the_root_collection (end)
-
-\subsection{Uploading files} % (fold)
-\label{sub:uploading_files}
-
-% subsection uploading_files (end)
-
-\subsection{Creating collections} % (fold)
-\label{sub:creating_collections}
-
-% subsection creating_collections (end)
-
-\subsection{Moving files} % (fold)
-\label{sub:moving_files}
-
-% subsection moving_files (end)
-
-\subsection{Creating hardlinks} % (fold)
-\label{sub:creating_hardlinks}
-
-% subsection creating_hardlinks (end)
-
-\subsection{Removing files} % (fold)
-\label{sub:removing_files}
-
-% subsection removing_files (end)
-
-\subsection{Removing collections} % (fold)
-\label{sub:removing_collections}
-
-% subsection removing_collections (end)
-
-\subsection{Close collections} % (fold)
-\label{sub:close_collections}
-
-% subsection close_collections (end)
-
-\subsection{Modify access policies} % (fold)
-\label{sub:modify_access_policies}
-
-% subsection modify_access_policies (end)
-
-\subsection{Working with mount points} % (fold)
-\label{sub:working_with_mount_points}
-
-% subsection working_with_mount_points (end)
-
-% section prototype_command_line_tool (end)
+% section grid_jobs (end)
 
@@
 
-\subsection{Configuring the FUSE module} % (fold)
-\label{sub:configuring_the_fuse_module}
-
-% subsection configuring_the_fuse_module (end)
-
-\subsection{Mounting the storage system} % (fold)
-\label{sub:mounting_the_storage_system}
-
-% subsection mounting_the_storage_system (end)
-
-\subsection{Supported operations} % (fold)
-\label{sub:supported_operations}
-
-% subsection supported_operations (end)
-
-\subsection{Limitations} % (fold)
-\label{sub:limitations}
-
-% subsection limitations (end)
-
 % section accessing_the_system_with_fuse (end)
 
-% chapter using_the_system (end)
+% chapter user_s_manual (end)
 

### 2009-02-23 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Completed the content of time service (text isnot readable yet) - started content of echo service

@@
-\chapter{The echo webservice}
+\chapter{The echo service}
 
-\section{Webservice definition}
-wsdl
+The next service which will be presented is the echo service. It simply returns the received message back to the client.
+Depending on the operation requested by the client, the letters of the message will be reordered reverse or not.
+The goal of this chapter is to offer a deeper knowledge in message processing. Furthermore the \textit{WSDL} (Web Services Description Language) shall be introduced along with this example.
 
-In case of webservices, which are supported by the KnowArc architecture and used in the present tutorial, the interface of the requests and responses may be definied by a \textit{wsdl} file (WebService Definition Language).
+\section{Web Services Description Language}
+
+Web Services are often reachable for a large set of users in order to easily access complex information.
+In many cases the Web Service are capable to replace HTML pages i.e.
+access to databases like Pubmed or the NLM catalog.
+% http://www.ncbi.nlm.nih.gov/entrez/query/static/esoap_help.html
+In contrary to HTML the interfaces of the data are well defined and easier to find for machines. bla...
+Due to that reason, the way to access the web service has to be defined in an own language: \textit{WSDL} (Web Services Description Language).
+WSDL is a meta language which determines the structure of the messages along with its permitted elements, the accepted operations, the supported protocols and the address to reach the service. 
+The WSDL for the echo service which shall be implemented in this chapter is to be seen in Listing~\ref{lst:echo_wsdl}. There are five main elements to describe a service:
+\begin{itemize}
+	\item \textbf{Types} --- Definition of the data types used by the messages, line~\ref{lst_code:echo_wsdl_types}.
+	\item \textbf{Message} --- Assignments which data types are representing a message, line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
+	\item \textbf{Port type} --- \parbox[t]{13cm}{Assignment of the interface: One-way (input), request-response (input, output), solicit-response (input, output, fault), notification (output).
+The request-response interface is used in line~\ref{lst_code:echo_wsdl_portType}.}
+	\item \textbf{Binding} --- Defines the concret protocol and data format used for the message transmission, line~\ref{lst_code:echo_wsdl_binding}.
+	\item \textbf{Service} --- Specifies the address to bind a service, line~\ref{lst_code:echo_wsdl_service}.
+\end{itemize}
+%\textcolor{white}{newline}
+\lstsetJUSTXML
+\lstinputlisting
+	[
+	label=lst:echo_wsdl,
+	caption={[WSDL file describing the echo service. Filename: echo.wsdl]
+	\textbf{HWSDL file describing the echo service. Filename: echo.wsdl}}
+	]
+{../src/services/echoservice/echo.wsdl}
+
+More information about WSDL may be found at \href{http://www.w3.org/TR/wsdl}{http://www.w3.org/TR/wsdl}.
+
+% a valid service request and the
+
+% WSDL ist eine Metasprache, mit deren Hilfe die angebotenen Funktionen, Daten, Datentypen und Austauschprotokolle eines Web Service beschrieben werden können.
+Es werden im Wesentlichen die Operationen definiert, die von außen zugänglich sind, sowie die Parameter und Rückgabewerte dieser Operationen.
+Im Einzelnen beinhaltet ein WSDL-Dokument funktionelle Angaben zu:
+
+%    * der Schnittstelle
+%    * Zugangsprotokoll und Details zum Deployment
+%    * Alle notwendigen Informationen zum Zugriff auf den Service, in maschinenlesbarem Format
+
+%Nicht enthalten sind hingegen:
+
+%    * Quality-of-Service-Informationen
+%    * Taxonomien/Ontologien zur semantischen Einordnung des Services
+
+\section{Service}
+
+The service defined by the WSDL file in the previous section shall now be implemented.
+The source code of the C++ file is shown in Listing~\ref{lst:echo_service_cpp}.
+Again the header file will be set aside for it contains to much redundant information.
+
+\lstinputlisting
+	[
+	label=lst:echo_service_cpp,
+	caption={[C++ implementation of the echo service. Filename: echoservice.cpp]
+	\textbf{C++ implementation of the echo service. Filename: echoservice.cpp}}
+	]
+{../src/services/echoservice/echoservice.cpp}
+
+\lstsetARCHEDXML
+
+\lstinputlisting
+	[
+	label=lst:arcecho_arched_xml, float=htb,
+	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
+	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	]
+{../src/services/echoservice/arched_echoservice.xml}
+
+\lstsetKSH
+\begin{lstlisting}[
+label=lst:invokation_arched_timeservice,float=htb,
+caption={[Transformation in eine uniforme konzentrische Verteilung.]
+         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+$ arched -c arched_echoservice.xml  && echo jo ||echo n
+\end{lstlisting}
+
+\section{Client}
 
@@
 
-\lstinputlisting
-	[
-	label=lst:arcecho_arched_xml,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
-	]
-{../src/services/echoservice/echoservice.cpp}
-
-\lstsetARCHEDXML
-\begin{minipage}[t]{\textwidth}
-\lstinputlisting
-	[
-	label=lst:arcecho_arched_xml,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
-	]
-{../src/services/echoservice/arched_echoservice.xml}
-\end{minipage}
-
 \lstsetKSH
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
+label=lst:invokation_arched_timeservice, float=htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ rm -f /var/log/arched.log
-$ arched -c arched_echoservice.xml  && echo jo ||echo n
-$ tail -n100 -f /var/log/arched.log
-$ killall arched
+$ ./echoclient http://localhost:60000/echo ordinary text_to_be_transmitted
+[ text_to_be_transmitted ]
+$ ./echoclient http://localhost:60000/echo reverse text_to_be_transmitted
+[ dettimsnart_eb_ot_txet ]
 \end{lstlisting}
-\end{minipage}
-
-\lstsetKSH
-\begin{minipage}[t]{\textwidth}
-\begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ ./echoclient
-Tue Feb 17 17:04:08 2009
-\end{lstlisting}
-\end{minipage}
 
 \lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source,
+label=lst:timeservice_cpp_source, float=htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
   <soap-env:Body>
-		<echoRequest:echo>
-			<echo:say>Der_Berg_ruft</echo:say>
-		</echoRequest:echo>
+    <echo:echoRequest>
+      <echo:say operation="reverse">text_to_be_transmitted</echo:say>
+    </echo:echoRequest>
   </soap-env:Body>
@@
 \end{lstlisting}
-\end{minipage}
 
 \lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source,
+label=lst:timeservice_cpp_source, float=htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
     <echo:echoResponse>
-      <echo:hear>[ Der_Berg_ruft ]</echo:hear>
+      <echo:hear>[ dettimsnart_eb_ot_txet ]</echo:hear>
     </echo:echoResponse>

### 2009-02-23 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Completed the content of time service (text isnot readable yet) - started content of echo service

@@
 %returned
-\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}\\
+\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}\task{remark in \textbf{introduction}, what the plexer is for.
+In our case: to have several services running on one port!!!!!!!!}\\
 

### 2009-02-23 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Completed the content of time service (text isnot readable yet) - started content of echo service

@@
 %	In the first line of the service the object is downcasted with the operator dynamic_cast
-The \textit{ServicePluginArgument} provides access to the HED configuration and the message chain of the service.
+The \textit{ServicePluginArgument} provides access to the HED 
 %       operator Config* (void) { return config_; };
@@
 %
-The time service uses the namespace \textit{ArcService}. The constructor can be found in line~\ref{lst_code:time_cpp_constructor} and the deconstructor in line~\ref{lst_code:time_cpp_deconstructor}.%
+ \textcolor{white}{newline}
 \lstsetCPP%
@@
 {../src/services/timeservice/timeservice.cpp}%
+configuration and the message chain of the service.
+The time service uses the namespace \textit{ArcService}.
+The constructor can be found in line~\ref{lst_code:time_cpp_constructor} and the deconstructor in line~\ref{lst_code:time_cpp_deconstructor}.
 In the context of the simple service they don't contribute with any functionality.
+Only the constructor is defining the XML null-namespace for the time service.
 The method \textit{process} which is required by the inherited class \textit{Service} is defined in line~\ref{lst_code:time_cpp_process} and realises the desired service. 
@@
 %\task{\#ifdef HAVE\_CONFIG\_H\#include <config.h>\#endif - habe ich entfernt... macht keinen Sinn, da keine weiteren Macros im Source Code enthalten sind... oder?}
-In the next step a simple client needs to be implemented whose task is to interprete the returned message of the client.
-The client needs to have the same protocol stack as the daemon but in contrast to it, it does not need the ability to be loaded as a plugin.
-As one will see, the creation of this protocol stack on client side is very simple.
-The source code of the client consists primarily of the main function and is to be seen in Listing~\ref{lst:time_client_cpp}.
-In line~\ref{lst_code:time_client_cpp_logger} the logger is initialized and redirected to the standard error stream.
-For the components of the protocol stack are likewise loaded dynamically, one need to specify the installation directory in line~\ref{lst_code:time_client_cpp_init} too.
-\task{remark in \textbf{introduction}, what the plexer is for. In our case: to have several services running on one port!!!!!!!!}
-The URL of the service is defined in line~\ref{lst_code:time_client_cpp_url}.
-It fits to the HED configuration file which was shown in Listing~\ref{lst:time_arched_xml}.
-The daemon is listing on port 60000 and is running on the localhost.
-The service itself uses HTTP and can be reached by using the path \textit{time}.
-Line~\ref{lst_code:time_client_cpp_config} loads the standard config file which is usually almost empty.
-As a result, no additional security procedures will later be linked into the client sided stack.
-The protocol stack is created in line~\ref{lst_code:time_client_cpp_client}.
-Thanks to the URL the information for the configuration can be completed.
-The constructor of the class \textit{ClientSOAP} creates a SOAP client which uses HTTP, addresses the port 60000 on localhost.
-If i.e.
-HTTPS would have been used, the stack would have been extended by a TLS layer.
-Now the client stack is ready, one can create a request to the service.
-In line~\ref{lst_code:time_client_cpp_message_reqeust} a message which corresponds to the Listing~\ref{lst:time_client_request} is created.
-The message gets processed by the command in line~\ref{lst_code:time_client_cpp_message_response}.
-If everything was fine, the message will finally be send to the standard out in line~\ref{lst_code:time_client_cpp_message_answer}.
-
+This section is introducing the implementation of the client whose main task is to interprete the returned message.
+In order to be able to communicate with the service providing daemon the client needs to use an identic protocol stack.
+Likewise to the HED it will also be created dynamically using the MCCs.
+But as one will see, the creation of the protocol stack on client side is very simple.
+%
+%
+%
+The source code of the client is to be seen in Listing~\ref{lst:time_client_cpp} and consists basically of one main function.
+Within the line~\ref{lst_code:time_client_cpp_logger} the object \textit{Logger} gets initialized.
+Due to the configuration of the the \textit{Logger} in the next lines, the messages passed to it will be redirected to the standard error stream.
+%
+For the components of the protocol stack are likewise loaded \textcolor{white}{newline}
 \lstsetCPP
@@
 {../src/clients/timeclient/timeclient.cpp}
+ dynamically, one need to specify the installation directory of ARC in line~\ref{lst_code:time_client_cpp_init}.
+%
+The URL of the service is defined in line~\ref{lst_code:time_client_cpp_url}. It fits to the HED configuration file which was shown in Listing~\ref{lst:time_arched_xml}. 
+The daemon is listing on port 60000 and is running on the localhost. The service itself uses HTTP and can be reached by using the path \textit{time}.
+Line~\ref{lst_code:time_client_cpp_config} loads the standard config file which is usually almost empty.
+As a result, no additional security procedures will later be linked into the client sided stack.
+The protocol stack is created in line~\ref{lst_code:time_client_cpp_client}. 
+Thanks to the service URL the information for the configuration is complete.
+The constructor of the class \textit{ClientSOAP} creates a SOAP client which uses HTTP and addresses the port 60000 on localhost.
+If i.e.
+HTTPS would have been used, the stack would have been extended by an additional TLS protocol.
+Once the client stack is ready, the request to the service can be created which is done in line~\ref{lst_code:time_client_cpp_message_reqeust}.
+Even the apperance of the message is unimportant to the service, the message is composed of two convoluted elements \textit{time} and \textit{timeReqest}.
+%
+The request will be processed in the next line by the method \textit{process} of the object \textit{client}.%in line~\ref{lst_code:time_client_cpp_message_response}. 
+The object \textit{MCCStatus} which is returned by the method indicates the success of the processing.
+After verifying the response, the message will passed to the standard out in line~\ref{lst_code:time_client_cpp_message_answer}.\\
 
-Once the HED is running, the service can
-
-The data is enveloped by an element called \textit{Envelope} .. See german wiki
-
+The client has to be compilied to an ordinary binary so it can directly be executed from the command line.
+Assure the service is running and use the client as to be seen in Listing~\ref{lst:time_client_ksh}.
+\lstsetKSH
+\begin{lstlisting}[
+label=lst:time_client_ksh,float=htb,
+caption={[Invokation of the time client.]
+         \textbf{Invokation of the time client}}]
+$ ./timeclient
+Wed Feb 18 11:20:30 2009
+\end{lstlisting}
+The structure of the messages on SOAP level which are transmitted are shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
+The enclosing element is \textit{Envelope} which may contain an element \textit{Header} and must contain an element \textit{Body}.
+%The namespace \textit{soap-env} is defined by the element \textit{envelope}
+The data created by the object \textit{PayloadSOAP} is to be found within the \textit{Body}.
 \lstsetJUSTXML
 \begin{lstlisting}[
-label=lst:time_client_request,float,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-<soap-env:Envelope xmlns:time="urn:time" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
+label=lst:time_client_request,float=htb,
+caption={[Request message created by the client.]
+         \textbf{Request message created by the client}}]
+<soap-env:Envelope xmlns="urn:time" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
   <soap-env:Body>
@@
 \begin{lstlisting}[
-label=lst:time_client_response,float,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-<soap-env:Envelope xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
+label=lst:time_client_response,float=htb,
+caption={[Responce message of the service.]
+         \textbf{Responce message of the service}}]
+<soap-env:Envelope xmlns="urn:time" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
   <soap-env:Body>
@@
 \end{lstlisting}
-
-\lstsetKSH
-\begin{minipage}[t]{\textwidth}
-\begin{lstlisting}[
-label=st:time_client_ksh,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ ./timeclient
-Wed Feb 18 11:20:30 2009
-\end{lstlisting}
-\end{minipage}

### 2009-02-23 doc/ws-programming-tutorial/src/clients/echoclient/echoclient.cpp

@@ -1,7 +1,3 @@
-#ifdef HAVE_CONFIG_H
-#include <config.h>
-#endif
-
 #include <arc/ArcLocation.h>
 #include <arc/message/MCC.h>
 #include <arc/client/ClientInterface.h>
@@ -12,41 +8,69 @@
 
 using namespace Arc;
 
-int main(int argc, char** argv) {
+int main(int argc, char** argv)
+{
 
-  std::string urlstring("http://localhost:60000/echo");
-  std::string message("She's crazy like a fool");
+	/** Get arguments passed by the command line*/
+	if(argc != 4){
+		printf("Usage:\n");
+		printf("  echoclient url type message\n");
+		printf("\n");
+		printf("The first argument defines the URL of the service.\n");
+		printf("The second argument the type of echo which shall be performed (ordinary or reverse).\n");
+		printf("The third agrument holds the message to be transmitted.\n");
+		printf("\n");
+		printf("Example:\n");
+		printf("  ./echoclient http://localhost:60000/echo ordinary text_to_be_transmitted\n");
+		return -1;
+	}
+	
+	std::string urlstring(argv[1]);
+	std::string type(argv[2]);
+	std::string message(argv[3]);
+	/** */
 
-  // Initiate logger to get the output of ARC classes
-  Arc::Logger logger(Arc::Logger::getRootLogger(), "arcecho");
-  Arc::LogStream logcerr(std::cerr);
-  Arc::Logger::getRootLogger().addDestination(logcerr);
-  Arc::Logger::rootLogger.setThreshold(Arc::WARNING);
+	// Initiate the Logger and set it to the standard error stream
+	Arc::Logger::rootLogger.setThreshold(Arc::WARNING);
+	Arc::Logger logger(Arc::Logger::getRootLogger(), "echo");
+	Arc::LogStream logcerr(std::cerr);
+	Arc::Logger::getRootLogger().addDestination(logcerr);
+	Arc::Logger::rootLogger.setThreshold(Arc::WARNING);
 
 
-  // Set ARC installation directory for being able to 
-  // loaded modules dynamically
-  Arc::ArcLocation::Init("/usr/lib/arc");  
+	// Set the ARC installation directory
+	Arc::ArcLocation::Init("/usr/lib/arc");
+	setlocale(LC_ALL, "");   
 
-  // Creates the counterpart to the HED provided by the arched server.
[…]

### 2009-02-23 doc/ws-programming-tutorial/src/services/echoservice/echoservice.h

@@ -51,7 +51,10 @@ class EchoService: public Arc::Service
 		* @param outmsg outgoing message
 		* @return Status of the result achieved
 		*/
-		Arc::MCC_Status make_fault(Arc::Message& outmsg);
+		Arc::MCC_Status makeFault(Arc::Message& outmsg);
+
+		#define ECHO_TYPE_ORDINARY "ordinary"
+		#define ECHO_TYPE_REVERSE  "reverse"
 
 	public:
 

### 2009-02-23 doc/storage_system/arc-storage-documentation.tex
updating the arc storage documentation

@@
 The replicas of the files are stored on different storage nodes.
-A storage node here is a network-accessible computer having storage space to share, and one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.).
+A storage node here is a network-accessible computer having storage space to share, and a storage element service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.).
 For each storage node one of the services of the ARC storage system is needed to manage it and to integrate it into the system.
-And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC storage system.
+There is a way on the client side to access third-party storage solutions through the namespace of the ARC storage system.
 The main services of the storage system are the following (see Figure~\ref{fig:services}):
 \begin{itemize}
-    \item the \textbf{A-Hash} service, which is a currently centralized, later distributed database;
-    \item the \textbf{Librarian} service, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
+    \item the \textbf{A-Hash} service, which is a replicated database which is used by the Librarian to store metadata;
+    \item the \textbf{Librarian} service, which handles the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
     \item the \textbf{Bartender} service, which provides a high-level interface for the users and for other services;
-    \item the \textbf{Shepherd} service, which provides a simple interface for storing files on storage nodes.
+    \item the \textbf{Shepherd} service, which manages storage element services, and provides a simple interface for storing files on storage nodes.
 \end{itemize}
@@
 
-The descriptions in Chapter~\ref{cha:overview} and Chapter~\ref{cha:use_cases} are containing some elements which are not implemented yet in the first prototype.
-See Chapter~\ref{cha:technical_description_and_implementation_status} for the details about what is implemented and what is missing.
+\section{Files and collections} % (fold)
+\label{sec:files_and_collections}
 
-\section{IDs used in the system}
-\label{sec:IDs}
+The storage system is capable of storing files which can be grouped in collections and sub-collections, etc.
+Every file and collection has a unique ID in the sysem called the \emph{GUID}.
+Compared to the well-known structure of local file systems, these GUIDs are very similar to the concept of \emph{inode}s.
+And as a directory on a local filesystem is basically just a list of name and inode pairs, a collection on the ARC storage is just a list of name and GUID pairs.
+There is a dedicated collection which is the \emph{root collection}.
+This makes the namespace of the ARC storage system a hierarchical namespace where you can start at the root collection, and go to sub-collections and sub-sub-collections to get to a file.
+This path is called the \emph{Logical Name} (LN).
+For example if there is a sub-collection called \verb!saturn!
+in the root collection, and there is a file called \verb!rings!
+in this sub-collection, then the LN of this file is \verb!/saturn/rings!.
 
-There are a number of IDs used in the ARC storage system, such as:
-\begin{itemize}
-    \item Each file and collection has a globally unique ID called \textbf{GUID}.
-    \item The files and collections are organized into a global hierarchical namespace and can be referred to using paths of this namespace called \textbf{Logical Name}s (\textbf{LN}s).
-    \item Each service has a unique \textbf{serviceID} which can be used to get an endpoint reference from the information system.
-We need an endpoint reference which is an address (URL) which we could connect to.
-    \item The Shepherds in the system identify their files using a \textbf{referenceID}.
-    \item The \textbf{location} of a replica consists of two IDs: the ID of the Shepherd and the ID of the file within the Shepherd: (serviceID, referenceID).
-\end{itemize}
+Besides the Logical Names we can refer to a file or collection by simply its GUID, or we can use GUIDs and Logical Names together, as seen on Figure~\ref{fig:namespace}.
 
-% section IDs (end)
-
-\section{GUIDs and Logical Names (LNs)} % (fold)
-\label{sec:GUIDsLNs}
-
-The syntax of Logical Names is \verb#/[path]# or \verb#<GUID>[/<path>]# where [...] indicates optional parts.
-
-Each file and collection has a GUID which is globally unique, so they can be unambiguously referred using this GUID, that's why a single GUID is a Logical Name itself.
-
-A collection is basically just a list of entries, where each entry has a name unique within this collection, and an entry can be a file or a sub-collection, in which there are files and sub-sub-collections, etc.
+The full syntax of Logical Names is \verb#/[path]# or \verb#<GUID>[/<path>]# where [...] indicates optional parts.
 
@@
 
-Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file as \verb#/green/orange/huge# if we know the GUID of the starting collection, so let's prefix the path with it: \verb#1234/green/orange/huge#.
-This is the Logical Name of that file.
-If there is a well-known system-wide root collection (its GUID could be e.g.~\verb#0#), then if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g.~\verb#/why/blue# means \verb#0/why/blue#.
+Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file with the Logical Name \verb#1234/green/orange/huge#, which means that from the collection called \verb!1234!
+we have to follow along the path: \verb!green!, \verb!orange!, \verb!huge!.
 
-If a client wants to find the file called \verb#/why/blue#, the system knows where to start the search: the GUID of the root collection.
+There is a dedicated root collection (which has the GUID \verb#0#), and if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g.~\verb#/why/blue# means \verb#0/why/blue#.
+If a user wants to find the file called \verb#/why/blue#, the system knows where to start the search: the GUID of the root collection.
 The root collection knows the GUID of \verb#why#, and the (sub-)collection \verb#why# knows the GUID of \verb#blue#.
-If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, so it's a hard link.
-To count the references (the number of hard links) of a file or collection, the GUID of the parent collection(s) is stored as a metadata for each file and collection.
+If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, which concept is very similar to a hardlink in a regular local file system.
 
-One possible usage scenario would be if each VO (Virtual Organization)\footnote{Virtual Organization, \url{http://www.globus.org/grid\_software/security/voms.php}} would create a VO-wide root collection, and put it in the generic root collection, e.g.~if a VO called \verb#vo1# creates a collection called \verb#vo1# as a sub-collection of the root collection (which has the GUID \verb#0#), then it can be referred as \verb#0/vo1# or just \verb#/vo1#.
-Then this VO can create some files, and put them in this \verb#/vo1# collection, e.g.~\verb#/vo1/file1#, etc.
-Or sub-collections, e.g.~\verb#/vo1/col1#, \verb#/vo1/col2/file3#, etc.
-For this the VO does not need the install any service.
-These files and collections can be created using a Bartender service.
+% section files_and_collections (end)
 
-% section GUIDsLNs (end)
+\section{Storage nodes and replicas} % (fold)
+\label{sec:storage_nodes_and_replicas}
+
+The collections in the ARC storage are logical entities, the content of a collection is stored as metadata of the collection, which means the a collection actually has no physical data.
+A file however has both metadata and real physicial data (the actual bytes of the file).
+The metadata of a file is stored in the same database where the collections are stored, but the physical data of a file is stored on storage nodes as multiple replicated copies.
+
+A storage node consists of two things: a storage element service which is capable of storing and serving files through a specific protocol (e.g.
+a web server, an FTP server, a GridFTP server, etc.) and a Shepherd service which provides a simple interface to access the storage node, and which can initiate and manage file transfers through the storage element service.
+The Shepherd has different backends for the supported storage element services which made it possible the communicate with them.
+
+So we have logical files, which are part of the hierarchical namespace and have a GUID and other metadata, and a logical file has one or more physical replicas.
+The physical replicas stored on seperate storage nodes.
+In order to connect the logical file to its replicas, we need to have some pointers.
+Each storage node has a URL and each replica has a unique ID within the storage node called \emph{referenceID}, the URL and the referenceID together is called a \emph{Location}, a Location unambiguously points to one specific replica.
+So to connect the logical files to the physical ones, each logical file has a list of Locations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% section storage_nodes_and_replicas (end)
 

### 2009-02-24 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Started outline for the TLS echo service

@@
 
-The next service which will be presented is the echo service. It simply returns the received message back to the client.
-Depending on the operation requested by the client, the letters of the message will be reordered reverse or not.
-The goal of this chapter is to offer a deeper knowledge in message processing. Furthermore the \textit{WSDL} (Web Services Description Language) shall be introduced along with this example.
+The next service to be presented is the echo service which returns the received message back to the client.
+Furthermore the client has to choose between the operations ``ordinary'' or ``reverse''.
+Depending on that operation the message will be returned directly or its letters will be reordered reverse. % without manipulation
+The goal of this chapter is to offer a deeper knowledge about message processing. Furthermore the \textit{WSDL} (Web Services Description Language) shall be introduced along with this example.
 
@@
 In many cases the Web Service are capable to replace HTML pages i.e.
-access to databases like Pubmed or the NLM catalog.
+to access to databases like Pubmed or the NLM catalog.
 % http://www.ncbi.nlm.nih.gov/entrez/query/static/esoap_help.html
-In contrary to HTML the interfaces of the data are well defined and easier to find for machines. bla...
-Due to that reason, the way to access the web service has to be defined in an own language: \textit{WSDL} (Web Services Description Language).
-WSDL is a meta language which determines the structure of the messages along with its permitted elements, the accepted operations, the supported protocols and the address to reach the service. 
-The WSDL for the echo service which shall be implemented in this chapter is to be seen in Listing~\ref{lst:echo_wsdl}. There are five main elements to describe a service:
+In contrary to HTML which describes a human readable document the interfaces of Web Services are well defined and easier to interpret for machines.
+In general the client is not familiar with the functionality of the service.
+Due to that reason, the access to a web service can be defined in a \textit{WSDL} (Web Services Description Language) file.
+WSDL is a meta language which describes the structure of the messages, the intended use of the message, the supported protocols and the endpoint of the service (URL).
+The WSDL file which specifies the echo service of this chapter is to be seen in Listing~\ref{lst:echo_wsdl}. 
+It consisits of five main WSDL elements:
+\newcommand{\parboxWidth}{13cm}
 \begin{itemize}
-	\item \textbf{Types} --- Definition of the data types used by the messages, line~\ref{lst_code:echo_wsdl_types}.
-	\item \textbf{Message} --- Assignments which data types are representing a message, line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
-	\item \textbf{Port type} --- \parbox[t]{13cm}{Assignment of the interface: One-way (input), request-response (input, output), solicit-response (input, output, fault), notification (output).
-The request-response interface is used in line~\ref{lst_code:echo_wsdl_portType}.}
-	\item \textbf{Binding} --- Defines the concret protocol and data format used for the message transmission, line~\ref{lst_code:echo_wsdl_binding}.
-	\item \textbf{Service} --- Specifies the address to bind a service, line~\ref{lst_code:echo_wsdl_service}.
+	\item \textbf{Types} --- The element \textit{types} contains the description of the message structures which is written in the language XSD (XML Schema Definition).
+The applicable data types of the echo service are defined in  line~\ref{lst_code:echo_wsdl_types}.
+
+	\item \textbf{Message} --- Within the element \textit{message} a subset of the previously defined types are assigned to be a message.
+In case of the echo service two messages are designated: echoRequest and echo Response, see lines line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
+
+	\item \textbf{Port type} --- The element \textit{portType} is used to assign the messages to an interface.
+Four interface types can be distinguished: One-way (input), request-response (input, output), solicit-response (input, output, fault) and notification (output).
+	The echo service realises a simple request-response interface which is declared subsequent line~\ref{lst_code:echo_wsdl_portType}. % in the lines followed by the line X
+
+	\item \textbf{Binding} --- The protocol and the data format used for the message transmission are denoted in the element \textit{binding}.
+Possible style attributes are \textit{document} or \textit{rpc} (Remote Procedure Call).
+The transport attribute defines the protocol which is regularly HTTP.
+For each operation a corresponding SOAP action has to be defined.
+Additionally the input and output has to be specified to be either \textit{literal} or \textit{encoded}~\cite{BUTEK_2009}.
+The service implemented in this chapter will use the binding \textit{document}/\textit{literal} and the protocol HTTP which is to be seen in line~\ref{lst_code:echo_wsdl_binding}.
+
+% The style attribute can be "rpc" or "document". In this case we use document. The transport attribute defines the SOAP protocol to use. In this case we use HTTP.
+%The operation element defines each operation that the port exposes.
+%For each operation the corresponding SOAP action has to be defined. You must also specify how the input and output are encoded. In this case we use "literal".
+%  http://www.ibm.com/developerworks/webservices/library/ws-whichwsdl/
+%   1. RPC/encoded    (Remote Procedure Call)
+%   2. RPC/literal
+%   3. Document/encoded
+%   4. Document/literal
+% The terminology here is very unfortunate: RPC versus document.
+These terms imply that the RPC style should be used for RPC programming models and that the document style should be used for document or messaging programming models.
+That is not the case at all.
+The style has nothing to do with a programming model.
+It merely dictates how to translate a WSDL binding to a SOAP message.
+Nothing more.
+You can use either style with any programming model.
+%
+% RPC/literal SOAP message for myMethod
+%
+%<soap:envelope>
+%    <soap:body>
+%        <myMethod>
+%            <x>5</x>
+%            <y>5.0</y>
+%        </myMethod>
+%    </soap:body>
+%</soap:envelope>
+%
+%  RPC/encoded SOAP message for myMethod
+% 
+% <soap:envelope>
+%     <soap:body>
+%         <myMethod>
+%             <x xsi:type="xsd:int">5</x>
+%             <y xsi:type="xsd:float">5.0</y>
+%         </myMethod>
+%     </soap:body>
+% </soap:envelope>
+
+	\item \textbf{Service} --- Within the last WSDL element the endpoint of the service is specified.
+As to be seen in line~\ref{lst_code:echo_wsdl_service} the endpoint of the echo service is assigned to \textit{http://localhost:60000/echo}.
+
 \end{itemize}
@@
 	[
-	label=lst:echo_wsdl,
+	label=lst:echo_wsdl,float=p,
 	caption={[WSDL file describing the echo service. Filename: echo.wsdl]
-	\textbf{HWSDL file describing the echo service. Filename: echo.wsdl}}
+	\textbf{WSDL file describing the echo service. Filename: echo.wsdl}}
 	]
@@
 %    * Taxonomien/Ontologien zur semantischen Einordnung des Services
+\clearpage
 
@@
 
-The service defined by the WSDL file in the previous section shall now be implemented.
+The implementation of the service defined in the WSDL file of the previous section will blub.
 The source code of the C++ file is shown in Listing~\ref{lst:echo_service_cpp}.
@@
 
+\lstsetCPP
 \lstinputlisting

### 2009-02-24 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
Started outline for the TLS echo service

@@
 export ARC\_BARTENDER\_URL
+
+if bartender_url.startswith('https'):
+    key_file = os.environ.get('ARC_KEY_FILE', None)
+    cert_file = os.environ.get('ARC_CERT_FILE', None)
+    proxy_file = os.environ.get('ARC_PROXY_FILE', None)
+    ca_file = os.environ.get('ARC_CA_FILE', None)
+    ca_dir = os.environ.get('ARC_CA_DIR', None)
+
+export ARC_KEY_FILE=/local/glodek/clientKey.pem
+export ARC_CA_FILE=/local/glodek/clientCA.pem
+export ARC_CERT_FILE=/local/glodek/clientCert.pem
+
+does the daemon runs the service single threaded
+
+race conditions
+\begin{itemize}
+ \item Transaktionen
+ \item nur lesen
+ \item jeder prozess schreibt in seine eigene Datei
+ \item Gibt es einen     // andere Zugriffe abblocken
+    flock (\$fp, LOCK\_EX);  (atomare befehle)
+ 	wie bei php??
+ \item Peterson Algorithmus
+ \item Was passiert beim Ausfall (Time out - Transaktion zurück nehmen)
+\end{itemize}
+
+\section{Schlusswort}
+
+Other tutorials are:\\
+\\
+%Oxana reported the status of OGF tutorials:
+\begin{itemize}
+ \item first tutorial will be given by Oxana and will deal with how to add the 
+resources to Nordugrid, .... (will be related mostly to production ARC)
+ \item second tutorial will be given by Gabor and he will show the 
+interoperability of ARC (video demo will be shown). A-REX - Unicore 
+interoperability.
+ \item there will be also a tutorial related to ETICS system where the 
+possibility of ETICS system to submit a job to ARC and Unicore will be 
+presented. The ETICS team has not contacted so far.
+\end{itemize}

### 2009-02-24 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Started outline for the TLS echo service

--- 2009-02-19 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
@@
 
-TLS provides security and data integrity
+The web service presented in the previous chapter has the mayor drawback that the communication between the two endpoints is unsecure.
+Due to that reason the service  shall be extended by an additional layer called TLS (Transport Layer Security) to provide a certain degree of securtity on \task{Grid-level <-- explain that}.
+%
+In general the internet security is often subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{verfiy citation}
+\begin{itemize}
+	\item \textbf{Confidentiality} --- Protection of the data against passiv attacks (such as publicise message content)
 
-Diffrent kind of security is intended;
-\begin{enumerate}
- \item Integrity (data is not manipulated along the transmission)
- \item Confidentiality 
- \item Authenticity (Server and client sided)
- \item Authorization (Server and client sided)
-\end{enumerate}
+	\item \textbf{Authentication} --- Verification of the Authenticity  of the communications partner.
 
-The first two goals can be reached by encrypting the channel/connection between client and server. An approach 
-%sich bewähren / behaupten
-which stand the test of time is to introduce a TLS (Transport Layer Security) also known as SSL (Secure Sockets Layer).
-Authenticity, Confidentiality, Integrity
-The client wants to reassure himself to send his vulnerable data only to the server who can prove its correct identity.
-Furthermore a prove of authenticity is needed. % Sich versichern, dass das Gegenüber der ist, für welchen er sich ausgibt
-It .... does fancy stuff ...
+	\item \textbf{Integrity} --- Protection against interception and manipulation,replay,insertion, etc.  of a message.
+
+	\item \textbf{Non-repudiation} --- Preventing the sender to repudiate the message transmitted by him.
+
+	\item \textbf{Access control} --- Restrict the access to ressources.
+
+	\item \textbf{Availability} --- Ensure a system is always usable which is challenged by several attacks.
+
+\end{itemize}
+%1-21   Gebräuchliche Dienste
+%     Identifizierung           Vermerk
+%     Berechtigung              Zugriff
+%     Lizenz/Zertifizierung     Gültigkeitsprüfung
+%     Unterschrift              Zeitpunkt des Auftretens
+%     Bezeugung                 Abstimmung
+%     Übereinstimmung           Eigentum
+%     Zuverlässigkeit           Registrierung
+%     Quittungen                Genehmigung/Verbot
+%     Bestätigung des Ursprungs Privatsphäre
+%
+% Nach Sicherheit in verteilten Systemen, ITM Lübeck
+The TLS to be introduced provides: Confidentiality, authentication, integrity, non-repudiation and access control.
+It is introduced after the TCP layer. 
+%
+Before presenting the implementation of the new secure echo service the basic concept of a secure certifcate based connection shall be explained.
+Its not the goal to give an exact technical desciption but to give a first inside what is required to ensure the security of the layer.
+
+% It is recommand to satifiy as much categories as possible. 
+%authentication , confidentiality and integrity protection of TCP-based communication --- non-repudiation due to private key, access control to our service.
+
+\section{Transport Layer Security}
+
+The TLS is a protocol which provides a secure connection between two endpoints.
+It is located directly upon the TCP/IP layers and uses typically an asymmetric encryption to establish the connection (alternatively a symmetric pre-shared key may be used).
+After the communication has successfully established, both participants are switching to another encryption method which is based on a new negotiated symmetric key.
+%    1. Peer negotiation for algorithm support
+%    2. Key exchange and authentication
+%    3. Symmetric cipher encryption and message authentication   - from wikipedia
 %
 %
-Authorization:  The server on the other side takes an interest in limitate its ressources to a close circle of users. Above all no foreign clients shall be able to use the services.\\
+Symmetric and asymmetric encryption are the two main classes of cryptography algorithms.
+In case of a symmetric encryption both participants are holding the same secret key. 
+Symmetric encryption can be ranked as being very secure but its downside is the key distribution which is very difficult in practice. 
+Ideally the secret key is transmitted over a secure channel which is either a complete diffrent medium (i.e. a letter) or an channel encrypted with another secret key.
+Asymmetric encryption follows a diffrent functional principle.
+While the symmetric encryption utilise one shared secret, the asymmetric ecryption is based on two keys: a private key and public key.
+As the names imply the private key is kept secret and never shared while the public key is available for everyone.
+Messages encrypted with the public key can only be decrypted with the private key and the other way round, messages encrypted with the private key can only be decrypted with the public key.
+The figure~\ref{fig:asymmetric_encryption} illustrates the circumstances of the case.\\
+\begin{figure}[htb]
+	\centering%epstopdf async.eps
+	\subfloat[Texttext\label{fig:cgd}]
+		{\includegraphics[width=12cm]{tex_tls_echoservice/async.pdf}}\\
+	\subfloat[Texttext\label{fig:cgd}]
+		{\includegraphics[width=12cm]{tex_tls_echoservice/async2.pdf}}
+	\label{fig:asymmetric_encryption}
+	\mycaption{specific}{general}
+\end{figure}
 
-One can add the TLS in client and server by simply extending the arched configuration file and a small modification of the client.
+At startup of the communication process A (Alice) who desires a secure connection transmits her public key to B (Bob).
+Bob is now able to encrypt the messages such that only A is able to decrypt them.
+That way confidentiality and integrity of messages are easily realised.
+Non-repudiation, authentication and due to that access control are not given because the identity of Alice is not proven.
+In order to be able to check identities so called certificates have been introduced.
+Certificates render the possibility to check the identity of its owner.
+They are signed by a certificate authority (CA) which's identity can be resolved to another CA or is  pre-configured to be trusted.
+A certificate is composed of its owner's public key, his identity, the name of the signing CA and a signiture which is a hash-value of the certificate encrypted with the private key of the CA. 
+That way the CA confirms the identity of the certificate owner. 
+To obtain a certificate a request has to be submitted to a CA which verifies the identity of the requesting person and signs it, if it comes to the result that the authenticity is valid.
+Figure~\ref{fig:certificate_request} shows the procedure to gain a certificate.
+\begin{figure}[htb]
+	\centering%epstopdf certificates.eps
+ 	\includegraphics[width=13cm]{tex_tls_echoservice/certificates.pdf}
+	\mycaption{specific}{general}
+	\label{fig:certificate_request}
+\end{figure}
+Figure~\ref{fig:verification_of_certificates} shows a usecase where Alice wants to resolve the identity of Bob.
+Alice submitts her request to Bob along with a challange consisting of a random number $n$.
+Bob encrypts the number with his private key and returns it together with his certificate.
+With the usage of the public key, Alice is able to decrypt the challange which implies that, if the certificate is to be trusted, the identity of Bob is to be trusted.
+Unfortunately Alice is unable to verify the certificate because it is signed by an unknown CA called CA 1.
+She establishes the connection to CA 1 in order to resolve the identity of that certificate authority.
+Again Alice sends a challenge --- this time to the certificate authority CA 1.
+CA 1 encrypts the challange with its private key and submits the result along with its certificate.
+Again Alice validates the certificate sucessfully.
+The certificate was signed by a CA called CA 2 which she is able to locate as a pre-configured certificate. Due to the reason she trusts CA 2, she can trust CA 1 and finally Bob.
 
+\begin{figure}[htb]
+	\centering%epstopdf verification.eps 
+	\includegraphics[width=13cm]{tex_tls_echoservice/verification.pdf}
+	\mycaption{specific}{general}
+	\label{fig:verification_of_certificates}
+\end{figure}
+
+The advantages of certificates are
+o easy scalable to a larger set of users
+o VOMs are realisable
+o Infrastructure is given and easily to maintain
+o In general the pre-configured organistion take money to sign certifcates, but they can also be created locally.
+
+TLS also supports the more secure bilateral connection mode (typically used in enterprise applications), in which both ends of the "conversation" can be ensured with whom they are communicating
+
+o What do we need to get a service running
 clientCA   - Authority guarantee for identity
@@
 Its  clientCERT clientKEY and some kind of root CA knowing the server
+
+%The first two goals can be reached by encrypting the channel/connection between client and server. An approach 
+%sich bewähren / behaupten
+%which stand the test of time is to introduce a TLS (Transport Layer Security) also known as SSL (Secure Sockets Layer).
+%Authenticity, Confidentiality, Integrity
+%The client wants to reassure himself to send his vulnerable data only to the server who can prove its correct identity.
+%Furthermore a prove of authenticity is needed. % Sich versichern, dass das Gegenüber der ist, für welchen er sich ausgibt
+%It .... does fancy stuff ...
+%
+%
+%Authorization:  The server on the other side takes an interest in limitate its ressources to a close circle of users. Above all no %foreign clients shall be able to use the services.\\
+
+\clearpage
+\section{Service}
+
+One can add the TLS in client and server by simply extending the arched configuration file and a small modification of the client.
 
@@
 
+4 types:
+                <KeyPath>./clientKey.pem</KeyPath>
+                <CertificatePath>./clientCert.pem</CertificatePath>
+                <CACertificatePath>./serviceCA.pem</CACertificatePath>
+
 \lstsetCPP
@@
 	[
-	label=lst:arcecho_arched_xml,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
-	]
-{../src/clients/tlsechoclient/tlsechoclient.cpp}
-
-\lstsetCPP
-
-\lstinputlisting
-	[
-	label=lst:arcecho_arched_xml,
+	label=lst:tls_echo_service_cpp,
 	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
@@
 \lstsetARCHEDXML
-\begin{minipage}[t]{\textwidth}
 \lstinputlisting
 	[
-	label=lst:arcecho_arched_xml,
+	label=lst:tls_echo_arched_xml,float=htb,
 	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
@@
 {../src/services/tlsechoservice/arched_tls_echoservice.xml}
-\end{minipage}
 
 \lstsetKSH
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
+label=lst:tls_echo_arched_invoke,float=htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{lstlisting}
-\end{minipage}
+
+\clearpage
+\section{Client}
+
+\lstsetCPP
+\lstinputlisting
+	[
+	label=lst:tls_echo_client_cpp,
+	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
+	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	]
+{../src/clients/tlsechoclient/tlsechoclient.cpp}
 
 \lstsetKSH
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
+label=lst:tls_echo_client_invoke,float=htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \end{lstlisting}
-\end{minipage}
 
 \lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source,
+label=lst:tls_echo_request_XML,float=htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
+<soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
   <soap-env:Body>
-		<echoRequest:echo>
-			<echo:say>Der_Berg_ruft</echo:say>
-		</echoRequest:echo>
+    <tlsecho:tlsechoRequest>
+      <tlsecho:say operation="ordinary">text_to_be_transmitted</tlsecho:say>
+    </tlsecho:tlsechoRequest>
   </soap-env:Body>
@@
 \end{lstlisting}
-\end{minipage}
 
 \lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source,
+label=lst:tls_echo_response_XML,float=htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
+<soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
   <soap-env:Body>
-    <echo:echoResponse>
-      <echo:hear>[ Der_Berg_ruft ]</echo:hear>
-    </echo:echoResponse>
+    <tlsecho:tlsechoResponse>
+      <tlsecho:hear>[ text_to_be_transmitted ]</tlsecho:hear>
+    </tlsecho:tlsechoResponse>
   </soap-env:Body>
@@
 \end{lstlisting}
-\end{minipage}

### 2009-02-24 doc/ws-programming-tutorial/src/clients/tlsechoclient/tlsechoclient.cpp

@@ -18,17 +18,24 @@ using namespace Arc;
 
 int main(int argc, char** argv) {
 
-	if(argc != 3){
+
+	/** Get arguments passed by the command line*/
+	if(argc != 4){
 		printf("Usage:\n");
-		printf("  tlsechoclient clientHED.xml message\n");
+		printf("  tlsechoclient configuration type message\n");
 		printf("\n");
-		printf("The first argument (clientHED.xml) specifies the HED of the client.\n");
-		printf("The argument message contains the text to be transmitted.\n");
+		printf("The first argument specifies the HED configuration file to be loaded\n");
+		printf("The second argument the type of echo which shall be performed (ordinary or reverse).\n");
+		printf("The third agrument holds the message to be transmitted.\n");
 		printf("\n");
 		printf("Example:\n");
-		printf("  ./tlsechoclient clientHED.xml text_to_be_transmitted\n");
+		printf("  ./tlsechoclient clientHED.xml ordinary text_to_be_transmitted\n");
 		return -1;
 	}
+	
+	std::string type(argv[2]);
+	std::string message(argv[3]);
+	/** */
 
 	std::string xmlstring;
 	std::ifstream file(argv[1]);
@@ -51,34 +58,27 @@ int main(int argc, char** argv) {
 		return -1;
 	}
 
-	std::string message(argv[2]);
 
 
-	// Initiate logger to get the output of ARC classes
+	// Initiate the Logger and set it to the standard error stream
 	Arc::Logger logger(Arc::Logger::getRootLogger(), "arcecho");
 	Arc::LogStream logcerr(std::cerr);
 	Arc::Logger::getRootLogger().addDestination(logcerr);
 	Arc::Logger::rootLogger.setThreshold(Arc::WARNING);
 
-	// Set ARC installation directory for being able to 
-	// loaded modules dynamically
+	// Set the ARC installation directory
 	std::string arclib("/usr/lib/arc");
 	Arc::ArcLocation::Init(arclib);  
 
-
-	// Loading client.xml file which is the counterpart of the arc 
-	// configuration file
+	// Load configuration 
 	Arc::XMLNode clientXml(xmlstring);
-
-	// Wrap the xml for hed client configuration
 	Arc::Config clientConfig(clientXml);
+
 	if(!clientConfig) {
[…]

### 2009-02-24 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
Small modifications...

@@
 
-if bartender_url.startswith('https'):
-    key_file = os.environ.get('ARC_KEY_FILE', None)
-    cert_file = os.environ.get('ARC_CERT_FILE', None)
-    proxy_file = os.environ.get('ARC_PROXY_FILE', None)
-    ca_file = os.environ.get('ARC_CA_FILE', None)
-    ca_dir = os.environ.get('ARC_CA_DIR', None)
+    key\_file = os.environ.get('ARC\_KEY\_FILE', None)
+    cert\_file = os.environ.get('ARC\_CERT\_FILE', None)
+    proxy\_file = os.environ.get('ARC\_PROXY\_FILE', None)
+    ca\_file = os.environ.get('ARC\_CA\_FILE', None)
+    ca\_dir = os.environ.get('ARC\_CA\_DIR', None)
 
-export ARC_KEY_FILE=/local/glodek/clientKey.pem
-export ARC_CA_FILE=/local/glodek/clientCA.pem
-export ARC_CERT_FILE=/local/glodek/clientCert.pem
+export ARC\_KEY\_FILE=/local/glodek/clientKey.pem
+export ARC\_CA\_FILE=/local/glodek/clientCA.pem
+export ARC\_CERT\_FILE=/local/glodek/clientCert.pem
 

### 2009-02-24 doc/storage_system/arc-storage-documentation.tex
updating the ARC storage documentation

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The user can specify for each file how many replicas are needed.
+The ARC storage system periodically checks the number of replicas, and automatically creates new replicas if there are fewer then needed, and removes replicas if there are more.
+
+The different replicas of a file could be in different states, e.g.
+the replica could be valid and alive or just in the process of creation or it could be corrupt or a whole storage node could be offline.
+This state is always stored as metadata next to the Location of the given replica.
+For each file there is a checksum calculated, and this checksum is used to detect if a replica gets corrupted.
+If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state.
+And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat triggers the modification of metadata as well.
 
 % section storage_nodes_and_replicas (end)
+
+\section{The A-Hash} % (fold)
+\label{sec:the_a_hash}
+
+The A-Hash is a replicated metadata store, which is capable of consistently storing `objects' where an object contains key-value pairs organized in sections.
+All metadata about files and collections are stored in the A-Hash, and some other information (e.g.
+about A-Hash replication, about Shepherd services, etc.) is stored in it as well.
+The A-Hash itself does not interpret the data, it basically just stores tuples of strings.
+
+% section the_a_hash (end)
+
+\section{The Librarians} % (fold)
+\label{sec:the_librarians}
+
+The Librarian is capable of managing the hierarchy and metadata of files and collections, and health information of the Shepherd services.
+It can traverse Logical Names and return the corresponding metadata.
+It can receive heartbeat messages from Shepherd services, and it automatically modify the states of files if needed.
+The Librarian itself is a stateless service, it uses the A-Hash to actually store and retrieve the metadata, that’s why there could be any number of independent Librarian services (all using the same A-Hashes) which provides high-availability and load-balancing.
+
+% section the_librarians (end)
 
@@
 The Bartender service provides a high-level interface for the storage system to the clients.
+Every interaction between a client and the ARC storage system begins with a request to a Bartender.
 You can create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names.
-The Bartender communicates with the Librarian and Shepherd services to accomplish the client’s requests.
+The Bartender authorizes users and force access policies of files and collections.
+It communicates with the Librarian and Shepherd services to accomplish the client’s requests.
 The actual file data does not go through the Bartender; file transfers are directly performed between the storage nodes and the clients.
 There could be any number of independent Bartender services in the system which provides high-availability and load-balancing.
+The Bartender also provides a way to access files on third-party storage solutions through its interface by mounting the namespace of the third-party storage into the namespace of the ARC storage (this is accomplished by so called `gateway' modules).
 
 % section the_bartenders (end)
-
-\section{The Librarians} % (fold)
-\label{sec:the_librarians}
-The Librarian is capable of managing the hierarchy and metadata of files and collections, and health information of the Shepherd services.
-Each file and collection in the Librarian has a globally unique ID (GUID).
-A collection contains files and other collections, and each of these entries has a name unique within the collection very much like entries in a usual directory on a local filesystem.
-Besides files and collections the Librarian stores a third type of entries called Mount Points which are references to external services which will be in the future used to mount the namespace of third-party storages to our global namespace and make the files on a third-party storage available through the interface of the ARC storage system.
-
-The Librarian also manages information about registered Shepherd services which are associated with a storage node, and receives heartbeat messages from them and change replica states automatically if needed.
-
-The Librarian uses the A-Hash as database, that’s why there could be any number of independent Librarian services (all using the same A-Hash) which provides high-availability and load-balancing.
-
-% section the_librarians (end)
-
-\section{The A-Hash} % (fold)
-\label{sec:the_a_hash}
-The A-Hash is a currently centralized, but later distributed service capable of consistently storing objects containing property-value pairs organized in sections.
-All metadata about files and collections are stored in the A-Hash, and some information about Shepherd services is stored in it as well.
-The A-Hash itself does not interpret the data, it just stores tuples of strings.
-
-% section the_a_hash (end)
 
@@
 \label{sec:the_shepherds}
-When a new file is put into the system the number of needed replicas is given for the file.
-The file replicas are stored on different storage nodes, for each storage node there is Shepherd service which manages the storage node, reports its health state to the Librarian and provides the interface for initiating file transfer.
-
-The file-naming used by a Shepherd has nothing to do with the the hierarchy of collections, or Logical Names.
-When a new replica upload is initiated, the Shepherd generates an ID which refers to it within that Shepherd.
-Each Shepherd has a unique ID itself, so with these IDs the replica can be unambiguously referenced, this is called a \emph{Location}.
-The namespace of these Locations has nothing to do with the namespace of GUIDs or the namespace of Logical Names.
-It consists of two IDs: the ID of the Shepherd and the ID of the file within the Shepherd: \verb!(serviceID, referenceID)!
+
+The files in the ARC storage system are usually replicated on different storage nodes.
+For each storage node there is a Shepherd service which manages the storage element service on the node, reports its health state to a Librarian and provides the interface for initiating file transfers.
+For each kind of storage element service (e.g.
+a HTTP server, an FTP server, a storage solution with a GridFTP interface, etc.) it is needed to have a Shepherd backend which is capable of managing the given storage element service.
+The Shepherd service periodically checks the health of the replicas based on their checksums, and if a replica is deleted or corrupted, the Shepherd tries to recover it by downloading a valid copy from an other storage node.
+The Shepherds also check if a file has fewer replicas in the system than needed, and they initiate replication if needed.
 
 % section the_shepherds (end)
-
-\section{Heartbeats and replication} % (fold)
-\label{sec:heartbeats_and_replication}
-Each Shepherd should periodically send heartbeats to a Librarian service with information about replicas whose state changed since the last heartbeat, the Librarian stores these file lists (which contains the GUIDs of the files as well), and if it doesn’t receive a heartbeat for a Shepherd in a given time, it invalidates all the replicas the Shepherd stores.
-This invalidating means that the state of that location will be \verb!offline!.
-
-If a Shepherd finds out that a file is missing or has a bad checksum, it reports that the file is \verb#invalid# to the Librarian immediately, and the Librarian alters the state of the given replica of the file.
-The Shepherd tries to recover its replica by downloading it from another Shepherd.
-In order to do this the Shepherd contacts a Bartender and asks for the file.
-The Bartender chooses a valid replica, initiate file transfer by a Shepherd having a valid replica, and returns the TURL to the Shepherd with the invalid replica.
-The Shepherd with the invalid replica downloads the file from the other Shepherd, and if everything is OK, signals to the Librarian that the replica is \verb!alive!
-again.
-
-The Shepherds periodically ask the Librarian whether the files they store have enough replicas.
-If a Shepherd finds that one of the files has not enough replica it turns to a Bartender offering replication.
-The Bartender chooses a Shepherd, initiates a put request then returns the TURL to the offering Shepherd which could upload the replica.
-The Shepherd who now has the new replica notifies the Librarian that the file is \verb!alive!.
-The Librarian sets the state of this new replica.
-
-% section heartbeats_and_replication (end)
 
@@
 
-The current version of the ARC storage system has no implemented solution for security issues.
-Designing and implementing a powerful, flexible, fine-grained solution for authentication and authorization is very important.
-Currently we have plans how to do that, and the design of the storage services makes it easy to integrate the chosen solution into the system.
-
-For additional details, see Section~\ref{sec:plans_for_security}.
+The ARC storage system consists of several services.
+Most of the services (A-Hash, Librarian, Shepherd) are `internal' services in a way that the end-user of the storage system never communicates with them directly.
+But these internal services are communicating with eachother, so they have to know who to trust.
+We call this aspect of the security architecture of the ARC storage `inter-service authorization'.
+
+The end users always connect to one of the Bartender services, which will decide if the user has permissions to do something or not.
+This is the `high-level authorization' part of the security architecture.
+
+For transfering the actual file data, the users have to connect to storage element services which are sitting on storage nodes.
+These services also have their own authentication and authorization methods.
+Managing these aspects is the `transfer-level authorization' part of the security architucture of the ARC storage.
+
+\subsection{Inter-service authorization} % (fold)
+\label{sub:inter_service_authorization}
+
+In a deployment of the ARC storage system, we could have several A-Hash, Librarian, Shepherd and Bartender services.
+The Bartenders send requests to the Librarians and the Shepherds, the Shepherds communciate with the Librarians, the Librarians talk with the A-Hashes.
+If any of these services get compromised or a new rouge service gets inserted in the system, we loose security completely.
+That's why it is vital for each service to authorize the services before sending or accepting requests.
+The services commincates via HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and they can examine the other service's certificates.
+Because of these X.509 certificates each service has Distinguish Name (DN).
+We can use these DNs to exactly specify which services we trust.
+We can configure a list of trusted DNs into each service, or we can store this list on a remote location.
+The services will only accept connections if the DN of the other end is listed in this list of trusted DNs.
+However the Bartender services will accept any incoming connection, which are from the users, because the users are authenticated differently.
+
+% subsection inter_service_authorization (end)
+
+\subsection{High-level authorization} % (fold)
+\label{sub:high_level_authorization}
+
+The Librarian component of the ARC storage system stores all the metadata about files and collections.
+For each file and collection there are access policies in the form of access control rules, and these are stored among these metadata.
+The users are identified by their DNs, and an access control rule specify the rights of the given user.
+One rule can be represented like this:
+
+\begin{verbatim}
+    DN +action +action -action
+\end{verbatim}
+
+This contains a list of actions, each prefixed with a \verb!+! or \verb!-! character which indicates that the given action is allowed or not allowed for the given DN.
+
+Besides specifying only one user with a DN, there are other types of access control rules: we can have a rule for a whole VO (Virtual Organization) or for ALL users, like this:
+
+\begin{verbatim}
+    ALL +action
+    VOMS:knowarc.eu +action -action 
+\end{verbatim}
+
+These are the actions which can be used for access control:
+\begin{itemize}
+    \item \emph{read}: user can get the list of entries in the collection; user can download the file
+    \item \emph{addEntry}: user can add a new entry to the collection;
+    \item \emph{removeEntry}: user can remove any entry from the collection 
+    \item \emph{delete}: user can delete the collection if it is empty; user can delete a file
+    \item \emph{modifyPolicy}: user can modify the policy of the file/collection
+    \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
+    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are key-value pairs)
+\end{itemize}
+
+Additionaly, each file and collection has an `owner' which is a user who always can modify the access control rules.
+
+% subsection high_level_authorization (end)
+
+\subsection{Transfer-level authorization} % (fold)
+\label{sub:transfer_level_authorization}
+
+Currently the transfer-level authorization is kept very simple.
+When the Bartender decides that a user has permission to download a file, then the Bartender chooses a replica, and initiates the transfer.
+The result of this initiation is a URL which is colled the transfer URL (TURL).
+This TURL is unique for each request, even for request to the same replica, and this TURL is only valid for one download.
+Currently we configure the storage element services to not do any authorization, and we use these one-time URLs to ensure that only the authorized users can access the contents of the storage elements.
+
+% subsection transfer_level_authorization (end)
 

### 2009-02-25 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Started outline for the TLS echo service
echo.wsdl

@@
 
+\task{Provide a list of modules to be loaded! Along with the implemented functionality which can be addressed with XML elements!!}
+    mcctcp\\
+\begin{verbatim}
+<tcp:Listen>
+	<tcp:Port>60000</tcp:Port>
+	<tcp:Version>4</tcp:Version>
+</tcp:Listen>
+\end{verbatim}
+    mcctls\\
+\begin{verbatim}
+ 			<KeyPath>./clientKey.pem</KeyPath>
+			<CertificatePath>./clientCert.pem</CertificatePath>
+			<CACertificatePath>./clientCA.pem</CACertificatePath><!-- CACertificatesDir Should be /etc/... blub mention that-->
+
+\end{verbatim}
+
+    mcchttp\\
+\begin{verbatim}
+                      <next id="soap">POST</next>
+                      <next id="plexer">GET</next>
+                      <next id="plexer">PUT</next>
+\end{verbatim}
+    mccsoap\\
+\begin{verbatim}
+	???<ClientSSLConfig FromFile='filename'/>  
+\end{verbatim}
+
 \subsection{The arc echo service}

### 2009-02-25 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Small changes concerning the echo client

@@
 
-The next service to be presented is the echo service which returns the received message back to the client.
-Furthermore the client has to choose between the operations ``ordinary'' or ``reverse''.
-Depending on that operation the message will be returned directly or its letters will be reordered reverse. % without manipulation
-The goal of this chapter is to offer a deeper knowledge about message processing. Furthermore the \textit{WSDL} (Web Services Description Language) shall be introduced along with this example.
+The next service to be presented is the echo service which's task is to return the received message back to the client.
+Furthermore the service provides two additional operations called ``ordinary'' and ``reverse''.
+Depending on the operation chosen by the client, the message will returned unmodified or its letters will be reordered reverse. % without manipulation
+The goal of this chapter is to offer a deeper knowledge about message processing. Furthermore the \textit{WSDL} (Web Services Description Language) shall be introduced.
 
@@
 
-Web Services are often reachable for a large set of users in order to easily access complex information.
-In many cases the Web Service are capable to replace HTML pages i.e.
-to access to databases like Pubmed or the NLM catalog.
-% http://www.ncbi.nlm.nih.gov/entrez/query/static/esoap_help.html
-In contrary to HTML which describes a human readable document the interfaces of Web Services are well defined and easier to interpret for machines.
-In general the client is not familiar with the functionality of the service.
-Due to that reason, the access to a web service can be defined in a \textit{WSDL} (Web Services Description Language) file.
-WSDL is a meta language which describes the structure of the messages, the intended use of the message, the supported protocols and the endpoint of the service (URL).
-The WSDL file which specifies the echo service of this chapter is to be seen in Listing~\ref{lst:echo_wsdl}. 
-It consisits of five main WSDL elements:
+Web services can be specified by a meta language called \textit{WSDL} (Web Services Description Language). 
+Beside the description of the service, WSDL renders the possibility that services can be discovered by service brokers.
+The discovery of services may be useful for an automatic adaption of a running client i.e. if the endpoint of the service changes or to share the load of demands.
+Furthermore the service specification enables the programmer to implement a suitable client which interfaces the service in a proper way.\\
+WSDL describes the structure of the messages, the intended use of the messages, the supported protocols and the endpoint of the service (URL).
+The WSDL file which specifies the echo service presented in this chapter is to be seen in listing~\ref{lst:echo_wsdl}. 
+It is composed of five main elements:
 \newcommand{\parboxWidth}{13cm}
 \begin{itemize}
-	\item \textbf{Types} --- The element \textit{types} contains the description of the message structures which is written in the language XSD (XML Schema Definition).
-The applicable data types of the echo service are defined in  line~\ref{lst_code:echo_wsdl_types}.
+	\item \textbf{Types} --- The element \textit{types} contains the description of the message structures and is written in the language XSD (XML Schema Definition).
+The data types used for the messages of the echo service are defined starting with the  line~\ref{lst_code:echo_wsdl_types}.
 
 	\item \textbf{Message} --- Within the element \textit{message} a subset of the previously defined types are assigned to be a message.
-In case of the echo service two messages are designated: echoRequest and echo Response, see lines line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
+In case of the echo service two messages are designated: echoRequest and echoResponse, compare lines line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
 
@@
 Four interface types can be distinguished: One-way (input), request-response (input, output), solicit-response (input, output, fault) and notification (output).
-	The echo service realises a simple request-response interface which is declared subsequent line~\ref{lst_code:echo_wsdl_portType}. % in the lines followed by the line X
+	The echo service realises a solicit-response interface which is declared subsequent to line~\ref{lst_code:echo_wsdl_portType}. % in the lines followed by the line X
 
@@
 The transport attribute defines the protocol which is regularly HTTP.
-For each operation a corresponding SOAP action has to be defined.
-Additionally the input and output has to be specified to be either \textit{literal} or \textit{encoded}~\cite{BUTEK_2009}.
-The service implemented in this chapter will use the binding \textit{document}/\textit{literal} and the protocol HTTP which is to be seen in line~\ref{lst_code:echo_wsdl_binding}.
+For each operation a corresponding SOAP action has to be defined which assignes a messages to be either \textit{literal} or \textit{encoded}~\cite{BUTEK_2009}.
+The service implemented in this chapter will use the binding \textit{document}/\textit{literal} and the protocol HTTP which is to be seen in the lines followed by  line~\ref{lst_code:echo_wsdl_binding}.
 
@@
 
-	\item \textbf{Service} --- Within the last WSDL element the endpoint of the service is specified.
+	\item \textbf{Service} --- Within the last WSDL element \textit{service} the endpoint of the service gets specified.
 As to be seen in line~\ref{lst_code:echo_wsdl_service} the endpoint of the echo service is assigned to \textit{http://localhost:60000/echo}.
@@
 {../src/services/echoservice/echo.wsdl}
+\task{Validate WSDL}
+
+The SOAP Fault Element
+
+The optional SOAP Fault element is used to indicate error messages.
+
+If a Fault element is present, it must appear as a child element of the Body element. A Fault element can only appear once in a SOAP message.
+
+The SOAP Fault element has the following sub elements:
+Sub Element 	Description
+<faultcode> 	A code for identifying the fault
+<faultstring> 	A human readable explanation of the fault
+<faultactor> 	Information about who caused the fault to happen
+<detail>
+
+Holds application specific error information related to the Body element
+SOAP Fault Codes
+
+The faultcode values defined below must be used in the faultcode element when describing faults:
+Error 	Description
+VersionMismatch 	Found an invalid namespace for the SOAP Envelope element
+MustUnderstand 	An immediate child element of the Header element, with the mustUnderstand attribute set to "1", was not understood
+Client 	The message was incorrectly formed or contained incorrect information
+Server 	There was a problem with the server so the message could not proceed
 
@@
 
-The implementation of the service defined in the WSDL file of the previous section will blub.
-The source code of the C++ file is shown in Listing~\ref{lst:echo_service_cpp}.
-Again the header file will be set aside for it contains to much redundant information.
+In the current section, the implementation of the service defined in the WSDL file of the previous section will be presented.
+The source code of the C++ file is shown in listing~\ref{lst:echo_service_cpp}.
+Again the header file will be set aside for it contains only redundant information.
+It can be reread in the source directory accompanying the tutorial.
+Within the source code of echo service basically three changes has been done.
+The first change concerns the constructor and is to be found in line~\ref{lst_code:echo_cpp_prefix}.
+Two strings will be extracted of the HED configuration file which are later used to envelope the message before it is returned.
+Furthermore a new method \textit{makeFault} in line~\ref{lst_code:echo_cpp_makeFault} has been introduced which is able to create fault messages and is called by the method \textit{process}.
+Within that method two new code fragments are now extracting and analyzing the incoming message, to be seen in line~\ref{lst_code:echo_cpp_extracting} and \ref{lst_code:echo_cpp_analyzing}.
+In case something unexpected happens both will create a SOAP fault message.\\
+% Vielleicht noch etwas über das extrahieren der Nachricht schreiben
 
@@
 
+Again the HED is configured by a XML file and have to be lunched as usual.
+The HED configuration file is shown in Listing~\ref{lst:echo_hed_configuration_xml}.
+The only change concerns the lines~\ref{lst_code:echo_hed_configuration_xml_pre} and \ref{lst_code:echo_hed_configuration_xml_pre} which are specifing the \textit{prefix} and \textit{suffix}.
+The parameters will be extracted in the constructor of the service and provide a good example how settings can be passed from the HED configuration file towards the service.
+
 \lstsetARCHEDXML
-
 \lstinputlisting
 	[
-	label=lst:arcecho_arched_xml, float=htb,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	label=lst:echo_hed_configuration_xml, float=htb,
+	caption={[HED configuration file for the echo service. Filename: arched\_echoservice.xml]
+	\textbf{HED configuration file for the echo service. Filename: arched\_echoservice.xml}}
 	]
@@
 
-\lstsetKSH
-\begin{lstlisting}[
-label=lst:invokation_arched_timeservice,float=htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ arched -c arched_echoservice.xml  && echo jo ||echo n
-\end{lstlisting}
+%\lstsetKSH
+%\begin{lstlisting}[
+%label=lst:invokation_arched_timeservice,float=htb,
+%caption={[Transformation in eine uniforme konzentrische Verteilung.]
+%         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+%$ arched -c arched_echoservice.xml  && echo jo ||echo n
+%\end{lstlisting}
 
 \section{Client}
+
+Likewise the client has become also more complex. The parameters are now passed by the command line, to be seen in the lines followed by line~\ref{lst_code:echo_client_cpp_arguments}.
+The creation of the SOAP client is unchanged. The request is created in lines subsequent line~\ref{lst_code:echo_client_cpp_request}.
+It is processed in the code after line~\ref{lst_code:echo_client_cpp_response} in which also the response is received.
+The checks of the response have expanded by an addition condition which tests for a fault message, see line~\ref{lst_code:echo_client_cpp_fault}.
 
@@
 
+The usage of the client is presented in Listing~\ref{lst:invokation_echoservice}.
+Three cases of echo service responses are displayed: ordinary operation, reverse operation and SOAP fault due to an unknown operation.
+
 \lstsetKSH
 \begin{lstlisting}[
-label=lst:invokation_arched_timeservice, float=htb,
+label=lst:invokation_echoservice, float=!htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 [ dettimsnart_eb_ot_txet ]
+$ ./echoclient http://localhost:60000/echo re2verse text_to_be_transmitted
+A SOAP fault occured:
+  Fault code:   4
+  Fault string: "Unknown operation. Please use "ordinary" or "reverse""
 \end{lstlisting}
+
+A few examples of the exchanged messages are shown in the Listings~\ref{lst:echoservice_request}, \ref{lst:echoservice_response} and \ref{lst:echoservice_fault}.
+Listing~\ref{lst:echoservice_request} illustrates a request message using the operation \textit{reverse}.
+Listing~\ref{lst:echoservice_response} illustrates the corresponding response message.
+In case the operation requested by the client is unknown, the server replies with the fault message to be seen in Listing~\ref{lst:echoservice_fault}.
 
@@
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source, float=htb,
+label=lst:echoservice_request, float=!htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 \begin{lstlisting}[
-label=lst:timeservice_cpp_source, float=htb,
+label=lst:echoservice_response, float=!htb,
 caption={[Transformation in eine uniforme konzentrische Verteilung.]
@@
 
-Folgender XML Aufruf und Antwort soll \textit{automatisch} generiert werden:
-
-%[caption={[]WegDamit},language=XML,basicstyle=\scriptsize,breaklines=true,label=lst:request]
-
-Zertifikate
-Zustände (Nutzer wiedererkennen, arbeit aufnehmen)
+\lstsetJUSTXML
+\begin{lstlisting}[
+label=lst:echoservice_fault, float=!htb,
+caption={[Transformation in eine uniforme konzentrische Verteilung.]
+         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
+  <soap-env:Body>
+    <soap-env:Fault>
+      <soap-env:faultcode>soap-env:Client</soap-env:faultcode>
+      <soap-env:faultstring>Unknown operation. Please use "ordinary" or "reverse"</soap-env:faultstring>
+    </soap-env:Fault>
+  </soap-env:Body>
+</soap-env:Envelope>
+\end{lstlisting}

### 2009-02-25 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
Small changes concerning the echo client

@@
 
-\section{Schlusswort}
+\chapter{Appendix}
+
+\section{More tutorials}
 
@@
 \end{itemize}
+
+\section{Important XSD files}
+
+Link to current XSD files.
+
+\task{Provide a list of modules to be loaded! Along with the implemented functionality which can be addressed with XML elements!!}
+    mcctcp\\
+\begin{verbatim}
+<tcp:Listen>
+	<tcp:Port>60000</tcp:Port>
+	<tcp:Version>4</tcp:Version>
+</tcp:Listen>
+\end{verbatim}
+    mcctls\\
+\begin{verbatim}
+ 			<KeyPath>./clientKey.pem</KeyPath>
+			<CertificatePath>./clientCert.pem</CertificatePath>
+			<CACertificatePath>./clientCA.pem</CACertificatePath><!-- CACertificatesDir Should be /etc/... blub mention that-->
+
+\end{verbatim}
+
+    mcchttp\\
+\begin{verbatim}
+                      <next id="soap">POST</next>
+                      <next id="plexer">GET</next>
+                      <next id="plexer">PUT</next>
+\end{verbatim}
+    mccsoap\\
+\begin{verbatim}
+	???<ClientSSLConfig FromFile='filename'/>  
+\end{verbatim}

### 2009-02-25 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Small changes concerning the echo client

@@
 	label=lst:tls_echo_client_cpp,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	caption={[HEC configuration file]
+	\textbf{HEC configuration file\textcolor{white}{hmf}}}
 	]
 {../src/clients/tlsechoclient/tlsechoclient.cpp}
+
+Folgender XML Aufruf und Antwort soll \textit{automatisch} generiert werden:
+
+%[caption={[]WegDamit},language=XML,basicstyle=\scriptsize,breaklines=true,label=lst:request]
+
+Zertifikate
+Zustände (Nutzer wiedererkennen, arbeit aufnehmen)
 

### 2009-02-25 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Small changes concerning the echo client
Reviewed introduction.

--- 2009-02-23 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
@@
 
-In general, the implementation of a client is easier \task{that} the implementation of the server. 
-ARC is designed as a middleware which encapsulates typical challenges of server-client infrastructures (security, exception handling, extensibility) and abstracts from underlying architecture (heterogeneity of computers, computer location, protocols).
-The core of Arc is the prior mentioned HED (Hosting Environment Daemon).
+In general, the implementation of a client is easier than the implementation of the server. 
+ARC is designed as a middleware which encapsulates typical challenges of server-client infrastructures (security, exception handling, extensibility) and abstracts from the underlying computer architecture (heterogeneity of computers, computer location, protocols).
+The core of ARC is the prior mentioned HED (Hosting Environment Daemon).
 To the administrator HED is mostly visible as the single binary that is started and configured by a file that describes services that shall be prepared by HED.
-Conceptionally, the HED determines how services shall be organised both as a principle, and for any given ARC-run server, HED determines the very explicit presence or absence of a service at a particular address.
-Its many configurable layers are described later in this tutorial.
+Conceptionally, the HED determines how services shall be organised both as a principle, and for any given ARC-run server.
+HED determines the very explicit presence or absence of a service at a particular address. Its many configurable layers are described later in this tutorial. 
 %
@@
 
-The HED which was already mentioned in the section above as an essential part of the ARC middleware provides hosting of various services at application level, as well as a number of modules.
-Figure~\ref{fig:HED_internal} illustrates a typical setup of a server sided HED.
-The HED consists of a few compononents called Message Chain Components (MCC) which are ordered within a layered structure and several \textcolor{urgent}{modules} which shall aid the programmer to simplify the development of the Web Services (Config, Loader, Logging, XMLNode).
-%
+The HED which was already mentioned in the section above. As an essential part of the ARC middleware it provides the hosting of various services at the application level.
+Its modularity allows its dynamic extension.
+Figure~\ref{fig:HED_internal} illustrates a typical setup of a server-sided HED.
+The HED consists of a few compononents called Message Chain Components (MCC).
+These are ordered within a layered structure and several \textcolor{urgent}{modules}, which shall aid the programmer to simplify the development of the web services: Config, Loader, Logging, XMLNode.
+
 \begin{figure}[htb]
@@
 \end{figure}
-The MCCs are passing the incoming and outgoing messages to the upper or the respective lower layer.
-% AUFZÄHLUNG???
+
+The MCCs are passing the incoming and outgoing messages to the upper or the respective lower layer. \task{which means? SM}
 The lowest layer has to be a transport protocol like TCP which enables the transmission between the two applications, the client and the server.
-(The Internet Protocol (IP) beneath the TCP layer enables a transmission between two computers.
-IP is needed but not part of the HED.) TCP is well suited for the application because it provides a reliable and ordered delivery of a byte stream which is important to ensure security and a requirement for the usage of HTTP.
-The TCP layer is followed by the Hypertext Transfer Protocol (HTTP). 
-In case a security layer is desired, the TLS (Transport Layer Security formerly known as SSL, Secure Sockets Layer) can be inserted between the two layers TCP and HTTP.
-The HTTP provides the client server architecture.
+Technically, there are lower leves, e.g. the Internet Protocol (IP) enables the transmission between two computers and is a technical requirement for the TCP layer.
+But IP is not directly modelled by the HED.
+TCP is well suited for the application because it provides a reliable and ordered delivery of a byte stream which is important to ensure security.
+Above the TCP layer is the Hypertext Transfer Protocol (HTTP). 
+In case a security layer is desired for authentication, the TLS (Transport Layer Security formerly known as SSL, Secure Sockets Layer) is inserted between the TCP and HTTP layers.
+The HTTP provides the client-server architecture.
 It is stateless but offers several extensions for requests, header information and status codes.
-Furthermore HTTP enables the usage of Uniform Resource Locators (URL) which is used by Arc to multiplex between different services.
+Furthermore HTTP enables the usage of Uniform Resource Locators (URL) which is used by ARC to multiplex between different services.
 The multiplexing layer can also be defined within the HED and is called Plexer.
-Depending on the path of the URL the Plexer passes messages to a defined SOAP service which again passes the message to the Web Service.\\
+Depending on the path of the URL the Plexer passes messages to a defined SOAP service which again passes the message to the Web Service.
 
@@
 A first impression how the HED can be configured shall be given in the following example which corresponds to the HED shown in figure~\ref{fig:HED_internal}.
-It uses the internal ARC echo service~\cite{QIANG_2005}.
+It uses the ARC echo service~\cite{QIANG_2005} that is shipped for testing purposes with the ARC source code.
 
-\subsection{The arc echo service}
+\subsection{The ARC echo service}
 
 \task{Where is the xsd file for HED configuration files??}
-The first example shall give a basic understanding in how to configure the HED.
-The task is to setup the ARC intern Echo Web Service which has to be done by invoking the \textit{arched} daemon with a fitting HED configuration file.
-Such a fitting HED configuration file is shown in Listing~\ref{lst:arched_arcecho_xml}.
+The first example shall give a basic understanding in how to configure the HED for setting up a novel service.
+Here, the task is to setup the regular ARC Echo Web Service.
+On the command line, all that needs to be done is to invoking the \textit{arched}\footnote{read: arc-hed} daemon with a fitting HED configuration file.
+An example is shown in Listing~\ref{lst:arched_arcecho_xml}.
 It is written in XML and its structure is based on the \textcolor{urgent}{??} schema.
-\lstsetARCHEDXML\\
+
+\lstsetARCHEDXML
 
@@
 {../src/services/arched_arcecho.xml}
-The first line of the XML file contains the XML declartion.
+The first line of the XML file contains the XML declaration.
 Several attributes can be defined here but at least the XML version should be specified.
-\task{What is this ArcConfig for??} % configuration namespaces
+The configuration as a whole is encapsulated by the \textit{<ArcConfig>} tag.
+It also performs the setting of namespace abbreviations, which are expected to be mostly invariant across all server installations.
 The \textit{Server} element, to be found at line~\ref{lst_code:arched_arcecho_Server}, provides basic settings for the daemon such as location of the Pid-file or the Log-file.
@@
 The path should be at least assigned to the ARC installation directory.
-% for the standard MCC are defined here
+Otherwise, the standard set of MCC modules might not be found.
 The next set of elements are holding the name of the plugins to be loaded (line~\ref{lst_code:arched_arcecho_Plugins}).
 The names have to correspond to the name of the dynamic libraries within the path defined in the \textit{ModuleManager}.
-Finally the chain of the components are declared within the element \textit{Chain}.
-Each component inside the chain is represented by an element called \textit{Component} which has at least a name attribute.
-Its important that a plugin has been loaded which provides the functionality of it.
-Depending on the component certain additional elements may be defined inside of it.
-\task{Internet Protocol is required.}
-The first component of the chain is the \textit{tcp.service}. As to be seen the port 60000 is assigned to the server to listen to.
+Finally, the chain of the components is declared within the element \textit{Chain}.
+Every component inside the chain is represented by an element called \textit{Component} which has at least a \textit{name} attribute.
+Its important that a plugin (*.so file found in the paths previously specified) has been loaded which provides the implementation of the component's functionality.
+The further configuration of the individual web service depends on the individual component's functionality.
+The \textit{Service} tag may be used to specify additional elements that the web service may then inspect.
+
+%\task{Internet Protocol is required.}
+
+We need to distinguish between the plugins and the components.
+The plugins refer to shared libraries that can be integrated as modules with the HED to extend its functionality.
+The Components refer to the implementation of web services within any of those libraries.
+A \textit{Chain} will determine the path that an event takes to be noticed.
+The first component of the chain is the \textit{tcp.service}.
+As to be seen, the port 60000 is assigned to the server to listen to.
+The port can be an arbitrary number.
+However, the number 60000 is commonly used for a HED providing the A-REX web service (the one organising the computation on a site) and the number 50000 is found commonly used for a HED running web services for storage.
+This way, one can organise a single server to offer multiple HED instances.
 The stream which is received by this component is passed to the component defined by the \textit{next} tag.
-In the present case it will be passed to the component with the \textit{id} atrribute named \textit{http}.
-According to this manner the message will be passed to higher layers until it reaches the \textit{Plexer}  Depending on the path of the URL which was demanded by the HTTP the Plexer multiplexes the request to the right Web Service.
+In the present case, it will be passed to the component with the \textit{id} attribute named \textit{http}.
+In this way, the message is be passed to higher layers until it reaches the \textit{Plexer}.
+Depending on the path of the URL which was demanded by the HTTP, the Plexer multiplexes the request to the right web service, the right function to execute according to the component IDs and plugin names.
+The plexer will thus allow to have multiple web service listen to the same port.
 The path, declared within the \textit{next} element, corresponds to a regular expression and leads to the service with the id named \textit{echo}. 
@@
 The Echo Web Service itself is provided by the ARC package.
-Additional fix parameters are passed by the elements \textit{prefix} and \textit{suffix}.
+Additional invariant parameters are passed by the elements \textit{prefix} and \textit{suffix}.
 They set the characters which will enclose the echoed string.
-%returned
-\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}\\
+\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}
 
-To create the daemon corresponding to the defined HED configuration file ARC provides the program called \textit{arched}.
+To create the daemon corresponding to the defined HED configuration file, ARC provides the program called \textit{arched}.
 The manner of invokation is to been seen in Listing~\ref{lst:arcecho_arched_invokation}.
@@
 %
-The name of the program is followed by the parameter \textit{c} which assignes, that the next string contains the full path of configuration file.
+The name of the program is followed by the parameter \textit{-c}.
+The string following that argument contains the full path of configuration file.
 Once the server is running, the client can be used as described in Listing~\ref{lst:arcecho_client_invokation}.
@@
 label=lst:arcecho_client_invokation,float,
-caption={[Usage of the Arc echo client.]
+caption={[Usage of the ARC echo client.]
          \textbf{Usage of the Arc echo client.\textcolor{white}{hmf}}}]
@@
 
-In this example the server and the client are running on the same computer so that the hostname \textit{localhost} can be utilized. 
-For debugging it is advisable to check the Log-file which was assigned inside the HED configuration file.\\
+In this example, the server and the client are running on the same computer such that the hostname \textit{localhost} can be utilized. 
+For debugging it is advisable to check the logfile which was assigned inside the HED configuration file.
+It is commonly found at /var/log/arched.log
 
-In the following chapters own Web Services will be developed. The first one will be the simple Time Web Service.
+The following chapters explain how to develop custom web services.
+The first one will be the simple Time Web Service.

### 2009-02-25 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
commited echo.wsdl and changes of the tls echo service

@@
 
-The web service presented in the previous chapter has the mayor drawback that the communication between the two endpoints is unsecure.
-Due to that reason the service  shall be extended by an additional layer called TLS (Transport Layer Security) to provide a certain degree of securtity on \task{Grid-level <-- explain that}.
-%
-In general the internet security is often subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{verfiy citation}
+The web services which were presented in the previous chapters have one mayor drawback: the communication between the two endpoints is unsecure.
+Due to that reason the current and the next chapter will discuss how to create a secure web service.
+In general security in a network can be subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{verfiy citation}
 \begin{itemize}
@@
 % Nach Sicherheit in verteilten Systemen, ITM Lübeck
-The TLS to be introduced provides: Confidentiality, authentication, integrity, non-repudiation and access control.
-It is introduced after the TCP layer. 
-%
-Before presenting the implementation of the new secure echo service the basic concept of a secure certifcate based connection shall be explained.
-Its not the goal to give an exact technical desciption but to give a first inside what is required to ensure the security of the layer.
+Ideally all six categories should have been achieved by such an application (though the last category, availability, is difficult to realise).
+In the present chapter, the security will be put into practice by extending the protocol stack with an additional security layer called TLS (Transport Layer Security).
+Due to the location after the TCP and the HTTP layer, the layer is only capable to limitate the access to the grid to a certain set of users.
+Nevertheless it provides the security categories: Confidentiality, authentication, integrity, non-repudiation and access control.\\
+
+In the next section gives a basic understanding of the working principale of the TLS.
+Emphasis is put on the security concept and not on an exact description of the technical realisation in order to enable newcomers an entry to  this area.
 
@@
 It is located directly upon the TCP/IP layers and uses typically an asymmetric encryption to establish the connection (alternatively a symmetric pre-shared key may be used).
-After the communication has successfully established, both participants are switching to another encryption method which is based on a new negotiated symmetric key.
+After the communication has successfully established, both participants are switching to another encryption method which is based on a new negotiated symmetric key.\\
 %    1. Peer negotiation for algorithm support
@@
 %
+
 Symmetric and asymmetric encryption are the two main classes of cryptography algorithms.
 In case of a symmetric encryption both participants are holding the same secret key. 
-Symmetric encryption can be ranked as being very secure but its downside is the key distribution which is very difficult in practice. 
-Ideally the secret key is transmitted over a secure channel which is either a complete diffrent medium (i.e. a letter) or an channel encrypted with another secret key.
+Symmetric encryption can be ranked as being very secure but its downside is that the key distribution is very difficult in practice.
+The key has to be shared previously, ideally over a secure channel which is may either be a complete diffrent medium (i.e.
+a letter, speech) or a channel which has been encrypted with another secret key.
 Asymmetric encryption follows a diffrent functional principle.
 While the symmetric encryption utilise one shared secret, the asymmetric ecryption is based on two keys: a private key and public key.
-As the names imply the private key is kept secret and never shared while the public key is available for everyone.
+As the names imply the private key is kept secret and is never shared while the public key is available to everyone.
 Messages encrypted with the public key can only be decrypted with the private key and the other way round, messages encrypted with the private key can only be decrypted with the public key.
-The figure~\ref{fig:asymmetric_encryption} illustrates the circumstances of the case.\\
+The Figure~\ref{fig:asymmetric_encryption} illustrates the circumstances of the case.\\
 \begin{figure}[htb]
@@
 Bob is now able to encrypt the messages such that only A is able to decrypt them.
-That way confidentiality and integrity of messages are easily realised.
+As one can see this first approach already provides confidentiality and integrity of messages.
+%
 Non-repudiation, authentication and due to that access control are not given because the identity of Alice is not proven.
-In order to be able to check identities so called certificates have been introduced.
-Certificates render the possibility to check the identity of its owner.
-They are signed by a certificate authority (CA) which's identity can be resolved to another CA or is  pre-configured to be trusted.
+In order to be able to check identities so called certificates must be introduced.
+Certificates render the possibility to validate the identity of its owner.
+They are signed by a certificate authority (CA) which's identity can be resolved to another CA or which is  pre-configured to be trusted.
 A certificate is composed of its owner's public key, his identity, the name of the signing CA and a signiture which is a hash-value of the certificate encrypted with the private key of the CA. 
-That way the CA confirms the identity of the certificate owner. 
-To obtain a certificate a request has to be submitted to a CA which verifies the identity of the requesting person and signs it, if it comes to the result that the authenticity is valid.
-Figure~\ref{fig:certificate_request} shows the procedure to gain a certificate.
+If a certificate can be resolved to a CA with is pre-configured to be trusted, the identity of the certificate owner is considered as to be proven. 
+In general a set of trusted CAs is already defined in the operation system.
+To obtain a certificate a certificate request has to be submitted to a CA.
+After the CA has verified the identity of the requesting person and the requested certificate will be signed.
+Figure~\ref{fig:certificate_request} shows the procedure of gaining a certificate.
 \begin{figure}[htb]
@@
 \end{figure}
+A private and a public key is generated by the requester and a request containing an unsigned certificate is created.
+The unsigned certificate is submitted to a CA which validates the authentity.
+If the CA decides to sign the request, a hash will be created for the request with a negotiated algorithm.
+(A fixed-sized short string which is generated by an algorithm based on a given text.
+A small change in the text causes the hash value to change almost like a random function such that the hash value can be used a fingerprint of the text.) Than the certificate will be signed by encrypted the hash value with the private key.
+%
+%
+In case a third person wants to verify the certificate owner, the encrypted hash value have to be
+
 Figure~\ref{fig:verification_of_certificates} shows a usecase where Alice wants to resolve the identity of Bob.

### 2009-02-25 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

@@
 
+\section{Listing the contents of a collection} % (fold)
+\label{sec:listing_the_contents_of_a_collection}
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{arc-storage-listing.pdf}}}
+\caption{\label{fig:listing}Listing the contents of a collection} }
+\end{figure}
+
+We want to list the contents of a collection, which has a Logical Name of \verb!/ourvo/users/we!.
+In order to do this, we have to contact a Bartender, and send a request to it containing the Logical Name we want to list, and the response from the Bartender will contain the list of entries.
+The steps are represented on Figure~\ref{fig:listing}.
+
+\begin{enumerate}
+    \item We need to know the URL of a Bartender.
+This could be preconfigured on the client side or in future releases it could be acquired from an information system.
+When we have the URL, we send a `list' request which contains the Logical Name \verb!/ourvo/users/we!.
+    \item The Bartender tries to get the metadata of the given LN by sending a `traverseLN' request to a Librarian.
+    \item The Librarian service starts the traversing by asking an A-Hash service about the first part of the LN, which is the \verb!/!
+root collection.
+The A-Hash service only knows about GUIDs and not about LNs, but the GUID of the root collection is well-known, so the A-Hash can return the metadata of it which contains the list of files and sub-collections in the root collection.
+    \item Hopefully the \verb!ourvo! collection can be found in the root collection, which means that new the Librarian knows its GUID, and can ask for its metadata from the A-Hash.
+    \item After the A-Hash returns the metadata of the \verb!/ourvo! collection, the Librarain finds the GUID of \verb!users! in it, then gets its metadata.
+    \item The A-Hash returns the metadata of \verb!/ourvo/users! which contains the GUID of \verb!we!, so the Librarian can ask for its metadata.
+    \item At last the A-Hash returns the metadata of \verb!/ourvo/users/we!
+to the Librarian, and the Librarian returns it to the Bartender.
+This metadata contains the list of entries within this collection, and it also contains the access policies for this collection.
+    \item The Bartender first checks if based on our DN (or our VO membership) and the access policies of this collection do we have rights to get the contents of this collection or not.
+If we are approved, then because the `list' request should return additional metadata about each entry in the collection, the Bartender send a `get' message to a Librarian requesting metadata of all the entries in this collection.
+    \item The Librarian gets the data from the A-Hash and returns it to the Bartender.
+    \item Now the Bartender has all the needed information, so it could return the proper response to us, the user. Our client tool formats and prints the results nicely.
+\end{enumerate}
+
+% section listing_the_contents_of_a_collection (end)
+
 \section{Downloading a file} % (fold)
@@
 
-We want to download a file about which we know that it is somewhere in our home collection on the storage (see Figure~\ref{fig:downloading}).
-The LN of our home collection is e.g.~\verb#/ourvo/users/we# (here a `home collection' could be just a collection which is given to us by our VO).
-We can get a list of entries in this collection from any Bartender.
+In this use case we want to download a file which has a Logical Name of \verb#/ourvo/users/we/thefilewewant# (see Figure~\ref{fig:downloading}).
 
 \begin{enumerate}
-\item We need to find a Bartender.
-Maybe we have a cached list of recently used Bartenders or we can get one from the information system.
-When we have an endpoint reference of a Bartender, we could call its list method with the LN \verb#/ourvo/users/we#.
-\item The Bartender has to find a Librarian service, again using its cache of recently used Librarian services or get a new one from the information system.
-When the Bartender has an endpoint reference of a Librarian service, it could ask the Librarian to traverse the LN \verb#/ourvo/users/we#.
-\item The Librarian needs the A-Hash service to access the stored data, when it has the endpoint reference of the A-Hash service, it could get the information about the root collection, which contains the GUID of the \verb#ourvo# sub-collection.
-Then the Librarian gets the entries of this \verb#ourvo# collection, and in it it can find the GUID of \verb#users#, and in the entries of \verb#users# there is the GUID of \verb#we#, which the Librarian returns to the Bartender with all the metadata.
-\item The Bartender now has the GUID and the metadata of the collection \verb#/ourvo/users/we#, including the list of its entries. This is returned to us.
-\item So we get the list of our \verb#/ourvo/users/we# collection, and now we realize that the file we want has the LN \verb#/ourvo/users/we/thefilewewant# and we know the GUID of it as well: e.g.~\verb#a4b2e#.
-(Of course we know the GUID of the \verb#/ourvo/users/we# collection too, which is e.g.~\verb#13245# and using this we could refer to our file as \verb#13245/thefilewewant# which means the entry called \verb#thefilewewant# in the collection with a GUID \verb#13245#.) We connect a Bartender again (the same one or maybe another one) to get the file with any of these LNs, the \verb#a4b2e# is the fastest solution because the Bartender need not to look up the whole LN again in the Librarian, a well-written client API should use this.
-With the get request we give the Bartender the list of transfer protocols we are able to use.
-\item The Bartender contacts the Librarian to get the locations of the replicas of this file.
-\item The Librarian returns all the metadata of the requested LN.
-\item The Bartender chooses a `location' which consists of the ID of a Shepherd, and the referenceID of the file within the Shepherd.
-Using the information system or its local cache it could get the endpoint reference of the Shepherd.
-The Bartender initiates a transfer by the Shepherd, if the Shepherd supports one of the transfer protocols we give, it can create a transfer URL (TURL) with a protocol we can download.
-\item The Shepherd returns the TURL to the Bartender.
+    \item We connect a Bartender and send a `getFile' request with the LN of our file.
+    \item The Bartender contacts a Librarian to traverse the Logical Name and to get the metadata of our file.
+    \item The Librarian do the step by step traversing and gets all the data from the A-Hash and returns the metadata of our file to the Bartender.
+This metadata contains the location of the file's replicas and the access policies of this file.
+    \item The Bartender checks based on the access policies and our identity if we are allowed to get the file, and if we are good to go, then it chooses a replica location.
+A location consists of the URL of a Shepherd service, and the ID of the replica within that Shepherd (which is called a `referenceID').
+The Bartender sends a `get' request to the chosen Shepherd.
+    \item The Shepherd prepares the file transfer by asking the storage element service to create a new one-time URL for this replica.
+This URL called the transfer URL, and it will only be valid for one download.
+The Shepherd returns the TURL to the Bartender.
     \item The Bartender returns the TURL to us.
-\item Now we have a TURL from which we can download it.
+    \item Now we use this TURL to get the file directly from the storage element service on the storage node.
 \end{enumerate}
@@
 % section downloading_a_file (end)
+
+\section{Creating a collection} % (fold)
+\label{sec:creating_a_collection}
+\begin{figure}[ht]
+\centering{{\scalebox{0.7}{\includegraphics{arc-storage-makecollection.pdf}}}
+\caption{\label{fig:makecollection}Creating a collection} }
+\end{figure}
+
+We want to create a new (empty) collection as sub-collection of \verb#/ourvo/common#, and we want to call it \verb!docs! (see Figure~\ref{fig:makecollection})
+
+\begin{enumerate}
+    \item We contact a Bartender and send a `makeCollection' request with the LN \verb!/ourvo/common/docs!.
+    \item The Bartender asks a Librarian to traverse this LN.
+    \item The Librarian try to traverse the Logical Name and it stops at the last possible point and returns the metadata of the last element.
+Because we want to put our collection to a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed.
+If the Librarian could traverse the whole LN that would mean that there is already an existing file or collection by that name.
+If the parent collection does exist then the Librarian can traverse it which means that the parent collection's metadata is returned to the Bartender.
+    \item The Bartender checks the access policies to decide if we have permissions to put something into this collection. Then it asks the Librarian to create a new collection.
+    \item The Librarian creates the collection, and returns its GUID.
+At this point this new collection has no real Logical Name yet, it only has a GUID, but it is not yet put into its parent collection.
+    \item The Bartender now asks the Librarian to add this new entry into the parent collection, which means that the new GUID and the name \verb!docs! are added as a pair.
+    \item The Librarian returns with a status message.
+    \item Finally the Bartender tells us if everything went OK or not.
+\end{enumerate}
+
+% section creating_a_collection (end)
 
@@
 \begin{enumerate}
-\item We contact a Bartender to put the file, we give the size and checksum and other metadata, and the transfer protocols we want to use.
-And of course we give the Logical Name we want to be the name of the file, which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
+    \item We contact a Bartender to put the file, we give the size and checksum and other metadata.
+And of course we give the Logical Name where we want to put the file, which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
     \item The Bartender ask a Librarian to traverse this LN.
-\item If the Librarian can traverse the whole LN then this LN is already exists, but if the LN is still available and the parent exists, the Librarian’s response contains the GUID and metadata of the parent collection.
-\item Then the Bartender creates a new file entry within the Librarian with all the information we gave. 
-\item The Librarian returns the GUID of this new entry.
-\item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf# which points to a file which has no replica at all.
-If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message `try again later'.
-\item The Bartender (using the information system) chooses a Shepherd and gets its endpoint reference.
-Then the Bartender initiates uploading of the file to the Shepherd: the request includes the size and checksum of the file, the GUID, and the protocols we are able to use.
-\item The Shepherd creates a transfer URL and a referenceID for this file and registers the GUID of the file in its own database and reports to the Librarian that there is a new replica with state \verb#creating#.
-The Librarian gets the message from the Shepherd and creates a new entry in the locations list of the given file with the serviceID and the referenceID the Shepherd have just reported.
-If someone tries to download this file now, still gets a `try again later' error message, because this new replica is still not \verb#alive#.
+    \item The Librarian traverses the Logical Name, and we expect it to stop at the \verb#/ourvo/common/docs# part of the LN, because that means that the name is available and the parent collection exists.
+If everything is fine, the metadata of the parent collection is returned to the Bartender.
+    \item The Bartender checks the access policies to decide if we have permissions to put something into this collection. Then it asks the Librarian to create a new file entry. 
+    \item The Librarian creates the entry and returns its GUID.
+    \item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf#.
+However this LN points to a file which has currently no replica at all.
+If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message.
+    \item The Bartender asks the Librarian about Shepherd services (which are sitting on storage nodes).
+    \item The Librarian returns a list of Shepherd services
+    \item The Bartender chooses a Shepherd service and sends it a `put' request to initiate the file upload. 
+    \item The Shepherd communicates with the storage element service on the same node to create a new transfer URL (TURL).
+Then it creates a `referenceID' for this file and then reports to the Librarian that there is a new replica in \verb#creating# state.
+The Librarian gets the message from the Shepherd and adds the new replica to our new file.
+Now the file has one replica, which is not uploaded yet into the system.
+If someone tries to download this file now, still gets an error message.
     \item The Shepherd returns the the TURL to the Bartender.
@@
     \item Then we can upload the file to this TURL.
-\item The Shepherd detects that the file is arrived and reports the change of state to \verb#alive# to the  Librarian who alters the state in the given file-entry.
-At this point the file has only one replica.
+    \item The Shepherd detects that the file is arrived.
+It checks the checksum of the file, and if it is OK, then it reports to the Librarian, that this replica is now \verb#alive#.
+The Librarian alters the state of this location, and now finally the file has one valid replica.
 \end{enumerate}
-\begin{itemize}
-\item The Shepherd periodically checks the Librarian if this is less than the needed replica number, and if it is then it initiates creating a new replica by a Bartender.
-\item The Bartender chooses another Shepherd, initiates the transfer then returns the TURL to the  first Shepherd which uploads the file to the new Shepherd.
-\item All the Shepherds check periodically whether their files have enough replica, and if any of them find that there is more replica needed, it initiates creating a new.
-If more than one Shepherd of course could cause that there will be more replicas than needed.
-If a Shepherd finds out that a file has more replicas than needed it notifies a Bartender about it.
-\item The Bartender ask the Librarian about all Shepherds this file has replicas on, flags this file as `removing a replica' which prevents other Bartenders to remove an other replica accidentally, then make a decision of which one is to be removed, then contacts the chosen Shepherd and asks it to remove the replica.
-The Shepherd then notify the Librarian, and the Librarian removes the replica, and removes the flag `removing a replica' as well.
-\item If the client cannot upload the file to the given TURL for some reason, it is possible to call \verb#addReplica# to get a new TURL without removing and recreating the file.
-\end{itemize}
-
+í
 % section uploading_a_file (end)
@@
     \item If we want to remove a file, we should connect to a Bartender with the LN of the file we want to remove.
-    \item The Bartender asks the Librarian to traverse the LN and return the list of parent collections of this file.
-    \item The Librarian returns the data.
-    \item If the file has only parent, it asks the Librarian to remove the file entry itself, and the entry from the parent collection.
-If the file has more parent collections (hard links) the Bartender only removes the entry from the parent collection and removes the parent from the list of parent  collections of this file.
-    \item Next time the Shepherd does its periodic check, it asks the Librarian about each of its stored replicas, and finds out that one of them no longer exists, so it removes the replica from the storage node.
+    \item The Bartender asks the Librarian to traverse the LN.
+    \item The Librarian returns the metadata of the file. The metadata contains information about all the hardlinks which points to this file if there are more than one.
+    \item Now the Bartender asks the Librarian to remove the file.
+    \item After that, the Bartender asks the Librarian to remove the links to this file from all the parent collections.
+    \item Next time a Shepherd which has a replica of this file does its periodic check, it asks the Librarian about the file, and notices that the file does not exist anymore, so it removes the replica itself from the storage node.
 \end{enumerate}

### 2009-02-26 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Enhanced the chapter about the service using TLS

@@
 \end{figure}
-A private and a public key is generated by the requester and a request containing an unsigned certificate is created.
-The unsigned certificate is submitted to a CA which validates the authentity.
-If the CA decides to sign the request, a hash will be created for the request with a negotiated algorithm.
-(A fixed-sized short string which is generated by an algorithm based on a given text.
-A small change in the text causes the hash value to change almost like a random function such that the hash value can be used a fingerprint of the text.) Than the certificate will be signed by encrypted the hash value with the private key.
+In the first step, a private and a public key will be generated by the requester.
+The public key together with a textual description of the identity  depict the request which will be transmitted to a CA.
+After the CA successfully varified the validity of the request, the identity of the CA and the name of used the encryption algorithms are added.
+The collected data will represent the plain text of the certificate. In order to ensure the integrity of the text, 
+a hash value will be created.
+(A fixed-sized short string which is generated by an algorithm based on an arbitrary long given text.
+The algorithm is designed such that a small change in the text will cause the hash value to change almost like a random function.
+Thus, the hash value can be used a fingerprint of the text.) The CA signs the plain certificate by encrypting its hash value with its private key.
+Finally the signed certifcate will be returned to the requester.
 %
 %
-In case a third person wants to verify the certificate owner, the encrypted hash value have to be
+If an outside person wants to verify the certificate owner, the encrypted hash value has to be decrypted by the public key of the CA and a hash value has to be created with the same algorithm in order to compare them.\\
 
-Figure~\ref{fig:verification_of_certificates} shows a usecase where Alice wants to resolve the identity of Bob.
+Figure~\ref{fig:verification_of_certificates} shows a usecase in which Alice wants to resolve the identity of Bob.
 Alice submitts her request to Bob along with a challange consisting of a random number $n$.
-Bob encrypts the number with his private key and returns it together with his certificate.
-With the usage of the public key, Alice is able to decrypt the challange which implies that, if the certificate is to be trusted, the identity of Bob is to be trusted.
-Unfortunately Alice is unable to verify the certificate because it is signed by an unknown CA called CA 1.
-She establishes the connection to CA 1 in order to resolve the identity of that certificate authority.
-Again Alice sends a challenge --- this time to the certificate authority CA 1.
-CA 1 encrypts the challange with its private key and submits the result along with its certificate.
-Again Alice validates the certificate sucessfully.
-The certificate was signed by a CA called CA 2 which she is able to locate as a pre-configured certificate. Due to the reason she trusts CA 2, she can trust CA 1 and finally Bob.
-
+The random number ensures that it is not possible to repeat a message (integrity).
+Bob encrypts the number with his private key and returns it together with his certificate. Alice uses the public key of Bob to decrypt the challange. 
+If the decrypted number is equal to the transmitted number, Alice can reason that the identity of Bob is to be trusted in case she is trusting the CA which has signed Bobs certificate. 
+%
+In the given example Alice only trusts the CA 2 but not CA 1.
+In order to resolve the identity, Alice establishes a connection to CA 1 and submits another random number $n$ as a challenge.
+The CA 1 encrypts the challange with its private key and send the result along with its certificate back to Alice.
+Alice succefully validates the certificate of CA 1.
+Due to the certificate was signed by CA 2 which is pre-configured to be trusted she can follow to trust CA 1 and in conclusion the identity of B.
+In summary, Alice knows she is talking to Bob.
+If Bob also wants to be sure how he's talking to the same procedure has to be done for the identity of Alice.
+The TLS mode in in which both endpoints are ensuring themself to whom they are talking to is called bilateral connection.
+Due to the client is aware of the service authentity, the confidential treatment of his data is given such that it won't be revealed to a third party (service sided confidentiality).
+Moreover the access to the service can be controled (access control) and the results are only passed back to the right person (client sided confidentiality).
+As one can see confidentiality, authentication, integrity and non-repudiation are now given.\\
 \begin{figure}[htb]
@@
 
-The advantages of certificates are
-o easy scalable to a larger set of users
+Certificates have several advantages.
+They are scaling very good with the amount of users because CAs can create new sub-CAs such that the load is balanced (certificates of a CA may be cached).
+Furthermore no critical information is exchanged which may gets intercepted - in particular no secret key is shared.
+Instead a set of pre-configured CAs which will be in general provided by the operation system is needed.
+In respect of ARC it is obvious, that the access to the grid should be under control of its administrator.
+To do so, the administrator have to build up his own CA to be in charge of creating certificates for a set of users.
+The created CA have to be shipped into the directory of pre-configured CAs.
+(Within the source code there are two shell scripts to create a CA and related certificates). Due to there's a bunch of tools for certificates (like \textit{openssl}) it is easy to maintain them.
+
+o Infrastructure is given and easily to maintain
+
+o In general the pre-configured organistion take money to sign certifcates, but they can also be created locally.
+
 o VOMs are realisable
-o Infrastructure is given and easily to maintain
-o In general the pre-configured organistion take money to sign certifcates, but they can also be created locally.
+
+ARC:
 
@@
 o What do we need to get a service running
+
 clientCA   - Authority guarantee for identity
@@
 Its  clientCERT clientKEY and some kind of root CA knowing the server
+
+\task{RENAMING into~ARC service configuration file}
+\task{echo service client configuration file}
 

### 2009-02-26 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

@@
 
-\chapter{Design Overview}
+\chapter{Design Overview} % (fold)
 \label{cha:overview}
@@
 % section security (end)
+
+% chapter design_overview (end)
 
@@
 
-\chapter{Technical description and implementation status} % (fold)
-\label{cha:technical_description_and_implementation_status}
+\chapter{Technical description} % (fold)
+\label{cha:technical_description}
+
+\section{Framework and language} % (fold)
+\label{sec:framework_and_language}
 
@@
 
-The current version of the prototype has no information system and no security, these are soon to be integrated to the system.
-
-The information system is needed to discover services, and to translate \emph{serviceID}s to endpoint references (URLs).
-Currently the URLs are written in the configuration files, and the Shepherd services are reporting their URLs to the Librarian, so the Bartender could ask for all alive Shepherds.
-
-Some self-healing mechanisms are not implemented yet, e.g.
-the files and collections contains their parent-collections as their metadata, and of course a collection contains a list of its files and sub-collections, and if somehow this information became inconsistent, the system should detect and repair it.
-This is not implemented yet.
-
-\section{Plans for security} % (fold)
-\label{sec:plans_for_security}
-
-Security is needed to do proper authorization of the users, and to manage access policies of files and collections.
-ARC has its own policy language, for each file and collection there will be a policy XML document stored as a metadata.
-The storage services will use these policies and the properties extracted from the communication channel to make authorization decisions.
-If the properties and the policies are present, the decision will be actually made by the security framework of HED.
-
-These are the planned actions which can be used for access control:
-\begin{itemize}
-    \item \emph{read}: user can get the list of entries in the collection; user can download the file
-    \item \emph{addEntry}: user can add a new entry to the collection;
-    \item \emph{removeEntry}: user can remove any entry from the collection 
-    \item \emph{delete}: user can delete the collection if it is empty; user can delete a file (if you want to remove a file/collection, then the Bartender needs to remove the entry from the parent collection, and then delete the file/collection itself, so you need to have both permissions)
-    \item \emph{modifyPolicy}: user can modify the policy of the file/collection
-    \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
-    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are key-value pairs)
-\end{itemize}
-When a user has the permission in the Librarian to download a file then the user should have permission to access at least one of the file's replica, so there should be a Shepherd which allows the user the get the file.
-If the Bartender has permission to access the Shepherd, then the Bartender should create an assertion which allows the user to access the file.
-This could be a signed token which contain a policy defining access to particular file.
-But all these are currently just plans.
-
-% section plans_for_security (end)
-
-Further prototype statuses and plans can be found below within each section about the services.
-
-% chapter technical_description_and_implementation_status (end)
+Because the next-generation information system of ARC is currently under development, we cannot use it to discover services, that's why currently the URLs of almost all the services are hardcoded in the configuration files of the service.
+There is one exception: the Shepherd services are reporting their URLs to the Librarians, so a Bartender always could get a very current list of alive Shepherd services.
+
+The HED also has a security framework which the Bartenders use the make access policy decision . The details of this is described in section X.
+
+% section framework_and_language (end)
+
+\section{Data model} % (fold)
+\label{sec:data_model}
+
+The ARC storage system stores different kinds of metadata about files, collections, mount point, Shepherd services, etc. Each of these has a unique ID which we call `GUID'.
+
+The A-Hash services provide a functionality to store `objects', where each object has a unique ID, and contains key-value pairs organized in sections.
+The properties, the values and the section names too are simple character strings.
+The A-Hash services provide a simple interface to interact with these objects, e.g.~to do conditinal and atomic changes.
+
+The Librarian services use the A-Hash services to store the metadata of all kinds, using the GUIDs as unique IDs.
+For this, we have to represent the metadata as key-value pairs organized in section.
+The files, the collections and the mount points have some common attributes: the GUID, the type and the owner which are in the \emph{entry} section; the creation and modification timestamps in the \emph{timestamps} section, a list of access control rules in the \emph{policy} section; the list of parent collections where this entry has a `hardlink'; and each entry could have arbitrary key-value pairs in the \emph{metadata} section.
+
+\begin{description}
+    \item [entry] section 
+    \begin{itemize}
+        \item \emph{type}: the type of the entry: `collection', `file' or `mountpoint'
+        \item \emph{GUID}: the unique ID of the entry
+        \item \emph{owner}: the DN of the user how ownes this entry
+    \end{itemize}
+    \item [timestamps] section 
+    \begin{itemize}
+        \item \emph{created}: timestamp of creation 
+        \item \emph{modified}: timestamp of last modification (the support for this is not implemented yet)
+    \end{itemize}
+    \item [policy] section 
+    \begin{itemize}
+        \item \emph{(identity, action list) pairs}: a list of allowed and not allowed actions for different identites (users, VOs, etc.). 
+    \end{itemize}
+    \item [parents] section
+    \begin{itemize}
+        \item list of GUIDs of collection where this entry is located, and the name of this entry within the given collections
+    \end{itemize}
+    \item [metadata] section 
+    \begin{itemize}
+        \item any other arbitrary key-value pairs
+    \end{itemize}
+\end{description}
+
+\subsection{Files} % (fold)
+\label{sub:files}
+
+A file in the ARC storage has a logical entry in the namespace, and a couple of physical replicas on the storage nodes.
+We store the locations of the replicas in the \emph{location} section.
+A location contains the URL of the Shepherd service, the local ID of the file within the storage node and the state of the replica.
+This state could be `\textbf{alive}' (if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the Shepherd claims it has no such file), `\textbf{offline}' (if the Shepherd is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{thirdwheel}' (if the replica is marked for deletion because of too many replicas).
+We also store (in the \emph{states} section) the size of the file, the number of needed replicas, and a checksum of the file, which is created by some kind of checksuming algorithm (currently only md5 is supported).
+
+\begin{description}
+    \item [locations] section 
+    \begin{itemize}
+        \item \emph{(location, state) pairs}, where a location is a (\emph{URL}, \emph{referenceID}) pair serialized as a string, where \emph{URL} is the address of the Shepherd service storing this replica, \emph{referenceID} is the ID of the file within that Shepherd service.
+    \end{itemize}
+    \item [states] section 
+    \begin{itemize}
+        \item \emph{size}: the file size in bytes
+        \item \emph{checksum}: checksum of the file
+        \item \emph{checksumType}: the name of the checksum method
+        \item \emph{neededReplicas}: how many valid replicas should this file have 
+    \end{itemize}
+\end{description}
+
+% subsection files (end)
+
+% The Librarian uses the A-Hash to store all the data about files and collections.
+The A-Hash is capable of storing key-value pairs organized in sections, which actually means that it stores (\emph{section}, \emph{key}, \emph{value}) tuples where each member is simply a string, e.g.~(`entry', `type', `collection') or (`ACL', `johnsmith', `owner') or (`timestamps', `created', `1196265901') or (`locations', `64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB:9A293F27DC86', `sentenced').
+There could be only one \emph{value} for a (\emph{section}, \emph{key}) pair.
+
+\subsection{Collections} % (fold)
+\label{sub:collections}
+
+A \emph{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-hierarchy.
+Each entry has a unique name within a collection and a GUID which points to corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section.
+The collections could be in a closed state, which means that its contents should not be changed.
+If you close a collection then it cannot be opened again.
+However it cannot be garanteed that the contents of a closed collection remains the same (e.g.
+if we include in our collection a file which we don't own, then the owner of the file could remove it nevertheless), but if a closed collection is changed, its state will be broken, and this state could never be changed again, which means you will always know if a closed collection is not intact anymore.
+This state is stored in the \emph{states} section, and its values could be `\textbf{no}' (if the collection is not closed), `\textbf{yes}' (if the collection is closed) and `\textbf{broken}' (if the collection was closed, but something happened, and its contents have been changed).
+
+\begin{description}
+    \item [entries] section 
+    \begin{itemize}
+        \item \emph{(name, GUID) pairs}: a collection is basically a list of name-GUID pairs. 
+    \end{itemize}
+    \item [states] section 
+    \begin{itemize}
+        \item \emph{closed}: this indicates if the collection is closed or broken. 
+    \end{itemize}
+\end{description}
+
+% subsection collections (end)
+
+\subsection{Mount Points} % (fold)
+\label{sub:mount_points}
+
+The mount point entry is a reference to an external third-party storage. Its metadata is basically just the URL of the external storage.
+
+\begin{description}
+    \item [mountpoint] section 
+    \begin{itemize}
+        \item \emph{externalURL}: the URL of the external storage.
+    \end{itemize}
+\end{description}
+
+% subsection mount_points (end)
+
+\subsection{Shepherds} % (fold)
+\label{sub:shepherds}
+
+The Librarian stores information about the registered Shepherd services.
+Each Shepherd reports its URL and the list of its files to a Librarian, and for each Shepherd a GUID is created.
+There is a special entry (with GUID `1' by default) which stores a list of the registered Shepherd services, storing their GUID and the timestamp of the last heartbeat message from them.
+
+\begin{description}
+	\item [nextHeartBeat] section
+	\begin{itemize}
+		\item (URL, timestamp) pairs contains when was the last heartbeat of this service
+	\end{itemize}
+	\item [serviceGUID] section
+    \begin{itemize}
+        \item (URL, GUID) pairs connects the URL of the Shepherd to the GUID where the information is stored about the Shepherd's stored files
+    \end{itemize}
+\end{description}
+
+And for each Shepherd there is a separate entry with the list of all files stored on the given storage nodes:
+
+\begin{description}
+	\item [entry] section
+    \begin{itemize}
+        \item \emph{type}: `shepherd'
+    \end{itemize}
+	\item [files] section
+	\begin{itemize}
+	    \item \emph{(referenceID, GUID) pairs} for each replica stored on the Shepherd
+	\end{itemize}
+\end{description}
+
+% subsection shepherds (end)
+
+% section data_model (end)
+
+\section{Security implementation} % (fold)
+\label{sec:security_implementation}
+
+The ARC HED hosting environment has a security framework, and it has its own policy language for describing access policies and requests.
+While the storage system use its own format to store the access policies, but one the time comes to make decision, it converts the stored policies and the incoming requests to this ARC native policy format, and uses the security framework to decide.
+
+%%%%%%%%%%%%%%%%
+
+% section security_implementation (end)
+
 \newpage
@@
 Currently it only has a centralized implementation.
-It stores \emph{objects}, where each object has an arbitrary string \emph{ID}, and contains any number of \emph{property}-\emph{value} pairs grouped in \emph{section}s, where \emph{property}, \emph{value} and \emph{section} are arbitrary strings.
-There could only be a single \emph{value} for a \emph{property} in a \emph{section}.
-
-If you have an ID, you can get all property-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties do you need.
-You can add or remove property-value pairs of an object or delete all occurrences of a property or create a new object with the \emph{change} method, and you can specify conditions, which means the change is only applied if the given conditions are met.
+It stores \emph{objects}, where each object has an arbitrary string \emph{ID}, and contains any number of \emph{key}-\emph{value} pairs grouped in \emph{section}s, where \emph{key}, \emph{value} and \emph{section} are arbitrary strings.
+There could only be a single \emph{value} for a \emph{key} in a \emph{section}.
+
+If you have an ID, you can get all key-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties do you need.
+You can add or remove key-value pairs of an object or delete all occurrences of a key or create a new object with the \emph{change} method, and you can specify conditions, which means the change is only applied if the given conditions are met.
 
@@
     \item \emph{ID} is an arbitrary string
-    \item \emph{object} contains property-value pairs in sections, technically it is a list of key-\emph{value} pairs where the key is a (\emph{section}, \emph{property}) tuple
+    \item \emph{object} contains key-value pairs in sections, technically it is a list of key-\emph{value} pairs where the key is a (\emph{section}, \emph{key}) tuple
 \end{itemize}
@@
 
-    The \emph{ids} is a list of string \emph{ID}s, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs.
-For each \emph{ID} it returns all the \emph{value}s for each \emph{property} in each \emph{section} (filtered by \emph{neededMetadata}), so \emph{object} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+    The \emph{ids} is a list of string \emph{ID}s, \emph{neededMetadata} is a list of (\emph{section}, \emph{key}) pairs.
+For each \emph{ID} it returns all the \emph{value}s for each \emph{key} in each \emph{section} (filtered by \emph{neededMetadata}), so \emph{object} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples.
 
@@
 
-    \emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}, \emph{conditions}), where \emph{changeID} is an arbitrary ID to identify in the response which change was successful; \emph{ID} points to the object we want to change; \emph{changeType} can be `\textbf{set}' (set the property within the section to value), `\textbf{unset}' (remove the property from the section regardless of the value), `\textbf{delete}' (removes the whole object), conditions is a list of (\emph{conditionID}, \emph{type}, \emph{section}, \emph{property}, \emph{value}) tuples, where \emph{type} could be `\textbf{is}' (the property in the section is set to the value), `\textbf{isnot}' (the property in the section is not set to the value), `\textbf{isset}' (the property of the section is set to any value), `\textbf{unset}' (the property of the section is not set at all).
+    \emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{key}, \emph{value}, \emph{conditions}), where \emph{changeID} is an arbitrary ID to identify in the response which change was successful; \emph{ID} points to the object we want to change; \emph{changeType} can be `\textbf{set}' (set the key within the section to value), `\textbf{unset}' (remove the key from the section regardless of the value), `\textbf{delete}' (removes the whole object), conditions is a list of (\emph{conditionID}, \emph{type}, \emph{section}, \emph{key}, \emph{value}) tuples, where \emph{type} could be `\textbf{is}' (the key in the section is set to the value), `\textbf{isnot}' (the key in the section is not set to the value), `\textbf{isset}' (the key of the section is set to any value), `\textbf{unset}' (the key of the section is not set at all).
     If all conditions are met, tries to apply changes to the objects, creates a new object if a previously non-existent ID is given.
@@
 \label{sub:data_model}
-Each librarian entry has a unique ID called \emph{GUID}.
-
-The Librarian uses the A-Hash to store all the data about files and collections.
-The A-Hash is capable of storing property-value pairs organized in sections, which actually means that it stores (\emph{section}, \emph{property}, \emph{value}) tuples where each member is simply a string, e.g.~(`entry', `type', `collection') or (`ACL', `johnsmith', `owner') or (`timestamps', `created', `1196265901') or (`locations', `64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB:9A293F27DC86', `sentenced').
-There could be only one \emph{value} for a (\emph{section}, \emph{property}) pair.
-
-\begin{itemize}
-    \item A \textbf{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-hierarchy.
-Each entry has a name which is only valid within this collection, and it is unique within the collection.
-Each entry is referenced by its GUID.
-So the metadata sections of a collection are as follows:
-    \begin{description}
-        \item [entry] section 
-        \begin{itemize}
-            \item \emph{type}: `collection' 
-        \end{itemize}
-        \item [entries] section 
-        \begin{itemize}
-            \item \emph{(name, GUID) pairs}: a Collection is basically a list of name-GUID pairs. 
-        \end{itemize}
-        \item [timestamps] section 
-        \begin{itemize}
-            \item \emph{created}: timestamp of creation 
-            \item \emph{modified}: timestamp of last modification 
-        \end{itemize}
-        \item [states] section 
-        \begin{itemize}
-            \item \emph{closed}: if the collection is closed, then nothing can be added to its contents 
-        \end{itemize}
-        \item [policies] section 
-        \begin{itemize}
-            \item XML representations of access policies 
-        \end{itemize}
-        \item [metadata] section 
-        \begin{itemize}
-            \item any other arbitrary metadata 
-        \end{itemize}
-    \end{description}
-    \item A \textbf{file} entry contains the following sections: 
-    \begin{description}
-        \item [entry] section 
-        \begin{itemize}
-            \item \emph{type}: `file' 
-        \end{itemize}
-        \item [locations] section 
-        \begin{itemize}
-            \item \emph{(location, state) pairs}, where a location is a (\emph{serviceID}, \emph{referenceID}) pair serialized as a  string, where \emph{serviceID} is the ID of the Shepherd service storing this replica, \emph{referenceID} is the ID of the file within that Shepherd service, and state could be `\textbf{alive}' (if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the Shepherd claims it has no such file), `\textbf{offline}' (if the Shepherd is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{sentenced}' (if the replica is marked for deletion)
-        \end{itemize}
-        \item [timestamps] section 
-        \begin{itemize}
-            \item \emph{created}: timestamp of creation 
-            \item \emph{modified}: timestamp of last modification (e.g.~modification of metadata)
-        \end{itemize}
-        \item [states] section 
-        \begin{itemize}
-            \item \emph{size}: the file size in bytes
-            \item \emph{checksum}: checksum of the file
-            \item \emph{checksumType}: the name of the checksum method
-            \item \emph{neededReplicas}: how many valid replicas should this file have 
-        \end{itemize}
-        \item [policies] section 
-        \begin{itemize}
-            \item XML representations of access policies 
-        \end{itemize}
-        \item [metadata] section 
-        \begin{itemize}
-            \item any other arbitrary metadata 
-        \end{itemize}
-    \end{description}
-    \item There is one more type of Librarian entries called \textbf{mount point} which is a reference to a service which is capable of handling a subtree of the namespace.
-The properties of a mount point in sections:
-    \begin{description}
-        \item [entry] section 
-        \begin{itemize}
-            \item \emph{type}: `mountpoint' 
-        \end{itemize}
-        \item [mount] section 
-        \begin{itemize}
-            \item \emph{target}: the ID of the service
-            \item \emph{ID}: an ID within the service (optional)
-        \end{itemize}
-        \item [timestamps] section 
-        \begin{itemize}
-            \item \emph{created}: timestamp of creation 
-            \item \emph{modified}: timestamp of last modification (e.g.~modification of metadata)
-        \end{itemize}
-        \item [policies] section 
-        \begin{itemize}
-            \item XML representations of access policies 
-        \end{itemize}
-        \item [metadata] section 
-        \begin{itemize}
-            \item any other arbitrary metadata 
-        \end{itemize}
-    \end{description}
-
-    \item The Librarian stores information about the Shepherds, so each Shepherd has a GUID as well.
-There is an entry (with GUID `1' by default) which contains the GUID and the timestamp of the last heartbeat for each registered Shepherd:
-    \begin{description}
-    	\item [nextHeartBeat] section
-    	\begin{itemize}
-    		\item (ID, timestamp) pairs
-    	\end{itemize}
-    	\item [serviceGUID] section
-        \begin{itemize}
-            \item (ID, GUID) pairs
-        \end{itemize}
-    \end{description}
-
-    \item For each Shepherd there is a separate entry with the list of files:
-    \begin{description}
-    	\item [entry] section
-        \begin{itemize}
-            \item \emph{type}: `shepherd'
-        \end{itemize}
-    	\item [files] section
-    	\begin{itemize}
-    	    \item \emph{(referenceID, GUID) pairs} for each replica stored on the Shepherd
-    	\end{itemize}
-    \end{description}
-\end{itemize}
 
@@
 
-    \emph{newRequestList} is a list of (\emph{requestID}, \emph{metadata}) where \emph{requestID} is an arbitrary ID used to identify this request in the list of responses; \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+    \emph{newRequestList} is a list of (\emph{requestID}, \emph{metadata}) where \emph{requestID} is an arbitrary ID used to identify this request in the list of responses; \emph{metadata} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples.
     This method generates a \emph{GUID} for each request, and inserts the new entry (with the given metadata) into the A-Hash, then returns the GUIDs of the newly created entries.
-The (`entry', `type') property of the metadata contains whether it is a file or a collection.
+The (`entry', `type') key of the metadata contains whether it is a file or a collection.
 
@@
 
-    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the property in the section to value), `\textbf{unset}' (remove the property-value pair from the section), `\textbf{add}' (set the property in the section to value only if it is not exists already).
+    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{key}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the key in the section to value), `\textbf{unset}' (remove the key-value pair from the section), `\textbf{add}' (set the key in the section to value only if it is not exists already).
 
@@
 
-    \emph{GUIDs} is a list of GUIDs, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs indicating only which properties we need, \emph{getResponse} is a list of (\emph{GUID}, \emph{metadata}) where metadata is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+    \emph{GUIDs} is a list of GUIDs, \emph{neededMetadata} is a list of (\emph{section}, \emph{key}) pairs indicating only which properties we need, \emph{getResponse} is a list of (\emph{GUID}, \emph{metadata}) where metadata is a list of (\emph{section}, \emph{key}, \emph{value}) tuples.
 This method returns the metadata of all the \emph{GUID}s filtered with \emph{neededMetadata}.
@@
     \begin{description}
-        \item[metadata] is all the metadata of the of traversedLN in the form of (\emph{section}, \emph{property}, \emph{value}) tuples
+        \item[metadata] is all the metadata of the of traversedLN in the form of (\emph{section}, \emph{key}, \emph{value}) tuples
         \item[GUID] is the \emph{GUID} of the \emph{traversedLN}
@@
 
-    \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).
-
-    \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.
+    \emph{getRequestData} is a list of (\emph{key}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).
+
+    \emph{getResponseData} is a list of (\emph{key}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.
 
@@
 
-    \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (for additional access policy).
-
-    \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.
+    \emph{putRequestData} is a list of (\emph{key}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (for additional access policy).
+
+    \emph{putResponseData} is a list of (\emph{key}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.
 
@@
 simply refers to the root collection.
-The term `\emph{metadata}' here refers to a list of property-value pairs organized in sections, see the data model description in Section~\ref{sub:data_model}.
+The term `\emph{metadata}' here refers to a list of key-value pairs organized in sections, see the data model description in Section~\ref{sub:data_model}.
 
@@
 
-    \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other property-value pairs in the `\textbf{metadata}' section.
+    \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other key-value pairs in the `\textbf{metadata}' section.
 The returned \emph{TURL} is a URL with a chosen \emph{protocol} to upload the file itself, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.
@@
 
-    \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sub:data_model})
+    \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{key}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sub:data_model})
 
@@
 
-    \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in key-value pairs.
+    \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' key (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in key-value pairs.
 The \emph{success} in the response could be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.
@@
 
-    \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
-
-    \emph{listResponse} is a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (Section~\ref{sub:data_model}), the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}' (because only collections can be listed).
+    \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{key}) pairs which filters the returned metadata.
+
+    \emph{listResponse} is a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples according to the data model of the Librarian (Section~\ref{sub:data_model}), the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}' (because only collections can be listed).
 
@@
 
-    \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).
+    \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{key}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{key} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{key}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{key} in the \emph{section} to \emph{value} only if it is not exists already).
 \emph{success} could be `\textbf{no such LN}', `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}', `\textbf{failed}: reason'.
@@
 
+% chapter technical_description (end)
+
 % \bibliography{grid}

### 2009-02-26 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Revised 2nd section and the first half of the third.

@@
 
-The next service to be presented is the echo service which's task is to return the received message back to the client.
-Furthermore the service provides two additional operations called ``ordinary'' and ``reverse''.
-Depending on the operation chosen by the client, the message will returned unmodified or its letters will be reordered reverse. % without manipulation
-The goal of this chapter is to offer a deeper knowledge about message processing. Furthermore the \textit{WSDL} (Web Services Description Language) shall be introduced.
+The next service to be presented is the echo service.
+Its task is to return the content of the received message back to the client.
+Furthermore, the service provides two additional operations called ``ordinary'' and ``reverse''.
+Depending on the operation chosen by the client, the message will returned unmodified or it is shown with its reversed.
+The goal of this chapter is to offer a deeper knowledge about message processing.
+Also, the \textit{WSDL} (Web Services Description Language) will be introduced, with which the section now starts.
 
@@
 
-Web services can be specified by a meta language called \textit{WSDL} (Web Services Description Language). 
+Web services can be specified (not implemented) by a meta language called \textit{WSDL} (Web Services Description Language). 
 Beside the description of the service, WSDL renders the possibility that services can be discovered by service brokers.
-The discovery of services may be useful for an automatic adaption of a running client i.e. if the endpoint of the service changes or to share the load of demands.
-Furthermore the service specification enables the programmer to implement a suitable client which interfaces the service in a proper way.\\
+The discovery of services may be useful for an automatic adaption of a running client, i.e. if the endpoint of the service changes or to share the load of demands.
+The WSDL service specification enables the programmer to implement a suitable client which interfaces the service in a proper way.
 WSDL describes the structure of the messages, the intended use of the messages, the supported protocols and the endpoint of the service (URL).
-The WSDL file which specifies the echo service presented in this chapter is to be seen in listing~\ref{lst:echo_wsdl}. 
+For that last reason, every instance of a service must be produced a distinct WSDL file for - or - it is automatically prepared from the information that the server has available once it is implemented.
+
+The WSDL file that specifies the here presented echo service is shown in listing~\ref{lst:echo_wsdl}. 
 It is composed of five main elements:
@@
 
-The SOAP Fault Element
+\subsection{The SOAP Fault Element}
 
 The optional SOAP Fault element is used to indicate error messages.
-
-If a Fault element is present, it must appear as a child element of the Body element. A Fault element can only appear once in a SOAP message.
+If a \textit{Fault} element is present, it must appear as a child element of the \textit{Body} element. A \textit{Fault} element can only appear once in a SOAP message.
 
 The SOAP Fault element has the following sub elements:
-Sub Element 	Description
-<faultcode> 	A code for identifying the fault
-<faultstring> 	A human readable explanation of the fault
-<faultactor> 	Information about who caused the fault to happen
-<detail>
+\begin{tabular}{lp{9cm}}
+Sub Element 	& Description \\
+<faultcode> 	& A code for identifying the fault \\
+<faultstring> 	& A human readable explanation of the fault \\
+<faultactor> 	& Information about who caused the fault to happen \\
+<detail> 	& Holds application specific error information related to the Body element\\
+\end{tabular}
 
-Holds application specific error information related to the Body element
 SOAP Fault Codes
@@
 The faultcode values defined below must be used in the faultcode element when describing faults:
-Error 	Description
-VersionMismatch 	Found an invalid namespace for the SOAP Envelope element
-MustUnderstand 	An immediate child element of the Header element, with the mustUnderstand attribute set to "1", was not understood
-Client 	The message was incorrectly formed or contained incorrect information
-Server 	There was a problem with the server so the message could not proceed
+\begin{tabular}{lp{9cm}}
+Error &	Description \\
+VersionMismatch & Found an invalid namespace for the SOAP Envelope element \\
+MustUnderstand 	& n immediate child element of the Header element, with the mustUnderstand attribute set to "1", was not understood \\
+Client 	& he message was incorrectly formed or contained incorrect information \\
+Server 	& here was a problem with the server so the message could not proceed \\
+\end{tabular}
+
+\task{Source of the above two tables?}
 

### 2009-02-26 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Revised 2nd section and the first half of the third.

@@
 
-The first implementation of a Web Service will be a simple time service. 
-In case that the service receives a request of a client, it will respond with a string containing the system time.
+The first implementation of a web service will be a simple time service. 
+When the service receives a request of a client, it will respond with a string containing the system time.
 For simplicity the service will not parse the request but always return the current time.
+It is therefore easier than the example of an Echo web service, that also needs to learn (by interpreting the request)
+about what should be echoed.
 In the following subsections the implementation of the service and the client will be presented.
@@
 
-In Arc the services are implementated as plugins such that they can easily be included or excluded by modifiying the HED configuration file as it was to be seen in Listing~\ref{lst:arched_arcecho_xml}.
+In ARC the services are implementated as plugins such that they can easily be included or excluded by modifiying the HED configuration file as it was to be seen in Listing~\ref{lst:arched_arcecho_xml}.
 Listing~\ref{lst:time_arched_cpp} contains the C++ implementation of the time service plugin.
 The proper class definition of the object \textit{TimeService} is defined in the corresponding header file which is not shown here.
-It inherits from an object called \textit{Service} defined in the Arc library.
+\task{show the header file, or at least the important bits of it - sm}
+It inherits from an object called \textit{Service} which itself is defined in the ARC library.
 The key element for each plugin is the struct \textit{PluginDescriptor} named \textit{PLUGINS\_TABLE\_NAME} to be seen in line~\ref{lst_code:time_cpp_ptn}. 
 %The struct defines basic entities of the service and provides a pointer to the method which is able to create an instance of the time service plugin.
-Within the struct a unique plugin name, the type of the plugin, its version and a pointer to a function which returns the instance of the time service has to be defined.
+Within the struct, a unique plugin name, the type of the plugin, its version and a pointer to a function which returns the instance of the time service has to be defined.
 The function which returns the pointer of the instance is desclared in line~\ref{lst_code:time_cpp_get_service}. 
-The interface demands an object \textit{PluginArgument} as a parameter but because we are writing a plugin which inherits from \textit{Service}, we are able to expect the \textit{PluginArgument} to be a \textit{ServicePluginArgument}.
+The interface demands an object \textit{PluginArgument} as a parameter.
+But since the plugin that is now written inherits from the class \textit{Service}, one can expect the \textit{PluginArgument} to be of a special subclass \textit{ServicePluginArgument}.
+
 %	In the first line of the service the object is downcasted with the operator dynamic_cast
@@
 {../src/services/timeservice/timeservice.cpp}%
-configuration and the message chain of the service.
-The time service uses the namespace \textit{ArcService}.
+configuration and the message chain of the service. The time service uses the namespace \textit{ArcService}.
 The constructor can be found in line~\ref{lst_code:time_cpp_constructor} and the deconstructor in line~\ref{lst_code:time_cpp_deconstructor}.
-In the context of the simple service they don't contribute with any functionality.
+In the context of the simple service they don't contribute any functionality.
 Only the constructor is defining the XML null-namespace for the time service.
-The method \textit{process} which is required by the inherited class \textit{Service} is defined in line~\ref{lst_code:time_cpp_process} and realises the desired service. 
+
+The method \textit{process} is required by the inherited class \textit{Service}.
+It is defined in line~\ref{lst_code:time_cpp_process} and realises the desired service,
+here: the calling of functions provided by system libraries to determine the current time. 
 The parameters \textit{inmsg} and \textit{outmsg} are references to the incoming and outgoing messages.
-Furthermore the method \textit{process} returns the object \textit{MCCStatus} which is representing the result of the service achieved.
-Within the method \textit{process}, the service first determines the current time and than creates a SOAP payload using the object \textit{PayloadSOAP} on line~\ref{lst_code:time_cpp_process_payloadSOAP}.
-The object \textit{PayloadSOAP} provides enough functionality to create a XML message which is conform to the SOAP protocol 
-(The apperance of the generated request and response will be discussed later on page~\pageref{lst:time_client_request}). 
-%
-%
-In the current case the payload consists of two elements nested into one another.
+Furthermore, the method \textit{process} returns the object \textit{MCCStatus} that is representing the result of the service achieved. 
+Within the method \textit{process}, the service first determines the current time and then creates a SOAP payload using the object \textit{PayloadSOAP} on line~\ref{lst_code:time_cpp_process_payloadSOAP}.
+The object \textit{PayloadSOAP} provides enough functionality to create an XML message that is conform to the SOAP protocol.
+The appearance of the generated request and its response will be discussed later on page~\pageref{lst:time_client_request}.
+
+In the current case, the payload consists of two elements, one nested into the other.
 The element \textit{time} holds the element \textit{timeResponse} which again holds the string containing the current time.
 Within the line~\ref{lst_code:time_cpp_process_message_payload} the payload is stored in the message. 
-In case nothing unexpected happened during the message is processed, the function returns with an \textit{MCCStatus} set to \textit{OK} as to be seen in line~\ref{lst_code:time_cpp_return}.\\
+When nothing unexpected happened while the message is processed, the function returns an \textit{MCCStatus} set to \textit{OK} as seen in line~\ref{lst_code:time_cpp_return}.
 
@@
 
-In order to run the service, it has to be compiled into a dynamic library named \textit{libtimeservice.so}.
-The library has be to be copied into the directory \textit{/tmp/arc/tutorial/lib} such that it can be found by the \textit{ModuleManager} of the HED.
-A suitable HED configuration file which engages the time service is to be seen in Listing~\ref{lst:time_arched_xml}.A new path element has been introduced into the element \textit{ModuleMananger} in line~\ref{lst_code:arched_time_moduleManager}.
+To run the service, it needs to be compiled into a dynamic library. Here this is named \textit{libtimeservice.so}, which is arbitrary but needs to be in sync with the configuration file of the HED.
+For the installation, the library needs to be copied to a directory that is accessible for the HED.
+Any would do, but most likely one will select one that is locally mounted, in order to reduce dependencies on external hardware.
+For his example, the installation direction was set to \textit{/tmp/arc/tutorial/lib}.
+It was also specified by the \textit{ModuleManager} of the HED configuration file.
+When this document was prepared, the paths cannot be changed at the runtime at the server.
+
+A suitable HED configuration file which engages the time service is to be seen in Listing~\ref{lst:time_arched_xml}.
+A new \textit{Path} element has been introduced into the element \textit{ModuleMananger} in line~\ref{lst_code:arched_time_moduleManager}.
 It assigns the location of the created dynamic library.
-On line~\ref{lst_code:arched_time_plugin} the library is explict mentioned to be loaded as a plugin.
-Another important change has been done within the \textit{Plexer} element in line~\ref{lst_code:arched_time_plexer_next}.
-A rule has been introduced which redirects the request to the time service, if the path of URL is \textit{time}.\\
+On line~\ref{lst_code:arched_time_plugin} the library is explictly mentioned to be loaded as a plugin.
+Another important change has been done within the \textit{Plexer} element in line~\ref{lst_code:arched_time_plexer_next}:
+a rule has been introduced which redirects the request to the time service, if the path of URL is \textit{time}. 
+Please note that the URL setting is case sensitive.
+
 \lstsetARCHEDXML
@@
 
-To start the service the \textit{arched} command which is shown in Listing~\ref{lst:arched_timeservice_ksh} has to be used.
-It is recommended to examine the log file \textit{/var/log/arched.log} if the HED configuration file has been modified in order to  assert the start was successful
-%To check the start of the HED it is recommended to examine the log file \textit{/var/log/arched.log}.\\
+To start the service, the \textit{arched} command which as shown in Listing~\ref{lst:arched_timeservice_ksh} has to be used.
+It is recommended to examine the log file \textit{/var/log/arched.log} if the HED configuration file has been modified to  assert the start was successful
+
 \lstsetKSH
@@
 
+Up to now we have implemented a server program, the code that runs on a particular machine and waits for other programs that ask it for the time.
+What is still needed is the client.
+If the server spoke the http protocol, then we could use our web browser.
+But it does not - wait, stop - in a way it does, but the server's SOAP envelope renders the reply rather awkward to read.
+\task{someone please check - SM}  The server exists only once.
+But there may be many many clients requesting it in parallel - on the same machine or across the internet.
+The client may be directly controlled by a human (like the web browser) or be some cron job or other software initiating the request.
+
 %\task{\#ifdef HAVE\_CONFIG\_H\#include <config.h>\#endif - habe ich entfernt... macht keinen Sinn, da keine weiteren Macros im Source Code enthalten sind... oder?}
-This section is introducing the implementation of the client whose main task is to interprete the returned message.
-In order to be able to communicate with the service providing daemon the client needs to use an identic protocol stack.
-Likewise to the HED it will also be created dynamically using the MCCs.
+This section is introducing the implementation of the client.
+Its main task is to interprete the returned message.
+In order to be able to communicate with the service providing daemon, the client needs to use an identical protocol stack.
+Likewise to the HED, it will also be created dynamically using the MCCs.
 But as one will see, the creation of the protocol stack on client side is very simple.
-%
-%
-%
+
 The source code of the client is to be seen in Listing~\ref{lst:time_client_cpp} and consists basically of one main function.
-Within the line~\ref{lst_code:time_client_cpp_logger} the object \textit{Logger} gets initialized.
+Within the line~\ref{lst_code:time_client_cpp_logger} the object \textit{Logger} is initialized.
 Due to the configuration of the the \textit{Logger} in the next lines, the messages passed to it will be redirected to the standard error stream.
-%
-For the components of the protocol stack are likewise loaded \textcolor{white}{newline}
+For the components of the protocol stack, which  are likewise loaded \textcolor{white}{newline}
 \lstsetCPP
@@
 	label=lst:time_client_cpp,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	caption={[HED configuration file for the ARC intern echo service. Filename: arcecho\_no\_ssl.xml]
+	\textbf{HED configuration file for the ARC intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
 	]
 {../src/clients/timeclient/timeclient.cpp}
- dynamically, one need to specify the installation directory of ARC in line~\ref{lst_code:time_client_cpp_init}.
-%
-The URL of the service is defined in line~\ref{lst_code:time_client_cpp_url}. It fits to the HED configuration file which was shown in Listing~\ref{lst:time_arched_xml}. 
+ dynamically, one needs to specify the installation directory of ARC in line~\ref{lst_code:time_client_cpp_init}.
+
+The URL of the service is defined in line~\ref{lst_code:time_client_cpp_url}.
+It fits to the HED configuration file which was shown in Listing~\ref{lst:time_arched_xml}.
+Actually, if available, in many cases the client may use the configuration file of the server. \task{seen for the arcecho, but is that true? sm}
 The daemon is listing on port 60000 and is running on the localhost. The service itself uses HTTP and can be reached by using the path \textit{time}.
-Line~\ref{lst_code:time_client_cpp_config} loads the standard config file which is usually almost empty.
-As a result, no additional security procedures will later be linked into the client sided stack.
+Line~\ref{lst_code:time_client_cpp_config} loads the standard configuration file, which is usually almost empty.
+As a result, no additional security procedures will later be linked into the client-sided stack. 
 The protocol stack is created in line~\ref{lst_code:time_client_cpp_client}. 
 Thanks to the service URL the information for the configuration is complete.
+
 The constructor of the class \textit{ClientSOAP} creates a SOAP client which uses HTTP and addresses the port 60000 on localhost.
-If i.e.
-HTTPS would have been used, the stack would have been extended by an additional TLS protocol.
-Once the client stack is ready, the request to the service can be created which is done in line~\ref{lst_code:time_client_cpp_message_reqeust}.
-Even the apperance of the message is unimportant to the service, the message is composed of two convoluted elements \textit{time} and \textit{timeReqest}.
+If HTTPS would have been used, the stack would have been extended by an additional TLS protocol.
+Once the client stack is ready, the request to the service can be created.
+This is done in line~\ref{lst_code:time_client_cpp_message_reqeust}.
+Even the appearance of the message is unimportant to the service, the message is composed of two convoluted elements \textit{time} and \textit{timeReqest}.
 %
@@
 The object \textit{MCCStatus} which is returned by the method indicates the success of the processing.
-After verifying the response, the message will passed to the standard out in line~\ref{lst_code:time_client_cpp_message_answer}.\\
+After verifying the response, the message will passed to the standard out in line~\ref{lst_code:time_client_cpp_message_answer}.
 
 The client has to be compilied to an ordinary binary so it can directly be executed from the command line.
-Assure the service is running and use the client as to be seen in Listing~\ref{lst:time_client_ksh}.
+Assure that the service is running and use the client as to be seen in Listing~\ref{lst:time_client_ksh}.
 \lstsetKSH
@@
 \end{lstlisting}
-The structure of the messages on SOAP level which are transmitted are shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
+The messages are transmitted in SOAP format and thus reach the SOAP level of the server. They shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
 The enclosing element is \textit{Envelope} which may contain an element \textit{Header} and must contain an element \textit{Body}.

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Enhanced the content of the TLS echo chapter

@@
 
-Symmetric and asymmetric encryption are the two main classes of cryptography algorithms.
-In case of a symmetric encryption both participants are holding the same secret key. 
+Symmetric and asymmetric encryption are the two main classes of cryptographic algorithms.
+In case of symmetric encryption both participants are holding the same secret key to encrypt and decrypt a plain text. The principle is to reverse the process of encryption for decrypting the text.
 Symmetric encryption can be ranked as being very secure but its downside is that the key distribution is very difficult in practice.
@@
 \end{figure}
+\task{Maybe the term service has to be replaced by server}
 
 At startup of the communication process A (Alice) who desires a secure connection transmits her public key to B (Bob).
-Bob is now able to encrypt the messages such that only A is able to decrypt them.
+Bob is now able to encrypt his messages such that only Alice is able to decrypt them.
 As one can see this first approach already provides confidentiality and integrity of messages.
@@
 Non-repudiation, authentication and due to that access control are not given because the identity of Alice is not proven.
-In order to be able to check identities so called certificates must be introduced.
+In order to be able to check identities so called certificates have to be introduced.
 Certificates render the possibility to validate the identity of its owner.
-They are signed by a certificate authority (CA) which's identity can be resolved to another CA or which is  pre-configured to be trusted.
-A certificate is composed of its owner's public key, his identity, the name of the signing CA and a signiture which is a hash-value of the certificate encrypted with the private key of the CA. 
+They are signed by a certificate authority (CA) which's identity can be verified by another CA or which is  pre-configured to be trusted.
+A certificate is composed of the identity of the owner, his public key, the name of the CA who signed the certificate and the signature of the CA which is a hash-value of the certificate encrypted with the private key of the CA.
+Furthermore the algorithms used to create the certificate are added.
 If a certificate can be resolved to a CA with is pre-configured to be trusted, the identity of the certificate owner is considered as to be proven. 
@@
 To obtain a certificate a certificate request has to be submitted to a CA.
-After the CA has verified the identity of the requesting person and the requested certificate will be signed.
+The CA verifies the identity of the requesting person and the requested certificate will be signed.
 Figure~\ref{fig:certificate_request} shows the procedure of gaining a certificate.
@@
 In the given example Alice only trusts the CA 2 but not CA 1.
-In order to resolve the identity, Alice establishes a connection to CA 1 and submits another random number $n$ as a challenge.
+To resolve the identity of CA 1, Alice establishes a connection to CA 1 and submits another random number $n$ as a challenge.
 The CA 1 encrypts the challange with its private key and send the result along with its certificate back to Alice.
@@
 If Bob also wants to be sure how he's talking to the same procedure has to be done for the identity of Alice.
-The TLS mode in in which both endpoints are ensuring themself to whom they are talking to is called bilateral connection.
-Due to the client is aware of the service authentity, the confidential treatment of his data is given such that it won't be revealed to a third party (service sided confidentiality).
-Moreover the access to the service can be controled (access control) and the results are only passed back to the right person (client sided confidentiality).
+The TLS mode in in which both endpoints are ensuring themself to whom they are talking to is called bilateral connection mode.\\
+
+In case the client is aware of the service authentity, the confidential treatment of his data is given such that one can be sure the data won't be revealed to a third party (service sided confidentiality).
+On the other side, if the service is aware of the client authentity, the access to the service can be controlled (access control) and the service can verify that the results will only be passed back to the right person (client sided confidentiality).
 As one can see confidentiality, authentication, integrity and non-repudiation are now given.\\
@@
 They are scaling very good with the amount of users because CAs can create new sub-CAs such that the load is balanced (certificates of a CA may be cached).
-Furthermore no critical information is exchanged which may gets intercepted - in particular no secret key is shared.
+Furthermore no critical information is exchanged during the establishment of a connection (in particular no secret key is shared which may get intercepted),
 Instead a set of pre-configured CAs which will be in general provided by the operation system is needed.
-In respect of ARC it is obvious, that the access to the grid should be under control of its administrator.
-To do so, the administrator have to build up his own CA to be in charge of creating certificates for a set of users.
-The created CA have to be shipped into the directory of pre-configured CAs.
-(Within the source code there are two shell scripts to create a CA and related certificates). Due to there's a bunch of tools for certificates (like \textit{openssl}) it is easy to maintain them.
+In respect of ARC it is obvious, that the access to the grid shall be under control of its administrator.
+To do so, the administrator have to build up his own CA such that he is in charge of creating certificates for a set of users.
+The identity of the created CA have to be shipped into the directory of pre-configured CAs.
+(Within the source code there are two shell scripts to create a CA and related certificates). Due to there's a bunch of tools for certificates (like \textit{openssl}) they are easy to maintain.\\
+ % Root CAs are self-signed ... is that interesting?
 
-o Infrastructure is given and easily to maintain
+A CA which controls the access to a grid can be considered as a virtual organsiation (VO).
+It may involve organisations close to the leading organisation as well as structure which are not formally associated with it.
+In most cases the basic idea is to share ressources.
 
-o In general the pre-configured organistion take money to sign certifcates, but they can also be created locally.
+\clearpage
+\section{Service}
 
-o VOMs are realisable
+In order to create a secure bilateral connection we hence need two certificates: a client certificate and a service certificate.
+The CA for the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key. Altogether six files are needed:
+\begin{itemize}
+ \item clientCA   - The CA which guarantees for the identity of the client
+ \item clientCERT - The certificate of the client.
+ \item clientKEY  - Secret key of the client to read messages, encrypted by the public key in the certificate!
 
-ARC:
-
-TLS also supports the more secure bilateral connection mode (typically used in enterprise applications), in which both ends of the "conversation" can be ensured with whom they are communicating
-
-o What do we need to get a service running
-
-clientCA   - Authority guarantee for identity
-clientCERT - Certificate given
-clientKEY  - Secret key of the client to read messages, encrypted by the public key in the certificate!
-
-serverCA   - Authority guarantee for cert. CA is known by the server. If not, the encrypted cert can be read be its CA.
-serverCERT - Certificate given by the CA. Contains a snippet which can be read by the Public key of the CA.
-serverKEY  - Secret key of the client to read messages, encrypted by the public key in the certificate!
+ \item serverCA   - The CA which guarantees for the identity of the server
+ \item serverCERT - The certificate of the server.
+ \item serverKEY  - Secret key of the client to read messages, encrypted by the public key in the certificate!
+\end{itemize}
 
@@
 %Authorization:  The server on the other side takes an interest in limitate its ressources to a close circle of users. Above all no %foreign clients shall be able to use the services.\\
-
-\clearpage
-\section{Service}
 

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Series of changes.

@@
 
-In the current section, the implementation of the service defined in the WSDL file of the previous section will be presented.
+This section presents the implementation of the service defined in the WSDL file of the previous section.
 The source code of the C++ file is shown in listing~\ref{lst:echo_service_cpp}.
-Again the header file will be set aside for it contains only redundant information.
-It can be reread in the source directory accompanying the tutorial.
-Within the source code of echo service basically three changes has been done.
-The first change concerns the constructor and is to be found in line~\ref{lst_code:echo_cpp_prefix}.
+Again, the header file will be set aside for it contains only redundant information.
+It can be reread in the source directory that is accompanying the tutorial.
+For the source code of echo service basically three changes are required for on the code for the timeservice webservice of the previous chapter.
+The first change concerns the constructor and is found at line~\ref{lst_code:echo_cpp_prefix}.
 Two strings will be extracted of the HED configuration file which are later used to envelope the message before it is returned.
-Furthermore a new method \textit{makeFault} in line~\ref{lst_code:echo_cpp_makeFault} has been introduced which is able to create fault messages and is called by the method \textit{process}.
-Within that method two new code fragments are now extracting and analyzing the incoming message, to be seen in line~\ref{lst_code:echo_cpp_extracting} and \ref{lst_code:echo_cpp_analyzing}.
+Furthermore, a new method \textit{makeFault} in line~\ref{lst_code:echo_cpp_makeFault} has been introduced which is able to create fault messages and is called by the method \textit{process}.
+Within that method two new code fragments are now extracting and analyzing the incoming message, to be seen at lines~\ref{lst_code:echo_cpp_extracting} and \ref{lst_code:echo_cpp_analyzing}.
 In case something unexpected happens both will create a SOAP fault message.\\
@@
 
-Again the HED is configured by a XML file and have to be lunched as usual.
+Again, the HED is configured by an XML file and is launched as introduced earlier.
 The HED configuration file is shown in Listing~\ref{lst:echo_hed_configuration_xml}.
-The only change concerns the lines~\ref{lst_code:echo_hed_configuration_xml_pre} and \ref{lst_code:echo_hed_configuration_xml_pre} which are specifing the \textit{prefix} and \textit{suffix}.
-The parameters will be extracted in the constructor of the service and provide a good example how settings can be passed from the HED configuration file towards the service.
+The only change concerns the lines~\ref{lst_code:echo_hed_configuration_xml_pre} and \ref{lst_code:echo_hed_configuration_xml_pre} which specify the \textit{prefix} and \textit{suffix}.
+The parameters will be extracted in the constructor of the service and explain nicely how settings can be passed from the HED configuration file towards the service.
 
@@
 
-Likewise the client has become also more complex. The parameters are now passed by the command line, to be seen in the lines followed by line~\ref{lst_code:echo_client_cpp_arguments}.
-The creation of the SOAP client is unchanged. The request is created in lines subsequent line~\ref{lst_code:echo_client_cpp_request}.
+With the interface of the server gaining complexity, so is the code for the client - for two reasons.
+Firstly, because the additional arguments need to be filled.
+Secondly, because of the extra effort to interact with the (here human) user to collect the data to forward to the echo service.
+The parameters are now passed by the command line, to be seen in the lines followed by line~\ref{lst_code:echo_client_cpp_arguments}.
+That code is independent from the ARC library.
+The creation of the SOAP client is unchanged.
+The request is created in lines subsequent line~\ref{lst_code:echo_client_cpp_request}.
 It is processed in the code after line~\ref{lst_code:echo_client_cpp_response} in which also the response is received.
-The checks of the response have expanded by an addition condition which tests for a fault message, see line~\ref{lst_code:echo_client_cpp_fault}.
+The checks of the response have expanded by an additional condition which tests for a fault message, see line~\ref{lst_code:echo_client_cpp_fault}.
 
@@
 	label=lst:arcecho_arched_xml,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	caption={[HED configuration file for the ARC echo service. Filename: arcecho\_no\_ssl.xml]
+	\textbf{HED configuration file for the ARC echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
 	]
@@
 
-A few examples of the exchanged messages are shown in the Listings~\ref{lst:echoservice_request}, \ref{lst:echoservice_response} and \ref{lst:echoservice_fault}.
+A few examples for the exchanged messages are shown in the Listings~\ref{lst:echoservice_request}, \ref{lst:echoservice_response} and \ref{lst:echoservice_fault}.
 Listing~\ref{lst:echoservice_request} illustrates a request message using the operation \textit{reverse}.
 Listing~\ref{lst:echoservice_response} illustrates the corresponding response message.
-In case the operation requested by the client is unknown, the server replies with the fault message to be seen in Listing~\ref{lst:echoservice_fault}.
+When the operation requested by the client is unknown, then the server replies with the fault message shown in Listing~\ref{lst:echoservice_fault}.
 

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Series of changes.

@@
 
-The web services which were presented in the previous chapters have one mayor drawback: the communication between the two endpoints is unsecure.
-Due to that reason the current and the next chapter will discuss how to create a secure web service.
-In general security in a network can be subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{verfiy citation}
+The web services of the previous chapters have one mayor drawback: the communication between the two endpoints is insecure.
+For that reason the current and the next chapter will discuss how to create a secure web service.
+In general, security in a network can be subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{verfiy citation}
 \begin{itemize}
@@
 % Nach Sicherheit in verteilten Systemen, ITM Lübeck
-Ideally all six categories should have been achieved by such an application (though the last category, availability, is difficult to realise).
+Ideally, all six categories should have been achieved by such an application (though the last category, availability, is difficult to realise).
 In the present chapter, the security will be put into practice by extending the protocol stack with an additional security layer called TLS (Transport Layer Security).
-Due to the location after the TCP and the HTTP layer, the layer is only capable to limitate the access to the grid to a certain set of users.
-Nevertheless it provides the security categories: Confidentiality, authentication, integrity, non-repudiation and access control.\\
+Due to the location after the TCP and the HTTP layer, the layer is only capable to limit the access to the grid to a certain set of users.\task{What do you mean here? SM}
+Nevertheless, it provides the security categories: confidentiality, authentication, integrity, non-repudiation and access control.
 
@@
 The TLS is a protocol which provides a secure connection between two endpoints.
-It is located directly upon the TCP/IP layers and uses typically an asymmetric encryption to establish the connection (alternatively a symmetric pre-shared key may be used).
-After the communication has successfully established, both participants are switching to another encryption method which is based on a new negotiated symmetric key.\\
+It is based directly on the TCP/IP layers and uses typically an asymmetric encryption to establish the connection (alternatively a symmetric pre-shared key may be used).
+After the communication has been successfully established, both participants are switching to another encryption method which is based on a new negotiated symmetric key.
 %    1. Peer negotiation for algorithm support
@@
 Symmetric and asymmetric encryption are the two main classes of cryptographic algorithms.
-In case of symmetric encryption both participants are holding the same secret key to encrypt and decrypt a plain text. The principle is to reverse the process of encryption for decrypting the text.
+In case of symmetric encryption, both participants are holding the same secret key to encrypt and decrypt a plain text.
+The principle is to reverse the process of encryption for decrypting the text.
 Symmetric encryption can be ranked as being very secure but its downside is that the key distribution is very difficult in practice.
@@
 While the symmetric encryption utilise one shared secret, the asymmetric ecryption is based on two keys: a private key and public key.
-As the names imply the private key is kept secret and is never shared while the public key is available to everyone.
-Messages encrypted with the public key can only be decrypted with the private key and the other way round, messages encrypted with the private key can only be decrypted with the public key.
-The Figure~\ref{fig:asymmetric_encryption} illustrates the circumstances of the case.\\
+As the names imply, the private key is kept secret and is never shared while the public key is available to everyone.
+Messages encrypted with the public key can only be decrypted with the private key and conversely, messages encrypted with the private key can only be decrypted with the public key. 
+The Figure~\ref{fig:asymmetric_encryption} illustrates the circumstances of the case.
 \begin{figure}[htb]
@@
 
-At startup of the communication process A (Alice) who desires a secure connection transmits her public key to B (Bob).
+At startup of the communication process, A (Alice) who desires a secure connection transmits her public key to B (Bob).
 Bob is now able to encrypt his messages such that only Alice is able to decrypt them.
-As one can see this first approach already provides confidentiality and integrity of messages.
-%
-Non-repudiation, authentication and due to that access control are not given because the identity of Alice is not proven.
-In order to be able to check identities so called certificates have to be introduced.
+As one can see, this first approach already provides confidentiality and integrity of messages.
+
+Neither Non-repudiation, authentication and consequently nor access control are established because the identity of Alice is not proven.
+In order to be able to check identities, so called certificates have been introduced.
 Certificates render the possibility to validate the identity of its owner.
-They are signed by a certificate authority (CA) which's identity can be verified by another CA or which is  pre-configured to be trusted.
+They are signed by a certificate authority (CA), the identity of which can be verified by another CA or that is pre-configured to be trusted.
+
 A certificate is composed of the identity of the owner, his public key, the name of the CA who signed the certificate and the signature of the CA which is a hash-value of the certificate encrypted with the private key of the CA.
-Furthermore the algorithms used to create the certificate are added.
-If a certificate can be resolved to a CA with is pre-configured to be trusted, the identity of the certificate owner is considered as to be proven. 
-In general a set of trusted CAs is already defined in the operation system.
-To obtain a certificate a certificate request has to be submitted to a CA.
+Furthermore, the algorithms used to create the certificate are added.
+If a certificate can be resolved to a trusted CA, the identity of the certificate owner is considered to be proven. 
+In general, a set of trusted CAs is already defined in the operation system.
+To obtain a certificate, a certificate request has to be submitted to a CA.
 The CA verifies the identity of the requesting person and the requested certificate will be signed.
 Figure~\ref{fig:certificate_request} shows the procedure of gaining a certificate.
+For instructions on the actual commands on the UNIX shell, you may follow instructions on \url{http://ca.nordugrid.org}.
 \begin{figure}[htb]
@@
 The public key together with a textual description of the identity depict the request which will be transmitted to a CA.
-After the CA successfully varified the validity of the request, the identity of the CA and the name of used the encryption algorithms are added.
-The collected data will represent the plain text of the certificate. In order to ensure the integrity of the text, 
-a hash value will be created.
-(A fixed-sized short string which is generated by an algorithm based on an arbitrary long given text.
+After the CA successfully verified the validity of the request, the identity of the CA and the name of the used encryption algorithms are added.
+The collected data will represent the plain text of the certificate.
+In order to ensure the integrity of the text, a hash value will be created.
+This is a fixed-sized short string which is generated by an algorithm based on an arbitrary long given text.
 The algorithm is designed such that a small change in the text will cause the hash value to change almost like a random function.
-Thus, the hash value can be used a fingerprint of the text.) The CA signs the plain certificate by encrypting its hash value with its private key.
+Thus, the hash value can be used a fingerprint of the text.
+The CA signs the plain certificate by encrypting its hash value with its private key.
 Finally the signed certifcate will be returned to the requester.
-%
-%
-If an outside person wants to verify the certificate owner, the encrypted hash value has to be decrypted by the public key of the CA and a hash value has to be created with the same algorithm in order to compare them.\\
+If an outside person wants to verify the certificate owner, the encrypted hash value has to be decrypted by the public key of the CA and a hash value has to be created with the same algorithm in order to compare them.
 
 Figure~\ref{fig:verification_of_certificates} shows a usecase in which Alice wants to resolve the identity of Bob.
-Alice submitts her request to Bob along with a challange consisting of a random number $n$.
+Alice submits her request to Bob along with a challange consisting of a random number $n$. 
 The random number ensures that it is not possible to repeat a message (integrity).
-Bob encrypts the number with his private key and returns it together with his certificate. Alice uses the public key of Bob to decrypt the challange. 
+Bob encrypts the number with his private key and returns it together with his certificate.
+Alice uses the public key of Bob to decrypt the challenge. 
 If the decrypted number is equal to the transmitted number, Alice can reason that the identity of Bob is to be trusted in case she is trusting the CA which has signed Bobs certificate.
-%
-In the given example Alice only trusts the CA 2 but not CA 1.
+
+In the given example, Alice only trusts the CA 2 but not CA 1.
 To resolve the identity of CA 1, Alice establishes a connection to CA 1 and submits another random number $n$ as a challenge.
-The CA 1 encrypts the challange with its private key and send the result along with its certificate back to Alice.
-Alice succefully validates the certificate of CA 1.
-Due to the certificate was signed by CA 2 which is pre-configured to be trusted she can follow to trust CA 1 and in conclusion the identity of B.
+The CA 1 encrypts the challenge with its private key and sends the result along with its certificate back to Alice.
+Alice successfully validates the certificate of CA 1.
+Since the certificate was signed by CA 2 which is pre-configured to be trusted, she als trusts CA 1 and in conclusion the identity of B.
 In summary, Alice knows she is talking to Bob.
-If Bob also wants to be sure how he's talking to the same procedure has to be done for the identity of Alice.
-The TLS mode in in which both endpoints are ensuring themself to whom they are talking to is called bilateral connection mode.\\
+If Bob also wants to be sure who he's talking to, then the same procedure has to be done to confirm the identity of Alice.
+The TLS mode in which both endpoints are ensuring themselves to whom they are talking to is called \textit{bilateral connection mode}.
 
-In case the client is aware of the service authentity, the confidential treatment of his data is given such that one can be sure the data won't be revealed to a third party (service sided confidentiality).
-On the other side, if the service is aware of the client authentity, the access to the service can be controlled (access control) and the service can verify that the results will only be passed back to the right person (client sided confidentiality).
-As one can see confidentiality, authentication, integrity and non-repudiation are now given.\\
+When the client is aware of the service authentity, the confidential treatment of his data is given such that one can be sure the data won't be revealed to a third party (service-sided confidentiality).
+On the other side, if the service is aware of the client authentity, the access to the service can be controlled (access control) and the service can verify that the results will only be passed back to the right person (client-sided confidentiality).
+As one can see confidentiality, authentication, integrity and non-repudiation are now given.
+
 \begin{figure}[htb]
@@
 They are scaling very good with the amount of users because CAs can create new sub-CAs such that the load is balanced (certificates of a CA may be cached).
-Furthermore no critical information is exchanged during the establishment of a connection (in particular no secret key is shared which may get intercepted),
-Instead a set of pre-configured CAs which will be in general provided by the operation system is needed.
-In respect of ARC it is obvious, that the access to the grid shall be under control of its administrator.
-To do so, the administrator have to build up his own CA such that he is in charge of creating certificates for a set of users.
-The identity of the created CA have to be shipped into the directory of pre-configured CAs.
-(Within the source code there are two shell scripts to create a CA and related certificates). Due to there's a bunch of tools for certificates (like \textit{openssl}) they are easy to maintain.\\
- % Root CAs are self-signed ... is that interesting?
+Furthermore no critical information is exchanged during the establishment of a connection (in particular no secret key is shared which may get intercepted).
+Instead, a set of pre-configured CAs which will be in general provided by the operation system is needed.
+In respect of ARC it is obvious that the access to the grid shall be under control of its administrator.
+To do so, the administrators have to establish their own CA such which is in charge of creating certificates for a set of users.
+Many universities have prepared local certificate authorities.
+And every user can establish his own.
+
+The identity of the created CA have to be added to the directory of pre-configured CAs.
+The ARC source code comes with two shell scripts to create a CA and related certificates.
+Several publicly availble tools for certificate handling haven been prepared, e.g. (like \textit{openssl}),
+which allow the certificate handling also for the advanced UNIX users.
+Certificates in general, however, have found entry to regular computer users, for instance with the FernUniversity Hagen (https://ca.fernuni-hagen.de/) who use that technology for their remote students to hand in exercises.
 
@@
 It may involve organisations close to the leading organisation as well as structure which are not formally associated with it.
-In most cases the basic idea is to share ressources.
+In most cases the basic idea is to share ressources. \task{this para will need some overhaul - SM}
 
@@
 In order to create a secure bilateral connection we hence need two certificates: a client certificate and a service certificate.
-The CA for the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key. Altogether six files are needed:
-\begin{itemize}
- \item clientCA   - The CA which guarantees for the identity of the client
- \item clientCERT - The certificate of the client.
- \item clientKEY  - Secret key of the client to read messages, encrypted by the public key in the certificate!
+The CA for the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key. 
+Altogether six files are needed:
+\begin{tabular}{l@{-}p{9cm}}
+clientCA   & The CA which guarantees for the identity of the client \\
+clientCERT & The certificate of the client. \\
+clientKEY  & Secret key of the client to read messages, encrypted by the public key in the certificate! \\
+\\
+serverCA   & The CA which guarantees for the identity of the server \\
+serverCERT & The certificate of the server. \\
+serverKEY  & Secret key of the client to read messages, encrypted by the public key in the certificate! \\
+\end{tabular}
 
- \item serverCA   - The CA which guarantees for the identity of the server
- \item serverCERT - The certificate of the server.
- \item serverKEY  - Secret key of the client to read messages, encrypted by the public key in the certificate!
-\end{itemize}
-
-Server has got have:
-Its  serverCERT serverKEY and some kind of root CA knowing the client
-Client has got to have
-Its  clientCERT clientKEY and some kind of root CA knowing the server
+The server needs: its  serverCERT and serverKEY and some kind of root CA knowing the client.
+Client needs its clientCERT and clientKEY and some kind of root CA knowing the server.
 
@@
 
-WARNING: If the programm fails, check if one of your certifacte is expired. If so, create new one using the script files in the directory 'certFactory'
+WARNING: If the program fails, check if one of your certificates has expired.
+If so, create new one using the script files in the directory 'certFactory'.
 
-For HED is build up to  % entsrprechend einer Mod Structur
-be modular, the security can easily be extend to our given example.
+Chalked up to the modularity of HED, the security can easily be extend to our given example.
 Each Message Chain Component (MCC) or service has a common interface for implementing various pluggable components (plug-ins) called SecHandler.
 The SecHandler components provide a method for processing messages traveling through Message Chains of the HED.
-Each MCC or Service usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called “incoming” and “outgoing” respectively.
+
+Each MCC or Service usually implements two queues of SecHandlers – one for incoming messages and one for outgoing, which are called “incoming” and “outgoing”, respectively.
 All SecHandler components attached to the queue are executed sequentially. 
 If any of them fails, message processing fails as well. 
-Each SecHandler is configured inside the arched configuration file used for configuring whole chain of MCCs 
+Each SecHandler is configured inside the arched configuration file that is used for configuring whole chain of MCCs 
 Some of the currently implemented SecHandler components make use of pluggable and configurable sub-modules called Policy Decision Point (PDP).
-PDP can process ARC specific Request and Policy documents which are as well written in XML format~\cite{QIANG_2008}..
+PDPs can process ARC specific Request and Policy documents which are also written in XML format~\cite{QIANG_2008}..
 

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_appendix/appendix.tex
Express checkout... Finished the basic TLS chapter content and started writting about the Sec echo client...

--- 2009-02-25 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
@@
- 
-\chapter{Persistent Webservice}
-
-export ARC\_BARTENDER\_URL
-
-    key\_file = os.environ.get('ARC\_KEY\_FILE', None)
-    cert\_file = os.environ.get('ARC\_CERT\_FILE', None)
-    proxy\_file = os.environ.get('ARC\_PROXY\_FILE', None)
-    ca\_file = os.environ.get('ARC\_CA\_FILE', None)
-    ca\_dir = os.environ.get('ARC\_CA\_DIR', None)
-
-export ARC\_KEY\_FILE=/local/glodek/clientKey.pem
-export ARC\_CA\_FILE=/local/glodek/clientCA.pem
-export ARC\_CERT\_FILE=/local/glodek/clientCert.pem
-
-does the daemon runs the service single threaded
-
-race conditions
-\begin{itemize}
- \item Transaktionen
- \item nur lesen
- \item jeder prozess schreibt in seine eigene Datei
- \item Gibt es einen     // andere Zugriffe abblocken
-    flock (\$fp, LOCK\_EX);  (atomare befehle)
- 	wie bei php??
- \item Peterson Algorithmus
- \item Was passiert beim Ausfall (Time out - Transaktion zurück nehmen)
-\end{itemize}
-
 \chapter{Appendix}

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
Express checkout... Finished the basic TLS chapter content and started writting about the Sec echo client...

--- 2009-02-24 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
@@
- 
 \chapter{Persistent Webservice}
 
+\task{Some more comments: the <LogLevel> thing is not working at service level, you can just remove those lines - proposed by Zsombor}
 export ARC\_BARTENDER\_URL
@@
 \end{itemize}
-
-\section{Schlusswort}
-
-Other tutorials are:\\
-\\
-%Oxana reported the status of OGF tutorials:
-\begin{itemize}
- \item first tutorial will be given by Oxana and will deal with how to add the 
-resources to Nordugrid, .... (will be related mostly to production ARC)
- \item second tutorial will be given by Gabor and he will show the 
-interoperability of ARC (video demo will be shown). A-REX - Unicore 
-interoperability.
- \item there will be also a tutorial related to ETICS system where the 
-possibility of ETICS system to submit a job to ARC and Unicore will be 
-presented. The ETICS team has not contacted so far.
-\end{itemize}

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
Started outline for the TLS echo service
Express checkout... Finished the basic TLS chapter content and started writting about the Sec echo client...

@@
 \chapter{SEC echoservice}
+
+The TLS echo service, presented in the previous chapter enables the adminstrator the control the access to the grid.
+Due to a certificate owner may be a member of a large VO it is sometimes recommended to have a more fine grained security within the daemon.
+The goal to be reached in this chapter is to enhance the possibilities to limitate the access to particular users and to narrow their rights on the workstation.\\
+
+The security framework of ARC is implemented based on the same concept of modularity as HED. 
+To extend the security of a MCC one only has to modify the service configuration file.
+This is to be done by including a \textit{SecHandler} element into a component or service which again has to contain a \textit{pdp} element.
+A small example of a chain component is shown in Listing~\ref{lst:sec_service_component_xml}.
+The \textit{SecHandler} is able to intervene the message flow
+by: manipulating the message content, the attributes or just disrupt the message. 
+% Attributes done by: identity.map, delegation.collector
+% manipulating content: usernametoken.handler
+% disrupt by: arc.authz
+\lstsetJUSTXML
+\begin{lstlisting}[
+label=lst:sec_service_component_xml,float=htb,
+caption={[Example of a SecHandler within the component \textit{http.service}.]
+         \textbf{Example of a SecHandler within the component \textit{http.service}}}]
+<Component name="http.service" id="http">
+	<next id="soap">POST</next>
+	<SecHandler name="arc.authz" id="auth" event="incoming">
+		<PDP name="simplelist.pdp" location="./clientlist.txt"/>
+	</SecHandler>
+</Component>
+\end{lstlisting}
+Each MCC or Service usually implement two queues of SecHandlers --- one for incoming messages and one for outgoing called
+``incoming'' and ``outgoing'' respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well~\cite{QIANG_2008}.
+In order to decide which message shall be modified by the \textit{SecHandler} so called PDP (Policy Decision Point) are used.
+They come in a bunch of flavours such that the variabilty of security which can be realised is big.\\
+
+In this tutorial only a small set of \textit{SecHandler}s and \textit{PDP} will be introduced.
+For further reading it is recommended to consult the document ``Security framework of ARC1'' written by Weizhong Qiang and Alexandr Konstantinov.
+The present chapter is closly related to it.\\
+
+  \begin{table}[htb]
+  \centering
+  \caption{List of SecHandlers}
+  \begin{tabular*}{\textwidth}[t]{lp{12cm}}
+	\hline
+ 	\textbf{identity.map}          & Maps the identity of the user to an identity of the workstation.\\
+ 	\textbf{arc.authz}             & Restricts the access to the next layer.\\
+ 	\textbf{delegation.collector}  & Collects the information for the delegation policy (Related to \textit{delegate.pdp}). \\
+ 	\textbf{usernametoken.handler} & Generates the WS-Security username-token and adds it into the SOAP header of outgoing
+					 messages. In case of incoming messages the SecHandler extracts the WS-Security username-token out of it. \task{Figure out what the username-token does...}\\
+	\hline
+  \end{tabular*}
+  \end{table}
+
+  \begin{table}[htb]
+  \centering
+  \caption{List of PDPs}
+  \begin{tabular*}{\textwidth}[t]{lp{12cm}}
+	\hline
+ 	\textbf{allow.pdp}           & bla \\
+ 	\textbf{simplelist.pdp}      & blub.\\
+ 	\textbf{arc.pdp}             &\\
+ 	\textbf{delegate.pdp}        & \task{Explain that well}\\
+ 	\textbf{pdpservice.invoker}  &\\
+	\hline
+  \end{tabular*}
+  \end{table}
+
+In order to realise security within an MCC 
+\begin{verbatim}
+SecHandler 
+==========
+	PDP
+	===	
+		elements/attributes required
+		============================
+identity.map						--- maps the identity of users to identity within the local workstation
+	allow.pdp					--- means concering all identities
+							  (I suppose it can be used in every MCC)
+		<LocalName>nobody</LocalName>		  (Due to the SecHandler within the PDP the local name has to be spechified)
+		???LocalList???
+		???LocalSimplePool???
+	simplelist.pdp					--- maps only persons listed within the file
+								(TLS identity is used here, thus it can be used only after the TLS)
+		<PDP name="simplelist.pdp" location="./clientlist.txt"/>
+	arc.pdp						--- Complexer policy to permit or deny someone. 
+		<PolicyStore>					(Is it possible to add parameters for the SOAP service here?)
+		<Location type="file">policy.xml</Location>
+		</PolicyStore>
+	[...]
+
+arc.authz					--- restricts access to next layer
+	arc.pdp					--- using policy.xml
+	simplelist.pdp				--- using a file with listed identities
+	delegate.pdp				--- like arc.pdp but poicy is parsed from a remote proxy
+	pdpservice.invoker			--- Is a client which can be used to invoke the PDP Service which 
+							implements the same functionality as Arc PDP, except that the
+							evaluation request and response are carried by SOAP
+							message. The benefit of implementing PDP Service and PDP Service 
+							Invoker is that the policy evaluation engine can be accessed remotely
+							and maintained centrally.
+							- wasn't that also possible with the delegate.pdp??
+
+delegation.collector 				--- Delegation SecHandler
+						--- is responsible for collecting the delegation policy information and 
+						putting this policy into message’s security attribute for the 
+						usage of other components, such as delegation.pdp
+usernametoken.handler				--- The task is to generate the WS-Security Username-Token and add it into
+						header of SOAP message which is the payload of outgoing message. 
+						It can also extract the WS-Security Username-Token from the header of 
+						SOAP message which is the payload of incoming message.
+						why should one want to do that???
+
+\end{verbatim}
+
+- have fine grained but complex authorization system (policy evaluator) which uses information stored on the server (arc.pdp) or an an proxy (delegation.pdp) or which is evaluated using a SOAP based service such that the request can be evaluated remotly (pdpservce.invoker).
+- have a simple list which matches identities (simplelist.pdp)
+- map users to an identity within the local workstation
+
+have a smaller set of users,
+have a policy which may map certain users to certain settings
+
+For HED is build up to  % entsrprechend einer Mod Structur
+be modular, the security can easily be extend to our given example.
+Each Message Chain Component (MCC) or service has a common interface for implementing various pluggable components (plug-ins) called SecHandler.
+The SecHandler components provide a method for processing messages traveling through Message Chains of the HED.
+Each MCC or Service usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called “incoming” and “outgoing” respectively.
+All SecHandler components attached to the queue are executed sequentially.
+If any of them fails, message processing fails as well. 
+Each SecHandler is configured inside the arched configuration file used for configuring whole chain of MCCs 
+Some of the currently implemented SecHandler components make use of pluggable and configurable sub-modules called Policy Decision Point (PDP).
+PDP can process ARC specific Request and Policy documents which are as well written in XML format~\cite{QIANG_2008}..
 
@@
 And what are PDP's?
-
-Currently, the ARC1 comes in severalfour policy decision point implementations:
-\begin{itemize}
- \item arc.pdp – Arc PDP\\
-The Arc PDP will organize the security attributes into the ARC specific authorization request, call the
-policy evaluator to evaluate the request against the policy (which is in ARC specific format) repository,
-and get back the evaluation result. See paragraph 3 for detail information about request schema and policy
-schema.
-\item delegtion.pdp – Delegation PDP\\
-The Delegation PDP is basically similar to Arc PDP, except it uses the delegation policy parsed from
-remote proxy credential by delegation.collector, and evaluates the request against delegation policy. See
-section 6for the design idea and use case of delegation policy in fine-grained identity delegation.
-\item simplelist.pdp – Simplelist PDP\\
-The Simplelist PDP is a simplest implementation of policy decision point. It will match the identity
-extracted from the remote credential (or proxy credential) with local list of permitted identities.
-\item pdpservice.invoker – PDP Service Invoker\\
-The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the
-same functionality as Arc PDP, except that the evaluation request and response are carried by SOAP
-message. The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation
-engine can be accessed remotely and maintained centrally.
- \item allow.pdp\\
- Does something unknown!
-\end{itemize}
 
@@
 \end{minipage}
+
+\section{Other PDPs}
+
+In order the adminisitrator wants to maintain the grid access using one workstation the delegation.pdp can be used.
+The client requests access to a service. In case the request is valid a certificate will be deployed on the desired server such that it can be accessed now \task{Am I right?} \cite{QIANG_2008}\\

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Express checkout... Finished the basic TLS chapter content and started writting about the Sec echo client...

@@
 \section{Service}
+\task{RENAMING HED configuration file into~ARC service configuration file - proposed by Marek}
+\task{(echo service) client configuration file - proposed by Marek}
 
@@
 The CA for the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key.
+For the given example X.509-certificates will be used which are standardised by the ITU-T.
 Altogether six files are needed:
-\begin{tabular}{l@{-}p{9cm}}
-clientCA   & The CA which guarantees for the identity of the client \\
-clientCERT & The certificate of the client. \\
-clientKEY  & Secret key of the client to read messages, encrypted by the public key in the certificate! \\
+\begin{tabular}{l@{~---~}p{13cm}}
+\textbf{clientCA}   & The CA which guarantees for the identity of the client \\
+\textbf{clientCERT} & The certificate of the client. \\
+\textbf{clientKEY}  & The private key of the client (should be kept in safe custody).\\
 \\
-serverCA   & The CA which guarantees for the identity of the server \\
-serverCERT & The certificate of the server. \\
-serverKEY  & Secret key of the client to read messages, encrypted by the public key in the certificate! \\
+\textbf{serverCA}   & The CA which guarantees for the identity of the server \\
+\textbf{serverCERT} & The certificate of the server. \\
+\textbf{serverKEY}  & The private key of the server (should be kept in safe custody).\\
 \end{tabular}
+\forcelinebreak
+As one can see, PEM has been choosen for the file format.\\
 
-The server needs: its  serverCERT and serverKEY and some kind of root CA knowing the client.
-Client needs its clientCERT and clientKEY and some kind of root CA knowing the server.
+To get this service running, the six files have to be distributed.
+The server should hold the files: serverCERT.pem, serverKEY.pem and clientCA.pem.
+The client on the other side should own the files: clientCERT.pem, clientKEY.pem and serverCA.pem.
+Within the source code directory certificates with corresponding names are prepared.
+Due to certificates have a limited period of validity, these certificates may no longer be accepted by the communication parter.
+In order to create a first basic set of certificates signed by a self created CA, the shell scripts \textit{createMasterCA.sh} and \textit{createSlaveCert.sh} to be found in the directory \textit{src/services/tlsechoservice/certFactory} may be used.
+The script \textit{createMasterCA.sh} will create a new CA and overwrite the old on which was stored in the directory.
+The script \textit{createSlaveCert.sh} create an arbitary amount of certificates signed by the CA which has been create by the other script.\\
+%WARNING: If the programm fails, check if one of your certifacte is expired. If so, create new one using the script files in the directory 'certFactory'
 
-\task{RENAMING into~ARC service configuration file}
-\task{echo service client configuration file}
-
-%The first two goals can be reached by encrypting the channel/connection between client and server. An approach 
-%sich bewähren / behaupten
-%which stand the test of time is to introduce a TLS (Transport Layer Security) also known as SSL (Secure Sockets Layer).
-%Authenticity, Confidentiality, Integrity
-%The client wants to reassure himself to send his vulnerable data only to the server who can prove its correct identity.
-%Furthermore a prove of authenticity is needed. % Sich versichern, dass das Gegenüber der ist, für welchen er sich ausgibt
-%It .... does fancy stuff ...
-%
-%
-%Authorization:  The server on the other side takes an interest in limitate its ressources to a close circle of users. Above all no %foreign clients shall be able to use the services.\\
-
-One can add the TLS in client and server by simply extending the arched configuration file and a small modification of the client.
-
-WARNING: If the program fails, check if one of your certificates has expired.
-If so, create new one using the script files in the directory 'certFactory'.
-
-Chalked up to the modularity of HED, the security can easily be extend to our given example.
-Each Message Chain Component (MCC) or service has a common interface for implementing various pluggable components (plug-ins) called SecHandler.
-The SecHandler components provide a method for processing messages traveling through Message Chains of the HED.
-
-Each MCC or Service usually implements two queues of SecHandlers – one for incoming messages and one for outgoing, which are called “incoming” and “outgoing”, respectively.
-All SecHandler components attached to the queue are executed sequentially. 
-If any of them fails, message processing fails as well. 
-Each SecHandler is configured inside the arched configuration file that is used for configuring whole chain of MCCs 
-Some of the currently implemented SecHandler components make use of pluggable and configurable sub-modules called Policy Decision Point (PDP).
-PDPs can process ARC specific Request and Policy documents which are also written in XML format~\cite{QIANG_2008}..
-
-4 types:
-                <KeyPath>./clientKey.pem</KeyPath>
-                <CertificatePath>./clientCert.pem</CertificatePath>
-                <CACertificatePath>./serviceCA.pem</CACertificatePath>
-
-\lstsetCPP
-\lstinputlisting
-	[
-	label=lst:tls_echo_service_cpp,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
-	]
-{../src/services/tlsechoservice/tlsechoservice.cpp}
+The TLS echo service implementation is almost the same as the one presented in the chapter before.
+Merely the Class name and the namespace of the service has changed.
+The only modification has to be done within the ARC service configuration file which is displayed in Listing~\ref{lst:tls_echo_arched_xml}.
 
@@
 {../src/services/tlsechoservice/arched_tls_echoservice.xml}
+The first change is to be found in lin~\ref{lst_code:tls_echo_arched_xml_plugin}.
+As to be seen, another plugin called \textit{mcctls} will be loaded by the \textit{ModuleManager}.
+The second change is the new element which has been placed between the TCP and the HTTP component, see line~\ref{lst_code:tls_echo_arched_xml_component}.
+Within the element the files are declared using the elements
+\begin{verbatim}
+<KeyPath>./clientKey.pem</KeyPath>
+<CertificatePath>./clientCert.pem</CertificatePath>
+<CACertificatePath>./serviceCA.pem</CACertificatePath>
+<CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+\end{verbatim}
+or
+\begin{verbatim}
+<ClientSSLConfig FromFile='filename'/>
+\end{verbatim}
+\task{You can put the ClientSSLConfig into a separate file, then use <ClientSSLConfig FromFile='filename'/> in each service, and then you don't need to change everywhere - proposed by Zsombor}
 
-\lstsetKSH
-\begin{lstlisting}[
-label=lst:tls_echo_arched_invoke,float=htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ rm -f /var/log/arched.log
-$ arched -c arched_echoservice.xml  && echo jo ||echo n
-$ tail -n100 -f /var/log/arched.log
-$ killall arched
-\end{lstlisting}
+%\lstsetCPP
+%\lstinputlisting
+%	[
+%	label=lst:tls_echo_service_cpp,
+%	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
+%	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+%	]
+%{../src/services/tlsechoservice/tlsechoservice.cpp}
+%
+% Lassen wir mal weg....
+%
+%
+% \lstsetKSH
+% \begin{lstlisting}[
+% label=lst:tls_echo_arched_invoke,float=htb,
+% caption={[Transformation in eine uniforme konzentrische Verteilung.]
+%          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+% $ rm -f /var/log/arched.log
+% $ arched -c arched_echoservice.xml  && echo jo ||echo n
+% $ tail -n100 -f /var/log/arched.log
+% $ killall arched
+% \end{lstlisting}
+% 
+%        ...das auch.....
 
@@
 
+In this section a new way to implement the client will be introduced.
+As already mentioned in section~\ref{sec:timeservice_client}, the protocol stack of the client is build up in almost the same way like the service.
+Indeed, the class~\textit{ClientSOAP} composes the stack with the same MCC used for the service.
+Once this is known, its self-evident to create the client in a simlar manner like the service with an own configuration file.
+Listing~\ref{lst:tls_echo_client_cpp} shows the C++ source code of the client implementation.
+The main difference between the simple echo client presented in the chapter before is that now the information about the client protocal stack is encapsulated within a configuration file.
 \lstsetCPP
@@
 {../src/clients/tlsechoclient/tlsechoclient.cpp}
+The file containing the client configuration is specified within the command line. 
+Its content will be loaded in the code fragment starting in line~\ref{lst_code:tls_echo_client_cpp_getfilecontent}. 
+Later, in line~\ref{lst_code:tls_echo_client_cpp_loadconfiguration}, the configuration will be parsed into XML and than transfered into a chain of component which is wrapped by the object \textit{loader}.
+In order to process a message, an entry point within the chain has to be disclosed.
+For we are implementing a SOAP client, we are looking for the SOAP component.
+To get the proper component an operator of MCCLoader is used in line~\ref{lst_code:tls_echo_client_cpp_cliententry}.
+The previously used class \textit{ClientSOAP} has eased some workload which now has to be done within the method.
+The payload now have to be wrapped by the message class which have to be initialised with an \textit{MessageAttribute} object and a \textit{MessageContext} object.
+After the payload has been wrapped within the message it gets transmitted in almost the same way like before, see lines followed by line~\ref{lst_code:tls_echo_client_cpp_createandprocess}.
+The rest of the source code is identical to the echo client.\\
 
-Folgender XML Aufruf und Antwort soll \textit{automatisch} generiert werden:
+% Folgender XML Aufruf und Antwort soll \textit{automatisch} generiert werden:
 
-%[caption={[]WegDamit},language=XML,basicstyle=\scriptsize,breaklines=true,label=lst:request]
-
-Zertifikate
-Zustände (Nutzer wiedererkennen, arbeit aufnehmen)
-
-\lstsetKSH
-\begin{lstlisting}[
-label=lst:tls_echo_client_invoke,float=htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ ./echoclient
-Tue Feb 17 17:04:08 2009
-\end{lstlisting}
-
-\lstsetJUSTXML
-\begin{lstlisting}[
-label=lst:tls_echo_request_XML,float=htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-<soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
-  <soap-env:Body>
-    <tlsecho:tlsechoRequest>
-      <tlsecho:say operation="ordinary">text_to_be_transmitted</tlsecho:say>
-    </tlsecho:tlsechoRequest>
-  </soap-env:Body>
-</soap-env:Envelope>
-\end{lstlisting}
-
-\lstsetJUSTXML
-\begin{lstlisting}[
-label=lst:tls_echo_response_XML,float=htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-<soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
-  <soap-env:Body>
-    <tlsecho:tlsechoResponse>
-      <tlsecho:hear>[ text_to_be_transmitted ]</tlsecho:hear>
-    </tlsecho:tlsechoResponse>
-  </soap-env:Body>
-</soap-env:Envelope>
-\end{lstlisting}
+% 
+% \lstsetJUSTXML
+% \begin{lstlisting}[
+% label=lst:tls_echo_request_XML,float=htb,
+% caption={[Transformation in eine uniforme konzentrische Verteilung.]
+%          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+% <soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
+%   <soap-env:Body>
+%     <tlsecho:tlsechoRequest>
+%       <tlsecho:say operation="ordinary">text_to_be_transmitted</tlsecho:say>
+%     </tlsecho:tlsechoRequest>
+%   </soap-env:Body>
+% </soap-env:Envelope>
+% \end{lstlisting}
+% 
+% 
+% 
+% 
+% 
+% 
+% \lstsetJUSTXML
+% \begin{lstlisting}[
+% label=lst:tls_echo_response_XML,float=htb,
+% caption={[Transformation in eine uniforme konzentrische Verteilung.]
+%          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+% <soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
+%   <soap-env:Body>
+%     <tlsecho:tlsechoResponse>
+%       <tlsecho:hear>[ text_to_be_transmitted ]</tlsecho:hear>
+%     </tlsecho:tlsechoResponse>
+%   </soap-env:Body>
+% </soap-env:Envelope>
+% \end{lstlisting}
+%

### 2009-03-01 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

@@
 %
-\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
+\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
 \hypersetup{
@@
 
-The A-Hash is a replicated metadata store, which is capable of consistently storing `objects' where an object contains key-value pairs organized in sections.
+The A-Hash is a replicated metadata store, which is capable of consistently storing `objects' where an object contains property-value pairs organized in sections.
 All metadata about files and collections are stored in the A-Hash, and some other information (e.g.
@@
     \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
-    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are key-value pairs)
+    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are property-value pairs)
 \end{itemize}
@@
 
-The A-Hash services provide a functionality to store `objects', where each object has a unique ID, and contains key-value pairs organized in sections.
+The A-Hash services provide a functionality to store `objects', where each object has a unique ID, and contains property-value pairs organized in sections.
 The properties, the values and the section names too are simple character strings.
@@
 The Librarian services use the A-Hash services to store the metadata of all kinds, using the GUIDs as unique IDs.
-For this, we have to represent the metadata as key-value pairs organized in section.
-The files, the collections and the mount points have some common attributes: the GUID, the type and the owner which are in the \emph{entry} section; the creation and modification timestamps in the \emph{timestamps} section, a list of access control rules in the \emph{policy} section; the list of parent collections where this entry has a `hardlink'; and each entry could have arbitrary key-value pairs in the \emph{metadata} section.
+For this, we have to represent the metadata as property-value pairs organized in section.
+The files, the collections and the mount points have some common attributes: the GUID, the type and the owner which are in the \emph{entry} section; the creation and modification timestamps in the \emph{timestamps} section, a list of access control rules in the \emph{policy} section; the list of parent collections where this entry has a `hardlink'; and each entry could have arbitrary property-value pairs in the \emph{metadata} section.
 
@@
     \begin{itemize}
-        \item any other arbitrary key-value pairs
+        \item any other arbitrary property-value pairs
     \end{itemize}
@@
 % The Librarian uses the A-Hash to store all the data about files and collections.
-The A-Hash is capable of storing key-value pairs organized in sections, which actually means that it stores (\emph{section}, \emph{key}, \emph{value}) tuples where each member is simply a string, e.g.~(`entry', `type', `collection') or (`ACL', `johnsmith', `owner') or (`timestamps', `created', `1196265901') or (`locations', `64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB:9A293F27DC86', `sentenced').
-There could be only one \emph{value} for a (\emph{section}, \emph{key}) pair.
+The A-Hash is capable of storing property-value pairs organized in sections, which actually means that it stores (\emph{section}, \emph{property}, \emph{value}) tuples where each member is simply a string, e.g.~(`entry', `type', `collection') or (`ACL', `johnsmith', `owner') or (`timestamps', `created', `1196265901') or (`locations', `64CDF45F-DDFA-4C1D-8D08-BCF7810CB2AB:9A293F27DC86', `sentenced').
+There could be only one \emph{value} for a (\emph{section}, \emph{property}) pair.
 
@@
 The ARC HED hosting environment has a security framework, and it has its own policy language for describing access policies and requests.
-While the storage system use its own format to store the access policies, but one the time comes to make decision, it converts the stored policies and the incoming requests to this ARC native policy format, and uses the security framework to decide.
-
-%%%%%%%%%%%%%%%%
+The storage system use a different internal format to store the access policies, but when the time comes to make decisions, it converts the stored policies and the incoming requests to the ARC native policy format, then asks the security framework to decide.
+
+The internal representation of the access policies is based on access rules. A rule contains an identity and a list of actions. Currently the identity could be one of the following:
+\begin{itemize}
+    \item a Distinguish Name (DN) of a user, e.g. `\verb!/DC=eu/DC=KnowARC/O=NIIFI/CN=james!'
+    \item a name of a Virtual Organization (VO) prefixed with `\verb!VOMS:!', e.g. `\verb!VOMS:knowarc.eu!'
+    \item `\verb!ALL!' to specify that this rule will be applied to everyone
+    \item `\verb!ANONYMOUS!' for a rule which will be applied to unauthenticated users (e.g. when non-secure HTTP is used and there is no DN)
+\end{itemize}
+
+Then for each rule there is list of actions, each action is either allowed or denied.
+In the internal representation of an access rule the list of actions is one single string, where the name of actions are separated with a space, and each name is prefixed with a `\verb!+!' or `\verb!-!' sign indicating if the action is allowed or denied for the given identity.
+The list of actions can be found in section~\ref{sub:high_level_authorization}.
+
+When there is an incoming connection to a Bartender, the security framework of the HED extract the identity information of the connecting client (DN of client, DN of the issuer CA, extended attributes, etc.), and all these information is accessible by the Bartender.
+When the given method of the Bartender gets to the point where the access rules are present, and it is clear that what kind of action the user wants to do, then either the request and the policy is converted to the native ARC policy language, and then the policy evaluator of the security framework is called to make the decision.
+Within the native ARC policy language XML representation of the requests and policies the storage actions are put into an `Attribute' element with the ID `\verb!http://www.nordugrid.org/schemas/policy-arc/types/storage/action!'.
+
+Because of the heavy use of X509 certificates, of course, all the services should have their own certificates, which they can use when connecting to other services.
+Currently it is needed to put the paths of the certificate and private key files (and the trusted CAs as well) into the configuration of each services, like this:
+
+\begin{verbatim}
+    <ClientSSLConfig>
+        <KeyPath>certs/hostkey-epsilon.pem</KeyPath>
+        <CertificatePath>certs/hostcert-epsilon.pem</CertificatePath>
+        <CACertificatesDir>certs/CA</CACertificatesDir>
+    </ClientSSLConfig>
+\end{verbatim}
 
@@
 
-The A-Hash will be a distributed service capable of storing tuples of strings in a scalable manner.
-Currently it only has a centralized implementation.
-It stores \emph{objects}, where each object has an arbitrary string \emph{ID}, and contains any number of \emph{key}-\emph{value} pairs grouped in \emph{section}s, where \emph{key}, \emph{value} and \emph{section} are arbitrary strings.
-There could only be a single \emph{value} for a \emph{key} in a \emph{section}.
-
-If you have an ID, you can get all key-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties do you need.
-You can add or remove key-value pairs of an object or delete all occurrences of a key or create a new object with the \emph{change} method, and you can specify conditions, which means the change is only applied if the given conditions are met.
+The A-Hash is a database for storing string tuples, and it provides conditional and atomic modification of them.
+The A-Hash can be used as a centralized service, or it could be deployed in a distributed way, using multiple nodes, where all the data is replicated on all the nodes.
+
+The A-Hash stores \emph{objects}, where each object has an arbitrary string \emph{ID}, and contains any number of \emph{property}-\emph{value} pairs grouped in \emph{section}s, where \emph{property}, \emph{value} and \emph{section} are arbitrary strings.
+There could only be a single \emph{value} for a \emph{property} in a \emph{section}.
+
+If you have an ID, you can get all property-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties do you need.
+You can add or remove property-value pairs of an object or delete all occurrences of a property or create a new object with the \emph{change} method, and you can specify conditions, which means the change is only applied if the given conditions are met.
 
@@
 
-\subsection{Prototype status and plans} % (fold)
-
-The A-Hash service currently implemented as a single central service, which stores the data on disk in separate files per \emph{object}.
-Fall of 2008 it will be reimplemented using a distributed hash table (DHT) algorithm, one possible candidate is the Chord\footnote{The Chord project - \url{http://pdos.csail.mit.edu/chord/}} algorithm with a consistency solution called Etna\footnote{Etna: a Fault-tolerant Algorithm for Atomic Mutable DHT Data - \url{http://pdos.csail.mit.edu/\~athicha/papers/etna.ps}} on top of it.
-This reimplementation hopefully won’t change the interface of the service.
-
-% subsection prototype_status_and_plans (end)
-
-\subsection{Data model} % (fold)
-
-\begin{itemize}
-    \item \emph{ID} is an arbitrary string
-    \item \emph{object} contains key-value pairs in sections, technically it is a list of key-\emph{value} pairs where the key is a (\emph{section}, \emph{key}) tuple
-\end{itemize}
-
-% subsection data_model (end)
-
 \subsection{Interface} % (fold)
@@
 \begin{description}
-    \item [get(ids, neededMetadata)] returns \emph{getResponse} which is a list of (\emph{ID}, \emph{object}) pairs.
-
-    The \emph{ids} is a list of string \emph{ID}s, \emph{neededMetadata} is a list of (\emph{section}, \emph{key}) pairs.
-For each \emph{ID} it returns all the \emph{value}s for each \emph{key} in each \emph{section} (filtered by \emph{neededMetadata}), so \emph{object} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples.
-
-    \item [change(changeRequest)] returns \emph{changeResponse} which is a list of (\emph{changeID}, \emph{success}, \emph{failedConditionID}) tuples.
-
-    \emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{key}, \emph{value}, \emph{conditions}), where \emph{changeID} is an arbitrary ID to identify in the response which change was successful; \emph{ID} points to the object we want to change; \emph{changeType} can be `\textbf{set}' (set the key within the section to value), `\textbf{unset}' (remove the key from the section regardless of the value), `\textbf{delete}' (removes the whole object), conditions is a list of (\emph{conditionID}, \emph{type}, \emph{section}, \emph{key}, \emph{value}) tuples, where \emph{type} could be `\textbf{is}' (the key in the section is set to the value), `\textbf{isnot}' (the key in the section is not set to the value), `\textbf{isset}' (the key of the section is set to any value), `\textbf{unset}' (the key of the section is not set at all).
+    \item [get(IDs, neededMetadataList)] returns all or some of the property-value pairs of the requested objects.
+
+    The \emph{IDs} is a list of string \emph{ID}s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs.
+If the \emph{neededMetadataList} is empty, then for each \emph{ID} it returns all the \emph{value}s for all the \emph{property} in all the \emph{section} in that given object.
+If there are sections and properties specified in \emph{neededMetadataList}, then only those values are returned.
+
+    The response contains a list of \emph{object}s, where an \emph{object} is an (\emph{ID}, \emph{metadataList}) pair, where \emph{metadataList} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+
+    \item [change(changeRequestList)] modifies, removes or creates objects if certain conditions are met, and returns information about the success of the modification requests.
+
+    The \emph{changeRequestList} is a list of \emph{changeRequest}s, where a \emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}, \emph{conditionList}), where \emph{changeID} is an arbitrary ID which is used in the response to refer to this part of the request; \emph{ID} points to the object we want to change; \emph{changeType} can be `\textbf{set}' (to set the property within the section to value), `\textbf{unset}' (to remove the property from the section regardless of the value), `\textbf{delete}' (to remove the whole object); \emph{conditionList} is a list of \emph{condition}s, where a \emph{condition} is a tuple of (\emph{conditionID}, \emph{conditionType}, \emph{section}, \emph{property}, \emph{value}), where \emph{conditionType} could be `\textbf{is}' (which will be true if the property in the section is set to the value), `\textbf{isnot}' (which will be true if the property in the section is not set to the value), `\textbf{isset}' (which will be true if the property of the section is set, regardless of the value), `\textbf{unset}' (which will be true if the property of the section is not set at all).
     If all conditions are met, tries to apply changes to the objects, creates a new object if a previously non-existent ID is given.
-If one of the conditions is not met, returns the ID of the failed condition.
+
+    The response contains the \emph{changeResponseList} which is a list of (\emph{changeID}, \emph{success}, \emph{failedConditionID}) tuples.
+The \emph{success} of the change could be \textbf{set}, \textbf{unset}, \textbf{deleted}, \textbf{failed}, \textbf{condition not met} (in this case the ID of the failed condition is put into \emph{failedConditionID}), \textbf{invalid change type} or \textbf{unknown}.
+
 \end{description}
@@
 % subsection interface (end)
+
+\subsection{Implementation} % (fold)
+
+The A-Hash service has a modular architecture which means that you can decide if you want to deploy it a centralized or a distributed way by simply specifying different modules in the service's configuration.
+There are different low-level modules for storing the data on disk, e.g.
+serialized into `pickle' format.
+And there is a module which stores the data replicated on multiple nodes.
+
+TODO: here comes a fairly detailed description of how the replicated A-Hash works.
+
+% subsection implementation (end)
+
+\subsection{Configuration} % (fold)
+
+The A-Hash has the following configuration variables:
+
+\begin{description}
+    \item[AHashClass] tells the A-Hash service the name of the business-logic A-Hash class.
+Currently there are two implementation, the \verb!CentralAHash!
+is a centralized version, and the \verb!ReplicatedAHash!
+is a replicated version.
+    \item[StoreClass] specifies the name of the store class.
+This class is used to store the data on disk.
+Currently there are several versions (\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore!
+is suitable for using in a replicated deployemant, all the others are only for a centralized deployement.
+    \item[StoreCfg] could contain some parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash should save its files.
+\end{description}
+
+TODO: describe here the configuration needed for the replicated A-Hash.
+
+An example configuration:
+
+\begin{verbatim}
+    <Service name="pythonservice" id="ahash">
+        <ClassName>storage.ahash.ahash.AHashService</ClassName>
+        <AHashClass>storage.ahash.ahash.CentralAHash</AHashClass>
+        <StoreClass>storage.store.cachedpicklestore.CachedPickleStore</StoreClass>
+        <StoreCfg>
+            <DataDir>ahash_data</DataDir>
+        </StoreCfg>
+    </Service>
+\end{verbatim}
+
+% subsection configuration (end)
 
@@
 %
-The Librarian manages a tree-hierarchy of files, grouping them into collections.
-There is a root collection with a well-known GUID which can be used as starting point when resolving Logical Names.
-If you create a new collection with the method \emph{new}, the Librarian generates a new GUID, but does not insert it into the tree-hierarchy which can be done by adding this GUID as a new entry to one of the existing collection using the \emph{modifyMetadata} method of the existing collection which makes it the parent of the new collection.
-A collection can be closed via metadata modification which cannot be undone and prevents files to be added or removed from this collection.
-A new file  also can be created with the \emph{new} method which returns the newly generated GUID of the new file entry which should be added to a parent collection to insert it into the global namespace.
-A file has a list of locations where its replicas are stored, this list too can be manipulated with \emph{modifyMetadata}.
-The access policies of the files and collections are also stored as metadata.
-The \emph{remove} method deletes an entry from the Librarian.
-The \emph{traverseLN} method try to traverse Logical Names by walking the hierarchy of the namespace and to return the GUID of the entry pointed by the LN.
-After you have a GUID of file, collection or mount point, you can get all the information using the \emph{get} method.
+
+The Librarian knows about the type of possible entries in the namespace of the ARC storage system: files, collections and mount points.
+It knows that these are organized in a tree-hiearchy with grouping them into collections, and it knows about Logical Names, and how to traverse a Logical Name to find the metadata of the file or collection (or mount point) which it refers to.
+However, the Librarian does not maintain the namespace of the ARC storage, e.g.
+it does not put new files into collections, or remove the link from its parent collection when a file is removed - this is the job of the Bartender.
+The only Librarian method which deals with Logical Names is the \emph{traverseLN} method, which traverses the Logical Name as long as possible, and returns the GUID and metadata of the last found entry in the path of the Logical Name.
+All the other methods work with GUIDs, and do not care about the hierarchical namespace.
+In this regard the Librarian can be treated as a higher level application-specifc interface to the metadata store (the A-Hash), because while the A-Hash does not care about what it stores, the Librarian knowa exactly that it stores metadata about files, collections, etc.
+With this interface you can create new files, collections and mount points with the \emph{new} method, you can get or modify the metadata of any entry with the \emph{get} and the \emph{modifyMetadata} methods, and you can remove an entry with the \emph{remove} method.
+
+Besides being an interface to the metadata store, the Librarian has an other importat functionality: it provides a way for Shepherds to register themselves, and to periodically send reports about the state of their stored files.
+A Shepherd send a \emph{report} message to a Librarian if e.g.
+a new file arrived, or an old file gets corrupted, thus the Librarian can keep the states of the replicas up-to-date.
+The Librarian monitors the registered Shepherds, and if one of them stops sending reports, the Librarian will assume, that it is offline, and modify the states of the all the replicas which are on the Shepherd's storage node.
+In order to do this, the Librarian stores for each Shepherd a list of files on the given storage nodes, this make it possible to know which files have replicas on a given storage nodes without checking all the files in the system.
+This list of files gets updated every time a Shepherd sends a report.
+
+The end-user of the ARC storage system never communicates directly with a Librarian, the Bartenders will contact the Librarians in order to fulfil the user's requests.
 
@@
 
-\subsection{Prototype status and plans} % (fold)
-
-The Librarian service currently implements all the methods below, but doesn't do very much error checking.
-This should be changed, the Librarian should check the validity of metadata, and forbid some cases, e.g.~reopen a closed collection.
-
-% subsection prototype_status_and_plans (end)
-
-\subsection{Data model} % (fold)
-\label{sub:data_model}
-
-% subsection data_model (end)
-
 \subsection{Interface} % (fold)
@@
 \begin{description}
-    \item[new(newRequestList)] returns a list of (\emph{requestID}, \emph{GUID}, \emph{success})
-
-    \emph{newRequestList} is a list of (\emph{requestID}, \emph{metadata}) where \emph{requestID} is an arbitrary ID used to identify this request in the list of responses; \emph{metadata} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples.
+
+    \item [get(getRequestList, neededMetadataList)] returns all or some of the metadata of the the requested GUIDs.
+
+    The \emph{getRequestList} is a list of string \emph{GUID}s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs.
+If the \emph{neededMetadataList} is empty, then for each \emph{GUID} it returns all the metadata \emph{value}s for all the \emph{property}s in all the \emph{section}s.
+If there are sections and properties specified in \emph{neededMetadataList}, then only those values are returned.
+
+    The response contains a list of \emph{GUID, metadatList} pairs, where \emph{metadataList} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+
+    \item[new(newRequestList)] creates a new entry in the Librarian with the given metadata, and returns the GUID of the new entry.
+
+    The \emph{newRequestList} is a list of (\emph{requestID}, \emph{metadataList}) where \emph{requestID} is an arbitrary ID used to identify this request in the list of responses; \emph{metadataList} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
     This method generates a \emph{GUID} for each request, and inserts the new entry (with the given metadata) into the A-Hash, then returns the GUIDs of the newly created entries.
-The (`entry', `type') key of the metadata contains whether it is a file or a collection.
-
-    \item[modifyMetadata(modifyMetadataRequestList)] returns a list of (\emph{changeID}, \emph{success})
-
-    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{key}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the key in the section to value), `\textbf{unset}' (remove the key-value pair from the section), `\textbf{add}' (set the key in the section to value only if it is not exists already).
-
-    \item [get(GUIDs, neededMetadata)] returns \emph{getResponse}
-
-    \emph{GUIDs} is a list of GUIDs, \emph{neededMetadata} is a list of (\emph{section}, \emph{key}) pairs indicating only which properties we need, \emph{getResponse} is a list of (\emph{GUID}, \emph{metadata}) where metadata is a list of (\emph{section}, \emph{key}, \emph{value}) tuples.
-This method returns the metadata of all the \emph{GUID}s filtered with \emph{neededMetadata}.
-
-    \item [remove(removeRequestList)] returns a list of (\emph{requestID}, \emph{success}) pairs
-
-    \emph{removeRequestList} is a list of (\emph{requestID}, \emph{GUID}) pairs. \emph{success} could be `\textbf{removed}' or `\textbf{failed}: reason'.
-
-    \item [traverseLN(traverseRequestList)] returns \emph{traverseResponseList}
-
-    \emph{traverseRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names to be traversed
-
-    \emph{traverseResponseList} is a list of (\emph{requestID}, \emph{metadata}, \emph{GUID}, \emph{traversedLN}, \emph{restLN}, \emph{wasComplete}, \emph{traversedList}) where:
-    \begin{description}
-        \item[metadata] is all the metadata of the of traversedLN in the form of (\emph{section}, \emph{key}, \emph{value}) tuples
+In the metadata of the new entry the `type' property in the `entry' section defines whether it is a file, a collection or a mount point.
+The `GUID' property can contain a GUID if we want to specify the GUID of the new entry, and don't want the Librarian to generate a random one.
+
+    The response contains a list of (\emph{requestID}, \emph{GUID}, \emph{success}) tuples.
+
+    \item[modifyMetadata(modifyMetadataRequestList)] modifies the metadata of the given entries.
+
+    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) tuples where \emph{changeType} can be `\textbf{set}' (set the property in the section to value), `\textbf{unset}' (remove the property-value pair from the section), `\textbf{add}' (set the property in the section to value only if it does not exist already) or `\textbf{setifvalue}=value' (only set the property in the section to value if it currently equals to the `value' in the \emph{changeType}).
+
+    The response is a list of (\emph{changeID}, \emph{success}) where \emph{success} can be `\textbf{set}', `\textbf{unset}', `\textbf{condition failed}', `\textbf{failed}: reason'.
+
+    \item [remove(removeRequestList)] removes the given entries.
+
+    The \emph{removeRequestList} is a list of (\emph{requestID}, \emph{GUID}) pairs.
+
+    The response is a list of (\emph{requestID}, \emph{success}) pairs where \emph{success} could be `\textbf{removed}' or `\textbf{failed}: reason'.
+
+    \item [traverseLN(traverseRequestList)] traverses the given Logical Names and returns extensive information about them.
+
+    The \emph{traverseRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names to be traversed
+
+    The response is a list of (\emph{requestID}, \emph{traversedList}, \emph{wasComplete}, \emph{traversedLN}, \emph{GUID}, \emph{metadataList}, \emph{restLN}) where:
+    \begin{description}
+        \item[traversedList] is a list of (\emph{LNpart}, \emph{GUID}) pairs, where \emph{LNpart} is a part of the \emph{LN}, \emph{GUID} is the GUID of the Librarian-entry referenced by that part of the \emph{LN}, the first element of this list is the shortest prefix of the \emph{LN}, the last element is the \emph{traversedLN} without its last part
+        \item[wasComplete] indicates whether the full \emph{LN} was traversed
+        \item[traversedLN] is the part of the \emph{LN} which was traversed, if \emph{wasComplete} is true, this should be the full \emph{LN}
         \item[GUID] is the \emph{GUID} of the \emph{traversedLN}
-        \item[traversedLN] is the part of the \emph{LN} which was traversed, if \emph{wasComplete} is true, this should be the full \emph{LN}
+        \item[metadataList] is all the metadata of the of traversedLN in the form of (\emph{section}, \emph{property}, \emph{value}) tuples
         \item[restLN] is the postfix of the \emph{LN} which was not traversed for some reason, if \emph{wasComplete} is true, this should be an empty string
-        \item[wasComplete] indicates whether the full \emph{LN} was traversed
-        \item[traversedList] is a list of (\emph{LNpart}, \emph{GUID}) pairs, where \emph{LNpart} is a part of the \emph{LN}, \emph{GUID} is the GUID of the Librarian-entry referenced by that part of the \emph{LN}, the first element of this list is the shortest prefix of the \emph{LN}, the last element is the \emph{LN} without its last part
-    \end{description}
-
-    \item [report(serviceID, filelist)] returns in \emph{nextReportTime} a number of seconds, which is the timeframe within the Librarian expects the next heartbeat from the Shepherd
-
-    \emph{filelist} is a list of (\emph{GUID}, \emph{referenceID}, \emph{state}) tuples containing the state of changed or new files, where \emph{state} could be `\textbf{invalid}' (if the periodic self-check of the Shepherd found a non-matching checksum or missing file), `\textbf{creating}' (if this is a new file not uploaded yet) or `\textbf{alive}' (if the file is uploaded and the checksum is OK).
+    \end{description}
+
+    \item [report(serviceID, filelist)] is a report message from a Shepherd to a Librarian, contains the ID of the Shepherd (currently the URL), and a list of changed files on the Shepherd's storage node.
+
+    The \emph{filelist} is a list of (\emph{GUID}, \emph{referenceID}, \emph{state}) tuples containing the state of changed or new files, where referenceID is the Sheperd-local ID of the given replica where GUID refers to the logical file of this replica.
+The \emph{state} could be `\textbf{invalid}' (if the periodic self-check of the Shepherd found a non-matching checksum or missing file), `\textbf{creating}' (if this is a new file not uploaded yet) or `\textbf{alive}' (if the file is uploaded and the checksum is OK).
+
+    The response is \emph{nextReportTime}, a number of seconds, which is the timeframe within the Librarian expects the next heartbeat from the Shepherd.
 
@@
 % subsection interface (end)
+
+\subsection{Implementation} % (fold)
+
+% subsection implementation (end)
+
+\subsection{Configuration} % (fold)
+
+% subsection configuration (end)
 
@@
 
-    \emph{getRequestData} is a list of (\emph{key}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).
-
-    \emph{getResponseData} is a list of (\emph{key}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.
+    \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).
+
+    \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.
 
@@
 
-    \emph{putRequestData} is a list of (\emph{key}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (for additional access policy).
-
-    \emph{putResponseData} is a list of (\emph{key}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.
+    \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (for additional access policy).
+
+    \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.
 
@@
 simply refers to the root collection.
-The term `\emph{metadata}' here refers to a list of key-value pairs organized in sections, see the data model description in Section~\ref{sub:data_model}.
+The term `\emph{metadata}' here refers to a list of property-value pairs organized in sections, see the data model description in Section~\ref{sec:data_model}.
 
@@
 
-    \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other key-value pairs in the `\textbf{metadata}' section.
+    \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other property-value pairs in the `\textbf{metadata}' section.
 The returned \emph{TURL} is a URL with a chosen \emph{protocol} to upload the file itself, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.
@@
 
-    \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{key}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sub:data_model})
+    \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sec:data_model})
 
@@
 
-    \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' key (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in key-value pairs.
+    \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in property-value pairs.
 The \emph{success} in the response could be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.
@@
 
-    \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{key}) pairs which filters the returned metadata.
-
-    \emph{listResponse} is a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{key}, \emph{value}) tuples according to the data model of the Librarian (Section~\ref{sub:data_model}), the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}' (because only collections can be listed).
+    \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
+
+    \emph{listResponse} is a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (Section~\ref{sec:data_model}), the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}' (because only collections can be listed).
 
@@
 
-    \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{key}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{key} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{key}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{key} in the \emph{section} to \emph{value} only if it is not exists already).
+    \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).
 \emph{success} could be `\textbf{no such LN}', `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}', `\textbf{failed}: reason'.
@@
 To integrate existing files on a third-party storage system to our namespace thus make them accessible through the interface of the ARC storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
-It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC storage (Section~\ref{sub:data_model}).
+It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC storage (Section~\ref{sec:data_model}).
 The files here would be referenced by a path which is local in the namespace of the third-party storage.

### 2009-03-02 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

@@
 \subsection{Implementation} % (fold)
+
+The Librarian service uses the A-Hash to store all the metadata, and it uses the GUIDs as IDs in the database.
+It has the URL of one or more A-Hash services from its configuration, and it can acquire more A-Hash URLs from these first A-Hashes in case of a replicated A-Hash deployement.
+
+TODO: here comes some sentences about how the Librarian reads from any A-Hash and finds the master A-Hash for writing.
+
+The Librarian also accepts report messages from Shepherds, and stores the contents of these messages in the A-Hash.
+There is one A-Hash object which contains the ID and the expected time of the next heartbeat of the registered Shepherds, and there is one A-Hash object for each Shepherd which stores the states and GUIDs of the files on the Shepherd's storage node.
+The Librarian checks periodically if there is a Shepherd which is late with its heartbeat, and it changes the states of all the replicas if needed.
+
+Because the Librarian services store everything in the A-Hash, the Librarian itself is a stateless service.
+We can deploy multiple Librarians, if we configure them to use the same replicated group of A-Hashes (or the same central A-Hash) then the Bartenders could use any Librarian to achieve the same results, and the Shepherds also could report to any of them.
+One Shepherd should only report to one Librarian, but if that Librarian is offline, it should find another.
+The reports will be registered in the A-Hash as well, so it does not matter which Librarian gets a report.
 

### 2009-03-02 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

@@
 
+For inter-service authorization we can configure the services to only accept connections from trusted services.
+We trust a service if we know its DN.
+For each service we can write trusted DNs in the configuration, or we can periodically get the list from an A-Hash.
+This can be configured with a `\emph{TrustManager}' section in the service configuration, which contains these entries:
+
+\begin{description}
+    \item[DN] is a trusted DN
+    \item[CA] is a DN of a trusted CA: we will trust all certificates issued by this CA
+    \item[DNsFromAHash] is a list of A-Hash URLs (in \textbf{AHashURL}) from where we periodically get a list of trusted DNs (there can be multiple A-Hash URLs listed here in case of one is offline).
+This has some attributes: \textbf{CheckingInterval} specifies how frequently we should get the list from the A-Hash (default: 600); \textbf{ID} specifies the ID of the A-Hash objects, where the list is stored (default: `3').
+\end{description}
+
+The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.
+
+Example TrustManager configuration:
+
+\begin{verbatim}
+    <TrustManager Force="yes">
+        <DN>/DC=eu/DC=KnowARC/O=NIIFI/CN=host/epsilon</DN>
+        <DN>/DC=eu/DC=KnowARC/O=NIIFI/CN=host/phi</DN>
+        <CA>/DC=eu/DC=KnowARC/CN=storage-1233659377.11</CA>
+        <DNsFromAHash CheckingInterval="10">
+            <AHashURL>https://localhost:60000/AHash</AHashURL>
+        </DNsFromAHash>
+    </TrustManager>
+\end{verbatim}
+
 Because of the heavy use of X509 certificates, of course, all the services should have their own certificates, which they can use when connecting to other services.
@@
 \end{verbatim}
+
+This can be put in a seperate file, and then use the \textbf{FromFile} attribute of the ClientSSLConfig section to specify where the file is.
 
@@
 The Librarian also accepts report messages from Shepherds, and stores the contents of these messages in the A-Hash.
-There is one A-Hash object which contains the ID and the expected time of the next heartbeat of the registered Shepherds, and there is one A-Hash object for each Shepherd which stores the states and GUIDs of the files on the Shepherd's storage node.
+There is one A-Hash object which contains the ID and the expected time of the next heartbeat of the registered Shepherds, and there is one A-Hash object for each Shepherd which stores the states and GUIDs of the files on the Shepherd's storage node (see Section~\ref{sub:shepherds} for details).
 The Librarian checks periodically if there is a Shepherd which is late with its heartbeat, and it changes the states of all the replicas if needed.
@@
 \subsection{Configuration} % (fold)
+
+The Librarian has these configuration variables:
+
+\begin{description}
+    \item[AHashURL] is the URL of an A-Hash which the Librarian should use. There could be multiple URLs specified here.
+    \item[HeartbeatTimeout] specifies in seconds how frequently the Sheperds should send reports to the Librarian.
+    \item[CheckPeriod] specifies in seconds how frequently the Librarian should check for late heartbeats.
+\end{description}
+
+An example configuration:
+
+\begin{verbatim}
+    <Service name="pythonservice" id="librarian">
+        <ClassName>storage.librarian.librarian.LibrarianService</ClassName>
+        <AHashURL>https://localhost:60000/AHash</AHashURL>
+        <HeartbeatTimeout>30</HeartbeatTimeout>
+        <CheckPeriod>20</CheckPeriod>
+        <ClientSSLConfig FromFile="clientsslconfig.xml"/>
+        <TrustManager FromFile='librarian_trust_manager.xml'/>
+    </Service>
+\end{verbatim}
 
@@
 
-A Shepherd service is always connected to a file transfer service (e.g.~`HTTP(S)', `FTP(S)', `ByteIO', `GridFTP', etc.).
-For each supported file transfer service we need a backend module which makes the Shepherd capable of communicating with the file transfer service to initiate file transfers, to detect whether a transfer was successful or not, to generate local IDs and checksums, etc.
+A Shepherd service is always connected to a storage element service (e.g.~a web server).
+For each supported storage element service we need a backend module which makes the Shepherd capable of communicating with it to initiate file transfers, to detect whether a transfer was successful or not, to generate local IDs and checksums, etc.
 
 A file in a storage node could be identified with a \emph{referenceID} which is unique within that node.
-If we know the \emph{location} of a file, which is the ID of the Shepherd service (\emph{serviceID}) and the \emph{referenceID}, we could get the endpoint reference (URL) of the Shepherd from the information system, then we could call its \emph{get} method with the \emph{referenceID} and a list of transfer protocols we can use (e.g.~`HTTP', `FTP'), the Shepherd chooses a protocol from this list which it can provide, and create a transfer URL (\emph{TURL}) and returns it along with the \emph{checksum} of the file.
+If we know the \emph{location} of a file, which is the ID of the Shepherd service (\emph{serviceID}, currently the URL of the Shepherd) and the \emph{referenceID}, we could call the Shepherds \emph{get} method with the \emph{referenceID} and a list of transfer protocols we can use, the Shepherd chooses a protocol from this list which it can provide, and create a transfer URL (\emph{TURL}) and returns it along with the \emph{checksum} of the file.
 We could download the file from this \emph{TURL}, and verify it with the \emph{checksum}.
@@
 
-With the \emph{stat} method we can get some information about a replica, e.g.~checksum, GUID, state (`creating', `alive' or `invalid'), etc. The \emph{delete} method removes the file.
+With the \emph{stat} method we can get some information about a replica, e.g.~checksum, GUID, state, etc. The \emph{delete} method removes the replica.
 
@@
 
+The replicas have a state, which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, or `\textbf{invalid}' if it does not exists anymore or has a bad checksum.
+
 % subsection functionality (end)
 
-\subsection{Prototype status and plans} % (fold)
-
-The current implementation of the Shepherd service has a working \emph{get}, \emph{put}, \emph{stat}, \emph{delete} methods, and a method called \emph{toggleReport} which can be used the simulate storage node failure with the Shepherd not reporting to a Librarian.
-There is a separate service which provide a subset of the ByteIO interface, and there is an other separate service which is a basic HTTP server, these are both could be used as file transfer services, both have its backend module for the Shepherd.
-Currently both file transfer services have the problem of using too much memory while transferring files.
-Further plans include better file transfer services and backend modules for third-party file transfer services.
-
-% subsection prototype_status_and_plans (end)
-
-\subsection{Data model} % (fold)
-
-A file of a Shepherd service is referenced by its \emph{referenceID}.
-Each file has a \emph{state} which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, or `\textbf{invalid}' if it does not exists anymore or has a bad checksum.
-Each file has a \emph{localID} which is used in the backend modules.
-
-% subsection data_model (end)
-
 \subsection{Interface} % (fold)
@@
 \begin{description}
-    \item[get(getRequestList)] returns list of (\emph{requestID}, \emph{getResponseData})
-
-    \emph{getRequestList} is a list of (\emph{requestID}, \emph{getRequestData}) where \emph{requestID} is an arbitrary ID used in the reply
-
-    \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).
-
-    \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.
-
-    \item[put(putRequestList)] returns a list of (\emph{requestID}, \emph{putResponseData})
-
-    \emph{putRequestList} is a list of (\emph{requestID}, \emph{putRequestData}) where \emph{requestID} is an ID used for the response
-
-    \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (for additional access policy).
-
-    \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.
-
-    \item[delete(deleteRequestList)] returns a list of (\emph{requestID}, \emph{status})
-
-    \emph{deleteRequestList} is a list of (\emph{requestID}, \emph{referenceID}) pairs selecting the files to remove. The status could be `\textbf{deleted}' or `\textbf{nosuchfile}'
+
+    \item[get(getRequestList)] initiates a download and returns the TURL
+
+    The \emph{getRequestList} is a list of (\emph{requestID}, \emph{getRequestData}) where \emph{requestID} is an arbitrary ID used in the reply, \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).
+
+    The response is a list of (\emph{requestID}, \emph{getResponseData}), where \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.
+
+    \item[put(putRequestList)] initiates an upload and returns the TURL
+
+    The \emph{putRequestList} is a list of (\emph{requestID}, \emph{putRequestData}) where \emph{requestID} is an ID used for the response,     \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (which is currently not used).
+
+    The response is a list of (\emph{requestID}, \emph{putResponseData}), where \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.
+
+    \item[delete(deleteRequestList)] removes a replica.
+
+    The \emph{deleteRequestList} is a list of (\emph{requestID}, \emph{referenceID}) pairs containing the IDs of the files to remove.
+
+    The response is a list of (\emph{requestID}, \emph{status}), where \emph{status} could be `\textbf{deleted}' or `\textbf{nosuchfile}'.
     .
-    \item[stat(statRequestList)] returns a list of (\emph{requestID}, \emph{referenceID}, \emph{state}, \emph{checksumType}, \emph{checksum}, \emph{acl}, \emph{size}, \emph{GUID}, \emph{localID})
-
-    \emph{statRequestList} is a list of (\emph{requestID}, \emph{referenceID}) where \emph{referenceID} points to the file whose data we want to get.
-The method returns all the data the Shepherd know about the replica.
-
+    \item[stat(statRequestList)] returns information about replicas.
+
+    The \emph{statRequestList} is a list of (\emph{requestID}, \emph{referenceID}) where \emph{referenceID} points to the file whose data we want to get.
+
+    The response is a list of (\emph{requestID}, \emph{referenceID}, \emph{state}, \emph{checksumType}, \emph{checksum}, \emph{acl}, \emph{size}, \emph{GUID}, \emph{localID})    
 \end{description}
@@
 
-\subsection{Backend modules} % (fold)
-\label{sub:backend_modules}
-
-The Shepherd could communicate with the file transfer services via backend modules.
-Currently there are two kinds of backend modules, one for the \emph{byteio} service (which is a simple implementation of a subset of the ByteIO interface) and one for the \emph{Hopi} service (which is a simple HED-based HTTP server).
+\subsection{Implementation} % (fold)
+
+The Shepherd is communicating with the storage element services via backend modules.
+Currently there are two backend modules implemented, one for the \emph{Hopi} service (which is a simple HED-based HTTP server), one for the \emph{apache} webserver.
 
@@
 
-In case of the byteio service, there is some small differences.
-The byteio service does not removes the files from the transfer directory, but it calls the backend module via SOAP to notify it that something is happened.
-The byteio backend has one SOAP method called `\textbf{notify}':
-\begin{description}
-    \item[notify(subject, state)] returns `\textbf{OK}' in \emph{notifyResponse}.
-\end{description}
-When this method is called, the byteio backend module notifies the Shepherd that the file is arrived.
-
-All the backend modules should have this common interface which the Shepherd can use to communicate with the file transfer service:
+All the backend modules should have this common interface which the Shepherd can use to communicate with the storage element services:
+
 \begin{description}
@@
     \item[matchProtocols(protocols)] only leave that protocols in the list \emph{protocols} which are supported by this file transfer service.
-    \item[checksum(localID, checksumType)] returns the checksum of the file referenced by \emph{localID}, which checksum is generated by the method \emph{checksumType}.
-
-\end{description}
-
-% subsection backend_modules (end)
+    \item[checksum(localID, checksumType)] generates a \emph{checksumType} checksum of the file referenced by \emph{localID}.
+\end{description}
+
+% subsection implementation (end)
+
+\subsection{Configuration} % (fold)
+
+The Shepherd has these configuration variables:
+
+\begin{verbatim}
+    <Service name="pythonservice" id="shepherd">
+        <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
+        <ServiceID>https://localhost:60000/Shepherd</ServiceID>
+        <CheckPeriod>20</CheckPeriod>
+        <MinCheckInterval>0.1</MinCheckInterval>
+        <StoreClass>storage.store.cachedpicklestore.CachedPickleStore</StoreClass>
+        <StoreCfg>
+            <DataDir>./shepherd_data1</DataDir>
+        </StoreCfg>
+        <BackendClass>storage.shepherd.hardlinkingbackend.HopiBackend</BackendClass>
+        <BackendCfg>
+            <DataDir>./shepherd_store</DataDir>
+            <TransferDir>./shepherd_transfer</TransferDir>
+            <TURLPrefix>https://localhost:60000/hopi/</TURLPrefix>
+        </BackendCfg>
+        <LibrarianURL>https://localhost:60000/Librarian</LibrarianURL>
+        <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
+        <ClientSSLConfig FromFile="clientsslconfig.xml"/>
+    </Service>    
+\end{verbatim}
+
+% subsection configuration (end)
 

### 2009-03-02 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

@@
         <ClientSSLConfig FromFile="clientsslconfig.xml"/>
-        <TrustManager FromFile='librarian_trust_manager.xml'/>
     </Service>
@@
 \subsection{Implementation} % (fold)
+\label{sub:shepherd_implementation}
 
@@
 The Shepherd has these configuration variables:
+
+\begin{description}
+    \item[ServiceID] is the ID of the service, currently it should be the URL of the service, because there is no information system where we can discover services by IDs.
+    \item[CheckPeriod] specifies in seconds how frequently should the Shepherd check the existence and the checksum of all the files on the storage node.
+    \item[MinCheckInterval] specifies in seconds how much time should the Shepherd wait between two subsequent file checks.
+This ensures that the Shepherd does not use up all the resources of the hosting machine.
+After checking the cheksum of a file, the Shepherd will always wait this long before checking the next file, even if this means that checking all the files takes longer than the `CheckPeriod'.
+    \item[StoreClass] specifies which type of store the Shepherd should use for storing its metadata on disk.
+    \item[StoreCfg] can have parameters for the store class, e.g. the local directory to be used.
+    \item[BackendClass] specifes which backend the Shepherd should use for communicating with the storage element service.
+    \item[BackendCfg] has some configuration parameters for the backend class: \textbf{DataDir} and \textbf{TransferDir} is the directories the backend will use (see Section~\ref{sub:shepherd_implementation}), \textbf{TURLPrefix} should be the URL of the storage element service ending with a `/'.
+    \item[LibrarianURL] is the URL of a Librarian where the Shepherd should send the reports.
+    \item[BartenderULR] is the URL of a Bartender whom the Shepherd should connect if it wants to offer a replica for replication or recover a corrupted replica.
+\end{description}
 
@@
 You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}.
+You can remove the link to a file or sub-collection from its parent collection without removing the file or sub-collection itself with the \emph{unlink} method.
 The metadata of a file or collection (e.g.~whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}.
-A \emph{stat} call gives all the information about a file or collection, and you can move (or hardlink) collections and files within the namespace with \emph{move}.
+A \emph{stat} call gives all the information about a file or collection, and you can move (or create hardlinks of) collections and files within the namespace with \emph{move}.
 You can upload an entirely new replica to a file (e.g.~if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.
+You can create and remove mount points with the \emph{makeMountpoint} and \emph{unmakeMountpoint} methods.
+
+The Bartender communicates with the Librarians to get and modify the metadata of files and collections, and it communicates with the Shepherds to initiate file transfers.
+The Bartender also has \emph{gateway modules} which are capable of communicating different kinds of thrid-party storage solutions.
+With these gateway modules it is possible to create mount points within the namespace of the ARC storage, and to access the namespace of the third-party storage through these mount points, which means the user can use the Bartender to get listing of directories on the third-party storage or to get files from the third-party storage.
+Thus the user can use a single client tool to access different kind of storages easily.
+
+The high-level authorization is done by the Bartender.
+It makes decisions based on the identity of the connecting client and the access policy rules of the files and collections.
+The Bartender uses the security framework of the HED to evaluate the requests and the policies, which are genereted by the Bartender from the internal representation of the access rules.
+For more details see Section~\ref{sec:security_implementation}.
 
@@
 
-\subsection{Prototype status and plans} % (fold)
-
-The methods mentioned in the above section are all implemented, but need more error-checking and metadata-checking.
-There are plans of adding new methods, e.g.~a \emph{copy} method or a \emph{glob} method for file pattern matching.
-The current version does not support closed (unmodifiable) collections yet.
-
-% subsection protoype_status_and_plans (end)
-
-\subsection{Data model} % (fold)
-
-The Bartender interface uses mostly Logical Names (\emph{LN}s), which have the syntax of: \verb!<GUID>/<path>!
-where both sides can be omitted (and in the case of a sole GUID we don't need the slash either), e.g.~\verb!afg342/foo!
-is an entry called \verb!foo!
-in the collection with GUID \verb!afg342!; the LN \verb!f36a7481!
-refers to the a file or collection with GUID \verb!f36a7481!; \verb!/vo/dir/stg!
-points to the entry which is reachable from the root collection using the given path; and \verb!/!
-simply refers to the root collection.
-The term `\emph{metadata}' here refers to a list of property-value pairs organized in sections, see the data model description in Section~\ref{sec:data_model}.
-
-% subsection data_model (end)
-
 \subsection{Interface} % (fold)
@@
 \begin{description}
-    \item[putFile(putFileRequestList)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})
-
-    \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the chosen Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where properties could be in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}', policy documents in the `\textbf{policies}' section and any other property-value pairs in the `\textbf{metadata}' section.
-The returned \emph{TURL} is a URL with a chosen \emph{protocol} to upload the file itself, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.
-
-    \item[getFile(getFileRequestList)] returns a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol})
-
-    \emph{getFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the file we want to get, \emph{protocols} is a list of transfer protocols the client supports.
-    In the response \emph{TURL} is the transfer URL using \emph{protocol}, with which we can download the file, \emph{success} could be `\textbf{done}', `\textbf{not found}', `\textbf{is not a file}', `\textbf{file has no valid replica}', `\textbf{error while getting TURL:} reason', etc.
-
-    \item[delFile(delFileRequestList)] returns a list of (\emph{requestID}, \emph{status})
-
-    \emph{delFileRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file we want to delete. The status in response could be `\textbf{deleted}' or `\textbf{nosuchLN}'.
-
-    \item[stat(statRequestList)] returns a list of (\emph{requestID}, \emph{metadata})
-
-    \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the file or collection we want to get information about, and it returns \emph{metadata} which is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (see Section~\ref{sec:data_model})
-
-    \item[makeCollection(makeCollectionRequestList)] returns a list of (\emph{requestID}, \emph{success})
-
-    \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name), in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is true then no more files can be added or removed later), in the `\textbf{policies}' section there could be some access policies, and in the `\textbf{metadata}' section there could be any other metadata in property-value pairs.
-The \emph{success} in the response could be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.
-    \item[unmakeCollection(unmakeCollectionRequestList)] returns a list of (\emph{requestID}, \emph{success}).
-
-    \emph{unmakeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names of the collections we want to remove.
-\emph{success} could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}', `\textbf{failed}: reason'.
-
-    \item[list(listRequestList, neededMetadata)] returns \emph{listResponse}.
-
-    \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection (or file) we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
-
-    \emph{listResponse} is a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples according to the data model of the Librarian (Section~\ref{sec:data_model}), the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}' (because only collections can be listed).
-
-    \item[move(moveRequestList)] returns a list of (\emph{requestID}, \emph{status}).
-
-    \emph{moveRequestList} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) where \emph{sourceLN} is the Logical Name referring to the file or collection we want to move (or just rename) and \emph{targetLN} is the new path, and if \emph{preserveOriginal} is true the \emph{sourceLN} would not be removed, so with \emph{preserveOriginal} we actually creating a hard link.
-The status could be `\textbf{moved}', `\textbf{nosuchLN}', `\textbf{targetexists}', `\textbf{invalidtarget}', `\textbf{failed adding child to parent}', `\textbf{failed removing child from parent}'
-
-    \item[modify(modifyRequestList)] returns a list of (\emph{changeID}, \emph{success})
-
-    \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).
-\emph{success} could be `\textbf{no such LN}', `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}', `\textbf{failed}: reason'.
+
+    \item[putFile(putFileRequestList)] creates new files at the requested Logical Names, chooses a Shepherd and initiates the file transfer, and returns a TURL for each new file.
+
+    The \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the requested Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples which should contain some mandatory properties in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}'.
+
+    The response is a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol}), where \emph{TURL} is a URL with a chosen \emph{protocol} which we can use to upload the file, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.
+
+    \item[getFile(getFileRequestList)] finds the files referenced by the given Logical Name, chooses a replica and initiates the download, and returns a TURL for each file.
+
+    The \emph{getFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the file we want to get, \emph{protocols} is a list of transfer protocols we support.
+
+    The response is a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol}), where \emph{TURL} is the transfer URL using \emph{protocol}, with which we can download the file, \emph{success} could be `\textbf{done}', `\textbf{not found}', `\textbf{is not a file}', `\textbf{file has no valid replica}', `\textbf{error while getting TURL:} reason', etc.
+
+    \item[delFile(delFileRequestList)] removes the files references by the given Logical Names.
+
+    The \emph{delFileRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the files we want to delete.
+
+    The response is a list of (\emph{requestID}, \emph{status}), where the \emph{status} could be `\textbf{deleted}' or `\textbf{nosuchLN}'.
+
+    \item[unlink(unlinkRequestList)] remove a link from a collection without deleting the file or sub-collection itself.
+
+    The \emph{unlinkRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the files and sub-collections we want to unlink.
+
+    The response is a list of (\emph{requestID}, \emph{success}) where \emph{success} can be `\textbf{unset}', `\textbf{no such LN}', `\textbf{denied}' or `\textbf{nothing to unlink}'
+
+    TODO: makeMountpoint, unmakeMountpoint
+
+    \item[stat(statRequestList)] returns all metadata of files, collections or mount points
+
+    The \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names we want to get information about.
+
+    The response is a list of (\emph{requestID}, \emph{metadata}), where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples (see the data model in Section~\ref{sec:data_model}).
+
+    \item[makeCollection(makeCollectionRequestList)] creates new collections.
+
+    The \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name) and in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is `yes' then no more files can be added or removed later).
+
+    The response is a list of (\emph{requestID}, \emph{success}), where \emph{success} can be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.
+
+    \item[unmakeCollection(unmakeCollectionRequestList)] deletes empty collections.
+
+    The \emph{unmakeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names of the collections we want to remove.
+
+    The response is a list of (\emph{requestID}, \emph{success}), where \emph{success} could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}', `\textbf{failed}: reason'.
+
+    \item[list(listRequestList, neededMetadata)] returns the contents of the requested collections.
+
+    The \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
+
+    The response is a a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples (see the data model in Section~\ref{sec:data_model}); the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}'.
+
+    \item[move(moveRequestList)] moves file or collections within the namespace which changes the Logical Name of the file or collection.
+This method only alters metadata, does not move real file data.
+
+    The \emph{moveRequestList} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) where \emph{sourceLN} is the Logical Name referring to the file or collection we want to move (rename) and \emph{targetLN} is the new path, and if \emph{preserveOriginal} is true the \emph{sourceLN} would not be removed, so with \emph{preserveOriginal} we are actually creating a hard link.
+
+    The response is a list of (\emph{requestID}, \emph{status}), where status could be `\textbf{moved}', `\textbf{nosuchLN}', `\textbf{targetexists}', `\textbf{invalidtarget}', `\textbf{failed adding child to parent}', `\textbf{failed removing child from parent}'.
+
+    \item[modify(modifyRequestList)] modifies the metadata of files, collections and mount poins.
+Only the `states', the `policy' and the `metadata' sections can be modified, and there are seperate access control actions for the three.
+
+    The \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).
+
+    The response is a list of (\emph{changeID}, \emph{success}), where \emph{success} could be  `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}' (for an `add' request), `\textbf{denied}', `\textbf{no such LN}', `\textbf{failed}: reason'.
 
@@
 % subsection interface (end)
+
+\subsection{Implementation} % (fold)
+
+For managing collections the Bartender needs only to communicate with a Librarian service, for managing files it needs to communicate with Shepherd services as well.
+When a Bartender wants to create a new replica of a file (or the first replica of a new file) it should find available Shepherd services.
+Fortunately the Librarian has a list of registered Shepherd services, so the Bartender can use that.
+
+TODO: The Bartender can utilize gateway modules for accessing file listings and transfer URL on third-party storages.
+Curretly we have one implementation of a gateway module [...] A gateway module should provide these methods: [...]
+
+TODO: a few sentences about the proxy delegation.
+
+% subsection implementation (end)
+
+\subsection{Configuration} % (fold)
+
+The Bartender has these configuration variables:
+
+\begin{description}
+    \item[LibrarianURL] specifies the URL of one or more Librarian services.
+    \item[ProxyStore] is a local directory where the Bartender will store the delegated proxies.
+    \item[GatewayClass] is the name of the gateway module we want to use in this Bartender.
+    \item[GatewayCfg] contains the configuration parameters of the gateway class: the \textbf{ProxyStore} is the directory where the gateway module could find delegated proxies, and the \textbf{CACertificatesDir} is a directory where the certificates of the trusted CAs are.
+\end{description}
+
+An example configuration:
+
+\begin{verbatim}
+    <Service name="pythonservice" id="bartender">
+        <ClassName>storage.bartender.bartender.BartenderService</ClassName>
+        <LibrarianURL>https://sal1.uppmax.uu.se:60000/Librarian</LibrarianURL>
+        <ProxyStore>/var/arc/spool/proxy_store</ProxyStore>
+        <GatewayClass>storage.bartender.gateway.gateway.Gateway</GatewayClass>
+        <GatewayCfg>
+            <ProxyStore>/var/arc/spool/proxy_store</ProxyStore>
+            <CACertificatesDir>/etc/grid-security/certificates/
+        </GatewayCfg>
+        <ClientSSLConfig FromFile='clientsslconfig.xml'/>
+    </Service>
+\end{verbatim}
+
+% subsection configuration (end)
 

### 2009-03-02 doc/storage_system/arc-storage-documentation.tex
update of the storage documentation

@@
 
+\subsection{Prototype CLI tool} % (fold)
+\label{sub:prototype_cli_tool}
+
 In the first prototype release there is one client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run.
 It is capable of communicating with a given Bartender service, and uploading and downloading TURLs via HTTP.
-
-The methods can be listed with:
+However in order to use proxy certificates or to transfer files with GridFTP protocol, you need to have the ARC client libraries installed.
+
+The credentials of the user and the URL of the Bartender can be configured either with an XML file located at \verb!~/.arc/client.xml! or with environment variables.
+
+Example for the \verb!client.xml! file:
+
+\begin{verbatim}
+    <ArcConfig>
+      <ProxyPath>/Users/zsombor/Development/arc/proxy</ProxyPath>
+      <CACertificatesDir>/Users/zsombor/Development/arc/certs/CA</CACertificatesDir>
+      <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
+    </ArcConfig>
+\end{verbatim}
+
+Example for the environment variables:
+
+\begin{verbatim}
+    export ARC_BARTENDER_URL=https://localhost:60000/Bartender
+    export ARC_KEY_FILE=/Users/zsombor/Development/arc/certs/userkey-john.pem
+    export ARC_CERT_FILE=/Users/zsombor/Development/arc/certs/usercert-john.pem 
+    export ARC_CA_DIR=/Users/zsombor/Development/arc/certs/CA    
+\end{verbatim}
+
+You can specify the number of needed replicas for new file with the \verb!ARC_NEEDED_REPLICAS! environment variable.
+
+The \verb!arc_storage_cli! has its built-in help. The methods can be listed with:
+
 \begin{verbatim}
@@
   Supported methods: stat, make[Collection], unmake[Collection], list, move,
-    put[File], get[File], del[File]
+    put[File], get[File], del[File], pol[icy], unlink,
+    credentialsDelegation, removeCredentials
 \end{verbatim}
@@
 
-Uploading, downloading and stat files:
+Here is an example of uploading, downloading and stat files:
 
@@
 - The md5 checksum of the file is 9a9dffa22d227afe0f1959f936993a80
-- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
 - Calling the Bartender's putFile method...
@@
 $ arc_storage_cli stat /tmp/testfile
-- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
 - Calling the Bartender's stat method...
@@
 $ arc_storage_cli get /tmp/testfile newfile
-- ARC_BARTENDER_URL environment variable not found, using http://localhost:60000/Bartender
 - Calling the Bartender's getFile method...
@@
 
-You can find more examples in the SVN\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage/README}}.
-
-There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, \url{http://fuse.sourceforge.net/}} module and Windows Shell Extensions\footnote{\url{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx}} to be able to mount the ARC storage namespace into the local filesystem namespace, and use it with the commands of the operating system.
+% subsection prototype_cli_tool (end)
+
+\subsection{FUSE module} % (fold)
+\label{sub:fuse_module}
+
+TODO: what is the FUSE module, how does it work, how can we use it
+
+% subsection fuse_module (end)
 
@@
 
-\section{Integrating third-party storage solutions} % (fold)
-\label{sec:integrating_third_party_storage_solutions}
-
-To integrate existing files on a third-party storage system to our namespace thus make them accessible through the interface of the ARC storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages.
-It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC storage (Section~\ref{sec:data_model}).
-The files here would be referenced by a path which is local in the namespace of the third-party storage.
-The interface could be something like this:
-\begin{description}
-    \item[get(getRequest)] returns list of (\emph{requestID}, \emph{getResponseData})
-
-    very similar to the \emph{get} method of the Shepherd
-
-    \item[put(putRequest)] returns a list of (\emph{requestID}, \emph{putResponseData})
-
-    very similar to the \emph{put} method of the Shepherd
-
-    \item[delete(deleteRequest)] returns a list of (\emph{requestID}, \emph{success})
-
-    very similar to the \emph{delete} method of the Shepherd
-
-    \item[stat(statRequest)] returns a list of (\emph{requestID}, \emph{statResponse})
-
-    very similar to the \emph{stat} method of the Shepherd
-
-    \item[list(listRequest)] returns \emph{listResponse}
-
-    very similar to the \emph{list} method of the Bartender
-
-    \item[move(moveRequest)] returns a list of (\emph{requestID}, \emph{status})
-
-    very similar to the \emph{move} method of the Bartender
-
-\end{description}
-This interface has some methods similar to the Shepherd and some other methods similar to the Bartender.
-% section integrating_third_party_storage_solutions (end)
+\section{Grid integration} % (fold)
+\label{sec:grid_integration}
+
+TODO: what is the ARC DMC, how does it work
+
+% section grid_integration (end)
 

### 2009-03-02 doc/storage_system/arc-storage-documentation.tex
Added description of how Librarian reads from and writes to replicated A-Hash

@@
 
-TODO: here comes some sentences about how the Librarian reads from any A-Hash and finds the master A-Hash for writing.
+%TODO: here comes some sentences about how the Librarian reads from any A-Hash and finds the master A-Hash for writing.
+The replicated A-Hash deployment is based on a single master, multiple clients scenario (see Section~\ref{sec:a_hash} for details).
+This means that the Librarian can read from any A-Hash, while it can only write to the master A-Hash.
+A list of A-Hashes, master and clients, is stored in an A-Hash object.
+The Librarian periodically gets this list to maintain its own list of A-Hash services.
+If the Librarian tries to write to the replicated A-Hash and the master is down, the Librarian will try to get a new list from one of the clients and retry the write request once before giving up.
+In case of reading from the replicated A-Hash, the Librarian will randomly loop through the list of clients until the read request is successful or until there are no more clients available.
 

### 2009-03-02 doc/ws-programming-tutorial/doc/tex_appendix/appendix.tex
Checked the whole document up to the Sec Echo Web Service. Started writting more about the Sec Echo Web Service. Added XSD of the XML configuration file to the appendix.

@@
 
+Questions:
+\begin{itemize}
+ \item The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - Web Service   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - Web Service
+ \item Does HED realises a replication Server which delegates incoming request to several threads which are communicating to the client on other free ports (I guess that not).
+i.e.
+Front end server is running on port 60000.
+Request will be delegated to thread which creates a new socket on port 54305.
+The communication get delegated and the front end is ready for new requests.
+\end{itemize}
+
 \section{Important XSD files}
+\lstsetJUSTXML
+\lstinputlisting
+	[
+	label=lst:arched_arcecho_xml,
+	caption={[HED configuration. Filename: arcecho\_no\_ssl.xml]
+	\textbf{HED configuration. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	]
+{../../../src/hed/libs/loader/mcc.xsd}
+
+\task{RENAMING HED configuration file into~ARC service configuration file - proposed by Marek}
+\task{(echo service) client configuration file - proposed by Marek}
 

### 2009-03-02 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Checked the whole document up to the Sec Echo Web Service. Started writting more about the Sec Echo Web Service. Added XSD of the XML configuration file to the appendix.

@@
-\chapter{The echo service}
+\chapter{The Echo Web Service}
 
-The next service to be presented is the echo service.
+The next service to be presented is the Echo Web Service.
 Its task is to return the content of the received message back to the client.
 Furthermore, the service provides two additional operations called ``ordinary'' and ``reverse''.
-Depending on the operation chosen by the client, the message will returned unmodified or it is shown with its reversed.
+Depending on the operation chosen by the client, the message gets returned unmodified or reversed.
 The goal of this chapter is to offer a deeper knowledge about message processing.
@@
 
-Web services can be specified (not implemented) by a meta language called \textit{WSDL} (Web Services Description Language). 
+Web Services can be specified (not implemented) by a meta language called \textit{WSDL} (Web Services Description Language).
 Beside the description of the service, WSDL renders the possibility that services can be discovered by service brokers.
@@
 WSDL describes the structure of the messages, the intended use of the messages, the supported protocols and the endpoint of the service (URL).
-For that last reason, every instance of a service must be produced a distinct WSDL file for - or - it is automatically prepared from the information that the server has available once it is implemented.
+\task{For that last reason, every instance of a service must be produced a distinct WSDL file for - or - it is automatically prepared from the information that the server has available once it is implemented.
+- ?MG }\\
 
-The WSDL file that specifies the here presented echo service is shown in listing~\ref{lst:echo_wsdl}. 
+The WSDL file that specifies the here presented Echo Web Service is shown in listing~\ref{lst:echo_wsdl}. 
 It is composed of five main elements:
@@
 	\item \textbf{Types} --- The element \textit{types} contains the description of the message structures and is written in the language XSD (XML Schema Definition).
-The data types used for the messages of the echo service are defined starting with the  line~\ref{lst_code:echo_wsdl_types}.
+The data types used for the messages of the Echo Web Service are defined starting with the  line~\ref{lst_code:echo_wsdl_types}.
 
 	\item \textbf{Message} --- Within the element \textit{message} a subset of the previously defined types are assigned to be a message.
-In case of the echo service two messages are designated: echoRequest and echoResponse, compare lines line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
+In case of the Echo Web Service  two messages are designated: echoRequest and echoResponse, compare lines line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
 
@@
 	\item \textbf{Service} --- Within the last WSDL element \textit{service} the endpoint of the service gets specified.
-As to be seen in line~\ref{lst_code:echo_wsdl_service} the endpoint of the echo service is assigned to \textit{http://localhost:60000/echo}.
+As to be seen in line~\ref{lst_code:echo_wsdl_service} the endpoint of the Echo Web Service  is assigned to \textit{http://localhost:60000/echo}.
 
@@
 
-The optional SOAP Fault element is used to indicate error messages.
+The optional SOAP Fault element is used to indicate error messages in case of the port type \textit{solicit-response} (input, output, fault).
 If a \textit{Fault} element is present, it must appear as a child element of the \textit{Body} element. A \textit{Fault} element can only appear once in a SOAP message.
@@
 
-\task{Source of the above two tables?}
+\task{Source of the above two tables? Extension of the second table? Numbers??}
 
@@
 The source code of the C++ file is shown in listing~\ref{lst:echo_service_cpp}.
-Again, the header file will be set aside for it contains only redundant information.
+The header file will be set aside for it contains only redundant information. 
 It can be reread in the source directory that is accompanying the tutorial.
-For the source code of echo service basically three changes are required for on the code for the timeservice webservice of the previous chapter.
-The first change concerns the constructor and is found at line~\ref{lst_code:echo_cpp_prefix}.
-Two strings will be extracted of the HED configuration file which are later used to envelope the message before it is returned.
-Furthermore, a new method \textit{makeFault} in line~\ref{lst_code:echo_cpp_makeFault} has been introduced which is able to create fault messages and is called by the method \textit{process}.
-Within that method two new code fragments are now extracting and analyzing the incoming message, to be seen at lines~\ref{lst_code:echo_cpp_extracting} and \ref{lst_code:echo_cpp_analyzing}.
+Regarding to the source code of the Echo Web Service basically three changes of the Time Web Service source code will be required.
+The first change concerns the constructor and can be found at line~\ref{lst_code:echo_cpp_prefix}.
+Two strings are extracted of the server configuration file which later are used to envelope the message before it is returned.
+Furthermore, a new method \textit{makeFault} in line~\ref{lst_code:echo_cpp_makeFault} has been introduced which is able to create the fault messages and is called by the method \textit{process}.
+Within the method \textit{process} two new code fragments are now extracting and analyzing the incoming message, to be seen at line~\ref{lst_code:echo_cpp_extracting} and \ref{lst_code:echo_cpp_analyzing}.
 In case something unexpected happens both will create a SOAP fault message.\\
@@
 Again, the HED is configured by an XML file and is launched as introduced earlier.
-The HED configuration file is shown in Listing~\ref{lst:echo_hed_configuration_xml}.
+The server configuration file is shown in Listing~\ref{lst:echo_hed_configuration_xml}.
 The only change concerns the lines~\ref{lst_code:echo_hed_configuration_xml_pre} and \ref{lst_code:echo_hed_configuration_xml_pre} which specify the \textit{prefix} and \textit{suffix}.
-The parameters will be extracted in the constructor of the service and explain nicely how settings can be passed from the HED configuration file towards the service.
+The parameters will be extracted in the constructor of the service and explain nicely how settings can be passed from the server configuration file towards the service.
 
@@
 	label=lst:echo_hed_configuration_xml, float=htb,
-	caption={[HED configuration file for the echo service. Filename: arched\_echoservice.xml]
-	\textbf{HED configuration file for the echo service. Filename: arched\_echoservice.xml}}
+	caption={[Server configuration file for the echo service. Filename: arched\_echoservice.xml]
+	\textbf{Server configuration file for the echo service. Filename: arched\_echoservice.xml}}
 	]
@@
 Firstly, because the additional arguments need to be filled.
-Secondly, because of the extra effort to interact with the (here human) user to collect the data to forward to the echo service.
+Secondly, because of the extra effort to interact with the (here human) user to collect the data to forward to the Echo Web Service.
 The parameters are now passed by the command line, to be seen in the lines followed by line~\ref{lst_code:echo_client_cpp_arguments}.
 That code is independent from the ARC library.
-The creation of the SOAP client is unchanged.
-The request is created in lines subsequent line~\ref{lst_code:echo_client_cpp_request}.
+The creation of the SOAP client is unchanged and the request is created in the lines subsequent line~\ref{lst_code:echo_client_cpp_request}.
 It is processed in the code after line~\ref{lst_code:echo_client_cpp_response} in which also the response is received.
@@
 The usage of the client is presented in Listing~\ref{lst:invokation_echoservice}.
-Three cases of echo service responses are displayed: ordinary operation, reverse operation and SOAP fault due to an unknown operation.
+Three cases of Echo Web Service responses are displayed: ordinary operation, reverse operation and SOAP fault due to an unknown operation.
 
@@
 
-A few examples for the exchanged messages are shown in the Listings~\ref{lst:echoservice_request}, \ref{lst:echoservice_response} and \ref{lst:echoservice_fault}.
+A few examples of exchanged messages are shown in the Listings~\ref{lst:echoservice_request}, \ref{lst:echoservice_response} and \ref{lst:echoservice_fault}.
 Listing~\ref{lst:echoservice_request} illustrates a request message using the operation \textit{reverse}.

### 2009-03-02 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Checked the whole document up to the Sec Echo Web Service. Started writting more about the Sec Echo Web Service. Added XSD of the XML configuration file to the appendix.

@@
 It is however beneficial to have some basic comprehension of the programming language C++, the file format XML and any regular UNIX shell.
-To start, the user shall have an installation of ARC-1, as it is performed with current RPM or .deb packages\footnote{see http://wiki.nordugrid.org for explicit instructions for your setup}.
+To start, the user shall have an installation of ARC-1, as it is performed with current RPM or .deb packages\footnote{see http://wiki.nordugrid.org for explicit instructions for setup}.
 
@@
 At the time of writing, the endpoint of the HED will be a SOAP based webservice.
-SOAP is a long established XML-based standard for web service communication.
+SOAP is a long established XML-based standard for Web Service communication.
 ARC abstracts from this format, but nevertheless it is useful to be aware of
@@
 
-The HED which was already mentioned in the section above. As an essential part of the ARC middleware it provides the hosting of various services at the application level.
-Its modularity allows its dynamic extension.
+The HED which was already mentioned in the section above is an essential part of the ARC middleware. 
+Its task is to provide the hosting of various services at the application level and is based on the idea modularity.% allows its dynamic extension.
+The HED consists of compononents called Message Chain Components (MCC), services and several modules, which shall aid the programmer to simplify the development of the Web Services: Config, Loader, Logging, XMLNode.
+The MCCs are ordered within a layered structure. 
+Each MCC provides a certain functionality such as communication between two applications (MCC TCP), secure data transfer (MCC TLS) or client-server architecture (MCC HTTP).
 Figure~\ref{fig:HED_internal} illustrates a typical setup of a server-sided HED.
-The HED consists of a few compononents called Message Chain Components (MCC).
-These are ordered within a layered structure and several \textcolor{urgent}{modules}, which shall aid the programmer to simplify the development of the web services: Config, Loader, Logging, XMLNode.
 
@@
 	\includegraphics[width=14cm]{tex_introduction/HED.pdf}
-	\mycaption{Example of HED structure with an Arex service and a file service.\task{ Replace image by an easier one!} }{\task{Write more stuff}\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}}
+	\mycaption{Example of HED structure with an Arex service and a file service.\task{ Replace image by an easier one!} }{\task{Write more stuff}\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - Web Service   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - Web Service}}
 \label{fig:HED_internal}
@@
 
-The MCCs are passing the incoming and outgoing messages to the upper or the respective lower layer. \task{which means? SM}
-The lowest layer has to be a transport protocol like TCP which enables the transmission between the two applications, the client and the server.
-Technically, there are lower leves, e.g. the Internet Protocol (IP) enables the transmission between two computers and is a technical requirement for the TCP layer.
-But IP is not directly modelled by the HED.
-TCP is well suited for the application because it provides a reliable and ordered delivery of a byte stream which is important to ensure security.
+As to be seen, the MCCs are connected which each other and passing incoming and outgoing message to the upper or the respective lower layer.
+In case of outgoing messages a MCC is wrapping the data of the upper layer as a payload into their own protocol (while incomming message will be unwrapped).
+The lowest layer has to be a transport protocol like TCP which enables the transmission between the two applications.
+Tehchnically there even lower layers required needed for the communication but they are realised within the operating system and the  network card
+TCP is well suited for the application because it provides a reliable and ordered delivery of a byte stream.
 Above the TCP layer is the Hypertext Transfer Protocol (HTTP). 
-In case a security layer is desired for authentication, the TLS (Transport Layer Security formerly known as SSL, Secure Sockets Layer) is inserted between the TCP and HTTP layers.
+In case a security layer is desired (i.e. for authentication), the TLS (Transport Layer Security formerly known as SSL, Secure Sockets Layer) has to be inserted between the TCP and HTTP layers.
 The HTTP provides the client-server architecture.
@@
 The multiplexing layer can also be defined within the HED and is called Plexer.
-Depending on the path of the URL the Plexer passes messages to a defined SOAP service which again passes the message to the Web Service.
+Depending on the path of the URL the Plexer passes messages to a defined SOAP service which again passes the message to the Web Service~\cite{QIANG_2005}. \task{Add text here}\\
+%In my understanding (because plexer is not my initial idea), plexer (as a sort of hub) can be put at both after http or soap.
+If you put after soap,then you are doing switching for different web services.
+While I put it after http, because somehow, I need to switch the message after http-processing, in more detail, one branch for soap and then web service/s, one branch for a non-soap service/server (the SP Service is not a service based on SOAP, actually it is based on http, so it is kind of web application).
 
-The structure of the HED can be configured freely by modifying the HED configuration file.
+The structure of the HED can be configured freely by modifying the server configuration file.
 A first impression how the HED can be configured shall be given in the following example which corresponds to the HED shown in figure~\ref{fig:HED_internal}.
-It uses the ARC echo service~\cite{QIANG_2005} that is shipped for testing purposes with the ARC source code.
+It uses the ARC echo service that is shipped for testing purposes with the ARC source code.
 
@@
 
-\task{Where is the xsd file for HED configuration files??}
 The first example shall give a basic understanding in how to configure the HED for setting up a novel service.
 Here, the task is to setup the regular ARC Echo Web Service.
-On the command line, all that needs to be done is to invoking the \textit{arched}\footnote{read: arc-hed} daemon with a fitting HED configuration file.
-An example is shown in Listing~\ref{lst:arched_arcecho_xml}.
-It is written in XML and its structure is based on the \textcolor{urgent}{??} schema.
+On the command line, all that needs to be done is to invoking the \textit{arched}%
+\footnote{\task{refer to the appendix in which the documentation to arc-hed is specified}}
+daemon with a fitting server configuration file.
+Such a fitting server configuration file is shown in Listing~\ref{lst:arched_arcecho_xml}.
+It is written in XML and its structure is based on within the xsd-schema \textcolor{urgent}{URL?} .
 
 \lstsetARCHEDXML
-
 \lstinputlisting
@@
 Several attributes can be defined here but at least the XML version should be specified.
-The configuration as a whole is encapsulated by the \textit{<ArcConfig>} tag.
-It also performs the setting of namespace abbreviations, which are expected to be mostly invariant across all server installations.
+The configuration as a whole is encapsulated by the element \textit{ArcConfig}.
+It also performs the setting of namespace abbreviations, which are expected to be mostly invariant across all server installations.% ABBREVIATIONS??
+- MG
 The \textit{Server} element, to be found at line~\ref{lst_code:arched_arcecho_Server}, provides basic settings for the daemon such as location of the Pid-file or the Log-file.
@@
 The names have to correspond to the name of the dynamic libraries within the path defined in the \textit{ModuleManager}.
-Finally, the chain of the components is declared within the element \textit{Chain}.
-Every component inside the chain is represented by an element called \textit{Component} which has at least a \textit{name} attribute.
-Its important that a plugin (*.so file found in the paths previously specified) has been loaded which provides the implementation of the component's functionality.
-The further configuration of the individual web service depends on the individual component's functionality.
-The \textit{Service} tag may be used to specify additional elements that the web service may then inspect.
-
-%\task{Internet Protocol is required.}
-
-We need to distinguish between the plugins and the components.
-The plugins refer to shared libraries that can be integrated as modules with the HED to extend its functionality.
-The Components refer to the implementation of web services within any of those libraries.
+Finally, the chain is declared within the element \textit{Chain} which creates the layered structure.
+It is composed out of the elements \textit{Component}, \textit{Plexer} and \textit{Service}.
+Every element inside has at least a \textit{name} attribute.
+Regarding the components and the services a plugin (*.so file found in the paths previously specified) has to be loaded which provides the implementation of its functionality.
+The further configuration depends on the individual functionality.\\
+%
+% within the chain ... 3 types of elements
+%
+%Its needed to distinguish between the plugins and the components.
+%The plugins refer to shared libraries that can be integrated as modules with the HED to extend its functionality.
+%The Components refer to the implementation of Web Services within any of those libraries.
+%
+%
 A \textit{Chain} will determine the path that an event takes to be noticed.
@@
 The port can be an arbitrary number.
-However, the number 60000 is commonly used for a HED providing the A-REX web service (the one organising the computation on a site) and the number 50000 is found commonly used for a HED running web services for storage.
+However, the number 60000 is commonly used for a HED providing the A-REX Web Service (the one organising the computation on a site) and the number 50000 is found commonly used for a HED running Web Services for storage.
 This way, one can organise a single server to offer multiple HED instances.
-The stream which is received by this component is passed to the component defined by the \textit{next} tag.
+The stream which is received by this component is passed to the component defined by the \textit{next} attribute.
 In the present case, it will be passed to the component with the \textit{id} attribute named \textit{http}.
 In this way, the message is be passed to higher layers until it reaches the \textit{Plexer}.
-Depending on the path of the URL which was demanded by the HTTP, the Plexer multiplexes the request to the right web service, the right function to execute according to the component IDs and plugin names.
-The plexer will thus allow to have multiple web service listen to the same port.
-The path, declared within the \textit{next} element, corresponds to a regular expression and leads to the service with the id named \textit{echo}. 
+Depending on the path of the URL which was demanded by the HTTP, the Plexer multiplexes the request to the right Web Service, the right function to execute according to the component IDs and plugin names.
+The plexer will thus allow to have multiple Web Service listen to the same port.
+The path, declared within the \textit{next} attribute, corresponds to a regular expression and leads to the service with the id named \textit{echo}. 
 Hence, the service will now be reachable under URL \textit{http://localhost:60000/echo}.
@@
 They set the characters which will enclose the echoed string.
-\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - WEB SERVICE   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - WEB SERVICE}
+\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - Web Service   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - Web Service}
 
-To create the daemon corresponding to the defined HED configuration file, ARC provides the program called \textit{arched}.
+To create the daemon corresponding to the defined server configuration file, ARC provides the program called \textit{arched}.
 The manner of invokation is to been seen in Listing~\ref{lst:arcecho_arched_invokation}.
@@
 In this example, the server and the client are running on the same computer such that the hostname \textit{localhost} can be utilized. 
-For debugging it is advisable to check the logfile which was assigned inside the HED configuration file.
+For debugging it is advisable to check the logfile which was assigned inside the server configuration file.
 It is commonly found at /var/log/arched.log
 
-The following chapters explain how to develop custom web services.
+The following chapters explain how to develop custom Web Services.
 The first one will be the simple Time Web Service.

### 2009-03-02 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
Checked the whole document up to the Sec Echo Web Service. Started writting more about the Sec Echo Web Service. Added XSD of the XML configuration file to the appendix.

@@
-\chapter{SEC echoservice}
+\chapter{SEC Echo Web Service}
 
-The TLS echo service, presented in the previous chapter enables the adminstrator the control the access to the grid.
+The TLS echo service, presented in the previous chapter enables the adminstrator to control the access to the grid.
 Due to a certificate owner may be a member of a large VO it is sometimes recommended to have a more fine grained security within the daemon.
@@
 The security framework of ARC is implemented based on the same concept of modularity as HED. 
-To extend the security of a MCC one only has to modify the service configuration file.
-This is to be done by including a \textit{SecHandler} element into a component or service which again has to contain a \textit{pdp} element.
+To extend the security of a MCC one only has to modify the server configuration file.
+This is to be done by including a \textit{SecHandler} element into a component or service which again has to contain a \textit{PDP} element.
 A small example of a chain component is shown in Listing~\ref{lst:sec_service_component_xml}.
-The \textit{SecHandler} is able to intervene the message flow
-by: manipulating the message content, the attributes or just disrupt the message. 
+The \textit{SecHandler} is able to intervene the message flow by: Verifying, parsing the message attribute, and using the attributes for authorization.
+
+%
+% Weizhong: Verifying, parsing the message attribute, and using the attributes for authorization
+%
 % Attributes done by: identity.map, delegation.collector
@@
 \end{lstlisting}
-Each MCC or Service usually implement two queues of SecHandlers --- one for incoming messages and one for outgoing called
-``incoming'' and ``outgoing'' respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well~\cite{QIANG_2008}.
+Each MCC or Service usually implement two queues of SecHandlers --- one for incoming messages and one for outgoing messages.
+% called ``incoming'' and ``outgoing'' respectively. 
+All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well~\cite{QIANG_2008}.
 In order to decide which message shall be modified by the \textit{SecHandler} so called PDP (Policy Decision Point) are used.
@@
 	\hline
- 	\textbf{identity.map}          & Maps the identity of the user to an identity of the workstation.\\
- 	\textbf{arc.authz}             & Restricts the access to the next layer.\\
- 	\textbf{delegation.collector}  & Collects the information for the delegation policy (Related to \textit{delegate.pdp}). \\
- 	\textbf{usernametoken.handler} & Generates the WS-Security username-token and adds it into the SOAP header of outgoing
-					 messages. In case of incoming messages the SecHandler extracts the WS-Security username-token out of it. \task{Figure out what the username-token does...}\\
+ 	\textbf{identity.map}         & Maps the identity of the user to an identity of the workstation which is realised 
+					by a message attribut.\\
+ 	\textbf{arc.authz}            & Grants access to the next layer.\\
+ 	\textbf{delegation.collector} & Processes a request for another resource with proxy certificate that has delegation policy embedded i.e.
+a client establishes a connection to that service and requests access to a certain resource.
+In case the service is willing to grant the access, the delegation policy inside the proxy certificate from the client will be stored on the server which is providing the requested resource such that the client is now able to use it.\\% Question how does the client creates the request
+ 	\textbf{usernametoken.handler}& Generates the WS-Security username-token and adds it into the SOAP header of outgoing
+					 messages.
+In case of incoming messages the SecHandler extracts the WS-Security username-token out of it.
+WS-Security extends SOAP and describes i.e.
+how to attach signatures and encrypton header to SOAP messages.\\
 	\hline
@@
   \end{table}
+% WS-Security describes how to attach signatures and encryption headers to SOAP messages.
+In addition, it describes how to attach security tokens, including binary security tokens such as X.509 certificates and Kerberos tickets, to messages.
 
@@
 	\hline
- 	\textbf{allow.pdp}           & bla \\
- 	\textbf{simplelist.pdp}      & blub.\\
- 	\textbf{arc.pdp}             &\\
+ 	\textbf{allow.pdp}           & Always permits the underlying SecHandler to be applied.\\
+ 	\textbf{simplelist.pdp}      & Applies the SecHandler in case the identity corresponds with the listed ones.\\
+ 	\textbf{arc.pdp}             & \\
  	\textbf{delegate.pdp}        & \task{Explain that well}\\

### 2009-03-02 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Express checkout... Finished the basic TLS chapter content and started writting about the Sec echo client...
Checked the whole document up to the Sec Echo Web Service. Started writting more about the Sec Echo Web Service. Added XSD of the XML configuration file to the appendix.

@@
-\chapter{The time webservice}
+\chapter{The Time Web Service}
 
-The first implementation of a web service will be a simple time service. 
+The first implementation of a Web Service will be a simple time service. 
 When the service receives a request of a client, it will respond with a string containing the system time.
 For simplicity the service will not parse the request but always return the current time.
-It is therefore easier than the example of an Echo web service, that also needs to learn (by interpreting the request)
+It is therefore easier than the example of an Echo Web Service, that also needs to learn (by interpreting the request)
 about what should be echoed.
@@
 In ARC the services are implementated as plugins such that they can easily be included or excluded by modifiying the HED configuration file as it was to be seen in Listing~\ref{lst:arched_arcecho_xml}.
-Listing~\ref{lst:time_arched_cpp} contains the C++ implementation of the time service plugin.
-The proper class definition of the object \textit{TimeService} is defined in the corresponding header file which is not shown here.
-\task{show the header file, or at least the important bits of it - sm}
-It inherits from an object called \textit{Service} which itself is defined in the ARC library.
+The class which is needs to be implemented for the Web Service inherits from the class \textit{Service} which itself is defined in the ARC library.
+The proper class definition of the object \textit{TimeService} is defined in the corresponding header file which is shown in Listing~\ref{lst:time_arched_h}.
+\lstsetCPP%
+\lstinputlisting[
+	label=lst:time_arched_h,
+	caption={[Header file of the Time Web Service. Filename: timeservice.h]
+	\textbf{Header file of the Time Web Service. Filename: timeservice.h}}
+	]%
+{../src/services/timeservice/timeservice.h}%
+As to be seen, the class \textit{Service} renders the possibility to implement the method \textit{process}.
+Listing~\ref{lst:time_arched_cpp} contains the C++ implementation of the Time Web Service plugin.
 The key element for each plugin is the struct \textit{PluginDescriptor} named \textit{PLUGINS\_TABLE\_NAME} to be seen in line~\ref{lst_code:time_cpp_ptn}. 
@@
 The function which returns the pointer of the instance is desclared in line~\ref{lst_code:time_cpp_get_service}. 
-The interface demands an object \textit{PluginArgument} as a parameter.
+Within the interface an object \textit{PluginArgument} is demanded as a parameter. 
 But since the plugin that is now written inherits from the class \textit{Service}, one can expect the \textit{PluginArgument} to be of a special subclass \textit{ServicePluginArgument}.
+%	In the first line of the service the object is downcasted with the operator dynamic_cast
+The \textit{ServicePluginArgument} provides access to the server configuration and to the message chain of the service which later may be useful.\\
 
-%	In the first line of the service the object is downcasted with the operator dynamic_cast
-The \textit{ServicePluginArgument} provides access to the HED 
-%       operator Config* (void) { return config_; };
-%       operator ChainContext*   			an interface to manipulate chains stored in Loader
-%
-%
-%
-%
- \textcolor{white}{newline}
-\lstsetCPP%
-\lstinputlisting[
-	label=lst:time_arched_cpp,
-	caption={[C++ implementation of the time service. Filename: timeservice.cpp]
-	\textbf{C++ implementation of the time service. Filename: timeservice.cpp}}
-	]%
-{../src/services/timeservice/timeservice.cpp}%
-configuration and the message chain of the service. The time service uses the namespace \textit{ArcService}.
+The Time Web Service uses the namespace \textit{ArcService}.
 The constructor can be found in line~\ref{lst_code:time_cpp_constructor} and the deconstructor in line~\ref{lst_code:time_cpp_deconstructor}.
@@
 
-The method \textit{process} is required by the inherited class \textit{Service}.
-It is defined in line~\ref{lst_code:time_cpp_process} and realises the desired service,
+%
+%
+% \textcolor{white}{newline}
+\lstsetCPP%
+\lstinputlisting[
+	label=lst:time_arched_cpp,
+	caption={[Implementation of the Time Web Service. Filename: timeservice.cpp]
+	\textbf{Implementation of the Time Web Service. Filename: timeservice.cpp}}
+	]%
+{../src/services/timeservice/timeservice.cpp}%
+
+The method \textit{process} is defined in line~\ref{lst_code:time_cpp_process} and realises the desired service,
 here: the calling of functions provided by system libraries to determine the current time. 
@@
 The object \textit{PayloadSOAP} provides enough functionality to create an XML message that is conform to the SOAP protocol.
-The appearance of the generated request and its response will be discussed later on page~\pageref{lst:time_client_request}.
+The appearance of the generated request and its response will be discussed later on page~\pageref{lst:time_client_request}.\\
 
@@
 Within the line~\ref{lst_code:time_cpp_process_message_payload} the payload is stored in the message. 
-When nothing unexpected happened while the message is processed, the function returns an \textit{MCCStatus} set to \textit{OK} as seen in line~\ref{lst_code:time_cpp_return}.
+When nothing unexpected happened while the message is processed, the function returns an \textit{MCCStatus} set to \textit{OK} as seen in line~\ref{lst_code:time_cpp_return}.\\
 
@@
 For his example, the installation direction was set to \textit{/tmp/arc/tutorial/lib}.
-It was also specified by the \textit{ModuleManager} of the HED configuration file.
-When this document was prepared, the paths cannot be changed at the runtime at the server.
+It was also specified by the \textit{ModuleManager} of the server configuration file.
+Once the server is running, the paths cannot be changed.\\
 
-A suitable HED configuration file which engages the time service is to be seen in Listing~\ref{lst:time_arched_xml}.
+A suitable server configuration file which engages the time service is to be seen in Listing~\ref{lst:time_arched_xml}.
 A new \textit{Path} element has been introduced into the element \textit{ModuleMananger} in line~\ref{lst_code:arched_time_moduleManager}.
@@
 a rule has been introduced which redirects the request to the time service, if the path of URL is \textit{time}. 
-Please note that the URL setting is case sensitive.
+Please note that the URL setting is case sensitive and corresponds to a regular expression.\\
 
@@
 
-To start the service, the \textit{arched} command which as shown in Listing~\ref{lst:arched_timeservice_ksh} has to be used.
-It is recommended to examine the log file \textit{/var/log/arched.log} if the HED configuration file has been modified to  assert the start was successful
+In order to start the service, the \textit{arched} command which as shown in Listing~\ref{lst:arched_timeservice_ksh} has to be used.
+It is recommended to examine the log file \textit{/var/log/arched.log} if the server configuration file has been modified to  assert the start was successful
 
@@
 
-\section{Client}
+\section{Client}\label{sec:timeservice_client}
 
-Up to now we have implemented a server program, the code that runs on a particular machine and waits for other programs that ask it for the time.
-What is still needed is the client.
-If the server spoke the http protocol, then we could use our web browser.
+Up to now we have implemented a server program, the code that runs on a particular machine and waits for other programs to ask it for the time.
+What is still needed is the client to send the request to the server and interprete the response of it.
+%If the server spoke the http protocol, then we could use our web browser.
 But it does not - wait, stop - in a way it does, but the server's SOAP envelope renders the reply rather awkward to read.
-\task{someone please check - SM}  The server exists only once.
+\task{someone please check - SM; I think you are wrong (I like to read SOAP messages) :-) - MG}
+%     -- Even a Web Browser is operating with HTTP it is not able to produce the required SOAP body and/or header...
+% \task{The server exists only once.
 But there may be many many clients requesting it in parallel - on the same machine or across the internet.
-The client may be directly controlled by a human (like the web browser) or be some cron job or other software initiating the request.
+The client may be directly controlled by a human (like the web browser) or be some cron job or other software initiating the request.} - I think that's not correct...
 
@@
 Likewise to the HED, it will also be created dynamically using the MCCs.
-But as one will see, the creation of the protocol stack on client side is very simple.
+But as one will see, the creation of the protocol stack on client side is very simple.\\
 
@@
 Due to the configuration of the the \textit{Logger} in the next lines, the messages passed to it will be redirected to the standard error stream.
-For the components of the protocol stack, which  are likewise loaded \textcolor{white}{newline}
+For the components of the protocol stack, which  are likewise loaded
+%\textcolor{white}{newline}
 \lstsetCPP
@@
 The URL of the service is defined in line~\ref{lst_code:time_client_cpp_url}.
-It fits to the HED configuration file which was shown in Listing~\ref{lst:time_arched_xml}.
-Actually, if available, in many cases the client may use the configuration file of the server. \task{seen for the arcecho, but is that true? sm}
+It fits to the server configuration file which was shown in Listing~\ref{lst:time_arched_xml}.
 The daemon is listing on port 60000 and is running on the localhost. The service itself uses HTTP and can be reached by using the path \textit{time}.
@@
 The protocol stack is created in line~\ref{lst_code:time_client_cpp_client}. 
-Thanks to the service URL the information for the configuration is complete.
+Thanks to the service URL the information for the configuration is complete.\\
 
@@
 The request will be processed in the next line by the method \textit{process} of the object \textit{client}.%in line~\ref{lst_code:time_client_cpp_message_response}. 
-The object \textit{MCCStatus} which is returned by the method indicates the success of the processing.
-After verifying the response, the message will passed to the standard out in line~\ref{lst_code:time_client_cpp_message_answer}.
+The object of the class \textit{MCCStatus} which is returned by the method indicates the success of the processing.
+After verifying the response, the message will be passed to the standard out in line~\ref{lst_code:time_client_cpp_message_answer}. \\
 
@@
 \end{lstlisting}
-The messages are transmitted in SOAP format and thus reach the SOAP level of the server. They shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
+The messages are transmitted in SOAP format and thus reach the SOAP level of the server. They are shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
 The enclosing element is \textit{Envelope} which may contain an element \textit{Header} and must contain an element \textit{Body}.

### 2009-03-02 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Checked the whole document up to the Sec Echo Web Service. Started writting more about the Sec Echo Web Service. Added XSD of the XML configuration file to the appendix.

@@
-\chapter{TLS echoservice}
+\chapter{The TLS Echo Web Service}
 
-The web services of the previous chapters have one mayor drawback: the communication between the two endpoints is insecure.
-For that reason the current and the next chapter will discuss how to create a secure web service.
+The Web Services of the previous chapters have one mayor drawback: the communication between the two endpoints is insecure.
+For that reason the current and the next chapter will discuss how to create a secure Web Service.
 In general, security in a network can be subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{verfiy citation}
@@
 
-	\item \textbf{Integrity} --- Protection against interception and manipulation,replay,insertion, etc.  of a message.
+	\item \textbf{Integrity} --- Protection against interception and manipulation, replay, insertion, etc.  of a message.
 
@@
 In the present chapter, the security will be put into practice by extending the protocol stack with an additional security layer called TLS (Transport Layer Security).
-Due to the location after the TCP and the HTTP layer, the layer is only capable to limit the access to the grid to a certain set of users.\task{What do you mean here? SM}
-Nevertheless, it provides the security categories: confidentiality, authentication, integrity, non-repudiation and access control.
+Due to the location between the TCP and the HTTP layer, the layer is only capable to limit the access to the grid to a certain set of users.
+Once the user has passed the layer sucessfully the connection has been established and he may use all resources provided by the HED.
+This kind of behaviouor can be seen as to be very coarse grained.
+Nevertheless, it provides the security categories: confidentiality, authentication, integrity, non-repudiation and access control.\\
 
-In the next section gives a basic understanding of the working principale of the TLS.
+The next section gives a basic understanding of the working principale of the TLS.
 Emphasis is put on the security concept and not on an exact description of the technical realisation in order to enable newcomers an entry to  this area.
@@
 It is based directly on the TCP/IP layers and uses typically an asymmetric encryption to establish the connection (alternatively a symmetric pre-shared key may be used).
-After the communication has been successfully established, both participants are switching to another encryption method which is based on a new negotiated symmetric key.
+After the communication has been successfully established, both participants are switching to another encryption method which is based on a new negotiated symmetric key.\\
 %    1. Peer negotiation for algorithm support
@@
 In case of symmetric encryption, both participants are holding the same secret key to encrypt and decrypt a plain text.
-The principle is to reverse the process of encryption for decrypting the text.
+The principle is to reverse the process of encryption in order to decrypt the text.
 Symmetric encryption can be ranked as being very secure but its downside is that the key distribution is very difficult in practice.
-The key has to be shared previously, ideally over a secure channel which is may either be a complete diffrent medium (i.e.
-a letter, speech) or a channel which has been encrypted with another secret key.
+The key has to be shared previously, ideally over a secure channel which may either be a complete diffrent medium (i.e.
+a letter or speech) or a channel which has been encrypted with another secret key.
 Asymmetric encryption follows a diffrent functional principle.
@@
 \end{figure}
-\task{Maybe the term service has to be replaced by server}
 
@@
 As one can see, this first approach already provides confidentiality and integrity of messages.
-
 Neither Non-repudiation, authentication and consequently nor access control are established because the identity of Alice is not proven.
@@
 Certificates render the possibility to validate the identity of its owner.
-They are signed by a certificate authority (CA), the identity of which can be verified by another CA or that is pre-configured to be trusted.
+They are signed by a certificate authority (CA), the identity of which can be verified by another CA or that is pre-configured to be trusted.\\
 
@@
 Finally the signed certifcate will be returned to the requester.
-If an outside person wants to verify the certificate owner, the encrypted hash value has to be decrypted by the public key of the CA and a hash value has to be created with the same algorithm in order to compare them.
+If an outside person wants to verify the certificate owner, the encrypted hash value has to be decrypted by the public key of the CA and a hash value has to be created with the same algorithm in order to compare them.\\
 
@@
 Alice submits her request to Bob along with a challange consisting of a random number $n$. 
-The random number ensures that it is not possible to repeat a message (integrity).
 Bob encrypts the number with his private key and returns it together with his certificate.
+The random number ensures that it is not possible to repeat a message (integrity) and to prove that the communication partner holds the private key.
 Alice uses the public key of Bob to decrypt the challenge. 
-If the decrypted number is equal to the transmitted number, Alice can reason that the identity of Bob is to be trusted in case she is trusting the CA which has signed Bobs certificate.
+If the decrypted number is equal to the transmitted number, Alice can reason that the identity of Bob is to be trusted in case she is trusting the CA which has signed Bobs certificate.\\
 
@@
 The CA 1 encrypts the challenge with its private key and sends the result along with its certificate back to Alice.
-Alice successfully validates the certificate of CA 1.
+Alice successfully compares the returned number. Additionally she validates the certificate of CA 1 with the pre-configured certificate of CA 2..
 Since the certificate was signed by CA 2 which is pre-configured to be trusted, she als trusts CA 1 and in conclusion the identity of B.
@@
 If Bob also wants to be sure who he's talking to, then the same procedure has to be done to confirm the identity of Alice.
-The TLS mode in which both endpoints are ensuring themselves to whom they are talking to is called \textit{bilateral connection mode}.
+The TLS mode in which both endpoints are ensuring themselves to whom they are talking to is called \textit{bilateral connection mode}.\\
 
-When the client is aware of the service authentity, the confidential treatment of his data is given such that one can be sure the data won't be revealed to a third party (service-sided confidentiality).
-On the other side, if the service is aware of the client authentity, the access to the service can be controlled (access control) and the service can verify that the results will only be passed back to the right person (client-sided confidentiality).
-As one can see confidentiality, authentication, integrity and non-repudiation are now given.
+When the client is aware of the server authentity, the confidential treatment of its data is given such that one can be sure the data won't be revealed to a third party (server-sided confidentiality).
+On the other side, if the server is aware of the client authentity, the access to services and resources can be controlled (access control).
+Thus, in case of services, the results will only be passed back to the right person (client-sided confidentiality).
+As one can see confidentiality, authentication, integrity and non-repudiation are now given.\\
 
@@
 Instead, a set of pre-configured CAs which will be in general provided by the operation system is needed.
-In respect of ARC it is obvious that the access to the grid shall be under control of its administrator.
-To do so, the administrators have to establish their own CA such which is in charge of creating certificates for a set of users.
+In respect of ARC it is obvious that the access to the grid shall be under the control of its administrator.
+To do so, the administrators have to establish their own CA which is in charge of creating certificates for a set of users.
 Many universities have prepared local certificate authorities.
-And every user can establish his own.
+And every user can establish his own.\\
 
 The identity of the created CA have to be added to the directory of pre-configured CAs.
-The ARC source code comes with two shell scripts to create a CA and related certificates.
-Several publicly availble tools for certificate handling haven been prepared, e.g. (like \textit{openssl}),
+\task{The ARC source code comes with two shell scripts to create a CA and related certificates. - really? MG}
+Several publicly availble tools for certificate handling have been prepared, e.g. (like \textit{openssl}),
 which allow the certificate handling also for the advanced UNIX users.
-Certificates in general, however, have found entry to regular computer users, for instance with the FernUniversity Hagen (https://ca.fernuni-hagen.de/) who use that technology for their remote students to hand in exercises.
+Certificates in general, however, have found entry to regular computer users, for instance with the FernUniversity Hagen (https://ca.fernuni-hagen.de/) who use that technology for their remote students to hand in exercises.\\
 
@@
 It may involve organisations close to the leading organisation as well as structure which are not formally associated with it.
-In most cases the basic idea is to share ressources. \task{this para will need some overhaul - SM}
+In most cases the basic idea is to share ressources.
+\task{this para will need some overhaul - SM; jup, that a bit short...
+unfortunatly I am not very familiar with that stuff, but I am on it :-( - MG}
 
@@
 \section{Service}
-\task{RENAMING HED configuration file into~ARC service configuration file - proposed by Marek}
-\task{(echo service) client configuration file - proposed by Marek}
 
@@
 The CA for the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key.
-For the given example X.509-certificates will be used which are standardised by the ITU-T.
+For the given example X.509-certificates in the PEM format will be used which are standardised by the ITU-T.
 Altogether six files are needed:
@@
 \forcelinebreak
-As one can see, PEM has been choosen for the file format.\\
+\\
 
@@
 The script \textit{createMasterCA.sh} will create a new CA and overwrite the old on which was stored in the directory.
-The script \textit{createSlaveCert.sh} create an arbitary amount of certificates signed by the CA which has been create by the other script.\\
+The script \textit{createSlaveCert.sh} can create several certificates signed by the CA which has been create by the other script.\\
 %WARNING: If the programm fails, check if one of your certifacte is expired. If so, create new one using the script files in the directory 'certFactory'
 
-The TLS echo service implementation is almost the same as the one presented in the chapter before.
+The TLS Echo Web Service implementation is almost the same as the one presented in the chapter before.
 Merely the Class name and the namespace of the service has changed.
-The only modification has to be done within the ARC service configuration file which is displayed in Listing~\ref{lst:tls_echo_arched_xml}.
+The only modification has to be done within the server configuration file which is displayed in Listing~\ref{lst:tls_echo_arched_xml}.
 
@@
 {../src/services/tlsechoservice/arched_tls_echoservice.xml}
-The first change is to be found in lin~\ref{lst_code:tls_echo_arched_xml_plugin}.
+The first change is to be found in line~\ref{lst_code:tls_echo_arched_xml_plugin}.
 As to be seen, another plugin called \textit{mcctls} will be loaded by the \textit{ModuleManager}.
@@
 \end{verbatim}
-\task{You can put the ClientSSLConfig into a separate file, then use <ClientSSLConfig FromFile='filename'/> in each service, and then you don't need to change everywhere - proposed by Zsombor}
+\task{You can put the ClientSSLConfig into a separate file, then use <ClientSSLConfig FromFile='filename'/> in each service, and then you don't need to change everywhere - proposed by Zsombor - but I can't find any example *hrmlgrmpf*...}
 
@@
 In this section a new way to implement the client will be introduced.
-As already mentioned in section~\ref{sec:timeservice_client}, the protocol stack of the client is build up in almost the same way like the service.
-Indeed, the class~\textit{ClientSOAP} composes the stack with the same MCC used for the service.
+As already mentioned in section~\ref{sec:timeservice_client}, the protocol stack of the client is build up in almost the same way like the one of the service.
+Indeed, the class~\textit{ClientSOAP} composes the stack with the same MCCs used for the service.
 Once this is known, its self-evident to create the client in a simlar manner like the service with an own configuration file.
 Listing~\ref{lst:tls_echo_client_cpp} shows the C++ source code of the client implementation.
-The main difference between the simple echo client presented in the chapter before is that now the information about the client protocal stack is encapsulated within a configuration file.
+The main difference between the simple Echo Client presented in the chapter before is that now the information about the client protocal stack is encapsulated within a configuration file.
 \lstsetCPP
@@
 	label=lst:tls_echo_client_cpp,
-	caption={[HEC configuration file]
-	\textbf{HEC configuration file\textcolor{white}{hmf}}}
+	caption={[Client configuration file]
+	\textbf{Client configuration file\textcolor{white}{hmf}}}
 	]
 {../src/clients/tlsechoclient/tlsechoclient.cpp}
-The file containing the client configuration is specified within the command line. 
+The file containing the client configuration is specified within the commandline. 
 Its content will be loaded in the code fragment starting in line~\ref{lst_code:tls_echo_client_cpp_getfilecontent}. 
-Later, in line~\ref{lst_code:tls_echo_client_cpp_loadconfiguration}, the configuration will be parsed into XML and than transfered into a chain of component which is wrapped by the object \textit{loader}.
+Later, in line~\ref{lst_code:tls_echo_client_cpp_loadconfiguration}, the configuration will be parsed into XML and than transfered into a chain of components which is wrapped by the object \textit{loader}.
 In order to process a message, an entry point within the chain has to be disclosed.
@@
 After the payload has been wrapped within the message it gets transmitted in almost the same way like before, see lines followed by line~\ref{lst_code:tls_echo_client_cpp_createandprocess}.
-The rest of the source code is identical to the echo client.\\
+The rest of the source code is identical to the Echo Client.\\
 

### 2009-03-03 doc/storage_system/arc-storage-documentation.tex
Started documenting replicated ahash

@@
 
-TODO: here comes a fairly detailed description of how the replicated A-Hash works.
+The replicated A-Hash module is built on Oracle Berkeley DB High Availability, an open source database library with a replication API.
+The replication is based on a single master, multiple clients framework where all clients can read from the database, while only the master is allowed to write to the database.
+This ensures that all database transactions are ACID (atomic, consistent, isolated and durable).
+The replicated A-Hash module consists of three modules, \verb!ReplicatedAHash!, \verb!ReplicationStore!
+and \verb!ReplicationManager!.
+
+\begin{description}
+    \item[ReplicatedAHash] TODO: describe ReplicatedAHash, sendMessage, processMessage
+    \item[ReplicationStore] TODO: describe ReplicationStore, connection with TransDBStore, config, etc.
+    \item[ReplicationManager] TODO: describe ReplicationManager, elections, communication (interservice communication, security, callback functions)
+\end{description}
 
@@
 is suitable for using in a replicated deployemant, all the others are only for a centralized deployement.
+This variable is only configurable for the centralized deployement.
     \item[StoreCfg] could contain some parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash should save its files.
-\end{description}
-
-TODO: describe here the configuration needed for the replicated A-Hash.
+This variable is only configurable for the centralized deployement.
+\end{description}
+
+For the replicated A-Hash there are some additional required configuration variables:
+
+\begin{description}
+    \item[LocalDir] specifies a directory on the local filesystem where this A-Hash replica should be stored.
+This directory must be unique for this replica, i.e., no two A-Hash replicas can share the same \verb!LocalDir!, and doing so will most likely leave both replicas broken beyond repair.
+    \item[MyURL] is the URL to this A-Hash service and is used as an ID for this service in the replication group.
+    \item[OtherURL] is the URL of another A-Hash in the replication group. More than one URL can be specified here.
+\end{description}
+
+Even though the default settings are useable, there are some optional variables available for the replicated A-Hash:
+TODO: Put in more optional variables.
+\begin{description}
+    \item[Priority] is related to the master election.
+If a master dies, there will be an election to decide who will be the new master.
+Higher priority means a higher chance to be elected.
+A priority of 0 means that this replica can never be elected.
+Defaults to 10.
+    \item[CheckPeriod] specifies in seconds how frequently the A-Hash should update the replica list. Defaults to 10 seconds.
+    \item[CacheSize] configures the amount of the storage to be cached in memory. Defaults to 10MB
+\end{description}
 
@@
         </StoreCfg>
+    </Service>
+\end{verbatim}
+
+Example of replicated A-Hash:
+\begin{verbatim}
+    <Service name="pythonservice" id="ahash1">
+        <ClassName>storage.ahash.ahash.AHashService</ClassName>
+        <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
+        <LocalDir>ahash_data1</LocalDir>
+        <MyURL>http://localhost:60000/AHash1</MyURL>
+        <OtherURL>http://localhost:60001/AHash2</OtherURL>
+        <Priority>50</Priority>
+        <ClientSSLConfig FromFile='clientsslconfig.xml'/>
     </Service>

### 2009-03-05 doc/ws-programming-tutorial/doc/tex_appendix/appendix.tex
Finished the Secure Echo Web Service and made first correction of the chapters.

--- 2009-02-27 doc/ws-programming-tutorial/doc/tex_appendix/appendix.tex
@@
 
-\section{More tutorials}
+\section{Useful tutorials and documentations}
 
-Other tutorials are:\\
-\\
-%Oxana reported the status of OGF tutorials:
+In order to learn more about ARC several other tutorials and documentations have been written:
 \begin{itemize}
- \item first tutorial will be given by Oxana and will deal with how to add the 
-resources to Nordugrid, .... (will be related mostly to production ARC)
- \item second tutorial will be given by Gabor and he will show the 
-interoperability of ARC (video demo will be shown). A-REX - Unicore 
-interoperability.
- \item there will be also a tutorial related to ETICS system where the 
-possibility of ETICS system to submit a job to ARC and Unicore will be 
-presented. The ETICS team has not contacted so far.
+ \item ARC Web Services Quick Usage Guide~\cite{2008_UNKNOWN}
+ \item The Hosting Environment of the Advanced Resource Connector middleware~\cite{2008_Cameron}.
+ \item Security framework of ARC1~\cite{QIANG_2008}.
+ \item Documentation of the ARC storage system~\cite{Nagy_2008}.
+
 \end{itemize}
 
-\section{Important XSD files}
+\section{Important XSD files}\label{sec:impXSD}
 
-Link to current XSD files.
+\lstsetJUSTXML
+\lstinputlisting
+	[
+	label=lst:mcc.xsd,
+	caption={[XML schema of configuration files.]\textbf{XML schema of configuration files.}}
+	]{../../../src/hed/libs/loader/mcc.xsd}
 
-\task{Provide a list of modules to be loaded! Along with the implemented functionality which can be addressed with XML elements!!}
-    mcctcp\\
-\begin{verbatim}
-<tcp:Listen>
-	<tcp:Port>60000</tcp:Port>
-	<tcp:Version>4</tcp:Version>
-</tcp:Listen>
-\end{verbatim}
-    mcctls\\
-\begin{verbatim}
- 			<KeyPath>./clientKey.pem</KeyPath>
-			<CertificatePath>./clientCert.pem</CertificatePath>
-			<CACertificatePath>./clientCA.pem</CACertificatePath><!-- CACertificatesDir Should be /etc/... blub mention that-->
+\lstsetJUSTXML
+\lstinputlisting
+	[
+	label=lst:tcp.xsd,
+	caption={[XML schema of the TCP component.]\textbf{XML schema of the TCP component.}}
+	]{../../../src/hed/mcc/tcp/tcp.xsd}
 
-\end{verbatim}
+\lstsetJUSTXML
+\lstinputlisting
+	[
+	label=lst:tls.xssd,
+	caption={[XML schema of the TLS component.]\textbf{XML schema of  the TLS component.}}
+	]{../../../src/hed/mcc/tls/tls.xsd}
 
-    mcchttp\\
-\begin{verbatim}
-                      <next id="soap">POST</next>
-                      <next id="plexer">GET</next>
-                      <next id="plexer">PUT</next>
-\end{verbatim}
-    mccsoap\\
-\begin{verbatim}
-	???<ClientSSLConfig FromFile='filename'/>  
-\end{verbatim}
+\lstsetJUSTXML
+\lstinputlisting
+	[
+	label=lst:http_xsd,
+	caption={[XML schema of the HTTP component.]\textbf{XML schema of the HTTP component.}}
+	]{../../../src/hed/mcc/http/http.xsd}
+
+% 
+% Questions:
+% \begin{itemize}
+%  \item The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - Web Service   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - Web Service\\
+% --- ANSWER BY WEIZHONG: In my understanding (because plexer is not my initial idea), plexer (as a sort of hub) can be put at both after http or soap.
+If you put after soap,then you are doing switching for different web services.
+While I put it after http, because somehow, I need to switch the message after http-processing, in more detail, one branch for soap and then web service/s,
+%  one branch for a non-soap service/server (the SP Service is not a service based on SOAP, actually it is based on http, so it is kind of web application).
+% 
+% \item The HED does not act as a front end server which delegates its task to new instances acting on a diffrent port such that the daemon is always available. 
+% \end{itemize}
+
+% 
+% 
+ \task{RENAMING HED configuration file into~ARC service configuration file - proposed by Marek}
+ \task{(echo service) client configuration file - proposed by Marek}
+ \task{Check commas using \url{http://leo.stcloudstate.edu/punct/comma.html}}
+
+ \task{Bisher noch nicht beachtet: \\
+\\\textcolor{urgent}{
+In general, a comma is used when the
+subordinate clause precedes the main clause, like this:\\
+\\
+``When subjects were aware of the purpose of the experiment, no effect was
+observed.''\\
+\\
+Furthermore:\\
+\\
+      ``We discarded specimens that had become contaminated.''\\
+\\
+In this case, the relative clause is restrictive because it deﬁnes exactly which specimens
+were discarded. Restrictive clauses should not be preceded by a comma.\\
+\\
+  In contrast, a non-restrictive clause gives additional information about an object that
+has already been identiﬁed; for example:\\
+\\
+      ``The ﬁle server uses the XFS ﬁle system, which supports journaling and pro-
+      vides excellent performance.''\\
+\\
+In this case, the relative clause does not restrict or deﬁne the particular ﬁle system being
+talked about, but merely provides additional information. Non-restrictive clauses should
+be preceded by a comma.}}

### 2009-03-05 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Finished the Secure Echo Web Service and made first correction of the chapters.

@@
 
-The next service to be presented is the Echo Web Service.
+The next service presented is the Echo Web Service.
 Its task is to return the content of the received message back to the client.
@@
 The discovery of services may be useful for an automatic adaption of a running client, i.e. if the endpoint of the service changes or to share the load of demands.
-The WSDL service specification enables the programmer to implement a suitable client which interfaces the service in a proper way.
+The WSDL service specification enables a foreign programmer to implement a suitable client which interfaces the service in a proper way.
 WSDL describes the structure of the messages, the intended use of the messages, the supported protocols and the endpoint of the service (URL).
 \task{For that last reason, every instance of a service must be produced a distinct WSDL file for - or - it is automatically prepared from the information that the server has available once it is implemented.
-- ?MG }\\
+- ??MG }\\
 
-The WSDL file that specifies the here presented Echo Web Service is shown in listing~\ref{lst:echo_wsdl}. 
+The WSDL file that specifies the here presented Echo Web Service is shown in Listing~\ref{lst:echo_wsdl}. 
 It is composed of five main elements:
@@
 Four interface types can be distinguished: One-way (input), request-response (input, output), solicit-response (input, output, fault) and notification (output).
-	The echo service realises a solicit-response interface which is declared subsequent to line~\ref{lst_code:echo_wsdl_portType}. % in the lines followed by the line X
+	The Echo Web Service realises a solicit-response interface which is declared subsequent to line~\ref{lst_code:echo_wsdl_portType}. % in the lines followed by the line X
 
@@
 	[
-	label=lst:echo_wsdl,float=p,
+	label=lst:echo_wsdl,
 	caption={[WSDL file describing the echo service. Filename: echo.wsdl]
@@
 {../src/services/echoservice/echo.wsdl}
-\task{Validate WSDL}
+\task{TODO: Validate the WSDL Listing}
 
-\subsection{The SOAP Fault Element}
+\subsection{The SOAP Fault Message}
 
-The optional SOAP Fault element is used to indicate error messages in case of the port type \textit{solicit-response} (input, output, fault).
-If a \textit{Fault} element is present, it must appear as a child element of the \textit{Body} element. A \textit{Fault} element can only appear once in a SOAP message.
+A SOAP fault message is always located within the body element and can appear only one time. 
+It indicates error messages in case the port type \textit{solicit-response} (input, output, fault) has been chosen.
+An element called \textit{fault} encapsulates diffrent types of elements which are describing the fault.
+These elements are listed in Table~\ref{tbl:fault_elements}.
+\begin{table}[h]
+ \centering
+ \caption{Elements of the fault message.}
+ \label{tbl:fault_elements}
+	\begin{tabular*}{\textwidth}[t]{p{2cm}p{13cm}}
+	\hline
+	\textbf{Element}        & \textbf{Description}\\
+	\hline
+	\textbf{faultcode} 	& Code which shall identify the fault.\\
+	\textbf{faultstring} 	& A human readable explanation of the fault.\\
+	\textbf{faultactor} 	& Information about who caused the fault to happen.\\
+	\textbf{detail}         & Holds application specific error information.\\
+	\hline\\
+	\end{tabular*}
+\end{table}
+The fault code describes the generic class of the fault. The ARC implementation assigns each fault a certain number.
+The fault code along with its description is described within the Table~\ref{tbl:faultcode}
+\begin{table}[h]
+ \centering
+ \caption{Possible fault codes.}
+ \label{tbl:faultcode}
+	\begin{tabular*}{\textwidth}[t]{p{2cm}p{4cm}p{9cm}}
+	\hline
+	\textbf{Fault code} & \textbf{Error} & \textbf{Description}\\
+	\hline
+	0 & \textbf{undefined}       & --- \\
+	1 & \textbf{unknown}         & Reason for failure couldn't be identified.\\
+	2 & \textbf{VersionMismatch} & Found an invalid namespace for the SOAP Envelope element.\\
+	3 & \textbf{MustUnderstand}  & The processing of the SOAP header is optional unless the flag mustUnderstand is set ``true''.
+If, according to that case, the SOAP service is not able to understand the header a fault message with the faultcode MustUnderstand is returned.
+\\
+	4 & \textbf{Client}          & The request message was incorrectly formed or contained incorrect information.\\
+	5 & \textbf{Server}          & There was a problem with the server so the message could not proceed \\
+	6 & \textbf{DataEncodingUnknown} & The message couldn't be processed due to encoding problems\\
+	\hline
+	\end{tabular*}
+\end{table}
 
-The SOAP Fault element has the following sub elements:
-\begin{tabular}{lp{9cm}}
-Sub Element 	& Description \\
-<faultcode> 	& A code for identifying the fault \\
-<faultstring> 	& A human readable explanation of the fault \\
-<faultactor> 	& Information about who caused the fault to happen \\
-<detail> 	& Holds application specific error information related to the Body element\\
-\end{tabular}
-
-SOAP Fault Codes
-
-The faultcode values defined below must be used in the faultcode element when describing faults:
-\begin{tabular}{lp{9cm}}
-Error           & Description \\
-VersionMismatch & Found an invalid namespace for the SOAP Envelope element \\
-MustUnderstand  & n immediate child element of the Header element, with the mustUnderstand attribute set to "1", was not understood \\
-Client          & he message was incorrectly formed or contained incorrect information \\
-Server          & here was a problem with the server so the message could not proceed \\
-\end{tabular}
-
-\task{Source of the above two tables? Extension of the second table? Numbers??}
-
-More information about WSDL may be found at \href{http://www.w3.org/TR/wsdl}{http://www.w3.org/TR/wsdl}.
+For morde information about WSDL it is recommended to check the internet address~\href{http://www.w3.org/TR/wsdl}{http://www.w3.org/TR/wsdl}.
 
@@
 This section presents the implementation of the service defined in the WSDL file of the previous section.
-The source code of the C++ file is shown in listing~\ref{lst:echo_service_cpp}.
+The source code is shown in Listing~\ref{lst:echo_service_cpp}.
 The header file will be set aside for it contains only redundant information. 
 It can be reread in the source directory that is accompanying the tutorial.
-Regarding to the source code of the Echo Web Service basically three changes of the Time Web Service source code will be required.
+Regarding to the source code of the Echo Web Service basically three changes of the Time Web Service source code are required.
 The first change concerns the constructor and can be found at line~\ref{lst_code:echo_cpp_prefix}.
-Two strings are extracted of the server configuration file which later are used to envelope the message before it is returned.
-Furthermore, a new method \textit{makeFault} in line~\ref{lst_code:echo_cpp_makeFault} has been introduced which is able to create the fault messages and is called by the method \textit{process}.
+Two strings are extracted of the server configuration file which are later used to envelope the message before it is returned.
+Furthermore, a new method \textit{makeFault} in line~\ref{lst_code:echo_cpp_makeFault} has been introduced which is able to create a fault messages.
 Within the method \textit{process} two new code fragments are now extracting and analyzing the incoming message, to be seen at line~\ref{lst_code:echo_cpp_extracting} and \ref{lst_code:echo_cpp_analyzing}.
@@
 	label=lst:echo_service_cpp,
-	caption={[C++ implementation of the echo service. Filename: echoservice.cpp]
-	\textbf{C++ implementation of the echo service. Filename: echoservice.cpp}}
+	caption={[Implementation of the Echo Web Service.]
+	\textbf{Implementation of the Echo Web Service.}}
 	]
@@
 
-Again, the HED is configured by an XML file and is launched as introduced earlier.
+Again, the HED is configured by an XML file and launched in the same manner as introduced before.
 The server configuration file is shown in Listing~\ref{lst:echo_hed_configuration_xml}.
@@
 	label=lst:echo_hed_configuration_xml, float=htb,
-	caption={[Server configuration file for the echo service. Filename: arched\_echoservice.xml]
-	\textbf{Server configuration file for the echo service. Filename: arched\_echoservice.xml}}
+	caption={[Server configuration file of the Echo Web Service]
+	\textbf{Server configuration file of the Echo Web Service}}
 	]
@@
 Firstly, because the additional arguments need to be filled.
-Secondly, because of the extra effort to interact with the (here human) user to collect the data to forward to the Echo Web Service.
+Secondly, because of the extra effort to interact with the (here human) user to collect the data which needs to be forwarded to the Echo Web Service.
+The source code of the client is shown Listing~\ref{lst:client_echo_cpp}.
 The parameters are now passed by the command line, to be seen in the lines followed by line~\ref{lst_code:echo_client_cpp_arguments}.
 That code is independent from the ARC library.
-The creation of the SOAP client is unchanged and the request is created in the lines subsequent line~\ref{lst_code:echo_client_cpp_request}.
-It is processed in the code after line~\ref{lst_code:echo_client_cpp_response} in which also the response is received.
+The creation of the SOAP client is unchanged and the request is created in the lines subsequent to line~\ref{lst_code:echo_client_cpp_request}.
+It is processed in the code after line~\ref{lst_code:echo_client_cpp_response} in which is also the response is received.
 The checks of the response have expanded by an additional condition which tests for a fault message, see line~\ref{lst_code:echo_client_cpp_fault}.
@@
 	[
-	label=lst:arcecho_arched_xml,
-	caption={[HED configuration file for the ARC echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the ARC echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	label=lst:client_echo_cpp,
+	caption={[Implementation of the client program.]
+	\textbf{Implementation of the client program.}}
 	]
@@
 label=lst:invokation_echoservice, float=!htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+caption={[Three diffrent invokations of the client program.]
+         \textbf{Three diffrent invokations of the client program.}}]
 $ ./echoclient http://localhost:60000/echo ordinary text_to_be_transmitted
@@
 A few examples of exchanged messages are shown in the Listings~\ref{lst:echoservice_request}, \ref{lst:echoservice_response} and \ref{lst:echoservice_fault}.
-Listing~\ref{lst:echoservice_request} illustrates a request message using the operation \textit{reverse}.
-Listing~\ref{lst:echoservice_response} illustrates the corresponding response message.
-When the operation requested by the client is unknown, then the server replies with the fault message shown in Listing~\ref{lst:echoservice_fault}.
+Listing~\ref{lst:echoservice_request} illustrates a request message using the operation \textit{reverse} while Listing~\ref{lst:echoservice_response} illustrates the corresponding response message.
+The last Listing~\ref{lst:echoservice_fault} shows the fault message which is created by the service after the client requested an unknown operation.
 
@@
 label=lst:echoservice_request, float=!htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+caption={[Request message created by the client.]
+         \textbf{Request message created by the client.}}]
 <soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
@@
 label=lst:echoservice_response, float=!htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+caption={[Response message created by the service.]
+         \textbf{Response message created by the service.}}]
 <soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
@@
 label=lst:echoservice_fault, float=!htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+caption={[Fault message created by the service..]
+         \textbf{Fault message created by the service.}}]
 <soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">

### 2009-03-05 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Finished the Secure Echo Web Service and made first correction of the chapters.

@@
 
-This document gently introduces to preparation of standalone webservices and cognate clients with the Advance Resource Connector (ARC).
-The reader is guided through a series of the practically oriented examples.
-Each are accompanied by explanations of the main concepts of the software architecture.
+This document gently introduces to the preparation of standalone Web Services and cognate clients with the Advance Resource Connector (ARC).
+The reader is guided through a series of practically oriented examples.
+Each is accompanied by explanations of the main concepts of the software architecture.
 No particular skills are required to follow the presented steps.
@@
  \item Transmit the request to the server which provides the desired service.
+ \item Wait for the reply
  \item Receive the response from the server.
@@
 %
-At the time of writing, the endpoint of the HED will be a SOAP based webservice.
+At the time of writing, the endpoint of the HED will be a SOAP based  Web Service.
 SOAP is a long established XML-based standard for Web Service communication.
@@
 
-The source code of the examples given in this tutorial are available in the subdirectory \textit{src}\footnote{If this was not shipped with the document you are reading, then please inspect\\ http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/ws-programming-tutorial}.
-
 \section{Hosting Environment Daemon}
@@
 The HED which was already mentioned in the section above is an essential part of the ARC middleware. 
-Its task is to provide the hosting of various services at the application level and is based on the idea modularity.% allows its dynamic extension.
-The HED consists of compononents called Message Chain Components (MCC), services and several modules, which shall aid the programmer to simplify the development of the Web Services: Config, Loader, Logging, XMLNode.
+Its task is to provide the hosting of various services at the application level and is based on the idea modularity. % allows its dynamic extension.
+The HED consists of components called Message Chain Components (MCC), Plexer, Services, and several modules, which shall aid the programmer to simplify the development of the Web Services: Config, Loader, Logging, XMLNode et cetera.
 The MCCs are ordered within a layered structure. 
@@
 	\centering
-	\includegraphics[width=14cm]{tex_introduction/HED.pdf}
-	\mycaption{Example of HED structure with an Arex service and a file service.\task{ Replace image by an easier one!} }{\task{Write more stuff}\task{The images of the HED structure always are ordered like: TCP - TLS - HTTP - PLEXER - SOAP - Web Service   but the source code examples are: TCP - TLS - HTTP - SOAP - PLEXER - Web Service}}
+	\includegraphics[width=14cm]{tex_introduction/HED2.pdf}
+	\mycaption{Example of a HED structure providing a Web Service}{The HED consists of several layers which are connected with each other.
+The arrow indicate the message flow of the used configuration.
+The MCCs and Services may hold a SecHandler in order to give additional security.
+Additional modules (on the left) are used within the HED.
+\task{Why is the Web Service in all images outside of the HED?}}
 \label{fig:HED_internal}
@@
 In case of outgoing messages a MCC is wrapping the data of the upper layer as a payload into their own protocol (while incomming message will be unwrapped).
-The lowest layer has to be a transport protocol like TCP which enables the transmission between the two applications.
-Tehchnically there even lower layers required needed for the communication but they are realised within the operating system and the  network card
+The lowest layer has to be a transport protocol like TCP which enables the transmission between two applications.
+Technically there even lower layers required needed for the communication but they are already realised within the operating system and the  network card.
 TCP is well suited for the application because it provides a reliable and ordered delivery of a byte stream.
@@
 The multiplexing layer can also be defined within the HED and is called Plexer.
-Depending on the path of the URL the Plexer passes messages to a defined SOAP service which again passes the message to the Web Service~\cite{QIANG_2005}. \task{Add text here}\\
+Depending on the path of the URL the Plexer passes messages to a defined SOAP service which again passes the message to the Web Service~\cite{QIANG_2005}.\\
+% QUESTION: Why is the plexer always illustrated  being above the SOAP MCC. There's no config file example like that in the tree.
 %In my understanding (because plexer is not my initial idea), plexer (as a sort of hub) can be put at both after http or soap.
@@
 The structure of the HED can be configured freely by modifying the server configuration file.
-A first impression how the HED can be configured shall be given in the following example which corresponds to the HED shown in figure~\ref{fig:HED_internal}.
+A first impression how the HED can be configured shall be given in the following example which corresponds to the HED shown in Figure~\ref{fig:HED_internal}.
 It uses the ARC echo service that is shipped for testing purposes with the ARC source code.
@@
 The first example shall give a basic understanding in how to configure the HED for setting up a novel service.
-Here, the task is to setup the regular ARC Echo Web Service.
-On the command line, all that needs to be done is to invoking the \textit{arched}%
-\footnote{\task{refer to the appendix in which the documentation to arc-hed is specified}}
-daemon with a fitting server configuration file.
+The task will be to setup the regular ARC Echo Web Service.
+On the command line, all that needs to be done is to invoking the \textit{arched} daemon with a fitting server configuration file.
 Such a fitting server configuration file is shown in Listing~\ref{lst:arched_arcecho_xml}.
-It is written in XML and its structure is based on within the xsd-schema \textcolor{urgent}{URL?} .
+It is written in XML. Its structure is based on the XSD schema listed in~\ref{sec:impXSD} in Listing~\ref{lst:mcc.xsd}.\\
 
@@
 \lstinputlisting
-	[
-	label=lst:arched_arcecho_xml,float,
-	caption={[HED configuration. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
-	]
+	[label=lst:arched_arcecho_xml,float=htb,
+	caption={[HED configuration file.]
+	\textbf{HED configuration file.}}]
 {../src/services/arched_arcecho.xml}
@@
 Finally, the chain is declared within the element \textit{Chain} which creates the layered structure.
-It is composed out of the elements \textit{Component}, \textit{Plexer} and \textit{Service}.
+It is composed of the elements \textit{Component}, \textit{Plexer} and \textit{Service}.
 Every element inside has at least a \textit{name} attribute.
@@
 In this way, the message is be passed to higher layers until it reaches the \textit{Plexer}.
-Depending on the path of the URL which was demanded by the HTTP, the Plexer multiplexes the request to the right Web Service, the right function to execute according to the component IDs and plugin names.
+Depending on the path of the URL which was demanded by the HTTP, the Plexer multiplexes the request to the right Web Service\task{, the right function to execute according to the component IDs and plugin names.
+-- ??MG}
 The plexer will thus allow to have multiple Web Service listen to the same port.
@@
 \begin{lstlisting}[
-label=lst:arcecho_arched_invokation,float,
-caption={[Invokation of the Arched Daemon.]
+label=lst:arcecho_arched_invokation,float=htb,
+caption={[Invokation of the arched daemon in order to start the ARC Echo Web Service.]
          \textbf{Invokation of the Arched Daemon.\textcolor{white}{hmf}}}]
@@
 \begin{lstlisting}[
-label=lst:arcecho_client_invokation,float,
-caption={[Usage of the ARC echo client.]
+label=lst:arcecho_client_invokation,float=htb,
+caption={[Usage of the ARC Echo client.]
          \textbf{Usage of the Arc echo client.\textcolor{white}{hmf}}}]

### 2009-03-05 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
Finished the Secure Echo Web Service and made first correction of the chapters.

--- 2009-02-23 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
@@
-\chapter{SEC echoservice}
+\chapter{Secure Echo Web Service}
 
-\section{SecureHandler}
-And what are PDP's?
+The TLS Echo Web Service, which was presented in the previous chapter, was able to control the access to the HED on a very low layer. 
+For the administration of a large VO the kind of security may be too general.
+Thus, it is recommended to have a more fine grained security. 
+This chapter will discuss how to install a more complex security such that it is possible to limitate the access of certain resources to particular users and to narrow their rights on the workstation.\\
 
-Currently, the ARC1 comes in severalfour policy decision point implementations:
-\begin{itemize}
- \item arc.pdp – Arc PDP\\
-The Arc PDP will organize the security attributes into the ARC specific authorization request, call the
-policy evaluator to evaluate the request against the policy (which is in ARC specific format) repository,
-and get back the evaluation result. See paragraph 3 for detail information about request schema and policy
-schema.
-\item delegtion.pdp – Delegation PDP\\
-The Delegation PDP is basically similar to Arc PDP, except it uses the delegation policy parsed from
-remote proxy credential by delegation.collector, and evaluates the request against delegation policy. See
-section 6for the design idea and use case of delegation policy in fine-grained identity delegation.
-\item simplelist.pdp – Simplelist PDP\\
-The Simplelist PDP is a simplest implementation of policy decision point. It will match the identity
-extracted from the remote credential (or proxy credential) with local list of permitted identities.
-\item pdpservice.invoker – PDP Service Invoker\\
-The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the
-same functionality as Arc PDP, except that the evaluation request and response are carried by SOAP
-message. The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation
-engine can be accessed remotely and maintained centrally.
- \item allow.pdp\\
- Does something unknown!
-\end{itemize}
+The security framework of ARC is implemented based on the same concept of modularity as HED. 
+To extend the security of a MCC or a service one only has to modify the server configuration file.
+That is done by including a \textit{SecHandler} element into the component or service which again has to contain a PDP element.
+A small example of a chain component containing a \textit{SecHandler} is shown in Listing~\ref{lst:sec_service_component_xml}.
+Within the element \textit{Component}, which realises a HTTP layer, an element \textit{SecHandler} is located.
+It contains three attributes.
+The  attribute \textit{name} is \textit{arc.authz}, the \textit{event} attribute is \textit{incoming} and the \textit{id} attribute which is \textit{auth}.
+While \textit{id} is only an optional parameter, the attribute \textit{name} causes a certain \textit{SecHandler} to be loaded.%, which again has a certain functionality. 
+In case the attribute \textit{name} is assigned to \textit{arc.authz}, the SecHandler provides the functionality of authentication and is able to permit or to deny the access to the next layer. 
+Each MCC or Service usually implements two queues of \textit{SecHandlers} --- one for incoming messages and one for outgoing messages.
+The attribute \textit{event} attaches the present one to the incoming queue.
+\textit{SecHandlers} are able to intervene the message flow by: manipulating the message content, manipulating the message attributes or to disrupt messages.
+Within one component, \textit{SecHandlers} are executed sequentially in the order in which they were attached to the queue.
+If any of them fails, message processing fails as well~\cite{QIANG_2008}.
+In order to decide how the \textit{SecHandler} shall modify the message so called PDP (Policy Decision Point) are used.
+Within the example a PDP named \textit{simplelist.pdp} is inserted within the \textit{SecHandler} element.
+The mechanism of this PDP uses a file which contains a list of identities.
+If the current identity is on the list, the PDP returns a positive result to the \textit{SecHandler}.
+In general more than one PDP can be defined within one \textit{SecHandler}.
+The default behaviour in that case is to execute all PDPs sequentially until one fails such that a negativ result will be returned or all return a negativ result such that also a positive result will be passed to the \textit{SecHandler}.\\
+%
+%
+% Weizhong: Verifying, parsing the message attribute, and using the attributes for authorization
+%
+% Attributes done by: identity.map, delegation.collector
+% manipulating content: usernametoken.handler
+% disrupt by: arc.authz
+\lstsetJUSTXML
+\begin{lstlisting}[
+label=lst:sec_service_component_xml,float=htb,
+caption={[Example of a SecHandler within the component \textit{http.service}.]
+         \textbf{Example of a SecHandler within the component \textit{http.service}}}]
+<Component name="http.service" id="http">
+	<next id="soap">POST</next>
+	<SecHandler name="arc.authz" event="incoming" id="auth">
+		<PDP name="simplelist.pdp" location="./clientlist.txt"/>
+	</SecHandler>
+</Component>
+\end{lstlisting}
 
-Following Flavors:
-\subsection{arc.authz – Authorization SecHandler}
-The arc.authz is responsible for calling the interface of policy decision point and getting back the
-authorization result, and then making decision according to this authorization result. There is one simple
-interface (see Figure 4) defined in PDP, which will be called by arc.authz if configured inside once there
-is message (incoming or outgoing) need to be processed.
+  \begin{table}[htb]
+  \centering
+  \caption{List of buildin SecHandlers}
+  \label{tbl:list_of_sechandler}
+  \begin{tabular*}{\textwidth}[t]{p{4cm}p{11cm}}
+	\hline
+ 	\textbf{identity.map}         & Maps the identity of the user to an identity of the workstation.\\
+%					realised by a message attribut.\\
+ 	\textbf{arc.authz}            & Grants the access to the next layer.\\
+ 	\textbf{delegation.collector} & Processes a request for another service located on a different HED.
+                                        A certificate, which has a delegation policy embedded, is transfered to the service running that \textit{SecHandler}.
+Once the has been extracted and verified it will be transfered
+                                        to the requested service.
+                                        The policy is now available within the HED of the service such that the client is now able to access the service.\\
+                                        % Question how does the client creates the request - with the SecHandler?
+ 	\textbf{usernametoken.handler}& Generates the WS-Security username-token and adds it into the SOAP header of outgoing
+					 messages.
+In case of incoming messages the SecHandler extracts the WS-Security username-token out of it.
+The WS-Security standard extends SOAP and describes i.e.
+how to attach signatures and encryption header to SOAP messages.\\
+	\hline
+  \end{tabular*}
+  \end{table}
+% WS-Security describes how to attach signatures and encryption headers to SOAP messages.
+In addition, it describes how to attach security tokens, including binary security tokens such as X.509 certificates and Kerberos tickets, to messages.
+%
+  \begin{table}[htb]
+  \centering
+  \caption{List of buildin PDPs}
+\label{tbl:list_of_pdps}
+  \begin{tabular*}{\textwidth}[t]{p{4cm}p{11cm}}
+	\hline
+ 	\textbf{allow.pdp}           & Always returns a positive result.\\
+	\textbf{deny.pdp}            & Always returns a negative result.\\
+ 	\textbf{simplelist.pdp}      & Renders a positive result in case the identity corresponds to the listed ones and additionally 
+                                       maps the identity to an user of the workstation.\\
+ 	\textbf{arc.pdp}             & Evaluates the result based on a policy XML file.\\
+ 	\textbf{delegate.pdp}        & Similar to \textit{arc.pdp} a policy file will be evaluated. But here, the policy document is 
+                                       extracted and transfered
+                                       by delegation.collector.\\
+ 	\textbf{pdpservice.invoker}  & Establishes a connection to special service in order to get a policy decision. The PDP service 
+                                       implements the same functionality as ARC PDP, except that the evaluation request and 
+                                       response is carried by SOAP messages. The benefit of the PDP Service and the PDP Service
+                                       Invoker is that the policy evaluation engine can be accessed remotely and maintained centrally.
+\\
+	\hline
+  \end{tabular*}
+  \end{table}
+%
+%
+Due to the fact, that presenting all \textit{SecHandlers} and PDPs would go beyond the scope of an introducing tutorial, only a small set of them will be introduced here in detail. 
+Nevertheless, the Tables~\ref{tbl:list_of_sechandler} and~\ref{tbl:list_of_pdps} shall give a first overview about existing \textit{SecHandlers} and \textit{PDPs}.
+For further reading it is recommended to consult the document ``Security framework of ARC1'' written by Weizhong Qiang and Alexandr Konstantinov.
+The present chapter is closly related to it.\\
 
-\begin{minipage}[t]{\textwidth}
-\begin{lstlisting}[
-label=lst:secHandler_arc.authz,
-language=XML,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-...
-<Service name="sececho" id="sececho">
-	<next id="sececho"/>
-	<echo:prefix>[ </echo:prefix>
-	<echo:suffix> ]</echo:suffix>
-	<SecHandler name="arc.authz" id="pdps" event="incoming">
-		<PDP name="simplelist.pdp" location="/etc/grid-security/grid-mapfile"/>
-	</SecHandler>
-</SecHandler>
-...
-\end{lstlisting}
-\end{minipage}
+The Secure Echo Web Service which will be presented in this chapter, is using the \textit{SecHandlers}: \textit{indentity.map} and \textit{arc.authz}, and the \textit{PDPs}: \textit{allow.pdp}, \textit{simplelist.pdp} and \textit{arc.pdp}.
+\\
 
-\begin{lstlisting}[
-label=lst:invokation,
-language=ksh,
-frame=single,
-numbers=none,
-backgroundcolor=\color{lightgray},
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ tail -n5 /etc/grid-security/grid-mapfile
-"/C=Au/O=Science/OU=ARC/CN=Baldur Meier" griduser
-"/C=UK/O=MoreScience/OU=ARC/CN=Thor Schulz" griduser
-"/C=Au/O=Organelle/OU=ARC/CN=Loki Svensson" griduser
-"/C=SK/O=Grid.Org/OU=ARC/CN=Iduna Apple" griduser
-"/C=RU/O=MegaGrid/O=ARC/CN=Fenrir Wodaneater" griduser
-\end{lstlisting}
-Add your client certificate into that list!
+\section{Service}
 
-Latest: Policy.xsd
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Policy.xsd
-
-\subsection{identity.map – Identity Mapping SecHandler}
-
-The identity.map is a specific authorization oriented security handler. It will map the global identity in
-the message into local identity like system username based on the result returned by Policy Decision Point
-components.
-
-\subsection{delegation.collector – Delegation SecHandler}
-
-The delegation.collector is responsible for collecting the delegation policy information from the remote
-proxy credential (proxy certificate is compatible to RFC3820) inside the message, and putting this policy
-into message’s security attribute for the usage of other components, such as delegation.pdp.
-
-\subsection{ usernametoken.handler – UsernameToken SecHandler}
-
-The task of the usernametoken.handler is to generate the WS-Security Username-Token and add it into
-header of SOAP message which is the payload of outgoing message. It can also extract the WS-Security
-Username-Token from the header of SOAP message which is the payload of incoming message.
-
-POLICY STUFF:
-copy the request out of the LOGGEr and use that to write your policy.xml
-%/C=DE/ST=Germany/L=L\xFCbeck City/O=Instite of Neuro- and Bioinformatics/OU=Mitarbeiter
-
-Deny-Overrides - this is default if no algorithm specified.
-...
-Administrator of service can configure Authorization SecHandler - arc.authz - for each MCC and Service
-and define reasonable and meaningful policy. While defining policy the administrator must take into
-account that the attributes defined in the policy should be already collected by previous components in a
-chain. For instance, policy with AttributeId “http://www.nordugrid.org/schemas/policy-
-arc/types/http/path” should not be configured inside SecHandler attached to MCCTLS.
-
+Listing~\ref{lst:sececho_service_cpp} contains the source code of the Secure Echo Web Service.
+As one can see, the only diffrence, compared to the source code of the previous section, is the additional code to process the \textit{SecHandler} starting at line~\ref{lst_code:sececho_service_cpp_process_secH}.
+Obviously only the incoming queue is processed and in case the SecHandler fails, the SOAP service returns a fault message.
+The file, which is revealing a lot more information, is the server configuration file which is shown in Listing~\ref{lst:sececho_server_configuration_file_xml}.
+In order to be able to have \textit{SecHandlers} working with identies, it is almost essential to utilise the TLS layer.
+% Fixed IP are also possible but not recommended..
+The TLS layer extracts the identity such that it later can be used to process the \textit{SecHandler}.
+In line~\ref{lst_code:arched_sec_conffile_plugins} and line~\ref{lst_code:arched_sec_conffile_plugins2} two additional plugins 
+are named to be loaded which are implementing of the \textit{SecHandlers} and \textit{PDPs}.
+%
+%  FIRST ONE
+%
+The first \textit{SecHandler} is declared within line~\ref{lst_code:arched_sec_conffile_secH2}. 
+Due to the PDP \textit{allow.pdp} returns always a positive result, everyone who passes the \textit{SecHandler} is mapped to the user \textit{nobody}. 
+%
+% SECOND ONE
+%
+The second \textit{SecHandler} limits the access to the SOAP services using the \textit{arc.authz} \textit{SecHandler} and the \textit{simplelist} policy, see line~\ref{lst_code:arched_sec_conffile_secH1}.
+(A similar example was already shown in Listing~\ref{lst:sec_service_component_xml}).
+The \textit{simplelist} policy returns a positive result if the identity of the certificate corresponds to an identity within the file \textit{clientlist.txt}.
+A small example of such a file is displayed in Listing~\ref{lst:simplelist_example}.
 \lstsetCPP
@@
 	[
-	label=lst:arcecho_arched_xml,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	label=lst:sececho_service_cpp,
+	caption={[Source code of the Secure Echo Service.]
+	\textbf{Source code of the Secure Echo Service.}}
 	]
-{../src/clients/secechoclient/secechoclient.cpp}
-
-\lstsetCPP
-
+{../src/services/secechoservice/secechoservice.cpp}
+%
+%
+The first part of an entry needs to be the distinguished name (DN) of the current client.
+A DN is a string which represents a user uniquely and is created using the identity of the certificate.
+To get the correct DN of a client one can use the TLS Echo Web Service and check the log-file.
+The second part of an entry specifies the user the identity shall be mapped to.
+In the present case, the identity will be mapped to \textit{griduser}.
+In case the policy is unable to match the identity, the PDP returns a negativ result.
+The \textit{SecHandler} within the SOAP component limits the access of SOAP to a certain group of users.
+%
+%
+% The third ONE
+The last \textit{SecHandler}, used in the server configuration file, is to be seen starting with line~\ref{lst_code:arched_sec_conffile_secH3}.
+The \textit{SecHandler} is chosen to be \textit{arc.authz} which leads to the fact that the access to the Echo Web Service will be limited one more time to a subset of the group.
+The policy \textit{arc.pdp} returns a positive result in case the rules encapsulated within the file \textit{policy.xml} will return a positive result.\\
+\lstsetARCHEDXML
 \lstinputlisting
 	[
-	label=lst:arcecho_arched_xml,
-	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
-	]
-{../src/services/secechoservice/secechoservice.cpp}
-
-\lstsetARCHEDXML
-\begin{minipage}[t]{\textwidth}
-\lstinputlisting
-	[
-	label=lst:arcecho_arched_xml,
+	label=lst:sececho_server_configuration_file_xml, float=htb
 	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
@@
 {../src/services/secechoservice/arched_sec_echoservice.xml}
-\end{minipage}
 
 \lstsetKSH
-\begin{minipage}[t]{\textwidth}
 \begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ rm -f /var/log/arched.log
-$ arched -c arched_echoservice.xml  && echo jo ||echo n
-$ tail -n100 -f /var/log/arched.log
-$ killall arched
+label=lst:simplelist_example, float=htb,
+caption={[Appearence of the file used by \textit{simplelist.pdp}.]
+         \textbf{Appearence of the file used by \textit{simplelist.pdp}.}}]
+$ cat clientlist.txt
+"/C=US/S=Maine/OU=Literary character/O=Ka-Tet Corp./CN=Roland Deschain" griduser
+"/C=US/S=Maine/OU=Literary character/O=Ka-Tet Corp./CN=Susannah Dean" griduser
+$
 \end{lstlisting}
-\end{minipage}
 
-\lstsetKSH
-\begin{minipage}[t]{\textwidth}
-\begin{lstlisting}[
-label=lst:invokation_arched_timeservice,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-$ ./echoclient
-Tue Feb 17 17:04:08 2009
-\end{lstlisting}
-\end{minipage}
+If the PDP \textit{arc.authz} is used, a bunch of requests will be generated for the client. The requests are compared with the rules defined in the policy and for each rule a decision is made.
+A sample policy is shown in Listing~\ref{lst:sececho_policy_xml}.
+The main element is \textit{policy} which encloses three elements called \textit{Rule}\footnote{The latest XML schema for defining an \textit{arc.pdp} policy can be found at \url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Policy.xsd}}.
+Two types of rules are possible: a permitting or a denying rule.
+A Rule may have four different decision states:
+\begin{itemize}
+ \item PERMIT --- A permitting rule returns the decision permit if one request fits to the rule and 
+the request is fullfilled by the rule. (A denying rule would return deny in that case)
+ \item DENY --- A permitting rule returns the decision deny if one request fits to the rule  and the request is not fullfilled by the rule. (A denying rule would return permit that case)
+ \item INDETERMINATE --- Requests and rule have some parts which not fits
+(i.e. shibboleth identity is checked, but user has identified himself via X.509 and additionally an existing resource is requested)
+ \item NOT\_APPLICABLE --- If the request doesn't fit with the rule at all. (i.e. if the 
+request is for the HTTP method GET, but the rule checks the method POST)
+\end{itemize}
+\forcelinebreak
 
+Once all rules are evaluated, a decision for the policy has to be generated which will depend on the results of the rules.
+The policy element attribute \textit{CombiningAlg} determines how the final result will be generated. Two possibility are given:
+\begin{itemize}
+ \item \textbf{Deny-Overrides} (default) 
+ \begin{itemize}
+	\item If there is at least one DENY in results final result is DENY. 
+	\item Otherwise if there is at least one PERMIT, the final result is PERMIT.
+	\item Otherwise if there is at least one NOT\_APPLICABLE final result is NOT\_APPLICABLE.
+	\item Otherwise final result is INDETERMINATE.
+ \end{itemize}
+  \item \textbf{Permit-Overrides}
+ \begin{itemize}
+ 	\item If there is at least one PERMIT in results final result is PERMIT.
+ 	\item Otherwise if there is at least one DENY the final result is DENY.
+ 	\item Otherwise if there is at least one NOT\_APPLICABLE final result is NOT\_APPLICABLE.
+ 	\item Otherwise final result is INDETERMINATE.
+ \end{itemize}
+\end{itemize}
+\forcelinebreak
+
+A rule is devided into four sub-elements: \textit{Subjects}, \textit{Resources}, \textit{Actions} and \textit{Conditions}. 
+The element \textit{Subjects} defines who shall be affected by that rule.
+The identities can be specified using the DN of the certificate.
+The second element determines the \textit{Resources} which shall be examined by the rule (i.e. using a certain HTTP path).
+The \textit{Actions} are specifing the intended usage of the resource the rule shall survey.
+The last element \textit{Condition} defines somewhat like Time, Duration, namespace of a SOAP message etc. which is needed to be fulfilled.
+For the description of conditions is complicated, they will not be explained in this tutorial.\\
+
+Within the sample policy, which was shown in Listing~\ref{lst:sececho_policy_xml}, three different examples of rule states are prepared: PERMIT, INDETERMINATE and NOT APPLICABLE.
+The first rule, to be seen starting at line~\ref{lst_code:policy.xml_rule1}, is fitting to the server configuration file such that a user with a suitable identity will produce the state PERMIT. 
+Due to there is no rule within the policy which will have the DENY state, this rule let the client pass the PDP.
+The second rule, begins after line~\ref{lst_code:policy.xml_rule2}.
+It is arranged in a manner such that the state INDETERMINATE will be induced.
+Two things are checked within it: The identity based on Shibboleth authentication and the path of the requested HTTP URL.
+Due to the fact, that the HTTP path is fitting to the requested one, but the authentication was not done by Shibboleth, the rule is not able to decide for a DENY or a PERMIT state --- the rule produces an INDETERMINATE state.
+The last rule is defined in line~\ref{lst_code:policy.xml_rule3}. It is constructed such that is permits the access to the HTTP method GET. 
+One can see, that within the server configuration file the Web Service uses the method POST.
+If a client wants to access the Service and reaches the PDP, a couple of requests are generated which are describing the manner the client wants to access the serivice.
+Some of these request will concern the HTTP method GET.
+In these cases, the rule returns NOT APPLICABLE.
+In the cases where the requests for a HTTP method is missing (i.e.
+in the request which only concerns the SOAP access) the state of the rule will be INDETERMINATE.\\
 \lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
-\begin{lstlisting}[
-label=lst:timeservice_cpp_source,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
-	<soap-env:Body>
-		<echoRequest:echo>
-			<echo:say>Der_Berg_ruft</echo:say>
-		</echoRequest:echo>
-	</soap-env:Body>
-</soap-env:Envelope>
-\end{lstlisting}
-\end{minipage}
+\lstinputlisting
+	[
+	label=lst:sececho_policy_xml, float=htb,
+	caption={[Policy file for the \textit{arc.pdp} policy decision point.]
+	\textbf{Policy file for the \textit{arc.pdp} policy decision point.}}
+	]
+{../src/services/secechoservice/policy_example.xml}
 
-\lstsetJUSTXML
-\begin{minipage}[t]{\textwidth}
-\begin{lstlisting}[
-label=lst:timeservice_cpp_source,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
-  <soap-env:Body>
-    <echo:echoResponse>
-      <echo:hear>[ Der_Berg_ruft ]</echo:hear>
-    </echo:echoResponse>
-  </soap-env:Body>
-</soap-env:Envelope>
-\end{lstlisting}
-\end{minipage}
+To get an impression in how the requests look like, one can check the log-file (the log level needs to set to VERBOSE).
+
+\section{Client}
+
+The security introduced to the server has no effect on the client, such that the source code is absolutly identical. Even the client configuration file doesn't need to be changed.

### 2009-03-05 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Finished the Secure Echo Web Service and made first correction of the chapters.

@@
 For simplicity the service will not parse the request but always return the current time.
-It is therefore easier than the example of an Echo Web Service, that also needs to learn (by interpreting the request)
-about what should be echoed.
-In the following subsections the implementation of the service and the client will be presented.
+It is therefore easier than the example of an Echo Web Service which additionally has to extract the incoming message in order to get the string which shall be echoed.
+In the following subsections the implementation of the service and the client will be presented.\\
+
+The source code of the examples given in this tutorial are available in the subdirectory \textit{src}\footnote{If this was not shipped with the document you are reading, then please inspect\\ http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/ws-programming-tutorial}.
 
@@
 
-In ARC the services are implementated as plugins such that they can easily be included or excluded by modifiying the HED configuration file as it was to be seen in Listing~\ref{lst:arched_arcecho_xml}.
+In ARC the services are implementated as plugins such that they can easily be included or excluded by modifiying the server configuration file as it was seen in Listing~\ref{lst:arched_arcecho_xml}.
 The class which is needs to be implemented for the Web Service inherits from the class \textit{Service} which itself is defined in the ARC library.
@@
 	label=lst:time_arched_h,
-	caption={[Header file of the Time Web Service. Filename: timeservice.h]
-	\textbf{Header file of the Time Web Service. Filename: timeservice.h}}
+	caption={[Header file of the Time Web Service.]
+	\textbf{Header file of the Time Web Service.}}
 	]%
 {../src/services/timeservice/timeservice.h}%
-As to be seen, the class \textit{Service} renders the possibility to implement the method \textit{process}.
+The class \textit{Service} renders the possibility to implement the method \textit{process}.
 Listing~\ref{lst:time_arched_cpp} contains the C++ implementation of the Time Web Service plugin.
@@
 
-The Time Web Service uses the namespace \textit{ArcService}.
+The Time Web Service is implemented using the namespace \textit{ArcService}.
 The constructor can be found in line~\ref{lst_code:time_cpp_constructor} and the deconstructor in line~\ref{lst_code:time_cpp_deconstructor}.
@@
 	label=lst:time_arched_cpp,
-	caption={[Implementation of the Time Web Service. Filename: timeservice.cpp]
-	\textbf{Implementation of the Time Web Service. Filename: timeservice.cpp}}
+	caption={[Implementation of the Time Web Service.]
+	\textbf{Implementation of the Time Web Service.}}
 	]%
@@
 
-The method \textit{process} is defined in line~\ref{lst_code:time_cpp_process} and realises the desired service,
-here: the calling of functions provided by system libraries to determine the current time. 
+The method \textit{process}, defined in line~\ref{lst_code:time_cpp_process}, realises the desired service. 
 The parameters \textit{inmsg} and \textit{outmsg} are references to the incoming and outgoing messages.
 Furthermore, the method \textit{process} returns the object \textit{MCCStatus} that is representing the result of the service achieved. 
-Within the method \textit{process}, the service first determines the current time and then creates a SOAP payload using the object \textit{PayloadSOAP} on line~\ref{lst_code:time_cpp_process_payloadSOAP}.
-The object \textit{PayloadSOAP} provides enough functionality to create an XML message that is conform to the SOAP protocol.
+Within the method \textit{process}, the service first determines the current time using a system library and then creates a SOAP payload using the object \textit{PayloadSOAP} on line~\ref{lst_code:time_cpp_process_payloadSOAP}.
+The object \textit{PayloadSOAP} provides enough functionality to create a XML message that is conform to the SOAP protocol.
 The appearance of the generated request and its response will be discussed later on page~\pageref{lst:time_client_request}.\\
@@
 Within the line~\ref{lst_code:time_cpp_process_message_payload} the payload is stored in the message. 
-When nothing unexpected happened while the message is processed, the function returns an \textit{MCCStatus} set to \textit{OK} as seen in line~\ref{lst_code:time_cpp_return}.\\
+When nothing unexpected happened while the message is processed, the function returns an \textit{MCCStatus} which was instanciated with the state \textit{OK}, as to be seen in line~\ref{lst_code:time_cpp_return}.\\
 
@@
 	label=lst:time_arched_xml,float=htb,
-	caption={[HED configuration for the time service. Filename: arched\_timeservice.xml]
-	\textbf{HED configuration for the time service. Filename: arched\_timeservice.xml}}
+	caption={[Configuration for the Time Web Service.]
+	\textbf{Configuration for the Time Web Service.}}
 	]
@@
 
-In order to start the service, the \textit{arched} command which as shown in Listing~\ref{lst:arched_timeservice_ksh} has to be used.
-It is recommended to examine the log file \textit{/var/log/arched.log} if the server configuration file has been modified to  assert the start was successful
+In order to start the service, the \textit{arched} command, shown in Listing~\ref{lst:arched_timeservice_ksh}, has to be used.
+It is recommended to examine the log file \textit{/var/log/arched.log} if the server configuration file has been modified to  assert the start was successful.
 
@@
 label=lst:arched_timeservice_ksh,float=htb,
-caption={[Transformation in eine uniforme konzentrische Verteilung.]
-         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
+caption={[Invokation if the arched daemon containing the Time Web Service.]
+         \textbf{Invokation if the arched daemon containing the Time Web Service}}]
 $ arched -c arched_timeservice.xml && echo Daemon started || echo Daemon start failed
@@
 Due to the configuration of the the \textit{Logger} in the next lines, the messages passed to it will be redirected to the standard error stream.
-For the components of the protocol stack, which  are likewise loaded
+For the components of the protocol stack, which  are likewise loaded dynamically, one needs to specify the installation directory of ARC in line~\ref{lst_code:time_client_cpp_init}.\\
 %\textcolor{white}{newline}
@@
 	label=lst:time_client_cpp,
-	caption={[HED configuration file for the ARC intern echo service. Filename: arcecho\_no\_ssl.xml]
-	\textbf{HED configuration file for the ARC intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
+	caption={[Souce code of the client program.]
+	\textbf{Souce code of the client program}}
 	]
 {../src/clients/timeclient/timeclient.cpp}
- dynamically, one needs to specify the installation directory of ARC in line~\ref{lst_code:time_client_cpp_init}.
 
@@
 The protocol stack is created in line~\ref{lst_code:time_client_cpp_client}. 
-Thanks to the service URL the information for the configuration is complete.\\
+Thanks to the service URL the information for the configuration is completed.\\
 
 The constructor of the class \textit{ClientSOAP} creates a SOAP client which uses HTTP and addresses the port 60000 on localhost.
-If HTTPS would have been used, the stack would have been extended by an additional TLS protocol.
+If HTTPS would have been denoted within the URL, the stack would have been extended by an additional TLS protocol.
 Once the client stack is ready, the request to the service can be created.
@@
 Even the appearance of the message is unimportant to the service, the message is composed of two convoluted elements \textit{time} and \textit{timeReqest}. 
-%
-The request will be processed in the next line by the method \textit{process} of the object \textit{client}.%in line~\ref{lst_code:time_client_cpp_message_response}. 
+The request will be processed in the next line by the method \textit{process} of the object \textit{client}. %in line~\ref{lst_code:time_client_cpp_message_response}. 
 The object of the class \textit{MCCStatus} which is returned by the method indicates the success of the processing.
@@
 
-The client has to be compilied to an ordinary binary so it can directly be executed from the command line.
-Assure that the service is running and use the client as to be seen in Listing~\ref{lst:time_client_ksh}.
+The client program has to be compilied to an ordinary binary such that it can directly be executed from the command line.
+Assure that the service is running and use the client like in Listing~\ref{lst:time_client_ksh}.
 \lstsetKSH
@@
 \end{lstlisting}
-The messages are transmitted in SOAP format and thus reach the SOAP level of the server. They are shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
+the transmitted messages are shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
 The enclosing element is \textit{Envelope} which may contain an element \textit{Header} and must contain an element \textit{Body}.

### 2009-03-05 doc/ws-programming-tutorial/doc/tex_titlepage/titlepage.tex
Finished the Secure Echo Web Service and made first correction of the chapters.

@@
 %%%%---- The title ----
-{\centering \textsc{\Large Webservice programming Tutorial}
-\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under continous development, but don't hesitate to send comments and suggestions to glodek@inb.uni-luebeck.de, I is much appreciated.} % DELETE ONLY THAT LINE!!!
+{\centering \textsc{\Large Web Service programming Tutorial}
+\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
 \Large \par}

### 2009-03-05 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Finished the Secure Echo Web Service and made first correction of the chapters.

@@
 For that reason the current and the next chapter will discuss how to create a secure Web Service.
-In general, security in a network can be subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{verfiy citation}
+In general, security in a network can be subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{TODO: verfiy this citation}
 \begin{itemize}
@@
 In the present chapter, the security will be put into practice by extending the protocol stack with an additional security layer called TLS (Transport Layer Security).
-Due to the location between the TCP and the HTTP layer, the layer is only capable to limit the access to the grid to a certain set of users.
-Once the user has passed the layer sucessfully the connection has been established and he may use all resources provided by the HED.
+Due to the location between the TCP and the HTTP layer, the layer is only capable to limit the access of the grid to a certain set of users.
+Once the user has passed the layer sucessfully, all resources provided by the HED are available.
 This kind of behaviouor can be seen as to be very coarse grained.
@@
 The next section gives a basic understanding of the working principale of the TLS.
-Emphasis is put on the security concept and not on an exact description of the technical realisation in order to enable newcomers an entry to  this area.
+In order to enable newcomers an entry to  this area, emphasis is put on the security concept and not on an exact description of the technical realisation.
 
@@
 Symmetric encryption can be ranked as being very secure but its downside is that the key distribution is very difficult in practice.
-The key has to be shared previously, ideally over a secure channel which may either be a complete diffrent medium (i.e.
+The key has to be shared previously, ideally over a secure channel which may either be a complete different medium (i.e.
 a letter or speech) or a channel which has been encrypted with another secret key.
-Asymmetric encryption follows a diffrent functional principle.
-While the symmetric encryption utilise one shared secret, the asymmetric ecryption is based on two keys: a private key and public key.
-As the names imply, the private key is kept secret and is never shared while the public key is available to everyone.
+Asymmetric encryption follows a different functional principle.
+While the symmetric encryption utilise one kay, the asymmetric ecryption is based on two keys: a private key and public key.
+As the names imply, the private key is kept secret and is never revealed to others while the public key is available for everyone.
 Messages encrypted with the public key can only be decrypted with the private key and conversely, messages encrypted with the private key can only be decrypted with the public key. 
-The Figure~\ref{fig:asymmetric_encryption} illustrates the circumstances of the case.
+The Figure~\ref{fig:asymmetric_encryption} illustrates the circumstances of the case.\\
 \begin{figure}[htb]
 	\centering%epstopdf async.eps
-	\subfloat[Texttext\label{fig:cgd}]
+	\subfloat[\textbf{A plain text which has been encrypted with the private key can only be decrypted with the public key.}
+		If the receivers is sure that the public key is related to a certain identity, 
+			the private key can be used to sign the messages.
+		Only the owner of the private key is able to create encrypted messages which can be decrypted with the public key. \label{fig:cgd}]
 		{\includegraphics[width=12cm]{tex_tls_echoservice/async.pdf}}\\
-	\subfloat[Texttext\label{fig:cgd}]
+	\subfloat[\textbf{A plain text which has been encrypted with the public key can only be decrypted with the private key.}
+		Using asymmetric keys in that manner offers the possiblity to send encrypted messages to the owner of the private key.
+		\label{fig:cgd2}]
 		{\includegraphics[width=12cm]{tex_tls_echoservice/async2.pdf}}
 	\label{fig:asymmetric_encryption}
-	\mycaption{specific}{general}
+	\mycaption{Usage of asymmetric encryption.}{}
 \end{figure}
 
-At startup of the communication process, A (Alice) who desires a secure connection transmits her public key to B (Bob).
+At startup of a communication process, A (Alice) who desires a secure connection transmits her public key to B (Bob).
 Bob is now able to encrypt his messages such that only Alice is able to decrypt them.
 As one can see, this first approach already provides confidentiality and integrity of messages.
-Neither Non-repudiation, authentication and consequently nor access control are established because the identity of Alice is not proven.
-In order to be able to check identities, so called certificates have been introduced.
+But neither Non-repudiation, authentication and consequently nor access control are established because the identity of Alice is not proven.
+In order to be able to check identities, so called certificates have been invented.
 Certificates render the possibility to validate the identity of its owner.
@@
 
-A certificate is composed of the identity of the owner, his public key, the name of the CA who signed the certificate and the signature of the CA which is a hash-value of the certificate encrypted with the private key of the CA.
+A certificate is composed of the identity of its owner, its public key, the name of the CA who signed the certificate and the signature of the CA which is a hash-value of the certificate encrypted with the private key of the CA.
 Furthermore, the algorithms used to create the certificate are added.
@@
 
-\begin{figure}[htb]
+\begin{figure}[tbh]
 	\centering%epstopdf verification.eps 
 	\includegraphics[width=13cm]{tex_tls_echoservice/verification.pdf}
-	\mycaption{specific}{general}
+	\mycaption{Usecase in which Alice is resolving the identity of Bob.}{In order to resolve the identity of Bob, Alice sends a request for Bob's identity along with a random number to Bob.
+Bob encrypts the received number using his private key and returns his public key to Alice.
+Due to the random number can only be encrypted by the owner of the private key, Alice is able ot verify that the certificate was transmitted by his owner.
+Since the certificate was signed by the CA 1 which she is not per-configured to be trust, Alice needs to resolve also the identity of CA 1.
+The request is done in the same manner, but this time the certificate is signed by the trusted CA 2.
+Due to the fact she is now able to trust CA 1, Alice is know able to trust Bob.}
 	\label{fig:verification_of_certificates}
@@
 They are scaling very good with the amount of users because CAs can create new sub-CAs such that the load is balanced (certificates of a CA may be cached).
-Furthermore no critical information is exchanged during the establishment of a connection (in particular no secret key is shared which may get intercepted).
+Furthermore no critical information is exchanged during the establishment of a connection (in particular: no secret key is shared which may get intercepted).
 Instead, a set of pre-configured CAs which will be in general provided by the operation system is needed.
 In respect of ARC it is obvious that the access to the grid shall be under the control of its administrator.
-To do so, the administrators have to establish their own CA which is in charge of creating certificates for a set of users.
+To do so, the administrators have to create their own CAs which is in charge of creating certificates for new users.
+The identity of the created CA has to be added to the directory of pre-configured CAs.
 Many universities have prepared local certificate authorities.
-And every user can establish his own.\\
-
-The identity of the created CA have to be added to the directory of pre-configured CAs.
-\task{The ARC source code comes with two shell scripts to create a CA and related certificates. - really? MG}
-Several publicly availble tools for certificate handling have been prepared, e.g. (like \textit{openssl}),
-which allow the certificate handling also for the advanced UNIX users.
+And every user can establish his own.
 Certificates in general, however, have found entry to regular computer users, for instance with the FernUniversity Hagen (https://ca.fernuni-hagen.de/) who use that technology for their remote students to hand in exercises.\\
@@
 
-In order to create a secure bilateral connection we hence need two certificates: a client certificate and a service certificate.
-The CA for the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key.
+In order to create a secure connection we hence need two certificates: a client certificate and a service certificate.
+The CA of the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key.
 For the given example X.509-certificates in the PEM format will be used which are standardised by the ITU-T.
-Altogether six files are needed:
+Altogether six files are needed:\\
+\\
 \begin{tabular}{l@{~---~}p{13cm}}
@@
 Within the source code directory certificates with corresponding names are prepared.
-Due to certificates have a limited period of validity, these certificates may no longer be accepted by the communication parter.
-In order to create a first basic set of certificates signed by a self created CA, the shell scripts \textit{createMasterCA.sh} and \textit{createSlaveCert.sh} to be found in the directory \textit{src/services/tlsechoservice/certFactory} may be used.
-The script \textit{createMasterCA.sh} will create a new CA and overwrite the old on which was stored in the directory.
-The script \textit{createSlaveCert.sh} can create several certificates signed by the CA which has been create by the other script.\\
-%WARNING: If the programm fails, check if one of your certifacte is expired. If so, create new one using the script files in the directory 'certFactory'
+Due to certificates have a limited period of validity, these certificates may no longer be accepted by the communication parter.\\
+
+Several tools are available which allow the certificate handling for more advanced UNIX users.
+\task{The ARC source code comes with two shell scripts to create a CA and related certificates. - ??MG}
+In order to create a first basic set of certificates signed by a self created CA, the shell scripts \textit{createMasterCA.sh} and \textit{createSlaveCert.sh} can  be found within the directory \textit{src/services/tlsechoservice/certFactory} may be used.
+The script \textit{createMasterCA.sh} will create a new CA and will overwrite older ones.
+The script \textit{createSlaveCert.sh} creates several certificates signed by the CA.\\
 
 The TLS Echo Web Service implementation is almost the same as the one presented in the chapter before.
-Merely the Class name and the namespace of the service has changed.
-The only modification has to be done within the server configuration file which is displayed in Listing~\ref{lst:tls_echo_arched_xml}.
+Merely the Class name and the namespace of the service have changed.
+The only modification is done within the server configuration file which is displayed in Listing~\ref{lst:tls_echo_arched_xml}.
 
@@
 {../src/services/tlsechoservice/arched_tls_echoservice.xml}
-The first change is to be found in line~\ref{lst_code:tls_echo_arched_xml_plugin}.
+The first change is in line~\ref{lst_code:tls_echo_arched_xml_plugin}.
 As to be seen, another plugin called \textit{mcctls} will be loaded by the \textit{ModuleManager}.
 The second change is the new element which has been placed between the TCP and the HTTP component, see line~\ref{lst_code:tls_echo_arched_xml_component}.
-Within the element the files are declared using the elements
-\begin{verbatim}
-<KeyPath>./clientKey.pem</KeyPath>
-<CertificatePath>./clientCert.pem</CertificatePath>
-<CACertificatePath>./serviceCA.pem</CACertificatePath>
-<CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
-\end{verbatim}
-or
-\begin{verbatim}
-<ClientSSLConfig FromFile='filename'/>
-\end{verbatim}
-\task{You can put the ClientSSLConfig into a separate file, then use <ClientSSLConfig FromFile='filename'/> in each service, and then you don't need to change everywhere - proposed by Zsombor - but I can't find any example *hrmlgrmpf*...}
+Within the TLS element the certificates can be declared using the elements \textit{KeyPath}, \textit{CertificatePath},\textit{CACertificatePath} and \textit{CACertificatesDir}.
+% \begin{verbatim}
+% <KeyPath>./clientKey.pem</KeyPath>
+% 
+% <CertificatePath>./clientCert.pem</CertificatePath>
+% 
+% <CACertificatePath>./serviceCA.pem</CACertificatePath>
+% 
+% <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+% \end{verbatim}
+% or
+% \begin{verbatim}
+% <ClientSSLConfig FromFile='filename'/>
+% \end{verbatim}
+% \task{You can put the ClientSSLConfig into a separate file, then use <ClientSSLConfig FromFile='filename'/> in each service, and then you don't need to change everywhere - proposed by Zsombor - but I can't find any example *hrmlgrmpf*...}
 
@@
 In this section a new way to implement the client will be introduced.
-As already mentioned in section~\ref{sec:timeservice_client}, the protocol stack of the client is build up in almost the same way like the one of the service.
+As already mentioned in section~\ref{sec:timeservice_client}, the protocol stack of the client is build up in almost the same way as the one of the service.
 Indeed, the class~\textit{ClientSOAP} composes the stack with the same MCCs used for the service.
 Once this is known, its self-evident to create the client in a simlar manner like the service with an own configuration file.
-Listing~\ref{lst:tls_echo_client_cpp} shows the C++ source code of the client implementation.
-The main difference between the simple Echo Client presented in the chapter before is that now the information about the client protocal stack is encapsulated within a configuration file.
+Listing~\ref{lst:tls_echo_client_cpp} shows the source code of the client program.
+The main difference between the simple Echo Client presented in the chapter before is that now the information about the client protocol stack is encapsulated within a configuration file.\\
 \lstsetCPP
@@
 	label=lst:tls_echo_client_cpp,
-	caption={[Client configuration file]
-	\textbf{Client configuration file\textcolor{white}{hmf}}}
+	caption={[Client programm which is capable to load a configuration file]
+	\textbf{Client programm which is capable to load a configuration file}}
 	]
 {../src/clients/tlsechoclient/tlsechoclient.cpp}
+
 The file containing the client configuration is specified within the commandline. 
-Its content will be loaded in the code fragment starting in line~\ref{lst_code:tls_echo_client_cpp_getfilecontent}. 
+Its content will be loaded in the code fragment starting at line~\ref{lst_code:tls_echo_client_cpp_getfilecontent}. 
 Later, in line~\ref{lst_code:tls_echo_client_cpp_loadconfiguration}, the configuration will be parsed into XML and than transfered into a chain of components which is wrapped by the object \textit{loader}.
@@
 To get the proper component an operator of MCCLoader is used in line~\ref{lst_code:tls_echo_client_cpp_cliententry}.
-The previously used class \textit{ClientSOAP} has eased some workload which now has to be done within the method.
-The payload now have to be wrapped by the message class which have to be initialised with an \textit{MessageAttribute} object and a \textit{MessageContext} object.
+The previously used class \textit{ClientSOAP} has eased some workload which now has to be done within the client program.
+The payloads now have to be wrapped by message objects which have to be initialised with an \textit{MessageAttribute} object and a \textit{MessageContext} object.
 After the payload has been wrapped within the message it gets transmitted in almost the same way like before, see lines followed by line~\ref{lst_code:tls_echo_client_cpp_createandprocess}.

### 2009-03-05 doc/ws-programming-tutorial/src/services/tlsechoservice/tlsechoservice.cpp

@@ -67,16 +67,14 @@ namespace ArcService
 	/**
 	* Method which creates a fault payload 
 	*/
-	Arc::MCC_Status TLSEchoService::makeFault(Arc::Message& outmsg) 
+	Arc::MCC_Status TLSEchoService::makeFault(Arc::Message& outmsg, const std::string &reason) 
 	{
-
-		// The boolean true indicates that inside of PayloadSOAP, 
-		// an object SOAPFault will be created inside.
+		logger.msg(Arc::WARNING, "Creating fault! Reason: \"%s\"",reason);
 		Arc::PayloadSOAP* outpayload = new Arc::PayloadSOAP(ns_,true);
 		Arc::SOAPFault* fault = outpayload->Fault();
 		if(fault) {
 			fault->Code(Arc::SOAPFault::Sender);
-			fault->Reason("Failed processing request");
+			fault->Reason(reason);
 		};
 		outmsg.Payload(outpayload);
 		return Arc::MCC_Status(Arc::STATUS_OK);
@@ -92,7 +90,8 @@ namespace ArcService
 	*/
 	Arc::MCC_Status TLSEchoService::process(Arc::Message& inmsg, Arc::Message& outmsg) 
 	{
-		logger.msg(Arc::DEBUG, "TLS Echoservice has been started...");
+		logger.msg(Arc::DEBUG, "TLS Echo Web Service has been started...");
+		logger.msg(Arc::DEBUG, "TLS Echo Web Service: Current Client has the DN: \"%s\"",inmsg.Attributes()->get("TLS:PEERDN"));//(@*\label{lst_code:TLS_service_cpp_DN}*@)
 
 		/** Both input and output are supposed to be SOAP */
 		Arc::PayloadSOAP* inpayload  = NULL;
@@ -105,32 +104,7 @@ namespace ArcService
 		} catch(std::exception& e) { };
 		if(!inpayload) {
 			logger.msg(Arc::ERROR, "Input is not SOAP");
-			return makeFault(outmsg);
-		};
-		/** */
-
-		/**Export the formated policy-decision request**/
-		MessageAuth* mauth = inmsg.Auth();
-		MessageAuth* cauth = inmsg.AuthContext();
-		if((!mauth) && (!cauth)) {
-			logger.msg(ERROR,"Missing security object in message");
-			return Arc::MCC_Status();
-		};
-		NS ns;
-		XMLNode requestxml(ns,"");
-		if(mauth) {
[…]

### 2009-03-09 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Correction of the first pages.

@@
 These are executed on the same machine that the server is installed on.
-Some services, like the a-rex service, will delegate computational efforts to other machines
+Some services, like the A-REX service, will delegate computational efforts to other machines
 via additional software like a local queueing system.
-But the a-rex service itself does not migrate away to other machines - at least this
+But the A-REX service itself does not migrate away to other machines - at least this
 has not been implemented yet.
 Services may comprise computational tasks (e.g., some secret and strongly
-patented algorithm), allow arbitrary computations (like ARC's a-rex or
+patented algorithm), allow arbitrary computations (like ARC's A-REX or
 Amazon's clouds) or provide access to other resources like a particular
@@
 Servers provide access to a single service or a set of services.
-For ARC, there is one service, named Hosting Environment Demon (HED), more special than other services: HED organises the assignment of other services on the same server to the network.
+For ARC, there is one service, named Hosting Environment Demon (HED), that is more special than other services: HED organises the assignment of other services on the same server to the network.
 But this does not break the concept, it is only a commodity.
@@
 The core of ARC is the prior mentioned HED (Hosting Environment Daemon).
-To the administrator HED is mostly visible as the single binary that is started and configured by a file that describes services that shall be prepared by HED. 
+To the administrator HED is mostly visible as the single binary that is started and configured by a file that describes services' that shall be prepared by HED. 
 Conceptionally, the HED determines how services shall be organised both as a principle, and for any given ARC-run server.
@@
 	\centering
-	\includegraphics[width=14cm]{tex_introduction/HED2.pdf}
+	\includegraphics[width=13cm]{tex_introduction/HED2.pdf}
 	\mycaption{Example of a HED structure providing a Web Service}{The HED consists of several layers which are connected with each other.
@@
 The lowest layer has to be a transport protocol like TCP which enables the transmission between two applications.
-Technically there even lower layers required needed for the communication but they are already realised within the operating system and the  network card.
+Technically there are even lower layers required needed for the communication but they are already realised within the operating system and the  network card.
 TCP is well suited for the application because it provides a reliable and ordered delivery of a byte stream.
@@
 
-\subsection{The ARC echo service}
+\subsection{The ARC Echo Web Service}
 
 The first example shall give a basic understanding in how to configure the HED for setting up a novel service.
-The task will be to setup the regular ARC Echo Web Service.
-On the command line, all that needs to be done is to invoking the \textit{arched} daemon with a fitting server configuration file.
+The task is to setup the regular ARC Echo Web Service.
+On the command line, all that needs to be done is to invoke the \textit{arched} daemon with a suitable server configuration file.
 Such a fitting server configuration file is shown in Listing~\ref{lst:arched_arcecho_xml}.
@@
 The configuration as a whole is encapsulated by the element \textit{ArcConfig}.
-It also performs the setting of namespace abbreviations, which are expected to be mostly invariant across all server installations.% ABBREVIATIONS??
+It also performs the setting of namespaces, which are expected to be mostly invariant across all server installations.
+% ABBREVIATIONS??
 - MG
-The \textit{Server} element, to be found at line~\ref{lst_code:arched_arcecho_Server}, provides basic settings for the daemon such as location of the Pid-file or the Log-file.
-The \textit{ModuleManager} holds the path to the plugin libraries.
+The \textit{Server} element, to be found at line~\ref{lst_code:arched_arcecho_Server}, provides basic settings for the daemon such as the location of the Pid-file or ofthe Log-file.
+The \textit{ModuleManager} holds the \textit{path} to the plugin libraries.
 Several paths may be defined here.
-Due to the fact that the first library fitting to the plugin name is loaded, the order is relevant.
+Due to the fact that the first library matching the plugin name is loaded, the order is relevant.
 The path should be at least assigned to the ARC installation directory.
-Otherwise, the standard set of MCC modules might not be found.
-The next set of elements are holding the name of the plugins to be loaded (line~\ref{lst_code:arched_arcecho_Plugins}).
-The names have to correspond to the name of the dynamic libraries within the path defined in the \textit{ModuleManager}.
-Finally, the chain is declared within the element \textit{Chain} which creates the layered structure.
+Otherwise, the set of ARC MCC plugins might not be found.
+The next elements are holding the name of the plugins to be loaded (line~\ref{lst_code:arched_arcecho_Plugins}).
+The names have to correspond to the names of the dynamic libraries within the path defined in the \textit{ModuleManager}.
+Finally, the chain, which creates the layered structure, is declared within the element \textit{Chain}.
 It is composed of the elements \textit{Component}, \textit{Plexer} and \textit{Service}.
-Every element inside has at least a \textit{name} attribute.
-Regarding the components and the services a plugin (*.so file found in the paths previously specified) has to be loaded which provides the implementation of its functionality.
-The further configuration depends on the individual functionality.\\
+Due to a MCC plugin may contain several components, one has to specify more exact which one to load.
+This is to be done using the \textit{name} attribute i.e.
+the plugin \textit{mcchttp} realises two components: \textit{http.service} and \textit{http.client}.
+Since we are defining a configuration file for a service, the \textit{http.service} should be used.
+Furthermore the elements may have the attribute \textit{id} as a unique identifier within the file.
+Regarding the components and the services a plugin (implemented as a *.so file that is searched in the paths previously specified) has to be loaded which provides the implementation of its functionality.
+The further configuration depends on the plugin's individual functionality.\\
 %
@@
 %The Components refer to the implementation of Web Services within any of those libraries.
-%
-%
+
 A \textit{Chain} will determine the path that an event takes to be noticed.
@@
 
-$ arched -c arcecho_no_ssl.xml && echo Daemon started || echo Daemon start failed
+$ arched -c arcecho_no_ssl.xml && echo Daemon started || echo Daemon failed
  Daemon started
@@
 The name of the program is followed by the parameter \textit{-c}.
-The string following that argument contains the full path of configuration file.
+The string following that argument contains the path to the configuration file.
 Once the server is running, the client can be used as described in Listing~\ref{lst:arcecho_client_invokation}.
@@
 For debugging it is advisable to check the logfile which was assigned inside the server configuration file.
-It is commonly found at /var/log/arched.log
+It is commonly found at /var/log/arched.log.
 

### 2009-03-10 doc/storage_system/arc-storage-documentation.tex
Description of replicated A-Hash Some random spelling corrections

@@
 The storage system is capable of storing files which can be grouped in collections and sub-collections, etc.
-Every file and collection has a unique ID in the sysem called the \emph{GUID}.
+Every file and collection has a unique ID in the system called the \emph{GUID}.
 Compared to the well-known structure of local file systems, these GUIDs are very similar to the concept of \emph{inode}s.
@@
 The collections in the ARC storage are logical entities, the content of a collection is stored as metadata of the collection, which means the a collection actually has no physical data.
-A file however has both metadata and real physicial data (the actual bytes of the file).
+A file however has both metadata and real physical data (the actual bytes of the file).
 The metadata of a file is stored in the same database where the collections are stored, but the physical data of a file is stored on storage nodes as multiple replicated copies.
@@
 So we have logical files, which are part of the hierarchical namespace and have a GUID and other metadata, and a logical file has one or more physical replicas.
-The physical replicas stored on seperate storage nodes.
+The physical replicas stored on separate storage nodes.
 In order to connect the logical file to its replicas, we need to have some pointers.
@@
 Most of the services (A-Hash, Librarian, Shepherd) are `internal' services in a way that the end-user of the storage system never communicates with them directly.
-But these internal services are communicating with eachother, so they have to know who to trust.
+But these internal services are communicating with each other, so they have to know who to trust.
 We call this aspect of the security architecture of the ARC storage `inter-service authorization'.
@@
 
-For transfering the actual file data, the users have to connect to storage element services which are sitting on storage nodes.
+For transferring the actual file data, the users have to connect to storage element services which are sitting on storage nodes.
 These services also have their own authentication and authorization methods.
-Managing these aspects is the `transfer-level authorization' part of the security architucture of the ARC storage.
+Managing these aspects is the `transfer-level authorization' part of the security architecture of the ARC storage.
 
@@
 In a deployment of the ARC storage system, we could have several A-Hash, Librarian, Shepherd and Bartender services.
-The Bartenders send requests to the Librarians and the Shepherds, the Shepherds communciate with the Librarians, the Librarians talk with the A-Hashes.
+The Bartenders send requests to the Librarians and the Shepherds, the Shepherds communicate with the Librarians, the Librarians talk with the A-Hashes.
 If any of these services get compromised or a new rouge service gets inserted in the system, we loose security completely.
 That's why it is vital for each service to authorize the services before sending or accepting requests.
-The services commincates via HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and they can examine the other service's certificates.
+The services communicates via HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and they can examine the other service's certificates.
 Because of these X.509 certificates each service has Distinguish Name (DN).
@@
 
-Additionaly, each file and collection has an `owner' which is a user who always can modify the access control rules.
+Additionally, each file and collection has an `owner' which is a user who always can modify the access control rules.
 
@@
 When the Bartender decides that a user has permission to download a file, then the Bartender chooses a replica, and initiates the transfer.
-The result of this initiation is a URL which is colled the transfer URL (TURL).
+The result of this initiation is a URL which is called the transfer URL (TURL).
 This TURL is unique for each request, even for request to the same replica, and this TURL is only valid for one download.
@@
     \item Hopefully the \verb!ourvo! collection can be found in the root collection, which means that new the Librarian knows its GUID, and can ask for its metadata from the A-Hash.
-    \item After the A-Hash returns the metadata of the \verb!/ourvo! collection, the Librarain finds the GUID of \verb!users! in it, then gets its metadata.
+    \item After the A-Hash returns the metadata of the \verb!/ourvo! collection, the Librarian finds the GUID of \verb!users! in it, then gets its metadata.
     \item The A-Hash returns the metadata of \verb!/ourvo/users! which contains the GUID of \verb!we!, so the Librarian can ask for its metadata.
@@
 
-Because the next-generation information system of ARC is currently under development, we cannot use it to discover services, that's why currently the URLs of almost all the services are hardcoded in the configuration files of the service.
+Because the next-generation information system of ARC is currently under development, we cannot use it to discover services, that's why currently the URLs of almost all the services are hard-coded in the configuration files of the service.
 There is one exception: the Shepherd services are reporting their URLs to the Librarians, so a Bartender always could get a very current list of alive Shepherd services.
@@
 The properties, the values and the section names too are simple character strings.
-The A-Hash services provide a simple interface to interact with these objects, e.g.~to do conditinal and atomic changes.
+The A-Hash services provide a simple interface to interact with these objects, e.g.~to do conditional and atomic changes.
 
@@
         \item \emph{GUID}: the unique ID of the entry
-        \item \emph{owner}: the DN of the user how ownes this entry
+        \item \emph{owner}: the DN of the user how owns this entry
     \end{itemize}
@@
     \begin{itemize}
-        \item \emph{(identity, action list) pairs}: a list of allowed and not allowed actions for different identites (users, VOs, etc.). 
+        \item \emph{(identity, action list) pairs}: a list of allowed and not allowed actions for different identities (users, VOs, etc.). 
     \end{itemize}
@@
 This state could be `\textbf{alive}' (if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the Shepherd claims it has no such file), `\textbf{offline}' (if the Shepherd is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{thirdwheel}' (if the replica is marked for deletion because of too many replicas).
-We also store (in the \emph{states} section) the size of the file, the number of needed replicas, and a checksum of the file, which is created by some kind of checksuming algorithm (currently only md5 is supported).
+We also store (in the \emph{states} section) the size of the file, the number of needed replicas, and a checksum of the file, which is created by some kind of checksumming algorithm (currently only md5 is supported).
 
@@
 If you close a collection then it cannot be opened again.
-However it cannot be garanteed that the contents of a closed collection remains the same (e.g.
+However it cannot be guaranteed that the contents of a closed collection remains the same (e.g.
 if we include in our collection a file which we don't own, then the owner of the file could remove it nevertheless), but if a closed collection is changed, its state will be broken, and this state could never be changed again, which means you will always know if a closed collection is not intact anymore.
@@
 
-This can be put in a seperate file, and then use the \textbf{FromFile} attribute of the ClientSSLConfig section to specify where the file is.
+This can be put in a separate file, and then use the \textbf{FromFile} attribute of the ClientSSLConfig section to specify where the file is.
 
@@
 \begin{description}
-    \item[ReplicatedAHash] TODO: describe ReplicatedAHash, sendMessage, processMessage
-    \item[ReplicationStore] TODO: describe ReplicationStore, connection with TransDBStore, config, etc.
-    \item[ReplicationManager] TODO: describe ReplicationManager, elections, communication (interservice communication, security, callback functions)
+    \item[ReplicatedAHash] %TODO: describe ReplicatedAHash, sendMessage, processMessage
+	is a subclass of the \verb!CentralizedAHash!
+business logic class and implements the extra logic needed for communication between A-Hash instances.
+The \verb!ReplicatedAHash!
+adds the method \verb!sendMessage!
+to the A-Hash interface and has a method \verb!processMessage!
+for processing messages from \verb!sendMessage!.
+These two methods are needed for the communication framework of the \verb!ReplicationManager!.
+The actual replication is managed by the \verb!ReplicationStore!
+and \verb!ReplicationManager!, meaning that the \verb!ReplicatedAHash!, with the exception of the above mentioned methods, is actually just a centralized A-Hash with a specific storage module.
+    \item[ReplicationStore] %TODO: describe ReplicationStore, connection with TransDBStore, config, etc.
+	is a specialized replicated DB.
+It is a subclass of \verb!TransDBStore!\footnote{A module for storing data on disk transactionally using the Berkeley DB library.} with some additional features for handling replication.
+The \verb!ReplicationStore!
+periodically updates an A-Hash object containing information about the current master A-Hash and available clients.
+In addition it contains a \verb!ReplicationManager!
+object which governs the replication framework.
+    \item[ReplicationManager] %TODO: describe ReplicationManager, elections, communication (interservice communication, security, callback functions)
+	handles all aspects of the Berkeley DB API and takes care of replicas, elections and replication messages.
+	\begin{itemize}
+	\item The replicated A-Hash consists of a set of replicas; one master and multiple clients.
+This means that the entire database is fully replicated on all replicas.
+When the master receives a write message, it will broadcast this message to all clients.
+To make sure that the db is consistent and that there will always be possible to elect a master, the master requires an acknowledgement message from a quorum of clients before continuing.
+	\item In the replicated A-Hash there is one master and multiple clients.
+When the A-Hash starts up or if a master goes offline, we need to find a new master.
+This is solved by the clients holding an election, according to the Paxos algorithm.
+Simply told, all clients sends a vote to all other clients, and the eligible client that gets the most votes wins the election.
+An eligible client is a client which has received the latest updates.
+(See the Berkeley DB documentation for more details.) There are cases where some servers are more suitable to be master than others (better hardware, faster Ethernet, etc.).
+In such cases you can influence the election by setting a higher priority on the best servers.
+While an election is being held, it is not possible to write or read from the A-Hash, as consistency cannot be guaranteed without a master.
+	\item As the replicas in the replicated A-Hash are services running within HED, the replication messages needs to be sent through the communication framework of HED.
+To be able to use this framework, the Berkeley DB needs a callback function for sending messages and the \verb!ReplicationManager!
+needs to call Berkeley DB to process incoming messages.
+The \verb!ReplicationMessage!
+takes a send function as one of the initialization arguments, and calls this function in the callback method.
+By providing the \verb!ReplicatedAHash.sendMessage!
+to the \verb!ReplicationManager!
+messages will be sent through the HED communication services just as in regular inter service communication.
+When \verb!ReplicatedAHash!
+receives a replication message from \verb!sendMessage!
+it will call \verb!ReplicationManager.processMessage!
+which in turn will call the Berkeley DB processMessage.
+This communication framework is both used by Berkeley DB and directly by the \verb!ReplicationManager!
+for, e.g., bootstrapping the A-Hash replicas and to communicate newly discovered clients.
 \end{description}
@@
 Currently there are several versions (\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore!
-is suitable for using in a replicated deployemant, all the others are only for a centralized deployement.
+is suitable for using in a replicated deployment, all the others are only for a centralized deployement.
 This variable is only configurable for the centralized deployement.
@@
     \item[modify(modifyRequestList)] modifies the metadata of files, collections and mount poins.
-Only the `states', the `policy' and the `metadata' sections can be modified, and there are seperate access control actions for the three.
+Only the `states', the `policy' and the `metadata' sections can be modified, and there are separate access control actions for the three.
 

### 2009-03-12 doc/client/client_technical.tex
Updating resubmission part

@@
 \newcommand{\TargetRetriever}{\texttt{TargetRetriever}}
+\newcommand{\Sandbox}{\texttt{Sandbox}}
 \newcommand{\Submitter}{\texttt{Submitter}}
@@
 Once a job is submitted job related information (job identification
-string, cluster etc) is stored in a local XML file which stores this 
+string, cluster etc.) is stored in a local XML file which stores this 
 information for all active jobs. This file may contain jobs running on
@@
 specialized {\Broker} and it implements the method for reducing a list of resources found by resource discovery (the list of 
-{\ExecutionTarget}s residing witin the {\TargetGenerator}) to a list of resources capable of running a certain job:
+{\ExecutionTarget}s residing within the {\TargetGenerator}) to a list of resources capable of running a certain job:
 
@@
 The {\texttt PreFilterTargets} method compares every hardware and software requirement given in the job description against the 
-computing resource (cluster) specifications stored in the {\ExecutionTarget}. If the{\ExecutionTarget} doesn't fulfill 
+computing resource (cluster) specifications stored in the {\ExecutionTarget}. If the{\ExecutionTarget} doesn't fulfil 
 the requirements imposed by the job description, or it is impossible to carry out the matchmaking due to incomplete information 
@@
 for job submission is known the parsed and (if needed) translated job description 
-can be retrieved through the {\texttt GetProduct} funcion as shown below:
+can be retrieved through the {\texttt GetProduct} function as shown below:
 
@@
 
-If the inner representation is empty or the source job descrition is not
+If the inner representation is empty or the source job description is not
 valid, then the {\texttt GetProduct} output is empty too and the method 
-returns with false. When the product format is the same as the source format 
-and not XRSL, than the product string is set to the source string. The XRSL 
-output is always generated from the inner representation, as it includes 
-automatic job description modifications similar to those of the production 
-ARC client. If the product format differs from the source format the 
-{\texttt GetProduct} method calls upon the \texttt{JobDescriptionOrderer} 
-class to carry out the translation.
+returns with false.  In order to obtain the automatic job description 
+modifications similar to those of the production ARC client, the 
+{\texttt GetProduct} output is always generated from the internal 
+representation using the \texttt{JobDescriptionOrderer} class to carry 
+out the translation. The original source string can be regained by
+calling the \texttt{getSourceString} method:
+
+\begin{shaded}
+\begin{verbatim}
+bool getSourceString(std::string source) const;
+\end{verbatim}
+\end{shaded}
 
@@
 attribute has no equivalent in the output language. In these events the {\texttt GetProduct} 
-method returns with a false and an empty product stringt. For limitations in the translation
+method returns with a false and an empty product string. For limitations in the translation
 see Section~\ref{sec:jobinnerrepresentation}.
@@
 attributes are not implemented.
+
+In case of job resubmission the internal representation will contain a list of 
+the jobid's of the previous jobs. This element will only be parsed to the 
+actually job description in case JSDL-POSIX output is requested. A new URL 
+can be added to the list of old jobid's by using the method:
+
+\begin{shaded}
+\begin{verbatim}
+bool addOldJobID(const URL& oldjobid);
+\end{verbatim}
+\end{shaded}
 
@@
 which is now ready for handling jobs. The public functions of the
-{\JobController} offer to get (download), clean, cancel, etc one or
+{\JobController} offer to get (download), clean, cancel, etc. one or
 more jobs and uses the private specializations for issuing the
@@
 
+The other public functions of the {\JobController} class are:
+The \texttt{Get} which is used to download jobs
+\begin{shaded}
+\begin{verbatim}
+bool Get(const std::list<std::string>& status, const std::string& downloaddir,
+         const bool keep, const int timeout);
+\end{verbatim}
+\end{shaded}
+\begin{shaded}
+\begin{verbatim}
+bool Kill(const std::list<std::string>& status, const bool keep, const int timeout);
+\end{verbatim}
+\end{shaded}
+\begin{shaded}
+\begin{verbatim}
+bool Clean(const std::list<std::string>& status, const bool force, const int timeout);
+\end{verbatim}
+\end{shaded}
+\begin{shaded}
+\begin{verbatim}
+bool Cat(const std::list<std::string>& status, const std::string& whichfile, 
+         const int timeout);
+\end{verbatim}
+\end{shaded}
+\begin{shaded}
+\begin{verbatim}
+bool RemoveJobs(const std::list<URL>& jobids);
+\end{verbatim}
+\end{shaded}
+\begin{shaded}
+\begin{verbatim}
+std::list<std::string> GetDownloadFiles(const URL& dir);
+\end{verbatim}
+\end{shaded}
+\begin{shaded}
+\begin{verbatim}
+bool CopyFile(const URL& src, const URL& dst);
+\end{verbatim}
+\end{shaded}
+The \texttt{GetJobDescriptions} method 
+\begin{shaded}
+\begin{verbatim}
+std::list<Job> GetJobDescriptions(const std::list<std::string>& status, 
+                                  const bool getlocal, const int timeout);
+\end{verbatim}
+\end{shaded}
+which is used to get the job descriptions of {\Job}s in the \texttt{JobStore} 
+with the specified status'. In case \texttt{getlocal} is true the job description 
+will first be searched for the users local XML file. If a local copy of 
+the job description is found the checksum of the local input files will be 
+computed and compared with the ones listed in the users local XML file. If 
+all checksums match the job description is returned.\\
+In case \texttt{getlocal} is false or the retrieval of the local job description 
+fails the job description will be tried retrieved from the cluster. This has 
+currently only been implemented for ARC(both generations) clusters. Note 
+that the job descriptions retried from a cluster might will most likely have 
+been changed since the submission either by the cluster\footnote{For both 
+generations of ARC it is possible to recover the initial job description send 
+to the cluster. However modifications can still have been made to the 
+job description before sending it to the cluster.} or by the submission 
+client. Therefore the only way of retrieving the original job description is 
+through the users local XML file.\\
+The \texttt{GetJobDescriptions} returns a list of the  {\Job}s where job description 
+have been found. The found job descriptions are kept in the 
+{\Job}.\texttt{JobDescription} attribute.
+
 \subsection{{\Job}}
@@
 const std::string& JobListFile() const;
-const XMLNode& ConfTree() const;
+const & ConfTree() const;
 bool ApplySecurity(XMLNode& ccfg) const;
@@
 
+\subsection{{\Sandbox}}
+The {\Sandbox} class, consisting of the two methods \texttt{Add} and 
+\texttt{GetCkSum}, have been created to assure a reliable resubmission 
+of jobs. 
+\begin{shaded}
+\begin{verbatim}
+static bool Add(JobDescription& jobdesc, XMLNode& info);
+static std::string GetCksum(const std::string& file);
+\end{verbatim}
+\end{shaded}
+The \texttt{Add} method  adds the job description and the checksum of 
+the local input files to a {\XMLNode}. Currently this {\XMLNode} is the one 
+stored in the user's local XML file upon job submission. The job 
+description which is stored is the initial job description before any 
+parsing has been performed. To calculate the checksum of the input 
+files the \texttt{GetCkSum} method is used. The \texttt{GetCkSum} 
+method returns the MD5 checksum of a given file
+
 \section{Specialized Classes (Grid Flavour and Broker plugins)}
@@
 they have. E.g. if two {\ExecutionTarget}s have zero queue and they have 90 and 10 free slots respectively, then the first
-{\ExecutionTarget} has a 90\% propability of becoming the top ranked cluster.}
+{\ExecutionTarget} has a 90\% probability of becoming the top ranked cluster.}
 \end{enumerate}
@@
   running grid job's std out or std error file.}
+\item{\texttt{arcsync} - for synchronising the information in the user's
+  local XML file about active jobs.}
+\item{\texttt{arcresub} - for redoing an entire job on the same or a new
+  computing resource.}
+\item{\texttt{arcmigrate} - for moving a queuing job from an ARC1 
+  computing resource to another.}
 \end{enumerate}

### 2009-03-13 doc/client/client_technical.tex
spelling corrections

@@
 
-This document describes from a technical viewpoint a plugin based
+This document describes from a technical viewpoint a plugin-based
 client library for the new Web Service (WS) based Advanced Resource
@@
 \item{handling proxy, user and host certificates,}
-\item{performing resource discovery and information retrieval,}
-\item{filtering and brokering of found resources,}
+\item{performing computing resource discovery and information retrieval,}
+\item{filtering and brokering of found computing resources,}
 \item{job submission and management and}
@@
 extensions to support additional middlewares involve plugin
-compilation only i.e.\ no recompilation of main libraries or clients
+compilation only i.e., no recompilation of main libraries or clients
 is necessary.
@@
 in Section~\ref{sec:Implementation}. Section~\ref{sec:cli} will
-through examples show how command line interfaces can be built upon
+show through examples how command line interfaces can be built upon
 the library.
@@
 \item{If while processing the query result the {\TargetRetriever}
-  finds an other registered index server or a registered computing
+  finds another registered index server or a registered computing
   service it creates a new {\TargetRetriever} for the found server or
@@
 preparation as outlined in the Section~\ref{sec:TargetDiscovery}. Not
-until a list of possible targets (which allows the user) is available
+until a list of possible targets (which authorize the user) is available
 is the job description read in order to enable bulk job submission of
@@
 
-Once a job is submitted job related information (job identification
+Once a job is submitted, job related information (job identification
 string, cluster etc.) is stored in a local XML file which stores this 
@@
 
-The {\TargetRetriever} is base class for {\TargetRetriever} grid
-middleware specializations and inherits from the {\ACC} base class in
+The {\TargetRetriever} is the base class for grid middleware specialized 
+{\TargetRetriever}s and inherits from the {\ACC} base class in
 order to be loadable by the ARC \texttt{Loader}. It is designed to
@@
 
-Following a broker ranking jobs are to be submitted. Since all
-information about the selected computing service resides within the
-selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of
-returning a {\Submitter} capable of submitting a job to the service it
-represents.
+All information needed to submit a job is stored in the {\ExecutionTarget}, 
+and the {\ExecutionTarget}'s method
 
@@
 
-The {\UserConfig} object is used to configure the security related
-information needed by the {\Submitter}.
+returns a {\Submitter} which is used to submit jobs. The {\UserConfig} object 
+is used to configure the security related information needed by the {\Submitter}.
 
@@
 published by the computing resource, the {\ExecutionTarget} will be removed from the list of possible targets. A detailed overview 
-of the matchmaking between {\JobDescription} and {\ExecutionTarget} attributes is given in Appendix~\ref{sec:broker-mapping}.
-
-Once prefiltered the remaining {\ExecutionTarget}s should be ranked in order to return the ``best'' {\ExecutionTarget} for 
+of the matchmaking between {\JobDescription} and {\ExecutionTarget} attributes is given in Appendix~\ref{app:broker-mapping}.
+
+Once prefiltered, the remaining {\ExecutionTarget}s should be ranked in order to return the ``best'' {\ExecutionTarget} for 
 job submission. Different ranking methods are implemented by the specialized brokers, but for usability and consistency reasons 
@@
 
-The identification of the source description format is internally handled by a 
-\texttt{JobDescriptionOrderer} class which identifies the format by pattern matching.
-Once the format is determined from the elements are mapped from the original format (XRSL, JDL or JSDL) 
-into the internal representation (the \texttt{JobInnerRepresentation} class). Mapping details 
-are given in Section~\ref{sec:jobinnerrepresentation}.
+The setSource method takes a source job description string as input and identifies it's format 
+(language) through pattern matching (the \texttt{JobDescriptionOrderer} class).
+Once the format is known, the source elements are mapped from the original format (XRSL, JDL or JSDL) 
+into an internal representation (the \texttt{JobInnerRepresentation} class). Mapping details 
+are given in Section~\ref{app:jobinnerrepresentation}.
 
@@
 Once the job description language required by the {\ExecutionTarget} chosen 
-for job submission is known the parsed and (if needed) translated job description 
+for job submission is known, the parsed and (if needed) translated job description 
 can be retrieved through the {\texttt GetProduct} function as shown below:
@@
 of the job description from one language to another. Thus the {\JobDescription}
-class can do nothing in those cases where there is not enough data is available 
+class can do nothing in those cases where there is not enough data available 
 to assemble a given attribute or when information can be lost because the received 
-attribute has no equivalent in the output language. In these events the {\texttt GetProduct} 
-method returns with a false and an empty product string. For limitations in the translation
-see Section~\ref{sec:jobinnerrepresentation}.
+attribute has no equivalent in the output language. In these cases the {\texttt GetProduct} 
+method returns with a false value and an empty product string. For limitations in the translation
+see Section~\ref{app:jobinnerrepresentation}.
 
@@
 
-{\Submitter} is base class for grid specific specializations (plugin).
-It submits job(s) to the service it represents and uploads (by the job
-needed) local input files.
+{\Submitter} is the base class for grid middleware specialized {\Submitter}s (plugins).
+It submits job(s) to the service it represents and uploads (needed by the job) local 
+input files.
 
@@
 the {\JobSupervisor} makes extensive use of the local joblist file
-housing information about all active jobs. Thus the {\JobSupervisor}
+which house information about all active jobs. Thus the {\JobSupervisor}
 constructor becomes
@@
 
-Although being loaded by the {\JobSupervisor} the {\JobController}
-objects truly resides within the ARC {\Loader} which is a member of
+Although loaded by the {\JobSupervisor} the {\JobController}
+objects actually reside within the ARC {\Loader} which is a member of
 the {\JobSupervisor} class. In order to get handles on the
-{\JobController}s the inline method
+{\JobController}s, the inline method
 
@@
 file containing information about all active jobs (flavour
-independent). Next step is the filling of the {\JobController}'s job
+independent). The next step is the filling of the {\JobController}'s job
 pool \texttt{JobStore} which is the pool of jobs that the
@@
 \item{If the \texttt{jobids} list has entries, fill \texttt{JobStore}
-  with the by user requested jobs.}
+  with the jobs requested by the user.}
 \item{If the \texttt{clusterselect} list has entries, fill
@@
   clusters.}
-\item{If \texttt{jobs} and \texttt{clusterselect} are both empty
+\item{If \texttt{jobids} and \texttt{clusterselect} are both empty
   lists, fill \texttt{JobStore} with all jobs except those running on
-  possible rejected clusters.}
+  rejected clusters if there are any.}
 \end{enumerate}
 
-The steps above completes the initialization of the {\JobController}
+The steps above complete the initialization of the {\JobController}
 which is now ready for handling jobs. The public functions of the
 {\JobController} offer to get (download), clean, cancel, etc. one or
-more jobs and uses the private specializations for issuing the
-command. Here exemplified by the \texttt{Stat} command:
+more jobs and use the virutal methods implemented by the grid middleware 
+specialized {\JobController}s for issuing the command. Here exemplified 
+by the \texttt{Stat} command:
 
@@
 and thus a mapping is needed for non Glue2 compliant grid middlewares.
-Appendix~\ref{app:job} shows the present mapping schema.
-
-All attributes of the {\Job} can be printed by the method
+
+All attributes of the {\Job} can be printed by the method (longlist = true 
+will give more details (long listing))
 
@@
 joblist location. Upon initialization (constructor) the {\UserConfig}
-locates the user files
+locates and reads the user files
 
@@
 \item{Merge system and user configuration by adding all system
-  configuration not already listed in the user configuration to the
+  configuration not already listed in the user configuration items to the
   latter.}
@@
 \item{If the \texttt{\$X509\_USER\_PROXY} environment variable is set,
-  use its value to define a \texttt{ProxyPath} tag in \texttt{cfg}. If
+  use its value to define a \texttt{ProxyPath} tag in \texttt{ccfg}. If
   the file does not exist or can't be read exit with an error.}
@@
   \texttt{\$X509\_USER\_KEY} environment variable to define
-  \texttt{CertificatePath} and \texttt{KeyPath} tags in \texttt{cfg}.
+  \texttt{CertificatePath} and \texttt{KeyPath} tags in \texttt{ccfg}.
   If \texttt{\$X509\_USER\_KEY} is not set or either file does not
@@
 \item{Otherwise, if the merged user configuration tree (see
-  \texttt{ConfTree} contains a \texttt{ProxyPath} tag, copy it to
-  \texttt{cfg}. If the file does not exist or can't be read exit with
+  \texttt{ConfTree}) contains a \texttt{ProxyPath} tag, copy it to
+  \texttt{ccfg}. If the file does not exist or can't be read exit with
   an error.}
 \item{Otherwise, if the merged user configuration tree contains a
-  \texttt{CertificatePath} tag, copy it to \texttt{cfg} along with the
+  \texttt{CertificatePath} tag, copy it to \texttt{ccfg} along with the
   accompanying \texttt{KeyPath} tag. If the \texttt{KeyPath} tag is
@@
 \item{Otherwise, if the file \texttt{/tmp/x509up\_u + userid} exists
-  add a \texttt{ProxyPath} tag in \texttt{cfg} pointing to it.}
+  add a \texttt{ProxyPath} tag in \texttt{ccfg} pointing to it.}
 \item{Otherwise, add \texttt{CertificatePath} and \texttt{KeyPath}
-  tags to \texttt{cfg} pointing to
+  tags to \texttt{ccfg} pointing to
   \texttt{\$HOME/.globus/usercert.pem} and
@@
   use its value to define a \texttt{CACertificatesDir} tag in
-  \texttt{cfg}. If the directory does not exist, exit with an error.}
+  \texttt{ccfg}. If the directory does not exist, exit with an error.}
 \item{Otherwise, if the user ID is not zero and the directory
   \texttt{\$HOME/.globus/certificates} exists add a
-  \texttt{CACertificatesDir} tag in \texttt{cfg} pointing to it.}
-\item{Otherwise, add a \texttt{CACertificatesDir} tag in \texttt{cfg}
+  \texttt{CACertificatesDir} tag in \texttt{ccfg} pointing to it.}
+\item{Otherwise, add a \texttt{CACertificatesDir} tag in \texttt{ccfg}
   pointing to \texttt{/etc/grid-security/certificates}. If the
@@
 The ARC 0.x local information system uses the {\GT}~\cite{globus} GRIS
-with a purpose made ARC schema. The information index server used is the
-{\GT} GIIS. Both these servers are using the LDAP~\cite{ldap}
+with a custom made ARC schema. The information index server used is the
+{\GT} GIIS. Both these servers use the LDAP~\cite{ldap}
 protocol. The specialization of the {\TargetRetriever} class for ARC0
@@
 A-Rex implements the BES~\cite{ogsa-bes} standard interface. Since
-this is a SOAP~\cite{soap} based interface the specializations of the
+this is a SOAP-based~\cite{soap} interface, the specializations of the
 {\TargetRetriever}, {\Submitter} and {\JobController} classes make use
@@
 
-The A-Rex service uses the https protocol put and get methods for
+The A-Rex service uses the https protocol \texttt{put} and \texttt{get} methods for
 stage-in and stage-out of input and output files. Therefore, the ARC1
@@
 \begin{itemize}
-\item CREAM2 has a Web Service interface that fits the Web Service
+\item CREAM2 has a Web Service interface that is very similar to the Web Service
   based ARC.
@@
 
-CREAM has its own SOAP based interface. The CREAM
+CREAM has its own SOAP-based interface. The CREAM
 specializations of the {\Submitter} and {\JobController} classes
@@
 \item{The {\ExecutionTarget}s are ranked according to their queue length relative to the number of slots/CPUs. Zero queue 
-{\ExecutionTarget}s is ranked top.}
-\item{If more than one {\ExecutionTarget} has zero queue the targets are randomly reshuffled based upon how many free slots 
-they have. E.g. if two {\ExecutionTarget}s have zero queue and they have 90 and 10 free slots respectively, then the first
+{\ExecutionTarget}s is ranked first.}
+\item{If more than one {\ExecutionTarget} has zero queue, the targets are randomly reshuffled based upon how many free slots 
+they have, e.g. if two {\ExecutionTarget}s have zero queue and they have 90 and 10 free slots respectively, then the first
 {\ExecutionTarget} has a 90\% probability of becoming the top ranked cluster.}
@@
 interface {\texttt CacheCheck} for querying for the presence of the file in the cache directory. This interface has a limitation 
-in the current implementation as it can only use cache directory of format
-
-\begin{shaded}
-\begin{verbatim}
-cachedir="/tmp/cache"
-\end{verbatim}
-\end{shaded}
-
-If the cachedir parameter involves {\%}U or other substitute element then the {\texttt CacheCheck} will not work.
+in the current implementation and does not support per user caches.
 
@@
 \begin{enumerate}
-\item{Only the A-REX service has {\texttt CacheCheck} method, thus {\ExecutionTarget}s not running A-REX are removed.
-therefore the other clusters will be ignored.}
-\item{Information about input files is collected from the JobInnerRepresentation.}
+\item{Only the A-REX service has {\texttt CacheCheck} method, thus {\ExecutionTarget}s not running A-REX are removed.}
+\item{Information about input files requested in the job description is collected from the \texttt{JobInnerRepresentation.}}
 \item{Each {\ExecutionTarget} is queried (through the {\texttt CacheCheck} method) for the existence of input files. 
-A single query is used for achieving all the necessary information and file sizes are summarized. If there is problems 
+A single query is used for achieving all the necessary information and file sizes are summarized. If there are problems 
 determining file sizes, then the summarized size will be zero.}
-\item{The possible {\ExecutionTarget}s are ranked in a descending order according to the amount of input data they have in cache.}
+\item{The possible {\ExecutionTarget}s are ranked in a descending order according to the amount of input data they have in their cache.}
 \end{enumerate}
@@
   queries.}
-\item{\texttt{arcsub} - for grid job submission}
+\item{\texttt{arcsub} - for grid job submission.}
 \item{\texttt{arcget} - for downloading output of finished, cancelled
@@
 \item{\texttt{arcclean} - for removing a grid job's session directory
-  including all contents}
+  including all contents.}
 \item{\texttt{arccat} - for performing the \texttt{cat} command to a

### 2009-03-13 doc/client/client_technical.tex
fixed small bug

@@
          <ExistInTheCache>true</ExistInTheCache>
-         <FileSize>30</FileSize>
+         <FileSize>190</FileSize>
    </Result>
    <Result>
-         <FileURL>http://knowarc1.grid.niif.hu/storage/Makefile</FileURL>
+         <FileURL>ftp://download.nordugrid.org/test/README</FileURL>
          <ExistInTheCache>true</ExistInTheCache>
-         <FileSize>30</FileSize>
+         <FileSize>176</FileSize>
    </Result>

### 2009-03-16 doc/storage_system/arc-storage-documentation.tex
itemize has to end!
Include the gateway related interfaces in the Bartender

@@
 for, e.g., bootstrapping the A-Hash replicas and to communicate newly discovered clients.
+\end{itemize}
 \end{description}
@@
     The response is a list of (\emph{changeID}, \emph{success}), where \emph{success} could be  `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}' (for an `add' request), `\textbf{denied}', `\textbf{no such LN}', `\textbf{failed}: reason'.
+
+    \item[makeMountpoint(makeMountpointRequestList)] used to create the mountpoint
+
+    \emph{makeMountpointRequestList}, list of (\emph{requestID}, \emph{status})
+
+    \emph{makeMountpoint} is quite similar to the \emph{makeCollection}.
+It is used to provide the access of the external storage system within the global namespace.
+\emph{makeMountpointRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{URL}, \emph{metadata} ) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name).
+\emph{LN} will be requested Logical name and the \emph{URL} is the path to the external point that is required to access.
+In the policies section there could be some access policies, and in the metadata section there could be any other \emph{metadata} in key-value pairs.
+
+    The \emph{status} in the response could be \emph{done}, \emph{mountpoint exists}, \emph{parent does not exist}, \emph{internal error}, \emph{'failed: credentials not found'}, \emph{'cannot create anything in mountpoint'}
+
+    \item[unmakeMountPoint(unmakeMountpointRequestList)] delete the mountpoint
+
+    \emph{unmakeMountpointRequestList}, list of (\emph{requestID}, \emph{status})
+
+    \emph{unmakeMountpointRequestList} is a list of (requestID, LN) with the Logical Names of the mountpoints.
+\emph{status} could be \emph{removed}, \emph{no such LN}, \emph{Failed: Credential not available}.
+
+    \item[CredentialDelegation(CredentialDelegationRequest)] returns (\emph{status}, \emph{ID})
+
+    This interface is use to pass the user's limited credentials to the system in order to access the \emph{third-party} store for methods like \emph{get}, \emph{put} and \emph{list}.
+For listing files and directories, \emph{gateway} uses the user credential, whereas in the \emph{get} and \emph{put} methods system only uses the credentials to check the validity of the URLs before sending it to the user.
+
+    From the client side \emph{createDelegation()} is called which interns initiate the \emph{CredentialDelegationInitRequest} for the service.
+After the successful handshake this method initiates an other request of \emph{UpdateCredentials( )}, which completes this process.
+Response from this interface could be \emph{done}, \emph{cannot delegate credentials, requested bartender does not supports the gateway}
+
+    \item[RemoveCredentials(RemoveCredentialRequest)] returns (\emph{message}, \emph{status})
+
+    \emph{RemoveCredentialRequest} initiates the process of deleting the user credential from the service side.
+The response could be \emph{Credential removed successfully}, \emph{cannot access proxyFile or ProxyStore problem at service side}
 

### 2009-03-16 doc/client/client_technical.tex
Adding documentation for arcmigrate

@@
 \hypersetup{
-  pdfauthor = {KnorARC collaboration},
+  pdfauthor = {KnowARC collaboration},
   pdftitle = {libarcclient},
@@
 \begin{verbatim}
-void PreFilterTargets(Arc::TargetGenerator& targen, Arc::JobDescription jd);
+void PreFilterTargets(const TargetGenerator& targen, const JobDescription& jd);
 \end{verbatim}
@@
 bool setSource(const std::string source);
+bool setSource(const XMLNode& source);
 \end{verbatim}
@@
 \begin{verbatim}
-bool getProduct(std::string& product, std::string format = "POSIXJSDL")
+bool getProduct(std::string& product, std::string format = "POSIXJSDL") const;
 \end{verbatim}
@@
 \begin{verbatim}
-virtual bool Submit(JobDescription& jobdesc, XMLNode& info) = 0;
+virtual bool Submit(const JobDescription& jobdesc, XMLNode& info) const = 0;
 \end{verbatim}
@@
 by the target.
+
+\begin{shaded}
+\begin{verbatim}
+virtual bool Migrate(const URL& url,
+                     const JobDescription& jobdesc,
+                     bool forcemigration,
+                     XMLNode& info) const = 0;
+\end{verbatim}
+\end{shaded}
+
+The \texttt{Migrate} method behave in much the same way as the \texttt{Submit}
+method, but is conceptually different from that method. The \texttt{Migrate}
+method takes two additional arguments. A {\URL} \texttt{url} which is
+the jobid of a job which should be migrated, and a boolean
+\texttt{forcemigration} which will indicate if the job, specified in the
+{\JobDescription} \texttt{jobdesc}, should persist if the cluster
+({\ExecutionTarget}) does not succeed submitting a kill/terminate request
+for the job specified by \texttt{url}. The \texttt{Migrate} method will
+also fill the {\XMLNode} \texttt{info} variable with information about the
+job for later job management, which additionally include the old jobid
+(\texttt{url}) of the migrated job.
 
@@
 \begin{verbatim}
-const std::list<Arc::JobController*>& GetJobControllers();
+const std::list<JobController*>& GetJobControllers();
 \end{verbatim}
@@
 
-The other public functions of the {\JobController} class are:
+The other public methods of the {\JobController} class are:
 The \texttt{Get} which is used to download jobs
@@
 {\Job}.\texttt{JobDescription} attribute.
+
+The \texttt{Migrate} method:
+\begin{shaded}
+\begin{verbatim}
+bool Migrate(TargetGenerator& targetGen,
+             Broker* broker,
+             const bool forcemigration,
+             const int timeout);
+\end{verbatim}
+\end{shaded}
+will try to migrate the jobs contained in the object ({\JobController}). The jobs
+will be migrated to the targets specified in {\TargetGenerator} \texttt{targetGen}
+by using the broker specified by {\Broker} \texttt{broker}. The boolean
+\texttt{forcemigration} specifies whether a migrated job should persist if the
+new cluster ({\ExecutionTarget}) chosen by the {\TargetGenerator} does not
+succeed submitting a kill/terminate request for the job. The integer
+\texttt{timeout} argument have not yet been implemented.
 

### 2009-03-16 doc/client/client_technical.tex
Updated for the internal release of D2.2-1

@@
 
-The \texttt{Migrate} method behave in much the same way as the \texttt{Submit}
-method, but is conceptually different from that method. The \texttt{Migrate}
-method takes two additional arguments. A {\URL} \texttt{url} which is
-the jobid of a job which should be migrated, and a boolean
-\texttt{forcemigration} which will indicate if the job, specified in the
-{\JobDescription} \texttt{jobdesc}, should persist if the cluster
-({\ExecutionTarget}) does not succeed submitting a kill/terminate request
-for the job specified by \texttt{url}. The \texttt{Migrate} method will
-also fill the {\XMLNode} \texttt{info} variable with information about the
-job for later job management, which additionally include the old jobid
-(\texttt{url}) of the migrated job.
+The \texttt{Migrate} method behaves in much the same way as the \texttt{Submit}
+method, but instead of sending a new job request to the cluster it sends a 
+migration request of a given job, currently this request is only supported 
+for ARC1 clusters. The \texttt{url} and \texttt{jobdesc} specifies the job-id 
+and {\JobDescription} of the job which should be migrated.\\
+The migration request will create a new job where the local input files are 
+transferred from old cluster. Once the input files have been transferred the 
+new cluster will send a request to kill the old job. If the kill request is 
+unsuccessful the new job will be terminated unless the \texttt{forcemigration} 
+boolean is \texttt{true}.\\
+The \texttt{Migrate} method will also fill the \texttt{info} {\XMLNode} with 
+information about the job the same way as the \texttt{Submit} method. However 
+for the \texttt{Migrate} method this \texttt{XMLNode} also contain the job-id 
+of the migrated job.
 
@@
 new cluster ({\ExecutionTarget}) chosen by the {\TargetGenerator} does not
-succeed submitting a kill/terminate request for the job. The integer
-\texttt{timeout} argument have not yet been implemented.
+succeed sending a kill/terminate request for the job. The integer
+\texttt{timeout} argument has not yet been implemented.
 

### 2009-03-17 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Corrections

@@
 For each operation a corresponding SOAP action has to be defined which assignes a messages to be either \textit{literal} or \textit{encoded}~\cite{BUTEK_2009}.
-The service implemented in this chapter will use the binding \textit{document}/\textit{literal} and the protocol HTTP which is to be seen in the lines followed by  line~\ref{lst_code:echo_wsdl_binding}.
+The service implemented in this chapter will use the binding \textit{document}/\textit{literal} and the protocol HTTP which is to be seen in the lines follow line~\ref{lst_code:echo_wsdl_binding}.
 
@@
 
-A SOAP fault message is always located within the body element and can appear only one time. 
+A SOAP fault message is always located within the body element and can appear only once. 
 It indicates error messages in case the port type \textit{solicit-response} (input, output, fault) has been chosen.
-An element called \textit{fault} encapsulates diffrent types of elements which are describing the fault.
+An element called \textit{fault} encapsulates different types of elements that describe the fault.
 These elements are listed in Table~\ref{tbl:fault_elements}.
@@
 
-For morde information about WSDL it is recommended to check the internet address~\href{http://www.w3.org/TR/wsdl}{http://www.w3.org/TR/wsdl}.
+For more information additional tutorials can be found in the appendix or check the internet address~\href{http://www.w3.org/TR/wsdl}{http://www.w3.org/TR/wsdl}.
 
@@
 The source code is shown in Listing~\ref{lst:echo_service_cpp}.
-The header file will be set aside for it contains only redundant information. 
+The header file will be set aside since it contains only redundant information. 
 It can be reread in the source directory that is accompanying the tutorial.
-Regarding to the source code of the Echo Web Service basically three changes of the Time Web Service source code are required.
+In comparsion with Time Web Service, basically three changes are required.
 The first change concerns the constructor and can be found at line~\ref{lst_code:echo_cpp_prefix}.
@@
 Within the method \textit{process} two new code fragments are now extracting and analyzing the incoming message, to be seen at line~\ref{lst_code:echo_cpp_extracting} and \ref{lst_code:echo_cpp_analyzing}.
-In case something unexpected happens both will create a SOAP fault message.\\
+In case something unexpected happens either will create a SOAP fault message.\\
 % Vielleicht noch etwas über das extrahieren der Nachricht schreiben
@@
 The server configuration file is shown in Listing~\ref{lst:echo_hed_configuration_xml}.
-The only change concerns the lines~\ref{lst_code:echo_hed_configuration_xml_pre} and \ref{lst_code:echo_hed_configuration_xml_pre} which specify the \textit{prefix} and \textit{suffix}.
+The only change concerns the lines~\ref{lst_code:echo_hed_configuration_xml_pre} and \ref{lst_code:echo_hed_configuration_xml_suf} which specify the \textit{prefix} and \textit{suffix}.
 The parameters will be extracted in the constructor of the service and explain nicely how settings can be passed from the server configuration file towards the service.
@@
 The source code of the client is shown Listing~\ref{lst:client_echo_cpp}.
-The parameters are now passed by the command line, to be seen in the lines followed by line~\ref{lst_code:echo_client_cpp_arguments}.
+The parameters are now passed by the command line, to be seen in the lines following line~\ref{lst_code:echo_client_cpp_arguments}.
 That code is independent from the ARC library.
-The creation of the SOAP client is unchanged and the request is created in the lines subsequent to line~\ref{lst_code:echo_client_cpp_request}.
-It is processed in the code after line~\ref{lst_code:echo_client_cpp_response} in which is also the response is received.
+The creation of the SOAP client is unchanged from the Echo client's implementation and the request to the server is created in the lines subsequent to line~\ref{lst_code:echo_client_cpp_request}.
+It is processed in the code after line~\ref{lst_code:echo_client_cpp_response} in which also the response is received.
 The checks of the response have expanded by an additional condition which tests for a fault message, see line~\ref{lst_code:echo_client_cpp_fault}.

### 2009-03-17 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Corrections

@@
 
-This document gently introduces to the preparation of standalone Web Services and cognate clients with the Advance Resource Connector (ARC).
+This document gently introduces to the preparation of standalone Web Services and cognated clients with the Advance Resource Connector (ARC).
 The reader is guided through a series of practically oriented examples.
@@
 It is however beneficial to have some basic comprehension of the programming language C++, the file format XML and any regular UNIX shell.
-To start, the user shall have an installation of ARC-1, as it is performed with current RPM or .deb packages\footnote{see http://wiki.nordugrid.org for explicit instructions for setup}.
+To start, the user shall have an installation of ARC-1, as it is performed with current RPM or .deb packages\footnote{see http://wiki.nordugrid.org for explicit instructions for setup}.\\
 
@@
 Amazon's clouds) or provide access to other resources like a particular
-database or to arbitrary disk space (like ARC's hopi or bartender).
+database or to arbitrary disk space (like ARC's hopi or bartender).\\
 
-The users of computational services are software tools, too.
-They are referred to as clients.
-It is those clients that perform the interactions with a human user.
-However, when a service requests a performance of another service at the same or another server, then this first service is in the role of a client, too.
-
-Servers provide access to a single service or a set of services.
-For ARC, there is one service, named Hosting Environment Demon (HED), that is more special than other services: HED organises the assignment of other services on the same server to the network.
-But this does not break the concept, it is only a commodity.
-If services are not busy reacting to a client's request, then they are waiting.
-The task of servers can be subdivided into:
+Services are invoked using software programs, which are referred to as clients. 
+These programs may directly perform the interactions of human users or request a performance of other services of another server (here the first service is in the role of a client, too).
+The tasks of servers and clients are well defined. If servers are not busy reacting to a client's request, then they are waiting. The task of servers can be subdivided into:
 \begin{itemize}
@@
  \end{itemize}
+Servers provide access to a single or a set of services.
+In ARC services are provided by the Hosting Environment Demon (HED) which enables the installation of several services on one network access point.
 \forcelinebreak
@@
 Additional modules (on the left) are used within the HED.
-\task{Why is the Web Service in all images outside of the HED?}}
+\task{Why is the Web Service in all images outside of the HED?} } % TODO: Reverse order of layers, s.t.
+the highest layers are above and the lowest layer are at the bottom
 \label{fig:HED_internal}
@@
 
-As to be seen, the MCCs are connected which each other and passing incoming and outgoing message to the upper or the respective lower layer.
+The MCCs are connected which each other and passing incoming and outgoing message to the upper or the respective lower layer.
 In case of outgoing messages a MCC is wrapping the data of the upper layer as a payload into their own protocol (while incomming message will be unwrapped).

### 2009-03-17 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
Corrections

@@
 For the administration of a large VO the kind of security may be too general.
-Thus, it is recommended to have a more fine grained security. 
-This chapter will discuss how to install a more complex security such that it is possible to limitate the access of certain resources to particular users and to narrow their rights on the workstation.\\
+Thus, it is recommended to have a more fine-grained security. 
+This chapter discusses how to install a more complex security such that it is possible to limit the access of certain resources to particular users and to constrain their rights on the workstation.\\
 
-The security framework of ARC is implemented based on the same concept of modularity as HED. 
-To extend the security of a MCC or a service one only has to modify the server configuration file.
-That is done by including a \textit{SecHandler} element into the component or service which again has to contain a PDP element.
+The security framework of ARC is based on the same concept of modularity as HED. 
+To extend the security of a MCC or a service one only needs to modify the server configuration file.
+That is done by including a \textit{SecHandler} element into the component or service which contains a PDP element.
 A small example of a chain component containing a \textit{SecHandler} is shown in Listing~\ref{lst:sec_service_component_xml}.
@@
 The  attribute \textit{name} is \textit{arc.authz}, the \textit{event} attribute is \textit{incoming} and the \textit{id} attribute which is \textit{auth}.
-While \textit{id} is only an optional parameter, the attribute \textit{name} causes a certain \textit{SecHandler} to be loaded.%, which again has a certain functionality. 
+While \textit{id} is only an optional parameter, the attribute \textit{name} causes a certain \textit{SecHandler} to be loaded. %, which again has a certain functionality. 
 In case the attribute \textit{name} is assigned to \textit{arc.authz}, the SecHandler provides the functionality of authentication and is able to permit or to deny the access to the next layer. 
@@
 \textit{SecHandlers} are able to intervene the message flow by: manipulating the message content, manipulating the message attributes or to disrupt messages.
-Within one component, \textit{SecHandlers} are executed sequentially in the order in which they were attached to the queue.
-If any of them fails, message processing fails as well~\cite{QIANG_2008}.
-In order to decide how the \textit{SecHandler} shall modify the message so called PDP (Policy Decision Point) are used.
+Within one component, \textit{SecHandlers} are executed sequentially in the order in which they were attached to the queue. If any of them fails, message processing fails as well~\cite{QIANG_2008}.\\
+
+In order to decide how the \textit{SecHandler} shall modify the message, so called PDP (Policy Decision Point) are used.
 Within the example a PDP named \textit{simplelist.pdp} is inserted within the \textit{SecHandler} element.
@@
   \centering
-  \caption{List of buildin SecHandlers}
+  \caption{List of built-in SecHandlers}
   \label{tbl:list_of_sechandler}
@@
 %					realised by a message attribut.\\
- 	\textbf{arc.authz}            & Grants the access to the next layer.\\
+ 	\textbf{arc.authz}            & Grants access to the next layer.\\
  	\textbf{delegation.collector} & Processes a request for another service located on a different HED.
@@
                                         to the requested service.
-                                        The policy is now available within the HED of the service such that the client is now able to access the service.\\
+                                        The policy is now available within the HED of the service such that the client is enabled to access the service.\\
                                         % Question how does the client creates the request - with the SecHandler?
@@
   \centering
-  \caption{List of buildin PDPs}
+  \caption{List of built-in PDPs}
 \label{tbl:list_of_pdps}
@@
 Listing~\ref{lst:sececho_service_cpp} contains the source code of the Secure Echo Web Service.
-As one can see, the only diffrence, compared to the source code of the previous section, is the additional code to process the \textit{SecHandler} starting at line~\ref{lst_code:sececho_service_cpp_process_secH}.
-Obviously only the incoming queue is processed and in case the SecHandler fails, the SOAP service returns a fault message.
+As one can see, the only difference, compared to the source code of the previous section, is the additional code to process the \textit{SecHandler} starting at line~\ref{lst_code:sececho_service_cpp_process_secH}.
+Obviously only the incoming queue is processed and in case the \textit{SecHandler} fails, the SOAP service returns a fault message.
 The file, which is revealing a lot more information, is the server configuration file which is shown in Listing~\ref{lst:sececho_server_configuration_file_xml}.
-In order to be able to have \textit{SecHandlers} working with identies, it is almost essential to utilise the TLS layer.
+To enable \textit{SecHandlers} to work with identities, it is almost essential to utilise the TLS layer.
 % Fixed IP are also possible but not recommended..
@@
 %
-The first \textit{SecHandler} is declared within line~\ref{lst_code:arched_sec_conffile_secH2}. 
-Due to the PDP \textit{allow.pdp} returns always a positive result, everyone who passes the \textit{SecHandler} is mapped to the user \textit{nobody}. 
+The first \textit{SecHandler} is declared at line~\ref{lst_code:arched_sec_conffile_secH2}. 
+The PDP \textit{allow.pdp} always returns a positive result, such that everyone who passes the \textit{SecHandler} is mapped to the user \textit{nobody}. 
 %
@@
 The second \textit{SecHandler} limits the access to the SOAP services using the \textit{arc.authz} \textit{SecHandler} and the \textit{simplelist} policy, see line~\ref{lst_code:arched_sec_conffile_secH1}.
-(A similar example was already shown in Listing~\ref{lst:sec_service_component_xml}).
+A similar example was already shown in Listing~\ref{lst:sec_service_component_xml}.
 The \textit{simplelist} policy returns a positive result if the identity of the certificate corresponds to an identity within the file \textit{clientlist.txt}.
@@
 Once all rules are evaluated, a decision for the policy has to be generated which will depend on the results of the rules.
-The policy element attribute \textit{CombiningAlg} determines how the final result will be generated. Two possibility are given:
+The policy element attribute \textit{CombiningAlg} determines how the final result will be generated. Two possibilities are given:
 \begin{itemize}
@@
 The \textit{Actions} are specifing the intended usage of the resource the rule shall survey.
-The last element \textit{Condition} defines somewhat like Time, Duration, namespace of a SOAP message etc. which is needed to be fulfilled.
-For the description of conditions is complicated, they will not be explained in this tutorial.\\
+The last element \textit{Condition} defines somewhat like Time, Duration, namespace of a SOAP message etc.
+which are needed to be fulfilled.
+The detailed description of these conditions is rather extensive and beyond the scope of this tutorial.
+Please refer to the further going tutorials which are listed in the appendix of this tutorial.\\
 
 Within the sample policy, which was shown in Listing~\ref{lst:sececho_policy_xml}, three different examples of rule states are prepared: PERMIT, INDETERMINATE and NOT APPLICABLE.
-The first rule, to be seen starting at line~\ref{lst_code:policy.xml_rule1}, is fitting to the server configuration file such that a user with a suitable identity will produce the state PERMIT. 
-Due to there is no rule within the policy which will have the DENY state, this rule let the client pass the PDP.
-The second rule, begins after line~\ref{lst_code:policy.xml_rule2}.
+The first rule, starting at line~\ref{lst_code:policy.xml_rule1}, is fitting to the server configuration file such that a user with a suitable identity will produce the state PERMIT. 
+Since there is no rule within the policy which will have the DENY state, in particular no rule that could have matched earlier, this rule lets the client pass the PDP.
+The second rule begins after line~\ref{lst_code:policy.xml_rule2}.
 It is arranged in a manner such that the state INDETERMINATE will be induced.
-Two things are checked within it: The identity based on Shibboleth authentication and the path of the requested HTTP URL.
+Two things are checked within it: the identity based on authentication method Shibboleth\footnote{Shibboleth is an open-source implementation for authentication and authorization based on the XML standard SAML (Security Assertion Markup Language).}
+and the path of the requested HTTP URL.
 Due to the fact, that the HTTP path is fitting to the requested one, but the authentication was not done by Shibboleth, the rule is not able to decide for a DENY or a PERMIT state --- the rule produces an INDETERMINATE state.
@@
 
-To get an impression in how the requests look like, one can check the log-file (the log level needs to set to VERBOSE).
+To get an impression in how the requests look like, one can check the log-file (the log level needs to be set to VERBOSE).
 
@@
 
-The security introduced to the server has no effect on the client, such that the source code is absolutly identical. Even the client configuration file doesn't need to be changed.
+The security introduced to the server has no effect on the client, such that the source code is completely identical. Even the client configuration file doesn't need to be changed.

### 2009-03-17 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Corrections

@@
 
-The Web Services of the previous chapters have one mayor drawback: the communication between the two endpoints is insecure.
-For that reason the current and the next chapter will discuss how to create a secure Web Service.
+The Web Services of the previous chapters have one major drawback: the communication between the two endpoints is insecure.
+For that reason the current and the next chapters will discuss how to create a secure Web Service.
 In general, security in a network can be subdivided into the following categories:~\cite{TANNENBAUM_2001}\task{TODO: verfiy this citation}
 \begin{itemize}
-	\item \textbf{Confidentiality} --- Protection of the data against passiv attacks (such as publicise message content)
+	\item \textbf{Confidentiality} --- Protection of the data against passive attacks (such as publicise message content)
 
-	\item \textbf{Authentication} --- Verification of the Authenticity  of the communications partner.
+	\item \textbf{Authentication} --- Verification of the identity of the communication partners.
 
@@
 
-	\item \textbf{Availability} --- Ensure a system is always usable which is challenged by several attacks.
+	\item \textbf{Availability} --- Ensure a system is always usable which is challenged by various attacks.
 
@@
 
-The next section gives a basic understanding of the working principale of the TLS.
+The next section gives a basic understanding of the working principle of the TLS.
 In order to enable newcomers an entry to  this area, emphasis is put on the security concept and not on an exact description of the technical realisation.
@@
 Asymmetric encryption follows a different functional principle.
-While the symmetric encryption utilise one kay, the asymmetric ecryption is based on two keys: a private key and public key.
+While the symmetric encryption utilise one key, the asymmetric ecryption is based on two keys: a private key and public key.
 As the names imply, the private key is kept secret and is never revealed to others while the public key is available for everyone.
@@
 As one can see, this first approach already provides confidentiality and integrity of messages.
-But neither Non-repudiation, authentication and consequently nor access control are established because the identity of Alice is not proven.
+But neither non-repudiation, authentication and consequently nor access control are established because the identity of Alice is not proven.
 In order to be able to check identities, so called certificates have been invented.
@@
  	\includegraphics[width=13cm]{tex_tls_echoservice/certificates.pdf}
-	\mycaption{specific}{general}
+	\mycaption{Procedure of certificate signing.}{}
 	\label{fig:certificate_request}
@@
 Alice successfully compares the returned number. Additionally she validates the certificate of CA 1 with the pre-configured certificate of CA 2..
-Since the certificate was signed by CA 2 which is pre-configured to be trusted, she als trusts CA 1 and in conclusion the identity of B.
+Since the certificate was signed by CA 2 which is pre-configured to be trusted, she also trusts CA 1 and in conclusion the identity of B.
 In summary, Alice knows she is talking to Bob.
@@
 
-When the client is aware of the server authentity, the confidential treatment of its data is given such that one can be sure the data won't be revealed to a third party (server-sided confidentiality).
+When the client is aware of the server's authentity, the confidential treatment of its data is guaranteed such that one can be sure the data won't be revealed to a third party (server-sided confidentiality).
 On the other side, if the server is aware of the client authentity, the access to services and resources can be controlled (access control).
 Thus, in case of services, the results will only be passed back to the right person (client-sided confidentiality).
-As one can see confidentiality, authentication, integrity and non-repudiation are now given.\\
+As one can see confidentiality, authentication, integrity and non-repudiation are now established.\\
 
@@
 	\includegraphics[width=13cm]{tex_tls_echoservice/verification.pdf}
-	\mycaption{Usecase in which Alice is resolving the identity of Bob.}{In order to resolve the identity of Bob, Alice sends a request for Bob's identity along with a random number to Bob.
+	\mycaption{Usecase in which Alice resolves the identity of Bob.}{In order to resolve the identity of Bob, Alice sends a request for Bob's identity along with a newly created random number to Bob.
 Bob encrypts the received number using his private key and returns his public key to Alice.
 Due to the random number can only be encrypted by the owner of the private key, Alice is able ot verify that the certificate was transmitted by his owner.
-Since the certificate was signed by the CA 1 which she is not per-configured to be trust, Alice needs to resolve also the identity of CA 1.
+Since the certificate was signed by the CA 1, which she is not per-configured to be trust, Alice needs to resolve also the identity of CA 1.
 The request is done in the same manner, but this time the certificate is signed by the trusted CA 2.
-Due to the fact she is now able to trust CA 1, Alice is know able to trust Bob.}
+Due to the fact she is now able to trust CA 1, Alice can now also trust the identity of Bob.}
 	\label{fig:verification_of_certificates}
@@
 They are scaling very good with the amount of users because CAs can create new sub-CAs such that the load is balanced (certificates of a CA may be cached).
-Furthermore no critical information is exchanged during the establishment of a connection (in particular: no secret key is shared which may get intercepted).
+Furthermore, no critical information is exchanged for the establishment of a connection (in particular: no secret key is shared which may get intercepted).
 Instead, a set of pre-configured CAs which will be in general provided by the operation system is needed.
 In respect of ARC it is obvious that the access to the grid shall be under the control of its administrator.
-To do so, the administrators have to create their own CAs which is in charge of creating certificates for new users.
+To do so, the administrators have to create their own CAs which are in charge of creating certificates for new users.
 The identity of the created CA has to be added to the directory of pre-configured CAs.
@@
 
-A CA which controls the access to a grid can be considered as a virtual organsiation (VO).
+A CA which controls access to a grid can be considered as a virtual organsiation (VO).
 It may involve organisations close to the leading organisation as well as structure which are not formally associated with it.
@@
 \textbf{clientCERT} & The certificate of the client. \\
-\textbf{clientKEY}  & The private key of the client (should be kept in safe custody).\\
-\\
+\textbf{clientKEY}  & The private key of the client (should be kept in safe custody).\vspace{2ex}\\
 \textbf{serverCA}   & The CA which guarantees for the identity of the server \\
@@
 The client on the other side should own the files: clientCERT.pem, clientKEY.pem and serverCA.pem.
-Within the source code directory certificates with corresponding names are prepared.
-Due to certificates have a limited period of validity, these certificates may no longer be accepted by the communication parter.\\
+Within the source code directory that accompanies this document certificates with corresponding names are prepared.
+Due to certificates have a limited period of validity, these certificates may no longer be accepted by the communication parter.
+One has to create new certificates to ensure the proper work of the TLS Web Service.\\
 
-Several tools are available which allow the certificate handling for more advanced UNIX users.
-\task{The ARC source code comes with two shell scripts to create a CA and related certificates. - ??MG}
+Several tools are available which allow the handling for more advanced UNIX users.
 In order to create a first basic set of certificates signed by a self created CA, the shell scripts \textit{createMasterCA.sh} and \textit{createSlaveCert.sh} can  be found within the directory \textit{src/services/tlsechoservice/certFactory} may be used.
@@
 
-The TLS Echo Web Service implementation is almost the same as the one presented in the chapter before.
-Merely the Class name and the namespace of the service have changed.
+The TLS Echo Web Service implementation is almost the same as the one presented in the previous chapter.
+Merely the Class name and the namespace of the service have been changed.
 The only modification is done within the server configuration file which is displayed in Listing~\ref{lst:tls_echo_arched_xml}.
-
 \lstsetARCHEDXML
@@
 The second change is the new element which has been placed between the TCP and the HTTP component, see line~\ref{lst_code:tls_echo_arched_xml_component}.
-Within the TLS element the certificates can be declared using the elements \textit{KeyPath}, \textit{CertificatePath},\textit{CACertificatePath} and \textit{CACertificatesDir}.
+Within the TLS element the certificates can be declared using the elements \textit{KeyPath}, \textit{CertificatePath}, \textit{CACertificatePath} and \textit{CACertificatesDir}.
 % \begin{verbatim}
@@
 In this section a new way to implement the client will be introduced.
-As already mentioned in section~\ref{sec:timeservice_client}, the protocol stack of the client is build up in almost the same way as the one of the service.
+As already mentioned in section~\ref{sec:timeservice_client}, the protocol stack of the client is built up in almost the same way as the one of the service.
 Indeed, the class~\textit{ClientSOAP} composes the stack with the same MCCs used for the service.
-Once this is known, its self-evident to create the client in a simlar manner like the service with an own configuration file.
+Once this is known, it is self-evident to create the client in a simlar manner like the service with an own configuration file.
 Listing~\ref{lst:tls_echo_client_cpp} shows the source code of the client program.
-The main difference between the simple Echo Client presented in the chapter before is that now the information about the client protocol stack is encapsulated within a configuration file.\\
+The main difference to the simple Echo Client presented in the previous chapter is that now the information about the client protocol stack is encapsulated within a configuration file.\\
 \lstsetCPP
@@
 
-The file containing the client configuration is specified within the commandline. 
+The file containing the client configuration is specified within the command line. 
 Its content will be loaded in the code fragment starting at line~\ref{lst_code:tls_echo_client_cpp_getfilecontent}. 
@@
 In order to process a message, an entry point within the chain has to be disclosed.
-For we are implementing a SOAP client, we are looking for the SOAP component.
+To implement a SOAP client, we are looking for the SOAP component.
 To get the proper component an operator of MCCLoader is used in line~\ref{lst_code:tls_echo_client_cpp_cliententry}.
@@
 The payloads now have to be wrapped by message objects which have to be initialised with an \textit{MessageAttribute} object and a \textit{MessageContext} object.
-After the payload has been wrapped within the message it gets transmitted in almost the same way like before, see lines followed by line~\ref{lst_code:tls_echo_client_cpp_createandprocess}.
-The rest of the source code is identical to the Echo Client.\\
+After the payload has been wrapped within the message it gets transmitted in almost the same way like before, see lines following by line~\ref{lst_code:tls_echo_client_cpp_createandprocess}.
+The remainder of the source code is identical to the Echo Client.\\
 

### 2009-03-17 doc/ws-programming-tutorial/src/services/secechoservice/secechoservice.h

@@ -5,8 +5,8 @@
  * was send to it.
  */
 
-#ifndef __ARCSERVICE_ECHO_H__
-#define __ARCSERVICE_ECHO_H__
+#ifndef __SECECHOSERVICE_H__
+#define __SECECHOSERVICE_H__
 
 #if HAVE_CONFIG_H
 	#include <config.h>

### 2009-03-18 doc/storage_system/arc-storage-manual.tex
starting an administrator's and user's manual

@@
 
+In this section we will follow the installation of the ARC storage system on a freshly installed debian linux machine.
+
+Get the latest ARC code from the subversion (\verb!http://svn.nordugrid.org/!):
+
 \begin{verbatim}
-http://svn.nordugrid.org/
-svn co http://svn.nordugrid.org/repos/nordugrid/arc1/trunk arc1   
-README file for dependencies
+$ svn co http://svn.nordugrid.org/repos/nordugrid/arc1/trunk arc1
+\end{verbatim}
+
+Let's check the README file for all the dependencies, and install them:
+
+\begin{verbatim}
+$ sudo aptitude install build-essential
+$ sudo aptitude install autoconf
+$ sudo aptitude install gettext
+$ sudo aptitude install cvs
+$ sudo aptitude install libtool
+$ sudo aptitude install pkg-config
+$ sudo aptitude install libglibmm-2.4-dev
+$ sudo aptitude install python-dev
+$ sudo aptitude install swig
+$ sudo aptitude install libxml2-dev
+$ sudo aptitude install libssl-dev
+\end{verbatim}
+
+Run the \verb!autogen! and \verb!configure! scripts (you may want to specify target directories for the installation with the \verb!--prefix! option):
+
+\begin{verbatim}
+$ cd arc1
 $ ./autogen.sh
@@
 
-RLS:                yes
-GridFTP:            yes
+RLS:                no
+GridFTP:            no
 LFC:                no
-RSL:                yes
-SAML:               yes
+RSL:                no
+SAML:               no
 MYSQL CLIENT LIB:   no
@@
 SRM client (DMC):   no
-GSI channel (MCC):  yes
-
-$ make
-
-$ sudo make install
+GSI channel (MCC):  no
 
 \end{verbatim}
+
+We can start to compile it:
+
+\begin{verbatim}
+$ make
+\end{verbatim}
+
+And then install it (you need to be root if you want to install it to the default location):
+
+\begin{verbatim}
+$ sudo make install
+\end{verbatim}
+
+Maybe we should run \verb!ldconfig! to refresh the list of libraries:
+
+\begin{verbatim}
+$ ldconfig
+\end{verbatim}
+
+We can check if the python packages are installed correctly:
+
+\begin{verbatim}
+$ python
+Python 2.5.2 (r252:60911, Jan  4 2009, 17:40:26) 
+[GCC 4.3.2] on linux2
+Type "help", "copyright", "credits" or "license" for more information.
+>>> import arc
+>>> import storage
+>>> import arcom
+>>> 
+\end{verbatim}
+
+If there is no error on importing any of the packages, we are good to go.
 

### 2009-03-19 doc/storage_system/arc-storage-manual.tex
starting a quick start manual to install and configure the storage system

@@
 \tableofcontents                          %Comment if use article style
-\begin{verbatim}
-    Administrator's and user's manual
-    ---------------------------------
-
-    Administrator's manual
-    - Quick start guide
-        - Checking the code out from the subversion
-        - Compiling and installing
-        - Modifying the configuration template
-        - Running the daemon
-        - Using the CLI tool
-    - Installation:
-        - Getting the source code from subversion
-        - Compilation details
-        - Installation
-    - Configuration:
-        - HED service config
-        - Centralized A-Hash
-        - Distributed A-Hash
-        - Librarian
-        - Shepherd
-        - Hopi
-        - Bartender
-        - Configuring inter-service trust: local listing of DNs, or using an A-Hash to store them
-        - Virtual organizations: how to configure HED to trust a VO
-    - Deployment scenarios:
-        - Single node: putting every service on one single node (including a storage node with Hopi)
-        - Centralized management: using seperate multiple storage nodes with Hopi, but leaving a central A-Hash, a Librarian and a Bartender on the central node
-        - Simple distribution: split the central node to three nodes, each of them has an A-Hash, a Librarian and a Bartender, and we have separate storage nodes
-        - One service per node: we have seperate nodes for all the services, any number of A-Hash nodes, any number of Librarian nodes, Bartender nodes, storage nodes
-
-    User's manual
-    - Overview: what do you need to access a running ARC storage system
-    - Prototype command line tool
-        - Configuring the command line tool
-        - Creating the root collection
-        - Uploading files
-        - Creating collections
-        - Moving files
-        - Creating hardlinks
-        - Removing files
-        - Removing collections
-        - Close collections
-        - Modify access policies
-        - Working with mount points
-    - Grid jobs
-        - Configuring the ARC storage DMC
-        - Specifying files in job descriptions
-    - Accessing the system with FUSE
-        - Configuring the FUSE module
-        - Mounting the storage system
-        - Supported operations
-        - Limitations
-
-\end{verbatim}
+% \begin{verbatim}
+%     Administrator's and user's manual
+%     ---------------------------------
+% 
+%     Administrator's manual
+%     - Quick start guide
+%         - Checking the code out from the subversion
+%         - Compiling and installing
+%         - Modifying the configuration template
+%         - Running the daemon
+%         - Using the CLI tool
+%     - Installation:
+%         - Getting the source code from subversion
+%         - Compilation details
+%         - Installation
+%     - Configuration:
+%         - HED service config
+%         - Centralized A-Hash
+%         - Distributed A-Hash
+%         - Librarian
+%         - Shepherd
+%         - Hopi
+%         - Bartender
+%         - Configuring inter-service trust: local listing of DNs, or using an A-Hash to store them
+%         - Virtual organizations: how to configure HED to trust a VO
+%     - Deployment scenarios:
+%         - Single node: putting every service on one single node (including a storage node with Hopi)
+%         - Centralized management: using seperate multiple storage nodes with Hopi, but leaving a central A-Hash, a Librarian and a Bartender on the central node
+%         - Simple distribution: split the central node to three nodes, each of them has an A-Hash, a Librarian and a Bartender, and we have separate storage nodes
+%         - One service per node: we have seperate nodes for all the services, any number of A-Hash nodes, any number of Librarian nodes, Bartender nodes, storage nodes
+% 
+%     User's manual
+%     - Overview: what do you need to access a running ARC storage system
+%     - Prototype command line tool
+%         - Configuring the command line tool
+%         - Creating the root collection
+%         - Uploading files
+%         - Creating collections
+%         - Moving files
+%         - Creating hardlinks
+%         - Removing files
+%         - Removing collections
+%         - Close collections
+%         - Modify access policies
+%         - Working with mount points
+%     - Grid jobs
+%         - Configuring the ARC storage DMC
+%         - Specifying files in job descriptions
+%     - Accessing the system with FUSE
+%         - Configuring the FUSE module
+%         - Mounting the storage system
+%         - Supported operations
+%         - Limitations
+%     
+% \end{verbatim}
 
@@
 
-In this section we will follow the installation of the ARC storage system on a freshly installed debian linux machine.
+In this section we will follow the installation of the ARC storage system on a freshly installed debian lenny machine.
 
@@
 $ ./autogen.sh
-$ ./configure --disable-java --disable-a-rex-service --disable-isi-service --disable-charon-service --disable-compiler-service --disable-paul-service --disable-sched-service
---prefix
+$ ./configure --disable-java --disable-a-rex-service --disable-isi-service \
+    --disable-charon-service --disable-compiler-service \
+    --disable-paul-service --disable-sched-service
 
@@
 
+It is important to syncronize the time of the machine - the storage system will make decisions based on differences between timestamps, when we deploy it on multiple machines, different times would cause problems.
+
+We need host (and user) certificates to run the system.
+For testing purposes we can use the NorduGrid InstantCA solution which is capable of creating a demo-CA with short lifetime.
+The URL of the InstantCA service is \verb!https://vls.grid.upjs.sk/CA/instantCA!, let's create 3 user and 3 host certificates.
+(e.g.
+we can set the Organization Name to `knowarc', the Common name of the CA to `storage-test', the name of the users could be `penny', `billy' and `hammer', the name of the hosts could be `upsilon', `omicron' and `theta', and we need a password for the CA and passwords for the user certificates, both should be at least 4 characters long).
+
+Let's download the generated certificates and untar the archive file to see its contents:
+
+\begin{verbatim}
+$ ls -lR
+.:
+total 28
+drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 CA
+-rw-r--r-- 1 zsombor zsombor  963 2009-03-19 12:11 ca.key
+-rw-r--r-- 1 zsombor zsombor  786 2009-03-19 12:11 ca.pem
+drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 hostCerts
+-rw-r--r-- 1 zsombor zsombor 1044 2009-03-19 12:11 readme
+-rw-r--r-- 1 zsombor zsombor    3 2009-03-19 12:11 serial.srl
+drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 userCerts
+
+./CA:
+total 8
+-rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 66fe707c.0
+-rw-r--r-- 1 zsombor zsombor 139 2009-03-19 12:11 66fe707c.signing_policy
+
+./hostCerts:
+total 24
+-rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 hostcert-omicron.pem
+-rw-r--r-- 1 zsombor zsombor 782 2009-03-19 12:11 hostcert-theta.pem
+-rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 hostcert-upsilon.pem
+-rw-r--r-- 1 zsombor zsombor 887 2009-03-19 12:11 hostkey-omicron.pem
+-rw-r--r-- 1 zsombor zsombor 891 2009-03-19 12:11 hostkey-theta.pem
+-rw-r--r-- 1 zsombor zsombor 887 2009-03-19 12:11 hostkey-upsilon.pem
+
+./userCerts:
+total 24
+-rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-billy.pem
+-rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-hammer.pem
+-rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-penny.pem
+-rw-r--r-- 1 zsombor zsombor 963 2009-03-19 12:11 userkey-billy.pem
+-rw-r--r-- 1 zsombor zsombor 963 2009-03-19 12:11 userkey-hammer.pem
+-rw-r--r-- 1 zsombor zsombor 951 2009-03-19 12:11 userkey-penny.pem
+\end{verbatim}
+
+We have certificate and key files for all the hosts and users, and we have the CA file with the proper hashed name.
+There is no rule where to put these certificate, but now we will put them in \verb!/etc/grid-security!.
+Let's create this directory, and a subdirectory called \verb!certificates!.
+We will use the `theta' host certificates for this machine, let's copy it there:
+
+\begin{verbatim}
+$ sudo cp hostCerts/hostcert-theta.pem /etc/grid-security/hostcert.pem
+$ sudo cp hostCerts/hostkey-theta.pem /etc/grid-security/hostkey.pem
+$ sudo cp CA/* /etc/grid-security/certificates/
+$ ls -lR /etc/grid-security/
+/etc/grid-security/:
+total 12
+drwxr-xr-x 2 root root 4096 2009-03-19 12:36 certificates
+-rw-r--r-- 1 root root  782 2009-03-19 12:35 hostcert.pem
+-rw-r--r-- 1 root root  891 2009-03-19 12:36 hostkey.pem
+
+/etc/grid-security/certificates:
+total 8
+-rw-r--r-- 1 root root 786 2009-03-19 12:36 66fe707c.0
+-rw-r--r-- 1 root root 139 2009-03-19 12:36 66fe707c.signing_policy
+\end{verbatim}
+
+And I choose a user certificate and put it into the \verb!~/.arc! directory, conveniently removing the password from the key file:
+
+\begin{verbatim}
+$ mkdir ~/.arc
+$ cp userCerts/usercert-billy.pem ~/.arc/usercert.pem
+$ openssl rsa -in userCerts/userkey-billy.pem -out ~/.arc/userkey.pem
+$ chmod 600 ~/.arc/userkey.pem
+$ ls -l ~/.arc
+total 8
+-rw-r--r-- 1 zsombor zsombor 778 2009-03-19 13:06 usercert.pem
+-rw------- 1 zsombor zsombor 891 2009-03-19 13:09 userkey.pem
+\end{verbatim}
+
+The storage system runs withen the \verb!arched!
+hosting environment daemon, which needs a configuration file which describes which services we want to run, and on which ports do we want to listen, etc.
+Let's copy the template configuration files to the \verb!/etc/arc! directory:
+
+\begin{verbatim}
+$ sudo mkdir /etc/arc
+$ sudo cp arc1/src/services/storage/storage_service.xml.example /etc/arc/storage_service.xml
+$ sudo cp arc1/src/services/storage/clientsslconfig.xml /etc/arc
+$ ls -l /etc/arc
+total 12
+-rw-r--r-- 1 root root  259 2009-03-19 13:26 clientsslconfig.xml
+-rw-r--r-- 1 root root 4606 2009-03-19 13:25 storage_service.xml
+\end{verbatim}
+
+The template configuration file specifies several directories where the services can store their data. We will create these directories at \verb!/var/spool/arc!, but they can be created anywhere else.
+
+\begin{verbatim}
+$ sudo mkdir /var/spool/arc/ahash_data
+$ sudo mkdir /var/spool/arc/shepherd_data
+$ sudo mkdir /var/spool/arc/shepherd_store
+$ sudo mkdir /var/spool/arc/shepherd_transfer
+$ ls -l /var/spool/arc
+total 16
+drwxr-xr-x 2 root root 4096 2009-03-19 13:42 ahash_data
+drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_data
+drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_store
+drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_transfer
+\end{verbatim}
+
+In this deployment we plan to run the daemon as root, so it is OK that the root owns these directories. 
+If we create these directories somewhere else, we would need to modify the paths in the \verb!storage_service.xml! file.
+
+If we put the host certificates and the CA certificate somewhere else than \verb!/etc/grid-security!
+then we would need to modify the \verb!storage_service.xml!
+(look for \verb!KeyPath!, \verb!CertificatePath!
+and \verb!CACertificatesDir!), and we would need to modify \verb!/etc/arc/clientsslconfig.xml!
+as well.
+This file contains the credentials which is used by the services when they connect to an other service.
+If we put \verb!clientsslconfig.xml!
+somewhere else than \verb!/etc/arc!
+then we would need to modify all the \verb!ClientSSLConfig!
+tags in the \verb!storage_service.xml!.
+
+Let's run the \verb!arched! daemon with this config (specified by the \verb!-c! option) first in the foreground (using the \verb!-f! option) to see immediately if everything is right:
+
+\begin{verbatim}
+$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service.xml -f
+\end{verbatim}
+
+The template configuration file specifies a loglevel which only prints error messages, so nothing should be written on the screen now.
+The \verb!arched! daemon should now listen on the ports: 60001 for our data transfer service called Hopi, and 60000 for all the other services:
+
+\begin{verbatim}
+$ netstat -at | grep LISTEN
+tcp        0      0 *:60000                 *:*                     LISTEN     
+tcp        0      0 *:60001                 *:*                     LISTEN     
+\end{verbatim}
+
+Now it is time to set up or prototype client tool to access the system.
+It is currently called \verb!arc_storage_cli!
+and it is by default installed to \verb!/usr/local/bin!.
+We have to tell this CLI tool where it can find our user credentials, and a URL of a Bartender service.
+The Bartender service is the front-end of the storage system.
+We can specify these things by environment variables, or by a configuration file called \verb!~/.arc/client.xml!.
+Let's create this \verb!client.xml!
+file:
+
+\begin{verbatim}
+$ cat ~/.arc/client.xml 
+<ArcConfig>
+  <KeyPath>~/.arc/userkey.pem</KeyPath>
+  <CertificatePath>~/.arc/usercert.pem</CertificatePath>
+  <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+  <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
+</ArcConfig>
+\end{verbatim}
+
+Now we can use the \verb!arc_storage_cli! to access our local deployment. Let's do a `list' on the root collection (`/'):
+
+\begin{verbatim}
+$ arc_storage_cli
+Usage:
+  arc_storage_cli <method> [<arguments>]
+Supported methods: stat, make[Collection], unmake[Collection], list, move, put[File], get[File], del[File], pol[icy], unlink, credentialsDelegation, removeCredentials
+Without arguments, each method prints its own help.
+
+$ arc_storage_cli list
+- The URL of the Bartender: https://localhost:60000/Bartender
+- The key file: ~/.arc/userkey.pem
+- The cert file: ~/.arc/usercert.pem
+- The CA dir: /etc/grid-security/certificates
+Usage: list <LN> [<LN> ...]
+
+$ arc_storage_cli list /
+- The URL of the Bartender: https://localhost:60000/Bartender
+- The key file: ~/.arc/userkey.pem
+- The cert file: ~/.arc/usercert.pem
+- The CA dir: /etc/grid-security/certificates
+- Calling the Bartender's list method...
+[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component tcp.client(tcp) could not be created
+[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component tls.client(tls) could not be created
+[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component http.client(http) could not be created
+[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component soap.client(soap) could not be created
+ERROR: Wrong status from server.
+Maybe wrong URL: 'https://localhost:60000/Bartender'
+
+\end{verbatim}
+
 % section quick_start_guide (end)
 
-\section{Installation} % (fold)
-\label{sec:installation}
-
-% section installation (end)
-
-\section{Configuration} % (fold)
-\label{sec:configuration}
-
-% section configuration (end)
-
-\section{Deployment scenarios} % (fold)
-\label{sec:deployment_scenarios}
-
-% section deployment_scenarios (end)
+% \section{Installation} % (fold)
+% \label{sec:installation}
+% 
+% % section installation (end)
+% 
+% \section{Configuration} % (fold)
+% \label{sec:configuration}
+% 
+% % section configuration (end)
+% 
+% \section{Deployment scenarios} % (fold)
+% \label{sec:deployment_scenarios}
+% 
+% % section deployment_scenarios (end)
 
@@
 
-\section{Overview} % (fold)
-\label{sec:overview}
-
-% section overview (end)
-
-\section{Prototype command line tool} % (fold)
-\label{sec:prototype_command_line_tool}
-
-% section prototype_command_line_tool (end)
-
-\section{Grid jobs} % (fold)
-\label{sec:grid_jobs}
-
-% section grid_jobs (end)
-
-\section{Accessing the system with FUSE} % (fold)
-\label{sec:accessing_the_system_with_fuse}
-
-% section accessing_the_system_with_fuse (end)
+% \section{Overview} % (fold)
+% \label{sec:overview}
+% 
+% % section overview (end)
+% 
+% \section{Prototype command line tool} % (fold)
+% \label{sec:prototype_command_line_tool}
+% 
+% % section prototype_command_line_tool (end)
+% 
+% \section{Grid jobs} % (fold)
+% \label{sec:grid_jobs}
+% 
+% % section grid_jobs (end)
+% 
+% \section{Accessing the system with FUSE} % (fold)
+% \label{sec:accessing_the_system_with_fuse}
+% 
+% % section accessing_the_system_with_fuse (end)
 

### 2009-03-20 doc/client/client_technical.tex
Update migration and resubmission part

@@
 migration request of a given job, currently this request is only supported 
-for ARC1 clusters. The \texttt{url} and \texttt{jobdesc} specifies the job-id 
-and {\JobDescription} of the job which should be migrated.\\
+for ARC1 clusters. The \texttt{URL} and \texttt{jobdesc} specifies the job-id 
+and {\JobDescription} of the job which should be migrated.
+
+Note that whereas the \texttt{Migrate} method make provisions for automatically 
+getting files from the migration source cluster, this does not apply to the job 
+description which instead should be passed to \texttt{Migrate} explicitly in the 
+form of a {\JobDescription} object.
+
 The migration request will create a new job where the local input files are 
@@
 unsuccessful the new job will be terminated unless the \texttt{forcemigration} 
-boolean is \texttt{true}.\\
+boolean is \texttt{true}.
+
 The \texttt{Migrate} method will also fill the \texttt{info} {\XMLNode} with 
@@
 computed and compared with the ones listed in the users local XML file. If 
-all checksums match the job description is returned.\\
+all checksums match the job description is returned.
+
 In case \texttt{getlocal} is false or the retrieval of the local job description 
-fails the job description will be tried retrieved from the cluster. This has 
+fails an attempt will be made to retrieved the job description from the cluster. This has 
 currently only been implemented for ARC(both generations) clusters. Note 
-that the job descriptions retried from a cluster might will most likely have 
-been changed since the submission either by the cluster\footnote{For both 
+that the job descriptions retrieved from a cluster will most likely have 
+been changed since the submission -- either by the cluster\footnote{For both 
 generations of ARC it is possible to recover the initial job description send 
@@
 job description before sending it to the cluster.} or by the submission 
-client. Therefore the only way of retrieving the original job description is 
-through the users local XML file.\\
+client. Therefore, the only way of retrieving the original job description is 
+through the users local XML file.
+
 The \texttt{GetJobDescriptions} returns a list of the  {\Job}s where job description 
@@
 
-The \texttt{Migrate} method:
+The \texttt{Migrate} method
 \begin{shaded}
@@
 management. {\libarcclient} offers the following native commands:
-
 \begin{enumerate}
@@
 \item{\texttt{arcsync} - for synchronising the information in the user's 
-  local XML file about active jobs.}
-\item{\texttt{arcresub} - for redoing an entire job on the same or a new
-  computing resource.}
-\item{\texttt{arcmigrate} - for moving a queuing job from an ARC1 
-  computing resource to another.}
+local XML file\footnote{By default \$HOME/.arc/jobs.xml} about active jobs.}
+\item{\texttt{arcresub} - for starting an entire job over on either the 
+same or a new computing resource.}
+\item{\texttt{arcmigrate} - for moving a queuing job from one A-REX 
+server to another.}
 \end{enumerate}

### 2009-03-20 doc/storage_system/arc-storage-documentation.tex
Some spelling fixes

@@
 
-The A-Hash is a replicated metadata store, which is capable of consistently storing `objects' where an object contains property-value pairs organized in sections.
-All metadata about files and collections are stored in the A-Hash, and some other information (e.g.
+The A-Hash is the replicated\footnote{Depends on system configuration.
+It is possible to configure the storage system with only one central A-Hash too} metadata store of the storage system.
+It consistently stores `objects' which contain property-value pairs organized in sections.
+All metadata about files and collections in the storage system are stored in the A-Hash.
+Additionall system configutation information (e.g.
 about A-Hash replication, about Shepherd services, etc.) is stored in it as well.
-The A-Hash itself does not interpret the data, it basically just stores tuples of strings.
+The A-Hash itself does not interpret any data.
 
@@
 
-The Librarian is capable of managing the hierarchy and metadata of files and collections, and health information of the Shepherd services.
+The Librarian manages the storage system hierarchy of files and collections including their metadata.
 It can traverse Logical Names and return the corresponding metadata.
-It can receive heartbeat messages from Shepherd services, and it automatically modify the states of files if needed.
-The Librarian itself is a stateless service, it uses the A-Hash to actually store and retrieve the metadata, that’s why there could be any number of independent Librarian services (all using the same A-Hashes) which provides high-availability and load-balancing.
+The Shephards send their heartbeat messages to the Librarians which if needed respond by modifying the states of files stored on a certain storage node.
+The Librarian itself is a stateless service, it uses the A-Hash to store and retrieve the metadata and thus there can be any number of independent Librarian services (all using the same A-Hashes).
+This ensures high-availability and load-balancing of the storage system.
 
@@
 
-The Bartender service provides a high-level interface for the storage system to the clients.
-Every interaction between a client and the ARC storage system begins with a request to a Bartender.
-You can create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names.
-The Bartender authorizes users and force access policies of files and collections.
+The Bartender service provides the high-level interface of the storage system to the clients.
+Every interaction between a client and the ARC storage system starts by sending a request to a Bartender.
+Using the Bartender it is possible to create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names, etc.
+The Bartender authorizes users and apply access policies of files and collections.
 It communicates with the Librarian and Shepherd services to accomplish the client’s requests.
-The actual file data does not go through the Bartender; file transfers are directly performed between the storage nodes and the clients.
-There could be any number of independent Bartender services in the system which provides high-availability and load-balancing.
+The file data itself does however not go through the Bartender; file transfers are directly performed between the storage nodes and the clients.
+There could be any number of independent Bartender services in the system which ensure high-availability and load-balancing.
 The Bartender also provides a way to access files on third-party storage solutions through its interface by mounting the namespace of the third-party storage into the namespace of the ARC storage (this is accomplished by so called `gateway' modules).
@@
 
-The files in the ARC storage system are usually replicated on different storage nodes.
-For each storage node there is a Shepherd service which manages the storage element service on the node, reports its health state to a Librarian and provides the interface for initiating file transfers.
+The Shepherd service manages the storage element service on the storage node.
+It reports its health state to a Librarian and provides the interface for initiating file transfers.
 For each kind of storage element service (e.g.
-a HTTP server, an FTP server, a storage solution with a GridFTP interface, etc.) it is needed to have a Shepherd backend which is capable of managing the given storage element service.
-The Shepherd service periodically checks the health of the replicas based on their checksums, and if a replica is deleted or corrupted, the Shepherd tries to recover it by downloading a valid copy from an other storage node.
-The Shepherds also check if a file has fewer replicas in the system than needed, and they initiate replication if needed.
+a HTTP server, an FTP server, a storage solution with a GridFTP interface, etc.) a Shepherd backend is needed which is capable of managing the given storage element service.
+The Shepherd service periodically checks the health of the replicas using their checksums, and if a replica is deleted or corrupted, the Shepherd tries to recover it by downloading a valid copy from another storage node.
+The Shepherd also periodicallys check if a file has the correct number of replicas in the system.
+If too few or too many replicas are found, new replicas will be created or redundant replicas removed.
 
@@
 
-The ARC storage system consists of several services.
-Most of the services (A-Hash, Librarian, Shepherd) are `internal' services in a way that the end-user of the storage system never communicates with them directly.
-But these internal services are communicating with each other, so they have to know who to trust.
+The ARC storage system consists of several services (both different kinds of services and multiple instances of the same kind).
+The A-Hash, Librarian and Shepherd services are `internal' services in the sense that the end-user of the storage system never communicates with them directly.
+But these internal services are communicating with each other, thus they have to know which other services to trust.
 We call this aspect of the security architecture of the ARC storage `inter-service authorization'.
 
-The end users always connect to one of the Bartender services, which will decide if the user has permissions to do something or not.
-This is the `high-level authorization' part of the security architecture.
-
-For transferring the actual file data, the users have to connect to storage element services which are sitting on storage nodes.
+The end users always connect to one of the Bartender services, which will decide which permission the user has. This is the `high-level authorization' part of the security architecture.
+
+For transferring the actual file data, the users have to connect to storage element services running on storage nodes.
 These services also have their own authentication and authorization methods.
@@
 
-In a deployment of the ARC storage system, we could have several A-Hash, Librarian, Shepherd and Bartender services.
+In a deployment of the ARC storage system it is possible to have several A-Hash, Librarian, Shepherd and Bartender services.
 The Bartenders send requests to the Librarians and the Shepherds, the Shepherds communicate with the Librarians, the Librarians talk with the A-Hashes.
-If any of these services get compromised or a new rouge service gets inserted in the system, we loose security completely.
-That's why it is vital for each service to authorize the services before sending or accepting requests.
-The services communicates via HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and they can examine the other service's certificates.
-Because of these X.509 certificates each service has Distinguish Name (DN).
-We can use these DNs to exactly specify which services we trust.
-We can configure a list of trusted DNs into each service, or we can store this list on a remote location.
+If any of these services gets compromised or a new rouge service is inserted in the system, security is lost.
+That's why it is vital for each service to authorize other services before sending or accepting requests.
+The services communicate via the HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and that they can examine eachother's certificates.
+Because of these X.509 certificates each service has Distinguish Name (DN), which can be used for identifying trusted services.
+A list of trusted DNs can be configured into each service, or a list of trusted services can be stored on a remote location.
 The services will only accept connections if the DN of the other end is listed in this list of trusted DNs.
@@
 The Librarian component of the ARC storage system stores all the metadata about files and collections.
-For each file and collection there are access policies in the form of access control rules, and these are stored among these metadata.
+For each file and collection there are access policies in the form of access control rules, and these are stored among the metadata.
 The users are identified by their DNs, and an access control rule specify the rights of the given user.
-One rule can be represented like this:
+A rule can be represented like this:
 

### 2009-03-20 doc/storage_system/arc-storage-manual.tex
adding some user's manual

@@
 
-If there is no error on importing any of the packages, we are good to go.
+If there is no error on importing any of the packages, we are good to go. If the packages cannot be imported, we should set the PYTHONPATH environment variable to point to their location.
 
@@
 <ArcConfig>
-  <KeyPath>~/.arc/userkey.pem</KeyPath>
-  <CertificatePath>~/.arc/usercert.pem</CertificatePath>
+  <KeyPath>/home/<username>/.arc/userkey.pem</KeyPath>
+  <CertificatePath>/home/<username>/.arc/usercert.pem</CertificatePath>
   <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
@@
   arc_storage_cli <method> [<arguments>]
-Supported methods: stat, make[Collection], unmake[Collection], list, move, put[File], get[File], del[File], pol[icy], unlink, credentialsDelegation, removeCredentials
+Supported methods:
+  stat - get detailed information about an entry
+  makeCollection | make | mkdir - create a collection
+  unmakeCollection | unmake | rmdir - remove an empty collection
+  list | ls - list the content of a collection
+  move | mv - move entries within the namespace
+  putFile | put - upload a file
+  getFile | get - download a file
+  delFile | del | rm - remove a file
+  modify | mod - modify metadata
+  policy | pol - modify access policy rules
+  unlink - remove a link to an entry from a collection without removing the entry itself
+  credentialsDelegation | cre - delegate credentials for using gateway
+  removeCredentials | rem - remove previously delegated credentials
 Without arguments, each method prints its own help.
@@
 - The URL of the Bartender: https://localhost:60000/Bartender
-- The key file: ~/.arc/userkey.pem
-- The cert file: ~/.arc/usercert.pem
+- The key file: /home/zsombor/.arc/userkey.pem
+- The cert file: /home/zsombor/.arc/usercert.pem
 - The CA dir: /etc/grid-security/certificates
@@
 - The URL of the Bartender: https://localhost:60000/Bartender
-- The key file: ~/.arc/userkey.pem
-- The cert file: ~/.arc/usercert.pem
+- The key file: /home/zsombor/.arc/userkey.pem
+- The cert file: /home/zsombor/.arc/usercert.pem
 - The CA dir: /etc/grid-security/certificates
 - Calling the Bartender's list method...
-[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component tcp.client(tcp) could not be created
-[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component tls.client(tls) could not be created
-[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component http.client(http) could not be created
-[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544] Component soap.client(soap) could not be created
+[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544]
+    Component tcp.client(tcp) could not be created
+[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544]
+    Component tls.client(tls) could not be created
+[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544]
+    Component http.client(http) could not be created
+[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544]
+    Component soap.client(soap) could not be created
 ERROR: Wrong status from server.
 Maybe wrong URL: 'https://localhost:60000/Bartender'
+\end{verbatim}
 
+If you get this error, where the `Component [...] could not be created', we should reinforce the python environment about the location of our python packages:
+
+\begin{verbatim}
+$ export PYTHONPATH=/usr/local/lib/python2.5/site-packages/
 \end{verbatim}
+
+Without arguments, the \verb!arc_storage_cli!
+prints the list of available methods.
+Specifying only a method name prints the syntex of the given method.
+The \verb!list!
+method requires Logical Names (LN).
+In the storage system, every file and collection has a Logical Name, which is a path within the namespace.
+
+Now we have a running (centralized) storage service including a Hopi service, which is a basic HTTP server, for transfering files.
+We have a Bartender service which is the front-end of the system to the users, so our client always connect to the Bartender service.
+There are several directories configured in the config file where the services store their data.
+These directories can always be emptied, which resets the whole system to a clean state.
+If we run the storage system with clean data directories, then it is completely empty, which means that there is not even a root collection.
+
+\begin{verbatim}
+$ arc_storage_cli list /
+- The URL of the Bartender: https://localhost:60000/Bartender
+- The key file: /home/zsombor/.arc/userkey.pem
+- The cert file: /home/zsombor/.arc/usercert.pem
+- The CA dir: /etc/grid-security/certificates
+- Calling the Bartender's list method...
+- done in 0.34 seconds.
+'/': not found
+\end{verbatim}
+
+(From now on I will omit the first few lines of the \verb!arc_storage_cli!'s output.)
+
+The root collection has the Logical Name of `\verb!/!', and it's role to provide a starting point for all the other Logical Names. We have to create this root collection first:
+
+\begin{verbatim}
+$ arc_storage_cli mkdir /
+- Calling the Bartender's makeCollection method...
+- done in 0.81 seconds.
+Creating collection '/': done
+
+$ arc_storage_cli list /
+- Calling the Bartender's list method...
+- done in 0.55 seconds.
+'/': collection
+    empty.
+\end{verbatim}
+
+Now we do have an empty root collection.
 
@@
 \label{cha:user_s_manual}
+
+Assume that we have a running storage system and we want to access it with the \verb!arc_storage_cli!
+client.
+We have the \verb!~/.arc/client.xml!
+set properly, so we can list the contents of the root collection:
+
+\begin{verbatim}
+$ arc_storage_cli list /
+- Calling the Bartender's list method...
+- done in 0.55 seconds.
+'/': collection
+    empty.
+\end{verbatim}
+
+Let's create a small file and upload it:
+
+\begin{verbatim}
+$ cat > orange
+orange
+$ cat orange 
+orange
+
+$ arc_storage_cli put orange /
+- The size of the file is 7 bytes
+- Calculating md5 checksum...
+- done in 0.0003 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
+- Calling the Bartender's putFile method...
+- done in 1.26 seconds.
+- Got transfer URL: http://localhost:60001/hopi/7fcef91e-ac9c-658d-4def-17d52ba33cfb
+- Uploading from 'orange'
+    to 'http://localhost:60001/hopi/7fcef91e-ac9c-658d-4def-17d52ba33cfb' with http...
+    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+- done in 0.0182 seconds.
+'orange' (7 bytes) uploaded as '/orange'.
+
+$ arc_storage_cli list /
+- Calling the Bartender's list method...
+- done in 0.51 seconds.
+'/': collection
+  orange              <file>
+
+\end{verbatim}
+
+We can get additional information about the file with the stat method:
+
+\begin{verbatim}
+$ arc_storage_cli stat /orange
+- Calling the Bartender's stat method...
+- done in 0.40 seconds.
+'/orange': found
+  states
+    checksumType: md5
+    neededReplicas: 1
+    size: 7
+    checksum: 6aefd2842be62cd470709b27aedc7db7
+  entry
+    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
+    GUID: 12ce831b-eab8-1f47-29f6-19fb85ebf46a
+    type: file
+  parents
+    0/orange: parent
+  locations
+    https://localhost:60000/Shepherd 3e9ea612-87c0-1e81-58f3-cec6b0b125c1: alive
+  timestamps
+    created: 1237504821.91
+\end{verbatim}
+
+Here we can see the owner of the file (Billy), the location of the replicas (currently only one), the number of needed replicas (which is also: one), the checksum and size of the file, and the timestamp of the creation.
+
+We can download this file:
+
+\begin{verbatim}
+$ arc_storage_cli get /orange /tmp
+- Calling the Bartender's getFile method...
+- done in 0.51 seconds.
+- Got transfer URL: http://localhost:60001/hopi/521c5869-4c93-8e28-aa3a-03f477f34f63
+- Downloading from 'http://localhost:60001/hopi/521c5869-4c93-8e28-aa3a-03f477f34f63'
+    to '/tmp/orange' with http...
+    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+- done in 1.0071 seconds.
+'/orange' (7 bytes) downloaded as '/tmp/orange'.
+$ cat /tmp/orange 
+orange
+\end{verbatim}
+
+We can create collections which can contain files and other collections forming a tree-hierarchy. We can move files and collection around within this namespace.
+
+\begin{verbatim}
+$ arc_storage_cli mkdir /fruits
+- Calling the Bartender's makeCollection method...
+- done in 0.79 seconds.
+Creating collection '/fruits': done
+
+$ arc_storage_cli mkdir /fruits/apple
+- Calling the Bartender's makeCollection method...
+- done in 0.91 seconds.
+Creating collection '/fruits/apple': done
+
+$ arc_storage_cli move /orange /fruits/
+- Calling the Bartender's move method...
+- done in 0.99 seconds.
+Moving '/orange' to '/fruits/': moved
+
+$ arc_storage_cli list /fruits
+- Calling the Bartender's list method...
+- done in 0.60 seconds.
+'/fruits': collection
+  orange              <file>
+  apple               <collection>
+\end{verbatim}
+
+We can remove files, and remove collections (but only if the collection is empty):
+
+\begin{verbatim}
+$ arc_storage_cli rmdir /fruits
+- Calling the Bartender's unmakeCollection method...
+- done in 0.41 seconds.
+Removing collection '/fruits': collection is not empty
+
+$ arc_storage_cli rmdir /fruits/apple
+- Calling the Bartender's unmakeCollection method...
+- done in 0.89 seconds.
+Removing collection '/fruits/apple': removed
+
+$ arc_storage_cli rm /fruits/orange
+- Calling the Bartender's delFile method...
+- done in 0.89 seconds.
+/fruits/orange: deleted
+
+$ arc_storage_cli rmdir /fruits
+- Calling the Bartender's unmakeCollection method...
+- done in 0.79 seconds.
+Removing collection '/fruits': removed
+\end{verbatim}
+
+Let's introduce a new user into the system.
+If we have some other user certificate and key files, then we can tell \verb!arc_storage_cli!
+to use them with modifying the \verb!client.xml!
+or by specifying the \verb!ARC_CERT_FILE!
+and \verb!ARC_KEY_FILE!
+environment variables.
+
+\begin{verbatim}
+$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
+arc_storage_cli list /
+- Calling the Bartender's list method...
+- done in 0.31 seconds.
+'/': denied
+
+$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
+arc_storage_cli put /tmp/orange /pennys-orange
+- The size of the file is 7 bytes
+- Calculating md5 checksum...
+- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
+- Calling the Bartender's putFile method...
+- done in 0.81 seconds.
+/pennys-orange: failed to add child to parent
+
+\end{verbatim}
+
+Listing the root collection or uploading files into it for Penny is denied, because the root collection is owned by Billy, and nobody else has permissions to it.
+Let's ask Billy to allow for everyone to access the root collection:
+
+\begin{verbatim}
+$ arc_storage_cli pol / change ALL +read +addEntry
+- Calling the Bartender's stat method...
+- done in 0.30 seconds.
+old action list for the user ALL was ''
+the new action list of user ALL for LN / will be '+read +addEntry'
+- Calling the Bartender's modify method...
+- done in 0.50 seconds.
+set.
+\end{verbatim}
+
+The syntax of the \verb!policy! method is the following:
+
+\begin{verbatim}
+$ arc_storage_cli pol
+Usage: policy <LN> <changeType> <identity> <action list>
+  <changeType> could be 'set', 'change' or 'clear'
+    'set': sets the action list to the given user overwriting the old one
+    'change': modify the current action list with adding and removing actions
+    'clear': clear the action list of the given user
+  <identity> could be a '<UserDN>' or a 'VOMS:<VO name>'
+  <action list> is a list actions prefixed with '+' or '-'
+    e.g. '+read +addEntry -delete'
+    possible actions:
+        read addEntry removeEntry delete
+        modifyPolicy modifyStates modifyMetadata
+\end{verbatim}
+
+So we have changed the action list for user `ALL' to `+read +addEntry', which means that everybody can list the contents of this collection, create subcollection in it, upload a file into it, or move some existing entry into it.
+
+Now we can upload a file with Penny:
+
+\begin{verbatim}
+$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
+arc_storage_cli put /tmp/orange /pennys-orange
+- The size of the file is 7 bytes
+- Calculating md5 checksum...
+- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
+- Calling the Bartender's putFile method...
+- done in 1.18 seconds.
+- Got transfer URL: http://localhost:60001/hopi/4145891d-0e4c-b6bd-e4d6-dc80554b9d35
+- Uploading from '/tmp/orange'
+    to 'http://localhost:60001/hopi/4145891d-0e4c-b6bd-e4d6-dc80554b9d35' with http...
+    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+- done in 1.0071 seconds.
+'/tmp/orange' (7 bytes) uploaded as '/pennys-orange'.
+\end{verbatim}
+
+This file now only can be downloaded by Penny, so let's allow for Billy to download it as well:
+
+\begin{verbatim}
+$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
+arc_storage_cli pol /pennys-orange change /DC=eu/DC=KnowARC/O=knowarc/CN=billy +read
+- Calling the Bartender's stat method...
+- done in 0.41 seconds.
+old action list for the user /DC=eu/DC=KnowARC/O=knowarc/CN=billy was ''
+the new action list of user /DC=eu/DC=KnowARC/O=knowarc/CN=billy for LN /pennys-orange
+    will be '+read'
+- Calling the Bartender's modify method...
+- done in 0.59 seconds.
+set.
+
+$ arc_storage_cli stat /pennys-orange
+- Calling the Bartender's stat method...
+- done in 0.43 seconds.
+'/pennys-orange': found
+  locations
+    https://localhost:60000/Shepherd 1cc3b115-2350-8334-d206-a7b891912c00: alive
+  states
+    checksumType: md5
+    neededReplicas: 1
+    size: 7
+    checksum: 6aefd2842be62cd470709b27aedc7db7
+  parents
+    0/pennys-orange: parent
+  timestamps
+    created: 1237509633.17
+  policy
+    /DC=eu/DC=KnowARC/O=knowarc/CN=billy: +read
+  entry
+    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=penny
+    GUID: 0e6a4c9a-d4e7-e0dc-e083-16b83d460d82
+    type: file
+\end{verbatim}
+
+We can see among the metadata of the file, that Billy has `read' rights. Billy can now download this file:
+
+\begin{verbatim}
+$ arc_storage_cli get /pennys-orange /tmp
+- Calling the Bartender's getFile method...
+- done in 0.50 seconds.
+- Got transfer URL: http://localhost:60001/hopi/c16d7e12-42b0-7882-c09e-1cac31dbdbcd
+- Downloading from 'http://localhost:60001/hopi/c16d7e12-42b0-7882-c09e-1cac31dbdbcd'
+    to '/tmp/pennys-orange' with http...
+    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+- done in 1.0065 seconds.
+'/pennys-orange' (7 bytes) downloaded as '/tmp/pennys-orange'.
+$ cat /tmp/pennys-orange 
+orange
+\end{verbatim}
+
+We can specify the needed number of replicas for a new file with the \verb!ARC_NEEDED_REPLICAS! environment variable.
+
+\begin{verbatim}
+$ ARC_NEEDED_REPLICAS=2 arc_storage_cli put orange /new-orange
+- The size of the file is 7 bytes
+- Calculating md5 checksum...
+- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
+- Calling the Bartender's putFile method...
+- done in 1.11 seconds.
+- Got transfer URL: http://localhost:60001/hopi/0ad54182-7618-a2af-c490-dea7c4d6fc31
+- Uploading from 'orange' to
+    'http://localhost:60001/hopi/0ad54182-7618-a2af-c490-dea7c4d6fc31' with http...
+    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+- done in 1.0057 seconds.
+'orange' (7 bytes) uploaded as '/new-orange'.
+
+$ arc_storage_cli stat /new-orange
+- Calling the Bartender's stat method...
+- done in 0.40 seconds.
+'/new-orange': found
+  states
+    checksumType: md5
+    neededReplicas: 2
+    size: 7
+    checksum: 6aefd2842be62cd470709b27aedc7db7
+  entry
+    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
+    GUID: 4e673e28-4769-50af-f02e-d7a67632f87a
+    type: file
+  parents
+    0/new-orange: parent
+  locations
+    https://localhost:60000/Shepherd 4aacd276-653c-3ca5-c1a2-78a876c129c9: alive
+  timestamps
+    created: 1237509989.43
+
+\end{verbatim}
+
+We can see that the `neededReplica' metadata says 2, but if we deployed the system in a centralized way for testing purposes, then we only have one storage element (which is a Shepherd and a Hopi service together), so the file can only have one replica, that's why there is only one locations.
+
+We can modify the number of needed replicas of an existing file with the `modify' method:
+
+\begin{verbatim}
+$ arc_storage_cli modify /new-orange set states neededReplicas 1
+- Calling the Bartender's modify method...
+- done in 0.63 seconds.
+set
+
+$ arc_storage_cli stat /new-orange | grep neededReplicas
+    neededReplicas: 1
+\end{verbatim}
 

### 2009-03-20 doc/storage_system/arc-storage-documentation.tex
arc dmc description started

@@
 
-TODO: what is the ARC DMC, how does it work
+To access data through the ARC middleware client tools, one needs to go through Data Manager Components (DMCs).
+These are protocol specific plugins to the client tools.
+E.g, if you need to access data from a HTTPS service, you will go through the HTTP DMC using a URL starting with \verb!https://!and if you want to access data from a SRM service, you will use the SRM DMC using a URL starting with \verb!srm://!.
+Similarly, if you want access the ARC storage system, you will use the ARC DMC with protocol using a URL starting with \verb!arc://!.
+
+As with the \verb!arc_storage_cli!, the credentials of the user and the URL of the Bartender can be configured with an XML file located at \verb!~/.arc/client.xml!.
+When the ARC DMC is initiated it will create an instance of \verb!DataPointARC!
+which provides functionality for listing collections and file IO.
+When downloading/uploading files from/to the storage system, the data point will use the provided Bartender URL to acquire a TURL from a Bartender.
+This TURL will be given to a \verb!DataTransfer!.
+Calling the constructor of the DataHandle class with a URL will trigger the iGetDataPoint class of all registered DMCs to be called until a non-NULL pointer is returned.
+This means that as long as there exists a data point plugin for the protocol of the returned TURL, the transfer can proceed.
+
+TODO: Some example of uploading, downloading and stat files, anything more?
 

### 2009-03-20 doc/storage_system/arc-storage-manual.tex
manual about adding more storage elements
TODO: manual about adding more Librarians and Bartenderds
TODO: manual about changing the CentralAHash to ReplicatedAHash

@@
 
-\chapter{Administrator's manual} % (fold)
-\label{cha:administrator_s_manual}
-
-\section{Quick start guide} % (fold)
-\label{sec:quick_start_guide}
+\chapter{Quick start guide} % (fold)
+
+\section{Installing a centralized system} % (fold)
 
@@
 
-% section quick_start_guide (end)
-
-% \section{Installation} % (fold)
-% \label{sec:installation}
-% 
-% % section installation (end)
-% 
-% \section{Configuration} % (fold)
-% \label{sec:configuration}
-% 
-% % section configuration (end)
-% 
-% \section{Deployment scenarios} % (fold)
-% \label{sec:deployment_scenarios}
-% 
-% % section deployment_scenarios (end)
-
-% chapter administrator_s_manual (end)
-
-\chapter{User's manual} % (fold)
-\label{cha:user_s_manual}
+\section{Using the prototype client tool} % (fold)
 
@@
 \begin{verbatim}
-$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
+$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
 arc_storage_cli list /
@@
 
-$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
+$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
 arc_storage_cli put /tmp/orange /pennys-orange
@@
 \begin{verbatim}
-$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
+$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
 arc_storage_cli put /tmp/orange /pennys-orange
@@
 \begin{verbatim}
-$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
+$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
 arc_storage_cli pol /pennys-orange change /DC=eu/DC=KnowARC/O=knowarc/CN=billy +read
@@
 
-We can see that the `neededReplica' metadata says 2, but if we deployed the system in a centralized way for testing purposes, then we only have one storage element (which is a Shepherd and a Hopi service together), so the file can only have one replica, that's why there is only one locations.
+We can see that the `neededReplica' metadata says 2, but if we deployed the system in a centralized way for testing purposes, then we only have one storage element (which is a Shepherd service and a Hopi service together), so the file can only have one replica, that's why there is only one locations.
 
@@
 
-% \section{Overview} % (fold)
-% \label{sec:overview}
-% 
-% % section overview (end)
-% 
-% \section{Prototype command line tool} % (fold)
-% \label{sec:prototype_command_line_tool}
-% 
-% % section prototype_command_line_tool (end)
-% 
-% \section{Grid jobs} % (fold)
-% \label{sec:grid_jobs}
-% 
-% % section grid_jobs (end)
-% 
-% \section{Accessing the system with FUSE} % (fold)
-% \label{sec:accessing_the_system_with_fuse}
-% 
-% % section accessing_the_system_with_fuse (end)
-
-% chapter user_s_manual (end)
-
-% \bibliography{grid}
+\section{Adding more storage elements}
+
+Let's add a new storage element to the system.
+We need to deploy a Shepherd service and a storage element service, which this time will be a `byteio' service.
+The `byteio' service implements a small part of the ByteIO interface, which means that it just put the entire file data into the SOAP message, so it is only suitable for small files.
+
+But first modify the configuration file of our original server.
+Currently all the transfer URLs which were generated when we upload and download files contained the hostname `localhost'.
+This is not OK if we want to access it from an other machine, we should provide transfer URLs which contains the external hostname of the machine.
+So lets change this line in \verb!/etc/arc/storage_service.xml!:
+
+\begin{verbatim}
+    <TURLPrefix>http://localhost:60001/hopi/</TURLPrefix>
+\end{verbatim}
+
+to the real hostname of the machine.
+But this is not enough.
+Currently the Shepherd services uses their URLs as IDs, and the other services use these IDs to connect to a Shepherd service, so we need to change to ID of the Shepherd service to a real external URL as well.
+And for the sake of clarity let's change all `localhost' entries in the config file to our real hostname.
+In our case the name of this machine is just `storage' so the Shepherd section should look like this:
+
+\begin{verbatim}
+<Service name="pythonservice" id="shepherd">
+    <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
+    <ServiceID>https://storage:60000/Shepherd</ServiceID>
+    <CheckPeriod>20</CheckPeriod>
+    <MinCheckInterval>0.1</MinCheckInterval>
+    <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
+    <StoreCfg><DataDir>/var/spool/arc/shepherd_data</DataDir></StoreCfg>
+    <BackendClass>storage.shepherd.hardlinkingbackend.HopiBackend</BackendClass>
+    <BackendCfg>
+        <DataDir>/var/spool/arc/shepherd_store</DataDir>
+        <TransferDir>/var/spool/arc/shepherd_transfer</TransferDir>
+        <TURLPrefix>http://storage:60001/hopi/</TURLPrefix>
+    </BackendCfg>
+    <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
+    <BartenderURL>https://storage:60000/Bartender</BartenderURL>
+    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
+</Service>
+\end{verbatim}
+
+Let's check first if our centralized deployment is still working, stop the \verb!arched! daemon remove the data directories to make it a clean start, and restart it with the modified config.
+
+\begin{verbatim}
+$ sudo rm -r /var/spool/arc/*
+$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service.xml -f    
+\end{verbatim}
+
+Let's see if we can still use the client:
+
+\begin{verbatim}
+$ arc_storage_cli list /
+- Calling the Bartender's list method...
+- done in 0.31 seconds.
+'/': not found
+
+$ arc_storage_cli mkdir /
+- Calling the Bartender's makeCollection method...
+- done in 0.63 seconds.
+Creating collection '/': done
+
+$ arc_storage_cli list /
+- Calling the Bartender's list method...
+- done in 0.50 seconds.
+'/': collection
+    empty.
+\end{verbatim}
+
+Let's upload a file and ask for two replicas:
+
+\begin{verbatim}
+$ ARC_NEEDED_REPLICAS=2 arc_storage_cli put orange /
+- The size of the file is 7 bytes
+- Calculating md5 checksum...
+- done in 0.0005 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
+- Calling the Bartender's putFile method...
+- done in 1.10 seconds.
+- Got transfer URL: http://storage:60001/hopi/139dc57d-8c12-c188-a4a6-973449553811
+- Uploading from 'orange'
+    to 'http://storage:60001/hopi/139dc57d-8c12-c188-a4a6-973449553811' with http...
+    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+- done in 1.0090 seconds.
+'orange' (7 bytes) uploaded as '/orange'.
+
+$ arc_storage_cli stat /orange
+- Calling the Bartender's stat method...
+- done in 0.41 seconds.
+'/orange': found
+  states
+    checksumType: md5
+    neededReplicas: 2
+    size: 7
+    checksum: 6aefd2842be62cd470709b27aedc7db7
+  entry
+    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
+    GUID: 3fdad88a-944f-79c5-4f23-c5b3d3bc981e
+    type: file
+  parents
+    0/orange: parent
+  locations
+    https://storage:60000/Shepherd 0b46f53c-419d-ce4e-181a-ec95d16dcc95: alive
+  timestamps
+    created: 1237519441.96
+\end{verbatim}
+
+Currently it has only one replica, because we still don't have more than one storage elements.
+While \verb!arched!
+keeps running, let's create a new config for the other machine, which is called \verb!storage2!:
+
+\begin{verbatim}
+<?xml version="1.0"?>
+<ArcConfig 
+  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
+  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
+>
+    <Server>
+        <Pidfile>/var/run/arched2.pid</Pidfile>
+        <Logger level="ERROR">/var/log/arched2.log</Logger>
+    </Server>
+    <ModuleManager>
+        <Path>/usr/local/lib/arc/</Path>
+    </ModuleManager>
+    <Plugins><Name>mcctcp</Name></Plugins>
+    <Plugins><Name>mcctls</Name></Plugins>
+    <Plugins><Name>mcchttp</Name></Plugins>
+    <Plugins><Name>mccsoap</Name></Plugins>
+    <Chain>
+        <Component name="tcp.service" id="tcp">
+            <next id="tls"/> 
+            <tcp:Listen>
+                <tcp:Port>60006</tcp:Port>
+                <tcp:Version>4</tcp:Version>
+            </tcp:Listen>
+        </Component>
+        <Component name="tls.service" id="tls">
+            <next id="http"/>
+            <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
+            <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
+            <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+        </Component>
+        <Component name="http.service" id="http">
+            <next id="soap">POST</next>
+        </Component>
+        <Component name="soap.service" id="soap">
+            <next id="plexer"/>
+        </Component>
+        <Plexer name="plexer.service" id="plexer">
+            <next id="shepherd">^/Shepherd$</next>
+            <next id="byteio">^/byteio/</next>
+        </Plexer>
+        <Service name="pythonservice" id="shepherd">
+            <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
+            <ServiceID>https://storage2:60006/Shepherd</ServiceID>
+            <CheckPeriod>20</CheckPeriod>
+            <MinCheckInterval>0.1</MinCheckInterval>
+            <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
+            <StoreCfg>
+                <DataDir>/var/spool/arc/shepherd_store2</DataDir>
+            </StoreCfg>
+            <BackendClass>storage.shepherd.byteio.ByteIOBackend</BackendClass>
+            <BackendCfg>
+                <DataDir>/var/spool/arc/shepherd_data2</DataDir>
+                <TransferDir>/var/spool/arc/shepherd_transfer2</TransferDir>
+                <TURLPrefix>https://storage2:60006/byteio/</TURLPrefix>
+            </BackendCfg>
+            <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
+            <BartenderURL>https://storage:60000/Bartender</BartenderURL>
+            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml"/>
+        </Service>
+        <Service name="pythonservice" id="byteio">
+            <ClassName>storage.shepherd.byteio.ByteIOService</ClassName>
+            <TransferDir>/var/spool/arc/shepherd_transfer2</TransferDir>
+            <NotifyURL>https://storage2:60006/Shepherd</NotifyURL>
+            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml"/>
+        </Service>
+    </Chain>
+</ArcConfig>    
+\end{verbatim}
+
+This configuration contains only two services: a Shepherd, and byteio service.
+The byteio service is just a very simple storage element service which moves files within SOAP messages.
+The Shepherd service needs to have a backend for all supported storage element services, so it has a ByteIOBackend configured.
+This Shepherd connects to our original server's Bartender and Librarian services.
+(See the technical documentation for details about the services.)
+
+While the original (\verb!storage!) server still runs the services, let's start this new server:
+
+\begin{verbatim}
+$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service_2.xml -f
+\end{verbatim}
+
+And after a minute or two, let's check our file:
+
+\begin{verbatim}
+$ arc_storage_cli stat /orange
+- Calling the Bartender's stat method...
+- done in 0.41 seconds.
+'/orange': found
+  states
+    checksumType: md5
+    neededReplicas: 2
+    size: 7
+    checksum: 6aefd2842be62cd470709b27aedc7db7
+  entry
+    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
+    GUID: 24b442eb-ab1a-680e-1f79-4c20479d3ea9
+    type: file
+  parents
+    0/orange: parent
+  locations
+    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
+    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
+  timestamps
+    created: 1237519828.4
+\end{verbatim}
+
+It indeed has two replicas now, one on each server.
+
+Let's download it multiple times:
+
+\begin{verbatim}
+$ arc_storage_cli get /orange /tmp/orange1
+- Calling the Bartender's getFile method...
+- done in 0.50 seconds.
+- Got transfer URL: http://storage:60001/hopi/de0d951f-7379-1ce4-0813-48a384c11982
+- Downloading from 'http://storage:60001/hopi/de0d951f-7379-1ce4-0813-48a384c11982'
+    to '/tmp/orange1' with http...
+    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+- done in 1.0060 seconds.
+'/orange' (7 bytes) downloaded as '/tmp/orange1'.
+
+$ arc_storage_cli get /orange /tmp/orange2
+- Calling the Bartender's getFile method...
+- done in 0.51 seconds.
+- Got transfer URL: https://storage2:60006/byteio/52b6bffa-4812-03a5-93a1-834e38077ac5
+- Downloading from 'https://storage2:60006/byteio/52b6bffa-4812-03a5-93a1-834e38077ac5'
+    to '/tmp/orange2' with byteio...
+- done in 0.2822 seconds.
+'/orange' (7 bytes) downloaded as '/tmp/orange2'.
+\end{verbatim}
+
+First it was downloaded from the \verb!storage! server via HTTP, second time it was downloaded from the \verb!storage2! server via ByteIO.
+
+Let's create a third storage element on a machine called \verb!storage3! this time using a Hopi service as storage element service like on our main \verb!storage! server:
+
+\begin{verbatim}
+zsombor@storage:~$ cat /etc/arc/storage_service_3.xml 
+<?xml version="1.0"?>
+<ArcConfig 
+  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
+  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
+>
+    <Server>
+        <Pidfile>/var/run/arched3.pid</Pidfile>
+        <Logger level="ERROR">/var/log/arched3.log</Logger>
+    </Server>
+    <ModuleManager>
+        <Path>/usr/local/lib/arc/</Path>
+    </ModuleManager>
+    <Plugins><Name>mcctcp</Name></Plugins>
+    <Plugins><Name>mcctls</Name></Plugins>
+    <Plugins><Name>mcchttp</Name></Plugins>
+    <Plugins><Name>mccsoap</Name></Plugins>
+    <Chain>
+        <Component name="tcp.service" id="tcp">
+            <next id="tls"/> 
+            <tcp:Listen>
+		<tcp:Port>60010</tcp:Port>
+		<tcp:Version>4</tcp:Version>
+	    </tcp:Listen>
+        </Component>
+        <Component name="tls.service" id="tls">
+            <next id="http"/>
+            <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
+            <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
+            <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+        </Component>
+        <Component name="http.service" id="http">
+            <next id="soap">POST</next>
+        </Component>
+        <Component name="soap.service" id="soap">
+            <next id="plexer"/>
+        </Component>
+        <Plexer name="plexer.service" id="plexer">
+            <next id="shepherd">^/Shepherd$</next>
+        </Plexer>
+        <Service name="pythonservice" id="shepherd">
+            <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
+            <ServiceID>https://storage3:60010/Shepherd</ServiceID>
+            <CheckPeriod>20</CheckPeriod>
+            <MinCheckInterval>0.1</MinCheckInterval>
+            <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
+            <StoreCfg><DataDir>/var/spool/arc/shepherd_data3</DataDir></StoreCfg>
+            <BackendClass>storage.shepherd.hardlinkingbackend.HopiBackend</BackendClass>
+            <BackendCfg>
+                <DataDir>/var/spool/arc/shepherd_store3</DataDir>
+                <TransferDir>/var/spool/arc/shepherd_transfer3</TransferDir>
+                <TURLPrefix>http://storage3:60011/hopi/</TURLPrefix>
+            </BackendCfg>
+            <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
+            <BartenderURL>https://storage:60000/Bartender</BartenderURL>
+            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
+        </Service>
+    </Chain>
+    <Chain>
+        <Component name="tcp.service" id="tcp2">
+            <next id="http2"/> 
+            <tcp:Listen>
+		<tcp:Port>60011</tcp:Port>
+		<tcp:Version>4</tcp:Version>
+	    </tcp:Listen>
+        </Component>
+        <Component name="http.service" id="http2">
+            <next id="plexer2">GET</next>
+            <next id="plexer2">PUT</next>
+        </Component>
+        <Plexer name="plexer.service" id="plexer2">
+            <next id="hopi">^/hopi/</next>
+        </Plexer>
+        <Service name="hopi" id="hopi">
+            <DocumentRoot>/var/spool/arc/shepherd_transfer3</DocumentRoot>
+            <SlaveMode>1</SlaveMode>
+        </Service>
+    </Chain>
+</ArcConfig>    
+\end{verbatim}
+
+This configuration has two separate ports, one with TLS security and a Shepherd service, and the other without TLS security and with the Hopi service. Let's run this server as well:
+
+\begin{verbatim}
+$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service_3.xml -f
+\end{verbatim}
+
+Now when checking our file again there should be no change at all:
+
+\begin{verbatim}
+$ arc_storage_cli stat /orange
+  locations
+    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
+    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
+\end{verbatim}
+
+Let's stop our server on \verb!storage2!, and watch what happens of our file:
+
+\begin{verbatim}
+$ arc_storage_cli stat /orange
+  locations
+    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
+    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline
+
+$ arc_storage_cli stat /orange
+  locations
+    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
+    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline
+    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: creating
+
+$ arc_storage_cli stat /orange
+  locations
+    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
+    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline
+    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
+\end{verbatim}
+
+A new replica have been created on the third server to maintain the needed number of 2.
+
+Let's start the services on \verb!storage2! again, and let's see what is happening with our file:
+
+\begin{verbatim}
+$ arc_storage_cli stat /orange
+  locations
+    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
+    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
+    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
+
+$ arc_storage_cli stat /orange
+  locations
+    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
+    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: thirdwheel
+    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
+
+$ arc_storage_cli stat /orange
+  locations
+    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
+    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
+\end{verbatim}
+
+One of the replicas have been disappeared to maintain the needed number of 2.
+
 \end{document}

### 2009-03-21 doc/storage_system/arc-storage-documentation.tex
some sentences about gateway module

@@
 
-    TODO: makeMountpoint, unmakeMountpoint
-
     \item[stat(statRequestList)] returns all metadata of files, collections or mount points
@@
 
-    This interface is use to pass the user's limited credentials to the system in order to access the \emph{third-party} store for methods like \emph{get}, \emph{put} and \emph{list}.
-For listing files and directories, \emph{gateway} uses the user credential, whereas in the \emph{get} and \emph{put} methods system only uses the credentials to check the validity of the URLs before sending it to the user.
-
     From the client side \emph{createDelegation()} is called which interns initiate the \emph{CredentialDelegationInitRequest} for the service.
-After the successful handshake this method initiates an other request of \emph{UpdateCredentials( )}, which completes this process.
+After the successful handshake this method initiates an other request of \emph{UpdateCredentials(..)}, which completes this process.
+For complete details see ARC-Security-Document.
 Response from this interface could be \emph{done}, \emph{cannot delegate credentials, requested bartender does not supports the gateway}
@@
 
-TODO: The Bartender can utilize gateway modules for accessing file listings and transfer URL on third-party storages.
-Curretly we have one implementation of a gateway module [...] A gateway module should provide these methods: [...]
-
-TODO: a few sentences about the proxy delegation.
-
+The Bartender can utilize gateway modules for accessing file listings and transfer URL from third-party storages.
+It works as an optional backend component of the Bartender service.
+\emph{Gateway} module is based on four basic methods \emph{get}, \emph{put}, \emph{list} and \emph{remove}.
+Since the task is to contact external store, which requires user to delegate the credentials so that it can communicate on behelf of the user.
+
+In order to access the \emph{files} and \emph{directories} from the external storage system, limited user credentials are passed to the \emph{Bartender} service.
+Users can also remove their \emph{credentials} from the service side.
+These credentials are used in the methods like \emph{get}, \emph{put}, \emph{remove} and \emph{list} of the \emph{Gateway} module.
+These limited credentials are not used in any type of transfer, while listing files/directories, \emph{Gateway} uses the user credential, whereas in the \emph{get} and \emph{put} methods, it only uses the credentials to check the validity of the URL(s) before sending it to the user.
 % subsection implementation (end)

### 2009-03-21 doc/storage_system/arc-storage-documentation.tex
Added description of fuse module and examples for fuse module and arcdmc

@@
 
-TODO: what is the FUSE module, how does it work, how can we use it
+The \textbf{FUSE module} provides a high-level access to the storage system.
+Filesystem in Userspace (FUSE) provides a simple library and a kernel-userspace interface.
+Using FUSE and the ARC Python interface, the FUSE module allows users to mount the storage namespace into the local namespace, enabling the use of, e.g., graphical file browsers.
+Currently no delegation handling is implemented in the fuse module, so that to change ownership, group access, etc., you need to use the \verb!arc_storage_cli!.
+
+As for the \verb!arc_storage_cli!
+the credentials of the user and the URL of the Bartender can be configured either with an XML file located at \verb!~/.arc/client.xml!
+or with environment variables.
+See Section \ref{sub:prototype_cli_tool} for user configuration.
+
+An example of uploading, downloading and list files with FUSE:
+
+\begin{verbatim}
+$ python arcfs.py ./mnt
+$ cat testfile
+This is a test file
+$ mkdir mnt/tmp
+$ cp testfile mnt/tmp
+$ ls -l mnt/tmp/testfile
+-rw-r--r-- 1 jonkni jonkni 20 2009-03-21 23:27 mnt/tmp/testfile
+$ cp mnt/tmp/testfile newfile
+$ cat newfile
+This is a test file
+$ cat mnt/tmp/testfile
+This is a test file
+$ fusermount -u ./mnt
+\end{verbatim}
 
@@
 
-TODO: Some example of uploading, downloading and stat files, anything more?
+Example of uploading, downloading and list files:
+
+\begin{verbatim}
+$ cat testfile
+This is a test file
+$ arccp testfile arc:///tmp/testfile
+$ arcls -l arc:///tmp/testfile
+testfile file * * * *
+$ arccp arc:///tmp/testfile newfile
+$ cat newfile
+This is a test file
+\end{verbatim}
 

### 2009-03-22 doc/ws-programming-tutorial/doc/tex_wsdl/wsdl.tex
Increased verbosity of WSDL introduction, not yet integrated with remaining document.

\subsection{How to prepare a WSDL document}

Firstly, it should be said that the preparation of WSDL documents 
is nothing that is ultimate fun to do. And it is tricky in the sense that
one does mistakes easily that are difficult to spot.

The community hence started to use computational tools to draft these documents.
The most prominent of these is probably Eclipse.
Since version 3.4 (you might need to install it directly from
\url{http://www.eclipse.org} 
if your distribution is a bit behind or if you are working with Windows)
it offers a web tools platform (WTP) module with which WSDL files can be crafted
graphically.  To install it, point the url http://download.eclipse.org/webtools/updates/
to the Eclipse updater, if it is not already listed.

\subsubsection{Preparation of the Echo WSDL file with Eclipse}

The application starts with a minimal setup, which basically is 
much like the echo web service already. We basically only need to add the
documentation of the service and name the parameters. To do so, just click on
the items displayed in the graphical representation and/or find respective
entries in the tabs below.

\begin{figure}
\includegraphics[width=.4\columnwidth]{wsdl_NewFileMenu.png}
\caption{Eclipse dialog to create a new file. WSDL is selected.}
\end{figure}

\begin{figure}
\includegraphics[width=.4\columnwidth]{wsdl_NewWSDLfile.png}
\caption{The default settings are fine for a WSDL service that communicates
with ARC.}
\end{figure}

\begin{figure}
\includegraphics[width=.4\columnwidth]{wsdl_defaultSettings.png}
\caption{The project first shown has already a strong similarity to the Echo
web service that is planned for.}
\end{figure}

\begin{figure}
\includegraphics[width=.4\columnwidth]{wsdl_defaultProject.png}
\caption{A first usable web service description is achieved solely by
specifying the address and passing the arguments the known type
\textit{string}.}
\end{figure}

\begin{figure}
\includegraphics[width=.4\columnwidth]{wsdl_EchoSimplest.png}
\caption{The WSDL for echo in a very simplistic manner.}
\label{fig:simpleEchoService}
\end{figure}

One will fine many specifications that are needed by WSDL to glue the different
parts of the WSDL file together. The service references the bindings and the
bindings reference the port and so on. One should not become overly nervous
about it. The regular programmer does not see those internal names. A good
sanity check for the programmer's view on a WSDL file is to load it from within
the workflow suite Taverna\footnote{Taverna is available from
\url{http://taverna.sf.net}. ARC offers a plugin that allows the direct
communication of Taverna with resources runnign ARC-0. An interface to ARC-1
is still pending. In principle, there is none needed. In practice, we still need
more practice to decide.}.

\begin{figure}
\includegraphics[width=.7\columnwidth]{wsdl_echo_taverna_minimal.png}
\caption{Taverna learned about the Echo Web Service and presents it. No
internal names of WSDL appear, only the name of operations and paramters are
shown.}
\label{fig:wsdlEchoTavernaMinimal}
\end{figure}

The \textit{activities} menu in Taverna offers the item \textit{New Activity}.
It will request the specification of an URL or a local path that points to a
WSDL file. The services presented in there will then be added to the list of
WSDL resources. So will the \textit{echo} workflow that was just prepared with
Eclipse. When it is dragged to the canvas in the middle, as shown in figure
\ref{fig:wsdlEchoTavernaMinimal}, it appears as an interactive object. To the
right, the parameters of the operation are displayed.

\begin{figure}
\includegraphics[width=.7\columnwidth]{wsdl_echo_taverna_context.png}
\caption{A completely functional (albeit small) workflow in Taverna that
addresses the Echo Web Service prepared with HED.}
\label{fig:wsdlEchoTavernaContext}
\end{figure}

With a string constant as a constant input added, alternatively one could link a
configurable workflow input as input to the echo Web Service, and the
workflow's output linked as a sink to the output of the workflow, the workflow
is complete. It can now be executed, presuming that the localhost is indeed
running arched with the echo service as described in section
\ref{sec:arch_echo_web_service}.

When comparing the GUI representation of Eclipse as shown with in figure
\ref{fig:simpleEchoService} or that of the just shown Taverna representation
with the raw WSDL of listing \ref{lst:wsdl_eclipse_echo_xml}, it becomes
apparent, that the raw form of the WSDL, as shown below, is more complicated. It seems to invent nested structures that are just not present, since it is only simple strings that are exchanged.
The good news here is, that one would have reached exactly that more
complicated looks when following the original setup without assignment of the
simple type 'string'. WSDL needs special types for every operation. And eclipse
provides them, even when the input types are shown in a more simple manner.

\lstsetECHOECLIPSEXML
\lstinputlisting
	[label=lst:wsdl_eclipse_echo_xml,float=htb,
	caption={[Minimalistic WSDL file for echo created with Eclipse.]
	\textbf{WSDL file for Echo created with Eclipse.}}]
{echo.wsdl}

This has the advantage, that you can use this WSDL file to describe the regular
ARC echo service, even though the original programmer might not have used
Eclipse to create it. Or has he? Gabor?

### 2009-03-22 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
First integration of WSDL description into remainder of document. Michael's scrutiny is still pending.

@@
 
-The next service presented is the Echo Web Service.
+The Echo Web Service was already presented in the introduction, 
+basically as the ground truth to learn if the installation of
+the HED is functional. And it was frequently referenced in the previous chapter
+on WSDL. In this section, we will implement our own version of the Echo
+Web Service.
+
 Its task is to return the content of the received message back to the client.
-Furthermore, the service provides two additional operations called ``ordinary'' and ``reverse''.
-Depending on the operation chosen by the client, the message gets returned unmodified or reversed.
-The goal of this chapter is to offer a deeper knowledge about message processing.
-Also, the \textit{WSDL} (Web Services Description Language) will be introduced, with which the section now starts.
+As a special feature that we introduce solely for educatory purposes, the
+service provides two additional operations called ``ordinary'' and ``reverse''.
+Depending on the operation chosen by the client, the message gets returned
+unmodified or reversed. The goal of this chapter is to deepen the knowledge
+on message processing.
 
-\section{Web Services Description Language}
+The WSDL file that specifies the here presented Echo Web Service is shown in
+Listing~\ref{lst:echo_wsdl}.
 
-Web Services can be specified (not implemented) by a meta language called \textit{WSDL} (Web Services Description Language).
-Beside the description of the service, WSDL renders the possibility that services can be discovered by service brokers.
-The discovery of services may be useful for an automatic adaption of a running client, i.e. if the endpoint of the service changes or to share the load of demands.
-The WSDL service specification enables a foreign programmer to implement a suitable client which interfaces the service in a proper way.
-WSDL describes the structure of the messages, the intended use of the messages, the supported protocols and the endpoint of the service (URL).
-\task{For that last reason, every instance of a service must be produced a distinct WSDL file for - or - it is automatically prepared from the information that the server has available once it is implemented.
-- ??MG }\\
-
-The WSDL file that specifies the here presented Echo Web Service is shown in Listing~\ref{lst:echo_wsdl}. 
-It is composed of five main elements:
-\newcommand{\parboxWidth}{13cm}
-\begin{itemize}
-	\item \textbf{Types} --- The element \textit{types} contains the description of the message structures and is written in the language XSD (XML Schema Definition).
-The data types used for the messages of the Echo Web Service are defined starting with the  line~\ref{lst_code:echo_wsdl_types}.
-
-	\item \textbf{Message} --- Within the element \textit{message} a subset of the previously defined types are assigned to be a message.
-In case of the Echo Web Service  two messages are designated: echoRequest and echoResponse, compare lines line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
-
-	\item \textbf{Port type} --- The element \textit{portType} is used to assign the messages to an interface.
-Four interface types can be distinguished: One-way (input), request-response (input, output), solicit-response (input, output, fault) and notification (output).
-	The Echo Web Service realises a solicit-response interface which is declared subsequent to line~\ref{lst_code:echo_wsdl_portType}. % in the lines followed by the line X
-
-	\item \textbf{Binding} --- The protocol and the data format used for the message transmission are denoted in the element \textit{binding}.
-Possible style attributes are \textit{document} or \textit{rpc} (Remote Procedure Call).
-The transport attribute defines the protocol which is regularly HTTP.
-For each operation a corresponding SOAP action has to be defined which assignes a messages to be either \textit{literal} or \textit{encoded}~\cite{BUTEK_2009}.
-The service implemented in this chapter will use the binding \textit{document}/\textit{literal} and the protocol HTTP which is to be seen in the lines follow line~\ref{lst_code:echo_wsdl_binding}.
-
-% The style attribute can be "rpc" or "document". In this case we use document. The transport attribute defines the SOAP protocol to use. In this case we use HTTP.
-%The operation element defines each operation that the port exposes.
-%For each operation the corresponding SOAP action has to be defined. You must also specify how the input and output are encoded. In this case we use "literal".
-%  http://www.ibm.com/developerworks/webservices/library/ws-whichwsdl/
-%   1. RPC/encoded    (Remote Procedure Call)
-%   2. RPC/literal
-%   3. Document/encoded
-%   4. Document/literal
-% The terminology here is very unfortunate: RPC versus document.
-These terms imply that the RPC style should be used for RPC programming models and that the document style should be used for document or messaging programming models.
-That is not the case at all.
-The style has nothing to do with a programming model.
-It merely dictates how to translate a WSDL binding to a SOAP message.
-Nothing more.
-You can use either style with any programming model.
-%
-% RPC/literal SOAP message for myMethod
-%
-%<soap:envelope>
-%    <soap:body>
-%        <myMethod>
-%            <x>5</x>
-%            <y>5.0</y>
-%        </myMethod>
-%    </soap:body>
-%</soap:envelope>
-%
-%  RPC/encoded SOAP message for myMethod
-% 
-% <soap:envelope>
-%     <soap:body>
-%         <myMethod>
-%             <x xsi:type="xsd:int">5</x>
-%             <y xsi:type="xsd:float">5.0</y>
-%         </myMethod>
-%     </soap:body>
-% </soap:envelope>
-
-	\item \textbf{Service} --- Within the last WSDL element \textit{service} the endpoint of the service gets specified.
-As to be seen in line~\ref{lst_code:echo_wsdl_service} the endpoint of the Echo Web Service  is assigned to \textit{http://localhost:60000/echo}.
-
-\end{itemize}
-%\textcolor{white}{newline}
-\lstsetJUSTXML
-\lstinputlisting
-	[
-	label=lst:echo_wsdl,
-	caption={[WSDL file describing the echo service. Filename: echo.wsdl]
-	\textbf{WSDL file describing the echo service. Filename: echo.wsdl}}
-	]
-{../src/services/echoservice/echo.wsdl}
-\task{TODO: Validate the WSDL Listing}
+% moved WSDL description into separate Chapter. See folder tex_wsdl.
 

### 2009-03-22 doc/ws-programming-tutorial/doc/tex_wsdl/wsdl.tex
First integration of WSDL description into remainder of document. Michael's scrutiny is still pending.

@@
-\subsection{How to prepare a WSDL document}
+\chapter{Web Services Description Language}
 
-Firstly, it should be said that the preparation of WSDL documents 
-is nothing that is ultimate fun to do. And it is tricky in the sense that
-one does mistakes easily that are difficult to spot.
+For Web Services, the mere communication, as for instance we know if from
+traditional CGI programming, is not everything. The communication is performed
+via the SOAP protocol, and there are several implementations of SOAP on top of
+CGI. What the here explained WSDL (Web Services Description Language) is about
+is the \textit{talking} about Web Services. In a now standardised XML-based format,
+one can emply WSDL to describe both technically and for regular users the
+performance of the Web Service.
+
+WSDL may be imagined as a meta language. Beside the
+description of the service, WSDL renders the possibility that services can be
+discovered by service brokers\footnote{The here referenced brokering is meant
+to act on a global registry of services and has little practical meaning. It
+is not to be confused with the brokering of grid resources, for which ARC-1
+provides a web service on its own, which is a key functionality of any
+computational grid.}. The discovery of services may be useful for an automatic
+adaption of a running client, i.e.
+if the endpoint of the service changes or to share the load of demands.
+The WSDL service specification enables a foreign programmer to implement a suitable client which interfaces the service in a proper way.
+WSDL describes the structure of the messages, the intended use of the messages, the supported protocols and the endpoint
+of the service (URL).
+\task{For that last reason, every instance of a service must be produced a
+distinct WSDL file for - or - it is automatically prepared from the information
+that the server has available once it is implemented. - ??MG }\\
+
+WSDL documents are composed of five main elements:
+\newcommand{\parboxWidth}{13cm}
+\begin{itemize}
+	\item \textbf{Types} --- The element \textit{types} contains the description of the message structures and is written in the language XSD (XML Schema Definition).
+The data types used for the messages of the Echo Web Service are defined starting with the  line~\ref{lst_code:echo_wsdl_types}.
+
+	\item \textbf{Message} --- Within the element \textit{message} a subset of the previously defined types are assigned to be a message.
+In case of the Echo Web Service  two messages are designated: echoRequest and echoResponse, compare lines line~\ref{lst_code:echo_wsdl_message1} and \ref{lst_code:echo_wsdl_message2}.
+
+	\item \textbf{Port type} --- The element \textit{portType} is used to assign the messages to an interface.
+Four interface types can be distinguished: One-way (input), request-response (input, output), solicit-response (input, output, fault) and notification (output).
+	The Echo Web Service realises a solicit-response interface which is declared subsequent to line~\ref{lst_code:echo_wsdl_portType}. % in the lines followed by the line X
+
+	\item \textbf{Binding} --- The protocol and the data format used for the message transmission are denoted in the element \textit{binding}.
+Possible style attributes are \textit{document} or \textit{rpc} (Remote Procedure Call).
+The transport attribute defines the protocol which is regularly HTTP.
+For each operation a corresponding SOAP action has to be defined which assignes a messages to be either \textit{literal} or \textit{encoded}~\cite{BUTEK_2009}.
+The service implemented in this chapter will use the binding \textit{document}/\textit{literal} and the protocol HTTP which is to be seen in the lines follow line~\ref{lst_code:echo_wsdl_binding}.
+
+% The style attribute can be "rpc" or "document". In this case we use document. The transport attribute defines the SOAP protocol to use. In this case we use HTTP.
+%The operation element defines each operation that the port exposes.
+%For each operation the corresponding SOAP action has to be defined. You must also specify how the input and output are encoded. In this case we use "literal".
+%  http://www.ibm.com/developerworks/webservices/library/ws-whichwsdl/
+%   1. RPC/encoded    (Remote Procedure Call)
+%   2. RPC/literal
+%   3. Document/encoded
+%   4. Document/literal
+% The terminology here is very unfortunate: RPC versus document.
+These terms imply that the RPC style should be used for RPC programming models and that the document style should be used for document or messaging programming models.
+That is not the case at all.
+The style has nothing to do with a programming model.
+It merely dictates how to translate a WSDL binding to a SOAP message.
+Nothing more.
+You can use either style with any programming model.
+%
+% RPC/literal SOAP message for myMethod
+%
+%<soap:envelope>
+%    <soap:body>
+%        <myMethod>
+%            <x>5</x>
+%            <y>5.0</y>
+%        </myMethod>
+%    </soap:body>
+%</soap:envelope>
+%
+%  RPC/encoded SOAP message for myMethod
+% 
+% <soap:envelope>
+%     <soap:body>
+%         <myMethod>
+%             <x xsi:type="xsd:int">5</x>
+%             <y xsi:type="xsd:float">5.0</y>
+%         </myMethod>
+%     </soap:body>
+% </soap:envelope>
+
+	\item \textbf{Service} --- Within the last WSDL element \textit{service} the
+	endpoint of the service gets specified. As to be seen in
+	line~\ref{lst_code:echo_wsdl_service} the endpoint of the Echo Web Service is
+	assigned to \textit{http://localhost:60000/echo}.
+\end{itemize}
+
+To understand an WSDL document, one best starts from the Service entry,
+commonly located at the end. It will describe what the service is all about.
+From there, move upwards to learn about the operations and their parameters.
+Most items in the WSDL document have their own tags that provide names or
+descriptions.
+%\textcolor{white}{newline}
+\lstsetJUSTXML
+\lstinputlisting
+	[
+	label=lst:echo_wsdl,
+	caption={[WSDL file describing the echo service. Filename: echo.wsdl]
+	\textbf{WSDL file describing the echo service. Filename: echo.wsdl}}
+	]
+{../src/services/echoservice/echo.wsdl}
+\task{TODO: Validate the WSDL Listing}
+
+\section{How to prepare a WSDL document}
+
+It should be said that the preparation of WSDL documents not ultimate fun to
+do. And it is tricky in the sense that one does mistakes easily that are
+difficult to spot.
 
 The community hence started to use computational tools to draft these documents.
-The most prominent of these is probably Eclipse.
-Since version 3.4 (you might need to install it directly from
-\url{http://www.eclipse.org} 
-if your distribution is a bit behind or if you are working with Windows)
-it offers a web tools platform (WTP) module with which WSDL files can be crafted
-graphically.  To install it, point the url http://download.eclipse.org/webtools/updates/
-to the Eclipse updater, if it is not already listed.
+The most prominent of these is probably Eclipse. Since its version 3.4 (you
+might need to install it directly from \url{http://www.eclipse.org} if your
+distribution is a bit behind or if you are working with Windows) it offers a
+web tools platform (WTP) module with which WSDL files can be crafted graphically.
+To install it, point the url http://download.eclipse.org/webtools/updates/ to
+the Eclipse updater, if it is not already listed.
 
-\subsubsection{Preparation of the Echo WSDL file with Eclipse}
-
-The application starts with a minimal setup, which basically is 
-much like the echo web service already. We basically only need to add the
-documentation of the service and name the parameters. To do so, just click on
-the items displayed in the graphical representation and/or find respective
-entries in the tabs below.
+\subsection{Preparation of the Echo WSDL file with Eclipse}
 
 \begin{figure}
-\includegraphics[width=.4\columnwidth]{wsdl_NewFileMenu.png}
+\includegraphics[width=.4\columnwidth]{tex_wsdl/wsdl_NewFileMenu.png}
 \caption{Eclipse dialog to create a new file. WSDL is selected.}
+\label{fig:wsdl_newFileMenu}
 \end{figure}
 
+To prepare a WSDL document with Eclipse, one should find the right file type in
+the typical Eclipse dialog window. This is shown in figure
+\ref{fig:wsdl_newFileMenu}. The empty WSDL document is not empty, but the
+application starts with a minimal setup: a service that expects some input and
+provides some immediate return value. This is already much like the echo web
+service. One basically only needs to add the documentation of the service and
+to name the parameters. To do so, just click on the items displayed in the
+graphical representation and/or find respective entries in the tabs below.
+
 \begin{figure}
-\includegraphics[width=.4\columnwidth]{wsdl_NewWSDLfile.png}
+\includegraphics[width=.4\columnwidth]{tex_wsdl/wsdl_NewWSDLfile.png}
 \caption{The default settings are fine for a WSDL service that communicates
@@
 \begin{figure}
-\includegraphics[width=.4\columnwidth]{wsdl_defaultSettings.png}
+\includegraphics[width=.4\columnwidth]{tex_wsdl/wsdl_defaultSettings.png}
 \caption{The project first shown has already a strong similarity to the Echo
@@
 \begin{figure}
-\includegraphics[width=.4\columnwidth]{wsdl_defaultProject.png}
+\includegraphics[width=.4\columnwidth]{tex_wsdl/wsdl_defaultProject.png}
 \caption{A first usable web service description is achieved solely by
@@
 \begin{figure}
-\includegraphics[width=.4\columnwidth]{wsdl_EchoSimplest.png}
+\includegraphics[width=.4\columnwidth]{tex_wsdl/wsdl_EchoSimplest.png}
 \caption{The WSDL for echo in a very simplistic manner.}
@@
 \end{figure}
+
+\subsection{Inspection of Web Services with Taverna}
 
@@
 \begin{figure}
-\includegraphics[width=.7\columnwidth]{wsdl_echo_taverna_minimal.png}
+\includegraphics[width=.7\columnwidth]{tex_wsdl/wsdl_echo_taverna_minimal.png}
 \caption{Taverna learned about the Echo Web Service and presents it. No
@@
 \begin{figure}
-\includegraphics[width=.7\columnwidth]{wsdl_echo_taverna_context.png}
+\includegraphics[width=.7\columnwidth]{tex_wsdl/wsdl_echo_taverna_context.png}
 \caption{A completely functional (albeit small) workflow in Taverna that
@@
 	\textbf{WSDL file for Echo created with Eclipse.}}]
-{echo.wsdl}
+{tex_wsdl/echo.wsdl}
 
@@
 Eclipse to create it. Or has he? Gabor?
+
+\section{What else to do with WSDL}
+
+There are several efforts to organise the world of web services and their
+interactions. Taverna was already mentioned above. Amongst the most advanced
+scientific community with respect to Web Services are the Bioinformaticians.
+Topics that come to mind are:
+
+\begin{itemize}
+  \item BioMoby \url{www.biomoby.org}
+  \item $^my$Grid and $^my$Experiment \url{www.myexperiment.org}
+  \item Distributed Annotation System (DAS) \url{http://www.biodas.org} 
+\end{itemize}
+
+But this is left for a future version of this tutorial\ldots and/or possibly
+Hajo's Bachelor thesis.

### 2009-03-22 doc/storage_system/arc-storage-documentation.tex
documentation about the STALLED state

@@
 A location contains the URL of the Shepherd service, the local ID of the file within the storage node and the state of the replica.
-This state could be `\textbf{alive}' (if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the Shepherd claims it has no such file), `\textbf{offline}' (if the Shepherd is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{thirdwheel}' (if the replica is marked for deletion because of too many replicas).
+This state could be `\textbf{alive}' (if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the Shepherd claims it has no such file), `\textbf{offline}' (if the Shepherd is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{thirdwheel}' (if the replica is marked for deletion because of too many replicas), `\textbf{stalled}' (if the the replica was in the state if uploading for a long time but hasn't arrived).
 We also store (in the \emph{states} section) the size of the file, the number of needed replicas, and a checksum of the file, which is created by some kind of checksumming algorithm (currently only md5 is supported).
@@
     The \emph{filelist} is a list of (\emph{GUID}, \emph{referenceID}, \emph{state}) tuples containing the state of changed or new files, where referenceID is the Sheperd-local ID of the given replica where GUID refers to the logical file of this replica.
-The \emph{state} could be `\textbf{invalid}' (if the periodic self-check of the Shepherd found a non-matching checksum or missing file), `\textbf{creating}' (if this is a new file not uploaded yet) or `\textbf{alive}' (if the file is uploaded and the checksum is OK).
+The \emph{state} could be `\textbf{invalid}' (if the periodic self-check of the Shepherd found a non-matching checksum or missing file), `\textbf{creating}' (if this is a new file not uploaded yet), `\textbf{thirdwheel}' (if the replica is marked for deletion because of too many replicas), `\textbf{alive}' (if the file is uploaded and the checksum is OK), or `\textbf{stalled}' (if the file was in `creating' state for a long time but hasn't arrived).
 
@@
 
-The replicas have a state, which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, or `\textbf{invalid}' if it does not exists anymore or has a bad checksum.
+The replicas have a state, which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, `\textbf{invalid}' if it does not exists anymore or has a bad checksum, `\textbf{thirdwheel}' if the replica is marked for deletion because the file has too many replicas, or `\textbf{stalled}' if it was in a `creating' state for a long time but hasn't arrived.
 
@@
 After checking the cheksum of a file, the Shepherd will always wait this long before checking the next file, even if this means that checking all the files takes longer than the `CheckPeriod'.
+    \item[CreatingTimeout] specifies in seconds after how much time the Shepherd decideds that a started file upload won't happen.
+This is not fatal, if the file is uploaded eventually, it will be a valid replica, but after this much time the system starts creating new replicas just in case.
     \item[StoreClass] specifies which type of store the Shepherd should use for storing its metadata on disk.
@@
         <MinCheckInterval>0.1</MinCheckInterval>
+        <CreatingTimeout>600</CreatingTimeout>
         <StoreClass>storage.store.cachedpicklestore.CachedPickleStore</StoreClass>

### 2009-03-23 doc/a-rex/arex_tech_doc.tex
Fixing joboption names
Adding MigrateActivity documentation

@@
 does not work in current implementation).
+\item \textit{activityid} -- Job-id of previous job in case the job has been 
+resubmitted or migrated. This value can appear multiple times if a job has been 
+resubmitted or migrate more than once.
+\item \textit{forcemigration} -- This boolean is only used for migration of jobs.
+It determines whether the job should persists if the termination of the previous 
+job fails.
 \end{itemize}
@@
 
-A-REX introduces a new operation in addition to those provided by
-BES. It does that by defining its own port-type with the single operation
-\emph{ChangeActivityStatus} (see Appendix~\ref{annex:arex-wsdl}). This operation provides
-a way to request simple transfers between states of jobs and corresponding
+A-REX introduces two new operations in addition to those provided by
+BES. It does that by defining its own port-type with new operations
+\emph{ChangeActivityStatus} and \emph{MigrateActivity}(see 
+Appendix~\ref{annex:arex-wsdl}).
+
+The \emph{ChangeActivityStatus} operation provides a way to request 
+simple transfers between states of jobs and corresponding
 actions.
@@
 \end{itemize}
-\end{itemize}
 
@@
 state. If such a procedure cannot be performed immediately then the
-corresponding sequence is initiated and fault OperationWillBeAppliedEventuallyFault
-will be returned.
+corresponding sequence is initiated and fault 
+OperationWillBeAppliedEventuallyFault will be returned.
 
@@
 described in Section~\ref{Section:Job Flow}.
+
+The \emph{MigrateActivity} operation generates a request to migrate a 
+grid job from another ARC1 cluster, i.e. the operation will get input 
+files and possibly job description from the cluster currently holding 
+the job and create the job as a new activity at the present cluster.
+Currently only migration of queuing jobs is supported.
+
+\item \emph{MigrateActivity}
+\begin{itemize}
+\item Input
+\begin{itemize}
+\item \emph{wsa:EndpointReferenceType ActivityIdentifier}: This 
+element should contain the \emph{wsa:EndpointReference} of the job to 
+be migrated.
+\item \emph{ActivityDocument}: JSDL document of the job to be 
+migrated. This element is optional.
+\item \emph{Boolean ForceMigration}: Boolean that determines whether 
+the job will persist on the new cluster if the termination of the 
+previous job fails.
+\end{itemize}
+\item Output
+\begin{itemize}
+\item \emph{wsa:EndpointReferenceType ActivityIdentifier}: This 
+element should contain the \emph{wsa:EndpointReference} of the new 
+activity.
+\item \emph{ActivityDocument}: Contains the JSDL document of the new 
+activity.
+\end{itemize}
+\item Fault(s)
+\begin{itemize}
+\item \emph{NotAuthorizedFault}: Indicates that the client is not allowed
+to do this operation.
+\item \emph{NotAcceptingNewActivitiesFault}: A fault that indicates that A-REX 
+currently is not accepting new activities.
+\item \emph{UnsupportedFeatureFault}: This fault indicates that an sub-element in the 
+JDSL document is not supported or the ActivityDocument has not been recognised as JSDL. 
+\item \emph{InvalidRequestMessageFault}: This fault indicates that an element in the 
+request is either missing or has an invalid format. Typically this would mean that 
+the job-id cannot be located in the ActivityIdentifier of the old job.
+\end{itemize}
+\end{itemize}
+\end{itemize}
+
+The \emph{ActiviterIdentifier} specifies the URL of the job which will be migrated. In case the 
+\emph{ActivityDocument} is filled this document will be used to create a new activity otherwise 
+an attempt will be made to retrieve the job description through the BES operation
+\emph{GetActivityDocument}.
+
+Once the input files have been downloaded from the other cluster, a request will be send 
+to terminate the old job. If this request fails the new activity at the present cluster 
+will be terminate unless the \emph{ForceMigration} is true. This is to prevent the job 
+from being executed at two different places at the same time.
 
@@
 of a computing resource.
+\item [{\emph{OldJobID}}] -- specifies the previous job-ids in case the job 
+has been resubmitted or migrated.
 \end{list}
@@
 \item joboption\_stderr -- same for stderr.
-\item joboption\_maxcputime -- amount of CPU time requested (minutes).
-\item joboption\_maxmemory -- amount of memory requested (megabytes).
+\item joboption\_cputime -- amount of CPU time requested (minutes).
+\item joboption\_memory -- amount of memory requested (megabytes).
 \item joboption\_count -- number of processors requested.
@@
                    type=\char`\"{}a-rex:ChangeActivityStatusResponseType\char`\"{}/>
+      <xsd:complexType name="MigrateActivityType">
+        <xsd:sequence>
+	  <xsd:element name=\char`\"{}ActivityIdentifier\char`\"{} 
+	  	           type=\char`\"{}wsa:EndpointReferenceType\char`\"{} 
+	  	           minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}1\char`\"{}/>
+          <xsd:element ref=\char`\"{}bes-factory:ActivityDocument\char`\"{} minOccurs=\char`\"{}0\char`\"{}/>
+	  <xsd:element name=\char`\"{}ForceMigration\char`\"{} type=\char`\"{}xsd:boolean\char`\"{} minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}1\char`\"{}/>    
+          <xsd:any namespace=\char`\"{}##other\char`\"{} processContents=\char`\"{}lax\char`\"{} minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>        
+        </xsd:sequence>
+        <xsd:anyAttribute namespace=\char`\"{}##other\char`\"{} processContents=\char`\"{}lax\char`\"{}/>
+      </xsd:complexType>
+
+      <xsd:complexType name=\char`\"{}MigrateActivityResponseType\char`\"{}>
+        <xsd:sequence>
+          <xsd:element name=\char`\"{}ActivityIdentifier\char`\"{} type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
+          <xsd:element ref=\char`\"{}bes-factory:ActivityDocument\char`\"{} minOccurs=\char`\"{}0\char`\"{}/>
+          <xsd:any namespace=\char`\"{}##other\char`\"{} processContents=\char`\"{}lax\char`\"{} minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>        
+        </xsd:sequence>
+        <xsd:anyAttribute namespace=\char`\"{}##other\char`\"{} processContents=\char`\"{}lax\char`\"{}/>
+      </xsd:complexType>
+
+      <xsd:element name=\char`\"{}MigrateActivity\char`\"{}
+        type=\char`\"{}a-rex:MigrateActivityType\char`\"{}/>
+      <xsd:element name=\char`\"{}MigrateActivityResponse\char`\"{}
+        type=\char`\"{}a-rex:MigrateActivityResponseType\char`\"{}/>
+
     </xsd:schema>
@@
   </wsdl:message>
+
+  <wsdl:message name=\char`\"{}MigrateActivityRequest\char`\"{}>
+    <wsdl:part name=\char`\"{}MigrateActivityRequest\char`\"{} 
+	element=\char`\"{}a-rex:MigrateActivity\char`\"{}/>
+  </wsdl:message>
+
+  <wsdl:message name=\char`\"{}MigrateActivityResponse\char`\"{}>
+    <wsdl:part name=\char`\"{}MigrateActivityResponse\char`\"{} 
+        element=\char`\"{}a-rex:MigrateActivityResponse\char`\"{}/>
+  </wsdl:message>
+
   <wsdl:portType name=\char`\"{}a-rex\char`\"{}>
@@
     </wsdl:operation>
+    <wsdl:operation name=\char`\"{}MigrateActivity\char`\"{}>
+      <wsdl:input 
+         name=\char`\"{}MigrateActivity\char`\"{}
+         message=\char`\"{}a-rex:MigrateActivityRequest\char`\"{}  
+      <wsdl:output
+         name=\char`\"{}MigrateActivityResponse\char`\"{}
+         message=\char`\"{}a-rex:MigrateActivityResponse\char`\"{}
+      <wsdl:fault name=\char`\"{}NotAuthorizedFault\char`\"{} 
+         message=\char`\"{}bes-factory:NotAuthorizedFault\char`\"{}
+        wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}NotAcceptingNewActivitiesFault\char`\"{} 
+         message=\char`\"{}bes-factory:NotAcceptingNewActivitiesFault\char`\"{}
+         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}UnsupportedFeatureFault\char`\"{} 
+         message=\char`\"{}bes-factory:UnsupportedFeatureFault\char`\"{}
+         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+      <wsdl:fault name=\char`\"{}InvalidRequestMessageFault\char`\"{}
+         message=\char`\"{}bes-factory:InvalidRequestMessageFault\char`\"{}
+         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+    </wsdl:operation>
   </wsdl:portType>
@@
         <soap:fault name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+    </wsdl:operation>
+    <wsdl:operation name=\char`\"{}MigrateActivity\char`\"{}>
+      <soap:operation soapAction=\char`\"{}MigrateActivity\char`\"{} /> 
+      <wsdl:input name=\char`\"{}MigrateActivity\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{} />
+      </wsdl:input>
+      <wsdl:output name=\char`\"{}MigrateActivityResponse\char`\"{}>
+        <soap:body use=\char`\"{}literal\char`\"{} />
+      </wsdl:output>
+      <wsdl:fault name=\char`\"{}NotAuthorizedFault\char`\"{}>
+        <soap:fault name=\char`\"{}NotAuthorizedFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}NotAcceptingNewActivitiesFault\char`\"{}>
+        <soap:fault name=\char`\"{}NotAcceptingNewActivitiesFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}UnsupportedFeatureFault\char`\"{}>
+        <soap:fault name=\char`\"{}UnsupportedFeatureFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      </wsdl:fault>
+      <wsdl:fault name=\char`\"{}InvalidRequestMessageFault\char`\"{}>
+        <soap:fault name=\char`\"{}InvalidRequestMessageFault\char`\"{} use=\char`\"{}literal\char`\"{} />
       </wsdl:fault>

### 2009-03-23 doc/storage_system/arc-storage-manual.tex
Added fuse howto

@@
 
+Besides the \verb!arc_storage_cli! tool we can access the storage system with the use of a FUSE modul. We can install the ARC storage FUSE modul following these steps:
+
+\begin{itemize}
+\item Install ARC1 storage system.
+
+\item Install \verb!fuse >= 2.7.3! and \verb!fuse-python >= 0.2!:
+\begin{verbatim}
+$ sudo aptitude install fuse-python
+\end{verbatim}
+
+\item Set up \verb!~/.arc/client.xml!:
+
+\begin{verbatim}
+$ cat ~/.arc/client.xml 
+<ArcConfig>
+  <KeyPath>/home/<username>/.arc/userkey.pem</KeyPath>
+  <CertificatePath>/home/<username>/.arc/usercert.pem</CertificatePath>
+  <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+  <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
+</ArcConfig>
+\end{verbatim}
+
+\item Create a mountpoint and mount arcfs.py:
+\begin{verbatim}
+   $ pwd
+   /home/me/arc
+   $ ln -s arc1/src/services/storage/fuse/arcfs.py arcfs.py
+   $ mkdir mnt
+   $ python arcfs.py ./mnt 
+\end{verbatim}
+This mounts ARCFS in \verb!/home/me/arc/mnt!, creates directory
+\verb!/home/me/arc/fuse_transfer! and a log file
+\verb!/home/me/arc/arcfsmessages!.
+
+\item Make a collection:
+\begin{verbatim}   
+   $ mkdir -p mnt/home/me
+   $ stat mnt/home/me
+     File: `mnt/home/me'
+     Size: 0               Blocks: 0          IO Block: 4096   directory
+   Device: 19h/25d Inode: 3           Links: 2
+   Access: (0755/drwxr-xr-x)  Uid: (  500/  me)   Gid: (  100/   users)
+   Access: 1970-01-01 01:00:00.000000000 +0100
+   Modify: 1970-01-01 01:00:00.000000000 +0100
+   Change: 1970-01-01 01:00:00.000000000 +0100
+\end{verbatim}
+Note that so far there is no security handling implemented, so
+that all files have mode 0644, all collections have mode 0755,
+everything is owned by the user that mounts the system.
+
+\item Create some entries in your collection:
+\begin{verbatim}
+   $ cd mnt/home/me
+   $ emacs -nw fish
+   catfish
+   $ cat fish
+   catfish
+   $ cp fish fisk2
+   $ ls -la
+   drwxr-xr-x 5 me users  0 1970-01-01 01:00 .
+   drwxr-xr-x 3 me users  0 1970-01-01 01:00 ..
+   -rw-r--r-- 1 me users  8 1970-01-01 01:00 fish
+   -rw-r--r-- 1 me users 28 1970-01-01 01:00 fish~
+   -rw-r--r-- 1 me users  8 1970-01-01 01:00 fish2
+\end{verbatim}
+
+\item You probably want to remove that anoying emacs backup file
+\begin{verbatim}   
+   $ rm fish~
+   $ ls 
+   fish  fish2
+\end{verbatim}
+
+\item Maybe you want fish2 in a separate collection
+\begin{verbatim}   
+   $ mkdir sea_creatures
+   $ mv fish2 sea_creatures
+\end{verbatim}
+
+\item Maybe you're not happy with the collection name
+\begin{verbatim}
+   $ mv sea_creatures creatures
+   $ ls creatures
+   fish2
+\end{verbatim}
+
+\item Moving the collection out of ARC storage:
+\begin{verbatim}
+   $ cd ../../..
+   $ mv mnt/home/me/creatures .
+\end{verbatim}
+
+\item The collection is no longer in the ARC storage:
+\begin{verbatim}
+   $ find mnt/home/me/
+   mnt/home/me/
+   mnt/home/me/fish
+\end{verbatim}
+
+\item Unmounting ARCFS; To unmount, refer to fuse documentation. Usually you can do
+\begin{verbatim}   
+   $ fusermount -u mnt
+\end{verbatim}
+
+If this for some reason this doesn't work, you can do
+\begin{verbatim}   
+   $ sudo umount -f mnt
+\end{verbatim}
+
+or even
+\begin{verbatim}   
+   $ sudo umount -l mnt
+\end{verbatim}
+
+Now, remounting the system, everything should still be there:
+\begin{verbatim}
+   $ python arcfs.py ./mnt
+
+   $ find mnt/home/me/
+   mnt/home/me/
+   mnt/home/me/fish
+
+   $ fusermount -u mnt
+\end{verbatim}
+\end{itemize}
+
 \end{document}

### 2009-03-23 doc/storage_system/arc-storage-documentation.tex
some updates

@@
 %%%%---- The title ----
-{\centering \textsc{\Large Documentation of the ARC storage system}\Large \par}
+{\centering \textsc{\Large Chelonia - Self-healing distributed storage system}\Large \par}
 \vspace*{0.5cm}
@@
 %%%%---- A subtitle, if necessary ----
-{\centering \textit{\large First prototype status and plans}\large \par}
+{\centering \textit{\large Documentation and developer's guide}\large \par}
 
@@
 \begin{description}
-    \item[ReplicatedAHash] %TODO: describe ReplicatedAHash, sendMessage, processMessage
+    \item[ReplicatedAHash] 
 	is a subclass of the \verb!CentralizedAHash!
@@
 and \verb!ReplicationManager!, meaning that the \verb!ReplicatedAHash!, with the exception of the above mentioned methods, is actually just a centralized A-Hash with a specific storage module.
-    \item[ReplicationStore] %TODO: describe ReplicationStore, connection with TransDBStore, config, etc.
+    \item[ReplicationStore] 
 	is a specialized replicated DB.
@@
 object which governs the replication framework.
-    \item[ReplicationManager] %TODO: describe ReplicationManager, elections, communication (interservice communication, security, callback functions)
+    \item[ReplicationManager] 
 	handles all aspects of the Berkeley DB API and takes care of replicas, elections and replication messages.
@@
 
-%TODO: here comes some sentences about how the Librarian reads from any A-Hash and finds the master A-Hash for writing.
 The replicated A-Hash deployment is based on a single master, multiple clients scenario (see Section~\ref{sec:a_hash} for details).

### 2009-03-23 doc/storage_system/arc-storage-manual.tex
some updates

@@
     <MinCheckInterval>0.1</MinCheckInterval>
+    <CreatingTimeout>600</CreatingTimeout>
     <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
@@
             <MinCheckInterval>0.1</MinCheckInterval>
+            <CreatingTimeout>600</CreatingTimeout>
             <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
             <StoreCfg>
-                <DataDir>/var/spool/arc/shepherd_store2</DataDir>
+                <DataDir>/var/spool/arc/shepherd_data2</DataDir>
             </StoreCfg>
@@
             <BackendCfg>
-                <DataDir>/var/spool/arc/shepherd_data2</DataDir>
+                <DataDir>/var/spool/arc/shepherd_store2</DataDir>
                 <TransferDir>/var/spool/arc/shepherd_transfer2</TransferDir>
@@
             <MinCheckInterval>0.1</MinCheckInterval>
+            <CreatingTimeout>600</CreatingTimeout>
             <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
@@
 
-Besides the \verb!arc_storage_cli! tool we can access the storage system with the use of a FUSE modul. We can install the ARC storage FUSE modul following these steps:
+\section{Make the A-Hash service replicated} % (fold)
+\label{sec:make_the_a_hash_service_replicated}
+
+Now we have a deployement, where one machine has the A-Hash, Librarian and Bartender services, and all three machines have a Shepherd and a storage element service.
+The A-Hash service stores all the metadata of all the files and collections, so it is critical.
+If the first machine goes offline then the whole system dies.
+We can deploy the A-Hash service on all the machines and still can have one consistent metadata database.
+The replicated version of the A-Hash uses the Berkeley DB to provide a distributed database.
+
+Let's install Berkeley DB as root (but you can use the \verb!--prefix! option of \verb!configure! to install it somewhere else than the default location).
+
+Get the latest Berkeley DB and its patches:
+\begin{verbatim}
+$ wget http://download.oracle.com/berkeley-db/db-4.7.25.tar.gz
+$ wget http://www.oracle.com/technology/products/berkeley-db/db/update/4.7.25/patch.4.7.25.1
+$ wget http://www.oracle.com/technology/products/berkeley-db/db/update/4.7.25/patch.4.7.25.2
+$ wget http://www.oracle.com/technology/products/berkeley-db/db/update/4.7.25/patch.4.7.25.3
+\end{verbatim}
+
+Extract it and apply the patches:
+\begin{verbatim}
+$ tar zxf db-4.7.25.tar.gz
+$ cd db-4.7.25
+$ patch -p 0 < ../patch.4.7.25.1
+$ patch -p 0 < ../patch.4.7.25.2
+$ patch -p 0 < ../patch.4.7.25.3
+\end{verbatim}
+
+Configure, make and install it (you can use the \verb!--prefix! option for \verb!configure!)
+\begin{verbatim}
+$ cd build_unix/
+$ ../dist/configure
+\end{verbatim}
+
+Now compile and install it:
+\begin{verbatim}
+$ make
+$ sudo make install
+\end{verbatim}
+
+Now we need the python wrapper for the Berkeley DB:
+
+\begin{verbatim}
+$ wget http://pypi.python.org/packages/source/b/bsddb3/bsddb3-4.7.5.tar.gz
+\end{verbatim}
+
+Let's install it to the default location:
+
+\begin{verbatim}
+$ tar zxf bsddb3-4.7.5.tar.gz
+$ cd bsddb3-4.7.5
+$ sudo python setup.py install
+\end{verbatim}
+
+If you've installed Berkeley DB somewhere else, or want to install the python wrapper to somewhere else than the default:
+
+\begin{verbatim}
+$ BERKELEYDB_DIR=/path/to/berkeleydb python setup.py install --prefix=/where/to/install
+\end{verbatim}
+
+Let's test the installation with the \verb!test.py! included in the directory of the python wrapper for the Berkeley DB:
+\begin{verbatim}
+$ sudo python test.py
+[...]
+Ran 383 tests in 29.858s
+
+OK
+\end{verbatim}
+
+\section{Using the FUSE module} % (fold)
+\label{sec:using_the_fuse_module}
+
+Besides the \verb!arc_storage_cli! tool we can access the storage system with the use of a FUSE module. We can install the ARC storage FUSE module following these steps:
 

### 2009-03-24 doc/storage_system/arc-storage-manual.tex
update of the manual about the replicated ahash and multiple librarians and bartenders
TODO: mount points

@@
 $ arc_storage_cli list
-- The URL of the Bartender: https://localhost:60000/Bartender
-- The key file: /home/zsombor/.arc/userkey.pem
-- The cert file: /home/zsombor/.arc/usercert.pem
-- The CA dir: /etc/grid-security/certificates
 Usage: list <LN> [<LN> ...]
@@
 $ arc_storage_cli list /
-- The URL of the Bartender: https://localhost:60000/Bartender
-- The key file: /home/zsombor/.arc/userkey.pem
-- The cert file: /home/zsombor/.arc/usercert.pem
-- The CA dir: /etc/grid-security/certificates
 - Calling the Bartender's list method...
@@
 \end{verbatim}
-
-(From now on I will omit the first few lines of the \verb!arc_storage_cli!'s output.)
 
@@
 
+Let's configure the our first server (\verb!storage!). Here is the AHash part of the config:
+
+\begin{verbatim}
+<Service name="pythonservice" id="ahash">
+    <ClassName>storage.ahash.ahash.AHashService</ClassName>
+    <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
+    <LocalDir>/var/spool/arc/ahash_data</LocalDir>
+    <MyURL>https://storage:60000/AHash</MyURL>
+    <OtherURL>https://storage2:60006/AHash</OtherURL>
+    <Priority>50</Priority>
+    <ClientSSLConfig FromFile='/etc/arc/clientsslconfig.xml'/>
+</Service>
+\end{verbatim}
+
+In this config we specify the second server as a peer. Let's configure the second server as well:
+
+\begin{verbatim}
+<Plexer name="plexer.service" id="plexer">
+    <next id="ahash">^/AHash$</next>
+    <next id="shepherd">^/Shepherd$</next>
+    <next id="byteio">^/byteio/</next>
+</Plexer>
+<Service name="pythonservice" id="ahash">
+    <ClassName>storage.ahash.ahash.AHashService</ClassName>
+    <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
+    <LocalDir>/var/spool/arc/ahash_data2</LocalDir>
+    <MyURL>https://storage2:60006/AHash</MyURL>
+    <OtherURL>https://storage:60000/AHash</OtherURL>
+    <Priority>50</Priority>
+    <ClientSSLConfig FromFile='/etc/arc/clientsslconfig.xml'/>
+</Service>
+\end{verbatim}
+
+In this config we specify the first server as a peer. So the two together will be able to form a replicated network with two nodes. But let's configure the thirs one:
+
+\begin{verbatim}
+<Plexer name="plexer.service" id="plexer">
+    <next id="ahash">^/AHash$</next>
+    <next id="shepherd">^/Shepherd$</next>
+</Plexer>
+<Service name="pythonservice" id="ahash">
+    <ClassName>storage.ahash.ahash.AHashService</ClassName>
+    <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
+    <LocalDir>/var/spool/arc/ahash_data3</LocalDir>
+    <MyURL>https://storage3:60010/AHash</MyURL>
+    <OtherURL>https://storage:60000/AHash</OtherURL>
+    <Priority>50</Priority>
+    <ClientSSLConfig FromFile='/etc/arc/clientsslconfig.xml'/>
+</Service>
+\end{verbatim}
+
+Here we specify the first server as the peer.
+
+It is important that we need to make a fresh start, and remove the existing database, which was created by a different store:
+\begin{verbatim}
+$ sudo rm -r /var/spool/arc/*
+\end{verbatim}
+
+Let's start the first server. It will complain that it cannot find its peer:
+\begin{verbatim}
+ERROR connecting to https://storage2:60006/AHash
+[2009-03-24 10:34:18] [Arc.Storage.ReplicatedAHash] [ERROR] [10827/156026112]
+    failed to send to 2 of [1, 2]
+[2009-03-24 10:34:20] [Arc.Storage.ReplicatedAHash] [ERROR] [10827/158713512]
+    Couldn't run election
+\end{verbatim}
+
+With only one node of the replicated A-Hash the system is not ready to work, if we try to use it, we get an error message:
+\begin{verbatim}
+$ arc_storage_cli mkdir /
+- Calling the Bartender's makeCollection method...
+- done in 0.55 seconds.
+Creating collection '/': failed to create new librarian entry
+\end{verbatim}
+
+Let's start the second server.
+It may first print some error messages, but then both servers should be quiet, and now we can use the system.
+Let's upload a file, and set the number of needed replicas to 3:
+
+\begin{verbatim}
+$ arc_storage_cli mkdir /
+- Calling the Bartender's makeCollection method...
+- done in 1.16 seconds.
+Creating collection '/': done
+$ arc_storage_cli put orange /
+- The size of the file is 7 bytes
+- Calculating md5 checksum...
+- done in 0.0214 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
+- Calling the Bartender's putFile method...
+- done in 2.22 seconds.
+- Got transfer URL: http://storage:60001/hopi/3f903c50-0424-2891-5730-521cc00d96dc
+- Uploading from 'orange'
+    to 'http://storage:60001/hopi/3f903c50-0424-2891-5730-521cc00d96dc' with http...
+    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+- done in 1.0293 seconds.
+'orange' (7 bytes) uploaded as '/orange'.
+$ arc_storage_cli modify /orange set states neededReplicas 3
+- Calling the Bartender's modify method...
+- done in 0.74 seconds.
+set
+$ arc_storage_cli stat /orange
+[...]
+  locations
+    https://storage:60000/Shepherd f4ad8530-354d-5066-04ae-c54c94fd525b: alive
+[...]
+zsombor@storage:~$ arc_storage_cli stat /orange
+[...]
+  locations
+    https://storage:60000/Shepherd f4ad8530-354d-5066-04ae-c54c94fd525b: alive
+    https://storage2:60006/Shepherd 97edebb5-e72d-098a-55a7-14595f668770: alive
+[...]    
+\end{verbatim}
+
+It cannot create 3 replicas, because are third server is not running. Let's start it. After a minute or two, or file has three replicas:
+
+\begin{verbatim}
+$ arc_storage_cli stat /orange
+[...]
+  locations
+    https://storage:60000/Shepherd f4ad8530-354d-5066-04ae-c54c94fd525b: alive
+    https://storage3:60010/Shepherd 3c55924b-73ad-c226-0f4a-4c056c967314: alive
+    https://storage2:60006/Shepherd 97edebb5-e72d-098a-55a7-14595f668770: alive
+[...]
+\end{verbatim}
+
+The fact that this file has three replicas is not at all related to the number of AHashes, but to the number of storage elements (Shepherds and storage element services), so this is no different then what we saw in the last section.
+The benefit of having more then one AHashes in a replicated way is that if one of the machines goes offline, the AHash is still working.
+But in our current case if we shutdown the first server, while the AHash will still work but we still have only one instance of the other two important service: the Librarian and the Bartender, so the whole system would be unusable.
+Let's add more Librarians and Bartenders to the system.
+
+\section{Additional Bartenders and Librarians} % (fold)
+\label{sec:additional_bartenders_and_librarians}
+
+Let's add a Bartender and a Librarian to each servers.
+We don't need to modify our first server now.
+For the other two we need to change the \verb!Plexer!
+section, add the twos services, and modify the Shepherd to use the server's own Bartender and Librarian for replication and heartbeats.
+Let's modify the config of our second server: (\verb![...]!
+indicates unchanged parts)
+
+\begin{verbatim}
+[...]
+<Plexer name="plexer.service" id="plexer">
+    <next id="ahash">^/AHash$</next>
+    <next id="shepherd">^/Shepherd$</next>
+    <next id="librarian">^/Librarian$</next>
+    <next id="bartender">^/Bartender$</next>
+    <next id="byteio">^/byteio/</next>
+</Plexer>
+<Service name="pythonservice" id="librarian">
+    <ClassName>storage.librarian.librarian.LibrarianService</ClassName>
+    <AHashURL>https://storage2:60006/AHash</AHashURL>
+    <HeartbeatTimeout>30</HeartbeatTimeout>
+    <CheckPeriod>20</CheckPeriod>
+    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
+</Service>
+<Service name="pythonservice" id="bartender">
+    <ClassName>storage.bartender.bartender.BartenderService</ClassName>
+    <LibrarianURL>https://storage2:60006/Librarian</LibrarianURL>
+    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
+</Service>
+[...]
+<Service name="pythonservice" id="shepherd">
+[...]
+    <LibrarianURL>https://storage2:60006/Librarian</LibrarianURL>
+    <BartenderURL>https://storage2:60006/Bartender</BartenderURL>
+[...]
+\end{verbatim}
+
+And the third server:
+
+\begin{verbatim}
+[...]
+<Plexer name="plexer.service" id="plexer">
+    <next id="ahash">^/AHash$</next>
+    <next id="shepherd">^/Shepherd$</next>
+    <next id="librarian">^/Librarian$</next>
+    <next id="bartender">^/Bartender$</next>
+</Plexer>
+<Service name="pythonservice" id="librarian">
+    <ClassName>storage.librarian.librarian.LibrarianService</ClassName>
+    <AHashURL>https://storage3:60010/AHash</AHashURL>
+    <HeartbeatTimeout>30</HeartbeatTimeout>
+    <CheckPeriod>20</CheckPeriod>
+    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
+</Service>
+<Service name="pythonservice" id="bartender">
+    <ClassName>storage.bartender.bartender.BartenderService</ClassName>
+    <LibrarianURL>https://storage3:60010/Librarian</LibrarianURL>
+    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
+</Service>
+[...]
+<Service name="pythonservice" id="shepherd">
+[...]
+    <LibrarianURL>https://storage3:60010/Librarian</LibrarianURL>
+    <BartenderURL>https://storage3:60010/Bartender</BartenderURL>
+[...]
+\end{verbatim}
+
+We need to modify our \verb!client.xml! to add all the other Bartenders:
+
+\begin{verbatim}
+$ cat ~/.arc/client.xml 
+<ArcConfig>
+  <KeyPath>/home/zsombor/.arc/userkey.pem</KeyPath>
+  <CertificatePath>/home/zsombor/.arc/usercert.pem</CertificatePath>
+  <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+  <BartenderURL>https://storage:60000/Bartender</BartenderURL>
+  <BartenderURL>https://storage2:60006/Bartender</BartenderURL>
+  <BartenderURL>https://storage3:60010/Bartender</BartenderURL>
+</ArcConfig>
+\end{verbatim}
+
+Now we can start all three servers.
+The replicated A-Hash sometimes prints great amount of error messages, but if it stops printing them, then everything is OK.
+Now we can use the system as intended.
+There is a flag for the \verb!arc_storage_cli!
+which makes it very verbose.
+We can check the used Bartender with it:
+
+\begin{verbatim}
+$ arc_storage_cli -v stat /orange
+- The URL of the Bartender(s): https://storage:60000/Bartender,
+    https://storage2:60006/Bartender, https://storage3:60010/Bartender
+- The key file: /home/zsombor/.arc/userkey.pem
+- The cert file: /home/zsombor/.arc/usercert.pem
+- The CA dir: /etc/grid-security/certificates
+- Calling the Bartender's stat method...
+- trying https://storage:60000/Bartender
+[...]
+- done in 1.27 seconds.
+'/orange': found
+\end{verbatim}
+
+Now we can stop our first server. The system will be unaccessable for a few seconds while the other two A-Hash figure out what is happening, but after that it will work again:
+
+\begin{verbatim}
+$ arc_storage_cli -v stat /orange
+- The URL of the Bartender(s): https://storage:60000/Bartender,
+    https://storage2:60006/Bartender, https://storage3:60010/Bartender
+- The key file: /home/zsombor/.arc/userkey.pem
+- The cert file: /home/zsombor/.arc/usercert.pem
+- The CA dir: /etc/grid-security/certificates
+- Calling the Bartender's stat method...
+- trying https://storage:60000/Bartender
+[...]
+- (-1, "Wrong status from server.\nMaybe wrong URL: 'https://storage:60000/Bartender'")
+- trying https://storage2:60006/Bartender
+[...]
+- done in 0.45 seconds.
+'/orange': found
+[...]
+\end{verbatim}
+
+If we start again our first server, after a few seconds the system will work again with the three fully-functional node.
+
+\begin{verbatim}
+zsombor@storage:~$ arc_storage_cli -v stat /orange
+- The URL of the Bartender(s): https://storage:60000/Bartender,
+    https://storage2:60006/Bartender, https://storage3:60010/Bartender
+- The key file: /home/zsombor/.arc/userkey.pem
+- The cert file: /home/zsombor/.arc/usercert.pem
+- The CA dir: /etc/grid-security/certificates
+- Calling the Bartender's stat method...
+- trying https://storage:60000/Bartender
+[...]
+- done in 0.49 seconds.
+'/orange': found
+  states
+    checksumType: md5
+    neededReplicas: 3
+    size: 7
+    checksum: 6aefd2842be62cd470709b27aedc7db7
+  entry
+    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
+    GUID: 0080d8be-1f6f-2fc7-ad71-732f899326c8
+    type: file
+  parents
+    0/orange: parent
+  locations
+    https://storage3:60010/Shepherd 03312843-d3b3-2d15-4732-ad673df958ae: alive
+    https://storage2:60006/Shepherd b0057765-4532-d0b4-fded-182499a91450: alive
+    https://storage:60000/Shepherd 532ec0f8-dd83-297b-d7b6-dc5dacea34e0: alive
+  timestamps
+    created: 1237887474.95    
+\end{verbatim}
+
 \section{Using the FUSE module} % (fold)

### 2009-03-24 doc/storage_system/arc-storage-documentation.tex
small update

@@
             <ProxyStore>/var/arc/spool/proxy_store</ProxyStore>
-            <CACertificatesDir>/etc/grid-security/certificates/
+            <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
         </GatewayCfg>

### 2009-03-24 doc/storage_system/arc-storage-manual.tex
small update

@@
 
+\section{Accessing external storage solutions} % (fold)
+\label{sec:using_the_gateway}
+
+install globus libraries
+
+\begin{verbatim}
+$ sudo mkdir /var/spool/arc/proxy_store
+\end{verbatim}
+
+\begin{verbatim}
+<Service name="pythonservice" id="bartender">
+    <ClassName>storage.bartender.bartender.BartenderService</ClassName>
+    <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
+    <ProxyStore>/var/spool/arc/proxy_store</ProxyStore>
+    <GatewayClass>storage.bartender.gateway.gateway.Gateway</GatewayClass>
+    <GatewayCfg>
+        <ProxyStore>/var/spool/arc/proxy_store</ProxyStore>
+        <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+    </GatewayCfg>
+    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
+</Service>
+\end{verbatim}
+
 \section{Using the FUSE module} % (fold)

### 2009-03-25 doc/storage_system/arc-storage-manual.tex
added gateway to manual and a small change in the documentation

--- 2009-03-24 doc/storage_system/arc-storage-manual.tex
@@
   pdfauthor = {Zsombor Nagy},
-  pdftitle = {User's manual of the ARC storage system},
+  pdftitle = {Administrator's manual of the ARC storage system},
   pdfsubject = {Paper subject},
@@
 %%%%---- The title ----
-{\centering \textsc{\Large Administrator's and user's manual of the ARC storage system}\Large \par}
+{\centering \textsc{\Large Administrator's manual of the ARC storage system}\Large \par}
 \vspace*{0.5cm}
@@
 
+\section{Accessing external storage solutions} % (fold)
+\label{sec:using_the_gateway}
+
+If we have permissions to access some third-party storage solutions (e.g.
+dCache), we can create a mount point within the namespace of the ARC storage system which points to a third-party URL, and then we can use the client interface of the ARC storage to access the third-party storage as well.
+This is achieved by a module of the Bartender, called the `Gateway'.
+Currently only GridFTP access is supported, and for that we need to install the Globus libraries to our machines.
+Then we need to recompile our source for GridFTP support.
+The output of the \verb!configure!
+script should look like this:
+
+\begin{verbatim}
+    Available third-party features:
+
+    RLS:                yes
+    GridFTP:            yes
+    LFC:                no
+    RSL:                yes
+    SAML:               yes
+    MYSQL CLIENT LIB:   no
+    gSOAP:              yes
+
+    Included components:
+    A-Rex service:      no
+    ISI service:        no
+    CHARON service:     no
+    HOPI service:       yes
+    SCHED service:      no
+    STORAGE service:    yes
+    PAUL service:       no
+    SRM client (DMC):   yes
+    GSI channel (MCC):  yes
+\end{verbatim}
+
+After compilation and installation we can use the \verb!arcls! tool to list the content of a third-party storage. Let's check if our Billy user has access to this directory:
+
+\begin{verbatim}
+$ arcls gsiftp://sal1.uppmax.uu.se:2811/pnfs/uppmax.uu.se/data
+data1
+data2
+other
+\end{verbatim}
+
+We can see some files in there, so we have permissions to list that directory.
+
+Now let's configure the Bartenders on all three machines to use the Gateway module:
+
+\begin{verbatim}
+<Service name="pythonservice" id="bartender">
+    <ClassName>storage.bartender.bartender.BartenderService</ClassName>
+    <LibrarianURL>[...]</LibrarianURL>
+    <ProxyStore>/var/spool/arc/proxy_store</ProxyStore>
+    <GatewayClass>storage.bartender.gateway.gateway.Gateway</GatewayClass>
+    <GatewayCfg>
+        <ProxyStore>/var/spool/arc/proxy_store</ProxyStore>
+        <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+    </GatewayCfg>
+    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
+</Service>
+\end{verbatim}
+
+Let's start all three servers. Then create a mount point with the Logical Name \verb!/salmount! which points to this URL:
+
+\begin{verbatim}
+$ arc_storage_cli makemount /salmount \
+    gsiftp://sal1.uppmax.uu.se:2811/pnfs/uppmax.uu.se/data
+- Calling the Bartender's makeMountpoint method...
+- done in 3.07 seconds.
+Creating mountpoint '/salmount': done
+
+$ arc_storage_cli stat /salmount
+- Calling the Bartender's stat method...
+- done in 0.47 seconds.
+'/salmount': found
+  states
+    closed: 0
+  entry
+    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
+    type: mountpoint
+    GUID: b21d9d22-e885-502a-ee0e-bd0a7b967543
+  parents
+    0/salmount: parent
+  mountpoint
+    externalURL: gsiftp://sal1.uppmax.uu.se:2811/pnfs/uppmax.uu.se/data
+  timestamps
+    created: 1237990168.09
+\end{verbatim}
+
+Let's list the content of this mountpoint:
+
+\begin{verbatim}
+$ arc_storage_cli list /salmount- Calling the Bartender's list method...
+- done in 0.49 seconds.
+'/salmount': Your proxy cannot be found. Please delegate your credentials!
+\end{verbatim}
+
+Yes, we first need to delegate our credentials with the \verb!credentialsDelegation! method, then try to list the mount point:
+
+\begin{verbatim}
+$ arc_storage_cli cre
+- Calling the Bartender's credentialsDelegation method...
+[...]
+- done in 0.42 seconds.
+Successful delegation.
+Proxy ID: eNCMDmi8i6YnPSAtDmVmuSEmABFKDmABFKDm0zKKDmFBFKDmeOzdBo
+
+$ arc_storage_cli list /salmount
+- Calling the Bartender's list method...
+- done in 3.19 seconds.
+'/salmount': collection
+  data1  <unknown>
+  data2  <unknown>
+  other  <unknown>
+\end{verbatim}
+
+Let's download a file:
+
+\begin{verbatim}
+$ arc_storage_cli get /salmount/data1 /tmp
+- Calling the Bartender's getFile method...
+- done in 3.56 seconds.
+- Got transfer URL: gsiftp://sal1.uppmax.uu.se:2811/pnfs/uppmax.uu.se/data/data1
+- Downloading from 'gsiftp://sal1.uppmax.uu.se:2811/pnfs/uppmax.uu.se/data/data1'
+    to '/tmp/data1' with gridftp...
+   2 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
+   3 s:        1.3 kB       0.4 kB/s       0.4 kB/s    . . .       
+- done in 2.7828 seconds.
+'/salmount/data1' (1290 bytes) downloaded as '/tmp/data1'.
+\end{verbatim}
+
 \section{Using the FUSE module} % (fold)

### 2009-03-25 doc/storage_system/arc-storage-manual.tex
manual about ARC DMC

@@
 \section{Adding more storage elements}
+
+This version of the \verb!arc_storage_cli! tool currently does not support recursive methods or working with multiple files at once.
 
@@
 
+\section{Using the ARC DMC} % (fold)
+\label{sec:using_the_arc_dmc}
+
+ARC has its own multi-protocol storage client tools which provide a unified way to access different kind of storage.
+For all the supported types there is a DMC module.
+We created a DMC module for the ARC storage as well.
+If it is compiled and installed, then you can use the \verb!arcls!, \verb!arccp!, \verb!arcrm!, etc.
+tools with the \verb!arc://!
+protocol to access the ARC storage system.
+You need to have a Bartender URL configured in the \verb!client.xml!
+user config.
+
+\begin{verbatim}
+$ arcls arc:///
+$ arccp orange arc:///orange
+$ arcls arc:///
+orange
+$ arccp arc:///orange /tmp/orange
+$ cat /tmp/orange 
+orange
+$ arcrm arc:///orange
+$ arcls arc:///
+$ 
+\end{verbatim}
+
+TODO: if Bartender returns a ByteIO URL - not good, need to fix it.
+
 \section{Using the FUSE module} % (fold)

### 2009-03-26 doc/storage_system/arc-storage-manual.tex
some small (last) updates

@@
 \begin{verbatim}
-$ arc_storage_cli mkdir /
+$ arc_storage_cli makeCollection /
 - Calling the Bartender's makeCollection method...
@@
 \begin{verbatim}
-$ arc_storage_cli mkdir /fruits
+$ arc_storage_cli makeCollection /fruits
 - Calling the Bartender's makeCollection method...
@@
 
-$ arc_storage_cli mkdir /fruits/apple
+$ arc_storage_cli makeCollecion /fruits/apple
 - Calling the Bartender's makeCollection method...
@@
 
-$ arc_storage_cli mkdir /
+$ arc_storage_cli makeCollection /
 - Calling the Bartender's makeCollection method...
@@
 protocol to access the ARC storage system.
-You need to have a Bartender URL configured in the \verb!client.xml!
+You can specify the URL of the Bartender with the \verb!BartenderURL!
+URL option, or you can have a Bartender URL configured in the \verb!client.xml!
 user config.
@@
 $ arcls arc:///
+orange
+$ arcls arc:///?BartenderURL=https://storage2:60006/Bartender
+orange
+$ arc_storage_cli list /
+- Calling the Bartender's list method...
+- done in 0.56 seconds.
+'/': collection
+  orange <file>
+$ arc_storage_cli makeCollection /fruits
+- Calling the Bartender's makeCollection method...
+- done in 2.38 seconds.
+Creating collection '/fruits': done
+$ arccp arc:///orange arc:///fruits/orange
+$ arcls arc:///
+orange
+fruits
+$ arcls arc:///fruits
 orange
@@
 $ arcls arc:///
-$ 
-\end{verbatim}
-
-TODO: if Bartender returns a ByteIO URL - not good, need to fix it.
+fruits
+\end{verbatim}
 

### 2009-04-10 doc/storage_system/arc-storage-documentation.tex
added gateway to manual and a small change in the documentation
TODO done
Alex comments up to section 3.4 + Style file for tracking changes

@@
 \usepackage{longtable}                             %for breaking tables
+\usepackage{trackchanges}
 %
@@
 
-The ARC storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.
-The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems), and a collection can contain sub-collections and sub-sub-collections in any depth.
+Chelonia is a distributed system for storing replicated \emph{file}s on several file storage nodes and managing them in a global namespace.
+The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems).
+A collection can contain sub-collections and sub-sub-collections in any depth.
 There is a dedicated \emph{root collection} to gather all collections to the global namespace.
-This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}s).
-The users can use this global namespace as they were using a local filesystem.
-Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user.
+This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}'s).
+The users can use this global namespace as if they were using a local filesystem.
+Files can be transfered by several different transfer protocols.
+The client side tools hide this from the user.
 The replicas of the files are stored on different storage nodes.
-A storage node here is a network-accessible computer having storage space to share, and a storage element service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.).
-For each storage node one of the services of the ARC storage system is needed to manage it and to integrate it into the system.
-There is a way on the client side to access third-party storage solutions through the namespace of the ARC storage system.
+A storage node here is a network-accessible computer having storage space to share, and a running storage element service (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.).
+\note[JKN]{TODO: Rephrase this sentence; "one of the services" seems vague}For each storage node one of the services of the ARC storage system is needed to manage it and to integrate it into the system.
+The client side provides access to third-party storage solutions through the namespace of the ARC storage system.
 The main services of the storage system are the following (see Figure~\ref{fig:services}):
@@
 \begin{figure}[ht]
-\centering{{\scalebox{0.9}{\includegraphics{arc-storage-services.pdf}}}
-\caption{\label{fig:services}The components of the ARC storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
+\centering{{\scalebox{0.8}{\includegraphics{arc-storage-services.pdf}}}
+\caption{\label{fig:services}{\bf TODO:} explain figure, since components are not yet introduced.
+The components of the ARC storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
 \end{figure}
@@
 Every file and collection has a unique ID in the system called the \emph{GUID}.
-Compared to the well-known structure of local file systems, these GUIDs are very similar to the concept of \emph{inode}s.
+Compared to the well-known structure of local file systems, these GUID's are very similar to the concept of \emph{inode}s.
 And as a directory on a local filesystem is basically just a list of name and inode pairs, a collection on the ARC storage is just a list of name and GUID pairs.
@@
 
-Besides the Logical Names we can refer to a file or collection by simply its GUID, or we can use GUIDs and Logical Names together, as seen on Figure~\ref{fig:namespace}.
+In addition to the Logical Names we can refer to a file or collection simply by its GUID, or we can use GUID's and Logical Names together, as seen in Figure~\ref{fig:namespace}.
 
@@
 
-Example on Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file with the Logical Name \verb#1234/green/orange/huge#, which means that from the collection called \verb!1234!
+An example of the hierarchy of the global namespace is shown in Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file with the Logical Name \verb#1234/green/orange/huge#, which means that from the collection called \verb!1234!
 we have to follow along the path: \verb!green!, \verb!orange!, \verb!huge!.
 
-There is a dedicated root collection (which has the GUID \verb#0#), and if a LN starts with no GUID prefix, it is implicitly prefixed with the GUID of this well-known root collection, e.g.~\verb#/why/blue# means \verb#0/why/blue#.
+There is a dedicated root collection (which has the GUID \verb#0#), and if a LN starts without a GUID prefix, it is implicitly prefixed with the GUID of the root collection, e.g.~\verb#/why/blue# means \verb#0/why/blue#.
 If a user wants to find the file called \verb#/why/blue#, the system knows where to start the search: the GUID of the root collection.
 The root collection knows the GUID of \verb#why#, and the (sub-)collection \verb#why# knows the GUID of \verb#blue#.
-If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, which concept is very similar to a hardlink in a regular local file system.
+If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, a concept very similar to a hardlink in a \note[JKN]{Todo: Posix file system?}regular local file system.
 
@@
 
-The collections in the ARC storage are logical entities, the content of a collection is stored as metadata of the collection, which means the a collection actually has no physical data.
-A file however has both metadata and real physical data (the actual bytes of the file).
+The collections in Chelonia are logical entities.
+The content of a collection is stored as metadata of the collection, which means that a collection actually has no physical data.
+A file, however, has both metadata and real physical data (the actual content of the file).
 The metadata of a file is stored in the same database where the collections are stored, but the physical data of a file is stored on storage nodes as multiple replicated copies.
@@
 a web server, an FTP server, a GridFTP server, etc.) and a Shepherd service which provides a simple interface to access the storage node, and which can initiate and manage file transfers through the storage element service.
-The Shepherd has different backends for the supported storage element services which made it possible the communicate with them.
-
-So we have logical files, which are part of the hierarchical namespace and have a GUID and other metadata, and a logical file has one or more physical replicas.
-The physical replicas stored on separate storage nodes.
-In order to connect the logical file to its replicas, we need to have some pointers.
+The Shepherd has different backends for the supported storage element services which made it possible to communicate with them.
+
+A logical file, which are part of the hierarchical namespace, has a GUID and other metadata, and one or more physical replicas.
+The physical replicas are stored on separate storage nodes.
+In order to connect the logical file to its replicas, the system needs to have global pointers.
 Each storage node has a URL and each replica has a unique ID within the storage node called \emph{referenceID}, the URL and the referenceID together is called a \emph{Location}, a Location unambiguously points to one specific replica.
-So to connect the logical files to the physical ones, each logical file has a list of Locations.
-
-The user can specify for each file how many replicas are needed.
-The ARC storage system periodically checks the number of replicas, and automatically creates new replicas if there are fewer then needed, and removes replicas if there are more.
-
-The different replicas of a file could be in different states, e.g.
+To connect the logical files to the physical ones, each logical file has a list of Locations.
+
+The user can specify how many replicas are required for each file.
+The ARC storage system periodically checks the number of replicas\note[JKN]{TODO: How will this scale?}, and automatically creates new replicas if there are fewer than required, and removes replicas if there are more.
+
+The different replicas of a file could be in different \note[JKN]{TODO: which states?
+a reference to definition of states maybe?}states, e.g.
 the replica could be valid and alive or just in the process of creation or it could be corrupt or a whole storage node could be offline.
-This state is always stored as metadata next to the Location of the given replica.
+This state is always stored as metadata of the given replica.
 For each file there is a checksum calculated, and this checksum is used to detect if a replica gets corrupted.
@@
 
-The A-Hash is the replicated\footnote{Depends on system configuration.
-It is possible to configure the storage system with only one central A-Hash too} metadata store of the storage system.
+The A-Hash is the metadata store of the storage system.
+The A-Hash may be deployed as either a centralized or a replicated service.
 It consistently stores `objects' which contain property-value pairs organized in sections.
@@
 It can traverse Logical Names and return the corresponding metadata.
-The Shephards send their heartbeat messages to the Librarians which if needed respond by modifying the states of files stored on a certain storage node.
+The Shepherds (see Sect.
+\ref{sec:the_shepherds}) send their heartbeat messages to the Librarians and, if needed, the Librarian responds by modifying the states of files stored on a certain storage node.
 The Librarian itself is a stateless service, it uses the A-Hash to store and retrieve the metadata and thus there can be any number of independent Librarian services (all using the same A-Hashes).
@@
 
-The Bartender service provides the high-level interface of the storage system to the clients.
+The Bartender service provides the high-level interface of the storage system to clients.
 Every interaction between a client and the ARC storage system starts by sending a request to a Bartender.
 Using the Bartender it is possible to create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names, etc.
-The Bartender authorizes users and apply access policies of files and collections.
-It communicates with the Librarian and Shepherd services to accomplish the client’s requests.
-The file data itself does however not go through the Bartender; file transfers are directly performed between the storage nodes and the clients.
-There could be any number of independent Bartender services in the system which ensure high-availability and load-balancing.
-The Bartender also provides a way to access files on third-party storage solutions through its interface by mounting the namespace of the third-party storage into the namespace of the ARC storage (this is accomplished by so called `gateway' modules).
+The Bartender authorizes users and applies access policies of files and collections.
+It communicates with the Librarian and Shepherd services to execute the requests of the clients.
+The file data itself does not go through the Bartender; file transfers are performed directly between the storage nodes and the clients.
+Any number of independent Bartender services may be deployed in the system to ensure high-availability and load-balancing.
+The Bartender also provides access to files in third-party storage solutions through its interface by mounting the namespace of the third-party storage into the namespace of the ARC storage (this is accomplished by so-called `gateway' modules).
 
@@
 For each kind of storage element service (e.g.
-a HTTP server, an FTP server, a storage solution with a GridFTP interface, etc.) a Shepherd backend is needed which is capable of managing the given storage element service.
+an HTTP server, an FTP server, a storage solution with a GridFTP interface, etc.) a Shepherd backend is needed which is capable of managing the given storage element service.
 The Shepherd service periodically checks the health of the replicas using their checksums, and if a replica is deleted or corrupted, the Shepherd tries to recover it by downloading a valid copy from another storage node.
-The Shepherd also periodicallys check if a file has the correct number of replicas in the system.
+The Shepherd also periodically checks if a file has the correct number of replicas in the system.
 If too few or too many replicas are found, new replicas will be created or redundant replicas removed.
@@
 
-The ARC storage system consists of several services (both different kinds of services and multiple instances of the same kind).
+Chelonia consists of several services (both different kinds of services and multiple instances of the same kind).
 The A-Hash, Librarian and Shepherd services are `internal' services in the sense that the end-user of the storage system never communicates with them directly.
 But these internal services are communicating with each other, thus they have to know which other services to trust.
-We call this aspect of the security architecture of the ARC storage `inter-service authorization'.
-
-The end users always connect to one of the Bartender services, which will decide which permission the user has. This is the `high-level authorization' part of the security architecture.
+We call this aspect of the security architecture of the ARC storage system `inter-service authorization'.
+
+The end users always connect to one of the Bartender services, which will decide which permissions the user has. This is the `high-level authorization' part of the security architecture.
 
 For transferring the actual file data, the users have to connect to storage element services running on storage nodes.
-These services also have their own authentication and authorization methods.
-Managing these aspects is the `transfer-level authorization' part of the security architecture of the ARC storage.
+These services have their own authentication and authorization methods.
+Managing these aspects is the `transfer-level authorization' part of the security architecture of Chelonia.
 
@@
 
-In a deployment of the ARC storage system it is possible to have several A-Hash, Librarian, Shepherd and Bartender services.
-The Bartenders send requests to the Librarians and the Shepherds, the Shepherds communicate with the Librarians, the Librarians talk with the A-Hashes.
-If any of these services gets compromised or a new rouge service is inserted in the system, security is lost.
+In a deployment of Chelonia, it is possible to have several A-Hash, Librarian, Shepherd and Bartender services.
+The Bartenders send requests to the Librarians and the Shepherds, the Shepherds communicate with the Librarians and the Librarians with the A-Hashes.
+If any of these services gets compromised or a new rogue service is inserted in the system, security is lost.
 That's why it is vital for each service to authorize other services before sending or accepting requests.
-The services communicate via the HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and that they can examine eachother's certificates.
-Because of these X.509 certificates each service has Distinguish Name (DN), which can be used for identifying trusted services.
-A list of trusted DNs can be configured into each service, or a list of trusted services can be stored on a remote location.
-The services will only accept connections if the DN of the other end is listed in this list of trusted DNs.
-However the Bartender services will accept any incoming connection, which are from the users, because the users are authenticated differently.
+The services communicate via the HTTPS protocol, which means that they should provide an X.509 certificate for each connection, and that they can examine each other's certificates.
+Each service is required to have an X.509 certificate and is thus identified by its Distinguish Name (DN).
+A list of trusted DN's can be configured into each service, or a list of trusted services can be stored on a remote location.
+The services will only accept connections if the DN of the other end is listed in this list of trusted DN's.
+However, the Bartender services will accept any incoming connection, which are from the users, because the users are authenticated differently.
 
@@
 For each file and collection there are access policies in the form of access control rules, and these are stored among the metadata.
-The users are identified by their DNs, and an access control rule specify the rights of the given user.
+The users are identified by their DN's, and an access control rule specifies the rights of the given user.
 A rule can be represented like this:
@@
 
-Besides specifying only one user with a DN, there are other types of access control rules: we can have a rule for a whole VO (Virtual Organization) or for ALL users, like this:
-
-\begin{verbatim}
+Besides specifying only one user with a DN, there are other types of access control rules: we can have a rule for a whole VO (Virtual Organization) or for all users, like this:
+
+\begin{verbatim}
+    VOMS:knowarc.eu +action -action 
     ALL +action
-    VOMS:knowarc.eu +action -action 
 \end{verbatim}
@@
 
-Currently the transfer-level authorization is kept very simple.
+Currently the transfer-level authorization is very simple.
 When the Bartender decides that a user has permission to download a file, then the Bartender chooses a replica, and initiates the transfer.
-The result of this initiation is a URL which is called the transfer URL (TURL).
-This TURL is unique for each request, even for request to the same replica, and this TURL is only valid for one download.
-Currently we configure the storage element services to not do any authorization, and we use these one-time URLs to ensure that only the authorized users can access the contents of the storage elements.
+The result of this initiation is the transfer URL (TURL).
+This TURL is unique for each request, even for requests for the same replica, and this TURL is only valid for one download.
+Currently we configure the storage element services to not do any authorization, and we use these one-time TURL's to ensure that only authorized users can access the contents of the storage elements.
 
@@
 
-\chapter{Use cases} % (fold)
+\chapter{Implementation of use cases} % (fold)
 \label{cha:use_cases}
+
+\note[JKN]{TODO: Short intro to this chapter mentioning that this chapter contains a partial list of use cases and that not all features are covered.}
 
@@
 
-We want to list the contents of a collection, which has a Logical Name of \verb!/ourvo/users/we!.
-In order to do this, we have to contact a Bartender, and send a request to it containing the Logical Name we want to list, and the response from the Bartender will contain the list of entries.
-The steps are represented on Figure~\ref{fig:listing}.
+A user wants to list the contents of a collection, which has a Logical Name of \verb!/ourvo/users/user!.
+In order to do this, the client tool has to contact a Bartender, and send a request to it containing the Logical Name the user wants to list, and the response from the Bartender will contain the list of entries.
+The steps are represented in Figure~\ref{fig:listing}.
 
 \begin{enumerate}
-    \item We need to know the URL of a Bartender.
+    \item The client tool needs to know the URL of a Bartender.
 This could be preconfigured on the client side or in future releases it could be acquired from an information system.
-When we have the URL, we send a `list' request which contains the Logical Name \verb!/ourvo/users/we!.
+When the client tool has the URL, it sends a `list' request which contains the Logical Name \verb!/ourvo/users/user!.
     \item The Bartender tries to get the metadata of the given LN by sending a `traverseLN' request to a Librarian.
@@
 root collection.
-The A-Hash service only knows about GUIDs and not about LNs, but the GUID of the root collection is well-known, so the A-Hash can return the metadata of it which contains the list of files and sub-collections in the root collection.
-    \item Hopefully the \verb!ourvo! collection can be found in the root collection, which means that new the Librarian knows its GUID, and can ask for its metadata from the A-Hash.
+The A-Hash service only knows about GUID's and not about LN's, but the GUID of the root collection is well-known, so the A-Hash can return the metadata of the list of files and sub-collections in the root collection.
+    \item Hopefully, the \verb!ourvo! collection can be found in the root collection, which means that the Librarian knows its GUID, and can ask for its metadata from the A-Hash.
     \item After the A-Hash returns the metadata of the \verb!/ourvo! collection, the Librarian finds the GUID of \verb!users! in it, then gets its metadata.
-    \item The A-Hash returns the metadata of \verb!/ourvo/users! which contains the GUID of \verb!we!, so the Librarian can ask for its metadata.
-    \item At last the A-Hash returns the metadata of \verb!/ourvo/users/we!
+    \item The A-Hash returns the metadata of \verb!/ourvo/users! which contains the GUID of \verb!user!, so the Librarian can ask for its metadata.
+    \item At last the A-Hash returns the metadata of \verb!/ourvo/users/user!
 to the Librarian, and the Librarian returns it to the Bartender.
 This metadata contains the list of entries within this collection, and it also contains the access policies for this collection.
-    \item The Bartender first checks if based on our DN (or our VO membership) and the access policies of this collection do we have rights to get the contents of this collection or not.
-If we are approved, then because the `list' request should return additional metadata about each entry in the collection, the Bartender send a `get' message to a Librarian requesting metadata of all the entries in this collection.
-    \item The Librarian gets the data from the A-Hash and returns it to the Bartender.
-    \item Now the Bartender has all the needed information, so it could return the proper response to us, the user. Our client tool formats and prints the results nicely.
+    \item The Bartender first checks, based on the user's DN (or his VO membership) and the access policies of this collection, if the user has permission to get the contents of this collection or not.
+If the user is approved, then because the `list' request should return additional metadata about each entry in the collection, the Bartender sends a `get' message to a Librarian which requests the metadata of all the entries in this collection.
+    \item The Librarian gets the metadata from the A-Hash and returns it to the Bartender.
+    \item Now the Bartender has all the needed information and returns the proper response to the client tool which formats and prints the results nicely for the user.
 \end{enumerate}
@@
 
-In this use case we want to download a file which has a Logical Name of \verb#/ourvo/users/we/thefilewewant# (see Figure~\ref{fig:downloading}).
+In this use case a user wants to download a file which has a Logical Name of \verb#/ourvo/users/user/thefilehewants# (see Figure~\ref{fig:downloading}).
 
 \begin{enumerate}
-    \item We connect a Bartender and send a `getFile' request with the LN of our file.
+    \item The client tool connects to a Bartender and sends a `getFile' request with the LN of the file.
     \item The Bartender contacts a Librarian to traverse the Logical Name and to get the metadata of our file.
-    \item The Librarian do the step by step traversing and gets all the data from the A-Hash and returns the metadata of our file to the Bartender.
+    \item The Librarian does the step-by-step traversing of the name space and gets all the metadata from the A-Hash and returns the metadata of our file to the Bartender.
 This metadata contains the location of the file's replicas and the access policies of this file.
-    \item The Bartender checks based on the access policies and our identity if we are allowed to get the file, and if we are good to go, then it chooses a replica location.
+    \item The Bartender checks, based on the access policies and the user's identity, if he is allowed to get the file, and if allowed, it chooses \note[JKN]{TODO: We need to define the choosing algorithm somewhere} a replica location.
 A location consists of the URL of a Shepherd service, and the ID of the replica within that Shepherd (which is called a `referenceID').
 The Bartender sends a `get' request to the chosen Shepherd.
-    \item The Shepherd prepares the file transfer by asking the storage element service to create a new one-time URL for this replica.
-This URL called the transfer URL, and it will only be valid for one download.
-The Shepherd returns the TURL to the Bartender.
-    \item The Bartender returns the TURL to us.
-    \item Now we use this TURL to get the file directly from the storage element service on the storage node.
+    \item The Shepherd prepares the file transfer by asking the storage element service to create a new one-time TURL for this replica. The Shepherd returns the TURL to the Bartender.
+    \item The Bartender returns the TURL to the client tool.
+    \item The client tool use this TURL to get the file directly from the storage element service on the storage node.
 \end{enumerate}
@@
 
-We want to create a new (empty) collection as sub-collection of \verb#/ourvo/common#, and we want to call it \verb!docs! (see Figure~\ref{fig:makecollection})
+Here, the user wants to create a new (empty) collection as sub-collection of \verb#/ourvo/common#, and he wants to call it \verb!docs! (see Figure~\ref{fig:makecollection})
 
 \begin{enumerate}
-    \item We contact a Bartender and send a `makeCollection' request with the LN \verb!/ourvo/common/docs!.
+    \item The client tool contacts a Bartender and send a `makeCollection' request with the LN \verb!/ourvo/common/docs!.
     \item The Bartender asks a Librarian to traverse this LN.
-    \item The Librarian try to traverse the Logical Name and it stops at the last possible point and returns the metadata of the last element.
-Because we want to put our collection to a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed.
+    \item The Librarian tries to traverse the Logical Name and it stops at the last possible point and returns the metadata of the last element.
+Because the user wants to put his collection in a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed.
 If the Librarian could traverse the whole LN that would mean that there is already an existing file or collection by that name.
-If the parent collection does exist then the Librarian can traverse it which means that the parent collection's metadata is returned to the Bartender.
-    \item The Bartender checks the access policies to decide if we have permissions to put something into this collection. Then it asks the Librarian to create a new collection.
-    \item The Librarian creates the collection, and returns its GUID.
-At this point this new collection has no real Logical Name yet, it only has a GUID, but it is not yet put into its parent collection.
+If LN does not exist but the the parent collection does, then the parent collection's metadata is returned to the Bartender.
+    \item The Bartender checks the access policies to decide if the user has permissions to put something into this collection. Then it asks the Librarian to create a new collection.
+    \item The Librarian creates the collection, and returns its GUID. At this point this new collection has no real Logical Name, it only has a GUID, but it is not yet put into its parent collection.
     \item The Bartender now asks the Librarian to add this new entry into the parent collection, which means that the new GUID and the name \verb!docs! are added as a pair.
     \item The Librarian returns with a status message.
-    \item Finally the Bartender tells us if everything went OK or not.
+    \item Finally the Bartender tells the client tool if everything went OK or not.
 \end{enumerate}
@@
 
-We have a file on our local disk we want to upload to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
+The user has a file on his local disk and wants to upload the file to a collection called \verb#/ourvo/common/docs#. (See Figure~\ref{fig:uploading}.)
 \begin{enumerate}
-    \item We contact a Bartender to put the file, we give the size and checksum and other metadata.
-And of course we give the Logical Name where we want to put the file, which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
-    \item The Bartender ask a Librarian to traverse this LN.
-    \item The Librarian traverses the Logical Name, and we expect it to stop at the \verb#/ourvo/common/docs# part of the LN, because that means that the name is available and the parent collection exists.
-If everything is fine, the metadata of the parent collection is returned to the Bartender.
-    \item The Bartender checks the access policies to decide if we have permissions to put something into this collection. Then it asks the Librarian to create a new file entry. 
+    \item The client tool contact a Bartender to upload the file, providing the file's metadata, including the the Logical Name (LN), which in this case will be \verb#/ourvo/common/docs/proposal.pdf#
+    \item The Bartender asks a Librarian to traverse this LN.
+    \item The Librarian traverses the Logical Name, and stopping at the \verb#/ourvo/common/docs# part of the LN, which means that the name is available and the parent collection exists.
+If everything is OK, the metadata of the parent collection is returned to the Bartender.
+    \item The Bartender checks the access policies to decide if the user has permissions to put anything into this collection. Then it asks the Librarian to create a new file entry. 
     \item The Librarian creates the entry and returns its GUID.
-    \item Then the Bartender add the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf#.
+    \item Next, the Bartender adds the name \verb#proposal.pdf# and the new GUID to the collection \verb#/ourvo/common/docs# and from now on there will be a valid LN \verb#/ourvo/common/docs/proposal.pdf#.
 However this LN points to a file which has currently no replica at all.
-If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, would get an error message.
+If someone tried to download the file called \verb#/ourvo/common/docs/proposal.pdf# now, they would get an error message.
     \item The Bartender asks the Librarian about Shepherd services (which are sitting on storage nodes).
-    \item The Librarian returns a list of Shepherd services
-    \item The Bartender chooses a Shepherd service and sends it a `put' request to initiate the file upload. 
+    \item The Librarian returns a list of Shepherd services.
+    \item The Bartender randomly chooses a Shepherd service and sends it a `put' request to initiate the file upload. 
     \item The Shepherd communicates with the storage element service on the same node to create a new transfer URL (TURL).
-Then it creates a `referenceID' for this file and then reports to the Librarian that there is a new replica in \verb#creating# state.
-The Librarian gets the message from the Shepherd and adds the new replica to our new file.
+Then it creates a `referenceID' for this file and reports to the Librarian that there is a new replica in the \verb#creating# state.
+The Librarian gets the message from the Shepherd and adds the new replica to the new file.
 Now the file has one replica, which is not uploaded yet into the system.
-If someone tries to download this file now, still gets an error message.
+If someone tries to download this file now, they still get an error message.
     \item The Shepherd returns the the TURL to the Bartender.
-    \item The Bartender returns the TURL to us.
-    \item Then we can upload the file to this TURL.
-    \item The Shepherd detects that the file is arrived.
+    \item The Bartender returns the TURL to the client tool.
+    \item Then client tool can upload the file to this TURL.
+    \item The Shepherd detects\note[JKN]{TODO: How does it detect it?
+Callback from the transfer service?} that the file has been uploaded.
 It checks the checksum of the file, and if it is OK, then it reports to the Librarian, that this replica is now \verb#alive#.
-The Librarian alters the state of this location, and now finally the file has one valid replica.
+The Librarian alters the state of this location, and now the file has one valid replica.
 \end{enumerate}
-í
+
 % section uploading_a_file (end)
@@
 \end{figure}
+In Chelonia, files are deleted through so called {\it lazy deletion}, meaning that at first only the Logcial Name (LN) is deleted, while the physical replicas are removed at a later stage.
 \begin{enumerate}
-    \item If we want to remove a file, we should connect to a Bartender with the LN of the file we want to remove.
+    \item If the user wants to remove a file, his client tool should connect to a Bartender with the LN of the file he wants to remove.
     \item The Bartender asks the Librarian to traverse the LN.
@@
     \item After that, the Bartender asks the Librarian to remove the links to this file from all the parent collections.
-    \item Next time a Shepherd which has a replica of this file does its periodic check, it asks the Librarian about the file, and notices that the file does not exist anymore, so it removes the replica itself from the storage node.
+    \item The next time a Shepherd which has a replica of this file does its periodic check, it asks the Librarian about the file, and notices that the file does not exist anymore, so it removes the replica itself from the storage node.
 \end{enumerate}
@@
 
-The services are written in Python and running in the HED\footnote{The ARC container - \url{https://www.knowarc.eu/documents/Knowarc\_D1.2-2\_07.pdf}} hosting environment.
-The HED itself is written in C++, but there are language bindings which allow us to write services in other languages, e.g.~in Python or Java.
-The source code of the storage services are in the NorduGrid Subversion repository\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage}}.
-
-Because the next-generation information system of ARC is currently under development, we cannot use it to discover services, that's why currently the URLs of almost all the services are hard-coded in the configuration files of the service.
-There is one exception: the Shepherd services are reporting their URLs to the Librarians, so a Bartender always could get a very current list of alive Shepherd services.
-
-The HED also has a security framework which the Bartenders use the make access policy decision . The details of this is described in section X.
+The Chelonia services are written in Python and execute in the HED\footnote{The ARC container - \url{https://www.knowarc.eu/documents/Knowarc\_D1.2-2\_07.pdf}} hosting environment.
+The HED itself is written in C++, but there are language bindings which allow services to be written in other languages, e.g.~in Python or Java.
+The source code of the storage services is in the NorduGrid Subversion repository\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/storage}}.
+
+Because the next-generation information system of ARC is currently under development, it cannot be used to discover services.
+Hence, currently the URL's of almost all the services are hard-coded in the configuration files of the service.
+There is one exception: the Shepherd services are reporting their URL's to the Librarians, so a Bartender can always get the most current list of accessible Shepherd services.
+
+The HED has a security framework which the Bartenders use the make access policy decisions. The details of this is described in Section \ref{sec:security_implementation}.
 
@@
 
-The ARC storage system stores different kinds of metadata about files, collections, mount point, Shepherd services, etc. Each of these has a unique ID which we call `GUID'.
+The ARC storage system stores different kinds of metadata about files, collections, mount points, Shepherd services, etc. Each of these has a unique ID; a GUID.
 
 The A-Hash services provide a functionality to store `objects', where each object has a unique ID, and contains property-value pairs organized in sections.
-The properties, the values and the section names too are simple character strings.
+The properties, the values and the section names are simple character strings.
 The A-Hash services provide a simple interface to interact with these objects, e.g.~to do conditional and atomic changes.
 
-The Librarian services use the A-Hash services to store the metadata of all kinds, using the GUIDs as unique IDs.
+The Librarian services use the A-Hash services to store the metadata, using the GUID's as unique ID's.
 For this, we have to represent the metadata as property-value pairs organized in section.
-The files, the collections and the mount points have some common attributes: the GUID, the type and the owner which are in the \emph{entry} section; the creation and modification timestamps in the \emph{timestamps} section, a list of access control rules in the \emph{policy} section; the list of parent collections where this entry has a `hardlink'; and each entry could have arbitrary property-value pairs in the \emph{metadata} section.
+The files, the collections and the mount points have some common attributes: the GUID, the type and the owner which are in the \emph{entry} section; the creation and modification timestamps in the \emph{timestamps} section, a list of access control rules in the \emph{policy} section; the list of parent collections where this entry has a `hardlink' in the \emph{parents} section; and optional, arbitrary property-value pairs in the \emph{metadata} section.
 
@@
         \item \emph{GUID}: the unique ID of the entry
-        \item \emph{owner}: the DN of the user how owns this entry
+        \item \emph{owner}: the DN of the user who owns this entry
     \end{itemize}
@@
     \begin{itemize}
-        \item \emph{(identity, action list) pairs}: a list of allowed and not allowed actions for different identities (users, VOs, etc.). 
+        \item \emph{(identity, action list) pairs}: a list of allowed and not allowed actions for different identities (users, VO's, etc., see Section \ref{sub:high_level_authorization} for details). 
     \end{itemize}
@@
     \begin{itemize}
-        \item list of GUIDs of collection where this entry is located, and the name of this entry within the given collections
+        \item list of GUID's of collection where this entry is located, and the name of this entry within the given collections
     \end{itemize}
@@
     \begin{itemize}
-        \item any other arbitrary property-value pairs
+        \item optional arbitrary property-value pairs
     \end{itemize}
@@
 
-A file in the ARC storage has a logical entry in the namespace, and a couple of physical replicas on the storage nodes.
-We store the locations of the replicas in the \emph{location} section.
+A file in Chelonia has a logical entry in the namespace, and one or more physical replicas on the storage nodes.
+We store the locations of the replicas in the \emph{locations} section.
 A location contains the URL of the Shepherd service, the local ID of the file within the storage node and the state of the replica.
-This state could be `\textbf{alive}' (if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy), `\textbf{invalid}' (if the replica has wrong checksum, or the Shepherd claims it has no such file), `\textbf{offline}' (if the Shepherd is not reachable, but may have a valid replica), `\textbf{creating}' (if the replica is in the state of uploading), `\textbf{thirdwheel}' (if the replica is marked for deletion because of too many replicas), `\textbf{stalled}' (if the the replica was in the state if uploading for a long time but hasn't arrived).
-We also store (in the \emph{states} section) the size of the file, the number of needed replicas, and a checksum of the file, which is created by some kind of checksumming algorithm (currently only md5 is supported).
+This state is `\textbf{alive}' if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy, `\textbf{invalid}' if the replica has wrong checksum, or the Shepherd claims it has no such file, `\textbf{offline}' if the Shepherd is not reachable, but may have a valid replica, `\textbf{creating}' if the replica is in the state of uploading, `\textbf{thirdwheel}' if the replica is marked for deletion because of too many replicas, or `\textbf{stalled}' if the the replica was in the state if uploading for a long time but hasn't arrived.
+The \emph{states} section contains the size of the file, the number of required replicas, and a checksum of the file, which is created by some kind of checksumming algorithm (currently only md5 is supported).
 
@@
     \begin{itemize}
-        \item \emph{(location, state) pairs}, where a location is a (\emph{URL}, \emph{referenceID}) pair serialized as a string, where \emph{URL} is the address of the Shepherd service storing this replica, \emph{referenceID} is the ID of the file within that Shepherd service.
+        \item \emph{(location, state) pairs}: where a location is a (\emph{URL}, \emph{referenceID}) pair serialized as a string, where \emph{URL} is the address of the Shepherd service storing this replica, \emph{referenceID} is the ID of the file within that Shepherd service.
     \end{itemize}
@@
 A \emph{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-hierarchy.
-Each entry has a unique name within a collection and a GUID which points to corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section.
+Each entry has a unique name within a collection and a GUID which points to a corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section (see below).
 The collections could be in a closed state, which means that its contents should not be changed.
@@
 if we include in our collection a file which we don't own, then the owner of the file could remove it nevertheless), but if a closed collection is changed, its state will be broken, and this state could never be changed again, which means you will always know if a closed collection is not intact anymore.
-This state is stored in the \emph{states} section, and its values could be `\textbf{no}' (if the collection is not closed), `\textbf{yes}' (if the collection is closed) and `\textbf{broken}' (if the collection was closed, but something happened, and its contents have been changed).
+This state is stored in the \emph{states} section, and its possible values are `\textbf{no}' if the collection is not closed, `\textbf{yes}' if the collection is closed and `\textbf{broken}' if the collection was closed, but something happened, and its contents have been changed.
 
@@
 
-The mount point entry is a reference to an external third-party storage. Its metadata is basically just the URL of the external storage.
+The mount point entry is a reference to an external third-party storage. Its metadata is basically just the URL of the external storage end point.
 
@@
     \begin{itemize}
-        \item \emph{externalURL}: the URL of the external storage.
+        \item \emph{externalURL}: the URL of the external storage end point \note[JKN]{TODO: an example URL maybe?}.
     \end{itemize}
@@
 Each Shepherd reports its URL and the list of its files to a Librarian, and for each Shepherd a GUID is created.
-There is a special entry (with GUID `1' by default) which stores a list of the registered Shepherd services, storing their GUID and the timestamp of the last heartbeat message from them.
+There is a special entry (with GUID `1' by default) which stores a list of the registered Shepherd services, with their GUID and the timestamp of the last heartbeat message from them.
 
@@
 	\begin{itemize}
-		\item (URL, timestamp) pairs contains when was the last heartbeat of this service
+		\item \emph{(URL, timestamp) pairs}: contains when was the last heartbeat of this service
 	\end{itemize}
@@
     \begin{itemize}
-        \item (URL, GUID) pairs connects the URL of the Shepherd to the GUID where the information is stored about the Shepherd's stored files
+        \item \emph{(URL, GUID) pairs}: connects the URL of the Shepherd to the GUID where the information is stored about the Shepherd's stored files
     \end{itemize}
@@
 	\begin{itemize}
-	    \item \emph{(referenceID, GUID) pairs} for each replica stored on the Shepherd
+	    \item \emph{(referenceID, GUID) pairs}: for each replica stored on the Shepherd
 	\end{itemize}
@@
 The ARC HED hosting environment has a security framework, and it has its own policy language for describing access policies and requests.
-The storage system use a different internal format to store the access policies, but when the time comes to make decisions, it converts the stored policies and the incoming requests to the ARC native policy format, then asks the security framework to decide.
+The storage system currently use a different internal format to store the access policies, but when the time comes to make decisions, it converts the stored policies and the incoming requests to the ARC native policy format, then asks the security framework to decide.
 
@@
 
-When there is an incoming connection to a Bartender, the security framework of the HED extract the identity information of the connecting client (DN of client, DN of the issuer CA, extended attributes, etc.), and all these information is accessible by the Bartender.
-When the given method of the Bartender gets to the point where the access rules are present, and it is clear that what kind of action the user wants to do, then either the request and the policy is converted to the native ARC policy language, and then the policy evaluator of the security framework is called to make the decision.
-Within the native ARC policy language XML representation of the requests and policies the storage actions are put into an `Attribute' element with the ID `\verb!http://www.nordugrid.org/schemas/policy-arc/types/storage/action!'.
-
-For inter-service authorization we can configure the services to only accept connections from trusted services.
-We trust a service if we know its DN.
-For each service we can write trusted DNs in the configuration, or we can periodically get the list from an A-Hash.
+When there is an incoming connection to a Bartender, the security framework of the HED extracts the identity information of the connecting client (DN of client, DN of the issuer CA, extended attributes, etc.), and all this information is accessible by the Bartender.
+When the given method of the Bartender gets to the point where the access rules are present, and it is clear that what kind of action the user wants to do, then either\note[JKN]{TODO: Where is the 'or' that should follow the either?} the request and the policy is converted to the native ARC policy language, and then the policy evaluator of the security framework is called to make the decision.
+Within the native ARC policy language, the XML representation of the requests and policies the storage actions are put into an `Attribute' element with the ID `\verb!http://www.nordugrid.org/schemas/policy-arc/types/storage/action!'\note[JKN]{TODO: This sentence needs more explanation; too technical.}.
+
+For inter-service authorization the services can be configured to only accept connections from trusted services.
+A service is trusted if we know its DN.
+For each service we can write trusted DN's in the configuration, or we can periodically get the list from an A-Hash.
 This can be configured with a `\emph{TrustManager}' section in the service configuration, which contains these entries:
@@
     \item[DN] is a trusted DN
-    \item[CA] is a DN of a trusted CA: we will trust all certificates issued by this CA
-    \item[DNsFromAHash] is a list of A-Hash URLs (in \textbf{AHashURL}) from where we periodically get a list of trusted DNs (there can be multiple A-Hash URLs listed here in case of one is offline).
-This has some attributes: \textbf{CheckingInterval} specifies how frequently we should get the list from the A-Hash (default: 600); \textbf{ID} specifies the ID of the A-Hash objects, where the list is stored (default: `3').
-\end{description}
-
-The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.
+    \item[CA] is a DN of a trusted CA: all certificates issued by this CA are trusted
+    \item[DNsFromAHash] is a list of A-Hash URL's (in \textbf{AHashURL}) from where a list of trusted DN's are periodically obtained (there can be multiple A-Hash URL's listed here in case of one is offline).
+The following attributes are obtained: \textbf{CheckingInterval} specifies how frequently the list from the A-Hash should be updated (default: 600 seconds); \textbf{ID} specifies the GUID where the list is stored (default: `3').
+\end{description}
+
+\note[JKN]{TODO: This paragraph is very unclear.
+All the 'could's makes it look like the TrustManager is just something planed in the far future}The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.
 
@@
 
-Because of the heavy use of X509 certificates, of course, all the services should have their own certificates, which they can use when connecting to other services.
-Currently it is needed to put the paths of the certificate and private key files (and the trusted CAs as well) into the configuration of each services, like this:
+All the services must have their own X.509 certificates, which they use when connecting to other services.
+Currently, it is needed to set the paths of the certificate and private key files (and the trusted CAs as well) in the configuration of each service, like this:
 
@@
 
-    The \emph{IDs} is a list of string \emph{ID}s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs.
+    The \emph{IDs} is a list of string \emph{ID}'s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs.
 If the \emph{neededMetadataList} is empty, then for each \emph{ID} it returns all the \emph{value}s for all the \emph{property} in all the \emph{section} in that given object.
@@
 Even though the default settings are useable, there are some optional variables available for the replicated A-Hash:
-TODO: Put in more optional variables.
 \begin{description}
@@
 The only Librarian method which deals with Logical Names is the \emph{traverseLN} method, which traverses the Logical Name as long as possible, and returns the GUID and metadata of the last found entry in the path of the Logical Name.
-All the other methods work with GUIDs, and do not care about the hierarchical namespace.
+All the other methods work with GUID's, and do not care about the hierarchical namespace.
 In this regard the Librarian can be treated as a higher level application-specifc interface to the metadata store (the A-Hash), because while the A-Hash does not care about what it stores, the Librarian knowa exactly that it stores metadata about files, collections, etc.
@@
 
-    \item [get(getRequestList, neededMetadataList)] returns all or some of the metadata of the the requested GUIDs.
-
-    The \emph{getRequestList} is a list of string \emph{GUID}s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs.
+    \item [get(getRequestList, neededMetadataList)] returns all or some of the metadata of the the requested GUID's.
+
+    The \emph{getRequestList} is a list of string \emph{GUID}'s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs.
 If the \emph{neededMetadataList} is empty, then for each \emph{GUID} it returns all the metadata \emph{value}s for all the \emph{property}s in all the \emph{section}s.
@@
     The \emph{newRequestList} is a list of (\emph{requestID}, \emph{metadataList}) where \emph{requestID} is an arbitrary ID used to identify this request in the list of responses; \emph{metadataList} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
-    This method generates a \emph{GUID} for each request, and inserts the new entry (with the given metadata) into the A-Hash, then returns the GUIDs of the newly created entries.
+    This method generates a \emph{GUID} for each request, and inserts the new entry (with the given metadata) into the A-Hash, then returns the GUID's of the newly created entries.
 In the metadata of the new entry the `type' property in the `entry' section defines whether it is a file, a collection or a mount point.
@@
 
-The Librarian service uses the A-Hash to store all the metadata, and it uses the GUIDs as IDs in the database.
-It has the URL of one or more A-Hash services from its configuration, and it can acquire more A-Hash URLs from these first A-Hashes in case of a replicated A-Hash deployement.
+The Librarian service uses the A-Hash to store all the metadata, and it uses the GUID's as ID's in the database.
+It has the URL of one or more A-Hash services from its configuration, and it can acquire more A-Hash URL's from these first A-Hashes in case of a replicated A-Hash deployement.
 
@@
 The Librarian also accepts report messages from Shepherds, and stores the contents of these messages in the A-Hash.
-There is one A-Hash object which contains the ID and the expected time of the next heartbeat of the registered Shepherds, and there is one A-Hash object for each Shepherd which stores the states and GUIDs of the files on the Shepherd's storage node (see Section~\ref{sub:shepherds} for details).
+There is one A-Hash object which contains the ID and the expected time of the next heartbeat of the registered Shepherds, and there is one A-Hash object for each Shepherd which stores the states and GUID's of the files on the Shepherd's storage node (see Section~\ref{sub:shepherds} for details).
 The Librarian checks periodically if there is a Shepherd which is late with its heartbeat, and it changes the states of all the replicas if needed.
@@
 \begin{description}
-    \item[AHashURL] is the URL of an A-Hash which the Librarian should use. There could be multiple URLs specified here.
+    \item[AHashURL] is the URL of an A-Hash which the Librarian should use. There could be multiple URL's specified here.
     \item[HeartbeatTimeout] specifies in seconds how frequently the Sheperds should send reports to the Librarian.
@@
 A Shepherd service is capable of managing a storage node.
-It keeps track all the files it stores with their GUIDs and checksums.
+It keeps track all the files it stores with their GUID's and checksums.
 It periodically checks each file to detect corruption, and send reports to a Librarian indicating that the storage node is up and running, and whether some file's state has been changed.
-If a file goes missing or has a bad checksum then the Librarian is notified about the error (here the Shepherd refers to the file with its GUID, that's why it needs to store the GUIDs of its files).
+If a file goes missing or has a bad checksum then the Librarian is notified about the error (here the Shepherd refers to the file with its GUID, that's why it needs to store the GUID's of its files).
 It periodically asks the Librarian how many replicas its files have, and if a file has fewer replicas than needed, the Shepherd offers its copy for replication by calling the Bartender.
@@
 A Shepherd service is always connected to a storage element service (e.g.~a web server).
-For each supported storage element service we need a backend module which makes the Shepherd capable of communicating with it to initiate file transfers, to detect whether a transfer was successful or not, to generate local IDs and checksums, etc.
+For each supported storage element service we need a backend module which makes the Shepherd capable of communicating with it to initiate file transfers, to detect whether a transfer was successful or not, to generate local ID's and checksums, etc.
 
@@
 
-These \emph{TURL}s are one-time URLs which means that after the client uploads or downloads the file these \emph{TURL}s cannot be used again to do the same.
-If we want to download the same file twice, we have to initiate the transfer twice, and will get two different \emph{TURL}s.
+These \emph{TURL}'s are one-time URL's which means that after the client uploads or downloads the file these \emph{TURL}'s cannot be used again to do the same.
+If we want to download the same file twice, we have to initiate the transfer twice, and will get two different \emph{TURL}'s.
 
@@
 
-    The \emph{deleteRequestList} is a list of (\emph{requestID}, \emph{referenceID}) pairs containing the IDs of the files to remove.
+    The \emph{deleteRequestList} is a list of (\emph{requestID}, \emph{referenceID}) pairs containing the ID's of the files to remove.
 
@@
 
-    Initiate transfer with \emph{protocol} for the file which has these IDs: \emph{localID} and \emph{referenceID}.
+    Initiate transfer with \emph{protocol} for the file which has these ID's: \emph{localID} and \emph{referenceID}.
 The reason for including here the referenceID as well is that this information could be used by the backend module later, e.g.~when the transfer finished and the state of the file needs to be changed.
@@
 
-    Initiate transfer with \emph{protocol} for the file which has these IDs: \emph{localID} and \emph{referenceID}.
+    Initiate transfer with \emph{protocol} for the file which has these ID's: \emph{localID} and \emph{referenceID}.
 
@@
 
-    \item[list()] returns a list of \emph{localID}s currently in the store directory.
+    \item[list()] returns a list of \emph{localID}'s currently in the store directory.
     \item[getAvailableSpace()] returns the available disk space in bytes.
@@
 \begin{description}
-    \item[ServiceID] is the ID of the service, currently it should be the URL of the service, because there is no information system where we can discover services by IDs.
+    \item[ServiceID] is the ID of the service, currently it should be the URL of the service, because there is no information system where we can discover services by ID's.
     \item[CheckPeriod] specifies in seconds how frequently should the Shepherd check the existence and the checksum of all the files on the storage node.
@@
 The Bartender provides an easy to use interface of the ARC storage system to the users.
-You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}.
+You can put, get and delete files using their logical names (\emph{LN}'s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}.
 You can remove the link to a file or sub-collection from its parent collection without removing the file or sub-collection itself with the \emph{unlink} method.
@@
 
-    The \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name) and in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is `yes' then no more files can be added or removed later).
+    The \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUID's with the given name) and in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is `yes' then no more files can be added or removed later).
 
@@
 It is used to provide the access of the external storage system within the global namespace.
-\emph{makeMountpointRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{URL}, \emph{metadata} ) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUIDs with the given name).
+\emph{makeMountpointRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{URL}, \emph{metadata} ) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUID's with the given name).
 \emph{LN} will be requested Logical name and the \emph{URL} is the path to the external point that is required to access.
@@
 In the first prototype release there is one client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run.
-It is capable of communicating with a given Bartender service, and uploading and downloading TURLs via HTTP.
+It is capable of communicating with a given Bartender service, and uploading and downloading TURL's via HTTP.
 However in order to use proxy certificates or to transfer files with GridFTP protocol, you need to have the ARC client libraries installed.

### 2009-04-11 doc/sec/papers/ISC09/ISC09.tex
Modification for last version

@@
 Mutual authentication is required by GSI, and is the default configuration for GSI based Grid deployment.
-The X.509 certificates are required for both the client and service sides in order to achieve mutual authentication.
-The X.509 certificates are issued by trusted third parties called certificate authorities (CA), and the CAs constitute trust federation which guarantees that two different X.509 certificates from different CAs can accomplish authentication against each other.
-Thus in the GSI based Grid system, if a user would like to access a Grid system, he/she should own a X.509 certificate which is issued by a CA that is trusted by other entities.
 
 In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS) ~\cite{AlfieriR05} have partly solved them, to a different extent.
-The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided by creating the proxy certificate.
-As the most successful Grid authorization model, VOMS adopts the Attribute Certificate~\cite{RFC3821link} approach.
-The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.Attributes include the VO membership data which is associated with the proxy certificate’s identity.
+The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided.
+VOMS adopts the Attribute Certificate~\cite{RFC3821link} approach.
+The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.
+Attributes include the VO membership data which is associated with the proxy certificate’s identity.
 Eventually the resources can make access control decisions based on parsing the VO membership attributes from the Attribute Certificate.
@@
 This goal can be achieved using Web Service Architecture (WSA).
-This way, the Web Service technology has been utilized in Grid computing middlewares.
-The Web Service technology has been adopted by the Globus Toolkit~\cite{GTlink}, gLite~\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)~\cite{ARClink}.
-In fact, utilizing Web Service technology and providing seamless interoperability with other middlewares is one of the main goals of the new version of ARC middleware.
+This way, the Web Service technology has been utilized in Grid computing middlewares, including Globus Toolkit~\cite{GTlink}, gLite~\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)~\cite{ARClink}.
 
@@
 \begin{itemize}
-\item The GSS API based confidential communication has not been adopted by Web Service implementations.
+\item The GSS API based confidential communication and identity delegation has not been adopted by Web Service implementations.
 \item The Attribute Certificate has not been adopted by any aspects Web Service implementation at all.
 \item Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
-\item Finally, identity delegation is a good solution for user or user's representative (acting on behalf of the user) to move/travel from one resource to another.
-But since the identity delegation has been completely coupled with the GSI confidential communication process, it is not possible for the existing identity delegation solution to be used by Web Service applications.
 \end{itemize}
@@
 As the key part of the implementation of new ARC middleware, there is a lightweight Web Service container called Hosting Environment Daemon (HED) which provides hosting for various services at application level, as well as a number of modules to support flexible, interoperable, and efficient communication mechanism for building SOAP based Web Services.
-The whole design of the HED is built around the idea of flexibility and modularity, such that the application service developers can simply concentrate on the application level Web Service implementation by only using the core minimum amount of components.
-It also simplifies work on the middleware level, e.g.
-making possible to implement another communication protocol or authentication mechanism by using the core minimum amount of components and external dependencies.
-A system administrator can easily configure and deploy the middleware and application for different kinds of requirements without having to know much about the implementation.
+The whole design of the HED is built around the idea of flexibility and modularity.
 
@@
 For instance, as shown in the example message flow, the HTTP MCC will process stream from the TLS MCC to parse the HTTP message and pass its body to the SOAP MCC, and also process the SOAP response from the SOAP MCC to generate the HTTP message for the TLS MCC.
-
-Dotted line in Figure \ref{fig:HED} shows an alternative path for the information to propagate among MCCs.
-A service administrator can configure the MCCs according to the interoperability requirements with the other part.
-For instance, the configuration marked with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP message mechanism (see WSE's SoapSender and SoapReceiver)~\cite{WSElink}.
-Another configuration could be SOAP over HTTPg (HTTP over GSI communication protocol) which is needed to interoperate with services like the Storage Resource Manager (SRM)~\cite{AShoshani03} service.
-This shows the flexibility of HED in terms of protocols support.
 
@@
 Each MCC or Service is usually configured with two queues of SecHandler -- one for the incoming message and one for the outgoing message respectively.
-All SecHandler components attached to the queue are executed sequentially.
-If any of them fails, message processing fails as well.
 In Figure \ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules inside MCCs and Services are examples of SecHandler.
-
-As explained above, HED is supposed to act as a Web/Grid Service container, but on the client side, from implementation perspective, there is a similar architecture implemented for processing messages from different protocols and handling security functionality; also there is a specific application programming interface (API) for developers to easily write Web Service client code.
 
@@
 The Shibboleth provides cross-domain single sign-on and attribute-based authorization while preserving user privacy.
-It is based on the OASIS Security Assertion Markup Language (SAML), specifically, the new version of Shibboleth supports SAML 2.0 specification.
+It is based on the OASIS Security Assertion Markup Language (SAML) 2.0 specifications.
 For the authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines two functional components, an Identity Provider and a Service Provider.
 The Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity, while the Service Provider (SP) is responsible for controlling access to services and resources by analyzing the SAML assertions produced and issued by the IdP upon request.
-In order to discover which home community does a user come from, Shibboleth specifies an optional third component called ``Where Are You From?'' (WAYF) service to aid in the process of IdP discovery.
-This IdP discovery process is also standardized and defined in SAML 2.0 specification as ``Identity Provider Discovery Profile''.
 
@@
 \begin{figure}
-\includegraphics[width=0.9\columnwidth]{SAML2SSO.png}
+\includegraphics[width=0.7\columnwidth]{SAML2SSO.png}
 \caption{SAML2.0 SSO profile in ARC}
@@
 The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is ``saml2sp'' which is specific for pointing to the SP Service.
-Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
-\item The SP Service searches the metadata (we use the same metadata format as defined in the Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself).
+Here Identity Provider (IdP) name is sent in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.
+\item The SP Service searches the metadata and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself).
 Then SP Service uses obtained information to compose and issue the SAML authentication request also by utilizing own X.509 certificate (note that in the SAML2.0 SSO profile, X.509 certificates are still needed for IdP and SP) and sends the response back to the user agent.
@@
 The client authentication may be switched off, so that client doesn't need to use any X.509 credential.
-Only the trusted CAs of SP and IdP services need to be configured on the client side to ensure SP and IdP can authentication.
+Only the trusted CAs of SP and IdP services need to be configured on the client side to ensure SP and IdP can be authenticated.
 
@@
 The SLCS client sends request to the SLCS service with the generated X.509 certificate request embedded.
-Upon reception of the request the SLCS service  composes a distinguished name (DN), andissues a certificate (short lived, 12 hours by default) with the SAML attribute (from the SAML2.0 SSO profile) as the X.509 certificate extension and puts the certificate in to the Web Service response.
+Upon reception of the request the SLCS service composes a distinguished name (DN), and issues a certificate (short lived, 12 hours by default) with the SAML attribute (from the SAML2.0 SSO profile) as the X.509 certificate extension and puts the certificate in to the Web Service response.
 The SLCS client gets the response and stores the X.509 certificate into local repository along with corresponding private key.
@@
 \begin{figure}
-\includegraphics[width=0.9\columnwidth]{Delegation_protocol.png}
+\includegraphics[width=0.7\columnwidth]{Delegation_protocol.png}
 \caption{Credential delegation protocol}
@@
 There are three kinds of WS-Security version 1.1 specifications which are completely or partly implemented: Username Token profile, X.509 Token profile, and SAML Token profile.
-
 For those WS clients on behalf of a user, at any step of service invocation, the WS client can augment the SOAP message with WS-Security token by simply configuring the WS-Security SecHandler into container’s configuration, if the relying party (target service) requires.
@@
 When it receives a message from the asserting entity, it will check the SAML assertion based on its predetermined trust relationship with the SAML issuing authority, and check the signature of the SOAP message based on the public key in the SAML assertion without directly trust relationship with attesting entity (holder of this public key).
-In terms of verification, since the proxy certificates have been introduced in IETF (RFC 3820)~\cite{RFC3820link}, and OpenSSL has also implemented RFC3820, it is not difficult to use the proxy certificate for the X.509 Token profile to issue an X.509 token.
 
@@
 
-GSI plug-in for gSOAP~\cite{GAloisio05} adds GSI support for gSOAP Web Service toolkit.
-The GSI MCC implemented in ARC provides samilar functionality.
-Moreover, the ARC solution is more flexible because GSI MCC of both service and client sides is pluggable, which means service and client developers of the ARC do not need to modify the service or client implementation in order to add or replace the GSI functionality.
-
 \section{Conclusion And Future Work}
@@
 It must be noted that due to the standard specifications utilized (SAML and WS-Security) which are independent from any particular middleware, the WS clients and Web/Grid Services developed within the ARC middleware can interoperate with WS clients and Web/Grid Services developed as part of some other middlewares in a standardized way.
-ARC also supports interoperability with Grid middlewares that require GSI based communication.
 
@@
 
-\bibitem{WSElink}
-Web Services Enhancements 2.0 Service Pack 2. http://msdn.microsoft.com/en-us/library/
-
-\bibitem{AShoshani03}
-A. Shoshani, A. Sim, and J. Gu, Storage Resource Managers: Essential Components for the Grid, Grid Resource Management: State of the Art and Future Trends, Kluwer Publishing, 2003.
-
 \bibitem{Shiblink}
@@
 eduPerson and eduOrg Object shema. http://middleware.internet2.edu/eduperson/
-
-\bibitem{RFC3820link}
-Internet X.509 Public Key Infrastructure (PKI) Proxy Certificate Profile. ttp://www.ietf.org/rfc/rfc3820.txt
 
@@
 
-\bibitem{GAloisio05}
-G.
-Aloisio, M.
-Cafaro, I.
-Epicoco, D.
-Lezzi, and R.
-van Engelen, The GSI plug-in for gSOAP: Enhanced Security, Performance, and Reliability, International Conference on Information Technology: Coding and Computing (ITCC'05), IEEE Press, Volume I, pp.
-304-309.
-
 \end{thebibliography}

### 2009-04-14 doc/storage_system/arc-storage-documentation.tex
Went through all of Alex' comments

@@
 
-The different replicas of a file could be in different \note[JKN]{TODO: which states?
-a reference to definition of states maybe?}states, e.g.
-the replica could be valid and alive or just in the process of creation or it could be corrupt or a whole storage node could be offline.
+The different replicas of a file can be in different \note[JKN]{TODO: which states?
+a reference to definition of states maybe?}states.
+The replica can be valid and alive or just in the process of creation or it may be corrupt or a whole storage node may be offline.
 This state is always stored as metadata of the given replica.
@@
     \item The client tool needs to know the URL of a Bartender.
-This could be preconfigured on the client side or in future releases it could be acquired from an information system.
+This can be preconfigured on the client side or, in future releases it may be acquired from an information system.
 When the client tool has the URL, it sends a `list' request which contains the Logical Name \verb!/ourvo/users/user!.
@@
 Because the user wants to put his collection in a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed.
-If the Librarian could traverse the whole LN that would mean that there is already an existing file or collection by that name.
+If the Librarian can traverse the whole LN it means that there is already an existing file or collection by that name.
 If LN does not exist but the the parent collection does, then the parent collection's metadata is returned to the Bartender.
@@
 Each entry has a unique name within a collection and a GUID which points to a corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section (see below).
-The collections could be in a closed state, which means that its contents should not be changed.
+The collections can be in a closed state, which means that its contents should not be changed.
 If you close a collection then it cannot be opened again.
 However it cannot be guaranteed that the contents of a closed collection remains the same (e.g.
-if we include in our collection a file which we don't own, then the owner of the file could remove it nevertheless), but if a closed collection is changed, its state will be broken, and this state could never be changed again, which means you will always know if a closed collection is not intact anymore.
+if we include in our collection a file which we don't own, then the owner of the file can still remove it), but if a closed collection is changed, its state will be broken, and this state can never be changed again, which means you will always know if a closed collection is not intact anymore.
 This state is stored in the \emph{states} section, and its possible values are `\textbf{no}' if the collection is not closed, `\textbf{yes}' if the collection is closed and `\textbf{broken}' if the collection was closed, but something happened, and its contents have been changed.
@@
 
-The internal representation of the access policies is based on access rules. A rule contains an identity and a list of actions. Currently the identity could be one of the following:
+The internal representation of the access policies is based on access rules. A rule contains an identity and a list of actions. Currently the identity can be one of the following:
 \begin{itemize}
@@
 \note[JKN]{TODO: This paragraph is very unclear.
-All the 'could's makes it look like the TrustManager is just something planed in the far future}The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.
+All the 'could's makes it look like the TrustManager is just something planed in the far future.
+Are the attributes mandatory or optional?}The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.
 
@@
 The A-Hash is a database for storing string tuples, and it provides conditional and atomic modification of them.
-The A-Hash can be used as a centralized service, or it could be deployed in a distributed way, using multiple nodes, where all the data is replicated on all the nodes.
+The A-Hash can be used as a centralized service, or it can be deployed in a distributed way, using multiple nodes, where all the data is replicated on all the nodes.
 
 The A-Hash stores \emph{objects}, where each object has an arbitrary string \emph{ID}, and contains any number of \emph{property}-\emph{value} pairs grouped in \emph{section}s, where \emph{property}, \emph{value} and \emph{section} are arbitrary strings.
-There could only be a single \emph{value} for a \emph{property} in a \emph{section}.
-
-If you have an ID, you can get all property-value pairs of the corresponding object with the \emph{get} method, or you could specify only which sections or properties do you need.
-You can add or remove property-value pairs of an object or delete all occurrences of a property or create a new object with the \emph{change} method, and you can specify conditions, which means the change is only applied if the given conditions are met.
+It can only be one \emph{value} for a \emph{property} in a \emph{section}.
+
+Given an ID, the \emph{get} method returns property-value pairs of the corresponding object.
+By specifying a section or property, the \emph{get} method returns the corresponding values.
+Using the \emph{change} method, property-value pairs can be added and removed from an object, all occurrences of a property can be deleted and new objects can be created.
+By specifying conditions, the change is only applied if the given conditions are met.
 
@@
     The \emph{IDs} is a list of string \emph{ID}'s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs.
-If the \emph{neededMetadataList} is empty, then for each \emph{ID} it returns all the \emph{value}s for all the \emph{property} in all the \emph{section} in that given object.
+If the \emph{neededMetadataList} is empty, then for each \emph{ID} it returns all the \emph{values} for all the \emph{properties} in all the \emph{section} in that given object.
 If there are sections and properties specified in \emph{neededMetadataList}, then only those values are returned.
@@
 
-    The \emph{changeRequestList} is a list of \emph{changeRequest}s, where a \emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}, \emph{conditionList}), where \emph{changeID} is an arbitrary ID which is used in the response to refer to this part of the request; \emph{ID} points to the object we want to change; \emph{changeType} can be `\textbf{set}' (to set the property within the section to value), `\textbf{unset}' (to remove the property from the section regardless of the value), `\textbf{delete}' (to remove the whole object); \emph{conditionList} is a list of \emph{condition}s, where a \emph{condition} is a tuple of (\emph{conditionID}, \emph{conditionType}, \emph{section}, \emph{property}, \emph{value}), where \emph{conditionType} could be `\textbf{is}' (which will be true if the property in the section is set to the value), `\textbf{isnot}' (which will be true if the property in the section is not set to the value), `\textbf{isset}' (which will be true if the property of the section is set, regardless of the value), `\textbf{unset}' (which will be true if the property of the section is not set at all).
-    If all conditions are met, tries to apply changes to the objects, creates a new object if a previously non-existent ID is given.
+    The \emph{changeRequestList} is a list of \emph{changeRequest}s, where a \emph{changeRequest} is a tuple of (\emph{changeID}, \emph{ID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}, \emph{conditionList}):
+\begin{itemize}
+\item \emph{changeID} is an arbitrary ID which is used in the response to refer to this part of the request
+\item \emph{ID} points to the object we want to change
+\item \emph{changeType} can be `\textbf{set}' to set the property within the section to value, `\textbf{unset}' to remove the property from the section regardless of the value, `\textbf{delete}' to remove the whole object
+\item \emph{conditionList} is a list of \emph{condition}s, where a \emph{condition} is a tuple of (\emph{conditionID}, \emph{conditionType}, \emph{section}, \emph{property}, \emph{value}), where \emph{conditionType} can be `\textbf{is}' which will be true if the property in the section is set to the value, `\textbf{isnot}' which will be true if the property in the section is not set to the value, `\textbf{isset}' which will be true if the property of the section is set regardless of the value or `\textbf{unset}' which will be true if the property of the section is not set at all.
+\end{itemize}
+    If all conditions are met, the A-Hash tries to apply changes to the objects, and creates a new object if a previously non-existent ID is given.
 
     The response contains the \emph{changeResponseList} which is a list of (\emph{changeID}, \emph{success}, \emph{failedConditionID}) tuples.
-The \emph{success} of the change could be \textbf{set}, \textbf{unset}, \textbf{deleted}, \textbf{failed}, \textbf{condition not met} (in this case the ID of the failed condition is put into \emph{failedConditionID}), \textbf{invalid change type} or \textbf{unknown}.
+The \emph{success} of the change is either \textbf{set}, \textbf{unset}, \textbf{deleted}, \textbf{failed}, \textbf{condition not met} (in which case the ID of the failed condition is put into \emph{failedConditionID}), \textbf{invalid change type} or \textbf{unknown}.
 
@@
 
-The A-Hash service has a modular architecture which means that you can decide if you want to deploy it a centralized or a distributed way by simply specifying different modules in the service's configuration.
-There are different low-level modules for storing the data on disk, e.g.
-serialized into `pickle' format.
-And there is a module which stores the data replicated on multiple nodes.
+The A-Hash service has a modular architecture which means that it can be deployed in a centralized or a distributed way by simply specifying different modules in the service's configuration.
+There are different low-level modules for storing the metadata on disk, e.g.
+serialized into the `pickle' format.
+In addition there is a module which stores the metadata replicated on multiple nodes.
 
@@
 The actual replication is managed by the \verb!ReplicationStore!
-and \verb!ReplicationManager!, meaning that the \verb!ReplicatedAHash!, with the exception of the above mentioned methods, is actually just a centralized A-Hash with a specific storage module.
+and \verb!ReplicationManager!, meaning that the \verb!ReplicatedAHash!, with the exception of the above-mentioned methods, is actually just a centralized A-Hash with a specific storage module.
     \item[ReplicationStore] 
-	is a specialized replicated DB.
+	is a specialized replicated database.
 It is a subclass of \verb!TransDBStore!\footnote{A module for storing data on disk transactionally using the Berkeley DB library.} with some additional features for handling replication.
 The \verb!ReplicationStore!
-periodically updates an A-Hash object containing information about the current master A-Hash and available clients.
+periodically updates an A-Hash object which contains information about the current master A-Hash and available clients.
 In addition it contains a \verb!ReplicationManager!
@@
     \item[ReplicationManager] 
-	handles all aspects of the Berkeley DB API and takes care of replicas, elections and replication messages.
+	handles all aspects of the Berkeley DB API and takes care of replicas, elections of the master replica and replication messages.
 	\begin{itemize}
@@
 This means that the entire database is fully replicated on all replicas.
-When the master receives a write message, it will broadcast this message to all clients.
-To make sure that the db is consistent and that there will always be possible to elect a master, the master requires an acknowledgement message from a quorum of clients before continuing.
+When the master receives a write message, it will broadcast this message to all the clients.
+To make sure that the database is consistent and that it will always be possible to elect a master, the master requires an acknowledgement message from a quorum of clients before continuing.
 	\item In the replicated A-Hash there is one master and multiple clients.
-When the A-Hash starts up or if a master goes offline, we need to find a new master.
-This is solved by the clients holding an election, according to the Paxos algorithm.
-Simply told, all clients sends a vote to all other clients, and the eligible client that gets the most votes wins the election.
+When the A-Hash starts up or if a master goes offline, a new master must be elected.
+This is solved by the clients holding an election, according to the Paxos algorithm\footnote{http://research.microsoft.com/users/lamport/pubs/pubs.html\#lamport-paxos}.
+Simply told, all clients send a vote to all other clients, and the eligible client that gets the most votes wins the election.
 An eligible client is a client which has received the latest updates.
-(See the Berkeley DB documentation for more details.) There are cases where some servers are more suitable to be master than others (better hardware, faster Ethernet, etc.).
-In such cases you can influence the election by setting a higher priority on the best servers.
-While an election is being held, it is not possible to write or read from the A-Hash, as consistency cannot be guaranteed without a master.
-	\item As the replicas in the replicated A-Hash are services running within HED, the replication messages needs to be sent through the communication framework of HED.
+See the Berkeley DB documentation for more details.
+There are cases where some servers are more suitable to be master than others (better hardware, faster Ethernet, etc.).
+In such cases the election can be influenced by setting a higher priority on the best servers.
+While an election is being held, it is not possible to write to or read from the A-Hash, as consistency cannot be guaranteed without a master.
+	\item As the replicas in the replicated A-Hash are services running within HED, the replication messages need to be sent through the communication framework of HED.
 To be able to use this framework, the Berkeley DB needs a callback function for sending messages and the \verb!ReplicationManager!
@@
 which in turn will call the Berkeley DB processMessage.
-This communication framework is both used by Berkeley DB and directly by the \verb!ReplicationManager!
-for, e.g., bootstrapping the A-Hash replicas and to communicate newly discovered clients.
+This communication framework is both used by Berkeley DB and directly by the \verb!ReplicationManager!, e.g., for bootstrapping the A-Hash replicas and to communicate newly discovered clients.
 \end{itemize}
@@
 \subsection{Configuration} % (fold)
+\label{sub:a_hash_configuration}
 
@@
 is a replicated version.
-    \item[StoreClass] specifies the name of the store class.
-This class is used to store the data on disk.
-Currently there are several versions (\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore!
-is suitable for using in a replicated deployment, all the others are only for a centralized deployement.
+    \item[StoreClass] specifies the name of the store class, used to store the data on disk.
+Currently there are several store classes \note[JKN]{TODO: describe the stores and how to choose among them}(\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore!
+class is suitable for use in a replicated deployment.
+All the others are only for a centralized deployement.
 This variable is only configurable for the centralized deployement.
-    \item[StoreCfg] could contain some parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash should save its files.
-This variable is only configurable for the centralized deployement.
-\end{description}
-
-For the replicated A-Hash there are some additional required configuration variables:
+    \item[StoreCfg] contains parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash can save its files.
+This variable is only configurable for the centralized deployment.
+\end{description}
+
+For the replicated A-Hash there are some additional mandatory configuration variables:
 
@@
     \item[LocalDir] specifies a directory on the local filesystem where this A-Hash replica should be stored.
-This directory must be unique for this replica, i.e., no two A-Hash replicas can share the same \verb!LocalDir!, and doing so will most likely leave both replicas broken beyond repair.
+This directory must be unique for this replica, i.e., no two A-Hash replicas can share the same \verb!LocalDir!
+since doing so will most likely leave both replicas broken beyond repair.
     \item[MyURL] is the URL to this A-Hash service and is used as an ID for this service in the replication group.
@@
     \item[Priority] is related to the master election.
-If a master dies, there will be an election to decide who will be the new master.
+If a master dies, there will be an election to decide which client will be the new master.
 Higher priority means a higher chance to be elected.
 A priority of 0 means that this replica can never be elected.
-Defaults to 10.
-    \item[CheckPeriod] specifies in seconds how frequently the A-Hash should update the replica list. Defaults to 10 seconds.
-    \item[CacheSize] configures the amount of the storage to be cached in memory. Defaults to 10MB
-\end{description}
-
-An example configuration:
+The defaults is 10.
+    \item[CheckPeriod] specifies in seconds how frequently the A-Hash should update the replica list. The default is 10 seconds.
+    \item[CacheSize] configures the amount of the storage to be cached in memory. The default is 10MB
+\end{description}
+
+Example of a centralized A-Hash configuration:
 
@@
 
-Example of replicated A-Hash:
+Example of a replicated A-Hash configuration:
 \begin{verbatim}
@@
 
-\section{Librarians} % (fold)
+\section{Librarian} % (fold)
 \label{sec:librarian}
@@
 
-The Librarian knows about the type of possible entries in the namespace of the ARC storage system: files, collections and mount points.
-It knows that these are organized in a tree-hiearchy with grouping them into collections, and it knows about Logical Names, and how to traverse a Logical Name to find the metadata of the file or collection (or mount point) which it refers to.
-However, the Librarian does not maintain the namespace of the ARC storage, e.g.
-it does not put new files into collections, or remove the link from its parent collection when a file is removed - this is the job of the Bartender.
-The only Librarian method which deals with Logical Names is the \emph{traverseLN} method, which traverses the Logical Name as long as possible, and returns the GUID and metadata of the last found entry in the path of the Logical Name.
+The Librarian knows about the type of possible entries in the namespace of Chelonia: files, collections and mount points.
+It knows that these are organized in a tree-hiearchy with entries grouped into collections, and it knows about Logical Names, and how to traverse a Logical Name to find the metadata of the file or collection (or mount point) which it refers to.
+However, the Librarian does not maintain the namespace of Chelonia.
+It does not put new files into collections, nor remove the link from its parent collection when a file is removed - this is the job of the Bartender.
+The only Librarian method which deals with Logical Names is the \emph{traverseLN} method, which traverses the Logical Name as far as possible, and returns the GUID and metadata of the last found entry in the path of the Logical Name.
 All the other methods work with GUID's, and do not care about the hierarchical namespace.
-In this regard the Librarian can be treated as a higher level application-specifc interface to the metadata store (the A-Hash), because while the A-Hash does not care about what it stores, the Librarian knowa exactly that it stores metadata about files, collections, etc.
-With this interface you can create new files, collections and mount points with the \emph{new} method, you can get or modify the metadata of any entry with the \emph{get} and the \emph{modifyMetadata} methods, and you can remove an entry with the \emph{remove} method.
-
-Besides being an interface to the metadata store, the Librarian has an other importat functionality: it provides a way for Shepherds to register themselves, and to periodically send reports about the state of their stored files.
-A Shepherd send a \emph{report} message to a Librarian if e.g.
+In this regard the Librarian can be treated as a higher level application-specifc interface to the metadata store (the A-Hash), because while the A-Hash does not care about what it stores, the Librarian knows exactly that it stores metadata about files, collections, etc.
+With this interface new files, collections and mount points can be created with the \emph{new} method, the metadata of any entry can be retrieved and modified with the \emph{get} and \emph{modifyMetadata} methods, and entries can be removed with the \emph{remove} method.
+
+Besides being an interface to the metadata store, the Librarian has another important functionality: it provides a way for Shepherds to register themselves and to periodically send reports about the state of their stored files.
+A Shepherd sends a \emph{report} message to a Librarian if e.g.
 a new file arrived, or an old file gets corrupted, thus the Librarian can keep the states of the replicas up-to-date.
-The Librarian monitors the registered Shepherds, and if one of them stops sending reports, the Librarian will assume, that it is offline, and modify the states of the all the replicas which are on the Shepherd's storage node.
+The Librarian monitors the registered Shepherds, and if one of them stops sending reports, the Librarian will assume, that it is offline, and modify the states of all the replicas which are on the Shepherd's storage node.
 In order to do this, the Librarian stores for each Shepherd a list of files on the given storage nodes, this make it possible to know which files have replicas on a given storage nodes without checking all the files in the system.
@@
 
-    \item [get(getRequestList, neededMetadataList)] returns all or some of the metadata of the the requested GUID's.
+    \item [get(getRequestList, neededMetadataList)] returns all or some of the metadata of the requested GUID's.
 
     The \emph{getRequestList} is a list of string \emph{GUID}'s, \emph{neededMetadataList} is a list of (\emph{section}, \emph{property}) pairs.
-If the \emph{neededMetadataList} is empty, then for each \emph{GUID} it returns all the metadata \emph{value}s for all the \emph{property}s in all the \emph{section}s.
+If \emph{neededMetadataList} is empty, then for each \emph{GUID} all the metadata values for all the properties in all the sections are returned.
 If there are sections and properties specified in \emph{neededMetadataList}, then only those values are returned.
 
-    The response contains a list of \emph{GUID, metadatList} pairs, where \emph{metadataList} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
-
-    \item[new(newRequestList)] creates a new entry in the Librarian with the given metadata, and returns the GUID of the new entry.
+    The response contains a list of \emph{GUID, metadataList} pairs, where \emph{metadataList} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+
+    \item[new(newRequestList)] creates a new entry in the Librarian with the given metadata, and returns the GUID's of the new entries.
 
@@
 
-    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) tuples where \emph{changeType} can be `\textbf{set}' (set the property in the section to value), `\textbf{unset}' (remove the property-value pair from the section), `\textbf{add}' (set the property in the section to value only if it does not exist already) or `\textbf{setifvalue}=value' (only set the property in the section to value if it currently equals to the `value' in the \emph{changeType}).
-
-    The response is a list of (\emph{changeID}, \emph{success}) where \emph{success} can be `\textbf{set}', `\textbf{unset}', `\textbf{condition failed}', `\textbf{failed}: reason'.
+    \emph{modifyMetadataRequestList} is a list of (\emph{changeID}, \emph{GUID}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) tuples where \emph{changeType} can be `\textbf{set}' to set the property in the section to value, `\textbf{unset}' to remove the property-value pair from the section, `\textbf{add}' to set the property in the section to \emph{value} only if it does not exist already or `\textbf{setifvalue}=value' to only set the property in the section to value if it currently equals to the `value' in the \emph{changeType}.
+
+    The response is a list of (\emph{changeID}, \emph{success}) pairs where \emph{success} can be `\textbf{set}', `\textbf{unset}', `\textbf{condition failed}', `\textbf{failed}: reason'.
 
@@
 
-    The response is a list of (\emph{requestID}, \emph{success}) pairs where \emph{success} could be `\textbf{removed}' or `\textbf{failed}: reason'.
+    The response is a list of (\emph{requestID}, \emph{success}) pairs where \emph{success} is either `\textbf{removed}' or `\textbf{failed}: reason'.
 
@@
 
-    The \emph{traverseRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names to be traversed
-
-    The response is a list of (\emph{requestID}, \emph{traversedList}, \emph{wasComplete}, \emph{traversedLN}, \emph{GUID}, \emph{metadataList}, \emph{restLN}) where:
+    The \emph{traverseRequestList} is a list of (\emph{requestID}, \emph{LN}) pairs with the Logical Names to be traversed
+
+    The response is a list of (\emph{requestID}, \emph{traversedList}, \emph{wasComplete}, \emph{traversedLN}, \emph{GUID}, \emph{metadataList}, \emph{restLN}) tuples where:
     \begin{description}
@@
         \item[metadataList] is all the metadata of the of traversedLN in the form of (\emph{section}, \emph{property}, \emph{value}) tuples
-        \item[restLN] is the postfix of the \emph{LN} which was not traversed for some reason, if \emph{wasComplete} is true, this should be an empty string
-    \end{description}
-
-    \item [report(serviceID, filelist)] is a report message from a Shepherd to a Librarian, contains the ID of the Shepherd (currently the URL), and a list of changed files on the Shepherd's storage node.
-
-    The \emph{filelist} is a list of (\emph{GUID}, \emph{referenceID}, \emph{state}) tuples containing the state of changed or new files, where referenceID is the Sheperd-local ID of the given replica where GUID refers to the logical file of this replica.
-The \emph{state} could be `\textbf{invalid}' (if the periodic self-check of the Shepherd found a non-matching checksum or missing file), `\textbf{creating}' (if this is a new file not uploaded yet), `\textbf{thirdwheel}' (if the replica is marked for deletion because of too many replicas), `\textbf{alive}' (if the file is uploaded and the checksum is OK), or `\textbf{stalled}' (if the file was in `creating' state for a long time but hasn't arrived).
-
-    The response is \emph{nextReportTime}, a number of seconds, which is the timeframe within the Librarian expects the next heartbeat from the Shepherd.
+        \item[restLN] is the remainders of the \emph{LN} which was not traversed for some reason, if \emph{wasComplete} is true, this should be an empty string
+    \end{description}
+
+    \item [report(serviceID, filelist)] is a report message from a Shepherd to a Librarian which contains the ID of the Shepherd, and a list of changed files on the Shepherd's storage node.
+
+    The \emph{filelist} is a list of (\emph{GUID}, \emph{referenceID}, \emph{state}) tuples containing the state of changed or new files, where referenceID is the Sheperd-local ID of the given replica and GUID refers to the logical file of this replica.
+The \emph{state} is one of `\textbf{invalid}' (if the periodic self-check of the Shepherd found a non-matching checksum or missing file), `\textbf{creating}' (if this is a new file not uploaded yet), `\textbf{thirdwheel}' (if the replica is marked for deletion because of too many replicas), `\textbf{alive}' (if the file is uploaded and the checksum is OK), or `\textbf{stalled}' (if the file was in `creating' state for a long time but hasn't arrived).
+
+    The response is \emph{nextReportTime}, a number of seconds, which is the timeframe within which the Librarian expects to receive the next heartbeat from the Shepherd.
 
@@
 The Librarian service uses the A-Hash to store all the metadata, and it uses the GUID's as ID's in the database.
-It has the URL of one or more A-Hash services from its configuration, and it can acquire more A-Hash URL's from these first A-Hashes in case of a replicated A-Hash deployement.
+It has the URL of one or more A-Hash services in its configuration, and it can acquire more A-Hash URL's from these first A-Hashes in case of a replicated A-Hash deployment.
 
@@
 A list of A-Hashes, master and clients, is stored in an A-Hash object.
-The Librarian periodically gets this list to maintain its own list of A-Hash services.
+The Librarian periodically retrieves this list to maintain its own list of A-Hash services.
 If the Librarian tries to write to the replicated A-Hash and the master is down, the Librarian will try to get a new list from one of the clients and retry the write request once before giving up.
@@
 Because the Librarian services store everything in the A-Hash, the Librarian itself is a stateless service.
-We can deploy multiple Librarians, if we configure them to use the same replicated group of A-Hashes (or the same central A-Hash) then the Bartenders could use any Librarian to achieve the same results, and the Shepherds also could report to any of them.
+Multiple Librarians may be deployed.
+If they are configured to use the same replicated group of A-Hashes (or the same central A-Hash) then the Bartenders can use any Librarian to achieve the same results, and the Shepherds can report to any of them.
 One Shepherd should only report to one Librarian, but if that Librarian is offline, it should find another.
-The reports will be registered in the A-Hash as well, so it does not matter which Librarian gets a report.
+The reports will be registered in the A-Hash as well, so it does not matter which Librarian receives a report.
 
@@
 
-The Librarian has these configuration variables:
-
-\begin{description}
-    \item[AHashURL] is the URL of an A-Hash which the Librarian should use. There could be multiple URL's specified here.
+The Librarian has the following configuration variables:
+
+\begin{description}
+    \item[AHashURL] is the URL of an A-Hash which the Librarian should use. Multiple A-Hashes may be specified.
     \item[HeartbeatTimeout] specifies in seconds how frequently the Sheperds should send reports to the Librarian.
@@
 
-\section{Shepherds} % (fold)
+\section{Shepherd} % (fold)
 \label{sec:shepherds}
@@
 A Shepherd service is capable of managing a storage node.
-It keeps track all the files it stores with their GUID's and checksums.
-It periodically checks each file to detect corruption, and send reports to a Librarian indicating that the storage node is up and running, and whether some file's state has been changed.
-If a file goes missing or has a bad checksum then the Librarian is notified about the error (here the Shepherd refers to the file with its GUID, that's why it needs to store the GUID's of its files).
+It keeps track of all the files it stores with their GUID's and checksums.
+It periodically checks each file to detect corruptions, and sends reports to a Librarian indicating that the storage node is up and running, and whether the state of any files have been changed.
+If a file goes missing or has a bad checksum then the Librarian is notified about the error\footnote{Here the Shepherd refers to the file with its GUID, that's why it needs to store the GUID's of its files.}.
 It periodically asks the Librarian how many replicas its files have, and if a file has fewer replicas than needed, the Shepherd offers its copy for replication by calling the Bartender.
 
-A Shepherd service is always connected to a storage element service (e.g.~a web server).
-For each supported storage element service we need a backend module which makes the Shepherd capable of communicating with it to initiate file transfers, to detect whether a transfer was successful or not, to generate local ID's and checksums, etc.
-
-A file in a storage node could be identified with a \emph{referenceID} which is unique within that node.
-If we know the \emph{location} of a file, which is the ID of the Shepherd service (\emph{serviceID}, currently the URL of the Shepherd) and the \emph{referenceID}, we could call the Shepherds \emph{get} method with the \emph{referenceID} and a list of transfer protocols we can use, the Shepherd chooses a protocol from this list which it can provide, and create a transfer URL (\emph{TURL}) and returns it along with the \emph{checksum} of the file.
-We could download the file from this \emph{TURL}, and verify it with the \emph{checksum}.
-An end user of the storage system does not need to call this \emph{get} method, because the Bartender service will do it, the user just asks the Bartender and gets the TURL.
-
-Storing a file starts with initiating the transfer with the \emph{put} method of the Shepherd, we should give the \emph{size} and \emph{checksum} of the file and its \emph{GUID} as well.
-We also specify a list of transfer protocols we are able to use, and the Shepherd chooses a \emph{protocol}, creates a \emph{TURL} for uploading and generates a \emph{referenceID}, then we can upload the file to the TURL.
-Again, the end user just asks the Bartender, and gets the TURL, the user does not need to call the \emph{put} method of the Shepherd directly.
-
-These \emph{TURL}'s are one-time URL's which means that after the client uploads or downloads the file these \emph{TURL}'s cannot be used again to do the same.
-If we want to download the same file twice, we have to initiate the transfer twice, and will get two different \emph{TURL}'s.
-
-With the \emph{stat} method we can get some information about a replica, e.g.~checksum, GUID, state, etc. The \emph{delete} method removes the replica.
-
-In normal operation the \emph{put} and \emph{get} calls is made by a Bartender but the actual uploading and downloading is done by the user's client.
-In the case of replication a Shepherd with a valid replica initiates the replication, this Shepherd asks the Bartender to choose a new Shepherd, the Bartender initiates putting the new replica on a chosen Shepherd and receives the TURL, then the Bartender returns the TURL to the initiator Shepherd, which uploads its replica to the given TURL.
-
-The replicas have a state, which could be `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, `\textbf{invalid}' if it does not exists anymore or has a bad checksum, `\textbf{thirdwheel}' if the replica is marked for deletion because the file has too many replicas, or `\textbf{stalled}' if it was in a `creating' state for a long time but hasn't arrived.
+A Shepherd service is always connected to a storage element service (e.g., a web server).
+For each supported storage element service there is a backend module which makes the Shepherd capable of communicating with the storage element to initiate file transfers, to detect whether a transfer was successful or not, to generate local ID's and checksums and so forth.
+
+A file in a storage node is identified with a \emph{referenceID} which is unique within that node.
+The \emph{location} of a file consists of the \emph{serviceID} of the Shepherd and the \emph{referenceID}.
+If the location is known, the Shepherds \emph{get} method can be called with the \emph{referenceID} and a list of transfer protocols, the Shepherd chooses a protocol from this list, and creates a transfer URL (\emph{TURL}).
+This TURL is returned together with the \emph{checksum} of the file.
+A client tool may then download the file from this \emph{TURL}, and verify it with the \emph{checksum}.
+The client tool does not need to call the \emph{get} method.
+The client tool simply queries the Bartender, which in turn calls the \emph{get} method and returns the TURL.
+
+Storing a file starts with initiating the transfer with the \emph{put} method of the Shepherd, providing the file \emph{size}, the \emph{checksum} of the file and its \emph{GUID}.
+The client tool also specifies a list of transfer protocols it is able to use, and the Shepherd chooses a \emph{protocol}, creates a \emph{TURL} for uploading and generates a \emph{referenceID}, so that the client tool can upload the file to the TURL.
+Again, the client tool just asks the Bartender, and gets the TURL; the client tool does not need to call the \emph{put} method of the Shepherd directly.
+
+These \emph{TURL}'s are one-time URL's which means that after the client tool uploads or downloads the file these \emph{TURL}'s cannot be used again to do the same.
+If the same file is to be downloaded twice, the transfer has to be initiated twice, yielding in two different \emph{TURL}'s.
+
+The \emph{stat} method provides information about a replica, e.g., checksum, GUID, state, etc. The \emph{delete} method removes the replica.
+
+In normal operation the \emph{put} and \emph{get} calls are made by a Bartender but the actual upload and download is done by the client tool.
+In the case of replication a Shepherd with a valid replica initiates the replication.
+This Shepherd asks the Bartender to choose a new Shepherd, the Bartender initiates putting the new replica on a chosen Shepherd and receives the TURL, and the Bartender returns the TURL to the initiator Shepherd, which uploads its replica to the given TURL.
+
+The replicas have a state, which is one of `\textbf{creating}' when the transfer is initiated but the file is not uploaded yet, `\textbf{alive}' if the file is uploaded and has a proper checksum, `\textbf{invalid}' if it does not exists anymore or has a bad checksum, `\textbf{thirdwheel}' if the replica is marked for deletion because the file has too many replicas, or `\textbf{stalled}' if it has been in the `creating' state for a long time but hasn't arrived.
 
@@
 
-    The \emph{getRequestList} is a list of (\emph{requestID}, \emph{getRequestData}) where \emph{requestID} is an arbitrary ID used in the reply, \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use (there could be multiple protocols in getRequestData).
-
-    The response is a list of (\emph{requestID}, \emph{getResponseData}), where \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, such as: `\textbf{TURL}' is a transfer URL which can be used by the client to download the file; `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica; `\textbf{checksumType}' is the name of the checksum method and `\textbf{error}' could contain an error message if there is one.
-
-    \item[put(putRequestList)] initiates an upload and returns the TURL
-
-    The \emph{putRequestList} is a list of (\emph{requestID}, \emph{putRequestData}) where \emph{requestID} is an ID used for the response,     \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs such as `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (which is currently not used).
-
-    The response is a list of (\emph{requestID}, \emph{putResponseData}), where \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs such as: `\textbf{TURL}' is the transfer URL where the client can upload the file, `\textbf{protocol}' is the chosen protocol of the TURL and `\textbf{referenceID}' is the generated ID for this new replica, `\textbf{error}' could contain an error message.
-
+    The \emph{getRequestList} is a list of (\emph{requestID}, \emph{getRequestData}) pairs where \emph{requestID} is an arbitrary ID used in the reply, \emph{getRequestData} is a list of (\emph{property}, \emph{value}) pairs, where mandatory properties are: `\textbf{referenceID}' which refers to the file to get and `\textbf{protocol}' indicates a protocol the client can use.
+getRequestData may contain multiple protocols.
+
+    The response is a list of (\emph{requestID}, \emph{getResponseData}), where \emph{getResponseData} is a list of (\emph{property}, \emph{value}) pairs, where \emph{property} may be one of the following:
+    \begin{itemize}
+      \item `\textbf{TURL}' is a transfer URL which can be used by the client to download the file
+      \item `\textbf{protocol}' is the protocol of the TURL; `\textbf{checksum}' is the checksum of the replica
+      \item `\textbf{checksumType}' is the name of the checksum method
+      \item `\textbf{error}' contains an error message if there is one.
+    \end{itemize}
+    \item[put(putRequestList)] initiates an upload and returns the TURL.
+
+    \emph{putRequestList} is a list of (\emph{requestID}, \emph{putRequestData}) pairs where \emph{requestID} is an ID used for the response,     \emph{putRequestData} is a list of (\emph{property}, \emph{value}) pairs where property can be one of `\textbf{GUID}', `\textbf{checksum}', `\textbf{checksumType}', `\textbf{size}' (the size of the file in bytes), `\textbf{protocol}' (a protocol the client can use, can be multiple) and `\textbf{acl}' (which is currently not used).
+
+    The response is a list of (\emph{requestID}, \emph{putResponseData}) pairs, where \emph{putResponseData} is a list of (\emph{property}, \emph{value}) pairs where \emph{property} may be one of the following:
+    \begin{itemize}
+      \item `\textbf{TURL}' is the transfer URL where the client can upload the file
+      \item `\textbf{protocol}' is the chosen protocol of the TURL
+      \item `\textbf{referenceID}' is the generated ID for this new replica
+      \item `\textbf{error}' contains an error message.
+    \end{itemize}
     \item[delete(deleteRequestList)] removes a replica.
@@
 
-    The response is a list of (\emph{requestID}, \emph{status}), where \emph{status} could be `\textbf{deleted}' or `\textbf{nosuchfile}'.
+    The response is a list of (\emph{requestID}, \emph{status}) pairs, where \emph{status} can be either `\textbf{deleted}' or `\textbf{nosuchfile}'.
     .
@@
 
-    The \emph{statRequestList} is a list of (\emph{requestID}, \emph{referenceID}) where \emph{referenceID} points to the file whose data we want to get.
-
-    The response is a list of (\emph{requestID}, \emph{referenceID}, \emph{state}, \emph{checksumType}, \emph{checksum}, \emph{acl}, \emph{size}, \emph{GUID}, \emph{localID})    
+    \emph{statRequestList} is a list of (\emph{requestID}, \emph{referenceID}) pairs where \emph{referenceID} points to the file whose metadata are requested.
+
+    The response is a list of (\emph{requestID}, \emph{referenceID}, \emph{state}, \emph{checksumType}, \emph{checksum}, \emph{acl}, \emph{size}, \emph{GUID}, \emph{localID}) tuples.
 \end{description}
@@
 
-The Shepherd is communicating with the storage element services via backend modules.
-Currently there are two backend modules implemented, one for the \emph{Hopi} service (which is a simple HED-based HTTP server), one for the \emph{apache} webserver.
+The Shepherd communicates with storage element services via backend modules.
+Currently there are two backend modules implemented, one for the \emph{Hopi} service\footnote{Note that the Hopi service needs to be configured in a special `slave' mode to be used as a backend for the Shepherd.} (which is a simple HED-based HTTP server), one for the \emph{Apache} webserver\footnote{\url{http://www.apache.org/}}.
 
@@
 
-Let's see the scenario for the Hopi service which should be in a special `slave' mode for this kind of operation: if a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), then the Shepherd service creates a hardlink into the transfer directory (e.g.~\verb!./transfer/abc!) and sets this file read-only.
-If the Hopi service is configured that way that it handles the HTTP path \verb!/prb!
+If a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), the Shepherd service creates a hardlink into the transfer directory (\verb!./transfer/abc!) and sets this file read-only.
+If the Hopi service is configured to serve the HTTP path \verb!/prb!
 and it is serving files from the directory \verb!./transfer!
-then after the hardlink is created, we have this URL for this file: \verb!http://localhost:60000/prb/abc!.
-Now we can give this URL to the client.
-Then the client \verb!GET!s this URL and gets the file.
+then after the hardlink is created, the transfer URL for this file reads \verb!http://localhost:60000/prb/abc!.
+Now this URL is passed to the client.
+Then the client \verb!GET!'s this URL and gets the file.
 The Hopi service removes (unlinks) this file immediately after the \verb!GET!
-request arrived, which makes this \verb!http://localhost:60000/prb/abc!
-URL invalid (so this is a one-time URL), but because of the hardlink the file is still there in the store directory, it is just removed from the transfer directory.
-Now if some other user wants this file, the Shepherd creates an other hardlink, e.g.~\verb!./transfer/qwe!
-and now we have an URL \verb!http://localhost:60000/prb/qwe!.
-
-If a client wants to upload a new file, then the Shepherd creates an empty file in the store directory, e.g.~\verb!./store/file2!
-and creates a hardlink into the transfer directory, e.g.~\verb!./transfer/oiu!
-and makes it writable, and now we have a URL \verb!http://localhost:60000/prb/oiu!, and the client is able to do a \verb!PUT!
-to this URL.
-When the client \verb!PUT!s the file there, the Hopi service immediately removes the uploaded file from the transfer directory, but because it has a hardlink in the store directory, the file is stored there as \verb!./store/file2!.
-The backend module for the Hopi service periodically checks whether a new file has two or just one hard links.
-If it has only one that means that a file is uploaded, so it could notify the Shepherd that the file is arrived.
-In order to do that, all the backend modules get a callback method `file\_arrived' from the Shepherd.
-
-All the backend modules should have this common interface which the Shepherd can use to communicate with the storage element services:
-
-\begin{description}
-    \item[prepareToGet(referenceID, localID, protocol)] returns the \emph{TURL}.
-
-    Initiate transfer with \emph{protocol} for the file which has these ID's: \emph{localID} and \emph{referenceID}.
-The reason for including here the referenceID as well is that this information could be used by the backend module later, e.g.~when the transfer finished and the state of the file needs to be changed.
-
-    \item[prepareToPut(referenceID, localID, protocol)] returns the \emph{TURL}.
-
-    Initiate transfer with \emph{protocol} for the file which has these ID's: \emph{localID} and \emph{referenceID}.
+request arrived, which makes \verb!http://localhost:60000/prb/abc!
+invalid (so this is a one-time URL).
+However, because of the hardlink the file is still in the store directory, it is just removed from the transfer directory.
+Now if some other user wants this file, the Shepherd creates an other hardlink, e.g., \verb!./transfer/qwe!
+with the URL \verb!http://localhost:60000/prb/qwe!.
+
+If a client wants to upload a new file, the Shepherd creates an empty file in the store directory, e.g.~\verb!./store/file2!
+and creates a hardlink in the transfer directory, e.g., \verb!./transfer/oiu!
+and makes it writable.
+The transfer URL is then \verb!http://localhost:60000/prb/oiu!, and the client can do a \verb!PUT!
+on this URL.
+When the client tool \verb!PUT!s the file there, the Hopi service immediately removes the uploaded file from the transfer directory, but because it has a hardlink in the store directory, the file is stored there as \verb!./store/file2!.
+The backend module for the Hopi service periodically checks whether a new file has two hardlinks  or just.
+If it has only one hardlink that means that the file is uploaded, so it will notify the Shepherd that the file has arrived.
+In order to do that, the Shepherd needs to provide a callback method `file\_arrived' to the backend module.
+
+All the backend modules should have the following common interface which the Shepherd can use to communicate with the storage element services:
+
+\begin{description}
+    \item[prepareToGet(referenceID, localID, protocol)] returns a one-time \emph{TURL}.
+
+    Initialize transfer with \emph{protocol} for the file identified by (\emph{localID}, \emph{referenceID}).
+The reason for including here the referenceID as well is that this information may be used by the backend module later, e.g., when the transfer has finished and the state of the file needs to be changed.
+
+    \item[prepareToPut(referenceID, localID, protocol)] returns a one-time \emph{TURL}.
+
+    Initialize transfer with \emph{protocol} for the file identified by (\emph{localID}, \emph{referenceID}).
 
@@
     \item[generateLocalID()] returns a new unique \emph{localID}.
-    \item[matchProtocols(protocols)] only leave that protocols in the list \emph{protocols} which are supported by this file transfer service.
+    \item[matchProtocols(protocols)] only leave those protocols in the list \emph{protocols} which are supported by this file transfer service.
     \item[checksum(localID, checksumType)] generates a \emph{checksumType} checksum of the file referenced by \emph{localID}.
@@
 
-The Shepherd has these configuration variables:
-
-\begin{description}
-    \item[ServiceID] is the ID of the service, currently it should be the URL of the service, because there is no information system where we can discover services by ID's.
-    \item[CheckPeriod] specifies in seconds how frequently should the Shepherd check the existence and the checksum of all the files on the storage node.
-    \item[MinCheckInterval] specifies in seconds how much time should the Shepherd wait between two subsequent file checks.
+The Shepherd has the following configuration variables:
+
+\begin{description}
+    \item[ServiceID] is the ID of the service, currently it should be the URL of the service, because there is no information system where the services can be discovered by ID's.
+    \item[CheckPeriod] specifies in seconds how frequently the Shepherd should check the existence and checksum of all the files on the storage node.
+    \item[MinCheckInterval] specifies in seconds how much time the should Shepherd wait between two subsequent file checks.
 This ensures that the Shepherd does not use up all the resources of the hosting machine.
-After checking the cheksum of a file, the Shepherd will always wait this long before checking the next file, even if this means that checking all the files takes longer than the `CheckPeriod'.
-    \item[CreatingTimeout] specifies in seconds after how much time the Shepherd decideds that a started file upload won't happen.
+After checking the checksum of a file, the Shepherd will always wait this long before checking the next file, even if this means that checking all the files takes longer than the `CheckPeriod'\note[JKN]{TODO: Which takes precedense; CheckPeriod or MinCheckInterval?}.
+    \item[CreatingTimeout] specifies in seconds the time before the Shepherd decides that a started file upload has failed.
 This is not fatal, if the file is uploaded eventually, it will be a valid replica, but after this much time the system starts creating new replicas just in case.
-    \item[StoreClass] specifies which type of store the Shepherd should use for storing its metadata on disk.
-    \item[StoreCfg] can have parameters for the store class, e.g. the local directory to be used.
-    \item[BackendClass] specifes which backend the Shepherd should use for communicating with the storage element service.
-    \item[BackendCfg] has some configuration parameters for the backend class: \textbf{DataDir} and \textbf{TransferDir} is the directories the backend will use (see Section~\ref{sub:shepherd_implementation}), \textbf{TURLPrefix} should be the URL of the storage element service ending with a `/'.
-    \item[LibrarianURL] is the URL of a Librarian where the Shepherd should send the reports.
-    \item[BartenderULR] is the URL of a Bartender whom the Shepherd should connect if it wants to offer a replica for replication or recover a corrupted replica.
-\end{description}
+    \item[StoreClass] specifies which type of store the Shepherd should use for storing its metadata on disk (see Section \ref{sub:a_hash_configuration} for a list of choises).
+    \item[StoreCfg] specifies configuration parameters for the store class.
+These parameters depends on the StoreClass, but will in most cases include \emph{DataDir}, i.e., the local directory to be used.
+    \item[BackendClass] specifes which backend the Shepherd should use to communicate with the storage element service.
+    \item[BackendCfg] specifies configuration parameters for the backend class: \textbf{DataDir} and \textbf{TransferDir} are the directories the backend will use (see Section~\ref{sub:shepherd_implementation}), \textbf{TURLPrefix} is the URL of the storage element service ending with a `/'.
+    \item[LibrarianURL] is the URL of a Librarian where the Shepherd should send the reports\note[JKN]{TODO: Which reports?}.
+    \item[BartenderURL] is the URL of a Bartender to which the Shepherd should connect if it wants to offer a replica for replication or recover a corrupted replica.
+\end{description}
+
+An example configuration:
 
@@
 
-\section{Bartenders} % (fold)
+\section{Bartender} % (fold)
 \label{sec:bartenders}
@@
 
-The Bartender provides an easy to use interface of the ARC storage system to the users.
-You can put, get and delete files using their logical names (\emph{LN}'s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}.
-You can remove the link to a file or sub-collection from its parent collection without removing the file or sub-collection itself with the \emph{unlink} method.
-The metadata of a file or collection (e.g.~whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}.
-A \emph{stat} call gives all the information about a file or collection, and you can move (or create hardlinks of) collections and files within the namespace with \emph{move}.
-You can upload an entirely new replica to a file (e.g.~if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.
-You can create and remove mount points with the \emph{makeMountpoint} and \emph{unmakeMountpoint} methods.
-
-The Bartender communicates with the Librarians to get and modify the metadata of files and collections, and it communicates with the Shepherds to initiate file transfers.
-The Bartender also has \emph{gateway modules} which are capable of communicating different kinds of thrid-party storage solutions.
-With these gateway modules it is possible to create mount points within the namespace of the ARC storage, and to access the namespace of the third-party storage through these mount points, which means the user can use the Bartender to get listing of directories on the third-party storage or to get files from the third-party storage.
-Thus the user can use a single client tool to access different kind of storages easily.
+The Bartender provides an easy to use interface of Chelonia to the users.
+The user can put, get and delete files using their logical names (\emph{LN}'s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}.
+The link to a file or sub-collection can be removed from its parent collection without removing the file or sub-collection itself with the \emph{unlink} method.
+The metadata of a file or collection (e.g., whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}.
+A \emph{stat} call gives all the information about a file or collection, and collections and files can be moved (or hardlinks be created) within the namespace with \emph{move}.
+An entirely new replica of a file can be uploaded (e.g., if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.
+Mount points can be created and removed with the \emph{makeMountpoint} and \emph{unmakeMountpoint} methods.
+
+The Bartender communicates with Librarians to get and modify the metadata of files and collections, and it communicates with the Shepherds to initiate file transfers.
+The Bartender also has \emph{gateway modules} which are capable of communicating with different kinds of third-party storage solutions.
+With these gateway modules, it is possible to create mount points within the namespace of Chelonia, and to access the namespace of the third-party storage element through these mount points, which means the user can use the Bartender to get listing of directories on the third-party storage element or to get files from the third-party storage element.
+Thus the user can use a single client tool to easily access different kind of storage elements.
 
@@
 It makes decisions based on the identity of the connecting client and the access policy rules of the files and collections.
-The Bartender uses the security framework of the HED to evaluate the requests and the policies, which are genereted by the Bartender from the internal representation of the access rules.
+The Bartender uses the security framework of the HED to evaluate the requests and the policies which are generated by the Bartender from the internal representation of the access rules.
 For more details see Section~\ref{sec:security_implementation}.
@@
 
-    \item[putFile(putFileRequestList)] creates new files at the requested Logical Names, chooses a Shepherd and initiates the file transfer, and returns a TURL for each new file.
-
-    The \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}), where \emph{requestID} is an arbitrary ID which will be used in the response; \emph{LN} is the requested Logical Name of the new file, \emph{protocols} is a list of protocols we can use for uploading, \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples which should contain some mandatory properties in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}'.
-
-    The response is a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol}), where \emph{TURL} is a URL with a chosen \emph{protocol} which we can use to upload the file, the \emph{success} string could be `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}', `\textbf{internal error:} reason', etc.
-
-    \item[getFile(getFileRequestList)] finds the files referenced by the given Logical Name, chooses a replica and initiates the download, and returns a TURL for each file.
-
-    The \emph{getFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the file we want to get, \emph{protocols} is a list of transfer protocols we support.
-
-    The response is a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol}), where \emph{TURL} is the transfer URL using \emph{protocol}, with which we can download the file, \emph{success} could be `\textbf{done}', `\textbf{not found}', `\textbf{is not a file}', `\textbf{file has no valid replica}', `\textbf{error while getting TURL:} reason', etc.
-
-    \item[delFile(delFileRequestList)] removes the files references by the given Logical Names.
-
-    The \emph{delFileRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the files we want to delete.
-
-    The response is a list of (\emph{requestID}, \emph{status}), where the \emph{status} could be `\textbf{deleted}' or `\textbf{nosuchLN}'.
+    \item[putFile(putFileRequestList)] creates new files at the requested Logical Names, chooses a Shepherd and initializes the file transfer, and returns a TURL for each new file.
+The actual file transfer is left to the client tool.
+
+    The \emph{putFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}, \emph{protocols}) tuples, where \emph{requestID} is an arbitrary ID which will be used in the response, \emph{LN} is the requested Logical Name of the new file, \emph{protocols} is a list of supported protocols for uploading and \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples which should contain the following mandatory properties in the `\textbf{states}' section: `\textbf{size}', `\textbf{checksum}', `\textbf{checksumType}', and  `\textbf{neededReplicas}'.
+
+    The response is a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol}) tuples, where \emph{TURL} is a URL with a chosen \emph{protocol} which can be used to upload the file, the \emph{success} string can be one of `\textbf{done}', `\textbf{missing metadata}', `\textbf{parent does not exists}' and `\textbf{internal error:} reason'.
+
+    \item[getFile(getFileRequestList)] finds the files referenced by the given logical name, chooses a replica and initializes the download, and returns a TURL for each file.
+
+    The \emph{getFileRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{protocols}) tuples where \emph{requestID} is used in the response, \emph{LN} is the Logical Name referring to the requested file, \emph{protocols} is a list of supported transfer protocols.
+
+    The response is a list of (\emph{requestID}, \emph{success}, \emph{TURL}, \emph{protocol}), where \emph{TURL} is the transfer URL using \emph{protocol}, with which the file can be downloaded, \emph{success} is one of `\textbf{done}', `\textbf{not found}', `\textbf{is not a file}', `\textbf{file has no valid replica}' and `\textbf{error while getting TURL:} reason'.
+
+    \item[delFile(delFileRequestList)] removes the files references by the given logical names.
+
+    The \emph{delFileRequestList} is a list of (\emph{requestID}, \emph{LN}) pairs where the \emph{LN}'s are the logical names of the files to be deleted.
+
+    The response is a list of (\emph{requestID}, \emph{status}), where the \emph{status} is either `\textbf{deleted}' or `\textbf{nosuchLN}'.
 
@@
 
-    The \emph{unlinkRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Name of the files and sub-collections we want to unlink.
-
-    The response is a list of (\emph{requestID}, \emph{success}) where \emph{success} can be `\textbf{unset}', `\textbf{no such LN}', `\textbf{denied}' or `\textbf{nothing to unlink}'
-
-    \item[stat(statRequestList)] returns all metadata of files, collections or mount points
-
-    The \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names we want to get information about.
-
-    The response is a list of (\emph{requestID}, \emph{metadata}), where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples (see the data model in Section~\ref{sec:data_model}).
+    The \emph{unlinkRequestList} is a list of (\emph{requestID}, \emph{LN}) pairs with the logical names of the files and sub-collections to be unlinked.
+
+    The response is a list of (\emph{requestID}, \emph{success}) pairs where \emph{success} is one of `\textbf{unset}', `\textbf{no such LN}', `\textbf{denied}' or `\textbf{nothing to unlink}'.
+
+    \item[stat(statRequestList)] returns all metadata of files, collections or mount points.
+
+    \emph{statRequestList} is a list of (\emph{requestID}, \emph{LN}) pairs with the logical names of enntries whose metadata should be retrieved.
+
+    The response is a list of (\emph{requestID}, \emph{metadata}) pairs, where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples (see the data model in Section~\ref{sec:data_model}).
 
@@
 
-    The \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the `\textbf{entries}' section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUID's with the given name) and in the `\textbf{states}' section there is the `\textbf{closed}' property (if it is `yes' then no more files can be added or removed later).
-
-    The response is a list of (\emph{requestID}, \emph{success}), where \emph{success} can be `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}', `\textbf{internal error}', etc.
+    \emph{makeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{metadata}) tuples where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+The `\textbf{entries}' section may contain the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUID's with the given name) and in the `\textbf{states}' section there is a `\textbf{closed}' property (if it is `yes' then no more files can be added or removed later).
+
+    The response is a list of (\emph{requestID}, \emph{success}) pairs, where \emph{success} is one of `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}' or `\textbf{internal error}'.
 
@@
 
-    The \emph{unmakeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}) with the Logical Names of the collections we want to remove.
-
-    The response is a list of (\emph{requestID}, \emph{success}), where \emph{success} could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}', `\textbf{failed}: reason'.
+    \emph{unmakeCollectionRequestList} is a list of (\emph{requestID}, \emph{LN}) pairs with the logical names of the collections to be removed.
+
+    The response is a list of (\emph{requestID}, \emph{success}) pairs, where \emph{success} is one of `\textbf{removed}', `\textbf{no such LN}', `\textbf{collection is not empty}' or `\textbf{failed}: reason'.
 
@@
 
-    The \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) where \emph{LN} is the Logical Name of the collection we want to list, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
-
-    The response is a a list of (\emph{requestID}, \emph{entries}, \emph{status}) where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples (see the data model in Section~\ref{sec:data_model}); the \emph{status} could be `\textbf{found}', `\textbf{not found}', `\textbf{is a file}'.
+    \emph{listRequestList} is a list of (\emph{requestID}, \emph{LN}) pairs where \emph{LN} is the logical mame of the collection to be listed, \emph{neededMetadata} is a list of (\emph{section}, \emph{property}) pairs which filters the returned metadata.
+
+    The response is a a list of (\emph{requestID}, \emph{entries}, \emph{status}) tuples where entries is a list of (\emph{name}, \emph{GUID}, \emph{metadata}) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples (see the data model in Section~\ref{sec:data_model}).
+The \emph{status} is one of `\textbf{found}', `\textbf{not found}' or `\textbf{is a file}'.
 
     \item[move(moveRequestList)] moves file or collections within the namespace which changes the Logical Name of the file or collection.
-This method only alters metadata, does not move real file data.
-
-    The \emph{moveRequestList} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) where \emph{sourceLN} is the Logical Name referring to the file or collection we want to move (rename) and \emph{targetLN} is the new path, and if \emph{preserveOriginal} is true the \emph{sourceLN} would not be removed, so with \emph{preserveOriginal} we are actually creating a hard link.
-
-    The response is a list of (\emph{requestID}, \emph{status}), where status could be `\textbf{moved}', `\textbf{nosuchLN}', `\textbf{targetexists}', `\textbf{invalidtarget}', `\textbf{failed adding child to parent}', `\textbf{failed removing child from parent}'.
-
-    \item[modify(modifyRequestList)] modifies the metadata of files, collections and mount poins.
-Only the `states', the `policy' and the `metadata' sections can be modified, and there are separate access control actions for the three.
-
-    The \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} can be `\textbf{set}' (set the \emph{property} in the \emph{section} to \emph{value}), `\textbf{unset}' (remove the \emph{property}-\emph{value} pair from the \emph{section}), `\textbf{add}' (set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already).
-
-    The response is a list of (\emph{changeID}, \emph{success}), where \emph{success} could be  `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}' (for an `add' request), `\textbf{denied}', `\textbf{no such LN}', `\textbf{failed}: reason'.
-
-    \item[makeMountpoint(makeMountpointRequestList)] used to create the mountpoint
-
-    \emph{makeMountpointRequestList}, list of (\emph{requestID}, \emph{status})
+This method only alters metadata, it does not move real file data.
+
+    \emph{moveRequestList} is a list of (\emph{requestID}, \emph{sourceLN}, \emph{targetLN}, \emph{preserveOriginal}) tuples where \emph{sourceLN} is the logical name referring to the file or collection to be moved (renamed) and \emph{targetLN} is the new path.
+If \emph{preserveOriginal} is true the \emph{sourceLN} is kept, so with \emph{preserveOriginal} a hard link is in fact created.
+
+    The response is a list of (\emph{requestID}, \emph{status}) pairs, where status is one of `\textbf{moved}', `\textbf{nosuchLN}', `\textbf{targetexists}', `\textbf{invalidtarget}', `\textbf{failed adding child to parent}' or `\textbf{failed removing child from parent}'.
+
+    \item[modify(modifyRequestList)] modifies the metadata of files, collections and mount points.
+Only the `states', the `policy' and the `metadata' sections may be modified, and there are separate access control actions for each of the three.
+
+    The \emph{modifyRequestList} is a list of (\emph{changeID}, \emph{LN}, \emph{changeType}, \emph{section}, \emph{property}, \emph{value}) where \emph{changeType} is one of `\textbf{set}', set the \emph{property} in the \emph{section} to \emph{value}, `\textbf{unset}', remove the \emph{property}-\emph{value} pair from the \emph{section} or `\textbf{add}', set the \emph{property} in the \emph{section} to \emph{value} only if it is not exists already.
+
+    The response is a list of (\emph{changeID}, \emph{success}), where \emph{success} is on of `\textbf{set}', `\textbf{unset}', `\textbf{entry exists}' (for an `add' request), `\textbf{denied}', `\textbf{no such LN}' or `\textbf{failed}: reason'.
+
+    \item[makeMountpoint(makeMountpointRequestList)] creates a mountpoint to an external storage element.
 
     \emph{makeMountpoint} is quite similar to the \emph{makeCollection}.
-It is used to provide the access of the external storage system within the global namespace.
-\emph{makeMountpointRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{URL}, \emph{metadata} ) where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples where in the section there could be the initial content of the catalog in the form of name-GUID pairs (these entries will be hard links to the given GUID's with the given name).
-\emph{LN} will be requested Logical name and the \emph{URL} is the path to the external point that is required to access.
-In the policies section there could be some access policies, and in the metadata section there could be any other \emph{metadata} in key-value pairs.
-
-    The \emph{status} in the response could be \emph{done}, \emph{mountpoint exists}, \emph{parent does not exist}, \emph{internal error}, \emph{'failed: credentials not found'}, \emph{'cannot create anything in mountpoint'}
-
-    \item[unmakeMountPoint(unmakeMountpointRequestList)] delete the mountpoint
-
-    \emph{unmakeMountpointRequestList}, list of (\emph{requestID}, \emph{status})
-
-    \emph{unmakeMountpointRequestList} is a list of (requestID, LN) with the Logical Names of the mountpoints.
-\emph{status} could be \emph{removed}, \emph{no such LN}, \emph{Failed: Credential not available}.
-
-    \item[CredentialDelegation(CredentialDelegationRequest)] returns (\emph{status}, \emph{ID})
-
+It is used to provide access to an external storage system within the global namespace.
+\emph{makeMountpointRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{URL}, \emph{metadata}) pairs where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
+\emph{section} contains the initial content of the catalog in the form of name-GUID pairs (these entries are hard links to the given GUID's with the given name).
+\emph{LN} is the requested logical name and \emph{URL} is the path to the external storage element to be accessed.
+The \emph{policies} section contains access policies, and the metadata section contains any other \emph{metadata} in key-value pairs.
+
+    The \emph{status} in the response is one of `\textbf{done}', `\textbf{mountpoint exists}', `\textbf{parent does not exist}', `\textbf{internal error}', `\textbf{failed}: credentials not found' or `\textbf{cannot create anything in mountpoint}'.
+
+    \item[unmakeMountPoint(unmakeMountpointRequestList)] deletes mountpoints.
+
+    \emph{unmakeMountpointRequestList} is a list of (requestID, LN) pairs with the logical names of the mountpoints.
+\emph{status} is one of `\textbf{removed}', `\textbf{no such LN}', `\textbf{Failed:} Credential not available'.
+
+    \item[CredentialDelegation(CredentialDelegationRequest)] delegates credentials.
+\note[JKN]{TODO: This item has new style. Rewrite to make it similar to all the previous items. Also, second possible response looks very strange!}    
     From the client side \emph{createDelegation()} is called which interns initiate the \emph{CredentialDelegationInitRequest} for the service.
@@
     \item[RemoveCredentials(RemoveCredentialRequest)] returns (\emph{message}, \emph{status})
-
+\note[JKN]{TODO: This item has new style. Rewrite to make it similar to all the previous items. Also, the responses look very strange!}    
     \emph{RemoveCredentialRequest} initiates the process of deleting the user credential from the service side.
@@
 
-For managing collections the Bartender needs only to communicate with a Librarian service, for managing files it needs to communicate with Shepherd services as well.
+To manage collections the Bartender only needs to communicate with a Librarian service.
+To manage files it needs to communicate with Shepherd services as well.
 When a Bartender wants to create a new replica of a file (or the first replica of a new file) it should find available Shepherd services.
-Fortunately the Librarian has a list of registered Shepherd services, so the Bartender can use that.
-
-The Bartender can utilize gateway modules for accessing file listings and transfer URL from third-party storages.
+The Bartender uses the list of Shepherd services registered to the Librarian.
+
+The Bartender utilizes gateway modules to access file listings and transfer URL's from third-party storage elements.
 It works as an optional backend component of the Bartender service.
-\emph{Gateway} module is based on four basic methods \emph{get}, \emph{put}, \emph{list} and \emph{remove}.
-Since the task is to contact external store, which requires user to delegate the credentials so that it can communicate on behelf of the user.
-
-In order to access the \emph{files} and \emph{directories} from the external storage system, limited user credentials are passed to the \emph{Bartender} service.
+The \emph{Gateway} module is based on four basic methods; \emph{get}, \emph{put}, \emph{list} and \emph{remove}.
+Since the task is to contact external storage element, which requires the user to delegate the credentials so that it can communicate on behelf of the user\note[JKN]{TODO: What does this sentence mean?
+Please rewrite.}.
+
+In order to access the files and directories from the external storage element, limited user credentials are passed to the \emph{Bartender} service.
 Users can also remove their \emph{credentials} from the service side.
-These credentials are used in the methods like \emph{get}, \emph{put}, \emph{remove} and \emph{list} of the \emph{Gateway} module.
-These limited credentials are not used in any type of transfer, while listing files/directories, \emph{Gateway} uses the user credential, whereas in the \emph{get} and \emph{put} methods, it only uses the credentials to check the validity of the URL(s) before sending it to the user.
+These credentials are used in the methods \emph{get}, \emph{put}, \emph{remove} and \emph{list} of the \emph{Gateway} module.
+These limited credentials are not used in any type of transfer.
+While listing files and directories, the Gateway uses the user credential, whereas in the \emph{get} and \emph{put} methods, it only uses the credentials to check the validity of URL's before sending them to the user.
 % subsection implementation (end)
@@
 
-The Bartender has these configuration variables:
+The Bartender has the following configuration variables:
 
@@
     \item[LibrarianURL] specifies the URL of one or more Librarian services.
-    \item[ProxyStore] is a local directory where the Bartender will store the delegated proxies.
-    \item[GatewayClass] is the name of the gateway module we want to use in this Bartender.
-    \item[GatewayCfg] contains the configuration parameters of the gateway class: the \textbf{ProxyStore} is the directory where the gateway module could find delegated proxies, and the \textbf{CACertificatesDir} is a directory where the certificates of the trusted CAs are.
+    \item[ProxyStore] is a local directory where the Bartender will store the delegated proxies. This directory should be as secure as possible.
+    \item[GatewayClass] is the name of the gateway module to be used in this Bartender.
+    \item[GatewayCfg] contains the configuration parameters of the gateway class: \textbf{ProxyStore} is the directory where the gateway module can find delegated proxies, and \textbf{CACertificatesDir} is a directory with the certificates of the trusted CAs.
 \end{description}
@@
 
-In the first prototype release there is one client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run.
-It is capable of communicating with a given Bartender service, and uploading and downloading TURL's via HTTP.
-However in order to use proxy certificates or to transfer files with GridFTP protocol, you need to have the ARC client libraries installed.
-
-The credentials of the user and the URL of the Bartender can be configured either with an XML file located at \verb!~/.arc/client.xml! or with environment variables.
-
-Example for the \verb!client.xml! file:
+In the first prototype release\note[JKN]{TODO: Mention the arc\_storage\_cli in the intro of this document} there is one client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run.
+It is capable of communicating with a given Bartender service, and can upload and download files via HTTP.
+However, in order to use proxy certificates or to transfer files with the GridFTP protocol, the ARC client libraries must also be installed.
+
+The credentials of the user and the URL of the Bartender must be configured either with an XML file located at \verb!~/.arc/client.xml! or with environment variables.
+
+An example of user credentials and Bartender configuration in the \verb!~/.arc/client.xml! file:
 
@@
 
-Example for the environment variables:
+An example of user credentials and Bartender configuration with environment variables:
 
@@
 
-You can specify the number of needed replicas for new file with the \verb!ARC_NEEDED_REPLICAS! environment variable.
-
-The \verb!arc_storage_cli! has its built-in help. The methods can be listed with:
+The number of needed replicas for new files can be specified with the \verb!ARC_NEEDED_REPLICAS! environment variable. The default is one replica.
+
+The \verb!arc_storage_cli! has its built-in help. The methods can be listed by giving the command without arguments:
 
@@
     put[File], get[File], del[File], pol[icy], unlink,
-    credentialsDelegation, removeCredentials
+    credentialsDelegation, removeCredenstials
 \end{verbatim}
-Without arguments, each method prints its own help:
+Without arguments, each method prints its own help, e.g., for the move method:
 \begin{verbatim}
@@
 
-Here is an example of uploading, downloading and stat files:
+Here is an example of uploading, stating and downloading a file:
 
@@
 
-The \textbf{FUSE module} provides a high-level access to the storage system.
-Filesystem in Userspace (FUSE) provides a simple library and a kernel-userspace interface.
-Using FUSE and the ARC Python interface, the FUSE module allows users to mount the storage namespace into the local namespace, enabling the use of, e.g., graphical file browsers.
-Currently no delegation handling is implemented in the fuse module, so that to change ownership, group access, etc., you need to use the \verb!arc_storage_cli!.
+\textbf{ARCFS}, a Filesystem in Userspace (FUSE) module, provides high-level access to the storage system.
+FUSE provides a simple library and a kernel-userspace interface.
+Using FUSE and the ARC Python interface, ARCFS allows users to mount the storage namespace into the local namespace, enabling the use of operating system commands and tools such as graphical file browsers.
+Delegation handling is not yet implemented in the fuse module, so that to change ownership, group access, etc., the \verb!arc_storage_cli!
+tool must be used.
 
@@
 or with environment variables.
-See Section \ref{sub:prototype_cli_tool} for user configuration.
-
-An example of uploading, downloading and list files with FUSE:
+See Section \ref{sub:prototype_cli_tool} for more details.
+
+ARCFS consists of a Python script, \verb!arcfs.py!.
+It requires the ARC Python interface, the FUSE library and pyFUSE, a Python wrapper to FUSE.
+\verb!arcfs.py!
+takes only one input argument which is the mount point where Chelonia should be mounted.
+
+An example of uploading, downloading and listing files in Chelonia mounted by FUSE:
 
@@
 $ fusermount -u ./mnt
+$ arc_storage_cli stat /tmp/testfile
+- Calling the Bartender's stat method...
+- done in 0.05 seconds.
+'/tmp/testfile': found
+  states
+    checksumType: md5
+    neededReplicas: 1
+    size: 20
+    checksum: 9a9dffa22d227afe0f1959f936993a80
+  timestamps
+    created: 1210232135.57
+  parents
+    51e12fab-fd3d-43ec-9bc5-17041da3f0b2/testfile: parent
+  locations
+    http://localhost:60000/Shepherd fc0d3d99-6406-4c43-b2eb-c7ec6d6ab7fe: alive
+  entry
+    type: file
 \end{verbatim}
@@
 
-To access data through the ARC middleware client tools, one needs to go through Data Manager Components (DMCs).
+To access data through the ARC middleware client tools, one needs to go through Data Manager Components (DMC's).
 These are protocol specific plugins to the client tools.
-E.g, if you need to access data from a HTTPS service, you will go through the HTTP DMC using a URL starting with \verb!https://!and if you want to access data from a SRM service, you will use the SRM DMC using a URL starting with \verb!srm://!.
-Similarly, if you want access the ARC storage system, you will use the ARC DMC with protocol using a URL starting with \verb!arc://!.
-
-As with the \verb!arc_storage_cli!, the credentials of the user and the URL of the Bartender can be configured with an XML file located at \verb!~/.arc/client.xml!.
-When the ARC DMC is initiated it will create an instance of \verb!DataPointARC!
+For example, to access data from a HTTPS service, the HTTP DMC will be used with a URL starting with \verb!https://!, to access data from an SRM service, the SRM DMC will be used with a URL starting with \verb!srm://!.
+Similarly, to access Chelonia, the ARC DMC will be used with a URL starting with \verb!arc://!.
+
+As with the \verb!arc_storage_cli!
+command the credentials of the user and the URL of the Bartender can be configured with an XML file located at \verb!~/.arc/client.xml!.
+When the ARC DMC is initiated, it will create an instance of \verb!DataPointARC!
 which provides functionality for listing collections and file IO.
 When downloading/uploading files from/to the storage system, the data point will use the provided Bartender URL to acquire a TURL from a Bartender.
-This TURL will be given to a \verb!DataTransfer!.
-Calling the constructor of the DataHandle class with a URL will trigger the iGetDataPoint class of all registered DMCs to be called until a non-NULL pointer is returned.
-This means that as long as there exists a data point plugin for the protocol of the returned TURL, the transfer can proceed.
-
-Example of uploading, downloading and list files:
+This TURL will be given to a \verb!DataTransfer!
+handle.
+Calling the constructor of the DataHandle class with a URL will trigger the iGetDataPoint class of all registered DMC's to be called until a non-NULL pointer is returned.
+This means that as long as there exists a data point plugin for the protocol of the returned TURL, the transfer will proceed.
+
+An example of uploading, downloading and listing a file with the ARC client tools:
 
 \begin{verbatim}
-$ cat testfile
+$ arccp arc:///tmp/testfile fusefile
+$ cat fusefile
 This is a test file
-$ arccp testfile arc:///tmp/testfile
-$ arcls -l arc:///tmp/testfile
+$ arccp testfile arc:///tmp/testfile2
+$ arcls -l arc:///tmp/testfile2
 testfile file * * * *
-$ arccp arc:///tmp/testfile newfile
-$ cat newfile
+$ arccp arc:///tmp/testfile2 newfile2
+$ cat newfile2
 This is a test file
@@
 
+Note here that long listing (\verb!arcls -l!)! with the \verb!arc! protocol not yet supports time stamps. Hence, only the type of the entry (here "file") is shown.
+
+\note[JKN]{TODO: There is nothing about status or plans. Either add some status and plans or change subtitle of document.}
+
 % section grid_integration (end)

### 2009-04-16 doc/sec/papers/GRID09/paper.tex
paper draft, not finished yet

\documentclass[conference]{IEEEtran}
\usepackage[pdftex]{graphicx}
%\usepackage{epstopdf}
\graphicspath{{./png/}{./eps/}}
\DeclareGraphicsExtensions{.pdf,.png}
\usepackage[caption=false,font=footnotesize]{subfig}
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}
\title{The design and implementation of standard-based Grid single sign-on using federated identity}

\author{\IEEEauthorblockN{Weizhong Qiang, Aleksandr Konstantinov}
\IEEEauthorblockA{Department of Physics, University of Oslo}
}

\maketitle

\begin{abstract}
When pursuing the task of making access to Grids as simple as possible, security is one of the most important challenges in production Grid infrastructures, especially when the Grid applications span multiple administrative domains as well as heterogeneous Grid middlewares. A typical example is wide scale e-Science applications which need to coordinate resources shared among a number of independent institutions with different Grid middlewares deployed on these resources. In this paper, we describe security implementation and considerations used in the upcoming version of the Advanced Resource Connector (ARC) middleware, where the heterogeneity issue has been addressed. The main goal of ARC implementation in terms of security is to let the middleware be capable of interoperating with other Grid middlewares by leveraging on standard specifications. The key aspect of the work is to enhance the current proxy certificate based authentication and single sign-on by utilizing and enhancing the standardized Web Service specifications such as Security Assertion Markup Languages (SAML), single sign-on (SSO) profile and Web Services Security in order to achieve cross-middleware authentication and single sign-on.

and more importantly, this solution can utilize the widely deployed SAML based authentication implementation in Grid systems avoiding the use of X.509 certificate on the client side.

user can get a short-lived X.509 certificate without being bothered to contact any registration authority (RA) or certificate authority (CA).

\end{abstract}

%-------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
Grid provides the technology for wide-scale, cross-domain collaboration where users work with resources and share data which are distributed across administrative domains. This kind of collaboration is called Virtual Organization (VO)~\cite{IFoster01}. The resources that contribute to the VO can come from independent administrative domains, and even require different security approaches to access control due to usage of different underlying middlewares. The same heterogeneous situation applies to users who need to access the VO resources, because users holding one kind of a credential may need to utilize multiple resources with different security approaches. To solve the above issues, Grid middlewares should provide the following abilities:
\begin{itemize}
\item The user which participates a VO should be able to access data and resources without being required to constantly do authentication at each resource or service;
\item The VO should be able to easily manage the privileges and memberships of the users;
\item The VO together with its resources should be able to easily manage the authorization, given the condition that the resources are owned by different administrative domains.
\end{itemize}
In the current Grid community, the Grid Security Infrastructure (GSI)~\cite{IFoster98} is the de facto standard for authentication and transport level security. It builds on the Public Key Infrastructure (PKI), an architecture based on the X.509 public key certificate. GSI implements some enhancement (such as certificate delegation) based on the standard SSL/TLS protocol. Mutual authentication is required by GSI, and is the default configuration for GSI based Grid deployment.

In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS) ~\cite{AlfieriR05} have partly solved them, to a different extent. The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided. VOMS adopts the Attribute Certificate~\cite{RFC3821link} approach. The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client. Attributes include the VO membership data which is associated with the proxy certificate’s identity. Eventually the resources can make access control decisions based on parsing the VO membership attributes from the Attribute Certificate.

The VO is essentially supposed to be able to provide resource sharing capability with resources running under heterogeneous systems/middlewares. This goal can be achieved using Web Service Architecture (WSA). This way, the Web Service technology has been utilized in Grid computing middlewares, including Globus Toolkit~\cite{GTlink}, gLite~\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)~\cite{ARClink}.

Unfortunately, the existing Grid security solutions such as GSI and VOMS do not apply to Web Service based applications very well due to the following reasons:
\begin{itemize}
\item The GSS API based confidential communication and identity delegation has not been adopted by Web Service implementations.
\item The Attribute Certificate has not been adopted by any aspects Web Service implementation at all.
\item Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
\end{itemize}

Moreover, the Grid security solutions require each user to possess a X.509 certificate. However, the process to get a certificate may take quite long time, since the requester needs to be checked and approved by the Registration Authority (RA) and only then the Certificate Authority (CA) can issue the certificate according to the approval from the RA. Meanwhile, the process to use the certificate is also not so easy for the non-IT advanced research community to deal with, since the user has to generate proxy certificates by using Grid-specific command lines, such as grid-proxy-init or voms-proxy-init. On the other hand, users could often have some more familiar local community or institutional credentials such as username/password. Enabling users to use their own community credentials instead of the X.509 certificate to access Grids is a promising solution to make the Grids more easily accessible.

The above analysis shows that in order to achieve interoperability and transparency for user to access Grid system which is established on different middlewares, we need to address the mismatch between the different security solutions from these middlewares in order to ensure the interoperability as wide as possible; it is also necessary to address the mismatch between users' daily experience and middleware's requirements in order to ensure the access as simple as possible.

The approach in the new version of ARC is to utilize the existing Web Service standards, such as SAML and WS-Security, to achieve the interoperability and accessibility without breaking existing Grid specific security protocols.

The rest of this paper is organized as follows: Section 2 presents the ARC Grid middleware, especially the architecture of the new version of ARC. Section 3 describes the solution for cross-middleware authentication and single sign-on, including implementations for bridging the mismatch in terms of authentication and single sign-on. Section 4 discusses the related research, and Section 5 presents conclusions and future work outlook.

We investigated the authentication and single sign-on issue in the scenario when a user or user's representative (acting on behalf of this user) needs to access resources across different middlewares. Our work focuses on the following five major aspects:
\begin{itemize}
	\item Integration of community authentication -- to utilize the standard-compatible community authentication mechanism from the Shibboleth~\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
	\item Short lived credentials service -- to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
	\item Credential delegation service -- to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
\end{itemize}

%-------------------------------------------------------------------------
\section{ARC Grid middleware}
\label{sec:arcmiddleware}

ARC (Advanced Resource Connector) is an open source Grid middleware solution released under Apache license. ARC middleware aims at developing self-organized, fault-tolerant, non-intrusive, easy-manageable Grid middleware~\cite{MEllert07}. The current production version of ARC provides Grid services for job submission and management, resource characterization, resource aggregation and discovery, basic data management, integration of Grid security solution, and so on. This release has been deployed and used in production environment, and is one of the widely deployed Grid middlewares in Europe.

The next generation of ARC components is developed by the KnowARC project~\cite{KnowARClink}, based on the functionality and capabilities of the classic ARC middleware. It aims at implementing a Web Service oriented Grid middleware which will provide higher levels of resource and user abstraction through well-defined Web Service interface~\cite{KnowARCDesignlink} in order to provide interoperability with other service-based Grid middlewares, as well as other Web Service compatible applications.

As the key part of the implementation of new ARC middleware, there is a lightweight Web Service container called Hosting Environment Daemon (HED) which provides hosting for various services at application level, as well as a number of modules to support flexible, interoperable, and efficient communication mechanism for building SOAP based Web Services. The whole design of the HED is built around the idea of flexibility and modularity.

The architecture of the HED is illustrated in Figure \ref{fig:HED}. In general, there are few components called Message Chain Component (MCC) which are in charge of implementing different protocol levels. For instance, as shown in the example message flow, the HTTP MCC will process stream from the TLS MCC to parse the HTTP message and pass its body to the SOAP MCC, and also process the SOAP response from the SOAP MCC to generate the HTTP message for the TLS MCC.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{HED.png}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}
HED contains a framework for implementing and enforcing authentication and authorization. Each MCC or Service has a common interface for implementing various authentication and authorization functionality. Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service. Each MCC or Service is usually configured with two queues of SecHandler -- one for the incoming message and one for the outgoing message respectively. In Figure \ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules inside MCCs and Services are examples of SecHandler.

%-------------------------------------------------------------------------
\section{Grid authentication using federated identity: Integrating SAML2SSO profile}
\label{sec:intergrationSAML2SSO}
Identity federation is an emerging technology which enables the identity information to be trustily transfered across autonomous security domains. By using identity federation, users from one domain can access another domain without the need for directly trust relationship between users and accessed domain, i.e., users can use accounts from their host domains to access external domains, and accessed domain does not need to maintain accounts for those external users. Identity federation doesn't enforce specific authentication mechanisms, so that various types of authentication can be supported between user and its host domain, and user is able to use his frequently used account/credential to accomplish authentication. Considering the Grid use cases, we see that utilizing the existing account to access Grid without being bothered to applying and manipulating a X.509 credential should be attractive for Grid users.

Shibboleth is the one of the several implementations of the identity federation, which is based on the OASIS Security Assertion Markup Language (SAML) 2.0 specifications. In terms of authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines two functional components, an Identity Provider and a Service Provider. The Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity, while the Service Provider (SP) is responsible for controlling access to services and resources by analyzing the SAML assertions produced and issued by the IdP upon request.

In the implementation of ARC middleware, SAML2.0 web browser SSO profile is supported by implementing the functionality of service provider and web browser's user agent, and utilizing Shibboleth IdP implementation for the functionality of identity provider. The user agent functionality is for mimicing the behavior of web browser's user agent for authentication, such as the HTTP redirection and the HTTP cookie processing. The implementation of the user agent is based on the the HTTPS client interface of ARC, since the client interface of ARC can support different protocols which are incarnated by different MCCs.

\begin{figure}
\includegraphics[width=1.0\columnwidth]{SAML2SSO_UML.pdf}
\caption{Sequence diagram of SAML2.0 SSO profile integration in SOAP invocation 
between ARC WS Client and Web Service}
\label{fig:SAML2SSOUML}
\end{figure}

Figure \ref{fig:SAML2SSOUML} shows how is SAML2.0 SSO profile integrated into the SOAP 
invocation between ARC WS Client and Web Service. Step 1 to step 5 in Figure \ref{fig:SAML2SSOUML} 
is similar to those steps depicted in Web Browser SSO profile[ref], with the differences that
we specify the way about how does the service provider determine which identity provider to use (identity
provider discovery), by enabling the client to send the IdP name to service provider; as well 
as specify the way about how does the client/user agent authenticate against service provider, 
by implementing the form-based HTTP authentication which uses username and password for authentication between 
user agent and service provider (rather than X.509 client authentication) and is compatible 
with the Username/Password login handler of Shibboleth IdP.

In step 6, the service provider (SP Service) verifies and checks the SAML response, decrypts and stores 
the SAML assertion into session/connection context. That assertion includes the AuthnStatement and 
AttributeStatement elements. From step 7 to step 9, the WS client launches the SOAP invocation via 
the same connection as the one which is used by user agent to contact SP Service, then Grid/Web Service 
checks the AuthnStatement stored in the session context to see if the AuthnStatement is valid or expired. 
If valid, service handles the SOAP request and returns the SOAP response to the WS client.

Note that in the current implementation the service requires WS client to reuse the same 
TCP connection as the one used by user agent in step 5, in order to guarantee that the validity of 
the SAML2SSO result is applicable to the SOAP invocation process. On the service side, 
the sesion reusing is accomplish by deloying the SP service and other functional Grid/Web Service(s) 
together in the same service container.

There are a few benifits about integrating SAML2.0 SSO profile for Grid/Web service authentication. 
Firstly, the client certificate authentication can be switched off, so that users do not need 
to apply and maintain a X.509 credential. Secondly, some implementation of identity provider, such as 
Shibboleth IdP, can cache the authentication result through session management once the user
has succeeded to authenticate, and for a short period this authentication result is valid, so that 
the next time the user authenticates against IdP, the user agent can just feed IdP with the last 
successfully authenticated session's identity rather than feed IdP with username and password again. So 
the user (or the client on behalf of that user) can travel across multiple security domains with 
only providing his name and password once, which is also the characteristic of single sign-on.
Lastly, even though there are a few implementations of identity providers, most of them are 
standard-compliant, so the solution implemented in ARC middleware can interoperate with other 
identity provider implementations with minimum change.

%-------------------------------------------------------------------------
\section{Bridging federated identity and X.509 credential}
\label{sec:fedidtoX509}
Many widely used Grid middlewares are based on GSI which requires mutual X.509 authentication. 
Meanwhile, Web Service based Grid applications mostly do require client certificate authentication.  
To bridge the gap beween federated identity and X.509 credential, a short lived credential 
service (SLCS) is implemented through which a user can get a short-lived X.509 certificate only by
providing his username/password and authenticating through the SAML2SSO profile.

Also, in order to provide single sign-on capability for services that act on behalf of user to 
invoke other services, a X.509 credential delegation service is implemented, which is based on 
SOAP communication rather than the proprietary GSI communication implemented in Globus Toolkit.

\subsection{Short-lived credential service}
\label{sec:slcs}
The sequence diagram of SLCS process specializes the diagram shown in Figure \ref{fig:SAML2SSOUML}, 
with the SOAP request including X.509 request and SOAP response including X.509 response. 
In detail, the SLCS client firstly accomplishes the SAML2SSO profile, then sends 
SOAP request to the SLCS service with the X.509 request embedded. The SLCS service makes
access control decision according to the SAML assertion stored in session context, and issues 
a certificate (with 12 hours lifetime) with the SAML attributes as the X.509 certificate 
extension. The SLCS client then gets the SOAP response with the X.509 certificate embedded.

Since the lifetime of the short lived credential has a short lifetime, it is permissible to protect 
the private key by using the local file system permissions instead of encrypting it by 
using a pass-phrase. Therefore, no manual interactivity is required to uncrypted the private
key. Hence single sign-on is provided, since the user only needs to input his password when 
he starts SAML2SSO profile.

How to compose a distinguished name (DN) for the certificate is a critical issue for 
the SLCS service. We pick up the relatively identical information ``eduPersonPrincipalName'' 
(see eduPerson schema~\cite{eduSchemalink}) from SAML attributes that returned from Shibboleth
IdP, and use it to compose the DN. A typical example of the eduPersonPrincipalName 
could be alice@example.org. In such case composed DN is for example\\``/O=knowarc/OU=example.org/CN=alice''.

By using the SLCS service, the user can easily access the Grid services that requires 
X.509 credential on client side, at anywhere, simply by running the SLCS client command 
and providing his username and password.

\subsection{X.509 credential delegation service}
\label{sec:creddeleg}

\begin{figure}
\includegraphics[width=1.0\columnwidth]{Delegation_UML.pdf}
\caption{Sequence diagram of the usage of delegation service}
\label{fig:DelegUML}
\end{figure}

The sequence diagram is shown in Figure \ref{fig:DelegUML}. Firstly, the client needs to delegate
the proxy certificate or short-lived certificate to a delegation service through step 1 to step 7.
Then this client invokes the peer Web Service by appending the delegation information (URL of delegation
service and the delegation ID corresponding to the above delegation) into the SOAP body as out-of-band 
message. The Web Service then sends a X.509 request and the delegation ID to the delegation service 
which is specified by client side, and gets back a delegated certificate which can be composed together
with the private key for a proxy certificate. The Web Service could use this proxy certificate 
to represent the user to invoke another Web Service, which either could cause one more process of 
delegation (step 1 to step 13) if another delegation service is specified, or could directly start
from step 8 if the same delegation service is specified as the one in the former delegation process.

The client only needs to delegate the certificate to delegation service one time per session, it does
not need to do delegation for subsequent SOAP invokations after the first one.

The functionality of both client and service sides are implemented as pluggable handlers which can be 
configured into the client and service configuration, so that the service or client implementation itself
does not need to be changed.

We suppose the short-lived certificate is used by client, so that the SLCS service together with
the delegation service can provide single sign-on from user's federated identity, while keeping the 
interoperability to those Grid/Web services that requires mutual authentication using certificate.

%-------------------------------------------------------------------------
\section{Performace Evaluation}
\label{sec:perfeval}
The following section evaluate the performance of the

\subsection{Performance comparison between HED and Axis2/C}
\label{sec:perhedandaxis}
Since the implementation of HED includes an alternative SOAP implementation which is the base of 
the security related web servies described in this paper, it is useful to compare the performance 
of HED and other SOAP implementations. Although there are many SOAP implementations such as 
AxisJava, gSOAP, XSOAP4, etc., the goal of this paper is not about complete comparison between
HED and all the other SOAP implementations, so we use the Axis2/C (v1.5.0) to demostrate the performance
HED.

For the service side, we implement the simple ``echoString'' service in HED, and use the ``echoString''
service in Axis2/C. For client side, we develope a client by using API of ARC, to invoke the ``echoString''
services from Axis2/C and HED. The service invocation is just sending a short string to service, and 
getting back the same string. One test/experiment consists of a set of clients which run simultaneously. 
Each client runs for a fixed period of time, and invokes the service as many times as it can during 
the test period. Multiple clients are launched by creating multiple threads in one test.
Two values are measured: average response time and throughput. Average response time is computed
by counting the average value of the response time for each invocation. Throughput is computed
by counting the number of invocation during each second.

The two ``echoString'' services run in a test machine equipped with 3GHz Intel Pentium dual-core 
CPU and 2GB memory, Red Hat Enterprise Linux 5. The client runs in another machine equipped 
with 1.8GHz Intel Pentium CPU and 2GB memory, Ubuntu 8.10. The Axis2/C service is run using 
the httpd module on Apache Http Server version 2.2.11 instance. We keep all the configurations 
default in Apache Http Server. The client and service are configured with mutual TLS authentication.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{HED2Axis_thread10.pdf}
\caption{Performance results for HED and Axis2C: 1 service and 10 clients}
\label{fig:HED2Axis_thread10}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{HED2Axis_thread40.pdf}
\caption{Performance results for HED and Axis2C: 1 service and 40 clients}
\label{fig:HED2Axis_thread40}
\end{figure}

Figure \ref{fig:HED2Axis_thread10} and Figure \ref{fig:HED2Axis_thread40} are the results of 
performance comparison. 
Two types of experiments are shown: session reused, session no-reused.
``Session reused'' means all of the SOAP invocations on each client share the same session 
(tcp connection); while ``Session no-reused'' means each invocation launches a new session.

For ``Session reused'' experiments, we see that the average response time in HED is about 81ms
which is twice of the value in Axis2/C, and the throughput in HED is about half of the value in Axis2/C, 
when the client number is 10. When the client number is increased to 40, the average response 
time in HED is close to the value in Axis2/C when the duration is more than 60 seconds, and the 
similar result appears for throughput. One obersavation is that, the performance results of HED is 
quite stable, with the average response time almost kept unchanged under different time points, 
as well as under different number of clients, and throughput kept unchanged under different 
time points, while increased linearly with number of clients.

For ``Session no-reused'' experiments, the average response time in HED is almost the same as
the value in Axis2/C, in both cases of client number is 10 (the value is 15ms) and 40 (the value
 is 53ms) correspondingly. On the other hand the throughput in HED is also the same as the value 
in Axis2/C, in both cases of client number is 10 and 40. One interesting obersabation is that the 
throughput is close to a fixed value (around 500 request/s) for both HED and Axis2/C in both 
cases of client number, when the duration value becomes bigger. Also, in case of client number is
40, the throughput from ``Session reused'' experiments also is close to the same fixed value.
We speculate that, with a relatively big value of client number, no matter the session is reused or
not, the throughput is becoming a fixed value which can be explained as the SOAP processing capability
of the SOAP implementation. In this case, we can believe that HED has the similar performance 
capability as Axis2/C.

\subsection{Performance of the integration SAML2 SSO profile}
\label{sec:perfsaml2sso}
In our next experiment, we evaluate how does the integration of SAML2SSO profile impact on the 
performance of Web Service invocation. We configure a ``echoString'' service together with a SP (
Service Provider) service, and develope a client by using SAML2SSO related API of ARC, to invoke 
the ``echoString'', as well as interact with the IdP (Identity Provider). Services and client run on
the same set of machine as section \ref{sec:perhedandaxis}. Meanwhile, we deploy the Shibboleth IdP 
implementation on a machine equipped with 2.8GHz Intel Pentium CPU and 1GB memory, Red Hat Enterprise 
Linux 5.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread1.pdf}
\caption{Performance results for SAML2SSO profile: 1 service and 1 client}
\label{fig:SAML2SSO_thread1}
\end{figure}

Figure \ref{fig:SAML2SSO_thread1} depicts the average response time and throughputs for one client 
running against the service. Since the success of authentication from IdP can be regarded as 
valid for the whole session between client and service, the client with ``Session reused'' 
need to accomplish the SAML2SSO profile only once for all SOAP invocations on this client, 
while the client with ``Session no-reused'' will need to accomplish this profile once per 
SOAP invocation.

The average response time of ``Session no-reused'' experiments changes little with the execution 
duration, and the value is about 730ms; while the throughput is around 82 requests per minutes.
Considering that the average response time in case of 
``Session no-reused'' and pure SOAP invocation is around 15ms when client number is 10 (see section
\ref{sec:perhedandaxis}), not demostrated in section \ref{sec:perhedandaxis}, the same value applies 
when client number is only 1, therefore we can calculate that the consumption of time for 
SAML2SSO profile is about 98\% of the whole SOAP invocation with SAML2SSO profile integrated.
Deeper investigation about recording the time of each step of SAML2SSO profile supports the above
obersavation. In order to improve the performance of SAML2SSO profile, optimizing the process between
client and IdP is the correct direction, and probably one of the easiest way is to enhance the 
hardware configuration.

One the other hand, the average response time of ``Session reused'' experiments is about 82 ms,
which is almost as short as the value in the pure SOAP invocation experiment in \ref{sec:perhedandaxis}.
The result is resonable because the time consumption caused by only once of SAML2SSO execution on one 
client is relatively small comparing with the time consumption caused by the following hundreds 
of pure SOAP invocations.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread5.pdf}
\caption{Performance results for SAML2SSO profile: 1 service and 5 clients}
\label{fig:SAML2SSO_thread5}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10.pdf}
\caption{Performance results for SAML2SSO profile: 1 service and 10 clients}
\label{fig:SAML2SSO_thread10}
\end{figure}

Figure \ref{fig:SAML2SSO_thread5} and Figure \ref{fig:SAML2SSO_thread10} depicts the average 
response time and throughputs for 5 and 10 clients running against one service simultaneously,
with the session not being reused. The throughput seems to change little with the number of 
clients increased from five to ten, with the value being abroud 200 requests per minutes;
while the average response time doubles, with the value increasing from 1.6s to 3.1s. We conclude
that in order to get acceptable response time, there should be at least less than five concurrent
clients if they need to continuously participate one SAML2SSO profile.

However for some typical grid application scenarios, normally the client does not need 
to continuously accomplish SAML2SSO profile. For instance, a user could submit a job 
which needs a period of time (normally this period is much longer than a few seconds)
to execute, which launches an execution of SAML2SSO profile; 
afterwards, this user could query the information as well as the result about this job, which 
launches another execution of SAML2SSO profile. On the other hand, statistically, users will most
probably not access grid system in a concurrent way, rather a random way. Therefore, if we 
suppose a user could access grid once per minute, in our test environment, we can claim that 
around 200 clients randomly can be supported with acceptable performance.

\subsection{Performance of SLCS service}
\label{sec:perfslcsserv}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SLCS.pdf}
\caption{Performance results for SLCS service: 1 service and 1 client}
\label{fig:SLCS}
\end{figure}

Figure \ref{fig:SLCS} is the performance about the SLCS service. Since the SLCS is based on the SAML2SSO
profile, the reponse time should be the accumulation of time consumption about the SAML2SSO processing 
and the SLCS processing. In this experiment, SLCS service and client run on the same set of machine 
as section \ref{sec:perhedandaxis}. This figure shows that the average response time changes little with
the execution time with the value being around 1.3s which is nearly double as the value (730ms) of one client 
experiment about ``echoString'' service invocation with SAML2SSO integrated. Not demostrated in 
Figure \ref{fig:SLCS}, the more consumption of time is caused by the creating of X.509 request on the
 client side, which takes about roughly 310ms in average, as well as by the signing of X.509 certificate 
on the SLCS service side, which takes roughly 33ms in average, together with the time consumption for 
communication.

Even though the average response time of SLCS service is not very short, considering the short-lived 
credential has a relatively long lifetime (12 hours by default) and therefore client only needs to invoke
SLCS service once per 12 hours, the performance is completely acceptable. Also considering the time 
consumption for the creation of X.509 request (1024-bit RSA key pair generation) is on the client side,
we speculate the SLCS service side will not be affected much by the time-consuming key pair generation, and
therefore its performance resemble the performance of SAML2SSO profile in section \ref{sec:perfsaml2sso}.

\subsection{Performance of delegation service}
\label{sec:perfdelegserv}

In the last experiment, we evaluate how does the usage of delegation service impact on the 
performance of Web Service invocation. We deployed a ``echoString'' service plugged with delegation
security handler on a machine equipped with 2.80GHz Intel Pentium CPU and 1GB memory, Red Hat 
Enterprise Linux 5, and also configure a client together with delegation security handler plugged 
on a machine equipped with 1.8GHz Intel Pentium CPU and 2GB memory, Ubuntu 8.10. Aside from 
``echoString'' service and client, a delegation service is deployed on a machine equipped with 
3GHz Intel Pentium dual-core CPU and 2GB memory, Red Hat Enterprise Linux 5.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{Delegation_thread_to_perf.pdf}
\caption{Performance results of delegation service: from 1 client to 15 clients}
\label{fig:Deleg}
\end{figure}

Initial experiments which are not demonstrated in this paper show that both the response time and
throughput almost do not change with the time. Therefore, we measure the performance as the function 
of the client numbers from 1 to 15. Figure \ref{fig:Deleg} is performance results. We see that 
the response time increases in a linear manner against client numbers; and the throughput gradually 
increases, and stops increasing and almost keeps unchanged when the client number is bigger than eight.
We conclude that if the client number is bigger than eight, in our test environment, there is no 
throughput benefit can be achieved by adding clients.

%-------------------------------------------------------------------------
\section{Related Work}
\label{sec:relatedwork}
Similar work that integrates Shibboleth and Grid middleware does exist. GridShib~\cite{VWelch05,TBarton06} uses Shibboleth to get entity's attribute for the authorization of services developed in Globus Tookit. Unlike the work described in this paper, the GridShib does not use Shibboleth's native login handlers for authentication to Shibboleth IdP, instead, a new login handler specific for X.509 based authentication is developed for the Shibboleth IdP, and a credential service is developed for acting as an online-CA, authenticating the user, and issuing short-lived X.509 credential. Moreover, this paper provides a single sign-on solution for the user with community credentials authenticating to Grid services by directly using SAML2.0 SSO profile.

SWITCH (The Swiss Education and Research Network) short-lived credential service (SLCS)~\cite{switchslcslink} provides a SLCS service and a client. The advantage of our work comparing to the SWTCH's solution is that the SLCS service and client are based on Web Services, so that they can be implemented in any Web Service container while keeping interoperability.

In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98,VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way. Gridsite project implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.

%-------------------------------------------------------------------------
\section{Conclusion And Future Work}
\label{sec:conclusion}
In this paper we presented the solution built into the ARC Grid middleware, aiming towards the cross-middle-ware authentication and single sign-on. The goal of our work is to develop an interoperable and accessible single sign-on solution to be used for a VO scenario in production Grids, where users can easily authenticate against the Grids with any of their credentials (not only X.509) from anywhere, and move inside Grid middleware infrastructure and even across heterogeneous Grids  boundaries, without re-authentication within a reasonable period. In order to do so, we implemented some standards based Web Services (including the SLCS service, the X.509 credential delegation service, and the SAML Token delegation service), as well as the user agent and the Service Provider in the context of the SAML2 SSO profile.

It must be noted that due to the standard specifications utilized (SAML and WS-Security) which are independent from any particular middleware, the WS clients and Web/Grid Services developed within the ARC middleware can interoperate with WS clients and Web/Grid Services developed as part of some other middlewares in a standardized way.

Though only authentication issue is discussed in this paper, as the future work, we will implement a solution for cross-middleware attribute-based authorization by using the standard specifications such as SAML.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

%\bibliographystyle{IEEEtran}

\begin{thebibliography}{1}

\bibitem{IFoster01}
I. Foster, C. Kesselman, S. Tuecke. The anatomy of the Grid: Enabling scalable virtual organizations, International Journal of Supercomputer Applications, 15(3), 200-222 (2001)

\bibitem{IFoster98}
I. Foster, C. Kesselman, G. Tsudik, and S. Tuecke, A Security Architecture for Computational Grids, ACM Conference on Computers and Security, 1998, 83-91.

\bibitem{AlfieriR05}
R. Alfieri, R. Cecchini, V. Ciaschini, L. dell’Agnello, A. Frohner, K. Lorentey, and F. Spataro. From gridmap-file to voms: managing authorization in a Grid environment. Future Generation Comp. Syst., 21(4), 549-558 (2005)

\bibitem{RFC3821link}
RFC 3821- An Internet Attribute Certificate Profile for Authorization. http://www.faqs.org/rfcs/rfc3281.html

\bibitem{GTlink}
Globus Toolkit. http://www.globus.org/toolkit/

\bibitem{gLitelink}
gLite: lightweight middleware for Grid computing. http://glite.web.cern.ch

\bibitem{ARClink}
Advanced Resource Connector. http://www.nordugrid.org/middleware/

\bibitem{WSSeclink}
OASIS Web Services Security. www.oasis-open.org/committees/wss/

\bibitem{SAMLlink}
OASIS Security Assertion Markup Languages (SAML). www.oasis-open.org/committees/security/

\bibitem{MEllert07}
M. Ellert et al. Advanced resource connector middleware for lightweight computational grids, Future Generation computer systems, 23(2), 219-240 (2007)

\bibitem{KnowARClink}
KnowARC project.  https://www.knowarc.eu/

\bibitem{KnowARCDesignlink}
Design document of new version ARC. https://www.knowarc.eu/documents/Knowarc\_D1.1-1\_07.pdf

\bibitem{Shiblink}
The Shibboleth Project. http://shibboleth.internet2.edu/

\bibitem{eduSchemalink}
eduPerson and eduOrg Object shema. http://middleware.internet2.edu/eduperson/

\bibitem{VWelch05}
V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. 4th Annual PKI R\&D Workshop, 2005.

\bibitem{TBarton06}
T. Barton et al. Identity Federation and Attribute-based Authorization through the Globus Toolkit, Shibboleth, Gridshib, and MyProxy. 5th Annual PKI R\&D Workshop, 2006.

\bibitem{switchslcslink}
SWITCH Short Lived Credential Service. http://www.switch.ch/Grid/slcs/

\bibitem{VWelch04}
V. Welch et al. X.509 proxy certificate for dynamic delegation, Proceeding of the 3rd Annual KI R\&D Workshop, 2004.

\bibitem{MAhsant04}
M. Ahsant, J. Basney, and O. Mulmo. Grid Delegation Protocol, UK Workshop on Grid Security Experiences, Oxford. 2004.

\bibitem{WSTrustlink}
OASIS WS-Trust specification. http://docs.oasis-open.org/ws-sx/ws-trust/200512

\bibitem{GridSitelink}
Gridsite delegation service. http://www.gridsite.org/wiki/Delegation\_protocol

\end{thebibliography}

\end{document}

### 2009-04-16 doc/sec/papers/GRID09/paper.tex
missing figure
More content, not yet finished still

@@
 \begin{document}
-\title{The design and implementation of standard-based Grid single sign-on using federated identity}
+\title{The design and implementation of standards-based Grid single sign-on using federated identity}
 
@@
 \begin{abstract}
-When pursuing the task of making access to Grids as simple as possible, security is one of the most important challenges in production Grid infrastructures, especially when the Grid applications span multiple administrative domains as well as heterogeneous Grid middlewares.
-A typical example is wide scale e-Science applications which need to coordinate resources shared among a number of independent institutions with different Grid middlewares deployed on these resources.
-In this paper, we describe security implementation and considerations used in the upcoming version of the Advanced Resource Connector (ARC) middleware, where the heterogeneity issue has been addressed.
-The main goal of ARC implementation in terms of security is to let the middleware be capable of interoperating with other Grid middlewares by leveraging on standard specifications.
-The key aspect of the work is to enhance the current proxy certificate based authentication and single sign-on by utilizing and enhancing the standardized Web Service specifications such as Security Assertion Markup Languages (SAML), single sign-on (SSO) profile and Web Services Security in order to achieve cross-middleware authentication and single sign-on.
-
-and more importantly, this solution can utilize the widely deployed SAML based authentication implementation in Grid systems avoiding the use of X.509 certificate on the client side.
-
-user can get a short-lived X.509 certificate without being bothered to contact any registration authority (RA) or certificate authority (CA).
-
+Security infrastructure is one of the most challenging tasks in the developement, integration and
+deployment of Grid middlewares. Even though the Grid community addresses the security issue 
+through public key infrastructures (PKI) to support mutual authentication using X.509 certificates,
+maintaining the X.509 credential is not that easy for non-IT specialists, and is proved to be an 
+obstacle for the more widely deployment of Grid technology. Considering the identity federation 
+technology is a emergingly widely used technology which can facilitate cross-domain single sign-on,
+without requiring the users to maintain the X.509 credentials rather than their own institutional
+accounts, we believe utilizing identity federation for Grid middlwares is a promising direction 
+for Grid technology to be more widely used.
+This paper describes a single sign-on infrastructure developed as part of the NorduGrid/ARC grid 
+middleware. It adopts the identity federation standard (SAML), as well as Web Service standand. 
+It focus on single sign-on solution on the middleware level for users to access Grids by only 
+using their frequently used accounts, without being bothered to maintain X.509 credentials. 
+Unlike other research which utilize identity federation on the Grid portal level, solution on
+the middleware level can provide more flexibility.
+Also, the performance of single sign-on solution is measured. We identify the performance limitation 
+of security-related services inside this solution, and analyze the ways to avoid or improve the 
+limitation. 
+To our knowledge, the work described in this paper is the first evaluated implementation that utilizes 
+identity federation for Grid usage on the middlware level.
 \end{abstract}
@@
 \label{sec:intro}
-Grid provides the technology for wide-scale, cross-domain collaboration where users work with resources and share data which are distributed across administrative domains.
-This kind of collaboration is called Virtual Organization (VO)~\cite{IFoster01}.
-The resources that contribute to the VO can come from independent administrative domains, and even require different security approaches to access control due to usage of different underlying middlewares.
-The same heterogeneous situation applies to users who need to access the VO resources, because users holding one kind of a credential may need to utilize multiple resources with different security approaches.
-To solve the above issues, Grid middlewares should provide the following abilities:
-\begin{itemize}
-\item The user which participates a VO should be able to access data and resources without being required to constantly do authentication at each resource or service;
-\item The VO should be able to easily manage the privileges and memberships of the users;
-\item The VO together with its resources should be able to easily manage the authorization, given the condition that the resources are owned by different administrative domains.
-\end{itemize}
-In the current Grid community, the Grid Security Infrastructure (GSI)~\cite{IFoster98} is the de facto standard for authentication and transport level security.
-It builds on the Public Key Infrastructure (PKI), an architecture based on the X.509 public key certificate.
-GSI implements some enhancement (such as certificate delegation) based on the standard SSL/TLS protocol.
-Mutual authentication is required by GSI, and is the default configuration for GSI based Grid deployment.
+Grid infrastructures facilitate the interoperability between wide-scale, cross-domain heterogeneous 
+resources, as well as the accessibility provided to users. In terms of security issue, the Grid security 
+infrastruture should provide accessibility as much as possible without loosing the security benifits.
 
-In terms of the Virtual Organization issues mentioned earlier, GSI and some other related solutions like the Virtual Organization Management Service (VOMS) ~\cite{AlfieriR05} have partly solved them, to a different extent.
-The identity delegation and the X.509 proxy certificate initiated by GSI can achieve single sign-on, so that the need to re-enter the user’s pass phrase for authentication can be avoided.
-VOMS adopts the Attribute Certificate~\cite{RFC3821link} approach.
-The Attribute Certificate is signed by the VOMS server and then bound to proxy certificates by the client.
-Attributes include the VO membership data which is associated with the proxy certificate’s identity.
-Eventually the resources can make access control decisions based on parsing the VO membership attributes from the Attribute Certificate.
+The current Grid community uses Grid Security Infrastructure (GSI)~\cite{IFoster98} as the de facto 
+standard for authentication and transport level security, which builds on the Public Key Infrastructure
+(PKI) to support authentication using X.509 certificate. Mutual authentication is required 
+by GSI based Grid deployment. Mutual authentication means that Grid users have to possess X.509 
+credentials. To obtain a X.509 credential, a user needs to contact the Registration Authority (RA)
+which will check the user's information and approve his certificate request; afterwards,
+the Certificate Authority (CA) can issue the certificate according to the approval from 
+the RA. The whole process is arduous and may take some time.
+Meanwhile, to maintain a X.509 certificate is also not so easy for the non-IT community to deal 
+with, since users have to periodically create proxy certificates by using Grid-specific command line 
+utility, such as grid-proxy-init or voms-proxy-init. Also maintaing the CAs or RAs is also not a  
+easy task, especially when the community of Grid users is getting much wider.
 
-The VO is essentially supposed to be able to provide resource sharing capability with resources running under heterogeneous systems/middlewares.
-This goal can be achieved using Web Service Architecture (WSA).
-This way, the Web Service technology has been utilized in Grid computing middlewares, including Globus Toolkit~\cite{GTlink}, gLite~\cite{gLitelink}, as well as the Advanced Resource Connector (ARC)~\cite{ARClink}.
+On the other hand, users normally could already have had some frequently-used campus or institutional 
+accounts/credentials such as username/password. We believe that enabling Grid users to use their home 
+credentials rather than the X.509 certificate to access Grids is a promising way 
+to make the Grids more easily accessible, as well as to extend the user community of Grids
+to a much larger scale.
 
-Unfortunately, the existing Grid security solutions such as GSI and VOMS do not apply to Web Service based applications very well due to the following reasons:
-\begin{itemize}
-\item The GSS API based confidential communication and identity delegation has not been adopted by Web Service implementations.
-\item The Attribute Certificate has not been adopted by any aspects Web Service implementation at all.
-\item Web Service implementations commonly use the Web Services Security (WS-Security)~\cite{WSSeclink}, Security Assertion Markup Languages (SAML)~\cite{SAMLlink}, and standard SSL/TLS as well as standard X.509 certificate when considering security, which make them impossible to interoperate with the GSI based Grid middlewares.
-\end{itemize}
+To enable users to utilize the institutional credentials to access Grid, some issues needs to be 
+addressed. Firstly, the Grid middleware needs to be improved to support the authentication based on
+institutional credentials instead of X.509 certificates; Secondly, in order to interoperate with
+those Grid infrastructures which require X.509 certificate based mutual authentication, an approach
+should be provided for obtaining X.509 certificate based on the institutional credentials; Thirdly,
+the single sign-on characteristic should be provided so that users can login once and then transverse 
+from Grid resources to Grid resources without being prompted to log in again at each of the resources.
 
-Moreover, the Grid security solutions require each user to possess a X.509 certificate.
-However, the process to get a certificate may take quite long time, since the requester needs to be checked and approved by the Registration Authority (RA) and only then the Certificate Authority (CA) can issue the certificate according to the approval from the RA.
-Meanwhile, the process to use the certificate is also not so easy for the non-IT advanced research community to deal with, since the user has to generate proxy certificates by using Grid-specific command lines, such as grid-proxy-init or voms-proxy-init.
-On the other hand, users could often have some more familiar local community or institutional credentials such as username/password.
-Enabling users to use their own community credentials instead of the X.509 certificate to access Grids is a promising solution to make the Grids more easily accessible.
+The approach in the new version of ARC grid middleware is to utilize the identity federation 
+standards, specifically, the Security Assertion Markup Language (SAML) specification, and Web 
+Service standards, to make the Grid single sign-on be possible.
 
-The above analysis shows that in order to achieve interoperability and transparency for user to access Grid system which is established on different middlewares, we need to address the mismatch between the different security solutions from these middlewares in order to ensure the interoperability as wide as possible; it is also necessary to address the mismatch between users' daily experience and middleware's requirements in order to ensure the access as simple as possible.
-
-The approach in the new version of ARC is to utilize the existing Web Service standards, such as SAML and WS-Security, to achieve the interoperability and accessibility without breaking existing Grid specific security protocols.
-
-The rest of this paper is organized as follows: Section 2 presents the ARC Grid middleware, especially the architecture of the new version of ARC.
-Section 3 describes the solution for cross-middleware authentication and single sign-on, including implementations for bridging the mismatch in terms of authentication and single sign-on.
-Section 4 discusses the related research, and Section 5 presents conclusions and future work outlook.
-
-We investigated the authentication and single sign-on issue in the scenario when a user or user's representative (acting on behalf of this user) needs to access resources across different middlewares.
-Our work focuses on the following five major aspects:
-\begin{itemize}
-	\item Integration of community authentication -- to utilize the standard-compatible community authentication mechanism from the Shibboleth~\cite{Shiblink} Identity Provider software when users authenticate against the Grid services.
-	\item Short lived credentials service -- to issue the short lived X.509 certificates for users authenticated using the community authentication, in order to interoperate with the services that depend on PKI based authentication.
-	\item Credential delegation service -- to process the X.509 credential delegation as a normal Web Service, in order to provide single sign-on ability.
-\end{itemize}
+The rest of this paper is organized as follows: Section \ref{sec:arcmiddleware} presents the 
+ARC Grid middleware, especially the architecture of the new version of ARC. Section 
+\ref{sec:intergrationSAML2SSO} describes the solution of using federated identity for 
+Grid authentication by using SAML2SSO profile; Setion \ref{sec:fedidtoX509} describes
+how to obtain a X.509 credential based on federated identity, and how to use it for 
+accessing multiple Grid resources; Section \ref{sec:perfeval} is performance evaluation;
+Section \ref{sec:relatedwork} describes the related work; Section \ref{sec:conclusion} 
+contains conclusions and future work outlook.
 
@@
 \begin{figure}
-\includegraphics[width=0.9\columnwidth]{HED.png}
+\includegraphics[width=0.9\columnwidth]{HED.pdf}
 \caption{The example of Host Environment Daemon deployed with A-REX and File services}

### 2009-04-17 doc/sec/papers/GRID09/paper.tex
more content; have not applied the comments yet

@@
 Also, the performance of single sign-on solution is measured. We identify the performance limitation 
-of security-related services inside this solution, and analyze the ways to avoid or improve the 
-limitation. 
+of security-related services inside this solution, and analyze the ways to avoid the limitation.
 To our knowledge, the work described in this paper is the first evaluated implementation that utilizes 
@@
 with, since users have to periodically create proxy certificates by using Grid-specific command line 
-utility, such as grid-proxy-init or voms-proxy-init. Also maintaing the CAs or RAs is also not a  
+utilities, such as grid-proxy-init or voms-proxy-init. Also maintaing the CAs or RAs is not a  
 easy task, especially when the community of Grid users is getting much wider.
@@
 On the other hand, users normally could already have had some frequently-used campus or institutional 
-accounts/credentials such as username/password. We believe that enabling Grid users to use their home 
+accounts/credentials such as username/password. We believe that enabling Grid users to use their institutional 
 credentials rather than the X.509 certificate to access Grids is a promising way 
@@
 should be provided for obtaining X.509 certificate based on the institutional credentials; Thirdly,
-the single sign-on characteristic should be provided so that users can login once and then transverse 
-from Grid resources to Grid resources without being prompted to log in again at each of the resources.
+the single sign-on characteristic should be provided so that users can authenticate once and then transverse 
+from Grid resources to Grid resources without being prompted to authenticate again at each of the resources.
 
@@
 standards, specifically, the Security Assertion Markup Language (SAML) specification, and Web 
-Service standards, to make the Grid single sign-on be possible.
+Service standards, to facilitate the Grid single sign-on.
 
@@
 
-ARC (Advanced Resource Connector) is an open source Grid middleware solution released under Apache license.
-ARC middleware aims at developing self-organized, fault-tolerant, non-intrusive, easy-manageable Grid middleware~\cite{MEllert07}.
-The current production version of ARC provides Grid services for job submission and management, resource characterization, resource aggregation and discovery, basic data management, integration of Grid security solution, and so on.
-This release has been deployed and used in production environment, and is one of the widely deployed Grid middlewares in Europe.
+ARC (Advanced Resource Connector) is an open source Grid middleware solution released 
+under Apache license. ARC middleware aims at developing self-organized, 
+fault-tolerant, non-intrusive, easy-manageable Grid middleware~\cite{MEllert07}. 
+The current production version of ARC provides Grid services for job submission and 
+management, resource characterization, resource aggregation and discovery, basic data 
+management, integration of Grid security solution, and so on. This release has been 
+deployed and used in production environment, and is one of the widely deployed Grid 
+middlewares in Europe.
 
-The next generation of ARC components is developed by the KnowARC project~\cite{KnowARClink}, based on the functionality and capabilities of the classic ARC middleware.
-It aims at implementing a Web Service oriented Grid middleware which will provide higher levels of resource and user abstraction through well-defined Web Service interface~\cite{KnowARCDesignlink} in order to provide interoperability with other service-based Grid middlewares, as well as other Web Service compatible applications.
+The development version of ARC components is developed by the KnowARC project~\cite{KnowARClink}, 
+based on the functionality of the current production ARC middleware. It aims at 
+implementing a Web Service oriented Grid middleware which will provide higher levels of 
+resource and user abstraction through well-defined Web Service interface~\cite{KnowARCDesignlink} 
+in order to provide interoperability with other Web Service oriented Grid middlewares, as well as 
+other Web Service compatible applications.
 
-As the key part of the implementation of new ARC middleware, there is a lightweight Web Service container called Hosting Environment Daemon (HED) which provides hosting for various services at application level, as well as a number of modules to support flexible, interoperable, and efficient communication mechanism for building SOAP based Web Services.
-The whole design of the HED is built around the idea of flexibility and modularity.
+As the key part of the implementation of new ARC middleware, there is a lightweight Web 
+Service container called Hosting Environment Daemon (HED) which provides hosting environment 
+for various services at application level, as well as a number of modules to support flexible, 
+interoperable, and efficient communication mechanism for building SOAP based Web Services. 
+The design of the HED is built around the idea of flexibility and modularity,
+such that the application service developers can simply concentrate on the application 
+level Web Service implementation by only using the core minimum amount of components. It 
+also simplifies work on the middleware level, e.g. making possible to implement another 
+communication protocol or authentication mechanism. A system administrator can easily 
+configure and deploy the middleware and application for different kinds of requirements 
+without having to know much about the implementation.
 
-The architecture of the HED is illustrated in Figure \ref{fig:HED}.
-In general, there are few components called Message Chain Component (MCC) which are in charge of implementing different protocol levels.
-For instance, as shown in the example message flow, the HTTP MCC will process stream from the TLS MCC to parse the HTTP message and pass its body to the SOAP MCC, and also process the SOAP response from the SOAP MCC to generate the HTTP message for the TLS MCC.
+The architecture of the HED is illustrated in Figure \ref{fig:HED}. In general, there 
+are few components called Message Chain Component (MCC) which are in charge of implementing 
+different protocol levels. For instance, as shown in the example message flow, the HTTP MCC 
+will process stream from the TLS MCC to parse the HTTP message and pass its body to the SOAP 
+MCC, and also process the SOAP response from the SOAP MCC to generate the HTTP message 
+for the TLS MCC.
+
+Dotted line in Figure \ref{fig:HED} shows an alternative path for the information 
+to propagate among MCCs. A service administrator can configure the MCCs according to the 
+interoperability requirements with the other part. For instance, the configuration marked 
+with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP 
+message mechanism (see WSE’s SoapSender and SoapReceiver)~\cite{WSElink}. Another configuration 
+could be SOAP over HTTPG (HTTP over GSI) which is needed to interoperate with services 
+like the Storage Resource Manager (SRM)~\cite{A.Shoshan03} service. This shows the 
+flexibility of HED in terms of protocols support.
 
@@
 \end{figure}
-HED contains a framework for implementing and enforcing authentication and authorization.
-Each MCC or Service has a common interface for implementing various authentication and authorization functionality.
-Each functionality can be implemented as a pluggable and configurable component (plug-ins) called SecHandler, which is implemented as a C++ class and provides a method for processing message that travels through MCC or Service.
-Each MCC or Service is usually configured with two queues of SecHandler -- one for the incoming message and one for the outgoing message respectively.
-In Figure \ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules inside MCCs and Services are examples of SecHandler.
+
+HED contains a flexible security framework for implementing and enforcing security 
+related functionality, such as authentication and authorization. 
+Each security related functionality can be implemented as a pluggable and configurable 
+component (plug-ins) called SecHandler. Each MCC or Service is usually configured with 
+two queues of SecHandler -- one for the incoming message and one for the outgoing message 
+respectively. In Figure \ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules inside 
+MCCs and Services are examples of SecHandler.
 
@@
 \label{sec:intergrationSAML2SSO}
-Identity federation is an emerging technology which enables the identity information to be trustily transfered across autonomous security domains.
-By using identity federation, users from one domain can access another domain without the need for directly trust relationship between users and accessed domain, i.e., users can use accounts from their host domains to access external domains, and accessed domain does not need to maintain accounts for those external users.
-Identity federation doesn't enforce specific authentication mechanisms, so that various types of authentication can be supported between user and its host domain, and user is able to use his frequently used account/credential to accomplish authentication.
-Considering the Grid use cases, we see that utilizing the existing account to access Grid without being bothered to applying and manipulating a X.509 credential should be attractive for Grid users.
+Identity federation is an emerging technology which enables the identity information to be 
+trustily transfered across autonomous security domains. By using identity federation, users 
+from one domain can access another domain without the need for directly trust relationship 
+between users and accessed domain, i.e., users can use accounts from their host domains to 
+access external domains, and accessed domain does not need to maintain accounts for those 
+external users. Identity federation doesn't enforce specific authentication mechanisms, so 
+that various types of authentication can be supported between user and its host domain, and 
+user is able to use his frequently used account/credential to accomplish authentication. 
+Considering the Grid use cases, we see that utilizing the existing account to access Grid 
+without being bothered to applying and manipulating a X.509 credential should be attractive 
+for Grid users.
 
-Shibboleth is the one of the several implementations of the identity federation, which is based on the OASIS Security Assertion Markup Language (SAML) 2.0 specifications.
-In terms of authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines two functional components, an Identity Provider and a Service Provider.
-The Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity, while the Service Provider (SP) is responsible for controlling access to services and resources by analyzing the SAML assertions produced and issued by the IdP upon request.
+Shibboleth is the one of the several implementations of the identity federation, which is 
+based on the OASIS Security Assertion Markup Language (SAML) 2.0 specifications. In terms 
+of authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines 
+two functional components, an Identity Provider and a Service Provider. The Identity Provider 
+(IdP) is responsible for creating, maintaining, and managing user identity, while the Service 
+Provider (SP) is responsible for controlling access to services and resources by analyzing 
+the SAML assertions produced and issued by the IdP upon request.
 
-In the implementation of ARC middleware, SAML2.0 web browser SSO profile is supported by implementing the functionality of service provider and web browser's user agent, and utilizing Shibboleth IdP implementation for the functionality of identity provider.
-The user agent functionality is for mimicing the behavior of web browser's user agent for authentication, such as the HTTP redirection and the HTTP cookie processing.
-The implementation of the user agent is based on the the HTTPS client interface of ARC, since the client interface of ARC can support different protocols which are incarnated by different MCCs.
+In the implementation of ARC middleware, SAML2.0 web browser SSO profile is supported by 
+implementing the functionality of service provider and web browser's user agent, and utilizing 
+Shibboleth IdP implementation for the functionality of identity provider. The user agent 
+functionality is for mimicing the behavior of web browser's user agent for authentication, 
+such as the HTTP redirection and the HTTP cookie processing. The implementation of the user 
+agent is based on the the HTTPS client interface of ARC, since the client interface of ARC 
+can support different protocols which are incarnated by different MCCs.
 
@@
 \label{sec:perfeval}
-The following section evaluate the performance of the
+The following sections evaluate the performance of the HED versus Axis2/C, as 
+well as the performance about the SAML2SSO profile, short-lived credential
+service, and delegation service.
 
@@
 performance comparison. 
-Two types of experiments are shown: session reused, session no-reused.
+Two types of experiments are shown: session reused, session non-reused.
 ``Session reused'' means all of the SOAP invocations on each client share the same session 
-(tcp connection); while ``Session no-reused'' means each invocation launches a new session.
+(tcp connection); while ``Session non-reused'' means each invocation launches a new session.
 
@@
 
-For ``Session no-reused'' experiments, the average response time in HED is almost the same as
+For ``Session non-reused'' experiments, the average response time in HED is almost the same as
 the value in Axis2/C, in both cases of client number is 10 (the value is 15ms) and 40 (the value
@@
 need to accomplish the SAML2SSO profile only once for all SOAP invocations on this client, 
-while the client with ``Session no-reused'' will need to accomplish this profile once per 
+while the client with ``Session non-reused'' will need to accomplish this profile once per 
 SOAP invocation.
 
-The average response time of ``Session no-reused'' experiments changes little with the execution 
+The average response time of ``Session non-reused'' experiments changes little with the execution 
 duration, and the value is about 730ms; while the throughput is around 82 requests per minutes.
 Considering that the average response time in case of 
-``Session no-reused'' and pure SOAP invocation is around 15ms when client number is 10 (see section
+``Session non-reused'' and pure SOAP invocation is around 15ms when client number is 10 (see section
 \ref{sec:perhedandaxis}), not demostrated in section \ref{sec:perhedandaxis}, the same value applies 
@@
 \label{sec:relatedwork}
-Similar work that integrates Shibboleth and Grid middleware does exist.
-GridShib~\cite{VWelch05,TBarton06} uses Shibboleth to get entity's attribute for the authorization of services developed in Globus Tookit.
-Unlike the work described in this paper, the GridShib does not use Shibboleth's native login handlers for authentication to Shibboleth IdP, instead, a new login handler specific for X.509 based authentication is developed for the Shibboleth IdP, and a credential service is developed for acting as an online-CA, authenticating the user, and issuing short-lived X.509 credential.
-Moreover, this paper provides a single sign-on solution for the user with community credentials authenticating to Grid services by directly using SAML2.0 SSO profile.
+There are other researches that utilize Shibboleth for protecting usage of Grid resources. 
+The work in ~\cite{J.Watt06} uses Shibboleth to protect a GridSphere portal and then 
+indirectly protect Grid services which are deployed in this GridSphere protal. Another 
+work in ~\cite{R.O.Sinnott06} also uses Shibboleth to protect Grid portal, moreover, it 
+maintains a few common used server certificates which can then be used to access Grid 
+resources after users have successfully authenticated themselves at Grid protal through
+Shibboleth. Unlike these two solutions that use Shibboleth to protect Grid portal, the solution
+in this paper uses Shibboleth or similar standards-based identity federation to directly protect 
+Grid resource/services, which provide possibility for those Grid applications that are directly
+based on Grid interface (e.g., applications that invokes SOAP operations of remote Grid services),
+to benefit from identity federation.
 
-SWITCH (The Swiss Education and Research Network) short-lived credential service (SLCS)~\cite{switchslcslink} provides a SLCS service and a client.
-The advantage of our work comparing to the SWTCH's solution is that the SLCS service and client are based on Web Services, so that they can be implemented in any Web Service container while keeping interoperability.
+GridShib~\cite{T.Scavo07,VWelch05} provides one solution that users can use X.509 
+certificate to authenticate to Shibbolth IdP, get back SAML assertion and embed it in 
+the proxy certificate. The Grid service will then use this SAML assertion for access control.
+Unlike the work described in this paper, GridShib still requires users to possess a X.509 
+credential, rather than username/password. GridShib also provides another 
+solution~\cite{TBarton06} in which a credential service (MyProxy server) is deployed together
+with Shibboleth IdP for acting as an online-CA, authenticating the user through username/password
+rather than X.509 certificate, and issuing short-lived X.509 credential. The second solution
+is similar to the short-lived credential service in our solution, while our solution
+also provides the option for users directly authenticating to Grid services via SAML2.0 SSO 
+profile using their username/password.
 
-In terms of delegation, apart from the credential delegation mechanism of GSI~\cite{IFoster98,VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol based on the WS-Trust~\cite{WSTrustlink} in order to provide a standard and interoperable protocol for the delegation in Grids.
-WS-Trust will also be adopted by the ARC middleware to express the specifications required to define delegation protocol in a standard way.
-Gridsite project implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web Service, with which ARC delegation client can be made easily interoperable.
+Another short-lived credential service is also provided by ~\cite{switchslcslink}. 
+The advantage of our work comparing to the work in ~\cite{switchslcslink} is that 
+our SLCS service is based on Web Services standard, so that it is easier to achieve 
+interoperability.
+
+In the aspect of credential delegation, apart from the credential delegation mechanism of 
+GSI~\cite{IFoster98,VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol 
+based on the WS-Trust~\cite{WSTrustlink} to provide a standard and interoperable protocol 
+for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the 
+specifications required to define delegation protocol in a standard way. Gridsite project also
+implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web 
+Service, with which ARC delegation client can be easily made interoperate.
 
@@
 \label{sec:conclusion}
-In this paper we presented the solution built into the ARC Grid middleware, aiming towards the cross-middle-ware authentication and single sign-on.
-The goal of our work is to develop an interoperable and accessible single sign-on solution to be used for a VO scenario in production Grids, where users can easily authenticate against the Grids with any of their credentials (not only X.509) from anywhere, and move inside Grid middleware infrastructure and even across heterogeneous Grids  boundaries, without re-authentication within a reasonable period.
-In order to do so, we implemented some standards based Web Services (including the SLCS service, the X.509 credential delegation service, and the SAML Token delegation service), as well as the user agent and the Service Provider in the context of the SAML2 SSO profile.
+In this paper we propose a Grid single sign-on framework that can utilize users' 
+federated identity instead of X.509 mutual authentication for authentication, which
+we believe can make further uptake of Grid resources as well as Grid middleware by 
+wider scale of users.
 
-It must be noted that due to the standard specifications utilized (SAML and WS-Security) which are independent from any particular middleware, the WS clients and Web/Grid Services developed within the ARC middleware can interoperate with WS clients and Web/Grid Services developed as part of some other middlewares in a standardized way.
+We implement the SAML2 SSO profile in the ARC Grid middleware, so that the Grid applications
+that developed by using the application interface provided by this middleware can easily 
+be switched to benefit from federated identity based authentication. 
+The performance evaluation shows even though the introducing of SAML2 SSO profile causes 
+much performance downgrade, this downgrade can be avoided by reusing the authentication 
+result for multiple SOAP invocations.
+Meanwhile, in order to be compatible with those Grid applications that require X.509 
+certificate on the client side, we also implement one Web Service which can be used for 
+obtaining an X.509 credential based on federated identity, and the other Web Service 
+which can be used for delegating X.509 credential. We believe that the evaluated 
+performance is enough for creating X.509 certificate and proxy certificate, considering that 
+the certificates should have a relaively much longer lifetime comparing to the response time of 
+these two services.
 
-Though only authentication issue is discussed in this paper, as the future work, we will implement a solution for cross-middleware attribute-based authorization by using the standard specifications such as SAML.
+Although only authentication issue has been discussed in this paper, we know single
+sign-in should also include the authorization issue. The SAML assertion which includes
+users' attributes can be used on the serrvice side to achiece attribute-based authorization.
+Therefore providing plugin on the service side to enforce access control using SAML attribute
+is considered one part of the future work.
+
+SimpleSAMLphp~\cite{simplesamllink} provides another implementation of SAML2 SSO profile, and it 
+has been widely deployed around nordic countries for identity federation. Hence integrating
+ARC based Grid applications with simpleSAMLphp is considered as another of future work.
 
@@
 
-\bibitem{IFoster01}
-I. Foster, C. Kesselman, S. Tuecke. The anatomy of the Grid: Enabling scalable virtual organizations, International Journal of Supercomputer Applications, 15(3), 200-222 (2001)
-
 \bibitem{IFoster98}
-I. Foster, C. Kesselman, G. Tsudik, and S. Tuecke, A Security Architecture for Computational Grids, ACM Conference on Computers and Security, 1998, 83-91.
+I. Foster, C. Kesselman, G. Tsudik, and S. Tuecke, A Security Architecture for 
+Computational Grids, ACM Conference on Computers and Security, 1998, 83-91.
 
 \bibitem{AlfieriR05}
-R.
-Alfieri, R.
-Cecchini, V.
-Ciaschini, L.
-dell’Agnello, A.
-Frohner, K.
-Lorentey, and F.
-Spataro.
-From gridmap-file to voms: managing authorization in a Grid environment.
-Future Generation Comp.
-Syst., 21(4), 549-558 (2005)
+R. Alfieri, R. Cecchini, V. Ciaschini, L. dell’Agnello, A. Frohner, K. Lorentey, 
+and F. Spataro. From gridmap-file to voms: managing authorization in a Grid environment. 
+Future Generation Comp. Syst., 21(4), 549-558 (2005)
 
@@
 \bibitem{MEllert07}
-M. Ellert et al. Advanced resource connector middleware for lightweight computational grids, Future Generation computer systems, 23(2), 219-240 (2007)
+M. Ellert et al. Advanced resource connector middleware for lightweight computational 
+grids, Future Generation computer systems, 23(2), 219-240 (2007)
 
@@
 
+\bibitem{WSElink}
+Web Services Enhancements 2.0 Service Pack 2. http://msdn.microsoft.com/en-us/library/
+
+\bibitem{A.Shoshan03}
+A. Shoshani, A. Sim, and J. Gu, Storage Resource Managers: Essential Components for the Grid, 
+Grid Resource Management: State of the Art and Future Trends, Kluwer Publishing, 2003.
+
 \bibitem{Shiblink}
@@
 
+\bibitem{R.O.Sinnott06}
+R.O. Sinnott, J. Jiang, J. Watt, and O. Ajayi. Shibboleth-based access to and 
+usage of grid resources. Proceeding of 7th IEEE/ACM International Conference on 
+Grid Computing, September 2006, Barcelona.
+
+\bibitem{J.Watt06}
+J. Watt, O. Ajayi, J. Jiang, J. Koetsier, R.O. Sinnott. A Shibboleth-Protected 
+Privilege Management Infrastructure for e-Science Education. in proceeding of Sixth 
+IEEE International Symposium on Cluster Computing and the Grid (CCGRID'06), May 2006,
+Singapore.
+
+\bibitem{T.Scavo07}
+T. Scavo and V. Welch. A Grid Authorization Model for Science Gateways. International 
+Workshop on Grid Computing Environments, 2007.
+
 \bibitem{VWelch05}
-V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. 4th Annual PKI R\&D Workshop, 2005.
+V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and 
+Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. in 
+proceeding of 4th Annual PKI R\&D Workshop, 2005.
 
 \bibitem{TBarton06}
-T. Barton et al. Identity Federation and Attribute-based Authorization through the Globus Toolkit, Shibboleth, Gridshib, and MyProxy. 5th Annual PKI R\&D Workshop, 2006.
+T. Barton et al. Identity Federation and Attribute-based Authorization 
+through the Globus Toolkit, Shibboleth, Gridshib, and MyProxy. in proceeding 
+of 5th Annual PKI R\&D Workshop, 2006.
 
@@
 \bibitem{VWelch04}
-V. Welch et al. X.509 proxy certificate for dynamic delegation, Proceeding of the 3rd Annual KI R\&D Workshop, 2004.
+V. Welch et al. X.509 proxy certificate for dynamic delegation. in proceeding 
+of the 3rd Annual KI R\&D Workshop, 2004.
 
 \bibitem{MAhsant04}
-M. Ahsant, J. Basney, and O. Mulmo. Grid Delegation Protocol, UK Workshop on Grid Security Experiences, Oxford. 2004.
+M. Ahsant, J. Basney, and O. Mulmo. Grid Delegation Protocol. UK Workshop on 
+Grid Security Experiences, Oxford. 2004.
 
@@
 
+\bibitem{simplesamllink}
+http://rnd.feide.no/simplesamlphp
+
 \end{thebibliography}

### 2009-04-20 doc/a-rex/arex_tech_doc.tex
Documented extra lines in job.ID.local and extra config lines

@@
 \item \textit{jobname} -- name of the job as supplied by the user
+\item \textit{projectname} -- name of the project as supplied by the user
 \item \textit{lrms} -- name of the LRMS backend to be used for local submission
@@
 reached state \textbf{InLRMS})
+\item \textit{globalid} -- BES ActivityIdentifier XML tree, the global
+identifier of the job
 \item \textit{args} -- executable name followed by a list of command-line
@@
 The A-REX will also send job records to this service in addition to
-the default logger service configured in the configuration (This feature
-does not work in current implementation).
+the default logger service configured in the configuration.
 \item \textit{activityid} -- Job-id of previous job in case the job has been 
@@
 is specified in days. Last specified value becomes effective.
+\item \textbf{\textit{jobreport\_credentials}}\textit{={[}key\_file
+    {[}cert\_file {[}ca\_dir]]]} -- specifies the credentials for
+  accessing the authentication service.
+\item \textbf{\textit{jobreport\_options}}\textit{={[}options]}
+  -- specifies additional options for Usage Reporter and/or
+  accounting service. The \textit{options} string is interpreted by Usage
+  Reporter, its format is described in the corresponding technical document.
 \item \textbf{\textit{securetransfer}}\textit{=yes|no} -- specifies whether

### 2009-04-20 doc/usagereporter/aur-tech-draft.tex
Created document for Usage Reporter

--- 2008-08-05 doc/client/arc-ws-clients.tex
@@
 \usepackage{longtable}                             %for breaking tables
-%\usepackage{bibtex}
-\usepackage{moreverb}
 %
@@
 \hypersetup{
-  pdfauthor = {A.U. Thors},
-  pdftitle = {WS-Based ARC Clients},
+  pdfauthor = {Author Name},
+  pdftitle = {Paper title},
   pdfsubject = {Paper subject},
@@
 \newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
-\newcommand{\uicommand}[1]{\textbf{\sffamily #1}}
 \definecolor{shadecolor}{rgb}{1,1,0.6}
@@
 
-%-------- Change this to NORDUGRID-XXXXXXX-NN
+%-------- Change this to NORDUGRID-TECH-NN
 
-{\raggedleft NORDUGRID-TECH-20\par}
+{\raggedleft NORDUGRID-TECH-NN\par}
 
@@
 %%%%---- The title ----
-{\centering \textsc{\Large WS-Based ARC Clients}\Large \par}
+{\centering \textsc{\Large ARC accounting component}\Large \par}
 \vspace*{0.5cm}
@@
 %%%%---- A subtitle, if necessary ----
-%{\centering \textit{\large Subtitile}\large \par}
+{\centering \textit{\large Technical document}\large \par}
 
@@
 %%%%---- A list of authors ----
-{\centering \large A.U. Thors\footnote{authors@gmail.com} \large \par}
+    {\centering \large P\'eter D\'ob\'e \footnote{dobe@iit.bme.hu} \large \par}
 
 %%%%---- An abstract - if style is article ----
-\begin{abstract}
-This NorduGrid Technical document gives an introduction on how to use
-the client tools developed for WS-based ARC solution.
-\end{abstract}
+%\begin{abstract}
+%The abstract
+%\end{abstract}
 \end{titlepage}
 
-\tableofcontents                          %Comment if use article style
+%\tableofcontents                          %Comment if use article style
 \newpage
-%\chapter{Preface}
-\section{Introduction}                    %Use Sections for articles
-\label{sec:intro}
 
-This document describes a set of simple client tools for the new Web
-Service~\cite{ws} (WS) based Advanced Resource Connector~\cite{arc}
-(ARC) middleware. These tools are prototypes and are intended to be a
-proof-of-concept for the new ARC.
+\section{Introduction}
 
-So far, the only service that is available for the new ARC is the ARC
-Resource-Coupled Execution Service~\cite{arex} (A-REX). Therefore,
-this prototype release of WS clients consists of job handling tools
-with the following functionality:
-\begin{itemize}
-\item submission of jobs;
-\item querying the status of jobs; and
-\item termination of jobs.
-\end{itemize}
+\section{Purpose}
+  metering resource usage
+  record it in a database
+    SGAS service used
+    standard OGF Usage Records
+  possibly billing
+  job-level accounting
+  user-based accounting
+  project name
 
-The names of the command line tools begin with ``ap'', which shall be
-interpreted as ``ARC Prototype'' to distinguish them from the
-full-fledged production tools.
+\section{Architecture}
+  (diagram)
 
-Support for staging of data would also have been desirable, but this
-is not yet fully implemented on the server side. Therefore it could
-not be included in this client prototype either. Thus, it is not
-possible to run any jobs that depend on local files on the computer on
-which the submitting client runs. Neither is it possible to retrieve
-any form of output (not even stdout/stderr) by using the new client.
+  A-REX
+  logger
+  SGAS service
 
-Behind the command line tools, there is a minimalistic library which
-is documented in Appendix~\ref{app:library}. The library will serve as
-a basis for the development of a generalized ARClib. This will be used
-by the new client tools of the production releases of the next
-generation of ARC.
+\section{Operation}
 
-\section{Configuration}
-\label{sec:config}
+  \subsection{Invocation}
+    stand-alone exe
+    run by A-REX hourly
+    command line arguments
+    no config, everything from A-REX
 
-The tools described here do not perform any brokering, but submit jobs
-to a specific A-REX service. Which A-REX service to use is specified
-in a configuration file. The configuration file also specifies how to
-set up the communication chain for the client. The location of the
-configuration file is specified by the environment variable
-\verb:ARC_AREX_CONFIG:. If there is no such environment variable, the
-configuration file is assumed to be \verb:arex_client.xml: in the
-current working directory. An example configuration file is shown in
-Appendix~\ref{app:config}.
+  \subsection{Parsing job log files}
+    ``id.randomletters''
+    UR generated from them
 
-\section{Job Submission}
-\label{sec:apsub}
+  \subsection{Accessing LUTS}
+    insert method
+    certificates from:...
 
-The \verb#apsub# command is used to submit jobs to an A-REX service.
-
-\begin{shaded}
-\paragraph{Usage} \verb:apsub <JSDL-file> <JobID-file>:
-\end{shaded}
-
-\paragraph{Arguments}
-\begin{description}
-\item[\texttt{<JSDL-file>}] The name of a file that contains
-the job specification in JSDL format.
-\item[\texttt{<JobID-file>}] The name of a file in which the
-Job ID will be stored. This file is created when the command is
-used. The reason for storing the Job ID in a file is its length
-(approximately 500 characters) which makes it unsuitable to use as a
-command line argument for the other tools.
-\end{description}
-
-\paragraph{Output} A message that the job was submitted and in which
-file the Job ID is stored if the submission succeded. If the
-submission failed for some reason, an error message is displayed.
-
-\section{Job Status Query}
-\label{sec:apstat}
-
-The \verb#apstat# command is used to query the status of a job that
-has been submitted to an A-REX service.
-
-\begin{shaded}
-\paragraph{Usage} \verb:apstat <JobID-file>:
-\end{shaded}
-
-\paragraph{Arguments}
-\begin{description}
-\item[\texttt{<JobID-file>}] The name of a file in which the
-Job ID is stored. This file was created by \verb:apsub: when the job
-was submitted.
-\end{description}
-
-\paragraph{Output} The status of the job, e.g. ``Pending/Accepted'',
-or an error message if the status request failed for some reason.
-
-\section{Job Termination}
-\label{sec:apkill}
-
-The \verb#apkill# command is used to terminate a job that has been
-submitted to an A-REX service.
-
-\begin{shaded}
-\paragraph{Usage} \verb:apkill <JobID-file>:
-\end{shaded}
-
-\paragraph{Arguments}
-\begin{description}
-\item[\texttt{<JobID-file>}] The name of a file in which the
-Job ID is stored. This file was created by \verb:apsub: when the job
-was submitted.
-\end{description}
-
-\paragraph{Output} A message that the job was terminated, or an error
-message if the termination failed for some reason.
+\section{Implementation}
+  c++
+  sw dependencies
+    full arc
+  install location
 
@@
 
-\include{ws_client_lib}
+\section{Configuration}
+  jobreport
+  jobreport\_options
+  jobreport\_credentials
 
-\section{Configuration Example}
-\label{app:config}
-Figure~\ref{fig:config} shows an example configuration file, which is
-an XML file. All configuration information is enclosed within an
-\verb:ArcConfig: element. That element also specifies which XML
-manespaces to use.
+\section{Usage Record properties}
+  input file format
+  filled properties
+  missing properties
 
-The \verb:ModuleManager: element contains paths to directories where
-loadable plugins can be found. In this case, there is only one path --
-\verb:/usr/lib/arc/: -- but there may be many.
-
-The \verb:Plugins: elements specify which plugins to load -- in this
-case message chain components (MCCs) for TCP, TLS, HTTP and SOAP.
-
-The \verb:Chain: element, finally, defines how to combine the MCCs
-that have been loaded into a communication chain. In this case, a TCP
-MCC that connects to port 6000 on the local host is connected to a TLS
-MCC, which is connected to a HTTP that uses the POST method, and
-finally a SOAP MCC is connected to the HTTP MCC.
-
-The client tools will use the SOAP MCC as an entry point when sending
-messages. The messages will then propagate through the MCCs in the
-chain as specified in the configuration file.
-
-\begin{figure}
-\begin{shaded}
-\verbatiminput{arex_client.xml}
-\end{shaded}
-\caption{An examle of an \texttt{arex\_client.xml} configuration file.}
-\label{fig:config}
-\end{figure}
-
-\clearpage
 \bibliography{grid}
-
 \end{document}

### 2009-04-21 doc/client/UnicoreSupport.tex
Adding documents on Unicore

\documentclass[11pt,oneside,a4paper,english]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{graphicx,epic,eepic}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{a4wide}
\usepackage{xspace}

\newcommand{\arcmigrate}{\texttt{arcmigrate}\xspace}

\begin{document}
%\maketitle
\section*{Support for Unicore resource discovery and job submission/management}
The next generation ARC client library will offer support for discovering grid resource running Unicore middleware and to submit and manage grid jobs (basic functionality) at these resources. This writeup documents the present status of the Unicore support, what is missing and what is the work plan to implement the missing functionality.
\subsection*{Versions of Unicore supported}
Text here about the supported versions
\subsection*{Status of ongoing work}
\subsubsection*{Resource discovery}
Resource discovery. Can we find all resources? Are there defaults in the config file? Can we get detailed information from the clusters?
\subsubsection*{Job submission and management}
Can we submit jobs? Can we get jobs? Can we kill jobs?

\subsection*{Work to be done}
I guess this section is is to write once the present status is documented. Any depencies? i.e. work that has to be done be others?

\subsection*{Work Plan}
Assuming that the Unicore work can start fulltime again first week of May. How much time would be needed in order to finalize the work listed above in the work to be done section. What should be done first etc.

\end{document}

### 2009-04-21 doc/usagereporter/aur-tech-draft.tex
Added to Usage Recorder doc

@@
 
-\section{Introduction}
+\section{Purpose}
 
-\section{Purpose}
-  metering resource usage
-  record it in a database
-    SGAS service used
-    standard OGF Usage Records
-  possibly billing
-  job-level accounting
-  user-based accounting
-  project name
+The ARC Usage Reporter tool is a component implementing accounting
+functionality in the ARC middleware. Its objective is to gather
+metered resource usage data for each job and submit it to an
+accounting service along with the job submitter's identity and
+miscellaneous job-related metadata.
+
+The accounting service stores the received usage data in a database,
+and provides an interface for querying it.  Queries can be made by the
+consumers of the accounting data, such as a billing component.  The
+service itself is a third party application, separate from the
+middleware distribution. The ARC Usage Reporter is currently capable
+of using the logging service of the SweGrid Accounting System (SGAS),
+but maintaining the possibility to enable utilizing other services
+have been kept in mind during design.
+
+Before the usage data collected from the resource manager is
+submitted, it is transformed into records of job-level granularity. To
+every job corresponds exactly one Grid user, therefore reports over a
+time period (e.g.~an invoice) can be generated per-user, or
+alternatively on a larger scale such as job project or VO level.
 
@@
   (diagram)
+
+\begin{figure}[ht]
+\centering{{\scalebox{0.9}{\includegraphics{usagereporting.png}}}
+\caption{\label{fig:myfigure1}Components participating in accounting.} }
+\end{figure}
 

### 2009-04-22 doc/jura/jura-tech-doc.tex
Usage Reporter renamed to JURA in doc + progress in text

--- 2009-04-21 doc/usagereporter/aur-tech-draft.tex
@@
 %%%%---- The title ----
-{\centering \textsc{\Large ARC accounting component}\Large \par}
+{\centering \textsc{\Large ARC Accounting Component -- JURA}\Large \par}
 \vspace*{0.5cm}
@@
 
-The ARC Usage Reporter tool is a component implementing accounting
-functionality in the ARC middleware. Its objective is to gather
-metered resource usage data for each job and submit it to an
-accounting service along with the job submitter's identity and
-miscellaneous job-related metadata.
+The \textit{Job Usage Reporter for ARC} (JURA) is a component
+implementing accounting functionality in the ARC middleware. Its
+objective is to gather metered resource usage data for each job and
+submit it to accounting services along with the job submitter's
+identity and miscellaneous job-related metadata.
 
@@
 service itself is a third party application, separate from the
-middleware distribution. The ARC Usage Reporter is currently capable
-of using the logging service of the SweGrid Accounting System (SGAS),
-but maintaining the possibility to enable utilizing other services
-have been kept in mind during design.
+middleware distribution. JURA is currently capable of using the
+logging service of the SweGrid Accounting System (SGAS)\cite{sgas},
+but maintaining the possibility to enable utilizing other services has
+been kept in mind during design.
 
@@
 \section{Architecture}
-  (diagram)
 
 \begin{figure}[ht]
-\centering{{\scalebox{0.9}{\includegraphics{usagereporting.png}}}
-\caption{\label{fig:myfigure1}Components participating in accounting.} }
+\centering{{{\includegraphics[width=0.9\textwidth]{components.pdf}}}
+\caption{\label{fig:components}Components participating in accounting.} }
 \end{figure}
 
-  A-REX
-  logger
-  SGAS service
+JURA offers a complete replacement in functionality for the old
+\textit{logger} utility\cite{logger}, a job metadata logging tool with
+a purpose very closely related to that of this new tool. However,
+backwards compatibility is maintained, so the old logger can still be
+deployed.
+
+The ARC execution manager, A-REX\cite{arex} initiates JURA in a
+similar way to invoking logger. JURA reads job log files provided by
+A-REX. These files have the same format as those meant for logger,
+with additional lines which are necessary for accounting scenarios.
+
+It also acts as a client for one or more accounting services,
+specifically SGAS Logging and Usage Tracking Services (LUTS's),
+inserting the generated records in batches. (See Figure
+\ref{fig:components}.)
 
@@
 \subsection{Invocation}
-    stand-alone exe
-    run by A-REX hourly
-    command line arguments
-    no config, everything from A-REX
+
+JURA is a stand-alone executable application, executed hourly by
+A-REX. It has no separate configuration file; it gets all necessary
+configuration options from A-REX, in part through command-line
+arguments, but mostly via lines in the job log files (see Appendix
+\ref{config} for details). The source of the latter are lines in the
+grid-manager configuration file.
+
+The command line format is a subset of that of logger:
+
+\verb|jura [-E <expiration_time>] <control_dir>|
+
+where \textit{expiration\_time} is the validity length of job log
+files in days, after which time they are considered invalid;
+\textit{control\_dir} is the A-REX control directory for a mapped
+local UNIX user.
 
 \subsection{Parsing job log files}
-    ``id.randomletters''
-    UR generated from them
+\label{joblogs}
+The job logs generated by A-REX reside under the directory
+\verb|<control_dir>/logs|. They have file name format
+\verb|<ngjobid>.<random>|, where \textit{ngjobid} is the identifier
+created for the job by A-REX, \textit{random} is a randomly generated
+sequence of alphanumeric characters to avoid collision of different
+files pertaining to the same job.
+
+2 for each logging destination: when submitted, when finished
+
+UR generated from them: only from finished
 
 \subsection{Accessing LUTS}
+
 insert method
+
 certificates from:...
+
+\section{Security}
+  runs as: same user as a-rex, typically root
+
+  accesses sensitive data in job logs
+
+  sgas security:
+
+    standard ssl cert, possibly proxy
+
+    no other attributes but DN considered
+
+    access control: flat-file based + XACML-like(?) configured thru
+    SAM porttype
+
+  in this case: service credentials, no proxy
 
@@
   c++
+
   sw dependencies
+
     full arc
+
+    ssl
+
   install location
@@
 \section{Configuration}
+\label{config}
   jobreport
+
   jobreport\_options
+
   jobreport\_credentials
@@
 \section{Usage Record properties}
+\label{log2ur}
   input file format
+
   filled properties
+
   missing properties

### 2009-04-22 doc/storage_system/arc-storage-documentation.tex
updating the storage documentation

@@
 A storage node here is a network-accessible computer having storage space to share, and a running storage element service (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.).
-\note[JKN]{TODO: Rephrase this sentence; "one of the services" seems vague}For each storage node one of the services of the ARC storage system is needed to manage it and to integrate it into the system.
-The client side provides access to third-party storage solutions through the namespace of the ARC storage system.
+On each storage node a Chelonia service is needed to be installed to manage and integrate it into the system.
+The client side provides access to third-party storage solutions through the namespace of Chelonia.
 The main services of the storage system are the following (see Figure~\ref{fig:services}):
@@
 \centering{{\scalebox{0.8}{\includegraphics{arc-storage-services.pdf}}}
-\caption{\label{fig:services}{\bf TODO:} explain figure, since components are not yet introduced.
-The components of the ARC storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
+\caption{\label{fig:services} The components of Chelonia: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
 \end{figure}
@@
 Compared to the well-known structure of local file systems, these GUID's are very similar to the concept of \emph{inode}s.
-And as a directory on a local filesystem is basically just a list of name and inode pairs, a collection on the ARC storage is just a list of name and GUID pairs.
+And as a directory on a local filesystem is basically just a list of name and inode pairs, a collection in Chelonia is just a list of name and GUID pairs.
 There is a dedicated collection which is the \emph{root collection}.
-This makes the namespace of the ARC storage system a hierarchical namespace where you can start at the root collection, and go to sub-collections and sub-sub-collections to get to a file.
+This makes the namespace of Chelonia a hierarchical namespace where you can start at the root collection, and go to sub-collections and sub-sub-collections to get to a file.
 This path is called the \emph{Logical Name} (LN).
@@
 The root collection knows the GUID of \verb#why#, and the (sub-)collection \verb#why# knows the GUID of \verb#blue#.
-If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, a concept very similar to a hardlink in a \note[JKN]{Todo: Posix file system?}regular local file system.
+If the GUID of this file is \verb#5678#, and somebody makes another entry in collection \verb#/why# (= \verb#0/why#) with name \verb#red# and GUID \verb#5678#, then the \verb#/why/red# LN points to the same file as \verb#/why/blue#, a concept very similar to a hardlink in a regular local (POSIX-like) file system.
 
@@
 The user can specify how many replicas are required for each file.
-The ARC storage system periodically checks the number of replicas\note[JKN]{TODO: How will this scale?}, and automatically creates new replicas if there are fewer than required, and removes replicas if there are more.
-
-The different replicas of a file can be in different \note[JKN]{TODO: which states?
-a reference to definition of states maybe?}states.
-The replica can be valid and alive or just in the process of creation or it may be corrupt or a whole storage node may be offline.
+The storage nodes periodically check the number of replicas for each file they store, and automatically creates new replicas if there are fewer than required, and removes replicas if there are more.
+(Currently if a storage node stores a great amount of files, then the checking period could be longer.)
+
+The file replicas can be in the following states: the replica can be valid and alive or just in the process of creation or it may be corrupt or a whole storage node may be offline.
 This state is always stored as metadata of the given replica.
@@
 If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state.
-And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat triggers the modification of metadata as well.
+And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat also triggers the modification of metadata.
 
@@
 All metadata about files and collections in the storage system are stored in the A-Hash.
-Additionall system configutation information (e.g.
+Additional system configutation information (e.g.
 about A-Hash replication, about Shepherd services, etc.) is stored in it as well.
@@
 The Bartender service provides the high-level interface of the storage system to clients.
-Every interaction between a client and the ARC storage system starts by sending a request to a Bartender.
+Every interaction between a client and the storage system starts by sending a request to a Bartender.
 Using the Bartender it is possible to create and remove collections, create, get and remove files, move files and collections within the namespace using Logical Names, etc.
@@
 Any number of independent Bartender services may be deployed in the system to ensure high-availability and load-balancing.
-The Bartender also provides access to files in third-party storage solutions through its interface by mounting the namespace of the third-party storage into the namespace of the ARC storage (this is accomplished by so-called `gateway' modules).
+The Bartender also provides access to files in third-party storage solutions through its interface by mounting the namespace of the third-party storage into the namespace of Chelonia (this is accomplished by so-called `gateway' modules).
 
@@
 But these internal services are communicating with each other, thus they have to know which other services to trust.
-We call this aspect of the security architecture of the ARC storage system `inter-service authorization'.
+We call this aspect of the security architecture of Chelonia `inter-service authorization'.
 
@@
 
-The Librarian component of the ARC storage system stores all the metadata about files and collections.
+The Librarian component of Chelonia manages all the metadata about files and collections.
 For each file and collection there are access policies in the form of access control rules, and these are stored among the metadata.
@@
 
-\note[JKN]{TODO: Short intro to this chapter mentioning that this chapter contains a partial list of use cases and that not all features are covered.}
+This chapter contains some examples to demonstrate the internal mechanism of Chelonia.
 
@@
 This metadata contains the location of the file's replicas and the access policies of this file.
-    \item The Bartender checks, based on the access policies and the user's identity, if he is allowed to get the file, and if allowed, it chooses \note[JKN]{TODO: We need to define the choosing algorithm somewhere} a replica location.
+    \item The Bartender checks, based on the access policies and the user's identity, if he is allowed to get the file, and if allowed, it chooses a replica location (currently randomly).
 A location consists of the URL of a Shepherd service, and the ID of the replica within that Shepherd (which is called a `referenceID').
@@
     \item Then client tool can upload the file to this TURL.
-    \item The Shepherd detects\note[JKN]{TODO: How does it detect it?
-Callback from the transfer service?} that the file has been uploaded.
-It checks the checksum of the file, and if it is OK, then it reports to the Librarian, that this replica is now \verb#alive#.
+    \item The Shepherd detects that the file has been uploaded.
+In order to do this the storage element service does post processing (e.g.
+remove a hardlink, do a SOAP call, etc.) after it got the file.
+The Shepherd checks the checksum of the file, and if it is OK, then it reports to the Librarian, that this replica is now \verb#alive#.
 The Librarian alters the state of this location, and now the file has one valid replica.
@@
 
-The ARC storage system stores different kinds of metadata about files, collections, mount points, Shepherd services, etc. Each of these has a unique ID; a GUID.
+The storage system stores different kinds of metadata about files, collections, mount points, Shepherd services, etc. Each of these has a unique ID; a GUID.
 
@@
     \begin{itemize}
-        \item \emph{externalURL}: the URL of the external storage end point \note[JKN]{TODO: an example URL maybe?}.
+        \item \emph{externalURL}: the URL of the external storage end point.
     \end{itemize}
@@
 When there is an incoming connection to a Bartender, the security framework of the HED extracts the identity information of the connecting client (DN of client, DN of the issuer CA, extended attributes, etc.), and all this information is accessible by the Bartender.
-When the given method of the Bartender gets to the point where the access rules are present, and it is clear that what kind of action the user wants to do, then either\note[JKN]{TODO: Where is the 'or' that should follow the either?} the request and the policy is converted to the native ARC policy language, and then the policy evaluator of the security framework is called to make the decision.
-Within the native ARC policy language, the XML representation of the requests and policies the storage actions are put into an `Attribute' element with the ID `\verb!http://www.nordugrid.org/schemas/policy-arc/types/storage/action!'\note[JKN]{TODO: This sentence needs more explanation; too technical.}.
+When the given method of the Bartender gets to the point where the access rules are present, and it is clear that what kind of action the user wants to do, then both the request and the policy are converted to the native ARC policy language, and then the policy evaluator of the security framework is called to make the decision.
+The native ARC policy language defines attribute types for attributes within requests and policies.
+The XML representation of an ARC policy or request contains several `Attribute' elements, each have a type and a value.
+The storage actions are also put into Attribute elements with the type `\verb!http://www.nordugrid.org/schemas/policy-arc/types/storage/action!'.
 
@@
 
-\note[JKN]{TODO: This paragraph is very unclear.
-All the 'could's makes it look like the TrustManager is just something planed in the far future.
-Are the attributes mandatory or optional?}The TrustManager section itself could have some attributes: \textbf{FromFile} could have a filename, if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.
+The TrustManager section itself have the following optional attributes: \textbf{FromFile} can specify a filename if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no', default: `yes') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.
 
@@
     \item[StoreClass] specifies the name of the store class, used to store the data on disk.
-Currently there are several store classes \note[JKN]{TODO: describe the stores and how to choose among them}(\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore!
+Currently there are several store classes each has the same interface but uses different mechanisms (\verb!PickelStore!, \verb!CachedPickleStore!, \verb!StringStore!, \verb!ZODBStore!, \verb!TransDBStore!), however only the \verb!TransDBStore!
 class is suitable for use in a replicated deployment.
-All the others are only for a centralized deployement.
-This variable is only configurable for the centralized deployement.
+All the others are only for a centralized deployement, so this variable is only configurable for the centralized deployement.
+Based on our tests, currently the \verb!CachedPickleStore!
+has the best performance.
     \item[StoreCfg] contains parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash can save its files.
@@
 The Librarian knows about the type of possible entries in the namespace of Chelonia: files, collections and mount points.
-It knows that these are organized in a tree-hiearchy with entries grouped into collections, and it knows about Logical Names, and how to traverse a Logical Name to find the metadata of the file or collection (or mount point) which it refers to.
+It knows that these are organized in a tree-hierarchy with entries grouped into collections, and it knows about Logical Names, and how to traverse a Logical Name to find the metadata of the file or collection (or mount point) which it refers to.
 However, the Librarian does not maintain the namespace of Chelonia.
@@
 
-The end-user of the ARC storage system never communicates directly with a Librarian, the Bartenders will contact the Librarians in order to fulfil the user's requests.
+The end-user of Chelonia never communicates directly with a Librarian, the Bartenders will contact the Librarians in order to fulfil the user's requests.
 
@@
 This ensures that the Shepherd does not use up all the resources of the hosting machine.
-After checking the checksum of a file, the Shepherd will always wait this long before checking the next file, even if this means that checking all the files takes longer than the `CheckPeriod'\note[JKN]{TODO: Which takes precedense; CheckPeriod or MinCheckInterval?}.
+After checking the checksum of a file, the Shepherd will always wait this long before checking the next file, even if this means that checking all the files takes longer than the `CheckPeriod' (in other words `MinCheckInterval' overrides `CheckPeriod' if the number of stored file is greater.)
     \item[CreatingTimeout] specifies in seconds the time before the Shepherd decides that a started file upload has failed.
@@
     \item[BackendCfg] specifies configuration parameters for the backend class: \textbf{DataDir} and \textbf{TransferDir} are the directories the backend will use (see Section~\ref{sub:shepherd_implementation}), \textbf{TURLPrefix} is the URL of the storage element service ending with a `/'.
-    \item[LibrarianURL] is the URL of a Librarian where the Shepherd should send the reports\note[JKN]{TODO: Which reports?}.
+    \item[LibrarianURL] is the URL of a Librarian where the Shepherd should send the reports about the health of the stored files.
     \item[BartenderURL] is the URL of a Bartender to which the Shepherd should connect if it wants to offer a replica for replication or recover a corrupted replica.
@@
 
-    \item[makeMountpoint(makeMountpointRequestList)] creates a mountpoint to an external storage element.
-
-    \emph{makeMountpoint} is quite similar to the \emph{makeCollection}.
-It is used to provide access to an external storage system within the global namespace.
-\emph{makeMountpointRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{URL}, \emph{metadata}) pairs where \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples.
-\emph{section} contains the initial content of the catalog in the form of name-GUID pairs (these entries are hard links to the given GUID's with the given name).
-\emph{LN} is the requested logical name and \emph{URL} is the path to the external storage element to be accessed.
-The \emph{policies} section contains access policies, and the metadata section contains any other \emph{metadata} in key-value pairs.
-
-    The \emph{status} in the response is one of `\textbf{done}', `\textbf{mountpoint exists}', `\textbf{parent does not exist}', `\textbf{internal error}', `\textbf{failed}: credentials not found' or `\textbf{cannot create anything in mountpoint}'.
+    \item[makeMountpoint(makeMountpointRequestList)] creates a mountpoint which is used to provide access to a third-party storage system within the global namespace.
+
+    \emph{makeMountpointRequestList} is a list of (\emph{requestID}, \emph{LN}, \emph{URL}, \emph{metadata}) tuples where \emph{LN} is the requested logical name, \emph{URL} is the URL to the third-party storage system to be accessed, and \emph{metadata} is a list of (\emph{section}, \emph{property}, \emph{value}) tuples which could contain additional metadata..
+
+    The response is a list of (\emph{requestID}, \emph{success}) pairs, where \emph{success} is one of `\textbf{done}', `\textbf{LN exists}', `\textbf{parent does not exist}', `\textbf{failed to create new catalog entry}', `\textbf{failed to add child to parent}' or `\textbf{internal error}' or `\textbf{cannot create anything in mountpoint}'.
 
@@
 
-    \emph{unmakeMountpointRequestList} is a list of (requestID, LN) pairs with the logical names of the mountpoints.
-\emph{status} is one of `\textbf{removed}', `\textbf{no such LN}', `\textbf{Failed:} Credential not available'.
-
-    \item[CredentialDelegation(CredentialDelegationRequest)] delegates credentials.
-\note[JKN]{TODO: This item has new style. Rewrite to make it similar to all the previous items. Also, second possible response looks very strange!}    
-    From the client side \emph{createDelegation()} is called which interns initiate the \emph{CredentialDelegationInitRequest} for the service.
-After the successful handshake this method initiates an other request of \emph{UpdateCredentials(..)}, which completes this process.
-For complete details see ARC-Security-Document.
-Response from this interface could be \emph{done}, \emph{cannot delegate credentials, requested bartender does not supports the gateway}
-
-    \item[RemoveCredentials(RemoveCredentialRequest)] returns (\emph{message}, \emph{status})
-\note[JKN]{TODO: This item has new style. Rewrite to make it similar to all the previous items. Also, the responses look very strange!}    
-    \emph{RemoveCredentialRequest} initiates the process of deleting the user credential from the service side.
-The response could be \emph{Credential removed successfully}, \emph{cannot access proxyFile or ProxyStore problem at service side}
-
-\end{description}
+    \emph{unmakeMountpointRequestList} is a list of (requestID, LN) pairs where \emph{LN} is the logical names of the mountpoint.
+
+    The response is a list of (\emph{requestID}, \emph{status}), where status could be `\textbf{removed}', `\textbf{no such LN}', `\textbf{denied}'.
+
+    \item[removeCredentials()] removes a previously delegated proxy certificate.
+
+    This method has no arguments, it removes the proxy certificate which is previously delegated by the same user.
+
+    The response is a (\emph{message}, \emph{status}) pair, where \emph{status} could be `\textbf{successful}' or `\textbf{failed}', and \emph{message} contains additional details.
+\end{description}
+
+The Bartender implements the \textbf{DelegateCredentialsInit} and \textbf{UpdateCredentials} methods as well to be able to accept credentials delegation.
+These two operations are described in this WSDL document: http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/libs/delegation/delegation.wsdl
 
@@
 
-The Bartender utilizes gateway modules to access file listings and transfer URL's from third-party storage elements.
+The Bartender utilizes gateway modules to access file listings and transfer URL's from third-party storage systems.
 It works as an optional backend component of the Bartender service.
-The \emph{Gateway} module is based on four basic methods; \emph{get}, \emph{put}, \emph{list} and \emph{remove}.
-Since the task is to contact external storage element, which requires the user to delegate the credentials so that it can communicate on behelf of the user\note[JKN]{TODO: What does this sentence mean?
-Please rewrite.}.
-
-In order to access the files and directories from the external storage element, limited user credentials are passed to the \emph{Bartender} service.
-Users can also remove their \emph{credentials} from the service side.
-These credentials are used in the methods \emph{get}, \emph{put}, \emph{remove} and \emph{list} of the \emph{Gateway} module.
-These limited credentials are not used in any type of transfer.
-While listing files and directories, the Gateway uses the user credential, whereas in the \emph{get} and \emph{put} methods, it only uses the credentials to check the validity of URL's before sending them to the user.
+The \emph{Gateway} module is based on four basic methods: \emph{get}, \emph{put}, \emph{list} and \emph{remove}.
+This module should connect to the third-party storages on behalf of the user, so the user has to delegate his or her credentials to the Bartender using proxy delegation.
+In order to access the files and directories on the external storage system, limited user credentials are passed to the Bartender service.
+Users can also remove their credentials from the service side.
 % subsection implementation (end)
@@
 
-In the first prototype release\note[JKN]{TODO: Mention the arc\_storage\_cli in the intro of this document} there is one client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run.
+In the prototype release there is a command-line client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run.
 It is capable of communicating with a given Bartender service, and can upload and download files via HTTP.
@@
 
-Note here that long listing (\verb!arcls -l!)! with the \verb!arc! protocol not yet supports time stamps. Hence, only the type of the entry (here "file") is shown.
-
-\note[JKN]{TODO: There is nothing about status or plans. Either add some status and plans or change subtitle of document.}
+Note here that long listing (\verb!arcls -l!)! with the \verb!arc! protocol not yet supports time stamps. Hence, only the type of the entry (here `file') is shown.
 

### 2009-04-23 doc/storage_system/arc-storage-manual.tex
implementing the requested changes in the internal review comments

@@
 \tableofcontents                          %Comment if use article style
-% \begin{verbatim}
-%     Administrator's and user's manual
-%     ---------------------------------
-% 
-%     Administrator's manual
-%     - Quick start guide
-%         - Checking the code out from the subversion
-%         - Compiling and installing
-%         - Modifying the configuration template
-%         - Running the daemon
-%         - Using the CLI tool
-%     - Installation:
-%         - Getting the source code from subversion
-%         - Compilation details
-%         - Installation
-%     - Configuration:
-%         - HED service config
-%         - Centralized A-Hash
-%         - Distributed A-Hash
-%         - Librarian
-%         - Shepherd
-%         - Hopi
-%         - Bartender
-%         - Configuring inter-service trust: local listing of DNs, or using an A-Hash to store them
-%         - Virtual organizations: how to configure HED to trust a VO
-%     - Deployment scenarios:
-%         - Single node: putting every service on one single node (including a storage node with Hopi)
-%         - Centralized management: using seperate multiple storage nodes with Hopi, but leaving a central A-Hash, a Librarian and a Bartender on the central node
-%         - Simple distribution: split the central node to three nodes, each of them has an A-Hash, a Librarian and a Bartender, and we have separate storage nodes
-%         - One service per node: we have seperate nodes for all the services, any number of A-Hash nodes, any number of Librarian nodes, Bartender nodes, storage nodes
-% 
-%     User's manual
-%     - Overview: what do you need to access a running ARC storage system
-%     - Prototype command line tool
-%         - Configuring the command line tool
-%         - Creating the root collection
-%         - Uploading files
-%         - Creating collections
-%         - Moving files
-%         - Creating hardlinks
-%         - Removing files
-%         - Removing collections
-%         - Close collections
-%         - Modify access policies
-%         - Working with mount points
-%     - Grid jobs
-%         - Configuring the ARC storage DMC
-%         - Specifying files in job descriptions
-%     - Accessing the system with FUSE
-%         - Configuring the FUSE module
-%         - Mounting the storage system
-%         - Supported operations
-%         - Limitations
-%     
-% \end{verbatim}
 
@@
 
-Let's check the README file for all the dependencies, and install them:
+Let's check the README file for all the dependencies, and install them!
+We use `aptitude' here, but of course on other linux distribution you may use a different package management software, and the name of the packages may differ as well.
 
@@
 
-The storage system runs withen the \verb!arched!
+The storage system runs within the \verb!arched!
 hosting environment daemon, which needs a configuration file which describes which services we want to run, and on which ports do we want to listen, etc.
@@
 
+This version of the \verb!arc_storage_cli! tool currently does not support recursive methods or working with multiple files at once.
+
 \section{Adding more storage elements}
 
-This version of the \verb!arc_storage_cli! tool currently does not support recursive methods or working with multiple files at once.
-
 Let's add a new storage element to the system.
+A new storage element should be on a different physical machine, preferably in a different building to minimize the possibility of both storage elements being offline at the same time.
+But for testing purposes it is of course possible to run it on the same machine using different ports.
 We need to deploy a Shepherd service and a storage element service, which this time will be a `byteio' service.
@@
 
-Let's install Berkeley DB as root (but you can use the \verb!--prefix! option of \verb!configure! to install it somewhere else than the default location).
+Let's install Berkeley DB as root (but you can use the \verb!--prefix! option of \verb!configure! to install it somewhere else than the default location). This should be done on all the machines:
 
@@
 
-Let's configure the our first server (\verb!storage!). Here is the AHash part of the config:
+Let's configure our first server (\verb!storage!). Here is the AHash part of the config:
 
@@
 
-In this config we specify the second server as a peer. Let's configure the second server as well:
+In this config we specify the second server as a peer. Let's configure the second server (\verb!storage2!) as well:
 
@@
 
-In this config we specify the first server as a peer. So the two together will be able to form a replicated network with two nodes. But let's configure the thirs one:
+In this config we specify the first server as a peer. So the two together will be able to form a replicated network with two nodes. But let's configure the third one (\verb!storage3!):
 
@@
 
-It is important that we need to make a fresh start, and remove the existing database, which was created by a different store:
+It is important that we need to make a fresh start on all the servers, and remove the existing database, which was created by a different store:
 \begin{verbatim}
@@
 If we have permissions to access some third-party storage solutions (e.g.
-dCache), we can create a mount point within the namespace of the ARC storage system which points to a third-party URL, and then we can use the client interface of the ARC storage to access the third-party storage as well.
+dCache via the GridFTP protocol), we can create a mount point within the namespace of the ARC storage system which points to a third-party URL, and then we can use the client interface of the ARC storage to access the third-party storage as well.
 This is achieved by a module of the Bartender, called the `Gateway'.
@@
    catfish
-   $ cp fish fisk2
+   $ cp fish fish2
    $ ls -la

### 2009-04-23 doc/storage_system/arc-storage-documentation.tex
implementing requested changes in internal review comments

@@
 If a storage node (more precisely: the Shepherd service on the storage node) detects that a file is invalid, it reports this so the metadata will be in sync with the real state.
-And the storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat also triggers the modification of metadata.
+The storage nodes sends heartbeat messages periodically, and if a storage node goes offline, the missing heartbeat also triggers the modification of metadata.
 
@@
     \item The Bartender returns the TURL to the client tool.
-    \item The client tool use this TURL to get the file directly from the storage element service on the storage node.
+    \item The client tool use this TURL to get the file directly from the storage element service (SES) on the storage node.
 \end{enumerate}
@@
     \item The Bartender randomly chooses a Shepherd service and sends it a `put' request to initiate the file upload. 
-    \item The Shepherd communicates with the storage element service on the same node to create a new transfer URL (TURL).
+    \item The Shepherd communicates with the storage element service (SES) on the same node to create a new transfer URL (TURL).
 Then it creates a `referenceID' for this file and reports to the Librarian that there is a new replica in the \verb#creating# state.
@@
     \item The Bartender returns the TURL to the client tool.
-    \item Then client tool can upload the file to this TURL.
+    \item Then client tool can upload the file to this TURL (to the storage element service, SES)
     \item The Shepherd detects that the file has been uploaded.
@@
 
-The HED has a security framework which the Bartenders use the make access policy decisions. The details of this is described in Section \ref{sec:security_implementation}.
+The HED has a security framework which the Bartenders use to make access policy decisions. The details of this is described in Section \ref{sec:security_implementation}.
 
@@
 A location contains the URL of the Shepherd service, the local ID of the file within the storage node and the state of the replica.
-This state is `\textbf{alive}' if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy, `\textbf{invalid}' if the replica has wrong checksum, or the Shepherd claims it has no such file, `\textbf{offline}' if the Shepherd is not reachable, but may have a valid replica, `\textbf{creating}' if the replica is in the state of uploading, `\textbf{thirdwheel}' if the replica is marked for deletion because of too many replicas, or `\textbf{stalled}' if the the replica was in the state if uploading for a long time but hasn't arrived.
+This state is `\textbf{alive}' if the replica passed the checksum test, and the Shepherd reports that the storage node is healthy, `\textbf{invalid}' if the replica has wrong checksum, or the Shepherd claims it has no such file, `\textbf{offline}' if the Shepherd is not reachable, but may have a valid replica, `\textbf{creating}' if the replica is in the state of uploading, `\textbf{thirdwheel}' if the replica is marked for deletion because of too many replicas, or `\textbf{stalled}' if the the replica was in the state of uploading for a long time but hasn't arrived.
 The \emph{states} section contains the size of the file, the number of required replicas, and a checksum of the file, which is created by some kind of checksumming algorithm (currently only md5 is supported).
@@
 
-A \emph{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-hierarchy.
+A \emph{collection} is a list of files and other collections, which are in parent-children relationships forming a tree-like hierarchy (with possible interconnection due to hardlinks).
 Each entry has a unique name within a collection and a GUID which points to a corresponding file or collection, so a collection is basically a list of name-GUID pairs, which list is stored in the \emph{entries} section (see below).
@@
     \item[AHashClass] tells the A-Hash service the name of the business-logic A-Hash class.
-Currently there are two implementation, the \verb!CentralAHash!
+Currently there are two implementations, the \verb!CentralAHash!
 is a centralized version, and the \verb!ReplicatedAHash!
@@
 
-If a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file!), the Shepherd service creates a hardlink into the transfer directory (\verb!./transfer/abc!) and sets this file read-only.
+If a client asks for a file called \verb!file1!, and this file is in the store directory (\verb!./store/file1!), the Shepherd service creates a hardlink into the transfer directory (\verb!./transfer/abc!) and sets this file read-only.
 If the Hopi service is configured to serve the HTTP path \verb!/prb!
@@
 
-The Bartender utilizes gateway modules to access file listings and transfer URL's from third-party storage systems.
-It works as an optional backend component of the Bartender service.
-The \emph{Gateway} module is based on four basic methods: \emph{get}, \emph{put}, \emph{list} and \emph{remove}.
-This module should connect to the third-party storages on behalf of the user, so the user has to delegate his or her credentials to the Bartender using proxy delegation.
-In order to access the files and directories on the external storage system, limited user credentials are passed to the Bartender service.
-Users can also remove their credentials from the service side.
+The Bartender utilizes \emph{Gateway} modules to access file listings and transfer URLs from third-party storage systems.
+The gateway module has four basic methods: \emph{get}, \emph{put}, \emph{list} and \emph{remove}.
+This module connects to the third-party storages on behalf of the user, in order to do this, the Bartender needs to have proxy certificates from the user.
+A proxy certificate can be passed to the Bartender through the process of credential delegation.
+Users can also remove their proxy certificate from the service side.
 % subsection implementation (end)
@@
 
-The credentials of the user and the URL of the Bartender must be configured either with an XML file located at \verb!~/.arc/client.xml! or with environment variables.
+The credentials of the user and the URL of the Bartender must be configured either with an XML file located at \verb!~/.arc/client.xml!
+or with environment variables.
+(The environment variables overwrite the values from the XML file.)
 

### 2009-04-23 doc/client/UnicoreSupport.tex
Update

@@
 The next generation ARC client library will offer support for discovering grid resource running Unicore middleware and to submit and manage grid jobs (basic functionality) at these resources.
-This writeup documents the present status of the Unicore support, what is missing and what is the work plan to implement the missing functionality.
+This write-up documents the present status of the Unicore support, what is missing and what is the work plan to implement the missing functionality.
 \subsection*{Versions of Unicore supported}
-Text here about the supported versions
+Unicore versions of 6.2 (release, not candiate) and later are supported.
+Earlier versions are not supported due to incomplete information about the computing resources.
+Some functionality will still be present for earlier versions but not intentionally.
 \subsection*{Status of ongoing work}
 \subsubsection*{Resource discovery}
-Resource discovery. Can we find all resources? Are there defaults in the config file? Can we get detailed information from the clusters?
+The Unicore resource discovery starts by asking Unicore index servers for BES interfaces to Unicore execution services.
+Thereby all resources with a BES interfaces that register with that server are discovered.
+Presently it is not possible to get detailed information from the cluster, but this will be implemented using the Glue2 enabled information providers in the recent versions of Unicore (6.2 and above).
+
+The new ARC client doesn't come with any default Unicore index servers.
+Thus, in order to make use of the new ARC client's Unicore interoperability it is required that the user provide the URL to an Unicore index server or Unicore BES interface.
 \subsubsection*{Job submission and management}
-Can we submit jobs? Can we get jobs? Can we kill jobs?
+The following ARC client commands will be interoperable with Unicore: 
+\begin{itemize}
+\item{\texttt{arcsub, arcstat, arcinfo, arcget, arcclean, arckill, arcresub(sandbox), arcsync and arctest.}} 
+\end{itemize}
+
+These ARC client commands are either not implementable (according to present knowledge of Unicore) or not relevant for Unicore:
+\begin{itemize}
+\item{\texttt{arccat, arcmigrate, arcrenew and arcresume}}
+\end{itemize}
+
+Presently it is possible to use the \texttt{arcsub, arcstat} and \texttt{arcinfo} commands, though with some limited functionality (it is not possible to upload input files and the returned job/cluster job information is sparse).
 
 \subsection*{Work to be done}
-I guess this section is is to write once the present status is documented. Any depencies? i.e. work that has to be done be others?
+In light of the present status the following to do's can be identified:
 
-\subsection*{Work Plan}
-Assuming that the Unicore work can start fulltime again first week of May.
-How much time would be needed in order to finalize the work listed above in the work to be done section.
-What should be done first etc.
-
+\begin{itemize}
+\item{Retrieve detailed cluster information using the new Glue2 information providers and populate the ARC internal ExecutionTarget with this information.
+Input/fixes from the Unicore team may be needed.}
+\item{Retrieve detailed job information (similarly to the cluster information).}
+\item{Enable job input files and get the remaining interoperable commands to work. The latter will in many cases only require small fixes to already existing code.}
+\end{itemize}
+\subsection*{Time estimate}
+It is estimated that the remaining work for Unicore interoperability is of the order of 1 PM.
+Thus if Unicore work can resume again early May the work should be completed early June if everything progresses according to plan.
 \end{document}

### 2009-04-23 doc/storage_system/arc-storage-manual.tex
De-we-fication

@@
-\documentclass{book}
+\documentclass{article}
+%\documentclass{book}
 \usepackage{graphicx}                              %for PNG images (pdflatex)
@@
 %%%%---- The title ----
-{\centering \textsc{\Large Administrator's manual of the ARC storage system}\Large \par}
+{\centering \textsc{\Large Chelonia Administrator's manual}\Large \par}
 \vspace*{0.5cm}
@@
 
-\chapter{Quick start guide} % (fold)
+%\chapter{Quick start guide} % (fold)
 
@@
 
-In this section we will follow the installation of the ARC storage system on a freshly installed debian lenny machine.
+This section shows how to install Chelonia on a freshly installed debian lenny machine.
 
@@
 
-Let's check the README file for all the dependencies, and install them!
-We use `aptitude' here, but of course on other linux distribution you may use a different package management software, and the name of the packages may differ as well.
+Check the README file for all the dependencies, and install them.
+Here `aptitude' is used, but of course on other Linux distribution different package management software may be used, and the name of the packages may differ as well.
 
@@
 
-Run the \verb!autogen! and \verb!configure! scripts (you may want to specify target directories for the installation with the \verb!--prefix! option):
+Now, run the \verb!autogen! and \verb!configure! scripts (target directories for the installation may be specified with the \verb!--prefix! option):
 
@@
 
-We can start to compile it:
+Start with compiling it:
 
@@
 
-Maybe we should run \verb!ldconfig! to refresh the list of libraries:
+It may be a good idea to run \verb!ldconfig! to refresh the list of libraries:
 
@@
 
-We can check if the python packages are installed correctly:
+Check if the python packages are installed correctly:
 
@@
 
-If there is no error on importing any of the packages, we are good to go. If the packages cannot be imported, we should set the PYTHONPATH environment variable to point to their location.
-
-It is important to syncronize the time of the machine - the storage system will make decisions based on differences between timestamps, when we deploy it on multiple machines, different times would cause problems.
-
-We need host (and user) certificates to run the system.
-For testing purposes we can use the NorduGrid InstantCA solution which is capable of creating a demo-CA with short lifetime.
-The URL of the InstantCA service is \verb!https://vls.grid.upjs.sk/CA/instantCA!, let's create 3 user and 3 host certificates.
-(e.g.
-we can set the Organization Name to `knowarc', the Common name of the CA to `storage-test', the name of the users could be `penny', `billy' and `hammer', the name of the hosts could be `upsilon', `omicron' and `theta', and we need a password for the CA and passwords for the user certificates, both should be at least 4 characters long).
-
-Let's download the generated certificates and untar the archive file to see its contents:
+If there is no error on importing any of the packages, everything is OK. If the packages cannot be imported, try to set the PYTHONPATH environment variable to point to location of the packages.
+
+It is important to syncronize the time of the machine - the storage system will make decisions based on differences between timestamps.
+When deploying multiple machines, different times will cause problems.
+
+Host (and user) certificates are required to run the system.
+For testing purposes the NorduGrid InstantCA solution can be used.
+The InstantCA is capable of creating a demo-CA with short lifetime.
+The URL of the InstantCA service is \verb!https://vls.grid.upjs.sk/CA/instantCA!.
+Now, create 3 user and 3 host certificates.
+(for this example, set the Organization Name to `knowarc', the Common name of the CA to `storage-test', the name of the users to `penny', `billy' and `hammer' and the name of the hosts to `upsilon', `omicron' and `theta'.
+A password for the CA and passwords for the user certificates are needed as well, these should be at least 4 characters long).
+
+Now, download the generated certificates and untar the archive file to see its contents:
 
@@
 
-We have certificate and key files for all the hosts and users, and we have the CA file with the proper hashed name.
-There is no rule where to put these certificate, but now we will put them in \verb!/etc/grid-security!.
-Let's create this directory, and a subdirectory called \verb!certificates!.
-We will use the `theta' host certificates for this machine, let's copy it there:
+You can see the certificate and key files for all the hosts and users, and the CA file with the proper hashed name.
+There is no rule about where to put these certificate, but it is common to put them in \verb!/etc/grid-security!.
+Now, create this directory, and a subdirectory called \verb!certificates!.
+To use the `theta' host certificates for this machine, copy the corresponding certificate and key there:
 
@@
 
-And I choose a user certificate and put it into the \verb!~/.arc! directory, conveniently removing the password from the key file:
+Choose a user certificate and put it into the \verb!~/.arc! directory, removing the password from the key file for convenience:
 
@@
 
-The storage system runs within the \verb!arched!
-hosting environment daemon, which needs a configuration file which describes which services we want to run, and on which ports do we want to listen, etc.
-Let's copy the template configuration files to the \verb!/etc/arc! directory:
+The storage system runs within the \verb!arched! hosting environment daemon, which needs a configuration file describing which services to be run, on which ports do to listen, etc.
+Copy the template configuration files to the \verb!/etc/arc! directory:
 
@@
 
-The template configuration file specifies several directories where the services can store their data. We will create these directories at \verb!/var/spool/arc!, but they can be created anywhere else.
+The template configuration file specifies several directories where the services can store their data. Create these directories at, e.g., \verb!/var/spool/arc!.
 
@@
 
-In this deployment we plan to run the daemon as root, so it is OK that the root owns these directories. 
-If we create these directories somewhere else, we would need to modify the paths in the \verb!storage_service.xml! file.
-
-If we put the host certificates and the CA certificate somewhere else than \verb!/etc/grid-security!
-then we would need to modify the \verb!storage_service.xml!
+In this deployment the daemon will be run as root, so it is OK that the root owns these directories. 
+If creating these directories somewhere else, then modify the paths in the \verb!storage_service.xml! file.
+
+If putting the host certificates and the CA certificate somewhere else than \verb!/etc/grid-security!
+then modify the \verb!storage_service.xml!
 (look for \verb!KeyPath!, \verb!CertificatePath!
-and \verb!CACertificatesDir!), and we would need to modify \verb!/etc/arc/clientsslconfig.xml!
-as well.
-This file contains the credentials which is used by the services when they connect to an other service.
-If we put \verb!clientsslconfig.xml!
-somewhere else than \verb!/etc/arc!
-then we would need to modify all the \verb!ClientSSLConfig!
+and \verb!CACertificatesDir!).
+Also, modify \verb!/etc/arc/clientsslconfig.xml!.
+This file contains the credentials which are used by the services when they connect to an other service.
+If putting \verb!clientsslconfig.xml!
+somewhere else than \verb!/etc/arc!, you need to modify all the \verb!ClientSSLConfig!
 tags in the \verb!storage_service.xml!.
 
-Let's run the \verb!arched! daemon with this config (specified by the \verb!-c! option) first in the foreground (using the \verb!-f! option) to see immediately if everything is right:
+Now, run the \verb!arched! daemon with this config (specified by the \verb!-c! option) first in the foreground (using the \verb!-f! option) to see immediately if everything is right:
 
@@
 The template configuration file specifies a loglevel which only prints error messages, so nothing should be written on the screen now.
-The \verb!arched! daemon should now listen on the ports: 60001 for our data transfer service called Hopi, and 60000 for all the other services:
+The \verb!arched! daemon should now listen on the ports: 60001 for the data transfer service called Hopi, and 60000 for all the other services:
 
@@
 
-Now it is time to set up or prototype client tool to access the system.
+Now it is time to set up the prototype client tool to access the system.
 It is currently called \verb!arc_storage_cli!
 and it is by default installed to \verb!/usr/local/bin!.
-We have to tell this CLI tool where it can find our user credentials, and a URL of a Bartender service.
+The CLI tool needs to be told where it can find the user credentials, and a URL of a Bartender service.
 The Bartender service is the front-end of the storage system.
-We can specify these things by environment variables, or by a configuration file called \verb!~/.arc/client.xml!.
-Let's create this \verb!client.xml!
+This can be specified either by environment variables, or by a configuration file called \verb!~/.arc/client.xml!.
+Create the following \verb!client.xml!
 file:
@@
 
-Now we can use the \verb!arc_storage_cli! to access our local deployment. Let's do a `list' on the root collection (`/'):
+Now you can use the \verb!arc_storage_cli! to access your local deployment. Try to do a `list' on the root collection (`/'):
 
@@
 
-If you get this error, where the `Component [...] could not be created', we should reinforce the python environment about the location of our python packages:
+If you get the error `Component [...] could not be created', try to specify to the python environment the location of the python packages:
 
@@
 prints the list of available methods.
-Specifying only a method name prints the syntex of the given method.
+Specifying only a method name prints the syntax of the given method.
 The \verb!list!
@@
 
-Now we have a running (centralized) storage service including a Hopi service, which is a basic HTTP server, for transfering files.
-We have a Bartender service which is the front-end of the system to the users, so our client always connect to the Bartender service.
+Now a (centralized) storage service is running, including a Hopi service, which is a basic HTTP server, for transferring files.
+The Bartender service is the front-end of the system to the users, so the clients always connect to the Bartender service.
 There are several directories configured in the config file where the services store their data.
 These directories can always be emptied, which resets the whole system to a clean state.
-If we run the storage system with clean data directories, then it is completely empty, which means that there is not even a root collection.
+If Chelonia is running with clean data directories, then it is completely empty, meaning that there is not even a root collection:
 
@@
 
-The root collection has the Logical Name of `\verb!/!', and it's role to provide a starting point for all the other Logical Names. We have to create this root collection first:
+The root collection has the Logical Name of `\verb!/!', and it's role to provide a starting point for all the other Logical Names. This root collection needs to be created first:
 
@@
 
-Now we do have an empty root collection.
+Now there is an empty root collection.
 
@@
 
-Assume that we have a running storage system and we want to access it with the \verb!arc_storage_cli!
-client.
-We have the \verb!~/.arc/client.xml!
-set properly, so we can list the contents of the root collection:
+Assume that Chelonia is running and you want to access it with the \verb!arc_storage_cli! client. The \verb!~/.arc/client.xml! is set properly, the contents of the root collection can be listed:
 
@@
 
-Let's create a small file and upload it:
+Now, create a small file and upload it:
 
@@
 
-We can get additional information about the file with the stat method:
+Additional information about the file can be found with the stat method:
 
@@
 
-Here we can see the owner of the file (Billy), the location of the replicas (currently only one), the number of needed replicas (which is also: one), the checksum and size of the file, and the timestamp of the creation.
-
-We can download this file:
+Here, the owner of the file (Billy), the location of the replicas (currently only one), the number of needed replicas (which is also only one), the checksum and size of the file, and the timestamp of the creation is shown.
+
+This file can be downloaded:
 
@@
 
-We can create collections which can contain files and other collections forming a tree-hierarchy. We can move files and collection around within this namespace.
+You can create collections which can contain files and other collections forming a tree-hierarchy. You can move files and collection around within this namespace.
 
@@
 
-We can remove files, and remove collections (but only if the collection is empty):
+You can remove files, and remove collections (but only if the collection is empty):
 
@@
 
-Let's introduce a new user into the system.
-If we have some other user certificate and key files, then we can tell \verb!arc_storage_cli!
-to use them with modifying the \verb!client.xml!
+Now, introduce a new user, Penny, into the system.
+Again, \verb!arc_storage_cli!
+can be told to use user certificate and key files for this user with modifying the \verb!client.xml!
 or by specifying the \verb!ARC_CERT_FILE!
@@
 Listing the root collection or uploading files into it for Penny is denied, because the root collection is owned by Billy, and nobody else has permissions to it.
-Let's ask Billy to allow for everyone to access the root collection:
+Now, Billy allows for everyone to access the root collection:
 
@@
 
-So we have changed the action list for user `ALL' to `+read +addEntry', which means that everybody can list the contents of this collection, create subcollection in it, upload a file into it, or move some existing entry into it.
-
-Now we can upload a file with Penny:
+So Billy has changed the action list for user `ALL' to `+read +addEntry', which means that everybody can list the contents of this collection, create subcollections in it, upload files into it, or move existing entries into it.
+
+Now, Penny can upload a file:
 
@@
 
-This file now only can be downloaded by Penny, so let's allow for Billy to download it as well:
+This file now only can be downloaded by Penny, so she should allow for Billy to download it as well:
 
@@
 
-We can see among the metadata of the file, that Billy has `read' rights. Billy can now download this file:
+The metadata of the file shows that Billy has `read' rights. Billy can now download this file:
 
@@
 
-We can specify the needed number of replicas for a new file with the \verb!ARC_NEEDED_REPLICAS! environment variable.
+The needed number of replicas for a new file can be specified with the \verb!ARC_NEEDED_REPLICAS! environment variable.
 
@@
 
-We can see that the `neededReplica' metadata says 2, but if we deployed the system in a centralized way for testing purposes, then we only have one storage element (which is a Shepherd service and a Hopi service together), so the file can only have one replica, that's why there is only one locations.
-
-We can modify the number of needed replicas of an existing file with the `modify' method:
+Here, the `neededReplica' metadata says 2, but the system is deployed in a centralized way for testing purposes, so there is only one storage element (which is a Shepherd service and a Hopi service together), so the file can only have one replica and there can only be one location.
+
+The number of needed replicas of an existing file can be modified with the `modify' method:
 
@@
 
-This version of the \verb!arc_storage_cli! tool currently does not support recursive methods or working with multiple files at once.
+It should be noted that the current version of the \verb!arc_storage_cli! tool does not support recursive methods or working with multiple files at once.
 
@@
 
-Let's add a new storage element to the system.
+The following shows how to add a new storage element to the system.
 A new storage element should be on a different physical machine, preferably in a different building to minimize the possibility of both storage elements being offline at the same time.
 But for testing purposes it is of course possible to run it on the same machine using different ports.
-We need to deploy a Shepherd service and a storage element service, which this time will be a `byteio' service.
+A storage element consists of a Shepherd service and a storage element service, which this time will be a `byteio' service.
 The `byteio' service implements a small part of the ByteIO interface, which means that it just put the entire file data into the SOAP message, so it is only suitable for small files.
 
-But first modify the configuration file of our original server.
-Currently all the transfer URLs which were generated when we upload and download files contained the hostname `localhost'.
-This is not OK if we want to access it from an other machine, we should provide transfer URLs which contains the external hostname of the machine.
-So lets change this line in \verb!/etc/arc/storage_service.xml!:
+Now, modify the configuration file of the original server.
+Currently all the transfer URLs which were generated when uploading and downloading files contained the hostname `localhost'.
+This is not OK if access from an other machine is needed.
+In this case, transfer URL's must contain the external hostname of the machine.
+To do this, change this line in \verb!/etc/arc/storage_service.xml!:
 
@@
 But this is not enough.
-Currently the Shepherd services uses their URLs as IDs, and the other services use these IDs to connect to a Shepherd service, so we need to change to ID of the Shepherd service to a real external URL as well.
-And for the sake of clarity let's change all `localhost' entries in the config file to our real hostname.
-In our case the name of this machine is just `storage' so the Shepherd section should look like this:
+Currently the Shepherd services uses their URLs as IDs, and the other services use these IDs to connect to a Shepherd service, so the ID of the Shepherd service needs to be changed to a real external URL as well.
+For the sake of clarity, change all `localhost' entries in the config file to the real hostname.
+In this example, the name of this machine is just `storage' so the Shepherd section should look like this:
 
@@
 
-Let's check first if our centralized deployment is still working, stop the \verb!arched! daemon remove the data directories to make it a clean start, and restart it with the modified config.
+Now, check if the centralized deployment is still working, stop the \verb!arched! daemon, remove the data directories to make it a clean start, and restart it with the modified config.
 
@@
 
-Let's see if we can still use the client:
+See if the client is still usable:
 
@@
 
-Let's upload a file and ask for two replicas:
+Now, upload a file and ask for two replicas:
 
@@
 
-Currently it has only one replica, because we still don't have more than one storage elements.
-While \verb!arched!
-keeps running, let's create a new config for the other machine, which is called \verb!storage2!:
+Currently it has only one replica, because there is still only one storage element. Now, create a new config for the other machine, here called \verb!storage2!:
 
@@
 The Shepherd service needs to have a backend for all supported storage element services, so it has a ByteIOBackend configured.
-This Shepherd connects to our original server's Bartender and Librarian services.
+This Shepherd connects to the first server's Bartender and Librarian services.
 (See the technical documentation for details about the services.)
 
-While the original (\verb!storage!) server still runs the services, let's start this new server:
+While the original (\verb!storage!) server still runs the services, start the new server:
 
@@
 
-And after a minute or two, let's check our file:
+After a minute or two, check the file again:
 
@@
 
-Let's download it multiple times:
+Now, try to download it multiple times:
 
@@
 
-Let's create a third storage element on a machine called \verb!storage3! this time using a Hopi service as storage element service like on our main \verb!storage! server:
+Now, create a third storage element on a machine called \verb!storage3! this time using a Hopi service as storage element service like on the main \verb!storage! server:
 
@@
 
-This configuration has two separate ports, one with TLS security and a Shepherd service, and the other without TLS security and with the Hopi service. Let's run this server as well:
+This configuration has two separate ports, one with TLS security and a Shepherd service, and the other without TLS security and with the Hopi service. Run this server as well:
 
@@
 
-Now when checking our file again there should be no change at all:
+Now when checking the file again there should be no change at all:
 
@@
 
-Let's stop our server on \verb!storage2!, and watch what happens of our file:
+Stop the server on \verb!storage2!, and watch what happens with the file:
 
@@
 
-A new replica have been created on the third server to maintain the needed number of 2.
-
-Let's start the services on \verb!storage2! again, and let's see what is happening with our file:
+A new replica has been created on the third server to maintain the needed number of 2.
+
+Now, start the services on \verb!storage2! again, and see what is happening with the file:
 
@@
 
-One of the replicas have been disappeared to maintain the needed number of 2.
+One of the replicas has been removed to maintain the needed number of 2.
 
@@
 
-Now we have a deployement, where one machine has the A-Hash, Librarian and Bartender services, and all three machines have a Shepherd and a storage element service.
+Now the deployment is as following: One machine has the A-Hash, Librarian and Bartender services, and all three machines have a Shepherd and a storage element service.
 The A-Hash service stores all the metadata of all the files and collections, so it is critical.
 If the first machine goes offline then the whole system dies.
-We can deploy the A-Hash service on all the machines and still can have one consistent metadata database.
+However, it is possible to deploy the A-Hash service on all the machines and still have one consistent metadata database.
 The replicated version of the A-Hash uses the Berkeley DB to provide a distributed database.
 
-Let's install Berkeley DB as root (but you can use the \verb!--prefix! option of \verb!configure! to install it somewhere else than the default location). This should be done on all the machines:
+In the following, Berkeley DB is installed as root (the \verb!--prefix!
+option of \verb!configure!
+can be used to install it somewhere else than the default location).
+This needs to be done on all the machines.
 
@@
 
-Configure, make and install it (you can use the \verb!--prefix! option for \verb!configure!)
+Configure, make and install it (and if needed, use the \verb!--prefix! option for \verb!configure!)
 \begin{verbatim}
@@
 
-Now we need the python wrapper for the Berkeley DB:
+Next, get the python wrapper for the Berkeley DB:
 
@@
 
-Let's install it to the default location:
+Install it to the default location:
 
@@
 
-If you've installed Berkeley DB somewhere else, or want to install the python wrapper to somewhere else than the default:
+If Berkeley DB was installed in a non-default location, or to install the python wrapper to somewhere else than the default:
 
@@
 
-Let's test the installation with the \verb!test.py! included in the directory of the python wrapper for the Berkeley DB:
+Test the installation with the \verb!test.py! included in the directory of the python wrapper for the Berkeley DB:
 \begin{verbatim}
@@
 
-Let's configure our first server (\verb!storage!). Here is the AHash part of the config:
+Configure the first server (\verb!storage!). Here is the AHash part of the config:
 
@@
 
-In this config we specify the second server as a peer. Let's configure the second server (\verb!storage2!) as well:
+In this config the second server is specified as a peer. Configure the second server (\verb!storage2!) as well:
 
@@
 
-In this config we specify the first server as a peer. So the two together will be able to form a replicated network with two nodes. But let's configure the third one (\verb!storage3!):
+In this config the first server is configured as a peer. So the two together will be able to form a replicated network with two nodes. Then configure the third one (\verb!storage3!):
 
@@
 
-Here we specify the first server as the peer.
-
-It is important that we need to make a fresh start on all the servers, and remove the existing database, which was created by a different store:
+Here the first server is specified as the peer.
+
+It is important to make a fresh start on all the servers, and remove the existing database, which was created by a different store:
 \begin{verbatim}
@@
 
-Let's start the first server. It will complain that it cannot find its peer:
+Now, start the first server. It will complain that it cannot find its peer:
 \begin{verbatim}
@@
 
-With only one node of the replicated A-Hash the system is not ready to work, if we try to use it, we get an error message:
+With only one node of the replicated A-Hash the system is not ready to work, and trying to use it will result in an error message:
 \begin{verbatim}
@@
 
-Let's start the second server.
-It may first print some error messages, but then both servers should be quiet, and now we can use the system.
-Let's upload a file, and set the number of needed replicas to 3:
+Start the second server. It may first print some error messages, but then both servers should be quiet, and the system is ready for use. Now, upload a file, and set the number of needed replicas to 3:
 
@@
 
-It cannot create 3 replicas, because are third server is not running. Let's start it. After a minute or two, or file has three replicas:
+It cannot create 3 replicas, because the third server is not running. Start it. After a minute or two, the file has three replicas:
 
@@
 
-The fact that this file has three replicas is not at all related to the number of AHashes, but to the number of storage elements (Shepherds and storage element services), so this is no different then what we saw in the last section.
+The fact that this file has three replicas is not at all related to the number of AHashes, but to the number of storage elements (Shepherds and storage element services), so this is no different then what was seen in the previous section.
 The benefit of having more then one AHashes in a replicated way is that if one of the machines goes offline, the AHash is still working.
-But in our current case if we shutdown the first server, while the AHash will still work but we still have only one instance of the other two important service: the Librarian and the Bartender, so the whole system would be unusable.
-Let's add more Librarians and Bartenders to the system.
+However, with the setup above, if the first server is shutdown, the AHash will still work but there is only one instance of the other two important service: the Librarian and the Bartender, so the whole system will still be unusable.
+Hence, more Librarians and Bartenders needs to be added to the system.
+The next section explains how.
 
@@
 
-Let's add a Bartender and a Librarian to each servers.
-We don't need to modify our first server now.
-For the other two we need to change the \verb!Plexer!
-section, add the twos services, and modify the Shepherd to use the server's own Bartender and Librarian for replication and heartbeats.
-Let's modify the config of our second server: (\verb![...]!
+In this section, a Bartender and a Librarian is added to each server.
+The first server does not need modification now.
+For the other two, the \verb!Plexer!
+section needs to be altered, the twos services must be added, and the Shepherd must be modified to use the server's own Bartender and Librarian for replication and heartbeats.
+Modify the config of the second server as follows: (\verb![...]!
 indicates unchanged parts)
@@
 
-We need to modify our \verb!client.xml! to add all the other Bartenders:
+\verb!client.xml! must be modified to add all the other Bartenders:
 
@@
 
-Now we can start all three servers.
+Now, all three servers can be started.
 The replicated A-Hash sometimes prints great amount of error messages, but if it stops printing them, then everything is OK.
-Now we can use the system as intended.
+Now the system can be used as intended.
 There is a flag for the \verb!arc_storage_cli!
 which makes it very verbose.
-We can check the used Bartender with it:
+You can check the used Bartender with it:
 
@@
 
-Now we can stop our first server. The system will be unaccessable for a few seconds while the other two A-Hash figure out what is happening, but after that it will work again:
+Now, stop the first server. The system will be unaccessible for a few seconds while the other two A-Hash figure out what is happening, but after that it will work again:
 
@@
 
-If we start again our first server, after a few seconds the system will work again with the three fully-functional node.
+If the first server is started again, after a few seconds the system will work again with the three fully-functional node.
 
@@
 
-If we have permissions to access some third-party storage solutions (e.g.
-dCache via the GridFTP protocol), we can create a mount point within the namespace of the ARC storage system which points to a third-party URL, and then we can use the client interface of the ARC storage to access the third-party storage as well.
+If you have permissions to access some third-party storage solutions (e.g.
+dCache via the GridFTP protocol), you can create a mount point within the namespace of the ARC storage system which points to a third-party URL, and then you can use the client interface of the ARC storage to access the third-party storage as well.
 This is achieved by a module of the Bartender, called the `Gateway'.
-Currently only GridFTP access is supported, and for that we need to install the Globus libraries to our machines.
-Then we need to recompile our source for GridFTP support.
+Currently only GridFTP access is supported, and for that the Globus libraries needs to be installed.
+Next, ARC needs to be recompiled for GridFTP support.
 The output of the \verb!configure!
@@
 
-After compilation and installation we can use the \verb!arcls! tool to list the content of a third-party storage. Let's check if our Billy user has access to this directory:
+After compilation and installation the \verb!arcls! tool can be used to list the content of a third-party storage. The user Billy can see if has access to this directory:
 
@@
 
-We can see some files in there, so we have permissions to list that directory.
-
-Now let's configure the Bartenders on all three machines to use the Gateway module:
+Since he can see some files in there, so he have permissions to list that directory.
+
+Now, configure the Bartenders on all three machines to use the Gateway module:
 
@@
 
-Let's start all three servers. Then create a mount point with the Logical Name \verb!/salmount! which points to this URL:
+Next, start all three servers. Then create a mount point with the Logical Name \verb!/salmount! which points to this URL:
 
@@
 
-Let's list the content of this mountpoint:
+List the content of this mountpoint:
 
@@
 
-Yes, we first need to delegate our credentials with the \verb!credentialsDelegation! method, then try to list the mount point:
+OK, First you need to delegate your credentials with the \verb!credentialsDelegation! method.
 
@@
 Proxy ID: eNCMDmi8i6YnPSAtDmVmuSEmABFKDmABFKDm0zKKDmFBFKDmeOzdBo
-
+\end{verbatim}
+
+Then try to list the mount point again:
+
+\begin{verbatim}
 $ arc_storage_cli list /salmount
@@
 
-Let's download a file:
+Now, download a file:
 
@@
 For all the supported types there is a DMC module.
-We created a DMC module for the ARC storage as well.
+For Chelonia, the ARC DMC module is used.
 If it is compiled and installed, then you can use the \verb!arcls!, \verb!arccp!, \verb!arcrm!, etc.
@@
 
-Besides the \verb!arc_storage_cli! tool we can access the storage system with the use of a FUSE module. We can install the ARC storage FUSE module following these steps:
+Besides the \verb!arc_storage_cli! tool you can access the storage system with the use of a FUSE module. The following steps installs the ARC storage FUSE module:
 

### 2009-04-23 doc/client/UnicoreSupport.tex
Replaced presently with a less ambiguous word in two places and some other minor edits.

@@
 \subsection*{Versions of Unicore supported}
-Unicore versions of 6.2 (release, not candiate) and later are supported.
+Unicore versions of 6.2 and later are supported.
 Earlier versions are not supported due to incomplete information about the computing resources.
@@
 Thereby all resources with a BES interfaces that register with that server are discovered.
-Presently it is not possible to get detailed information from the cluster, but this will be implemented using the Glue2 enabled information providers in the recent versions of Unicore (6.2 and above).
+Currently it is not possible to get detailed information from the cluster, but this will be implemented using the Glue2 enabled information providers in the recent versions of Unicore (6.2 and above).
 
@@
 
-Presently it is possible to use the \texttt{arcsub, arcstat} and \texttt{arcinfo} commands, though with some limited functionality (it is not possible to upload input files and the returned job/cluster job information is sparse).
+Currently it is possible to use the \texttt{arcsub, arcstat} and \texttt{arcinfo} commands, though with some limited functionality (it is not possible to upload input files and the returned job/cluster job information is sparse).
 

### 2009-04-26 doc/jura/jura-tech-doc.tex
Added to JURA doc

@@
 \usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
-\usepackage{color}                                 %for defining custom colors
+\usepackage{color}                                 %for defining custom colours
 \usepackage{framed}                                %for shaded and framed paragraphs
@@
 
-The \textit{Job Usage Reporter for ARC} (JURA) is a component
+The \textit{Job Usage Reporter of ARC} (JURA) is a component
 implementing accounting functionality in the ARC middleware. Its
@@
 logging service of the SweGrid Accounting System (SGAS)\cite{sgas},
-but maintaining the possibility to enable utilizing other services has
+but maintaining the possibility to enable utilising other services has
 been kept in mind during design.
@@
 \label{joblogs}
-The job logs generated by A-REX reside under the directory
-\verb|<control_dir>/logs|. They have file name format
-\verb|<ngjobid>.<random>|, where \textit{ngjobid} is the identifier
+The job log files generated by A-REX reside under the directory
+\textit{<control\_dir>/logs}. They have file name format
+\textit{<ngjobid>.<random>}, where \textit{ngjobid} is the identifier
 created for the job by A-REX, \textit{random} is a randomly generated
 sequence of alphanumeric characters to avoid collision of different
-files pertaining to the same job.
+files pertaining to the same job. For each job and for each reporting
+destination (i.e. accounting service to be contacted), at least two
+log files are written for each job: one at the time of job submission,
+and another one after the job finishes, and possibly others at each
+start and stop event.
 
-2 for each logging destination: when submitted, when finished
+A job log file consists of ``\textit{name=value}'' lines.  These
+make up some of the configuration options for JURA, such as the URL of
+an accounting destination in the ``\textit{loggerurl=}'' line.  The
+file also contains detailed resource usage data and related metadata
+to be reported about the job.(See Appendix \ref{config} for more
+details.)  The accuracy of the metered data may depend on the type of
+batch system used.
 
-UR generated from them: only from finished
+Based on the usage data in the processed job log file, JURA assembles
+records in the format proposed by the Open Grid Forum (OGF), called
+the \textit{Usage Record} (UR)\cite{ur}. This is an XML representation
+holding consumption information for all commonly used resources and
+metrics. It can be extended by custom elements for non-standard
+resources and/or other types of job metadata. For a list of UR
+properties used by JURA, see Appendix \ref{log2ur}.
+
+Some elements of UR are mandatory, these must all be present in the
+job log file to be able to generate a UR. For example, the job log
+file generated upon job submission contains no \textit{status} entry,
+so this file is ignored, and no UR is generated from it.
 
 \subsection{Accessing LUTS}
+After the necessary URs to be inserted are filled and valid, they are
+submitted to the accounting service through its insertion
+interface. The service in particular, SGAS LUTS has a simple custom
+web service interface loosely based on WS-ResourceProperties.
 
-insert method
-
-certificates from:...
+To increase communication efficiency, LUTS accepts a batch of several
+URs within a single request. The batch is an XML element called
+\textit{UsageRecords}, containing elements representing URs. The
+maximal number of URs in a batch is limited, it can be set in the
+``\textit{jobreport\_options=}'' line of the A-REX configuration file
+(see Appendix \ref{config}).
 
 \section{Security}
-  runs as: same user as a-rex, typically root
+The JURA executable runs as the same user as A-REX does, typically as
+\textit{root}. The owner of a job log file is the local user mapped
+for the submitter entity of the corresponding job. These files contain
+confidential data, access to which must be restricted, therefore read
+access is limited to the owner and the super user. If JURA is executed
+by A-REX, it can read data from these files, and delete expired files.
 
-  accesses sensitive data in job logs
+The authentication towards the SGAS LUTS is done via the standard
+X.509 certificate mechanism over SSL protocol: a chain of valid
+(i.e. not expired and/or revoked) certificates with a trusted root
+certification authority is accepted as authentic identification of the
+client. This means that the client can access the service using a
+proxy certificate as well. In the scenario involving A-REX and JURA,
+all usage records are submitted using credentials given in the
+\textit{jobreport\_credentials} line of the A-REX configuration file
+(see Appendix \ref{config}), and no proxies are used. Normally the
+credentials for the A-REX service should be used.
 
-  sgas security:
-
-    standard ssl cert, possibly proxy
-
-    no other attributes but DN considered
-
-    access control: flat-file based + XACML-like(?) configured thru
-    SAM porttype
-
-  in this case: service credentials, no proxy
+The access control policy for LUTS can be configured in two ways. One
+is a plain text configuration file where users, identified by their
+DNs, can be granted or denied publish and/or query rights for all
+URs. The other option is a service port type called \textit{Service
+  Authorization Management} (SAM). One can set access control rules
+through the SAM port of the service in a format based on XACML.
 
 \section{Implementation}
-  c++
+JURA is written merely in C++, and is tested in a GNU/Linux
+environment, built with standard GNU tools. It depends on the HED
+libraries, and thus also on all mandatory dependencies thereof. The
+\textit{openssl} library is also required, since sensitive accounting
+data needs to be sent in a secure manner.
 
-  sw dependencies
+%TODO about other service types
 
-    full arc
-
-    ssl
-
-  install location
+When deployed, the executable called ``\textit{jura}'' is placed into the
+\textit{libexec} directory of the ARC install location. There is a
+symbolic link %TODO make it so!
+to the executable in the same directory, called ``\textit{logger}'',
+through which A-REX executes JURA. This is to keep compatibility with
+the old logger client. No other executables or wrapper scripts are
+needed.
 
@@
 \label{config}
-  jobreport
 
-  jobreport\_options
+%   loggerurl: jobreport, JSDL
 
-  jobreport\_credentials
+%   (lifetime)
+
+%   jobreport
+
+%   accounting\_options: jobreport\_options
+
+%   key\_path, certificate\_path, ca\_certificates\_dir: jobreport\_credentials
 
@@
 \label{log2ur}
-  input file format
+%   input file format
 
-  filled properties
+%   filled properties
 
-  missing properties
+%   missing properties
 

### 2009-04-27 doc/internal_notes/UnicoreSupport/UnicoreSupport.tex
Starting cleaning up documentation

--- 2009-04-23 doc/client/UnicoreSupport.tex
@@
 
-Currently it is possible to use the \texttt{arcsub, arcstat} and \texttt{arcinfo} commands, though with some limited functionality (it is not possible to upload input files and the returned job/cluster job information is sparse).
+Currently it is possible to use the \texttt{arcsub, arcstat, arcinfo} and \texttt{arckill} commands, though with some limited functionality (it is not possible to upload input files and the returned job/cluster job information is sparse).
 

### 2009-04-27 doc/internal_notes/client/GUI_LAP_plans.tex
Starting cleaning up documentation

\documentclass[11pt,oneside,a4paper,english]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{graphicx,epic,eepic}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{a4wide}
\usepackage{xspace}

\newcommand{\arcmigrate}{\texttt{arcmigrate}\xspace}

\begin{document}
%\maketitle
\section*{The Lunarc Application Portal - A graphical grid user interface using ARC}
The Lunarc Application Portal (LAP) is an existing web interface for submission and management of grid jobs. As reflected in the name the LAP is an application oriented portal, providing easy to use interfaces for commonly available applications. The LAP is also a framework for implementing web interfaces for user applications.

A new and updated version of LAP is presently under development which will make the LAP to Web 2.0 compliant and which will interface to the next generation ARC client library. The latter interfacing is done in collaboration between LAP and the KnowARC project and will allow the KnowARC project to use the next generation LAP for it's deliverable D2.2-2 "Advanced graphical client with easy-to-use job creation module and support for bulk job management and monitoring".

\section*{Functionality to be offered though LAP}
Below functionality to be offered through the LAP and which will be a part of the KnowARC deliverable D2.2-2 is listed and discussed:

\begin{itemize}
\item{The portal will support bulk job handling, i.e. it will support listing and sorting of (several) hundred(s) of jobs using a java table. Using the same table bulk handling will be enabled making it possible to select groups of jobs to execute commands on like e.g. kill, clean, get, resume etc. Will it be possible to group jobs together?}
\item{Job statuses will be updated automatically by a background processes, but a "manual update" button will also be provided.}
\item{Job "babysitting" will offer to automatically download finished jobs or resubmit jobs that have failed. Resubmission will be limited to only 2(3?) attempts. It will be possible to activate "babysitting" at any time for any kind of job.}
\item{Users will be able to create and maintain lists of favorite clusters. Blacklisting of clusters will also be possible.}
\item{The LAP is application orieented and new grid jobs are created through an application specific interface. The new and updated version of LAP will offer an application unspecific interface allowing users to create any kind of grid job. The interface will offer the user an easy to use job description creation module and the possibility to upload executable and job input files.}
\item{Occasionally users would like to use existing job description files, but still submit their grid job through the LAP. To support this the LAP will offer an interface for uploading and submitting job description files.}
\end{itemize}

\end{document}

### 2009-04-27 doc/manuals/user_interface/ui-job.tex
Starting cleaning up documentation

\subsection{arcsub}\label{sec:arcsub}
The \texttt{arcsub}\index{arcsub}\index{commands:arcsub} command is the
most essential one, as it is used for submitting jobs to the Grid
resources. \index{submit job}. \texttt{arcsub} matches user's job
description to the information collected from the Grid, and the
optimal site is being selected for job submission. The job description
is then being forwarded to that site, in order to be submitted to the
Local Resource Management System (LRMS), which can be, e.g., PBS or
Condor or SGE etc.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcsub [options]}%$<$task ...$>$
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{ -c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject a specific site (cluster)\\
   \texttt{ -i, --indexurl}&\textit{url}&URL of an index server\\
   \texttt{ -e, --jobdescrstring}&\textit{filename}&string describing the job to be submitted\\
   \texttt{ -f, --jobdescrfile}&\textit{filename}&file describing the job to be submitted\\
   \texttt{ -j, --joblist}&\textit{filename}&file where the job IDs will be stored\\
   \texttt{ -D, --dryrun}&&add dryrun option to the job description\\
   \texttt{ -x, --dumpdescription}&&do not submit -- dump transformed job description to stdout\\
   \texttt{ -t, --timeout}&\textit{time}&timeout for queries in seconds (default 20)\\
   \texttt{ -d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
   \texttt{ -U, --unknownattr}&&allow unknown attributes in the job description\\
   \texttt{ -v, --version}&&print version information\\
   \texttt{ -?, --help}&&print help page\\
%   Arguments:&&\\
%   \texttt{ task ...}&&strings or files describing the jobs to be submitted\\
\end{longtable}

\begin{framed}
The -c and -i arguments accept ``urls'' of the format GRID:URL,

e.g. for index servers:

ARC0:ldap://grid.tsl.uu.se:2135/mds-vo-name=sweden,O=grid

CREAM:ldap://cream.grid.upjs.sk:2170/o=grid

or clusters:

ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid
\end{framed}

As a shorthand \texttt{-f} can be omitted if the job description file is put last on the commandline.

A simple \textit{"Hello World"} job could look like:

\begin{shaded}
\uicommand{arcsub -c ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid -f job.jsdl}
\end{shaded}

Assuming that the url represents an ARC0 cluster that the user is authorized to submit to and job.jsdl is the file below.

\begin{framed}
\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?>
<JobDefinition
 xmlns="http://schemas.ggf.org/jsdl/2005/11/jsdl"
 xmlns:posix="http://schemas.ggf.org/jsdl/2005/11/jsdl-posix">
 <JobDescription>
   <JobIdentification>
     <JobName>Hello World job</JobName>
   </JobIdentification>
   <Application>
     <posix:POSIXApplication>
       <posix:Executable>/bin/echo</posix:Executable>
       <posix:Argument>'Hello World'</posix:Argument>
       <posix:Output>out.txt</posix:Output>
       <posix:Error>err.txt</posix:Error>
     </posix:POSIXApplication>
   </Application>
   <DataStaging>
     <FileName>out.txt</FileName>
     <CreationFlag>overwrite</CreationFlag>
     <DeleteOnTermination>false</DeleteOnTermination>
   </DataStaging>
   <DataStaging>
     <FileName>err.txt</FileName>
     <CreationFlag>overwrite</CreationFlag>
     <DeleteOnTermination>false</DeleteOnTermination>
   </DataStaging>
 </JobDescription>
</JobDefinition>
\end{verbatim}
\end{framed}

\subsection{arcstat}
\label{sec:arcstat}

\begin{shaded}
   \uicommand{arcstat [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, --all}& & all jobs\\
   \texttt{-i, --joblist}& \textit{filename}& file containing a list of jobIDs\\
   \texttt{-c, --cluster}& & show information about a site (cluster)\\
   \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-i, --indexurl}& \textit{url} &URL of an index service\\
   \texttt{-q, --queues}&&show information about clusters and queues\\
   \texttt{-l, --long}& & long format (extended information)\\
   \texttt{-t, --timeout}& \textit{time}& timeout for queries (default 20 sec)\\
   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
   \texttt{-v, --version}& & print version information\\
   \texttt{-h, --help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

The arcstat command returns the status of jobs submitted to the Grid. Then \texttt{-c} and \texttt{-i} accept arguments in the GRID:URL notation explained in the description of \texttt{arcsub}.

Different sites may report slightly different job states, depending on
the installed software version.% A summary of essential job
%states\index{job:states} is:

%  ---  Does the table below need updating? ---

% \begin{longtable}{l|l|p{8cm}}
% ARC 0.3, ARC 0.4 & ARC 0.5, ARC0.6 & Description\\ \hline
%                    &\texttt{ACCEPTING} &job has reached the site\\
% \texttt{ACCEPTED}  &\texttt{ACCEPTED}  &job submitted but not yet processed\\
% \texttt{PREPARING} &\texttt{PREPARING} &input files are being retreived\\
%                    &\texttt{PREPARED}  &input files are retreived\\
% \texttt{SUBMITTING}&\texttt{SUBMITTING}&interaction with LRMS ongoing\\
% \texttt{INLRMS: Q} &\texttt{INLRMS:Q}  &job is queued by LRMS\\
% \texttt{INLRMS: R} &\texttt{INLRMS:R}  &job is running\\
%                    &\texttt{INLRMS:S}  &job is suspended\\
%                    &\texttt{INLRMS:E}  &job is finishing in LRMS\\
%                    &\texttt{INLRMS:O}  &job is in any other LRMS state\\
% \texttt{CANCELING} &\texttt{KILLING}   &job is being cancelled by user request\\
%                    &\texttt{EXECUTED}  &job is completed in LRMS\\
% \texttt{FINISHING} &\texttt{FINISHING} &output files are being transferred\\
% \texttt{FINISHED}  &\texttt{FINISHED}  &job is finished\\
%                    &\texttt{FAILED}    &job is finished with an error\\
%                    &\texttt{KILLED}    &job is cancelled by user request\\
% \texttt{DELETED}   &\texttt{DELETED}   &job is removed due to expiration time\\
% \end{longtable}

\subsection{arccat}
\label{sec:arccat}

It is often useful to monitor the job progress by checking what it
prints on the standard output or error. The command \texttt{arccat}
\index{arccat}\index{commands:arccat} assists here, extracting the
corresponding information from the execution cluster and pasting it
on the user's screen. It works both for running tasks and for the
finished ones. This allows a user to check the output of the
finished task without actually retreiving it.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arccat [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, --all}& & all jobs\\
   \texttt{-i, --joblist}& \textit{filename} & file containing a list of job IDs\\
   \texttt{-c, --cluster}& & show information about clusters\\
   \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-o, --stdout}& & show the stdout of the job (default)\\
   \texttt{-e, --stderr}& & show the stderr of the job\\
   \texttt{-l, --gmlog}& & show the grid manager's error log of the job\\
   \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
   \texttt{-v, --version}& & print version information\\
   \texttt{-h, --help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

The \texttt{arccat} command can return the standard output of a job
(\texttt{-o} option), the standard error (\texttt{-e} option) and the
errors reported by the Grid Manager (\texttt{-l} option).

\begin{framed}

\end{framed}

\subsection{arcget}
\label{sec:arcget}

To retrieve the results of a finished job, the \texttt{arcget}
\index{arcget}\index{commands:arcget} command should be used. It
will download the files specified by the \texttt{outputfiles}
attribute of job description to the user's computer.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcget [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, --all}& & all jobs\\
   \texttt{-i, --joblist}& \textit{filename} & file containing a list of jobIDs\\
   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
   \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-D, --dir} & \textit{dirname} & download directory (the job directory will be created in this directory)\\
   \texttt{-k, --keep}& & keep files on gatekeeper (do not clean)\\
   \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
   \texttt{-v, --version}& & print version information\\
   \texttt{-h, --help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

Only the results of jobs that have finished can be downloaded. The job
can be referred to either by the \texttt{jobID} that was returned by
\texttt{arcsub} at submission time, or by its name, if the job
description contained a job name attribute.

\subsection{arckill}
\label{sec:arckill}

It happens that a user may wish to cancel a job. This is done by using
the \texttt{arckill} \index{arckill}\index{commands:arckill} command. A
job can be killed amost on any stage of processing through the Grid.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arckill [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, --all}& & all jobs\\
   \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
   \texttt{-c, --cluster}& & show information about clusters\\
   \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-k, --keep}& & keep files on gatekeeper (do not clean)\\
   \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
   \texttt{-v, --version}& & print version information\\
   \texttt{-h, --help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

\begin{framed}
   Job cancellation is an asynchronous process, such that it
   may take a few minutes before the job is actually cancelled.
\end{framed}

% \subsection{arcresub}
% \label{sec:arcresub}
% 
% Quite often it happens that a user would like to re-submit a job, but
% has difficulties recovering the original job description xRSL file.
% This happens when xRSL files are created by scripts on-fly, and
% matching of xRSL to the job ID is not straightforward. The utility
% called \texttt{arcresub}\index{arcresub}\index{commands:arcresub} helps
% in such situations, allowing users to resubmit jobs known only by
% their job IDs.
% 
% \begin{framed}
%    Only jobs where the \verb#gmlog#\index{gmlog} attribute was
%    specified in the job description can be resubmitted.
% \end{framed}
% 
% \hspace*{0.5cm}
% \begin{shaded}
%    \uicommand{arcresub [options] [job ...]}
% \end{shaded}
% \versions{ARC 0.9}
% \begin{longtable}{llp{8cm}}
%    Options:&&\\
%    \texttt{-a, -all}& & all jobs\\
%    \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
%    \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
%    \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
%    \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
%    \texttt{-k, -kluster}&\verb#[-]#textem{name}&explicitly select or
%    reject a specific site (cluster) as re-submission target\\
%    \texttt{-K, -Klustlist}&\verb#[-]#textem{filename}&list of
%    clusters to select or reject as re-submission target\\
%    \texttt{-g, -giisurl}& \textit{url} &URL of a central Information System server\\
%    \texttt{-G, -giislist}&\textit{filename}&list of GIIS URLs\\
%    \texttt{-o, -joblist}&\textit{filename}&file where the job IDs will be stored\\
%    \texttt{~~~~-dryrun}&&add dryrun option to the xRSL\\
%    \texttt{~~~~-dumpxrsl}&&do not submit -- dump transformed xRSL to stdout\\
%    \texttt{~~~~-keep}& & keep files on gatekeeper (do not clean)\\
%    \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
%    \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
%    \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
%    \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
%    \texttt{-v, -version}& & print version information\\
%    \texttt{-h, -help}& & print help page\\
%    Arguments:&&\\
%    \texttt{job ...} && list of job IDs and/or jobnames\\
% \end{longtable}

\subsection{arcclean}
\label{sec:arcclean}

If a job fails, or you are not willing to retrieve the results for
some reasons, a good practice for users is not to wait for the Grid
Manager to clean up the job leftovers, but to use \texttt{arcclean}
\index{arcclean}\index{commands:arcclean} to release the disk space
and to remove the job ID from the list of submitted jobs and from
the Information System.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcclean [options] [job ...]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-a, --all}& & all jobs\\
   \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
   \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
   \texttt{-f, --force} & & removes the job ID from the local list\\
   & &~~~even if the job is not found on the Grid\\
   \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
   \texttt{-v, --version}& & print version information\\
   \texttt{-h, --help}& & print help page\\
   Arguments:&&\\
   \texttt{job ...} && list of job IDs and/or jobnames\\
\end{longtable}

Only jobs that have finished can be cleaned.

% \subsection{arcrenew}
% \label{sec:arcrenew}
% 
% Quite often, the user proxy expires while the job is still running (or
% waiting in a queue). In case such job has to upload output files to a
% Grid location (Storage Element), it will fail. By using the \texttt{arcrenew}
% \index{arcrenew}\index{commands:arcrenew} command, users can upload
% a new proxy to the job. This can be done while a job is still running,
% thus preventing it from failing, or whithin 24 hours (or whatever is
% the expiration time set by the site) after the job
% end. In the latter case, the Grid Manager will attempt to finalize
% the job by uploading the output files to the desired location.
% 
% \hspace*{0.5cm}
% \begin{shaded}
%    \uicommand{arcrenew [options] [job ...]}
% \end{shaded}
% \versions{ARC 0.9}
% \begin{longtable}{llp{8cm}}
%    Options:&&\\
%    \texttt{-a, -all}& & all jobs\\
%    \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
%    \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
%    \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
%    \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
%    \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
%    \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
%    \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
%    \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
%    \texttt{-v, -version}& & print version information\\
%    \texttt{-h, -help}& & print help page\\
%    Arguments:&&\\
%    \texttt{job ...} && list of job IDs and/or jobnames\\
% \end{longtable}
% 
% Prior to using \texttt{arcrenew}, be sure to actually create the
% new proxy!
% 
% \subsection{arcsync}
% \label{sec:arcsync}
% 
% If you are using User Interface installations on different machines,
% your local lists of submitted jobs will be different. To synchronise
% these lists with the information in the Information System, use the
% \texttt{arcsync} \index{arcsync}\index{commands:arcsync} command.
% 
% \hspace*{0.5cm}
% \begin{shaded}
%    \uicommand{arcsync [options]}
% \end{shaded}
% \versions{ARC 0.9}
% \begin{longtable}{llp{8cm}}
%    Options:&&\\
%    \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
%    \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
%    \texttt{-g, -giisurl}& \textit{url} &URL of a central Information System server\\
%    \texttt{-G, -giislist}&\textit{filename}&list of GIIS URLs\\
%    \texttt{-f, -force}&&don't ask for confirmation\\
%    \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
%    \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
%    \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
%    \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
%    \texttt{-v, -version}& & print version information\\
%    \texttt{-h, -help}& & print help page\\
% \end{longtable}
% 
% The ARC User Interface keeps a local list of jobs in the user's home
% directory (see section~\ref{sec:joblists}). If this file is lost,
% corrupt, or the user wants to recreate the file on a different
% workstation, the \texttt{arcsync} command will recreate this file from
% the information available in the Information System.

% \subsection{Auxilliary files}\label{sec:joblists}
% 
% User Interface keeps local job lists in two files:
% {\$}\verb#HOME/.ngjobs# and {\$}\verb#HOME/.arc/history#\footnote{In
%   ARC $\leq$ 0.5.48, {\$}\texttt{HOME/.nghistory}}.
% 
% \textbf{{\$}\texttt{HOME/.ngjobs}} is a local list of the user's active
% jobs. When a job is successfully submitted, it is added to this list,
% and when it is removed from the remote site, it is removed from this
% list. This list is used as the list of all active jobs when the user
% specifies \verb#-a# option to the various ARC user interface commands.
% For information about how to reconstruct this file in case it is
% damaged or you relocate to a different workstation, see
% section~\ref{sec:arcsync} about the \texttt{arcsync} command.
% 
% \textbf{{\$}\texttt{HOME/.arc/history}} contains the \texttt{jobID}s of
% the jobs the user has submitted together with the time of
% submission. This file is purely informational.

### 2009-04-28 doc/internal_notes/dmc/dmc.tex
Reorganizing doc catalogue and enabling make doc

--- 2008-01-22 doc/dmc/dmc.tex
@@
 
-  {\raggedleft NORDUGRID-TECH-16\par}
+  {\raggedleft NORDUGRID-TECH-23\par}
 
@@
 
-  {\centering \large Mattias Ellert\footnote{mattias.ellert@tsl.uu.se} \large \par}
+  {\centering \large Mattias Ellert\footnote{mattias.ellert@fysast.uu.se} \large \par}
 
@@
     virtual ~DataPointCustomDirect();
-    virtual bool start_reading(DataBufferPar& buffer);
-    virtual bool start_writing(DataBufferPar& buffer,
+    virtual DataStatus StartReading(DataBuffer& buffer);
+    virtual DataStatus StartWriting(DataBuffer& buffer,
                                     DataCallback *space_cb = NULL);
-    virtual bool stop_reading(void);
-    virtual bool stop_writing(void);
-    virtual bool analyze(analyze_t& arg);
-    virtual bool check(void);
-    virtual bool remove(void);
-    virtual bool list_files(std::list<FileInfo>& files, bool resolve = true);
+    virtual DataStatus StopReading();
+    virtual DataStatus StopWriting();
+    virtual DataStatus Check();
+    virtual DataStatus Remove();
+    virtual DataStatus ListFiles(std::list<FileInfo>& files,
+                                 bool long_list = false, bool resolve = false);
   };
@@
     ~DataPointCustom2() {};
-    virtual bool meta_resolve(bool source);
-    virtual bool meta_preregister(bool replication, bool force = false);
-    virtual bool meta_postregister(bool replication);
-    virtual bool meta_preunregister(bool replication);
-    virtual bool meta_unregister(bool all);
-    virtual bool list_files(std::list<FileInfo> &files, bool resolve = true);
+    virtual DataStatus Resolve(bool source);
+    virtual DataStatus PreRegister(bool replication, bool force = false);
+    virtual DataStatus PostRegister(bool replication);
+    virtual DataStatus PreUnregister(bool replication);
+    virtual DataStatus Unregister(bool all);
+    virtual DataStatus ListFiles(std::list<FileInfo>& files,
+                                 bool long_list = false, bool resolve = false);
   };
@@
 
-\bibliography{grid}
-\printindex
+%\bibliography{grid}
+%\printindex
 

### 2009-04-28 doc/manuals/user_interface/ui.tex
Reorganizing doc catalogue and enabling make doc

\documentclass{book}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=3cm,lmargin=2cm,rmargin=3cm}
\usepackage{makeidx}
\makeindex
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color,framed}
\usepackage{longtable}                             %for breaking tables

%%%%---- Use EPS figures and graphics with LaTeX ----
%\usepackage{graphics}
%%%%---- USE PNG or JPG figures and graphicx with PDFLaTeX ----
\usepackage{graphicx}
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref}
\hypersetup{
  pdfauthor = {},
  pdftitle = {ARC Clients},
  pdfsubject = {User's manual},
  pdfkeywords = {Grid,KnowARC,ARC,client},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}

\bibliographystyle{IEEEtran}

\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}[1]{\colorbox{shadecolor}{\texttt #1}}
\newcommand{\xrsl}[1]{\textbf{\sffamily #1}}
\newcommand{\uicommand}[1]{\textbf{\sffamily #1}}
\newcommand{\versions}[1]{\vspace{-0.3cm}(#1)\par}
\newcommand{\GT}{Globus Toolkit$\mathrm{^{TM}}$}
\newcommand{\globus}{Globus\textsuperscript{\textregistered}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

   \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
   \end{tabular}

   \hrulefill

   {\raggedleft NORDUGRID-MANUAL-?\par}

   {\raggedleft \today\par}

   \vspace*{2cm}

   %%%% ---- The title ----

   \begin{center}

      \textsc{\Large ARC Clients}
      \Large \par \textit{User's Manual}

      \vspace*{2cm}

      % {\centering \large Refers to ARC release series 0.4 and up \large \par}

   \end{center}

   \vspace*{2cm}

   % \abstract{
   %   This is the user's  manual to ARC User Interface.
   % }

\end{titlepage}
% \thispagestyle{empty} $ $
% \newpage
% $\ $
\tableofcontents
\newpage

\chapter{Introduction} \label{sec:intro}

The command line user interface of ARC consists of a set of commands
necessary for job submission and manipulation and data management. A
special utility also exists for test purposes.

\chapter{Commands} \label{sec:cli}

\section{Job submission and management} \label{sec:ui} 
\index{job management}
The following commands are used for job submission and management,
such as status check, results retrieval, cancellation, re-submission
and such. The jobs must be described using a job description language. ARC supports the languages JSDL, xRSL and JDL~\cite{xrsl}.

\input{ui-job}

\section{Data manipulation} \label{sec:dm} 
\index{data management} ARC provides basic data management tools,
which are simple commands for file copy and removal, with eventual use
of data indexing services.

\input{ui-data}

\section{Test suite} \label{sec:testsuite}
%\input{arctest}

\section{Third-party commands} \label{sec:ui-other}
%\input{ui-other}

\chapter{URLs}\label{sec:urls}
\index{URLs}
\input{xrsl-urls}

\chapter{Configuration} \label{sec:client.conf} 
\section{ARC Client Configuration}
%\input{ui-conf}

\bibliography{grid}
\printindex

\end{document}

### 2009-04-28 doc/manuals/ws-quick-guide/ws-quick-guide.tex
Reorganizing doc catalogue and enabling make doc

--- 2008-09-09 doc/ws-quick-guide/ws-quick-guide.tex
@@
 Now you may use the command line utility ``\texttt{arcinfo}'' to obtain a service
-description. Make sure user key and certificate and CA certificate(s) are
-readable from account you use, and run something like:
+description. You can do something like:
 
 \begin{lstlisting}
-./arcinfo -d VERBOSE -C $HOME/usercert.pem -K $HOME/userkey.pem
- -A /etc/grid-security/certificates https://localhost:60000/arex 2>error.log
+./arcinfo -c ARC1:https://localhost:60000/arex -l
 \end{lstlisting}
 
-This should produce XML output describing resource A-REX represents.
-Listing~\ref{lst:arcinfo} shows an example of proper output. In case of error see
-\texttt{error.log} file and report problems to ARC development team. Include
-content of \texttt{error.log} and server log.
+This should produce a description list~\ref{lst:arcinfo} of the resources A-REX
+represents. Below you can see an example of proper output.
 
-\begin{lstlisting}[caption={Example of a good arcinfo output},label=lst:arcinfo]
-<n:nordugrid xmlns:n="urn:nordugrid" xmlns:M0="urn:Mds"
-xmlns:nc0="urn:nordugrid-cluster">
-  <nc0:name>mobile2.home.net</nc0:name>
-  <nc0:aliasname>MINIMAL Computing Element</nc0:aliasname>
-  <nc0:comment>This is a minimal out-of-box CE setup</nc0:comment>
-  <nc0:issuerca>/O=Grid/O=Test/CN=CA</nc0:issuerca>
-  <nc0:issuerca-hash>4457e417</nc0:issuerca-hash>
-  <nc0:trustedca>/O=Grid/O=NorduGrid/CN=NorduGrid Certification
-Authority</nc0:trustedca>
-  <nc0:trustedca>/O=Grid/O=Test/CN=CA</nc0:trustedca>
-  <nc0:contactstring>gsiftp://mobile2.home.net:2811/jobs</nc0:contactstring>
-  <nc0:lrms-type>fork</nc0:lrms-type>
-  <nc0:architecture>i686</nc0:architecture>
-  <nc0:homogeneity>TRUE</nc0:homogeneity>
-  <nc0:nodeaccess>inbound</nc0:nodeaccess>
-  <nc0:nodeaccess>outbound</nc0:nodeaccess>
-  <nc0:totalcpus>1</nc0:totalcpus>
-  <nc0:usedcpus>0</nc0:usedcpus>
-  <nc0:cpudistribution>1cpu:1</nc0:cpudistribution>
-  <nc0:prelrmsqueued>0</nc0:prelrmsqueued>
-  <nc0:totaljobs>0</nc0:totaljobs>
-  <nc0:sessiondir-free>8469</nc0:sessiondir-free>
-  <nc0:sessiondir-total>43865</nc0:sessiondir-total>
-  <nc0:sessiondir-lifetime>10080</nc0:sessiondir-lifetime>
-  <M0:validfrom>20080220142836Z</M0:validfrom>
-  <M0:validto>20080220143036Z</M0:validto>
-  <nq0:name xmlns:nq0="urn:nordugrid-queue" name="fork">
-    <nq0:name>fork</nq0:name>
-    <nq0:status>inactive, grid-manager is down</nq0:status>
-    <nq0:comment>This queue is nothing more than a fork host</nq0:comment>
-    <nq0:schedulingpolicy>FIFO</nq0:schedulingpolicy>
-    <nq0:homogeneity>TRUE</nq0:homogeneity>
-    <nq0:nodecpu>Intel(R) Celeron(R) M CPU        420  @ 1.60GHz @ 1596.048
-MHz</nq0:nodecpu>
-    <nq0:architecture>i686</nq0:architecture>
-    <nq0:maxcputime>unlimited</nq0:maxcputime>
-    <nq0:maxwalltime>unlimited</nq0:maxwalltime>
-    <nq0:running>0</nq0:running>
-    <nq0:gridrunning>0</nq0:gridrunning>
-    <nq0:localqueued>0</nq0:localqueued>
-    <nq0:gridqueued>0</nq0:gridqueued>
-    <nq0:prelrmsqueued>0</nq0:prelrmsqueued>
-    <nq0:totalcpus>1</nq0:totalcpus>
-    <IM0:validfrom>20080220142836Z</M0:validfrom>
-    <M0:validto>20080220143036Z</M0:validto>
-    <nig0:name xmlns:nig0="urn:nordugrid-info-group" name="jobs">
-      <nig0:name>jobs</nig0:name>
-      <M0:validfrom>20080220142836Z</M0:validfrom>
-      <M0:validto>20080220143036Z</M0:validto>
-      <nj0:globalid xmlns:nj0="urn:nordugrid-job"
-name="gsiftp://mobile2.home.net:2811/jobs/636312034687631804
-289383">
+\begin{lstlisting}[caption={Example of proper arcinfo output},label=lst:arcinfo]
+Cluster: localhost
+ Health State: ok
 
-<nj0:globalid>gsiftp://mobile2.home.net:2811/jobs/636312034687631804289383</nj0:
-globalid>
-        <nj0:globalowner>/O=Grid/O=Test/CN=user</nj0:globalowner>
-        <nj0:jobname>TEST</nj0:jobname>
-        <nj0:execcluster>mobile2.home.net</nj0:execcluster>
-        <nj0:execqueue>fork</nj0:execqueue>
-        <nj0:submissionui>127.0.0.1:59790</nj0:submissionui>
-        <nj0:submissiontime>20080220005243Z</nj0:submissiontime>
-        <nj0:sessiondirerasetime>20080227005843Z</nj0:sessiondirerasetime>
-        <nj0:completiontime>20080220005843Z</nj0:completiontime>
-        <nj0:cpucount>1</nj0:cpucount>
-        <nj0:comment>GM: The grid-manager is down</nj0:comment>
-        <nj0:usedcputime>0</nj0:usedcputime>
-        <nj0:usedwalltime>0</nj0:usedwalltime>
-        <nj0:errors>Job submission to LRMS failed</nj0:errors>
-        <nj0:status>FAILED</nj0:status>
-        <nj0:rerunable>SUBMIT</nj0:rerunable>
-        <M0:validfrom>20080220142836Z</M0:validfrom>
-        <M0:validto>20080220143036Z</M0:validto>
-      </nj0:globalid>
-    </nig0:name>
-    <nig0:name xmlns:nig0="urn:nordugrid-info-group" name="users">
-      <nig0:name>users</nig0:name>
-      <M0:validfrom>20080220142836Z</M0:validfrom>
-      <M0:validto>20080220143036Z</M0:validto>
-    </nig0:name>
-  </nq0:name>
-</n:nordugrid>
+Location information:
+
+Domain information:
+
+Service information:
+ Service Name: MINIMAL Computing Element
+ Service Type: org.nordugrid.arex
+
+Endpoint information:
+ URL: https://localhost:60000/arex
+ Capabilities:
+  executionmanagement.jobexecution
+ Technology: webservice
+ Interface Name: OGSA-BES
+ Supported Profiles:
+  WS-I 1.0
+  HPC-BP
+ Implementor: NorduGrid
+ Implementation Name: A-REX
+ Implementation Version: 0.9
+ QualityLevel: development
+ Health State: ok
+ Serving State: production
+ Issuer CA: /O=Grid/O=NorduGrid/CN=NorduGrid Certification Authority
+ Trusted CAs:
+  /C=BE/O=BELNET/OU=BEGrid/CN=BEGrid CA/emailAddress=gridca@belnet.be
+  /C=FR/O=CNRS/CN=CNRS2-Projets
+  /DC=org/DC=ugrid/CN=UGRID CA
+  /C=BR/O=ICPEDU/O=UFF BrGrid CA/CN=UFF Brazilian Grid Certification Authority
+  /C=DE/O=DFN-Verein/OU=DFN-PKI/CN=DFN-Verein PCA Grid - G01
+  /C=PT/O=LIPCA/CN=LIP Certification Authority
+  /C=FR/O=CNRS/CN=GRID-FR
+  /C=FR/O=CNRS/CN=CNRS2
+  /C=TR/O=TRGrid/CN=TR-Grid CA
+  /C=NL/O=NIKHEF/CN=NIKHEF medium-security certification auth
+  /DC=org/DC=DOEGrids/OU=Certificate Authorities/CN=DOEGrids CA 1
+  /DC=ch/DC=cern/CN=CERN Trusted Certification Authority
+  /C=AU/O=APACGrid/OU=CA/CN=APACGrid/emailAddress=camanager@vpac.org
+  /C=IE/O=Grid-Ireland/CN=Grid-Ireland Certification Authority
+  /O=Grid/O=NorduGrid/CN=NorduGrid Certification Authority
+  /DC=RO/DC=RomanianGRID/O=ROSA/OU=Certification Authority/CN=RomanianGRID CA
+  /DC=bg/DC=acad/CN=BG.ACAD CA
+  /C=MX/O=UNAMgrid/OU=UNAM/CN=CA
+  /C=GR/O=HellasGrid/OU=Certification Authorities/CN=HellasGrid Root CA 2006
+  /C=CL/O=REUNACA/CN=REUNA Certification Authority
+  /DC=org/DC=balticgrid/CN=Baltic Grid Certification Authority
+  /C=IT/O=INFN/CN=INFN CA
+  /DC=me/DC=ac/DC=MREN/CN=MREN-CA
+  /C=FR/O=CNRS/CN=CNRS-Projets
+  /C=DE/O=DFN-Verein/OU=DFN-PKI/CN=DFN-Verein User CA Grid - G01
+  /C=UK/O=eScienceCA/OU=Authority/CN=UK e-Science CA
+  /C=RS/O=AEGIS/CN=AEGIS-CA
+  /C=SI/O=SiGNET/CN=SiGNET CA
+  /C=VE/O=Grid/O=Universidad de Los Andes/OU=CeCalCULA/CN=ULAGrid Certification Authority
+  /DC=ORG/DC=SEE-GRID/CN=SEE-GRID CA
+  /C=CH/O=Switch - Teleinformatikdienste fuer Lehre und Forschung/CN=SWITCH Personal CA
+  /C=RU/O=RDIG/CN=Russian Data-Intensive Grid CA
+  /C=HU/O=KFKI RMKI CA/CN=KFKI RMKI CA
+  /C=JP/O=KEK/OU=CRC/CN=KEK GRID Certificate Authority
+  /DC=EDU/DC=UTEXAS/DC=TACC/O=UT-AUSTIN/CN=TACC Root CA
+  /C=AT/O=AustrianGrid/OU=Certification Authority/CN=Certificate Issuer
+  /C=IL/O=IUCC/CN=IUCC/emailAddress=ca@mail.iucc.ac.il
+  /DC=TW/DC=ORG/DC=NCHC/CN=NCHC CA
+  /C=KR/O=KISTI/O=GRID/CN=KISTI Grid Certificate Authority
+  /DC=LV/DC=latgrid/CN=Certification Authority for Latvian Grid
+  /DC=NET/DC=PRAGMA-GRID/CN=PRAGMA-UCSD CA
+  /C=CH/O=SwissSign/CN=SwissSign CA (RSA IK May 6 1999 18:00:58)/emailAddress=ca@SwissSign.com
+  /C=MA/O=MaGrid/CN=MaGrid CA
+  /C=MK/O=MARGI/CN=MARGI-CA
+  /C=GR/O=HellasGrid/OU=Certification Authorities/CN=HellasGrid CA 2006
+  /C=TH/O=NECTEC/OU=GOC/CN=NECTEC GOC CA
+  /C=PL/O=GRID/CN=Polish Grid CA
+  /C=UK/O=eScienceRoot/OU=Authority/CN=UK e-Science Root
+  /DC=cz/DC=cesnet-ca/CN=CESNET CA
+  /C=TW/O=AS/CN=Academia Sinica Grid Computing Certification Authority Mercury
+  /DC=es/DC=irisgrid/CN=IRISGridCA
+  /C=JP/O=AIST/OU=GRID/CN=Certificate Authority
+  /C=JP/O=National Research Grid Initiative/OU=CGRD/CN=NAREGI CA
+  /DC=BR/DC=UFF/DC=IC/O=UFF LACGrid CA/CN=UFF Latin American and Caribbean Catch-all Grid CA
+  /C=CY/O=CyGrid/O=HPCL/CN=CyGridCA
+  /DC=CN/DC=Grid/CN=Root Certificate Authority at CNIC
+  /C=AR/O=e-Ciencia/OU=UNLP/L=CeSPI/CN=PKIGrid
+  /C=CN/O=HEP/CN=gridca-cn/emailAddress=gridca@ihep.ac.cn
+  /C=CA/O=Grid/CN=Grid Canada Certificate Authority
+  /CN=SWITCH CA/emailAddress=switch.ca@switch.ch/O=Switch - Teleinformatikdienste fuer Lehre und Forschung/C=CH
+  /DC=CN/DC=Grid/DC=SDG/CN=Scientific Data Grid CA
+  /C=HU/O=NIIF/OU=Certificate Authorities/CN=NIIF Root CA
+  /C=IR/O=IPM/O=IRAN-GRID/CN=IRAN-GRID CA
+  /C=FR/O=CNRS/CN=CNRS
+  /C=CH/O=Switch - Teleinformatikdienste fuer Lehre und Forschung/CN=SWITCHgrid Root CA
+  /C=AM/O=ArmeSFo/CN=ArmeSFo CA
+  /C=FR/O=CNRS/CN=GRID2-FR
+  /DC=net/DC=ES/O=ESnet/OU=Certificate Authorities/CN=ESnet Root CA 1
+  /DC=ch/DC=cern/CN=CERN Root CA
+  /DC=IN/DC=GARUDAINDIA/CN=Indian Grid Certification Authority
+  /C=DE/O=GermanGrid/CN=GridKa-CA
+  /C=SK/O=SlovakGrid/CN=SlovakGrid CA
+  /CN=SwissSign Bronze CA/emailAddress=bronze@swisssign.com/O=SwissSign/C=CH
+  /DC=EDU/DC=UTEXAS/DC=TACC/O=UT-AUSTIN/CN=TACC Classic CA
+  /C=BE/OU=BEGRID/O=BELNET/CN=BEgrid CA
+  /CN=SwissSign Silver CA/emailAddress=silver@swisssign.com/O=SwissSign/C=CH
+  /C=CH/O=Switch - Teleinformatikdienste fuer Lehre und Forschung/CN=SWITCH Server CA
+  /C=PK/O=NCP/CN=PK-GRID-CA
+  /C=DE/O=DFN-Verein/OU=DFN-PKI/CN=DFN-Verein Server CA Grid - G01
+  /C=HR/O=edu/OU=srce/CN=SRCE CA
+ Staging: staginginout
+ Job Descriptions:
+  ogf:jsdl:1.0
+
+Queue information:
+ Mapping Queue: default
+ Max Total Jobs: 100
+ Max Running Jobs: 10
+ Max Waiting Jobs: 99
+ Max Pre LRMS Waiting Jobs: 0
+ Max User Running Jobs: 5
+ Max Slots Per Job: 1
+ Doesn't Support Preemption
+ Total Jobs: 0
+ Running Jobs: 0
+ Waiting Jobs: 0
+ Suspended Jobs: 0
+ Staging Jobs: 0
+ Pre-LRMS Waiting Jobs: 0
+ Free Slots: 10
+ Free Slots With Duration:
+  P68Y1M5DT3H14M7S: 10
+ Used Slots: 0
+ Requested Slots: 0
+
+Manager information:
+ Resource Manager: torque
+ Doesn't Support Advance Reservations
+ Doesn't Support Bulk Submission
+ Total Physical CPUs: 10
+ Total Logical CPUs: 10
+ Total Slots: 10
+ Non-homogeneous Resource
+ Working area is nor shared among jobs
+ Working Area Total Size: 15
+ Working Area Free Size: 4
+ Working Area Life Time: P7D
+ Cache Area Total Size: 15
+ Cache Area Free Size: 4
+
+Execution Environment information:
+ Execution environment is a physical machine
+ Execution environment does not support inbound connections
+ Execution environment does not support outbound connections
 \end{lstlisting}
@@
 Please note that you can run similar \texttt{arcinfo} request against any ARC1
-service except Echo service. Output should always be XML description of the service.
+service except echo service.
 
-A-REX accepts jobs described in JSDL language. Example JSDL jobs are provided in
-\verb|$ARC_LOCATION/share/doc/| in files \verb|jsdl_simple.xml| and
-\verb|jsdl_stage.xml|. To submit job to A-REX service one may use
+A-REX accepts jobs described in the JSDL language. Example of JSDL jobs are
+provided in \verb|$ARC_LOCATION/share/doc/| in the files \verb|jsdl_simple.xml|
+and \verb|jsdl_stage.xml|. To submit job to A-REX service one may use the
 \texttt{arcsub} command:
@@
 \begin{lstlisting}
-$ARC_LOCATION/bin/arcsub https://localhost:60000/arex
- /usr/local/share/doc/arc/jsdl_simple.xml id.xml
+$ARC_LOCATION/bin/arcsub -c ARC1:https://localhost:60000/arex -f $ARC_LOCATION/share/doc/arc/jsdl_simple.xml -j id.xml
 \end{lstlisting}
+
+If everything goes properly somewhere in it's output there should be a message
+"Job submitted!", and a job identifier is obtained which will be stored
+in 'id.xml' file. One can then query job state with the \texttt{arcstat} utility:
 
@@
 \begin{lstlisting}
-$ARC_LOCATION/bin/arcstat id.xml 2>/dev/null
-Job status: Running/Submiting
+$ARC_LOCATION/bin/arcstat id.xml
+Job status: Running/Submitting
 
-$ARC_LOCATION/bin/arcstat id.xml 2>/dev/null
+$ARC_LOCATION/bin/arcstat id.xml
 Job status: Running/Finishing
 
-$ARC_LOCATION/bin/arcstat id.xml 2>/dev/null
+$ARC_LOCATION/bin/arcstat id.xml
 Job status: Finished/Finished
@@
 
-Full set of A-REX client tools consists of \texttt{arcsub}, \texttt{arcstat},
-\texttt{arckill}, \texttt{arcget} and \texttt{arcclean} commands.
-For more information please see man pages ot those utilities.
+Some of the of A-REX client tools consists of \texttt{arcsub}, \texttt{arcstat},
+\texttt{arckill}, \texttt{arcget} and \texttt{arcclean} commands. For more
+information please see the man pages of those utilities.
 

### 2009-04-28 doc/tech_doc/hed/counter.tex
Reorganizing doc catalogue and enabling make doc

--- 2007-04-24 doc/hed/counter.tex
@@
 \title{ARC Counter Library\\
-Requirements and Design}
+Requirements and Design
+\footnote{This document was used as input for discussions during
+NorduGrid Technical Meeting in L\"ubeck May 2--4, 2007. For the final
+design, see the Doxygen documentation of http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/libs/counter/Counter.h}}
 \author{Markus Nord\'en}

### 2009-04-28 doc/tech_doc/storage_system/arc-storage-documentation.tex
Reorganizing doc catalogue and enabling make doc

--- 2009-04-23 doc/storage_system/arc-storage-documentation.tex
@@
 
-In addition to the Logical Names we can refer to a file or collection simply by its GUID, or we can use GUID's and Logical Names together, as seen in Figure~\ref{fig:namespace}.
+In addition to the Logical Names a file or collection can be refered to simply by its GUID, or you can use GUID's and Logical Names together, as seen in Figure~\ref{fig:namespace}.
 
@@
 
-An example of the hierarchy of the global namespace is shown in Figure~\ref{fig:namespace}: if we have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then we can refer to this file with the Logical Name \verb#1234/green/orange/huge#, which means that from the collection called \verb!1234!
-we have to follow along the path: \verb!green!, \verb!orange!, \verb!huge!.
+An example of the hierarchy of the global namespace is shown in Figure~\ref{fig:namespace}: if you have a collection with GUID \verb#1234#, and there is a collection called \verb#green# in it, and in \verb#green# there is another collection called \verb#orange#, and in \verb#orange# there is a file called \verb#huge#, then you can refer to this file with the Logical Name \verb#1234/green/orange/huge#, which means that from the collection called \verb!1234!
+you have to follow along the path: \verb!green!, \verb!orange!, \verb!huge!.
 
@@
 But these internal services are communicating with each other, thus they have to know which other services to trust.
-We call this aspect of the security architecture of Chelonia `inter-service authorization'.
+This aspect of the security architecture of Chelonia is called `inter-service authorization'.
 
@@
 
-Besides specifying only one user with a DN, there are other types of access control rules: we can have a rule for a whole VO (Virtual Organization) or for all users, like this:
+Besides specifying only one user with a DN, there are other types of access control rules. For example, you can have a rule for a whole VO (Virtual Organization) or for all users, like this:
 
@@
 This TURL is unique for each request, even for requests for the same replica, and this TURL is only valid for one download.
-Currently we configure the storage element services to not do any authorization, and we use these one-time TURL's to ensure that only authorized users can access the contents of the storage elements.
+Currently the storage element services are  configured to not do any authorization, and these one-time TURL's are used to ensure that only authorized users can access the contents of the storage elements.
 
@@
     \item The Librarian tries to traverse the Logical Name and it stops at the last possible point and returns the metadata of the last element.
-Because the user wants to put his collection in a new path but into an existing collection, we expect that only the \verb#/ourvo/common# part of the LN can be traversed.
+Because the user wants to put his collection in a new path but into an existing collection, only the \verb#/ourvo/common# part of the LN is expected to be traversed.
 If the Librarian can traverse the whole LN it means that there is already an existing file or collection by that name.
@@
 The Librarian services use the A-Hash services to store the metadata, using the GUID's as unique ID's.
-For this, we have to represent the metadata as property-value pairs organized in section.
+For this, the metadata have to be represented as property-value pairs organized in section.
 The files, the collections and the mount points have some common attributes: the GUID, the type and the owner which are in the \emph{entry} section; the creation and modification timestamps in the \emph{timestamps} section, a list of access control rules in the \emph{policy} section; the list of parent collections where this entry has a `hardlink' in the \emph{parents} section; and optional, arbitrary property-value pairs in the \emph{metadata} section.
@@
 A file in Chelonia has a logical entry in the namespace, and one or more physical replicas on the storage nodes.
-We store the locations of the replicas in the \emph{locations} section.
+The locations of the replicas have to be stored in the \emph{locations} section.
 A location contains the URL of the Shepherd service, the local ID of the file within the storage node and the state of the replica.
@@
 However it cannot be guaranteed that the contents of a closed collection remains the same (e.g.
-if we include in our collection a file which we don't own, then the owner of the file can still remove it), but if a closed collection is changed, its state will be broken, and this state can never be changed again, which means you will always know if a closed collection is not intact anymore.
+if a user include in his collection a file which he does not own, then the owner of the file can still remove it), but if a closed collection is changed, its state will be broken, and this state can never be changed again, which means you will always know if a closed collection is not intact anymore.
 This state is stored in the \emph{states} section, and its possible values are `\textbf{no}' if the collection is not closed, `\textbf{yes}' if the collection is closed and `\textbf{broken}' if the collection was closed, but something happened, and its contents have been changed.
@@
 For inter-service authorization the services can be configured to only accept connections from trusted services.
-A service is trusted if we know its DN.
-For each service we can write trusted DN's in the configuration, or we can periodically get the list from an A-Hash.
+A service is trusted if its DN is known.
+For each service the trusted DN's can be inserted to the configuration, or the list can periodically be retrieved from an A-Hash.
 This can be configured with a `\emph{TrustManager}' section in the service configuration, which contains these entries:
@@
 
-The TrustManager section itself have the following optional attributes: \textbf{FromFile} can specify a filename if we want to get the whole TrustManager configuration from a separate file; and \textbf{Force} indicates (with `yes' or `no', default: `yes') that we want to ensure that our service does not even get the request if the client of the incoming connection is not trusted.
+The TrustManager section itself have the following optional attributes: \textbf{FromFile} can specify a filename if the whole TrustManager configuration should be taken from a separate file; and \textbf{Force} indicates (with `yes' or `no', default: `yes') that you want to ensure that your service does not even get the request if the client of the incoming connection is not trusted.
 
@@
 \item \emph{changeID} is an arbitrary ID which is used in the response to refer to this part of the request
-\item \emph{ID} points to the object we want to change
+\item \emph{ID} points to the object to be change
 \item \emph{changeType} can be `\textbf{set}' to set the property within the section to value, `\textbf{unset}' to remove the property from the section regardless of the value, `\textbf{delete}' to remove the whole object
@@
 In the metadata of the new entry the `type' property in the `entry' section defines whether it is a file, a collection or a mount point.
-The `GUID' property can contain a GUID if we want to specify the GUID of the new entry, and don't want the Librarian to generate a random one.
+The `GUID' property can contain a GUID if you want to specify the GUID of the new entry, and don't want the Librarian to generate a random one.
 

### 2009-04-28 doc/tech_doc/xrsl/jsdl-arc-schema.tex
Reorganizing doc catalogue and enabling make doc

  \begin{verbatim}

<?xml version="1.0" encoding="UTF-8"?>
<xsd:schema xmlns:xsd="http://www.w3.org/2001/XMLSchema"
	    xmlns="http://www.nordugrid.org/ws/schemas/jsdl-arc"
	    xmlns:jsdl-arc="http://www.nordugrid.org/ws/schemas/jsdl-arc"
	    targetNamespace="http://www.nordugrid.org/ws/schemas/jsdl-arc">

 <xsd:simpleType name="GMState_Type">
  <xsd:restriction base="xsd:string">
   <xsd:enumeration value="ACCEPTED"/>
   <xsd:enumeration value="PREPARING"/>
   <xsd:enumeration value="SUBMIT"/>
   <xsd:enumeration value="INLRMS"/>
   <xsd:enumeration value="FINISHING"/>
   <xsd:enumeration value="FINISHED"/>
   <xsd:enumeration value="DELETED"/>
   <xsd:enumeration value="CANCELING"/>
  </xsd:restriction>
 </xsd:simpleType>

 <xsd:complexType name="Version_Type">
  <xsd:sequence>
   <xsd:element name="UpperExclusive" type="xsd:string" minOccurs="0"/>
   <xsd:element name="LowerExclusive" type="xsd:string" minOccurs="0"/>
   <xsd:element name="Exact" type="xsd:string" minOccurs="0" maxOccurs="unbounded"/>
   <xsd:element name="Exclusive" type="xsd:boolean" minOccurs="0"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:simpleType name="SessionType_Type">
  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
  <!-- xsd:element ref="SessionType" minOccurs="0"/ -->
  <xsd:restriction base="xsd:string">
   <xsd:enumeration value="INTERNAL"/>
   <xsd:enumeration value="LIMITED"/>
   <xsd:enumeration value="READONLY"/>
   <xsd:enumeration value="FULL"/>
  </xsd:restriction>
 </xsd:simpleType>

 <xsd:simpleType name="IsExecutable_Type">
  <xsd:documentation> For jsdl:DataStaging_Type (default: false) </xsd:documentation>
  <!-- xsd:element ref="IsExecutable" minOccurs="0"/ -->
  <xsd:restriction base="xsd:boolean"/>
 </xsd:simpleType>

 <xsd:simpleType name="FileParameters_Type">
  <xsd:documentation> For jsdl:DataStaging_Type </xsd:documentation>
  <!-- xsd:element ref="IsExecutable" minOccurs="0"/ -->
  <xsd:restriction base="xsd:string"/>
 </xsd:simpleType>

 <xsd:simpleType name="JoinOutputs_Type">
  <xsd:documentation> For jsdl:JobDescription_Type (default: false) </xsd:documentation>
  <!-- xsd:element ref="JoinOutputs" minOccurs="0"/ -->
  <xsd:restriction base="xsd:boolean"/>
 </xsd:simpleType>

 <xsd:simpleType name="Reruns_Type">
  <xsd:documentation> For jsdl:JobDescription_Type (default: false) </xsd:documentation>
  <!-- xsd:element ref="Reruns" minOccurs="0"// -->
  <xsd:restriction base="xsd:integer"/>
 </xsd:simpleType>

 <xsd:complexType name="RunTimeEnvironment_Type">
  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
  <!-- xsd:element ref="RunTimeEnvironment" minOccurs="0" maxOccurs="unbounded"/ -->
  <xsd:sequence>
   <xsd:element name="Name" type="xsd:string"/>
   <xsd:element name="Version" type="Version_Type" minOccurs="0"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:complexType name="Middleware_Type">
  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
  <!-- xsd:element ref="Middleware" minOccurs="0" maxOccurs="unbounded"/ -->
  <xsd:sequence>
   <xsd:element name="Name" type="xsd:string"/>
   <xsd:element name="Version" type="Version_Type" minOccurs="0"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:complexType name="RemoteLogging_Type">
  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
  <!-- xsd:element ref="RemoteLogging" minOccurs="0" maxOccurs="3"/ -->
  <xsd:sequence>
   <xsd:element name="URL" minOccurs="1" maxOccurs="1" type="xsd:anyURI"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:complexType name="CredentialServer_Type">
  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
  <!-- xsd:element ref="CredentialServer" minOccurs="0"/ -->
  <xsd:sequence>
   <xsd:element name="URL" minOccurs="1" maxOccurs="1" type="xsd:anyURI"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:complexType name="LocalLogging_Type">
  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
  <!-- xsd:element ref="LocalLogging" minOccurs="0" maxOccurs="1"/ -->
  <xsd:sequence>
   <xsd:element name="Directory" minOccurs="1" maxOccurs="1" type="xsd:string"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:simpleType name="AccessControlType_Type">
  <xsd:restriction base="xsd:string">
   <xsd:enumeration value="GACL"/>
  </xsd:restriction>
 </xsd:simpleType>
 <xsd:complexType name="AccessControl_Type">
  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
  <!-- xsd:element ref="AccessControl" minOccurs="0"/ -->
  <xsd:sequence>
   <xsd:element name="OwnerAlwaysAllowed" type="xsd:boolean" minOccurs="0"/>
   <xsd:element name="Type" type="AccessControlType_Type" minOccurs="0"/>
   <xsd:element name="Content" minOccurs="0" type="xsd:string"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:simpleType name="NotificationType_Type">
  <xsd:restriction base="xsd:string">
   <xsd:enumeration value="Email"/>
  </xsd:restriction>
 </xsd:simpleType>
 <xsd:complexType name="Notify_Type">
  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
  <!-- xsd:element ref="Notify" minOccurs="0" maxOccurs="3"/ -->
  <xsd:sequence>
   <xsd:element name="Type" type="NotificationType_Type" minOccurs="0"/>
   <xsd:element name="Endpoint" minOccurs="0" type="xsd:string"/>
   <xsd:element name="State" minOccurs="1" maxOccurs="unbounded" type="GMState_Type"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:simpleType name="SessionLifeTime_Type">
  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
  <!-- xsd:element ref="SessionLifeTime" minOccurs="0" maxOccurs="1"/ -->
  <xsd:restriction base="xsd:long"/>
 </xsd:simpleType>

 <xsd:simpleType name="GridTimeLimit_Type">
  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
  <!-- xsd:element ref="GridTimeLimit" minOccurs="0" maxOccurs="1"/ -->
  <xsd:restriction base="xsd:positiveInteger"/>
 </xsd:simpleType>

 <xsd:complexType name="CandidateTarget_Type">
  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
  <!-- xsd:element ref="jsdl-arc:CandidateTarget" minOccurs="0" maxOccurs="1"/ -->
  <xsd:sequence>
   <xsd:element name="HostName" minOccurs="0" type="xsd:string"/>
   <xsd:element name="QueueName" minOccurs="0" type="xsd:string"/>
  </xsd:sequence>
 </xsd:complexType>

 <xsd:simpleType name="Time_Type">
  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
  <!-- xsd:element ref="ProcessingStartTime" minOccurs="0" maxOccurs="1"/ -->
  <xsd:restriction base="xsd:dateTime"/>
 </xsd:simpleType>

 <!--=========================================================-->
 <xsd:element name="IsExecutable" type="IsExecutable_Type"/>
 <xsd:element name="FileParameters" type="FileParameters_Type"/>
 <xsd:element name="RunTimeEnvironment" type="RunTimeEnvironment_Type"/>
 <xsd:element name="Middleware" type="Middleware_Type"/>
 <xsd:element name="RemoteLogging" type="RemoteLogging_Type"/>
 <xsd:element name="LocalLogging" type="LocalLogging_Type"/>
 <xsd:element name="AccessControl" type="AccessControl_Type"/>
 <xsd:element name="Notify" type="Notify_Type"/>
 <xsd:element name="SessionLifeTime" type="SessionLifeTime_Type"/>
 <xsd:element name="SessionType" type="SessionType_Type"/>
 <xsd:element name="JoinOutputs" type="JoinOutputs_Type"/>
 <xsd:element name="Reruns" type="Reruns_Type"/>
 <xsd:element name="CredentialServer" type="CredentialServer_Type"/>
 <xsd:element name="GridTimeLimit" type="GridTimeLimit_Type"/>
 <xsd:element name="CandidateTarget" type="CandidateTarget_Type"/>
 <xsd:element name="ProcessingStartTime" type="Time_Type"/>

</xsd:schema>

  \end{verbatim}

### 2009-04-28 doc/tech_doc/xrsl/xrsl-gm.tex
Reorganizing doc catalogue and enabling make doc

  \phantomsection
  \index{XRSL:attribute:sstdin}\addcontentsline{toc}{subsection}{sstdin}
  \hspace*{0.5cm}
  \begin{shaded}
    \xrsl{sstdin}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC $\leq$ 0.5.25}
  \begin{tabular}{lp{13cm}}
    User input:&\verb##\\
    GM input:&\verb#(sstdin=<filename>)#\\
    Example:&\verb#(sstdin="myinput.dat")#\\
  \end{tabular}

  Internal attribute for the standard input. Can also be spelled \texttt{stdinput}.
  Only needed for GRAM compatibility, not used by ARC as such.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{filename} & standard input file name\\
  \end{tabular}

  \phantomsection
  \index{XRSL:attribute:action}\addcontentsline{toc}{subsection}{action}
  \hspace*{0.5cm}
  \begin{shaded}
    \xrsl{action}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb##\\
    GM input:&\verb#(action="request"|"cancel"|"clean")#\\
    Example:&\verb#(action="request")#\\
  \end{tabular}

  Action to be taken by the gatekeeper: submit the job, cancel job
  execution, or clear the results of the job (also cancels the job).

  \phantomsection
  \index{XRSL:attribute:savestate}\addcontentsline{toc}{subsection}{savestate}
  \hspace*{0.5cm}
  \begin{shaded}
    \xrsl{savestate}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC $\leq$ 0.5.25}
  \begin{tabular}{lp{13cm}}
    User input:&\verb##\\
    GM input:&\verb#(savestate="yes"|"no")#\\
    Example:&\verb#(savestate="yes")#\\
  \end{tabular}

  If "yes", input RSL is stored in a temporary file at the
  gatekeeper. Must be always set as "yes" in the current implementation.
  Only needed for GRAM compatibility, not used by ARC as such.

  \phantomsection
  \index{XRSL:attribute:lrmstype}\addcontentsline{toc}{subsection}{lrmstype}
  \hspace*{0.5cm}
  \begin{shaded}
    \xrsl{lrmstype}
  \end{shaded}
  \versions{ARC $\leq$ 0.3.15}
  \begin{tabular}{lp{13cm}}
    User input:&\verb##\\
    GM input:&\verb#(lrmstype=<string>)#\\
    Example:&\verb#(lrmstype="pbs")#\\
  \end{tabular}

  LRMS type, indicating which submission script is to be invoked.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & LRMS type \\
  \end{tabular}

  \phantomsection
  \index{XRSL:attribute:hostName}\addcontentsline{toc}{subsection}{hostName}
  \hspace*{0.5cm}
  \begin{shaded}
    \xrsl{hostName}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb##\\
    GM input:&\verb#(hostname=<string>)#\\
    Example:&\verb#(hostName="grid.quark.lu.se")#\\
  \end{tabular}

  Submission host name.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & host name, as passed by the UI\\
  \end{tabular}

  \phantomsection
  \index{XRSL:attribute:jobid}\addcontentsline{toc}{subsection}{jobid}
  \hspace*{0.5cm}
  \begin{shaded}
    \xrsl{jobid}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb##\\
    GM input:&\verb#(jobid=<string>)#\\
    Example:&\verb#(jobid="grid.quark.lu.se:2119/jobmanager-ng/157111017133827")#\\
  \end{tabular}

  Unique job identification string, needed for cancellation and clean-up.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & global job ID\\
  \end{tabular}

  It can also be provided during submission of the job and should be
  unique to a computing element (cluster).

  \phantomsection
  \index{XRSL:attribute:clientxrsl}\addcontentsline{toc}{subsection}{clientxrsl}
  \hspace*{0.5cm}
  \begin{shaded}
    \xrsl{clientxrsl}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.28, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb##\\
    GM input:&\verb#(clientxrsl=<string>)#\\
    Example:&\verb#(clientxrsl="&(executable=/bin/echo)(arguments=boo)")#\\
  \end{tabular}

  Job description XRSL string as submitted by the user, before being pre-processed by the client.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & original XRSL description submitted by the user\\
  \end{tabular}

  This attribute is added by the User Interface during pre-processing, and is used for job re-submission in order to repeat brokering and matchmaking.

  \phantomsection
  \index{XRSL:attribute:clientsoftware}\addcontentsline{toc}{subsection}{clientsoftware}
  \hspace*{0.5cm}
  \begin{shaded}
    \xrsl{clientsoftware}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.35, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb##\\
    GM input:&\verb#(clientsoftware=<string>)#\\
    Example:&\verb#(clientsoftware="nordugrid-arc-0.5.39")#\\
  \end{tabular}

  Version of ARC client used to submit the job.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & \\
  \end{tabular}

  This attribute is added by the User Interface during pre-processing.

### 2009-04-28 doc/tech_doc/xrsl/xrsl-urls.tex
Reorganizing doc catalogue and enabling make doc

File locations in ARC can be specified both as local file
names, and as Internet standard \textit{Uniform Resource Locators
  (URL)}\index{URL}. There are also some additional URL
\textit{options} that can be used.

The following transfer protocols and metadata servers are supported:

\begin{tabular}{lp{10cm}}
   \texttt{ftp} & ordinary \textit{File Transfer Protocol (FTP)}\\
   \texttt{gsiftp} & GridFTP, the \globus\ -enhanced FTP protocol with
security, encryption, etc. developed by The Globus Alliance \cite{globus}\\
   \texttt{http} & ordinary \textit{Hyper-Text Transfer Protocol (HTTP)} with PUT and GET methods using multiple streams\\
   \texttt{https} & HTTP with SSL v3\\
   \texttt{httpg} & HTTP with \globus\  GSI\\
   \texttt{ldap} & ordinary \textit{Lightweight Data Access Protocol (LDAP)}~\cite{ldap}\\
   \texttt{rc} & \globus\  \textit{Replica Catalog (RC)}~\cite{rc}\\
   \texttt{rls} & \globus\  \textit{Replica Location Service (RLS)}~\cite{rls}\\
   \texttt{fireman} & Fireman indexing service of EGEE gLite~\cite{glite}\\
   \texttt{lfc} & LFC catalog and indexing service of EGEE gLite~\cite{glite}\\
   \texttt{se} & ARC Smart Storage Element service~\cite{se}\\
   \texttt{srm} & Storage Resource Manager (SRM) service~\cite{srm}\\
   \texttt{file} & local to the host file name with a full path\\
\end{tabular}

An URL can be used in a standard form, i.e.
\begin{shaded}
   \verb#<protocol>://host[:port]/<file>#
\end{shaded}

Or, to enhance the performance, it can have additional options:
\begin{shaded}
   \verb#<protocol>://host[:port][;option[;option[...]]]/<file>#
\end{shaded}

For a metadata service URL, construction is the following:
\begin{shaded}
   \verb#rc://rc://[location[|location[...]]@]<host>[:port]/<DN>/<lfn>#\\
   \verb#rls://[url[|url[...]]@]<host>[:port]/<lfn>#\\
   \verb#fireman://[url[|url[...]]@]<host>[:port]/<service_path>?<lfn>#\\
   \verb#lfc://[url[|url[...]]@]<host>[:port]/<lfn>#
\end{shaded}

For the Smart Storage Element service, the syntax is
\begin{shaded}
   \verb#se://host[:port][;options]/path[?file_id]#
\end{shaded}

For the SRM service, the syntax is
\begin{shaded}
   \verb#srm://<host>[:port][;options]/[service_path?SFN=]<file_id>#
\end{shaded}

Versions 1.1 and 2.2 of the SRM protocol are supported. The
default \emph{service\_path} is srm/managerv2 when the server supports
v2.2, srm/managerv1 otherwise.

The URL components are:

\begin{tabular}{lp{10cm}}
   \verb#location#&\verb#<location_name_in_RC>[;option[;option[...]]]#\\
   \verb#host[:port]#&IP address of a server\\
   \verb#DN#&Distinguished Name (as in LDAP) of an RC collection\\
   \verb#lfn#&Logical File Name\\
   \verb#url#&URL of the file as registered in RLS/Fireman\\
   \verb#service_path#&End-point path of the Web service\\
   \verb#file#&local to the host file name with a full path\\
\end{tabular}

The following options are supported for location URLs:\index{URL:options}

\begin{tabular}{lp{10cm}}
   \verb#threads=<number># & specifies number of parallel
   streams to be used by GridFTP or HTTP(s,g); default value is 1,
   maximal value is 10\\
   \verb#cache=yes|no|renew# & indicates whether the GM should
   cache the file; default for input files is \verb#yes#. \verb#renew# forces a download of the file, even if the cached copy is still valid. (NB: cached files can not be modified by jobs)\\
   \verb#readonly=yes|no# & for transfers to \verb#file://# destinations, specifies whether the file should be
   read-only (unmodifiable) or not; default is \verb#yes# \\
   \verb#secure=yes|no# & indicates whether the GridFTP data
   channel should be encrypted; default is \verb#no#\\
   \verb#blocksize=<number># & specifies size of
   chunks/blocks/buffers used in GridFTP or HTTP(s,g) transactions;
   default is protocol dependent\\ 
   \verb#checksum=cksum|md5|adler32# & specifies the algorithm for checksum to be
   computed (ev. provided to the indexing server). This is overridden
   by any metadata options specified (see below). If this option is
   not provided, the default for the protocol is used. \verb#checksum=no#
   in a source URL will disable checksum verification.\\ 
   \verb#exec=yes|no# & means the file should be treated as executable\\
   \verb#preserve=yes|no# & specify if file must be uploaded to this
   destination even if job processing failed (default is \verb#no#)\\
   \verb#pattern=<pattern># & defines file matching pattern; currently
   works for file listing requests sent to an \verb#se://# endpoint\\
   \verb#guid=yes|no# & make software use GUIDs instead of LFNs while
   communicating to indexing services; meaningful for \verb#rls://#
   only\\
   \verb#overwrite=yes|no# & make software try to overwrite existing
   file(s), i.e. before writing to destination, tools will try to remove
   any information/content associated with specified URL\\
   \verb#protocol=gsi|gssapi# & to distinguish between two kinds of 
   \verb#httpg#. \verb#gssapi# stands for implemention using only GSSAPI 
   functions to wrap data and \verb#gsi# uses additional headers as 
   implmented in Globus IO\\
   \verb#spacetoken=<pattern># & specify the space token to be used for
   uploads to SRM storage elements supporting SRM version 2.2 or higher\\
   \verb#autodir=yes|no# & specify if before writing to specified location
   software should try to create all directories mentioned in specified
   URL. Currently this applies to FTP and GridFTP only. Default for those
   protocols is \verb#yes#\\
\end{tabular}

Local files are referred to by specifying either a location relative
to the job submission working directory, or by an absolute path (the
one that starts with "/"), preceded with a \verb#file://# prefix.

Metadata service URLs also support metadata options which can be used
for register additional metadata attributes or query the service using
metadata attributes. These options are specified at the end of the LFN
and consist of name and value pairs separated by colons. The following
attributes are supported:

\begin{tabular}{lp{10cm}}
   \verb#guid# & GUID of the file in the metadata service \\
   \verb#checksumtype# & Type of checksum. Supported values are cksum
   (default), md5 and ad (adler32 checksum) \\
   \verb#checksumvalue# & The checksum of the file \\
\end{tabular}

Currently these metadata options are only supported for lfc:// URLs.

\begin{framed}
   Examples of URLs are:\\
   \\
   \verb#http://grid.domain.org/dir/script.sh#\\
   \verb#gsiftp://grid.domain.org:2811;threads=10;secure=yes/dir/input_12378.dat#\\
   \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
   \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#\\
   \verb#rls://gsiftp://se.domain.org/datapath/file25.dat@grid.domain.org:61238/myfile02.dat#$^1$\\
   \verb#fireman://fireman_host:8443/glite-data-catalog-interface/FiremanCatalog?data.root#\\
   \verb#file:///home/auser/griddir/steer.cra#\\
   \verb#lfc://srm://srm.domain.org/griddir@lfc.domain.org//user/file1:guid=\# \\
   \verb#    bc68cdd0-bf94-41ce-ab5a-06a1512764dc:checksumtype=ad:checksumvalue=12345678#$^2$\\
   \verb#lfc://;cache=no@lfc.domain.org/:guid=bc68cdd0-bf94-41ce-ab5a-06a1512764d#$^3$\\
\end{framed}

$^1$This is a destination URL. The file will be copied to the GridFTP
server at se.domain.org with the path datapath/file25.dat and
registered in the RLS indexing service at grid.domain.org with the LFN
myfile02.dat.

$^2$This is a destination URL. The file will be copied to
srm.domain.org at the path griddir/file1 and registered to the LFC
service at lfc.domain.org with the LFN /user/file1. The given GUID
and checksum attributes will be registered.

$^3$This is a source URL. The file is registered in the LFC service at
lfc.domain.org with the given GUID and can be copied or queried by
this URL. Note that as URL options are part of the location (physical)
URL, in meta service URLs the options must be part of the location
URL, even if the location URL is empty.

### 2009-04-28 doc/tech_doc/xrsl/xrsl-user.tex
Reorganizing doc catalogue and enabling make doc

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:executable}\addcontentsline{toc}{subsection}{executable}
  \begin{shaded}
    \xrsl{executable}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(executable=<string>)#\\
    GM input:& for ARC $\leq$ 0.5.25:  \verb#(executable="/bin/echo")#\\
    Example:&\verb#(executable="local_to_job.exe")#\\
  \end{tabular}

  The executable to be submitted as a main task to a Local Resource
  Management System (LRMS).

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string}&file name (including path), local to the computing
    element (CE)\\
  \end{tabular}

  Executable is a file that has to be executed as the main process
  of the task. It could be either a pre-compiled binary, or a
  script. Users may transfer their own executables, or use the ones
  known to be already installed on the remote system (CE).

  If an executable has to be transferred to the destination site (CE) from some source, it
  has to be specified in the \texttt{inputFiles} list. If it is not specified in \texttt{inputFiles},
  the source is expected to be local to the user (client) and
  will be added as such to the \texttt{inputFiles} list by the UI.

  If the file name starts with a leading slash ("\verb#/#"), it is considered
  to be \textbf{the full path to the executable at the destination site (CE)}; otherwise
  the location of the file is \textbf{relative} to the session
  directory (where job input and files are stored).

  If the xRSL string is entered from the command line and is enclosed
  in double quotes, standard shell expansion of variables takes place.
  That is, if the file name
  contains an environment variable ("\verb#$...#"), %$
  the value of this variable is resolved locally, but if the name
  itself is also
  enclosed in double quotes, it will be resolved at the remote
  computing element:\\
  \verb#(executable=$ROOT_DIR/myprog.exe)# -- \verb#$ROOT_DIR# is resolved
  locally (\textit{will cause errors if the path does not exist at the
  execution machine})\\
  \verb#(executable="$ROOT_DIR/myprog.exe")# -- \verb#$ROOT_DIR# will be
  resolved remotely\\

  For ARC $\leq$ 0.5.25, UI created a dummy executable attribute for internal GM use,
  with e.g., "\verb#/bin/echo#" as an argument (unless the
  specified executable itself started with "\verb#/#"). The actual
  executable was transferred by the UI to
  \texttt{arguments}, \texttt{inputFiles} and \texttt{executables} lists
  in the pre-processed xRSL script.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:arguments}\addcontentsline{toc}{subsection}{arguments}
  \begin{shaded}
    \xrsl{arguments}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(arguments=<string> [string] ... )#\\
    GM input:&\verb#(arguments=<executable> <string> [string] ... )#\\
    Example:&\verb#(arguments="10000" $(ATLAS)/input.dat)#\\ %$
  \end{tabular}

  List of the arguments for the executable.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string}&an argument\\
    \hspace*{1cm}&\texttt{executable}&the executable to be run by LRMS, taken by the
    UI from the user-specified \texttt{executable} attribute\\
  \end{tabular}

  For ARC $\leq$ 0.5.25, UI added the name specified in \texttt{executable} to the 
  \texttt{arguments} list in the pre-processed xRSL script.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:inputFiles}\addcontentsline{toc}{subsection}{inputFiles}
  \begin{shaded}
    \xrsl{inputFiles}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{12cm}}
    User input:&\verb#(inputFiles=(<filename> <location>) ... )#\\
    GM input:&\verb#(inputFiles=(<filename> <URL>)#\\
    &\verb#        (<filename> [size][.checksum]) ... )#\\
    Example:&\verb#(inputFiles=("local_to_job" "gsiftp://se1.lu.se/p1/remote_one")#\\
    &\verb#        ("local_to_job.dat" "/scratch/local_to_me.dat")#\\
    &\verb#        ("same_name_as_in_my_current_dir" ""))#\\
  \end{tabular}

  List of files to be copied to the computing element before the
  execution.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{filename} & file name, local to the computing element and
    always relative to the session directory\\
    \hspace*{1cm}&\texttt{location} & location of the file (\texttt{gsiftp},
    \texttt{https}, \texttt{ftp}, \texttt{http} URLs, or a path, local
    to the submission node). If void ("", use the quotes!), the input file is taken
    from the submission directory.\\
    \hspace*{1cm}&\texttt{URL} & URL of the file (see Section~\ref{sec:url})\\
    \hspace*{1cm}&\texttt{size} & file size in bytes\\
    \hspace*{1cm}&\texttt{checksum} & file checksum (as returned by \texttt{cksum})\\
  \end{tabular}

  If the list does not contain the standard input file (as specified
  by \texttt{stdin}) and/or the executable file (as specified by
  \texttt{executable} if the given name), the UI appends these files
  to the list. If the \verb#<location># is a URL, it is passed by the
  UI to the GM without changes; if it is a local path (or void, ""),
  the UI converts it to \verb#<size># and \verb#<checksum># and
  uploads those files to the execution cluster.

  \begin{framed}
  Please note that the \texttt{inputFiles} attribute is not meant to
  operate with directories: you must specify a pair
  \verb#("<local_to_job>" "[source]")# for each file.
  \end{framed}

  For ARC $\leq$ 0.5.25, UI added the name specified in \texttt{executable} to the 
  \texttt{inputFiles} list in the pre-processed xRSL script.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:executables}\addcontentsline{toc}{subsection}{executables}
  \begin{shaded}
    \xrsl{executables}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(executables=<string> [string] ...)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(executables="myscript.sh" "myjob.exe")#\\
  \end{tabular}

  List of files from the \texttt{inputFiles} set, which will be given
  executable permissions.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string}& file name, local to the computing element and
    relative to the session directory\\
  \end{tabular}

  If the executable file (as specified in \texttt{executable} and is
  relative to the session directory)  is not listed, it will be added
  to the list by the UI.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:cache}\addcontentsline{toc}{subsection}{cache}
  \begin{shaded}
    \xrsl{cache}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.34, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(cache="yes"|"no")#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(cache="yes")#\\
  \end{tabular}

  Specifies whether input files specified in the \texttt{inputFiles}
  should be placed by default in the cache or not. This does not
  affect files described by the \texttt{executables}, which will be
  placed in the session directory always.

  If not specified, default value is "yes".

  \begin{framed}
    Cached files can not be modified by jobs by default. If your job has to 
    modify input files, please specify \texttt{(cache="no")} or use
    \texttt{(readonly="no")} URL option for those files.
  \end{framed}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:outputFiles}\addcontentsline{toc}{subsection}{outputFiles} 
  \begin{shaded}
    \xrsl{outputFiles}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(outputFiles=(<string> <URL>) ... )#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(outputFiles=("local_to_job.dat" "gsiftp://se1.uo.no/stored.dat")#\\
    &\verb#    ("local_to_job.dat" "rc://rc1.uc.dk/set1/stored_and_indexed.dat")#\\  
    &\verb#    ("local_to_job_dir/" ""))#\\  
  \end{tabular}

  List of files to be retrieved by the user or uploaded by the GM and
  indexed (registered) in a data indexing service, e.g. \globus\  RLS or Fireman.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string}& file name, local to the
    \textit{Computing Element (CE)}\index{Computing Element}. If this
    string ends with a backslash "/" and \verb#<URL># is empty, the
    entire directory is being kept at the execution site. If however
    this string ends with a backslash "/" but the \verb#<URL>#
    starts with \texttt{gsiftp} or \texttt{ftp}, the whole directory
    is transfered to the destination.\\
    \hspace*{1cm}&\texttt{URL} & URL of the remote file (see
    Section~\ref{sec:url}); if void ("", use the quotes!), the file is
    kept for manual retrieval.  Note that this can not be a local
    \texttt{file://} URL.\\
  \end{tabular}

  Using a Replica Catalog pseudo-URL (see Section~\ref{sec:url}), you should make
  sure that the location is already defined in the Replica Catalog. If more
  than one such location is specified, only those found in the Replica Catalog
  for this collection will be used. Grid Manager will store output files in
  \textbf{one} location only: the first randomly picked location that
  exists. If no locations are specified, all those found in the
  Replica Catalog for this collection will be tried.

  If in a \verb#rc://# pseudo-URL the server component \texttt{host[:port]/DN} is not
  specified, server specified in the \texttt{replicaCollection} attribute will be used.

  If the list does not contain standard output, standard error
  file names and GM log-files directory name (as specified by \texttt{stdout},
  \texttt{stderr} and \texttt{gmlog}), the UI appends these
  items to the \texttt{outputFiles} list. If the \verb#<URL># is not specified (void,
  "", use the quotes!), files will be kept on SE and should be downloaded 
  by the user via the UI. If specified name of file ends with "/", the 
  entire directory is kept.

  A convenient way to keep the entire job directory at the remote site
  for a manual retrieval is to specify \texttt{(outputfiles=("/" ""))}.

  In some cases, the list of output files may only be known after the
  job has completed. ARC allows a user to specify a list of output
  files dynamically in a file or files in the session directory as
  part of their job. The file(s) containing the output file
  information can be defined in the xRSL script as the path to the
  file relative to the session directory preceeded by '@'. The format
  of these files is lines of 2 values separated by a space. The first
  value contains name of the output file relative to the session
  directory and the second value is a URL to which the file will be
  uploaded.

  \begin{tabular}{lp{13cm}}
    Example:&\verb#(outputFiles=("@output.files" "")#\\
    & \emph{output.files} is generated by the user and contains \\
    &\verb#   file1 gsiftp://grid.domain.org/file1#\\
    &\verb#   file2 gsiftp://grid.domain.org/file2#\\
  \end{tabular}

  After the job completes, the file output.files in the session
  directory will be read and any files described within will be
  uploaded to the given URLs.

  \begin{framed}
    Please note that the \texttt{outputFiles} attribute is not meant to
    operate with directories: you must specify a pair
    \verb#("<local_to_job>" "[destination]")# for each file. One
    exception is when you want to preserve an entire directory at the
    remote computer for later \textbf{manual} download via
    \texttt{ngget}. In that case, simply add the trailing backslash
    "/" at the end of the remote directory name. You can not upload a
    directory to a URL location, only to your local disk.
  \end{framed}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:cpuTime}\addcontentsline{toc}{subsection}{cpuTime}
  \begin{shaded}
    \xrsl{cpuTime}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.17\footnote{Was called \texttt{maxCpuTime} for ARC $\leq$ 0.3.16.}, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(cpuTime=<time>)#\\
    GM input:&\verb#(cpuTime=<tttt>)#\\
    Example:&\verb#(cpuTime="240")#\\
  \end{tabular}

  Maximal CPU time request for the job.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{time} & time (minutes)\\
    \hspace*{1cm}&\texttt{tttt} & time converted by the UI from \texttt{time} to 
    minutes (ARC $\leq$ 0.5.26) or seconds (ARC $\geq$ 0.5.27)\\
  \end{tabular}

  If only number is specified, the time is assumed to be minutes (ARC
  $\leq$ 0.5.26) or seconds (ARC $\geq$ 0.5.27). Otherwise, a free
  format is accepted, i.e., any of the following will be interpreted
  properly (make sure to enclose such strings in quotes!):

  \begin{tabular}{ll}
    \hspace*{1cm}&\texttt{"1 week"}\\
    \hspace*{1cm}&\texttt{"3 days"}\\
    \hspace*{1cm}&\texttt{"2 days, 12 hours"}\\
    \hspace*{1cm}&\texttt{"1 hour, 30 minutes"}\\
    \hspace*{1cm}&\texttt{"36 hours"}\\
    \hspace*{1cm}&\texttt{"9 days"}\\
    \hspace*{1cm}&\texttt{"240 minutes"}\\
  \end{tabular}

  If both \texttt{cpuTime} and \texttt{wallTime} are specified, the UI converts them both to 
  seconds. \texttt{cpuTime} can not be specified together with \texttt{gridTime}
  or \texttt{benchmarks}.

  \begin{framed}
    This attribute should be used to direct the job to a system with
    sufficient CPU resources, typically, a batch queue with the
    sufficient upper time limit. Jobs exceeding this maximum most
    likely will be \textbf{terminated} by remote systems! If
    time limits are not specified, the limit is not set and jobs
   can run as long as the system settings allow (note that in this
    case you can not avoid queues with too short time limits).
  \end{framed}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:wallTime}\addcontentsline{toc}{subsection}{wallTime}
  \begin{shaded}
    \xrsl{wallTime}
  \end{shaded}
  \versions{ARC $\geq$ 0.5.27, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(wallTime=<time>)#\\
    GM input:&\verb#(wallTime=<tttt>)#\\
    Example:&\verb#(wallTime="240")#\\
  \end{tabular}

  Maximal wall clock time request for the job.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{time} & time (minutes)\\
    \hspace*{1cm}&\texttt{tttt} & time converted by the UI from \texttt{time} to seconds\\
  \end{tabular}

  If only number is specified, the time is assumed to be
  seconds. Otherwise, a free format is accepted, i.e., any of the
  following will be interpreted properly (make sure to enclose such
  strings in quotes!):

  \begin{tabular}{ll}
    \hspace*{1cm}&\texttt{"1 week"}\\
    \hspace*{1cm}&\texttt{"3 days"}\\
    \hspace*{1cm}&\texttt{"2 days, 12 hours"}\\
    \hspace*{1cm}&\texttt{"1 hour, 30 minutes"}\\
    \hspace*{1cm}&\texttt{"36 hours"}\\
    \hspace*{1cm}&\texttt{"9 days"}\\
    \hspace*{1cm}&\texttt{"240 minutes"}\\
  \end{tabular}

  If both \texttt{cpuTime} and \texttt{wallTime} are specified, the UI converts them both to 
  seconds. \texttt{wallTime} can not be specified together with \texttt{gridTime}
  or \texttt{benchmarks}. If only \texttt{wallTime} is specified, but not \texttt{cpuTime},
  the corresponding  \texttt{cpuTime} value is evaluated by the UI and added to the job description.

  \begin{framed}
    This attribute should be used to direct the job to a system with
    sufficient CPU resources, typically, a batch queue with the
    sufficient upper time limit. Jobs exceeding this maximum most
    likely will be \textbf{terminated} by remote systems! If 
    time limits are not specified, the limit is not set and jobs
    can run as long as the system settings allow (note that in this
    case you can not avoid queues with too short time limits).
  \end{framed}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:gridTime}\addcontentsline{toc}{subsection}{gridTime}
  \begin{shaded}
    \xrsl{gridTime}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.31, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(gridTime=<time>)#\\
    GM input:& none\\
    Example:&\verb#(gridTime="2 h")#\\
  \end{tabular}

  Maximal CPU time request for the job scaled to the 2.8~GHz
  Intel\textsuperscript{\textregistered}
  Pentium\textsuperscript{\textregistered}~4 processor.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{time} & time (minutes)\\
  \end{tabular}

  The attribute is completely analogous to \texttt{cpuTime}, except
  that it will be recalculated to the actual CPU time request for each
  queue, depending on the published processor clock speed.

  \texttt{gridTime} can not be specified together with \texttt{cpuTime}
  or \texttt{wallTime}. If only \texttt{gridTime} is specified, but not \texttt{cpuTime},
  the corresponding  \texttt{cpuTime} value is evaluated by the UI and added to the job description.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:benchmarks}\addcontentsline{toc}{subsection}{benchmarks}
  \begin{shaded}
    \xrsl{benchmarks}
  \end{shaded}
  \versions{ARC $\geq$ 0.5.14, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(benchmarks=(<string> <value> <time>) ... )#\\
    GM input:&\\
    Example:&\verb#(benchmarks=("mybenchmark" "10" "1 hour, 30 minutes"))#\\
  \end{tabular}

  Evaluate a job's \texttt{cpuTime} based on benchmark values.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & benchmark name\\
    \hspace*{1cm}&\texttt{value} & benchmark value of reference machine\\
    \hspace*{1cm}&\texttt{time} & the \texttt{cpuTime} the job requires on the reference machine\\
  \end{tabular}

  \texttt{benchmarks} can not be specified together with \texttt{cpuTime}
  or \texttt{wallTime}. If only \texttt{benchmarks} is specified, but not \texttt{cpuTime},
  the corresponding  \texttt{cpuTime} value is evaluated by the UI and added to the job description.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:memory}\addcontentsline{toc}{subsection}{memory}
  \begin{shaded}
    \xrsl{memory}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.17\footnote{Was called \texttt{maxMemory} for ARC $\leq$ 0.3.16}, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(memory=<integer>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(memory>="500")#\\
  \end{tabular}

  Memory required for the job.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{integer} & size (Mbytes)\\
  \end{tabular}

  \begin{framed}
    Just like \texttt{cpuTime}, this attribute should be used to
    direct a job to a resource with a sufficient capacity. Jobs
    exceeding this memory limit will most likely be \textbf{terminated}
    by the remote system.
  \end{framed}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:disk}\addcontentsline{toc}{subsection}{disk}
  \begin{shaded}
    \xrsl{disk}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.17\footnote{Was called \texttt{maxDisk} for ARC $\leq$ 0.3.16}, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(disk=<integer>)#\\
    GM input:&none\\
    Example:&\verb#(disk="500")#\\
  \end{tabular}

  Disk space required for the job.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{integer}  & disk space, Mbytes\\
  \end{tabular}

  \begin{framed}
    This attribute is used at the job submission time to find a system
    with sufficient disk space. However, it \textbf{does not
    guarantee} that this space will be available at the end of the
    job, as most known systems do not allow for disk space
    allocation. Eventually, a remote system can terminate a job that
    exceeds the requested disk space.
  \end{framed}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:runTimeEnvironment}\addcontentsline{toc}{subsection}{runTimeEnvironment}
  \begin{shaded}
    \xrsl{runTimeEnvironment}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(runTimeEnvironment=<string>)(runTimeEnvironment=<string>)#\\
    GM input:&For ARC $\leq$ 0.5.39: \verb#(runTimeEnvironment="<string>" "<string>")#\\
    Example:&\verb#(runTimeEnvironment>="APPS/HEP/ATLAS-10.0.1")#\\
  \end{tabular}

  Required runtime environment.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & environment name\\
  \end{tabular}

  The site to submit the job to will be chosen by the UI among those
  advertising specified runtime environments. Before starting the job,
  the GM will set up environment variables and paths according to
  those requested. Runtime environment names are defined by Virtual 
  Organizations, and tend to be organized in name spaces.

  To request several environments, repeat the attribute string:\\
  \verb#(runTimeEnvironment="ENV1")(runTimeEnvironment="ENV2")# etc.   
  In ARC $\leq$ 0.5.39, before being submitted to the GM, a grouping was done by the UI, such that GM input was,
  e.g., \verb#(runTimeEnvironment="ENV1" "ENV2")#.

  To make a disjunct-request, use a boolean expression:\\
  \verb#(|(runTimeEnvironment="env1")(runTimeEnvironment="env2"))#.

  You can use "\verb#>=#" or "\verb#<=#" operators: job will be submitted to
  any suitable site that satisfies such requirements, and among the available at the sites
  runtime environments, the highest version satisfying a requirement will be requested
  in the pre-processed xRSL script.

  Runtime environment string interpretation is
  case-insensitive.  If a runtime environment string consists of a name
  and a version number, a partial specification is possible: it is sufficient
  to request only the name and use "\verb#>#" or "\verb#>=#" operators to select the highest version.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:middleware}\addcontentsline{toc}{subsection}{middleware}
  \begin{shaded}
    \xrsl{middleware}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(middleware=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(middleware="nordugrid-arc-0.5.99")#\\
  \end{tabular}

  Required middleware.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & Grid middleware name.\\
  \end{tabular}

  The site to submit the job to will be chosen by the UI among those
  advertising specified middleware. Usage is identical to that of the
  \texttt{runTimeEnvironment}. Use the "\verb#>=#" operator to
  request a version "equal or higher". Request
  \texttt{(middleware="nordugrid-arc")} defaults to
  \verb#(middleware>="nordugrid-arc-0.0.0.0")#.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:opsys}\addcontentsline{toc}{subsection}{opsys}
  \begin{shaded}
    \xrsl{opsys}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.21, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(opsys=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(opsys="FC3")#\\
  \end{tabular}

  Required operating system.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & Operating system name and version.\\
  \end{tabular}

  The site to submit the job to will be chosen by the UI among those
  advertising specified operating system. Usage is identical to that of
  \texttt{runTimeEnvironment} and \texttt{middleware}. Use the "\verb#>=#" operator to
  request a version "equal or higher".

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:stdin}\addcontentsline{toc}{subsection}{stdin}
  \begin{shaded}
    \xrsl{stdin}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(stdin=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(stdin="myinput.dat")#\\
  \end{tabular}

  The standard input file.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & file name, local to the computing element\\
  \end{tabular}

  The standard input file should be listed in the \texttt{inputFiles}
  attribute; otherwise it will be forced to that list by the UI.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:stdout}\addcontentsline{toc}{subsection}{stdout}
  \begin{shaded}
    \xrsl{stdout}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(stdout=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(stdout="myoutput.txt")#\\
  \end{tabular}

  The standard output file.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & file name, local to the computing element and
    relative to the session directory.\\
  \end{tabular}

  The standard output file should be listed in the
  \texttt{outputFiles} attribute; otherwise it will be forced to that
  list by the UI. If the standard output is not defined, UI assigns a
  name.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:stderr}\addcontentsline{toc}{subsection}{stderr}
  \begin{shaded}
    \xrsl{stderr}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(stderr=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(stderr="myjob.err")#\\
  \end{tabular}

  The standard error file.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & file name, local to the computing element and
    relative to the session directory.\\
  \end{tabular}

  The standard error file should be listed as an \texttt{outputFiles}
  attribute; otherwise it will be forced to that list by the UI. If
  the standard error is not defined, UI assigns a name. If \texttt{join} is 
  specified with value "yes", UI adds \texttt{stderr} to the 
  pre-processed xRSL script with the same value as \texttt{stdout}.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:join}\addcontentsline{toc}{subsection}{join}
  \begin{shaded}
    \xrsl{join}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(join="yes"|"no")#\\
    GM input:none\\
    Example:&\verb#(join="yes")#\\
  \end{tabular}

  If "yes", joins \texttt{stderr} and  \texttt{stdout} files into
  the \texttt{stdout} one. Default is \verb#no#.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:gmlog}\addcontentsline{toc}{subsection}{gmlog}
  \begin{shaded}
    \xrsl{gmlog}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(gmlog=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(gmlog="myjob.log")#\\
  \end{tabular}

  A name of the directory containing grid-specific diagnostics per job.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & a directory, local to the computing element and
    relative to the session directory\\
  \end{tabular}

%  The job log file should be listed as an \texttt{outputFiles} attribute;
%  otherwise it will be forced to that list by the UI.
	This directory is kept in the session directory to be
	available for retrieval (UI forces it to the list if
	\texttt{outputFiles})

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:jobName}\addcontentsline{toc}{subsection}{jobName}
  \begin{shaded}
    \xrsl{jobName}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(jobName=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(jobName="My Job nr. 1")#\\
  \end{tabular}

  User-specified job name.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & job name\\
  \end{tabular}

  This name is meant for convenience of the user. It can be used to
  select the job while using the UI. It is also available through the
  Information System.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:ftpThreads}\addcontentsline{toc}{subsection}{ftpThreads}
  \begin{shaded}
    \xrsl{ftpThreads}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(ftpThreads=<integer>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(ftpThreads="4")#\\
  \end{tabular}

  Defines how many parallel streams will be used by the GM during
  \texttt{gsiftp} and \texttt{http(s|g)}transfers of files.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{integer} & a number from 1 to 10\\
  \end{tabular}

  If not specified, parallelism is not used.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:acl}\addcontentsline{toc}{subsection}{acl}
  \begin{shaded}
    \xrsl{acl}
  \end{shaded}
  \versions{ARC $\geq$ 0.5.12, ARC 0.6}

  \textit{Available in ARC v0.5.12 and higher}.

  \begin{tabular}{lp{13cm}}
    User input:&\verb#(acl=<xml>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(acl="<?xml version=""1.0""?>#\\
    &\verb#    <gacl version=""0.0.1""><entry><any-user></any-user>#\\
    &\verb#    <allow><write/><read/><list/><admin/></allow></entry></gacl>")#\\
  \end{tabular}

  Makes use of GACL~\cite{gacl} rules to list users who are allowed to
  access and control job in addition to job's owner. Access and
  control levels are specified per user. \texttt{any-user} tag refers
  to any user authorized at the execution cluster. To get more
  information about GACL please refer to \url{http://www.gridsite.org}.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{xml} & a GACL-compliant XML string defining
    access control list\\
  \end{tabular}

  Following job control levels can be specified via \texttt{acl}:

  \begin{tabular}{llll}
    \hspace*{1cm}&\texttt{write}&--&\parbox[t]{11cm}{allows to modify contents of job data (job directory) and control job flow (cancel, clean, etc.)}\\
    \hspace*{1cm}&\texttt{read}&--&\parbox[t]{11cm}{allows to read content of job data
    (contents of job directory)}\\
    \hspace*{1cm}&\texttt{list}&--&\parbox[t]{11cm}{allows to list files available for
    the job (contents of job directory)}\\
    \hspace*{1cm}&\texttt{admin}&--&\parbox[t]{11cm}{allows to do everything --
    full equivalence to job ownership}\\
  \end{tabular}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:cluster}\addcontentsline{toc}{subsection}{cluster}
  \begin{shaded}
    \xrsl{cluster}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(cluster=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(cluster="cluster1.site2.org")#\\
  \end{tabular}

  The name of the execution cluster.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & known cluster name, or a substring of it\\
  \end{tabular}

  Use this attribute to explicitly force job submission to a cluster,
  or to avoid such. The job will not be submitted if the cluster does
  not satisfy other requirements of the job. Disjunct-requests of the
  kind \verb#(|(cluster="clus1")(cluster="clus2"))# are supported. To
  exclude a cluster, use \verb#(cluster!="clus3")#.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:queue}\addcontentsline{toc}{subsection}{queue}
  \begin{shaded}
    \xrsl{queue}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(queue=<string>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(queue="pclong")#\\
  \end{tabular}

  The name of the remote batch queue. Use only if you are sure that
  the queue by this name does exist.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & known queue name\\
  \end{tabular}

  Use this attribute to explicitly force job submission to a queue.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:startTime}\addcontentsline{toc}{subsection}{startTime}
  \begin{shaded}
    \xrsl{startTime}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(startTime=<time>)#\\
    GM input:&\verb#(startTime=<tttt>)#\\
    Example:&\verb#(startTime="2002-05-25 21:30")#\\
  \end{tabular}

  Time to start job processing by the Grid Manager, such as e.g. start downloading input files.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{time} & time string, YYYY-MM-DD hh:mm:ss\\ 
    \hspace*{1cm}&\texttt{tttt} & time string, YYYYMMDDhhmmss[Z] (converted by the UI from
    \texttt{time})\\
  \end{tabular}

  \begin{framed}
   Actual job processing on a worker node starts depending on local scheduling mechanisms, but not sooner than \texttt{startTime}. 
   \end{framed}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:lifeTime}\addcontentsline{toc}{subsection}{lifeTime}
  \begin{shaded}
    \xrsl{lifeTime}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(lifeTime=<time>)#\\
    GM input:&\verb#(lifeTime=<tttt>)#\\
    Example:&\verb#(lifeTime="60")#\\
  \end{tabular}

  Maximal time to keep job files (the session directory) on the
  gatekeeper upon job completion.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{time} & time (days)\\
    \hspace*{1cm}&\texttt{tttt} & time (seconds, converted by the UI from \texttt{time})\\
  \end{tabular}

  Typical life time is 1 day (24 hours). Specified life time can not
  exceed local settings.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:notify}\addcontentsline{toc}{subsection}{notify}
  \begin{shaded}
    \xrsl{notify}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(notify=<string> [string] ... )#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(notify="be your.name@your.domain.com")#\\
  \end{tabular}

  Request e-mail notifications on job status change.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & string of the format:
    \verb#[b][q][f][e][c] user1@domain1 [user2@domain2] ...#\\
    \hspace*{1cm}&&here flags indicating the job status are:\\
    \hspace*{1cm}&&\texttt{b} -- begin (PREPARING)\\
    \hspace*{1cm}&&\texttt{q} -- queued (INLRMS)\\
    \hspace*{1cm}&&\texttt{f} -- finalizing (FINISHING)\\
    \hspace*{1cm}&&\texttt{e} -- end (FINISHED)\\
    \hspace*{1cm}&&\texttt{c} -- cancellation (CANCELLED)\\
    \hspace*{1cm}&&\texttt{d} -- deleted (DELETED)\\
  \end{tabular}

  No more than 3 e-mail addresses per status change accepted.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:replicaCollection}\addcontentsline{toc}{subsection}{replicaCollection}
  \begin{shaded}
    \xrsl{replicaCollection}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(replicaCollection=<URL>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(replicaCollection="ldap://grid.uio.no:389/lc=TestCollection,#\\
    &\verb#          rc=NorduGrid,nordugrid,dc=org")#\\
  \end{tabular}

  Location of a logical collection in the Replica Catalog.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{URL} & LDAP directory specified as an URL (\verb#ldap://host[:port]/dn#)\\
  \end{tabular}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:rerun}\addcontentsline{toc}{subsection}{rerun}
  \begin{shaded}
    \xrsl{rerun}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(rerun=<integer>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(rerun="2")#\\
  \end{tabular}

  Number of reruns (if a system failure occurs).

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{integer}  & an integer number\\
  \end{tabular}

  If not specified, the default is 0. Default maximal allowed value is
  5. Before 0.5.29 jobs were rerun automatically and only in case of
  failure in INLRMS state. Since 0.5.29 job may be rerun after failure
  in any state for which reruning has sense. To initiate rerun user
  has to use \texttt{ngresume} command.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:architecture}\addcontentsline{toc}{subsection}{architecture}
  \begin{shaded}
    \xrsl{architecture}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(architecture=<string>)#\\
    GM input:&\\
    Example:&\verb#(architecture="i686")#\\
  \end{tabular}

  Request a specific architecture.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string} & architecture  (e.g., as produced by uname -a)\\
  \end{tabular}

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:nodeAccess}\addcontentsline{toc}{subsection}{nodeAccess}
  \begin{shaded}
    \xrsl{nodeAccess}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.24, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(nodeAccess="inbound"|"outbound")#\\
    GM input:&\\
    Example:&\verb#(nodeAccess="inbound")#\\
  \end{tabular}

  Request cluster nodes with inbound or outbound IP connectivity. If
  both are needed, a conjunct request should be specified.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:dryRun}\addcontentsline{toc}{subsection}{dryRun}
  \begin{shaded}
    \xrsl{dryRun}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(dryRun="yes"|"no")#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(dryRun="yes")#\\
  \end{tabular}

  If "yes", do dry-run: RSL is parsed, but no job submission to LRMS
  is made. Should be used for xRSL and communication validation.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:rsl\_substitution}\addcontentsline{toc}{subsection}{rsl\_substitution}
  \begin{shaded}
    \xrsl{rsl\_substitution}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(rsl_substitution=(<string1> <string2>))#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(rsl_substitution=("ATLAS" "/opt/atlas"))#\\
  \end{tabular}

  Substitutes \verb#<string2># with \verb#<string1># for
  \textbf{internal} RSL use.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{string1} & new internal RSL variable\\
    \hspace*{1cm}&\texttt{string2} & any string, e.g., existing combination of
    variables or a path\\
  \end{tabular}

  Use this attribute to define variables that simplify xRSL editing,
  e.g. when same path is used in several values, typically in
  \texttt{inputFiles}. Only one pair per substitution is allowed. To
  request several substitution, concatenate such requests. Bear in
  mind that substitution must be defined \textbf{prior} to actual use
  of a new variable \texttt{string1}.

  After the substitution is defined, it should be used in a way
  similar to shell variables in scripts: enclosed in round brackets,
  preceded with a dollar sign, \textbf{without quotes}:\\
  \verb#(inputfiles=("myfile" $(ATLAS)/data/somefile))#

  Unlike the \texttt{environment} attribute, \texttt{rsl\_substitution}
  definition is only used by the client and is valid inside xRSL
  script. It can not be used to define environment or shell variable
  at the execution site.

  \hspace*{0.5cm}\phantomsection
  \index{XRSL:attribute:environment}\addcontentsline{toc}{subsection}{environment}
  \begin{shaded}
    \xrsl{environment}
  \end{shaded}
  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(environment=(<VAR> <string>) [(<VAR> <string>)] ... )#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(environment=("ATLSRC" "/opt/atlas/src")#\\
    &\verb#         ("ALISRC" "/opt/alice/src"))#\\
  \end{tabular}

  Defines execution shell environment variables.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{VAR} & new variable name\\
    \hspace*{1cm}&\texttt{string} & any string, e.g., existing combination of
    variables or a path\\
  \end{tabular}

  Use this to define variables at an execution site. Unlike the
  \texttt{rsl\_substitution} attribute, it can not be used to define
  variables on the client side.

%%   \hspace*{0.5cm}\phantomsection
%%   \index{XRSL:attribute:jobtype}\addcontentsline{toc}{subsection}{jobtype}
%%   \begin{shaded}
%%     \xrsl{jobtype}
%%   \end{shaded}
%%   \versions{ARC $\geq$ 0.3.12, ARC 0.4}
%%   \begin{tabular}{lp{13cm}}
%%     User input:&\verb#(jobtype= "single"|"multiple"|"mpi"|"condor" )#\\
%%     Example:&\verb#(jobtype="mpi")#\\
%%   \end{tabular}

%%   Specifies a type of job (default is \texttt{single}).

  \hspace*{0.5cm}\phantomsection
  \addcontentsline{toc}{subsection}{count}\index{XRSL:attribute:count}
  \begin{shaded}
    \xrsl{count}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.12, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(count=<integer>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(count="4")#\\
  \end{tabular}

  Specifies amount of sub-jobs to be submitted for parallel tasks.

  \hspace*{0.5cm}\phantomsection
  \addcontentsline{toc}{subsection}{jobreport}\index{XRSL:attribute:jobreport}
  \begin{shaded}
    \xrsl{jobreport}
  \end{shaded}
  \versions{ARC $\geq$ 0.3.34, ARC 0.4, ARC 0.5, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(jobreport=<URL>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(jobreport="https://grid.uio.no:8001/logger")#\\
  \end{tabular}

  Specifies an URL for a logging service to send reports about job
  to. The default is set up in the cluster configuration.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{URL} & URL\\
  \end{tabular}

  It is up to a user to make sure the requested logging service
  accepts reports from the set of clusters she intends to use.

  \hspace*{0.5cm}\phantomsection
  \addcontentsline{toc}{subsection}{credentialserver}\index{XRSL:attribute:credentialserver}
  \begin{shaded}
    \xrsl{credentialserver}
  \end{shaded}
  \versions{ARC $\geq$ 0.5.45, ARC 0.6}
  \begin{tabular}{lp{13cm}}
    User input:&\verb#(credentialserver=<URL>)#\\
    GM input:&\verb#-"-#\\
    Example:&\verb#(credentialserver="myproxy://myproxy.nordugrid.org;username=user")#\\
  \end{tabular}

  Specifies an URL which Grid Manager may contact to renew/extend delegated
  proxy of job. Only MyProxy servers are supported.

  \begin{tabular}{llp{10cm}}
    \hspace*{1cm}&\texttt{URL} & URL of MyProxy server\\
  \end{tabular}

  It is up to a user to make sure the specified MyProxy server will accept
  requests from Grid Manager to renew expired credentials.
  \texttt{URL} may contain options \texttt{username} and \texttt{credname}
  to specify user name and credentials name which Grid Manager should pass
  to MyProxy server. If \texttt{username} is not specified DN of user 
  credentials is used instead.

### 2009-04-28 doc/tech_doc/xrsl/xrsl.tex
Reorganizing doc catalogue and enabling make doc

\documentclass{book}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=3cm,lmargin=2cm,rmargin=3cm}
\usepackage{makeidx}
\makeindex
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color,framed}
\usepackage{longtable}                             %for breaking tables

%%%%---- Use EPS figures and graphics with LaTeX ----
%\usepackage{graphics}
%%%%---- USE PNG or JPG figures and graphicx with PDFLaTeX ----
\usepackage{graphicx}
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref}
\hypersetup{
  pdfauthor = {Oxana Smirnova},
  pdftitle = {Extended Resource Specification Language},
  pdfsubject = {Reference manual},
  pdfkeywords = {Grid,NorduGrid,ARC,Globus,RSL},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}

\bibliographystyle{IEEEtran}

\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\xrsl}[1]{\textbf{\sffamily #1}}
\newcommand{\versions}[1]{\vspace{-0.3cm}(#1)\par}
\newcommand{\GT}{Globus Toolkit$\mathrm{^{TM}}$}
\newcommand{\globus}{Globus\textsuperscript{\textregistered}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
  \def\today{\number\day/\number\month/\number\year}

  \begin{titlepage}

    \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
    \end{tabular}

    \hrulefill

    {\raggedleft NORDUGRID-MANUAL-4\par}

    {\raggedleft \today\par}

    \vspace*{2cm}

%%%%---- The title ----

    \begin{center}

    \textsc{\Large Extended Resource Specification Language}
     \Large \par \textit{Reference Manual}

    \vspace*{2cm}

%    {\centering \large Refers to ARC release series 0.4 and up \large \par}

    \end{center}

    \vspace*{2cm}

%    \abstract{
%      The paper represents an overview and a reference manual for the
%      NorduGrid Extended RSL, developed on the basis of the \globus\ 
%      RSL.
%    }

  \end{titlepage}
  %\thispagestyle{empty} $ $
  %\newpage
  %$\ $
  \tableofcontents
  \newpage

  \chapter{Introduction}
  \label{sec:intro}

  Execution of jobs at batch clusters is the major part of the data
  analysis, production and other computational tasks and applications
  in many research areas. In the Grid infrastructure~\cite{grid}, such
  clusters are not necessarily local to a job submission site, but can
  be widely distributed. This implies extra requirements for the
  proper description of job options. The \globus\ \index{Globus}
  project~\cite{globus} developed a Grid middleware toolkit, which
  makes use of the \textit{Resource Specification Language
  (RSL)}~\cite{rsl}\index{RSL} to parse job options and definitions to
  resource management systems. The NorduGrid project~\cite{nordugrid}
  \index{NorduGrid} developed the Advanced Resource Connector (ARC),
  -- a 
  solution~\cite{para02,acat02} for a Grid facility, suitable for
  complex tasks, like, for example, a High Energy Physics
  production. To match the complexity of such tasks, this solution
  requires certain extensions to the RSL.

  To describe a task to be submitted to the ARC-enabled resources, an
  extended version of the \globus\  RSL is used. The extensions concern
  not only introduction of new attributes, but also differentiation
  between two levels of the job options specifications:

  \begin{itemize}
  \item[] \textbf{User-side RSL}, i.e., the set of attributes
    specified by a user in a job-specific file. This file is
    interpreted by the \textit{User Interface (UI)}~\cite{ui},
    \index{User Interface} and after the necessary modifications is
    parsed to the \textit{Grid Manager (GM)}~\cite{gm}\index{Grid
    Manager}
  \item[] \textbf{GM-side RSL}, i.e., the set of attributes
    pre-processed by the UI, and ready to be interpreted by the GM
  \end{itemize}

  A user only has to know the user-side part, and utilize it to
  describe the Grid tasks. The Grid Manager, however, uses slightly
  different notations, supplied by the User Interface.

  In what follows, the description of the NorduGrid-extended RSL,
  further denoted as \textbf{xRSL}\index{xRSL}, is given, using the
  following notations:

  \begin{tabular}{ll}
    \verb#<xxxx>#&parameter to be substituted with a corresponding string or a number\\
    \verb#[xxxx]# & optional parameter\\
    \verb#xxx|yyy|zzz# &list of possible values of a parameter\\
    \verb#-"-# &"same as above"\\
  \end{tabular}

  \chapter{xRSL syntax and rules}
  \label{sec:remarks}

  For a complete description of \globus\  RSL, see
  reference~\cite{rsl}. xRSL uses the same syntax conventions,
  although changes the meaning and interpretation of some attributes.

  \section{Syntax}
  \label{sec:syntax}

  A Grid task is described by the mean of xRSL attributes, which can
  be either passed via a command-line, or, more conveniently, be
  collected a so-called xRSL-file (suggested extension
  \textit{.xrsl}). Such a file contains a plain list of attribute-value pairs
  and boolean operands "\&" (for AND) and "\verb#|#" (for
  OR). Attribute names are case-insensitive.

  If the attributes are specified from the command line, the entire
  description must be enclosed either in single or in double quotes.
  Single quotes enclosure is completely analogous to xRSL-file usage,
  while double quotes imply standard shell expansion of the enclosed
  string. This has implications when strings contain local shell
  variables: they \textbf{will not be expanded} unless the task description is
  entered from the command line and is enclosed in double quotes.

  An attribute-value pair is a key element of specification. It
  consists of an expression that assigns one or more values to an
  attribute, and is enclosed in round brackets:

  \begin{shaded}
    \verb#(attribute="value")#\\
    \verb#(attribute="value1" "value2")# (for multi-valued attributes)\\
  \end{shaded}

  Certain attributes do not have assigned value; instead, they contain
  a list of values that should be put in proper correspondence with
  each other:

  \begin{shaded}
    \verb#(attribute=("value1" "value2")("value3" "value4"))#
  \end{shaded}

  In the example above, \verb#value1# and \verb#value3# are put in correspondence to
  \verb#value2# and \verb#value4# respectively, according to the
  context of the attribute.

  Values should be enclosed in quotes if they contain blank
  spaces or special characters. The \textbf{special characters} are:
  \begin{shaded}    
    \verb. +   &   |   (   )   =   <   >   !   "   '   ^   #   $.%$
  \end{shaded}

  To quote a string containing special characters, you can use 
  either single or double quotes. If your string, however, contains
  both such characters, you can define any character as an own
  delimiter, by preceding it with the "carat" (\verb#^#) character:
  \verb#attribute=^*My "good" value^*# makes use of a carat-escaped
  asterisk as a delimiter.

  Typically, an xRSL job description starts with an ampersand
  ("\&") , to indicate implicit \textbf{conjunction} of all the
  attributes:
  \begin{shaded}
    \verb#&(attribute1=value1)(attribute2="value 2")...#
  \end{shaded}

  Whenever a \textbf{disjunct}-request of two or more attributes is
  needed, the following construction can be used:
  \begin{shaded}
    \verb#(|(attribute="value1")(attribute="value2")...)#
  \end{shaded}

  In expressions, the following operands are allowed:
  \begin{shaded}
    \verb#=   !=   >   <   >=   <=#
  \end{shaded}

  \textbf{Commented} lines should start with "(*" and be
  closed with "*)":
  \begin{shaded}
    \verb#(*attribute="value1"*)#
  \end{shaded}
  Comments can not be nested.

  \textbf{Multiple job} description in one file is realized via a standard
  \globus\  RSL multi-request operand "+", which should precede
  multiple job description: 
  \begin{shaded}
    \verb#+(&(...))(&(...))(&(...))#
  \end{shaded}

  The xRSL attributes can be written in a single string, or split in
  lines arbitrary; blank spaces between and inside
  \verb#(attribute="value")# relations are ignored.

  \section{URLs}
  \label{sec:url}

  \input{xrsl-urls}

  \chapter{Attributes}
  \label{sec:attributes}

  Most of the attributes \index{attributes} introduced originally by \globus\ RSL~1.0 are
  supported, some with modifications as indicated in this
  document. Many new attributes are introduced, of which some are
  to be specified in the user's script, and others are internal for
  the GM (are added or modified by the UI).

  Attribute names are case-insensitive, although assigned values may
  well be case-sensitive, if they represent file names, environment
  variables etc..

  \section{User-side attributes}
  \label{sec:user}

  The following attributes \index{attributes:user-side} can be
  specified in a user's xRSL script. Some are to be modified by the UI
  before being passed to the GM. If this is the case, corresponding GM input is described.

  \input{xrsl-user}

  \section{GM-side attributes}
  \label{sec:gm}

  The following attributes are constructed by the UI and are parsed to
  the GM, although users may specify them manually (\textbf{not
  advised!}).

  \input{xrsl-gm}

  \section{Unsupported Globus RSL attributes}
  \label{sec:globus}

  The following \globus\ attributes are not supported by the ARC
  middleware. Whenever they are specified, a warning is issued by the
  UI, and the corresponding attribute is ignored.

  \subsection{Unsupported RSL 1.0 attributes}

  \begin{itemize}
    \item \verb#(resourceManagerContact=<string>)#
    \item \verb#(directory=<string>)#
    \item \verb#(maxCpuTime=<time>)#
    \item \verb#(maxWallTime=<time>)#
    \item \verb#(maxTime=<time>)#
    \item \verb#(maxMemory=<memory>)#
    \item \verb#(minMemory=<memory>)#
    \item \verb#(gramMyJob=independent|collective)#
    \item \verb#(project=<string>)#
    \item \verb#(hostCount=<number>)#
    \item \verb#(label=<string>)#
    \item \verb#(subjobCommsType=blocking-join|independent)#
    \item \verb#(subjobStartType=strict-barrier|loose-barrier|no-barrier)#
  \end{itemize}

  \subsection{Unsupported GRAM RSL attributes}

  \begin{itemize}
    \item \verb#(directory=<string>)#
    \item \verb#(fileCleanUp=<array>)#
    \item \verb#(fileStageIn=<array>)#
    \item \verb#(fileStageInShared=<array>)#
    \item \verb#(fileStageOut=<array>)#
    \item \verb#(gassCache=<path>)#
    \item \verb#(gramMyJob=independent|collective)#
    \item \verb#(hostCount=<number>)#
    \item \verb#(jobType=<string>)#
    \item \verb#(libraryPath=<path>)#
    \item \verb#(maxCpuTime=<time>)#
    \item \verb#(maxWallTime=<time>)#
    \item \verb#(maxTime=<time>)#
    \item \verb#(maxMemory=<memory>)#
    \item \verb#(minMemory=<memory>)#
    \item \verb#(project=<string>)#
    \item \verb#(remoteIoUrl=<string>)#
    \item \verb#(scratchDir=<string>)#
  \end{itemize}

  \chapter{XRSL and JSDL}
  \label{sec:jsdl}

  \input{xrsl-jsdl}

  \appendix

  \chapter{Examples}
  \label{sec:examples}

  \section{User-side xRSL script}

  \begin{verbatim}
&
(* test run: if "yes", only submits RSL without actual job start *)
     (dryRun="no")
(* some local variables defined for further convenience *)
     (rsl_substitution=("TOPDIR" "/home/johndoe")) 
     (rsl_substitution=("NGTEST" $(TOPDIR)/ngtest))
     (rsl_substitution=("BIGFILE" "/scratch/johndoe/100mb.tmp"))
(* some environment variables, to be used by the job *)
     (environment=("ATLAS" "/opt/atlas") ("CERN" "/cern"))
(* the main executable file to be staged in and submitted to the PBS *)
     (executable="checkall.sh")
(* the arguments for the executable above *)
     (arguments="pal")
(* files to be staged in before the execution *)
     (inputFiles = ("be_kaons" "")
     ("file1" gsiftp://grid.uio.no$(TOPDIR)/remfile.txt) 
      ("bigfile.dat" $(BIGFILE) )    )
(* files to be given executable permissions after staging in *)
     (executables="be_kaons")
(* files to be staged out after the execution *)
     (outputFiles=
        ("file1" "gsiftp://grid.tsl.uu.se/tmp/file1.tmp")
        ("100mb.tmp" "rls://rls.nordugrid.org:39281/test/bigfile")
        ("be_kaons.hbook"  gsiftp://ce1.grid.org$(NGTEST)/kaons.hbook)   )
(* user-specified job name *)
     (jobName="NGtest")
(* standard input file *)
     (stdin="myinput.dat")
(* standard output file *)
     (stdout="myoutput.dat")
(* standard error file *)
     (stderr="myerror.dat")
(* GM logs directory name *)
     (gmlog="gmlog")
(* flag whether to merge stdout and stderr *)
     (join="no")
(* request e-mail notification on status change *)
     (notify="bqfe john.doe@gmail.com jane.doe@mail.org")
(* specific queue to submit the job *)
     (queue="atlas")
(* maximal CPU time required for the job, minutes for PBS*)
     (CpuTime="60")
(* maximal time for the session directory to exist on the remote node, days *)
     (lifeTime="7")
 (* memory required for the job, Mbytes *)
     (Memory="200")
(* wall time to start job processing *)
     (startTime="2002-04-28 17:15:00")
(* disk space required for the job, Mbytes *)
     (Disk="500")
(* required architecture of the execution node *)
     (architecture="i686")
(* required run-time environment *)
     (runTimeEnvironment="Atlas-1.1")
(* number of re-runs, in case of a system failure *)
     (rerun="2")
  \end{verbatim}

  \section{GM-side xRSL script}

Note that the client tool expands all the \texttt{rsl\_substitution}
values and adds double quotes to \textbf{all} strings. It also
streamlines blank spaces.

Comment lines are removed from the GM-side xRSL; below they are shown
to explain details.

  \begin{verbatim}
&
(* saves RSL in a temporary file if "yes" *)
     ("savestate" = "yes" )
(* job submission to be performed if action is "request" *)
     ("action" = "request" )
(* submission host name *)
     ("hostname" = "ce1.grid.org" )
(* client software version *)
     ("clientsoftware" = "nordugrid-arc-0.6.0.3" )
(* walltime value added by the client, in seconds *)
     ("walltime" = "3600" )
(* test run: if "yes", only submits RSL without actual job start *)
     ("dryRun" = "no" )
(* some local variables defined for further convenience *)
     ("rsl_substitution" = ("TOPDIR" "/home/johndoe" ) ) 
     ("rsl_substitution" = ("NGTEST" "/home/johndoe/ngtest" ) )
     ("rsl_substitution" = ("BIGFILE" "/scratch/johndoe/100mb.tmp" ) )
(* some environment variables, to be used by the job *)
     ("environment" = ("ATLAS" "/opt/atlas" ) ("CERN" "/cern" ) )
(* executable *)
     ("executable" = "checkall.sh" )
(* arguments *)
     ("arguments" = "pal" )
(* files to be staged in before the execution *)
     ("inputfiles" = ("checkall.sh" "279320" )
        ("myinput.dat" "39806" )
        ("be_kaons" "8807" )
        ("file1" "gsiftp://grid.uio.no/home/johndoe/remfile.txt" ) 
        ("bigfile.dat" "104857600" )
     )
(* files to be given executable permissions after staging in *)
     ("executables" = "checkall.sh" "be_kaons" )
(* files to be staged out after the execution *)
     ("outputfiles" = ("file1" "gsiftp://grid.tsl.uu.se/tmp/file1.tmp" )
        ("100mb.tmp" "rls://rls.nordugrid.org:39281/test/bigfile" )
        ("be_kaons.hbook" "gsiftp://ce1.grid.org/home/johndoe/ngtest/kaons.hbook" )
        ("myoutput.dat" "" )
        ("myerror.dat" "" )
    )
(* user-specified job name *)
     ("jobName" = "NGtest" )
(* standard input file *)
     ("stdin" = "myinput.dat" )
(* standard output file *)
     ("stdout"="myoutput.dat" )
(* standard error file *)
     ("stderr" = "myerror.dat" )
(* flag whether to merge stdout and stderr *)
     ("join" = "no" )
(* request e-mail notification on status change *)
     ("notify" = "bqfe john.doe@gmail.com jane.doe@mail.org" )
(* specific queue to submit the job *)
     ("queue" = "atlas" )
(* CPU time required for the job, converted into seconds *)
     ("cputime" = "3600" )
(* maximal time for the session directory to exist on the remote node, seconds *)
     ("lifetime" = "604800" )
(* memory required for the job, Mbytes *)
     ("memory" = "200" )
(* wall time to start job processing *)
     ("startTime" = "20020128171500" )
(* disk space required for the job, Mbytes *)
     ("disk" = "500" )
(* required architecture of the execution node *)
     ("architecture" = "i686" )
(* required run-time environment *)
     ("runtimeenvironment" = "Atlas-1.1" )
(* number of re-runs, in case of a system failure *)
     ("rerun" = "2" )
(* original client xRSL with expanded string substitutions; shortened here *)
     ("clientxrsl" = "&(""dryrun"" = ""no"" )(... )(""rerun"" = ""2"" )" )
  \end{verbatim}

  \section{JSDL script}

  \begin{verbatim}

<?xml version="1.0" encoding="UTF-8"?>
<jsdl:JobDefinition
    xmlns:jsdl="http://schemas.ggf.org/jsdl/2005/11/jsdl"
    xmlns:jsdl-posix="http://schemas.ggf.org/jsdl/2005/11/jsdl-posix"
    xmlns:jsdl-arc="http://www.nordugrid.org/ws/schemas/jsdl-arc">
 <jsdl:JobDescription>
    <!-- Generic job description (jobname in XRSL) -->
    <jsdl:JobIdentification>
        <jsdl:JobName>JSDL Hello World job</jsdl:JobName>
        <jsdl:Description>
          This is a simple Hello World job test
        </jsdl:Description>
    </jsdl:JobIdentification>
    <!-- Standard application (same as executable, arguments in XRSL) -->
    <jsdl:Application>
        <jsdl:ApplicationName>Echo</jsdl:ApplicationName>
        <jsdl-posix:POSIXApplication>
          <jsdl-posix:Executable>/bin/echo</jsdl-posix:Executable>
          <jsdl-posix:Argument>Hello world</jsdl-posix:Argument>
        </jsdl-posix:POSIXApplication>
    </jsdl:Application>
    <!-- Get the GM diagnostics (same as gmlog in XRSL) -->
    <jsdl-arc:LocalLogging>
	      <jsdl-arc:Directory>gmlog</jsdl-arc:Directory>
    </jsdl-arc:LocalLogging>
 </jsdl:JobDescription>
</jsdl:JobDefinition>

  \end{verbatim}

  \chapter{JSDL-ARC schema}
  \label{sec:jsdl-arc-schema}

  \input{jsdl-arc-schema}

  \bibliography{grid}
  \printindex

\end{document}

### 2009-04-28 doc/tutorials/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Reorganizing doc catalogue and enabling make doc

--- 2009-03-22 doc/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
@@
 
+The section starts with another roundup on the WSDL for the Echo service,
+continues with the implementation of the server and finally describes the
+client.
+
+\section{Description with WSDL}
+
 The WSDL file that specifies the here presented Echo Web Service is shown in
 Listing~\ref{lst:echo_wsdl}.
+\lstsetJUSTXML
+\lstinputlisting
+	[
+	label=lst:echo_wsdl,
+	caption={[WSDL file describing the echo service. Filename: echo.wsdl]
+	\textbf{WSDL file describing the echo service. Filename: echo.wsdl}}
+	]
+{../src/services/echoservice/echo.wsdl}
+\task{TODO: Validate the WSDL Listing}
 
 % moved WSDL description into separate Chapter. See folder tex_wsdl.
+\section{Implementation of the Server}
 
-\subsection{The SOAP Fault Message}
+This section presents the implementation of the service defined in the WSDL
+file of the previous section and shown in
+Listing~\ref{lst:echo_service_cpp}. For the user of a web service, it is not
+important to know the programming language that this was implemented in. This
+tutorial uses the C++ libraries of HED. Though ARC also provides everything
+that one would need to implement the service with a scripting language like
+Python. A later version of this tutorial hopefully describes both APIs side by
+side.
+
+\subsection{Communication with SOAP: The SOAP Fault Message}
 
@@
 %    * Taxonomien/Ontologien zur semantischen Einordnung des Services
-\clearpage
 
-\section{Service}
+\subsection{The Service in C++}
 
-This section presents the implementation of the service defined in the WSDL file of the previous section.
-The source code is shown in Listing~\ref{lst:echo_service_cpp}.
 The header file will be set aside since it contains only redundant information. 
@@
 
-\section{Client}
+\section{Implementation of the Client}
 
@@
 The checks of the response have expanded by an additional condition which tests for a fault message, see line~\ref{lst_code:echo_client_cpp_fault}.
+
+\subsection{The Client in C++}
 

### 2009-04-28 doc/tutorials/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Reorganizing doc catalogue and enabling make doc

--- 2009-03-05 doc/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
@@
 When the service receives a request of a client, it will respond with a string containing the system time.
-For simplicity the service will not parse the request but always return the current time.
+For simplicity, the service will not parse the request but always return the current time.
 It is therefore easier than the example of an Echo Web Service which additionally has to extract the incoming message in order to get the string which shall be echoed.
@@
 
-The source code of the examples given in this tutorial are available in the subdirectory \textit{src}\footnote{If this was not shipped with the document you are reading, then please inspect\\ http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/ws-programming-tutorial}.
+The source code of the examples given in this tutorial are available in the directory \textit{src}\footnote{If this was not shipped with the document you are reading, then please inspect\\ http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/ws-programming-tutorial.} that accompanies the electronic version of this document.
 
@@
 
-In ARC the services are implementated as plugins such that they can easily be included or excluded by modifiying the server configuration file as it was seen in Listing~\ref{lst:arched_arcecho_xml}.
-The class which is needs to be implemented for the Web Service inherits from the class \textit{Service} which itself is defined in the ARC library.
-The proper class definition of the object \textit{TimeService} is defined in the corresponding header file which is shown in Listing~\ref{lst:time_arched_h}.
+In ARC, the services are implementated as plugins such that they can easily be included or excluded by modifiying the server configuration file as it was seen in Listing~\ref{lst:arched_arcecho_xml}.
+The class needed to be implemented for the Web Service inherits from the class \textit{Service}, which itself is defined in the ARC library.
+The proper class definition of the object \textit{TimeService} is defined in the corresponding header as shown in Listing~\ref{lst:time_arched_h}.
 \lstsetCPP%
@@
 Within the line~\ref{lst_code:time_cpp_process_message_payload} the payload is stored in the message. 
-When nothing unexpected happened while the message is processed, the function returns an \textit{MCCStatus} which was instanciated with the state \textit{OK}, as to be seen in line~\ref{lst_code:time_cpp_return}.\\
+When nothing unexpected happened while the message was processed, the function returns an \textit{MCCStatus} which was instanciated with the state \textit{OK}, as to be seen in line~\ref{lst_code:time_cpp_return}.\\
 
@@
 
-To run the service, it needs to be compiled into a dynamic library. Here this is named \textit{libtimeservice.so}, which is arbitrary but needs to be in sync with the configuration file of the HED.
+To run the service, it needs to be compiled into a dynamic library. Here this is named \textit{libtimeservice.so}. The name is arbitrary but needs to be in sync with the configuration file of the HED.
 For the installation, the library needs to be copied to a directory that is accessible for the HED.
@@
 
-A suitable server configuration file which engages the time service is to be seen in Listing~\ref{lst:time_arched_xml}.
+A suitable server configuration file which engages the time service is shown in Listing~\ref{lst:time_arched_xml}.
 A new \textit{Path} element has been introduced into the element \textit{ModuleMananger} in line~\ref{lst_code:arched_time_moduleManager}.
@@
 
-Up to now we have implemented a server program, the code that runs on a particular machine and waits for other programs to ask it for the time.
-What is still needed is the client to send the request to the server and interprete the response of it.
+Up to now we have implemented a server program which is running on a particular machine and waiting for other programs to ask for the time.
+What is still needed is the client to send the request to the server and interprete the response.
 %If the server spoke the http protocol, then we could use our web browser.
@@
 
-The source code of the client is to be seen in Listing~\ref{lst:time_client_cpp} and consists basically of one main function.
+The source code of the client can be inspected in Listing~\ref{lst:time_client_cpp} and consists basically of one main function.
 Within the line~\ref{lst_code:time_client_cpp_logger} the object \textit{Logger} is initialized.
-Due to the configuration of the the \textit{Logger} in the next lines, the messages passed to it will be redirected to the standard error stream.
+Due to the configuration of the the \textit{Logger} in the following lines, the messages passed to it will be redirected to the standard error stream.
 For the components of the protocol stack, which  are likewise loaded dynamically, one needs to specify the installation directory of ARC in line~\ref{lst_code:time_client_cpp_init}.\\
@@
 This is done in line~\ref{lst_code:time_client_cpp_message_reqeust}.
-Even the appearance of the message is unimportant to the service, the message is composed of two convoluted elements \textit{time} and \textit{timeReqest}. 
+Even the appearance of the request is unimportant to the service, the message is composed of two convoluted elements \textit{time} and \textit{timeReqest}. 
 The request will be processed in the next line by the method \textit{process} of the object \textit{client}. %in line~\ref{lst_code:time_client_cpp_message_response}. 
@@
 The client program has to be compilied to an ordinary binary such that it can directly be executed from the command line.
-Assure that the service is running and use the client like in Listing~\ref{lst:time_client_ksh}.
+Assure that the service is running and use the client as demonstrated in Listing~\ref{lst:time_client_ksh}.
 \lstsetKSH
@@
 \end{lstlisting}
-the transmitted messages are shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
+The transmitted messages are shown in the Listings~\ref{lst:time_client_request} and \ref{lst:time_client_response}.
 The enclosing element is \textit{Envelope} which may contain an element \textit{Header} and must contain an element \textit{Body}.

### 2009-04-28 doc/sec/papers/GRID09/paper.tex
Redo some performance testing; change the text

@@
 Provider (SP) is responsible for controlling access to services and resources by analyzing 
-the SAML assertions produced and issued by the IdP upon request.
+the SAML assertions produced and issued by the IdP upon request (request from client application 
+to the service/resource that is protected by SP).
 
@@
 the SAML assertion into session/connection context. That assertion includes the AuthnStatement and 
-AttributeStatement elements. From step 7 to step 9, the WS client launches the SOAP invocation via 
+AttributeStatement elements. From step 8 to step 10, the WS client launches the SOAP invocation via 
 the same connection as the one which is used by user agent to contact SP Service, then Grid/Web Service 
 checks the AuthnStatement stored in the session context to see if the AuthnStatement is valid or expired. 
-If valid, service handles the SOAP request and returns the SOAP response to the WS client.
+If valid, service handles the SOAP request and returns the SOAP response to the WS client. Note
+the ``connection'' and ``session'' are interchangeably used in this paper, with the same meaning for 
+TCP connection.
 
@@
 invoke other services, a X.509 credential delegation service is implemented, which is based on 
-SOAP communication rather than the proprietary GSI communication implemented in Globus Toolkit.
+SOAP communication rather than the proprietary GSI communication which is used in the delegation
+services such as MyProxy.
 
@@
 \label{sec:slcs}
-The sequence diagram of SLCS process specializes the diagram shown in Figure \ref{fig:SAML2SSOUML}, 
-with the SOAP request including X.509 request and SOAP response including X.509 response. 
+The sequence diagram of SLCS process is the same as the diagram shown in 
+Figure \ref{fig:SAML2SSOUML}, with SOAP request (the message of step 8 of 
+Figure \ref{fig:SAML2SSOUML}) including X.509 request 
+and SOAP response (the message of step 10 of Figure \ref{fig:SAML2SSOUML}) 
+including an signed X.509 certificate and the CA certificate that is used to 
+sign this X.509 certificate. 
 In detail, the SLCS client firstly accomplishes the SAML2SSO profile, then sends 
@@
 How to compose a distinguished name (DN) for the certificate is a critical issue for 
-the SLCS service. We pick up the relatively identical information ``eduPersonPrincipalName'' 
-(see eduPerson schema~\cite{eduSchemalink}) from SAML attributes that returned from Shibboleth
-IdP, and use it to compose the DN. A typical example of the eduPersonPrincipalName 
+the SLCS service. Since the ``eduPersonPrincipalName'' (see eduPerson schema~\cite{eduSchemalink})
+is identical to the user that accomplishes SAML2SSO profile. We pick up ``eduPersonPrincipalName'' 
+from SAML attributes that returned from Shibboleth IdP, and use it to compose the DN. A 
+typical example of the eduPersonPrincipalName 
 could be alice@example.org. In such case composed DN is for example\\``/O=knowarc/OU=example.org/CN=alice''.
@@
 Then this client invokes the peer Web Service by appending the delegation information (URL of delegation
-service and the delegation ID corresponding to the above delegation) into the SOAP body as out-of-band 
-message. The Web Service then sends a X.509 request and the delegation ID to the delegation service 
+service and the delegation ID corresponding to the above delegation) into the SOAP body. The Web 
+Service then sends a X.509 request and the delegation ID to the delegation service 
 which is specified by client side, and gets back a delegated certificate which can be composed together
@@
 
-For the service side, we implement the simple ``echoString'' service in HED, and use the ``echoString''
-service in Axis2/C. For client side, we develope a client by using API of ARC, to invoke the ``echoString''
+For the service side, we implement two simple ``echoString'' services in HED and Axis2/C. The 
+``echoString'' provides a simple SOAP portType which will output the same value as it gets from
+input.
+For client side, we develope a client by using the client API of ARC, to invoke the ``echoString''
 services from Axis2/C and HED. The service invocation is just sending a short string to service, and 
@@
 CPU and 2GB memory, Red Hat Enterprise Linux 5. The client runs in another machine equipped 
-with 1.8GHz Intel Pentium CPU and 2GB memory, Ubuntu 8.10. The Axis2/C service is run using 
+with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat Enterprise Linux 5. The Axis2/C service is run using 
 the httpd module on Apache Http Server version 2.2.11 instance. We keep all the configurations 
-default in Apache Http Server. The client and service are configured with mutual TLS authentication.
+default in Apache Http Server.
 
@@
 \begin{figure}
-\includegraphics[width=0.9\columnwidth]{HED2Axis_thread40.pdf}
-\caption{Performance results for HED and Axis2C: 1 service and 40 clients}
-\label{fig:HED2Axis_thread40}
+\includegraphics[width=0.9\columnwidth]{TCP_thread_all.pdf}
+\caption{Performance of HED and Axis2C echoString services without TLS: performance
+changes with client numbers; session non-reused }
+\label{fig:TCP_thread_all}
 \end{figure}
 
-Figure \ref{fig:HED2Axis_thread10} and Figure \ref{fig:HED2Axis_thread40} are the results of 
-performance comparison. 
+\begin{figure}
+\includegraphics[width=0.9\columnwidth]{TLS_thread_all.pdf}
+\caption{Performance of HED and Axis2C echoString services with TLS: performance
+changes with client numbers; session non-reused}
+\label{fig:TLS_thread_all}
+\end{figure}
+
+Figure \ref{fig:HED2Axis_thread10} is the results of performance comparison with ten 
+clients being launched.
 Two types of experiments are shown: session reused, session non-reused.
@@
 For ``Session reused'' experiments, we see that the average response time in HED is about 81ms
-which is twice of the value in Axis2/C, and the throughput in HED is about half of the value in Axis2/C, 
-when the client number is 10. When the client number is increased to 40, the average response 
-time in HED is close to the value in Axis2/C when the duration is more than 60 seconds, and the 
-similar result appears for throughput. One obersavation is that, the performance results of HED is 
-quite stable, with the average response time almost kept unchanged under different time points, 
-as well as under different number of clients, and throughput kept unchanged under different 
-time points, while increased linearly with number of clients.
+which is twice of the value in Axis2/C, and the throughput in HED is about half of the value in Axis2/C. 
+One obersavation is that, the performance results of HED is quite stable, with the average response time 
+and throughput almost kept unchanged under different time points.
+For ``Session non-reused'' experiments, the average response time in HED is almost the same as
+the value in Axis2/C, with the value 15ms. On the other hand, the throughput in HED is close to
+ the value in Axis2/C.
 
-For ``Session non-reused'' experiments, the average response time in HED is almost the same as
-the value in Axis2/C, in both cases of client number is 10 (the value is 15ms) and 40 (the value
- is 53ms) correspondingly. On the other hand the throughput in HED is also the same as the value 
-in Axis2/C, in both cases of client number is 10 and 40. One interesting obersabation is that the 
-throughput is close to a fixed value (around 500 request/s) for both HED and Axis2/C in both 
-cases of client number, when the duration value becomes bigger. Also, in case of client number is
-40, the throughput from ``Session reused'' experiments also is close to the same fixed value.
-We speculate that, with a relatively big value of client number, no matter the session is reused or
-not, the throughput is becoming a fixed value which can be explained as the SOAP processing capability
-of the SOAP implementation. In this case, we can believe that HED has the similar performance 
-capability as Axis2/C.
+Since the echoString service is a simple service which occupies little and stable time for processing
+service logic itself, and also the experiments are done under local network, we can speculate 
+the response time we get in Figure \ref{fig:HED2Axis_thread10} is mainly about the processing time 
+in container itself.
+With the stability characteristic of performance, we can speculate that the service container 
+itself will stablly impact on the performance of other services.
+
+To more deeply compare the performance of services from HED and Axis2/C, some tests have
+been done to investigate how does the number of simultaneous clients impact on the performance.
+Figure \ref{fig:TCP_thread_all} is the performance results for echoString services with client number
+being changed. In this set of experiments, the session is not reused, and secure communication is not
+used. Each experiment is executed for a relatively long duration (two minutes), so that the there are
+enough times of service invokations and then the performance results are reliable.
+We see that both the response time of services in HED and Axis2/C starts from similar value 
+(when client number is five) and changes linearly with number of clients, while service in HED 
+provides shorter response time with the value being around two thirds of the response time of 
+service in Axis2/C when client number becomes bigger.
+In case of throughput, service in HED provides stable throughtput, while service in Axis2/C provides 
+throughput which slightly becomes smaller when client number becomes bigger.
+
+Figure \ref{fig:TLS_thread_all} is the performance results when secure communication is used.
+For Axis2/C, we configure the http server (that hosts Axis2/C) with SSL module enabled.
+Each experiment is executed for a duration of two minutes.
+We see that, similar to Figure \ref{fig:TCP_thread_all}, the both the response time of services 
+in HED and Axis2/C also starts from similar value (when client number is five) and changes linearly 
+with number of clients, while the service in HED also provides much shorter response time 
+with the value being around sightly more than one third of the response time of service in Axis2/C 
+when client number becomes bigger.
+In case of throughput, the throughput of service in HED increases linearly with client numbers intially,
+however, it starts to keep stable to be a fixed value (around 13000 requests per minutes) when client 
+number is bigger than 30. On the other hand, the throughput of service in Axis2/C starts to keep 
+stable when client number is bigger than 10, with the value being around sightly more than one 
+third of the throughput of service in HED.
+
+We conclude that the HED can provide better performance than Axis2/C, in both cases of secure 
+communication being enabled and disabled, especially when the number of simultaneous clients 
+is bigger.
 
@@
 Service Provider) service, and develope a client by using SAML2SSO related API of ARC, to invoke 
-the ``echoString'', as well as interact with the IdP (Identity Provider). Services and client run on
+the ``echoString'', as well as interact with the IdP (Identity Provider). Service and clients run on
 the same set of machine as section \ref{sec:perhedandaxis}. Meanwhile, we deploy the Shibboleth IdP 
@@
 \begin{figure}
-\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread1.pdf}
-\caption{Performance results for SAML2SSO profile: 1 service and 1 client}
-\label{fig:SAML2SSO_thread1}
+\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10.pdf}
+\caption{Performance of HED echoString services with SAMLSSO profile: 1 service and 10 clients}
+\label{fig:SAML2SSO_thread10}
 \end{figure}
 
-Figure \ref{fig:SAML2SSO_thread1} depicts the average response time and throughputs for one client 
-running against the service. Since the success of authentication from IdP can be regarded as 
-valid for the whole session between client and service, the client with ``Session reused'' 
-need to accomplish the SAML2SSO profile only once for all SOAP invocations on this client, 
-while the client with ``Session non-reused'' will need to accomplish this profile once per 
-SOAP invocation.
+\begin{figure}
+\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread_all.pdf}
+\caption{Performance of HED echoString services with SAMLSSO profile:
+performance changes with client numbers; session non-reused }
+\label{fig:SAML2SSO_thread_all}
+\end{figure}
+
+Figure \ref{fig:SAML2SSO_thread10} depicts the average response time and throughputs for 10 clients
+running simultaneously against the ``echoString'' service enhanced with SAML2SSO profile. Since 
+the success of authentication from IdP can be regarded as valid for the whole session between client 
+and service, the client with ``Session reused'' need to accomplish the SAML2SSO profile only once 
+for all SOAP invocations on this client, while the client with ``Session non-reused'' will need to 
+accomplish this profile once per SOAP invocation.
 
 The average response time of ``Session non-reused'' experiments changes little with the execution 
-duration, and the value is about 730ms; while the throughput is around 82 requests per minutes.
-Considering that the average response time in case of 
-``Session non-reused'' and pure SOAP invocation is around 15ms when client number is 10 (see section
-\ref{sec:perhedandaxis}), not demostrated in section \ref{sec:perhedandaxis}, the same value applies 
-when client number is only 1, therefore we can calculate that the consumption of time for 
-SAML2SSO profile is about 98\% of the whole SOAP invocation with SAML2SSO profile integrated.
-Deeper investigation about recording the time of each step of SAML2SSO profile supports the above
-obersavation. In order to improve the performance of SAML2SSO profile, optimizing the process between
+duration, and the value is about 3000ms; while the throughput is around 200 requests per minutes.
+Considering that the average response time in case of ``Session non-reused'' and pure SOAP invocation 
+is around 15ms when client number is 10 (see section \ref{sec:perhedandaxis}), therefore we can 
+calculate that the consumption of time for SAML2SSO profile is about 99.5\% of the whole SOAP 
+invocation with SAML2SSO profile integrated.
+
+Deeper investigation about recording the time of each step of SAML2SSO profile supports the 
+above obersavation. The average time cost for getting an authencation request from SP service 
+is about 60ms; the average time cost for authenticating against IdP is about 2750ms; the 
+average time cost for sending the SAML response (from IdP) to SP service is about 20ms;
+the average time cost for verifying (by SP service) the signature of SAML response is 
+around 130ms.
+
+In order to improve the performance of SAML2SSO profile, optimizing the process between 
 client and IdP is the correct direction, and probably one of the easiest way is to enhance the 
@@
 One the other hand, the average response time of ``Session reused'' experiments is about 82 ms,
-which is almost as short as the value in the pure SOAP invocation experiment in \ref{sec:perhedandaxis}.
-The result is resonable because the time consumption caused by only once of SAML2SSO execution on one 
-client is relatively small comparing with the time consumption caused by the following hundreds 
-of pure SOAP invocations.
+which is almost as short as the value in the pure SOAP invocation experiment in section 
+\ref{sec:perhedandaxis}. The result is resonable because the time consumption caused by 
+only once of SAML2SSO execution on one client is relatively small comparing with the time 
+consumption caused by the following hundreds of pure SOAP invocations on the same client.
 
-\begin{figure}
-\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread5.pdf}
-\caption{Performance results for SAML2SSO profile: 1 service and 5 clients}
-\label{fig:SAML2SSO_thread5}
-\end{figure}
+Figure \ref{fig:SAML2SSO_thread_all} depicts the performance results for echoString services
+that are enhanced with SAML2SSO profile, with client number being changed. In this set of experiments, 
+the session is not reused, and secure communication without client certificate authentication is used.
+The results show the average response time increases linearly with the client number, while the 
+throughput increases initially and then keeps in a fixed value (around 215 requests per minute) 
+when the client number is bigger than 4. When the client number is four, the response time is 
+around one second, while the throughput of the service is 215 requests per minutes which is almost
+the fixed value.
+We speculate that in order to get acceptable response time, there should be less than four (including four) 
+concurrent clients if they need to continuously participate one SAML2SSO profile.
 
-\begin{figure}
-\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10.pdf}
-\caption{Performance results for SAML2SSO profile: 1 service and 10 clients}
-\label{fig:SAML2SSO_thread10}
-\end{figure}
-
-Figure \ref{fig:SAML2SSO_thread5} and Figure \ref{fig:SAML2SSO_thread10} depicts the average 
-response time and throughputs for 5 and 10 clients running against one service simultaneously,
-with the session not being reused. The throughput seems to change little with the number of 
-clients increased from five to ten, with the value being abroud 200 requests per minutes;
-while the average response time doubles, with the value increasing from 1.6s to 3.1s. We conclude
-that in order to get acceptable response time, there should be at least less than five concurrent
-clients if they need to continuously participate one SAML2SSO profile.
+Considering that the throughput of ``echoString'' service enhanced with mutual TLS is 13000, and 
+the concurrent client number when arriving fixed throughput is 30, the performance of
+``echoString'' service enhanced with SAML2SSO profile is much worse, with the values being 215
+and 4 correspondingly.
 
@@
 launches another execution of SAML2SSO profile. On the other hand, statistically, users will most
-probably not access grid system in a concurrent way, rather a random way. Therefore, if we 
+probably not access grid system in a concurrent way, rather than a random way. Therefore, if we 
 suppose a user could access grid once per minute, in our test environment, we can claim that 
-around 200 clients randomly can be supported with acceptable performance.
+around 215 clients randomly can be supported with acceptable performance.
 
@@
 \begin{figure}
-\includegraphics[width=0.9\columnwidth]{SLCS.pdf}
-\caption{Performance results for SLCS service: 1 service and 1 client}
-\label{fig:SLCS}
+\includegraphics[width=0.9\columnwidth]{SLCS_thread_all.pdf}
+\caption{Performance of SLCS service: performance changes with client numbers}
+\label{fig:SLCS_thread_all}
 \end{figure}
 
-Figure \ref{fig:SLCS} is the performance about the SLCS service. Since the SLCS is based on the SAML2SSO
-profile, the reponse time should be the accumulation of time consumption about the SAML2SSO processing 
-and the SLCS processing. In this experiment, SLCS service and client run on the same set of machine 
-as section \ref{sec:perhedandaxis}. This figure shows that the average response time changes little with
-the execution time with the value being around 1.3s which is nearly double as the value (730ms) of one client 
-experiment about ``echoString'' service invocation with SAML2SSO integrated. Not demostrated in 
-Figure \ref{fig:SLCS}, the more consumption of time is caused by the creating of X.509 request on the
- client side, which takes about roughly 310ms in average, as well as by the signing of X.509 certificate 
-on the SLCS service side, which takes roughly 33ms in average, together with the time consumption for 
-communication.
+Figure \ref{fig:SLCS_thread_all} is the performance about the SLCS service. Since the SLCS is 
+based on the SAML2SSO profile, the reponse time should be the accumulation of time consumption 
+about the SAML2SSO processing and the SLCS processing. In this experiment, SLCS service and 
+client run on the same set of machine as section \ref{sec:perhedandaxis}.
+
+The results show that the throughput of SLCS service has the same characteristic as the throughput
+of ``echoString'' service enhanced with SAML2SSO profile, with the value being around 215 
+requests per minute; and the average response time resembles the value of ``echoString'' service as 
+well, with the value being increased for about 80ms. The increased time consumption is caused by 
+the creating of X.509 request (1024-bit RSA key pair generation) on the client side.
 
@@
 credential has a relatively long lifetime (12 hours by default) and therefore client only needs to invoke
-SLCS service once per 12 hours, the performance is completely acceptable. Also considering the time 
-consumption for the creation of X.509 request (1024-bit RSA key pair generation) is on the client side,
-we speculate the SLCS service side will not be affected much by the time-consuming key pair generation, and
-therefore its performance resemble the performance of SAML2SSO profile in section \ref{sec:perfsaml2sso}.
+SLCS service once per 12 hours, the performance is completely acceptable.
 
@@
 performance of Web Service invocation. We deployed a ``echoString'' service plugged with delegation
-security handler on a machine equipped with 2.80GHz Intel Pentium CPU and 1GB memory, Red Hat 
-Enterprise Linux 5, and also configure a client together with delegation security handler plugged 
-on a machine equipped with 1.8GHz Intel Pentium CPU and 2GB memory, Ubuntu 8.10. Aside from 
+security handler on a machine equipped with 3GHz Intel Pentium dual-core CPU and 2GB memory, Red 
+Hat Enterprise Linux 5, and also configure a client together with delegation security handler plugged 
+on a machine equipped with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat Enterprise Linux 5. Aside from 
 ``echoString'' service and client, a delegation service is deployed on a machine equipped with 
@@
 \includegraphics[width=0.9\columnwidth]{Delegation_thread_to_perf.pdf}
-\caption{Performance results of delegation service: from 1 client to 15 clients}
-\label{fig:Deleg}
+\caption{Performance of HED echoString service with credential delegation: performance 
+changes with client numbers; echoString service contacts delegation service upon each service 
+invokation from client}
+\label{fig:Delegation_thread_to_perf}
 \end{figure}
 
-Initial experiments which are not demonstrated in this paper show that both the response time and
-throughput almost do not change with the time. Therefore, we measure the performance as the function 
-of the client numbers from 1 to 15. Figure \ref{fig:Deleg} is performance results. We see that 
-the response time increases in a linear manner against client numbers; and the throughput gradually 
-increases, and stops increasing and almost keeps unchanged when the client number is bigger than eight.
-We conclude that if the client number is bigger than eight, in our test environment, there is no 
-throughput benefit can be achieved by adding clients.
+In the experiment, the client launches a new credential delegation for each service invokation, 
+and ``echoString'' service acquires delegated credential from delegation service upon each 
+service invokation.
+Figure \ref{fig:Delegation_thread_to_perf} depicts the average response time and throughput of 
+the ``echoString'' service with delegation security handler embedded. The average response time 
+inreases linearly with the client number, while the throughput starts to keep at fixed value
+(around 400 request per minutes) when the client number is 8, after initially increasing.
+We conclude that if the client number is bigger than eight, in our test environment, there 
+is no throughput benefit can be achieved by adding clients.
 

### 2009-04-30 doc/sec/papers/GRID09/paper.tex
Editorial corrections, some new references; still not finished (~30% left)

@@
 \begin{abstract}
-Security infrastructure is one of the most challenging tasks in the developement, integration and
+Security infrastructure is one of the most challenging tasks in the development, integration and
 deployment of Grid middlewares. Even though the Grid community addresses the security issue 
 through public key infrastructures (PKI) to support mutual authentication using X.509 certificates,
-maintaining the X.509 credential is not that easy for non-IT specialists, and is proved to be an 
-obstacle for the more widely deployment of Grid technology. Considering the identity federation 
-technology is a emergingly widely used technology which can facilitate cross-domain single sign-on,
-without requiring the users to maintain the X.509 credentials rather than their own institutional
-accounts, we believe utilizing identity federation for Grid middlwares is a promising direction 
-for Grid technology to be more widely used.
-This paper describes a single sign-on infrastructure developed as part of the NorduGrid/ARC grid 
-middleware. It adopts the identity federation standard (SAML), as well as Web Service standand. 
-It focus on single sign-on solution on the middleware level for users to access Grids by only 
+maintaining X.509 credentials is not that easy for non-IT-experts, and has proved to be an 
+obstacle for a more wide deployment of Grid technologies. The identity federation 
+is an increasingly popular technology that can facilitate cross-domain single sign-on
+without requiring the users to maintain any credentials additional to their own institutional
+accounts. We believe that utilizing identity federation for Grid middlewares is a promising path 
+for the Grid technology to get more widely used.
+This paper describes a single sign-on infrastructure developed as a part of the NorduGrid ARC Grid 
+middleware. It adopts the identity federation standard (SAML), as well as Web Service approach. 
+It focuses on a single sign-on solution at the middleware level for users to access Grids by only 
 using their frequently used accounts, without being bothered to maintain X.509 credentials. 
@@
 the middleware level can provide more flexibility.
-Also, the performance of single sign-on solution is measured. We identify the performance limitation 
-of security-related services inside this solution, and analyze the ways to avoid the limitation.
-To our knowledge, the work described in this paper is the first evaluated implementation that utilizes 
-identity federation for Grid usage on the middlware level.
+In addition, the performance of single sign-on solution is measured. We identify performance limitations 
+of security-related services inside this solution, and analyse the ways to avoid the limitations.
+To our knowledge, the work presented in this paper is the first evaluated implementation that utilizes 
+identity federation for Grid usage on the middleware level.
 \end{abstract}
@@
 \label{sec:intro}
-Grid infrastructures facilitate the interoperability between wide-scale, cross-domain heterogeneous 
-resources, as well as the accessibility provided to users. In terms of security issue, the Grid security 
-infrastruture should provide accessibility as much as possible without loosing the security benifits.
+Grid infrastructures facilitate interoperability between wide-scale, cross-domain heterogeneous 
+resources, as well as the user access to these resources. In terms of security issues, a Grid security 
+infrastructure should provide as simple accessibility as possible without loosing the security benefits.
 
-The current Grid community uses Grid Security Infrastructure (GSI)~\cite{IFoster98} as the de facto 
-standard for authentication and transport level security, which builds on the Public Key Infrastructure
-(PKI) to support authentication using X.509 certificate. Mutual authentication is required 
-by GSI based Grid deployment. Mutual authentication means that Grid users have to possess X.509 
-credentials. To obtain a X.509 credential, a user needs to contact the Registration Authority (RA)
-which will check the user's information and approve his certificate request; afterwards,
-the Certificate Authority (CA) can issue the certificate according to the approval from 
-the RA. The whole process is arduous and may take some time.
-Meanwhile, to maintain a X.509 certificate is also not so easy for the non-IT community to deal 
+Many current Grid communities predominantly use the Grid Security Infrastructure (GSI)~\cite{IFoster98} as the \textit{de facto} 
+standard for authentication and transport level security. It builds on the Public Key Infrastructure
+(PKI) to support authentication using X.509 certificates. GSI-based Grid deployment requires mutual authentication.
+Mutual authentication means that Grid users have to possess X.509 
+credentials. To obtain an X.509 credential, a user needs to contact a designated Registration Authority (RA)
+which will check the user's personal data and eventually approve her certificate request; afterwards,
+the respective Certificate Authority (CA) can issue the certificate following the approval from 
+the RA. The whole process is arduous and may take quite some time.
+Meanwhile, to maintain an X.509 certificate is also not so easy for the non-IT-skilled community to deal 
 with, since users have to periodically create proxy certificates by using Grid-specific command line 
-utilities, such as grid-proxy-init or voms-proxy-init. Also maintaing the CAs or RAs is not a  
-easy task, especially when the community of Grid users is getting much wider.
+utilities, such as \texttt{grid-proxy-init} or \texttt{voms-proxy-init}. Also, operating CAs or RAs is not an  
+easy task, especially when the community of Grid users is getting much larger.
 
-On the other hand, users normally could already have had some frequently-used campus or institutional 
+On the other hand, users normally would already have some frequently-used campus or company 
 accounts/credentials such as username/password. We believe that enabling Grid users to use their institutional 
@@
 
-To enable users to utilize the institutional credentials to access Grid, some issues needs to be 
-addressed. Firstly, the Grid middleware needs to be improved to support the authentication based on
-institutional credentials instead of X.509 certificates; Secondly, in order to interoperate with
-those Grid infrastructures which require X.509 certificate based mutual authentication, an approach
-should be provided for obtaining X.509 certificate based on the institutional credentials; Thirdly,
-the single sign-on characteristic should be provided so that users can authenticate once and then transverse 
-from Grid resources to Grid resources without being prompted to authenticate again at each of the resources.
+To enable users to utilize the institutional credentials to access Grid, some issues need to be 
+addressed. Firstly, the Grid middleware needs to be extended to support the authentication based on
+institutional credentials instead of X.509 certificates; secondly, in order to interoperate with
+those Grid infrastructures which require X.509 certificate-based mutual authentication, an approach
+should be provided for obtaining X.509 certificate based on the institutional credentials; thirdly,
+the single sign-on characteristic should be provided so that users can authenticate once and then traverse 
+from Grid resources to Grid resources without being prompted to authenticate again at each resource.
 
-The approach in the new version of ARC grid middleware is to utilize the identity federation 
-standards, specifically, the Security Assertion Markup Language (SAML) specification, and Web 
+The new approach in the ARC~\cite{MEllert07} Grid middleware is to utilize the identity federation 
+standards, specifically, the Security Assertion Mark-up Language (SAML) specification, and Web 
 Service standards, to facilitate the Grid single sign-on.
 
-The rest of this paper is organized as follows: Section \ref{sec:arcmiddleware} presents the 
-ARC Grid middleware, especially the architecture of the new version of ARC. Section 
-\ref{sec:intergrationSAML2SSO} describes the solution of using federated identity for 
-Grid authentication by using SAML2SSO profile; Setion \ref{sec:fedidtoX509} describes
-how to obtain a X.509 credential based on federated identity, and how to use it for 
-accessing multiple Grid resources; Section \ref{sec:perfeval} is performance evaluation;
-Section \ref{sec:relatedwork} describes the related work; Section \ref{sec:conclusion} 
+The rest of this paper is organized as follows: Section~\ref{sec:arcmiddleware} presents the 
+ARC Grid middleware, and in particular its new architecture. Section~\ref{sec:intergrationSAML2SSO} 
+describes the solution of using federated identity for 
+Grid authentication by using \texttt{SAML2SSO} profile; Section~\ref{sec:fedidtoX509} describes
+how to obtain X.509 credentials based on federated identity, and how to use such for 
+accessing multiple Grid resources; Section~\ref{sec:perfeval} is dedicated to performance evaluation;
+Section~\ref{sec:relatedwork} describes the related work; Section~\ref{sec:conclusion} 
 contains conclusions and future work outlook.
@@
 
-ARC (Advanced Resource Connector) is an open source Grid middleware solution released 
+ARC (Advanced Resource Connector)~\cite{MEllert07} is an open source Grid middleware solution released 
 under Apache license. ARC middleware aims at developing self-organized, 
-fault-tolerant, non-intrusive, easy-manageable Grid middleware~\cite{MEllert07}. 
+fault-tolerant, non-intrusive, easy-manageable Grid middleware. 
 The current production version of ARC provides Grid services for job submission and 
 management, resource characterization, resource aggregation and discovery, basic data 
-management, integration of Grid security solution, and so on. This release has been 
-deployed and used in production environment, and is one of the widely deployed Grid 
-middlewares in Europe.
+management, integration of Grid security solution, and so on. It has been 
+deployed and used in production environment, and is one of the most widely deployed Grid
+solutions in Europe.
 
-The development version of ARC components is developed by the KnowARC project~\cite{KnowARClink}, 
+Further development of new ARC components is pursued by the EU KnowARC project~\cite{KnowARClink}, 
 based on the functionality of the current production ARC middleware. It aims at 
 implementing a Web Service oriented Grid middleware which will provide higher levels of 
-resource and user abstraction through well-defined Web Service interface~\cite{KnowARCDesignlink} 
+resource and user abstraction through a well-defined Web Service interface~\cite{KnowARCDesignlink} 
 in order to provide interoperability with other Web Service oriented Grid middlewares, as well as 
@@
 
-As the key part of the implementation of new ARC middleware, there is a lightweight Web 
-Service container called Hosting Environment Daemon (HED) which provides hosting environment 
+The key part of the new approach is a lightweight Web 
+Service container called Hosting Environment Daemon (HED), which provides hosting environment 
 for various services at application level, as well as a number of modules to support flexible, 
-interoperable, and efficient communication mechanism for building SOAP based Web Services. 
+interoperable, and efficient communication mechanism for building SOAP-based Web Services. 
 The design of the HED is built around the idea of flexibility and modularity,
@@
 level Web Service implementation by only using the core minimum amount of components. It 
-also simplifies work on the middleware level, e.g. making possible to implement another 
+also simplifies work on the middleware level: for example, it makes possible to implement another 
 communication protocol or authentication mechanism. A system administrator can easily 
@@
 
-The architecture of the HED is illustrated in Figure \ref{fig:HED}. In general, there 
-are few components called Message Chain Component (MCC) which are in charge of implementing 
-different protocol levels. For instance, as shown in the example message flow, the HTTP MCC 
-will process stream from the TLS MCC to parse the HTTP message and pass its body to the SOAP 
+The architecture of HED is illustrated in Figure~\ref{fig:HED}. In general, there 
+are several components called Message Chain Component (MCC) which are in charge of implementing 
+different protocol levels. For example, as shown in the message flow, the HTTP MCC 
+will process a stream from the TLS MCC to parse the HTTP message and pass its body to the SOAP 
 MCC, and also process the SOAP response from the SOAP MCC to generate the HTTP message 
@@
 
-Dotted line in Figure \ref{fig:HED} shows an alternative path for the information 
+Dotted line in Figure~\ref{fig:HED} shows an alternative path for the information 
 to propagate among MCCs. A service administrator can configure the MCCs according to the 
-interoperability requirements with the other part. For instance, the configuration marked 
+interoperability requirements with a counterpart. For instance, the configuration marked 
 with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP 
-message mechanism (see WSE’s SoapSender and SoapReceiver)~\cite{WSElink}. Another configuration 
+message mechanism (c.f. WSE SoapSender and SoapReceiver)~\cite{WSElink}. Another configuration 
 could be SOAP over HTTPG (HTTP over GSI) which is needed to interoperate with services 
@@
 
-HED contains a flexible security framework for implementing and enforcing security 
-related functionality, such as authentication and authorization. 
-Each security related functionality can be implemented as a pluggable and configurable 
-component (plug-ins) called SecHandler. Each MCC or Service is usually configured with 
-two queues of SecHandler -- one for the incoming message and one for the outgoing message 
-respectively. In Figure \ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules inside 
-MCCs and Services are examples of SecHandler.
+HED contains a flexible security framework for implementing and enforcing
+security-related functionality, such as authentication and authorization. 
+Each security-related functionality can be implemented as a pluggable and
+configurable component (plug-in) called \textit{SecHandler}. Each MCC or Service is usually
+configured with two queues of SecHandler -- one for incoming messages and one for
+outgoing messages, respectively. In Figure~\ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules
+inside MCCs and Services are examples of SecHandler.
 
 %-------------------------------------------------------------------------
-\section{Grid authentication using federated identity: Integrating SAML2SSO profile}
+\section{Grid authentication using federated identity: Integrating SAML2SSO
+profile}
 \label{sec:intergrationSAML2SSO}
-Identity federation is an emerging technology which enables the identity information to be 
-trustily transfered across autonomous security domains. By using identity federation, users 
-from one domain can access another domain without the need for directly trust relationship 
-between users and accessed domain, i.e., users can use accounts from their host domains to 
-access external domains, and accessed domain does not need to maintain accounts for those 
-external users. Identity federation doesn't enforce specific authentication mechanisms, so 
-that various types of authentication can be supported between user and its host domain, and 
-user is able to use his frequently used account/credential to accomplish authentication. 
-Considering the Grid use cases, we see that utilizing the existing account to access Grid 
-without being bothered to applying and manipulating a X.509 credential should be attractive 
-for Grid users.
+Identity federation is an emerging technology which enables the identity
+information to be trustily transferred across autonomous security domains. By using identity
+federation, users from one domain can access another domain without the need for a direct trust
+relationship between users and accessed domains, i.e., users can use accounts from their host
+domains to access external domains, and accessed domain does not need to maintain accounts
+for those external users. Identity federation does not enforce specific authentication
+mechanisms, therefore various types of authentication can be supported
+between users and their host domains, and users are able to use their frequently used accounts or credentials to
+accomplish authentication. Considering Grid use cases, we see that utilizing an existing account to
+access Grid without being bothered to apply for and manipulate a X.509 credential
+would be attractive for Grid users.
 
-Shibboleth is the one of the several implementations of the identity federation, which is 
-based on the OASIS Security Assertion Markup Language (SAML) 2.0 specifications. In terms 
-of authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines 
-two functional components, an Identity Provider and a Service Provider. The Identity Provider 
-(IdP) is responsible for creating, maintaining, and managing user identity, while the Service 
-Provider (SP) is responsible for controlling access to services and resources by analyzing 
-the SAML assertions produced and issued by the IdP upon request (request from client application 
-to the service/resource that is protected by SP).
+Shibboleth~\cite{Shiblink} is the one of the several implementations of the
+identity federation, which is based on the OASIS Security Assertion Markup Language (SAML) 2.0
+specifications~\cite{SAMLlink}. In terms of authentication, Shibboleth implements SAML2.0 web browser SSO
+profile~\cite{SAMLprofiles}, which  defines two functional components, an Identity Provider and a Service Provider. The
+Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity,
+while the Service Provider (SP) is responsible for controlling access to services and resources by
+analysing the SAML assertions produced and issued by the IdP upon request (request from
+client application to the service/resource that is protected by SP).
 
-In the implementation of ARC middleware, SAML2.0 web browser SSO profile is supported by 
-implementing the functionality of service provider and web browser's user agent, and utilizing 
-Shibboleth IdP implementation for the functionality of identity provider. The user agent 
-functionality is for mimicing the behavior of web browser's user agent for authentication, 
-such as the HTTP redirection and the HTTP cookie processing. The implementation of the user 
-agent is based on the the HTTPS client interface of ARC, since the client interface of ARC 
-can support different protocols which are incarnated by different MCCs.
+In the implementation of ARC middleware, SAML2.0 web browser SSO profile is
+supported by implementing the functionality of service provider and web browser's user agent,
+and utilizing Shibboleth IdP implementation for the functionality of identity provider. The
+user agent functionality is for mimicking the behaviour of web browser's user agent for
+authentication, such as the HTTP redirection and the HTTP cookie processing. Implementation
+of the user agent is based on the the HTTPS client interface of ARC, since the client
+interface of ARC can support different protocols which are incarnated by different MCCs.
 
@@
 
-Figure \ref{fig:SAML2SSOUML} shows how is SAML2.0 SSO profile integrated into the SOAP 
-invocation between ARC WS Client and Web Service. Step 1 to step 5 in Figure \ref{fig:SAML2SSOUML} 
-is similar to those steps depicted in Web Browser SSO profile[ref], with the differences that
-we specify the way about how does the service provider determine which identity provider to use (identity
-provider discovery), by enabling the client to send the IdP name to service provider; as well 
-as specify the way about how does the client/user agent authenticate against service provider, 
-by implementing the form-based HTTP authentication which uses username and password for authentication between 
-user agent and service provider (rather than X.509 client authentication) and is compatible 
-with the Username/Password login handler of Shibboleth IdP.
+Figure~\ref{fig:SAML2SSOUML} shows how the SAML2.0 SSO profile is integrated
+into the SOAP invocation between ARC WS Client and Web Service. Steps 1 to 5 in
+Figure~\ref{fig:SAML2SSOUML} are similar to those steps depicted in the Web Browser SSO profile, with
+the difference that we specify the way how does the service provider determine which identity
+provider to use (identity provider discovery) by enabling the client to send the IdP name to the service
+provider. We also specify the way how does the client (user agent)
+authenticate against a service provider by implementing the form-based HTTP authentication which uses username and
+password for authentication between the user agent and a service provider (rather than the X.509 client authentication),
+which is compatible with the Username/Password login handler of the Shibboleth IdP.
 
-In step 6, the service provider (SP Service) verifies and checks the SAML response, decrypts and stores 
-the SAML assertion into session/connection context. That assertion includes the AuthnStatement and 
-AttributeStatement elements. From step 8 to step 10, the WS client launches the SOAP invocation via 
-the same connection as the one which is used by user agent to contact SP Service, then Grid/Web Service 
-checks the AuthnStatement stored in the session context to see if the AuthnStatement is valid or expired. 
-If valid, service handles the SOAP request and returns the SOAP response to the WS client. Note
-the ``connection'' and ``session'' are interchangeably used in this paper, with the same meaning for 
-TCP connection.
+In step 6, the service provider (SP Service) verifies and checks the SAML
+response, decrypts and stores the SAML assertion into session/connection context. That assertion includes the
+AuthnStatement and AttributeStatement elements. From step 8 to step 10, the WS client first
+launches the SOAP invocation via the same connection as the one which is used by the user agent to contact an SP
+Service, and then the Grid/Web Service checks the AuthnStatement stored in the session context to see whether the
+AuthnStatement is valid or expired. If valid, the service handles the SOAP request and returns the SOAP response to
+the WS client. Note that the ``connection'' and ``session'' are interchangeably used in this paper, with
+the same meaning for a TCP connection.
 
-Note that in the current implementation the service requires WS client to reuse the same 
-TCP connection as the one used by user agent in step 5, in order to guarantee that the validity of 
-the SAML2SSO result is applicable to the SOAP invocation process. On the service side, 
-the sesion reusing is accomplish by deloying the SP service and other functional Grid/Web Service(s) 
-together in the same service container.
+Note also that in the current implementation the service requires the WS client
+to re-use the same TCP connection as the one used by the user agent in step 5, in order to
+guarantee that the validity of the SAML2SSO result is applicable to the SOAP invocation process. On the service
+side, the session re-using is accomplished by deploying the SP service and other
+functional Grid/Web Service(s) together in the same service container.
 
-There are a few benifits about integrating SAML2.0 SSO profile for Grid/Web service authentication. 
-Firstly, the client certificate authentication can be switched off, so that users do not need 
-to apply and maintain a X.509 credential. Secondly, some implementation of identity provider, such as 
-Shibboleth IdP, can cache the authentication result through session management once the user
-has succeeded to authenticate, and for a short period this authentication result is valid, so that 
-the next time the user authenticates against IdP, the user agent can just feed IdP with the last 
-successfully authenticated session's identity rather than feed IdP with username and password again. So 
-the user (or the client on behalf of that user) can travel across multiple security domains with 
-only providing his name and password once, which is also the characteristic of single sign-on.
-Lastly, even though there are a few implementations of identity providers, most of them are 
-standard-compliant, so the solution implemented in ARC middleware can interoperate with other 
-identity provider implementations with minimum change.
+There are a several benefits with integrating the SAML2.0 SSO profile for
+Grid/Web service authentication. Firstly, the client certificate authentication can be switched off, so that
+users would not need to apply and maintain X.509 credentials. Secondly, some implementations of
+identity provider, such as Shibboleth IdP, can cache the authentication result through session management
+once the user has succeeded to authenticate, and for a short period this authentication result
+is valid, so that the next time the user authenticates against IdP, the user agent can just feed
+IdP with the last successfully authenticated session's identity rather than feed IdP with
+the username and password again. So the user (or the client on behalf of that user) can travel across multiple
+security domains by only providing his name and password once, which is also a characteristic of
+single sign-on. Lastly, even though there are several implementations of identity providers,
+most of them are standard-compliant, so the solution implemented in ARC middleware can
+interoperate with other identity provider implementations with minimum change.
 
@@
 \label{sec:fedidtoX509}
-Many widely used Grid middlewares are based on GSI which requires mutual X.509 authentication. 
-Meanwhile, Web Service based Grid applications mostly do require client certificate authentication.  
-To bridge the gap beween federated identity and X.509 credential, a short lived credential 
-service (SLCS) is implemented through which a user can get a short-lived X.509 certificate only by
-providing his username/password and authenticating through the SAML2SSO profile.
+Many widely used Grid middlewares are based on GSI which requires mutual X.509
+authentication. Meanwhile, Web Service based Grid applications mostly require client
+certificate authentication. To bridge the gap between federated identities and X.509 credentials, a short
+lived credential service (SLCS) is implemented through which a user can get a short-lived X.509
+certificate only by providing his username and password and authenticating through the SAML2SSO
+profile.
 
-Also, in order to provide single sign-on capability for services that act on behalf of user to 
-invoke other services, a X.509 credential delegation service is implemented, which is based on 
-SOAP communication rather than the proprietary GSI communication which is used in the delegation
-services such as MyProxy.
+Also, in order to provide single sign-on capability for services that act on
+behalf of the user to invoke other services, a X.509 credential delegation service is implemented,
+which is based on SOAP communication rather than the proprietary GSI communication which is used
+in the delegation services such as MyProxy~\cite{myproxy}.
 
@@
 \label{sec:slcs}
-The sequence diagram of SLCS process is the same as the diagram shown in 
-Figure \ref{fig:SAML2SSOUML}, with SOAP request (the message of step 8 of 
-Figure \ref{fig:SAML2SSOUML}) including X.509 request 
-and SOAP response (the message of step 10 of Figure \ref{fig:SAML2SSOUML}) 
-including an signed X.509 certificate and the CA certificate that is used to 
-sign this X.509 certificate. 
-In detail, the SLCS client firstly accomplishes the SAML2SSO profile, then sends 
-SOAP request to the SLCS service with the X.509 request embedded. The SLCS service makes
-access control decision according to the SAML assertion stored in session context, and issues 
-a certificate (with 12 hours lifetime) with the SAML attributes as the X.509 certificate 
-extension. The SLCS client then gets the SOAP response with the X.509 certificate embedded.
+The sequence diagram of a SLCS process is the same as the diagram shown in 
+Figure~\ref{fig:SAML2SSOUML}, with the SOAP request (the message of step 8 of 
+Figure~\ref{fig:SAML2SSOUML}) including an X.509 request, and the SOAP response (the message of step 10 
+of Figure~\ref{fig:SAML2SSOUML}) including a signed X.509 certificate and a CA certificate that is used 
+to sign this X.509 certificate. In detail, the SLCS client firstly accomplishes the SAML2SSO profile, 
+then sends the SOAP request to the SLCS service with the X.509 request embedded. The SLCS
+service makes access control decision according to the SAML assertion stored in the session
+context, and issues a certificate (with 12 hours lifetime) with SAML attributes as the X.509
+certificate extension. The SLCS client then gets the SOAP response with the X.509
+certificate embedded.
 
-Since the lifetime of the short lived credential has a short lifetime, it is permissible to protect 
-the private key by using the local file system permissions instead of encrypting it by 
-using a pass-phrase. Therefore, no manual interactivity is required to uncrypted the private
-key. Hence single sign-on is provided, since the user only needs to input his password when 
-he starts SAML2SSO profile.
+Since the lifetime of the short lived credential is short, it is
+permissible to protect the private key by using the local file system permissions instead of encrypting
+it by using a pass-phrase. Therefore, no manual interactivity is required to encrypt
+the private key. Hence a single sign-on is provided, since the user only needs to enter his
+password when starting a SAML2SSO profile.
 
-How to compose a distinguished name (DN) for the certificate is a critical issue for 
-the SLCS service. Since the ``eduPersonPrincipalName'' (see eduPerson schema~\cite{eduSchemalink})
-is identical to the user that accomplishes SAML2SSO profile. We pick up ``eduPersonPrincipalName'' 
-from SAML attributes that returned from Shibboleth IdP, and use it to compose the DN. A 
-typical example of the eduPersonPrincipalName 
-could be alice@example.org. In such case composed DN is for example\\``/O=knowarc/OU=example.org/CN=alice''.
+A critical issue for the SLCS service is how to compose a distinguished name
+(DN) for the certificate. Since the ``eduPersonPrincipalName'' (see eduPerson schema~\cite{eduSchemalink})
+is identical to the user that accomplishes SAML2SSO profile, we pick up
+``eduPersonPrincipalName'' from SAML attributes that returned from Shibboleth IdP, and use it to compose
+the DN. A typical example of the eduPersonPrincipalName 
+could be \textit{alice@example.org}. In such a case the composed DN would be,
+for example,\\``/O=knowarc/OU=example.org/CN=alice''.
 
-By using the SLCS service, the user can easily access the Grid services that requires 
-X.509 credential on client side, at anywhere, simply by running the SLCS client command 
-and providing his username and password.
+By using the SLCS service, a user can easily access Grid services that
+require X.509 credentials on the client side from anywhere, simply by running the SLCS
+client command and providing his username and password.
 
@@
 
-The sequence diagram is shown in Figure \ref{fig:DelegUML}. Firstly, the client needs to delegate
-the proxy certificate or short-lived certificate to a delegation service through step 1 to step 7.
-Then this client invokes the peer Web Service by appending the delegation information (URL of delegation
-service and the delegation ID corresponding to the above delegation) into the SOAP body. The Web 
-Service then sends a X.509 request and the delegation ID to the delegation service 
-which is specified by client side, and gets back a delegated certificate which can be composed together
-with the private key for a proxy certificate. The Web Service could use this proxy certificate 
-to represent the user to invoke another Web Service, which either could cause one more process of 
-delegation (step 1 to step 13) if another delegation service is specified, or could directly start
-from step 8 if the same delegation service is specified as the one in the former delegation process.
+The delegation service usage sequence diagram is shown in
+Figure~\ref{fig:DelegUML}. Firstly, the client needs to delegate
+the proxy certificate or the short-lived certificate to a delegation service
+through steps 1 to 7.
+Then this client invokes the peer Web Service by appending the delegation
+information (URL of delegation service and the delegation ID corresponding to the above delegation) to the
+SOAP body. The Web Service then sends a X.509 request and the delegation ID to the delegation
+service which is specified by the client side, and gets back a delegated certificate
+which can be composed together with the private key for a proxy certificate. The Web Service could use this
+proxy certificate to represent the user to invoke another Web Service, which either could cause
+one more process of delegation (steps 1 to 13) if another delegation service is specified, or
+could directly start from step 8 if the same delegation service is specified as the one in the
+previous delegation process.
 
-The client only needs to delegate the certificate to delegation service one time per session, it does
-not need to do delegation for subsequent SOAP invokations after the first one.
+The client only needs to delegate the certificate to the delegation service once
+per session; it does not have to do delegation for subsequent SOAP invocations after the first one.
 
-The functionality of both client and service sides are implemented as pluggable handlers which can be 
-configured into the client and service configuration, so that the service or client implementation itself
-does not need to be changed.
+The functionalities of both client and service sides are implemented as
+pluggable handlers which can be configured into the client and service configuration, so that the service or
+client implementation itself does not need to be changed.
 
-We suppose the short-lived certificate is used by client, so that the SLCS service together with
-the delegation service can provide single sign-on from user's federated identity, while keeping the 
-interoperability to those Grid/Web services that requires mutual authentication using certificate.
+We assume that the short-lived certificate is used by the client, so that the
+SLCS service together with the delegation service can provide single sign-on from user's federated
+identity, while keeping the interoperability to those Grid/Web services that require mutual authentication
+using certificate.
 
 %-------------------------------------------------------------------------
-\section{Performace Evaluation}
+\section{Performance Evaluation}
 \label{sec:perfeval}
 The following sections evaluate the performance of the HED versus Axis2/C, as 
-well as the performance about the SAML2SSO profile, short-lived credential
+well as the performance related to the SAML2SSO profile, short-lived credential
 service, and delegation service.
@@
 \label{sec:perhedandaxis}
-Since the implementation of HED includes an alternative SOAP implementation which is the base of 
-the security related web servies described in this paper, it is useful to compare the performance 
-of HED and other SOAP implementations. Although there are many SOAP implementations such as 
-AxisJava, gSOAP, XSOAP4, etc., the goal of this paper is not about complete comparison between
-HED and all the other SOAP implementations, so we use the Axis2/C (v1.5.0) to demostrate the performance
-HED.
+Since the implementation of HED includes an alternative SOAP implementation
+which is the base of the security related web services described in this paper, it is useful to
+compare the performance of HED and other SOAP implementations. Although there are many SOAP
+implementations such as AxisJava, gSOAP, XSOAP4, etc.., it is not the goal of this paper to do a
+complete comparison between HED and all the other SOAP implementations, so we use Axis2/C (v1.5.0) to
+demonstrate the performance HED.
 
-For the service side, we implement two simple ``echoString'' services in HED and Axis2/C. The 
-``echoString'' provides a simple SOAP portType which will output the same value as it gets from
-input.
-For client side, we develope a client by using the client API of ARC, to invoke the ``echoString''
-services from Axis2/C and HED. The service invocation is just sending a short string to service, and 
-getting back the same string. One test/experiment consists of a set of clients which run simultaneously. 
-Each client runs for a fixed period of time, and invokes the service as many times as it can during 
-the test period. Multiple clients are launched by creating multiple threads in one test.
-Two values are measured: average response time and throughput. Average response time is computed
-by counting the average value of the response time for each invocation. Throughput is computed
-by counting the number of invocation during each second.
+For the service side, we implement two simple ``echoString'' services in HED and
+Axis2/C. The ``echoString'' provides a simple SOAP portType which will output the same value
+as it gets from input. For the client side, we developed a client by using the client API of ARC, to
+invoke ``echoString'' services from Axis2/C and HED. Service invocation just sends a short
+string to a service, and gets back the same string. One test involves a set of clients
+which run simultaneously. Each client runs for a fixed period of time, and invokes the service as many
+times as it can during the test period. Multiple clients are launched by creating multiple threads in
+one test. Two values are measured: average response time and throughput. Average response
+time is computed by counting the average value of the response time for each invocation.
+Throughput is computed by counting number of invocations per second.
 
-The two ``echoString'' services run in a test machine equipped with 3GHz Intel Pentium dual-core 
-CPU and 2GB memory, Red Hat Enterprise Linux 5. The client runs in another machine equipped 
-with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat Enterprise Linux 5. The Axis2/C service is run using 
-the httpd module on Apache Http Server version 2.2.11 instance. We keep all the configurations 
-default in Apache Http Server.
+The two ``echoString'' services run in a test machine equipped with a 3GHz Intel
+Pentium dual-core CPU and 2GB memory, Red Hat Enterprise Linux 5. The client runs in another
+machine equipped with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat Enterprise Linux 5. The
+Axis2/C service is run using the httpd module on Apache Http Server version 2.2.11 instance. We use all the
+default configurations in the Apache Http Server.
 
@@
 \includegraphics[width=0.9\columnwidth]{TCP_thread_all.pdf}
-\caption{Performance of HED and Axis2C echoString services without TLS: performance
-changes with client numbers; session non-reused }
+\caption{Performance of HED and Axis2C echoString services without TLS:
+performance
+changes with client numbers; session not re-used }
 \label{fig:TCP_thread_all}
@@
 \caption{Performance of HED and Axis2C echoString services with TLS: performance
-changes with client numbers; session non-reused}
+changes with client numbers; session not re-used}
 \label{fig:TLS_thread_all}
@@
 
-Figure \ref{fig:HED2Axis_thread10} is the results of performance comparison with ten 
+Figure~\ref{fig:HED2Axis_thread10} shows results of performance comparison with
+ten 
 clients being launched.
-Two types of experiments are shown: session reused, session non-reused.
-``Session reused'' means all of the SOAP invocations on each client share the same session 
-(tcp connection); while ``Session non-reused'' means each invocation launches a new session.
+Two types of experiments are shown: session re-used and session not re-used.
+``Session re-used'' means that all the SOAP invocations on each client share
+the same session (TCP connection); while ``Session not re-used'' means that each invocation
+launches a new session.
 
-For ``Session reused'' experiments, we see that the average response time in HED is about 81ms
-which is twice of the value in Axis2/C, and the throughput in HED is about half of the value in Axis2/C. 
-One obersavation is that, the performance results of HED is quite stable, with the average response time 
-and throughput almost kept unchanged under different time points.
-For ``Session non-reused'' experiments, the average response time in HED is almost the same as
-the value in Axis2/C, with the value 15ms. On the other hand, the throughput in HED is close to
- the value in Axis2/C.
+For the ``Session re-used'' experiments, we see that the average response time
+in HED is about 81ms which is twice the value for Axis2/C, and the throughput in HED is about half
+of the value for Axis2/C. One observation is that the performance results of HED are quite stable, with
+the average response time and throughput almost unchanged for different time points.
+For the ``Session not re-used'' experiments, the average response time in HED is
+almost the same as the value for Axis2/C, with the value 15ms. On the other hand, the throughput in
+HED is close to the value for Axis2/C.
 
-Since the echoString service is a simple service which occupies little and stable time for processing
-service logic itself, and also the experiments are done under local network, we can speculate 
-the response time we get in Figure \ref{fig:HED2Axis_thread10} is mainly about the processing time 
-in container itself.
-With the stability characteristic of performance, we can speculate that the service container 
-itself will stablly impact on the performance of other services.
+Since the echoString service is a simple service which occupies little
+resources and has stable time for the processing service logic itself, and also since the experiments are done over the local
+network, we can speculate that the response time shown in Figure~\ref{fig:HED2Axis_thread10} mainly
+corresponds to the processing time in the container itself.
+With the stability characteristic of the performance, we can speculate that the
+service container itself will have a stable impact on the performance of other services.
 
-To more deeply compare the performance of services from HED and Axis2/C, some tests have
-been done to investigate how does the number of simultaneous clients impact on the performance.
-Figure \ref{fig:TCP_thread_all} is the performance results for echoString services with client number
-being changed. In this set of experiments, the session is not reused, and secure communication is not
-used. Each experiment is executed for a relatively long duration (two minutes), so that the there are
-enough times of service invokations and then the performance results are reliable.
-We see that both the response time of services in HED and Axis2/C starts from similar value 
-(when client number is five) and changes linearly with number of clients, while service in HED 
-provides shorter response time with the value being around two thirds of the response time of 
-service in Axis2/C when client number becomes bigger.
-In case of throughput, service in HED provides stable throughtput, while service in Axis2/C provides 
-throughput which slightly becomes smaller when client number becomes bigger.
+To compare more deeply performance of the services in HED and Axis2/C, some
+tests have been done to investigate how does the number of simultaneous clients affect
+the performance. Figure~\ref{fig:TCP_thread_all} shows performance results for echoString
+services with the changing number of clients. In this set of experiments, the session is not re-used, and secure
+communication is not used. Each experiment is executed for a relatively long duration (two minutes),
+so that there are enough instances of service invocation and thus the performance measurements are
+reliable. We see that the response time of services in both HED and Axis2/C starts from a
+similar value (when the client number is five) and changes linearly with the number of
+clients, while the service in HED provides shorter response time with the value being around two thirds of the
+response time of the service in Axis2/C when the number of clients becomes higher.
+In the case of throughput, the service in HED provides stable throughput, while
+the service in Axis2/C provides throughput which  becomes slightly smaller when number of clients becomes
+higher.
 
-Figure \ref{fig:TLS_thread_all} is the performance results when secure communication is used.
-For Axis2/C, we configure the http server (that hosts Axis2/C) with SSL module enabled.
-Each experiment is executed for a duration of two minutes.
-We see that, similar to Figure \ref{fig:TCP_thread_all}, the both the response time of services 
-in HED and Axis2/C also starts from similar value (when client number is five) and changes linearly 
-with number of clients, while the service in HED also provides much shorter response time 
-with the value being around sightly more than one third of the response time of service in Axis2/C 
-when client number becomes bigger.
-In case of throughput, the throughput of service in HED increases linearly with client numbers intially,
-however, it starts to keep stable to be a fixed value (around 13000 requests per minutes) when client 
-number is bigger than 30. On the other hand, the throughput of service in Axis2/C starts to keep 
-stable when client number is bigger than 10, with the value being around sightly more than one 
-third of the throughput of service in HED.
+Figure~\ref{fig:TLS_thread_all} shows performance results when secure
+communication is used. For Axis2/C, we configure the http server (that hosts Axis2/C) with the SSL
+module enabled. Each experiment is executed for a duration of two minutes.
+We see that, similarly to Figure~\ref{fig:TCP_thread_all}, response
+time of services both in HED and Axis2/C also starts from a similar value (when the number of
+clients is five) and changes linearly with the number of clients, while the service in HED also provides much shorter
+response time with the value being sightly more than about one third of the response time of
+the service in Axis2/C when  number of clients becomes higher.
+In the case of throughput, the throughput of service in HED increases linearly
+with the client numbers initially, however, it saturates at around 13000 requests per
+minutes when the number  of client number is higher than 30. On the other hand, the throughput of service in
+Axis2/C saturates already when the number of clients reaches 10, with
+the value being around sightly more than one third of the throughput of service in HED.
 
-We conclude that the HED can provide better performance than Axis2/C, in both cases of secure 
-communication being enabled and disabled, especially when the number of simultaneous clients 
-is bigger.
+We conclude that the HED can provide better performance than Axis2/C, in both
+cases of secure communication being enabled and disabled, especially when the number of
+simultaneous clients is bigger.
 
-\subsection{Performance of the integration SAML2 SSO profile}
+\subsection{Performance of the integration of the SAML2 SSO profile}
 \label{sec:perfsaml2sso}
-In our next experiment, we evaluate how does the integration of SAML2SSO profile impact on the 
-performance of Web Service invocation. We configure a ``echoString'' service together with a SP (
-Service Provider) service, and develope a client by using SAML2SSO related API of ARC, to invoke 
-the ``echoString'', as well as interact with the IdP (Identity Provider). Service and clients run on
-the same set of machine as section \ref{sec:perhedandaxis}. Meanwhile, we deploy the Shibboleth IdP 
-implementation on a machine equipped with 2.8GHz Intel Pentium CPU and 1GB memory, Red Hat Enterprise 
-Linux 5.
+In our next experiment, we evaluate how does the integration of the SAML2SSO
+profile affect the performance of Web Service invocation. We configure an ``echoString'' service
+together with as SP (Service Provider) service, and develop a client by using
+SAML2SSO related API of ARC, to invoke the ``echoString'', as well as interact with the IdP (Identity Provider).
+Service and clients run on the same set of machine as Section~\ref{sec:perhedandaxis}. Meanwhile, we deploy
+the Shibboleth IdP implementation on a machine equipped with 2.8GHz Intel Pentium CPU and 1GB
+memory, Red Hat Enterprise Linux 5.
 
@@
 \includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10.pdf}
-\caption{Performance of HED echoString services with SAMLSSO profile: 1 service and 10 clients}
+\caption{Performance of HED echoString services with SAMLSSO profile: 1 service
+and 10 clients}
 \label{fig:SAML2SSO_thread10}
@@
 
-Figure \ref{fig:SAML2SSO_thread10} depicts the average response time and throughputs for 10 clients
-running simultaneously against the ``echoString'' service enhanced with SAML2SSO profile. Since 
-the success of authentication from IdP can be regarded as valid for the whole session between client 
-and service, the client with ``Session reused'' need to accomplish the SAML2SSO profile only once 
-for all SOAP invocations on this client, while the client with ``Session non-reused'' will need to 
-accomplish this profile once per SOAP invocation.
+Figure \ref{fig:SAML2SSO_thread10} depicts the average response time and
+throughputs for 10 clients running simultaneously against the ``echoString'' service enhanced with SAML2SSO
+profile. Since the success of authentication from IdP can be regarded as valid for the whole
+session between client and service, the client with ``Session reused'' need to accomplish the SAML2SSO
+profile only once for all SOAP invocations on this client, while the client with ``Session
+non-reused'' will need to accomplish this profile once per SOAP invocation.
 
-The average response time of ``Session non-reused'' experiments changes little with the execution 
-duration, and the value is about 3000ms; while the throughput is around 200 requests per minutes.
-Considering that the average response time in case of ``Session non-reused'' and pure SOAP invocation 
-is around 15ms when client number is 10 (see section \ref{sec:perhedandaxis}), therefore we can 
-calculate that the consumption of time for SAML2SSO profile is about 99.5\% of the whole SOAP 
-invocation with SAML2SSO profile integrated.
+The average response time of ``Session non-reused'' experiments changes little
+with the execution duration, and the value is about 3000ms; while the throughput is around 200
+requests per minutes. Considering that the average response time in case of ``Session non-reused'' and
+pure SOAP invocation is around 15ms when client number is 10 (see section \ref{sec:perhedandaxis}),
+therefore we can calculate that the consumption of time for SAML2SSO profile is about 99.5\% of
+the whole SOAP invocation with SAML2SSO profile integrated.
 
-Deeper investigation about recording the time of each step of SAML2SSO profile supports the 
-above obersavation. The average time cost for getting an authencation request from SP service 
-is about 60ms; the average time cost for authenticating against IdP is about 2750ms; the 
-average time cost for sending the SAML response (from IdP) to SP service is about 20ms;
-the average time cost for verifying (by SP service) the signature of SAML response is 
-around 130ms.
+Deeper investigation about recording the time of each step of SAML2SSO profile
+supports the above observation. The average time cost for getting an authentication request
+from SP service is about 60ms; the average time cost for authenticating against IdP is about
+2750ms; the average time cost for sending the SAML response (from IdP) to SP service is
+about 20ms; the average time cost for verifying (by SP service) the signature of SAML
+response is around 130ms.
 
-In order to improve the performance of SAML2SSO profile, optimizing the process between 
-client and IdP is the correct direction, and probably one of the easiest way is to enhance the 
-hardware configuration.
+In order to improve the performance of SAML2SSO profile, optimizing the process
+between client and IdP is the correct direction, and probably one of the easiest way is
+to enhance the hardware configuration.
 
-One the other hand, the average response time of ``Session reused'' experiments is about 82 ms,
-which is almost as short as the value in the pure SOAP invocation experiment in section 
-\ref{sec:perhedandaxis}. The result is resonable because the time consumption caused by 
-only once of SAML2SSO execution on one client is relatively small comparing with the time 
-consumption caused by the following hundreds of pure SOAP invocations on the same client.
+One the other hand, the average response time of ``Session reused'' experiments
+is about 82 ms, which is almost as short as the value in the pure SOAP invocation experiment in
+section \ref{sec:perhedandaxis}. The result is reasonable because the time consumption
+caused by only once of SAML2SSO execution on one client is relatively small comparing with
+the time consumption caused by the following hundreds of pure SOAP invocations on the
+same client.
 
-Figure \ref{fig:SAML2SSO_thread_all} depicts the performance results for echoString services
-that are enhanced with SAML2SSO profile, with client number being changed. In this set of experiments, 
-the session is not reused, and secure communication without client certificate authentication is used.
-The results show the average response time increases linearly with the client number, while the 
-throughput increases initially and then keeps in a fixed value (around 215 requests per minute) 
-when the client number is bigger than 4. When the client number is four, the response time is 
-around one second, while the throughput of the service is 215 requests per minutes which is almost
-the fixed value.
-We speculate that in order to get acceptable response time, there should be less than four (including four) 
-concurrent clients if they need to continuously participate one SAML2SSO profile.
+Figure~\ref{fig:SAML2SSO_thread_all} depicts the performance results for
+echoString services that are enhanced with SAML2SSO profile, with client number being changed. In
+this set of experiments, the session is not reused, and secure communication without client certificate
+authentication is used. The results show the average response time increases linearly with the client
+number, while the throughput increases initially and then keeps in a fixed value (around 215
+requests per minute) when the client number is bigger than 4. When the client number is four, the
+response time is around one second, while the throughput of the service is 215 requests per
+minutes which is almost the fixed value.
+We speculate that in order to get acceptable response time, there should be less
+than four (including four) concurrent clients if they need to continuously participate one SAML2SSO
+profile.
 
-Considering that the throughput of ``echoString'' service enhanced with mutual TLS is 13000, and 
-the concurrent client number when arriving fixed throughput is 30, the performance of
-``echoString'' service enhanced with SAML2SSO profile is much worse, with the values being 215
-and 4 correspondingly.
+Considering that the throughput of the ``echoString'' service enhanced with mutual
+TLS is 13000, and the concurrent client number when arriving fixed throughput is 30, the
+performance of ``echoString'' service enhanced with SAML2SSO profile is much worse, with the
+values being 215 and 4 correspondingly.
 
-However for some typical grid application scenarios, normally the client does not need 
-to continuously accomplish SAML2SSO profile. For instance, a user could submit a job 
-which needs a period of time (normally this period is much longer than a few seconds)
-to execute, which launches an execution of SAML2SSO profile; 
-afterwards, this user could query the information as well as the result about this job, which 
-launches another execution of SAML2SSO profile. On the other hand, statistically, users will most
-probably not access grid system in a concurrent way, rather than a random way. Therefore, if we 
-suppose a user could access grid once per minute, in our test environment, we can claim that 
-around 215 clients randomly can be supported with acceptable performance.
+However for some typical grid application scenarios, normally the client does
+not need to continuously accomplish SAML2SSO profile. For instance, a user could submit a
+job which needs a period of time (normally this period is much longer than a few
+seconds) to execute, which launches an execution of SAML2SSO profile; 
+afterwards, this user could query the information as well as the result about
+this job, which launches another execution of SAML2SSO profile. On the other hand,
+statistically, users will most probably not access grid system in a concurrent way, rather than a random way.
+Therefore, if we suppose a user could access grid once per minute, in our test environment, we
+can claim that around 215 clients randomly can be supported with acceptable performance.
 
-\subsection{Performance of SLCS service}
+\subsection{Performance of the SLCS service}
 \label{sec:perfslcsserv}
@@
 
-Figure \ref{fig:SLCS_thread_all} is the performance about the SLCS service. Since the SLCS is 
-based on the SAML2SSO profile, the reponse time should be the accumulation of time consumption 
-about the SAML2SSO processing and the SLCS processing. In this experiment, SLCS service and 
-client run on the same set of machine as section \ref{sec:perhedandaxis}.
+Figure \ref{fig:SLCS_thread_all} is the performance about the SLCS service.
+Since the SLCS is based on the SAML2SSO profile, the response time should be the accumulation of
+time consumption about the SAML2SSO processing and the SLCS processing. In this experiment, SLCS
+service and client run on the same set of machine as section~\ref{sec:perhedandaxis}.
 
-The results show that the throughput of SLCS service has the same characteristic as the throughput
-of ``echoString'' service enhanced with SAML2SSO profile, with the value being around 215 
-requests per minute; and the average response time resembles the value of ``echoString'' service as 
-well, with the value being increased for about 80ms. The increased time consumption is caused by 
-the creating of X.509 request (1024-bit RSA key pair generation) on the client side.
+The results show that the throughput of SLCS service has the same characteristic
+as the throughput of ``echoString'' service enhanced with SAML2SSO profile, with the value being
+around 215 requests per minute; and the average response time resembles the value of
+``echoString'' service as well, with the value being increased for about 80ms. The increased time
+consumption is caused by the creating of X.509 request (1024-bit RSA key pair generation) on the client
+side.
 
-Even though the average response time of SLCS service is not very short, considering the short-lived 
-credential has a relatively long lifetime (12 hours by default) and therefore client only needs to invoke
-SLCS service once per 12 hours, the performance is completely acceptable.
+Even though the average response time of SLCS service is not very short,
+considering the short-lived credential has a relatively long lifetime (12 hours by default) and therefore
+client only needs to invoke SLCS service once per 12 hours, the performance is completely acceptable.
 
-\subsection{Performance of delegation service}
+\subsection{Performance of the delegation service}
 \label{sec:perfdelegserv}
 
-In the last experiment, we evaluate how does the usage of delegation service impact on the 
-performance of Web Service invocation. We deployed a ``echoString'' service plugged with delegation
-security handler on a machine equipped with 3GHz Intel Pentium dual-core CPU and 2GB memory, Red 
-Hat Enterprise Linux 5, and also configure a client together with delegation security handler plugged 
-on a machine equipped with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat Enterprise Linux 5. Aside from 
-``echoString'' service and client, a delegation service is deployed on a machine equipped with 
-3GHz Intel Pentium dual-core CPU and 2GB memory, Red Hat Enterprise Linux 5.
+In the last experiment, we evaluate how does the usage of delegation service
+impact on the performance of Web Service invocation. We deployed a ``echoString'' service
+plugged with delegation security handler on a machine equipped with 3GHz Intel Pentium dual-core CPU and
+2GB memory, Red Hat Enterprise Linux 5, and also configure a client together with delegation
+security handler plugged on a machine equipped with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat
+Enterprise Linux 5. Aside from ``echoString'' service and client, a delegation service is deployed on a machine
+equipped with 3GHz Intel Pentium dual-core CPU and 2GB memory, Red Hat Enterprise Linux 5.
 
@@
 \includegraphics[width=0.9\columnwidth]{Delegation_thread_to_perf.pdf}
-\caption{Performance of HED echoString service with credential delegation: performance 
-changes with client numbers; echoString service contacts delegation service upon each service 
-invokation from client}
+\caption{Performance of HED echoString service with credential delegation:
+performance changes with client numbers; echoString service contacts delegation service upon
+each service invocation from client}
 \label{fig:Delegation_thread_to_perf}
@@
 
-In the experiment, the client launches a new credential delegation for each service invokation, 
-and ``echoString'' service acquires delegated credential from delegation service upon each 
-service invokation.
-Figure \ref{fig:Delegation_thread_to_perf} depicts the average response time and throughput of 
-the ``echoString'' service with delegation security handler embedded. The average response time 
-inreases linearly with the client number, while the throughput starts to keep at fixed value
-(around 400 request per minutes) when the client number is 8, after initially increasing.
-We conclude that if the client number is bigger than eight, in our test environment, there 
-is no throughput benefit can be achieved by adding clients.
+In the experiment, the client launches a new credential delegation for each
+service invocation, and ``echoString'' service acquires delegated credential from delegation service
+upon each service invocation.
+Figure~\ref{fig:Delegation_thread_to_perf} depicts the average response time and
+throughput of the ``echoString'' service with delegation security handler embedded. The
+average response time increases linearly with the client number, while the throughput starts to keep at
+fixed value (around 400 request per minutes) when the client number is 8, after initially
+increasing.
+We conclude that if the client number is bigger than eight, in our test
+environment, there is no throughput benefit can be achieved by adding clients.
 
@@
 \label{sec:relatedwork}
-There are other researches that utilize Shibboleth for protecting usage of Grid resources. 
-The work in ~\cite{J.Watt06} uses Shibboleth to protect a GridSphere portal and then 
-indirectly protect Grid services which are deployed in this GridSphere protal. Another 
-work in ~\cite{R.O.Sinnott06} also uses Shibboleth to protect Grid portal, moreover, it 
-maintains a few common used server certificates which can then be used to access Grid 
-resources after users have successfully authenticated themselves at Grid protal through
-Shibboleth. Unlike these two solutions that use Shibboleth to protect Grid portal, the solution
-in this paper uses Shibboleth or similar standards-based identity federation to directly protect 
-Grid resource/services, which provide possibility for those Grid applications that are directly
-based on Grid interface (e.g., applications that invokes SOAP operations of remote Grid services),
-to benefit from identity federation.
+There are other researches that utilize Shibboleth for protecting usage of Grid
+resources. The work in ~\cite{J.Watt06} uses Shibboleth to protect a GridSphere portal and
+then indirectly protect Grid services which are deployed in this GridSphere portal.
+Another work in ~\cite{R.O.Sinnott06} also uses Shibboleth to protect Grid portal,
+moreover, it maintains a few common used server certificates which can then be used to access
+Grid resources after users have successfully authenticated themselves at Grid portal
+through Shibboleth. Unlike these two solutions that use Shibboleth to protect Grid
+portal, the solution in this paper uses Shibboleth or similar standards-based identity federation to
+directly protect Grid resource/services, which provide possibility for those Grid applications
+that are directly based on Grid interface (e.g., applications that invokes SOAP operations of
+remote Grid services), to benefit from identity federation.
 
-GridShib~\cite{T.Scavo07,VWelch05} provides one solution that users can use X.509 
-certificate to authenticate to Shibbolth IdP, get back SAML assertion and embed it in 
-the proxy certificate. The Grid service will then use this SAML assertion for access control.
-Unlike the work described in this paper, GridShib still requires users to possess a X.509 
-credential, rather than username/password. GridShib also provides another 
-solution~\cite{TBarton06} in which a credential service (MyProxy server) is deployed together
-with Shibboleth IdP for acting as an online-CA, authenticating the user through username/password
-rather than X.509 certificate, and issuing short-lived X.509 credential. The second solution
-is similar to the short-lived credential service in our solution, while our solution
-also provides the option for users directly authenticating to Grid services via SAML2.0 SSO 
-profile using their username/password.
+GridShib~\cite{T.Scavo07,VWelch05} provides one solution that users can use
+X.509 certificate to authenticate to Shibboleth IdP, get back SAML assertion and embed
+it in the proxy certificate. The Grid service will then use this SAML assertion for
+access control. Unlike the work described in this paper, GridShib still requires users to
+possess a X.509 credential, rather than username/password. GridShib also provides another 
+solution~\cite{TBarton06} in which a credential service (MyProxy server) is
+deployed together with Shibboleth IdP for acting as an online-CA, authenticating the user through
+username/password rather than X.509 certificate, and issuing short-lived X.509 credential. The
+second solution is similar to the short-lived credential service in our solution, while our
+solution also provides the option for users directly authenticating to Grid services via
+SAML2.0 SSO profile using their username/password.
 
-Another short-lived credential service is also provided by ~\cite{switchslcslink}. 
-The advantage of our work comparing to the work in ~\cite{switchslcslink} is that 
-our SLCS service is based on Web Services standard, so that it is easier to achieve 
-interoperability.
+Another short-lived credential service is also provided by SWITCH~\cite{switchslcslink}. 
+The advantage of our work comparing to the work of SWITCH is that 
+our SLCS service is based on the Web Services standard, thus that it is easier to
+achieve interoperability.
 
-In the aspect of credential delegation, apart from the credential delegation mechanism of 
-GSI~\cite{IFoster98,VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol 
-based on the WS-Trust~\cite{WSTrustlink} to provide a standard and interoperable protocol 
-for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the 
-specifications required to define delegation protocol in a standard way. Gridsite project also
-implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web 
-Service, with which ARC delegation client can be easily made interoperate.
+In the aspect of credential delegation, apart from the credential delegation
+mechanism of GSI~\cite{IFoster98,VWelch04}, several authors~\cite{MAhsant04} propose a
+delegation protocol based on the WS-Trust~\cite{WSTrustlink} to provide a standard and interoperable
+protocol for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware
+to express the specifications required to define delegation protocol in a standard way.
+Gridsite project also implements a X.509 credential delegation solution~\cite{GridSitelink} based on
+the Web Service, with which ARC delegation client can be easily made interoperate.
 
@@
 \label{sec:conclusion}
-In this paper we propose a Grid single sign-on framework that can utilize users' 
-federated identity instead of X.509 mutual authentication for authentication, which
-we believe can make further uptake of Grid resources as well as Grid middleware by 
-wider scale of users.
+In this paper we propose a Grid single sign-on framework that can utilize
+users' federated identity instead of X.509 mutual authentication for authentication,
+which we believe can make further uptake of Grid resources as well as Grid middleware
+by wider scale of users.
 
-We implement the SAML2 SSO profile in the ARC Grid middleware, so that the Grid applications
-that developed by using the application interface provided by this middleware can easily 
-be switched to benefit from federated identity based authentication. 
-The performance evaluation shows even though the introducing of SAML2 SSO profile causes 
-much performance downgrade, this downgrade can be avoided by reusing the authentication 
-result for multiple SOAP invocations.
-Meanwhile, in order to be compatible with those Grid applications that require X.509 
-certificate on the client side, we also implement one Web Service which can be used for 
-obtaining an X.509 credential based on federated identity, and the other Web Service 
-which can be used for delegating X.509 credential. We believe that the evaluated 
-performance is enough for creating X.509 certificate and proxy certificate, considering that 
-the certificates should have a relaively much longer lifetime comparing to the response time of 
-these two services.
+We implement the SAML2 SSO profile in the ARC Grid middleware, so that the Grid
+applications that developed by using the application interface provided by this middleware
+can easily be switched to benefit from federated identity based authentication. 
+The performance evaluation shows even though the introducing of SAML2 SSO
+profile causes much performance downgrade, this downgrade can be avoided by reusing the
+authentication result for multiple SOAP invocations.
+Meanwhile, in order to be compatible with those Grid applications that require
+X.509 certificate on the client side, we also implement one Web Service which can be
+used for obtaining an X.509 credential based on federated identity, and the other Web
+Service which can be used for delegating X.509 credential. We believe that the
+evaluated performance is enough for creating X.509 certificate and proxy certificate,
+considering that the certificates should have a relatively much longer lifetime comparing to the
+response time of these two services.
 
-Although only authentication issue has been discussed in this paper, we know single
-sign-in should also include the authorization issue. The SAML assertion which includes
-users' attributes can be used on the serrvice side to achiece attribute-based authorization.
-Therefore providing plugin on the service side to enforce access control using SAML attribute
-is considered one part of the future work.
+Although only authentication issue has been discussed in this paper, we know
+single sign-in should also include the authorization issue. The SAML assertion which
+includes users' attributes can be used on the service side to achieve attribute-based
+authorization. Therefore providing plug-in on the service side to enforce access control using
+SAML attribute is considered one part of the future work.
 
-SimpleSAMLphp~\cite{simplesamllink} provides another implementation of SAML2 SSO profile, and it 
-has been widely deployed around nordic countries for identity federation. Hence integrating
-ARC based Grid applications with simpleSAMLphp is considered as another of future work.
+SimpleSAMLphp~\cite{simplesamllink} provides another implementation of SAML2 SSO
+profile, and it has been widely deployed around Nordic countries for identity federation. Hence
+integrating ARC based Grid applications with simpleSAMLphp is considered as another of
+future work.
 
@@
 
-\begin{thebibliography}{1}
+\begin{thebibliography}{99}
 
@@
 Computational Grids, ACM Conference on Computers and Security, 1998, 83-91.
-
-\bibitem{AlfieriR05}
-R. Alfieri, R. Cecchini, V. Ciaschini, L. dell’Agnello, A. Frohner, K. Lorentey, 
-and F. Spataro. From gridmap-file to voms: managing authorization in a Grid environment. 
-Future Generation Comp. Syst., 21(4), 549-558 (2005)
-
-\bibitem{RFC3821link}
-RFC 3821- An Internet Attribute Certificate Profile for Authorization. http://www.faqs.org/rfcs/rfc3281.html
-
-\bibitem{GTlink}
-Globus Toolkit. http://www.globus.org/toolkit/
-
-\bibitem{gLitelink}
-gLite: lightweight middleware for Grid computing. http://glite.web.cern.ch
-
-\bibitem{ARClink}
-Advanced Resource Connector. http://www.nordugrid.org/middleware/
-
-\bibitem{WSSeclink}
-OASIS Web Services Security. www.oasis-open.org/committees/wss/
-
-\bibitem{SAMLlink}
-OASIS Security Assertion Markup Languages (SAML). www.oasis-open.org/committees/security/
 
@@
 
+\bibitem{SAMLlink}
+OASIS Security Assertion Markup Languages (SAML).
+www.oasis-open.org/committees/security/
+
+\bibitem{SAMLprofiles}
+J. Hughes et al., Profiles for the OASIS Security Assertion Markup
+Language (SAML) V2.0, OASIS Standard, 2005
+
+\bibitem{myproxy}
+MyProxy Credential Management Service. http://grid.ncsa.uiuc.edu/myproxy/
+
 \bibitem{eduSchemalink}
 eduPerson and eduOrg Object shema. http://middleware.internet2.edu/eduperson/
+
+% \bibitem{AlfieriR05}
+% R. Alfieri, R. Cecchini, V. Ciaschini, L. dell'Agnello, A. Frohner, K. Lorentey, 
+% and F. Spataro. From gridmap-file to voms: managing authorization in a Grid environment. 
+% Future Generation Comp. Syst., 21(4), 549-558 (2005)
+% 
+% \bibitem{RFC3821link}
+% RFC 3821- An Internet Attribute Certificate Profile for Authorization. http://www.faqs.org/rfcs/rfc3281.html
+% 
+% \bibitem{GTlink}
+% Globus Toolkit. http://www.globus.org/toolkit/
+% 
+% \bibitem{gLitelink}
+% gLite: lightweight middleware for Grid computing. http://glite.web.cern.ch
+% 
+% \bibitem{ARClink}
+% Advanced Resource Connector. http://www.nordugrid.org/middleware/
+% 
+% \bibitem{WSSeclink}
+% OASIS Web Services Security. www.oasis-open.org/committees/wss/
+% 
+\bibitem{J.Watt06}
+J. Watt, O. Ajayi, J. Jiang, J. Koetsier, R.O. Sinnott. A Shibboleth-Protected 
+Privilege Management Infrastructure for e-Science Education. in proceeding of Sixth 
+IEEE International Symposium on Cluster Computing and the Grid (CCGRID'06), May 2006,
+Singapore.
 
@@
 Grid Computing, September 2006, Barcelona.
-
-\bibitem{J.Watt06}
-J. Watt, O. Ajayi, J. Jiang, J. Koetsier, R.O. Sinnott. A Shibboleth-Protected 
-Privilege Management Infrastructure for e-Science Education. in proceeding of Sixth 
-IEEE International Symposium on Cluster Computing and the Grid (CCGRID'06), May 2006,
-Singapore.
 

### 2009-04-30 doc/tech_doc/storage_system/arc-storage-documentation.tex
Implemented Marek's comments

@@
 When a Bartender wants to create a new replica of a file (or the first replica of a new file) it should find available Shepherd services.
-The Bartender uses the list of Shepherd services registered to the Librarian.
+The Bartender uses the list of Shepherd services registered to the Librarian, and chooses the locations of the replicas randomly from this list according a uniform distribution.
 
@@
 
+The ARC DMC allows Grid jobs to access Chelonia directly.
+As long as A-REX, the job execution service of ARC, and ARC DMC are installed on a site, files can be both downloaded and uploaded by specifying the corresponding URL's in the job description.
+In this case, the Bartender URL needs to be embedded in the URL as a URL option.
+For example, if a job requires an input file \verb!/user/me/input.dat!, and a Bartender \verb!https://storage/Bartender!
+the URL specified in the job description will be as follows:
+\begin{verbatim}
+arc:///user/me/input.dat?BartenderURL=https://storage/Bartender
+\end{verbatim}
 % section grid_integration (end)

### 2009-04-30 doc/sec/papers/GRID09/paper.tex
a little change to the paper

@@
 describes the solution of using federated identity for 
-Grid authentication by using \texttt{SAML2SSO} profile; Section~\ref{sec:fedidtoX509} describes
+Grid authentication by using \texttt{SAML2 Web Browser SSO Profile (SAML2SSO)} profile; 
+Section~\ref{sec:fedidtoX509} describes
 how to obtain X.509 credentials based on federated identity, and how to use such for 
@@
 also simplifies work on the middleware level: for example, it makes possible to implement another 
-communication protocol or authentication mechanism. A system administrator can easily 
+communication protocol or authentication mechanism. Meanwhile, a service administrator can easily 
 configure and deploy the middleware and application for different kinds of requirements 
@@
 \includegraphics[width=1.0\columnwidth]{SAML2SSO_UML.pdf}
-\caption{Sequence diagram of SAML2.0 SSO profile integration in SOAP invocation 
+\caption{Sequence diagram of the integration of SAML2.0 SSO profile in SOAP invocation 
 between ARC WS Client and Web Service}
@@
 into the SOAP invocation between ARC WS Client and Web Service. Steps 1 to 5 in
-Figure~\ref{fig:SAML2SSOUML} are similar to those steps depicted in the Web Browser SSO profile, with
+Figure~\ref{fig:SAML2SSOUML} are similar to those steps depicted in the SAML2 Web Browser SSO profile, with
 the difference that we specify the way how does the service provider determine which identity
@@
 IdP with the last successfully authenticated session's identity rather than feed IdP with
-the username and password again. So the user (or the client on behalf of that user) can travel across multiple
+the username and password again. So the user (or the client on behalf of that user) can access multiple
 security domains by only providing his name and password once, which is also a characteristic of
@@
 Figure~\ref{fig:SAML2SSOUML}) including an X.509 request, and the SOAP response (the message of step 10 
-of Figure~\ref{fig:SAML2SSOUML}) including a signed X.509 certificate and a CA certificate that is used 
+of Figure~\ref{fig:SAML2SSOUML}) including an X.509 certificate and a CA certificate that is used 
 to sign this X.509 certificate. In detail, the SLCS client firstly accomplishes the SAML2SSO profile, 
@@
 certificate extension. The SLCS client then gets the SOAP response with the X.509
-certificate embedded.
+certificate and the CA certificate embedded.
 
@@
 \includegraphics[width=1.0\columnwidth]{Delegation_UML.pdf}
-\caption{Sequence diagram of the usage of delegation service}
+\caption{Sequence diagram of the usage of delegation service in SOAP invocation 
+between ARC WS Client and Web Service}
 \label{fig:DelegUML}

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
minor changes

@@
 \label{sec:fedidtoX509}
-Many widely used Grid middlewares are based on GSI which requires mutual X.509
+The integration of SAML2SSO profile in Section~\ref{sec:intergrationSAML2SSO}
+provides complete single sign-on in the Grid middleware level without requiring X.509
+certificate on the client side.
+
+However, many widely used Grid middlewares are based on GSI which requires mutual X.509
 authentication. Meanwhile, Web Service based Grid applications mostly require client
@@
 The following sections evaluate the performance of the HED versus Axis2/C, as 
-well as the performance related to the SAML2SSO profile, short-lived credential
-service, and delegation service.
+well as the performance related to the integration of SAML2SSO profile, 
+short-lived credential service, and delegation service.
 
@@
 complete comparison between HED and all the other SOAP implementations, so we use Axis2/C (v1.5.0) to
-demonstrate the performance HED.
+demonstrate the performance of HED.
 
@@
 Axis2/C. The ``echoString'' provides a simple SOAP portType which will output the same value
-as it gets from input. For the client side, we developed a client by using the client API of ARC, to
+as it gets from input. For the client side, we develope a client by using the client API of ARC, to
 invoke ``echoString'' services from Axis2/C and HED. Service invocation just sends a short
@@
 time is computed by counting the average value of the response time for each invocation.
-Throughput is computed by counting number of invocations per second.
+Throughput is computed by counting number of invocations per second or per minute.
 

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
More plot; tex change comes soon

@@
 \begin{figure}
-\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10.pdf}
+\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10_nosso.pdf}
 \caption{Performance of HED echoString services with SAMLSSO profile: 1 service
 and 10 clients}
-\label{fig:SAML2SSO_thread10}
+\label{fig:SAML2SSO_thread10_nosso}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10_nosso_and_sso.pdf}
+\caption{Performance of HED echoString services with SAMLSSO profile: 1 service
+and 10 clients}
+\label{fig:SAML2SSO_thread10_nosso_and_sso}
 \end{figure}
@@
 
-Figure \ref{fig:SAML2SSO_thread10} depicts the average response time and
+Figure \ref{fig:SAML2SSO_thread10_nosso} depicts the average response time and
 throughputs for 10 clients running simultaneously against the ``echoString'' service enhanced with SAML2SSO

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
More text

@@
 \includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10_nosso.pdf}
-\caption{Performance of HED echoString services with SAMLSSO profile: 1 service
-and 10 clients}
+\caption{Performance of HED echoString services with SAML2SSO profile: 1 service
+and 10 clients; single sign-on is disabled; session used and session not re-used are compared}
 \label{fig:SAML2SSO_thread10_nosso}
@@
 \includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10_nosso_and_sso.pdf}
-\caption{Performance of HED echoString services with SAMLSSO profile: 1 service
-and 10 clients}
+\caption{Performance of HED echoString services with SAML2SSO profile: 1 service
+and 10 clients; session not re-used; single sign-on enabled and single sign-on disabled are compared}
 \label{fig:SAML2SSO_thread10_nosso_and_sso}
@@
 \includegraphics[width=0.9\columnwidth]{SAML2SSO_thread_all.pdf}
-\caption{Performance of HED echoString services with SAMLSSO profile:
-performance changes with client numbers; session non-reused }
+\caption{Performance of HED echoString services with SAML2SSO profile:
+performance changes with client numbers; session not re-used; single sign-on enabled and 
+single sign-on disabled are compared}
 \label{fig:SAML2SSO_thread_all}
@@
 
-Figure \ref{fig:SAML2SSO_thread10_nosso} depicts the average response time and
-throughputs for 10 clients running simultaneously against the ``echoString'' service enhanced with SAML2SSO
-profile. Since the success of authentication from IdP can be regarded as valid for the whole
-session between client and service, the client with ``Session reused'' need to accomplish the SAML2SSO
+We consider two situations in terms of the authenticating between user agent (client) and IdP: 
+SSO enabled, and SSO disabled. ``SSO disabled'' means the successful authentication
+session between user agent and IdP is not regarded as valid for the subsequent authentication, and
+user is required to authenticate by inputing his username and password on each request.
+``SSO enabled'' means the successful authentication session is regarded as valid for the 
+subsequent authentication, in a short period (we setup this period as 30 minutes).
+
+Considering the success of authentication from IdP can be regarded as valid for the whole
+session between client and service, we also consider two situation in terms of the 
+using of authentication result between client and service: session re-used, and session non-reused.
+The client with ``Session reused'' need to accomplish the SAML2SSO
 profile only once for all SOAP invocations on this client, while the client with ``Session
@@
 
+Figure \ref{fig:SAML2SSO_thread10_nosso} depicts the average response time and
+throughputs for 10 clients running simultaneously against the ``echoString'' service enhanced with SAML2SSO
+profile, with single sign-on disabled.
+
 The average response time of ``Session non-reused'' experiments changes little
-with the execution duration, and the value is about 3000ms; while the throughput is around 200
+with the execution duration, and the value is about 3000ms; while the throughput is around 215
 requests per minutes. Considering that the average response time in case of ``Session non-reused'' and
@@
 the time consumption caused by the following hundreds of pure SOAP invocations on the
-same client.
+same client. So besides improving the performance of SAML2SSO profile, reusing 
+the session as longer as possbile is also a way to improve the performance of 
+Web Service invokation with SAML2SSO profile integrated as authentication approach.
+
+Figure \ref{fig:SAML2SSO_thread10_nosso_and_sso} depicts the performance comparison between singled 
+sign-on disabled and singled sign-on enabled, for 10 clients running simultaneously 
+against the ``echoString'' service enhanced with SAML2SSO profile, with session not re-used.
+
+The average response time of ``SSO enabled'' is about 1700ms, while the throughput is about
+340 requests per minutes. The average response time of ``SSO disabled'' is about 3000ms, while 
+the through put is about 215 requests per minutes. The comparison shows that the performance 
+of ``SSO enabled'' is better than that of ``SSO disabled'', but still on the same quantitative
+level.
 
@@
 echoString services that are enhanced with SAML2SSO profile, with client number being changed. In
-this set of experiments, the session is not reused, and secure communication without client certificate
-authentication is used. The results show the average response time increases linearly with the client
-number, while the throughput increases initially and then keeps in a fixed value (around 215
-requests per minute) when the client number is bigger than 4. When the client number is four, the
-response time is around one second, while the throughput of the service is 215 requests per
-minutes which is almost the fixed value.
-We speculate that in order to get acceptable response time, there should be less
-than four (including four) concurrent clients if they need to continuously participate one SAML2SSO
-profile.
+this set of experiments, the session is not reused, and single sign-on enabled and singled sign-on
+disabled are compared.
+
+For the situation of singled sign-on disabled, the average response time 
+increases linearly with the client number, while the throughput increases initially and then 
+keeps in a fixed value (around 215 requests per minute) when the client number is bigger 
+than 4. When the client number is four, the response time is around one second, while the 
+throughput of the service is 215 requests per minutes which is almost the fixed value.
+
+For the situation of singled sign-on enabled, the average response time 
+also increases linearly with the client number, while the throughput also increases initially and then 
+keeps in a fixed value (around 340 requests per minute) when the client number is bigger 
+than 5. When the client number is five, the response time is around 880ms, while the 
+throughput of the service is 340 requests per minutes which is almost the fixed value.
+
+We speculate that in order to get acceptable response time, for the case of single sign-on disabled, 
+there should be less than four (including four) concurrent clients if they need to continuously 
+participate one SAML2SSO profile; for the case of single sign-on enabled, 
+there should be less than five (including five) concurrent clients if they need to continuously 
+participate one SAML2SSO profile.
 
@@
 TLS is 13000, and the concurrent client number when arriving fixed throughput is 30, the
-performance of ``echoString'' service enhanced with SAML2SSO profile is much worse, with the
-values being 215 and 4 correspondingly.
+performance of ``echoString'' service enhanced with SAML2SSO profile is much worse, both 
+with single sign-on disabled and enabled, with the values being 215, 4, and 340, 5, correspondingly.
 
@@
 Therefore, if we suppose a user could access grid once per minute, in our test environment, we
-can claim that around 215 clients randomly can be supported with acceptable performance.
+can claim that for the case of of single sign-on disabled, around 215 clients randomly can be 
+supported with acceptable performance, while for the case of single sign-on enabled, around 340 
+clients randomly can be supported with acceptable performance.
 
@@
 \includegraphics[width=0.9\columnwidth]{SLCS_thread_all.pdf}
-\caption{Performance of SLCS service: performance changes with client numbers}
+\caption{Performance of SLCS service: performance changes with client numbers; single sign-on disabled}
 \label{fig:SLCS_thread_all}
@@
 service and client run on the same set of machine as section~\ref{sec:perhedandaxis}.
+
+Considering the time interval of the usage of SLCS client should be similar to the lifetime of SLCS
+credential which is 12 hours by default, and the lifetime of successful authentication session is 
+30 minutes by default, enabling single sign-on is not useful in real deployment because the successful 
+authentication session will be expired before the next time the SLCS client is used, therefore, 
+we do not enable single sign-on when testing the SLCS service.
 
@@
 \section*{Acknowledgment}
-
-The authors would like to thank...
+We would thank all of the ARC middleware developers for
+their intelligent work. We wish to thank Oxana Smirnova for
+vital comments and proof-reading.
 

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
Done with proof-reading

--- 2009-05-01 doc/sec/papers/GRID09/paper.tex
@@
 Axis2/C. The ``echoString'' provides a simple SOAP portType which will output the same value
-as it gets from input. For the client side, we develope a client by using the client API of ARC, to
+as it gets from input. For the client side, we develop a client by using the client API of ARC, to
 invoke ``echoString'' services from Axis2/C and HED. Service invocation just sends a short
@@
 Figure~\ref{fig:HED2Axis_thread10} shows results of performance comparison with
-ten 
-clients being launched.
+ten clients being launched.
 Two types of experiments are shown: session re-used and session not re-used.
@@
 
-Figure \ref{fig:SAML2SSO_thread10_nosso} depicts the average response time and
-throughputs for 10 clients running simultaneously against the ``echoString'' service enhanced with SAML2SSO
-profile. Since the success of authentication from IdP can be regarded as valid for the whole
-session between client and service, the client with ``Session reused'' need to accomplish the SAML2SSO
+We consider two scenarios of authenticating between user agent (client) and IdP: 
+SSO enabled, and SSO disabled. ``SSO disabled'' means the successful authentication
+session between user agent and IdP is not regarded as valid for the subsequent authentication, and
+user is required to authenticate by inputing his username and password on each request.
+``SSO enabled'' means the successful authentication session is regarded as valid for the 
+subsequent authentication during a short period (we set up this period at 30 minutes).
+
+Since the success of authentication from IdP can be regarded as valid for the whole
+session between client and service, we also consider two scenaros in 
+usage of the authentication result between client and service: session re-used, and session non-reused.
+The client with ``Session reused'' need to accomplish the SAML2SSO
 profile only once for all SOAP invocations on this client, while the client with ``Session
@@
 
+Figure~\ref{fig:SAML2SSO_thread10_nosso} shows the average response time and
+throughputs for 10 clients running simultaneously against the ``echoString'' service enhanced with SAML2SSO
+profile, with single sign-on disabled.
+
 The average response time of ``Session non-reused'' experiments changes little
-with the execution duration, and the value is about 3000ms; while the throughput is around 200
-requests per minutes. Considering that the average response time in case of ``Session non-reused'' and
-pure SOAP invocation is around 15ms when client number is 10 (see section \ref{sec:perhedandaxis}),
-therefore we can calculate that the consumption of time for SAML2SSO profile is about 99.5\% of
+with the execution duration, and the value is about 3000ms; while the throughput is around 215
+requests per minute. Given that the average response time in case of ``Session non-reused'' and
+pure SOAP invocation is around 15ms for 10 clients (see Section~\ref{sec:perhedandaxis}),
+we can evaluate that the consumption of time for SAML2SSO profile is about 99.5\% of
 the whole SOAP invocation with SAML2SSO profile integrated.
 
-Deeper investigation about recording the time of each step of SAML2SSO profile
+Deeper investigation involving recording the time for each step of SAML2SSO profile
 supports the above observation. The average time cost for getting an authentication request
@@
 In order to improve the performance of SAML2SSO profile, optimizing the process
-between client and IdP is the correct direction, and probably one of the easiest way is
+between client and IdP is the correct direction, and probably one of the easiest ways is
 to enhance the hardware configuration.
@@
 is about 82 ms, which is almost as short as the value in the pure SOAP invocation experiment in
-section \ref{sec:perhedandaxis}. The result is reasonable because the time consumption
+Section~\ref{sec:perhedandaxis}. The result is reasonable because the time consumption
 caused by only once of SAML2SSO execution on one client is relatively small comparing with
 the time consumption caused by the following hundreds of pure SOAP invocations on the
-same client.
+same client. So besides improving the performance of SAML2SSO profile, reusing 
+the session as long as possbile is also a way to improve the performance of 
+Web Service invocation with SAML2SSO profile integrated as authentication approach.
 
-Figure~\ref{fig:SAML2SSO_thread_all} depicts the performance results for
+Figure~\ref{fig:SAML2SSO_thread10_nosso_and_sso} depicts performance comparison between single 
+sign-on disabled and single sign-on enabled scenarios, for 10 clients running simultaneously 
+against the ``echoString'' service enhanced with the SAML2SSO profile, with session not re-used.
+
+The average response time in the ``SSO enabled'' case is about 1700ms, while the throughput is about
+340 requests per minutes. The average response time in the ``SSO disabled'' case is about 3000ms, while 
+the throughput is about 215 requests per minutes. The comparison shows that the performance 
+of ``SSO enabled'' is better than that of ``SSO disabled'', but still on the same quantitative
+level.
+
+Figure~\ref{fig:SAML2SSO_thread_all} illustrates performance results for
 echoString services that are enhanced with SAML2SSO profile, with client number being changed. In
-this set of experiments, the session is not reused, and secure communication without client certificate
-authentication is used. The results show the average response time increases linearly with the client
-number, while the throughput increases initially and then keeps in a fixed value (around 215
-requests per minute) when the client number is bigger than 4. When the client number is four, the
-response time is around one second, while the throughput of the service is 215 requests per
-minutes which is almost the fixed value.
-We speculate that in order to get acceptable response time, there should be less
-than four (including four) concurrent clients if they need to continuously participate one SAML2SSO
-profile.
+this set of experiments, the session is not re-used, and single sign-on enabled and single sign-on
+disabled scenarios are compared.
 
-Considering that the throughput of the ``echoString'' service enhanced with mutual
-TLS is 13000, and the concurrent client number when arriving fixed throughput is 30, the
-performance of ``echoString'' service enhanced with SAML2SSO profile is much worse, with the
-values being 215 and 4 correspondingly.
+For the situation when single sign-on is disabled, the average response time 
+increases linearly with the number of clients, while the throughput increases initially and then 
+stays constant (around 215 requests per minute) when the client number is bigger 
+than 4. When the client number is 4, the response time is around one second, while the 
+throughput of the service is 215 requests per minute, which is almost constant.
 
-However for some typical grid application scenarios, normally the client does
+For the situation of single sign-on enabled, the average response time 
+also increases linearly with the client number, while the throughput also increases initially and then 
+saturates at around 340 requests per minute when the client number is bigger 
+than 5. When the client number is five, the response time is around 880ms, while the 
+throughput of the service is 340 requests per minute, which is almost constant.
+
+We speculate that in order to get acceptable response time, for the case when single sign-on is disabled, 
+there should be less than four (including four) concurrent clients if they need to continuously 
+participate one SAML2SSO profile; for the case when single sign-on is enabled, 
+there should be less than five (including five) concurrent clients if they need to continuously 
+participate one SAML2SSO profile.
+
+Given that the throughput of the ``echoString'' service enhanced with mutual
+TLS is 13000, and the concurrent client number when throughput reaches saturation is 30, the
+performance of ``echoString'' service enhanced with SAML2SSO profile is much worse, both 
+with single sign-on disabled and enabled, with the values being 215, 4, and 340, 5, correspondingly.
+
+However, for some typical grid application scenarios, the client normally does
 not need to continuously accomplish SAML2SSO profile. For instance, a user could submit a
 job which needs a period of time (normally this period is much longer than a few
-seconds) to execute, which launches an execution of SAML2SSO profile; 
-afterwards, this user could query the information as well as the result about
-this job, which launches another execution of SAML2SSO profile. On the other hand,
-statistically, users will most probably not access grid system in a concurrent way, rather than a random way.
-Therefore, if we suppose a user could access grid once per minute, in our test environment, we
-can claim that around 215 clients randomly can be supported with acceptable performance.
+seconds) to execute, which launches an execution of SAML2SSO profile. 
+This user could afterwards query the information as well as the result of
+this job, which would launch another execution of SAML2SSO profile. On the other hand,
+statistically, users will most probably not access grid system in a concurrent way, rather than in a random way.
+Therefore, if we suppose that a user could access grid once per minute, in our test environment, we
+can claim that for the case of disabled single sign-on, around 215 clients randomly can be 
+supported with acceptable performance, while for the case of single sign-on enabled, around 340 
+clients randomly can be supported with acceptable performance.
 
@@
 
-Figure \ref{fig:SLCS_thread_all} is the performance about the SLCS service.
-Since the SLCS is based on the SAML2SSO profile, the response time should be the accumulation of
-time consumption about the SAML2SSO processing and the SLCS processing. In this experiment, SLCS
-service and client run on the same set of machine as section~\ref{sec:perhedandaxis}.
+Figure~\ref{fig:SLCS_thread_all} shows performance of the SLCS service.
+Since the SLCS is based on the SAML2SSO profile, the response time should be the combination of
+time consumption for the SAML2SSO processing and the SLCS processing. In this experiment, SLCS
+service and client run on the same set of machine as in Section~\ref{sec:perhedandaxis}.
+
+Since the time interval of the usage of an SLCS client should be similar to the lifetime of the SLCS
+credential, which is 12 hours by default, and the lifetime of a successful authentication session is 
+30 minutes by default, enabling single sign-on is not useful in a real deployment because a successful 
+authentication session will expire before the next time the SLCS client is used. Therefore, 
+we do not enable single sign-on when testing the SLCS service.
 
@@
 as the throughput of ``echoString'' service enhanced with SAML2SSO profile, with the value being
-around 215 requests per minute; and the average response time resembles the value of
-``echoString'' service as well, with the value being increased for about 80ms. The increased time
+around 215 requests per minute; and the average response time is close to the value of
+``echoString'' service as well, with the value being increased to about 80ms. The increased time
 consumption is caused by the creating of X.509 request (1024-bit RSA key pair generation) on the client
@@
 In the last experiment, we evaluate how does the usage of delegation service
-impact on the performance of Web Service invocation. We deployed a ``echoString'' service
+affect performance of Web Service invocation. We deployed an ``echoString'' service
 plugged with delegation security handler on a machine equipped with 3GHz Intel Pentium dual-core CPU and
-2GB memory, Red Hat Enterprise Linux 5, and also configure a client together with delegation
+2GB memory, Red Hat Enterprise Linux 5, and also configured a client together with delegation
 security handler plugged on a machine equipped with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat
-Enterprise Linux 5. Aside from ``echoString'' service and client, a delegation service is deployed on a machine
+Enterprise Linux 5. Aside from ``echoString'' service and client, a delegation service was deployed on a machine
 equipped with 3GHz Intel Pentium dual-core CPU and 2GB memory, Red Hat Enterprise Linux 5.
@@
 \caption{Performance of HED echoString service with credential delegation:
-performance changes with client numbers; echoString service contacts delegation service upon
+performance changes with the number of clients; echoString service contacts delegation service upon
 each service invocation from client}
@@
 
-In the experiment, the client launches a new credential delegation for each
+In this experiment, the client launches a new credential delegation for each
 service invocation, and ``echoString'' service acquires delegated credential from delegation service
@@
 throughput of the ``echoString'' service with delegation security handler embedded. The
-average response time increases linearly with the client number, while the throughput starts to keep at
-fixed value (around 400 request per minutes) when the client number is 8, after initially
-increasing.
+average response time increases linearly with the client number, while the throughput starts to saturate
+at around 400 request per minutes when the client number reaches 8.
 We conclude that if the client number is bigger than eight, in our test
@@
 \label{sec:relatedwork}
-There are other researches that utilize Shibboleth for protecting usage of Grid
-resources. The work in ~\cite{J.Watt06} uses Shibboleth to protect a GridSphere portal and
+
+There are other efforts that utilize Shibboleth for protecting usage of Grid
+resources. The work of Watt and colleagues~\cite{J.Watt06} uses Shibboleth to protect a GridSphere portal and
 then indirectly protect Grid services which are deployed in this GridSphere portal.
-Another work in ~\cite{R.O.Sinnott06} also uses Shibboleth to protect Grid portal,
-moreover, it maintains a few common used server certificates which can then be used to access
-Grid resources after users have successfully authenticated themselves at Grid portal
+Another work of Sinott and colleagues~\cite{R.O.Sinnott06} also uses Shibboleth to protect a Grid portal,
+moreover, it maintains a few commonly used server certificates which can then be used to access
+Grid resources after users have successfully authenticated themselves at a Grid portal
 through Shibboleth. Unlike these two solutions that use Shibboleth to protect Grid
 portal, the solution in this paper uses Shibboleth or similar standards-based identity federation to
-directly protect Grid resource/services, which provide possibility for those Grid applications
-that are directly based on Grid interface (e.g., applications that invokes SOAP operations of
+directly protect Grid resource/services, which provides possibility for those Grid applications
+that are directly based on Grid interface (e.g., applications that invoke SOAP operations of
 remote Grid services), to benefit from identity federation.
 
-GridShib~\cite{T.Scavo07,VWelch05} provides one solution that users can use
-X.509 certificate to authenticate to Shibboleth IdP, get back SAML assertion and embed
-it in the proxy certificate. The Grid service will then use this SAML assertion for
+GridShib~\cite{T.Scavo07,VWelch05} provides a solution where users can use
+X.509 certificates to authenticate to Shibboleth IdP, get back SAML assertion and embed
+it in the proxy certificates. The Grid service will then use this SAML assertion for
 access control. Unlike the work described in this paper, GridShib still requires users to
@@
 username/password rather than X.509 certificate, and issuing short-lived X.509 credential. The
-second solution is similar to the short-lived credential service in our solution, while our
-solution also provides the option for users directly authenticating to Grid services via
+second solution is similar to the short-lived credential service in our work, while our
+solution also provides the option for users to directly authenticate to Grid services via
 SAML2.0 SSO profile using their username/password.
@@
 The advantage of our work comparing to the work of SWITCH is that 
-our SLCS service is based on the Web Services standard, thus that it is easier to
+our SLCS service is based on the Web Services standard, thus it is easier to
 achieve interoperability.
@@
 users' federated identity instead of X.509 mutual authentication for authentication,
-which we believe can make further uptake of Grid resources as well as Grid middleware
-by wider scale of users.
+which we believe can enhance uptake of Grid technlogies by resources owners as well as
+by wider range of users.
 
 We implement the SAML2 SSO profile in the ARC Grid middleware, so that the Grid
-applications that developed by using the application interface provided by this middleware
-can easily be switched to benefit from federated identity based authentication. 
-The performance evaluation shows even though the introducing of SAML2 SSO
-profile causes much performance downgrade, this downgrade can be avoided by reusing the
+applications developed by using the application interface provided by this middleware
+can easily be adapted to benefit from the federated identity based authentication. 
+The performance evaluation shows that even though introduction of SAML2 SSO
+profile causes notable performance downgrade, this downgrade can be avoided by reusing the
 authentication result for multiple SOAP invocations.
@@
 X.509 certificate on the client side, we also implement one Web Service which can be
-used for obtaining an X.509 credential based on federated identity, and the other Web
+used for obtaining an X.509 credential based on federated identity, and another Web
 Service which can be used for delegating X.509 credential. We believe that the
-evaluated performance is enough for creating X.509 certificate and proxy certificate,
+observed performance is sufficient for creating X.509 certificates and proxy certificates,
 considering that the certificates should have a relatively much longer lifetime comparing to the
@@
 
-Although only authentication issue has been discussed in this paper, we know
-single sign-in should also include the authorization issue. The SAML assertion which
+Although only the authentication issue has been discussed in this paper, we know that
+single sign-in should also include the authorization aspect. The SAML assertion which
 includes users' attributes can be used on the service side to achieve attribute-based
 authorization. Therefore providing plug-in on the service side to enforce access control using
-SAML attribute is considered one part of the future work.
+SAML attribute is considered an important part of the future work.
 
 SimpleSAMLphp~\cite{simplesamllink} provides another implementation of SAML2 SSO
-profile, and it has been widely deployed around Nordic countries for identity federation. Hence
+profile, and it has been widely deployed e.g. around Nordic countries for identity federation. Hence
 integrating ARC based Grid applications with simpleSAMLphp is considered as another of

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
little more text

@@
 which is the base of the security related web services described in this paper, it is useful to
-compare the performance of HED and other SOAP implementations. Although there are many SOAP
-implementations such as AxisJava, gSOAP, XSOAP4, etc.., it is not the goal of this paper to do a
-complete comparison between HED and all the other SOAP implementations, so we use Axis2/C (v1.5.0) to
-demonstrate the performance of HED.
+compare the performance of HED and other SOAP implementations. Here we compare 
+the performance of HED and Axis2/C (v1.5.0), because we need to demostrate that the performance of HED
+is at least comparable to other implementations. We do not compare the performance of HED 
+with other SOAP implementations such as AxisJava, gSOAP, XSOAP4, etc., because a complete 
+performance comparison between HED and all the other SOAP implementations is not the main 
+goal of this paper to achieve, and a partial performance comparison is enough. Another reason
+of choosing Axis2/C is because it is also implemented in C/C++ programming language.
 
@@
 machine equipped with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat Enterprise Linux 5. The
-Axis2/C service is run using the httpd module on Apache Http Server version 2.2.11 instance. We use all the
-default configurations in the Apache Http Server.
+Axis2/C service is run using the httpd module on Apache Http Server version 2.2.11 instance. 
+We use all the default configurations in the Apache Http Server, and complie Axis2/C
+with default options except specifying the ``--with-apache2'' option.
 

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
little change again

@@
 Since the implementation of HED includes an alternative SOAP implementation
-which is the base of the security related web services described in this paper, it is useful to
+which is the base of the security related Grid/Web Services described in this paper, it is useful to
 compare the performance of HED and other SOAP implementations. Here we compare 
@@
 
-However, for some typical grid application scenarios, the client normally does
+However, for some typical Grid application scenarios, the client normally does
 not need to continuously accomplish SAML2SSO profile. For instance, a user could submit a
@@
 this job, which would launch another execution of SAML2SSO profile. On the other hand,
-statistically, users will most probably not access grid system in a concurrent way, rather than in a random way.
-Therefore, if we suppose that a user could access grid once per minute, in our test environment, we
+statistically, users will most probably not access Grid system in a concurrent way, rather than in a random way.
+Therefore, if we suppose that a user could access Grid once per minute, in our test environment, we
 can claim that for the case of disabled single sign-on, around 215 clients randomly can be 

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
little text again

@@
 the Shibboleth IdP implementation on a machine equipped with 2.8GHz Intel Pentium CPU and 1GB
-memory, Red Hat Enterprise Linux 5.
+memory, Red Hat Enterprise Linux 5. We run Shibboleth IdP (2.0.0) on Apache Tomcat(6.0.18), and
+configure the JVM memory for Tomcat by setting ``CATALINA\_OPTS=-Xms16m -Xmx256m''. The Java
+version is 1.6.0_07.
 
@@
 In order to improve the performance of SAML2SSO profile, optimizing the process
-between client and IdP is the correct direction, and probably one of the easiest ways is
-to enhance the hardware configuration.
+between client and IdP is the correct direction. Considerting most of the performance 
+characteristics for the Shibboleth IdP are determined by the Servlet Container (that is Tomcat 
+in our test environment) and Java VM running it, we also increase the JVM memory for Tomcat 
+by setting ``CATALINA\_OPTS=-Xms256m -Xmx1024m'', and the performance result after setting
+shows little improvement.
+Therefore we conclude that the easiest way is to enhance the hardware configuration.
 

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
little change

@@
 configure the JVM memory for Tomcat by setting ``CATALINA\_OPTS=-Xms16m -Xmx256m''. The Java
-version is 1.6.0_07.
+version is 1.6.0\_07.
 
@@
 In order to improve the performance of SAML2SSO profile, optimizing the process
-between client and IdP is the correct direction. Considerting most of the performance 
+between client and IdP is the correct direction. Considering most of the performance 
 characteristics for the Shibboleth IdP are determined by the Servlet Container (that is Tomcat 
@@
 by setting ``CATALINA\_OPTS=-Xms256m -Xmx1024m'', and the performance result after setting
-shows little improvement.
+shows little improvement in terms of throughput and reponse time.
 Therefore we conclude that the easiest way is to enhance the hardware configuration.
@@
 consumption is caused by the creating of X.509 request (1024-bit RSA key pair generation) on the client
-side.
+side. We conclude that the performance of SLCS service is mostly influenced by the SAML2SSO profile,
+while the performance consumption for RSA key pair generation is very little comparing with the 
+consumption for SAML2SSO profile.
 
@@
 environment, there is no throughput benefit can be achieved by adding clients.
+
+In the delegation process (showing in Figure~\ref{fig:DelegUML}), the delegation service needs
+to generate RSA pair key (step 2 in Figure~\ref{fig:DelegUML}), and the ``echoString'' service
+needs to generate RSA pair key (step 10 in Figure~\ref{fig:DelegUML}). Also the delegation service
+needs to sign the X.509 request from ``echoString'' service.
+We speculate the throughput of ``echoString'' service that configured with credential delegation
+is mainly influenced by these cryptographic operations.
 

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
little change again

@@
 with the client numbers initially, however, it saturates at around 13000 requests per
-minutes when the number  of client number is higher than 30. On the other hand, the throughput of service in
+minute when the number  of client number is higher than 30. On the other hand, the throughput of service in
 Axis2/C saturates already when the number of clients reaches 10, with
@@
 
-We conclude that the HED can provide better performance than Axis2/C, in both
-cases of secure communication being enabled and disabled, especially when the number of
-simultaneous clients is bigger.
+We conclude that the HED can provide comparable performance 
+as Axis2/C, in both cases of secure communication being enabled and disabled.
 

### 2009-05-05 doc/tech_doc/jura/jura-tech-doc.tex
Documenting JURA

--- 2009-04-26 doc/jura/jura-tech-doc.tex
@@
 \centering{{{\includegraphics[width=0.9\textwidth]{components.pdf}}}
-\caption{\label{fig:components}Components participating in accounting.} }
+\caption{\label{fig:components}The usage reporting mechanism.} }
 \end{figure}
@@
 
-The command line format is a subset of that of logger:
+The command line format is a subset of that of logger: %TODO make it so!
 
@@
 and another one after the job finishes, and possibly others at each
-start and stop event.
+start and stop event. Reporting destination URLs are acquired from
+``\textit{jobreport=}'' lines in the A-REX configuration file, and a
+limited number of additional destinations can be supplied by the
+submitter in the job description.
 
@@
 \subsection{Accessing LUTS}
+\label{accessing}
 After the necessary URs to be inserted are filled and valid, they are
-submitted to the accounting service through its insertion
-interface. The service in particular, SGAS LUTS has a simple custom
-web service interface loosely based on WS-ResourceProperties.
+submitted to the accounting service specified by the URL in the
+``\textit{loggerurl=}'' line of the job log file, through the
+insertion interface of the service. The service in particular, SGAS
+LUTS has a simple custom web service interface loosely based on
+WS-ResourceProperties\cite{wsrp}.
 
@@
 ``\textit{jobreport\_options=}'' line of the A-REX configuration file
-(see Appendix \ref{config}).
+(see Sec. \ref{config}).
 
@@
 \textit{jobreport\_credentials} line of the A-REX configuration file
-(see Appendix \ref{config}), and no proxies are used. Normally the
-credentials for the A-REX service should be used.
+(see A-REX Description and Administration Manual\cite{arex}), and no
+proxies are used. Normally the credentials for the A-REX service
+should be used.
 
@@
 
+\section{Configuration}
+\label{config}
+
+JURA can ne set up through the configuration file of A-REX. Three
+commands are used by JURA: \textit{jobreport},
+\textit{jobreport\_credentials} and \textit{jobreport\_options}. See the
+A-REX Description and Administrator's Manual\cite{arex}, Section
+``Configuration'' for details.
+
+The \textit{jobreport\_options} command allows passing a generic option
+string to JURA verbatim. This option can be used for example for
+service-related settings. Currently only SGAS LUTS service is
+supported, for which the UR batch size can be set with the following
+command format:
+
+\verb|jobreport_options="urbatch:size"|
+
+%TODO implement!
+where \textit{size} is a positive integer, the number of URs sent
+within one request. If not given, the default is 50.
+
 \section{Implementation}
-JURA is written merely in C++, and is tested in a GNU/Linux
-environment, built with standard GNU tools. It depends on the HED
-libraries, and thus also on all mandatory dependencies thereof. The
+JURA is written merely in C++, built with standard GNU tools, and
+tested in a GNU/Linux environment. It depends on the HED libraries,
+and thus also on all mandatory dependencies thereof. The
 \textit{openssl} library is also required, since sensitive accounting
@@
 
-%TODO about other service types
+%TODO enable other service types, write about them
 
-When deployed, the executable called ``\textit{jura}'' is placed into the
-\textit{libexec} directory of the ARC install location. There is a
-symbolic link %TODO make it so!
+\section{Installation and deployment}
+
+Upon \verb|make install|, the executable called ``\textit{jura}'' is
+placed into the \textit{libexec} directory of the configured ARC
+install location. There is a symbolic link %TODO make it so!
 to the executable in the same directory, called ``\textit{logger}'',
@@
 the old logger client. No other executables or wrapper scripts are
-needed.
+installed.
+
+\section{Limitations and issues}
+Although stable, JURA still has minor imperfections, some of them
+stemming from flaws in A-REX or the \textit{arcsub} client:
+
+\begin{itemize}
+\item The time frequency of running JURA is not configurable. It is a
+  hardcoded value in A-REX: 3600 seconds, i.e. one hour. 
+\item The number of user-supplied reporting destinations is limited
+  for the sake of robustness. This upper limit is hardcoded: max. 3
+  destinations are parsed from JSDL, and max. 1 from RSL.
+\item The \textit{arcsub} job submission client removes all but one
+  reporting destination URL from the job description, further limiting
+  the number of user-supplied destinations.
+\item The expiration time of job log files can be given in each
+  ``\textit{jobreport=}'' line in the configuration file. However,
+  only one single value is used for all reporting destinations; the
+  last one given overrides the previous ones.
+\item Missing UR properties: charge, processors, network etc. see
+  App. \ref{log2ur}
+\item The GNU time memory measurement bug
+\end{itemize}
 
@@
 
-\section{Configuration}
-\label{config}
-
-%   loggerurl: jobreport, JSDL
-
-%   (lifetime)
-
-%   jobreport
-
-%   accounting\_options: jobreport\_options
-
-%   key\_path, certificate\_path, ca\_certificates\_dir: jobreport\_credentials
-
 \section{Usage Record properties}
 \label{log2ur}
-%   input file format
+The following table shows which properties in OGF UR are filled, what
+data source was used for them, which properties are missing, and what
+extensions have been added to UR.
 
-%   filled properties
+\begin{tabular}{|p{0.2\columnwidth}|p{0.5\columnwidth}|p{0.2\columnwidth}|}
+\hline
+\textbf{UR Property}&
+\textbf{Description}&
+\textbf{Source\newline(job log entry)}\\
+\hline\hline
+RecordId&
+Globally unique identifier for UR&
+nodename,\newline
+ngjobid\\
+\hline
+GlobalJobId&
+Globally unique identifier of job&
+globalid\\
+\hline
+LocalJobId&
+CE-specific identifier of job&
+localid\\
+\end{tabular}
 
-%   missing properties
+%  filled:
+% RecordId: nodename+ngjobid
+% GlobalJobId: globalid
+% LocalJobId: localid
+% GlobalUserName: usersn
+% LocalUserId: localuser
+% JobName: jobname
+% Status: status
+% WallDuration: usedwalltime
+% CpuDuration: usedcputime
+% StartTime: submissiontime (?)
+% EndTime: endtime
+% MachineName: nodename/1
+% Host: nodename
+% SubmitHost: clienthost
+% Queue: lrms
+% ProjectName: projectname
+
+%  unfilled:
+% ProcessId (localjobid?)
+% Charge
+
+%  extra:
+% UserVO(?) userproxy
+% RuntimeEnvironment: runtimeenvironment
+
+\section{Additional job log file entries}
+Table \ref{table:joblog} describes entries of the job log file not
+present in Table \ref{table:ur}.
 

### 2009-05-08 doc/tech_doc/infosys/infosys_technical.tex
Starting technical documentation of ISIS

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {},
  pdftitle = {Documentation of the ARC Information System},
  pdfsubject = {},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-TECH-XX\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large ARC Information System}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Documentation and developer's guide}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large \footnote{@} \large \par}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}

\chapter{Design Overview} % (fold)
\label{cha:design_overview}

\chapter{ISIS} % (fold)
\label{cha:isis}

\section{Registration handling} % (fold)
\label{sec:registration_handling}

\subsection{Functionality} % (fold)
\label{sub:isis_functionality}

 Main functionality of ISIS service visible from outside of ISIS cloud is to accept registration and provide collected information to clients. For that ISIS implements operations described in following section.
The seingle ISIS service accepts Registration Records pushed to it by other services (including ISIS services too) and stores them in local XML database. Stored records can be queried by clients using mandatory and service-specific attributes for selection criteria.

 In case of multiple ISIS services they form a peer-to-peer network called \texttt{Cloud of ISISes}. Inside that cloud Registration Records are propagated between services in such a way that all ISIS instances continiously try to synchronise their databases.

% TODO: put figure here

% subsection functionality (end)

\subsection{Interface} % (fold)
\label{sub:isis_interface}

\subsubsection{Operation Register}

\begin{description}

  \item[Input]~\begin{description}
    \item[Header]~\begin{description}
      \item[RequesterID] Identifier of the client.
      \item[MessageGenerationTime] Time when following set of RegEntry was generated. There may be multiple RegEntry elements.
    \end{description}
    \item[RegEntry]~\begin{description}
      \item[SrcAdv]~\begin{description}
        \item[Type] Type of service being registered. This element is opaque string for now. There shall be service types defined later.
        \item[EPR] Endpoint Reference of service being registered in terms of WS-Addressing.
        \item[SSPair] Set of key/value pairs representing service specific information.
      \end{description}
      \item[MetaSrcAdv]~\begin{description}
        \item[ServiceID] Globally unique and persistent identifier of the service.
        \item[GenTime] Time when this RegEntry was generated
        \item[Expiration] Validity period of this registration record.
      \end{description}
    \end{description}
  \end{description}

  \item[Output]~\begin{description}
    \item[Fault] Optional element describing fault which occured while performing registration. If missing registration succeeded.
  \end{description}

  \item[Faults]~\begin{description}
    \item[none]No specific faults are defined
  \end{description}

\end{description}

This operation is usually called by service which wants to register it's presence in ISIS. Service must supply mandatory information. \texttt{Endpoint Reference} is used to contact service. Only required element is contact URL of service. \texttt{Type} specifies kind of service and is used to find out functionality nad interface of service. \texttt{ServiceID} is used to distinguish between registered services and to deal with case of service changing it's contact URL.

As result of this operation new Registration Record is stored inside ISIS internal database and eventually propagated to other ISISes. If registration record of same ID already existed it will be renewed.

This operation is also used by ISIS services to propagate Registration Records inside ISIS cloud.

\subsubsection{Operation RemoveRegistrations}

\begin{description}

  \item[Input]~\begin{description}
    \item[MessageGenerationtime]
    \item[ServiceID] Multiple identifiers of services, whose recordds has to be removed.
  \end{description}

  \item[Output]~\begin{description}
    \item[RemoveRegistrationResponseElement]~\begin{description}
      \item[ServiceID] Identifier of service whose record was not removed
      \item[Fault] Description of failure reason
    \end{description}
  \end{description}

  \item[Faults]~\begin{description}
    \item[none]No specific faults are defined
  \end{description}

\end{description}

This operation is used to explicitely erase Registration Records associated with specified ServiceID values. If corresponding record does not exist it's identifier will be present in response message together with corresponding Fault element.

\subsubsection{Operation GetISISList}

\begin{description}

  \item[Input]~\begin{description}
    \item[none]
  \end{description}

  \item[Output]~\begin{description}
    \item[EPR] Multiple Endpoint References of known ISIS services
  \end{description}

  \item[Faults]~\begin{description}
    \item[none]No specific faults are defined
  \end{description}

\end{description}

In response to this operation the EndpointReferences to all known ISIS services are returned.

\subsubsection{Operation Query}

\begin{description}

  \item[Input]~\begin{description}
    \item[QueryString] XPath query expression
  \end{description}

  \item[Output]~\begin{description}
    \item[any] Result of query
  \end{description}

  \item[Faults]~\begin{description}
    \item[none]No specific faults are defined
  \end{description}

\end{description}

This operation allows any XPath queries to be performed on stored Registration Records. The records are treated as merged in one XML document with each record being equivalent to \texttt{RegEntry} element of \texttt{Register} operation. In response all elements produced by XPath query are returned.

% subsection interface (end)

% section registration_handling (end)

\section{Peer-to-Peer} % (fold)
\label{sec:peer_to_peer}

\subsection{Functionality} % (fold)
\label{sub:functionality}

% subsection functionality (end)

\subsection{Interface} % (fold)
\label{sub:interface}

% subsection interface (end)

% section peer_to_peer (end)

% chapter isis (end)

\chapter{Service} % (fold)
\label{cha:service}

\section{Information} % (fold)
\label{sec:information}

% section information (end)

\section{Registration} % (fold)
\label{sec:registration}

% section registration (end)

% chapter service (end)

% \bibliography{grid}
\end{document}

### 2009-05-09 doc/tech_doc/jura/jura-tech-doc.tex
JURA documentation complemented

--- 2009-04-22 doc/jura/jura-tech-doc.tex
@@
 
-%-------- Change this to NORDUGRID-TECH-NN
+%-------- Change this to NORDUGRID-TECH-24
 
-{\raggedleft NORDUGRID-TECH-NN\par}
+{\raggedleft NORDUGRID-TECH-24\par}
 
@@
 %%%%---- The title ----
-{\centering \textsc{\Large ARC Accounting Component -- JURA}\Large \par}
+{\centering \textsc{\Large Job Usage Reporter of ARC -- JURA}\Large \par}
 \vspace*{0.5cm}
@@
 %%%%---- A subtitle, if necessary ----
-{\centering \textit{\large Technical document}\large \par}
+{\centering \textit{\large Technical description}\large \par}
 
@@
 
-\section{Purpose}
+\section{Introduction}
 
-The \textit{Job Usage Reporter for ARC} (JURA) is a component
-implementing accounting functionality in the ARC middleware. Its
-objective is to gather metered resource usage data for each job and
-submit it to accounting services along with the job submitter's
-identity and miscellaneous job-related metadata.
+The \textit{Job Usage Reporter of ARC} (JURA) is a component
+implementing a part of the accounting functionality in the ARC
+middleware. Its objective is to gather metered resource usage data for
+each job and submit it to accounting services along with the job
+submitter's identity and miscellaneous job-related metadata.  The
+collected usage data is transformed into records of job-level
+granularity and a format corresponding to an OGF specification.
 
-The accounting service stores the received usage data in a database,
-and provides an interface for querying it. Queries can be made by the
-consumers of the accounting data, such as a billing component. The
-service itself is a third party application, separate from the
-middleware distribution. JURA is currently capable of using the
-logging service of the SweGrid Accounting System (SGAS)\cite{sgas},
-but maintaining the possibility to enable utilizing other services has
-been kept in mind during design.
+JURA submits the records to an accounting service.  The accounting
+service stores the received usage data in a database, and provides an
+interface for querying it. Queries can be made by the consumers of the
+accounting data, such as a billing component. JURA is currently
+capable of reporting to the logging service of the SweGrid Accounting
+System (SGAS)\cite{sgas}. Maintaining the possibility to utilise other
+accounting services has been kept in mind during design, the JURA
+modular architecture enables easy creation of adapters for accounting
+services.
 
-Before the usage data collected from the resource manager is
-submitted, it is transformed into records of job-level granularity. To
-every job corresponds exactly one Grid user, therefore reports over a
-time period (e.g.~an invoice) can be generated per-user, or
-alternatively on a larger scale such as job project or VO level.
+JURA offers a functionality similar to that of the
+\textit{logger}\cite{logger} client, with improvements and extensions.
+It also serves as a complete replacement for the JARM component of
+SGAS.
 
-\section{Architecture}
+\section{Overview of functionality}
 
@@
 \centering{{{\includegraphics[width=0.9\textwidth]{components.pdf}}}
-\caption{\label{fig:components}Components participating in accounting.} }
+\caption{\label{fig:components}The usage reporting mechanism.} }
 \end{figure}
 
-JURA offers a complete replacement in functionality for the old
-\textit{logger} utility\cite{logger}, a job metadata logging tool with
-a purpose very closely related to that of this new tool. However,
-backwards compatibility is maintained, so the old logger can still be
-deployed.
+JURA runs as a stand-alone binary application called by A-REX (see
+Figure \ref{fig:components}). There is no designated configuration
+file for JURA, nor is the configuration file of A-REX read directly by
+the application. Instead, options related to reporting are included
+within the job log files generated by A-REX. The primary purpose of
+these job log files is holding job-related metadata, the main input of
+JURA. Its format is described in detail in Section \ref{joblogs}.
 
-The ARC execution manager, A-REX\cite{arex} initiates JURA in a
-similar way to invoking logger. JURA reads job log files provided by
-A-REX. These files have the same format as those meant for logger,
-with additional lines which are necessary for accounting scenarios.
+The application is run periodically. First, it processes the resource
+usage content of job log files, and transforms it, generating a
+standard XML representation of usage data, called Usage Records
+(URs)\cite{ur}. Then these records are sent to one or more accounting
+services, referred to as \textit{reporting destinations} in this
+document. Several reporting destinations are supported, these can be
+configured by the system administrator in the A-REX configuration
+file, and in addition, the user submitting the job can specify
+destinations in the job description.
 
-It also acts as a client for one or more accounting services,
-specifically SGAS Logging and Usage Tracking Services (LUTS's),
-inserting the generated records in batches. (See Figure
-\ref{fig:components}.)
+Currently, the SGAS Logging and Usage Tracking Service (LUTS) is the
+only supported reporting destination. Communication with a LUTS server
+is done via a web service interface. JURA is securely authenticated by
+the server using X.509 certificates.
+
+% runs as a standalone binary called by the a/rex. config values are
+% also provided by AREX through the joblogfiles (the JURA does not read
+% config file)
+
+% processe job log data of AREX
+% transfrorsm that to UR, generates XML
+% periodically submits records to reporting destinations *accounting
+% service)
+% supports multiple reporting desti
+% reporting destination can be set by sysadmin *in config( AND by user
+% in her jobdescription
+% currently SGASLUTS is the only reporting destination
+% Communicastion with SGASluts is done via WS interface
+% Securely identifies ^ authenticates itself with credentials x509
 
@@
 \subsection{Invocation}
+\label{exec}
+JURA is a stand-alone executable application, executed by A-REX hourly
+(currently a hardcoded time interval, see Section \ref{future}). It
+has no separate configuration file, and does not process the A-REX
+configuration file. It receives all necessary options from A-REX in
+part through command-line arguments and mostly via variables inserted
+into the job log files (See Section \ref{joblogs}). The following
+configuration variables can be present in the job log files:
 
-JURA is a stand-alone executable application, executed hourly by
-A-REX. It has no separate configuration file; it gets all necessary
-configuration options from A-REX, in part through command-line
-arguments, but mostly via lines in the job log files (see Appendix
-\ref{config} for details). The source of the latter are lines in the
-grid-manager configuration file.
+\begin{itemize}
+\item \textbf{\textit{key\_path}} -- Path to the private key file used when
+  submitting records.
+\item \textbf{\textit{certificate\_path}} -- Path to the certificate file used
+  when submitting records.
+\item \textbf{\textit{ca\_certificates\_dir}} -- Directory holding the
+  certificates of trusted CAs.
+\item \textbf{\textit{accounting\_options}} -- Additional configuration options
+  for JURA.
+\end{itemize}
 
-The command line format is a subset of that of logger:
+The source of these variables is the ``\textit{grid-manager}'' block
+of the A-REX configuration file (see Section \ref{config}).
 
-\verb|jura [-E <expiration_time>] <control_dir>|
+The command line format of JURA is the following: %TODO make it so!
+
+\verb|jura [-E <expiration_time>] <control_dir> [<control_dir> [...]]|
 
@@
 \textit{control\_dir} is the A-REX control directory for a mapped
-local UNIX user.
+local UNIX user. The ``\textit{logs}'' subdirectory of each control
+directory is traversed by JURA separately, in sequence.
 
-\subsection{Parsing job log files}
+\subsection{Processing job log files}
 \label{joblogs}
-The job logs generated by A-REX reside under the directory
-\verb|<control_dir>/logs|. They have file name format
-\verb|<ngjobid>.<random>|, where \textit{ngjobid} is the identifier
-created for the job by A-REX, \textit{random} is a randomly generated
-sequence of alphanumeric characters to avoid collision of different
-files pertaining to the same job.
 
-2 for each logging destination: when submitted, when finished
+Job log files contain practically all input data (except those passed
+as command line arguments) for JURA. A-REX generates these files, at
+least two for each job and for each reporting destination: one at the
+time of job submission, another one after the job finishes, and
+possibly others at each start and stop event. Job log files are the
+main and only source of detailed resource usage information.
+Furthermore, they are used to communicate configuration parameters of
+JURA (see Section \ref{exec}).
 
-UR generated from them: only from finished
+The job log files generated by A-REX reside under the directory
+\textit{$<$control\_dir$>$/logs}\cite{arex}. They have file name format
+\textit{$<$ngjobid$>$.$<$random$>$}, where \textit{ngjobid} is the
+identifier created for the job by A-REX, \textit{random} is a randomly
+generated sequence of alphanumeric characters to avoid collision of
+different files pertaining to the same job.
 
-\subsection{Accessing LUTS}
+A file consists of ``\textit{name=value}'' lines, where
+``\textit{value}'' is either a job-related resource usage data or a
+configuration parameter. The URL of the reporting destination
+corresponding to the job log file is acquired from a
+``\textit{jobreport=}'' line in the A-REX configuration file. In
+addition to this server-side configuration, a limited number of
+destinations can be supplied by the submitter in the job
+description.
 
-insert method
+% For each job
+% and for each reporting destination (i.e. accounting service to be
+% contacted), at least two log files are written for each job: one at
+% the time of job submission, another one after the job finishes,
+% and possibly others at each start and stop event.
 
-certificates from:...
+% A job log file consists of ``\textit{name=value}'' lines.  These make
+% up some of the configuration options for JURA (see Sec.~\ref{config}),
+% such as the URL of an accounting destination in the
+% ``\textit{loggerurl=}'' line. The file also contains detailed resource
+% usage data and related metadata to be reported about the job. (See
+% Appendix \ref{log2ur} and \ref{joblog-extra} for more details.)
+% The accuracy of the metered data may depend on the type of batch
+% system used.
+
+JURA generates records in the Usage Record (UR) format\cite{ur}
+proposed by the Open Grid Forum (OGF), using the information stored in
+the job log files. The generated UR is an XML representation holding
+consumption information for all commonly used resources and
+metrics. It can be extended by custom elements for non-standard
+resources and/or other types of job metadata. For a list of UR
+properties and their sources in the job log file, see Appendix
+\ref{log2ur}.
+
+Some elements of UR are mandatory, these must all be present in the
+job log file to be able to generate a UR. For example, the job log
+file generated upon job submission contains no \textit{status} entry,
+so this file is ignored, and no UR is generated from it.
+
+After successful submission to a reporting destination, the job log
+file is deleted, preventing multiple insertion of usage records. If
+submission fails, the log files are kept, so another attempt is made
+upon a subsequent run of JURA. This can repeat until the expiration
+time passes (see ``\textit{-E}'' command line switch in Section
+\ref{exec}), at which point the next execution of JURA removes the
+file without processing.
+%TODO archiving
+
+%minden reporting desztihez van egy sajat job logfile  file
+
+%mi lesz a fajlokkal? expiration_time?
+
+\subsection{Reporting to LUTS}
+\label{accessing}
+The generated URs are submitted to the accounting services specified
+by the reporting destination configuration parameters and if present,
+to the destinations specified in the job description as
+well. Reporting URs to several destinations is possible, but currently
+only SGAS LUTS destinations are supported.
+
+LUTS has a simple custom web service interface loosely based on
+WS-ResourceProperties\cite{wsrp}. JURA uses the insertion method of
+this interface to report URs. The corresponding job log files are
+deleted after receiving a non-fault response from the service.
+
+% archive???
+
+%Currently several reporting desti possible.
+
+%The supported desti service az csak LUTS.
+
+%communikacio hogy megy lutsal.
+
+To increase communication efficiency JURA can send URs in batches
+provided that the server side supports this feature. LUTS accepts a
+batch of URs in a single request. The batch is an XML element called
+\textit{UsageRecords}, containing elements representing URs.
+
+The process of handling batches is the following: JURA does not send
+all usage records immediately after generating, but instead collects
+them in a batch until reaching the maximal number of elements or until
+running out of job log files. This means that all batches are of
+maximal size except the last one. The maximal number of URs in a batch
+can be set as a configuration parameter of JURA
+(``\textit{jobreport\_options}'', see Section \ref{config}).
+
+%ird le hogy megy a btch kuldes.
 
 \section{Security}
-  runs as: same user as a-rex, typically root
+The JURA executable runs with the same user priviliges as the A-REX,
+typically as \textit{root}. The owner of a job log file is the local
+user mapped for the submitter entity of the corresponding job. These
+files contain confidential data, access to which must be restricted,
+therefore read access is limited to the owner and the super user. If
+JURA is executed by A-REX, it can read data from these files, and
+delete expired files.
 
-  accesses sensitive data in job logs
+The authentication towards the SGAS LUTS is done via the standard
+X.509 certificate mechanism over SSL protocol: a chain of valid
+(i.e. not expired and/or revoked) certificates with a trusted root
+certification authority is accepted as authentic identification of the
+client. In the scenario involving A-REX and JURA,
+all usage records are submitted using credentials given in the
+``\textit{jobreport\_credentials=}'' line of the A-REX configuration file
+(see A-REX Description and Administration Manual\cite{arex}), and no
+proxies are used. Normally the credentials for the A-REX service
+should be used.
 
-  sgas security:
+Access control to the LUTS service can only be controlled by very
+simple means: the Distinguished Name (DN) of the client (in this case
+JURA) is checked against configured rights. Policies distinguish two
+rights: publishing and querying. Clients with publishing right can
+insert any UR, regardless of content. By default, querying right only
+allows retrieving URs pertaining to jobs submitted by the querying
+entity. In addition, there is a super user role allowing publishing
+and querying of any record.
 
-    standard ssl cert, possibly proxy
+% For every DN
 
-    no other attributes but DN considered
+% *DN( rights of
+% wrirte.tyjytukuolpo
 
-    access control: flat-file based + XACML-like(?) configured thru
-    SAM porttype
+% The access control within LUTS is based upon DN of the reported *JURA
+% LUTS can control which client *JURA( can submit records and who can
+% read records.
 
-  in this case: service credentials, no proxy
+% can be configured in two ways.
+% mik az policik??
 
 \section{Implementation}
-  c++
+\label{implement}
+JURA is written merely in C++, built with standard GNU tools, and
+tested in a GNU/Linux environment. It depends on HED libraries: the
+common utilities library, the message handling library and the Message
+Chain Components for the TCP, TLS, SOAP and HTTP protocols. It also
+depends on all mandatory dependencies of ARC: gthread-2.0, glibmm-2.4,
+libxml-2.0, openssl, e2fsprogs and GNU gettext.
 
-  sw dependencies
+%TODO enable other service types, write about them
 
-    full arc
+The usage reporting part of the JURA code has a modular design in
+order to enable adding other types of accounting services besides the
+curently only supported LUTS. An abstract interface class called
+``\textit{Destination}'' represents a usage reporting destination
+service. Different inherited classes of ``\textit{Destination}''
+handle different types of services. Currently the only inherited class
+is ``\textit{LutsDestination}'', implementing support for LUTS. If one
+wishes to develop a module for another service type, he or she has to
+create a new descendant class of ``\textit{Destination}'', implement
+its ``\textit{report()}'' method to submit the contents of the job log
+file given as an argument, and adapt the static
+``\textit{createDestination()}'' method of ``\textit{Destination}'' to
+instantiate the appropriate class.
 
-    ssl
+\section{Installation and deployment}
 
-  install location
+JURA is distributed as part of the ARC technology preview release
+source tarballs \footnote{http://download.nordugrid.org/software/nordugrid-arc1/trunk/}.
+The component can be built from source tarball using the standard autotools.
 
-\appendix
+The README and INSTALL of the source tarball provide full
+instructions about building ARC including JURA. The files also list
+required dependencies (see Section \ref{implement}).
+
+Upon \verb|make install|, the executable called ``\textit{jura}'' is
+placed into the \textit{bin} directory of the configured ARC
+install location. No other executables or wrapper scripts are
+installed.
+
+A-REX executes the reporting executable through the file ``logger'' in
+the \textit{libexec/arc} directory under the install location (this is
+an A-REX legacy kept for backwards compatibility with the old
+logger). Therefore, to enable automatic periodic reporting, a symbolic
+link called ``logger'' should be placed in \textit{libexec}, pointing
+to the JURA executable.
+
+The usage reporting can also be performed manually provided that
+access to the credentials are granted, by executing JURA with the
+proper command line arguments (see Section \ref{exec}). The example
+command below will send generated usage records from the job log files
+in the standard location, ``\textit{/tmp/jobstatus/logs}'' and send
+them to LUTS services. Files older than a week are deleted without
+processing.
+
+\begin{verbatim}
+jura -E 7 /tmp/jobstatus
+\end{verbatim}
 
@@
 \label{config}
-  jobreport
 
-  jobreport\_options
+JURA can be configured through the configuration file of
+A-REX\cite{arex}. As it was already mentioned in Section \ref{exec},
+JURA does not process the A-REX configuration file directly; the
+configuration values are propagated to JURA through the job log
+files. The following variables in the ``\textit{grid-manager}'' block
+of the A-REX configuration file are relevant for JURA:
 
-  jobreport\_credentials
+\begin{itemize}
+\item \textbf{\textit{jobreport}}\textit{={[}URL ... number]} --
+  specifies reporting destination URLs. Multiple entries and multiple
+  URLs are allowed. \textit{number} specifies how long old records
+  have to be kept if failed to be reported. That value is specified in
+  days. Last specified value becomes effective.
+\item \textbf{\textit{jobreport\_credentials}}\textit{={[}key\_file
+    {[}cert\_file {[}ca\_dir]]]} -- specifies the credentials for
+  accessing the accounting service.
+\item \textbf{\textit{jobreport\_options}}\textit{={[}options]}
+  -- specifies additional options for JURA.
+\end{itemize}
 
-\section{Usage Record properties}
+The \textit{jobreport\_options} variable allows passing a generic
+option string to JURA verbatim. This string can be used for example
+for service-related settings. Currently only SGAS LUTS service is
+supported, for which the UR batch size can be set with the following
+configuration line:
+
+\verb|jobreport_options="urbatch:size"|
+
+%TODO implement!
+where \textit{size} is a positive integer, the number of URs sent
+within one request. If not given, the default is 50. Zero value means
+unlimited batch size.
+
+%EXAMPLE!
+
+%TODO archiving option!!!
+
+\section{Limitations and future plans}
+\label{future}
+Although complete, JURA still has minor imperfections, some of them
+stemming from flaws in A-REX or the \textit{arcsub} client:
+
+\begin{itemize}
+\item The time frequency of running JURA is not configurable. It is a
+  hardcoded value in A-REX: 3600 seconds, i.e. one hour. 
+\item The number of user-supplied reporting destinations is limited
+  for the sake of robustness. This upper limit is hardcoded in A-REX:
+  max. 3 destinations are parsed from JSDL, and max. 1 from RSL.
+\item The \textit{arcsub} job submission client removes all but one
+  reporting destination URL from the job description, further limiting
+  the number of user-supplied destinations.
+\item The current implementation of JURA and A-REX supports only one
+  expiration time for all the reporting destinations. Even though the
+  configuration enables the specification of different expiration
+  values per reporting destination, it is not taken into account by
+  the system, the last value is used as the common expiration time
+  value.
+\item It is not possible to use different credentials per destinations.
+\item Some optional UR properties are not supported (see
+  App.~\ref{log2ur}).
+\item Memory is not reported correctly. A bug in GNU ``time'' results
+  in all memory usage set incorrectly as zero.
+\item Some necessary extensions to the generated UR are not yet filled
+  though the information is already collected in the job log files.
+\item Handling of runtime environment string is missing.
+\item Detailed user identity information based on proxy or other
+  submitted credentials is missing from URs.
+\end{itemize}
+
+There is also plan to extend JURA with ther following features:
+
+\begin{itemize}
+\item An archiving functionality, storing generated URs in a local
+  directory specified in the configuration.
+\item Support for other accounting systems besides LUTS, especially
+  the EGEE Accounting Service. However, appropriate documentation on
+  the service interfaces is missing as of now.
+\item Extend JURA to be able to read all necessary parameters
+  (credentials, reporting targets) from the command line so that it
+  can serve as a stand-alone reporting utility invoked independently
+  from A-REX.
+\item Enable generating coarser-grained ``\textit{Aggregate URs}''
+  from multiple URs.
+\item Investigate the subject of project-related charging: who is
+  responsible for determining the ``\textit{charge}'' value; what rules should
+  be applied?
+\end{itemize}
+
+%//archive
+
+%//supoort for other accounting systems (glite) felteve hogy ezek
+%interface elerheto lesz.
+
+%extend JURA to be able to read all the parameters (credentials,
+%reporting targets) from the command
+%line so that it can serve as a standalone reporting utility.
+
+%/aggregated records
+
+%//investigate project related charging, who should set it reliably, how...
+
+\newpage
+\appendix
+
+\section{Generated Usage Record}
 \label{log2ur}
-  input file format
+The following table shows which properties in OGF UR\cite{ur} are
+filled, what data source was used for them, and which properties are
+missing.
 
-  filled properties
+%, and what extensions have been added to UR.
 
-  missing properties
+\begin{tabular}{|p{0.22\columnwidth}|p{0.22\columnwidth}|p{0.54\columnwidth}|}
+\hline
+\textbf{Generated\newline UR Property}&
+\textbf{Source\newline(job log entry)}&
+\textbf{Information content}\\
+\hline\hline
+RecordId&
+nodename,\newline
+ngjobid&
+Globally unique identifier for UR\\
+\hline
+GlobalJobId&
+globalid&
+Globally unique identifier of job:\newline
+XML element as defined by BES\\
+\hline
+LocalJobId&
+localid&
+CE-specific identifier of job\\
+\hline
+GlobalUserName&
+usersn&
+DN of submitting user's certificate\\
+\hline
+LocalUserId&
+localuser&
+POSIX user on CE executing the job\\
+\hline
+JobName&
+jobname&
+Name of job, as given in job description\\
+\hline
+Status&
+status&
+Status of job\\
+\hline
+WallDuration\newline $[$ISO 8601 duration$]$&
+usedwalltime $[$s$]$&
+Wall-clock time used by job\\
+\hline
+CpuDuration\newline $[$ISO 8601 duration$]$&
+usedcputime $[$s$]$&
+CPU time used by job\\
+\hline
+StartTime\newline $[$ISO 8601 time stamp$]$&
+submissiontime\newline $[$ISO 8601 time stamp$]$&
+Time instant the job started\\
+\hline
+EndTime\newline $[$ISO 8601 time stamp$]$&
+endtime\newline $[$ISO 8601 time stamp$]$&
+Time instant the job ended\\
+\hline
+MachineName&
+nodename&
+Name of the machine where the job ran\newline
+(first node from colon-separated list put into element)\\
+\hline
+Host&
+nodename&
+System hostname(s) where the job ran\newline
+(nodes from colon-separated list put into separate elements)\\
+\hline
+SubmitHost&
+clienthost&
+System hostname the job was submitted from\\
+\hline
+Queue&
+lrms&
+Name of the queue from which the job was executed\\
+\hline
+ProjectName&
+projectname&
+Name of project, as given in job description\\
+\hline
+Memory\newline
+$[$average virtual, kB$]$&
+usedmemory $[$kB$]$&
+Average total memory used by job\\
+\hline
+Memory\newline
+$[$max physical, kB$]$&
+usedmaxresident $[$kB$]$&
+Maximal resident memory used by job\\
+\hline
+Memory\newline
+$[$average physical, kB$]$&
+usedaverageresident $[$kB$]$&
+Average resident memory used by job\\
+\hline
+NodeCount&
+nodecount&
+Number of nodes (physical machines) involved in the job.\\
+\hline
+ProcessID&
+\textbf{MISSING}&
+The process ID(s) of the job\\
+\hline
+Charge&
+\textbf{MISSING}&
+Total charge of the job (money or abstract credits)\\
+\hline
+Network&
+\textbf{MISSING}&
+Network usage of job\\
+\hline
+Disk&
+\textbf{MISSING}&
+Disk usage of job\\
+\hline
+Swap&
+\textbf{MISSING}&
+Swap usage of job\\
+\hline
+Processors&
+\textbf{MISSING}&
+Number of processors used or requested\\
+\hline
+TimeDuration&
+\textbf{MISSING}&
+Additionally measured time duration(s)\\
+\hline
+TimeInstant&
+\textbf{MISSING}&
+Additionally identified time instant(s)\\
+\hline
+ServiceLevel&
+\textbf{MISSING}&
+Quality of service associated with usage\\
+
+%\hline
+%\hline
+%\textbf{Extra UR Property}&
+%\textbf{Source\newline(job log entry)}&
+%\textbf{Description}\\
+%\hline\hline
+%UserProxy&
+%userproxy&
+%Proxy of user, including VOMS info\\
+%\hline
+%RuntimeEnvironment&
+%runtimeenvironment&
+%Runtime environment info\\
+
+\hline
+\end{tabular}
+
+%  filled:
+%% RecordId: nodename+ngjobid
+%% GlobalJobId: globalid
+%% LocalJobId: localid
+%% GlobalUserName: usersn
+%% LocalUserId: localuser
+%% JobName: jobname
+%% Status: status
+%% WallDuration: usedwalltime
+%% CpuDuration: usedcputime
+%% StartTime: submissiontime (?)
+%% EndTime: endtime
+%% MachineName: nodename/1
+%% Host: nodename
+%% SubmitHost: clienthost
+%% Queue: lrms
+%% ProjectName: projectname
+
+%  unfilled:
+%% ProcessId (localjobid?)
+%% Charge
+
+%  extra:
+% UserVO(?) userproxy
+% RuntimeEnvironment: runtimeenvironment
+
+%%%% Extended UR schema!
 

### 2009-05-10 doc/tech_doc/infosys/infosys_technical.tex
Porting part or Technical Handbook to ISIS tech doc.

@@
 
-{\raggedleft NORDUGRID-TECH-XX\par}
+{\raggedleft NORDUGRID-TECH-21\par}
 
@@
 
+ The functionality of ISIS is made of two parts. On the one hand they are working as ordinary Web Services, and on the other hand maintain a peer-to-peer network - \texttt{ISIS cloud}.
+
  Main functionality of ISIS service visible from outside of ISIS cloud is to accept registration and provide collected information to clients.
 For that ISIS implements operations described in following section.
-The seingle ISIS service accepts Registration Records pushed to it by other services (including ISIS services too) and stores them in local XML database.
+The single ISIS service accepts Registration Records pushed to it by other services (including ISIS services too) and stores them in local XML database.
 Stored records can be queried by clients using mandatory and service-specific attributes for selection criteria.
 
- In case of multiple ISIS services they form a peer-to-peer network called \texttt{Cloud of ISISes}.
+ In case of multiple ISIS services they form a peer-to-peer network.
 Inside that cloud Registration Records are propagated between services in such a way that all ISIS instances continiously try to synchronise their databases.
@@
         \item[ServiceID] Globally unique and persistent identifier of the service.
-        \item[GenTime] Time when this RegEntry was generated
-        \item[Expiration] Validity period of this registration record.
+        \item[GenTime] (Generation Time) The actual timestamp of information (called wake\_up\_time in the pseudo code below)
+        \item[Expiration] Validity period of Service Advertisement record.
       \end{description}
@@
 This operation is usually called by service which wants to register it's presence in ISIS.
+This message consists of one or more \textbf{Registration Entry} and at most one \textbf{Registration Header}.
+The \textbf{Registration Entry}(RegEntry) contains a \textbf{Service Advertisement}(SrcAdv) and a corresponding \textbf{Service Advertisement Metadata}(MetaSrcAdv).
+This structure is shown on Figure \ref{fig:RegistrationMessage}.
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{RegistrationMessage.png}}}
+\caption{\label{fig:RegistrationMessage}Embedded structure of Registration Message} }
+\end{figure}
+
 Service must supply mandatory information.
-\texttt{Endpoint Reference} is used to contact service.
-Only required element is contact URL of service.
+Among those the \texttt{Endpoint Reference} is used to contact service - only required element is contact URL of service.
 \texttt{Type} specifies kind of service and is used to find out functionality nad interface of service.
 \texttt{ServiceID} is used to distinguish between registered services and to deal with case of service changing it's contact URL.
+For more information about mandatory and optional information see Section \ref{service_advertisement}
 
@@
     \item[MessageGenerationtime]
-    \item[ServiceID] Multiple identifiers of services, whose recordds has to be removed.
+    \item[ServiceID] Multiple identifiers of services, whose records has to be removed.
   \end{description}
@@
 
-This operation is used to explicitely erase Registration Records associated with specified ServiceID values.
+This operation is used to explicitely requesting the removal of zero or more Registration Entries associated with specified ServiceID values stored in the Information System.
 If corresponding record does not exist it's identifier will be present in response message together with corresponding Fault element.
@@
 In response to this operation the EndpointReferences to all known ISIS services are returned.
+The operation is used for obtaining a list of known ISIS instances from any particular ISIS.
+Clients can then use the obtained list to run direct queries against the ISIS instances.
+The operation is provided for fault tolerance and for providing optional performance boost.
+This operation returns the known peer-to-peer neighbors.
+If client is interested in all ISIS instances they can be obtained using the Query operation because they are ordinary services registered into the Information System.
 
@@
 In response all elements produced by XPath query are returned.
+The purpose of this operation is to make it possible to obtain any kind of information related to the Indexing Database.
 
@@
 \subsection{Functionality} % (fold)
-\label{sub:functionality}
+\label{sub:p2p_functionality}
 
@@
 \subsection{Interface} % (fold)
-\label{sub:interface}
+\label{sub:p2p_interface}
 
@@
 \section{Information} % (fold)
-\label{sec:information}
+\label{sec:service_information}
+
+The service developers have to ensure that the services are providing the necessary information about themselves.
+This is done by implementing the subclass of the Arc::Service class - the RegistrationCollector function has to provide up-to-date status information about the service and anything else it wants to be advertised.
+This information package is called \textbf{Service Advertisement}.
+\label{service_advertisement}
+The \textbf{Service Advertisement} can contain any information service wants to advertise but the mandatory elements have to be always present:
+\begin{itemize}
+  \item Service ID: A globally unique identifier of the service.
+  \item Service Type: The Glue2 type of service.
+  \item Endpoint URL: The URL where the service can be contacted provided as part of EPR element.
+\end{itemize}
 
@@
 \section{Registration} % (fold)
-\label{sec:registration}
+\label{sec:service_registration}
+
+The registration of service is carried out by internal module called Registrant.
+The Registrant is active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes.
+In practice, the configuration part of the Registrant contains exactly one ISIS to bind, and the Registrant will collect the necessary information about the other ISISes belonging to the same network.
+
+To register services to more than one ISIS network multiple Registrant instances has to be configured.
+In this case, the default Registrant will be used for registering every services unless configured expilicitely.
+The registration of service can be done once or periodically based either on the configuration of the Registrant or overwritten for every service separately.
+The Registrant is also performing message aggregation of all services linked to it if possible.
+The simplified algorithm of the Registrant is presented below.
+
+\begin{framed}
+  Registrant - pseudo algorithm\\
+  \\
+  \verb#// Initialize phase#\\
+  \verb#Read the configuration and store the information about the services in a list#\\
+  \verb#do { // Cyclic phase in a different Thread#\\
+  \verb# wake_up_time = now();#\\
+  \verb# messages = null;#\\
+  \verb#  if ( 0 < count(service where service.next_run <= wake_up_time)) {#\\
+  \verb#    foreach( service where service.next_run <= wake_up_time) {#\\
+  \verb#      messages.add(service.RegistrationCollector);#\\
+  \verb#      service.next_run = wake_up_time + service.period;#\\
+  \verb#    }#\\
+  \verb#    if (0 < count(messages)) {#\\
+  \verb#      sent_message = assemble message with headers(messages);#\\
+  \verb#      send(sent_message);#\\
+  \verb#    }#\\
+  \verb#  } else {#\\
+  \verb#    sleep(min(service.next_run) - now()); #\\
+  \verb#  }#\\
+  \verb#} while(true)#\\
+\end{framed}
+Current implementation does not allow value of the period to be less than 2 minutes.
+
+Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{isis_interface}).
+Before sending this information the Registrant extends it with additional data (\textbf{Service Advertisement Metadata}).
 

### 2009-05-11 doc/tech_doc/storage_system/StoragePaper/storagepaper.tex
First draft

%%
%% This is file `elsarticle-template-num.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% elsarticle.dtx  (with options: `numtemplate')
%% 
%% Copyright 2007, 2008 Elsevier Ltd.
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% -------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Future Generation Computing Systems}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Chelonia - distributed cloud storage}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author[UiO1,UiO2]{Jon Kerr Nilsen}
\ead{j.k.nilsen@usit.uio.no}

\author[UU1]{Salman Toor}
\ead{salman.toor@it.uu.se}

\author[NIIF]{Zsombor Nagy}
\ead{zsombor@niif.hu}

\author[UU2]{Bjarte Mohn}
\ead{bjarte.mohn@fysast.uu.se}

\address[UiO1]{University of Oslo, Dept. of Physics,  P. O. Box 1048, Blindern, N-0316 Oslo, Norway}
\address[UiO2]{University of Oslo, Center for Information Technology, P. O. Box 1059, Blindern, N-0316 Oslo, Norway}
\address[UU1]{Dept. Information Technology, Div. of Scientific Computing Uppsala University, Box 256, SE-751 05 Uppsala, Sweden}
\address[NIIF]{Institute of National Information and Infrastructure Development NIIF/HUNGARNET, Victor Hugo 18-22, H-1132 Budapest, Hungary}
\address[UU2]{Dept. of Physics and Astronomy, Div. of Nuclear and Particle Physics, Uppsala University, Box 535, SE-75121 Uppsala, Sweden}

\begin{abstract}
Nice abstract. 
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Grid, Storage, Cloud
%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%% \linenumbers

%% main text
\section{Introduction}
\label{sec:introduction}
The introduction should introduce the field of research, the "problem"
we are trying to solve and briefly explain our solution (and layout of
this paper). Key words: Distributed storage, clouds, grid enabled.

\section{System Architecture Overview}
\label{sec:systemoverview}
Blah blah here (web services, written in HED etc) and a figure showing
all the services (the usual one).
\subsection{Core Services}
\label{sec:coreservices}
Below all main services are described. ADD MORE TEXT HERE.

\subsubsection{The A-Hash}
\label{sec:theahash}
The A-Hash service is a database storing objects containing key-value
pairs. It can be deployed on multiple nodes to get a replicated
database. The Librarian service uses the A-Hash as database.

\subsubsection{The Librarian}
\label{sec:librarian}
The Librarian service manages the hierarchy and metadata of files and
collections, handling of Logical Names and monitoring of the Shepherd
services. The Librarian service is stateless, it stores all the
persistent information in the A-Hash. This makes it possible to deploy
any number of independent Librarian services to provide
high-availability and load balancing. In this case all the Librarians
should communicate with the same set of A-Hashes in order to use the
same database of metadata.

\subsubsection{The Shephard}
\label{sec:shephard}
The Shepherd service manages a particular storage node and provides
uniform interface for storing and accessing file replicas. The
Shepherd service itself cannot transfer files, so on a storage node
there must be at least one independent storage element service (with
an interface such as HTTP(S), ByteIO, etc.) which is capable of
performing the actual file transfer. So a storage node consists of a
Shepherd service and a storage element service connected
together. Storage element services can either be KnowARC or
third-party services. For each kind of storage element service, a
Shepherd backend module is needed which enables the Shepherd service
to communicate with the storage element service, e.g. to initiate file
uploads, downloads and removal, and to detect whether a file transfer
was successful or not. Currently there are three Shepherd backends:
for the ARC native HTTP(S) server called Hopi; for the Apache web
server; and for a service which implements a subset of the ByteIO
interface.

\subsubsection{The Bartender}
\label{sec:bartender}
The Bartender service provides a high-level interface of the storage
system for the clients (other services or users). The clients can
create and remove collections (directories), create, get and remove
files, move files and collections within the namespace using Logical
Names. There can be access policies associated with files and
collections, which are evaluated every time a user wants to access
them. The Bartender communicates with the Librarian and Shepherd
services to accomplish the client’s requests.  The file content
itself does not go through the Bartender; file transfers are directly
performed between the storage nodes and the clients. The Bartender
also has so-called gateway modules which make it possible to
communicate with third-party storage solutions, thus the user can
access multiple storage systems through a single Bartender
client. Currently there is only one gateway module, for third-party
storages with GridFTP 3 interface. It is possible to deploy multiple
Bartenders for high-availability and load balancing. In this case all
the Bartenders show the same namespace and work with the same files
and collections. They can use different set of Librarians as long as
all the Librarians are using the same set of A-Hashes.

\subsection{Security}
\label{sec:security}
As is the case for all openly accessible web services, the security
model is of crucial importance for the NorduGrid storage. The security
architecture of the storage can be split into three parts; the
inter-service authorization; the transfer-level authorization; and the
high-level authorization:

The communication with and within the storage system is realized
through HTTPS with standard X.509 authentication.

\subsubsection{Inter-Service Authorization}
The inter-service authorization maintains the integrity of the
internal communication between services. There are several
communication paths between the services in the storage system. The
Bartenders send requests to the Librarians and the Shepherds, the
Shepherds communicate with the Librarians and the Librarians talk with
the A-Hash. If any of these services is compromised or a new rogue
service gets inserted in the system, the security of the entire system
is compromised. To enable trust between the services, they need to
know each other's Distinguished Names (DNs). This way, a rogue service
would need to obtain a certificate with that exact DN from some
trusted Certificate Authority (CA).

\subsubsection{Transfer-Level Authorization}
The transfer-level authorization handles the authorization in the
cases of uploading and downloading files. When a transfer is
requested, the Shepherd will provide a one-time Transfer URL (TURL) to
which the client can connect. In the current architecture, this TURL
is world-accessible. This may not seem very secure at first. However,
provided that the TURL has a very long, unguessable name, that it is
transfered to the user in a secure way and that it can only be
accessed once before it is deleted, the chance of being compromised is
very low.

\subsubsection{High-Level Authorization}
The high-level authorization considers the access policies for the
files and collections in the system. These policies are stored in
A-Hash, in the metadata of the corresponding file or collection,
providing a fine-grained security in the system.

\subsection{Client Tools}
\label{sec:clienttools}
Being the only part a user will (and should) see from a storage
system, the client tools are an important part of the NorduGrid
storage. Currently ARC supports two ways of accessing the storage
solution. The \textbf{Command-line Interface} (CLI) provides access to
the storage through the methods \texttt{stat},
\texttt{makeCollection}, \texttt{unmakeCollection}, \texttt{putFile},
\texttt{getFile}, \texttt{delFile}, \texttt{list} and
\texttt{move}. Methods for modifying access and ownership will be
available in the near future.

The \textbf{FUSE module} provides a high-level access to the storage
system. Filesystem in Userspace (FUSE)~\cite{FUSE} provides a simple
library and a kernel-userspace interface. Using FUSE and the ARC
Python interface, the FUSE module allows users to mount the storage
namespace into the local namespace, enabling the use of graphical file
browsers.

It is worth mentioning that the client tools queries the storage
system through the Bartender only. Currently upload and download is
realized through HTTP(S), but there are plans to add support for other
protocols, such as SRM and GridFTP.

\section{System Characteristics}
\label{sec:systemcharateristics}
Describe here what characterizes the system. What it offers. What it
is, what it can be used for. Anything from automatic file recovery
till that the system can behave as virtual file system should be mentioned
here.

\section{System Performance}
\label{sec:systemperformance}
Describe here the results of all our tests divided into catagories
depending on if the test was a stability/usabilty test, or if it was a
performance test only. Key word: Scalability.

\section{Related Work}
\label{sec:systemperformance}
There are a number of Grid storage solutions on the market, but there
is no direct analogue to Chelonia.

dCache \cite{} mainly differs from Chelonia in that dCache has a
centralized set of core services while Chelonia is distributed by
design. dCache is a service oriented storage system which combines
heterogeneous storage elements to collect several hundreds of
terabytes in a single namespace. Originally designed to work on a
local area network, dCache has proven to be useful also in a Grid
environment, with the NDGF (Nordic Data Grid Facility) dCache
installation as the largest example. There, the core components, such
as the metadata catalogue, indexing service and protocol doors are run
in a centralized manner, while the storage pools are
distributed. Chelonia, designed to have multiple instances of all
services running in a Grid environment, will not need a centralized
set of core services. Additionally, dCache is relatively difficult to
deploy and integrate with new applications. Being a more light-weight
and flexible storage solution, Chelonia aims more towards new user
groups, less familiar with Grid solutions.

While Chelonia can run multiple instances of every service, the Disk
Pool Manager \cite{} (DPM) needs single instances of all services but the
storage pools. Most notably DPM has a centralized data base for
storing metadata. DPM is a lightweight solution for disk storage
management. It mainly consists of a DPM server, a DPM name server to
handle the namespace, SRM and GridFTP servers, and disk servers.
While serving as a lightweight alternative to dCache, the limited
distributed services may cause scaling challenges when it comes to the
usage rates.

Scalla \cite{} differs from Chelonia in that Scalla is designed for
use on centralized clusters, while Chelonia is designed for a
distributed environment. Scalla is a widely used software suite
consisting of an xrootd server for data access and an olbd server for
building scalable xrootd clusters. Originally developed for use with
the physics analysis tool ROOT 11 , xrootd offers data access both
through the specialized xroot protocol and through other third-party
protocols. The combination of the xrootd and olbd components offers a
cluster storage designed for low latency, high bandwidth
environments. In contrast, Chelonia is optimized for high latency and
is more suitable for the Grid environment.

As opposed to Chelonia, iRODS \cite{} does not provide any storage
itself but is more an interface to other, third-party storage
systems. Based on the client-server model, iRODS provides a flexible
data grid management system. It allows uniform access to heterogeneous
storage resources over a wide area network. Its functionality, with a
uniform namespace for several Data Grid Managers and file systems, is
quite similar to the functionality offered by our Gateway
service. However, iRODS uses a database system for maintaining the
attributes and states of data and operations. This is not needed with
Chelonia’s Gateway service.

While Chelonia is designed for geographically distributed users and
data storage, Hadoop \cite{} with its file system HDFS is directed
towards physically closely grouped clusters. HDFS builds on the
master-slave architecture where a single NameNode works as a master
and responsible for the metadata where as DataNodes are used to store
the actual data. Though similar to Chelonia’s metadata service, the
NameNode cannot be replicated and may become a bottleneck in the
system. Additionally, HDFS uses non-standard protocols for
communication and security while Chelonia uses standard protocols like
HTTP(S), GridFTP and X509.

XTtreemFS \cite{} has a very different standpoint when compared to
Chelonia, in that it is designed to work only in a specific operating
system while our Chelonia is non-intrusive by design. XTreemFS is an
object-based file system designed for the Grid environment of the
operating system XTreemOS. Being object-based, XTreemFS have separated
the metadata of files and the file content, or objects in the same
fashion as Chelonia. Being a tightly integrated part of an operating
system enables XTreemFS to implement functionality very similar to
that of local file systems while being a distributed Grid
storage. However, this also enforces a specific operating system,
requiring dedicated hardware resources.

\section{Future Work}
\label{sec:futurework}
I'm not sure if this section should be here.

\section{Conclusions}
The best system ever built.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

\begin{thebibliography}{00}

%% \bibitem{label}
%% Text of bibliographic item

\bibitem{}

\end{thebibliography}
\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.

### 2009-05-12 doc/tech_doc/infosys/infosys_technical.tex
Moving more from handbook. Started adding security parts.

@@
 
-\chapter{Design Overview} % (fold)
+\chapter{Design Overview}
 \label{cha:design_overview}
 
-\chapter{ISIS} % (fold)
+\chapter{ISIS}
 \label{cha:isis}
 
-\section{Registration handling} % (fold)
-\label{sec:registration_handling}
+\section{Registration handling}
+\label{sec:isis_registration_handling}
 
-\subsection{Functionality} % (fold)
+\subsection{Functionality}
 \label{sub:isis_functionality}
@@
 
-\subsection{Interface} % (fold)
+\subsection{Interface} 
 \label{sub:isis_interface}
@@
 The \textbf{Registration Entry}(RegEntry) contains a \textbf{Service Advertisement}(SrcAdv) and a corresponding \textbf{Service Advertisement Metadata}(MetaSrcAdv).
-This structure is shown on Figure \ref{fig:RegistrationMessage}.
+This structure is shown on Figure \ref{fig:registration_message}.
 \begin{figure}[ht]
 \centering{{\scalebox{0.6}{\includegraphics{RegistrationMessage.png}}}
-\caption{\label{fig:RegistrationMessage}Embedded structure of Registration Message} }
+\caption{\label{fig:registration_message}Embedded structure of Registration Message} }
 \end{figure}
@@
 
-\section{Peer-to-Peer} % (fold)
-\label{sec:peer_to_peer}
+\section{Peer-to-Peer}
+\label{sec:isis_peer_to_peer}
 
-\subsection{Functionality} % (fold)
-\label{sub:p2p_functionality}
+\subsection{Functionality}
+\label{sub:peer_to_peer_functionality}
+
+%\subsubsection{Peer-to-Peer functions}
+%\label{ISIS Peer-to-Peer functions}
+%\subsection{Client}
+%\label{Client}
+
+%The client part of the information service is for querying the data stored in the database of the ISIS 
+%services. This is implemented as a part of the ARCLib and using the ISIS service's interface for the 
+%various data queries. For further details see Section \ref{Query data}.
+
+%\section{Operations}
+%\label{Operations}
+%The different operations are performed by different Information system components. These cases are described below in the dedicated subsections.
 
@@
 
-\subsection{Interface} % (fold)
-\label{sub:p2p_interface}
+\subsection{Interface}
+\label{sub:isis_peer_to_peer_interface}
 
@@
 
+\section{Authorization}
+\label{sec:isis_authorization}
+
+\subsection{Client Authorzation}
+\label{sub:isis_authorization_client_authorization}
+
+% TODO figure with mutiple service types
+
+To ensure information stored in ISIS cloud can't be tampered and only available to proper clients following authorization framework is implemented.
+All actions perfored by ISIS clients are divided into three following groups:
+
+\begin{itemize}
+ \item Operations initiated by other ISISes in the cloud. Those include:
+ \begin{itemize}
+  \item Register with Registration Message containing information not about contacting client
+  \item RemoveRegistrations with request to remove Registration Message representing not contacting client
+  \item Connect % TODO: describe this operation
+ \end{itemize}
+
+ Those operations may cause uncontrollable changes in colleccted information and must be granted only for higly trusted entities like ISISes themselves.
+
+ \item Operations initiated by the Services registering to Information System. Those are:
+ \begin{itemize}
+  \item Register with Registration Message containing information about contacting client
+  \item RemoveRegistrations with request to remove Registration Message representing contacting client
+ \end{itemize}
+
+ \item Operations which are allowed for any liable client of particular Grid infrastructure.
+ \begin{itemize}
+  \item Query
+  \item GetISISList
+ \end{itemize}
+\end{itemize}
+
+% subsection client_authorization (end)
+
+\subsection{Information Authorzation}
+\label{sub:isis_authorization_information_authorization}
+
+%* Clasification of information
+%
+% 1. From acccessibility point of view information can be devided
+%into 3 groups:
+% a. public
+% b. available to trusted clients (like broker or scheduler service)
+% c. private - available to owner of information object and entities
+%trusted by owner.
+%Distinction between 2 and 3 is not clear. One may accept that 2 is
+%imposed by architecture/deployment of system and 3 is controlled by
+%entities which are users of the system.
+%
+ 2. From storage/access location point of view information is divided
+%into:
+% a. local - stored at endpoint representing resource which is
+%described by information. This information is obtainable through
+%ALIS interface.
+% b. propagated to dedicated services (ISIS) which collect information
+%needed for pre-selection and initial lookup of services
+%representing resources.
+%
+%NOTE: Definition of propagation of information does not exactly belong
+%to security architectire of information system. But it affects it significantly
+%and hence at least minimal set of definitions is needed.
+%
+%OUT OF BAND NOTE: Those services could provide pre-selection capability.
+%
+% Sets 1 and 2 may intersect in any way. In first approach to simplify
+%implementation 2b may be considered to always belong to 1a.
+%
+%
+%* Implementaion guidelines
+%
+% To implement set 1 information objects need to have Access Control
+%Rules (ACR) assigned (usually used term Access Control List seems to
+%be inadequate due to meaning of word List).
+%
+% To implement set 2 in similar way information objects have Propagation
+%Control Rules (PCR) assigned. Because curently information system is made
+%of one layer of ISIS PCRs may consist of one boolean property.
+%
+% Information is structured in tree-like fashion - XML document.
+%Access control is defined at node and follows all branches - similar
+%to file systems. Propagation control is similar to access control (if
+%needed at all). Children nodes can't widen access defined at parent
+%nodes. This should decrease amount of ACR to be evaluated.
+%
+%EXAMPLE: if parent node A allows access only to VO_1 then children
+%node B can narrow access to Adminsitrator of VO_1 and can't open
+%access for VO_2.
+%
+% By default all nodes are public and are propagatable. Properties
+%assigned to clildren nodes limit access and propagation. Access
+%permission to particular information node is evaluated by traversing
+%all nodes prom parent to child. At first node which gives negative
+%result evaluation is stopped.
+%
+% Because evaluation of permission may be costly (CPU intensive,
+%causing big latency if access to external service is needed, etc.)
+%other tricks to reduce cost are needed.
+%
+% To reduce access control evaluation cost ACRs may/should/must(?)
+%be assigned to information nodes in one-to-many way - one ACR per
+%multiple information nodes. This grouping may be initiated either
+%interactively (by clients explicitely specifying that same ACR is
+%applicable to multiple nodes) or automatically (by comapring new
+%ACR to already existing).
+%
+% If information nodes containing ACRs are propagated to ISIS (second
+%implementation step) propagation algorithm must take ACR into account.
+%For informtion propagation between ISISes this rule also applies.
+%EXAMPLE: Information with access open to VO_1 must not be propagated
+%to ISIS belonging to VO_2.
+%
+%
+%** Ideas for in-memory XML documents
+%
+% Because informational documents are normally produced dynamically and
+%have short lifetime there seems to be reasonable to keep them in memory.
+% Each node with assigned ACR contains attribute (even if that breaks schema)
+%containing id of ACR. ACR is defined as XML element (even if that breaks
+%schema) located at higher, same or child level of considered node.
+% Because informational documents are produced dynamically there are few
+%ways for processing ACRs:
+% 1. Right after document is generated it is traversed and ACRs applied.
+%Parts of document which do not pass ACR evaluation are cut off. ACRs are
+%cut off too.
+% 2. While calling document generation procedure identity of client is
+%supplied and ACRs are applied immediately.
+% 3. While calling document generation procedure hooks/callbacks are supplied
+%which are used and implement ACR evaluation.
+% Combination of 1 and 3 looks most promising because it is flexible and
+%keeps information system and policy evaluation code separated.
+%
+%
+%* References
+%"Information access control use cases" http://wiki.nordugrid.org/index.php/ARC1/Infosys/Access_Use_cases
+%
+
+% subsection information_authorization (end)
+
+% section authorization (end)
+
 % chapter isis (end)
 
-\chapter{Service} % (fold)
+\chapter{Service}
 \label{cha:service}
 
-\section{Information} % (fold)
-\label{sec:service_information}
+\section{Information generation}
+\label{sec:service_information_generation}
 
@@
 
+Because there may be multiple registration processes running in parallel it is important to ensure that implementation RegistrationCollector is thread safe or there are internal locks implemented.
+
 % section information (end)
 
-\section{Registration} % (fold)
+\section{Registration}
 \label{sec:service_registration}
@@
 
-Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{isis_interface}).
+Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{sub:isis_interface}).
 Before sending this information the Registrant extends it with additional data (\textbf{Service Advertisement Metadata}).
 
+An example layout of services is shown on Figure \ref{fig:registration_overview}. In this configuration example 
+there are two Registrants configured in one HED container for three services. The \textit{Registrant A} is the 
+default Registrant and the services configured in the following way:
+
+\begin{itemize}
+  \item Service 1: There is no Registrant configured so the default one will register it.
+  \item Service 2: The \textit{Registrant A} is configured explicitly.
+  \item Service 3: The \textit{Registrant B} is configured explicitly.
+\end{itemize}
+
+So the Service 1 and 2 are handled both by \textit{Registrant A} and Service 3 by \textit{Registrant B}.
+In the first step each Registrant performs information collection from all assigned services sequentially.
+Registrants themselves are executing in parallel.
+For Figure \ref{fig:registration_overview} the approximate sequence of information collection is (\textit{A1}-\textit{A2}, \textit{B1}-\textit{B2}) followed by (\textit{A3}, \textit{A4}).
+After collection  the aggregated information is sent to the corresponding ISISes independently by each other Registrant (\textit{A5}, \textit{B5}).
+
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{Registration.png}}}
+\caption{\label{fig:registration_overview}Overview of the registration process} }
+\end{figure}
+
+This registration operation is done once during the start-up phase and periodically according to (per service) configured periods.
+
 % section registration (end)
+
+\section{Configuration}
+\label{sec:service_configuration}
+
+% The registration operation is done by the Registrants and described in Section \ref{service_registrantion}.
+
+% section configuration (end)
 

### 2009-05-12 doc/tech_doc/jura/jura-tech-doc.tex
added to/changed tech doc

@@
 \label{implement}
-JURA is written merely in C++, built with standard GNU tools, and
+JURA is written merely in C++, built with widely used GNU tools, and
 tested in a GNU/Linux environment. It depends on HED libraries: the
@@
 
-%TODO implement!
 where \textit{size} is a positive integer, the number of URs sent
@@
 
+The example below is a part of the ``\textit{grid-manager}'' block of
+the A-REX configuration. It enables logging of URs to two hosts, using
+the host credential files (placed in the standard locations), with a
+maximum of 50 URs per batch. Job log files expire after a week.
 %EXAMPLE!
+\begin{verbatim}
+...
+jobreport="https://luts1.nordugrid.org:8443/wsrf/services/sgas/LUTS"
+jobreport="https://luts2.nordugrid.org:8443/wsrf/services/sgas/LUTS 7"
+jobreport_credentials="/etc/grid-security/hostkey.pem /etc/grid-security/hostcert.pem 
+     /etc/grid-security/certificates"
+jobreport_options="urbatch:50"
+...
+\end{verbatim}
 
@@
 Although complete, JURA still has minor imperfections, some of them
-stemming from flaws in A-REX or the \textit{arcsub} client:
+stemming from limitations or operational characteristics of A-REX or
+the \textit{arcsub} client:
 
@@
 \item Handling of runtime environment string is missing.
-\item Detailed user identity information based on proxy or other
-  submitted credentials is missing from URs.
+\item Detailed user identity information based on the X.509 proxy
+  certificate content or other submitted credentials is missing from
+  URs.
 \end{itemize}

### 2009-05-14 doc/tutorials/ws-programming-tutorial/doc/tex_appendix/appendix.tex
Reorganizing doc catalogue and enabling make doc
Added a modified doc folder refering to the new folder structure of the examples

--- 2009-03-05 doc/ws-programming-tutorial/doc/tex_appendix/appendix.tex
@@
  \item Documentation of the ARC storage system~\cite{Nagy_2008}.
+ \item Eclipse WTP 1.5.1, Introduction to the WSDL Editor, \url{http://wiki.eclipse.org/index.php/Introduction_to_the_WSDL_Editor}
 
@@
 	caption={[XML schema of configuration files.]\textbf{XML schema of configuration files.}}
-	]{../../../src/hed/libs/loader/mcc.xsd}
+	]{../../../../src/hed/libs/loader/mcc.xsd}
 
@@
 	caption={[XML schema of the TCP component.]\textbf{XML schema of the TCP component.}}
-	]{../../../src/hed/mcc/tcp/tcp.xsd}
+	]{../../../../src/hed/mcc/tcp/tcp.xsd}
 
@@
 	caption={[XML schema of the TLS component.]\textbf{XML schema of  the TLS component.}}
-	]{../../../src/hed/mcc/tls/tls.xsd}
+	]{../../../../src/hed/mcc/tls/tls.xsd}
 
@@
 	caption={[XML schema of the HTTP component.]\textbf{XML schema of the HTTP component.}}
-	]{../../../src/hed/mcc/http/http.xsd}
+	]{../../../../src/hed/mcc/http/http.xsd}
 

### 2009-05-14 doc/tutorials/ws-programming-tutorial/doc/tex_echoservice/echoservice.tex
Added a modified doc folder refering to the new folder structure of the examples

@@
 	]
-{../src/services/echoservice/echo.wsdl}
+{../examples/src/services/echoservice/echo.wsdl}
 \task{TODO: Validate the WSDL Listing}
@@
 	]
-{../src/services/echoservice/echoservice.cpp}
+{../examples/src/services/echoservice/echoservice.cpp}
 
@@
 	]
-{../src/services/echoservice/arched_echoservice.xml}
+{../examples/src/services/echoservice/arched_echoservice.xml}
 
@@
 	]
-{../src/clients/echoclient/echoclient.cpp}
+{../examples/src/clients/echoclient/echoclient.cpp}
 

### 2009-05-14 doc/tutorials/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Added a modified doc folder refering to the new folder structure of the examples

@@
 	]%
-{../src/services/timeservice/timeservice.h}%
+{../examples/src/services/timeservice/timeservice.h}%
 The class \textit{Service} renders the possibility to implement the method \textit{process}.
@@
 	]%
-{../src/services/timeservice/timeservice.cpp}%
+{../examples/src/services/timeservice/timeservice.cpp}%
 
@@
 	]
-{../src/services/timeservice/arched_timeservice.xml}
+{../examples/src/services/timeservice/arched_timeservice.xml}
 
@@
 	]
-{../src/clients/timeclient/timeclient.cpp}
+{../examples/src/clients/timeclient/timeclient.cpp}
 

### 2009-05-14 doc/tutorials/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Added a modified doc folder refering to the new folder structure of the examples

--- 2009-03-17 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
@@
 	]
-{../src/services/tlsechoservice/arched_tls_echoservice.xml}
+{../examples/src/services/tlsechoservice/arched_tls_echoservice.xml}
 The first change is in line~\ref{lst_code:tls_echo_arched_xml_plugin}.
@@
 %	]
-%{../src/services/tlsechoservice/tlsechoservice.cpp}
+%{../examples/src/services/tlsechoservice/tlsechoservice.cpp}
 %
@@
 	]
-{../src/clients/tlsechoclient/tlsechoclient.cpp}
+{../examples/src/clients/tlsechoclient/tlsechoclient.cpp}
 

### 2009-05-14 doc/tech_doc/jura/jura-tech-doc.tex
Documenting new features

@@
 
+An archiving functionality allows to store generated URs in a
+specified directory (see Section \ref{config}) on the disk. If
+enabled, valid UR XMLs are written to files named
+``\textit{usagerecord.$<$ngjobid$>$.$<$random$>$}'', where
+``\textit{ngjobid}'' and ``\textit{random}'' match those of the source
+job log file. If a job log file is processed repeatedly -- for example
+because of temporary connection failures to a LUTS service -- and a
+respective UR archive file already exists, then the UR is not
+generated again. Instead, the contents of the archive file are used
+without change (NB: the creation time stamp is also retained).
+
 After successful submission to a reporting destination, the job log
@@
 \section{Security}
-The JURA executable runs with the same user priviliges as the A-REX,
+The JURA executable runs with the same user privileges as the A-REX,
 typically as \textit{root}. The owner of a job log file is the local
@@
 order to enable adding other types of accounting services besides the
-curently only supported LUTS. An abstract interface class called
+currently only supported LUTS. An abstract interface class called
 ``\textit{Destination}'' represents a usage reporting destination
@@
 The \textit{jobreport\_options} variable allows passing a generic
-option string to JURA verbatim. This string can be used for example
-for service-related settings. Currently only SGAS LUTS service is
-supported, for which the UR batch size can be set with the following
-configuration line:
+option string to JURA verbatim. This string is interpreted by JURA as
+a comma-separated list of ``\textit{name:value}'' pairs (note the
+colon!), which represent service-related settings and extended
+reporting parameters. The job reporting options currently defined are:
 
-\verb|jobreport_options="urbatch:size"|
-
-where \textit{size} is a positive integer, the number of URs sent
-within one request. If not given, the default is 50. Zero value means
-unlimited batch size.
+\begin{itemize}
+\item \textbf{\textit{urbatch}}\textit{:size} -- sets the maximal
+  number of URs in the batch sent within one request. Zero value means
+  unlimited batch size. Default is 50.
+\item \textbf{\textit{archiving}}\textit{:dir} -- enables archiving of
+  generated URs in the given directory. If the directory does not
+  exist, an attempt is made to create it. If this option is absent, no
+  archiving is performed.
+\end{itemize}
 
@@
 the host credential files (placed in the standard locations), with a
-maximum of 50 URs per batch. Job log files expire after a week.
+maximum of 50 URs per batch. Generated URs are archived in the
+directory ``\textit{/var/urs}''.
+Job log files expire after a week.
 %EXAMPLE!
@@
      /etc/grid-security/certificates"
-jobreport_options="urbatch:50"
+jobreport_options="urbatch:50,archiving:/var/urs"
 ...
@@
   though the information is already collected in the job log files.
-\item Handling of runtime environment string is missing.
+%\item Handling of runtime environment string is missing.
 \item Detailed user identity information based on the X.509 proxy
@@
 
-There is also plan to extend JURA with ther following features:
+There is also plan to extend JURA with the following features:
 
 \begin{itemize}
-\item An archiving functionality, storing generated URs in a local
-  directory specified in the configuration.
+%\item An archiving functionality, storing generated URs in a local
+%  directory specified in the configuration.
 \item Support for other accounting systems besides LUTS, especially
@@
 
-\begin{tabular}{|p{0.22\columnwidth}|p{0.22\columnwidth}|p{0.54\columnwidth}|}
+\begin{tabular}{|p{0.21\columnwidth}|p{0.21\columnwidth}|p{0.56\columnwidth}|}
 \hline
@@
 nodecount&
-Number of nodes (physical machines) involved in the job.\\
+Number of nodes (physical machines) involved in the job\\
 \hline
@@
 Quality of service associated with usage\\
-
-%\hline
-%\hline
-%\textbf{Extra UR Property}&
-%\textbf{Source\newline(job log entry)}&
-%\textbf{Description}\\
-%\hline\hline
+\hline
+\hline
+\textbf{Extended\newline UR Property}&
+\textbf{Source\newline(job log entry)}&
+\textbf{Description}\\
+\hline\hline
 %UserProxy&
@@
 %\hline
-%RuntimeEnvironment&
-%runtimeenvironment&
-%Runtime environment info\\
+RuntimeEnvironment&
+runtimeenvironment&
+Requested runtime environment, specified in job description\newline
+(RTEs from space-separated list put into separate elements)\\
 

### 2009-05-14 doc/manuals/janitor/images/dependencies.tex
Introduced a manual for Janitor.

\begin{tikzpicture}[>=latex,line join=bevel,]
  \pgfsetlinewidth{1bp}
%%
\pgfsetcolor{black}
  % Edge: Installer -> Util
  \draw [->] (711bp,144bp) .. controls (710bp,136bp) and (710bp,127bp)  .. (710bp,108bp);
  % Edge: Janitor -> Filefetcher
  \draw [->] (732bp,221bp) .. controls (750bp,211bp) and (776bp,194bp)  .. (805bp,175bp);
  % Edge: Catalog -> Tag
  \draw [->] (390bp,153bp) .. controls (417bp,143bp) and (463bp,125bp)  .. (501bp,108bp) .. controls (503bp,107bp) and (504bp,107bp)  .. (515bp,101bp);
  % Edge: BaseSystem -> Note
  \draw [->] (458bp,72bp) .. controls (458bp,64bp) and (458bp,55bp)  .. (458bp,36bp);
  % Edge: Installer -> Execute
  \draw [->] (727bp,147bp) .. controls (737bp,137bp) and (749bp,124bp)  .. (766bp,105bp);
  % Edge: Janitor -> Catalog
  \draw [->] (686bp,229bp) .. controls (625bp,216bp) and (472bp,184bp)  .. (392bp,167bp);
  % Edge: Catalog -> MetaPackage
  \draw [->] (341bp,155bp) .. controls (315bp,147bp) and (274bp,131bp)  .. (245bp,108bp) .. controls (224bp,90bp) and (207bp,65bp)  .. (192bp,36bp);
  % Edge: Janitor -> TarPackage
  \draw [->] (685bp,231bp) .. controls (622bp,225bp) and (462bp,207bp)  .. (330bp,180bp) .. controls (214bp,155bp) and (185bp,147bp)  .. (73bp,108bp) .. controls (71bp,107bp) and (69bp,106bp)  .. (57bp,101bp);
  % Edge: Filefetcher -> Execute
  \draw [->] (815bp,145bp) .. controls (810bp,137bp) and (803bp,126bp)  .. (792bp,107bp);
  % Edge: Janitor -> Config
  \draw [->] (739bp,232bp) .. controls (793bp,226bp) and (913bp,211bp)  .. (938bp,180bp) .. controls (953bp,160bp) and (938bp,132bp)  .. (917bp,105bp);
  % Edge: Janitor -> Execute
  \draw [->] (724bp,217bp) .. controls (731bp,207bp) and (741bp,193bp)  .. (748bp,180bp) .. controls (758bp,160bp) and (767bp,135bp)  .. (776bp,108bp);
  % Edge: Catalog -> TarPackage
  \draw [->] (339bp,159bp) .. controls (287bp,153bp) and (169bp,137bp)  .. (73bp,108bp) .. controls (71bp,107bp) and (68bp,106bp)  .. (56bp,102bp);
  % Edge: Catalog -> DebianPackage
  \draw [->] (392bp,156bp) .. controls (431bp,147bp) and (509bp,128bp)  .. (573bp,108bp) .. controls (576bp,107bp) and (579bp,106bp)  .. (592bp,102bp);
  % Edge: Installer -> Config
  \draw [->] (736bp,153bp) .. controls (770bp,140bp) and (831bp,117bp)  .. (878bp,99bp);
  % Edge: Tag -> Note
  \draw [->] (521bp,75bp) .. controls (509bp,65bp) and (494bp,51bp)  .. (474bp,33bp);
  % Edge: MetaPackageCollection -> MetaPackage
  \draw [->] (151bp,72bp) .. controls (155bp,64bp) and (162bp,54bp)  .. (172bp,36bp);
  % Edge: Common -> Config
  \draw [->] (902bp,144bp) .. controls (902bp,136bp) and (902bp,127bp)  .. (902bp,108bp);
  % Edge: Janitor -> DebianPackage
  \draw [->] (697bp,219bp) .. controls (688bp,208bp) and (676bp,194bp)  .. (667bp,180bp) .. controls (654bp,160bp) and (642bp,135bp)  .. (630bp,108bp);
  % Edge: Catalog -> Package
  \draw [->] (349bp,148bp) .. controls (336bp,137bp) and (319bp,123bp)  .. (298bp,104bp);
  % Edge: Catalog -> NoteCollection
  \draw [->] (366bp,144bp) .. controls (366bp,136bp) and (366bp,127bp)  .. (366bp,108bp);
  % Edge: Janitor -> Installer
  \draw [->] (712bp,216bp) .. controls (712bp,208bp) and (712bp,199bp)  .. (712bp,180bp);
  % Edge: Janitor -> Util
  \draw [->] (698bp,219bp) .. controls (690bp,208bp) and (680bp,194bp)  .. (676bp,180bp) .. controls (671bp,164bp) and (671bp,159bp)  .. (676bp,144bp) .. controls (679bp,134bp) and (684bp,123bp)  .. (695bp,106bp);
  % Edge: Execute -> InstallLogger
  \draw [->] (781bp,72bp) .. controls (781bp,64bp) and (781bp,55bp)  .. (781bp,36bp);
  % Edge: Janitor -> Common
  \draw [->] (737bp,226bp) .. controls (767bp,217bp) and (821bp,199bp)  .. (866bp,180bp) .. controls (868bp,179bp) and (870bp,178bp)  .. (881bp,173bp);
  % Edge: NoteCollection -> Note
  \draw [->] (386bp,74bp) .. controls (399bp,63bp) and (418bp,49bp)  .. (440bp,32bp);
  % Edge: Catalog -> BaseSystem
  \draw [->] (384bp,148bp) .. controls (397bp,138bp) and (415bp,123bp)  .. (439bp,105bp);
  % Edge: Catalog -> MetaPackageCollection
  \draw [->] (341bp,154bp) .. controls (306bp,142bp) and (238bp,121bp)  .. (181bp,103bp);
  % Node: Note
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (458bp,18bp) ellipse (27bp and 18bp);
  \draw (458bp,18bp) node {Note};
\end{scope}
  % Node: Execute
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (781bp,90bp) ellipse (27bp and 18bp);
  \draw (781bp,90bp) node {Execute};
\end{scope}
  % Node: Package
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (281bp,90bp) ellipse (27bp and 18bp);
  \draw (281bp,90bp) node {Package};
\end{scope}
  % Node: DebianPackage
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (623bp,90bp) ellipse (41bp and 18bp);
  \draw (623bp,90bp) node {DebianPackage};
\end{scope}
  % Node: Config
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (902bp,90bp) ellipse (27bp and 18bp);
  \draw (902bp,90bp) node {Config};
\end{scope}
  % Node: MetaPackage
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (183bp,18bp) ellipse (37bp and 18bp);
  \draw (183bp,18bp) node {MetaPackage};
\end{scope}
  % Node: NoteCollection
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (366bp,90bp) ellipse (40bp and 18bp);
  \draw (366bp,90bp) node {NoteCollection};
\end{scope}
  % Node: Util
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (709bp,90bp) ellipse (27bp and 18bp);
  \draw (709bp,90bp) node {Util};
\end{scope}
  % Node: Catalog
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (366bp,162bp) ellipse (27bp and 18bp);
  \draw (366bp,162bp) node {Catalog};
\end{scope}
  % Node: Tag
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (537bp,90bp) ellipse (27bp and 18bp);
  \draw (537bp,90bp) node {Tag};
\end{scope}
  % Node: Filefetcher
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (826bp,162bp) ellipse (31bp and 18bp);
  \draw (826bp,162bp) node {Filefetcher};
\end{scope}
  % Node: Common
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (902bp,162bp) ellipse (27bp and 18bp);
  \draw (902bp,162bp) node {Common};
\end{scope}
  % Node: MetaPackageCollection
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (140bp,90bp) ellipse (58bp and 18bp);
  \draw (140bp,90bp) node {MetaPackageCollection};
\end{scope}
  % Node: Installer
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (712bp,162bp) ellipse (27bp and 18bp);
  \draw (712bp,162bp) node {Installer};
\end{scope}
  % Node: Janitor
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (712bp,234bp) ellipse (27bp and 18bp);
  \draw (712bp,234bp) node {Janitor};
\end{scope}
  % Node: TarPackage
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (32bp,90bp) ellipse (32bp and 18bp);
  \draw (32bp,90bp) node {TarPackage};
\end{scope}
  % Node: BaseSystem
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (458bp,90bp) ellipse (34bp and 18bp);
  \draw (458bp,90bp) node {BaseSystem};
\end{scope}
  % Node: InstallLogger
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (781bp,18bp) ellipse (35bp and 18bp);
  \draw (781bp,18bp) node {InstallLogger};
\end{scope}
%
\end{tikzpicture}

### 2009-05-14 doc/manuals/janitor/tex_appendix/appendix.tex
Introduced a manual for Janitor.

\chapter{Appendix}

\section{Useful tutorials and documentations}

\begin{itemize}
 \item \textbf{Another document describing Janitor.}\\
    D2.5-1 RDF Based Semantic Runtime Environment (RE) Description And Dynamic RE Management Framework Including Creating Proof Of Concept Bioinformatics REs, Daniel Bayer and Steffen Möller and Frederik Orellana\cite{BAYER_2007}

\end{itemize}

% * Literature
% ** www.nordugrid.org/documents/xrsl.pdf
% ** Literature for JSDL is missing
% ** ``"D2.5-1 RDF Based Semantic Runtime Environment (RE) Description And Dynamic RE Management Framework Including Creating Proof Of Concept Bioinformatics REs" (see also prototype code and code documentation), May 2007''
%    http://www.knowarc.eu/documents/Knowarc_D2.5-1_07.pdf
% ** Code documentation http://www.knowarc.eu/download/D2.5-1_documentation.pdf

### 2009-05-14 doc/manuals/janitor/tex_concept/concept.tex
Introduced a manual for Janitor.

\chapter{Programming concept}

 The main language for the implementation of the functionality of the
 dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
 integration with the Grid Manager and the Web server) for the Janitor. In
 the pre-web-service implementation the Catalog remains a static web page.
 The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
 The modules can be separated into two
 functional groups. One addresses the retrieval of information from
 the Catalog RDF file in the left major branch of the figure. The other
 addresses the process of fetching and installing the packages.

 \begin{landscape}
\begin{figure}[!h]
\vspace{4cm}
 \begin{center} 
    \resizebox{24cm}{!}{\input{images/dependencies}}
    \mycaption{Modules of the Janitor and their dependencies}{}
 \end{center} 
\vfill
 \label{fig:janitorDependencies} 
\end{figure}
 \end{landscape}

In order to get a more detailed view on the full functionality of the envisioned system it is suggested to consult the 
Design Document\footnote{\href{http://www.knowarc.eu/documents/Knowarc_D1.1-1_07.pdf}
{http://www.knowarc.eu/documents/Knowarc\_D1.1-1\_07.pdf}}.

\section{Runtime environments states}
A main motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.
With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

These states are specific for every CE and communicated
between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
possible states, while Figure~\ref{fig:RE_states}
displays the transitions between the states that a Runtime Environment may
be in during its life time at a particular CE. \task{?at a particular CE?}

\begin{table}[!h]
 \begin{center}
 \begin{tabular}{lp{12cm}}
 State&Description\\
 \hline
 UNAVAILABLE&
	The RE is not available for the BaseSystem (see \ref{sec:rdfschema}) the site uses. \\
 INSTALLABLE&
	The RE is available for the BaseSystem the site uses and it
	will be automatically installed once a job requests it. \\
 INSTALLING/a&
	A job requested the RE and it is currently being installed \\
 INSTALLING/m&
	The RE-adminstrator requested the installation of the RE. Its
	currently being installed. \\
 FAILED&
	The installation process failed. \\
 INSTALLED/a&
	The RE is installed dynamically. \\
 INSTALLED/m&
	The RE ist installed manually by the RE-administrator \\
 BROKEN/m&
	The RE is installed but failed tests of the RE-administrator \\
 VALIDATED/m&
	The RE is installed and successfully passed the tests of the
	RE-administrator \\
 REMOVAL PENDING&
	The RE is still installed but will be removed as soon as
	possible. It is not available to new jobs. \\
 REMOVING&
	The RE is currently being removed. \\
 INSTALLED/s&
	The RE was installed in the traditional way by the site administrator. \\
 BROKEN/s&
	The RE was installed in the traditional way and failed
	validation by the RE-administrator, \\
 VALIDATED/s&
	The RE was installed in the traditional way and was
	successfully verified.
 \\
 \end{tabular}
 \end{center}
 \mycaption{States a Runtime Environment can possibly be in.}{}
 \label{tab:states}
 \end{table}

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=10cm]{images/RE_states.png}
    \mycaption{Relationships between the possible states
 of Runtime Environments.}{Red arcs represent human interaction.
 The distinction between {\bf /a}, {\bf /m} and {\bf /s}
 states does not need to be visible for all clients.}
    \label{fig:RE_states}
  \end{center}
\end{figure}

The manually induced transitions are marked in red, he automated transitions in black.
A transition between states can be induced automatically (i.\,e.\ by the
advent of a job requesting a particular dynamic RE) or manually by the
site's supervisor or an individual with respective rights to use the
Janitor's web service.

Upon presentation of a package to a Catalog, a CE may
classify a package to be \texttt{INSTALLABLE} if all the dependencies
are installable or already \texttt{INSTALLED}. The installation can be
performed manually (\texttt{INSTALLING}/{\bf m}) or in an automated fashion
(\ldots/{\bf a}). Should the installation process return an error, then
the installation has \texttt{FAILED}. Once the installation succeeded,
the installed package is validated for its correctness. Should that
process fail, then the package's state it is said to be \texttt{BROKEN}.

Automatically installed packages can be removed by the automatism. A
manually installed package or one that has failed to be installed,
can only be removed upon manual induction.  The \ldots/{\bf s} states
represent those Runtime Environments that are installed in the original
manual way of RE installation in ARC 0.6.

% Runtime Environment are installed for job idenitified by a certain id.

\section{Job states}

Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
in Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached in Janitor. 
After Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
an unforeseen exception occures during that process, Janitor will drop the job from its database and set the affected runtime
environments to the state \texttt{FAILED}.

\section{Integration into AREX}

\newenvironment{note}
{\rule{1ex}{1ex}\hspace{\stretch{1}}}
{\hspace{\stretch{1}}\rule{1ex}{1ex}\\}
\begin{note}
The integration into AREX is not completed yet!
\end{note}
\begin{note}
Thus, there will be bigger changes here in this section\dots
\end{note}

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=5cm]{images/arex-stages.pdf}
    \mycaption{}{}
    \label{fig:RE_states}
  \end{center}
\end{figure}

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=12cm]{images/janitor_integration_2nd_edition.pdf}
    \mycaption{}{}
    \label{fig:RE_states}
  \end{center}
\end{figure}

\section{WebService Interface}

Default port number: 55555\\
Client command equal, except assignment of HED.xml\\
 (from /arc1/trunk:12561)

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=12cm]{images/WS_structure.pdf}
    \mycaption{}{\task{To be translated and beautificated. SVG file is missing!}}
    \label{fig:RE_states}
  \end{center}
\end{figure}

Proposal for SOAP messages:\\
 namespace: dynamicruntime or janitor\\
\task{Create WSDL files for that}
\task{Permission concepts: Depending on certificates. Certain certificates may \texttt{sweep}. Defined in service\_HED.xml. Evaluated in:?? }
\lstsetJUSTXML
\begin{lstlisting}[
        label=lst:SOAPrequest,
        caption={ [Example \textit{arc.conf} settings for janitor.]
                  \textbf{Example \textit{arc.conf} settings for janitor.}}
        ]
<Request action="SEARCH|SWEEP|LIST|DEPLOY|REMOVE|CHECK|REGISTER">
	<Initiator jobid="1234"/>  <!-- Needed for: CHECK|REGISTER|DEPLOY|REMOVE -->
                                  <!-- May contain no jobID in this case a new one will 
                                       be created and returned via the response message-->

	<Runtimeenvironment type="dynamic">      <!-- Needed for: SEARCH|REGISTER-->
		<Package name="APPS/BIO/WEKA-3.4.10"/>
		<Package name="APPS/BIO/WEKA-3.4.11"/>
	</Runtimeenvironment>
                                  <!-- SWEEP and LIST only works, if the TLS-adminstrator identity 
                                       (which is assigned in the arched configuration file) is
                                       to be found by the SecHandler. Both need neither initiator 
                                       nor runtimeenvironment elements -->
</Request>
\end{lstlisting}

\lstsetJUSTXML
\begin{lstlisting}[
        label=lst:SOAPrequest,
        caption={ [Example \textit{arc.conf} settings for janitor.]
                  \textbf{Example \textit{arc.conf} settings for janitor.}}
        ]
<response action="SEARCH|SWEEP|LIST|DEPLOY|REMOVE|CHECK|REGISTER">
	<initiator jobid="1234"/>  <!-- Needed for REMOVE|REGISTER|DEPLOY|CHECK-->
	<result code="0" message="Sucessfully initailized job."> <!-- -->
	<jobs> <!--LIST|CHECK-->
		<job jobid="1234">
			<created>1234567890</created>    <!-- in unix time-->
			<age>0</age>           <!-- in seconds-->
			<runtimeenvironment>
				<package>APPS/BIO/WEKA-3.4.10</package>
			</runtimeenvironment>
			<state>INITIALIZED</state>
		</job>
		<job jobid="4321">
			<created>1234567891</created>
			<age>0</age>
			<package>APPS/BIO/WEKA-3.4.10</package>
			<state>INITIALIZED</state>
			<runtimeenvironmentkey>APPS_BIO_WEKA_3_4_10-835614b62c98c4eb6cb03d74d3161b5d</runtimeenvironmentkey> <!-- at least CHECK -->
			<uses>/nfshome/knowarc/dredesign/src/services/dRE3/perl/spool/runtime/jre__57T1ke1UVz/runtime</uses> <!-- at least CHECK -->
			<uses>/nfshome/knowarc/dredesign/src/services/dRE3/perl/spool/runtime/weka_wHfyytarlE/runtime</uses> <!-- at least CHECK -->
		</job>
	</jobs>

	<runtimeenvironment type="local">      <!-- Needed for: LIST|SEARCH -->
		<package name="APPS/BIO/MUSTANG-3.0-1"/>
		<package name="APPS/BIO/EXONERATE-2.1.0-1"/>
	</runtimeenvironment>

	<runtimeenvironment type="dynamic">      <!-- Needed for: LIST -->
		<package name="APPS/BIO/WEKA-3.4.11">
			<state>INSTALLED_A</state>
			<lastused>1234567890</lastused>
			<jobid>1234</jobid>
		</package>
		<package name="APPS/BIO/WEKA-3.4.10">
			<state>INSTALLED_A</state>
			<lastused>1234567890</lastused>
			<jobid>1234</jobid>
			<jobid>4321</jobid>
		</package>
	</runtimeenvironment>

	<runtimeenvironment type="installable">      <!-- Needed for: LIST -->
		<package name="APPS/GRAPH/POVRAY-3.6">
			<description>The Persistence of Vision Raytracer</description>
			<lastupdate>1234567890</lastupdate>
		</package>
		<package name="APPS/BIO/WEKA-3.4.8A">
			<description>WEKA Machine Learning Software</description>
			<lastupdate>1234567890</lastupdate>
		</package>
	</runtimeenvironment>

</response>
\end{lstlisting}

\section{Janitor file system permissions}

Currently the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
a setuid wrapper has been written. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s}
to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\

Thus, Janitor is not executed directly. Instead a dynamic link will be created in \task{libexec or sbin} which is pointing to
the wrapper.\\

It is suggested to create a user and a group "janitor" for the Janitor.\\

% Currently the Janitor must be executed as root. For this a suid root helper
% is needed. It is provided in /opt/janitor/rjanitor.c. If you installed
% the Janitor in another directory then /opt/janitor change the parameter
% of the execv command in rjanitor.c. Then compile the file with\\
% \texttt{\textdollar gcc -o rjanitor rjanitor.c}\\
% and run as root\\
% \texttt{\textdollar chown root:root rjanitor}\\
% \texttt{\textdollar chmod u+s rjanitor}\\
% to make the binary suid root. A future version of the Janitor will get rid
% of the suid root helper.\\
% 
% The Grid Manager does not call the Janitor directly in /opt/janitor but
% calls /opt/nordugrid/libexec/janitor. So a link is needed (as root):\\
% \texttt{\textdollar ln -s /opt/janitor/rjanitor /opt/nordugrid/libexec/janitor}\\
% \\
% It is suggested to create a user and a group "janitor" for the Janitor.\\
% \\
% The Janitor needs a directory for storing information on installed REs:\\
% \texttt{\textdollar mkdir -p /var/lib/nordugrid/janitor}\\
% \texttt{\textdollar chown janitor:janitor /var/lib/nordugrid/janitor}\\
% \\
% It also needs a directory on the shared volume for installing the REs:\\
% % \texttt{\textdollar mkdir -p $<$shared_volume$>$/janitor}\\
% % \texttt{\textdollar chown janitor:janitor $<$shared_volume$>$/janitor}

\subsection{Security Consideration}

Security is a major concern for the grid systems. An automatic software
installation inheritently introduces security threats. This section
addresses those and describes the available solutions to limit
security risks.

In the current installation, every user authorised to execute a job is
also authorised to install a REs. Restrictions are
only imposed on the set of DRE that are available for installation.
Restrictions are imposed by the site admininistrators on the descriptions
that are given by the Catalog that is offering the package. These
descriptions may explicitly mention DRE names, a regular expression on
these, or refer to tags of packages that categorise these. However,
the core of these controls lie with the maintainers of the Catalog,
who needs to be trusted.

All dynamic REs are installed in separate directories. The provisioning of
disk space is the duty of the site administrator.  In the current
implementation, the installation is completely transparent to the user:

\begin{itemize}
\item DREs are not distinguished between {\em installed} and
  {\em installable} in the information system.
\item No status information is given at the time an DRE is installed.
\end{itemize}

Malevolent regular users with respective training in using system exploits
to gain root access are likely to find security holes by regularly
submitted scripts.  The authentication and authentification of users,
together with respective logging, is the major defense against such
attacks. What is consequently left to be protected against are unwanted
sideeffects by the installation of software.

The worst case scenario would be the installation of a RE
that overwrites system files. With the current implementation, which is
based solely on tar files, this is barely possible, unless such is
performed by the install scripts that accompany the tar files. However,
hereto the installation would have to be performed by a user with system
priviledges, for which there is no technical requirement.

The installation of packages from the Debian distribution (or other
packages of mainstream Linux distributions) is seeked to reduce the
complexity and burden in the maintenance for DRE. In the
current implementation Debian packages may be installed only by their
transformation into tar files. With the advent of the interface for
the virtualisation of the grid infrastructure, it is anticipated to
work with native packages of the Debian Linux distribution. The reuse
of packages that passed many eyeballs - as it is the case with packages
from major Linux distribution - security is further increased or becomes
as high as with the operating system underneath virtual clients.

Summarising, there is general concern about the security of grid
computing.  Dynamic REs introduce new dangers since a
manual control at the grid site is substituted by a remote process that
is out the direct supervision of a local site administrator. The signing
of packages by known and directly or indirectly trusted developers is
a good indicator that no malevolent individuals have tampered with the
binary. The site administrators can limit the sources of packages and
specify packages that are eligible or excluded from installations.

% * Programming concept
% ** design decisions: Perl, RDF, Two classes entities with states: jobs and rtes
% ** State transitions of RTEs
% ** State transitions of Jobs
% ** RDF based knowledge base for runtime environments
% ** File based recovery of Jobs and RTEs
% ** Class structure
% ** Interaction with A-REX
%

### 2009-05-14 doc/manuals/janitor/tex_future/future.tex
Introduced a manual for Janitor.

\chapter{Outlook}
% 
% * Future work
% ** API and WebService-Interface for Janitor
% ** Glue2 specification for RTE states
%

\section{Representation of dynamic REs in the information model}
Dynamic REs require an extended representation in the
information model.  The Application Software description should be able
to distinguish installed REs from installable REs, potentially offer
description of extended RE state-like information. This work is planned
to be carried out as part of the Glue-2.0 effort of OGF\footnote{OGF
GLUE: https://forge.gridforum.org/sf/projects/glue-wg}.

\section{Integration with Workflow Management}

Future development of ARC aims at integrating grid computing with
workflow tools for the web services that have a growing user base in
bioinformatics. The challenge is to prepare REs for
programs or databases and to offer such concisely to users of the
workflow environments.  In the bioinformatics community, such are
today offered as web services.  This anticipated development instead
fosters the dynamic installation on the grid whenever appropriate to
allow for special computational demands in high-throughput analyses.
Conversely, because of the increased complexity of workflows with respect
to the already today not manually manageable number of REs, without an
automatism for the automated installation of software packages on the
grid, the use of workflows in grid computing seems mute.

\section{Implementation of a Catalog service}

A Catalog service is planed to be implemented on top of the ARC
HED component.  This service will render the currently used locally
accessible RDF file externally accessible. Selected users are then allowed
to remotely add/edit/remove REs to/from to it.  The Janitor will access
the content of the Catalog through a well-defined Web Service interface.

\section{Integration with the Virtualization work}

The RDF schema nicely prepares for the upcoming virtualisation of worker
nodes.  How exactly the dynamics are integrated will depend on how
dynamic the virtualisation of the nodes is. In the simplest scenario,
a worker node's CPU will only be occupied by a single virtual machine
and that will not be changed. In this case, there is no difference to
the setup of the Janitor with today's static setups.

However, if the BaseSystems can be substituted dynamically, then a RE
can possibly be offered via multiple BaseSystems. The RDF
Schema describes BaseSystems as separate instances and as such differs
from the current RE registry.  Heuristics that prefer one BaseSystem for
another can make direct use of the data that is presented in the schema.
The integration of packages from Linux distributions in the description
of REs is essential to have a means to decide for the equivalence of
manual additions and the functionality that comes with BaseSystem.

TODO:\\
\begin{itemize}
 \item There is a arc.conf file in which all possiblie flags are listed... ADD THE JANITOR FLAGS!!!
 \item Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change that
 \item Verfication
\end{itemize}

### 2009-05-14 doc/manuals/janitor/tex_installation/installation.tex
Introduced a manual for Janitor.

\chapter{Installation}

% ** Dependencies
% *** Ordianary: log4perl and ``Redland RDF Library - Perl Interface''
% *** WebService: libperl-dev
% *** Protege: For maintaining the knowledge base

In order to use Janitor, the two perl packages listed in table~\ref{tab:install_dependencies} are required. To have the WebService interface for Janitor, the packages listed in the table~\ref{tab:install_dependencies_optional} 
have to be installed before the build process is.
\begin{table}[!h]
   \begin{center}
        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
                   Redland RDF libraray is used for accessing the knowledge base of RTEs.}
        \label{tab:install_dependencies}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   liblog-log4perl-perl & \textit{Log4perl is a port of the log4j logging package}\\
	\hline
	   librdf-perl          & \textit{Perl language bindings for the Redland RDF library}\\
	\hline
	\end{tabular} 
   \end{center}
\end{table}
\begin{table}[!h]
   \begin{center}
        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
                   WebService to the Perl interpreter.}
        \label{tab:install_dependencies_optional}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   libperl-dev & \textit{ Perl library: development files}\\
	\hline
	\end{tabular}
   \end{center}
\end{table}
\forcelinebreak

Janitor will be installed by default along with ARC1. If desired it is possible to disable the shipping of Janitor 
using the \textit{configure} flags \textit{--disable-janitor-service} for the complete janitor or 
\textit{--disable-janitor-webservice} for only the Web Service support.
Furthermore it is recommended to install the ontology editor 
Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
knowledge database of installable packages.

\section{Configuration}\label{sec:janitor_configuration}

The current version of Janitor can be configured using the common file \textit{arc.conf} which is to be found in the 
configuration directory \textit{etc}. Janitor is using the environment variable \textit{NORDUGRID\_CONFIG} to determine the location 
of the corresponding file. If the variable is not set, the default location \textit{/etc/arc.conf} will be used. 
The configuration is assigned by the section \lbrack janitor\rbrack. The table~\ref{tab:arc_conf_tags} contains the available tags
for janitor configuration.
\begin{landscape}
\begin{table}[!h]
   \begin{center}
        \mycaption{Tags usable in \textit{arc.conf} within the section janitor.}{Tags usable in \textit{arc.conf} within the section janitor.}
	\label{tab:arc_conf_tags}
	\begin{longtable}{|p{3cm}|p{10cm}||p{10cm}|}
	\hline
	   \textbf{tag}    & \textbf{example}                      & \textbf{description}\\
        \hline
           enabled         & "1"                                   & Boolean flag which enables or disables janitor in A-REX.\\
	   uid             & "root"                                & The effective uid. \\
	   gid             & "0"                                   & The effective gid. \\
	   registrationdir & "/var/spool/nordugrid/janitor"        & Directory where we the current states of jobs are kept. \\
	   catalog         & "/var/spool/nordugrid/janitor/catalog/knowarc.rdf"& URL of the catalog containg the package information.\\
	   downloaddir     & "/var/spool/nordugrid/janitor/download" & Directory for downloads \\
	   installationdir & "/var/spool/nordugrid/janitor/runtime"& Directory for installation of packages                   \\
	   jobexpirytime   & "7200"                                & If a job is older than this, it is considered dead and assigned to be removal pending.\\
	   rteexpirytime   & "36"                                  & If a runtime environment was not used for this time, it will be assigned to be removal pending.\\
	   allow\_base     & "*"                                   & Allow rule for base packages. \\
	   deny\_base      & "debian::etch"                        & Deny rule for base packages.\\
	   allow\_rte      & "*"                                   & Allow rule for base packages. \\
	   deny\_rte       & "APPS/MATH/ELMER-5.0.2"               & Deny rule for base packages. \\
	   logconf         & "/opt/nordugrid/etc/log.conf"         & Location of the logging configuration file for janitor.\\
	\hline
	\end{longtable}
   \end{center}
\end{table}
\end{landscape}

The parameter \texttt{enable} defines whether Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
In these directories the states of the jobs and the runtime environments will be stored. 
The knowledge base of installable packages is specified by the parmeter \texttt{catalog}. 
Its value can be any kind of URL pointing to an file written in the Resource Description Framework (RDF) format.
The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
The parameter \texttt{downloaddir} assignes the directory in which the installation files will be saved after they have been 
downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
directory into which all packages will be installed. Instead of the other directories the \texttt{installationdir} should be 
available for all computing elements i.e. by using a shared volume.
If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, Janitor will also
create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by Janitor.
The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintain by
a higher organization.
The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
An examples how to configure arc and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf} and.

\lstsetCONFIGURE
\begin{lstlisting}[
        label=lst:arc_conf,
        caption={ [Example \textit{arc.conf} settings for janitor.]
                  \textbf{Example \textit{arc.conf} settings for janitor.}}
        ]
[janitor]
enabled="1"
logconf="/opt/nordugrid/etc/log.conf"
registrationdir="/var/spool/nordugrid/janitor"
installationdir="/var/spool/nordugrid/janitor/runtime"
downloaddir="/var/spool/nordugrid/janitor/download"
jobexpirytime="7200"
rteexpirytime="36"
uid="root"
gid="0"
allow_base="*"
allow_rte="*"

[janitor/nordugrid]
catalog="/var/spool/nordugrid/janitor/catalog/knowarc.rdf"
\end{lstlisting}

\lstsetCONFIGURE
\begin{lstlisting}[
        label=lst:log_conf,
        caption={ [Example \textit{log.conf} settings for janitor.]
                  \textbf{Example \textit{log.conf} settings for janitor.}}
        ]
# Master Loglevel
# [OFF | DEBUG | INFO | WARN | ERROR | FATAL]
#log4perl.threshold = OFF

log4perl.rootLogger = WARN, DebugLog, MainLog, ErrorLog
log4perl.appender.DebugLog = Log::Log4perl::Appender::Screen
log4perl.appender.DebugLog.layout = PatternLayout
log4perl.appender.DebugLog.layout.ConversionPattern = [%C] %d %p> %m%n

log4perl.appender.MainLog = Log::Log4perl::Appender::File
log4perl.appender.MainLog.Threshold = VERBOSE
log4perl.appender.MainLog.filename = /var/log/janitor.log
log4perl.appender.MainLog.layout = PatternLayout
log4perl.appender.MainLog.layout.ConversionPattern = %d %p> %m%n

log4perl.appender.ErrorLog = Log::Log4perl::Appender::File
log4perl.appender.ErrorLog.Threshold = ERROR
log4perl.appender.ErrorLog.filename = /var/log/janitor_error.log

log4perl.appender.ErrorLog.layout = PatternLayout
log4perl.appender.ErrorLog.layout.ConversionPattern = %d %p> %m%n
\end{lstlisting}

% ** Configuring arc.conf
% *** Where to store the data of janitor
% ** Configuring log.conf
% *** Where to store the log of janitor

\section{Limitations}

Janitor is designed for the usage on Linux distributions.

% Is designed for linux distributions. Windows or Mac won't work (Reason wget! maybe also the Executer).

### 2009-05-14 doc/manuals/janitor/tex_usage/usage.tex
Introduced a manual for Janitor.

\chapter{Usage}

Janitor can be used either with or without A-REX. In case A-REX is used, installations will be maintained automatically the runtime environments.

\section{Janitor with A-REX}

Runtime Environments can be specified using the supported job description languages.
Representative two common languages shall be explained at this point: xRSL and JSDL.
Listing~\ref{lst:xrsl_job} shows the xRSL example in which two runtime environments are requested.

\lstsetXRSL
\begin{lstlisting}[
        label=lst:xrsl_job,
        caption={ [Job submission using the xRSL job description language.]
                  \textbf{Job submission using the xRSL job description language.}}
        ]
 &
 (executable = "run.sh" )
 (arguments = "weka.classifiers.trees.J48" "-t" "weather.arff")
 ("inputfiles" = ("weather.arff" "" ))
 ("stderr" = "stderr" )
 ("stdout" = "stdout" )
 ("gmlog" = "gmlog" )
 ("runtimeenvironment" = "APPS/BIO/WEKA-3.4.10")
 ("runtimeenvironment" = "APPS/BIO/WISE-2.4.1-5")
\end{lstlisting}

A comprehensive reference manual of the Extended Resource Specification Language (XRSL) can be
found at \href{www.nordugrid.org/documents/xrsl.pdf}{www.nordugrid.org/documents/xrsl.pdf}~\cite{NORDUGRID_MANUAL_4}.
Within Listing~\ref{lst:jsdl_job} an example using JSDL is provided. The specification of assigning runtime environments in JSDL is 
currently only defined within the nordugrid jsdl-arc schema~
\href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/a-rex/grid-manager/jobdesc/jsdl/jsdl_arc.xsd}
     {http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/a-rex/grid-manager/jobdesc/jsdl/jsdl\_arc.xsd}.
\lstsetJUSTXML
\begin{lstlisting}[
        label=lst:jsdl_job,
        caption={ [Job submission using JSDL.]
                  \textbf{Job submission using JSDL.}}
        ]
<?xml version="1.0" encoding="UTF-8"?>
<JobDefinition
  xmlns="http://schemas.ggf.org/jsdl/2005/11/jsdl"
  xmlns:posix="http://schemas.ggf.org/jsdl/2005/11/jsdl-posix"
  xmlns:arc="http://www.nordugrid.org/ws/schemas/jsdl-arc">
  <JobDescription>
    <Application>
      <posix:POSIXApplication>
        <posix:Executable>/bin/sleep</posix:Executable>
        <posix:Argument>120</posix:Argument>
      </posix:POSIXApplication>
    </Application>
    <DataStaging>
      <FileName>test.sh</FileName>
      <Source/>
      <Target/>
    </DataStaging>
    <DataStaging>
      <FileName>transferGSI-small</FileName>
      <Source>
        <URI>gsiftp://pgs02.grid.upjs.sk:2811/unixacl/transferGSI-small</URI>
      </Source>
      <Target/>
    </DataStaging>
    <Resources>
      <arc:RunTimeEnvironment>
        <arc:Name>APPS/BIO/WISE-2.4.1-5</arc:Name>
        <arc:Version><Exact>2.4.1</Exact></arc:Version>
      </arc:RunTimeEnvironment>
      <arc:RunTimeEnvironment>
        <arc:Name>APPS/BIO/APPS/BIO/WEKA-3.4.10</arc:Name>
        <arc:Version><Exact>3.4</Exact></arc:Version>
      </arc:RunTimeEnvironment>
    </Resources>
  </JobDescription>
</JobDefinition>
\end{lstlisting}

\section{Janitor without A-REX}

In addition of using Janitor together with A-REX, Janitor can also be used as a standalone commandline tool.
The available commands are listed in the Table~\ref{janitor_commandline_man}.
\begin{table}[!h]
   \begin{center}
        \mycaption{Overview about the available commands in Janitor.}{}
        \label{tab:janitor_commandline_man}
	\begin{tabular}{p{0.5cm}p{2cm}p{11cm}}
	\multicolumn{3}{l}{\textbf{janitor [COMMAND] [JOB-ID] [RTE] \dots}} \\
	\multicolumn{3}{l}{\textbf{Command:}}\\
	&	register			& Registers a job and a set of runtime environments in the Janitor database. Requires the parameters [JOB-ID] and a list of [RTE]s.\\
	&	deploy				& Downloads and installs the desired runtime environments. Requires the name of an already registered [JOB-ID].\\
	&	remove				& Removes the placeholder of the job on the runtime environments. If no more jobs are using the runtime environment and the lifespan of the runtime environment has be expired, the runtime environment can be removed using the \texttt{sweep} command. Requires the [JOB-ID] to be removed.\\
	&					&\\
	&	sweep				& Removes unused runtime environments. No further arguements are required. Using the option \texttt{--force} enforces the removal of all unused runtime environments. Runtime environments having the state FAILED will not be removed.\\
	&	setstate			& Changes the state of a dynamically installed runtime environment. This might be useful in case a runtime environment with a state FAILED shall be removed (new state might be REMOVAL\_PENDING). Requires the argument [STATE] followd by a list of [RTE]s.\\
	&					&\\
	&	search				& Performs a simple search in the catalog and the manually installed runtime environments (\texttt{runtimedir}). Requires no [JOB-ID] nor [RTE]s, but only a list of string to be searched for.\\
	&	list				& Lists all information about jobs, automatically installed runtime environments and manually installed runtime environments. No additional parameters have to be passed.\\
	&	info				& Renders information about a job. Requires the parameter [JOB-ID].\\
	\multicolumn{3}{l}{\textbf{Job id:}}\\
	&					& A unique sequence of numbers. Once Janitor registered a job id it cannot register a second job having the same job id.\\
	\multicolumn{3}{l}{\textbf{Runtime environments:}}\\
	&					& Runtime environments are defined by a continuous string. The name of valid runtime environment names can be investigated using the \texttt{list} or the \texttt{search} commands. They are defined in the catalog or by the directories and scripts of the \texttt{runtimedir} of the \texttt{grid-manager}.\\
	\end{tabular} 
   \end{center}
\end{table}
The most important commands for Janitor are \texttt{register}, \texttt{deploy} and \texttt{remove}. In order to 
register a job along with a set of runtime environments in Janitor, the first command \texttt{register} followed
by a job identifier and a list of runtime environments has to be used.
A job is identified by sequence of numbers. Runtime environments are specified by a string containing the name as it is defined 
within the catalog (resp. the runtime directory of the grid-manager).
The command \texttt{deploy} extracts the necessary dependcies of the desired runtime environments and then downloads and installs 
the required packages. %How dependecies can are defined will be explained later in the section~\ref{sec:catalog}.

% PROCEED HERE!!!!!!!! !! ! ! !
In order to remove jobs registered in Janitor the command \texttt{remove} has to be used. 
The command only removes the job entry and the lock on the runtime environment. If there are no more locks on the runtime environment
it might be deleted for real.

Easy commandline examples are provided in Listing~\ref{lst:janitor_example}.

Each command has a certain behaviour for its exit status. The Table~\ref{tab:janitor_commandline_exit_status} lists the possible
outcomes.

\begin{table}[!h]
   \begin{center}
        \mycaption{Possible exit states of Janitor}{}
        \label{tab:janitor_commandline_exit_status}
	\begin{tabular}{p{0.5cm}p{2cm}p{0.5cm}p{11cm}}
	\multicolumn{3}{l}{\textbf{Exit status:}} \\
	&\multicolumn{3}{l}{The exit status of Janitor depends on the used command.} \\
	&	register			& 0 & Registration was successful. No noteworthy occurrences.\\
	&					& 1 & Registration was successful but some runtime environments aren't installed yet. Deploy is mandatory.\\ 
	&					& 2 & An error occured.\\
	&					&   &\\
	&	deploy				& 0 & Sucessfully initialized job.\\
	&					& 1 & Can't provide requested runtime environments.\\ 
	&					&   &\\
	&	remove				& 0 & Sucessfully removed job or no such job.\\
	&					& 1 & Can't provide requested runtime environments.\\ 
	&					&   &\\
	&	sweep				& 0 & Always returns this exit code.\\
	&					&   &\\
	&	setstate			& 0 & Changing the state was successful.\\
	&					& 1 & Can not change the state.\\ 
	&					&   &\\
	&	search				& 0 & Search sucessfully finished.\\
	&					&   &\\
	&	list				& 0 & Successfully retrieved information.\\
	&					&   &\\
	&	info				& 0 & Successfully retrieved job information.\\
	&					& 1 & No such job.\\ 
	&					& 2 & Error while retrieving job information.\\
	\end{tabular} 
   \end{center}
\end{table}

\lstsetKSH
\begin{lstlisting}[
        label=lst:janitor_example,
        caption={ [Example \textit{log.conf} settings for janitor.]
                  \textbf{Example \textit{log.conf} settings for janitor.}}
        ]
# janitor register 1999 APP/BIO/JASPAR-CORE-1.0 APPS/BIO/APPS/BIO/WEKA-3.4.10
# janitor deploy 1999
# janitor remove 1999

# janitor sweep --force
# janitor setstate REMOVAL_PENDING APP/BIO/JASPAR-CORE-1.0 APPS/BIO/APPS/BIO/WEKA-3.4.10

# janitor search JASPAR WEKA
# janitor list
# janitor info 1999
\end{lstlisting}

% 
% * Usage
% ** Without A-REX
% *** Commandline (Advantage/Disadvantages)
% *** WebService (Advantage/Disadvantages)
% 
% * With A-REX
% ** Example:
% *** Simple:     runtimedir
% *** Simple:     RDF
% *** Simple job: JSDL
% *** Simple job: XRSL

\section{Example}
% 
%

### 2009-05-18 doc/tech_doc/infosys/infosys_technical.tex
Security and configuration (partial) for infosys document

@@
 
-%\subsubsection{Peer-to-Peer functions}
-%\label{ISIS Peer-to-Peer functions}
-%\subsection{Client}
-%\label{Client}
-
-%The client part of the information service is for querying the data stored in the database of the ISIS 
-%services. This is implemented as a part of the ARCLib and using the ISIS service's interface for the 
-%various data queries. For further details see Section \ref{Query data}.
-
-%\section{Operations}
-%\label{Operations}
-%The different operations are performed by different Information system components. These cases are described below in the dedicated subsections.
-
 % subsection functionality (end)
@@
 % TODO figure with mutiple service types
+
+\textit{This section extensively uses terms defined in "Security Framework of ARC1" document"~\cite{arcsec}.}
 
@@
 
+Those 3 action groups are handled using Security Framework of ARC~\cite{arcsec}. For each group corresponding 
+Action is defined for ARC policy language as descriped in table \ref{tab:policy_actions}.
+
+\begin{table}{h}
+\caption{ARC Policy Actions of ISIS}
+\label{tab:policy_actions}
+\begin{tabular}{|c|c|l|}
+\hline
+\textbf{Group}&\textbf{Action}&\textbf{AttributeId}\\
+\hline
+Request from ISIS&\textit{isis}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
+\hline
+Request from Service&\textit{service}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
+\hline
+Request from generic client&\textit{client}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
+\hline
+\end{tabular}
+\end{table}
+
+Access restrictions for clients are defined using ARC Policies which are specified and 
+processed using generic Security Handlers approach. Corresponding Security Handlers have to be 
+configured inside configuration block of corresponding ISIS service and attached 
+\textit{incoming} queue.
+
 % subsection client_authorization (end)
@@
 \label{sub:isis_authorization_information_authorization}
+
+Because all ISIS services trust each other they freely exchanging collected information. But 
+not all information stored inside ISIS cloud is public or readable by clients authorized according 
+to procedure described in section \ref{sub:isis_authorization_client_authorization}. There
+may be some pieces of information available only for specific clients. An example could be 
+information about resources serving particular Virtual Organization (VO) could be visible only to
+members of that VO.
+
+To implement functionality described above each node in agregated XML documents of information
+collected by ISIS cloud may have Access Control Policy associated with it. Access control is 
+defined at level of XML node and propagates all children nodes - similar to file systems.
+Children nodes can only additionally restrict access control imposed by parent node. For example
+if parent node A allows access only to VO1 then children node B can narrow access to Adminsitrator 
+of VO1 and can't grant access for VO2 members.
+
+By default all XML nodes are public. Access Control Policies are embedded into XML document
+as XML nodes (see Appendix \ref{annex:info_policies_schema} for schema) even if that violates schema 
+of document. Then nodes are assigned policies aby adding XML attributes refering to defined Policies.
+Access permission to particular information node is evaluated by traversing all nodes prom 
+parent to childen. At first node which gives negative result evaluation is stopped and
+this node including all its children is removed from document.
+
+Before providing results of query operation ISIS runs procedure described above on results and 
+also removes Access Control Policies. Reduced document obtained in this way is returned to
+requesting client.
 
@@
 %entities which are users of the system.
-%
- 2. From storage/access location point of view information is divided
+% 2. From storage/access location point of view information is divided
 %into:
@@
 
-The service developers have to ensure that the services are providing the necessary information about themselves.
-This is done by implementing the subclass of the Arc::Service class - the RegistrationCollector function has to provide up-to-date status information about the service and anything else it wants to be advertised.
-This information package is called \textbf{Service Advertisement}.
+The service developers have to ensure that the services are providing the necessary information about 
+themselves when registering to ISISes. This is done by implementing the subclass of the Arc::Service class - 
+the RegistrationCollector function has to provide up-to-date status information about the service and anything 
+else it wants to be advertised. This information package is called \textbf{Service Advertisement}.
 \label{service_advertisement}
-The \textbf{Service Advertisement} can contain any information service wants to advertise but the mandatory elements have to be always present:
+The \textbf{Service Advertisement} can contain any information service wants to advertise but the mandatory 
+elements have to be always present:
 \begin{itemize}
@@
 
-Because there may be multiple registration processes running in parallel it is important to ensure that implementation RegistrationCollector is thread safe or there are internal locks implemented.
+Because there may be multiple registration processes running in parallel it is important to ensure that 
+implementation of RegistrationCollector is thread safe or there are internal locks implemented.
+
+Every service registering to ISIS should also provide an interface for direct querying of information 
+describing service. Normally this information should be more detailed than one sent to ISIS. For this 
+purpose LIDI interface is defined which is a susbset of WS Resource Properties (WSRP)~\cite{wsrf-rp}. Following WSRP
+operations must be supported - \textit{GetResourcePropertDocument}, \textit{GetResourceProperty},
+\textit{GetMultipleResourceProperties} and \textit{QueryResourceProperties}.
+
+%TODO: define names of mandatory properties - at least Glue2 property
 
@@
 
-This registration operation is done once during the start-up phase and periodically according to (per service) configured periods.
+This registration operation is done once during the start-up phase and periodically according to 
+(per service) configured periods.
 
 % section registration (end)
+
+\section{Authorization}
+\label{sec:service_authorization}
+
+For each information provided by service through the LIDI interface same procedure as described in 
+Section \ref{sub:isis_authorization_information_authorization} should be applied.
+
+% section authorization (end)
 
@@
 
-% The registration operation is done by the Registrants and described in Section \ref{service_registrantion}.
+The registration operation is done by the Registrants as described in Section \ref{sec:service_registration}.
+Each Registrant is instantiated by corresponding InfoRegistrar XML element in configuration file as defined in Section \ref{annex:service_configuration_schema}.
+First InfoRegistrar defines Registrant which is used to register every service by default.
+
+Registration parameters per service are defined by InfoRegister element located inside corresponding service configuration element.
+To assign service to specific Registrant(s) ona may use Registrar configuration elements situated inside InfoRegister.
+If none Registrar element is defined service will be registed using first (default) Registrant.
+To make not register itself special configration element NoRegister has to be used.
+
+Following elements are defined inside configuration element if the Registrant:
+\begin{description}
+\item{URL}            <!-- URL specifies contact endpoint of a bootstrap Information
+                 Registration service. Further ISIS adresses will be queried
+                 from this service. -->
+\item{KeyPath, CertificatePath, ProxyPath and CACertificatesDir}
+            <!-- Optional KeyPath, CertificatePath, ProxyPath and CACertificatesDir are
+                 paths to files storing X509 credentials used for establishing connections. -->
+\item{Retry}            <!-- Retry count. Specifie how many times have an ISIS to timeout to handle it as
+                 unavailable. If there is no value specified the default value: 5 will be used. -->
+            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1"/>
+
+\end{description}
+
+% section configuration (end)
 
@@
 
-% \bibliography{grid}
+\chapter{Appendices}
+
+\section{WSDL of ISIS Specific Interface}
+\label{annex:isis_wsdl}
+\begin{verbatim}
+
+?xml version="1.0" encoding="UTF-8"?>
+<!--
+    Data types of Information Indexing Service
+-->
+
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns:isis="http://www.nordugrid.org/schemas/isis/2007/06"
+  xmlns:wsa="http://www.w3.org/2005/08/addressing"
+  targetNamespace="http://www.nordugrid.org/schemas/isis/2007/06"
+  elementFormDefault="qualified" attributeFormDefault="unqualified"
+>
+
+    <xsd:import namespace="http://www.w3.org/2005/08/addressing" schemaLocation="http://www.w3.org/2006/03/addressing/ws-addr.xsd"/>
+
+<!-- This is an initial and incomplete DRAFT which mainly concentrates on the structure but not on the actual names. Final version will use G
+LUE-2.0 ternminology. -->
+
+    <!-- ==== Input and output types for IIS operations ==== -->
+    <!-- Input type for Register operation -->
+    <xsd:complexType name="RegisterType">
+        <xsd:sequence>
+            <xsd:element name="Header" type="isis:HeaderType" minOccurs="0" maxOccurs="1"/>
+            <xsd:element name="RegEntry" type="isis:RegistrationEntryType" minOccurs="1" maxOccurs="unbounded"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="Register" type="isis:RegisterType"/>
+
+    <!-- Output type for Register operation -->
+    <xsd:complexType name="RegisterResponseType">
+        <xsd:sequence>
+            <xsd:element name="Fault" type="isis:FaultType" minOccurs="0" maxOccurs="1"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="RegisterResponse" type="isis:RegisterResponseType"/>
+
+    <!-- Input type for RemoveRegistrations operation -->
+    <xsd:complexType name="RemoveRegistrationsType">
+        <xsd:sequence>
+            <xsd:element name="MessageGenerationTime" type="xsd:string" minOccurs="1" maxOccurs="1"/>
+            <xsd:element name="ServiceID" type="xsd:string" minOccurs="1" maxOccurs="unbounded"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="RemoveRegistrations" type="isis:RemoveRegistrationsType"/>
+
+    <!-- Output type for RemoveRegistrations operation -->
+    <xsd:complexType name="RemoveRegistrationResponseElementType">
+        <xsd:sequence>
+            <xsd:element name="ServiceID" type="xsd:string" minOccurs="1" maxOccurs="1"/>
+            <xsd:element name="Fault" type="isis:FaultType" minOccurs="1" maxOccurs="1"/>
+        </xsd:sequence>
+    </xsd:complexType>
+
+    <xsd:complexType name="RemoveRegistrationsResponseType">
+        <xsd:sequence>
+            <xsd:element name="RemoveRegistrationResponseElement" type="isis:RemoveRegistrationResponseElementType" minOccurs="0" maxOccurs="
+unbounded" />
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="RemoveRegistrationsResponse" type="isis:RemoveRegistrationsResponseType"/>
+
+    <!-- Input type for GetISISList operation -->
+    <xsd:complexType name="GetISISListType">
+        <xsd:sequence>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="GetISISList" type="isis:GetISISListType"/>
+
+    <!-- Output type for GetISISList operation -->
+    <xsd:complexType name="GetISISListResponseType">
+        <xsd:sequence>
+            <xsd:element name="EPR" type="wsa:EndpointReferenceType" minOccurs="1" maxOccurs="unbounded"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="GetISISListResponse" type="isis:GetISISListResponseType"/>
+
+    <!-- Input type for Query operation -->
+    <xsd:complexType name="QueryType">
+        <xsd:sequence>
+            <xsd:element name="QueryString" type="xsd:string" minOccurs="1" maxOccurs="1"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="Query" type="isis:QueryType"/>
+    <!-- Output type for Query operation -->
+    <xsd:complexType name="QueryResponseType">
+        <xsd:sequence>
+            <xsd:any minOccurs="0" maxOccurs="unbounded"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="QueryResponse" type="isis:QueryResponseType"/>
+
+    <!-- === Helper type definitions === -->
+    <xsd:simpleType name="StatusType">
+        <xsd:restriction base="xsd:string">
+            <xsd:enumeration value="1"/>
+            <xsd:enumeration value="2"/>
+        </xsd:restriction>
+    </xsd:simpleType>
+
+    <xsd:complexType name="HeaderType">
+        <xsd:sequence>
+            <!-- Identifier of the source HED -->
+            <xsd:element name="RequesterID" type="xsd:string" minOccurs="0" maxOccurs="1"/>
+            <!-- Time of the aggregated message generation -->
+            <xsd:element name="MessageGenerationTime" type="xsd:dateTime"/>
+        </xsd:sequence>
+    </xsd:complexType>
+
+    <xsd:complexType name="RegistrationEntryType">
+        <xsd:sequence>
+            <xsd:element name="SrcAdv" type="isis:ServiceAdvertisementType" minOccurs="1" maxOccurs="1"/>
+            <xsd:element name="MetaSrcAdv" type="isis:ServiceAdvertisementMetadataType" minOccurs="1" maxOccurs="1"/>
+        </xsd:sequence>
+    </xsd:complexType>
+
+    <xsd:complexType name="ServiceAdvertisementType">
+        <xsd:sequence>
+            <!-- General part of the Service Advertisment -->
+            <xsd:element name="Type" type="isis:ServiceTypeType"/>
+            <xsd:element name="EPR" type="wsa:EndpointReferenceType"/>
+            <!-- Service specific part of the Service Advertisment -->
+            <xsd:element name="SSPair" type="isis:NameValuePairType" minOccurs="0" maxOccurs="unbounded"/>
+        </xsd:sequence>
+    </xsd:complexType>
+
+    <xsd:complexType name="ServiceAdvertisementMetadataType">
+        <xsd:sequence>
+            <!-- Globally unique and persistent ID of the service -->
+            <xsd:element name="ServiceID" type="xsd:string" minOccurs="1" maxOccurs="1"/>
+            <!-- Time of information generation or collection -->
+            <xsd:element name="GenTime" type="xsd:dateTime" minOccurs="1" maxOccurs="1"/>
+            <xsd:element name="Expiration" type="xsd:duration"/>
+        </xsd:sequence>
+    </xsd:complexType>
+
+    <xsd:simpleType name="FaultTypeType">
+        <xsd:restriction base="xsd:string">
+            <xsd:enumeration value="1"/>
+            <xsd:enumeration value="2"/>
+        </xsd:restriction>
+    </xsd:simpleType>
+
+    <!-- description of fault -->
+    <xsd:complexType name="FaultType">
+        <xsd:sequence>
+            <xsd:element name="Name"        type="xsd:string"/>
+            <xsd:element name="Type"        type="isis:FaultTypeType"/>
+            <xsd:element name="Description" type="xsd:string"/>
+        </xsd:sequence>
+    </xsd:complexType>
+
+    <!-- List of the service types will be provided by the GLUE-2.0 -->
+    <xsd:simpleType name="ServiceTypeType">
+        <xsd:restriction base="xsd:string">
+        </xsd:restriction>
+    </xsd:simpleType>
+
+    <xsd:complexType name="NameValuePairType">
+        <xsd:sequence>
+            <xsd:element name="Name" type="xsd:string"/>
+            <xsd:element name="Value" type="xsd:string"/>
+        </xsd:sequence>
+    </xsd:complexType>
+
+</xsd:schema>
+
+<?xml version="1.0" encoding="utf-8"?>
+<wsdl:definitions name="InformationIndexing"
+  xmlns:SOAP-ENV="http://schemas.xmlsoap.org/soap/envelope/"
+  xmlns:SOAP-ENC="http://schemas.xmlsoap.org/soap/encoding/"
+  xmlns:wsdl="http://schemas.xmlsoap.org/wsdl/"
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns:soap="http://schemas.xmlsoap.org/wsdl/soap/"
+  xmlns:wsa="http://www.w3.org/2005/08/addressing"
+  xmlns:isis="http://www.nordugrid.org/schemas/isis/2007/06"
+  xmlns:wsrf-bf="http://docs.oasis-open.org/wsrf/bf-2"
+  xmlns:wsrf-rp="http://docs.oasis-open.org/wsrf/rp-2"
+  xmlns:wsrf-rpw="http://docs.oasis-open.org/wsrf/rpw-2"
+  xmlns:wsrf-rw="http://docs.oasis-open.org/wsrf/rw-2"
+  targetNamespace="http://www.nordugrid.org/schemas/isis/2007/06"
+>
+<!--
+Interface of Information Indexing Service
+-->
+
+    <wsdl:import namespace="http://docs.oasis-open.org/wsrf/rpw-2" location="http://docs.oasis-open.org/wsrf/rpw-2.wsdl"/>
+
+    <!-- ======= Type definitions ====== -->
+    <wsdl:types>
+        <xsd:schema
+            xmlns:isis="http://www.nordugrid.org/schemas/isis/2007/06"
+            targetNamespace="http://www.nordugrid.org/schemas/isis/2007/06"
+            attributeFormDefault="unqualified"
+            elementFormDefault="qualified"
+            >
+            <xsd:include schemaLocation="./isis.xsd"/>
+        </xsd:schema>
+    </wsdl:types>
+
+<!-- ====== Messages definitions ====== -->
+
+<!-- ====== Register ====== -->
+<wsdl:message name="RegisterRequest">
+    <wsdl:part name="Register"  element="isis:Register"/>
+</wsdl:message>
+
+<wsdl:message name="RegisterResponse">
+    <wsdl:part name="RegisterResponse" element="isis:RegisterResponse"/>
+</wsdl:message>
+
+<!-- ====== RemoveRegistrations ====== -->
+
+<wsdl:message name="RemoveRegistrationsRequest">
+    <wsdl:part name="RemoveRegistrations" element="isis:RemoveRegistrations"/>
+</wsdl:message>
+
+<wsdl:message name="RemoveRegistrationsResponse">
+    <wsdl:part name="RemoveRegistrationsResponse" element="isis:RemoveRegistrationsResponse"/>
+</wsdl:message>
+
+<!-- ====== GetISISList ====== -->
+<wsdl:message name="GetISISListRequest">
+    <wsdl:part name="GetISISList" element="isis:GetISISList"/>
+</wsdl:message>
+
+<wsdl:message name="GetISISListResponse">
+    <wsdl:part name="GetISISListResponse" element="isis:GetISISListResponse"/>
+</wsdl:message>
+
+<!-- ====== Query ====== -->
+<wsdl:message name="QueryRequest">
+    <wsdl:part name="Query" element="isis:Query"/>
+</wsdl:message>
+
+<wsdl:message name="QueryResponse">
+    <wsdl:part name="QueryResponse" element="isis:QueryResponse"/>
+</wsdl:message>
+
+<!-- ====== PortType definitions ====== -->
+<wsdl:portType name="ISISPortType">
+    <wsdl:operation name="Register">
+        <wsdl:input name="RegisterRequest" message="isis:RegisterRequest"/>
+        <wsdl:output name="RegisterResponse" message="isis:RegisterResponse"/>
+    </wsdl:operation>
+    <wsdl:operation name="RemoveRegistrations">
+        <wsdl:input name="RemoveRegistrationsRequest" message="isis:RemoveRegistrationsRequest"/>
+        <wsdl:output name="RemoveRegistrationsResponse" message="isis:RemoveRegistrationsResponse"/>
+    </wsdl:operation>
+    <wsdl:operation name="GetISISList">
+        <wsdl:input name="GetISISListRequest" message="isis:GetISISListRequest"/>
+        <wsdl:output name="GetISISListResponse" message="isis:GetISISListResponse"/>
+    </wsdl:operation>
+    <wsdl:operation name="Query">
+        <wsdl:input name="QueryRequest" message="isis:QueryRequest"/>
+        <wsdl:output name="QueryResponse" message="isis:QueryResponse"/>
+    </wsdl:operation>
+</wsdl:portType>
+
+<!-- ====== Bindings ====== -->
+
+<wsdl:binding name="isis" type="isis:ISISPortType">
+    <soap:binding style="document" transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="Register">
+        <soap:operation soapAction="Register"/>
+        <wsdl:input name="RegisterRequest">
+            <soap:body use="literal"/>
+        </wsdl:input>
+        <wsdl:output name="RegisterResponse">
+            <soap:body use="literal"/>
+        </wsdl:output>
+    </wsdl:operation>
+    <wsdl:operation name="RemoveRegistrations">
+        <soap:operation soapAction="RemoveRegistrations"/>
+        <wsdl:input name="RemoveRegistrationsRequest">
+            <soap:body use="literal"/>
+        </wsdl:input>
+        <wsdl:output name="RemoveRegistrationsResponse">
+            <soap:body use="literal"/>
+        </wsdl:output>
+    </wsdl:operation>
+    <wsdl:operation name="GetISISList">
+        <soap:operation soapAction="GetISISList"/>
+        <wsdl:input name="GetISISListRequest">
+            <soap:body use="literal"/>
+        </wsdl:input>
+        <wsdl:output name="GetISISListResponse">
+            <soap:body use="literal"/>
+        </wsdl:output>
+    </wsdl:operation>
+    <wsdl:operation name="Query">
+        <soap:operation soapAction="Query"/>
+        <wsdl:input name="QueryRequest">
+            <soap:body use="literal"/>
+        </wsdl:input>
+        <wsdl:output name="QueryResponse">
+            <soap:body use="literal"/>
+        </wsdl:output>
+    </wsdl:operation>
+</wsdl:binding>
+<wsdl:binding name="GetResourcePropertyDocument" type="wsrf-rpw:GetResourcePropertyDocument">
+    <soap:binding style="document" transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="GetResourcePropertyDocument">
+       <soap:operation soapAction="GetResourcePropertyDocument"/>
+       <wsdl:input name="GetResourcePropertyDocumentRequest">
+         <soap:body use="literal"/>
+        </wsdl:input>
+        <wsdl:output name="GetResourcePropertyDocumentResponse">
+            <soap:body use="literal"/>
+        </wsdl:output>
+        <wsdl:fault name="ResourceUnknownFault">
+            <soap:fault name="ResourceUnknownFault" use="literal"/>
+        </wsdl:fault>
+        <wsdl:fault name="ResourceUnavailableFault">
+            <soap:fault name="ResourceUnavailableFault" use="literal"/>
+        </wsdl:fault>
+    </wsdl:operation>
+</wsdl:binding>
+
+<wsdl:binding name="GetResourceProperty" type="wsrf-rpw:GetResourceProperty">
+    <soap:binding style="document" transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="GetResourceProperty">
+        <soap:operation soapAction="GetResourceProperty"/>
+        <wsdl:input name="GetResourcePropertyRequest">
+            <soap:body use="literal"/>
+        </wsdl:input>
+        <wsdl:output name="GetResourcePropertyResponse">
+            <soap:body use="literal"/>
+        </wsdl:output>
+        <wsdl:fault name="ResourceUnknownFault">
+            <soap:fault name="ResourceUnknownFault" use="literal"/>
+        </wsdl:fault>
+        <wsdl:fault name="ResourceUnavailableFault">
+            <soap:fault name="ResourceUnavailableFault" use="literal"/>
+        </wsdl:fault>
+        <wsdl:fault name="InvalidResourcePropertyQNameFault">
+            <soap:fault name="InvalidResourcePropertyQNameFault" use="literal"/>
+        </wsdl:fault>
+    </wsdl:operation>
+</wsdl:binding>
+
+<wsdl:binding name="QueryResourceProperties" type="wsrf-rpw:QueryResourceProperties">
+    <soap:binding style="document" transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="QueryResourceProperties">
+        <soap:operation soapAction="QueryResourceProperties"/>
+        <wsdl:input name="QueryResourcePropertiesRequest">
+            <soap:body use="literal"/>
+        </wsdl:input>
+        <wsdl:output name="QueryResourcePropertiesResponse">
+            <soap:body use="literal"/>
+        </wsdl:output>
+        <wsdl:fault name="ResourceUnknownFault">
+            <soap:fault name="ResourceUnknownFault" use="literal"/>
+        </wsdl:fault>
+        <wsdl:fault name="ResourceUnavailableFault">
+            <soap:fault name="ResourceUnavailableFault" use="literal"/>
+        </wsdl:fault>
+        <wsdl:fault name="InvalidResourcePropertyQNameFault">
+            <soap:fault name="InvalidResourcePropertyQNameFault" use="literal"/>
+        </wsdl:fault>
+        <wsdl:fault name="UnknownQueryExpressionDialectFault">
+            <soap:fault name="UnknownQueryExpressionDialectFault" use="literal"/>
+        </wsdl:fault>
+        <wsdl:fault name="InvalidQueryExpressionFault">
+            <soap:fault name="InvalidQueryExpressionFault" use="literal"/>
+        </wsdl:fault>
+        <wsdl:fault name="QueryEvaluationErrorFault">
+            <soap:fault name="QueryEvaluationErrorFault" use="literal"/>
+        </wsdl:fault>
+    </wsdl:operation>
+</wsdl:binding>
+
+<!-- ====== Service definition ====== -->
+<wsdl:service name="isis">
+    <wsdl:port name="GetResourcePropertyDocument" binding="isis:GetResourcePropertyDocument">
+        <soap:address location="mailto:test@test.com"/>
+    </wsdl:port>
+    <wsdl:port name="GetResourceProperty" binding="isis:GetResourceProperty">
+        <soap:address location="mailto:test@test.com"/>
+    </wsdl:port>
+    <wsdl:port name="QueryResourceProperties" binding="isis:QueryResourceProperties">
+        <soap:address location="mailto:test@test.com"/>
+    </wsdl:port>
+    <wsdl:port name="isis" binding="isis:isis">
+        <soap:address location="mailto:test@test.com"/>
+    </wsdl:port>
+</wsdl:service>
+
+</wsdl:definitions>
+\end{verbatim}
+
+% section isis_wsdl (end)
+
+\section{Schema of ISIS Configuration}
+\label{annex:isis_configuration_schema}
+\begin{verbatim}
+
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2009/isis"
+  xmlns:icfg="http://www.nordugrid.org/schemas/ArcConfig/2009/isis"
+  targetNamespace="http://www.nordugrid.org/schemas/ArcConfig/2009/isis"
+  elementFormDefault="qualified" attributeFormDefault="unqualified"
+>
+
+  <xsd:element name="endpoint" type="xsd:anyURI"/>
+
+  <xsd:element name="retry" type="xsd:nonNegativeInteger" minOccurs="0"/>
+
+  <xsd:element name="sparsity" type="xsd:nonNegativeInteger" minOccurs="0"/>
+
+  <xsd:simpleType name="Logger_Type">
+    <xsd:restriction base="xsd:string"/>
+    <xsd:attribute name="level" type="xsd:string" use="optional"/>
+  </xsd:simpleType>
+  <xsd:element name="Logger" type="icfg:Logger_Type" minOccurs="0"/>
+
+  <xsd:element name="ETValid" type="xsd:nonNegativeInteger" minOccurs="0"/>
+
+  <xsd:element name="ETRemove" type="xsd:nonNegativeInteger" minOccurs="0"/>
+
+  <xsd:element name="DBPath" type="xsd:string"/>
+
+  <xsd:complexType name="InfoProvider_Type">
+    <xsd:sequence>
+      <xsd:element name="URL" type="xsd:anyURI"/>
+    </xsd:sequence>
+  </xsd:complexType>
+  <xsd:element name="InfoProvider" type="icfg:InfoProvider_Type" minOccurs="0" maxOccurs="unbounded"/>
+
+</xsd:schema>
+
+\end{verbatim}
+
+% section isis_configuration_schema (end)
+
+\section{Schema of Service Registartion Configuration}
+\label{annex:service_configuration_schema}
+\begin{verbatim}
+
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns:iregc="http://www.nordugrid.org/schemas/InfoRegisterConfig/2008"
+  targetNamespace="http://www.nordugrid.org/schemas/InfoRegisterConfig/2008"
+  elementFormDefault="qualified">
+
+    <xsd:complexType name="InfoRegistrar_Type">
+        <!-- This element defines configuration of Information Registration
+             active element. This element is located at top level configuration
+             document. Each instance if such element instatiates one instance
+             of Information Registrant active element. -->
+        <xsd:sequence>
+            <!-- URL specifies contact endpoint of a bootstrap Information
+                 Registration service. Further ISIS adresses will be queried
+                 from this service. -->
+            <xsd:element name="URL" type="xsd:string"/> <!-- maxOccurs="unbounded" -->
+            <!-- Optional KeyPath, CertificatePath, ProxyPath and CACertificatesDir are
+                 paths to files storing X509 credentials used for establishing connections. -->
+            <xsd:element name="KeyPath" type="xsd:string" minOccurs="0"/>
+            <xsd:element name="CertificatePath" type="xsd:string" minOccurs="0"/>
+            <xsd:element name="ProxyPath" type="xsd:string" minOccurs="0"/>
+            <xsd:element name="CACertificatesDir" type="xsd:string" minOccurs="0"/>
+            <!-- Retry count. Specifie how many times have an ISIS to timeout to handle it as
+                 unavailable. If there is no value specified the default value: 5 will be used. -->
+            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1"/>
+        </xsd:sequence>
+        <!-- Attribute id may be used to refer to this element -->
+        <xsd:attribute name="id" type="xsd:string" use="optional"/>
+    </xsd:complexType>
+    <xsd:element name="InfoRegistrar" type="iregc:InfoRegistrar_Type"/>
+
+    <xsd:complexType name="Registrar_Type">
+        <xsd:sequence>
+            <!-- Missing elements means that the default values will be used. -->
+
+            <!-- Period specifies how often registration to be done in minutes.
+                 The 0 value means single registration. -->
+            <xsd:element name="Period" type="xsd:nonNegativeInteger" minOccurs="0" maxOccurs="1"/>
+            <!-- Element defines the unique id of the service propagated outside. -->
+            <xsd:element name="ServiceID" type="xsd:string" minOccurs="0" maxOccurs="1"/>
+            <!-- This element defines URL of the service as seen from outside. -->
+            <xsd:element name="endpoint" type="xsd:string" minOccurs="0" maxOccurs="1"/>
+            <!-- This element defines the expiration time of the information provided by the service. -->
+            <xsd:element name="expiration" type="xsd:string" minOccurs="0" maxOccurs="1"/>
+        </xsd:sequence>
+        <!-- Attribute id is to define the previously configured InfoRegistrar to be refered -->
+        <xsd:attribute name="id" type="xsd:string" use="required"/>
+    </xsd:complexType>
+
+    <xsd:complexType name="InfoRegister_Type">
+    <!-- Element for Service element to link it to InfoRegistrar
+         elements. It may also override some configuration parameters.
+         Presence of this element means that service will be registered
+         to ISISes. -->
+        <xsd:sequence>
+            <!-- This elements specify which registrars must be used
+                 for registering services. If there is no such element
+                 then registration is done using all registrars. -->
+            <xsd:element name="Registrar" type="iregc:Registrar_Type" minOccurs="0" maxOccurs="unbounded"/>
+            <!-- Default values for every InfoRegistrar -->
+
+            <!-- Period specifies how often registration to be done in minutes.
+                 The 0 value means single registration. -->
+            <xsd:element name="Period" type="xsd:nonNegativeInteger" minOccurs="1" maxOccurs="1"/>
+            <!-- Element defines the unique id of the service propagated outside. -->
+            <xsd:element name="ServiceID" type="xsd:string" minOccurs="0" maxOccurs="1"/>
+            <!-- This element defines URL of the service as seen from outside. -->
+            <xsd:element name="endpoint" type="xsd:string" minOccurs="1" maxOccurs="1"/>
+            <!-- This element defines the expiration time of the information provided by the service. -->
+            <xsd:element name="expiration" type="xsd:string" minOccurs="1" maxOccurs="1"/>
+        </xsd:sequence>
+    </xsd:complexType>
+    <xsd:element name="InfoRegister" type="iregc:InfoRegister_Type"/>
+
+    <!-- Configuration element force skipping the Self-Registration -->
+    <xsd:element name="NoRegister">
+        <xsd:complexType />
+    </xsd:element>
+
+</xsd:schema>
+
+\end{verbatim}
+
+% section service_configuration_schema (end)
+
+\section{Schema of Information Document Policies}
+\label{annex:info_policies_schema}
+\begin{verbatim}
+
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema
+  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+  xmlns:if="http://www.nordugrid.org/schemas/InfoFilter/2008"
+  xmlns="http://www.nordugrid.org/schemas/InfoFilter/2008"
+  xmlns:policy="http://www.nordugrid.org/schemas/policy-arc"
+  targetNamespace="http://www.nordugrid.org/schemas/InfoFilter/2008"
+  elementFormDefault="qualified">
+
+    <xsd:complexType name="InfoFilterDefinition_Type">
+        <!-- This element is a container for single policy element of any kind.
+             Attached mandatory attribute defines identifier which is then used
+             to link XML elements with defined policy.
+             Currently only supported policy is one defined in
+             http://www.nordugrid.org/schemas/policy-arc namespace. -->
+        <xsd:any namespace="##other" processContents="lax" minOccurs="1" maxOccurs="1"/>
+        <xsd:attribute name="id" type="xsd:string" use="required"/>
+    </xsd:complexType>
+    <xsd:element name="InfoFilterDefinition" type="if:InfoFilterDefinition_Type" minOccurs="0" maxOccurs="unbounded"/>
+
+    <xsd:attribute name="InfoFilterTag">
+        <!-- This attribute is to be used in element for which authorization
+             policy to be applied. It specifies identifier of policy defined
+             in InfoFilterDefinition element. -->
+        <xsd:simpleType>
+            <xsd:restriction base="xsd:string"/>
+        </xsd:simpleType>
+    </xsd:attribute>
+
+</xsd:schema>
+
+\end{verbatim}
+
+% section info_policies_schema (end)
+
+% chapter appendices (end)
+
+\bibliography{grid}
 \end{document}

### 2009-05-18 doc/tutorials/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
These are also not executables

--- 2009-03-17 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
@@
 	]
-{../src/services/secechoservice/secechoservice.cpp}
+{../examples/src/services/secechoservice/secechoservice.cpp}
 %
@@
 	]
-{../src/services/secechoservice/arched_sec_echoservice.xml}
+{../examples/src/services/secechoservice/arched_sec_echoservice.xml}
 
@@
 	]
-{../src/services/secechoservice/policy_example.xml}
+{../examples/src/services/secechoservice/policy_example.xml}
 

### 2009-05-19 doc/manuals/janitor/images/dependencies.tex
Temporary solution to save the nightlies - use old pgf style

@@
-\begin{tikzpicture}[>=latex,line join=bevel,]
+\begin{tikzpicture}[>=latex,join=bevel,]
   \pgfsetlinewidth{1bp}
@@
   % Edge: Installer -> Util
-  \draw [->] (711bp,144bp) .. controls (710bp,136bp) and (710bp,127bp)  .. (710bp,108bp);
+  \draw [->] (649bp,144bp) .. controls (648bp,136bp) and (648bp,127bp)  .. (647bp,108bp);
   % Edge: Janitor -> Filefetcher
-  \draw [->] (732bp,221bp) .. controls (750bp,211bp) and (776bp,194bp)  .. (805bp,175bp);
+  \draw [->] (670bp,221bp) .. controls (687bp,210bp) and (714bp,193bp)  .. (742bp,175bp);
   % Edge: Catalog -> Tag
-  \draw [->] (390bp,153bp) .. controls (417bp,143bp) and (463bp,125bp)  .. (501bp,108bp) .. controls (503bp,107bp) and (504bp,107bp)  .. (515bp,101bp);
+  \draw [->] (352bp,152bp) .. controls (376bp,141bp) and (416bp,124bp)  .. (449bp,108bp) .. controls (451bp,107bp) and (452bp,107bp)  .. (463bp,101bp);
   % Edge: BaseSystem -> Note
-  \draw [->] (458bp,72bp) .. controls (458bp,64bp) and (458bp,55bp)  .. (458bp,36bp);
+  \draw [->] (410bp,72bp) .. controls (410bp,64bp) and (410bp,55bp)  .. (410bp,36bp);
   % Edge: Installer -> Execute
-  \draw [->] (727bp,147bp) .. controls (737bp,137bp) and (749bp,124bp)  .. (766bp,105bp);
+  \draw [->] (665bp,146bp) .. controls (674bp,136bp) and (686bp,123bp)  .. (703bp,105bp);
   % Edge: Janitor -> Catalog
-  \draw [->] (686bp,229bp) .. controls (625bp,216bp) and (472bp,184bp)  .. (392bp,167bp);
+  \draw [->] (625bp,228bp) .. controls (568bp,216bp) and (430bp,184bp)  .. (355bp,168bp);
   % Edge: Catalog -> MetaPackage
-  \draw [->] (341bp,155bp) .. controls (315bp,147bp) and (274bp,131bp)  .. (245bp,108bp) .. controls (224bp,90bp) and (207bp,65bp)  .. (192bp,36bp);
+  \draw [->] (304bp,155bp) .. controls (279bp,146bp) and (241bp,131bp)  .. (215bp,108bp) .. controls (195bp,90bp) and (180bp,65bp)  .. (166bp,36bp);
   % Edge: Janitor -> TarPackage
-  \draw [->] (685bp,231bp) ..
-controls (622bp,225bp) and (462bp,207bp)  ..
-(330bp,180bp) ..
-controls (214bp,155bp) and (185bp,147bp)  ..
-(73bp,108bp) ..
-controls (71bp,107bp) and (69bp,106bp)  ..
-(57bp,101bp);
+  \draw [->] (624bp,231bp) ..
+controls (564bp,225bp) and (415bp,208bp)  ..
+(293bp,180bp) ..
+controls (191bp,156bp) and (166bp,146bp)  ..
+(67bp,108bp) ..
+controls (65bp,107bp) and (63bp,107bp)  ..
+(52bp,101bp);
   % Edge: Filefetcher -> Execute
-  \draw [->] (815bp,145bp) .. controls (810bp,137bp) and (803bp,126bp)  .. (792bp,107bp);
+  \draw [->] (751bp,145bp) .. controls (746bp,137bp) and (739bp,126bp)  .. (728bp,107bp);
   % Edge: Janitor -> Config
-  \draw [->] (739bp,232bp) .. controls (793bp,226bp) and (913bp,211bp)  .. (938bp,180bp) .. controls (953bp,160bp) and (938bp,132bp)  .. (917bp,105bp);
+  \draw [->] (678bp,231bp) .. controls (731bp,226bp) and (845bp,210bp)  .. (869bp,180bp) .. controls (884bp,160bp) and (869bp,132bp)  .. (848bp,105bp);
   % Edge: Janitor -> Execute
-  \draw [->] (724bp,217bp) .. controls (731bp,207bp) and (741bp,193bp)  .. (748bp,180bp) .. controls (758bp,160bp) and (767bp,135bp)  .. (776bp,108bp);
+  \draw [->] (663bp,218bp) .. controls (671bp,207bp) and (680bp,193bp)  .. (687bp,180bp) .. controls (696bp,160bp) and (704bp,136bp)  .. (712bp,108bp);
   % Edge: Catalog -> TarPackage
-  \draw [->] (339bp,159bp) .. controls (287bp,153bp) and (169bp,137bp)  .. (73bp,108bp) .. controls (71bp,107bp) and (68bp,106bp)  .. (56bp,102bp);
+  \draw [->] (302bp,159bp) .. controls (255bp,152bp) and (151bp,135bp)  .. (67bp,108bp) .. controls (65bp,107bp) and (63bp,106bp)  .. (51bp,102bp);
   % Edge: Catalog -> DebianPackage
-  \draw [->] (392bp,156bp) .. controls (431bp,147bp) and (509bp,128bp)  .. (573bp,108bp) .. controls (576bp,107bp) and (579bp,106bp)  .. (592bp,102bp);
+  \draw [->] (355bp,156bp) .. controls (392bp,146bp) and (463bp,128bp)  .. (521bp,108bp) .. controls (523bp,107bp) and (526bp,106bp)  .. (538bp,102bp);
   % Edge: Installer -> Config
-  \draw [->] (736bp,153bp) .. controls (770bp,140bp) and (831bp,117bp)  .. (878bp,99bp);
+  \draw [->] (674bp,153bp) .. controls (707bp,140bp) and (764bp,117bp)  .. (810bp,99bp);
   % Edge: Tag -> Note
-  \draw [->] (521bp,75bp) .. controls (509bp,65bp) and (494bp,51bp)  .. (474bp,33bp);
+  \draw [->] (469bp,75bp) .. controls (458bp,65bp) and (444bp,51bp)  .. (425bp,33bp);
   % Edge: MetaPackageCollection -> MetaPackage
-  \draw [->] (151bp,72bp) .. controls (155bp,64bp) and (162bp,54bp)  .. (172bp,36bp);
+  \draw [->] (131bp,72bp) .. controls (135bp,64bp) and (140bp,54bp)  .. (150bp,36bp);
   % Edge: Common -> Config
-  \draw [->] (902bp,144bp) .. controls (902bp,136bp) and (902bp,127bp)  .. (902bp,108bp);
+  \draw [->] (833bp,144bp) .. controls (833bp,136bp) and (833bp,127bp)  .. (833bp,108bp);
   % Edge: Janitor -> DebianPackage
-  \draw [->] (697bp,219bp) .. controls (688bp,208bp) and (676bp,194bp)  .. (667bp,180bp) .. controls (654bp,160bp) and (642bp,135bp)  .. (630bp,108bp);
+  \draw [->] (636bp,219bp) .. controls (627bp,208bp) and (615bp,194bp)  .. (606bp,180bp) .. controls (593bp,160bp) and (582bp,135bp)  .. (571bp,108bp);
   % Edge: Catalog -> Package
-  \draw [->] (349bp,148bp) .. controls (336bp,137bp) and (319bp,123bp)  .. (298bp,104bp);
+  \draw [->] (313bp,147bp) .. controls (302bp,137bp) and (287bp,123bp)  .. (267bp,105bp);
   % Edge: Catalog -> NoteCollection
-  \draw [->] (366bp,144bp) .. controls (366bp,136bp) and (366bp,127bp)  .. (366bp,108bp);
+  \draw [->] (329bp,144bp) .. controls (329bp,136bp) and (329bp,127bp)  .. (329bp,108bp);
   % Edge: Janitor -> Installer
-  \draw [->] (712bp,216bp) .. controls (712bp,208bp) and (712bp,199bp)  .. (712bp,180bp);
+  \draw [->] (651bp,216bp) .. controls (651bp,208bp) and (651bp,199bp)  .. (651bp,180bp);
   % Edge: Janitor -> Util
-  \draw [->] (698bp,219bp) ..
-controls (690bp,208bp) and (680bp,194bp)  ..
-(676bp,180bp) ..
-controls (671bp,164bp) and (671bp,159bp)  ..
-(676bp,144bp) ..
-controls (679bp,134bp) and (684bp,123bp)  ..
-(695bp,106bp);
+  \draw [->] (637bp,219bp) ..
+controls (629bp,208bp) and (619bp,194bp)  ..
+(615bp,180bp) ..
+controls (610bp,164bp) and (610bp,159bp)  ..
+(615bp,144bp) ..
+controls (617bp,134bp) and (622bp,124bp)  ..
+(633bp,106bp);
   % Edge: Execute -> InstallLogger
-  \draw [->] (781bp,72bp) .. controls (781bp,64bp) and (781bp,55bp)  .. (781bp,36bp);
+  \draw [->] (717bp,72bp) .. controls (717bp,64bp) and (717bp,55bp)  .. (717bp,36bp);
   % Edge: Janitor -> Common
-  \draw [->] (737bp,226bp) .. controls (767bp,217bp) and (821bp,199bp)  .. (866bp,180bp) .. controls (868bp,179bp) and (870bp,178bp)  .. (881bp,173bp);
+  \draw [->] (675bp,226bp) .. controls (704bp,216bp) and (755bp,198bp)  .. (797bp,180bp) .. controls (799bp,179bp) and (800bp,179bp)  .. (811bp,173bp);
   % Edge: NoteCollection -> Note
-  \draw [->] (386bp,74bp) .. controls (399bp,63bp) and (418bp,49bp)  .. (440bp,32bp);
+  \draw [->] (347bp,74bp) .. controls (358bp,64bp) and (373bp,51bp)  .. (393bp,33bp);
   % Edge: Catalog -> BaseSystem
-  \draw [->] (384bp,148bp) .. controls (397bp,138bp) and (415bp,123bp)  .. (439bp,105bp);
+  \draw [->] (345bp,147bp) .. controls (357bp,137bp) and (372bp,123bp)  .. (393bp,105bp);
   % Edge: Catalog -> MetaPackageCollection
-  \draw [->] (341bp,154bp) .. controls (306bp,142bp) and (238bp,121bp)  .. (181bp,103bp);
+  \draw [->] (305bp,154bp) .. controls (271bp,142bp) and (208bp,120bp)  .. (156bp,102bp);
   % Node: Note
@@
   \pgfsetstrokecolor{black}
-  \draw (458bp,18bp) ellipse (27bp and 18bp);
-  \draw (458bp,18bp) node {Note};
+  \draw (410bp,18bp) ellipse (27bp and 18bp);
+  \draw (410bp,18bp) node {Note};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (781bp,90bp) ellipse (27bp and 18bp);
-  \draw (781bp,90bp) node {Execute};
+  \draw (717bp,90bp) ellipse (27bp and 18bp);
+  \draw (717bp,90bp) node {Execute};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (281bp,90bp) ellipse (27bp and 18bp);
-  \draw (281bp,90bp) node {Package};
+  \draw (251bp,90bp) ellipse (27bp and 18bp);
+  \draw (251bp,90bp) node {Package};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (623bp,90bp) ellipse (41bp and 18bp);
-  \draw (623bp,90bp) node {DebianPackage};
+  \draw (565bp,90bp) ellipse (35bp and 18bp);
+  \draw (565bp,90bp) node {DebianPackage};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (902bp,90bp) ellipse (27bp and 18bp);
-  \draw (902bp,90bp) node {Config};
+  \draw (833bp,90bp) ellipse (27bp and 18bp);
+  \draw (833bp,90bp) node {Config};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (183bp,18bp) ellipse (37bp and 18bp);
-  \draw (183bp,18bp) node {MetaPackage};
+  \draw (159bp,18bp) ellipse (32bp and 18bp);
+  \draw (159bp,18bp) node {MetaPackage};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (366bp,90bp) ellipse (40bp and 18bp);
-  \draw (366bp,90bp) node {NoteCollection};
+  \draw (329bp,90bp) ellipse (33bp and 18bp);
+  \draw (329bp,90bp) node {NoteCollection};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (709bp,90bp) ellipse (27bp and 18bp);
-  \draw (709bp,90bp) node {Util};
+  \draw (645bp,90bp) ellipse (27bp and 18bp);
+  \draw (645bp,90bp) node {Util};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (366bp,162bp) ellipse (27bp and 18bp);
-  \draw (366bp,162bp) node {Catalog};
+  \draw (329bp,162bp) ellipse (27bp and 18bp);
+  \draw (329bp,162bp) node {Catalog};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (537bp,90bp) ellipse (27bp and 18bp);
-  \draw (537bp,90bp) node {Tag};
+  \draw (485bp,90bp) ellipse (27bp and 18bp);
+  \draw (485bp,90bp) node {Tag};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (826bp,162bp) ellipse (31bp and 18bp);
-  \draw (826bp,162bp) node {Filefetcher};
+  \draw (761bp,162bp) ellipse (27bp and 18bp);
+  \draw (761bp,162bp) node {Filefetcher};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (902bp,162bp) ellipse (27bp and 18bp);
-  \draw (902bp,162bp) node {Common};
+  \draw (833bp,162bp) ellipse (27bp and 18bp);
+  \draw (833bp,162bp) node {Common};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (140bp,90bp) ellipse (58bp and 18bp);
-  \draw (140bp,90bp) node {MetaPackageCollection};
+  \draw (122bp,90bp) ellipse (46bp and 18bp);
+  \draw (122bp,90bp) node {MetaPackageCollection};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (712bp,162bp) ellipse (27bp and 18bp);
-  \draw (712bp,162bp) node {Installer};
+  \draw (651bp,162bp) ellipse (27bp and 18bp);
+  \draw (651bp,162bp) node {Installer};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (712bp,234bp) ellipse (27bp and 18bp);
-  \draw (712bp,234bp) node {Janitor};
+  \draw (651bp,234bp) ellipse (27bp and 18bp);
+  \draw (651bp,234bp) node {Janitor};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (32bp,90bp) ellipse (32bp and 18bp);
-  \draw (32bp,90bp) node {TarPackage};
+  \draw (29bp,90bp) ellipse (29bp and 18bp);
+  \draw (29bp,90bp) node {TarPackage};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (458bp,90bp) ellipse (34bp and 18bp);
-  \draw (458bp,90bp) node {BaseSystem};
+  \draw (410bp,90bp) ellipse (30bp and 18bp);
+  \draw (410bp,90bp) node {BaseSystem};
 \end{scope}
@@
   \pgfsetstrokecolor{black}
-  \draw (781bp,18bp) ellipse (35bp and 18bp);
-  \draw (781bp,18bp) node {InstallLogger};
+  \draw (717bp,18bp) ellipse (33bp and 18bp);
+  \draw (717bp,18bp) node {InstallLogger};
 \end{scope}

### 2009-05-20 doc/tech_doc/infosys/infosys_technical.tex
Few words about design/architecture of infosys

@@
 
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{InfosysArchitecture.png}}}
+\caption{\label{fig:infosys_architecture}Architecture of Information System} }
+\end{figure}
+
+Information System of ARC middleware is composed of 3 main parts.
+The Information System Indexing Services (ISIS) for a set of information containers.
+Every generic ARC Service pushes information about itself into nearby ISIS container/service hence registering itself to Information System.
+The information stored in every ISIS is propagated among all ISIS services.
+Clients can query any nearby ISIS service for registered information in order to perform discovery of services which posses specific properties.
+
 \chapter{ISIS}
@@
  In case of multiple ISIS services they form a peer-to-peer network.
-Inside that cloud Registration Records are propagated between services in such a way that all ISIS instances continiously try to synchronise their databases.
+Inside that cloud Registration Records are propagated between services in such a way that all ISIS instances continuously try to synchronize their databases.
 
@@
   \item[Output]~\begin{description}
-    \item[Fault] Optional element describing fault which occured while performing registration. If missing registration succeeded.
+    \item[Fault] Optional element describing fault which occurred while performing registration. If missing registration succeeded.
   \end{description}
@@
 Among those the \texttt{Endpoint Reference} is used to contact service - only required element is contact URL of service.
-\texttt{Type} specifies kind of service and is used to find out functionality nad interface of service.
+\texttt{Type} specifies kind of service and is used to find out functionality and interface of service.
 \texttt{ServiceID} is used to distinguish between registered services and to deal with case of service changing it's contact URL.
@@
 
-This operation is used to explicitely requesting the removal of zero or more Registration Entries associated with specified ServiceID values stored in the Information System.
+This operation is used to explicitly requesting the removal of zero or more Registration Entries associated with specified ServiceID values stored in the Information System.
 If corresponding record does not exist it's identifier will be present in response message together with corresponding Fault element.
@@
 
-\subsection{Client Authorzation}
+\subsection{Client Authorization}
 \label{sub:isis_authorization_client_authorization}
@@
 To ensure information stored in ISIS cloud can't be tampered and only available to proper clients following authorization framework is implemented.
-All actions perfored by ISIS clients are divided into three following groups:
+All actions performed by ISIS clients are divided into three following groups:
 
@@
 
- Those operations may cause uncontrollable changes in colleccted information and must be granted only for higly trusted entities like ISISes themselves.
+ Those operations may cause uncontrollable changes in collected information and must be granted only for highly trusted entities like ISISes themselves.
 
@@
 Those 3 action groups are handled using Security Framework of ARC~\cite{arcsec}. For each group corresponding 
-Action is defined for ARC policy language as descriped in table \ref{tab:policy_actions}.
+Action is defined for ARC policy language as described in table \ref{tab:policy_actions}.
 
@@
 
-\subsection{Information Authorzation}
+\subsection{Information Authorization}
 \label{sub:isis_authorization_information_authorization}
@@
 
-To implement functionality described above each node in agregated XML documents of information
+To implement functionality described above each node in aggregated XML documents of information
 collected by ISIS cloud may have Access Control Policy associated with it. Access control is 
@@
 Children nodes can only additionally restrict access control imposed by parent node. For example
-if parent node A allows access only to VO1 then children node B can narrow access to Adminsitrator 
+if parent node A allows access only to VO1 then children node B can narrow access to Administrator 
 of VO1 and can't grant access for VO2 members.
@@
 as XML nodes (see Appendix \ref{annex:info_policies_schema} for schema) even if that violates schema 
-of document. Then nodes are assigned policies aby adding XML attributes refering to defined Policies.
+of document. Then nodes are assigned policies by adding XML attributes referring to defined Policies.
 Access permission to particular information node is evaluated by traversing all nodes prom 
-parent to childen. At first node which gives negative result evaluation is stopped and
+parent to children. At first node which gives negative result evaluation is stopped and
 this node including all its children is removed from document.
@@
 describing service. Normally this information should be more detailed than one sent to ISIS. For this 
-purpose LIDI interface is defined which is a susbset of WS Resource Properties (WSRP)~\cite{wsrf-rp}. Following WSRP
+purpose LIDI interface is defined which is a subset of WS Resource Properties (WSRP)~\cite{wsrf-rp}. Following WSRP
 operations must be supported - \textit{GetResourcePropertDocument}, \textit{GetResourceProperty},
@@
 To register services to more than one ISIS network multiple Registrant instances has to be configured.
-In this case, the default Registrant will be used for registering every services unless configured expilicitely.
+In this case, the default Registrant will be used for registering every services unless configured explicitly.
 The registration of service can be done once or periodically based either on the configuration of the Registrant or overwritten for every service separately.
@@
 Registration parameters per service are defined by InfoRegister element located inside corresponding service configuration element.
-To assign service to specific Registrant(s) ona may use Registrar configuration elements situated inside InfoRegister.
-If none Registrar element is defined service will be registed using first (default) Registrant.
-To make not register itself special configration element NoRegister has to be used.
+To assign service to specific Registrant(s) one may use Registrar configuration elements situated inside InfoRegister.
+If none Registrar element is defined service will be registered using first (default) Registrant.
+To make not register itself special configuration element NoRegister has to be used.
 
@@
 \item{URL}            <!-- URL specifies contact endpoint of a bootstrap Information
-                 Registration service. Further ISIS adresses will be queried
+                 Registration service. Further ISIS addresses will be queried
                  from this service. -->
@@
                  paths to files storing X509 credentials used for establishing connections. -->
-\item{Retry}            <!-- Retry count. Specifie how many times have an ISIS to timeout to handle it as
+\item{Retry}            <!-- Retry count. Specifies how many times have an ISIS to timeout to handle it as
                  unavailable. If there is no value specified the default value: 5 will be used. -->

### 2009-05-21 doc/tech_doc/infosys/infosys_technical.tex
Authorization configuration with examples. Use architecture figure from Ivan. Better name of document.

@@
   pdfauthor = {},
-  pdftitle = {Documentation of the ARC Information System},
+  pdftitle = {Documentation of the Web Service based ARC Information System},
   pdfsubject = {},
@@
 \begin{figure}[ht]
-\centering{{\scalebox{0.6}{\includegraphics{InfosysArchitecture.png}}}
-\caption{\label{fig:infosys_architecture}Architecture of Information System} }
+ \centering{
+  \scalebox{0.6}{\includegraphics{SystemOverview.png}}
+  \caption{\label{fig:infosys_architecture}Architecture of Information System}
+ }
 \end{figure}
@@
 Information System of ARC middleware is composed of 3 main parts.
-The Information System Indexing Services (ISIS) for a set of information containers.
+The Information System Indexing Services (ISIS) form a set of information containers.
 Every generic ARC Service pushes information about itself into nearby ISIS container/service hence registering itself to Information System.
@@
 Clients can query any nearby ISIS service for registered information in order to perform discovery of services which posses specific properties.
+Clients can also query ARC Services to obtain more detailed information about Service properties.
 
@@
 
-\begin{table}{h}
+\begin{table}[h]
 \caption{ARC Policy Actions of ISIS}
@@
 %
+
+% subsection information_authorization (end)
+
+\subsection{Configuration}
+\label{sub:isis_authorization_configuration}
+
+\textit{For information about sophisticated authorization policies and how to deploy various Policy Decision Point entities please see "Security Framework of ARC1" document"~\cite{arcsec}.}
+
+To restrict set of clients allowed to perform operations in ISIS service proper authorization policy is needed.
+Let's assume ISIS is operating over TLS connection and all participants posses X.509 certificates with following subject names:
+\begin{itemize}
+\item /O=Grid/O=Test/CN=CLIENT - generic client entities
+\item /O=Grid/O=Test/CN=SERVICE - generic service entities
+\item /O=Grid/O=Test/CN=ISIS - all ISIS belonging to ISIS cloud
+\end{itemize}
+
+\textit{In NO way we suggest to use such setup. Real installation should use more sophisticated way to 
+identify clients contacting ISIS service. For an infrastructure with quite static roles distribution for 
+example we suggest to use Virtual Organization Management Service (VOMS) attributes embedded into X.509 
+certificates representing participating entities.}
+
+Below is an example policy made of 3 rules defined in lines 4-17, 18-29 and 30-39. Those define allowed 
+behavior for ISIS, generic service and generic client. Lines 6-7, 20-21 and 32-33 specify attributes 
+used to recognise type of connecting client. In this case those are subjects of X.509 certificates with 
+values defined above. Lines 10-15, 24-27 and 36-37 specify allowed actions. One can see that this policy 
+allows all operations to be performed by ISIS client. It limits operations allowed for generic service 
+to "service" and "client" types. And client entities are allowed to perform "client" type operations only.
+
+\begin{verbatim}
+ 1: <?xml version="1.0" encoding="UTF-8"?>
+ 2: <Policy xmlns="http://www.nordugrid.org/schemas/policy-arc" PolicyId="policy1"
+ 3:           CombiningAlg="Deny-Overrides">
+ 4:  <Rule RuleId="isis_to_isis" Effect="Permit">
+ 5:    <Subjects>
+ 6:      <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/identity"
+ 7:              >/O=Grid/O=Test/CN=ISIS</Subject>
+ 8:    </Subjects>
+ 9:    <Actions>
+10:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
+11:             >isis</Action>
+12:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
+13:             >service</Action>
+14:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
+15:             >client</Action>
+16:    </Actions>
+17:  </Rule>
+18:  <Rule RuleId="service_to_isis" Effect="Permit">
+19:    <Subjects>
+20:      <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/identity"
+21:              >/O=Grid/O=Test/CN=SERVICE</Subject>
+22:    </Subjects>
+23:    <Actions>
+24:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
+25:             >service</Action>
+26:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
+27:             >client</Action>
+28:    </Actions>
+29:  </Rule>
+30:  <Rule RuleId="client_to_isis" Effect="Permit">
+31:    <Subjects>
+32:      <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/identity"
+33:              >/O=Grid/O=Test/CN=CLIENT</Subject>
+34:    </Subjects>
+35:    <Actions>
+36:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
+37:             >client</Action>
+38:    </Actions>
+39:  </Rule>
+40: </Policy>
+
+\end{verbatim}
+
+This policy is not as restrictive as it could be in order to allow ISIS register themselves as ordinary 
+services and to allow services which perform discovery of other services to behave like ordinary clients.
+
+In order to activate policy it must be linked to ISIS service through Security Handle and Policy 
+Decision Point emedded into configuration of ISIS. Below is minimal example configuration element of 
+ISIS service. In lines 7-17 a set of Security Handler and Policy Decision Point is configured to handle
+policies written in ARC Policy Language (described in ~\cite{arcsec}). Line 14 specifies that policy is 
+read from file \textit{/opt/arc/etc/isis/policy.xml} every time new request is arrived. If policy is not 
+satisfied then ISIS returns SOAP Fault instead of usual informative response.
+
+\begin{verbatim}
+ 1: <Service 
+ 2:   xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
+ 3:   xmlns:isis="http://www.nordugrid.org/schemas/ArcConfig/2009/isis"
+ 4:   name="isis" id="isis">
+ 5:  <isis:endpoint>https://localhost:60000/isis</isis:endpoint>
+ 6:  <isis:DBPath>/opt/arc/share/isisdb</isis:DBPath>
+ 7:  <SecHandler
+ 8:    xmlsns:auth="http://www.nordugrid.org/schemas/SimpleListAuthZ"
+ 9:    name="arc.authz" id="authz" event="incoming">
+10:   <auth:PDP
+11:     xmlns:pdp="http://www.nordugrid.org/schemas/ArcPDP"
+12:     name="arc.pdp">
+13:    <pdp:PolicyStore>
+14:     <pdp:Location type="file">/opt/arc/etc/isis/policy.xml</pdp:Location>
+15:    </pdp:PolicyStore>
+16:   </auth:PDP>
+17:  </SecHandler>
+18: </Service>
+
+\end{verbatim}
 
@@
 The registration operation is done by the Registrants as described in Section \ref{sec:service_registration}.
-Each Registrant is instantiated by corresponding InfoRegistrar XML element in configuration file as defined in Section \ref{annex:service_configuration_schema}.
-First InfoRegistrar defines Registrant which is used to register every service by default.
-
-Registration parameters per service are defined by InfoRegister element located inside corresponding service configuration element.
-To assign service to specific Registrant(s) one may use Registrar configuration elements situated inside InfoRegister.
-If none Registrar element is defined service will be registered using first (default) Registrant.
-To make not register itself special configuration element NoRegister has to be used.
-
-Following elements are defined inside configuration element if the Registrant:
+Each Registrant is instantiated by corresponding InfoRegistrar XML element in configuration file as 
+defined in Section \ref{annex:service_configuration_schema}. First InfoRegistrar defines Registrant 
+which is used to register every service by default.
+
+Registration parameters per service are defined by InfoRegister element located inside corresponding 
+service configuration element. To assign service to specific Registrant(s) one may use Registrar 
+configuration elements situated inside InfoRegister. If none Registrar element is defined service will 
+be registered using first (default) Registrant. To make not register itself special configuration element 
+NoRegister has to be used.
+
+Following elements are defined inside configuration element of the Registrant:
 \begin{description}
-\item{URL}            <!-- URL specifies contact endpoint of a bootstrap Information
-                 Registration service. Further ISIS addresses will be queried
-                 from this service. -->
-\item{KeyPath, CertificatePath, ProxyPath and CACertificatesDir}
-            <!-- Optional KeyPath, CertificatePath, ProxyPath and CACertificatesDir are
-                 paths to files storing X509 credentials used for establishing connections. -->
-\item{Retry}            <!-- Retry count. Specifies how many times have an ISIS to timeout to handle it as
-                 unavailable. If there is no value specified the default value: 5 will be used. -->
-            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1"/>
-
+\item{URL} specifies contact endpoint of a bootstrap ISIS. If neede further ISIS addresses will be 
+queried from this service. This element is mandatory.
+\item{KeyPath, CertificatePath, ProxyPath and CACertificatesDir} are paths to files storing X509 
+credentials used for establishing connections. Those elements are optional and needed only if 
+TLS communication is used.
+\item{Retry} specifies how many times communication to ISIS have to fail/timeout to start treating is
+as unavailable. The default value is 5.
 \end{description}

### 2009-05-22 doc/tech_doc/infosys/infosys_technical.tex
First version of P2P documentation. Mistakes will be fixed soon.

@@
 \label{sub:peer_to_peer_functionality}
-
+The ISISs are able to organize themselves into a peer-to-peer network by synchronizing the stored replicated database and maintain this synchronicity and network topology.
+Every node of this peer-to-peer cloud (thatÕs an ISIS service) has an own PeerID (an identifier only used in this network) made as a hash of its endpoint URL.
+The members are ordered by this identifier where there is also a sequence defined between them where the follower of the entity is the ISIS with the smallest PeerID of the greater PeerIDs and the smallest in case of the greatest value.
+This metric is used for define the neighbor relation thatÕs the base of the inter-peer-to-peer network communication.
+A strong condition that only registered ISIS can have any role in the cloud.
+
+\subsubsection{Connecting to the peer-to-peer network}
+Every ISIS tries to connect to the network at its startup by accessing the InfoProvider ISISs  known from the configuration.
+The order of attempting is random to achieve as much load-balancing as possible.
+(By design we are counting of a huge amount of ISISs configured in a very similar way, where the InfoProvider ISIS are mainly the same.) If there is no InfoProvider available or nothing is known from the configuration, the node decides to be the fist member and to build a new network.
+
+If there is any InfoProvider available the list of member ISIS will be queried and the ÒfollowerÓ will be defined using the previously mentioned metric.
+This ÒfollowerÓ node will help the ISIS to connect to the cloud.
+The new member will use the ÒConnectÓ Web Service operation to get the newest version of database stored in the peer-to-peer network, and place the newer entries to the cloud by register them.
+An initial database synchronization can be achieved with this process.
+
+We can prevent to overload the InfoProvider ISISs by balancing the operation of database synchronizing thatÕs the most costly procedure in the connection phase.
+
+There is also an other extension used in connection with the InfoProvider ISISs.
+If a node wasn't able to connect any of its InfoProviders then it will redo that later.
+This achieves that the network is able to fuse two disjunct parts of network when the missing InfoProvider ISIS is appearing.
+
+\subsubsection{Routing in the peer-to-peer network}
+We are keeping the databases in synch by passing every Registration and RemoveRegistrations  to every node acting in the network.
+This set the redundancy and the fault tolerant message routing very important between any two ISISs.
+
+The most important parameter of the network is the {\it sparsity}.
+This an integer not less then 2 that determines the number of neighbors in the function of the actual number of member nodes of the network, where the number of neighbors can be growing when new members connecting to the network or members leave.
+A spares graph can be shaped by choosing a great value to sparsity or a dense by choosing the most less.
+This value is defined during the design and determines the main properties of the peer-to-peer network.
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{Circle2-1.png}}}
+\caption{\label{fig:p2p_neighbors_2} The neighbor connections in the peer-to-peer network in case of sparsity=2} }
+\end{figure}
+
+Every node has always exactly n = $\lceil \log_s{N} \rceil$ neighbors where the $N$ is the number of the ISIS registered into the network.
+The node has as neighbor the ISISs with PeerID $s^0$-th, $s^1$-th ...
+$s^{n-1}$-th greater than the own.
+(For example: There is 19 nodes in a network with sparsity=9.
+Then the 1st node has as neighbors the 2nd, 3rd, 5th, 9th, and 17th node as seen on figure \ref{fig:p2p_neighbors_2}.)
+
+There can be exactly n redundancy achieved in the system so at most $n-1$ node can be disappear without any serious communication problem.
+Since we are also using a method that tries to deliver the message even every neighbor is unavailable, this fault tolerance is much greater.
+With this extension, we will pass the message to at least one other available ISIS so the message wont stick if the node is not the only working node in the network.
+This solution provides a secure and quite fast solution for the message deliver but has a high communication cost because there is $n*N$ message necessary in general for every Registration or RemoveRegistrations.
+
+If there is a much greater {\it sparsity} used that's even greater then the number of the expected number of members then every node will have only one neighbor.
+This situation can be seen on Figure \ref{fig:p2p_neighbors_40}.
+In this case it can be such a functioning achieved where is only N (the theoretically minimum) message in the network at the expense of a slower propagation.
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{Circle40-40.png}}}
+\caption{\label{fig:p2p_neighbors_40} The neighbor connections in the peer-to-peer network in case of sparsity>N} }
+\end{figure}
+
+There is a data-driven routing used in the peer-to-peer cloud.
+It means that the given nodes examines the received Registration Entries and store and forward only the newer ones than the already known version belonging to the same service identifier.
+If the information is out-of-date then it will be simple dropped.
+(A Coordinated Universal Time (UTC) standard time and date format is used for the time stamps that is also able to handle the differences originated from the different time zones.) We managed to achieve using this method not to store the already deliver nodes list in the message but deciding about the forwarding locally based on locally available information.
+Since the routing is based on time stamps, it is very important to keep the nodes clock in synch with each other or with an outer reference for example by using NTP.
+
+An other advantage of this solution is the ability of handling the case of medley Register and RemoveRegistrations messages.
+It is possible for a RemoveRegistrations message to forerun a previously generated Register message.
+This cause just a temporary imprecise entry in the database but the mistake can be quickly fixed and won't be much further propagated.
+
+Since we are base our routing to the state of the messages belonging to the service identifier it is important to keep the fact of message deleting for a while.
+If we are receiving a RemoveRegistrations message then do not remove the named entry from the database just change its state to {\it deleted} by keeping only a piece of data about it (the service identifier and the deletion time).
+
+\subsubsection{Maintenance of the peer-to-peer network}
+In the network there is two different maintenance needed. On the one hand the database have to be periodically kept up and on the other hand the list of neighbors have always to be up-to-date.
+
+The database applied in the system is a soft-state database so the entries are {\it valid} just for a limited time.
+Without any registration freshening it's validity will expire and the entry's state will change to {\it expired}.
+The time of expiration have to be defined to every singe message by the registrar entity.
+The RemoveRegistrations message turn the state of registration entry {\it deleted} as it was previously mentioned.
+The {\it expired} or {\it deleted} messages will purged for good and all in the course of the periodical database maintenance.
+The states of registration entries can be shown on Figure \ref{fig:RegEntry states}.
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{RegEntryStates.png}}}
+\caption{\label{fig:RegEntry states} Registration entry states} }
+\end{figure}
+
+The second kind of maintenance is keeping the list of neighbors up-to-date.
+While receiving even Registration or RemoveRegistrations messages or found an outdated entry during the periodical database cleaning we are doing the necessary steps to update our list.
+Since every leaving or entering node has influence to the every node's neighbor registry with a high probability it will be rebuilt in these cases.
+The modification are passing locally because this connections are asymmetric so there is no compliance or network communication is needed.
+
+There is no special action done and no additional network traffic needed during the neighbor checking.
+The fact the node sends periodically messages to every neighbor can be exploited.
+If the node do not success to send its message to the neighbor after some tries, then the unavailable node will be marked.
+If every neighbor will be getting marked then the node change its PeerID and reconnect to the network if any InfoProvider ISIS is available.
+
+When a neighbor is unavailable we keep also trying to deliver the message to the followers of the neighbor until we manged to find one to pass the information or diagnose that every nodes are unavailable.
+Here the follower means the nodes that have greater PeerID than the neighbor has but less then our next neighbor.
 % subsection functionality (end)
@@
 \label{sub:isis_peer_to_peer_interface}
+\subsubsection{Operation Connect}
+
+\begin{description}
+
+  \item[Input]~\begin{description}
+    \item[none]
+  \end{description}
+
+  \item[Output]~\begin{description}
+    \item[RegEntry] Multiple \texttt{RegEntries} locally stored in the ISIS database
+  \end{description}
+
+  \item[Faults]~\begin{description}
+    \item[none]No specific faults are defined
+  \end{description}
+
+\end{description}
+
+This operation allows the ISIS to get the already existing database from the ISIS that helps it to connect the peer-to-peer network.
+As a second step of the connection the newer  \texttt{RegEntries} that are stored at the connecting entity will be propagated through \texttt{Register} operation of the standard service specific interface.
 
@@
   \item RemoveRegistrations with request to remove Registration Message representing not contacting client
-  \item Connect % TODO: describe this operation
+  \item Connect indicating to provide every information about the stored database
  \end{itemize}

### 2009-05-25 doc/tech_doc/infosys/infosys_technical.tex
Technical documentation corrections

--- 2009-05-21 doc/tech_doc/infosys/infosys_technical.tex
@@
 \label{sub:peer_to_peer_functionality}
-
+The ISIS nodes are able to organize themselves into a peer-to-peer network by synchronizing the stored replicated database and maintain this synchronicity and the network topology.
+Every node of this peer-to-peer cloud (that is an ISIS service) has an own PeerID, an identifier only used in this network, made as a hash of its endpoint URL.
+The members are ordered by this identifier where there is also a sequence defined between them where the successor entity is the ISIS with the smallest PeerID of the greater PeerIDs.
+This metric is used for defining the neighbour relation that is the base of the inter-peer-to-peer network communication.
+It is presumed that only registered ISIS nodes can have any role in the cloud.
+
+\subsubsection{Connecting to the peer-to-peer network}
+Every ISIS tries to connect to the network at its startup by accessing the InfoProvider ISIS nodes known from its configuration.
+The choice from among the few pre-configured InfoProviders is random to achieve as much load-balancing as possible.
+(In the design it is assumed that a huge amount of ISIS nodes are configured in very similar way, where the InfoProvider ISIS nodes are mainly the same.) If there is no InfoProvider available or nothing is known from the configuration, the node decides to be the fist member and to build a new network.
+
+If there are any InfoProvider nodes available the list of member ISIS nodes will be queried and the ÒsuccessorÓ will be defined using the previously mentioned metric.
+This ÒsuccessorÓ node will help the entering ISIS to connect to the cloud.
+The new member will use ÒConnectÓ Web Service operation to get the newest version of database stored in the peer-to-peer network, and place its own entries to the cloud by registering them.
+An initial database synchronization can be achieved using this method.
+
+We can prevent the InfoProvider ISIS nodes from overloading by balancing the operation of database synchronizing being the most costly procedure in the connection phase.
+
+There is also an other extension used regarding the InfoProvider ISIS nodes.
+If a node is not able to connect any of the InfoProviders then it tries to repeat it later.
+This results that the network is able to fuse two disjoint parts of network when the missing InfoProvider ISIS appears again.
+
+\subsubsection{Routing in the peer-to-peer network}
+We are keeping the databases in synch by passing every Registration and RemoveRegistrations to every node acting in the network.
+This yields redundancy and the fault tolerant message routing is very important between any two ISIS nodes.
+
+The most important parameter of the network is the {\it sparsity}.
+This is an integer not less then 2.
+It determines the number of neighbors in the function of the actual number of member nodes of the network, where the number of neighbors can change when new members connecting to the network or members leave.
+A sparse graph can be shaped by choosing a great value for sparsity or a dense one by choosing the less.
+This value is defined during the design and determines the main properties of the peer-to-peer network.
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{Circle2-1.png}}}
+\caption{\label{fig:p2p_neighbors_2} The neighbor connections in the peer-to-peer network in case of sparsity=2} }
+\end{figure}
+
+Every node has always exactly n = $\lceil \log_s{N} \rceil$ neighbors where the $N$ is the number of ISIS nodes registered into the network.
+The node has as many neighbors as the ISIS nodes with PeerID $s^0$-th, $s^1$-th ...
+$s^{n-1}$-th greater than its own.
+(For example: There are 19 nodes in this network with sparsity=9.
+Then the first node has the following neighbors: the 2nd, 3rd, 5th, 9th, and 17th node as seen in figure \ref{fig:p2p_neighbors_2}.)
+
+There can be exactly n-way redundancy achieved in the system so at most $n-1$ node can disappear without any serious communication impair.
+Since there is also an emergency method used trying to deliver messages even if all neighbors are unavailable, this fault tolerance is greater.
+With this extension, messages are passed to one of the non-neighbor, yet available ISIS, so the message will surely be delivered to one of the operational node of the network.
+This solution provides a secure and quite fast solution for the message delivery but has a high communication cost because there are $n*N$ messages necessary in general for every Registration or RemoveRegistrations.
+
+If there is a much greater {\it sparsity} used, then the number of the expected members decrease, even only to one neighbor.
+This situation can be seen in Figure \ref{fig:p2p_neighbors_40}.
+In this case such functioning is achieved when there are only N (the theoretically minimum) messages traversing in the network at the expense of a slower data propagation.
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{Circle40-40.png}}}
+\caption{\label{fig:p2p_neighbors_40} The neighbor connections in the peer-to-peer network in case of sparsity>N} }
+\end{figure}
+
+There is a data-driven routing used in the peer-to-peer cloud.
+It means that the given nodes examine the received Registration Entries then store and forward only those messages that are newer than the already stored version belonging to the same service identifier.
+If the information in the message is out-of-date then it will be simple dropped.
+(A Coordinated Universal Time (UTC) standard time and date format is used for the time stamps that is also able to handle the differences originated from the different time zones.) By using this method it is achieved that it is not necessary to store the formerly seen nodes in the messages, but the routing decision that is who to send the message to is based on local information.
+Since the routing is based on time stamps, it is very important to keep the nodes clock in synch with each other or with an outer reference, say, by using NTP.
+
+Another advantage of this solution is the capability of handling the case of swapped Register and RemoveRegistrations messages.
+It is possible for a RemoveRegistrations message to forerun a previously generated Register message.
+This causes just a temporary imprecise entry in the database but the mistake can be quickly fixed and is not much further propagated.
+
+Since the applied routing is based on the state of the messages belonging to some service identifier it is important to keep the fact of message deleting for a while.
+If a RemoveRegistrations message is received, then the named entry from the database is not removed immediately, just its state changes to {\it deleted} and a piece of data about it (the service identifier and the deletion time) is kept.
+
+\subsubsection{Maintenance of the peer-to-peer network}
+In the network there are two different maintenance methods needed.
+On one hand the database have to be periodically cleaned and on the other hand the list of neighbors have always got to be up-to-date.
+
+The database applied in the system is a soft-state database so the entries are {\it valid} just for a limited time.
+Without any registration freshening its validity will expire and the state of the entry will change to {\it expired}.
+The time of expiration is set in every singe message by the registrar entity.
+The RemoveRegistrations message turn the state of registration entry to {\it deleted} as it was previously mentioned.
+The {\it expired} or {\it deleted} messages will be purged in the course of periodical database maintenance.
+The states of registration entries are shown in Figure \ref{fig:RegEntry states}.
+\begin{figure}[ht]
+\centering{{\scalebox{0.6}{\includegraphics{RegEntryStates.png}}}
+\caption{\label{fig:RegEntry states} Registration entry states} }
+\end{figure}
+
+The second kind of maintenance is keeping the list of neighbors up-to-date.
+While receiving even Registration or RemoveRegistrations messages or finding an outdated entry during the periodical database cleaning nodes perform necessary steps to update our list.
+Since every leaving or entering node infulences the neighbor registry of all the other nodes at a high probability the local set of neighbors will be rebuilt in these cases.
+The modifications are passing locally because these connections are asymmetric so there is no compliance or network communication is needed.
+
+There is no special action done and no additional network traffic needed during the neighbor checking.
+The fact the node sends periodically messages to every neighbor can be exploited.
+If the node is not successful in sending its messages to the neighbors after some tries, then the unavailable node will be marked.
+If every neighbor has been marked then the node changes its PeerID and reconnects to the network if there are any InfoProvider ISIS nodes available.
+
+When a neighbor is unavailable ISIS keeps trying to deliver the message to the successor of the neighbor until either the information is passed to at least one available node or it is diagnosed that all the nodes are unavailable.
+Here the successor means thhose nodes that have greater PeerID than the neighbor but less then the next ISIS neighbor.
 % subsection functionality (end)

### 2009-05-25 doc/manuals/janitor/tex_concept/concept.tex
Use pregenerated graphics instead of fancy LaTeX graphics style
Series of updates to the Janitor documentation.

@@
  functional groups. One addresses the retrieval of information from
- the Catalog RDF file in the left major branch of the figure. The other
+ the Catalog's RDF file in the left major branch of the figure. The other
  addresses the process of fetching and installing the packages.
@@
  \begin{center} 
-    \resizebox{24cm}{!}{\input{images/dependencies}}
+%    \resizebox{24cm}{!}{\input{images/dependencies}}
+    \includegraphics[width=24cm]{images/dependencies.png}
     \mycaption{Modules of the Janitor and their dependencies}{}
@@
 
-\section{Runtime environments states}
-A main motivation for the managed, manual initiation of dynamic RE
+\section{States of runtime environments}
+A major motivation for the managed, manual initiation of dynamic RE
 installation is the subsequent manual verification of the installed
@@
 
-These states are specific for every CE and communicated
+These states are specific for every compute element (CE) and communicated
 between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
@@
 
-Upon presentation of a package to a Catalog, a CE may
+Upon presentation of a the package name to a Catalog, from which details about
+the package are retrieved, a CE may
 classify a package to be \texttt{INSTALLABLE} if all the dependencies
@@
 
-Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
-in Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached in Janitor. 
-After Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
-an unforeseen exception occures during that process, Janitor will drop the job from its database and set the affected runtime
+The Janitor manages the states that the runtime environments at a particular
+compute element are in. However, it is also most important for the Janitor
+to be aware of the jobs that depend on the installtion of a RE. REs still in use
+should not be removed until the respective job has completed its computations.
+The installation or removal of REs by the Janitor is perceived as a mere
+consequene of jobs demanding a RE or not, thus, the communcation between
+the job-manager AREX and the Janitor will be performed on that 'job level'.
+
+The Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
+in the Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached. 
+After the Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
+an unforeseen exception occures during that process, the Janitor will drop the job from its database and set the affected runtime
 environments to the state \texttt{FAILED}.
 
-\section{Integration into AREX}
+\section{Integration with AREX}
 
@@
 
-Currently the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
-a setuid wrapper has been written. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s}
-to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\
+In the current implementation the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
+a setuid wrapper has been prepared. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s},
+i.e. the sticky bit is set to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\
 
-Thus, Janitor is not executed directly. Instead a dynamic link will be created in \task{libexec or sbin} which is pointing to
+Thus, Janitor is not executed directly. Instead, a dynamic link will be created in \task{libexec or sbin} which is pointing to
 the wrapper.\\
@@
 
-Security is a major concern for the grid systems. An automatic software
+Security is a major concern for grid systems. Any additional feature
+and especially an automatic software
 installation inheritently introduces security threats. This section
@@
 also authorised to install a REs. Restrictions are
-only imposed on the set of DRE that are available for installation.
+only imposed on the set of dynamic REs that are available for installation.
 Restrictions are imposed by the site admininistrators on the descriptions
 that are given by the Catalog that is offering the package. These
-descriptions may explicitly mention DRE names, a regular expression on
+descriptions may explicitly mention dynamic REs' names, e.g. a regular expression on
 these, or refer to tags of packages that categorise these. However,
-the core of these controls lie with the maintainers of the Catalog,
+the core of these controls lies with the maintainers of the Catalog,
 who needs to be trusted.
@@
 \begin{itemize}
-\item DREs are not distinguished between {\em installed} and
+\item Dynamic REs are not distinguished between {\em installed} and
   {\em installable} in the information system.
-\item No status information is given at the time an DRE is installed.
+\item No status information is given at the time a dynamic RE is installed.
 \end{itemize}
@@
 attacks. What is consequently left to be protected against are unwanted
-sideeffects by the installation of software.
+side-effects by the installation of software.
 
@@
 packages of mainstream Linux distributions) is seeked to reduce the
-complexity and burden in the maintenance for DRE. In the
-current implementation Debian packages may be installed only by their
+complexity and burden in the maintenance for dynamic REs. In the
+current implementation, Debian packages may be installed only by their
 transformation into tar files. With the advent of the interface for

### 2009-05-25 doc/manuals/janitor/tex_installation/installation.tex
Series of updates to the Janitor documentation.

@@
 % ** Dependencies
-% *** Ordianary: log4perl and ``Redland RDF Library - Perl Interface''
+% *** Ordinary: log4perl and ``Redland RDF Library - Perl Interface''
 % *** WebService: libperl-dev
@@
 
-In order to use Janitor, the two perl packages listed in table~\ref{tab:install_dependencies} are required.
-To have the WebService interface for Janitor, the packages listed in the table~\ref{tab:install_dependencies_optional}
-have to be installed before the build process is.
+In order to use the Janitor, the two perl packages listed in
+table~\ref{tab:install_dependencies} are required. To have
+the WebService interface for the Janitor, the packages listed in the
+table~\ref{tab:install_dependencies_optional} need to be installed before
+the build process is.
+
 \begin{table}[!h]
@@
         \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
-                   Redland RDF libraray is used for accessing the knowledge base of RTEs.}
+                   Redland RDF libraray is used for accessing the knowledge base of Runtime Environments.}
         \label{tab:install_dependencies}
@@
 
-Janitor will be installed by default along with ARC1. If desired it is possible to disable the shipping of Janitor 
-using the \textit{configure} flags \textit{--disable-janitor-service} for the complete janitor or 
-\textit{--disable-janitor-webservice} for only the Web Service support.
+The Janitor will be installed by default along with ARC1. Support for
+the dynamic REs can be completely disabled at compile time by using
+the \textit{configure} flags \textit{--disable-janitor-service} for
+the complete janitor or \textit{--disable-janitor-webservice} to only
+disable the Janitor's Web Service.
+
 Furthermore it is recommended to install the ontology editor 
@@
 
-The current version of Janitor can be configured using the common file \textit{arc.conf} which is to be found in the 
-configuration directory \textit{etc}. Janitor is using the environment variable \textit{NORDUGRID\_CONFIG} to determine the location 
-of the corresponding file. If the variable is not set, the default location \textit{/etc/arc.conf} will be used. 
+The current version of the Janitor can be configured using the common file \textit{arc.conf}. It is expected in the 
+configuration directory \textit{etc}. The Janitor is using the environment variable \textit{NORDUGRID\_CONFIG}
+to determine the location of the corresponding file. If that variable is not set, the default location \textit{/etc/arc.conf} will be used. 
 The configuration is assigned by the section \lbrack janitor\rbrack. The table~\ref{tab:arc_conf_tags} contains the available tags
 for janitor configuration.
+
 \begin{landscape}
@@
 
-The parameter \texttt{enable} defines whether Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
+The parameter \texttt{enable} defines whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
 \texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
@@
 In these directories the states of the jobs and the runtime environments will be stored. 
-The knowledge base of installable packages is specified by the parmeter \texttt{catalog}. 
-Its value can be any kind of URL pointing to an file written in the Resource Description Framework (RDF) format.
+The knowledge base of installable packages is specified by the parameter \texttt{catalog}. 
+Its value can be any kind of URL pointing to a file written in the Resource Description Framework (RDF) format.
 The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
-The parameter \texttt{downloaddir} assignes the directory in which the installation files will be saved after they have been 
+The parameter \texttt{downloaddir} assigns the directory in which the installation files will be saved after they have been 
 downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
-directory into which all packages will be installed. Instead of the other directories the \texttt{installationdir} should be 
-available for all computing elements i.e. by using a shared volume.
-If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, Janitor will also
-create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by Janitor.
+directory into which all packages will be installed. Instead of the other directories, the \texttt{installationdir} should be 
+available for all computing elements, i.e. by using a shared volume.
+If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, the Janitor will also
+create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by the Janitor.
 The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
@@
 The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
-or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintain by
+or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintained by
 a higher organization.
@@
 
-Janitor is designed for the usage on Linux distributions.
+The Janitor is designed for the usage on Linux distributions.
+Its porting to other platforms has not yet been addressed.
+
+The ARC middleware is not ultimately essential for dynamic Runtime Environments.
+All the Perl code would be functional with any Grid middleware.
 

### 2009-05-25 doc/manuals/janitor/tex_introduction/introduction.tex
Series of updates to the Janitor documentation.

\chapter{Introduction}

Janitor is a service for the automated installation of runtime environment for grid computing elements.
It is addressed transparently via the A-REX service used for job submission but can also be used as a standalone tool.

\section{Motivation}

A major motivation for grid projects is to stimulate new communities to adopt computational grids for their
causes. From the current grid user's viewpoint, the admission of users of a very different education will
suddenly impose difficulties in the communication between site maintainers. One will not even understand
the respective other side's research aims. Hence, the proper installation of non-standard software (Runtime
Environments) is not guaranteed and the priorities of manual labour will be mostly disjunctive.

A core problem remains to distribute a locally working solution, the Know-How, quickly across all
contributing sites, i. e., without manual interference. Every scientific discipline has its respective own set of
technologies for the distribution of work load. E.g. research in bioinformatics requires access to so many
different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
a commodity, with all their intrinsic problems as there are bottlenecks and restrictions of repeated access.
The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges
with the here presented work extends the NorduGrid's
Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
automated installation of software packages.

An automation of the software installation, referred to as dynamic Runtime Environments, seems the only
approach to use the computational grid to its full potential. Components of workflows shall be spawned
as jobs in a computational grid using dynamic Runtime Environments rather than as shared web services.
The grid introduces an extra level of parallelism that web services cannot provide. The required short
response times and the heterogeneous education of site-administrators on a grid demand an automatism for
the installation of software and databases without manual interference~\cite{BAYER_2007}.

\section{Overview}

This document will start with a chapter on how to set-up the Janitor locally. The following chapter will then give furhter instructions on how to 
use the Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
which is basically covering the method how to prepare new runtime environments. Deeper insights on the design of the Janitor will 
be given by the subsequent forth chapter. In the fifth chapter, an outlook to anticipated future developments will given.

% BASIC CONCEPT

% * Installation

% * Usage

% * Maintainance

% * Programming concept

% * Future work

% * APPENDIX

### 2009-05-25 doc/manuals/janitor/tex_maintenance/maintenance.tex
Series of updates to the Janitor documentation.

\chapter{Maintenance}

This chapter explains how to maintain the Catalog and the Janitor itself. 
In order to administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e. The first
section will explain how this is done.
The second section presents a detailed explanation how to create new packages for the Janitor.
In the last section, a typical use case in maintaining Janitor will be listed.

% * Maintainance
% ** Protege and RDF files
% ** Creating own tar balls

\section{Catalog}\label{sec:catalog}

The Catalog describes runtime environments and is either served through a web server or is distributed together with the Janitor as a regular file. 
It is specifed by a (Resource Description Framework) RDF file assigned to 
Janitor using the tag \texttt{catlog} within the configuration file (see~\ref{sec:janitor_configuration}). 
The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\

In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
    \label{fig:protege_example}
  \end{center}
\end{figure}
On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
It has one or more instances of the class \texttt{Package} and is
described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
(i.e. here etch or sid), i.e. the name of a common installation or a virtual image.
The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
they can be found more easily.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
classes are interacting with each other the Tables~\ref{tab:main_class_meta},\ref{tab:main_class_base}, \ref{tab:main_class_tag},
\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Metapackage.}{}
         \label{tab:main_class_meta}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	homepage       & single                & String       \\
	instance       & multiple              & Instance of Package       \\
	lastupdated    & single                & String       \\
	name           & required single       & String       \\
	tag            & multiple              & Instance of Tag       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Basesystem.}{}
         \label{tab:main_class_base}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	distribution   & required single       & String       \\
	name           & required single       & String       \\
	short\_description & required single   & String       \\
	url            & required single       & String       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Tag.}{}
         \label{tab:main_class_tag}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	name           & required single       & String       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class DebianPackage.}{}
         \label{tab:main_class_debian}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	basesystem     & required single       & Instance of BaseSystem       \\
	debconf        & multiple              & String       \\
	depends        & multiple              & Instance of MetaPackage  or Package      \\
	package        & required multiple              & String      \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class TarPackage.}{}
         \label{tab:main_class_tar}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	basesystem     & required single       & Instance of BaseSystem       \\
	depends        & multiple              & Instance of MetaPackage or Package      \\
	environ        & multiple              & String       \\
	url            & required multiple     & String      \\
	\end{tabular} 
   \end{center}
\end{table}

\section{HTML interface of the catalog}

The dynamic Runtime Environments stored in the Catalog are presented on the aforementioned dedicated web
page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
RDF syntax and an automated transformation to
HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
This HTML page listing the manually or automatically installable REs is prepared by the script web/list.pl.
This script is meant to be run by a mod\_perl-enabled Apache. The script itself is only loosely coupled
with the Janitor. In the first lines of the script some variables specific to the site are set. To configure the
script these have to be changed~\cite[p. 9]{BAYER_2007}.

\section{Introducing new packages}\label{sec:catalog}

This section describes how to add new packages to the Catalog. In the current implementation, only tar based packages are processed by the Janitor.
Within the example they are assigned to be used together with Debian Etch. This limitation is only literal. 
There is no limitiation to newer Debian distributions.

\subsection{Debian Etch (tar based)}

At the time of writing, only the tape archive (tar) file format is accepted for dynamic Runtime Environment installation, 
a well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files. 
Subdirectories are visualised in Figure~\ref{fig:tar_folder}.

\begin{figure}
  \begin{center}
    \includegraphics[width=4cm]{images/tar_folder.pdf}
    \mycaption{Directory structure in the tar files for automated installation.}{}
    \label{fig:tar_folder}
  \end{center}
\end{figure}

The tar file contains two directories, named {\tt control} and {\tt data}. Software is stored in the latter subdirectory,
while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
the content of the data directory is extracted to some directory \textdollar BAR. 
After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
To be offered to computing elements for an installation, the such prepared runtime environment must be
announced to a Catalog to which the Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.

\subsection{Protoypes}

In order to have an impression how the tar files are created, several prototypes are provided at \task{URL required!!}.:\\

The WEKA package for machine learning~\cite{FRANK_20074} and the Java Runtime Environment
are available as dynamic Runtime Environments. Further packages for bioinformatics comprise dynamic
variants of tools for the analysis of transcription factor binding sites. These are already offered for manual
installation via the prior mentioned traditional page representing Runtime Environments for ARC. 
The corresponding tar file is named \task{somewhat??}. The \texttt{data} directory simply contains a ZIP file
which needs to be unzipped in the installation directory. For that reason, the \texttt{control/install} script is 
written as follows:
% dollar signs made trouble in listings *hrmlgrmpf* ... Costs too much time to figure out why *ARC!*
\begin{verbatim} 
#!/bin/sh
set -e  # Makes the script to terminate at the first line it fails.

WEKA_ZIP="weka-3-4-8a.zip"
unzip $WEKA_ZIP
rm -f $WEKA_ZIP
\end{verbatim}
The runtime script sets the environment variable of the Java Classpath:
\begin{verbatim}
#!/bin/sh

WEKA_JAR="weka-3-4-8a/weka.jar"
case "\$1" in
	0)	# Just before job submission
		# none
	;;
	1)	# Just before job execution
		# Initialize the java environment
		CLASSPATH="%BASEDIR%/$WEKA_JAR:$CLASSPATH"
		export CLASSPATH
	;;
	2)	# After job termination
		# none
	;;
	*)
		return 1
	;;
esac	
\end{verbatim}
The remove script, which will be executed right before WEKA is deinstalled, is empty. Janitor will delete the directory, so there
is nothing more to be done.\\

\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\

To address the concerns of the physicists using ARC, a dynamic runtime environment for the
ATLAS software suite was prepared. It extends prior work on an automated installation that is available at
\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}. The preparation comprised the following steps:
\begin{itemize}
    \item The file system path specifications in the automated installation scripts were modified using the Janitor
       path variables.
    \item A tarball was prepared containing a directory structure as illustrated in Figure~\ref{fig:tar_folder}. The data directory
       was empty, since the automatic installation script downloads the software from a remote server.
    \item An entry was added to the Catalog file.
\end{itemize}

What sets High Energy Physics software apart is it's sheer size. The package in question takes up more than
5 GB. This was a test illustrating the feasibility of using dynamic REs in High Energy Physics. The application
of the dynamic RE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
service, e.g. a software manager of a big experiment will be able to deploy software packages on production
sites simply by creating a tarball and adding an entry to the Catalog.

\subsection{Debian Etch}

These kind of packages are not yet supported.

\subsection{Debian Sid}

These kind of packages are not yet supported.

\subsection{Adminstrating the Catalog}

\section{Typical use cases}\label{sec:catalog}

% ** setstate REMOVAL_PENDING
% no warranty interferences in the file syste,

### 2009-05-25 doc/manuals/janitor/tex_titlepage/titlepage.tex
Series of updates to the Janitor documentation.

--- 2009-03-05 doc/ws-programming-tutorial/doc/tex_titlepage/titlepage.tex
@@
 \begin{tabular}{rl}
-\resizebox*{3cm}{!}{\includegraphics{images/ng-logo.png}}
+\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
 &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
@@
 
-{\raggedleft NORDUGRID-TECH-19\par}
+{\raggedleft NORDUGRID-D2.5-2 ???\par}
 
@@
 %%%%---- The title ----
-{\centering \textsc{\Large Web Service programming Tutorial}
+{\centering \textsc{\Large Dynamic Runtime Environment Installation with Janitor}
 \\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
@@
 %%%%---- A list of authors ----
-    {\centering \large Michael Glodek, Steffen M\"oller \par}
+    {\centering \large Michael Glodek\footnote{glodek@inb.uni-luebeck.de}, Daniel Bayer,  Steffen M\"oller\footnote{moeller@inb.uni-luebeck.de} \par}
 

### 2009-05-25 doc/tech_doc/infosys/infosys_technical.tex
Fixing sparsity value typo
Adding example ISIS configuration to the technical documentation as an appendix.

@@
 $s^{n-1}$-th greater than its own.
-(For example: There are 19 nodes in this network with sparsity=9.
+(For example: There are 19 nodes in this network with sparsity=2.
 Then the first node has the following neighbors: the 2nd, 3rd, 5th, 9th, and 17th node as seen in figure \ref{fig:p2p_neighbors_2}.)
@@
 
+\section{Example ISIS service configuration}
+\label{annex:example_isis_configuration}
+\begin{verbatim}
+<ArcConfig
+  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
+  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
+  xmlns:isis="http://www.nordugrid.org/schemas/isis/2008/08"
+  xmlns:infosys="http://www.nordugrid.org/schemas/InfoRegisterConfig/2008"
+>
+  <!-- Various server settings -->
+  <infosys:InfoRegistrar id="NorduGridInfosys">
+    <!-- Contact and behavior details of the InfoRegsitrar element-->
+    <infosys:URL><!-- Available endpoint URL of an ISIS representing the cloud --></infosys:URL>
+    <infosys:Retry>4</infosys:Retry>
+  </infosys:InfoRegistrar>
+
+  <Plugins><Name>isis</Name></Plugins>
+
+  <Chain>
+
+    <!-- Chain configuration goes here! -->
+    <Plexer name="plexer.service" id="plexer">
+      <next id="isis1">^/isis1$</next>
+    </Plexer>
+
+    <Service name="isis" id="isis1">
+        <isis:DBPath>
+            <!-- File location where the database is locally stored -->
+        </isis:DBPath>
+        <isis:endpoint>
+            <!-- The URL where the service can be accessed from outside -->
+        </isis:endpoint>
+        <isis:InfoProviderISIS>
+            <isis:URL>
+                <!-- The URL of one or more ISIS where the other ISIS's can be queried. -->
+            </isis:URL>
+        </isis:InfoProviderISIS>
+        <infosys:InfoRegister>
+            <infosys:Period>
+                <!-- How often should the service be registered. 
+                For example: PT75S for 75 seconds -->
+            </infosys:Period>
+            <infosys:Endpoint>
+                <!-- The URL where the service can be accessed from outside -->
+            <infosys:Endpoint>
+            <infosys:Expiration>
+                <!-- The lifetime of the RegEntry provided by the service about itself.
+                For example: PT5H for 5 hours -->
+            </infosys:Expiration>
+            <!-- If there is no specific Registrar defined then every InfoRegistrar
+             will be use used during the registration. The period of the repeating 
+             registration will be the value given here in every ISIS cloud. -->
+        </infosys:InfoRegister>
+    </Service>
+  </Chain>
+</ArcConfig>
+
+\end{verbatim}
+% section example_ISIS_configuration (end)
+
 % chapter appendices (end)

### 2009-05-25 doc/manuals/janitor/tex_usage/usage.tex
Series of updates to the Janitor documentation.
Some typos fixed.

@@
 
-Janitor can be used either with or without A-REX. In case A-REX is used, installations will be maintained automatically the runtime environments.
+The Janitor can be used either with or without A-REX. In case A-REX is used, installations will be maintained automatically the runtime environments.
 
@@
 Runtime Environments can be specified using the supported job description languages.
-Representative two common languages shall be explained at this point: xRSL and JSDL.
+The most representative two common languages shall be explained at this point: xRSL and JSDL.
 Listing~\ref{lst:xrsl_job} shows the xRSL example in which two runtime environments are requested.
@@
 
-In addition of using Janitor together with A-REX, Janitor can also be used as a standalone commandline tool.
+In addition to using the Janitor together with A-REX, the Janitor can also be used as a standalone commandline tool.
 The available commands are listed in the Table~\ref{janitor_commandline_man}.
@@
 \end{table}
-The most important commands for Janitor are \texttt{register}, \texttt{deploy} and \texttt{remove}. In order to 
+The most important commands for the Janitor are \texttt{register}, \texttt{deploy} and \texttt{remove}. To 
 register a job along with a set of runtime environments in Janitor, the first command \texttt{register} followed
 by a job identifier and a list of runtime environments has to be used.
-A job is identified by sequence of numbers. Runtime environments are specified by a string containing the name as it is defined 
-within the catalog (resp. the runtime directory of the grid-manager).
-The command \texttt{deploy} extracts the necessary dependcies of the desired runtime environments and then downloads and installs 
+A job is identified by a sequence of numbers. Runtime environments are specified by a string containing the name as it is defined 
+within the Catalog (resp. the runtime directory of the grid-manager).
+The command \texttt{deploy} extracts the necessary dependencies of the desired runtime environments to then download and install
 the required packages. %How dependecies can are defined will be explained later in the section~\ref{sec:catalog}.
@@
 
-Each command has a certain behaviour for its exit status. The Table~\ref{tab:janitor_commandline_exit_status} lists the possible
+Every command has a certain behaviour for its exit status. The Table~\ref{tab:janitor_commandline_exit_status} lists the possible
 outcomes.

### 2009-05-26 doc/internal_notes/client/GUI_LAP_plans.tex
Update of text

@@
 \section*{The Lunarc Application Portal - A graphical grid user interface using ARC}
-The Lunarc Application Portal (LAP) is an existing web interface for submission and management of grid jobs.
-As reflected in the name the LAP is an application oriented portal, providing easy to use interfaces for commonly available applications.
-The LAP is also a framework for implementing web interfaces for user applications.
+The Lunarc Application Portal (LAP) is an existing web interface for
+submission and management of grid jobs. As reflected in the name the
+LAP is an application oriented portal, providing easy to use
+interfaces for commonly available applications. The LAP is also a
+framework for implementing web interfaces for user applications.
 
-A new and updated version of LAP is presently under development which will make the LAP to Web 2.0 compliant and which will interface to the next generation ARC client library.
-The latter interfacing is done in collaboration between LAP and the KnowARC project and will allow the KnowARC project to use the next generation LAP for it's deliverable D2.2-2 "Advanced graphical client with easy-to-use job creation module and support for bulk job management and monitoring".
+A new and updated version of LAP is presently under development which
+will make use of Web 2.0 features and which will interface to the next
+generation ARC client library. The latter interfacing is done in
+collaboration between LAP and the KnowARC project and will allow the
+KnowARC project to use the next generation LAP for it's deliverable
+D2.2-2 "Advanced graphical client with easy-to-use job creation module
+and support for bulk job management and monitoring".
 
 \section*{Functionality to be offered though LAP}
-Below functionality to be offered through the LAP and which will be a part of the KnowARC deliverable D2.2-2 is listed and discussed:
+Below functionality to be offered through the LAP and which will be a
+part of the KnowARC deliverable D2.2-2 is listed and discussed:
 
 \begin{itemize}
-\item{The portal will support bulk job handling, i.e.
-it will support listing and sorting of (several) hundred(s) of jobs using a java table.
-Using the same table bulk handling will be enabled making it possible to select groups of jobs to execute commands on like e.g.
-kill, clean, get, resume etc.
-Will it be possible to group jobs together?}
-\item{Job statuses will be updated automatically by a background processes, but a "manual update" button will also be provided.}
-\item{Job "babysitting" will offer to automatically download finished jobs or resubmit jobs that have failed.
-Resubmission will be limited to only 2(3?) attempts.
-It will be possible to activate "babysitting" at any time for any kind of job.}
-\item{Users will be able to create and maintain lists of favorite clusters. Blacklisting of clusters will also be possible.}
-\item{The LAP is application orieented and new grid jobs are created through an application specific interface.
-The new and updated version of LAP will offer an application unspecific interface allowing users to create any kind of grid job.
-The interface will offer the user an easy to use job description creation module and the possibility to upload executable and job input files.}
-\item{Occasionally users would like to use existing job description files, but still submit their grid job through the LAP.
-To support this the LAP will offer an interface for uploading and submitting job description files.}
+\item{The portal will support bulk job handling, i.e. it will support
+  listing and sorting of (several) hundred(s) of jobs using a
+  table. Using the same table bulk handling will be enabled making it
+  possible to select groups of jobs to execute commands on like
+  e.g. kill, clean, get, resume etc. }
+\item{Job statuses will be updated automatically by a background
+  processes, but a "manual update" button will also be provided.}
+\item{Job "babysitting" will offer to automatically download finished
+  jobs or resubmit jobs that have failed. Resubmission will be limited
+  to only 2(3?) attempts. It will be possible to activate
+  "babysitting" at any time for any kind of job.}
+\item{Users will be able to create and maintain lists of favorite
+  clusters. Blacklisting of clusters will also be possible.}
+\item{The LAP is application orieented and new grid jobs are created
+  through an application specific interface. The new and updated
+  version of LAP will offer an application unspecific interface
+  allowing users to create any kind of grid job. The interface will
+  offer the user an easy to use job description creation module and
+  the possibility to upload executable and job input files.}
+\item{Occasionally users would like to use existing job description
+  files, but still submit their grid job through the LAP. To support
+  this the LAP will offer an interface for uploading and submitting
+  job description files.}
 \end{itemize}

### 2009-06-08 doc/manuals/janitor/tex_installation/installation.tex
Michael's additions and some pending corrections by Steffen.

@@
 
-In order to use the Janitor, the two perl packages listed in
-table~\ref{tab:install_dependencies} are required. To have
+The Janitor requires the two perl packages listed in
+table~\ref{tab:install_dependencies}. To have
 the WebService interface for the Janitor, the packages listed in the
 table~\ref{tab:install_dependencies_optional} need to be installed before
-the build process is.
+the build process is. The Perl modules are available on CPAN and ship
+with all major Linux distributions.
 
@@
         \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
-                   Redland RDF libraray is used for accessing the knowledge base of Runtime Environments.}
+                   Redland RDF library is used for accessing the knowledge base of Runtime Environments.}
         \label{tab:install_dependencies}
@@
 Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
-knowledge database of installable packages.
+knowledge database of installable packages. At the time of writing, no Linux distribution is offering packages for this fine tool.
 
@@
 to determine the location of the corresponding file. If that variable is not set, the default location \textit{/etc/arc.conf} will be used. 
-The configuration is assigned by the section \lbrack janitor\rbrack. The table~\ref{tab:arc_conf_tags} contains the available tags
-for janitor configuration.
+The configuration is assigned by the section \lbrack janitor\rbrack. Table~\ref{tab:arc_conf_tags} describes the available tags
+for Janitor's configuration.
 
@@
 
-The parameter \texttt{enable} defines whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
+The parameter \texttt{enable} specifies whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
 \texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
@@
 downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
-directory into which all packages will be installed. Instead of the other directories, the \texttt{installationdir} should be 
-available for all computing elements, i.e. by using a shared volume.
+directory into which all packages will be installed and should be 
+available for all computing elements, i.e.\ by using a shared volume.
 If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, the Janitor will also
@@
 The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
-An examples how to configure arc and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf} and.
+An example how to configure ARC and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.
 

### 2009-06-08 doc/manuals/janitor/tex_introduction/introduction.tex
Michael's additions and some pending corrections by Steffen.

@@
 Janitor is a service for the automated installation of runtime environment for grid computing elements.
-It is addressed transparently via the A-REX service used for job submission but can also be used as a standalone tool.
+It is addressed transparently via the A-REX service for job submission but can also be used as a standalone tool.
 
@@
 A core problem remains to distribute a locally working solution, the Know-How, quickly across all
-contributing sites, i. e., without manual interference. Every scientific discipline has its respective own set of
+contributing sites, i.e., without manual interference. Every scientific discipline has its respective own set of
 technologies for the distribution of work load. E.g. research in bioinformatics requires access to so many
@@
 a commodity, with all their intrinsic problems as there are bottlenecks and restrictions of repeated access.
-The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges
+The EU project KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges
 with the here presented work extends the NorduGrid's
@@
 approach to use the computational grid to its full potential. Components of workflows shall be spawned
-as jobs in a computational grid using dynamic Runtime Environments rather than as shared web services.
-The grid introduces an extra level of parallelism that web services cannot provide. The required short
+as jobs in a computational grid using dynamic Runtime Environments rather than accessing a web service
+at one particular machine that is shared amongst all users.
+The grid introduces an extra level of parallelism that web services cannot provide. The demands for short
 response times and the heterogeneous education of site-administrators on a grid demand an automatism for
@@
 
-This document will start with a chapter on how to set-up the Janitor locally. The following chapter will then give furhter instructions on how to 
+This document first dedicates a chapter on how to set-up the Janitor locally. It is followed by a chapter that gives furhter instructions on how to 
 use the Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
 which is basically covering the method how to prepare new runtime environments. Deeper insights on the design of the Janitor will 
-be given by the subsequent forth chapter. In the fifth chapter, an outlook to anticipated future developments will given.
+be given by the forth chapter. The document ends with an outlook to anticipated future developments and opportunities.
+
+\section*{Abbreviations}
+
+\begin{tabular}{l@{--}p{150mm}}
+RTE&Runtime Environment\\
+dRTE&dynamic Runtime Environment\\
+RDF&Resource Description Framework (supporting the RTE Catalog)\\
+\end{tabular}
 

### 2009-06-08 doc/manuals/janitor/tex_maintenance/maintenance.tex
Michael's additions and some pending corrections by Steffen.

@@
 
-This chapter explains how to maintain the Catalog and the Janitor itself. 
-In order to administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e. The first
-section will explain how this is done.
-The second section presents a detailed explanation how to create new packages for the Janitor.
-In the last section, a typical use case in maintaining Janitor will be listed.
+This chapter explains how to maintain the {\em Catalog} and the {\em Janitor} itself. 
+It gives detailed instructions how to create new packages for the Janitor.
+To administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e,
+as explained in the next section.
+In the last section, a typical use case in maintaining the Janitor will be shown.
 
@@
 It is specifed by a (Resource Description Framework) RDF file assigned to 
-Janitor using the tag \texttt{catlog} within the configuration file (see~\ref{sec:janitor_configuration}). 
+Janitor using the tag \texttt{catalog} within the configuration file (see~\ref{sec:janitor_configuration}). 
 The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
@@
 They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
-classes are interacting with each other the Tables~\ref{tab:main_class_meta},\ref{tab:main_class_base}, \ref{tab:main_class_tag},
+classes are interacting with each other the tables~\ref{tab:main_class_meta}, \ref{tab:main_class_base}, \ref{tab:main_class_tag},
 \ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.
@@
 page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
-RDF syntax and an automated transformation to
+RDF syntax and its automated transformation to
 HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
@@
 This HTML page listing the manually or automatically installable REs is prepared by the script web/list.pl.
-This script is meant to be run by a mod\_perl-enabled Apache. The script itself is only loosely coupled
-with the Janitor. In the first lines of the script some variables specific to the site are set. To configure the
+This script is meant to be run by a mod\_perl-enabled Apache. The script itself does not contribute to the 
+core functionality of the Janitor.
+In the first lines of the script some variables specific to the site are set. To configure the
 script these have to be changed~\cite[p. 9]{BAYER_2007}.
@@
 At the time of writing, only the tape archive (tar) file format is accepted for dynamic Runtime Environment installation, 
-a well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files. 
+a well accepted file format throughout the UNIX community. The reflects the traditional manual approach towards RTE in ARC,
+for which one directory is made available to all compute nodes. This section explains the inner structure of the tar files. 
 Subdirectories are visualised in Figure~\ref{fig:tar_folder}.
@@
 installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
-To be offered to computing elements for an installation, the such prepared runtime environment must be
-announced to a Catalog to which the Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.
+Once the tar file was prepared, it must its entry to a RTE Catalogue~\cite[p. 10]{BAYER_2007}.
+
+From such Catalogues, the Janitor finds all information to install packages that possibly have never been installed
+on the site before. The offers of RTEs in a catalogue are cross-checked against the local infrastructure and a subset
+of the available packages will be accepted as "installable". This list of installable RTEs is forwarded to the
+grid information system.
 
@@
 
-In order to have an impression how the tar files are created, several prototypes are provided at \task{URL required!!}.:\\
+In order to have an impression how the tar files are created, several
+prototypes are provided at \url{http://dre.knowarc.eu}.
 
@@
 
-\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
-provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
-dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
-thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
-projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\
+%\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
+%provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
+%dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
+%thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
+%projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\
 
@@
 
-\subsection{Debian Etch}
-
-These kind of packages are not yet supported.
-
-\subsection{Debian Sid}
-
-These kind of packages are not yet supported.
-
-\subsection{Adminstrating the Catalog}
+%\subsection{Debian Etch}
+%
+%These kind of packages are not yet supported.
+%
+%\subsection{Debian Sid}
+%
+%These kind of packages are not yet supported.
+%
+%\subsection{Adminstrating the Catalog}
 

### 2009-06-08 doc/manuals/janitor/tex_usage/usage.tex
Michael's additions and some pending corrections by Steffen.

@@
 \end{lstlisting}
+\task{The runtime environment names are composed out a directory name, the package name and the version number.}
 
@@
 
-In addition to using the Janitor together with A-REX, the Janitor can also be used as a standalone commandline tool.
-The available commands are listed in the Table~\ref{janitor_commandline_man}.
+In addition to using the Janitor together with A-REX, the Janitor can also be used as a standalone command line tool.
+The available commands are listed in table~\ref{janitor_commandline_man}.
 \begin{table}[!h]
@@
 	\multicolumn{3}{l}{\textbf{Job id:}}\\
-	&					& A unique sequence of numbers. Once Janitor registered a job id it cannot register a second job having the same job id.\\
+	&					& A unique sequence of numbers. Once Janitor registered a job id, it cannot register a second job having the same job id.\\
 	\multicolumn{3}{l}{\textbf{Runtime environments:}}\\
@@
 % PROCEED HERE!!!!!!!! !! ! ! !
-In order to remove jobs registered in Janitor the command \texttt{remove} has to be used. 
-The command only removes the job entry and the lock on the runtime environment. If there are no more locks on the runtime environment
-it might be deleted for real.
+In order to remove jobs registered in Janitor, the command \texttt{remove}
+has to be used.  The command only removes the job entry and the lock
+on the runtime environment. If there are no more locks on the runtime
+environment it might be deleted for real. The demand to pass a job
+number for the removal of a RTE is irritating at first.  This shall
+prevent the removal of runtime envrironments that are still being used
+by jobs in the system. Instead, the janitor is informed about a job's
+termination and is requested to remove the assignment of that job to the
+runtime environment. Only those RTEs with no job-assignment are eligible
+for being sweeped. RTEs come with an expiry time or the command may be
+performed via the command line.
 
-Easy commandline examples are provided in Listing~\ref{lst:janitor_example}.
+Easy command line examples are provided in Listing~\ref{lst:janitor_example}.
 

### 2009-06-12 doc/tech_doc/infosys/infosys_technical_internal.tex
Keep infosys technical documentation status during the internal review

--- 2009-05-25 doc/tech_doc/infosys/infosys_technical.tex
@@
   \scalebox{0.6}{\includegraphics{SystemOverview.png}}
-  \caption{\label{fig:infosys_architecture}Architecture of Information System}
+  \caption{\label{fig:infosys_architecture}Architecture of the Information System}
  }
@@
 
-Information System of ARC middleware is composed of 3 main parts.
+The Information System of the ARC middleware is composed of 3 main parts.
 The Information System Indexing Services (ISIS) form a set of information containers.
-Every generic ARC Service pushes information about itself into nearby ISIS container/service hence registering itself to Information System.
+Every generic ARC Service pushes information about itself into a nearby ISIS container/service hence registering itself to the Information System.
 The information stored in every ISIS is propagated among all ISIS services.
-Clients can query any nearby ISIS service for registered information in order to perform discovery of services which posses specific properties.
+Clients can query any nearby ISIS service for registered information in order to perform discovery of services that possess specific properties.
 Clients can also query ARC Services to obtain more detailed information about Service properties.
@@
 
- The functionality of ISIS is made of two parts. On the one hand they are working as ordinary Web Services, and on the other hand maintain a peer-to-peer network - \texttt{ISIS cloud}.
-
- Main functionality of ISIS service visible from outside of ISIS cloud is to accept registration and provide collected information to clients.
-For that ISIS implements operations described in following section.
-The single ISIS service accepts Registration Records pushed to it by other services (including ISIS services too) and stores them in local XML database.
-Stored records can be queried by clients using mandatory and service-specific attributes for selection criteria.
-
- In case of multiple ISIS services they form a peer-to-peer network.
-Inside that cloud Registration Records are propagated between services in such a way that all ISIS instances continuously try to synchronize their databases.
+ The functionality of the ISIS services consists of two parts. On the one hand they are working as ordinary Web Services, and on the other hand they maintain a peer-to-peer network --- the ISIS cloud.
+
+ The main functionality of the ISIS service visible from the outside of the ISIS cloud is to accept registrations from and provide collected information to clients.
+For that ISIS implements the operations described in following section.
+A single ISIS service accepts Registration Records pushed to it by other services (including ISIS services) and stores them in a local XML database.
+Stored records can be queried by clients using mandatory and service-specific attributes as selection criteria.
+
+ When multiple ISIS services are used they form a peer-to-peer network.
+Inside this cloud Registration Records are propagated between services in such a way that all ISIS instances continuously try to synchronize their databases.
 
@@
       \item[RequesterID] Identifier of the client.
-      \item[MessageGenerationTime] Time when following set of RegEntry was generated. There may be multiple RegEntry elements.
+      \item[MessageGenerationTime] Time when the following set of RegEntries was generated. There may be multiple RegEntry elements.
     \end{description}
@@
       \item[SrcAdv]~\begin{description}
-        \item[Type] Type of service being registered. This element is opaque string for now. There shall be service types defined later.
+        \item[Type] Type of service being registered. This element is an opaque string for now. There shall be service types defined later.
         \item[EPR] Endpoint Reference of service being registered in terms of WS-Addressing.
@@
 
-This operation is usually called by service which wants to register it's presence in ISIS.
-This message consists of one or more \textbf{Registration Entry} and at most one \textbf{Registration Header}.
+This operation is usually called by a service that wants to register its presence in ISIS.
+This message consists of one or more \textbf{Registration Entries} and at most one \textbf{Registration Header}.
 The \textbf{Registration Entry}(RegEntry) contains a \textbf{Service Advertisement}(SrcAdv) and a corresponding \textbf{Service Advertisement Metadata}(MetaSrcAdv).
@@
 
-Service must supply mandatory information.
-Among those the \texttt{Endpoint Reference} is used to contact service - only required element is contact URL of service.
-\texttt{Type} specifies kind of service and is used to find out functionality and interface of service.
-\texttt{ServiceID} is used to distinguish between registered services and to deal with case of service changing it's contact URL.
-For more information about mandatory and optional information see Section \ref{service_advertisement}
-
-As result of this operation new Registration Record is stored inside ISIS internal database and eventually propagated to other ISISes.
-If registration record of same ID already existed it will be renewed.
-
-This operation is also used by ISIS services to propagate Registration Records inside ISIS cloud.
+The service must supply mandatory information.
+This includes the \texttt{Endpoint Reference} used to contact the service --- the only required element is the contact URL of the service.
+\texttt{Type} specifies the kind of service and is used to find out functionality and interface of the service.
+\texttt{ServiceID} is used to distinguish between registered services and to deal with the case when the service changes its contact URL.
+For more information about mandatory and optional information see Section \ref{service_advertisement}.
+
+As a result of this operation a new Registration Record is stored inside the ISIS internal database and eventually propagated to other ISISes.
+If a registration record with the same ID already existed it will be renewed.
+
+This operation is also used by ISIS services to propagate Registration Records inside the ISIS cloud.
 
@@
 
-This operation is used to explicitly requesting the removal of zero or more Registration Entries associated with specified ServiceID values stored in the Information System.
-If corresponding record does not exist it's identifier will be present in response message together with corresponding Fault element.
+This operation is used to explicitly request the removal of zero or more Registration Entries associated with specified ServiceID values stored in the Information System.
+If the corresponding record does not exist its identifier will be present in the response message together with a corresponding Fault element.
 
@@
 This operation returns the known peer-to-peer neighbors.
-If client is interested in all ISIS instances they can be obtained using the Query operation because they are ordinary services registered into the Information System.
+If a client is interested in all ISIS instances they can be obtained using the Query operation because they are ordinary services registered into the Information System.
 
@@
 This operation allows any XPath queries to be performed on stored Registration Records.
-The records are treated as merged in one XML document with each record being equivalent to \texttt{RegEntry} element of \texttt{Register} operation.
-In response all elements produced by XPath query are returned.
+The records are treated as merged in one XML document with each record being equivalent to a \texttt{RegEntry} element of a \texttt{Register} operation.
+In the response all elements produced by the XPath query are returned.
 The purpose of this operation is to make it possible to obtain any kind of information related to the Indexing Database.
@@
 \label{sub:peer_to_peer_functionality}
-The ISIS nodes are able to organize themselves into a peer-to-peer network by synchronizing the stored replicated database and maintain this synchronicity and the network topology.
-Every node of this peer-to-peer cloud (that is an ISIS service) has an own PeerID, an identifier only used in this network, made as a hash of its endpoint URL.
-The members are ordered by this identifier where there is also a sequence defined between them where the successor entity is the ISIS with the smallest PeerID of the greater PeerIDs.
-This metric is used for defining the neighbour relation that is the base of the inter-peer-to-peer network communication.
+The ISIS nodes are able to organize themselves into a peer-to-peer network by synchronizing the stored replicated database and maintaining this synchronicity and the network topology.
+Every node of this peer-to-peer cloud (that is an ISIS service) has its own PeerID, an identifier only used in this network, made as a hash of its endpoint URL.
+The members are ordered by this identifier in a ring topology, where the successor entity of the ISIS with the largest PeerID is the ISIS with the smallest PeerID.
+This metric is used to define the neighbor relations that are the basis of the inter-peer-to-peer network communication.
 It is presumed that only registered ISIS nodes can have any role in the cloud.
@@
 The choice from among the few pre-configured InfoProviders is random to achieve as much load-balancing as possible.
-(In the design it is assumed that a huge amount of ISIS nodes are configured in very similar way, where the InfoProvider ISIS nodes are mainly the same.) If there is no InfoProvider available or nothing is known from the configuration, the node decides to be the fist member and to build a new network.
-
-If there are any InfoProvider nodes available the list of member ISIS nodes will be queried and the ÒsuccessorÓ will be defined using the previously mentioned metric.
-This ÒsuccessorÓ node will help the entering ISIS to connect to the cloud.
-The new member will use ÒConnectÓ Web Service operation to get the newest version of database stored in the peer-to-peer network, and place its own entries to the cloud by registering them.
+(In the design it is assumed that a huge amount of ISIS nodes are configured in very similar way, where the InfoProvider ISIS nodes are mainly the same.) If there is no InfoProvider available or nothing is known from the configuration, the node decides to be the first member and to build a new network.
+
+If there are any InfoProvider nodes available the list of member ISIS nodes will be queried and the `successor' will be defined using the previously mentioned metric.
+This `successor' node will help the entering ISIS to connect to the cloud.
+The new member will use the `Connect' Web Service operation to get the newest version of the database stored in the peer-to-peer network, and place its own entries in the cloud by registering them.
 An initial database synchronization can be achieved using this method.
 
-We can prevent the InfoProvider ISIS nodes from overloading by balancing the operation of database synchronizing being the most costly procedure in the connection phase.
+We can prevent the InfoProvider ISIS nodes from overloading by balancing the operation of database synchronizing --- this being the most costly procedure in the connection phase.
 
 There is also an other extension used regarding the InfoProvider ISIS nodes.
-If a node is not able to connect any of the InfoProviders then it tries to repeat it later.
-This results that the network is able to fuse two disjoint parts of network when the missing InfoProvider ISIS appears again.
+If a node is not able to connect to any of the InfoProviders then it tries to repeat it later.
+This results in that the network is able to fuse two disjoint parts of the network when the missing InfoProvider ISIS appears again.
 
@@
 We are keeping the databases in synch by passing every Registration and RemoveRegistrations to every node acting in the network.
-This yields redundancy and the fault tolerant message routing is very important between any two ISIS nodes.
+This yields redundancy and fault tolerant message routing between any two ISIS nodes.
 
 The most important parameter of the network is the {\it sparsity}.
-This is an integer not less then 2.
-It determines the number of neighbors in the function of the actual number of member nodes of the network, where the number of neighbors can change when new members connecting to the network or members leave.
-A sparse graph can be shaped by choosing a great value for sparsity or a dense one by choosing the less.
+This is an integer not less than 2.
+It determines the number of neighbors as a function of the actual number of member nodes of the network, where the number of neighbors can change when new members connect to the network or old members leave.
+A sparse graph can be shaped by choosing a great value for sparsity or a dense one by choosing a smaller one.
 This value is defined during the design and determines the main properties of the peer-to-peer network.
@@
 There can be exactly n-way redundancy achieved in the system so at most $n-1$ node can disappear without any serious communication impair.
-Since there is also an emergency method used trying to deliver messages even if all neighbors are unavailable, this fault tolerance is greater.
-With this extension, messages are passed to one of the non-neighbor, yet available ISIS, so the message will surely be delivered to one of the operational node of the network.
-This solution provides a secure and quite fast solution for the message delivery but has a high communication cost because there are $n*N$ messages necessary in general for every Registration or RemoveRegistrations.
-
-If there is a much greater {\it sparsity} used, then the number of the expected members decrease, even only to one neighbor.
+Since there is also an emergency method used trying to deliver messages even if all neighbors are unavailable, the fault tolerance is even greater.
+With this extension, messages are passed to one of the non-neighbor, yet available, ISIS, so the message will surely be delivered to one of the operational node of the network.
+This solution provides a secure and quite fast solution for the message delivery but has a high communication cost because there are $n \times N$ messages necessary in general for every Registration or RemoveRegistrations.
+
+If there is a much greater {\it sparsity} used, then the number of expected members decreases, in the extreme case to only one neighbor.
 This situation can be seen in Figure \ref{fig:p2p_neighbors_40}.
-In this case such functioning is achieved when there are only N (the theoretically minimum) messages traversing in the network at the expense of a slower data propagation.
+In this case there are only N (the theoretically minimum) messages traversing in the network at the expense of a slower data propagation.
 \begin{figure}[ht]
 \centering{{\scalebox{0.6}{\includegraphics{Circle40-40.png}}}
-\caption{\label{fig:p2p_neighbors_40} The neighbor connections in the peer-to-peer network in case of sparsity>N} }
+\caption{\label{fig:p2p_neighbors_40} The neighbor connections in the peer-to-peer network in case of sparsity $>$ N} }
 \end{figure}
@@
 There is a data-driven routing used in the peer-to-peer cloud.
-It means that the given nodes examine the received Registration Entries then store and forward only those messages that are newer than the already stored version belonging to the same service identifier.
-If the information in the message is out-of-date then it will be simple dropped.
-(A Coordinated Universal Time (UTC) standard time and date format is used for the time stamps that is also able to handle the differences originated from the different time zones.) By using this method it is achieved that it is not necessary to store the formerly seen nodes in the messages, but the routing decision that is who to send the message to is based on local information.
+This means that the nodes examine the received Registration Entries and then store and forward only those messages that are newer than the already stored version belonging to the same service identifier.
+If the information in the message is out-of-date it will be simply dropped.
+(A Coordinated Universal Time (UTC) standard time and date format is used for the time stamps that is also able to handle the differences originated from the different time zones.) By using this method it is not necessary to store the formerly seen nodes in the messages, but the routing decision --- that is who to send the message to --- is based on local information.
 Since the routing is based on time stamps, it is very important to keep the nodes clock in synch with each other or with an outer reference, say, by using NTP.
@@
 Since the applied routing is based on the state of the messages belonging to some service identifier it is important to keep the fact of message deleting for a while.
-If a RemoveRegistrations message is received, then the named entry from the database is not removed immediately, just its state changes to {\it deleted} and a piece of data about it (the service identifier and the deletion time) is kept.
+If a RemoveRegistrations message is received, then the named entry from the database is not removed immediately, instead its state changes to {\it deleted} and a piece of data about it (the service identifier and the deletion time) is kept.
 
 \subsubsection{Maintenance of the peer-to-peer network}
-In the network there are two different maintenance methods needed.
-On one hand the database have to be periodically cleaned and on the other hand the list of neighbors have always got to be up-to-date.
+In the network there are two different maintenance methods needed. On one hand the database has to be periodically cleaned and on the other hand the list of neighbors has always to be up-to-date.
 
 The database applied in the system is a soft-state database so the entries are {\it valid} just for a limited time.
-Without any registration freshening its validity will expire and the state of the entry will change to {\it expired}.
+Without any registration freshening the entry's validity will expire and its state will change to {\it expired}.
 The time of expiration is set in every singe message by the registrar entity.
-The RemoveRegistrations message turn the state of registration entry to {\it deleted} as it was previously mentioned.
+The RemoveRegistrations message turns the state of the registration entry to {\it deleted} as was previously mentioned.
 The {\it expired} or {\it deleted} messages will be purged in the course of periodical database maintenance.
@@
 The second kind of maintenance is keeping the list of neighbors up-to-date.
-While receiving even Registration or RemoveRegistrations messages or finding an outdated entry during the periodical database cleaning nodes perform necessary steps to update our list.
+While receiving Registration or RemoveRegistrations messages or when finding an outdated entry during the periodical database cleaning nodes perform the necessary steps to update the list.
 Since every leaving or entering node infulences the neighbor registry of all the other nodes at a high probability the local set of neighbors will be rebuilt in these cases.
-The modifications are passing locally because these connections are asymmetric so there is no compliance or network communication is needed.
+These modifications are passed locally because these connections are asymmetric so there is no compliance or network communication needed.
 
 There is no special action done and no additional network traffic needed during the neighbor checking.
-The fact the node sends periodically messages to every neighbor can be exploited.
+The fact that the node periodically sends messages to every neighbor can be exploited.
 If the node is not successful in sending its messages to the neighbors after some tries, then the unavailable node will be marked.
@@
 When a neighbor is unavailable ISIS keeps trying to deliver the message to the successor of the neighbor until either the information is passed to at least one available node or it is diagnosed that all the nodes are unavailable.
-Here the successor means thhose nodes that have greater PeerID than the neighbor but less then the next ISIS neighbor.
+Here the successor means those nodes that have greater PeerID than the neighbor but less then the next ISIS neighbor.
 % subsection functionality (end)
@@
 
-To ensure information stored in ISIS cloud can't be tampered and only available to proper clients following authorization framework is implemented.
-All actions performed by ISIS clients are divided into three following groups:
+To ensure that the information stored in the ISIS cloud can't be tampered with and only is available to proper clients the following authorization framework is implemented.
+All actions performed by ISIS clients are divided into the three following groups:
 
 \begin{itemize}
- \item Operations initiated by other ISISes in the cloud. Those include:
+ \item Operations initiated by other ISIS services in the cloud. Those include:
  \begin{itemize}
@@
 
- \item Operations initiated by the Services registering to Information System. Those are:
+ \item Operations initiated by the Services registering to the Information System. Those are:
  \begin{itemize}
@@
 
- \item Operations which are allowed for any liable client of particular Grid infrastructure.
+ \item Operations which are allowed for any liable client of a particular Grid infrastructure.
  \begin{itemize}
@@
 
-Those 3 action groups are handled using Security Framework of ARC~\cite{arcsec}. For each group corresponding 
-Action is defined for ARC policy language as described in table \ref{tab:policy_actions}.
+Those 3 action groups are handled using Security Framework of ARC~\cite{arcsec}. For each group a corresponding
+Action is defined for the ARC policy language as described in table \ref{tab:policy_actions}.
 
@@
 Access restrictions for clients are defined using ARC Policies which are specified and 
-processed using generic Security Handlers approach. Corresponding Security Handlers have to be 
-configured inside configuration block of corresponding ISIS service and attached 
+processed using a generic Security Handlers approach. The corresponding Security Handlers have to be 
+configured inside the configuration block of the corresponding ISIS service and attached to the
 \textit{incoming} queue.
@@
 
-Because all ISIS services trust each other they freely exchanging collected information. But 
-not all information stored inside ISIS cloud is public or readable by clients authorized according 
-to procedure described in section \ref{sub:isis_authorization_client_authorization}. There
-may be some pieces of information available only for specific clients. An example could be 
-information about resources serving particular Virtual Organization (VO) could be visible only to
+Because all ISIS services trust each other they freely exchange collected information. But 
+not all information stored inside the ISIS cloud is public or readable by clients authorized according 
+to the procedure described in section \ref{sub:isis_authorization_client_authorization}. There
+may be some pieces of information available only for specific clients. An example could be that
+information about resources serving a particular Virtual Organization (VO) might be visible only to
 members of that VO.
 
-To implement functionality described above each node in aggregated XML documents of information
-collected by ISIS cloud may have Access Control Policy associated with it. Access control is 
-defined at level of XML node and propagates all children nodes - similar to file systems.
-Children nodes can only additionally restrict access control imposed by parent node. For example
+To implement the functionality described above each node in the aggregated XML documents of information
+collected by the ISIS cloud may have an Access Control Policy associated with it. Access control is 
+defined at the level of an XML node and propagates to all children nodes --- similar to file systems.
+Children nodes can only additionally restrict access control imposed by the parent node. For example
 if parent node A allows access only to VO1 then children node B can narrow access to Administrator 
@@
 By default all XML nodes are public. Access Control Policies are embedded into XML document
-as XML nodes (see Appendix \ref{annex:info_policies_schema} for schema) even if that violates schema 
-of document. Then nodes are assigned policies by adding XML attributes referring to defined Policies.
-Access permission to particular information node is evaluated by traversing all nodes prom 
-parent to children. At first node which gives negative result evaluation is stopped and
+as XML nodes (see Appendix \ref{annex:info_policies_schema} for schema) even if that violates the schema 
+of the document. Then nodes are assigned policies by adding XML attributes referring to defined Policies.
+Access permission to a particular information node is evaluated by traversing all nodes from 
+parent to children. At the first node that gives a negative result the evaluation is stopped and
 this node including all its children is removed from document.
 
-Before providing results of query operation ISIS runs procedure described above on results and 
-also removes Access Control Policies. Reduced document obtained in this way is returned to
+Before providing results of a query operation ISIS runs the procedure described above on the results and 
+also removes Access Control Policies. The reduced document obtained in this way is returned to the
 requesting client.
@@
 % b. available to trusted clients (like broker or scheduler service)
-% c. private - available to owner of information object and entities
+% c. private --- available to owner of information object and entities
 %trusted by owner.
@@
 %into:
-% a. local - stored at endpoint representing resource which is
+% a. local --- stored at endpoint representing resource which is
 %described by information. This information is obtainable through
@@
 %
-% Information is structured in tree-like fashion - XML document.
-%Access control is defined at node and follows all branches - similar
+% Information is structured in tree-like fashion --- XML document.
+%Access control is defined at node and follows all branches --- similar
 %to file systems. Propagation control is similar to access control (if
@@
 % To reduce access control evaluation cost ACRs may/should/must(?)
-%be assigned to information nodes in one-to-many way - one ACR per
+%be assigned to information nodes in one-to-many way --- one ACR per
 %multiple information nodes. This grouping may be initiated either
@@
 
-To restrict set of clients allowed to perform operations in ISIS service proper authorization policy is needed.
-Let's assume ISIS is operating over TLS connection and all participants posses X.509 certificates with following subject names:
+To restrict the set of clients allowed to perform operations on the ISIS service a proper authorization policy is needed.
+Let's assume ISIS is operating over a TLS connection and all participants possess X.509 certificates with the following subject names:
 \begin{itemize}
-\item /O=Grid/O=Test/CN=CLIENT - generic client entities
-\item /O=Grid/O=Test/CN=SERVICE - generic service entities
-\item /O=Grid/O=Test/CN=ISIS - all ISIS belonging to ISIS cloud
+\item /O=Grid/O=Test/CN=CLIENT --- generic client entities
+\item /O=Grid/O=Test/CN=SERVICE --- generic service entities
+\item /O=Grid/O=Test/CN=ISIS --- all ISIS belonging to ISIS cloud
 \end{itemize}
 
-\textit{In NO way we suggest to use such setup. Real installation should use more sophisticated way to 
-identify clients contacting ISIS service. For an infrastructure with quite static roles distribution for 
+\textit{In NO way do we suggest to use such setup. A real installation should use more sophisticated ways to 
+identify clients contacting the ISIS service. For an infrastructure with a quite static roles distribution for 
 example we suggest to use Virtual Organization Management Service (VOMS) attributes embedded into X.509 
-certificates representing participating entities.}
-
-Below is an example policy made of 3 rules defined in lines 4-17, 18-29 and 30-39. Those define allowed 
-behavior for ISIS, generic service and generic client. Lines 6-7, 20-21 and 32-33 specify attributes 
-used to recognise type of connecting client. In this case those are subjects of X.509 certificates with 
-values defined above. Lines 10-15, 24-27 and 36-37 specify allowed actions. One can see that this policy 
-allows all operations to be performed by ISIS client. It limits operations allowed for generic service 
+certificates representing the participating entities.}
+
+Below is an example policy made of 3 rules defined in lines 4--17, 18--29 and 30--39. Those define the allowed 
+behavior for the ISIS, generic services and generic clients. Lines 6--7, 20--21 and 32--33 specify attributes 
+used to recognise the type of the connecting client. In this case those are subjects of X.509 certificates with 
+values defined above. Lines 10--15, 24--27 and 36--37 specify allowed actions. One can see that this policy 
+allows all operations to be performed by the ISIS client. It limits operations allowed for generic service 
 to "service" and "client" types. And client entities are allowed to perform "client" type operations only.
@@
 
-This policy is not as restrictive as it could be in order to allow ISIS register themselves as ordinary 
+This policy is not as restrictive as it could be in order to allow ISIS services to register themselves as ordinary 
 services and to allow services which perform discovery of other services to behave like ordinary clients.
 
-In order to activate policy it must be linked to ISIS service through Security Handle and Policy 
-Decision Point emedded into configuration of ISIS. Below is minimal example configuration element of 
-ISIS service. In lines 7-17 a set of Security Handler and Policy Decision Point is configured to handle
-policies written in ARC Policy Language (described in ~\cite{arcsec}). Line 14 specifies that policy is 
-read from file \textit{/opt/arc/etc/isis/policy.xml} every time new request is arrived. If policy is not 
-satisfied then ISIS returns SOAP Fault instead of usual informative response.
+In order to activate the policy it must be linked to the ISIS service through the Security Handle and the Policy 
+Decision Point emedded into the configuration of ISIS. Below is a minimal example configuration element of 
+an ISIS service. In lines 7--17 a set of Security Handlers and Policy Decision Points is configured to handle
+policies written in the ARC Policy Language (described in ~\cite{arcsec}). Line 14 specifies that the policy is 
+read from file \textit{/opt/arc/etc/isis/policy.xml} every time a new request arrives. If the policy is not 
+satisfied then the ISIS returns SOAP Fault instead of the usual informative response.
 
@@
 The service developers have to ensure that the services are providing the necessary information about 
-themselves when registering to ISISes. This is done by implementing the subclass of the Arc::Service class - 
+themselves when registering to ISISes. This is done by implementing a subclass of the Arc::Service class ---
 the RegistrationCollector function has to provide up-to-date status information about the service and anything 
@@
 \label{service_advertisement}
-The \textbf{Service Advertisement} can contain any information service wants to advertise but the mandatory 
-elements have to be always present:
+The \textbf{Service Advertisement} can contain any information the service wants to advertise but the mandatory 
+elements have to always be present:
 \begin{itemize}
@@
 
-Because there may be multiple registration processes running in parallel it is important to ensure that 
-implementation of RegistrationCollector is thread safe or there are internal locks implemented.
+Because there may be multiple registration processes running in parallel it is important to ensure that the
+implementation of the RegistrationCollector is thread safe or there are internal locks implemented.
 
 Every service registering to ISIS should also provide an interface for direct querying of information 
-describing service. Normally this information should be more detailed than one sent to ISIS. For this 
-purpose LIDI interface is defined which is a subset of WS Resource Properties (WSRP)~\cite{wsrf-rp}. Following WSRP
-operations must be supported - \textit{GetResourcePropertDocument}, \textit{GetResourceProperty},
+describing the service. Normally this information should be more detailed than the one sent to ISIS. For this 
+purpose the LIDI interface is defined which is a subset of WS Resource Properties (WSRP)~\cite{wsrf-rp}. The following WSRP
+operations must be supported --- \textit{GetResourcePropertDocument}, \textit{GetResourceProperty},
 \textit{GetMultipleResourceProperties} and \textit{QueryResourceProperties}.
 
-%TODO: define names of mandatory properties - at least Glue2 property
+%TODO: define names of mandatory properties --- at least Glue2 property
 
@@
 
-The registration of service is carried out by internal module called Registrant.
-The Registrant is active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes.
+The registration of a service is carried out by an internal module called Registrant.
+The Registrant is an active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes.
 In practice, the configuration part of the Registrant contains exactly one ISIS to bind, and the Registrant will collect the necessary information about the other ISISes belonging to the same network.
@@
 To register services to more than one ISIS network multiple Registrant instances has to be configured.
-In this case, the default Registrant will be used for registering every services unless configured explicitly.
-The registration of service can be done once or periodically based either on the configuration of the Registrant or overwritten for every service separately.
+In this case, the default Registrant will be used for registering every service unless configured explicitly.
+The registration of a service can be done once or periodically based either on the configuration of the Registrant or overwritten for every service separately.
 The Registrant is also performing message aggregation of all services linked to it if possible.
@@
 \begin{framed}
-  Registrant - pseudo algorithm\\
+  Registrant --- pseudo algorithm\\
   \\
@@
 \end{framed}
-Current implementation does not allow value of the period to be less than 2 minutes.
-
-Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{sub:isis_interface}).
+The current implementation does not allow the value of the period to be less than 2 minutes.
+
+The service provides the \textbf{Service Advertisement} part of the information sent to ISIS (see Section \ref{sub:isis_interface}).
 Before sending this information the Registrant extends it with additional data (\textbf{Service Advertisement Metadata}).
@@
 
-So the Service 1 and 2 are handled both by \textit{Registrant A} and Service 3 by \textit{Registrant B}.
+So Service 1 and 2 are both handled by \textit{Registrant A} and Service 3 by \textit{Registrant B}.
 In the first step each Registrant performs information collection from all assigned services sequentially.
@@
 
-For each information provided by service through the LIDI interface same procedure as described in 
+For every piece of information provided by the service through the LIDI interface the same procedure as described in 
 Section \ref{sub:isis_authorization_information_authorization} should be applied.
@@
 The registration operation is done by the Registrants as described in Section \ref{sec:service_registration}.
-Each Registrant is instantiated by corresponding InfoRegistrar XML element in configuration file as 
-defined in Section \ref{annex:service_configuration_schema}. First InfoRegistrar defines Registrant 
-which is used to register every service by default.
-
-Registration parameters per service are defined by InfoRegister element located inside corresponding 
-service configuration element. To assign service to specific Registrant(s) one may use Registrar 
-configuration elements situated inside InfoRegister. If none Registrar element is defined service will 
-be registered using first (default) Registrant. To make not register itself special configuration element 
+Each Registrant is instantiated by the corresponding InfoRegistrar XML element in the configuration file as 
+defined in Section \ref{annex:service_configuration_schema}. The first InfoRegistrar defines the Registrant 
+that is used to register every service by default.
+
+Registration parameters per service are defined by the InfoRegister element located inside the corresponding 
+service configuration element. To assign a service to specific Registrant(s) one may use Registrar 
+configuration elements situated inside InfoRegister. If no Registrar element is defined the service will 
+be registered using first (default) Registrant. To make a service not register itself the special configuration element 
 NoRegister has to be used.
 
-Following elements are defined inside configuration element of the Registrant:
+The Following elements are defined inside the configuration element of the Registrant:
 \begin{description}
-\item{URL} specifies contact endpoint of a bootstrap ISIS. If neede further ISIS addresses will be 
+\item{URL} specifies the contact endpoint of a bootstrap ISIS. If needed further ISIS addresses will be 
 queried from this service. This element is mandatory.
@@
 TLS communication is used.
-\item{Retry} specifies how many times communication to ISIS have to fail/timeout to start treating is
+\item{Retry} specifies how many times communication to ISIS have to fail/timeout to start treating it
 as unavailable. The default value is 5.

### 2009-06-14 doc/tech_doc/infosys/infosys_technical.tex
Temporary changes for the internal review

@@
 \usepackage{longtable}                             %for breaking tables
+\usepackage{soul}				       %package for highlight text
 %\usepackage{trackchanges}
@@
 
-\vspace*{1.5cm}
+%%%%\vspace*{1.5cm}
 %%%%---- A list of authors ----
-    {\centering \large \footnote{@} \large \par}
+%%%%    {\centering \large \footnote{@} \large \par}
 \end{titlepage}
@@
 We are keeping the databases in synch by passing every Registration and RemoveRegistrations to every node acting in the network.
-This yields redundancy and the fault tolerant message routing is very important between any two ISIS nodes.
+This yields redundancy and fault tolerant message routing between any two ISIS nodes.
 
 The most important parameter of the network is the {\it sparsity}.
-This is an integer not less then 2.
-It determines the number of neighbors in the function of the actual number of member nodes of the network, where the number of neighbors can change when new members connecting to the network or members leave.
-A sparse graph can be shaped by choosing a great value for sparsity or a dense one by choosing the less.
-This value is defined during the design and determines the main properties of the peer-to-peer network.
+This is an integer not less than 2.
+It determines the number of neighbors as a function of the actual number of member nodes of the network, where the number of neighbors can change when new members connect to the network or old members leave.
+A sparse graph can be shaped by choosing a great value for sparsity or a dense one by choosing a smaller one.
+This value is defined during the service configuration and determines the main properties of the peer-to-peer network.
 \begin{figure}[ht]
@@
 
-The registration of service is carried out by internal module called Registrant.
-The Registrant is active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes.
-In practice, the configuration part of the Registrant contains exactly one ISIS to bind, and the Registrant will collect the necessary information about the other ISISes belonging to the same network.
-
-To register services to more than one ISIS network multiple Registrant instances has to be configured.
-In this case, the default Registrant will be used for registering every services unless configured explicitly.
-The registration of service can be done once or periodically based either on the configuration of the Registrant or overwritten for every service separately.
-The Registrant is also performing message aggregation of all services linked to it if possible.
-The simplified algorithm of the Registrant is presented below.
+The registration of a service is carried out by an internal module called \hl{Registrant}.
+The \hl{Registrant} is an active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes.
+In practice, the configuration part of the \hl{Registrant} contains exactly one ISIS to bind, and the \hl{Registrant} will collect the necessary information about the other ISISes belonging to the same network.
+
+To register services to more than one ISIS network multiple \hl{Registrant} instances has to be configured.
+In this case, the default \hl{Registrant} will be used for registering every service unless configured explicitly.
+The registration of a service can be done once or periodically based either on the configuration of the \hl{Registrant} or overwritten for every service separately.
+The \hl{Registrant} is also performing message aggregation of all services linked to it if possible.
+The simplified algorithm of the \hl{Registrant} is presented below.
 
 \begin{framed}
-  Registrant - pseudo algorithm\\
+  \hl{Registrant} --- pseudo algorithm\\
   \\
@@
 \end{framed}
-Current implementation does not allow value of the period to be less than 2 minutes.
-
-Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{sub:isis_interface}).
-Before sending this information the Registrant extends it with additional data (\textbf{Service Advertisement Metadata}).
+The current implementation does not allow the value of the period to be less than 2 minutes.
+
+The service provides the \textbf{Service Advertisement} part of the information sent to ISIS (see Section \ref{sub:isis_interface}).
+Before sending this information the \hl{Registrant} extends it with additional data (\textbf{Service Advertisement Metadata}).
 
 An example layout of services is shown on Figure \ref{fig:registration_overview}. In this configuration example 
-there are two Registrants configured in one HED container for three services. The \textit{Registrant A} is the 
-default Registrant and the services configured in the following way:
+there are two \hl{Registrant}s configured in one HED container for three services. The \textit{\hl{Registrant} A} is the 
+default \hl{Registrant} and the services configured in the following way:
 
 \begin{itemize}
-  \item Service 1: There is no Registrant configured so the default one will register it.
-  \item Service 2: The \textit{Registrant A} is configured explicitly.
-  \item Service 3: The \textit{Registrant B} is configured explicitly.
+  \item Service 1: There is no \hl{Registrant} configured so the default one will register it.
+  \item Service 2: The \textit{\hl{Registrant} A} is configured explicitly.
+  \item Service 3: The \textit{\hl{Registrant} B} is configured explicitly.
 \end{itemize}
 
-So the Service 1 and 2 are handled both by \textit{Registrant A} and Service 3 by \textit{Registrant B}.
-In the first step each Registrant performs information collection from all assigned services sequentially.
-Registrants themselves are executing in parallel.
+So Service 1 and 2 are both handled by \textit{\hl{Registrant} A} and Service 3 by \textit{\hl{Registrant} B}.
+In the first step each \hl{Registrant} performs information collection from all assigned services sequentially.
+\hl{Registrant}s themselves are executing in parallel.
 For Figure \ref{fig:registration_overview} the approximate sequence of information collection is (\textit{A1}-\textit{A2}, \textit{B1}-\textit{B2}) followed by (\textit{A3}, \textit{A4}).
-After collection  the aggregated information is sent to the corresponding ISISes independently by each other Registrant (\textit{A5}, \textit{B5}).
+After collection  the aggregated information is sent to the corresponding ISISes independently by each other \hl{Registrant} (\textit{A5}, \textit{B5}).
 
@@
 
-The registration operation is done by the Registrants as described in Section \ref{sec:service_registration}.
-Each Registrant is instantiated by corresponding InfoRegistrar XML element in configuration file as 
-defined in Section \ref{annex:service_configuration_schema}. First InfoRegistrar defines Registrant 
-which is used to register every service by default.
-
-Registration parameters per service are defined by InfoRegister element located inside corresponding 
-service configuration element. To assign service to specific Registrant(s) one may use Registrar 
-configuration elements situated inside InfoRegister. If none Registrar element is defined service will 
-be registered using first (default) Registrant. To make not register itself special configuration element 
+The registration operation is done by the \hl{Registrant}s as described in Section \ref{sec:service_registration}.
+Each \hl{Registrant} is instantiated by the corresponding InfoRegistrar XML element in the configuration file as 
+defined in Section \ref{annex:service_configuration_schema}. The first InfoRegistrar defines the \hl{Registrant} 
+that is used to register every service by default.
+
+Registration parameters per service are defined by the InfoRegister element located inside the corresponding 
+service configuration element. To assign a service to specific \hl{Registrant}(s) one may use Registrar 
+configuration elements situated inside InfoRegister. If no Registrar element is defined the service will 
+be registered using first (default) \hl{Registrant}. To make a service not register itself the special configuration element 
 NoRegister has to be used.
 
-Following elements are defined inside configuration element of the Registrant:
+The Following elements are defined inside the configuration element of the \hl{Registrant}:
 \begin{description}
-\item{URL} specifies contact endpoint of a bootstrap ISIS. If neede further ISIS addresses will be 
+\item{URL} specifies the contact endpoint of a bootstrap ISIS. If needed further ISIS addresses will be 
 queried from this service. This element is mandatory.
@@
              document. Each instance if such element instatiates one instance
-             of Information Registrant active element. -->
+             of Information \hl{Registrant} active element. -->
         <xsd:sequence>

### 2009-06-16 doc/tech_doc/infosys/infosys_technical.tex
Final changes for the internal review

@@
 %%%%---- The title ----
-{\centering \textsc{\Large ARC Information System}\Large \par}
+{\centering \textsc{\Large ARC peer-to-peer information system}\Large \par}
 \vspace*{0.5cm}
@@
 
-The registration of a service is carried out by an internal module called \hl{Registrant}.
-The \hl{Registrant} is an active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes.
-In practice, the configuration part of the \hl{Registrant} contains exactly one ISIS to bind, and the \hl{Registrant} will collect the necessary information about the other ISISes belonging to the same network.
-
-To register services to more than one ISIS network multiple \hl{Registrant} instances has to be configured.
-In this case, the default \hl{Registrant} will be used for registering every service unless configured explicitly.
-The registration of a service can be done once or periodically based either on the configuration of the \hl{Registrant} or overwritten for every service separately.
-The \hl{Registrant} is also performing message aggregation of all services linked to it if possible.
-The simplified algorithm of the \hl{Registrant} is presented below.
+The registration of a service is carried out by an internal module called InfoRegistrar.
+The InfoRegistrar is an active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes.
+In practice, the configuration part of the InfoRegistrar contains exactly one ISIS to bind, and the InfoRegistrar will collect the necessary information about the other ISISes belonging to the same network.
+
+To register services to more than one ISIS network multiple InfoRegistrar instances has to be configured.
+In this case, every InfoRegistrar will be used parallel for registering every service unless configured explicitly.
+The registration of a service can be done once or periodically based either on the configuration of the service or overwritten for every InfoRegistrar per service separately.
+The InfoRegistrar is also performing message aggregation of all services linked to it if possible.
+The simplified algorithm of the InfoRegistrar is presented below.
 
 \begin{framed}
-  \hl{Registrant} --- pseudo algorithm\\
+  InfoRegistrar --- simplified pseudo algorithm\\
   \\
@@
 The service provides the \textbf{Service Advertisement} part of the information sent to ISIS (see Section \ref{sub:isis_interface}).
-Before sending this information the \hl{Registrant} extends it with additional data (\textbf{Service Advertisement Metadata}).
+Before sending this information the InfoRegistrar extends it with additional data (\textbf{Service Advertisement Metadata}).
 
 An example layout of services is shown on Figure \ref{fig:registration_overview}. In this configuration example 
-there are two \hl{Registrant}s configured in one HED container for three services. The \textit{\hl{Registrant} A} is the 
-default \hl{Registrant} and the services configured in the following way:
+there are two InfoRegistrars configured in one HED container for three services. The services are configured in the following way:
 
 \begin{itemize}
-  \item Service 1: There is no \hl{Registrant} configured so the default one will register it.
-  \item Service 2: The \textit{\hl{Registrant} A} is configured explicitly.
-  \item Service 3: The \textit{\hl{Registrant} B} is configured explicitly.
+  \item Service 1: The \textit{InfoRegistrar A} is configured explicitly.
+  \item Service 2: There is no InfoRegistrar configured so all existing InfoRegistrar will register it parallel.
+  \item Service 3: The \textit{InfoRegistrar B} is configured explicitly.
 \end{itemize}
 
-So Service 1 and 2 are both handled by \textit{\hl{Registrant} A} and Service 3 by \textit{\hl{Registrant} B}.
-In the first step each \hl{Registrant} performs information collection from all assigned services sequentially.
-\hl{Registrant}s themselves are executing in parallel.
-For Figure \ref{fig:registration_overview} the approximate sequence of information collection is (\textit{A1}-\textit{A2}, \textit{B1}-\textit{B2}) followed by (\textit{A3}, \textit{A4}).
-After collection  the aggregated information is sent to the corresponding ISISes independently by each other \hl{Registrant} (\textit{A5}, \textit{B5}).
+So Service 1 and 2 are both handled by \textit{InfoRegistrar A} since Service 2 and Service 3 by \textit{InfoRegistrar B}.
+In the first step each InfoRegistrar performs information collection from all assigned services sequentially.
+InfoRegistrars themselves are executing in parallel.
+After collection  the aggregated information is sent to the corresponding ISISes independently by each other InfoRegistrar.
 
@@
 
-The registration operation is done by the \hl{Registrant}s as described in Section \ref{sec:service_registration}.
-Each \hl{Registrant} is instantiated by the corresponding InfoRegistrar XML element in the configuration file as 
-defined in Section \ref{annex:service_configuration_schema}. The first InfoRegistrar defines the \hl{Registrant} 
-that is used to register every service by default.
-
-Registration parameters per service are defined by the InfoRegister element located inside the corresponding 
-service configuration element. To assign a service to specific \hl{Registrant}(s) one may use Registrar 
+The registration operation is done by the InfoRegistrars as described in Section \ref{sec:service_registration}.
+Each InfoRegistrar is instantiated by the corresponding InfoRegistrar XML element in the configuration file as 
+defined in Section \ref{annex:service_configuration_schema}.
+
+The default registration parameters per service are defined by the InfoRegister elements located inside the corresponding 
+service configuration element. To assign a service to specific InfoRegistrar(s) one may use Registrar 
 configuration elements situated inside InfoRegister. If no Registrar element is defined the service will 
-be registered using first (default) \hl{Registrant}. To make a service not register itself the special configuration element 
+be registered using every InfoRegistrars. To make a service not register itself the special configuration element 
 NoRegister has to be used.
 
-The Following elements are defined inside the configuration element of the \hl{Registrant}:
+The Following elements are defined inside the configuration element of the InfoRegistrar:
 \begin{description}

### 2009-07-06 doc/tech_doc/sec/SecurityFrameworkofARC1.tex
move sec doc into tech_doc directory

--- 2009-01-19 doc/sec/SecurityFrameworkofARC1.tex
@@
-% This file was converted to LaTeX by Writer2LaTeX ver. 0.9.3
+% This file was converted to LaTeX by Writer2LaTeX ver. 0.5
 % see http://www.hj-gym.dk/~hj/writer2latex for more info
-\documentclass[a4paper]{article}
+\documentclass{article}
 \usepackage[ascii]{inputenc}
@@
 \usepackage[spanish,english]{babel}
-\usepackage{amsmath}
-\usepackage{amssymb,amsfonts,textcomp}
+\usepackage{amsmath,amssymb,amsfonts,textcomp}
 \usepackage{color}
@@
 \usepackage{hyperref}
-\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=Aleksandr Konstantinov, pdfsubject=, pdfkeywords=}
+\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=qiangwz, pdfsubject=, pdfkeywords=}
 \usepackage[pdftex]{graphicx}
@@
 \setcounter{secnumdepth}{4}
-\renewcommand\thesection{\arabic{section}}
-\renewcommand\thesubsection{\arabic{subsection}}
-\renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection}}
-\renewcommand\theparagraph{\arabic{subsection}.\arabic{subsubsection}.\arabic{paragraph}}
+\renewcommand\thesection{\arabic{section}.}
+\renewcommand\thesubsection{\arabic{subsection}.}
+\renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection}.}
+\renewcommand\theparagraph{\arabic{subsection}.\arabic{subsubsection}.\arabic{paragraph}.}
 \makeatletter
@@
 }
-\newcommand\liststyleLii{%
+\newcommand\liststyleLi{%
 \renewcommand\theenumi{\arabic{enumi}}
@@
 % Page layout (geometry)
+\setlength\paperwidth{8.2673in}
+\setlength\paperheight{11.6925in}
 \setlength\voffset{-1in}
@@
 \title{}
-\author{Aleksandr Konstantinov}
-\date{2009-01-19}
 \begin{document}
@@
 \bigskip
-\clearpage\clearpage\pagestyle{Convertv}
+\clearpage\clearpage\setcounter{page}{1}\pagestyle{Convertv}
 
@@
 \textit{delegation.collector}, and evaluates the request against
-delegation policy. See section \ref{bkm:Ref204009393}for the design
-idea and use case of delegation policy in fine-grained identity
-delegation.}
+delegation policy. See section 6for the design idea and use case of
+delegation policy in fine-grained identity delegation.}
 
@@
 {\selectlanguage{english}\itshape\color{black}
-\ \ Rule (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ Subjects (1)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ Subject (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ \ \ Attribute (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ Resources (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ Resource (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ Actions (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ Action (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ Conditions (0-1)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ Condition (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ \ \ Attribute (1-)}
+ Rule (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+  Subjects (1)}
+
+{\selectlanguage{english}\itshape\color{black}
+   Subject (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+    Attribute (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+   Resources (0-1)}
+
+{\selectlanguage{english}\itshape\color{black}
+   Resource (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+  Actions (0-1)}
+
+{\selectlanguage{english}\itshape\color{black}
+   Action (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+  Conditions (0-1)}
+
+{\selectlanguage{english}\itshape\color{black}
+   Condition (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+    Attribute (1-)}
 
@@
 {\selectlanguage{english}\itshape\color{black}
-\ \ RequestItem (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ Subject (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ SubjectAttribute (1-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ Resource (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ Action (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ Context (0-)}
-
-{\selectlanguage{english}\itshape\color{black}
-\ \ \ \ \ \ ContextAttribute (1-)}
+ RequestItem (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+  Subject (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+   SubjectAttribute (1-)}
+
+{\selectlanguage{english}\itshape\color{black}
+  Resource (0-)}
+
+{\selectlanguage{english}\itshape\color{black}
+  Action (0-)}
+
+{\selectlanguage{english}\itshape\color{black}
+  Context (0-)}
+
+{\selectlanguage{english}\itshape\color{black}
+   ContextAttribute (1-)}
 
@@
 method. Currently only supported format is ARC Policy/Request XML
-document described in Section \ref{bkm:Ref204009564}and
-\ref{bkm:Ref204009595}}
+document described in Section 7.1and 7.2}
 
@@
 into existing MCCs or implemented as separate SecHandler plugins. See
-Section \ref{bkm:Ref204009636}for available Collectors and
-corresponding Security Attributes. }
+Section 5.2for available Collectors and corresponding Security
+Attributes. }
 
@@
 \item {\selectlanguage{english}\color{black}
-Delegation PDP is described below in Section \ref{bkm:Ref204009695}}
+Delegation PDP is described below in Section 6.3}
 \end{itemize}
@@
 described except those used for Delegation Restrictions. Those are
-described in Section \ref{bkm:Ref204009734}}
+described in Section 6.2}
 
@@
 {\selectlanguage{english}\color{black}
-Each set of attributes is acompanied by identifier of service which
+Each set of attributes is accompanied by identifier of service which
 provided those attributes. It is made of \textit{voname} element with
@@
 /voname=knowarc.eu/hostname=arthur.hep.lu.se:15001/knowarc.eu:UniqueKnowarcAttribute=False}
+
+{\selectlanguage{english}\color{black}
+Which is needed to describe is the configuration of trusted certificates
+to constrain the Attribute Certificate (AC) which the service would
+accept. The trusted certificates should be organized in a chunk which
+contains a suit of DN of trusted certificate in a chain
+(\textit{{\textless}VOMSCertTrustDNChain{\textgreater}}), in which the
+first line is the DN of the VOMS server certificate (cert0) which is
+used to sign the Attribute Certificate (AC), the second line is the DN
+of the corresponding issuer certificate(cert1) which is used to sign
+cert0. So if there are one or more intermediate issuers, then there
+should be 3 or more than 3 lines of DN in this chunk (considering cert0
+and the root CA certificate, plus the intermediate certificate).
+Regular expression is supported as well for match-making, in this case,
+only one line of regular expression is needed under one chain. Also,
+the trusted certificate chain can also be acquired from an independent
+text file including the same format of content.}
+
+{\selectlanguage{english}\color{black}
+The chain (\textit{{\textless}VOMSCertTrustDNChain{\textgreater}})
+should be configured under the configuration of TLS MCC. There also
+could be a few chains, if there are a few VOMS servers which are
+trusted by this service container.}
+
+\begin{center}
+\tablehead{}\begin{supertabular}{|m{5.6886597in}|}
+\hline
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChain{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDN{\textgreater}/O=Grid/O=NorduGrid/CN=host/arthur.hep.lu.se{\textless}/tls:VOMSCertTrustDN{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDN{\textgreater}/O=Grid/O=NorduGrid/CN=NorduGrid
+Certification Authority{\textless}/tls:VOMSCertTrustDN{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/tls:VOMSCertTrustDNChain{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChain{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDN{\textgreater}/DC=ch/DC=cern/OU=computers/CN=voms.cern.ch{\textless}/tls:VOMSCertTrustDN{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDN{\textgreater}/DC=ch/DC=cern/CN=CERN
+Trusted Certification
+Authority{\textless}/tls:VOMSCertTrustDN{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/tls:VOMSCertTrustDNChain{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChain{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustRegex{\textgreater}\^{}/O=Grid/O=NorduGrid{\textless}/tls:VOMSCertTrustRegex{\textgreater}}
+
+{\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}/tls:VOMSCertTrustDNChain{\textgreater}}
+
+\selectlanguage{english}\itshape\color{black}
+\ \ \ \ \ \ \ \ \ \ \ \ {\textless}tls:VOMSCertTrustDNChainsLocation{\textgreater}./{\textless}/tls:VOMSCertTrustDNChainsLocation{\textgreater}\\\hline
+\end{supertabular}
+\end{center}
+
+\bigskip
 
@@
 behalf provides Proxy Certificate to the service using Web Service
-based Delegation interface described in Section \ref{bkm:Ref204009766}}
+based Delegation interface described in Section 6.4}
 
@@
 Each deployment implementing Delegation Restrictions must use dedicated
-Security Handler plugin (see section \ref{bkm:Ref204009818}) to collect
-all Policy documents from Proxy Certificates used for establishing
-secure connection. Then those documents must be processed by dedicated
-Policy Decision Point plugin (see section \ref{bkm:Ref204009855}) to
-make a final decision based on collected Policies and various
-information about client{\textquotesingle}s identity and requested
-operation. Service or MCC chain supporting Delegation Restrictions must
-accept negative decision of this PDP as final and do not override it
-with any other decision based on other policies.}
+Security Handler plugin (see section 5.1) to collect all Policy
+documents from Proxy Certificates used for establishing secure
+connection. Then those documents must be processed by dedicated Policy
+Decision Point plugin (see section 2.3) to make a final decision based
+on collected Policies and various information about
+client{\textquotesingle}s identity and requested operation. Service or
+MCC chain supporting Delegation Restrictions must accept negative
+decision of this PDP as final and do not override it with any other
+decision based on other policies.}
 
@@
 stored inside X509 certificate proxy extension as defined in RFC3820
-and described in Section \ref{bkm:Ref204009914}All proxy certificates
-in a chain provided by client are examined and all available policies
-are extracted. }
+and described in Section 6.1All proxy certificates in a chain provided
+by client are examined and all available policies are extracted. }
 
@@
 Here is how delegation procedure works (also shown in figures
-\ref{seq:refFigure7} and \ref{seq:refFigure8}) :}
+\ref{seq:refFigure5} and \ref{seq:refFigure6}) :}
 
@@
 {\centering\selectlanguage{english}\itshape\color{black}
-Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure7}}. The
-flow diaram of delegation procedure with multiple second step and
+Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure5}}. The
+flow diagram of delegation procedure with multiple second step and
 session expiration. 
@@
 {\centering\selectlanguage{english}\itshape\color{black}
-Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure8}}. The
+Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure6}}. The
 flow diagram of delegation procedure with certificate transfered as
@@
 {\selectlanguage{english}\upshape\color{black}
-See Section \ref{bkm:Ref204009974}for the explanation of ARC Policy.}
+See Section 7.7for the explanation of ARC Policy.}
 
@@
 {\selectlanguage{english}\upshape\color{black}
-There is an example policy for echo service below. See Section
-\ref{bkm:Ref204009564}for the policy schema. The example policy is
-\ made of following elements: }
+There is an example policy for echo service below. See Section 7.1for
+the policy schema. The example policy is \ made of following elements:
+}
 
@@
 \end{enumerate}
-\subsubsection{PDP Service Invoker configuration}
+\subsubsection[PDP Service Invoker configuration]{PDP Service Invoker
+configuration}
 {\selectlanguage{english}\upshape\color{black}
@@
 Configuration XML schema with comments available at
-\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/x509tokensh/X509TokenSH.xsd}}.}
+\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/x509tokensh/X509TokenSH.xs}}\href{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/usernametokensh/UsernameTokenSH.xsd}{\textstyleInternetlink{d}}.}
 
@@
 
-\subsection[Using Shibboleth IdP for Authentication and
-Attribute{}-based Authorization]{Using
-\foreignlanguage{english}{Shibboleth} IdP for Authentication and
-Attribute-based Authorization}
+\subsection[Using Shibboleth IdP for Authentication and Attribute-based
+Authorization]{Using \foreignlanguage{english}{Shibboleth} IdP for
+Authentication and Attribute-based Authorization}
 {\selectlanguage{english}\color{black}
@@
 Figure
-{\refstepcounter{Figure}\theFigure\label{seq:refFigure5}}.\textbf{
+{\refstepcounter{Figure}\theFigure\label{seq:refFigure7}}.\textbf{
 }SAML2.0 SSO profile in ARC
@@
 {\selectlanguage{english}\color{black}
-\ The steps shows in Figure \ref{seq:refFigure5} are described as
+\ The steps shows in Figure \ref{seq:refFigure7} are described as
 follows:}
 
-\liststyleLii
+\liststyleLi
 \begin{enumerate}
@@
 
-\subsection[\ Short{}-Lived Credential
+\subsection[\ Short-Lived Credential
 Service]{\foreignlanguage{english}{\ }Short-Lived Credential Service}
@@
 WS client. The whole process of SLCS invocation is showed in Figure
-\ref{seq:refFigure6} (from step 1 to step 8), which is the same as in
-Figure \ref{seq:refFigure5}, except that step 7 and step 8 are invoked
+\ref{seq:refFigure8} (from step 1 to step 8), which is the same as in
+Figure \ref{seq:refFigure7}, except that step 7 and step 8 are invoked
 for the SLCS certificate request and response.}
@@
 {\centering\selectlanguage{english}\itshape\color{black}
-Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure6}}.Short
+Figure {\refstepcounter{Figure}\theFigure\label{seq:refFigure8}}.Short
 lived credential service
@@
 {\selectlanguage{english}\color{black}
-\stepcounter{Bibitem}{\theBibitem}.\foreignlanguage{english}{Web Service
-Security Username Token Profile 1.1.
-}\url{http://www.oasis-open.org/committees/download.php/16782/wss-v1.1-spec-os-UsernameTokenProfile.pdf}}
+\textstyleInternetlink{\textcolor{black}{\stepcounter{Bibitem}{\theBibitem}}}\textstyleInternetlink{\textcolor{black}{.}}\textstyleInternetlink{\foreignlanguage{english}{\textcolor{black}{Web
+Service Security Username Token Profile 1.1.
+}}}\url{http://www.oasis-open.org/committees/download.php/16782/wss-v1.1-spec-os-UsernameTokenProfile.pdf}}
 
@@
 {\selectlanguage{english}\color{black}
-\stepcounter{Bibitem}{\theBibitem}. OASIS Web Services Security (WSS)
-TC.
-\url{http://www.oasis-open.org/committees/tc_home.php?wg_abbrev=wss}}
+\textstyleInternetlink{\textcolor{black}{\stepcounter{Bibitem}{\theBibitem}}}\textstyleInternetlink{\textcolor{black}{.
+OASIS Web Services Security (WSS) TC.
+}}\url{http://www.oasis-open.org/committees/tc_home.php?wg_abbrev=wss}}
 \end{document}

### 2009-08-07 doc/tech_doc/sec/arc-security-documentation.tex
sec document: convert odt into tex, not finished yet, arc1 only

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {Weizhong Qiang},
  pdftitle = {Documentation of the ARC security framework},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-TECH-16\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\ SECURITY FRAMEWORK OF ARC1}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Documentation and developer's guide}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Weizhong Qiang\footnote{weizhong.qiang@fys.uio.no} \large \par}
    {\centering \large Aleksandr Konstantinov\footnote{aleksandr.konstantinov@fys.uio.no} \large \par}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}

\chapter{Technical Description} % (fold)
\label{cha:tech_description}

\section{Introduction} % (fold)
\label{sec:introduction}

The security framework of the ARC1 includes two parts of capabilities: security capability embedded in hosting environment, and security capability implemented as plug-ins with well-defined interfaces which can be accessed by hosting environment and applications. The following design concerns were employed when designing:

\begin{itemize}
    \item Interoperability and standardlization. In consistent with the main design concern of the ARC1, interoperability and standardization is considered in security framework. For example, in terms of authentication, PKI infrastructure and proxy certificate (RFC3820 [1]) is used as most of the other grid middle-wares do. Since supporting of standardization is a way for implementing interoperability, some standard specifications have been implemented as prototype and tested, such as SAML specification.

    \item Modularity and extensibility. Besides the security functionality which is embedded in hosting environment, the other security functionality is implemented as plug-ins which has well-defined interfaces, and is configurable and dynamically loadable. Since the interoperation interface between security plug-in and hosting environment or applications is predefined, it is easy to extend the security functionality in order to support some other security capability by implementing the interface.

    \item Backward compatibility. The GSI (Grid Security Infrastructure) based mechanism has been a de-facto solution for grid security. The design of security framework should be compatible to it.

\end{itemize}

%section introduction (end)

\section{Security architecture in HED: SecHandler and PDP} % (fold)
\label{sec:sec_architecture}

\subsection{Structure of SecHandler and PDP} % (fold)
\label{subsec:structure_sechandler}

In the implementation of the ARC1, there is a Service Container – the Hosting Environment Daemon (HED) (D1.2-2, [2]) which provides a hosting place for various services in application level, as well as a flexible and efficient communication mechanism.
HED contains a framework for implementing and enforcing authentication and authorization. Each Message Chain Component (MCC) or service has a common interface for implementing various authentication and authorization functionality. This functionality is implemented by using pluggable components (plug-ins) called SecHandler. The SecHandler components are C++ classes and provide method for processing messages traveling through Message Chains of the HED. Each MCC or Service  usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called ``incoming'' and ``outgoing'' respectively. It is possible for MCC or Service to implement other set of queues. Please check documentation of particular component for that particular information. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well.
Each SecHandler is configured inside same configuration file used for configuring whole chain of MCCs. Some of implemented SecHandler components also make use of pluggable and configurable sub-modules which specifically handle various security functionalities, such as authorization, authentication, etc. The currently implemented sub-modules used by some SecHandlers are Policy Decision Point (PDP) components such as Arc PDP which can process ARC specific Request and Policy documents. Figure 1 gives the structure of a MCC/Service, and the message sequence inside it. And Figure 2 shows the configuration of SecHandler components for an example ``Echo'' service.

\subsection{Interface of SecHandler} % (fold)
\label{subsec:interface_sechandler}

When one component (MCC or service) is loaded according to the configuration information, the SecHandler under the component and the plug-ins like PDP which are attached to the SecHandler will be loaded as well. 
There is one simple interface (see Figure 3) defined in class SecHandler, which will be called by the containing MCC/Service once there is message (incoming or outgoing) need to be processed.

Currently, the ARC1 comes with the following four security handler implemented:

\begin{itemize}
    \item arc.authz – Authorization SecHandler
The arc.authz and serves as container for he Policy Decision Point components. It is responsible for calling their interface and getting back the authorization result. Then obtained results are processed and   combined decision is made. Description of configuration and examples can be found in section 8.7. Usually the Authorization SecHandler and included PDPs are used on the service side of communication channel. Although it is also possible to use them on the client side.

    \item identity.map – Identity Mapping SecHandler
The identity.map is a specific authorization oriented security handler. It will map the global identity in the message into local identity like system username based on the result returned by Policy Decision Point components.

    \item delegation.collector – Delegation SecHandler
The delegation.collector is responsible for collecting the delegation policy information from the remote proxy credential (proxy certificate is compatible to RFC3820) inside the message, and putting this policy into message’s security attribute for the usage of other components, such as  delegation.pdp.

    \item usernametoken.handler – UseranemToken SecHandler
The task of the  usernametoken.handler is to generate the WS-Security Username-Token and add it into header of SOAP message which is the payload of outgoing message. It can also extract the WS-Security Username-Token from the header of SOAP message which is the payload of incoming message.

\end{itemize}

\subsection{Interface of PDP} % (fold)
\label{subsec:interface_pdp}

Figure 4 shows the definition of abstract class PDP. The implementation could be some function which implements the interface by composing the policy evaluation request, evaluating this request against some policy, and returning the evaluation result, or just by composing the policy evaluation request, invoking some remote policy decision web service and getting back the evaluation result.

Currently, the ARC1 comes with the following four policy decision point implementation:

\begin{itemize}
    \item arc.pdp – Arc PDP
The Arc PDP will organize the security attributes into the ARC specific authorization request, call the policy evaluator to evaluate the request against the policy (which is in ARC specific format) repository, and get back the evaluation result. See paragraph 3 for detail information about request schema and policy schema.

    \item delegation.pdp – Delegation PDP
The Delegation PDP is basically similar to Arc PDP, except it uses the delegation policy parsed from remote proxy credential by delegation.collector, and evaluates the request against delegation policy. See section 6for the design idea and use case of delegation policy in fine-grained identity delegation.

    \item simplelist.pdp – Simplelist PDP
The Simplelist PDP is a simplest implementation of policy decision point. It will match the identity extracted from the remote credential (or proxy credential) with local list of permitted identities.

    \item pdpservice.invoker – PDP Service Invoker
The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the same functionality as Arc PDP, except that the evaluation request and response are carried by SOAP message. The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation engine can be accessed remotely and maintained centrally.

\end{itemize}

%section sec_architecture (end)

\section{Policy Evaluation Engine} % (fold)
\label{sec:policy_eval}

\subsection{Design of policy evaluation engine} % (fold)
\label{subsec:design_policyengine}

The ARC1 defines specific evaluation request and policy schema. Based on the schema definition, one policy evaluation engine is implemented. The design principal of policy evaluation engine is generality by which the implementation of the policy evaluation engine can be easily extended to adopt some other policy schema, such as XACML policy schema.

Figure 5 shows the UML class diagram about the policy evaluation engine. It shows all classes and relations simultaneously for getting the overall picture. 
The Evaluator class is the key class for policy evaluation. It accepts request evaluates it against loaded policy and returns evaluation response.

Three abstract factories - FnFactory, AlgFactory, AttributeFactory - are responsible for creating the Function, CombiningAlg and AttributeValue objects correspondingly. The classes inherited from CombiningAlg class take care of implementing various combining algorithms which define relations between \textit{$<$Rule/$>$} elements in policy. The AttributeValue type of classes are used for processing different types of \textit{$<$Attribute/$>$} and similar elements. The Function classes take care of comparing \textit{$<$Attribute/$>$} elements of request and policy.

The Policy class parses \textit{$<$Policy/$>$} or \textit{$<$Rule/$>$} elements and creates  CombingAlg objects according to the \textit{$<$RuleCombiningAlg/$>$} attribute of \textit{$<$Policy/$>$}, Function objects according to the \textit{$<$Function/$>$} attribute of \textit{$<$Attribute/$>$} and AttributeValue objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}. Those objects will be used when evaluating the request.

The Request class is responsible for parsing \textit{$<$Request/$>$} element and creates corresponding AttributeValue objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}. When evaluating, each AttributeValue in request will be evaluated against corresponding AttributeValue in the policy by using relevant Function.

Due to extensible architecture of code it is relatively easy to add support for new types of  AttributeValue, Function and CombingAlg objects in this way supporting various types of XML based policy languages.

\subsection{Schemas for policy evaluation engine} % (fold)
\label{subsec:policy_schema}
The schema for ARC Policy is available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd .

The hierarchy tree of ARC Policy is show below (numbers show multiplicity of elements)

\begin{verbatim}
    Policy (1)
        Rule (1-)
            Subjects (1)
                Subject (1-)
                    Attribute (1-)
            Resources (0-1)
                Resource (1-)
            Actions (0-1)
                Action (1-)
            Conditions (0-1)
                Condition (1-)
                    Attribute (1-)
\end{verbatim}

The schema for ARC Request is available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd .

The hierarchy tree of ARC Request is show below (numbers show multiplicity of elements)

\begin{verbatim}
    Request (1)
        RequestItem (1-)
            Subject (1-)
	        SubjectAttribute (1-)
            Resource (0-)
            Action (0-)
            Context (0-)
                ContextAttribute (1-)
\end{verbatim}

The schema for ARC Response is available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Response.xsd .

The ARC Response is not used directly in code. It is in use by PDP Service which provides remote evaluation of policies.

\subsection{Basic Elements of Policy} % (fold)
\label{subsec:poicy_element}
There are 2 basic objects - "policy" and "request". There is 1 main actor - Evaluator. Curretly there are two types of elements in policy: Policy and Rule. Policy element is made of Rule elements.
Evaluator matches request to policy and produces one of 4 following results:

\begin{itemize}
    \item  \emph{PERMIT} - policy explicitely permits activity specified in request because request matches some part of policy and corresponding effect specified in policy is PERMIT.

        \emph{Example}:

            \emph{Rule}: PERMIT person ALICE to PLAY in place called WONDERLAND

            \emph{Request}: person ALICE wants to PLAY in place called WONDERLAND

    \item  \emph{DENY} - policy explicitely denies activity specified in Request because Request matches some part of policy and corresponding effect specified in policy is DENY.

        \emph{Example}:

            \emph{Rule}: DENY fruit APPLE to GROW on PEACH tree

            \emph{Request}: fruit APPLE to be GROWN on PEACH tree

    \item  \emph{INDETERMINATE} - request has some part which does not correspond to policy.

        \emph{Example}:

            \emph{Rule}: DENY fruit APPLE to GROW on PEACH tree

            \emph{Request}: fruit APPLE to be GROWN on WHEAT ground

            \emph{Request}: flower SUNFLOWER to be grown on PEACH tree

        \emph{Explaination}: Here, it is not possible to obtain any matching result - neither positive (DENY or PERMIT) nor negative (NOT\_APPLICABLE, see below)

In the request, the ``ground'' is completely uncomparable to the ``tree'' in policy. One can compare ``PEACH tree'' and ``APPLE tree'' because they are both ``tree''; But it is impossible to compare ``PEACH tree'' and ``WHEAT ground'' becaue they of different kind (Policy is about tree and Request is about ground).

In a similar way one can't compare ``fruit APPLE'' and ``flower SUNFLOWER'' (here policy is about fruits and Request is about flower).

Any other situation which makes it impossibile to compare two attributes will also cause ``INDETERMINATE''.

    \item  \emph{NOT\_APPLICABLE} - all parts of the Request have corresponding parts in the Policy, but some value of those parts are not the same. Hence request does not match policy.

        \emph{Example}:

            \emph{Rule}: DENY fruit APPLE to GROW on PEACH tree

            \emph{Request}: fruit APPLE to be GROWN on APPLE tree

            \emph{Request}: fruit ORANGE to be GROWN on PEACH tree

            \emph{Request}: fruit ORANGE to be GROWN on APPLE tree

        \emph{Explanation}: for each part of the Request evaluator can find relevant part in the Policy - both Policy and Request are about fruit and tree. But the values do not match.

\end{itemize}

If it is required to reduce evaluation results to boolean value PERMIT maps to TRUE and rest of results to FALSE.

\emph{Note}: It would be useful to make it possible to specify secondary effect which would become active in case Request is NOT\_APPLICABLE. For example:

DENY fruit APPLE to GROW on PEACH tree otherwise PERMIT

But one should be careful because example above would allow fruit PLUMS to grow on APPLE trees :)

This kind of requirement can be supported by using the algorithm between policies.  For example, in case of above scenario, we can use some algorithm like "Permit-if-notapplicable". See below the "Policy matching" part for more explaination.

\subsection{Policy Matching} % (fold)
\label{subsec:poicy_matching}
Policy is made of Rule elements. Request is evaluated against each Rule. Each evaluation produces same results as policy evaluation described above. The results from all Rules are then combined in order to produce final result for whole policy. Results Combining Algorithm is specified in Policy. There are 26 algorithms currently:

\begin{itemize}
    \item  \emph{Deny-Overrides} - this is default if no algorithm specified.

    \begin{itemize}
        \item If there is at least one DENY in results final result is DENY.
        \item Otherwise if there is at least one PERMIT, the final result is PERMIT.
        \item Otherwise if there is at least one NOT\_APPLICABLE final result is NOT\_APPLICABLE.
        \item Otherwise final result is INDETERMINATE.
    \end{itemize}

    Special case is Policy with no rules. Probably such policy should be treated as always producing DENY.

    \item  \emph{Permit-Overrides}

    \begin{itemize}
        \item If there is at least one PERMIT in results final result is PERMIT.
        \item Otherwise if there is at least one DENY the final result is DENY.
        \item Otherwise if there is at least one NOT\_APPLICABLE final result is NOT\_APPLICABLE.
        \item Otherwise final result is INDETERMINATE.
    \end{itemize}

    Special case is Policy wih no rules. Probably such policy should be treated as always producing DENY.

    \item  \emph{Ordered algorithms}

These specify priorities for all four possible results. Their names look like Result1-Result2-Result3-Result4 with Result\# naming result types, for example Permit-Deny-NotApplicable-Indeterminate. The results are combined in following way:

    \begin{itemize}
        \item If there is at least one result of Result1 type then final result is Result1.
        \item Otherwise if there is at least one result of Result2 type then final result is Result2.
        \item Otherwise if there is at least one result of Result3 type then final result is Result3.
        \item Otherwise final result is Result4.
    \end{itemize}

    There are 24 possible combinations of those algorithms.

\emph{Note}: It would be useful to have more combining algorithms. For example

    \item  \emph{Permit-if-notapplicable} - the use case could be ``DENY fruit APPLE to GROW on PEACH tree otherwise PERMIT''. In this case there is only one Rule under Policy, and this Rule is with ``Deny'' effect.

    \begin{itemize}
        \item If this Rule gives DENY in results, final result is DENY.
        \item Otherwise if this Rule gives NOT\_APPLICABLE, final result is PERMIT.
        \item Otherwise final result is INDETERMINATE. 
    \end{itemize}

    \item  \emph{Permit-if-allPermit} - Permit if all the Rules gives Permit, this algorithm is useful in case if we are collecting different policies from a few sources, and we want the request to satisfy all of them.

    \begin{itemize}
        \item If all of the Rule give PERMIT, the final result is PERMIT.
        \item Otherwise if there is at least one DENY the final result is DENY.
        \item Otherwise if there is at least one NOT\_APPLICABLE final result is NOT\_APPLICABLE.
        \item Otherwise final result is INDETERMINATE.
    \end{itemize}

    \item  \emph{OnlyOneApplicable}

    \begin{itemize}
        \item If there is one gives INDETERMINATE, final result INDETERMINATE is given immediately.
        \item Otherwise if there is exactly only one gives applicable result (DENY or PERMIT), final result is as this result.
        \item Otherwise if there is more than one gives applicable result, final result is INDTERMINATE.
        \item Otherwise final result is NOT\_APPLICABLE.
    \end{itemize}

This algorithm makes sure that only one Rule is selected when making decision.

    \item  \emph{FirstApplicable}

    \begin{itemize}
        \item If there is one give DENY, PERMIT or INDETERMINATE result, final result is given immediately as this result.
        \item Otherwise final result is NOT\_APPLICABLE.
    \end{itemize}

\end{itemize}

\subsection{Request Structure} % (fold)
\label{subsec:request_structure}

Request is made of RequestItem elements. Each RequestItem is evaluated against Policy Rule and for each evaluation separate result is generated as described above.RequestItem is made of 4 elements:
    \begin{itemize}
        \item \emph{Subject} - represents entity requesting specified action
        \item \emph{Resource} - destination/object of the action
        \item \emph{Action} - specifies what has to be done on resource
        \item \emph{Context} - for additional information which does not fit anywhere else, like the current time.
    \end{itemize}

Effectively RequestItem may have only one Subject, one Resource, one Action and one Context. If there are more than one element of any kind of sub-elelemt, then in the evaluator this RequestItem is split into several items containing all possible permutations and results are obtained for every item separately. How results are combined will be explained later.
Additionally Subject could contain sub-elements SubjectAttribute. Those are meant to represent different kinds of requesters' identities. Example:

    \begin{itemize}
        \item Subject
        \begin{itemize}     
            \item SubjectAttribute: name is ALICE
            \item SubjectAttribute: age is YOUNG
            \item SubjectAttribute: gender is GIRL
        \end{itemize}
    \end{itemize}
Context could also be made of ContextAttribute elements in the same way as Subject.

The following is an example of the Request:

\begin{verbatim}
    <Request xmlns="http://www.nordugrid.org/schemas/request-arc">
        <RequestItem>
            <Subject>
                <SubjectAttribute AttributeId="urn:knowarc:x509:identity">
                    /O=KnowARC/OU=UiO/CN=Physicist
                </SubjectAttribute>
                <SubjectAttribute AttributeId="urn:knowarc:voms:attribute>
                    knowarc:atlasuser
                </SubjectAttribute>
            </Subject>
            <Subject AttributeId="urn:knowarc:shibboleth:attribute">member</Subject>
            <Action AttributeId="urn:knowarc:fileoperation">Read</Action>
            <Resource AttributeId="urn:knowarc:fileidentity>file:///home/test</Resource>
            <Context AttributeId="urn:knowarc:time" Type="time">2008-09-15T20:30:20</Context>
        </RequestItem>
    </Request>
\end{verbatim}

While evaluating this RequestItem will be split into two RequestItems:

\begin{verbatim}
    <Request xmlns="http://www.nordugrid.org/schemas/request-arc">
        <RequestItem>
            <Subject>
                <SubjectAttribute AttributeId="urn:knowarc:x509:identity">
                    /O=KnowARC/OU=UiO/CN=Physicist
                </SubjectAttribute>
                <SubjectAttribute AttributeId="urn:knowarc:voms:attribute>
                    knowarc:atlasuser
                </SubjectAttribute>
            </Subject>
            <Action AttributeId="urn:knowarc:fileoperation">Read</Action>
            <Resource AttributeId="urn:knowarc:fileidentity>file:///home/test</Resource>
            <Context AttributeId="urn:knowarc:time" Type="time">2008-09-15T20:30:20</Context>
        </RequestItem>
        <RequestItem>
            <Subject AttributeId="urn:knowarc:shibboleth:attribute">member</Subject>
            <Action AttributeId="urn:knowarc:fileoperation">Read</Action>
            <Resource AttributeId="urn:knowarc:fileidentity>file:///home/test</Resource>
            <Context AttributeId="urn:knowarc:time" Type="time">2008-09-15T20:30:20</Context>
        </RequestItem>
    </Request>
\end{verbatim}

The following means this Subject possesses both of these Attributes.

\begin{verbatim}
    <Subject>
        <SubjectAttribute AttributeId="urn:knowarc:x509:identity">
            /O=KnowARC/OU=UiO/CN=Physicist
        </SubjectAttribute>
        <SubjectAttribute AttributeId="urn:knowarc:voms:attribute>
            knowarc:atlasuser
        </SubjectAttribute>
    </Subject>
\end{verbatim}

However, the following means two Subject each of which possesses one Attribute.

\begin{verbatim}
    <Subject AttributeId="urn:knowarc:x509:identity">
        /O=KnowARC/OU=UiO/CN=Physicist
    </Subject>
    <Subject AttributeId="urn:knowarc:voms:attribute>
        knowarc:atlasuser
    </Subject>
\end{verbatim}

The ``Type'' xml-attribute is for distinguishing how to process the xml-node value, which is critical when evaluate two value from request side and policy side because different type requires different evaluating/comparing approach. The default ``Type'' is ``string'', in this case (also with the ``Function'' xml-attribute on the policy side is ``equal'', which will be explained later), each letters of these two values will be compared one by one when evaluating them.

The ``AttributeId'' xml-attribute is for evaluator to find the Attribute with AttributeId from the request side which corresponds to the Attribute with the same AttributeId on the policy side. Only if two Attributes' AttributeId are equal, the evaluator will then compare the value.

Each RequestItem will be sequencialy and independently evaluated against policy/policies. So for one Request (including few RequestItems), some RequestItem could get positive evaluation result (PERMIT) from policy engine, others could get negative evaluation result (DENY, NOT\_APPLICABLE, INDETERMINATE).

It is up to policy decision point to make final decision according to the evaluation results returned by evaluator, and the evaluator itself can not give this kind of final decision.

Basically the policy decision point will feed policy engine with request, get back evaluation results, 
and make final decision.

\subsection{Rule Composition and Matching} % (fold)
\label{subsec:rule_comp_match}

Policy rule is made of 4 elements - Subjects, Resources, Actions, Conditions (See the following example). Those are only used to group multiple elements Subject, Resource, Action, Condition. For instance, you can merge two Rules with the same Resources, Actions, Conditions, and the same ``Effect'' but different Subjects into one Rule.

There is no logical relationship between Subject, which means you can split one Rule into two Subject (under Subjects) into two Rule (each of which has one Subject (under Subjects)).From now only later ones (Subjects with only one Subject as sub-element, and the same for others) are described. Their meaning is same as in request with Condition corresponding to Context. Subject and Condition elements are also made of Attributes. All elements may be present more than one time. During procedure of matching each element in RequestItem is matched against all elements of same kind in Policy - Subject is matched to Subject, Resource to Resource, etc. For every combination 3 possible results are produced:

    \begin{itemize}
        \item \emph{MATCHED} - element from RequestItem matched element in Policy Rule. 
            \emph{Example}:
            \emph{RequestItem Resource}: \textit{place called WONDERLAND}
            \emph{PolicyItem Resource}: \textit{place called WONDERLAND}
        \item \emph{NOT MATCHED} - element from RequestItem did not match element in Policy Rule. 
            \emph{Example}:
            \emph{RequestItem Resource}: \textit{place called WONDERLAND}
            \emph{PolicyItem Resource}: \textit{place called PLAYGROUND}
        \item \emph{INDETERMINATE} - element from RequestItem could not be compared to element in Policy  Rule because they are of incompatible ids/belong to different namespaces. 
            \emph{Example}:
            \emph{RequestItem Resource}: \textit{place called WONDERLAND (with namespace ``place'')}
            \emph{PolicyItem Resource}: \textit{LEMON tree (with namespace ``tree'')}
    \end{itemize}

The produced results then combined to produce final 4 types of results in following way:

    \begin{itemize}
        \item If for every element in RequestItem there is at least one MATCHED result then result for this Policy Rule is as specified in the corresponding Effect (Deny or Permit).
        \item Otherwise if for every element in RequestItem there is at least one gets INDETERMINATE result then result for Policy Rule is INDETERMINATE.
        \item Otherwise result is NOT\_APPLICABLE.
    \end{itemize}

Special case is then RequestItem does not have the element(s) of some kind (Subject, Action, Resource or Context/Condition). If there are elements of corresponding kind in the Policy Rule then such situation should be considered as INDETERMINATE.

The following is an example of the Policy:

\begin{verbatim}
<Policy xmlns="http://www.nordugrid.org/schemas/policy-arc" CombiningAlg="Permit-Overrides">
    <Rule Effect="Permit">
        <Subjects>
            <Subject>
                <Attribute AttributeId="urn:knowarc:x509:identity">
                    /O=KnowARC/OU=UiO/CN=Physicist
                </Attribute>
                <Attribute AttributeId="urn:knowarc:voms:attribute>
                    knowarc:atlasuser
                </Attribute>
            </Subject>
            <Subject AttributeId="urn:knowarc:shibboleth:attribute">member</Subject>
        </Subjects>
        <Actions>
            <Action AttributeId="urn:knowarc:fileoperation">Read</Action>
            <Action AttributeId="urn:knowarc:fileoperation">Delete</Action>   
        </Actions>
        <Resources>
            <Resource AttributeId="urn:knowarc:fileidentity">file:///home/test</Resource>
        </Resources>
        <Conditions>
            <Condition AttributeId="urn:knowarc:period" Type="period" Function="Inrange">
                2008-09-10T20:30:20/P1Y1M
            </Condition>
        </Conditions>
    </Rule>
</Policy>
\end{verbatim}

For the Subject which includes two Attributes in this example:

\begin{verbatim}
    <Subject>
        <Attribute AttributeId="urn:knowarc:x509:identity">
            /O=KnowARC/OU=UiO/CN=Physicist
        </Attribute>
        <Attribute AttributeId="urn:knowarc:voms:attribute>knowarc:atlasuser</Attribute>
    </Subject>
\end{verbatim}

These two attributes mean the Rule requires the request should possess both of these two attributes.

However, if You put these above two Attribute into two Subject elements:

\begin{verbatim}
    <Subject AttributeId="urn:knowarc:x509:identity">/O=KnowARC/OU=UiO/CN=Physicist</Subject>
    <Subject AttributeId="urn:knowarc:voms:attribute>knowarc:atlasuser</Subject>
\end{verbatim}

Then it means the Rule requires the request to possess at least one of these two attributes.

For the xml-attribute ``Type'' and ``AttributeId'', the explaination for Request example also applies here.
The ``Function'' xml-attribute is for distinguishing different comparison algorithm when comparing these two xml-node value. If Function is absent, ``equal'' will be used as default.

\subsection{Rule Elements Matching} % (fold)
\label{subsec:rule_element_match}
For elements without attributes those elements have:

    \begin{itemize}
        \item Kind specified by AttributeId XML attribute. There is no default.
        \item Matching algorithm specified by Id XML attribute. By default string-equal matching is used.
        \item Content

        \textit{Example}: LEMON tree

        \textbf{Kind}: tree

        \textbf{Matching algorithm}: default

        \textbf{Content}: LEMON

    \end{itemize}

Matching procedure consists of following steps:
    \begin{itemize}
        \item Kinds are compared using simple string equal matching. If those do not match then result is INDETERMINATE.
        \item Matching algorithm is used to compare content of elements. Result is either MATCH or NO\_MATCH according to matching algorithm.
    \end{itemize}

Each element on the RequestItem must satisfy corresponding element in Rule.In detail, for Subjects element under Rule, if there is at least one Subject (with one Attribute or a few Attribute) which is matched by a Subject on this RequestItem, we say this Subjects is matched by the RequestItem; and also the same for the other elements (Actions, Resources, Conditions).

For elements with multiple Attribute sub-elements the way to judging whether elements match is if and only if all of the Attribute under the Rule have matching Attributes at RequestItem side.

Example of the Subject with three Attributes:

    \textbf{Subject}:

    \textbf{SubjectAttribute}: name is ALICE

    \textbf{SubjectAttribute}: age is YOUNG

    \textbf{SubjectAttribute}: gender is GIRL

In XML that is:

\begin{verbatim}
    <Subject>
        <Attribute AttributeId="name">Alice</Attribute>
        <Attribute AttributeId="age>YOUNG</Attribute>
        <Attribute AttributeId="gender>GIRL</Attribute>
    </Subject>
\end{verbatim}

That requires the Subject in the RequestItem to possess at least these three Attributes.

\begin{verbatim}
    <RequestItem>
        <Subject>
            <Attribute AttributeId="name">Alice</Attribute>
            <Attribute AttributeId="age">YOUNG</Attribute>
            <Attribute AttributeId="gender">GIRL</Attribute>
        <!--Some other Attribute-->
        </Subject>
    </RequestItem>
\end{verbatim}

The above example shows that the Subject in the RequestItem ``MATCH'' one Subject on the Rule side.

If the Subject in the RequestItem is like this:

\begin{verbatim}
    <Subject>
        <Attribute AttributeId="name">Alice</Attribute>
        <Attribute AttributeId="age>YOUNG</Attribute>
        <Attribute AttributeId="from">OSLO</Attribute>
        <!--Some other Attribute, buts not a "gender"-->
    </Subject>
\end{verbatim}

Then evaluator will produce INDETERMINATE as the match-making result of this Subject.

If the Subject in the RequestItem is like this:

\begin{verbatim}
    <Subject>
      <Attribute AttributeId="name">Bob</Attribute>
      <Attribute AttributeId="age>YOUNG</Attribute>
      <Attribute AttributeId="gender">BOY</Attribute>
      <!--Some other Attribute-->
    </Subject>
\end{verbatim}

Then evaluator will give NO\_MATCH as the match-making result of this Subject.

Finally if and only if all of the elemens (Subjects, Actions, Resources, Conditions) which are not empty under the Rule have been matched (gets MATCH) to the RequestItem, then the whole Rule is considered to be matched (produces MATCH result). MATCH is then mapped to final evaluation result depending on the specified Effect. If Effect is set to Deny then DENY decision will be produced for this Rule; if Effect is Permit then PERMIT.

Otherwise if any of the element (Subjects, Actions, Resources, Conditions) of RequestItem got INDETERMINATE decision then the INDETERMINATE decision will be made for this Rule.

Otherwise the NOT\_APPLICABLE decision will be made for this Rule. In other words that means at least one of the elements of this Rule got NO\_MATCH and the other elements got MATCH.

\subsection{Interface for using the policy evaluation engine} % (fold)
\label{subsec:interface_policy_engine}

For making usage of policy evaluation engine more convenient basic Evaluator class is complemented by additional interfaces. Below are examples of steps needed to carry out policy evaluation and corresponding helper interfaces.

a)Create the policy evaluation object:

\begin{verbatim}
    // Create object which provides an interface 
    // for loading other objects
    ArcSec::EvaluatorLoader eval_loader; 
    //Load the Evaluator 
    ArcSec::Evaluator* eval = NULL;
    // Define name of policy evaluator.
    // This one is for evaluation ARC policies
    std::string evaluator = "arc.evaluator"; 
    eval = eval_loader.getEvaluator(evaluator);
\end{verbatim}
b)Create the policy object:

\begin{verbatim}
    ArcSec::Policy* policy = NULL; 
    // Define type of policy – ARC policy in this case
    std::string policyclassname = "arc.policy"; 
    // Define source from which policy to be taken
    ArcSec::SourceFile policy_source("Policy_Example.xml"); 
    // Load and parse policy
    policy = eval_loader.getPolicy(policyclassname, policy_source); 
\end{verbatim}

c)Create the request:
\begin{verbatim}
    ArcSec::Request* request = NULL;
    // Define type of request – ARC request in this case
    std::string requestclassname = "arc.request";
    // Define source from which request to be taken
    ArcSec::SourceFile request_source("Request.xml");
    // Load and parse request
    request = eval_loader.getRequest(requestclassname, request_source);
\end{verbatim}

d)Add the policy into Evaluator object:
\begin{verbatim}
    eval->addPolicy(policy);
\end{verbatim}

e)Evaluate the request object:
\begin{verbatim}
    ArcSec::Response *resp = NULL;
    resp = eval->evaluate(request);
\end{verbatim}

The steps d) and e) can also be replaced by:

\begin{verbatim}
    resp = eval->evaluate(request, policy);
\end{verbatim}

The Evalutor::evaluate() method can also be feed up with both Policy/Request objects and their sources in any combination. See example code at http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/testinterface.cpp for more details about usage of the interface.

The description of mentioned classes and their methods are avaialble in  API document at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/KnowARC-API.pdf?format=raw .

%section policy_eval (end)

\section{Policy Decision Service} % (fold)
\label{sec:policy_decision_service}
Policy decision service is a service implementation which contains the functionality of ArcPDP. It will accept the soap request containing policy decision request and return soap response containing policy decision response.

The WSDL description of policy decision service is available at

 http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl.

It's configuration is presented in Section 7.5.

%section policy_decision_service (end)

\section{Security Attributes} % (fold)
\label{sec:sec_attributes}

\subsection{Infrastructure} % (fold)
\label{subsec:sec_attr_infrastructure}

Security Attributes represent security related information inside HED framework and store information representing various aspects needed to perform authorization decison - identity of client, requested action, targeted resource, constraint policies. 
Each kind of Security Attribute is represented by own class inherited from parent SecAttr class <arc/message/SecAttr.h>. Each Security Attribute stores it's information in internal format and is capable to export it to one of predefined formats using Export() method. Currently only supported format is ARC Policy/Request XML document described in Section 8.1and 8.2
Collectors of Security Attributes instantiate corresponding classes and link them to Secuirity Attributes containers - MessageAuth <arc/message/MessageAuth.h> and MessageAuthContext <arc/message/Message.h> storing collected attributes per request and per session correspondingly. Each attribute is assigned a name. Current implementations of Security Attributes Collectors are either integrated into existing MCCs or implemented as separate SecHandler plugins. See Section 5.2for available Collectors and corresponding Security Attributes. 
Note for service developers: Services may implement own authorization algorithms. But they may use Security Atributes as well by providing instances of classes inherited from SecAttr and running them through either configured or hardcoded processors/PDPs.
Processors of Security Attributes are implemented as Policy Decision Point components. Currently there are 2 PDP components available:
Arc PDP makes use of Security Attributes containing identities of client, resource and requested action. It evaluates either all or selected set of attributes against specified Policy documents thus making it possible to enforce policies defined/selected by service providers. 
Delegation PDP is described below in Section 6.3.

\subsection{Available collectors} % (fold)
\label{subsec:sec_attr_avail_collectors}

Here Security Attribute collectors distributed as part of the ARC1 are described except those used for Delegation Restrictions. Those are described in Section 6.2

\subsubsection{TCP} % (fold)
\label{subsubsec:sec_attr_TCP}
Information is collected inside TCP MCC. The Security Attribute is stored under name 'TCP' and exports ARC Request with following attributes:

\begin{table}[ht]
\caption{Security Attributes collected at TCP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/localendpoint & service\_ip[:service\_port] \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/remoteendpoint & client\_ip[:client\_port] \\ \hline
\end{tabular}
\label{table:tcp_attr}
\end{table}

\subsubsection{TLS and VOMS} % (fold)
\label{subsubsec:sec_attr_TLS_VOMS}
Information is collected inside TLS MCC. Generated Security Attribute class is stored under name 'TLS' and exports ARC Request with following attributes:

\begin{table}[ht]
\caption{Security Attributes collected at TLS MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/ca & signer of first certificate in client's chain
 \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/chain & Subject of certificate in client's chain - multiple items \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/subject & Subject of last certificate in client's chain \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/identity & Subject of last non-proxy certificate in client's chain \\ \hline
SubjectAttribute & http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute & VOMS attributes extracted from whole client's chain of certificates \\ \hline
\end{tabular}
\label{table:tls_attr}
\end{table}

As one can see in addition to information extractable from generic TLS/SSL session this collector also understands attribute certificates provided by VOMS server and embedded into X.509 certificate

The VOMS attributes are presented in format similar to VOMS FQAN with slight modifications. Differently from FQAN all values are prepended with their names like VO and Group. Missing elements are reported as having string value NULL. Example of VOMS attribute looks like:

\emph{/VO=knowarc.eu/Group=testers/Role=admin/Capability=NULL}

Each set of attributes is accompanied by identifier of service which provided those attributes. It is made of \textit{voname} element with name of VO followed by optional element \textit{hostname} with hostname and port of service. Here is an example:

\emph{/voname=knowarc.eu/hostname=arthur.hep.lu.se:15001}

If VOMS extensions contain user definable attributes those are presented together with the information of their grantor. They consist of \textit{voname} and \textit{hostname} elements presented above (if \textit{hostname} is missing it is assigned string value NULL) followed by user attribute. It's pattern is \textit{qualifier:name=value}. Here qualifier acts as namespace and is usually same as VO name. Below is an example of such attribute:

\emph{/voname=knowarc.eu/hostname=arthur.hep.lu.se:15001/knowarc.eu:UniqueKnowarcAttribute=False}

Configuration of the TLS MCC is described in section 8.5.

\subsubsection{HTTP} % (fold)
\label{subsubsec:sec_attr_HTTP}
Information is collected inside HTTP MCC. The Security Attribute is stored under name 'HTTP' and exports ARC Request with following attributes:

\begin{table}[ht]
\caption{Security Attributes collected at HTTP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/http/path & HTTP path without host and port part \\ \hline
Action & http://www.nordugrid.org/schemas/policy-arc/types/http/method & HTTP method \\ \hline
\end{tabular}
\label{table:http_attr}
\end{table}

\subsubsection{SOAP} % (fold)
\label{subsubsec:sec_attr_SOAP}
Information is collected inside SOAP MCC. Security Attribute is stored under name 'SOAP' and exports ARC Request with following attributes:

\begin{table}[ht]
\caption{Security Attributes collected at SOAP MCC}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{Element} & \textbf{AttributeId} & \textbf{Content} \\ \hline
Resource & http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint & To element of WS-Addressing structure \\ \hline
Action & http://www.nordugrid.org/schemas/policy-arc/types/soap/operation & SOAP top level element name without namespace prefix \\ \hline
Context & http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace & Namespace of SOAP top level element \\ \hline
\end{tabular}
\label{table:soap_attr}
\end{table}

%section sec_attributes (end)

\section{Delegation} % (fold)
\label{sec:delegation}

\subsection{Delegation Architecture} % (fold)
\label{subsec:delegation_arch}
In current implementation delegation is achieved through Identity Delegation implemented using X509 Proxy Certificates as defined in RFC 3820 [1]. Client wishing to allow service to act on it's behalf provides Proxy Certificate to the service using Web Service based Delegation interface described in Section 6.4.

For limiting the scope of delegated credentials along with usually used time constraints it is possible to attach Policy document to Proxy Certificate. According to RFC 3820 Policy is stored in \textit{ProxyPolicy} extension. In order not to introduce new type of object Policy is assigned \textit{id-ppl-anyLanguage} identifier. RFC 3820 allows any octet string associated with such object. We are using textual representation of ARC Policy XML document.

Each deployment implementing Delegation Restrictions must use dedicated Security Handler plugin (see section 5.1) to collect all Policy documents from Proxy Certificates used for establishing secure connection. Then those documents must be processed by dedicated Policy Decision Point plugin (see section 2.3) to make a final decision based on collected Policies and various information about client's identity and requested operation. Service or MCC chain supporting Delegation Restrictions must accept negative decision of this PDP as final and do not override it with any other decision based on other policies.

To have Delegation Restriction working their processing must be enabled in all participating clients and services. Because Delegated Restrictions are marked as critical extension in X.509 proxy certificate any  service which does not support such extension will fail to autheticate client presenting such certificate.

\subsection{Delegation Collector} % (fold)
\label{subsec:delegation_collector}
This Security Attribute is collected by dedicated Security Handler plugin named ``delegation.collector'' avaialble as part of the ARC1 distribution. It extracts policy document stored inside X509 certificate proxy extension as defined in RFC3820 and described in Section 6.1. All proxy certificates in a chain provided by client are examined and all available policies are extracted. Configuration of Delegation SecHandler is described in section 8.12.

Extracted content is converted into XML document. Then document is checked to be of ARC Policy kind. If policy is not recognized as ARC Policy procedure fails and that causes failure of communication.

Proxy certificates with id-ppl-inheritAll [5. RFC3820. http://www.faqs.org/rfcs/rfc3820.html] property are passed through and no policy document is generated for them. Proxies with other type of policies including id-ppl-independent are not accepted and generate immediate failure.

\subsection{Delegation PDP} % (fold)
\label{subsec:delegation_pdp}
The Delegation PDP is similar to the Arc PDP described above except that it takes it's Policy documents directly from Security Attributes. Differently from Arc PDP it is meant to be used for enforcing policies defined by client. Configuration of Delegation PDP is described in section 8.11.

\subsection{Delegation Interface} % (fold)
\label{subsec:delegation_interface}
Delegation interface in the ARC1 is implemented using Web Service approach. Each ARC1 service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials. Here is how delegation procedure works (also shown in figures 6 and 7) :

    \begin{itemize}
        \item \emph{Step 1}
            \begin{itemize}
                \item Client contacts service requesting operation DelegateCredentialsInit. This operation has no arguments.
                \item Service responds with DelegateCredentialsInitResponse message with element TokenRequest. That element contains credentials request generated by service in Value. Type of request is defined by attribute Format. Currently only supported format is x509. Along with Value service provides identifier Id which is used in second step.
            \end{itemize}
        \item \emph{Step 2}
            \begin{itemize}        
                \item Client requests UpdateCredentials operation with DelegatedToken argument. This element contains Value with serialized delegated credentials and Id which links it to first step. Delegated token element may also contain multiple Reference elements. Reference refers to the object which these credentials should be applied to in a way specific to the service. The DelegatedToken element may also be used for delegating credentials when Step 2 is combined with other operations on service in service specific way.
                \item Service responds with empty UpdateCredentialsResponse message.
            \end{itemize}
    \end{itemize}
Optionally step 2 can be skipped and the DelegatedToken element provided to Service as additional payload of other service specific message.

The Id element obtained in the step 1 can be reused multiple times with different content of the Value element.

WSDL of portType implementing delegation functionality can be found at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/libs/delegation/delegation.wsdl .

\subsection{Delegated Credentials (Proxy) Generation Utility} % (fold)
\label{subsec:delegation_client}
Command line utility “arcproxy” can be used to generate X.509 Proxy Certificate with (or without)  Policy embedded.
The arcproxy my be used in following way:

\emph{approxy -P proxy.pem -C cert.pem -K key.pem -c constraint}

Here options -P, -C and -K specify path to files containing generated Proxy, user's credentials and user's private key respectively. By using argument ``-c'', some constraints can be specified for proxy certificate. Each constraint string is a key and value pair with key representing type of contraint. There may be multiple -c options specified.  Currently suported contraint types are:

    \begin{itemize}
        \item validityStart, validityEnd and validityPeriod  specify when Proxy becomes valid, when it's validity finishes or for how long the Proxy is valid respectively. For example

-c validityStart=2008-05-29T10:20:30Z

-c validityEnd=2008-06-29T10:20:30Z

        \item proxyPolicy and proxyPolicyFile specify the Policy document to be embedded into the Proxy either directly or by pointing to the file which contains that document. Like

-c proxyPolicyFile=delegation\_policy.xml
    \end{itemize}

The Policy maybe any of any type supported by ARC1 middleware (or third-party plugins) installed on the services where that policy is processed. Currenlty supported Policies include ARC Policy (described in section 8) and GACL Policy [2].

Simple example below renders delegated credentials usable only for contacting service attached to HTTP communication channel under path /arex (line 5)  and allows HTTP operation POST (line 8) on it.

\begin{verbatim}
1.<?xml version="1.0" encoding="UTF-8"?>
2.<Policy xmlns="http://www.nordugrid.org/schemas/policy-arc" PolicyId="sm-example:policy1" 
      CombiningAlg="Deny-Overrides">
3.    <Rule RuleId="rule1" Effect="Permit">
4.        <Resources>
5.            <Resource Type="string" 
                  AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/http/path">
                 /arex
              </Resource>
6.        </Resources>
7.        <Actions>
8.            <Action Type="string" 
                  AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/http/method">
                  POST
              </Action>
9.        </Actions>
10.   </Rule>
11.</Policy>
\end{verbatim}

Another example of the delegation policy is presented below. This policy restricts usage of delegated credentials to SOAP operation CreateActivity (line 5) of Basic Execution Service (BES) [4] namespace (line 9). Such policy could be embedded into credentials delegated to high level Brokering service performing Grid job submission to low level BES on behalf of user.

\begin{verbatim}
1.<?xml version="1.0" encoding="UTF-8"?>
2.<Policy xmlns="http://www.nordugrid.org/schemas/policy-arc" PolicyId="sm-example:policy1" 
    CombiningAlg="Deny-Overrides">
3.    <Rule RuleId="rule1" Effect="Permit">
4.        <Actions>
5.            <Action Type="string" 
                AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/soap/operation">
                CreateActivity
              </Action>
6.        </Actions>
7.        <Conditions>
8.            <Condition>
9.                <Attribute Type="string" 
                      AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace">
                      http://schemas.ggf.org/bes/2006/08/bes-factory
                  </Attribute>
10.           </Condition>
11.       </Conditions>
12.   </Rule>
13.</Policy>
\end{verbatim}

\subsubsection{Delegated Credentials with VOMS Attributes} % (fold)
\label{subsec:delegation_voms}
Currently the proxy creation utility arcproxy can also be used for creating VOMS Proxy Certificate, as the way to replace the ``voms-proxy-init'' utility.

%section delegation (end)

\section{Web Service Security Support} % (fold)
\label{sec:webservice}

\subsection{UsernameToken SecHandler} % (fold)
\label{subsec:username_token}
The UsernameToken SecHandler is meant for processing - generating and extracting - WS-Security [4] UsernameToken in the SOAP header. Hence it must be attached to the MCC which processes SOAP payloads - like SOAP MCC or Service accepting SOAP messages. For description of configuration see section 8.13.

For the incoming message this SecHandler authorizes SOAP message according to specified configuration.

For the outgoing message this SecHandler creates adds proper token into SOAP header according to configuration.

\subsection{X509Token SecHandler} % (fold)
\label{subsec:x509_token}
The X509Token SecHandler is meant for processing - generating and extracting - WS-Security [4] X509Token from SOAP header. Hence it must be attached to the MCC which processes SOAP payloads – like SOAP MCC or Service accepting SOAP messages. For description of configuration see section 8.14.

For the incoming message this SecHandler decrypts and checks signature of SOAP message using attached public key and verifies that key against specified CA certificate.

For the outgoing message this SecHandler creates X509Token in SOAP header. SOAP message body is encrypted and signed.

%section webservice (end)

\section{Schemas, descriptions and examples} % (fold)
\label{sec:schema_description_example}

\subsection{Authorization Policy} % (fold)
\label{subsec:authz_policy}
XML schema with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd .

\subsection{Authorization Request} % (fold)
\label{subsec:authz_request}
XML schema with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd .

\subsection{Authorization Response} % (fold)
\label{subsec:authz_response}
XML schema with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Response.xsd .

\subsection{Interface of policy decision service} % (fold)
\label{subsec:interface_pds}
WSDL with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl .

\subsection{TLS MCC configuration} % (fold)
\label{subsec:tlsmcc_conf}
For full description of TLS MCC configuration please read ``The Hosting Environment of the Advanced Resource Connector middleware''. Here only part related to VOMS attributes extraction is provided for convenience. Configuration schema can be found at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/mcc/tls/tls.xsd .

While processing VOMS extension of X.509 certificate only attributes which can be verified are extracted and collected in the list of Security Attributes. To ensure proper authentication trusted VOMS services must to be configured. The VOMS services are identified by certificates which they use to sign AC with VOMS related information. And also by whole certificates chain used to sign VOMS service certificate.

The TLS MCC configuration makes it possible to specify DNs of all certificates in such chains. Each chain is stored in separate \textit{$<$VOMSCertTrustDNChain$>$} element. Each such element is composed either of multiple \textit{$<$VOMSCertTrustDN$>$} elements or single \textit{$<$VOMSCertTrustRegex$>$}

Each \textit{$<$VOMSCertTrustDN$>$} element defined DN of one certificate in a chain starting from certificate of VOMS service and down to DN of last CA in the chain.

The \textit{$<$VOMSCertTrustRegex$>$} element defines regular expression which is applied to every certificate in the chain.

Along with \textit{$<$VOMSCertTrustDNChain$>$} it is also possible to specify \textit{$<$VOMSCertTrustDNChainsLocation$>$}.

The \textit{$<$VOMSCertTrustDNChainsLocation$>$} specifies path to file containing XML document with single \textit{$<$VOMSCertTrustDNChain$>$} element.

Below is an example presenting all possible options.

\begin{verbatim}
<tls:VOMSCertTrustDNChain> 
  <tls:VOMSCertTrustDN>/O=Grid/O=NorduGrid/CN=host/arthur.hep.lu.se</tls:VOMSCertTrustDN>
  <tls:VOMSCertTrustDN>/O=Grid/O=NorduGrid/CN=NorduGrid CA</tls:VOMSCertTrustDN>
</tls:VOMSCertTrustDNChain>
<tls:VOMSCertTrustDNChain>
  <tls:VOMSCertTrustDN>/DC=ch/DC=cern/OU=computers/CN=voms.cern.ch</tls:VOMSCertTrustDN>
  <tls:VOMSCertTrustDN>/DC=ch/DC=cern/CN=CERN CA</tls:VOMSCertTrustDN>
</tls:VOMSCertTrustDNChain>
<tls:VOMSCertTrustDNChain>
  <tls:VOMSCertTrustRegex>^/O=Grid/O=NorduGrid</tls:VOMSCertTrustRegex>
</tls:VOMSCertTrustDNChain>
<tls:VOMSCertTrustDNChainsLocation>./voms_trust.xml</tls:VOMSCertTrustDNChainsLocation>
\end{verbatim}

\subsection{Configuration of PDP service} % (fold)
\label{subsec:pdpservice_conf}
XML schema with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.xsd .

Below is an example configuration of PDP service which can evaluate ARC Request against ARC Policy stored in local file.

\begin{verbatim}
<Service name="pdp.service" id="pdp_service">
<!--The element <Evaluator/>, <Policy/> and <Request/> configuration
       are supposed to be used to load object; element <PolicyStore/> is
       supposed to be used to get the location of policy-->
   <pdp:PDPConfig>
       <pdp:PolicyStore>
           <Location Type="file">Policy_Example.xml</Location>
           <!-- other policy location-->
       </pdp:PolicyStore>
       <pdp:Evaluator name="arc.evaluator" />
       <pdp:Policy name="arc.policy" />
       <pdp:Request name="arc.request" />
   </pdp:PDPConfig>
</Service>
\end{verbatim}

See Section 8.9for the explanation of ARC Policy.

\subsection{Authorization SecHandler configuration} % (fold)
\label{subsec:authzhandler_conf}
XML schema with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcauthzsh/SimpleListAuthZ.xsd.

Default behavior of Authorization SecHandler is to execute all PDPs corresponding to elements \textit{$<$PDP$>$} in configuration sequentially till either one of them fails or all produced positive results. This behavior may be modified by attribute ``action'' of embedded \textit{$<$PDP/$>$} elements. Following options are supported:

    \begin{itemize}
        \item breakOnAllow - if PDP returned positive result stop PDPs processing and return positive result. Otherwise continue to next PDP or return negative result if no more PDPs to process. That is a default behavior.
        \item BreakOnDeny - if PDP returned negative result stop PDPs processing and return negative result. Otherwise continue to next PDP or return negative result if no more PDPs to process.
        \item BreakAlways – stop processing PDPs and return result which this PDP returned.
        \item BreakNever – continue to next PDP. If there are no more PDPs to process then return result which this PDP returned.
    \end{itemize}

Below is an example of the Authorization SecHandler with 4 PDPs in the list:

simplelist.pdp – for comparing client's credentilals to list of DNs

arc.pdp – for comparing collected information to specified policies

pdpservice.invoker – for contacting external PDP service.xml

delegation.pdp – for evaluating restrictions embedded into X.509 proxy certificates

\begin{verbatim}
<SecHandler name="arc.authz" id="authz" event="incoming">
  <PDP name="simplelist.pdp" location="simplelistfile"/>
  <PDP name="arc.pdp">
    <PolicyStore>
      <Location type="file">Policy_Example.xml</Location>
    </PolicyStore>
  </PDP>
  <PDP name="pdpservice.invoker">
    <ServiceEndpoint>https://127.0.0.1:60001/pdp.service</ServiceEndpoint>
    <KeyPath>./testkey-nopass.pem</KeyPath>
    <CertificatePath>./testcert.pem</CertificatePath>
    <CACertificatePath>./cacert.pem</CACertificatePath>
  </PDP>
  <PDP name="delegation.pdp"/>
</SecHandler>
\end{verbatim}

\subsection{SimpleList PDP configuration and Policy Example} % (fold)
\label{subsec:simplepdp_conf}
XML schema with comments is available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/simplelistpdp/SimpleListPDP.xsd .

Below is an example configuration of SimpleList PDP inside ``echo'' service.

\begin{verbatim}
<Service name="echo" id="echo">
  <SecHandler name="arc.authz" id="authz" event="incoming">
    <PDP name="simplelist.pdp" location="simplelist"/>
  </SecHandler>
  <echo:prefix>[ </echo:prefix>
  <echo:suffix> ]</echo:suffix>
</Service>
\end{verbatim}

The attribute ``name'' of \textit{$<$PDP/$>$} is critical for loading the object. Specifically, the name ``simplelist.pdp'' is for loading the SimpleList PDP object.

The policy file ``simplelist'' is a local file which contains the list of X509 subjects of authorized entities. It the peer certificate is proxy certificate, the identity in this list should only include the original DN of users's certificate.
For example content of simplelist file may look like this:

\emph{/C=NO/O=UiO/CN=test1}

\emph{/C=NO/O=UiO/CN=test2}

\subsection{Arc PDP configuration and Policy Example} % (fold)
\label{subsec:arcpdp_conf}
XML schema with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/ArcPDP.xsd .

Below is an example of configuration of Arc PDP inside ``echo'' service.
\begin{verbatim}
<Service name="echo" id="echo">
  <SecHandler name="arc.authz" id="authz" event="incoming">
    <PDP name="arc.pdp">
      <PolicyStore>
        <Location type="file">Policy_Example.xml</Location>
        <!--other policy location-->
      </PolicyStore>
    </PDP>
  </SecHandler>
  <echo:prefix>[ </echo:prefix>
  <echo:suffix> ]</echo:suffix>
</Service>
\end{verbatim}

The name ``arc.pdp'' is for loading the ArcPDP object.

There could be a few policy files under \textit{$<$PolicyStore/$>$}. The request will be checked against all of the policies.

There is an example policy for echo service below. See Section 8.1for the policy schema. The example policy is  made of following elements:

  \newcounter{policy_count}
  \begin{list}{\arabic{policy_count}.}
  {\usecounter{policy_count}\setlength{\rightmargin}{\leftmargin}}
    \item Line 14  defines resource being protected. In this it is everything located under HTTP path ``/Echo''.
    \item Lines 17 and 18 define allowed HTTP operations to be ``POST'' and ``GET''. Line 19 also defines SOAP operation ``echo'' to be applied to service at path defined above.
    \item Lines 10 and 9 require the requester to present X509 certificate with specified identity and signed by specified Certification Authority.
    \item No \textit{$<$Conditions/$>$} defined.
    \item Line 3 defines that if and only if all of the above constraints have been satisfied by requester, the \textit{$<$Rule/$>$} evaluates to Permit decision.
  \end{list}

The Secuirity Attributes used by Arc PDP are collected by different MCCs. It is possible for service to collect some application-specific attributes by implementing class inherited from SecAtt. And that should be the task of application developer.

Administrator of service can configure Authorization SecHandler - arc.authz - for each MCC and Service and define reasonable and meaningful policy. While defining policy the administrator must take into account that the attributes defined in the policy should be already collected by previous components in a chain. For instance, policy with AttributeId ``http://www.nordugrid.org/schemas/policy-arc/types/http/path'' should not be configured inside SecHandler attached to MCCTLS.

\begin{verbatim}
1.<?xml version="1.0" encoding="UTF-8"?>
2.<Policy xmlns="http://www.nordugrid.org/schemas/policy-arc" PolicyId="sm-example:arcpdppolicy" 
     CombiningAlg="Deny-Overrides">
3.   <Rule Effect="Permit">
4.     <Description>
5.       Example policy for echo service
6.     </Description>
7.      <Subjects>
8.         <Subject>
9.            <Attribute 
               AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/ca" 
               Type="string">
               /C=NO/ST=Oslo/O=UiO/CN=CA
              </Attribute>
10.           <Attribute 
               AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/identity" 
               Type="string">
               /C=NO/ST=Oslo/O=UiO/CN=test
              </Attribute>
11.         </Subject>
12.      </Subjects>
13.      <Resources>
14.         <Resource 
             AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/http/path" 
             Type="string">
             /Echo
            </Resource>
15.      </Resources>
16.      <Actions>
17.         <Action 
             AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/http/method" 
             Type="string">
             POST
            </Action>
18.         <Action 
             AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/http/method" 
             Type="string">
             GET
            </Action>
19.         <Action 
             AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/soap/operation" 
             Type="string">
             echo
            </Action>
20.      </Actions>
21.      <Conditions/>
22.   </Rule>
23.</Policy>
\end{verbatim}

\subsection{PDP Service Invoker configuration} % (fold)
\label{subsec:pdpservice_invoker_conf}
Configuration XML schema with comments is available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/pdpserviceinvoker/ArcPDPServiceInvoker.xsd .

Below is an example of configuration of PDP Service Invoker inside ``echo'' service.

\begin{verbatim}
<Service name="echo" id="echo">
  <SecHandler name="arc.authz" id="authz" event="incoming">
    <!--Remote pdp service invoking-->
      <PDP name="pdpservice.invoker">
        <ServiceEndpoint>https://127.0.0.1:60001/pdp.service</ServiceEndpoint>
        <KeyPath>./key.pem</KeyPath>
        <CertificatePath>./cert.pem</CertificatePath>
        <CACertificatePath>./ca.pem</CACertificatePath>
      </PDP>
  </SecHandler>
  <next id="echo"/>
  <echo:prefix>[ </echo:prefix>
  <echo:suffix> ]</echo:suffix>
</Service>
\end{verbatim}

The name ``pdpservice.invoker'' defines the PDP Service Invoker object.

The PDP Service Invoker is a client of PDP Service. The configuration options include endpoint of service and credentials to be used for establishing secure connection.

\subsection{Delegation PDP configuration} % (fold)
\label{subsec:delegpdp_conf}
Configuration XML schema with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/delegationsh/DelegationPDP.xsd .

Below is an example of configuration of Delegation PDP inside ``echo'' service.

\begin{verbatim}
<Service name="echo" id="echo">
  <SecHandler name="arc.authz" id="authz" event="incoming">
    <PDP name="delegation.pdp"/>
  </SecHandler>
  <next id="echo"/>
  <echo:prefix>[ </echo:prefix>
  <echo:suffix> ]</echo:suffix>
</Service>
\end{verbatim}

For Delegation PDP, no specific configuration is needed. It is enough to switch it on by adding \textit{$<$PDP name=``delegation.pdp''/$>$} under \textit{$<$SecHandler$>$} which supports processing of PDPs (currently arc.authz).
The precondition for using Delegation PDP is that there must be Delegation SecHandler instantiated earlier in the chain.

\subsection{Delegation SecHandler Configuration} % (fold)
\label{subsec:deleg_sechandler_conf}
Below is an example of configuration of Delegation SecHandler inside TLS MCC component.

\begin{verbatim}
         <Component name="tls.service" id="tls"> <next id="http"/>
            <tls:KeyPath>./key.pem</tls:KeyPath>
            <tls:CertificatePath>./cert.pem</tls:CertificatePath>
            <tls:CACertificatePath>./ca.pem</tls:CACertificatePath>
            <!--delegation.collector must be inside tls MCC-->
            <SecHandler name="delegation.collector" 
                id="delegation" event="incoming"></SecHandler>
         </Component>
\end{verbatim}

Current implementation of Delegation SecHandler must be attached to TLS MCC.

\subsection{UsernameToken SecHandler Configuration} % (fold)
\label{subsec:ut_sechandler_conf}
Configuration XML schema with comments is available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/usernametokensh/UsernameTokenSH.xsd .

Below is an example of configuration of UsernameToken SecHandler inside MCCSOAP component of the service side. This example processes UsernameToken related information in SOAP message and returns failure if any problem found. In detail, this SecHandler will check header of the incoming SOAP message for the presence of UserName WS-Security token and compare provided password value to those stored in the local file password.txt.

\begin{verbatim}
<Component name="soap.service" id="soap">
  <next id="echo"/>
  <SecHandler name="usernametoken.handler" id="usernametoken" event="incoming">
    <Process>extract</Process>
    <PasswordSource>password.txt</PasswordSource>
  </SecHandler>
</Component>
\end{verbatim}

This example will add token with username ``user'' and password ``pass'' (using ``digest'' encoding algorithm) into outgoing SOAP message.

For the client side, the developer should add the configuration information about X.509 security handler into client's configuration, in order to generate X509 Token into any SOAP message. Below is an example of how to use it:

\begin{verbatim}
    Arc::XMLNode sechanlder_nd_ut("\ 
        <SecHandler name='usernametoken.handler' id='usernametoken' event='outgoing'>\ 
            <Process>generate</Process>\ 
            <PasswordEncoding>digest</PasswordEncoding>\ 
            <Username>user</Username>\ 
            <Password>passwd</Password>\ 
        </SecHandler>"); 
  Arc::ClientSOAP *client; 
  client = new Arc::ClientSOAP(mcc_cfg,url); 
  client->AddSecHandler(sechanlder_nd_ut, "arcshc");
\end{verbatim}

\subsection{X509Token SecHandler configuration} % (fold)
\label{subsec:xt_sechandler_conf}
Configuration XML schema with comments is available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/x509tokensh/X509TokenSH.xsd.

Below is an example of configuration of X509Token SecHandler inside MCCSOAP component of the service side. This example processes X509Token related information in SOAP message and returns failure if any problem found.

\begin{verbatim}
<Component name="soap.service" id="soap">
  <next id="echo"/>
  <SecHandler name="x509token.handler" id="x509token" event="incoming">
    <Process>extract</Process>
    <CACertificatePath>ca.pem</CACertificatePath>
  </SecHandler>
</Component>
\end{verbatim}

For the client side, the developer should add the configuration information about X.509 security handler into client's configuration, in order to generate X509 Token into any SOAP message. Below is an example of how to use it:

\begin{verbatim}
  Arc::XMLNode sechanlder_nd_xt("\ 
        <SecHandler name='x509token.handler' id='x509token' event='outgoing'>\ 
            <Process>generate</Process>\ 
            <CertificatePath>./testcert.pem</CertificatePath>\ 
            <KeyPath>./testkey-nopass.pem</KeyPath>\ 
        </SecHandler>"); 
  Arc::ClientSOAP *client; 
  client = new Arc::ClientSOAP(mcc_cfg,url); 
  client->AddSecHandler(sechanlder_nd_xt, "arcshc");
\end{verbatim}

%section schema_description_example (end)

\section{Using SAML2 Web SSO for Authentication and Attribute-based Authorization} % (fold)
\label{sec:samlsso}
   In ARC1, besides that GSI is supported for talking with external grid services which are based on GSI, the standard TLS/SSL is also supported. No matter standard SSL/TLS or GSI is used, mutual authentication applies for both of them, and is the default configuration for grid deployment; and X.509 certificate is required for both of the client and service sides. X.509 certificates is issued by certificate authorities (CA), and then CAs constitute trust federation and guarantee two different X.509 certificates from different CAs can accomplish authentication to each other. So if a user would access grid system, he/she should own a X.509 certificate which is issued by a CA that is trusted by other's entity in the grid system.

   ARC1 provides another option by which users are not required to possess a X.509 certificate for authentication. It is based on AAI(Authentication and Authorization Infrastructure) and utilize users' existing community credential (such as username password) for authentication rather than X.509 credential.

   SAML 2.0 specification defines the Web browser SSO profile which is a standardization for single sign-on (SSO). SSO enables the establishment of federated environment in which clients can sign in the home organization once and access services offered by different external organizations. By using SAML2.0 Web browser SSO profile, users can access external organizations by only authenticating to their own organization. Also instead of requiring users to possess X.509 credential, different authentication mechanisms, such as username/password authentication, IP address authentication, etc.

  The essential idea of SAML2.0 Web browser SSO profile is that uses can access external resources by only depending on their own existing credentials, instead of applying a new credential for accessing external organization. Concerning the use case of Grids which commonly require users to obtain a X.509 credential for the access to Grid, even if these users already have possessed the credential from their host organizations, integrating SAML2.0 SSO profile into Grid usage is considered to be a promising direction for the widely take up of Grids. The SAML2.0 web browser SSO profile is implemented in ARC middleware as another option for authentication, beside the mutual SSL/TLS authentication.

  SAML2.0 SSO profile defines three functional components, an Identity Provider, a Service Provider, and a user agent. The Identity Provider (IdP) is responsible for creating, maintaining, and managing user identity.  The Service Provider (SP) is responsible for controlling access to services and resources by using the SAML assertion produced and issued by IdP upon request. The user agent is responsible for initiating the access request to SP and redirecting the message between SP and IdP, and user agent is commonly hosted inside Web browser. There are several implementations of SAML2.0 SSO profile, among which Shibboleth is one implementation which has been widely deployed.

    In terms of implementation, since the SSO profile is primarily supposed to protect Web applications and provide authentication for Web users, some external code on the client and service side is implemented to integrate the SSO profile. On the client side, apart from the client interface for writing Web Service client, the user agent functionality of the Web browser is implemented in order to mimic its behavior, such as HTTP redirection and HTTP cookie processing. In fact, implementation of the user agent is also based on the client interface of ARC, specifically, the HTTPs client interface, since the client interface of ARC can support different protocols which are incarnated by different MCCs. The client developers who would use SAML2.0 SSO profile should call the user agent interface and then the Web Service client interface. On the service side, the Service Provider functionality (based on the HTTP MCC configured together with TLS MCC) is implemented by SP Service. For Identity Provider, the Shibboleth IdP implementation is used. Figure 8 shows the process of SAML2.0 SSO integrated in ARC client and service.

   Normally SP service will not be deployed independently, instead, it should be deployed together with other services. The SLCS service is the typical deployment about SP service. On the client side, client developer should use the ClientSAML2SSO instead of ClientInterface to call the client functionality. The SLCS client (arcslcs) is the typical usage of  ClientSAML2SSO interface.

    Once the SP Service has cooperated with user agent and Identity Provider (external) and succeeded to accomplish the SAML2 SSO profile,  the SP service will get the saml authentication assertion which asserts that the authentication has succeeded, and then SP service will store this assertion into session context.

    SAML2 assertion consumer handler (saml2ssoassertionconsumer.handler) is the security handler which will understand the authentication assertion and attribute assertion from IdP, and make authorization decision according to the attribute values inside these assertions. Currently the SAML2 assertion consumer handler is an empty security handler and does not effect the services.

The steps shows in Figure 8 are described as follows:

  \newcounter{sso_count}
  \begin{list}{\arabic{sso_count}.}
  {\usecounter{sso_count}\setlength{\rightmargin}{\leftmargin}}
    \item The client uses the user agent interface to launch a HTTP request including the IdP name (to which the user belongs) to the service side. The endpoint of the SP (Service Provider) service is the same as that of the other target services, except the last part of the endpoint is ``saml2sp'' which is specific for pointing to the SP Service. Note that we use Identity Provider (IdP) name here to simplify the IdP discovery process in order to avoid the IdP discovery process, because we suppose that the user who would access the target services should better know where is he from initially.

    \item The SP Service searches the metadata (we use the same metadata format as defined in Shibboleth) and gets the location of the single sign-on service (hosted in IdP) and also the location of assertion consuming service (hosted in this SP itself) in order to compose the SAML $<$samlp:AuthnRequest$>$ message. Then SP Service issues this $<$samlp:AuthnRequest$>$ message by using its own X.509 certificate (note that in the SAML SSO profile, X.509 certificates are still needed for IdP and SP) and sends back to user agent.

    \item User agent sends the $<$samlp:AuthnRequest$>$ message to the Identity Provider.

    \item Identity Provider requires an act of authentication. The authentication mechanism is outside of the SAML2.0 SSO profile. Shibboleth IdP implementation chooses some login handlers for authentication. The current user agent implementation is compatible with the Username/Password login handler of Shibboleth IdP. Through the HTTP protocol, the user agent will feed IdP with the username/password which has been given by the caller of user agent interface.

    \item Once the authentication has been succeeded, the IdP issues a SAML response including an encrypted (encrypted by destination SP’s public key) SAML assertion, and then this SAML response will be delivered by the user agent to the Service Provider.

    \item The SP Service verifies and checks the SAML response, decrypts and stores the SAML assertion into session/connection context. The SAML assertion includes the $<$saml:AuthnStatement$>$ and $<$saml:AttributeStatement$>$.

    \item The WS client launches the Grid/Web Service request via the same connection as the one which is used by user agent to contact SP Service.

    \item The Grid/Web Service checks the $<$saml:AuthnStatement$>$ from the session context to see if the session is still valid through the SecHandler called ``SAML assertion consumer SecHandler''. If valid, service handles the service processing and returns the response to WS client. Note that service requires that WS client is from the same connection as the one on which user agent contact SP service, in order to guarantee that the validity of SSO profile result effects the WS client/Web Service interaction.

    The SP service and other functional service(s) are hosted by the same container, and they use the same X.509 credential. The client authentication is switched off, so that client doesn’t need to use any X.509 credential. Only the trusted certificates (CA certificates for both SP and IdP) need to be configured for the client side, so that SP and IdP can authenticate themselves to the client. As required by the SAML2.0 profile, the SP and IdP should have trust relationship to each other.
   \end{list}

    A typical configuration for service that is enhanced by SAML2SSO profile is as follows:

\begin{verbatim}
    <Chain>
        <Component name="tcp.service" id="tcp"> 
            <next id="tls"/> 
            <tcp:Listen><tcp:Port>60000</tcp:Port></tcp:Listen> 
        </Component> 
        <Component name="tls.service" id="tls"> 
            <next id="http"/> 
            <tls:KeyPath>./testkey-nopass.pem</tls:KeyPath> 
            <tls:CertificatePath>./testcert.pem</tls:CertificatePath>
           <!--Since the client side's certificate is not required to exist, the trusted CA certificates are not needed here-->
            <!--tls:CACertificatePath>./cacert.pem</tls:CACertificatePath--> 
           <!--Client side authentication is not required-->
            <tls:ClientAuthn>false</tls:ClientAuthn> 
        </Component> 
        <Component name="http.service" id="http"> 
            <next id="plexer">POST</next> 
        </Component> 
        <Plexer name="plexer.service" id="plexer"> 
            <next id="samlsp">/saml2sp</next> 
            <next id="soap">/echo</next>
            <!--Other services ID can be put here-->
            <!--next id="soap">/aotherservice</next-->
        </Plexer> 
        <Component name="soap.service" id="soap"> 
            <next id="echo"/> 
           <!--The security handler used for doing, e.g., access control based on the SAML assertion from the SAML2.0 SSO profile; currently it is just a prototype implementation-->
            <SecHandler name="saml2ssoassertionconsumer.handler" id="saml2ssosp" event="incoming"/> 
        </Component> 
        <Service name='saml.sp' id='samlsp'>
            <!--Metadata that is according to SAML metadata standard. urn:oasis:names:tc:SAML:2.0:metadata-->
            <MetaDataLocation>./test_metadata.xml</MetaDataLocation> 
           <!--The entity ID of this Service Provider-->
            <ServiceProviderName>https://squark.uio.no/shibboleth-sp</ServiceProviderName>
           <!--The certificate and key file which will be used by Service Provide to sign the AuthnRequest-->
            <KeyPath>./testkey-nopass.pem</KeyPath> 
            <CertificatePath>./testcert.pem</CertificatePath> 
        </Service> 
        <Service name="echo" id="echo"> 
            <next id="echo"/> 
            <echo:prefix>[ </echo:prefix> 
            <echo:suffix> ]</echo:suffix> 
        </Service> 
    </Chain> 
\end{verbatim}

    The client developer should use the ClientSOAPwithSAML2SSO  or ClientHTTPwithSAML2SSO instead of ClientInterface for the integration of SAML2SSO profile. A typical usage of ClientSOAPwithSAML2SSO is shown as follows. Note the IdP name and the username/password should be feed as the input argument for ``process'' method.

\begin{verbatim}
   Arc::ClientSOAPwithSAML2SSO *client = NULL; 
   client = new Arc::ClientSOAPwithSAML2SSO(mcc_cfg,url); 
   Arc::MCC_Status status = client->process(&req,&resp, idp_name, username, password); 
\end{verbatim}

    The message flow of interaction between Client, ARC SP Service, ARC Service and Shibboleth IdP is also shown in Figure.9.

  One benefit of the SAML2.0 SSO profile that is worth mentioning is: the Identity Provider could cache the authentication result through session management once the user agent has succeeded to authenticate; then for a short period this authentication result is valid so that the user agent doesn’t need to feed IdP with user’s username and password the next time (if this point of time is not out of the scope of valid period) it authenticates against IdP. So user (or the client on behalf of this user) can access multiple security  domains with only providing his name and password once, which is the characteristic of single sign-on.

    Since the Shibboleth implementation of SAML is standard-compliant and widely deployed, the solution implemented in ARC can easily interoperate with other SAML implementations with minimum change, and more importantly, this solution can succeed to utilize the widely deployed SAML implementation for authentication in Grid systems by avoiding the usage of X.509 certificate.

%section samlsso (end)

\section{Short-Lived Credential Service} % (fold)
\label{sec:slcs}
In order to create X.509 credential for inter-operating with Grid services that requires client side certificate authentication, based on the solution described in Section 9, a short lived credential service (SLCS) is implemented by which user can get a short-lived X.509 certificate without being bothered to contact any registration authority (RA) or certificate authority (CA).

    SLCS service is for signing short-lived x509 credential based on user's username/password credential. Then the user can use this short-lived x509 credential to access grid services/resources where x509 credential is required by default. SLCS service should depend on the SP (Service provider) service which is one of the participants of SAML2 SSO profile (SAML2 SSO profile is used for authenticating based on username/password credential and getting SAML authentication assertion; and SAML authentication assertion is then used as basis for signing short-lived x509 credential).

    The SLCS service is also a Web Service (standard Web Service implemented by using ARC service interface), and the SLCS client is a specific command-line interface (CLI) which includes the user agent and WS client. The whole process of SLCS invocation is showed in Figure 10 (from step 1 to step 8), which is the same as in Figure 8, except that step 7 and step 8 are invoked for the SLCS certificate request and response.

 SLCS service is supposed to run together with SP service. A typical configuration for SLCS service is as follows:

\begin{verbatim}
    <Chain>
        <Component name="tcp.service" id="tcp"> 
            <next id="tls"/> 
            <tcp:Listen><tcp:Port>60000</tcp:Port></tcp:Listen> 
        </Component> 
        <Component name="tls.service" id="tls"> 
            <next id="http"/> 
            <tls:KeyPath>./testkey-nopass.pem</tls:KeyPath> 
            <tls:CertificatePath>./testcert.pem</tls:CertificatePath> 
            <!--tls:CACertificatePath>./cacert.pem</tls:CACertificatePath--> 
            <tls:ClientAuthn>false</tls:ClientAuthn> 
        </Component> 
        <Component name="http.service" id="http"> 
            <next id="plexer">POST</next> 
        </Component> 
        <Plexer name="plexer.service" id="plexer"> 
            <next id="samlsp">/saml2sp</next> 
            <next id="soap">/slcs</next> 
        </Plexer> 
        <Component name="soap.service" id="soap"> 
            <next id="slcs"/> 
            <SecHandler name="saml2ssoassertionconsumer.handler" id="saml2ssosp" event="incoming"/> 
        </Component> 
        <Service name='saml.sp' id='samlsp'> 
            <MetaDataLocation>./test_metadata.xml</MetaDataLocation> 
            <ServiceProviderName>https://squark.uio.no/shibboleth-sp</ServiceProviderName> 
            <KeyPath>./testkey-nopass.pem</KeyPath> 
            <CertificatePath>./testcert.pem</CertificatePath> 
        </Service> 
        <Service name="slcs.service" id="slcs"> 
            <next id="slcs"/> 
            <slcs:CACertificate>./CAcert.pem</slcs:CACertificate> 
            <slcs:CAKey>./CAkey.pem</slcs:CAKey> 
            <slcs:CASerial>./CAserial</slcs:CASerial> 
        </Service> 
    </Chain> 
\end{verbatim}

    SLCS service and SP service should together be configured by using the Plexer. SP service is directly based on http and SLCS is directly based on soap. As shown in the above table, a plexer dispatches the message flow (outgoing from http.service) into two SP service (saml.sp) and SOAP service (soap.service).

    Since the client authentication is switched off, it is not necessary to configure $<$CACertificatePath$>$ or $<$CACertificateDir$>$ for TLS MCC.

    SP service needs to be configured with $<$MetaDataLocation$>$ and $<$ServiceProviderName$>$.

    SLCS service should specifically include the CA credential (certificate and key file) and the serial number file (which includes the serial number of each signed certificate), so SLCS service needs to be configured with $<$CACertificate$>$ $<$CAKey$>$ and $<$CASerial$>$. Therefore, the service administrator should firstly create a CA credential (note for interoperability purpose in the production grid deployment, the CA credential should be trusted by others).

Note: When deploying the SLCS service, service administrator should deploy a dedicated IdP (Identity Provider) which can be assigned as authentication/attribute authority, or the users should have already had his own IdP. And the IdP information (authentication URL and attribute authority URL) should have already been included into the metadata of above configuration (e.g. test\_metadata.xml). On the other hand, the SP (SP service) information (assertion consuming URL) should also have been included into the metadata of the IdP.

    After the whole process, user only needs to authentication with his existing IdP and can get back the x509 credential.

    Shibboleth IdP (http://shibboleth.internet2.edu) is used for the current solution of IdP, since it is widely deployed for other AAI (Authentication and Authority Infrastructure). There is a test IdP deployed in https://squark.uio.no:8443 (idpname: https://squark.uio.no/idp/shibboleth, username: staff, password: 123456), you can use the SP from https://sp.testshib.org/ to test the validity of the test IdP. And also you can installed your own IdP which is supposed to authenticates the users from your own organization.

    SLCS client generates  a X.509 certificate  request, launches a Web Service request which includes the certificate request; SLCS service then gets the certificate request, composes a distinguished name (DN), issues a certificate (short lived, 12 hours by default) with the SAML attribute (from the SAML2.0 SSO profile) as the X.509 certificate extension, and puts the certificate in to the Web Service response; SLCS client get the response and stores the X.509 certificate into local repository.

The CLI for the SLCS client is like this:

\textit{./arcslcs -S https://127.0.0.1:60000/slcs -I https://idp.testshib.org/idp/shibboleth -U myself -P myself}

    Since the lifetime of the short lived credential is normally short, it is not a must to protect the private key by a pass phrase. As illustrated in steps (a) and (b) in Figure 10, if the private key is not protected through the Web Service client, the user can use the X.509 certificate to access Grid Service or Web Service from any kind of middleware. If the private key is protected, she can use the X.509 certificate to generate a proxy certificate (by using a command-line interface utility such as grid-proxy-init , voms-proxy-init, or arcproxy ), and then use the proxy certificate to access a Grid/Web Service.

    Composing the distinguished name (DN) for the certificate is a critical issue for the SLCS service. Since the Shibboleth Identity Provider uses the eduPerson schema for the definition of $<$saml:Attribute$>$ in $<$saml:AttributeStatement$>$, we pick the relatively distinguishable attribute \textit{``eduPersonPrincipalName''} for the DN. A typical eduPersonPrincipalName value could be \textit{alice@example.org}, then the DN is

\textit{``/O=knowarc/OU=example.org/CN=alice''}.

    The obvious benefit of the SLCS service is that: If a user passes the authentication to her home Identity Provider, she can get the X.509 credential anywhere simply by running the SLCS client command together with providing her username and password to this home IdP, and then access the Grid system conveniently.

\subsection{The Usage of SLCS client} % (fold)
\label{subsec:slcs_client}

    On the SLCS client side, there is a client utility called ``arcslcs''.  The command option for arcslcs is as following:
Application Options:

  -S, --url=url                URL of SLCS service

  -I, --idp=string             IdP name

  -U, --user=string            User account to IdP

  -P, --password=string        Password for user account to IdP

  -Z, --keysize=number         Key size of the private key (512, 1024, 2048)

  -K, --keypass=passphrase     Private key passphrase

  -L, --lifetime=period        Lifetime of the certificate, start with current time, hour as unit

  -D, --storedir=directory     Store directory for key and signed certificate

  -c, --conffile=filename      configuration file (default ~/.arc/client.xml)

An example is:

\textit{./arcslcs -S https://127.0.0.1:60000/slcs -I https://squark.uio.no/idp/shibboleth -U root -P aa1122 -D   ~/arc-0.9/src/clients/credentials -c client.xml}

    Note user should input the ``IdP name'' is the corresponding Identity Provider (one of the participants of SAML2 SSO profile) name to which the user would authenticate against by using its username/password credential. And the name is stored inside the metadata on both SP service and IdP provider, and it is used by SP service to get the authentication URL for username/password based authentication.

    And the ``user'' and ``password'' is the credential which will be used to authentication against IdP.
The ``conffile'' can be another option for all of the above options, see the following as an example:

\begin{verbatim}
<ArcConfig> 
 <!-- change the paths below to the location of your user certs --> 
 <!-- here only trusted certificates are set. Because this arcslcs is 
 not supposed to have x509 credential before the samlsso profile has been 
 (by using client's username/password credential) passed and the SLCS 
 service succeeded to repond this client with the signed x509 credential. 
 On the SLCS service's configuration: ClientAuthn should be off.--> 
 <!--KeyPath>./testkey-nopass.pem</KeyPath--> 
 <!--CertificatePath>./testcert.pem</CertificatePath--> 
 <CACertificatesDir>./certificates</CACertificatesDir> 
 <SLCSURL>https://127.0.0.1:60000/slcs</SLCSURL> 
 <IdPname>https://squark.uio.no/idp/shibboleth</IdPname> 
 <Username>root</Username> 
 <Password>aa1122</Password> 
 <Keysize>1024</Keysize> 
 <Keypass>123456</Keypass> 
 <CertLifetime>24</CertLifetime> 
 <StoreDir>./</StoreDir> 
 <Debug>INFO</Debug> 
</ArcConfig> 
\end{verbatim}

    There is a temporary Identity Provider (IdP) deployed on squark.uio.no for test with the following test username and password: staff, researcher, librarian, binduser ; with the same password ``123456''

\textit{./arcslcs -S https://127.0.0.1:60000/slcs -I https://squark.uio.no/idp/shibboleth -U staff -P 123456 -D /home/wzqiang/arc-0.9/src/clients/credentials -c client.xml}

    The short-lived credential issued by SLCS service will include the SAML assertion as the extension of X.509 certificate as a proof of passing SAML2 SSO profile.

%section slcs (end)

\section{X.509 Credential Delegation Service} % (fold)
\label{sec:delegation_service}
Besides the delegation interface described in Section 6.4, which provides interface for X.509 credential delegation. A dedicated Web Service is also developed for credential delegation, with the intention of providing a general approach for using delegation by changing the configuration information, rather than a specific approach which requires directly changing the code.

    As shown in Figure 11, there is a specific WS client and Web Service for processing delegation: delegation client, and delegation service. The WSDL (Web Service Description Language) of the delegation service has three main operations, the first one for processing the delegation initiation, the second one for storing the signed proxy certificate, and the last one for processing acquisition of the delegated credential. The sequential diagram of the delegation is detailed in Figure 12.

    For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure 11), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential. For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure 11), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.

    Since each WS client accesses the target service (delegation service or other general services) on behalf of the user, the trust relationship between the user’s certificates (X.509 certificate and proxy certificate) and services is required, while the trust relationship between service’s certificate and another service’s certificate is not required.

Below is the configuration of the delegation service, and the ``echo'' service which contacts the delegation service to acquire the delegated certificate.

The configuration of the delegation service which is located at ``glueball.uio.no''.

\begin{verbatim}
    <Chain>
        <Component name="tcp.service" id="tcp"> 
            <next id="tls"/> 
            <tcp:Listen><tcp:Port>60005</tcp:Port></tcp:Listen> 
        </Component> 
        <Component name="tls.service" id="tls"> 
            <next id="http"/> 
            <tls:KeyPath>./testkey-nopass.pem</tls:KeyPath> 
            <tls:CertificatePath>./testcert.pem</tls:CertificatePath> 
            <tls:CACertificatePath>./cacert.pem</tls:CACertificatePath> 
              <tls:CACertificatesDir>./certificates</tls:CACertificatesDir> 
        </Component> 
        <Component name="http.service" id="http"> 
            <next id="soap">POST</next> 
        </Component> 
        <Component name="soap.service" id="soap"> 
            <next id="delegation"/> 
        </Component> 
        <Service name="delegation.service" id="delegation"> 
            <next id="delegation"/> 
            <deleg:CACertificatesDir>./certificates</deleg:CACertificatesDir> 
            <deleg:CACertificatePath>./testcacert.pem</deleg:CACertificatePath> 
        </Service> 
    </Chain> 
\end{verbatim}

The configuration of the ``echo'' service (e.g., located at ``selectron.uio.no'') which contacts the delegation service that is located at ``glueball.uio.no''.

\begin{verbatim}
    <Chain>
        <Component name="tcp.service" id="tcp"> 
            <next id="tls"/> 
            <tcp:Listen><tcp:Port>60000</tcp:Port></tcp:Listen> 
        </Component> 
        <Component name="tls.service" id="tls"> 
            <next id="http"/> 
            <tls:KeyPath>./testkey-nopass.pem</tls:KeyPath> 
            <tls:CertificatePath>./testcert.pem</tls:CertificatePath> 
            <tls:CACertificatePath>./cacert.pem</tls:CACertificatePath> 
              <tls:CACertificatesDir>./certificates</tls:CACertificatesDir> 
        </Component> 
        <Component name="http.service" id="http"> 
            <next id="soap">POST</next> 
        </Component> 
        <Component name="soap.service" id="soap"> 
            <next id="echo"/> 
        </Component> 
        <Service name="echo" id="echo"> 
            <SecHandler name='delegation.handler' id='delegation' event='incoming'> 
              <deleg:Type>x509</deleg:Type> 
              <deleg:Role>delegatee</deleg:Role> 
              <deleg:DelegationServiceEndpoint>https://glueball.uio.no:60000/delegation</deleg:DelegationServiceEndpoint> 
              <!--The credential and trusted certificates which will be used by this security handler to contact with delegation service -->
              <deleg:KeyPath>./testkey-nopass.pem</deleg:KeyPath> 
              <deleg:CertificatePath>./testcert.pem</deleg:CertificatePath> 
              <deleg:CACertificatesDir>./certificates</deleg:CACertificatesDir> 
              <deleg:CACertificatePath>./testcacert.pem</deleg:CACertificatePath> 
            </SecHandler> 
            <next id="echo"/> 
            <echo:prefix>[ </echo:prefix> 
            <echo:suffix> ]</echo:suffix> 
        </Service> 
    </Chain> 
\end{verbatim}

On the client side of ``echo'' service, the client developer is supposed to add the following piece of configuration into the client configuration information, by adding one line of code:

\begin{verbatim}
client->AddSecHandler(sechanlder_nd, "arcshc");
\end{verbatim}

Some explanation about the configuration is described.

\begin{verbatim}
     Arc::XMLNode sechanlder_nd("\ 
        <SecHandler name='delegation.handler' id='delegation' event='outgoing'>\ 
          <Type>x509</Type>\ 
          <Role>delegator</Role>\ 
          <!--The endpoint of delegation service, to which this security handler will delegate the credential-->\
          <DelegationServiceEndpoint>https://glueball.uio.no:60005/delegation</DelegationServiceEndpoint>\ 
          <!--The endpoint of the peer service, to which this WS client will access-->\
          <PeerServiceEndpoint>https://selectron.uio.no:60000/echo</PeerServiceEndpoint>\ 
         <!--The key and certificate of the credential, which is be delegated to the delegation service-->\
          <KeyPath>./testkey-nopass.pem</KeyPath>\ 
          <CertificatePath>./testcert.pem</CertificatePath>\ 
         <!--Another option of defining the location of credential-->\
          <!--ProxyPath>/tmp/5612d050.pem</ProxyPath-->\ 
         <!--Yet another option of defining how to get the credential; this option is useful for the service which has already acquired the delegated credential (this delegated credential is stored under /tmp/hash_value.pem) from the delegation service, yet still needs to delegate this credential to another delegation service; the DelegationCredentialIdentity can be parsed from the tls connection from client side, and then used to calculate the hash value to get the delegated credential from local repository-->\
          <!--DelegationCredIdentity>/O=KnowARC/OU=UiO/CN=squark.uio.no</DelegationCredIdentity-->\ 
          <CACertificatePath>./cacert.pem</CACertificatePath>\ 
          <CACertificatesDir>./certificates</CACertificatesDir>\ 
        </SecHandler>"); 
\end{verbatim}

%section delegation_service (end)

\end{document}

### 2009-08-10 doc/tech_doc/a-rex/arex_tech_doc.tex
Reorganizing doc catalogue and enabling make doc
bring up to date with new features in code

--- 2009-04-20 doc/a-rex/arex_tech_doc.tex
@@
 
-Before giving access to a file already in the cache, the A-REX contacts
-the initial file source to check if the user has read permission on
-the file. Also file creation or validity times from the original source
-are checked to make sure the cached file is fresh enough. If it is
-impossible to obtain creation and invalidation times for the file,
-it is invalidated 24 hours after download.
+Before giving access to a file already in the cache, the A-REX
+contacts the initial file source to check if the user has read
+permission on the file. In order to prevent repeated checks on source
+files, this authentication information is cached for a limited
+time. On passing the check for a cached file, the user's DN is stored
+in the \emph{.meta} file, with an expiry time equivalent to the
+lifetime remaining for the user's proxy certificate. This means that
+the permission check is not performed for this user for this file
+until this time is up (usually several hours). File creation and
+validity times from the original source are also checked to make sure
+the cached file is fresh enough. If the modification time of the
+source is later than that of the cached file, the file will be
+downloaded again. The file will also be downloaded again if the
+modification date of the source is not available, as it is assumed the
+cache file is out of date. These checks are not performed if the DN is
+cached and is still valid.
 
@@
 about accepted jobs. It must have full permissions there.\\
-A subdirectory called \textit{log} is created there. It is used to
+A subdirectory called \textit{logs} is created there. It is used to
 accumulate information about started and finished jobs. This information
@@
     {[}cert\_file {[}ca\_dir]]]} -- specifies the credentials for
-  accessing the authentication service.
+  accessing the accounting service.
 \item \textbf{\textit{jobreport\_options}}\textit{={[}options]}
@@
 Missing value or -1 means no limit.
-\item \textbf{\textit{maxlod}}\textit{={[}max\_frontend\_jobs {[}emergency\_frontend\_jobs
+\item \textbf{\textit{maxload}}\textit{={[}max\_frontend\_jobs {[}emergency\_frontend\_jobs
 {[}max\_transferred\_files]]]} -- specifies maximum load caused by
@@
 the JD).
-\item \textbf{\textit{session}}\textit{=path} -- specifies path to the directory
+\item \textbf{\textit{sessiondir}}\textit{=path} -- specifies path to the directory
 in which the SD is created. If the path is {*} the default one is
-used -- \textit{\$HOME/.jobs}. In new configuration file this command
-is called \textbf{\textit{sessiondir}}.
+used -- \textit{\$HOME/.jobs}.
 \item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} - specifies
@@
 files are not soft-linked, nor are per-job links created, but files
-are copied to the session directory. In old configuration file this
-command is called \textbf{\textit{cache}}.
+are copied to the session directory.
 \item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} - specifies

### 2009-08-10 doc/manuals/user_interface/ui-data.tex
Starting cleaning up documentation
clean up arcls/cp/rm doc

--- 2008-08-14 doc/client/arcdata.tex
@@
     \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level: 0 = some, 1 = more, 2 = a lot \\
+    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
     \texttt{-l} &  & detailed listing\\
@@
     be downloaded and\\ & & temporary cached locations\\
+    \texttt{-m} && display all available metadata\\
     Arguments:&&\\
@@
 \begin{verbatim}
-    ngls rls://rc.host:38203/logical_file_name
-    ngls -l gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535
-    ngls -L srm://grid.uio.no:58000/srm/managerv1/johndoe/log2
+    arcls rls://rc.host:38203/logical_file_name
+    arcls -l gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535
+    arcls -L srm://grid.uio.no:58000/srm/managerv1/johndoe/log2
 \end{verbatim}
@@
 
-The \texttt{arccp}\footnote{In ARC $\leq$ 0.5.28 was called
-  \texttt{ngcopy}} \index{arccp}\index{commands:arccp} is a powerful
-tool to copy files over the Grid. It is a part of the Grid Manager,
+The \texttt{arccp} \index{arccp}\index{commands:arccp} is a powerful
+tool to copy files over the Grid. It is a part of the A-REX,
 but can be used by the User Interface as well.
@@
 \end{shaded}
-\versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
+\versions{ARC 0.9}
 \begin{longtable}{llp{8cm}}
@@
     \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level: from -3 = quiet to 3 = verbose\\
+    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
     \texttt{-y} & \textit{cache\_path} & path to local cache (use to put file into cache)\\
-    \texttt{-Y} & \textit{cache\_data\_path} & path for cache data (if different from -y)\\
     \texttt{-p} && use passive transfer (does not work if secure is on, default if secure is not requested)\\
@@
 
-\begin{framed}
-   Since the job ID is in fact a \verb#gsiftp://# URL of the job
-   top directory, you can use \texttt{arccp} to copy files from the
-   job directory at any time.
-\end{framed}
-
 Usage examples of \texttt{arccp} are:
@@
 \begin{verbatim}
-    ngcp gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535/job.out \
+    arccp gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535/job.out \
               file:///home/myname/job2.out
-    ngcp gsiftp://aftpexp.bnl.gov;threads=10/rep/my.file \
+    arccp gsiftp://aftpexp.bnl.gov;threads=10/rep/my.file \
               rc://;threads=4@grid.uio.no/lc=Collection,rc=Catalog/zebra4.f
-    ngcp http://www.nordugrid.org/data/somefile gsiftp://hathi.hep.lu.se/data/
+    arccp http://www.nordugrid.org/data/somefile gsiftp://hathi.hep.lu.se/data/
 \end{verbatim}
@@
 
-The \texttt{arcrm}\footnote{In ARC $\leq$ 0.5.28 was called
-  \texttt{arcremove}} \index{arcrm}\index{commands:arcrm}
+The \texttt{arcrm} \index{arcrm}\index{commands:arcrm}
 command allows users to erase files at any location specified by a
@@
 \begin{shaded}
-   \uicommand{ngrm [options] $<$source$>$}
+   \uicommand{arcrm [options] $<$source$>$}
 \end{shaded}
-\versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
+\versions{ARC 0.9}
 \begin{longtable}{llp{8cm}}
@@
     \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level: 0 = some, 1 = more, 2 = a lot \\
+    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
     \texttt{-c} & &continue with meta-data even if it failed to delete real file\\
-    \texttt{-C} & \textit{cache\_data\_path} & store cached data\\
     Arguments:&&\\
@@
 \begin{framed}
-   A convenient use for \texttt{ngarc} is to erase the files in a data
+   A convenient use for \texttt{arcrm} is to erase the files in a data
    indexing catalog (RC, RLS or such), as it will not only remove the
@@
 \begin{verbatim}
-    ngrm rc://grid.uio.no/lc=Collection,rc=Catalog/badfile#\\
+    arcrm rc://grid.uio.no/lc=Collection,rc=Catalog/badfile
 \end{verbatim}
@@
    Options:&&\\
-    \texttt{-d, -debug} & \textit{debuglevel} &debug level: 0 = some, 1 = more, 2 = a lot \\
+    \texttt{-d, -debug} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
     \texttt{-v} && print version information\\
@@
 \begin{verbatim}
-    ngacl get gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename
-    ngacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename $<$ myacl
+    arcacl get gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename
+    arcacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename $<$ myacl
 \end{verbatim}

### 2009-08-12 doc/tech_doc/sec/arc-security-documentation.tex
more change for security documentation

@@
-\documentclass{book}
+%\documentclass{book}
+\documentclass{article}                            %for shorter notes
 \usepackage{graphicx}                              %for PNG images (pdflatex)
+\graphicspath{{./figures/}}
 \usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
@@
 \tableofcontents                          %Comment if use article style
+% Yet we are not using documentclass--book (article instead) here,
+% we still need to use TOC.
 \newpage
-
 \renewcommand{\thefootnote}{\arabic{footnote}}
 
-\chapter{Technical Description} % (fold)
-\label{cha:tech_description}
-
-\section{Introduction} % (fold)
+%\chapter{Technical Description} % (fold)
+%\label{cha:tech_description}
+
+\section{Introduction}
 \label{sec:introduction}
@@
 In the implementation of the ARC1, there is a Service Container – the Hosting Environment Daemon (HED) (D1.2-2, [2]) which provides a hosting place for various services in application level, as well as a flexible and efficient communication mechanism.
+
 HED contains a framework for implementing and enforcing authentication and authorization.
@@
 If any of them fails, message processing fails as well.
+
+\begin{figure}[ht]
+\centering{{{\includegraphics[width=1.0\textwidth]{MCC_Component.pdf}}}
+\caption{\label{fig:messageinMCC}\textit{There are usually two chains of SecHandlers inside the MCC or service.
+Each SecHandler will parse the Security Attributes which are generated by the upstream MCC/services or probably upstream SecHandlers in the same or other MCC/Service, and do message processing or authenticate or authorize the incoming/outgoing message based on the collected information.
+The SecHandler can also change the payload and attributes of Messsage itself.
+For example, the Username-Token SecHandler will insert the WSS Username Token [3] into header part of SOAP message.
+The PDPs are called by the SecHandlers and are supposed to make authorization decision.
+Here the two local PDP and one remote PDP service is just for demonstration, and one or any number of PDPs can be configured under each SecHandler.
+}} }
+\end{figure}
+
 Each SecHandler is configured inside same configuration file used for configuring whole chain of MCCs.
@@
 The currently implemented sub-modules used by some SecHandlers are Policy Decision Point (PDP) components such as Arc PDP which can process ARC specific Request and Policy documents.
-Figure 1 gives the structure of a MCC/Service, and the message sequence inside it.
-And Figure 2 shows the configuration of SecHandler components for an example ``Echo'' service.
+Figure \ref{fig:messageinMCC} gives the structure of a MCC/Service, and the message sequence inside it.
+
+The configuration of SecHandler components for an example ``Echo'' service is shown as below.
+Example Echo service is configured to use two SecHandlers, both responsible for authorization.
+First SecHandler uses the identity of client extracted from the incoming message to map it into local identity like local Linux username.
+In this case all clients are mapped to local account “test”.
+The second one uses two PDPs: one will compose ARC specific authorization request based on the Security Attributes collected from “incoming” message and evaluate it against the ARC specific authorization policy which is defined in local file “policy.xml”; the other will compare the X509 identity of client extracted from the incoming message against list of identities stored locally.
+
+\begin{verbatim}
+<Service name="echo" id="echo">
+    <SecHandler name="identity.map" id="map" event="incoming">
+        <PDP name="allow.pdp"><LocalName>test</LocalName></PDP>
+    </SecHandler>
+    <SecHandler name="arc.authz" id="authz" event="incoming">
+        <PDP name="arc.pdp">
+            <PolicyStore>
+                <Location type="file">policy.xml</Location>
+                <!-- other policy location-->
+            </PolicyStore>
+        </PDP>
+        <PDP name="simplelist.pdp" location="pemittedlist.txt"/>
+    </SecHandler>
+</Service>
+\end{verbatim}
 
@@
 When one component (MCC or service) is loaded according to the configuration information, the SecHandler under the component and the plug-ins like PDP which are attached to the SecHandler will be loaded as well.
-There is one simple interface (see Figure 3) defined in class SecHandler, which will be called by the containing MCC/Service once there is message (incoming or outgoing) need to be processed.
+
+There is one simple interface (see below) defined in class SecHandler, which will be called by the containing MCC/Service once there is message (incoming or outgoing) need to be processed.
+
+\begin{verbatim}
+class SecHandler {
+   public:
+    SecHandler(Arc::Config*) {};
+    virtual ~SecHandler() {};
+    virtual bool Handle(Arc::Message *msg) = 0;
+  };
+\end{verbatim}
+
+Class \textit{SecHandler} is an abstract class which includes a general interface called \textit{Handle} which uses Message object as argument.
+Any security handler implementation should inherit class \textit{SecHandler} and implement the interface according to the actual functionality.
+The interface only return simple \textit{Boolean} value, and any useful information generated during the calling of this interface should be put into the security attribute of the message, or put into the payload itself.
 
@@
 
-Figure 4 shows the definition of abstract class PDP.
+Below shows the definition of abstract class PDP.
 The implementation could be some function which implements the interface by composing the policy evaluation request, evaluating this request against some policy, and returning the evaluation result, or just by composing the policy evaluation request, invoking some remote policy decision web service and getting back the evaluation result.
+
+\begin{verbatim}
+class PDP {
+   public:
+    PDP(Arc::Config* cfg) { };
+    virtual ~PDP() {};
+    virtual bool isPermitted(Arc::Message *msg) = 0;
+  };
+\end{verbatim}
+
+Class \textit{PDP} is an abstract class which includes a general interface called \textit{isPermitted} which uses Message object as argument.
+Any policy decision point implementation should inherit class PDP and implement the interface according to the actual functionality.
+The interface only return simple \textit{Boolean} value, and any useful information generated during the calling of this interface should be put into the security attribute of the message, or put into the payload itself.
 
@@
 
+    \item xacml.pdp – XACML PDP
+The XACML PDP will organize the security attributes into the XACML authorization request (see: http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-context-schema-os.xsd), call the policy evaluator to evaluate the request against the policy (which is in XACML format: http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-policy-schema-os.xsd) repository, and get back the evaluation result.
+
     \item delegation.pdp – Delegation PDP
@@
     \item pdpservice.invoker – PDP Service Invoker
-The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the same functionality as Arc PDP, except that the evaluation request and response are carried by SOAP message.
+The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the same functionality as Arc PDP or XACML PDP, except that the evaluation request and response are carried by SOAP message.
 The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation engine can be accessed remotely and maintained centrally.
@@
 
-Figure 5 shows the UML class diagram about the policy evaluation engine. It shows all classes and relations simultaneously for getting the overall picture. 
-The Evaluator class is the key class for policy evaluation. It accepts request evaluates it against loaded policy and returns evaluation response.
-
-Three abstract factories - FnFactory, AlgFactory, AttributeFactory - are responsible for creating the Function, CombiningAlg and AttributeValue objects correspondingly.
-The classes inherited from CombiningAlg class take care of implementing various combining algorithms which define relations between \textit{$<$Rule/$>$} elements in policy.
-The AttributeValue type of classes are used for processing different types of \textit{$<$Attribute/$>$} and similar elements.
-The Function classes take care of comparing \textit{$<$Attribute/$>$} elements of request and policy.
-
-The Policy class parses \textit{$<$Policy/$>$} or \textit{$<$Rule/$>$} elements and creates  CombingAlg objects according to the \textit{$<$RuleCombiningAlg/$>$} attribute of \textit{$<$Policy/$>$}, Function objects according to the \textit{$<$Function/$>$} attribute of \textit{$<$Attribute/$>$} and AttributeValue objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}.
+\begin{figure}[ht]
+\centering{{{\includegraphics[width=1.0\textwidth]{Evaluator_ARC.pdf}}}
+\caption{\label{fig:policyengine_ARC}The UML class diagram of the classes inside policy evaluation engine that support ARC policy} }
+\end{figure}
+
+\begin{figure}[ht]
+\centering{{{\includegraphics[width=1.0\textwidth]{Evaluator_XACML.pdf}}}
+\caption{\label{fig:policyengine_XACML}The UML class diagram of the classes inside policy evaluation engine that support XACML policy} }
+\end{figure}
+
+Figure \ref{fig:policyengine_ARC} and \ref{fig:policyengine_XACML} respectively show the UML class diagram about the policy evaluation engine for ARC policy and XACML policy.
+They show all classes and relations simultaneously for getting the overall picture.
+
+The \textbf{Evaluator} class is the key class for policy evaluation. It accepts request evaluates it against loaded policy and returns evaluation response.
+
+Three abstract factories - \textit{FnFactory}, \textit{AlgFactory}, \textit{AttributeFactory} - are responsible for creating the \textit{Function}, \textit{CombiningAlg} and \textit{AttributeValue} objects correspondingly.
+The classes inherited from \textit{CombiningAlg} class take care of implementing various combining algorithms which define relations between \textit{$<$Rule/$>$} elements in policy.
+The \textit{AttributeValue} type of classes are used for processing different types of \textit{$<$Attribute/$>$} and similar elements.
+The \textit{Function} classes take care of comparing \textit{$<$Attribute/$>$} elements of request and policy.
+
+The \textit{Policy} class parses \textit{$<$Policy/$>$} or \textit{$<$Rule/$>$} elements and creates \textit{CombingAlg} objects according to the \textit{$<$RuleCombiningAlg/$>$} attribute of \textit{$<$Policy/$>$}, \textit{Function} objects according to the \textit{$<$Function/$>$} attribute of \textit{$<$Attribute/$>$} and \textit{AttributeValue} objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}.
 Those objects will be used when evaluating the request.
 
-The Request class is responsible for parsing \textit{$<$Request/$>$} element and creates corresponding AttributeValue objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}.
-When evaluating, each AttributeValue in request will be evaluated against corresponding AttributeValue in the policy by using relevant Function.
-
-Due to extensible architecture of code it is relatively easy to add support for new types of  AttributeValue, Function and CombingAlg objects in this way supporting various types of XML based policy languages.
-
-\subsection{Schemas for policy evaluation engine} % (fold)
-\label{subsec:policy_schema}
+The \textit{Request} class is responsible for parsing \textit{$<$Request/$>$} element and creates corresponding \textit{AttributeValue} objects according to the \textit{$<$Type/$>$} attribute of \textit{$<$Attribute/$>$}.
+When evaluating, each \textit{AttributeValue} in request will be evaluated against corresponding \textit{AttributeValue} in the policy by using relevant \textit{Function}.
+
+Due to extensible architecture of code it is relatively easy to add support for new types of \textit{AttributeValue}, \textit{Function} and \textit{CombingAlg} objects in this way supporting various types of XML based policy languages.
+
+\subsection{Policy evaluation engine --- Support of ARC policy and request} % (fold)
+\label{subsec:policy_engine_arc}
+
+\subsubsection{Schemas for ARC policy and request} % (fold)
+\label{subsubsec:arc_policy_schema}
 The schema for ARC Policy is available at
@@
 
-The hierarchy tree of ARC Policy is show below (numbers show multiplicity of elements)
+The hierarchy tree of ARC Policy is shown below (numbers show multiplicity of elements)
 
@@
 
-\subsection{Basic Elements of Policy} % (fold)
-\label{subsec:poicy_element}
-There are 2 basic objects - "policy" and "request". There is 1 main actor - Evaluator. Curretly there are two types of elements in policy: Policy and Rule. Policy element is made of Rule elements.
-Evaluator matches request to policy and produces one of 4 following results:
+\subsubsection{Basic Elements of Policy} % (fold)
+\label{subsubsec:poicy_element}
+There are 2 basic objects - ``policy'' and ``request''.
+There is 1 main actor - Evaluator.
+Curretly there are two types of elements in policy: \textit{Policy} and \textit{Rule}.
+\textit{Policy} element is made of \textit{Rule} elements.
+\textit{Evaluator} matches request to policy and produces one of 4 following results:
 
@@
 
-\subsection{Policy Matching} % (fold)
-\label{subsec:poicy_matching}
+\subsubsection{Policy Matching} % (fold)
+\label{subsubsec:poicy_matching}
 Policy is made of Rule elements.
@@
 
-\subsection{Request Structure} % (fold)
-\label{subsec:request_structure}
+\subsubsection{Request Structure} % (fold)
+\label{subsubsec:request_structure}
 
@@
 
-\subsection{Rule Composition and Matching} % (fold)
-\label{subsec:rule_comp_match}
+\subsubsection{Rule Composition and Matching} % (fold)
+\label{subsubsec:rule_comp_match}
 
@@
 
-\subsection{Rule Elements Matching} % (fold)
-\label{subsec:rule_element_match}
+\subsubsection{Rule Elements Matching} % (fold)
+\label{subsubsec:rule_element_match}
 For elements without attributes those elements have:
@@
 
+\subsection{Policy evaluation engine --- Support of XACML policy and request} % (fold)
+\label{subsec:policy_engine_xacml}
+Currently, XACML specification is partially supported/implemented in ARC. More specifically, except the \textit{$<$Obligation$>$} element, other elements are supported.
+
+http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-core-spec-os.pdf
+
+http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-policy-schema-os.xsd
+
+http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-context-schema-os.xsd
+
 \subsection{Interface for using the policy evaluation engine} % (fold)
@@
 
-For making usage of policy evaluation engine more convenient basic Evaluator class is complemented by additional interfaces.
+For making usage of policy evaluation engine more convenient basic \textit{Evaluator} class is complemented by additional interfaces.
 Below are examples of steps needed to carry out policy evaluation and corresponding helper interfaces.
@@
     std::string evaluator = "arc.evaluator"; 
+    // If xacml evaluation engine is used, 
+    // std::string evaluator = "xacml.evaluator";
+
     eval = eval_loader.getEvaluator(evaluator);
@@
     std::string policyclassname = "arc.policy"; 
+    // If xacml policy is used, 
+    // std::string policyclassname = "xacml.policy";
+
     // Define source from which policy to be taken
@@
     std::string requestclassname = "arc.request";
+    // If xacml request is used, 
+    // std::string requestclassname = "xacml.request";
+
     // Define source from which request to be taken
@@
 
-The Evalutor::evaluate() method can also be feed up with both Policy/Request objects and their sources in any combination.
-See example code at http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/testinterface.cpp for more details about usage of the interface.
+The Evalutor::evaluate() method can also be feed up with both \textit{Policy}/\textit{Request} objects and their sources in any combination.
+See example code at http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/testinterface\_arc.cpp or http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/testinterface\_xacml.cpp for more details about usage of the interface.
 
@@
 
-\section{Policy Decision Service} % (fold)
+\section{Policy Decision Service (Charon Service)} % (fold)
 \label{sec:policy_decision_service}
-Policy decision service is a service implementation which contains the functionality of ArcPDP.
+Policy decision service is a service implementation which configurably contains the functionality of ArcPDP or XACMLPDP.
 It will accept the soap request containing policy decision request and return soap response containing policy decision response.
@@
 
- http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl.
+ http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/charon/charon.wsdl
 
@@
 
-\section{Security Attributes} % (fold)
+\section{Security Attributes -- how to compose policy decision request to policy evaluation engine} % (fold)
 \label{sec:sec_attributes}
@@
 Security Attributes represent security related information inside HED framework and store information representing various aspects needed to perform authorization decison - identity of client, requested action, targeted resource, constraint policies.
-Each kind of Security Attribute is represented by own class inherited from parent SecAttr class <arc/message/SecAttr.h>.
-Each Security Attribute stores it's information in internal format and is capable to export it to one of predefined formats using Export() method.
+
+Each kind of Security Attribute is represented by own class inherited from parent SecAttr class $<$arc/message/SecAttr.h$>$.
+
+Each Security Attribute stores it's information in internal format and is capable to export it to one of predefined formats using \textit{Export()} method.
 Currently only supported format is ARC Policy/Request XML document described in Section 8.1and 8.2
-Collectors of Security Attributes instantiate corresponding classes and link them to Secuirity Attributes containers - MessageAuth <arc/message/MessageAuth.h> and MessageAuthContext <arc/message/Message.h> storing collected attributes per request and per session correspondingly.
+Collectors of Security Attributes instantiate corresponding classes and link them to Secuirity Attributes containers - MessageAuth $<$arc/message/MessageAuth.h$>$ and MessageAuthContext $<$arc/message/Message.h$>$ storing collected attributes per request and per session correspondingly.
 Each attribute is assigned a name.
@@
 See Section 5.2for available Collectors and corresponding Security Attributes.
+
 Note for service developers: Services may implement own authorization algorithms.
@@
 \label{subsec:delegation_interface}
+
+\begin{figure}
+\centering{{{\includegraphics[width=0.9\textwidth]{delegation_flow.pdf}}}
+\caption{\label{fig:delegation_flow}The flow diagram of delegation procedure with multiple second step and session expiration} }
+\end{figure}
+
+\begin{figure}
+\centering{{{\includegraphics[width=0.9\textwidth]{delegation_flow1.pdf}}}
+\caption{\label{fig:delegation_flow1}The flow diagram of delegation procedure with certificate transfered as payload of service specific message} }
+\end{figure}
+
 Delegation interface in the ARC1 is implemented using Web Service approach.
 Each ARC1 service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials.
-Here is how delegation procedure works (also shown in figures 6 and 7) :
+Here is how delegation procedure works (also shown in Figure \ref{fig:delegation_flow} and Figure \ref{fig:delegation_flow1}) :
 
@@
 For the outgoing message this SecHandler creates X509Token in SOAP header. SOAP message body is encrypted and signed.
+
+\subsection{SAMLToken SecHandler} % (fold)
+\label{subsec:saml_token}
 
@@
             <tls:CertificatePath>./testcert.pem</tls:CertificatePath>
-           <!--Since the client side's certificate is not required to exist, the trusted CA certificates are not needed here-->
+           <!--Since the client side's certificate is not required to exist, 
+            the trusted CA certificates are not needed here-->
             <!--tls:CACertificatePath>./cacert.pem</tls:CACertificatePath--> 
@@
             <next id="echo"/> 
-           <!--The security handler used for doing, e.g., access control based on the SAML assertion from the SAML2.0 SSO profile; currently it is just a prototype implementation-->
+           <!--The security handler used for doing, e.g., access control based on 
+           the SAML assertion from the SAML2.0 SSO profile; currently it is just a 
+            prototype implementation-->
             <SecHandler name="saml2ssoassertionconsumer.handler" id="saml2ssosp" event="incoming"/> 
@@
         <Service name='saml.sp' id='samlsp'>
-            <!--Metadata that is according to SAML metadata standard. urn:oasis:names:tc:SAML:2.0:metadata-->
+            <!--Metadata that is according to SAML metadata standard. 
+             urn:oasis:names:tc:SAML:2.0:metadata-->
             <MetaDataLocation>./test_metadata.xml</MetaDataLocation> 
@@
             <ServiceProviderName>https://squark.uio.no/shibboleth-sp</ServiceProviderName>
-           <!--The certificate and key file which will be used by Service Provide to sign the AuthnRequest-->
+            <!--The certificate and key file which will be used by Service Provide 
+             to sign the AuthnRequest-->
             <KeyPath>./testkey-nopass.pem</KeyPath> 
@@
               <deleg:Role>delegatee</deleg:Role> 
-              <deleg:DelegationServiceEndpoint>https://glueball.uio.no:60000/delegation</deleg:DelegationServiceEndpoint> 
-              <!--The credential and trusted certificates which will be used by this security handler to contact with delegation service -->
+              <deleg:DelegationServiceEndpoint>
+                https://glueball.uio.no:60000/delegation
+              </deleg:DelegationServiceEndpoint> 
+              <!--The credential and trusted certificates which will be used by this 
+              security handler to contact with delegation service -->
               <deleg:KeyPath>./testkey-nopass.pem</deleg:KeyPath> 
@@
           <Role>delegator</Role>\ 
-          <!--The endpoint of delegation service, to which this security handler will delegate the credential-->\
-          <DelegationServiceEndpoint>https://glueball.uio.no:60005/delegation</DelegationServiceEndpoint>\ 
+          <!--The endpoint of delegation service, to which this security handler will 
+          delegate the credential-->\
+          <DelegationServiceEndpoint>\
+            https://glueball.uio.no:60005/delegation
+          </DelegationServiceEndpoint>\ 
           <!--The endpoint of the peer service, to which this WS client will access-->\
           <PeerServiceEndpoint>https://selectron.uio.no:60000/echo</PeerServiceEndpoint>\ 
-         <!--The key and certificate of the credential, which is be delegated to the delegation service-->\
+         <!--The key and certificate of the credential, which is be delegated to the 
+         delegation service-->\
           <KeyPath>./testkey-nopass.pem</KeyPath>\ 
@@
           <!--ProxyPath>/tmp/5612d050.pem</ProxyPath-->\ 
-         <!--Yet another option of defining how to get the credential; this option is useful for the service which has already acquired the delegated credential (this delegated credential is stored under /tmp/hash_value.pem) from the delegation service, yet still needs to delegate this credential to another delegation service; the DelegationCredentialIdentity can be parsed from the tls connection from client side, and then used to calculate the hash value to get the delegated credential from local repository-->\
+         <!--Yet another option of defining how to get the credential; this option is 
+         useful for the service which has already acquired the delegated credential 
+         (this delegated credential is stored under /tmp/hash_value.pem) from the 
+         delegation service, yet still needs to delegate this credential to another delegation 
+         service; the DelegationCredentialIdentity can be parsed from the tls connection from 
+         client side, and then used to calculate the hash value to get the delegated credential 
+         from local repository-->\
           <!--DelegationCredIdentity>/O=KnowARC/OU=UiO/CN=squark.uio.no</DelegationCredIdentity-->\ 
@@
 
+\section{SAML Attribute Authority Service} % (fold)
+\label{sec:saml_aa_service}
+
+%section saml_aa_service (end)
+
 \end{document}

### 2009-08-13 doc/tech_doc/sec/arc-security-documentation.tex
more changes for sec documentation

@@
 
-It's configuration is presented in Section 7.5.
+It's configuration is presented in Section \ref{subsec:pdpservice_conf}.
 
@@
 Hence it must be attached to the MCC which processes SOAP payloads - like SOAP MCC or Service accepting SOAP messages.
-For description of configuration see section 8.13.
+For description of configuration see section \ref{subsec:ut_sechandler_conf}.
 
@@
 
-For the outgoing message this SecHandler creates adds proper token into SOAP header according to configuration.
+For the outgoing message this SecHandler creates and adds proper token into SOAP header according to configuration.
 
@@
 Hence it must be attached to the MCC which processes SOAP payloads – like SOAP MCC or Service accepting SOAP messages.
-For description of configuration see section 8.14.
+For description of configuration see section \ref{subsec:xt_sechandler_conf}.
 
@@
 \label{subsec:saml_token}
+The SAMLToken SecHandler is meant for processing - generating and extracting - WS-Security [4] SAMLToken from SOAP header.
+Hence it must be attached to the MCC which processes SOAP payloads – like SOAP MCC or Service accepting SOAP messages.
+For description of configuration see section \ref{subsec:st_sechandler_conf}.
+
+For the incoming message this SecHandler decrypts and checks signature of SOAP message using attached public key and verifies that key against specified CA certificate.
+
+For the outgoing message this SecHandler creates SAMLToken in SOAP header. SOAP message body is encrypted and signed.
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Policy.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Policy.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Request.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Request.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/Response.xsd .
-
-\subsection{Interface of policy decision service} % (fold)
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Response.xsd
+
+\subsection{Interface of policy decision service (Charon service)} % (fold)
 \label{subsec:interface_pds}
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/charon/charon.wsdl
+
+The following is the configuration about the Charon service, which is configured to use xacml policy engine.
+The policy engine is configurable by changing the ``name'' attribute of three elements: \textit{$<$Evaluator$>$}, \textit{$<$Policy$>$}, \textit{$<$Request$>$} into ``arc.evaluator'', ``arc.policy'', ``arc.request''.
+Also the policy should also be changed to the one with ARC specific format.
+
+\begin{verbatim}
+    <Service name="charon" id="charon_service">
+        <!--The element <Evaluator/>, <Policy/> and <Request/> configuration
+        are supposed to be used to load object; element <PolicyStore/> is
+        supposed to be used to get the location of policy-->
+        <charon:PDPConfig>
+            <charon:PolicyStore>
+                <charon:Location Type="file">charon_policy_xacml.xml.example</charon:Location>
+                <!-- other policy location-->
+            </charon:PolicyStore>
+            <charon:Evaluator name="xacml.evaluator" />
+            <charon:Policy name="xacml.policy" />
+            <charon:Request name="xacml.request" />
+        </charon:PDPConfig>
+    </Service>
+\end{verbatim}
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/mcc/tls/tls.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/mcc/tls/tls.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/charon/charon.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcauthzsh/SimpleListAuthZ.xsd.
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcauthzsh/SimpleListAuthZ.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/simplelistpdp/SimpleListPDP.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/simplelistpdp/SimpleListPDP.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/ArcPDP.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/ArcPDP.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/pdpserviceinvoker/ArcPDPServiceInvoker.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/pdpserviceinvoker/PDPServiceInvoker.xsd
 
@@
         <CACertificatePath>./ca.pem</CACertificatePath>
+        <RequestFormat>XACML</RequestFormat>
+        <TransferProtocol>SAML</TransferProtocol>
       </PDP>
@@
 
-The PDP Service Invoker is a client of PDP Service. The configuration options include endpoint of service and credentials to be used for establishing secure connection.
+The PDP Service Invoker is a client of PDP Service.
+The configuration options include endpoint of service and credentials to be used for establishing secure connection.
+In addition, the $<$RequestFormat$>$ element is for specifying the format of the request, and the $<$TransferProtocol$>$ element is for specifying the protocol of tranfering the request.
+
+Table \ref{table:support_pdp_request_protocol} shows the support of request and protocol in remote policy decision making.
+Note that if ``SAML 2.0 profile of XACML v2.0'' is configured in ``pdpservice invoker'', besides the authorizaton service (called charon service) implemented in ARC, it can interact with external authorization services, such as gLite authorization service (Yet the interoperation test has not been done).
+
+\begin{table}[ht]
+\caption{Support of request and protocol in remote policy decision making}
+\centering
+\begin{tabular}{| l | p{7cm} | p{5cm} |}
+\hline
+\textbf{  } & \textbf{ARC Request/Response} & \textbf{XACML Request/Response} \\ \hline
+\textbf{ARC protocol} & support & support \\ \hline
+\textbf{SAML protocol} & not support & support (SAML 2.0 profile of XACML v2.0) \\ \hline
+\end{tabular}
+\label{table:support_pdp_request_protocol}
+\end{table}
+
+\textbf{ARC protocol}:
+
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/charon/charon.wsdl
+
+\textbf{SAML protocol}:
+
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/charon/charon.wsdl
+
+http://www.oasis-open.org/committees/download.php/11475/access\_control-xacml-2.0-saml-protocol-schema-os.xsd
+
+\textbf{ARC request/response}:
+
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Request.xsd
+
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Response.xsd
+
+\textbf{Note} that the format of ``Response'' is supposed to correspond with the format of ``Request''.
+
+\textbf{XACML request/response}:
+
+http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-context-schema-os.xsd
+
+http://www.oasis-open.org/committees/download.php/11474/access\_control-xacml-2.0-saml-assertion-schema-os.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/delegationsh/DelegationPDP.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/delegationpdp/DelegationPDP.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/usernametokensh/UsernameTokenSH.xsd .
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/usernametokensh/UsernameTokenSH.xsd
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/x509tokensh/X509TokenSH.xsd.
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/x509tokensh/X509TokenSH.xsd
 
@@
 \end{verbatim}
+
+\subsection{SAMLToken SecHandler Configuration} % (fold)
+\label{subsec:st_sechandler_conf}
+Configuration XML schema with comments is available at
+
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/samltokensh/SAMLTokenSH.xsd
 
@@
 \label{sec:saml_aa_service}
+SAML Attribute Authority Service (SAML AA Service) is implemented for accepting SAML \textit{$<$samlp:AttributeQuery$>$} from the requester, and issuing SAML \textit{$<$samlp:Response$>$} which is supposed to include \textit{$<$saml:Assertion$>$} that contains the attribute information about the requester.
+
+SAML AA service is implemented with the same functionality as VOMS SAML service. More specifically, the interface of SAML AA service is the same as VOMS SAML service.
+
+The client functionality of \textbf{SAML AA service} exists in two places:
+
+\begin{itemize}
+    \item The specific client utility (saml\_assertion\_init), where the SAML AA service is contacted to get back the SAML assertion.
+
+    \item The security handler about SAML token, where the SAML AA service is contacted and then the responded SAML assertion is inserted into SOAP message according to WS-Security SAML Token profile.
+
+\end{itemize}
+
+The same as VOMS SAML Service, the VOMS Admin service is used in ARC for managing the attributes. Therefore the database schema from VOMS is also introduced into ARC.
+
+However the implementation of SAML AA Service in ARC is designed not to be coupled with any specific backend database.
+Therefore even if the database schema is changed, the implementation is not necessarily going to be changed.
+The way about decoupling the service implementation with database schema is through the service configuration.
+
+\begin{verbatim}
+        <Service name="aa.service" id="aa_service">
+            <aa:KeyPath>../../tests/echo/testkey-nopass.pem</aa:KeyPath>
+            <aa:CertificatePath>../../tests/echo/testcert.pem</aa:CertificatePath>
+            <aa:CACertificatesDir>../../tests/echo/certificates</aa:CACertificatesDir>
+            <aa:CACertificatePath>../../tests/echo/testcacert.pem</aa:CACertificatePath>
+
+            <aa:Database type="mysql" ip="localhost" port="3305" dbname="voms_knowarc" user="root" 
+              password="123456">
+              <!--The sql following sentences and their combination are introduced
+              according to the code of voms' mysql plugin.
+              The intention is that if some other database schema is used, you can
+              simply change the sql sentences here, without changing the code. But still,
+              the input argument to these queries should be with the same meaning,
+              and the same sequence as the following; also the output of these queries
+              should also be with the same meaning and the same sequence as the following-->
+              <aa:SQLSet name="UID">
+                <aa:SQL name="GetUID">SELECT userid FROM usr WHERE usr.dn = ?</aa:SQL>
+              </aa:SQLSet>
+
+              <aa:SQLSet name="Role">
+                <aa:SQL name="GetRole">SELECT groups.dn, role FROM groups, 
+                  m LEFT JOIN roles ON roles.rid = m.rid WHERE groups.gid = m.gid 
+                  AND m.userid = ? AND roles.role = ?</aa:SQL>
+                <aa:SQL name="GetGroup">SELECT groups.dn, NULL FROM groups, m 
+                  WHERE groups.gid = m.gid AND m.userid = ?</aa:SQL>
+              </aa:SQLSet>
+
+              <aa:SQLSet name="Group">
+                <aa:SQL name="GetGroup">SELECT groups.dn, NULL FROM groups, m 
+                  WHERE groups.gid = m.gid AND m.userid = ?</aa:SQL>
+              </aa:SQLSet>
+
+              <aa:SQLSet name="GroupAndRole">
+                <aa:SQL name="GetGroup">SELECT groups.dn, NULL FROM groups, m 
+                  WHERE groups.gid = m.gid AND m.userid = ?</aa:SQL>
+                <aa:SQL name="GetGroupAndRole">SELECT groups.dn, role FROM groups, 
+                  m LEFT JOIN roles ON roles.rid = m.rid WHERE groups.gid = m.gid 
+                  AND m.userid = ? AND roles.role = ? AND groups.dn = ?</aa:SQL>
+              </aa:SQLSet>
+
+              <aa:SQLSet name="All">
+                <aa:SQL name="GetAll">SELECT groups.dn, role FROM groups, 
+                  m LEFT JOIN roles ON roles.rid = m.rid WHERE groups.gid = m.gid 
+                  AND m.userid = ?</aa:SQL>
+              </aa:SQLSet>
+
+              <aa:SQLSet name="GroupAndRoleAttribute">
+                <aa:SQL name="GetUserAttribute">SELECT attributes.a_name, 
+                  usr_attrs.a_value, NULL, NULL FROM attributes, usr_attrs 
+                  WHERE attributes.a_id = usr_attrs.a_id AND usr_attrs.u_id = ?
+                </aa:SQL>
+                <aa:SQL name="GetGroupAttribute">SELECT attributes.a_name, 
+                  group_attrs.a_value, groups.dn, NULL FROM attributes, group_attrs, 
+                  groups, m WHERE attributes.a_id = group_attrs.a_id AND groups.gid = m.gid 
+                  AND m.userid = ? AND m.rid is NULL AND group_attrs.g_id = m.gid
+                </aa:SQL>
+                <aa:SQL name="GetGroupAndRoleAttribute">SELECT attributes.a_name, 
+                  role_attrs.a_value, groups.dn, roles.role FROM attributes, role_attrs, 
+                  groups, roles, m WHERE attributes.a_id = role_attrs.a_id 
+                  AND groups.gid = m.gid AND m.userid = ? AND m.rid = roles.rid 
+                  AND roles.role = ? AND groups.dn = ? AND role_attrs.g_id = m.gid 
+                  AND role_attrs.r_id = m.rid
+                </aa:SQL>
+              </aa:SQLSet>
+
+              <aa:SQLSet name="GroupAttribute">
+                <aa:SQL name="GetUserAttribute">SELECT attributes.a_name, 
+                  usr_attrs.a_value, NULL, NULL FROM attributes, usr_attrs 
+                  WHERE attributes.a_id = usr_attrs.a_id AND usr_attrs.u_id = ?
+                </aa:SQL>
+                <aa:SQL name="GetGroupAttribute">SELECT attributes.a_name, 
+                  group_attrs.a_value, groups.dn, NULL FROM attributes, group_attrs, 
+                  groups, m WHERE attributes.a_id = group_attrs.a_id AND groups.gid = m.gid 
+                  AND m.userid = ? AND m.rid is NULL AND group_attrs.g_id = m.gid
+                </aa:SQL>
+              </aa:SQLSet>
+
+              <aa:SQLSet name="RoleAttribute">
+                <aa:SQL name="GetUserAttribute">SELECT attributes.a_name, 
+                  usr_attrs.a_value, NULL, NULL FROM attributes, usr_attrs 
+                  WHERE attributes.a_id = usr_attrs.a_id AND usr_attrs.u_id = ?
+                </aa:SQL>
+                <aa:SQL name="GetRoleAttribute">SELECT attributes.a_name, 
+                  role_attrs.a_value, groups.dn, roles.role FROM m 
+                  INNER JOIN groups ON m.gid = groups.gid LEFT JOIN roles 
+                  ON roles.rid = m.rid INNER JOIN role_attrs on groups.gid = role_attrs.g_id 
+                  INNER JOIN attributes on attributes.a_id = role_attrs.a_id 
+                  WHERE role_attrs.r_id = roles.rid AND m.userid = ? AND roles.role = ?
+                </aa:SQL>
+              </aa:SQLSet>
+
+              <aa:SQLSet name="AllAttribute">
+                <aa:SQL name="GetUserAttribute">SELECT attributes.a_name, 
+                  usr_attrs.a_value, NULL, NULL FROM attributes, usr_attrs 
+                  WHERE attributes.a_id = usr_attrs.a_id AND usr_attrs.u_id = ?
+                </aa:SQL>
+                <aa:SQL name="GetGroupAttribute">SELECT attributes.a_name, 
+                  group_attrs.a_value, groups.dn, NULL FROM attributes, group_attrs, 
+                  groups, m WHERE attributes.a_id = group_attrs.a_id AND groups.gid = m.gid 
+                  AND m.userid = ? AND m.rid is NULL AND group_attrs.g_id = m.gid
+                </aa:SQL>
+                <aa:SQL name="GetGroupAndRoleAttributeAll">SELECT attributes.a_name, 
+                  role_attrs.a_value, groups.dn, roles.role FROM attributes, role_attrs, 
+                  groups, roles, m WHERE attributes.a_id = role_attrs.a_id 
+                  AND groups.gid = m.gid AND m.userid = ? AND m.rid = roles.rid 
+                  AND role_attrs.g_id = m.gid AND role_attrs.r_id = m.rid
+                </aa:SQL>
+              </aa:SQLSet>
+
+              <!--Other SQLSet-->
+
+            </aa:Database>
+        </Service>
+\end{verbatim}
 

### 2009-08-17 doc/tech_doc/sec/arc-security-documentation.tex
More in security documentation

@@
 For description of configuration see section \ref{subsec:st_sechandler_conf}.
+
+\begin{figure}
+\centering{{{\includegraphics[width=0.9\textwidth]{samltoken_aa.pdf}}}
+\caption{\label{fig:samltoken_aa} Interaction among Client, Grid/Web Service and Attribute Authority} }
+\end{figure}
+
+Figure \ref{fig:samltoken_aa} shows the interaction among Client, Grid/Web Service and Attribute Authority when SAMLToken security hanlder is deployed.
+The Attribute Authority service is described in Section \ref{sec:saml_aa_service}.
+For the Grid/Web service, if the service is hosted on ARC middleware, then the SAMLToken security handler should be deployed; if the service is hosted on other middlewares, then there should be other ways for supporting SAML Token authentication, e.g., Rampart (WS-Security modlue for Axis2).
 
@@
 SAML Attribute Authority Service (SAML AA Service) is implemented for accepting SAML \textit{$<$samlp:AttributeQuery$>$} from the requester, and issuing SAML \textit{$<$samlp:Response$>$} which is supposed to include \textit{$<$saml:Assertion$>$} that contains the attribute information about the requester.
+The implementation is according to the SAML 2.0 ``Assertion Query/Request Profile'', see paragraph 6 of the ``Profiles for the OASIS Security Assertion Markup Language (SAML) V2.0'':
+
+http://docs.oasis-open.org/security/saml/v2.0/saml-profiles-2.0-os.pdf
 
@@
 \begin{itemize}
-    \item The specific client utility (saml\_assertion\_init), where the SAML AA service is contacted to get back the SAML assertion.
+    \item The specific client utility (saml\_assertion\_init), where the SAML AA service is contacted to get back the SAML assertion; Also it is planned to put this client functionality into the client utility ``arcproxy'', so that the the returned SAML assertion will be one of the \textit{``extentions''} of proxy certificate.
 
@@
 
-The same as VOMS SAML Service, the VOMS Admin service is used in ARC for managing the attributes. Therefore the database schema from VOMS is also introduced into ARC.
-
-However the implementation of SAML AA Service in ARC is designed not to be coupled with any specific backend database.
+As a way of managing the attributes related to the requesters, the VOMS Admin service is used. Also the database schema that is used in VOMS is also introduced into ARC.
+
+Therefore, the following packages about VOMS Admin service are currently required for the running of SAML AA Service:
+
+glite-security-voms-admin-server-2.0.15-1
+
+glite-security-voms-admin-interface-2.0.2-1
+
+glite-security-voms-admin-client-2.0.8-1
+
+\textit{Note} that the version of these pakages could be updated according to the updating of gLite middlware.
+
+\begin{figure}[ht]
+\centering{{{\includegraphics[width=1.0\textwidth]{arc_voms.pdf}}}
+\caption{\label{fig:arc_voms}The relationship between ARC components and gLite (VOMS) components} }
+\end{figure}
+
+Howerver, the running of ARC's SAML AA Service does not neccessarily depend on VOMS Admin Service.
+As long as there is another attribute administration service (in parallel to VOMS Admin Service) which can nicely be used to manage the attributes, the SAML AA Service can be configured to adopt that service.
+The relationship between ARC's implementation and VOMS implementation is shown in \ref{figure:arc_voms}.
+
+In terms of implementation, SAML AA Service in ARC is designed not to be coupled with any specific backend database schema.
 Therefore even if the database schema is changed, the implementation is not necessarily going to be changed.
 The way about decoupling the service implementation with database schema is through the service configuration.
+
+\begin{verbatim}
+    <samlp:AttributeQuery>
+        <saml:Issuer/>
+        <saml:Subject/>
+        <saml:Attribute Name = "All">
+        <saml:Attribute Name = "Role">
+    <samlp:AttributeQuery>
+\end{verbatim}
+
+The following is the configuration about SAML AA Service, which uses the VOMS database schema when quering attributes.
+The database query is distinguished by the ``\textit{name}'' attribute of $<$SQLSet$>$ element.
+If the ``\textit{Name}'' of $<$saml:Attribute$>$ (see the above for an example about \textit{$<$samlp:AttributeQuery$>$}) matches the ``\textit{name}'' attribute of $<$SQLSet$>$ element, then the corresponding \textit{SQL} under this $<$SQLSet$>$ element will be executed, and the result will be put into the SAML asserion.
 
@@
 
+              <aa:SQLSet name="Default">
+                <aa:SQL name="GetDefault">SELECT groups.dn, role FROM groups, 
+                  m LEFT JOIN roles ON roles.rid = m.rid WHERE groups.gid = m.gid 
+                  AND m.userid = ?</aa:SQL>
+              </aa:SQLSet>
               <!--Other SQLSet-->
@@
 
+The configuration of SAML AA service is according to the schema:
+
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/saml/aaservice.xsd
+
 %section saml_aa_service (end)

### 2009-08-25 doc/tutorials/ws-programming-tutorial/doc/tex_appendix/appendix.tex
Update schema locations in tutorial document

@@
 	caption={[XML schema of configuration files.]\textbf{XML schema of configuration files.}}
-	]{../../../../src/hed/libs/loader/mcc.xsd}
+	]{../../../../src/hed/libs/loader/schema/loader.xsd}
 
@@
 	caption={[XML schema of the TCP component.]\textbf{XML schema of the TCP component.}}
-	]{../../../../src/hed/mcc/tcp/tcp.xsd}
+	]{../../../../src/hed/mcc/tcp/schema/tcp.xsd}
 
@@
 	caption={[XML schema of the TLS component.]\textbf{XML schema of  the TLS component.}}
-	]{../../../../src/hed/mcc/tls/tls.xsd}
+	]{../../../../src/hed/mcc/tls/schema/tls.xsd}
 
@@
 	caption={[XML schema of the HTTP component.]\textbf{XML schema of the HTTP component.}}
-	]{../../../../src/hed/mcc/http/http.xsd}
+	]{../../../../src/hed/mcc/http/schema/http.xsd}
 

### 2009-08-28 doc/tech_doc/xrsl/xrsl-urls.tex
document new cache=copy option

@@
    maximal value is 10\\
-   \verb#cache=yes|no|renew# & indicates whether the GM should
-   cache the file; default for input files is \verb#yes#. \verb#renew# forces a download of the file, even if the cached copy is still valid. (NB: cached files can not be modified by jobs)\\
-   \verb#readonly=yes|no# & for transfers to \verb#file://# destinations, specifies whether the file should be
+   \verb#cache=yes|no|renew|copy# & indicates whether the GM should
+   cache the file; default for input files is \verb#yes#. \verb#renew#
+   forces a download of the file, even if the cached copy is still valid.
+   \verb#copy# forces the cached file to be copied (rather than linked) to
+   the session dir, this is useful if for example the file is to be modified.\\
+   \verb#readonly=yes|no# & for transfers to \verb#file://# destinations,
+   specifies whether the file should be
    read-only (unmodifiable) or not; default is \verb#yes# \\

### 2009-09-02 doc/tech_doc/janitor/tex_introduction/introduction.tex
Adding Janitor documentation from janitor branch.

--- 2009-05-25 doc/manuals/janitor/tex_introduction/introduction.tex
@@
 
-Janitor is a service for the automated installation of runtime environment for grid computing elements.
-It is addressed transparently via the A-REX service used for job submission but can also be used as a standalone tool.
+Janitor is a service for automated runtime environment installation for grid computing elements.
+It is integrated in the A-REX service used for job submission but also can be used as a standalone tool.
 
@@
 
-A core problem remains to distribute a locally working solution, the Know-How, quickly across all
-contributing sites, i. e., without manual interference. Every scientific discipline has its respective own set of
-technologies for the distribution of work load. E.g. research in bioinformatics requires access to so many
+A core problem remains to distribute a locally working solution, the Know-How, quickly across all contributing sites, i.
+e., without manual interference.
+Every scientific discipline has its respective own set of
+technologies for the distribution of work load. E. g. research in bioinformatics requires access to so many
 different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
-a commodity, with all their intrinsic problems as there are bottlenecks and restrictions of repeated access.
-The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges
-with the here presented work extends the NorduGrid's
+a commodity, with all the problems with respect to bottlenecks and restrictions of repeated access.
+The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges with the here presented work extends the NorduGrid?s
 Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
@@
 
-This document will start with a chapter on how to set-up the Janitor locally. The following chapter will then give furhter instructions on how to 
-use the Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
-which is basically covering the method how to prepare new runtime environments. Deeper insights on the design of the Janitor will 
-be given by the subsequent forth chapter. In the fifth chapter, an outlook to anticipated future developments will given.
+This document will start with a chapter of how to set up janitor locally. The following chapter will give an instruction on how to 
+use Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
+which is basically covering the method how to prepare new runtime environments. A deeper knowledge about the design of Janitor will 
+be given by the subsequent forth chapter. In the fifth chapter, an outlook to future developments will given.
 

### 2009-09-02 doc/tech_doc/janitor/tex_maintenance/maintenance.tex
Adding Janitor documentation from janitor branch.

--- 2009-05-25 doc/manuals/janitor/tex_maintenance/maintenance.tex
@@
 
-This chapter explains how to maintain the Catalog and the Janitor itself. 
-In order to administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e. The first
+This chapter explains how to maintain the catalog and Janitor itself. 
+In order to administrate the catalog, it is absolutly recommened (but not required) to use the ontology editor Prot\`eg\`e. The first
 section will explain how this is done.
-The second section presents a detailed explanation how to create new packages for the Janitor.
-In the last section, a typical use case in maintaining Janitor will be listed.
+The second section a detailed explanation how to create new packages for Janitor will be given.
+In the last section typical use case in maintaining Janitor will be listed.
 
@@
 
-The Catalog describes runtime environments and is either served through a web server or is distributed together with the Janitor as a regular file. 
+The Catalog describes runtime environments and is either served through a web server or distributed together with the Janitor. 
 It is specifed by a (Resource Description Framework) RDF file assigned to 
@@
 and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
-(i.e. here etch or sid), i.e. the name of a common installation or a virtual image.
-The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
-they can be found more easily.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
+(i.e. here etch or sid). The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
+they can be easier found.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
 They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
@@
 
-The dynamic Runtime Environments stored in the Catalog are presented on the aforementioned dedicated web
+The dynamic Runtime Environments stored in the catalog are presented on the aforementioned dedicated web
 page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
@@
 That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
-This HTML page listing the manually or automatically installable REs is prepared by the script web/list.pl.
-This script is meant to be run by a mod\_perl-enabled Apache. The script itself is only loosely coupled
-with the Janitor. In the first lines of the script some variables specific to the site are set. To configure the
+This html page listing the manually or automatically installable RTEs is prepared by the script web/list.pl.
+This script is meant to be run by a mod-perl enabled Apache. The script itself is only loosely integrated
+within the Janitor. In the first lines of the script some variables specific to the site are set. To configure the
 script these have to be changed~\cite[p. 9]{BAYER_2007}.
@@
 
-This section describes how to add new packages to the Catalog. In the current implementation, only tar based packages are processed by the Janitor.
+This section describes how to add new packages to the catalog. Currently only tar based packages are processable by Janitor.
 Within the example they are assigned to be used together with Debian Etch. This limitation is only literal. 
-There is no limitiation to newer Debian distributions.
+There should be no limitiation
+to newer Debian distributions.
 
@@
 
-At the time of writing, only the tape archive (tar) file format is accepted for dynamic Runtime Environment installation, 
+At the time of writing, only the tape archive (tar) file format is accepted for dynamic runtime environment installation, 
 a well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files. 
@@
 
-The tar file contains two directories, named {\tt control} and {\tt data}. Software is stored in the latter subdirectory,
+The tar file contains two directories named control and data. Software is stored in the latter subdirectory,
 while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
 the content of the data directory is extracted to some directory \textdollar BAR. 
-After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
+After this unpacking of the tar file, the Janitor executes the install skript provided in the control directory. 
 It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
@@
 The remove script, which will be executed right before WEKA is deinstalled, is empty. Janitor will delete the directory, so there
-is nothing more to be done.\\
+is nothing to be done.\\
 
@@
 \end{itemize}
-
 What sets High Energy Physics software apart is it's sheer size. The package in question takes up more than
-5 GB. This was a test illustrating the feasibility of using dynamic REs in High Energy Physics. The application
-of the dynamic RE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
+5 GB. This was a test illustrating the feasibility of using DREs in High Energy Physics. The application
+of the DRE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
 service, e.g. a software manager of a big experiment will be able to deploy software packages on production

### 2009-09-03 doc/tech_doc/janitor/tex_concept/concept.tex
Making Janitor documentation compile in usual (almost) way
Janitor documetation fixes: build fixed, tex fixed, image recovered

--- 2009-05-14 doc/manuals/janitor/tex_concept/concept.tex
@@
  \begin{center} 
-    \resizebox{24cm}{!}{\input{images/dependencies}}
+    %\resizebox{24cm}{!}{\input{images/dependencies}}
+    \includegraphics[width=24cm]{images/dependencies.pdf}
     \mycaption{Modules of the Janitor and their dependencies}{}
@@
  UNAVAILABLE&
-	The RE is not available for the BaseSystem (see \ref{sec:rdfschema}) the site uses. \\
+	The RE is not available for the BaseSystem (see \ref{sec:catalog}) the site uses. \\
  INSTALLABLE&
@@
     \mycaption{}{}
-    \label{fig:RE_states}
+    \label{fig:arex_states}
   \end{center}
@@
     \mycaption{}{}
-    \label{fig:RE_states}
+    \label{fig:janitor_integration_2}
   \end{center}
@@
     \mycaption{}{\task{To be translated and beautificated. SVG file is missing!}}
-    \label{fig:RE_states}
+    \label{fig:WS_structure}
   \end{center}
@@
 \begin{lstlisting}[
-        label=lst:SOAPrequest,
+        label=lst:arc_conf2,
         caption={ [Example \textit{arc.conf} settings for janitor.]
@@
 \begin{lstlisting}[
-        label=lst:SOAPrequest,
+        label=lst:arc_conf3,
         caption={ [Example \textit{arc.conf} settings for janitor.]

### 2009-09-03 doc/tech_doc/janitor/tex_maintenance/maintenance.tex
Janitor documetation fixes: build fixed, tex fixed, image recovered

@@
 
-\section{Introducing new packages}\label{sec:catalog}
+\section{Introducing new packages}\label{sec:introducing_new_packages}
 
@@
 
-The WEKA package for machine learning~\cite{FRANK_20074} and the Java Runtime Environment
+The WEKA package for machine learning~\cite{FRANK_2004} and the Java Runtime Environment
 are available as dynamic Runtime Environments. Further packages for bioinformatics comprise dynamic
@@
 
-\section{Typical use cases}\label{sec:catalog}
+\section{Typical use cases}\label{sec:typical_use_cases}
 

### 2009-09-08 doc/tech_doc/xrsl/xrsl-urls.tex
add new URL option tcpnodelay; correct checksum options

@@
    computed (ev. provided to the indexing server). This is overridden
-   by any metadata options specified (see below). If this option is
-   not provided, the default for the protocol is used. \verb#checksum=no#
-   in a source URL will disable checksum verification.\\ 
+   by any metadata options specified (see below).\\ 
    \verb#exec=yes|no# & means the file should be treated as executable\\
@@
    protocols is \verb#yes#\\
+   \verb#tcpnodelay=yes|no# & controls the use of the TCP\_NODELAY
+   socket option (which disables the Nagle algorithm). Applies to
+   http(s) only. Default is \verb#no#\\ 
 \end{tabular}
@@
    \verb#checksumtype# & Type of checksum. Supported values are cksum
-   (default), md5 and ad (adler32 checksum) \\
+   (default), md5 and adler32 \\
    \verb#checksumvalue# & The checksum of the file \\
@@
    \verb#lfc://srm://srm.domain.org/griddir@lfc.domain.org//user/file1:guid=\# \\
-   \verb#    bc68cdd0-bf94-41ce-ab5a-06a1512764dc:checksumtype=ad:checksumvalue=12345678#$^2$\\
+   \verb#    bc68cdd0-bf94-41ce-ab5a-06a1512764dc:checksumtype=adler32:checksumvalue=12345678#$^2$\\
    \verb#lfc://;cache=no@lfc.domain.org/:guid=bc68cdd0-bf94-41ce-ab5a-06a1512764d#$^3$\\

### 2009-09-10 doc/manuals/chelonia/arc-storage-manual.tex
Moved Chelonia documents around
renaming arc_storage_cli to chelonia

--- 2009-04-23 doc/storage_system/arc-storage-manual.tex
@@
 
-{\raggedleft NORDUGRID-XXXXXXX-NN\par}
+{\raggedleft NORDUGRID-MANUAL-10\par}
 
@@
 Now it is time to set up the prototype client tool to access the system.
-It is currently called \verb!arc_storage_cli!
+It is called \verb!chelonia!
 and it is by default installed to \verb!/usr/local/bin!.
@@
 
-Now you can use the \verb!arc_storage_cli! to access your local deployment. Try to do a `list' on the root collection (`/'):
-
-\begin{verbatim}
-$ arc_storage_cli
+Now you can use the tool \verb!chelonia! to access your local deployment. Try to do a `list' on the root collection (`/'):
+
+\begin{verbatim}
+$ chelonia
 Usage:
-  arc_storage_cli <method> [<arguments>]
+  chelonia <method> [<arguments>]
 Supported methods:
@@
 
-$ arc_storage_cli list
+$ chelonia list
 Usage: list <LN> [<LN> ...]
 
-$ arc_storage_cli list /
+$ chelonia list /
 - Calling the Bartender's list method...
@@
 
-Without arguments, the \verb!arc_storage_cli!
-prints the list of available methods.
+Without arguments, the \verb!chelonia!
+tool prints the list of available methods.
 Specifying only a method name prints the syntax of the given method.
@@
 \begin{verbatim}
-$ arc_storage_cli list /
+$ chelonia list /
 - Calling the Bartender's list method...
@@
 \begin{verbatim}
-$ arc_storage_cli makeCollection /
+$ chelonia makeCollection /
 - Calling the Bartender's makeCollection method...
@@
 
-$ arc_storage_cli list /
+$ chelonia list /
 - Calling the Bartender's list method...
@@
 
-Assume that Chelonia is running and you want to access it with the \verb!arc_storage_cli! client. The \verb!~/.arc/client.xml! is set properly, the contents of the root collection can be listed:
-
-\begin{verbatim}
-$ arc_storage_cli list /
+Assume that Chelonia is running and you want to access it with the \verb!chelonia! tool. The \verb!~/.arc/client.xml! is set properly, the contents of the root collection can be listed:
+
+\begin{verbatim}
+$ chelonia list /
 - Calling the Bartender's list method...
@@
 
-$ arc_storage_cli put orange /
+$ chelonia put orange /
 - The size of the file is 7 bytes
@@
 \begin{verbatim}
-$ arc_storage_cli stat /orange
+$ chelonia stat /orange
 - Calling the Bartender's stat method...
@@
 \begin{verbatim}
-$ arc_storage_cli get /orange /tmp
+$ chelonia get /orange /tmp
 - Calling the Bartender's getFile method...
@@
 \begin{verbatim}
-$ arc_storage_cli makeCollection /fruits
+$ chelonia makeCollection /fruits
 - Calling the Bartender's makeCollection method...
@@
 
-$ arc_storage_cli makeCollecion /fruits/apple
+$ chelonia makeCollecion /fruits/apple
 - Calling the Bartender's makeCollection method...
@@
 
-$ arc_storage_cli move /orange /fruits/
+$ chelonia move /orange /fruits/
 - Calling the Bartender's move method...
@@
 
-$ arc_storage_cli list /fruits
+$ chelonia list /fruits
 - Calling the Bartender's list method...
@@
 \begin{verbatim}
-$ arc_storage_cli rmdir /fruits
+$ chelonia rmdir /fruits
 - Calling the Bartender's unmakeCollection method...
@@
 
-$ arc_storage_cli rmdir /fruits/apple
+$ chelonia rmdir /fruits/apple
 - Calling the Bartender's unmakeCollection method...
@@
 
-$ arc_storage_cli rm /fruits/orange
+$ chelonia rm /fruits/orange
 - Calling the Bartender's delFile method...
@@
 
-$ arc_storage_cli rmdir /fruits
+$ chelonia rmdir /fruits
 - Calling the Bartender's unmakeCollection method...
@@
 Now, introduce a new user, Penny, into the system.
-Again, \verb!arc_storage_cli!
+Again, \verb!chelonia!
 can be told to use user certificate and key files for this user with modifying the \verb!client.xml!
@@
 $ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
-arc_storage_cli list /
+chelonia list /
 - Calling the Bartender's list method...
@@
 $ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
-arc_storage_cli put /tmp/orange /pennys-orange
+chelonia put /tmp/orange /pennys-orange
 - The size of the file is 7 bytes
@@
 \begin{verbatim}
-$ arc_storage_cli pol / change ALL +read +addEntry
+$ chelonia pol / change ALL +read +addEntry
 - Calling the Bartender's stat method...
@@
 \begin{verbatim}
-$ arc_storage_cli pol
+$ chelonia pol
 Usage: policy <LN> <changeType> <identity> <action list>
@@
 $ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
-arc_storage_cli pol /pennys-orange change /DC=eu/DC=KnowARC/O=knowarc/CN=billy +read
+chelonia pol /pennys-orange change /DC=eu/DC=KnowARC/O=knowarc/CN=billy +read
 - Calling the Bartender's stat method...
@@
 
-$ arc_storage_cli stat /pennys-orange
+$ chelonia stat /pennys-orange
 - Calling the Bartender's stat method...
@@
 \begin{verbatim}
-$ arc_storage_cli get /pennys-orange /tmp
+$ chelonia get /pennys-orange /tmp
 - Calling the Bartender's getFile method...
@@
 \begin{verbatim}
-$ ARC_NEEDED_REPLICAS=2 arc_storage_cli put orange /new-orange
+$ ARC_NEEDED_REPLICAS=2 chelonia put orange /new-orange
 - The size of the file is 7 bytes
@@
 
-$ arc_storage_cli stat /new-orange
+$ chelonia stat /new-orange
 - Calling the Bartender's stat method...
@@
 \begin{verbatim}
-$ arc_storage_cli modify /new-orange set states neededReplicas 1
+$ chelonia modify /new-orange set states neededReplicas 1
 - Calling the Bartender's modify method...
@@
 
-$ arc_storage_cli stat /new-orange | grep neededReplicas
+$ chelonia stat /new-orange | grep neededReplicas
     neededReplicas: 1
@@
 
-It should be noted that the current version of the \verb!arc_storage_cli! tool does not support recursive methods or working with multiple files at once.
+It should be noted that the current version of the \verb!chelonia! tool does not support recursive methods or working with multiple files at once.
 
@@
 
-$ arc_storage_cli makeCollection /
+$ chelonia makeCollection /
 - Calling the Bartender's makeCollection method...
@@
 \begin{verbatim}
-$ ARC_NEEDED_REPLICAS=2 arc_storage_cli put orange /
+$ ARC_NEEDED_REPLICAS=2 chelonia put orange /
 - The size of the file is 7 bytes
@@
 
-$ arc_storage_cli stat /orange
+$ chelonia stat /orange
 - Calling the Bartender's stat method...
@@
 \begin{verbatim}
-$ arc_storage_cli get /orange /tmp/orange1
+$ chelonia get /orange /tmp/orange1
 - Calling the Bartender's getFile method...
@@
 
-$ arc_storage_cli get /orange /tmp/orange2
+$ chelonia get /orange /tmp/orange2
 - Calling the Bartender's getFile method...
@@
 \begin{verbatim}
-$ arc_storage_cli stat /orange
+$ chelonia stat /orange
   locations
@@
 
-$ arc_storage_cli stat /orange
+$ chelonia stat /orange
   locations
@@
 Creating collection '/': done
-$ arc_storage_cli put orange /
+$ chelonia put orange /
 - The size of the file is 7 bytes
@@
 'orange' (7 bytes) uploaded as '/orange'.
-$ arc_storage_cli modify /orange set states neededReplicas 3
+$ chelonia modify /orange set states neededReplicas 3
 - Calling the Bartender's modify method...
@@
 set
-$ arc_storage_cli stat /orange
+$ chelonia stat /orange
 [...]
@@
 [...]
-zsombor@storage:~$ arc_storage_cli stat /orange
+zsombor@storage:~$ chelonia stat /orange
 [...]
@@
 \begin{verbatim}
-$ arc_storage_cli stat /orange
+$ chelonia stat /orange
 [...]
@@
 Now the system can be used as intended.
-There is a flag for the \verb!arc_storage_cli!
-which makes it very verbose.
+There is a flag for the \verb!chelonia!
+tool which makes it very verbose.
 You can check the used Bartender with it:
@@
 \begin{verbatim}
-$ arc_storage_cli -v stat /orange
+$ chelonia -v stat /orange
 - The URL of the Bartender(s): https://storage:60000/Bartender,
@@
 \begin{verbatim}
-zsombor@storage:~$ arc_storage_cli -v stat /orange
+zsombor@storage:~$ chelonia -v stat /orange
 - The URL of the Bartender(s): https://storage:60000/Bartender,
@@
 \begin{verbatim}
-$ arc_storage_cli makemount /salmount \
+$ chelonia makemount /salmount \
     gsiftp://sal1.uppmax.uu.se:2811/pnfs/uppmax.uu.se/data
@@
 
-$ arc_storage_cli stat /salmount
+$ chelonia stat /salmount
 - Calling the Bartender's stat method...
@@
 \begin{verbatim}
-$ arc_storage_cli list /salmount- Calling the Bartender's list method...
+$ chelonia list /salmount- Calling the Bartender's list method...
 - done in 0.49 seconds.
@@
 \begin{verbatim}
-$ arc_storage_cli cre
+$ chelonia cre
 - Calling the Bartender's credentialsDelegation method...
@@
 \begin{verbatim}
-$ arc_storage_cli list /salmount
+$ chelonia list /salmount
 - Calling the Bartender's list method...
@@
 \begin{verbatim}
-$ arc_storage_cli get /salmount/data1 /tmp
+$ chelonia get /salmount/data1 /tmp
 - Calling the Bartender's getFile method...
@@
 orange
-$ arc_storage_cli list /
+$ chelonia list /
 - Calling the Bartender's list method...
@@
   orange <file>
-$ arc_storage_cli makeCollection /fruits
+$ chelonia makeCollection /fruits
 - Calling the Bartender's makeCollection method...
@@
 
-Besides the \verb!arc_storage_cli! tool you can access the storage system with the use of a FUSE module. The following steps installs the ARC storage FUSE module:
+Besides the \verb!chelonia! tool you can access the storage system with the use of a FUSE module. The following steps installs the ARC storage FUSE module:
 

### 2009-09-10 doc/tech_doc/chelonia/arc-storage-documentation.tex
renaming arc_storage_cli to chelonia

--- 2009-04-30 doc/tech_doc/storage_system/arc-storage-documentation.tex
@@
 
-In the prototype release there is a command-line client tool called \verb!arc_storage_cli!, which is written in Python, and only need a basic Python installation to run.
+In the prototype release there is a command-line client tool called \verb!chelonia!, which is written in Python, and only need a basic Python installation to run.
 It is capable of communicating with a given Bartender service, and can upload and download files via HTTP.
@@
 
-The \verb!arc_storage_cli! has its built-in help. The methods can be listed by giving the command without arguments:
-
-\begin{verbatim}
-$ arc_storage_cli
+The \verb!chelonia! has its built-in help. The methods can be listed by giving the command without arguments:
+
+\begin{verbatim}
+$ chelonia
 Usage:
-  arc_storage_cli <method> [<arguments>]
+  chelonia <method> [<arguments>]
   Supported methods: stat, make[Collection], unmake[Collection], list, move,
@@
 \begin{verbatim}
-$ arc_storage_cli move
+$ chelonia move
 Usage: move <sourceLN> <targetLN>
@@
 This is a testfile.
-$ arc_storage_cli put testfile /tmp/
+$ chelonia put testfile /tmp/
 - The size of the file is 20 bytes
@@
 'testfile' (20 bytes) uploaded as '/tmp/testfile'.
-$ arc_storage_cli stat /tmp/testfile
+$ chelonia stat /tmp/testfile
 - Calling the Bartender's stat method...
@@
     type: file
-$ arc_storage_cli get /tmp/testfile newfile
+$ chelonia get /tmp/testfile newfile
 - Calling the Bartender's getFile method...
@@
 Using FUSE and the ARC Python interface, ARCFS allows users to mount the storage namespace into the local namespace, enabling the use of operating system commands and tools such as graphical file browsers.
-Delegation handling is not yet implemented in the fuse module, so that to change ownership, group access, etc., the \verb!arc_storage_cli!
+Delegation handling is not yet implemented in the fuse module, so that to change ownership, group access, etc., the \verb!chelonia!
 tool must be used.
 
-As for the \verb!arc_storage_cli!
+As for the \verb!chelonia!
 the credentials of the user and the URL of the Bartender can be configured either with an XML file located at \verb!~/.arc/client.xml!
@@
 $ fusermount -u ./mnt
-$ arc_storage_cli stat /tmp/testfile
+$ chelonia stat /tmp/testfile
 - Calling the Bartender's stat method...
@@
 
-As with the \verb!arc_storage_cli!
+As with the \verb!chelonia!
 command the credentials of the user and the URL of the Bartender can be configured with an XML file located at \verb!~/.arc/client.xml!.

### 2009-09-14 doc/tech_doc/xrsl/xrsl-urls.tex
tidy up and remove unsupported protocols

@@
    \texttt{ldap} & ordinary \textit{Lightweight Data Access Protocol (LDAP)}~\cite{ldap}\\
-   \texttt{rc} & \globus\  \textit{Replica Catalog (RC)}~\cite{rc}\\
    \texttt{rls} & \globus\  \textit{Replica Location Service (RLS)}~\cite{rls}\\
-   \texttt{fireman} & Fireman indexing service of EGEE gLite~\cite{glite}\\
    \texttt{lfc} & LFC catalog and indexing service of EGEE gLite~\cite{glite}\\
-   \texttt{se} & ARC Smart Storage Element service~\cite{se}\\
    \texttt{srm} & Storage Resource Manager (SRM) service~\cite{srm}\\
@@
 \begin{shaded}
-   \verb#<protocol>://host[:port]/<file>#
+   \verb#protocol://[host[:port]]/file#
 \end{shaded}
@@
 \begin{shaded}
-   \verb#<protocol>://host[:port][;option[;option[...]]]/<file>#
+   \verb#protocol://[host[:port]][;option[;option[...]]]/file#
 \end{shaded}
@@
 \begin{shaded}
-   \verb#rc://rc://[location[|location[...]]@]<host>[:port]/<DN>/<lfn>#\\
-   \verb#rls://[url[|url[...]]@]<host>[:port]/<lfn>#\\
-   \verb#fireman://[url[|url[...]]@]<host>[:port]/<service_path>?<lfn>#\\
-   \verb#lfc://[url[|url[...]]@]<host>[:port]/<lfn>#
-\end{shaded}
-
-For the Smart Storage Element service, the syntax is
-\begin{shaded}
-   \verb#se://host[:port][;options]/path[?file_id]#
+   \verb#protocol://[url[|url[...]]@]host[:port][;option[;option[...]]/#\\
+   \verb#     lfn[:metadataoption[:metadataoption[...]]#
 \end{shaded}
@@
 \begin{shaded}
-   \verb#srm://<host>[:port][;options]/[service_path?SFN=]<file_id>#
+   \verb#srm://host[:port][;options]/[service_path?SFN=]file#
 \end{shaded}
@@
 \begin{tabular}{lp{10cm}}
-   \verb#location#&\verb#<location_name_in_RC>[;option[;option[...]]]#\\
-   \verb#host[:port]#&IP address of a server\\
-   \verb#DN#&Distinguished Name (as in LDAP) of an RC collection\\
-   \verb#lfn#&Logical File Name\\
-   \verb#url#&URL of the file as registered in RLS/Fireman\\
-   \verb#service_path#&End-point path of the Web service\\
-   \verb#file#&local to the host file name with a full path\\
+   \verb#host[:port]# & Hostname or IP address [and port] of a server\\
+   \verb#lfn# & Logical File Name\\
+   \verb#url# & URL of the file as registered in indexing service\\
+   \verb#service_path# & End-point path of the web service\\
+   \verb#file# & File name with full path\\
+   \verb#option# & URL option\\
+   \verb#metadataoption# & Metadata option for indexing service\\
 \end{tabular}
@@
    destination even if job processing failed (default is \verb#no#)\\
-   \verb#pattern=<pattern># & defines file matching pattern; currently
-   works for file listing requests sent to an \verb#se://# endpoint\\
    \verb#guid=yes|no# & make software use GUIDs instead of LFNs while
@@
    \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
-   \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#\\
    \verb#rls://gsiftp://se.domain.org/datapath/file25.dat@grid.domain.org:61238/myfile02.dat#$^1$\\
-   \verb#fireman://fireman_host:8443/glite-data-catalog-interface/FiremanCatalog?data.root#\\
    \verb#file:///home/auser/griddir/steer.cra#\\
-   \verb#lfc://srm://srm.domain.org/griddir@lfc.domain.org//user/file1:guid=\# \\
+   \verb#lfc://srm://srm.domain.org/griddir@lfc.domain.org/user/file1:guid=\# \\
    \verb#    bc68cdd0-bf94-41ce-ab5a-06a1512764dc:checksumtype=adler32:checksumvalue=12345678#$^2$\\
-   \verb#lfc://;cache=no@lfc.domain.org/:guid=bc68cdd0-bf94-41ce-ab5a-06a1512764d#$^3$\\
+   \verb#lfc://lfc.domain.org;cache=no/:guid=bc68cdd0-bf94-41ce-ab5a-06a1512764d#$^3$\\
 \end{framed}
@@
 lfc.domain.org with the given GUID and can be copied or queried by
-this URL. Note that as URL options are part of the location (physical)
-URL, in meta service URLs the options must be part of the location
-URL, even if the location URL is empty.
+this URL.

### 2009-09-18 doc/manuals/python/SWIG-wrapped-ARC-Python-API.tex
Documentation about the HED's python API and the arcom utility package - written by Tamás Kazinczy

\documentclass{article}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%

%---CUSTOM FLOAT---
\usepackage{float}
\floatstyle{ruled}
\newfloat{example}{H}{loe}
\floatname{example}{Example}
\floatstyle{plain}
\newfloat{illustration}{htp}{loi}
\floatname{illustration}{Illustration}

%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Tamas Kazinczy},
  pdftitle = {Documentation of ARC1 Python API},
  pdfsubject = {},
  pdfkeywords = {API,ARC1,arcom,Python,SWIG},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
%\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://download.nordugrid.org/}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{fig/logo-knowarc.png}}
\end{tabular}

\hrulefill

%-------- Change this to Knowarc_D<w>.<d>-<n>_<yy>

{\raggedleft Knowarc\_D0.0-0\_00\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large The SWIG wrapped ARC Python API and the arcom utility package\footnote{as of rev. 13724}}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large Paper subtitle}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Tam\'as Kazinczy\footnote{kazy@niif.hu} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

\tableofcontents
\newpage
%\listoffigures
%\newpage

\section{Preface}
When it comes to creating new functionality in ARC\footnote{Advanced Resource Connector; \url{http://www.nordugrid.org/middleware/}} 
it is nice if one could do that in a simple, quick and neat way.
Python\footnote{\url{http://www.python.org/}} is known for its versatility, simple and clean design, that
made it a good choice for rapid prototyping and gluing.
As the new ARC has moved towards the service oriented architecture, it has become a viable 
option to use Python both for creating new services and gluing existing ones together.
To be able to work with the functionality provided by ARC HED\footnote{Hosting Environment Daemon; \url{http://www.nordugrid.org/documents/ARCHED\_article.pdf}}, 
a decision was made to wrap its API with SWIG\footnote{Simplified Wrapper and Interface Generator; \url{http://www.swig.org/}}.
Furthermore, the arcom utility package was created on the basis of Python modules 
that had been created to ease developers' lives. (It contains reusable components to support various parts of development, including but not limited to the following categories: client development, logging, security and service development.)

\subsection{Purpose of this document}

Objectives of this document are:

\begin{itemize}

  \item{to demonstrate differences of original and wrapped API}

  \item{to comment about the use of ARC Python API}

  \item{to document the arcom utility package}

\end{itemize}

\subsection{Structure}

The document has the following structure:

\begin{itemize}

  \item{In the first part we provide an overview of how the API is wrapped by SWIG. 
    Here the use of SWIG interface files will be discussed along with the main 
    differences resulting from them and some thoughts will be given about usage.}

  \item{In the second part the arcom utility package will be discussed.
    Here the functionality found in the package will be described and some
    examples of use will be shown.}

\end{itemize}

\section{The SWIG generated API}

In this part the following topics will be covered:

\begin{itemize}

  \item{Python specific parts of SWIG interface files}

  \item{main differences between the original and the SWIG generated API resulting 
    from the style of wrapping}

  \item{the use of the generated API from Python}

\end{itemize}

\subsection{Python specific parts of SWIG interface files}

SWIG interface files are found in the swig subdirectory under the ARC source directory.

Specific parts for Python in SWIG interface files are enclosed in SWIGPYTHON blocks, 
starting with \textit{\#ifdef SWIGPYTHON} and ending with \textit{\#endif}.

There are seven interface files that contain such blocks:

\begin{itemize}
  \item{Arc.i}

  \item{common.i}

  \item{message.i}

  \item{client.i}

  \item{data.i}

  \item{delegation.i}

  \item{security.i}
\end{itemize}

When not in a specific block (neither Python nor another - e.g. SWIGJAVA), instructions
are applied to all languages.

\subsubsection{Arc.i}
\label{iArc}
\begin{flushleft}
\begin{itemize}
  \item{\verb$%module arc$} \linebreak
    As stated in the SWIG documentation\footnote{\url{http://www.swig.org/Doc1.3/SWIGDocumentation.html}}: ''The \verb$%module$ directive defines the name of the module that will be created by SWIG.''
    That means, this module could be imported in Python code with: \textit{import arc}
\end{itemize}
\begin{itemize}
  \item{\verb$%include <stl.i>$} \linebreak
    Include the Standard Template Library, so that templates could be used.
\end{itemize}
\begin{itemize}
  \item{
\begin{verbatim}
  #ifdef SWIGPYTHON
  %include <std_list.i>
  #endif\end{verbatim}
    If target language is Python include \verb$std_list$.
  } \linebreak
\end{itemize}
\begin{itemize}
  \item{\verb$%template(StringList) std::list<std::string>;$\linebreak
    Define StringList as a template list of strings. An example of handling a list of strings is shown below. \linebreak
    \begin{example}
      \caption{Handling a list of strings}\label{listex}
\begin{verbatim}
>>> import arc
>>> #create an empty list
... sl = arc.StringList()
>>> #append strings to the list
... #size increases
... sl.size()
0
>>> sl.append('apple')
>>> sl.size()
1
>>> sl.append('banana')
>>> sl.size()
2
>>> sl.append('lemon')
>>> sl.size()
3
>>> sl.append('orange')
>>> sl.size()
4
>>> #list members
... sl[0], sl[1], sl[2], sl[3]
('apple', 'banana', 'lemon', 'orange')
>>> #do some slicing
... tmp = sl[:2]
>>> #now tmp contains the first two strings
... tmp.size()
2
>>> tmp[0], tmp[1]
('apple', 'banana')
>>> #do some more slicing
... tmp = sl[1:]
>>> #now tmp contains all the strings except the first
... tmp.size()
3
>>> tmp[0], tmp[1], tmp[2]
('banana', 'lemon', 'orange')
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{\verb$%template(StringStringMap) std::map<std::string, std::string>;$ \linebreak
    Define StringStringMap as a template map where both key and value are strings.\linebreak
    \begin{example}
      \caption{arc.StringStringMap}\label{mapex}
\begin{verbatim}
>>> import arc
>>> #create an empty map
... ssm = arc.StringStringMap()
>>> #add mapping
... ssm['key1'] = 'value1'
>>> ssm['key2'] = 'value2'
>>> #get keys
... ssm.keys()
['key1', 'key2']
>>> #get value for 'key1'
... ssm['key1']
'value1'
>>> #get value for 'key2'
... ssm['key2']
'value2'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{
\begin{verbatim}
  %include "common.i"
  %include "message.i"
  %include "client.i"
  %include "data.i"
  %include "delegation.i"\end{verbatim}
  Include other interface files.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  %include "security.i"
  #endif\end{verbatim}
  If target language is Python include interface file: security.i
  }
\end{itemize}
\end{flushleft}

%\newpage
\subsubsection{common.i}
\label{icommon}
\begin{flushleft}
\begin{itemize}
  \item{ \verb$%include <typemaps.i>$} \linebreak
  Typemaps will be used. 
\end{itemize}
\begin{itemize}
  \item{\begin{verbatim}
  %ignore operator !;
  %ignore operator [];
  %ignore operator =;
  %ignore operator ++;
  %ignore operator <<;

  %ignore *::operator [];
  %ignore *::operator ++;
  %ignore *::operator --;
  %ignore *::operator =;

  %ignore Arc::MatchXMLName;
  %ignore Arc::MatchXMLNamespace;

  %template(XMLNodeList) std::list<Arc::XMLNode>;
  %template(URLList) std::list<Arc::URL>;
  \end{verbatim}
  }
  As stated in SWIG documentation: ''\verb$%ignore$ instructs SWIG to ignore 
  declarations that match a given identifier.
  Any function, variable, etc. which matches \verb$%ignore$ will not be wrapped 
  and therefore will not be available from the target language.''
  According to this, ignored operators and methods above will not be accessible 
  from Python. \linebreak
  Define XMLNodeList as a template list of Arc::XMLNode and URLList as a template list of 
  Arc::URL objects. (See Example\ref{listex} for an example of using a list of objects - 
  list of strings in that particular case - in Python.)\linebreak
\end{itemize}
\begin{itemize}
  \item{
\begin{verbatim}
  %rename(toBool) operator bool;
  %rename(__str__) operator std::string;
\end{verbatim} 
    If target language is Python then apply the following:
    \begin{itemize}
      \item{
      The operator ''bool'' is being renamed to ''toBool'' and is thus accessible from Python that way. 
      However, ''toBool'' does not appear in the Python API, that is, there is no trace of ''toBool'' 
      either in the generated code or in error messages.
      }
      \item{
      The operator ''std::string'' is being renamed to ''\verb$__str__$'' and is thus accessible from 
      Python that way.
      }
    \end{itemize}
    \begin{example}
      \caption{String operator - simple node}\label{strex1}
\begin{verbatim}
>>> import arc
>>> #create an XMLNode
... mynode = arc.XMLNode(arc.NS({'me':'http://example.com/myExample'}),'me:myNode')
>>> #set text content
... mynode.Set('Hello, World!')
>>> #representation of the node
... mynode.GetXML()
'<me:myNode xmlns:me="http://example.com/myExample">Hello, World!</me:myNode>'
>>> #String operator in action
... str(mynode)
'Hello, World!'
\end{verbatim}
    \end{example}
    \begin{example}
      \caption{String operator - complex node}\label{strex2}
\begin{verbatim}
>>> # String operator of XMLNode works with the text content.
... # It does not care about child nodes or text contents of those nodes.
>>> 
>>> # build a tree and set text content in nodes
... #       r
... #     / | \
... #    /  |  \
... #   /   |   \
... # c01  c02  c03
... #      /\    |
... #     /  \  c06  
... #   c04  c05
... 
>>> # create the root node
... r = arc.XMLNode(arc.NS(),'r')
>>> # set text content
... r.Set('R')
>>> # create rest of the tree
... c01 = r.NewChild('c01')
>>> c01.Set('C-01')
>>> c02 = r.NewChild('c02')
>>> c02.Set('C-02')
>>> c03 = r.NewChild('c03')
>>> c03.Set('C-03')
>>> c04 = c02.NewChild('c04')
>>> c04.Set('C-04')
>>> c05 = c02.NewChild('c05')
>>> c05.Set('C-05')
>>> c06 = c03.NewChild('c06')
>>> c06.Set('C-06')
>>> # String operator does not care about child nodes
... str(c03)
'C-03'
>>> str(c02)
'C-02'
>>> str(r)
'R'
>>> # XML representation of a complex node
... c02.GetXML()
'<c02>C-02<c04>C-04</c04><c05>C-05</c05></c02>'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{\verb$%rename(_print) Arc::Config::print;$ } \linebreak
  ''Arc::Config::print'' is being renamed to ''\verb$_print$'' and is thus accessible from 
  Python that way.
\end{itemize}
\begin{itemize}
  \item{\begin{verbatim}
  %apply std::string& OUTPUT { std::string& out_xml_str };
  %include "../src/hed/libs/common/XMLNode.h"
  %clear std::string& out_xml_str;\end{verbatim}
  It is common that a function provides return values through parameters (pointers or references).
  In Python, such functions should have multiple return values. This typemap tells SWIG that all 
  occurrences of ''\verb$out_xml_str$'' of type ''\verb$std::string&$'' as a parameter in XMLNode.h
  should be turned into ''\verb$std::string&$'' output values instead. An example regarding the 
  GetDoc function is shown below.
    \begin{example}
      \caption{Occurence of out\_xml\_str as an output value}\label{outxml}
\begin{verbatim}
>>> import arc
>>> #create node
... n = arc.XMLNode(arc.NS({'me':'myNS'}),'me:myNode')
>>> #set content
... n.Set('Hello, World!')
>>> #get document
... #GetDoc according to API:
... #void GetDoc(std::string &out_xml_str, bool user_friendly=false) const
... #out_xml_str is turned into an output value and user_friendly is an optional 
... #parameter so there are no necessary parameters this time
... #result will hold the output value
... result = n.GetDoc()
>>> #show result
... result
'<?xml version="1.0"?>\n<me:myNode xmlns:me="myNS">Hello, World!</me:myNode>\n'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{\verb$%rename(LogStream_ostream) LogStream;$ \linebreak
  LogStream is being renamed to ''\verb$LogStream_ostream$''.
  It is so, because a function with the same name will be defined later.
  (See below.)
  }
\end{itemize}
\begin{itemize}
  \item{
\begin{verbatim}
%inline %{ 
class CPyOutbuf : public std::streambuf {
 ...
};

class CPyOstream : public std::ostream {
 ...
};

%}
\end{verbatim}
  Classes CPyOutBuf and CPyOstream are created. These are to be used when creating LogStreams.
  (Python does not have stream objects - it uses files instead - so these classes are defined 
   to be able to represent a Python file as an ostream that is required by the C++ code.)
  Although they will be accessible from Python and thus could be used to create LogStreams,
  it is much more comfortable to take advantage of the function defined right after.
  (Complete source code can be found in Appendix \ref{LogStreamHelperFullCode} of this document.)
  }
\end{itemize}
\begin{itemize}
  \item{\begin{verbatim}
%pythoncode %{
    def LogStream(file):
        os = CPyOstream(file)
        os.thisown = False
        ls = LogStream_ostream(os)
        ls.thisown = False
        return ls
%}
\end{verbatim}
  Provides an easy-to-use function to create LogStreams with.
    \begin{example}
      \caption{Using LogStream to add new destination to the root logger}\label{logstream}
\begin{verbatim}
>>> import arc
>>> import sys
>>> #get root logger
... root_logger = arc.Logger_getRootLogger()
>>> #create a LogStream; sys.stdout would be OK
... stream = arc.LogStream(sys.stdout)
>>> #add destination to root logger
... root_logger.addDestination(stream)
>>> #log a message
... #result immediately appears on sys.stdout
... root_logger.msg(arc.INFO, 'Hello, World!')
[2009-05-12 15:39:07] [Arc] [INFO] [28021/161022392] Hello, World!
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

%\newpage
\subsubsection{message.i}
\label{imessage}
\begin{flushleft}
\begin{itemize}
  \item{ \verb$%include <typemaps.i>$} \linebreak
  Typemaps will be used. 
\end{itemize}
\begin{itemize}
  \item{\verb$%rename(next) Arc::AttributeIterator::operator++;$ \linebreak
  Rename ''operator++'' of Arc::AttributeIterator to ''next''.
  For an example see the wrapping of Arc::MessageAttributes::getAll below. 
  }
\end{itemize}
\begin{itemize}
  \item{\begin{verbatim}
  #ifdef SWIGPYTHON
  %pythonappend Arc::MessageAttributes::getAll %{
      d = dict()
      while val.hasMore():
          d[val.key()] = val.__ref__()
          val.next()
      return d
  %}
  #endif\end{verbatim} 
  If target language is Python then append the above code to ''Arc::MessageAttributes::getAll''. 
  Basically this means that MessageAttributes are rather returned as a dictionary.
  (The original result is an iterator. This iterator is used in the appended code to collect 
   MessageAttributes and put them in a dictionary which is more natural to use in Python.)
    \begin{example}
      \caption{MessageAttributes - getAll}\label{msgattrex}
\begin{verbatim}
>>> import arc
>>> #create MessageAttributes object
... mas = arc.MessageAttributes()
>>> #add some key-value pairs
... mas.add('key1', 'value1')
>>> mas.add('key3', 'value3')
>>> mas.add('key2', 'value2')
>>> #get all attributes
... all = mas.getAll()
>>> #show result
... all
{'key3': 'value3', 'key2': 'value2', 'key1': 'value1'}
>>> #function returns a dictionary
... type(all)
<type 'dict'>
>>> #show keys
... all.keys()
['key3', 'key2', 'key1']
>>> #show values
... all.values()
['value3', 'value2', 'value1']
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %apply std::string& OUTPUT { std::string &val };
  %include "../src/hed/libs/message/SecAttr.h"
  %clear std::string &val;\end{verbatim} 
  This typemap tells SWIG that all occurrences of ''val'' of type ''\verb$std::string&$'' as a parameter
  in SecAttr.h should be turned into ''\verb$std::string&$'' output values instead.
  An example could be the Export functionality of SecAttr. However, SecAttr is a base class that
  should be extended, and such classes are not yet available for the Python API.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  %pythonprepend Arc::MessageAuth::Export %{
        x = XMLNode("<dummy/>")
        args = args[:-1] + (args[-1].fget(), x)
  %}
  %pythonappend Arc::MessageAuth::Export %{
        return x
  %}
  #endif \end{verbatim} 
  If target language is Python then ''Arc::MessageAuth::Export'' is modified in a way that: 
    \begin{itemize}
    \item{
      A dummy node is created that will be replaced with the actual output.
    }
    \item{
      The argument tuple (args) of the method is recreated to contain the dummy node mentioned above.
      (It will be the last element of the tuple.) Note that the fget() call is required to get the
      SecAttrFormat from the property object as current SecAttrFormats are defined as property objects
      in the SecAttr class.
    }
    \item{
      Finally, the new output - that used to be the dummy node - is returned.
    }
    \end{itemize}
  An example would require the use of SecAttr here, therefore one is not yet available for the Python API.
  (An own extension of SecAttr on the Python side would not be enough here because it would not be used by MessageAuth.)
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %apply std::string& OUTPUT { std::string& out_xml_str }; 
  %include "../src/hed/libs/message/SOAPEnvelope.h"
  %clear std::string& out_xml_str;
  \end{verbatim}
  This tells SWIG that all occurrences of ''\verb$out_xml_str$'' of type ''\verb$std::string&$'' as a parameter in SOAPEnvelope.h
  should be turned into ''\verb$std::string&$'' output values instead. An example regarding the GetXML function is shown below.
    \begin{example}
      \caption{SOAPEnvelope - out\_xml\_str}\label{outxmlex}
\begin{verbatim}
>>> import arc
>>> #create a namespace
... ns = arc.NS({'me':'http://example.com/myExample'})
>>> #create an empty SOAPEnvelope
... #use the namespace created above
... se = arc.SOAPEnvelope(ns,False)
>>> #show it
... se.GetXML()
'<soap-env:Envelope xmlns:me="http://example.com/myExample" xmlns:soap-enc="http
://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.or
g/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http:/
/www.w3.org/2001/XMLSchema-instance"><soap-env:Body/></soap-env:Envelope>'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

\subsubsection{client.i}
\label{iclient}
\begin{flushleft}
\begin{itemize}
  \item{ \begin{verbatim}
  %template(ExecutionTargetList) std::list<Arc::ExecutionTarget>;
  %template(JobControllerList) std::list<Arc::JobController *>;
  %template(JobList) std::list<Arc::Job>;
  %template(JobStateList) std::list<Arc::JobState>;
  %template(SourceTypeList) std::list<Arc::DataSourceType>;
  %template(TargetTypeList) std::list<Arc::DataTargetType>;
  %template(FileTypeList) std::list<Arc::FileType>;
  %template(DirectoryTypeList) std::list<Arc::DirectoryType>;
  %template(ApplicationEnvironmentList) std::list<Arc::ApplicationEnvironment>;
  %template(SoftwareVersionList) std::list<Arc::SoftwareVersion>;
  %template(SoftwareRequirementList) std::list<Arc::SoftwareRequirement>;
  \end{verbatim}
  }
  Define ExecutionTargetList as a template list of Arc::ExecutionTarget objects, 
  JobControllerList as a template list of Arc::JobController pointers, 
  JobList as a template list of Arc::Job, JobStateList as a template list of 
  Arc::JobState, SourceTypeList as a template list of Arc::DataSourceType, 
  TargetTypeList as a template list of Arc::DataTargetType, FileTypeList as a 
  template list of Arc::FileType, DirectoryTypeList as a template list of 
  Arc::DirectoryType, ApplicationEnvironmentList as a template list of 
  Arc::ApplicationEnvironment, SoftwareVersionList as a template list of 
  Arc::SoftwareVersion and SoftwareRequirementList as a template list of 
  Arc::SoftwareRequirement objects. \linebreak
  (See Example\ref{listex} for an example of using a list of objects in Python.)
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  namespace Arc {

  %typemap(in, numinputs=0) PayloadSOAP ** response (PayloadSOAP *temp) {
      $1 = &temp;
  }
  \end{verbatim}
  If target language is Python then this input typemap tells SWIG that ''response'' 
  arguments of type ''PayloadSOAP **'' found in ''Arc'' namespace are to be ignored. 
  As stated in the SWIG documentation: ''When numinputs is set to 0, the argument is 
  effectively ignored and cannot be supplied from the target language. '' 
  However the argument is still required when making the C/C++ call. This is solved 
  by providing a locally declared variable called temp from which the value used is 
  obtained. Statement ''\verb#$1 = &temp;#'' sets the input argument to point to this 
  temporary variable.
  \begin{verbatim}
  %typemap(argout) PayloadSOAP ** response {
      PyObject *o, *tuple;
      o = SWIG_NewPointerObj(*$1, SWIGTYPE_p_Arc__PayloadSOAP, SWIG_POINTER_OWN | 0 );
      tuple = PyTuple_New(2);
      PyTuple_SetItem(tuple,0,o);
      PyTuple_SetItem(tuple,1,$result);
      $result = tuple;
  }
  \end{verbatim}
  As the SWIG documentation tells: ''The \verb$"argout"$ typemap is used to return values from arguments.''
  With the previous typemap in mind this one indicates that the result will be changed as the following:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the argument.
    }
    \item{
      The second one comes from the original result.
    }
    \item{
      The tuple will be returned as the new result.
    }
    \end{itemize}
  When processing a SOAP request it is required to return \verb$MCC_Status$ along with
  the SOAP response. This is achieved by returning a tuple that contains both of them.
  An example of calling the Echo service is shown below.
    \begin{example}
      \caption{ClientSOAP - process}\label{procex}
\begin{verbatim}
>>> import arc
>>> #create default config
... cfg = arc.MCCConfig()
>>> #create URL
... url = arc.URL('http://localhost:50000/Echo')
>>> #create payload
... payload = arc.PayloadSOAP(arc.NS({'echo':'urn:echo'}))
>>> #create payload content
... #and set echo message
... payload.NewChild('echo:echo').NewChild('echo:say').Set('Hello, World!')
>>> #create client
... client = arc.ClientSOAP(cfg,url)
>>> #let the client call the service
... response, status = client.process(payload)
>>> #see response
... print response.GetXML()
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.
org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" x
mlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XM
LSchema-instance"><soap-env:Body><echo:echoResponse><echo:hear>[ Hello, World! ]
</echo:hear></echo:echoResponse></soap-env:Body></soap-env:Envelope>
>>> #check status
... status.getKind() == arc.STATUS_OK
True
\end{verbatim}
    \end{example}

  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
%extend ACCLoader {
  Arc::Broker * getBroker(const std::string& brokername) { 
      return dynamic_cast<Arc::Broker*>( self->getACC(brokername) ); 
  }
  Arc::Broker * loadBroker(const std::string& brokername, const XMLNode * cfg) { 
      return dynamic_cast<Arc::Broker*>( self->loadACC(brokername, cfg) ); 
  }
  Arc::JobController * getJobController(const std::string& jobcontrollername) { 
      return dynamic_cast<Arc::JobController*>( self->getACC(jobcontrollername) ); 
  }
  Arc::JobController * loadJobController(const std::string& jobcontrollername, 
                                         const XMLNode * cfg) { 
      return dynamic_cast<Arc::JobController*>( 
                                         self->loadACC(jobcontrollername, cfg) ); 
  }
}
  \end{verbatim}
  SWIG methods added to get/load Broker and JobController objects from an ACCLoader object.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %apply std::string& INOUT { std::string& delegation_id };
  %include "../src/hed/libs/client/ClientX509Delegation.h"
  %clear std::string& delegation_id;
  \end{verbatim}
  This tells SWIG that all occurrences of ''\verb$delegation_id$'' of type ''\verb$std::string&$'' 
  as a parameter in ClientX509Delegation.h should be used as input and turned into 
  ''\verb$std::string&$'' output values as well. In fact, it depends on the type of delegation
  whether ''\verb$delegation_id$'' is used as an input or is created and returned as an output value.
  (See example below.)
    \begin{example}
      \caption{ClientX509Delegation - delegation\_id}\label{delegex}
\begin{verbatim}
For example the class ClientX509Delegation have a method createDelegation that makes 
use of delegation_id: 
  "For gridsite delegation service, the delegation_id
   is supposed to be created by client side, and sent to service side;
   for ARC delegation service, the delegation_id is supposed to be created
   by service side, and returned back. So for gridsite delegation service,
   this parameter is treated as input, while for ARC delegation service,
   it is treated as output."
However, this implies that one should take great care when dealing with delegation_id.
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

\subsubsection{data.i}
\label{idata}
\begin{flushleft}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  namespace Arc {
  \end{verbatim}
  If target language is Python then proceed with the following rules in ''Arc'' namespace.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %typemap(in, numinputs=0) std::list<FileInfo> & files (std::list<FileInfo> temp) {
    $1 = &temp;
  }
  \end{verbatim}
  This input typemap tells SWIG that ''files'' arguments of type ''\verb$std::list<FileInfo>&$''
  are to be ignored. However the argument is still required when making the C/C++ call. 
  This is solved by providing a locally declared variable called temp from which the value 
  used is obtained. Statement ''\verb#$1 = &temp;#'' sets the input argument to point to this temporary variable.

  \begin{verbatim}
  %typemap(argout) std::list<FileInfo> & files {
    PyObject *o, *tuple;
    o = PyList_New(0);
    std::list<Arc::FileInfo>::iterator it;
    for (it = (*$1).begin(); it != (*$1).end(); ++it) {
        PyList_Append(o, SWIG_NewPointerObj(new Arc::FileInfo(*it), 
            SWIGTYPE_p_Arc__FileInfo, SWIG_POINTER_OWN | 0 ));
    }
    tuple = PyTuple_New(2);
    PyTuple_SetItem(tuple,0,o);
    PyTuple_SetItem(tuple,1,$result);
    $result = tuple;
  }
  \end{verbatim}
  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as the following:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the argument.
    }
    \item{
      The second one comes from the original result.
    }
    \item{
      The tuple will be returned as the new result.
    }
    \end{itemize}
    \begin{example}
      \caption{DataPointARC - ListFiles}\label{lfex}
\begin{verbatim}
One possible use of the FileInfo list could be the ListFiles funcionality of the 
DataPointARC class. This returns DataStatus along with the collected files. An example 
of using ListFiles will be shown later.
\end{verbatim} See Example\ref{url2dp} for details.
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %typemap(in, numinputs=0) std::string& failure_description (std::string temp) {
    $1 = &temp;
  }
  \end{verbatim}
  This input typemap tells SWIG that \verb$failure_description$ arguments of type \verb$std::string&$
  are to be ignored. However the argument is still required when making the C/C++ call. 
  This is solved by providing a locally declared variable called temp from which the value 
  used is obtained from. Statement ''\verb#$1 = &temp;#'' sets the input argument to point to this temporary variable.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %typemap(argout) std::string & failure_description {
    PyObject *o, *tuple;
    tuple = PyTuple_New(2);
    o = PyString_FromString($1->c_str());
    PyTuple_SetItem(tuple,0,$result);
    PyTuple_SetItem(tuple,1,o);
    $result = tuple;
  }
  \end{verbatim}
  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as the following:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the original result.
    }
    \item{
      The second one comes from the argument.
    }
    \item{
      The tuple will be returned as the new result.
    }

  One possible use of \verb$failure_description$ could be the Transfer funcionality of the 
  DataMover class. This returns a failure description along with the corresponding 
  DataStatus. This description may contain more details about the error if transfer 
  fails for some reason.
  (See example below.)
    \end{itemize}
    \begin{example}
      \caption{DataMover - Transfer}\label{dmex}
\begin{verbatim}
>>> import arc
>>> # create DataMover
... mover = arc.DataMover()
>>> # import datapoint_from_url from arcom package
... # DataPoints are required to use DataMover
... from arcom import datapoint_from_url
>>> # create source DataPoint
... src = datapoint_from_url('file://./file1')
>>> # create destination DataPoint
... dst = datapoint_from_url('file://./file2')
>>> # do the transfer
... status, failure = mover.Transfer(src, dst, arc.FileCache(), arc.URLMap())
>>> # check status
... str(status)
'Operation completed successfully'
>>> # check for failure message
... failure
''
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{\verb$%ignore Arc::DataHandle::operator->;$} \linebreak
  Ignored operator will not be accessible from Python.
\end{itemize}
\end{flushleft}
\subsubsection{delegation.i}
\label{idelegation}
\begin{flushleft}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  %ignore Arc::DelegationConsumer::Acquire(std::string&, std::string&);
  %ignore Arc::DelegationConsumerSOAP::UpdateCredentials(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationConsumerSOAP::DelegatedToken(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationConsumerSOAP::DelegatedToken(std::string&, std::string&, 
    const XMLNode&);
  %ignore Arc::DelegationContainerSOAP::UpdateCredentials(std::string&, 
    std::string&, const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationContainerSOAP::DelegatedToken(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationContainerSOAP::DelegatedToken(std::string&, std::string&, 
    const XMLNode&);
  #endif
  \end{verbatim}
  If target language is Python the selected methods are to be ignored and thus 
  will not be accessible.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
    %apply std::string& OUTPUT { std::string& credentials };
    %apply std::string& OUTPUT { std::string& identity };
    %ignore DelegationContainerSOAP::UpdateCredentials(
          std::string &,const SOAPEnvelope &, SOAPEnvelope &);
    %include "../src/hed/libs/delegation/DelegationInterface.h"
    %clear std::string& identity;
    %clear std::string& credentials;
  \end{verbatim}
  These tell SWIG that all occurrences of ''credentials'' and ''identity'' of type 
  ''\verb$std::string&$'' in DelegationInterface.h should be turned into 
  ''\verb$std::string&$'' output values.
  Furthermore, ''DelegationContainerSOAP::UpdateCredentials'' - with the parameters above - 
  will not be available from Python.
  Note that this is required by SWIG to properly generate Python code because applying the 
  previous typemap would result in two UpdateCredentials method having exactly the same parameter list.
    \begin{example}
      \caption{DelegationConsumerSOAP - UpdateCredentials}\label{dcsucex}
\begin{verbatim}
UpdateCredentials functionality of DelegationConsumerSOAP contains both credentials
and identity according to the C++ API documentation:

bool Arc::DelegationConsumerSOAP::UpdateCredentials (std::string & credentials,
std::string & identity, const SOAPEnvelope & in, SOAPEnvelope & out)

Occurences are turned into output values, that is, Python API will have UpdateCredentials 
take two parameters (both of them being of type SOAPEnvelope) and return a list.
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

\subsubsection{security.i}
This interface file differs from previous ones in that it does not contain implicit wrapping. 
That is, content will be wrapped according to explicit statements alone found in this file.
Examples regarding security will be shown later in the discussion of the ''arcom.security'' module.
\begin{flushleft}
\begin{itemize}
  \item{\verb$namespace ArcSec$} \linebreak
  Proceed with the following rules in ''ArcSec'' namespace.
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
    %nodefaultctor Policy;
    class Policy  {};

    %nodefaultctor Request;
    class Request {};
  \end{verbatim}
  No default constructor will be generated for classes Policy and Request.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
    typedef enum {
      DECISION_PERMIT = 0,
      DECISION_DENY = 1,
      DECISION_INDETERMINATE = 2,
      DECISION_NOT_APPLICABLE = 3
    } Result;

    typedef struct {
      Result res;
    } ResponseItem;
  \end{verbatim}
  }
  Define Result and ResponseItem.
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
    class ResponseList {
    public:
      int size() ;
      ResponseItem* getItem(int n);
      ResponseItem* operator[](int n);
      bool empty();
    };

    class Response {
    public:
      ResponseList& getResponseItems ();
    };

    class Source {
     public:
      Source(const Source& s):node(s.node);
      Source(Arc::XMLNode& xml);
      Source(std::istream& stream);
      Source(Arc::URL& url);
      Source(const std::string& str);
    };

    class SourceFile: public Source {
     public:
      SourceFile(const SourceFile& s):Source(s),stream(NULL);
      SourceFile(const char* name);
      SourceFile(const std::string& name);
    };

    class SourceURL: public Source {
     public:
      SourceURL(const SourceURL& s):Source(s),url(NULL);
      SourceURL(const char* url);
      SourceURL(const std::string& url);
    };
  \end{verbatim}
  }
  These tells SWIG the classes and their methods to be wrapped.
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
    %nodefaultctor Evaluator;
    %newobject Evaluator::evaluate;
    class Evaluator {
    public:
        void addPolicy(const Source& policy,const std::string& id = "");
        %apply SWIGTYPE *DISOWN {Policy *policy};
        void addPolicy(Policy* policy,const std::string& id = "");
        %clear Policy *policy;
        Response* evaluate(Request* request);
        Response* evaluate(const Source& request);
        Response* evaluate(const Source& request, const Source& policy);
        Response* evaluate(const Source& request, Policy* policyobj);
        Response* evaluate(Request* request, Policy* policyobj);
        Response* evaluate(Request* request, const Source& policy);
    };

    %newobject EvaluatorLoader::getEvaluator;
    %newobject EvaluatorLoader::getRequest;
    %newobject EvaluatorLoader::getPolicy;
    class EvaluatorLoader {
    public:
        EvaluatorLoader();
        Evaluator* getEvaluator(const std::string& classname);
        Request* getRequest(const std::string& classname, 
                            const Source& requestsource);
        Policy* getPolicy(const std::string& classname, 
                          const Source& policysource);
    };
  \end{verbatim}
  No default constructor will be generated for the Evaluator class. \linebreak
  SWIG is also told about additional classes to be wrapped. 
  Furthermore, a hint is given to SWIG about ownership of objects in libarcsecurity.
  }
\end{itemize}
\end{flushleft}

\newpage

\section{The arcom utility package}

This part is dedicated to documenting the arcom utility package.
As it is meant to ease developers' lives, the arcom package comes
with a handful of helper classes and utility functions.
The package has the following structure:
\begin{itemize}

  \item{\verb$__init__$\newline
  Provides a set of readily available utility functions. These functions provide a way 
  of e.g. importing classes or better handling of XML structures.
  }

  \item{client\newline
  Provides a simple class (Client) for sending SOAP messages to services.
  }

  \item{logger\newline
  Defines log levels and provides a function to get a logger of choice.
  }

  \item{security\newline
  Collection of classes and functions that helps when dealing with authorization requests,
  policies and decision making.
  }

  \item{service\newline
  Provides a base class for service definition. This base class should be extended then.
  }

  \item{threadpool\newline
  Thread pool support. Consists of classes ThreadPool, ThreadPoolThread and ReadWriteLock.
  The latter is a lock object that allows many simultaneous ''read locks'', but only one 
  ''write lock.''
  }

  \item{xmltree\newline
  Provides the XMLTree class which provides a way to convert from XML to native Python 
  structures and vice versa.
  }

  \item{store\newline
  Provides classes for storing arbitrary objects locally.
  }

  \begin{itemize}

    \item{basestore\newline
    Provides a base class for stores. This base class should be extended.
    }

    \item{cachedpicklestore\newline
    Class for storing objects to files using the Python Pickle module\footnote{\url{http://docs.python.org/library/pickle.html}}. 
    This class keeps all the objects in memory.
    }

    \item{cachedstringstore\newline
    Class for storing objects to files as strings. This class keeps all the objects in memory.
    }

    \item{picklestore\newline
    Class for storing objects to files using the Python Pickle module.
    }

    \item{stringstore\newline
    Class for storing objects to files as strings.
    }

    \item{transdbstore\newline
    Class for storing objects to a transactional Berkeley DB\footnote{\url{http://www.oracle.com/database/berkeley-db/db/index.html}}.
    }

    \item{zodbstore\newline
    Class for storing objects to a Zope Object Database\footnote{\url{http://www.zope.org/Products/StandaloneZODB}}.
    }

  \end{itemize}

\end{itemize}

\subsection{\_\_init\_\_.py}
\label{ainit}
''\verb$__init__.py$'' contains those parts of the arcom package that are
readily accessible without any further action (apart from importing the 
arcom package itself).
\begin{flushleft}
\begin{itemize}
  \item{\verb$import_class_from_string(str)$\newline
  Imports a class given as a string parameter.\newline
  The string parameter has the format: ''[package.]*module.classname'',
  that is, package or packages (separated by periods) followed by 
  module name and finally the class name.
    \begin{example}
      \caption{Importing the Logger class from module ''logger'' of ''arcom'' package}\label{loglog}
\begin{verbatim}
>>> import arcom
>>> # From the logger module within the arcom package
... # import the Logger class
... Logger = arcom.import_class_from_string('arcom.logger.Logger')
>>> # Now the class could be access through Logger
... Logger
<class arcom.logger.Logger at 0x89081dc>
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \verb$get_attributes(node)$\newline
  Returns all attributes of the supplied XMLNode (node) in a dictionary
  where attribute names will serve as keys.
    \begin{example}
      \caption{Getting attributes from an XMLNode}\label{getattr}
\begin{verbatim}
>>> import arc
>>> import arcom
>>> # Create XMLNode
... n = arc.XMLNode(arc.NS({'me':'http://example.com/myExample'}),'me:myExample')
>>> # Add attributes to node and set their values
... n.NewAttribute('foo').Set('Hello')
>>> n.NewAttribute('moo').Set('World')
>>> # Show XML
... n.GetXML()
'<me:myExample xmlns:me="http://example.com/myExample" foo="Hello" moo="World"/>'
>>> # get attributes of this XMLNode
... a = arcom.get_attributes(n)
>>> # result is a dictionary
... type(a)
<type 'dict'>
>>> # show dictionary
... a
{'foo': 'Hello', 'moo': 'World'}
>>> # extracting an attribute value
... a['moo']
'World'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\newpage
\begin{itemize}
  \item{ \verb$get_child_nodes(node)$\newline
  Gets children of the supplied XMLNode (node).
  Result will be returned as a list of XMLNodes.
\begin{verbatim}
In the example below, the following structure will be created:

<a>A
    <b>B</b>
    <c>C
        <e>E</e>
        <f>F</f>
    </c>
    <d>D
        <g>G</g>
    </d>
<a>

Relationships:

        a       children: (b, c, d)
       /|\      
      / | \     
     b  c  d    children: () (e, f) (g)
       / \  \   
      e   f  g  children: () () ()
\end{verbatim}
    \begin{example}
      \caption{Getting child nodes}\label{getchld}
\begin{verbatim}
>>> import arc
>>> import arcom
>>> # Create root node
... a = arc.XMLNode(arc.NS(),'a')
>>> # Create child nodes for 'a'
... b = a.NewChild('b')
>>> c = a.NewChild('c')
>>> d = a.NewChild('d')
>>> # Create child nodes for 'c'
... e = c.NewChild('e')
>>> f = c.NewChild('f')
>>> # Create child node for 'd'
... g = d.NewChild('g')
>>> # show XML
... a.GetXML()
'<a><b/><c><e/><f/></c><d><g/></d></a>'
>>> # get child nodes for 'a'
... tmp = arcom.get_child_nodes(a)
>>> # result is a list
... type(tmp)
<type 'list'>
>>> # 'a' has 3 children
... len(tmp)
3
>>> # show name for each
... ', '.join(x.Name() for x in tmp)
'b, c, d'
>>> # 'b' has no children
... tmp = arcom.get_child_nodes(b)
>>> len(tmp)
0
>>> # 'c' has 2 children: 'e' and 'f'
... tmp = arcom.get_child_nodes(c)
>>> len(tmp)
2
>>> ', '.join(x.Name() for x in tmp)
'e, f'
>>> # 'd' has 1 child: 'g'
... tmp = arcom.get_child_nodes(d)
>>> len(tmp)
1
>>> ', '.join(x.Name() for x in tmp)
'g'
>>> # nodes 'e', 'f' and 'g' have no children
... tmp = arcom.get_child_nodes(e)
>>> len(tmp)
0
>>> tmp = arcom.get_child_nodes(f)
>>> len(tmp)
0
>>> tmp = arcom.get_child_nodes(g)
>>> len(tmp)
0
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \verb$get_child_values_by_name(node, name)$\newline
  Gets values from children of a supplied XMLNode (node),
  where those children have the supplied name (name).
    \begin{example}
      \caption{Get values of specified children}\label{getchdval}
\begin{verbatim}
In the example below, the following structure will be created:

<node>
    <same>firstEQ</same>
    <same>secondEQ</same>
    <different>DIFF</different>
</node>

>>> import arc
>>> import arcom
>>> # Create XMLNode
... n = arc.XMLNode(arc.NS(),'node')
>>> # Create 3 child nodes (two of which get the same name)
... x = n.NewChild('same')
>>> y = n.NewChild('same')
>>> z = n.NewChild('different')
>>> # Set values for nodes
... x.Set('firstEQ')
>>> y.Set('secondEQ')
>>> z.Set('DIFF')
>>> # Show XML
... n.GetXML()
'<node><same>firstEQ</same><same>secondEQ</same><different>DIFF</different></node>'
>>> # Get child values where name is 'same'
... tmp = arcom.get_child_values_by_name(n,'same')
>>> # Result is a list
... type(tmp)
<type 'list'>
>>> # Show result
... tmp
['firstEQ', 'secondEQ']
>>> # Get child values where name is 'different'
... tmp = arcom.get_child_values_by_name(n,'different')
>>> # Show result
... tmp
['DIFF']
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \verb$datapoint_from_url(url_string)$\newline
  Creates DataPoint from specified URL (\verb$url_string$).
    \begin{example}
      \caption{Creating DataPoint from URL}\label{url2dp}
\begin{verbatim}
>>> import arc
>>> import arcom
>>> tmpList = []
>>> status = ''
>>> # create DataPoint from a local directory
... dp = arcom.datapoint_from_url('file:///usr/local/share/arc')
>>> # list files
... (files, stat) = dp.ListFiles()
>>> # if it is not empty
... if files:
>>>     status = 'found'
>>>     # for all the entries, get type and name
...     for f in files:
>>>         if (f.GetType() == arc.FileInfo.file_type_file):
>>>             type = 'file'
>>>         elif (f.GetType() == arc.FileInfo.file_type_dir):
>>>             type = 'dir'
>>>         else:
>>>             type = 'unknown'
>>>         # get results together in a list
...         tmpList.append(f.GetName() + ' (' + type + ')\n')
>>> else:
>>>     status = 'Could not access data. Reason: %s' % str(stat)
>>> 
>>> # see result
... str(stat)
'Operation completed successfully'
>>> # show list
... tmpList
['nordugrid.schema (unknown)\n']
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \verb$parse_url(url)$\newline
  Parses a URL. Gets protocol, host, port and path.
    \begin{example}
      \caption{Parsing a URL}\label{parseurl}
\begin{verbatim}
>>> import arcom
>>> proto, host, port, path = arcom.parse_url('boo://no.one.here:123/foo')
>>> proto
'boo'
>>> host
'no.one.here'
>>> port
123
>>> path
'foo'
\end{verbatim}
    \end{example}
  }
\end{itemize}

\end{flushleft}

\subsection{arcom.client}
\label{aclient}
Module ''arcom.client'' contains a base Client class for sending SOAP messages to services.
\begin{flushleft}
\begin{itemize}
  \item{ \verb$Client(url, ns, print_xml = False, xmlnode_class = arc.XMLNode, ssl_config = {})$\newline
  This is the constructor of the Client class.
  \begin{itemize}
    \item{url is the URL of the service, it could be a list of URLs}
    \item{ns contains the namespaces we want to use with each message}
    \item{\verb$print_xml$ is for debugging, prints all the SOAP messages to the screen}
    \item{\verb$xmlnode_class$ is the XML node class to be used; by default, it is ''arc.XMLNode''}
    \item{\verb$ssl_config$ is the SSL configuration to be used for secure connection; it is a dictionary that contains information about client (\verb$proxy_file$ or \verb$key_file$ and \verb$cert_file$) and CAs (\verb$ca_file$ or \verb$ca_dir$)}
  \end{itemize}
  }
\end{itemize}
\begin{itemize}
  \item{ \verb$call(tree, return_tree_only = False)$\newline
    Creates a SOAP message from an XMLTree and sends it to the service. \newline
    - tree is an XMLTree object containing the content of the request \newline
    - \verb$return_tree_only$ indicates that we only need to put the response into an XMLTree
    \begin{example}
      \caption{Creating a client and calling the echo service (XMLTree)}\label{clientex1}
\begin{verbatim}
>>> import arc
>>> import arcom
>>> # Import the Client class
... Client = arcom.import_class_from_string('arcom.client.Client')
>>> # Import the XMLTree class
... XMLTree = arcom.import_class_from_string('arcom.xmltree.XMLTree')
>>> # Create namespace - it will be used for the message sent
... ns = arc.NS({'echo':'urn:echo'})
>>> # Create client
... c = Client('http://your.server.example.com:50000/Echo',ns,print_xml=True)
>>> # Create message
... msg = XMLTree(from_tree = ('echo:echo',[('echo:say', 'Hello, World!')]))
>>> # Let the client do what it is meant for
... # Note that we created the client with print_xml=True
... # so both request and response will be displayed in an easy-to-read form
... c.call(msg)
Request:
        #   <soap-env:Body>
        #       <echo:echo>
        #           <echo:say>Hello, World!</echo:say>
        #       </echo:echo>
        #   </soap-env:Body>

Response:
        #   <soap-env:Envelope>
        #       <soap-env:Body>
        #           <echo:echoResponse>
        #               <echo:hear>[ Hello, World! ]</echo:hear>
        #           </echo:echoResponse>
        #       </soap-env:Body>
        #   </soap-env:Envelope>

'<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap
/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://
www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"><soap-en
v:Body><echo:echoResponse><echo:hear>[ Hello, World! ]</echo:hear></echo:echoResponse></s
oap-env:Body></soap-env:Envelope>'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \verb$call_raw(outpayload)$\newline
  Send a POST request with the SOAP XML message.\newline
  - outpayload is an XMLNode with the SOAP message
    \begin{example}
      \caption{Creating a client and calling the echo service (SOAP)}\label{clientex2}
\begin{verbatim}
>>> import arc
>>> import arcom
>>> # Import the Client class
... Client = arcom.import_class_from_string('arcom.client.Client')
>>> # Create namespace - it will be used for the message sent
... ns = arc.NS({'echo':'urn:echo'})
>>> # Create client
... c = Client('http://arctest.ki.iif.hu:50000/Echo',ns,print_xml=True)
>>> # Create SOAP Payload
... pl = arc.PayloadSOAP(ns)
>>> # Create message structure and set content
... pl.NewChild('echo:echo',ns).NewChild('echo:say',ns).Set('Hello, World!')
>>> # Let client do the call; see response
... c.call_raw(pl)
'<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap
/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://
www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"><soap-en
v:Body><echo:echoResponse><echo:hear>[ Hello, World! ]</echo:hear></echo:echoResponse></s
oap-env:Body></soap-env:Envelope>'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

\subsection{arcom.logger}
\label{alogger}
Module ''arcom.logger'' provides simple access to the logging capability of ARC.
\begin{flushleft}
\begin{itemize}
  \item{ \verb$log_levels$\newline
  There are six log levels at the moment:
    \begin{itemize}
      \item{arc.VERBOSE}
      \item{arc.DEBUG}
      \item{arc.INFO}
      \item{arc.WARNING}
      \item{arc.ERROR}
      \item{arc.FATAL}
    \end{itemize}
  }
\end{itemize}
\begin{itemize}
  \item{ \verb$get_logger(system = '<UNKNOWN>')$\newline
  Creates a logger - with root logger as a parent - on a given subdomain (system). 
  If no such is given, it will be ''\verb$<UNKNOWN>$'' by default.
    \begin{example}
      \caption{Using the Logger}\label{loggerex}
\begin{verbatim}
>>> import arc
>>> import arcom
>>> import sys
>>> # Import get_logger function
... from arcom.logger import get_logger
>>> # Create logger
... rl = get_logger()
>>> rl.logger
<arc.Logger; proxy of <Swig Object of type 'Arc::Logger *' at 0x9ef3f60> >
>>> arclogger = rl.logger
>>> # Add new destination
... arclogger.addDestination(arc.LogStream(sys.stdout))
>>> # Log a message
... # Note that the message is also returned as an output
... rl.msg(arc.INFO, 'Hello, World!')
[2009-06-04 11:00:42] [Arc.<UNKNOWN>] [INFO] [13120/148748400] Hello, World!
'Hello, World!'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

\subsection{arcom.security}
\label{asecurity}
Module ''arcom.security'' provides tools for managing simple authorisation policies and requests.
This includes policy format conversion (between storage and ARC policy), utility functions 
for accessing the decision making mechanism and retrieving credential information.
\begin{flushleft}

\begin{itemize}
  \item{ \verb$class AuthRequest$\newline
  Represents simple authorisation request.
  Conversion is limited to ARCAuth format at the moment.
  \begin{itemize}
    \item{ \verb$__init__ - construction of AuthRequest$\newline
      AuthRequest is created from the incoming message.
      Before the message reaches a service it is passed through several message chain components.
      (See ARC HED documentation\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/tech\_doc/hed/ARCHED\_article.pdf}} for details.)
      During this, information specific to the MCC is probably added to the message.
      Specifically, when it is passed through the TLS MCC, then security information is added.
      (See security documentation\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/tech\_doc/sec/SecurityFrameworkofARC1.tex}} 
       for details about security design in ARC.)
      This information is then retrieved when AuthRequest is created.
      \begin{example}
        \caption{Identity part of subject retrieved when TLS is not used}\label{asubnotls}
\begin{verbatim}
     <ra:Subject xmlns:ra="http://www.nordugrid.org/schemas/request-arc">
       ...
       <ra:SubjectAttribute 
         AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/identity" 
         Type="string">ANONYMOUS</ra:SubjectAttribute>
     </ra:Subject>
\end{verbatim}
        When TLS MCC is not included in the message chain, identity retrieval will fail,
        thus showing up with an anonymous user.
      \end{example}
    }
  \end{itemize}
  \begin{itemize}
    \item{ \verb$get_request(self, action, format = 'ARCAuth')$\newline
      Returns a request in a XML string for the given action. This action should be one of the storage actions. 
      These are 'read', 'addEntry', 'removeEntry', 'delete', 'modifyPolicy', 'modifyStates' and 'modifyMetadata'.
      Format should be ARCAuth at the moment, as other formats are not supported yet.
    }
  \end{itemize}
  \begin{itemize}
    \item{ \verb$get_identity(self)$\newline
      Returns identity information.
    }
  \end{itemize}
  \begin{itemize}
    \item{ \verb$get_identity_and_ca(self)$\newline
      Returns identity and CA information.
    }
  \end{itemize}
  }
\end{itemize}

\begin{itemize}
  \item{ \verb$class AuthPolicy$\newline
  Class to set / retrieve / convert simple policies.
    \begin{example}
      \caption{Creating policy}\label{policyex}
\begin{verbatim}
>>> from arcom.security import AuthPolicy
>>> # create policy object
... p = AuthPolicy()
>>> # set policy for Alice and Bob
... # Alice is allowed to read and add entries (addEntry)
... # Bob is only allowed to read
... AlicePol = ('Alice','+read +addEntry')
>>> BobPol = ('Bob','+read')
>>> p.set_policy([AlicePol,BobPol])
>>> # see Policy document
... print p.get_policy('ARCAuth')
<Policy xmlns="http://www.nordugrid.org/schemas/policy-arc" 
        CombiningAlg="Deny-Overrides">
  <Rule Effect="Permit">
    <Description>Alice is allowed to read, addEntry</Description>
    <Subjects>
      <Subject>
        <Attribute AttributeId="http://www.nordugrid.org/schemas/policy-arc/
types/tls/identity" Type="string">Alice</Attribute>
      </Subject>
    </Subjects>
    <Actions>
      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/
types/storage/action" Type="string">read</Action>
      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/
types/storage/action" Type="string">addEntry</Action>
    </Actions>
  </Rule>
  <Rule Effect="Permit">
    <Description>Bob is allowed to read</Description>
    <Subjects>
      <Subject>
        <Attribute AttributeId="http://www.nordugrid.org/schemas/policy-arc/
types/tls/identity" Type="string">Bob</Attribute>
      </Subject>
    </Subjects>
    <Actions>
      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/
types/storage/action" Type="string">read</Action>
    </Actions>
  </Rule>
</Policy>
\end{verbatim}
    \end{example}
  }
\end{itemize}

\begin{itemize}
  \item{ \verb$make_decision(policy, request)$\newline
  Method for accessing decision making functionality of ARC.
  ARC Evaluator is used to make decision about the request
  according to supplied policy.
    \begin{example}
      \caption{Decision making}\label{decisionex}
\begin{verbatim}

... 
>>> dsa = DummySecAttr('Alice')
>>> # set 'thisown' to False to avoid problems
... dsa.thisown = False
>>> 
>>> # export XML in ARCAuth format
... ex = dsa.Export()
>>> 
>>> from arcom.security import AuthPolicy
>>> # create policy object
... p = AuthPolicy()
>>> # set policy for Alice
... # Alice is allowed to read and add entries (addEntry)
... AlicePol = ('Alice','+read +addEntry')
>>> p.set_policy([AlicePol])
>>> 
>>> from arcom.security import make_decision
>>> 
>>> # get policy XML
... px = p.get_policy()
>>> rx = ex.GetXML()
>>> 
>>> decision = make_decision(px, rx)
>>> 
>>> # Possible results:
... # arc.DECISION_PERMIT -- 0
... # arc.DECISION_DENY -- 1
... # arc.DECISION_INDETERMINATE -- 2
... # arc.DECISION_NOT_APPLICABLE -- 3
... 
>>> decision
0
\end{verbatim}
      In this example, Alice requests read for a temporary file.
      This request is made via DummySecAttr, when Export is called.
      (Details about DummySecAttr can be found in Appendix \ref{DummySecAttrCode} of this document.)
    \end{example}
  }
\end{itemize}

\begin{itemize}
  \item{ \verb$parse_ssl_config(cfg)$\newline
  Method for processing an XML node and retrieve SSL configuration information.
  First it looks for a ClientSSLConfig child of the supplied node, 
  then it either collects information from a file given in FromFile attribute
  or from other children of ClientSSLConfig, namely KeyPath, CertificatePath 
  and CACertificatesDir.\newline
  So if our ARC config looked like
\begin{verbatim}
<?xml version="1.0"?>
<ArcConfig 
  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
>
    ...
    <Chain>
        ...
        <Service name="pythonservice" id="bartender">
            <ClassName>storage.bartender.bartender.BartenderService</ClassName>
            <LibrarianURL>https://localhost:60000/Librarian</LibrarianURL>
            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
        </Service>
        ...
    </Chain>
</ArcConfig>
\end{verbatim}
  and ''/etc/arc/clientsslconfig.xml'' contained the following lines:
\begin{verbatim}
<?xml version="1.0"?>
<ClientSSLConfig>
    <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
    <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
    <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
</ClientSSLConfig>
\end{verbatim}
  then parsing would produce results like in the example below.
    \begin{example}
      \caption{SSL config example}\label{sslconfex}
\begin{verbatim}
>>> import arc
>>> import arcom
>>> from arcom.security import parse_ssl_config
>>> 
>>> xml_str = file('/etc/arc/ssl_config_example.xml').read()
>>> configNode = arc.XMLNode(xml_str)
>>> 
>>> serviceNode = configNode.Get('Chain').Get('Service')
>>> 
>>> print parse_ssl_config(serviceNode)
{'key_file': '/etc/grid-security/hostkey.pem', 
 'cert_file': '/etc/grid-security/hostcert.pem', 
 'ca_file': '/etc/grid-security/certificates'}
\end{verbatim}
    \end{example}
  }
\end{itemize}

\end{flushleft}

\subsection{arcom.service}
\label{aservice}
Module ''arcom.service'' provides the Service class and other tools for
service development. The Service class deals with Trust Manager
and SSL configuration and has the process method to get requests from 
incoming messages and create outgoing ones from the results.
This class should be extended when creating a new service.
Note that security related parts will not be discussed here; for those please see 
the security documentation\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/tech\_doc/sec/SecurityFrameworkofARC1.tex}} 
of ARC.
\begin{flushleft}

\begin{itemize}
  \item{ \verb$Service$ \linebreak
  Service class will be presented through an example, in which DummyService will provide
  access to the funcionality of the Dummy class. (Complete source code can be found in Appendix 
  \ref{DummyServiceCode} of this document.)\linebreak
  Note that a small change in AuthRequest class (found in ''security.py'') is required for 
  this example to work because the incoming message in our example could not be processed
  in a way like on the server side. Therefore it is assumed that the \verb$__init__$ method
  of AuthRequest has:
\begin{verbatim}
        auth = message.Auth()
        import arc
        try:
            xml = auth.Export(arc.SecAttr.ARCAuth)
            subject = xml.Get('RequestItem').Get('Subject')
        except:
            subject = arc.XMLNode(arc.NS({'ra':request_ns}), 'ra:Subject')
\end{verbatim}
  instead of \linebreak
\begin{verbatim}
        auth = message.Auth()
        import arc
        xml = auth.Export(arc.SecAttr.ARCAuth)
        subject = xml.Get('RequestItem').Get('Subject')
\end{verbatim}
  so an empty Subject is available if auth is not present.
    \begin{example}
      \caption{Using DummyService - an example service based on arcom.service.Service}\label{srvex}
\begin{verbatim}
>>> # Reminder:
... # import arc
... #
... # class Dummy:
... #   def foo(self, foomsg = ''):
... #     return ''.join(['Foo message is: ',str(foomsg)])
... # 
... # from arcom.service import Service
... # 
... # class DummyService(Service):
... #   """ DummyService class based on Service class in arcom.service """
... #   ...
... 
>>> # Create instance of DummyService
... ds = DummyService(None)
>>> # Create namespace - it will be used for the message sent
... ns = arc.NS({'dmmy':'urn:foo'})
>>> # Create SOAP Envelope
... se = arc.SOAPEnvelope(ns)
>>> # Create content
... se.NewChild('dmmy:foo').Set('Foo!')
>>> # Create SOAP Message
... msg = arc.SOAPMessage()
>>> # Payload
... pls = arc.PayloadSOAP(se)
>>> # Set Message Payload
... msg.Payload(pls)
>>> # create outgoing message that will hold response
... outse = arc.SOAPEnvelope(ns)
>>> outmsg = arc.SOAPMessage()
>>> outpls = arc.PayloadSOAP(outse)
>>> outmsg.Payload(pls)
>>> # call service
... ds.process(msg, outmsg)
<arc.MCC_Status; proxy of <Swig Object of type 'Arc::MCC_Status *' at 0x8a39078> >
>>> # show result
... print outmsg.Payload().GetXML()
<soap-env:Envelope xmlns:dmmy="urn:foo" xmlns:soap-enc="http://schemas.xmlsoap.o
rg/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xm
lns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XML
Schema-instance"><soap-env:Body><dmmy:fooResponse>Foo message is: Foo!</dmmy:foo
Response></soap-env:Body></soap-env:Envelope>
\end{verbatim}
    \end{example}
  }
\end{itemize}

\begin{itemize}
  \item{ \verb$parse_node(node, names, single = False, string = True)$ \linebreak
  Calls \verb$node_to_data()$ for each child of the given node.
  \begin{itemize}
    \item{\verb$node$ is the XMLNode whose children we want to convert}
    \item{\verb$names$ is a list of tag names which will be returned in the specified order}
    \item{\verb$single$ indicates whether only one value is needed beside the key; if so, do not put it into a list}
    \item{\verb$string$ indicates whether string contents of nodes are needed - not the nodes themselves}
  \end{itemize}
    \begin{example}
      \caption{parse\_node}\label{pnex}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.service import parse_node
>>> 
>>> # create xml node
... xml = XMLNode('''
... <myList>
...   <myElement>
...     <myID>0</myID>
...     <LN>/</LN>
...   </myElement>
...   <myElement>
...     <myID>1</myID>
...     <LN>/testfile</LN>
...   </myElement>
... </myList>''')
>>> 
>>> # default: single = False and string = True
... parse_node(xml, ['myID','LN'])
{'1': ['/testfile'], '0': ['/']}
>>> # single = True
... parse_node(xml, ['myID','LN'], single = True)
{'1': '/testfile', '0': '/'}
>>> parse_node(xml, ['myID','LN'], True)
{'1': '/testfile', '0': '/'}
>>> # string = False
... parse_node(xml, ['myID','LN'], string = False)
{<arc.XMLNode; proxy of <Swig Object of type 'Arc::XMLNode *' at 0x996f040> >: [
<arc.XMLNode; proxy of <Swig Object of type 'Arc::XMLNode *' at 0x996f050> >], <
arc.XMLNode; proxy of <Swig Object of type 'Arc::XMLNode *' at 0x9997590> >: [<a
rc.XMLNode; proxy of <Swig Object of type 'Arc::XMLNode *' at 0x99975a0> >]}
\end{verbatim}
    \end{example}
  }
\end{itemize}

\begin{itemize}
  \item{ \verb$parse_to_dict(node, names)$ \linebreak
  Converts the children of the node to a dictionary of dictionaries.
  \begin{itemize}
    \item{\verb$node$ is the XMLNode whose children we want to convert}
    \item{\verb$names$ is a list of tag names; for each child only these names will be included in the dictionary}
  \end{itemize}
  Note that the first element of ''names'' is considered as sthe child whose value will serve as the key of dictionary.
  Other elements are put in an inner dictionary with their names being the key and their content being the value for that key.
    \begin{example}
      \caption{parse\_to\_dict}\label{p2dex}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.service import parse_to_dict
>>> 
>>> # create xml node
... xml = XMLNode('''
... <myList>
...   <myElement>
...     <myID>123</myID>
...     <refID>abc</refID>
...     <state>alive</state>
...     <size>123456</size>
...   </myElement>
...   <myElement>
...     <myID>456</myID>
...     <refID>fed</refID>
...     <state>alive</state>
...     <size>987</size>
...   </myElement>
... </myList>''')
>>> 
>>> # example 1 - 'myID' as key
... # 'state' and 'size' goes to the inner dictionary
... parse_to_dict(xml, ['myID','state','size'])
{'123': {'state': 'alive', 'size': '123456'}, '456': {'state': 'alive', 'size': 
'987'}}
>>> 
>>> # example 2 - 'myID' as key
... # 'myID' also goes to the inner dictionary
... parse_to_dict(xml, ['myID','myID','state','size'])
{'123': {'myID': '123', 'state': 'alive', 'size': '123456'}, '456': {'myID': '45
6', 'state': 'alive', 'size': '987'}}
\end{verbatim}
    \end{example}
  }
\end{itemize}

\begin{itemize}
  \item{ \verb$create_response(method_name, tag_names, elements, payload, single = False)$ \linebreak
  Creates a SOAP XML payload from a dictionary of tag names and list of values.
  \begin{itemize}
    \item{\verb$method_name$ is the name of the method which will be used as a prefix in the name of the ''Response'' tag}
    \item{\verb$tag_names$ is a list of names which will be used in the specified order as tag names}
    \item{\verb$elements$ is a dictionary where the key will be tagged as the first tag name 
and the value is a list whose items will be tagged in the order of the \verb$tag_names$ list}
    \item{\verb$payload$ is an XMLNode that the response will be added to}
    \item{\verb$single$ indicates whether there is only one value per key}
  \end{itemize}  
    \begin{example}
      \caption{create\_response}\label{crespex}
\begin{verbatim}
>>> import arc
>>> from arcom.service import create_response
>>> 
>>> method_name = 'dummy'
>>> # create tag_names; 'myID' will be the key
... tag_names = ['myID', 'state', 'size']
>>> # create elements
... # elements of the lists in this dictionary are values and will be tagged 
... # according to names found in tag_names except for the first element
... # e.g.: 'alive' will be tagged with 'state'
... elements = {'123': ['alive', '123456'], '456': ['alive', '987']}
>>> # create empty payload
... payload = arc.PayloadSOAP(arc.NS())
>>> 
>>> # create response
... response = create_response(method_name, tag_names, elements, payload)
>>> 
>>> # show it
... print response.GetXML(True)
<soap-env:Envelope xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xm
lns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w
3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <dummyResponse>
      <dummyResponseList>
        <dummyResponseElement>
          <myID>123</myID>
          <state>alive</state>
          <size>123456</size>
        </dummyResponseElement>
        <dummyResponseElement>
          <myID>456</myID>
          <state>alive</state>
          <size>987</size>
        </dummyResponseElement>
      </dummyResponseList>
    </dummyResponse>
  </soap-env:Body>
</soap-env:Envelope>
\end{verbatim}
    \end{example}
  }
\end{itemize}

\begin{itemize}
  \item{ \verb$node_to_data(node, names, single = False, string = True)$ \linebreak
  Get some children of an XMLNode and return them in a list in the specified order using the first one as a key.
  \begin{itemize}
    \item{\verb$node$ is an XMLNode that has some children}
    \item{\verb$names$ is a list of the names of those children we want to extract; first string in this list will always be the key}
    \item{\verb$single$ indicates whether only one value is needed; if so, do not put it into a list}
    \item{\verb$string$ indicates whether string contents of nodes are needed - not the nodes themselves}
  \end{itemize}  
    \begin{example}
      \caption{node\_to\_data - 1}\label{n2dex1}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.service import node_to_data
>>> 
>>> # create node
... xml = XMLNode('''
... <myElement>
...   <myID>0</myID>
...   <refID>abc</refID>
...   <section>states</section>
...   <property>spares</property>
...   <value>2</value>
...   <dummy>foo</dummy>
...   <changeType>set</changeType>
... </myElement>
... ''')
>>> 
>>> # create names
... # 'myID' will be the key
... # all the other names will be in a list in the specified order
... # 'dummy' will be left out
... names = ['myID', 'refID', 'changeType', 'section', 'property', 'value']
>>> 
>>> # call node_to_data
... node_to_data(xml,names)
('0', ['abc', 'set', 'states', 'spares', '2'])
\end{verbatim}
    \end{example}
    \begin{example}
      \caption{node\_to\_data - 2}\label{n2dex2}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.service import node_to_data
>>> 
>>> # create node
... xml = XMLNode('''
... <getRequest>
...   <GUID>11</GUID>
...   <myID>99</myID>
... </getRequest>
... ''')
>>> 
>>> # create names
... # 'myID' will be the key
... names = ['myID', 'GUID']
>>> 
>>> # call node_to_data
... # GUID is the single value so do not put it in a list
... node_to_data(xml,names, True)
('99', '11')
\end{verbatim}
    \end{example}
  }
\end{itemize}

\begin{itemize}
  \item{ \verb$get_data_node(node)$ \linebreak
  Gets the data node from the \verb$Body$ of a SOAP message where the first child node 
  of the \verb$Body$ refers to the method being called. E.g.:
\begin{verbatim}
Request:
        <soap-env:Envelope>
          <soap-env:Body>
            <method>
              <data>Value</data>
            </method>
          </soap-env:Body>
        </soap-env:Envelope>

Data node:
          <data>Value</data>
\end{verbatim}
    \begin{example}
      \caption{get\_data\_node}\label{gdnex}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.service import get_data_node
>>> 
>>> # create node
... xml = XMLNode('''
... <Envelope>
...   <Body>
...     <method>
...       <data>Value</data>
...     </method>
...   </Body>
... </Envelope>
... ''')
>>> 
>>> # call get_data_node
... dn = get_data_node(xml)
>>> 
>>> # show result
... dn.GetXML()
'<data>Value</data>'
\end{verbatim}
    \end{example}
  }
\end{itemize}

\end{flushleft}

\subsection{arcom.threadpool}
\label{athreadpool}
Module ''arcom.threadpool'' provides simple thread pool support.
Consists of classes ThreadPool, ThreadPoolThread and ReadWriteLock.
ThreadPool could be used to create a pool of threads, then it accepts 
tasks that will be dispatched to the next available thread.
ThreadPoolThread could be used to retrieve a task and execute it.
ReadWriteLock is a lock object that allows many simultaneous ''read locks'', 
but only one ''write lock.''

\begin{flushleft}

\begin{itemize}
  \item{ \verb$class ThreadPool$ \linebreak
  ThreadPool is a flexible thread pool class. It creates a pool of threads, 
  then accepts tasks that will be dispatched to the next available thread.
  }
  \begin{itemize}
    \item{ \verb$setThreadCount(self, newNumThreads)$ \linebreak
    Sets the current pool size.  It does so by acquiring the resizing lock, 
    then calling the private version of this method to grow or shrink the pool.
    }
  \end{itemize}

  \begin{itemize}
    \item{ \verb$getThreadCount(self)$ \linebreak
    Returns the number of threads in the pool.
    }
  \end{itemize}

  \begin{itemize}
    \item{ \verb$queueTask(self, task, args=None, taskCallback=None)$ \linebreak
    Inserts a task into the queue. Task must be callable; 
    args and taskCallback may be None.
    }
  \end{itemize}

  \begin{itemize}
    \item{ \verb$getNextTask(self)$ \linebreak
    Retrieves the next task from the task queue.  For use only by ThreadPoolThread 
    objects contained in the pool.
    }
  \end{itemize}

  \begin{itemize}
    \item{ \verb$joinAll(self, waitForTasks = True, waitForThreads = True)$ \linebreak
    Clears the task queue and terminates all pooled threads,
    optionally allowing the tasks and threads to finish.
    }
  \end{itemize}

\end{itemize}

\begin{itemize}
  \item{ \verb$class ThreadPoolThread(threading.Thread)$ \linebreak
  ThreadPoolThread is a pooled thread class.
  }
  \begin{itemize}
    \item{ \verb$run(self)$ \linebreak
    Until told to quit, retrieve the next task and execute it, calling the callback if any.
    }
  \end{itemize}

  \begin{itemize}
    \item{ \verb$goAway(self)$ \linebreak
    Exit the run loop next time through.
    }
  \end{itemize}

\end{itemize}

\begin{itemize}
  \item{ \verb$class ReadWriteLock$ \linebreak
  ReadWriteLock is a lock object that allows many simultaneous ''read locks'', 
  but only one ''write lock''.
  }
  \begin{itemize}
    \item{ \verb$acquire_read(self)$ \linebreak
    Acquires a read lock. Blocks only if a thread has acquired the write lock.
    }
  \end{itemize}

  \begin{itemize}
    \item{ \verb$release_read(self)$ \linebreak
    Releases a read lock.
    }
  \end{itemize}

  \begin{itemize}
    \item{ \verb$acquire_write(self)$ \linebreak
    Acquires a write lock. Blocks until there are no acquired read or write locks.
    }
  \end{itemize}

  \begin{itemize}
    \item{ \verb$release_write(self)$ \linebreak
    Releases a write lock.
    }
  \end{itemize}

\end{itemize}

  Two examples are pressented below. In the first one, four waitTasks with different 
  wait times are queued while there are three threads. In the second one, a counter
  is incremented by four addTasks on four threads. (Here a ReadWriteLock is used to
  make sure that only one thread has write access to the counter at a time.)
  (Complete source code can be found in Appendix \ref{ThreadPoolTestCode} of this document.)
  \begin{example}
    \caption{arcom.threadpool}\label{arctpex}
\begin{verbatim}
$./threadpooltest.py

Example1 - waitTask
(1): WaitTask starting
WaitTask sleeping for 4 seconds
(2): WaitTask starting
WaitTask sleeping for 1 seconds
(3): WaitTask starting
WaitTask sleeping for 6 seconds
Callback called for Waiter (2)
(4): WaitTask starting
WaitTask sleeping for 2 seconds
Callback called for Waiter (4)
Callback called for Waiter (1)
Callback called for Waiter (3)

Example2 - addTask
COUNTER before join:  0
(5): Added 10 to counter, counter is now 10
(6): Added 10 to counter, counter is now 20
(7): Added 10 to counter, counter is now 30
(8): Added 10 to counter, counter is now 40
(6): Added 10 to counter, counter is now 50
(5): Added 10 to counter, counter is now 60
(7): Added 10 to counter, counter is now 70
(8): Added 10 to counter, counter is now 80
(5): Added 10 to counter, counter is now 90
(7): Added 10 to counter, counter is now 100
(6): Added 10 to counter, counter is now 110
(8): Added 10 to counter, counter is now 120
Callback called for ('addTask', 10)
Callback called for ('addTask', 10)
Callback called for ('addTask', 10)
Callback called for ('addTask', 10)
COUNTER after join:  120
\end{verbatim}
  \end{example}

\end{flushleft}

\subsection{arcom.XMLTree}
\label{axtree}
Module ''arcom.XMLTree'' provides the XMLTree class to convert from XML to native Python structures and vice versa.
Furthermore, it provides utility functions and some basic query methods for XMLTree structures.
\begin{flushleft}

\begin{itemize}
  \item{ \verb$class XMLTree$ \linebreak
    With the help of the XMLTree class XML can be converted to native Python structures and vice versa.
    It also provides some useful functions to handle these structures.
    \begin{itemize}
      \item{\begin{verbatim}
XMLTree(from_node = None, from_string = '', from_tree = None, rewrite = {}, 
        forget_namespace = False) \end{verbatim}
        \begin{itemize}
          \item{\verb$from_node$ is an XMLNode that could be converted to XMLTree}
          \item{\verb$from_string$ is a string representation of an XMLNode that could be converted to XMLTree}
          \item{\verb$from_tree$ is a tree structure or an XMLTree object that could be converted to XMLTree}
          \item{\verb$rewrite$ is a dictionary; if an XML node has a name which is a key 
                in this dictionary then it will be renamed as the value of that key; note though, 
                when \verb$from_tree$ is used, this parameter is ignored}
          \item{\verb$forget_namespace$ tells whether the XMLTree should not contain the namespace prefixes; 
                note though, when \verb$from_tree$ is used, this parameter is ignored}
        \end{itemize}
        The parameter \verb$from_tree$ has the highest priority; if it is not None, then \verb$from_string$ and 
        \verb$from_node$ are ignored. If \verb$from_tree$ is None but \verb$from_node$ is given, 
        then \verb$from_string$ is ignored. If only \verb$from_string$ is given, then it will be the choosen one.
      }
    \end{itemize}
    Some examples of creating an XMLTree are shown below.
    \begin{example}
      \caption{Creating an XMLTree}\label{crxt}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create XMLTree sources
... # xmlstr - will be used to create an XMLTree from string
... # xmlnodestr - will be used to create 'node' (an XMLNode)
... # node - will be used to create an XMLTree from XMLNode
... # treestruct - for creating XMLTree from tree structure
... xmlstr = '<echo><say>Hello!</say></echo>'
>>> xmlnodestr = '<node><subnode>Subnode</subnode></node>'
>>> node = XMLNode(xmlnodestr)
>>> treestruct = ('root', [('leaf','Leaf')])
>>> 
>>> # from_tree has the highest priority
... t1 = XMLTree(from_node = node, from_string = xmlstr, from_tree = treestruct) 
>>> t1.get()
[('root', [('leaf', 'Leaf')])]
>>> 
>>> # no from_tree supplied; from_node will be used
... t2 = XMLTree(from_node = node, from_string = xmlstr)
>>> t2.get()
[('node', [('subnode', 'Subnode')])]
>>> 
>>> # from_string only
... t3 = XMLTree(from_string = xmlstr)
>>> t3.get()
[('echo', [('say', 'Hello!')])]
\end{verbatim}
    \end{example}
    \begin{example}
      \caption{XMLTree - forget namespace}\label{xtfn}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.xmltree import XMLTree
>>> 
>>> # Create XMLTree objects to demonstrate
... # the use of forget namespace feature
... # node - is an XMLNode
... # tree - is an XMLTree that is created from 'node'
... node = XMLNode('<soap-env:Envelope xmlns:hash="urn:hash" \
... xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" \
... xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" \
... xmlns:xsd="http://www.w3.org/2001/XMLSchema" \
... xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">\
... <soap-env:Body>\
... <hash:get>\
... <hash:IDs>\
... <hash:ID>0</hash:ID>\
... <hash:ID>1</hash:ID>\
... <hash:ID>2</hash:ID>\
... </hash:IDs>\
... </hash:get>\
... </soap-env:Body>\
... </soap-env:Envelope>')
>>> 
>>> tree = XMLTree(node)
>>> # show tree structure
... print tree
('soap-env:Envelope', [('soap-env:Body', [('hash:get', [('hash:IDs', [('hash:ID
', '0'), ('hash:ID', '1'), ('hash:ID', '2')])])])])
>>> 
>>> # forget_namespace is ignored when from_tree is used
... tree2 = XMLTree(from_tree = tree, forget_namespace = True)
... print tree2
('soap-env:Envelope', [('soap-env:Body', [('hash:get', [('hash:IDs', [('hash:ID
', '0'), ('hash:ID', '1'), ('hash:ID', '2')])])])])
>>> 
>>> # forget_namespace with from_node
... tree3 = XMLTree(from_node = node, forget_namespace = True)
>>> # it works as expected
... print tree3
('Envelope', [('Body', [('get', [('IDs', [('ID', '0'), ('ID', '1'), ('ID', '2')
])])])])
>>> 
>>> # forget_namespace with from_string
... strnode = '<myns:foo xmlns:myns="urn:foo"><myns:dummy>Hello, World!</myns:dummy>\
... </myns:foo>'
>>> tree4 = XMLTree(from_string = strnode, forget_namespace = True)
>>> # it also works
... print tree4
('foo', [('dummy', 'Hello, World!')])
\end{verbatim}
    \end{example}
    Some examples of creating an XMLTree with the rewrite feature are shown below.
    \begin{example}
      \caption{XMLTree - rewrite}\label{xtrew}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.xmltree import XMLTree
>>> 
>>> # Create XMLTree objects to demonstrate
... # the use of rewrite feature
>>> strnode = '<myns:foo xmlns:myns="urn:foo"><myns:dummy>Hello, World!</myns:dummy>\
... </myns:foo>'
>>> tree = XMLTree(from_string = strnode,\
...                rewrite = {'myns:foo':'myns:boo'},\
...                forget_namespace = False)
>>> 
>>> # 'myns:foo' is turned into 'myns:boo'
... print tree
('myns:boo', [('myns:dummy', 'Hello, World!')])
>>> 
>>> # combine with forget_namespace
... tree2 = XMLTree(from_string = strnode,\
...                 rewrite = {'foo':'boo'},\
...                 forget_namespace = True)
>>> # 'forget_namespace' is True so namespace prefix 'myns' is being stripped off
... # ('myns:foo' -> 'foo'; 'myns:dummy' -> 'dummy')
... # then 'foo' is turned into 'boo' 
... print tree2
('boo', [('dummy', 'Hello, World!')])
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$add_to_node(self, node, path = None)$ \linebreak
      Adds a tree structure to an XMLNode. Structure is added as a child of the target XMLNode.
        \begin{itemize}
          \item{\verb$node$ is the target XMLNode}
          \item{\verb$path$ selects the part of the XMLTree to be added}
        \end{itemize}
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - add\_to\_node}\label{xtadd2n}
\begin{verbatim}
>>> from arc import XMLNode
>>> from arcom.xmltree import XMLTree
>>> 
>>> node = XMLNode('<node/>')
>>> 
>>> # create XMLTree
... tree = XMLTree(from_string = '<root><dummy/><hello>World</hello></root>')
>>> 
>>> # from 'tree' add 'dummy' to 'node'
... tree.add_to_node(node, '/root/dummy')
>>> 
>>> # show result
... print node.GetXML(True)
<node>
  <dummy></dummy>
</node>
>>> 
>>> # from 'tree' add 'hello' to 'dummy' in 'node'
... dummy = node.Get('dummy')
>>> tree.add_to_node(dummy, '/root/hello')
>>> 
>>> # show result
... print node.GetXML(True)
<node>
  <dummy><hello>World</hello></dummy>
</node>
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$pretty_xml(self, indent = ' ', path = None, prefix = '')$ \linebreak
      Returns a nicely formatted XML representation of the structure.
        \begin{itemize}
          \item{\verb$indent$ is a string that child nodes are indented with}
          \item{\verb$path$ selects the part of the XMLTree to be shown. Full structure is shown when \verb$path$ is None.}
          \item{\verb$prefix$ is a string that is put to the beginning of each line}
        \end{itemize}
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - pretty\_xml}\label{xtpretty}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create XMLTree
... tree = XMLTree(from_string = '<root><Hello>World</Hello></root>')
>>> 
>>> # select entire tree
... # indent children with space character
... # apply prefix for every new line
... print tree.pretty_xml(indent=' ', path='', prefix='# ')
# <root>
#  <Hello>World</Hello>
# </root>
>>> 
>>> # note that it is possible to select multiple parts
... # create an example XMLTree to demonstrate that
... tree2 = XMLTree(from_string = '<root><sub><Hello>World</Hello></sub>\
... <sub><dummy>foo</dummy></sub></root>')
>>> 
>>> # show '/root/sub'
... print tree2.pretty_xml(indent=' ', path='/root/sub', prefix='# ')
# <sub>
#  <Hello>World</Hello>
# </sub>
# <sub>
#  <dummy>foo</dummy>
# </sub>
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$__str__(self)$ \linebreak
      Returns string representation of the structure. Each node is represented by a dictionary where the name
      of the node serves as the key and the value for this key is a list of child nodes of that node.
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - \_\_str\_\_}\label{xtstr}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create XMLTree
... tree = XMLTree(from_string = '<root><dummy>foo</dummy></root>')
>>> 
>>> # show tree
... # print will result in __str__ being called
... # when trying to convert XMLTree to string
... print tree
('root', [('dummy', 'foo')])
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$get(self, path = None)$ \linebreak
      Returns the parts of the XMLTree that match \verb$path$.
      If \verb$path$ is not given, it defaults to the root node.
      This function always returns a list.
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - get}\label{xtget}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create tree structure
... tree = \
... ('root',\
...   [('trunk',\
...     [('leaf','1'),\
...      ('leaf','2')]\
...   )]\
... )
>>> 
>>> # create XMLTree from tree structure
... xt = XMLTree(from_tree = tree)
>>> 
>>> # if called with path being None
... # get() selects the root node
... xt.get()
[('root', [('trunk', [('leaf', '1'), ('leaf', '2')])])]
>>> 
>>> # Path is just a plain path.
... # Empty tag name matches everything, so - in this particular case -
... # all expressions below will produce the same result.
... xt.get('/root/trunk/leaf')
[('leaf', '1'), ('leaf', '2')]
>>> xt.get('//trunk/leaf')
[('leaf', '1'), ('leaf', '2')]
>>> xt.get('/root//leaf')
[('leaf', '1'), ('leaf', '2')]
>>> xt.get('/root/trunk/')
[('leaf', '1'), ('leaf', '2')]
>>> xt.get('///leaf')
[('leaf', '1'), ('leaf', '2')]
>>> xt.get('///')
[('leaf', '1'), ('leaf', '2')]
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$get_trees(self, path = None)$ \linebreak
      Returns XMLTree object for each subtree that match \verb$path$.
      This function always returns a list.
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - get\_trees}\label{xtgettrees}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create tree structure
... tree = \
... ('trunk',\
...   [( 'branch',\
...      [('leaf','1'),\
...       ('leaf','2')]\
...    ),\
...    ( 'branch',\
...      [('leaf','3')]\
...    )]\
... )
>>> 
>>> # create XMLTree from tree structure
... xt = XMLTree(from_tree = tree)
>>> 
>>> # get branches ('/trunk/branch') in a list
... blist = xt.get_trees('/trunk/branch')
>>> 
>>> # show branches in blist
... for b in blist:
...     print b
... 
('branch', [('leaf', '1'), ('leaf', '2')])
('branch', [('leaf', '3')])
>>> # get leaves ('/trunk/branch/leaf') in a list
... llist = xt.get_trees('/trunk/branch/leaf')
>>> 
>>> # show leaves in llist
... for l in llist:
...     print l
... 
('leaf', '1')
('leaf', '2')
('leaf', '3')
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$get_value(self, path = None, *args)$ \linebreak
      Returns the value of the selected part. This means that if node N is
      the first node matched by \verb$path$ then the value of N (i.e. the list of
      its child nodes or its content if it has no child) is returned. If there is no match, 
      and a default value is given, it will be returned instead.
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - get\_value}\label{xtgetval}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create tree structure
... ts = ('branch',[('leaf','1')])
>>> 
>>> # create XMLTree from tree structure
... tree = XMLTree(from_tree = ts)
>>> 
>>> # get_value for '/branch/leaf'
... # will return the value from 'leaf'
... tree.get_value('/branch/leaf')
'1'
>>> 
>>> # get_value for '/branch'
... # will return the value from 'branch'
... # ie the list of its children
... tree.get_value('/branch')
[('leaf', '1')]
>>> 
>>> # get_value for a path that does not match
... # provide a default value of 'N/A'
... tree.get_value('/dummy','N/A')
'N/A'
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$add_tree(self, tree, path = None)$ \linebreak
      Adds a new subtree to a path. This will actually add \verb$tree$ to the first node that matches \verb$path$.
      (Root node of \verb$tree$ will be a new child of the node matched by \verb$path$.)
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - add\_tree}\label{xtaddtree}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create tree structure
... ts = \
... ('root',\
...   [ ('branch',\
...       [ ('leaf','1'),\
...         ('leaf','2')\
...       ]\
...     ),\
...     ('branch',\
...       [ ('leaf','3'),\
...         ('leaf','4')\
...       ]\
...     )\
...   ]\
... )
>>> 
>>> # create XMLTree from tree structure
... tree = XMLTree(from_tree = ts)
>>> 
>>> # add a new leaf ('leaf','5') to '/root/branch'
... # the new leaf will be added to the first node 
... # matched by path ('/root/branch') 
... tree.add_tree(('leaf','5'), '/root/branch')
>>> 
>>> # show tree
... print tree
('root', [('branch', [('leaf', '1'), ('leaf', '2'), ('leaf', '5')]), ('branch', 
[('leaf', '3'), ('leaf', '4')])])
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$get_values(self, path = None)$ \linebreak
      Returns the value (i.e. the list of child nodes or the content if one has no child) 
      for all matched nodes selected by \verb$path$. This function always returns a list.
      Note that unlike \verb$get_value$ this one takes all matched nodes into account.
      Furthermore this one does not accept a default value to return.
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - get\_values}\label{xtgetvals}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create tree structure
... ts = \
... ('root',\
...   [ ('branch',\
...       [ ('leaf','1'),\
...         ('leaf','2')\
...       ]\
...     ),\
...     ('branch',\
...       [ ('leaf','3'),\
...         ('leaf','4')\
...       ]\
...     )\
...   ]\
... )
>>> 
>>> # create XMLTree from tree structure
... tree = XMLTree(from_tree = ts)
>>> 
>>> # get values of branches ('/root/branch') in a list
... # this will return the lists of children of the branches
... tree.get_values('/root/branch')
[[('leaf', '1'), ('leaf', '2')], [('leaf', '3'), ('leaf', '4')]]
>>> 
>>> # get values of leaves ('/root/branch/leaf') in a list
... # this will return the values of leaves
... tree.get_values('/root/branch/leaf')
['1', '2', '3', '4']
>>> 
>>> # try a path that does not match
... # this will return an empty list
... tree.get_values('/dummy')
[]
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$get_dict(self, path = None, keys = {})$ \linebreak
      This method is designed to restore a set of \verb$(key, value)$ pairs stored 
      in an XML structure. It returns a dictionary from the first node matched by \verb$path$.
      \verb$keys$ is a dictionary which filters and translates the keys
      (e.g. if \verb$keys$ is {'my:node':'node'}, it will only return
      'my:node', and will call it 'node'). The value for this key will be 
      the text content of that node or the list of its child nodes.
      (Note though that \verb$get_dict$ was designed with simple text content in mind.)
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - get\_dict}\label{xtgetdict}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create tree structure
... ts =\
... ('person',\
...   [ ('name', 'Alice'),\
...     ('id', '11'),\
...     ('job', 'Librarian')\
...   ]
... )
>>> 
>>> # create XMLTree from tree structure
... tree = XMLTree(from_tree = ts)
>>> 
>>> # get dictionary from '/person'
... # only name and job is to be returned
... tree.get_dict('/person',{'name':'name','job':'job'})
{'job': 'Librarian', 'name': 'Alice'}
\end{verbatim}
    \end{example}

    \begin{itemize}
      \item{ \verb$get_dicts(self, path = None, keys = {})$ \linebreak
      This method is designed to restore multiple sets of \verb$(key, value)$ pairs stored 
      in an XML structure. It returns a list of dictionaries from nodes matched by \verb$path$.
      \verb$keys$ is a dictionary which filters and translates the keys
      (e.g. if \verb$keys$ is {'my:node':'node'}, it will only return
      'my:node', and will call it 'node'). The value for this key will be 
      the text content of that node or the list of its child nodes.
      (Note though that \verb$get_dict$ was designed with simple text content in mind.)
      }
    \end{itemize}

    \begin{example}
      \caption{XMLTree - get\_dicts}\label{xtgetdicts}
\begin{verbatim}
>>> from arcom.xmltree import XMLTree
>>> 
>>> # create tree structure
... ts =\
... ('persons',\
...   [ ('person',\
...       [ ('name', 'Alice'),\
...         ('id', '11'),\
...         ('job', 'Librarian')\
...       ]\
...     ),\
...     ('person',\
...       [ ('name', 'Bob'),\
...         ('id', '12'),\
...         ('job', 'Bartender')\
...       ]\
...     )\
...   ]\
... )
>>> 
>>> # create XMLTree from tree structure
... tree = XMLTree(from_tree = ts)
>>> 
>>> # get dictionaries for '/persons/person'
... tree.get_dicts('/persons/person')
[{'job': 'Librarian', 'name': 'Alice', 'id': '11'}, 
{'job': 'Bartender', 'name': 'Bob', 'id': '12'}]
\end{verbatim}
    \end{example}

  }
\end{itemize}

\end{flushleft}

\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}
\appendix
\section{Helper classes for LogStream function}
\label{LogStreamHelperFullCode}
\begin{verbatim}
%inline %{ 
class CPyOutbuf : public std::streambuf 
{ 
public: 
     CPyOutbuf(PyObject* obj) { 
         m_PyObj = obj; 
         Py_INCREF(m_PyObj); 
     } 
     ~CPyOutbuf() { 
         Py_DECREF(m_PyObj); 
     } 
protected: 
     int_type overflow(int_type c) { 
         // Call to PyGILState_Ensure ensures there is Python 
         // thread state created/assigned.
         PyGILState_STATE gstate = PyGILState_Ensure();
         PyObject_CallMethod(m_PyObj, (char*) "write", (char*) "c", c); 
         PyGILState_Release(gstate);
         return c; 
     } 
     std::streamsize xsputn(const char* s, std::streamsize count) { 
         // Call to PyGILState_Ensure ensures there is Python 
         // thread state created/assigned.
         PyGILState_STATE gstate = PyGILState_Ensure();
         PyObject_CallMethod(m_PyObj, (char*) "write", (char*) "s#", s, int(count)); 
         PyGILState_Release(gstate);
         return count; 
     } 
     PyObject* m_PyObj; 
};

class CPyOstream : public std::ostream 
{ 
public: 
     CPyOstream(PyObject* obj) : m_Buf(obj), std::ostream(&m_Buf) {} 
private: 
     CPyOutbuf m_Buf; 
};

%}
\end{verbatim}

\section{Dummy SecAttr made for use with make\_decision}
\label{DummySecAttrCode}
\begin{verbatim}
import arc
from arc import SecAttr
import sys

class DummySecAttr (SecAttr):
  "Minimalistic implementation of SecAttr."
  def __init__(self, identity = 'Anonymous'):
    SecAttr.__init__(self)
    self.identity = identity
  def Export(self, format = SecAttr.ARCAuth):
    if format != SecAttr.ARCAuth:
      return None
    else:
      ns = arc.NS({'ra':'http://www.nordugrid.org/schemas/request-arc'})
      res = arc.XMLNode(ns, 'ra:Request')
      ritem = res.NewChild('ra:RequestItem')
      sub = ritem.NewChild('ra:Subject')
      subattr = sub.NewChild('ra:SubjectAttribute')
      subattr.Set(self.identity)
      subattr.NewAttribute('Type').Set('string')
      subattr.NewAttribute('AttributeId').Set(\
      'http://www.nordugrid.org/schemas/policy-arc/types/tls/identity'\
      )
      # request read for temporary file
      rsrc = ritem.NewChild('ra:Resource')
      rsrc.Set('file:///home/test')
      rsrc.NewAttribute('Type').Set('string')
      rsrc.NewAttribute('AttributeId').Set('urn:arc:resource:file')
      ac = ritem.NewChild('ra:Action')
      ac.Set('read')
      ac.NewAttribute('Type').Set('string')
      ac.NewAttribute('AttributeId').Set(\
      'http://www.nordugrid.org/schemas/policy-arc/types/storage/action'\
      )
      return res
\end{verbatim}

\section{Dummy and DummyService - an example service based on arcom.service.Service}
\label{DummyServiceCode}
\begin{verbatim}
import arc

class Dummy:
  """ Dummy class that provides functionality """
  def foo(self, foomsg = ''):
      return ''.join(['Foo message is: ',str(foomsg)])

from arcom.service import Service

class DummyService(Service):
  """ DummyService class based on Service class in arcom.service 
      Functionality of Dummy is accessible through this service."""
  def __init__(self, cfg):
    """ Constructor of the DummyService

        DummyService(cfg)

        'cfg' is an XMLNode which contains the config of this service."""

    # set service name
    self.service_name = 'Dummy'

    # set name of provided method
    self.request_names = ['foo']

    # call the Service's constructor
    Service.__init__(self,\
                     [{'request_names' : self.request_names,\
                       'namespace_prefix': 'dmmy',\
                       'namespace_uri': 'urn:foo'}],\
                     cfg)

    self.dummy = Dummy()

  def foo(self, inpayload):
    # get request
    req = inpayload.Child()
    fooreq = str(req)

    # get the job done    
    foomsg = self.dummy.foo(fooreq)

    # create response
    out = self._new_soap_payload()
    response_node = out.NewChild('dmmy:fooResponse')    
    response_node.Set(foomsg)

    return out
\end{verbatim}

\section{arcom.threadpool test - threadpooltest.py}
\label{ThreadPoolTestCode}
\begin{verbatim}
#!/usr/bin/python
# arcom.threadpool test

import threading, traceback

from time import sleep

from arcom.threadpool import ThreadPool, ThreadPoolThread, ReadWriteLock

COUNTER = 0

# waitTask: just sleep for a number of seconds
def waitTask(data):
  num = data[0]
  time = data[1]
  print "(%d): WaitTask starting" % num
  print "WaitTask sleeping for %d seconds" % time
  sleep(time)
  return "Waiter (%d)" % num

locker = ReadWriteLock()

# addTask: increment COUNTER with inc
def addTask(data):
  num = data[0]
  inc = data[1]
  global COUNTER
  nadds = 3
  for i in range(nadds):
    # re-acquiring lock for every add
    # don't do this at home...
    locker.acquire_write()
    COUNTER += inc
    print "(%d): Added %d to counter, counter is now %d"%(num,inc,COUNTER)
    locker.release_write()
    # sleep a bit to give other threads enough time to acquire the lock 
    sleep(1)
  return "addTask", inc

# Both tasks use the same callback
def taskCallback(data):
  print "Callback called for", data

# Create a pool with three worker threads
pool = ThreadPool(3)

print "\nExample1 - waitTask"

# Insert tasks into the queue and let them run
pool.queueTask(waitTask, (1,4), taskCallback)
pool.queueTask(waitTask, (2,1), taskCallback)
pool.queueTask(waitTask, (3,6), taskCallback)
pool.queueTask(waitTask, (4,2), taskCallback)

pool.joinAll()

pool = ThreadPool(4)

print "\nExample2 - addTask"

pool.queueTask(addTask, (5,10), taskCallback)
pool.queueTask(addTask, (6,10), taskCallback)
pool.queueTask(addTask, (7,10), taskCallback)
pool.queueTask(addTask, (8,10), taskCallback)

print "COUNTER before join: ", COUNTER
# When all tasks are finished, allow the threads to terminate
pool.joinAll()
print "COUNTER after join: ", COUNTER
\end{verbatim}

\input{app/ex-2-1-1.tex}
\input{app/ex-2-1-2.tex}
\input{app/ex-2-1-3.tex}
\input{app/ex-2-1-4.tex}
\input{app/ex-2-1-5.tex}
\input{app/ex-3-1.tex}
\input{app/ex-3-2.tex}
\input{app/ex-3-3.tex}
\input{app/ex-3-4.tex}
\input{app/ex-3-5.tex}
\input{app/ex-3-6.tex}
\input{app/ex-3-7.tex}

\end{document}

### 2009-09-18 doc/manuals/python/app/ex-2-1-2.tex
Documentation about the HED's python API and the arcom utility package - written by Tamás Kazinczy

\section{Examples of \ref{icommon}}
\subsection{String operator - simple node}
(Example \ref{strex1})
\label{cstrex1}
\begin{verbatim}
import arc
#create an XMLNode
mynode = arc.XMLNode(arc.NS({'me':'http://example.com/myExample'}),'me:myNode')
#set text content
mynode.Set('Hello, World!')
#representation of the node
mynode.GetXML()
#String operator in action
str(mynode)
\end{verbatim}

\subsection{String operator - complex node}
(Example \ref{strex2})
\label{cstrex2}
\begin{verbatim}
# String operator of XMLNode works with the text content.
# It does not care about child nodes or text contents of those nodes.

# build a tree and set text content in nodes
#       r
#     / | \
#    /  |  \
#   /   |   \
# c01  c02  c03
#      /\    |
#     /  \  c06  
#   c04  c05

# create the root node
r = arc.XMLNode(arc.NS(),'r')
# set text content
r.Set('R')
# create rest of the tree
c01 = r.NewChild('c01')
c01.Set('C-01')
c02 = r.NewChild('c02')
c02.Set('C-02')
c03 = r.NewChild('c03')
c03.Set('C-03')
c04 = c02.NewChild('c04')
c04.Set('C-04')
c05 = c02.NewChild('c05')
c05.Set('C-05')
c06 = c03.NewChild('c06')
c06.Set('C-06')
# String operator does not care about child nodes
str(c03)
str(c02)
str(r)
# XML representation of a complex node
c02.GetXML()
\end{verbatim}

\subsection{Occurence of out\_xml\_str as an output value}
(Example \ref{outxml})
\label{coutxml}
\begin{verbatim}
import arc
#create node
n = arc.XMLNode(arc.NS({'me':'myNS'}),'me:myNode')
#set content
n.Set('Hello, World!')
#get document
#GetDoc according to API:
#void GetDoc(std::string &out_xml_str, bool user_friendly=false) const
#out_xml_str is turned into an output value and user_friendly is an optional 
#parameter so there are no necessary parameters this time
#result will hold the output value
result = n.GetDoc()
#show result
result
\end{verbatim}

\subsection{Using LogStream to add new destination to the root logger}
(Example \ref{logstream})
\label{clogstream}
\begin{verbatim}
import arc
import sys
#get root logger
root_logger = arc.Logger_getRootLogger()
#create a LogStream; sys.stdout would be OK
stream = arc.LogStream(sys.stdout)
#add destination to root logger
root_logger.addDestination(stream)
#log a message
#result immediately appears on sys.stdout
root_logger.msg(arc.INFO, 'Hello, World!')
\end{verbatim}

### 2009-09-18 doc/manuals/python/app/ex-2-1-4.tex
Documentation about the HED's python API and the arcom utility package - written by Tamás Kazinczy

\section{Examples of \ref{iclient}}
\subsection{ClientSOAP - process}
(Example \ref{procex})
\label{cprocex}
\begin{verbatim}
import arc
#create default config
cfg = arc.MCCConfig()
#create URL
url = arc.URL('http://localhost:50000/Echo')
#create payload
payload = arc.PayloadSOAP(arc.NS({'echo':'urn:echo'}))
#create payload content
#and set echo message
payload.NewChild('echo:echo').NewChild('echo:say').Set('Hello, World!')
#create client
client = arc.ClientSOAP(cfg,url)
#let the client call the service
response, status = client.process(payload)
#see response
print response.GetXML()
#check status
status.getKind() == arc.STATUS_OK
\end{verbatim}

### 2009-09-18 doc/manuals/python/app/ex-2-1-5.tex
Documentation about the HED's python API and the arcom utility package - written by Tamás Kazinczy

\section{Examples of \ref{idata}}
\subsection{DataMover - Transfer}
(Example \ref{dmex})
\label{cdmex}
\begin{verbatim}
import arc
# create DataMover
mover = arc.DataMover()
# import datapoint_from_url from arcom package
# DataPoints are required to use DataMover
from arcom import datapoint_from_url
# create source DataPoint
src = datapoint_from_url('file://./file1')
# create destination DataPoint
dst = datapoint_from_url('file://./file2')
# do the transfer
status, failure = mover.Transfer(src, dst, arc.FileCache(), arc.URLMap())
# check status
str(status)
'Operation completed successfully'
# check for failure message
failure
\end{verbatim}

### 2009-09-18 doc/manuals/python/app/ex-3-1.tex
Documentation about the HED's python API and the arcom utility package - written by Tamás Kazinczy

\section{Examples of \ref{ainit}}
\subsection{Importing the Logger class from module ''logger'' of ''arcom'' package}
(Example \ref{loglog})
\label{cloglog}
\begin{verbatim}
import arcom
# From the logger module within the arcom package
# import the Logger class
Logger = arcom.import_class_from_string('arcom.logger.Logger')
# Now the class could be access through Logger
Logger
\end{verbatim}

\subsection{Getting attributes from an XMLNode}
(Example \ref{getattr})
\label{cgetattr}
\begin{verbatim}
import arc
import arcom
# Create XMLNode
n = arc.XMLNode(arc.NS({'me':'http://example.com/myExample'}),'me:myExample')
# Add attributes to node and set their values
n.NewAttribute('foo').Set('Hello')
n.NewAttribute('moo').Set('World')
# Show XML
n.GetXML()
# get attributes of this XMLNode
a = arcom.get_attributes(n)
# result is a dictionary
type(a)
# show dictionary
a
# extracting an attribute value
a['moo']
\end{verbatim}

\subsection{Getting child nodes}
(Example \ref{getchld})
\label{cgetchld}
\begin{verbatim}
import arc
import arcom
# Create root node
a = arc.XMLNode(arc.NS(),'a')
# Create child nodes for 'a'
b = a.NewChild('b')
c = a.NewChild('c')
d = a.NewChild('d')
# Create child nodes for 'c'
e = c.NewChild('e')
f = c.NewChild('f')
# Create child node for 'd'
g = d.NewChild('g')
# show XML
a.GetXML()
# get child nodes for 'a'
tmp = arcom.get_child_nodes(a)
# result is a list
type(tmp)
# 'a' has 3 children
len(tmp)
# show name for each
', '.join(x.Name() for x in tmp)
# 'b' has no children
tmp = arcom.get_child_nodes(b)
len(tmp)
# 'c' has 2 children: 'e' and 'f'
tmp = arcom.get_child_nodes(c)
len(tmp)
', '.join(x.Name() for x in tmp)
# 'd' has 1 child: 'g'
tmp = arcom.get_child_nodes(d)
len(tmp)
', '.join(x.Name() for x in tmp)
# nodes 'e', 'f' and 'g' have no children
tmp = arcom.get_child_nodes(e)
len(tmp)
tmp = arcom.get_child_nodes(f)
len(tmp)
tmp = arcom.get_child_nodes(g)
len(tmp)
\end{verbatim}

\subsection{Get values of specified children}
(Example \ref{getchdval})
\label{cgetchdval}
\begin{verbatim}
import arc
import arcom
# Create XMLNode
n = arc.XMLNode(arc.NS(),'node')
# Create 3 child nodes (two of which get the same name)
x = n.NewChild('same')
y = n.NewChild('same')
z = n.NewChild('different')
# Set values for nodes
x.Set('firstEQ')
y.Set('secondEQ')
z.Set('DIFF')
# Show XML
n.GetXML()
# Get child values where name is 'same'
tmp = arcom.get_child_values_by_name(n,'same')
# Result is a list
type(tmp)
# Show result
tmp
# Get child values where name is 'different'
tmp = arcom.get_child_values_by_name(n,'different')
# Show result
tmp
\end{verbatim}

\subsection{Creating DataPoint from URL}
(Example \ref{url2dp})
\label{curl2dp}
\begin{verbatim}
import arc
import arcom
tmpList = []
status = ''
# create DataPoint from a local directory
dp = arcom.datapoint_from_url('file:///usr/local/share/arc')
# list files
(files, stat) = dp.ListFiles()
# if it is not empty
if files:
    status = 'found'
    # for all the entries, get type and name
    for f in files:
        if (f.GetType() == arc.FileInfo.file_type_file):
            type = 'file'
        elif (f.GetType() == arc.FileInfo.file_type_dir):
            type = 'dir'
        else:
            type = 'unknown'
        # get results together in a list
        tmpList.append(f.GetName() + ' (' + type + ')\n')
else:
    status = 'Could not access data. Reason: %s' % str(stat)

# see result
str(stat)
# show list
tmpList
\end{verbatim}

\subsection{Parsing a URL}
(Example \ref{parseurl})
\label{cparseurl}
\begin{verbatim}
import arcom
proto, host, port, path = arcom.parse_url('boo://no.one.here:123/foo')
proto
host
port
path
\end{verbatim}

### 2009-09-18 doc/manuals/python/app/ex-3-2.tex
Documentation about the HED's python API and the arcom utility package - written by Tamás Kazinczy

\section{Examples of \ref{aclient}}
\subsection{Creating a client and calling the echo service (XMLTree)}
(Example \ref{clientex1})
\label{cclientex1}
\begin{verbatim}
import arc
import arcom
# Import the Client class
Client = arcom.import_class_from_string('arcom.client.Client')
# Import the XMLTree class
XMLTree = arcom.import_class_from_string('arcom.xmltree.XMLTree')
# Create namespace - it will be used for the message sent
ns = arc.NS({'echo':'urn:echo'})
# Create client
c = Client('http://your.server.example.com:50000/Echo',ns,print_xml=True)
# Create message
msg = XMLTree(from_tree = ('echo:echo',[('echo:say', 'Hello, World!')]))
# Let the client do what it is meant for
# Note that we created the client with print_xml=True
# so both request and response will be displayed in an easy-to-read form
c.call(msg)
\end{verbatim}

\subsection{Creating a client and calling the echo service (SOAP)}
(Example \ref{clientex2})
\label{cclientex2}
\begin{verbatim}
import arc
import arcom
# Import the Client class
Client = arcom.import_class_from_string('arcom.client.Client')
# Create namespace - it will be used for the message sent
ns = arc.NS({'echo':'urn:echo'})
# Create client
c = Client('http://arctest.ki.iif.hu:50000/Echo',ns,print_xml=True)
# Create SOAP Payload
pl = arc.PayloadSOAP(ns)
# Create message structure and set content
pl.NewChild('echo:echo',ns).NewChild('echo:say',ns).Set('Hello, World!')
# Let client do the call; see response
c.call_raw(pl)
\end{verbatim}

### 2009-09-18 doc/manuals/python/app/ex-3-5.tex
Documentation about the HED's python API and the arcom utility package - written by Tamás Kazinczy

\section{Examples of \ref{aservice}}
\subsection{Using DummyService - an example service based on arcom.service.Service}
(Example \ref{srvex})
\label{csrvex}
\begin{verbatim}
# Reminder:
# import arc
#
# class Dummy:
#   def foo(self, foomsg = ''):
#     return ''.join(['Foo message is: ',str(foomsg)])
# 
# from arcom.service import Service
# 
# class DummyService(Service):
#   """ DummyService class based on Service class in arcom.service """
#   ...

# Create instance of DummyService
ds = DummyService(None)
# Create namespace - it will be used for the message sent
ns = arc.NS({'dmmy':'urn:foo'})
# Create SOAP Envelope
se = arc.SOAPEnvelope(ns)
# Create content
se.NewChild('dmmy:foo').Set('Foo!')
# Create SOAP Message
msg = arc.SOAPMessage()
# Payload
pls = arc.PayloadSOAP(se)
# Set Message Payload
msg.Payload(pls)
# create outgoing message that will hold response
outse = arc.SOAPEnvelope(ns)
outmsg = arc.SOAPMessage()
outpls = arc.PayloadSOAP(outse)
outmsg.Payload(pls)
# call service
ds.process(msg, outmsg)
# show result
print outmsg.Payload().GetXML()
\end{verbatim}

\subsection{parse\_node}
(Example \ref{pnex})
\label{cpnex}
\begin{verbatim}
from arc import XMLNode
from arcom.service import parse_node

# create xml node
xml = XMLNode('''
<myList>
  <myElement>
    <myID>0</myID>
    <LN>/</LN>
  </myElement>
  <myElement>
    <myID>1</myID>
    <LN>/testfile</LN>
  </myElement>
</myList>''')

# default: single = False and string = True
parse_node(xml, ['myID','LN'])
# single = True
parse_node(xml, ['myID','LN'], single = True)
parse_node(xml, ['myID','LN'], True)
# string = False
parse_node(xml, ['myID','LN'], string = False)
\end{verbatim}

\subsection{parse\_to\_dict}
(Example \ref{p2dex})
\label{cp2dex}
\begin{verbatim}
from arc import XMLNode
from arcom.service import parse_to_dict

# create xml node
xml = XMLNode('''
<myList>
  <myElement>
    <myID>123</myID>
    <refID>abc</refID>
    <state>alive</state>
    <size>123456</size>
  </myElement>
  <myElement>
    <myID>456</myID>
    <refID>fed</refID>
    <state>alive</state>
    <size>987</size>
  </myElement>
</myList>''')

# example 1 - 'myID' as key
# 'state' and 'size' goes to the inner dictionary
parse_to_dict(xml, ['myID','state','size'])

# example 2 - 'myID' as key
# 'myID' also goes to the inner dictionary
parse_to_dict(xml, ['myID','myID','state','size'])
\end{verbatim}

\subsection{create\_response}
(Example \ref{crespex})
\label{ccrespex}
\begin{verbatim}
import arc
from arcom.service import create_response

method_name = 'dummy'
# create tag_names; 'myID' will be the key
tag_names = ['myID', 'state', 'size']
# create elements
# elements of the lists in this dictionary are values and will be tagged 
# according to names found in tag_names except for the first element
# e.g.: 'alive' will be tagged with 'state'
elements = {'123': ['alive', '123456'], '456': ['alive', '987']}
# create empty payload
payload = arc.PayloadSOAP(arc.NS())

# create response
response = create_response(method_name, tag_names, elements, payload)

# show it
print response.GetXML(True)
\end{verbatim}

\subsection{node\_to\_data - 1}
(Example \ref{n2dex1})
\label{cn2dex1}
\begin{verbatim}
from arc import XMLNode
from arcom.service import node_to_data

# create node
xml = XMLNode('''
<myElement>
  <myID>0</myID>
  <refID>abc</refID>
  <section>states</section>
  <property>spares</property>
  <value>2</value>
  <dummy>foo</dummy>
  <changeType>set</changeType>
</myElement>
''')

# create names
# 'myID' will be the key
# all the other names will be in a list in the specified order
# 'dummy' will be left out
names = ['myID', 'refID', 'changeType', 'section', 'property', 'value']

# call node_to_data
node_to_data(xml,names)
\end{verbatim}

\subsection{node\_to\_data - 2}
(Example \ref{n2dex2})
\label{cn2dex2}
\begin{verbatim}
from arc import XMLNode
from arcom.service import node_to_data

# create node
xml = XMLNode('''
<getRequest>
  <GUID>11</GUID>
  <myID>99</myID>
</getRequest>
''')

# create names
# 'myID' will be the key
names = ['myID', 'GUID']

# call node_to_data
# GUID is the single value so do not put it in a list
node_to_data(xml,names, True)
\end{verbatim}

\subsection{get\_data\_node}
(Example \ref{gdnex})
\label{cgdnex}
\begin{verbatim}
from arc import XMLNode
from arcom.service import get_data_node

# create node
xml = XMLNode('''
<Envelope>
  <Body>
    <method>
      <data>Value</data>
    </method>
  </Body>
</Envelope>
''')

# call get_data_node
dn = get_data_node(xml)

# show result
dn.GetXML()
\end{verbatim}

### 2009-09-18 doc/manuals/python/app/ex-3-7.tex
Documentation about the HED's python API and the arcom utility package - written by Tamás Kazinczy

\section{Examples of \ref{axtree}}
\subsection{Creating an XMLTree}
(Example \ref{crxt})
\label{ccrxt}
\begin{verbatim}
from arc import XMLNode
from arcom.xmltree import XMLTree

# create XMLTree sources
# xmlstr - will be used to create an XMLTree from string
# xmlnodestr - will be used to create 'node' (an XMLNode)
# node - will be used to create an XMLTree from XMLNode
# treestruct - for creating XMLTree from tree structure
xmlstr = '<echo><say>Hello!</say></echo>'
xmlnodestr = '<node><subnode>Subnode</subnode></node>'
node = XMLNode(xmlnodestr)
treestruct = ('root', [('leaf','Leaf')])

# from_tree has the highest priority
t1 = XMLTree(from_node = node, from_string = xmlstr, from_tree = treestruct) 
t1.get()

# no from_tree supplied; from_node will be used
t2 = XMLTree(from_node = node, from_string = xmlstr)
t2.get()

# from_string only
t3 = XMLTree(from_string = xmlstr)
t3.get()
\end{verbatim}

\subsection{XMLTree - forget namespace}
(Example \ref{xtfn})
\label{cxtfn}
\begin{verbatim}
from arc import XMLNode
from arcom.xmltree import XMLTree

# Create XMLTree objects to demonstrate
# the use of forget namespace feature
# node - is an XMLNode
# tree - is an XMLTree that is created from 'node'
node = XMLNode('<soap-env:Envelope xmlns:hash="urn:hash" \
xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" \
xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" \
xmlns:xsd="http://www.w3.org/2001/XMLSchema" \
xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">\
<soap-env:Body>\
<hash:get>\
<hash:IDs>\
<hash:ID>0</hash:ID>\
<hash:ID>1</hash:ID>\
<hash:ID>2</hash:ID>\
</hash:IDs>\
</hash:get>\
</soap-env:Body>\
</soap-env:Envelope>')

tree = XMLTree(node)
# show tree structure
print tree

# forget_namespace is ignored when from_tree is used
tree2 = XMLTree(from_tree = tree, forget_namespace = True)
print tree2

# forget_namespace with from_node
tree3 = XMLTree(from_node = node, forget_namespace = True)
# it works as expected
print tree3

# forget_namespace with from_string
strnode = '<myns:foo xmlns:myns="urn:foo"><myns:dummy>Hello, World!</myns:dummy>\
</myns:foo>'
tree4 = XMLTree(from_string = strnode, forget_namespace = True)
# it also works
print tree4
\end{verbatim}

\subsection{XMLTree - rewrite}
(Example \ref{xtrew})
\label{cxtrew}
\begin{verbatim}
from arc import XMLNode
from arcom.xmltree import XMLTree

# Create XMLTree objects to demonstrate
# the use of rewrite feature
strnode = '<myns:foo xmlns:myns="urn:foo"><myns:dummy>Hello, World!</myns:dummy>\
</myns:foo>'
tree = XMLTree(from_string = strnode,\
               rewrite = {'myns:foo':'myns:boo'},\
               forget_namespace = False)

# 'myns:foo' is turned into 'myns:boo'
print tree

# combine with forget_namespace
tree2 = XMLTree(from_string = strnode,\
                rewrite = {'foo':'boo'},\
                forget_namespace = True)
# 'forget_namespace' is True so namespace prefix 'myns' is being stripped off
# ('myns:foo' -> 'foo'; 'myns:dummy' -> 'dummy')
# then 'foo' is turned into 'boo' 
print tree2
\end{verbatim}

\subsection{XMLTree - add\_to\_node}
(Example \ref{xtadd2n})
\label{cxtadd2n}
\begin{verbatim}
from arc import XMLNode
from arcom.xmltree import XMLTree

node = XMLNode('<node/>')

# create XMLTree
tree = XMLTree(from_string = '<root><dummy/><hello>World</hello></root>')

# from 'tree' add 'dummy' to 'node'
tree.add_to_node(node, '/root/dummy')

# show result
print node.GetXML(True)

# from 'tree' add 'hello' to 'dummy' in 'node'
dummy = node.Get('dummy')
tree.add_to_node(dummy, '/root/hello')

# show result
print node.GetXML(True)
\end{verbatim}

\subsection{XMLTree - pretty\_xml}
(Example \ref{xtpretty})
\label{cxtpretty}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create XMLTree
tree = XMLTree(from_string = '<root><Hello>World</Hello></root>')

# select entire tree
# indent children with space character
# apply prefix for every new line
print tree.pretty_xml(indent=' ', path='', prefix='# ')

# note that it is possible to select multiple parts
# create an example XMLTree to demonstrate that
tree2 = XMLTree(from_string = '<root><sub><Hello>World</Hello></sub>\
<sub><dummy>foo</dummy></sub></root>')

# show '/root/sub'
print tree2.pretty_xml(indent=' ', path='/root/sub', prefix='# ')
\end{verbatim}

\subsection{XMLTree - \_\_str\_\_}
(Example \ref{xtstr})
\label{cxtstr}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create XMLTree
tree = XMLTree(from_string = '<root><dummy>foo</dummy></root>')

# show tree
# print will result in __str__ being called
# when trying to convert XMLTree to string
print tree
\end{verbatim}

\subsection{XMLTree - get}
(Example \ref{xtget})
\label{cxtget}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
tree = \
('root',\
  [('trunk',\
    [('leaf','1'),\
     ('leaf','2')]\
  )]\
)

# create XMLTree from tree structure
xt = XMLTree(from_tree = tree)

# if called with path being None
# get() selects the root node
xt.get()

# Path is just a plain path.
# Empty tag name matches everything, so - in this particular case -
# all expressions below will produce the same result.
xt.get('/root/trunk/leaf')
xt.get('//trunk/leaf')
xt.get('/root//leaf')
xt.get('/root/trunk/')
xt.get('///leaf')
xt.get('///')
\end{verbatim}

\subsection{XMLTree - get\_trees}
(Example \ref{xtgettrees})
\label{cxtgettrees}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
tree = \
('trunk',\
  [( 'branch',\
     [('leaf','1'),\
      ('leaf','2')]\
   ),\
   ( 'branch',\
     [('leaf','3')]\
   )]\
)

# create XMLTree from tree structure
xt = XMLTree(from_tree = tree)

# get branches ('/trunk/branch') in a list
blist = xt.get_trees('/trunk/branch')

# show branches in blist
for b in blist:
    print b

# get leaves ('/trunk/branch/leaf') in a list
llist = xt.get_trees('/trunk/branch/leaf')

# show leaves in llist
for l in llist:
    print l

\end{verbatim}

\subsection{XMLTree - get\_value}
(Example \ref{xtgetval})
\label{cxtgetval}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts = ('branch',[('leaf','1')])

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# get_value for '/branch/leaf'
# will return the value from 'leaf'
tree.get_value('/branch/leaf')

# get_value for '/branch'
# will return the value from 'branch'
# ie the list of its children
tree.get_value('/branch')

# get_value for a path that does not match
# provide a default value of 'N/A'
tree.get_value('/dummy','N/A')
\end{verbatim}

\subsection{XMLTree - add\_tree}
(Example \ref{xtaddtree})
\label{cxtaddtree}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts = \
('root',\
  [ ('branch',\
      [ ('leaf','1'),\
        ('leaf','2')\
      ]\
    ),\
    ('branch',\
      [ ('leaf','3'),\
        ('leaf','4')\
      ]\
    )\
  ]\
)

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# add a new leaf ('leaf','5') to '/root/branch'
# the new leaf will be added to the first node 
# matched by path ('/root/branch') 
tree.add_tree(('leaf','5'), '/root/branch')

# show tree
print tree
\end{verbatim}

\subsection{XMLTree - get\_values}
(Example \ref{xtgetvals})
\label{cxtgetvals}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts = \
('root',\
  [ ('branch',\
      [ ('leaf','1'),\
        ('leaf','2')\
      ]\
    ),\
    ('branch',\
      [ ('leaf','3'),\
        ('leaf','4')\
      ]\
    )\
  ]\
)

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# get values of branches ('/root/branch') in a list
# this will return the lists of children of the branches
tree.get_values('/root/branch')

# get values of leaves ('/root/branch/leaf') in a list
# this will return the values of leaves
tree.get_values('/root/branch/leaf')

# try a path that does not match
# this will return an empty list
tree.get_values('/dummy')
\end{verbatim}

\subsection{XMLTree - get\_dict}
(Example \ref{xtgetdict})
\label{cxtgetdict}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts =\
('person',\
  [ ('name', 'Alice'),\
    ('id', '11'),\
    ('job', 'Librarian')\
  ]
)

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# get dictionary from '/person'
# only name and job is to be returned
tree.get_dict('/person',{'name':'name','job':'job'})
\end{verbatim}

\subsection{XMLTree - get\_dicts}
(Example \ref{xtgetdicts})
\label{cxtgetdicts}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts =\
('persons',\
  [ ('person',\
      [ ('name', 'Alice'),\
        ('id', '11'),\
        ('job', 'Librarian')\
      ]\
    ),\
    ('person',\
      [ ('name', 'Bob'),\
        ('id', '12'),\
        ('job', 'Bartender')\
      ]\
    )\
  ]\
)

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# get dictionaries for '/persons/person'
tree.get_dicts('/persons/person')
\end{verbatim}

### 2009-09-24 doc/tech_doc/a-rex/arex_tech_doc.tex
add SLURM backend (copied from arc0/trunk)
port gm-jobs utility from arc0. I will let the release coordinators decide whether to put on the final release branch

@@
 \item \textit{submit-{*}-job} -- Here {*} stands for the name of the LRMS.
-Currently supported LRMS are PBS/Torque, Condor, LoadLeveler, LSF
+Currently supported LRMS are PBS/Torque, Condor, LoadLeveler, LSF, SLURM,
 and SGE. Also \emph{fork} pseudo-LRMS is supported for testing purposes.
@@
 
-\begin{comment}In addition, there is also an administration utility:
+In addition, there is also an administration utility:
 
@@
 and the number of jobs in each state.\\
-\hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name]\\
+\hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name] {[}-c conf\_file] {[}-d control\_dir]\\
+\hspace*{0.5cm}-h -- print short help,\\
 \hspace*{0.5cm}-l -- print more information about each job,\\
 \hspace*{0.5cm}-u -- pretend utility is run by user with id \emph{uid},\\
-\hspace*{0.5cm}-U -- pretend utility is run by user with name \emph{name}.
-\end{itemize}
-\end{comment}
+\hspace*{0.5cm}-U -- pretend utility is run by user with name \emph{name},\\
+\hspace*{0.5cm}-c -- use specified configuration file,\\
+\hspace*{0.5cm}-d -- read information from specified control dir.
+\end{itemize}
 

### 2009-10-07 doc/tech_doc/infosys/infosys_technical.tex
Minor language fix
Adapting ISIS documentation to the alternated schema

@@
 
-So Service 1 and 2 are both handled by \textit{InfoRegistrar A} since Service 2 and Service 3 by \textit{InfoRegistrar B}.
+So Service 1 and Service 2 are handled by \textit{InfoRegistrar A} and Service 2 and Service 3 by \textit{InfoRegistrar B}.
 In the first step each InfoRegistrar performs information collection from all assigned services sequentially.
@@
   xmlns:xsd="http://www.w3.org/2001/XMLSchema"
-  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2009/isis"
-  xmlns:icfg="http://www.nordugrid.org/schemas/ArcConfig/2009/isis"
-  targetNamespace="http://www.nordugrid.org/schemas/ArcConfig/2009/isis"
+  xmlns="http://www.nordugrid.org/schemas/isis/2009/08"
+  xmlns:icfg="http://www.nordugrid.org/schemas/isis/2009/08"
+  targetNamespace="http://www.nordugrid.org/schemas/isis/2009/08"
   elementFormDefault="qualified" attributeFormDefault="unqualified"
@@
 
-  <xsd:element name="endpoint" type="xsd:anyURI"/>
-
-  <xsd:element name="retry" type="xsd:nonNegativeInteger" minOccurs="0"/>
-
-  <xsd:element name="sparsity" type="xsd:nonNegativeInteger" minOccurs="0"/>
-
-  <xsd:simpleType name="Logger_Type">
-    <xsd:restriction base="xsd:string"/>
-    <xsd:attribute name="level" type="xsd:string" use="optional"/>
+  <xsd:element name="KeyPath" type="xsd:string">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              Optional KeyPath for SSL connection
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="CertificatePath" type="xsd:string">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              Optional CertificatePath for SSL connection
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="ProxyPath" type="xsd:string">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              Optional ProxyPath for SSL connection
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="CACertificatesDir" type="xsd:string">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              Optional CACertificatesDir for SSL connection
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="endpoint" type="xsd:anyURI">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              The URL where the service can be accessed from outside.
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="retry" type="xsd:nonNegativeInteger" default="5">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              Retry specifies how many times communication to ISIS have to fail/timeout to start treating it as unavailable.
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="sparsity" type="xsd:nonNegativeInteger" default="2">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              It determines the number of neighbors as a function of the actual number of member nodes of the network.
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:complexType name="Logger_Type">
+    <xsd:simpleContent>
+      <xsd:extension base="xsd:string">
+        <xsd:attribute name="level" default="INFO">
+          <xsd:simpleType>
+            <xsd:restriction base="xsd:string">
+              <xsd:enumeration value="VERBOSE"/>
+              <xsd:enumeration value="DEBUG"/>
+              <xsd:enumeration value="INFO"/>
+              <xsd:enumeration value="WARNING"/>
+              <xsd:enumeration value="ERROR"/>
+              <xsd:enumeration value="FATAL"/>
+            </xsd:restriction>
           </xsd:simpleType>
-  <xsd:element name="Logger" type="icfg:Logger_Type" minOccurs="0"/>
-
-  <xsd:element name="ETValid" type="xsd:nonNegativeInteger" minOccurs="0"/>
-
-  <xsd:element name="ETRemove" type="xsd:nonNegativeInteger" minOccurs="0"/>
-
-  <xsd:element name="DBPath" type="xsd:string"/>
+        </xsd:attribute>
+      </xsd:extension>
+    </xsd:simpleContent>
+  </xsd:complexType>
+
+  <xsd:element name="Logger" type="icfg:Logger_Type">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              File location where the log message is locally stored.
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="ETValid" type="xsd:nonNegativeInteger" default="43200">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              Soft state database parameter. Time of the valid messages checking period.
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="ETRemove" type="xsd:nonNegativeInteger" default="43200">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              Soft state database parameter. Time of the expired messages checking period.
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
+
+  <xsd:element name="DBPath" type="xsd:string">
+      <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+              File location where the database is locally stored.
+          </xsd:documentation>
+      </xsd:annotation>
+  </xsd:element>
 
@@
     <xsd:sequence>
-      <xsd:element name="URL" type="xsd:anyURI"/>
-    </xsd:sequence>
-  </xsd:complexType>
-  <xsd:element name="InfoProvider" type="icfg:InfoProvider_Type" minOccurs="0" maxOccurs="unbounded"/>
+      <xsd:element name="URL" type="xsd:anyURI">
+          <xsd:annotation>
+              <xsd:documentation xml:lang="en">
+                  The URL of one or more ISIS where the other ISIS's can be queried.
+              </xsd:documentation>
+          </xsd:annotation>
+      </xsd:element>
+    </xsd:sequence>
+  </xsd:complexType>
+
+  <xsd:element name="InfoProvider" type="icfg:InfoProvider_Type"/>
 
@@
   xmlns:xsd="http://www.w3.org/2001/XMLSchema"
-  xmlns:iregc="http://www.nordugrid.org/schemas/InfoRegisterConfig/2008"
-  targetNamespace="http://www.nordugrid.org/schemas/InfoRegisterConfig/2008"
+  xmlns:iregc="http://www.nordugrid.org/schemas/infosys/2009/08"
+  targetNamespace="http://www.nordugrid.org/schemas/infosys/2009/08"
   elementFormDefault="qualified">
 
-    <xsd:complexType name="InfoRegistrar_Type">
-        <!-- This element defines configuration of Information Registration
-             active element. This element is located at top level configuration
-             document. Each instance if such element instatiates one instance
-             of Information \hl{Registrant} active element. -->
-        <xsd:sequence>
-            <!-- URL specifies contact endpoint of a bootstrap Information
-                 Registration service. Further ISIS adresses will be queried
-                 from this service. -->
-            <xsd:element name="URL" type="xsd:string"/> <!-- maxOccurs="unbounded" -->
-            <!-- Optional KeyPath, CertificatePath, ProxyPath and CACertificatesDir are
-                 paths to files storing X509 credentials used for establishing connections. -->
-            <xsd:element name="KeyPath" type="xsd:string" minOccurs="0"/>
-            <xsd:element name="CertificatePath" type="xsd:string" minOccurs="0"/>
-            <xsd:element name="ProxyPath" type="xsd:string" minOccurs="0"/>
-            <xsd:element name="CACertificatesDir" type="xsd:string" minOccurs="0"/>
-            <!-- Retry count. Specifie how many times have an ISIS to timeout to handle it as
-                 unavailable. If there is no value specified the default value: 5 will be used. -->
-            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1"/>
-        </xsd:sequence>
-        <!-- Attribute id may be used to refer to this element -->
-        <xsd:attribute name="id" type="xsd:string" use="optional"/>
-    </xsd:complexType>
-    <xsd:element name="InfoRegistrar" type="iregc:InfoRegistrar_Type"/>
-
     <xsd:complexType name="Registrar_Type">
-        <xsd:sequence>
-            <!-- Missing elements means that the default values will be used. -->
-
-            <!-- Period specifies how often registration to be done in minutes.
-                 The 0 value means single registration. -->
-            <xsd:element name="Period" type="xsd:nonNegativeInteger" minOccurs="0" maxOccurs="1"/>
-            <!-- Element defines the unique id of the service propagated outside. -->
-            <xsd:element name="ServiceID" type="xsd:string" minOccurs="0" maxOccurs="1"/>
-            <!-- This element defines URL of the service as seen from outside. -->
-            <xsd:element name="endpoint" type="xsd:string" minOccurs="0" maxOccurs="1"/>
-            <!-- This element defines the expiration time of the information provided by the service. -->
-            <xsd:element name="expiration" type="xsd:string" minOccurs="0" maxOccurs="1"/>
-        </xsd:sequence>
-        <!-- Attribute id is to define the previously configured InfoRegistrar to be refered -->
-        <xsd:attribute name="id" type="xsd:string" use="required"/>
+        <xsd:annotation>
+            <xsd:documentation xml:lang="en">
+             This element defines configuration of Information Registration.
+             active element.
+            </xsd:documentation>
+        </xsd:annotation>
+        <xsd:sequence>
+
+            <xsd:element name="URL" type="xsd:string" minOccurs="1" maxOccurs="unbounded">.
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     URL specifies contact endpoint of a bootstrap Information.
+                     Registration service. Further ISIS addresses will be queried.
+                     from this service.
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+
+            <xsd:element name="ServiceID" type="xsd:string" minOccurs="0" maxOccurs="1">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     Element defines the unique id of the service propagated outside.
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+
+            <xsd:element name="KeyPath" type="xsd:string"  minOccurs="0" maxOccurs="1">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     Optional KeyPath for SSL connection
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+
+            <xsd:element name="CertificatePath" type="xsd:string"  minOccurs="0" maxOccurs="1">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     Optional CertificatePath for SSL connection
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+             <xsd:element name="ProxyPath" type="xsd:string"  minOccurs="0" maxOccurs="1">
+                 <xsd:annotation>
+                     <xsd:documentation xml:lang="en">
+                      Optional ProxyPath for SSL connection
+                     </xsd:documentation>
+                 </xsd:annotation>
+             </xsd:element>
+
+            <xsd:element name="CACertificatesDir" type="xsd:string"  minOccurs="0" maxOccurs="1">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     Optional CACertificatesDir for SSL connection
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+
+            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1" default="5">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     Retry count. Specified how many times have an ISIS to timeout to handle it as.
+                     unavailable. If there is no value specified the default value: 5 will be used..
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+        </xsd:sequence>
     </xsd:complexType>
@@
     <xsd:complexType name="InfoRegister_Type">
-    <!-- Element for Service element to link it to InfoRegistrar
+        <xsd:annotation>
+            <xsd:documentation xml:lang="en">
+             Element for Service element to link it to Registrar.
              elements. It may also override some configuration parameters.
              Presence of this element means that service will be registered
-         to ISISes. -->
-        <xsd:sequence>
-            <!-- This elements specify which registrars must be used
+             to ISISes.
+            </xsd:documentation>
+        </xsd:annotation>
+        <xsd:sequence>
+            <xsd:annotation>
+                <xsd:documentation xml:lang="en">
+                 This elements specify which registrars must be used
                  for registering services. If there is no such element
-                 then registration is done using all registrars. -->
+                 then registration is done using all registrars.
+                </xsd:documentation>
+            </xsd:annotation>
+
+            <xsd:element name="Period" type="xsd:duration" minOccurs="1" maxOccurs="1" default="PT3M">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     Period specifies how often registration to be done in minutes.
+                     The 0 value means single registration.
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+
+            <xsd:element name="Endpoint" type="xsd:string" minOccurs="1" maxOccurs="1">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     This element defines URL of the service as seen from outside.
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+
+            <xsd:element name="Expiration" type="xsd:string" minOccurs="1" maxOccurs="1" default="PT10M">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     This element defines the expiration time of the information provided by the service.
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
             <xsd:element name="Registrar" type="iregc:Registrar_Type" minOccurs="0" maxOccurs="unbounded"/>
-            <!-- Default values for every InfoRegistrar -->
-
-            <!-- Period specifies how often registration to be done in minutes.
-                 The 0 value means single registration. -->
-            <xsd:element name="Period" type="xsd:nonNegativeInteger" minOccurs="1" maxOccurs="1"/>
-            <!-- Element defines the unique id of the service propagated outside. -->
-            <xsd:element name="ServiceID" type="xsd:string" minOccurs="0" maxOccurs="1"/>
-            <!-- This element defines URL of the service as seen from outside. -->
-            <xsd:element name="endpoint" type="xsd:string" minOccurs="1" maxOccurs="1"/>
-            <!-- This element defines the expiration time of the information provided by the service. -->
-            <xsd:element name="expiration" type="xsd:string" minOccurs="1" maxOccurs="1"/>
+            <!-- Default values for every Registrar -->
         </xsd:sequence>
@@
 
-    <!-- Configuration element force skipping the Self-Registration -->
     <xsd:element name="NoRegister">
+        <xsd:annotation>
+            <xsd:documentation xml:lang="en">
+             Configuration element force skipping the Self-Registration
+            </xsd:documentation>
+        </xsd:annotation>
         <xsd:complexType />
@@
   <!-- Various server settings -->
-  <infosys:InfoRegistrar id="NorduGridInfosys">
-    <!-- Contact and behavior details of the InfoRegsitrar element-->
-    <infosys:URL><!-- Available endpoint URL of an ISIS representing the cloud --></infosys:URL>
-    <infosys:Retry>4</infosys:Retry>
-  </infosys:InfoRegistrar>
 
@@
             </infosys:Expiration>
-            <!-- If there is no specific Registrar defined then every InfoRegistrar
-             will be use used during the registration. The period of the repeating 
+            <!-- The period of the repeating 
              registration will be the value given here in every ISIS cloud. -->
+            <infosys:Registrar>
+                <!-- Contact and behavior details of the Registrar element-->
+                <infosys:URL><!-- Available endpoint URL of an ISIS representing the cloud --></infosys:URL>
+                <infosys:Retry>4</infosys:Retry>        
+            </infosys:Registrar>
         </infosys:InfoRegister>

### 2009-10-15 doc/tech_doc/client/job_description_mapping.tex
Converting document to LaTeX. Please port to final.

\documentclass{article}
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage[english]{babel}
\usepackage{multirow}

\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}

\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}

\newcommand{\todo}{TODO}

\newcommand{\same}{{\textquotesingle}{\textquotesingle}}
\newcommand{\subsubsubsection}[1]{\paragraph{#1}}
\newcommand{\subsubsubsubsection}[1]{\subparagraph{#1}}

\newenvironment{inouttabular}%
{\begin{center}\begin{tabular}{l>{\ttfamily\footnotesize}c>{\ttfamily\footnotesize}c}%
\toprule
& \textnormal{\normalsize{In}} & \textnormal{\normalsize{Out}}\\ \cmidrule{2-3}}
{\bottomrule\end{tabular}\end{center}}

\makeatletter
\newcommand\arraybslash{\let\\\@arraycr}
\makeatother

%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

\raggedleft {NORDUGRID-XXXXXXX-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large The ARC Job Description}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Internal Representation Mapping}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large M. S. Andersen\footnote{skou@nbi.dk} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage

\section{Introduction}
This document contains information about how XRSL, JDL and the various
supported JSDL flavours are mapped to and from the job description
internal representation in ARC{}-lib.

\section{Data Structure}
The internal representation of the job description is contained in the
JobDescription class, which consist of the following objects:
JobIdentification, Application, Resources, DataStaging and JobMetaData.
JobIdentification contains data used to identify the job description
through various types, tags and names. The Application object is used
to describe explicitly the executable which should be executed at the
Computing Endpoint (CE), environment variables, logging, standard in
and out, credential service, etc. The Resources class contains
information about requirements and the type of execution node preferred
by the user. The DataStaging class contains information about input
files and output files/directories created by the job. Finally the job
description may contain meta data, describing the context in which the
description was generated, and it is stored in the JobMetaData class.

\subsection{JobIdentification}
\subsubsection{JobName}
\begin{table}[h]
\begin{inouttabular}
XRSL & jobname & jobname\\
JDL & No matching attribute & Not mapped\\
JSDL & JobIdentification/JobName & JobIdentification/JobName\\
ARC-JSDL &  & \\
\end{inouttabular}
\end{table}

\subsubsection{Description}
\begin{table}[h]
\begin{inouttabular}
ARC-JSDL & JobIdentification/Description & JobIdentification/Description\\
\end{inouttabular}
\end{table}

\subsubsection{JobType}
\begin{inouttabular}
ARC-JSDL & JobIdentification/JobType & JobIdentification/JobType\\
\end{inouttabular}
JDL has a jobtype and a type attribute, however a mapping has not been
made.

\subsubsection{UserTag}
\begin{inouttabular}
JDL & usertags & usertags\\
ARC-JSDL & JobIdentification/UserTag & JobIdentification/UserTag\\
\end{inouttabular}

\subsection{JobVOName}
\begin{inouttabular}
XRSL & jobname & jobname\\
ARC-JSDL & JobIdentification/JobVOName & JobIdentification/JobVOName\\
\end{inouttabular}

\subsection{Application}
\subsubsection{Executable}
\begin{inouttabular}
XRSL & executable, arguments & executable, arguments\\
JDL & Executable, Arguments & Executable, Arguments\\
\multirow{2}{*}{Posix-JSDL} & Application/POSIXApplication/Executable, & Application/POSIXApplication/Executable,\\
&  Application/POSIXApplication/Arguments & Application/POSIXApplication/Arguments\\
\multirow{2}{*}{HPCP-JSDL} & Application/HPCProfileApplication/Executable, & Application/HPCProfileApplication/Executable,\\
&  Application/HPCProfileApplication/Arguments & Application/HPCProfileApplication/Arguments\\
\multirow{2}{*}{ARC-JSDL} & Application/Executable/Path, & Application/Executable/Path,\\
 & Application/Executable/Argument & Application/Executable/Argument\\
\end{inouttabular}
The Executable data member is a composite structure of a string and a
list of strings, where the string holds the executable and the list
holding the arguments.

\subsubsection{Input}
\begin{inouttabular}
XRSL & stdin & stdin, inputFiles\\
JDL & StdInput & StdInput, InputSandbox\\
Posix-JSDL & Application/POSIXApplication/Input & Application/POSIXApplication/Input\\
HPCP-JSDL & Application/HPCProfileApplication/Input & Application/HPCProfileApplication/Input\\
ARC-JSDL & Application/Input & \\
\end{inouttabular}
If the Input data member does not exist in the file list, it will be
added to it.

\subsubsection{Output}
\begin{inouttabular}
XRSL & stdout & stdout, outputFiles\\
JDL & StdOutput & StdOutput, OutputSandbox, OutputSandboxDestURI\\
Posix-JSDL & Application/POSIXApplication/Output & \\
HPCP-JSDL & Application/HPCProfileApplication/Output & \\
ARC-JSDL & Application/Output & \\
\end{inouttabular}
If the Output data member does not exist in the File
list, it will be added to it.

\subsubsection{Error}
\begin{inouttabular}
XRSL & stderr & stderr, outputFiles\\
JDL & StdError & StdError, OutputSandbox, OutputSandboxDestURI\\
Posix-JSDL & Application/POSIXApplication/Error &\\
HPCP-JSDL & Application/HPCProfileApplication/Error &\\
ARC-JSDL & Application/Error & \\
\end{inouttabular}
If the Error data member does not exist in the File list, it will be
added to it.

\subsubsection{Join}
\begin{inouttabular}
XRSL & join & Not mapped\todo{}\\
ARC-JSDL & Application/Join & Application/Join\\
\end{inouttabular}

\subsubsection{Environment}
\subsubsubsection{Name}
\begin{inouttabular}
XRSL & environment* & environment*\\
JDL & Environment** & Environment**\\
Posix-JSDL & Application/POSIXApplication/Environment.name & Application/POSIXApplication/Environment.name\\
HPCP-JSDL & Application/HPCProfileApplication/Environment.name & Application/HPCProfileApplication/Environment.name\\
ARC-JSDL & Application/Environment/Name & ~\\
\end{inouttabular}
The XRSL attribute environment is a list of
pairs. The first element of each pair is mapped to the Name data member.
The JDL attribute Environment is a list of equalities.

\subsubsubsection{Value}
\begin{inouttabular}
XRSL & environment* & environment*\\
JDL & Environment** & Environment**\\
JSDL & No matching attribute & Not mapped\\
Posix-JSDL & Application/POSIXApplication/Environment & Application/POSIXApplication/Environment\\
HPCP-JSDL & Application/HPCProfileApplication/Environment & Applicaion/HPCProfileApplication/Environment\\
ARC-JSDL & Application/Environment/Value & ~\\
\end{inouttabular}

\subsubsection{Prologue}
\begin{inouttabular}
JDL & Prologue, PrologueArgument & Prologue, PrologueArgument\\
\multirow{2}{*}{ARC-JSDL} & Application/Prologue/Path, & Application/Prologue/Path,\\
& Application/Prologue/Argument & Application/Prologue/Argument\\
\end{inouttabular}

\subsubsection{Epilogue}
\begin{inouttabular}
JDL & Epilogue, EpilogueArgument & Epilogue, EpilogueArgument\\
\multirow{2}{*}{ARC-JSDL} & Application/Epilogue/Path, & Application/Epilogue/Path,\\
& Application/Epilogue/Argument & Application/Epilogue/Argument\\
\end{inouttabular}

\subsubsection{LogDir}
\begin{inouttabular}
XRSL & gmlog & gmlog outputFiles\\
ARC-JSDL & Application/LogDir & Application/LogDir\\
\end{inouttabular}

\subsection{RemoteLogging}
\begin{inouttabular}
XRSL & jobreport & jobreport\\
ARC-JSDL & Application/RemoteLogging & Application/RemoteLogging\\
\end{inouttabular}

\subsubsection{Rerun}
\begin{inouttabular}
XRSL & rerun & rerun\\
JDL & RetryCount ShallowRetryCount* & RetryCount ShallowRetryCount**\\
ARC-JSDL & Application/Rerun & Application/Rerun\\
\end{inouttabular}
* Only the maximum value between RetryCount
and ShallowRetryCount is stored.

** Both RetryCount and ShallowRetryCount will be set if
Rerun is set, and they will both have the value of Rerun.

\subsubsection{ExpiryTime}
\begin{inouttabular}
JDL & ExpiryTime & ExpiryTime\\
ARC-JSDL & Application/ExpiryTime & Application/ExpiryTime\\
\end{inouttabular}

\subsubsection{StartOnHold}
\begin{inouttabular}
\end{inouttabular}

\subsubsection{ProcessingStartTime}
\begin{inouttabular}
ARC-JSDL & Application/ProcessingStartTime & Application/ProcessingStartTime\\
\end{inouttabular}

\subsubsection{Notification}
\begin{inouttabular}
XRSL & notify & notify\\
ARC-JSDL & Application/Notification & Application/Notification\\
\end{inouttabular}

\subsubsection{CredentialService}
\begin{inouttabular}
XRSL & credentialserver & credentialserver\\
JDL & MyProxyServer & MyProxyServer\\
ARC-JSDL & Application/CredentialService & Application/CredentialService\\
\end{inouttabular}

\subsubsection{AccessControl}
\begin{inouttabular}
XRSL & acl & acl\\
ARC-JSDL & Application/AccessControl & Application/AccessControl\\
\end{inouttabular}

\subsection{Resources}
\subsubsection{OperatingSystem}
\begin{inouttabular}
XRSL & opsys & opsys\\
ARC-JSDL & Resources/OperatingSystem & Resources/OperatingSystem\\
\end{inouttabular}

\subsubsection{Platform}
\begin{inouttabular}
XRSL & architecture & architecture\\
JSDL & Resources/CPUAchitecture/CPUArchitectureName & ~\\
ARC-JSDL & Resources/Platform &~\\
\end{inouttabular}

\subsubsection{NetworkInfo}
\begin{inouttabular}
JSDL & Resources/IndividualNetworkBandwidth &\\
ARC-JSDL & Resources/NetworkInfo &\\
\end{inouttabular}

\subsubsection{NodeAccess}
\begin{inouttabular}
XRSL & nodeAccess & nodeAccess\\
ARC-JSDL & Resources/NodeAccess & ~\\
\end{inouttabular}

\subsubsection{IndividualPhysicalMemory}
\begin{inouttabular}
XRSL & memory & memory\\
Posix-JSDL & Application/POSIXApplication/MemoryLimit &\\
ARC-JSDL & Resources/IndividualPhysicalMemory &\\
\end{inouttabular}

\subsubsection{IndividualVirtualMemory}
\begin{inouttabular}
Posix-JSDL & Application/POSIXApplication/VirtualMemoryLimit &\\
ARC-JSDL & Resources/IndividualVirtualMemory &\\
\end{inouttabular}

\subsubsection{DiskSpaceRequirement}
\subsubsubsection{DiskSpace}
\begin{inouttabular}
XRSL & disk & disk\\
JSDL & Application/FileSystem/DiskSpace &\\
ARC-JSDL & Resources/DiskSpaceRequirement/DiskSpace &\\
\end{inouttabular}

\subsubsubsection{CacheDiskSpace}
\begin{inouttabular}
ARC-JSDL & Resources/CacheDiskSpace & Resources/CacheDiskSpace\\
\end{inouttabular}

\subsubsubsection{SessionDiskSpace}
\begin{inouttabular}
ARC-JSDL & Resources/SessionDiskSpace & Resources/SessionDiskSpace\\
\end{inouttabular}

\subsubsection{SessionLifetime}
\begin{inouttabular}
XRSL & lifeTime & lifeTime\\
ARC-JSDL & Resources/SessionLifeTime & Resources/SessionLifeTime\\
\end{inouttabular}

\subsubsection{SessionAccessMode}
\begin{inouttabular}
\end{inouttabular}

\subsubsection{IndividualCPUTime}
\begin{inouttabular}
JSDL & Resources/IndividualCPUTime* &\\
ARC-JSDL & Resources/IndividualCPUTime &\\
\end{inouttabular}
* The JSDL element is mapped to the Range part of
IndividualCPUTime.

\subsubsection{TotalCPUTime}
\begin{inouttabular}
XRSL & cpuTime & cpuTime\\
JSDL & Resources/TotalCPUTime* &\\
Posix-JSDL & Application/POSIXApplication/CPUTimeLimit* &\\
ARC-JSDL & Resources/TotalCPUTime &\\
\end{inouttabular}
* The JSDL and POSIX-JSDL elements is mapped to the
Range part of TotalCPUTime.

\subsubsection{IndividualWallTime}
\begin{inouttabular}
ARC-JSDL & Resources/IndividualWallTime &\\
\end{inouttabular}

\subsubsection{TotalWallTime}
\begin{inouttabular}
XRSL & wallTime & wallTime\\
Posix-JSDL & Application/POSIXApplication/WallTimeLimit &\\
ARC-JSDL & Resources/TotalWallTime &\\
\end{inouttabular}

\subsubsection{Homogeneous}
\begin{inouttabular}
\end{inouttabular}

\subsubsection{Benchmark}
\subsubsubsection{Type}
\begin{inouttabular}
\end{inouttabular}

\subsubsubsection{Value}
\begin{inouttabular}
\end{inouttabular}

\subsubsection{CEType}
\begin{inouttabular}
XRSL & middleware & middleware\\
ARC-JSDL & Application/CEType & Application/CEType\\
\end{inouttabular}

\subsubsection{SlotRequirement}
\subsubsubsection{NumberOfSlots}
\begin{inouttabular}
Posix-JSDL & Application/POSIXApplication/ProcessCountLimit &\\
ARC-JSDL & Resources/SlotRequirement/NumberOfSlots &\\
\end{inouttabular}

\subsubsubsection{ProcessPerHost}
\begin{inouttabular}
XRSL & count & count\\
JSDL & Resources/TotalCPUCount & \\
ARC-JSDL & Resources/SlotRequirement/ProcessPerHost &\\
\end{inouttabular}

\subsubsubsection{ThreadsPerProcesses}
\begin{inouttabular}
Posix-JSDL & Application/POSIXApplication/ThreadCountLimit &\\
ARC-JSDL & Resources/SlotRequirement/ThreadsPerProcesses &\\
\end{inouttabular}

\subsubsubsection{SPMDVariation}
\begin{inouttabular}
ARC-JSDL & Resources/SlotRequirement/SPMDVariation &\\
\end{inouttabular}

\subsubsection{CandidateTarget}
\subsubsubsection{EndPointURL}
\begin{inouttabular}
XRSL & cluster & Not mapped\\
JSDL & Resources/CandidateHosts/HostName &\\
ARC-JSDL & Resources/CandidateTarget/EndPointURL & Resources/CandidateTarget/EndPointURL\\
\end{inouttabular}

\subsubsubsection{QueueName}
\begin{inouttabular}
XRSL & queue & queue\\
JDL & QueueName & QueueName\\
ARC-JSDL & Resources/CandidateTarget/QueueName & Resources/CandidateTarget/QueueName\\
\end{inouttabular}

\subsubsection{RunTimeEnvironment}
\begin{inouttabular}
XRSL & runTimeEnvironment & runTimeEnvironment\\
ARC-JSDL & Application/RunTimeEnvironment & Application/RunTimeEnvironment\\
\end{inouttabular}

\subsection{DataStaging}
\subsubsection{File}
\subsubsubsection{Name}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{Source}
\subsubsubsubsection{URL}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{Threads}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{DataIndexingService}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{Target}
\subsubsubsubsection{URL}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{Threads}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{Mandatory}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{DataIndexingService}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{NeededReplica}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{KeepData}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{IsExecutable}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{DownloadToCache}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsection{Directory}

\subsection{JobMetaData}
\subsubsection{Author}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsection{DocumentExpiration}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsection{Rank}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsection{FuzzyRank}
\begin{inouttabular}
XRSL & ~ & ~\\
JDL & ~ & ~\\
JSDL & ~ & ~\\
Posix-JSDL & ~ & ~\\
HPCP-JSDL & ~ & ~\\
ARC-JSDL & ~ & ~\\
\end{inouttabular}

\section{XRSL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
XRSL input  & Internal representaion & XRSL output\\\hline
executable & Application.Executable.Name & executable\\
arguments & Application.Executable.Arguments & arguments\\
inputFiles & DataStaging.File.Name DataStaging.File.Source & inputFiles\\
executables & DataStaging.File.IsExecutable & executables\\
cache & TODO & TODO\\
outputFiles & DataStaging.File.Name DataStaging.File.Target & outputFiles\\
cpuTime & Resources.TotalCPUTime & cpuTime\\
wallTime & Resources.TotalWallTime & wallTime\\
gridTime & Not mapped & Not mapped\\
benchmarks & Not mapped & Not mapped\\
memory & Resources.IndividualPhysicalMemory & memory\\
disk & Resources.DiskSpaceRequirement.DiskSpace & disk\\
runTimeEnvironment & Resources.RunTimeEnvironment & runTimeEnvironment\\
middleware & Resources.CEType & middleware\\
opsys & Resources.OperatingSystem & opsys\\
stdin & Application.Input & stdin\\
stdout & Application.Output & stdout outputFiles\\
stderr & Application.Error & stderr outputFiles\\
join & Application.Join & Not mapped\\
gmlog & Application.LogDir & gmlog\\
jobName & Identification.JobName & jobName\\
ftpThreads & DataStaging.File.Source.Threads DataStaging.File.Target.Threads & Not mapped\\
acl & Application.AccessControl & acl\\
cluster & Resources.CandidateTarget.EndPointURL & Not mapped\\
queue & Resources.CandidateTarget.QueueName & queue\\
startTime & Application.ProcessingStartTime & Not mapped\\
lifeTime & Resources.SessionLifeTime & lifeTime\\
notify & Application.Notification & notify\\
replicaCollection & DataStaging.File.DataIndexingService & Not mapped\\
rerun & Application.Rerun & rerun\\
architecture & Resources.Platform & architecture\\
nodeAccess & Resources.NodeAccess & Not mapped\\
dryRun & Not mapped & Not mapped\\
rsl\_substitution & Processed internally by the parser & Not mapped\\
environment & Application.Environment & environment\\
count & Resources.SlotRequirement.ProcessPerHost & count\\
jobreport & Application.RemoteLogging & jobreport\\
credentialserver & Application.CredentialService & credentialserver\\
\end{tabular}
\end{center}
\end{table}

\section{JDL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
JDL input  & Internal representaion & JDL output\\\hline
JobType & Not mapped & Not mapped\\
Executable & Application.Executable.Name & Executable InputSandbox\\
Arguments & Application.Executable.Arguments & Arguments\\
StdInput & Application.Input & StdInput InputSandbox\\
StdOutput & Application.Output & StdOutput OutputSandbox OutputSandboxDestURI\\
StdError & Application.Error & StdError OutputSandbox OutputSandboxDestURI\\
InputSandbox & DataStaging.File.Name DataStaging.File.Source & InputSandbox\\
InputSandboxBaseURI & DataStaging.File.Source.URI & InputSandbox\\
OutputSandbox & DataStaging.File.Name DataStaging.File.Target.URI & OutputSandbox OutputSandboxDestURI\\
OutputSandboxDestURI & DataStaging.File.Target DataStaging.File.KeepData & OutputSandbox OutputSandboxDestURI\\
OutputSandboxBaseDestURI & DataStaging.File.Target.URI & OutputSandbox OutputSandboxDestURI\\
Prologue & Application.Prologue.Name & Prologue\\
PrologueArguments & Application.Prologue.Arguments & PrologueArguments\\
Epilogue & Application.Epilogue.Name & Epilogue\\
EpilogueArguments & Application.Epilogue.Arguments & EpilogueArguments\\
AllowZippedISB & Stored internally, not processed & AllowZippedISB\\
ZippedISB & Stored internally, not processed & ZippedISB\\
ExpiryTime & Application.ExpiryTime & ExpiryTime\\
Environment & Application.Environment & Environment\\
PerusalFileEnable & Stored internally, not processed & PerusalFileEnable\\
PerusalTimeInterval & Stored internally, not processed & PerusalTimeInterval\\
PerusalFilesDestURI & Stored internally, not processed & PerusalFilesDestURI\\
InputData (Deprecated) & Not mapped & Not mapped\\
OutputData (Deprecated) & Not mapped & Not mapped\\
StorageIndex (Deprecated) & Not mapped & Not mapped\\
DataCatalog (Deprecated) & Not mapped & Not mapped\\
DataRequirements & Stored internally, not processed & DataRequirements\\
DataAccessProtocol & Stored internally, not processed & DataAccessProtocol\\
OutputSE & Stored internally, not processed & OutputSE\\
VirtualOrganisation & Identification.JobVOName & VirtualOrganisation\\
RetryCount & Application.Rerun & RetryCount ShallowRetryCount\\
ShallowRetryCount & Application.Rerun & RetryCount ShallowRetryCount\\
LBAddress & Stored internally, not processed & LBAddress\\
MyProxyServer & Application.CredentialService & MyProxyServer\\
HLRLocation & Stored internally, not processed & HLRLocation\\
JobProvenance & Stored internally, not processed & JobProvenance\\
NodeNumber & Stored internally, not processed & NodeNumber\\
JobSteps (Deprecated) & Not mapped & Not mapped\\
CurrentStep (Deprecated) & Not mapped & Not mapped\\
JobState (Deprecated) & Not mapped & Not mapped\\
ListenerPort & Stored internally, not processed & ListenerPort\\
ListenerHost & Stored internally, not processed & ListenerHost\\
ListenerPipeName & Stored internally, not processed & ListenerPipeName\\
Requirements & Stored internally, not processed & Requirements\\
Rank & JobMeta.Rank & Rank\\
FuzzyRank & JobMeta.FuzzyRank & FuzzyRank\\
UserTags & Identification.UserTag & UserTags\\
BatchSystem & Stored internally, not processed & BatchSystem\\
QueueName & Resources.CandidateTarget.QueueName & QueueName\\
\end{tabular}
\end{center}
\end{table}

\section{JSDL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
JSDL input  & Internal representaion & JSDL output\\\hline
\bfseries JobIdentification & &\\
JobName & &\\
JobAnnotation & &\\
JobProject & &\\
\bfseries Application & &\\
ApplicationName & &\\
ApplicationVersion & &\\
\bfseries Resources & &\\
CandidateHosts/HostName & &\\
FileSystem/FileSystemType & &\\
FileSystem/Description & &\\
FileSystem/MountPoint & &\\
FileSystem/DiskSpace & &\\
ExclusiveExecution & &\\
OperatingSystem/OperatingSystemType/OperatingSystemName & &\\
OperatingSystem/OperatingSystemVersion & &\\
CPUArchitecture/CPUArchitectureName & &\\
IndividualCPUSpeed & &\\
IndividualCPUTime & &\\
IndividualCPUCount & &\\
IndividualNetworkBandwidth & &\\
IndividualPhysicalMemory & &\\
IndividualVirtualMemory & &\\
IndividualDiskSpace & &\\
TotalCPUTime & &\\
TotalCPUCount & &\\
TotalPhysicalMemory & &\\
TotalVirtualMemory & &\\
TotalDiskSpace & &\\
TotalResourceCount & &\\
\bfseries DataStaging & {}-{}- & {}-{}-\\
FileName & &\\
FileSystemName & &\\
CreationFlag & &\\
DeleteOnTermination & &\\
Source/URI & &\\
Target/URI & &\\
\end{tabular}
\end{center}
\end{table}

\section{POSIX-JSDL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
POSIX-JSDL input  & Internal representation & POSIX-JSDL output\\\hline
Executable & &\\
Argument & &\\
Input & &\\
Output & &\\
Error & &\\
WorkingDirectory & &\\
Environment & &\\
WallTimeLimit & &\\
FileSizeLimit & &\\
CoreDumpLimit & &\\
DataSegmentLimit & &\\
LockedMemoryLimit & &\\
OpenDescriptorsLimit & &\\
PipeSizeLimit & &\\
StackSizeLimit & &\\
CPUTimeLimit & &\\
ProcessCountLimit & &\\
VirtualMemoryLimit & &\\
ThreadCountLimit & &\\
UserName & &\\
GroupName & &\\
\end{tabular}
\end{center}
\end{table}

\section{HPCProfile-JSDL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
HPCProfile-JSDL input  & Internal representaion & HPCProfile-JSDL output\\\hline
Executable & & \\
Argument & & \\
Input & & \\
Output & & \\
Error & & \\
WorkingDirectory & &\\
Environment & &\\
UserName & &\\
\end{tabular}
\end{center}
\end{table}
\end{document}

### 2009-10-15 doc/tech_doc/chelonia/arc-storage-documentation.tex
rename confusing configuration attribute in Shepherd's backend config (DataDir -> StoreDir)
please port to final

@@
     \item[BackendClass] specifes which backend the Shepherd should use to communicate with the storage element service.
-    \item[BackendCfg] specifies configuration parameters for the backend class: \textbf{DataDir} and \textbf{TransferDir} are the directories the backend will use (see Section~\ref{sub:shepherd_implementation}), \textbf{TURLPrefix} is the URL of the storage element service ending with a `/'.
+    \item[BackendCfg] specifies configuration parameters for the backend class: \textbf{StoreDir} and \textbf{TransferDir} are the directories the backend will use (see Section~\ref{sub:shepherd_implementation}), \textbf{TURLPrefix} is the URL of the storage element service ending with a `/'.
     \item[LibrarianURL] is the URL of a Librarian where the Shepherd should send the reports about the health of the stored files.
@@
         <BackendCfg>
-            <DataDir>./shepherd_store</DataDir>
+            <StoreDir>./shepherd_store</StoreDir>
             <TransferDir>./shepherd_transfer</TransferDir>

### 2009-10-15 src/services/storage/shepherd/byteio.py

@@ -19,7 +19,7 @@ class ByteIOBackend:
         self.file_arrived = file_arrived
         self.ns = arc.NS('she', ns_uri)
         self.ssl_config = ssl_config
-        self.datadir = str(backendcfg.Get('DataDir'))
+        self.datadir = str(backendcfg.Get('StoreDir'))
         self.transferdir = str(backendcfg.Get('TransferDir'))
         self.turlprefix = str(backendcfg.Get('TURLPrefix'))
         if not os.path.exists(self.datadir):

### 2009-10-17 doc/tech_doc/janitor/tex_concept/concept.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

@@
 dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
- integration with the Grid Manager and the Web server) for the Janitor. In
+integration with the Grid Manager and the Web service) for the Janitor. In
 the pre-web-service implementation the Catalog remains a static web page.
@@
 \section{Runtime environments states}
+
 A main motivation for the managed, manual initiation of dynamic RE
@@
 packages -- prior to their use in production.
+
 With an automation of the installation, the verification of
@@
 These states are specific for every CE and communicated
-between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
+between the Janitor and the execution service \AREX. Table~\ref{tab:states} shows all
 possible states, while Figure~\ref{fig:RE_states}
 displays the transitions between the states that a Runtime Environment may
-be in during its life time at a particular CE. \task{?at a particular CE?}
+be in during its life time.
 
@@
 
+The concept of ARC prevails to have a single directory into which to
+install software to be execute on all compute elements of the site.
+Otherwise, the installation of a runtime environment would be
+required to be performed just when the jobs initiates its computation.
+
+In principle this would be doable, particularly for those runtime 
+environments that are part of a trusted Linux distribution already.
+But this would also mean that the compute nodes become inhomogeneous,
+and at the time of writing this was considered undesireable.
+
 \begin{figure}[!h]
@@
 The manually induced transitions are marked in red, he automated transitions in black.
-A transition between states can be induced automatically (i.\,e.\ by the
+A transition between states can be induced automatically (\ie by the
 advent of a job requesting a particular dynamic RE) or manually by the
-site's supervisor or an individual with respective rights to use the
-Janitor's web service.
+site's administrator or an individual with respective rights to use the
+Janitor's command line.
 
@@
 
-% Runtime Environment are installed for job idenitified by a certain id.
+\subsection{Dependencies on Job IDs}
+
+The Janitor was designed with the job execution in mind. As such, all
+actions it performs are driven by the need of a job. And those jobs
+should then be specified, to learn when the demand for a particular
+runtime environment has ended.
+
+This concept partially conflicts with the idea to use the Janitor
+as an aid for the manual invocation of an installation -- there is no
+job ID that could be assigned for that process. Today, the administrator
+is suggested to use a special number, \eg 0, to indicate such manual
+installs. This will change in near future.
 
@@
 
-Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
-in Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached in Janitor. 
-After Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
-an unforeseen exception occures during that process, Janitor will drop the job from its database and set the affected runtime
-environments to the state \texttt{FAILED}.
+The Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}.
+After a job has been succesfully registered with the Janitor, its state
+will be set to \texttt{PREPARED}. Invalid jobs are not cached by the Janitor. 
+After the Janitor is requested to deploy the runtime environment, the
+state of the job will change to \texttt{INITIALIZED}. If an unforeseen
+exception occures during that process, Janitor will drop the job from
+its database and set the affected runtime environments to the state
+\texttt{FAILED}.
 
@@
 {\hspace{\stretch{1}}\rule{1ex}{1ex}\\}
-\begin{note}
-The integration into AREX is not completed yet!
-\end{note}
-\begin{note}
-Thus, there will be bigger changes here in this section\dots
-\end{note}
 
@@
 
-Currently the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
-a setuid wrapper has been written. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s}
-to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\
-
-Thus, Janitor is not executed directly. Instead a dynamic link will be created in \task{libexec or sbin} which is pointing to
-the wrapper.\\
-
-It is suggested to create a user and a group "janitor" for the Janitor.\\
-
-% Currently the Janitor must be executed as root. For this a suid root helper
-% is needed. It is provided in /opt/janitor/rjanitor.c. If you installed
-% the Janitor in another directory then /opt/janitor change the parameter
-% of the execv command in rjanitor.c. Then compile the file with\\
-% \texttt{\textdollar gcc -o rjanitor rjanitor.c}\\
-% and run as root\\
-% \texttt{\textdollar chown root:root rjanitor}\\
-% \texttt{\textdollar chmod u+s rjanitor}\\
-% to make the binary suid root. A future version of the Janitor will get rid
-% of the suid root helper.\\
-% 
-% The Grid Manager does not call the Janitor directly in /opt/janitor but
-% calls /opt/nordugrid/libexec/janitor. So a link is needed (as root):\\
-% \texttt{\textdollar ln -s /opt/janitor/rjanitor /opt/nordugrid/libexec/janitor}\\
-% \\
-% It is suggested to create a user and a group "janitor" for the Janitor.\\
-% \\
-% The Janitor needs a directory for storing information on installed REs:\\
-% \texttt{\textdollar mkdir -p /var/lib/nordugrid/janitor}\\
-% \texttt{\textdollar chown janitor:janitor /var/lib/nordugrid/janitor}\\
-% \\
-% It also needs a directory on the shared volume for installing the REs:\\
-% % \texttt{\textdollar mkdir -p $<$shared_volume$>$/janitor}\\
-% % \texttt{\textdollar chown janitor:janitor $<$shared_volume$>$/janitor}
+At the time of writing, the Janitor needs to be executed as root. 
+This may change over time, but as long as all other components
+of ARC also demand superuser privileges, one can expect the
+demand for the Janitor to be non-problematic.
 
@@
 based solely on tar files, this is barely possible, unless such is
-performed by the install scripts that accompany the tar files. However,
-hereto the installation would have to be performed by a user with system
-priviledges, for which there is no technical requirement.
+performed by the install scripts that accompany the tar files.
+%However,
+%hereto the installation would have to be performed by a user with system
+%priviledges, for which there is no technical requirement.
 
@@
 
-Summarising, there is general concern about the security of grid
-computing.  Dynamic REs introduce new dangers since a
-manual control at the grid site is substituted by a remote process that
+As a conclusion, there is general concern about the security of grid
+computing.  Dynamic REs introduce new dangers since a manual control
+at the grid site is substituted by a remote process that
 is out the direct supervision of a local site administrator. The signing

### 2009-10-17 doc/tech_doc/janitor/tex_future/future.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

--- 2009-05-14 doc/manuals/janitor/tex_future/future.tex
@@
 to distinguish installed REs from installable REs, potentially offer
-description of extended RE state-like information. This work is planned
+descriptions of extended RE state-like information. This work is planned
 to be carried out as part of the Glue-2.0 effort of OGF\footnote{OGF
@@
 
-TODO:\\
+\section{Use of RDF}
+
+The RDF is "correctly" used within the Janitor via the Redland libraries.
+What is still missing is a semantical reasoning on what packages to
+allow or disallow. For instance: allow all packages associated with
+bioinformatics. This shall wait a bit longer until more experiences
+on the community's demand for this service have been made.
+
+\section{Manual Verification}
+
+There is yet no explicit notion of a concept on how users can report
+on the reliability of individual compute nodes and use such information
+for the decision making on where to sent their jobs. And with an increasing
+complexity of software installed, being used across versions, one will
+rather trust one's very own experiences for individual sites than some
+external repository or other pieces of information that may be weeks
+or months old.
+
+Consequently, no means for a manual verification and the communication
+of results of such have yet been implemented.
+
+\section{Known issues}
 \begin{itemize}
@@
  \item Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change that
- \item Verfication
 \end{itemize}

### 2009-10-17 doc/tech_doc/janitor/tex_installation/installation.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

--- 2009-05-14 doc/manuals/janitor/tex_installation/installation.tex
@@
 
-In order to use Janitor, the two perl packages listed in table~\ref{tab:install_dependencies} are required.
-To have the WebService interface for Janitor, the packages listed in the table~\ref{tab:install_dependencies_optional}
-have to be installed before the build process is.
+The Janitor depends on two perl packages listed in table~\ref{tab:install_dependencies}. To have the WebService interface for Janitor,
+the packages listed in the table~\ref{tab:install_dependencies_optional} as build requirements.
 \begin{table}[!h]
    \begin{center}
-        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
+        \mycaption{Required perl packages for the Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
                    Redland RDF libraray is used for accessing the knowledge base of RTEs.}
@@
 
-Janitor will be installed by default along with ARC1. If desired it is possible to disable the shipping of Janitor 
-using the \textit{configure} flags \textit{--disable-janitor-service} for the complete janitor or 
-\textit{--disable-janitor-webservice} for only the Web Service support.
+If you are using regular Debian or Ubuntu packages, then the Janitor
+can be installed as root by "apt-get install nordugrid-arc1-janitor".
+Installing it will not drag other components of ARC with it since the
+Janitor can be used in its own right or in conjunction with another
+grid system, possibly. Packages for Redhat/Fedora and SuSE/OpenSuSE are
+also provided, the redland library however may not yet be available for
+those systems.  If you are compiling the sources yourself, the Janitor
+will be enabled by default. If desired, it is possible to disable it
+with the \textit{configure} flags \textit{--disable-janitor-service}
+for the complete janitor or \textit{--disable-janitor-webservice} for
+only the Web Service support.
+
 Furthermore it is recommended to install the ontology editor 
@@
 
-The parameter \texttt{enable} defines whether Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
-\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
-The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
-In these directories the states of the jobs and the runtime environments will be stored. 
-The knowledge base of installable packages is specified by the parmeter \texttt{catalog}. 
-Its value can be any kind of URL pointing to an file written in the Resource Description Framework (RDF) format.
-The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
-The parameter \texttt{downloaddir} assignes the directory in which the installation files will be saved after they have been 
-downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
-directory into which all packages will be installed. Instead of the other directories the \texttt{installationdir} should be 
-available for all computing elements i.e. by using a shared volume.
-If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, Janitor will also
-create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by Janitor.
-The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
-The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
-The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
-or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintain by
-a higher organization.
-The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
-An examples how to configure arc and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf} and.
+The parameter \texttt{enable} defines whether the Janitor shall be
+used within \AREX or not. Use the value \texttt{"0"} to disable the
+Janitor. The \texttt{uid} and the \texttt{gid} are defining which
+effective user and group id shall be used for Janitor. Setting these to
+values different from root will increase security.
+
+The \texttt{registrationdir} describes the directory in which
+the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
+In these directories the states of the jobs and the runtime environments
+are stored. Please recall that the Janitor does not use a database as
+a backend, but all communication between invocations are performed via
+files in those folders.
+
+The knowledge base of installable packages is specified by the parmeter
+\texttt{catalog}.  Its value can be any kind of URL pointing to a file
+written in the Resource Description Framework (RDF) format. One should
+not light-heartedly use a remote address for this purpose. Such a remote
+source needs to be trusted.
+
+The specification of the RDF file will be explained in detail in
+section~\ref{sec:catalog}.  The parameter \texttt{downloaddir} assignes
+the directory in which the installation files will be saved after they
+have been downloaded or copied from the repository which was specified by
+the catalog. Please remember: the URL in arc.conf indicates the location
+of the catalog. And the URLs somehow specified in the catalog specify
+the location from where to download the runtime environment.
+
+The \texttt{installationdir} finally specifies the directory into which
+all packages will be installed. This \texttt{installationdir} directory
+needs to be available for all computing nodes for the execution of 
+arbitrary programs, mostl likely by using it as a shared NFS volume.
+
+If the configuration file furthermore contains the \texttt{runtimedir}
+tag within the section \texttt{grid-manager}, the Janitor will
+also create a symbolic link in the \texttt{runtimedir} pointing to
+the configuration script of the installation done by Janitor. The
+tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used
+for automated cleanup and are defined in seconds.  The default
+value for thew\texttt{jobexpirytime} is seven days and for
+the \texttt{rteexpirytime} three days.  The additional tags
+\texttt{allow\_base}, \texttt{deny\_base}, \texttt{allow\_rte} and
+\texttt{deny\_rte} are used to include or exclude certain base packages
+or runtime environments of the catalog. This feature is useful, if the
+catalog is maintain by a higher organization. But again: you need to
+trust it.
+
+The path to the log4perl configuration file is defined by the tag
+\texttt{logconf}.  Examples how to configure arc and log4perl are
+provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.
 
@@
 
-Janitor is designed for the usage on Linux distributions.
-
-% Is designed for linux distributions. Windows or Mac won't work (Reason wget! maybe also the Executer).
+The Janitor is only available for UNIX-based servers. While not tested,
+it should also be functional on MacOS X and Windows with Cygwin.

### 2009-10-17 doc/tech_doc/janitor/tex_introduction/introduction.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

@@
 
-Janitor is a service for automated runtime environment installation for grid computing elements.
-It is integrated in the A-REX service used for job submission but also can be used as a standalone tool.
+The {\em Janitor} is a service for the automated installation of runtime environments for grid computing elements.
+Its command line interface allows for a direct interaction with site administrators. However, the main stimulus
+for its development was the idea integrate such a service with the regular handling of compute jobs. For ARC this
+is performed by the \AREX module.
+
+From the programmer's view, the Janitor is mostly a Perl script and the routines within \AREX to invoke it. The
+site administrator will also associate with it also the Catalog files, that describe the availability of runtime
+environments, and the repository of installable runtime envionments themselves, which are regular tar archives obeying
+to particular structure and reside in a separate folder. In order to minimise the latency for the invocation of
+the Perl script and the associated parsing of files, a Janitor web service was developed, which still is a Perl
+script.
 
@@
 A major motivation for grid projects is to stimulate new communities to adopt computational grids for their
-causes. From the current grid user's viewpoint, the admission of users of a very different education will
-suddenly impose difficulties in the communication between site maintainers. One will not even understand
-the respective other side's research aims. Hence, the proper installation of non-standard software (Runtime
-Environments) is not guaranteed and the priorities of manual labour will be mostly disjunctive.
+causes. From the current grid user's viewpoint, the admission of users of a very different scientific disciplines
+or compute skills will impose difficulties in the communication between site maintainers. One will not even understand
+the respective other side's research aims. Hence, the proper installation of non-standard software (read Runtime
+Environments) is not guaranteed.
 
-A core problem remains to distribute a locally working solution, the Know-How, quickly across all contributing sites, i.
-e., without manual interference.
-Every scientific discipline has its respective own set of
-technologies for the distribution of work load. E. g. research in bioinformatics requires access to so many
+A core problem remains to distribute a locally working solution, the Know-How, quickly across all contributing
+sites, i. e., without manual interference. Every scientific discipline has its respective own set of
+technologies for the distribution of work load. For instance, research in bioinformatics requires access to so many
 different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
 a commodity, with all the problems with respect to bottlenecks and restrictions of repeated access.
-The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges with the here presented work extends the NorduGrid?s
-Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
+The EU project KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}},  amongst other challenges,
+with the here presented work extends the NorduGrid?s
+Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an the
 automated installation of software packages.
@@
 This document will start with a chapter of how to set up janitor locally. The following chapter will give an instruction on how to 
-use Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
-which is basically covering the method how to prepare new runtime environments. A deeper knowledge about the design of Janitor will 
-be given by the subsequent forth chapter. In the fifth chapter, an outlook to future developments will given.
+use Janitor with \AREX and/or without \AREX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
+which is basically covering the preparation of runtime environments. Deeper knowledge about the design of the Janitor will 
+be given by the subsequent forth chapter. In the fifth chapter, an outlook on ongoing or future developments will presented.
 

### 2009-10-17 doc/tech_doc/janitor/tex_maintenance/maintenance.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

@@
 
-This chapter explains how to maintain the catalog and Janitor itself. 
-In order to administrate the catalog, it is absolutly recommened (but not required) to use the ontology editor Prot\`eg\`e. The first
+This chapter explains how to maintain the catalog and the Janitor itself. 
+In order to administrate the catalog, it is absolutly recommended (but not required) to use the ontology editor Prot\`eg\`e. The first
 section will explain how this is done.
-The second section a detailed explanation how to create new packages for Janitor will be given.
-In the last section typical use case in maintaining Janitor will be listed.
+The second section gives a detailed explanation how to create new packages for the Janitor.
+In the last section, a typical use case in maintaining Janitor will be demonstrated.
 
@@
 
-The Catalog describes runtime environments and is either served through a web server or distributed together with the Janitor. 
-It is specifed by a (Resource Description Framework) RDF file assigned to 
-Janitor using the tag \texttt{catlog} within the configuration file (see~\ref{sec:janitor_configuration}). 
-The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
-which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
+The Catalog describes runtime environments and is either served through a web server or distributed
+together with the Janitor.  It is specifed by a (Resource Description Framework) RDF file assigned to 
+the Janitor using the tag \texttt{catlog} within the configuration file (see~\ref{sec:janitor_configuration}). 
+The format of the RDF file is defined by an RDF schema file \texttt{knowarc.rdfs}
+which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory
+\href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
 {http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\
@@
 \end{figure}
-On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
-The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
-It has one or more instances of the class \texttt{Package} and is
-described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
-and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
-(i.e. here etch or sid). The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
-they can be easier found.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
-They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
-classes are interacting with each other the Tables~\ref{tab:main_class_meta},\ref{tab:main_class_base}, \ref{tab:main_class_tag},
+On the left side of the editor the class browser is placed. Three
+main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and
+\texttt{Package}.  The \texttt{Metapackage} is a general platform
+independent description of a \texttt{Package}. It can be understood as
+a reference to a functionality that should be implemented at the remote
+site. But the exact instruction on how to install it is not given.
+The "instruction-level" comes with instances of its subclass
+\texttt{Package}.
+
+The \texttt{MetaPackage} is described by the subclasses of \texttt{Note}.
+The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
+and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem}
+describes the Debian release a \texttt{Package} refers to
+(i.e. here etch or sid). The class \texttt{Tag} provides small keywords
+which can be assigned to \texttt{MetaPackages} such that
+they can be easier found.  \texttt{TarPackage} and \texttt{DebianPackage} or
+currently the only subclasses of \texttt{Package}.
+
+They are representing the necessary information (i.e. URL or Packagename)
+for the installation. To provide an overview on how the classes are
+interacting with each other the Tables~\ref{tab:main_class_meta},\ref{tab:main_class_base}, \ref{tab:main_class_tag},
 \ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.
@@
    \begin{center}
-        \mycaption{Specification of the class Metapackage.}{}
+        \mycaption{Specification of class \texttt{Metapackage}.}{}
          \label{tab:main_class_meta}
@@
    \begin{center}
-        \mycaption{Specification of the class Basesystem.}{}
+        \mycaption{Specification of class \texttt{BaseSystem}.}{}
          \label{tab:main_class_base}
@@
    \begin{center}
-        \mycaption{Specification of the class Tag.}{}
+        \mycaption{Specification of class \texttt{Tag}.}{}
          \label{tab:main_class_tag}
@@
    \begin{center}
-        \mycaption{Specification of the class DebianPackage.}{}
+        \mycaption{Specification of class \texttt{DebianPackage}.}{}
          \label{tab:main_class_debian}
@@
    \begin{center}
-        \mycaption{Specification of the class TarPackage.}{}
+        \mycaption{Specification of the class \texttt{TarPackage}.}{}
          \label{tab:main_class_tar}
@@
 
+\subsection{Debian packages - dysfunctional in current implementation}
+
+The problem with Debian packages is that these are available only for the
+local machine and not immediately also for the whole network. This feature
+was meant for setups that use virtual machines for the execution of jobs.
+
+Those entries can also be used to help the specification of further
+dependencies of runtime environments. It would then be left to the
+responsibility of the system administrator to manually (or assisted with
+scripts) distribute a series of extra Debian packages throughout the
+compute nodes and use the catalog entry merely to indicate their presence.
+
+One can as such interpret the catalog to represent an interface between
+software distributed via Linux distributions and independently from these
+via grid communities. It should be noted that the Linux distributions
+have now all started to accept communities to maintain packages, which
+may bringt many scientific packages away from being traditional ARC 
+runtime environments towards becoming regular packages of some Linux
+distribution. For Debian, the Debian-Science and
+Debian-Med\footnote{\href{http://debian-med.alioth.debian.org}{http://debian-med.alioth.debian.org}}
+communities are known to be very open to grid and cloud computing.
+
 \section{HTML interface of the catalog}
 
-The dynamic Runtime Environments stored in the catalog are presented on the aforementioned dedicated web
-page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
-RDF syntax and an automated transformation to
-HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
-Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
-That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
-This html page listing the manually or automatically installable RTEs is prepared by the script web/list.pl.
-This script is meant to be run by a mod-perl enabled Apache. The script itself is only loosely integrated
-within the Janitor. In the first lines of the script some variables specific to the site are set. To configure the
-script these have to be changed~\cite[p. 9]{BAYER_2007}.
+The dynamic Runtime Environments stored in the
+catalog are presented on the aforementioned dedicated web
+page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also
+links to both the formal Catalog in RDF syntax and an automated
+transformation to HTML. The latter mimics the traditional
+site describing Runtime Environments in the Runtime Environment
+Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues
+with an eventual transition to the new system.
+
+That traditional page collects descriptions for Runtime Environments to
+encourage human site administrators to install these.  This html page
+listing the manually or automatically installable RTEs is prepared by
+the script web/list.pl.  This script is meant to be run by a mod-perl
+enabled Apache web server. The script itself is only loosely integrated
+with the remaining functionality of the Janitor, used only for the
+presentation to users. In the first lines of the script some variables
+specific to the site are set. To configure the script these have to be
+changed~\cite[p. 9]{BAYER_2007}.
 
@@
 
-This section describes how to add new packages to the catalog. Currently only tar based packages are processable by Janitor.
-Within the example they are assigned to be used together with Debian Etch. This limitation is only literal. 
-There should be no limitiation
-to newer Debian distributions.
+This section describes how to add new packages to the catalog. Currently
+only tar based packages are processable by the Janitor.  Within the here
+presented example the packages are assigned to be used together with
+Debian Etch. This limitation is only literal. 
+There should be no limitiation to newer Debian distributions.
 
@@
 
-At the time of writing, only the tape archive (tar) file format is accepted for dynamic runtime environment installation, 
-a well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files. 
+At the time of writing, only the tape archive (tar) file format is accepted for dynamic runtime environment installation. 
+It is well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files
+for the representation of dynamic runtime environments. 
 Subdirectories are visualised in Figure~\ref{fig:tar_folder}.
@@
 
-The tar file contains two directories named control and data. Software is stored in the latter subdirectory,
-while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
-the content of the data directory is extracted to some directory \textdollar BAR. 
-After this unpacking of the tar file, the Janitor executes the install skript provided in the control directory. 
+The tar file contains two directories:
+\begin{description}
+\item [data/] contains all software that the grid-job may need
+\item [control/] contains files formally specifying how to deal with the information in the data/ directory.
+\end{description}
+
+While installing such tar-based runtime environments, the content of the data directory is extracted to some directory \textdollar BAR. 
+After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
 It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
-The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
-before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
-Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
-installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
-To be offered to computing elements for an installation, the such prepared runtime environment must be
-announced to a Catalog to which the Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.
+But the working directory shall not be moved. All post-processing needs to be performed \textit{in situ}.
 
-\subsection{Protoypes}
+The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install}, just
+before the tar-package is removed. In most cases \texttt{control/remove} will be empty, implying that the working directory
+\textdollar BAR shall be remove and no other action is required.
 
-In order to have an impression how the tar files are created, several prototypes are provided at \task{URL required!!}.:\\
+The remainder actions are regular actions performed upon execution of every grid-job. Upon submission,
+the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. Every
+ARC runtime environment must specify such a runtime script, new is only its specific location as control/runtime.
+Since the directory \textdollar BAR is not known for the individual preparing the runtime environment, that
+file will instead use the placeholder \%BASEDIR\%.  After installing the package, the Janitor changes all
+occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.  To be offered to computing elements
+for an installation, the such prepared runtime environment must be announced to a Catalog to which the
+Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.
+
+\subsection{Example: WEKA machine learning Java library}
+
+%For an impression how the tar files are created, several prototypes are provided at \task{URL required!!}.:\\
 
 The WEKA package for machine learning~\cite{FRANK_2004} and the Java Runtime Environment
-are available as dynamic Runtime Environments. Further packages for bioinformatics comprise dynamic
+are made available as dynamic Runtime Environments. Further packages for bioinformatics comprise dynamic
 variants of tools for the analysis of transcription factor binding sites. These are already offered for manual
@@
 
-\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
+\subsection{Example: R packages in CRAN and BioConductor.org}
+
+To demonstrate the technical proximity to scientific communities that
 provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
@@
 thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
-projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\
+projects.
+
+This effort has not found practical applications beyond this proof of concept. R packages are now
+frequently found in regular Linux distributions. And for most packages, it is reasonably well possible
+to install the packages at runtime.
+
+\subsection{Example: ATLAS for High Energy Physics}
 
@@
 
-\subsection{Adminstrating the Catalog}
-
-\section{Typical use cases}\label{sec:typical_use_cases}
-
-% ** setstate REMOVAL_PENDING
-% no warranty interferences in the file syste,
+\section{Adminstrating the Catalog}

### 2009-10-17 doc/tech_doc/janitor/tex_titlepage/titlepage.tex
Adding Janitor documentation from janitor branch.
Round of documentation updates - mostly semantics-invariant linguistic changes.

--- 2009-05-25 doc/manuals/janitor/tex_titlepage/titlepage.tex
@@
 {\centering \textsc{\Large Dynamic Runtime Environment Installation with Janitor}
-\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
+\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{The Janitor and this document with it is under continuous development. Your comments and suggestions are appreciated.} % DELETE ONLY THAT LINE!!!
 \Large \par}
@@
 %%%%---- A list of authors ----
-    {\centering \large Michael Glodek\footnote{glodek@inb.uni-luebeck.de}, Daniel Bayer,  Steffen M\"oller\footnote{moeller@inb.uni-luebeck.de} \par}
+    {\centering \large Michael Glodek, Daniel Bayer,  Steffen M\"oller\footnote{moeller@inb.uni-luebeck.de} \par}
 

### 2009-10-17 doc/tech_doc/janitor/tex_usage/usage.tex
Janitor documetation fixes: build fixed, tex fixed, image recovered
Round of documentation updates - mostly semantics-invariant linguistic changes.

--- 2009-05-14 doc/manuals/janitor/tex_usage/usage.tex
@@
 
-Janitor can be used either with or without A-REX. In case A-REX is used, installations will be maintained automatically the runtime environments.
+The Janitor can be used either with or without an integration with \AREX. The Janitor's installation will not be affected
+by this decision pro or cons and integration with \AREX. Both can be installed in parallel. If \AREX is not allowed to
+install runtime environments upon demand, such automated installations can still be invoked manually via the Janitor's
+command line  interface.
 
@@
 found at \href{www.nordugrid.org/documents/xrsl.pdf}{www.nordugrid.org/documents/xrsl.pdf}~\cite{NORDUGRID_MANUAL_4}.
-Within Listing~\ref{lst:jsdl_job} an example using JSDL is provided. The specification of assigning runtime environments in JSDL is 
-currently only defined within the nordugrid jsdl-arc schema~
+Within Listing~\ref{lst:jsdl_job} an example using JSDL is provided. The specification of how
+to assign runtime environments in JSDL is currently only defined within the nordugrid jsdl-arc schema~
 \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/a-rex/grid-manager/jobdesc/jsdl/jsdl_arc.xsd}
@@
 
-In addition of using Janitor together with A-REX, Janitor can also be used as a standalone commandline tool.
-The available commands are listed in the Table~\ref{janitor_commandline_man}.
+On Linux systems, the Janitor's standalone commandline tool is available as /usr/lib/arc/janitor.
+Some Linux distributions may prefer /usr/libexec or similar paths.
+The script is only functional as root\footnote{Should you find that constraint unbearable for your purpose,
+please investigate the file rjanitor.cc in the ARC source tree. It wraps the janitor application
+and as a C binary can be configured to attract root privileges.}.
+To find that binary directly, you may decide to add that location to your \$PATH environment variable.
+
+The available commands to the Janitor, implemented as options to the janitor script, are listed
+in the Table~\ref{tab:janitor_commandline_man}.
 \begin{table}[!h]
    \begin{center}
-        \mycaption{Overview about the available commands in Janitor.}{}
+        \mycaption{Overview about the available commands to the Janitor.}{}
         \label{tab:janitor_commandline_man}
@@
 	&	remove				& Removes the placeholder of the job on the runtime environments.
-If no more jobs are using the runtime environment and the lifespan of the runtime environment has be expired, the runtime environment can be removed using the \texttt{sweep} command.
-Requires the [JOB-ID] to be removed.\\
+If there is no longer any job using the runtime environment and the lifespan of the runtime environment has be expired, the runtime environment can be removed using the \texttt{sweep} command.
+To indicate that a job has terminated, send the Janitor the remove command on that job's ID.\\
 	&					&\\
@@
 	\multicolumn{3}{l}{\textbf{Job id:}}\\
-	&					& A unique sequence of numbers. Once Janitor registered a job id it cannot register a second job having the same job id.\\
+	&					& A unique sequence of numbers. Once the Janitor registered a job ID, it cannot register a second job having the same job id.\\
 	\multicolumn{3}{l}{\textbf{Runtime environments:}}\\
@@
 The most important commands for Janitor are \texttt{register}, \texttt{deploy} and \texttt{remove}. In order to 
-register a job along with a set of runtime environments in Janitor, the first command \texttt{register} followed
+register a job along with a set of runtime environments in the Janitor, the first command \texttt{register} followed
 by a job identifier and a list of runtime environments has to be used.
@@
 within the catalog (resp. the runtime directory of the grid-manager).
+
 The command \texttt{deploy} extracts the necessary dependcies of the desired runtime environments and then downloads and installs 
@@
 % PROCEED HERE!!!!!!!! !! ! ! !
-In order to remove jobs registered in Janitor the command \texttt{remove} has to be used. 
+The disassociation of a job with a runtime environment is indicated as a "removal" of that job from the runtime environment.
+This is performed with the Janitor the command \texttt{remove}. 
 The command only removes the job entry and the lock on the runtime environment. If there are no more locks on the runtime environment
@@
 
-Easy commandline examples are provided in Listing~\ref{lst:janitor_example}.
+Easy commandline examples are provided in Listing~\ref{lst:janitor_example}. You may also want to inspect the janitor(8) man page.
 
-Each command has a certain behaviour for its exit status. The Table~\ref{tab:janitor_commandline_exit_status} lists the possible
-outcomes.
+Every command has a certain behaviour for its exit status. The Table~\ref{tab:janitor_commandline_exit_status} lists the possible
+outcomes. A value of 0 always indicates that no error occurred.
 
@@
    \begin{center}
-        \mycaption{Possible exit states of Janitor}{}
+        \mycaption{Possible exit states of the janitor application}{}
         \label{tab:janitor_commandline_exit_status}
@@
 
+Once a dynamic runtime environment is installed, it looks completely
+indistinguishable from traditionally installed runtime environments. This
+also means that the general concept to have one installation performed
+for all compute nodes in the network is kept.
+
 % 
@@
 
-\section{Example}
+%\section{Example}
 % 
 %
+
+\section{Janitor with \AREX}
+
+The motivation to have a runtime environment available comes from the submitters
+of the grid jobs that depends on that runtime environment for their execution.
+The site administrator's sole responsibility is to have the dynamic runtime
+environment at the site's disposal. No more. With \AREX allowed to initiated
+the commands to the Janitor, no further interaction from the site administrator
+is required. An exception may be to confirm the consistency of the system when
+the machine has crahsed and the Janitor may still find jobs assinged to runtime
+environments that are no longer running.
+
+Another exception for an active involvement of the site administrator is the
+initial configuration of the Janitor and the updating of runtime environments
+that are eligible to be installed.

### 2009-10-19 doc/tech_doc/client/job_description_mapping.tex
Updating job description mapping document.

@@
 \usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
-\usepackage{color}                                 %for defining custom colors
+%~ \usepackage{color}                                 %for defining custom colors
 \usepackage{framed}                                %for shaded and framed paragraphs
@@
 \usepackage{geometry}                              %for defining page size
-\usepackage{longtable}                             %for breaking tables
 \usepackage{array}
 \usepackage{booktabs}
-\usepackage{tabularx}
+\usepackage{ltablex}
 \usepackage[english]{babel}
 \usepackage{multirow}
+\usepackage[table]{xcolor} % Note: The rowcolors command does not work properly with the tabularx environment.
+
+\colorlet{tableheadcolor}{gray!25}
+\colorlet{tablerowcolor}{gray!12.5}
+
+\newcolumntype{C}{>{\ttfamily\footnotesize}c}
+\newcolumntype{T}{>{\ttfamily\footnotesize\centering\arraybackslash}X}
+
+\newcommand{\TODO}{\normalsize{\textnormal{TODO}}}
+\newcommand{\NM}{\normalsize{\textnormal{Not mapped}}}
+\newcommand{\NST}[1]{\normalsize{\textnormal{#1}}}
+\newcommand{\defineheadfoot}{\endfirsthead %
+\midrule %
+\multicolumn{3}{c}{Continued from previous page}\\%
+\midrule %
+\endhead %
+\midrule %
+\multicolumn{3}{c}{Continues at next page}\\%
+\midrule %
+\endfoot %
+\bottomrule %
+\endlastfoot}
+
+\renewcommand{\tabularxcolumn}[1]{m{#1}}
 
@@
 \section{XRSL mapping}
-\begin{table}
 \begin{center}
-\begin{tabular}{l|c|c}
-XRSL input  & Internal representaion & XRSL output\\\hline
+\rowcolors{3}{tablerowcolor}{}
+\begin{tabularx}{400pt}{CTC}
+\toprule
+\NST{XRSL input} & \NST{Internal representaion} & \NST{XRSL output}\\
+\midrule
+\defineheadfoot
 executable & Application.Executable.Name & executable\\
 arguments & Application.Executable.Arguments & arguments\\
-inputFiles & DataStaging.File.Name DataStaging.File.Source & inputFiles\\
+inputFiles& DataStaging.File.Name,\linebreak DataStaging.File.Source & inputFiles\\
 executables & DataStaging.File.IsExecutable & executables\\
-cache & TODO & TODO\\
-outputFiles & DataStaging.File.Name DataStaging.File.Target & outputFiles\\
+cache & \TODO & \TODO \\
+outputFiles & DataStaging.File.Name,\linebreak DataStaging.File.Target & outputFiles \\
 cpuTime & Resources.TotalCPUTime & cpuTime\\
 wallTime & Resources.TotalWallTime & wallTime\\
-gridTime & Not mapped & Not mapped\\
-benchmarks & Not mapped & Not mapped\\
+gridTime & Resources.TotalCPUTime & \NM \\
+benchmarks & Resources.TotalCPUTime & \NM \\
 memory & Resources.IndividualPhysicalMemory & memory\\
@@
 stdin & Application.Input & stdin\\
-stdout & Application.Output & stdout outputFiles\\
-stderr & Application.Error & stderr outputFiles\\
-join & Application.Join & Not mapped\\
+stdout & Application.Output & stdout, outputFiles\\
+stderr & Application.Error & stderr, outputFiles\\
+join & Application.Join & \NM\\
 gmlog & Application.LogDir & gmlog\\
 jobName & Identification.JobName & jobName\\
-ftpThreads & DataStaging.File.Source.Threads DataStaging.File.Target.Threads & Not mapped\\
+ftpThreads & DataStaging.File.Source.Threads,\linebreak DataStaging.File.Target.Threads & \NM\\
 acl & Application.AccessControl & acl\\
-cluster & Resources.CandidateTarget.EndPointURL & Not mapped\\
+cluster & Resources.CandidateTarget.EndPointURL & \NM\\
 queue & Resources.CandidateTarget.QueueName & queue\\
-startTime & Application.ProcessingStartTime & Not mapped\\
+startTime & Application.ProcessingStartTime & \NM\\
 lifeTime & Resources.SessionLifeTime & lifeTime\\
 notify & Application.Notification & notify\\
-replicaCollection & DataStaging.File.DataIndexingService & Not mapped\\
+replicaCollection & DataStaging.File.DataIndexingService & \NM\\
 rerun & Application.Rerun & rerun\\
 architecture & Resources.Platform & architecture\\
-nodeAccess & Resources.NodeAccess & Not mapped\\
-dryRun & Not mapped & Not mapped\\
-rsl\_substitution & Processed internally by the parser & Not mapped\\
+nodeAccess & Resources.NodeAccess & \NM\\
+dryRun & \NM & \NM\\
+rsl\_substitution & \NST{Processed internally by the parser} & \NM\\
 environment & Application.Environment & environment\\
@@
 credentialserver & Application.CredentialService & credentialserver\\
-\end{tabular}
+\end{tabularx}
 \end{center}
-\end{table}
 
 \section{JDL mapping}
-\begin{table}
 \begin{center}
-\begin{tabular}{l|c|c}
-JDL input  & Internal representaion & JDL output\\\hline
-JobType & Not mapped & Not mapped\\
-Executable & Application.Executable.Name & Executable InputSandbox\\
+\rowcolors{3}{}{tablerowcolor}
+\begin{tabularx}{\textwidth}{CTT}
+\toprule
+\rowcolor{white}
+\NST{JDL input} & \NST{Internal representaion} & \NST{JDL output}\\
+\midrule
+\defineheadfoot
+JobType & \NM & \NM\\
+Executable & Application.Executable.Name & Executable, InputSandbox\\
 Arguments & Application.Executable.Arguments & Arguments\\
-StdInput & Application.Input & StdInput InputSandbox\\
-StdOutput & Application.Output & StdOutput OutputSandbox OutputSandboxDestURI\\
-StdError & Application.Error & StdError OutputSandbox OutputSandboxDestURI\\
-InputSandbox & DataStaging.File.Name DataStaging.File.Source & InputSandbox\\
+StdInput & Application.Input & StdInput,\linebreak InputSandbox\\
+StdOutput & Application.Output & StdOutput,\linebreak OutputSandbox,\linebreak OutputSandboxDestURI\\
+StdError & Application.Error & StdError,\linebreak OutputSandbox,\linebreak OutputSandboxDestURI\\
+InputSandbox & DataStaging.File.Name,\linebreak DataStaging.File.Source & InputSandbox\\
 InputSandboxBaseURI & DataStaging.File.Source.URI & InputSandbox\\
-OutputSandbox & DataStaging.File.Name DataStaging.File.Target.URI & OutputSandbox OutputSandboxDestURI\\
-OutputSandboxDestURI & DataStaging.File.Target DataStaging.File.KeepData & OutputSandbox OutputSandboxDestURI\\
-OutputSandboxBaseDestURI & DataStaging.File.Target.URI & OutputSandbox OutputSandboxDestURI\\
+OutputSandbox & DataStaging.File.Name,\linebreak DataStaging.File.Target.URI & OutputSandbox,\linebreak OutputSandboxDestURI\\
+OutputSandboxDestURI & DataStaging.File.Target,\linebreak DataStaging.File.KeepData & OutputSandbox,\linebreak OutputSandboxDestURI\\
+OutputSandboxBaseDestURI & DataStaging.File.Target.URI & OutputSandbox,\linebreak OutputSandboxDestURI\\
 Prologue & Application.Prologue.Name & Prologue\\
@@
 Environment & Application.Environment & Environment\\
-PerusalFileEnable & Stored internally, not processed & PerusalFileEnable\\
-PerusalTimeInterval & Stored internally, not processed & PerusalTimeInterval\\
-PerusalFilesDestURI & Stored internally, not processed & PerusalFilesDestURI\\
-InputData (Deprecated) & Not mapped & Not mapped\\
-OutputData (Deprecated) & Not mapped & Not mapped\\
-StorageIndex (Deprecated) & Not mapped & Not mapped\\
-DataCatalog (Deprecated) & Not mapped & Not mapped\\
-DataRequirements & Stored internally, not processed & DataRequirements\\
-DataAccessProtocol & Stored internally, not processed & DataAccessProtocol\\
-OutputSE & Stored internally, not processed & OutputSE\\
+PerusalFileEnable & \NST{Stored internally, not processed} & PerusalFileEnable\\
+PerusalTimeInterval & \NST{Stored internally, not processed} & PerusalTimeInterval\\
+PerusalFilesDestURI & \NST{Stored internally, not processed} & PerusalFilesDestURI\\
+InputData \NST{(Deprecated)} & \NM & \NM\\
+OutputData \NST{(Deprecated)} & \NM & \NM\\
+StorageIndex \NST{(Deprecated)} & \NM & \NM\\
+DataCatalog \NST{(Deprecated)} & \NM & \NM\\
+DataRequirements & \NST{Stored internally, not processed} & DataRequirements\\
+DataAccessProtocol & \NST{Stored internally, not processed} & DataAccessProtocol\\
+OutputSE & \NST{Stored internally, not processed} & OutputSE\\
 VirtualOrganisation & Identification.JobVOName & VirtualOrganisation\\
@@
 ShallowRetryCount & Application.Rerun & RetryCount ShallowRetryCount\\
-LBAddress & Stored internally, not processed & LBAddress\\
+LBAddress & \NST{Stored internally, not processed} & LBAddress\\
 MyProxyServer & Application.CredentialService & MyProxyServer\\
-HLRLocation & Stored internally, not processed & HLRLocation\\
-JobProvenance & Stored internally, not processed & JobProvenance\\
-NodeNumber & Stored internally, not processed & NodeNumber\\
-JobSteps (Deprecated) & Not mapped & Not mapped\\
-CurrentStep (Deprecated) & Not mapped & Not mapped\\
-JobState (Deprecated) & Not mapped & Not mapped\\
-ListenerPort & Stored internally, not processed & ListenerPort\\
-ListenerHost & Stored internally, not processed & ListenerHost\\
-ListenerPipeName & Stored internally, not processed & ListenerPipeName\\
-Requirements & Stored internally, not processed & Requirements\\
+HLRLocation & \NST{Stored internally, not processed} & HLRLocation\\
+JobProvenance & \NST{Stored internally, not processed} & JobProvenance\\
+NodeNumber & \NST{Stored internally, not processed} & NodeNumber\\
+JobSteps \NST{(Deprecated)} & \NM & \NM\\
+CurrentStep \NST{(Deprecated)} & \NM & \NM\\
+JobState \NST{(Deprecated)} & \NM & \NM\\
+ListenerPort & \NST{Stored internally, not processed} & ListenerPort\\
+ListenerHost & \NST{Stored internally, not processed} & ListenerHost\\
+ListenerPipeName & \NST{Stored internally, not processed} & ListenerPipeName\\
+Requirements & \NST{Stored internally, not processed} & Requirements\\
 Rank & JobMeta.Rank & Rank\\
@@
 UserTags & Identification.UserTag & UserTags\\
-BatchSystem & Stored internally, not processed & BatchSystem\\
-QueueName & Resources.CandidateTarget.QueueName & QueueName\\
-\end{tabular}
+BatchSystem & \NST{Stored internally, not processed} & BatchSystem\\
+QueueName & Resources.CandidateTarget.QueueName & QueueName
+\end{tabularx}
 \end{center}
-\end{table}
 
 \section{JSDL mapping}
-\begin{table}
 \begin{center}
-\begin{tabular}{l|c|c}
-JSDL input  & Internal representaion & JSDL output\\\hline
-\bfseries JobIdentification & &\\
+\rowcolors{3}{}{tablerowcolor}
+\begin{tabularx}{400pt}{TCC}
+\toprule
+\NST{JSDL input} & \NST{Internal representaion} & \NST{JSDL output}\\
+\midrule
+\defineheadfoot
+\multicolumn{3}{c}{\bfseries JobIdentification}\\
+\midrule
 JobName & &\\
@@
 JobProject & &\\
-\bfseries Application & &\\
+\midrule
+\multicolumn{3}{c}{\bfseries Application}\\
+\midrule
 ApplicationName & &\\
 ApplicationVersion & &\\
-\bfseries Resources & &\\
+\midrule
+\multicolumn{3}{c}{\bfseries Resources}\\
+\midrule
 CandidateHosts/HostName & &\\
@@
 ExclusiveExecution & &\\
-OperatingSystem/OperatingSystemType/OperatingSystemName & &\\
-OperatingSystem/OperatingSystemVersion & &\\
-CPUArchitecture/CPUArchitectureName & &\\
+OperatingSystem/\linebreak OperatingSystemType/\linebreak OperatingSystemName & &\\
+OperatingSystem/\linebreak OperatingSystemVersion & &\\
+CPUArchitecture/\linebreak CPUArchitectureName & &\\
 IndividualCPUSpeed & &\\
@@
 TotalResourceCount & &\\
-\bfseries DataStaging & {}-{}- & {}-{}-\\
+\midrule
+\multicolumn{3}{c}{\bfseries DataStaging}\\
+\midrule
 FileName & &\\
@@
 Target/URI & &\\
-\end{tabular}
+\end{tabularx}
 \end{center}
-\end{table}
 
 \section{POSIX-JSDL mapping}
-\begin{table}
 \begin{center}
-\begin{tabular}{l|c|c}
-POSIX-JSDL input  & Internal representation & POSIX-JSDL output\\\hline
+\rowcolors{3}{}{tablerowcolor}
+\begin{tabularx}{400pt}{CCC}
+\toprule
+\NST{POSIX-JSDL input} & \NST{Internal representation} & \NST{POSIX-JSDL output}\\
+\midrule
+\defineheadfoot
 Executable & &\\
@@
 GroupName & &\\
-\end{tabular}
+\end{tabularx}
 \end{center}
-\end{table}
 
 \section{HPCProfile-JSDL mapping}
-\begin{table}
 \begin{center}
-\begin{tabular}{l|c|c}
-HPCProfile-JSDL input  & Internal representaion & HPCProfile-JSDL output\\\hline
+\rowcolors{3}{tablerowcolor}{}
+\begin{tabularx}{400pt}{CCC}
+\toprule
+\NST{HPCProfile-JSDL input}  & \NST{Internal representaion} & \NST{HPCProfile-JSDL output}\\
+\midrule
+\defineheadfoot
 Executable & & \\
@@
 UserName & &\\
-\end{tabular}
+\end{tabularx}
 \end{center}
-\end{table}
 \end{document}

### 2009-10-20 doc/tech_doc/a-rex/arex_tech_doc.tex
fix latex problems in xml schemas

@@
 \begin{verbatim}
-    <!-{}- A-Rex service -{}->
-    <Service name=\char`\"{}a-rex\char`\"{} id=\char`\"{}a-rex\char`\"{}>
-      <!-{}- Optional endpoint element is advised in case of multiple IP adresses -{}->
+    <!-- A-Rex service -->
+    <Service name="a-rex" id="a-rex">
+      <!-- Optional endpoint element is advised in case of multiple IP adresses -->
       <arex:endpoint>https://localhost:60000/arex</arex:endpoint>
-      <!-{}- Use information generated by identity.map plugin or default provided below -{}->
+      <!-- Use information generated by identity.map plugin or default provided below -->
       <arex:usermap><arex:defaultLocalName>nobody</arex:defaultLocalName></arex:usermap>
-      <!-{}- grid-manager part of a-rex requires legacy configuration file.
-           Use arc\_arex.conf example or write own. -{}->
-      <arex:gmconfig>/etc/arc\_arex.conf</arex:gmconfig>
+      <!-- grid-manager part of a-rex requires legacy configuration file.
+           Use arc.conf example or write own. -->
+      <arex:gmconfig>/etc/arc.conf</arex:gmconfig>
     </Service>
@@
 \begin{verbatim}
-<?xml version=\char`\"{}1.0\char`\"{} encoding=\char`\"{}UTF-8\char`\"{}?>
+<?xml version="1.0" encoding="UTF-8"?>
 <xsd:schema
-xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-xmlns=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
-xmlns:arc=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
-targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/ArcConfig/2007/arex\char`\"{}
-elementFormDefault=\char`\"{}qualified\char`\"{}>
-  <xsd:complexType name=\char`\"{}endpoint\_Type\char`\"{}>
-    <!-{}-
+xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007/arex"
+xmlns:arc="http://www.nordugrid.org/schemas/ArcConfig/2007/arex"
+targetNamespace="http://www.nordugrid.org/schemas/ArcConfig/2007/arex"
+elementFormDefault="qualified">
+  <xsd:complexType name="endpoint_Type">
+    <!--
     This element defines URL of A-REX service as seen from outside.
-    -{}->
+    -->
     <xsd:simpleContent>
-      <xsd:extension base=\char`\"{}xsd:string\char`\"{}> 
+      <xsd:extension base="xsd:string"> 
       </xsd:extension>
@@
   </xsd:complexType>
-  <xsd:element name=\char`\"{}endpoint\char`\"{} type=\char`\"{}endpoint\_Type\char`\"{}/>
-  <xsd:complexType name=\char`\"{}gmconfig\_Type\char`\"{}>
-    <!-{}-
+  <xsd:element name="endpoint" type="endpoint_Type"/>
+  <xsd:complexType name="gmconfig_Type">
+    <!--
     This element defines path to arc0 Grid Manager configuartion file.
     By default it is /etc/arc.conf.
-    -{}->
+    -->
     <xsd:simpleContent>
-      <xsd:extension base=\char`\"{}xsd:string\char`\"{}>
+      <xsd:extension base="xsd:string">
       </xsd:extension>
@@
   </xsd:complexType>
-  <xsd:element name=\char`\"{}gmconfig\char`\"{} type=\char`\"{}gmconfig\_Type\char`\"{}/>
-  <xsd:simpleType name=\char`\"{}gmrun\_Type\char`\"{}>
-    <!-{}-
+  <xsd:element name="gmconfig" type="gmconfig_Type"/>
+  <xsd:simpleType name="gmrun_Type">
+    <!--
     This element defines how grid-manager part of A-Rex is run.
@@
     Default is 'internal'.
-    -{}->
-    <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
-      <xsd:enumeration value=\char`\"{}internal\char`\"{}/>
-      <xsd:enumeration value=\char`\"{}external\char`\"{}/>
-      <xsd:enumeration value=\char`\"{}none\char`\"{}/>
+    -->
+    <xsd:restriction base="xsd:string">
+      <xsd:enumeration value="internal"/>
+      <xsd:enumeration value="external"/>
+      <xsd:enumeration value="none"/>
     </xsd:restriction>
   </xsd:simpleType>
-  <xsd:element name=\char`\"{}gmrun\char`\"{} type=\char`\"{}gmrun\_Type\char`\"{}/>
-  <xsd:complexType name=\char`\"{}usermap\_Type\char`\"{}>
-    <xsd:sequence>
-      <xsd:element name=\char`\"{}defaultLocalName\char`\"{} type=\char`\"{}xsd:string\char`\"{}
-                   minOccurs=\char`\"{}0\char`\"{}/>
-    </xsd:sequence>
-  </xsd:complexType>
-  <xsd:element name=\char`\"{}usermap\char`\"{} type=\char`\"{}usermap\_Type\char`\"{}/>
-  <!-{}- CommonName attribute of bes-factory. -{}->
-  <xsd:element name=\char`\"{}commonName\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-  <!-{}- LongDescription attribute of bes-factory. -{}->
-  <xsd:element name=\char`\"{}longDescription\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-  <!-{}- Name of Local Resource Management System. -{}->
-  <xsd:element name=\char`\"{}LRMSName\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-  <!-{}-
+  <xsd:element name="gmrun" type="gmrun_Type"/>
+  <xsd:complexType name="usermap_Type">
+    <xsd:sequence>
+      <xsd:element name="defaultLocalName" type="xsd:string"
+                   minOccurs="0"/>
+    </xsd:sequence>
+  </xsd:complexType>
+  <xsd:element name="usermap" type="usermap_Type"/>
+  <!-- CommonName attribute of bes-factory. -->
+  <xsd:element name="commonName" type="xsd:string"/>
+  <!-- LongDescription attribute of bes-factory. -->
+  <xsd:element name="longDescription" type="xsd:string"/>
+  <!-- Name of Local Resource Management System. -->
+  <xsd:element name="LRMSName" type="xsd:string"/>
+  <!--
   Name of Operating System.
-  The values are based on the OSType field of the CIM\_OperatingSystem model:
-  http://www.dmtf.org/standards/cim/cim\_schema\_v29
-  -{}->
-  <xsd:element name=\char`\"{}OperatingSystem\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
+  The values are based on the OSType field of the CIM_OperatingSystem model:
+  http://www.dmtf.org/standards/cim/cim_schema_v29
+  -->
+  <xsd:element name="OperatingSystem" type="xsd:string"/>
 </xsd:schema>
@@
 \begin{footnotesize}\begin{verbatim}
-<?xml version=\char`\"{}1.0\char`\"{} encoding=\char`\"{}UTF-8\char`\"{}?>
-<wsdl:definitions targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}
- xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
- xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
- xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
- xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
- xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
- xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
- xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
- xmlns:bes-factory=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}
- xmlns:bes-mgmt=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}
- xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
- xmlns:wsrf-rpw=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}
- xmlns:a-rex=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
-  <wsdl:import namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory\char`\"{}
-               location=\char`\"{}./bes-factory.wsdl\char`\"{}/>
-  <wsdl:import namespase=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-management\char`\"{}
-               location=\char`\"{}./bes-management.wsdl\char`\"{}/>
-  <wsdl:import namespase=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
-               location=\char`\"{}../schemas/delegation.wsdl\char`\"{}/>
-  <wsdl:import namespase=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2\char`\"{}
-               location=\char`\"{}http://docs.oasis-open.org/wsrf/rpw-2.wsdl\char`\"{}/>
+<?xml version="1.0" encoding="UTF-8"?>
+<wsdl:definitions targetNamespace="http://www.nordugrid.org/schemas/a-rex"
+ xmlns:SOAP-ENV="http://schemas.xmlsoap.org/soap/envelope/"
+ xmlns:SOAP-ENC="http://schemas.xmlsoap.org/soap/encoding/"
+ xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"
+ xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+ xmlns:soap="http://schemas.xmlsoap.org/wsdl/soap/"
+ xmlns:wsdl="http://schemas.xmlsoap.org/wsdl/"
+ xmlns:wsa="http://www.w3.org/2005/08/addressing"
+ xmlns:bes-factory="http://schemas.ggf.org/bes/2006/08/bes-factory"
+ xmlns:bes-mgmt="http://schemas.ggf.org/bes/2006/08/bes-management"
+ xmlns:deleg="http://www.nordugrid.org/schemas/delegation"
+ xmlns:wsrf-rpw="http://docs.oasis-open.org/wsrf/rpw-2"
+ xmlns:a-rex="http://www.nordugrid.org/schemas/a-rex">
+  <wsdl:import namespase="http://schemas.ggf.org/bes/2006/08/bes-factory"
+               location="./bes-factory.wsdl"/>
+  <wsdl:import namespase="http://schemas.ggf.org/bes/2006/08/bes-management"
+               location="./bes-management.wsdl"/>
+  <wsdl:import namespase="http://www.nordugrid.org/schemas/delegation"
+               location="../schemas/delegation.wsdl"/>
+  <wsdl:import namespase="http://docs.oasis-open.org/wsrf/rpw-2"
+               location="http://docs.oasis-open.org/wsrf/rpw-2.wsdl"/>
   <wsdl:types>
-    <xsd:schema targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/a-rex\char`\"{}>
-      <xsd:import namespace=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
-                  schemaLocation=\char`\"{}./ws-addr.xsd\char`\"{}/>
-      <xsd:simpleType name=\char`\"{}ActivitySubStateType\char`\"{}>
-        <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
-          <xsd:enumeration value=\char`\"{}Accepting\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Accepted\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Preparing\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Prepared\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Submitting\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Executing\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Killing\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Executed\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Finishing\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Finished\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Failed\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Deleted\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Pending\char`\"{}/>
-          <xsd:enumeration value=\char`\"{}Held\char`\"{}/>
+    <xsd:schema targetNamespace="http://www.nordugrid.org/schemas/a-rex">
+      <xsd:import namespace="http://www.w3.org/2005/08/addressing"
+                  schemaLocation="./ws-addr.xsd"/>
+      <xsd:simpleType name="ActivitySubStateType">
+        <xsd:restriction base="xsd:string">
+          <xsd:enumeration value="Accepting"/>
+          <xsd:enumeration value="Accepted"/>
+          <xsd:enumeration value="Preparing"/>
+          <xsd:enumeration value="Prepared"/>
+          <xsd:enumeration value="Submitting"/>
+          <xsd:enumeration value="Executing"/>
+          <xsd:enumeration value="Killing"/>
+          <xsd:enumeration value="Executed"/>
+          <xsd:enumeration value="Finishing"/>
+          <xsd:enumeration value="Finished"/>
+          <xsd:enumeration value="Failed"/>
+          <xsd:enumeration value="Deleted"/>
+          <xsd:enumeration value="Pending"/>
+          <xsd:enumeration value="Held"/>
         </xsd:restriction>
       </xsd:simpleType>
-      <xsd:element name=\char`\"{}State\char`\"{} type=\char`\"{}a-rex:ActivitySubStateType\char`\"{}/>
-      <xsd:complexType name=\char`\"{}ResourceInformationDocumentType\char`\"{}>
-        <xsd:sequence>
-           <xsd:element name=\char`\"{}BESFactory\char`\"{}
-                        type=\char`\"{}bes-factory:FactoryResourceAttributesDocumentType\char`\"{}/>
-          <xsd:complexType name=\char`\"{}Glue2Resource\char`\"{} minOccurs='0'>
-            <xsd:sequence>
-              <xsd:any namespace=\char`\"{}\#\#other\char`\"{} processContents=\char`\"{}lax\char`\"{}
-                       minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>
-            </xsd:sequence>
-          </xsd:complexType>
-          <xsd:complexType name=\char`\"{}Activities\char`\"{} minOccurs='0'>
-            <xsd:sequence>
-              <xsd:complexType name=\char`\"{}Activity\char`\"{} minOccurs='0' maxOccurs='unbounded'>
-                <xsd:sequence>
-                  <xsd:element name=\char`\"{}ActivityIdentifier\char`\"{}
-                               type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
-                  <xsd:element ref=\char`\"{}bes-factory:ActivityDocument\char`\"{} minOccurs='0'/>
-                  <xsd:complexType name=\char`\"{}Glue2Job\char`\"{} minOccurs='0'>
-                    <xsd:sequence>
-                      <xsd:any namespace=\char`\"{}\#\#other\char`\"{} processContents=\char`\"{}lax\char`\"{}
-                               minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>
-                    </xsd:sequence>
-                  </xsd:complexType>
-                </xsd:sequence>
-              </xsd:complexType>
-            </xsd:sequence>
-          </xsd:complexType>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:complexType name=\char`\"{}ChangeActivityStatusRequestType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}ActivityIdentifier\char`\"{} type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
-          <xsd:element name=\char`\"{}OldStatus\char`\"{} type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}
-                       minOccurs=\char`\"{}0\char`\"{}/>
-          <xsd:element name=\char`\"{}NewStatus\char`\"{} type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}ChangeActivityStatus\char`\"{}
-                   type=\char`\"{}a-rex:ChangeActivityStatusRequestType\char`\"{}/>
-      <xsd:complexType name=\char`\"{}ChangeActivityStatusResponseType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}NewStatus\char`\"{} type=\char`\"{}bes-factory:ActivityStatusType\char`\"{}/>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
-                   type=\char`\"{}a-rex:ChangeActivityStatusResponseType\char`\"{}/>
+      <xsd:element name="State" type="a-rex:ActivitySubStateType"/>
+      <xsd:complexType name="ResourceInformationDocumentType">
+        <xsd:sequence>
+           <xsd:element name="BESFactory"
+                        type="bes-factory:FactoryResourceAttributesDocumentType"/>
+          <xsd:complexType name="Glue2Resource" minOccurs='0'>
+            <xsd:sequence>
+              <xsd:any namespace="##other" processContents="lax"
+                       minOccurs="0" maxOccurs="unbounded"/>
+            </xsd:sequence>
+          </xsd:complexType>
+          <xsd:complexType name="Activities" minOccurs='0'>
+            <xsd:sequence>
+              <xsd:complexType name="Activity" minOccurs='0' maxOccurs='unbounded'>
+                <xsd:sequence>
+                  <xsd:element name="ActivityIdentifier"
+                               type="wsa:EndpointReferenceType"/>
+                  <xsd:element ref="bes-factory:ActivityDocument" minOccurs='0'/>
+                  <xsd:complexType name="Glue2Job" minOccurs='0'>
+                    <xsd:sequence>
+                      <xsd:any namespace="##other" processContents="lax"
+                               minOccurs="0" maxOccurs="unbounded"/>
+                    </xsd:sequence>
+                  </xsd:complexType>
+                </xsd:sequence>
+              </xsd:complexType>
+            </xsd:sequence>
+          </xsd:complexType>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:complexType name="ChangeActivityStatusRequestType">
+        <xsd:sequence>
+          <xsd:element name="ActivityIdentifier" type="wsa:EndpointReferenceType"/>
+          <xsd:element name="OldStatus" type="bes-factory:ActivityStatusType"
+                       minOccurs="0"/>
+          <xsd:element name="NewStatus" type="bes-factory:ActivityStatusType"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="ChangeActivityStatus"
+                   type="a-rex:ChangeActivityStatusRequestType"/>
+      <xsd:complexType name="ChangeActivityStatusResponseType">
+        <xsd:sequence>
+          <xsd:element name="NewStatus" type="bes-factory:ActivityStatusType"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="ChangeActivityStatusResponse"
+                   type="a-rex:ChangeActivityStatusResponseType"/>
       <xsd:complexType name="MigrateActivityType">
         <xsd:sequence>
-	  <xsd:element name=\char`\"{}ActivityIdentifier\char`\"{} 
-	  	           type=\char`\"{}wsa:EndpointReferenceType\char`\"{} 
-	  	           minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}1\char`\"{}/>
-          <xsd:element ref=\char`\"{}bes-factory:ActivityDocument\char`\"{} minOccurs=\char`\"{}0\char`\"{}/>
-	  <xsd:element name=\char`\"{}ForceMigration\char`\"{} type=\char`\"{}xsd:boolean\char`\"{} minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}1\char`\"{}/>    
-          <xsd:any namespace=\char`\"{}##other\char`\"{} processContents=\char`\"{}lax\char`\"{} minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>        
-        </xsd:sequence>
-        <xsd:anyAttribute namespace=\char`\"{}##other\char`\"{} processContents=\char`\"{}lax\char`\"{}/>
-      </xsd:complexType>
-
-      <xsd:complexType name=\char`\"{}MigrateActivityResponseType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}ActivityIdentifier\char`\"{} type=\char`\"{}wsa:EndpointReferenceType\char`\"{}/>
-          <xsd:element ref=\char`\"{}bes-factory:ActivityDocument\char`\"{} minOccurs=\char`\"{}0\char`\"{}/>
-          <xsd:any namespace=\char`\"{}##other\char`\"{} processContents=\char`\"{}lax\char`\"{} minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>        
-        </xsd:sequence>
-        <xsd:anyAttribute namespace=\char`\"{}##other\char`\"{} processContents=\char`\"{}lax\char`\"{}/>
-      </xsd:complexType>
-
-      <xsd:element name=\char`\"{}MigrateActivity\char`\"{}
-        type=\char`\"{}a-rex:MigrateActivityType\char`\"{}/>
-      <xsd:element name=\char`\"{}MigrateActivityResponse\char`\"{}
-        type=\char`\"{}a-rex:MigrateActivityResponseType\char`\"{}/>
+	  <xsd:element name="ActivityIdentifier" 
+	  	           type="wsa:EndpointReferenceType" 
+	  	           minOccurs="1" maxOccurs="1"/>
+          <xsd:element ref="bes-factory:ActivityDocument" minOccurs="0"/>
+	  <xsd:element name="ForceMigration" type="xsd:boolean" minOccurs="0" maxOccurs="1"/>    
+          <xsd:any namespace="##other" processContents="lax" minOccurs="0" maxOccurs="unbounded"/>        
+        </xsd:sequence>
+        <xsd:anyAttribute namespace="##other" processContents="lax"/>
+      </xsd:complexType>
+
+      <xsd:complexType name="MigrateActivityResponseType">
+        <xsd:sequence>
+          <xsd:element name="ActivityIdentifier" type="wsa:EndpointReferenceType"/>
+          <xsd:element ref="bes-factory:ActivityDocument" minOccurs="0"/>
+          <xsd:any namespace="##other" processContents="lax" minOccurs="0" maxOccurs="unbounded"/>        
+        </xsd:sequence>
+        <xsd:anyAttribute namespace="##other" processContents="lax"/>
+      </xsd:complexType>
+
+      <xsd:element name="MigrateActivity"
+        type="a-rex:MigrateActivityType"/>
+      <xsd:element name="MigrateActivityResponse"
+        type="a-rex:MigrateActivityResponseType"/>
 
@@
   </wsdl:types>
-  <wsdl:message name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
-    <wsdl:part name=\char`\"{}ChangeActivityStatusRequest\char`\"{}
-               element=\char`\"{}a-rex:ChangeActivityStatus\char`\"{}/>
+  <wsdl:message name="ChangeActivityStatusRequest">
+    <wsdl:part name="ChangeActivityStatusRequest"
+               element="a-rex:ChangeActivityStatus"/>
   </wsdl:message>
-  <wsdl:message name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
-    <wsdl:part name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
-               element=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
+  <wsdl:message name="ChangeActivityStatusResponse">
+    <wsdl:part name="ChangeActivityStatusResponse"
+               element="a-rex:ChangeActivityStatusResponse"/>
   </wsdl:message>
 
-  <wsdl:message name=\char`\"{}MigrateActivityRequest\char`\"{}>
-    <wsdl:part name=\char`\"{}MigrateActivityRequest\char`\"{} 
-	element=\char`\"{}a-rex:MigrateActivity\char`\"{}/>
+  <wsdl:message name="MigrateActivityRequest">
+    <wsdl:part name="MigrateActivityRequest" 
+	element="a-rex:MigrateActivity"/>
   </wsdl:message>
 
-  <wsdl:message name=\char`\"{}MigrateActivityResponse\char`\"{}>
-    <wsdl:part name=\char`\"{}MigrateActivityResponse\char`\"{} 
-        element=\char`\"{}a-rex:MigrateActivityResponse\char`\"{}/>
+  <wsdl:message name="MigrateActivityResponse">
+    <wsdl:part name="MigrateActivityResponse" 
+        element="a-rex:MigrateActivityResponse"/>
   </wsdl:message>
 
-  <wsdl:portType name=\char`\"{}a-rex\char`\"{}>
-    <wsdl:operation name=\char`\"{}ChangeActivityStatus\char`\"{}>
+  <wsdl:portType name="a-rex">
+    <wsdl:operation name="ChangeActivityStatus">
       <wsdl:documentation>
@@
       </wsdl:documentation>
-      <wsdl:input name=\char`\"{}ChangeActivityStatusRequest\char`\"{}
-        message=\char`\"{}a-rex:ChangeActivityStatusRequest\char`\"{}/>
-      <wsdl:output name=\char`\"{}ChangeActivityStatusResponse\char`\"{}
-        message=\char`\"{}a-rex:ChangeActivityStatusResponse\char`\"{}/>
-      <wsdl:fault name=\char`\"{}NotAuthorizedFault\char`\"{}
-        message=\char`\"{}bes-factory:NotAuthorizedFault\char`\"{}
-        wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}
-        message=\char`\"{}bes-factory:InvalidActivityIdentifierFault\char`\"{}
-        wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}
-         message=\char`\"{}bes-factory:CantApplyOperationToCurrentStateFault\char`\"{}
-         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}
-         message=\char`\"{}bes-factory:OperationWillBeAppliedEventuallyFault\char`\"{}
-         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+      <wsdl:input name="ChangeActivityStatusRequest"
+        message="a-rex:ChangeActivityStatusRequest"/>
+      <wsdl:output name="ChangeActivityStatusResponse"
+        message="a-rex:ChangeActivityStatusResponse"/>
+      <wsdl:fault name="NotAuthorizedFault"
+        message="bes-factory:NotAuthorizedFault"
+        wsa:Action="http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault"/>
+      <wsdl:fault name="InvalidActivityIdentifierFault"
+        message="bes-factory:InvalidActivityIdentifierFault"
+        wsa:Action="http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault"/>
+      <wsdl:fault name="CantApplyOperationToCurrentStateFault"
+         message="bes-factory:CantApplyOperationToCurrentStateFault"
+         wsa:Action="http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault"/>
+      <wsdl:fault name="OperationWillBeAppliedEventuallyFault"
+         message="bes-factory:OperationWillBeAppliedEventuallyFault"
+         wsa:Action="http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault"/>
     </wsdl:operation>
-    <wsdl:operation name=\char`\"{}MigrateActivity\char`\"{}>
+    <wsdl:operation name="MigrateActivity">
       <wsdl:input 
-         name=\char`\"{}MigrateActivity\char`\"{}
-         message=\char`\"{}a-rex:MigrateActivityRequest\char`\"{}  
+         name="MigrateActivity"
+         message="a-rex:MigrateActivityRequest"  
       <wsdl:output
-         name=\char`\"{}MigrateActivityResponse\char`\"{}
-         message=\char`\"{}a-rex:MigrateActivityResponse\char`\"{}
-      <wsdl:fault name=\char`\"{}NotAuthorizedFault\char`\"{} 
-         message=\char`\"{}bes-factory:NotAuthorizedFault\char`\"{}
-        wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}NotAcceptingNewActivitiesFault\char`\"{} 
-         message=\char`\"{}bes-factory:NotAcceptingNewActivitiesFault\char`\"{}
-         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}UnsupportedFeatureFault\char`\"{} 
-         message=\char`\"{}bes-factory:UnsupportedFeatureFault\char`\"{}
-         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}InvalidRequestMessageFault\char`\"{}
-         message=\char`\"{}bes-factory:InvalidRequestMessageFault\char`\"{}
-         wsa:Action=\char`\"{}http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault\char`\"{}/>
+         name="MigrateActivityResponse"
+         message="a-rex:MigrateActivityResponse"
+      <wsdl:fault name="NotAuthorizedFault" 
+         message="bes-factory:NotAuthorizedFault"
+        wsa:Action="http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault"/>
+      <wsdl:fault name="NotAcceptingNewActivitiesFault" 
+         message="bes-factory:NotAcceptingNewActivitiesFault"
+         wsa:Action="http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault"/>
+      <wsdl:fault name="UnsupportedFeatureFault" 
+         message="bes-factory:UnsupportedFeatureFault"
+         wsa:Action="http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault"/>
+      <wsdl:fault name="InvalidRequestMessageFault"
+         message="bes-factory:InvalidRequestMessageFault"
+         wsa:Action="http://schemas.ggf.org/bes/2006/08/bes-factory/BESFactoryPortType/Fault"/>
     </wsdl:operation>
   </wsdl:portType>
-  <wsdl:binding name=\char`\"{}a-rex\char`\"{} type=\char`\"{}a-rex:a-rex\char`\"{}>
-    <soap:binding style=\char`\"{}document\char`\"{}
-                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-    <wsdl:operation name=\char`\"{}ChangeActivityStatus\char`\"{}>
-      <soap:operation soapAction=\char`\"{}ChangeActivityStatus\char`\"{}/>
-      <wsdl:input name=\char`\"{}ChangeActivityStatusRequest\char`\"{}>
-         <soap:body use=\char`\"{}literal\char`\"{}/>
+  <wsdl:binding name="a-rex" type="a-rex:a-rex">
+    <soap:binding style="document"
+                  transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="ChangeActivityStatus">
+      <soap:operation soapAction="ChangeActivityStatus"/>
+      <wsdl:input name="ChangeActivityStatusRequest">
+         <soap:body use="literal"/>
       </wsdl:input>
-      <wsdl:output name=\char`\"{}ChangeActivityStatusResponse\char`\"{}>
-         <soap:body use=\char`\"{}literal\char`\"{}/>
+      <wsdl:output name="ChangeActivityStatusResponse">
+         <soap:body use="literal"/>
       </wsdl:output>
-      <wsdl:fault name=\char`\"{}NotAuthorizedFault\char`\"{}>
-        <soap:fault name=\char`\"{}NotAuthorizedFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="NotAuthorizedFault">
+        <soap:fault name="NotAuthorizedFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}InvalidActivityIdentifierFault\char`\"{}>
-        <soap:fault name=\char`\"{}InvalidActivityIdentifierFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="InvalidActivityIdentifierFault">
+        <soap:fault name="InvalidActivityIdentifierFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{}>
-        <soap:fault name=\char`\"{}CantApplyOperationToCurrentStateFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="CantApplyOperationToCurrentStateFault">
+        <soap:fault name="CantApplyOperationToCurrentStateFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{}>
-        <soap:fault name=\char`\"{}OperationWillBeAppliedEventuallyFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="OperationWillBeAppliedEventuallyFault">
+        <soap:fault name="OperationWillBeAppliedEventuallyFault" use="literal" />
       </wsdl:fault>
     </wsdl:operation>
-    <wsdl:operation name=\char`\"{}MigrateActivity\char`\"{}>
-      <soap:operation soapAction=\char`\"{}MigrateActivity\char`\"{} /> 
-      <wsdl:input name=\char`\"{}MigrateActivity\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{} />
+    <wsdl:operation name="MigrateActivity">
+      <soap:operation soapAction="MigrateActivity" /> 
+      <wsdl:input name="MigrateActivity">
+        <soap:body use="literal" />
       </wsdl:input>
-      <wsdl:output name=\char`\"{}MigrateActivityResponse\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{} />
+      <wsdl:output name="MigrateActivityResponse">
+        <soap:body use="literal" />
       </wsdl:output>
-      <wsdl:fault name=\char`\"{}NotAuthorizedFault\char`\"{}>
-        <soap:fault name=\char`\"{}NotAuthorizedFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="NotAuthorizedFault">
+        <soap:fault name="NotAuthorizedFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}NotAcceptingNewActivitiesFault\char`\"{}>
-        <soap:fault name=\char`\"{}NotAcceptingNewActivitiesFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="NotAcceptingNewActivitiesFault">
+        <soap:fault name="NotAcceptingNewActivitiesFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}UnsupportedFeatureFault\char`\"{}>
-        <soap:fault name=\char`\"{}UnsupportedFeatureFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="UnsupportedFeatureFault">
+        <soap:fault name="UnsupportedFeatureFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}InvalidRequestMessageFault\char`\"{}>
-        <soap:fault name=\char`\"{}InvalidRequestMessageFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="InvalidRequestMessageFault">
+        <soap:fault name="InvalidRequestMessageFault" use="literal" />
       </wsdl:fault>
@@
   </wsdl:binding>
-  <wsdl:binding name=\char`\"{}GetResourcePropertyDocument\char`\"{}
-                type=\char`\"{}wsrf-rpw:GetResourcePropertyDocument\char`\"{}>
-    <soap:binding style=\char`\"{}document\char`\"{}
-                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-    <wsdl:operation name=\char`\"{}GetResourcePropertyDocument\char`\"{}>
-      <soap:operation soapAction=\char`\"{}GetResourcePropertyDocument\char`\"{}/>
-      <wsdl:input name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentRequest\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{}/>
+  <wsdl:binding name="GetResourcePropertyDocument"
+                type="wsrf-rpw:GetResourcePropertyDocument">
+    <soap:binding style="document"
+                  transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="GetResourcePropertyDocument">
+      <soap:operation soapAction="GetResourcePropertyDocument"/>
+      <wsdl:input name="wsrf-rpw:GetResourcePropertyDocumentRequest">
+        <soap:body use="literal"/>
       </wsdl:input>
-      <wsdl:output name=\char`\"{}wsrf-rpw:GetResourcePropertyDocumentResponse\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{}/>
+      <wsdl:output name="wsrf-rpw:GetResourcePropertyDocumentResponse">
+        <soap:body use="literal"/>
       </wsdl:output>
-      <wsdl:fault name=\char`\"{}ResourceUnknownFault\char`\"{}>
-        <soap:fault name=\char`\"{}ResourceUnknownFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="ResourceUnknownFault">
+        <soap:fault name="ResourceUnknownFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-        <soap:fault name=\char`\"{}ResourceUnavailabbleFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="ResourceUnavailableFault">
+        <soap:fault name="ResourceUnavailabbleFault" use="literal" />
       </wsdl:fault>
@@
   </wsdl:binding>
-  <wsdl:binding name=\char`\"{}GetResourceProperty\char`\"{} type=\char`\"{}wsrf-rpw:GetResourceProperty\char`\"{}>
-    <soap:binding style=\char`\"{}document\char`\"{}
-                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-    <wsdl:operation name=\char`\"{}GetResourceProperty\char`\"{}>
-      <soap:operation soapAction=\char`\"{}GetResourceProperty\char`\"{}/>
-      <wsdl:input name=\char`\"{}wsrf-rpw:GetResourcePropertyRequest\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{}/>
+  <wsdl:binding name="GetResourceProperty" type="wsrf-rpw:GetResourceProperty">
+    <soap:binding style="document"
+                  transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="GetResourceProperty">
+      <soap:operation soapAction="GetResourceProperty"/>
+      <wsdl:input name="wsrf-rpw:GetResourcePropertyRequest">
+        <soap:body use="literal"/>
       </wsdl:input>
-      <wsdl:output name=\char`\"{}wsrf-rpw:GetResourcePropertyResponse\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{}/>
+      <wsdl:output name="wsrf-rpw:GetResourcePropertyResponse">
+        <soap:body use="literal"/>
       </wsdl:output>
-      <wsdl:fault name=\char`\"{}ResourceUnknownFault\char`\"{}>
-        <soap:fault name=\char`\"{}ResourceUnknownFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="ResourceUnknownFault">
+        <soap:fault name="ResourceUnknownFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-        <soap:fault name=\char`\"{}ResourceUnavailabbleFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="ResourceUnavailableFault">
+        <soap:fault name="ResourceUnavailabbleFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
-        <soap:fault name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="InvalidResourcePropertyQNameFault">
+        <soap:fault name="InvalidResourcePropertyQNameFault" use="literal" />
       </wsdl:fault>
@@
   </wsdl:binding>
-  <wsdl:binding name=\char`\"{}QueryResourceProperties\char`\"{} type=\char`\"{}wsrf:QueryResourceProperties\char`\"{}>
-    <soap:binding style=\char`\"{}document\char`\"{}
-                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-    <wsdl:operation name=\char`\"{}QueryResourceProperties\char`\"{}>
-      <soap:operation soapAction=\char`\"{}QueryResourceProperties\char`\"{}/>
-      <wsdl:input name=\char`\"{}wsrf-rpw:QueryResourcePropertiesRequest\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{}/>
+  <wsdl:binding name="QueryResourceProperties" type="wsrf:QueryResourceProperties">
+    <soap:binding style="document"
+                  transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="QueryResourceProperties">
+      <soap:operation soapAction="QueryResourceProperties"/>
+      <wsdl:input name="wsrf-rpw:QueryResourcePropertiesRequest">
+        <soap:body use="literal"/>
       </wsdl:input>
-      <wsdl:output name=\char`\"{}wsrf-rpw:QueryResourcePropertiesResponse\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{}/>
+      <wsdl:output name="wsrf-rpw:QueryResourcePropertiesResponse">
+        <soap:body use="literal"/>
       </wsdl:output>
-      <wsdl:fault name=\char`\"{}ResourceUnknownFault\char`\"{}>
-        <soap:fault name=\char`\"{}ResourceUnknownFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="ResourceUnknownFault">
+        <soap:fault name="ResourceUnknownFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}ResourceUnavailableFault\char`\"{}>
-        <soap:fault name=\char`\"{}ResourceUnavailabbleFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="ResourceUnavailableFault">
+        <soap:fault name="ResourceUnavailabbleFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{}>
-        <soap:fault name=\char`\"{}InvalidResourcePropertyQNameFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="InvalidResourcePropertyQNameFault">
+        <soap:fault name="InvalidResourcePropertyQNameFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{}>
-        <soap:fault name=\char`\"{}UnknownQueryExpressionDialectFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="UnknownQueryExpressionDialectFault">
+        <soap:fault name="UnknownQueryExpressionDialectFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}InvalidQueryExpressionFault\char`\"{}>
-        <soap:fault name=\char`\"{}InvalidQueryExpressionFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="InvalidQueryExpressionFault">
+        <soap:fault name="InvalidQueryExpressionFault" use="literal" />
       </wsdl:fault>
-      <wsdl:fault name=\char`\"{}QueryEvaluationErrorFault\char`\"{}>
-        <soap:fault name=\char`\"{}QueryEvaluationErrorFault\char`\"{} use=\char`\"{}literal\char`\"{} />
+      <wsdl:fault name="QueryEvaluationErrorFault">
+        <soap:fault name="QueryEvaluationErrorFault" use="literal" />
       </wsdl:fault>
@@
   </wsdl:binding>
-  <wsdl:service name=\char`\"{}a-rex\char`\"{}>
-    <wsdl:port name=\char`\"{}delegation\char`\"{} binding=\char`\"{}deleg:DelegationBinding\char`\"{}>
+  <wsdl:service name="a-rex">
+    <wsdl:port name="delegation" binding="deleg:DelegationBinding">
     </wsdl:port>
-    <wsdl:port name=\char`\"{}bes-factory\char`\"{} binding=\char`\"{}bes-factory:BESFactoryBinding\char`\"{}>
+    <wsdl:port name="bes-factory" binding="bes-factory:BESFactoryBinding">
     </wsdl:port>
-    <wsdl:port name=\char`\"{}bes-mgmt\char`\"{} binding=\char`\"{}bes-mgmt:BESManagementBinding\char`\"{}>
+    <wsdl:port name="bes-mgmt" binding="bes-mgmt:BESManagementBinding">
     </wsdl:port>
-    <wsdl:port name=\char`\"{}GetResourcePropertyDocument\char`\"{}
-               binding=\char`\"{}a-rex:GetResourcePropertyDocument\char`\"{}>
+    <wsdl:port name="GetResourcePropertyDocument"
+               binding="a-rex:GetResourcePropertyDocument">
     </wsdl:port>
-    <wsdl:port name=\char`\"{}GetResourceProperty\char`\"{} binding=\char`\"{}a-rex:GetResourceProperty\char`\"{}>
+    <wsdl:port name="GetResourceProperty" binding="a-rex:GetResourceProperty">
     </wsdl:port>
-    <wsdl:port name=\char`\"{}QueryResourceProperties\char`\"{}
-               binding=\char`\"{}a-rex:QueryResourceProperties\char`\"{}>
+    <wsdl:port name="QueryResourceProperties"
+               binding="a-rex:QueryResourceProperties">
     </wsdl:port>
-    <wsdl:port name=\char`\"{}a-rex\char`\"{} binding=\char`\"{}a-rex:a-rex\char`\"{}>
+    <wsdl:port name="a-rex" binding="a-rex:a-rex">
     </wsdl:port>
@@
 \begin{footnotesize}\begin{verbatim}
-<?xml version=\char`\"{}1.0\char`\"{} encoding=\char`\"{}UTF-8\char`\"{}?>
-<wsdl:definitions targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}
- xmlns:SOAP-ENV=\char`\"{}http://schemas.xmlsoap.org/soap/envelope/\char`\"{}
- xmlns:SOAP-ENC=\char`\"{}http://schemas.xmlsoap.org/soap/encoding/\char`\"{}
- xmlns:xsi=\char`\"{}http://www.w3.org/2001/XMLSchema-instance\char`\"{}
- xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
- xmlns:soap=\char`\"{}http://schemas.xmlsoap.org/wsdl/soap/\char`\"{}
- xmlns:wsdl=\char`\"{}http://schemas.xmlsoap.org/wsdl/\char`\"{}
- xmlns:wsa=\char`\"{}http://www.w3.org/2005/08/addressing\char`\"{}
- xmlns:deleg=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
+<?xml version="1.0" encoding="UTF-8"?>
+<wsdl:definitions targetNamespace="http://www.nordugrid.org/schemas/delegation"
+ xmlns:SOAP-ENV="http://schemas.xmlsoap.org/soap/envelope/"
+ xmlns:SOAP-ENC="http://schemas.xmlsoap.org/soap/encoding/"
+ xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"
+ xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+ xmlns:soap="http://schemas.xmlsoap.org/wsdl/soap/"
+ xmlns:wsdl="http://schemas.xmlsoap.org/wsdl/"
+ xmlns:wsa="http://www.w3.org/2005/08/addressing"
+ xmlns:deleg="http://www.nordugrid.org/schemas/delegation">
   <wsdl:types>
-    <xsd:schema targetNamespace=\char`\"{}http://www.nordugrid.org/schemas/delegation\char`\"{}>
-      <!-{}- Common types -{}->
-      <xsd:simpleType name=\char`\"{}TokenFormatType\char`\"{}>
-        <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
-          <xsd:enumeration value=\char`\"{}x509\char`\"{}/>
+    <xsd:schema targetNamespace="http://www.nordugrid.org/schemas/delegation">
+      <!-- Common types -->
+      <xsd:simpleType name="TokenFormatType">
+        <xsd:restriction base="xsd:string">
+          <xsd:enumeration value="x509"/>
         </xsd:restriction>
       </xsd:simpleType>
-      <xsd:complexType name=\char`\"{}ReferenceType\char`\"{}>
-        <xsd:sequence>
-          <xsd:any namespace=\char`\"{}\#\#other\char`\"{} processContents=\char`\"{}lax\char`\"{}
-                   minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:complexType name=\char`\"{}DelegatedTokenType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}Id\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-          <xsd:element name=\char`\"{}Value\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-          <xsd:element name=\char`\"{}Reference\char`\"{} type=\char`\"{}deleg:ReferenceType\char`\"{}
-                       minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}/>
-        </xsd:sequence>
-        <xsd:attribute name=\char`\"{}Format\char`\"{} type=\char`\"{}deleg:TokenFormatType\char`\"{}
-                       use=\char`\"{}required\char`\"{}/>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}DelegatedToken\char`\"{} type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
-      <xsd:complexType name=\char`\"{}TokenRequestType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}Id\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-          <xsd:element name=\char`\"{}Value\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-        </xsd:sequence>
-        <xsd:attribute name=\char`\"{}Format\char`\"{} type=\char`\"{}deleg:TokenFormatType\char`\"{}
-                       use=\char`\"{}required\char`\"{}/>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}TokenRequest\char`\"{} type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
-      <!-{}- Types for messages -{}->
-      <xsd:complexType name=\char`\"{}DelegateCredentialsInitRequestType\char`\"{}>
-        <xsd:sequence>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}DelegateCredentialsInit\char`\"{}
-                   type=\char`\"{}deleg:DelegateCredentialsInitRequestType\char`\"{}/>
-      <xsd:complexType name=\char`\"{}DelegateCredentialsInitResponseType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}TokenRequest\char`\"{} type=\char`\"{}deleg:TokenRequestType\char`\"{}/>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
-                   type=\char`\"{}deleg:DelegateCredentialsInitResponseType\char`\"{}/>
-      <xsd:complexType name=\char`\"{}UpdateCredentialsRequestType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}DelegatedToken\char`\"{} type=\char`\"{}deleg:DelegatedTokenType\char`\"{}/>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}UpdateCredentials\char`\"{}
-                   type=\char`\"{}deleg:UpdateCredentialsRequestType\char`\"{}/>
-      <xsd:complexType name=\char`\"{}UpdateCredentialsResponseType\char`\"{}>
-        <xsd:sequence>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}UpdateCredentialsResponse\char`\"{}
-                   type=\char`\"{}deleg:UpdateCredentialsResponseType\char`\"{}/>
-      <!-{}- Faults -{}->
-      <xsd:complexType name=\char`\"{}UnsupportedFaultType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}Description\char`\"{} type=\char`\"{}xsd:string\char`\"{}
-                       minOccurs=\char`\"{}0\char`\"{}/>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}UnsupportedFault\char`\"{} type=\char`\"{}deleg:UnsupportedFaultType\char`\"{}/>
-      <xsd:complexType name=\char`\"{}ProcessingFaultType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}Description\char`\"{} type=\char`\"{}xsd:string\char`\"{}
-                       minOccurs=\char`\"{}0\char`\"{}/>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}ProcessingFault\char`\"{} type=\char`\"{}deleg:ProcessingFaultType\char`\"{}/>
-      <xsd:complexType name=\char`\"{}WrongReferenceFaultType\char`\"{}>
-        <xsd:sequence>
-          <xsd:element name=\char`\"{}Description\char`\"{} type=\char`\"{}xsd:string\char`\"{}
-                       minOccurs=\char`\"{}0\char`\"{}/>
-        </xsd:sequence>
-      </xsd:complexType>
-      <xsd:element name=\char`\"{}WrongReferenceFault\char`\"{} type=\char`\"{}deleg:WrongReferenceFaultType\char`\"{}/>
+      <xsd:complexType name="ReferenceType">
+        <xsd:sequence>
+          <xsd:any namespace="##other" processContents="lax"
+                   minOccurs="0" maxOccurs="unbounded"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:complexType name="DelegatedTokenType">
+        <xsd:sequence>
+          <xsd:element name="Id" type="xsd:string"/>
+          <xsd:element name="Value" type="xsd:string"/>
+          <xsd:element name="Reference" type="deleg:ReferenceType"
+                       minOccurs="0" maxOccurs="unbounded"/>
+        </xsd:sequence>
+        <xsd:attribute name="Format" type="deleg:TokenFormatType"
+                       use="required"/>
+      </xsd:complexType>
+      <xsd:element name="DelegatedToken" type="deleg:DelegatedTokenType"/>
+      <xsd:complexType name="TokenRequestType">
+        <xsd:sequence>
+          <xsd:element name="Id" type="xsd:string"/>
+          <xsd:element name="Value" type="xsd:string"/>
+        </xsd:sequence>
+        <xsd:attribute name="Format" type="deleg:TokenFormatType"
+                       use="required"/>
+      </xsd:complexType>
+      <xsd:element name="TokenRequest" type="deleg:TokenRequestType"/>
+      <!-- Types for messages -->
+      <xsd:complexType name="DelegateCredentialsInitRequestType">
+        <xsd:sequence>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="DelegateCredentialsInit"
+                   type="deleg:DelegateCredentialsInitRequestType"/>
+      <xsd:complexType name="DelegateCredentialsInitResponseType">
+        <xsd:sequence>
+          <xsd:element name="TokenRequest" type="deleg:TokenRequestType"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="DelegateCredentialsInitResponse"
+                   type="deleg:DelegateCredentialsInitResponseType"/>
+      <xsd:complexType name="UpdateCredentialsRequestType">
+        <xsd:sequence>
+          <xsd:element name="DelegatedToken" type="deleg:DelegatedTokenType"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="UpdateCredentials"
+                   type="deleg:UpdateCredentialsRequestType"/>
+      <xsd:complexType name="UpdateCredentialsResponseType">
+        <xsd:sequence>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="UpdateCredentialsResponse"
+                   type="deleg:UpdateCredentialsResponseType"/>
+      <!-- Faults -->
+      <xsd:complexType name="UnsupportedFaultType">
+        <xsd:sequence>
+          <xsd:element name="Description" type="xsd:string"
+                       minOccurs="0"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="UnsupportedFault" type="deleg:UnsupportedFaultType"/>
+      <xsd:complexType name="ProcessingFaultType">
+        <xsd:sequence>
+          <xsd:element name="Description" type="xsd:string"
+                       minOccurs="0"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="ProcessingFault" type="deleg:ProcessingFaultType"/>
+      <xsd:complexType name="WrongReferenceFaultType">
+        <xsd:sequence>
+          <xsd:element name="Description" type="xsd:string"
+                       minOccurs="0"/>
+        </xsd:sequence>
+      </xsd:complexType>
+      <xsd:element name="WrongReferenceFault" type="deleg:WrongReferenceFaultType"/>
     </xsd:schema>
   </wsdl:types>
-  <wsdl:message name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
-    <wsdl:part name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}
-               element=\char`\"{}deleg:DelegateCredentialsInit\char`\"{}/>
+  <wsdl:message name="DelegateCredentialsInitRequest">
+    <wsdl:part name="DelegateCredentialsInitRequest"
+               element="deleg:DelegateCredentialsInit"/>
   </wsdl:message>
-  <wsdl:message name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
-    <wsdl:part name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
-               element=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
+  <wsdl:message name="DelegateCredentialsInitResponse">
+    <wsdl:part name="DelegateCredentialsInitResponse"
+               element="deleg:DelegateCredentialsInitResponse"/>
   </wsdl:message>
-  <wsdl:message name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
-    <wsdl:part name=\char`\"{}UpdateCredentialsRequest\char`\"{} element=\char`\"{}deleg:UpdateCredentials\char`\"{}/>
+  <wsdl:message name="UpdateCredentialsRequest">
+    <wsdl:part name="UpdateCredentialsRequest" element="deleg:UpdateCredentials"/>
   </wsdl:message>
-  <wsdl:message name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
-    <wsdl:part name=\char`\"{}UpdateCredentialsResponse\char`\"{}
-               element=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
+  <wsdl:message name="UpdateCredentialsResponse">
+    <wsdl:part name="UpdateCredentialsResponse"
+               element="deleg:UpdateCredentialsResponse"/>
   </wsdl:message>
-  <wsdl:message name=\char`\"{}UnsupportedFault\char`\"{}>
-    <wsdl:part name=\char`\"{}Detail\char`\"{} element=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
+  <wsdl:message name="UnsupportedFault">
+    <wsdl:part name="Detail" element="deleg:UnsupportedFault"/>
   </wsdl:message>
-  <wsdl:message name=\char`\"{}ProcessingFault\char`\"{}>
-    <wsdl:part name=\char`\"{}Detail\char`\"{} element=\char`\"{}deleg:ProcessingFault\char`\"{}/>
+  <wsdl:message name="ProcessingFault">
+    <wsdl:part name="Detail" element="deleg:ProcessingFault"/>
   </wsdl:message>
-  <wsdl:message name=\char`\"{}WrongReferenceFault\char`\"{}>
-    <wsdl:part name=\char`\"{}Detail\char`\"{} element=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
+  <wsdl:message name="WrongReferenceFault">
+    <wsdl:part name="Detail" element="deleg:WrongReferenceFault"/>
   </wsdl:message>
-  <wsdl:portType name=\char`\"{}DelegationPortType\char`\"{}>
-   <wsdl:operation name=\char`\"{}DelegateCredentialsInit\char`\"{}>
+  <wsdl:portType name="DelegationPortType">
+   <wsdl:operation name="DelegateCredentialsInit">
       <wsdl:documentation>
       </wsdl:documentation>
-      <wsdl:input name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}
-        message=\char`\"{}deleg:DelegateCredentialsInitRequest\char`\"{}/>
-      <wsdl:output name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}
-        message=\char`\"{}deleg:DelegateCredentialsInitResponse\char`\"{}/>
-      <wsdl:fault name=\char`\"{}UnsupportedFault\char`\"{}
-        message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}ProcessingFault\char`\"{}
-        message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
+      <wsdl:input name="DelegateCredentialsInitRequest"
+        message="deleg:DelegateCredentialsInitRequest"/>
+      <wsdl:output name="DelegateCredentialsInitResponse"
+        message="deleg:DelegateCredentialsInitResponse"/>
+      <wsdl:fault name="UnsupportedFault"
+        message="deleg:UnsupportedFault"/>
+      <wsdl:fault name="ProcessingFault"
+        message="deleg:ProcessingFault"/>
     </wsdl:operation>
-   <wsdl:operation name=\char`\"{}UpdateCredentials\char`\"{}>
+   <wsdl:operation name="UpdateCredentials">
       <wsdl:documentation>
       </wsdl:documentation>
-      <wsdl:input name=\char`\"{}UpdateCredentialsRequest\char`\"{}
-        message=\char`\"{}deleg:UpdateCredentialsRequest\char`\"{}/>
-      <wsdl:output name=\char`\"{}UpdateCredentialsResponse\char`\"{}
-        message=\char`\"{}deleg:UpdateCredentialsResponse\char`\"{}/>
-      <wsdl:fault name=\char`\"{}UnsupportedFault\char`\"{}
-        message=\char`\"{}deleg:UnsupportedFault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}ProcessingFault\char`\"{}
-        message=\char`\"{}deleg:ProcessingFault\char`\"{}/>
-      <wsdl:fault name=\char`\"{}WrongReferenceFault\char`\"{}
-        message=\char`\"{}deleg:WrongReferenceFault\char`\"{}/>
+      <wsdl:input name="UpdateCredentialsRequest"
+        message="deleg:UpdateCredentialsRequest"/>
+      <wsdl:output name="UpdateCredentialsResponse"
+        message="deleg:UpdateCredentialsResponse"/>
+      <wsdl:fault name="UnsupportedFault"
+        message="deleg:UnsupportedFault"/>
+      <wsdl:fault name="ProcessingFault"
+        message="deleg:ProcessingFault"/>
+      <wsdl:fault name="WrongReferenceFault"
+        message="deleg:WrongReferenceFault"/>
     </wsdl:operation>
   </wsdl:portType>
-  <wsdl:binding name=\char`\"{}DelegationBinding\char`\"{} type=\char`\"{}deleg:DelegationPortType\char`\"{}>
-    <soap:binding style=\char`\"{}document\char`\"{}
-                  transport=\char`\"{}http://schemas.xmlsoap.org/soap/http\char`\"{}/>
-    <wsdl:operation name=\char`\"{}DelegateCredentialsInit\char`\"{}>
-      <soap:operation soapAction=\char`\"{}DelegateCredentialsInit\char`\"{}/>
-      <wsdl:input name=\char`\"{}DelegateCredentialsInitRequest\char`\"{}>
-         <soap:body use=\char`\"{}literal\char`\"{}/>
+  <wsdl:binding name="DelegationBinding" type="deleg:DelegationPortType">
+    <soap:binding style="document"
+                  transport="http://schemas.xmlsoap.org/soap/http"/>
+    <wsdl:operation name="DelegateCredentialsInit">
+      <soap:operation soapAction="DelegateCredentialsInit"/>
+      <wsdl:input name="DelegateCredentialsInitRequest">
+         <soap:body use="literal"/>
       </wsdl:input>
-      <wsdl:output name=\char`\"{}DelegateCredentialsInitResponse\char`\"{}>
-         <soap:body use=\char`\"{}literal\char`\"{}/>
+      <wsdl:output name="DelegateCredentialsInitResponse">
+         <soap:body use="literal"/>
       </wsdl:output>
     </wsdl:operation>
-    <wsdl:operation name=\char`\"{}UpdateCredentials\char`\"{}>
-      <soap:operation soapAction=\char`\"{}UpdateCredentials\char`\"{}/>
-      <wsdl:input name=\char`\"{}UpdateCredentialsRequest\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{}/>
+    <wsdl:operation name="UpdateCredentials">
+      <soap:operation soapAction="UpdateCredentials"/>
+      <wsdl:input name="UpdateCredentialsRequest">
+        <soap:body use="literal"/>
       </wsdl:input>
-      <wsdl:output name=\char`\"{}UpdateCredentialsResponse\char`\"{}>
-        <soap:body use=\char`\"{}literal\char`\"{}/>
+      <wsdl:output name="UpdateCredentialsResponse">
+        <soap:body use="literal"/>
       </wsdl:output>
@@
 \begin{footnotesize}\begin{verbatim}
-<?xml version=\char`\"{}1.0\char`\"{} encoding=\char`\"{}UTF-8\char`\"{}?>
-<xsd:schema xmlns:xsd=\char`\"{}http://www.w3.org/2001/XMLSchema\char`\"{}
-            xmlns=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
-            xmlns:jsdl-arc=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}
-            targetNamespace=\char`\"{}http://www.nordugrid.org/ws/schemas/jsdl-arc\char`\"{}>
- <xsd:simpleType name=\char`\"{}GMState\_Type\char`\"{}>
-  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
-   <xsd:enumeration value=\char`\"{}ACCEPTED\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}PREPARING\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}SUBMIT\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}INLRMS\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}FINISHING\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}FINISHED\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}DELETED\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}CANCELING\char`\"{}/>
+<?xml version="1.0" encoding="UTF-8"?>
+<xsd:schema xmlns:xsd="http://www.w3.org/2001/XMLSchema"
+            xmlns="http://www.nordugrid.org/ws/schemas/jsdl-arc"
+            xmlns:jsdl-arc="http://www.nordugrid.org/ws/schemas/jsdl-arc"
+            targetNamespace="http://www.nordugrid.org/ws/schemas/jsdl-arc">
+ <xsd:simpleType name="GMState_Type">
+  <xsd:restriction base="xsd:string">
+   <xsd:enumeration value="ACCEPTED"/>
+   <xsd:enumeration value="PREPARING"/>
+   <xsd:enumeration value="SUBMIT"/>
+   <xsd:enumeration value="INLRMS"/>
+   <xsd:enumeration value="FINISHING"/>
+   <xsd:enumeration value="FINISHED"/>
+   <xsd:enumeration value="DELETED"/>
+   <xsd:enumeration value="CANCELING"/>
   </xsd:restriction>
  </xsd:simpleType>
- <xsd:complexType name=\char`\"{}Version\_Type\char`\"{}>
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}UpperExclusive\char`\"{} type=\char`\"{}xsd:string\char`\"{}
-                minOccurs=\char`\"{}0\char`\"{}/>
-   <xsd:element name=\char`\"{}LowerExclusive\char`\"{} type=\char`\"{}xsd:string\char`\"{}
-                minOccurs=\char`\"{}0\char`\"{}/>
-   <xsd:element name=\char`\"{}Exact\char`\"{} type=\char`\"{}xsd:string\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                maxOccurs=\char`\"{}unbounded\char`\"{}/>
-   <xsd:element name=\char`\"{}Exclusive\char`\"{} type=\char`\"{}xsd:boolean\char`\"{}
-                minOccurs=\char`\"{}0\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:simpleType name=\char`\"{}SessionType\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}SessionType\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
-  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
-   <xsd:enumeration value=\char`\"{}INTERNAL\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}LIMITED\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}READONLY\char`\"{}/>
-   <xsd:enumeration value=\char`\"{}FULL\char`\"{}/>
+ <xsd:complexType name="Version_Type">
+  <xsd:sequence>
+   <xsd:element name="UpperExclusive" type="xsd:string"
+                minOccurs="0"/>
+   <xsd:element name="LowerExclusive" type="xsd:string"
+                minOccurs="0"/>
+   <xsd:element name="Exact" type="xsd:string" minOccurs="0"
+                maxOccurs="unbounded"/>
+   <xsd:element name="Exclusive" type="xsd:boolean"
+                minOccurs="0"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name="SessionType_Type">
+  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
+  <!-- xsd:element ref="SessionType" minOccurs="0"/ -->
+  <xsd:restriction base="xsd:string">
+   <xsd:enumeration value="INTERNAL"/>
+   <xsd:enumeration value="LIMITED"/>
+   <xsd:enumeration value="READONLY"/>
+   <xsd:enumeration value="FULL"/>
   </xsd:restriction>
  </xsd:simpleType>
- <xsd:simpleType name=\char`\"{}IsExecutable\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:DataStaging\_Type (default: false) </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}IsExecutable\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
-  <xsd:restriction base=\char`\"{}xsd:boolean\char`\"{}/>
+ <xsd:simpleType name="IsExecutable_Type">
+  <xsd:documentation> For jsdl:DataStaging_Type (default: false) </xsd:documentation>
+  <!-- xsd:element ref="IsExecutable" minOccurs="0"/ -->
+  <xsd:restriction base="xsd:boolean"/>
  </xsd:simpleType>
- <xsd:simpleType name=\char`\"{}FileParameters\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:DataStaging\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}IsExecutable\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
-  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}/>
+ <xsd:simpleType name="FileParameters_Type">
+  <xsd:documentation> For jsdl:DataStaging_Type </xsd:documentation>
+  <!-- xsd:element ref="IsExecutable" minOccurs="0"/ -->
+  <xsd:restriction base="xsd:string"/>
  </xsd:simpleType>
- <xsd:simpleType name=\char`\"{}JoinOutputs\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:JobDescription\_Type (default: false) </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}JoinOutputs\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
-  <xsd:restriction base=\char`\"{}xsd:boolean\char`\"{}/>
+ <xsd:simpleType name="JoinOutputs_Type">
+  <xsd:documentation> For jsdl:JobDescription_Type (default: false) </xsd:documentation>
+  <!-- xsd:element ref="JoinOutputs" minOccurs="0"/ -->
+  <xsd:restriction base="xsd:boolean"/>
  </xsd:simpleType>
- <xsd:simpleType name=\char`\"{}Reruns\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:JobDescription\_Type (default: false) </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}Reruns\char`\"{} minOccurs=\char`\"{}0\char`\"{}// -{}->
-  <xsd:restriction base=\char`\"{}xsd:integer\char`\"{}/>
+ <xsd:simpleType name="Reruns_Type">
+  <xsd:documentation> For jsdl:JobDescription_Type (default: false) </xsd:documentation>
+  <!-- xsd:element ref="Reruns" minOccurs="0"// -->
+  <xsd:restriction base="xsd:integer"/>
  </xsd:simpleType>
- <xsd:complexType name=\char`\"{}RunTimeEnvironment\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}RunTimeEnvironment\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                     maxOccurs=\char`\"{}unbounded\char`\"{}/ -{}->
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}Name\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-   <xsd:element name=\char`\"{}Version\char`\"{} type=\char`\"{}Version\_Type\char`\"{}
-                minOccurs=\char`\"{}0\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:complexType name=\char`\"{}Middleware\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}Middleware\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                     maxOccurs=\char`\"{}unbounded\char`\"{}/ -{}->
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}Name\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-   <xsd:element name=\char`\"{}Version\char`\"{} type=\char`\"{}Version\_Type\char`\"{}
-                minOccurs=\char`\"{}0\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:complexType name=\char`\"{}RemoteLogging\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}RemoteLogging\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                     maxOccurs=\char`\"{}3\char`\"{}/ -{}->
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}URL\char`\"{} minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}1\char`\"{}
-                type=\char`\"{}xsd:anyURI\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:complexType name=\char`\"{}CredentialServer\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}CredentialServer\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}URL\char`\"{} minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}1\char`\"{}
-                type=\char`\"{}xsd:anyURI\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:complexType name=\char`\"{}LocalLogging\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}LocalLogging\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}Directory\char`\"{} minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}1\char`\"{}
-                type=\char`\"{}xsd:string\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:simpleType name=\char`\"{}AccessControlType\_Type\char`\"{}>
-  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
-   <xsd:enumeration value=\char`\"{}GACL\char`\"{}/>
+ <xsd:complexType name="RunTimeEnvironment_Type">
+  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
+  <!-- xsd:element ref="RunTimeEnvironment" minOccurs="0"
+                     maxOccurs="unbounded"/ -->
+  <xsd:sequence>
+   <xsd:element name="Name" type="xsd:string"/>
+   <xsd:element name="Version" type="Version_Type"
+                minOccurs="0"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:complexType name="Middleware_Type">
+  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
+  <!-- xsd:element ref="Middleware" minOccurs="0"
+                     maxOccurs="unbounded"/ -->
+  <xsd:sequence>
+   <xsd:element name="Name" type="xsd:string"/>
+   <xsd:element name="Version" type="Version_Type"
+                minOccurs="0"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:complexType name="RemoteLogging_Type">
+  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
+  <!-- xsd:element ref="RemoteLogging" minOccurs="0"
+                     maxOccurs="3"/ -->
+  <xsd:sequence>
+   <xsd:element name="URL" minOccurs="1" maxOccurs="1"
+                type="xsd:anyURI"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:complexType name="CredentialServer_Type">
+  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
+  <!-- xsd:element ref="CredentialServer" minOccurs="0"/ -->
+  <xsd:sequence>
+   <xsd:element name="URL" minOccurs="1" maxOccurs="1"
+                type="xsd:anyURI"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:complexType name="LocalLogging_Type">
+  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
+  <!-- xsd:element ref="LocalLogging" minOccurs="0"
+                     maxOccurs="1"/ -->
+  <xsd:sequence>
+   <xsd:element name="Directory" minOccurs="1" maxOccurs="1"
+                type="xsd:string"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name="AccessControlType_Type">
+  <xsd:restriction base="xsd:string">
+   <xsd:enumeration value="GACL"/>
   </xsd:restriction>
  </xsd:simpleType>
- <xsd:complexType name=\char`\"{}AccessControl\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}AccessControl\char`\"{} minOccurs=\char`\"{}0\char`\"{}/ -{}->
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}OwnerAlwaysAllowed\char`\"{} type=\char`\"{}xsd:boolean\char`\"{}
-                minOccurs=\char`\"{}0\char`\"{}/>
-   <xsd:element name=\char`\"{}Type\char`\"{} type=\char`\"{}AccessControlType\_Type\char`\"{}
-                minOccurs=\char`\"{}0\char`\"{}/>
-   <xsd:element name=\char`\"{}Content\char`\"{} minOccurs=\char`\"{}0\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:simpleType name=\char`\"{}NotificationType\_Type\char`\"{}>
-  <xsd:restriction base=\char`\"{}xsd:string\char`\"{}>
-   <xsd:enumeration value=\char`\"{}Email\char`\"{}/>
+ <xsd:complexType name="AccessControl_Type">
+  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
+  <!-- xsd:element ref="AccessControl" minOccurs="0"/ -->
+  <xsd:sequence>
+   <xsd:element name="OwnerAlwaysAllowed" type="xsd:boolean"
+                minOccurs="0"/>
+   <xsd:element name="Type" type="AccessControlType_Type"
+                minOccurs="0"/>
+   <xsd:element name="Content" minOccurs="0" type="xsd:string"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name="NotificationType_Type">
+  <xsd:restriction base="xsd:string">
+   <xsd:enumeration value="Email"/>
   </xsd:restriction>
  </xsd:simpleType>
- <xsd:complexType name=\char`\"{}Notify\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}Notify\char`\"{} minOccurs=\char`\"{}0\char`\"{} maxOccurs=\char`\"{}3\char`\"{}/
-    -{}->
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}Type\char`\"{} type=\char`\"{}NotificationType\_Type\char`\"{}
-                minOccurs=\char`\"{}0\char`\"{}/>
-   <xsd:element name=\char`\"{}Endpoint\char`\"{} minOccurs=\char`\"{}0\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-   <xsd:element name=\char`\"{}State\char`\"{} minOccurs=\char`\"{}1\char`\"{} maxOccurs=\char`\"{}unbounded\char`\"{}
-                type=\char`\"{}GMState\_Type\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:simpleType name=\char`\"{}SessionLifeTime\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}SessionLifeTime\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
-  <xsd:restriction base=\char`\"{}xsd:long\char`\"{}/>
+ <xsd:complexType name="Notify_Type">
+  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
+  <!-- xsd:element ref="Notify" minOccurs="0" maxOccurs="3"/
+    -->
+  <xsd:sequence>
+   <xsd:element name="Type" type="NotificationType_Type"
+                minOccurs="0"/>
+   <xsd:element name="Endpoint" minOccurs="0" type="xsd:string"/>
+   <xsd:element name="State" minOccurs="1" maxOccurs="unbounded"
+                type="GMState_Type"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name="SessionLifeTime_Type">
+  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
+  <!-- xsd:element ref="SessionLifeTime" minOccurs="0"
+                     maxOccurs="1"/ -->
+  <xsd:restriction base="xsd:long"/>
  </xsd:simpleType>
- <xsd:simpleType name=\char`\"{}GridTimeLimit\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}GridTimeLimit\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
-  <xsd:restriction base=\char`\"{}xsd:positiveInteger\char`\"{}/>
+ <xsd:simpleType name="GridTimeLimit_Type">
+  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
+  <!-- xsd:element ref="GridTimeLimit" minOccurs="0"
+                     maxOccurs="1"/ -->
+  <xsd:restriction base="xsd:positiveInteger"/>
  </xsd:simpleType>
- <xsd:complexType name=\char`\"{}CandidateTarget\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:Resources\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}jsdl-arc:CandidateTarget\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
-  <xsd:sequence>
-   <xsd:element name=\char`\"{}HostName\char`\"{} minOccurs=\char`\"{}0\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-   <xsd:element name=\char`\"{}QueueName\char`\"{} minOccurs=\char`\"{}0\char`\"{} type=\char`\"{}xsd:string\char`\"{}/>
-  </xsd:sequence>
- </xsd:complexType>
- <xsd:simpleType name=\char`\"{}Time\_Type\char`\"{}>
-  <xsd:documentation> For jsdl:JobDescription\_Type </xsd:documentation>
-  <!-{}- xsd:element ref=\char`\"{}ProcessingStartTime\char`\"{} minOccurs=\char`\"{}0\char`\"{}
-                     maxOccurs=\char`\"{}1\char`\"{}/ -{}->
-  <xsd:restriction base=\char`\"{}xsd:dateTime\char`\"{}/>
+ <xsd:complexType name="CandidateTarget_Type">
+  <xsd:documentation> For jsdl:Resources_Type </xsd:documentation>
+  <!-- xsd:element ref="jsdl-arc:CandidateTarget" minOccurs="0"
+                     maxOccurs="1"/ -->
+  <xsd:sequence>
+   <xsd:element name="HostName" minOccurs="0" type="xsd:string"/>
+   <xsd:element name="QueueName" minOccurs="0" type="xsd:string"/>
+  </xsd:sequence>
+ </xsd:complexType>
+ <xsd:simpleType name="Time_Type">
+  <xsd:documentation> For jsdl:JobDescription_Type </xsd:documentation>
+  <!-- xsd:element ref="ProcessingStartTime" minOccurs="0"
+                     maxOccurs="1"/ -->
+  <xsd:restriction base="xsd:dateTime"/>
  </xsd:simpleType>
- <!-{}-=========================================================-{}->
- <xsd:element name=\char`\"{}IsExecutable\char`\"{} type=\char`\"{}IsExecutable\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}FileParameters\char`\"{} type=\char`\"{}FileParameters\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}RunTimeEnvironment\char`\"{} type=\char`\"{}RunTimeEnvironment\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}Middleware\char`\"{} type=\char`\"{}Middleware\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}RemoteLogging\char`\"{} type=\char`\"{}RemoteLogging\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}LocalLogging\char`\"{} type=\char`\"{}LocalLogging\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}AccessControl\char`\"{} type=\char`\"{}AccessControl\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}Notify\char`\"{} type=\char`\"{}Notify\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}SessionLifeTime\char`\"{} type=\char`\"{}SessionLifeTime\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}SessionType\char`\"{} type=\char`\"{}SessionType\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}JoinOutputs\char`\"{} type=\char`\"{}JoinOutputs\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}Reruns\char`\"{} type=\char`\"{}Reruns\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}CredentialServer\char`\"{} type=\char`\"{}CredentialServer\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}GridTimeLimit\char`\"{} type=\char`\"{}GridTimeLimit\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}CandidateTarget\char`\"{} type=\char`\"{}CandidateTarget\_Type\char`\"{}/>
- <xsd:element name=\char`\"{}ProcessingStartTime\char`\"{} type=\char`\"{}Time\_Type\char`\"{}/>
+ <!--=========================================================-->
+ <xsd:element name="IsExecutable" type="IsExecutable_Type"/>
+ <xsd:element name="FileParameters" type="FileParameters_Type"/>
+ <xsd:element name="RunTimeEnvironment" type="RunTimeEnvironment_Type"/>
+ <xsd:element name="Middleware" type="Middleware_Type"/>
+ <xsd:element name="RemoteLogging" type="RemoteLogging_Type"/>
+ <xsd:element name="LocalLogging" type="LocalLogging_Type"/>
+ <xsd:element name="AccessControl" type="AccessControl_Type"/>
+ <xsd:element name="Notify" type="Notify_Type"/>
+ <xsd:element name="SessionLifeTime" type="SessionLifeTime_Type"/>
+ <xsd:element name="SessionType" type="SessionType_Type"/>
+ <xsd:element name="JoinOutputs" type="JoinOutputs_Type"/>
+ <xsd:element name="Reruns" type="Reruns_Type"/>
+ <xsd:element name="CredentialServer" type="CredentialServer_Type"/>
+ <xsd:element name="GridTimeLimit" type="GridTimeLimit_Type"/>
+ <xsd:element name="CandidateTarget" type="CandidateTarget_Type"/>
+ <xsd:element name="ProcessingStartTime" type="Time_Type"/>
 </xsd:schema>

### 2009-10-20 doc/tech_doc/a-rex/arex_tech_doc.tex
add hed configuration info and clean up the rest

@@
 
-Each job gets a directory on the CE called the \emph{session directory}
-(SD). Input files are gathered in the SD. The job is supposed to produce
-new data files in the SD as well. The A-REX does not guarantee the
-availability of any other places accessible by the job other than
+Each job gets a directory on the CE called the \emph{session
+  directory} (SD). Input files are gathered in the SD. The job may
+also produce new data files in the SD. The A-REX does not guarantee
+the availability of any other places accessible by the job other than
 SD (unless such a place is part of a requested Runtime Environment).
-The SD is also the only place which is controlled by the A-REX. It
-is accessible by the user from outside through the HTTP(S) protocol.
-Any file created outside the SD is not controlled by the A-REX. Any
+The SD is also the only place which is controlled by the A-REX. It is
+accessible by the user from outside through the HTTP(S) protocol.  Any
+file created outside the SD is not controlled by the A-REX. Any
 exchange of data between client and A-REX (including also program
 modules) is performed via HTTP(S). An URL for accessing input/output
-files is  obtained through the WS Local Information Description Interface
-(LIDI) of A-REX.
+files is obtained through the WS Local Information Description
+Interface (LIDI) of A-REX.
 
@@
 data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
-http, https, lfc} and \emph{rls.} For more information please see
+http, https, lfc, rls} and \emph{srm.} For more information please see
 {}``The Hosting Environment of the Advanced Resource Connector middleware''
@@
 \textit{control directory}. In most common configuration case A-REX
-serves all users defined by Operating System and stores their control
-file in same directory. To make it easier to recover in case of failure,
+serves all users defined by the Operating System and stores their control
+files in the same directory. To make it easier to recover in case of failure,
 the A-REX stores most information in files rather than in memory.
@@
 \item \textit{forcemigration} -- This boolean is only used for migration of jobs.
-It determines whether the job should persists if the termination of the previous 
+It determines whether the job should persist if the termination of the previous 
 job fails.
-\end{itemize}
 \end{itemize}
@@
 
-\begin{itemize}
 \item \textit{job.ID.input} -- list of input files. Each line contains 2
 values separated by a space. First value contains name of the file
-relative to the SD. Second value is an URL or a file description.
+relative to the SD. Second value is a URL or a file description.
 Example:
@@
 
-\emph{URL} represents location from which file can be downloaded.
+A URL represents a location from which a file can be downloaded.
 Each URL can contain additional options.
 
-\textit{file description} -- {[}size]{[}.checksum].
-
-\hspace*{1cm}\textit{size} -- size of the file in bytes.
-
-\hspace*{1cm}\textit{checksum} -- checksum of the file identical to
+A file description refers to a file uploaded from the UI and consists
+of {[}size]{[}.checksum] where
+
+\hspace*{1cm}\textit{size} - size of the file in bytes.
+
+\hspace*{1cm}\textit{checksum} - checksum of the file identical to
 the one produced by \textbf{\textit{cksum}} (1).
 
-Both size and checksum can be left out. Special kind of file description
-{*}.{*} is used to specify files which are \textbf{not} required to
-exist.
-\end{itemize}
+These values are used to verify the transfer of the uploaded
+file. Both size and checksum can be left out. A special kind of file
+description {*}.{*} is used to specify files which are \textbf{not}
+required to exist.
 
@@
 
-\begin{itemize}
 \item \textit{job.ID.output} -- list of output files. Each line contains
@@
 URLs are the same as those supported by job.ID.input.
-\end{itemize}
 
@@
 
-\begin{itemize}
 \item \textit{job.ID.failed} -- the existence of this file marks the failure
@@
 
-The A-REX can cache input files. Caching is enabled if one or more cache
-directories are specified in the configuration file. The A-REX does not
-cache files marked as executable in a job. Caching can also be explicitly
-turned off by the user for each file by using the\textit{ cache=no}
-option in the URL (for URL options read {}``Protocols, Uniform Resource
-Locators (URL) and extensions supported in ARC'' \cite{urls}). The
-disk space occupied by the cache is controlled by removing old files.
-For more information look in section \ref{SubSection:ConfigFile}.
+The A-REX can cache input files. Caching is enabled if one or more
+cache directories are specified in the configuration file. Caching can
+also be explicitly turned off by the user for each file by using
+the\textit{ cache=no} option in the URL (for URL options read
+{}``Protocols, Uniform Resource Locators (URL) and extensions
+supported in ARC'' \cite{urls}). The disk space occupied by the cache
+is controlled by removing old files.  For more information on
+configuration look in section \ref{SubSection:ConfigFile}.
 
@@
 
-\subsection{Administration tools}
+\subsection{Cache Administration}
+
+The cache is cleaned automatically periodically (every 2 minutes) by
+the A-REX to keep the size of each cache within the configured limits.
+Files are removed from the cache if the total size of the cache is
+greater than the configured limit. Files which are not locked are
+removed in order of access time, starting with the earliest, until the
+size is lower than the configured lower limit. If the lower limit
+cannot be reached (because too many files are locked, or other files
+outside the cache are taking up space on the file system), the
+cleaning will stop before the lower limit is reached.
+
+Since the limits on cache size are given as a percentage of space used
+on the filesystem on which the cache is located, it is recommended
+that each cache has its own dedicated file system. If the cache shares
+space with other data on a file system, changes in the amount of
+non-cache data will result in changes in the available cache space.
+
+With large caches mounted over NFS and an A-REX heavily loaded with
+data transfer processes, cache cleaning can become slow, leading to
+caches filling up beyond their configured limits. For performance
+reasons it may be advantageous to disable cache cleaning by the A-REX,
+and run the \emph{cache-clean} tool independently on the machine
+hosting the file system.
 
@@
 \begin{itemize}
-\item \emph{cache-clean} - This tool is used periodically (every 2 minutes)
-by the A-REX to keep the size of each cache within the configurated limits.
-It removes files from the cache if the total size of the cache is
-greater than the configured limit. It will attempt to remove files
-which are not locked in order of access time, starting with the earliest,
-until the size is lower than the configured lower limit. If the lower
-limit cannot be reached (because too many files are locked, or other
-files outside the cache are taking up space on the file system), the
-tool will exit before the lower limit is reached.\\
-\emph{cache-clean -h} gives a list of options. The most useful option
-for administrators is \emph{-s}, which does not delete anything, but
-gives summary information on the files in the cache, including information
-on the ages of the files in the cache.\\
-It is not recommended to run \emph{cache-clean }manually to clean
-up the cache, unless it is desired to temporarily clean up the cache
-with different size limits to those specified in the configuration.
-\item \emph{cache-list} - This tool is used to list all files present in
-each cache. It simply reads through all the \emph{.meta }files and
-prints to stdout a list of all URLs stored in each cache and their
-corresponding cache filename, one per line.
+\item \emph{cache-clean} - This tool is used periodically by the A-REX to
+  keep the size of each cache within the configured
+  limits.\\
+  \emph{cache-clean -h} gives a list of options. The most
+  useful option for administrators is \emph{-s}, which does not delete
+  anything, but gives summary information on the files in the cache,
+  including information on the ages of the files in the cache.\\
+  It is not recommended to run \emph{cache-clean} manually to clean up
+  the cache, unless it is desired to temporarily clean up the cache with
+  different size limits to those specified in the configuration, or to
+  improve performance by running it on the file system's local node as
+  mentioned above.
+\item \emph{cache-list} - This tool is used to list all files present
+  in each cache. It simply reads through all the \emph{.meta }files
+  and prints to stdout a list of all URLs stored in each cache and
+  their corresponding cache filename, one per line.
 \end{itemize}
@@
 Several (or even all) users may share the same session root directory.\\
-The A-REX need to have permission to create new files and directories
+The A-REX needs to have permission to create new files and directories
 in the session root directory. If A-REX is run under a dedicated user
@@
 \begin{shaded}
-\verb|name=''arguments string''|
+\verb|name=``arguments string''|
 \end{shaded}
@@
 and after job was deleted (\emph{ttr}) due to \emph{ttl}. Defaults
-are 7days for \emph{ttl} and 30 days for \emph{ttr}.
+are 7 days for \emph{ttl} and 30 days for \emph{ttr}.
 \item \textbf{\textit{lrms}}\textit{=default\_lrms\_name default\_queue\_name}
@@
 \item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} - specifies
-a directory to store cached data (see \ref{sec:cache}).\emph{ }Multiple
+a directory to store cached data (see \ref{sec:cache}). Multiple
 cache directories may be specified by specifying multiple \emph{cachedir}
-commands. Cached data will be distributed evenly over the caches.
+commands. Cached files will be distributed evenly between the caches.
 Specifying no \emph{cachedir }command or commands with an empty path
@@
 
-The A-REX service is a pluggable module of the HED. So first You need
-to setup your HED -- see HED documentation.
-
-Then You need to add A-REX to the HED configuration. Add new \verb|<Name>|
-element inside \verb|<Plugins>| containing string \emph{arex}. That will
-make HED load libarex plugin library.
-
-Then add new \verb|<Service>| element with attribute \verb|name=``a-rex''|. That
-will instantiate A-REX service. Now to make service accessible extend
-\verb|<Plexer>| element with new \verb|<next>| referring to an id of the service.
-Write \verb|<Service>| element carefully -- here is an example:
+The A-REX service is a pluggable module of the HED, so it is first
+required to set up HED configuration, and add the A-REX elements. To
+add A-REX to an existing HED configuration, add a new \verb|<Name>|
+element inside \verb|<Plugins>| containing the string
+\emph{arex}. This will make HED load the libarex plugin library.
+
+Then add a new \verb|<Service>| element with attribute
+\verb|name="a-rex"|. That will instantiate the A-REX service. Now to
+make service accessible extend the \verb|<Plexer>| element with new
+\verb|<next>| referring to an id of the service. Take care to write
+the \verb|<Service>| element carefully.
+
+Here is an example of the full HED configuration file including the
+A-REX service definition:
 
@@
 \begin{verbatim}
+<?xml version="1.0"?>
+<ArcConfig
+  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
+  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
+  xmlns:arex="http://www.nordugrid.org/schemas/a-rex/Config"
+>
+  <!-- Common configuration of the daemon -->
+  <Server>
+    <PidFile>/var/run/arched.pid</PidFile>
+    <Logger>
+      <Level>DEBUG</Level>
+      <File>/var/log/arched.log</File>
+      <Backups>10</Backups>
+      <Maxsize>100000000</Maxsize>
+    </Logger>
+  </Server>
+  <!-- Where to find plugins -->
+  <ModuleManager>
+    <Path>/usr/local/lib/arc/</Path>
+  </ModuleManager>
+  <!-- Simply load all needed plugins -->
+  <Plugins><Name>mcctcp</Name></Plugins>
+  <Plugins><Name>mcctls</Name></Plugins>
+  <Plugins><Name>mcchttp</Name></Plugins>
+  <Plugins><Name>mccsoap</Name></Plugins>
+  <Plugins><Name>arcpdc</Name></Plugins>
+  <Plugins><Name>identitymap</Name></Plugins>
+  <Plugins><Name>arex</Name></Plugins>
+  <!-- Create a chain -->
+  <Chain>
+    <!-- TCP listening socket -->
+    <Component name="tcp.service" id="tcp">
+      <next id="tls"/>
+      <tcp:Listen><tcp:Port>60000</tcp:Port></tcp:Listen>
+    </Component>
+    <!-- Transport-level security -->
+    <Component name="tls.service" id="tls">
+      <next id="http"/>
+      <!-- Location of server's security keys -->
+      <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
+      <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
+      <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+      <!-- Evaluate requestor's identity into local identity.
+           Comment it if no user mapping is needed. -->
+      <SecHandler name="identity.map" id="map" event="incoming">
+        <!-- Safe choice if all other rules failed -->
+        <PDP name="allow.pdp"><LocalName>nobody</LocalName></PDP>
+      </SecHandler>
+    </Component>
+    <!-- HTTP processing is done here -->
+    <Component name="http.service" id="http">
+      <next id="soap">POST</next>
+      <next id="plexer">GET</next>
+      <next id="plexer">PUT</next>
+    </Component>
+    <!-- This one parses content into XML tree -->
+    <Component name="soap.service" id="soap">
+      <next id="plexer"/>
+    </Component>
+    <!-- Directing messages to proper service -->
+    <Plexer name="plexer.service" id="plexer">
+      <!-- RegExp pattern is matched to path part of endpoint.
+           Unmatched part of path is propagated to service in
+           PLEXER:EXTENSION attribute. -->
+      <next id="a-rex">^/arex</next>
+    </Plexer>
     <!-- A-Rex service -->
@@
     </Service>
+  </Chain>
+</ArcConfig>
 \end{verbatim}
@@
 
-Use a template arc\_arex.conf or write own A-REX configuration file.
-For information about format and available configuration commands
-see Section~\ref{SubSection:ConfigFile}.
+For A-REX configuration, either use a template arc.conf or write a new
+A-REX configuration file.  For information about format and available
+configuration commands see Section~\ref{SubSection:ConfigFile}. It is
+also possible to specify the A-REX configuration in XML format in the
+HED configuration file, instead of a separate file. For more
+information on this see the examples included in the documentation
+bundled with the release.
 
@@
 
-Please read \emph{User Guide}~\cite{userguide1} for usage instructions.
+For a quick start, simply run
+
+\begin{shaded}
+\begin{verbatim}
+$ARC_LOCATION/sbin/arched -c <HED configuration file>
+\end{verbatim}
+\end{shaded}
+
+For more information please read the \emph{User Guide}~\cite{userguide1}.
 

### 2009-10-20 doc/manuals/chelonia/arc-storage-manual.tex
Chelonia administrator's manual is more or less updated (using INI configs, and the latest trunk).

@@
 
-This section shows how to install Chelonia on a freshly installed debian lenny machine.
-
-Get the latest ARC code from the subversion (\verb!http://svn.nordugrid.org/!):
-
-\begin{verbatim}
-$ svn co http://svn.nordugrid.org/repos/nordugrid/arc1/trunk arc1
-\end{verbatim}
-
-Check the README file for all the dependencies, and install them.
-Here `aptitude' is used, but of course on other Linux distribution different package management software may be used, and the name of the packages may differ as well.
-
-\begin{verbatim}
-$ sudo aptitude install build-essential
-$ sudo aptitude install autoconf
-$ sudo aptitude install gettext
-$ sudo aptitude install cvs
-$ sudo aptitude install libtool
-$ sudo aptitude install pkg-config
-$ sudo aptitude install libglibmm-2.4-dev
-$ sudo aptitude install python-dev
-$ sudo aptitude install swig
-$ sudo aptitude install libxml2-dev
-$ sudo aptitude install libssl-dev
-\end{verbatim}
-
-Now, run the \verb!autogen! and \verb!configure! scripts (target directories for the installation may be specified with the \verb!--prefix! option):
-
-\begin{verbatim}
-$ cd arc1
-$ ./autogen.sh
-$ ./configure --disable-java --disable-a-rex-service --disable-isi-service \
-    --disable-charon-service --disable-compiler-service \
-    --disable-paul-service --disable-sched-service
-
-Unit testing:       yes
-Java binding:       no
-Python binding:     yes (2.5)
-
-Available third-party features:
-
-RLS:                no
-GridFTP:            no
-LFC:                no
-RSL:                no
-SAML:               no
-MYSQL CLIENT LIB:   no
-gSOAP:              no
-
-Included components:
-A-Rex service:      no
-ISI service:        no
-CHARON service:     no
-HOPI service:       yes
-SCHED service:      no
-STORAGE service:    yes
-PAUL service:       no
-SRM client (DMC):   no
-GSI channel (MCC):  no
-
-\end{verbatim}
-
-Start with compiling it:
-
-\begin{verbatim}
-$ make
-\end{verbatim}
-
-And then install it (you need to be root if you want to install it to the default location):
-
-\begin{verbatim}
-$ sudo make install
-\end{verbatim}
-
-It may be a good idea to run \verb!ldconfig! to refresh the list of libraries:
-
-\begin{verbatim}
-$ ldconfig
-\end{verbatim}
-
-Check if the python packages are installed correctly:
+The Chelonia storage system can be installed from binary packages, or from compiling the source packages.
+Here we assume that it is installed on every machine we want to use.
+But we can do a quick check if the ARC client python libraries are installed properly.
+It is recommended to set the PYTHONPATH variable, e.g.:
+
+\begin{verbatim}
+$ export PYTHONPATH=/usr/local/lib/python2.5/site-packages
+\end{verbatim}
+
+Then we can check if the packages can be imported:
 
@@
 
-If there is no error on importing any of the packages, everything is OK. If the packages cannot be imported, try to set the PYTHONPATH environment variable to point to location of the packages.
-
-It is important to syncronize the time of the machine - the storage system will make decisions based on differences between timestamps.
+If there is no error on importing any of the packages, everything is OK.
+
+It is very important to syncronize the time of the machines (at least within a few seconds) - the storage system will make decisions based on differences between timestamps.
 When deploying multiple machines, different times will cause problems.
+
+In this document we will use a debian lenny system, and we will run multiple versions of the ARC HED hosting environment to simulate multiple machines.
+We will use three different hostnames , which will in this case point to the same machine, but in real deployment they would of course point to different machines (\verb!chelonia[123].knowarc.eu!).
 
@@
 Now, create 3 user and 3 host certificates.
-(for this example, set the Organization Name to `knowarc', the Common name of the CA to `storage-test', the name of the users to `penny', `billy' and `hammer' and the name of the hosts to `upsilon', `omicron' and `theta'.
+(for this example, we set the Organization Name to `knowarc', the Common name of the CA to `chelonia', the name of the users to `penny', `billy' and `hammer' and the name of the hosts to `chelonia1', `chelonia2' and `chelonia3'.
 A password for the CA and passwords for the user certificates are needed as well, these should be at least 4 characters long).
@@
 total 28
-drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 CA
--rw-r--r-- 1 zsombor zsombor  963 2009-03-19 12:11 ca.key
--rw-r--r-- 1 zsombor zsombor  786 2009-03-19 12:11 ca.pem
-drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 hostCerts
--rw-r--r-- 1 zsombor zsombor 1044 2009-03-19 12:11 readme
--rw-r--r-- 1 zsombor zsombor    3 2009-03-19 12:11 serial.srl
-drwxr-xr-x 2 zsombor zsombor 4096 2009-03-19 12:11 userCerts
+drwxr-xr-x 2 zsombor zsombor 4096 2009-10-16 13:31 CA
+-rw-r--r-- 1 zsombor zsombor  951 2009-10-16 13:31 ca.key
+-rw-r--r-- 1 zsombor zsombor  778 2009-10-16 13:31 ca.pem
+drwxr-xr-x 2 zsombor zsombor 4096 2009-10-16 13:31 hostCerts
+-rw-r--r-- 1 zsombor zsombor 1044 2009-10-16 13:31 readme
+-rw-r--r-- 1 zsombor zsombor    3 2009-10-16 13:31 serial.srl
+drwxr-xr-x 2 zsombor zsombor 4096 2009-10-16 13:31 userCerts
 
@@
 total 8
--rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 66fe707c.0
--rw-r--r-- 1 zsombor zsombor 139 2009-03-19 12:11 66fe707c.signing_policy
+-rw-r--r-- 1 zsombor zsombor 778 2009-10-16 13:31 afc5f10f.0
+-rw-r--r-- 1 zsombor zsombor 135 2009-10-16 13:31 afc5f10f.signing_policy
 
@@
 total 24
--rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 hostcert-omicron.pem
--rw-r--r-- 1 zsombor zsombor 782 2009-03-19 12:11 hostcert-theta.pem
--rw-r--r-- 1 zsombor zsombor 786 2009-03-19 12:11 hostcert-upsilon.pem
--rw-r--r-- 1 zsombor zsombor 887 2009-03-19 12:11 hostkey-omicron.pem
--rw-r--r-- 1 zsombor zsombor 891 2009-03-19 12:11 hostkey-theta.pem
--rw-r--r-- 1 zsombor zsombor 887 2009-03-19 12:11 hostkey-upsilon.pem
+-rw-r--r-- 1 zsombor zsombor 782 2009-10-16 13:31 hostcert-chelonia1.pem
+-rw-r--r-- 1 zsombor zsombor 782 2009-10-16 13:31 hostcert-chelonia2.pem
+-rw-r--r-- 1 zsombor zsombor 782 2009-10-16 13:31 hostcert-chelonia3.pem
+-rw-r--r-- 1 zsombor zsombor 887 2009-10-16 13:31 hostkey-chelonia1.pem
+-rw-r--r-- 1 zsombor zsombor 887 2009-10-16 13:31 hostkey-chelonia2.pem
+-rw-r--r-- 1 zsombor zsombor 887 2009-10-16 13:31 hostkey-chelonia3.pem
 
@@
 total 24
--rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-billy.pem
--rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-hammer.pem
--rw-r--r-- 1 zsombor zsombor 778 2009-03-19 12:11 usercert-penny.pem
--rw-r--r-- 1 zsombor zsombor 963 2009-03-19 12:11 userkey-billy.pem
--rw-r--r-- 1 zsombor zsombor 963 2009-03-19 12:11 userkey-hammer.pem
--rw-r--r-- 1 zsombor zsombor 951 2009-03-19 12:11 userkey-penny.pem
-\end{verbatim}
-
-You can see the certificate and key files for all the hosts and users, and the CA file with the proper hashed name.
+-rw-r--r-- 1 zsombor zsombor 769 2009-10-16 13:31 usercert-billy.pem
+-rw-r--r-- 1 zsombor zsombor 774 2009-10-16 13:31 usercert-hammer.pem
+-rw-r--r-- 1 zsombor zsombor 769 2009-10-16 13:31 usercert-penny.pem
+-rw-r--r-- 1 zsombor zsombor 963 2009-10-16 13:31 userkey-billy.pem
+-rw-r--r-- 1 zsombor zsombor 963 2009-10-16 13:31 userkey-hammer.pem
+-rw-r--r-- 1 zsombor zsombor 963 2009-10-16 13:31 userkey-penny.pem
+\end{verbatim}
+
+Here can be seen the certificate and key files for all the hosts and users, and the CA file with the proper hashed name.
 There is no rule about where to put these certificate, but it is common to put them in \verb!/etc/grid-security!.
-Now, create this directory, and a subdirectory called \verb!certificates!.
-To use the `theta' host certificates for this machine, copy the corresponding certificate and key there:
-
-\begin{verbatim}
-$ sudo cp hostCerts/hostcert-theta.pem /etc/grid-security/hostcert.pem
-$ sudo cp hostCerts/hostkey-theta.pem /etc/grid-security/hostkey.pem
-$ sudo cp CA/* /etc/grid-security/certificates/
-$ ls -lR /etc/grid-security/
-/etc/grid-security/:
-total 12
-drwxr-xr-x 2 root root 4096 2009-03-19 12:36 certificates
--rw-r--r-- 1 root root  782 2009-03-19 12:35 hostcert.pem
--rw-r--r-- 1 root root  891 2009-03-19 12:36 hostkey.pem
-
-/etc/grid-security/certificates:
+
+In these examples we assume that the CA certificate is put into \verb!/etc/grid-security/certificates!,
+
+and the host's certificate and key is at \verb!/etc/grid-security/hostcert-chelonia[123].pem!
+
+and \verb!/etc/grid-security/hostkey-chelonia[123].pem!.
+
+\begin{verbatim}
+$ ls -l /etc/grid-security/certificates/
 total 8
--rw-r--r-- 1 root root 786 2009-03-19 12:36 66fe707c.0
--rw-r--r-- 1 root root 139 2009-03-19 12:36 66fe707c.signing_policy
+-rw-r--r-- 1 root root 778 2009-10-16 14:55 afc5f10f.0
+-rw-r--r-- 1 root root 135 2009-10-16 14:55 afc5f10f.signing_policy
 \end{verbatim}
@@
 $ mkdir ~/.arc
-$ cp userCerts/usercert-billy.pem ~/.arc/usercert.pem
-$ openssl rsa -in userCerts/userkey-billy.pem -out ~/.arc/userkey.pem
-$ chmod 600 ~/.arc/userkey.pem
+$ cp userCerts/usercert-billy.pem ~/.arc
+$ openssl rsa -in userCerts/userkey-billy.pem -out ~/.arc/userkey-billy.pem
+$ chmod 600 ~/.arc/userkey-billy.pem
 $ ls -l ~/.arc
 total 8
--rw-r--r-- 1 zsombor zsombor 778 2009-03-19 13:06 usercert.pem
--rw------- 1 zsombor zsombor 891 2009-03-19 13:09 userkey.pem
-\end{verbatim}
-
-The storage system runs within the \verb!arched! hosting environment daemon, which needs a configuration file describing which services to be run, on which ports do to listen, etc.
-Copy the template configuration files to the \verb!/etc/arc! directory:
-
-\begin{verbatim}
-$ sudo mkdir /etc/arc
-$ sudo cp arc1/src/services/storage/storage_service.xml.example /etc/arc/storage_service.xml
-$ sudo cp arc1/src/services/storage/clientsslconfig.xml /etc/arc
-$ ls -l /etc/arc
-total 12
--rw-r--r-- 1 root root  259 2009-03-19 13:26 clientsslconfig.xml
--rw-r--r-- 1 root root 4606 2009-03-19 13:25 storage_service.xml
-\end{verbatim}
-
-The template configuration file specifies several directories where the services can store their data. Create these directories at, e.g., \verb!/var/spool/arc!.
-
-\begin{verbatim}
-$ sudo mkdir /var/spool/arc/ahash_data
-$ sudo mkdir /var/spool/arc/shepherd_data
-$ sudo mkdir /var/spool/arc/shepherd_store
-$ sudo mkdir /var/spool/arc/shepherd_transfer
-$ ls -l /var/spool/arc
-total 16
-drwxr-xr-x 2 root root 4096 2009-03-19 13:42 ahash_data
-drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_data
-drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_store
-drwxr-xr-x 2 root root 4096 2009-03-19 13:42 shepherd_transfer
-\end{verbatim}
-
-In this deployment the daemon will be run as root, so it is OK that the root owns these directories. 
-If creating these directories somewhere else, then modify the paths in the \verb!storage_service.xml! file.
-
-If putting the host certificates and the CA certificate somewhere else than \verb!/etc/grid-security!
-then modify the \verb!storage_service.xml!
-(look for \verb!KeyPath!, \verb!CertificatePath!
-and \verb!CACertificatesDir!).
-Also, modify \verb!/etc/arc/clientsslconfig.xml!.
-This file contains the credentials which are used by the services when they connect to an other service.
-If putting \verb!clientsslconfig.xml!
-somewhere else than \verb!/etc/arc!, you need to modify all the \verb!ClientSSLConfig!
-tags in the \verb!storage_service.xml!.
-
-Now, run the \verb!arched! daemon with this config (specified by the \verb!-c! option) first in the foreground (using the \verb!-f! option) to see immediately if everything is right:
+-rw-r--r-- 1 zsombor zsombor 778 2009-10-16 14:55 usercert-billy.pem
+-rw------- 1 zsombor zsombor 891 2009-10-16 14:55 userkey-billy.pem
+\end{verbatim}
+
+The storage system runs within the \verb!arched!
+hosting environment daemon, which needs a configuration file describing which services to be run, on which ports do to listen, etc.
+There are several profiles for different deployment scenarios, these profiles are installed by default in the \verb!/usr/local/share/arc/profiles!
+directory\footnote{or they can be acquired from http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/profiles}.
+First we will use the profile called \verb!CheloniaAllServicesCentralizedAHashWithISIS!.
+After we have chosen a profile, we need to create a config file which contains the deployment-specific attributes.
+Let's create a file in \verb!/etc/arc!
+called \verb!chelonia1.ini!:
+
+\begin{verbatim}
+profile=/usr/local/share/arc/profiles/CheloniaAllServicesCentralizedAHashWithISIS.xml
+
+pidfile=/tmp/chelonia1.pid
+logfile=/var/log/chelonia1.log
+debug=ERROR
+
+port=60001
+
+cacert=/etc/grid-security/certificates
+host_cert=/etc/grid-security/hostcert-chelonia1.pem
+host_key=/etc/grid-security/hostkey-chelonia1.pem
+client_cert=/etc/grid-security/hostcert-chelonia1.pem
+client_key=/etc/grid-security/hostkey-chelonia1.pem
+
+[register]
+index_server1=https://chelonia1.knowarc.eu:60001/ISIS
+
+[ahash]
+endpoint=https://chelonia1.knowarc.eu:60001/AHash
+datadir=/var/spool/arc/chelonia1_ahash_data
+
+[bartender]
+endpoint=https://chelonia1.knowarc.eu:60001/Bartender
+
+[librarian]
+endpoint=https://chelonia1.knowarc.eu:60001/Librarian
+
+[shepherd]
+endpoint=https://chelonia1.knowarc.eu:60001/Shepherd
+turlprefix=https://chelonia1.knowarc.eu:60001/Hopi
+datadir=/var/spool/arc/chelonia1_shepherd_data
+storedir=/var/spool/arc/chelonia1_shepherd_store
+transferdir=/var/spool/arc/chelonia1_shepherd_transfer
+
+[isis]
+endpoint=https://chelonia1.knowarc.eu:60001/ISIS
+dbpath=/var/spool/arc/chelonia1_isis_data
+peerurl1=https://chelonia1.knowarc.eu:60001/ISIS
+\end{verbatim}
+
+In this config we configure several services, and we have some common parts as well.
+The first line points to the profile which we want to use (please make sure that it is there or specify the correct path).
+The common parts contains the logfile, the level of debugging, the port to listen, the path to the certificates to be used on the service side, and the certificates to be used when a service wants to connect to an other service (in current case they are the same).
+Then we have an ISIS service which stores information about the running services and it is needed for the services to find eachother.
+The register section tells all the other services where to find the ISIS service.
+And then we have all the storage services: the Bartender, the Librarian, the Shepherd and the A-Hash.
+The endpoints of the services should be specified for all services.
+There is one additional service which doesn't need extra configuration: the Hopi service is a lightweight HTTP server.
+The Shepherd service is connected to the Hopi service, that's why it is needed to specify the Hopi's URL in the Shepherd's section (\verb!turlprefix!).
+Some of these services need to have directories to store data.
+We will use subdirectories in \verb!/var/spool/arc!, so we have to create this directory first:
+
+\begin{verbatim}
+$ sudo mkdir /var/spool/arc
+\end{verbatim}
+
+Now, run the \verb!arched!
+daemon with this INI config which is specified with the \verb!-i!
+option.
+(In this document we will run the ARC HED daemon as root.) We can first run it in the foreground (with the \verb!-f!
+flag), to see the possible initial misconfiguration errors immediately:
 
@@
 
-The template configuration file specifies a loglevel which only prints error messages, so nothing should be written on the screen now.
-The \verb!arched! daemon should now listen on the ports: 60001 for the data transfer service called Hopi, and 60000 for all the other services:
-
-\begin{verbatim}
-$ netstat -at | grep LISTEN
-tcp        0      0 *:60000                 *:*                     LISTEN     
+When we start the services, they are trying hard to find eachother, which means lots of connection related error messages in the beginning, but after a minute or so, the system should stabilize.
+(This is the case with shutdown as well, it is normal to get lots of connection related error messages).
+The \verb!arched!
+daemon should now listen no port 60001:
+
+\begin{verbatim}
+$ netstat -at | grep 60001
 tcp        0      0 *:60001                 *:*                     LISTEN     
@@
 
-Now it is time to set up the prototype client tool to access the system.
+Now it is time to set up the client tool to access the system.
 It is called \verb!chelonia!
 and it is by default installed to \verb!/usr/local/bin!.
-The CLI tool needs to be told where it can find the user credentials, and a URL of a Bartender service.
+The CLI tool needs to be told where it can find the user credentials, and a URL of one or more Bartender services or preferably the URL of one or more ISIS services, where Bartender services are registered.
 The Bartender service is the front-end of the storage system.
-This can be specified either by environment variables, or by a configuration file called \verb!~/.arc/client.xml!.
-Create the following \verb!client.xml!
-file:
-
-\begin{verbatim}
-$ cat ~/.arc/client.xml 
-<ArcConfig>
-  <KeyPath>/home/<username>/.arc/userkey.pem</KeyPath>
-  <CertificatePath>/home/<username>/.arc/usercert.pem</CertificatePath>
-  <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
-  <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
-</ArcConfig>
-\end{verbatim}
-
-Now you can use the tool \verb!chelonia! to access your local deployment. Try to do a `list' on the root collection (`/'):
-
-\begin{verbatim}
-$ chelonia
-Usage:
-  chelonia <method> [<arguments>]
-Supported methods:
-  stat - get detailed information about an entry
-  makeCollection | make | mkdir - create a collection
-  unmakeCollection | unmake | rmdir - remove an empty collection
-  list | ls - list the content of a collection
-  move | mv - move entries within the namespace
-  putFile | put - upload a file
-  getFile | get - download a file
-  delFile | del | rm - remove a file
-  modify | mod - modify metadata
-  policy | pol - modify access policy rules
-  unlink - remove a link to an entry from a collection without removing the entry itself
-  credentialsDelegation | cre - delegate credentials for using gateway
-  removeCredentials | rem - remove previously delegated credentials
-Without arguments, each method prints its own help.
-
-$ chelonia list
-Usage: list <LN> [<LN> ...]
-
-$ chelonia list /
-- Calling the Bartender's list method...
-[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544]
+These can be specified with a config file, which is called \verb!~/.arc/client.conf!
+by default:
+
+\begin{verbatim}
+defaultservices=index:ARC1:https://chelonia1.knowarc.eu:60001/ISIS
+keypath=/home/zsombor/.arc/userkey-billy.pem
+certificatepath=/home/zsombor/.arc/usercert-billy.pem
+cacertificatesdirectory=/etc/grid-security/certificates
+\end{verbatim}
+
+Now we can use the tool \verb!chelonia!
+to access the local deployment.
+If our client.conf is not at the default directory, then we can specify the path with the \verb!-z!
+flag.
+First let's try to do a `list' on the root collection (`/'):
+
+\begin{verbatim}
+zsombor@chelonia:~$ chelonia list /
+ERROR: No Bartender URL found.
+(use the '-v' flag for more details, e.g. 'chelonia -v list /')
+\end{verbatim}
+
+If we get this error message, we should call it again with the \verb!-v! verbose flag:
+\begin{verbatim}
+zsombor@chelonia:~$ chelonia -v list /
+- Using default user config.
+- The key file: /home/zsombor/.arc/userkey-billy.pem
+- The cert file: /home/zsombor/.arc/usercert-billy.pem
+- The CA dir: /etc/grid-security/certificates
+- No Bartender URL found in the config or in the environment, try to get one from ISIS:
+[2009-10-20 14:11:34] [Arc.Loader] [ERROR] [19139/149632560]
     Component tcp.client(tcp) could not be created
-[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544]
+[2009-10-20 14:11:34] [Arc.Loader] [ERROR] [19139/149632560]
     Component tls.client(tls) could not be created
-[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544]
+[2009-10-20 14:11:34] [Arc.Loader] [ERROR] [19139/149632560] 
     Component http.client(http) could not be created
-[2009-03-19 18:37:12] [Arc.Loader] [ERROR] [12491/152991544]
+[2009-10-20 14:11:34] [Arc.Loader] [ERROR] [19139/149632560] 
     Component soap.client(soap) could not be created
-ERROR: Wrong status from server.
-Maybe wrong URL: 'https://localhost:60000/Bartender'
-\end{verbatim}
-
-If you get the error `Component [...] could not be created', try to specify to the python environment the location of the python packages:
-
-\begin{verbatim}
+  - https://chelonia1.knowarc.eu:60001/ISIS - Failed to connect to ISIS.
+ERROR: No Bartender URL found.
+\end{verbatim}
+
+If we get the error `Component [...] could not be created', it is usually because python somehow doesn't find the ARC clien libraries, so let's check if we have the PYTHONPATH  variable set properly, and set it, if not:
+
+\begin{verbatim}
+$ echo $PYTHONPATH
+
 $ export PYTHONPATH=/usr/local/lib/python2.5/site-packages/
@@
 
-Now a (centralized) storage service is running, including a Hopi service, which is a basic HTTP server, for transferring files.
+Now a (centralized) Chelonia is running, including a Hopi service, which is a basic HTTP server, for transferring files.
 The Bartender service is the front-end of the system to the users, so the clients always connect to the Bartender service.
@@
 $ chelonia list /
-- Calling the Bartender's list method...
-- done in 0.34 seconds.
 '/': not found
@@
 \begin{verbatim}
-$ chelonia makeCollection /
-- Calling the Bartender's makeCollection method...
-- done in 0.81 seconds.
+$ chelonia mkdir /
 Creating collection '/': done
@@
 $ chelonia list /
-- Calling the Bartender's list method...
-- done in 0.55 seconds.
 '/': collection
@@
 
-\section{Using the prototype client tool} % (fold)
-
-Assume that Chelonia is running and you want to access it with the \verb!chelonia! tool. The \verb!~/.arc/client.xml! is set properly, the contents of the root collection can be listed:
+\section{Using the chelonia client tool} % (fold)
+
+We assume that Chelonia is running and can be accessed with the \verb!chelonia! tool. The \verb!~/.arc/client.conf! is set properly, the contents of the root collection can be listed:
 
@@
 \begin{verbatim}
-$ cat > orange
-orange
+$ echo ORANGE > orange
 $ cat orange 
-orange
+ORANGE
 
 $ chelonia put orange /
-- The size of the file is 7 bytes
-- Calculating md5 checksum...
-- done in 0.0003 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
-- Calling the Bartender's putFile method...
-- done in 1.26 seconds.
-- Got transfer URL: http://localhost:60001/hopi/7fcef91e-ac9c-658d-4def-17d52ba33cfb
-- Uploading from 'orange'
-    to 'http://localhost:60001/hopi/7fcef91e-ac9c-658d-4def-17d52ba33cfb' with http...
-    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
-- done in 0.0182 seconds.
 'orange' (7 bytes) uploaded as '/orange'.
@@
 $ chelonia list /
-- Calling the Bartender's list method...
-- done in 0.51 seconds.
 '/': collection
   orange <file>
-
 \end{verbatim}
@@
 $ chelonia stat /orange
-- Calling the Bartender's stat method...
-- done in 0.40 seconds.
 '/orange': found
@@
     checksumType: md5
-    neededReplicas: 1
+    neededReplicas: 3
     size: 7
-    checksum: 6aefd2842be62cd470709b27aedc7db7
+    checksum: a7992bd22ce079983da12c0172983ff4
   entry
     owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
-    GUID: 12ce831b-eab8-1f47-29f6-19fb85ebf46a
+    GUID: eb973cb4-7924-0663-2b3e-13c01203fcd1
     type: file
@@
   locations
-    https://localhost:60000/Shepherd 3e9ea612-87c0-1e81-58f3-cec6b0b125c1: alive
+    https://chelonia1.knowarc.eu:60001/Shepherd 630b9bf0-7b70-8e04-1098-01769af3f8ea: alive
   timestamps
-    created: 1237504821.91
-\end{verbatim}
-
-Here, the owner of the file (Billy), the location of the replicas (currently only one), the number of needed replicas (which is also only one), the checksum and size of the file, and the timestamp of the creation is shown.
+    created: 1256041232.28
+\end{verbatim}
+
+Here, the owner of the file (Billy), the location of the replicas (currently only one), the number of needed replicas (which is three), the checksum and size of the file, and the timestamp of the creation is shown.
 
@@
 $ chelonia get /orange /tmp
-- Calling the Bartender's getFile method...
-- done in 0.51 seconds.
-- Got transfer URL: http://localhost:60001/hopi/521c5869-4c93-8e28-aa3a-03f477f34f63
-- Downloading from 'http://localhost:60001/hopi/521c5869-4c93-8e28-aa3a-03f477f34f63'
-    to '/tmp/orange' with http...
-    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
-- done in 1.0071 seconds.
 '/orange' (7 bytes) downloaded as '/tmp/orange'.
+
 $ cat /tmp/orange 
-orange
-\end{verbatim}
-
-You can create collections which can contain files and other collections forming a tree-hierarchy. You can move files and collection around within this namespace.
-
-\begin{verbatim}
-$ chelonia makeCollection /fruits
-- Calling the Bartender's makeCollection method...
-- done in 0.79 seconds.
+ORANGE
+\end{verbatim}
+
+We can create collections which can contain files and other collections forming a tree-hierarchy. We can move files and collection around within this namespace.
+
+\begin{verbatim}
+$ chelonia mkdir /fruits
 Creating collection '/fruits': done
 
-$ chelonia makeCollecion /fruits/apple
-- Calling the Bartender's makeCollection method...
-- done in 0.91 seconds.
+$ chelonia mkdir /fruits/apple
 Creating collection '/fruits/apple': done
 
-$ chelonia move /orange /fruits/
-- Calling the Bartender's move method...
-- done in 0.99 seconds.
+$ chelonia mv /orange /fruits
+Moving '/orange' to '/fruits': targetexists
+\end{verbatim}
+
+Note that if we want to move something into a colleciton, we have to end the collection's logical name with a slash:
+
+\begin{verbatim}
+$ chelonia mv /orange /fruits/
 Moving '/orange' to '/fruits/': moved
 
-$ chelonia list /fruits
-- Calling the Bartender's list method...
-- done in 0.60 seconds.
+$ chelonia ls /fruits
 '/fruits': collection
@@
 
-You can remove files, and remove collections (but only if the collection is empty):
+We can remove files, and remove collections (but only if the collection is empty):
 
@@
 $ chelonia rmdir /fruits
-- Calling the Bartender's unmakeCollection method...
-- done in 0.41 seconds.
 Removing collection '/fruits': collection is not empty
@@
 $ chelonia rmdir /fruits/apple
-- Calling the Bartender's unmakeCollection method...
-- done in 0.89 seconds.
 Removing collection '/fruits/apple': removed
@@
 $ chelonia rm /fruits/orange
-- Calling the Bartender's delFile method...
-- done in 0.89 seconds.
 /fruits/orange: deleted
@@
 $ chelonia rmdir /fruits
-- Calling the Bartender's unmakeCollection method...
-- done in 0.79 seconds.
 Removing collection '/fruits': removed
@@
 
-Now, introduce a new user, Penny, into the system.
-Again, \verb!chelonia!
-can be told to use user certificate and key files for this user with modifying the \verb!client.xml!
-or by specifying the \verb!ARC_CERT_FILE!
-and \verb!ARC_KEY_FILE!
-environment variables.
-
-\begin{verbatim}
-$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
-chelonia list /
+Now, let's introduce a new user, Penny, into the system. The easiest would be to create a new client.conf for Penny:
+
+\begin{verbatim}
+$ cat /home/zsombor/.arc/penny.conf 
+defaultservices=index:ARC1:https://chelonia1.knowarc.eu:60001/ISIS
+keypath=/home/zsombor/.arc/userkey-penny.pem
+certificatepath=/home/zsombor/.arc/usercert-penny.pem
+cacertificatesdirectory=/etc/grid-security/certificates    
+\end{verbatim}
+
+We should copy her certificate and key the same way as we did with Billy's certificate and key.
+Then we can specify Penny's config with the \verb!-z!
+flag (and we can use the \verb!-v!
+flag to see what is happening:
+
+\begin{verbatim}
+$ chelonia -v -z ~/.arc/penny.conf list /
+- Config file specified: /home/zsombor/.arc/penny.conf
+- The key file: /home/zsombor/.arc/userkey-penny.pem
+- The cert file: /home/zsombor/.arc/usercert-penny.pem
+- The CA dir: /etc/grid-security/certificates
+- No Bartender URL found in the config or in the environment, try to get one from ISIS:
+  - https://chelonia1.knowarc.eu:60001/ISIS - Got Bartender URL from ISIS: https://chelonia1.knowarc.eu:60001/Bartender
+- The URL of the Bartender(s): https://chelonia1.knowarc.eu:60001/Bartender
 - Calling the Bartender's list method...
-- done in 0.31 seconds.
+- Trying https://chelonia1.knowarc.eu:60001/Bartender...
+- done in 0.15 seconds.
 '/': denied
 
-$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
-chelonia put /tmp/orange /pennys-orange
-- The size of the file is 7 bytes
-- Calculating md5 checksum...
-- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
-- Calling the Bartender's putFile method...
-- done in 0.81 seconds.
-/pennys-orange: failed to add child to parent
-
+$ chelonia -z ~/.arc/penny.conf put orange /pennys-orange
+/pennysorange: failed to add child to parent
 \end{verbatim}
@@
 \begin{verbatim}
-$ chelonia pol / change ALL +read +addEntry
-- Calling the Bartender's stat method...
-- done in 0.30 seconds.
-old action list for the user ALL was ''
-the new action list of user ALL for LN / will be '+read +addEntry'
-- Calling the Bartender's modify method...
-- done in 0.50 seconds.
-set.
+$ chelonia policy / change ALL +read +addEntry
+Setting action list of '/' for user ALL to +read +addEntry: set.
 \end{verbatim}
@@
 $ chelonia pol
+$ chelonia policy
 Usage: policy <LN> <changeType> <identity> <action list>
@@
     e.g. '+read +addEntry -delete'
-    possible actions:
-        read addEntry removeEntry delete
+    possible actions:  read addEntry removeEntry delete
                        modifyPolicy modifyStates modifyMetadata
@@
 \begin{verbatim}
-$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
-chelonia put /tmp/orange /pennys-orange
-- The size of the file is 7 bytes
-- Calculating md5 checksum...
-- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
-- Calling the Bartender's putFile method...
-- done in 1.18 seconds.
-- Got transfer URL: http://localhost:60001/hopi/4145891d-0e4c-b6bd-e4d6-dc80554b9d35
-- Uploading from '/tmp/orange'
-    to 'http://localhost:60001/hopi/4145891d-0e4c-b6bd-e4d6-dc80554b9d35' with http...
-    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
-- done in 1.0071 seconds.
-'/tmp/orange' (7 bytes) uploaded as '/pennys-orange'.
+$ chelonia -z ~/.arc/penny.conf put orange /pennys-orange
+'orange' (7 bytes) uploaded as '/pennys-orange'.
 \end{verbatim}
@@
 \begin{verbatim}
-$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
-chelonia pol /pennys-orange change /DC=eu/DC=KnowARC/O=knowarc/CN=billy +read
-- Calling the Bartender's stat method...
-- done in 0.41 seconds.
-old action list for the user /DC=eu/DC=KnowARC/O=knowarc/CN=billy was ''
-the new action list of user /DC=eu/DC=KnowARC/O=knowarc/CN=billy for LN /pennys-orange
-    will be '+read'
-- Calling the Bartender's modify method...
-- done in 0.59 seconds.
-set.
+$ chelonia get /pennys-orange /tmp
+/pennys-orange: denied
+
+$ chelonia -z ~/.arc/penny.conf pol /pennys-orange
+    change /DC=eu/DC=KnowARC/O=knowarc/CN=billy +read
+Setting action list of '/pennys-orange' for user /DC=eu/DC=KnowARC/O=knowarc/CN=billy
+    to +read: set.
 
 $ chelonia stat /pennys-orange
-- Calling the Bartender's stat method...
-- done in 0.43 seconds.
 '/pennys-orange': found
   locations
-    https://localhost:60000/Shepherd 1cc3b115-2350-8334-d206-a7b891912c00: alive
+    https://chelonia1.knowarc.eu:60001/Shepherd 22be0347-ab35-42c8-12a0-07a50160399a: alive
   states
     checksumType: md5
-    neededReplicas: 1
+    neededReplicas: 3
     size: 7
-    checksum: 6aefd2842be62cd470709b27aedc7db7
+    checksum: a7992bd22ce079983da12c0172983ff4
   parents
@@
   timestamps
-    created: 1237509633.17
+    created: 1256044171.04
   policy
@@
     owner: /DC=eu/DC=KnowARC/O=knowarc/CN=penny
-    GUID: 0e6a4c9a-d4e7-e0dc-e083-16b83d460d82
+    GUID: ea4fdb99-6a13-e926-0ec5-093e9fac524e
     type: file
@@
 $ chelonia get /pennys-orange /tmp
-- Calling the Bartender's getFile method...
-- done in 0.50 seconds.
-- Got transfer URL: http://localhost:60001/hopi/c16d7e12-42b0-7882-c09e-1cac31dbdbcd
-- Downloading from 'http://localhost:60001/hopi/c16d7e12-42b0-7882-c09e-1cac31dbdbcd'
-    to '/tmp/pennys-orange' with http...
-    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
-- done in 1.0065 seconds.
 '/pennys-orange' (7 bytes) downloaded as '/tmp/pennys-orange'.
+
 $ cat /tmp/pennys-orange 
-orange
-\end{verbatim}
-
-The needed number of replicas for a new file can be specified with the \verb!ARC_NEEDED_REPLICAS! environment variable.
-
-\begin{verbatim}
-$ ARC_NEEDED_REPLICAS=2 chelonia put orange /new-orange
-- The size of the file is 7 bytes
-- Calculating md5 checksum...
-- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
-- Calling the Bartender's putFile method...
-- done in 1.11 seconds.
-- Got transfer URL: http://localhost:60001/hopi/0ad54182-7618-a2af-c490-dea7c4d6fc31
-- Uploading from 'orange' to
-    'http://localhost:60001/hopi/0ad54182-7618-a2af-c490-dea7c4d6fc31' with http...
-    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
-- done in 1.0057 seconds.
-'orange' (7 bytes) uploaded as '/new-orange'.
-
-$ chelonia stat /new-orange
-- Calling the Bartender's stat method...
-- done in 0.40 seconds.
-'/new-orange': found
-  states
-    checksumType: md5
+ORANGE
+\end{verbatim}
+
+Currently we only have one storage element (a Shepherd service and a Hopi service together), that's why the file has only one replica, even though it should have 3.
+However the number of needed replicas of an existing file can be modified with the `modify' method:
+
+\begin{verbatim}
+$ chelonia stat /pennys-orange | grep neededReplicas
+    neededReplicas: 3
+
+$ chelonia -z ~/.arc/penny.conf modify /pennys-orange set states neededReplicas 2
+set
+
+$ chelonia stat /pennys-orange | grep neededReplicas
     neededReplicas: 2
-    size: 7
-    checksum: 6aefd2842be62cd470709b27aedc7db7
-  entry
-    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
-    GUID: 4e673e28-4769-50af-f02e-d7a67632f87a
-    type: file
-  parents
-    0/new-orange: parent
-  locations
-    https://localhost:60000/Shepherd 4aacd276-653c-3ca5-c1a2-78a876c129c9: alive
-  timestamps
-    created: 1237509989.43
-
-\end{verbatim}
-
-Here, the `neededReplica' metadata says 2, but the system is deployed in a centralized way for testing purposes, so there is only one storage element (which is a Shepherd service and a Hopi service together), so the file can only have one replica and there can only be one location.
-
-The number of needed replicas of an existing file can be modified with the `modify' method:
-
-\begin{verbatim}
-$ chelonia modify /new-orange set states neededReplicas 1
-- Calling the Bartender's modify method...
-- done in 0.63 seconds.
-set
-
-$ chelonia stat /new-orange | grep neededReplicas
-    neededReplicas: 1
 \end{verbatim}
@@
 But for testing purposes it is of course possible to run it on the same machine using different ports.
-A storage element consists of a Shepherd service and a storage element service, which this time will be a `byteio' service.
-The `byteio' service implements a small part of the ByteIO interface, which means that it just put the entire file data into the SOAP message, so it is only suitable for small files.
-
-Now, modify the configuration file of the original server.
-Currently all the transfer URLs which were generated when uploading and downloading files contained the hostname `localhost'.
-This is not OK if access from an other machine is needed.
-In this case, transfer URL's must contain the external hostname of the machine.
-To do this, change this line in \verb!/etc/arc/storage_service.xml!:
-
-\begin{verbatim}
-    <TURLPrefix>http://localhost:60001/hopi/</TURLPrefix>
-\end{verbatim}
-
-to the real hostname of the machine.
-But this is not enough.
-Currently the Shepherd services uses their URLs as IDs, and the other services use these IDs to connect to a Shepherd service, so the ID of the Shepherd service needs to be changed to a real external URL as well.
-For the sake of clarity, change all `localhost' entries in the config file to the real hostname.
-In this example, the name of this machine is just `storage' so the Shepherd section should look like this:
-
-\begin{verbatim}
-<Service name="pythonservice" id="shepherd">
-    <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
-    <ServiceID>https://storage:60000/Shepherd</ServiceID>
-    <CheckPeriod>20</CheckPeriod>
-    <MinCheckInterval>0.1</MinCheckInterval>
-    <CreatingTimeout>600</CreatingTimeout>
-    <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
-    <StoreCfg><DataDir>/var/spool/arc/shepherd_data</DataDir></StoreCfg>
-    <BackendClass>storage.shepherd.hardlinkingbackend.HopiBackend</BackendClass>
-    <BackendCfg>
-        <DataDir>/var/spool/arc/shepherd_store</DataDir>
-        <TransferDir>/var/spool/arc/shepherd_transfer</TransferDir>
-        <TURLPrefix>http://storage:60001/hopi/</TURLPrefix>
-    </BackendCfg>
-    <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
-    <BartenderURL>https://storage:60000/Bartender</BartenderURL>
-    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
-</Service>
-\end{verbatim}
-
-Now, check if the centralized deployment is still working, stop the \verb!arched! daemon, remove the data directories to make it a clean start, and restart it with the modified config.
-
-\begin{verbatim}
-$ sudo rm -r /var/spool/arc/*
-$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service.xml -f    
-\end{verbatim}
-
-See if the client is still usable:
-
-\begin{verbatim}
-$ chelonia list /
-- Calling the Bartender's list method...
-- done in 0.31 seconds.
-'/': not found
-
-$ chelonia makeCollection /
-- Calling the Bartender's makeCollection method...
-- done in 0.63 seconds.
-Creating collection '/': done
-
-$ chelonia list /
-- Calling the Bartender's list method...
-- done in 0.50 seconds.
-'/': collection
-    empty.
-\end{verbatim}
-
-Now, upload a file and ask for two replicas:
-
-\begin{verbatim}
-$ ARC_NEEDED_REPLICAS=2 chelonia put orange /
-- The size of the file is 7 bytes
-- Calculating md5 checksum...
-- done in 0.0005 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
-- Calling the Bartender's putFile method...
-- done in 1.10 seconds.
-- Got transfer URL: http://storage:60001/hopi/139dc57d-8c12-c188-a4a6-973449553811
-- Uploading from 'orange'
-    to 'http://storage:60001/hopi/139dc57d-8c12-c188-a4a6-973449553811' with http...
-    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
-- done in 1.0090 seconds.
-'orange' (7 bytes) uploaded as '/orange'.
-
-$ chelonia stat /orange
-- Calling the Bartender's stat method...
-- done in 0.41 seconds.
-'/orange': found
-  states
-    checksumType: md5
-    neededReplicas: 2
-    size: 7
-    checksum: 6aefd2842be62cd470709b27aedc7db7
-  entry
-    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
-    GUID: 3fdad88a-944f-79c5-4f23-c5b3d3bc981e
-    type: file
-  parents
-    0/orange: parent
+A storage element consists of a Shepherd service and a storage element service, which will be our lightweight HTTP server called Hopi.
+(But there is support for using apache as well.)
+
+For our second machine we will use the profile called \verb!CheloniaShepherdWithHopi!, which only contains the Shepherd service and the Hopi service. Here is the INI config for our chelonia2 server:
+
+\begin{verbatim}
+profile=/usr/local/share/arc/profiles/CheloniaShepherdWithHopi.xml
+
+pidfile=/tmp/chelonia2.pid
+logfile=/var/log/chelonia2.log
+debug=ERROR
+
+port=60002
+
+cacert=/etc/grid-security/certificates
+host_cert=/etc/grid-security/hostcert-chelonia2.pem
+host_key=/etc/grid-security/hostkey-chelonia2.pem
+client_cert=/etc/grid-security/hostcert-chelonia2.pem
+client_key=/etc/grid-security/hostkey-chelonia2.pem
+
+[register]
+index_server1=https://chelonia1.knowarc.eu:60001/ISIS
+
+[shepherd]
+endpoint=https://chelonia2.knowarc.eu:60002/Shepherd
+turlprefix=https://chelonia2.knowarc.eu:60002/Hopi
+datadir=/var/spool/arc/chelonia2_shepherd_data
+storedir=/var/spool/arc/chelonia2_shepherd_store
+transferdir=/var/spool/arc/chelonia2_shepherd_transfer
+\end{verbatim}
+
+The Hopi service doesn't need additional configuration, it is tied to the Shepherd here.
+We have to specify the ISIS running on the \verb!chelonia1!
+server in order to the services in the two container could find eachother.
+
+While our first server (\verb!chelonia1!) is still running, start this new server:
+
+\begin{verbatim}
+$ sudo arched -i /etc/arc/chelonia2.ini
+$ cat /tmp/chelonia2.pid
+21477
+\end{verbatim}
+
+Now in a few minutes the file should have a second replica:
+
+\begin{verbatim}
+$ chelonia stat /pennys-orange
+'/pennys-orange': found
   locations
-    https://storage:60000/Shepherd 0b46f53c-419d-ce4e-181a-ec95d16dcc95: alive
-  timestamps
-    created: 1237519441.96
-\end{verbatim}
-
-Currently it has only one replica, because there is still only one storage element. Now, create a new config for the other machine, here called \verb!storage2!:
-
-\begin{verbatim}
-<?xml version="1.0"?>
-<ArcConfig 
-  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
-  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
->
-    <Server>
-        <Pidfile>/var/run/arched2.pid</Pidfile>
-        <Logger level="ERROR">/var/log/arched2.log</Logger>
-    </Server>
-    <ModuleManager>
-        <Path>/usr/local/lib/arc/</Path>
-    </ModuleManager>
-    <Plugins><Name>mcctcp</Name></Plugins>
-    <Plugins><Name>mcctls</Name></Plugins>
-    <Plugins><Name>mcchttp</Name></Plugins>
-    <Plugins><Name>mccsoap</Name></Plugins>
-    <Chain>
-        <Component name="tcp.service" id="tcp">
-            <next id="tls"/> 
-            <tcp:Listen>
-                <tcp:Port>60006</tcp:Port>
-                <tcp:Version>4</tcp:Version>
-            </tcp:Listen>
-        </Component>
-        <Component name="tls.service" id="tls">
-            <next id="http"/>
-            <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
-            <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
-            <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
-        </Component>
-        <Component name="http.service" id="http">
-            <next id="soap">POST</next>
-        </Component>
-        <Component name="soap.service" id="soap">
-            <next id="plexer"/>
-        </Component>
-        <Plexer name="plexer.service" id="plexer">
-            <next id="shepherd">^/Shepherd$</next>
-            <next id="byteio">^/byteio/</next>
-        </Plexer>
-        <Service name="pythonservice" id="shepherd">
-            <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
-            <ServiceID>https://storage2:60006/Shepherd</ServiceID>
-            <CheckPeriod>20</CheckPeriod>
-            <MinCheckInterval>0.1</MinCheckInterval>
-            <CreatingTimeout>600</CreatingTimeout>
-            <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
-            <StoreCfg>
-                <DataDir>/var/spool/arc/shepherd_data2</DataDir>
-            </StoreCfg>
-            <BackendClass>storage.shepherd.byteio.ByteIOBackend</BackendClass>
-            <BackendCfg>
-                <DataDir>/var/spool/arc/shepherd_store2</DataDir>
-                <TransferDir>/var/spool/arc/shepherd_transfer2</TransferDir>
-                <TURLPrefix>https://storage2:60006/byteio/</TURLPrefix>
-            </BackendCfg>
-            <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
-            <BartenderURL>https://storage:60000/Bartender</BartenderURL>
-            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml"/>
-        </Service>
-        <Service name="pythonservice" id="byteio">
-            <ClassName>storage.shepherd.byteio.ByteIOService</ClassName>
-            <TransferDir>/var/spool/arc/shepherd_transfer2</TransferDir>
-            <NotifyURL>https://storage2:60006/Shepherd</NotifyURL>
-            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml"/>
-        </Service>
-    </Chain>
-</ArcConfig>    
-\end{verbatim}
-
-This configuration contains only two services: a Shepherd, and byteio service.
-The byteio service is just a very simple storage element service which moves files within SOAP messages.
-The Shepherd service needs to have a backend for all supported storage element services, so it has a ByteIOBackend configured.
-This Shepherd connects to the first server's Bartender and Librarian services.
-(See the technical documentation for details about the services.)
-
-While the original (\verb!storage!) server still runs the services, start the new server:
-
-\begin{verbatim}
-$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service_2.xml -f
-\end{verbatim}
-
-After a minute or two, check the file again:
-
-\begin{verbatim}
-$ chelonia stat /orange
-- Calling the Bartender's stat method...
-- done in 0.41 seconds.
-'/orange': found
-  states
-    checksumType: md5
-    neededReplicas: 2
-    size: 7
-    checksum: 6aefd2842be62cd470709b27aedc7db7
-  entry
-    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
-    GUID: 24b442eb-ab1a-680e-1f79-4c20479d3ea9
-    type: file
-  parents
-    0/orange: parent
+    https://chelonia1.knowarc.eu:60001/Shepherd 22be0347-ab35-42c8-12a0-07a50160399a: alive
+    https://chelonia2.knowarc.eu:60002/Shepherd 386aec8c-174c-c83c-e5ea-f0784f5ff79f: alive
+[...]
+\end{verbatim}
+
+Let's start a third server called \verb!chelonia3! with this config:
+
+\begin{verbatim}
+profile=/usr/local/share/arc/profiles/CheloniaShepherdWithHopi.xml
+
+pidfile=/tmp/chelonia3.pid
+logfile=/var/log/chelonia3.log
+debug=ERROR
+
+port=60003
+
+cacert=/etc/grid-security/certificates
+host_cert=/etc/grid-security/hostcert-chelonia3.pem
+host_key=/etc/grid-security/hostkey-chelonia3.pem
+client_cert=/etc/grid-security/hostcert-chelonia3.pem
+client_key=/etc/grid-security/hostkey-chelonia3.pem
+
+[register]
+index_server1=https://chelonia1.knowarc.eu:60001/ISIS
+
+[shepherd]
+endpoint=https://chelonia3.knowarc.eu:60003/Shepherd
+turlprefix=https://chelonia3.knowarc.eu:60003/Hopi
+datadir=/var/spool/arc/chelonia3_shepherd_data
+storedir=/var/spool/arc/chelonia3_shepherd_store
+transferdir=/var/spool/arc/chelonia3_shepherd_transfer
+\end{verbatim}
+
+It again has the same ISIS configured. Let's start it:
+
+\begin{verbatim}
+$ sudo arched -i /etc/arc/chelonia3.ini
+\end{verbatim}
+
+Now when checking the file again there should be no change at all:
+
+\begin{verbatim}
+$ chelonia stat /pennys-orange
+'/pennys-orange': found
   locations
-    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
-    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
-  timestamps
-    created: 1237519828.4
-\end{verbatim}
-
-It indeed has two replicas now, one on each server.
-
-Now, try to download it multiple times:
-
-\begin{verbatim}
-$ chelonia get /orange /tmp/orange1
-- Calling the Bartender's getFile method...
-- done in 0.50 seconds.
-- Got transfer URL: http://storage:60001/hopi/de0d951f-7379-1ce4-0813-48a384c11982
-- Downloading from 'http://storage:60001/hopi/de0d951f-7379-1ce4-0813-48a384c11982'
-    to '/tmp/orange1' with http...
-    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
-- done in 1.0060 seconds.
-'/orange' (7 bytes) downloaded as '/tmp/orange1'.
-
-$ chelonia get /orange /tmp/orange2
-- Calling the Bartender's getFile method...
-- done in 0.51 seconds.
-- Got transfer URL: https://storage2:60006/byteio/52b6bffa-4812-03a5-93a1-834e38077ac5
-- Downloading from 'https://storage2:60006/byteio/52b6bffa-4812-03a5-93a1-834e38077ac5'
-    to '/tmp/orange2' with byteio...
-- done in 0.2822 seconds.
-'/orange' (7 bytes) downloaded as '/tmp/orange2'.
-\end{verbatim}
-
-First it was downloaded from the \verb!storage! server via HTTP, second time it was downloaded from the \verb!storage2! server via ByteIO.
-
-Now, create a third storage element on a machine called \verb!storage3! this time using a Hopi service as storage element service like on the main \verb!storage! server:
-
-\begin{verbatim}
-zsombor@storage:~$ cat /etc/arc/storage_service_3.xml 
-<?xml version="1.0"?>
-<ArcConfig 
-  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
-  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
->
-    <Server>
-        <Pidfile>/var/run/arched3.pid</Pidfile>
-        <Logger level="ERROR">/var/log/arched3.log</Logger>
-    </Server>
-    <ModuleManager>
-        <Path>/usr/local/lib/arc/</Path>
-    </ModuleManager>
-    <Plugins><Name>mcctcp</Name></Plugins>
-    <Plugins><Name>mcctls</Name></Plugins>
-    <Plugins><Name>mcchttp</Name></Plugins>
-    <Plugins><Name>mccsoap</Name></Plugins>
-    <Chain>
-        <Component name="tcp.service" id="tcp">
-            <next id="tls"/> 
-            <tcp:Listen>
-		<tcp:Port>60010</tcp:Port>
-		<tcp:Version>4</tcp:Version>
-	    </tcp:Listen>
-        </Component>
-        <Component name="tls.service" id="tls">
-            <next id="http"/>
-            <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
-            <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
-            <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
-        </Component>
-        <Component name="http.service" id="http">
-            <next id="soap">POST</next>
-        </Component>
-        <Component name="soap.service" id="soap">
-            <next id="plexer"/>
-        </Component>
-        <Plexer name="plexer.service" id="plexer">
-            <next id="shepherd">^/Shepherd$</next>
-        </Plexer>
-        <Service name="pythonservice" id="shepherd">
-            <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
-            <ServiceID>https://storage3:60010/Shepherd</ServiceID>
-            <CheckPeriod>20</CheckPeriod>
-            <MinCheckInterval>0.1</MinCheckInterval>
-            <CreatingTimeout>600</CreatingTimeout>
-            <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
-            <StoreCfg><DataDir>/var/spool/arc/shepherd_data3</DataDir></StoreCfg>
-            <BackendClass>storage.shepherd.hardlinkingbackend.HopiBackend</BackendClass>
-            <BackendCfg>
-                <DataDir>/var/spool/arc/shepherd_store3</DataDir>
-                <TransferDir>/var/spool/arc/shepherd_transfer3</TransferDir>
-                <TURLPrefix>http://storage3:60011/hopi/</TURLPrefix>
-            </BackendCfg>
-            <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
-            <BartenderURL>https://storage:60000/Bartender</BartenderURL>
-            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
-        </Service>
-    </Chain>
-    <Chain>
-        <Component name="tcp.service" id="tcp2">
-            <next id="http2"/> 
-            <tcp:Listen>
-		<tcp:Port>60011</tcp:Port>
-		<tcp:Version>4</tcp:Version>
-	    </tcp:Listen>
-        </Component>
-        <Component name="http.service" id="http2">
-            <next id="plexer2">GET</next>
-            <next id="plexer2">PUT</next>
-        </Component>
-        <Plexer name="plexer.service" id="plexer2">
-            <next id="hopi">^/hopi/</next>
-        </Plexer>
-        <Service name="hopi" id="hopi">
-            <DocumentRoot>/var/spool/arc/shepherd_transfer3</DocumentRoot>
-            <SlaveMode>1</SlaveMode>
-        </Service>
-    </Chain>
-</ArcConfig>    
-\end{verbatim}
-
-This configuration has two separate ports, one with TLS security and a Shepherd service, and the other without TLS security and with the Hopi service. Run this server as well:
-
-\begin{verbatim}
-$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service_3.xml -f
-\end{verbatim}
-
-Now when checking the file again there should be no change at all:
-
-\begin{verbatim}
-$ chelonia stat /orange
+    https://chelonia1.knowarc.eu:60001/Shepherd 22be0347-ab35-42c8-12a0-07a50160399a: alive
+    https://chelonia2.knowarc.eu:60002/Shepherd 386aec8c-174c-c83c-e5ea-f0784f5ff79f: alive
+[...]
+\end{verbatim}
+
+Let's stop \verb!chelonia2! to see what is happening with the two replicas:
+
+\begin{verbatim}
+$ chelonia stat /pennys-orange
+'/pennys-orange': found
   locations
-    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
-    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
-\end{verbatim}
-
-Stop the server on \verb!storage2!, and watch what happens with the file:
-
-\begin{verbatim}
-$ chelonia stat /orange
+    https://chelonia1.knowarc.eu:60001/Shepherd 22be0347-ab35-42c8-12a0-07a50160399a: alive
+    https://chelonia2.knowarc.eu:60002/Shepherd 386aec8c-174c-c83c-e5ea-f0784f5ff79f: offline
+[...]
+
+$ chelonia stat /pennys-orange
+'/pennys-orange': found
   locations
-    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
-    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline
-
-$ chelonia stat /orange
+    https://chelonia1.knowarc.eu:60001/Shepherd 22be0347-ab35-42c8-12a0-07a50160399a: alive
+    https://chelonia3.knowarc.eu:60003/Shepherd 4a5cb7d5-bcec-0c84-e37f-3d2658ed3d4c: alive
+    https://chelonia2.knowarc.eu:60002/Shepherd 386aec8c-174c-c83c-e5ea-f0784f5ff79f: offline
+[...]
+
+\end{verbatim}
+
+A new replica has been created on the third server to maintain the needed number of 2.
+
+Let's start \verb!chelonia2! again, and check the state of the replicas:
+
+\begin{verbatim}
+$ chelonia stat /pennys-orange
+'/pennys-orange': found
   locations
-    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
-    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline
-    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: creating
-
-$ chelonia stat /orange
+    https://chelonia1.knowarc.eu:60001/Shepherd 22be0347-ab35-42c8-12a0-07a50160399a: alive
+    https://chelonia3.knowarc.eu:60003/Shepherd 4a5cb7d5-bcec-0c84-e37f-3d2658ed3d4c: alive
+    https://chelonia2.knowarc.eu:60002/Shepherd 386aec8c-174c-c83c-e5ea-f0784f5ff79f: alive
+
+$ chelonia stat /pennys-orange
+'/pennys-orange': found
   locations
-    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
-    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline
-    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
-\end{verbatim}
-
-A new replica has been created on the third server to maintain the needed number of 2.
-
-Now, start the services on \verb!storage2! again, and see what is happening with the file:
-
-\begin{verbatim}
-$ chelonia stat /orange
+    https://chelonia1.knowarc.eu:60001/Shepherd 22be0347-ab35-42c8-12a0-07a50160399a: alive
+    https://chelonia3.knowarc.eu:60003/Shepherd 4a5cb7d5-bcec-0c84-e37f-3d2658ed3d4c: alive
+    https://chelonia2.knowarc.eu:60002/Shepherd 386aec8c-174c-c83c-e5ea-f0784f5ff79f: thirdwheel
+
+$ chelonia stat /pennys-orange
+'/pennys-orange': found
   locations
-    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
-    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
-    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
-
-$ chelonia stat /orange
-  locations
-    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
-    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: thirdwheel
-    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
-
-$ chelonia stat /orange
-  locations
-    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
-    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
+    https://chelonia1.knowarc.eu:60001/Shepherd 22be0347-ab35-42c8-12a0-07a50160399a: alive
+    https://chelonia3.knowarc.eu:60003/Shepherd 4a5cb7d5-bcec-0c84-e37f-3d2658ed3d4c: alive
 \end{verbatim}
@@
 One of the replicas has been removed to maintain the needed number of 2.
-
-\section{Make the A-Hash service replicated} % (fold)
+Sometimes it takes a while while the Shepherds figure out how should remove the replica to make sure we don't loose the file accidentaly.
+
+\section{Make the full system replicated} % (fold)
 \label{sec:make_the_a_hash_service_replicated}
@@
 The replicated version of the A-Hash uses the Berkeley DB to provide a distributed database.
-
-In the following, Berkeley DB is installed as root (the \verb!--prefix!
-option of \verb!configure!
-can be used to install it somewhere else than the default location).
-This needs to be done on all the machines.
-
-Get the latest Berkeley DB and its patches:
-\begin{verbatim}
-$ wget http://download.oracle.com/berkeley-db/db-4.7.25.tar.gz
-$ wget http://www.oracle.com/technology/products/berkeley-db/db/update/4.7.25/patch.4.7.25.1
-$ wget http://www.oracle.com/technology/products/berkeley-db/db/update/4.7.25/patch.4.7.25.2
-$ wget http://www.oracle.com/technology/products/berkeley-db/db/update/4.7.25/patch.4.7.25.3
-\end{verbatim}
-
-Extract it and apply the patches:
-\begin{verbatim}
-$ tar zxf db-4.7.25.tar.gz
-$ cd db-4.7.25
-$ patch -p 0 < ../patch.4.7.25.1
-$ patch -p 0 < ../patch.4.7.25.2
-$ patch -p 0 < ../patch.4.7.25.3
-\end{verbatim}
-
-Configure, make and install it (and if needed, use the \verb!--prefix! option for \verb!configure!)
-\begin{verbatim}
-$ cd build_unix/
-$ ../dist/configure
-\end{verbatim}
-
-Now compile and install it:
-\begin{verbatim}
-$ make
-$ sudo make install
-\end{verbatim}
-
-Next, get the python wrapper for the Berkeley DB:
-
-\begin{verbatim}
-$ wget http://pypi.python.org/packages/source/b/bsddb3/bsddb3-4.7.5.tar.gz
-\end{verbatim}
-
-Install it to the default location:
-
-\begin{verbatim}
-$ tar zxf bsddb3-4.7.5.tar.gz
-$ cd bsddb3-4.7.5
-$ sudo python setup.py install
-\end{verbatim}
-
-If Berkeley DB was installed in a non-default location, or to install the python wrapper to somewhere else than the default:
-
-\begin{verbatim}
-$ BERKELEYDB_DIR=/path/to/berkeleydb python setup.py install --prefix=/where/to/install
-\end{verbatim}
-
-Test the installation with the \verb!test.py! included in the directory of the python wrapper for the Berkeley DB:
-\begin{verbatim}
-$ sudo python test.py
+But the A-Hash is just the database backend of the Librarian, so we need more than one Librarian as well.
+And the Bartender service is critical for the user to access the system, so we should have more than one Bartender services.
+And because of all the services are registered in the ISIS service, and they find eachother by querying the ISIS, that's why we need more than one ISIS as well.
+So let's deploy every service on each machine.
+
+While the Librarian, Bartender and ISIS services could be easily deployed on multiple sites (the ISIS services should have eachother URLs in the config), the A-Hash services should form a consistent replicated database together, and this is implemented in a different version of the A-Hash which has different configuration as well.
+So if we want to use the replicated A-Hash, we need a different configuration profile on each servers, called \verb!CheloniaAllServicesReplicatedAHashWithISIS!.
+
+Let's stop all the arched daemons, and remove everything from the /var/spool/arc directory, which clears all the metadata and files stored in the system (because the replicated A-Hash and the centralized A-Hash cannot be mixed together, and currently cannot be converted).
+
+We need to install at least the 4.6 version of Berkeley DB, and the python bindings as well.
+Maybe the easiest way to install the python bindings is to use \verb!easy_install!
+which is part of \verb!python-setuptools!:
+
+\begin{verbatim}
+$ sudo easy_install bsddb3    
+\end{verbatim}
+
+If we managed to install it, we should be able to run the unittests without a problem:
+
+\begin{verbatim}
+$ sudo python /usr/lib/python2.5/site-packages/\
+    bsddb3-4.8.0-py2.5-linux-i686.egg/bsddb3/tests/test_all.py
+\end{verbatim}
+
+Let's see the configuration of the first server:
+
+\begin{verbatim}
+profile=/usr/local/share/arc/profiles/CheloniaAllServicesReplicatedAHashWithISIS.xml
+
+pidfile=/tmp/chelonia1.pid
+logfile=/var/log/chelonia1.log
+debug=ERROR
+
+port=60001
+
+cacert=/etc/grid-security/certificates
+host_cert=/etc/grid-security/hostcert-chelonia1.pem
+host_key=/etc/grid-security/hostkey-chelonia1.pem
+client_cert=/etc/grid-security/hostcert-chelonia1.pem
+client_key=/etc/grid-security/hostkey-chelonia1.pem
+
+[register]
+index_server1=https://chelonia1.knowarc.eu:60001/ISIS
+
+[ahash]
+endpoint=https://chelonia1.knowarc.eu:60001/AHash
+datadir=/var/spool/arc/chelonia1_ahash_data
+peerurl1=https://chelonia1.knowarc.eu:60002/AHash
+
+[bartender]
+endpoint=https://chelonia1.knowarc.eu:60001/Bartender
+
+[librarian]
+endpoint=https://chelonia1.knowarc.eu:60001/Librarian
+
+[shepherd]
+endpoint=https://chelonia1.knowarc.eu:60001/Shepherd
+turlprefix=https://chelonia1.knowarc.eu:60001/Hopi
+datadir=/var/spool/arc/chelonia1_shepherd_data
+storedir=/var/spool/arc/chelonia1_shepherd_store
+transferdir=/var/spool/arc/chelonia1_shepherd_transfer
+
+[isis]
+endpoint=https://chelonia1.knowarc.eu:60001/ISIS
+dbpath=/var/spool/arc/chelonia1_isis_data
+peerurl1=https://chelonia1.knowarc.eu:60001/ISIS
+peerurl2=https://chelonia2.knowarc.eu:60002/ISIS
+peerurl3=https://chelonia3.knowarc.eu:60003/ISIS
+\end{verbatim}
+
+The main difference in the config is that there is a peer URL given for the A-Hash, and three peer URL is given for the ISIS.
+The A-Hash is replicated not because we set a peer URL here, but because of the different profile.
+This profile contains a replicated A-Hash, while the previous profile contained a centralized one.
+
+The config for \verb!chelonia2!:
+
+\begin{verbatim}
+profile=/usr/local/share/arc/profiles/CheloniaAllServicesReplicatedAHashWithISIS.xml
+
+pidfile=/tmp/chelonia2.pid
+logfile=/var/log/chelonia2.log
+debug=ERROR
+
+port=60002
+
+cacert=/etc/grid-security/certificates
+host_cert=/etc/grid-security/hostcert-chelonia2.pem
+host_key=/etc/grid-security/hostkey-chelonia2.pem
+client_cert=/etc/grid-security/hostcert-chelonia2.pem
+client_key=/etc/grid-security/hostkey-chelonia2.pem
+
+[register]
+index_server1=https://chelonia2.knowarc.eu:60002/ISIS
+
+[ahash]
+endpoint=https://chelonia2.knowarc.eu:60002/AHash
+datadir=/var/spool/arc/chelonia2_ahash_data
+peerurl1=https://chelonia1.knowarc.eu:60001/AHash
+
+[bartender]
+endpoint=https://chelonia2.knowarc.eu:60002/Bartender
+
+[librarian]
+endpoint=https://chelonia2.knowarc.eu:60002/Librarian
+
+[shepherd]
+endpoint=https://chelonia2.knowarc.eu:60002/Shepherd
+turlprefix=https://chelonia2.knowarc.eu:60002/Hopi
+datadir=/var/spool/arc/chelonia2_shepherd_data
+storedir=/var/spool/arc/chelonia2_shepherd_store
+transferdir=/var/spool/arc/chelonia2_shepherd_transfer
+
+[isis]
+endpoint=https://chelonia2.knowarc.eu:60002/ISIS
+dbpath=/var/spool/arc/chelonia2_isis_data
+peerurl1=https://chelonia1.knowarc.eu:60001/ISIS
+peerurl2=https://chelonia2.knowarc.eu:60002/ISIS
+peerurl3=https://chelonia3.knowarc.eu:60003/ISIS
+\end{verbatim}
+
+And for \verb!chelonia3!:
+
+\begin{verbatim}
+profile=/usr/local/share/arc/profiles/CheloniaAllServicesReplicatedAHashWithISIS.xml
+
+pidfile=/tmp/chelonia3.pid
+logfile=/var/log/chelonia3.log
+debug=ERROR
+
+port=60003
+
+cacert=/etc/grid-security/certificates
+host_cert=/etc/grid-security/hostcert-chelonia3.pem
+host_key=/etc/grid-security/hostkey-chelonia3.pem
+client_cert=/etc/grid-security/hostcert-chelonia3.pem
+client_key=/etc/grid-security/hostkey-chelonia3.pem
+
+[register]
+index_server1=https://chelonia3.knowarc.eu:60003/ISIS
+
+[ahash]
+endpoint=https://chelonia3.knowarc.eu:60003/AHash
+datadir=/var/spool/arc/chelonia3_ahash_data
+peerurl1=https://chelonia1.knowarc.eu:60001/AHash
+peerurl2=https://chelonia2.knowarc.eu:60002/AHash
+
+[bartender]
+endpoint=https://chelonia3.knowarc.eu:60003/Bartender
+
+[librarian]
+endpoint=https://chelonia3.knowarc.eu:60003/Librarian
+
+[shepherd]
+endpoint=https://chelonia3.knowarc.eu:60003/Shepherd
+turlprefix=https://chelonia3.knowarc.eu:60003/Hopi
+datadir=/var/spool/arc/chelonia3_shepherd_data
+storedir=/var/spool/arc/chelonia3_shepherd_store
+transferdir=/var/spool/arc/chelonia3_shepherd_transfer
+
+[isis]
+endpoint=https://chelonia3.knowarc.eu:60003/ISIS
+dbpath=/var/spool/arc/chelonia3_isis_data
+peerurl1=https://chelonia1.knowarc.eu:60001/ISIS
+peerurl2=https://chelonia2.knowarc.eu:60002/ISIS
+peerurl3=https://chelonia3.knowarc.eu:60003/ISIS
+\end{verbatim}
+
+Now we can start all three. It may take a while for the system to stabilize, but then we will see a clean system, with no root collection. We should create one, then upload a file:
+
+\begin{verbatim}
+$ chelonia list /
+'/': not found
+
+$ chelonia mkdir /
+Creating collection '/': done
+
+$ chelonia put orange /
+'orange' (7 bytes) uploaded as '/orange'.
+\end{verbatim}
+
+After a while the file would have 3 replicas:
+
+\begin{verbatim}
+$ chelonia stat /orange
+'/orange': found
+  locations
+    https://chelonia3.knowarc.eu:60003/Shepherd 46831763-8162-06c5-8c61-81ab8a910e28: alive
+    https://chelonia1.knowarc.eu:60001/Shepherd 69e08316-093a-a788-6aea-6592c769c1de: alive
+    https://chelonia2.knowarc.eu:60002/Shepherd 1332408b-55e9-54b9-f56f-0ea5d77a069b: alive
 [...]
-Ran 383 tests in 29.858s
-
-OK
-\end{verbatim}
-
-Configure the first server (\verb!storage!). Here is the AHash part of the config:
-
-\begin{verbatim}
-<Service name="pythonservice" id="ahash">
-    <ClassName>storage.ahash.ahash.AHashService</ClassName>
-    <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
-    <LocalDir>/var/spool/arc/ahash_data</LocalDir>
-    <MyURL>https://storage:60000/AHash</MyURL>
-    <OtherURL>https://storage2:60006/AHash</OtherURL>
-    <Priority>50</Priority>
-    <ClientSSLConfig FromFile='/etc/arc/clientsslconfig.xml'/>
-</Service>
-\end{verbatim}
-
-In this config the second server is specified as a peer. Configure the second server (\verb!storage2!) as well:
-
-\begin{verbatim}
-<Plexer name="plexer.service" id="plexer">
-    <next id="ahash">^/AHash$</next>
-    <next id="shepherd">^/Shepherd$</next>
-    <next id="byteio">^/byteio/</next>
-</Plexer>
-<Service name="pythonservice" id="ahash">
-    <ClassName>storage.ahash.ahash.AHashService</ClassName>
-    <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
-    <LocalDir>/var/spool/arc/ahash_data2</LocalDir>
-    <MyURL>https://storage2:60006/AHash</MyURL>
-    <OtherURL>https://storage:60000/AHash</OtherURL>
-    <Priority>50</Priority>
-    <ClientSSLConfig FromFile='/etc/arc/clientsslconfig.xml'/>
-</Service>
-\end{verbatim}
-
-In this config the first server is configured as a peer. So the two together will be able to form a replicated network with two nodes. Then configure the third one (\verb!storage3!):
-
-\begin{verbatim}
-<Plexer name="plexer.service" id="plexer">
-    <next id="ahash">^/AHash$</next>
-    <next id="shepherd">^/Shepherd$</next>
-</Plexer>
-<Service name="pythonservice" id="ahash">
-    <ClassName>storage.ahash.ahash.AHashService</ClassName>
-    <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
-    <LocalDir>/var/spool/arc/ahash_data3</LocalDir>
-    <MyURL>https://storage3:60010/AHash</MyURL>
-    <OtherURL>https://storage:60000/AHash</OtherURL>
-    <Priority>50</Priority>
-    <ClientSSLConfig FromFile='/etc/arc/clientsslconfig.xml'/>
-</Service>
-\end{verbatim}
-
-Here the first server is specified as the peer.
-
-It is important to make a fresh start on all the servers, and remove the existing database, which was created by a different store:
-\begin{verbatim}
-$ sudo rm -r /var/spool/arc/*
-\end{verbatim}
-
-Now, start the first server. It will complain that it cannot find its peer:
-\begin{verbatim}
-ERROR connecting to https://storage2:60006/AHash
-[2009-03-24 10:34:18] [Arc.Storage.ReplicatedAHash] [ERROR] [10827/156026112]
-    failed to send to 2 of [1, 2]
-[2009-03-24 10:34:20] [Arc.Storage.ReplicatedAHash] [ERROR] [10827/158713512]
-    Couldn't run election
-\end{verbatim}
-
-With only one node of the replicated A-Hash the system is not ready to work, and trying to use it will result in an error message:
-\begin{verbatim}
-$ chelonia makeCollection /
-- Calling the Bartender's makeCollection method...
-- done in 0.55 seconds.
-Creating collection '/': failed to create new librarian entry
-\end{verbatim}
-
-Start the second server. It may first print some error messages, but then both servers should be quiet, and the system is ready for use. Now, upload a file, and set the number of needed replicas to 3:
-
-\begin{verbatim}
-$ chelonia makeCollection /
-- Calling the Bartender's makeCollection method...
-- done in 1.16 seconds.
-Creating collection '/': done
-$ chelonia put orange /
-- The size of the file is 7 bytes
-- Calculating md5 checksum...
-- done in 0.0214 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
-- Calling the Bartender's putFile method...
-- done in 2.22 seconds.
-- Got transfer URL: http://storage:60001/hopi/3f903c50-0424-2891-5730-521cc00d96dc
-- Uploading from 'orange'
-    to 'http://storage:60001/hopi/3f903c50-0424-2891-5730-521cc00d96dc' with http...
-    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
-- done in 1.0293 seconds.
-'orange' (7 bytes) uploaded as '/orange'.
-$ chelonia modify /orange set states neededReplicas 3
-- Calling the Bartender's modify method...
-- done in 0.74 seconds.
-set
+\end{verbatim}
+
+Now we can stop \verb!chelonia1!, because all the services are replicated.
+
+\begin{verbatim}
+$ sudo kill `cat /tmp/chelonia1.pid`
+\end{verbatim}
+
+Let's check the file's replicas again:
+
+\begin{verbatim}
 $ chelonia stat /orange
+ERROR: No Bartender URL found.
+(use the '-v' flag for more details, e.g. 'chelonia -v list /')
+\end{verbatim}
+
+As always, let's use the \verb!-v! flag to get more details:
+
+\begin{verbatim}
+$ chelonia -v stat /orange
+- Using default user config.
+- The key file: /home/zsombor/.arc/userkey-billy.pem
+- The cert file: /home/zsombor/.arc/usercert-billy.pem
+- The CA dir: /etc/grid-security/certificates
+- No Bartender URL found in the config or in the environment, try to get one from ISIS:
+[2009-10-20 16:23:44] [Arc.MCC.TLS] [ERROR] [25243/148071368] Failed to establish SSL connection
+[2009-10-20 16:23:44] [Arc.MCC.TLS] [ERROR] [25243/148071368] SSL error: -1 - (empty):(empty):(empty)
+[2009-10-20 16:23:44] [Arc.MCC.TLS] [ERROR] [25243/148071368] Failed to send content of buffer
+  - https://chelonia1.knowarc.eu:60001/ISIS - Failed to connect to ISIS.
+ERROR: No Bartender URL found.
+\end{verbatim}
+
+It says failed to connect to ISIS on the first server, which is true, because we stopped it.
+But in the current client configuration there is only this one ISIS specified, so let's add the others (the first three line should be in one line, separated with space)
+
+\begin{verbatim}
+defaultservices=index:ARC1:https://chelonia1.knowarc.eu:60001/ISIS
+    index:ARC1:https://chelonia2.knowarc.eu:60002/ISIS
+    index:ARC1:https://chelonia3.knowarc.eu:60003/ISIS
+keypath=/home/zsombor/.arc/userkey-billy.pem
+certificatepath=/home/zsombor/.arc/usercert-billy.pem
+cacertificatesdirectory=/etc/grid-security/certificates
+\end{verbatim}
+
+And try again:
+
+\begin{verbatim}
+$ chelonia stat /orange
+'/orange': found
+  locations
+    https://chelonia3.knowarc.eu:60003/Shepherd 46831763-8162-06c5-8c61-81ab8a910e28: alive
+    https://chelonia1.knowarc.eu:60001/Shepherd 69e08316-093a-a788-6aea-6592c769c1de: offline
+    https://chelonia2.knowarc.eu:60002/Shepherd 1332408b-55e9-54b9-f56f-0ea5d77a069b: alive
 [...]
-  locations
-    https://storage:60000/Shepherd f4ad8530-354d-5066-04ae-c54c94fd525b: alive
-[...]
-zsombor@storage:~$ chelonia stat /orange
-[...]
-  locations
-    https://storage:60000/Shepherd f4ad8530-354d-5066-04ae-c54c94fd525b: alive
-    https://storage2:60006/Shepherd 97edebb5-e72d-098a-55a7-14595f668770: alive
-[...]    
-\end{verbatim}
-
-It cannot create 3 replicas, because the third server is not running. Start it. After a minute or two, the file has three replicas:
-
-\begin{verbatim}
-$ chelonia stat /orange
-[...]
-  locations
-    https://storage:60000/Shepherd f4ad8530-354d-5066-04ae-c54c94fd525b: alive
-    https://storage3:60010/Shepherd 3c55924b-73ad-c226-0f4a-4c056c967314: alive
-    https://storage2:60006/Shepherd 97edebb5-e72d-098a-55a7-14595f668770: alive
-[...]
-\end{verbatim}
-
-The fact that this file has three replicas is not at all related to the number of AHashes, but to the number of storage elements (Shepherds and storage element services), so this is no different then what was seen in the previous section.
-The benefit of having more then one AHashes in a replicated way is that if one of the machines goes offline, the AHash is still working.
-However, with the setup above, if the first server is shutdown, the AHash will still work but there is only one instance of the other two important service: the Librarian and the Bartender, so the whole system will still be unusable.
-Hence, more Librarians and Bartenders needs to be added to the system.
-The next section explains how.
-
-\section{Additional Bartenders and Librarians} % (fold)
-\label{sec:additional_bartenders_and_librarians}
-
-In this section, a Bartender and a Librarian is added to each server.
-The first server does not need modification now.
-For the other two, the \verb!Plexer!
-section needs to be altered, the twos services must be added, and the Shepherd must be modified to use the server's own Bartender and Librarian for replication and heartbeats.
-Modify the config of the second server as follows: (\verb![...]!
-indicates unchanged parts)
-
-\begin{verbatim}
-[...]
-<Plexer name="plexer.service" id="plexer">
-    <next id="ahash">^/AHash$</next>
-    <next id="shepherd">^/Shepherd$</next>
-    <next id="librarian">^/Librarian$</next>
-    <next id="bartender">^/Bartender$</next>
-    <next id="byteio">^/byteio/</next>
-</Plexer>
-<Service name="pythonservice" id="librarian">
-    <ClassName>storage.librarian.librarian.LibrarianService</ClassName>
-    <AHashURL>https://storage2:60006/AHash</AHashURL>
-    <HeartbeatTimeout>30</HeartbeatTimeout>
-    <CheckPeriod>20</CheckPeriod>
-    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
-</Service>
-<Service name="pythonservice" id="bartender">
-    <ClassName>storage.bartender.bartender.BartenderService</ClassName>
-    <LibrarianURL>https://storage2:60006/Librarian</LibrarianURL>
-    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
-</Service>
-[...]
-<Service name="pythonservice" id="shepherd">
-[...]
-    <LibrarianURL>https://storage2:60006/Librarian</LibrarianURL>
-    <BartenderURL>https://storage2:60006/Bartender</BartenderURL>
-[...]
-\end{verbatim}
-
-And the third server:
-
-\begin{verbatim}
-[...]
-<Plexer name="plexer.service" id="plexer">
-    <next id="ahash">^/AHash$</next>
-    <next id="shepherd">^/Shepherd$</next>
-    <next id="librarian">^/Librarian$</next>
-    <next id="bartender">^/Bartender$</next>
-</Plexer>
-<Service name="pythonservice" id="librarian">
-    <ClassName>storage.librarian.librarian.LibrarianService</ClassName>
-    <AHashURL>https://storage3:60010/AHash</AHashURL>
-    <HeartbeatTimeout>30</HeartbeatTimeout>
-    <CheckPeriod>20</CheckPeriod>
-    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
-</Service>
-<Service name="pythonservice" id="bartender">
-    <ClassName>storage.bartender.bartender.BartenderService</ClassName>
-    <LibrarianURL>https://storage3:60010/Librarian</LibrarianURL>
-    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
-</Service>
-[...]
-<Service name="pythonservice" id="shepherd">
-[...]
-    <LibrarianURL>https://storage3:60010/Librarian</LibrarianURL>
-    <BartenderURL>https://storage3:60010/Bartender</BartenderURL>
-[...]
-\end{verbatim}
-
-\verb!client.xml! must be modified to add all the other Bartenders:
-
-\begin{verbatim}
-$ cat ~/.arc/client.xml 
-<ArcConfig>
-  <KeyPath>/home/zsombor/.arc/userkey.pem</KeyPath>
-  <CertificatePath>/home/zsombor/.arc/usercert.pem</CertificatePath>
-  <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
-  <BartenderURL>https://storage:60000/Bartender</BartenderURL>
-  <BartenderURL>https://storage2:60006/Bartender</BartenderURL>
-  <BartenderURL>https://storage3:60010/Bartender</BartenderURL>
-</ArcConfig>
-\end{verbatim}
-
-Now, all three servers can be started.
-The replicated A-Hash sometimes prints great amount of error messages, but if it stops printing them, then everything is OK.
-Now the system can be used as intended.
-There is a flag for the \verb!chelonia!
-tool which makes it very verbose.
-You can check the used Bartender with it:
-
-\begin{verbatim}
-$ chelonia -v stat /orange
-- The URL of the Bartender(s): https://storage:60000/Bartender,
-    https://storage2:60006/Bartender, https://storage3:60010/Bartender
-- The key file: /home/zsombor/.arc/userkey.pem
-- The cert file: /home/zsombor/.arc/usercert.pem
-- The CA dir: /etc/grid-security/certificates
-- Calling the Bartender's stat method...
-- trying https://storage:60000/Bartender
-[...]
-- done in 1.27 seconds.
-'/orange': found
-\end{verbatim}
-
-Now, stop the first server. The system will be unaccessible for a few seconds while the other two A-Hash figure out what is happening, but after that it will work again:
-
-\begin{verbatim}
-$ chelonia -v stat /orange
-- The URL of the Bartender(s): https://storage:60000/Bartender,
-    https://storage2:60006/Bartender, https://storage3:60010/Bartender
-- The key file: /home/zsombor/.arc/userkey.pem
-- The cert file: /home/zsombor/.arc/usercert.pem
-- The CA dir: /etc/grid-security/certificates
-- Calling the Bartender's stat method...
-- trying https://storage:60000/Bartender
-[...]
-- (-1, "Wrong status from server.\nMaybe wrong URL: 'https://storage:60000/Bartender'")
-- trying https://storage2:60006/Bartender
-[...]
-- done in 0.45 seconds.
-'/orange': found
-[...]
-\end{verbatim}
-
-If the first server is started again, after a few seconds the system will work again with the three fully-functional node.
-
-\begin{verbatim}
-zsombor@storage:~$ chelonia -v stat /orange
-- The URL of the Bartender(s): https://storage:60000/Bartender,
-    https://storage2:60006/Bartender, https://storage3:60010/Bartender
-- The key file: /home/zsombor/.arc/userkey.pem
-- The cert file: /home/zsombor/.arc/usercert.pem
-- The CA dir: /etc/grid-security/certificates
-- Calling the Bartender's stat method...
-- trying https://storage:60000/Bartender
-[...]
-- done in 0.49 seconds.
-'/orange': found
-  states
-    checksumType: md5
-    neededReplicas: 3
-    size: 7
-    checksum: 6aefd2842be62cd470709b27aedc7db7
-  entry
-    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
-    GUID: 0080d8be-1f6f-2fc7-ad71-732f899326c8
-    type: file
-  parents
-    0/orange: parent
-  locations
-    https://storage3:60010/Shepherd 03312843-d3b3-2d15-4732-ad673df958ae: alive
-    https://storage2:60006/Shepherd b0057765-4532-d0b4-fded-182499a91450: alive
-    https://storage:60000/Shepherd 532ec0f8-dd83-297b-d7b6-dc5dacea34e0: alive
-  timestamps
-    created: 1237887474.95    
+\end{verbatim}
+
+So it is working, we have two replicas and we can download the file:
+
+\begin{verbatim}
+$ chelonia get /orange /tmp
+'/orange' (7 bytes) downloaded as '/tmp/orange'.
+
+$ cat /tmp/orange 
+ORANGE
+\end{verbatim}
+
+And if we restart \verb!chelonia1! all three replicas should be online again.
+
+\begin{verbatim}
+
 \end{verbatim}
@@
 \label{sec:using_the_gateway}
+
+\textbf{This section was not updated.}
 
@@
 \label{sec:using_the_arc_dmc}
+
+\textbf{This section was not updated.}
 
@@
 
+\textbf{This section was not updated.}
+
 Besides the \verb!chelonia! tool you can access the storage system with the use of a FUSE module. The following steps installs the ARC storage FUSE module:

### 2009-10-21 doc/tech_doc/jura/jura-tech-doc.tex
Documenting JURA interactive mode, correcting refs in PDF

@@
 
-\verb|jura [-E <expiration_time>] <control_dir> [<control_dir> [...]]|
+\verb|jura [-E <expiration_time>] [-u <url> [-u <url> [...]]] <control_dir> [<control_dir> [...]]|
 
@@
 directory is traversed by JURA separately, in sequence.
+
+The ``\textit{-u}'' option can be used for interactive execution
+(e.g.~from a terminal). In this case, usage data generated for each
+job is reported once to each of the specified destination URLs
+regardless of the content of the job log files, and no job log files
+are deleted.
 
@@
 
-After successful submission to a reporting destination, the job log
+If interactive mode is not activated by the ``\textit{-u}'' option
+(see Section \ref{exec}),
+after successful submission to a reporting destination, the job log
 file is deleted, preventing multiple insertion of usage records. If
@@
 \label{accessing}
-The generated URs are submitted to the accounting services specified
+In case of non-interactive invocation of JURA by A-REX, 
+the generated URs are submitted to the accounting services specified
 by the reporting destination configuration parameters and if present,
 to the destinations specified in the job description as
-well. Reporting URs to several destinations is possible, but currently
+well. Under interactive mode of operation, they are submitted to the
+services given via the ``\textit{-u}'' command line option. Reporting
+URs to several destinations is possible, but currently
 only SGAS LUTS destinations are supported.
@@
   the service interfaces is missing as of now.
-\item Extend JURA to be able to read all necessary parameters
-  (credentials, reporting targets) from the command line so that it
-  can serve as a stand-alone reporting utility invoked independently
-  from A-REX.
+\item Extend JURA to be able to read credentials from the command line
+  in interactive mode.
 \item Enable generating coarser-grained ``\textit{Aggregate URs}''

### 2009-10-25 doc/manuals/user_interface/ui-conf.tex
First approach at updating and synchronizing the new CLI manual

Default behaviour of an ARC client can be configured by specifying
alternative values for some parameters in the client configuration
file. The file is called \texttt{client.conf} and is located in
directory \texttt{.arc} in user's home area:
\begin{shaded}
 {\$}HOME/.arc/client.conf
\end{shaded}
If this file is not present or does not contain the relevant
configuration information, the global configuration files (if exist)
or default values are used instead. Some client tools may be able to create
the default \texttt{{\$}HOME/.arc/client.conf}, if it does not exist.

The ARC configuration file consists of several configuration blocks.
Each configuration block is identified by a keyword and contains
configuration options for a specific part of the ARC middleware.

The configuration file is written in a plain text format known as INI.
Configuration blocks start with identifying keywords inside square brackets.
Typically, first comes a common block: \verb#[common]#. Thereafter follows one
or more attribute-value pairs written one on each line in the following
format:

\begin{framed}
\begin{verbatim}
[common]
attribute1=value1
attribute2=value2
attribute3=value3 value4
# comment line 1
# comment line 2
attribute2=value5
...
\end{verbatim}
\end{framed}

%For multi-valued attributes, several elements or attribute-value pairs
%have to be specified -- one per each value.

The following blocks and respective attributes are recognized.

\section{Block \texttt{common}}

\phantomsection
\index{configuration:defaultservices}\addcontentsline{toc}{subsection}{defaultservices}
\hspace*{0.5cm}
\begin{shaded}
  \uicommand{defaultservices}
\end{shaded}
\textbf{This attribute is multi-valued.}

This attribute is used to specify default services to be used. Defining such in
the user configuration file will override the default services set in the system
configuration.

The value of this attribute should follow the format:
\begin{verbatim}
  service_type:flavour:service_url
\end{verbatim}

where \texttt{service\_type} is type of service (e.g. \texttt{computing} or
\texttt{index}), \texttt{flavour} specifies type of middleware plugin to use
when contacting the service (e.g. ARC0, ARC1, CREAM, UNICORE, etc.) and
\texttt{service\_url} is the URL used to contact the service.

Example:
\begin{itemize}
\item{plain text:}\\ 
\verb#defaultservices=index:ARC0:ldap://index1.nordugrid.org:2135/Mds-Vo-name=NorduGrid,o=grid index:ARC1:https://knowarc2.grid.niif.hu:50000/isis computing:ARC1:https://knowar c1.grid.niif.hu:60000/arex computing:CREAM:ldap://cream.grid.upjs.sk:2170/o=grid computing:UNICORE:https://testbed5.grid.upjs.sk:8080/KnowARC-testbed/services/BESFactory?res=default_bes_factory#
\end{itemize}

% \phantomsection
% \index{configuration:giis}\addcontentsline{toc}{subsection}{giis}
% \hspace*{0.5cm}
% \begin{shaded}
%   \uicommand{giis}
% \end{shaded}
% \textbf{This attribute is multi-valued.}
% 
% Configures which GIISes (site indices) to use to discover computing
% and storage resources. The default is to use the four top level
% GIISes:
% 
% \begin{verbatim}
%   ldap://index1.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid
%   ldap://index2.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid
%   ldap://index3.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid
%   ldap://index4.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid
% \end{verbatim}
% 
% Examples:
% \begin{itemize}
% \item{plain text:}\\ 
% \verb#    giis="ldap://atlasgiis.nbi.dk:2135/O=Grid/Mds-Vo-name=Atlas"#\\
% \verb#    giis="ldap://index1.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid"#\\
% \verb#    giis="ldap://grid.tsl.uu.se:2135/O=Grid/Mds-Vo-name=local"#
% \item{XML:}\\ 
% \verb#    <giis>ldap://odin.switch.ch:2135/O=Grid/Mds-Vo-name=Switzerland</giis>#\\
% \verb#    <giis>ldap://index1.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid</giis>#\\
% \verb#    <giis>ldap://grid.tsl.uu.se:2135/O=Grid/Mds-Vo-name=local</giis>#
% \end{itemize}
% 
% \begin{framed}
% Note that LDAP URLs containing "Mds-Vo-name=local" refer not to
% GIISes, but to individual sites. This feature can be used to add
% "hidden" sites that do not register to any GIIS, or to list explicitly
% preferred submission sites.
% \end{framed}
% 
% \phantomsection
% \index{configuration:debug}\addcontentsline{toc}{subsection}{debug}
% \hspace*{0.5cm}
% \begin{shaded}
%   \uicommand{debug}
% \end{shaded}
% Default debug level to use for the ARC clients. Corresponds to the \verb#-d#
% command line switch of the clients. Default value is 0, possible range
% is from -3 to 3.
% 
% Examples:
% \begin{itemize}
% \item{plain text:}\\ \verb#    debug="2"#
% \item{XML:}\\ \verb#    <debug>2</debug>#
% \end{itemize}
% 
% 
% \phantomsection
% \index{configuration:timeout}\addcontentsline{toc}{subsection}{timeout}
% \hspace*{0.5cm}
% \begin{shaded}
%   \uicommand{timeout}
% \end{shaded}
% Timeout to use for interaction with LDAP servers, gridftp servers,
% etc. If a server, during e.g. job submission, does not answer in the
% specified number of seconds, the connection is timed out and
% closed. Default value is 20 seconds.
% 
% Examples:
% \begin{itemize}
% \item{plain text:}\\ \verb#    timeout="20"#
% \item{XML:}\\ \verb#    <timeout>10</timeout>#
% \end{itemize}
% 
% 
% \phantomsection
%
\index{configuration:downloaddir}\addcontentsline{toc}{subsection}{downloaddir}
% \hspace*{0.5cm}
% \begin{shaded}
%   \uicommand{downloaddir}
% \end{shaded}
% Default download directory to download files to when retrieving output
% files from jobs using \texttt{ngget}.  Default is the current working
% directory.
% 
% Examples:
% \begin{itemize}
% \item{plain text:}\\ \verb#    downloaddir="/home/johndoe/arc-downloads"#
% \item{XML:}\\ \verb#    <downloaddir>/tmp/johndoe</downloaddir>#
% \end{itemize}
% 
% 
% \phantomsection
% \index{configuration:alias}\addcontentsline{toc}{subsection}{alias}
% \hspace*{0.5cm}
% \begin{shaded}
%   \uicommand{alias}
% \end{shaded}
% \textbf{This attribute is multi-valued.}
% 
% Alias substitutions to perform in connection with the \verb#-c# command line
% switch of the ARC clients. Alias definitions are recursive. Any alias
% defined in a block that is read before a given block can be used in
% alias definitions in that block. An alias defined in a block can also
% be used in alias definitions later in the same block.
% 
% Examples:
% \begin{itemize}
% \item{plain text:}\\ \verb#    alias="host1=somehost.nbi.dk"#\\
% \verb#    alias="host2=otherhost.uu.se"#\\
% \verb#    alias="myhosts=host1 host2"#
% \item{XML:}\\ \verb#    <alias>host1=somehost.nbi.dk</alias>#\\
% \verb#    <alias>host2=otherhost.uu.se</alias>#\\
% \verb#    <alias>myhosts=host1 host2</alias>#
% \end{itemize}
% 
% With the example above, \verb#ngsub -c myhosts# will resolve to \verb#ngsub -c somehost.nbi.dk -c otherhost.uu.se#.
% 
% \phantomsection
% \index{configuration:broker}\addcontentsline{toc}{subsection}{broker}
% \hspace*{0.5cm}
% \begin{shaded}
%   \uicommand{broker}
% \end{shaded}
% Configures which brokering algorithm to use during job submission. The
% default one is the \texttt{FastestCpus} broker that chooses, among the
% possible targets, the target with the fastest CPUs. Another
% possibility is, for example, the
% \texttt{RandomSort} broker that chooses the target randomly among the targets
% surviving the job description matchmaking.
% 
% Examples:
% \begin{itemize}
% \item{plain text:}\\ \verb#    broker="RandomSort"#
% \item{XML:}\\ \verb#    <broker>RandomSort</broker>#
% \end{itemize}

\index{configuration:deprecated files}\section{Deprecated configuration files}

ARC configuration file in releases 0.6 and 0.8 has the same name and the same
format; attributes unknown to newer ARC versions are ignored.

In ARC $\leq$ 0.5.48, configuration was done via files {\$}\texttt{HOME/.ngrc},
{\$}\texttt{HOME/.nggiislist} and {\$}\texttt{HOME/.ngalias}.

The main configuration file \textbf{{\$}\texttt{HOME/.ngrc}} could contain
user's default settings for the debug level, the information system
query timeout and the download directory used by \texttt{ngget}. A
sample file could be the following:
\begin{verbatim}
# Sample .ngrc file
# Comments starts with #
NGDEBUG=1
NGTIMEOUT=60
NGDOWNLOAD=/tmp
\end{verbatim}

If the environment variables NGDEBUG, NGTIMEOUT or NGDOWNLOAD were
defined, these took precedence over the values defined in this
configuration. Any command line options override the defaults.

The file \textbf{{\$}\texttt{HOME/.nggiislist}} was used to keep the
list of default GIIS server URLs, one line per GIIS (see \texttt{giis}
attribute description above).

The file \textbf{{\$}\texttt{HOME/.ngalias}} was used to keep the
list of site aliases, one line per alias (see \texttt{alias}
attribute description above).

### 2009-10-25 doc/manuals/user_interface/ui.tex
add note about previous incarnation of command names
First approach at updating and synchronizing the new CLI manual

@@
 \hypersetup{
-  pdfauthor = {},
+  pdfauthor = {ARC developers},
   pdftitle = {ARC Clients},
   pdfsubject = {User's manual},
-  pdfkeywords = {Grid,KnowARC,ARC,client},
+  pdfkeywords = {Grid,KnowARC,ARC,NorduGrid,client},
   pdfcreator = {PDFLaTeX with hyperref package},
@@
 
-   {\raggedleft NORDUGRID-MANUAL-?\par}
+   {\raggedleft NORDUGRID-MANUAL-1\par}
 
@@
 The command line user interface of ARC consists of a set of commands
-necessary for job submission and manipulation and data management. A
-special utility also exists for test purposes.
+necessary for job submission and manipulation and data management.
+This manual replaces the older version of \texttt{NORDUGRID-MANUAL-1} and is
+valid for ARC versions 0.9 and above. Command line tools semantics is the same
+as in earlier versions of ARC, roughly following that of basic Linux commands
+and most common batch system commands. One obvious difference is change of
+the legacy prefix from ``ng'' to the more appropriate ``arc''. This is not only
+a cosmetic change: \textbf{behaviour of the commands also have changed}, as did
+their functionalities and options.
+\begin{framed}
+  Users are strongly discouraged from modifying their old scripts by
+  simply replacing ``ng'' with ``arc'' -- results may be unpredictable.
+\end{framed}
 
@@
 such as status check, results retrieval, cancellation, re-submission
-and such. The jobs must be described using a job description language. ARC supports the languages JSDL, xRSL and JDL~\cite{xrsl}.
+and such. The jobs must be described using a job description language. ARC
+supports the following languages: JSDL~\cite{jsdl}, xRSL~\cite{xrsl} and
+JDL~\cite{jdl}.
 
@@
 
-\section{Test suite} \label{sec:testsuite}
+%\section{Test suite} \label{sec:testsuite}
 %\input{arctest}
 
-\section{Third-party commands} \label{sec:ui-other}
+%\section{Third-party commands} \label{sec:ui-other}
 %\input{ui-other}
@@
 
-\chapter{Configuration} \label{sec:client.conf} 
-\section{ARC Client Configuration}
-%\input{ui-conf}
+\chapter{ARC Client Configuration} \label{sec:client.conf} 
+\input{ui-conf}
 

### 2009-10-25 doc/tech_doc/xrsl/xrsl-urls.tex
First approach at updating and synchronizing the new CLI manual

@@
 
-\begin{tabular}{lp{10cm}}
+\begin{longtable}{lp{10cm}}
    \verb#threads=<number># & specifies number of parallel
@@
    default is protocol dependent\\
-   \verb#checksum=cksum|md5|adler32# & specifies the algorithm for checksum to be
-   computed (ev. provided to the indexing server). This is overridden
-   by any metadata options specified (see below).\\ 
+   \verb#checksum=cksum|md5|adler32|no# & specifies the algorithm for checksum to be
+   computed (for transfer verification or provided to the indexing server). This is overridden
+   by any metadata options specified (see below). If this option is
+   not provided, the default for the protocol is used. \verb#checksum=no#
+   disables checksum calculation.\\
    \verb#exec=yes|no# & means the file should be treated as executable\\
@@
    http(s) only. Default is \verb#no#\\ 
-\end{tabular}
+\end{longtable}
 
@@
 $^1$This is a destination URL. The file will be copied to the GridFTP
-server at se.domain.org with the path datapath/file25.dat and
-registered in the RLS indexing service at grid.domain.org with the LFN
-myfile02.dat.
+server at \texttt{se.domain.org} with the path \texttt{datapath/file25.dat} and
+registered in the RLS indexing service at \texttt{grid.domain.org} with the LFN
+\texttt{myfile02.dat}.
 
 $^2$This is a destination URL. The file will be copied to
-srm.domain.org at the path griddir/file1 and registered to the LFC
-service at lfc.domain.org with the LFN /user/file1. The given GUID
+\texttt{srm.domain.org} at the path \texttt{griddir/file1} and registered to the LFC
+service at \texttt{lfc.domain.org} with the LFN \texttt{/user/file1}. The given GUID
 and checksum attributes will be registered.
@@
 $^3$This is a source URL. The file is registered in the LFC service at
-lfc.domain.org with the given GUID and can be copied or queried by
+\texttt{lfc.domain.org} with the given GUID and can be copied or queried by
 this URL.

### 2009-10-25 doc/manuals/user_interface/ui-conf.tex
Added Proxy and Configuration sections

@@
 # comment line 2
-attribute2=value5
 ...
@@
 
-The following blocks and respective attributes are recognized.
+Most attributes have counterpart command line options. Command line options
+always overwrite configuration attributes.
 
-\section{Block \texttt{common}}
+Two blocks are currently recognized, \texttt{[common]} and
+\texttt{[alias]}. Following sections describe supported attributes per block.
+
+\section{Block \texttt{[common]}}
 
@@
 when contacting the service (e.g. ARC0, ARC1, CREAM, UNICORE, etc.) and
-\texttt{service\_url} is the URL used to contact the service.
+\texttt{service\_url} is the URL used to contact the service. Several services
+can be listed, separated with a blank space (no line breaks allowed).
 
 Example:
+\begin{verbatim*} 
+defaultservices=index:ARC0:ldap://index1.ng.org:2135/Mds-Vo-name=testvo,o=grid
+ index:ARC1:https://index2.ng.org:50000/isis
+ computing:ARC1:https://ce.arc.org:60000/arex
+ computing:CREAM:ldap://ce.glite.org:2170/o=grid
+ computing:UNICORE:https://ce.unicore.org:8080/test/services/BESFactory?res=default_bes_factory
+\end{verbatim*}
+
+\phantomsection
+\index{configuration:rejectservices}\addcontentsline{toc}{subsection}{rejectservices}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{rejectservices}
+\end{shaded}
+\textbf{This attribute is multi-valued.}
+
+This attribute can be used to indicate that a certain service should be
+rejected (``blacklisted''). Several services can be listed, separated with a
+blank space (no line breaks allowed).
+
+Example:
+\verb#    rejectservices=computing:ARC1:https://bad.service.org/arex#
+
+ \phantomsection
+ \index{configuration:verbosity}\addcontentsline{toc}{subsection}{verbosity}
+ \hspace*{0.5cm}
+ \begin{shaded}
+   \uicommand{verbosity}
+ \end{shaded}
+ Default verbosity (debug) level to use for the ARC clients. Corresponds to the
+\verb#-d# command line option of the clients. Default value is \texttt{WARNING},
+possible values are \texttt{FATAL}, \texttt{ERROR}, \texttt{WARNING},
+\texttt{INFO}, \texttt{DEBUG} or \texttt{VERBOSE}.
+
+Example:
+\verb#    verbosity=INFO#
+
+\phantomsection
+\index{configuration:timeout}\addcontentsline{toc}{subsection}{timeout}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{timeout}
+\end{shaded}
+Sets the period of time the client should wait for a service (information,
+computing, storage etc) to respond when communicating with it. The period
+should be given in seconds. Default value is 20 seconds. This attribute
+corresponds to the \verb#-t# command line option.
+
+Example:
+\verb#    timeout=10#
+
+\phantomsection
+\index{configuration:brokername}\addcontentsline{toc}{subsection}{brokername}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{brokername}
+\end{shaded}
+
+Configures which brokering algorithm to use during job submission. This attribute
+corresponds to the \verb#-b# command line option. The
+default one is the \texttt{Random} broker that chooses targets randomly.
+Another possibility is, for example, the \texttt{FastestQueue} broker that
+chooses the target with the shortest estimated queue waiting time. Possible
+values are:
 \begin{itemize}
-\item{plain text:}\\ 
-\verb#defaultservices=index:ARC0:ldap://index1.nordugrid.org:2135/Mds-Vo-name=NorduGrid,o=grid index:ARC1:https://knowarc2.grid.niif.hu:50000/isis computing:ARC1:https://knowar c1.grid.niif.hu:60000/arex computing:CREAM:ldap://cream.grid.upjs.sk:2170/o=grid computing:UNICORE:https://testbed5.grid.upjs.sk:8080/KnowARC-testbed/services/BESFactory?res=default_bes_factory#
+ \item[\texttt{Random}] -- ranks targets randomly (default)
+ \item[\texttt{FastestQueue}] -- ranks targets according to their queue length
+% \item[\texttt{FastestCPU}] -- ranks targets according to CPU performance
+ \item[\texttt{Data}] -- ranks target according the amount of megabytes of the
+requested input files that are already in the computing resource’s cache.
+% \item[\texttt{SampleBroker}] -- Ranks targets randomly (written in Python)
 \end{itemize}
 
-% \phantomsection
-% \index{configuration:giis}\addcontentsline{toc}{subsection}{giis}
-% \hspace*{0.5cm}
-% \begin{shaded}
-%   \uicommand{giis}
-% \end{shaded}
-% \textbf{This attribute is multi-valued.}
-% 
-% Configures which GIISes (site indices) to use to discover computing
-% and storage resources. The default is to use the four top level
-% GIISes:
-% 
-% \begin{verbatim}
-%   ldap://index1.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid
-%   ldap://index2.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid
-%   ldap://index3.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid
-%   ldap://index4.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid
-% \end{verbatim}
-% 
-% Examples:
-% \begin{itemize}
-% \item{plain text:}\\ 
-% \verb#    giis="ldap://atlasgiis.nbi.dk:2135/O=Grid/Mds-Vo-name=Atlas"#\\
-% \verb#    giis="ldap://index1.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid"#\\
-% \verb#    giis="ldap://grid.tsl.uu.se:2135/O=Grid/Mds-Vo-name=local"#
-% \item{XML:}\\ 
-% \verb#    <giis>ldap://odin.switch.ch:2135/O=Grid/Mds-Vo-name=Switzerland</giis>#\\
-% \verb#    <giis>ldap://index1.nordugrid.org:2135/O=Grid/Mds-Vo-name=NorduGrid</giis>#\\
-% \verb#    <giis>ldap://grid.tsl.uu.se:2135/O=Grid/Mds-Vo-name=local</giis>#
-% \end{itemize}
-% 
-% \begin{framed}
-% Note that LDAP URLs containing "Mds-Vo-name=local" refer not to
-% GIISes, but to individual sites. This feature can be used to add
-% "hidden" sites that do not register to any GIIS, or to list explicitly
-% preferred submission sites.
-% \end{framed}
-% 
-% \phantomsection
-% \index{configuration:debug}\addcontentsline{toc}{subsection}{debug}
-% \hspace*{0.5cm}
-% \begin{shaded}
-%   \uicommand{debug}
-% \end{shaded}
-% Default debug level to use for the ARC clients. Corresponds to the \verb#-d#
-% command line switch of the clients. Default value is 0, possible range
-% is from -3 to 3.
-% 
-% Examples:
-% \begin{itemize}
-% \item{plain text:}\\ \verb#    debug="2"#
-% \item{XML:}\\ \verb#    <debug>2</debug>#
-% \end{itemize}
-% 
-% 
-% \phantomsection
-% \index{configuration:timeout}\addcontentsline{toc}{subsection}{timeout}
-% \hspace*{0.5cm}
-% \begin{shaded}
-%   \uicommand{timeout}
-% \end{shaded}
-% Timeout to use for interaction with LDAP servers, gridftp servers,
-% etc. If a server, during e.g. job submission, does not answer in the
-% specified number of seconds, the connection is timed out and
-% closed. Default value is 20 seconds.
-% 
-% Examples:
-% \begin{itemize}
-% \item{plain text:}\\ \verb#    timeout="20"#
-% \item{XML:}\\ \verb#    <timeout>10</timeout>#
-% \end{itemize}
-% 
-% 
-% \phantomsection
-%
-\index{configuration:downloaddir}\addcontentsline{toc}{subsection}{downloaddir}
-% \hspace*{0.5cm}
-% \begin{shaded}
-%   \uicommand{downloaddir}
-% \end{shaded}
-% Default download directory to download files to when retrieving output
-% files from jobs using \texttt{ngget}.  Default is the current working
-% directory.
-% 
-% Examples:
-% \begin{itemize}
-% \item{plain text:}\\ \verb#    downloaddir="/home/johndoe/arc-downloads"#
-% \item{XML:}\\ \verb#    <downloaddir>/tmp/johndoe</downloaddir>#
-% \end{itemize}
-% 
-% 
-% \phantomsection
-% \index{configuration:alias}\addcontentsline{toc}{subsection}{alias}
-% \hspace*{0.5cm}
-% \begin{shaded}
-%   \uicommand{alias}
-% \end{shaded}
-% \textbf{This attribute is multi-valued.}
-% 
-% Alias substitutions to perform in connection with the \verb#-c# command line
-% switch of the ARC clients. Alias definitions are recursive. Any alias
-% defined in a block that is read before a given block can be used in
-% alias definitions in that block. An alias defined in a block can also
-% be used in alias definitions later in the same block.
-% 
-% Examples:
-% \begin{itemize}
-% \item{plain text:}\\ \verb#    alias="host1=somehost.nbi.dk"#\\
-% \verb#    alias="host2=otherhost.uu.se"#\\
-% \verb#    alias="myhosts=host1 host2"#
-% \item{XML:}\\ \verb#    <alias>host1=somehost.nbi.dk</alias>#\\
-% \verb#    <alias>host2=otherhost.uu.se</alias>#\\
-% \verb#    <alias>myhosts=host1 host2</alias>#
-% \end{itemize}
-% 
-% With the example above, \verb#ngsub -c myhosts# will resolve to \verb#ngsub -c somehost.nbi.dk -c otherhost.uu.se#.
-% 
-% \phantomsection
-% \index{configuration:broker}\addcontentsline{toc}{subsection}{broker}
-% \hspace*{0.5cm}
-% \begin{shaded}
-%   \uicommand{broker}
-% \end{shaded}
-% Configures which brokering algorithm to use during job submission. The
-% default one is the \texttt{FastestCpus} broker that chooses, among the
-% possible targets, the target with the fastest CPUs. Another
-% possibility is, for example, the
-% \texttt{RandomSort} broker that chooses the target randomly among the targets
-% surviving the job description matchmaking.
-% 
-% Examples:
-% \begin{itemize}
-% \item{plain text:}\\ \verb#    broker="RandomSort"#
-% \item{XML:}\\ \verb#    <broker>RandomSort</broker>#
-% \end{itemize}
+ Example:
+\verb#    brokername=Data#
+
+\phantomsection
+\index{configuration:brokerarguments}\addcontentsline{toc}{subsection}{brokerarguments}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{brokerarguments}
+\end{shaded}
+
+This attribute is used in case a broker comes with arguments.
+
+Example:
+\verb#    brokerarguments=cow#
+
+\phantomsection
+\index{configuration:joblist}\addcontentsline{toc}{subsection}{joblist}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{joblist}
+\end{shaded}
+
+Path to the job list file. This file will be used by commands such as \texttt{arcsub}, \texttt{arcstat},
+\texttt{arcsync} etc. to read and write information about jobs. This attribute
+corresponds to the \verb#-j# command line option. The default
+location of the file is in the {\$}HOME/.arc/client.conf directory with the
+name \texttt{jobs.xml}.
+
+Example:
+\begin{verbatim}
+    joblist=/home/user/run/jobs.xml
+    joblist=C:\\run\jobs.xml
+\end{verbatim}
+
+\phantomsection
+\index{configuration:bartender}\addcontentsline{toc}{subsection}{bartender}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{bartender}
+\end{shaded}
+
+Specifies default \textit{Bartender} services. Multiple Bartender URLs should
+be separated with a blank space.
+These URLs are used by the \texttt{chelonia} command line tool, the Chelonia
+FUSE plugin and by the data tool commands \texttt{arccp}, \texttt{arcls}, \texttt{arcrm}, etc..
+
+Example:
+\verb#    bartender=http://my.bar.com/tender#
+
+\phantomsection
+\index{configuration:proxypath}\addcontentsline{toc}{subsection}{proxypath}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{proxypath}
+\end{shaded}
+
+Specifies a non-standard location of proxy certificate. It is used by
+\texttt{arcproxy} or similar tools during proxy generation, and all other tools
+during establishing of a secure connection. This attribute
+corresponds to the \verb#-P# command line option of \texttt{arcproxy}.
+
+Example:
+\verb#    proxypath=/tmp/my-proxy#
+
+\phantomsection
+\index{configuration:keypath}\addcontentsline{toc}{subsection}{keypath}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{keypath}
+\end{shaded}
+
+Specifies a non-standard location of user's private key. It is used by
+\texttt{arcproxy} or similar tools during proxy generation. This attribute
+corresponds to the \verb#-K# command line option of \texttt{arcproxy}.
+
+Example:
+\verb#    keypath=/home/username/key.pem#
+
+\phantomsection
+\index{configuration:certificatepath}\addcontentsline{toc}{subsection}{certificatepath}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{certificatepath}
+\end{shaded}
+
+Specifies a non-standard location of user's public certificate. It is used by
+\texttt{arcproxy} or similar tools during proxy generation. This attribute
+corresponds to the \verb#-C# command line option of \texttt{arcproxy}.
+
+Example:
+\verb#    certificatepath=/home/username/cert.pem#
+
+\phantomsection
+\index{configuration:cacertificatesdirectory}\addcontentsline{toc}{subsection}{cacertificatesdirectory}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{cacertificatesdirectory}
+\end{shaded}
+
+Specifies non-standard location of the directory containing CA-certificates.
+This attribute
+corresponds to the \verb#-T# command line option of \texttt{arcproxy}.
+
+Example:
+\verb#    cacertificatesdirectory=/home/user/cacertificates#
+
+\phantomsection
+\index{configuration:cacertificatepath}\addcontentsline{toc}{subsection}{cacertificatepath}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{cacertificatepath}
+\end{shaded}
+
+Specifies an explicit path to the certificate of the CA that issued user's credentials.
+
+Example:
+\verb#    cacertificatepath=/home/user/myCA.0#
+
+\phantomsection
+\index{configuration:vomsserverpath}\addcontentsline{toc}{subsection}{vomsserverpath}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{vomsserverpath}
+\end{shaded}
+
+Specifies non-standard path to the file which contians list of VOMS services and
+associated configuration parameters. This attribute
+corresponds to the \verb#-V# command line option of \texttt{arcproxy}.
+
+Example:
+\verb#    vomsserverpath=/etc/voms/vomses#
+
+\phantomsection
+\index{configuration:username}\addcontentsline{toc}{subsection}{username}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{username}
+\end{shaded}
+
+Sets default username to be used for requesting credentials from Short Lived
+Credentials Service. This attribute
+corresponds to the \verb#-U# command line option of \texttt{arcslcs}.
+
+Example:
+\verb#    username=johndoe#
+
+\phantomsection
+\index{configuration:password}\addcontentsline{toc}{subsection}{password}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{password}
+\end{shaded}
+
+Sets default password to be used for requesting credentials from Short Lived
+Credentials Service. This attribute
+corresponds to the \verb#-P# command line option of \texttt{arcslcs}.
+
+Example:
+\verb#    password=secret#
+
+\phantomsection
+\index{configuration:keypassword}\addcontentsline{toc}{subsection}{keypassword}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{keypassword}
+\end{shaded}
+
+Sets default password to be used to encode the private key of credentials obtained
+from a Short Lived Credentials Service. This attribute
+corresponds to the \verb#-K# command line option of \texttt{arcslcs}.
+
+Example:
+\verb#    keypassword=secret2#
+
+\phantomsection
+\index{configuration:keysize}\addcontentsline{toc}{subsection}{keysize}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{keysize}
+\end{shaded}
+
+Sets size (strength) of the private key of credentials obtained from a Short
+Lived Credentials Service. Default value is 1024. This attribute
+corresponds to the \verb#-Z# command line option of \texttt{arcslcs}.
+
+Example:
+\verb#    keysize=2048#
+
+\phantomsection
+\index{configuration:certificatelifetime}\addcontentsline{toc}{subsection}{certificatelifetime}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{certificatelifetime}
+\end{shaded}
+
+Sets lifetime (in hours, starting from current time) of user certificate which
+will be obtained from a Short Lived Credentials Service. This attribute
+corresponds to the \verb#-L# command line option of \texttt{arcslcs}.
+
+Example:
+\verb#    certificatelifetime=12#
+
+\phantomsection
+\index{configuration:slcs}\addcontentsline{toc}{subsection}{slcs}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{slcs}
+\end{shaded}
+
+Sets the URL to the Short Lived Certificate Service. This attribute
+corresponds to the \verb#-S# command line option of \texttt{arcslcs}.
+
+Example:
+\verb#    slcs=https://127.0.0.1:60000/slcs#
+
+\phantomsection
+\index{configuration:storedirectory}\addcontentsline{toc}{subsection}{storedirectory}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{storedirectory}
+\end{shaded}
+
+Sets directory which will be used to store credentials obtained from a Short
+Lived Credential Servise. This attribute
+corresponds to the \verb#-D# command line option of \texttt{arcslcs}.
+
+Example:
+\verb#    storedirectory=/home/mycredentials#
+
+\phantomsection
+\index{configuration:idpname}\addcontentsline{toc}{subsection}{idpname}
+\hspace*{0.5cm}
+\begin{shaded}
+  \uicommand{idpname}
+\end{shaded}
+
+Sets Identity Provider name (Shibboleth) to which user belongs. It is used for
+contacting Short Lived Certificate Services. This attribute
+corresponds to the \verb#-I# command line option of \texttt{arcslcs}.
+
+Example:
+\verb#    idpname=https://idp.testshib.org/idp/shibboleth#
+
+\section{Block \texttt{[alias]}}
+
+Users often prefer to submit jobs to a specific site; since contact URLs (and
+especially end-point references) are very long, it is very convenient to replace
+them with aliases. Block \texttt{[alias]} simply contains a list of
+alias-value pairs.
+
+ Alias substitutions is performed in connection with the \verb#-c# command line
+ switch of the ARC clients.
+
+ Aliases can refer to a list of services (separated by a blank space).
+
+ Alias definitions can be recursive. Any alias
+ defined in a list that is read before a given list can be used in
+ alias definitions in that list. An alias defined in a list can also
+ be used in alias definitions later in the same list.
+
+ Examples:
+\begin{verbatim}
+[alias]
+
+arc0=computing:ARC0:ldap://ce.ng.org:2135/nordugrid-cluster-name=ce.ng.org,Mds-Vo-name=local,o=grid
+arc1=computing:ARC1:https://arex.ng.org:60000/arex
+cream=computing:CREAM:ldap://cream.glite.org:2170/o=grid
+unicore=computing:UNICORE:https://bes.unicore.org:8080/test/services/BESFactory?res=default_bes
+crossbrokering=arc0 arc1 cream unicore
+\end{verbatim}
 
@@
 ARC configuration file in releases 0.6 and 0.8 has the same name and the same
-format; attributes unknown to newer ARC versions are ignored.
+format. Only one attribute is preserved (\texttt{timeout}); other attributes
+unknown to newer ARC versions are ignored.
 

### 2009-10-25 doc/manuals/user_interface/ui-job.tex
Added Proxy and Configuration sections

@@
    Options:&&\\
-   \texttt{ -c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject a specific site (cluster)\\
-   \texttt{ -i, --indexurl}&\textit{url}&URL of an index server\\
-   \texttt{ -e, --jobdescrstring}&\textit{filename}&string describing the job to be submitted\\
-   \texttt{ -f, --jobdescrfile}&\textit{filename}&file describing the job to be submitted\\
-   \texttt{ -j, --joblist}&\textit{filename}&file where the job IDs will be stored\\
-   \texttt{ -D, --dryrun}&&add dryrun option to the job description\\
-   \texttt{ -x, --dumpdescription}&&do not submit -- dump transformed job description to stdout\\
-   \texttt{ -t, --timeout}&\textit{time}&timeout for queries in seconds (default 20)\\
-   \texttt{ -d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
-   \texttt{ -U, --unknownattr}&&allow unknown attributes in the job description\\
-   \texttt{ -v, --version}&&print version information\\
-   \texttt{ -?, --help}&&print help page\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject a specific site (cluster)\\
+   \texttt{-i, --indexurl}&\textit{url}&URL of an index server\\
+   \texttt{-e, --jobdescrstring}&\textit{filename}&string describing the job to be submitted\\
+   \texttt{-f, --jobdescrfile}&\textit{filename}&file describing the job to be submitted\\
+   \texttt{-j, --joblist}&\textit{filename}&file where the job IDs will be stored\\
+   \texttt{-D, --dryrun}&&add dryrun option to the job description\\
+   \texttt{-x, --dumpdescription}&&do not submit -- dump transformed job description to stdout\\
+   \texttt{-U, --unknownattr}&&allow unknown attributes in the job description\\
+   \texttt{-t, --timeout}&\textit{time}&timeout for queries in seconds (default 20)\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-v, --version}&&print version information\\
+   \texttt{-h, --help}&&print help page\\
 %   Arguments:&&\\
@@
    \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
    \texttt{-v, --version}& & print version information\\
@@
    \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
-   \texttt{-f, --force} & & removes the job ID from the local list\\
-   & &~~~even if the job is not found on the Grid\\
+   \texttt{-f, --force} & & removes the job ID from the local list even if the job is not found on the Grid\\
    \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
    \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
    \texttt{-v, --version}& & print version information\\

### 2009-10-25 doc/manuals/user_interface/ui-proxy.tex
Added Proxy and Configuration sections

\subsection{arcproxy}
\label{sec:arcproxy}

\index{arcproxy}\index{commands:arcproxy}In order to contact Grid services
(submit jobs, copy data, check information etc),
one has to present valid credentials. The are commonly formalized as so-called
``proxy'' certificates. There are many different types of proxy certificates,
with different Grids and different services having own preferences. \texttt{arcproxy}
is a powerful tool that can be used to generate most commonly used proxies. It supports
the following types:
\begin{itemize}
  \item pre-RFC GSI proxy
  \item RFC-compliant proxy (default)
  \item VOMS-extended proxy
  \item MyProxy delegation
\end{itemize}

\texttt{arcproxy} requires presence of user's private key and public certificate,
as well as the public certificate of their issuer CA.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcproxy [options]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-P, --proxy}& \textit{path} & path to the proxy file\\
   \texttt{-C, --cert}& \textit{path} & path to the certificate file\\
   \texttt{-K, --key}& \textit{path} & path to the key file\\
   \texttt{-T, --cadir}& \textit{path} & path to the trusted certificate directory, only needed for VOMS client functionality\\
   \texttt{-V, --vomses}& \textit{path} & path to the VOMS server configuration file\\
   \texttt{-S, --voms}& \textit{voms[:command]} & Specify VOMS server (more than one VOMS server can be specified like this:\\
   & &--voms VOa:command1 --voms VOb:command2)\\
   & &:command is optional, and is used to ask for specific attributes(e.g. roles). Command options are:\\
   & &all -- put all of this DN's attributes into AC;\\
   & &list -- list all of the DN's attribute,will not create AC extension;\\
   & &/Role=yourRole -- specify the role, if this DN has such a role, the role will be put into AC\\
   & &/voname/groupname/Role=yourRole -- specify the VO,group and role; if this DN has such a role, the role will be put into AC\\
   \texttt{-G, --gsicom}& & use GSI communication protocol for contacting VOMS services\\
   \texttt{-O, --old}& & use GSI proxy (default is RFC 3820 compliant proxy)\\
   \texttt{-I, --info}& & print all information about this proxy. In order to show the Identity (DN without CN as suffix for proxy) of the certificate, the 'trusted certdir' is needed.\\
   \texttt{-U, --user}& \textit{string} & username for MyProxy server\\
   \texttt{-L, --myproxysrv}& \textit{URL} & URL of MyProxy server\\
   \texttt{-M, --myproxycmd}& \textit{PUT|GET} & command to MyProxy server. The command can be PUT and GET.\\
   & &PUT/put -- put a delegated credential to MyProxy server;\\
   & &GET/get -- get a delegated credential from MyProxy server, credential (certificate and key) is not needed in this case.\\
   \texttt{-c, --constraint}& \textit{string} & proxy constraints\\
   \texttt{-t, --timeout}& \textit{seconds} & timeout in seconds (default 20 seconds)\\
   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
   \texttt{-v, --version}& & print version information\\
   \texttt{-h, --help}& & print help page\\
\end{longtable}

Supported constraints are:
\begin{itemize}
  \item \texttt{validityStart=time} -- e.g. 2008-05-29T10:20:30Z; time when certificate becomes valid. Default is now.
  \item \texttt{validityEnd=time} -- time when certificate becomes invalid. Default is 43200 (12 hours) from start.
  \item \texttt{validityPeriod=time} -- e.g. 43200 or 12h or 12H; for how long certificate is valid. If neither \texttt{validityPeriod} nor \texttt{validityEnd} are specified, default is 12 hours
  \item \texttt{vomsACvalidityPeriod=time} -- e.g. 43200 or 12h or 12H; for how long the AC is valid. Default is the same as \texttt{validityPeriod}.
  \item \texttt{proxyPolicy=policy content} -- assigns specified string to proxy prolicy to limit it's functionality.
  \item \texttt{proxyPolicyFile=policy file}
\end{itemize}

MyProxy functionality can be used together with VOMS functionality.

\subsection{arcslcs}
\label{sec:arcslcs}

\index{arcslcs}\index{commands:arcslcs}This utility generates  short-lived
credential  based  on the credential to IdP in SAML2SSO profile (normally the
username/password to Shibboleth IdP).

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcslcs [options]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-S, --ur;}& \textit{URL} & URL of SLCS Service (e.g. https://127.0.0.1:60000/slcs)\\
   \texttt{-I, --idp}& \textit{URL} & the name of IdP (e.g. https://idp.testshib.org/idp/shibboleth)\\
   \texttt{-U, --user}& \textit{string} & User account to IdP\\
   \texttt{-P, --password}& \textit{string} & password for user accoutn to IdP\\
   \texttt{-Z, --keysize}& \textit{integer} & size of the private key, default is 1024\\
   \texttt{-K, --keypass}& \textit{} & passphrase for protecting the private key; if not set, the private key file will not be protected by the passphrase.\\
   \texttt{-L, --lifetime}& \textit{hours} & life time of the credential (hours)), starting with current time\\
   \texttt{-D, --storedir}& \textit{path} & store directory of the credential\\
   \texttt{-t, --timeout}& \textit{seconds} & timeout in seconds (default 20 seconds)\\
   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
   \texttt{-c, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
   \texttt{-v, --version}& & print version information\\
   \texttt{-h, --help}& & print help page\\
\end{longtable}

### 2009-10-25 doc/manuals/user_interface/ui-data.tex
arcsync, arcsub and chelonia fixed

@@
 This command is not implemented.
+
+\subsection{chelonia}\label{sec:chelonia}
+
+\index{chelonia}\index{commands:chelonia}\texttt{chelonia} is a client tool for
+accessing the Chelonia storage system. With it you can
+create, remove and list file collections, upload, download and remove files,
+and move and stat collections and files, using Logical Names (LN).
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia [options] $<$method$>$ [arguments]}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+   Options:&&\\
+   \texttt{-b} & \textit{URL} & URL of Bartender to connect\\
+   \texttt{-x} && print SOAP XML messages\\
+   \texttt{-v} && verbose mode\\
+   \texttt{-z}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-w} && allow to run without the ARC python client libraries (with limited functionality)\\
+   Methods:&&\\
+   \texttt{stat} &\textit{LN [LN ...]}& get detailed information about an entry or several\\
+   \texttt{makeCollection, make, mkdir} &\textit{LN}& create a collection\\
+   \texttt{unmakeCollection, unmake, rmdir} &\textit{LN}& remove an empty collection\\
+   \texttt{list, ls} &\textit{LN}& list the content of a collection\\
+   \texttt{move, mv} &\textit{source target}& move entries within the namespace (both LNs)\\
+   \texttt{putFile, put} &\textit{source target}& upload a file from a \textit{source} to a \textit{target} (both specified as LNs))\\
+   \texttt{getFile, get} &\textit{source [target]}& download a file from a \textit{source} to a \textit{target}\\
+   \texttt{delFile, del, rm} &\textit{LN [LN ...]}& remove file(s))\\
+   \texttt{modify, mod} &\textit{string}& modify metadata\\
+   \texttt{policy, pol} &\textit{string}& modify access policy rules. The string has a form $<$LN$>$ $<$changeType$>$ $<$identity$>$ $<$action list$>$.\\
+   & &$<$changeType$>$ could be 'set', 'change' or 'clear'\\
+   & & 'set': sets the action list to the given user overwriting the old one\\
+   & & 'change': modify the current action list with adding and removing actions\\
+   & & 'clear': clear the action list of the given user\\
+   & &$<$identity$>$ could be a '$<$UserDN$>$' or a 'VOMS:$<$VO name$>$'\\
+   & &$<$action list$>$ is a list actions prefixed with '+' or '-', e.g.
+'+read +addEntry -delete'; possible actions are: \texttt{read}, \texttt{addEntry}, \texttt{removeEntry}, \texttt{delete}, \texttt{modifyPolicy}, \texttt{modifyStates}, \texttt{modifyMetadata}\\
+   \texttt{unlink} &\textit{string}& remove a link to an entry from a collection without removing the entry itself\\
+   \texttt{credentialsDelegation, cre} &\textit{string}& delegate credentials for using gateway\\
+   \texttt{removeCredentials, rem} &\textit{string}& remove previously delegated credentials\\
+   \texttt{makeMountPoint, makemount} &\textit{string}& create a mount point\\
+\end{longtable}
+
+Without arguments, each method prints its own help.
+
+Examples:
+\begin{verbatim}
+    chelonia list /
+    chelonia put orange /
+    chelonia stat /orange
+    chelonia get /orange /tmp
+    chelonia mkdir /fruits
+    chelonia mkdir /fruits/apple
+    chelonia mv /orange /fruits
+    chelonia ls /fruits
+    chelonia rmdir /fruits/apple
+    chelonia rmdir /fruits
+    chelonia rm /fruits/orange
+    chelonia policy / change ALL +read +addEntry
+    chelonia modify /pennys-orange set states neededReplicas 2
+\end{verbatim}

### 2009-10-25 doc/manuals/user_interface/ui-job.tex
arcsync, arcsub and chelonia fixed

@@
+\subsection{arcsync}
+\label{sec:arcsync}
+
+It is advised to start every grid session by running \texttt{arcsync},
+especially when changing workstations. The reason is that your job submission
+history is cached on your machine, and if you are using ARC client
+installations on different machines, your local lists of submitted jobs will be different. To synchronise
+these lists with the information in the Information System, use the
+\texttt{arcsync} \index{arcsync}\index{commands:arcsync} command.
+
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{arcsync [options]}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+   Options:&&\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site\\
+   \texttt{-i, --index}&\textit{url}&explicitly select or reject (-) a specific index server\\
+   \texttt{-j, --joblist}&\textit{filename}&file where user's job information will be stored\\
+   \texttt{-f, --force}& &don't ask for confirmation\\
+%   \texttt{-m, --merge}& &merge the found jobs with the jobs already in the joblist\\
+   \texttt{-T, --truncate}& &truncate the job list before synchronising\\
+   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-v, --version}&&print version information\\
+   \texttt{-h, --help}&&print help page\\
+\end{longtable}
+
+The ARC client keeps a local list of jobs in the user's home
+directory. If this file is lost,
+corrupt, or the user wants to recreate the file on a different
+workstation, the \texttt{arcsync} command will recreate this file from
+the information available in the Information System.
+
+Since  the  information  about  a job retrieved from a clsuter can be slightly out of date if the user very recently
+submitted or removed a job, a warning is issued when this command is run. The \verb#--force# option disables this warning.
+
+%If a specific cluster is selected for the syncronisation while the job list contains jobs  on  a  different  cluster,
+%running  the arcsync command will remove the old jobs from the joblist, unless the \verb#--merge# option is given, in which
+%case the new found jobs are merged with the jobs already in the joblist.
+
 \subsection{arcsub}\label{sec:arcsub}
@@
 most essential one, as it is used for submitting jobs to the Grid
-resources. \index{submit job}. \texttt{arcsub} matches user's job
+resources. \index{submit job}\texttt{arcsub} matches user's job
 description to the information collected from the Grid, and the
@@
 \begin{shaded}
-   \uicommand{arcsub [options]}%$<$task ...$>$
+   \uicommand{arcsub [options] [filename]}
 \end{shaded}
@@
    Options:&&\\
-   \texttt{-c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject a specific site (cluster)\\
-   \texttt{-i, --indexurl}&\textit{url}&URL of an index server\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject (-) a specific site\\
+   \texttt{-i, --index}&\textit{url}&explicitly select or reject (-) a specific index server\\
    \texttt{-e, --jobdescrstring}&\textit{filename}&string describing the job to be submitted\\
    \texttt{-f, --jobdescrfile}&\textit{filename}&file describing the job to be submitted\\
-   \texttt{-j, --joblist}&\textit{filename}&file where the job IDs will be stored\\
-   \texttt{-D, --dryrun}&&add dryrun option to the job description\\
+   \texttt{-j, --joblist}&\textit{filename}&file where user's job information will be stored\\
+%   \texttt{-D, --dryrun}&&add dryrun option to the job description\\
    \texttt{-x, --dumpdescription}&&do not submit -- dump transformed job description to stdout\\
-   \texttt{-U, --unknownattr}&&allow unknown attributes in the job description\\
-   \texttt{-t, --timeout}&\textit{time}&timeout for queries in seconds (default 20)\\
+%   \texttt{-U, --unknownattr}&&allow unknown attributes in the job description\\
+   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
    \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
    \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-b, --broker}&\textit{string}&select broker method (default is Random)\\
+   \texttt{-n, --dolocalsandbox}& &store job descriptions in local sandbox (useful for eventual resubmission/migration)\\
    \texttt{-v, --version}&&print version information\\
    \texttt{-h, --help}&&print help page\\
-%   Arguments:&&\\
-%   \texttt{ task ...}&&strings or files describing the jobs to be submitted\\
+   Arguments:&&\\
+   \texttt{filename}&&file describing the job(s) to be submitted\\
 \end{longtable}
@@
 \begin{framed}
-The -c and -i arguments accept ``urls'' of the format GRID:URL,
-
-e.g. for index servers:
-
-ARC0:ldap://grid.tsl.uu.se:2135/mds-vo-name=sweden,O=grid
-
-CREAM:ldap://cream.grid.upjs.sk:2170/o=grid
-
+The \verb#-c# and \verb#-i# arguments accept meta-URLs of the format \texttt{GRID:URL}, where \texttt{GRID} indicates a Grid middleware flavour.
+Possible flavours are \texttt{ARC0}, \texttt{ARC1}, \texttt{CREAM} and \texttt{UNICORE}.
+For example, for index servers:
+\begin{verbatim}
+ ARC0:ldap://index.ng.org:2135/mds-vo-name=sweden,O=grid
+ CREAM:ldap://cream.glite.org:2170/o=grid
+\end{verbatim}
 or clusters:
 
-ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid
+\verb# ARC0:ldap://ce.ng.eu:2135/nordugrid-cluster-name=ce.ng.eu,Mds-Vo-name=local,o=grid#
 \end{framed}
@@
 \begin{shaded}
-\uicommand{arcsub -c ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid -f job.jsdl}
+ arcsub -c ARC0:ldap://grid.ng.org:2135/nordugrid-cluster-name=grid.ng.org,Mds-Vo-name=local,o=grid -f job.jsdl
 \end{shaded}
 
-Assuming that the url represents an ARC0 cluster that the user is authorized to submit to and job.jsdl is the file below.
+A user has to have valid credentials (see Section~\ref{sec:proxy}) and be authorised at the specified site. The test file \texttt{job.jsdl} is shown below.
 
@@
 \end{framed}
+
+\begin{framed}
+   If a job is successfully submitted, a \textbf{job identifier}
+   (\textit{job ID})\index{job ID} is printed to standard output.
+\end{framed}
+
+The job ID uniquely identifies the job while it is being executed. Job IDs
+differ strongly between Grid flavours, but basically they have a form of a URL.
+You should use Job ID as a handle to refer to the job when doing other
+job manipulations, such as querying job status (\verb#arcstat#),
+killing it (\verb#arckill#), re-submitting (\verb#arcresub#), or
+retrieving the result (\verb#arcget#).
+
+\begin{framed}
+   Every job ID is a valid URL for the job session directory. You can
+   always use it to access the files related to the job, by using data
+   management tools (see Chapter~\ref{sec:dm}).
+\end{framed}
+
+The job description in xRSL or JSDL format can be given either as an
+argument on the command line, or can be read from a file. Several jobs can be
+requested at the same time by giving more than one filename
+argument, or by repeating the \verb#-f# or \verb#-e# options. It is
+possible to mix \verb#-e# and \verb#-f# options in the same
+\texttt{arcsub} command.
+
+The \verb#-c# option can be used to \textbf{force} a job to be
+submitted to a particular site (cluster), or to reject submission to a
+site. The matching is done by string match to the site name (i.e.
+hostname) as defined in the Information System, or to an alias, as
+defined in user configuration file (see
+Section~\ref{sec:client.conf}).  The \verb#-c# option can be repeated
+several times, for example:
+\begin{verbatim}
+    arcsub -c alias1 -c alias2 myjob.xrsl
+\end{verbatim}
+This will submit a job to either \verb#alias1# or
+\verb#alias2#. To submit a job to any site except
+\verb#badsite#, use \verb#-# sign in front of the name:
+\begin{verbatim}
+    arcsub -c -badsite myjob.xrsl
+\end{verbatim}
+
+If option \verb#-c# is not given, the \verb#arcsub# command locates the available sites by querying the
+Information System. Default index services for the Information System are specified in the
+configuration template distributed with the middleware, and can be overwritten both in the user's
+configuration (see Section~\ref{sec:client.conf}) and from the command line using option
+\verb#-i#. Different Grids use different notation for such index services.
+
+If you would like to get diagnostics of the process of resource
+discovery and requirements matching, a very useful option is
+\verb#-d#. The following command:
+\begin{verbatim}
+    arcsub -d DEBUG myjob.xrsl
+\end{verbatim}
+
+will print out the steps taken by the ARC client to find the
+best cluster satisfying your job requirements. A default value for
+the debug level can be set in the user's configuration file.
+
+It often happens that some sites that \verb#arcsub# has to contact
+are slow to answer, or are down altogether. This will not prevent
+you from submitting a job, but will slow down the submission. To
+speed it up, you may want to specify a shorter timeout (default is
+20 seconds) with the \verb#-t# option:
+\begin{verbatim}
+    arcsub -t 5 myjob.xrsl
+\end{verbatim}
+
+A default value for the timeout can be set in the user's configuration file.
+
+The user interface transforms input job description into a format
+that can be understood by the Grid services to which it is being
+submitted. By specifying the \verb#-dumpdescription# option, such transformed
+description is written to stdout instead of being submitted to the remote site.
+
+Possible broker values for the option \verb#-b# are:
+\begin{itemize}
+ \item[\texttt{Random}] -- ranks targets randomly (default)
+ \item[\texttt{FastestQueue}] -- ranks targets according to their queue length
+% \item[\texttt{FastestCPU}] -- ranks targets according to CPU performance
+ \item[\texttt{Data}] -- ranks target according the amount of megabytes of the
+requested input files that are already in the computing resource’s cache.
+% \item[\texttt{SampleBroker}] -- Ranks targets randomly (written in Python)
+\end{itemize}
 
@@
 %
-% \subsection{arcsync}
-% \label{sec:arcsync}
-% 
-% If you are using User Interface installations on different machines,
-% your local lists of submitted jobs will be different. To synchronise
-% these lists with the information in the Information System, use the
-% \texttt{arcsync} \index{arcsync}\index{commands:arcsync} command.
-% 
-% \hspace*{0.5cm}
-% \begin{shaded}
-%    \uicommand{arcsync [options]}
-% \end{shaded}
-% \versions{ARC 0.9}
-% \begin{longtable}{llp{8cm}}
-%    Options:&&\\
-%    \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
-%    \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
-%    \texttt{-g, -giisurl}& \textit{url} &URL of a central Information System server\\
-%    \texttt{-G, -giislist}&\textit{filename}&list of GIIS URLs\\
-%    \texttt{-f, -force}&&don't ask for confirmation\\
-%    \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
-%    \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
-%    \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
-%    \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
-%    \texttt{-v, -version}& & print version information\\
-%    \texttt{-h, -help}& & print help page\\
-% \end{longtable}
-% 
-% The ARC User Interface keeps a local list of jobs in the user's home
-% directory (see section~\ref{sec:joblists}). If this file is lost,
-% corrupt, or the user wants to recreate the file on a different
-% workstation, the \texttt{arcsync} command will recreate this file from
-% the information available in the Information System.
 

### 2009-10-28 doc/tech_doc/client/client_technical.tex
Broker name sync with the current broker loader name

--- 2009-03-20 doc/client/client_technical.tex
@@
 \newcommand{\Broker}{\texttt{Broker}}
-\newcommand{\RandomBroker}{\texttt{RandomBroker}}
-\newcommand{\FastestCPUBroker}{\texttt{FastestCPUBroker}}
+\newcommand{\Random}{\texttt{Random}}
+\newcommand{\Benchmark}{\texttt{Benchmark}}
 \newcommand{\FastestQueueBroker}{\texttt{FastestQueueBroker}}
 \newcommand{\PythonBroker}{\texttt{PythonBroker}}
-\newcommand{\DataBroker}{\texttt{DataBroker}}
+\newcommand{\Data}{\texttt{Data}}
 \newcommand{\ClientInterface}{\texttt{ClientInterface}}
@@
 
-\subsubsection{\RandomBroker}
-
-The {\RandomBroker} ranks the {\ExecutionTarget}s randomly.
-
-\subsubsection{\FastestCPUBroker}
-
-The {\FastestCPUBroker} ranks the {\ExecutionTarget}s according to their CPU performance. The most 
-reliable comparison is achieved through the usage of a benchmark scenario, and the {\FastestCPUBroker}
+\subsubsection{\Random}
+
+The {\Random} ranks the {\ExecutionTarget}s randomly.
+
+\subsubsection{\Benchmark}
+
+The {\Benchmark} ranks the {\ExecutionTarget}s according to their CPU performance. The most 
+reliable comparison is achieved through the usage of a benchmark scenario, and the {\Benchmark}
 uses the CINT2000 (Integer Component of SPEC CPU2000)\footnote{http://www.spec.org/cpu2000/CINT2000/} 
@@
 
-\subsubsection{\DataBroker}
-
-The {\DataBroker} ranks the {\ExecutionTarget}s according to how many megabytes of the requested input files that already 
+\subsubsection{\Data}
+
+The {\Data} ranks the {\ExecutionTarget}s according to how many megabytes of the requested input files that already 
 stored in the cache of the computing resource the {\ExecutionTarget} represents. The broker is motivated by that jobs should 

### 2009-10-28 doc/tech_doc/jura/jura-tech-doc.tex
Installation instructions: Win compatibility

@@
 
-A-REX executes the reporting executable through the file ``logger'' in
-the \textit{libexec/arc} directory under the install location (this is
-an A-REX legacy kept for backwards compatibility with the old
-logger). Therefore, to enable automatic periodic reporting, a symbolic
-link called ``logger'' should be placed in \textit{libexec}, pointing
-to the JURA executable.
+A-REX executes the reporting executable through the file
+``\textit{logger}'' in the \textit{libexec/arc} directory under the
+install location (this is an A-REX legacy kept for backwards
+compatibility with the old logger). Therefore, to enable automatic
+periodic reporting, the ``\textit{jura}'' executable file should be
+copied into \textit{libexec/arc}, or if possible, a symbolic link
+called ``\textit{logger}'' should be created there, pointing to the
+``\textit{jura}'' executable.
 

### 2009-10-28 doc/manuals/chelonia/chelonia-cli.tex
CLI manual: moved chelonia section under doc/chelonia, fixed Makefile

--- 2009-10-25 doc/manuals/user_interface/ui-data.tex
@@
-\subsection{arcls}\label{sec:arcls}
-The \texttt{arcls} \index{arcls}\index{commands:arcls} is a simple
-utility that allows to list contents and view some attributes of
-objects of a specified (by an URL) remote directory.
-
-\hspace*{0.5cm}
-\begin{shaded}
-   \uicommand{arcls [options] $<$URL$>$}
-\end{shaded}
-\versions{ARC 0.9}
-\begin{longtable}{llp{8cm}}
-    Options:&&\\
-    \texttt{-h} && short help\\
-    \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
-    \texttt{-l} &  & detailed listing\\
-    \texttt{-L} &  & detailed listing including URLs from which file can 
-    be downloaded and\\ & & temporary cached locations\\
-    \texttt{-m} && display all available metadata\\
-    Arguments:&&\\
-    \texttt{URL} && file or directory URL\\
-\end{longtable}
-
-This tool is very convenient not only because it allows to list files
-at a Storage Element or records in an indexing service, but also
-because it can give a quick overview of a job's working directory,
-which is explicitly given by job ID.
-
-Usage examples can be as follows:
-
-\begin{verbatim}
-    arcls rls://rc.host:38203/logical_file_name
-    arcls -l gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535
-    arcls -L srm://grid.uio.no:58000/srm/managerv1/johndoe/log2
-\end{verbatim}
-
-Examples of URLs accepted by this tool can be found in
-Section~\ref{sec:urls}, though \texttt{arcls} won't be able to list a
-directory at an HTTP server, as they normally do not return directory
-listings.
-
-\subsection{arccp}\label{sec:arccp}
-
-The \texttt{arccp} \index{arccp}\index{commands:arccp} is a powerful
-tool to copy files over the Grid. It is a part of the A-REX,
-but can be used by the User Interface as well.
-\hspace*{0.5cm}
-\begin{shaded}
-   \uicommand{arccp [options] $<$source$>$ $<$destination$>$}
-\end{shaded}
-\versions{ARC 0.9}
-\begin{longtable}{llp{8cm}}
-    Options:&&\\
-    \texttt{-h} && short help\\
-    \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
-    \texttt{-y} & \textit{cache\_path} & path to local cache (use to put file into cache)\\
-    \texttt{-p} && use passive transfer (does not work if secure is on, default if secure is not requested)\\
-    \texttt{-n} && do not try to force passive transfer\\
-    \texttt{-i} && show progress indicator\\
-    \texttt{-u} && use secure transfer (insecure by default)\\
-    \texttt{-r} & \textit{recursion\_level} & operate recursively (if possible) up to specified level (0 - no recursion)\\
-    \texttt{-R} & \textit{number} & how many times to retry transfer of every file before failing\\
-    \texttt{-t} & \textit{time} & timeout in seconds (default 20)\\
-    \texttt{-f} && if the destination is an indexing service and not the same as the source and the destination is already registered, then the copy is normally not done.
-However, if this option is specified the source is assumed to be a replica of the destination created in an uncontrolled way and the copy is done like in case of replication\\
-    \texttt{-T} && do not transfer file, just register it - destination must be non-existing meta-url\\
-    Arguments:&&\\
-    \texttt{source} && source URL\\
-    \texttt{destination} && destination URL\\
-\end{longtable}
-
-This command transfers contents of a file between 2 end-points.
-End-points are represented by URLs or meta-URLs. For supported
-endpoints please refer to Section~\ref{sec:urls}.
-
-\texttt{arccp} can perform multi-stream transfers if \texttt{threads}
-URL option is specified and server supports it.
-
-Source URL can end with \verb#"/"#. In that case, the whole fileset
-(directory) will be copied. Also, if the destination ends with
-\verb#"/"#, it is extended with part of source URL after last
-\verb#"/"#, thus allowing users to skip the destination file or
-directory name if it is meant to be identical to the source.
-
-Usage examples of \texttt{arccp} are:
-
-\begin{verbatim}
-    arccp gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535/job.out \
-              file:///home/myname/job2.out
-    arccp gsiftp://aftpexp.bnl.gov;threads=10/rep/my.file \
-              rc://;threads=4@grid.uio.no/lc=Collection,rc=Catalog/zebra4.f
-    arccp http://www.nordugrid.org/data/somefile gsiftp://hathi.hep.lu.se/data/
-\end{verbatim}
-
-\subsection{arcrm}\label{sec:arcrm}
-
-The \texttt{arcrm} \index{arcrm}\index{commands:arcrm}
-command allows users to erase files at any location specified by a
-valid URL.
-\hspace*{0.5cm}
-\begin{shaded}
-   \uicommand{arcrm [options] $<$source$>$}
-\end{shaded}
-\versions{ARC 0.9}
-\begin{longtable}{llp{8cm}}
-    Options:&&\\
-    \texttt{-h} && short help\\
-    \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
-    \texttt{-c} & &continue with meta-data even if it failed to delete real file\\
-    Arguments:&&\\
-    \texttt{source} && source URL\\
-\end{longtable}
-
-\begin{framed}
-   A convenient use for \texttt{arcrm} is to erase the files in a data
-   indexing catalog (RC, RLS or such), as it will not only remove the
-   physical instance, but also will clean up the database record.
-\end{framed}
-
-Here is an \texttt{arcrm} example:
-
-\begin{verbatim}
-    arcrm rc://grid.uio.no/lc=Collection,rc=Catalog/badfile
-\end{verbatim}
-
-\subsection{arcacl}\label{sec:arcacl}
-
-\index{arcacl}\index{commands:arcacl}This command retrieves or modifies
-access control information associated with a stored object if service
-supports GridSite GACL language~\cite{gacl} for access control.
-\hspace*{0.5cm}
-\begin{shaded}
-   \uicommand{arcacl [options] get$|$put $<$URL$>$}
-\end{shaded}
-\versions{ARC 0.9}
-\begin{longtable}{llp{8cm}}
-   Options:&&\\
-    \texttt{-d, -debug} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
-    \texttt{-v} && print version information\\
-    \texttt{-h} && short help\\
-   Arguments:&&\\
-    \texttt{get} &\textit{URL}& get Grid ACL for the object\\
-    \texttt{put} &\textit{URL}& set Grid ACL for the object\\
-    \texttt{URL} && object URL; curently only gsiftp and sse URLs are supported\\
-\end{longtable}
-
-ACL document (an XML file) is printed to standard output when
-\texttt{get} is requested, and is acquired from standard input when
-\texttt{set} is specified\footnote{In ARC $\leq$ 0.5.28, \texttt{set}
-  shoud be used instead of \texttt{put}}. Usage examples are:
-\begin{verbatim}
-    arcacl get gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename
-    arcacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename $<$ myacl
-\end{verbatim}
-
-\subsection{arctransfer}\label{sec:arctransfer}
-\index{arctransfer}\index{commands:arctransfer}The \texttt{arctransfer}
-This command is not implemented.
-
 \subsection{chelonia}\label{sec:chelonia}

### 2009-10-29 doc/manuals/user_interface/ui-brokers.tex
Added brokers description as a separate file, other small fixes

% text about brokers; may need to be extended and possibly re-used in another
% document, or moved as a separate subsection

Possible broker values for the \texttt{arcsub} command line option \verb#-b# are:
\begin{itemize}
 \item[--] \texttt{Random} -- ranks targets randomly (default)
 \item[--] \texttt{FastestQueue} -- ranks targets according to their queue length
 \item[--] \texttt{Benchmark[:name]} -- ranks targets according to a given benchmark, as specified by the \texttt{name}. If no benchmark is specified, CINT2000~\footnote{http://www.spec.org/cpu2000/CINT2000/} is used
 \item[--] \texttt{Data} -- ranks targets according the amount of megabytes of the
requested input files that are already in the computing resource’s cache.
 \item[--] \texttt{Python:$<$module$>$.$<$class$>$[:arguments]} -- ranks targets using any user-supplied custom Python broker module, optionally with broker arguments. Such module can reside anywhere in user's \texttt{PYTHONPATH}
 \item[--] \texttt{$<$otherbroker$>$[:arguments]} -- ranks targets using any user-supplied custom C++ broker plugin, optionally with broker arguments. Default location for broker plugins is \texttt{/usr/lib/arc} (may depend on the operating system), or the one specified by the \texttt{ARC\_PLUGIN\_PATH}.
\end{itemize}

To write a custom broker in C++ one has to write a new specialization of the \texttt{Broker} base class and implement the \texttt{SortTargets} method in the new class. The class should be compiled as a loadable
module that has the proper ARC plugin descriptor for the new broker. For example, to build a broker plugin ``MyBroker'' one executes:
\begin{lstlisting}
  g++ -I /arc-install/include \
      -L /arc-install/lib \
      `pkg-config --cflags glibmm-2.4 libxml-2.0` \
      -o libaccmybroker.so -shared MyBroker.cpp
\end{lstlisting}
For more details, refer to \textit{libarclib} documentation~\ref{cite:libarcclient}.

### 2009-10-29 doc/manuals/user_interface/ui-conf.tex
Broker name sync with the current broker loader name
Added brokers description as a separate file, other small fixes

@@
 Another possibility is, for example, the \texttt{FastestQueue} broker that
-chooses the target with the shortest estimated queue waiting time. Possible
-values are:
-\begin{itemize}
- \item[\texttt{Random}] -- ranks targets randomly (default)
- \item[\texttt{FastestQueue}] -- ranks targets according to their queue length
-% \item[\texttt{FastestCPU}] -- ranks targets according to CPU performance
- \item[\texttt{Data}] -- ranks target according the amount of megabytes of the
-requested input files that are already in the computing resource’s cache.
-% \item[\texttt{SampleBroker}] -- Ranks targets randomly (written in Python)
-\end{itemize}
+chooses the target with the shortest estimated queue waiting time. For an 
+overview of brokers, please refer to Section~\ref{sec:arcsub}.
 
@@
 
-This attribute is used in case a broker comes with arguments.
+This attribute is used in case a broker comes with arguments. This corresponds to the
+parameter that follows column in the \verb#-b# command line option.
 

### 2009-10-30 doc/manuals/chelonia/arc-storage-manual.tex
fixing the manual
updating the chelonia manual

@@
 \begin{verbatim}
-$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service.xml -f
+$ sudo /usr/sbin/arched -i /etc/arc/chelonia1.ini -f
 \end{verbatim}
@@
 \begin{verbatim}
-<Service name="pythonservice" id="bartender">
-    <ClassName>storage.bartender.bartender.BartenderService</ClassName>
-    <LibrarianURL>[...]</LibrarianURL>
-    <ProxyStore>/var/spool/arc/proxy_store</ProxyStore>
-    <GatewayClass>storage.bartender.gateway.gateway.Gateway</GatewayClass>
-    <GatewayCfg>
-        <ProxyStore>/var/spool/arc/proxy_store</ProxyStore>
-        <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
-    </GatewayCfg>
-    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
-</Service>
+[bartender]
+endpoint=https://chelonia3.knowarc.eu:60003/Bartender
+proxystore=/var/spool/arc/proxy_store
 \end{verbatim}
@@
 
-\textbf{This section was not updated.}
-
-ARC has its own multi-protocol storage client tools which provide a unified way to access different kind of storage.
-For all the supported types there is a DMC module.
-For Chelonia, the ARC DMC module is used.
-If it is compiled and installed, then you can use the \verb!arcls!, \verb!arccp!, \verb!arcrm!, etc.
-tools with the \verb!arc://!
-protocol to access the ARC storage system.
-You can specify the URL of the Bartender with the \verb!BartenderURL!
-URL option, or you can have a Bartender URL configured in the \verb!client.xml!
-user config.
+ARC has its own multi-protocol storage client tools which provide a
+unified way to access different kind of storage. For all the supported
+types there is a DMC module. For Chelonia, the ARC DMC module is
+used. If it is compiled and installed, then you can use the
+\verb!arcls!, \verb!arccp!, \verb!arcrm!, etc. tools with the
+\verb!arc://! protocol to access the ARC storage system. You can
+specify the URL of the Bartender with the \verb!BartenderURL! URL
+option, or you can have a Bartender URL configured in the
+\verb!client.ini! user config by adding an entry \verb!bartender!:
+
+\begin{verbatim}
+bartender= https://chelonia1/Bar https://chelonia2/Bar https://chelonia3/Bar
+\end{verbatim}
+
+If the \verb!client.ini! is not in the default path (for Linux,
+\verb!$HOME/.arc/client.conf!), you need to specify the path to the
+configuration file with the \verb!-z! option.
+
+% todo: make arc dmc look up isis
 
@@
 orange
-$ arcls arc:///?BartenderURL=https://storage2:60006/Bartender
+$ arcls arc:///?BartenderURL=https://chelonia1.knowarc.eu:60001/Bartender
 orange
@@
 
-\textbf{This section was not updated.}
-
 Besides the \verb!chelonia! tool you can access the storage system with the use of a FUSE module. The following steps installs the ARC storage FUSE module:
@@
 
-\item Set up \verb!~/.arc/client.xml!:
-
-\begin{verbatim}
-$ cat ~/.arc/client.xml 
-<ArcConfig>
-  <KeyPath>/home/<username>/.arc/userkey.pem</KeyPath>
-  <CertificatePath>/home/<username>/.arc/usercert.pem</CertificatePath>
-  <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
-  <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
-</ArcConfig>
-\end{verbatim}
-
-\item Create a mountpoint and mount arcfs.py:
+\item Make sure \verb!bartender! is set in
+  \verb!$HOME/.arc/client.conf!:
+\begin{verbatim}
+$ grep bartender ~/.arc/client.conf 
+bartender= https://chelonia1.knowarc.eu:60001/Bartender
+\end{verbatim}
+
+\item Create a mountpoint and mount arcfs:
 \begin{verbatim}
@@
    /home/me/arc
-   $ ln -s arc1/src/services/storage/fuse/arcfs.py arcfs.py
    $ mkdir mnt
-   $ python arcfs.py ./mnt 
+   $ arcfs ./mnt 
 \end{verbatim}
@@
 \begin{verbatim}
-   $ python arcfs.py ./mnt
+   $ arcfs ./mnt
 

### 2009-10-30 doc/tech_doc/chelonia/arc-storage-documentation.tex
client.xml removed

@@
 
-The credentials of the user and the URL of the Bartender must be configured either with an XML file located at \verb!~/.arc/client.xml!
-or with environment variables.
-(The environment variables overwrite the values from the XML file.)
-
-An example of user credentials and Bartender configuration in the \verb!~/.arc/client.xml! file:
-
-\begin{verbatim}
-    <ArcConfig>
-      <ProxyPath>/Users/zsombor/Development/arc/proxy</ProxyPath>
-      <CACertificatesDir>/Users/zsombor/Development/arc/certs/CA</CACertificatesDir>
-      <BartenderURL>https://localhost:60000/Bartender</BartenderURL>
-    </ArcConfig>
+The credentials of the user and the URL of the Bartender must be
+configured either with an ini configuration file located at \verb!~/.arc/client.conf! or with environment variables. (The environment variables overwrite the values from the ini file.)
+
+An example of user credentials and Bartender configuration in the \verb!~/.arc/client.ini! file:
+
+\begin{verbatim}
+[common]
+proxypath=/Users/zsombor/Development/arc/proxy
+cacertificatesdirectory=/Users/zsombor/Development/arc/certs/CA
+bartender=https://localhost:60000/Bartender
 \end{verbatim}
@@
 
-As for the \verb!chelonia!
-the credentials of the user and the URL of the Bartender can be configured either with an XML file located at \verb!~/.arc/client.xml!
-or with environment variables.
-See Section \ref{sub:prototype_cli_tool} for more details.
+As for the \verb!chelonia! the credentials of the user and the URL of
+the Bartender can be configured either with an ini configuration file located at \verb!~/.arc/client.conf! or with environment variables. See Section \ref{sub:prototype_cli_tool} for more details.
 
@@
 
-As with the \verb!chelonia!
-command the credentials of the user and the URL of the Bartender can be configured with an XML file located at \verb!~/.arc/client.xml!.
+As with the \verb!chelonia! command the credentials of the user and
+the URL of the Bartender can be configured with an ini configuration file located at \verb!~/.arc/client.conf!.
 When the ARC DMC is initiated, it will create an instance of \verb!DataPointARC!

### 2009-10-30 doc/manuals/user_interface/ui-job.tex
Broker name sync with the current broker loader name
Added brokers description as a separate file, other small fixes
client.xml was renamed to client.confg, correcting references. Please port to final.

@@
    \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
-   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
    \texttt{-v, --version}&&print version information\\
@@
    \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
-   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
    \texttt{-b, --broker}&\textit{string}&select broker method (default is Random)\\
@@
 
-Possible broker values for the option \verb#-b# are:
-\begin{itemize}
- \item[\texttt{Random}] -- ranks targets randomly (default)
- \item[\texttt{FastestQueue}] -- ranks targets according to their queue length
-% \item[\texttt{FastestCPU}] -- ranks targets according to CPU performance
- \item[\texttt{Data}] -- ranks target according the amount of megabytes of the
-requested input files that are already in the computing resource’s cache.
-% \item[\texttt{SampleBroker}] -- Ranks targets randomly (written in Python)
-\end{itemize}
+\input{ui-brokers}
 
@@
    \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
-   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
    \texttt{-v, --version}& & print version information\\

### 2009-10-30 doc/manuals/user_interface/ui-proxy.tex
arcsync, arcsub and chelonia fixed
client.xml was renamed to client.confg, correcting references. Please port to final.

@@
    \texttt{-L, --myproxysrv}& \textit{URL} & URL of MyProxy server\\
-   \texttt{-M, --myproxycmd}& \textit{PUT|GET} & command to MyProxy server. The command can be PUT and GET.\\
+   \texttt{-M, --myproxycmd}& \textit{PUT$|$GET} & command to MyProxy server. The command can be PUT and GET.\\
    & &PUT/put -- put a delegated credential to MyProxy server;\\
@@
    \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
-   \texttt{-c, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-c, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
    \texttt{-v, --version}& & print version information\\

### 2009-10-30 doc/manuals/ws-quick-guide/ws-quick-guide.tex
Keep documentation up-to-date to the GLUE2 schema
client.xml was renamed to client.confg, correcting references. Please port to final.

@@
 user configuration file with predefined values. Default location for this file
-is \texttt{~/.arc/client.xml}. There is an example installed in
-\verb|$ARC_LOCATION/share/doc/arc/client.xml|. Please edit before copying it
+is \texttt{~/.arc/client.conf}. There is an example installed in
+\verb|$ARC_LOCATION/share/arc/examples/client.conf.example|. Please edit before copying it
 into your home directory.
@@
 
-The configuration of the ARC Echo client is specified in an XML
-file. The location of the configuration file is specified by the
-environment variable \verb|$ARC_ECHO_CONFIG|. If there is no such
-environment
-variable, the configuration file is assumed to be \verb|echo_client.xml| in
-the current working directory.
+%%% This does not seem to be correct. The echo client seems to use a normal INI client configuration. Some one please verify this...
+%~ The configuration of the ARC Echo client is specified in an XML
+%~ file. The location of the configuration file is specified by the
+%~ environment variable \verb|$ARC_ECHO_CONFIG|. If there is no such
+%~ environment
+%~ variable, the configuration file is assumed to be \verb|echo_client.conf| in
+%~ the current working directory.
 
@@
  Service Name: MINIMAL Computing Element
- Service Type: org.nordugrid.arex
+ Service Type: org.nordugrid.execution.arex
 

### 2009-10-30 doc/tech_doc/sec/arc-security-documentation.tex
client.xml was renamed to client.confg, correcting references. Please port to final.

@@
 
-  -c, --conffile=filename      configuration file (default ~/.arc/client.xml)
+  -z, --conffile=filename      configuration file (default ~/.arc/client.conf)
 
@@
 
-\textit{./arcslcs -S https://127.0.0.1:60000/slcs -I https://squark.uio.no/idp/shibboleth -U root -P aa1122 -D   ~/arc-0.9/src/clients/credentials -c client.xml}
+\textit{./arcslcs -S https://127.0.0.1:60000/slcs -I https://squark.uio.no/idp/shibboleth -U root -P aa1122 -D   ~/arc-0.9/src/clients/credentials -z client.confg}
 
@@
 \begin{verbatim}
-<ArcConfig> 
-   <!-- change the paths below to the location of your user certs --> 
-   <!-- here only trusted certificates are set. Because this arcslcs is 
-   not supposed to have x509 credential before the samlsso profile has been 
-   (by using client's username/password credential) passed and the SLCS 
-   service succeeded to repond this client with the signed x509 credential. 
-   On the SLCS service's configuration: ClientAuthn should be off.--> 
-   <!--KeyPath>./testkey-nopass.pem</KeyPath--> 
-   <!--CertificatePath>./testcert.pem</CertificatePath--> 
-   <CACertificatesDir>./certificates</CACertificatesDir> 
-   <SLCSURL>https://127.0.0.1:60000/slcs</SLCSURL> 
-   <IdPname>https://squark.uio.no/idp/shibboleth</IdPname> 
-   <Username>root</Username> 
-   <Password>aa1122</Password> 
-   <Keysize>1024</Keysize> 
-   <Keypass>123456</Keypass> 
-   <CertLifetime>24</CertLifetime> 
-   <StoreDir>./</StoreDir> 
-   <Debug>INFO</Debug> 
-</ArcConfig> 
+[common]
+keypath=./testkey-nopass.pem
+certificatepath=./testcert.pem
+cacertificatesdirectory=./certificates
+slcsurl=https://127.0.0.1:60000/slcs
+idpname=https://squark.uio.no/idp/shibboleth
+username=root
+password=aa1122
+keysize=1024
+keypass=123456
+certlifetime=24
+storedir=./
+verbosity=INFO
 \end{verbatim}
@@
 
-\textit{./arcslcs -S https://127.0.0.1:60000/slcs -I https://squark.uio.no/idp/shibboleth -U staff -P 123456 -D /home/wzqiang/arc-0.9/src/clients/credentials -c client.xml}
+\textit{./arcslcs -S https://127.0.0.1:60000/slcs -I https://squark.uio.no/idp/shibboleth -U staff -P 123456 -D /home/wzqiang/arc-0.9/src/clients/credentials -z client.conf}
 

### 2009-10-30 doc/tutorials/ws-programming-tutorial/examples/src/clients/echoclient/echoclient.cpp

@@ -11,96 +11,96 @@ using namespace Arc;
 int main(int argc, char** argv)
 {
 
-	/** Get arguments passed by the command line*/
-	if(argc != 4){//(@*\label{lst_code:echo_client_cpp_arguments}*@)
-		printf("Usage:\n");
-		printf("  echoclient url type message\n");
-		printf("\n");
-		printf("The first argument defines the URL of the service.\n");
-		printf("The second argument the type of echo which shall be performed (ordinary or reverse).\n");
-		printf("The third agrument holds the message to be transmitted.\n");
-		printf("\n");
-		printf("Example:\n");
-		printf("  ./echoclient http://localhost:60000/echo ordinary text_to_be_transmitted\n");
-		return -1;
-	}
-	
-	std::string urlstring(argv[1]);
-	std::string type(argv[2]);
-	std::string message(argv[3]);
-	/** */
+  /** Get arguments passed by the command line*/
+  if(argc != 4){//(@*\label{lst_code:echo_client_cpp_arguments}*@)
+    printf("Usage:\n");
+    printf("  echoclient url type message\n");
+    printf("\n");
+    printf("The first argument defines the URL of the service.\n");
+    printf("The second argument the type of echo which shall be performed (ordinary or reverse).\n");
+    printf("The third agrument holds the message to be transmitted.\n");
+    printf("\n");
+    printf("Example:\n");
+    printf("  ./echoclient http://localhost:60000/echo ordinary text_to_be_transmitted\n");
+    return -1;
+  }
 
-	// Initiate the Logger and set it to the standard error stream
-	Arc::Logger::rootLogger.setThreshold(Arc::WARNING);
-	Arc::Logger logger(Arc::Logger::getRootLogger(), "echo");
-	Arc::LogStream logcerr(std::cerr);
-	Arc::Logger::getRootLogger().addDestination(logcerr);
-	Arc::Logger::rootLogger.setThreshold(Arc::WARNING);
+  std::string urlstring(argv[1]);
+  std::string type(argv[2]);
+  std::string message(argv[3]);
+  /** */
 
-	// Set the ARC installation directory
-	Arc::ArcLocation::Init("/usr/lib/arc");
-	setlocale(LC_ALL, "");   
+  // Initiate the Logger and set it to the standard error stream
+  Arc::Logger::rootLogger.setThreshold(Arc::WARNING);
[…]

### 2009-10-30 src/clients/arclib/arccat.cpp

@@ -83,7 +83,7 @@ int main(int argc, char **argv) {
 
   std::string conffile;
   options.AddOption('z', "conffile",
-                    istring("configuration file (default ~/.arc/client.xml)"),
+                    istring("configuration file (default ~/.arc/client.conf)"),
                     istring("filename"), conffile);
 
   std::string debug;

### 2009-10-30 src/clients/arclib/arcresub.cpp

@@ -97,7 +97,7 @@ int main(int argc, char **argv) {
 
   std::string conffile;
   options.AddOption('z', "conffile",
-                    istring("configuration file (default ~/.arc/client.xml)"),
+                    istring("configuration file (default ~/.arc/client.conf)"),
                     istring("filename"), conffile);
 
   std::string debug;

### 2009-10-30 doc/manuals/user_interface/ui-brokers.tex
Some clean-up in bibliography and layout

@@
 
-Possible broker values for the \texttt{arcsub} command line option \verb#-b# are:
+Possible \index{broker}broker values for the \texttt{arcsub} command line option \verb#-b# are:
 \begin{itemize}
@@
 module that has the proper ARC plugin descriptor for the new broker. For example, to build a broker plugin ``MyBroker'' one executes:
-\begin{lstlisting}
+\begin{lstlisting}[language=sh]
   g++ -I /arc-install/include \
@@
 \end{lstlisting}
-For more details, refer to \textit{libarclib} documentation~\ref{cite:libarcclient}.
+For more details, refer to \textit{libarclib} documentation~\cite{libarcclient}.

### 2009-10-30 doc/manuals/user_interface/ui-job.tex
Some clean-up in bibliography and layout

@@
 
-Since  the  information  about  a job retrieved from a clsuter can be slightly out of date if the user very recently
+Since  the  information  about  a job retrieved from a cluster can be slightly out of date if the user very recently
 submitted or removed a job, a warning is issued when this command is run. The \verb#--force# option disables this warning.
 
-%If a specific cluster is selected for the syncronisation while the job list contains jobs  on  a  different  cluster,
-%running  the arcsync command will remove the old jobs from the joblist, unless the \verb#--merge# option is given, in which
-%case the new found jobs are merged with the jobs already in the joblist.
+If the job list is not empty when invoking syncronisation, the old jobs will be merged with the new jobs, unless the \verb#--truncate# option is given, in which case the job list will first be truncated and then the new jobs will be added.
 
@@
 
-\begin{framed}
-\begin{verbatim}
+\begin{lstlisting}[language=xml]
 <?xml version="1.0" encoding="UTF-8"?>
@@
 </JobDefinition>
-\end{verbatim}
-\end{framed}
+\end{lstlisting}
 

### 2009-10-30 doc/manuals/user_interface/ui.tex
Added Proxy and Configuration sections
CLI manual: moved chelonia section under doc/chelonia, fixed Makefile
Added brokers description as a separate file, other small fixes
Some clean-up in bibliography and layout

@@
 \definecolor{cyan}{rgb}{0,1,1}
+\definecolor{grey}{rgb}{0.9,0.9,0.9}
+
+\usepackage{listings}
+ \lstset{
+         basicstyle=\small\ttfamily,
+         numbers=none,
+         numberstyle=\tiny,
+         stepnumber=1,
+         numbersep=5pt,
+         tabsize=2,
+         extendedchars=true,
+         breaklines=true,
+         keywordstyle=\color{red},
+                frame=b,         
+ %        keywordstyle=[1]\textbf,
+ %        keywordstyle=[2]\textbf,
+ %        keywordstyle=[3]\textbf,
+ %        keywordstyle=[4]\textbf,   \sqrt{\sqrt{}} %
+         stringstyle=\color{blue}\ttfamily,
+         showspaces=false,
+         showtabs=false,
+         xleftmargin=17pt,
+         frame=single,
+         framexleftmargin=17pt,
+         framexrightmargin=5pt,
+         framexbottommargin=4pt,
+         framextopmargin=4pt,
+         backgroundcolor=\color{grey},
+         showstringspaces=false     
+ }
+ \lstloadlanguages{
+         sh,
+         bash,
+         Python,
+         PHP,
+         Perl,
+         C,
+         C++,
+         XML,
+         HTML,
+         Java
+ }
 
@@
 
+\section{Proxy utilities} \label{sec:proxy}
+\index{security} ARC now comes complete with a set of utilities to create
+temporary user credentials (proxies) used to access Grid services.
+
+\input{ui-proxy}
+
 \section{Job submission and management} \label{sec:ui} 
@@
 
+\input{chelonia-cli}
+
 %\section{Test suite} \label{sec:testsuite}
@@
 
-\bibliography{grid}
+\bibliography{grid,nordugrid}
 \printindex

### 2009-11-02 doc/manuals/chelonia/arc-storage-manual.tex
arcfs is not included in final release -> commenting out fuse until solution is found

@@
 
-\section{Using the FUSE module} % (fold)
-\label{sec:using_the_fuse_module}
-
-Besides the \verb!chelonia! tool you can access the storage system with the use of a FUSE module. The following steps installs the ARC storage FUSE module:
-
-\begin{itemize}
-\item Install ARC1 storage system.
-
-\item Install \verb!fuse >= 2.7.3! and \verb!fuse-python >= 0.2!:
-\begin{verbatim}
-$ sudo aptitude install fuse-python
-\end{verbatim}
-
-\item Make sure \verb!bartender! is set in
-  \verb!$HOME/.arc/client.conf!:
-\begin{verbatim}
-$ grep bartender ~/.arc/client.conf 
-bartender= https://chelonia1.knowarc.eu:60001/Bartender
-\end{verbatim}
-
-\item Create a mountpoint and mount arcfs:
-\begin{verbatim}
-   $ pwd
-   /home/me/arc
-   $ mkdir mnt
-   $ arcfs ./mnt 
-\end{verbatim}
-This mounts ARCFS in \verb!/home/me/arc/mnt!, creates directory
-\verb!/home/me/arc/fuse_transfer! and a log file
-\verb!/home/me/arc/arcfsmessages!.
-
-\item Make a collection:
-\begin{verbatim}   
-   $ mkdir -p mnt/home/me
-   $ stat mnt/home/me
-     File: `mnt/home/me'
-     Size: 0               Blocks: 0          IO Block: 4096   directory
-   Device: 19h/25d Inode: 3           Links: 2
-   Access: (0755/drwxr-xr-x)  Uid: (  500/  me)   Gid: (  100/   users)
-   Access: 1970-01-01 01:00:00.000000000 +0100
-   Modify: 1970-01-01 01:00:00.000000000 +0100
-   Change: 1970-01-01 01:00:00.000000000 +0100
-\end{verbatim}
-Note that so far there is no security handling implemented, so
-that all files have mode 0644, all collections have mode 0755,
-everything is owned by the user that mounts the system.
-
-\item Create some entries in your collection:
-\begin{verbatim}
-   $ cd mnt/home/me
-   $ emacs -nw fish
-   catfish
-   $ cat fish
-   catfish
-   $ cp fish fish2
-   $ ls -la
-   drwxr-xr-x 5 me users  0 1970-01-01 01:00 .
-   drwxr-xr-x 3 me users  0 1970-01-01 01:00 ..
-   -rw-r--r-- 1 me users  8 1970-01-01 01:00 fish
-   -rw-r--r-- 1 me users 28 1970-01-01 01:00 fish~
-   -rw-r--r-- 1 me users  8 1970-01-01 01:00 fish2
-\end{verbatim}
-
-\item You probably want to remove that anoying emacs backup file
-\begin{verbatim}   
-   $ rm fish~
-   $ ls 
-   fish  fish2
-\end{verbatim}
-
-\item Maybe you want fish2 in a separate collection
-\begin{verbatim}   
-   $ mkdir sea_creatures
-   $ mv fish2 sea_creatures
-\end{verbatim}
-
-\item Maybe you're not happy with the collection name
-\begin{verbatim}
-   $ mv sea_creatures creatures
-   $ ls creatures
-   fish2
-\end{verbatim}
-
-\item Moving the collection out of ARC storage:
-\begin{verbatim}
-   $ cd ../../..
-   $ mv mnt/home/me/creatures .
-\end{verbatim}
-
-\item The collection is no longer in the ARC storage:
-\begin{verbatim}
-   $ find mnt/home/me/
-   mnt/home/me/
-   mnt/home/me/fish
-\end{verbatim}
-
-\item Unmounting ARCFS; To unmount, refer to fuse documentation. Usually you can do
-\begin{verbatim}   
-   $ fusermount -u mnt
-\end{verbatim}
-
-If this for some reason this doesn't work, you can do
-\begin{verbatim}   
-   $ sudo umount -f mnt
-\end{verbatim}
-
-or even
-\begin{verbatim}   
-   $ sudo umount -l mnt
-\end{verbatim}
-
-Now, remounting the system, everything should still be there:
-\begin{verbatim}
-   $ arcfs ./mnt
-
-   $ find mnt/home/me/
-   mnt/home/me/
-   mnt/home/me/fish
-
-   $ fusermount -u mnt
-\end{verbatim}
-\end{itemize}
+%% \section{Using the FUSE module} % (fold)
+%% \label{sec:using_the_fuse_module}
+
+%% Besides the \verb!chelonia! tool you can access the storage system with the use of a FUSE module. The following steps installs the ARC storage FUSE module:
+
+%% \begin{itemize}
+%% \item Install ARC1 storage system.
+
+%% \item Install \verb!fuse >= 2.7.3! and \verb!fuse-python >= 0.2!:
+%% \begin{verbatim}
+%% $ sudo aptitude install fuse-python
+%% \end{verbatim}
+
+%% \item Make sure \verb!bartender! is set in
+%%   \verb!$HOME/.arc/client.conf!:
+%% \begin{verbatim}
+%% $ grep bartender ~/.arc/client.conf 
+%% bartender= https://chelonia1.knowarc.eu:60001/Bartender
+%% \end{verbatim}
+
+%% \item Create a mountpoint and mount arcfs:
+%% \begin{verbatim}
+%%    $ pwd
+%%    /home/me/arc
+%%    $ mkdir mnt
+%%    $ arcfs ./mnt 
+%% \end{verbatim}
+%% This mounts ARCFS in \verb!/home/me/arc/mnt!, creates directory
+%% \verb!/home/me/arc/fuse_transfer! and a log file
+%% \verb!/home/me/arc/arcfsmessages!.
+
+%% \item Make a collection:
+%% \begin{verbatim}   
+%%    $ mkdir -p mnt/home/me
+%%    $ stat mnt/home/me
+%%      File: `mnt/home/me'
+%%      Size: 0               Blocks: 0          IO Block: 4096   directory
+%%    Device: 19h/25d Inode: 3           Links: 2
+%%    Access: (0755/drwxr-xr-x)  Uid: (  500/  me)   Gid: (  100/   users)
+%%    Access: 1970-01-01 01:00:00.000000000 +0100
+%%    Modify: 1970-01-01 01:00:00.000000000 +0100
+%%    Change: 1970-01-01 01:00:00.000000000 +0100
+%% \end{verbatim}
+%% Note that so far there is no security handling implemented, so
+%% that all files have mode 0644, all collections have mode 0755,
+%% everything is owned by the user that mounts the system.
+
+%% \item Create some entries in your collection:
+%% \begin{verbatim}
+%%    $ cd mnt/home/me
+%%    $ emacs -nw fish
+%%    catfish
+%%    $ cat fish
+%%    catfish
+%%    $ cp fish fish2
+%%    $ ls -la
+%%    drwxr-xr-x 5 me users  0 1970-01-01 01:00 .
+%%    drwxr-xr-x 3 me users  0 1970-01-01 01:00 ..
+%%    -rw-r--r-- 1 me users  8 1970-01-01 01:00 fish
+%%    -rw-r--r-- 1 me users 28 1970-01-01 01:00 fish~
+%%    -rw-r--r-- 1 me users  8 1970-01-01 01:00 fish2
+%% \end{verbatim}
+
+%% \item You probably want to remove that anoying emacs backup file
+%% \begin{verbatim}   
+%%    $ rm fish~
+%%    $ ls 
+%%    fish  fish2
+%% \end{verbatim}
+
+%% \item Maybe you want fish2 in a separate collection
+%% \begin{verbatim}   
+%%    $ mkdir sea_creatures
+%%    $ mv fish2 sea_creatures
+%% \end{verbatim}
+
+%% \item Maybe you're not happy with the collection name
+%% \begin{verbatim}
+%%    $ mv sea_creatures creatures
+%%    $ ls creatures
+%%    fish2
+%% \end{verbatim}
+
+%% \item Moving the collection out of ARC storage:
+%% \begin{verbatim}
+%%    $ cd ../../..
+%%    $ mv mnt/home/me/creatures .
+%% \end{verbatim}
+
+%% \item The collection is no longer in the ARC storage:
+%% \begin{verbatim}
+%%    $ find mnt/home/me/
+%%    mnt/home/me/
+%%    mnt/home/me/fish
+%% \end{verbatim}
+
+%% \item Unmounting ARCFS; To unmount, refer to fuse documentation. Usually you can do
+%% \begin{verbatim}   
+%%    $ fusermount -u mnt
+%% \end{verbatim}
+
+%% If this for some reason this doesn't work, you can do
+%% \begin{verbatim}   
+%%    $ sudo umount -f mnt
+%% \end{verbatim}
+
+%% or even
+%% \begin{verbatim}   
+%%    $ sudo umount -l mnt
+%% \end{verbatim}
+
+%% Now, remounting the system, everything should still be there:
+%% \begin{verbatim}
+%%    $ arcfs ./mnt
+
+%%    $ find mnt/home/me/
+%%    mnt/home/me/
+%%    mnt/home/me/fish
+
+%%    $ fusermount -u mnt
+%% \end{verbatim}
+%% \end{itemize}
 

### 2009-11-04 doc/manuals/user_interface/ui-data.tex
CLI manual: moved chelonia section under doc/chelonia, fixed Makefile
bring up to date arc data client examples

--- 2009-08-10 doc/manuals/user_interface/ui-data.tex
@@
 \subsection{arcls}\label{sec:arcls}
-The \texttt{arcls} \index{arcls}\index{commands:arcls} is a simple
+\texttt{arcls}\index{arcls}\index{commands:arcls} is a simple
 utility that allows to list contents and view some attributes of
-objects of a specified (by an URL) remote directory.
+objects of a specified (by a URL) remote directory.
 
@@
     \texttt{-L} &  & detailed listing including URLs from which file can 
-    be downloaded and\\ & & temporary cached locations\\
+    be downloaded\\
     \texttt{-m} && display all available metadata\\
@@
 \begin{verbatim}
-    arcls rls://rc.host:38203/logical_file_name
+    arcls -L rls://rls.nordugrid.org:38203/logical_file_name
     arcls -l gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535
-    arcls -L srm://grid.uio.no:58000/srm/managerv1/johndoe/log2
+    arcls srm://grid.uio.no:8446/srm/managerv2?SFN=/johndoe/log2
 \end{verbatim}
@@
 
-The \texttt{arccp} \index{arccp}\index{commands:arccp} is a powerful
+\texttt{arccp}\index{arccp}\index{commands:arccp} is a powerful
 tool to copy files over the Grid. It is a part of the A-REX,
@@
     arccp gsiftp://aftpexp.bnl.gov;threads=10/rep/my.file \
-              rc://;threads=4@grid.uio.no/lc=Collection,rc=Catalog/zebra4.f
+              rls://grid.uio.no/zebra4.f
     arccp http://www.nordugrid.org/data/somefile gsiftp://hathi.hep.lu.se/data/
@@
 
-The \texttt{arcrm} \index{arcrm}\index{commands:arcrm}
+The \texttt{arcrm}\index{arcrm}\index{commands:arcrm}
 command allows users to erase files at any location specified by a
@@
    A convenient use for \texttt{arcrm} is to erase the files in a data
-   indexing catalog (RC, RLS or such), as it will not only remove the
+   indexing catalog (LFC, RLS or such), as it will not only remove the
    physical instance, but also will clean up the database record.
@@
 \begin{verbatim}
-    arcrm rc://grid.uio.no/lc=Collection,rc=Catalog/badfile
+    arcrm lfc://grid.uio.no/grid/atlas/AOD_0947.pool.root
 \end{verbatim}
@@
 
-ACL document (an XML file) is printed to standard output when
+The ACL document (an XML file) is printed to standard output when
 \texttt{get} is requested, and is acquired from standard input when
 \texttt{set} is specified\footnote{In ARC $\leq$ 0.5.28, \texttt{set}
-  shoud be used instead of \texttt{put}}. Usage examples are:
+  was used instead of \texttt{put}}. Usage examples are:
 \begin{verbatim}
     arcacl get gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename
-    arcacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename $<$ myacl
+    arcacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename < myacl
 \end{verbatim}
@@
 \subsection{arctransfer}\label{sec:arctransfer}
-\index{arctransfer}\index{commands:arctransfer}The \texttt{arctransfer}
-This command is not implemented.
+\index{arctransfer}\index{commands:arctransfer}
+The \texttt{arctransfer} command is not implemented.

### 2009-11-05 doc/tech_doc/a-rex/arex_tech_doc.tex
Explicitly stating that reference is pointing to a section.
improvements to the cache section after feedback from Martin Skou

@@
 
-The A-REX can cache input files. Caching is enabled if one or more
-cache directories are specified in the configuration file. Caching can
-also be explicitly turned off by the user for each file by using
-the\textit{ cache=no} option in the URL (for URL options read
-{}``Protocols, Uniform Resource Locators (URL) and extensions
-supported in ARC'' \cite{urls}). The disk space occupied by the cache
-is controlled by removing old files.  For more information on
-configuration look in section \ref{SubSection:ConfigFile}.
+The A-REX can cache input files, so that subsequent jobs requiring the
+same files do not have to download them again. Caching is enabled if
+one or more cache directories are specified in the configuration
+file. All input files except files uploaded by the user during job
+submission are cached by default. This includes excutable files
+downloaded by the A-REX. Caching can be explicitly turned off by the
+user in the job description (see~\cite{userguide1}). The disk space
+occupied by the cache is controlled by removing files in the order of
+least recent access. For more information on configuration see
+Section~\ref{SubSection:ConfigFile}.
 
@@
 \item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} - specifies
-a directory to store cached data (see \ref{sec:cache}). Multiple
+a directory to store cached data (see section \ref{sec:cache}). Multiple
 cache directories may be specified by specifying multiple \emph{cachedir}

### 2009-11-05 doc/manuals/chelonia/chelonia-cli.tex
updating the chelonia manual
more documentation about the chelonia cli, not finished yet.

@@
 \index{chelonia}\index{commands:chelonia}\texttt{chelonia} is a client tool for
-accessing the Chelonia storage system. With it you can
+accessing the Chelonia storage system. With it it is possible to
 create, remove and list file collections, upload, download and remove files,
@@
    \texttt{-v} && verbose mode\\
-   \texttt{-z}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
+   \texttt{-z}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
    \texttt{-w} && allow to run without the ARC python client libraries (with limited functionality)\\
@@
    \texttt{policy, pol} &\textit{string}& modify access policy rules. The string has a form $<$LN$>$ $<$changeType$>$ $<$identity$>$ $<$action list$>$.\\
-   & &$<$changeType$>$ could be 'set', 'change' or 'clear'\\
-   & & 'set': sets the action list to the given user overwriting the old one\\
-   & & 'change': modify the current action list with adding and removing actions\\
-   & & 'clear': clear the action list of the given user\\
-   & &$<$identity$>$ could be a '$<$UserDN$>$' or a 'VOMS:$<$VO name$>$'\\
-   & &$<$action list$>$ is a list actions prefixed with '+' or '-', e.g.
-'+read +addEntry -delete'; possible actions are: \texttt{read}, \texttt{addEntry}, \texttt{removeEntry}, \texttt{delete}, \texttt{modifyPolicy}, \texttt{modifyStates}, \texttt{modifyMetadata}\\
+   & &$<$changeType$>$ could be `set', `change' or `clear'\\
+   & & `set': sets the action list to the given user overwriting the old one\\
+   & & `change': modify the current action list with adding and removing actions\\
+   & & `clear': clear the action list of the given user\\
+   & &$<$identity$>$ could be a `$<$UserDN$>$' or a `VOMS:$<$VO name$>$'\\
+   & &$<$action list$>$ is a list actions prefixed with `+' or `-', e.g.
+`+read +addEntry -delete'; possible actions are: \texttt{read}, \texttt{addEntry}, \texttt{removeEntry}, \texttt{delete}, \texttt{modifyPolicy}, \texttt{modifyStates}, \texttt{modifyMetadata}\\
    \texttt{unlink} &\textit{string}& remove a link to an entry from a collection without removing the entry itself\\
@@
 \end{verbatim}
+
+\subsubsection{stat} % (fold)
+\label{ssub:stat}
+With the \texttt{stat} method it is possible to get all the metadata about one or more entry (file, collection, etc.). The entries are specified with their Logical Name (LN).
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia stat $<$LN$>$ [$<$LN$>$ ...]}
+\end{shaded}
+
+The output contains key-value pairs grouped in sections.
+The `states' section contains the size and the checksum of a file, the number of needed replicas, and whether a collection is closed or not; the `entry' section contains the DN of the owner, the globally unique ID (GUID) of the entry, and the type of the entry (file, collection, etc.); the `parents' section contains the GUID of the parent collection(s) of this entry, and the name of this entry in that collection separated with a `/'; the `locations' sections contains the location of the replicas of a file, which contains of the ID (the URL) of the storage element, the ID of the replica within the storage element, and the state of the replica; the `timestamps' section contains the creation time of the entry; the `entries' section contains the name and GUID of the entries of a collection.
+Example stat of a file:
+\begin{verbatim}
+$ chelonia stat /thing
+'/thing': found
+  states
+    checksumType: md5
+    neededReplicas: 3
+    size: 6
+    checksum: a0186a90393bd4a639a1ce35d8ef85f6
+  entry
+    owner: /C=HU/O=NIIF CA/OU=GRID/OU=NIIF/CN=Nagy Zsombor
+    GUID: 398CBDEA-E282-4735-8DF6-2464CD00BE2D
+    type: file
+  parents
+    0/thing: parent
+  locations
+    https://localhost:60000/Shepherd D519F687-EF65-4AEA-9766-E6E2D42166C4: alive
+  timestamps
+    created: 1257351119.3
+\end{verbatim}
+Example stat of a collection:
+\begin{verbatim}
+$ chelonia stat /
+'/': found
+  states
+    closed: no
+  entry
+    owner: /C=HU/O=NIIF CA/OU=GRID/OU=NIIF/CN=Nagy Zsombor
+    GUID: 0
+    type: collection
+  timestamps
+    created: 1257351114.37
+  entries
+    thing: 398CBDEA-E282-4735-8DF6-2464CD00BE2D    
+\end{verbatim}
+% subsubsection stat (end)
+
+\subsubsection{makeCollection} % (fold)
+\label{ssub:makecollection}
+With the \texttt{makeCollection} or \texttt{mkdir} method it is possible to create a new empty collection. The requested Logical Name (LN) should be specified.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia makeCollection $<$LN$>$}
+\end{shaded}
+
+The parent collection of the requested Logical Name must exist.
+
+Example output of the method:
+\begin{verbatim}
+$ chelonia mkdir /newcoll
+Creating collection '/newcoll': done
+
+$ chelonia mkdir /nonexistent/newcoll
+Creating collection '/nonexistent/newcoll': parent does not exist
+\end{verbatim}
+% subsubsection makecollection (end)
+
+\subsubsection{unmakeCollection} % (fold)
+\label{ssub:unmakecollection}
+With the \texttt{unmakeCollection} or \texttt{rmdir} method it is possible to delete an empty collection which is specified by its Logical Name (LN).
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia unmakeCollection $<$LN$>$}
+\end{shaded}
+
+Example output of the method:
+\begin{verbatim}
+$ chelonia rmdir /newcoll
+Removing collection '/newcoll': removed
+
+$ chelonia rmdir /dir
+Removing collection '/dir': collection is not empty
+\end{verbatim}
+% subsubsection unmakecollection (end)
+
+\subsubsection{list} % (fold)
+\label{ssub:list}
+With the \texttt{list} or \texttt{ls} method it is possible to list the contents of one or more collections which are specified by their Logical Name (LN).
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia list $<$LN$>$ [$<$LN$>$ ...]}
+\end{shaded}
+
+Example output of the method:
+\begin{verbatim}
+$ chelonia list / /newcoll
+'/newcoll': collection
+    empty.
+'/': collection
+  thing   <file>
+  dir     <collection>
+  newcoll <collection>
+\end{verbatim}
+% subsubsection list (end)
+
+\subsubsection{move} % (fold)
+\label{ssub:move}
+With the \texttt{move} or \texttt{mv} method it is possible to move a file or collection within the namespace of chelonia (including renaming the entry).
+The source path and the target path should be specified as Logical Names
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia list $<$sourceLN$>$ $<$targetLN$>$}
+\end{shaded}
+
+Example output of the method:
+\begin{verbatim}
+$ chelonia mv /thing /newcoll/
+Moving '/thing' to '/newcoll/': moved
+
+$ chelonia mv /newcoll/thing /newcoll/othername
+Moving '/newcoll/thing' to '/newcoll/othername': moved   
+\end{verbatim}
+% subsubsection move (end)
+
+\subsubsection{putFile} % (fold)
+\label{ssub:putfile}
+With the \texttt{putFile} or \texttt{put} method it is possible to upload a new file into the system creating a new Logical Name (LN). It is possible the specify the number of needed replicas.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia putFile $<$source filename$>$ $<$target LN$>$ [$<$number of replicas needed$>$]}
+\end{shaded}
+
+Example output of the method:
+\begin{verbatim}
+$ chelonia put thing /newcoll/
+'thing' (6 bytes) uploaded as '/newcoll/thing'.    
+\end{verbatim}
+% subsubsection putfile (end)
+
+\subsubsection{getFile} % (fold)
+\label{ssub:getfile}
+With the \texttt{getFile} or \texttt{get} method it is possible to download a file specified with its Logical Name (LN).
+If the target local path is not given, then the file will be put into the local directory.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia getFile $<$source LN$>$ [$<$target filename$>$]}
+\end{shaded}
+
+Example output of the method:
+\begin{verbatim}
+$ chelonia get /newcoll/thing newlocalname
+'/newcoll/thing' (6 bytes) downloaded as 'newlocalname'.
+\end{verbatim}
+% subsubsection getfile (end)
+
+\subsubsection{delFile} % (fold)
+\label{ssub:delfile}
+With the \texttt{delFile} or \texttt{rm} method it is possible to delete one or more files from the system.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia delFile $<$LN$>$ [$<$LN$>$ ...]}
+\end{shaded}
+
+Example output of the method:
+\begin{verbatim}
+$ chelonia rm /newcoll/othername
+/newcoll/othername: deleted    
+\end{verbatim}
+% subsubsection delfile (end)
+
+\subsubsection{modify} % (fold)
+\label{ssub:modify}
+
+With the \texttt{modify} or \texttt{mod} method it is possible to modify some metadata of an entry.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia modify $<$LN$>$ $<$changeType$>$ $<$section$>$ $<$property$>$ $<$value$>$}
+\end{shaded}
+
+The possible values of `changeType' are `set' (sets the property to value within the given section), `unset' (removes the property from the given section - the `value' does not matter) and `add' (sets the property to value within the given section only if it does not exist yet).
+
+To change the number of needed replicas for a file:
+\begin{verbatim}
+chelonia modify <LN> set states neededReplicas <number of needed replicas>    
+\end{verbatim}
+
+To close a collection:
+\begin{verbatim}
+chelonia modify <LN> set states closed yes    
+\end{verbatim}
+
+To change metadata key-value pairs:
+\begin{verbatim}
+chelonia modify <LN> set|unset|add metadata <key> <value>    
+\end{verbatim}
+% subsubsection modify (end)

### 2009-11-05 doc/manuals/user_interface/ui-job.tex
bring up to date arc data client examples
Finished job commands

@@
-\subsection{arcsync}
-\label{sec:arcsync}
-
-It is advised to start every grid session by running \texttt{arcsync},
-especially when changing workstations. The reason is that your job submission
-history is cached on your machine, and if you are using ARC client
-installations on different machines, your local lists of submitted jobs will be different. To synchronise
-these lists with the information in the Information System, use the
-\texttt{arcsync} \index{arcsync}\index{commands:arcsync} command.
-
-\hspace*{0.5cm}
-\begin{shaded}
-   \uicommand{arcsync [options]}
-\end{shaded}
-\versions{ARC 0.9}
-\begin{longtable}{llp{8cm}}
-   Options:&&\\
-   \texttt{-c, --cluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site\\
-   \texttt{-i, --index}&\textit{url}&explicitly select or reject (-) a specific index server\\
-   \texttt{-j, --joblist}&\textit{filename}&file where user's job information will be stored\\
-   \texttt{-f, --force}& &don't ask for confirmation\\
-%   \texttt{-m, --merge}& &merge the found jobs with the jobs already in the joblist\\
-   \texttt{-T, --truncate}& &truncate the job list before synchronising\\
-   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
-   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
-   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
-   \texttt{-v, --version}&&print version information\\
-   \texttt{-h, --help}&&print help page\\
-\end{longtable}
-
-The ARC client keeps a local list of jobs in the user's home
-directory. If this file is lost,
-corrupt, or the user wants to recreate the file on a different
-workstation, the \texttt{arcsync} command will recreate this file from
-the information available in the Information System.
-
-Since  the  information  about  a job retrieved from a cluster can be slightly out of date if the user very recently
-submitted or removed a job, a warning is issued when this command is run. The \verb#--force# option disables this warning.
-
-If the job list is not empty when invoking syncronisation, the old jobs will be merged with the new jobs, unless the \verb#--truncate# option is given, in which case the job list will first be truncated and then the new jobs will be added.
-
 \subsection{arcsub}\label{sec:arcsub}
@@
 %   \texttt{-U, --unknownattr}&&allow unknown attributes in the job description\\
+   \texttt{-b, --broker}&\textit{string}&select broker method (default is Random)\\
+   \texttt{-n, --dolocalsandbox}& &store job descriptions in local sandbox (useful for eventual resubmission/migration)\\
    \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
@@
    \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
-   \texttt{-b, --broker}&\textit{string}&select broker method (default is Random)\\
-   \texttt{-n, --dolocalsandbox}& &store job descriptions in local sandbox (useful for eventual resubmission/migration)\\
    \texttt{-v, --version}&&print version information\\
@@
 \verb# ARC0:ldap://ce.ng.eu:2135/nordugrid-cluster-name=ce.ng.eu,Mds-Vo-name=local,o=grid#
+
+It is strongly recommended to use aliases for these long URLs. Aliases are specified in the configuration file (see Section~\ref{sec:client.conf}).
 \end{framed}
@@
 
-A simple \textit{"Hello World"} job could look like:
+A simple \textit{"Hello World"} job can look like:
 
 \begin{shaded}
- arcsub -c ARC0:ldap://grid.ng.org:2135/nordugrid-cluster-name=grid.ng.org,Mds-Vo-name=local,o=grid -f job.jsdl
+ arcsub -c my-test-site job.jsdl
 \end{shaded}
 
-A user has to have valid credentials (see Section~\ref{sec:proxy}) and be authorised at the specified site. The test file \texttt{job.jsdl} is shown below.
+The \verb#-c# option can be repeated several times, for example:
+\begin{verbatim}
+    arcsub -c alias1 -c alias2 job.xrsl
+\end{verbatim}
+This will submit a job to either \verb#alias1# or \verb#alias2#. To submit a job to any site except
+\verb#badsite#, use \verb#-# sign in front of the name:
+\begin{verbatim}
+    arcsub -c -badsite job.xrsl
+\end{verbatim}
+
+If option \verb#-c# is not given, the \verb#arcsub# command locates the available sites by querying the
+Information System. Default index services for the Information System are specified in the
+configuration template distributed with the middleware, and can be overwritten both in the user's
+configuration (see Section~\ref{sec:client.conf}) and from the command line using option
+\verb#-i#. Different Grids use different notation for such index services.
+
+A user has to have valid credentials (see Section~\ref{sec:proxy}) and be authorised at the specified site. A test file \texttt{job.jsdl} is shown below.
 
@@
 
-The \verb#-c# option can be used to \textbf{force} a job to be
-submitted to a particular site (cluster), or to reject submission to a
-site. The matching is done by string match to the site name (i.e.
-hostname) as defined in the Information System, or to an alias, as
-defined in user configuration file (see
-Section~\ref{sec:client.conf}).  The \verb#-c# option can be repeated
-several times, for example:
+In order to keep track of submitted jobs, ARC client stores information in a
+dedicated file, by default located in \texttt{{\$}HOME/.arc/jobs.xml}. It is
+sometimes convenient to keep separate lists (e.g., for different kinds of jobs),
+to be used later with e.g. \verb#arcstat#. This is achieved with the help of
+\verb#-j# command line option.
+
+The user interface transforms input job description into a format
+that can be understood by the Grid services to which it is being
+submitted. By specifying the \verb#-dumpdescription# option, such transformed
+description is written to stdout instead of being submitted to the remote site.
+
+% Description of brokers in a separate file
+\input{ui-brokers}
+
+If you plan to resubmit or migrate jobs, you will have to use command line
+option \verb#-n#, which will instruct the client to store complete job descriptions
+in a local sandbox, such that a resubmitted/migrated job will be identical to the
+original one.
+
+It often happens that some sites that \verb#arcsub# has to contact
+are slow to answer, or are down altogether. This will not prevent
+you from submitting a job, but will slow down the submission. To
+speed it up, you may want to specify a shorter timeout (default is
+20 seconds) with the \verb#-t# option:
 \begin{verbatim}
-    arcsub -c alias1 -c alias2 myjob.xrsl
-\end{verbatim}
-This will submit a job to either \verb#alias1# or
-\verb#alias2#. To submit a job to any site except
-\verb#badsite#, use \verb#-# sign in front of the name:
-\begin{verbatim}
-    arcsub -c -badsite myjob.xrsl
+    arcsub -t 5 myjob.jsdl
 \end{verbatim}
 
-If option \verb#-c# is not given, the \verb#arcsub# command locates the available sites by querying the
-Information System. Default index services for the Information System are specified in the
-configuration template distributed with the middleware, and can be overwritten both in the user's
-configuration (see Section~\ref{sec:client.conf}) and from the command line using option
-\verb#-i#. Different Grids use different notation for such index services.
+Default value for the timeout can be set in the user's configuration file.
 
@@
 will print out the steps taken by the ARC client to find the
-best cluster satisfying your job requirements. A default value for
-the debug level can be set in the user's configuration file.
+best cluster satisfying your job requirements. Possible diagnostics degrees, in the
+order of increasing verbosity, are: \texttt{FATAL}, \texttt{ERROR}, \texttt{WARNING},
+\texttt{INFO}, \texttt{DEBUG} and \texttt{VERBOSE}. Default is \texttt{WARNING}, and
+it can be set to another value in the user's configuration file.
 
-It often happens that some sites that \verb#arcsub# has to contact
-are slow to answer, or are down altogether. This will not prevent
-you from submitting a job, but will slow down the submission. To
-speed it up, you may want to specify a shorter timeout (default is
-20 seconds) with the \verb#-t# option:
-\begin{verbatim}
-    arcsub -t 5 myjob.xrsl
-\end{verbatim}
+Default configuration file is {\$}HOME/.arc/client.conf. However, a user can chose any
+other pre-defined configuration through option \verb#-z#.
 
-A default value for the timeout can be set in the user's configuration file.
-
-The user interface transforms input job description into a format
-that can be understood by the Grid services to which it is being
-submitted. By specifying the \verb#-dumpdescription# option, such transformed
-description is written to stdout instead of being submitted to the remote site.
-
-\input{ui-brokers}
+Command line option \verb#-v# prints out version of the installed ARC client, and option
+\verb#-h# provides a short help text.
 
@@
    \texttt{-a, --all}& & all jobs\\
-   \texttt{-i, --joblist}& \textit{filename}& file containing a list of jobIDs\\
-   \texttt{-c, --cluster}& & show information about a site (cluster)\\
+   \texttt{-j, --joblist}& \textit{filename}& file containing a list of jobIDs\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site\\
+%   \texttt{-c, --cluster}& & show information about a site (cluster)\\
    \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
-   \texttt{-i, --indexurl}& \textit{url} &URL of an index service\\
-   \texttt{-q, --queues}&&show information about clusters and queues\\
+%   \texttt{-i, --index}& \textit{url} &URL of an index service\\
    \texttt{-l, --long}& & long format (extended information)\\
@@
 
-The arcstat command returns the status of jobs submitted to the Grid. Then \texttt{-c} and \texttt{-i} accept arguments in the GRID:URL notation explained in the description of \texttt{arcsub}.
+The arcstat command returns the status of jobs in the Grid, and is typically issued with a
+job ID (as returned by \verb#arcsub#) as an argument. It is also possible to use job name instead of
+ID, but if several jobs have identical names, information will be collected about all of them. 
+More than one job ID and/or name can be given.
 
-Different sites may report slightly different job states, depending on
-the installed software version.% A summary of essential job
-%states\index{job:states} is:
+If the \verb#-l# option is given, extended information is printed.
 
-%  ---  Does the table below need updating? ---
+Options \verb#-a#, \verb#-c#, \verb#-s# and \verb#-j# do not use job ID or names. By
+specifying the \verb#-a# option, the status of all active jobs will be shown. If the \verb#-j# option
+is used, the list of jobs is read from a file with the specified filename, instead of
+the default one (\texttt{{\$}HOME/.arc/jobs.xml}).
 
-% \begin{longtable}{l|l|p{8cm}}
-% ARC 0.3, ARC 0.4 & ARC 0.5, ARC0.6 & Description\\ \hline
-%                    &\texttt{ACCEPTING} &job has reached the site\\
-% \texttt{ACCEPTED}  &\texttt{ACCEPTED}  &job submitted but not yet processed\\
-% \texttt{PREPARING} &\texttt{PREPARING} &input files are being retreived\\
-%                    &\texttt{PREPARED}  &input files are retreived\\
-% \texttt{SUBMITTING}&\texttt{SUBMITTING}&interaction with LRMS ongoing\\
-% \texttt{INLRMS: Q} &\texttt{INLRMS:Q}  &job is queued by LRMS\\
-% \texttt{INLRMS: R} &\texttt{INLRMS:R}  &job is running\\
-%                    &\texttt{INLRMS:S}  &job is suspended\\
-%                    &\texttt{INLRMS:E}  &job is finishing in LRMS\\
-%                    &\texttt{INLRMS:O}  &job is in any other LRMS state\\
-% \texttt{CANCELING} &\texttt{KILLING}   &job is being cancelled by user request\\
-%                    &\texttt{EXECUTED}  &job is completed in LRMS\\
-% \texttt{FINISHING} &\texttt{FINISHING} &output files are being transferred\\
-% \texttt{FINISHED}  &\texttt{FINISHED}  &job is finished\\
-%                    &\texttt{FAILED}    &job is finished with an error\\
-%                    &\texttt{KILLED}    &job is cancelled by user request\\
-% \texttt{DELETED}   &\texttt{DELETED}   &job is removed due to expiration time\\
-% \end{longtable}
+Option \verb#-c# accepts arguments in the \texttt{GRID:URL} notation,
+as explained in the description of \texttt{arcsub}, or their aliases as 
+specified in the configuration file.
+
+Different sites may report different job states, depending on
+the installed grid middleware version. Typical values can be e.g. 
+``Accepted'', ``Preparing'', ``Running'', ``Finished'' or ``Deleted''.
+Please refer to the respective middleware documentation for job state model
+description.
+
+Command line option \verb#-s# will instruct the client to display information
+of only those jobs which status matches the instruction. This option must be given
+together with either \verb#-a# or \verb#-c# ones, e.g.:
+\begin{verbatim}
+    arcstat -as Finished
+\end{verbatim}
+
+Other command line options are identical to those of \verb#arcsub#.
 
@@
 \index{arccat}\index{commands:arccat} assists here, extracting the
-corresponding information from the execution cluster and pasting it
+corresponding information from the execution cluster and dumping it
 on the user's screen. It works both for running tasks and for the
@@
    \texttt{-a, --all}& & all jobs\\
-   \texttt{-i, --joblist}& \textit{filename} & file containing a list of job IDs\\
-   \texttt{-c, --cluster}& & show information about clusters\\
+   \texttt{-j, --joblist}& \textit{filename} & file containing a list of job IDs\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject (-) a specific site\\
    \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
@@
 
-The \texttt{arccat} command can return the standard output of a job
-(\texttt{-o} option), the standard error (\texttt{-e} option) and the
-errors reported by the Grid Manager (\texttt{-l} option).
+The \texttt{arccat} command returns the standard output of a job
+(\texttt{-o} option), the standard error (\texttt{-e} option) or
+errors reported by either Grid Manager or A-REX (\texttt{-l} option).
 
-\begin{framed}
-
-\end{framed}
+Other command line options have the same meaning as in \verb#arcstat#.
 
@@
 \index{arcget}\index{commands:arcget} command should be used. It
-will download the files specified by the \texttt{outputfiles}
-attribute of job description to the user's computer.
+will transfer the files specified for download in job description
+to the user's computer.
 
@@
    \texttt{-a, --all}& & all jobs\\
-   \texttt{-i, --joblist}& \textit{filename} & file containing a list of jobIDs\\
+   \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
    \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
    \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
-   \texttt{-D, --dir} & \textit{dirname} & download directory (the job directory will be created in this directory)\\
-   \texttt{-k, --keep}& & keep files on gatekeeper (do not clean)\\
+   \texttt{-D, --dir} & \textit{dirname} & download path (the job directory will be created in that location)\\
+   \texttt{-k, --keep}& & keep files in the Grid (do not clean)\\
    \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
@@
 
-Only the results of jobs that have finished can be downloaded. The job
-can be referred to either by the \texttt{jobID} that was returned by
+Only the results of jobs that have finished can be downloaded. Just like in \verb#arstat# 
+and \verb#arccat# cases, the job can be referred to either by the \texttt{jobID} that was returned by
 \texttt{arcsub} at submission time, or by its name, if the job
 description contained a job name attribute.
+
+By default, the job is downloaded into a newly created directory in the current path, with the
+name typically being a large random number. In order to instruct \verb#arcget# to use another
+path, use option \verb#-D# (note the capital ``D''), e.g.
+\begin{verbatim}
+    arcget -D /tmp/myjobs "Test job nr 1"
+\end{verbatim}
+
+\begin{framed}
+ After downloading, your jobs will be erased from the execution site! Use command line option \verb#-k#
+ to keep finished jobs in the Grid.
+\end{framed}
+
+Other command line options are identical to those of e.g. \verb#arcstat#.
+
+\subsection{arcsync}
+\label{sec:arcsync}
+
+It is advised to start every grid session by running \texttt{arcsync},
+especially when changing workstations. The reason is that your job submission
+history is cached on your machine, and if you are using ARC client
+installations on different machines, your local lists of submitted jobs will be different. To synchronise
+these lists with the information in the Information System, use the
+\texttt{arcsync} \index{arcsync}\index{commands:arcsync} command.
+
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{arcsync [options]}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+   Options:&&\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site\\
+   \texttt{-i, --index}&\textit{url}&explicitly select or reject (-) a specific index server\\
+   \texttt{-j, --joblist}&\textit{filename}&file where user's job information will be stored\\
+   \texttt{-f, --force}& &don't ask for confirmation\\
+%   \texttt{-m, --merge}& &merge the found jobs with the jobs already in the joblist\\
+   \texttt{-T, --truncate}& &truncate the job list before synchronising\\
+   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}&&print version information\\
+   \texttt{-h, --help}&&print help page\\
+\end{longtable}
+
+The ARC client keeps a local list of jobs in the user's home
+directory. If this file is lost,
+corrupt, or the user wants to recreate the file on a different
+workstation, the \texttt{arcsync} command will recreate this file from
+the information available in the Information System.
+
+Since  the  information  about  a job retrieved from a cluster can be slightly out of date if the user very recently
+submitted or removed a job, a warning is issued when this command is run. The \verb#-f# option disables this warning.
+
+If the job list is not empty when invoking syncronisation, the old jobs will be merged with the new jobs, unless
+the \verb#-T# option is given (note the capital ``T''), in which case the job list will first be truncated and then the new jobs will be added.
 
@@
 the \texttt{arckill} \index{arckill}\index{commands:arckill} command. A
-job can be killed amost on any stage of processing through the Grid.
+job can be killed almost at any stage of processing through the Grid.
 
@@
    \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
-   \texttt{-c, --cluster}& & show information about clusters\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject (-) a specific site\\
    \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
-   \texttt{-k, --keep}& & keep files on gatekeeper (do not clean)\\
+   \texttt{-k, --keep}& & keep files in the Grid (do not clean)\\
    \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
@@
 
+If a job is killed, its traces are being cleaned from the Grid. If you wish to keep the killed job
+in the system, e.g. for a post-mortem analysis, use the \verb#-k# option.
+
 \begin{framed}
@@
 
-% \subsection{arcresub}
-% \label{sec:arcresub}
-% 
-% Quite often it happens that a user would like to re-submit a job, but
-% has difficulties recovering the original job description xRSL file.
-% This happens when xRSL files are created by scripts on-fly, and
-% matching of xRSL to the job ID is not straightforward. The utility
-% called \texttt{arcresub}\index{arcresub}\index{commands:arcresub} helps
-% in such situations, allowing users to resubmit jobs known only by
-% their job IDs.
-% 
-% \begin{framed}
-%    Only jobs where the \verb#gmlog#\index{gmlog} attribute was
-%    specified in the job description can be resubmitted.
-% \end{framed}
-% 
-% \hspace*{0.5cm}
-% \begin{shaded}
-%    \uicommand{arcresub [options] [job ...]}
-% \end{shaded}
-% \versions{ARC 0.9}
-% \begin{longtable}{llp{8cm}}
-%    Options:&&\\
-%    \texttt{-a, -all}& & all jobs\\
-%    \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
-%    \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
-%    \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
-%    \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
-%    \texttt{-k, -kluster}&\verb#[-]#textem{name}&explicitly select or
-%    reject a specific site (cluster) as re-submission target\\
-%    \texttt{-K, -Klustlist}&\verb#[-]#textem{filename}&list of
-%    clusters to select or reject as re-submission target\\
-%    \texttt{-g, -giisurl}& \textit{url} &URL of a central Information System server\\
-%    \texttt{-G, -giislist}&\textit{filename}&list of GIIS URLs\\
-%    \texttt{-o, -joblist}&\textit{filename}&file where the job IDs will be stored\\
-%    \texttt{~~~~-dryrun}&&add dryrun option to the xRSL\\
-%    \texttt{~~~~-dumpxrsl}&&do not submit -- dump transformed xRSL to stdout\\
-%    \texttt{~~~~-keep}& & keep files on gatekeeper (do not clean)\\
-%    \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
-%    \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
-%    \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
-%    \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
-%    \texttt{-v, -version}& & print version information\\
-%    \texttt{-h, -help}& & print help page\\
-%    Arguments:&&\\
-%    \texttt{job ...} && list of job IDs and/or jobnames\\
-% \end{longtable}
+Command line options have the same meaning as the corresponding ones of \verb#arcstat# and others.
 
@@
 
-If a job fails, or you are not willing to retrieve the results for
-some reasons, a good practice for users is not to wait for the Grid
-Manager to clean up the job leftovers, but to use \texttt{arcclean}
-\index{arcclean}\index{commands:arcclean} to release the disk space
-and to remove the job ID from the list of submitted jobs and from
-the Information System.
+If a job fails or gets killed with \verb#-k# option, or when you are not willing 
+to retrieve the results for some reasons, a good practice for users is not to wait 
+for the system to clean up the job leftovers, but to use 
+\texttt{arcclean}\index{arcclean}\index{commands:arcclean} to release the disk 
+space and to remove the job ID from the list of submitted jobs and from the Information System.
 
@@
 
-Only jobs that have finished can be cleaned.
+Only jobs that have finished or were cancelled can be cleaned.
 
-% \subsection{arcrenew}
-% \label{sec:arcrenew}
-% 
-% Quite often, the user proxy expires while the job is still running (or
-% waiting in a queue). In case such job has to upload output files to a
-% Grid location (Storage Element), it will fail. By using the \texttt{arcrenew}
-% \index{arcrenew}\index{commands:arcrenew} command, users can upload
-% a new proxy to the job. This can be done while a job is still running,
-% thus preventing it from failing, or whithin 24 hours (or whatever is
-% the expiration time set by the site) after the job
-% end. In the latter case, the Grid Manager will attempt to finalize
-% the job by uploading the output files to the desired location.
-% 
-% \hspace*{0.5cm}
-% \begin{shaded}
-%    \uicommand{arcrenew [options] [job ...]}
-% \end{shaded}
-% \versions{ARC 0.9}
-% \begin{longtable}{llp{8cm}}
-%    Options:&&\\
-%    \texttt{-a, -all}& & all jobs\\
-%    \texttt{-i, -joblist}& \textit{filename} & file containing a list of jobIDs\\
-%    \texttt{-c, -cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
-%    \texttt{-C, -clustlist}&\verb#[-]#textem{filename}&list of sites (clusters) to select or reject\\
-%    \texttt{-s, -status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
-%    \texttt{-t, -timeout}& \textit{time} & timeout for queries (default 40 sec)\\
-%    \texttt{-d, -debug}& \textit{debuglevel} &debug level, from -3 (quiet) to 3 (verbose) - default 0\\
-%    \texttt{-x, -anonymous}& & use anonymous bind for queries (default)\\
-%    \texttt{-X, -gsi}& & use GSI-GSSAPI bind for queries\\
-%    \texttt{-v, -version}& & print version information\\
-%    \texttt{-h, -help}& & print help page\\
-%    Arguments:&&\\
-%    \texttt{job ...} && list of job IDs and/or jobnames\\
-% \end{longtable}
-% 
-% Prior to using \texttt{arcrenew}, be sure to actually create the
-% new proxy!
-%
+It happens ever so often that the job is cleaned by the system, or is otherwise unreachable, and yet your 
+local job list file still has it listed. Use \verb#-f# option in this case to forcefully remove such stale job
+information from the local list.
 
-% \subsection{Auxilliary files}\label{sec:joblists}
-% 
-% User Interface keeps local job lists in two files:
-% {\$}\verb#HOME/.ngjobs# and {\$}\verb#HOME/.arc/history#\footnote{In
-%   ARC $\leq$ 0.5.48, {\$}\texttt{HOME/.nghistory}}.
-% 
-% \textbf{{\$}\texttt{HOME/.ngjobs}} is a local list of the user's active
-% jobs. When a job is successfully submitted, it is added to this list,
-% and when it is removed from the remote site, it is removed from this
-% list. This list is used as the list of all active jobs when the user
-% specifies \verb#-a# option to the various ARC user interface commands.
-% For information about how to reconstruct this file in case it is
-% damaged or you relocate to a different workstation, see
-% section~\ref{sec:arcsync} about the \texttt{arcsync} command.
-% 
-% \textbf{{\$}\texttt{HOME/.arc/history}} contains the \texttt{jobID}s of
-% the jobs the user has submitted together with the time of
-% submission. This file is purely informational.
+Other command line options have the same meaning as the corresponding ones of \verb#arcstat# and others.
+
+\subsection{arcrenew}
+\label{sec:arcrenew}
+
+Quite often, the user proxy expires while the job is still running (or
+waiting in a queue). In case such job has to upload output files to a
+Grid location (Storage Element), it will fail. By using the \texttt{arcrenew}
+\index{arcrenew}\index{commands:arcrenew} command, users can upload
+a new proxy to the job. This can be done while a job is still running,
+thus preventing it from failing
+
+If a job has failed in file upload due to expired proxy, \texttt{arcrenew}
+can be issued whithin 24 hours (or whatever is
+the expiration time set by the site) after the job
+end, which must be followed by \texttt{arcresume}. The Grid
+Manager or A-REX will then attempt to finalize
+the job by uploading the output files to the desired location.
+
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{arcrenew [options] [job ...]}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+   Options:&&\\
+   \texttt{-a, --all}& & all jobs\\
+   \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
+   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
+   \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
+   \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
+   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}& & print version information\\
+   \texttt{-h, --help}& & print help page\\
+   Arguments:&&\\
+   \texttt{job ...} && list of job IDs and/or jobnames\\
+\end{longtable}
+
+\begin{framed}
+ Prior to using \texttt{arcrenew}, be sure to actually create the
+new proxy by running \verb#arcproxy#!
+\end{framed}
+
+Command line options have the same meaning as the corresponding ones of \verb#arcstat# and others.
+
+\subsection{arcresume}
+\label{sec:arcresume}
+
+In some cases a user may want to restart a failed job, for example, when input
+files become available, or the storage element for the output files came back
+online, or when a proxy is renewed with \texttt{arcrenew}. This can be done using
+the \texttt{arcresume}\index{arcresume}\index{commands:arcresume} command.
+
+\begin{framed}
+Make sure your proxy is still valid, or when uncertain, run \verb#arcproxy# followed by 
+\verb#arcrenew# before \verb#arcresume#. The job will be resumed from the state where it has failed.
+\end{framed}
+
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{arcresume [options] [job ...]}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+   \texttt{-a, --all}& & all jobs\\
+   \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
+   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
+   \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
+   \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
+   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}& & print version information\\
+   \texttt{-h, --help}& & print help page\\
+   Arguments:&&\\
+   \texttt{job ...} && list of job IDs and/or jobnames\\
+\end{longtable}
+
+Command line options have the same meaning as the corresponding ones of \verb#arcstat# and others.
+
+\subsection{arcresub}
+\label{sec:arcresub}
+
+Quite often it happens that a user would like to re-submit a job, but
+has difficulties recovering the original job description xRSL file.
+This happens when xRSL files are created by scripts on-fly, and
+matching of xRSL to the job ID is not straightforward. The utility
+called \texttt{arcresub}\index{arcresub}\index{commands:arcresub} helps
+in such situations, allowing users to resubmit jobs.
+
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{arcresub [options] [job ...]}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+   Options:&&\\
+   \texttt{-a, --all}& & all jobs\\
+   \texttt{-i, --index}&\textit{url}&explicitly select or reject (-) a specific index server\\
+   \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
+   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific source site\\
+   \texttt{-q, --qluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site as re-submission target\\
+   \texttt{-m, --same}& &re-submit to the same site\\
+   \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
+   \texttt{-x, --dumpdescription}&&do not submit -- dump transformed job description to stdout\\
+   \texttt{-k, --keep}& & keep files in the Grid (do not clean)\\
+   \texttt{-b, --broker}&\textit{string}&select broker method (default is Random)\\
+   \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
+   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}& & print version information\\
+   \texttt{-h, --help}& & print help page\\
+   Arguments:&&\\
+   \texttt{job ...} && list of job IDs and/or jobnames\\
+\end{longtable}
+
+\begin{framed}
+   Only jobs where the \verb#gmlog#\index{gmlog} attribute was
+   specified in the job description can be resubmitted.
+\end{framed}
+
+More  than  one  jobid and/or jobname can be given. If several 
+jobs were submitted with the same jobname all those jobs will be resubmitted.
+
+Upon resubmission of a job the corresponding job description will be 
+fetched from the local job list file. If input files have changed since the
+original job submission, the job no longer remains the same job
+and will therefore not be resubmitted. To make sure the job is always resubmittable,
+submit it with \verb#arcsub -n#.
+
+In case the job description is not found in the joblist, an attempt will be made to
+retrieve it from the cluster holding the orignal job.  This however
+may fail, since both the submission client and
+the cluster can have made modifications to the job description.
+
+Upon resubmision the job will receive a new job ID. The old job ID will be kept
+in the local job list file, enabling future back tracing of the resubmitted job.
+
+Regarding command line options, \verb#arcresub# behaves much like \verb#arcsub#, except that
+\verb#-c# in this case indicates not the submission target site, but on the contrary, the \textbf{site 
+from which the jobs will be resubmitted}. Submission target site is specified with
+option \verb#-q#. If you wish to re-submit each job to the same site, use option \verb#-m#.
+
+If the original job was successfully killed, its traces will be removed from the execution site,
+unless the \verb#-k# option is specified.
+
+\subsection{arcmigrate}
+\label{sec:arcmigrate}
+
+Quite often jobs end up stuck in long queues, and users wish to migrate them to a better
+resource. Command \texttt{arcmigrate}\index{arcmigrate}\index{commands:arcmigrate} is
+triggering this migration. It applies only to jobs submitted to A-REX, as other Grid
+execution services do not support this functionality.
+
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{arcmigrate [options] [job ...]}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+   Options:&&\\
+   \texttt{-a, --all}& & all jobs\\
+   \texttt{-i, --index}&\textit{url}&explicitly select or reject (-) a specific index server\\
+   \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
+   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
+   \texttt{-q, --qluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site as re-submission target\\
+   \texttt{-f, --forcemigration}& & force migration, ignoring kill failure\\
+   \texttt{-b, --broker}&\textit{string}&select broker method (default is Random)\\
+   \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
+   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}& & print version information\\
+   \texttt{-h, --help}& & print help page\\
+   Arguments:&&\\
+   \texttt{job ...} && list of job IDs and/or jobnames\\
+\end{longtable}
+
+\begin{framed}
+Currently only jobs having the A-REX status ``Running'', ``Executing'' or ``Queuing'' can be migrated
+\end{framed}
+
+Command line options \verb#-c# and \verb #-q# are interpreted in the same way as in \verb#arcresub#, namely, 
+\verb#-c# indicates ``from'' and \verb #-q# -- ``to'' which site the job will be migrated.
+
+If the job(s) is successfully migrated, a new job ID(s) is printed out. This jobID uniquely 
+identifies the job while it is being executed.

### 2009-11-09 doc/tech_doc/a-rex/arex_tech_doc.tex
more clarifications in cache doc

@@
 it cannot be deleted and so that another download process cannot write
-the same file simultaneously. This done by creating a file with the
+the same file simultaneously. This is done by creating a file with the
 same name as the cache filename but with a \emph{.lock} suffix. This
@@
 by the local user running the job, and the cache directory is readable
-only by the A-REX user (usually root). This means that the local user
-cannot access any other users' cache files. It also means that cache
-files can be removed without needing to know whether they are in use
-by a currently running job.
+only by the A-REX user. This means that the local user cannot access
+any other users' cache files. It also means that cache files can be
+removed without needing to know whether they are in use by a currently
+running job. Note that when running A-REX under a non-privileged user
+account, all cache files will be owned and accessible by the same
+user, and therefore modifyable by running jobs. This is potentially
+dangerous and so cacheing should be used with caution in this case.
 
@@
 are copied to the session directory.
-\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} - specifies
-high and low watermarks for space used by each cache, as a percentage
-of the space on the file system on which the cache directory is located.
-When the max is exceeded, files will be deleted to bring the used
-space down to the min level. It is a good idea to have each cache
-on its own separate file system. To turn off this feature, \char`\"{}cachesize\char`\"{}
-without parameters can be specified. These cache settings apply to
-all caches specified by \emph{cachedir} commands.
+\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} -
+  specifies high and low watermarks for space used on the file system
+  on which the cache directory is located, as a percentage of total
+  file system capacity/ When the max is exceeded, files will be
+  deleted to bring the used space down to the min level. It is a good
+  idea to have each cache on its own separate file system. To turn off
+  this feature, \char`\"{}cachesize\char`\"{} without parameters can
+  be specified. These cache settings apply to all caches specified by
+  \emph{cachedir} commands.
 \item \textbf{\textit{maxrerun}}\textit{=number} -- specifies maximal number

### 2009-11-09 doc/tech_doc/a-rex/arex_tech_doc.tex
typo

@@
   on which the cache directory is located, as a percentage of total
-  file system capacity/ When the max is exceeded, files will be
+  file system capacity. When the max is exceeded, files will be
   deleted to bring the used space down to the min level. It is a good

### 2009-11-10 doc/tech_doc/infosys/infosys_technical.tex
Removing custom logger from ISIS documentation and fixing the line wraping at the scema file.

@@
 
-\textit{This section extensively uses terms defined in "Security Framework of ARC1" document"~\cite{arcsec}.}
+\textit{This section extensively uses terms defined in "Security Framework of WS-ARC" document"~\cite{arcsec}.}
 
@@
 
-\textit{For information about sophisticated authorization policies and how to deploy various Policy Decision Point entities please see "Security Framework of ARC1" document"~\cite{arcsec}.}
+\textit{For information about sophisticated authorization policies and how to deploy various Policy Decision Point entities please see "Security Framework of WS-ARC" document"~\cite{arcsec}.}
 
@@
 
-    <xsd:import namespace="http://www.w3.org/2005/08/addressing" schemaLocation="http://www.w3.org/2006/03/addressing/ws-addr.xsd"/>
-
-<!-- This is an initial and incomplete DRAFT which mainly concentrates on the structure but not on the actual names. Final version will use G
-LUE-2.0 ternminology. -->
+    <xsd:import namespace="http://www.w3.org/2005/08/addressing" 
+    schemaLocation="http://www.w3.org/2006/03/addressing/ws-addr.xsd"/>
+
+<!-- This is an initial and incomplete DRAFT which mainly concentrates on the structure
+ but not on the actual names. Final version will use GLUE-2.0 ternminology. -->
 
@@
             <xsd:element name="Header" type="isis:HeaderType" minOccurs="0" maxOccurs="1"/>
-            <xsd:element name="RegEntry" type="isis:RegistrationEntryType" minOccurs="1" maxOccurs="unbounded"/>
+            <xsd:element name="RegEntry" type="isis:RegistrationEntryType" minOccurs="1"
+            maxOccurs="unbounded"/>
         </xsd:sequence>
@@
         <xsd:sequence>
-            <xsd:element name="MessageGenerationTime" type="xsd:string" minOccurs="1" maxOccurs="1"/>
-            <xsd:element name="ServiceID" type="xsd:string" minOccurs="1" maxOccurs="unbounded"/>
+            <xsd:element name="MessageGenerationTime" type="xsd:string" minOccurs="1" 
+            maxOccurs="1"/>
+            <xsd:element name="ServiceID" type="xsd:string" minOccurs="1" 
+            maxOccurs="unbounded"/>
         </xsd:sequence>
@@
         <xsd:sequence>
-            <xsd:element name="RemoveRegistrationResponseElement" type="isis:RemoveRegistrationResponseElementType" minOccurs="0" maxOccurs="
+            <xsd:element name="RemoveRegistrationResponseElement" 
+            type="isis:RemoveRegistrationResponseElementType" minOccurs="0" maxOccurs="
 unbounded" />
@@
     </xsd:complexType>
-    <xsd:element name="RemoveRegistrationsResponse" type="isis:RemoveRegistrationsResponseType"/>
+    <xsd:element name="RemoveRegistrationsResponse" 
+    type="isis:RemoveRegistrationsResponseType"/>
 
@@
         <xsd:sequence>
-            <xsd:element name="EPR" type="wsa:EndpointReferenceType" minOccurs="1" maxOccurs="unbounded"/>
+            <xsd:element name="EPR" type="wsa:EndpointReferenceType" minOccurs="1" 
+            maxOccurs="unbounded"/>
         </xsd:sequence>
@@
         <xsd:sequence>
-            <xsd:element name="SrcAdv" type="isis:ServiceAdvertisementType" minOccurs="1" maxOccurs="1"/>
-            <xsd:element name="MetaSrcAdv" type="isis:ServiceAdvertisementMetadataType" minOccurs="1" maxOccurs="1"/>
+            <xsd:element name="SrcAdv" type="isis:ServiceAdvertisementType" minOccurs="1" 
+            maxOccurs="1"/>
+            <xsd:element name="MetaSrcAdv" type="isis:ServiceAdvertisementMetadataType" 
+            minOccurs="1" maxOccurs="1"/>
         </xsd:sequence>
@@
             <!-- Service specific part of the Service Advertisment -->
-            <xsd:element name="SSPair" type="isis:NameValuePairType" minOccurs="0" maxOccurs="unbounded"/>
+            <xsd:element name="SSPair" type="isis:NameValuePairType" minOccurs="0" 
+            maxOccurs="unbounded"/>
         </xsd:sequence>
@@
 
-    <wsdl:import namespace="http://docs.oasis-open.org/wsrf/rpw-2" location="http://docs.oasis-open.org/wsrf/rpw-2.wsdl"/>
+    <wsdl:import namespace="http://docs.oasis-open.org/wsrf/rpw-2" 
+    location="http://docs.oasis-open.org/wsrf/rpw-2.wsdl"/>
 
@@
 <wsdl:message name="RemoveRegistrationsResponse">
-    <wsdl:part name="RemoveRegistrationsResponse" element="isis:RemoveRegistrationsResponse"/>
+    <wsdl:part name="RemoveRegistrationsResponse" 
+    element="isis:RemoveRegistrationsResponse"/>
 </wsdl:message>
@@
     <wsdl:operation name="RemoveRegistrations">
-        <wsdl:input name="RemoveRegistrationsRequest" message="isis:RemoveRegistrationsRequest"/>
-        <wsdl:output name="RemoveRegistrationsResponse" message="isis:RemoveRegistrationsResponse"/>
+        <wsdl:input name="RemoveRegistrationsRequest" 
+        message="isis:RemoveRegistrationsRequest"/>
+        <wsdl:output name="RemoveRegistrationsResponse" 
+        message="isis:RemoveRegistrationsResponse"/>
     </wsdl:operation>
@@
 </wsdl:binding>
-<wsdl:binding name="GetResourcePropertyDocument" type="wsrf-rpw:GetResourcePropertyDocument">
+<wsdl:binding name="GetResourcePropertyDocument" 
+type="wsrf-rpw:GetResourcePropertyDocument">
     <soap:binding style="document" transport="http://schemas.xmlsoap.org/soap/http"/>
@@
 <wsdl:service name="isis">
-    <wsdl:port name="GetResourcePropertyDocument" binding="isis:GetResourcePropertyDocument">
+    <wsdl:port name="GetResourcePropertyDocument" 
+    binding="isis:GetResourcePropertyDocument">
         <soap:address location="mailto:test@test.com"/>
@@
           <xsd:documentation xml:lang="en">
-              Retry specifies how many times communication to ISIS have to fail/timeout to start treating it as unavailable.
+              Retry specifies how many times communication to ISIS have to 
+              fail/timeout to start treating it as unavailable.
           </xsd:documentation>
@@
           <xsd:documentation xml:lang="en">
-              It determines the number of neighbors as a function of the actual number of member nodes of the network.
-          </xsd:documentation>
-      </xsd:annotation>
-  </xsd:element>
-
-  <xsd:complexType name="Logger_Type">
-    <xsd:simpleContent>
-      <xsd:extension base="xsd:string">
-        <xsd:attribute name="level" default="INFO">
-          <xsd:simpleType>
-            <xsd:restriction base="xsd:string">
-              <xsd:enumeration value="VERBOSE"/>
-              <xsd:enumeration value="DEBUG"/>
-              <xsd:enumeration value="INFO"/>
-              <xsd:enumeration value="WARNING"/>
-              <xsd:enumeration value="ERROR"/>
-              <xsd:enumeration value="FATAL"/>
-            </xsd:restriction>
-          </xsd:simpleType>
-        </xsd:attribute>
-      </xsd:extension>
-    </xsd:simpleContent>
-  </xsd:complexType>
-
-  <xsd:element name="Logger" type="icfg:Logger_Type">
-      <xsd:annotation>
-          <xsd:documentation xml:lang="en">
-              File location where the log message is locally stored.
+              It determines the number of neighbors as a function of the actual 
+              number of member nodes of the network.
           </xsd:documentation>
@@
 
-            <xsd:element name="URL" type="xsd:string" minOccurs="1" maxOccurs="unbounded">.
+            <xsd:element name="URL" type="xsd:string" minOccurs="1" 
+            maxOccurs="unbounded">.
                 <xsd:annotation>
@@
 
-            <xsd:element name="ServiceID" type="xsd:string" minOccurs="0" maxOccurs="1">
+            <xsd:element name="ServiceID" type="xsd:string" minOccurs="0" 
+            maxOccurs="1">
                 <xsd:annotation>
@@
 
-            <xsd:element name="KeyPath" type="xsd:string"  minOccurs="0" maxOccurs="1">
+            <xsd:element name="KeyPath" type="xsd:string"  minOccurs="0" 
+            maxOccurs="1">
                 <xsd:annotation>
@@
 
-            <xsd:element name="CertificatePath" type="xsd:string"  minOccurs="0" maxOccurs="1">
+            <xsd:element name="CertificatePath" type="xsd:string"  minOccurs="0" 
+            maxOccurs="1">
                 <xsd:annotation>
@@
             </xsd:element>
-             <xsd:element name="ProxyPath" type="xsd:string"  minOccurs="0" maxOccurs="1">
+             <xsd:element name="ProxyPath" type="xsd:string"  minOccurs="0" 
+             maxOccurs="1">
                  <xsd:annotation>
@@
 
-            <xsd:element name="CACertificatesDir" type="xsd:string"  minOccurs="0" maxOccurs="1">
+            <xsd:element name="CACertificatesDir" type="xsd:string"  minOccurs="0" 
+            maxOccurs="1">
                 <xsd:annotation>
@@
 
-            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1" default="5">
-                <xsd:annotation>
-                    <xsd:documentation xml:lang="en">
-                     Retry count. Specified how many times have an ISIS to timeout to handle it as.
-                     unavailable. If there is no value specified the default value: 5 will be used..
+            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1" 
+            default="5">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     Retry count. Specified how many times have an ISIS to timeout to
+                     handle it as unavailable. If there is no value specified the default
+                     value: 5 will be used.
                     </xsd:documentation>
@@
 
-            <xsd:element name="Period" type="xsd:duration" minOccurs="1" maxOccurs="1" default="PT3M">
+            <xsd:element name="Period" type="xsd:duration" minOccurs="1" maxOccurs="1" 
+            default="PT3M">
                 <xsd:annotation>
@@
 
-            <xsd:element name="Expiration" type="xsd:string" minOccurs="1" maxOccurs="1" default="PT10M">
-                <xsd:annotation>
-                    <xsd:documentation xml:lang="en">
-                     This element defines the expiration time of the information provided by the service.
-                    </xsd:documentation>
-                </xsd:annotation>
-            </xsd:element>
-            <xsd:element name="Registrar" type="iregc:Registrar_Type" minOccurs="0" maxOccurs="unbounded"/>
+            <xsd:element name="Expiration" type="xsd:string" minOccurs="1" maxOccurs="1" 
+            default="PT10M">
+                <xsd:annotation>
+                    <xsd:documentation xml:lang="en">
+                     This element defines the expiration time of the information
+                     provided by the service.
+                    </xsd:documentation>
+                </xsd:annotation>
+            </xsd:element>
+            <xsd:element name="Registrar" type="iregc:Registrar_Type" minOccurs="0" 
+            maxOccurs="unbounded"/>
             <!-- Default values for every Registrar -->
@@
              http://www.nordugrid.org/schemas/policy-arc namespace. -->
-        <xsd:any namespace="##other" processContents="lax" minOccurs="1" maxOccurs="1"/>
+        <xsd:any namespace="##other" processContents="lax" minOccurs="1" 
+        maxOccurs="1"/>
         <xsd:attribute name="id" type="xsd:string" use="required"/>
     </xsd:complexType>
-    <xsd:element name="InfoFilterDefinition" type="if:InfoFilterDefinition_Type" minOccurs="0" maxOccurs="unbounded"/>
+    <xsd:element name="InfoFilterDefinition" type="if:InfoFilterDefinition_Type" 
+    minOccurs="0" maxOccurs="unbounded"/>
 
@@
                 <!-- Contact and behavior details of the Registrar element-->
-                <infosys:URL><!-- Available endpoint URL of an ISIS representing the cloud --></infosys:URL>
+                <infosys:URL>
+                    <!-- Available endpoint URL of an ISIS representing the cloud -->
+                </infosys:URL>
                 <infosys:Retry>4</infosys:Retry>        

### 2009-11-10 doc/manuals/chelonia/chelonia-cli.tex
documenting the remaining methods of the chelonia cli

@@
 \begin{shaded}
-   \uicommand{chelonia list $<$sourceLN$>$ $<$targetLN$>$}
+   \uicommand{chelonia move $<$sourceLN$>$ $<$targetLN$>$}
 \end{shaded}
@@
 % subsubsection modify (end)
+
+\subsubsection{policy} % (fold)
+\label{ssub:policy}
+
+With the \texttt{policy} or \texttt{pol} method it is possible to modify the policy of the entry
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia policy $<$LN$>$ $<$changeType$>$ $<$identity$>$ $<$action list$>$}   
+\end{shaded}
+
+The possible values of `changeType' are `set' (sets the action list to the given user overwriting the old one), `change' (modify the current action list with adding and removing actions) and `clear' (clear the action list of the given user).
+
+The `identity' could be currently three things: the DN of a user; the name of a VO (with the syntax: `VOMS:$<$VO name$>$'); or `ALL' for all users.
+
+The `action list' is a list of actions prefixed with `+' or `-', e.g. `+read +addEntry -delete'.
+
+These are the actions which can be used for access control:
+\begin{itemize}
+    \item \emph{read}: user can get the list of entries in the collection; user can download the file
+    \item \emph{addEntry}: user can add a new entry to the collection;
+    \item \emph{removeEntry}: user can remove any entry from the collection 
+    \item \emph{delete}: user can delete the collection if it is empty; user can delete a file
+    \item \emph{modifyPolicy}: user can modify the policy of the file/collection
+    \item \emph{modifyStates}: user can modify some special metadata of the file/collection (close the collection, change the number of needed replica of the file)
+    \item \emph{modifyMetadata}: user can modify the arbitrary metadata section of the file/collection (these are property-value pairs)
+\end{itemize}
+
+There is an implicit default policy: the owner always has all the rights.
+Checking the `stat' of new collections:
+\begin{verbatim}
+$ chelonia stat /newcoll
+'/newcoll': found
+  states
+    closed: no
+  entry
+    owner: /C=HU/O=NIIF CA/OU=GRID/OU=NIIF/CN=Nagy Zsombor
+    GUID: 41CBD461-09BE-46FD-8A1B-767C7D427AF9
+    type: collection
+  parents
+    0/newcoll: parent
+  timestamps
+    created: 1257435820.26
+  entries
+    thing: A63658B4-2C6E-46A3-8238-7D291F8F81C2
+\end{verbatim}
+shows no policies, but it shows the owner. This collection has no additional policies just the default one: the owner can do anything, noone else can do anything.
+
+Let's set it in a way that all users can read the contents of this collection:
+\begin{verbatim}
+$ chelonia policy /newcoll change ALL +read
+Setting action list of '/newcoll' for user ALL to +read: set.
+$ chelonia stat /newcoll
+'/newcoll': found
+  [...]
+  policy
+    ALL: +read
+  [...]
+\end{verbatim}
+
+Then we can set that all the members of the knowarc VO would be able to add entries to this collection:
+\begin{verbatim}
+$ chelonia policy /newcoll change VOMS:knowarc +addEntry
+Setting action list of '/newcoll' for user VOMS:knowarc to +addEntry: set.
+$ chelonia stat /newcoll
+'/newcoll': found
+  [...]
+  policy
+    ALL: +read
+    VOMS:knowarc: +addEntry
+  [...]
+\end{verbatim}
+
+And for example we can set a specific user to be able to remove entries from this collections:
+\begin{verbatim}
+$ chelonia policy /newcoll change \
+    "/C=HU/O=NIIF CA/OU=GRID/OU=NIIF/CN=TestUser" +removeEntry
+Setting action list of '/newcoll'
+    for user /C=HU/O=NIIF CA/OU=GRID/OU=NIIF/CN=TestUser to +removeEntry: set.
+$ chelonia stat /newcoll'/newcoll': found
+  [...]
+  policy
+    /C=HU/O=NIIF CA/OU=GRID/OU=NIIF/CN=TestUser: +removeEntry
+    ALL: +read
+    VOMS:knowarc: +addEntry
+  [...]
+\end{verbatim}
+
+% subsubsection policy (end)
+
+\subsubsection{unlink} % (fold)
+\label{ssub:unlink}
+
+With the \texttt{unlink} method it is possible to remove a file or collection just from its parent collection without removing the file or collection itself.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia unlink $<$LN$>$}   
+\end{shaded}
+
+If there is a file called `/newcoll/thing', it is in the listing of the `/newcoll' collection:
+\begin{verbatim}
+$ chelonia list /newcoll
+'/newcoll': collection
+  thing <file>
+\end{verbatim}    
+The file is in the entries of the collection:
+\begin{verbatim}
+$ chelonia stat /newcoll
+'/newcoll': found
+  entries
+    thing: A63658B4-2C6E-46A3-8238-7D291F8F81C2
+  [...]
+\end{verbatim}
+
+It is possible the `stat' the file with the Logical Name `/newcoll/thing':
+\begin{verbatim}
+jim:~ zsombor$ chelonia stat /newcoll/thing
+'/newcoll/thing': found
+  states
+    checksumType: md5
+    neededReplicas: 3
+    size: 6
+    checksum: a0186a90393bd4a639a1ce35d8ef85f6
+  [...]
+\end{verbatim}
+
+Now with the `unlink' method it is possible to remove the file from the `/newcoll' collection, but not from the system:
+\begin{verbatim}
+$ chelonia unlink /newcoll/thing
+Unlinking '/newcoll/thing': unset
+\end{verbatim}
+
+Now the file is not in the collection anymore:
+\begin{verbatim}
+$ chelonia list /newcoll
+'/newcoll': collection
+    empty.
+$ chelonia stat /newcoll/thing
+'/newcoll/thing': not found
+\end{verbatim}
+
+But with the GUID of the file, it can still be accessed:
+\begin{verbatim}
+$ chelonia stat A63658B4-2C6E-46A3-8238-7D291F8F81C2
+'A63658B4-2C6E-46A3-8238-7D291F8F81C2': found
+  states
+    checksumType: md5
+    neededReplicas: 3
+    size: 6
+    checksum: a0186a90393bd4a639a1ce35d8ef85f6
+  [...]
+\end{verbatim}
+% subsubsection unlink (end)
+
+\subsubsection{credentialDelegation} % (fold)
+\label{ssub:credentialdelegation}
+With the \texttt{credentialDelegation} or \texttt{cre} method it is possible to delegate credentials to the Bartender.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia credentialDelegation}   
+\end{shaded}
+% subsubsection credentialdelegation (end)
+
+\subsubsection{removeCredentials} % (fold)
+\label{ssub:removecredentials}
+With the \texttt{removeCredentials} or \texttt{rem} method it is possible to remove the previously delegated credentials.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia removeCredentials}   
+\end{shaded}
+% subsubsection removecredentials (end)
+
+\subsubsection{makeMountPoint} % (fold)
+\label{ssub:makemountpoint}
+With the \texttt{makeMountPoint} or \texttt{makemount} method it is possible to create a mount point within the namespace of Chelonia which points to a GridFTP server.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia makeMountPoint $<$LN$>$ $<$URL$>$}   
+\end{shaded}
+
+The `LN' is the requested Logical Name for the mount point, the `URL' points to the GridFTP server.
+% subsubsection makemountpoint (end)
+
+\subsubsection{unmakeMountPoint} % (fold)
+\label{ssub:unmakemountpoint}
+With the \texttt{unmakeMountPoint} or \texttt{unmount} method it is possible to remove a previously created mount point.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{chelonia unmakeMountPoint $<$LN$>$}   
+\end{shaded}
+
+The `LN' is the Logical Name of the mount point.
+% subsubsection unmakemountpoint (end)

### 2009-11-10 doc/tech_doc/infosys/infosys_technical.tex
Replacing 'Security Framework of ARC1' with 'ARC Security Framework (NORDUGRID-TECH-16)'

@@
 
-\textit{This section extensively uses terms defined in "Security Framework of WS-ARC" document"~\cite{arcsec}.}
+\textit{This section extensively uses terms defined in "ARC Security Framework (NORDUGRID-TECH-16)" document"~\cite{arcsec}.}
 
@@
 
-\textit{For information about sophisticated authorization policies and how to deploy various Policy Decision Point entities please see "Security Framework of WS-ARC" document"~\cite{arcsec}.}
+\textit{For information about sophisticated authorization policies and how to deploy various Policy Decision Point entities please see "ARC Security Framework (NORDUGRID-TECH-16)" document"~\cite{arcsec}.}
 

### 2009-11-11 doc/manuals/janitor/tex_installation/installation.tex
Interchange DEBUG and VERBOSE loglevels.

@@
 log4perl.appender.MainLog = Log::Log4perl::Appender::File
-log4perl.appender.MainLog.Threshold = VERBOSE
+log4perl.appender.MainLog.Threshold = DEBUG
 log4perl.appender.MainLog.filename = /var/log/janitor.log

### 2009-11-11 doc/manuals/user_interface/ui-data.tex
Interchange DEBUG and VERBOSE loglevels.

@@
     \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
     \texttt{-l} &  & detailed listing\\
@@
     \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
     \texttt{-y} & \textit{cache\_path} & path to local cache (use to put file into cache)\\
@@
     \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
     \texttt{-c} & &continue with meta-data even if it failed to delete real file\\
@@
    Options:&&\\
-    \texttt{-d, -debug} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+    \texttt{-d, -debug} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
     \texttt{-v} && print version information\\

### 2009-11-11 doc/manuals/user_interface/ui-job.tex
Interchange DEBUG and VERBOSE loglevels.

@@
    \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
-   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE - default WARNING\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG - default WARNING\\
    \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
@@
 \begin{verbatim}
-    arcsub -d DEBUG myjob.xrsl
+    arcsub -d VERBOSE myjob.xrsl
 \end{verbatim}
@@
 order of increasing verbosity, are: \texttt{FATAL}, \texttt{ERROR}, \texttt{WARNING},
-\texttt{INFO}, \texttt{DEBUG} and \texttt{VERBOSE}. Default is \texttt{WARNING}, and
+\texttt{INFO}, \texttt{VERBOSE} and \texttt{DEBUG}. Default is \texttt{WARNING}, and
 it can be set to another value in the user's configuration file.
@@
    \texttt{-t, --timeout}& \textit{time}& timeout for queries (default 20 sec)\\
-   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
    \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
@@
    \texttt{-t, --timeout}& \textit{time} & timeout for queries (default 20 sec)\\
-   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
    \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\

### 2009-11-11 doc/tech_doc/a-rex/arex_tech_doc.tex
Interchange DEBUG and VERBOSE loglevels.

@@
     <Logger>
-      <Level>DEBUG</Level>
+      <Level>VERBOSE</Level>
       <File>/var/log/arched.log</File>

### 2009-11-11 doc/tutorials/ws-programming-tutorial/examples/src/services/secechoservice/secechoservice.cpp

@@ -85,18 +85,18 @@ namespace ArcService
 	*/
 	Arc::MCC_Status SecEchoService::process(Arc::Message& inmsg, Arc::Message& outmsg) 
 	{
-		logger.msg(Arc::DEBUG, "Secure Echo Web Service has been started...");
+		logger.msg(Arc::VERBOSE, "Secure Echo Web Service has been started...");
 
 
 
 		/** Processes the SecHandlers of the service *///(@*\label{lst_code:sececho_service_cpp_process_secH}*@)
-		logger.msg(Arc::DEBUG, "Secure Echo Web Service: Current Client has the DN: \"%s\"",inmsg.Attributes()->get("TLS:PEERDN"));
+		logger.msg(Arc::VERBOSE, "Secure Echo Web Service: Current Client has the DN: \"%s\"",inmsg.Attributes()->get("TLS:PEERDN"));
 
 		if(!ProcessSecHandlers(inmsg,"incoming")) {
-			logger.msg(Arc::DEBUG, "Secure Echo Web Service: Didn't passes SecHandler!");
+			logger.msg(Arc::VERBOSE, "Secure Echo Web Service: Didn't passes SecHandler!");
 			return makeFault(outmsg, "Access to this service is not permitted!");
 		}else{
-			logger.msg(Arc::DEBUG, "Secure Echo Web Service: Passed SecHandler!");
+			logger.msg(Arc::VERBOSE, "Secure Echo Web Service: Passed SecHandler!");
 		}
 		/** */
 
@@ -124,7 +124,7 @@ namespace ArcService
 		std::string operation = (std::string) sayNode.Attribute("operation");
 		std::string say       = (std::string) sayNode;
 		std::string hear      = "";
-		logger.msg(Arc::DEBUG, "Say: \"%s\"  Operation: \"%s\"",say,operation);
+		logger.msg(Arc::VERBOSE, "Say: \"%s\"  Operation: \"%s\"",say,operation);
 
 		//Compare strings with constants defined in the header file
 		if(operation.compare(ECHO_TYPE_ORDINARY) == 0)
@@ -155,10 +155,10 @@ namespace ArcService
 		{
 			std::string str;
 			outpayload->GetDoc(str, true);
-			logger.msg(Arc::DEBUG, "process: response=%s",str);
+			logger.msg(Arc::VERBOSE, "process: response=%s",str);
 		}; 
 
[…]

### 2009-11-11 src/clients/credentials/arcslcs.cpp

@@ -191,7 +191,7 @@ int main(int argc, char *argv[]) {
 
   std::string debug;
   options.AddOption('d', "debug",
-                    istring("FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE"),
+                    istring("FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG"),
                     istring("debuglevel"), debug);
 
   bool version = false;

### 2009-11-11 src/clients/jura/LutsDestination.cpp

@@ -120,7 +120,7 @@ namespace Arc
     component.NewChild("next").NewAttribute("id")="http";
 
     clientchain.GetDoc(dump,true);
-    logger.msg(Arc::DEBUG, "Client chain configuration: %s",
+    logger.msg(Arc::VERBOSE, "Client chain configuration: %s",
                dump.c_str() );
 
     //Get Batch Size:
@@ -178,7 +178,7 @@ namespace Arc
     logger.msg(Arc::INFO, 
                "Logging UR set of %d URs.",
                urn);
-    logger.msg(Arc::VERBOSE, 
+    logger.msg(Arc::DEBUG, 
                "UR set dump: %s",
                urstr.c_str());
   

### 2009-11-11 src/clients/saml/saml_assertion_init.cpp

@@ -49,7 +49,7 @@ int main(int argc, char* argv[]){
 
     std::string debug;
     options.AddOption('d', "debug",
-                      istring("FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE"),
+                      istring("FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG"),
                       istring("debuglevel"), debug);
 
     bool version = false;

### 2009-11-11 src/clients/wsrf/arcwsrf.cpp

@@ -53,7 +53,7 @@ int main(int argc, char **argv) {
 
   std::string debug;
   options.AddOption('d', "debug",
-                    istring("FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE"),
+                    istring("FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG"),
                     istring("debuglevel"), debug);
 
   bool version = false;

### 2009-11-11 src/hed/acc/ARC0/SubmitterARC0.cpp

@@ -123,7 +123,7 @@ namespace Arc {
 
         if (it1->Name == it2->Name && (!it1->Source.empty() && !it2->Source.empty() ||
                                        !it1->Target.empty() && !it2->Target.empty())) {
-          logger.msg(DEBUG, "Two files have identical file name '%s'.", it1->Name);
+          logger.msg(VERBOSE, "Two files have identical file name '%s'.", it1->Name);
           return false;
         }
       }
@@ -190,21 +190,21 @@ namespace Arc {
     if (!jobdesc.Resources.RunTimeEnvironment.empty() &&
         !jobdesc.Resources.RunTimeEnvironment.selectSoftware(et.ApplicationEnvironments)) {
       // This error should never happen since RTE is checked in the Broker.
-      logger.msg(DEBUG, "Unable to select run time environment");
+      logger.msg(VERBOSE, "Unable to select run time environment");
       return false;
     }
 
     if (!jobdesc.Resources.CEType.empty() &&
         !jobdesc.Resources.CEType.selectSoftware(et.Implementation)) {
       // This error should never happen since Middleware is checked in the Broker.
-      logger.msg(DEBUG, "Unable to select middleware");
+      logger.msg(VERBOSE, "Unable to select middleware");
       return false;
     }
 
     if (!jobdesc.Resources.OperatingSystem.empty() &&
         !jobdesc.Resources.OperatingSystem.selectSoftware(et.Implementation)) {
       // This error should never happen since OS is checked in the Broker.
-      logger.msg(DEBUG, "Unable to select operating system.");
+      logger.msg(VERBOSE, "Unable to select operating system.");
       return false;
     }
 

### 2009-11-11 src/hed/acc/ARC1/JobControllerARC1.cpp

@@ -52,7 +52,7 @@ namespace Arc {
   bool JobControllerARC1::GetJob(const Job& job,
                                  const std::string& downloaddir) {
 
-    logger.msg(DEBUG, "Downloading job: %s", job.JobID.str());
+    logger.msg(VERBOSE, "Downloading job: %s", job.JobID.str());
 
     std::string path = job.JobID.Path();
     std::string::size_type pos = path.rfind('/');
@@ -115,7 +115,7 @@ namespace Arc {
       logger.msg(ERROR, "Job %s does not report a resumable state", job.JobID.str());
       //return false;
 
-    logger.msg(DEBUG, "Resuming job: %s at state: %s", job.JobID.str(), job.RestartState);
+    logger.msg(VERBOSE, "Resuming job: %s at state: %s", job.JobID.str(), job.RestartState);
 
     MCCConfig cfg;
     usercfg.ApplyToConfig(cfg);
@@ -124,7 +124,7 @@ namespace Arc {
     AREXClient::createActivityIdentifier(job.JobID, idstr);
     bool ok = ac.resume(idstr);
     if (ok)
-      logger.msg(DEBUG, "Job resuming successful");
+      logger.msg(VERBOSE, "Job resuming successful");
     return ok;
   }
 

### 2009-11-11 src/hed/acc/Broker/DataBroker.cpp

@@ -99,22 +99,22 @@ namespace Arc {
       iter++;
     }
 
-    logger.msg(DEBUG, "Matching against job description, following targets possible for DataBroker: %d", PossibleTargets.size());
+    logger.msg(VERBOSE, "Matching against job description, following targets possible for DataBroker: %d", PossibleTargets.size());
 
     iter = PossibleTargets.begin();
 
     for (int i = 1; iter != PossibleTargets.end(); iter++, i++)
-      logger.msg(DEBUG, "%d. Cluster: %s", i, (*iter)->DomainName);
+      logger.msg(VERBOSE, "%d. Cluster: %s", i, (*iter)->DomainName);
 
     CacheCheck();
     PossibleTargets.sort(DataCompare);
 
-    logger.msg(DEBUG, "Best targets are: %d", PossibleTargets.size());
+    logger.msg(VERBOSE, "Best targets are: %d", PossibleTargets.size());
 
     iter = PossibleTargets.begin();
 
     for (int i = 1; iter != PossibleTargets.end(); iter++, i++)
-      logger.msg(DEBUG, "%d. Cluster: %s", i, (*iter)->DomainName);
+      logger.msg(VERBOSE, "%d. Cluster: %s", i, (*iter)->DomainName);
 
     TargetSortingDone = true;
   }

### 2009-11-11 src/hed/acc/UNICORE/JobControllerUNICORE.cpp

@@ -85,7 +85,7 @@ namespace Arc {
   bool JobControllerUNICORE::GetJob(const Job& job,
                                     const std::string& downloaddir) {
     /*
-        logger.msg(DEBUG, "Downloading job: %s", job.JobID.str());
+        logger.msg(VERBOSE, "Downloading job: %s", job.JobID.str());
 
         std::string path = job.JobID.Path();
         std::string::size_type pos = path.rfind('/');

### 2009-11-11 src/hed/acc/UNICORE/UNICOREClient.cpp

@@ -153,7 +153,7 @@ namespace Arc {
       }
     }
     act_doc.GetXML(jsdl_str);
-    logger.msg(VERBOSE, "Job description to be sent: %s", jsdl_str);
+    logger.msg(DEBUG, "Job description to be sent: %s", jsdl_str);
 
     // Try to figure out which credentials are used
     // TODO: Method used is unstable beacuse it assumes some predefined
@@ -269,7 +269,7 @@ namespace Arc {
       std::string s;
       resp->GetXML(s);
       delete resp;
-      logger.msg(VERBOSE, "Submission returned failure: %s", s);
+      logger.msg(DEBUG, "Submission returned failure: %s", s);
       logger.msg(ERROR, "Submission failed, service returned: %s", faultstring);
       return false;
     }
@@ -342,7 +342,7 @@ namespace Arc {
       std::string s;
       resp->GetXML(s);
       delete resp;
-      logger.msg(VERBOSE, "Submission returned failure: %s", s);
+      logger.msg(DEBUG, "Submission returned failure: %s", s);
       logger.msg(ERROR, "Submission failed, service returned: %s", faultstring);
       return false;
     }
@@ -526,7 +526,7 @@ namespace Arc {
       return false;
     }
     XMLNode st;
-    logger.msg(VERBOSE, "Response:\n%s", (std::string)(*resp));
+    logger.msg(DEBUG, "Response:\n%s", (std::string)(*resp));
     (*resp)["GetFactoryAttributesDocumentResponse"]
     ["FactoryResourceAttributesDocument"].New(st);
     st.GetDoc(state, true);

### 2009-11-11 src/hed/dmc/srm/https/HTTPSClient.cpp

@@ -93,7 +93,7 @@ namespace Arc {
       if(!c->read(buf,&l)) return;
       if(!c->transfer(isread,iswritten,0)) { c->read(); return; };
       if(!isread) { c->read(); return; };
-      logger.msg(VERBOSE, "clear_input: %s", buf);
+      logger.msg(DEBUG, "clear_input: %s", buf);
     };
   }
   
@@ -163,7 +163,7 @@ namespace Arc {
         };
         line_buf[line_p]=0;
         if(line_p == 0) break; // end of header
-        logger.msg(VERBOSE, "read_response_header: line: %s", line_buf);
+        logger.msg(DEBUG, "read_response_header: line: %s", line_buf);
         analyze_response_line(line_buf);
         line_p=0;
       };
@@ -185,49 +185,49 @@ namespace Arc {
         return -1;
       };
     };
-    logger.msg(VERBOSE, "read_response_header: header finished");
+    logger.msg(DEBUG, "read_response_header: header finished");
     return 0;
   }
   
   int HTTPSClient::skip_response_entity(void) {
-    logger.msg(VERBOSE, "skip_response_entity");
+    logger.msg(DEBUG, "skip_response_entity");
     if(fields.haveContentLength() || fields.haveContentRange()) {
       unsigned long long int size = fields.ContentLength();
-      logger.msg(VERBOSE, "skip_response_entity: size: %llu", size);
+      logger.msg(DEBUG, "skip_response_entity: size: %llu", size);
       if(size<=answer_size) {
         memmove(answer_buf,answer_buf+size,answer_size-size);
         answer_size-=size;
-        logger.msg(VERBOSE, "skip_response_entity: already have all");
+        logger.msg(DEBUG, "skip_response_entity: already have all");
         return 0;
       };
       size-=answer_size;
-      logger.msg(VERBOSE, "skip_response_entity: size left: %llu", size);
+      logger.msg(DEBUG, "skip_response_entity: size left: %llu", size);
       // read size bytes
       char buf[1024];
       for(;size;) {
-        logger.msg(VERBOSE, "skip_response_entity:  to read: %llu", size);
+        logger.msg(DEBUG, "skip_response_entity:  to read: %llu", size);
         answer_size=sizeof(buf);
         if(!c->read(buf,&answer_size)) {
           disconnect(); return -1;
         };
         bool isread,iswritten;
[…]

### 2009-11-11 src/hed/dmc/srm/srmclient/SRM1Client.cpp

@@ -293,7 +293,7 @@
         };
         if(n<fstatus->__size) continue;
       };
-      logger.msg(Arc::DEBUG, "File could not be moved to Running state: %s", *f_url);
+      logger.msg(Arc::VERBOSE, "File could not be moved to Running state: %s", *f_url);
       file_id=file_ids.erase(file_id); f_url=urls.erase(f_url);
     };
     req.file_ids(file_ids);
@@ -412,7 +412,7 @@
         };
         if(n<fstatus->__size) continue;
       };
-      logger.msg(Arc::DEBUG, "File could not be moved to Done state");
+      logger.msg(Arc::VERBOSE, "File could not be moved to Done state");
       ++file_id;
     };
     req.file_ids(file_ids);

### 2009-11-11 src/hed/libs/client/ClientX509Delegation.cpp

@@ -295,7 +295,7 @@ namespace Arc {
         proxy_cred.OutputPrivatekey(delegation_cred);
         proxy_cred.OutputCertificateChain(delegation_cred);
 
-        logger.msg(VERBOSE,"Get delegated credential from delegation service: \n %s",delegation_cred.c_str());
+        logger.msg(DEBUG,"Get delegated credential from delegation service: \n %s",delegation_cred.c_str());
         return true;
       }
       else {

### 2009-11-11 src/hed/libs/client/JobSupervisor.cpp

@@ -35,7 +35,7 @@ namespace Arc {
 
     if (!jobs.empty()) {
 
-      logger.msg(DEBUG, "Identifying needed job controllers according to "
+      logger.msg(VERBOSE, "Identifying needed job controllers according to "
                  "specified jobs");
 
       for (std::list<std::string>::const_iterator it = jobs.begin();
@@ -60,7 +60,7 @@ namespace Arc {
 
           if (std::find(controllers.begin(), controllers.end(),
                         flavour) == controllers.end()) {
-            logger.msg(DEBUG, "Need job controller for grid flavour %s",
+            logger.msg(VERBOSE, "Need job controller for grid flavour %s",
                        flavour);
             controllers.push_back(flavour);
           }
@@ -70,14 +70,14 @@ namespace Arc {
 
     if (!usercfg.GetSelectedServices(COMPUTING).empty()) {
 
-      logger.msg(DEBUG, "Identifying needed job controllers according to "
+      logger.msg(VERBOSE, "Identifying needed job controllers according to "
                  "specified clusters");
 
       for (URLListMap::const_iterator it = usercfg.GetSelectedServices(COMPUTING).begin();
            it != usercfg.GetSelectedServices(COMPUTING).end(); it++)
         if (std::find(controllers.begin(), controllers.end(),
                       it->first) == controllers.end()) {
-          logger.msg(DEBUG, "Need job controller for grid flavour %s",
+          logger.msg(VERBOSE, "Need job controller for grid flavour %s",
                      it->first);
           controllers.push_back(it->first);
         }
@@ -85,7 +85,7 @@ namespace Arc {
 
     if (jobs.empty() && usercfg.GetSelectedServices(COMPUTING).empty()) {
 
-      logger.msg(DEBUG, "Identifying needed job controllers according to "
+      logger.msg(VERBOSE, "Identifying needed job controllers according to "
                  "all jobs present in job list");
 
       XMLNodeList xmljobs = jobstorage.XPathLookup("/ArcConfig/Job", NS());
@@ -100,7 +100,7 @@ namespace Arc {
         if (std::find(controllers.begin(), controllers.end(),
                       (std::string)(*it)["Flavour"]) == controllers.end()) {
           std::string flavour = (*it)["Flavour"];
-          logger.msg(DEBUG, "Need job controller for grid flavour %s",
+          logger.msg(VERBOSE, "Need job controller for grid flavour %s",
                      flavour);
           controllers.push_back(flavour);
         }

### 2009-11-11 src/hed/libs/client/XRSLParser.cpp

@@ -61,7 +61,7 @@ namespace Arc {
     RSLParser parser(source);
     const RSL *r = parser.Parse();
     if (!r) {
-      logger.msg(DEBUG, "RSL parsing error");
+      logger.msg(VERBOSE, "RSL parsing error");
       return JobDescription();
     }
 
@@ -461,15 +461,15 @@ namespace Arc {
         if (!ListValue(c, l))
           return false;
         if (l.size() < 2) {
-          logger.msg(DEBUG, "Syntax error in notify attribute. One or more job states and one or more email addresses must be specified.");
+          logger.msg(VERBOSE, "Syntax error in notify attribute. One or more job states and one or more email addresses must be specified.");
           return false;
         }
         if (l.front().find('@') != std::string::npos) {
-          logger.msg(DEBUG, "Syntax error in notify attribute. It cannot begin with an email address.");
+          logger.msg(VERBOSE, "Syntax error in notify attribute. It cannot begin with an email address.");
           return false;
         }
         if (l.back().find('@') == std::string::npos) {
-          logger.msg(DEBUG, "Syntax error in notify attribute. It must end with an email address.");
+          logger.msg(VERBOSE, "Syntax error in notify attribute. It must end with an email address.");
           return false;
         }
 
@@ -525,7 +525,7 @@ namespace Arc {
           else if (*it == "outbound")
             j.Resources.NodeAccess = (j.Resources.NodeAccess == NAT_INBOUND || j.Resources.NodeAccess == NAT_INOUTBOUND ? NAT_INOUTBOUND : NAT_OUTBOUND);
           else {
-            logger.msg(DEBUG, "Invalid nodeaccess value: %s", *it);
+            logger.msg(VERBOSE, "Invalid nodeaccess value: %s", *it);
             return false;
           }
         return true;
@@ -590,7 +590,7 @@ namespace Arc {
         if (!SingleValue(c, action))
           return false;
         if (action != "request" && action != "cancel" && action != "clean") {
-          logger.msg(DEBUG, "Invalid action value %s", action);
+          logger.msg(VERBOSE, "Invalid action value %s", action);
           return false;
         }
         j.XRSL_elements["action"] = action;
@@ -664,7 +664,7 @@ namespace Arc {
         return true;
       }
 
-      logger.msg(DEBUG, "Unknown XRSL attribute: %s - Ignoring it.", c->Attr());
+      logger.msg(VERBOSE, "Unknown XRSL attribute: %s - Ignoring it.", c->Attr());
       return true;
     }
     else {

### 2009-11-11 src/hed/libs/client/confusa/OAuthConsumer.cpp

@@ -76,11 +76,11 @@ namespace Arc {
 		// we need to pass that through an URL to have the port in it, so checking the SHA1 signature verify on the OAuth SP will not fail
 		URL request_url((*sso_pages_)["SimpleSAML"] + REQUEST_TOKEN_URL);
 		std::string request_url_str = request_url.fullstr();
-		logger.msg(DEBUG, "The request_url is %s", request_url_str);
+		logger.msg(VERBOSE, "The request_url is %s", request_url_str);
 
 		// produce an OAuth request token
 		char *req_url = oauth_sign_url(request_url_str.c_str(), NULL, OA_HMAC, consumer_key.c_str(), consumer_secret.c_str(), NULL, NULL);
-		logger.msg(DEBUG, "Sending oauth request to signed url %s", req_url);
+		logger.msg(VERBOSE, "Sending oauth request to signed url %s", req_url);
 
 		if (!req_url) {
 			return MCC_Status(GENERIC_ERROR, origin, "Could not sign the request url " + request_url);
@@ -219,7 +219,7 @@ namespace Arc {
 		std::string about_url = sp_url.fullstr();
 
 		char *req_url = oauth_sign_url(about_url.c_str(), NULL, OA_HMAC, consumer_key.c_str(), consumer_secret.c_str(), oauth_access_token.c_str(), oauth_access_secret.c_str());
-		logger.msg(DEBUG, "The about-you request url is %s", req_url);
+		logger.msg(VERBOSE, "The about-you request url is %s", req_url);
 		ClientHTTP about_client(cfg_,URL(req_url));
 		PayloadRaw about_request;
 		PayloadRawInterface *about_response = NULL;
@@ -259,7 +259,7 @@ namespace Arc {
 		logger.msg(INFO, "Approving the cert signing request at %s", approve_page);
 
 		char *req_url = oauth_sign_url(approve_page.c_str(), NULL, OA_HMAC, consumer_key.c_str(), consumer_secret.c_str(), oauth_access_token.c_str(), oauth_access_secret.c_str());
-		logger.msg(DEBUG, "The OAuth request url is %s", req_url);
+		logger.msg(VERBOSE, "The OAuth request url is %s", req_url);
 
 		ClientHTTP confusa_approve_client(cfg_, URL(req_url));
 		PayloadRaw confusa_approve_request;
@@ -297,7 +297,7 @@ namespace Arc {
[…]

### 2009-11-11 src/hed/libs/common/Thread.cpp

@@ -81,7 +81,7 @@ namespace Arc {
     if(n > INT_MAX) n = INT_MAX;
     max_count = (int)n;
     // TODO: can't use logger here because it will try to initilize pool
-    //threadLogger.msg(VERBOSE, "Maximum number of threads is %i",max_count);
+    //threadLogger.msg(DEBUG, "Maximum number of threads is %i",max_count);
   }
   int ThreadPool::CheckQueue(void) {
     Glib::Mutex::Lock lock(queue_lock, Glib::TRY_LOCK);

### 2009-11-11 src/hed/libs/credential/testeec.cpp

@@ -11,7 +11,7 @@
 int main(void) {
   Arc::LogStream cdest(std::cerr);
   Arc::Logger::getRootLogger().addDestination(cdest);
-  Arc::Logger::getRootLogger().setThreshold(Arc::DEBUG);
+  Arc::Logger::getRootLogger().setThreshold(Arc::VERBOSE);
 
   std::string CAcert("./CAcert.pem"); 
   std::string CAkey("./CAkey.pem");

### 2009-11-11 src/hed/libs/credential/testvoms.cpp

@@ -21,7 +21,7 @@ int main(void) {
 int main(void) {
   Arc::LogStream cdest(std::cerr);
   Arc::Logger::getRootLogger().addDestination(cdest);
-  Arc::Logger::getRootLogger().setThreshold(Arc::DEBUG);
+  Arc::Logger::getRootLogger().setThreshold(Arc::VERBOSE);
   std::string cert("../../../tests/echo/testcert.pem");
   std::string key("../../../tests/echo/testkey-nopass.pem");
   std::string cafile("../../../tests/echo/testcacert.pem");

### 2009-11-11 src/hed/libs/loader/test.cpp

@@ -11,7 +11,7 @@ int main(void)
 {
     Arc::LogStream ldest(std::cerr);
     Arc::Logger::getRootLogger().addDestination(ldest);
-    Arc::Logger::getRootLogger().setThreshold(Arc::DEBUG);
+    Arc::Logger::getRootLogger().setThreshold(Arc::VERBOSE);
     Arc::Config *c = new Arc::Config("test.xml");    
     Arc::Loader chains(c);
     return 0;

### 2009-11-11 src/hed/libs/message/Plexer.cpp

@@ -42,7 +42,7 @@ namespace Arc {
   MCC_Status Plexer::process(Message& request, Message& response){
     std::string ep = request.Attributes()->get("ENDPOINT");
     std::string path = getPath(ep);
-    logger.msg(DEBUG, "Operation on path \"%s\"",path);
+    logger.msg(VERBOSE, "Operation on path \"%s\"",path);
     std::list<PlexerEntry>::iterator iter;
     for (iter=services.begin(); iter!=services.end(); ++iter) {
       std::list<std::string> unmatched, matched;

### 2009-11-11 src/hed/libs/python/arcom/logger.py

@@ -1,7 +1,7 @@
 import arc
 import traceback
 
-log_levels = [arc.VERBOSE, arc.DEBUG, arc.INFO, arc.WARNING, arc.ERROR, arc.FATAL]
+log_levels = [arc.DEBUG, arc.VERBOSE, arc.INFO, arc.WARNING, arc.ERROR, arc.FATAL]
 
 def get_logger(system = '<UNKNOWN>'):
     return Logger(arc.Logger(arc.Logger_getRootLogger(), system))
@@ -16,11 +16,11 @@ class Logger:
         # initializing logging facility
         args = list(args)
         if not args:
-            severity = arc.DEBUG
+            severity = arc.VERBOSE
         elif args[0] in log_levels:
             severity = args.pop(0)
         else:
-            severity = arc.DEBUG
+            severity = arc.VERBOSE
         if not args:
             args = ['Python exception:\n', traceback.format_exc()]
             severity = arc.ERROR

### 2009-11-11 src/hed/libs/python/arcom/store/cachedpicklestore.py

@@ -19,7 +19,7 @@ class CachedPickleStore(PickleStore):
         'non_existent_object' will be returned if an object not found
         """
         PickleStore.__init__(self, storecfg, non_existent_object)
-        log.msg(arc.DEBUG, "PickleStore with datadir '%s' is a CachedPickleStore" % self.datadir)
+        log.msg(arc.VERBOSE, "PickleStore with datadir '%s' is a CachedPickleStore" % self.datadir)
         self.store = {}
         self._load_storage()
 

### 2009-11-11 src/hed/libs/python/arcom/store/transdbstore.py

@@ -31,8 +31,8 @@ class TransDBStore(BaseStore):
         """
 
         BaseStore.__init__(self, storecfg, non_existent_object)
-        log.msg(arc.DEBUG, "TransDBStore constructor called")
-        log.msg(arc.DEBUG, "datadir:", self.datadir)
+        log.msg(arc.VERBOSE, "TransDBStore constructor called")
+        log.msg(arc.VERBOSE, "datadir:", self.datadir)
 
         # db and transaction pointers
         self.dbp  = None
@@ -139,7 +139,7 @@ class TransDBStore(BaseStore):
                 # It can take a while for the master site to be found and
                 # synced, and no DB will be available until then
                 if errnum == errno.ENOENT:
-                    log.msg(arc.DEBUG, "No stock db available yet - retrying")
+                    log.msg(arc.VERBOSE, "No stock db available yet - retrying")
                     log.msg()
                     try:
                         dbp.close(0)
@@ -154,7 +154,7 @@ class TransDBStore(BaseStore):
                     time.sleep(self.sleeptime)
                     continue
                 elif errnum == db.DB_LOCK_DEADLOCK:
-                    log.msg(arc.DEBUG, "got deadlock - retrying")
+                    log.msg(arc.VERBOSE, "got deadlock - retrying")
                     try:
                         dbp.close(0)
                     except db.DBError, (errnum2, strerror2):
@@ -282,12 +282,12 @@ class TransDBStore(BaseStore):
                 # need to close transaction handle as well
                 time.sleep(0.2)
                 if retry_count < self.deadlock_retries:
-                    log.msg(arc.DEBUG, "got DBLockDeadlockError")
+                    log.msg(arc.VERBOSE, "got DBLockDeadlockError")
                     retry_count += 1
-                    log.msg(arc.DEBUG, "retrying transaction", retry_count)
+                    log.msg(arc.VERBOSE, "retrying transaction", retry_count)
                     retry = True
                 else:
-                    log.msg(arc.DEBUG, "Deadlock exception, giving up...")
+                    log.msg(arc.VERBOSE, "Deadlock exception, giving up...")
                     retry = False
             except db.DBRepHandleDeadError:
                 log.msg(arc.INFO, "Got rep_dead_handle error")            

### 2009-11-11 src/hed/libs/security/ArcPDP/attr/RequestAttribute.cpp

@@ -34,7 +34,7 @@ RequestAttribute::RequestAttribute(XMLNode& node, AttributeFactory* attrfactory)
   //Create the attribute value object according to the data type
   attrval = attrfactory->createValue(node, type);
 
-  logger.msg(DEBUG, "Id= %s,Type= %s,Issuer= %s,Value= %s",id, type, issuer, attrval?attrval->encode():"");
+  logger.msg(VERBOSE, "Id= %s,Type= %s,Issuer= %s,Value= %s",id, type, issuer, attrval?attrval->encode():"");
 
   if(attrval == NULL)
     logger.msg(ERROR,"No Attribute exists, which can deal with type: %s", type);

### 2009-11-11 src/hed/mcc/gsi/MCCGSI.cpp

@@ -157,7 +157,7 @@ namespace Arc {
     recv_tok.value = malloc(recv_tok.length);
     memcpy(recv_tok.value, readbuf, 5);
 
-    logger.msg(DEBUG, "Recieved token length: %i", recv_tok.length);
+    logger.msg(VERBOSE, "Recieved token length: %i", recv_tok.length);
 
     while (recv_tok.length > pos) {
       int len = recv_tok.length - pos;
@@ -189,7 +189,7 @@ namespace Arc {
 
       logger.msg(INFO, "GSS accept security context: %i/%i", majstat, minstat);
 
-      logger.msg(DEBUG, "Returned token length: %i", send_tok.length);
+      logger.msg(VERBOSE, "Returned token length: %i", send_tok.length);
 
       PayloadRaw *outpayload = new PayloadRaw;
       if (send_tok.length > 0)
@@ -216,7 +216,7 @@ namespace Arc {
 
       logger.msg(INFO, "GSS unwrap: %i/%i", majstat, minstat);
 
-      logger.msg(DEBUG, "Sent token length: %i", send_tok.length);
+      logger.msg(VERBOSE, "Sent token length: %i", send_tok.length);
 
       PayloadRaw payload;
       payload.Insert((const char*)send_tok.value, 0, send_tok.length);
@@ -340,7 +340,7 @@ namespace Arc {
       recv_tok.value = inpayload->Buffer(n);
       recv_tok.length = inpayload->BufferSize(n);
 
-      logger.msg(DEBUG, "Recieved token length: %i", recv_tok.length);
+      logger.msg(VERBOSE, "Recieved token length: %i", recv_tok.length);
 
       OM_uint32 majstat, minstat;
 
@@ -358,7 +358,7 @@ namespace Arc {
 
       logger.msg(INFO, "GSS wrap: %i/%i", majstat, minstat);
 
-      logger.msg(DEBUG, "Sent token length: %i", send_tok.length);
+      logger.msg(VERBOSE, "Sent token length: %i", send_tok.length);
 
       gsipayload.Insert((const char*)send_tok.value, size, send_tok.length);
       size += send_tok.length;
@@ -461,7 +461,7 @@ namespace Arc {
     namebuf.value = (void*)hostname.c_str();
     namebuf.length = hostname.size();
 
-    logger.msg(DEBUG, "Peer host name to which this client will access: %s", hostname.c_str());
[…]

### 2009-11-11 src/hed/mcc/tcp/PayloadTCPSocket.cpp

@@ -39,17 +39,17 @@ int PayloadTCPSocket::connect_socket(const char* hostname,int port)
   }
   int s = -1;
   for(struct addrinfo *info_ = info;info_;info_=info_->ai_next) {
-    logger.msg(DEBUG,"Trying to connect %s(%s):%d",
+    logger.msg(VERBOSE,"Trying to connect %s(%s):%d",
                      hostname,info_->ai_family==AF_INET6?"IPv6":"IPv4",port);
     s = ::socket(info_->ai_family, info_->ai_socktype, info_->ai_protocol);
     if(s == -1) {
       // TODO: print error description
-      logger.msg(DEBUG, "Failed to create socket to %s(%s):%d",
+      logger.msg(VERBOSE, "Failed to create socket to %s(%s):%d",
                         hostname,info_->ai_family==AF_INET6?"IPv6":"IPv4",port);
       continue;
     }
     if(::connect(s, info_->ai_addr, info_->ai_addrlen) == -1) {
-      logger.msg(DEBUG, "Failed to connect to %s(%s):%i",
+      logger.msg(VERBOSE, "Failed to connect to %s(%s):%i",
                         hostname,info_->ai_family==AF_INET6?"IPv6":"IPv4",port);
       close(s); s = -1;
       continue;
@@ -57,7 +57,7 @@ int PayloadTCPSocket::connect_socket(const char* hostname,int port)
     break;
   };
   if(s == -1) {
-    logger.msg(DEBUG, "Failed to establish connection to %s:%i", hostname, port);
+    logger.msg(VERBOSE, "Failed to establish connection to %s:%i", hostname, port);
   };
   freeaddrinfo(info);
   return s;

### 2009-11-11 src/hed/mcc/tls/MCCTLS.cpp

@@ -163,7 +163,7 @@ TLSSecAttr::TLSSecAttr(PayloadTLSStream& payload, ConfigTLSMCC& config, Logger& 
       buf[0]=0;
       X509_NAME_oneline(X509_get_subject_name(peercert),buf,sizeof buf);
       subject=buf;
-      //logger.msg(DEBUG, "Peer name: %s", peer_dn);
+      //logger.msg(VERBOSE, "Peer name: %s", peer_dn);
       subjects_.push_back(subject);
 #ifdef HAVE_OPENSSL_PROXY
       if(X509_get_ext_by_NID(peercert,NID_proxyCertInfo,-1) < 0) {
@@ -185,7 +185,7 @@ TLSSecAttr::TLSSecAttr(PayloadTLSStream& payload, ConfigTLSMCC& config, Logger& 
       buf[0]=0;
       X509_NAME_oneline(X509_get_subject_name(hostcert),buf,sizeof buf);
       target_=buf;
-      //logger.msg(DEBUG, "Host name: %s", peer_dn);
+      //logger.msg(VERBOSE, "Host name: %s", peer_dn);
    };
 }
 
@@ -364,11 +364,11 @@ MCC_Status MCC_TLS_Service::process(Message& inmsg,Message& outmsg) {
       TLSSecAttr* sattr = new TLSSecAttr(*tstream, config_, logger);
       nextinmsg.Auth()->set("TLS",sattr);
       //Getting the subject name of peer(client) certificate
-      logger.msg(DEBUG, "Peer name: %s", sattr->Subject());
+      logger.msg(VERBOSE, "Peer name: %s", sattr->Subject());
       nextinmsg.Attributes()->set("TLS:PEERDN",sattr->Subject());
-      logger.msg(DEBUG, "Identity name: %s", sattr->Identity());
+      logger.msg(VERBOSE, "Identity name: %s", sattr->Identity());
       nextinmsg.Attributes()->set("TLS:IDENTITYDN",sattr->Identity());
-      logger.msg(DEBUG, "CA name: %s", sattr->CA());
+      logger.msg(VERBOSE, "CA name: %s", sattr->CA());
       nextinmsg.Attributes()->set("TLS:CADN",sattr->CA());
 
       nextinmsg.Attributes()->set("TLS:PEERCERT",sattr->X509Str());
@@ -427,11 +427,11 @@ MCC_Status MCC_TLS_Client::process(Message& inmsg,Message& outmsg) {
[…]

### 2009-11-11 src/hed/shc/gaclpdp/GACLPDP.cpp

@@ -146,10 +146,10 @@ bool GACLPDP::isPermitted(Message *msg){
     };
     delete cauth;
   };
-  if(VERBOSE >= logger.getThreshold()) {
+  if(DEBUG >= logger.getThreshold()) {
     std::string s;
     requestxml.GetXML(s);
-    logger.msg(VERBOSE,"GACL Auth. request: %s",s);
+    logger.msg(DEBUG,"GACL Auth. request: %s",s);
   };
   if(requestxml.Size() <= 0) {
     logger.msg(ERROR,"No requested security information was collected");

### 2009-11-11 src/services/a-rex/arex.cpp

@@ -191,7 +191,7 @@ ARexConfigContext* ARexService::get_configuration(Arc::Message& inmsg) {
     logger_.msg(Arc::ERROR, "No local account name specified");
     return NULL;
   };
-  logger_.msg(Arc::VERBOSE,"Using local account '%s'",uname);
+  logger_.msg(Arc::DEBUG,"Using local account '%s'",uname);
   std::string grid_name = inmsg.Attributes()->get("TLS:IDENTITYDN");
   std::string endpoint = endpoint_;
   if(endpoint.empty()) {
@@ -242,8 +242,8 @@ Arc::MCC_Status ARexService::process(Arc::Message& inmsg,Arc::Message& outmsg) {
   std::string id = GetPath(inmsg,endpoint);
   std::string clientid = (inmsg.Attributes()->get("TCP:REMOTEHOST"))+":"+(inmsg.Attributes()->get("TCP:REMOTEPORT"));
   if((inmsg.Attributes()->get("PLEXER:PATTERN").empty()) && id.empty()) id=endpoint;
-  logger_.msg(Arc::DEBUG, "process: method: %s", method);
-  logger_.msg(Arc::DEBUG, "process: endpoint: %s", endpoint);
+  logger_.msg(Arc::VERBOSE, "process: method: %s", method);
+  logger_.msg(Arc::VERBOSE, "process: endpoint: %s", endpoint);
   while(id[0] == '/') id=id.substr(1);
   std::string subpath;
   {
@@ -254,8 +254,8 @@ Arc::MCC_Status ARexService::process(Arc::Message& inmsg,Arc::Message& outmsg) {
       while(subpath[0] == '/') subpath=subpath.substr(1);
     };
   };
-  logger_.msg(Arc::DEBUG, "process: id: %s", id);
-  logger_.msg(Arc::DEBUG, "process: subpath: %s", subpath);
+  logger_.msg(Arc::VERBOSE, "process: id: %s", id);
+  logger_.msg(Arc::VERBOSE, "process: subpath: %s", subpath);
 
   // Process grid-manager configuration if not done yet
   ARexConfigContext* config = get_configuration(inmsg);
@@ -278,7 +278,7 @@ Arc::MCC_Status ARexService::process(Arc::Message& inmsg,Arc::Message& outmsg) {
   // Using simplified algorithm - POST for SOAP messages,
   // GET and PUT for data transfer
   if(method == "POST") {
[…]

### 2009-11-11 src/services/a-rex/change_activity_status.cpp

@@ -32,7 +32,7 @@ Arc::MCC_Status ARexService::ChangeActivityStatus(ARexGMConfig& config,Arc::XMLN
   {
     std::string s;
     in.GetXML(s);
-    logger_.msg(Arc::DEBUG, "ChangeActivityStatus: request = \n%s", s);
+    logger_.msg(Arc::VERBOSE, "ChangeActivityStatus: request = \n%s", s);
   };
   Arc::WSAEndpointReference id(in["ActivityIdentifier"]);
   if(!(Arc::XMLNode)id) {
@@ -167,7 +167,7 @@ Arc::MCC_Status ARexService::ChangeActivityStatus(ARexGMConfig& config,Arc::XMLN
   {
     std::string s;
     out.GetXML(s);
-    logger_.msg(Arc::DEBUG, "ChangeActivityStatus: response = \n%s", s);
+    logger_.msg(Arc::VERBOSE, "ChangeActivityStatus: response = \n%s", s);
   };
   return Arc::MCC_Status(Arc::STATUS_OK);
 }

### 2009-11-11 src/services/a-rex/get_activity_documents.cpp

@@ -26,7 +26,7 @@ Arc::MCC_Status ARexService::GetActivityDocuments(ARexGMConfig& config,Arc::XMLN
   {
     std::string s;
     in.GetXML(s);
-    logger.msg(Arc::DEBUG, "GetActivityDocuments: request = \n%s", s);
+    logger.msg(Arc::VERBOSE, "GetActivityDocuments: request = \n%s", s);
   };
   for(int n = 0;;++n) {
     Arc::XMLNode id = in["ActivityIdentifier"][n];
@@ -69,7 +69,7 @@ Arc::MCC_Status ARexService::GetActivityDocuments(ARexGMConfig& config,Arc::XMLN
   {
     std::string s;
     out.GetXML(s);
-    logger_.msg(Arc::DEBUG, "GetActivityDocuments: response = \n%s", s);
+    logger_.msg(Arc::VERBOSE, "GetActivityDocuments: response = \n%s", s);
   };
   return Arc::MCC_Status(Arc::STATUS_OK);
 }

### 2009-11-11 src/services/a-rex/grid-manager/jobs/states.cpp

@@ -492,7 +492,7 @@ bool JobsList::state_loading(const JobsList::iterator &i,bool &state_changed,boo
     };
   } else {
     if(i->child->Running()) {
-      logger.msg(Arc::DEBUG,"%s: State: PREPARING/FINISHING: child is running",i->job_id);
+      logger.msg(Arc::VERBOSE,"%s: State: PREPARING/FINISHING: child is running",i->job_id);
       /* child is running - come later */
       return true;
     };
@@ -543,7 +543,7 @@ bool JobsList::state_loading(const JobsList::iterator &i,bool &state_changed,boo
           i->AddFailure("Failed in files download due to expired credentials - try to renew");
         };
       } else if(i->child->Result() == 4) { // retryable cache error
-        logger.msg(Arc::VERBOSE, "%s: State: PREPARING/FINISHING: retryable error", i->job_id);
+        logger.msg(Arc::DEBUG, "%s: State: PREPARING/FINISHING: retryable error", i->job_id);
         delete i->child; i->child=NULL;
         return true;
       } 
@@ -743,7 +743,7 @@ void JobsList::ActJobAccepted(JobsList::iterator &i,bool /*hard_job*/,
       /* accepted state - job was just accepted by jobmager-ng and we already
          know that it is accepted - now we are analyzing/parsing request,
          or it can also happen we are waiting for user specified time */
-        logger.msg(Arc::DEBUG,"%s: State: ACCEPTED",i->job_id);
+        logger.msg(Arc::VERBOSE,"%s: State: ACCEPTED",i->job_id);
         if(!GetLocalDescription(i)) {
           job_error=true; i->AddFailure("Internal error");
           return; /* go to next job */
@@ -792,7 +792,7 @@ void JobsList::ActJobPreparing(JobsList::iterator &i,bool /*hard_job*/,
         /* preparing state - means job is parsed and we are going to download or
            already downloading input files. process downloader is run for
            that. it also checks for files user interface have to upload itself*/
-        logger.msg(Arc::DEBUG,"%s: State: PREPARING",i->job_id);
+        logger.msg(Arc::VERBOSE,"%s: State: PREPARING",i->job_id);
         if(i->job_pending || state_loading(i,state_changed,false)) {
           if(i->job_pending || state_changed) {
             if((JOB_NUM_RUNNING<max_jobs_running) || (max_jobs_running==-1)) {
[…]

### 2009-11-11 src/services/a-rex/grid-manager/jobs/users.cpp

@@ -365,7 +365,7 @@ bool JobUserHelper::run(JobUser &user) {
       args[n]=strdup(arg_s.c_str());
     };
     args[n]=NULL;
-    logger.msg(Arc::DEBUG,"Starting helper process (%s): %s",
+    logger.msg(Arc::VERBOSE,"Starting helper process (%s): %s",
                user.UnixName().c_str(),command.c_str());
     std::string helper_id="helper."+user.UnixName();
     bool started=RunParallel::run(user,helper_id.c_str(),args,&proc);

### 2009-11-11 src/services/a-rex/grid-manager/loaders/janitor.cpp

@@ -40,12 +40,12 @@ Janitor::Janitor(const std::string& id,const std::string& cdir):
   // create path to janitor utility
   path_ = Glib::build_filename(nordugrid_libexec_loc(),"janitor");
   if(path_.empty()) {
-    logger.msg(Arc::DEBUG, "Failed to create path to janitor at %s",nordugrid_libexec_loc());
+    logger.msg(Arc::VERBOSE, "Failed to create path to janitor at %s",nordugrid_libexec_loc());
     return;
   };
   // check if utility exists
   if(!Glib::file_test(path_,Glib::FILE_TEST_IS_EXECUTABLE)) {
-    logger.msg(Arc::DEBUG, "Janitor executable not found at %s",path_);
+    logger.msg(Arc::VERBOSE, "Janitor executable not found at %s",path_);
     path_.resize(0);
     return;
   };
@@ -86,14 +86,14 @@ void Janitor::deploy_thread(void* arg) {
   {
     Arc::Run run(cmd);
     if(!run) {
-      logger.msg(Arc::VERBOSE, "Can't run %s", cmd);
+      logger.msg(Arc::DEBUG, "Can't run %s", cmd);
       it.completed_.signal();
       return;
     };
     run.KeepStdout(true);
     run.KeepStderr(true);
     if(!run.Start()) {
-      logger.msg(Arc::VERBOSE, "Can't start %s", cmd);
+      logger.msg(Arc::DEBUG, "Can't start %s", cmd);
       it.completed_.signal();
       return;
     };
@@ -103,12 +103,12 @@ void Janitor::deploy_thread(void* arg) {
       if(it.cancel_.wait(0)) { it.completed_.signal(); return; };
     };
     if(run.Result() == 0) {
-      logger.msg(Arc::VERBOSE, "janitor register returned 0 - no RTE needs to be deployed");
+      logger.msg(Arc::DEBUG, "janitor register returned 0 - no RTE needs to be deployed");
       it.result_=DEPLOYED; it.completed_.signal();
       return;
     };
     if(run.Result() == 3) {
-      logger.msg(Arc::VERBOSE, "janitor register returned 3 - no Janitor enabled in configuration");
+      logger.msg(Arc::DEBUG, "janitor register returned 3 - no Janitor enabled in configuration");
       it.result_=NOTENABLED; it.completed_.signal();
       return;
     };
@@ -117,7 +117,7 @@ void Janitor::deploy_thread(void* arg) {
[…]

### 2009-11-11 src/services/echo_python/EchoService.py

@@ -66,8 +66,8 @@ class EchoService:
         # time.sleep(10)
         # get the payload from the message
         inpayload = inmsg.Payload()
-        log.msg(arc.DEBUG, 'AuthRequest(inmsg) = ', AuthRequest(inmsg))
-        log.msg(arc.DEBUG, 'inmsg.Attributes().getAll() = ', inmsg.Attributes().getAll())
+        log.msg(arc.VERBOSE, 'AuthRequest(inmsg) = ', AuthRequest(inmsg))
+        log.msg(arc.VERBOSE, 'inmsg.Attributes().getAll() = ', inmsg.Attributes().getAll())
         log.msg(arc.INFO, "EchoService (python) got:", inpayload.GetXML())
         # the first child of the payload should be the name of the request
         request_node = inpayload.Child()

### 2009-11-11 src/services/hopi/hopi.cpp

@@ -73,7 +73,7 @@ time_t HopiFileChunks::last_timeout = time(NULL);
 void HopiFileChunks::Print(void) {
   int n = 0;
   for(chunks_t::iterator c = chunks.begin();c!=chunks.end();++c) {
-    Hopi::logger.msg(Arc::VERBOSE, "Chunk %u: %u - %u",n,c->first,c->second);
+    Hopi::logger.msg(Arc::DEBUG, "Chunk %u: %u - %u",n,c->first,c->second);
   };
 }
 
@@ -245,7 +245,7 @@ HopiFile::~HopiFile(void) {
     if(!for_read) {
       if(chunks.Complete()) {
         if(slave) {
-          Hopi::logger.msg(Arc::DEBUG, "Removing complete file in slave mode");
+          Hopi::logger.msg(Arc::VERBOSE, "Removing complete file in slave mode");
           ::unlink(path.c_str());
         }
         chunks.Remove();
@@ -488,7 +488,7 @@ static off_t GetEntitySize(Arc::MessagePayload &payload) {
 Arc::MCC_Status Hopi::Put(const std::string &path, Arc::MessagePayload &payload)
 {
     // XXX eliminate relative paths first
-    logger.msg(Arc::DEBUG, "PUT called");
+    logger.msg(Arc::VERBOSE, "PUT called");
     std::string full_path = Glib::build_filename(doc_root, path);
     if (slave_mode && (Glib::file_test(full_path, Glib::FILE_TEST_EXISTS) == false)) {
         logger.msg(Arc::ERROR, "Hopi SlaveMode is active, PUT is only allowed to existing files");        
@@ -499,7 +499,7 @@ Arc::MCC_Status Hopi::Put(const std::string &path, Arc::MessagePayload &payload)
         return Arc::MCC_Status();
     }
     fd.Size(GetEntitySize(payload));
-    logger.msg(Arc::VERBOSE, "File size is %u",fd.Size());
+    logger.msg(Arc::DEBUG, "File size is %u",fd.Size());
     try {
         Arc::PayloadStreamInterface& stream = dynamic_cast<Arc::PayloadStreamInterface&>(payload);
         char sbuf[1024*1024];
@@ -508,13 +508,13 @@ Arc::MCC_Status Hopi::Put(const std::string &path, Arc::MessagePayload &payload)
             off_t offset = stream.Pos();
             if(!stream.Get(sbuf,size)) {
                 if(!stream) {
-                    logger.msg(Arc::DEBUG, "error reading from HTTP stream");
[…]

### 2009-11-11 src/services/janitor/interfaces/webservice/dREWebService.cpp

@@ -85,22 +85,22 @@ namespace DREService
 	*/
 	DREWebService::~DREWebService(void) 
 	{
-		logger.msg(Arc::VERBOSE, " Deconstructing Web Service");
-		logger.msg(Arc::VERBOSE, " Flushing set and queue");
+		logger.msg(Arc::DEBUG, " Deconstructing Web Service");
+		logger.msg(Arc::DEBUG, " Flushing set and queue");
 		pTaskqueue->flush();
 		pTaskset->flush();
 
-		logger.msg(Arc::VERBOSE, " Deconstructing is waiting for PerlProcessor");
+		logger.msg(Arc::DEBUG, " Deconstructing is waiting for PerlProcessor");
 		delete pPerlProcessor; // holds taskqueue and taskset
 
 
-		logger.msg(Arc::VERBOSE, " Deconstructing is waiting for TaskQueue");
+		logger.msg(Arc::DEBUG, " Deconstructing is waiting for TaskQueue");
 		delete pTaskqueue;
-		logger.msg(Arc::VERBOSE, " Deconstructing is waiting for TaskSet");
+		logger.msg(Arc::DEBUG, " Deconstructing is waiting for TaskSet");
 		delete pTaskset;
 
 		
-		logger.msg(Arc::VERBOSE, " Deconstructing Web Service ... done");
+		logger.msg(Arc::DEBUG, " Deconstructing Web Service ... done");
 	}
 
 	/**
@@ -129,7 +129,7 @@ namespace DREService
 		Task::Task *pTask = new Task::Task(-1,&inmsg, &outmsg);
 		pTask->bla = 666;
 
-		logger.msg(Arc::VERBOSE, "DREWEBSERVICE 1  %d",pTask->bla);
+		logger.msg(Arc::DEBUG, "DREWEBSERVICE 1  %d",pTask->bla);
 		int taskID = pTaskqueue->pushTask(pTask);
 		pTaskset->removeTask(taskID);  // we don't want the return pointer. 
 		                               // the result should be also in our original one
@@ -137,7 +137,7 @@ namespace DREService
 		
 		// We don't have to extract the outmsg either
 		// because the object is already manipulated
-		logger.msg(Arc::VERBOSE, "DREWEBSERVICE 2  %d",pTask->bla);
+		logger.msg(Arc::DEBUG, "DREWEBSERVICE 2  %d",pTask->bla);
 	
 
[…]

### 2009-11-11 src/services/paul/info.cpp

@@ -123,7 +123,7 @@ bool PaulService::information_collector(Arc::XMLNode &doc)
 }
 
 bool PaulService::RegistrationCollector(Arc::XMLNode &doc) {
-  logger_.msg(Arc::VERBOSE,"Passing service's information from collector to registrator");
+  logger_.msg(Arc::DEBUG,"Passing service's information from collector to registrator");
   Arc::XMLNode empty(ns_, "RegEntry");
   empty.New(doc);
 

### 2009-11-11 src/services/saml/SPService.cpp

@@ -276,15 +276,15 @@ Arc::MCC_Status Service_SP::process(Arc::Message& inmsg,Arc::Message& outmsg) {
     bool must_signed = true; //TODO: get the information from metadata
     std::string authnRequestQuery;
     std::string query = "SAMLRequest=" + BuildDeflatedQuery(authn_request);
-    logger.msg(Arc::VERBOSE,"AuthnRequest after deflation: %s", query.c_str());
+    logger.msg(Arc::DEBUG,"AuthnRequest after deflation: %s", query.c_str());
     if(must_signed) {
       //SP service uses it's private key to sign the AuthnRequest,
       //then after User Agent redirecting AuthnRequest to IdP, 
       //IdP will verify the signature by picking up SP's certificate
       //from IdP's metadata.
-      logger.msg(Arc::VERBOSE,"Using private key file to sign: %s", privkey_file_.c_str());
+      logger.msg(Arc::DEBUG,"Using private key file to sign: %s", privkey_file_.c_str());
       authnRequestQuery = SignQuery(query, Arc::RSA_SHA1, privkey_file_);
-      logger.msg(Arc::VERBOSE,"After signature: %s", authnRequestQuery.c_str());
+      logger.msg(Arc::DEBUG,"After signature: %s", authnRequestQuery.c_str());
     }
     else authnRequestQuery = query;
 
@@ -319,7 +319,7 @@ Arc::MCC_Status Service_SP::process(Arc::Message& inmsg,Arc::Message& outmsg) {
       //Decrypt the encrypted saml assertion
       std::string saml_assertion;
       assertion_nd.GetXML(saml_assertion);
-      logger.msg(Arc::VERBOSE,"Encrypted saml assertion: %s", saml_assertion.c_str());
+      logger.msg(Arc::DEBUG,"Encrypted saml assertion: %s", saml_assertion.c_str());
 
       Arc::XMLSecNode sec_assertion_nd(assertion_nd);
       Arc::XMLNode decrypted_assertion_nd;
@@ -332,13 +332,13 @@ Arc::MCC_Status Service_SP::process(Arc::Message& inmsg,Arc::Message& outmsg) {
 
       std::string decrypted_saml_assertion;
       decrypted_assertion_nd.GetXML(decrypted_saml_assertion);
[…]

### 2009-11-11 src/services/saml/aaservice.cpp

@@ -420,7 +420,7 @@ bool Service_AA::query_db(Arc::QueryArrayResult& attributes, const std::string& 
   portstr = (std::string)(nd.Attribute("port"));
   port = atoi((portstr.c_str()));
 
-  logger_.msg(Arc::DEBUG, "Access database %s from server %s port %s, with user %s and password %s",
+  logger_.msg(Arc::VERBOSE, "Access database %s from server %s port %s, with user %s and password %s",
               dbname.c_str(), server.c_str(), portstr.c_str(), user.c_str(), password.c_str()); 
 
   //TODO: make the database and sql object dynamic loaded 
@@ -433,7 +433,7 @@ bool Service_AA::query_db(Arc::QueryArrayResult& attributes, const std::string& 
   }
 
   Arc::MySQLQuery myquery(&mydb);
-    logger_.msg(Arc::DEBUG, "Is connected to database? %s", mydb.isconnected()? "yes":"no");
+    logger_.msg(Arc::VERBOSE, "Is connected to database? %s", mydb.isconnected()? "yes":"no");
 
   std::string querystr;
   for(int i = 0;; i++) {
@@ -444,12 +444,12 @@ bool Service_AA::query_db(Arc::QueryArrayResult& attributes, const std::string& 
         Arc::XMLNode scn = cn["aa:SQL"][k];
         if(!scn) break;
         querystr = (std::string)scn;
-        logger_.msg(Arc::DEBUG, "Query: %s", querystr.c_str());
+        logger_.msg(Arc::VERBOSE, "Query: %s", querystr.c_str());
         myquery.get_array(querystr, attributes, sqlargs);
       }
     }
   }
-  logger_.msg(Arc::DEBUG, "Get result array with %d rows",attributes.size());
+  logger_.msg(Arc::VERBOSE, "Get result array with %d rows",attributes.size());
   return true;
 }
 

### 2009-11-11 src/services/sched/get_activity_documents.cpp

@@ -22,7 +22,7 @@ Arc::MCC_Status GridSchedulerService::GetActivityDocuments(Arc::XMLNode &in,Arc:
   {
     std::string s;
     in.GetXML(s);
-    logger.msg(Arc::DEBUG, "GetActivityDocuments: request = \n%s", s);
+    logger.msg(Arc::VERBOSE, "GetActivityDocuments: request = \n%s", s);
   };
   for (int n = 0;;++n) {
     Arc::XMLNode id = in["ActivityIdentifier"][n];
@@ -58,7 +58,7 @@ Arc::MCC_Status GridSchedulerService::GetActivityDocuments(Arc::XMLNode &in,Arc:
   {
     std::string s;
     out.GetXML(s);
-    logger_.msg(Arc::DEBUG, "GetActivityDocuments: response = \n%s", s);
+    logger_.msg(Arc::VERBOSE, "GetActivityDocuments: response = \n%s", s);
   };
   return Arc::MCC_Status(Arc::STATUS_OK);
 }

### 2009-11-11 src/services/storage/librarian/librarian.py

@@ -34,7 +34,7 @@ class Librarian:
         ahash_urls =  get_child_values_by_name(cfg, 'AHashURL')
         if ahash_urls:
             log.msg(arc.INFO,'Got AHash URLs:', ahash_urls)
-            log.msg(arc.DEBUG, 'AHash URL found in the configuration.')
+            log.msg(arc.VERBOSE, 'AHash URL found in the configuration.')
             self.master_ahash = arc.URL(ahash_urls[0])
             self.ahash_reader = AHashClient(ahash_urls, ssl_config = self.ssl_config)
             self.ahash_writer = AHashClient(ahash_urls, ssl_config = self.ssl_config)
@@ -207,7 +207,7 @@ class Librarian:
         ahash_response = self.ahash_change(ahash_request)
         #print '_set_next_heartbeat response', ahash_response
         if ahash_response['report'][0] != 'set':
-            log.msg(arc.DEBUG, 'ERROR setting next heartbeat time!')
+            log.msg(arc.VERBOSE, 'ERROR setting next heartbeat time!')
     
     def report(self, serviceID, filelist):
         try:

### 2009-11-11 src/services/storage/shepherd/shepherd.py

@@ -160,7 +160,7 @@ class Shepherd:
                         # get its local data (GUID, size, state, etc.)
                         localData = self.store.get(changed)
                         if not localData.has_key('GUID'):
-                            log.msg(arc.DEBUG, 'Error in shepherd.reportingThread()\n\treferenceID is in changed_states, but not in store')
+                            log.msg(arc.VERBOSE, 'Error in shepherd.reportingThread()\n\treferenceID is in changed_states, but not in store')
                         else:
                             # add to the filelist the GUID, the referenceID and the state of the file
                             filelist.append((localData.get('GUID'), changed, localData.get('state')))
@@ -169,7 +169,7 @@ class Shepherd:
                     try:
                         next_report = self.librarian.report(self.serviceID, filelist)
                     except:
-                        log.msg(arc.DEBUG, 'Error sending report message to the Librarian, reason:', traceback.format_exc())
+                        log.msg(arc.VERBOSE, 'Error sending report message to the Librarian, reason:', traceback.format_exc())
                         # if next_report is below zero, then we will send everything again
                         next_report = -1
                     # we should get the time of the next report
@@ -179,7 +179,7 @@ class Shepherd:
                     last_report = time.time()
                     # if the next report time is below zero it means:
                     if next_report < 0: # 'please send all'
-                        log.msg(arc.DEBUG, 'reporting - asked to send all file data again')
+                        log.msg(arc.VERBOSE, 'reporting - asked to send all file data again')
                         # add the full list of stored files to the changed_state list - all the files will be reported next time (which is immediately, see below)
                         self.changed_states.extend(self.store.list())
                     # let's wait until there is any changed file or the reporting time is up - we need to do report even if no file changed (as a heartbeat)
@@ -228,7 +228,7 @@ class Shepherd:
             # if the original and the current checksum is the same, then the replica is valid
             if state in [INVALID, CREATING, STALLED]:
                 # if it is currently INVALID or CREATING or STALLED, its state should be changed
-                log.msg(arc.DEBUG, '\nCHECKSUM OK', referenceID)
+                log.msg(arc.VERBOSE, '\nCHECKSUM OK', referenceID)
                 self.changeState(referenceID, ALIVE)
                 state = ALIVE
             return state
@@ -248,7 +248,7 @@ class Shepherd:
                 return state
             if state != INVALID:
                 # but if it is not CREATING, not DELETED or STALLED and not INVALID - then it was ALIVE: now its state should be changed to INVALID
[…]

### 2009-11-11 src/tests/echo/echo.cpp

@@ -95,7 +95,7 @@ Arc::MCC_Status Service_Echo::process(Arc::Message& inmsg,Arc::Message& outmsg) 
   {
       std::string str;
       inpayload->GetDoc(str, true);
-      logger.msg(Arc::DEBUG, "process: request=%s",str);
+      logger.msg(Arc::VERBOSE, "process: request=%s",str);
   }; 
 
   Arc::PayloadSOAP* outpayload = NULL;
@@ -163,7 +163,7 @@ Arc::MCC_Status Service_Echo::process(Arc::Message& inmsg,Arc::Message& outmsg) 
   {
       std::string str;
       outpayload->GetDoc(str, true);
-      logger.msg(Arc::DEBUG, "process: response=%s",str);
+      logger.msg(Arc::VERBOSE, "process: response=%s",str);
   }; 
   return Arc::MCC_Status(Arc::STATUS_OK);
 }

### 2009-11-11 src/tests/perf/perftest_deleg_bysechandler.cpp

@@ -188,7 +188,7 @@ int main(int argc, char* argv[]){
               << "         'soap' entry point and HOSTNAME, PORTNUMBER and PATH " << std::endl
               << "         keyword for hostname, port and HTTP path of 'echo' service." << std::endl
 	      << "debug    The textual representation of desired debug level. Available " << std::endl
-              << "         levels: VERBOSE, DEBUG, INFO, WARNING, ERROR, FATAL." << std::endl;
+              << "         levels: DEBUG, VERBOSE, INFO, WARNING, ERROR, FATAL." << std::endl;
     exit(EXIT_FAILURE);
   }
   url_str = std::string(argv[1]);

### 2009-11-11 src/tests/perf/perftest_msgsize.cpp

@@ -198,7 +198,7 @@ int main(int argc, char* argv[]){
               << "           'soap' entry point and HOSTNAME, PORTNUMBER and PATH " << std::endl
               << "            keyword for hostname, port and HTTP path of 'echo' service." << std::endl
               << "-d debug   The textual representation of desired debug level. Available " << std::endl
-              << "            levels: VERBOSE, DEBUG, INFO, WARNING, ERROR, FATAL." << std::endl
+              << "            levels: DEBUG, VERBOSE, INFO, WARNING, ERROR, FATAL." << std::endl
               << "-r         If specified close connection and reconnect after " << std::endl
               << "            every request." << std::endl
               << "-t         Toggles TCP_NODELAY option " << std::endl

### 2009-11-13 doc/manuals/chelonia/chelonia-cli.tex
Finalizing the new UI manual (ref. nr. NORDUGRID-MANUAL-13)

@@
    \texttt{modify, mod} &\textit{string}& modify metadata\\
-   \texttt{policy, pol} &\textit{string}& modify access policy rules. The string has a form $<$LN$>$ $<$changeType$>$ $<$identity$>$ $<$action list$>$.\\
-   & &$<$changeType$>$ could be `set', `change' or `clear'\\
-   & & `set': sets the action list to the given user overwriting the old one\\
-   & & `change': modify the current action list with adding and removing actions\\
-   & & `clear': clear the action list of the given user\\
-   & &$<$identity$>$ could be a `$<$UserDN$>$' or a `VOMS:$<$VO name$>$'\\
-   & &$<$action list$>$ is a list actions prefixed with `+' or `-', e.g.
-`+read +addEntry -delete'; possible actions are: \texttt{read}, \texttt{addEntry}, \texttt{removeEntry}, \texttt{delete}, \texttt{modifyPolicy}, \texttt{modifyStates}, \texttt{modifyMetadata}\\
+   \texttt{policy, pol} &\textit{string}& modify access policy rules\\
    \texttt{unlink} &\textit{string}& remove a link to an entry from a collection without removing the entry itself\\
@@
 
-Without arguments, each method prints its own help.
+Without arguments, each method prints its own help. Detailed explanation of each method is given below.
 
@@
 \label{ssub:stat}
-With the \texttt{stat} method it is possible to get all the metadata about one or more entry (file, collection, etc.). The entries are specified with their Logical Name (LN).
+With the \texttt{stat}\index{commands:chelonia:stat} method it is possible to get all the metadata about one or more entry (file, collection, etc.).
+The entries are specified with their Logical Name (LN).
 \hspace*{0.5cm}
@@
 \label{ssub:makecollection}
-With the \texttt{makeCollection} or \texttt{mkdir} method it is possible to create a new empty collection. The requested Logical Name (LN) should be specified.
+With the \texttt{makeCollection}\index{commands:chelonia:makeCollection} or \texttt{mkdir} method it is possible to create a new empty collection. The requested Logical Name (LN) should be specified.
 \hspace*{0.5cm}
@@
 \label{ssub:unmakecollection}
-With the \texttt{unmakeCollection} or \texttt{rmdir} method it is possible to delete an empty collection which is specified by its Logical Name (LN).
+With the \texttt{unmakeCollection}\index{commands:chelonia:unmakeCollection} or \texttt{rmdir} method it is possible to delete an empty collection which is specified by its Logical Name (LN).
 \hspace*{0.5cm}
@@
 \label{ssub:list}
-With the \texttt{list} or \texttt{ls} method it is possible to list the contents of one or more collections which are specified by their Logical Name (LN).
+With the \texttt{list}\index{commands:chelonia:list} or \texttt{ls} method it is possible to list the contents of one or more collections which are specified by their Logical Name (LN).
 \hspace*{0.5cm}
@@
 \label{ssub:move}
-With the \texttt{move} or \texttt{mv} method it is possible to move a file or collection within the namespace of chelonia (including renaming the entry).
+With the \texttt{move}\index{commands:chelonia:move} or \texttt{mv} method it is possible to move a file or collection within the namespace of chelonia (including renaming the entry).
 The source path and the target path should be specified as Logical Names
@@
 \label{ssub:putfile}
-With the \texttt{putFile} or \texttt{put} method it is possible to upload a new file into the system creating a new Logical Name (LN). It is possible the specify the number of needed replicas.
+With the \texttt{putFile}\index{commands:chelonia:putFile} or \texttt{put} method it is possible to upload a new file into the system creating a new Logical Name (LN).
+It is possible the specify the number of needed replicas.
 \hspace*{0.5cm}
@@
 \label{ssub:getfile}
-With the \texttt{getFile} or \texttt{get} method it is possible to download a file specified with its Logical Name (LN).
+With the \texttt{getFile}\index{commands:chelonia:getFile} or \texttt{get} method it is possible to download a file specified with its Logical Name (LN).
 If the target local path is not given, then the file will be put into the local directory.
@@
 \label{ssub:delfile}
-With the \texttt{delFile} or \texttt{rm} method it is possible to delete one or more files from the system.
+With the \texttt{delFile}\index{commands:chelonia:delFile} or \texttt{rm} method it is possible to delete one or more files from the system.
 \hspace*{0.5cm}
@@
 
-With the \texttt{modify} or \texttt{mod} method it is possible to modify some metadata of an entry.
+With the \texttt{modify}\index{commands:chelonia:modify} or \texttt{mod} method it is possible to modify some metadata of an entry.
 \hspace*{0.5cm}
@@
 
-With the \texttt{policy} or \texttt{pol} method it is possible to modify the policy of the entry
+With the \texttt{policy}\index{commands:chelonia:policy} or \texttt{pol} method it is possible to modify the policy of the entry
 \hspace*{0.5cm}
@@
 
-With the \texttt{unlink} method it is possible to remove a file or collection just from its parent collection without removing the file or collection itself.
+With the \texttt{unlink}\index{commands:chelonia:unlink} method it is possible to remove a file or collection just from its parent collection without removing the file or collection itself.
 \hspace*{0.5cm}
@@
 \label{ssub:credentialdelegation}
-With the \texttt{credentialDelegation} or \texttt{cre} method it is possible to delegate credentials to the Bartender.
+With the \texttt{credentialDelegation}\index{commands:chelonia:credentialDelegation} or \texttt{cre} method it is possible to delegate credentials to the Bartender.
 \hspace*{0.5cm}
@@
 \label{ssub:removecredentials}
-With the \texttt{removeCredentials} or \texttt{rem} method it is possible to remove the previously delegated credentials.
+With the \texttt{removeCredentials}\index{commands:chelonia:removeCredentials} or \texttt{rem} method it is possible to remove the previously delegated credentials.
 \hspace*{0.5cm}
@@
 \label{ssub:makemountpoint}
-With the \texttt{makeMountPoint} or \texttt{makemount} method it is possible to create a mount point within the namespace of Chelonia which points to a GridFTP server.
+With the \texttt{makeMountPoint}\index{commands:chelonia:makeMountPoint} or \texttt{makemount} method it is possible to create a mount point within the namespace of Chelonia which points to a GridFTP server.
 \hspace*{0.5cm}
@@
 \label{ssub:unmakemountpoint}
-With the \texttt{unmakeMountPoint} or \texttt{unmount} method it is possible to remove a previously created mount point.
+With the \texttt{unmakeMountPoint}\index{commands:chelonia:unmakeMountPoint} or \texttt{unmount} method it is possible to remove a previously created mount point.
 \hspace*{0.5cm}

### 2009-11-13 doc/manuals/user_interface/ui-data.tex
Finalizing the new UI manual (ref. nr. NORDUGRID-MANUAL-13)

@@
     Options:&&\\
-    \texttt{-h} && short help\\
-    \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
-    \texttt{-l} &  & detailed listing\\
-    \texttt{-L} &  & detailed listing including URLs from which file can 
-    be downloaded\\
-    \texttt{-m} && display all available metadata\\
+   \texttt{-l, --long} &  & detailed listing\\
+   \texttt{-L, --locations} &  & detailed listing including URLs from which files can be downloaded\\
+   \texttt{-m, --metadata} && display all available metadata\\
+   \texttt{-r, --recursive} & \textit{recursion\_level} & operate recursively (if possible) up to specified level (0 - no recursion)\\
+   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG - default WARNING\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}&&print version information\\
+   \texttt{-h, --help}&&print help page\\
     Arguments:&&\\
@@
     Options:&&\\
-    \texttt{-h} && short help\\
-    \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
-    \texttt{-y} & \textit{cache\_path} & path to local cache (use to put file into cache)\\
-    \texttt{-p} && use passive transfer (does not work if secure is on, default if secure is not requested)\\
-    \texttt{-n} && do not try to force passive transfer\\
-    \texttt{-i} && show progress indicator\\
-    \texttt{-u} && use secure transfer (insecure by default)\\
-    \texttt{-r} & \textit{recursion\_level} & operate recursively (if possible) up to specified level (0 - no recursion)\\
-    \texttt{-R} & \textit{number} & how many times to retry transfer of every file before failing\\
-    \texttt{-t} & \textit{time} & timeout in seconds (default 20)\\
-    \texttt{-f} && if the destination is an indexing service and not the same as the source and the destination is already registered, then the copy is normally not done.
+   \texttt{-p, --passive} && use passive transfer (does not work if secure is on, default if secure is not requested)\\
+   \texttt{-n, --nopassive} && do not try to force passive transfer\\
+   \texttt{-f, --force} && if the destination is an indexing service and not the same as the source and the destination is already registered, then the copy is normally not done.
 However, if this option is specified the source is assumed to be a replica of the destination created in an uncontrolled way and the copy is done like in case of replication\\
-    \texttt{-T} && do not transfer file, just register it - destination must be non-existing meta-url\\
+   \texttt{-i, --indicate} && show progress indicator\\
+   \texttt{-T, --notransfer} && do not transfer file, just register it - destination must be non-existing meta-url\\
+   \texttt{-u, --secure} && use secure transfer (insecure by default)\\
+   \texttt{-y, --cache} &\textit{path} & path to local cache (use to put file into cache). The \texttt{X509\_USER\_PROXY} and \texttt{X509\_CERT\_DIR} environment variables must be set correctly\\
+   \texttt{-r, --recursive} & \textit{recursion\_level} & operate recursively (if possible) up to specified level (0 - no recursion)\\
+   \texttt{-R, --retries} & \textit{number} & how many times to retry transfer of every file before failing\\
+   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG - default WARNING\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}&&print version information\\
+   \texttt{-h, --help}&&print help page\\
     Arguments:&&\\
@@
     Options:&&\\
-    \texttt{-h} && short help\\
-    \texttt{-v} && print version information\\
-    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
-    \texttt{-c} & &continue with meta-data even if it failed to delete real file\\
+   \texttt{-f, --force} & &remove logical file name registration even if not all physical instances were removed\\
+   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG - default WARNING\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}&&print version information\\
+   \texttt{-h, --help}&&print help page\\
     Arguments:&&\\
@@
 
-\subsection{arcacl}\label{sec:arcacl}
+% \subsection{arcacl}\label{sec:arcacl}
+% 
+% \index{arcacl}\index{commands:arcacl}This command retrieves or modifies
+% access control information associated with a stored object if service
+% supports GridSite GACL language~\cite{gacl} for access control.
+% \hspace*{0.5cm}
+% \begin{shaded}
+%    \uicommand{arcacl [options] get$|$put $<$URL$>$}
+% \end{shaded}
+% \versions{ARC 0.9}
+% \begin{longtable}{llp{8cm}}
+%    Options:&&\\
+%     \texttt{-d, -debug} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
+%     \texttt{-v} && print version information\\
+%     \texttt{-h} && short help\\
+%    Arguments:&&\\
+%     \texttt{get} &\textit{URL}& get Grid ACL for the object\\
+%     \texttt{put} &\textit{URL}& set Grid ACL for the object\\
+%     \texttt{URL} && object URL; curently only gsiftp and sse URLs are supported\\
+% \end{longtable}
+% 
+% The ACL document (an XML file) is printed to standard output when
+% \texttt{get} is requested, and is acquired from standard input when
+% \texttt{set} is specified\footnote{In ARC $\leq$ 0.5.28, \texttt{set}
+%   was used instead of \texttt{put}}. Usage examples are:
+% \begin{verbatim}
+%     arcacl get gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename
+%     arcacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename < myacl
+% \end{verbatim}
 
-\index{arcacl}\index{commands:arcacl}This command retrieves or modifies
-access control information associated with a stored object if service
-supports GridSite GACL language~\cite{gacl} for access control.
-\hspace*{0.5cm}
-\begin{shaded}
-   \uicommand{arcacl [options] get$|$put $<$URL$>$}
-\end{shaded}
-\versions{ARC 0.9}
-\begin{longtable}{llp{8cm}}
-   Options:&&\\
-    \texttt{-d, -debug} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
-    \texttt{-v} && print version information\\
-    \texttt{-h} && short help\\
-   Arguments:&&\\
-    \texttt{get} &\textit{URL}& get Grid ACL for the object\\
-    \texttt{put} &\textit{URL}& set Grid ACL for the object\\
-    \texttt{URL} && object URL; curently only gsiftp and sse URLs are supported\\
-\end{longtable}
-
-The ACL document (an XML file) is printed to standard output when
-\texttt{get} is requested, and is acquired from standard input when
-\texttt{set} is specified\footnote{In ARC $\leq$ 0.5.28, \texttt{set}
-  was used instead of \texttt{put}}. Usage examples are:
-\begin{verbatim}
-    arcacl get gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename
-    arcacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename < myacl
-\end{verbatim}
-
-\subsection{arctransfer}\label{sec:arctransfer}
-\index{arctransfer}\index{commands:arctransfer}
-The \texttt{arctransfer} command is not implemented.
+% \subsection{arctransfer}\label{sec:arctransfer}
+% \index{arctransfer}\index{commands:arctransfer}
+% The \texttt{arctransfer} command is not implemented.

### 2009-11-13 doc/manuals/user_interface/ui-job.tex
Finalizing the new UI manual (ref. nr. NORDUGRID-MANUAL-13)

@@
    \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
-   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site (cluster)\\
    \texttt{-s, --status}& \textit{statusstr} &only select jobs whose status is \textit{statusstr}\\
@@
    \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
-   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific source site\\
-   \texttt{-q, --qluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site as re-submission target\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific source site\\
+   \texttt{-q, --qluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site as re-submission target\\
    \texttt{-m, --same}& &re-submit to the same site\\
@@
    \texttt{-j, --joblist}& \textit{filename} & file containing a list of jobIDs\\
-   \texttt{-c, --cluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site (cluster)\\
-   \texttt{-q, --qluster}&\verb#[-]#textem{name}&explicitly select or reject a specific site as re-submission target\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site (cluster)\\
+   \texttt{-q, --qluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site as re-submission target\\
    \texttt{-f, --forcemigration}& & force migration, ignoring kill failure\\

### 2009-11-13 doc/manuals/echo/echo-manual.tex
very-very basic manual about hopi and echo - what else should go in these?

\documentclass{article}
%\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {The Hopi service},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-MANUAL-XXX\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large The Echo Service - The Missing Manual}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large First prototype status and plans}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large The Unknown Knowarcer\footnote{nordugrid-discuss@nordugrid.org} \large \par}
\end{titlepage}

\tableofcontents                          %Comment if use article style

\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}

%\chapter{Quick start guide} % (fold)

\section{Introduction} % (fold)
\label{sec:introduction}
The capability of the echo service is to accept SOAP messages like this:
\begin{verbatim}
<?xml version="1.0"?>
<soap-env:Envelope
	xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:echo="urn:echo">
    <soap-env:Body>
        <echo:echo>
            <echo:say>HELLO</echo:say>
        </echo:echo>
    </soap-env:Body>
</soap-env:Envelope>
\end{verbatim}
Then the Echo service adds a prefix and a suffix to the message and sends it back like this:
\begin{verbatim}
<?xml version="1.0"?> 
<soap-env:Envelope
	xmlns:echo="urn:echo" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/">
    <soap-env:Body>
        <echo:echoResponse>
            <echo:hear>hi!}}</echo:hear>
        </echo:echoResponse>
    </soap-env:Body>
</soap-env:Envelope>
\end{verbatim}
There are a C++ and a Python implementation of the Echo service, and there are configuration profiles for secure and non-secure deployments for both languages.
% section introduction (end)

\section{Clients} % (fold)
\label{sec:clients}
Currently in the nordugrid subversion there are at least two readily available client to test the Echo service:
\begin{itemize}
	\item \textbf{echo\_client.py}\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/tests/echo/echo_client.py}} is a python script which gets the credentials from the userconfig, and sends a single message to the given URL, then prints the reply
	\item \textbf{perftest}\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/tests/perf/perftest.cpp}} sends as many messages as possible within a given number of seconds using a given number of threads
\end{itemize}
% section clients (end)

\section{Configuration of the C++ Echo service} % (fold)
\label{sec:configuration}
Here is an example configuration of a secure C++ Echo service:

\begin{verbatim}
<?xml version="1.0"?>
<cfg:ArcConfig xmlns="http://www.nordugrid.org/schemas/loader/2009/08"
	xmlns:cfg="http://www.nordugrid.org/schemas/arcconfig/2009/08"
	xmlns:tcp="http://www.nordugrid.org/schemas/tcp/2009/08"
	xmlns:tls="http://www.nordugrid.org/schemas/tls/2009/08"
	xmlns:echo="http://www.nordugrid.org/schemas/echo/2009/08">
    <cfg:Server>
        <cfg:PidFile>tmp/arched.pid</cfg:PidFile>
        <cfg:Logger>
            <cfg:File>/var/log/arched.log</cfg:File>
            <cfg:Level>ERROR</cfg:Level>
        </cfg:Logger>
    </cfg:Server>
    <ModuleManager>
        <Path inisections="common" initag="libpath">/usr/local/lib/arc/</Path>
    </ModuleManager>
    <Plugins>
        <Name>mcctls</Name>
        <Name>mcchttp</Name>
        <Name>mccsoap</Name>
        <Name>mcctcp</Name>
    </Plugins>
    <Chain>
        <Component name="tcp.service" id="tcp">
            <next id="tls"/>
            <tcp:Listen>
               <tcp:Interface>0.0.0.0</tcp:Interface>
               <tcp:Port>50000</tcp:Port>
               <tcp:Version>4</tcp:Version>
            </tcp:Listen>
        </Component>
        <Component name="tls.service" id="tls">
            <next id="http"/>
            <tls:KeyPath>/etc/grid-security/hostkey.pem</tls:KeyPath>
            <tls:CertificatePath>/etc/grid-security/hostcert.pem</tls:CertificatePath>
            <tls:CACertificatesDir>/etc/grid-security/certificates</tls:CACertificatesDir>
        </Component>
        <Component name="http.service" id="http">
            <next id="soap">POST</next>
            <next id="plexer">GET</next>
            <next id="plexer">PUT</next>
        </Component>
        <Component name="soap.service" id="soap">
            <next id="plexer"/>
        </Component>
        <Plexer name="plexer.service" id="plexer">
            <next id="echo">^/Echo$</next>
        </Plexer>
        <Service name="echo" id="echo">
            <echo:prefix>[</echo:prefix>
            <echo:suffix>]</echo:suffix>
        </Service>
    </Chain>
</cfg:ArcConfig>
\end{verbatim}
% section configuration (end)

\section{Configuration of the Python Echo service} % (fold)
\label{sec:configuration_of_the_python_echo_service}
Here is an example configuration of a secure python Echo service:

\begin{verbatim}
<?xml version="1.0"?>
<cfg:ArcConfig xmlns="http://www.nordugrid.org/schemas/loader/2009/08"
	xmlns:cfg="http://www.nordugrid.org/schemas/arcconfig/2009/08"
	xmlns:tcp="http://www.nordugrid.org/schemas/tcp/2009/08"
	xmlns:tls="http://www.nordugrid.org/schemas/tls/2009/08"
	xmlns:echo="http://www.nordugrid.org/schemas/echo/2009/08"
	xmlns:py="http://www.nordugrid.org/schemas/pythonwrapper/2009/08">
    <cfg:Server>
        <cfg:PidFile>tmp/arched.pid</cfg:PidFile>
        <cfg:Logger>
            <cfg:File>/var/log/arched.log</cfg:File>
            <cfg:Level>ERROR</cfg:Level>
        </cfg:Logger>
    </cfg:Server>
    <ModuleManager>
        <Path inisections="common" initag="libpath">/usr/local/lib/arc/</Path>
    </ModuleManager>
    <Plugins>
        <Name>mcctls</Name>
        <Name>mcchttp</Name>
        <Name>mccsoap</Name>
        <Name>mcctcp</Name>
    </Plugins>
    <Chain>
        <Component name="tcp.service" id="tcp">
            <next id="tls"/>
            <tcp:Listen>
               <tcp:Interface>0.0.0.0</tcp:Interface>
               <tcp:Port>50000</tcp:Port>
               <tcp:Version>4</tcp:Version>
            </tcp:Listen>
        </Component>
        <Component name="tls.service" id="tls">
            <next id="http"/>
            <tls:KeyPath>/etc/grid-security/hostkey.pem</tls:KeyPath>
            <tls:CertificatePath>/etc/grid-security/hostcert.pem</tls:CertificatePath>
            <tls:CACertificatesDir>/etc/grid-security/certificates</tls:CACertificatesDir>
        </Component>
        <Component name="http.service" id="http">
            <next id="soap">POST</next>
            <next id="plexer">GET</next>
            <next id="plexer">PUT</next>
        </Component>
        <Component name="soap.service" id="soap">
            <next id="plexer"/>
        </Component>
        <Plexer name="plexer.service" id="plexer">
            <next id="echo">^/Echo$</next>
        </Plexer>
        <Service name="pythonservice" id="echo">
            <py:ClassName>echo_python.EchoService.EchoService</py:ClassName>
            <echo:prefix>[</echo:prefix>
            <echo:suffix>]</echo:suffix>
        </Service>
    </Chain>
</cfg:ArcConfig>
\end{verbatim}
% section configuration_of_the_python_echo_service (end)

\end{document}

### 2009-11-13 doc/manuals/hopi/hopi-manual.tex
very-very basic manual about hopi and echo - what else should go in these?

--- 2009-11-13 doc/manuals/echo/echo-manual.tex
@@
 %%%%---- The title ----
-{\centering \textsc{\Large The Echo Service - The Missing Manual}\Large \par}
+{\centering \textsc{\Large The Hopi manual}\Large \par}
 \vspace*{0.5cm}
@@
 \label{sec:introduction}
-The capability of the echo service is to accept SOAP messages like this:
-\begin{verbatim}
-<?xml version="1.0"?>
-<soap-env:Envelope
-	xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:echo="urn:echo">
-    <soap-env:Body>
-        <echo:echo>
-            <echo:say>HELLO</echo:say>
-        </echo:echo>
-    </soap-env:Body>
-</soap-env:Envelope>
-\end{verbatim}
-Then the Echo service adds a prefix and a suffix to the message and sends it back like this:
-\begin{verbatim}
-<?xml version="1.0"?> 
-<soap-env:Envelope
-	xmlns:echo="urn:echo" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/">
-    <soap-env:Body>
-        <echo:echoResponse>
-            <echo:hear>hi!}}</echo:hear>
-        </echo:echoResponse>
-    </soap-env:Body>
-</soap-env:Envelope>
-\end{verbatim}
-There are a C++ and a Python implementation of the Echo service, and there are configuration profiles for secure and non-secure deployments for both languages.
+The Hopi service is a HTTP(S) server supporting GET and PUT methods and capable of generating a HTML listing of files. It's main configuration parameter is the document root directory.
+
+It also has a slave mode which is used in the Chelonia storage system tied to the Shepherd service.
+In the slave mode: 1.
+there is no HTML listing generated; 2.
+the files are removed after download; 3.
+upload is only possible to an existing file, which will be removed after the upload.
+(The Shepherd service keeps a second hardlink of each file in a separate directory, so that's why it makes sense to remove files after transfer.)
+
+The Hopi service can be the only service in a message chain, in that case it does not need to have a Plexer, and the URLs will not contain any extra path; or the Hopi could be after the Plexer, in that case all the URLs contains a fixed path (e.g.
+\verb!https://localhost:60000/hopi/!).
 % section introduction (end)
@@
 \label{sec:clients}
-Currently in the nordugrid subversion there are at least two readily available client to test the Echo service:
+Because Hopi is a HTTP(S) server, any HTTP client can be used (for HTTPS maybe needs more configuration)
 \begin{itemize}
-	\item \textbf{echo\_client.py}\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/tests/echo/echo_client.py}} is a python script which gets the credentials from the userconfig, and sends a single message to the given URL, then prints the reply
-	\item \textbf{perftest}\footnote{\url{http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/tests/perf/perftest.cpp}} sends as many messages as possible within a given number of seconds using a given number of threads
+	\item web browser (needs importing the client credentials into the browser)
+	\item curl/wget (needs specifying the client credentials as command line argument)
+	\item arccp (the client credentials comes from the userconfig)
 \end{itemize}
@@
 
-\section{Configuration of the C++ Echo service} % (fold)
+\section{Configuration} % (fold)
 \label{sec:configuration}
-Here is an example configuration of a secure C++ Echo service:
+Here is an example configuration of a secure Hopi with Plexer:
 
@@
 	xmlns:tcp="http://www.nordugrid.org/schemas/tcp/2009/08"
-	xmlns:tls="http://www.nordugrid.org/schemas/tls/2009/08"
-	xmlns:echo="http://www.nordugrid.org/schemas/echo/2009/08">
+	xmlns:tls="http://www.nordugrid.org/schemas/tls/2009/08">
     <cfg:Server>
@@
     <ModuleManager>
-        <Path inisections="common" initag="libpath">/usr/local/lib/arc/</Path>
+        <Path>/usr/local/lib/arc/</Path>
     </ModuleManager>
@@
         <Plexer name="plexer.service" id="plexer">
-            <next id="echo">^/Echo$</next>
+            <next id="hopi">^/hopi/</next>
         </Plexer>
-        <Service name="echo" id="echo">
-            <echo:prefix>[</echo:prefix>
-            <echo:suffix>]</echo:suffix>
+        <Service name="hopi" id="hopi">
+            <DocumentRoot>/tmp/hopi</DocumentRoot>
+            <SlaveMode>0</SlaveMode>
         </Service>
@@
 
-\section{Configuration of the Python Echo service} % (fold)
-\label{sec:configuration_of_the_python_echo_service}
-Here is an example configuration of a secure python Echo service:
-
-\begin{verbatim}
-<?xml version="1.0"?>
-<cfg:ArcConfig xmlns="http://www.nordugrid.org/schemas/loader/2009/08"
-	xmlns:cfg="http://www.nordugrid.org/schemas/arcconfig/2009/08"
-	xmlns:tcp="http://www.nordugrid.org/schemas/tcp/2009/08"
-	xmlns:tls="http://www.nordugrid.org/schemas/tls/2009/08"
-	xmlns:echo="http://www.nordugrid.org/schemas/echo/2009/08"
-	xmlns:py="http://www.nordugrid.org/schemas/pythonwrapper/2009/08">
-    <cfg:Server>
-        <cfg:PidFile>tmp/arched.pid</cfg:PidFile>
-        <cfg:Logger>
-            <cfg:File>/var/log/arched.log</cfg:File>
-            <cfg:Level>ERROR</cfg:Level>
-        </cfg:Logger>
-    </cfg:Server>
-    <ModuleManager>
-        <Path inisections="common" initag="libpath">/usr/local/lib/arc/</Path>
-    </ModuleManager>
-    <Plugins>
-        <Name>mcctls</Name>
-        <Name>mcchttp</Name>
-        <Name>mccsoap</Name>
-        <Name>mcctcp</Name>
-    </Plugins>
-    <Chain>
-        <Component name="tcp.service" id="tcp">
-            <next id="tls"/>
-            <tcp:Listen>
-               <tcp:Interface>0.0.0.0</tcp:Interface>
-               <tcp:Port>50000</tcp:Port>
-               <tcp:Version>4</tcp:Version>
-            </tcp:Listen>
-        </Component>
-        <Component name="tls.service" id="tls">
-            <next id="http"/>
-            <tls:KeyPath>/etc/grid-security/hostkey.pem</tls:KeyPath>
-            <tls:CertificatePath>/etc/grid-security/hostcert.pem</tls:CertificatePath>
-            <tls:CACertificatesDir>/etc/grid-security/certificates</tls:CACertificatesDir>
-        </Component>
-        <Component name="http.service" id="http">
-            <next id="soap">POST</next>
-            <next id="plexer">GET</next>
-            <next id="plexer">PUT</next>
-        </Component>
-        <Component name="soap.service" id="soap">
-            <next id="plexer"/>
-        </Component>
-        <Plexer name="plexer.service" id="plexer">
-            <next id="echo">^/Echo$</next>
-        </Plexer>
-        <Service name="pythonservice" id="echo">
-            <py:ClassName>echo_python.EchoService.EchoService</py:ClassName>
-            <echo:prefix>[</echo:prefix>
-            <echo:suffix>]</echo:suffix>
-        </Service>
-    </Chain>
-</cfg:ArcConfig>
-\end{verbatim}
-% section configuration_of_the_python_echo_service (end)
-
 \end{document}

### 2009-11-13 doc/manuals/chelonia/chelonia-cli.tex
added missing arcsrmping, some more index entires

@@
 \label{ssub:stat}
-With the \texttt{stat}\index{commands:chelonia:stat} method it is possible to get all the metadata about one or more entry (file, collection, etc.).
+With the \texttt{stat}\index{commands:chelonia:stat}\index{chelonia:stat}  method it is possible to get all the metadata about one or more entry (file, collection, etc.).
 The entries are specified with their Logical Name (LN).
@@
 \label{ssub:makecollection}
-With the \texttt{makeCollection}\index{commands:chelonia:makeCollection} or \texttt{mkdir} method it is possible to create a new empty collection. The requested Logical Name (LN) should be specified.
+With the \texttt{makeCollection}\index{commands:chelonia:makeCollection}\index{chelonia:makeCollection} or \texttt{mkdir} method it is possible to create a new empty collection.
+The requested Logical Name (LN) should be specified.
 \hspace*{0.5cm}
@@
 \label{ssub:unmakecollection}
-With the \texttt{unmakeCollection}\index{commands:chelonia:unmakeCollection} or \texttt{rmdir} method it is possible to delete an empty collection which is specified by its Logical Name (LN).
+With the \texttt{unmakeCollection}\index{commands:chelonia:unmakeCollection}\index{chelonia:unmakeCollection} or \texttt{rmdir} method it is possible to delete an empty collection which is specified by its Logical Name (LN).
 \hspace*{0.5cm}
@@
 \label{ssub:list}
-With the \texttt{list}\index{commands:chelonia:list} or \texttt{ls} method it is possible to list the contents of one or more collections which are specified by their Logical Name (LN).
+With the \texttt{list}\index{commands:chelonia:list}\index{chelonia:list} or \texttt{ls} method it is possible to list the contents of one or more collections which are specified by their Logical Name (LN).
 \hspace*{0.5cm}
@@
 \label{ssub:move}
-With the \texttt{move}\index{commands:chelonia:move} or \texttt{mv} method it is possible to move a file or collection within the namespace of chelonia (including renaming the entry).
+With the \texttt{move}\index{commands:chelonia:move}\index{chelonia:move} or \texttt{mv} method it is possible to move a file or collection within the namespace of chelonia (including renaming the entry).
 The source path and the target path should be specified as Logical Names
@@
 \label{ssub:putfile}
-With the \texttt{putFile}\index{commands:chelonia:putFile} or \texttt{put} method it is possible to upload a new file into the system creating a new Logical Name (LN).
+With the \texttt{putFile}\index{commands:chelonia:putFile}\index{chelonia:putFile} or \texttt{put} method it is possible to upload a new file into the system creating a new Logical Name (LN).
 It is possible the specify the number of needed replicas.
@@
 \label{ssub:getfile}
-With the \texttt{getFile}\index{commands:chelonia:getFile} or \texttt{get} method it is possible to download a file specified with its Logical Name (LN).
+With the \texttt{getFile}\index{commands:chelonia:getFile}\index{chelonia:getFile} or \texttt{get} method it is possible to download a file specified with its Logical Name (LN).
 If the target local path is not given, then the file will be put into the local directory.
@@
 \label{ssub:delfile}
-With the \texttt{delFile}\index{commands:chelonia:delFile} or \texttt{rm} method it is possible to delete one or more files from the system.
+With the \texttt{delFile}\index{commands:chelonia:delFile}\index{chelonia:delFile} or \texttt{rm} method it is possible to delete one or more files from the system.
 \hspace*{0.5cm}
@@
 
-With the \texttt{modify}\index{commands:chelonia:modify} or \texttt{mod} method it is possible to modify some metadata of an entry.
+With the \texttt{modify}\index{commands:chelonia:modify}\index{chelonia:modify} or \texttt{mod} method it is possible to modify some metadata of an entry.
 \hspace*{0.5cm}
@@
 
-With the \texttt{policy}\index{commands:chelonia:policy} or \texttt{pol} method it is possible to modify the policy of the entry
+With the \texttt{policy}\index{commands:chelonia:policy}\index{chelonia:policy} or \texttt{pol} method it is possible to modify the policy of the entry
 \hspace*{0.5cm}
@@
 
-With the \texttt{unlink}\index{commands:chelonia:unlink} method it is possible to remove a file or collection just from its parent collection without removing the file or collection itself.
+With the \texttt{unlink}\index{commands:chelonia:unlink}\index{chelonia:unlink} method it is possible to remove a file or collection just from its parent collection without removing the file or collection itself.
 \hspace*{0.5cm}
@@
 \label{ssub:credentialdelegation}
-With the \texttt{credentialDelegation}\index{commands:chelonia:credentialDelegation} or \texttt{cre} method it is possible to delegate credentials to the Bartender.
+With the \texttt{credentialDelegation}\index{commands:chelonia:credentialDelegation}\index{chelonia:credentialDelegation} or \texttt{cre} method it is possible to delegate credentials to the Bartender.
 \hspace*{0.5cm}
@@
 \label{ssub:removecredentials}
-With the \texttt{removeCredentials}\index{commands:chelonia:removeCredentials} or \texttt{rem} method it is possible to remove the previously delegated credentials.
+With the \texttt{removeCredentials}\index{commands:chelonia:removeCredentials}\index{chelonia:removeCredentials} or \texttt{rem} method it is possible to remove the previously delegated credentials.
 \hspace*{0.5cm}
@@
 \label{ssub:makemountpoint}
-With the \texttt{makeMountPoint}\index{commands:chelonia:makeMountPoint} or \texttt{makemount} method it is possible to create a mount point within the namespace of Chelonia which points to a GridFTP server.
+With the \texttt{makeMountPoint}\index{commands:chelonia:makeMountPoint}\index{chelonia:makeMountPoint} or \texttt{makemount} method it is possible to create a mount point within the namespace of Chelonia which points to a GridFTP server.
 \hspace*{0.5cm}
@@
 \label{ssub:unmakemountpoint}
-With the \texttt{unmakeMountPoint}\index{commands:chelonia:unmakeMountPoint} or \texttt{unmount} method it is possible to remove a previously created mount point.
+With the \texttt{unmakeMountPoint}\index{commands:chelonia:unmakeMountPoint}\index{chelonia:unmakeMountPoint} or \texttt{unmount} method it is possible to remove a previously created mount point.
 \hspace*{0.5cm}

### 2009-11-13 doc/manuals/user_interface/ui-data.tex
added missing arcsrmping, some more index entires

@@
 
+\subsection{arcsrmping}\label{sec:arcsrmping}
+
+The \texttt{arcsrmping}\index{arcsrmping}\index{commands:arcsrmping}
+command is used to quickly test availability of an SRM service, similarly to the \textit{ping} tool in Unix.
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{arcsrmping [options] $<$service$>$}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+    Options:&&\\
+   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG - default WARNING\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}&&print version information\\
+   \texttt{-h, --help}&&print help page\\
+    Arguments:&&\\
+   \texttt{service} && A URL to an SRM service\\
+\end{longtable}
+
+The \texttt{arcsrmping} command is a \textit{ping} client for the SRM service. It sends an SRM ping request to the SRM service and displays the result.
+
 % \subsection{arcacl}\label{sec:arcacl}

### 2009-11-18 doc/tech_doc/sec/arc-security-documentation.tex
Switching to latest security document. Making it buildable.
Fixing references and citations. A bit of cleaning (not finished yet).

@@
 \usepackage{longtable}                             %for breaking tables
-\usepackage{trackchanges}
+%\usepackage{trackchanges}
 %
@@
 %%%%---- The title ----
-{\centering \textsc{\ SECURITY FRAMEWORK OF ARC1}\Large \par}
+{\centering \textsc{SECURITY FRAMEWORK OF ARC NOX}\Large \par}
 \vspace*{0.5cm}
@@
 
-The security framework of the ARC1 includes two parts of capabilities: security capability embedded in hosting environment, and security capability implemented as plug-ins with well-defined interfaces which can be accessed by hosting environment and applications.
-The following design concerns were employed when designing:
-
-\begin{itemize}
-    \item Interoperability and standardlization.
-In consistent with the main design concern of the ARC1, interoperability and standardization is considered in security framework.
-For example, in terms of authentication, PKI infrastructure and proxy certificate (RFC3820 [1]) is used as most of the other grid middle-wares do.
+The security framework of the ARC NOX includes two parts of capabilities: security capability embedded in hosting environment, and security capability implemented as plug-ins with well-defined interfaces which can be accessed by hosting environment and applications.
+The following concerns were employed when designing this framework:
+
+\begin{itemize}
+    \item Interoperability and standardization.
+In consistency with the main design concerns of the ARC middle-ware, interoperability and standardization is considered in security framework.
+For example, in terms of authentication, PKI infrastructure and X.509 proxy certificates (RFC3820 \cite{x509proxy}) are used as most of the other Grid middle-wares do.
 Since supporting of standardization is a way for implementing interoperability, some standard specifications have been implemented as prototype and tested, such as SAML specification.
@@
     \item Modularity and extensibility.
-Besides the security functionality which is embedded in hosting environment, the other security functionality is implemented as plug-ins which has well-defined interfaces, and is configurable and dynamically loadable.
-Since the interoperation interface between security plug-in and hosting environment or applications is predefined, it is easy to extend the security functionality in order to support some other security capability by implementing the interface.
-
-    \item Backward compatibility. The GSI (Grid Security Infrastructure) based mechanism has been a de-facto solution for grid security. The design of security framework should be compatible to it.
+Besides the security functionality which is embedded in hosting environment, a lot of functionality is implemented as plug-ins which has well-defined interfaces, and are configurable and dynamically loadable.
+Since the interoperation interface between security plug-ins and hosting environment or applications is predefined, it is easy to extend the security functionality in order to support other new security capabilites.
+
+    \item Backward compatibility.
+The GSI (Grid Security Infrastructure)\cite{gsi} based mechanism has been a de-facto solution for Grid security for long time alredy.
+Although it has drawback for compatibility reasons the security framework should include it as part of its capability.
 
@@
 
-\section{Security architecture in HED: SecHandler and PDP} % (fold)
+\section{Security architecture in HED: Security Handler and Policy Decision Point} % (fold)
 \label{sec:sec_architecture}
 
-\subsection{Structure of SecHandler and PDP} % (fold)
+\subsection{Structure of Security Handler and Policy Decision  Point} % (fold)
 \label{subsec:structure_sechandler}
 
-In the implementation of the ARC1, there is a Service Container – the Hosting Environment Daemon (HED) (D1.2-2, [2]) which provides a hosting place for various services in application level, as well as a flexible and efficient communication mechanism.
+In the implementation of the ARC NOX, there is a Service Container called Hosting Environment Daemon (HED) \cite{hed} which provides a hosting place for various services at application and protocol level, as well as a flexible and efficient communication mechanism.
 
 HED contains a framework for implementing and enforcing authentication and authorization.
-Each Message Chain Component (MCC) or service has a common interface for implementing various authentication and authorization functionality.
-This functionality is implemented by using pluggable components (plug-ins) called SecHandler.
+Each Message Chain Component (MCC) or Service has a common interface for implementing various authentication and authorization functionality.
+This functionality is implemented by using pluggable components (plug-ins) called Security Handlers (SecHandler).
 The SecHandler components are C++ classes and provide method for processing messages traveling through Message Chains of the HED.
-Each MCC or Service  usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called ``incoming'' and ``outgoing'' respectively.
+Each MCC or Service  usually implement two queues of SecHandlers one for incoming messages and one for outgoing called ``incoming'' and ``outgoing'' respectively.
 It is possible for MCC or Service to implement other set of queues.
@@
 
+Each SecHandler is configured inside same configuration file used for configuring whole chain of MCCs.
+Some of implemented SecHandler components also make use of other pluggable and configurable sub-modules which specifically handle various security functionalities, such as authorization, authentication, etc.
+The currently implemented sub-modules used by some SecHandlers are Policy Decision Point (PDP) components such as Arc PDP which can process ARC specific Request and Policy documents.
+Figure \ref{fig:messageinMCC} shows the structure of a MCC/Service, and the message processing sequence inside it.
+
 \begin{figure}[ht]
 \centering{{{\includegraphics[width=1.0\textwidth]{MCC_Component.pdf}}}
-\caption{\label{fig:messageinMCC}\textit{There are usually two chains of SecHandlers inside the MCC or service.
+\caption{
+\label{fig:messageinMCC}
+\textit{There are usually two chains of SecHandlers inside the MCC or service.
 Each SecHandler will parse the Security Attributes which are generated by the upstream MCC/services or probably upstream SecHandlers in the same or other MCC/Service, and do message processing or authenticate or authorize the incoming/outgoing message based on the collected information.
 The SecHandler can also change the payload and attributes of Messsage itself.
-For example, the Username-Token SecHandler will insert the WSS Username Token [3] into header part of SOAP message.
-The PDPs are called by the SecHandlers and are supposed to make authorization decision.
-Here the two local PDP and one remote PDP service is just for demonstration, and one or any number of PDPs can be configured under each SecHandler.
+For example, the Username-Token SecHandler will insert the WSS Username Token \cite{ws-security} into header part of SOAP message.
+The PDPs are called by the dedicated SecHandlers and are supposed to make authorization decision.
+In this example two local PDPs and one remote PDP service are presented, and any number of PDPs can be configured under corresponding SecHandler.
 }} }
@@
 
-Each SecHandler is configured inside same configuration file used for configuring whole chain of MCCs.
-Some of implemented SecHandler components also make use of pluggable and configurable sub-modules which specifically handle various security functionalities, such as authorization, authentication, etc.
-The currently implemented sub-modules used by some SecHandlers are Policy Decision Point (PDP) components such as Arc PDP which can process ARC specific Request and Policy documents.
-Figure \ref{fig:messageinMCC} gives the structure of a MCC/Service, and the message sequence inside it.
-
-The configuration of SecHandler components for an example ``Echo'' service is shown as below.
-Example Echo service is configured to use two SecHandlers, both responsible for authorization.
-First SecHandler uses the identity of client extracted from the incoming message to map it into local identity like local Linux username.
-In this case all clients are mapped to local account “test”.
-The second one uses two PDPs: one will compose ARC specific authorization request based on the Security Attributes collected from “incoming” message and evaluate it against the ARC specific authorization policy which is defined in local file “policy.xml”; the other will compare the X509 identity of client extracted from the incoming message against list of identities stored locally.
+The configuration of SecHandler components for an example ``Echo'' service is shown below.
+Example ``Echo'' service is configured to use two SecHandlers, both performing authorization.
+First SecHandler uses the X.509 identity of client (certificate subject) extracted from the incoming message to map it into local identity like Linux username.
+In this case all clients are mapped to local account ``test''.
+The second one uses two PDPs: one will compose ARC specific authorization request based on the Security Attributes collected from incoming message and evaluate it against the ARC specific authorization policy which is stored in local file ``policy.xml'', the other will compare the X.509 identity of client extracted from the incoming message against list of identities stored locally.
 
@@
 
-When one component (MCC or service) is loaded according to the configuration information, the SecHandler under the component and the plug-ins like PDP which are attached to the SecHandler will be loaded as well.
-
-There is one simple interface (see below) defined in class SecHandler, which will be called by the containing MCC/Service once there is message (incoming or outgoing) need to be processed.
+When either MCC or Service are loaded according to the configuration information, the SecHandler under the component and the plug-ins like PDP which are attached to the SecHandler will be loaded as well.
+
+Each SecHandler implements one simple interface (see below), which is called by the containing MCC/Service once there is message (incoming or outgoing) need to be processed.
 
@@
 
-Class \textit{SecHandler} is an abstract class which includes a general interface called \textit{Handle} which uses Message object as argument.
-Any security handler implementation should inherit class \textit{SecHandler} and implement the interface according to the actual functionality.
-The interface only return simple \textit{Boolean} value, and any useful information generated during the calling of this interface should be put into the security attribute of the message, or put into the payload itself.
-
-Currently, the ARC1 comes with the following four security handler implemented:
-
-\begin{itemize}
-    \item arc.authz – Authorization SecHandler
-The arc.authz and serves as container for he Policy Decision Point components.
+Class \textit{SecHandler} is an abstract class which includes a general interface method called \textit{Handle} which takes Message object as argument.
+Any security handler implementation must inherit from class \textit{SecHandler} and implement the interface according to the actual functionality.
+The method returns simple \textit{Boolean} value, and any useful information generated during the calling of this interface should be put into the security attributes of the message, or put into the payload itself.
+
+Currently, the ARC NOX comes with the following security handlers implemented:
+
+\begin{itemize}
+    \item arc.authz Authorization SecHandler
+The arc.authz and serves as container for the Policy Decision Point components.
 It is responsible for calling their interface and getting back the authorization result.
 Then obtained results are processed and combined decision is made.
-Description of configuration and examples can be found in section 8.7.
+Description of configuration and examples can be found in section \ref{subsec:authzhandler_conf}.
 Usually the Authorization SecHandler and included PDPs are used on the service side of communication channel.
@@
 
-    \item identity.map – Identity Mapping SecHandler
+    \item identity.map Identity Mapping SecHandler
 The identity.map is a specific authorization oriented security handler.
@@
 
-    \item delegation.collector – Delegation SecHandler
-The delegation.collector is responsible for collecting the delegation policy information from the remote proxy credential (proxy certificate is compatible to RFC3820) inside the message, and putting this policy into message’s security attribute for the usage of other components, such as  delegation.pdp.
-
-    \item usernametoken.handler – UseranemToken SecHandler
-The task of the  usernametoken.handler is to generate the WS-Security Username-Token and add it into header of SOAP message which is the payload of outgoing message.
+    \item delegation.collector Delegation Collector SecHandler
+The delegation.collector is responsible for collecting the delegation policy information from the remote proxy credential (proxy certificate compatible with RFC3820) inside the message, and putting this policy into the message security attribute for the usage of other components, such as the ``delegation.pdp''.
+
+    \item usernametoken.handler UseranemToken SecHandler
+The task of the  usernametoken.handler is to generate the WS-Security\cite{ws-security} Username Token and add it into header of SOAP message which is the payload of outgoing message.
 It can also extract the WS-Security Username-Token from the header of SOAP message which is the payload of incoming message.
+
+    \item x509token.handler X.509 Token SecHandler
+This SecHandler generates and process the WS-Security\cite{ws-security} X.509 Token inside the header of SOAP message.
+
+    \item samltoken.handler
+
+    \item saml2ssoassertionconsumer.handler
+
+    \item delegation.handler
 
@@
 
-Below shows the definition of abstract class PDP.
-The implementation could be some function which implements the interface by composing the policy evaluation request, evaluating this request against some policy, and returning the evaluation result, or just by composing the policy evaluation request, invoking some remote policy decision web service and getting back the evaluation result.
+Below is the definition of abstract class PDP.
+The implementation for example could implement method isPermitted() by composing the policy evaluation request, evaluating this request against some policy, and returning the evaluation result.
+Or it could compose the policy evaluation request, invoke some remote policy decision web service and return back the evaluation result.
 
@@
 
-Class \textit{PDP} is an abstract class which includes a general interface called \textit{isPermitted} which uses Message object as argument.
-Any policy decision point implementation should inherit class PDP and implement the interface according to the actual functionality.
-The interface only return simple \textit{Boolean} value, and any useful information generated during the calling of this interface should be put into the security attribute of the message, or put into the payload itself.
-
-Currently, the ARC1 comes with the following four policy decision point implementation:
-
-\begin{itemize}
-    \item arc.pdp – Arc PDP
-The Arc PDP will organize the security attributes into the ARC specific authorization request, call the policy evaluator to evaluate the request against the policy (which is in ARC specific format) repository, and get back the evaluation result.
-See paragraph 3 for detail information about request schema and policy schema.
-
-    \item xacml.pdp – XACML PDP
-The XACML PDP will organize the security attributes into the XACML authorization request (see: http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-context-schema-os.xsd), call the policy evaluator to evaluate the request against the policy (which is in XACML format: http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-policy-schema-os.xsd) repository, and get back the evaluation result.
-
-    \item delegation.pdp – Delegation PDP
-The Delegation PDP is basically similar to Arc PDP, except it uses the delegation policy parsed from remote proxy credential by delegation.collector, and evaluates the request against delegation policy.
-See section 6for the design idea and use case of delegation policy in fine-grained identity delegation.
-
-    \item simplelist.pdp – Simplelist PDP
-The Simplelist PDP is a simplest implementation of policy decision point. It will match the identity extracted from the remote credential (or proxy credential) with local list of permitted identities.
-
-    \item pdpservice.invoker – PDP Service Invoker
+Class \textit{PDP} is an abstract class which includes a general interface method called \textit{isPermitted} which uses Message object as argument.
+Any policy decision point implementation must inherit from class PDP and implement the interface according to the actual functionality.
+The interface method return simple \textit{Boolean} value, and any useful information generated during the calling of this interface should be put into the security attribute of the message, or put into the payload itself.
+
+Currently, the ARC NOX comes with the following PDP implementations:
+
+\begin{itemize}
+    \item arc.pdp Arc PDP
+The Arc PDP will organize the security attributes into the ARC specific authorization request, call the policy evaluator to evaluate the request against the policy (which is in ARC specific format) stored in local repository, and return back the evaluation result.
+See section \ref{sec:policy_eval} for detailed information about request and policy schema.
+
+    \item xacml.pdp XACML PDP
+The XACML PDP will organize the security attributes into the XACML authorization request (see: http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-context-schema-os.xsd), call the policy evaluator to evaluate the request against the policy (which is in XACML format: http://docs.oasis-open.org/xacml/2.0/access\_control-xacml-2.0-policy-schema-os.xsd) stored in local repository, and return back the evaluation result.
+
+    \item delegation.pdp Delegation PDP
+The Delegation PDP is basically similar to Arc PDP, except it uses the delegation policy parsed from remote proxy credential by delegation.collector, and evaluates the request against configured delegation policy.
+See section \ref{sec:delegation} for the design idea and use case of delegation policy in fine-grained identity delegation.
+
+    \item simplelist.pdp Simplelist PDP
+The Simplelist PDP is a simplest implementation of policy decision point. It will match the identity extracted from the remote credential (or proxy credential) to local list of permitted identities.
+
+    \item pdpservice.invoker PDP Service Invoker
 The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the same functionality as Arc PDP or XACML PDP, except that the evaluation request and response are carried by SOAP message.
@@
 
+    \item allow.pdp Allow PDP
+This PDP always returns positive result.
+
+    \item deny.pdp Deny PDP
+This PDP always returns negative result.
+
 \end{itemize}
@@
 
-The ARC1 defines specific evaluation request and policy schema.
+The ARC NOX defines specific evaluation request and policy schema.
 Based on the schema definition, one policy evaluation engine is implemented.
@@
 
-It's configuration is presented in Section \ref{subsec:pdpservice_conf}.
+It's configuration is presented in section \ref{subsec:pdpservice_conf}.
 
@@
 Each Security Attribute stores it's information in internal format and is capable to export it to one of predefined formats using \textit{Export()} method.
-Currently only supported format is ARC Policy/Request XML document described in Section 8.1and 8.2
+Currently only supported format is ARC Policy/Request XML document described in sections \ref{subsec:authz_policy} and \ref{subsec:authz_request}.
 Collectors of Security Attributes instantiate corresponding classes and link them to Secuirity Attributes containers - MessageAuth $<$arc/message/MessageAuth.h$>$ and MessageAuthContext $<$arc/message/Message.h$>$ storing collected attributes per request and per session correspondingly.
@@
 Current implementations of Security Attributes Collectors are either integrated into existing MCCs or implemented as separate SecHandler plugins.
-See Section 5.2for available Collectors and corresponding Security Attributes.
+See section \ref{subsec:sec_attr_avail_collectors} for available Collectors and corresponding Security Attributes.
 
@@
 It evaluates either all or selected set of attributes against specified Policy documents thus making it possible to enforce policies defined/selected by service providers.
-Delegation PDP is described below in Section 6.3.
+Delegation PDP is described below in section \ref{subsec:delegation_pdp}.
 
@@
 
-Here Security Attribute collectors distributed as part of the ARC1 are described except those used for Delegation Restrictions. Those are described in Section 6.2
+Here Security Attribute collectors distributed as part of the ARC NOX are described except those used for Delegation Restrictions. Those are described in section \ref{subsec:delegation_collector}
 
@@
 
-Configuration of the TLS MCC is described in section 8.5.
+Configuration of the TLS MCC is described in section \ref{subsec:tlsmcc_conf}.
 
@@
 \label{subsec:delegation_arch}
-In current implementation delegation is achieved through Identity Delegation implemented using X509 Proxy Certificates as defined in RFC 3820 [1].
-Client wishing to allow service to act on it's behalf provides Proxy Certificate to the service using Web Service based Delegation interface described in Section 6.4.
+In current implementation delegation is achieved through Identity Delegation implemented using X.509 Proxy Certificates as defined in RFC 3820 \cite{x509proxy}.
+Client wishing to allow service to act on it's behalf provides Proxy Certificate to the service using Web Service based Delegation interface described in section \ref{subsec:delegation_interface}.
 
@@
 
-Each deployment implementing Delegation Restrictions must use dedicated Security Handler plugin (see section 5.1) to collect all Policy documents from Proxy Certificates used for establishing secure connection.
-Then those documents must be processed by dedicated Policy Decision Point plugin (see section 2.3) to make a final decision based on collected Policies and various information about client's identity and requested operation.
+Each deployment implementing Delegation Restrictions must use dedicated Security Handler plugin (see section \ref{subsec:delegation_collector}) to collect all Policy documents from Proxy Certificates used for establishing secure connection.
+Then those documents must be processed by dedicated Policy Decision Point plugin (see section \ref{subsec:interface_pdp}) to make a final decision based on collected Policies and various information about client's identity and requested operation.
 Service or MCC chain supporting Delegation Restrictions must accept negative decision of this PDP as final and do not override it with any other decision based on other policies.
@@
 \label{subsec:delegation_collector}
-This Security Attribute is collected by dedicated Security Handler plugin named ``delegation.collector'' avaialble as part of the ARC1 distribution.
-It extracts policy document stored inside X509 certificate proxy extension as defined in RFC3820 and described in Section 6.1.
+This Security Attribute is collected by dedicated Security Handler plugin named ``delegation.collector'' avaialble as part of the ARC NOX distribution.
+It extracts policy document stored inside X.509 certificate proxy extension as defined in RFC3820 and described in section \ref{subsec:delegation_arch}.
 All proxy certificates in a chain provided by client are examined and all available policies are extracted.
-Configuration of Delegation SecHandler is described in section 8.12.
+Configuration of Delegation SecHandler is described in section \ref{subsec:deleg_sechandler_conf}.
 
@@
 
-Proxy certificates with id-ppl-inheritAll [5.
-RFC3820.
-http://www.faqs.org/rfcs/rfc3820.html] property are passed through and no policy document is generated for them.
+Proxy certificates with id-ppl-inheritAll \cite{x509proxy} property are passed through and no policy document is generated for them.
 Proxies with other type of policies including id-ppl-independent are not accepted and generate immediate failure.
@@
 Differently from Arc PDP it is meant to be used for enforcing policies defined by client.
-Configuration of Delegation PDP is described in section 8.11.
+Configuration of Delegation PDP is described in section \ref{subsec:deleg_pdp_conf}.
 
@@
 
-Delegation interface in the ARC1 is implemented using Web Service approach.
-Each ARC1 service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials.
+Delegation interface in the ARC NOX is implemented using Web Service approach.
+Each ARC NOX service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials.
 Here is how delegation procedure works (also shown in Figure \ref{fig:delegation_flow} and Figure \ref{fig:delegation_flow1}) :
@@
 
-The Policy maybe any of any type supported by ARC1 middleware (or third-party plugins) installed on the services where that policy is processed.
-Currenlty supported Policies include ARC Policy (described in section 8) and GACL Policy [2].
+The Policy maybe any of any type supported by ARC NOX middleware (or third-party plugins) installed on the services where that policy is processed.
+Currenlty supported Policies include ARC Policy (described in section \ref{subsec:authz_policy}) and GACL Policy \cite{gacl}.
 
@@
 Another example of the delegation policy is presented below.
-This policy restricts usage of delegated credentials to SOAP operation CreateActivity (line 5) of Basic Execution Service (BES) [4] namespace (line 9).
+This policy restricts usage of delegated credentials to SOAP operation CreateActivity (line 5) of Basic Execution Service (BES) \cite{ogsa-bes} namespace (line 9).
 Such policy could be embedded into credentials delegated to high level Brokering service performing Grid job submission to low level BES on behalf of user.
@@
 \label{subsec:username_token}
-The UsernameToken SecHandler is meant for processing - generating and extracting - WS-Security [4] UsernameToken in the SOAP header.
+The UsernameToken SecHandler is meant for processing - generating and extracting - WS-Security \cite{ws-security} UsernameToken in the SOAP header.
 Hence it must be attached to the MCC which processes SOAP payloads - like SOAP MCC or Service accepting SOAP messages.
@@
 \label{subsec:x509_token}
-The X509Token SecHandler is meant for processing - generating and extracting - WS-Security [4] X509Token from SOAP header.
+The X.509 Token SecHandler is meant for processing - generating and extracting - WS-Security \cite{ws-security} X.509 Token from SOAP header.
 Hence it must be attached to the MCC which processes SOAP payloads – like SOAP MCC or Service accepting SOAP messages.
@@
 
-For the outgoing message this SecHandler creates X509Token in SOAP header. SOAP message body is encrypted and signed.
+For the outgoing message this SecHandler creates X.509 Token in SOAP header. SOAP message body is encrypted and signed.
 
@@
 \label{subsec:saml_token}
-The SAMLToken SecHandler is meant for processing - generating and extracting - WS-Security [4] SAMLToken from SOAP header.
+The SAMLToken SecHandler is meant for processing - generating and extracting - WS-Security \cite{ws-security} SAMLToken from SOAP header.
 Hence it must be attached to the MCC which processes SOAP payloads – like SOAP MCC or Service accepting SOAP messages.
@@
 Figure \ref{fig:samltoken_aa} shows the interaction among Client, Grid/Web Service and Attribute Authority when SAMLToken security hanlder is deployed.
-The Attribute Authority service is described in Section \ref{sec:saml_aa_service}.
+The Attribute Authority service is described in section \ref{sec:saml_aa_service}.
 For the Grid/Web service, if the service is hosted on ARC middleware, then the SAMLToken security handler should be deployed; if the service is hosted on other middlewares, then there should be other ways for supporting SAML Token authentication, e.g., Rampart (WS-Security modlue for Axis2).
@@
 
-See Section 8.9for the explanation of ARC Policy.
+See section \ref{subsec:arcpdp_conf} for the explanation of ARC Policy.
 
@@
 
-The policy file ``simplelist'' is a local file which contains the list of X509 subjects of authorized entities.
+The policy file ``simplelist'' is a local file which contains the list of X.509 subjects of authorized entities.
 It the peer certificate is proxy certificate, the identity in this list should only include the original DN of users's certificate.
@@
 
-There is an example policy for echo service below. See Section 8.1for the policy schema. The example policy is  made of following elements:
+There is an example policy for echo service below. See section \ref{subsec:authz_policy} for the policy schema. The example policy is  made of following elements:
 
@@
     \item Lines 17 and 18 define allowed HTTP operations to be ``POST'' and ``GET''. Line 19 also defines SOAP operation ``echo'' to be applied to service at path defined above.
-    \item Lines 10 and 9 require the requester to present X509 certificate with specified identity and signed by specified Certification Authority.
+    \item Lines 10 and 9 require the requester to present X.509 certificate with specified identity and signed by specified Certification Authority.
     \item No \textit{$<$Conditions/$>$} defined.
@@
 \subsection{Delegation PDP configuration} % (fold)
-\label{subsec:delegpdp_conf}
+\label{subsec:deleg_pdp_conf}
 Configuration XML schema with comments available at
@@
 
-For the client side, the developer should add the configuration information about X.509 security handler into client's configuration, in order to generate X509 Token into any SOAP message.
+For the client side, the developer should add the configuration information about X.509 security handler into client's configuration, in order to generate X.509 Token into any SOAP message.
 Below is an example of how to use it:
@@
 
-Below is an example of configuration of X509Token SecHandler inside MCCSOAP component of the service side.
-This example processes X509Token related information in SOAP message and returns failure if any problem found.
+Below is an example of configuration of X.509 Token SecHandler inside MCCSOAP component of the service side.
+This example processes X.509 Token related information in SOAP message and returns failure if any problem found.
 
@@
 \label{sec:samlsso}
-   In ARC1, besides that GSI is supported for talking with external grid services which are based on GSI, the standard TLS/SSL is also supported.
-No matter standard SSL/TLS or GSI is used, mutual authentication applies for both of them, and is the default configuration for grid deployment; and X.509 certificate is required for both of the client and service sides.
+   In ARC NOX, besides that GSI is supported for talking with external Grid services which are based on GSI, the standard TLS/SSL is also supported.
+No matter standard SSL/TLS or GSI is used, mutual authentication applies for both of them, and is the default configuration for Grid deployment; and X.509 certificate is required for both of the client and service sides.
 X.509 certificates is issued by certificate authorities (CA), and then CAs constitute trust federation and guarantee two different X.509 certificates from different CAs can accomplish authentication to each other.
-So if a user would access grid system, he/she should own a X.509 certificate which is issued by a CA that is trusted by other's entity in the grid system.
-
-   ARC1 provides another option by which users are not required to possess a X.509 certificate for authentication.
+So if a user would access Grid system, he/she should own a X.509 certificate which is issued by a CA that is trusted by other's entity in the Grid system.
+
+   ARC NOX provides another option by which users are not required to possess a X.509 certificate for authentication.
 It is based on AAI(Authentication and Authorization Infrastructure) and utilize users' existing community credential (such as username password) for authentication rather than X.509 credential.
@@
 \label{sec:slcs}
-In order to create X.509 credential for inter-operating with Grid services that requires client side certificate authentication, based on the solution described in Section 9, a short lived credential service (SLCS) is implemented by which user can get a short-lived X.509 certificate without being bothered to contact any registration authority (RA) or certificate authority (CA).
+In order to create X.509 credential for inter-operating with Grid services that requires client side certificate authentication, based on the solution described in section \ref{sec:samlsso}, a short lived credential service (SLCS) is implemented by which user can get a short-lived X.509 certificate without being bothered to contact any registration authority (RA) or certificate authority (CA).
 
     SLCS service is for signing short-lived x509 credential based on user's username/password credential.
-Then the user can use this short-lived x509 credential to access grid services/resources where x509 credential is required by default.
+Then the user can use this short-lived x509 credential to access Grid services/resources where x509 credential is required by default.
 SLCS service should depend on the SP (Service provider) service which is one of the participants of SAML2 SSO profile (SAML2 SSO profile is used for authenticating based on username/password credential and getting SAML authentication assertion; and SAML authentication assertion is then used as basis for signing short-lived x509 credential).
@@
     SLCS service should specifically include the CA credential (certificate and key file) and the serial number file (which includes the serial number of each signed certificate), so SLCS service needs to be configured with $<$CACertificate$>$ $<$CAKey$>$ and $<$CASerial$>$.
-Therefore, the service administrator should firstly create a CA credential (note for interoperability purpose in the production grid deployment, the CA credential should be trusted by others).
+Therefore, the service administrator should firstly create a CA credential (note for interoperability purpose in the production Grid deployment, the CA credential should be trusted by others).
 
@@
 \label{sec:delegation_service}
-Besides the delegation interface described in Section 6.4, which provides interface for X.509 credential delegation.
+Besides the delegation interface described in section \ref{subsec:delegation_interface}, which provides interface for X.509 credential delegation.
 A dedicated Web Service is also developed for credential delegation, with the intention of providing a general approach for using delegation by changing the configuration information, rather than a specific approach which requires directly changing the code.
@@
 As long as there is another attribute administration service (in parallel to VOMS Admin Service) which can nicely be used to manage the attributes, the SAML AA Service can be configured to adopt that service.
-The relationship between ARC's implementation and VOMS implementation is shown in \ref{figure:arc_voms}.
+The relationship between ARC's implementation and VOMS implementation is shown in \ref{fig:arc_voms}.
 

### 2009-11-18 doc/tech_doc/sec/arc-security-documentation.tex
More references fixed. 2 teporary pictures added. Few mising pictures identified.

@@
 The ARC NOX defines specific evaluation request and policy schema.
-Based on the schema definition, one policy evaluation engine is implemented.
+Based on the schema definition the policy evaluation engine is implemented.
 The design principal of policy evaluation engine is generality by which the implementation of the policy evaluation engine can be easily extended to adopt some other policy schema, such as XACML policy schema.
@@
 
-\subsection{Policy evaluation engine --- Support of XACML policy and request} % (fold)
+\subsection{Policy evaluation engine. Support for XACML policy and request} % (fold)
 \label{subsec:policy_engine_xacml}
-Currently, XACML specification is partially supported/implemented in ARC. More specifically, except the \textit{$<$Obligation$>$} element, other elements are supported.
+Currently, XACML specification \cite{xacml} is partially supported/implemented in ARC. More specifically, except the \textit{$<$Obligation$>$} element, other elements are supported.
 
@@
 \label{sec:policy_decision_service}
-Policy decision service is a service implementation which configurably contains the functionality of ArcPDP or XACMLPDP.
-It will accept the soap request containing policy decision request and return soap response containing policy decision response.
+Policy decision service is a service implementatiing Arc PDP and XACML PDP depending on its configuration.
+It will accept the SOAP request containing policy decision request and return SOAP response containing policy decision response.
 
@@
 
-It's configuration is presented in section \ref{subsec:pdpservice_conf}.
+Its configuration is presented in section \ref{subsec:pdpservice_conf}.
 
@@
 
-\section{Security Attributes -- how to compose policy decision request to policy evaluation engine} % (fold)
+\section{Security Attributes. How to compose policy decision request to policy evaluation engine} % (fold)
 \label{sec:sec_attributes}
@@
 
-Each Security Attribute stores it's information in internal format and is capable to export it to one of predefined formats using \textit{Export()} method.
+Each Security Attribute stores its information in internal format and is capable to export it to one of predefined formats using \textit{Export()} method.
 Currently only supported format is ARC Policy/Request XML document described in sections \ref{subsec:authz_policy} and \ref{subsec:authz_request}.
@@
 \label{subsubsec:sec_attr_TCP}
-Information is collected inside TCP MCC. The Security Attribute is stored under name 'TCP' and exports ARC Request with following attributes:
-
-\begin{table}[ht]
+Information is collected inside TCP MCC. The Security Attribute is stored under name 'TCP' and exports ARC Request with attributes described in table \ref{table:tcp_attr}.
+
+\begin{table}[h]
 \caption{Security Attributes collected at TCP MCC}
@@
 \label{subsubsec:sec_attr_TLS_VOMS}
-Information is collected inside TLS MCC. Generated Security Attribute class is stored under name 'TLS' and exports ARC Request with following attributes:
-
-\begin{table}[ht]
+Information is collected inside TLS MCC. Generated Security Attribute class is stored under name 'TLS' and exports ARC Request with attributes described in table \ref{table:tls_attr}.
+
+\begin{table}[h]
 \caption{Security Attributes collected at TLS MCC}
@@
 They consist of \textit{voname} and \textit{hostname} elements presented above (if \textit{hostname} is missing it is assigned string value NULL) followed by user attribute.
-It's pattern is \textit{qualifier:name=value}.
+Its pattern is \textit{qualifier:name=value}.
 Here qualifier acts as namespace and is usually same as VO name.
@@
 \label{subsubsec:sec_attr_HTTP}
-Information is collected inside HTTP MCC. The Security Attribute is stored under name 'HTTP' and exports ARC Request with following attributes:
-
-\begin{table}[ht]
+Information is collected inside HTTP MCC. The Security Attribute is stored under name 'HTTP' and exports ARC Request with attributes described in table \ref{table:http_attr}.
+
+\begin{table}[h]
 \caption{Security Attributes collected at HTTP MCC}
@@
 \label{subsubsec:sec_attr_SOAP}
-Information is collected inside SOAP MCC. Security Attribute is stored under name 'SOAP' and exports ARC Request with following attributes:
-
-\begin{table}[ht]
+Information is collected inside SOAP MCC. Security Attribute is stored under name 'SOAP' and exports ARC Request with attributes described in table \ref{table:soap_attr}.
+
+\begin{table}[h]
 \caption{Security Attributes collected at SOAP MCC}
@@
 In current implementation delegation is achieved through Identity Delegation implemented using X.509 Proxy Certificates as defined in RFC 3820 \cite{x509proxy}.
-Client wishing to allow service to act on it's behalf provides Proxy Certificate to the service using Web Service based Delegation interface described in section \ref{subsec:delegation_interface}.
+Client wishing to allow service to act on its behalf provides Proxy Certificate to the service using Web Service based Delegation interface described in section \ref{subsec:delegation_interface}.
 
@@
 \label{subsec:delegation_pdp}
-The Delegation PDP is similar to the Arc PDP described above except that it takes it's Policy documents directly from Security Attributes.
+The Delegation PDP is similar to the Arc PDP described above except that it takes its Policy documents directly from Security Attributes.
 Differently from Arc PDP it is meant to be used for enforcing policies defined by client.
@@
 Each ARC NOX service wishing to act on behalf of client identity implements this interface in order to accept delegated credentials.
-Here is how delegation procedure works (also shown in Figure \ref{fig:delegation_flow} and Figure \ref{fig:delegation_flow1}) :
+Here is how delegation procedure works (also shown in figure \ref{fig:delegation_flow} and figure \ref{fig:delegation_flow1}) :
 
@@
     \begin{itemize}
-        \item validityStart, validityEnd and validityPeriod  specify when Proxy becomes valid, when it's validity finishes or for how long the Proxy is valid respectively. For example
+        \item validityStart, validityEnd and validityPeriod  specify when Proxy becomes valid, when its validity finishes or for how long the Proxy is valid respectively. For example
 
@@
 For Identity Provider, the Shibboleth IdP implementation is used.
-Figure 8 shows the process of SAML2.0 SSO integrated in ARC client and service.
+Figure \ref{fig:saml_sso} shows the process of SAML2.0 SSO integrated in ARC client and service.
 
@@
 
-The steps shows in Figure 8 are described as follows:
+\begin{figure}[ht]
+\centering{{{\includegraphics[width=0.5\textwidth]{saml_sso.png}}}
+\caption{\label{fig:saml_sso}SAML 2.0 SSO Profile in ARC} }
+\end{figure}
+
+The steps shows in figure \ref{fig:saml_sso} are described as follows:
 
@@
 
-    The message flow of interaction between Client, ARC SP Service, ARC Service and Shibboleth IdP is also shown in Figure.9.
+    The message flow of interaction between Client, ARC SP Service, ARC Service and Shibboleth IdP is also shown in figure \ref{fig:saml_message_flow}.
+
+\begin{figure}[ht]
+ \centering{
+  % {{\includegraphics[width=0.5\textwidth]{missing_picture.png}}}
+  \caption{\label{fig:saml_message_flow}SAML 2.0 SSO Profile message flow}
+}
+\end{figure}
 
@@
     The SLCS service is also a Web Service (standard Web Service implemented by using ARC service interface), and the SLCS client is a specific command-line interface (CLI) which includes the user agent and WS client.
-The whole process of SLCS invocation is showed in Figure 10 (from step 1 to step 8), which is the same as in Figure 8, except that step 7 and step 8 are invoked for the SLCS certificate request and response.
-
- SLCS service is supposed to run together with SP service. A typical configuration for SLCS service is as follows:
+The whole process of SLCS invocation is showed in figure \ref{fig:slcs} (from step 1 to step 8), which is the same as in figure \ref{fig:saml_sso}, except that step 7 and step 8 are invoked for the SLCS certificate request and response.
+
+\begin{figure}[ht]
+ \centering{
+  {{\includegraphics[width=0.7\textwidth]{slcs.png}}}
+  \caption{\label{fig:slcs}Short Lived Credential Service}
+}
+\end{figure}
+
+ The SLCS service is supposed to run together with SP service. A typical configuration for SLCS service is as follows:
 
@@
     SLCS service and SP service should together be configured by using the Plexer.
-SP service is directly based on http and SLCS is directly based on soap.
+SP service is directly based on http and SLCS is directly based on SOAP.
 As shown in the above table, a plexer dispatches the message flow (outgoing from http.service) into two SP service (saml.sp) and SOAP service (soap.service).
@@
     Since the lifetime of the short lived credential is normally short, it is not a must to protect the private key by a pass phrase.
-As illustrated in steps (a) and (b) in Figure 10, if the private key is not protected through the Web Service client, the user can use the X.509 certificate to access Grid Service or Web Service from any kind of middleware.
+As illustrated in steps (a) and (b) in figure \ref{fig:delegation_flow1}, if the private key is not protected through the Web Service client, the user can use the X.509 certificate to access Grid Service or Web Service from any kind of middleware.
 If the private key is protected, she can use the X.509 certificate to generate a proxy certificate (by using a command-line interface utility such as grid-proxy-init , voms-proxy-init, or arcproxy ), and then use the proxy certificate to access a Grid/Web Service.
@@
 
-    As shown in Figure 11, there is a specific WS client and Web Service for processing delegation: delegation client, and delegation service.
+    As shown in figure \ref{fig:delegation_service}, there is a specific WS client and Web Service for processing delegation: delegation client, and delegation service.
 The WSDL (Web Service Description Language) of the delegation service has three main operations, the first one for processing the delegation initiation, the second one for storing the signed proxy certificate, and the last one for processing acquisition of the delegated credential.
-The sequential diagram of the delegation is detailed in Figure 12.
-
-    For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in Figure 11), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential.
-For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in Figure 11), should also use the user’s X.509 credential or a delegated credential for authentication and secure communication.
-
-    Since each WS client accesses the target service (delegation service or other general services) on behalf of the user, the trust relationship between the user’s certificates (X.509 certificate and proxy certificate) and services is required, while the trust relationship between service’s certificate and another service’s certificate is not required.
+The sequential diagram of the delegation is detailed in figure \ref{fig:delegation_service_flow}.
+
+\begin{figure}[ht]
+ \centering{
+  % {{\includegraphics[width=0.7\textwidth]{missing_picture.png}}}
+  \caption{\label{fig:delegation_service}X.509 Credential Delegation Service}
+}
+\end{figure}
+
+\begin{figure}[ht]
+ \centering{
+  % {{\includegraphics[width=0.7\textwidth]{missing_picture.png}}}
+  \caption{\label{fig:delegation_service_flow}X.509 Credential Delegation message flow}
+}
+\end{figure}
+
+    For the delegation process, the delegation WS client functionality (inside each delegation service or the initial client) which is included to access the target delegation service (d1 or d2 in figure \ref{fig:delegation_service}), should use the user’s X.509 credential or a delegated credential (on behalf of the user) for authentication and secure communication, as well as delegate one more level of the delegated credential.
+For the other general  service invocation process, the WS client functionality (inside each service or the initial client) which is included to access the target service (step a1 or a2 in figure \ref{fig:delegation_service}), should also use the user's X.509 credential or a delegated credential for authentication and secure communication.
+
+    Since each WS client accesses the target service (delegation service or other general services) on behalf of the user, the trust relationship between the user's certificates (X.509 certificate and proxy certificate) and services is required, while the trust relationship between service's certificate and another service's certificate is not required.
 

### 2009-11-19 doc/tech_doc/hed/RegisteredServiceHowTo.tex
Adding RegisteredService howto to the HED documentation as an appendix (after a request of Balazs). The new part of document needs revision.)

\section{Step-by-step instructions to add registration ability to HED services}
\label{annex:service_registration}
\subsection{General knowledge}
\label{General knowledge}
The HED has a brand new self-register ability. If your Service is prepared to co-operate with this mechanism then there will be an InfoRegister connecting to it. There are other, so called InfoRegistrar objects present that are the HED self-register mechanism's active elements communicating with different ISIS clouds. (The InfoRegisterContainter is an instance that connects them, but this is just a good-to-know detail.) The system administrator can connect the Services with ISIS clouds connecting InfoRegisters with InfoRegistrars. The configuration (for example during testing your service) has to cover the both side. (It is possible to connect one InfoRegister to more InfoRegistrars and vice versa. See figure \ref{fig:hed_internal_registration_infrastructure}.)
\begin{figure}[ht]
 \centering{
  \scalebox{0.6}{\includegraphics{RegisterRegistrar.png}}
  \caption{\label{fig:hed_internal_registration_infrastructure}The HED internal registration infrastructure}
 }
\end{figure}

The configuration details will be shown on the section Modify your configuration\ref{Modify your configuration}.

The InfoRegistrar that are connected with your Service (by configuration) will periodically ask your Service for a Registration Entry what are stored in the information system. This "poll" is done by executing the RegistrationCollector function that provide an XML document in a given format. For further details see section Compose the Registration Entry\ref{Compose the Registration Entry}! Actually this interface can be accessed in every C++ or Python service.

The InfoRegistrar collect the simultaneous generated Registration Entries and compose a Registration Message by aggregating them (if any aggregation is possible).

The general "sandbox" echo services are already prepared to this functionality and will be referenced at the proper locations.

\subsection{Change your source}
\label{Change your source}

\subsubsection{C++ source}
\label{C++ source}
\begin{itemize}
\item Your Service have to extend the RegisteredService instead of the Service class. This class is implemented in the infosys library and not in the message library as before. (Of course the new library have also to be linked to your Service.)
In your constructor please call the RegisteredService class's constructor!
\item You have also to implement the RegistrationCollector function that will provide the Registration Entry XML document.
\end{itemize}
\textbf{Your\_Service.h}
\begin{verbatim}
...
#include <arc/infosys/RegisteredService.h>
...
class Your_Service: public Arc::RegisteredService {
...
    bool RegistrationCollector(Arc::XMLNode &doc);
...
}
\end{verbatim}
For example see: src/tests/echo/echo.h in the source tree.
\textbf{Your\_Service.cpp}
\begin{verbatim}
...
Your_Service::Your_Service(Arc::Config *cfg):RegisteredService(cfg) {
...
    bool Your_Service::RegistrationCollector(Arc::XMLNode &doc) {
        Arc::NS isis_ns; isis_ns["isis"] = "http://www.nordugrid.org/schemas/isis/2008/08";
        Arc::XMLNode regentry(isis_ns, "RegEntry");
        regentry.NewChild("SrcAdv").NewChild("Type") = "Your_Service_Type";
        regentry.New(doc);
        return true;
    }
...
}
\end{verbatim}
For example see: src/tests/echo/echo.cpp in the source tree.
\textbf{Makefile.am}
\begin{verbatim}
libyourservice_la_LIBADD = ... \
                           $(top_srcdir)/src/hed/libs/infosys/libinfosys.la
\end{verbatim}
For example see: src/tests/echo/Makefile.am in the source tree.

\subsubsection{Python source}
\begin{itemize}
\item The PythonService is already extending the RegisteredService class so you have nothing to do on this place.
\item Your only to-do is to implement the RegistrationCollector function and provide the Registration Entry XML document.
\end{itemize}
\textbf{Your\_Service.py}
\begin{verbatim}
def RegistrationCollector(self, doc):
    regentry = arc.XMLNode('<RegEntry />')
    regentry.NewChild('SrcAdv').NewChild('Type').Set('Your_Service_Type') 
    #Place the document into the doc attribute
    doc.Replace(regentry)
    return True
\end{verbatim}
For example see: src/services/echo\_python/EchoService.py in the source tree.

\subsection{Compose the Registration Entry}
\label{Compose the Registration Entry}
The Registration Entry is an XML document with a given format. The HED internal mechanism tries to aggregate these messages in the Registration Messages.

There are 6 mandatory element in a Registration Entry. At least these should be provided by the service developer. The Type, Endponint reference and the ID of the Service, and the Expiration and Generation time of the message. If the ID is not present then it will be deputized with the Endpoint reference. Finally the Generation time of the message will be also automatically filled if missing from the Registration Entry. (See figure \ref{fig:registration_message}.)
\begin{figure}[ht]
 \centering{
  \scalebox{0.6}{\includegraphics{RegistrationMessage.png}}
  \caption{\label{fig:registration_message}The structure of an aggregated Registration Message}
 }
\end{figure}
The schema of the Registration Entry:
\begin{verbatim}
    <!-- List of the service types will be provided by the GLUE-2.0 -->
    <xsd:simpleType name="ServiceTypeType">
        <xsd:restriction base="xsd:string">
        </xsd:restriction>
    </xsd:simpleType>

    <xsd:complexType name="NameValuePairType">
        <xsd:sequence>
            <xsd:element name="Name" type="xsd:string"/>
            <xsd:element name="Value" type="xsd:string"/>
        </xsd:sequence>
    </xsd:complexType>

    <xsd:complexType name="ServiceAdvertisementMetadataType">
        <xsd:sequence>
            <!-- Globally unique and persistent ID of the service -->
            <xsd:element name="ServiceID" type="xsd:string" minOccurs="1" 
            maxOccurs="1"/>
            <!-- Time of information generation or collection -->
            <xsd:element name="GenTime" type="xsd:dateTime" minOccurs="1" 
            maxOccurs="1"/>
            <xsd:element name="Expiration" type="xsd:duration"/>
        </xsd:sequence>
    </xsd:complexType>

    <xsd:complexType name="ServiceAdvertisementType">
        <xsd:sequence>
            <!-- General part of the Service Advertisment -->
            <xsd:element name="Type" type="isis:ServiceTypeType"/>
            <xsd:element name="EPR" type="wsa:EndpointReferenceType"/>
            <!-- Service specific part of the Service Advertisment -->
            <xsd:element name="SSPair" type="isis:NameValuePairType" minOccurs="0" 
            maxOccurs="unbounded"/>
        </xsd:sequence>
    </xsd:complexType>

    <xsd:complexType name="RegistrationEntryType">
        <xsd:sequence>
            <xsd:element name="SrcAdv" type="isis:ServiceAdvertisementType" 
            minOccurs="1" maxOccurs="1"/>
            <xsd:element name="MetaSrcAdv" type="isis:ServiceAdvertisementMetadataType" 
            minOccurs="1" maxOccurs="1"/>
        </xsd:sequence>
    </xsd:complexType>
\end{verbatim}
It can be also found in the source tree on the path: src/services/isis/schema/isis.xsd
An examle Registration Entry XML document:
\begin{verbatim}
<RegEntry>
  <MetaSrcAdv>
    <ServiceID>http://your.domain.com/echo<ServiceID>
    <GenTime>1994-11-05T13:15:30Z</GenTime>
    <Expiration>PT3H</Expiration>
  </MetaSrcAdv>
  <SrcAdv>
    <Type>org.nordugrid.tests.echo</Type>
    <EPR>
      <Address>http://your.domain.com/echo</Address>
    </EPR>
  </SrcAdv>
</RegEntry>
\end{verbatim}
The service type follows the GLUE2 naming convention and organize the services into categories based on their functionality. The following service types are already defined:

\begin{enumerate}
\item Storage:
\begin{itemize}
\item org.nordugrid.storage.ahash
\item org.nordugrid.storage.bartender
\item org.nordugrid.storage.librarian
\item org.nordugrid.storage.shepherd
\item org.nordugrid.storage.hopi
\end{itemize}
\item Security:
\begin{itemize}
\item org.nordugrid.security.charon
\item org.nordugrid.security.saml
\item org.nordugrid.security.slcs
\item org.nordugrid.security.delegation
\end{itemize}
\item Infosys:
\begin{itemize}
\item org.nordugrid.infosys.isis
\item org.nordugrid.infosys.eils
\item org.nordugrid.infosys.rte-catalog
\end{itemize}
\item Execution:
\begin{itemize}
\item org.nordugrid.execution.arex
\item org.nordugrid.execution.janitor
\item org.nordugrid.execution.sched
\item org.nordugrid.execution.paul
\end{itemize}
\item Accounting:
\begin{itemize}
\item org.nordugrid.accounting.mars
\end{itemize}
\item Tests:
\begin{itemize}
\item org.nordugrid.tests.echo
\item org.nordugrid.tests.echo\_java
\item org.nordugrid.tests.echo\_python
\item org.nordugrid.tests.isistest
\end{itemize}
\end{enumerate}

\subsection{Modify your configuration}
\label{Modify your configuration}
The configuration have to contain the connection between the InfoRegisters (Services) and InfoRegistrars (ISIS clouds). The schema of the configuration can be also found in the source tree (src/hed/libs/infosys/InfoRegisterConfig.xsd). There is also an example configuration between the configuration templates in the source tree (src/hed/profiles/SecureP2PIIS/SecureP2PIIS.xml).

The Service configuration should contain the InfoRegister and implicitly the InfoRegistrar configuration.
\begin{verbatim}
           <infosys:InfoRegister>
	                <infosys:Period>PT20S</infosys:Period>
	                <infosys:Endpoint>https://localhost:50000/example_service</infosys:Endpoint>
	                <infosys:Expiration>PT100S</infosys:Expiration>
	                <infosys:Registrar>
	                    <infosys:URL>some_url</infosys:URL>
	                    <infosys:KeyPath>some_local_path</infosys:KeyPath>
	                    <infosys:CertificatePath>some_local_path</infosys:CertificatePath>
	                    <infosys:CACertificatesDir>some_local_path</infosys:CACertificatesDir>
	                </infosys:Registrar>
	                <infosys:Registrar>
	                    <infosys:URL>	some_other_url</infosys:URL>
	                    <infosys:KeyPath>some_local_path</infosys:KeyPath>
	                    <infosys:CertificatePath>some_local_path</infosys:CertificatePath>
	                    <infosys:CACertificatesDir>some_local_path</infosys:CACertificatesDir>
	                </infosys:Registrar>
	            </infosys:InfoRegister>
\end{verbatim}
If the same InfoRegistrar would exist at more then one Serivce (the URL have to be the same) then the internal HED mechanism detect the similarity and tries to aggregate the messages if it's possible. You can also override the default Period, Endpoint, Expiration values inside the single Registrar elements and them too by written some value in the document in your source.

If the service configuration contains a \begin{verbatim}<NoRegister />\end{verbatim} element then it won't be registered.

### 2009-11-19 doc/manuals/janitor/tex_introduction/introduction.tex
Yet another review with a few rewordings, reformatting.

@@
 
-Janitor is a service for the automated installation of runtime environment for grid computing elements.
-It is addressed transparently via the A-REX service for job submission but can also be used as a standalone tool.
+The Janitor is a service for an automated installation of runtime
+environments (RTEs) for grid computing elements.  It is addressed
+transparently via the A-REX service for job submission but can also be
+used as a standalone tool.
 
@@
 
-A major motivation for grid projects is to stimulate new communities to adopt computational grids for their
-causes. From the current grid user's viewpoint, the admission of users of a very different education will
-suddenly impose difficulties in the communication between site maintainers. One will not even understand
-the respective other side's research aims. Hence, the proper installation of non-standard software (Runtime
-Environments) is not guaranteed and the priorities of manual labour will be mostly disjunctive.
+A major motivation for grid projects is to stimulate new communities
+to adopt the technology to start sharing their resources. From the
+current grid user's viewpoint, the admission of users of a very different
+education will suddenly impose difficulties in the communication between
+site maintainers. One will not even understand the respective other side's
+research aims. Hence, the proper installation of non-standard software
+(runtime environments, RTEs) is not guaranteed. And the available time for
+manual labour plus self-education is scarse.
 
-A core problem remains to distribute a locally working solution, the Know-How, quickly across all
-contributing sites, i.e., without manual interference. Every scientific discipline has its respective own set of
-technologies for the distribution of work load. E.g. research in bioinformatics requires access to so many
-different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
-a commodity, with all their intrinsic problems as there are bottlenecks and restrictions of repeated access.
-The EU project KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges
-with the here presented work extends the NorduGrid's
-Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
-automated installation of software packages.
+A core problem remains to distribute a locally working solution, the
+Know-How, quickly across all contributing sites, i.e., without manual
+interference. Every scientific discipline has its respective own set
+of technologies for the distribution of work load. E.g. research in
+bioinformatics requires access to so many different tools and databases,
+that few sites, if any, install them all. Instead, the use of web
+services became a commodity, with all their intrinsic problems as there
+are bottlenecks and restrictions of repeated access.  The EU project
+KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}}
+amongst other challenges with the here presented work
+extends the NorduGrid's Advanced Research Connector (ARC) grid
+middleware~\cite{ELLERT_2007} towards an infrastructure for the automated
+installation of software packages.
 
-An automation of the software installation, referred to as dynamic Runtime Environments, seems the only
-approach to use the computational grid to its full potential. Components of workflows shall be spawned
-as jobs in a computational grid using dynamic Runtime Environments rather than accessing a web service
-at one particular machine that is shared amongst all users.
-The grid introduces an extra level of parallelism that web services cannot provide. The demands for short
-response times and the heterogeneous education of site-administrators on a grid demand an automatism for
-the installation of software and databases without manual interference~\cite{BAYER_2007}.
+An automation of the software installation, referred to as dynamic
+Runtime Environments (dRTEs), seems the only approach to use the
+computational grid to its full potential. Components of workflows
+shall be spawned as jobs in a computational grid using dRTEs rather
+than accessing a web service at one particular machine that is shared
+amongst all users.  The grid introduces an extra level of parallelism
+that web services cannot provide. The demands for short response times
+and the heterogeneous education of site-administrators on a grid demand
+an automatism for the installation of software and databases without
+manual interference~\cite{BAYER_2007}.
 
@@
 
-This document first dedicates a chapter on how to set-up the Janitor locally. It is followed by a chapter that gives furhter instructions on how to 
-use the Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
-which is basically covering the method how to prepare new runtime environments. Deeper insights on the design of the Janitor will 
-be given by the forth chapter. The document ends with an outlook to anticipated future developments and opportunities.
+This document first dedicates a chapter on how to set-up the Janitor
+locally. It is followed by a chapter that gives further instructions on
+how to use the Janitor with A-REX and/or without A-REX. Afterwards, in the
+third chapter, the maintenance of the program will be presented, which is
+basically covering the method how to prepare new dRTEs. Deeper insights on
+the design of the Janitor will be given by the forth chapter. The document
+ends with an outlook to anticipated future developments and opportunities.
 

### 2009-11-19 doc/manuals/janitor/tex_installation/installation.tex
Review, reformatting.

@@
         \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
-                   Redland RDF library is used for accessing the knowledge base of Runtime Environments.}
+                   Redland RDF library is used for accessing the catalogue of Runtime Environments.}
         \label{tab:install_dependencies}
@@
    \begin{center}
-        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
+        \mycaption{Optional libraries for the Janitor.}{The library libperl-dev provides the required header files to link the 
                    WebService to the Perl interpreter.}
@@
 
-The Janitor will be installed by default along with ARC1. Support for
+The Janitor will be installed by default along with ARC-NOX. Support for
 the dynamic REs can be completely disabled at compile time by using
@@
 Furthermore it is recommended to install the ontology editor 
-Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
-knowledge database of installable packages. At the time of writing, no Linux distribution is offering packages for this fine tool.
+Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}}
+in order to be to easily maintain the knowledge database of installable packages.
+At the time of writing, no Linux distribution is offering packages for this fine tool.
 
@@
 
-The current version of the Janitor can be configured using the common file \textit{arc.conf}. It is expected in the 
-configuration directory \textit{etc}. The Janitor is using the environment variable \textit{NORDUGRID\_CONFIG}
-to determine the location of the corresponding file. If that variable is not set, the default location \textit{/etc/arc.conf} will be used. 
-The configuration is assigned by the section \lbrack janitor\rbrack. Table~\ref{tab:arc_conf_tags} describes the available tags
-for Janitor's configuration.
+The current version of the Janitor can be configured using the
+common file \textit{arc.conf}. It is expected in the configuration
+directory \textit{etc}. The Janitor is using the environment
+variable \textit{NORDUGRID\_CONFIG} to determine the location of the
+corresponding file. If that variable is not set, the default location
+\textit{/etc/arc.conf} will be used.  The configuration is assigned
+by the section \lbrack janitor\rbrack. Table~\ref{tab:arc_conf_tags}
+describes the available tags for Janitor's configuration.
 
@@
 
-The parameter \texttt{enable} specifies whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
-\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
-The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
-In these directories the states of the jobs and the runtime environments will be stored. 
-The knowledge base of installable packages is specified by the parameter \texttt{catalog}. 
-Its value can be any kind of URL pointing to a file written in the Resource Description Framework (RDF) format.
-The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
-The parameter \texttt{downloaddir} assigns the directory in which the installation files will be saved after they have been 
-downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
-directory into which all packages will be installed and should be 
-available for all computing elements, i.e.\ by using a shared volume.
-If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, the Janitor will also
-create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by the Janitor.
-The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
-The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
-The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
-or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintained by
-a higher organization.
-The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
-An example how to configure ARC and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.
+The parameter \texttt{enable} specifies whether the Janitor shall be used
+within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
+\texttt{uid} and the \texttt{gid} are defining which effective user id
+(uid) and group id (gid) shall be used for the Janitor.\footnote{For
+the time speaking, you need to have started the ARC services as root to
+allow the Janitor to change into those IDs.} The \texttt{registrationdir}
+describes the directory in which the subdirectories \texttt{jobs} and
+\texttt{rtes} will be created.  In these directories the states of the
+jobs and the runtime environments will be stored.  The knowledge base
+of installable packages is specified by the parameter \texttt{catalog}.
+Its value can be any kind of URL pointing to a file written in the
+Resource Description Framework (RDF) format.  The specification of the
+RDF file will be explained in detail in section~\ref{sec:catalog}.
+The parameter \texttt{downloaddir} assigns the directory in which
+the installation files will be saved after they have been downloaded
+or copied from the repository which was specified by the catalog. The
+\texttt{installationdir} finally specifies the directory into which all
+packages will be installed and should be available for all computing
+elements, i.e.\ by using a shared volume.  If the configuration file
+furthermore contains the \texttt{runtimedir} tag within the section
+\texttt{grid-manager}, the Janitor will also create a symbolic link
+in the \texttt{runtimedir} pointing to the configuration script of the
+installation performed by the Janitor.  The tags \texttt{jobexpirytime}
+and \texttt{rteexpirytime} are used for an automated cleanup and is
+defined in seconds.  The default value for the \texttt{jobexpirytime} is
+seven days and for the \texttt{rteexpirytime} three days.  The additional
+tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and
+\texttt{deny\_rte} are used to include or exclude certain base packages
+or runtime environments of the catalog. This feature is useful, if
+the catalog is maintained by a higher organization.  The path to the
+log4perl configuration file is defined by the tag \texttt{logconf}.
+An example how to configure ARC and log4perl is provided in the
+Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.
 
@@
 
-The Janitor is designed for the usage on Linux distributions.
+The Janitor is designed to be used on Linux distributions.
 Its porting to other platforms has not yet been addressed.

### 2009-11-19 doc/manuals/janitor/tex_titlepage/titlepage.tex
Michael's bits.
Smallish updates of names and document ID

@@
 
-{\raggedleft NORDUGRID-D2.5-2 ???\par}
+{\raggedleft NORDUGRID-MANUAL-11\par}
 
@@
 %%%%---- The title ----
-{\centering \textsc{\Large Dynamic Runtime Environment Installation with Janitor}
-\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
+{\centering \textsc{\Large Dynamic Runtime Environments with Janitor}
+\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to moeller@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
 \Large \par}

### 2009-11-19 doc/manuals/janitor/tex_usage/usage.tex
Reformatting of the usage section. Some smallish rewordings, fewer typos.

@@
 
-The Janitor can be used either with or without A-REX. In case A-REX is used, installations will be maintained automatically the runtime environments.
+The Janitor can be used either with or without the A-REX service.
+In case A-REX is used, the invocation of the Janitor will be performed in an automated manner.
+It is then triggered by incoming jobs that request a particular RTE for their execution.
+If it is not already installed, but available via the Janitor, then it will be installed
+without further manual intervention.
 
@@
 
-In addition to using the Janitor together with A-REX, the Janitor can also be used as a standalone command line tool.
+In addition to using the Janitor together with A-REX,
+the Janitor can also be used as a standalone command line tool.
 The available commands are listed in table~\ref{janitor_commandline_man}.
@@
 	\multicolumn{3}{l}{\textbf{Command:}}\\
-	&	register			& Registers a job and a set of runtime environments in the Janitor database. Requires the parameters [JOB-ID] and a list of [RTE]s.\\
-	&	deploy				& Downloads and installs the desired runtime environments. Requires the name of an already registered [JOB-ID].\\
+	&	register	& Registers a job and a set of runtime environments in the Janitor database.
+				  Requires the parameters [JOB-ID] and a list of [RTE]s.\\
+	&	deploy		& Downloads and installs the desired runtime environments.
+				  Requires the name of an already registered [JOB-ID].\\
 	&	remove		& Removes the placeholder of the job on the runtime environments.
-If no more jobs are using the runtime environment and the lifespan of the runtime environment has be expired, the runtime environment can be removed using the \texttt{sweep} command.
+				  If no more jobs are using the runtime environment and the
+				  lifespan of the runtime environment has be expired, the runtime
+				  environment can be removed using the \texttt{sweep} command.
 				  Requires the [JOB-ID] to be removed.\\
 	&			&\\
-	&	sweep				& Removes unused runtime environments.
-No further arguements are required.
-Using the option \texttt{--force} enforces the removal of all unused runtime environments.
-Runtime environments having the state FAILED will not be removed.\\
+	&	sweep		& Removes unused runtime environments. No further arguements are required.
+				  Using the option \texttt{--force} enforces the removal of all unused
+				  runtime environments. Runtime environments having the state FAILED will
+				  not be removed.\\
 	&	setstate	& Changes the state of a dynamically installed runtime environment.
-This might be useful in case a runtime environment with a state FAILED shall be removed (new state might be REMOVAL\_PENDING).
-Requires the argument [STATE] followd by a list of [RTE]s.\\
+				  This might be useful in case a runtime environment with a state FAILED
+				  shall be removed (new state might be REMOVAL\_PENDING). Requires the argument
+				  [STATE] followd by a list of [RTE]s.\\
 	&			&\\
-	&	search				& Performs a simple search in the catalog and the manually installed runtime environments (\texttt{runtimedir}).
-Requires no [JOB-ID] nor [RTE]s, but only a list of string to be searched for.\\
-	&	list				& Lists all information about jobs, automatically installed runtime environments and manually installed runtime environments. No additional parameters have to be passed.\\
+	&	search		& Performs a simple search in the catalog and the manually installed
+				  runtime environments (\texttt{runtimedir}). Requires no [JOB-ID] nor [RTE]s,
+				  but only a list of string to be searched for.\\
+	&	list		& Lists all information about jobs, automatically installed runtime
+				  environments and manually installed runtime environments.
+				  No additional parameters have to be passed.\\
 	&	info		& Renders information about a job. Requires the parameter [JOB-ID].\\
 	\multicolumn{3}{l}{\textbf{Job id:}}\\
-	&					& A unique sequence of numbers. Once Janitor registered a job id, it cannot register a second job having the same job id.\\
+	&			& A unique sequence of numbers. Once Janitor registered a job id, it
+				  cannot register a second job having the same job id.\\
 	\multicolumn{3}{l}{\textbf{Runtime environments:}}\\
 	&			& Runtime environments are defined by a continuous string.
-The name of valid runtime environment names can be investigated using the \texttt{list} or the \texttt{search} commands.
-They are defined in the catalog or by the directories and scripts of the \texttt{runtimedir} of the \texttt{grid-manager}.\\
+				  The name of valid runtime environment names can be investigated using
+				  the \texttt{list} or the \texttt{search} commands. They are defined in
+				  the catalog or by the directories and scripts of the \texttt{runtimedir}
+				  of the \texttt{grid-manager}.\\
 	\end{tabular} 
@@
 \end{table}
-The most important commands for the Janitor are \texttt{register}, \texttt{deploy} and \texttt{remove}. To 
-register a job along with a set of runtime environments in Janitor, the first command \texttt{register} followed
-by a job identifier and a list of runtime environments has to be used.
-A job is identified by a sequence of numbers. Runtime environments are specified by a string containing the name as it is defined 
-within the Catalog (resp. the runtime directory of the grid-manager).
-The command \texttt{deploy} extracts the necessary dependencies of the desired runtime environments to then download and install
-the required packages. %How dependecies can are defined will be explained later in the section~\ref{sec:catalog}.
 
-% PROCEED HERE!!!!!!!! !! ! ! !
-In order to remove jobs registered in Janitor, the command \texttt{remove}
-has to be used.  The command only removes the job entry and the lock
-on the runtime environment. If there are no more locks on the runtime
-environment it might be deleted for real. The demand to pass a job
-number for the removal of a RTE is irritating at first.  This shall
-prevent the removal of runtime envrironments that are still being used
-by jobs in the system. Instead, the janitor is informed about a job's
-termination and is requested to remove the assignment of that job to the
-runtime environment. Only those RTEs with no job-assignment are eligible
-for being sweeped. RTEs come with an expiry time or the command may be
-performed via the command line.
+The most important commands for the Janitor are \texttt{register},
+\texttt{deploy} and \texttt{remove}. To register a job along with a set
+of runtime environments in Janitor, the first command \texttt{register}
+followed by a job identifier and a list of runtime environments has to be
+used.  A job is identified by a sequence of numbers. Runtime environments
+are specified by a string containing the name as it is defined within
+the Catalog (resp. the runtime directory of the grid-manager).
+The command \texttt{deploy} extracts the necessary dependencies of the
+desired runtime environments to then download and install the required
+packages.
+
+%How dependencies can are defined will be explained later in the section~\ref{sec:catalog}.
+
+In order to remove jobs registered in the Janitor, the command
+\texttt{remove} has to be used.  The command only removes the job
+entry and the lock on the runtime environment. If there are no more
+locks on the runtime environment it is ok to be deleted also physically
+from the disk.  The demand to pass a job number for the removal of a
+RTE is irritating at first.  This shall prevent the removal of runtime
+envrironments that are still being used by jobs in the system. Instead,
+the janitor is informed about a job's termination and is requested to
+remove the assignment of that job to the runtime environment. Only those
+RTEs with no job-assignment are eligible for being sweeped. RTEs come
+with an expiry time or the command may be performed via the command line.
 
@@
 
-Every command has a certain behaviour for its exit status. The Table~\ref{tab:janitor_commandline_exit_status} lists the possible
-outcomes.
+Every command has a certain behaviour for its exit status.
+Table~\ref{tab:janitor_commandline_exit_status} lists the possible outcomes.
 

### 2009-11-19 doc/manuals/janitor/tex_maintenance/maintenance.tex
Some rewording, but otherwise a fine.

@@
 It gives detailed instructions how to create new packages for the Janitor.
-To administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e,
-as explained in the next section.
+To administrate the catalog, it is strongly recommended (but not
+ultimately required) to use the ontology editor Prot\`eg\`e, as explained
+in the next section.
 In the last section, a typical use case in maintaining the Janitor will be shown.
@@
 
-The Catalog describes runtime environments and is either served through a web server or is distributed together with the Janitor as a regular file. 
-It is specifed by a (Resource Description Framework) RDF file assigned to 
-Janitor using the tag \texttt{catalog} within the configuration file (see~\ref{sec:janitor_configuration}). 
-The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
-which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
-{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\
+The Catalog describes runtime environments and is either served
+through a web server or is distributed together with the Janitor as a
+regular file.  It is specifed by a (Resource Description Framework)
+RDF file assigned to Janitor using the tag \texttt{catalog} within
+the configuration file (see~\ref{sec:janitor_configuration}).
+The format of the RDF file is defined by an RDF schema file
+\texttt{knowarc.rdfs} which can be found along with an RDF
+example file \texttt{knowarc.rdf} in the Janitor source directory
+\href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
+{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.
 
-In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
-editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
+To edit the catalog, the ontology editor Prot\`eg\`e should be
+used. Figure~\ref{fig:protege_example} shows the editor while the
+MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has
+been selected.
+
 \begin{figure}
@@
 \end{figure}
-On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
-The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
-It has one or more instances of the class \texttt{Package} and is
-described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
-and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
-(i.e. here etch or sid), i.e. the name of a common installation or a virtual image.
-The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
-they can be found more easily.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
-They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
-classes are interacting with each other the tables~\ref{tab:main_class_meta}, \ref{tab:main_class_base}, \ref{tab:main_class_tag},
-\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.
+
+On the left side of the editor the class browser is placed.  Three
+main classes are prepared: \texttt{MetaPackage}, \texttt{Note}
+and \texttt{Package}.  The \texttt{Metapackage} is a general
+platform-independent description of a \texttt{Package}.  It has one
+or more instances of the class \texttt{Package} and is described
+by the subclasses of \texttt{Note}. The class \texttt{Note} has
+two subclasses: \texttt{BaseSystem} and \texttt{Tag} to describe
+the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the
+Debian release a \texttt{Package} refers to (i.e. here etch or
+sid), i.e. the name of a common installation or a virtual image.
+The class \texttt{Tag} provides small keywords which can be assigned
+to \texttt{MetaPackages} such that they can be found more easily.
+\texttt{TarPackage} and \texttt{DebianPackage} or currently the only
+subclasses of \texttt{Package}.  They are representing the necessary
+information (i.e. URL or Packagename) for the installation. In order
+to have an overview how the classes are interacting with each other
+the tables~\ref{tab:main_class_meta}, \ref{tab:main_class_base},
+\ref{tab:main_class_tag}, \ref{tab:main_class_debian} and
+\ref{tab:main_class_tar} are pictured.
 
@@
 
-The dynamic Runtime Environments stored in the Catalog are presented on the aforementioned dedicated web
-page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
-RDF syntax and its automated transformation to
-HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
-Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
-That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
-This HTML page listing the manually or automatically installable REs is prepared by the script web/list.pl.
-This script is meant to be run by a mod\_perl-enabled Apache. The script itself does not contribute to the 
-core functionality of the Janitor.
-In the first lines of the script some variables specific to the site are set. To configure the
-script these have to be changed~\cite[p. 9]{BAYER_2007}.
+The dynamic Runtime Environments stored in the
+Catalog are presented on the aforementioned dedicated web
+page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also
+links to both the formal Catalog in RDF syntax and its automated
+transformation to HTML.  The latter mimics the traditional
+site describing Runtime Environments in the Runtime Environment
+Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with
+an eventual transition to the new system.  That page collects descriptions
+for Runtime Environments to encourage human site administrators to
+install these.  This HTML page listing the manually or automatically
+installable REs is prepared by the script web/list.pl.  This script is
+meant to be run by a mod\_perl-enabled Apache. The script itself does
+not contribute to the core functionality of the Janitor.  In the first
+lines of the script some variables specific to the site are set. To
+configure the script these have to be changed~\cite[p. 9]{BAYER_2007}.
 
@@
 
-This section describes how to add new packages to the Catalog. In the current implementation, only tar based packages are processed by the Janitor.
-Within the example they are assigned to be used together with Debian Etch. This limitation is only literal. 
-There is no limitiation to newer Debian distributions.
+This section describes how to add new packages to the Catalog. In the
+current implementation, only tar based packages are processed by the
+Janitor.  Within the example they are assigned to be used together with
+Debian Etch. This limitation is only literal.  There is no restriction
+for newer Debian distributions.
 
@@
 
-At the time of writing, only the tape archive (tar) file format is accepted for dynamic Runtime Environment installation, 
-a well accepted file format throughout the UNIX community. The reflects the traditional manual approach towards RTE in ARC,
-for which one directory is made available to all compute nodes. This section explains the inner structure of the tar files. 
-Subdirectories are visualised in Figure~\ref{fig:tar_folder}.
+At the time of writing, only the tape archive (tar) file format is
+accepted for dynamic Runtime Environment installation, a well accepted
+file format throughout the UNIX community. The concept reflects the
+traditional manual approach towards RTE in ARC, for which one directory
+is made available to all compute nodes. This section explains the
+inner structure of the tar files.  Subdirectories are visualised in
+Figure~\ref{fig:tar_folder}.
 
@@
 
-The tar file contains two directories, named {\tt control} and {\tt data}. Software is stored in the latter subdirectory,
-while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
-the content of the data directory is extracted to some directory \textdollar BAR. 
-After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
-It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
-The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
-before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
-Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
-installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
-Once the tar file was prepared, it must its entry to a RTE Catalogue~\cite[p. 10]{BAYER_2007}.
+The tar file contains two directories, named {\tt control} and {\tt
+data}. Software is stored in the latter subdirectory, while the files
+formally specifying how to deal with such packages are stored in the
+prior. Upon installation, the content of the data directory is extracted
+to some directory \textdollar BAR.  After this unpacking of the tar file,
+the Janitor executes the install script provided in the control directory.
+It is executed within the working directory \textdollar BAR. The job of
+this skript is to perform any necessary post-processing.  The Janitor
+stores the file \texttt{control/remove}. It will be executed in the
+same way as \texttt{control/install} just before the tar-package is
+removed. In most cases \texttt{control/remove} will be empty.  Finally,
+the file \texttt{control/runtime} is sourced multiple times by the
+Grid Manager's job-submit script. After installing the package, the
+Janitor changes all occurences of \%BASEDIR\% in the runtime script to
+\textdollar BAR.  Once the tar file was prepared, it must its entry to
+a RTE Catalogue~\cite[p. 10]{BAYER_2007}.
 
-From such Catalogues, the Janitor finds all information to install packages that possibly have never been installed
-on the site before. The offers of RTEs in a catalogue are cross-checked against the local infrastructure and a subset
-of the available packages will be accepted as "installable". This list of installable RTEs is forwarded to the
-grid information system.
+From such Catalogs, the Janitor finds all information to install
+packages that possibly have never been installed on the site before. The
+offers of RTEs in a Catalog are cross-checked against the local
+infrastructure and a subset of the available packages will be accepted as
+"installable". This list of installable RTEs is forwarded to the grid
+information system.
 
@@
 
-The WEKA package for machine learning~\cite{FRANK_20074} and the Java Runtime Environment
-are available as dynamic Runtime Environments. Further packages for bioinformatics comprise dynamic
-variants of tools for the analysis of transcription factor binding sites. These are already offered for manual
-installation via the prior mentioned traditional page representing Runtime Environments for ARC. 
-The corresponding tar file is named \task{somewhat??}. The \texttt{data} directory simply contains a ZIP file
-which needs to be unzipped in the installation directory. For that reason, the \texttt{control/install} script is 
-written as follows:
+The WEKA package for machine learning~\cite{FRANK_20074} and the Java
+Runtime Environment are available as dynamic Runtime Environments. Further
+packages for bioinformatics comprise dynamic variants of tools for
+the analysis of transcription factor binding sites. These are already
+offered for manual installation via the prior mentioned traditional page
+representing Runtime Environments for ARC.  The corresponding tar file
+is named \task{somewhat??}. The \texttt{data} directory simply contains
+a ZIP file which needs to be unzipped in the installation directory. For
+that reason, the \texttt{control/install} script is written as follows:
+
 % dollar signs made trouble in listings *hrmlgrmpf* ... Costs too much time to figure out why *ARC!*
@@
 \end{verbatim}
-The remove script, which will be executed right before WEKA is deinstalled, is empty. Janitor will delete the directory, so there
-is nothing more to be done.\\
+
+The remove script, which will be executed right before WEKA is deinstalled, is empty.
+Janitor will delete the directory, so there is nothing more to be done.
 
@@
 
-To address the concerns of the physicists using ARC, a dynamic runtime environment for the
-ATLAS software suite was prepared. It extends prior work on an automated installation that is available at
-\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}. The preparation comprised the following steps:
+To address the concerns of the physicists using ARC, a dynamic runtime
+environment for the ATLAS software suite was prepared. It extends prior
+work on an automated installation that is available at
+\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}.
+The preparation comprised the following steps:
 \begin{itemize}
-    \item The file system path specifications in the automated installation scripts were modified using the Janitor
+    \item The file system path specifications in the automated
+       installation scripts were modified using the Janitor
        path variables.
-    \item A tarball was prepared containing a directory structure as illustrated in Figure~\ref{fig:tar_folder}. The data directory
-       was empty, since the automatic installation script downloads the software from a remote server.
+    \item A tarball was prepared containing a directory structure as
+       illustrated in Figure~\ref{fig:tar_folder}. The data directory
+       was empty, since the automatic installation script downloads the
+       software from a remote server.
     \item An entry was added to the Catalog file.
@@
 
-What sets High Energy Physics software apart is it's sheer size. The package in question takes up more than
-5 GB. This was a test illustrating the feasibility of using dynamic REs in High Energy Physics. The application
-of the dynamic RE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
-service, e.g. a software manager of a big experiment will be able to deploy software packages on production
-sites simply by creating a tarball and adding an entry to the Catalog.
+What sets High Energy Physics software apart is it's sheer size. The
+package in question takes up more than 5 GB. This was a test illustrating
+the feasibility of using dynamic REs in High Energy Physics. The
+application of the dynamic RE for ATLAS needs to wait for the planned web
+service extension of the Catalog. With such a service, e.g. a software
+manager of a big experiment will be able to deploy software packages
+on production sites simply by creating a tarball and adding an entry to
+the Catalog.
 
@@
 
-\section{Typical use cases}\label{sec:catalog}
+%\section{Typical use cases}\label{sec:catalog}
 

### 2009-11-19 doc/tech_doc/a-rex/arex_tech_doc.tex
correct location of cache tools
add passivetransfer option

@@
 
-The following tools (installed in \emph{\$ARC\_LOCATION/libexec})
+The following tools (installed in \emph{\$ARC\_LOCATION/libexec/arc})
 exist to help with administration of the cache:
@@
 URL options.
+\item \textbf{\textit{passivetransfer}}\textit{=yes|no} -- specifies
+  whether GridFTP transfers are passive. Setting this option to yes
+  can solve transfer problems caused by firewalls. Default is no.
+
 \item \textbf{\textit{localtransfer}}\textit{=yes|no} -- specifies whether

### 2009-11-20 doc/manuals/user_interface/ui-job.tex
a few typos corrected

@@
 
-Default configuration file is {\$}HOME/.arc/client.conf. However, a user can chose any
+Default configuration file is {\$}HOME/.arc/client.conf. However, a user can choose any
 other pre-defined configuration through option \verb#-z#.
@@
 
-Only the results of jobs that have finished can be downloaded. Just like in \verb#arstat# 
+Only the results of jobs that have finished can be downloaded. Just like in \verb#arcstat# 
 and \verb#arccat# cases, the job can be referred to either by the \texttt{jobID} that was returned by

### 2009-11-20 doc/manuals/user_interface/ui-proxy.tex
Interchange DEBUG and VERBOSE loglevels.
a few typos corrected

@@
 (submit jobs, copy data, check information etc),
-one has to present valid credentials. The are commonly formalized as so-called
+one has to present valid credentials. These are commonly formalized as so-called
 ``proxy'' certificates. There are many different types of proxy certificates,
@@
    \texttt{-t, --timeout}& \textit{seconds} & timeout in seconds (default 20 seconds)\\
-   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
    \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
@@
    \texttt{-t, --timeout}& \textit{seconds} & timeout in seconds (default 20 seconds)\\
-   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
+   \texttt{-d, --debug}& \textit{debuglevel}&debug level is one of  FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG\\
    \texttt{-c, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\

### 2009-11-23 doc/manuals/janitor/tex_installation/installation.tex
Added warning about non-existing directories.

@@
 	   registrationdir & "/var/spool/nordugrid/janitor"        & Directory where we the current states of jobs are kept. \\
-	   catalog         & "/var/spool/nordugrid/janitor/catalog/knowarc.rdf"& URL of the catalog containg the package information.\\
+	   catalog         & "/var/spool/nordugrid/janitor/catalog/knowarc.rdf"& URL of the catalog containing the package information.\\
 	   downloaddir     & "/var/spool/nordugrid/janitor/download" & Directory for downloads \\
@@
 \end{landscape}
+
+It should be noted that the downloaddir or the installationdir specified
+in arc.conf could be any directory. Those will not be prepared by the
+package for the Linux distribution but need be created by the administrator
+manually after the Janitor has been installed. This also holds for the
+catalog.
 

### 2009-11-25 doc/manuals/janitor/tex_concept/concept.tex
Michael's bits.
Series of updates, need to verify the janitor webservice, still. Please merge.

@@
-\chapter{Programming concept}
+\chapter{Technical Motivation and Concepts}
+
+The current implementation, using the rather advanced concepts of the
+semantic web, may seem unexpected. They were chosen since they are expected
+to scale with all the extra demands forseen for the Janitor.
+
+\section{Implementation}
+
+\subsection{Choice of Perl}
 
  The main language for the implementation of the functionality of the
- dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
- integration with the Grid Manager and the Web server) for the Janitor. In
+ dRTE's functionality is Perl. And it is solely required (exceptions are the
+ integration with the Grid Manager and the Web server) for the Janitor.
+ The language was perceived a side-issue, the complexity of the data
+ expected was of major concern.
+
+\subsection{Modular structure}
+
+ In
  the pre-web-service implementation the Catalog remains a static web page.
@@
 
-\section{States of runtime environments}
+\section{Concepts}
+
+In the following, some paragraphs fail render it difficult to 
+clearly distinguish between conceptional truths and the state
+of the current implementation. Please read carefully.
+
+\subsection{States of runtime environments}
+
 A major motivation for the managed, manual initiation of dynamic RE
 installation is the subsequent manual verification of the installed
-packages -- prior to their use in production.
+packages -- prior to their use in production. It would be nice to
+see the Janitor and/or the Catalog prepare for reviews by selected
+users. This has not yet been implemented.
+
 With an automation of the installation, the verification of
@@
 
+\subsection{Subset of that functionality as implemented today}
 Upon presentation of a the package name to a Catalog, from which details about
@@
 
-\section{Janitor file system permissions}
-
-In the current implementation the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
-a setuid wrapper has been prepared. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s},
-i.e. the sticky bit is set to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\
-
-Thus, Janitor is not executed directly. Instead, a dynamic link will be created in \task{libexec or sbin} which is pointing to
-the wrapper.\\
-
-It is suggested to create a user and a group "janitor" for the Janitor.\\
+%FIXME: needs further communication with Anders
+%\section{Janitor file system permissions}
+%
+%In the current implementation the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
+%a setuid wrapper has been prepared. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s},
+%i.e. the sticky bit is set to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\
+%
+%Thus, Janitor is not executed directly. Instead, a dynamic link will be created in \task{libexec or sbin} which is pointing to
+%the wrapper.\\
+%
+%It is suggested to create a user and a group "janitor" for the Janitor.\\
 
@@
 % % \texttt{\textdollar chown janitor:janitor $<$shared_volume$>$/janitor}
+
+\subsection{What happens during installation}
+
+Register
+\begin{enumerate}
+ \item Look up Catalog and find corresponding runtime environments.
+ \item Check if dependcies are available (installed or installable)
+ \item Store job in \textit{registrationdir}
+\end{enumerate}
+
+Deploy
+\begin{enumerate}
+ \item Load job out of \textit{registrationdir}
+ \item Look up Catalog and find corresponding runtime environments.
+ \item Check if dependcies are available (installed or installable)
+ \item Download RTEs into the \textit{downloaddir}
+ \item $\lbrack$ to be continued or skipped $\rbrack$
+\end{enumerate}
 

### 2009-11-25 doc/manuals/janitor/tex_future/future.tex
Series of updates to the Janitor documentation.
Michael's comments on future developments on the Janitor.
Series of updates, need to verify the janitor webservice, still. Please merge.

@@
 
-\section{Representation of dynamic REs in the information model}
-Dynamic REs require an extended representation in the
+\section{Representation of dynamic RTEs in the information model}
+Dynamic RTEs require an extended representation in the
 information model.  The Application Software description should be able
-to distinguish installed REs from installable REs, potentially offer
-description of extended RE state-like information. This work is planned
+to distinguish installed RTEs from installable RTEs, potentially offer
+description of extended RTE state-like information. This work is planned
 to be carried out as part of the Glue-2.0 effort of OGF\footnote{OGF
@@
 workflow tools for the web services that have a growing user base in
-bioinformatics. The challenge is to prepare REs for
+bioinformatics. The challenge is to prepare RTEs for
 programs or databases and to offer such concisely to users of the
@@
 Conversely, because of the increased complexity of workflows with respect
-to the already today not manually manageable number of REs, without an
+to the already today not manually manageable number of RTEs, without an
 automatism for the automated installation of software packages on the
@@
 The RDF schema nicely prepares for the upcoming virtualisation of worker
-nodes.  How exactly the dynamics are integrated will depend on how
+nodes. Hereto, the {\tt BaseSystem} indicates a virtual image to which
+further packages, the dynamic REs, would then be added.
+How exactly the dynamics are integrated will depend on how
 dynamic the virtualisation of the nodes is. In the simplest scenario,
@@
 
-TODO:\\
-\begin{itemize}
- \item There is a arc.conf file in which all possiblie flags are listed... ADD THE JANITOR FLAGS!!!
- \item Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change that
- \item Verfication
-\end{itemize}
+%TODO:\\
+%\begin{itemize}
+% \item There is a arc.conf file in which all possiblie flags are listed... ADD THE JANITOR FLAGS!!!
+% \item Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change that
+% \item Verfication
+%\end{itemize}
+
+%(1) The HostInfoJanitor.pm queries the catalog for available RTE names.
+%    It is fast, such that I assume a fork or a cache is currently not needed. 
+%    > Nevertheless I installed an alarm of 15 Seconds
+
+%(2) In order to enable the Grid-Manager to invoke Janitor, a wrapper rJanitor.c has
+%    (already) been written (by Daniel). 
+%    This wrapper gives Janitor the necessary rights using a suid bit. 
+%    >It is installed and used but not yet tested well
+%    >In case Daniel is right, it will work
+
+% (3) > ... ?? hmm... well seems there is some information loss here... point 3... suppose it must have been important.
+
+%(4) The function "check" (nowadays ``info'') was meant to return the information about a runtime 
+%      environment. It may possible to merge with the deploy functionality 
+%      if desired. (The Janitor source code will remain the same, only the interface will 
+%      be changed.)
+%      >Since this is mere beautification, this is postponed
+
+%(5) The information about the runtime environments is passed to the backend
+%    script. (It is not specifed yet how this has to be done)
+%       o Installation directories
+%       o Script to set environment variables
+%    > Currently a ``deploy'' action will set symbolic links for *ALL* dynamically installed
+%    > METAPACKAGES!  That means, nothing has to be done here. The backends script won't (hopefully)
+%    > except the symbolic links and execute the corresponding script delievered along with the tarball 
+%    > package. 
+%    > The symbolic link will be removed respictvely on ``remove''
+%
+%    > TODO: Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change *that*
+%
+%    > TODO: Think about the user rights of the symbolic link...
+
+%(6) Once the grid job has been finished, the token on the runtime environment will
+%      be removed. (If no tokens left, the environment may be deinstalled)
+%    > Well, that's nothing new for me, but maybe nice to know :-)
+
+%(7) GLUE2 specification: Balazs like to have an interface for Janitor. This task is 
+%   put on the agenda. It is still unclear how the interaction with that interface shall be 
+%   realised. Janitor can only access data concerning the class ApplicationEnvironment.
+%   Shall a XML snippet be returned?
+%  > Alright now, almost 1.5 month later I know what he asked for.
+%  >
+%  > Please examine following files in src/services/a-rex/infoproviders/:
+%    o ARC1ClusterInfo.pm:~744
+%    o HostInfo.pm:~218
+%    o HostInfoJanitor.pm
+%  >
+%  > The HostInfo.pm collects a list of manually and dynamically (HostInfoJanitor.pm) installed RTE *NAMES*.
+%  > ARC1ClusterInfo.pm adds some hard coded information (description, states, etc.) to it and puts it into the schema.
+%  >
+%  > WHAT HAS TO BE DONE?
+%  >
+%  > HostInfo.pm has to create a *HASH MAP* with must have as much information as possible (description, states, etc.).
+%  > Therefore HostInfoJanitor.pm has to be extended to gather more information out of the catalog.
+%  > ARC1ClusterInfo.pm needs to be stultified such that the *HASH MAP* received from HostInfo.pm will simply mapped
+%  > to the schema.
+
+%(8) Verfication
+%    o Of validity of tar-package (Store the catalog somewhere safe! Hash-value database?)
+%    o Of the executer of Janitor (This can be done using a WEB SERVICE!!! )
+%    o TODO: I experienced some problems in case some shell tools are not installed on clusters
+%            i.e.  unzip
+%            There should be some kind of check or RESTRICTION to what can be used
+
+%(9) The missing feature for dependencies 
+%    Assume, one wants to install package A which depends on package B.
+%    Both are getting installed dynamically.
+%    Right before job execution, the runtime scripts of package A is executed. 
+%    Still missing is the execution of the runtime script of B!
+%
+%    Simple workaround: Request package B in job description too
+%    Correct solution:  Dependencies have to be returned by Janitor and added to the job.description file!!!!!
+%                       (Janitor info 21 or *EXTEND* Janitor deploy + Get Janitor output in states.cpp and put
+%			it into the description file)

### 2009-11-25 doc/manuals/janitor/tex_maintenance/maintenance.tex
Series of updates, need to verify the janitor webservice, still. Please merge.

@@
 
+\subsection{Automated transformation of install directory to dRTE}
+
+The script 'prepareDRE.pl' was created to help with the transformation
+of a readily installed software into a dynamic runtime environment.
+I also prepares a complete catalog file that can be offered
+individually or next to other catalogs. See the associated man page
+prepareDRE(8) for details.
+
 \subsection{Protoypes}

### 2009-11-26 doc/manuals/python/SWIG-wrapped-ARC-Python-API.tex
Interchange DEBUG and VERBOSE loglevels.
Updated according to changes in code.

@@
 %%%%---- The title ----
-{\centering \textsc{\Large The SWIG wrapped ARC Python API and the arcom utility package\footnote{as of rev. 13724}}\Large \par}
+{\centering \textsc{\Large The SWIG wrapped ARC Python API and the arcom utility package}\Large \par}
 \vspace*{0.5cm}
@@
 \begin{itemize}
+  \item{ \begin{verbatim}
+#ifdef SWIGPYTHON
+namespace Arc {
+
+%typemap(in, numinputs=0) std::string& content (std::string str) { 
+  $1 = &str; 
+}
+\end{verbatim}
+  If target language is Python then this input typemap tells SWIG that ''content'' 
+  arguments of type ''\verb$std::string&$'' found in ''Arc'' namespace are to be ignored. 
+  As stated in the SWIG documentation: ''When numinputs is set to 0, the argument is 
+  effectively ignored and cannot be supplied from the target language. '' 
+  However the argument is still required when making the C/C++ call. This is solved 
+  by providing a locally declared variable called str from which the value used is 
+  obtained. Statement ''\verb#$1 = &str;#'' sets the input argument to point to this 
+  temporary variable.
+\begin{verbatim}
+%typemap(argout) std::string& content { 
+  PyObject *tuple; 
+  tuple = PyTuple_New(2); 
+  PyTuple_SetItem(tuple,0,$result); 
+  PyTuple_SetItem(tuple,1,Py_BuildValue("s",$1->c_str())); 
+  $result = tuple; 
+} 
+\end{verbatim}
+  As the SWIG documentation tells: ''The \verb$"argout"$ typemap is used to return values from arguments.''
+  With the previous typemap in mind this one indicates that the result will be changed as follows:
+    \begin{itemize}
+    \item{
+      A tuple will be created with two members.
+    }
+    \item{
+      The first member comes from the original result.
+    }
+    \item{
+      The second one comes from the argument.
+    }
+    \item{
+      The tuple will be returned as the new result.
+    }
+    \end{itemize}
+  }
+\end{itemize}
+\begin{itemize}
   \item{
@@
 \begin{itemize}
+  \item{ \verb$%include <std_vector.i>$} \linebreak
+  The \verb$std_vector$ library will be used.
+\end{itemize}
+\begin{itemize}
   \item{\begin{verbatim}
@@
   %template(URLList) std::list<Arc::URL>;
+  %template(URLVector) std::vector<Arc::URL>;
+  %template(URLListMap) std::map< std::string, std::list<Arc::URL> >;
   \end{verbatim}
@@
   Define XMLNodeList as a template list of Arc::XMLNode and URLList as a template list of 
-  Arc::URL objects. (See Example\ref{listex} for an example of using a list of objects - 
+  Arc::URL objects. Define URLVector as a template vector of Arc::URL objects and URLListMap 
+  as a template map where the key is a string and the value is a list of Arc::URL objects. 
+  (See Example\ref{listex} for an example of using a list of objects - 
   list of strings in that particular case - in Python.)\linebreak
@@
   %template(ApplicationEnvironmentList) std::list<Arc::ApplicationEnvironment>;
-  %template(SoftwareVersionList) std::list<Arc::SoftwareVersion>;
+  %template(SoftwareList) std::list<Arc::Software>;
   %template(SoftwareRequirementList) std::list<Arc::SoftwareRequirement>;
+  %template(ResourceTargetTypeList) std::list<Arc::ResourceTargetType>;
   \end{verbatim}
@@
   Arc::DirectoryType, ApplicationEnvironmentList as a template list of 
-  Arc::ApplicationEnvironment, SoftwareVersionList as a template list of 
-  Arc::SoftwareVersion and SoftwareRequirementList as a template list of 
-  Arc::SoftwareRequirement objects. \linebreak
+  Arc::ApplicationEnvironment, SoftwareList as a template list of 
+  Arc::Software, SoftwareRequirementList as a template list of 
+  Arc::SoftwareRequirement and ResourceTargetTypeList as a template list of 
+  Arc::ResourceTargetType objects. \linebreak
   (See Example\ref{listex} for an example of using a list of objects in Python.)
@@
   \end{verbatim}
-  As the SWIG documentation tells: ''The \verb$"argout"$ typemap is used to return values from arguments.''
-  With the previous typemap in mind this one indicates that the result will be changed as the following:
+  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as follows:
     \begin{itemize}
@@
 \end{itemize}
-\begin{itemize}
-  \item{ \begin{verbatim}
-%extend ACCLoader {
-  Arc::Broker * getBroker(const std::string& brokername) { 
-      return dynamic_cast<Arc::Broker*>( self->getACC(brokername) ); 
-  }
-  Arc::Broker * loadBroker(const std::string& brokername, const XMLNode * cfg) { 
-      return dynamic_cast<Arc::Broker*>( self->loadACC(brokername, cfg) ); 
-  }
-  Arc::JobController * getJobController(const std::string& jobcontrollername) { 
-      return dynamic_cast<Arc::JobController*>( self->getACC(jobcontrollername) ); 
-  }
-  Arc::JobController * loadJobController(const std::string& jobcontrollername, 
-                                         const XMLNode * cfg) { 
-      return dynamic_cast<Arc::JobController*>( 
-                                         self->loadACC(jobcontrollername, cfg) ); 
-  }
-}
-  \end{verbatim}
-  SWIG methods added to get/load Broker and JobController objects from an ACCLoader object.
-  }
-\end{itemize}
+% \begin{itemize}
+%   \item{ \begin{verbatim}
+% %extend ACCLoader {
+%   Arc::Broker * getBroker(const std::string& brokername) { 
+%       return dynamic_cast<Arc::Broker*>( self->getACC(brokername) ); 
+%   }
+%   Arc::Broker * loadBroker(const std::string& brokername, const XMLNode * cfg) { 
+%       return dynamic_cast<Arc::Broker*>( self->loadACC(brokername, cfg) ); 
+%   }
+%   Arc::JobController * getJobController(const std::string& jobcontrollername) { 
+%       return dynamic_cast<Arc::JobController*>( self->getACC(jobcontrollername) ); 
+%   }
+%   Arc::JobController * loadJobController(const std::string& jobcontrollername, 
+%                                          const XMLNode * cfg) { 
+%       return dynamic_cast<Arc::JobController*>( 
+%                                          self->loadACC(jobcontrollername, cfg) ); 
+%   }
+% }
+%   \end{verbatim}
+%   SWIG methods added to get/load Broker and JobController objects from an ACCLoader object.
+%   }
+% \end{itemize}
 \begin{itemize}
@@
   \end{verbatim}
-  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as the following:
+  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as follows:
     \begin{itemize}
@@
 \end{verbatim} See Example\ref{url2dp} for details.
-    \end{example}
-  }
-\end{itemize}
-\begin{itemize}
-  \item{ \begin{verbatim}
-  %typemap(in, numinputs=0) std::string& failure_description (std::string temp) {
-    $1 = &temp;
-  }
-  \end{verbatim}
-  This input typemap tells SWIG that \verb$failure_description$ arguments of type \verb$std::string&$
-  are to be ignored. However the argument is still required when making the C/C++ call. 
-  This is solved by providing a locally declared variable called temp from which the value 
-  used is obtained from. Statement ''\verb#$1 = &temp;#'' sets the input argument to point to this temporary variable.
-  }
-\end{itemize}
-\begin{itemize}
-  \item{ \begin{verbatim}
-  %typemap(argout) std::string & failure_description {
-    PyObject *o, *tuple;
-    tuple = PyTuple_New(2);
-    o = PyString_FromString($1->c_str());
-    PyTuple_SetItem(tuple,0,$result);
-    PyTuple_SetItem(tuple,1,o);
-    $result = tuple;
-  }
-  \end{verbatim}
-  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as the following:
-    \begin{itemize}
-    \item{
-      A tuple will be created with two members.
-    }
-    \item{
-      The first member comes from the original result.
-    }
-    \item{
-      The second one comes from the argument.
-    }
-    \item{
-      The tuple will be returned as the new result.
-    }
-
-  One possible use of \verb$failure_description$ could be the Transfer funcionality of the 
-  DataMover class. This returns a failure description along with the corresponding 
-  DataStatus. This description may contain more details about the error if transfer 
-  fails for some reason.
-  (See example below.)
-    \end{itemize}
-    \begin{example}
-      \caption{DataMover - Transfer}\label{dmex}
-\begin{verbatim}
->>> import arc
->>> # create DataMover
-... mover = arc.DataMover()
->>> # import datapoint_from_url from arcom package
-... # DataPoints are required to use DataMover
-... from arcom import datapoint_from_url
->>> # create source DataPoint
-... src = datapoint_from_url('file://./file1')
->>> # create destination DataPoint
-... dst = datapoint_from_url('file://./file2')
->>> # do the transfer
-... status, failure = mover.Transfer(src, dst, arc.FileCache(), arc.URLMap())
->>> # check status
-... str(status)
-'Operation completed successfully'
->>> # check for failure message
-... failure
-''
-\end{verbatim}
     \end{example}
@@
   }
+\end{itemize}
+\begin{itemize}
+  \item{ \begin{verbatim}
+    %template(ScalableTimeInt) Arc::ScalableTime<int>;
+    %template(RangeInt) Arc::Range<int>;
+    %template(RangeInt64) Arc::Range<int64_t>;
+  \end{verbatim}
+  }
+  These are template instantiations for template classes \verb$Arc::ScalableTime<int>$, 
+  \verb$Arc::Range<int>$ and \verb$Arc::Range<int64_t>$.
 \end{itemize}
@@
 \begin{itemize}
-  \item{ \verb$datapoint_from_url(url_string)$\newline
-  Creates DataPoint from specified URL (\verb$url_string$).
+  \item{ \verb$datapoint_from_url(url_string, ssl_config = None)$\newline
+  Creates DataPoint from specified URL (\verb$url_string$) with the 
+  specified SSL configuration (if \verb$ssl_config$ is present).
     \begin{example}
@@
     \begin{itemize}
+      \item{arc.DEBUG}
       \item{arc.VERBOSE}
-      \item{arc.DEBUG}
       \item{arc.INFO}
@@
   this example to work because the incoming message in our example could not be processed
-  in a way like on the server side. Therefore it is assumed that the \verb$__init__$ method
+  in a way like on the server side.\label{ModAuthReq} Therefore it is assumed that the \verb$__init__$ method
   of AuthRequest has:
@@
 ... se.NewChild('dmmy:foo').Set('Foo!')
+(None, 'Foo!')
 >>> # Create SOAP Message
@@
 >>> outpls = arc.PayloadSOAP(outse)
->>> outmsg.Payload(pls)
+>>> outmsg.Payload(outpls)
 >>> # call service
 ... ds.process(msg, outmsg)
-<arc.MCC_Status; proxy of <Swig Object of type 'Arc::MCC_Status *' at 0x8a39078> >
+<arc.MCC_Status; proxy of <Swig Object of type 'Arc::MCC_Status *' at 0x9cd8228> >
 >>> # show result
@@
 Response></soap-env:Body></soap-env:Envelope>
+\end{verbatim}
+    \end{example}
+  }
+\end{itemize}
+
+\begin{itemize}
+  \item{ \verb$ServiceState$ \linebreak
+  ServiceState is part of the basic LIDI\footnote{Local Information Description Interface; see: 
+  \url{http://www.knowarc.eu/documents/Knowarc_D1.2-1_07.pdf}} for Python services. This basic 
+  implementation provides information about whether a service is up and running. XML 
+  representation of state could be retrieved by calling the ''GetLocalInformation'' method 
+  found in the Service class. This will be presented through an example, in which DummyService 
+  will provide information regarding its state. (Complete source code can be found in Appendix 
+  \ref{DummyServiceCode} of this document.)\linebreak
+  Note that a change in AuthRequest class (found in ''security.py'') is required for 
+  this example to work (see \ref{ModAuthReq}).
+    \begin{example}
+      \caption{Retrieving state of DummyService}\label{RetDSstate}
+\begin{verbatim}
+>>> # Reminder:
+... # import arc
+... #
+... # ...
+... # 
+... # from arcom.service import Service
+... # 
+... # class DummyService(Service):
+... #   """ DummyService class based on Service class in arcom.service """
+... # 
+... #   ...
+... # 
+... #   def status(self, inpayload):
+... #       # get local info
+... #       fooinfo = self.GetLocalInformation()
+... #       
+... #       # get service status
+... #       foostatus = str(fooinfo.Get('AdminDomain').Get('Services').Get('Service').\
+... #       Get('Endpoint').Get('ServingState'))
+... #       ...
+... 
+>>> # Create instance of DummyService
+... ds = DummyService(None)
+>>> # Create namespace - it will be used for the message sent
+... ns = arc.NS({'dmmy':'urn:foo'})
+>>> # Create SOAP Envelope
+... se = arc.SOAPEnvelope(ns)
+>>> # Create content
+... se.NewChild('dmmy:status')
+<arc.XMLNode; proxy of <Swig Object of type 'Arc::XMLNode *' at 0x9cb8af8> >
+>>> # Create SOAP Message
+... msg = arc.SOAPMessage()
+>>> # Payload
+... pls = arc.PayloadSOAP(se)
+>>> # Set Message Payload
+... msg.Payload(pls)
+>>> # create outgoing message that will hold response
+... outse = arc.SOAPEnvelope(ns)
+>>> outmsg = arc.SOAPMessage()
+>>> outpls = arc.PayloadSOAP(outse)
+>>> outmsg.Payload(outpls)
+>>> # call service
+... ds.process(msg, outmsg)
+<arc.MCC_Status; proxy of <Swig Object of type 'Arc::MCC_Status *' at 0x9cdb090> >
+>>> # show result
+... print outmsg.Payload().GetXML()
+<soap-env:Envelope xmlns:dmmy="urn:foo" xmlns:soap-enc="http://schemas.xmlsoap.o
+rg/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xm
+lns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XML
+Schema-instance"><soap-env:Body><dmmy:statusResponse>Status: production</dmmy:st
+atusResponse></soap-env:Body></soap-env:Envelope>
 \end{verbatim}
@@
     # set name of provided method
-    self.request_names = ['foo']
+    self.request_names = ['foo','status']
 
@@
     response_node.Set(foomsg)
+
+    return out
+
+  def status(self, inpayload):
+    # get local info
+    dummyinfo = self.GetLocalInformation()
+
+    # get service status
+    dummystatus = str(dummyinfo.Get('AdminDomain').Get('Services').Get('Service').\
+    Get('Endpoint').Get('ServingState'))
+
+    # create response
+    out = self._new_soap_payload()
+    response_node = out.NewChild('dmmy:statusResponse')    
+    response_node.Set('Status: ' + dummystatus)
 
@@
 \input{app/ex-2-1-4.tex}
-\input{app/ex-2-1-5.tex}
 \input{app/ex-3-1.tex}

### 2009-11-26 doc/manuals/python/app/ex-3-5.tex
Updated according to changes in code.

@@
 outpls = arc.PayloadSOAP(outse)
-outmsg.Payload(pls)
+outmsg.Payload(outpls)
+# call service
+ds.process(msg, outmsg)
+# show result
+print outmsg.Payload().GetXML()
+\end{verbatim}
+
+\subsection{Retrieving state of DummyService}
+(Example \ref{RetDSstate})
+\label{cRetDSstate}
+\begin{verbatim}
+# Reminder:
+# import arc
+#
+# ...
+# 
+# from arcom.service import Service
+# 
+# class DummyService(Service):
+#   """ DummyService class based on Service class in arcom.service """
+# 
+#   ...
+# 
+#   def status(self, inpayload):
+#       # get local info
+#       fooinfo = self.GetLocalInformation()
+#       
+#       # get service status
+#       foostatus = str(fooinfo.Get('AdminDomain').Get('Services').Get('Service').\
+#       Get('Endpoint').Get('ServingState'))
+#       ...
+
+# Create instance of DummyService
+ds = DummyService(None)
+# Create namespace - it will be used for the message sent
+ns = arc.NS({'dmmy':'urn:foo'})
+# Create SOAP Envelope
+se = arc.SOAPEnvelope(ns)
+# Create content
+se.NewChild('dmmy:status')
+# Create SOAP Message
+msg = arc.SOAPMessage()
+# Payload
+pls = arc.PayloadSOAP(se)
+# Set Message Payload
+msg.Payload(pls)
+# create outgoing message that will hold response
+outse = arc.SOAPEnvelope(ns)
+outmsg = arc.SOAPMessage()
+outpls = arc.PayloadSOAP(outse)
+outmsg.Payload(outpls)
 # call service

### 2009-11-26 doc/manuals/janitor/tex_installation/installation.tex
Update for explaining how to list multiple catalog files.
fix typo

@@
 
-It should be noted that the downloaddir or the installationdir specified
-in arc.conf could be any directory. Those will not be prepared by the
-package for the Linux distribution but need be created by the administrator
-manually after the Janitor has been installed. This also holds for the
-catalog.
-
-The parameter \texttt{enable} specifies whether the Janitor shall be used
+The parameter \texttt{enabled} specifies whether the Janitor shall be used
 within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
@@
 
+It should be noted that the downloaddir or the installationdir specified
+in arc.conf could be any directory. Those will not be prepared by the
+package for the Linux distribution but need be created by the administrator
+manually after the Janitor has been installed. This also holds for the
+catalog.
+
+When working with several catalogs, then the multiple catalog lines can
+be placed into the same arc.conf file. But every must go into its own block
+as separated with \[janitor/someName\] directives.
+
 \lstsetCONFIGURE

### 2009-11-27 doc/tech_doc/client/broker.tex
Adding broker mapping document.

\documentclass[twoside,english]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{booktabs}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Document Title},
  pdfsubject = {Document Subtitle},
  pdfkeywords = {ARC,NorduGrid,manual,keywords},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%

\makeatother

\usepackage{babel}

\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl} \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}} &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}} \end{tabular}

\hrulefill

{\raggedleft NORDUGRID-XXXXXXX-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

\begin{center}
\textsc{\Large ARC Broker}
\par\end{center}{\Large \par}

\vspace*{0.5cm}

\begin{center}
\textit{\large Paper Subtitle}
\par\end{center}{\large \par}

\vspace*{1.5cm}

{\centering \large M.S. Andersen\footnote{skou@nbi.dk} \large \par}

\end{titlepage}

\thispagestyle{empty}

\newpage

\tableofcontents{}

\newpage

\section{Introduction}\label{sec:Introduction}

This document contains a description of the broker used in the Advanced
Resource Connector (ARC) middleware as of version 1.0. In section
\ref{sec:Attribute-matching} a description of how target attributes
are matched and compared to attributes specified in the job description.

\section{Attribute matching}\label{sec:Attribute-matching}

In the broker attributes published by the candidate target need to
be matched and compared to attributes specified in the job description
to be able to select targets which are suitable for accepting the
job description. Targets which do not pass a certain comparison will
be not be a candidate target for submission. Detailed information
about the attributes described below can be found in the ARC job description
document. First the broker will check some basics for the target being
processed. First, if the job description specifies submission to specific
targets and/or queues, the target will be checked and if it is not
among these targets and/or queues, it will be dropped. Next the number
of free slots on the target will be checked, and if no free slots
exist, then it will be dropped. After this, the health state of the
target is checked, and if any other value than "ok"
is published then it is dropped. Last, if middleware specific submission
is specified in the job description, the middleware type of the target
is checked, and if it does not satisfy the middleware requirements
then it is dropped. After these basic checks, certain attributes specified
in the job description are compared to the attributes published by
the target being processed. If the target does not publish a attribute
which is specified in the job description the target will dropped,
with some exceptions.

\begin{table}
\footnotesize{\texttt{
\begin{tabular}{cccc}
\toprule
\textnormal{\normalsize Job description attributes} & \textnormal{\normalsize Comparator} & \textnormal{\normalsize Target attribute} & \tabularnewline
\midrule
Resources.CandidateTarget[i].EndPointURL & $==$ & url & \tabularnewline
Resources.CandidateTarget[i].QueueName & $==$ & MappingQueue & \tabularnewline
\midrule
Application.ProcessingStartTime    & $<$ & DowntimeStarts & \multirow{2}{*}{AND} \tabularnewline
Application.ProcessingStartTime    & $>$ & DowntimeEnds &\tabularnewline
\midrule
Resources.CEType & isSatisfied( & Implementation & ) \tabularnewline
\midrule
Resources.TotalWallTime.range.max & $<$ & MaxWallTime & \tabularnewline
Resources.TotalWallTime.range.min & $>$ & MinWallTime & \tabularnewline
Resources.TotalCPUTime.range.max & $<$ & MaxCPUTime & \tabularnewline
Resources.TotalCPUTime.range.min & $>$ & MinCPUTime & \tabularnewline
\midrule
Resources.IndividualPhysicalMemory & $\leq$ & MainMemory & \multirow{2}{*}{OR} \tabularnewline
Resources.IndividualPhysicalMemory & $\leq$ & MaxMemory &\tabularnewline
Resources.IndividualVirtualMemory & $\leq$ & MaxVirtualMemory & \tabularnewline
\midrule
Resources.Platform & $!=$ & Platform & \tabularnewline
\midrule
Resources.OperatingSystem & isSatisfied( & OperatingSystem & )\tabularnewline
Resources.RunTimeEnvironment & isSatisfied( & ApplicationEnvironments & )\tabularnewline
\midrule
Resources.NetworkInfo & in & NetworkInfo & \tabularnewline
\midrule
Resources.DiskSpaceRequirement.SessionDiskSpace & $\leq$ & MaxDiskSpace & \multirow{2}{*}{OR} \tabularnewline
Resources.DiskSpaceRequirement.SessionDiskSpace & $\leq$ & WorkingAreaTotal & \tabularnewline
\midrule
Resources.DiskSpaceRequirement.DiskSpace & \multirow{2}{*}{$\leq$} & \multirow{2}{*}{MaxDiskSpace} & \multirow{4}{*}{OR} \tabularnewline
$-$ Resources.DiskSpaceRequirement.CacheDiskSpace & & & \tabularnewline
Resources.DiskSpaceRequirement.DiskSpace & \multirow{2}{*}{$\leq$} & \multirow{2}{*}{WorkingAreaTotal} & \tabularnewline
$-$ Resources.DiskSpaceRequirement.CacheDiskSpace & & & \tabularnewline
\midrule
Resources.DiskSpaceRequirement.DiskSpace & $\leq$ & MaxDiskSpace & \multirow{2}{*}{OR} \tabularnewline
Resources.DiskSpaceRequirement.DiskSpace & $\leq$ & WorkingAreaTotal & \tabularnewline
Resources.DiskSpaceRequirement.CacheDiskSpace & $\leq$ & CacheTotal & \tabularnewline
\midrule
Resources.SlotRequirement.NumberOfSlots & $\leq$ & TotalSlots & \tabularnewline
Resources.SlotRequirement.NumberOfSlots & $\leq$ & MaxSlotsPerJob & \tabularnewline
\midrule
Resources.SessionLifeTime & $\leq$ & WorkingAreaLifeTime & \tabularnewline
\midrule
Resources.NodeAccess is NAT\_{}INBOUND OR NAT\_{}INOUTBOUND & AND & ConnectivityIn & \tabularnewline
Resources.NodeAccess is NAT\_{}OUTBOUND OR NAT\_{}INOUTBOUND & AND & ConnectivityOut & \tabularnewline
\bottomrule
\end{tabular}
}}
\caption{}
\end{table}

\begin{thebibliography}{2}
\bibitem{key-1}NorduGrid middleware, the Advanced Resource Connector.
\url{http://www.nordugrid.org/middleware/}

\bibitem{Globus}The Globus Alliance. \texttt{\url{http://www.globus.org}}
\end{thebibliography}

\end{document}

### 2009-11-27 doc/tech_doc/client/job_description_mapping.tex
Updating job description mapping document.

@@
 XRSL & jobname & jobname\\
-JDL & No matching attribute & Not mapped\\
 JSDL & JobIdentification/JobName & JobIdentification/JobName\\
-ARC-JSDL &  & \\
+ARC-JSDL & JobIdentification/JobName & JobIdentification/JobName \\
 \end{inouttabular}
@@
 &  Application/HPCProfileApplication/Arguments & Application/HPCProfileApplication/Arguments\\
-\multirow{2}{*}{ARC-JSDL} & Application/Executable/Path, & Application/Executable/Path,\\
- & Application/Executable/Argument & Application/Executable/Argument\\
 \end{inouttabular}
@@
 HPCP-JSDL & Application/HPCProfileApplication/Input & Application/HPCProfileApplication/Input\\
-ARC-JSDL & Application/Input & \\
 \end{inouttabular}
@@
 JDL & StdOutput & StdOutput, OutputSandbox, OutputSandboxDestURI\\
-Posix-JSDL & Application/POSIXApplication/Output & \\
-HPCP-JSDL & Application/HPCProfileApplication/Output & \\
-ARC-JSDL & Application/Output & \\
+Posix-JSDL & Application/POSIXApplication/Output & Application/POSIXApplication/Output\\
+HPCP-JSDL & Application/HPCProfileApplication/Output & Application/HPCProfileApplication/Output\\
 \end{inouttabular}
@@
 JDL & StdError & StdError, OutputSandbox, OutputSandboxDestURI\\
-Posix-JSDL & Application/POSIXApplication/Error &\\
-HPCP-JSDL & Application/HPCProfileApplication/Error &\\
-ARC-JSDL & Application/Error & \\
+Posix-JSDL & Application/POSIXApplication/Error & Application/POSIXApplication/Error\\
+HPCP-JSDL & Application/HPCProfileApplication/Error & Application/HPCProfileApplication/Error\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & join & Not mapped\todo{}\\
+XRSL & join & \NM\\
 ARC-JSDL & Application/Join & Application/Join\\
@@
 HPCP-JSDL & Application/HPCProfileApplication/Environment.name & Application/HPCProfileApplication/Environment.name\\
-ARC-JSDL & Application/Environment/Name & ~\\
 \end{inouttabular}
@@
 HPCP-JSDL & Application/HPCProfileApplication/Environment & Applicaion/HPCProfileApplication/Environment\\
-ARC-JSDL & Application/Environment/Value & ~\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & gmlog & gmlog outputFiles\\
+XRSL & gmlog & gmlog, outputFiles\\
 ARC-JSDL & Application/LogDir & Application/LogDir\\
@@
 ARC-JSDL & Application/ExpiryTime & Application/ExpiryTime\\
-\end{inouttabular}
-
-\subsubsection{StartOnHold}
-\begin{inouttabular}
 \end{inouttabular}
@@
 XRSL & architecture & architecture\\
-JSDL & Resources/CPUAchitecture/CPUArchitectureName & ~\\
-ARC-JSDL & Resources/Platform &~\\
+JSDL & Resources/CPUAchitecture/CPUArchitectureName & Resources/CPUAchitecture/CPUArchitectureName\\
+ARC-JSDL & Resources/Platform & Resources/Platform\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-JSDL & Resources/IndividualNetworkBandwidth &\\
-ARC-JSDL & Resources/NetworkInfo &\\
+JSDL & Resources/IndividualNetworkBandwidth & Resources/IndividualNetworkBandwidth \\
+ARC-JSDL & Resources/NetworkInfo & \NM\\
 \end{inouttabular}
@@
 XRSL & nodeAccess & nodeAccess\\
-ARC-JSDL & Resources/NodeAccess & ~\\
+ARC-JSDL & Resources/NodeAccess & Resources/NodeAccess\\
 \end{inouttabular}
@@
 XRSL & memory & memory\\
-Posix-JSDL & Application/POSIXApplication/MemoryLimit &\\
-ARC-JSDL & Resources/IndividualPhysicalMemory &\\
+JSDL & Resources/IndividualPhysicalMemory & Resources/IndividualPhysicalMemory\\
+Posix-JSDL & Application/POSIXApplication/MemoryLimit & Application/POSIXApplication/MemoryLimit\\
+ARC-JSDL & Resources/IndividualPhysicalMemory & Resources/IndividualPhysicalMemory\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-Posix-JSDL & Application/POSIXApplication/VirtualMemoryLimit &\\
-ARC-JSDL & Resources/IndividualVirtualMemory &\\
+JSDL & Resources/IndividualVirtualMemory & Resources/IndividualVirtualMemory\\
+Posix-JSDL & Application/POSIXApplication/VirtualMemoryLimit & Application/POSIXApplication/VirtualMemoryLimit\\
+ARC-JSDL & Resources/IndividualVirtualMemory & Resources/IndividualVirtualMemory\\
 \end{inouttabular}
@@
 XRSL & disk & disk\\
-JSDL & Application/FileSystem/DiskSpace &\\
-ARC-JSDL & Resources/DiskSpaceRequirement/DiskSpace &\\
+JSDL & Resources/FileSystem/DiskSpace & Resources/FileSystem/DiskSpace\\
+ARC-JSDL & Resources/DiskSpaceRequirement/DiskSpace & Resources/DiskSpaceRequirement/DiskSpace\\
 \end{inouttabular}
@@
 
-\subsubsection{SessionAccessMode}
-\begin{inouttabular}
-\end{inouttabular}
-
 \subsubsection{IndividualCPUTime}
 \begin{inouttabular}
-JSDL & Resources/IndividualCPUTime* &\\
-ARC-JSDL & Resources/IndividualCPUTime &\\
+JSDL & Resources/IndividualCPUTime* & Resources/IndividualCPUTime\\
+ARC-JSDL & Resources/IndividualCPUTime & Resources/IndividualCPUTime\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & cpuTime & cpuTime\\
-JSDL & Resources/TotalCPUTime* &\\
-Posix-JSDL & Application/POSIXApplication/CPUTimeLimit* &\\
-ARC-JSDL & Resources/TotalCPUTime &\\
+XRSL & cpuTime, gridtime, benchmark & cpuTime\\
+JSDL & Resources/TotalCPUTime* & Resources/TotalCPUTime\\
+Posix-JSDL & Application/POSIXApplication/CPUTimeLimit* & Application/POSIXApplication/CPUTimeLimit\\
+ARC-JSDL & Resources/TotalCPUTime & Resources/TotalCPUTime\\
 \end{inouttabular}
-* The JSDL and POSIX-JSDL elements is mapped to the
-Range part of TotalCPUTime.
+The \texttt{XRSL} attribute \texttt{cputime} is mapped to the
+\texttt{Range} part of \texttt{TotalCPUTime}, where as for the
+\texttt{gridtime} attibute the pair ("ARC-gridtime", 1800) will
+be added to the benchmark part.
+The JSDL and POSIX-JSDL elements is mapped to the Range part of
+TotalCPUTime.
 
@@
 \begin{inouttabular}
-ARC-JSDL & Resources/IndividualWallTime &\\
+ARC-JSDL & Resources/IndividualWallTime & Resources/IndividualWallTime\\
 \end{inouttabular}
@@
 XRSL & wallTime & wallTime\\
-Posix-JSDL & Application/POSIXApplication/WallTimeLimit &\\
-ARC-JSDL & Resources/TotalWallTime &\\
-\end{inouttabular}
-
-\subsubsection{Homogeneous}
-\begin{inouttabular}
-\end{inouttabular}
-
-\subsubsection{Benchmark}
-\subsubsubsection{Type}
-\begin{inouttabular}
-\end{inouttabular}
-
-\subsubsubsection{Value}
-\begin{inouttabular}
+Posix-JSDL & Application/POSIXApplication/WallTimeLimit & Application/POSIXApplication/WallTimeLimit\\
+ARC-JSDL & Resources/TotalWallTime & Resources/TotalWallTime\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-Posix-JSDL & Application/POSIXApplication/ProcessCountLimit &\\
-ARC-JSDL & Resources/SlotRequirement/NumberOfSlots &\\
+Posix-JSDL & Application/POSIXApplication/ProcessCountLimit & Application/POSIXApplication/ProcessCountLimit\\
+ARC-JSDL & Resources/SlotRequirement/NumberOfSlots & Resources/SlotRequirement/NumberOfSlots\\
 \end{inouttabular}
@@
 XRSL & count & count\\
-JSDL & Resources/TotalCPUCount & \\
-ARC-JSDL & Resources/SlotRequirement/ProcessPerHost &\\
+JSDL & Resources/TotalCPUCount & Resources/TotalCPUCount\\
+ARC-JSDL & Resources/SlotRequirement/ProcessPerHost & Resources/SlotRequirement/ProcessPerHost\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-Posix-JSDL & Application/POSIXApplication/ThreadCountLimit &\\
-ARC-JSDL & Resources/SlotRequirement/ThreadsPerProcesses &\\
+Posix-JSDL & Application/POSIXApplication/ThreadCountLimit & Application/POSIXApplication/ThreadCountLimit\\
+ARC-JSDL & Resources/SlotRequirement/ThreadsPerProcesses & Resources/SlotRequirement/ThreadsPerProcesses\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-ARC-JSDL & Resources/SlotRequirement/SPMDVariation &\\
+ARC-JSDL & Resources/SlotRequirement/SPMDVariation & Resources/SlotRequirement/SPMDVariation\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & cluster & Not mapped\\
-JSDL & Resources/CandidateHosts/HostName &\\
+XRSL & cluster & \NM\\
+JSDL & Resources/CandidateHosts/HostName & \NM\\
 ARC-JSDL & Resources/CandidateTarget/EndPointURL & Resources/CandidateTarget/EndPointURL\\
@@
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+XRSL & inputFiles, outputFiles & inputFiles, outputFiles\\
+JDL & InputSandbox, OutputSandbox & InputSandbox, OutputSandbox\\
+JSDL & DataStaging/File/FileName & DataStaging/File/FileName\\
 \end{inouttabular}
@@
 \subsubsubsection{Source}
-\subsubsubsubsection{URL}
+\subsubsubsubsection{URI}
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+XRSL & inputFiles & inputFiles\\
+JDL & InputSandbox, InputSandboxURI & InputSandbox, InputSandboxURI\\
+JSDL & DataStaging/File/Source/URI & DataStaging/File/Source/URI\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+XRSL & ftpThreads & ftpThreads\\
+%~ ARC-JSDL & &\\
 \end{inouttabular}
 
-\subsubsubsubsection{DataIndexingService}
-\begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
-\end{inouttabular}
+%~ \subsubsubsubsection{DataIndexingService}
+%~ \begin{inouttabular}
+%~ XRSL & &\\
+%~ JDL & &\\
+%~ JSDL & &\\
+%~ Posix-JSDL & &\\
+%~ HPCP-JSDL & &\\
+%~ ARC-JSDL & &\\
+%~ \end{inouttabular}
 
 \subsubsubsection{Target}
-\subsubsubsubsection{URL}
+\subsubsubsubsection{URI}
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+XRSL & outputFiles & outputFiles\\
+JDL & OutputSandbox, OutputSandboxDestURI & OutputSandbox, OutputSandboxDestURI\\
+JSDL & DataStaging/File/Target/URI & DataStaging/File/Target/URI\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+XRSL & ftpThreads &ftpThreads\\
+%~ ARC-JSDL & &\\
 \end{inouttabular}
 
-\subsubsubsubsection{Mandatory}
-\begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
-\end{inouttabular}
+%~ \subsubsubsubsection{Mandatory}
+%~ \begin{inouttabular}
+%~ XRSL & &\\
+%~ JDL & &\\
+%~ JSDL & &\\
+%~ Posix-JSDL & &\\
+%~ HPCP-JSDL & &\\
+%~ ARC-JSDL & &\\
+%~ \end{inouttabular}
 
-\subsubsubsubsection{DataIndexingService}
-\begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
-\end{inouttabular}
+%~ \subsubsubsubsection{DataIndexingService}
+%~ \begin{inouttabular}
+%~ XRSL & &\\
+%~ JDL & &\\
+%~ JSDL & &\\
+%~ Posix-JSDL & &\\
+%~ HPCP-JSDL & &\\
+%~ ARC-JSDL & &\\
+%~ \end{inouttabular}
 
-\subsubsubsubsection{NeededReplica}
-\begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
-\end{inouttabular}
+%~ \subsubsubsubsection{NeededReplica}
+%~ \begin{inouttabular}
+%~ XRSL & &\\
+%~ JDL & &\\
+%~ JSDL & &\\
+%~ Posix-JSDL & &\\
+%~ HPCP-JSDL & &\\
+%~ ARC-JSDL & &\\
+%~ \end{inouttabular}
 
@@
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+JDL & OutputSandboxDestURI & OutputSandboxDestURI\\
+JSDL & DataStaging/File/DeleteOnTermination & DataStaging/File/DeleteOnTermination\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+XRSL & executables & executables\\
+ARC-JSDL & DataStaging/File/IsExecutable & DataStaging/File/IsExecutable\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+XRSL & \todo{} & \todo{}\\
+ARC-JSDL & DataStaging/File/DownloadToCache & DataStaging/File/DownloadToCache\\
 \end{inouttabular}
 
-\subsubsection{Directory}
+%~ \subsubsection{Directory}
 
 \subsection{JobMetaData}
-\subsubsection{Author}
-\begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
-\end{inouttabular}
-
-\subsubsection{DocumentExpiration}
-\begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
-\end{inouttabular}
+%~ \subsubsection{Author}
+%~ \begin{inouttabular}
+%~ XRSL & &\\
+%~ JDL & &\\
+%~ JSDL & &\\
+%~ Posix-JSDL & &\\
+%~ HPCP-JSDL & &\\
+%~ ARC-JSDL & &\\
+%~ \end{inouttabular}
+%~
+%~ \subsubsection{DocumentExpiration}
+%~ \begin{inouttabular}
+%~ XRSL & &\\
+%~ JDL & &\\
+%~ JSDL & &\\
+%~ Posix-JSDL & &\\
+%~ HPCP-JSDL & &\\
+%~ ARC-JSDL & &\\
+%~ \end{inouttabular}
 
@@
 \begin{inouttabular}
-XRSL & &\\
-JDL & &\\
-JSDL & &\\
-Posix-JSDL & &\\
-HPCP-JSDL & &\\
-ARC-JSDL & &\\
+JDL & Rank & Rank\\
+%~ ARC-JSDL & &\\
 \end{inouttabular}
@@
 \begin{inouttabular}
-XRSL & ~ & ~\\
-JDL & ~ & ~\\
-JSDL & ~ & ~\\
-Posix-JSDL & ~ & ~\\
-HPCP-JSDL & ~ & ~\\
-ARC-JSDL & ~ & ~\\
+JDL & FuzzyRank & FuzzyRank\\
+%~ ARC-JSDL & & \\
 \end{inouttabular}
@@
 notify & Application.Notification & notify\\
-replicaCollection & DataStaging.File.DataIndexingService & \NM\\
+replicaCollection & \NM & \NM\\
 rerun & Application.Rerun & rerun\\
@@
 OutputSandboxDestURI & DataStaging.File.Target,\linebreak DataStaging.File.KeepData & OutputSandbox,\linebreak OutputSandboxDestURI\\
-OutputSandboxBaseDestURI & DataStaging.File.Target.URI & OutputSandbox,\linebreak OutputSandboxDestURI\\
 Prologue & Application.Prologue.Name & Prologue\\
@@
 \rowcolors{3}{}{tablerowcolor}
-\begin{tabularx}{400pt}{TCC}
+\begin{tabularx}{\textwidth}{TCT}
 \toprule
@@
 \midrule
-JobName & &\\
-JobAnnotation & &\\
-JobProject & &\\
+JobName & Application.JobName & JobName\\
+JobAnnotation & \NM & \NM\\
+JobProject & \NM & \NM\\
 \midrule
@@
 \midrule
-ApplicationName & &\\
-ApplicationVersion & &\\
+ApplicationName & \NM & \NM\\
+ApplicationVersion & \NM & \NM\\
 \midrule
@@
 \midrule
-CandidateHosts/HostName & &\\
-FileSystem/FileSystemType & &\\
-FileSystem/Description & &\\
-FileSystem/MountPoint & &\\
-FileSystem/DiskSpace & &\\
-ExclusiveExecution & &\\
-OperatingSystem/\linebreak OperatingSystemType/\linebreak OperatingSystemName & &\\
-OperatingSystem/\linebreak OperatingSystemVersion & &\\
-CPUArchitecture/\linebreak CPUArchitectureName & &\\
-IndividualCPUSpeed & &\\
-IndividualCPUTime & &\\
-IndividualCPUCount & &\\
-IndividualNetworkBandwidth & &\\
-IndividualPhysicalMemory & &\\
-IndividualVirtualMemory & &\\
-IndividualDiskSpace & &\\
-TotalCPUTime & &\\
-TotalCPUCount & &\\
-TotalPhysicalMemory & &\\
-TotalVirtualMemory & &\\
-TotalDiskSpace & &\\
-TotalResourceCount & &\\
+CandidateHosts/HostName & Resources.CandidateTarget.EndPointURL & \NM\\
+FileSystem/FileSystemType & \NM & \NM\\
+FileSystem/Description & \NM & \NM\\
+FileSystem/MountPoint & \NM & \NM\\
+FileSystem/DiskSpace & Resources.DiskSpaceRequirement.DiskSpace & FileSystem/DiskSpace\\
+ExclusiveExecution & \NM & \NM\\
+OperatingSystem/\linebreak OperatingSystemType/\linebreak OperatingSystemName & Resources.OperatingSystem & OperatingSystem/\linebreak OperatingSystemType/\linebreak OperatingSystemName\\
+OperatingSystem/\linebreak OperatingSystemVersion & Resources.OperatingSystem & OperatingSystem/\linebreak OperatingSystemVersion\\
+CPUArchitecture/\linebreak CPUArchitectureName & Resources.Platform & CPUAchitecture/\linebreak CPUAchitectureName\\
+IndividualCPUSpeed & \NM & \NM\\
+IndividualCPUTime & job.Resources.IndividualCPUTime & IndividualCPUTime\\
+IndividualCPUCount & \NM & \NM\\
+IndividualNetworkBandwidth & Resources.NetworkInfo & IndividualNetworkBandwidth\\
+IndividualPhysicalMemory & job.Resources.IndividualPhysicalMemory & IndividualPhysicalMemory\\
+IndividualVirtualMemory & job.Resources.IndividualVirtualMemory & IndividualVirtualMemory\\
+IndividualDiskSpace & \NM & \NM\\
+TotalCPUTime & job.Resources.TotalCPUTime.range & TotalCPUTime\\
+TotalCPUCount & job.Resources.SlotRequirement.ProcessPerHost & TotalCPUCount\\
+TotalPhysicalMemory & \NM & \NM\\
+TotalVirtualMemory & \NM & \NM\\
+TotalDiskSpace & \NM & \NM\\
+TotalResourceCount & \NM & \NM\\
 \midrule
@@
 \midrule
-FileName & &\\
-FileSystemName & &\\
-CreationFlag & &\\
-DeleteOnTermination & &\\
-Source/URI & &\\
-Target/URI & &\\
+FileName & DataStaging.File.FileName & FileName\\
+FileSystemName & \NM & \NM\\
+CreationFlag & \NM & \NM\\
+DeleteOnTermination & DataStaging.File.KeepData & DeleteOnTermination\\
+Source/URI & DataStaging.File.Source.URI & Source/URI\\
+Target/URI & DataStaging.File.Target.URI & Target/URI\\
 \end{tabularx}
@@
 \rowcolors{3}{}{tablerowcolor}
-\begin{tabularx}{400pt}{CCC}
+\begin{tabularx}{\textwidth}{CCT}
 \toprule
@@
 \defineheadfoot
-Executable & &\\
-Argument & &\\
-Input & &\\
-Output & &\\
-Error & &\\
-WorkingDirectory & &\\
-Environment & &\\
-WallTimeLimit & &\\
-FileSizeLimit & &\\
-CoreDumpLimit & &\\
-DataSegmentLimit & &\\
-LockedMemoryLimit & &\\
-OpenDescriptorsLimit & &\\
-PipeSizeLimit & &\\
-StackSizeLimit & &\\
-CPUTimeLimit & &\\
-ProcessCountLimit & &\\
-VirtualMemoryLimit & &\\
-ThreadCountLimit & &\\
-UserName & &\\
-GroupName & &\\
+Executable & Application.Executable.Name & Executable,\linebreak DataStaging.File.Name,\linebreak DataStaging.File.Source\\
+Argument & Application.Executable.Arguments & Argument\\
+Input & Application.Input & Input,\linebreak DataStaging.File.Name,\linebreak DataStaging.File.Source\\
+Output & Application.Output & Output,\linebreak DataStaging.File.Name,\linebreak DataStaging.File.Target\\
+Error & Application.Error & Error,\linebreak DataStaging.File.Name,\linebreak DataStaging.File.Target\\
+WorkingDirectory & \NM & \NM\\
+Environment & Application.Environment & Environment\\
+WallTimeLimit & job.Resources.TotalWallTime.range.max & WallTimeLimit\\
+FileSizeLimit & \NM & \NM\\
+CoreDumpLimit & \NM & \NM\\
+DataSegmentLimit & \NM & \NM\\
+LockedMemoryLimit & \NM & \NM\\
+OpenDescriptorsLimit & \NM & \NM\\
+PipeSizeLimit & \NM & \NM\\
+StackSizeLimit & \NM & \NM\\
+CPUTimeLimit & job.Resources.TotalCPUTime.range.max & CPUTimeLimit\\
+ProcessCountLimit & job.Resources.SlotRequirement.NumberOfSlots.max & ProcessCountLimit\\
+VirtualMemoryLimit & job.Resources.IndividualVirtualMemory.max & VirtualMemoryLimit\\
+ThreadCountLimit & job.Resources.SlotRequirement.ThreadsPerProcesses.max & ThreadCountLimit\\
+UserName & \NM & \NM\\
+GroupName & \NM & \NM\\
 \end{tabularx}
@@
 \rowcolors{3}{tablerowcolor}{}
-\begin{tabularx}{400pt}{CCC}
+\begin{tabularx}{\textwidth}{CCT}
 \toprule
@@
 \defineheadfoot
-Executable & & \\
-Argument & & \\
-Input & & \\
-Output & & \\
-Error & & \\
-WorkingDirectory & &\\
-Environment & &\\
-UserName & &\\
+Executable & Application.Executable.Name & Executable,\linebreak DataStaging.File.Name,\linebreak DataStaging.File.Source\\
+Argument & Application.Executable.Arguments & Argument\\
+Input & Application.Input & Input,\linebreak DataStaging.File.Name,\linebreak DataStaging.File.Source\\
+Output & Application.Output & Output,\linebreak DataStaging.File.Name,\linebreak DataStaging.File.Target\\
+Error & Application.Error & Error,\linebreak DataStaging.File.Name,\linebreak DataStaging.File.Target\\
+WorkingDirectory & \NM & \NM \\
+Environment & Application.Environment & Environment\\
+UserName & \NM & \NM\\
 \end{tabularx}

### 2009-11-27 doc/tech_doc/a-rex/arex_tech_doc.tex
add note to wakeupperiod option
fix example config in a-rex manual

@@
 finished in LRMS, etc.). \textit{time} is a minimal time period specified
-in seconds. Default is \emph{3 minutes}.
+in seconds. Default is \emph{3 minutes}. The A-REX may also be woken up by
+external processes such as LRMS scripts before this time period expires.
 \item \textbf{\textit{authplugin}}\textit{=state options plugin} -- specifies
@@
 <ArcConfig
-  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
-  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
-  xmlns:arex="http://www.nordugrid.org/schemas/a-rex/Config"
+  xmlns="http://www.nordugrid.org/schemas/arcconfig/2009/08"
+  xmlns:loader="http://www.nordugrid.org/schemas/loader/2009/08"
+  xmlns:arex="http://www.nordugrid.org/schemas/a-rex/2009/08"
+  xmlns:tcp="http://www.nordugrid.org/schemas/tcp/2009/08"
+  xmlns:tls="http://www.nordugrid.org/schemas/tls/2009/08"
+  xmlns:authz="http://www.nordugrid.org/schemas/arcauthz/2009/08"
+  xmlns:idmap="http://www.nordugrid.org/schemas/identitymap/2009/10"
 >
@@
       <File>/var/log/arched.log</File>
+      <MaxSize>100000000</MaxSize>
       <Backups>10</Backups>
-      <Maxsize>100000000</Maxsize>
     </Logger>
@@
   <!-- Where to find plugins -->
-  <ModuleManager>
-    <Path>/usr/local/lib/arc/</Path>
-  </ModuleManager>
+  <loader:ModuleManager>
+    <loader:Path>/usr/local/lib/arc/</loader:Path>
+  </loader:ModuleManager>
   <!-- Simply load all needed plugins -->
-  <Plugins><Name>mcctcp</Name></Plugins>
-  <Plugins><Name>mcctls</Name></Plugins>
-  <Plugins><Name>mcchttp</Name></Plugins>
-  <Plugins><Name>mccsoap</Name></Plugins>
-  <Plugins><Name>arcpdc</Name></Plugins>
-  <Plugins><Name>identitymap</Name></Plugins>
-  <Plugins><Name>arex</Name></Plugins>
+  <loader:Plugins>
+    <loader:Name>mcctcp</loader:Name>
+    <loader:Name>mcctls</loader:Name>
+    <loader:Name>mcchttp</loader:Name>
+    <loader:Name>mccsoap</loader:Name>
+    <loader:Name>arcpdc</loader:Name>
+    <loader:Name>identitymap</loader:Name>
+    <loader:Name>arex</loader:Name>
+  </loader:Plugins>
   <!-- Create a chain -->
-  <Chain>
+  <loader:Chain>
     <!-- TCP listening socket -->
-    <Component name="tcp.service" id="tcp">
-      <next id="tls"/>
+    <loader:Component name="tcp.service" id="tcp">
+      <loader:next id="tls"/>
       <tcp:Listen><tcp:Port>60000</tcp:Port></tcp:Listen>
-    </Component>
+    </loader:Component>
     <!-- Transport-level security -->
-    <Component name="tls.service" id="tls">
-      <next id="http"/>
+    <loader:Component name="tls.service" id="tls">
+      <loader:next id="http"/>
       <!-- Location of server's security keys -->
-      <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
-      <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
-      <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
+      <tls:KeyPath>/etc/grid-security/hostkey.pem</tls:KeyPath>
+      <tls:CertificatePath>/etc/grid-security/hostcert.pem</tls:CertificatePath>
+      <tls:CACertificatesDir>/etc/grid-security/certificates</tls:CACertificatesDir>
       <!-- Evaluate requestor's identity into local identity.
       Comment it if no user mapping is needed. -->
-      <SecHandler name="identity.map" id="map" event="incoming">
+      <loader:SecHandler name="identity.map" id="map" event="incoming">
         <!-- Safe choice if all other rules failed -->
-        <PDP name="allow.pdp"><LocalName>nobody</LocalName></PDP>
-      </SecHandler>
-    </Component>
+        <idmap:PDP name="allow.pdp"><idmap:LocalName>nobody</idmap:LocalName></idmap:PDP>
+      </loader:SecHandler>
+    </loader:Component>
     <!-- HTTP processing is done here -->
-    <Component name="http.service" id="http">
-      <next id="soap">POST</next>
-      <next id="plexer">GET</next>
-      <next id="plexer">PUT</next>
-    </Component>
+    <loader:Component name="http.service" id="http">
+      <loader:next id="soap">POST</loader:next>
+      <loader:next id="plexer">GET</loader:next>
+      <loader:next id="plexer">PUT</loader:next>
+    </loader:Component>
     <!-- This one parses content into XML tree -->
-    <Component name="soap.service" id="soap">
-      <next id="plexer"/>
-    </Component>
+    <loader:Component name="soap.service" id="soap">
+      <loader:next id="plexer"/>
+    </loader:Component>
     <!-- Directing messages to proper service -->
-    <Plexer name="plexer.service" id="plexer">
+    <loader:Plexer name="plexer.service" id="plexer">
       <!-- RegExp pattern is matched to path part of endpoint.
@@
       PLEXER:EXTENSION attribute. -->
-      <next id="a-rex">^/arex</next>
-    </Plexer>
+      <loader:next id="a-rex">^/arex</loader:next>
+    </loader:Plexer>
     <!-- A-Rex service -->
-    <Service name="a-rex" id="a-rex">
+    <loader:Service name="a-rex" id="a-rex">
       <!-- Optional endpoint element is advised in case of multiple IP adresses -->
@@
       <arex:gmconfig>/etc/arc.conf</arex:gmconfig>
-    </Service>
-  </Chain>
+    </loader:Service>
+  </loader:Chain>
 </ArcConfig>

### 2009-11-27 doc/tech_doc/client/job_description_mapping.tex
Updating job description mapping document.

@@
 \begin{inouttabular}
-XRSL & \todo{} & \todo{}\\
+XRSL & cache & \NST{cache in URL-option}\\
 ARC-JSDL & DataStaging/File/DownloadToCache & DataStaging/File/DownloadToCache\\
@@
 executables & DataStaging.File.IsExecutable & executables\\
-cache & \TODO & \TODO \\
+cache & DataStaging.File.DownloadToCache & \NST{cache in URL-option} \\
 outputFiles & DataStaging.File.Name,\linebreak DataStaging.File.Target & outputFiles \\

### 2009-11-27 doc/tech_doc/client/client_technical.tex
Moved API documentation from client_technical.tex to appropriate header files as doxygen documentation

@@
 
-\chapter{Implementation}
-\label{sec:Implementation}
+\chapter{The ARC Brokering Module}
+\label{sec:brokering}
 
-In this section an overview of all important classes in the new
-{\libarcclient} is given. The overview is subdivided in two parts where
-first all the generic classes are presented,
-Section~\ref{sec:genclass}, before the grid flavour specializations
-are presented in Section~\ref{sec:plugins}. Classes are described both
-in words and by code.
-
-\section{Generic Classes}
-\label{sec:genclass}
-
-\subsection{{\ACC}}
-
-The Arc Client Component ({\ACC}) class is a base class needed for the
-{\Loader} in order to create loadable classes. It stores information
-about which flavour it supports and security related information
-needed by all ACC specializations.
-
-\begin{shaded}
-\begin{verbatim}
-class ACC {
-protected:
-  ACC(Config *cfg, const std::string& flavour);
-public:
-  virtual ~ACC();
-  const std::string& Flavour();
-protected:
-  std::string flavour;
-  std::string proxyPath;
-  std::string certificatePath;
-  std::string keyPath;
-  std::string caCertificatesDir;
-};
-\end{verbatim}
-\end{shaded}
-
-\subsection{{\TargetGenerator}}
-
-The {\TargetGenerator} class is the main class for resource discovery
-and information retrieval. It loads {\TargetRetriever} plugins in
-accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed}
-(e.g.\ if an {\URL} pointing to an ARC1 resource is given the
-TargetRetrieverARC1 is loaded).
-
-To perform a resource discovery, first construct a {\TargetGenerator}
-object using the {\TargetGenerator} constructor,
-
-\begin{shaded}
-\begin{verbatim}
-TargetGenerator(const UserConfig& usercfg,
-                const std::list<std::string>& clusters,
-                const std::list<std::string>& indexurls);
-\end{verbatim}
-\end{shaded}
-
-where \texttt{clusters} and \texttt{indexurls} are lists of strings
-where each string is either the name of a grid flavour followed by a
-colon and the URL of an information contact endpoint for a computing
-service or index server, respectively, of that grid flavour or an
-alias defined in the {\UserConfig} object passed as the first
-argument, e.g.:
-
-\begin{shaded}
-\begin{small}
-\begin{verbatim}
-ARC0:ldap://grid.tsl.uu.se:2135/Mds-Vo-name=Sweden,o=grid
-ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid
-\end{verbatim}
-\end{small}
-\end{shaded}
-
-where the former is the URL to an index server while the latter is an
-URL to a computing service. If a string in \texttt{clusters} or
-\texttt{indexurls} is prefixed by a minus sign, the corresponding
-computing service or index server is rejected and excluded during
-resource discovery.
-
-To prepare a list of {\ExecutionTarget}s use the {\TargetGenerator}
-object and invoke its method
-
-\begin{shaded}
-\begin{verbatim}
-GetTargets(int targetType, int detailLevel);
-\end{verbatim}
-\end{shaded}
-
-The {\TargetGenerator} will pass this request to the loaded
-{\TargetRetriever}s running them as individual threads for improved
-performance. The {\TargetGenerator} keeps records of index servers and
-computing services found by the {\TargetRetriever}s in order to avoid
-multiple identical queries. Accepted {\ExecutionTarget}s are stored in
-the FoundTargets array kept by the {\TargetGenerator}. See also
-Section~\ref{sec:TargetDiscovery} for a schematic drawing of the
-resource discovery process.
-
-Information about the found {\ExecutionTarget}s can be printed by
-
-\begin{shaded}
-\begin{verbatim}
-PrintTargetInfo(bool longlist) const;
-\end{verbatim}
-\end{shaded}
-
-\subsection{{\TargetRetriever}}
-
-The {\TargetRetriever} is the base class for grid middleware specialized 
-{\TargetRetriever}s and inherits from the {\ACC} base class in
-order to be loadable by the ARC \texttt{Loader}. It is designed to
-work in conjunction with the {\TargetGenerator} and contains the pure
-virtual method
-
-\begin{shaded}
-\begin{verbatim}
-virtual void GetTargets(TargetGenerator& mom, int targetType,
-                        int detailLevel) = 0;
-\end{verbatim}
-\end{shaded}
-
-which is to be implemented by the specialized class. While it is not
-mandatory it is recommended that the specialized class divides this
-method into two components: QueryIndex and InterrogateTarget. The
-former handles index server queries, and the latter the computing
-service queries and {\ExecutionTarget} preparation.
-
-If an index server query yields a URL to a different index server
-than the one queried, then the {\TargetRetriever} should call itself
-recursively creating a new {\TargetRetriever} for the discovered index
-server.
-
-\subsection{{\ExecutionTarget}}
-
-The {\ExecutionTarget} is the class representation of a computing
-resource (queue) capable of executing a grid job. It serves as input
-to the {\Broker} which is able to filter and rank different 
-{\ExecutionTarget}s from different grid flavours without a
-priori knowing their difference. The {\ExecutionTarget} class mimics
-the Glue2 information model (with a flattened structure), and thus a
-mapping between attributes from other information systems into the
-Glue2 format is needed. Appendix \ref{app:ExTarget} shows the current
-mapping for the production ARC, ARC1 and gLite middlewares.
-
-All attributes of the {\ExecutionTarget} can be printed by the method
-
-\begin{shaded}
-\begin{verbatim}
-void Print(bool longlist) const;
-\end{verbatim}
-\end{shaded}
-
-All information needed to submit a job is stored in the {\ExecutionTarget}, 
-and the {\ExecutionTarget}'s method
-
-\begin{shaded}
-\begin{verbatim}
-Submitter *GetSubmitter(const UserConfig& ucfg) const;
-\end{verbatim}
-\end{shaded}
-
-returns a {\Submitter} which is used to submit jobs. The {\UserConfig} object 
-is used to configure the security related information needed by the {\Submitter}.
-
-\subsection{{\Broker}}
-
-The {\Broker} inherits from the {\ACC} base class in order to be loadable by the ARC \texttt{Loader}. It is the base class of the 
-specialized {\Broker} and it implements the method for reducing a list of resources found by resource discovery (the list of 
-{\ExecutionTarget}s residing within the {\TargetGenerator}) to a list of resources capable of running a certain job:
+The {\Broker} inherits from the {\ACC} base class in order to be
+loadable by the ARC \texttt{Loader}. It is the base class of the
+specialized {\Broker} and it implements the method for reducing a list
+of resources found by resource discovery (the list of
+{\ExecutionTarget}s residing within the {\TargetGenerator}) to a list
+of resources capable of running a certain job:
 
@@
 
-The {\texttt PreFilterTargets} method compares every hardware and software requirement given in the job description against the 
-computing resource (cluster) specifications stored in the {\ExecutionTarget}. If the{\ExecutionTarget} doesn't fulfil 
-the requirements imposed by the job description, or it is impossible to carry out the matchmaking due to incomplete information 
-published by the computing resource, the {\ExecutionTarget} will be removed from the list of possible targets. A detailed overview 
-of the matchmaking between {\JobDescription} and {\ExecutionTarget} attributes is given in Appendix~\ref{app:broker-mapping}.
+The {\texttt PreFilterTargets} method compares every hardware and
+software requirement given in the job description against the
+computing resource (cluster) specifications stored in the
+{\ExecutionTarget}. If the{\ExecutionTarget} doesn't fulfil the
+requirements imposed by the job description, or it is impossible to
+carry out the matchmaking due to incomplete information published by
+the computing resource, the {\ExecutionTarget} will be removed from
+the list of possible targets. A detailed overview of the matchmaking
+between {\JobDescription} and {\ExecutionTarget} attributes is given
+in Appendix~\ref{app:broker-mapping}.
 
-Once prefiltered, the remaining {\ExecutionTarget}s should be ranked in order to return the ``best'' {\ExecutionTarget} for 
-job submission. Different ranking methods are implemented by the specialized brokers, but for usability and consistency reasons 
-these methods are encapsulated by the {\Broker} base class method
+Once prefiltered, the remaining {\ExecutionTarget}s should be ranked
+in order to return the ``best'' {\ExecutionTarget} for job
+submission. Different ranking methods are implemented by the
+specialized brokers, but for usability and consistency reasons these
+methods are encapsulated by the {\Broker} base class method
 
@@
 
-which invokes the {\texttt SortTargets} method implemented by the specialized broker (see Section~\ref{sec:brokerplugins}) 
-and returns the best target. The {\texttt GetBestTarget} method is ``incremented'' at each call, thus upon a second call 
-the second best {\ExecutionTarget} will be returned.
+which invokes the {\texttt SortTargets} method implemented by the
+specialized broker (see Section~\ref{sec:brokerplugins}) and returns
+the best target. The {\texttt GetBestTarget} method is ``incremented''
+at each call, thus upon a second call the second best
+{\ExecutionTarget} will be returned.
 
-\subsection{{\JobDescription}}
-
-When addressing interoperability it is of paramount importance to
-transparently address grid job descriptions written in different job
-description languages by translating them automatically. In the
-{\libarcclient} library this functionality is implemented in the
-{\JobDescription} class which parses job descriptions given in the 
-XRSL~\cite{xrsl}, JDL~\cite{jdl} or JSDL~\cite{jsdl} formats into an 
-internal representation based upon the upcoming PGI-JSDL proposal.
-
-\begin{shaded}
-\begin{verbatim}
-bool setSource(const std::string source);
-bool setSource(const XMLNode& source);
-\end{verbatim}
-\end{shaded}
-
-The setSource method takes a source job description string as input and identifies it's format 
-(language) through pattern matching (the \texttt{JobDescriptionOrderer} class).
-Once the format is known, the source elements are mapped from the original format (XRSL, JDL or JSDL) 
-into an internal representation (the \texttt{JobInnerRepresentation} class). Mapping details 
-are given in Section~\ref{app:jobinnerrepresentation}.
-
-The {\JobDescription} class has three back-end classes, sometimes
-referred to as back-end or translator modules, corresponding to the
-three supported job description languages. The {\JobDescription} class
-chooses the appropriate back-end module according to the pattern matching
-performed by the \texttt{JobDescriptionOrderer}, and uses the back-end
-module for parsing and translation (if needed) of the job descriptions.
-
-Once the job description language required by the {\ExecutionTarget} chosen 
-for job submission is known, the parsed and (if needed) translated job description 
-can be retrieved through the {\texttt GetProduct} function as shown below:
-
-\begin{shaded}
-\begin{verbatim}
-bool getProduct(std::string& product, std::string format = "POSIXJSDL") const;
-\end{verbatim}
-\end{shaded}
-
-If the inner representation is empty or the source job description is not
-valid, then the {\texttt GetProduct} output is empty too and the method 
-returns with false.  In order to obtain the automatic job description 
-modifications similar to those of the production ARC client, the 
-{\texttt GetProduct} output is always generated from the internal 
-representation using the \texttt{JobDescriptionOrderer} class to carry 
-out the translation. The original source string can be regained by
-calling the \texttt{getSourceString} method:
-
-\begin{shaded}
-\begin{verbatim}
-bool getSourceString(std::string source) const;
-\end{verbatim}
-\end{shaded}
-
-The translation of a job description basically means a syntactical transformation 
-of the job description from one language to another. Thus the {\JobDescription}
-class can do nothing in those cases where there is not enough data available 
-to assemble a given attribute or when information can be lost because the received 
-attribute has no equivalent in the output language. In these cases the {\texttt GetProduct} 
-method returns with a false value and an empty product string. For limitations in the translation
-see Section~\ref{app:jobinnerrepresentation}.
-
-For the JSDL output generation the {\JobDescription} class uses
-the core JSDL's capabilities amended with the JSDL-POSIX and JSDL-ARC
-extensions. The loss of information is minimal in case of generating
-such an output.
-
-For the JDL generation the latest version of JDL specification was
-used (v0.9) \cite{jdl}. Deprecated and backward compatibility
-attributes are not implemented.
-
-In case of job resubmission the internal representation will contain a list of 
-the jobid's of the previous jobs. This element will only be parsed to the 
-actually job description in case JSDL-POSIX output is requested. A new URL 
-can be added to the list of old jobid's by using the method:
-
-\begin{shaded}
-\begin{verbatim}
-bool addOldJobID(const URL& oldjobid);
-\end{verbatim}
-\end{shaded}
-
-\subsection{{\Submitter}}
-
-{\Submitter} is the base class for grid middleware specialized {\Submitter}s (plugins).
-It submits job(s) to the service it represents and uploads (needed by the job) local 
-input files.
-
-\begin{shaded}
-\begin{verbatim}
-virtual bool Submit(const JobDescription& jobdesc, XMLNode& info) const = 0;
-\end{verbatim}
-\end{shaded}
-
-The \texttt{Submit} method fills the {\XMLNode} \texttt{info} with all
-needed information about the job for later job management. The
-{\Submitter} is returned by the {\ExecutionTarget} selected for job
-execution and thus the {\ExecutionTarget} populates (through its
-{\XMLNode} config element) the {\Submitter} with information about
-submission endpoint ({\URL}) and job description languages understood
-by the target.
-
-\begin{shaded}
-\begin{verbatim}
-virtual bool Migrate(const URL& url,
-                     const JobDescription& jobdesc,
-                     bool forcemigration,
-                     XMLNode& info) const = 0;
-\end{verbatim}
-\end{shaded}
-
-The \texttt{Migrate} method behaves in much the same way as the \texttt{Submit}
-method, but instead of sending a new job request to the cluster it sends a 
-migration request of a given job, currently this request is only supported 
-for ARC1 clusters. The \texttt{URL} and \texttt{jobdesc} specifies the job-id 
-and {\JobDescription} of the job which should be migrated.
-
-Note that whereas the \texttt{Migrate} method make provisions for automatically 
-getting files from the migration source cluster, this does not apply to the job 
-description which instead should be passed to \texttt{Migrate} explicitly in the 
-form of a {\JobDescription} object.
-
-The migration request will create a new job where the local input files are 
-transferred from old cluster. Once the input files have been transferred the 
-new cluster will send a request to kill the old job. If the kill request is 
-unsuccessful the new job will be terminated unless the \texttt{forcemigration} 
-boolean is \texttt{true}.
-
-The \texttt{Migrate} method will also fill the \texttt{info} {\XMLNode} with 
-information about the job the same way as the \texttt{Submit} method. However 
-for the \texttt{Migrate} method this \texttt{XMLNode} also contain the job-id 
-of the migrated job.
-
-\subsection{{\JobSupervisor}}
-
-The {\JobSupervisor} is responsible for loading the appropriate
-{\JobController}s for managing jobs running on a certain grid
-flavour. Job manipulation can be performed either on individual jobs
-or on groups of jobs (e.g.\ all jobs running on certain cluster or all
-jobs with job state ``FINISHED''), and in order to translate the
-information given by the user into a set of loadable {\JobController}s
-the {\JobSupervisor} makes extensive use of the local joblist file
-which house information about all active jobs. Thus the {\JobSupervisor}
-constructor becomes
-
-\begin{shaded}
-\begin{verbatim}
-JobSupervisor(const UserConfig& usercfg,
-              const std::list<std::string>& jobs,
-              const std::list<std::string>& clusters,
-              const std::string& joblist);
-\end{verbatim}
-\end{shaded}
-
-where \texttt{jobs} is a list of job identifiers and job names,
-\texttt{clusters} is a list of selected and rejected computing
-services in the same format as described above for the
-{\TargetGenerator}, \texttt{joblist} is a string containing the name
-of the joblist file.
-
-Although loaded by the {\JobSupervisor} the {\JobController}
-objects actually reside within the ARC {\Loader} which is a member of
-the {\JobSupervisor} class. In order to get handles on the
-{\JobController}s, the inline method
-
-\begin{shaded}
-\begin{verbatim}
-const std::list<JobController*>& GetJobControllers();
-\end{verbatim}
-\end{shaded}
-
-returns a list of pointers to the loaded {\JobController}s.
-
-\subsection{{\JobController}}
-
-The {\JobController} is a base class for grid specific
-specializations, but also the implementer of all public functionality
-offered by the {\JobController}s. In other words all virtual functions
-of the {\JobController} are private. The initialization of a
-(specialized) {\JobController} object takes two steps. First the
-{\JobController} specialization for the required grid flavour must be
-loaded by the ARC {\Loader}, which sees to that the {\JobController}
-receives information about its flavour (grid) and the local joblist
-file containing information about all active jobs (flavour
-independent). The next step is the filling of the {\JobController}'s job
-pool \texttt{JobStore} which is the pool of jobs that the
-{\JobController} can manage.
-
-\begin{shaded}
-\begin{verbatim}
-void FillJobStore(const std::list<URL>& jobids,
-                  const std::list<URL>& clusterselect,
-                  const std::list<URL>& cluterreject);
-\end{verbatim}
-\end{shaded}
-
-Here \texttt{jobids}, \texttt{clusterselect} and
-\texttt{clusterreject} have been resolved for job names and aliases by
-the {\JobSupervisor}, and no further resolution is needed. The
-following rules are observed when filling the \texttt{JobStore}:
-
-\begin{enumerate}
-\item{If the \texttt{jobids} list has entries, fill \texttt{JobStore}
-  with the jobs requested by the user.}
-\item{If the \texttt{clusterselect} list has entries, fill
-  \texttt{JobStore} with the jobs running on the selected clusters.}
-\item{If clusters are rejected and \texttt{JobStore} has entries,
-  remove from \texttt{JobStore} all jobs running on rejected
-  clusters.}
-\item{If \texttt{jobids} and \texttt{clusterselect} are both empty
-  lists, fill \texttt{JobStore} with all jobs except those running on
-  rejected clusters if there are any.}
-\end{enumerate}
-
-The steps above complete the initialization of the {\JobController}
-which is now ready for handling jobs. The public functions of the
-{\JobController} offer to get (download), clean, cancel, etc. one or
-more jobs and use the virutal methods implemented by the grid middleware 
-specialized {\JobController}s for issuing the command. Here exemplified 
-by the \texttt{Stat} command:
-
-\begin{shaded}
-\begin{small}
-\begin{verbatim}
-bool JobController::Stat(const std::list<std::string>& status,
-                         const bool longlist,
-                         const int timeout) {
-
-  GetJobInformation();
-
-  for (std::list<Job>::iterator it = jobstore.begin();
-       it != jobstore.end(); it++) {
-    if (it->State.empty()) {
-      logger.msg(WARNING, "Job state information not found: %s",
-                 it->JobID.str());
-      Time now;
-      if (now - it->LocalSubmissionTime < 90)
-        logger.msg(WARNING, "This job was very recently "
-                   "submitted and might not yet"
-                   "have reached the information-system");
-      continue;
-    }
-    it->Print(longlist);
-  }
-  return true;
-}
-\end{verbatim}
-\end{small}
-\end{shaded}
-
-The \texttt{Stat} command prints the job information to screen and in
-order to do so the {\JobController} has to query local information
-server for the latest status. Due to different protocols used for
-different grid flavours (e.g.\ ldap for production ARC), the
-\texttt{GetJobInformation} has to be grid flavour specific and is only
-declared as a private virtual method within the {\JobController} base
-class. For details about the flavour specific implementations see
-Section~\ref{sec:plugins}.
-
-The other public methods of the {\JobController} class are:
-The \texttt{Get} which is used to download jobs
-\begin{shaded}
-\begin{verbatim}
-bool Get(const std::list<std::string>& status, const std::string& downloaddir,
-         const bool keep, const int timeout);
-\end{verbatim}
-\end{shaded}
-\begin{shaded}
-\begin{verbatim}
-bool Kill(const std::list<std::string>& status, const bool keep, const int timeout);
-\end{verbatim}
-\end{shaded}
-\begin{shaded}
-\begin{verbatim}
-bool Clean(const std::list<std::string>& status, const bool force, const int timeout);
-\end{verbatim}
-\end{shaded}
-\begin{shaded}
-\begin{verbatim}
-bool Cat(const std::list<std::string>& status, const std::string& whichfile, 
-         const int timeout);
-\end{verbatim}
-\end{shaded}
-\begin{shaded}
-\begin{verbatim}
-bool RemoveJobs(const std::list<URL>& jobids);
-\end{verbatim}
-\end{shaded}
-\begin{shaded}
-\begin{verbatim}
-std::list<std::string> GetDownloadFiles(const URL& dir);
-\end{verbatim}
-\end{shaded}
-\begin{shaded}
-\begin{verbatim}
-bool CopyFile(const URL& src, const URL& dst);
-\end{verbatim}
-\end{shaded}
-The \texttt{GetJobDescriptions} method 
-\begin{shaded}
-\begin{verbatim}
-std::list<Job> GetJobDescriptions(const std::list<std::string>& status, 
-                                  const bool getlocal, const int timeout);
-\end{verbatim}
-\end{shaded}
-which is used to get the job descriptions of {\Job}s in the \texttt{JobStore} 
-with the specified status'. In case \texttt{getlocal} is true the job description 
-will first be searched for the users local XML file. If a local copy of 
-the job description is found the checksum of the local input files will be 
-computed and compared with the ones listed in the users local XML file. If 
-all checksums match the job description is returned.
-
-In case \texttt{getlocal} is false or the retrieval of the local job description 
-fails an attempt will be made to retrieved the job description from the cluster. This has 
-currently only been implemented for ARC(both generations) clusters. Note 
-that the job descriptions retrieved from a cluster will most likely have 
-been changed since the submission -- either by the cluster\footnote{For both 
-generations of ARC it is possible to recover the initial job description send 
-to the cluster. However modifications can still have been made to the 
-job description before sending it to the cluster.} or by the submission 
-client. Therefore, the only way of retrieving the original job description is 
-through the users local XML file.
-
-The \texttt{GetJobDescriptions} returns a list of the  {\Job}s where job description 
-have been found. The found job descriptions are kept in the 
-{\Job}.\texttt{JobDescription} attribute.
-
-The \texttt{Migrate} method
-\begin{shaded}
-\begin{verbatim}
-bool Migrate(TargetGenerator& targetGen,
-             Broker* broker,
-             const bool forcemigration,
-             const int timeout);
-\end{verbatim}
-\end{shaded}
-will try to migrate the jobs contained in the object ({\JobController}). The jobs
-will be migrated to the targets specified in {\TargetGenerator} \texttt{targetGen}
-by using the broker specified by {\Broker} \texttt{broker}. The boolean
-\texttt{forcemigration} specifies whether a migrated job should persist if the
-new cluster ({\ExecutionTarget}) chosen by the {\TargetGenerator} does not
-succeed sending a kill/terminate request for the job. The integer
-\texttt{timeout} argument has not yet been implemented.
-
-\subsection{{\Job}}
-
-The {\Job} is a generic job class for storing all job related
-information. Attributes are derived from the Glue2 information model
-and thus a mapping is needed for non Glue2 compliant grid middlewares.
-
-All attributes of the {\Job} can be printed by the method (longlist = true 
-will give more details (long listing))
-
-\begin{shaded}
-\begin{verbatim}
-void Print(bool longlist) const;
-\end{verbatim}
-\end{shaded}
-
-\subsection{{\UserConfig}}
-
-The {\UserConfig} class handles the client setup i.e. proxy,
-certificate and key location, user and system configuration and local
-joblist location. Upon initialization (constructor) the {\UserConfig}
-locates and reads the user files
-
-\begin{shaded}
-\begin{verbatim}
-$HOME/.arc/client.xml
-$HOME/.arc/jobs.xml
-\end{verbatim}
-\end{shaded}
-
-and if either of them is non-existing a default (empty) one is created.
-
-The {\UserConfig} has three main public methods
-
-\begin{shaded}
-\begin{verbatim}
-const std::string& JobListFile() const;
-const & ConfTree() const;
-bool ApplySecurity(XMLNode& ccfg) const;
-\end{verbatim}
-\end{shaded}
-
-where \texttt{JobListFile()} returns the string pointing to the
-jobs.xml file while \texttt{ConfTree()} returns a configuration
-{\XMLNode} object which is the merge between the user and system
-configurations. In order to do this the method has to locate the
-system configuration and resolve possible conflicts with the user
-configuration. This proceeds through the following chain of actions:
-
-\begin{enumerate}
-\item{Try reading system configuration from \texttt{<ARC Install
-    Location>/etc/arcclient.xml}}
-\item{If the previous step failed try reading system configuration
-  from \texttt{/etc/arcclient.xml}}
-\item{Merge system and user configuration by adding all system
-  configuration not already listed in the user configuration items to the
-  latter.}
-\end{enumerate}
-
-The \texttt{ApplySecurity(XMLNode\& ccfg)} adds security tags to the
-ccfg {\XMLNode} as follows:
-
-\begin{itemize}
-\item{If the \texttt{\$X509\_USER\_PROXY} environment variable is set,
-  use its value to define a \texttt{ProxyPath} tag in \texttt{ccfg}. If
-  the file does not exist or can't be read exit with an error.}
-\item{Otherwise, if the \texttt{\$X509\_USER\_CERT} environment
-  variable is set, use its value and the value of the
-  \texttt{\$X509\_USER\_KEY} environment variable to define
-  \texttt{CertificatePath} and \texttt{KeyPath} tags in \texttt{ccfg}.
-  If \texttt{\$X509\_USER\_KEY} is not set or either file does not
-  exist or can't be read exit with an error.}
-\item{Otherwise, if the merged user configuration tree (see
-  \texttt{ConfTree}) contains a \texttt{ProxyPath} tag, copy it to
-  \texttt{ccfg}. If the file does not exist or can't be read exit with
-  an error.}
-\item{Otherwise, if the merged user configuration tree contains a
-  \texttt{CertificatePath} tag, copy it to \texttt{ccfg} along with the
-  accompanying \texttt{KeyPath} tag. If the \texttt{KeyPath} tag is
-  undefined or either file does not exist or can't be read exit with
-  an error.}
-\item{Otherwise, if the file \texttt{/tmp/x509up\_u + userid} exists
-  add a \texttt{ProxyPath} tag in \texttt{ccfg} pointing to it.}
-\item{Otherwise, add \texttt{CertificatePath} and \texttt{KeyPath}
-  tags to \texttt{ccfg} pointing to
-  \texttt{\$HOME/.globus/usercert.pem} and
-  \texttt{\$HOME/.globus/userkey.pem}, respectively. If either file
-  does not exist exit with error.}
-\item{If the \texttt{\$X509\_CERT\_DIR} environment variable is set,
-  use its value to define a \texttt{CACertificatesDir} tag in
-  \texttt{ccfg}. If the directory does not exist, exit with an error.}
-\item{Otherwise, if the user ID is not zero and the directory
-  \texttt{\$HOME/.globus/certificates} exists add a
-  \texttt{CACertificatesDir} tag in \texttt{ccfg} pointing to it.}
-\item{Otherwise, add a \texttt{CACertificatesDir} tag in \texttt{ccfg}
-  pointing to \texttt{/etc/grid-security/certificates}. If the
-  directory does not exist, exit with an error.}
-\end{itemize}
-
-\subsection{{\ACCConfig}}
-
-Locating Arc Client Components (plugins) is handled by the
-{\ACCConfig} class. It inherits from \texttt{BaseConfig} and
-implements only one method
-
-\begin{shaded}
-\begin{verbatim}
-virtual XMLNode MakeConfig(XMLNode cfg) const;
-\end{verbatim}
-\end{shaded}
-
-The \texttt{MakeConfig} method searches plugin paths for all libraries
-named \texttt{libacc*}. Matching libraries are added as plugins to the
-configuration object \texttt{cfg}.
-
-\subsection{{\ClientInterface}}
-
-The {\ClientInterface} class is a utility base class used for
-configuring a client side Message Chain Component (MCC) chain and
-loading it into memory. It has several specializations of increasing
-complexity of the MCC chains.
-
-\subsubsection{{\ClientTCP}}
-
-The {\ClientTCP} class is a specialization of the {\ClientInterface}
-which sets up a client MCC chain for TCP communication, and optionally
-with a TLS layer on top.
-
-\subsubsection{{\ClientHTTP}}
-
-The {\ClientHTTP} class inherits from the {\ClientTCP} class and adds
-an HTTP MCC to the chain.
-
-\subsubsection{{\ClientSOAP}}
-
-The {\ClientSOAP} class inherits from the {\ClientHTTP} class and adds
-a SOAP MCC to the chain. Specializations of the {\TargetRetriever},
-{\Submitter} and {\JobController} classes that communicate with SOAP
-based services make use of this class.
-
-\subsection{{\Sandbox}}
-The {\Sandbox} class, consisting of the two methods \texttt{Add} and 
-\texttt{GetCkSum}, have been created to assure a reliable resubmission 
-of jobs. 
-\begin{shaded}
-\begin{verbatim}
-static bool Add(JobDescription& jobdesc, XMLNode& info);
-static std::string GetCksum(const std::string& file);
-\end{verbatim}
-\end{shaded}
-The \texttt{Add} method  adds the job description and the checksum of 
-the local input files to a {\XMLNode}. Currently this {\XMLNode} is the one 
-stored in the user's local XML file upon job submission. The job 
-description which is stored is the initial job description before any 
-parsing has been performed. To calculate the checksum of the input 
-files the \texttt{GetCkSum} method is used. The \texttt{GetCkSum} 
-method returns the MD5 checksum of a given file
-
-\section{Specialized Classes (Grid Flavour and Broker plugins)}
-\label{sec:plugins}
-
-\subsection{ARC0 Plugins}
-
-The ARC0 plugins enables support for the interfaces used by computing
-elements running ARC version 0.x.
-
-The ARC 0.x local information system uses the {\GT}~\cite{globus} GRIS
-with a custom made ARC schema. The information index server used is the
-{\GT} GIIS. Both these servers use the LDAP~\cite{ldap}
-protocol. The specialization of the {\TargetRetriever} class for ARC0
-is implemented using the ARC LDAP Data Management Component (DMC) (see
-\cite{hed} for technical details).
-
-Jobs running on an ARC 0.x computing element are handled by the ARC
-grid-manager~\cite{gm}. Job submission and job control are done using
-the gridftp~\cite{gridftp} protocol. The specializations of the
-{\Submitter} and {\JobController} classes use the globus ftp control
-library.
-
-Stage-in and stage-out of input and output files are also done using
-the gridftp~\cite{gridftp} protocol. This means that proper
-functionality of the ARC0 plugins requires the gridftp DMC.
-
-\subsection{ARC1 Plugins}
-
-The computing element in ARC version 1.x is the A-Rex~\cite{arex}
-service running in a HED~\cite{hed} container.
-
-A-Rex implements the BES~\cite{ogsa-bes} standard interface. Since
-this is a SOAP-based~\cite{soap} interface, the specializations of the
-{\TargetRetriever}, {\Submitter} and {\JobController} classes make use
-of a chain of ARC Message Chain Components (MCC~\cite{hed}) ending
-with the SOAP client MCC.
-
-The A-Rex service uses the https protocol \texttt{put} and \texttt{get} methods for
-stage-in and stage-out of input and output files. Therefore, the ARC1
-plugins requires the http DMC.
-
-\subsection{gLite Plugins}
-
-The gLite computing element offers several interfaces, one of them
-being the Web Service based computing element interface known as the
-CREAM CE~\cite{cream}. The current revision of this interface (CREAM
-version 2) has been chosen for implementation within {\libarcclient} for the
-following reasons:
-
-\begin{itemize}
-\item CREAM2 has a Web Service interface that is very similar to the Web Service
-  based ARC.
-\item CREAM2 enables direct access to the gLite computing element
-  without having to go via the gLite workload management system.
-\item CREAM2 contains numerous improvements when compared to the
-  earlier CREAM versions.
-\item CREAM2 supports direct job status queries.
-\item CREAM2 offers a convenient way of handling input and output
-  files through accessing the input and output sandbox via GridFTP.
-\end{itemize}
-
-gLite resources are registered in top level and site BDIIs. The CREAM
-specialization of the {\TargetRetriever} therefore makes use of the
-LDAP DMC similarly to the ARC0 plugins.
-
-CREAM has its own SOAP-based interface. The CREAM
-specializations of the {\Submitter} and {\JobController} classes
-therefore use an MCC chain ending with the SOAP client MCC the same 
-way the ARC1 plugin does.
-
-Stage-in and stage-out of input and output files are done using the
-gridftp protocol. The gridftp DMC is therefore required.
-
-\subsection{Broker plugins}
+\section{Broker plugins}
 \label{sec:brokerplugins}
@@
 
-\subsubsection{\Benchmark}
+\subsection{\Benchmark}
 
@@
 
-\subsubsection{\FastestQueueBroker}
+\subsection{\FastestQueueBroker}
 
@@
 
-\subsubsection{\Data}
+\subsection{\Data}
 
@@
 
-\subsubsection{\PythonBroker}
+\subsection{\PythonBroker}
 
@@
 
-\chapter{Building Command Line Interfaces}
-\label{sec:cli}
+\chapter{Job Description}
+\label{sec:jobdesc}
 
-Given all components listed above it is possible to write versatile
-command line interfaces (cli) for grid job submission and
-management. {\libarcclient} offers the following native commands:
-\begin{enumerate}
-\item{\texttt{arcstat} - for computing resource or grid job status 
-queries.}
-\item{\texttt{arcsub} - for grid job submission.}
-\item{\texttt{arcget} - for downloading output of finished, cancelled 
-or failed grid jobs.}
-\item{\texttt{arckill} - for terminating grid jobs.}
-\item{\texttt{arcclean} - for removing a grid job's session directory 
-including all contents.}
-\item{\texttt{arccat} - for performing the \texttt{cat} command to a 
-running grid job's std out or std error file.}
-\item{\texttt{arcsync} - for synchronising the information in the user's 
-local XML file\footnote{By default \$HOME/.arc/jobs.xml} about active jobs.}
-\item{\texttt{arcresub} - for starting an entire job over on either the 
-same or a new computing resource.}
-\item{\texttt{arcmigrate} - for moving a queuing job from one A-REX 
-server to another.}
-\end{enumerate}
+\chapter{Grid Flavour Plugins)}
+\label{sec:plugins}
 
-Each of the commands above are encoded within one C++ file with the
-following structure, here exemplified with the arcget command:
+\section{ARC0 Plugins}
 
-\begin{shaded}
-\begin{verbatim}
-#ifdef HAVE_CONFIG_H
-#include <config.h>
-#endif
+The ARC0 plugins enables support for the interfaces used by computing
+elements running ARC version 0.x.
 
-#include <iostream>
-#include <list>
-#include <string>
+The ARC 0.x local information system uses the {\GT}~\cite{globus} GRIS
+with a custom made ARC schema. The information index server used is the
+{\GT} GIIS. Both these servers use the LDAP~\cite{ldap}
+protocol. The specialization of the {\TargetRetriever} class for ARC0
+is implemented using the ARC LDAP Data Management Component (DMC) (see
+\cite{hed} for technical details).
 
-#include <arc/ArcLocation.h>
-#include <arc/IString.h>
-#include <arc/Logger.h>
-#include <arc/OptionParser.h>
-#include <arc/client/JobController.h>
-#include <arc/client/JobSupervisor.h>
-#include <arc/client/UserConfig.h>
+Jobs running on an ARC 0.x computing element are handled by the ARC
+grid-manager~\cite{gm}. Job submission and job control are done using
+the gridftp~\cite{gridftp} protocol. The specializations of the
+{\Submitter} and {\JobController} classes use the globus ftp control
+library.
 
-int main(int argc, char **argv) {
+Stage-in and stage-out of input and output files are also done using
+the gridftp~\cite{gridftp} protocol. This means that proper
+functionality of the ARC0 plugins requires the gridftp DMC.
 
-  setlocale(LC_ALL, "");
+\section{ARC1 Plugins}
 
-  Arc::Logger logger(Arc::Logger::getRootLogger(), "arcget");
-  Arc::LogStream logcerr(std::cerr);
-  Arc::Logger::getRootLogger().addDestination(logcerr);
-  Arc::Logger::getRootLogger().setThreshold(Arc::WARNING);
+The computing element in ARC version 1.x is the A-Rex~\cite{arex}
+service running in a HED~\cite{hed} container.
 
-  Arc::ArcLocation::Init(argv[0]);
+A-Rex implements the BES~\cite{ogsa-bes} standard interface. Since
+this is a SOAP-based~\cite{soap} interface, the specializations of the
+{\TargetRetriever}, {\Submitter} and {\JobController} classes make use
+of a chain of ARC Message Chain Components (MCC~\cite{hed}) ending
+with the SOAP client MCC.
 
-  Arc::OptionParser options(istring("[job ...]"),
-                            istring("The arcget command is used for "
-                                    "retrieving the results from a job."),
-                            istring("Argument to -c has the format "
-                                    "Flavour:URL e.g.\n"
-                                    "ARC0:ldap://grid.tsl.uu.se:2135/"
-                                    "nordugrid-cluster-name=grid.tsl.uu.se,"
-                                    "Mds-Vo-name=local,o=grid"));
+The A-Rex service uses the https protocol \texttt{put} and \texttt{get} methods for
+stage-in and stage-out of input and output files. Therefore, the ARC1
+plugins requires the http DMC.
 
-  bool all = false;
-  options.AddOption('a', "all",
-                    istring("all jobs"),
-                    all);
+\section{gLite Plugins}
 
-  // Removed most of the option definition from this write-up to
-  // save space. See the real source file for the complete list.
-  // ...
+The gLite computing element offers several interfaces, one of them
+being the Web Service based computing element interface known as the
+CREAM CE~\cite{cream}. The current revision of this interface (CREAM
+version 2) has been chosen for implementation within {\libarcclient} for the
+following reasons:
 
-  bool version = false;
-  options.AddOption('v', "version", istring("print version information"),
-                    version);
+\begin{itemize}
+\item CREAM2 has a Web Service interface that is very similar to the Web Service
+  based ARC.
+\item CREAM2 enables direct access to the gLite computing element
+  without having to go via the gLite workload management system.
+\item CREAM2 contains numerous improvements when compared to the
+  earlier CREAM versions.
+\item CREAM2 supports direct job status queries.
+\item CREAM2 offers a convenient way of handling input and output
+  files through accessing the input and output sandbox via GridFTP.
+\end{itemize}
 
-  std::list<std::string> jobs = options.Parse(argc, argv);
+gLite resources are registered in top level and site BDIIs. The CREAM
+specialization of the {\TargetRetriever} therefore makes use of the
+LDAP DMC similarly to the ARC0 plugins.
 
-  if (!debug.empty())
-    Arc::Logger::getRootLogger().setThreshold(Arc::string_to_level(debug));
+CREAM has its own SOAP-based interface. The CREAM
+specializations of the {\Submitter} and {\JobController} classes
+therefore use an MCC chain ending with the SOAP client MCC the same 
+way the ARC1 plugin does.
 
-  Arc::UserConfig usercfg(conffile);
-  if (!usercfg) {
-    logger.msg(Arc::ERROR, "Failed configuration initialization");
-    return 1;
-  }
-
-  if (debug.empty() && usercfg.ConfTree()["Debug"]) {
-    debug = (std::string)usercfg.ConfTree()["Debug"];
-    Arc::Logger::getRootLogger().setThreshold(Arc::string_to_level(debug));
-  }
-
-  if (version) {
-    std::cout << Arc::IString("%s version %s", "arcget", VERSION)
-              << std::endl;
-    return 0;
-  }
-
-  if (jobs.empty() && joblist.empty() && !all) {
-    logger.msg(Arc::ERROR, "No jobs given");
-    return 1;
-  }
-
-  if (joblist.empty())
-    joblist = usercfg.JobListFile();
-
-  Arc::JobSupervisor jobmaster(usercfg, jobs, clusters, joblist);
-
-  std::list<Arc::JobController*> jobcont = jobmaster.GetJobControllers();
-
-  if (jobcont.empty()) {
-    logger.msg(Arc::ERROR, "No job controllers loaded");
-    return 1;
-  }
-
-  int retval = 0;
-  for (std::list<Arc::JobController*>::iterator it = jobcont.begin();
-       it != jobcont.end(); it++)
-    if (!(*it)->Get(status, downloaddir, keep, timeout))
-      retval = 1;
-
-  return retval;
-}
-\end{verbatim}
-\end{shaded}
+Stage-in and stage-out of input and output files are done using the
+gridftp protocol. The gridftp DMC is therefore required.
 
@@
 
-\chapter{Job Inner Representation}
-\label{app:jobinnerrepresentation}
+\chapter{Job Status mapping}
+\label{app:broker-mapping}
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/client/JobInnerRepresentation.ods
+http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/client/broker-mapping.ods
 

### 2009-11-27 src/hed/libs/client/JobController.h

@@ -20,6 +20,20 @@ namespace Arc {
   class UserConfig;
 
   //! Base class for the JobControllers
+  /** The JobController is the base class for middleware specialized
+   *  derived classes. The JobController base class is also the
+   *  implementer of all public functionality that should be offered
+   *  by the middleware specific specializations. In other words all
+   *  virtual functions of the JobController are private. The
+   *  initialization of a (specialized) JobController object takes two
+   *  steps. First the JobController specialization for the required
+   *  grid flavour must be loaded by the JobControllerLoader, which
+   *  sees to that the JobController receives information about its
+   *  Grid flavour and the local joblist file containing information
+   *  about all active jobs (flavour independent). The next step is
+   *  the filling of the JobController job pool (JobStore) which is
+   *  the pool of jobs that the JobController can manage.
+  **/
   /// Must be specialiced for each supported middleware flavour.
   class JobController
     : public Plugin {
@@ -29,6 +43,15 @@ namespace Arc {
   public:
     virtual ~JobController();
 
+    /// Fill jobstore
+    /**
+     * Method to fill the jobstore with jobs that should be managed. 
+     *
+     * @param jobids List of jobids to be loaded to the jobstore. If
+     * empty all jobs of the specialized grid flavour present in the
+     * joblist file (given through the usercfg to the constructor)
+     * will be loaded to the jobstore.
+     **/
     void FillJobStore(const std::list<URL>& jobids);
     void FillJobStore(const Job& job);
 
@@ -72,6 +95,19 @@ namespace Arc {
     bool PrintJobStatus(const std::list<std::string>& status,
                         const bool longlist);
 
+    /// Migrate job from cluster A to Cluster B
+    /**  Method to migrate the jobs contained in the jobstore.
+     *
+     * @param targetGen TargetGenerator with targets to migrate the
+     * job to.
+     *
+     * @param broker Broker to be used when selecting target.
+     *
+     * @param forcemigration boolean which specifies whether a
+     * migrated job should persist if the new cluster does not succeed
+     * sending a kill/terminate request for the job.
+     * 
+     **/
     bool Migrate(TargetGenerator& targetGen,
                  Broker *broker,
                  const UserConfig& usercfg,

### 2009-11-27 doc/tech_doc/client/client_technical.tex
fixing the broker section

@@
 All capabilities are enabled for three different grid flavours
-(Production ARC, ARC1 and gLite~\cite{glite}) through a modular design
-using plugins specialized for each supported middleware. Future
-extensions to support additional middlewares involve plugin
-compilation only i.e., no recompilation of main libraries or clients
-is necessary.
+(Production ARC, Web Service ARC and gLite~\cite{glite}) through a
+modular design using plugins specialized for each supported
+middleware. Future extensions to support additional middlewares
+involve plugin compilation only i.e., no recompilation of main
+libraries or clients is necessary.
 
 Using the library, a set of command line tools have been built which
-puts the library's functionality at the fingertips of users. While
-this documentation will illustrate how such command line tools can be
-built, the main documentation of the command line tools is given in
-the client user manual~\cite{ui}.
-
-In the following we will give a functionality overview in
-Section~\ref{sec:FuncOver} while all technical details will be given
-in Section~\ref{sec:Implementation}. Section~\ref{sec:cli} will
-show through examples how command line interfaces can be built upon
-the library.
+puts the library's functionality at the fingertips of users. This
+documentation is not intended to document the developed CLI but the
+concept of how to build command line tools will be
+demonstrated. Readers interested in the user interface are recommended
+to read the client user manual~\cite{ui}.
 
@@
 is given which also introduces the most important higher level
-classes.
+classes. Readers interested in the library API are recommended to look
+up the online API documentation~\cite{libarcclientAPI}.
 
@@
 \section{Job Management}
-
 Once a job is submitted, job related information (job identification
@@
 \label{sec:brokering}
-
-The {\Broker} inherits from the {\ACC} base class in order to be
-loadable by the ARC \texttt{Loader}. It is the base class of the
-specialized {\Broker} and it implements the method for reducing a list
+The ARC brokering module is made up of two kinds of classes: one
+brokering base class and specialized classes derived thereof. The
+brokering base class which implements the method for reducing a list
 of resources found by resource discovery (the list of
@@
 
-The {\Benchmark} ranks the {\ExecutionTarget}s according to their CPU performance. The most 
-reliable comparison is achieved through the usage of a benchmark scenario, and the {\Benchmark}
-uses the CINT2000 (Integer Component of SPEC CPU2000)\footnote{http://www.spec.org/cpu2000/CINT2000/} 
-benchmark for this purpose.
-
-The {\texttt SortTargets} method has two steps:
-
-\begin{enumerate}
-\item{{\ExecutionTarget}s not publishing the specint2000 benchmark information are removed from the list 
-  of possible targets as their performance is unknown.}
-\item{The {\ExecutionTarget}s are ranked such that the {\ExecutionTarget} with highest specint2000 value 
-  receives top ranking}
-\end{enumerate}
+The {\Benchmark} ranks the {\ExecutionTarget}s according to their
+benchmark performance. Through the command line tool (see the user
+manual\cite{ui} for reference) this specialized broker takes a user
+specified benchmark as input and ranks the {\ExecutionTarget}s
+according to their published benchmark performance. If no benchmark is
+specified the CINT2000 (Integer Component of SPEC
+CPU2000)\footnote{http://www.spec.org/cpu2000/CINT2000/} benchmark is
+used as default.
 
@@
 
-The {\FastestQueueBroker} ranks the {\ExecutionTarget}s according to their queue length measured in percentage 
-of the {\ExecutionTarget}'s size (i.e. the queue length divided by the number of total slots/CPUs). If more 
-than one {\ExecutionTarget} has zero queue, the {\FastestQueueBroker} makes use of a basic load balancing method 
-to rearrange the zero queue {\ExecutionTarget}s depending on their number of free slots/CPUs.
-
-The {\texttt SortTargets} method has three steps:
-
-\begin{enumerate}
-\item{{\ExecutionTarget}s not publishing WaitingJobs, TotalSlots and FreeSlots are removed from the list of possible targets 
-as it is impossible to determine their queue length and perform load balancing.}
-\item{The {\ExecutionTarget}s are ranked according to their queue length relative to the number of slots/CPUs. Zero queue 
-{\ExecutionTarget}s is ranked first.}
-\item{If more than one {\ExecutionTarget} has zero queue, the targets are randomly reshuffled based upon how many free slots 
-they have, e.g. if two {\ExecutionTarget}s have zero queue and they have 90 and 10 free slots respectively, then the first
-{\ExecutionTarget} has a 90\% probability of becoming the top ranked cluster.}
-\end{enumerate}
+The {\FastestQueueBroker} ranks the {\ExecutionTarget}s according to
+their queue length measured in percentage of the {\ExecutionTarget}'s
+size (i.e. the queue length divided by the number of total
+slots/CPUs). If more than one {\ExecutionTarget} has zero queue, the
+{\FastestQueueBroker} makes use of a basic load balancing method to
+rearrange the zero queue {\ExecutionTarget}s depending on their number
+of free slots/CPUs.
 
@@
 
-The {\Data} ranks the {\ExecutionTarget}s according to how many megabytes of the requested input files that already 
-stored in the cache of the computing resource the {\ExecutionTarget} represents. The broker is motivated by that jobs should 
-be submitted to {\ExecutionTarget}s where the data already is, thus reducing the network load on both the computing resource 
-and client side. The ranking method is based upon the A-REX\footnote{http://www.knowarc.eu/download/D1.2-3\_documentation.pdf} 
-interface {\texttt CacheCheck} for querying for the presence of the file in the cache directory. This interface has a limitation 
-in the current implementation and does not support per user caches.
+The {\Data} ranks the {\ExecutionTarget}s according to how many
+megabytes of the requested input files that already stored in the
+cache of the computing resource the {\ExecutionTarget} represents. The
+broker is motivated by that jobs should be submitted to
+{\ExecutionTarget}s where the data already is, thus reducing the
+network load on both the computing resource and client side. The
+ranking method is based upon the
+A-REX\footnote{http://www.knowarc.eu/download/D1.2-3\_documentation.pdf}
+interface {\texttt CacheCheck} for querying for the presence of the
+file in the cache directory. This interface has a limitation in the
+current implementation and does not support per user caches.
 
@@
 \begin{enumerate}
-\item{Only the A-REX service has {\texttt CacheCheck} method, thus {\ExecutionTarget}s not running A-REX are removed.}
-\item{Information about input files requested in the job description is collected from the \texttt{JobInnerRepresentation.}}
-\item{Each {\ExecutionTarget} is queried (through the {\texttt CacheCheck} method) for the existence of input files. 
-A single query is used for achieving all the necessary information and file sizes are summarized. If there are problems 
-determining file sizes, then the summarized size will be zero.}
-\item{The possible {\ExecutionTarget}s are ranked in a descending order according to the amount of input data they have in their cache.}
+\item{Only the A-REX service has {\texttt CacheCheck} method, thus
+  {\ExecutionTarget}s not running A-REX are removed.}
+\item{Information about input files requested in the job description
+  is collected from the \texttt{JobInnerRepresentation.}}
+\item{Each {\ExecutionTarget} is queried (through the {\texttt
+    CacheCheck} method) for the existence of input files.  A single
+  query is used for achieving all the necessary information and file
+  sizes are summarized. If there are problems determining file sizes,
+  then the summarized size will be zero.}
+\item{The possible {\ExecutionTarget}s are ranked in a descending
+  order according to the amount of input data they have in their
+  cache.}
 \end{enumerate}
@@
 
-\bibliography{grid}
+\bibliography{grid, libarcclient}
 

### 2009-11-27 doc/tech_doc/client/client_technical.tex
Updating client technical.

@@
 \usepackage{longtable}             %for breaking tables
+\usepackage{multirow}
+\usepackage{booktabs}
+\usepackage{ltablex}
+\usepackage[table]{xcolor} % Note: The rowcolors command does not work properly with the tabularx environment.
+
 %
@@
 \newcommand{\XMLNode}{\texttt{XMLNode}}
+\newcommand{\NST}[1]{\normalsize{\textnormal{#1}}}
+\newcommand{\defineheadfoot}{\endfirsthead %
+\midrule %
+\multicolumn{5}{c}{Continued from previous page}\\%
+\midrule %
+\endhead %
+\midrule %
+\multicolumn{5}{c}{Continues at next page}\\%
+\midrule %
+\endfoot %
+\bottomrule %
+\endlastfoot}
 \definecolor{shadecolor}{rgb}{1,1,0.6}
@@
 \definecolor{cyan}{rgb}{0,1,1}
+
+\colorlet{tableheadcolor}{gray!25}
+\colorlet{tablerowcolor}{gray!12.5}
+
+\newcolumntype{C}{>{\ttfamily\footnotesize}c}
+\newcolumntype{T}{>{\ttfamily\footnotesize\centering\arraybackslash}X}
+
 %
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/client/broker-mapping.ods
+This document contains a description of the broker used in the Advanced
+Resource Connector (ARC) middleware as of version 1.0. In section
+\ref{sec:Attribute-matching} a description of how target attributes
+are matched and compared to attributes specified in the job description.
+
+In the broker attributes published by the candidate target need to
+be matched and compared to attributes specified in the job description
+to be able to select targets which are suitable for accepting the
+job description. Targets which do not pass a certain comparison will
+be not be a candidate target for submission. Detailed information
+about the attributes described below can be found in the ARC job description
+document. First the broker will check some basics for the target being
+processed. First, if the job description specifies submission to specific
+targets and/or queues, the target will be checked and if it is not
+among these targets and/or queues, it will be dropped. Next the number
+of free slots on the target will be checked, and if no free slots
+exist, then it will be dropped. After this, the health state of the
+target is checked, and if any other value than "ok"
+is published then it is dropped. Last, if middleware specific submission
+is specified in the job description, the middleware type of the target
+is checked, and if it does not satisfy the middleware requirements
+then it is dropped. After these basic checks, certain attributes specified
+in the job description are compared to the attributes published by
+the target being processed. If the target does not publish a attribute
+which is specified in the job description the target will dropped,
+with some exceptions.
+
+\begin{table}
+\footnotesize{\texttt{
+\begin{tabular}{cccc}
+\toprule
+\textnormal{\normalsize Job description attributes} & \textnormal{\normalsize Comparator} & \textnormal{\normalsize Target attribute} & \tabularnewline
+\midrule
+Resources.CandidateTarget[i].EndPointURL & $==$ & url & \tabularnewline
+Resources.CandidateTarget[i].QueueName & $==$ & MappingQueue & \tabularnewline
+\midrule
+Application.ProcessingStartTime    & $<$ & DowntimeStarts & \multirow{2}{*}{AND} \tabularnewline
+Application.ProcessingStartTime    & $>$ & DowntimeEnds &\tabularnewline
+\midrule
+Resources.CEType & isSatisfied( & Implementation & ) \tabularnewline
+\midrule
+Resources.TotalWallTime.range.max & $<$ & MaxWallTime & \tabularnewline
+Resources.TotalWallTime.range.min & $>$ & MinWallTime & \tabularnewline
+Resources.TotalCPUTime.range.max & $<$ & MaxCPUTime & \tabularnewline
+Resources.TotalCPUTime.range.min & $>$ & MinCPUTime & \tabularnewline
+\midrule
+Resources.IndividualPhysicalMemory & $\leq$ & MainMemory & \multirow{2}{*}{OR} \tabularnewline
+Resources.IndividualPhysicalMemory & $\leq$ & MaxMemory &\tabularnewline
+Resources.IndividualVirtualMemory & $\leq$ & MaxVirtualMemory & \tabularnewline
+\midrule
+Resources.Platform & $==$ & Platform & \tabularnewline
+\midrule
+Resources.OperatingSystem & isSatisfied( & OperatingSystem & )\tabularnewline
+Resources.RunTimeEnvironment & isSatisfied( & ApplicationEnvironments & )\tabularnewline
+\midrule
+Resources.NetworkInfo & in & NetworkInfo & \tabularnewline
+\midrule
+Resources.DiskSpaceRequirement.SessionDiskSpace & $\leq$ & MaxDiskSpace & \multirow{2}{*}{OR} \tabularnewline
+Resources.DiskSpaceRequirement.SessionDiskSpace & $\leq$ & WorkingAreaTotal & \tabularnewline
+\midrule
+Resources.DiskSpaceRequirement.DiskSpace & \multirow{2}{*}{$\leq$} & \multirow{2}{*}{MaxDiskSpace} & \multirow{4}{*}{OR} \tabularnewline
+$-$ Resources.DiskSpaceRequirement.CacheDiskSpace & & & \tabularnewline
+Resources.DiskSpaceRequirement.DiskSpace & \multirow{2}{*}{$\leq$} & \multirow{2}{*}{WorkingAreaTotal} & \tabularnewline
+$-$ Resources.DiskSpaceRequirement.CacheDiskSpace & & & \tabularnewline
+\midrule
+Resources.DiskSpaceRequirement.DiskSpace & $\leq$ & MaxDiskSpace & \multirow{2}{*}{OR} \tabularnewline
+Resources.DiskSpaceRequirement.DiskSpace & $\leq$ & WorkingAreaTotal & \tabularnewline
+Resources.DiskSpaceRequirement.CacheDiskSpace & $\leq$ & CacheTotal & \tabularnewline
+\midrule
+Resources.SlotRequirement.NumberOfSlots & $\leq$ & TotalSlots & \tabularnewline
+Resources.SlotRequirement.NumberOfSlots & $\leq$ & MaxSlotsPerJob & \tabularnewline
+\midrule
+Resources.SessionLifeTime & $\leq$ & WorkingAreaLifeTime & \tabularnewline
+\midrule
+Resources.NodeAccess is NAT\_{}INBOUND OR NAT\_{}INOUTBOUND & AND & ConnectivityIn & \tabularnewline
+Resources.NodeAccess is NAT\_{}OUTBOUND OR NAT\_{}INOUTBOUND & AND & ConnectivityOut & \tabularnewline
+\bottomrule
+\end{tabular}
+}}
+\caption{}
+\end{table}
 
@@
 
-http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/doc/client/broker-mapping.ods
+The following job states comprises the state model in WS-ARC:
+\begin{description}
+\item[UNDEFINED] Job state could not be resolved,
+\item[ACCEPTED] Job was accepted by the service,
+\item[PREPARING] Job is preparing,
+\item[SUBMITTING] Job is being submitted to a computing share,
+\item[HOLD] Job is put on hold,
+\item[QUEUING] Job is on computing share waiting to run,
+\item[RUNNING] Job is running on computing share,
+\item[FINISHING] Job is finishing,
+\item[FINISHED] Job has finished,
+\item[KILLED] Job has been killed,
+\item[FAILED] Job failed,
+\item[DELETED] Job have been deleted,
+\item[OTHER] Any job state which does not fit the above states.
+\end{description}
+
+\begin{center}
+\rowcolors{3}{tablerowcolor}{}
+\begin{tabularx}{\textwidth}{CCTCT}
+\toprule
+\NST{Internal} & \NST{ARC0}  & \NST{ARC1} & \NST{BES} & \NST{CREAM}\\
+\midrule
+\defineheadfoot
+ACCEPTED & ACCEPTED & ACCEPTED & Pending & REGISTERED,\linebreak PENDING\\
+PREPARING & PREPARING & PREPARING & \NST{None} & \NST{None} \\
+SUBMITTING & SUBMIT & SUBMIT & \NST{None} & \NST{None} \\
+HOLD & \NST{None} & \NST{None} & \NST{None} & HELD \\
+QUEUING & INLRMS:Q & INLRMS:Q & \NST{None} & IDLE \\
+RUNNING & INLRMS:R & INLRMS:R,\linebreak INLRMS:EXECUTED,\linebreak INLRMS:S, INLRMS:E & Running & RUNNING,\linebreak REALLY-RUNNING\\
+FINISHING & FINISHING & FINISHING & \NST{None} & \NST{None}\\
+FINISHED & FINISHED & FINISHED & Finished & DONE-OK \\
+KILLED & KILLED & KILLED & \NST{None} & CANCELLED \\
+FAILED & FAILED & FAILED & Failed & DONE-FAILED,\linebreak ABORTED\\
+DELETED & DELETED & DELETED & \NST{None} & \NST{None} \\
+OTHER & \NST{Any other state} & \NST{Any other state} & \NST{None} & \NST{Any other state} \\
+\end{tabularx}
+\end{center}
 

### 2009-11-27 doc/tech_doc/client/client_technical.tex
updated brokering appendix

@@
 \label{app:broker-mapping}
+The table below shows the relation between {\JobDescription}
+attributes and the {\ExecutionTarget} attributes when the {\Broker}
+base class filter the available targets to find those suitable for job
+submission.
 
-This document contains a description of the broker used in the Advanced
-Resource Connector (ARC) middleware as of version 1.0. In section
-\ref{sec:Attribute-matching} a description of how target attributes
-are matched and compared to attributes specified in the job description.
-
-In the broker attributes published by the candidate target need to
-be matched and compared to attributes specified in the job description
-to be able to select targets which are suitable for accepting the
-job description. Targets which do not pass a certain comparison will
-be not be a candidate target for submission. Detailed information
-about the attributes described below can be found in the ARC job description
-document. First the broker will check some basics for the target being
-processed. First, if the job description specifies submission to specific
-targets and/or queues, the target will be checked and if it is not
-among these targets and/or queues, it will be dropped. Next the number
-of free slots on the target will be checked, and if no free slots
-exist, then it will be dropped. After this, the health state of the
-target is checked, and if any other value than "ok"
-is published then it is dropped. Last, if middleware specific submission
-is specified in the job description, the middleware type of the target
-is checked, and if it does not satisfy the middleware requirements
-then it is dropped. After these basic checks, certain attributes specified
-in the job description are compared to the attributes published by
-the target being processed. If the target does not publish a attribute
-which is specified in the job description the target will dropped,
-with some exceptions.
+First, if the job description specifies submission to specific targets
+and/or queues, the target will be checked and if it is not among these
+targets and/or queues, it will be dropped. Next the number of free
+slots on the target will be checked, and if no free slots exist, then
+it will be dropped. After this, the health state of the target is
+checked, and if any other value than "ok" is published then it is
+dropped. Last, if middleware specific submission is specified in the
+job description, the middleware type of the target is checked, and if
+it does not satisfy the middleware requirements then it is
+dropped. After these basic checks, certain attributes specified in the
+job description are compared to the attributes published by the target
+being processed. If the target does not publish a attribute which is
+specified in the job description the target will dropped, with some
+exceptions.
 

### 2009-11-27 doc/tech_doc/client/client_technical.tex
Updating client technical.

@@
 
-\chapter{Grid Flavour Plugins)}
+\chapter{Grid Flavour Plugins and Commons}
 \label{sec:plugins}
+With a library utilizing plugins a mapping to a common set of job states
+is needed in order to be able to treat jobs from different middlewares
+in the same way. In WS-ARC the following job states comprises the state
+model which plugins should map their job states:
+\begin{description}
+\item[UNDEFINED] Job state could not be resolved,
+\item[ACCEPTED] Job was accepted by the service,
+\item[PREPARING] Job is preparing,
+\item[SUBMITTING] Job is being submitted to a computing share,
+\item[HOLD] Job is put on hold,
+\item[QUEUING] Job is on computing share waiting to run,
+\item[RUNNING] Job is running on computing share,
+\item[FINISHING] Job is finishing,
+\item[FINISHED] Job has finished,
+\item[KILLED] Job has been killed,
+\item[FAILED] Job failed,
+\item[DELETED] Job have been deleted,
+\item[OTHER] Any job state which does not fit the above states.
+\end{description}
+See Appendix~\ref{app:job-status-mapping} for the actual job state
+mapping of each of the below described plugins
 
@@
 \chapter{Job Status mapping}
-\label{app:broker-mapping}
-
-The following job states comprises the state model in WS-ARC:
-\begin{description}
-\item[UNDEFINED] Job state could not be resolved,
-\item[ACCEPTED] Job was accepted by the service,
-\item[PREPARING] Job is preparing,
-\item[SUBMITTING] Job is being submitted to a computing share,
-\item[HOLD] Job is put on hold,
-\item[QUEUING] Job is on computing share waiting to run,
-\item[RUNNING] Job is running on computing share,
-\item[FINISHING] Job is finishing,
-\item[FINISHED] Job has finished,
-\item[KILLED] Job has been killed,
-\item[FAILED] Job failed,
-\item[DELETED] Job have been deleted,
-\item[OTHER] Any job state which does not fit the above states.
-\end{description}
-
+\label{app:job-status-mapping}
 \begin{center}

### 2009-11-27 doc/tech_doc/client/client_technical.tex
Updating client technical.

@@
 \label{sec:jobdesc}
+Since new WS-ARC utilizes a middleware plugin structure, it also need to
+support the various job description languages used by these middlewares.
+Therefore the JSDL, XRSL and JDL job description languages are supported.
+In order to provide support for the
+three languages listed above, an internal job description data strucuture
+is needed which is the union of the supported languages. The internal job
+description data structure of new WS-ARC is described in a separate document,
+the ARC Job Description Internal Representation~\cite{internaljobdescription}.
+%~ Additional
+%~ documentation on each of the supported job description languages can be
+%~ found in the respective language documentation~\cite{}, but also in the job
+%~ description mapping document which documents how the three languages are
+%~ mapped in and out from the ARC internal job description data structure~\cite{}.
 
@@
 
-\bibliography{grid, libarcclient}
+\bibliography{grid,libarcclient}
 

### 2009-11-29 doc/manuals/janitor/tex_installation/installation.tex
Merged janitor installation.tex of manual and tech_doc

@@
 The Janitor requires the two perl packages listed in
-table~\ref{tab:install_dependencies}. To have
-the WebService interface for the Janitor, the packages listed in the
-table~\ref{tab:install_dependencies_optional} need to be installed before
+table~\ref{tab:install_dependencies}. To have the WebService
+interface for the Janitor, the packages listed in
+table~\ref{tab:install_dependencies_optional}
+need to be installed before
 the build process is. The Perl modules are available on CPAN and ship
 with all major Linux distributions.  
-
 \begin{table}[!h]
    \begin{center}
-        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
-                   Redland RDF library is used for accessing the catalogue of Runtime Environments.}
+        \mycaption{Required perl packages for the Janitor.
+		   Log4perl is used for the internal logging of the
+		   Janitor, while the Redland RDF library is used
+		   for accessing the knowledge base (catalog) of
+		   Runtime Environments.}
         \label{tab:install_dependencies}
@@
    \begin{center}
-        \mycaption{Optional libraries for the Janitor.}{The library libperl-dev provides the required header files to link the 
+        \mycaption{Optional libraries for the Janitor.}{The
+		   library libperl-dev provides the required header files to link the 
                    WebService to the Perl interpreter.}
@@
 
-The Janitor will be installed by default along with ARC-NOX. Support for
-the dynamic REs can be completely disabled at compile time by using
-the \textit{configure} flags \textit{--disable-janitor-service} for
-the complete janitor or \textit{--disable-janitor-webservice} to only
-disable the Janitor's Web Service.
+If you are using regular Debian or Ubuntu packages, then the Janitor
+can be installed as root by "apt-get install nordugrid-arc1-janitor".
+Installing it will not drag other components of ARC with it since the
+Janitor can be used in its own right or in conjunction with another
+grid system, possibly. Packages for Redhat/Fedora and SuSE/OpenSuSE are
+also provided, the redland library however may not yet be available for
+those systems.
+
+The Janitor source code is shipped as a part of the regular ARC-NOX
+source tree.  If you are compiling the sources yourself, the default
+is to have the A-REX grid manager technically prepared to interact
+with the Janitor.
+The interaction can be prohibited
+with the \textit{configure} flags \textit{--disable-janitor-service}
+for the complete janitor or \textit{--disable-janitor-webservice} for
+only the Web Service support.
 
@@
 Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}}
-in order to be to easily maintain the knowledge database of installable packages.
-At the time of writing, no Linux distribution is offering packages for this fine tool.
+in order to be to easily maintain the knowledge database of installable
+packages.  At the time of writing, no Linux distribution is offering
+packages for this fine tool. The basic editing can also be performed
+fairly easily without that tool, but one is unlikely to use the technology
+to its full potential, then.
 
@@
 by the section \lbrack janitor\rbrack. Table~\ref{tab:arc_conf_tags}
-describes the available tags for Janitor's configuration.
+describes the available tags for the Janitor's configuration.
 
@@
 The parameter \texttt{enabled} specifies whether the Janitor shall be used
-within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
+within A-REX or not. This needs to be distinguished from the previously mentioned
+compilation flags of a similar name. Use the value \texttt{"0"} to disable Janitor. The
 \texttt{uid} and the \texttt{gid} are defining which effective user id
-(uid) and group id (gid) shall be used for the Janitor.\footnote{For
-the time speaking, you need to have started the ARC services as root to
-allow the Janitor to change into those IDs.} The \texttt{registrationdir}
-describes the directory in which the subdirectories \texttt{jobs} and
-\texttt{rtes} will be created.  In these directories the states of the
-jobs and the runtime environments will be stored.  The knowledge base
-of installable packages is specified by the parameter \texttt{catalog}.
-Its value can be any kind of URL pointing to a file written in the
-Resource Description Framework (RDF) format.  The specification of the
-RDF file will be explained in detail in section~\ref{sec:catalog}.
-The parameter \texttt{downloaddir} assigns the directory in which
-the installation files will be saved after they have been downloaded
-or copied from the repository which was specified by the catalog. The
-\texttt{installationdir} finally specifies the directory into which all
-packages will be installed and should be available for all computing
-elements, i.e.\ by using a shared volume.  If the configuration file
-furthermore contains the \texttt{runtimedir} tag within the section
-\texttt{grid-manager}, the Janitor will also create a symbolic link
-in the \texttt{runtimedir} pointing to the configuration script of the
-installation performed by the Janitor.  The tags \texttt{jobexpirytime}
-and \texttt{rteexpirytime} are used for an automated cleanup and is
-defined in seconds.  The default value for the \texttt{jobexpirytime} is
-seven days and for the \texttt{rteexpirytime} three days.  The additional
-tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and
+(uid) and group id (gid) shall be used for the Janitor.
+
+The \texttt{registrationdir} describes the directory in which the
+subdirectories \texttt{jobs} and \texttt{rtes} will be created.
+In these directories the states of the jobs and the runtime environments
+are stored. Please recall that the Janitor does not use a database as
+a backend, but all communication between invocations are performed via
+files in those folders.
+
+The knowledge base of installable packages is specified by the parameter
+\texttt{catalog}.  Its value can be any kind of URL pointing to a file
+written in the Resource Description Framework (RDF) format.  One should
+not light-heartedly use a remote address for this purpose. Such a remote
+source needs to be trusted, since any runtime environment specified in
+a catalog (if the package description matches constraints by the local
+site administrator) may possibly be installed by regular grid users.
+
+The specification of the RDF file will be explained in detail in
+section~\ref{sec:catalog}.  The parameter \texttt{downloaddir} assigns
+the directory to which the installation files will be saved after they
+have been downloaded or copied from the repository which was specified by
+the catalog. Please remember: the URL in arc.conf indicates the location
+of the catalog. And the URLs somehow specified in the catalog specify
+the location from where to download the runtime environment.
+
+The \texttt{installationdir} finally specifies the directory into which
+all packages will be installed. This directory needs to be available
+for all computing nodes for the execution of arbitrary programs, most
+commonly by using it as a shared NFS volume.
+
+If the configuration file furthermore contains the \texttt{runtimedir}
+tag within the section \texttt{grid-manager}, the Janitor will also
+create a symbolic link in the \texttt{runtimedir} pointing to the
+configuration script of the installation performed by the Janitor.
+The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used
+for an automated cleanup and is defined in seconds.  The default
+value for the \texttt{jobexpirytime} is seven days and for
+the \texttt{rteexpirytime} three days.  The additional tags
+\texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and
 \texttt{deny\_rte} are used to include or exclude certain base packages
-or runtime environments of the catalog. This feature is useful, if
-the catalog is maintained by a higher organization.  The path to the
-log4perl configuration file is defined by the tag \texttt{logconf}.
-An example how to configure ARC and log4perl is provided in the
-Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.
+or runtime environments of the catalog. This feature is useful, if the
+catalog is maintained by a higher organization. But again: you need to
+trust it.
+
+The path to the log4perl configuration file is defined by the tag
+\texttt{logconf}.  Examples on how to configure ARC and log4perl are
+provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.
 
@@
 
-The Janitor is designed to be used on Linux distributions.
-Its porting to other platforms has not yet been addressed.
+The Janitor was designed to be used for UNIX-compatible operating systems
+and tested for various Linux distributions. It should also be functional
+on MaxOS X and Windows with Cygwin / CoLinux.  The porting of the Janitor
+to other platforms has not yet been addressed.
 
-The ARC middleware is not ultimately essential for dynamic Runtime Environments.
-All the Perl code would be functional with any Grid middleware.
-
-% Is designed for linux distributions. Windows or Mac won't work (Reason wget! maybe also the Executer).
+The ARC middleware is not ultimately essential for dynamic Runtime
+Environments.  All the Perl code would be functional with any Grid
+middleware.

### 2009-11-29 doc/tech_doc/janitor/tex_installation/installation.tex
only invoke janitor if it's enabled in the config
Added symbolic link for compatibility until the tech-doc/janitor directory is removed.

--- 2009-06-08 doc/manuals/janitor/tex_installation/installation.tex
@@
-\chapter{Installation}
-
-% ** Dependencies
-% *** Ordinary: log4perl and ``Redland RDF Library - Perl Interface''
-% *** WebService: libperl-dev
-% *** Protege: For maintaining the knowledge base
-
-The Janitor requires the two perl packages listed in
-table~\ref{tab:install_dependencies}. To have
-the WebService interface for the Janitor, the packages listed in the
-table~\ref{tab:install_dependencies_optional} need to be installed before
-the build process is. The Perl modules are available on CPAN and ship
-with all major Linux distributions.
-
-\begin{table}[!h]
-   \begin{center}
-        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
-                   Redland RDF library is used for accessing the knowledge base of Runtime Environments.}
-        \label{tab:install_dependencies}
-	\begin{tabular}{|p{3cm}|p{7cm}|}
-	\hline
-	   liblog-log4perl-perl & \textit{Log4perl is a port of the log4j logging package}\\
-	\hline
-	   librdf-perl          & \textit{Perl language bindings for the Redland RDF library}\\
-	\hline
-	\end{tabular} 
-   \end{center}
-\end{table}
-\begin{table}[!h]
-   \begin{center}
-        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
-                   WebService to the Perl interpreter.}
-        \label{tab:install_dependencies_optional}
-	\begin{tabular}{|p{3cm}|p{7cm}|}
-	\hline
-	   libperl-dev & \textit{ Perl library: development files}\\
-	\hline
-	\end{tabular}
-   \end{center}
-\end{table}
-\forcelinebreak
-
-The Janitor will be installed by default along with ARC1. Support for
-the dynamic REs can be completely disabled at compile time by using
-the \textit{configure} flags \textit{--disable-janitor-service} for
-the complete janitor or \textit{--disable-janitor-webservice} to only
-disable the Janitor's Web Service.
-
-Furthermore it is recommended to install the ontology editor 
-Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
-knowledge database of installable packages. At the time of writing, no Linux distribution is offering packages for this fine tool.
-
-\section{Configuration}\label{sec:janitor_configuration}
-
-The current version of the Janitor can be configured using the common file \textit{arc.conf}. It is expected in the 
-configuration directory \textit{etc}. The Janitor is using the environment variable \textit{NORDUGRID\_CONFIG}
-to determine the location of the corresponding file. If that variable is not set, the default location \textit{/etc/arc.conf} will be used. 
-The configuration is assigned by the section \lbrack janitor\rbrack. Table~\ref{tab:arc_conf_tags} describes the available tags
-for Janitor's configuration.
-
-\begin{landscape}
-\begin{table}[!h]
-   \begin{center}
-        \mycaption{Tags usable in \textit{arc.conf} within the section janitor.}{Tags usable in \textit{arc.conf} within the section janitor.}
-	\label{tab:arc_conf_tags}
-	\begin{longtable}{|p{3cm}|p{10cm}||p{10cm}|}
-	\hline
-	   \textbf{tag}    & \textbf{example}                      & \textbf{description}\\
-        \hline
-           enabled         & "1"                                   & Boolean flag which enables or disables janitor in A-REX.\\
-	   uid             & "root"                                & The effective uid. \\
-	   gid             & "0"                                   & The effective gid. \\
-	   registrationdir & "/var/spool/nordugrid/janitor"        & Directory where we the current states of jobs are kept. \\
-	   catalog         & "/var/spool/nordugrid/janitor/catalog/knowarc.rdf"& URL of the catalog containg the package information.\\
-	   downloaddir     & "/var/spool/nordugrid/janitor/download" & Directory for downloads \\
-	   installationdir & "/var/spool/nordugrid/janitor/runtime"& Directory for installation of packages                   \\
-	   jobexpirytime   & "7200"                                & If a job is older than this, it is considered dead and assigned to be removal pending.\\
-	   rteexpirytime   & "36"                                  & If a runtime environment was not used for this time, it will be assigned to be removal pending.\\
-	   allow\_base     & "*"                                   & Allow rule for base packages. \\
-	   deny\_base      & "debian::etch"                        & Deny rule for base packages.\\
-	   allow\_rte      & "*"                                   & Allow rule for base packages. \\
-	   deny\_rte       & "APPS/MATH/ELMER-5.0.2"               & Deny rule for base packages. \\
-	   logconf         & "/opt/nordugrid/etc/log.conf"         & Location of the logging configuration file for janitor.\\
-	\hline
-	\end{longtable}
-   \end{center}
-\end{table}
-\end{landscape}
-
-The parameter \texttt{enable} specifies whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
-\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
-The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
-In these directories the states of the jobs and the runtime environments will be stored. 
-The knowledge base of installable packages is specified by the parameter \texttt{catalog}. 
-Its value can be any kind of URL pointing to a file written in the Resource Description Framework (RDF) format.
-The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
-The parameter \texttt{downloaddir} assigns the directory in which the installation files will be saved after they have been 
-downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
-directory into which all packages will be installed and should be 
-available for all computing elements, i.e.\ by using a shared volume.
-If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, the Janitor will also
-create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by the Janitor.
-The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
-The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
-The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
-or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintained by
-a higher organization.
-The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
-An example how to configure ARC and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.
-
-\lstsetCONFIGURE
-\begin{lstlisting}[
-        label=lst:arc_conf,
-        caption={ [Example \textit{arc.conf} settings for janitor.]
-                  \textbf{Example \textit{arc.conf} settings for janitor.}}
-        ]
-[janitor]
-enabled="1"
-logconf="/opt/nordugrid/etc/log.conf"
-registrationdir="/var/spool/nordugrid/janitor"
-installationdir="/var/spool/nordugrid/janitor/runtime"
-downloaddir="/var/spool/nordugrid/janitor/download"
-jobexpirytime="7200"
-rteexpirytime="36"
-uid="root"
-gid="0"
-allow_base="*"
-allow_rte="*"
-
-[janitor/nordugrid]
-catalog="/var/spool/nordugrid/janitor/catalog/knowarc.rdf"
-\end{lstlisting}
-
-\lstsetCONFIGURE
-\begin{lstlisting}[
-        label=lst:log_conf,
-        caption={ [Example \textit{log.conf} settings for janitor.]
-                  \textbf{Example \textit{log.conf} settings for janitor.}}
-        ]
-# Master Loglevel
-# [OFF | DEBUG | INFO | WARN | ERROR | FATAL]
-#log4perl.threshold = OFF
-
-log4perl.rootLogger = WARN, DebugLog, MainLog, ErrorLog
-log4perl.appender.DebugLog = Log::Log4perl::Appender::Screen
-log4perl.appender.DebugLog.layout = PatternLayout
-log4perl.appender.DebugLog.layout.ConversionPattern = [%C] %d %p> %m%n
-
-log4perl.appender.MainLog = Log::Log4perl::Appender::File
-log4perl.appender.MainLog.Threshold = VERBOSE
-log4perl.appender.MainLog.filename = /var/log/janitor.log
-log4perl.appender.MainLog.layout = PatternLayout
-log4perl.appender.MainLog.layout.ConversionPattern = %d %p> %m%n
-
-log4perl.appender.ErrorLog = Log::Log4perl::Appender::File
-log4perl.appender.ErrorLog.Threshold = ERROR
-log4perl.appender.ErrorLog.filename = /var/log/janitor_error.log
-
-log4perl.appender.ErrorLog.layout = PatternLayout
-log4perl.appender.ErrorLog.layout.ConversionPattern = %d %p> %m%n
-\end{lstlisting}
-
-% ** Configuring arc.conf
-% *** Where to store the data of janitor
-% ** Configuring log.conf
-% *** Where to store the log of janitor
-
-\section{Limitations}
-
-The Janitor is designed for the usage on Linux distributions.
-Its porting to other platforms has not yet been addressed.
-
-The ARC middleware is not ultimately essential for dynamic Runtime Environments.
-All the Perl code would be functional with any Grid middleware.
-
-% Is designed for linux distributions. Windows or Mac won't work (Reason wget! maybe also the Executer).
+../../../manuals/janitor/tex_installation/installation.tex

### 2009-11-29 doc/manuals/janitor/tex_introduction/introduction.tex
Merged manual/tech_doc janitor/introduction

@@
 
-The Janitor is a service for an automated installation of runtime
-environments (RTEs) for grid computing elements.  It is addressed
-transparently via the A-REX service for job submission but can also be
-used as a standalone tool.
+The {\em Janitor} is a service for the automated installation of runtime
+environments for grid computing elements.  Its command line interface
+allows for a direct interaction with site administrators. However, the
+main stimulus for its development was the idea integrate such a service
+with the regular handling of compute jobs. For ARC this is performed by
+the \AREX module.
+
+From the programmer's view, the Janitor is mostly a Perl script and the routines within \AREX to invoke it. The
+site administrator will also associate with it also the Catalog files, that describe the availability of runtime
+environments, and the repository of installable runtime envionments themselves, which are regular tar archives obeying
+to particular structure and reside in a separate folder. In order to minimise the latency for the invocation of
+the Perl script and the associated parsing of files, a Janitor web service was developed, which still is a Perl
+script.
 
@@
 to adopt the technology to start sharing their resources. From the
-current grid user's viewpoint, the admission of users of a very different
+current grid user's viewpoint, the admission of users with a very different
 education will suddenly impose difficulties in the communication between
@@
 research aims. Hence, the proper installation of non-standard software
-(runtime environments, RTEs) is not guaranteed. And the available time for
-manual labour plus self-education is scarse.
+(Runtime Environments, RTEs) is not guaranteed. And extra time for manual
+labour plus self-education is scarse.
 
 A core problem remains to distribute a locally working solution, the
-Know-How, quickly across all contributing sites, i.e., without manual
-interference. Every scientific discipline has its respective own set
-of technologies for the distribution of work load. E.g. research in
-bioinformatics requires access to so many different tools and databases,
-that few sites, if any, install them all. Instead, the use of web
-services became a commodity, with all their intrinsic problems as there
-are bottlenecks and restrictions of repeated access.  The EU project
-KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}}
-amongst other challenges with the here presented work
+Know-How, quickly across all contributing sites, \ie without manual
+interference. Every scientific discipline has its respective own
+set of technologies for the distribution of work load. For instance,
+research in bioinformatics requires access to so many different tools
+and databases, that few sites, if any, install them all. Instead, the use
+of web services became a commodity, with all their intrinsic problems as
+there are bottlenecks and restrictions of repeated access.  The EU project
+KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}},
+amongst other challenges, with the here presented work
 extends the NorduGrid's Advanced Research Connector (ARC) grid
-middleware~\cite{ELLERT_2007} towards an infrastructure for the automated
-installation of software packages.
+middleware~\cite{ELLERT_2007} towards an infrastructure for the
+automated installation of software packages.
 
@@
 shall be spawned as jobs in a computational grid using dRTEs rather
-than accessing a web service at one particular machine that is shared
-amongst all users.  The grid introduces an extra level of parallelism
-that web services cannot provide. The demands for short response times
-and the heterogeneous education of site-administrators on a grid demand
-an automatism for the installation of software and databases without
-manual interference~\cite{BAYER_2007}.
+than accessing a public web service at one particular machine that
+is shared amongst all users.  The grid introduces an extra level of
+parallelism that web services cannot provide. The demands for short
+response times and the heterogeneous education of site-administrators
+on a grid demand an automatism for the installation of software and
+databases without manual interference~\cite{BAYER_2007}.
 
@@
 
-This document first dedicates a chapter on how to set-up the Janitor
+This document starts with a chapter on how to set-up the Janitor
 locally. It is followed by a chapter that gives further instructions on

### 2009-11-29 doc/tech_doc/janitor/tex_introduction/introduction.tex
Added temporarily symbolic link from tech_doc to manual for janitor/introduction until transition is completed.

--- 2009-09-02 doc/tech_doc/janitor/tex_introduction/introduction.tex
@@
-\chapter{Introduction}
-
-Janitor is a service for automated runtime environment installation for grid computing elements.
-It is integrated in the A-REX service used for job submission but also can be used as a standalone tool.
-
-\section{Motivation}
-
-A major motivation for grid projects is to stimulate new communities to adopt computational grids for their
-causes. From the current grid user's viewpoint, the admission of users of a very different education will
-suddenly impose difficulties in the communication between site maintainers. One will not even understand
-the respective other side's research aims. Hence, the proper installation of non-standard software (Runtime
-Environments) is not guaranteed and the priorities of manual labour will be mostly disjunctive.
-
-A core problem remains to distribute a locally working solution, the Know-How, quickly across all contributing sites, i.
-e., without manual interference.
-Every scientific discipline has its respective own set of
-technologies for the distribution of work load. E. g. research in bioinformatics requires access to so many
-different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
-a commodity, with all the problems with respect to bottlenecks and restrictions of repeated access.
-The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges with the here presented work extends the NorduGrid?s
-Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
-automated installation of software packages.
-
-An automation of the software installation, referred to as dynamic Runtime Environments, seems the only
-approach to use the computational grid to its full potential. Components of workflows shall be spawned
-as jobs in a computational grid using dynamic Runtime Environments rather than as shared web services.
-The grid introduces an extra level of parallelism that web services cannot provide. The required short
-response times and the heterogeneous education of site-administrators on a grid demand an automatism for
-the installation of software and databases without manual interference~\cite{BAYER_2007}.
-
-\section{Overview}
-
-This document will start with a chapter of how to set up janitor locally. The following chapter will give an instruction on how to 
-use Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
-which is basically covering the method how to prepare new runtime environments. A deeper knowledge about the design of Janitor will 
-be given by the subsequent forth chapter. In the fifth chapter, an outlook to future developments will given.
-
-% BASIC CONCEPT
-
-% * Installation
-
-% * Usage
-
-% * Maintainance
-
-% * Programming concept
-
-% * Future work
-
-% * APPENDIX
+../../../manuals/janitor/tex_introduction/introduction.tex

### 2009-11-29 doc/manuals/janitor/tex_titlepage/titlepage.tex
Update of Michael's email address.
Merged tex_titlepage/titlepage.tex in manuals/janitor/ and tech_doc/janitor/

@@
 {\centering \textsc{\Large Dynamic Runtime Environments with Janitor}
-\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to moeller@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
+\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This
+	Janitor and this document with it is under continuous development.
+	Your comments and suggestions are appreciated.} % DELETE ONLY THAT LINE!!!
 \Large \par}

### 2009-11-29 doc/tech_doc/janitor/tex_titlepage/titlepage.tex
Having symbolic link helping the merge process.

--- 2009-03-05 doc/ws-programming-tutorial/doc/tex_titlepage/titlepage.tex
@@
-\def\today{\number\day/\number\month/\number\year}
-
-\begin{titlepage}
-
-\begin{tabular}{rl}
-\resizebox*{3cm}{!}{\includegraphics{images/ng-logo.png}}
-&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
-\end{tabular}
-
-\hrulefill
-
-{\raggedleft NORDUGRID-TECH-19\par}
-
-{\raggedleft \today\par}
-
-\vspace*{2cm}
-
-%%%%---- The title ----
-{\centering \textsc{\Large Web Service programming Tutorial}
-\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
-\Large \par}
-\vspace*{0.5cm}
-
-%%%%---- A subtitle, if necessary ----
-%{\centering \textit{\large Paper subtitle}\large \par}
-
-\vspace*{1.5cm}
-%%%%---- A list of authors ----
-    {\centering \large Michael Glodek, Steffen M\"oller \par}
-
-%%%%---- An abstract - if style is article ----
-%\begin{abstract}
-%The abstract
-%\end{abstract}
-\end{titlepage}
-
-\tableofcontents                          %Comment if use article style
-\newpage
-%\chapter{Preface}
-%\section{Introduction}                    %Use Sections for articles
-%\label{sec:intro}
-
-\sloppy
+../../../manuals/janitor/tex_titlepage/titlepage.tex

### 2009-11-29 doc/manuals/janitor/tex_maintenance/maintenance.tex
Merged maintenance.tex.

@@
 
-This chapter explains how to maintain the {\em Catalog} and the {\em Janitor} itself. 
-It gives detailed instructions how to create new packages for the Janitor.
-To administrate the catalog, it is strongly recommended (but not
-ultimately required) to use the ontology editor Prot\`eg\`e, as explained
-in the next section.
-In the last section, a typical use case in maintaining the Janitor will be shown.
+This chapter explains how to maintain the {\em Catalog} and the {\em Janitor}
+themselves.
+It gives detailed instructions how to create new packages for the
+Janitor.
+To administrate the catalog, it is strongly recommended
+(but not ultimately required) to use the ontology editor Prot\`eg\`e,
+as explained in the next section.
+In the last section, a typical use case in maintaining the Janitor will
+be presented.
 
@@
 through a web server or is distributed together with the Janitor as a
-regular file.  It is specifed by a (Resource Description Framework)
-RDF file assigned to Janitor using the tag \texttt{catalog} within
+regular file.
+It is specifed by a (Resource Description Framework) RDF file
+assigned to the Janitor using the tag \texttt{catalog} within
 the configuration file (see~\ref{sec:janitor_configuration}).
@@
 On the left side of the editor the class browser is placed. Three
-main classes are prepared: \texttt{MetaPackage}, \texttt{Note}
-and \texttt{Package}.  The \texttt{Metapackage} is a general
-platform-independent description of a \texttt{Package}.  It has one
-or more instances of the class \texttt{Package} and is described
-by the subclasses of \texttt{Note}. The class \texttt{Note} has
-two subclasses: \texttt{BaseSystem} and \texttt{Tag} to describe
-the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the
-Debian release a \texttt{Package} refers to (i.e. here etch or
-sid), i.e. the name of a common installation or a virtual image.
-The class \texttt{Tag} provides small keywords which can be assigned
-to \texttt{MetaPackages} such that they can be found more easily.
-\texttt{TarPackage} and \texttt{DebianPackage} or currently the only
-subclasses of \texttt{Package}.  They are representing the necessary
-information (i.e. URL or Packagename) for the installation. In order
-to have an overview how the classes are interacting with each other
-the tables~\ref{tab:main_class_meta}, \ref{tab:main_class_base},
-\ref{tab:main_class_tag}, \ref{tab:main_class_debian} and
-\ref{tab:main_class_tar} are pictured.
+main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and
+\texttt{Package}.
+
+The \texttt{Metapackage} is a general platform-independent description
+of a \texttt{Package}. It can be understood as a reference to a functionality
+that should be implemented at the remote site. But the exact instruction
+on how to install it is not given.  The "instruction-level" comes with
+instances of its subclass \texttt{Package}.
+
+MetaPackages contain one or more instances of the class
+\texttt{Package}, which are then tangible software packages providing
+the functionality that the MetaPackage references. The packages may
+be aiming for different versions of the operating system or be different
+in the way these are installed, but any package being assigned to the same
+MetaPackage needs to perform the same functionality.
+MetaPackages are described by the subclasses of \texttt{Note, which
+in turn has two subclasses: \texttt{BaseSystem} and
+\texttt{Tag}.
+
+The \texttt{MetaPackage} is described by the subclasses of \texttt{Note}.
+The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
+and \texttt{Tag} to describe the \texttt{MetaPackage}.
+
+The \texttt{BaseSystem} describes the Debian release a \texttt{Package}
+refers to (i.e. here etch or sid), i.e. the name of a common installation
+or a virtual image.
+The class \texttt{Tag} provides small keywords which can be assigned to
+\texttt{MetaPackages} such that they can be found more easily.
+\texttt{TarPackage} and \texttt{DebianPackage}
+or currently the only subclasses of \texttt{Package}.
+
+They are representing the necessary information (i.e. URL or Packagename)
+for the installation.
+To provide an overview on how the classes are
+interacting with each other the Tables~\ref{tab:main_class_meta},
+\ref{tab:main_class_base}, \ref{tab:main_class_tag},
+\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.
 
@@
 
+\subsection{Debian packages - dysfunctional in current implementation}
+
+The problem with Debian packages is that these are available only for the
+local machine and not immediately also for the whole network. This feature
+was meant for setups that use virtual machines for the execution of jobs.
+
+Those entries can also be used to help the specification of further
+dependencies of runtime environments. It would then be left to the
+responsibility of the system administrator to manually (or assisted with
+scripts) distribute a series of extra Debian packages throughout the
+compute nodes and use the catalog entry merely to indicate their presence.
+
+One can as such interpret the catalog to represent an interface between
+software distributed via Linux distributions and independently from these
+via grid communities. It should be noted that the Linux distributions
+have now all started to accept communities to maintain packages, which
+may bringt many scientific packages away from being traditional ARC 
+runtime environments towards becoming regular packages of some Linux
+distribution. For Debian, the Debian-Science and
+Debian-Med\footnote{\href{http://debian-med.alioth.debian.org}{http://debian-med.alioth.debian.org}}
+communities are known to be very open to grid and cloud computing.
+
 \section{HTML interface of the catalog}
@@
 site describing Runtime Environments in the Runtime Environment
-Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with
-an eventual transition to the new system.  That page collects descriptions
-for Runtime Environments to encourage human site administrators to
-install these.  This HTML page listing the manually or automatically
-installable REs is prepared by the script web/list.pl.  This script is
-meant to be run by a mod\_perl-enabled Apache. The script itself does
-not contribute to the core functionality of the Janitor.  In the first
-lines of the script some variables specific to the site are set. To
-configure the script these have to be changed~\cite[p. 9]{BAYER_2007}.
+Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues
+with an eventual transition to the new system.
 
-\section{Introducing new packages}\label{sec:catalog}
+That page, in a look resembling the classical description of RTE,
+collects descriptions for Runtime Environments to encourage human site
+administrators to install these.  This HTML page listing the manually or
+automatically installable RTEs is prepared by the script web/list.pl.
+This script is meant to be run by a mod-perl enabled Apache. The
+script itself does not contribute to the core functionality of the
+Janitor. It only performs the human-readable presentation of a catalog's
+RDF file to users. In the first lines of the script some variables
+specific to the site are set. To configure the script these have to be
+changed~\cite[p. 9]{BAYER_2007}.
 
-This section describes how to add new packages to the Catalog. In the
-current implementation, only tar based packages are processed by the
-Janitor.  Within the example they are assigned to be used together with
-Debian Etch. This limitation is only literal.  There is no restriction
-for newer Debian distributions.
+\section{Introducing new packages}\label{sec:catalog}\label{sec:introducing_new_packages}
+
+This section describes how to add new packages to the Catalog. In
+the current implementation, only tar based packages are processed by
+the Janitor.  Within the here presented example they are assigned to
+be used together with Debian Etch. This limitation is only literal.
+There is no restriction for newer Debian distributions.
 
@@
 accepted for dynamic Runtime Environment installation, a well accepted
-file format throughout the UNIX community. The concept reflects the
+file format throughout the UNIX community.
+The concept reflects the
 traditional manual approach towards RTE in ARC, for which one directory
-is made available to all compute nodes. This section explains the
-inner structure of the tar files.  Subdirectories are visualised in
-Figure~\ref{fig:tar_folder}.
+is made available to all compute nodes.
+This section explains the inner structure of the tar files for the
+representation of dynamic runtime environments.  Subdirectories are
+visualised in Figure~\ref{fig:tar_folder}.
 
@@
 The tar file contains two directories, named {\tt control} and {\tt
-data}. Software is stored in the latter subdirectory, while the files
-formally specifying how to deal with such packages are stored in the
-prior. Upon installation, the content of the data directory is extracted
-to some directory \textdollar BAR.  After this unpacking of the tar file,
-the Janitor executes the install script provided in the control directory.
-It is executed within the working directory \textdollar BAR. The job of
-this skript is to perform any necessary post-processing.  The Janitor
-stores the file \texttt{control/remove}. It will be executed in the
-same way as \texttt{control/install} just before the tar-package is
-removed. In most cases \texttt{control/remove} will be empty.  Finally,
-the file \texttt{control/runtime} is sourced multiple times by the
-Grid Manager's job-submit script. After installing the package, the
-Janitor changes all occurences of \%BASEDIR\% in the runtime script to
-\textdollar BAR.  Once the tar file was prepared, it must its entry to
-a RTE Catalogue~\cite[p. 10]{BAYER_2007}.
+data}:
+
+\begin{description}
+\item [data/] contains all software that the grid-job may need
+\item [control/] contains files formally specifying how to deal
+	with the information in the data/ directory.
+\end{description}
+
+Upon installation of such tar-based runtime environments,
+the content of the data directory is extracted to some directory
+\textdollar BAR.  After this unpacking of the tar file, the Janitor
+executes the install script provided in the control directory.  It is
+executed within the working directory \textdollar BAR. The job of
+this skript is to perform any necessary post-processing.
+The Janitor stores the file \texttt{control/remove}. It will be executed
+in the same way as \texttt{control/install}, just before the tar-package
+is removed. In most cases \texttt{control/remove} will be empty, implying
+that the working directory \textdollar BAR shall be removed and no other
+action is required.
+Finally, the file \texttt{control/runtime} is sourced multiple times
+by the Grid Manager's job-submit script. After installing the package,
+the Janitor changes all occurences of \%BASEDIR\% in the runtime script
+to \textdollar BAR.  Once the tar file was prepared, it must its entry
+to a RTE Catalogue~\cite[p. 10]{BAYER_2007}.
+But the working directory shall not be moved. All post-processing needs
+to be performed \textit{in situ}.
 
@@
 information system.
+
+The remainder actions are regular actions performed upon execution of every grid-job. Upon submission,
+the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. Every
+ARC runtime environment must specify such a runtime script, new is only its specific location as control/runtime.
+Since the directory \textdollar BAR is not known for the individual preparing the runtime environment, that
+file will instead use the placeholder \%BASEDIR\%.  After installing the package, the Janitor changes all
+occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.  To be offered to computing elements
+for an installation, the such prepared runtime environment must be announced to a Catalog to which the
+Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.
 
@@
 prototypes are provided at \url{http://dre.knowarc.eu}.
+
+\subsubsection{Example: WEKA machine learning Java library}
 
@@
 
-The remove script, which will be executed right before WEKA is deinstalled, is empty.
-Janitor will delete the directory, so there is nothing more to be done.
+The remove script, which will be executed right before WEKA is
+deinstalled, is empty. The Janitor will delete the whole directory,
+so there remains nothing to be removed in addition.
+The remove script may \eg be used to remove indices or other files
+in /tmp.
 
+%\subsection{Example: R packages in CRAN and BioConductor.org}
 %\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
@@
 
+\subsection{Example: ATLAS for High Energy Physics}
+
 To address the concerns of the physicists using ARC, a dynamic runtime
@@
 \href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}.
+
 The preparation comprised the following steps:
@@
 package in question takes up more than 5 GB. This was a test illustrating
-the feasibility of using dynamic REs in High Energy Physics. The
-application of the dynamic RE for ATLAS needs to wait for the planned web
+the feasibility of using dynamic RTEs in High Energy Physics. The
+application of the dRTEs for ATLAS needs to wait for the planned web
 service extension of the Catalog. With such a service, e.g. a software
@@
 %
-%\subsection{Adminstrating the Catalog}
+%\section{Adminstrating the Catalog}
 

### 2009-11-29 doc/tech_doc/janitor/tex_maintenance/maintenance.tex
Added symbolic link to help transition.

--- 2009-09-02 doc/tech_doc/janitor/tex_maintenance/maintenance.tex
@@
-\chapter{Maintenance}
-
-This chapter explains how to maintain the catalog and Janitor itself. 
-In order to administrate the catalog, it is absolutly recommened (but not required) to use the ontology editor Prot\`eg\`e. The first
-section will explain how this is done.
-The second section a detailed explanation how to create new packages for Janitor will be given.
-In the last section typical use case in maintaining Janitor will be listed.
-
-% * Maintainance
-% ** Protege and RDF files
-% ** Creating own tar balls
-
-\section{Catalog}\label{sec:catalog}
-
-The Catalog describes runtime environments and is either served through a web server or distributed together with the Janitor. 
-It is specifed by a (Resource Description Framework) RDF file assigned to 
-Janitor using the tag \texttt{catlog} within the configuration file (see~\ref{sec:janitor_configuration}). 
-The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
-which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
-{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\
-
-In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
-editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
-    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
-    \label{fig:protege_example}
-  \end{center}
-\end{figure}
-On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
-The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
-It has one or more instances of the class \texttt{Package} and is
-described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
-and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
-(i.e. here etch or sid). The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
-they can be easier found.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
-They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
-classes are interacting with each other the Tables~\ref{tab:main_class_meta},\ref{tab:main_class_base}, \ref{tab:main_class_tag},
-\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.
-
-\begin{table}[!h]
-   \begin{center}
-        \mycaption{Specification of the class Metapackage.}{}
-         \label{tab:main_class_meta}
-	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
-	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
-	\hline
-	description    & single                & String       \\
-	homepage       & single                & String       \\
-	instance       & multiple              & Instance of Package       \\
-	lastupdated    & single                & String       \\
-	name           & required single       & String       \\
-	tag            & multiple              & Instance of Tag       \\
-	\end{tabular} 
-   \end{center}
-\end{table}
-
-\begin{table}[!h]
-   \begin{center}
-        \mycaption{Specification of the class Basesystem.}{}
-         \label{tab:main_class_base}
-	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
-	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
-	\hline
-	description    & single                & String       \\
-	distribution   & required single       & String       \\
-	name           & required single       & String       \\
-	short\_description & required single   & String       \\
-	url            & required single       & String       \\
-	\end{tabular} 
-   \end{center}
-\end{table}
-
-\begin{table}[!h]
-   \begin{center}
-        \mycaption{Specification of the class Tag.}{}
-         \label{tab:main_class_tag}
-	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
-	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
-	\hline
-	description    & single                & String       \\
-	name           & required single       & String       \\
-	\end{tabular} 
-   \end{center}
-\end{table}
-
-\begin{table}[!h]
-   \begin{center}
-        \mycaption{Specification of the class DebianPackage.}{}
-         \label{tab:main_class_debian}
-	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
-	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
-	\hline
-	basesystem     & required single       & Instance of BaseSystem       \\
-	debconf        & multiple              & String       \\
-	depends        & multiple              & Instance of MetaPackage  or Package      \\
-	package        & required multiple              & String      \\
-	\end{tabular} 
-   \end{center}
-\end{table}
-
-\begin{table}[!h]
-   \begin{center}
-        \mycaption{Specification of the class TarPackage.}{}
-         \label{tab:main_class_tar}
-	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
-	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
-	\hline
-	basesystem     & required single       & Instance of BaseSystem       \\
-	depends        & multiple              & Instance of MetaPackage or Package      \\
-	environ        & multiple              & String       \\
-	url            & required multiple     & String      \\
-	\end{tabular} 
-   \end{center}
-\end{table}
-
-\section{HTML interface of the catalog}
-
-The dynamic Runtime Environments stored in the catalog are presented on the aforementioned dedicated web
-page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
-RDF syntax and an automated transformation to
-HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
-Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
-That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
-This html page listing the manually or automatically installable RTEs is prepared by the script web/list.pl.
-This script is meant to be run by a mod-perl enabled Apache. The script itself is only loosely integrated
-within the Janitor. In the first lines of the script some variables specific to the site are set. To configure the
-script these have to be changed~\cite[p. 9]{BAYER_2007}.
-
-\section{Introducing new packages}\label{sec:catalog}
-
-This section describes how to add new packages to the catalog. Currently only tar based packages are processable by Janitor.
-Within the example they are assigned to be used together with Debian Etch. This limitation is only literal. 
-There should be no limitiation
-to newer Debian distributions.
-
-\subsection{Debian Etch (tar based)}
-
-At the time of writing, only the tape archive (tar) file format is accepted for dynamic runtime environment installation, 
-a well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files. 
-Subdirectories are visualised in Figure~\ref{fig:tar_folder}.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=4cm]{images/tar_folder.pdf}
-    \mycaption{Directory structure in the tar files for automated installation.}{}
-    \label{fig:tar_folder}
-  \end{center}
-\end{figure}
-
-The tar file contains two directories named control and data. Software is stored in the latter subdirectory,
-while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
-the content of the data directory is extracted to some directory \textdollar BAR. 
-After this unpacking of the tar file, the Janitor executes the install skript provided in the control directory. 
-It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
-The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
-before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
-Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
-installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
-To be offered to computing elements for an installation, the such prepared runtime environment must be
-announced to a Catalog to which the Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.
-
-\subsection{Protoypes}
-
-In order to have an impression how the tar files are created, several prototypes are provided at \task{URL required!!}.:\\
-
-The WEKA package for machine learning~\cite{FRANK_20074} and the Java Runtime Environment
-are available as dynamic Runtime Environments. Further packages for bioinformatics comprise dynamic
-variants of tools for the analysis of transcription factor binding sites. These are already offered for manual
-installation via the prior mentioned traditional page representing Runtime Environments for ARC. 
-The corresponding tar file is named \task{somewhat??}. The \texttt{data} directory simply contains a ZIP file
-which needs to be unzipped in the installation directory. For that reason, the \texttt{control/install} script is 
-written as follows:
-% dollar signs made trouble in listings *hrmlgrmpf* ... Costs too much time to figure out why *ARC!*
-\begin{verbatim} 
-#!/bin/sh
-set -e  # Makes the script to terminate at the first line it fails.
-
-WEKA_ZIP="weka-3-4-8a.zip"
-unzip $WEKA_ZIP
-rm -f $WEKA_ZIP
-\end{verbatim}
-The runtime script sets the environment variable of the Java Classpath:
-\begin{verbatim}
-#!/bin/sh
-
-WEKA_JAR="weka-3-4-8a/weka.jar"
-case "\$1" in
-	0)	# Just before job submission
-		# none
-	;;
-	1)	# Just before job execution
-		# Initialize the java environment
-		CLASSPATH="%BASEDIR%/$WEKA_JAR:$CLASSPATH"
-		export CLASSPATH
-	;;
-	2)	# After job termination
-		# none
-	;;
-	*)
-		return 1
-	;;
-esac	
-\end{verbatim}
-The remove script, which will be executed right before WEKA is deinstalled, is empty. Janitor will delete the directory, so there
-is nothing to be done.\\
-
-\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
-provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
-dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
-thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
-projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\
-
-To address the concerns of the physicists using ARC, a dynamic runtime environment for the
-ATLAS software suite was prepared. It extends prior work on an automated installation that is available at
-\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}. The preparation comprised the following steps:
-\begin{itemize}
-    \item The file system path specifications in the automated installation scripts were modified using the Janitor
-       path variables.
-    \item A tarball was prepared containing a directory structure as illustrated in Figure~\ref{fig:tar_folder}. The data directory
-       was empty, since the automatic installation script downloads the software from a remote server.
-    \item An entry was added to the Catalog file.
-\end{itemize}
-What sets High Energy Physics software apart is it's sheer size. The package in question takes up more than
-5 GB. This was a test illustrating the feasibility of using DREs in High Energy Physics. The application
-of the DRE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
-service, e.g. a software manager of a big experiment will be able to deploy software packages on production
-sites simply by creating a tarball and adding an entry to the Catalog.
-
-\subsection{Debian Etch}
-
-These kind of packages are not yet supported.
-
-\subsection{Debian Sid}
-
-These kind of packages are not yet supported.
-
-\subsection{Adminstrating the Catalog}
-
-\section{Typical use cases}\label{sec:catalog}
-
-% ** setstate REMOVAL_PENDING
-% no warranty interferences in the file syste,
+../../../manuals/janitor/tex_maintenance/maintenance.tex

### 2009-11-29 doc/manuals/janitor/tex_concept/concept.tex
Merged concept.tex

@@
 The main language for the implementation of the functionality of the
- dRTE's functionality is Perl. And it is solely required (exceptions are the
- integration with the Grid Manager and the Web server) for the Janitor.
+dRTEs' functionality is Perl. And it is solely required (exceptions
+are the integration with the Grid Manager and the Web service) for the
+Janitor.
 The language was perceived a side-issue, the complexity of the data
- expected was of major concern.
+expected was of major concern. This led to the choice of RDF with its
+intrinsic query engine and Perl offering access to the Redland library.
 
@@
 
- In
- the pre-web-service implementation the Catalog remains a static web page.
- The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
- The modules can be separated into two
- functional groups. One addresses the retrieval of information from
- the Catalog's RDF file in the left major branch of the figure. The other
- addresses the process of fetching and installing the packages.
+In the pre-web-service implementation the Catalog remains a static
+web page. The Perl code is split into multiple modules as depicted
+in Figure~\ref{fig:janitorDependencies}.  The modules can be separated
+into two functional groups. One addresses the retrieval of information
+from the Catalog's RDF file in the left major branch of the figure. The
+other addresses the process of fetching and installing the packages.
 
@@
  \begin{center} 
-%    \resizebox{24cm}{!}{\input{images/dependencies}}
-    \includegraphics[width=24cm]{images/dependencies.png}
+ %\resizebox{24cm}{!}{\input{images/dependencies}}
+ %\includegraphics[width=24cm]{images/dependencies.png}
+ \includegraphics[width=24cm]{images/dependencies.pdf}
  \mycaption{Modules of the Janitor and their dependencies}{}
@@
 
-\subsection{States of runtime environments}
+\subsection{States of Runtime Environments}
 
@@
 These states are specific for every compute element (CE) and communicated
-between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
+between the Janitor and the Execution Service. Table~\ref{tab:states}
+shows all
 possible states, while Figure~\ref{fig:RE_states}
 displays the transitions between the states that a Runtime Environment may
-be in during its life time at a particular CE. \task{?at a particular CE?}
+be in during its life time at a particular CE.
 
@@
 The manually induced transitions are marked in red, he automated transitions in black.
-A transition between states can be induced automatically (i.\,e.\ by the
+A transition between states can be induced automatically (\ie by the
 advent of a job requesting a particular dynamic RE) or manually by the
-site's supervisor or an individual with respective rights to use the
-Janitor's web service.
+site's administrator or an individual with respective rights to use the
+Janitor's command line.
 
 \subsection{Subset of that functionality as implemented today}
-Upon presentation of a the package name to a Catalog, from which details about
-the package are retrieved, a CE may
+Upon presentation of a the package name to a Catalog, from which
+details about the package are retrieved, a CE may
 classify a package to be \texttt{INSTALLABLE} if all the dependencies
@@
 
-The Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
-in the Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached. 
-After the Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
-an unforeseen exception occures during that process, the Janitor will drop the job from its database and set the affected runtime
-environments to the state \texttt{FAILED}.
+The Janitor has two states for jobs: \texttt{PREPARED} and
+\texttt{INITIALIZED}. After a job has been succesfully registered with
+the Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs
+are not cached.  After the Janitor is requested to deploy the runtime
+environment, the state of the job will change to \texttt{INITIALIZED}. If
+an unforeseen exception occures during that process, The Janitor will drop
+the job from its database and set the affected runtime environments to
+the state \texttt{FAILED}.
 
@@
 Security is a major concern for grid systems. Any additional feature
-and especially an automatic software
+and especially an automated software
 installation inheritently introduces security threats. This section
@@
 also authorised to install a REs. Restrictions are
-only imposed on the set of dynamic REs that are available for installation.
+only imposed on the set of dynamic RTEs that are available for installation.
 Restrictions are imposed by the site admininistrators on the descriptions
@@
 \begin{itemize}
-\item Dynamic REs are not distinguished between {\em installed} and
+\item Dynamic RTEs are not distinguished between {\em installed} and
   {\em installable} in the information system.
@@
 based solely on tar files, this is barely possible, unless such is
-performed by the install scripts that accompany the tar files. However,
-hereto the installation would have to be performed by a user with system
-priviledges, for which there is no technical requirement.
+performed by the install scripts that accompany the tar files.
+\task{Clarify root vs non-root execution of Janitor.}
+%However,
+%hereto the installation would have to be performed by a user with system
+%priviledges, for which there is no technical requirement.
 

### 2009-11-29 doc/manuals/janitor/tex_usage/usage.tex
Merged usage.tex for the Janitor manual.

@@
 The Janitor can be used either with or without the A-REX service.
-In case A-REX is used, the invocation of the Janitor will be performed in an automated manner.
-It is then triggered by incoming jobs that request a particular RTE for their execution.
-If it is not already installed, but available via the Janitor, then it will be installed
-without further manual intervention.
+In case
+A-REX is used, the invocation of the Janitor will be performed in an
+automated manner.  It is then triggered by incoming jobs that request
+a particular RTE for their execution.  If it is not already installed,
+but
+\begin{enumerate}
+\item found as a MetaPackage in a Catalog that the site supports
+\item with a package that first the BaseSystem of the site
+\end{enumerate}
+then it will be installed without further manual intervention by the
+Janitor - triggered by the A-REX that received the compute request.
+
+The Janitor's installation will not be affected by this decision
+pro or cons and integration with \AREX. Both can be installed in
+parallel. If \AREX is not allowed to install runtime environments upon
+demand, such automated installations can still be invoked manually via
+the Janitor's command line  interface.
 
@@
 
-Runtime Environments can be specified using the supported job description languages.
-The most representative two common languages shall be explained at this point: xRSL and JSDL.
-Listing~\ref{lst:xrsl_job} shows the xRSL example in which two runtime environments are requested.
+Runtime Environments can be specified using the supported job description
+languages.  The most representative two common languages shall be
+explained at this point: xRSL and JSDL.  Listing~\ref{lst:xrsl_job}
+shows the xRSL example in which two runtime environments are requested.
 
@@
 
-A comprehensive reference manual of the Extended Resource Specification Language (XRSL) can be
-found at \href{www.nordugrid.org/documents/xrsl.pdf}{www.nordugrid.org/documents/xrsl.pdf}~\cite{NORDUGRID_MANUAL_4}.
-Within Listing~\ref{lst:jsdl_job} an example using JSDL is provided. The specification of assigning runtime environments in JSDL is 
-currently only defined within the nordugrid jsdl-arc schema~
+A comprehensive reference manual of the Extended
+Resource Specification Language (XRSL) can be found at
+\href{www.nordugrid.org/documents/xrsl.pdf}{www.nordugrid.org/documents/xrsl.pdf}~\cite{NORDUGRID_MANUAL_4}.
+Within Listing~\ref{lst:jsdl_job} an example using JSDL is provided. The
+specification of how to assign runtime environments in JSDL is currently
+only defined within the nordugrid jsdl-arc schema~
 \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/a-rex/grid-manager/jobdesc/jsdl/jsdl_arc.xsd}
@@
 
-In addition to using the Janitor together with A-REX,
-the Janitor can also be used as a standalone command line tool.
-The available commands are listed in table~\ref{janitor_commandline_man}.
+On Linux systems, the Janitor's standalone commandline tool is
+available as /usr/lib/arc/janitor.  Some Linux distributions may
+prefer /usr/libexec or similar paths.  The script is only functional as
+root\footnote{Should you find that constraint unbearable for your purpose,
+please investigate the file rjanitor.cc in the ARC source tree. It wraps
+the janitor application and as a C binary can be configured to attract
+root privileges.}.  To find that binary directly, you may decide to add
+that location to your \$PATH environment variable.
+
+The available commands to the Janitor, implemented as options to the janitor script, are listed
+in the Table~\ref{tab:janitor_commandline_man}.
+
 \begin{table}[!h]
    \begin{center}
-        \mycaption{Overview about the available commands in Janitor.}{}
+        \mycaption{Overview about the available commands to the Janitor.}{}
         \label{tab:janitor_commandline_man}
@@
 	\multicolumn{3}{l}{\textbf{Command:}}\\
-	&	register	& Registers a job and a set of runtime environments in the Janitor database.
+	&	register & Registers a job and a set of runtime
+			   environments in the Janitor database.
 			   Requires the parameters [JOB-ID] and a list of [RTE]s.\\
-	&	deploy		& Downloads and installs the desired runtime environments.
-				  Requires the name of an already registered [JOB-ID].\\
+	&	deploy	 & Downloads and installs the desired
+			   runtime environments. Requires the name
+			   of an already registered [JOB-ID].\\
 	&	remove	 & Removes the placeholder of the job on the runtime environments.
@@
 	\multicolumn{3}{l}{\textbf{Job id:}}\\
-	&			& A unique sequence of numbers. Once Janitor registered a job id, it
-				  cannot register a second job having the same job id.\\
+	&		 & A unique sequence of numbers. Once Janitor
+			   registered a job id, it cannot register a
+			   second job having the same job id.\\
 	\multicolumn{3}{l}{\textbf{Runtime environments:}}\\
-	&			& Runtime environments are defined by a continuous string.
-				  The name of valid runtime environment names can be investigated using
-				  the \texttt{list} or the \texttt{search} commands. They are defined in
-				  the catalog or by the directories and scripts of the \texttt{runtimedir}
+	&		& Runtime environments are defined by a continuous
+			  string.  The name of valid runtime environment
+			  names can be investigated using
+			  the \texttt{list} or the \texttt{search}
+			  commands. They are defined in the
+			  catalog or by the directories and
+			  scripts of the \texttt{runtimedir}
 			  of the \texttt{grid-manager}.\\
@@
 \texttt{deploy} and \texttt{remove}. To register a job along with a set
-of runtime environments in Janitor, the first command \texttt{register}
+of runtime environments in the Janitor, the first command \texttt{register}
 followed by a job identifier and a list of runtime environments has to be
@@
 are specified by a string containing the name as it is defined within
-the Catalog (resp. the runtime directory of the grid-manager).
-The command \texttt{deploy} extracts the necessary dependencies of the
-desired runtime environments to then download and install the required
-packages.
+the Catalog (resp. the runtime directory of the grid-manager).  The command
+\texttt{deploy} extracts the necessary dependencies of the desired
+dRTEs and then downloads and installs the required packages.
 
@@
 Easy command line examples are provided in Listing~\ref{lst:janitor_example}.
+You may also want to inspect the janitor(8) man page.
 
 Every command has a certain behaviour for its exit status.
-Table~\ref{tab:janitor_commandline_exit_status} lists the possible outcomes.
+Table~\ref{tab:janitor_commandline_exit_status} lists the possible
+outcomes. A value of 0 always indicates that no error occurred.
 

### 2009-11-29 doc/manuals/janitor/tex_future/future.tex
Merged future.tex.

@@
 to distinguish installed RTEs from installable RTEs, potentially offer
-description of extended RTE state-like information. This work is planned
+insights on the state that an dRTE is in. This work is planned
 to be carried out as part of the Glue-2.0 effort of OGF\footnote{OGF
@@
 manual additions and the functionality that comes with BaseSystem.
+
+\section{Use of RDF}
+
+The RDF is "correctly" used within the Janitor via the Redland libraries.
+What is still missing is a semantical reasoning on what packages to
+allow or disallow. For instance: allow all packages associated with
+bioinformatics. This shall wait a bit longer until more experiences
+on the community's demand for this service have been made.
+
+\section{Manual Verification}
+
+There is yet no explicit notion of a concept on how users can report
+on the reliability of individual compute nodes and use such information
+for the decision making on where to sent their jobs. And with an increasing
+complexity of software installed, being used across versions, one will
+rather trust one's very own experiences for individual sites than some
+external repository or other pieces of information that may be weeks
+or months old.
+
+Consequently, no means for a manual verification and the communication
+of results of such have yet been implemented.
 

### 2009-11-29 doc/manuals/janitor/tex_installation/installation.tex
janitor.pdf builds again. Many thanks to Anders for spotting the issue.

@@
 The Janitor requires the two perl packages listed in
-table~\ref{tab:install_dependencies}. To have the WebService
+table~\ref{tab:installDependencies}. To have the WebService
 interface for the Janitor, the packages listed in
-table~\ref{tab:install_dependencies_optional}
+table~\ref{tab:installDependenciesOptional}
 need to be installed before
@@
 with all major Linux distributions.
+
 \begin{table}[!h]
@@
 		   Runtime Environments.}
-        \label{tab:install_dependencies}
+        \label{tab:installDependencies}
 	\begin{tabular}{|p{3cm}|p{7cm}|}
@@
                    WebService to the Perl interpreter.}
-        \label{tab:install_dependencies_optional}
+        \label{tab:installDependenciesOptional}
 	\begin{tabular}{|p{3cm}|p{7cm}|}
@@
 
-\section{Configuration}\label{sec:janitor_configuration}
+\section{Configuration}\label{sec:janitorConfiguration}
 
@@
 \textit{/etc/arc.conf} will be used.  The configuration is assigned
-by the section \lbrack janitor\rbrack. Table~\ref{tab:arc_conf_tags}
+by the section \lbrack janitor\rbrack. Table~\ref{tab:arcConfTags}
 describes the available tags for the Janitor's configuration.
@@
         \mycaption{Tags usable in \textit{arc.conf} within the section janitor.}{Tags usable in \textit{arc.conf} within the section janitor.}
-	\label{tab:arc_conf_tags}
+	\label{tab:arcConfTags}
 	\begin{longtable}{|p{3cm}|p{10cm}||p{10cm}|}
@@
 \texttt{logconf}.  Examples on how to configure ARC and log4perl are
-provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.
+provided in the Listings~\ref{lst:arcConf} and ~\ref{lst:logConf}.
 
@@
 \begin{lstlisting}[
-        label=lst:arc_conf,
+        label=lst:arcConf,
         caption={ [Example \textit{arc.conf} settings for janitor.]
@@
 \begin{lstlisting}[
-        label=lst:log_conf,
+        label=lst:logConf,
         caption={ [Example \textit{log.conf} settings for janitor.]

### 2009-11-29 doc/manuals/janitor/tex_appendix/appendix.tex
deuglification

@@
  \item \textbf{Another document describing Janitor.}\\
-    D2.5-1 RDF Based Semantic Runtime Environment (RE) Description And Dynamic RE Management Framework Including Creating Proof Of Concept Bioinformatics REs, Daniel Bayer and Steffen Möller and Frederik Orellana\cite{BAYER_2007}
+    D2.5-1 RDF Based Semantic Runtime Environment (RE) Description And Dynamic RE Management Framework Including Creating Proof Of Concept Bioinformatics REs, Daniel Bayer and Steffen M\"oller and Frederik Orellana\cite{BAYER_2007}
 

### 2009-11-29 doc/manuals/janitor/tex_installation/installation.tex
deuglification

@@
    \begin{center}
-        \mycaption{Required perl packages for the Janitor.
-		   Log4perl is used for the internal logging of the
+        \mycaption{Required perl packages for the Janitor.}
+		   {Log4perl is used for the internal logging of the
 		   Janitor, while the Redland RDF library is used
@@
 and tested for various Linux distributions. It should also be functional
-on MaxOS X and Windows with Cygwin / CoLinux.  The porting of the Janitor
+on MaxOS X and Windows with Cygwin or coLinux.  The porting of the Janitor
 to other platforms has not yet been addressed.

### 2009-11-30 doc/manuals/echo/echo-manual.tex
Synchronising Echo and Hopi documentation with Nox

@@
   pdfauthor = {Zsombor Nagy},
-  pdftitle = {The Hopi service},
+  pdftitle = {The Echo service},
   pdfsubject = {Paper subject},
@@
 
-{\raggedleft NORDUGRID-MANUAL-XXX\par}
+% {\raggedleft NORDUGRID-MANUAL-XXX\par}
 
@@
 %%%%---- The title ----
-{\centering \textsc{\Large The Echo Service - The Missing Manual}\Large \par}
+{\centering \textsc{\Large The Echo Service Manual}\Large \par}
 \vspace*{0.5cm}
@@
 %%%%---- A list of authors ----
-    {\centering \large The Unknown Knowarcer\footnote{nordugrid-discuss@nordugrid.org} \large \par}
+    {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
 \end{titlepage}

### 2009-11-30 doc/manuals/configuration/configuration_manual.tex
Copying configuration documentation over from Nox

\documentclass{article}
%\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Martin Savko},
  pdftitle = {ARC configuration manual},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

% {\raggedleft NORDUGRID-MANUAL-10\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large ARC configuration manual}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
%{\centering \textit{\large First prototype status and plans}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
    {\centering \large Ferenc Szalai\footnote{szferi@wsbricks.com} \large \par}
    {\centering \large Martin Savko\footnote{martin.savko@gmail.com} \large \par}
\end{titlepage}

\tableofcontents                          %Comment if use article style

\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}

%\chapter{Quick start guide} % (fold)

\section{Introduction} % (fold)

The arched daemon's configuration file contains general information about the daemon's environment (pid file, log file, log level, library paths, plugins), and it contains one (or more) message chains, which specifies how the different services are connected to the outside world. The daemon's listening interface, certificates used, and the HTTP path of the services are among the things specified in the chain. This part contains also the configuration specifics for each service.

A chain could contain multiple types of components: message chain components (called MCCs) which provide TCP, TLS, HTTP and SOAP support; the plexer, which connects different HTTP paths to different services; and the services themselves with all their specific configuration details.

The configuration file is in XML format, with XML schemas and namespaces given and documented.

For typical deployment scenarios there is an additional mechanism to create an XML configuration on-the-fly based on XML templates (called “profiles”) and more administrator-friendly INI configuration files.

Configuration related arched command line switches:

\begin{verbatim}
	-c <xml config file>
	-i <ini config file>
	-d
	-s <xml schema>
\end{verbatim}

By default the daemon looks for the configuration files in the following order: first /etc/arc/server.ini, then if it is not found: /etc/arc/server.xml. However user can define which configuration  should be used with the -c or -i command line argument. If accidentally both of them is defined then the -i has greater priority.

In order the check the generated XML configuration the -d options can be used. If used, arched will  print the XML configuration to the standard output and then stop.

\section{XML level configuration}

\begin{itemize}
		\item The first part of the config is the “Server” part, which contains the path of the pid file, and the path of the log file, and the verbosity of the logging.
		\item The second part of the config is the “ModuleManager” part, which contains the path to the directory with the ARC libraries.
		\item The third part is the “Plugins” part with the list of 'plugins' (e.g. message chain components, services) used in this config.
		\item The fourth part is the “Chain” part, which contains the description of the message chain, which connects the services to the network. (It is possible to have more than one chain in one configuration.) A chain could contain the message chain components, the plexer and the services.
		\begin{itemize}
			\item The TCP message chain component does the listening on the network interface. This is the place to set the IP version to use, the IP and port to listen to.
			\item The TLS message chain component implements the security layer. This is the place to set the server-side key and the certificate, and the directory of the trusted CA certificates.
			\item The configuration of the HTTP and the SOAP message components most probably don't need to be changed.
			\item The Plexer component has a list of all available services, and has a HTTP path for each of them. The path is a regular expression, the incoming HTTP requests will be sent to the service which has a matching path in the Plexer config.
			\item Then come the specific configurations of all the services which will be running in this arched.
		\end{itemize}
\end{itemize}

\subsection{The service only gets its own config}

It is important to know that the services will not be able to get the whole configuration, they will only get their part, so everything a service should know have to be specified in its configuration part, e.g. a service would not know the IP or the port of the server, the HTTP path to itself or the path of the key and certificate file unless these are specified in the service configuration part too.

Because some services would establish some network connection while running (e.g. to send SOAP messages to other services), and these connections would most probably be secure HTTPS connection, these services may have a part where a client-side key and certificate file have to be specified which can be used to establish connection.

A service most probably would want to advertise its URL by registering itself to an information indexing service (called ISIS), in order to do this, the full URL of the service should be put into the configuration part of the service. This means that if the daemon's interface (port or hostname) would be changed, all these URLs should be modified too.

\subsection{Summary of where to change things}
\begin{itemize}
	\item pidfile, logfile, loglevel: the “Server” part
	\item path to the ARC libraries/modules/plugins: the “ModuleManager” part
	\item listening interface (port, IP, IP version): the TCP Component in the “Chain” part; and maybe the URLs of the services should be changed also in the configuration of the services.
	\item path to the server-side key and certificate files and CA dir: the TLS Component in the “Chain” part
	\item path to the client-side key and certificate files and CA dir: in the configuration of the individual services in the “Chain” part
	\item the HTTP path of the services: the Plexer in the “Chain” part, the paths are specified in regular expressions; and maybe the URL of the service should be changed also in the configuration of the service itself.
	\item the URL of the information indexing service (ISIS) where a service should advertise itself: in the config of each services in the “Chain” part
\end{itemize}

There are documentations and XML schemas for the possible configuration options of the message chain components (MCCs) and for all the services (TODO: link to some docs and stuff)

\section{INI style configuration}

Because the parameters which a system administrator would want to change while configuring arched are spread across the whole XML configuration file and sometimes even repeated several times, and because there are lots of things in the XML configuration file which would be always the same for a given deployment-scenario, it is possible to create different configuration profiles for different scenarios, and then configure these profiles with configuration files which are in a more user-friendly INI format containing key-value pairs grouped in sections. Then the XML config is generated from the given profile and the INI config. (A profile is basically a full XML configuration, but has some extra details about which part of the XML should be overwritten with which INI configuration parameter.)

For example there is a profile for a secure Echo service which contains a working set of message chain components and the Echo service with reasonable defaults for the paths of libraries, logfile, pidfile, certificates, the port to listen, the HTTP path of the service, etc. Then the system administrator can create an INI config which only contains the things which need to be non-default, e.g. if the key and certificate files are in a non-standard directory, the correct paths could be specified in the INI config, and these paths then would be put into the profile everywhere where needed (even multiple times).

Or if there is a profile with several services, all of them registering to an information index service (ISIS), which means that each service has the URL of the ISIS service in its config, this URL could be changed by adding one line to the INI config with the new URL, which makes this URL inserted into the config of each services.

\subsection{Issues}

Using the INI config is much easier for the system administrator than figuring out where to modify the XML config in order to change some paths or URLs. However using the INI config is much less flexible then using the XML config. Everything depends on the profiles. If the system administrator would need a specific set of services, and there is no profile for that, then the INI config can't be used. Or if there is profile but a given configuration parameter is not given in the profile, then it cannot be set in the INI config.

Another issue is that for each individual profile a separate documentation would be needed, because a profile can use arbitrary INI tags. e.g. the INI tag for setting the log level could be “loglevel” in one profile and it could be “debug” in another, or something completely different in the third one. The way to set the path to the directory of the trusted CAs' certificates could be “cacert=path” or “cacertificatesdir=path” depending which profile is in use. Currently there are example INI configs for most of the existing profiles, which should give hints about the possible INI tags.

\subsection{INI section}

The key-value pairs in an INI file are grouped into sections, like this:

\begin{verbatim}
	key1=value1
	key2=value2

	[section1]
	key1=value11
	key3=value3

	[section2]
	key5=value5
\end{verbatim}

The key-value pairs before the first section name are in the “common” section. The sections are used to selectively override arguments. e.g. the profile could say that it needs the value for “key1”, and then it could specify that it needs the value from the “common” section only, or from the |”section1” section only, or it could say that it needs the value from either of these two sections and the value from “section1” overrides the value from the “common” section, or vice versa.

For a more specific example, if there is a client-side key file given in the common section, then the profile could specify for each service, that it should use this key file unless there is a different key file specified in the section of the given service. This means that if there is an INI section for the service, and it contains a path to the key file, then that path will be used, but if it does not contain a path, then the common key path will be used from the “common” section.

\subsection{Profiles}

The profile is a full XML service configuration with some additional XML attributes about which part of the XML should be overwritten by which INI parameter. Let's look at this part of a profile:

\begin{verbatim}
	<infosys:URL inisections='register ahash' initag=\verb!index_server1!/>
\end{verbatim}

The initag attribute tells the key of the parameter, the inisections attribute specifies a list of sections with the last one having the highest priority. This XML tag here is empty, which means that it's default value is an empty string.

\begin{itemize}
	\item If the INI contains a key \verb!index_server1! somewhere in the 'register' or 'ahash' sections, then the value from that key would be put into this XML node.
	\item If there is an \verb!index_server1! key in both the 'register' and the 'ahash' section, then the value from the 'ahash' section would be used, it has the higher priority.
	\item If there is no \verb!index_server1! key in none of these two sections in the INI config, then this 'infosys:URL' tag will still be in the generated XML config, but with an empty content.
\end{itemize}

This part has a non-empty default:

\begin{verbatim}
	<cfg:File inisections='common' initag='logfile'>/var/log/arched.log</cfg:File>
\end{verbatim}

This means if that would be a 'logfile' key in the 'common' section of the INI configuration file, then this default value will be overwritten by the value of that key from the INI file. If there is no key called 'logfile' in the 'common' section of the INI file, then the generated XML config will contain the given default value.

\subsection{Limitations}

Currently overwriting XML attributes are not supported. e.g. the id attribute of:

\begin{verbatim}
	<next id='tls'/>
\end{verbatim}

can not be overwritten by the INI config.

Currently there is no support for XML tags with variable number of occurrences. This means that if there would be a part in the XML config like this:

\begin{verbatim}
	<ISISURL>the URL of an ISIS service</ISISURL>
	<ISISURL>the URL of another ISIS service</ISISURL>
	<ISISURL>the URL of a third ISIS service</ISISURL>
\end{verbatim}

It is not possible to control the number of ISISURL tags from the INI config. If there are three such lines in the profile, then there will be three lines in the generated XML config, no matter how many URLs are specified in the INI config. A workaround could be using empty defaults, like this:

\begin{verbatim}
	<lib:ISISURL inisections='register librarian' initag=\verb!index_server1!/>
	<lib:ISISURL inisections='register librarian' initag='index_server2'/>
	<lib:ISISURL inisections='register librarian' initag='index_server3'/>
\end{verbatim}

Then the INI config could be like this:

\begin{verbatim}
	[register]
	index_server1=the URL of an ISIS service
	index_server2=the URL of another ISIS service
\end{verbatim}

Which means that the first and second ISISURL line would be filled with data from the INI, and the third one would be left empty. For this to work, it should be supported by the service itself, e.g. the service should throw out empty ISISURLs.

Here it was needed to use different initags for the three ISISURLs, because if same initags would be used like this:

\begin{verbatim}
	<lib:ISISURL inisections='register librarian' initag='index_server'/>
	<lib:ISISURL inisections='register librarian' initag='index_server'/>
	<lib:ISISURL inisections='register librarian' initag='index_server'/>
\end{verbatim}

Then this INI config would cause unexpected results:

\begin{verbatim}
	[register]
	index_server=the URL of an ISIS service
	index_server=the URL of another ISIS service
\end{verbatim}

This INI config would generate this XML config from that profile:

\begin{verbatim}
	<lib:ISISURL>the URL of another ISIS service</lib:ISISURL>
	<lib:ISISURL>the URL of another ISIS service</lib:ISISURL>
	<lib:ISISURL>the URL of another ISIS service</lib:ISISURL>
\end{verbatim}

Which is the same URL three times, because the last value of the \verb!index_server! tag would be inserted into each places.

There are future plans for adding support for this kind of simple multiplicity, but there are more complex cases which are not trivial to handle. e.g. in order to register to multiple ISIS services, the path to the key and certificate files should be given for each ISIS service even though they most probably would be the same. This part would register to one ISIS service:

\begin{verbatim}
	<infosys:Registrar>
		<infosys:URL inisections='register shepherd' initag='index_server'/>
		<infosys:KeyPath inisections='common shepherd' initag='client_key'/>
		<infosys:CertificatePath inisections='common shepherd' initag='client_cert'/>
		<infosys:CACertificatesDir inisections='common shepherd' initag='cacert'/>
	</infosys:Registrar>
\end{verbatim}

It is possible that we want to register only to one ISIS or two or maybe three or more, then for each of them this section should be repeated. Currently the workaround looks like this:

\begin{verbatim}
	<infosys:Registrar>
		<infosys:URL inisections='register shepherd' initag=\verb!index_server1!/>
		<infosys:KeyPath inisections='common shepherd' initag='client_key'/>
		<infosys:CertificatePath inisections='common shepherd' initag='client_cert'/>
		<infosys:CACertificatesDir inisections='common shepherd' initag='cacert'/>
	</infosys:Registrar>
	<infosys:Registrar>
		<infosys:URL inisections='register shepherd' initag='index_server2'/>
		<infosys:KeyPath inisections='common shepherd' initag='client_key'/>
		<infosys:CertificatePath inisections='common shepherd' initag='client_cert'/>
		<infosys:CACertificatesDir inisections='common shepherd' initag='cacert'/>
	</infosys:Registrar>
	<infosys:Registrar>
		<infosys:URL inisections='register shepherd' initag='index_server3'/>
		<infosys:KeyPath inisections='common shepherd' initag='client_key'/>
		<infosys:CertificatePath inisections='common shepherd' initag='client_cert'/>
		<infosys:CACertificatesDir inisections='common shepherd' initag='cacert'/>
	</infosys:Registrar>
\end{verbatim}

Which allows to register to up to three ISIS services, and has empty defaults in order to be able to configure only one or two or zero ISIS services. However, the ISIS registration part should handle the empty defaults, and throw them away.

\section{Examples}

The most up to date Profiles can be found in the Nordugrid SVN:

\verb!http://svn.nordugrid.org/trac/nordugrid/browser/arc1/branches/nox/src/hed/profiles!

Upon a successful installation (nordugrid-arc-nox-hed package when installing from binaries), all the example configuration files coming with the release can be found under

\verb!<instllation_directory>/share/arc/!

\end{document}

### 2009-12-01 doc/tech_doc/a-rex/arex_tech_doc.tex
cleaning up and clarifying configuration options

@@
 
-Due to historical reasons configuration of the A-REX is split into
-2 parts. The small part is located inside HED configuration (see Appendix~\ref{annex:arex-conf}) for schema and description of supported elements. It refers to
-file containing main part of configuration. The default location of
-main configuration file is \textit{/etc/arc.conf}.
+Due to historical reasons configuration of the A-REX is split into 2
+parts. The small part is located inside HED configuration (see
+Appendix~\ref{annex:arex-conf} for schema and description of supported
+elements). This configuration refers to the file containing the main
+part of the configuration. The default location of the main
+configuration file is \textit{/etc/arc.conf}.
 
@@
 starting with \#. It is separated into sections. Each section starts
-from string containing pattern
+with a string containing
 
@@
 
-Each section continues till next section or end of file. One configuration
-file can have commands for multiple services/modules/programs. Each
-service get it's own section named after it. The A-REX uses section
-\emph{{[}grid-manager]}. Some services can make use of multiple subsections
-to reflect their internal modular structure. Commands in section \emph{{[}common]}
-apply to all services. Command lines have format
+Each section continues until the next section or until the end of the
+file. The configuration file can have commands for multiple
+services/modules/programs. Each service has its own section named
+after it. The A-REX uses the \emph{{[}grid-manager]} section. Some
+services can make use of multiple subsections to reflect their
+internal modular structure. Commands in section \emph{{[}common]}
+apply to all services. Command lines have the format
 
@@
 \end{shaded}
-Following commands are defined:
-
-\begin{itemize}
-\item \textbf{\textit{joblog}}\textit{={[}path]} -- specifies where to store
+The following commands are defined:
+
+Commands affecting the A-REX process and logging:
+
+\begin{itemize}
+\item \textbf{\textit{pidfile}}\textit{=path} -- specifies file where
+  process id of A-REX process will be stored. Defaults to
+  \emph{/var/log/arched.pid} if running as root and
+  \emph{\$HOME/arched.pid} otherwise.
+\item \textbf{\textit{logfile}}\textit{=path} -- specifies name of file
+  for logging debug/informational output. Defaults to
+  \emph{/var/log/grid-manager.log}.
+\item \textbf{\textit{logsize}}\textit{=size number} -- restricts
+  log file size to \emph{size} and keeps \emph{number} archived log
+  files.
+\item \textbf{\textit{debug}}\textit{=number} -- specifies level of
+  debug information. More information is printed for higher
+  levels. Currently the highest effective number is 5 (DEBUG) and
+  lowest 0 (FATAL). Defaults to 2 (WARNING).
+\item \textbf{\textit{user}}\textit{=username} -- specifies username to
+  which the A-REX must switch after reading configuration. Defaults to
+  \emph{not switch}.
+\end{itemize}
+
+Commands setting limits and options for how the A-REX handles jobs and
+files:
+
+\begin{itemize}
+\item \textbf{\textit{joblog}}\textit{=path} -- specifies where to store
 log file containing information about started and finished jobs.
-\item \textbf{\textit{jobreport}}\textit{={[}URL ... number]} -- specifies
+\item \textbf{\textit{jobreport}}\textit{=URL ... number} -- specifies
 that A-REX has to report information about jobs being processed (started,
-finished) to centralized service running at given \textit{URL}. Multiple
+finished) to a centralized service running at the given \textit{URL}. Multiple
 entries and multiple URLs are allowed. \textit{number} specifies how
-long old records have to be kept if failed to be reported. That value
-is specified in days. Last specified value becomes effective.
-\item \textbf{\textit{jobreport\_credentials}}\textit{={[}key\_file
-    {[}cert\_file {[}ca\_dir]]]} -- specifies the credentials for
+long (in days) old records have to be kept if failed to be
+reported. The last specified value becomes effective.
+\item \textbf{\textit{jobreport\_credentials}}\textit{=key\_file
+    {[}cert\_file {[}ca\_dir]]} -- specifies the credentials for
   accessing the accounting service.
-\item \textbf{\textit{jobreport\_options}}\textit{={[}options]}
+\item \textbf{\textit{jobreport\_options}}\textit{=options}
   -- specifies additional options for Usage Reporter and/or
@@
 to use encryption while transferring data. Currently works for GridFTP
-only. Default is \emph{no}. It is overridden by value specified in
+only. Default is \emph{no}. It is overridden by values specified in
 URL options.
@@
   can solve transfer problems caused by firewalls. Default is no.
-
 \item \textbf{\textit{localtransfer}}\textit{=yes|no} -- specifies whether
 to pass file downloading/uploading task to computing node. If set
-to yes the A-REX won't initiate download/upload files. Instead it
-composes script submitted to LRMS in way to make it do that. This
+to yes the A-REX will not download/upload files but compose script
+submitted to the LRMS in order that the LRMS can execute file tranfer. This
 requires installation of A-REX and all related software to be accessible
@@
 \item \textbf{\textit{maxjobs}}\textit{={[}max\_processed\_jobs {[}max\_running\_jobs]]}
-- specifies maximum number of jobs being processed by the A-REX at
+-- specifies maximum number of jobs being processed by the A-REX at
 different stages:\\
-\textit{max\_processed\_jobs} -- maximal amount of jobs being processed
-by A-REX. This does not limit amount of jobs, which can be submitted
-to A-REX\\
-\textit{max\_running jobs} -- maximal amount of jobs passed to Local
+\textit{max\_processed\_jobs} -- maximum number of concurrent jobs
+processed by A-REX. This does not limit the amount of jobs, which can
+be submitted to the cluster.\\
+\textit{max\_running jobs} -- maximum number of jobs passed to Local
 Resource Management System\\
@@
 jobs being processed on frontend:\\
-\textit{max\_frontend\_jobs} -- maximal amount of jobs heavily using
-resources of frontend (applied before moving job to PREPARING and
-FINISHING states)\textit{}\\
+\textit{max\_frontend\_jobs} -- maximum number of jobs in PREPARING and
+FINISHING states (downloading and uploading files). Jobs in these
+states can cause a heavy load on the A-REX host. This limit is applied
+before moving jobs to PREPARING and FINISHING states.\\
 \textit{emergency\_frontend\_jobs} -- if limit of \textit{max\_frontend\_jobs}
-is used only by PREPARING or by FINISHING jobs aforementioned number
-of jobs can be moved to another state. This is used to avoid case
-then jobs can't finish due to big amount of recently submitted jobs.\textit{}\\
-\textit{max\_transfered\_files} -- maximal number of files being transferred
+is used only by PREPARING or by FINISHING jobs, aforementioned number
+of jobs can be moved to another state. This is used to avoid the case
+where jobs cannot finish due to a large number of recently submitted jobs.\\
+\textit{max\_transferred\_files} -- maximum number of files being transferred
 in parallel by every job. Used to decrease load on not so powerful
@@
 \item \textbf{\textit{authplugin}}\textit{=state options plugin} -- specifies
-\emph{plugin} (external executable) to be run every time job is going
-to switch to \emph{state}. Following states are allowed: ACCEPTED,
+\emph{plugin} (external executable) to be run every time job is about
+to switch to \emph{state}. The following states are allowed: ACCEPTED,
 PREPARING, SUBMIT, FINISHING, FINISHED and DELETED. If exit code is
-not 0 job is canceled by default. \textit{Options} consist of \textit{name}=\textit{value}
-pairs separated by a comma. Following \textit{name}s are supported:\\
+not 0 job is canceled by default. \textit{Options} consists of \textit{name}=\textit{value}
+pairs separated by commas. The following \textit{name}s are supported:\\
 \textit{timeout} -- specifies how long in seconds execution of the
@@
 or if average rate is lower than \emph{min\_average\_speed}, or no
-data is received for longer than \textit{max\_inactivity} seconds.
+data is received for longer than \textit{max\_inactivity} seconds. To
+allow statistics to build up, no transfers will be stopped within the
+first 3 minutes.
 \item \textbf{\textit{copyurl}}\textit{=template replacement} -- specifies
-that URLs, starting from template should be accessed in a different
+that URLs starting from \emph{template} should be accessed in a different
 way (most probably Unix open). The \textit{template} part of the URL
@@
 \item \textbf{\textit{linkurl}}\textit{=template replacement {[}node\_path]}
--- mostly identical to \textit{copyurl} but file won't be copied. Instead
-soft-link will be created. \textit{replacement} specifies the way
+-- mostly identical to \textit{copyurl} but file will not be copied. Instead
+a soft-link will be created. \textit{replacement} specifies the way
 to access the file from the frontend, and is used to check permissions.
@@
 computing nodes, and will be used for soft-link creation. If \textit{node\_path}
-is missing, \textit{local\_path} will be used instead. Both \textit{node\_path}
-and \textit{replacement} should not be URLs.
+is missing, \textit{local\_path} will be used instead. Neither \textit{node\_path}
+nor \textit{replacement} should be URLs.
 \begin{framed}
@@
 treated as more easily accessible than other URLs. That means if A-REX
-has to choose between few URLs from which should it download input
+has to choose between several URLs from which should it download input
 file, these will be tried first.
@@
 
-Per UNIX user commands:
+Per UNIX user commands
 
@@
 \item \textbf{\textit{mail}}\textit{=e-mail\_address} -- specifies an email
-address \textbf{from} which the notification mails are sent.
+address \textbf{from} which notification mails are sent.
 \item \textbf{\textit{defaultttl}}\textit{=ttl {[}ttr]} -- specifies the
-time in seconds for the SD to be available after job finished (\emph{ttl})
+time in seconds for the SD to be available after job finishes (\emph{ttl})
 and after job was deleted (\emph{ttr}) due to \emph{ttl}. Defaults
-are 7 days for \emph{ttl} and 30 days for \emph{ttr}.
+are 7 days for \emph{ttl} and 30 days for \emph{ttr}. The minumum value
+for both parameters is 2 hours.
 \item \textbf{\textit{lrms}}\textit{=default\_lrms\_name default\_queue\_name}
-- specifies names for the LRMS and queue. Queue name can also be specified
-in the JD (currently it is not allowed to override LRMS by usingon
+-- specifies names for the LRMS and queue. Queue name can also be specified
+in the JD (currently it is not allowed to override LRMS by using
 the JD).
-\item \textbf{\textit{sessiondir}}\textit{=path} -- specifies path to the directory
-in which the SD is created. If the path is {*} the default one is
-used -- \textit{\$HOME/.jobs}.
-\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} - specifies
+\item \textbf{\textit{sessiondir}}\textit{=path} -- specifies path to
+  the directory in which the SD is created. If \emph{path} is {*} the
+  default sessiondir is used -- \textit{\$HOME/.jobs}.
+\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} -- specifies
 a directory to store cached data (see section \ref{sec:cache}). Multiple
@@
 are copied to the session directory.
-\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} -
+\item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} --
   specifies high and low watermarks for space used on the file system
@@
 of times job will be allowed to rerun after it failed at any stage.
-Default value is \emph{5}. This only specifies a upper limit. Actual
+Default value is \emph{5}. This only specifies a upper limit. The actual
 number is provided in job description and defaults to 0.
@@
 
-All per-user commands should be put before \textit{control} command
+All per-user commands should be put before the \textit{control} command
 which initiates serviced user.
@@
 \begin{itemize}
-\item \textbf{\textit{control}}\textit{=path username {[}username {[}...]]}
-- This option initiates UNIX user as being serviced by the A-REX.
-\textit{path} refers to the control directory (see Section~\ref{section:internals}
-for the description of control directory). If the path is {*} the
-default one is used -- \verb|$HOME/.jobstatus|. \textit{username} stands
-for UNIX name of the local user. Multiple names can be specified.
-If the name is {*} it is substituted by all names found in file \texttt{/etc/grid-security/grid-mapfile}
-(for the format of this file one should study the Globus project~\cite{globus}).\\
-Also the special name '.'(dot) can be used. Corresponding control
-directory will be used for \textbf{any} user. This option should be
-the last one in the configuration file. There is also command \textbf{\textit{controldir}}\textit{=path}.
-It uses special username '.' and is always executed last independent
-of placement in file.
-\item \textbf{\textit{helper}}\textit{=username command {[}argument {[}argument
-{[}...]]]} -- associates external program with the local UNIX user.
-This program will be kept running under account of the specified user.
-\textit{username} stands for the name of the user. Special names can
-be used: '{*}' -- all names from /etc/grid-security/grid-mapfile, '.'
-- root user. The user should be already configured with \textit{control}
+\item \textbf{\textit{control}}\textit{=path username {[}username
+    {[}...]]} -- This option initiates UNIX user as being serviced by
+  the A-REX.  \textit{path} refers to the control directory (see
+  Section~\ref{section:internals} for the description of control
+  directory). If the path is {*} the default one is used --
+  \verb|$HOME/.jobstatus|. \textit{username} stands for UNIX name of
+  the local user. Multiple names can be specified.  If the name is {*}
+  it is substituted by all names found in file
+  \texttt{/etc/grid-security/grid-mapfile} (for the format of this
+  file one should study the Globus project~\cite{globus}).\\ Also the
+  special name '.'(dot) can be used. Corresponding control directory
+  will be used for \textbf{any} user. This option should be the last
+  one in the configuration file. There is also command
+  \textbf{\textit{controldir}}\textit{=path}.  It uses special
+  username '.' and is always executed last independent of its
+  placement in file.
+\item \textbf{\textit{helper}}\textit{=username command {[}argument
+    {[}argument {[}...]]]} -- associates an external program with the
+  local UNIX user.  This program will be kept running under account of
+  the user specified by \textit{username}. Special names can be used:
+  '{*}' -- all names from /etc/grid-security/grid-mapfile, '.'  - root
+  user. The user should be already configured with \textit{control}
   option (except root, who is always configured). \textit{command} is
@@
 
+The following are global commands specific to communication with the underlying
+LRMS.
+
 \begin{comment}
-
-Following are global commands specific to communication with underlying
-LRMS (PBS in this case).
 
@@
 \begin{itemize}
-\item \textbf{\textit{gnu\_time}}\textit{=path} -- path to \emph{time} utility.
+\item \textbf{\textit{gnu\_time}}\textit{=path} -- path to \emph{time}
+  utility.
 \item \textbf{\textit{tmpdir}}\textit{=path} -- path to directory for temporary
@@
 \begin{itemize}
-\item [{\%R}] -- session root -- see command \emph{session}
+\item [{\%R}] -- session root -- see command \emph{sessiondir}
 \item [{\%C}] -- control dir -- see command \emph{control}
@@
 \item [{\%H}] -- home dir -- home specified in /etc/passwd
-\item [{\%Q}] -- default queue -- look command 'defaultlrms'
-\item [{\%L}] -- default lrms -- look command 'defaultlrms'
+\item [{\%Q}] -- default queue -- see command \emph{lrms}
+\item [{\%L}] -- default lrms -- ses command \emph{lrms}
 \item [{\%W}] -- installation path -- \$\{ARC\_LOCATION\}
@@
 \item [{\%c}] -- list of all control directories
-\item [{\%I}] -- job's ID (for plugins only, substituted in runtime)
-\item [{\%S}] -- job's state (for \emph{authplugin} plugins only, substituted
+\item [{\%I}] -- job ID (for plugins only, substituted in runtime)
+\item [{\%S}] -- job state (for \emph{authplugin} plugins only, substituted
 in runtime)

### 2009-12-01 doc/tech_doc/a-rex/arex_tech_doc.tex
Getting rid of symbolic links (they don't work on Windows!), correcting Makefiles

@@
 For information about supported LRMSes and their specific features
-and configuration options please read dedicated documentation~\cite{backends-arc1}.
+and configuration options please read dedicated documentation~\cite{arc1-backends}.
 
@@
 
-\bibliography{grid}
+\bibliography{grid,nordugrid}
 \end{document}

### 2009-12-06 doc/manuals/user_interface/ui-job.tex
Flag dolocalsandbox is not used. Default is to add sandbox to job list file.
Editing and updating the user interface manual.

@@
 \begin{shaded}
-   \uicommand{arcsub [options] [filename]}
+   \uicommand{arcsub [options] [filename ...]}
 \end{shaded}
@@
    \texttt{-c, --cluster}&\verb#[-]#\textit{url}&explicitly select or reject (-) a specific site\\
-   \texttt{-i, --index}&\textit{url}&explicitly select or reject (-) a specific index server\\
+   \texttt{-i, --index}&\verb#[-]#\textit{url}&explicitly select or reject (-) a specific index server\\
    \texttt{-e, --jobdescrstring}&\textit{filename}&string describing the job to be submitted\\
@@
    \texttt{-b, --broker}&\textit{string}&select broker method (default is Random)\\
-   \texttt{-n, --dolocalsandbox}& &store job descriptions in local sandbox (useful for eventual resubmission/migration)\\
+%   \texttt{-n, --dolocalsandbox}& &store job descriptions in local sandbox (useful for eventual resubmission)\\
    \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
@@
    Arguments:&&\\
-   \texttt{filename}&&file describing the job(s) to be submitted\\
+   \texttt{filename ...}&&file(s) describing the job(s) to be submitted\\
 \end{longtable}
@@
 
+%%% This might depend on the version of glibmm installed. In recent versions atleast it
+%%% does not matter where the job description filename occurs on the command line. Any argument
+%%% without a corresponding flag is taken to be a job description filename.
 As a shorthand \texttt{-f} can be omitted if the job description file is put last on the commandline.
@@
    </Application>
-   <DataStaging>
-     <FileName>out.txt</FileName>
-     <CreationFlag>overwrite</CreationFlag>
-     <DeleteOnTermination>false</DeleteOnTermination>
-   </DataStaging>
-   <DataStaging>
-     <FileName>err.txt</FileName>
-     <CreationFlag>overwrite</CreationFlag>
-     <DeleteOnTermination>false</DeleteOnTermination>
-   </DataStaging>
  </JobDescription>
@@
 \begin{framed}
+%% This depends on the middleware flavour. Some flavours do not accept listing
+%% the job session directory (CREAM).
    Every job ID is a valid URL for the job session directory. You can
@@
 
-The job description in xRSL or JSDL format can be given either as an
+The job description in one of the supported languages can be given, either as an
 argument on the command line, or can be read from a file. Several jobs can be
-requested at the same time by giving more than one filename
-argument, or by repeating the \verb#-f# or \verb#-e# options. It is
-possible to mix \verb#-e# and \verb#-f# options in the same
-\texttt{arcsub} command.
+requested at the same time by giving more than one filename argument, or by
+repeating the \verb#-f# or \verb#-e# options. It is possible to mix \verb#-e#
+and \verb#-f# options in the same \texttt{arcsub} command.
 
+%%% The below path is OS dependend.
 In order to keep track of submitted jobs, ARC client stores information in a
@@
 that can be understood by the Grid services to which it is being
-submitted. By specifying the \verb#-dumpdescription# option, such transformed
+submitted. By specifying the \verb#--dumpdescription# option, such transformed
 description is written to stdout instead of being submitted to the remote site.
@@
 
-If you plan to resubmit or migrate jobs, you will have to use command line
-option \verb#-n#, which will instruct the client to store complete job descriptions
-in a local sandbox, such that a resubmitted/migrated job will be identical to the
-original one.
+% By default local sandbox is enabled, and it is not possible atm to turn it
+% off.
+%If you plan to resubmit jobs, you will have to use command line option
+%\verb#-n#, which will instruct the client to store complete job descriptions
+%in a local sandbox, such that a resubmitted job will be identical to the
+%original one.
 
@@
 
-Default configuration file is {\$}HOME/.arc/client.conf. However, a user can choose any
+Default configuration file is \texttt{{\$}HOME/.arc/client.conf}. However, a user can choose any
 other pre-defined configuration through option \verb#-z#.

### 2009-12-07 doc/manuals/user_interface/ui-job.tex
Added arcinfo, notes on OR in arcstat/arccat/arcget etc, some more translations

@@
 
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand returns information for \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arcstat -s Finished -c mycluster <jobid># will return information about all finished
+  jobs on the Grid, plus about all jobs (in any state) on the cluster \verb#mycluster#, plus about the job \verb#<jobid>#.
+\end{framed}
+
 If the \verb#-l# option is given, extended information is printed.
@@
 
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand prints logs for \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arccat -s Finished -c mycluster <jobid># will print logs of all finished
+  jobs on the Grid, plus of all jobs (in any state) on the cluster \verb#mycluster#, plus of the job \verb#<jobid>#.
+\end{framed}
+
 \subsection{arcget}
@@
 
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand retrieves \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arcget -s Finished -c mycluster <jobid># will retrieve all finished
+  jobs on the Grid, plus all jobs (in any state) on the cluster \verb#mycluster#, plus the job \verb#<jobid>#.
+\end{framed}
+
 \subsection{arcsync}
@@
 
+\subsection{arcinfo}
+\label{sec:arcinfo}
+
+The \texttt{arcinfo} \index{arcinfo}\index{commands:arcinfo} command is used to obtain status information about clusters on the Grid.
+
+\hspace*{0.5cm}
+\begin{shaded}
+   \uicommand{arcinfo [options]}
+\end{shaded}
+\versions{ARC 0.9}
+\begin{longtable}{llp{8cm}}
+   Options:&&\\
+   \texttt{-c, --cluster}&\verb#[-]#\textit{name}&explicitly select or reject a specific site\\
+   \texttt{-i, --index}&\textit{url}&explicitly select or reject (-) a specific index server\\
+   \texttt{-l, --long}& & long format (extended information)\\
+   \texttt{-t, --timeout}&\textit{seconds}&timeout in seconds (default 20)\\
+   \texttt{-d, --debug}&\textit{debuglevel}&debug level, FATAL, ERROR, WARNING, INFO, VERBOSE or DEBUG - default WARNING\\
+   \texttt{-z, --conffile}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.conf)\\
+   \texttt{-v, --version}&&print version information\\
+   \texttt{-h, --help}&&print help page\\
+\end{longtable}
+
+The  \texttt{arcinfo}  command is used to obtain information about clusters and queues (\textit{targets})
+available on the Grid. Either the \verb#--cluster# or \verb#--index# flag should be used to specify the
+target(s) which should be queried for information. Both of these flags take a service endpoint as argument.
+See \texttt{arcsub} and the configuration notes in Section~\ref{sec:client.conf} for description of these.
+
+Detailed information about queried computing services can be obtained by specifying the \verb#--long# flag.
+
+When  specifying  the  \verb#--index#  flag,  the  information about the computing services registered at the index server will
+be queried, rather than the status of the index server itself.
+
 \subsection{arckill}
@@
 
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand kills \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arckill -s INLRMS:R -c mycluster <jobid># will kill all running
+  jobs on the Grid, plus all jobs (in any state) on the cluster \verb#mycluster#, plus the job \verb#<jobid>#.
+\end{framed}
+
 \subsection{arcclean}
@@
 Other command line options have the same meaning as the corresponding ones of \verb#arcstat# and others.
+
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand cleans \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arcclean -s FAILED -c mycluster <jobid># will clean all failed
+  jobs on the Grid, plus all jobs (in any state) on the cluster \verb#mycluster#, plus the job \verb#<jobid>#.
+\end{framed}
 
@@
 
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand renews proxies for \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arcrenew -s FAILED -c mycluster <jobid># will renew proxies of all failed
+  jobs on the Grid, plus of all jobs (in any state) on the cluster \verb#mycluster#, plus of the job \verb#<jobid>#.
+\end{framed}
+
 \subsection{arcresume}
@@
 Command line options have the same meaning as the corresponding ones of \verb#arcstat# and others.
+
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand resumes \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arcresume -s FAILED -c mycluster <jobid># will resume all failed
+  jobs on the Grid, plus all jobs (in any state) on the cluster \verb#mycluster#, plus the job \verb#<jobid>#.
+\end{framed}
 
@@
 
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand resubmits \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arcresub -s FAILED -c mycluster <jobid># will resubmit all failed
+  jobs on the Grid, plus all jobs (in any state) on the cluster \verb#mycluster#, plus the job \verb#<jobid>#.
+\end{framed}
+
 \subsection{arcmigrate}
@@
 identifies the job while it is being executed.
+
+\begin{framed}
+  When several of the \verb#-a#,  \verb#-j#, \verb#-c#, \verb#-s# and \verb#[job...]#
+  command line options are specified, the comand migrates \textbf{ALL} jobs that
+  match either of the criteria (logical OR).
+
+  For example, \verb#arcmigrate -s FAILED -c mycluster <jobid># will migrate all failed
+  jobs on the Grid, plus all jobs (in any state) on the cluster \verb#mycluster#, plus the job \verb#<jobid>#.
+\end{framed}

### 2009-12-15 doc/tech_doc/a-rex/arex_tech_doc.tex
Information about transfer shares added to the A-REX manual

@@
 the job moves to \textbf{Finishing} state. It is possible to put a
-limit on the number of simultaneous \textbf{Preparing} jobs. If this
-limit is exceeded, jobs ready to enter the \textbf{Preparing} state
+limit on the number of simultaneous \textbf{Preparing} jobs. 
+If this limit is exceeded, jobs ready to enter the \textbf{Preparing} state
  will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
@@
 not specified as output files are removed from the SD at very beginning
-of this state. It is possible to limit number of simultaneous jobs
-in this state.
+of this state.It is possible to limit the number of simultaneous jobs in this state.
 \item \textbf{Finished} -- No more processing is performed by the A-REX.
@@
 job fails.
+\item \textit{transfershare} -- name of share used in \textbf{Preparing}
+and \textbf{Finishing} states.
 \end{itemize}
@@
 Missing value or -1 means no limit.
+\item \textbf{\textit{maxloadshare}}\textit{=max\_share share\_type} -- specifies a 
+sharing mechanism for data transfer. \emph{max\_share} is the maximum
+number of processes that can run per transfer share and \emph{share\_type} is 
+the scheme used to assign jobs to transfer shares. See Section~\ref{sub:Transfershare}
+for possible values and more details.
 \item \textbf{\textit{wakeupperiod}}\textit{=time} -- specifies how often
@@
 \end{itemize}
+
+\subsection{Transfer shares\label{sub:Transfershare}}
+
+For many jobs, large amounts of input and output data can mean
+significant time is spent in the PREPARING and FINISHING states
+gathering input data and writing output data. With FIFO processing,
+this can lead to one user or group of users blocking the queue for
+others. The A-REX implements a sharing system to avoid this problem, by
+assigning each user or group of users to a ``transfer share'' and
+specifying a limit on the number of data transfer processes per
+share. If one user's jobs' transfer share is using the maximum number
+of processes and another user submits jobs which are assigned to a
+different share, the second user's jobs can immediately go to
+PREPARING, up to the same maximum limit of processes. This means that
+no matter how many jobs the first user submits, the second user's jobs
+are not blocked. Assuming the bandwidth from the sources of input data
+for both users' jobs is similar, the available throughput will then be
+split evenly between the two users' jobs.
+
+If a limit on the total number of data transfer processes is set in
+the \emph{maxload} option, the maximum number of processes per
+transfer share is set by splitting the total maximum evenly among all
+the shares with jobs in data transfer states, up to the maximum
+allowed per share.
+
+The scheme used to assign jobs to transfer shares can be set in the
+\emph{maxloadshare} option. Possible values are:
+
+\begin{itemize}
+\item \emph{dn} - each job is assigned to a share based on the DN of
+  the user sumbitting the job.
+\item \emph{voms:vo} - if the user's proxy is a VOMS~\cite{voms} proxy
+  the job is assigned to a share based on the VO specified in the
+  proxy. If the proxy is not a VOMS proxy a default share is used.
+\item \emph{voms:role} - if the user's proxy is a VOMS proxy the job
+  is assigned to a share based on the role specified in the first
+  attribute found in the proxy. If the proxy is not a VOMS proxy a
+  default share is used.
+\item \emph{voms:group} - if the user's proxy is a VOMS proxy the job is
+  assigned to a share based on the group specified in the first
+  attribute found in the proxy. If the proxy is not a VOMS proxy a
+  default share is used.
+\end{itemize}
+
+If VOMS is not supported, the \emph{dn} scheme is the only option that
+should be used, as using a VOMS-based scheme will lead to all jobs
+being assigned to the default share.
+
+\textbf{Important:} If a sharing mechanism based on VOMS is used,
+server certificates for each supported VO must be installed. It is
+possible to either download the public key of each VOMS server, or
+create a special file for each VO containing the server's DN and its
+CA DN. Instructions are given on NorduGrid's web site at
+\url{http://www.nordugrid.org/documents/voms-notes.html}.
 

### 2009-12-16 doc/tech_doc/a-rex/arex_tech_doc.tex
gm-jobs -s switch ported from ARC0 to ARC1 with proper documentation update

@@
 and the number of jobs in each state.\\
-\hspace*{0.5cm}gm-jobs {[}-h] {[}-l] {[}-u uid] {[}-U name] {[}-c conf\_file] {[}-d control\_dir]\\
+\hspace*{0.5cm}gm-jobs {[}-h] {[}-s] {[}-l] {[}-u uid] {[}-U name] {[}-c conf\_file] {[}-d control\_dir]\\
 \hspace*{0.5cm}-h -- print short help,\\
+\hspace*{0.5cm}-s -- print summary of jobs in each transfer share,\\
 \hspace*{0.5cm}-l -- print more information about each job,\\
@@
 should be used, as using a VOMS-based scheme will lead to all jobs
-being assigned to the default share.
+being assigned to the default share. The current number of jobs
+processing and pending processing for each share can be seen with the
+command \emph{gm-jobs -s}.
 

### 2009-12-17 doc/tech_doc/infosys/infosys_technical.tex
Fixed bibliography

--- 2009-11-10 doc/tech_doc/infosys/infosys_technical.tex
@@
 
-\textit{This section extensively uses terms defined in "Security Framework of WS-ARC" document"~\cite{arcsec}.}
+\textit{This section extensively uses terms defined in "ARC Security Framework (NORDUGRID-TECH-16)" document"~\cite{arc1-security}.}
 
@@
 
-Those 3 action groups are handled using Security Framework of ARC~\cite{arcsec}. For each group a corresponding
+Those 3 action groups are handled using Security Framework of ARC~\cite{arc1-security}. For each group a corresponding
 Action is defined for the ARC policy language as described in table \ref{tab:policy_actions}.
@@
 
-\textit{For information about sophisticated authorization policies and how to deploy various Policy Decision Point entities please see "Security Framework of WS-ARC" document"~\cite{arcsec}.}
+\textit{For information about sophisticated authorization policies and how to deploy various Policy Decision Point entities please see "ARC Security Framework (NORDUGRID-TECH-16)" document"~\cite{arc1-security}.}
 
@@
 an ISIS service. In lines 7--17 a set of Security Handlers and Policy Decision Points is configured to handle
-policies written in the ARC Policy Language (described in ~\cite{arcsec}). Line 14 specifies that the policy is 
+policies written in the ARC Policy Language (described in~\cite{arc1-security}). Line 14 specifies that the policy is
 read from file \textit{/opt/arc/etc/isis/policy.xml} every time a new request arrives. If the policy is not 

### 2009-12-18 doc/tech_doc/a-rex/arex_tech_doc.tex
Small documentation and example update to reflect the configuration of transfer shares through XML
add new cache features

@@
 file. All input files except files uploaded by the user during job
-submission are cached by default. This includes excutable files
+submission are cached by default. This includes executable files
 downloaded by the A-REX. Caching can be explicitly turned off by the
@@
 
-Cached files are stored in sub-direcories under the \emph{data} directory
-in each main cache directory. Filenames are constructed from an SHA-1
-hash of the URL of the file, and split into subdirectories based on
-the two initial characters of the hash. This enables the cached files
-to be evenly split over a number of subdirectories. If
-more than one cache directory is used the initial letter of the hash
-also determines which cache is to be used. The algorithm is simply
-to use the cache directory at index \emph{i} in the list of directories,
-where \emph{i} is found from the initial letter of the hash mod the
-number of caches. This limits the maximum number of cache directories
-to 16, as SHA-1 hashes use the character set {[}\emph{0-9,a-f}]. In
-the extremely unlikely event of a collision between two URLs having
-the same SHA-1 hash, caching will not be used for the second file.
-
-There is no indexing system for the cache, and any cache filename
-can be easily determined from a URL by using the same hashing algorithm
-as the A-REX, e.g. the standard command line tool \emph{sha1sum. }The
-cache directory used by a file can be determined from the initial
-letter of the hash and the order of cache directories in the configuration
-file. Some associated metadata (the corresponding URL and an expiry
-time, if available) are stored in a file with the same name as the
+Cached files are stored in sub-directories under the \emph{data}
+directory in each main cache directory. Filenames are constructed from
+an SHA-1 hash of the URL of the file and split into subdirectories
+based on the two initial characters of the hash. In the extremely
+unlikely event of a collision between two URLs having the same SHA-1
+hash, caching will not be used for the second file.
+
+When multiple caches are used, a new cache file goes to a randomly
+selected cache, where each cache is weighted according to the size of
+the file system on which it is located.  For example: if there are two
+caches of 1TB and 9TB then on average 10\% of input files will go to
+the first cache and 90\% will go to the second cache.
+
+Some associated metadata including the corresponding URL and an expiry
+time, if available, are stored in a file with the same name as the
 cache file, with a \emph{.meta} suffix.
@@
 containing the cache is not accessible from worker nodes. The latter
-option is the default option. The per-job directory is only readable
-by the local user running the job, and the cache directory is readable
-only by the A-REX user. This means that the local user cannot access
-any other users' cache files. It also means that cache files can be
-removed without needing to know whether they are in use by a currently
-running job. Note that when running A-REX under a non-privileged user
-account, all cache files will be owned and accessible by the same
-user, and therefore modifyable by running jobs. This is potentially
-dangerous and so cacheing should be used with caution in this case.
+option is the default option.  Files marked as executable in the job
+will be stored in the cache without executable permissions, but they
+will be copied to the SD and the appropriate permissions applied to
+the copy.
+
+The per-job directory is only readable by the local user running the
+job, and the cache directory is readable only by the A-REX user. This
+means that the local user cannot access any other users' cache
+files. It also means that cache files can be removed without needing
+to know whether they are in use by a currently running
+job. \textbf{IMPORTANT:} If a cache is mounted from an NFS server and
+the GM is run by the root user, the server must have the
+\emph{no\_root\_squash} option set for the GM host in the
+\emph{/etc/exports} file, otherwise the GM will not be able to create
+the required directories. Note that when running A-REX under a
+non-privileged user account, all cache files will be owned and
+accessible by the same user, and therefore modifyable by running
+jobs. This is potentially dangerous and so cacheing should be used
+with caution in this case.
 
@@
 
+\subsection{Remote Caches}
+\label{sec:remotecaches}
+
+If a site has multiple A-REXs running, an A-REX can be configured to
+have its own caches and have read-only access to caches under the
+control of other A-REXs (remote caches). An efficient way to reduce
+network traffic within a site is to configure A-REXs to be under
+control of caches on their local disks and have caches on other hosts
+as remote caches. If an A-REX wishes to cache a file and it is not
+available on the local cache, it searches for the file in remote
+caches. If the file is found in a remote cache, the actions the A-REX
+takes depends on the policy for the remote cache. The file may be
+replicated to the local cache to decrease the load on the remote file
+system caused by many jobs accessing the file. However, this will
+decrease the total number of cache files that can be stored. The other
+policy is to use the file in the remote cache, creating a per-job
+directory for the hard link in the remote cache. Then the link is
+created from the session dir to that directory, bypassing the local
+cache completely. The usual permission and validity checks are
+performed for the remote file. Note that no creation or deletion of
+remote cache data is done - cache cleaning is only performed on local
+caches.
+
 \subsection{Cache Administration}
@@
 hosting the file system.
+
+Caches can be added to and removed from the configuration as required
+without affecting any cached data, but after changing the
+configuration file, the A-REX should be restarted. If a cache is to be
+removed and all data erased, it is recommended that the cache be put
+in a \emph{draining} state until all currently running jobs possibly
+accessing files in this cache have finished. In this state the cache
+will not be used by any new jobs, but the hard links in the
+\emph{joblinks} directory will be cleaned up as each job
+finishes. Once this directory is empty it is safe to delete the entire
+cache. See the \emph{cachedir} option in
+Section~\ref{SubSection: ConfigFile} for how to set a cache to a
+draining state.
 
@@
   default sessiondir is used -- \textit{\$HOME/.jobs}.
-\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} -- specifies
-a directory to store cached data (see section \ref{sec:cache}). Multiple
-cache directories may be specified by specifying multiple \emph{cachedir}
-commands. Cached files will be distributed evenly between the caches.
-Specifying no \emph{cachedir }command or commands with an empty path
-disables caching. Optional \textit{link\_path} specifies the path
-at which \emph{path} is accessible on computing nodes, if it is different
-from the path on the A-REX host. If \textit{link\_path} is set to '.'
-files are not soft-linked, nor are per-job links created, but files
-are copied to the session directory.
+\item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} --
+  specifies a directory to store cached data (see section
+  \ref{sec:cache}). Multiple cache directories may be specified by
+  specifying multiple \emph{cachedir} commands. Cached data will be
+  distributed over multiple caches according to free space in
+  each. Specifying no \emph{cachedir} command or commands with an
+  empty path disables caching. The optional \textit{link\_path} specifies
+  the path at which \emph{path} is accessible on computing nodes, if
+  it is different from the path on the A-REX host. If
+  \textit{link\_path} is set to '.'  files are not soft-linked, nor
+  are per-job links created, but files are copied to the session
+  directory. If a cache directory needs to be drained, then
+  \emph{cachedir} should specify \emph{``drain''} as the
+  \emph{link\_path}.
+\item \textbf{\textit{remotecachedir}}\textit{=path {[}link\_path]} -
+  specifies caches which are under the control of other A-REXs, but which
+  this A-REX can have read-only access to (see
+  Section~\ref{sec:remotecaches}).  Multiple remote cache directories may be
+  specified by specifying multiple \emph{remotecachedir} commands. If
+  a file is not available in paths specified by cachedir, the GM looks
+  in remote caches. \emph{link\_path} has the same meaning as in
+  \emph{cachedir}, but the special path ``replicate'' means files will
+  be replicated from remote caches to local caches when they are requested.
 \item \textbf{\textit{cachesize}}\textit{=high\_mark {[}low\_mark]} --
@@
 \url{http://www.nordugrid.org/documents/voms-notes.html}.
+
+When XML file only is used to configure the A-REX, the transfer shares
+can be implemented by defining \emph{maxLoadShare} (the limit itself)
+and \emph{loadShareType} (the scheme used) elements inside 
+\emph{loadLimits} block.
 

### 2009-12-23 doc/tech_doc/xrsl/xrsl-urls.tex
put some info about the ARC DMC - but I am not really sure if this is the good place for it or not
new httpg protocol logic

@@
    \texttt{file} & local to the host file name with a full path\\
+   \texttt{arc} & for the Chelonia storage service, communicates with Bartenders, the path should be a Logical Name (LN)\\
 \end{tabular}
@@
 \end{shaded}
+
+For Chelonia, the syntax is
+\begin{shaded}
+    \verb#arc://<LogicalName>[?BartenderURL=<URL>]#
+\end{shaded}
+where the BartenderURL could come from the `bartender' parameter of the client configuration file.
 
@@
    functions to wrap data and \verb#gsi# uses additional headers as 
-   implmented in Globus IO\\
+   implmented in Globus IO. The default is \verb#gssapi#. In case
+   this fails, \verb#gsi# is then tried.\\
    \verb#spacetoken=<pattern># & specify the space token to be used for

### 2010-01-14 src/services/a-rex/grid-manager/conf/conf_cache.cpp
use common logging package in cache-clean, and add conf option to specify log level. Partially fixes bug 1383

@@ -3,7 +3,8 @@
 #include "conf_cache.h"
 
 CacheConfig::CacheConfig(std::string username): _cache_max(100),
-                                                _cache_min(100) {
+                                                _cache_min(100),
+                                                _log_level("INFO") {
   // open conf file
   std::ifstream cfile;
   if(nordugrid_config_loc().empty()) read_env_vars(true);
@@ -158,6 +159,31 @@ void CacheConfig::parseINIConf(std::string username, ConfigSections* cf) {
         throw CacheConfigException("max cache size must be greater than min size");
       _cache_min = min_i;
     }
+    else if(command == "cacheloglevel") {
+      std::string log_level = config_next_arg(rest);
+      if(log_level.length() == 0)
+        throw CacheConfigException("No value specified in cacheloglevel");
+      off_t level_i;
+      if(!Arc::stringto(log_level, level_i))
+        throw CacheConfigException("bad number in cacheloglevel parameter");
+      // manual conversion from int to log level
+      switch (level_i) {
+        case 0: { _log_level = "FATAL"; };
+          break;
+        case 1: { _log_level = "ERROR"; };
+          break;
+        case 2: { _log_level = "WARNING"; };
+          break;
+        case 3: { _log_level = "INFO"; };
+          break;
+        case 4: { _log_level = "VERBOSE"; };
+          break;
+        case 5: { _log_level = "DEBUG"; };
+          break;
+        default: { _log_level = "INFO"; };
+          break;
+      } 
+    }
     else if(command == "control") {
       // if the user specified here matches the one given, exit the loop
       config_next_arg(rest);
@@ -196,6 +222,7 @@ void CacheConfig::parseXMLConf(std::string username, Arc::XMLNode cfg) {
         link
       highWatermark
       lowWatermark
+      cacheLogLevel
     defaultTTL
     defaultTTR
     maxReruns
@@ -281,6 +308,9 @@ void CacheConfig::parseXMLConf(std::string username, Arc::XMLNode cfg) {
           throw CacheConfigException("highWatermark must be greater than lowWatermark");
         _cache_min = min_i;
       }
+      std::string cache_log_level = cache_node["cacheLogLevel"];
+      if (!cache_log_level.empty())
+        _log_level = cache_log_level;
       Arc::XMLNode remote_location_node = cache_node["remotelocation"];
       for(;remote_location_node;++remote_location_node) {
         std::string cache_dir = remote_location_node["path"];

### 2010-01-14 src/services/a-rex/grid-manager/conf/conf_cache.h

@@ -44,6 +44,10 @@ class CacheConfig {
     * Cache directories that are needed to be drained
     **/
    std::vector<std::string> _draining_cache_dirs;
+   /**
+    * cache-clean log level
+    */
+   std::string _log_level;
  public:
    /**
     * Create a new CacheConfig instance. Read the config file and fill in
@@ -68,6 +72,7 @@ class CacheConfig {
   int getCacheMax() { return _cache_max; };
   int getCacheMin() { return _cache_min; };
   bool cleanCache() { return _cache_max < 100; };
+  std::string getLogLevel() { return _log_level; };
 };
 
 #endif /*__GM_CONFIG_CACHE_H__*/

### 2010-01-25 doc/tech_doc/a-rex/arex_tech_doc.tex
use common logging package in cache-clean, and add conf option to specify log level. Partially fixes bug 1383
dont delete cache files that have hard links (bug 1748)

@@
 files. It also means that cache files can be removed without needing
-to know whether they are in use by a currently running
-job. \textbf{IMPORTANT:} If a cache is mounted from an NFS server and
-the GM is run by the root user, the server must have the
-\emph{no\_root\_squash} option set for the GM host in the
-\emph{/etc/exports} file, otherwise the GM will not be able to create
-the required directories. Note that when running A-REX under a
+to know whether they are in use by a currently running job. However,
+as deleting a file which has hard links does not free space on the
+disk, cache files are not deleted until all per-job hard links are
+deleted. \textbf{IMPORTANT:} If a cache is mounted from an NFS server
+and the A-REX is run by the root user, the server must have the
+\emph{no\_root\_squash} option set for the A-REX host in the
+\emph{/etc/exports} file, otherwise the A-REX will not be able to
+create the required directories. Note that when running A-REX under a
 non-privileged user account, all cache files will be owned and
@@
 cache. See the \emph{cachedir} option in
-Section~\ref{SubSection: ConfigFile} for how to set a cache to a
+Section~\ref{SubSection:ConfigFile} for how to set a cache to a
 draining state.
@@
   specified by specifying multiple \emph{remotecachedir} commands. If
-  a file is not available in paths specified by cachedir, the GM looks
+  a file is not available in paths specified by cachedir, the A-REX looks
   in remote caches. \emph{link\_path} has the same meaning as in
@@
   \emph{cachedir} commands.
+\item \textbf{\textit{cacheloglevel}}\textit{=number} -- specified the
+  level of logging by the \emph{cache-clean} tool, between 0 (FATAL)
+  and 5 (DEBUG). Defaults to 3 (INFO).
 \item \textbf{\textit{maxrerun}}\textit{=number} -- specifies maximal number

### 2010-01-29 doc/manuals/configuration/configuration_manual.tex
Rewriting service configuration documentation. Not complete yet. Please come with feedback.

@@
 \usepackage{longtable}                             %for breaking tables
+\usepackage{array}
+\usepackage{booktabs}
 %
@@
 \hypersetup{
-  pdfauthor = {Martin Savko},
-  pdftitle = {ARC configuration manual},
+  pdfauthor = {Martin Skou Andersen},
+  pdftitle = {WS-ARC service configuration manual},
   pdfsubject = {Paper subject},
@@
 \definecolor{cyan}{rgb}{0,1,1}
+\definecolor{grey}{rgb}{0.9,0.9,0.9}
+
+\usepackage{listings}
+ \lstset{
+         basicstyle=\small\ttfamily,
+         numbers=none,
+         numberstyle=\tiny,
+         stepnumber=1,
+         numbersep=5pt,
+         tabsize=2,
+         extendedchars=true,
+         breaklines=true,
+         keywordstyle=\color{red},
+ %        keywordstyle=[1]\textbf,
+ %        keywordstyle=[2]\textbf,
+ %        keywordstyle=[3]\textbf,
+ %        keywordstyle=[4]\textbf,   \sqrt{\sqrt{}} %
+         stringstyle=\color{blue}\ttfamily,
+         showspaces=false,
+         showtabs=false,
+         xleftmargin=17pt,
+         frame=single,
+         framexleftmargin=17pt,
+         framexrightmargin=5pt,
+         framexbottommargin=4pt,
+         framextopmargin=4pt,
+         backgroundcolor=\color{grey},
+         showstringspaces=false
+ }
+ \lstloadlanguages{
+  sh,
+  XML
+ }
+
+% Make lineshift after paragrah...
+\makeatletter
+\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
+  {-3.25ex\@plus -1ex \@minus -.2ex}%
+  {1.5ex \@plus .2ex}%
+  {\normalfont\normalsize\bfseries}}
+\makeatother
+
  %
@@
 %%%%---- The title ----
-{\centering \textsc{\Large ARC configuration manual}\Large \par}
+{\centering \textsc{\Large WS-ARC service configuration manual}\Large \par}
 \vspace*{0.5cm}
@@
 %%%%---- A list of authors ----
-    {\centering \large Zsombor Nagy\footnote{zsombor@niif.hu} \large \par}
-    {\centering \large Ferenc Szalai\footnote{szferi@wsbricks.com} \large \par}
-    {\centering \large Martin Savko\footnote{martin.savko@gmail.com} \large \par}
+    {\centering \large Martin Skou Andersen\footnote{skou@nbi.dk} \large \par}
 \end{titlepage}
@@
 
-The arched daemon's configuration file contains general information about the daemon's environment (pid file, log file, log level, library paths, plugins), and it contains one (or more) message chains, which specifies how the different services are connected to the outside world.
-The daemon's listening interface, certificates used, and the HTTP path of the services are among the things specified in the chain.
-This part contains also the configuration specifics for each service.
+The Web Service (WS) Advanced Resource Connector (ARC) provides a great deal of
+services which all share a common configuration interface. One or more services
+can be specified in a configuration and the \texttt{arched} daemon should then
+be used for starting the service(s). The configuration should be
+specified in a file using either the relatively simple INI format, or the more
+complex XML format.
 
-A chain could contain multiple types of components: message chain components (called MCCs) which provide TCP, TLS, HTTP and SOAP support; the plexer, which connects different HTTP paths to different services; and the services themselves with all their specific configuration details.
+INI configuration is the recommended way of configuring services for people who
+are new to ARC or unfamiliar with WS-ARC configuration. However it is limited in
+the sense that it relies on profiles. A profile is a configuration
+template setup for one or more services, and common use case profiles exist for
+most services and are installed per default. The limitation lies in the fact
+that only services and options which are specified in the profile can be
+configured in INI configuration, however the installed profiles should give
+enough flexibility for common use cases. For a detailed description on INI
+configuration see section~\ref{sec:INI configuration}. Profiles are further
+discussed in section~\ref{sec:Profiles}.
 
-The configuration file is in XML format, with XML schemas and namespaces given and documented.
+XML configuration is the core format of WS-ARC configuration -- it
+is used internally -- and thus it gives full flexibility when configuring
+services. A detailed description of XML configuration can be found in
+section~\ref{sec:XML configuration}.
 
-For typical deployment scenarios there is an additional mechanism to create an XML configuration on-the-fly based on XML templates (called “profiles”) and more administrator-friendly INI configuration files.
+The \texttt{arched} daemon, which is used to start WS-ARC services according to
+the configuration is briefly discussed in section~\ref{sec:arched}.
 
-Configuration related arched command line switches:
+\subsection{How to read this manual}
+This manual is divided into the three independent sections INI configuration,
+XML configuration and Profiles, and these can be read independent of each other.
+Additionally the \texttt{arched} section describe the command with the same
+name.
 
-\begin{verbatim}
-	-c <xml config file>
-	-i <ini config file>
-	-d
-	-s <xml schema>
-\end{verbatim}
+It should be noted that words typeset with the \texttt{typewriter} font should
+be typed as written.
 
-By default the daemon looks for the configuration files in the following order: first /etc/arc/server.ini, then if it is not found: /etc/arc/server.xml.
-However user can define which configuration  should be used with the -c or -i command line argument.
-If accidentally both of them is defined then the -i has greater priority.
+\section{INI configuration}\label{sec:INI configuration}
+As mentioned in the introduction, INI configuration directly relies on profiles,
+and a profile is a configuration template setup for one or more services. So
+when creating a INI configuration file a choice of profile must be made. Only
+one profile can be used in a given INI configuration. A standard WS-ARC
+installation comes with a set of profiles and the list in
+appendix~\ref{app:List of profiles} gives a short description of these. The
+profiles will per default be installed into the \texttt{/usr/share/arc/profiles}
+folder, where it is also possible to find an example INI configuration file for
+each of the profiles.
 
-In order the check the generated XML configuration the -d options can be used. If used, arched will  print the XML configuration to the standard output and then stop.
+A INI configuration file basically consist of sections, attribute-value pairs
+and maybe comments and blank lines. In the following code-listing the syntax of
+an INI configuration file is shown:
+\begin{lstlisting}
+# A comment
+[ common ]
+profile = <path-to-profile>
 
-\section{XML level configuration}
+<attribute> = <value>
+...
+...
 
-\begin{itemize}
-		\item The first part of the config is the “Server” part, which contains the path of the pid file, and the path of the log file, and the verbosity of the logging.
-		\item The second part of the config is the “ModuleManager” part, which contains the path to the directory with the ARC libraries.
-		\item The third part is the “Plugins” part with the list of 'plugins' (e.g. message chain components, services) used in this config.
-		\item The fourth part is the “Chain” part, which contains the description of the message chain, which connects the services to the network.
-(It is possible to have more than one chain in one configuration.) A chain could contain the message chain components, the plexer and the services.
-		\begin{itemize}
-			\item The TCP message chain component does the listening on the network interface. This is the place to set the IP version to use, the IP and port to listen to.
-			\item The TLS message chain component implements the security layer. This is the place to set the server-side key and the certificate, and the directory of the trusted CA certificates.
-			\item The configuration of the HTTP and the SOAP message components most probably don't need to be changed.
-			\item The Plexer component has a list of all available services, and has a HTTP path for each of them.
-The path is a regular expression, the incoming HTTP requests will be sent to the service which has a matching path in the Plexer config.
-			\item Then come the specific configurations of all the services which will be running in this arched.
-		\end{itemize}
-\end{itemize}
+# Another comment
+[ <section-name> ]
+<attribute> = <value>
+...
+...
+[ <another-section> ]
+<attribute> = <value>
+...
+...
+\end{lstlisting}
 
-\subsection{The service only gets its own config}
+A section is specified by putting the name of the section inside brackets
+(\texttt{[} and \texttt{]}), where the left bracket (\texttt{[}) must be the
+first non white space character on the line. Any white space character after the
+left bracket and before the right will not be part of the section name. Every
+attribute-value pair belongs to the section which has be been defined
+previously. If the file starts without declaring a section, then following
+attributes will belong to the common section. Multiple sections with identical
+names are different sections, except for the common section, which is special in
+this respect and if multiple common sections exist these will be considered as
+one.
 
-It is important to know that the services will not be able to get the whole configuration, they will only get their part, so everything a service should know have to be specified in its configuration part, e.g.
-a service would not know the IP or the port of the server, the HTTP path to itself or the path of the key and certificate file unless these are specified in the service configuration part too.
+The \texttt{profile} attribute is also a special, as it is used to indicate
+which profile the INI configuration uses, and its value should be the
+absolute path to the profile file. The \texttt{profile} attribute must be
+located in the common section. If multiple \texttt{profile} attributes exist
+only the first attribute will be used.
 
-Because some services would establish some network connection while running (e.g.
-to send SOAP messages to other services), and these connections would most probably be secure HTTPS connection, these services may have a part where a client-side key and certificate file have to be specified which can be used to establish connection.
+Attribute-value pairs are defined by specifying the name of the attribute
+followed by an equal sign and then the value the attribute should take. It is
+the first equal sign which separates the attribute from the value. If a line,
+not defining a section, does not contain an equal then it will be ignored and if
+it contains multiple equal signs then all but first equal sign will be part of
+the value. Any white space before and after the attribute and value will be
+ignored. Only attributes which have been defined in the selected profile will be
+consideren by \texttt{arched}. For each profile in WS-ARC a corresponding INI
+configuration file exist which defines sensible values and at the time list all
+configurable attributes. When it comes to multiple attributes with identical
+names defined in the same section, it is the selected profile which determines
+if multiple attributes is allowed and will be interpreted or not. It is possible
+that attributes from one section can override attributes from another section,
+however this behaviour is completely determined by the profile.
 
-A service most probably would want to advertise its URL by registering itself to an information indexing service (called ISIS), in order to do this, the full URL of the service should be put into the configuration part of the service.
-This means that if the daemon's interface (port or hostname) would be changed, all these URLs should be modified too.
+Comments in the INI configuration file are specified by letting the first non
+white space character be the sharp character (\texttt{\#}), and these will be
+ignored by \texttt{arched}.
 
-\subsection{Summary of where to change things}
-\begin{itemize}
-	\item pidfile, logfile, loglevel: the “Server” part
-	\item path to the ARC libraries/modules/plugins: the “ModuleManager” part
-	\item listening interface (port, IP, IP version): the TCP Component in the “Chain” part; and maybe the URLs of the services should be changed also in the configuration of the services.
-	\item path to the server-side key and certificate files and CA dir: the TLS Component in the “Chain” part
-	\item path to the client-side key and certificate files and CA dir: in the configuration of the individual services in the “Chain” part
-	\item the HTTP path of the services: the Plexer in the “Chain” part, the paths are specified in regular expressions; and maybe the URL of the service should be changed also in the configuration of the service itself.
-	\item the URL of the information indexing service (ISIS) where a service should advertise itself: in the config of each services in the “Chain” part
-\end{itemize}
+\subsection{Examples}
+\subsubsection{Echo service}
 
-There are documentations and XML schemas for the possible configuration options of the message chain components (MCCs) and for all the services (TODO: link to some docs and stuff)
+\begin{lstlisting}
+profile = /usr/share/arc/profiles/EchoService.xml
 
-\section{INI style configuration}
+pidfile = /tmp/arched.pid
+logfile = /tmp/arched.log
+loglevel = INFO
 
-Because the parameters which a system administrator would want to change while configuring arched are spread across the whole XML configuration file and sometimes even repeated several times, and because there are lots of things in the XML configuration file which would be always the same for a given deployment-scenario, it is possible to create different configuration profiles for different scenarios, and then configure these profiles with configuration files which are in a more user-friendly INI format containing key-value pairs grouped in sections.
-Then the XML config is generated from the given profile and the INI config.
-(A profile is basically a full XML configuration, but has some extra details about which part of the XML should be overwritten with which INI configuration parameter.)
+port = 60000
 
-For example there is a profile for a secure Echo service which contains a working set of message chain components and the Echo service with reasonable defaults for the paths of libraries, logfile, pidfile, certificates, the port to listen, the HTTP path of the service, etc.
-Then the system administrator can create an INI config which only contains the things which need to be non-default, e.g.
-if the key and certificate files are in a non-standard directory, the correct paths could be specified in the INI config, and these paths then would be put into the profile everywhere where needed (even multiple times).
+cacert = /etc/grid-security/certificates
+host_cert = /etc/grid-security/hostcert.pem
+host_key = /etc/grid-security/hostkey.pem
 
-Or if there is a profile with several services, all of them registering to an information index service (ISIS), which means that each service has the URL of the ISIS service in its config, this URL could be changed by adding one line to the INI config with the new URL, which makes this URL inserted into the config of each services.
+[ echo ]
+prefix = {{
+suffix = }}
+\end{lstlisting}
 
-\subsection{Issues}
-
-Using the INI config is much easier for the system administrator than figuring out where to modify the XML config in order to change some paths or URLs.
-However using the INI config is much less flexible then using the XML config.
-Everything depends on the profiles.
-If the system administrator would need a specific set of services, and there is no profile for that, then the INI config can't be used.
-Or if there is profile but a given configuration parameter is not given in the profile, then it cannot be set in the INI config.
-
-Another issue is that for each individual profile a separate documentation would be needed, because a profile can use arbitrary INI tags.
-e.g.
-the INI tag for setting the log level could be “loglevel” in one profile and it could be “debug” in another, or something completely different in the third one.
-The way to set the path to the directory of the trusted CAs' certificates could be “cacert=path” or “cacertificatesdir=path” depending which profile is in use.
-Currently there are example INI configs for most of the existing profiles, which should give hints about the possible INI tags.
-
-\subsection{INI section}
-
-The key-value pairs in an INI file are grouped into sections, like this:
-
-\begin{verbatim}
-	key1=value1
-	key2=value2
-
-	[section1]
-	key1=value11
-	key3=value3
-
-	[section2]
-	key5=value5
-\end{verbatim}
-
-The key-value pairs before the first section name are in the “common” section.
-The sections are used to selectively override arguments.
-e.g.
-the profile could say that it needs the value for “key1”, and then it could specify that it needs the value from the “common” section only, or from the |”section1” section only, or it could say that it needs the value from either of these two sections and the value from “section1” overrides the value from the “common” section, or vice versa.
-
-For a more specific example, if there is a client-side key file given in the common section, then the profile could specify for each service, that it should use this key file unless there is a different key file specified in the section of the given service.
-This means that if there is an INI section for the service, and it contains a path to the key file, then that path will be used, but if it does not contain a path, then the common key path will be used from the “common” section.
-
-\subsection{Profiles}
-
-The profile is a full XML service configuration with some additional XML attributes about which part of the XML should be overwritten by which INI parameter. Let's look at this part of a profile:
-
-\begin{verbatim}
-	<infosys:URL inisections='register ahash' initag=\verb!index_server1!/>
-\end{verbatim}
-
-The initag attribute tells the key of the parameter, the inisections attribute specifies a list of sections with the last one having the highest priority.
-This XML tag here is empty, which means that it's default value is an empty string.
-
-\begin{itemize}
-	\item If the INI contains a key \verb!index_server1! somewhere in the 'register' or 'ahash' sections, then the value from that key would be put into this XML node.
-	\item If there is an \verb!index_server1! key in both the 'register' and the 'ahash' section, then the value from the 'ahash' section would be used, it has the higher priority.
-	\item If there is no \verb!index_server1! key in none of these two sections in the INI config, then this 'infosys:URL' tag will still be in the generated XML config, but with an empty content.
-\end{itemize}
-
-This part has a non-empty default:
-
-\begin{verbatim}
-	<cfg:File inisections='common' initag='logfile'>/var/log/arched.log</cfg:File>
-\end{verbatim}
-
-This means if that would be a 'logfile' key in the 'common' section of the INI configuration file, then this default value will be overwritten by the value of that key from the INI file.
-If there is no key called 'logfile' in the 'common' section of the INI file, then the generated XML config will contain the given default value.
+\subsubsection{Other service}
 
 \subsection{Limitations}
+Configuring WS-ARC services in INI format may pose some limitations, since this
+configuration format depend on profiles. If one wants to configure a service
+setup which is not covered by any of the available profiles, it is simply not
+possible to configure that service setup in INI format and one has to use the
+XML format described in section~\ref{sec:XML configuration}. It should however
+be noted that it is easy to switch from INI to XML format since the
+\texttt{arched} command provide a way to convert a INI configuration into XML.
+See section~\ref{sec:arched} for further details.
 
-Currently overwriting XML attributes are not supported. e.g. the id attribute of:
+\section{XML configuration}\label{sec:XML configuration}
+\subsection{Overview}
+XML configuration in WS-ARC provides full flexibility for configuring services,
+and to construct an adequate configuration one need to address the following
+subjects. If not done already, a choice on which service(s) being setup need to
+be made and a short discussion of the available WS-ARC services is given in
+section~\ref{sec:Service configuration}. Upon the choice of service(s) the
+proper message chain need to be configured and a discussion about message chains
+is given in section~\ref{sec:Message Chain}. Most people also want some kind of
+security setup capable of allowing/denying certain users, which can be done with
+security handlers and is discussed in section~\ref{sec:Security Handlers}.
+Common configuration options like specifying plugins required for instance by
+the message chain, and configuring logging, is discussed in
+section~\ref{sec:General Daemon Configuration}. In
+section~\ref{sec:XML examples} example XML configuration files are given.
 
-\begin{verbatim}
-	<next id='tls'/>
-\end{verbatim}
+\subsection{Service configuration}\label{sec:Service configuration}
+The following services are part of WS-ARC and specific configuration details for
+these are given in the respective references below:
+\begin{description}
+\item[A-REX]
+  A-REX
+\item[Chelonia]
+\item[ISIS]
+\item[Charon]
+\item[hopi]
+  hopi, a lightweight HTTP webserver, is described in details in \%\% REFERENCE.
 
-can not be overwritten by the INI config.
+\end{description}
 
-Currently there is no support for XML tags with variable number of occurrences. This means that if there would be a part in the XML config like this:
+\subsection{Message Chain}\label{sec:Message Chain}
+To configure WS-ARC services, a basic knowledge about how these are
+communicating with clients or other services is needed. WS-ARC uses a chain of
+message components (MCCs) which serves different levels of functionality. In one
+end it accepts messages from an interface and in the other a service processes
+them and produce a response. Two types of MCCs exist, and they differ in that
+the first type should be able to listen on some interface for messages, while
+the other simply
+receives messages from another MCC. The MCC then process the message and pass it
+on to a single or multiple other MCCs or services. MCCs should be configured
+using the XML element \texttt{Component}, where the \texttt{name} attribute
+specifies the MCC to use. The following MCCs is part of WS-ARC:
+\begin{description}
+\item[\texttt{tcp.service}] Listens on the TCP interface, see
+  section~\ref{sec:tcp-mcc},
+\item[\texttt{tls.service}] Adds encryption layer to communication, see
+  section~\ref{sec:tls-mcc},
+\item[\texttt{http.service}] Processes HTTP messages, see
+  section~\ref{sec:http-mcc},
+%~ \item[\texttt{gsi.service}] Processes GSI messages, see
+  %~ section~\ref{sec:gsi-mcc},
+\item[\texttt{soap.service}] Processes SOAP messages, see
+  section~\ref{sec:soap-mcc}.
+\end{description}
+Additionally a special MCC,
+the Plexer, exist which is able to pass messages to different components simply
+based on the endpoint the messages was sent to, and it should should be
+configured using the \texttt{Plexer} element, see section~\ref{sec:plexer}. In
+the end of the chain a service should be specified, using the  \texttt{Service}
+element, where the \texttt{name} attribute indicates the given service.
 
-\begin{verbatim}
-	<ISISURL>the URL of an ISIS service</ISISURL>
-	<ISISURL>the URL of another ISIS service</ISISURL>
-	<ISISURL>the URL of a third ISIS service</ISISURL>
-\end{verbatim}
+A working configuration should then link a set of MCCs and services, starting
+with a MCC that receives messages from some interface. The MCCs and services
+should be uniquely identifiable, by setting the \texttt{id} attributes of these
+elements. A MCC links to another or several MCCs or services using the
+\texttt{next} element, where the \texttt{id} attribute specifies the \texttt{id}
+of that MCC or service. The message chain should be contained in the root
+element \texttt{ArcConfig}, which is shown in the conceptual configuration
+example below:
+\begin{lstlisting}[language=xml]
+...
+<Component name="mcc1.service" id="mcc1">
+  <next id="mcc2"/>
+</Component>
+<Component name="mcc2.service" id="mcc2">
+  <next id="service"/>
+</Component>
+<Service name="service" id="service"/>
+...
+\end{lstlisting}
+For readability the message chain can be grouped into a \texttt{Chain} element
+in one or more levels.
 
-It is not possible to control the number of ISISURL tags from the INI config.
-If there are three such lines in the profile, then there will be three lines in the generated XML config, no matter how many URLs are specified in the INI config.
-A workaround could be using empty defaults, like this:
+\subsubsection{TCP MCC}\label{sec:tcp-mcc}
+The TCP message component is named \texttt{tcp.service} and should be the first
+component in a chain. It is used to setup a server socket which will listen for
+incoming connections. The socket can be configured using the \texttt{Listen}
+element, which can occur multiple times and under each element the following
+child elements can be specified:
+\begin{description}
+\item[\texttt{Interface}] This element specifies the interface which socket
+  should listen on. If this element is not specified then the socket will listen
+  on all available interfaces.
+\item[\texttt{Port}] The port which the socket should bind to should be
+  specified with this element. If no port is specified an error will be reported
+  and this \texttt{Listen} element will be ignored.
+\item[\texttt{Version}] The IP version can be specified by this element, and by
+  default both version 4 and 6 will be used. It is possible to specify version 4
+  or 6 only.
+\item[\texttt{NoDelay}] By default Nagle's algorithm is used for sending data,
+  however if the \texttt{NoDelay} element is set to \texttt{true} data will be
+  send when ready.
+\item[\texttt{Timeout}] Specifies how much time, in seconds, there should be
+  used on waiting for a client response. If no client response is obtained
+  before timeout the connection is dropped. The default value is 60 seconds.
+\end{description}
+It is also possible to put a limit on incoming connections, and how these should
+be handled if the limit is reached. The limit can only be put on all configured
+interfaces, and not per interface. The \texttt{Limit} element specifies the
+limit on incoming connections, and the \texttt{drop} attribute specifies whether
+new connections, above the limit, will be put on hold (default) or if they
+should be dropped, which is specified by setting the attribute to \texttt{true}.
 
-\begin{verbatim}
-	<lib:ISISURL inisections='register librarian' initag=\verb!index_server1!/>
-	<lib:ISISURL inisections='register librarian' initag='index_server2'/>
-	<lib:ISISURL inisections='register librarian' initag='index_server3'/>
-\end{verbatim}
+An example of a TCP MCC configuration using the described elements is shown
+below:
+\begin{lstlisting}[language=xml]
+...
+<Component name="tcp.service" id="tcp">
+  <Listen>
+    <Interface>w.x.y.z</Interface>
+    <Port>44444</Port>
+    <Version>4</Version>
+    <NoDelay>true</NoDelay>
+    <Timeout>120</Timeout>
+  </Listen>
+  <Listen>
+    <Interface>a:b:c:d:e:f:g:h</Interface>
+    <Port>55555</Port>
+    <Version>6</Version>
+    <NoDelay>false</NoDelay>
+    <Timeout>20</Timeout>
+  </Listen>
+  <Limit drop="true">30</Limit>
+</Component>
+...
+\end{lstlisting}
 
-Then the INI config could be like this:
+\subsubsection{TLS MCC}\label{sec:tls-mcc}
+A security layer can be configured with the TLS MCC, named \texttt{tls.service},
+and it should be put on top of a TCP MCC. By default it is able to understand
+the SSLv2, SSLv3 and TLSv1 protocols, however the \texttt{Handshake} element can
+be used to specify another behaviour. The supported values for the
+\texttt{Handshake} element are \texttt{TLS} (default) and \texttt{SSLv3}, where
+specifying the latter will configure a MCC which only understand the SSLv3
+protocol. If other values than the mentioned ones are given, then \texttt{TLS}
+will be used.
 
-\begin{verbatim}
-	[register]
-	index_server1=the URL of an ISIS service
-	index_server2=the URL of another ISIS service
-\end{verbatim}
+For the TLS MCC to function, it should have access to a host key-certificate
+pair. The key should be the private one, and it should not be password
+protected. The path to the key-certificate pair can be specified by using the
+\texttt{KeyPath} and \texttt{CertificatePath} elements. If the \texttt{KeyPath}
+element is not specified the default path
+\texttt{/etc/grid-security/hostkey.pem} will be used, and likewise if the
+\texttt{CertificatePath} element is not specified the default path
+\texttt{/etc/grid-security/hostcert.pem} will be used. The supported key and
+certificate formats are PEM and DER.
 
-Which means that the first and second ISISURL line would be filled with data from the INI, and the third one would be left empty.
-For this to work, it should be supported by the service itself, e.g.
-the service should throw out empty ISISURLs.
+When establishing a connection with a client, the client certificate are per
+default verified against the Certificate Authority (CA) certificates, known to
+the TLS MCC. The CA certificates used for verification can be specified with
+either the \texttt{CACertificatesDir} or the \texttt{CACertificatePath} element,
+or both elements can be used. The \texttt{CACertificatesDir} element should
+point to a directory containing one or more CA certificates (in PEM format). The
+files should each contain exactly one certificate, and should be named by the
+hash value of the CA subject name. CA certificates provided by NorduGrid uses
+this option. The \texttt{CACertificatePath} element should point to a file
+containing one or more CA certificates (in PEM format) identified by:
+\begin{lstlisting}
+  -----BEGIN CERTIFICATE-----
+  ... (CA certificate in base64 encoding) ...
+  -----END CERTIFICATE-----
+\end{lstlisting}
+sequences. If both elements have been specified, CA certificates will first be
+looked up in the file pointed to by the \texttt{CACertificatePath} element and
+then in the directory pointed to by the \texttt{CACertificatesDir} element. If
+none of the two elements have been specified, then the
+\texttt{CACertificatesDir} element will be initialised to the
+\texttt{/etc/grid-security/certificates} location. Client certificate
+verification can also be skipped, which is done by setting the
+\texttt{ClientAuthn} element to \texttt{false}.
 
-Here it was needed to use different initags for the three ISISURLs, because if same initags would be used like this:
+\begin{lstlisting}[language=xml]
+...
+<Component name="tls.service" id="tls">
+  <KeyPath></KeyPath>
+  <CertificatePath></CertificatePath>
+  <CACertificatePath></CACertificatePath>
+  <CACertificatesDir></CACertificatesDir>
+  <ClientAuthn></ClientAuthn>
+  <Handshake>SSLv3</Handshake>
+</Component>
+...
+\end{lstlisting}
 
-\begin{verbatim}
-	<lib:ISISURL inisections='register librarian' initag='index_server'/>
-	<lib:ISISURL inisections='register librarian' initag='index_server'/>
-	<lib:ISISURL inisections='register librarian' initag='index_server'/>
-\end{verbatim}
+%~ \paragraph{Globus Policy}
+%~ attribute \texttt{PolicyGlobus} (CACertificatePath).
+%~
+%~ \paragraph{Trusted VOMS Credentials}
+%~ \texttt{VOMSCertTrustDNChain}
+%~ <xsd:sequence>
+  %~ \texttt{VOMSCertTrustDN}
+  %~ \texttt{VOMSCertTrustRegex}
+%~ </xsd:sequence>
+%~ The DN list of the trusted voms server credential;
+%~ in the AC part of voms proxy certificate, voms proxy
+%~ certificate comes with the server certificate which is
+%~ used to sign the AC. So when verifying the AC on the
+%~ AC-consuming side (in ARC1, it is the MCCTLS which will
+%~ consumes the AC), the server certificate will be checked
+%~ against a trusted DN list. Only if the DN and issuer's
+%~ DN of server certificate exactly matches the DN list
+%~ in the configuration under TLS component, the AC can be trusted
+%~
+%~ \texttt{VOMSCertTrustDNChainsLocation}
+%~ DN list in an external file, which is in the same format as VOMSCertTrustDNChain
+%~
+%~ \begin{lstlisting}[language=xml]
+%~ ...
+%~ <Component name="tls.service" id="tls">
+  %~ <VOMSCertTrustDNChain>
+    %~ <VOMSCertTrustDN></VOMSCertTrustDN>
+    %~ <VOMSCertTrustDN></VOMSCertTrustDN>
+    %~ <VOMSCertTrustDN></VOMSCertTrustDN>
+    %~ <VOMSCertTrustRegex></VOMSCertTrustRegex>
+  %~ </VOMSCertTrustDNChain>
+%~ </Component>
+%~ ...
+%~ \end{lstlisting}
 
-Then this INI config would cause unexpected results:
+\subsubsection{HTTP MCC}\label{sec:http-mcc}
+The HTTP protocol is supported by the HTTP MCC named \texttt{http.service}, and
+it should either be put on top of the TCP or TLS MCC. A working configuration of
+this MCC is simple, since it should only use the \texttt{next} element which
+must have a value equal to one of the HTTP request methods. An example of a HTTP
+MCC configuration is shown below.
 
-\begin{verbatim}
-	[register]
-	index_server=the URL of an ISIS service
-	index_server=the URL of another ISIS service
-\end{verbatim}
+\begin{lstlisting}[language=xml]
+...
+<Component name="http.service">
+  <next id="component1">GET</next>
+  <next id="component2">POST</next>
+  <next id="component3">PUT</next>
+</Component>
+...
+\end{lstlisting}
 
-This INI config would generate this XML config from that profile:
+%~ \subsubsection{GSI MCC}\label{sec:gsi-mcc}
+%~ \texttt{KeyPath} Location of private key.
+%~
+%~ \texttt{CertificatePath} Location of public certificate.
+%~
+%~ \texttt{ProxyPath} Location of proxy credentials - includes certificates, key
+%~ and chain of involved certificates. Overwrites elements \texttt{KeyPath} and
+%~ \texttt{CertificatePath}.
 
-\begin{verbatim}
-	<lib:ISISURL>the URL of another ISIS service</lib:ISISURL>
-	<lib:ISISURL>the URL of another ISIS service</lib:ISISURL>
-	<lib:ISISURL>the URL of another ISIS service</lib:ISISURL>
-\end{verbatim}
+\subsubsection{SOAP MCC}\label{sec:soap-mcc}
+SOAP is supported by the SOAP MCC, named \texttt{soap.service}. It parses and
+processes incoming messages as SOAP messages. This component should be put right
+after the HTTP MCC in the message chain. The component is needed by most of the
+services in WS-ARC. The only option which should be configured for the SOAP MCC
+is to which component or service the valid SOAP messages should be directed to,
+and it is done with the \texttt{next} element as for the other components.
 
-Which is the same URL three times, because the last value of the \verb!index_server! tag would be inserted into each places.
+\subsubsection{Plexer}\label{sec:plexer}
+The Plexer is a special MCC which filters messages based on their URL endpoints.
+One or more regular expressions (POSIX extended) should be defined using one or
+more \texttt{next} elements, and these will then be used to filter incoming
+messages, thus forwarding messages, where the path of the endpoint URL matched a
+given regular expression, to the component or service specified by the
+\texttt{id} attribute of the corresponding \texttt{next} element. At the first
+match the message will be forwarded to the respective component or service and
+no more expressions will be tried. If a message did not match any of the regular
+expressions then it is discarded and an error code is returned to the client. An
+example of a Plexer configuration is shown below.
+%% What is the order testing the regular expressions. First listed first tested or Last listed firsted tested?
 
-There are future plans for adding support for this kind of simple multiplicity, but there are more complex cases which are not trivial to handle.
-e.g.
-in order to register to multiple ISIS services, the path to the key and certificate files should be given for each ISIS service even though they most probably would be the same.
-This part would register to one ISIS service:
+\begin{lstlisting}[language=xml]
+...
+<Plexer>
+  <next>^/arex$</next>
+  <next>^/[eE][cC][hH][oO]$</next>
+</Plexer>
+...
+\end{lstlisting}
 
-\begin{verbatim}
-	<infosys:Registrar>
-		<infosys:URL inisections='register shepherd' initag='index_server'/>
-		<infosys:KeyPath inisections='common shepherd' initag='client_key'/>
-		<infosys:CertificatePath inisections='common shepherd' initag='client_cert'/>
-		<infosys:CACertificatesDir inisections='common shepherd' initag='cacert'/>
-	</infosys:Registrar>
-\end{verbatim}
+\subsection{Security Handlers}\label{sec:Security Handlers}
+\subsubsection{Identity Mapping}% identity.map
+\subsubsection{ARC Authorization}% arc.authz
+\subsubsection{Delegation Collector}% delegation.collector
+\subsubsection{Username Token}% usernametoken.handler
+%\subsubsection{Delegation Handler}% delegation.handler
+%\subsubsection{X509 Token}% x509token.handler
+%\subsubsection{SAML Token}% samltoken.handler
+%\subsubsection{SAML Single Sign-On}% saml2ssoassertionconsumer.handler
 
-It is possible that we want to register only to one ISIS or two or maybe three or more, then for each of them this section should be repeated. Currently the workaround looks like this:
+\subsection{General Daemon Configuration}\label{sec:General Daemon Configuration}
+The following options configures the \texttt{arched} daemon, and thus applies to
+all services. Some of the options can also be set using command line arguments,
+and if used these will override the configuration file options. See
+section~\ref{sec:arched} for a description of the \texttt{arched} daemon.
 
-\begin{verbatim}
-	<infosys:Registrar>
-		<infosys:URL inisections='register shepherd' initag=\verb!index_server1!/>
-		<infosys:KeyPath inisections='common shepherd' initag='client_key'/>
-		<infosys:CertificatePath inisections='common shepherd' initag='client_cert'/>
-		<infosys:CACertificatesDir inisections='common shepherd' initag='cacert'/>
-	</infosys:Registrar>
-	<infosys:Registrar>
-		<infosys:URL inisections='register shepherd' initag='index_server2'/>
-		<infosys:KeyPath inisections='common shepherd' initag='client_key'/>
-		<infosys:CertificatePath inisections='common shepherd' initag='client_cert'/>
-		<infosys:CACertificatesDir inisections='common shepherd' initag='cacert'/>
-	</infosys:Registrar>
-	<infosys:Registrar>
-		<infosys:URL inisections='register shepherd' initag='index_server3'/>
-		<infosys:KeyPath inisections='common shepherd' initag='client_key'/>
-		<infosys:CertificatePath inisections='common shepherd' initag='client_cert'/>
-		<infosys:CACertificatesDir inisections='common shepherd' initag='cacert'/>
-	</infosys:Registrar>
-\end{verbatim}
+Under the \texttt{Server} element the file to store the PID in, can be specified
+using the \texttt{PidFile} element, if not specified the PID will be stored in
+the file \texttt{/var/run/arched.pid}. This options can also be set using a
+command line argument. If the \texttt{Foreground} element is
+present under the \texttt{Server} element, \texttt{arched} will run in
+foreground and thus the \texttt{PidFile} element will be ignored. It is also
+posible to specify that \texttt{arched} should run in foreground on the command
+line, however it is not posible to specify that it should run as a daemon, if
+configured to run in foreground in the configuration file. Also under the
+\texttt{Server} the user and group owning the \texttt{arched} process can be
+specified, which is accomplished with the \texttt{User} and \texttt{Group}
+elements. These can also be set on the command line.
 
-Which allows to register to up to three ISIS services, and has empty defaults in order to be able to configure only one or two or zero ISIS services.
-However, the ISIS registration part should handle the empty defaults, and throw them away.
+Logging options should be specified under the \texttt{Logger} element under the
+\texttt{Server} element. The \texttt{File} element specifies which file to write
+log messages to and if not present log messages will be written to the
+\texttt{/var/log/arched.log} file. Each log message is associated with a log
+level which indicates the serverity of the message. The following log levels
+exist in WS-ARC, ordered by serverity: \texttt{FATAL}, \texttt{ERROR},
+\texttt{WARNING}, \texttt{INFO}, \texttt{VERBOSE} and \texttt{DEBUG}. With
+the \texttt{Level} element the level of messages being reported can be
+specified. Log messages with a higher or same level as specified by the
+\texttt{Level} element will be written to the log file. The default log level is
+\texttt{WARNING}. \newline
+By default the log file is not limited in size, neither
+will any log rotation be done. The \texttt{Maxsize} and \texttt{Backups}
+elements can be used to change this behaviour. The \texttt{Maxsize} element sets
+the maximum file size of the log file in bytes, while the \texttt{Backups}
+element specifies the number of files used for log rotation. When the size of
+the log file exceeds the limit specified by the \texttt{Maxsize} element the log
+file is renamed to \texttt{<log-file>.1}. If this file exist it is renamed to
+\texttt{<log-file>.2} and so forth up to the number specified by
+\texttt{Backups}, which means that the last file will be deleted. The log file
+size is only allowed to exceed the specified limit by the size of one log
+message. If the \texttt{Maxsize} element have been specified and no
+\texttt{Backups} element is specified then log rotation will not be carried out,
+thus the log file is truncated when the limit is exceeded.
 
-\section{Examples}
+Plugins which should be used later in the configuration should be specified
+under the \texttt{Plugins} element where the \texttt{Name} element specifies the
+name of the plugin. The \texttt{Path} element under the \texttt{ModuleManager}
+element specifies where to locate the plugins.
 
-The most up to date Profiles can be found in the Nordugrid SVN:
+The following example show the use of the above configuration options:
+\begin{lstlisting}[language=xml]
+<?xml version="1.0"?>
+<ArcConfig>
+  <Server>
+    <PidFile>/tmp/arched.pid</PidFile>
+    <User></User>
+    <Group></Group>
+    <Logger>
+      <File>/tmp/arched.log</File>
+      <Level>WARNING</Level>
+      <MaxSize>1000000</MaxSize>
+      <Backups>5</Backups>
+    </Logger>
+  </Server>
+  <ModuleManager>
+    <Path>/usr/lib/arc</Path>
+  </ModuleManager>
+  <Plugins>
+    <Name>plugin1</Name>
+    <Name>plugin2</Name>
+    ...
+  </Plugins>
+  ...
+</ArcConfig>
+\end{lstlisting}
 
-\verb!http://svn.nordugrid.org/trac/nordugrid/browser/arc1/branches/nox/src/hed/profiles!
+\subsection{Examples}\label{sec:XML examples}
+\subsubsection{Echo Service}
+\begin{lstlisting}[language=xml]
+<?xml version="1.0"?>
+<cfg:ArcConfig xmlns="http://www.nordugrid.org/schemas/loader/2009/08"
+               xmlns:cfg="http://www.nordugrid.org/schemas/arcconfig/2009/08"
+               xmlns:tcp="http://www.nordugrid.org/schemas/tcp/2009/08"
+               xmlns:tls="http://www.nordugrid.org/schemas/tls/2009/08"
+               xmlns:echo="http://www.nordugrid.org/schemas/echo/2009/08"
+               xmlns:infosys="http://www.nordugrid.org/schemas/infosys/2009/08">
+  <cfg:Server>
+    <cfg:PidFile>/tmp/arched.pid</cfg:PidFile>
+    <cfg:Logger>
+      <cfg:File>/var/log/arched.log</cfg:File>
+      <cfg:Level>WARNING</cfg:Level>
+    </cfg:Logger>
+  </cfg:Server>
+  <ModuleManager>
+    <Path>/usr/lib/arc</Path>
+  </ModuleManager>
+  <Plugins>
+    <Name>mcctls</Name>
+    <Name>mcchttp</Name>
+    <Name>mccsoap</Name>
+    <Name>mcctcp</Name>
+  </Plugins>
+  <Chain>
+    <Component name="tcp.service" id="tcp">
+      <next id="http"/>
+      <tcp:Listen>
+        <tcp:Interface>0.0.0.0</tcp:Interface>
+        <tcp:Version>4</tcp:Version>
+      </tcp:Listen>
+    </Component>
+    <Component name="http.service" id="http">
+      <next id="soap">POST</next>
+      <next id="plexer">GET</next>
+    </Component>
+    <Component name="soap.service" id="soap">
+      <next id="plexer"/>
+    </Component>
+    <Plexer id="plexer">
+      <next id="echo">/Echo</next>
+    </Plexer>
+    <Service name="echo" id="echo">
+      <echo:prefix>[</echo:prefix>
+      <echo:suffix>]</echo:suffix>
+    </Service>
+  </Chain>
+</cfg:ArcConfig>
+\end{lstlisting}
 
-Upon a successful installation (nordugrid-arc-nox-hed package when installing from binaries), all the example configuration files coming with the release can be found under
+\section{Profiles}\label{sec:Profiles}
+In WS-ARC the concept of profiles have been introduced. A profile is a complete
+XML configuration template, for a specific service setup and it is used when
+configuring services in INI format. The core of WS-ARC
+uses an XML struture for service initialization, which makes the XML format
+the apparent choice for configuration. However the XML language is not simple
+and easy to use, which is why WS-ARC utilizes profiles, namely to provide a
+configuration format (the INI format) easy to use.
 
-\verb!<instllation_directory>/share/arc/!
+To write a profile, some knowledge on the INI (see
+section~\ref{sec:INI configuration}) and XML configuration formats (see
+section~\ref{sec:XML configuration}) is needed. Not necessaray to say, knowledge
+on the desired service setup is also needed. A XML configuration file is only a
+profile if it defines a option mapping from INI format to XML, thus making it
+possible to utilise the XML configuration through the INI format.
 
+Two XML attributes have been introduced to be able to create a mapping, and
+these are \texttt{inisections} and \texttt{initag}. XML elements which should be
+configurable in INI format must specify these attributes. The \texttt{initag}
+attribute defines the name of the tag which represent the respective XML element
+in the INI configuration file, and the \texttt{inisections} attribute is used to
+specify, as a space separated list, which sections the tag can occur in. The
+list is ordered, in the sense that the listed sections is searched, in the
+specified order, for the tag specified by the \texttt{initag} attribute, and if
+found in a section no further sections will be searched. Only leaf XML elements
+(elements with no child elements) should be mapped, mapping non leaf element
+result in a undefined behaviour. A profile example of a XML element mapping is
+shown in the listing below:
+\begin{lstlisting}[language=xml,title=ElementMapping.xml]
+...
+<Server>
+  ...
+  <Logger>
+    <File inisections="special common"
+          initag="logfile">/var/log/arched.log</File>
+    ...
+  </Logger>
+</Server>
+...
+\end{lstlisting}
+And this example profile can then be utilised in a INI configuration file as:
+\begin{lstlisting}[language=xml,title=ElementMapping.ini]
+profile=ElementMapping.xml
+logfile=/etc/arc/arched.log
+...
+[ special ]
+logfile=/var/arc/arched.log
+...
+\end{lstlisting}
+And the resulting XML configuration will look like:
+\begin{lstlisting}[language=xml]
+...
+<Server>
+  ...
+  <Logger>
+    <File>/var/arc/arched.log</File>
+    ...
+  </Logger>
+</Server>
+...
+\end{lstlisting}
+
+To be able to map XML attributes the use of a new XML element
+\texttt{AttributeRepresentation} have been introduced. The element should have
+three attributes set, the two attributes \texttt{inisections} and
+\texttt{initag} used as described above, and a new attribute \texttt{id} which
+specifies which attribute of the parent element is being mapped. A profile
+example of a XML attribute is shown in the listing below:
+\begin{lstlisting}[language=xml,title=AttributeMapping.xml]
+...
+<Component name="tcp.service">
+  ...
+  <Limit drop="false">
+    <AttributeRepresentation id="drop" inisections="common" initag="drop"/>
+    100
+  </Limit>
+</Component>
+...
+\end{lstlisting}
+And this example profile can then be utilised in a INI configuration file as:
+\begin{lstlisting}[language=xml,title=AttributeMapping.ini]
+profile=AttributeMappingExample.xml
+drop=true
+...
+\end{lstlisting}
+And the resulting XML configuration will look like:
+\begin{lstlisting}[language=xml]
+...
+<Component name="tcp.service">
+  ...
+  <Limit drop="true">100</Limit>
+</Component>
+...
+\end{lstlisting}
+
+\section{The \texttt{arched} daemon}\label{sec:arched}
+The \texttt{arched} daemon is the daemon which is used to start every service in
+WS-ARC. It will load a configuration file, and start the services specified
+herein.
+
+\hspace*{0.5cm}
+\begin{shaded}
+\textbf{\sffamily arched [OPTION...]}
+\end{shaded}
+\vspace{-0.3cm}(ARC 0.9)\par
+\begin{longtable}{llp{8cm}}
+  \texttt{-f, --foreground} && run daemon in foreground\\
+  \texttt{-c, --xml-config} & \textit{path} & full path of XML configuration file\\
+  \texttt{-i, --ini-config} & \textit{path} & fill path of INI configuration file\\
+  \texttt{-d, --config-dump} && dump generated XML configuration\\
+  \texttt{-p, --pid-file} & \textit{path} & full path of PID file\\
+  \texttt{-u, --user} & \textit{user} & user name\\
+  \texttt{-g, --group} & \textit{group} & group name\\
+  \texttt{-s, --schema} & \textit{path} & full path of XML schema file\\
+  \texttt{-h, --help} && Show help options
+\end{longtable}
+
+If no configuration file is specified, \texttt{arched} first looks for the
+\texttt{/etc/arc/server.ini} INI configuration file if it does not exist then
+the \texttt{/etc/arc/server.xml} XML configuration file is tried. If none of
+these is found \texttt{arched} will abort with an error and exit code 1.
+
+A INI configuration file can be specified with the \texttt{-i} option,  while
+the \texttt{-c} option makes it possible to specify a XML configuration file. If
+both options are given, the INI configuration will be chosen.
+
+\texttt{arched} also has the capability of printing the XML configuration which
+will be generated from a given INI configuration. By specifying the \texttt{-d}
+option along with the \texttt{-i} option and the path to the INI configuration
+\texttt{arched} will transform the INI configuration to XML and dump the output
+to standard out and exit.
+
+\appendix
+\section{List of standard profiles in WS-ARC}\label{app:List of profiles}
+
+\subsection{A-REX}
+\begin{description}
+%~ \item[ComputingElementWithFork]
+\item[ComputingElementWithFork] MINIMAL configuration template for a secure
+standalone Computing Element (CE) with a fork queue.
+%~ \item[NonSecureComputingElement]
+\item[InsecureComputingElement]
+%~ \item[SecureComputingElementWithMapfile]
+\item[ComputingElementWithMapfile]
+%~ \item[SecureComputingElementWithMapfileAndPBSBackend]
+\item[ComputingElementWithMapfileAndPBSBackend]
+%~ \item[SecureComputingElement\_DNlist]
+\item[ComputingElement\_DNlist]
+%~ \item[SecureComputingElement\_VOMS]
+\item[ComputingElement\_VOMS]
+\end{description}
+
+\subsection{Chelonia}
+\begin{description}
+%~ \item[CheloniaAllServicesCentralizedAHash]
+\item[CheloniaAllCentralizedAHash]
+%~ \item[CheloniaCentralizedAHashWithISIS]
+\item[CheloniaAllCentralizedAHashWithISIS]
+%~ \item[CheloniaAllServicesReplicatedAHash]
+\item[CheloniaAllReplicatedAHash]
+%~ \item[CheloniaAllServicesReplicatedAHashGatewayVOMSWithISIS]
+\item[CheloniaAllReplicatedAHashGatewayVOMSWithISIS]
+%~ \item[CheloniaAllServicesReplicatedAHashWithISIS]
+\item[CheloniaAllReplicatedAHashWithISIS]
+%~ \item[CheloniaReplicatedAHash]
+\item[CheloniaReplicatedAHash]
+%~ \item[CheloniaShepherdWithHopi]
+\item[CheloniaShepherdWithHopi]
+%~ \item[NonSecureNonDistributedStorageElement]
+\item[InsecureNonDistributedStorageElement]
+%~ \item[SecureNonDistributedStorageElement\_DNlist]
+\item[NonDistributedStorageElement\_DNlist]
+%~ \item[SecureNonDistributedStorageElement\_VOMS]
+\item[NonDistributedStorageElement\_VOMS]
+\end{description}
+
+\subsection{ISIS}
+\begin{description}
+%~ \item[NonSecureP2PIIS]
+\item[InsecureP2PIIS]
+%~ \item[NonSecureStandaloneIIS]
+\item[InsecureStandaloneIIS]
+%~ \item[SecureP2PIIS]
+\item[P2PIIS]
+%~ \item[SecureStandaloneIIS]
+\item[StandaloneIIS]
+\end{description}
+
+\subsection{Charon}
+\begin{description}
+%~ \item[CharonService]
+\item[Charon]
+\end{description}
+
+\subsection{Hopi}
+\begin{description}
+%~ \item[HopiService]
+\item[Hopi]
+%~ \item[HopiServiceNonSecure]
+\item[HopiInsecure]
+%~ \item[HopiServiceWithPlexer]
+\item[HopiWithPlexer]
+%~ \item[HopiServiceWithPlexerNonSecure]
+\item[HopiWithPlexerInsecure]
+\end{description}
+
+\subsection{Combined services}
+\begin{description}
+%~ \item[NonSecureComputingAndStorageElement]
+\item[InsecureComputingAndStorageElement]
+%~ \item[SecureComputingAndStorageElement\_DNlist]
+\item[ComputingAndStorageElement\_DNlist]
+%~ \item[SecureComputingAndStorageElement\_VOMS]
+\item[ComputingAndStorageElement\_VOMS]
+\end{description}
+
+\subsection{Testing}
+\begin{description}
+%~ \item[EchoService]
+\item[Echo]
+%~ \item[EchoInsecure]
+\item[EchoInsecure]
+%~ \item[EchoServicePython]
+\item[EchoPython]
+%~ \item[EchoServicePythonNonSecure]
+\item[EchoPythonInsecure]
+%~ \item[EchoServiceSAMLToken]
+\item[EchoSAMLToken]
+%~ \item[EchoServiceUsernameToken]
+\item[EchoUsernameToken]
+%~ \item[EchoServiceX509Token]
+\item[EchoX509Token]
+%~ \item[NonSecureHopiService]
+\item[InsecureHopiService]
+\end{description}
+
+\section{Profile attributes naming convention}
+Since profiles defines a mapping of XML elements or attributes to INI tags, they
+control the naming of the tags. So for consistency tags in different profiles
+reprensenting the same XML element or attribute should not be named differently.
+A naming convention list for the XML elements and attributes described in
+section~\ref{sec:XML configuration} is presented below. Note that there are only
+conventions for leaf XML elements (element with no children).
+
+\begin{tabular}{>{\ttfamily}l>{\ttfamily}l}
+\toprule
+\textrm{\textbf{XML element/attribute}} & \textrm{\textbf{INI tag}} \\
+\midrule
+/Server/PidFile & pidfile \\
+/Server/User & user \\
+/Server/Group & group \\
+/Server/Logger/File & logfile \\
+/Server/Logger/Level & loglevel \\
+/Server/Logger/Maxsize & logmaxsize \\
+/Server/Logger/Backups & logbackups \\
+\midrule
+\multicolumn{2}{c}{\textbf{TCP MCC}}\\
+\midrule
+/Component/Listen/Interface & interface \\
+/Component/Listen/Port & port \\
+/Component/Listen/Version & ipversion \\
+/Component/Listen/NoDelay & nodelay \\
+/Component/Listen/Timeout & timeout \\
+/Component/Limit & limit \\
+/Component/Limit.drop & drop \\
+\midrule
+\multicolumn{2}{c}{\textbf{TLS MCC}}\\
+\midrule
+/Component/KeyPath & x509\_user\_key \\
+/Component/CertificatePath & x509\_user\_cert \\
+/Component/CACertificatePath & x509\_cacert\_path \\
+/Component/CACertificatesDir & x509\_cacert\_dir \\
+/Component/CACertificatesDir.PolicyGlobus & policyglobus \\
+/Component/ClientAuthn & clientauthn \\
+/Component/VOMSCertTrustDNChain/VOMSCertTrustDN &  \\
+/Component/VOMSCertTrustDNChain/VOMSCertTrustRegex &  \\
+/Component/VOMSCertTrustDNChainsLocation & \\
+/Component/Handshake & handshake \\
+\bottomrule
+\end{tabular}
+
+\section{XML Schemas}
+\subsection{General Daemon Schema}
+\lstinputlisting[language=xml]{base.xsd}
+
+\subsection{Loader Schema}
+\lstinputlisting[language=xml]{loader.xsd}
+
+\subsection{TCP Schema}
+\lstinputlisting[language=xml]{tcp.xsd}
+
+\subsection{TLS Schema}
+\lstinputlisting[language=xml]{tls.xsd}
+
+%%% REFERENCES
 \end{document}

### 2010-02-01 doc/manuals/user_interface/ui-conf.tex
Interchange DEBUG and VERBOSE loglevels.
try different SRM parameters and cache on success, to fix bug 1431

@@
-Default behaviour of an ARC client can be configured by specifying
+The default behaviour of an ARC client can be configured by specifying
 alternative values for some parameters in the client configuration
@@
 possible values are \texttt{FATAL}, \texttt{ERROR}, \texttt{WARNING},
-\texttt{INFO}, \texttt{DEBUG} or \texttt{VERBOSE}.
+\texttt{INFO}, \texttt{VERBOSE} or \texttt{DEBUG}.
 
@@
 
+\index{configuration:srms.conf}\section{srms.conf}
+
+If any data management commands are used with the Storage Resource
+Management (SRM)~\cite{srm} protocol, the file
+
+\begin{shaded}
+ {\$}HOME/.arc/srms.conf
+\end{shaded}
+
+may be created to store cached information on these services. For more
+information see the description inside this file.
+
 \index{configuration:deprecated files}\section{Deprecated configuration files}

### 2010-02-01 doc/tech_doc/a-rex/arex_tech_doc.tex
Information about ported upload/download retries included into the documentation
try different SRM parameters and cache on success, to fix bug 1431

@@
 event must be sent, a separate file is written. Once an event is sent,
-the corresponding file is deleted.
+the corresponding file is deleted.\\
+Some utility files, which contain for example cached information on remote
+services, may also be found in the control directory. These files have
+a \textit{.conf} suffix.
 \end{itemize}
@@
 number is provided in job description and defaults to 0.
+\item \textbf{\textit{maxtransfertries}}\textit{=number} -- specifies
+  the maximum number of times download and upload will be attempted
+  per job (retries are only performed if an error is judged to be
+  temporary). This number must be greater than 0 and defaults to 10.
 \end{itemize}

### 2010-02-01 src/clients/data/arccp.cpp

@@ -212,6 +212,7 @@ bool arccp(const Arc::URL& source_url_,
   std::string cache_path;
   std::string cache_data_path;
   std::string id = "<ngcp>";
+  
   if (timeout <= 0)
     timeout = 300; // 5 minute default
   if (tries < 0)
@@ -437,7 +438,6 @@ bool arccp(const Arc::URL& source_url_,
   if (!cache_dir.empty()) cache = Arc::FileCache(cache_dir+" .", "", cache_user.get_uid(), cache_user.get_gid());
   if (verbose)
     mover.set_progress_indicator(&progress);
-  std::string failure;
   Arc::DataStatus res = mover.Transfer(*source, *destination, cache, Arc::URLMap(),
                                        0, 0, 0, timeout);
   if (!res.Passed()) {
@@ -557,6 +557,7 @@ int main(int argc, char **argv) {
     logger.msg(Arc::ERROR, "Failed configuration initialization");
     return 1;
   }
+  usercfg.UtilsDirPath(Arc::UserConfig::ARCUSERDIRECTORY);
 
   if (debug.empty() && !usercfg.Verbosity().empty())
     Arc::Logger::getRootLogger().setThreshold(Arc::string_to_level(usercfg.Verbosity()));

### 2010-02-01 src/clients/data/arcls.cpp

@@ -186,6 +186,7 @@ int main(int argc, char **argv) {
     logger.msg(Arc::ERROR, "Failed configuration initialization");
     return 1;
   }
+  usercfg.UtilsDirPath(Arc::UserConfig::ARCUSERDIRECTORY);
 
   if (debug.empty() && !usercfg.Verbosity().empty())
     Arc::Logger::getRootLogger().setThreshold(Arc::string_to_level(usercfg.Verbosity()));

### 2010-02-01 src/hed/dmc/srm/DataPointSRM.cpp

@@ -17,6 +17,7 @@
 #include <arc/StringConv.h>
 #include <arc/Logger.h>
 #include <arc/URL.h>
+#include <arc/UserConfig.h>
 #include <arc/data/DataBuffer.h>
 #include <arc/data/DataCallback.h>
 #include <arc/data/CheckSum.h>
@@ -71,7 +72,7 @@ namespace Arc {
 
   DataStatus DataPointSRM::Check() {
 
-    SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout);
+    SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout, usercfg.UtilsDirPath());
     if (!client) {
       if (timeout)
         return DataStatus::CheckErrorRetryable;
@@ -125,7 +126,7 @@ namespace Arc {
 
   DataStatus DataPointSRM::Remove() {
 
-    SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout);
+    SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout, usercfg.UtilsDirPath());
     if (!client) {
       if (timeout)
         return DataStatus::DeleteErrorRetryable;
@@ -173,7 +174,7 @@ namespace Arc {
     reading = true;
     buffer = &buf;
 
-    SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout, buffer->speed.get_max_inactivity_time());
+    SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout, usercfg.UtilsDirPath(), buffer->speed.get_max_inactivity_time());
     if (!client) {
       reading = false;
       if (timeout)
@@ -312,7 +313,7 @@ namespace Arc {
     }
     
     if (srm_request) {
-      SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout, buffer->speed.get_max_inactivity_time());
+      SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout, usercfg.UtilsDirPath(), buffer->speed.get_max_inactivity_time());
       if (client) {
         if(buffer->error_read() || srm_request->status() == SRM_REQUEST_SHOULD_ABORT) {
           client->abort(*srm_request);
@@ -341,7 +342,7 @@ namespace Arc {
     writing = true;
     buffer = &buf;
 
-    SRMClient *client = SRMClient::getInstance(url.fullstr(), timeout, buffer->speed.get_max_inactivity_time());
[…]

### 2010-02-01 src/hed/dmc/srm/srmclient/SRMClient.h

@@ -347,14 +347,14 @@
      * this SURL. All operations with a client instance must use SURLs with
      * the same host as this one.
      * @param timedout Whether the connection timed out
+     * @param utils_dir Path to a directory contining files with SRM info
      * @param timeout Connection timeout.
-     * @param version If this is non-NULL a client instance of this version
      * is returned.
      */
     static SRMClient* getInstance(std::string url,
                                   bool& timedout,
-                                  time_t timeout=300,
-                                  SRMVersion srm_version=SRM_VNULL);
+                                  std::string utils_dir,
+                                  time_t timeout=300);
     /**
      * empty destructor
      */

### 2010-02-01 src/hed/dmc/srm/srmclient/SRMURL.cpp

@@ -5,12 +5,15 @@
 //namespace Arc {
   
   SRMURL::SRMURL(std::string url) try: URL(url) {
+    portdefined = false;
     if(protocol != "srm") { 
       valid=false;
       return; 
     };
     valid=true;
     if(port <= 0) port=8443;
+    else portdefined = true;
+    srm_version = SRM_URL_VERSION_2_2; // v2.2 unless explicitly made otherwise
     if(HTTPOption("SFN", "") != "") {
       filename=HTTPOption("SFN");
       isshort=false;
@@ -20,13 +23,11 @@
         path.erase(0,1);
       };
       if (path[path.size()-1] == '1') srm_version = SRM_URL_VERSION_1;
-      else if (path[path.size()-1] == '2') srm_version = SRM_URL_VERSION_2_2;
     }
     else {
       if(path.length() > 0) filename=path.c_str()+1; // Skip leading '/'
-      path = "";
+      path = "/srm/managerv2";
       isshort=true;
-      srm_version = SRM_URL_VERSION_UNKNOWN;
     }
   } catch (std::exception e) {
     valid=false;
@@ -69,13 +70,19 @@
     return (protocol+"://"+host+":"+Arc::tostring(port)+"/"+filename);
   }
   
+  void SRMURL::GSSAPI(bool gssapi) {
+    if (gssapi)
+      AddOption("protocol", "gssapi", true);
+    else
+      AddOption("protocol", "gsi", true);
+  }
+
   bool SRMURL::GSSAPI(void) const {
     try {
-      const std::string proto_val =
-            ((std::map<std::string,std::string>&)(Options()))["protocol"];
-      if(proto_val == "gssapi") return true;
+      const std::string proto_val = Option("protocol");
+      if(proto_val == "gsi") return false;
     } catch (std::exception e) { };
-    return false;
+    return true;
   }
 
 //} // namespace Arc

### 2010-02-01 src/hed/dmc/srm/srmclient/SRMURL.h

@@ -57,7 +57,11 @@
   
     enum SRM_URL_VERSION SRMVersion() { return srm_version; };
     bool Short(void) const { return isshort; };
+    void GSSAPI(bool gssapi);
     bool GSSAPI(void) const;
+    void SetPort(int portno) { port = portno; };
+    /** Was the port number given in the constructor? */
+    bool PortDefined() { return portdefined; };
     operator bool(void) { return valid; };
     bool operator!(void) { return !valid; };
 
@@ -66,6 +70,7 @@
     std::string filename;
     bool isshort;
     bool valid;
+    bool portdefined;
     enum SRM_URL_VERSION srm_version;
     
 

### 2010-02-02 doc/manuals/configuration/configuration_manual.tex
Few words about GSI MCC. Small enhancements.

@@
   however if the \texttt{NoDelay} element is set to \texttt{true} data will be
-  send when ready.
-\item[\texttt{Timeout}] Specifies how much time, in seconds, there should be
-  used on waiting for a client response. If no client response is obtained
-  before timeout the connection is dropped. The default value is 60 seconds.
+  sent when ready. This is mainly used for optimizing connections for communication
+  patterns which involve a lot of very short messages exchanged by service 
+  and client.
+\item[\texttt{Timeout}] Specifies how much time, in seconds, data sending or
+  receiving may take. If time is exceeded MCC reports error and in most cases 
+  connection is dropped. The default value is 60 seconds.
 \end{description}
-It is also possible to put a limit on incoming connections, and how these should
-be handled if the limit is reached. The limit can only be put on all configured
-interfaces, and not per interface. The \texttt{Limit} element specifies the
-limit on incoming connections, and the \texttt{drop} attribute specifies whether
-new connections, above the limit, will be put on hold (default) or if they
-should be dropped, which is specified by setting the attribute to \texttt{true}.
+It is also possible to put a limit on number of incoming connections, and how 
+these should be handled if the limit is reached. The limit can only be put on all 
+configured interfaces, and not per interface. The \texttt{Limit} element specifies 
+the limit on incoming connections, and the \texttt{drop} attribute specifies whether 
+new connections, above the limit, will be put on hold (default) or if they should 
+be dropped, which is specified by setting the attribute to \texttt{true}.
 
@@
 and it should be put on top of a TCP MCC. By default it is able to understand
-the SSLv2, SSLv3 and TLSv1 protocols, however the \texttt{Handshake} element can
-be used to specify another behaviour. The supported values for the
+the SSLv2 (hadshake only), SSLv3 and TLSv1 protocols, however the \texttt{Handshake} 
+element can be used to specify another behaviour. The supported values for the
 \texttt{Handshake} element are \texttt{TLS} (default) and \texttt{SSLv3}, where
@@
 
-%~ \paragraph{Globus Policy}
-%~ attribute \texttt{PolicyGlobus} (CACertificatePath).
+If element \texttt{CACertificatesDir} have attribute \texttt{PolicyGlobus} set
+to "true" additional verification of client credentials is performed. Client
+credentials are restricted to be signed by specific Certification Authorities
+as specified in additional configuration files. Those files are stored in the
+location specified by the element \texttt{CACertificatesDir} and have names
+made of the hash of the subject of CA certificate followed by \texttt{.signing\_policy}
+suffix. Information about format of those files may be found at 
+http://dev.globus.org/wiki/Signing\_Policy.
 %~
@@
 
+\subsubsection{GSI MCC}\label{sec:gsi-mcc}
+At its purpose MCC is similar to TLS described in section \ref{sec:tls-mcc}. It 
+provides security layer using Globus implementation of GSS-API and corresponding 
+communication protocol.
+
+Its configuration element \texttt{KeyPath} and \texttt{CertificatePath} are 
+identical to those of TLS MCC. This MCC is implemented using libraries of 
+Globus Toolkit and its functionality and requirements are better to be 
+described by descrition of Globus gssapi library.
+
+Use of this MCC is not recommended due to non-standard origin of underlying
+protocol.
+
 \subsubsection{HTTP MCC}\label{sec:http-mcc}
@@
 \end{lstlisting}
-
-%~ \subsubsection{GSI MCC}\label{sec:gsi-mcc}
-%~ \texttt{KeyPath} Location of private key.
-%~
-%~ \texttt{CertificatePath} Location of public certificate.
-%~
-%~ \texttt{ProxyPath} Location of proxy credentials - includes certificates, key
-%~ and chain of involved certificates. Overwrites elements \texttt{KeyPath} and
-%~ \texttt{CertificatePath}.
 
@@
 \subsection{Security Handlers}\label{sec:Security Handlers}
-\subsubsection{Identity Mapping}% identity.map
-\subsubsection{ARC Authorization}% arc.authz
-\subsubsection{Delegation Collector}% delegation.collector
-\subsubsection{Username Token}% usernametoken.handler
-%\subsubsection{Delegation Handler}% delegation.handler
-%\subsubsection{X509 Token}% x509token.handler
+\subsubsection{Identity Mapping}\label{sec:identitymap-shc}
+\subsubsection{ARC Authorization}\label{sec:arcautz-shc}
+\subsubsection{Delegation Collector}\label{sec:delegcollector-shc}
+\subsubsection{Delegation Handler}\label{sec:deleghandler-shc}
+\subsubsection{Username Token}\label{sec:unametoken-shc}
+\subsubsection{Username Token}\label{sec:x509token-shc}
 %\subsubsection{SAML Token}% samltoken.handler
 %\subsubsection{SAML Single Sign-On}% saml2ssoassertionconsumer.handler
+
+\subsection{Policy Decision Points}\label{sec:Policy Decision Points}
 

### 2010-02-03 doc/manuals/configuration/configuration_manual.tex
Started adding information about sechandlers

@@
 \subsection{Security Handlers}\label{sec:Security Handlers}
+Security Handlers are components which extract security related properties of
+the message, collect then and use for making authorization decisions.
+
+The Security Handlers are assigned MCCs and are arranged in queues. Each queue
+is processing specific messages. Usually MCCs implement two queues named 
+"incoming" and "outgoing" for the messages going through chain to service
+and returning back respectively.
+
+The Security Handlers in each queue are applied sequentially and if any
+produces negative result processing stops are error is raised.
+
+The Security Handlers are added to MCCs through \texttt{SecHandler} configuration
+element put inside \texttt{Component} or \texttt{Service} elements. The
+\texttt{name} attribute specifies which plugin to be used. And queue is 
+specified by \texttt{event} attribute. Each configured instance can be 
+optionally assigned identifier through \texttt{id} attribute. Tne content
+of \texttt{SecHandler} is specific for every kind of plugin.
+Example is below:
+\begin{lstlisting}[language=xml]
+...
+<Component name="mcc1.service" id="mcc1">
+  <next id="mcc2"/>
+  <SecHandler name="handler1" event="incoming" id="id1">
+  </SecHandler>
+</Component>
+...
+\end{lstlisting}
+
+\subsubsection{ARC Authorization}\label{sec:arcautz-shc}
+This Security Handler applies set of Policy Decision Point (PDP) element to the
+message and then produces decision result using combining algorithm specified in 
+\texttt{action} attribute of \texttt{SecHandler} element. For desciption of PDP
+elements see section \ref{sec:Policy Decision Points}.
+
+Available combining algorithms are:
+\begin{description}
+\item[breakOnDeny] - processing stops when PDP returns Deny decision. The SecHandler
+produces negative result. This is a default algorith.
+\item[breakOnAllow] - processing stopr when PDP return Allow decision. The SecHandler
+produces positive result.
+\item[breakAlways] - processing stops after first PDP. Produced result is used as
+result of SecHandler.
+\item[breakNever] - all PDPs are applied. Result of last PDP is used as result of 
+SecHandler.
+\end{description}
+
 \subsubsection{Identity Mapping}\label{sec:identitymap-shc}
-\subsubsection{ARC Authorization}\label{sec:arcautz-shc}
+
 \subsubsection{Delegation Collector}\label{sec:delegcollector-shc}

### 2010-02-04 doc/manuals/configuration/configuration_manual.tex
Few more sec handlers described

@@
 
-\subsubsection{ARC Authorization}\label{sec:arcautz-shc}
+\subsubsection{ARC Authorization}\label{sec:arcauthz-shc}
 This Security Handler applies set of Policy Decision Point (PDP) element to the
@@
 \texttt{action} attribute of \texttt{SecHandler} element. For desciption of PDP
-elements see section \ref{sec:Policy Decision Points}.
+elements see section \ref{sec:Policy Decision Points}. The plugin name of this 
+Secirity hadler is \texttt{arc.authz} and it resides in \texttt{arcshc} loadable module.
 
@@
 \end{description}
+Here is an example of of ARC Authorization Security Handler configuration.
+\begin{lstlisting}[language=xml]
+...
+  <SecHandler name="arc.authz" event="incoming" id="authz">
+    <PDP name="arc.pdp">
+      ...
+    </PDP>
+    <PDP name="xacml.pdp">
+      ...
+    </PDP>
+  </SecHandler>
+...
+\end{lstlisting}
 
 \subsubsection{Identity Mapping}\label{sec:identitymap-shc}
+The Identity Mapping plugin applies set of associated PDPs to the message
+being processed and depending on result assigns local account to the 
+message. Its plugin name is \texttt{identity.map} and it resides in 
+\texttt{identitymap} loading module.
+
+This Security Handler works in a similar way to ARC Authorization (see
+section \ref{sec:arcauthz-shc}). But instead of passing results of PDPs as
+final processing result it applies local account mapping algorithm 
+associated with particular PDP.
+
+Configuration is also similar with only difference being that inside 
+each \texttt{PDP} configuration element there is additional element
+representing mapping algorithm used for producing local account name.
+Available are following mapping algorithms:
+\begin{description}
+\item[\texttt{LocalName}] - maps to specified account name.
+\begin{lstlisting}[language=xml]
+...
+    <PDP name="arc.pdp">
+      <LocalName>nobody</LocalName>
+      ...
+    </PDP>
+...
+\end{lstlisting}
+
+\item[\texttt{LocalList}] - uses subject of X.509 certificate presented by client
+to perform mapping. Element content provides path to local file containing pairs
+of subjects and account names. One pair per line separated by blank space.
+\begin{lstlisting}[language=xml]
+...
+    <PDP name="arc.pdp">
+      <LocalList>/etc/grid-security/grid-mapfile</LocalList>
+      ...
+    </PDP>
+...
+\end{lstlisting}
+
+\item[\texttt{LocalSimplePool}] - manages dynamic mapping of X.509 client subject
+to a set of local accont names. Content of element point to directory containing
+file named \texttt{pool}. This file contains list of local account available for
+mapping one per line. This directory is also used fro storing currently active
+mappings.
+\begin{lstlisting}[language=xml]
+...
+    <PDP name="arc.pdp">
+      <LocalSimplePool>/etc/grid-security/mappool</LocalSimplePool>
+      ...
+    </PDP>
+...
+\end{lstlisting}
+\end{description}
 
 \subsubsection{Delegation Collector}\label{sec:delegcollector-shc}
-\subsubsection{Delegation Handler}\label{sec:deleghandler-shc}
-\subsubsection{Username Token}\label{sec:unametoken-shc}
-\subsubsection{Username Token}\label{sec:x509token-shc}
+This plugin processes chain of X.509 certificates and extracts embedded
+policies. Policies are attached to the message and may be later used by
+Delegation PDP described in section \ref{sec:delegation-pdp}. The
+plugin name of this Security Handler is \texttt{delegation.collector} 
+and it resides in \texttt{mcctls} loadable module.
+It has no configuration parameters.
+
+% \subsubsection{Delegation Handler}\label{sec:deleghandler-shc}
+
+\subsubsection{UsernameToken}\label{sec:unametoken-shc}
+The UsernameToken Security Handler processes information in the SOAP Header
+stored according to "Web Services Security UsernameToken Profile 1.0"
+\cite{ws-security-usernametoken}. The plugin name of this Security Handler
+is \texttt{usernametoken.handler} and it resides in \texttt{arcshc} loadable
+module.
+
+The configuration element \texttt{Process} defines type of processing to 
+be performed. Possible values are:
+\begin{description}
+\item[extract] - parse and process elements in SOAP Header. This option
+shoudl be used for service side configuration.
+\item[generate] - create UsernameToken. This option is only used for client
+side.
+\end{description}
+
+Only another configuration element relevant for service side processing is
+\texttt{PasswordSource}. It specifies path to local file containing
+list of usernames and passwords (one pair per line) which are accepted 
+by plugin. If there is no pair in file matching one SOAP Header this 
+Security Handler will raise an error.
+
+\subsubsection{X.509 Token}\label{sec:x509token-shc}
+The X.509 Token Security Handler processes information in the SOAP Header
+stored according to "Web Services Security X.509 Token Profile 1.0"
+\cite{ws-security-usernametoken}. The plugin name of this Security Handler
+is \texttt{x509token.handler} and it resides in \texttt{arcshc} loadable
+module.
+
+The configuration element \texttt{Process} defines type of processing to 
+be performed. Possible values are:
+\begin{description}
+\item[extract] - parse and process elements in SOAP Header. This option
+shoudl be used for service side configuration.
+\item[generate] - create X.509 Token. This option is only used for client
+side.
+\end{description}
+
+In \texttt{extract} mode this plugin process X.509 Token from SOAP Header
+and validates contained credetials against locally stored CA certificates.
+Following configuration elements specify location of trusted CA certificates:
+\begin{description}
+\item[CACertificatePath] - path to file containing CA certificate.
+\item[CACertificatesDir] - path to directory containing files with CA certificates.
+\end{description}
+If provided credentials failed verification this Security Handler will raise 
+an error.
+
 %\subsubsection{SAML Token}% samltoken.handler
@@
 
-%%% REFERENCES
+\bibliography{grid}
 \end{document}

### 2010-02-04 doc/manuals/configuration/configuration_manual.tex
Added ARC PDP and Delegation PDP

@@
 stored according to "Web Services Security X.509 Token Profile 1.0"
-\cite{ws-security-usernametoken}. The plugin name of this Security Handler
+\cite{ws-security-x509token}. The plugin name of this Security Handler
 is \texttt{x509token.handler} and it resides in \texttt{arcshc} loadable
@@
 
-%\subsubsection{SAML Token}% samltoken.handler
-%\subsubsection{SAML Single Sign-On}% saml2ssoassertionconsumer.handler
+\subsubsection{SAML Token}
+
+\subsubsection{SAML Single Sign-On}
 
 \subsection{Policy Decision Points}\label{sec:Policy Decision Points}
+The Policy Decision Point component is responsible for applying particular
+policy or any kind of authorization evaluation algorithm to properties 
+of the message and providing authorization decision which then used by
+other components - usually by the Security Handlers implementing 
+functionality of Policy Enforcement Point.
+
+\subsubsection{Allow}
+This PDP always returns positive result. Its plugin name is \texttt{allow.pdp}
+and it resides in \texttt{arcshc} loadable module.
+
+\subsubsection{Deny}
+This PDP always returns negative result. Its plugin name is \texttt{deny.pdp}
+and it resides in \texttt{arcshc} loadable module.
+
+\subsection{Simple List}
+This PDP matches X.509 subject of client certificate to list of predefined values.
+Its plugin name is \texttt{simplelist.pdp} and it resides in \texttt{arcshc} loadable
+module.
+
+Multiple configuration elements \texttt{DN} specify accepted subjects. Also attribute
+\texttt{location} may be used to point to a file containing list of subjects, one
+per line. If client's subject is not listed neither in \texttt{DN} nor in 
+\texttt{location} PDP returns negative result. Otherwise result is positive. Below is
+configuration example for simple authorization by grid-mapfile:
+\begin{lstlisting}[language=xml]
+...
+    <PDP name="simplelist.pdp" location="/etc/grid-security/grid-mapfile">
+      <DN>/O=Grid/O=NorduGrid/CN=Service Admin</DN>
+    </PDP>
+...
+\end{lstlisting}
+
+\subsubsection{ARC}\label{sec:arc-pdp}
+This plugin applies authorization policy written in ARC Policy language to attributes
+of the message. Its plugin name is \texttt{arc.pdp} and it resides in \texttt{arcshc}
+loadable module. For more information about ARC Policy language please "Security 
+Framework of ARC NOX" \cite{arc1-sec}.
+
+The configuration element \texttt{PolicyStore} specifies locations of policy. Its
+\texttt{Location} subelement provides path to file storing policy. Also policy can be 
+directly embedded into configuration by using \texttt{Policy} configuration element.
+
+The \texttt{PolicyCombiningAlg} element specifies how results from multiple policies
+are combined. For detailed explanation please refer to \cite{arc1-sec}.
+
+It is also possible to limit number of message attributes which are matched to policy
+using \texttt{Filter} configuration element. This may be needed to achieve better
+evaluation performance. The multiple subelements \texttt{Select} and \texttt{Reject}
+allow to either use only specific attributes or remove specific attributes. If neither 
+\texttt{Select} not \texttt{Reject} are specified then all attribites are used.
+
+Example of simple ARC PDP reading policy from external file is below:
+\begin{lstlisting}[language=xml]
+...
+    <PDP name="arc.pdp">
+      <PolicyStore>
+        <Location>/etc/grid-security/policy.xml</Location>
+      </PolicyStore>
+    </PDP>
+...
+\end{lstlisting}
+
+\subsubsection{GACL}
+
+\subsubsection{XACML}
+
+\subsubsection{Delegation}\label{sec:delegation-pdp}
+This plugin applies authorization policy written in ARC Policy language and obtained
+from cahin of X.509 Proxy Certificates of client to attributes of the message. Its 
+plugin name is \texttt{delegation.pdp} and it resides in \texttt{arcshc} loadable module.
+
+Functionality wise it is very similar to ARC PDP described in section \ref{sec:arc-pdp}
+but uses policy internally obtained from message. Hence it has only \texttt{Filter}
+configuration element.
+
+For detailed explanation how policy document is embedded into X.509 Proxy Certificate 
+please refer to \cite{arc1-sec}.
+
+\subsubsection{Service Invoker}
 

### 2010-02-05 doc/manuals/configuration/configuration_manual.tex
Added GACL, XACL and service PDPs

@@
 
-\subsection{Simple List}
+\subsubsection{Simple List}
 This PDP matches X.509 subject of client certificate to list of predefined values.
@@
 of the message. Its plugin name is \texttt{arc.pdp} and it resides in \texttt{arcshc}
-loadable module. For more information about ARC Policy language please "Security 
-Framework of ARC NOX" \cite{arc1-sec}.
+loadable module. For more information about ARC Policy language please see "Security 
+Framework of ARC NOX" \cite{arc1-sec}. That docment also describes security
+attributes produced by different components of ARC which can be included in policies.
 
@@
 
-\subsubsection{GACL}
-
-\subsubsection{XACML}
+\subsubsection{GACL}\label{gacl-pdp}
+This PDP shares configuration parameters and provides functionality similar to ARC PDP
+described in the section \ref{sec:arc-pdp}. Its plugin name is \texttt{gacl.pdp} and it
+resides in the same \texttt{arcshc} loadable module.
+
+This plugin accepts policies written in GACL instead of ARC language. For more information
+about GACL prease check "The Gridsite Web" \cite{gacl} and "GACL Mini-Howto" \cite{gacl-mini}.
+
+\subsubsection{XACML}\label{xacml-pdp}
+This plugin's configuration and functionality are similar to previously described ARC and
+GACL PDPs.  Its plugin name is \texttt{xacml.pdp} and it resides in the same \texttt{arcshc}
+loadable module.
+
+This plugin accepts policies written in subset of XACML language \cite{xacml}. Currently 
+XACML functionality is not complete yet. For more information see \cite{arc1-sec}.
 
@@
 
-\subsubsection{Service Invoker}
+\subsubsection{Service Invoker}\label{service-pdp}
+This plugin comunicates with possibly remote authorization service by passing it authorization
+request and getting back authorization decision. Its plugin name is \texttt{pdpservice.invoker}
+and it resides in \texttt{arcshc} loadable module.
+
+The configuration element \texttt{Filter} is similar to those in other PDPs.
+
+To specify service to be contacted \texttt{ServiceEndpoint} element is used. It contains
+contact URL of service. Service is expected to accept SOAP over HTTP(S) messages conforming
+to protocol defined in \texttt{TransferProtocol} configuration element. Possible options
+are \texttt{arc} for ARC own protocol and \texttt{saml} for "SAML 2.0 Profile of the XACML 2.0"
+kind of exchange. The format of payload - authorization request and response - is controlled
+using \texttt{RequestFormat} element. Implemented are \texttt{xacml} and \texttt{arc} options
+for XACML conforming or ARC specific formats respectively.
+
+For establishing TLS communication channel to the service PDP needs to apply X.509 credentials.
+For that configuration elements \texttt{CertificatePath} and \texttt{KeyPath} are used.
+Alternatively if PDP uses X.509 Proxy Certificate path to file containing proxy may be
+specified in \texttt{ProxyPath}.
+
+For verifying credentials presented by service trusted CA certificates are specified through
+\texttt{CACertificatePath} and \texttt{CACertificatesDir} for single file and directory 
+containing multiple files respectively.
 

### 2010-02-08 doc/tech_doc/client/client_technical.tex
Removing dependency on ltablex latex style.

@@
 \usepackage{booktabs}
-\usepackage{ltablex}
 \usepackage[table]{xcolor} % Note: The rowcolors command does not work properly with the tabularx environment.
@@
 \newcommand{\NST}[1]{\normalsize{\textnormal{#1}}}
-\newcommand{\defineheadfoot}{\endfirsthead %
-\midrule %
-\multicolumn{5}{c}{Continued from previous page}\\%
-\midrule %
-\endhead %
-\midrule %
-\multicolumn{5}{c}{Continues at next page}\\%
-\midrule %
-\endfoot %
-\bottomrule %
-\endlastfoot}
+% parbox with centered text and vertical padding. It provides fixed width table-cells holding multiline centered text. \mlccell{<width>}{<text>}
+% Use in a tabular environment as \mlccell{10em}{line 1 \linebreak line 2 \linebreak line 3, etc.}
+\newcommand{\mlccell}[2]{\parbox{#1}{\vspace{0.4\baselineskip}\centering #2 \vspace{0.4\baselineskip}}}
+\newcommand{\n}{\linebreak}
 \definecolor{shadecolor}{rgb}{1,1,0.6}
@@
 
-\newcolumntype{C}{>{\ttfamily\footnotesize}c}
-\newcolumntype{T}{>{\ttfamily\footnotesize\centering\arraybackslash}X}
-
-%
 %----- DON'T CHANGE HEADER MATTER
@@
 \rowcolors{3}{tablerowcolor}{}
-\begin{tabularx}{\textwidth}{CCTCT}
+\texttt{
+\begin{tabular}{ccccc}
 \toprule
@@
 \midrule
-\defineheadfoot
-ACCEPTED & ACCEPTED & ACCEPTED & Pending & REGISTERED,\linebreak PENDING\\
+ACCEPTED & ACCEPTED & ACCEPTED & Pending & \mlccell{10em}{REGISTERED,\n PENDING}\\
 PREPARING & PREPARING & PREPARING & \NST{None} & \NST{None} \\
@@
 QUEUING & INLRMS:Q & INLRMS:Q & \NST{None} & IDLE \\
-RUNNING & INLRMS:R & INLRMS:R,\linebreak INLRMS:EXECUTED,\linebreak INLRMS:S, INLRMS:E & Running & RUNNING,\linebreak REALLY-RUNNING\\
+RUNNING & INLRMS:R & \mlccell{10em}{INLRMS:R,\n INLRMS:EXECUTED,\n INLRMS:S,\n INLRMS:E} & Running & \mlccell{10em}{RUNNING, \n REALLY-RUNNING}\\
 FINISHING & FINISHING & FINISHING & \NST{None} & \NST{None}\\
@@
 KILLED & KILLED & KILLED & \NST{None} & CANCELLED \\
-FAILED & FAILED & FAILED & Failed & DONE-FAILED,\linebreak ABORTED\\
+FAILED & FAILED & FAILED & Failed & \mlccell{10em}{DONE-FAILED,\n ABORTED}\\
 DELETED & DELETED & DELETED & \NST{None} & \NST{None} \\
 OTHER & \NST{Any other state} & \NST{Any other state} & \NST{None} & \NST{Any other state} \\
-\end{tabularx}
+\bottomrule
+\end{tabular}
+}
 \end{center}

### 2010-02-09 doc/manuals/configuration/configuration_manual.tex
Adding more schemas and reducing their font size
Removing HTTP schema from configuration manual.
Added references to another manuals and description of SAML Token SHC

@@
 \begin{description}
-\item[A-REX]
-  A-REX
-\item[Chelonia]
-\item[ISIS]
-\item[Charon]
-\item[hopi]
-  hopi, a lightweight HTTP webserver, is described in details in \%\% REFERENCE.
-
+\item[A-REX] - configuration manual in "The ARC Computational Job Management
+Module - A-REX" \cite{a-rex},
+\item[Chelonia] storage multi-service - "Chelonia Adminstrator's Manual"
+\cite{storage-manual},
+\item[ISIS] distributed P2P information service - "ARC Information System"
+\cite{isis},
+\item[Charon] authorization service - currently described in "Security
+Framework of ARC NOX" \cite{arc1-security},
+\item[HOPI] lightweight HTTP webserver - "The HOPI Manual" \cite{hopi} 
+and also partially described in "Chelonia Administrator's Manual"
+\cite{storage-manual}.
 \end{description}
@@
 the first type should be able to listen on some interface for messages, while
-the other simply
-receives messages from another MCC. The MCC then process the message and pass it
-on to a single or multiple other MCCs or services. MCCs should be configured
-using the XML element \texttt{Component}, where the \texttt{name} attribute
-specifies the MCC to use. The following MCCs is part of WS-ARC:
+the other simply receives messages from another MCC. The MCC then process
+the message and pass it on to a single or multiple other MCCs or services.
+MCCs should be configured using the XML element \texttt{Component}, where the
+\texttt{name} attribute specifies the MCC to use. The following MCCs are part
+of WS-ARC:
 \begin{description}
@@
   section~\ref{sec:http-mcc},
-%~ \item[\texttt{gsi.service}] Processes GSI messages, see
-  %~ section~\ref{sec:gsi-mcc},
+\item[\texttt{gsi.service}] Processes GSI messages as implemented in 
+Globus toolkit \cite{globus}, see section~\ref{sec:gsi-mcc},
 \item[\texttt{soap.service}] Processes SOAP messages, see
@@
 Additionally a special MCC,
-the Plexer, exist which is able to pass messages to different components simply
-based on the endpoint the messages was sent to, and it should should be
+the Plexer, exists which is able to pass messages to different components simply
+based on the endpoint the messages was sent to, and it should be
 configured using the \texttt{Plexer} element, see section~\ref{sec:plexer}. In
@@
 \texttt{next} element, where the \texttt{id} attribute specifies the \texttt{id}
-of that MCC or service. The message chain should be contained in the root
-element \texttt{ArcConfig}, which is shown in the conceptual configuration
-example below:
+of that MCC or service. Some MCCs support multiple \texttt{next} elements with
+textual label inside. Those are used by such MCCs to perform internal routing
+of processed message. But most MCCs do not do any routing and hence require
+single \texttt{next} without label.
+
+The message chain should be contained in the root element \texttt{ArcConfig},
+which is shown in the conceptual configuration example below:
 \begin{lstlisting}[language=xml]
@@
 in one or more levels.
+
+More technical information about available MCCs and the messages their process
+may be found in the Hosting Environment technical manual \cite{hed}.
 
@@
 it should either be put on top of the TCP or TLS MCC. A working configuration of
-this MCC is simple, since it should only use the \texttt{next} element which
-must have a value equal to one of the HTTP request methods. An example of a HTTP
-MCC configuration is shown below.
+this MCC is simple, since it should only use the \texttt{next} element.
+Differently from most other MCCs this one may have multiple \texttt{next} with
+content being name of HTTP method in uppercase. Depending on method client
+requested message may be routed to different components. Most services 
+of WS-ARC support GET and PUT for data communication and POST for SOAP. Some
+also support HEAD.
+
+An example of a HTTP MCC configuration is shown below.
 
@@
 The Plexer is a special MCC which filters messages based on their URL endpoints.
-One or more regular expressions (POSIX extended) should be defined using one or
-more \texttt{next} elements, and these will then be used to filter incoming
-messages, thus forwarding messages, where the path of the endpoint URL matched a
-given regular expression, to the component or service specified by the
-\texttt{id} attribute of the corresponding \texttt{next} element. At the first
-match the message will be forwarded to the respective component or service and
-no more expressions will be tried. If a message did not match any of the regular
-expressions then it is discarded and an error code is returned to the client. An
-example of a Plexer configuration is shown below.
-%% What is the order testing the regular expressions. First listed first tested or Last listed firsted tested?
+It supports multiple \texttt{next} elements with content being regular expressions
+(POSIX extended) used to filter incoming messages.
+The message is forwarded to the component or service specified by the \texttt{id}
+attribute of the corresponding \texttt{next} element which mathed the endpoint 
+URL of the message. At the first match the message will be forwarded to the
+respective component or service and no more expressions will be tried. The
+\texttt{next} elements are tried in order they are specified in configuration.
+If the message did not match any of the regular expressions then it is discarded
+and an error is returned to the client. An example of a Plexer configuration is
+shown below.
 
@@
 \end{lstlisting}
+
+For more technical information about security-related attributes of the
+message and the Security Handlers please see \cite{arc1-security}.
 
@@
 The X.509 Token Security Handler processes information in the SOAP Header
-stored according to "Web Services Security X.509 Token Profile 1.0"
+stored according to "Web Services Security: X.509 Token Profile 1.0"
 \cite{ws-security-x509token}. The plugin name of this Security Handler
@@
 \subsubsection{SAML Token}
+The SAML Token Security Handler processes information in the SOAP Header
+stored according to "Web Services Security: SAML Token Profile 1.1"
+\cite{ws-security-samltoken}. The plugin name of this Security Handler
+is \texttt{samltoken.handler} and it resides in \texttt{arcshc} loadable
+module.
+
+The configuration of this plugin is identical to previously described
+X.509 Token Security Handler. It supports configuration elements 
+\texttt{Process}, \texttt{CACertificatePath} and \texttt{CACertificatesDir}.
+
+The information collected by this plugin may then be used in ARC Policy 
+Decision Point and Service Invoker Policy Decision Point plugins.
 
@@
 
+ \lstset{
+         basicstyle=\footnotesize\ttfamily,
+ }
+
 \section{XML Schemas}
@@
 
-\subsection{TCP Schema}
+\subsection{TCP MCC Schema}
 \lstinputlisting[language=xml]{tcp.xsd}
 
-\subsection{TLS Schema}
+\subsection{TLS MCC Schema}
 \lstinputlisting[language=xml]{tls.xsd}
 
-\bibliography{grid}
+\subsection{GSI MCC Schema}
+\lstinputlisting[language=xml]{gsi.xsd}
+
+\subsection{ARC Authorization Security Handler Schema}
+\lstinputlisting[language=xml]{ArcAuthZ.xsd}
+
+\subsection{UsernameToken Security Handler Schema}
+\lstinputlisting[language=xml]{UsernameTokenSH.xsd}
+
+\subsection{X.509 Token Security Handler Schema}
+\lstinputlisting[language=xml]{X509TokenSH.xsd}
+
+\subsection{SAML Token Security Handler Schema}
+\lstinputlisting[language=xml]{SAMLTokenSH.xsd}
+
+\subsection{ARC PDP Schema}
+\lstinputlisting[language=xml]{ArcPDP.xsd}
+
+\subsection{XACML PDP Schema}
+\lstinputlisting[language=xml]{XACMLPDP.xsd}
+
+\subsection{Delegation PDP Schema}
+\lstinputlisting[language=xml]{DelegationPDP.xsd}
+
+\subsection{Service Invoker PDP Schema}
+\lstinputlisting[language=xml]{PDPServiceInvoker.xsd}
+
+\subsection{Simple List PDP Schema}
+\lstinputlisting[language=xml]{SimpleListPDP.xsd}
+
+\bibliography{grid,nordugrid}
 \end{document}

### 2010-02-11 doc/manuals/user_interface/ui-data.tex
arccp -f can be used to skip validation checks

@@
    \texttt{-f, --force} && if the destination is an indexing service and not the same as the source and the destination is already registered, then the copy is normally not done.
-However, if this option is specified the source is assumed to be a replica of the destination created in an uncontrolled way and the copy is done like in case of replication\\
+However, if this option is specified the source is assumed to be a replica of the destination created in an uncontrolled way and the copy is done like in case of replication.
+Using this option also skips validation of completed transfers.\\
    \texttt{-i, --indicate} && show progress indicator\\

### 2010-02-11 doc/manuals/configuration/configuration_manual.tex
Making it possible to load a profile by only specifying its name.

@@
 [ common ]
-profile = <path-to-profile>
+profile = <profilename> or <path-to-profile>
 
@@
 
-The \texttt{profile} attribute is also a special, as it is used to indicate
-which profile the INI configuration uses, and its value should be the
-absolute path to the profile file. The \texttt{profile} attribute must be
+The \texttt{profile} attribute is also special, as it is used to indicate
+which profile the INI configuration uses. Its value should either be the name of
+a WS-ARC profile (no \texttt{xml} suffix), see
+appendix~\ref{app:List of profiles} or it should be the absolute path to a
+profile (with \texttt{xml} suffix). The \texttt{profile} attribute must be
 located in the common section. If multiple \texttt{profile} attributes exist

### 2010-02-15 doc/manuals/janitor/tex_installation/installation.tex
Smallish series of updates on Janitor documentation.

@@
 can be installed as root by "apt-get install nordugrid-arc1-janitor".
-Installing it will not drag other components of ARC with it since the
-Janitor can be used in its own right or in conjunction with another
+Installing it will not drag other components of ARC with it, since the
+Janitor can be used in its own right -- or in conjunction with another
 grid system, possibly. Packages for Redhat/Fedora and SuSE/OpenSuSE are
@@
 
-Furthermore it is recommended to install the ontology editor 
+Furthermore many users will want to consider installing the ontology editor 
 Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}}
-in order to be to easily maintain the knowledge database of installable
+to easily maintain the knowledge database of installable
 packages.  At the time of writing, no Linux distribution is offering
-packages for this fine tool. The basic editing can also be performed
-fairly easily without that tool, but one is unlikely to use the technology
-to its full potential, then.
+packages for this fine tool. However, the basic editing can also be
+performed fairly easily without that tool, and everyone is working on
+simplifying that process.
 

### 2010-02-15 doc/manuals/janitor/tex_maintenance/maintenance.tex
janitor.pdf builds again. Many thanks to Anders for spotting the issue.
Further update on the Janitor docs.

@@
 Janitor.
-To administrate the catalog, it is strongly recommended
-(but not ultimately required) to use the ontology editor Prot\`eg\`e,
-as explained in the next section.
+To administrate the catalog, one can either use an ontology editor 
+like Prot\`eg\`e, as explained in the next section, edit the documents
+manually.
 In the last section, a typical use case in maintaining the Janitor will
@@
 
-To edit the catalog, the ontology editor Prot\`eg\`e should be
-used. Figure~\ref{fig:protege_example} shows the editor while the
+To edit the catalog, the ontology editor Prot\`eg\`e may be
+used. With some experience gained, one is likely to prefer a manual editing.
+Figure~\ref{fig:protege_example} shows the editor while the
 MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has
@@
 main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and
-\texttt{Package}.
+\texttt{Package}. The RDF file is kept in the RDF:XML format, and
+when inspecting the entry, one will fine a direct correspondence 
+to the data stored in RDF:
+
+\begin{verbatim}
+<rdfs:Class rdf:about="&kb;MetaPackage"
+         rdfs:label="MetaPackage">
+  <rdfs:subClassOf rdf:resource="&rdfs;Resource"/>
+  <rdfs:comment>
+   Reference to a piece of software that shall be made available -
+   somehow, and to the expectation to the grid user.  The traditional
+   way to achieve the installation is via tarballs. Debian packages may
+   be an alternative.
+  </rdfs:comment>
+</rdfs:Class>
+<rdfs:Class rdf:about="&kb;Note"
+   rdfs:label="Note">
+   <rdfs:subClassOf rdf:resource="&rdfs;Resource"/>
+</rdfs:Class>
+<rdfs:Class rdf:about="&kb;Package"
+   rdfs:label="Package">
+  <rdfs:subClassOf rdf:resource="&rdfs;Resource"/>
+  <rdfs:comment>
+   Superclass of "TarPackage" and "DebianPackage".
+  </rdfs:comment>
+</rdfs:Class>
+\end{verbatim}
 
@@
 
+Links between resources are referred to as {\em Properties}. The following
+specifies the dependencies that any package may have on other resources, i.e.
+on other packages. The dependency on a base system is expressed by the
+property {\em basesystem}, which not to be confused with the resource
+{\em BaseSystem}.
+
+\begin{verbatim}
+<rdf:Property rdf:about="&kb;depends"
+         rdfs:comment="lists dependencies of this package"
+         rdfs:label="depends">
+        <rdfs:domain rdf:resource="&kb;Package"/>
+        <rdfs:range rdf:resource="&rdfs;Resource"/>
+\end{verbatim}
+
 MetaPackages contain one or more instances of the class
@@
 MetaPackage needs to perform the same functionality.
-MetaPackages are described by the subclasses of \texttt{Note, which
+MetaPackages are described by the subclasses of \texttt{Note}, which
 in turn has two subclasses: \texttt{BaseSystem} and
@@
 offered for manual installation via the prior mentioned traditional page
-representing Runtime Environments for ARC.  The corresponding tar file
-is named \task{somewhat??}. The \texttt{data} directory simply contains
+representing Runtime Environments for ARC.
+The \texttt{data} directory simply contains
 a ZIP file which needs to be unzipped in the installation directory. For

### 2010-02-16 doc/tech_doc/a-rex/arex_tech_doc.tex
add explanation of download checks and retry strategy, and improve formatting

@@
 
-Each job gets an identifier (\textit{jobid}). This is a handle -- WS-Addressing~\cite{ws-addr-soap} XML document -- which identifies the job in the
-A-REX and in the Information Interface.
+Each job gets an identifier (\textit{jobid}). This is a handle --
+WS-Addressing~\cite{ws-addr-soap} XML document -- which identifies the
+job in the A-REX and in the Information Interface.
 
 Jobs are initiated and controlled through the WS interface. Complete
-job descriptions (JD) are passed to the A-REX through WS in JSDL~\cite{jsdl}
-coded description. Input data files and job executables are transferred
-separately through the same interface, as described in Section~\ref{sec:input/output data}.
+job descriptions (JD) are passed to the A-REX through WS in
+JSDL~\cite{jsdl} coded description. Input data files and job
+executables are transferred separately through the same interface, as
+described in Section~\ref{sec:input/output data}.
 
@@
 
-One of the most important tasks of the A-REX is to take care of processing of the
-input and output data (files) of the job. Input files are gathered
-in the SD or in the associated cache area. There are two ways to put a file
-into the SD:
-
-\begin{itemize}
-\item Download is initiated by the A-REX -- This is the case for files defined
-in the JD (with name and source). The A-REX alone is responsible to
-ensure that all required files will be available in the SD.\\
-The supported protocols for sources at the moment are (in case of
-full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
-some less standard sources are supported. Those are described below.
-The A-REX fully relies on HED framework~\cite{hed} for data transferring
-capabilities. Hence actual set of supported protocols depends on installed
-Data Management Components of the HED.
-\item Upload is initiated by the user directly or through the User Interface
-(UI). Because the SD becomes available immediately at the time of
-submission of JD, UI can (and should) use that to upload data files
-which are not otherwise accessible by the A-REX. Examples of such
-files are the main executable of the job, the job's input files, etc.
-These files can (and should) also be specified in the JD.
-\end{itemize}
-\textbf{There is no} other reliable way for a job to obtain input
-data on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and
-any other remote data transport during execution of a job is not guaranteed
-(at least not by A-REX).
+One of the most important tasks of the A-REX is to take care of
+processing of the input and output data (files) of the job. Input
+files are gathered in the SD or in the associated cache area. There
+are two ways to put a file into the SD:
+
+\begin{itemize}
+
+\item Download is initiated by the A-REX -- This is the case for files
+  defined in the JD (with name and source). The A-REX alone is
+  responsible to ensure that all required files will be available in
+  the SD.\\ The supported protocols for sources at the moment are (in
+  case of full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over
+  SSLv3). Some less standard sources and \emph{Indexing Services} are
+  also supported, these are described in Section~\ref{sec:urls}. The
+  A-REX fully relies on the HED framework~\cite{hed} for data
+  transferring capabilities and so the actual set of supported
+  protocols depends on the installed Data Management Components of the
+  HED.
+
+\item Upload is initiated by the user directly or through the User
+  Interface (UI). Because the SD becomes available immediately at the
+  time of submission of the JD, the UI can (and should) use that to
+  upload data files which are not otherwise accessible by the
+  A-REX. Examples of such files are the main executable of the job,
+  the job's input files, etc.  These files can (and should) also be
+  specified in the JD.
+\end{itemize}
+
+\textbf{There is no} other reliable way for a job to obtain input data
+on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and any
+other remote data transport during execution of a job is not
+guaranteed (at least not by A-REX).
+
+At the start of a download, a dummy file with the same name and size
+as the source file is created in the SD to check that enough space is
+available. As the transfer proceeds this dummy file is over-written
+with the real data. If the file is to be cached (see
+Section~\ref{sec:cache}) then the pre-allocation of space is not done,
+as cache space is managed by the A-REX. In the case of caching two
+extra validation checks are also performed: firstly, if the source is
+an Indexing Service then metadata (file size, checksum) reported by
+the Indexing Service and the service hosting the physical replica are
+compared, if available. If they differ then that replica is not
+downloaded. Secondly, on completion of the download the file size is
+compared to that reported by the source - if they differ then the
+download is failed.
 
@@
 \begin{itemize}
-\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
-and optionally registered in some \emph{Indexing Service} like the
+
+\item Files which are supposed to be moved to a \emph{Storage Element}
+  (SE) and optionally registered in some Indexing Service like the
   Globus \emph{Replica Location Service} (RLS) -- The A-REX takes care
-of these files. They have to be specified in the JD. If the job fails
-during any stage of processing, no attempt is made to transfer those
-files to their final destination, unless the option \emph{preserve=yes}
-is specified in their URLs.
-\item Files which are supposed to be fetched by the user -- The user has
-to use a tool like the UI to obtain these files. They \textbf{must}
-also be specified in the JD.
-\end{itemize}
+  of these files. They have to be specified in the JD. If the job
+  fails during any stage of processing, no attempt is made to transfer
+  those files to their final destination, unless the option
+  \emph{preserve=yes} is specified in their URLs.
+
+\item Files which are supposed to be fetched by the user -- The user
+  has to use a tool like the UI to obtain these files. They
+  \textbf{must} also be specified in the JD.
+\end{itemize}
+
 All files not specified in the JD are deleted after job execution
@@
 \begin{itemize}
-\item \textbf{Accepted} -- In this state the job has been submitted to a
-CE but is not processed yet. The A-REX will analyze the JD and move
-to the next stage. If the JD can not be processed the job will be
-canceled and moved to the state \textbf{Finishing}.
-\item \textbf{Preparing} -- The input data is being gathered in the SD (stage-in).
-The A-REX is downloading the files specified in the JD and is waiting
-for files which are supposed to be uploaded by the UI. If all files
-are successfully gathered the job moves to the next state. If \textbf{any}
-file can't be downloaded or it takes the UI too long to upload a file,
-the job moves to \textbf{Finishing} state. It is possible to put a
-limit on the number of simultaneous \textbf{Preparing} jobs. 
-If this limit is exceeded, jobs ready to enter the \textbf{Preparing} state
- will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
-mark. Exceptions are jobs which have no files to be downloaded. These
-are processed out of limits.
-\item \textbf{Submitting} -- The job is being passed for execution to the
-\emph{Local Resource Management System} (LRMS). The corresponding
-backends for many LRMSs are provided with the default installation.
-If the local job submission is successful the job moves to the \textbf{Executing}
-state. Otherwise it moves to \textbf{Finishing}. It is possible to
-limit the aggregate number of jobs in \textbf{Submitting} and \textbf{Executing}
-states.
-\item \textbf{Executing} -- The job is queued or being executed in the LRMS.
-The A-REX takes no actions except waiting until the job finishes.
-\item \textbf{Killing} -- Necessary action to cancel the job in the LRMS
-is being taken.
-\item \textbf{Finishing} -- The output data is being processed (stage-out).
-Specified data files are moved to the specified SEs and are optionally
-registered at an Indexing Service. The user can download data files
-from the SD by using the UI or other adequate tool. All the files
-not specified as output files are removed from the SD at very beginning
-of this state.It is possible to limit the number of simultaneous jobs in this state.
-\item \textbf{Finished} -- No more processing is performed by the A-REX.
-The user can continue to download data files from the SD. The SD is
-kept available for some time (default is 1 week). After that the job
-is moved to the state \textbf{Deleted}. The 'deletion' time can be
-obtained by querying the Information Interface of the A-REX. If a
-job was moved to \textbf{Finished} because of failure, it may be restarted
-on request of a client. When restarted, a job is moved to the state
-previous to the one in which it failed and is assigned mark PENDING.
-This is needed in order to not break the configuration limits. Exception
-is a job failed in \textbf{Executing} state and lacking input files
-specified in JD. Such a job is treated like failed in \textbf{Preparing}
-state.
-\item \textbf{Deleted} -- The job is moved to this state if the user have
-not requested job to be cleaned before the SD's lifetime expires.
-Only minimal subset of information about such job is kept. The SD
-is not available anymore.
-\end{itemize}
+
+\item \textbf{Accepted} -- In this state the job has been submitted to
+  a CE but is not processed yet. The A-REX will analyze the JD and
+  move to the next stage. If the JD can not be processed the job will
+  be canceled and moved to the state \textbf{Finishing}.
+
+\item \textbf{Preparing} -- The input data is being gathered in the SD
+  (stage-in).  The A-REX is downloading the files specified in the JD
+  and is waiting for files which are supposed to be uploaded by the
+  UI. If all files are successfully gathered the job moves to the next
+  state. If \textbf{any} file can't be downloaded or it takes the UI
+  too long to upload a file, the job moves to \textbf{Finishing}
+  state. It is possible to put a limit on the number of simultaneous
+  \textbf{Preparing} jobs.  If this limit is exceeded, jobs ready to
+  enter the \textbf{Preparing} state will stay in the
+  \textbf{Accepted} state, but prefixed with the PENDING:
+  mark. Exceptions are jobs which have no files to be
+  downloaded. These are processed out of limits. If the A-REX fails to
+  download a file due to a temporary error with a remote service, the
+  job will be moved back to the \textbf{Accepted} state, but prefixed
+  with the PENDING: mark, and will wait in that state for some time
+  before being retried. The waiting time increases exponentially with
+  each attempt.
+
+\item \textbf{Submitting} -- The job is being passed for execution to
+  the \emph{Local Resource Management System} (LRMS). The
+  corresponding backends for many LRMSs are provided with the default
+  installation.  If the local job submission is successful the job
+  moves to the \textbf{Executing} state. Otherwise it moves to
+  \textbf{Finishing}. It is possible to limit the aggregate number of
+  jobs in \textbf{Submitting} and \textbf{Executing} states.
+
+\item \textbf{Executing} -- The job is queued or being executed in the
+  LRMS.  The A-REX takes no actions except waiting until the job
+  finishes.
+
+\item \textbf{Killing} -- Necessary action to cancel the job in the
+  LRMS is being taken.
+
+\item \textbf{Finishing} -- The output data is being processed
+  (stage-out).  Specified data files are moved to the specified SEs
+  and are optionally registered at an Indexing Service. The user can
+  download data files from the SD by using the UI or other adequate
+  tool. All the files not specified as output files are removed from
+  the SD at very beginning of this state. It is possible to limit the
+  number of simultaneous jobs in this state.  If the A-REX fails to
+  upload a file due to a temporary error with a remote service, the
+  job will be moved back to the \textbf{Executing} state, but prefixed
+  with the PENDING: mark, and will wait in that state for some time
+  before being retried. The waiting time increases exponentially with
+  each attempt.
+
+\item \textbf{Finished} -- No more processing is performed by the
+  A-REX.  The user can continue to download data files from the
+  SD. The SD is kept available for some time (default is 1
+  week). After that the job is moved to the state
+  \textbf{Deleted}. The 'deletion' time can be obtained by querying
+  the Information Interface of the A-REX. If a job was moved to
+  \textbf{Finished} because of failure, it may be restarted on request
+  of a client. When restarted, a job is moved to the state previous to
+  the one in which it failed and is assigned mark PENDING.  This is
+  needed in order to not break the configuration limits. Exception is
+  a job failed in \textbf{Executing} state and lacking input files
+  specified in JD. Such a job is treated like failed in
+  \textbf{Preparing} state.
+
+\item \textbf{Deleted} -- The job is moved to this state if the user
+  have not requested job to be cleaned before the SD's lifetime
+  expires.  Only minimal subset of information about such job is
+  kept. The SD is not available anymore.
+
+\end{itemize}
+
 In case of a failure special processing is applied to output files.
-All specified output files are treated as \textbf{downloadable by
-the user}. No files will be moved to their destination SE.
+All specified output files are treated as \textbf{downloadable by the
+  user}. No files will be moved to their destination SE.
 

### 2010-02-17 doc/manuals/janitor/tex_maintenance/maintenance.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.

@@
 
-\section{Introducing new packages}\label{sec:catalog}\label{sec:introducing_new_packages}
+\section{Introducing new packages}\label{sec:introducing_new_packages}
 
@@
 
-The WEKA package for machine learning~\cite{FRANK_20074} and the Java
+The WEKA package for machine learning~\cite{FRANK_2004} and the Java
 Runtime Environment are available as dynamic Runtime Environments. Further

### 2010-02-18 doc/tech_doc/a-rex/arex_tech_doc.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
add info on running a-rex under GridFTP interface

@@
 %
-\bibliographystyle{IEEEtran}                       %a nice bibliography style
+\usepackage[numbers]{natbib}
+\bibliographystyle{plainnat}
 %
@@
 
-%-------- Change this to NORDUGRID-XXXXXXX-NN
-
 {\raggedleft NORDUGRID-TECH-14\par}
@@
 
-%\tableofcontents                          %Comment if use article style
+\tableofcontents                          %Comment if use article style
 \newpage
@@
 
-A-REX is one of ARC middleware components that implements functions
+The A-REX is an ARC middleware component that implements functions
 of the so-called \emph{Computing Element} (CE). Here Computing Element
-is a service accepting requests containing description of generic
-computational job and executing it in the underlying local batch system.
-
-It takes care of job pre- and post-processing. This comprises
-stage-in of files containing input data or program modules from wide
-range of sources and transfer or storing of output results.
-
-A-REX implements a Web Service (WS) interface which provides a
-way to submit and control computational tasks (jobs) to be executed
-by A-REX and the underlying Local Resource Management System.
+is a service accepting requests containing a description of generic
+computational jobs and executing it in the underlying local batch system.
+
+It takes care of job pre- and post-processing, i.e.\ stage-in of
+files containing input data or program modules from a wide range of
+sources and transfer or storing of output results.
+
+In previous versions of ARC, computational tasks (jobs) were submitted
+to a resource through a GridFTP service running on the CE, and
+processed by a Grid Manager (GM). The A-REX replaces both these
+components, by implementing a Web Service (WS) interface which
+provides a way to submit jobs, and a revised version of the GM to
+process jobs and interact with the underlying Local Resource
+Management System. The A-REX can also be set up to process jobs
+submitted through the traditional ARC GridFTP service, as well as or
+instead of those using the WS interface. Unless stated otherwise, the
+remainder of this document assumes running only the WS interface.
 
 \begin{framed}You should use this document for advanced configuration purposes
-and understanding of internals of the aforementioned tools. For general
-installation and configuration of whole system refer to other documents
-available at \url{http://www.nordugrid.org/papers.html}.
+and understanding of the internals of the aforementioned tools. For
+general installation and configuration of the whole system please
+refer to other documents available at
+\url{http://www.nordugrid.org/papers.html}.
 \end{framed}
@@
 exchange of data between client and A-REX (including also program
-modules) is performed via HTTP(S). An URL for accessing input/output
+modules) is performed via HTTP(S). A URL for accessing input/output
 files is obtained through the WS Local Information Description
@@
 
-At the start of a download, a dummy file with the same name and size
-as the source file is created in the SD to check that enough space is
-available. As the transfer proceeds this dummy file is over-written
-with the real data. If the file is to be cached (see
-Section~\ref{sec:cache}) then the pre-allocation of space is not done,
-as cache space is managed by the A-REX. In the case of caching two
-extra validation checks are also performed: firstly, if the source is
-an Indexing Service then metadata (file size, checksum) reported by
-the Indexing Service and the service hosting the physical replica are
-compared, if available. If they differ then that replica is not
-downloaded. Secondly, on completion of the download the file size is
-compared to that reported by the source - if they differ then the
-download is failed.
+At the start of a download initiated by the A-REX, a dummy file with
+the same name and size as the source file is created in the SD to
+check that enough space is available. As the transfer proceeds this
+dummy file is over-written with the real data. If the file is to be
+cached (see Section~\ref{sec:cache}) then the pre-allocation of space
+is not done, as cache space is managed by the A-REX. In the case of
+caching two extra validation checks are also performed: firstly, if
+the source is an Indexing Service then metadata (file size, checksum)
+reported by the Indexing Service and the service hosting the physical
+replica are compared, if available. If they differ then that replica
+is not downloaded. Secondly, on completion of the download the file
+size is compared to that reported by the source - if they differ then
+the download is failed.
 
@@
 Figure~\ref{job states diagram} presents a diagram of the possible
-states of a job.%
+states of a job.
 
@@
 
-\item \textbf{Executing} -- The job is queued or being executed in the
+\item \textbf{Executing (InLRMS)} -- The job is queued or being executed in the
   LRMS.  The A-REX takes no actions except waiting until the job
@@
 
-\item \textbf{Killing} -- Necessary action to cancel the job in the
+\item \textbf{Killing (Canceling)} -- Necessary action to cancel the job in the
   LRMS is being taken.
@@
 
-In case of a failure special processing is applied to output files.
+In case of failure, special processing is applied to output files.
 All specified output files are treated as \textbf{downloadable by the
@@
 
-In full installation the A-REX and it's components support the following
-data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
-http, https, lfc, rls} and \emph{srm.} For more information please see
-{}``The Hosting Environment of the Advanced Resource Connector middleware''
-document~\cite{hed}.
+In a full installation, the A-REX and its components support the
+following data transfer protocols and corresponding URLs: \emph{ftp,
+  gsiftp, http, https, lfc, rls} and \emph{srm.} For more information
+please see ``The Hosting Environment of the Advanced Resource
+Connector middleware'' document~\cite{hed}.
 
@@
 about jobs belonging to that user. Multiple users can share the same
-\textit{control directory}. In most common configuration case A-REX
+\textit{control directory}. In the most common configuration case, the A-REX
 serves all users defined by the Operating System and stores their control
@@
 and other information provided by GNU \emph{time} utility. Note that
-some implementation of \emph{time} insert unrequested information
+some implementations of \emph{time} insert unrequested information
 in their output. Hence some lines can have broken format.
-\item \textit{job.ID.proxy} -delegated X509 credentials.
+\item \textit{job.ID.proxy} -- delegated X509 credentials.
 \item \textit{job.ID.proxy.tmp} -- temporary X509 credentials with different
 UNIX ownership used by processes run with effective \emph{user id}
-different form job owner's \emph{id}.
+different from job owner's \emph{id}.
 \end{itemize}
@@
 The \emph{MigrateActivity} operation generates a request to migrate a 
-grid job from another ARC1 cluster, i.e. the operation will get input 
+grid job from another A-REX, i.e. the operation will get input 
 files and possibly job description from the cluster currently holding 
@@
 \begin{itemize}
-\item \textit{control directory} -- In this directory A-REX stores an information
+\item \textit{control directory} -- In this directory A-REX stores information
 about accepted jobs. It must have full permissions there.\\
@@
 
-Due to historical reasons configuration of the A-REX is split into 2
-parts. The small part is located inside HED configuration (see
-Appendix~\ref{annex:arex-conf} for schema and description of supported
-elements). This configuration refers to the file containing the main
-part of the configuration. The default location of the main
-configuration file is \textit{/etc/arc.conf}.
+Due to historical reasons, configuration of the A-REX is split into 2
+parts: HED configuration for the WS interface and a legacy
+configuration file for the rest. For more information on the HED
+configuration see instructions in Section~\ref{sec:wssetup}, and
+Appendix~\ref{annex:arex-conf} for a full schema and description of
+supported elements. The HED configuration refers to the legacy file,
+which is identical to the configuration file used for the GridFTP/Grid
+Manager services in previous versions of ARC. The default location of
+this legacy file is \textit{/etc/arc.conf}. If the A-REX is run
+without the WS interface, then only the legacy file is needed (to
+configure the GridFTP interface see~\cite{gm}). This section describes
+the content of the legacy file relating to running the A-REX.
 
@@
 
-The A-REX is installed as an ARC1 component of ARC middleware and is available
-in NorduGrid download area at ARC1 Download \url{http://download.nordugrid.org/software/nordugrid-arc1/}.There are packages available for various Linux distributions. The
-A-REX comes in the \texttt{nordugrid-arc1-arex} package. Source code ready for
-compilation is available too.
+The A-REX is installed as a component of the ARC middleware and
+packages for various distributions are available from NorduGrid
+repositories, or directly from the download area of the NorduGrid
+website. Source code ready for compilation is available too.
 
@@
 
-\subsection{Setup of the A-REX}
+\subsection{Setup of the A-REX with WS Interface}
+\label{sec:wssetup}
 
@@
 
-\subsection{Usage}
-
 For a quick start, simply run
@@
 For more information please read the \emph{User Guide}~\cite{userguide1}.
+
+\subsection{Setup of the A-REX with GridFTP Interface}
+
+The A-REX can process jobs submitted through a traditional GridFTP
+server instead of, or as well as, the WS interface. The GridFTP/Grid
+Manager server packages can be installed alongside A-REX packages, so
+it is easy to switch between the two. The configuration of the GridFTP
+server is in the same legacy \emph{arc.conf} configuration file as is
+used by the A-REX, and configuration instructions are given
+in~\cite{gm}. In general no configuration changes are necessary to use
+a Grid Manager-configured GridFTP server with the A-REX.
+
+If it is desired to run both WS and GridFTP interfaces at the same
+time, simply setup and start the A-REX as described in the previous
+section, then start the GridFTP server. If only the GridFTP interface
+is to be used, then the special \emph{a-rex} command should be used
+(rather than \emph{arched}).
+
+\begin{shaded}
+\begin{verbatim}
+$ARC_LOCATION/etc/init.d/a-rex start
+\end{verbatim}
+\end{shaded}
+
+This command automatically creates a HED configuration for running the
+A-REX without the WS interface, and so no manual HED configuration
+needs to be done. The GridFTP server should then be started.
+
+\begin{framed}
+IMPORTANT: Do not run the A-REX and Grid Manager services at the same
+time, as they will interfere with each other with unpredictable
+results.
+\end{framed}
 

### 2010-03-01 doc/tech_doc/a-rex/arex_tech_doc.tex
Limited shares explained in A-REX technical documentation

@@
 for possible values and more details.
+\item \textbf{\textit{share\_limit}}\textit{=name limit} -- specifies a transfer
+share that has a number of processes different from the default value in
+maxloadshare. \emph{name} is the name of the share and \emph{limit} is the 
+number of processes for this share. In the configuration should appear after 
+maxloadshare. Can be repeated several times for different shares.
+See Section~\ref{sub:Transfershare} for how to compose shares' names and more
+details.
 \item \textbf{\textit{wakeupperiod}}\textit{=time} -- specifies how often
@@
 
+It's possible to distinguish some transfer shares and assign them a limit
+different from what's specified in \emph{maxloadshare}. It's done by
+\emph{share\_limit} option. \emph{share\_limit} can only be used if 
+\emph{maxloadshare} has been already set before. Depending on the 
+sharing mechanism used by \emph{maxloadshare}, the proper name for the share
+should be specified, as illustrated by the following examples (note, that
+in dn case spaces are allowed, the configuration parser will take care of them):
+
+\begin{itemize}
+\item \emph{dn}: /O=Grid/O=NorduGrid/OU=domainname.com/CN=Jane Doe
+\item \emph{voms:vo}: voname
+\item \emph{voms:role}: voname:rolename
+\item \emph{voms:group}: /voname/groupname
+\end{itemize}
+
+The specific shares, specified in \emph{share\_limit}, are processed
+differently from the other shares. A-REX reserves an indicated number of 
+processes for each specific share. The number of unreserved processes is then
+split evenly between the ordinary shares, as determined by \emph{maxloadshare}. 
+So the specific shares have a strict, non-decreaseable limit, unlike all the 
+ordinary shares, whose limit can be decreased while A-REX tries to split the 
+load evenly. However, A-REX reserves processes only for active 
+specific shares, i.e. shares to which at least one active job on the resource 
+belongs to. If the share is not active, its slots are used in overall splitting
+between ordinary transfer shares.
+
+A particular case is when A-REX reserves more processes than specified in
+\emph{maxjobs}. A-REX will process jobs from specific share at FIFO-basis and
+stop at reaching \emph{maxjobs} number of processes, even if some specific
+shares haven't reached their limits. Also in this situation each ordinary
+share is allowed to launch only one upload and download process.
+
 If VOMS is not supported, the \emph{dn} scheme is the only option that
@@
 and \emph{loadShareType} (the scheme used) elements inside 
-\emph{loadLimits} block.
+\emph{loadLimits} block. For defining the specific shares, \emph{shareLimit}
+sub-blocks with \emph{name} and \emph{limit} elements can be used after 
+\emph{maxLoadShare}.
 

### 2010-03-03 doc/tech_doc/a-rex/arex_tech_doc.tex
add support for multiple session directories

@@
 the JD).
-\item \textbf{\textit{sessiondir}}\textit{=path} -- specifies path to
-  the directory in which the SD is created. If \emph{path} is {*} the
-  default sessiondir is used -- \textit{\$HOME/.jobs}.
+\item \textbf{\textit{sessiondir}}\textit{=path [drain]} - specifies
+  the path to the directory in which the SD is created. Multiple
+  session directories may be specified by specifying multiple
+  \emph{sessiondir} commands. In this case jobs are spread evenly over
+  the session directories. If the path is {*} the default sessiondir
+  is used - \textit{\$HOME/.jobs}. When adding a new session
+  directory, ensure to restart the A-REX so that jobs assigned there
+  are processed. A session directory can be drained prior to removal
+  by adding the \emph{``drain''} option (no restart is required in
+  this case). No new jobs will be assigned to this session directory
+  but running jobs will still be accessible. When all jobs are
+  processed and the session directory is empty, it can be removed and
+  the A-REX should be restarted.
 \item \textbf{\textit{cachedir}}\textit{=path {[}link\_path]} --

### 2010-03-03 src/services/a-rex/grid-manager/conf/conf_pre.cpp

@@ -22,7 +22,8 @@
 Arc::Logger& logger = Arc::Logger::getRootLogger();
 
 bool configure_user_dirs(const std::string &my_username,
-                std::string &control_dir,std::string &session_root,
+                std::string &control_dir,std::vector<std::string> &session_roots,
+                std::vector<std::string> &session_roots_non_draining,
                 std::string &default_lrms,std::string &default_queue,
                 std::list<std::string>& queues,
                 ContinuationPlugins &plugins,RunPlugin &cred_plugin,
@@ -135,7 +136,23 @@ bool configure_user_dirs(const std::string &my_username,
         };
         if(user_match) {
           std::string control_dir_ = tmp_node["controlDir"];
-          std::string session_root_ = tmp_node["sessionRootDir"];
+          Arc::XMLNode session_node = tmp_node["sessionRootDir"];
+          for (;session_node; ++session_node) {
+            std::string session_root = session_node;
+            if(session_root.empty()) {
+              logger.msg(Arc::ERROR,"sessionRootDir is missing"); return false;
+            };
+            std::string::size_type i = session_root.find(' ');
+            if (i != std::string::npos) {
+              if (session_root.substr(i+1) != "drain") {
+                session_roots_non_draining.push_back(session_root.substr(0, i));
+              }
+              session_root = session_root.substr(0, i);
+            } else {
+              session_roots_non_draining.push_back(session_root);
+            }
+            session_roots.push_back(session_root);
+          }
           if(username == ".") username = "";
           JobUser user(username);
           if(!user.is_valid()) { config_close(cfile); return false; };
@@ -143,10 +160,17 @@ bool configure_user_dirs(const std::string &my_username,
           elementtobool(tmp_node,"noRootPower",strict_session,&logger);
           user.SetLRMS(default_lrms,default_queue);
           user.substitute(control_dir_);
-          user.substitute(session_root_);
-          user.SetSessionRoot(session_root_);
           user.SetControlDir(control_dir_);
-          session_root=user.SessionRoot();
+          for(std::vector<std::string>::iterator i = session_roots_non_draining.begin(); i != session_roots_non_draining.end(); i++) {
+            user.substitute(*i);
+          }
+          user.SetSessionRoot(session_roots_non_draining);
+          session_roots_non_draining = user.SessionRoots();
+          for(std::vector<std::string>::iterator i = session_roots.begin(); i != session_roots.end(); i++) {
+            user.substitute(*i);
+          }
[…]

### 2010-03-03 src/services/a-rex/grid-manager/conf/conf_pre.h

@@ -23,7 +23,8 @@ class RunPlugin;
     look description of grid-manager and configuration template.
 */
 bool configure_user_dirs(const std::string &my_username,
-                std::string &control_dir,std::string &session_root,
+                std::string &control_dir,std::vector<std::string> &session_roots,
+                std::vector<std::string> &session_roots_non_draining,
                 std::string &default_lrms,std::string &default_queue,
                 std::list<std::string>& queues,
                 ContinuationPlugins &plugins,RunPlugin& cred,

### 2010-03-03 src/services/a-rex/grid-manager/files/info_files.cpp

@@ -385,7 +385,7 @@ bool job_diagnostics_mark_move(const JobDescription &desc,JobUser &user) {
   if(h2==-1) return false;
   fix_file_owner(fname2,desc,user);
   fix_file_permissions(fname2,user);
-  std::string fname1 = user.SessionRoot() + "/" + desc.get_id() + sfx_diag;
+  std::string fname1 = user.SessionRoot(desc.get_id()) + "/" + desc.get_id() + sfx_diag;
   if(user.StrictSession()) {
     JobUser tmp_user(user.get_uid()==0?desc.get_uid():user.get_uid());
     job_file_read_t arg; arg.h=h2; arg.fname=&fname1;
@@ -998,10 +998,10 @@ bool job_clean_deleted(const JobDescription &desc,JobUser &user,std::list<std::s
   fname = user.ControlDir()+"/job."+id+".input"; remove(fname.c_str());
   fname = user.ControlDir()+"/job."+id+".rte"; remove(fname.c_str());
   fname = user.ControlDir()+"/job."+id+".grami_log"; remove(fname.c_str());
-  fname = user.SessionRoot()+"/"+id+sfx_lrmsoutput; remove(fname.c_str());
+  fname = user.SessionRoot(id)+"/"+id+sfx_lrmsoutput; remove(fname.c_str());
   /* remove session directory */
   std::list<FileData> flist;
-  std::string dname = user.SessionRoot()+"/"+id;
+  std::string dname = user.SessionRoot(id)+"/"+id;
   if(user.StrictSession()) {
     JobUser tmp_user(user.get_uid()==0?desc.get_uid():user.get_uid());
     job_dir_remove_t arg; arg.dname=&dname; arg.flist=&flist;

### 2010-03-03 src/services/a-rex/grid-manager/jobs/states.cpp

@@ -98,7 +98,7 @@ bool JobsList::AddJobNoCheck(const JobId &id,uid_t uid,gid_t gid){
 }
 
 bool JobsList::AddJobNoCheck(const JobId &id,JobsList::iterator &i,uid_t uid,gid_t gid){
-  i=jobs.insert(jobs.end(),JobDescription(id,user->SessionRoot() + "/" + id));
+  i=jobs.insert(jobs.end(),JobDescription(id,user->SessionRoot(id) + "/" + id));
   i->keep_finished=user->KeepFinished();
   i->keep_deleted=user->KeepDeleted();
   i->set_uid(uid,gid);
@@ -110,7 +110,7 @@ bool JobsList::AddJob(const JobId &id,uid_t uid,gid_t gid){
   if(FindJob(id) != jobs.end()) return false;
   logger.msg(Arc::INFO,"%s: Added",id);
   iterator i=jobs.insert(jobs.end(),
-         JobDescription(id,user->SessionRoot() + "/" + id));
+         JobDescription(id,user->SessionRoot(id) + "/" + id));
   i->keep_finished=user->KeepFinished();
   i->keep_deleted=user->KeepDeleted();
   i->set_uid(uid,gid);

### 2010-03-03 src/services/a-rex/grid-manager/jobs/users.h

@@ -43,8 +43,8 @@ class JobUser {
  private:
   /* directory where files explaining jobs are stored */
   std::string control_dir;
-  /* directory where directories used to run jobs are created */
-  std::string session_root;
+  /* directories where directories used to run jobs are created */
+  std::vector<std::string> session_roots;
   /* cache information */
   CacheConfig * cache_params;
   /* default LRMS and queue to use */
@@ -88,6 +88,7 @@ class JobUser {
   /* Set and get corresponding private values */
   void SetControlDir(const std::string &dir);
   void SetSessionRoot(const std::string &dir);
+  void SetSessionRoot(const std::vector<std::string> &dirs);
   void SetCacheParams(CacheConfig* params);
   void SetLRMS(const std::string &lrms_name,const std::string &queue_name);
   void SetKeepFinished(time_t ttl) { keep_finished=ttl; };
@@ -99,7 +100,8 @@ class JobUser {
   bool CreateDirectories(void);
   bool is_valid(void) { return valid; };
   const std::string & ControlDir(void) const { return control_dir; };
-  const std::string & SessionRoot(void) const { return session_root; };
+  const std::string & SessionRoot(std::string job_id = "") const;
+  const std::vector<std::string> & SessionRoots() const { return session_roots; };
   CacheConfig * CacheParams(void) const { return cache_params; };
   bool CachePrivate(void) const { return false; };
   const std::string & DefaultLRMS(void) const { return default_lrms; };
@@ -147,7 +149,7 @@ class JobUsers {
   const_iterator end(void) const { return users.end(); };
   size_t size(void) const { return users.size(); };
   /* Create new user object and push into list */
-  iterator AddUser(const std::string &unix_name,RunPlugin* cred_plugin = NULL,const std::string &control_dir = "",const std::string &session_root = "");
+  iterator AddUser(const std::string &unix_name,RunPlugin* cred_plugin = NULL,const std::string &control_dir = "",const std::vector<std::string> *session_root = NULL);
   iterator DeleteUser(iterator &user) { return users.erase(user); };
   /* Check existance of user of given unix name */
   bool HasUser(const std::string &name) {

### 2010-03-03 src/services/a-rex/grid-manager/run/run_parallel.cpp

@@ -54,7 +54,7 @@ bool RunParallel::run(JobUser& user,const JobDescription& desc,const char *const
     JobUser tmp_user(desc.get_uid());
     if(!tmp_user.is_valid()) return false;
     tmp_user.SetControlDir(user.ControlDir());
-    tmp_user.SetSessionRoot(user.SessionRoot());
+    tmp_user.SetSessionRoot(user.SessionRoot(desc.get_id()));
     return run(tmp_user,desc.get_id().c_str(),args,ere,su,
                                         true,cred,&job_subst,&subs);
   };

### 2010-03-11 doc/tech_doc/a-rex/arex_tech_doc.tex
add cache lifetime feature, part of fix for 1851

@@
 \item \emph{cache-list} - This tool is used to list all files present
-  in each cache. It simply reads through all the \emph{.meta }files
-  and prints to stdout a list of all URLs stored in each cache and
-  their corresponding cache filename, one per line.
+  in each cache or, given a list of URLs as arguments, shows the
+  location of each URL in the cache if present. In the first case it
+  simply reads through all the \emph{.meta }files and prints to stdout
+  a list of all URLs stored in each cache and their corresponding
+  cache filename, one per line. In the second case the cache filename
+  of each URL is calculated and then each cache is checked for the
+  existence of the file.
 \end{itemize}
@@
   \emph{cachedir} commands.
+\item \textbf{\textit{cachelifetime}}\textit{=lifetime} - if cache
+  cleaning is enabled, files accessed less recently than the
+  \emph{lifetime} time period will be deleted. Example values of this
+  option are 1800, 90s, 24h, 30d. When no suffix is given the unit is
+  seconds.
 \item \textbf{\textit{cacheloglevel}}\textit{=number} -- specified the

### 2010-03-11 src/services/a-rex/grid-manager/grid_manager.cpp

@@ -30,7 +30,7 @@
 /* do job cleaning every 2 hours */
 #define HARD_JOB_PERIOD 7200
 
-/* cache cleaning and registration every 5 minutes */
+/* cache cleaning every 5 minutes */
 #define CACHE_CLEAN_PERIOD 300
 
 #define DEFAULT_LOG_FILE "/var/log/grid-manager.log"
@@ -64,16 +64,18 @@ static void* cache_func(void* arg) {
         
         // get the cache dirs
         std::vector<std::string> cache_info_dirs = cache_info->getCacheDirs();
-
+        if (cache_info_dirs.empty()) continue;
+        
         // in arc.conf % of used space is given, but cleanbyage uses % of free space
         std::string minfreespace = Arc::tostring(100-cache_info->getCacheMax());
         std::string maxfreespace = Arc::tostring(100-cache_info->getCacheMin());  
-
+        std::string cachelifetime = cache_info->getLifeTime();
+        
         // set log file location - controldir/job.cache-clean.errors
         // TODO: use GM log?
         gmuser.SetControlDir(cacheuser->ControlDir()); // Should this requirement be removed ?
         int argc=0;
-        char* args[7+cache_info_dirs.size()+1];
+        char* args[9+cache_info_dirs.size()+1];
         
         // do cache-clean -h for explanation of options
         std::string cmd = nordugrid_libexec_loc() + "/cache-clean";
@@ -82,6 +84,8 @@ static void* cache_func(void* arg) {
         args[argc++]=(char*)minfreespace.c_str();
         args[argc++]=(char*)"-M";
         args[argc++]=(char*)maxfreespace.c_str();
+        args[argc++]=(char*)"-E";
+        args[argc++]=(char*)cachelifetime.c_str();        
         args[argc++]=(char*)"-D";
         args[argc++]=(char*)cache_info->getLogLevel().c_str();
         std::vector<std::string> cache_dirs;

### 2010-03-19 doc/manuals/configuration/configuration_manual.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
Minor corrections.

@@
 A security layer can be configured with the TLS MCC, named \texttt{tls.service},
-and it should be put on top of a TCP MCC. By default it is able to understand
+which is based on the OpenSSL library~\cite{openssl}, and it should be put on
+top of a TCP MCC. By default it is able to understand
 the SSLv2 (hadshake only), SSLv3 and TLSv1 protocols, however the \texttt{Handshake}
@@
 \section{The \texttt{arched} daemon}\label{sec:arched}
-The \texttt{arched} daemon is the daemon which is used to start every service in
-WS-ARC. It will load a configuration file, and start the services specified
+The \texttt{arched} daemon is the daemon which is used to host every service in
+WS-ARC. It will load a configuration file, and host the services specified
 herein.

### 2010-03-19 doc/manuals/configuration/configuration_manual.tex
Removing schemas from configuration manual, as these can now be found at http://www.nordugrid.org/schemas/.
Specifying default values for configuration options.

@@
 
-\begin{tabular}{>{\ttfamily}l>{\ttfamily}l}
+\begin{center}
+\small
+\begin{tabular}{>{\ttfamily}l>{\ttfamily}l>{\ttfamily}l}
 \toprule
-\textrm{\textbf{XML element/attribute}} & \textrm{\textbf{INI tag}} \\
+\textrm{\textbf{XML element/attribute}} & \textrm{\textbf{INI tag}} & \textrm{\textbf{Default value}}\\
 \midrule
-/Server/PidFile & pidfile \\
-/Server/User & user \\
-/Server/Group & group \\
-/Server/Logger/File & logfile \\
-/Server/Logger/Level & loglevel \\
-/Server/Logger/Maxsize & logmaxsize \\
-/Server/Logger/Backups & logbackups \\
+/Server/PidFile & pidfile & /var/run/arched.pid\\
+/Server/User & user & \textrm{None} \\
+/Server/Group & group & \textrm{None} \\
+/Server/Logger/File & logfile & /var/log/arched.log \\
+/Server/Logger/Level & loglevel & WARNING \\
+/Server/Logger/Maxsize & logmaxsize & \textrm{None} \\
+/Server/Logger/Backups & logbackups & \textrm{None} \\
+/ModuleManager/Path & libpath & \textrm{None} \\
+/Plexer/next & endpointpattern & \textrm{None} \\
 \midrule
-\multicolumn{2}{c}{\textbf{TCP MCC}}\\
+\multicolumn{3}{c}{\textbf{TCP MCC}}\\
 \midrule
-/Component/Listen/Interface & interface \\
-/Component/Listen/Port & port \\
-/Component/Listen/Version & ipversion \\
-/Component/Listen/NoDelay & nodelay \\
-/Component/Listen/Timeout & timeout \\
-/Component/Limit & limit \\
-/Component/Limit.drop & drop \\
+/Component/Listen/Interface & interface & 0.0.0.0 \\
+/Component/Listen/Port & port & 60000\\
+/Component/Listen/Version & ipversion & \textrm{None} \\
+/Component/Listen/NoDelay & nodelay & false\\
+/Component/Listen/Timeout & timeout & \textrm{None}\\
+/Component/Limit & limit & \textrm{None}\\
+/Component/Limit.drop & drop & \textrm{None}\\
 \midrule
-\multicolumn{2}{c}{\textbf{TLS MCC}}\\
+\multicolumn{3}{c}{\textbf{TLS MCC}}\\
 \midrule
-/Component/KeyPath & x509\_user\_key \\
-/Component/CertificatePath & x509\_user\_cert \\
-/Component/CACertificatePath & x509\_cacert\_path \\
-/Component/CACertificatesDir & x509\_cacert\_dir \\
-/Component/CACertificatesDir.PolicyGlobus & policyglobus \\
-/Component/ClientAuthn & clientauthn \\
-/Component/VOMSCertTrustDNChain/VOMSCertTrustDN &  \\
-/Component/VOMSCertTrustDNChain/VOMSCertTrustRegex &  \\
-/Component/VOMSCertTrustDNChainsLocation & \\
-/Component/Handshake & handshake \\
+/Component/KeyPath & x509\_host\_key & /etc/grid-security/hostkey.pem\\
+/Component/CertificatePath & x509\_host\_cert & /etc/grid-security/hostcert.pem\\
+/Component/CACertificatePath & x509\_cacert\_path & \textrm{None}\\
+/Component/CACertificatesDir & x509\_cacert\_dir & /etc/grid-security/certificates\\
+%/Component/CACertificatesDir.PolicyGlobus & policyglobus \\
+/Component/ClientAuthn & clientauthn & true\\
+%/Component/VOMSCertTrustDNChain/VOMSCertTrustDN &  \\
+%/Component/VOMSCertTrustDNChain/VOMSCertTrustRegex &  \\
+%Component/VOMSCertTrustDNChainsLocation & \\
+/Component/Handshake & handshake & TLS\\
 \bottomrule
 \end{tabular}
-
- \lstset{
-         basicstyle=\footnotesize\ttfamily,
- }
-
-\section{XML Schemas}
-\subsection{General Daemon Schema}
-\lstinputlisting[language=xml]{base.xsd}
-
-\subsection{Loader Schema}
-\lstinputlisting[language=xml]{loader.xsd}
-
-\subsection{TCP MCC Schema}
-\lstinputlisting[language=xml]{tcp.xsd}
-
-\subsection{TLS MCC Schema}
-\lstinputlisting[language=xml]{tls.xsd}
-
-\subsection{GSI MCC Schema}
-\lstinputlisting[language=xml]{gsi.xsd}
-
-\subsection{ARC Authorization Security Handler Schema}
-\lstinputlisting[language=xml]{ArcAuthZ.xsd}
-
-\subsection{UsernameToken Security Handler Schema}
-\lstinputlisting[language=xml]{UsernameTokenSH.xsd}
-
-\subsection{X.509 Token Security Handler Schema}
-\lstinputlisting[language=xml]{X509TokenSH.xsd}
-
-\subsection{SAML Token Security Handler Schema}
-\lstinputlisting[language=xml]{SAMLTokenSH.xsd}
-
-\subsection{ARC PDP Schema}
-\lstinputlisting[language=xml]{ArcPDP.xsd}
-
-\subsection{XACML PDP Schema}
-\lstinputlisting[language=xml]{XACMLPDP.xsd}
-
-\subsection{Delegation PDP Schema}
-\lstinputlisting[language=xml]{DelegationPDP.xsd}
-
-\subsection{Service Invoker PDP Schema}
-\lstinputlisting[language=xml]{PDPServiceInvoker.xsd}
-
-\subsection{Simple List PDP Schema}
-\lstinputlisting[language=xml]{SimpleListPDP.xsd}
+\end{center}
 

### 2010-03-28 doc/tech_doc/a-rex/arex_tech_doc.tex
Removing direct references to Globus from A-REX. Related to #1904.

@@
     {[}...]]} -- This option initiates UNIX user as being serviced by
-  the A-REX.  \textit{path} refers to the control directory (see
+  the A-REX. The \textit{path} refers to the control directory (see
   Section~\ref{section:internals} for the description of control
   directory). If the path is {*} the default one is used --
-  \verb|$HOME/.jobstatus|. \textit{username} stands for UNIX name of
-  the local user. Multiple names can be specified.  If the name is {*}
-  it is substituted by all names found in file
-  \texttt{/etc/grid-security/grid-mapfile} (for the format of this
-  file one should study the Globus project~\cite{globus}).\\ Also the
+  \texttt{\$HOME/.jobstatus}. The \textit{username} stands for UNIX name of
+  the local user. Multiple names can be specified. If the name starts from
+  {@} rest is treated as path to file containing list serviced users.
+  Usernames are specified one per line and may be optionally prepended
+  with Grid identity of user - last one is ignored. That is done for
+  compatibility with so-called grid-mapfile (for more information 
+  please see the description of Globus project~\cite{globus}). Also the
   special name '.'(dot) can be used. Corresponding control directory
@@
   one in the configuration file. There is also command
-  \textbf{\textit{controldir}}\textit{=path}.  It uses special
+  \textbf{\textit{controldir}}\textit{=path}.  It presumes special
   username '.' and is always executed last independent of its
@@
 \item [{\%C}] -- control dir -- see command \emph{control}
-\item [{\%U}] -- username
+\item [{\%U}] -- username (as specified in configuration, hence empty for
+'.' control directories)
 \item [{\%u}] -- userid -- numerical
 \item [{\%g}] -- groupid -- numerical
-\item [{\%H}] -- home dir -- home specified in /etc/passwd
+\item [{\%H}] -- home dir -- home of username as specified in /etc/passwd
 \item [{\%Q}] -- default queue -- see command \emph{lrms}
@@
 \item [{\%W}] -- installation path -- \$\{ARC\_LOCATION\}
-\item [{\%G}] -- globus path -- \$\{GLOBUS\_LOCATION\}
 \item [{\%c}] -- list of all control directories
@@
 \item joboption\_jobname -- name of the job as given by user.
-\item joboption\_rsl -- whole RSL for very clever submission scripts.
-\item joboption\_rsl\_\emph{name} -- RSL attributes and values (like joboption\_rsl\_executable=''/bin/echo'')
 \end{itemize}
@@
 
-\section{Error messages of A-REX\label{annex:arex-errors}}
+\section{Error messages of A-REX (outdated)\label{annex:arex-errors}}
 
@@
 Could not process RSL&
-Job description could not be processed to syntax errors or missing
+Job description could not be processed due to syntax errors or missing
 elements\tabularnewline

### 2010-03-28 src/services/a-rex/grid-manager/conf/environment.h

@@ -30,10 +30,6 @@ std::string operator+(const char*,const prstring&);
 std::string operator+(const std::string&,const prstring&);
 */
 
-/// Globus installation path - $GLOBUS_LOCATION, /opt/globus
-std::string globus_loc(void);
-// Various Globus scripts - $GLOBUS_LOCATION/libexec
-std::string globus_scripts_loc(void);
 /// ARC installation path - $ARC_LOCATION, executable path
 std::string nordugrid_loc(void);
 /// ARC system tools - $ARC_LOCATION/libexec/arc, $ARC_LOCATION/libexec
@@ -58,9 +54,6 @@ void runtime_config_dir(const std::string&);
 /// grid.manager@hostname, it can also be set from configuration file 
 std::string support_mail_address(void);
 void support_mail_address(const std::string&);
-/// Global gridmap files with welcomed users' DNs and UNIX names
-/// $GRIDMAP, default /etc/grid-security/grid-mapfile
-std::string globus_gridmap(void);
 
 ///  Read environment, check files and set variables
 ///  Accepts:

### 2010-03-28 src/services/a-rex/grid-manager/conf/gridmap.h

@@ -5,14 +5,14 @@
 #include <list>
 
 /*
-  Read gridmap file (specified by globus_gridmap global variable).
+  Read file specified by path argument.
   Returns:
     true - success
     false - error (most probaly file is missing)
-    'ulist' contains unix usernames found in gridmap file separted by
-    blank spaces.
+    'ulist' contains unix usernames found in file - one per line or
+    in gridmap-like format - separted by blank spaces.
 */
-bool gridmap_user_list(std::string &ulist);
-bool gridmap_user_list(std::list<std::string> &ulist);
+bool file_user_list(const std::string& path,std::string &ulist);
+bool file_user_list(const std::string& path,std::list<std::string> &ulist);
 
 #endif // __GM_GRIDMAP_H__

### 2010-03-28 src/services/a-rex/grid-manager/jobs/users.cpp

@@ -168,7 +168,9 @@ bool JobUser::substitute(std::string& param) const {
       case 'u': to_put=Arc::tostring(get_uid()); break;
       case 'g': to_put=Arc::tostring(get_gid()); break;
       case 'W': to_put=nordugrid_loc(); break;
-      case 'G': to_put=globus_loc(); break;
+      case 'G': 
+        logger.msg(Arc::ERROR,"Globus location variable substitution is not supported anymore. Please specify path directly.");
+        break;
       default: to_put=param.substr(pos-1,2);
     };
     curpos=pos+1+(to_put.length() - 2);

### 2010-04-14 doc/tech_doc/client/client_technical.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
knowarc1 changes in every files
New testing machine: https://interop.grid.niif.hu:2010/arex-x509

@@
 %
-\bibliographystyle{IEEEtran}       %a nice bibliography style
+\usepackage[numbers]{natbib}
+\bibliographystyle{plainnat}
 %
@@
    <TheseFilesNeedToCheck>
-       <FileURL>http://knowarc1.grid.niif.hu/storage/Makefile</FileURL>
+       <FileURL>http://example.org/storage/Makefile</FileURL>
        <FileURL>ftp://download.nordugrid.org/test/README</FileURL>
@@
    <Result>
-         <FileURL>http://knowarc1.grid.niif.hu/storage/Makefile</FileURL>
+         <FileURL>http://example.org/storage/Makefile</FileURL>
          <ExistInTheCache>true</ExistInTheCache>

### 2010-04-14 src/clients/credentials/test2myproxyserver_get.cpp

@@ -73,7 +73,7 @@ int main(void) {
   cfg.AddCADir("$HOME/.globus/certificates/");
 
   //Arc::ClientTCP client(cfg, "127.0.0.1", 7512, Arc::GSISec, 60);
-  Arc::ClientTCP client(cfg, "knowarc1.grid.niif.hu", 7512, Arc::GSISec, 60);
+  Arc::ClientTCP client(cfg, "example.org", 7512, Arc::GSISec, 60);
 
   Arc::PayloadRaw request;
   request.Insert(send_msg.c_str(), 0, send_msg.length());

### 2010-04-14 src/clients/credentials/test2myproxyserver_put.cpp

@@ -72,7 +72,7 @@ int main(void) {
   cfg.AddCADir("$HOME/.globus/certificates/");
 
   //Arc::ClientTCP client(cfg, "127.0.0.1", 7512, Arc::GSISec, 60);
-  Arc::ClientTCP client(cfg, "knowarc1.grid.niif.hu", 7512, Arc::GSISec, 60);
+  Arc::ClientTCP client(cfg, "example.org", 7512, Arc::GSISec, 60);
 
   Arc::PayloadRaw request;
   request.Insert(send_msg.c_str(), 0, send_msg.length());

### 2010-05-26 doc/tech_doc/xrsl/xrsl-urls.tex
document new checksum functionality

@@
 
-Metadata service URLs also support metadata options which can be used
-for register additional metadata attributes or query the service using
-metadata attributes. These options are specified at the end of the LFN
-and consist of name and value pairs separated by colons. The following
+URLs also support metadata options which can be used for registering
+additional metadata attributes or querying the service using metadata
+attributes. These options are specified at the end of the LFN and
+consist of name and value pairs separated by colons. The following
 attributes are supported:
@@
 
-Currently these metadata options are only supported for lfc:// URLs.
+The checksum attributes may also be used to validate files that were
+uploaded to remote storage.
 
@@
 \texttt{srm.domain.org} at the path \texttt{griddir/file1} and registered to the LFC
-service at \texttt{lfc.domain.org} with the LFN \texttt{/user/file1}. The given GUID
-and checksum attributes will be registered.
+service at \texttt{lfc.domain.org} with the
+LFN \texttt{/user/file1}. The checksum will be compared to what is
+reported by the SRM service after the transfer. The given GUID and
+checksum attributes will also be registered in the LFC.
 

### 2010-06-02 doc/manuals/configuration/configuration_manual.tex
Adding description of new Logger features - Domain and Reopen

@@
 \texttt{Level} element will be written to the log file. The default log level is
-\texttt{WARNING}. \newline
+\texttt{WARNING}.
+
+The log messages are also assigned different domains. Domain is a string
+indicating where message came from. Domains follow hierarchical layout with
+top level called \texttt{Arc} and subdomains added to it separated by dot -
+like \texttt{Arc.MCC.TCP}. It is possible to assign different logging levels
+to different domains by using multiple \texttt{Level} elemets each with
+\texttt{Domain} attribute containing affected domain name. The \texttt{Level}
+without \texttt{Domain} is assigned to \texttt{Arc} domain and is inherited
+by all other domains unless they have theirs levels specified explicitely.
+
 By default the log file is not limited in size, neither
@@
 
+By default log file is kept continiously open. For compatibility with external
+log rotating utilities it is possible to force log file being closed after
+every log record is written and re-opened for new one. For that optional
+\texttt{Reopen} element is used. If set to \texttt{true} it turns compatibility
+mode on.
+
 Plugins which should be used later in the configuration should be specified
@@
       <Level>WARNING</Level>
+      <Level Domain="Arc.MCC.TCP">DEBUG</Level>
       <MaxSize>1000000</MaxSize>
       <Backups>5</Backups>
+      <Reopen>false</Reopen>
     </Logger>

### 2010-06-04 doc/manuals/user_interface/ui-job.tex
Added description of sort options to arcstat

@@
    \texttt{-l, --long}& & long format (extended information)\\
+   \texttt{-S, --sort}& \textit{criterion} & sort jobs according to jobid, submissiontime or jobname\\
+   \texttt{-R, --rsort}& \textit{criterion} & reverse sorting of jobs according to jobid, submissiontime or jobname\\
+
    \texttt{-t, --timeout}& \textit{time}& timeout for queries (default 20 sec)\\
@@
 If the \verb#-l# option is given, extended information is printed.
+
+Jobs can be sorted according to the \texttt{jobid},
+\texttt{submissiontime} or \texttt{jobname}, either in normal or
+reverse order. By using the \verb#--sort# or \verb#--rsort# option
+followed by the desired ordering (\texttt{jobid},
+\texttt{submissiontime} or \texttt{jobname}), jobs will be sorted in
+normal or reverse order. Note that the options \verb#--sort# and
+\verb#--rsort# cannot be used at the same time.
 

### 2010-08-09 doc/tech_doc/gridftpd/gridftpd.tex
doc for arc1 gridftpd service

\documentclass{article}
\usepackage{mathptmx}
\usepackage{helvet}
\usepackage{courier}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{url}
\usepackage{amsmath}
%
\usepackage{setspace}
%\onehalfspacing
\usepackage{amssymb}
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {D.Cameron, A.Konstantinov},
  pdftitle = {The NorduGrid GridFTP Server.},
  pdfsubject = {},
  pdfkeywords = {},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}

\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}

&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}

\end{tabular}

\hrulefill

{\raggedleft NORDUGRID-TECH-XX\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large The NorduGrid GridFTP Server}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Description and Administrator's Manual}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
{\centering \large A. Konstantinov\footnote{aleks@fys.uio.no}, D. Cameron \large \par}

\end{titlepage}

\tableofcontents

\newpage

\section{Introduction\label{sec:intro}}

The NorduGrid~\cite{nordugrid} GridFTP service (GFS) has historically
been the gateway between Grid users and their jobs running on
ARC-enabled resources. It provides an interface for secure job
submission and retrieval with authorization and authentication based
on the Grid Security Infrastructure~\cite{gsi}, and consists of a
standard GridFTP server with NorduGrid modifications on top to handle
Grid jobs. These jobs are then processed and sent to the local batch
system by the Grid Manager (GM).

With the evolution of Grid technologies, and the need for more
standardized interfaces to allow different Grid products to
communicate with each other, the Hosting Environment Daemon
(HED)~\cite{HED} framework and associated Advanced Resource Execution
Service (A-REX) web service were developed to replace the aging GFS
and GM with a modern standards-compliant product. However, the
transistion to a new framework takes time and so the legacy GFS is
still supported and maintained until such time as it becomes
obsolete. This document is designed for administrators who wish to run
the GFS alongside A-REX and describes how to configure and run the GFS
service. For installation of the GFS, up to date instructions may be
found on the NorduGrid website \textbf{\url{http://www.nordugrid.org}}.

\section{Main Concepts}

A detailed summary of job workflow in ARC may be found in the A-REX
technical documentation~\cite{arex}. The following instructions
assume knowledge of concepts such as the \emph{session directory},
job states and how data management is performed by A-REX.

The GFS provides a means to map GSI identities to local usernames, and
thus can expose a local filesystem to the Grid using a highly
configurable set of authentication policies. It also allows Grid job
submission by providing a way for clients to upload a job description
to a Grid resource, and for that job to be executed on a resource
under a mapped local username.

Local file access in the GFS is implemented through plugins (shared
libraries). There are 3 plugins provided: \textit{fileplugin.so,
  gaclplugin.so} and \textit{jobplugin.so} . The
\textit{fileplugin.so} is intended to be used for plain file access
with the configuration sensitive to the user subject and is not
necessary for setting up a NorduGrid compatible site. The
\textit{gaclplugin.so} uses GACL~\cite{gacl} to control access to the
local file system. The \textit{jobplugin.so} uses information about
jobs being controlled by A-REX and provides access to session
directories of the jobs owned by the user. It also provides an
interface (virtual directory and virtual operations) to submit,
cancel, clean, renew credentials and obtain information about the job.

To make GFS to interoperate with other parts of ARC only one
\emph{jobplugin.so} needs to be configured.

\section{Configuration}

The GFS configuration is done through a single INI-style configuration
file. The XML-style configuration supported by A-REX are not supported
by the GFS, and so care must be taken if using the GFS with A-REX
configured using XML-style configuration. The safest option is to use
the same INI-style file for A-REX and the GFS. The default location of
the GFS configuration file is

\begin{itemize}
\item \textit{/etc/arc.conf}
\end{itemize}

A different configuration file location can be specified by the
environment variable ARC\_CONFIG. The configuration file
consists of empty lines, lines containing comments (lines starting
with \#) or configuration commands. It is separated into
sections. Each section starts with a string containing

\begin{itemize}
\item \emph{{[}section name/subsection name/subsubsection name]}. 
\end{itemize}

Each section continues until the next section or until the end of the
file. The configuration file can have commands for multiple
services/modules/programs. Each service has its own section named
after it. The GFS uses the \emph{{[}gridftpd]} section and
sub-sections, along with other authorization-related
sections. Commands in section \emph{{[}common]} apply to all services
configured in the configuration file. Command lines have the format

\begin{itemize}
\item \emph{name=''arguments string''.}
\end{itemize}

\subsection{General Configuration Parameters}

The following parameters are defined in the \emph{{[}gridftpd]}
section of the configuration file.

\begin{itemize}
\item \textbf{\textit{port}}\textit{=number} - specifies TCP/IP port
  number.  Default is 2811.
\item \textbf{\textit{include}}\textit{=path} - include contents of
  another file. Generic commands cannot be specified there.
\item \textbf{\textit{encryption}}\textit{=yes|no} - specifies if
  server will allow data transfer to be encrypted. Default is yes.
\item \textbf{\textit{pluginpath}}\textit{=path} - specifies the path
  where plugin libraries are installed. In a normal installation this
  is \emph{\$(ARC\_LOCATION)/lib/arc}.
\item \textbf{\textit{allowunknown}}\textit{=yes|no} - if set to
  \emph{yes}, clients are not checked against the grid-mapfile. Hence
  only access rules specified in this configuration file will be
  applied.
\item \textbf{\textit{firewall}}\textit{=hostname} - use IP address of
  the \textit{hostname} in response to PASV command instead of IP
  address of a network interface of the computer. An IP address can be
  used instead of \textit{hostname}. This command may be useful if the
  server is situated behind a NAT.
\item \textbf{\textit{unixgroup}}\textit{=group rule} - define local
  UNIX user and optionally UNIX group to which user belonging to
  specified authorization \textit{group} is mapped (see
  Section~\ref{sec:Authorization} for definition of group). Local
  names are obtained from the specified \textit{rule}. If the
  specified rule could not produce any mapping, the next command is
  used. Mapping stops at first matched rule. The following rules are
  supported:

\begin{itemize}
\item \textbf{\textit{mapfile}} \textit{file} - the user's subject is
  matched against a list of subjects stored in the specified file, one
  per line followed by a local UNIX name.
\item \textbf{\textit{simplepool}} \textit{directory} - the user is
  assigned one of the local UNIX names stored in a file
  \textit{directory/pool}, one per line. Used names are stored in
  other files placed in the same \textit{directory}. If a UNIX name
  was not used for 10 days, it may be reassigned to another user.
\item \textbf{\textit{lcmaps}} \textit{library directory database} -
  call LCMAPS functions to do mapping. Here \textit{library} is the
  path to the shared library of LCMAPS, either absolute or relative to
  \textit{directory}; \textit{directory} is the path to the LCMAPS
  installation directory, equivalent to the LCMAPS\_DIR variable;
  \textit{database} is the path to the LCMAPS database, equivalent to
  the LCMAPS\_DB\_FILE variable. Each argument except \textit{library}
  is optional and may be either skipped or replaced with '{*}'. It is
  important to ensure that no configured LCMAPS plugin performs switch
  of local user identity (setuid). That may interfere with way the GFS
  handles local user identities.
\item \textbf{\textit{mapplugin}} \textit{timeout plugin}
  {[}\textit{arg1} {[}\textit{arg2} {[}...]]] - run external
  \emph{plugin} executable with specified arguments. Execution of
  \emph{plugin} may not last longer than \emph{timeout} seconds. A
  rule matches if the exit code is 0 and there is a UNIX name printed
  on \emph{stdout}. A name may be optionally followed by a UNIX group
  separated by ':'. In arguments the following substitutions are
  applied before the plugin is started:

\begin{itemize}
\item \%D - subject of users's cerificate,
\item \%P - name of credentials' proxy file.
\end{itemize}

\end{itemize}

\item \textbf{\textit{unixvo}}\textit{=vo rule} - same as
  \textbf{\textit{unixgroup}} for users belonging to Virtual
  Organization (VO) \textit{vo}.
\item \textbf{\textit{unixmap}}\textit{={[}unixname]{[}:unixgroup]
  rule} - define a local UNIX user and optionally group used to
  represent connected client. \textit{rule} is one of those allowed
  for \textbf{\textit{authorization groups}} (see
  Section~\ref{sec:Authorization}) and for
  \textbf{\textit{unixgroup}}/\textbf{\textit{unixvo}}.  In case of a
  mapping rule, username is the one provided by the rule. Otherwise
  the specified \textit{unixname:unixgroup} is taken. Both
  \textit{unixname} and \textit{unixgroup} may be either omitted or
  set to '{*}' to specify missing value.
\item \textbf{\textit{groupcfg}}\textit{=name} - is put into
  subsections representing a plugin or {[}group] section and defines
  if that section is effective. The only unaffected option is
  \textbf{\textit{groupcfg}}.  If name is empty (or no groupcfg is
  used at all), following lines apply to all users.
\end{itemize}

\subsection{Plugin Configuration}

Subsections of the \emph{gridftpd} section specify plugins which serve
the virtual FTP path (similar to the UNIX mount command). The name of
the subsection is irrelevant but it is useful to use a name related to
the plugin, e.g.\ {[}gridftpd/jobs] for the \emph{jobplugin}. Inside
the subsection, the following commands are supported:

\begin{itemize}
\item \textbf{\textit{plugin}}\textit{=library\_name} - use plugin
  \textit{library\_name} to serve virtual path.
\item \textbf{\textit{path}}\textit{=path} - virtual path to serve.
\end{itemize}

The GFS comes with 3 plugins: \textit{fileplugin.so, gaclplugin.so} and
\textit{jobplugin.so}.

\subsubsection{JobPlugin}

\textit{jobplugin.so} supports the following options:

\begin{itemize}
\item \textbf{\textit{configfile}}\textit{=path} - defines
  non-standard location of the A-REX configuration file,
\item \textbf{\textit{allownew}}\textit{=yes|no} - specifies if new
  jobs can be submitted. Default is \emph{yes}.
\item
  \textbf{\textit{unixgroup}}/\textbf{\textit{unixvo}}/\textbf{\textit{unixmap}}
  - same options as in the top-level GFS configuration. If the mapping
  succeeds, the obtained local user will be used to run the submitted
  job.
\item \textbf{\textit{remotegmdirs}}\textit{=control\_dir session\_dir
  [drain]} - specifies control and session directories under the
  control of another A-REX to which jobs can be assigned (see
  Section~\ref{sub:Multiarex}). Remote directories can be added and
  removed without restarting the GFS. However, it may be desirable to
  drain them prior to removal by adding the \emph{``drain''}
  option. In this case no new jobs will be assigned to these
  directories but their contents will still be accessible.
\item \textbf{\textit{maxjobdesc}}\textit{=size} - specifies maximal
  allowed size of job description in bytes. Default value is 5MB. If
  value is missing or set to 0 no limit is applied.
\end{itemize}

\subsubsection{FilePlugin}

\textit{fileplugin.so} supports the following options:

\begin{itemize}
\item \textbf{\textit{mount}}\textit{=path} - defines the place on
  local filesystem to which file access operations apply.
\item \textbf{\textit{dir}}\textit{=path options} - specifies access
  rules for accessing files in \textit{path} (relative to virtual and
  real path) and all the files below. \\ \textit{options} is a list of
  the following keywords:

\begin{itemize}
\item \textbf{\textit{nouser}} - do not use local file system rights, only
use those specified in this line.
\item \textbf{\textit{owner}} - check only file owner access rights.
\item \textbf{\textit{group}} - check only group access rights.
\item \textbf{\textit{other}} - check only ``others'' access rights.
\end{itemize}

The options above are exclusive. If none of the above are specified, the usual
UNIX access rights are applied.

\begin{itemize}
\item \textbf{\textit{read}} - allow reading files.
\item \textbf{\textit{delete}} - allow deleting files.
\item \textbf{\textit{append}} - allow appending files (does not allow creation).
\item \textbf{\textit{overwrite}} - allow overwriting of existing files
(does not allow creation, file attributes are not changed).
\item \textbf{\textit{dirlist}} - allow obtaining list of the files.
\item \textbf{\textit{cd}} - allow to make this directory current.
\item \textbf{\textit{create}} \textit{owner:group permissions\_or:permissions\_and}
- allow creating new files. File will be owned by \textit{owner} and
owning group will be \textit{group}. If '{*}' is used, the user/group
to which connected user is mapped will be used. The permissions will
be set to \textit{permissions\_or} \& \textit{permissions\_and} (the second
number is reserved for future usage). 
\item \textbf{\textit{mkdir}} \textit{owner:group permissions\_or:permissions\_and}
- allow creating new directories.
\end{itemize}
\end{itemize}

\subsubsection{GACLPlugin}

\textit{gaclplugin.so} supports the following options:

\begin{itemize}
\item \textbf{\textit{gacl}}\textit{=gacl} - GACL XML.
\item \textbf{\textit{mount}}\textit{=path} - local path served by
plugin.
\end{itemize}

The GACL XML may contain variables which are replaced with values taken from
the client's credentials. The following variables are supported:

\begin{list}{--}{\setlength{\labelwidth}{0.5cm}\setlength{\rightmargin}{\leftmargin}}
\item [{\emph{\$subject}}] - subject of user's certificate (DN),
\item [{\emph{\$voms}}] - subject of VOMS\cite{voms} server (DN),
\item [{\emph{\$vo}}] - name of VO (from VOMS certificate),
\item [{\emph{\$role}}] - role (from VOMS certificate),
\item [{\emph{\$capability}}] - capabilities (from VOMS certificate),
\item [{\emph{\$group}}] - name of group (from VOMS certificate) .
\end{list}
Additionally, the root directory must contain a \emph{.gacl} file with initial
ACLs. Otherwise the rule will be {}``deny all for everyone''.

\subsection{Authorization}
\label{sec:Authorization}

% this section is taken from the Config_Auth document
% ``Configuration and Authorisation of ARC (NorduGrid) Services''

ARC services which have to authorize remote client applications use
the notion of \textit{group} for authorization purposes. Each
\textit{group} is made of \textit{rules} applied sequentially. If a
client's credentials pass \textit{all rules}, the client is treated as
belonging to the specified \textit{group}.

Each group is represented by a top level section named
{[}\textbf{\textit{group}}{]} or its subsection. Each such section
represents a separate authorization group and its name is given by the
\textbf{\textit{name}} command inside that section. If there is no
\textbf{\textit{name}} command then the name of the subsection is
used.

Authorization is performed by applying set of rules. Rules obey same
format as the rest of the configuration file. Each rules command
consists of a \textit{rule word} prepended with optional
\textit{modifiers} - {[}+|-{]}{[}!{]}

\begin{itemize}
\item [+]accept credential if matches the following rule (positive match, default
action),
\item [-]reject credential if matches the following rule (negative match),
\item [!]invert matching. Match is treated as non-match. Non-match is treated
as match, either positive ({}``+'' or nothing) or negative ({}``-'').
\end{itemize}

Processing of rules in every group stops after the first positive or
negative match, or failure is reached. If a rule does not match then
processing continues.  Failures are rule-dependant and may be caused
by conditions like a missing file, unsupported rule, etc.

The following \textit{rule words} and arguments are supported:

\begin{itemize}
\item {[}\textbf{\textit{subject}}{]}\textit{=subject}
  {[}\textit{subject} {[}...{]}{]} - match user with one of specified
  subjects
\item \textbf{\textit{file}}={[}\textit{filename} {[}...{]}{]} - read
  rules from specified files (format of file is similar to Globus
  grid-mapfile with user names ignored)
\item \textbf{\textit{remote}}={[}\textit{ldap://host:port/dn}
  {[}...{]}{]} - match user listed in one of specified LDAP
  directories (uses network connection hence can take time to process)
\item \textbf{\textit{voms}}\textit{=vo group role capabilities} -
  accept user with VOMS proxy with specified \textit{vo},
  \textit{group}, \textit{role} and \textit{capabilities}. '{*}' can
  be used to accept any value
\item \textbf{\textit{vo}}={[}\textit{vo} {[}...{]}{]} - match user
  belonging to one of specified Virtual Organizations as defined in
  \emph{vo} configuration section (see below).
\item \textbf{\textit{group}}={[}\textit{groupname}
  {[}\textit{groupname} {[}...{]}{]}{]} - match user already belonging
  to one of specified groups.
\item \textbf{\textit{plugin}}\textit{=timeout plugin}
  {[}\textit{arg1} {[}\textit{arg2} {[}...{]}{]}{]} - run external
  \emph{plugin} (executable or function in shared library) with
  specified arguments. Execution of \emph{plugin} may not last longer
  than \emph{timeout} seconds.  If \emph{plugin} looks like
  \emph{function@path} then function \emph{int
    function(char{*},char{*},char{*},...)} from shared library
  \emph{path} is called (\emph{timeout} is not functional in that
  case). Rule matches if exit code is 0. In arguments following
  substitions are applied before plugin is started:

\begin{itemize}
\item \%D - subject of users's cerificate,
\item \%P - name of credentials' proxy file.
\end{itemize}

\item \textbf{\textit{lcas}}\textit{=library directory database} -
  call LCAS functions to check rule. Here \textit{library} is path to
  shared library of LCAS, either absolute or relative to
  \textit{directory}; \textit{directory} is path to LCAS installation
  directory, equivalent of LCAS\_DIR variable; \textit{database} is
  path to LCAS database, equivalent to LCAS\_DB\_FILE variable. Each
  arguments except \textit{library} is optional and may be either
  skiped or replaced with '{*}'.
\item \textbf{\textit{all}} - accept any user
\end{itemize}

Here is an example of authorization group:

\begin{verbatim}
 [group/admins]

 -subject="/O=Grid/OU=Wrong Place/CN=Bad Person"
 file="/etc/grid-security/internal-staff"
 voms="nordugrid * * admin"
\end{verbatim}

In this example the following rules are applied to determine whether
the identity presented is part of the group ``admins'':

\begin{itemize}
\item If the identity is ``/O=Grid/OU=Wrong Place/CN=Bad Person'' it
  is rejected
\item If the identity is in the mapfile
  ``/etc/grid-security/internal-staff'' the identity is accepted
\item If the identity is a VOMS proxy with a nordugrid VO extension
  and admin capability, the identity is accepted
\end{itemize}

\subsubsection{Virtual Organizations}

VOs are defined in the \emph{vo} configuration section. The following
commands are supported:

\begin{itemize}
\item \textbf{\textit{vo}}\textit{=vo\_name} - specifies name of
  VO. Mandatory command.
\item \textbf{\textit{file}}\textit{=path} - path to file which
  contains list of users' DNs belonging to VO.
\item \textbf{\textit{source}}\textit{=URL} - specifies URL from which
  list of users may be obtained. May be multiple.
\end{itemize}

\section{Configuration Examples}

The examples presented below contain a full configuration examples for
the GridFTP server. The A-REX and information system configurations
are not shown - this is described in other documents.

\subsection{Simple Example}

In the following minimal example we use a single static mapfile which
contains all possible user mappings for this site.

\begin{verbatim}
[common]
hostname="myhost.org"
lrms="fork"
gridmap="/etc/grid-security/grid-mapfile"

[gridftpd]
debug="3"
logfile="/var/log/gridftpd.log"
logsize="10000000 2"
pidfile="/var/run/gridftpd.pid"
pluginpath="/usr/local/lib/arc"
encryption="no"
allowunknown="no"
maxconnections="200"

[gridftpd/jobs]
path="/jobs"
plugin="jobplugin.so"
\end{verbatim}

\subsection{Detailed Example}

Here we configure a simple PBS based cluster according to the
following use case.  John is member of the VO "smscg" where he belongs
to the group "atlas" and has been assigned the roles "production" and
"test". Since groups and roles are fully decoupled, John can request
proxies that can include one (or several) of the following different
group-role combinations (termed "Fully Qualified Names" (FQAN)):
\begin{itemize}
\item /smscg (notice it's the same as /smscg/Role=NULL)
\item /smscg/Role=production
\item /smscg/Role=test
\item /smscg/atlas
\item /smscg/atlas/Role=production
\item /smscg/atlas/Role=test
\end{itemize}

A-REX serves as front-end to a batch-system that provides a
"low\_prio\_queue" and a "high\_prio\_queue". Assignment to the
different queues is done via local user identites.  More precisely,
the local users "smscg001, smscg002, smscg003" will be assigned to the
low\_prio\_queue, whereas users "smscgP001, smscgP002, smscgP003" to
the high\_prio\_queue (the configration of the batch-system to support
this is out of scope of this example).

Users sending jobs to A-REX should be assigned to one of the queues
depending on the credentials they present in their proxy
certificate. The assigment shall look as follows:
\begin{itemize}
\item /smscg , /smscg/Role=test , /smscg/Role=production => shall map
  to one of the smscg00[1-3] local identities (thus low\_prio\_queue)
\item /smscg/atlas , /smscg/atlas/Role=test ,
  /smscg/atlas/Role=production => shall map to one of the
  smscgP00[1-3] local identities (thus high\_prio\_queue)
\end{itemize}

The following usage pattern is considered. User John first wants to
run a monitoring job on the high\_prio\_queue. He performs a
voms-proxy-init and specifies his "/smscg/atlas/Role=test" FQAN to be
used. When he submits his monitoring-job, John will be mapped to one
of the smscgP001, smscgP002, smscgP003 accounts. John's job will thus
run on the high\_prio\_queue.

After submitting the monitoring job, John submits regular jobs with
his FQAN "/smscg". These jobs will run on the low\_prio\_queue. Later
John switches back to the FQAN "/smscg/atlas/Role=test" to fetch the
result of his monitoring job.

The discrimination to what queue John is to be mapped is done with VO
information only and not on the basis of the DN of John's
certificate. Hence the choice to what queue to be mapped is under
control of John (we silently presumed John knows the mappings at the
source).

Notes:
\begin{itemize}
\item a DN based grid-mapfile is generated on the front-end with a
  default mapping entry for John. The grid-mapfile is only used by the
  information system (GIIS) to make the grid resource look eligible
  for jobs submitted by John.
\item the DN based grid-mapfile per se does not permit John to access
  the grid resource under different local identies (e.g. once as
  smscg001 and later as smscgP001), since the first matching DN
  defines the local identity John is to be mapped to. This is not a
  flaw since NorduGrid has support for lcmaps, which allows a
  're-mapping' of a user.
\item the mapping of the FQAN to the local user identity
  (e.g. "/smscg" to local user "smscg001") shall be done with lcmaps
  (in detail the lcmaps framework + lcmaps voms plugins). Direct VOMS
  based mapping is also possible.
\end{itemize}

If user John creates a proxy certificate with the "grid-proxy-init"
command instead of "voms-proxy-init", hence the proxy certificate will
not contain any VO information and submits a job to A-REX (the
matchmaking will still work, since it's done with John's DN) he shall
not be authorized.

Example configuration:

\begin{verbatim}
[common]
pbs_bin_path="/usr/bin"
pbs_log_path="/var/spool/pbs/server_logs"
hostname="myhost.org"
lrms="pbs"

[vo]
# We will use this configuration block for a few purposes.
# 1. To generate grid-mapfile needed for information system.
#    For that purpose nordugridmap utility will have to be 
#    run periodically.
# 2. To provide coarse-grained information to authorization 
#    rules used to define authorization groups. If needed of
#    course.
id="smscg_vo"
vo="smscg_vo"

# Here we define path to file to which nordugridmap will write DNs of
# users matching rules below. Because we are going to use it as
# grid-mapfile for other purposes it is going to reside at default
# location.  
file="/etc/grid-security/grid-mapfile"

# Now we tell nordugridmap to pull information from
# VOMRS/VOMSS/or_whatever_it_is_called_now service and to ask for
# users belonging to smscg VO.
source="vomss://voms.smscg.org:8443/voms/smscg"

# Now we specify default mapping to local *NIX id. It is possible to
# completely redefine mapping in [gridftpd] block. But this one will
# be used by information system to compute and present resources
# available to user.  Let's use one of lowest priority account defined
# in use-case.
mapped_unixid="smscg001"

[group]
# In this authorization group we are going to check if user presents
# any proof that he belongs to 'smscg' VO. We can use that information 
# later to explicitely limit access to resources. If such access 
# control is not needed this group can be removed.
name="smscg_auth"

# Here we can use internal support of ARC for VOMS attrbutes
# voms="smscg * * *"
# If we want to limit access to resources also by other VOMS
# attributes then other voms rules similar to those defined
# below in [gridftpd] section may be used.

# Or we can ask some external executable to analyze delegated
# credentials of user. In this example executable vomatch 
# is called with first argument containing path to delegated
# proxy certificate and second - required VO name.
# plugin="10 /opt/external/bin/vomatch %P smscg"

# Or - probably prefered way in this use case - we can use 
# LCAS to analyze delegated proxy.
# First element after '=' sign is path to LCAS library whatever
# it is called in current implementation. Second is LCAS installation
# path - it will be used to set environment variable LCAS_DIR.
# And third element is path to LCAS database file - it will be passed
# to environemnt variable LCAS_DB_FILE.
# Function 'lcas_get_fabric_authorization' of specified LCAS library
# will be called with following 3 arguments
#  1. char* pointer to string containing DN of user
#  2. gss_cred_id_t variable pointing at delegated credentials of user
#  3. char* pointer to empty string
# Returned 0 int value is treated as positive response
lcas="/opt/glite/lib/liblcas.so /opt/glite /opt/glite/share/lcas.db"

# As coarse grained solution it is also possible to check if user 
# belongs to one of defined VOs as specified in _previously_ defined
# [vo] group. Here we refer to VO group smscg_vo defined above.
#vo="smscg_vo"

[gridftpd]
debug="2"
logfile="/var/log/gridftpd.log"
logsize="100000 2"
pidfile="/var/run/gridftpd.pid"
port="2811"
pluginpath="/usr/local/lib/arc"
encryption="no"

# By specifying 'no' here we limit users allowed to exatblish
# connection to this server to those specified in grid-mapfile. This
# may be not necessary if additional authorization is applied as done
# below. But this provides additional layer of protection so let it
# be.
allowunknown="no"

maxconnections="200"

# Here we start fine-grained user mapping. Let's first define few VOMS
# mappings using embedded functionality of ARC.  These lines should
# map Grid users to high-priority and low-priority *NIX users smscg001
# and smscgP001. Mind order - those with more attributes defined come
# first.  I do not know if missing attribute is passed by VOMS as
# empty string or as string containing NULL keyword. Here I assume
# empty string. If it is NULL then "" has to be replaced with NULL.
#unixmap="smscgP001 voms smscg atlas test *
#unixmap="smscgP001 voms smscg atlas production *
#unixmap="smscgP001 voms smscg atlas "" *
# These 3 lines are not needed if grid-mapfile defines default mapping
# to smscg001 user. But we can have them for consistence and if mapping
# to nobody is defined below for safety reasons.
#unixmap="smscg001 voms smscg "" test *
#unixmap="smscg001 voms smscg "" production *
#unixmap="smscg001 voms smscg "" "" *

# Instead of using multiple unixmap commands above we may define 
# 2 authorization groups using [group] blocks. Let's say their
# names are smscg_low and smscg_high. Then 'group' matching rule
# may be used. 
#unixmap="smscgP001 group smscg_high"
#unixmap="smscg001 group smscg_low"

# Or if we want to use all 6 local accounts and let mapping choose
# randomly within 2 group accounts 'simplepool' may be used. In
# example below 'unixgroup' ensures proper choise of group and
# 'simplepool' makes a choise from acoounts in pool. Last argument
# specifies directory containing file named 'pool'. That file contains
# list of local user accounts. Also this directory will be used for
# writing information about current mappings.
#unixgroup="smscg_high simplepool /var/nordugrid/smscg_high"
#unixgroup="smscg_low simplepool /var/nordugrid/smscg_low"

# And mapping prefered in this use case - through LCMAPS.  First
# element after '=' sign is path to LCMAPS library whatever it is
# called in current implementation. Second is LCMAPS installation path
# - it will be used to set environment variable LCMAPS_DIR.  And third
# element is path to LCMAPS database file - it will be passed to
# environemnt variable LCMAPS_DB_FILE. Those 3 arguments are followed
# list of policy names.
# Function 'lcmaps_run_and_return_username' of specified LCMAPS library
# will be called with following arguments
#  1. char* pointer to string containing DN of user
#  2. gss_cred_id_t variable pointing at delegated credentials of user
#  3. char* pointer to empty string
#  4. char** pointer for chosen username.
#  5. int variable containing number of policies
#  6. char** list of policy names
# Expected 0 int value returned and argument 4 set. Value returned in
# 4th argument is used as username of local account.
unixmap="* lcmaps /opt/glite/lib/liblcmaps.so /opt/glite \
  /opt/glite/share/lcmaps.db policy1 policy2"

# Here we can specify mapping to some harmless local user account for
# safety reasons. If that account is not allowed to submit jobs to
# LRMS then this will also work as authorization effectively cuting
# off users without proper VOMS attributes.
unixmap="nobody:nobody all"

[gridftpd/jobs]
# This block defines job submission service
path="/jobs"
plugin="jobplugin.so"

# Line below specifies that this plugin/service is only available to
# users belonging to authorization group. If such behavior is not
# required then this line must be commented.
groupcfg="smscg_auth"

[queue/low_prio_queue]
name="low_prio_queue"
homogeneity="True"
scheduling_policy="FIFO"
comment="This queue is low priority"
nodecpu="adotf"
nodememory="512"
architecture="adotf"
opsys="Mandrake 8.0"
opsys="Linux-2.4.19"
benchmark="SPECINT2000 222"
benchmark="SPECFP2000 333"
cachetime="30"
timelimit="30"
sizelimit="5000"

[queue/high_prio_queue]
name="high_prio_queue"
homogeneity="True"
scheduling_policy="FIFO"
comment="This queue is high priority"
nodecpu="adotf"
nodememory="512"
architecture="adotf"
opsys="Mandrake 8.0"
opsys="Linux-2.4.19"
benchmark="SPECINT2000 222"
benchmark="SPECFP2000 333"
\end{verbatim}

\section{Running the service}

An initialization script \emph{a-rex-gridftpd} for the GFS is provided
in \emph{\$ARC\_LOCATION/etc/init.d} (or equivalent depending on
architecture).

\begin{verbatim}
Usage: a-rex-gridftpd {start|stop|status|restart|reload|condrestart}
\end{verbatim}

Upon starting and depending on the configured log level, messages will
be logged in the log file specified in the configuration file.

\section{Using Multiple A-REX services Under One GFS\label{sub:Multiarex}}

For large clusters, using a single machine for all input and output
file transfer, as well for the interaction with the LRMS and
Information System, can limit the job throughput of the
cluster. Running several A-REXs on separate hosts can help spread the
hardware and network load. A single GFS can feed jobs to several A-REXs,
hence a cluster with many A-REXs still appears as a single site to the
outside world. When a job is submitted, the GFS jobplugin assigns a
random control directory to use for the job from the main
\emph{controldir} specified in the A-REX configuration and any extra
\emph{remotegmdirs} specified in the jobplugin configuration.

Each control directory is used by a separate A-REX, therefore for every
\emph{remotegmdirs} command in the GFS configuration, there must be a
A-REX running which defines the corresponding \emph{controldir} and
\emph{sessiondir} in its configuration file. Each A-REX can run
independently on its own host, the only requirement is that the
control and session directories must be accessible on the GFS host,
and the GFS user must have write access to these directories. It is
recommended that these directories are local to the GFS host and
exported (via NFS for example) to the other hosts, rather than being
on a remote filesystem and exported to the GFS host. This means that
any glitches in the network do not cause the GFS host to hang. It is
also important that the local user accounts on each host must be
synchronised with the GFS host. A A-REX is not aware that any other A-REXs
are running, as they only see what the GFS decides should go into
their own control directory. All communication between the GFS and a
A-REX is through the A-REX's control directory. Note that in remote control
directories there is no way to specify control directories per user,
as with the \emph{control} command. Only the \emph{controldir} command
can be used and all users will use the same control directory.

One feature of this design is that multiple A-REXs can share the same
LRMS, and hence compete with each other to submit jobs. Therefore any
LRMS settings in the configuration files must be carefully matched in
order not to bias one A-REX over another. In most cases each host's
configuration file can be identical apart from the control and session
directories. Some configuration sections such as the GFS and infosys
sections will be ignored by the remote A-REX hosts as these services are
not running.

Cacheing can be set up in a variety of different ways. Each A-REX can
have its own cache, completely independent from any other, which will
lead to popular files being replicated in many caches. Or, all caches
can be shared with all A-REXs, which means no replication between
caches but heavy intra-site network traffic if the cache file systems
are hosted on different hosts. Another option is to give each A-REX
its own cache, but access to the other caches as remote caches. This
avoids replicating files and intra-site network traffic. Replication
can still be enabled by specifying ``replicate'' as the link\_path for
remote cache dirs, and the advantage of this is that files are copied
from the remote cache rather than being downloaded again from source.

When setting up an extra A-REX, it is important that no other services
(GFS, infosys) run on the host. The instances of these services
running on the ``main'' host take care of all the A-REXs. In other
words, the startup scripts for \emph{gridftpd} and \emph{grid-infosys}
should be removed from any place where they would be started
automatically (usually \emph{\$ARC\_LOCATION/etc/init.d/}). No host
certificates are required for hosts which only run a A-REX
instance. Note that in this multiple A-REX set up, there is a one to
one relationship between control and session directories, hence
multiple \emph{sessiondir} commands cannot be used in a A-REX
configuration.

\bibliography{grid}

\end{document}

### 2010-08-10 doc/manuals/chelonia/arc-storage-manual.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
ca_dir instead of cacert

@@
 
-cacert=/etc/grid-security/certificates
+ca_dir=/etc/grid-security/certificates
 host_cert=/etc/grid-security/hostcert-chelonia1.pem
@@
 
-cacert=/etc/grid-security/certificates
+ca_dir=/etc/grid-security/certificates
 host_cert=/etc/grid-security/hostcert-chelonia2.pem
@@
 
-cacert=/etc/grid-security/certificates
+ca_dir=/etc/grid-security/certificates
 host_cert=/etc/grid-security/hostcert-chelonia3.pem

### 2010-08-11 doc/tech_doc/a-rex/arex_tech_doc.tex
Adding description of %F substitution of a-rex
add preferredpattern option to allow ordering of replicas according to preference (bug 1933)

@@
 \item Download is initiated by the A-REX -- This is the case for files
-  defined in the JD (with name and source). The A-REX alone is
-  responsible to ensure that all required files will be available in
-  the SD.\\ The supported protocols for sources at the moment are (in
-  case of full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over
-  SSLv3). Some less standard sources and \emph{Indexing Services} are
-  also supported, these are described in Section~\ref{sec:urls}. The
-  A-REX fully relies on the HED framework~\cite{hed} for data
-  transferring capabilities and so the actual set of supported
-  protocols depends on the installed Data Management Components of the
-  HED.
+  defined in the JD (with name and source). The A-REX alone is responsible to
+  ensure that all required files will be available in the SD.\\ The supported
+  protocols for sources at the moment are (in case of full installation):
+  GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Some less standard sources and
+  \emph{Indexing Services} are also supported, these are described in
+  Section~\ref{sec:urls}. In the case where a file in an indexing service
+  resolves to multiple physical replicas, some selection of the ``best'' replica
+  may be performed based on any available information on the access latency of
+  those replicas. The A-REX fully relies on the HED framework~\cite{hed} for data
+  transferring capabilities and so the actual set of supported protocols depends
+  on the installed Data Management Components of the HED.
 
@@
 first 3 minutes.
+\item \textbf{\textit{preferredpattern}}\textit{=pattern} - specifies
+  a preferred pattern on which to sort multiple replicas of an input
+  file. It consists of one or more patterns separated by a pipe
+  character ($|$) listed in order of preference. Replicas will be
+  ordered by the earliest match. If the dollar character (\$) is used
+  at the end of a pattern, the pattern will be matched to the end of
+  the hostname of the replica.
 \item \textbf{\textit{copyurl}}\textit{=template replacement} -- specifies
@@
 \item [{\%c}] -- list of all control directories
+\item [{\%F}] -- path to configuration file of this instance
 \item [{\%I}] -- job ID (for plugins only, substituted in runtime)

### 2010-08-11 src/hed/libs/data/DataMover.cpp

@@ -349,6 +349,8 @@ namespace Arc {
       if (strcasecmp(value.c_str(), "no") != 0)
         destination_overwrite = true;
     }
+    // sort source replicas according to the expression supplied
+    source.SortLocations(preferred_pattern);
     if (destination_overwrite) {
       if ((destination.IsIndex() && destination_meta_initially_stored)
           || (!destination.IsIndex())) {

### 2010-08-11 src/services/a-rex/grid-manager/jobs/states.cpp

@@ -589,6 +589,8 @@ bool JobsList::state_loading(const JobsList::iterator &i,bool &state_changed,boo
       NULL, // (-d)
       NULL, // -C
       NULL, // (-C)
+      NULL, // -r
+      NULL, // (-r)
       NULL, // id
       NULL, // control
       NULL, // session
@@ -636,6 +638,10 @@ bool JobsList::state_loading(const JobsList::iterator &i,bool &state_changed,boo
     if (!user->Env().nordugrid_config_loc().empty()) {
       args[argn]="-C"; argn++;
       args[argn]=(char*)(cfg_path.c_str()); argn++;
+    }
+    if(!jcfg.preferred_pattern.empty()) {
+      args[argn]="-r"; argn++;
+      args[argn]=(char*)(jcfg.preferred_pattern.c_str()); argn++;
     }
     args[argn]=(char*)(i->job_id.c_str()); argn++;
     args[argn]=(char*)(user->ControlDir().c_str()); argn++;

### 2010-08-12 doc/tech_doc/gridftpd/gridftpd.tex
assign document number
add description of log and pid options and spellcheck

@@
 
-{\raggedleft NORDUGRID-TECH-XX\par}
+{\raggedleft NORDUGRID-TECH-26\par}
 
@@
 and GM with a modern standards-compliant product. However, the
-transistion to a new framework takes time and so the legacy GFS is
+transition to a new framework takes time and so the legacy GFS is
 still supported and maintained until such time as it becomes
@@
 
-To make GFS to interoperate with other parts of ARC only one
+To make GFS to inter-operate with other parts of ARC only one
 \emph{jobplugin.so} needs to be configured.
@@
 \begin{itemize}
+\item \textbf{\textit{pidfile}}\textit{=path} -- specifies file where
+  process id of the gridftpd process will be stored. Defaults to
+  \emph{/var/run/gridftpd.pid} if running as root and
+  \emph{\$HOME/gridftpd.pid} otherwise.
+\item \textbf{\textit{logfile}}\textit{=path} -- specifies name of file
+  for logging debug/informational output. Defaults to
+  \emph{/var/log/gridftpd.log}.
+\item \textbf{\textit{logsize}}\textit{=size number} -- restricts
+  log file size to \emph{size} and keeps \emph{number} archived log
+  files.
+\item \textbf{\textit{debug}}\textit{=number} -- specifies level of
+  debug information. More information is printed for higher
+  levels. Currently the highest effective number is 5 (DEBUG) and
+  lowest 0 (FATAL). Defaults to 3 (INFO).
 \item \textbf{\textit{port}}\textit{=number} - specifies TCP/IP port
@@
 \begin{itemize}
-\item \%D - subject of users's cerificate,
+\item \%D - subject of user's certificate,
 \item \%P - name of credentials' proxy file.
@@
   case). Rule matches if exit code is 0. In arguments following
-  substitions are applied before plugin is started:
+  substitutions are applied before plugin is started:
 
 \begin{itemize}
-\item \%D - subject of users's cerificate,
+\item \%D - subject of user's certificate,
 \item \%P - name of credentials' proxy file.
@@
   arguments except \textit{library} is optional and may be either
-  skiped or replaced with '{*}'.
+  skipped or replaced with '{*}'.
 \item \textbf{\textit{all}} - accept any user
@@
   ``/etc/grid-security/internal-staff'' the identity is accepted
-\item If the identity is a VOMS proxy with a nordugrid VO extension
+\item If the identity is a VOMS proxy with a NorduGrid VO extension
   and admin capability, the identity is accepted
@@
 
-The examples presented below contain a full configuration examples for
+The examples presented below contain full configuration examples for
 the GridFTP server. The A-REX and information system configurations
@@
 low\_prio\_queue, whereas users "smscgP001, smscgP002, smscgP003" to
-the high\_prio\_queue (the configration of the batch-system to support
+the high\_prio\_queue (the configuration of the batch-system to support
 this is out of scope of this example).
@@
 depending on the credentials they present in their proxy
-certificate. The assigment shall look as follows:
+certificate. The assignment shall look as follows:
 \begin{itemize}
@@
 \item the DN based grid-mapfile per se does not permit John to access
-  the grid resource under different local identies (e.g. once as
+  the grid resource under different local identities (e.g. once as
   smscg001 and later as smscgP001), since the first matching DN
@@
 # any proof that he belongs to 'smscg' VO. We can use that information 
-# later to explicitely limit access to resources. If such access 
+# later to explicitly limit access to resources. If such access 
 # control is not needed this group can be removed.
@@
 
-# Here we can use internal support of ARC for VOMS attrbutes
+# Here we can use internal support of ARC for VOMS attributes
 # voms="smscg * * *"
@@
 
-# Or - probably prefered way in this use case - we can use 
+# Or - probably preferred way in this use case - we can use 
 # LCAS to analyze delegated proxy.
@@
 # And third element is path to LCAS database file - it will be passed
-# to environemnt variable LCAS_DB_FILE.
+# to environment variable LCAS_DB_FILE.
 # Function 'lcas_get_fabric_authorization' of specified LCAS library
@@
 
-# By specifying 'no' here we limit users allowed to exatblish
+# By specifying 'no' here we limit users allowed to establish
 # connection to this server to those specified in grid-mapfile. This
@@
 # randomly within 2 group accounts 'simplepool' may be used. In
-# example below 'unixgroup' ensures proper choise of group and
-# 'simplepool' makes a choise from acoounts in pool. Last argument
+# example below 'unixgroup' ensures proper choice of group and
+# 'simplepool' makes a choise from accounts in pool. Last argument
 # specifies directory containing file named 'pool'. That file contains
@@
 
-# And mapping prefered in this use case - through LCMAPS.  First
+# And mapping preferred in this use case - through LCMAPS.  First
 # element after '=' sign is path to LCMAPS library whatever it is
@@
 # element is path to LCMAPS database file - it will be passed to
-# environemnt variable LCMAPS_DB_FILE. Those 3 arguments are followed
+# environment variable LCMAPS_DB_FILE. Those 3 arguments are followed
 # list of policy names.
@@
 # safety reasons. If that account is not allowed to submit jobs to
-# LRMS then this will also work as authorization effectively cuting
+# LRMS then this will also work as authorization effectively cutting
 # off users without proper VOMS attributes.

### 2010-09-20 doc/manuals/chelonia/arc-storage-manual.tex
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

@@
 pidfile=/tmp/chelonia1.pid
-logfile=/var/log/chelonia1.log
+logfile=/var/log/arc/chelonia1.log
 debug=ERROR
@@
 pidfile=/tmp/chelonia2.pid
-logfile=/var/log/chelonia2.log
+logfile=/var/log/arc/chelonia2.log
 debug=ERROR
@@
 pidfile=/tmp/chelonia3.pid
-logfile=/var/log/chelonia3.log
+logfile=/var/log/arc/chelonia3.log
 debug=ERROR

### 2010-09-20 doc/manuals/configuration/configuration_manual.tex
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

@@
 log messages to and if not present log messages will be written to the
-\texttt{/var/log/arched.log} file. Each log message is associated with a log
+\texttt{/var/log/arc/arched.log} file. Each log message is associated with a log
 level which indicates the serverity of the message. The following log levels
@@
     <cfg:Logger>
-      <cfg:File>/var/log/arched.log</cfg:File>
+      <cfg:File>/var/log/arc/arched.log</cfg:File>
       <cfg:Level>WARNING</cfg:Level>
@@
     <File inisections="special common"
-          initag="logfile">/var/log/arched.log</File>
+          initag="logfile">/var/log/arc/arched.log</File>
     ...
@@
 /Server/Group & group & \textrm{None} \\
-/Server/Logger/File & logfile & /var/log/arched.log \\
+/Server/Logger/File & logfile & /var/log/arc/arched.log \\
 /Server/Logger/Level & loglevel & WARNING \\

### 2010-09-20 doc/manuals/echo/echo-manual.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

@@
         <cfg:Logger>
-            <cfg:File>/var/log/arched.log</cfg:File>
+            <cfg:File>/var/log/arc/arched.log</cfg:File>
             <cfg:Level>ERROR</cfg:Level>

### 2010-09-20 doc/manuals/hopi/hopi-manual.tex
Synchronising Echo and Hopi documentation with Nox
Assigning a number to Hopi manual
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

@@
 
-{\raggedleft NORDUGRID-MANUAL-XXX\par}
+{\raggedleft NORDUGRID-MANUAL-15\par}
 

### 2010-09-20 doc/manuals/janitor/tex_installation/installation.tex
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

@@
 log4perl.appender.MainLog.Threshold = DEBUG
-log4perl.appender.MainLog.filename = /var/log/janitor.log
+log4perl.appender.MainLog.filename = /var/log/arc/janitor.log
 log4perl.appender.MainLog.layout = PatternLayout
@@
 log4perl.appender.ErrorLog.Threshold = ERROR
-log4perl.appender.ErrorLog.filename = /var/log/janitor_error.log
+log4perl.appender.ErrorLog.filename = /var/log/arc/janitor_error.log
 

### 2010-09-20 doc/tech_doc/a-rex/arex_tech_doc.tex
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

@@
   for logging debug/informational output. Defaults to
-  \emph{/var/log/grid-manager.log}.
+  \emph{/var/log/arc/grid-manager.log}. If installed from packages, the
+  default log is managed by logrotate.
 \item \textbf{\textit{logsize}}\textit{=size number} -- restricts
@@
       <Level>VERBOSE</Level>
-      <File>/var/log/arched.log</File>
+      <File>/var/log/arc/arched.log</File>
       <MaxSize>100000000</MaxSize>

### 2010-09-20 doc/tech_doc/gridftpd/gridftpd.tex
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

@@
   for logging debug/informational output. Defaults to
-  \emph{/var/log/gridftpd.log}.
+  \emph{/var/log/arc/gridftpd.log}. If installed from packages, the
+  default log is managed by logrotate.
 \item \textbf{\textit{logsize}}\textit{=size number} -- restricts
@@
 debug="3"
-logfile="/var/log/gridftpd.log"
+logfile="/var/log/arc/gridftpd.log"
 logsize="10000000 2"
@@
 debug="2"
-logfile="/var/log/gridftpd.log"
-logsize="100000 2"
+logfile="/var/log/arc/gridftpd.log"
+logsize="10000000 2"
 pidfile="/var/run/gridftpd.pid"

### 2010-09-20 doc/tutorials/ws-programming-tutorial/doc/tex_introduction/introduction.tex
Reorganizing doc catalogue and enabling make doc
Added a modified doc folder refering to the new folder structure of the examples
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

--- 2009-03-17 doc/ws-programming-tutorial/doc/tex_introduction/introduction.tex
@@
 \subsection{The ARC Echo Web Service}
+\label{sec:arc_echo_web_service}
 
@@
 	\textbf{HED configuration file.}}]
-{../src/services/arched_arcecho.xml}
+{../examples/src/services/arched_arcecho.xml}
 The first line of the XML file contains the XML declaration.
@@
 
-% $ rm -f /var/log/arched.log
-% $ tail -n100 -f /var/log/arched.log
+% $ rm -f /var/log/arc/arched.log
+% $ tail -n100 -f /var/log/arc/arched.log
 %...
@@
 For debugging it is advisable to check the logfile which was assigned inside the server configuration file.
-It is commonly found at /var/log/arched.log.
+It is commonly found at /var/log/arc/arched.log.
 

### 2010-09-20 doc/tutorials/ws-programming-tutorial/doc/tex_timeservice/timeservice.tex
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/

@@
 In order to start the service, the \textit{arched} command, shown in Listing~\ref{lst:arched_timeservice_ksh}, has to be used.
-It is recommended to examine the log file \textit{/var/log/arched.log} if the server configuration file has been modified to  assert the start was successful.
+It is recommended to examine the log file \textit{/var/log/arc/arched.log} if the server configuration file has been modified to  assert the start was successful.
 

### 2010-09-20 src/hed/daemon/win32/main_win32.cpp

@@ -74,17 +74,16 @@ static std::string init_logger(Arc::Config& cfg)
     if(log["Backups"]) {
       int backups;
       if(Arc::stringto((std::string)log["Backups"], backups)) {
-        sd->setBackups(backups);
+	sd->setBackups(backups);
       }   
     }   
     
     if(log["Maxsize"]) {
       int maxsize;
       if(Arc::stringto((std::string)log["Maxsize"], maxsize)) {
-        sd->setMaxSize(maxsize);
+	sd->setMaxSize(maxsize);
       }   
-    }
-
+    }    
     if(log["Reopen"]) {
       std::string reopen = (std::string)(log["Reopen"]);
       bool reopen_b = false;

### 2010-09-20 src/hed/libs/common/Logger.cpp

@@ -291,7 +291,6 @@ namespace Arc {
       return;
     }
   }
-
   void LogFile::setMaxSize(int newsize) {
     maxsize = newsize;
   }
@@ -299,7 +298,6 @@ namespace Arc {
   void LogFile::setBackups(int newbackups) {
     backups = newbackups;
   }
-
   void LogFile::setReopen(bool newreopen) {
     reopen = newreopen;
     if(reopen) {
@@ -353,7 +351,6 @@ namespace Arc {
     if (reopen) destination.close();
     backup();
   }
-
   void LogFile::backup(void) {
     if(maxsize <= 0) return;
     if(destination.tellp() < maxsize) return;

### 2010-09-20 src/hed/libs/common/Logger.h

@@ -341,6 +341,8 @@ namespace Arc {
      */
     void setBackups(int newbackup);
 
+
+
     //! Set file reopen on every write.
     /*! Set file reopen on every write. If set to true file is opened 
        before writing every log record and closed afterward.

### 2010-09-20 src/services/a-rex/grid-manager/grid_manager.cpp

@@ -34,7 +34,7 @@
 /* cache cleaning every 5 minutes */
 #define CACHE_CLEAN_PERIOD 300
 
-#define DEFAULT_LOG_FILE "/var/log/grid-manager.log"
+#define DEFAULT_LOG_FILE "/var/log/arc/grid-manager.log"
 #define DEFAULT_PID_FILE "/var/run/grid-manager.pid"
 
 namespace ARex {

### 2010-09-20 src/services/gridftpd/listener.cpp

@@ -16,7 +16,7 @@
 #define DEFAULT_BUFFER_SIZE (65536)
 #define DEFAULT_MAX_CONECTIONS (100)
 #define DEFAULT_GRIDFTP_PORT 2811
-#define DEFAULT_LOG_FILE "/var/log/gridftpd.log"
+#define DEFAULT_LOG_FILE "/var/log/arc/gridftpd.log"
 #define DEFAULT_PID_FILE "/var/run/gridftpd.pid"
 
 GridFTP_Commands *client;

### 2010-10-13 doc/tech_doc/a-rex/arex_tech_doc.tex
document new option for max jobs per DN

@@
 set accordingly. Default is \emph{no}.
-\item \textbf{\textit{maxjobs}}\textit{={[}max\_processed\_jobs {[}max\_running\_jobs]]}
--- specifies maximum number of jobs being processed by the A-REX at
-different stages:\\
+\item \textbf{\textit{maxjobs}}\textit{={[}max\_processed\_jobs
+{[}max\_running\_jobs {[}max\_jobs\_per\_dn]]]} -- specifies maximum number of
+jobs being processed by the A-REX at different stages:\\
 \textit{max\_processed\_jobs} -- maximum number of concurrent jobs
-processed by A-REX. This does not limit the amount of jobs, which can
+processed by A-REX. This does not limit the amount of jobs which can
 be submitted to the cluster.\\
-\textit{max\_running jobs} -- maximum number of jobs passed to Local
+\textit{max\_running\_jobs} -- maximum number of jobs passed to Local
 Resource Management System\\
-Missing value or -1 means no limit.
+\textit{max\_jobs\_per\_dn} -- maximum number of concurrent jobs processed by
+A-REX per user DN. If this option is used the total maximum number of jobs
+processed is still \textit{max\_processed\_jobs}.
+\\ Missing value or -1 means no limit.
 \item \textbf{\textit{maxload}}\textit{={[}max\_frontend\_jobs {[}emergency\_frontend\_jobs

### 2010-10-14 doc/tech_doc/chelonia/arc-storage-documentation.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
making the config tags more uniform

@@
     \item[StoreCfg] contains parameters for the different store classes, and it almost always contains a \emph{DataDir} parameter which specifies a directory on the (local) filesystem where the A-Hash can save its files.
-This variable is only configurable for the centralized deployment.
+For the TransDBStore and for the Replicated A-Hash also some additional options can be set here, e.g.
+\emph{CacheSize} configures the amount of the storage to be cached in memory.
 \end{description}
@@
 \begin{description}
-    \item[LocalDir] specifies a directory on the local filesystem where this A-Hash replica should be stored.
-This directory must be unique for this replica, i.e., no two A-Hash replicas can share the same \verb!LocalDir!
-since doing so will most likely leave both replicas broken beyond repair.
-    \item[MyURL] is the URL to this A-Hash service and is used as an ID for this service in the replication group.
-    \item[OtherURL] is the URL of another A-Hash in the replication group. More than one URL can be specified here.
+    \item[Endpoint] is the URL to this A-Hash service and is used as an ID for this service in the replication group.
+    \item[PeerURL] is the URL of another A-Hash in the replication group. More than one URL can be specified here.
 \end{description}
@@
     \item[CheckPeriod] specifies in seconds how frequently the A-Hash should update the replica list. The default is 10 seconds.
-    \item[CacheSize] configures the amount of the storage to be cached in memory. The default is 10MB
 \end{description}
@@
         <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
-        <LocalDir>ahash_data1</LocalDir>
-        <MyURL>http://localhost:60000/AHash1</MyURL>
-        <OtherURL>http://localhost:60001/AHash2</OtherURL>
+        <StoreCfg>
+            <DataDir>ahash_data1</DataDir>
+        </StoreCfg>
+        <Endpoint>http://localhost:60000/AHash1</Endpoint>
+        <PeerURL>http://localhost:60001/AHash2</PeerURL>
         <Priority>50</Priority>

### 2010-10-14 src/services/storage/ahash/replicatedahash.py

@@ -16,9 +16,11 @@ Sample configuration:
         <Service name="pythonservice" id="ahash">
             <ClassName>storage.ahash.ahash.AHashService</ClassName>
             <AHashClass>storage.ahash.replicatedahash.ReplicatedAHash</AHashClass>
-            <LocalDir>ahash_data</LocalDir>
-            <MyURL>http://localhost:60000/RepAHash</MyURL>
-            <OtherURL>http://otherhost:60000/RepAHash</OtherURL>
+            <StoreCfg>
+                <DataDir>ahash_data</DataDir>
+            </StoreCfg>
+            <Endpoint>http://localhost:60000/RepAHash</Endpoint>
+            <PeerURL>http://otherhost:60000/RepAHash</PeerURL>
         </Service>
 """
 import arc
@@ -168,16 +170,11 @@ class ReplicationStore(TransDBStore):
                       db.DB_INIT_MPOOL | \
                       db.DB_INIT_TXN    
 
-        storecfg = XMLTree(from_tree = 
-                           ['StoreCfg', 
-                            [('DataDir', 
-                              (str(cfg.Get('LocalDir')))),
-                              ('Priority', (int(str(cfg.Get('Priority'))))),
-                               ('CheckPeriod',(5)),
-                              ('DBEnvFlags',
-                               (str(dbenv_flags)))
-                            ]])
-        storecfg = arc.XMLNode(storecfg.pretty_xml()) 
+        storecfg = cfg.Get('StoreCfg')
+        if storecfg.Name() != 'StoreCfg': # StoreCfg missing, maybe old config, trying to create one from LocalDir
+            storecfg = arc.XMLNode('<StoreCfg><DataDir>%s</DataDir></StoreCfg>' % cfg.Get('LocalDir'))
+        if not str(storecfg.Get('DBEnvFlags')):
+            storecfg.NewChild('DBEnvFlags').Set(str(dbenv_flags))
 
         # db environment is opened in TransDBStore.__init__
         TransDBStore.__init__(self, storecfg, non_existent_object)
@@ -218,8 +215,10 @@ class ReplicationStore(TransDBStore):
 
 
     def __configureReplication(self, cfg):
-        self.my_url = str(cfg.Get('MyURL'))
-        self.other_urls = get_child_values_by_name(cfg, 'OtherURL')
+        self.my_url = str(cfg.Get('Endpoint'))
+        if not self.my_url: # backward compatibility with old config
[…]

### 2010-10-14 doc/tech_doc/sec/arc-security-documentation.tex
Enhancing delegation pictures. Fixing figures dependencies in makefile.
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
Spelling correction: It is not transfered but transferred.

@@
 \centering{{{\includegraphics[width=0.9\textwidth]{delegation_flow1.pdf}}}
-\caption{\label{fig:delegation_flow1}The flow diagram of delegation procedure with certificate transfered as payload of service specific message} }
+\caption{\label{fig:delegation_flow1}The flow diagram of delegation procedure with certificate transferred as payload of service specific message} }
 \end{figure}
@@
 \textbf{  } & \textbf{ARC Request/Response} & \textbf{XACML Request/Response} \\ \hline
-\textbf{ARC protocol} & support & support \\ \hline
-\textbf{SAML protocol} & not support & support (SAML 2.0 profile of XACML v2.0) \\ \hline
+\textbf{ARC protocol} & supported & supported \\ \hline
+\textbf{SAML protocol} & not supported & supported (SAML 2.0 profile of XACML v2.0) \\ \hline
 \end{tabular}

### 2010-10-14 doc/tech_doc/xrsl/xrsl-user.tex
Synchronising with ARC0 change - improved description of "notify" attribute
Spelling correction: It is not transfered but transferred.

@@
     starts with \texttt{gsiftp} or \texttt{ftp}, the whole directory
-    is transfered to the destination.\\
+    is transferred to the destination.\\
     \hspace*{1cm}&\texttt{URL} & URL of the remote file (see
@@
     \hspace*{1cm}&\texttt{string} & string of the format:
-    \verb#[b][q][f][e][c] user1@domain1 [user2@domain2] ...#\\
+    \verb#[b][q][f][e][c][d] user1@domain1 [user2@domain2] ...#\\
     \hspace*{1cm}&&here flags indicating the job status are:\\
@@
   \end{tabular}
+
+  When no notification flags are specified, default value of ``\verb#eb#''
+  will be used, i.e., notifications will be sent at the job's beginning
+  and at its end.
 

### 2010-10-14 doc/tutorials/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
Interchange DEBUG and VERBOSE loglevels.
Spelling correction: It is not transfered but transferred.

@@
  	\textbf{delegation.collector} & Processes a request for another service located on a different HED.
-                                        A certificate, which has a delegation policy embedded, is transfered to the service running that \textit{SecHandler}.
-Once the has been extracted and verified it will be transfered
+                                        A certificate, which has a delegation policy embedded, is transferred to the service running that \textit{SecHandler}.
+Once the has been extracted and verified it will be transferred
                                         to the requested service.
@@
  	\textbf{delegate.pdp}        & Similar to \textit{arc.pdp} a policy file will be evaluated. But here, the policy document is 
-                                       extracted and transfered
+                                       extracted and transferred
                                        by delegation.collector.\\
@@
 
-To get an impression in how the requests look like, one can check the log-file (the log level needs to be set to VERBOSE).
+To get an impression in how the requests look like, one can check the log-file (the log level needs to be set to DEBUG).
 

### 2010-10-14 doc/tutorials/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Move ARC to using Logrotate instead of built-in log rotation Move ARC logging default from /var/log to /var/log/arc/
Spelling correction: It is not transfered but transferred.

@@
 %          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
-% $ rm -f /var/log/arched.log
+% $ rm -f /var/log/arc/arched.log
 % $ arched -c arched_echoservice.xml  && echo jo ||echo n
-% $ tail -n100 -f /var/log/arched.log
+% $ tail -n100 -f /var/log/arc/arched.log
 % $ killall arched
@@
 Its content will be loaded in the code fragment starting at line~\ref{lst_code:tls_echo_client_cpp_getfilecontent}. 
-Later, in line~\ref{lst_code:tls_echo_client_cpp_loadconfiguration}, the configuration will be parsed into XML and than transfered into a chain of components which is wrapped by the object \textit{loader}.
+Later, in line~\ref{lst_code:tls_echo_client_cpp_loadconfiguration}, the configuration will be parsed into XML and than transferred into a chain of components which is wrapped by the object \textit{loader}.
 In order to process a message, an entry point within the chain has to be disclosed.

### 2010-10-14 src/services/a-rex/grid-manager/loaders/downloader.cpp

@@ -516,7 +516,7 @@ int main(int argc,char** argv) {
     mover.set_default_min_average_speed(min_average_speed);
   if(max_inactivity_time != 0)
     mover.set_default_max_inactivity_time(max_inactivity_time);
-  bool transfered = true;
+  bool transferred = true;
   bool credentials_expired = false;
   std::list<FileData> output_files;
 
@@ -665,10 +665,10 @@ int main(int argc,char** argv) {
     failure_reason+="Input file: "+i->lfn+" - "+(std::string)(i->res)+"\n";
     if(i->res == Arc::DataStatus::CredentialsExpiredError)
       credentials_expired=true;
-    transfered=false;
+    transferred=false;
   };
   // Check if all files have been properly downloaded
-  if(!transfered) {
+  if(!transferred) {
     logger.msg(Arc::INFO, "Some downloads failed"); res=2;
     if(credentials_expired) res=3;
     goto exit;

### 2010-10-14 src/services/a-rex/grid-manager/loaders/uploader.cpp

@@ -416,7 +416,7 @@ int main(int argc,char** argv) {
     mover.set_default_min_average_speed(min_average_speed);
   if(max_inactivity_time != 0)
     mover.set_default_max_inactivity_time(max_inactivity_time);
-  bool transfered = true;
+  bool transferred = true;
   bool credentials_expired = false;
   std::list<FileData>::iterator it = job_files_.begin();
   std::list<FileData>::iterator it2 = job_files_.begin();
@@ -570,10 +570,10 @@ int main(int argc,char** argv) {
     failure_reason+="Output file: "+i->lfn+" - "+(std::string)(i->res)+"\n";
     if(i->res == Arc::DataStatus::CredentialsExpiredError)
       credentials_expired=true;
-    transfered=false;
+    transferred=false;
   };
   // Check if all files have been properly uploaded
-  if(!transfered) {
+  if(!transferred) {
     logger.msg(Arc::INFO, "Some uploads failed"); res=2;
     if(credentials_expired) res=3;
     goto exit;

### 2010-10-14 doc/tech_doc/infosys/infosys_technical.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
Consolidating ISIS configuration while keeping it backward compatible

@@
 
-  <xsd:complexType name="InfoProvider_Type">
-    <xsd:sequence>
-      <xsd:element name="URL" type="xsd:anyURI">
+  <xsd:element name="InfoProvider" type="xsd:anyURI">
       <xsd:annotation>
@@
   </xsd:element>
-    </xsd:sequence>
-  </xsd:complexType>
-
-  <xsd:element name="InfoProvider" type="icfg:InfoProvider_Type"/>
 

### 2010-10-30 doc/tech_doc/a-rex/arex_tech_doc.tex
New limit on total number of jobs implemented

@@
 \item \textbf{\textit{maxjobs}}\textit{={[}max\_processed\_jobs
-{[}max\_running\_jobs {[}max\_jobs\_per\_dn]]]} -- specifies maximum number of
-jobs being processed by the A-REX at different stages:\\
+{[}max\_running\_jobs {[}max\_jobs\_per\_dn {[}max\_jobs\_total]]]]} -- specifies 
+maximum number of jobs being processed by the A-REX at different stages:\\
 \textit{max\_processed\_jobs} -- maximum number of concurrent jobs
@@
 A-REX per user DN. If this option is used the total maximum number of jobs
-processed is still \textit{max\_processed\_jobs}.
+processed is still \textit{max\_processed\_jobs}.\\
+\textit{max\_jobs\_total} -- total maximum number of jobs associated with 
+service. It is advised to use this limit only inexceptional case because
+it also accounts for finished jobs.\\
 \\ Missing value or -1 means no limit.

### 2010-11-09 doc/manuals/janitor/tex_installation/installation.tex
implement option to enable/disable janitor

@@
 corresponding file. If that variable is not set, the default location
-\textit{/etc/arc.conf} will be used.  The configuration is assigned
-by the section \lbrack janitor\rbrack. Table~\ref{tab:arcConfTags}
+\textit{/etc/arc.conf} will be used.  The parameter \texttt{use_janitor} in the
+\lbrack grid-manager\rbrack section has to be set in order to tell A-REX
+whether to use Janitor or not. By default, Janitor is not used. Use the value
+\texttt{"1"} to enable Janitor.
+
+Janitor is configured through parameters
+in the section \lbrack janitor\rbrack. Table~\ref{tab:arcConfTags}
 describes the available tags for the Janitor's configuration.
@@
         \hline
-           enabled         & "1"                                   & Boolean flag which enables or disables janitor in A-REX.\\
 	   uid             & "root"                                & The effective uid. \\
@@
 
-The parameter \texttt{enabled} specifies whether the Janitor shall be used
-within A-REX or not. This needs to be distinguished from the previously mentioned
-compilation flags of a similar name. Use the value \texttt{"0"} to disable Janitor. The
-\texttt{uid} and the \texttt{gid} are defining which effective user id
+The \texttt{uid} and the \texttt{gid} are defining which effective user id
 (uid) and group id (gid) shall be used for the Janitor.

### 2010-11-18 doc/tech_doc/a-rex/arex_tech_doc.tex
update instructions with how to run WS-interface through arc.conf

@@
 The A-REX service is a pluggable module of the HED, so it is first
-required to set up HED configuration, and add the A-REX elements. To
-add A-REX to an existing HED configuration, add a new \verb|<Name>|
-element inside \verb|<Plugins>| containing the string
-\emph{arex}. This will make HED load the libarex plugin library.
+required to set up HED configuration, and add the A-REX elements. HED
+configuration is in an XML format, however a special comand \emph{a-rex} has
+been written which automatically creates a HED configuration for running the
+A-REX from an existing arc.conf file used by the Grid Manager, and so no manual
+XML configuration editing needs to be done. To enable the WS interface the
+following line must be added to the arc.conf configuration file:
+
+\begin{itemize}
+  \item \textbf{\textit{arex\_mount\_point}}\textit{=path}
+\end{itemize}
+
+where \emph{path} is the A-REX service endpoint, for example
+\verb|https://your.host:60000/arex|. Then the A-REX can be started
+
+\begin{shaded}
+\begin{verbatim}
+$ARC_LOCATION/etc/init.d/a-rex start
+\end{verbatim}
+\end{shaded}
+
+It is also possible to write configuration directly in XML. To add A-REX to an
+existing HED configuration, add a new \verb|<Name>| element inside
+\verb|<Plugins>| containing the string \emph{arex}. This will make HED load the
+libarex plugin library.
 
@@
 
-The A-REX can process jobs submitted through a traditional GridFTP
-server instead of, or as well as, the WS interface. The GridFTP/Grid
-Manager server packages can be installed alongside A-REX packages, so
-it is easy to switch between the two. The configuration of the GridFTP
-server is in the same legacy \emph{arc.conf} configuration file as is
-used by the A-REX, and configuration instructions are given
-in~\cite{gm}. In general no configuration changes are necessary to use
-a Grid Manager-configured GridFTP server with the A-REX.
-
-If it is desired to run both WS and GridFTP interfaces at the same
-time, simply setup and start the A-REX as described in the previous
-section, then start the GridFTP server. If only the GridFTP interface
-is to be used, then the special \emph{a-rex} command should be used
-(rather than \emph{arched}).
-
-\begin{shaded}
-\begin{verbatim}
-$ARC_LOCATION/etc/init.d/a-rex start
-\end{verbatim}
-\end{shaded}
-
-This command automatically creates a HED configuration for running the
-A-REX without the WS interface, and so no manual HED configuration
-needs to be done. The GridFTP server should then be started.
+The A-REX can process jobs submitted through a traditional GridFTP interface
+instead of, or as well as, the WS interface. The A-REX version of the ARC GridFTP
+server (\verb|a-rex-gridftpd|) must be used instead of the Grid Manager server.
+However, the GridFTP/Grid Manager server packages can be installed alongside
+A-REX packages, so it is easy to switch between the two. The configuration of the
+GridFTP server is in the same legacy \emph{arc.conf} configuration file as is
+used by the A-REX, and configuration instructions are given in~\cite{gridftpd}.
+In general no configuration changes are necessary to use a Grid
+Manager-configured GridFTP server with the A-REX.
+
+If it is desired to run both WS and GridFTP interfaces at the same time, simply
+setup and start the A-REX as described in the previous section, then start the
+GridFTP server. If only the GridFTP interface is to be used, then the special
+\emph{a-rex} command should be used (rather than \emph{arched}) to start A-REX.
+The GridFTP server should then be started.
 

### 2010-12-07 doc/tech_doc/xrsl/xrsl-urls.tex
Synchronising changes in xrsl related changes

@@
    \texttt{ldap} & ordinary \textit{Lightweight Data Access Protocol (LDAP)}~\cite{ldap}\\
+   \texttt{srm} & Storage Resource Manager (SRM) service~\cite{srm}\\
+   \texttt{se} & ARC Smart Storage Element service~\cite{se}\\
+   \texttt{lfc} & LFC catalog and indexing service of EGEE gLite~\cite{glite}\\
+   \texttt{rc} & \globus\  \textit{Replica Catalog (RC)}~\cite{rc}\\
    \texttt{rls} & \globus\  \textit{Replica Location Service (RLS)}~\cite{rls}\\
-   \texttt{lfc} & LFC catalog and indexing service of EGEE gLite~\cite{glite}\\
-   \texttt{srm} & Storage Resource Manager (SRM) service~\cite{srm}\\
    \texttt{file} & local to the host file name with a full path\\
-   \texttt{arc} & for the Chelonia storage service, communicates with Bartenders, the path should be a Logical Name (LN)\\
 \end{tabular}
@@
 \begin{shaded}
-   \verb#protocol://[host[:port]]/file#
+   \verb#<protocol>://host[:port]/<file>#
 \end{shaded}
@@
 \begin{shaded}
-   \verb#protocol://[host[:port]][;option[;option[...]]]/file#
+   \verb#<protocol>://host[:port][;option[;option[...]]]/<file>#
 \end{shaded}
@@
 \begin{shaded}
-   \verb#protocol://[url[|url[...]]@]host[:port][;option[;option[...]]/#\\
-   \verb#     lfn[:metadataoption[:metadataoption[...]]#
+   \verb#lfc://[url[|url[...]]@]<host>[:port]/<lfn>#\\
+   \verb#rls://[url[|url[...]]@]<host>[:port]/<lfn>#\\
+   \verb#rc://rc://[location[|location[...]]@]<host>[:port]/<DN>/<lfn>#\\
+   \verb#fireman://[url[|url[...]]@]<host>[:port]/<service_path>?<lfn>#
 \end{shaded}
 
-For Chelonia, the syntax is
+For the Smart Storage Element service, the syntax is
 \begin{shaded}
-    \verb#arc://<LogicalName>[?BartenderURL=<URL>]#
+   \verb#se://host[:port][;options]/path[?file_id]#
 \end{shaded}
-where the BartenderURL could come from the `bartender' parameter of the client configuration file.
 
@@
 \begin{shaded}
-   \verb#srm://host[:port][;options]/[service_path?SFN=]file#
+   \verb#srm://<host>[:port][;options]/[service_path?SFN=]<file_id>#
 \end{shaded}
@@
 \begin{tabular}{lp{10cm}}
-   \verb#host[:port]# & Hostname or IP address [and port] of a server\\
-   \verb#lfn# & Logical File Name\\
-   \verb#url# & URL of the file as registered in indexing service\\
-   \verb#service_path# & End-point path of the web service\\
-   \verb#file# & File name with full path\\
-   \verb#option# & URL option\\
-   \verb#metadataoption# & Metadata option for indexing service\\
+   \verb#location#&\verb#<location_name_in_RC>[;option[;option[...]]]#\\
+   \verb#host[:port]#&IP address of a server\\
+   \verb#DN#&Distinguished Name (as in LDAP) of an RC collection\\
+   \verb#lfn#&Logical File Name (LFN)\\
+   \verb#url#&URL of the file as registered in LFC/RLS/Fireman\\
+   \verb#service_path#&End-point path of the Web service\\
+   \verb#file#&local to the host file name with a full path\\
 \end{tabular}
@@
    \verb#copy# forces the cached file to be copied (rather than linked) to
-   the session dir, this is useful if for example the file is to be modified.\\
+   the session dir, this is useful if for example the file is to be modified (\verb#copy# option is available in ARC 0.8.1 and above).\\
    \verb#readonly=yes|no# & for transfers to \verb#file://# destinations,
@@
    not provided, the default for the protocol is used. \verb#checksum=no#
-   disables checksum calculation.\\
+   disables checksum calculation (\verb#adler32# and \verb#no# options are available in ARC 0.8.1 and above).\\
    \verb#exec=yes|no# & means the file should be treated as executable\\
@@
    destination even if job processing failed (default is \verb#no#)\\
+   \verb#pattern=<pattern># & defines file matching pattern; currently
+   works for file listing requests sent to an \verb#se://# endpoint\\
    \verb#guid=yes|no# & make software use GUIDs instead of LFNs while
@@
    functions to wrap data and \verb#gsi# uses additional headers as
-   implmented in Globus IO. The default is \verb#gssapi#. In case
-   this fails, \verb#gsi# is then tried.\\
+   implmented in Globus IO\\
    \verb#spacetoken=<pattern># & specify the space token to be used for
@@
    protocols is \verb#yes#\\
-   \verb#tcpnodelay=yes|no# & controls the use of the TCP\_NODELAY
-   socket option (which disables the Nagle algorithm). Applies to
-   http(s) only. Default is \verb#no#\\ 
 \end{longtable}
@@
    \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
+   \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#\\
    \verb#rls://gsiftp://se.domain.org/datapath/file25.dat@grid.domain.org:61238/myfile02.dat#$^1$\\
+   \verb#fireman://fireman_host:8443/glite-data-catalog-interface/FiremanCatalog?data.root#\\
    \verb#file:///home/auser/griddir/steer.cra#\\
-   \verb#lfc://srm://srm.domain.org/griddir@lfc.domain.org/user/file1:guid=\# \\
+   \verb#lfc://srm://srm.domain.org/griddir@lfc.domain.org//user/file1:guid=\# \\
    \verb#   bc68cdd0-bf94-41ce-ab5a-06a1512764dc:checksumtype=adler32:checksumvalue=12345678#$^2$\\
-   \verb#lfc://lfc.domain.org;cache=no/:guid=bc68cdd0-bf94-41ce-ab5a-06a1512764d#$^3$\\
+   \verb#lfc://lfc.domain.org;cache=no/:guid=bc68cdd0-bf94-41ce-ab5a-06a1512764d#$^3$
 \end{framed}

### 2010-12-07 doc/tech_doc/xrsl/xrsl-user.tex
Synchronising changes in xrsl related changes

--- 2009-04-28 doc/tech_doc/xrsl/xrsl-user.tex
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(executable=<string>)#\\
@@
   the source is expected to be local to the user (client) and
-  will be added as such to the \texttt{inputFiles} list by the UI.
+  will be added as such to the \texttt{inputFiles} list by the ARC Client.
 
@@
 
-  For ARC $\leq$ 0.5.25, UI created a dummy executable attribute for internal GM use,
-  with e.g., "\verb#/bin/echo#" as an argument (unless the
-  specified executable itself started with "\verb#/#"). The actual
-  executable was transferred by the UI to
-  \texttt{arguments}, \texttt{inputFiles} and \texttt{executables} lists
-  in the pre-processed xRSL script.
-
   \hspace*{0.5cm}\phantomsection
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(arguments=<string> [string] ... )#\\
@@
     \hspace*{1cm}&\texttt{executable}&the executable to be run by LRMS, taken by the
-    UI from the user-specified \texttt{executable} attribute\\
-  \end{tabular}
-
-  For ARC $\leq$ 0.5.25, UI added the name specified in \texttt{executable} to the 
-  \texttt{arguments} list in the pre-processed xRSL script.
+    ARC Client from the user-specified \texttt{executable} attribute\\
+  \end{tabular}
 
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
   \begin{tabular}{lp{12cm}}
-    User input:&\verb#(inputFiles=(<filename> <location>) ... )#\\
+    Unique:&yes\\
+    Operators:&\verb#=#\\
+    User input:&\verb#(inputFiles=(<filename> <source>) ... )#\\
     GM input:&\verb#(inputFiles=(<filename> <URL>)#\\
@@
   \begin{tabular}{llp{10cm}}
-    \hspace*{1cm}&\texttt{filename} & file name, local to the computing element and
+    \hspace*{1cm}&\texttt{filename} & destination file name, local to the computing element and
     always relative to the session directory\\
-    \hspace*{1cm}&\texttt{location} & location of the file (\texttt{gsiftp},
+    \hspace*{1cm}&\texttt{source} & source of the file: (\texttt{gsiftp},
     \texttt{https}, \texttt{ftp}, \texttt{http} URLs, or a path, local
@@
 
+  \begin{framed}
+   ARC Client does not check whether the same destination name is specified for different source locations. Make sure all \texttt{filename} values are different!
+  \end{framed}
+
   If the list does not contain the standard input file (as specified
   by \texttt{stdin}) and/or the executable file (as specified by
-  \texttt{executable} if the given name), the UI appends these files
-  to the list. If the \verb#<location># is a URL, it is passed by the
-  UI to the GM without changes; if it is a local path (or void, ""),
-  the UI converts it to \verb#<size># and \verb#<checksum># and
+  \texttt{executable} if the given name), the ARC Client appends these files
+  to the list. If the \verb#<source># is a URL, it is passed by the
+  ARC Client to the GM without changes; if it is a local path (or void, ""),
+  the ARC Client converts it to \verb#<size># and \verb#<checksum># and
   uploads those files to the execution cluster.
@@
   Please note that the \texttt{inputFiles} attribute is not meant to
-  operate with directories: you must specify a pair
-  \verb#("<local_to_job>" "[source]")# for each file.
+  operate with directories, for reasons of access control and checksum verifications. You must specify a pair
+  \verb#("<local_to_job>" "<source>")# for each file.
   \end{framed}
 
-  For ARC $\leq$ 0.5.25, UI added the name specified in \texttt{executable} to the 
-  \texttt{inputFiles} list in the pre-processed xRSL script.
-
   \hspace*{0.5cm}\phantomsection
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(executables=<string> [string] ...)#\\
@@
   relative to the session directory)  is not listed, it will be added
-  to the list by the UI.
+  to the list by the ARC Client.
 
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.34, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.34, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(cache="yes"|"no")#\\
@@
   \begin{framed} 
-    Cached files can not be modified by jobs by default. If your job has to 
-    modify input files, please specify \texttt{(cache="no")} or use
-    \texttt{(readonly="no")} URL option for those files.
+  Cached files can not be modified by jobs by default. If your job has
+  to modify input files, please use the \texttt{(readonly="no")} URL
+  option for those files. This option does not affect whether or not
+  the file is cached.
   \end{framed}
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(outputFiles=(<string> <URL>) ... )#\\
@@
     is transfered to the destination.\\
-    \hspace*{1cm}&\texttt{URL} & URL of the remote file (see
+    \hspace*{1cm}&\texttt{URL} & destination URL of the remote file (see
     Section~\ref{sec:url}); if void ("", use the quotes!), the file is
@@
   file names and GM log-files directory name (as specified by \texttt{stdout},
-  \texttt{stderr} and \texttt{gmlog}), the UI appends these
+  \texttt{stderr} and \texttt{gmlog}), the ARC Client appends these
   items to the \texttt{outputFiles} list. If the \verb#<URL># is not specified (void,
   "", use the quotes!), files will be kept on SE and should be downloaded 
-  by the user via the UI. If specified name of file ends with "/", the 
+  by the user via the ARC Client. If specified name of file ends with "/", the 
   entire directory is kept.
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.17\footnote{Was called \texttt{maxCpuTime} for ARC $\leq$ 0.3.16.}, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.17, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(cpuTime=<time>)#\\
@@
 
-  Maximal CPU time request for the job.
-
-  \begin{tabular}{llp{10cm}}
-    \hspace*{1cm}&\texttt{time} & time (minutes)\\
-    \hspace*{1cm}&\texttt{tttt} & time converted by the UI from \texttt{time} to 
-    minutes (ARC $\leq$ 0.5.26) or seconds (ARC $\geq$ 0.5.27)\\
-  \end{tabular}
-
-  If only number is specified, the time is assumed to be minutes (ARC
-  $\leq$ 0.5.26) or seconds (ARC $\geq$ 0.5.27). Otherwise, a free
-  format is accepted, i.e., any of the following will be interpreted
+  Maximal CPU time request for the job. For a multi-processor job, this is a sum over all requested processors.
+
+  \begin{tabular}{llp{10cm}}
+    \hspace*{1cm}&\texttt{time} & time (in minutes if no unit is specified)\\
+    \hspace*{1cm}&\texttt{tttt} & time converted by the ARC Client from \texttt{time} to 
+    seconds.\\
+  \end{tabular}
+
+  The client converts time specified in the user-side XRSL file to seconds. 
+  If no time unit is specified, the client assumes the time given in minutes. 
+  Otherwise, a text format is accepted, i.e., any of the following will be interpreted
   properly (make sure to enclose such strings in quotes!):
@@
 
-  If both \texttt{cpuTime} and \texttt{wallTime} are specified, the UI converts them both to 
-  seconds. \texttt{cpuTime} can not be specified together with \texttt{gridTime}
+  If both \texttt{cpuTime} and \texttt{wallTime} are specified, the ARC Client converts them 
+  both. \texttt{cpuTime} can not be specified together with \texttt{gridTime}
   or \texttt{benchmarks}.
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.5.27, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.5.27, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(wallTime=<time>)#\\
@@
   \begin{tabular}{llp{10cm}}
-    \hspace*{1cm}&\texttt{time} & time (minutes)\\
-    \hspace*{1cm}&\texttt{tttt} & time converted by the UI from \texttt{time} to seconds\\
-  \end{tabular}
-
-  If only number is specified, the time is assumed to be
-  seconds. Otherwise, a free format is accepted, i.e., any of the
-  following will be interpreted properly (make sure to enclose such
-  strings in quotes!):
+    \hspace*{1cm}&\texttt{time} & time (in minutes if no unit is specified)\\
+    \hspace*{1cm}&\texttt{tttt} & time converted by the ARC Client to seconds\\
+  \end{tabular}
+
+  The client converts time specified in the user-side XRSL file seconds. If no time unit is specified,
+  the client assumes the time given in minutes. Otherwise, a text
+  format is accepted, i.e., any of the following will be interpreted
+  properly (make sure to enclose such strings in quotes!):
 
@@
 
-  If both \texttt{cpuTime} and \texttt{wallTime} are specified, the UI converts them both to 
-  seconds. \texttt{wallTime} can not be specified together with \texttt{gridTime}
+  If both \texttt{cpuTime} and \texttt{wallTime} are specified, the ARC Client converts them both.
+  \texttt{wallTime} can not be specified together with \texttt{gridTime}
   or \texttt{benchmarks}. If only \texttt{wallTime} is specified, but not \texttt{cpuTime},
-  the corresponding  \texttt{cpuTime} value is evaluated by the UI and added to the job description.
+  the corresponding  \texttt{cpuTime} value is evaluated by the ARC Client and added to the job description.
 
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.31, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.31, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(gridTime=<time>)#\\
@@
   \begin{tabular}{llp{10cm}}
-    \hspace*{1cm}&\texttt{time} & time (minutes)\\
+    \hspace*{1cm}&\texttt{time} & time (in minutes if no unit is specified)\\
   \end{tabular}
@@
   or \texttt{wallTime}. If only \texttt{gridTime} is specified, but not \texttt{cpuTime},
-  the corresponding  \texttt{cpuTime} value is evaluated by the UI and added to the job description.
+  the corresponding  \texttt{cpuTime} value is evaluated by the ARC Client and added to the job description.
 
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.5.14, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.5.14, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(benchmarks=(<string> <value> <time>) ... )#\\
@@
   or \texttt{wallTime}. If only \texttt{benchmarks} is specified, but not \texttt{cpuTime},
-  the corresponding  \texttt{cpuTime} value is evaluated by the UI and added to the job description.
+  the corresponding  \texttt{cpuTime} value is evaluated by the ARC Client and added to the job description.
 
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.17\footnote{Was called \texttt{maxMemory} for ARC $\leq$ 0.3.16}, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.17\footnote{Was called \texttt{maxMemory} for ARC $\leq$ 0.3.16}, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(memory=<integer>)#\\
@@
 
-  Memory required for the job.
+  Memory required for the job, per rank.
 
@@
   \begin{framed}
-    Just like \texttt{cpuTime}, this attribute should be used to
+    Similarly to \texttt{cpuTime}, this attribute should be used to
     direct a job to a resource with a sufficient capacity. Jobs
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.17\footnote{Was called \texttt{maxDisk} for ARC $\leq$ 0.3.16}, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.17\footnote{Was called \texttt{maxDisk} for ARC $\leq$ 0.3.16}, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=   !=   >   <   >=   <=#\\
     User input:&\verb#(disk=<integer>)#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=   !=   >   <   >=   <=#\\
     User input:&\verb#(runTimeEnvironment=<string>)(runTimeEnvironment=<string>)#\\
@@
 
-  The site to submit the job to will be chosen by the UI among those
+  The site to submit the job to will be chosen by the ARC Client among those
   advertising specified runtime environments. Before starting the job,
@@
   \verb#(runTimeEnvironment="ENV1")(runTimeEnvironment="ENV2")# etc.
-  In ARC $\leq$ 0.5.39, before being submitted to the GM, a grouping was done by the UI, such that GM input was,
-  e.g., \verb#(runTimeEnvironment="ENV1" "ENV2")#.
 
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=   !=   >   <   >=   <=#\\
     User input:&\verb#(middleware=<string>)#\\
@@
 
-  Required middleware.
+  Required middleware version. Make sure to specify full name and version number.
 
@@
 
-  The site to submit the job to will be chosen by the UI among those
+  The site to submit the job to will be chosen by the ARC Client among those
   advertising specified middleware. Usage is identical to that of the
   \texttt{runTimeEnvironment}. Use the "\verb#>=#" operator to
-  request a version "equal or higher". Request
-  \texttt{(middleware="nordugrid-arc")} defaults to
-  \verb#(middleware>="nordugrid-arc-0.0.0.0")#.
+  request a version "equal or higher".% Not true (see bug 1485): Request
+  %\texttt{(middleware="nordugrid-arc")} defaults to
+  %\verb#(middleware>="nordugrid-arc-0.0.0.0")#.
 
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.21, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.21, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=   !=   >   <   >=   <=#\\
     User input:&\verb#(opsys=<string>)#\\
@@
 
-  The site to submit the job to will be chosen by the UI among those
+  The site to submit the job to will be chosen by the ARC Client among those
   advertising specified operating system. Usage is identical to that of
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(stdin=<string>)#\\
@@
   The standard input file should be listed in the \texttt{inputFiles}
-  attribute; otherwise it will be forced to that list by the UI.
+  attribute; otherwise it will be forced to that list by the ARC Client.
 
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(stdout=<string>)#\\
@@
   \texttt{outputFiles} attribute; otherwise it will be forced to that
-  list by the UI. If the standard output is not defined, UI assigns a
+  list by the ARC Client. If the standard output is not defined, ARC Client assigns a
   name.
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(stderr=<string>)#\\
@@
   The standard error file should be listed as an \texttt{outputFiles}
-  attribute; otherwise it will be forced to that list by the UI. If
-  the standard error is not defined, UI assigns a name. If \texttt{join} is 
-  specified with value "yes", UI adds \texttt{stderr} to the 
+  attribute; otherwise it will be forced to that list by the ARC Client. If
+  the standard error is not defined, ARC Client assigns a name. If \texttt{join} is 
+  specified with value "yes", ARC Client adds \texttt{stderr} to the 
   pre-processed xRSL script with the same value as \texttt{stdout}.
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(join="yes"|"no")#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(gmlog=<string>)#\\
@@
 %  The job log file should be listed as an \texttt{outputFiles} attribute;
-%  otherwise it will be forced to that list by the UI.
+%  otherwise it will be forced to that list by the ARC Client.
 	This directory is kept in the session directory to be
-	available for retrieval (UI forces it to the list if
+	available for retrieval (ARC Client forces it to the list if
 	\texttt{outputFiles})
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(jobName=<string>)#\\
@@
   This name is meant for convenience of the user. It can be used to
-  select the job while using the UI. It is also available through the
+  select the job while using the ARC Client. It is also available through the
   Information System.
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(ftpThreads=<integer>)#\\
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.5.12, ARC 0.6}
-
-  \textit{Available in ARC v0.5.12 and higher}.
-
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.5.12, ARC 0.6, ARC 0.8}
+
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=#\\
     User input:&\verb#(acl=<xml>)#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=   !=#\\
     User input:&\verb#(cluster=<string>)#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=   !=#\\
     User input:&\verb#(queue=<string>)#\\
@@
 
-  The name of the remote batch queue. Use only if you are sure that
-  the queue by this name does exist.
+  The name of the remote batch queue.
+
+  \begin{framed}
+    Use only when you are sure that the queue by this name does exist.
+  \end{framed}
 
@@
 
-  Use this attribute to explicitly force job submission to a queue.
+  While users are not expected to specify \texttt{queue} in job descriptions, this attribute \textbf{must} be present in the GM-side xRSL.
+In fact, this is primarily an internal attribute, added to the job description by client tools after resource discovery and matchmaking.
+Still, users can specify this attribute to explicitly force job submission to a queue: when specified explicitly by the user, this value will not be overwritten by the ARC Client, and an attempt will be made to submit the job to the specified queue.
+
+  If for some reason (e.g.
+due to a client tool error) \texttt{queue} is absent from the GM-side xRSL, GM on the selected cluster will attempt to submit the job to the default queue if such is specified in the GM configuration.
 
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(startTime=<time>)#\\
@@
     \hspace*{1cm}&\texttt{time} & time string, YYYY-MM-DD hh:mm:ss\\ 
-    \hspace*{1cm}&\texttt{tttt} & time string, YYYYMMDDhhmmss[Z] (converted by the UI from
+    \hspace*{1cm}&\texttt{tttt} & time string, YYYYMMDDhhmmss[Z] (converted by the ARC Client from
     \texttt{time})\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(lifeTime=<time>)#\\
     GM input:&\verb#(lifeTime=<tttt>)#\\
-    Example:&\verb#(lifeTime="60")#\\
+    Example:&\verb#(lifeTime="2 weeks")#\\
   \end{tabular}
@@
   \begin{tabular}{llp{10cm}}
-    \hspace*{1cm}&\texttt{time} & time (days)\\
-    \hspace*{1cm}&\texttt{tttt} & time (seconds, converted by the UI from \texttt{time})\\
+    \hspace*{1cm}&\texttt{time} & time (in minutes if no unit is specified)\\
+    \hspace*{1cm}&\texttt{tttt} & time (seconds, converted by the ARC Client from \texttt{time})\\
   \end{tabular}
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(notify=<string> [string] ... )#\\
@@
 
+  When no notification flags are specified, default value of ``\verb#eb#''
+  will be used, i.e., notifications will be sent at the job's beginning
+  and at its end.
+
   No more than 3 e-mail addresses per status change accepted.
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=#\\
     User input:&\verb#(replicaCollection=<URL>)#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(rerun=<integer>)#\\
@@
   If not specified, the default is 0. Default maximal allowed value is
-  5. Before 0.5.29 jobs were rerun automatically and only in case of
-  failure in INLRMS state. Since 0.5.29 job may be rerun after failure
-  in any state for which reruning has sense. To initiate rerun user
-  has to use \texttt{ngresume} command.
+  5. The job may be rerun after failure in any state for which reruning 
+  has sense. To initiate rerun user has to use the \texttt{ngresume} command.
 
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=   !=#\\
     User input:&\verb#(architecture=<string>)#\\
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.24, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.24, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(nodeAccess="inbound"|"outbound")#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(dryRun="yes"|"no")#\\
@@
 
-  If "yes", do dry-run: RSL is parsed, but no job submission to LRMS
-  is made. Should be used for xRSL and communication validation.
+  If "yes", do dry-run: job description is sent to the optimal destination, input files are transferred, but no actual job submission to LRMS
+  is made. Typically used for xRSL and communication validation.
 
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=#\\
     User input:&\verb#(rsl_substitution=(<string1> <string2>))#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&no\\
+    Operators:&\verb#=#\\
     User input:&\verb#(environment=(<VAR> <string>) [(<VAR> <string>)] ... )#\\
@@
 %%   \begin{tabular}{lp{13cm}}
+%%    Unique:&yes\\
+%%    Operators:&\verb#=#\\
 %%     User input:&\verb#(jobtype= "single"|"multiple"|"mpi"|"condor" )#\\
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.12, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.12, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(count=<integer>)#\\
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.34, ARC 0.4, ARC 0.5, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.3.34, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(jobreport=<URL>)#\\
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.5.45, ARC 0.6}
-  \begin{tabular}{lp{13cm}}
+  \versions{ARC $\geq$ 0.5.45, ARC 0.6, ARC 0.8}
+  \begin{tabular}{lp{13cm}}
+    Unique:&yes\\
+    Operators:&\verb#=#\\
     User input:&\verb#(credentialserver=<URL>)#\\
@@
   credentials is used instead.
+
+%  
+% To uncomment when implemented in clients; more info needed
+%
+%   \hspace*{0.5cm}\phantomsection
+%   \addcontentsline{toc}{subsection}{sessiondirectorytype}\index{XRSL:attribute:sessiondirectorytype}
+%   \begin{shaded}
+%     \xrsl{sessiondirectorytype}
+%   \end{shaded}
+%   \versions{not supported yet}
+%   \begin{tabular}{lp{13cm}}
+%    Unique:&yes\\
+%    Operators:&\verb#=#\\
+%     User input:&\verb#(sessiondirectorytype="readonly"|"limited"|"internal"})#\\
+%     GM input:&\verb#-"-#\\
+%     Example:&\verb#(sessiondirectorytype="limited")#\\
+%   \end{tabular}
+% 
+%   Instructs Grid Manager to set specific access permissions for the session directory.
+Normally, session directory is read-only for external processes.
+In case of interactive jobs, however, it may be necessary to allow writing from an external client.
+% 
+%   \begin{tabular}{llp{10cm}}
+%     \hspace*{1cm}&\texttt{readonly} & Default value, session directory is read-only\\
+%     \hspace*{1cm}&\texttt{limited} & \\
+%     \hspace*{1cm}&\texttt{internal} & \\
+%   \end{tabular}

### 2010-12-09 doc/tech_doc/a-rex/arex_tech_doc.tex
Modified code and documentation about (semi-)static information published by AREX. By default this feature is disable.

@@
       <arex:gmconfig>/etc/arc.conf</arex:gmconfig>
+      <!-- This optional parameter can be used to enable publishing of additional
+      information to ISIS. The default is not to publish ("no"). The information,
+      which is considered in some degree to be static, includes HealthState, 
+      OSFamily, Platform, PhysicalCPUs, CPUMultiplicity, CPUModel and
+      ApplicationEnvironment.--> 
+      <arex:publishStaticInfo>yes<arex:publishStaticInfo>
+
     </loader:Service>
@@
   <xsd:element name="OperatingSystem" type="xsd:string"/>
+  <!--
+  Optional parameter which is enable to publish additional
+  information to ISIS. 
+  {*} yes - Allow publishing additional information.
+  {*} no  - Disallow additional information to publish. 
+            Registered with Only contact information.  
+  -->  
+    <xsd:element name="publishStaticInfo" default="no">
+      <xsd:simpleType>
+        <xsd:annotation>
+          <xsd:documentation xml:lang="en">
+            This optional parameter can be used to enable publishing
+            of additional information to ISIS. The default is not to
+            publish ("no"). The information, which is considered in
+            some degree to be static, includes HealthState, OSFamily,
+            Platform, PhysicalCPUs, CPUMultiplicity, CPUModel and
+            ApplicationEnvironment.
+          </xsd:documentation>
+        </xsd:annotation>
+        <xsd:restriction base="xsd:string">
+          <xsd:enumeration value="no">
+          <xsd:enumeration value="yes">
+        </xsd:restriction>
+      </xsd:simpleType>
+    </xsd:element>
+
 </xsd:schema>

### 2010-12-12 doc/tech_doc/a-rex/arex_tech_doc.tex
Explanation of GRID_GLOBAL_JOBID and few more words about job environment

@@
 reached state \textbf{InLRMS})
-\item \textit{globalid} -- BES ActivityIdentifier XML tree, the global
-identifier of the job
+\item \textit{globalid} -- BES ActivityIdentifier XML tree or GridFTP URL,
+ the global identifier of the job for accessing it from outside
 \item \textit{args} -- executable name followed by a list of command-line
@@
 For publicly available runtime environments please see the RTE repository at \url{http://gridrer.csc.fi/}.
+
+\section{Job environment\label{sec:job}}
+
+Because setup of resources represented by A-REX may vary a lot it is important to 
+understand which features of execution environment to expect. A-REX ensures 
+following:
+
+\begin{list}{--}{\setlength{\labelwidth}{0.5cm}\setlength{\rightmargin}{\leftmargin}}
+\item Specified executable is started with current directory set to session directory.
+\item Executable is started under account which owns session directory.
+\item Due to caching input files are not always writable by job's local account unless
+\emph{readonly=no} URL option is specified for those files.
+\item Environment variable GRID\_GLOBAL\_JOBID contains job's identifier usable for 
+accessing job from outside through A-REX. Currently that is either XML of Activity
+Identifier ot GridFTP URL depending on used interface.
+\item Subject to requested Runtime Environments corresponding environment variables
+are set and executable and data packages are made available.
+\end{list}
 

### 2010-12-13 doc/tech_doc/xrsl/xrsl-urls.tex
Recovering changes lost in previous commit, making a merged arc0-arc1 version of URLs

--- 2010-05-26 doc/tech_doc/xrsl/xrsl-urls.tex
@@
    \texttt{lfc} & LFC catalog and indexing service of EGEE gLite~\cite{glite}\\
+%   \texttt{rc} & \globus\  \textit{Replica Catalog (RC)}~\cite{rc}\\
    \texttt{srm} & Storage Resource Manager (SRM) service~\cite{srm}\\
+%   \texttt{se} & ARC Smart Storage Element service~\cite{se}\\
+   \texttt{arc} & for the Chelonia storage service, communicates with Bartenders, the path should be a Logical Name (LN) (supported only in \verb#arcls#, \verb#arcsub# and other \verb#arc*# tools.)\\
    \texttt{file} & local to the host file name with a full path\\
-   \texttt{arc} & for the Chelonia storage service, communicates with Bartenders, the path should be a Logical Name (LN)\\
 \end{tabular}
@@
 \begin{shaded}
+%   \verb#lfc://[url[|url[...]]@]<host>[:port]/<lfn>#\\
+%   \verb#rls://[url[|url[...]]@]<host>[:port]/<lfn>#\\
+%   \verb#rc://rc://[location[|location[...]]@]<host>[:port]/<DN>/<lfn>#\\
+%   \verb#fireman://[url[|url[...]]@]<host>[:port]/<service_path>?<lfn>#
    \verb#protocol://[url[|url[...]]@]host[:port][;option[;option[...]]/#\\
@@
 
-For Chelonia, the syntax is
+For %the Smart Storage Element Service
+Chelonia, the syntax is
 \begin{shaded}
+%   \verb#se://host[:port][;options]/path[?file_id]#
     \verb#arc://<LogicalName>[?BartenderURL=<URL>]#
 \end{shaded}
-where the BartenderURL could come from the `bartender' parameter of the client configuration file.
+where the \verb#BartenderURL# may come from the \verb#bartender# parameter of
+the client configuration file. Note that Chelonia URLs are \textbf{not}
+supported in \verb#ngls#, \verb#ngsub# and other \verb#ng*# tools.
 
@@
 \begin{tabular}{lp{10cm}}
-   \verb#host[:port]# & Hostname or IP address [and port] of a server\\
-   \verb#lfn# & Logical File Name\\
-   \verb#url# & URL of the file as registered in indexing service\\
-   \verb#service_path# & End-point path of the web service\\
-   \verb#file# & File name with full path\\
+%   \verb#location#&\verb#<location_name_in_RC>[;option[;option[...]]]#\\
+   \verb#host[:port]#&Hostname or IP address [and port] of a server\\
+%   \verb#DN#&Distinguished Name (as in LDAP) of an RC collection\\
+   \verb#lfn#&Logical File Name\\
+   \verb#url#&URL of the file as registered in indexing service\\
+   \verb#service_path#&End-point path of the web service\\
+   \verb#file#&File name with full path\\
    \verb#option# & URL option\\
@@
    \verb#copy# forces the cached file to be copied (rather than linked) to
-   the session dir, this is useful if for example the file is to be modified.\\
+   the session dir, this is useful if for example the file is to be modified
+   (\verb#copy# option is available in ARC 0.8.1 and above).\\
    \verb#readonly=yes|no# & for transfers to \verb#file://# destinations,
@@
    not provided, the default for the protocol is used. \verb#checksum=no#
-   disables checksum calculation.\\
+   disables checksum calculation (\verb#adler32# and \verb#no# options are
+   available in ARC 0.8.1 and above).\\
    \verb#exec=yes|no# & means the file should be treated as executable\\
@@
    destination even if job processing failed (default is \verb#no#)\\
+%   \verb#pattern=<pattern># & defines file matching pattern; currently
+%   works for file listing requests sent to an \verb#se://# endpoint\\
    \verb#guid=yes|no# & make software use GUIDs instead of LFNs while
@@
    socket option (which disables the Nagle algorithm). Applies to
-   http(s) only. Default is \verb#no#\\ 
+   http(s) only. Default is \verb#no# (supported only in \verb#arcls# and
+   other \verb#arc*# tools)\\
 \end{longtable}
@@
    \verb#ldap://grid.domain.org:389/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org#\\
+% \verb#rc://grid.domain.org/lc=collection1,rc=Nordugrid,dc=nordugrid,dc=org/zebra/f1.zebra#\\
    \verb#rls://gsiftp://se.domain.org/datapath/file25.dat@grid.domain.org:61238/myfile02.dat#$^1$\\
+% \verb#fireman://fireman_host:8443/glite-data-catalog-interface/FiremanCatalog?data.root#\\
    \verb#file:///home/auser/griddir/steer.cra#\\

### 2010-12-13 doc/tech_doc/xrsl/xrsl-user.tex
Recovering changes lost in previous commit, making a merged arc0-arc1 version of URLs

@@
     starts with \texttt{gsiftp} or \texttt{ftp}, the whole directory
-    is transfered to the destination.\\
+    is transferred to the destination.\\
     \hspace*{1cm}&\texttt{URL} & destination URL of the remote file (see

### 2010-12-16 doc/tech_doc/backend_interface/Backends-arc1.tex
Reorganizing doc catalogue and enabling make doc
Keep documentation up-to-date to the GLUE2 schema
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
Update backends doc: GLUE2 is well supported now

--- 2008-08-12 doc/backend_interface/Backends-arc1.tex
@@
   pdfauthor = {T.Frågåt},
-  pdftitle ={Arc batch system back-end interface guide with support for GLUE 2},
+  pdftitle ={Arc batch system back-end interface guide with support for GLUE2},
   pdfsubject = {Description and developer's guide for ARC1},
@@
 generates an XML output to \texttt{STDOUT} by one invocation of the Perl script named
-\texttt{cluster+qju.pl}. The XML output includes two representations of the output data,
+\texttt{CEinfo.pl}. The XML output includes two representations of the output data,
 namely the classic NorduGrid information schema and an incomplete GLUE2 schema representation.
@@
 \begin{itemize}
-\item \texttt{cluster+qju.pl} - driver for information collection.
+\item \texttt{CEinfo.pl} - driver for information collection.
 It calls all other information collectors and prints the results in XML.
@@
 ARC1 provides support for both the classic NorduGrid information
-schema~\cite{is} as well as an early minimal GLUE 2 implementation.
+schema~\cite{is} as well as GLUE2.
 The output is rendered in XML format.
 
-Currently, the the following GLUE2 entities are included in ARC1:
-\begin{itemize}
-\item ComputingService
-\item ComputingEndpoint
-\item ComputingShare
-\end{itemize}
-
-While the following entities are still missing:
-\begin{itemize}
-\item ApplicationEnvironment
-\item ExecutionEnvironment
-\item ComputingActivity
-\item Policy including MappingPolicy / AccessPolicy
-\end{itemize}
-
 The different entities with their attributes are described in their respective sections below.
@@
 \hline
-Type & \texttt{org.nordugrid.arex} \\
+Type & \texttt{org.nordugrid.execution.arex} \\
 \hline
@@
 \subsection{ExecutionEnvironment}
-Not yet implemented.
 
 \subsection{ApplicationEnvironment}
-Not yet implemented.
 
 \subsection{ComputingActivity}
-Not yet implemented.
 
 \subsection{MappingPolicy and AccessPolicy}
-Not yet implemented.
 

### 2010-12-18 doc/tech_doc/a-rex/arex_tech_doc.tex
Adding VOMS elements into A-REX configuration example and VOMS-based authorization policy example

@@
   xmlns:idmap="http://www.nordugrid.org/schemas/identitymap/2009/10"
+  xmlns:pdp="http://www.nordugrid.org/schemas/arcpdp/2009/08"
 >
@@
       <tls:CACertificatesDir>/etc/grid-security/certificates</tls:CACertificatesDir>
-      <!-- Evaluate requestor's identity into local identity.
-      Comment it if no user mapping is needed. -->
-      <loader:SecHandler name="identity.map" id="map" event="incoming">
-        <!-- Safe choice if all other rules failed -->
-        <idmap:PDP name="allow.pdp"><idmap:LocalName>nobody</idmap:LocalName></idmap:PDP>
-      </loader:SecHandler>
+      <!-- DNs of acceptable VOMS ACs -->
+      <tls:VOMSCertTrustDNChain>
+        <!-- Accept VOMS attributes from all sources -->
+        <tls:VOMSCertTrustRegex>^.*$</tls:VOMSCertTrustRegex>
+      </tls:VOMSCertTrustDNChain>
     </loader:Component>
@@
     <loader:Service name="a-rex" id="a-rex">
+      <!-- Performing authorization. Also see security documentation. -->
+      <loader:SecHandler name="arc.authz" event="incoming">
+        <authz:PDP name="arc.pdp">
+          <pdp:Filter>
+            <!-- Using only attributes comming from TLS (VOMS) and
+                AREX (requested operations) -->
+            <pdp:Select>TLS</pdp:Select>
+            <pdp:Select>AREX</pdp:Select>
+          </pdp:Filter>
+          <pdp:PolicyStore>
+            <!-- Type is currently not used. Only file is supported -->
+            <pdp:Type>file</pdp:Type>
+            <!-- Read and process policy from file. See policy 
+                example in appendix. -->
+            <pdp:Location>/opt/arc/etc/arex_policy.xml</pdp:Location>
+          </pdp:PolicyStore>
+        </authz:PDP>
+      </loader:SecHandler>
+      <!-- Evaluate requestor's Grid identity into local identity. -->
+      <loader:SecHandler name="identity.map" id="map" event="incoming">
+        <!-- Old bad mapfile -->
+        <idmap:PDP name="allow.pdp">
+          <idmap:LocalList>/etc/grid-security/grid-mapfile</idmap:LocalList>
+        </idmap:PDP>
+        <!-- Safe choice if all other rules failed -->
+        <idmap:PDP name="allow.pdp">
+          <idmap:LocalName>nobody</idmap:LocalName>
+        </idmap:PDP>
+      </loader:SecHandler>
       <!-- Optional endpoint element is advised in case of multiple IP adresses -->
@@
 
+\section{Example of authorization policy for A-REX\label{annex:policy-example}}
+
+\begin{footnotesize}\begin{verbatim}
+<?xml version="1.0" encoding="UTF-8"?>
+<Policy xmlns="http://www.nordugrid.org/schemas/policy-arc"
+        PolicyId="policy:arex:example"
+        CombiningAlg="Deny-Overrides">
+   <Rule RuleId="knowarc.eu:serviceinfo" Effect="Permit">
+      <Description>
+         All who belongs to knowarc.eu VO can query information about service.
+      </Description>
+      <Subjects>
+         <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute">/VO=knowarc.eu</Subject>
+      </Subjects>
+      <Actions>
+         <!-- AttributeId should be .../operation. But because of bug in A-REX this one is used. -->
+         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/joboperation">Info</Action>
+      </Actions>
+   </Rule>
+   <Rule RuleId="knowarc.eu:admin" Effect="Permit">
+      <Description>
+         All developers belonging to knowarc.eu VO can manage service.
+         This is useless rule because A-REX does not suport remote management.
+      </Description>
+      <Subjects>
+         <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute">/VO=knowarc.eu/Group=coredev</Subject>
+      </Subjects>
+      <Actions>
+         <!-- AttributeId should be .../operation. But because of bug in A-REX this one is used. -->
+         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/joboperation">Admin</Action>
+      </Actions>
+   </Rule>
+   <Rule RuleId="knowarc.eu:job" Effect="Permit">
+      <Description>
+         All testers of knowarc.eu VO can run jobs.
+      </Description>
+      <Subjects>
+         <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute">/VO=knowarc.eu/Group=testers</Subject>
+      </Subjects>
+      <Actions>
+         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/joboperation">Create</Action>
+         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/joboperation">Modify</Action>
+         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/joboperation">Read</Action>
+      </Actions>
+   </Rule>
+   <Rule RuleId="knowarc.eu:all" Effect="Permit">
+      <Description>
+         All superusers of knowarc.eu VO can do anything.
+      </Description>
+      <Subjects>
+         <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/vomsattribute">/VO=knowarc.eu/Group=superusers</Subject>
+      </Subjects>
+   </Rule>
+</Policy>
+\end{verbatim}
+\end{footnotesize}
+
 \section{Error messages of A-REX (outdated)\label{annex:arex-errors}}

### 2010-12-20 doc/tech_doc/a-rex/arex_tech_doc.tex
Fixing wrong attribute id in security attributes o a-rex

@@
       <Actions>
-         <!-- AttributeId should be .../operation. But because of bug in A-REX this one is used. -->
-         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/joboperation">Info</Action>
+         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/operation">Info</Action>
       </Actions>
@@
       <Actions>
-         <!-- AttributeId should be .../operation. But because of bug in A-REX this one is used. -->
-         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/joboperation">Admin</Action>
+         <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/a-rex/operation">Admin</Action>
       </Actions>

### 2010-12-22 doc/manuals/user_interface/ui-job.tex
Adding short description of GRID_GLOBAL_JOBID to UI manual. Fixing dependencies in Makefile.

@@
 \begin{framed}
-%% This depends on the middleware flavour. Some flavours do not accept listing
-%% the job session directory (CREAM).
-   Every job ID is a valid URL for the job session directory. You can
+   Usually job ID is a valid URL for the job session directory. You can
    always use it to access the files related to the job, by using data
    management tools (see Chapter~\ref{sec:dm}).
+
+   There may be exception for some computational services like CREAM 
+   which do not support listing job session directory.
+\end{framed}
+
+\begin{framed}
+   For advanced computational jobs which may need to communicate their 
+   presence to some service outside there may be need for knowing own
+   job ID. For jobs accepted by ARC computationsl services this information
+   may be found in local (for job executable) environment variable 
+   GRID\_GLOBAL\_JOBID. One need to take into account that this ID may be 
+   different from one provided by arcsub. Example is ID provided by
+   A-REX computational service. That service provides OGSA BES 
+   compatible interface for job management and ID contains XML document
+   according to OGSA BES specifications.
 \end{framed}

### 2010-12-28 doc/manuals/monitor/monitor.tex
Copying monitor documentation to arc1

\documentclass{article}
%
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{graphics}
%
\usepackage{graphicx}                              %for PNG images
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{makeidx}                               %for index page
\usepackage{geometry}                              %for defining page size
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref} %for \url tag
\usepackage{doc}                                   %for pfill used by gist index style
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\makeindex
\hypersetup{
  pdfauthor = {Oxana Smirnova},
  pdftitle = {The Grid Monitor},
  pdfsubject = {Usage Manual},
  pdfkeywords = {Grid,NorduGrid,monitor,user,guide,reference,manual},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
  \def\today{\number\day/\number\month/\number\year}

  \begin{titlepage}

    \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
    \end{tabular}

    \hrulefill

    {\raggedleft NORDUGRID-MANUAL-5\par}

    {\raggedleft \today\par}

    \vspace*{2cm}

%%%%---- The title ----
    {\centering \textsc{\Large The Grid Monitor}\Large \par}

    \vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
    {\centering \textit{\large Usage manual}\large \par}

    \vspace*{2cm}

%%%%---- A list of authors ----
    {\centering \large Oxana Smirnova\footnote{O.Smirnova@cern.ch} \large \par}

    \vspace*{1.5cm}

%%%%---- An abstract ----
\begin{abstract}
  The Grid Monitor is a Web client tool for the ARC Information
  System, allowing to browse all the published information about the
  system. It makes use of the hierarchical information organization
  and the PHP LDAP module to provide a real-time monitoring and
  primary debugging for ARC-based grids.
\end{abstract}

\end{titlepage}
\thispagestyle{empty} $ $
\newpage
$\ $

\section{Introduction}
\label{sec:intro}

Information services play a very important role in any computational
grid architecture, being a nerve system of the Grid. Resource
discovery, scheduling, monitoring and many other tasks are impossible
without a reliable and up-to-date information about system components.

The MDS-inspired~\cite{mds} ARC Information
System~\cite{is} provides a robust and dynamic model for accessing not
only quasi-static information about resources and services, but also
about such rapidly changing parameters like queue and job
status. Being based on OpenLDAP~\cite{ldap}, it can be easily
interfaced to any browsing or monitoring tool, giving thus a
user-friendly overview of all the testbed resources.

The Grid Monitor makes use of the LDAP module of PHP~\cite{php} to
provide a Web client tool to browse the ARC information infrastructure. It
is available in many human languages\index{language} in order to follow browser
localization settings\footnote{In order to change the Monitor's language, simply
change the preferred language of your browser}. This
document gives a summary of its capabilities and usage guidelines.

\section{Grid Monitor Modules}
\label{sec:modules}

The structure of the Grid Monitor to great extent follows that of the
ARC Information System~\cite{is}. The basic objects are defined
by the following schema's objectclasses\index{objectclass}:
\begin{list}{--}{\itemsep=-0.5mm}
\item \textsf{nordugrid-cluster}: a cluster \index{nordugrid-cluster}
\item \textsf{nordugrid-queue}: a queue at the cluster, accessible by
  the authorised users \index{nordugrid-queue}
\item \textsf{nordugrid-job}: a Grid job, associated with a queue
  \index{nordugrid-job}
\item \textsf{nordugrid-authuser}: a user, authorized to submit jobs
  to a given queue \index{nordugrid-authuser}
\end{list}

The Grid Monitor also uses the Virtual Organisation (VO)
\textsf{organisationalPerson} and Storage Element
\textsf{nordugrid-se} objectclasses, and their attributes.

For each objectclass, either an essential subset of attributes, or the
whole list of them, is presented in an easily accessible inter-linked
manner. This is realized as a set of browser windows, each being associated
with a corresponding module. There are nine major modules
\index{modules}: \newcounter{count1}
\begin{list}{\arabic{count1})}{\usecounter{count1} \itemsep=-0.5mm}
\item An overall Grid Monitor
\item Cluster Description
\item Queue Details
\item Job Information
\item User Information
\item Attributes Overview
\item Customizeable Display Tool (``Match-it-yourself'')
\item List of Storage Facilities
\item List of Users
\end{list}
Each module displays both dynamic and static information: for example,
a queue name is static, while the amount of running jobs in this queue
is dynamic. Most of the displayed objects are linked to appropriate
modules, such that with a simple mouse click, a user can launch
another module, expanding the information about the corresponding
object or attribute. Each such module opens in an own window, and
gives access to other modules in turn, providing thus a rather
intuitive browsing.

In what follows, these modules are described in details, giving an
overview of their functionality and usage hints.

\subsection{The Grid Monitor}
\label{sec:loadmon}

The basic module, \index{grid monitor} providing access to the most
required information, is the Grid Monitor, showing the overall status
of the system. It serves as a starting point for browsing the system
information. The purpose of this module is to give a quick overview of
the current status of the Grid infrastructure by showing the list of the
available clusters and the most essential information about them: an
alias, number of working processors, number of occupied processors and
number of queueing jobs. In the current implementation, the main Grid
Monitor window contains also the link to the user base of the infrastructure. 
Figure~\ref{fig:loadmon} shows a screenshot of the running
monitor. All the information shown is dynamic, including
organizational names (countries in this case).

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{loadmon.png}}}
\caption{\label{fig:loadmon}The Grid Monitor} }
\end{figure}

In Figure~\ref{fig:loadmon}, the numbered tags indicate clickable
objects as explained below:
\newcounter{count2}
\begin{list}{\arabic{count2})}{\usecounter{count2} \itemsep=-0.5mm}
\item \textsf{Cluster}: a cluster alias, \index{cluster>alias} linked
  to the cluster description module (Section~\ref{sec:clusdes}), which
  provides complete information about the current status of a cluster.
\item \textsf{Load}: a graphical and numeric representation of the
  cluster load, \index{cluster>load} showing both Grid- and non-Grid
  (submitted locally) running processes. Colored bar shows percentage of
  Grid processes, while the grey bar shows total relative
  occupancy of a cluster. Numbers indicate the absolute amount of
  running processes, with first figure corresponding to the Grid, and
  second - to the non-Grid ones. It should be noted that number of
  processes does not necessarily correspond to the number of running
  jobs: a parallel job \index{job>parallel} can occupy several
  processors. By clicking on a bar, a user accesses the list of all
  Grid jobs, \index{job>running} running on a cluster
  (Section~\ref{sec:jobstat}).
\item \textsf{Queueing}: number of queueing jobs, \index{job>queueing}
  which includes both jobs queued in an LRMS and those being
  pre-processed by the Grid Manager~\cite{gm}. Only jobs which can be
  potentially executed in a Grid queue are counted. The number is
  linked to the same module as the \textsf{Load} item, with the only
  difference that it displays the list of the Grid-queued jobs. Note
  that non-Grid jobs are counted in the total number of queued jobs,
  while they can not be listed by the Grid Monitor, as they are not
  providing any information in the ARC Information System.
\item \textsf{Match-it-yourself}: link to the ``Match-it-yourself''
  interface (Section~\ref{sec:match}) which allows users to compose
  non-standard monitor requests. \index{Match-it-yourself}
\item \textsf{Storage resources}: link to the list of available
  storage resources (Section~\ref{sec:storage}). \index{storage}
\item \textsf{User base}: several auxiliary links, providing an access
  to the VO-listing module (Section~\ref{sec:vo-users}). The main
  purpose of this link is to provide an easy access to the
  user-specific information, such as the list of submitted jobs and
  available resources.
\end{list}

\subsection{Cluster Description}
\label{sec:clusdes}

The cluster description \index{cluster} module displays all the
cluster attributes stored in the local information tree, as well as most
relevant
information about the queues, accessible by the Grid users. The
window thus contains two lists, as shown in Figure~\ref{fig:clusdes}:

\newcounter{count3}
\begin{list}{\arabic{count3})}{\usecounter{count3} \itemsep=-0.5mm}
\item \textsf{Attributes}: \index{cluster>attributes} this is a dump
  of all the attributes of the \textsf{nordugrid-cluster} objectclass,
  dynamic and static ones. Such attributes as cluster alias, or domain
  name, are static; others are dynamic, with the values obtained by
  the information providers: e.g., total CPU number,
  amount of jobs, or available disk space. More details about these
  attributes can be found in the ARC Information System
  description~\cite{is}. Each attribute (apart of the time stamps)
  is linked to the Attributes Overview module
  (Section~\ref{sec:attlist}), such that clicking on an attribute name
  brings the list of the values of this particular attribute on all
  the Grid clusters. For instance, this is the most convenient
  way to browse available disk space \index{disk space} or runtime
  environment\index{runtime environment} values over the system.
\item \textsf{Queues}: \index{queue>list} the list of queues at a
  given cluster, accessible by the Grid users. While the detailed
  list of queue attributes and corresponding jobs can be obtained by
  clicking on a queue name (see Queue Details module description,
  Section~\ref{sec:quelist}), the most essential parameters are listed
  already in the Cluster Description module. They are: queue name,
  \index{queue>name} queue status, queue length \index{queue>length}
  (minimal and maximal), number of CPUs assigned to a queue (if
  available), and number of running and queued jobs. Since queues can
  be shared between Grid and local users, the total number of jobs is
  shown, with the number of Grid jobs in parentheses.
\end{list}

The Cluster Description module is linked from most other modules (except
the List of Users one): clicking on a domain name of a cluster brings
the Cluster Description window.

\begin{figure}[ht]
\centering{
{\scalebox{0.6}{\includegraphics{clusdes.jpg}}}
\caption{\label{fig:clusdes}Grid cluster details} }
\end{figure}

\subsection{Queue Details}
\label{sec:quelist}

In the ARC Information System, \index{queue} the
\textsf{nordugrid-queue} objectclass is described by a set of
queue-specific attributes, and has two sub-trees: \textsf{nordugrid-job}
and \index{job} \textsf{nordugrid-authuser}. This structure reflects the
fact that users are not implicitly authorized to submit jobs to any
queue. However, the list of users allowed to a specific queue is a
fairly static information, and thus is beyond the scope of the Grid
Monitor\footnote{List of queues available for a given user can be
obtained through the User Information module}.

\begin{figure}[ht]
\centering{
{\scalebox{0.32}{\includegraphics{quelist.png}}} }
\caption{\label{fig:quelist}Grid queue details}
\end{figure}

The Queue Details module provides the list of the queue attributes and
of all the jobs scheduled (running or waiting) to this queue.
Figure~\ref{fig:quelist} shows the queue description window,
with clickable fields marked by numbered tags as follows:

\newcounter{count4}
\begin{list}{\arabic{count4})}{\usecounter{count4} \itemsep=-0.5mm}
\item \textsf{Attributes}: \index{queue>attributes} the dump of the
  queue attributes. Just like the cluster attributes
  (Section~\ref{sec:clusdes}), they can be both static and
  dynamic. Every attribute is linked to the Attributes Overview module
  (Section~\ref{sec:attlist}), which allows to browse the values of
  each attribute over all the Grid system.
\item \textsf{Cluster name}: each queue is associated with the
  cluster, which name is shown at the top of the window. Clicking the
  cluster name brings up the Cluster Description window
  (Section~\ref{sec:clusdes}).
\item \textsf{Job name}: from the Queue Details window, users can get
  access to detailed information about every job \index{job>in a
  queue} in the queue by clicking the job name.\index{job>name} Each
  job name is linked to the Job Information module, described in
  Section~\ref{sec:jobstat}.
\item \textsf{Owner}: \index{job>owner} The Grid authentication
  mechanism allows to associate every job with a corresponding user,
  even though an actual Unix account owner may be a generic
  "griduser". The Grid Monitor uses this feature to display explicitly
  each job owner. In the Queue Details window (as in all other
  modules), user's name is linked to the User Information module
  (Section~\ref{sec:userlist}), which displays all the resources
  available for a given user, as well as the list of user's jobs.
\end{list}

Queue Information module is accessible via links to queue names in the
Cluster Information (Section~\ref{sec:clusdes}), Job Information
(Section~\ref{sec:jobstat}), User Information
(Section~\ref{sec:userlist}) and Attributes Overview
(Sec~\ref{sec:attlist}) modules.

\subsection{Job Information}
\label{sec:jobstat}

The Job Information module is activated on three different occasions:
\begin{list}{--}{\itemsep=-0.5mm}
\item To display a list of all running Grid jobs
  \index{job>running} on a cluster
\item To display a list of all queued \index{job>queued} Grid
  jobs on a cluster
\item To show the full information on a given job
\end{list}

Lists of running and queued jobs \index{job>{on cluster}} are
accessible from the top Grid Monitor window
(Section~\ref{sec:loadmon}) by clicking the corresponding fields
(marked 2 and 3 in Figure~\ref{fig:loadmon}).  As shown in
Figure~\ref{fig:jobstat2}, such a list contains not only job names,
but also their respective owners, status (as returned by the Grid
Manager), execution time (in case of running jobs), and the submission
queue.

\begin{figure}[hb]
\centering{
{\scalebox{0.4}{\includegraphics{jobstat2.png}}} }
\caption{\label{fig:jobstat2}Grid job list}
\end{figure}

Most of the fields in a job list window are linked to the corresponding
monitor modules, giving access to more detailed information:
\newcounter{count5}
\begin{list}{\arabic{count5})}{\usecounter{count5} \itemsep=-0.5mm}
\item \textsf{Job name}: just like in the Queue Details window
  (Section~\ref{sec:quelist}), the job name is linked to the Job
  Information window, described below. However, while the Queue
  Details module lists the jobs in a given queue, the Job Information
  window gives an overview of all the Grid jobs on a cluster.
\item \textsf{Owner}: this field is also identical to the one in the
  Queue Details window: user's name is linked to the User Information
  module (Section~\ref{sec:userlist}), which displays all the
  resources available for a given user and the list of user's jobs.
\item \textsf{Queue}: the name of the queue is liked to the Queue
  Details window (Section~\ref{sec:quelist}), which gives a snapshot
  of the queue status, including al the Grid jobs submitted to a
  particular queue -- running or waiting.
\item \textsf{Cluster name}: clicking on the cluster name brings up
  the Cluster Description window (Section~\ref{sec:clusdes}), which
  gives a general overview of a given cluster and the status of its
  queues (those available for the Grid users).
\end{list}

\begin{figure}
\centering{
{\scalebox{0.6}{\includegraphics{jobstat1.jpg}}} }
\caption{\label{fig:jobstat1}Grid job statistics}
\end{figure}

The job \index{job>information} information window is invoked by
clicking on a job name in any Grid Monitor window which lists jobs. It
is handled by the same module which produces running/queued job list,
and contains simple dump of all the available job attributes (see
Figure~\ref{fig:jobstat1}). Just like in the Cluster Description and
Queue Description windows, each attribute is clickable (as indicated
by a tag numbered 1 in Figure~\ref{fig:jobstat1}), and is linked to
the Attributes Overview module (Section~\ref{sec:attlist}). This is a
convenient way to compare jobs that reside on the system.

\subsection{User Information}
\label{sec:userlist}

The User Information \index{user information} module of the Grid
Monitor gives access to all the available information, related to a
given user. This includes the list of available resources
\index{resources} (queues, processors and disk space), and the list of
user jobs, residing on the system at the time of query. To collect
this information, the whole system has to be queried, therefore
invocation of this module typically takes quite a bit of time (at
least comparing to most other modules).

\begin{figure}
\centering{
{\scalebox{0.4}{\includegraphics{userlist.png}}} }
\caption{\label{fig:userlist}Grid user information}
\end{figure}

Figure~\ref{fig:userlist} shows a typical User Information window, where
the numbered fields are linked to other Grid Monitor modules:
\newcounter{count6}
\begin{list}{\arabic{count6})}{\usecounter{count6} \itemsep=-0.5mm}
\item \textsf{Job name}: \index{job>by user} this field is linked to
  the Job Information window (Section~\ref{sec:jobstat}), providing
  access to the detailed information on a given job. Unlike of Job
  Information or Queue Information modules, which list local to a
  cluster jobs, the User Information module collects all the jobs
  submitted by a given user to the whole system.
\item \textsf{Cluster}: since the User Information window displays all
  the jobs associated with a given user, description of each
  respective cluster is available by clicking the cluster name. This
  brings up a cluster description window, described in
  Section~\ref{sec:clusdes}.
\item \textsf{Queue}: this field is linked to the Queue Details module
  (Section~\ref{sec:quelist}), thus giving access to the information
  about the status of the relevant queue.
\item \textsf{Cluster}: the upper part of the User Information window
  lists the Grid resources, available for a user. Each cluster,
  to which a user is authorized to submit jobs, is indicated by its
  name.  Cluster names are linked to the Cluster Description window
  (Section~\ref{sec:clusdes}), giving detailed information on
  available resources.
\item \textsf{Queue}: since users authorization may be not only
  cluster-based, but also queue-based, the allowed queue information
  can be accessed by clicking a queue name. This brings up the Queue
  Details window, described in Section~\ref{sec:quelist}.
\end{list}

The simplest way to access the User Information window is via the List
of Users (Section~\ref{sec:vo-users}), although it can be invoked from
any Grid Monitor window where a user name is displayed (e.g., a Job
Information or a Queue Details window).

\subsection{Attributes Overview}
\label{sec:attlist}

As it was mentioned above, every ARC objectclass attribute,
\index{attributes} appearing in a Grid Monitor window, is linked to
the Attributes Overview module, which queries all the relevant objects
on the system and delivers a comparative list of the
attributes. Similarly to the User Information module, querying all the
Grid resources takes somewhat long time, as the Grid Monitor does
not have an own cache.

This module can also be accessed via the ``Match-it-yourself''
interface (Section~\ref{sec:match}). In this case, it can list as many
attributes as specified by a user request, eventualy applying the user
selection criteria.

\begin{figure}[hb]
  \centering{
    {\scalebox{0.6}{\includegraphics{attlist.png}}} }
  \caption{\label{fig:attlist}Grid objects grouped by attribute}
\end{figure}

Figure~\ref{fig:attlist} shows a typical result of the Attributes
Overview query: in this example, the \textsf{nordugrid-cluster}
attribute "Jobs, total amount" was queried, and a comparative list of
results returned. The \textsf{Resource} field (indicated by the tag 1)
depends on the nature of the attribute, and can be either of:
\begin{list}{--}{\itemsep=-0.5mm}
\item cluster name, linked to the Cluster Description module,
\item cluster name and queue name, linked to the Cluster Description
  and Queue Details modules respectively,
\item job ID string (see ref.\cite{gm} for details), linked to the Job
  Information module.
\end{list}

\subsection{``Match-it-yourself''}
\label{sec:match}

The ``Match-it-yourself'' \index{Match it yourself} is a customizeable
interface to the Attributes Overview module
(Section~\ref{sec:attlist}). It allows users to chose which attributes
of an object to display, optionally applying filters. While the other
Monitor windows display a pre-defined set of data, this module gives
an advanced user a possibility to build a customized request to the
Information System.

An example use case for this interface could be a user desiring to
view a list of his running (but not queued or finished) jobs, complete
with used CPU and wall time, memory and job name. The
``Match-it-yourself'' tool would be then invoked for the \textsf{job}
object, and the display request would contain \textsf{Name},
\textsf{Used CPU time}, \textsf{Used wall time}, \textsf{Used memory
  (KB)}, and \textsf{Status} -- the latter with a filter
\fbox{\textsf{Status = INLRMS: R}}.

\begin{figure}[ht]
  \centering{
    {\scalebox{0.6}{\includegraphics{match1.png}}} }
  \caption{\label{fig:match1}Object class selection window}
\end{figure}

Figure~\ref{fig:match1} shows the first screen of the
``Match-it-yourself'' interface, which welcomes users to select the
object class to work with, and the amount of attributes to be
displayed. When not sure about the latter, users should specify a top
estimate -- unused fields will be ignored in further searches.

\begin{figure}[hb]
  \centering{
    {\scalebox{0.6}{\includegraphics{match2.png}}} }
  \caption{\label{fig:match2}Attribute selection window}
\end{figure}

Figure~\ref{fig:match2} is a snapshot of the screen where the
atttributes to display and their selection criteria are specified. If
a user wishes to display an attribute for all the objects,
independently of its value, the rightmost field may be either kept
empty, or filled with an asterisk (\verb#*#), while the middle field
should be set to ``$=$''. Whenever a filter has to be applied, an
operator should be selected in the middle column, and a match string
specified in the rightmost field. For example, if only clusters
containing ``NBI'' in their domain names have to be shown, the
attribute filter would be \fbox{\textsf{Front-end domain name $\sim$
nbi}}. Matches are case-insensitive.

 \begin{figure}[ht]
   \centering{
     {\scalebox{0.6}{\includegraphics{match3.png}}} }
   \caption{\label{fig:match3}Customized cluster information display}
 \end{figure}

Figure~\ref{fig:match3} is the result of the search according to the
criteria defined in the exampel in Figure~\ref{fig:match2}. Three
filters were applied: on operating system attribute, total number of
CPUs and postal code (in this case we were selecting any cluster which
is not in Sweden). Since we wanted to display each cluster's alias as
well, this attribute was added to the selection, but with a ``match
everything'' scope. The attribute matching method is exactly the same
as used by the Attributes Overview module (Section~\ref{sec:attlist}),
and it re-uses the screen layout shown in Figure~\ref{fig:attlist}.

\subsection{Storage Resources}
\label{sec:storage}

\index{Storage resources}Although there is no well-defined Storage
Element\index{Storage Element} concept in ARC, some
information about the storage resources can be found in the
Information System. The Storage Resources module, linked from the main
Monitor window, displays all the available information for those
Storage Elements which publish it. Particularly important is the base
URL, which specifies the Grid mount point that could be used in job
descriptions.

\begin{figure}[hb]
  \centering{
    {\scalebox{0.7}{\includegraphics{storage.png}}} }
  \caption{\label{fig:storage}List of storage elements}
\end{figure}

\subsection{List of Users}
\label{sec:vo-users}

The List of Users \index{user list} module is different from the rest
of the Grid Monitor modules because it does not deal with the
ARC information system. Instead, it retrieves lists of users from
VO databases~\cite{vo}. It serves as a link between different databases
(LDAP and VO), by interfacing each user record to the User Information
module (Section~\ref{sec:userlist}). Figure~\ref{fig:vo-users} shows a
screenshot of a typical VO user list, with numbered tags indicating
clickable links as follows: \newcounter{count7}
\begin{list}{\arabic{count7})}{\usecounter{count7} \itemsep=-0.5mm}
\item \textsf{Name}: user name as given in the corresponding Grid
  certificate field, linked to the User Information module
  (Section~\ref{sec:userlist}).
\item \textsf{E-mail}: E-mail \index{E-mail} address of a user, if
  available. It is linked to an e-mail URL, allowing to send a message
  to a user directly from the browser (if such an option is enabled in
  a browser).
\end{list}

\begin{figure}[ht]
  \centering{
    {\scalebox{0.6}{\includegraphics{vo-users.jpg}}} }
  \caption{\label{fig:vo-users}List of the Grid users}
\end{figure}

The List of Users is available only from the top Grid Monitor window.

\section{Implementation notes}
\label{sec:implement}

The Grid Monitor is implemented entirely in PHP, with optional usage
of client-side JavaScript. Since all the databases the Grid Monitor
has to deal with are hierarchical LDAP ones, the server-side LDAP
module of PHP is absolutely necessary to be enabled in order to make
the Grid Monitor functioning. The PHP LDAP module conveniently allows
parallel LDAP searches, -- the feature heavily used by the Grid
Monitor, since it speeds up the data retrieval.

The Grid Monitor uses only minimal disk caching (overview window only), storing
all the LDAP
query results in the memory. To minimize the memory usage, only the
attributes relevant to each query are retrieved.

In order to speed up the queries, the Grid Monitor makes as little use
of the original MDS information propagation mechanisms as possible. GIIS and
GRIS services are used only as link collections\footnote{Globus MDS2.2
provided a quick access to the lower level servers via a base scope
LDAP search for the "giisregistrationstatus" attribute}.

Since the ARC architecture makes use of several equivalent
top-level GIIS servers, the Grid Monitor queries all of them in order
to have a reliable access to all the system information. In some
cases, lower-level GRIS servers can also be duplicated, hence the Grid
Monitor contains a built-in mechanism to prevent
double-counting.

Discovery of lower-level GRIS servers is done recursively, starting
from the registration information in all the top-level indexes, and
ending at the local level. This recursive search method is invoked not
only to discover clusters in e.g., the main Monitor module, but also
to locate storage facilities.

All the Grid Monitor windows are automatically refreshed by the means
of the built-in browser HTML instructions. Every window can be
forcefully refreshed, printed and closed by using either standard
browser tools, or the provided JavaScript-enabled buttons.

In the top Grid Monitor window, clusters are automatically grouped by
respective second-level hierarchy VOs -- in the described above
case, this is nothing but countries.

Such fields as cluster aliases, user names and attribute names, are
customizable, and can be adjusted from the stored in the information system
values to
any more appropriate ones, depending on the actual requirements.

In general, the Grid Monitor was designed to be a cross-browser,
cross-platform tool, and have been shown to work properly with browsers
\index{browsers} ranging from Lynx to Konqueror to Microsoft Internet
Explorer.

%% \section{HOWTO}
%% \label{sec:howto}

%% This is a list of most common examples of the Grid Monitor usage.

%% \begin{description}
%% \item[How to list all the jobs submitted by a user?]$\ $\\ Click a
%%   user name in any window, e.g., in the List of Users
%%   (Figure~\ref{fig:vo-users}, item 1). The List of Users is available
%%   from the top Grid Monitor window (Figure~\ref{fig:loadmon}, item 4).
%% \item[Why the jobs finished two days ago do not show up?]$\ $\\
%%   Job results are being kept in the NorduGrid session directory
%%   only for a limited period, typically 24 hours. After a session
%%   directory is erased, -- either by a user request or after its
%%   lifetime \index{lifetime} expiration -- all the job information
%%   disappears from the system.
%% \item[How to list all jobs running on a cluster?]$\ $\\ Click a
%%   ``\textsf{Load}'' bar in the top Grid Monitor window
%%   (Figure~\ref{fig:loadmon}, item 2).
%% \item[How to list all jobs in a queue?]$\ $\\ Click a queue name in
%%   any window, e.g., the cluster information (Figure~\ref{fig:clusdes},
%%   item 2), the job information (Figure~\ref{fig:jobstat2}, item 3), or
%%   the user information (Figure~\ref{fig:userlist}, item 5).
%% \item[How to list available runtime environments?]$\ $\\ Click any
%%   ``\textsf{Cluster}'' field in the top Grid Monitor window
%%   (Figure~\ref{fig:loadmon}, item 1) to bring up the Cluster Details
%%   window. Then click "Runtime environment" link in the
%%   ``\textsf{Attribute}'' column (Figure~\ref{fig:clusdes}, item 1).
%% \end{description}

\bibliography{grid,nordugrid}
\printindex

\end{document}

### 2010-12-28 doc/tech_doc/ldap-arc-infosys/arc_infosys.tex
Copying LDAP-based ARC infosys documentation to arc1

\documentclass{article}
%
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{graphics}
%
\usepackage{graphicx}                              %for PNG images
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
%\usepackage{makeidx}                               %for index page
\usepackage{geometry}                              %for defining page size
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref} %for \url tag
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\makeindex
\hypersetup{
  pdfauthor = {Balazs Konya},
  pdftitle = {ARC Information System},
  pdfsubject = {Technical description},
  pdfkeywords = {Grid,NorduGrid,information systems},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
%\bibliographystyle{IEEEtran}
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\gt}{Globus Toolkit$\mathrm{^{TM}}$}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
  \def\today{\number\day/\number\month/\number\year}

  \begin{titlepage}

    \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
    \end{tabular}

    \hrulefill

    {\raggedleft NORDUGRID-TECH-4\par}

    {\raggedleft \today\par}

    \vspace*{2cm}

%%%%---- The title ----
    {\centering \textsc{\Large The NorduGrid/ARC Information System}\Large \par}

    \vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
    {\centering \textit{\large Technical Description and Reference Manual}\large \par}

    \vspace*{2cm}

%%%%---- A list of authors ----
    {\centering \large Bal\'azs K\'onya\footnote{Comments to:
     balazs.konya@hep.lu.se} \large \par}

    \vspace*{3.5cm}

    {\centering \large %Refers to ARC release series 0.4 and up 
    \it Version 0.9
    \large \par}

%%%%---- An abstract ----
%\begin{abstract}

%\end{abstract}

\end{titlepage}
\thispagestyle{empty} $ $
\newpage
$\ $

%%%%---- Table of Contents, uncomment if necessary
%\tableofcontents{}

%TODO make clear what is available in which arc-version, mention 0.5 
% as a pre-0.6

\section{Introduction}

A stable, robust, scalable, dynamic and reliable information system is
a cornerstone of any kind of Grid system. Without a properly working
information system it is not possible to construct a production
quality Grid. A scalable Grid information system is inherently
distributed, a centralized system is not able to cope with the
dynamism of the Grid.

The information system acts as a nervous system of the Grid and its
main tasks consist of
\begin{itemize}
\item Resource Description:
Characterization of Grid resources by specifying static, semi-static
and dynamic properties (e.g. information about Grid jobs and user
quotas are presented as dynamic local resource properties).

Grid clients are relying on the resource description functionality of
the information system during their matchmaking and brokering
process. Grid monitoring and job status queries also rely on resource
description.

\item Resource aggregation:
Individual resources are connected to an "information mesh" by
dynamically registering to some of the information index services. The
information index services are responsible for the resource
aggregation, they maintain a dynamic list of available Grid resources.
Furthermore, the index services are connected to each other following
a specific topological order. The resulting structure is the
"information mesh" of the Grid.

Grid clients performing resource discovery scan the '"information
mesh" utilizing its topological structure in order to find available
Grid resources.  Therefore resource discovery is delicately coupled to
the topological structure of the "information mesh" that is to the
resource aggregation process.
\end{itemize}

The ARC middleware implements a scalable, production quality
dynamic distributed information system. The ARC information system 
has been deployed and being used in a large scale production environment 
since May 2002. As of writing, the information system aggregates ~70 
resources providing \gt 40 thousand CPUs and serves \gt 400 thousand jobs per month.

The ARC information system is an OpenLDAP-based system\cite{Openldap}
which is derived from the Globus Monitoring and Discovery Services
framework \cite{mds}. It uses BDII\cite{bdii} together with a updated MDS LDAP Schema.

This document presents a technical overview of the ARC Grid information system. 
The document describes the architecture, the implementation of the main components 
and the NorduGrid/ARC Information model. The document is also intended to serve as  
reference manual by giving a detailed description of the available Grid information.

\section{Overview}

The ARC middleware implements a dynamic LDAP-based distributed information 
system via set of coupled resource lists (index services) and 
local LDAP databases. The system consists of three main components:
\begin{enumerate}
\item the Local Information Trees (LIT),
\item the Index Services  (IS),
\item and the Registration Processes (RP)
\end{enumerate}

The local information trees are responsible for resource (computing or storage)
description and characterization. The local information is generated on the 
resource, optionally it is cached. Upon client requests it is presented to the 
clients via LDAP interface.

The main task of the index services is to maintain a dynamic list of resources 
(LDAP URLs of the local information trees) and index services. 
The index services are further coupled together implementing a specific topology.

The local information trees make use of the registration processes
running locally on the resources in order to list themselves in some of the 
resource lists maintained by the index services. Registrations are always 
initiated by the registrants (bottom-up model).

Grid clients such as monitors, web portals or user interfaces
perform two type of queries:
\begin{enumerate}
\item During the resource discovery process
clients query index services in order to collect list of LDAP contact 
URLs of local information trees describing Grid-connected resources.

\item
During a direct resource query the clients directly contact the 
local information tree by making use of the obtained LDAP contact URLs.
\end{enumerate}

Both type of queries are carried out and served via LDAP protocol.

Figure~\ref{fig:overview} presents an overview of the ARC information system
components.

\begin{figure}[hb]
\centering
{ 
{\scalebox{0.35}{\includegraphics{infosys_overview.png}}}
}
\caption{\label{fig:overview}Overview of the ARC information system components.}
\end{figure}

\section{Local Information Tree}

The LIT component of the information system is responsible
for generating the dynamic state information, implementing the first-level caching
of the local information and providing the requested Grid information to the clients 
through the LDAP protocol. The LIT is basically nothing more but a 
specially populated and customized OpenLDAP database.

The dynamic resource state information is generated on the resource.
Small and efficient programs, called information providers, are used to 
collect local state information from the batch system, from the local Grid layer 
(e.g. Grid Manager or GridFTP server \cite{grid-manager}) or from the
local operating system (e.g. information available in the \texttt{/proc} area).
Currently, ARC is capable interfacing to the following batch systems 
(or local resource management system LRMS in the ARC terminology): 
UNIX fork, the PBS-family (OpenPBS, PBS-Pro, Torque), Condor, Sun Grid Engine, IBM LoadLeveler and SLURM.

The output of the information providers (generated in LDIF format) is
used to populate the local LDAP tree. This OpenLDAP back-end
implements two things: it is capable caching the providers output and
upon client query request it triggers the information providers unless
the data is already available in its cache.  The caching feature of
the OpenLDAP back-end provides protection against overloading the
local resource by continuously triggering the information providers.

The information stored in the LIT follows the NorduGrid-ARC
information model.  The section \ref{sec:infomodel} gives a detailed
technical account of the ARC information model.

\subsection{Security considerations}

The Local Information Tree is implemented via an LDAP database which implies the security and
confidentiality capabilities of the system.

OpenLDAP\cite{Openldap} contains two methods for specifying access control. 
The first is static, i.e. you define the rights in configuration files.
From an operational point of view, the problem of this method is that needs a 
server restart at every security configuration change.
The second method for access control, called as ACI (Access Control Information), 
inserts access control information inside the directory itself by augmenting 
every LDAP entry with a dynamically  modifiable ACL.
Unfortunately the ACI method is still considered to be experimental.

The current ARC setup makes use of the static LDAP access control, the trees are 
configured to be fully readable by anybody: ARC provides  anonymous read
access to every information stored in the local trees.

There are considerations to experiment with the ACI access control method  
or to modify the static configuration and require authentication
from the clients.

\section{The ARC information model \label{sec:infomodel}}

A Grid information model should be a result of a
delicate design process how to represent the resources and what is
the best way to structure this information.

ARC implements an LDAP-based information system. In an LDAP-based system the 
information is being stored as attribute-value pairs grouped 
together in entries which are organized into a hierarchical tree.
Therefore an LDAP-based information model is technically specified via 
an LDAP schema AND the structure of the LDAP-tree (DIT).

The ARC information model naturally describes the main Grid
components:
\begin{itemize} 
\item computing resources with Grid jobs and Grid users, 
\item storage elements,
\item and metadata catalogues
\end{itemize}
though the latter two are treated in a rather simplistic
manner.

\subsection{LDAP technicalities: namespace, OID, objectclasses, attributes}

The NorduGrid/ARC LDAP schema (available in appendix~\ref{appendix:schema})
makes use of the {\it nordugrid-} namespace, the objectclass
and attribute names starts with the {\it nordugrid-} prefix.

NorduGrid is assigned to the {\mbox 1.3.6.1.4.1.11604} Private Enterprise Number 
which is utilized according to the Table\ref{Table:OID1}.

%this is turned into a table
% 1.3.6.1.4.1.11604.1 security
% 1.3.6.1.4.1.11604.2 info
% 1.3.6.1.4.1.11604.3 data management
% 1.3.6.1.4.1.11604.4 user management

\providecommand{\tabularnewline}{\\}

\begin{table}
\centering
\begin{tabular}{|l|c|}
\hline
Object Identifier&
Service area\tabularnewline
\hline
\hline
1.3.6.1.4.1.11604.1&
security\tabularnewline
\hline
1.3.6.1.4.1.11604.2&
information system\tabularnewline
\hline
1.3.6.1.4.1.11604.3&
data management\tabularnewline
\hline
1.3.6.1.4.1.11604.4&
user management\tabularnewline
\hline
\end{tabular}
\caption{\label{Table:OID1} The OID space utilization within ARC}
\end{table}

The OID's used in the LDAP schema are shown in 
Table\ref{Table:OID2} and are taken from the range {\mbox 1.3.6.1.4.1.11604.2.*}
Table\ref{Table:OID2} also serves as a list of the NorduGrid objectclasses.
The ARC implementation follows a "one LDAP entry = one objectclass" approach,
The ARC information system objects such as Grid-enabled clusters, queues, storages, 
Grid users and Grid jobs are described by specific LDAP entry which utilizes
a single objectclass. As a result a one-to-one correspondence exists between
ARC LDAP entries and ARC objectclasses.

The detailed description of the objectclasses and attributes 
are given in the following subsections. First the main purpose  
behind the objectclass is outlined followed by the one-by-one 
description of the attributes. The attribute descriptions also contain 
information about the attributes role played in the brokering\cite{arc_brokering},
the job submission or the monitoring process. If applicable, the corresponding 
xRSL attribute\cite{xrsl_manual} is displayed. 
Please notice that the most of the attribute values documented below are not enforced, 
misconfigured or rough sites can publish incorrect information.

%TODO: think of attribute default values, unset/irrelevant attributes

%this is turned into a table
%1.3.6.1.4.1.11604.2.1.1  	cluster objectclass
%1.3.6.1.4.1.11604.2.1.1.x	cluster attributes
%1.3.6.1.4.1.11604.2.1.2  	nordugrid-info-group objectclass
%1.3.6.1.4.1.11604.2.1.2.x	nordugrid-info-group attributes
%1.3.6.1.4.1.11604.2.1.3  	nordugrid-queue objectclass
%1.3.6.1.4.1.11604.2.1.3.x 	nordugrid-queue attributes
%1.3.6.1.4.1.11604.2.1.4  	nordugrid-job
%1.3.6.1.4.1.11604.2.1.4.x	nordugrid-job attributes
%1.3.6.1.4.1.11604.2.1.5  	nordugrid-authuser
%1.3.6.1.4.1.11604.2.1.5.x	nordugrid-authuser attributes
%1.3.6.1.4.1.11604.2.1.6  	nordugrid-se
%1.3.6.1.4.1.11604.2.1.6.x	nordugrid-se attributes
%1.3.6.1.4.1.11604.2.1.7  	nordugrid-rc
%1.3.6.1.4.1.11604.2.1.7.x	nordugrid-rc attributes
%1.3.6.1.4.1.11604.2.1.7  	mds
%1.3.6.1.4.1.11604.2.1.7.x	mds attributes

\begin{table}
\centering
\begin{tabular}{|l|c|}
\hline 
1.3.6.1.4.1.11604.2.1.1&
cluster objectclass\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.1.x&
cluster attributes\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.2&
info-group objectclass\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.2.x&
info-group attributes\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.3 &
queue objectclass\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.3.x&
queue attributes\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.4&
job objectclass\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.4.x&
job attributes\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.5&
authuser objectclass\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.5.x&
authuser attributes\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.6&
se objectclass\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.6.x&
se attributes\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.7 &
rc objectclass\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.7.x&
rc attributes\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.8 &
Mds objectclass\tabularnewline
\hline 
1.3.6.1.4.1.11604.2.1.8.x&
Mds attributes\tabularnewline
\hline 
\end{tabular}
\caption{\label{Table:OID2} The OIDs from the NorduGrid/ARC schema}
\end{table}

\subsection{The structure of the local LDAP tree: Arc DIT}

The ARC local LDAP tree gives a natural representation 
of a Grid-enabled resource. 
The {\it mds-vo-name=local,o=grid} top level entry of the tree
carries no information, its role is purely structural. This entry
accommodates the subtrees of the different Grid services offered by the 
local machine. A Grid resource in ARC can host a computing service 
and several storage or data indexing services. A computing service is 
described by the cluster subtree, while the storage and data indexing services 
are characterised by the {\it se} and {\it rc} single-entry subtrees, respectively.

%
%TODO add a paragraph about the useability, performance implication of a bad DIT
% design

\begin{figure}[hb]
\centering
{ {\scalebox{0.4}{\includegraphics{lit.png}}} }
\caption{\label{fig:dit} The local information tree on two resources. 
The first machine bambi.hep.lu.se provides both computing, storage and data indexing services
while the second resource hathi.hep.lu.se hosts two storage elements.}
\end{figure}

Figure~\ref{fig:dit} shows the local LDAP tree of two Grid-enabled resources.
The first machine bambi.hep.lu.se offers both a computing service,
a storage service and a data indexing service, therefore the LIT of bambi.hep.lu.se
contains a cluster subtree under the {\it nordugrid-cluster-name=bambi.hep.lu.se} entry 
a storage {\it nordugrid-se-name=..} and a data indexing {\it nordugrid-rc-name=...}
entry. The second resource hathi.hep.lu.se serves as a dedicated storage hosting two storage
elements, therefore the LIT of hathi.hep.lu.se consists of the two storage entries.

The schematic structure of the cluster subtree is shown enlarged in 
Fig.~\ref{fig:dit_subcluster}. The {\it cluster} top entry of the 
subtree describes the hardware, software and middleware
properties of a cluster. Grid-enabled queues are represented by their 
{\it queue} entries. Active Grid jobs and authorized Grid users are 
described by their corresponding {\it job} and {\it authuser} 
entries which are located under their hosting queues.
The {\it job} and {\it authuser} entries belonging to the same 
queue are grouped in two distinct subtrees,
the branching is accomplished by structural {\it nordugrid-info-group=job} 
and {\it nordugrid-info-group=user} entries.

The storage and data indexing services are represented by their corresponding
single LDAP entries, currently no LDAP subtree is associated to them.

\begin{figure}[hb]
\centering
{ {\scalebox{0.4}{\includegraphics{cluster_dit.png}}} }
\caption{\label{fig:dit_subcluster} The schematic picture of an LDAP subtree representing a
computing resource. The cluster subtree is part of the LIT shown in Fig:~\ref{fig:dit}}
\end{figure}

\subsection{Globus Mds}
These are the globus-mds attributes that we have incorporated into
arc.  The schema contains special objectclasses: {\it Mds}, 
{\it MdsVo}, {\it MdsVoName}, {\it MdsServiceLdap} and {\it MdsService}
whose role it is to create structural entries in the LDAP tree.

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-validfrom}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-validfrom}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {time}#\\
    Example: &\verb# 20050307103026Z#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-validto}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-validto}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {time}#\\
    Example: &\verb# 20050307103026Z#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-keepto}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-keepto}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {time}#\\
    Example: &\verb# 20050307103026Z#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-vo-name}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-vo-name}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {Locally unique VO name}#\\
    Example: &\verb# local #\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-vo-op-name}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-vo-op-name}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {Locally unique Op name}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-type}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-type}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {Mds service type}#\\
    Example: &\verb# ldap #\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-protocol}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-protocol}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {Service protocol OID}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-port}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-port}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {Service TCP port}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-hn}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-hn}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {Service FQDN hostname}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-url}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-url}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {Service URL}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-ldap-suffix}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-ldap-suffix}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {DN suffix of service}#\\
    Example: &\verb# Mds-Vo-Name=local, o=grid #\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-ldap-timeout}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-ldap-timeout}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {time in minutes}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-ldap-sizelimit}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-ldap-sizelimit}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {size}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-service-ldap-cachettl}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-ldap-cachettl}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {time in minutes}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-sercice-ldap-ttl}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-service-ldap-ttl}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {time in minutes}#\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-reg-status}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-reg-status}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {status}#\\
    Example: &\verb# VALID #\\
    Related xRSL:&none \\
  \end{tabular}

\phantomsection
  \addcontentsline{toc}{subsubsection}{mds-bind-method-servers}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{mds-bind-method-servers}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {type of server}#\\
    Example: &\verb# ANONYM-ONLY #\\
    Related xRSL:&none \\
  \end{tabular}

\subsection{Grouping authuser and job entries: nordugrid-info-group objectclass}

The schema contains a special objectclass the {\it nordugrid-info-group}
whose role is to create structural entries in the LDAP tree.
The entries {\it nordugrid-info-group=jobs} and {\it nordugrid-info-group=users}
of the LIT are separating nordugrid-job and nordugrid-authuser entries of a 
grid queue by grouping them in two
separate LDAP branches under the queue entry (see Fig.~\ref{fig:dit} ).

The objectclass comes with a single attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-info-group-name}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-info-group-name}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&\verb# {users,jobs}#\\
    Related xRSL:&none \\
  \end{tabular}

The {\it nordugrid-info-group-name} attribute
is used to distinguish between jobs or users grouping:
nordugrid-job entries are grouped under the structural entry 
{\it nordugrid-info-group-name=jobs} while nordugrid-authuser entries are
grouped under the {\it nordugrid-info-group-name=users} entry.

\subsection{Computing resources: nordugrid-cluster and nordugrid-queue objectclass}
\label{section:cluster-queue}

The nordugrid schema provides two objectclasses for the description of 
a computing resource. The {\it nordugrid-cluster} is used to describe 
general properties such as ownership, name, location, contact URL, 
pre-installed software environments, Grid scratch space, batch system,
node properties such as architecture, CPUs, network connectivity. 
Dynamic cluster load information, such as number of queued/total jobs, 
is also part of the objectclass information.

The generalized concept of a 
computing queue plays a central role in ARC:
queues are the job submission targets in an ARC-based Grid, during the brokering
process clients select a Grid-enabled queue on a computing resource.
An ARC queue represents either a traditional batch queue of a local resource management
system (LRMS) such as the PBS or describes an entire LRMS when the 
LRMS does not support conventional queues
(Condor and SGE is handled this way). The very special  LRMS, the UNIX fork
is also described as queue. 
The {\it nordugrid-queue} objectclass is designed to describe the generalized concept of
a computing queue. Besides the usual queue-specific information (queue status
and limits, number of running/queueing jobs)  queue-level node attributes are also 
introduced to describe hardware/software 
characteristics of computing nodes assigned to a certain queue.  
Also notice that the schema makes possible the 
distinction of Grid and non-grid jobs being managed by the queue.

The ARC schema introduces cluster- and queue-level node attributes
together with two homogeneity attributes
in order to handle possible inhomogeneity within a computing resource.
The schema is designed to be capable describing inhomogeneous resources with a
queue-level homogeneity assumption but higher level inhomogeneity
can also be treated with less accuracy.

In case of homogeneous nodes the {\it nordugrid-cluster-homogeneity=true}
is set and the cluster-level node attributes carry the relevant information.
If the nodes are inhomogeneous the {\it nordugrid-cluster-homogeneity=no}
is set and the cluster-level node attributes are either not set or
their value refers to the smallest/slowest/least powerful node.
Suppose the nodes can be organized into homogeneous subgroups, 
this case the queue-level node attributes are used to 
describe the properties of the homogeneous nodes
assigned to the same queue. Clients should always treat the 
queue-level node attributes with higher priority than the 
cluster-level ones. The {\it nordugrid-queue-homogeneity=true}
attribute value is used to specify the node homogeneity within a queue.
The {\it nordugrid-queue-homogeneity=no} means that if 
a given queue-level node attribute is set it
refers to the smallest/slowest/least powerful node.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-name}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-name}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& FQDN\\
    Example:&  \verb#nordugrid-cluster-name: gate1.monstercluster.nordugrid.org#\\
    Related xRSL:& \verb#cluster#\\
    UI role:& used in matchmaking, monitoring, job manipulation \\

  \end{tabular}

Description: The fully qualified domain name of the front-end machine of the 
cluster. This attribute is used in the Distinguished Name of a cluster LDAP entry.

%DISCUSS several computing resource running on the same box: what makes the 
% cluster-name unique?

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-aliasname}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-aliasname}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-cluster-aliasname: Grid Monster#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: A free form text attribute displaying the alias name of the computing resource.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-contactstring}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-contactstring}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& URL\\
    Example:& \verb#nordugrid-cluster-contactstring: gsiftp://bambi.hep.lu.se:2811/jobs#\\
    Related xRSL:& none\\    
    UI role:& used during the job submission process\\
  \end{tabular}

Description: The URL of the job submission service running on the cluster front-end.
Clients must use this attribute to determine the URL of the job submission gateway
available on the cluster.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-support}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-support}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& RFC822 email address\\
    Example:& \verb#nordugrid-cluster-support: help@gridcluster.gridcenter.org#\\
    Related xRSL:& none\\
    UI role:& none \\
  \end{tabular}

Description: The support email address of the Grid-enabled computing resource,
users are suggested to use this address in case they need to contact the site.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-location}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-location}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& Postal ZIP code with two letter country prefix\\
    Example:& \verb#nordugrid-cluster-location: SE-22100#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: The geographical location of the cluster, preferably
specified as a postal code with a two letter country prefix.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-owner}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-owner}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-cluster-owner: Danish Center for Grid Computing#\\
    Example:& \verb#nordugrid-cluster-owner: Copenhagen University#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: The multivalued attribute is used to display the owner of the resource.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-acl}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-acl}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed form syntax\\
    Example:& \verb#nordugrid-cluster-acl: VO:ATLAS#\\
    Example:& \verb#nordugrid-cluster-acl: VO:developers.nordugrid.org#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: The multivalued attribute is used to display authorization rules 
applied on a cluster. The attribute value follows a {\it fixed form syntax}.
Current syntax is rather coarse-grained and primitive: a "VO:" prefix followed by a VO name 
means that the given VO is authorized on the cluster. 
Note that there are no standards for VO names.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-issuerca}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-issuerca}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& LDAP Distinguished Name\\
    Example:& \verb#nordugrid-cluster-issuerca: /C=DE/O=GermanGrid/CN=GridKa-CA#\\
    Related xRSL:& none\\
    UI role:& used during the job submission, matchmaking\\
  \end{tabular}

Description: The certificate issuer of the cluster, 
the DN of the CA which issued the host certificate is shown by the attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-issuerca-hash}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-issuerca-hash}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& Hash value of a CA certificate\\
    Example:& \verb#nordugrid-cluster-issuerca-hash: 1f0e8352#\\
    Related xRSL:& none\\
    UI role:& ???\\
  \end{tabular}

Description: The hash value of the certificate of the issuer CA of the cluster, 
the hash of the certificate of the CA which issued the host certificate 
used by the cluster is shown by the attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-trustedca}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-trustedca}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& LDAP Distinguished Name\\
    Example:& \verb#nordugrid-cluster-trustedca: /C=DE/O=GermanGrid/CN=GridKa-CA#\\
    Example:& \verb#nordugrid-cluster-trustedca: /DC=ORG/DC=SEE-GRID/CN=SEE-GRID CA#\\
    Related xRSL:& none\\
    UI role:& ???\\
  \end{tabular}

Description: The DNs trusted by the cluster are shown by this
multivalued attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-credentialexpirationtime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-credentialexpirationtime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}
    Attribute value:& GMT formatted time stamp\\
    Example:& \verb#nordugrid-job-credentialexpirationtime: 20050222120449Z#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The expiration date of the shortest living credential affecting
the cluster's x509 environment in GMT time format.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-lrms-type}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-lrms-type}
  \end{shaded}
  \begin{tabular}{lp{10cm}}
    Attribute value:& \verb#{PBSPro, OpenPBS, torque, SGE,SGEEE, fork, Condor,ll,SLURM}#\\
    Example:& \verb#nordugrid-cluster-lrms-type: torque#\\
    Related xRSL:& none\\
    UI role:& it will be used in the brokering (not implemented yet)\\
  \end{tabular}

Description:  The type of the Local Resource Management System (LRMS) running on the cluster.
ARC currently supports the PBS family, the SGE family, the Condor, IBM LoadLeveler, SLURM and the UNIX fork 
batch systems.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-lrms-version}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-lrms-version}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& version string\\
    Example:& \verb#nordugrid-cluster-lrms-version: 1.0.1p5#\\
    Related xRSL:& none\\
    UI role:& none \\
  \end{tabular}

Description: The vendor specific version string of the Local Resource Management System.
The original vendor-provided LRMS version string is displayed without any modification.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-lrms-config}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-lrms-config}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#Short parallel jobs are prioritised"#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: A free form text attribute for additional remarks on the LRMS setup of the 
cluster. The attribute is purely for 'human consumption'.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-homogeneity}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-homogeneity}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{True, False}#\\
    Example: \verb#nordugrid-cluster-homogeneity: False#\\    
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: A logical flag indicating the homogeneity of the cluster nodes. 
The front-end is not needed to be homogeneous with the nodes. If the nodes 
are declared inhomogeneous on the cluster-level, then the 
the cluster-level node attributes are referring to the properties of the 
slowest/smallest/least powerful node.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-architecture}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-architecture}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{i686, x86_64, alpha, sun4u}#\\
    Example:& \verb#nordugrid-cluster-architecture: i686#\\
    Related xRSL:& \verb#architecture#\\
    UI role:& used in matchmaking/brokering \\
  \end{tabular}

Description: This is a cluster-level node attribute describing the 'hardware type' of the
nodes of the cluster. The 'hardware type' is defined as the output of the \verb#uname -m# UNIX 
command.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-opsys}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-opsys}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-cluster-opsys: Redhat-7.2#\\
    Example:& \verb#nordugrid-cluster-opsys: Linux-2.4.21-mypatch#\\   
    Example:& \verb#nordugrid-cluster-opsys: glibc-2.3.1#\\ 
    Related xRSL:& none\\
    UI role:& not yet used in the brokering\\
  \end{tabular}

Description: The multivalued cluster-level node attribute is meant to describe 
the operating system of the computing nodes. The attribute describes the
operating system via the specification of the software distribution.
The same multivalued attribute is also used to specify the kernel or libc version
in case those differ from the originally shipped ones. 
The attribute value follows a {\it fixed form syntax}:  the distribution name
is given as distroname-version.number where spaces are not allowed.
{\it Kernel} and {\it libc} versions are specified according to a 
fixed form: kernelname-version.number, libcname-version.number.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-nodecpu}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-nodecpu}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-cluster-nodecpu: Dual AMD Athlon(tm) MP Processor 1800+ @ 1500 MHz#\\
    Related xRSL:& \verb#gridTime#\\
    UI role:& used in the brokering process\\
  \end{tabular}

Description: This cluster-level node attribute gives the CPU type information of the cluster nodes
in a fixed format. The string is constructed as \verb#CPU-model-name CPU-frequency MHZ#,
where CPU-model-name and CPU-frequency are vendor specified values (on Linux systems 
the data is taken from the /proc/cpuinfo).

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-benchmark}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-benchmark} ***
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-cluster-benchmark: SPECINT2000 @ 222#\\
    Example:& \verb#nordugrid-cluster-benchmark: SPECFP2000 @ 333#\\
    Related xRSL:& \verb#benchmark#\\
    UI role:& used in brokering\\
  \end{tabular}

Description\footnote{not in a stable release yet}: The multivalued cluster-level node attribute shows the 
performance of the computing nodes with respect to specified benchmarks.
The attribute value follows a fixed syntax: the benchmark name 
and value is separated by "@".

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-nodememory}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-nodememory}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& a number showing the amount in MBs\\
    Example:& \verb#nordugrid-cluster-nodememory: 900#\\
    Related xRSL:& \verb#memory#\\
    UI role:& used in matchmaking/brokering\\
  \end{tabular}

Description: The amount of node memory in MBs which can be guaranteed to be
available for the application running on the node. Please note in most cases
it is less than the physical memory installed in the nodes.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-totalcpus}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-totalcpus}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-cluster-totalcpus: 60#\\
    Related xRSL:& \verb#count#\\
    UI role:& used in matchmaking/brokering\\
  \end{tabular}

Description: The total number of CPUs of the computing resource being controlled by the
LRMS. It is possible that not all of them are available for Grid jobs (e.g. the cluster
has a non-grid queue with dedicated nodes).

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-cpudistribution}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-cpudistribution}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-cluster-cpudistribution: 1cpu:36 2cpu:7#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description:  The CPU distribution over the nodes given in the form of 
{\it n}cpu:{\it m} where {\it n} is the number of CPUs per machine and {\it m} 
is the number of such computers, an example: {\it 1cpu:3,2cpu:4,4cpu:1}
represents a cluster with 3 single CPU machines, 4 dual machines 
and one computer with 4 CPUs.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-sessiondir-free}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-sessiondir-free}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the amount in MBs\\
    Example: &\verb#nordugrid-cluster-sessiondir-free: 447870#\\
    Related xRSL:& \verb#disk#\\
    UI role:& used in matchmaking/brokering\\
  \end{tabular}

Description:  Each Grid job has a dedicated Grid scratch area called the 
session directory. This attribute shows the available free disk space in MBs for the 
session directories. As a minimum protection the broker compares the available 
disk space to the size of the uploadable input data and rejects the clusters 
with insufficient free space.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-sessiondir-total}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-sessiondir-total}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the amount in MBs\\
    Example: &\verb#nordugrid-cluster-sessiondir-total: 1447870#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The total disk space in MB allocated on the cluster to host the 
grid job's session directories.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-sessiondir-lifetime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-sessiondir-lifetime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& time interval specified in minutes\\
    Example: &\verb#nordugrid-cluster-sessiondir-lifetime: 10080#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The lifetime of the job's session directory after the job has completed.
The job's session directory together with all the user's data is erased when the 
{\it nordugrid-cluster-sessiondir-lifetime} has expired counted from the completion of the 
job.

%DISCUSS: should it be brokerable?

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-cache-free}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-cache-free}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&  number showing the amount in MBs\\
    Example: & \verb#nordugrid-cluster-cache-free: 2048#\\
    Related xRSL:& \verb#disk#\\
    UI role:& used in matchmaking/brokering process\\
  \end{tabular}

Description: ARC clusters can provide a cache area to store frequently used input
data. Upon user request the input data is placed into the cache
instead of the session directory of the job (input data in a session directory
is not accessible by a consequent jobs). This attribute shows the available 
space in the cache in MBs.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-cache-total}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-cache-total}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&  number showing the amount in MBs\\
    Example:& \verb#nordugrid-cluster-cache-total: 8048#\\
    Related xRSL:& none\\
    UI role:& none\\    
  \end{tabular}

Description: The total space in MBs allocated for the cache service.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-runtimeenvironment}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-runtimeenvironment}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& Runtime Environment string\cite{rer}\\
    Example:& \verb#nordugrid-cluster-runtimeenvironment: APPS/MODELCHECK/DUPPAAL#\\
    Related xRSL:& \verb#runtimeenvironment#\\
    UI role:& used in matchmaking\\
  \end{tabular}

Description: Runtime Environments are special pre-installed and pre-configured 
software packages provided in a standard way by the computing resources. 
A Runtime Environment Registry \cite{rer} maintains a list of available REs with
pointers to RE descriptions. The multivalued attribute is used to display the REs available 
and supported on the cluster.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-localse}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-localse}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& URL\\
    Example:& \verb#nordugrid-cluster-localse: gsiftp://hypatia.uio.no/scratch/#\\
    Related xRSL:& none\\
    UI role:& used in brokering\\ 
  \end{tabular}

Description: This multivalued parameter tells the broker that certain storage URLs
should be considered "locally" available on the cluster.
The attribute gives the URL of storage elements considered to be "local" to the cluster.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-middleware}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-middleware}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form string to represent a software package\\
    Example:& \verb#nordugrid-cluster-middleware: nordugrid-0.4.4#\\
    Example:& \verb#nordugrid-cluster-middleware: globus-2.4.3-15ng#\\
    Related xRSL:& \verb#middleware#\\
    UI role:& used in matchmaking\\
  \end{tabular}

Description: This multivalued attribute specifies the middleware packages 
installed on the cluster.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-totaljobs}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-totaljobs}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-cluster-totaljobs: 580#\\
    Related xRSL:& none\\
    UI role:& ? \\
  \end{tabular}

Description: The total number of non-completed jobs in the cluster.
Totaljobs includes both Grid and non-grid jobs, non-grid jobs are
those batch jobs which are directly submitted to the LRMS by a local user.
Grid jobs with {\it FINISHING, FINISHED, DELETED} status are discarded.

%DISCUSS: the real meaning of totaljobs in view of the new jobstates

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-usedcpus}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-usedcpus}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-cluster-usedcpus: 12#\\
    Related xRSL:& none\\
    UI role:& ?\\ 
  \end{tabular}

Description: The total number of occupied CPUs in the cluster. 
The attribute displays the number of busy/used CPUs regardless 
if the CPU is occupied by a Grid or a non-grid job.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-queuedjobs}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-queuedjobs}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-cluster-queuedjobs: 812#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The total number of jobs (grid and non-grid) not-yet running: preparing
(e.g. Grid stage-in process) or waiting to run on a cluster. 
A Grid job submitted to the cluster needs to complete several stages before 
it arrives to the LRMS. All these 'pre-LRMS Grid jobs' plus the LRMS queuing jobs 
are taken into account in the {\it nordugrid-cluster-queuedjobs} attribute.
WARNING: The attribute is DEPRECATED in the 0.6 release!

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-prelrmsqueued}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-prelrmsqueued}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-cluster-prelrmsqueued: 423#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The total number of grid jobs not-yet reached the
LRMS. These jobs are being processed or put on hold by the grid layer
of the cluster. A Grid job submitted to the cluster needs to complete several 
stages before it arrives to the LRMS. All these 'pre-LRMS Grid jobs' 
are taken into account in the {\it nordugrid-cluster-prelrmsqueued} attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-nodeaccess}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-nodeaccess}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{inbound,outbound}#\\
    Example:&  \verb#nordugrid-cluster-nodeaccess: inbound#\\
    Example:&  \verb#nordugrid-cluster-nodeaccess: outbound#\\
    Related xRSL:& \verb#nodeaccess#\\
    UI role:& used in matchmaking\\ 
  \end{tabular}

Description: The inbound/outbound network accessibility of the nodes
determines how the nodes can connect to the Internet: {\it outbound}
access means the nodes can connect to the outside world while {\it inbound}
access means the nodes can be connected from outside.
Specifying both {\it inbound, outbound} means the nodes are sitting on an 
open network. If a cluster has not set this attribute then the 
nodes are assumed to be sitting on a private isolated network.

%DISCUSS: what about introducing "none", also more flexible description is needed
%with respect to ports

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-comment}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-comment}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-cluster-comment: This cluster is dedicated for CMS calculations#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: The free form attribute displays some additional information about 
the cluster. Sometimes it contains an URL where more information can be read
about the cluster.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-cluster-interactive-contactstring}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-cluster-interactive-contactstring}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& URL\\
    Example:& \verb#nordugrid-cluster-interactive-contactstring: gsissh://atlas.hpc.unimelb.edu.au:2200#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The URL for interactive login to the cluster.
Some clusters offer GSI-enabled ssh services, this attribute presents the
URL of that service.

%DISCUSS: clients should be able to make use of this attribute!!

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-name}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-name}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string representing a queue name\\
    Example:& \verb#nordugrid-queue-name: longqueue#\\
    Related xRSL:&  \verb#queue#\\
    UI role:& used during job submission\\
  \end{tabular}

Description: The name of the Grid-enabled batch queue.
The special value {\it fork} is used for the 'UNIX fork' system.
This attribute constitutes the Distinguished Name of a queue LDAP entry.

%DISCUSS: special value for LRMS which does not support queues, like fork, condor, etc..

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-status}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-status}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{active#\\
    & \verb#inactive#\\
    & \verb#inactive, grid-manager does not accept new jobs#\\
    & \verb#inactive, grid-manager is down#\\
    & \verb#inactive, gridftp is down}#\\
    Example:& \verb#nordugrid-queue-status: inactive, grid-manager is down#\\ 
    Related xRSL:& none\\
    UI role:& used in brokering\\
  \end{tabular}

Description: The generalized status of the queue. Besides the usual 
batch system queue status the attribute also takes into account the 
status of the Grid services such as the {\it grid-manager} and the {\it gridftp server} 
serving the queue. Grid jobs are only submitted to queues with {\it active} status.

%TODO: review queue-status in view of glueue/cim service status flags
%DISCUSS: correlation between freecpus & queue-status

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-comment}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-comment}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-queue-comment: Special queue dedicated to BIO Apps#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: A free form attribute containing additional information about the 
queue.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-schedulingpolicy}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-schedulingpolicy}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-queue-schedulingpolicy: SIMPLE FIFO#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: 
The attribute is used to describe the implied scheduling policy of the queue (i.e. FIFO).

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-homogeneity}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-homogeneity}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{True, False}#\\
    Example:& \verb#nordugrid-queue-homogeneity: False#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: 
A logical flag indicating the homogeneity of the queue nodes
If the nodes are declared inhomogeneous on the queue-level, then the 
the queue-level node attributes are referring to the properties of the 
slowest/smallest/least powerful node within the queue.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-nodecpu}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-nodecpu}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-queue-nodecpu: Dual AMD Athlon(tm) MP Processor 1800+ @ 1500 MHz#\\
    Related xRSL:& \verb#gridTime#\\
    UI role:& used in brokering \\
  \end{tabular}

Description: This queue-level node attribute gives the CPU type information of the queue nodes
in a fixed format. The string is constructed as \verb#CPU-model-name CPU-frequency MHZ#,
where CPU-model-name and CPU-frequency are vendor specified values (on Linux systems 
the data is taken from the /proc/cpuinfo).

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-nodememory}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-nodememory}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& a number showing the amount in MBs\\
    Example:& \verb#nordugrid-queue-nodememory: 600#\\
    Related xRSL:& \verb#memory#\\
    UI role:& used in matchmaking/brokering\\
  \end{tabular}

Description: The queue-level node attribute shows the amount of node memory in MBs 
which can be guaranteed to be available for the application running on the node. 
Please note in most cases it is less than the physical memory installed in the nodes.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-architecture}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-architecture}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{i686, x86_64, alpha, sun4u}#\\
    Example:& \verb#nordugrid-queue-architecture: x86_64#\\
    Related xRSL:& \verb#architecture#\\
    UI role:& used in matchmaking \\
  \end{tabular}

Description: This is a queue-level node attribute describing the 'hardware type' of the
nodes of the queue. The 'hardware type' is defined as the output of the \verb#uname -m# unix 
command.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-opsys}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-opsys}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-queue-opsys: Redhat-7.2#\\
    Example:& \verb#nordugrid-queue-opsys: Linux-2.4.21-mypatch#\\   
    Example:& \verb#nordugrid-queue-opsys: glibc-2.3.1#\\     
    Related xRSL:& none\\
    UI role:& not yet used in brokering\\
  \end{tabular}

Description: The multivalued queue-level node attribute is meant to describe 
the operating system of the computing nodes. The attribute describes the
operating system via the specification of the software distribution.
The same multivalued attribute is also used to specify the kernel or libc version
in case those differ from the originally shipped ones. 
The attribute value follows a {\it fixed form syntax}:  the distribution name
is given as distroname-version.number where spaces are not allowed.
{\it Kernel} and {\it libc} versions are specified according to a 
fixed form: kernelname-version.number, libcname-version.number.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-benchmark}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-benchmark} ***
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-queue-benchmark: SPECINT2000 @ 111#\\
    Example:& \verb#nordugrid-queue-benchmark: SPECFP2000 @ 555#\\    
    Related xRSL:& \verb#benchmark#\\
    UI role:& used in brokering\\
  \end{tabular}

Description\footnote{not in a stable release yet}: The multivalued queue-level node attribute shows the 
performance of the computing nodes with respect to specified benchmarks.
The attribute value follows a fixed syntax: the benchmark name 
and value is separated by "@".

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-maxrunning}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-maxrunning}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-maxrunning: 120#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The batch queue limit indicating the 
maximum number of jobs allowed to run from this queue.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-maxqueuable}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-maxqueuable}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-maxqueuable: 500#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The batch queue limit indicating the 
maximum number of jobs allowed to reside in the queue (both queuing and running).

%TODO check this attribute in the providers

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-maxuserrun}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-maxuserrun}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-maxuserrun: 12#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The batch queue limit indicating the 
maximum number of jobs a user can run at the same time in the queue.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-maxtotalcputime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-maxtotalcputime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-queue-maxtotalcputime: 120#\\
    Related xRSL:& \verb#cpuTime#\\
    UI role:& used in matchmaking, relevant for parallel jobs\\
  \end{tabular}

Description: The batch queue limit giving the maximum total CPU time (in
minutes) a job can use/request within this queue. The total is calculated over
all processes belonging to the job (on all nodes, in case of parallel jobs).
Only published on clusters that support such a limit.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-maxcputime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-maxcputime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-queue-maxcputime: 120#\\
    Related xRSL:& \verb#cpuTime#\\
    UI role:& used in matchmaking\\
  \end{tabular}

Description: The batch queue limit giving the maximum CPU time per CPU (in
minutes) a job can use/request within this queue. If
nordugrid-queue-maxtotalcputime is also published, clients should ignore
nordugrid-queue-maxcputime.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-mincputime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-mincputime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-queue-mincputime: 10#\\
    Related xRSL:& \verb#cpuTime#\\
    UI role:& used in matchmaking\\ 
  \end{tabular}

Description: The queue limit giving the lower value of 
job CPU time requests in minutes allowed in the queue.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-defaultcputime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-defaultcputime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-queue-defaultcputime: 70#\\
    Related xRSL:& \verb#cpuTime#\\
    UI role:& ?\\
  \end{tabular}

Description: The default CPU time assigned to this queue in minutes.
Jobs not specifying their CPU time requests are set to this default
CPU time value by the LRMS.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-maxwalltime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-maxwalltime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-queue-maxwalltime: 140#\\
    Related xRSL:& \verb#cpuTime#\\
    UI role:& ???\\
  \end{tabular}

Description: The batch queue limit gives the maximum  walltime (in minutes) a job 
can use/request within this queue.

Comment: The {\it nordugrid-queue-maxwalltime} attribute is introduced
in the 0.6.1 ARC release.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-minwalltime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-minwalltime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-queue-minwalltime: 30#\\
    Related xRSL:& \verb#cpuTime#\\
    UI role:& ?\\ 
  \end{tabular}

Description: The queue limit giving the lower value of 
job walltime requests in minutes allowed in the queue.

Comment: The {\it nordugrid-queue-minwalltime} attribute is introduced
in the 0.6.1 ARC release.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-defaultwalltime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-defaultwalltime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-queue-defaultwalltime: 90#\\
    Related xRSL:& \verb#cpuTime#\\
    UI role:& ?\\
  \end{tabular}

Description: The default  walltime assigned to this queue in minutes.
Jobs not specifying their CPU time requests are set to this default
CPU time value by the LRMS.

Comment: The {\it nordugrid-queue-defaultwalltime } attribute is introduced
in the 0.6.1 ARC release.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-running}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-running}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-running: 14#\\
    Related xRSL:& none\\
    UI role:& ?\\ 
  \end{tabular}

Description: The attribute gives the number of 
CPUs being occupied by running jobs in the queue
including both the Grid and  non-Grid jobs.  Multi-node jobs are counted 
with their multiplicity: a four-node running job increases the value of 
{\it nordugrid-queue-running} by four.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-gridrunning}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-gridrunning}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-gridrunning: 6#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The attribute gives the number of CPUs currently 
being occupied by running Grid jobs in the queue.
Multi-node Grid jobs are counted with their multiplicity: 
a four-node running job increases the value of {\it nordugrid-queue-running}
by four.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-queued}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-queued}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-queued: 23#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The attribute gives the number of jobs, including both Grid and non-Grid, 
waiting in the queue. Each queuing job counts as one regardless their multiplicity. 
WARNING: The attribute is DEPRECATED in the 0.6 release!

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-gridqueued}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-gridqueued}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-gridqueued: 11#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The attribute gives the number of waiting Grid jobs in the batch queue. 
Each queuing job counts as one regardless their multiplicity.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-localqueued}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-localqueued}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-localqueued: 24#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The attribute gives the number of locally submitted non-Grid jobs
waiting in the queue. Each queuing job counts as one regardless their 
multiplicity.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-prelrmsqueued}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-prelrmsqueued}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-prelrmsqueued: 25#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The attribute gives the number of Grid jobs belonging to this 
queue being processed or waiting in the Grid-layer before the LRMS submission.
Each queuing job counts as one regardless their multiplicity.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-queue-totalcpus}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-queue-totalcpus}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-queue-totalcpus: 11#\\
    Related xRSL:& ?\\
    UI role:& ?\\    
  \end{tabular}

Description: 
Some of the batch systems provides the possibility of assigning 
nodes to queues. This attribute shows the total number of CPUs 
exclusively dedicated to the queue within such batch system.

\subsection{Grid jobs: nordugrid-job objectclass}

In the NorduGrid/ARC information system every Grid job submitted to
a Grid-enabled resource is represented by a \emph{job} entry.
Job entries are generated and optionally cached in the local LDAP tree
of the hosting resource. This implies that job information within ARC is 
coupled to the execution Grid resource, namely for job status query or
job monitoring the LDAP server of the hosting resource has to be contacted,
this way ARC implements a fully distributed job status monitoring system:
no central database or service is used for job status query/monitoring.

A job entry is generated and stored in the local LDAP tree for every 
existing Grid job on a resource. The job entry is kept in the local LDAP tree as long as the
job is handled by the resource, when a job is removed from a resource 
the corresponding  job entry is also deleted from the local LDAP tree.
This implies that the ARC information system contains no information about 
non-existing deleted Grid jobs, another  ARC service, the logging service is 
designed to store historical job information \cite{logger}.

Job monitoring and status query of active Grid jobs is entirely based upon the 
LDAP job entries stored in the local information trees. 
Job entries carry information collected from the 
grid layer running on the resource (read from the job control files
of the ARC Grid manager) and from the LRMS system.
The attributes of the {\it nordugrid-job} objectclass are designed to 
provide all the necessary information.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-globalid}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-globalid}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& URL\\
    Example:& \verb#gsiftp://farm.hep.lu.se:2811/jobs/243361109008699845213642#\\
    Related xRSL:& none\\
    UI role:&  used as a job handle in job management and as an URL in data movement\\
  \end{tabular}

Description: ARC uses a GridFTP URL as a unique global jobID. The globally 
unique GridFTP URL is used as a handle in job manipulations such as rerun,
kill or output retrieval. The GridFTP URL can also be used to access the 
session directory of the Grid job during the job's existence on the resource.
The {\it nordugrid-job-globalid} attribute constitutes to the DN of the job entry.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-globalowner}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-globalowner}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& LDAP Distinguished Name\\
    Example:& \verb#/O=Grid/O=NorduGrid/OU=nordugrid.org/CN=Lars Jenssen#\\
    Related xRSL:& none\\
    UI role:& used during the job discovery process of {\it ngsync}\\
  \end{tabular}

Description:  The LDAP Subject Name of the job owner as specified in his/her Grid credentials.
A Grid user or a client can easily find his/her own jobs on the Grid-enabled 
resource by issuing an LDAP search with a filter 
of {\it nordugrid-job-globalowner=his/her SN}.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-jobname}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-jobname}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-job-jobname: ngtest-job-80#\\
    Related xRSL:& \verb#jobname#\\
    UI role:& {\it ngget} optionally makes use of it\\ 
  \end{tabular}

Description:  The job name specified by the user with the {\it jobname} xRSL attribute.
The client tools optionally can use the user-specified job name as
the name of the local copy of the session directory of the job.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-execcluster}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-execcluster}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& FQDN\\
    Example:& \verb#nordugrid-job-execcluster: farm.hep.lu.se#\\
    Related xRSL:& \verb#cluster#\\
    UI role:& ?\\
  \end{tabular}

Description:  The name of the execution cluster specified as the fully 
qualified domain name of the front-end machine.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-execqueue}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-execqueue}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string representing a queue name\\
    Example:& \verb#nordugrid-job-execqueue: fastq#\\
    Related xRSL:&  \verb#queue#\\
    UI role:& ?\\    
  \end{tabular}

Description: The name of the execution queue hosting the Grid job.
Within ARC the queues are coupled to clusters and used together as 
submission targets. Therefore the execution queue is
selected together with the executing cluster during the brokering process, which means
that the value of the {\it nordugrid-job-execqueue} is known for all the accepted Grid jobs
even if they are not yet handed over to the local batch system.
Also recall that Grid job entries are linked under their hosting queue entries in the 
local LDAP tree.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-executionnodes}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-executionnodes}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string representing a node name\\
    Example:& \verb#nordugrid-job-executionnodes: n3#\\
    Example:& \verb#nordugrid-job-executionnodes: n4#\\
    Example:& \verb#nordugrid-job-executionnodes: n5#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: The multivalued attribute presents the local node names of the cluster
nodes which are occupied by the running Grid job. Every node being used by the job is 
listed with an attribute value pair. The shown example corresponds to a 3-node-job 
running on the nodes n3,n4,n5. Obviously, the {\it nordugrid-job-executionnodes}
attribute is only available for jobs being run or already completed 
in the local batch system.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-submissionui}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-submissionui}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-job-submissionui: 130.235.91.118:45447;guest4.hep.lu.se#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The attribute specifies client machine from where the job was submitted
in a fixed format string. The string contains the submission host's IP,
the port and the host name.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-submissiontime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-submissiontime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& GMT formatted time stamp\\
    Example:& \verb#nordugrid-job-submissiontime: 20050220155311Z#\\
    Related xRSL:& none\\
    UI role:& ? \\
  \end{tabular}

Description: The time stamp of the submission of the job specified 
in Globus MDS time format (GMT).
Job submission is the process when the client handles over the job request to the 
selected resource and the resource returns a job handle (the globally unique job ID).

%TODO find out what this GMT is, find/give a reference

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-sessiondirerasetime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-sessiondirerasetime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& GMT formatted time stamp\\
    Example:& \verb#nordugrid-job-sessiondirerasetime: 20050220165311Z#\\
    Related xRSL:& \verb#lifeTime#\\
    UI role:& none\\
  \end{tabular}

Description: Within an ARC Grid every Grid job is confined to a dedicated 
area on the execution cluster which is called the {\it session directory}. 
After job completion the {\it session directory} of the 
grid job contains all the job and debugging output which was not requested 
to be uploaded to a storage element. The {\it session directory} can be accessed 
and the output data within the directory be downloaded for a limited time 
after the job completion. The date when the {\it session directory} is removed 
from the cluster is given in GMT time format by the 
{\it nordugrid-job-sessiondirerasetime} attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-proxyexpirationtime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-proxyexpirationtime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& GMT formatted time stamp\\
    Example:& \verb#nordugrid-job-proxyexpirationtime: 20050222120449Z#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The expiration time of the proxy assigned to the job displayed in GMT
time format. A valid proxy is required for the stage-out phase of the Grid job if the
stage out target makes use of GSI-based authentication and authorization.
Intelligent clients can use this attribute to check if the job possesses a valid proxy 
and  automatically initiate proxy renewal in case a proxy expiration.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-completiontime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-completiontime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& GMT formatted time stamp\\
    Example:& \verb#nordugrid-job-completiontime: 20050222120449Z#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description:  The completion time of the Grid job expressed in GMT time format.
Job completion refers to the {\it FINISHED} job state when the job completed 
all the requested operations including both job execution and stage out.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-runtimeenvironment}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-runtimeenvironment}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string representing a valid RuntimeEnvironment\\
    Example:& \verb#nordugrid-job-runtimeenvironment: APPS/CHEM/DALTON-1.2.1-1.0#\\
    Related xRSL:& \verb#runtimeenvironment#\\
    UI role:& none\\
  \end{tabular}

Description: The multivalued attribute lists the RuntimeEnvironments requested by 
the job.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-gmlog}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-gmlog}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string representing a directory name\\
    Example:& \verb#nordugrid-job-gmlog: Grid_manager_logdir#\\
    Related xRSL:& \verb#gmlog#\\
    UI role:& optionally used for status monitoring\\
  \end{tabular}

Description: The name of the directory which contains the Grid session related logs
within the {\it session directory} of the job.
The {\it gmlog} directory contains plenty of useful information for tracking or debugging
the Grid job being processed on the execution site.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-clientsoftware}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-clientsoftware}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string\\
    Example:& \verb#nordugrid-job-clientsoftware: nordugrid-0.5.21#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description:  The client software which was used to submit the job.
The client software needs to be able to communicate its version to the Grid
layer of the resource in order to have this attribute set.

%TODO ask the web portals, GUIS to specially mark their jobs

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-stdout}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-stdout}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string representing a file name\\
    Example:& \verb#nordugrid-job-stdout: JG.out#\\
    Related xRSL:& \verb#stdout#\\
    UI role:& ?\\
  \end{tabular}

Description: The name of the file which contains the standard output
of the job.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-stderr}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-stderr}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string representing a file name\\
    Example:& \verb#nordugrid-job-stderr: JG.err#\\
    Related xRSL:& \verb#stderr#\\
    UI role:& ?\\
  \end{tabular}

Description: The name of the file which contains the standard error of the job.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-stdin}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-stdin}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string representing a file name\\
    Example:& \verb#nordugrid-job-stdin: my.job_input#\\
    Related xRSL:& \verb#stdin#\\
    UI role:& ?\\
  \end{tabular}

Description: The name of the file which is used as the standard input of the job.

%TODO check if it works, try to set it in xRSL

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-cpucount}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-cpucount}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-job-cpucount: 7#\\
    Related xRSL:& \verb#count#\\
    UI role:& none\\
  \end{tabular}

Description: The number of CPUs requested by the job.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-reqcputime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-reqcputime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-job-reqcputime: 146#\\
    Related xRSL:& \verb#cpuTime#\\
    UI role:& none\\
  \end{tabular}

Description: The CPU time request by the job specified in minutes.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-reqwalltime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-reqwalltime} ***
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-job-reqwalltime: 146#\\
    Related xRSL:& \verb#wallTime#\\
    UI role:& none\\
  \end{tabular}

Description\footnote{not in a stable release yet}: The wallclock time request of the job specified in minutes.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-queuerank}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-queuerank}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-job-queuerank: 13#\\
    Related xRSL:& none\\
    UI role:& the information can be used to initiate resubmission\\
  \end{tabular}

Description: The attribute displays the queue position of the Grid job being 
idle in a batch queue. Most of the cases the given value is rather 
approximate since the majority of schedulers are not able to provide 
accurate information.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-lrmscomment}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-lrmscomment}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-job-lrmscomment: Job is not running no available resources#\\
    Related xRSL:& none\\
    UI role:& none\\ 
  \end{tabular}

Description: The optional comment provided by the Local Resource Management System.
The attribute is only available in version 0.4.x, it was replaced by the more general
{\it nordugrid-job-comment}.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-comment}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-comment}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-job-comment: LRMS: Job is not running no available resources#\\
    Example:& \verb#nordugrid-job-comment: GM: The grid-manager is down#\\
    Related XRSL:& none\\
    UI role:& none\\ 
  \end{tabular}

Description: The multivalued attribute contains the optional job comments provided by either 
the Grid Layer (e.g. Grid Manager) or the Local Resource Management System. 
The attribute has been introduced as a more general replacement of the 
{\it nordugrid-job-lrmscomment} and available with the ARC 0.6 version.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-usedcputime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-usedcputime}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-job-usedcputime: 144#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: The consumed CPU time of the job in minutes as it was reported by the 
local batch system.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-usedwalltime}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-usedwalltime} ***
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the time interval in minutes\\
    Example:& \verb#nordugrid-job-usedwalltime: 166#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description\footnote{not in a stable release yet}: The consumed wall clock time of the job in minutes as it was reported by the 
local batch system.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-usedmem}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-usedmem}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number representing memory consumption in KBs\\
    Example:& \verb#nordugrid-job-usedmem: 4376#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: The memory usage of the job reported in KBs.

%TODO investigate when this attribute is set, only for LRMS:R or for FINISH* ?

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-exitcode}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-exitcode} ***
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-job-exitcode: 127#\\
    Related xRSL:& none\\
    UI role:& used in job status monitoring\\
  \end{tabular}

Description\footnote{not in a stable release yet}: The exit code of the executable of the job obtained from the 
Local Resource Management System.

%DISCUSS ask if 0 exitcode should be the default

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-errors}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-errors}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example v0.4.x:& \verb#nordugrid-job-errors: JOB FAILURE: Failed extracting LRMS ID due to internal error#\\
    Example v.0.6:& \verb#nordugrid-job-errors: Failed extracting LRMS ID due to internal error#\\
    Related xRSL:& none\\
    UI role:& used in job status monitoring\\
  \end{tabular}

Description:  Textual explanation of the job's failure, error message provided by the
Grid layer running on the resource.

0.4.x release implementation:
This attribute was/is used to determine the failure of the job, the  presence of the
attribute indicates a failure in the Grid job execution. 
The attribute text starts with the {\it JOB FAILURE: } prefix.

0.6 release implementation:
The {\it JOB FAILURE: } prefix is dropped from the attribute value,
new {\it FINISHED, KILLED, FAILED} final states were introduced, clients 
don't have to rely on the {\it nordugrid-job-errors} any longer to 
determine job failure.

%DISCUSS: shall we use arc error codes instead of texts?

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-status}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-status}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value v0.4.x:& \verb#{ACCEPTED,PREPARING,SUBMITTING,INLRMS: X,FINISHING,FINISHED,CANCELLING,#\\
                           & \verb#PENDING:ACCEPTED,PENDING:PREPARING,PENDING:INLRMS,DELETED}#\\
    Attribute value v.0.6:& \verb#{ACCEPTING,ACCEPTED,PREPARING,PREPARED,SUBMITTING,INLRMS:X#\\
    & \verb#KILLING,EXECUTED FINISHING,FINISHED,FAILED,KILLED,DELETED}#\\ 
    Example v0.4.x:&  \verb#nordugrid-job-status: FINISHED at: 20020402161013Z#\\
    Example v0.6:& \verb#nordugrid-job-status: FINISHED#\\ 
    Related xRSL:& none\\
    UI role:& used in job status monitoring 
  \end{tabular}

Description: The status of the Grid job. The job state representation 
is undergoing a major change with the upcoming 0.6 release.

{\emph 0.4.x release implementation:} 
The attribute fully exposes the internal Grid Manager job states,
for the explanation of the states {\it ACCEPTED, PREPARING, SUBMITTING, INLRMS, FINISHING, 
CANCELLING, FINISHED} consult the Grid manager manual\cite{grid-manager}. 
The internal {\it INLRMS} state, meaning the job is under the 
control of the Local Resource Management System, is expanded by the 
information system to display the batch system status of the job as well:
{\it INLRMS: R} or  {\it INLRMS: Q} states are used to represent Grid jobs running 
or queuing  in the local batch system.

Completed jobs are labeled by the {\it FINISHED} state regardless their
success or failure. The {\it nordugrid-job-errors} attribute is used to 
distinguish between failed and successfully completed jobs.
The {\it FINISHED} job state also carries information about the completion
time of the job expressed in the GMT time format (see the example v0.4.x above)

{\emph 0.6 release implementation:}
The Grid Manager internals are not fully exposed to the clients, 
new job states are introduced. The terminal job state is separated into
three new states {\it FINISHED, KILLED, FAILED}. 
Furthermore, the {\it nordugrid-job-completiontime} attribute was introduced to 
separate the completion time from the {\it FINISHED} state.

{\it ACCEPTING}
This is the initial job state
The job has reached the cluster, a session directory was created, 
the UI has optionally started to upload some files to the sessiondir,
the job waits to be detected by the Grid manager (GM). \\
internal state: ACCEPTED

{\it ACCEPTED}
The job has been detected by the GM but 
can't go to the next state due to an internal GM-limit. The job may also be in
the ACCEPTED state if the grid manager died during the stagein process 
(while the job was in the PREPARING state)\\
internal state: PENDING:ACCEPTED \\
internal state: PREPARING (in case of dead grid-manager)

{\it PREPARING}
The input data is being gathered into the session directory
(or to the cache), the GM downloads files specified in the user's xRSL.
This is the Grid-stage-in process to the cluster.
This is the latest state when the upload from the UI finishes. 
During this state, the UI can still upload files to the session directory. \\
internal state: PREPARING

{\it PREPARED}
The stage-in process has successfully completed, 
the job is held waiting in the GM's internal queue due to an  
exceeded internal GM limit. 
The job may also be in the PREPARED state if the grid manager died during the LRMS
job submission  process (while the job was in the SUBMITTING state)\\
internal state: PENDING:PREPARING \\
internal state: SUBMIT (in case of a dead grid-manager)

{\it SUBMITTING}
The GM prepares the LRMS job submission script
and submits the job to the LRMS. \\
internal state: SUBMIT
%TODO check with Aleksandr

{\it INLRMS:X}
The job is in the local batch system, the job is 
controlled, managed by the LRMS. This state has several sub-states
which are general mappings of native batch system states.
Currently implemented sub-states:

{\it INLRMS:R} 
The job is running in the LRMS, executing on a node controlled by 
the batch system. \\
internal state: INLRMS

{\it INLRMS:Q}
The job is queuing in the LRMS, waiting for a node, 
being put on hold, for some reason the
job is in a 'pending state' of the LRMS.\\
internal state: INLRMS

{\it INLRMS:S} 
An already running job is in a suspended state.\\
internal state: INLRMS

{\it INLRMS:E} 
The job is finishing in the LRMS, it is 'exiting' from the batch system
either because the job is completed or because it was cancelled.\\
internal state: INLRMS

{\it INLRMS:O} 
Any other native LRMS state which can not be mapped to the above
general states must be labeled as 'O', meaning "other"\\
internal state: INLRMS

{\it KILLING}
The job was requested to be killed and it is being killed by the GM, 
the GM interacts with the LRMS by running the job-cancel script \\
internal state: CANCELING

{\it EXECUTED}
The job has completed in the batch system.  
There are two internal states corresponding to this state:

The job left the LRMS but the GM has not yet recognized this fact.
The information system can't find the job in the batch system any longer 
but the GM still thinks the job is in the batch system (due to its latency). \\
internal state: INLRMS

The job has completed in the batch system
AND the GM scanning process has recognized the job left the batch system
BUT the job is held waiting in the GM internal queue due to an exceeded
GM limit. 
The job may also be in the EXECUTED state if the grid manager died during the 
stageout process (while the job was in the FINISHING state) \\
internal state: PENDING:INLRMS \\
internal state: FINISHING (in case of dead grid-manager)

{\it FINISHING}
The GM is doing the Grid stage-out process, specified output files
are moved to their locations, GM is cleaning up the session directory removing
everything which was not requested to be kept. \\
internal state: FINISHING

{\it FINISHED}
The job has finished ALL its activity on the cluster
AND no errors occurred during the Grid job's lifetime (no job.xx.failed file was 
created in the control-directory).
The state {\it FINISHED} corresponds to the successful Grid job completion. \\
internal state: FINISHED

{\it FAILED}
The job has finished ALL its activity on the cluster AND there occurred some problem during 
the lifetime of the Grid job. The {\it nordugrid-job-errors} and 
{\it nordugrid-job-exitcode} attributes contain more information about the job failure. \\
internal state: FINISHED

{\it KILLED}
The job has finished ALL its activity on the cluster as a result of being killed by
a client. \\
internal state: FINISHED

{\it DELETED}
The job's session directory has removed from the cluster due to the
expired session-directory-lifetime, only minimal set of info is kept about such a job. \\
internal state: DELETED

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-job-rerunable}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-job-rerunable}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{none, PREPARING, INLRMS, FINISHING}#\\
    Example:& \verb#nordugrid-job-rerunable: PREPARING#\\   
    Related xRSL:& \verb#rerun#\\
    UI role:& ngrerun utility makes use of it\\
  \end{tabular}

Description: The attribute is only set for FAILED jobs and its value is either 
{\it none} or the name of the Grid job state from which the job can be rerun
following a client request.

%DISCUSS rerun xrsl attribute and ngrerun

\subsection{Grid users: nordugrid-authuser objectclass}

Within the ARC information model every authorized Grid user of a 
resource is described by an \emph{authuser} entry in the local tree.
The user entries are used to present user-specific view of
the resource, information such as free CPUs and available disk space 
are shown for every authorized Grid user.
The existence of an {\it nordugrid-authuser} entry implies the 
granted access to the queue of the resource for the corresponding Grid identity.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-authuser-name}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-authuser-name}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& string\\
    Example:& \verb#nordugrid-authuser-name: Lars Jenssen...8#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: 
The Common Name of the authorized user appended by a local unique number.
The Common Name is determined from the Certificate of the user.
This {\it nordugrid-authuser-name}  attribute constitutes to the DN of the user entry.

%DISCUSS do we need this attribute? if not what shall be the DN attribute?

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-authuser-sn}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-authuser-sn}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& LDAP Distinguished Name\\
    Example:& \verb#nordugrid-authuser-sn: /O=Grid/O=NorduGrid/OU=nordugrid.org/CN=Lars Jenssen#\\
    Related xRSL:& none\\
    UI role:& used while searching for available resources of a user\\ 
  \end{tabular}

Description:  The LDAP Subject Name of the authorized Grid user
as specified in his/her Grid credentials.
A Grid user or a client can easily find the resources where he/she is authorized
by issuing an LDAP search with a filter 
of {\it nordugrid-authuser-sn=his/her SN}.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-authuser-freecpus}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-authuser-freecpus}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-authuser-freecpus: 2 4:25 5:180#\\
    Related xRSL:& \verb#count#\\
    UI role:& used in brokering\\ 
  \end{tabular}

Description:
The number of freely available CPUs with their time limits for the specific Grid identity 
is given by this attribute according to the following syntax: \\
{\it ncpus[:min] [ncpus:min] ...} where the pair {\it ncpus:min}  gives the number 
of free CPUs with their time limit in minutes. The time limit information is optional. 
When there are blocked grid jobs in the Grid layer on a cluster this attribute is set 
to zero regardless of available free slots on the cluster.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-authuser-diskspace}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-authuser-diskspace}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&  number showing the amount in MBs\\
    Example:& \verb#nordugrid-authuser-diskspace: 13964#\\
    Related xRSL:& \verb#disk#\\
    UI role:& used in the matchmaking\\
  \end{tabular}

Description: The free disk space available for the session directory of the 
user's Grid job given in MBs.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-authuser-queuelength}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-authuser-queuelength}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number\\
    Example:& \verb#nordugrid-authuser-queuelength: 0#\\
    Related xRSL:& none\\
    UI role:& used in the brokering\\
  \end{tabular}

Description: The number of queuing jobs of a particular user due to the
Grid layer and batch system queue. The attribute takes into account the user's 
jobs accumulated both in the Grid layer and in the LRMS and shows the sum of the
user's "queuing" jobs, gives the "length" of the user's personal queue.

\subsection{Storage Resources: the nordugrid-se objectclass}

The {\it nordugrid-se} objectclass is used to describe storage resources 
within the NorduGrid/ARC model. A storage resource consists of physical data
source (the storage space itself) plus the protocols, policies,
services and interfaces which make the storage space available to the clients.
The attributes of the objectclass are designed to describe all of these layers.

Some of the current attributes are not yet supported in the information system
implementation. Furthermore the following attributes are being discussed
to be added to the schema (see \url{http://bugzilla.nordugrid.org}, bug \#181):

\begin{itemize}
\item se-iospeed: The average IO capability of the storage (MB/s)

\item se-networkspeed: The network capability of the storage (MB/s)

\item se-architecture: The hardware architecture of the Storage \\
Enumeration values: disk, raid of disks, memory, raid of memory,
Tape, hierarchical storage, network storage

\item se-status: The status of the storage service. \\
Enumeration values following glue status info: OK, Warning, Critical, Unknown,
other. Most probable the OK and Critical is enough

\item se-load: a numeric value representing the load on the storage element.
load values from the UNIX top?

\item se-accessprotocol: The protocol supported to access/transform files, e.g. the SRM flavour.

\item se-backupfrequency: The backup service provided by the storage. \\
Enumeration values: never, occasionally, monthly, weekly, nightly.

\end{itemize}

%DISCUSS: the concrete definition and implementation of the above
% attributes

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-name}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-name}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed format string\\
    Example:& \verb#nordugrid-se-name: HEP:hathi.hep.lu.se#\\    
    Related xRSL:& none\\
  \end{tabular}

Description: The globally unique name of the Storage Element
composed as {\it local-name colon  FQDN}.
The {\it nordugrid-se-name} attribute constitutes to the DN of the se entry.

%DISCUSS  unique names, what shall be the SE name?

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-aliasname}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-aliasname}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-se-aliasname: Lund HEP SE#\\
    Related xRSL:& none\\
  \end{tabular}

Description:
A free form text attribute displaying the alias name of the storage resource.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-type}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-type}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{gridftp, SSE, other}#\\    
    Example:& \verb#nordugrid-se-type: gridftp#\\
    Related xRSL:& none\\
  \end{tabular}

Description: The type of the storage element. ARC currently comes with two
class of storage elements, the conventional GridFTP-based and the 
Smart Storage Element (SSE)\cite{smartse}

%DISCUSS what are the SE types???

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-freespace}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-freespace}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&  number showing the amount in MBs\\
    Example:& \verb#nordugrid-se-freespace: 253870#\\
    Related xRSL:& none\\
  \end{tabular}

Description: The total amount of free space available on the SE in MBs.
Not all of this space may be available for every Grid user.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-totalspace}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-totalspace}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& number showing the amount in MBs\\
    Example:& \verb#nordugrid-se-totalspace: 1531381#\\
    Related xRSL:& none\\
  \end{tabular}

Description:
The total capacity of the storage resource displayed in MBs.

%DISCUSS: is the MB a good unit?

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-url}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-url}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& URL\\
    Example:& \verb#nordugrid-se-url: gsiftp://hathi.hep.lu.se:2811/hep#\\
    Related xRSL:& none\\
  \end{tabular}

Description: The URL to contact the Storage Element.
Multivalued attribute, an SE can be accessed via several URLs.

%DISCUUSS: can it be multivalued?, currently it is single-valued,
% several URL to the same storage?
%TODO: the provider hardcodes the "gsiftp" !!!

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-location}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-location}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&  Postal ZIP code with two letter country prefix\\
    Example:& \verb#nordugrid-se-location: SE-22100#\\
    Related xRSL:& none\\
  \end{tabular}

Description: 
The geographical location of the storage resource, preferably
specified as a postal code with a two letter country prefix.
Not yet supported.
%TODO: implement in the providers!!!

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-owner}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-owner}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-se-owner: Danish Center for Grid Computing#\\
    Related xRSL:& none\\
  \end{tabular}

Description: The multivalued attribute is used to display the owner of the resource.
Not yet supported.
%TODO: implement in the providers!!!

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-acl}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-acl}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& fixed form syntax\\
    Example:& \verb#nordugrid-se-acl: VO:ATLAS#\\
    Example:& \verb#nordugrid-se-acl: VO:developers.nordugrid.org#\\
    Related xRSL:& none\\
    UI role:& none\\
  \end{tabular}

Description: The multivalued attribute is used to display authorization rules 
applied on an SE. The attribute value follows a {\it fixed form syntax}.
Current syntax is rather coarse-grained and primitive: a "VO:" prefix followed by a VO name 
means that the given VO is authorized on the SE. 
Note that there are no standards for VO names.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-issuerca}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-issuerca}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& LDAP Distinguished Name\\
    Example:& \verb#nordugrid-se-issuerca: /C=DE/O=GermanGrid/CN=GridKa-CA#\\
    Related xRSL:& none\\
    UI role:& ?\\ 
  \end{tabular}

Description: 
The certificate issuer of the storage resource.
The DN of the CA which issued the host certificate is shown by the attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-issuerca-hash}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-issuerca-hash}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& Hash value of a CA certificate\\
    Example:& \verb#nordugrid-cluster-se-hash: 1f0e8352#\\
    Related xRSL:& none\\
    UI role:& ???\\
  \end{tabular}

Description: The hash value of the certificate of the issuer CA of the SE, 
the hash of the certificate of the CA which issued the host certificate 
used by the SE is shown by the attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-trustedca}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-trustedca}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& LDAP Distinguished Name\\
    Example:& \verb#nordugrid-se-trustedca: /C=DE/O=GermanGrid/CN=GridKa-CA#\\
    Example:& \verb#nordugrid-se-trustedca: /DC=ORG/DC=SEE-GRID/CN=SEE-GRID CA#\\
    Related xRSL:& none\\
    UI role:& ???\\
  \end{tabular}

Description: The DNs trusted by the SE are shown by this
multivalued attribute.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-middleware}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-middleware}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form string representing a software package\\
    Example:& \verb#nordugrid-se-middleware: nordugrid-200501280505cvs#\\
    Example:& \verb#nordugrid-cluster-middleware: globus-2.4.3-15ng#\\
    Related xRSL:& none\\
  \end{tabular}

Description: The multivalued attribute specifies the middleware packages 
installed on the storage resource.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-comment}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-comment}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& free form text\\
    Example:& \verb#nordugrid-se-comment: Dedicated HEP storage#\\
    Related xRSL:& none\\
  \end{tabular}

Description: The free form attribute displays additional information about 
the storage. Sometimes it contains an URL where more information can be read
about the resource.

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-authuser}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-authuser}
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:&  LDAP Distinguished Name\\
    Example:& \verb#/O=Grid/O=NorduGrid/OU=nordugrid.org/CN=Lars Jenssen#\\
    Example:& \verb#/O=Grid/O=NorduGrid/OU=nordugrid.org/CN=Leif Jenssen#\\
    Related xRSL:& none\\
    UI role:& ?\\
  \end{tabular}

Description: The multivalued attribute lists the DNs of the authorized users.

%DISCUSS: what about replacing the attribute by a user-specific entry
% which makes use of the nordugrid-authuser objectclass?

\phantomsection
  \addcontentsline{toc}{subsubsection}{nordugrid-se-accesscontrol}
  \hspace*{0.5cm}
  \begin{shaded}
    \textbf{nordugrid-se-accesscontrol} ***
  \end{shaded}
  \begin{tabular}{lp{10cm}}  
    Attribute value:& \verb#{trivial, gacl, other}#\\
    Example:& \verb#nordugrid-se-accesscontrol: gacl#\\
    Related xRSL:& none\\
    UI role:& ? \\
  \end{tabular}

Description\footnote{not in a stable release yet}: The access control framework provided by the storage
element and can be utilized by the users. Currently ARC implements two
type of access control frameworks, the {\it GACL}~\cite{gacl} and a
trivial one \cite{grid-manager}.

\subsection{Other services: nordugrid-rc objectclass for data catalogues}

Originally, the {\it nordugrid-rc} objectclass was introduced to
describe the Replica Catalogue services. The objectclass was rarely
used and by now it has become obsolete. The {\it nordugrid-rc}
objectclass is to be deprecated and replaced by a more general concept
of Grid service.

For the sake of completeness the attributes of the {\it nordugrid-rc}
are listed without proper description, please refer to the Storage
Resource section or to the Appendix in case more information is
needed.

The list of the {\it nordugrid-rc} attributes: nordugrid-rc-name,
nordugrid-rc-aliasname, nordugrid-rc-baseurl, nordugrid-rc-authuser,
nordugrid-rc-location, nordugrid-rc-owner, nordugrid-rc-issuerca.

\subsection{Mapping to the GLUE Information model}

The GLUE Infomation model, or the "GLUE Schema" represents an
alternative information model\cite{glue1.2} developed
parallel to the ARC schema. The version 1.2 of
the GLUE Schema is widely used in large Grid deployments such as LCG/EGEE or 
the Open Science Grid. Below a mapping between the ARC and the
GLUE information model is presented, in particular the most 
commonly used GLUE 1.2 entities are populated using information stored 
in ARC attributes and objectclasses. Unfortunately,
the GLUE specification\cite{glue1.2} is incomplete, the relation 
between the GLUE entities, the DIT is not fully specified within 
the model. Therefore in our mapping we had to rely on a particular
implememtation, the mapping presented below follows the GLUE
implementation of the LCG/EGEE project.

The GLUE model aims to describe Grid entities such as {\it Site,
Service, Cluster, SubCluster, Host, ComputingElement, Job, StorageElement},
however not all of the above entities are used in production deployments.
The ARC - GLUE-1.2 mapping presented below deals with the most
relevant {\it ComputingElement, Cluster/SubCluster and Site} entities.

\subsubsection{The ComputingElement entity of the LCG Glue model}

The LCG implementation of the Glue-1.2 ComputingElement entity is
populated with information stored in ARC schema: 
attributes from the {\it nordugrid-cluster} and {\it nordugrid-queue} 
entries (objectclass) are used to fill an LCG-Glue ComputingElement.
The ARC schema is much richer than the Glue one, almost every Glue CE attribute
can be filled from ARC info while there are lots of ARC information
which has no place within the Glue-1.2 representation.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEUniqueID} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:&
    nordugrid-cluster-contactstring{\it ?queue=}nordugrid-queue-name\\
  Example:& \verb#GlueCEUniqueID: gsiftp://gridbox.nordugrid.org:2811/jobs?queue=gridqueue#\\
\end{tabular}

Comment: ARC computing elements are defined by cluster contact information
and queue information. Notice the {\it ?queue=} literal  introduced to
extend the {\it nordugrid-cluster-contactstring} with the
{\it nordugrid-queue-name}.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEName} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-name\\
  Example:& \verb#GlueCEName: gridqueue#\\
\end{tabular}

Comment:  -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInformationServiceURL} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& an infrastructure specific LDAP URL\\
  Example:& \verb#GlueCEInformationServiceURL: ldap://arc-bdii.cern.ch:2170/o=grid#\\
\end{tabular}

Comment: The attribute should contain the LDAP URL of the BDII service
providing the ARC-Glue mapping.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoLRMSType} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-lrms-type\\
  Example:& \verb#GlueCE: SGE#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoLRMSVersion} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-lrms-version\\
  Example:& \verb#GlueCEInfoLRMSVersion: 5.1#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoGRAMVersion} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& this attribute is irrelevant for ARC\\   
\end{tabular}

Comment: ARC does not use GRAM for job submission.
GlueCEInfoGRAMVersion Glue attribute is too Globus specific.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoHostname} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-name\\
  Example:& \verb#GlueCEInfoHostname: gridbox.nordugrid.org#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoGatekeeperPort} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& this attribute is irrelevant for ARC\\   
\end{tabular}

Comment: The Glue attribute is too much Globus specific, again. 
ARC runs no Globus gatekeeper.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoJobmanager} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& this attribute is irrelevant for ARC\\   
\end{tabular}

Comment: Too much Globus specific, again. No Globus jobmanager on ARC.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoContactString} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-contactstring{\it ?queue=}nordugrid-queue-name\\
  Example:& \verb#GlueCEInfoContactString: gsiftp://gridbox.nordugrid.org:2811/jobs?queue=gridqueue#\\
\end{tabular}

Comment: The attribute carries the same value as the GlueCEUniqueID: 
constructed from the {\it nordugrid-cluster-contactstring} and 
the {\it nordugrid-queue-name}. In order to access a CE in ARC you need to 
specify the queue name together with the gsiftp:// cluster contact string.
GlueCEInfoContactString attribute obsoletes the GlueCEInfo.HostName, 
GlueCEInfo.GatekeeperPort and GlueCEInfo.Jobmanager attributes.
The attribute name is rather misleading since it is not the contact string for
the Information service but the contact info for the job submission service
of the CE.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoTotalCPUs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& if defined use the {\it nordugrid-queue-totalcpus} otherwise the 
    {\it nordugrid-cluster-totalcpus}\\
  Example:& \verb#GlueCEInfoTotalCPUs: 66#\\
\end{tabular}

Comment: This Glue attribute is DEPRECATED in Glue version 1.2, better 
not to set.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoApplicationDir} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& {\it ApplicationDir} does not exist on ARC computing resource\\   
\end{tabular}

Comment: none-existing attributes shouldn't be set.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEInfoDataDir} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& {\it DataDir} does not exist on ARC computing resource\\    
\end{tabular}

Comment: none-existing attributes shouldn't be set.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEDefaultSE} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& This concept does not exist on ARC computing resource\\    
\end{tabular}

Comment:  none-existing attributes shouldn't be set.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEStateStatus} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& perform mapping on the {\it nordugrid-queue-status} values\\
  Example:& \verb#GlueCEtateStatus:production #\\
\end{tabular}

Comment: Use the following mapping of the {\it nordugrid-queue-status} values:
active (ARC) = Production (Glue); inactive (ARC) = Closed (Glue).

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEStateRunningjobs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-running\\
  Example:& \verb#GlueCEStateRunningjobs: 12#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEStateWaitingJobs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& sum of the {\it nordugrid-queue-gridqueued,
    nordugrid-queue-localqueued, nordugrid-queue-prelrmsqueued} \\
  Example:& \verb#GlueCEStateWaitingJobs: 12#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEStateTotaljobs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& sum of the {\it nordugrid-queue-running,
    nordugrid-queue-gridqueued, nordugrid-queue-localqueued,
    nordugrid-queue-prelrmsqueued}\\
  Example:& \verb#GlueCE: 54#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEStateEstimatedResponseTime} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& ARC does not use this attribute and recommends to set it to 
    a fix value\\
  Example:& \verb#GlueCEStateEstimatedResponseTime: 1000#\\
\end{tabular}

Comment: Without advanced reservation it is almost impossible to
come up with a reasonable value for queue "EStateEstimatedResponseTime".
NorduGrid does not believe in the usefulness of this attribute. 
Best is to set it for the same value for all CEs and let the broker decide 
using other attributes.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEStateWorstResponseTime} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& ARC does not use this attribute and recommends to set it to 
    a fix value\\
  Example:& \verb#GlueCEStateWorstResponseTime: 2000#\\
\end{tabular}

Comment: NorduGrid does not believe in the usefulness of this attribute. 
Best is to set it for the same value for all CEs and let the broker decide 
using other attributes.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEStateFreeJobSlots} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& calculate it from the {\it nordugrid-queue-totalcpus/
    nordugrid-cluster-totalcpus} and the {\it nordugrid-queue-running}\\
  Example:& \verb#GlueCEStateFreeJobSlots: 3#\\
\end{tabular}

Comment: Substract the {\it nordugrid-queue-running} from the 
{\it nordugrid-queue-totalcpus} or from the 
{\it nordugrid-cluster-totalcpus} in case the {\it nordugrid-queue-totalcpus}
attribute is not set.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEStateFreeCPUs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:&  - same as the GlueCEState.FreeJobSlots\\   
\end{tabular}

Comment: This attribute is DEPRECATED in glue-1.2, better not to 
set.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEPolicyMaxWallTime} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& use the value of nordugrid-queue-maxcputime \\
  Example:& \verb#GlueCEPolicyMaxWallTime: 120#\\
\end{tabular}

Comment: queue-maxwalltime attribute currently does not exist in ARC schema, 
the value of the queue-maxcputime can be used. nordugrid-queue-maxwalltime
will be added to the ARC schema.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEPolicyMaxCPUTime} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-maxcputime\\
  Example:& \verb#GlueCEPolicyMaxCPUTime: 160#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEPolicyMaxTotalJobs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-maxqueuable\\
  Example:& \verb#GlueCEPolicyMaxTotalJobs: 120#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEPolicyMaxRunningJobs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-maxrunning\\
  Example:& \verb#GlueCEPolicyMaxRunningJobs: 16#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEPolicyPriority} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& does not exist in ARC, set it to fix value\\
  Example:& \verb#GlueCE: 1#\\
\end{tabular}

Comment: Set it to 1 for every CE, it is a rather useless internal LRMS value.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEPolicyAssignedJobSlots} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& use one of the {\it nordugrid-queue-totalcpus, 
    nordugrid-queue-maxrunning, nordugrid-cluster-totalcpus} attributes\\
  Example:& \verb#GlueCEPolicyAssignedJobSlots: 30#\\
\end{tabular}

Comment: One of the above listed three nordugrid attributes in the 
given priority order can be used to set this strange Glue attribute: 
there is large overlap with the GlueCEPolicyMaxRunningjobs attribute, 
one of them should be enough.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueCEAccessControlBaseRule} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-acl\\
  Example:& \verb#GlueCEAccessControlBaseRule: VO:atlas#\\
\end{tabular}

Comment: The {\it nordugrid-cluster-acl} attribute is introduced
in the 0.6 ARC release.

\subsubsection{The Cluster and SubCluster entity of the LCG Glue model}

Originally, the Glue model offers the {\it Cluster, Subcluster and Host} 
entities for describing physical (hardware) properties of computing services.
Heterogeneity was supposed to be addressed by the {\it SubCluster} and
{\it Host} entities: clusters can be heterogeneous while 
subclusters are assumed to be homogeneous. The {\it Cluster} entity was 
designed to link the hardware description (SubCluster, Host) to the 
computing service view of the resource: the 
ComputingElement entity.

Unfortunately, in the actual LCG/EGEE deployment the Glue Cluster, SubCluster, 
Host entities got mixed up. The Cluster entry is rather empty, 
it carries almost zero information. Clusters have only one SubCluster
subentry (no heterogeneity is treated), SubCluster entries got merged with the 
Host entities, thus SubClusters carry all the Host properties. 
Furthermore, due to the flat LDAP tree the ComputingElement entries
are not linked to the Cluster/SubCluster entries.

Since the Glue model comes with no DIT, it does not specify the relation of 
particular entities. In the LCG/EGEE deployment it lead to an almost completely
flat structure where ComputingElement and Cluster entities (and even Site entities !)
are placed on the same level of the LDAP tree.

The ARC infosys representation has cluster and queue entries.
More importantly, it contains the notion of cluster- and queue-level attributes 
and a cluster-homogeneity flag. Populating the Glue-LCG Cluster/SubCluster 
entities using the ARC schema is rather strateforward since ARC offers cluster 
and queue-level view of computing resources to treat heterogeneity 
(see Section \ref{section:cluster-queue}).
Queue-level view is used for heterogeneous subset of computing resurces.
For homogeneus clusters (where the {\it nordugrid-cluster-homogeneity=true} 
is set) create only one Glue SubCluster and use the cluster-level attributes from
the ARC information model. For inhmogeneous clusters ( 
{\it nordugrid-cluster-homogeneity=false}) create a Glue SubCluster for every 
queue and populate the Glue SubCluster with the corresponding 
ARC queue-level attributes.

Final remark: as of writing, on the whole LCG/EGEE Grid 
there is not a single Cluster whith more than one SubCluster.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueClusterUniqueID}
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-name\\
  Example:& \verb#GlueClusterUniqueID: gridbox.nordugrid.org#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueClusterName} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-aliasname\\
  Example:& \verb#GlueClusterName: Gridbox cluster#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueClusterTmpDir} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& does not exist in ARC\\ 
\end{tabular}

Comment: The location of a temporary directory on the frontend is an internal 
cluster property, shouldn't be directly advertised on the Grid. 
Don't set this attribute.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueClusterWNTmpDir} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& does not exist in ARC\\
\end{tabular}

Comment: The location of a temporary directory on the worker node is an internal 
cluster property, shouldn't be directly advertised on the Grid. 
Don't set this attribute.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSubClusterUniqueID} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& {\it nordugrid-cluster-name/nordugrid-queue-name}\\
  Example:& \verb#GlueSubClusterUniqueID: gridbox.nordugrid.org/gridqueue#\\
\end{tabular}

Comment: notice the "Glue SubCluster" = "ARC queue" assumption.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSubClusterName} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& {\it nordugrid-cluster-aliasname/nordugrid-queue-name}\\
  Example:& \verb#GlueSubClusterName:Gridbox cluster/gridqueue #\\
\end{tabular}

Comment: notice the "Glue SubCluster" = "ARC queue" assumption.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSubClusterPhysicalCPUs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-totalcpus or nordugrid-cluster-totalcpus\\ 
\end{tabular}

Comment: The queue or cluster level attribute should be used depending on the 
cluster-homogeneity flag.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSubClusterLogicalCPUs} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& not available in ARC\\ 
\end{tabular}

Comment: No need to set this attribute.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSubClusterLocationName/Version/Path} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& does not exist in ARC\\
  Example:& \verb#GlueSubCluster:#\\
\end{tabular}

Comment: The name/version/path of an installed software is an internal cluster 
property, shouldn't be advertised on the grid. The 'Location' Glue entity 
which defines the above three attributes is a bad idea. Don't set these attributes.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostOperatingSystemName/OperatingSystemRelease/OperatingSystemVersion} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:&  nordugrid-queue-opsys or nordugrid-cluster-opsys\\
  Example:& \verb#GlueHostOperatingSystemName: Redhat#\\
  Example:& \verb#GlueHostOperatingVersion: 7.3#\\
\end{tabular}

Comment: The ARC schema has only two multivalued (!) attributes for describing the 
operatingsystem. Use the {\it nordugrid-queue-opsys} or the 
{\it nordugrid-cluster-opsys} multivalued (!) attribute,
depending on the cluster-homogeneity flag, to set some of the above Glue attributes.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostProcessorModel/ProcessorVersion/ProcessorVendor/ProcessorClockSpeed/
  ProcessorInstructionSet/ProcessorOtherDescription} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-nodecpu or nordugrid-cluster-nodecpu\\  
\end{tabular}

Comment: ARC has only two attributes for describing the node CPU, 
the {\it nordugrid-queue-nodecpu} or the {\it nordugrid-cluster-nodecpu}. 
Use one of them depending on the cluster-homogeneity flag.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostRAMSize} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-nodememory or nordugrid-cluster-nodememory\\
  Example:& \verb#GlueHostRAMSize: 1024#\\
\end{tabular}

Comment: The queue or cluster level attribute should be used depending on the 
cluster-homogeneity flag. Notice the different semantics of the 
LCG-Glue and ARC memory attributes:
ARC memory is 'amount of memory guaranteed for the application' while 
LCG-Glue memory  is 'amount of RAM'.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostVirtualSize} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& not available in ARC\\ 
\end{tabular}

Comment: Not available in the ARC model, don't set this attribute.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostNetworkAdapterOutboundIP} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-nodeaccess\\
  Example:& \verb#GlueHostNetworkAdapterOutboundIP: FALSE#\\
\end{tabular}

Comment: Set {\it TRUE} if 
{\it nordugrid-cluster-nodeaccess=outbound}, otherwise {\it FALSE}.
There is no queue-level nodeaccess property in ARC, 
it is defined on the cluster-level.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostNetworkAdapterInboundIP} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-nodeaccess\\
  Example:& \verb#GlueHostNetworkAdapterInboundIP: TRUE#\\
\end{tabular}

Comment: Set {\it TRUE} if {\it nordugrid-cluster-nodeaccess=inbound}, 
otherwise {\it FALSE}. There is no queue-level nodeaccess property in ARC, 
it is defined on the cluster-level.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostArchitecturePlatformType} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-architecture or nordugrid-cluster-architecture\\
  Example:& \verb#GlueHostArchitecturePlatformType: x86_64#\\
\end{tabular}

Comment: The queue or cluster level attribute should be used depending on the 
cluster-homogeneity flag.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostBenchmarkSI00/BenchmarkSF00} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-queue-benchmark or nordugrid-cluster-benchmark\\
  Example:& \verb#GlueHostBenchmarkSI00: 100#\\
\end{tabular}

Comment: Take the benchmark value  from the 
{\it nordugrid-queue-benchmark: SPECINT2000 @ 111}
or use the nordugrid-cluster-benchmark for homogeneous clusters.
The way Glue addresses (hardcodes) benchmark names are really bad.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueHostApplicationSoftwareRunTimeEnvironment}
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-runtimeenvironment\\
  Example:& \verb#GlueHost: APPS/CHEM/DALTON-1.2.1-1.0#\\
\end{tabular}

Comment: Runtimeenvironments are defined on cluster-level in ARC.

\subsubsection{The Site entity of the LCG Glue model}

The LCG-Glue Site entity is meant to be an administrative and structuring
entry in the information system representation. Its main purpose is to collect
resources belonging to the same "site" and their administrative description 
in a single entry. Unfortunately, in the LCG Glue deployment the Site entry is
NOT used in the LDAP tree to group the resources/entries 
belonging to the same site. The introduction of the 
Site concept in LCG-Glue has a consequence that
Glue-LCG resource entries (Cluster, ComputingElement, StorageElement) 
have no administrative attributes.

The ARC schema follows a different approach.
The ARC infosys representation has no Site object, resources (clusters, SEs) 
are not grouped on a site level, administrative information 
(owner, location, comments, etc) are stored within the resource entries 
themselves.

In order to create an LCG-Glue Site entry from the ARC information 
tree the best is to follow  the "one resource = one site" approach and take the 
administrative info present e.g. in the cluster ARC entry and use that to 
create a Glue Site entry.
In the proposed mapping below every ARC cluster corresponds to a Glue-LCG site.
Optionaly, resources having the same {\it owner} attribute can be
grouped under a common Site entry. Unfortunately there is no easy 
and reliable method to find out which resources would belong to the same 
administrative unit.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteUniqueID} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-name\\
  Example:& \verb#GlueSiteUniqueID: gridbox.nordugrid.org #\\
\end{tabular}

Comment: The proposed mapping expresss the "one cluster = one site" mapping
which provides real uniqueness. btw, the currently used  LCG Site names such as
"Budapest" don't guarantee uniqueness at all.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteName} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-aliasname\\
  Example:& \verb#GlueSiteName: Gridbox cluster#\\
\end{tabular}

Comment: Use the cluster alias or let site admins manually fill the
site names.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteDescription} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-comment\\
  Example:& \verb#GlueSiteDescription: General purpose Grid box#\\
\end{tabular}

Comment: -

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteUserSupportContact/SysAdminContact/SecurityContact} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& {\it mailto: nordugrid-cluster-support}\\
  Example:& \verb#GlueSiteSysAdminContact: mailto: contact@gridbox.nordugrid.org#\\
\end{tabular}

Comment: ARC schema has only one "contact" email address, that can be used 
to set all the three Glue contact attributes. Notice the {\it mailto:}
Glue prefix in front of the {\it nordugrid-cluster-support}.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteLocation} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-location\\  
\end{tabular}

Comment: Observe the different semantics! ARC location is a postal ZIP code while 
LCG-location is 'city,state,country'

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteLatitude/Longitude} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& these don't exist in ARC\\  
\end{tabular}

Comment: This information is not available in ARC schema.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteWeb} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& does not exist in ARC\\ 
\end{tabular}

Comment: This info (website) sometimes is available in the 
{\it nordugrid-cluster-comment} attribute.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteSponsor} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:& nordugrid-cluster-owner\\
  Example:& \verb#GlueSiteSponsor: SweGrid#\\
\end{tabular}

Comment: Note that the {\it nordugrid-cluster-owner} attribute 
is multivalued.

\hspace*{0.5cm}
\begin{shaded}
 \textbf{GlueSiteOtherInfo} 
\end{shaded}
\begin{tabular}{lp{10cm}}  
  ARC mapping:&  add the fixed text shown inthe example\\
  Example:& \verb#GlueSiteOtherInfo: This site is running on the ARC middleware#\\
\end{tabular}

Comment: Let us have this small marketing.
Btw, many LCG deployments use the GlueSiteOtherInfo to advertise their webpage
or a hostname, though there are attributes for those.

\section{Registration Processes, Index Services: Topology}

The individual local information trees need to be connected and organized 
into some sort of topological structure in order to create
a coherent information system. NorduGrid/ARC utilizes registration 
processes and Index Services to build a distributed information system 
out of the individual local information trees.

Connecting information sources together is usually referred to as the resource 
information aggregation. ARC implements a minimalistic aggregation process:
An ARC Index Service collects only the contact information of the 
information resources and no information is gathered or cached from the 
local information trees.

The local Information Trees and Index Services are linked together via the 
{\it registration processes}. During a registration process the registrant 
(lower level) sends its registration packet to an Index service.
A registration packet contains information about the host (the registrant) 
initiating the registration process and about the information service running 
on the registrant (either a local tree or an index service): the registration 
message is basically the LDAP contact URL of the information service 
running on the registrant
(additionally some timing parameter is also transfered indicating how long the 
registration information is expected to be kept in the Index Service).
The target Index Service can filter out registration processes coming from 
registrants, unfortunately the filtering capability of the current Index Service
is rather limited, it is based only on the FQDN and NOT on the
LDAP URL. Registrations are sent periodically to the target Index Services, 
thus the registration mechanism follows a periodic push model.
Technically, the registrations are implemented as periodic {\it ldapadd} operations.

{\it Index Services} are used to maintain dynamic
lists of available resources, containing the LDAP contact information of
the registrants. A registrant can either be a Local Information Tree or another 
Index Service. The content of the Index Service, that is the information on
the registrants, is periodically purged, this way maintaining a dynamic 
registrant list. The Index Services are implemented as a special purpose 
LDAP databases, the registration entries are stored as pseudo-LDAP entries 
in the Globus GIIS LDAP back-end. Although the registration 
information is presented in a valid LDIF format the registration entries
do not constitute a legitimate LDAP tree, therefore they are referred to as the
pseudo-LDAP registration entries of the Index Service.
The periodic purging of the 
registrant entries are done by NOT removing the obsoleted entry BUT
setting the {\it Mds-Reg-status: PURGED} attribute. 
ARC Index Services are purely used to maintain a set of pseudo-LDAP registration 
entries shown in Figures \ref{fig:regentry1},\ref{fig:regentry2}. No any other information
is stored in the ARC Indices.
It is planned to replace the currently abused GIIS LDAP back-end implementation 
of the ARC Index Service by something more suitable, the simplistic functionality 
of maintaining a list of registrant tables could be trivially done e.g. 
in native LDAP.

The pseudo-LDAP registration entries stored in an Index Service
running on {\it host, port} under the LDAP {\it basedn}  can be 
obtained by a \textbf{specially crafted LDAP query} executed against the Index Service:

\begin{figure}
\begin{verbatim}
ldapsearch -x
	   -h host 
	   -p port 	   
	   -b basedn 
	   -scope base
	   giisregistrationstatus 
\end{verbatim}
\caption{\label{fig:giisregistrationstatus} The special LDAP query to obtain the content
of an Index Service.}
\end{figure}

That is an anonymous (-x) ldapsearch against the LDAP server 
of the Index Service with search scope {\it base}  AND asking for the
{\it giisregistrationstatus} special attribute provides all the pseudo-LDAP 
registration entries stored in the Index Service.
Examples for Index Service queries can be found in the the Appendix
\ref{appendix:examples}.

A valid pseudo-LDAP registration entry, stored in an Index Service and obtained 
by the above described special LDAP query, 
is shown on Figure \ref{fig:regentry1}. The LDAP contact URL of the
Local Information Tree running on the {\it grid.tsl.uu.se} machine  is given by the 
attributes {\it Mds-Service-hn, Mds-Service-port, Mds-Service-Ldap-suffix}.
The {\it dn: nordugrid-cluster-name=grid.tsl.uu.se, Mds-Vo-name=Sweden,o=grid} 
of the registration entry indicates that the registration corresponds to 
computing resource registering to an Index Service called 
{\it  Mds-Vo-name=Sweden,o=grid}.
The {\it Mds-Reg-status} attribute shows the validity of the registration.
Figure \ref{fig:regentry2} shows another pseudo-LDAP registration entry 
corresponding to a PURGED state registration of an (lower level) Index Service 
to another (higher level) Index Service.

\begin{figure}
\begin{verbatim}
#grid.tsl.uu.se, Sweden, grid
dn: nordugrid-cluster-name=grid.tsl.uu.se, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: grid.tsl.uu.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: nordugrid-cluster-name=grid.tsl.uu.se, 
                         Mds-Vo-name=local, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 45
Mds-Service-Ldap-cachettl: 15
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: VALID
\end{verbatim}
\caption{\label{fig:regentry1} A pseudo-LDAP registration entry obtained from the Sweden Index Service 
describing the valid registration of a computing resource}
\end{figure}

\begin{figure}
\begin{verbatim}
# SweGrid, Sweden, grid
dn: Mds-Vo-name=SweGrid, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: hagrid.it.uu.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: Mds-Vo-name=SweGrid, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 120
Mds-Service-Ldap-cachettl: 30
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: PURGED
\end{verbatim}
\caption{\label{fig:regentry2} A pseudo-LDAP registration entry obtained 
from the Sweden Index Service 
describing the PURGED registration of the SWEGRID Index Service:
The SWEGRID Index Service is running on the {\it hagrid.it.uu.se} machine on port
{\it 2135} and with LDAP base {\it Mds-Vo-name=SweGrid, o=grid}.}
\end{figure}

The Local Information Trees and the Index Services of the NorduGrid/ARC Grid are
organized into a multi-level tree hierarchy. The Local Information Trees describing the
actual Storage or Computing resources represent the lowest level of the tree-like
topology. Resources are registering to first level Index Services which are registering to
Second level Services, so on so forth. The registration chain ends at 
the Top Level Indices which represent the root of the tree hierarchy. 
The structure is built from bottom to top: always the lower level registers to 
the higher one. The tree-like hierarchical structure is motivated by  the 
natural geographical organization where resources belonging to the same 
region register under a region index, region indices are registering to
the appropriate country index while country indices are grouped
together and register to the top level Grid index services. 
In order to avoid any single point of failure, NorduGrid/ARC operates a
multi-rooted tree with several top-level Indices (Table \ref{Table:topindices}
lists the LDAP contact URL of the Top Level Index Services).
Figure \ref{fig:index_tree} shows simplified schematic view of the multi-rooted 
tree topology of  ARC-connected Resources and Index Services.
Besides the geographical structuring there are some Index Services which 
group resources by specific application area or organization. These 
application/organization Index Services either link themselves to a country Index or 
register directly to a Top level Index.

\begin{figure}[hb]
\centering
{ 
{\scalebox{0.4}{\includegraphics{tree_hierarchy.png}}}
}
\caption{\label{fig:index_tree} Resources (hosts) and Index Services are
linked via the registration process creating a multi-rooted tree topology}
\end{figure}

%TODO mention the possibility of alternative topologies, the extension of registration 
% process

\begin{table}
\centering
\begin{tabular}{|l|l|l|}
\hline
host &
port&
LDAP baseDN \\
\hline
\hline
index1.nordugrid.org &
2135 &
mds-vo-name=nordugrid,o=grid \\
\hline
index2.nordugrid.org &
2135 &
mds-vo-name=nordugrid,o=grid \\
\hline
index3.nordugrid.org &
2135 &
mds-vo-name=nordugrid,o=grid \\
\hline
index4.nordugrid.org &
2135 &
mds-vo-name=nordugrid,o=grid \\
\hline
\end{tabular}
\caption{\label{Table:topindices} LDAP URL of the TOP Level ARC Index Services}
\end{table}

\subsection{Resource discovery}

Resource discovery is the process when clients walk through the 
Index Services organized in a tree topology (see Fig.\ref{fig:index_tree})
and collect  LDAP contact URLs of the Computing and Storage resources.
The discovery process usually starts at the top of the tree by querying some of the 
Top Level Index Services(Table \ref{Table:topindices} lists the Top Indices). 
A Top Level Index Service is queried by a specially 
crafted ldapsearch (Figure \ref{fig:giisregistrationstatus}) which results 
all the registration entries stored in the Index Service. 
Index Services contain registration information of both resources and 
other Index Services. In the second step the client 
has to separate the entries corresponding to Index Services from entries
describing registrations of actual resources (Local Trees). 
The Figures\ref{fig:regentry1},\ref{fig:regentry2} show a registration table of
a Computing resource and an Index Service, respectively. 
The {\it Ldap-suffix} of a resource always contains the special string 
{\it "mds-vo-name=local,o=grid"} referring to the fact that the resource runs
a Local Information Tree. Then, the client contact the the newly discovered Index
Services and obtains the registration tables. The tables are separated into 
Indices and Resources, again. The process is repeated until all the Index Services 
are queried and the full list of LDAP Contact URL of Resources are collected.

Once the client has collected a list of Resource LDAP Contact URLs 
from the Index Services,
the second phase of the information collection begins: 
The client directly contacts every Resource and initiates an LDAP query 
against the Local Information Tree.  
This is the real information gathering process in contrast to the first 
phase in which only the LDAP URLS were collected. Remember, 
unlike other systems (Globus GIIS, GT4 aggregator, R-GMA, LCG-BDII) 
ARC has no service  which caches or aggregates
Resource specific information on a higher level, 
ARC Index Services are not used to store local
information, Indices maintain only LDAP URLs.

\newpage
\appendix
\section{Clients of the ARC information system}

The entire content of the Information system including both the 
local trees and the Index Services are presented via an LDAP interface.
LDAP is a very well supported protocol, therefore it is very easy to 
construct clients making use of the ARC information system.

The Grid Monitor\cite{monitor} is a very powerful graphical interface 
to the Information System written in PHP. The monitor runs 
on a web-server and provides a browsable graphical representation of the 
information content of the Grid.

The matured command line client (UI)\cite{ui} or the 
Graphical UIs and  ARClib being developed  provide another examples
of straightforward interfaceability to the LDAP-based Information System.

%TODO ASK Oxana for the monitor - ldap attributes mapping
%reference to the ARCLIB/ new GUIS

\newpage

\section{LDAP Examples \label{appendix:examples}}

\begin{itemize}

\item
LDAP query against a local tree with a filter for nordugrid-cluster objectclass:

\tiny
\begin{verbatim}
ldapsearch -h bambi.hep.lu.se -p 2135 -x -b 'mds-vo-name=local,o=grid' 'objectclass=nordugrid-cluster'
version: 2

#
# filter: objectclass=nordugrid-cluster
# requesting: ALL
#

# bambi.hep.lu.se, local, grid
dn: nordugrid-cluster-name=bambi.hep.lu.se,Mds-Vo-name=local,o=grid
objectClass: Mds
objectClass: nordugrid-cluster
nordugrid-cluster-name: bambi.hep.lu.se
nordugrid-cluster-aliasname: Bambi Test Cluster
nordugrid-cluster-owner: EHEP
nordugrid-cluster-location: SE-221 00
nordugrid-cluster-issuerca: /O=Grid/O=NorduGrid/CN=NorduGrid Certification Aut
 hority
nordugrid-cluster-contactstring: gsiftp://bambi.hep.lu.se:2811/jobs
nordugrid-cluster-support: grid.support@mysite.org
nordugrid-cluster-support: grid.support@myproject.org
nordugrid-cluster-lrms-type: fork
nordugrid-cluster-lrms-config: single job per processor
nordugrid-cluster-architecture: i686
nordugrid-cluster-opsys: Mandrake-10.0
nordugrid-cluster-opsys: glibc-2.3.3-10mdk
nordugrid-cluster-benchmark: SPECINT2000 @ 222
nordugrid-cluster-benchmark: SPECFP2000 @ 333
nordugrid-cluster-homogeneity: True
nordugrid-cluster-nodecpu: Intel(R) Pentium(R) 4 CPU 3.00GHz @ 2993.100 MHz
nordugrid-cluster-nodeaccess: inbound
nordugrid-cluster-nodeaccess: outbound
nordugrid-cluster-totalcpus: 2
nordugrid-cluster-usedcpus: 0
nordugrid-cluster-cpudistribution: 2cpu:1
nordugrid-cluster-queuedjobs: 0
nordugrid-cluster-totaljobs: 0
nordugrid-cluster-sessiondir-free: 13958
nordugrid-cluster-sessiondir-total: 26403
nordugrid-cluster-cache-free: 9536
nordugrid-cluster-cache-total: 9536
nordugrid-cluster-middleware: nordugrid-0.5.20
nordugrid-cluster-middleware: globus-2.4.3-12ng
nordugrid-cluster-middleware: my grid software
nordugrid-cluster-runtimeenvironment: tt
Mds-validfrom: 20050307103026Z
Mds-validto: 20050307103029Z

# search result
search: 2
result: 0 Success

# numResponses: 2
# numEntries: 1
\end{verbatim}
\normalsize

\item Query for active Grid jobs stored in the local tree describing a computing resource:

\tiny
\begin{verbatim}
ldapsearch -h quark.hep.lu.se -p 2135 -x -b 'mds-vo-name=local,o=grid' 'objectclass=nordugrid-job'
version: 2

#
# filter: objectclass=nordugrid-job
# requesting: ALL
#

# gsiftp://quark.hep.lu.se:2811/jobs/131601109950874935622127, jobs, pc, quar
 k.hep.lu.se, local, grid
dn: nordugrid-job-globalid=gsiftp://quark.hep.lu.se:2811/jobs/1316011099508749
 35622127, nordugrid-info-group-name=jobs, nordugrid-queue-name=pc,nordugrid-c
 luster-name=quark.hep.lu.se,Mds-Vo-name=local,o=grid
objectClass: Mds
objectClass: nordugrid-job
nordugrid-job-globalid: gsiftp://quark.hep.lu.se:2811/jobs/1316011099508749356
 22127
nordugrid-job-globalowner: /O=Grid/O=NorduGrid/OU=cmm.ki.se/CN=Roxana Merino
nordugrid-job-jobname: TFBS2
nordugrid-job-submissiontime: 20050304154114Z
nordugrid-job-execcluster: quark.hep.lu.se
nordugrid-job-execqueue: pc
nordugrid-job-cpucount: 1
nordugrid-job-sessiondirerasetime: 20050305154232Z
nordugrid-job-stdout: tfbs.out
nordugrid-job-stderr: tfbs.out
nordugrid-job-gmlog: gmlog
nordugrid-job-runtimeenvironment: BIO-GEIJER-0.0.2
nordugrid-job-submissionui: 217.208.119.237:18236;10.0.0.1
nordugrid-job-clientsoftware: nordugrid-0.4.4
nordugrid-job-proxyexpirationtime: 20050304212024Z
nordugrid-job-status: DELETED
nordugrid-job-reqcputime: 2
nordugrid-job-executionnodes: node2
Mds-validfrom: 20050307110037Z
Mds-validto: 20050307110107Z

# gsiftp://quark.hep.lu.se:2811/jobs/5381111006551269125810, jobs, pclong, qu
 ark.hep.lu.se, local, grid
dn: nordugrid-job-globalid=gsiftp://quark.hep.lu.se:2811/jobs/5381111006551269
 125810, nordugrid-info-group-name=jobs, nordugrid-queue-name=pclong,nordugrid
 -cluster-name=quark.hep.lu.se,Mds-Vo-name=local,o=grid
objectClass: Mds
objectClass: nordugrid-job
nordugrid-job-globalid: gsiftp://quark.hep.lu.se:2811/jobs/5381111006551269125
 810
nordugrid-job-globalowner: /O=Grid/O=NorduGrid/OU=fys.ku.dk/CN=Brian Moller An
 dersen
nordugrid-job-jobname: selfconsistentMDSSAFNAF100slabonemu
nordugrid-job-submissiontime: 20050305233152Z
nordugrid-job-execcluster: quark.hep.lu.se
nordugrid-job-execqueue: pclong
nordugrid-job-cpucount: 1
nordugrid-job-stdout: output
nordugrid-job-stderr: output
nordugrid-job-submissionui: 130.225.102.149:33853;johansen.fys.ku.dk
nordugrid-job-clientsoftware: nordugrid-0.3.39
nordugrid-job-proxyexpirationtime: 20050306100826Z
nordugrid-job-status: INLRMS: R
nordugrid-job-usedmem: 30736
nordugrid-job-usedwalltime: 2127
nordugrid-job-usedcputime: 2125
nordugrid-job-reqcputime: 10080
nordugrid-job-executionnodes: node1/0
nordugrid-job-lrmscomment: Job started on Sun Mar 06 at 00:32
Mds-validfrom: 20050307110038Z
Mds-validto: 20050307110108Z

# gsiftp://quark.hep.lu.se:2811/jobs/52301110065467612188831, jobs, pclong, q
 uark.hep.lu.se, local, grid
dn: nordugrid-job-globalid=gsiftp://quark.hep.lu.se:2811/jobs/5230111006546761
 2188831, nordugrid-info-group-name=jobs, nordugrid-queue-name=pclong,nordugri
 d-cluster-name=quark.hep.lu.se,Mds-Vo-name=local,o=grid
objectClass: Mds
objectClass: nordugrid-job
nordugrid-job-globalid: gsiftp://quark.hep.lu.se:2811/jobs/5230111006546761218
 8831
nordugrid-job-globalowner: /O=Grid/O=NorduGrid/OU=fys.ku.dk/CN=Brian Moller An
 dersen
nordugrid-job-jobname: selfconsistentMDSSAFNAF100slabonemu
nordugrid-job-submissiontime: 20050305233107Z
nordugrid-job-execcluster: quark.hep.lu.se
nordugrid-job-execqueue: pclong
nordugrid-job-cpucount: 1
nordugrid-job-stdout: output
nordugrid-job-stderr: output
nordugrid-job-submissionui: 130.225.102.149:33846;johansen.fys.ku.dk
nordugrid-job-clientsoftware: nordugrid-0.3.39
nordugrid-job-proxyexpirationtime: 20050306100826Z
nordugrid-job-status: INLRMS: R
nordugrid-job-usedmem: 30736
nordugrid-job-usedwalltime: 2126
nordugrid-job-usedcputime: 2126
nordugrid-job-reqcputime: 10080
nordugrid-job-executionnodes: node3/0
nordugrid-job-lrmscomment: Job started on Sun Mar 06 at 00:32
Mds-validfrom: 20050307110038Z
Mds-validto: 20050307110108Z

# gsiftp://quark.hep.lu.se:2811/jobs/49831110065360908977417, jobs, pclong, q
 uark.hep.lu.se, local, grid
dn: nordugrid-job-globalid=gsiftp://quark.hep.lu.se:2811/jobs/4983111006536090
 8977417, nordugrid-info-group-name=jobs, nordugrid-queue-name=pclong,nordugri
 d-cluster-name=quark.hep.lu.se,Mds-Vo-name=local,o=grid
objectClass: Mds
objectClass: nordugrid-job
nordugrid-job-globalid: gsiftp://quark.hep.lu.se:2811/jobs/4983111006536090897
 7417
nordugrid-job-globalowner: /O=Grid/O=NorduGrid/OU=fys.ku.dk/CN=Brian Moller An
 dersen
nordugrid-job-jobname: selfconsistentMDSSAFNAF100slabonemu
nordugrid-job-submissiontime: 20050305232920Z
nordugrid-job-execcluster: quark.hep.lu.se
nordugrid-job-execqueue: pclong
nordugrid-job-cpucount: 1
nordugrid-job-stdout: output
nordugrid-job-stderr: output
nordugrid-job-submissionui: 130.225.102.149:33836;johansen.fys.ku.dk
nordugrid-job-clientsoftware: nordugrid-0.3.39
nordugrid-job-proxyexpirationtime: 20050306100826Z
nordugrid-job-status: INLRMS: R
nordugrid-job-usedmem: 30736
nordugrid-job-usedwalltime: 2128
nordugrid-job-usedcputime: 2128
nordugrid-job-reqcputime: 8640
nordugrid-job-executionnodes: node2/0
nordugrid-job-lrmscomment: Job started on Sun Mar 06 at 00:30
Mds-validfrom: 20050307110038Z
Mds-validto: 20050307110108Z

# search result
search: 2
result: 0 Success

# numResponses: 5
# numEntries: 4
\end{verbatim}
\normalsize

\item LDAP query to obtain the registration entries stored in an Index Service:

\tiny
\begin{verbatim}
ldapsearch -h quark.hep.lu.se -p 2135 -x -b 'mds-vo-name=Sweden,o=Grid' -s base giisregistrationstatus
version: 2

#
# filter: (objectclass=*)
# requesting: giisregistrationstatus
#

# se1:se1.hpc2n.umu.se, Sweden, grid
dn: nordugrid-se-name=se1:se1.hpc2n.umu.se, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: ido-i.hpc2n.umu.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: nordugrid-se-name=se1:se1.hpc2n.umu.se, Mds-Vo-name=l
 ocal, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 45
Mds-Service-Ldap-cachettl: 15
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: VALID

# bphysics:quark.hep.lu.se, Sweden, grid
dn: nordugrid-se-name=bphysics:quark.hep.lu.se, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: quark.hep.lu.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: nordugrid-se-name=bphysics:quark.hep.lu.se, Mds-Vo-na
 me=local, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 45
Mds-Service-Ldap-cachettl: 15
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: VALID

# Quark:quark.hep.lu.se, Sweden, grid
dn: nordugrid-se-name=Quark:quark.hep.lu.se, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: quark.hep.lu.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: nordugrid-se-name=Quark:quark.hep.lu.se, Mds-Vo-name=
 local, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 45
Mds-Service-Ldap-cachettl: 15
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: PURGED

# quark.hep.lu.se, Sweden, grid
dn: nordugrid-cluster-name=quark.hep.lu.se, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: quark.hep.lu.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: nordugrid-cluster-name=quark.hep.lu.se, Mds-Vo-name=l
 ocal, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 45
Mds-Service-Ldap-cachettl: 0
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: VALID

# files:grid.tsl.uu.se, Sweden, grid
dn: nordugrid-se-name=files:grid.tsl.uu.se, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: grid.tsl.uu.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: nordugrid-se-name=files:grid.tsl.uu.se, Mds-Vo-name=l
 ocal, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 45
Mds-Service-Ldap-cachettl: 15
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: PURGED

# toto7.lunarc.lu.se, Sweden, grid
dn: nordugrid-cluster-name=toto7.lunarc.lu.se, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: toto7.lunarc.lu.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: nordugrid-cluster-name=toto7.lunarc.lu.se, Mds-Vo-nam
 e=local, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 45
Mds-Service-Ldap-cachettl: 0
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: VALID

# Swegrid, Sweden, grid
dn: Mds-Vo-name=Swegrid, Mds-Vo-name=Sweden,o=grid
objectClass: Mds
objectClass: MdsVoOp
objectClass: MdsService
objectClass: MdsServiceLdap
Mds-Service-type: ldap
Mds-Service-hn: nexus.swegrid.se
Mds-Service-port: 2135
Mds-Service-Ldap-suffix: Mds-Vo-name=Swegrid, o=grid
Mds-Service-Ldap-sizelimit: 0
Mds-Service-Ldap-timeout: 120
Mds-Service-Ldap-cachettl: 30
Mds-Bind-Method-servers: ANONYM-ONLY
Mds-Reg-status: VALID

# search result
search: 2
result: 0 Success

# numResponses: 8
# numEntries: 7

\end{verbatim}
\normalsize

\end{itemize}

\newpage

\section{NorduGrid LDAP schema file\label{appendix:schema}}
\tiny

%
%TODO: Don't forget to insert the most recent schema file !!!!!!!!!
%
\begin{verbatim}

#---------------------------------------------------------
# These classes and attributes are imported from globus mds
# slightly to be proper LDAP schemas.

attributetype ( 1.3.6.1.4.1.11604.2.1.8.0.1
    NAME 'Mds-validfrom'
    DESC 'Object creation time'
    EQUALITY generalizedTimeMatch
    ORDERING generalizedTimeOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.24
    SINGLE-VALUE
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.0.2
    NAME 'Mds-validto'
    DESC 'Object expiration time'
    EQUALITY generalizedTimeMatch
    ORDERING generalizedTimeOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.24
    SINGLE-VALUE
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.0.3
    NAME 'Mds-keepto'
    DESC 'Object purge time'
    EQUALITY generalizedTimeMatch
    ORDERING generalizedTimeOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.24
    SINGLE-VALUE
 )

objectclass ( 1.3.6.1.4.1.11604.2.1.8
    NAME 'Mds'
    ABSTRACT
    MUST ( Mds-validfrom $ Mds-validto )
    MAY Mds-keepto
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.1.4.0.1
    NAME 'Mds-Vo-name'
    DESC 'Locally unique VO name'
    EQUALITY caseIgnoreMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE
 )

objectclass ( 1.3.6.1.4.1.11604.2.1.8.1.4
    NAME 'MdsVo'
    SUP 'Mds'
    STRUCTURAL
    MUST Mds-Vo-name
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.1.4.1.0.1
    NAME 'Mds-Vo-Op-name'
    DESC 'Locally unique Op name'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE
 )

objectclass ( 1.3.6.1.4.1.11604.2.1.8.1.4.1
    NAME 'MdsVoOp'
    SUP 'Mds'
    STRUCTURAL
    MUST Mds-Vo-Op-name
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.0.1
    NAME 'Mds-Service-type'
    DESC 'Service protocol'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.0.2
    NAME 'Mds-Service-protocol'
    DESC 'Service protocol OID'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.0.3
    NAME 'Mds-Service-port'
    DESC 'Service TCP port'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.0.4
    NAME 'Mds-Service-hn'
    DESC 'Service FQDN hostname'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.0.5
    NAME 'Mds-Service-url'
    DESC 'Service URL'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE
 )

objectclass ( 1.3.6.1.4.1.11604.2.1.8.2.7.1
    NAME 'MdsService'
    SUP 'Mds'
    AUXILIARY
    MUST ( Mds-Service-type $ Mds-Service-protocol $ Mds-Service-port $ Mds-Service-hn )
    MAY Mds-Service-url
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.1.0.1
    NAME 'Mds-Service-Ldap-suffix'
    DESC 'DN suffix of service'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.1.0.2
    NAME 'Mds-Service-Ldap-timeout'
    DESC 'suggested timeout'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.1.0.3
    NAME 'Mds-Service-Ldap-sizelimit'
    DESC 'suggested sizelimit'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.1.0.4
    NAME 'Mds-Service-Ldap-cachettl'
    DESC 'suggested cacheability'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.1.0.5
    NAME 'Mds-Service-Ldap-ttl'
    DESC 'suggested ttl'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.1.0.10
    NAME 'Mds-Reg-status'
    DESC 'VALID/INVALID/PURGED'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
 )

attributetype ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.1.0.11
    NAME 'Mds-Bind-Method-servers'
    DESC 'AUTHC-ONLY/AUTHC-FIRST/ANONYM-ONLY'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
 )

objectclass ( 1.3.6.1.4.1.11604.2.1.8.2.7.1.1
    NAME 'MdsServiceLdap'
    SUP 'MdsService'
    AUXILIARY
    MUST Mds-Service-Ldap-suffix
    MAY ( Mds-Service-Ldap-timeout $ Mds-Service-Ldap-sizelimit $ Mds-Service-Ldap-cachettl $ Mds-Service-Ldap-ttl $ Mds-Reg-status $ Mds-Bind-Method-servers )
 )

# attributes for the nordugrid-cluster objectclass 
#
attributetype ( 1.3.6.1.4.1.11604.2.1.1.1
    NAME 'nordugrid-cluster-name'
    DESC 'The name of the cluster specified as the domain name of the frontend'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.2
    NAME 'nordugrid-cluster-aliasname'
    DESC 'The alias name of the cluster'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE)

attributetype ( 1.3.6.1.4.1.11604.2.1.1.3
    NAME 'nordugrid-cluster-contactstring'
    DESC 'The URL of the job submission service running on the cluster frontend'
    EQUALITY caseExactIA5Match
    SUBSTR caseExactIA5SubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.26
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.4
    NAME 'nordugrid-cluster-support'
    DESC 'RFC822 email address of support'
    EQUALITY caseIgnoreIA5Match
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.26{256})

attributetype ( 1.3.6.1.4.1.11604.2.1.1.5
    NAME 'nordugrid-cluster-lrms-type'
    DESC 'The type of the Local Resource Management System'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.6
    NAME 'nordugrid-cluster-lrms-version'
    DESC 'The version of the Local Resource Management System'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.7
    NAME 'nordugrid-cluster-lrms-config'
    DESC 'Additional remark on the LRMS config'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.8
    NAME 'nordugrid-cluster-architecture'
    DESC 'The architecture of the cluster nodes'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.9
    NAME 'nordugrid-cluster-opsys'
    DESC 'The operating system of the machines of the cluster'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15)

attributetype ( 1.3.6.1.4.1.11604.2.1.1.10
    NAME 'nordugrid-cluster-homogeneity'
    DESC 'A logical flag indicating the homogeneity of the cluster nodes'
    EQUALITY booleanMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.7
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.11
    NAME 'nordugrid-cluster-nodecpu'
    DESC 'The cpu type of the nodes expressed in a fixed form (model name + MHz)'
    EQUALITY caseIgnoreMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.12
    NAME 'nordugrid-cluster-nodememory'
    DESC 'The amount of memory which can be guaranteed to be available on the node in MB'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.13
    NAME 'nordugrid-cluster-totalcpus'
    DESC 'The total number of cpus in the cluster'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.14
    NAME 'nordugrid-cluster-cpudistribution'
    DESC 'The cpu distribution of the nodes given in the form of 1cpu:3 2cpu:4 4cpu:1'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.15
    NAME 'nordugrid-cluster-sessiondir-free'
    DESC 'Free diskspace in MB of the sessiondirectory on the cluster'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.16
    NAME 'nordugrid-cluster-sessiondir-total'
    DESC 'Total diskspace in MB of the sessiondirectory on the cluster'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.17
    NAME 'nordugrid-cluster-cache-free'
    DESC 'Free diskspace in MB of the cache area on the cluster'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.18
    NAME 'nordugrid-cluster-cache-total'
    DESC 'Total diskspace in MB of the cache area on the cluster'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.19
    NAME 'nordugrid-cluster-runtimeenvironment'
    DESC 'preinstalled software packages of the cluster'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.20
    NAME 'nordugrid-cluster-localse'
    DESC 'The URL of a storage element considered to be local to the cluster'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.21
    NAME 'nordugrid-cluster-middleware'
    DESC 'The middleware packages on the cluster'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.22
    NAME 'nordugrid-cluster-totaljobs'
    DESC 'The total number of jobs (Grid + non-Grid) in the cluster'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.23
    NAME 'nordugrid-cluster-usedcpus'
    DESC 'The total number of occupied cpus in the cluster'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.24
    NAME 'nordugrid-cluster-queuedjobs'
    DESC 'The total number of jobs (grid and non-grid) not-yet running: preparing or waiting to run 
          on the cluster, either in the grid-manager or in the LRMS. 
	  The attribute is TO BE DEPRECATED'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.25
    NAME 'nordugrid-cluster-location'
    DESC 'The geographical location of the cluster expressed in terms of a Postal ZIP code'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44 
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.26
    NAME 'nordugrid-cluster-owner'
    DESC 'The owner of the resource'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.27
    NAME 'nordugrid-cluster-issuerca'
    DESC 'The DN of the Certificate Authority which issued the certificate of the cluster'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.28
    NAME 'nordugrid-cluster-nodeaccess'
    DESC 'The inbound/outbound network accessibility of the nodes'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44 )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.29
    NAME 'nordugrid-cluster-comment'
    DESC 'Free form comment'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.30
    NAME 'nordugrid-cluster-interactive-contactstring'
    DESC 'The URL for interactive login'
    EQUALITY caseExactIA5Match
    SUBSTR caseExactIA5SubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.26 )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.31
    NAME 'nordugrid-cluster-benchmark'
    DESC '@ separated benchmark_name, benchmark_value pair characterizing the cluster nodes'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.32
    NAME 'nordugrid-cluster-sessiondir-lifetime'
    DESC 'The lifetime of the sessiondir after the job has completed (in minutes)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.33
    NAME 'nordugrid-cluster-prelrmsqueued'
    DESC 'The total number of grid jobs not-yet reached the LRMS: preparing or queuing in the grid-layer'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.34
    NAME 'nordugrid-cluster-issuerca-hash'
    DESC 'The HASH of the Certificate Authority which issued the certificate for the cluster'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44 
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.35
    NAME 'nordugrid-cluster-trustedca'
    DESC 'The DN of a Certificate Authority trusted by the cluster'
    EQUALITY caseIgnoreMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.1.36
    NAME 'nordugrid-cluster-acl'
    DESC 'Cluster authorization information'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44 )

attributetype (1.3.6.1.4.1.11604.2.1.1.37
    NAME 'nordugrid-cluster-credentialexpirationtime'
    DESC 'The expiration date of the shortest living credential affecting the cluster's x509 environment in GMT'
    EQUALITY generalizedTimeMatch
    ORDERING generalizedTimeOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.24
    SINGLE-VALUE )

objectclass ( 1.3.6.1.4.1.11604.2.1.1
    NAME 'nordugrid-cluster'
    DESC 'Description of a Nordugrid cluster'
    SUP 'Mds'
    STRUCTURAL
    MUST ( nordugrid-cluster-name $ nordugrid-cluster-contactstring )
    MAY  ( nordugrid-cluster-aliasname $ nordugrid-cluster-support $
           nordugrid-cluster-lrms-type $ nordugrid-cluster-lrms-version $
	   nordugrid-cluster-lrms-config $ nordugrid-cluster-architecture $
	   nordugrid-cluster-opsys $ nordugrid-cluster-homogeneity $
	   nordugrid-cluster-nodecpu $ nordugrid-cluster-nodememory $
	   nordugrid-cluster-totalcpus $ nordugrid-cluster-cpudistribution $
	   nordugrid-cluster-sessiondir-free $ nordugrid-cluster-sessiondir-total $
	   nordugrid-cluster-cache-free $ nordugrid-cluster-cache-total $
	   nordugrid-cluster-runtimeenvironment $ nordugrid-cluster-localse $
	   nordugrid-cluster-middleware $ nordugrid-cluster-totaljobs $
	   nordugrid-cluster-usedcpus $	nordugrid-cluster-queuedjobs  $
	   nordugrid-cluster-location $ nordugrid-cluster-owner $ 
	   nordugrid-cluster-issuerca $ nordugrid-cluster-nodeaccess $
	   nordugrid-cluster-comment $ nordugrid-cluster-interactive-contactstring $
	   nordugrid-cluster-benchmark $ nordugrid-cluster-sessiondir-lifetime $
	   nordugrid-cluster-prelrmsqueued $ nordugrid-cluster-issuerca-hash $
	   nordugrid-cluster-trustedca $ nordugrid-cluster-acl $
	   nordugrid-cluster-credentialexpirationtime ))

#-----------------------------------------------------------------
# attributes for the nordugrid-info-group objectclass
#
attributetype ( 1.3.6.1.4.1.11604.2.1.2.1
    NAME 'nordugrid-info-group-name'
    DESC 'Locally unique info group name'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE)

objectclass ( 1.3.6.1.4.1.11604.2.1.2
    NAME 'nordugrid-info-group'
    DESC 'A general entry for grouping together MDS entries'
    SUP 'Mds'
    STRUCTURAL
    MUST ( nordugrid-info-group-name ))

#-----------------------------------------------------------------
# attributes for the nordugrid-queue objectclass
#
attributetype ( 1.3.6.1.4.1.11604.2.1.3.1
    NAME 'nordugrid-queue-name'
    DESC 'The queue name'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.2
    NAME 'nordugrid-queue-status'
    DESC 'The queue status'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.3
    NAME 'nordugrid-queue-running'
    DESC 'Number of running jobs (Grid + non-Grid) in the queue with multi-node jobs multiciplity'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.4
    NAME 'nordugrid-queue-queued'
    DESC 'The number of jobs (Grid + non-Grid) waiting in the queue.
          The attribute is TO BE DEPRECATED'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.5
    NAME 'nordugrid-queue-maxrunning'
    DESC 'The maximum number of jobs allowed to run from this queue'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.6
    NAME 'nordugrid-queue-maxqueuable'
    DESC 'The maximum number of jobs allowed to reside in the queue'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.7
    NAME 'nordugrid-queue-maxuserrun'
    DESC 'Maximum number of jobs a user can run at the same time'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.8
    NAME 'nordugrid-queue-maxcputime'
    DESC 'The maximum cputime allowed in this queue (in minutes)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.9
    NAME 'nordugrid-queue-mincputime'
    DESC 'The minimum possible cputime of this queue (in minutes)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.10
    NAME 'nordugrid-queue-defaultcputime'
    DESC 'The default cputime assigned to this queue (in minutes)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.11
    NAME 'nordugrid-queue-schedulingpolicy'
    DESC 'The scheduling policy of the queue (i.e. FIFO)'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.12
    NAME 'nordugrid-queue-totalcpus'
    DESC 'Total number of cpus assigned to the queue'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.13
    NAME 'nordugrid-queue-nodecpu'
    DESC 'The cpu type of the nodes assigned to the queue (model name + MHz)'
    EQUALITY caseIgnoreMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.14
    NAME 'nordugrid-queue-nodememory'
    DESC 'The installed memory of a node assigned to the queue in MB'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.15
    NAME 'nordugrid-queue-architecture'
    DESC 'The architecture of the machines in the queue'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.16
    NAME 'nordugrid-queue-opsys'
    DESC 'The operating system of the nodes in the queue'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15)

attributetype ( 1.3.6.1.4.1.11604.2.1.3.17
    NAME 'nordugrid-queue-gridrunning'
    DESC 'Number of running Grid jobs in the queue with multi-node jobs multiciplity'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.18
    NAME 'nordugrid-queue-gridqueued'
    DESC 'The number of Grid jobs waiting in the LRMS queue'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.19
    NAME 'nordugrid-queue-comment'
    DESC 'Free form comment'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.20
    NAME 'nordugrid-queue-benchmark'
    DESC 'Colon separated benchmark_name, benchmark_value pair characterizing the queue'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.21
    NAME 'nordugrid-queue-homogeneity'
    DESC 'A logical flag indicating the homogeneity of the queue nodes'
    EQUALITY booleanMatch  
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.7
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.22
    NAME 'nordugrid-queue-prelrmsqueued'
    DESC 'The number of Grid jobs belonging to this queue being processed
    	 or waiting in the Grid-layer before the LRMS submission.'      
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.23
    NAME 'nordugrid-queue-localqueued'
    DESC 'The number of non-Grid jobs waiting in the LRMS queue.'    
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.24
    NAME 'nordugrid-queue-maxwalltime'
    DESC 'The maximum walltime allowed in this queue (in minutes)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.25
    NAME 'nordugrid-queue-minwalltime'
    DESC 'The minimum possible walltime of this queue (in minutes)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.26
    NAME 'nordugrid-queue-defaultwalltime'
    DESC 'The default walltime assigned to this queue (in minutes)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.3.27
    NAME 'nordugrid-queue-maxtotalcputime'
    DESC 'The maximum total cputime allowed in this queue (in minutes)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

objectclass ( 1.3.6.1.4.1.11604.2.1.3
    NAME 'nordugrid-queue'
    DESC 'An LRMS queue'
    SUP 'Mds'
    STRUCTURAL
    MUST ( nordugrid-queue-name $ nordugrid-queue-status)    
    MAY  ( nordugrid-queue-running $ nordugrid-queue-queued $
    	   nordugrid-queue-maxrunning  $ nordugrid-queue-maxqueuable$
    	   nordugrid-queue-maxuserrun $ nordugrid-queue-maxcputime $
    	   nordugrid-queue-mincputime $ nordugrid-queue-defaultcputime $
    	   nordugrid-queue-schedulingpolicy $ nordugrid-queue-totalcpus $
    	   nordugrid-queue-nodecpu $ nordugrid-queue-nodememory $
	   nordugrid-queue-opsys $ nordugrid-queue-architecture $	   
	   nordugrid-queue-gridrunning $ nordugrid-queue-gridqueued $
	   nordugrid-queue-comment $ nordugrid-queue-benchmark $
	   nordugrid-queue-homogeneity $ nordugrid-queue-prelrmsqueued $
	   nordugrid-queue-localqueued $ nordugrid-queue-maxwalltime $
	   nordugrid-queue-minwalltime $ nordugrid-queue-defaultwalltime $
	   nordugrid-queue-maxtotalcputime ))

#-----------------------------------------------------------------
#attributes for the nordugrid-job objectclass
#
attributetype ( 1.3.6.1.4.1.11604.2.1.4.1
    NAME 'nordugrid-job-globalid'
    DESC 'The global job identifier string'
    EQUALITY caseExactIA5Match
    SUBSTR caseExactIA5SubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.26    
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.2
    NAME 'nordugrid-job-globalowner'
    DESC 'The SubjectName of the job owner'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44  
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.3
    NAME 'nordugrid-job-execcluster'
    DESC 'The name of the execution cluster'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE)

attributetype ( 1.3.6.1.4.1.11604.2.1.4.4
    NAME 'nordugrid-job-execqueue'
    DESC 'The name of the execution queue'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.5
    NAME 'nordugrid-job-stdout'
    DESC 'The name of the file which contains the stdout'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.6
    NAME 'nordugrid-job-stderr'
    DESC 'The name of the file which contains the stderr'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.7
    NAME 'nordugrid-job-stdin'
    DESC 'The name of the file which contains the stdin'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.8
    NAME 'nordugrid-job-reqcputime'
    DESC 'The cputime request by the job in minutes'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.9
    NAME 'nordugrid-job-status'
    DESC 'The status of the grid job'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.10
    NAME 'nordugrid-job-queuerank'
    DESC 'The queue position of the job'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.11
    NAME 'nordugrid-job-comment'
    DESC 'Free form comment about the job'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.12
    NAME 'nordugrid-job-submissionui'
    DESC 'The name of the UI from where the job was submitted'
    EQUALITY caseIgnoreMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.13
    NAME 'nordugrid-job-submissiontime'
    DESC 'The submission time of the job in GMT'
    EQUALITY generalizedTimeMatch
    ORDERING generalizedTimeOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.24
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.14
    NAME 'nordugrid-job-usedcputime'
    DESC 'The consumed cputime of the job in minutes'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.15
    NAME 'nordugrid-job-usedwalltime'
    DESC 'The consumed walltime of the job in minutes'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.16
    NAME 'nordugrid-job-sessiondirerasetime'
    DESC 'The date when the session dir will be deleted in GMT'
    EQUALITY generalizedTimeMatch
    ORDERING generalizedTimeOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.24
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.17
    NAME 'nordugrid-job-usedmem'
    DESC 'The memory usage of the job (in KB)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.18
    NAME 'nordugrid-job-errors'
    DESC 'Error mesages from the cluster'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.19
    NAME 'nordugrid-job-jobname'
    DESC 'The jobname specified by the user with the jobname RSL attribute'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.20
    NAME 'nordugrid-job-runtimeenvironment'
    DESC 'The runtimeenvironment requested by the job'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.21
    NAME 'nordugrid-job-cpucount'
    DESC 'The number of CPUs requested by the job'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.22
    NAME 'nordugrid-job-executionnodes'
    DESC 'The list of nodenames where the job is running'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.23
    NAME 'nordugrid-job-gmlog'
    DESC 'The name of the directory which contains the grid session related logs'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.24
    NAME 'nordugrid-job-clientsoftware'
    DESC 'The client software which submitted the job'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15)

attributetype ( 1.3.6.1.4.1.11604.2.1.4.25
    NAME 'nordugrid-job-proxyexpirationtime'
    DESC 'The expiration time of the proxy of the job in GMT'
    EQUALITY generalizedTimeMatch
    ORDERING generalizedTimeOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.24
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.26
    NAME 'nordugrid-job-completiontime'
    DESC 'The completion time of the grid job in GMT'
    EQUALITY generalizedTimeMatch
    ORDERING generalizedTimeOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.24
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.27
    NAME 'nordugrid-job-exitcode'
    DESC 'The exit code of the executable of the job obtained from the LRMS'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.28
    NAME 'nordugrid-job-rerunable'
    DESC 'Rerunability of the FAILED grid jobs'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.4.29
    NAME 'nordugrid-job-reqwalltime'
    DESC 'The request wallclock time of the job in minutes'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

objectclass ( 1.3.6.1.4.1.11604.2.1.4
    NAME 'nordugrid-job'
    DESC 'A Grid job'
    SUP 'Mds'
    STRUCTURAL
    MUST ( nordugrid-job-globalid $ nordugrid-job-globalowner $
           nordugrid-job-status ) 
    MAY  ( nordugrid-job-queuerank $ nordugrid-job-submissionui $
    	   nordugrid-job-submissiontime  $
    	   nordugrid-job-usedcputime $ nordugrid-job-usedwalltime $
    	   nordugrid-job-usedmem $ nordugrid-job-comment $
    	   nordugrid-job-execcluster $ nordugrid-job-execqueue $
    	   nordugrid-job-stdout $ nordugrid-job-stderr $    	   
    	   nordugrid-job-stdin $ 
    	   nordugrid-job-sessiondirerasetime $ nordugrid-job-reqcputime $
	   nordugrid-job-errors $ nordugrid-job-jobname $
	   nordugrid-job-runtimeenvironment $ nordugrid-job-cpucount $
	   nordugrid-job-executionnodes $ nordugrid-job-gmlog $
	   nordugrid-job-clientsoftware $ nordugrid-job-proxyexpirationtime $
	   nordugrid-job-completiontime $ nordugrid-job-exitcode $
	   nordugrid-job-rerunable $ nordugrid-job-reqwalltime))

#----------------------------------------------------------------
# attributes for the nordugrid-authuser objectclass
#

attributetype ( 1.3.6.1.4.1.11604.2.1.5.1
    NAME 'nordugrid-authuser-name'
    DESC 'The Common Name of the authorized user plus a local unique number'
    EQUALITY caseIgnoreMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.5.2
    NAME 'nordugrid-authuser-sn'
    DESC 'The SubjectName of the authorized user'
    EQUALITY caseIgnoreMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.5.3
    NAME 'nordugrid-authuser-freecpus'
    DESC 'The number of freely available cpus with their timelimits in minutes 
          for a user in the queue. Given in the form ncpus:min,
	  min is optional (example: 2 4:25 5:180)'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch 
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.5.4
    NAME 'nordugrid-authuser-diskspace'
    DESC 'The free diskspace available for the job (in MB)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.5.5
    NAME 'nordugrid-authuser-queuelength'
    DESC 'The number of queuing jobs of a particular user, 
          both queuing in the LRMS and in the Grid-layer'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

objectclass ( 1.3.6.1.4.1.11604.2.1.5
    NAME 'nordugrid-authuser'
    DESC 'An authorised user of a NorduGrid cluster'
    SUP 'Mds'
    STRUCTURAL
    MUST ( nordugrid-authuser-name $ nordugrid-authuser-sn ) 
    MAY  ( nordugrid-authuser-queuelength $ nordugrid-authuser-diskspace $
           nordugrid-authuser-freecpus ))

#----------------------------------------------------------------
# 
# nordugrid-se

attributetype ( 1.3.6.1.4.1.11604.2.1.6.1
    NAME 'nordugrid-se-name'
    DESC 'The name of the Storage Element'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.2
    NAME 'nordugrid-se-aliasname'
    DESC 'The alias name of the SE'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.3
    NAME 'nordugrid-se-type'
    DESC 'The type of the SE'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.4
    NAME 'nordugrid-se-freespace'
    DESC 'The free space available in the SE (in MB)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.5
    NAME 'nordugrid-se-url'
    DESC 'The URL to contact the Storage Element'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.6
    NAME 'nordugrid-se-authuser'
    DESC 'The DN of an authorized user'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.7
    NAME 'nordugrid-se-location'
    DESC 'The geographical location of the SE expressed in terms of a Postal ZIP code: SE-22363'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44 )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.8
    NAME 'nordugrid-se-owner'
    DESC 'The owner of the resource'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.9
    NAME 'nordugrid-se-issuerca'
    DESC 'The DN of the Certificate Authority which issued the certificate of the SE'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.10
    NAME 'nordugrid-se-totalspace'
    DESC 'The total capacity of the SE (in MB)'
    EQUALITY integerMatch
    ORDERING integerOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.27
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.11
    NAME 'nordugrid-se-middleware'
    DESC 'The middleware packages on the SE'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.12
    NAME 'nordugrid-se-comment'
    DESC 'Free form comment'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.13
    NAME 'nordugrid-se-accesscontrol'
    DESC 'The access control framework of the SE'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.14
    NAME 'nordugrid-se-issuerca-hash'
    DESC 'The HASH of the Certificate Authority which issued the certificate for the SE'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44 
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.15
    NAME 'nordugrid-se-trustedca'
    DESC 'The DN of a Certificate Authority trusted by the SE'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.6.16
    NAME 'nordugrid-se-acl'
    DESC 'Storage Element authorization information'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

objectclass ( 1.3.6.1.4.1.11604.2.1.6
    NAME 'nordugrid-se'
    DESC 'A storage element in the Nordugrid'
    SUP 'Mds'
    STRUCTURAL
    MUST ( nordugrid-se-name $ nordugrid-se-url)
    MAY  ( nordugrid-se-aliasname $ nordugrid-se-type $ 
    	   nordugrid-se-freespace $ nordugrid-se-authuser $
	   nordugrid-se-location $ nordugrid-se-owner $
	   nordugrid-se-issuerca $ nordugrid-se-totalspace $ 
	   nordugrid-se-middleware $ nordugrid-se-comment $
	   nordugrid-se-accesscontrol $ nordugrid-se-issuerca-hash $
	   nordugrid-se-trustedca $ nordugrid-se-acl ))

#--------------------------------------------------------------------
# nordugrid-rc
#
attributetype ( 1.3.6.1.4.1.11604.2.1.7.1
    NAME 'nordugrid-rc-name'
    DESC 'The domain name of the machine hosting the Replica Catalog'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.7.2
    NAME 'nordugrid-rc-aliasname'
    DESC 'The alias name of the rc'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.7.3
    NAME 'nordugrid-rc-baseurl'
    DESC 'The URL of the Replica Catalog'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15
    SINGLE-VALUE )

attributetype ( 1.3.6.1.4.1.11604.2.1.7.4
    NAME 'nordugrid-rc-authuser'
    DESC 'An authorized user of the replica catalog'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.7.5
    NAME 'nordugrid-rc-location'
    DESC 'The geographical location of the RC expressed in terms of a Postal ZIP code'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.44 )

attributetype ( 1.3.6.1.4.1.11604.2.1.7.6
    NAME 'nordugrid-rc-owner'
    DESC 'The owner of the resource'
    EQUALITY caseIgnoreMatch
    ORDERING caseIgnoreOrderingMatch
    SUBSTR caseIgnoreSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

attributetype ( 1.3.6.1.4.1.11604.2.1.7.7
    NAME 'nordugrid-rc-issuerca'
    DESC 'The DN of the Certificate Authority which issued the certificate of the RC'
    EQUALITY caseExactMatch
    ORDERING caseExactOrderingMatch
    SUBSTR caseExactSubstringsMatch
    SYNTAX 1.3.6.1.4.1.1466.115.121.1.15 )

objectclass ( 1.3.6.1.4.1.11604.2.1.7
    NAME 'nordugrid-rc'
    DESC 'A replica catalogue in the Nordugrid'
    SUP 'Mds'
    STRUCTURAL
    MUST ( nordugrid-rc-name $ nordugrid-rc-baseurl )
    MAY  ( nordugrid-rc-aliasname $ nordugrid-rc-authuser $
    	   nordugrid-rc-location $ nordugrid-rc-owner $
	   nordugrid-rc-issuerca ))

\end{verbatim}
\newpage

\normalsize
\begin{thebibliography}{99}

\bibitem{NorduGrid}
	NorduGrid project. \texttt{http://www.nordugrid.org}

\bibitem{Openldap} 
	Openldap. \texttt{http://www.openldap.org}

\bibitem{mds}
        Monitoring and Discovery Services. \texttt{http://www.globus.org/mds/mds2/}

\bibitem{grid-manager}
	The NorduGrid Grid Manager And GridFTP Server: Description And Administrator's
	Manual. \texttt{http://www.nordugrid.org/papers.html}

\bibitem{arc_brokering}	
	The NorduGrid Brokering Algorithm, M.Ellert,
	\texttt{http://www.nordugrid.org/papers.html}

\bibitem{xrsl_manual}
	xRSL (Extended Resource Specification Language), O.Smirnova,
	\texttt{http://www.nordugrid.org/papers.html}

\bibitem{rer} 
        Runtime Environment Registry,
	\texttt{http://www.csc.fi/grid/rer/}

\bibitem{logger} Usage statistics and usage patterns on the NorduGrid,
	K.Pajchel,
	\texttt{http://www.nordugrid.org/papers.html}

\bibitem{monitor} The Grid Monitor: Usage Manual,
	\texttt{http://www.nordugrid.org/documents/monitor.pdf}

\bibitem{glue1.2} The GLUE Information model versin 1.2
	\texttt{http://infnforge.cnaf.infn.it/glueinfomodel/}

\bibitem{ui} The NorduGrid Toolkit User Interface: User's Manual
        \texttt{http://www.nordugrid.org/documents/NorduGrid-UI.pdf}

\bibitem{smartse} The NorduGrid "Smart" Storage Element, A.Konstantinov,
	\texttt{http://www.nordugrid.org/papers.html}

\bibitem{gacl} The Gridsite,
	\texttt{http://www.gridsite.org}

\bibitem{bdii} BDII homepage,
	\texttt{https://twiki.cern.ch/twiki/bin/view/EGEE/BDIIv4}

\end{thebibliography}

\end{document}

% LocalWords:  objectclass OID's

### 2010-12-28 doc/tech_doc/urlogger/urlogger.tex
Copying URlogger documentation to arc1

\documentclass{article}
%
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{graphics}
%
\usepackage{graphicx}                              %for PNG images
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{makeidx}                               %for index page
\usepackage{geometry}                              %for defining page size
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref} %for \url tag
\usepackage{doc}                                   %for pfill used by gist index style
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\makeindex
\hypersetup{
  pdfauthor = {Henrik Thostrup Jensen},
  pdftitle = {ARC Usage Record Logger},
  pdfsubject = {Setup Guide},
  pdfkeywords = {Grid,ARC,usage},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
  \def\today{\number\day/\number\month/\number\year}

  \begin{titlepage}

    \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
    \end{tabular}

    \hrulefill

    {\raggedleft NORDUGRID-MANUAL-16\par}

    {\raggedleft \today\par}

    \vspace*{2cm}

%%%%---- The title ----
    {\centering \textsc{\Large ARC Usage Record Logger}\Large \par}

    \vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
    {\centering \textit{\large Setup Guide}\large \par}

    \vspace*{2cm}

%%%%---- A list of authors ----
    {\centering \large Henrik Thostrup Jensen \texttt{<htj@ndgf.org>} \large \par}

    \vspace*{1.5cm}

%%%%---- An abstract ----
%\begin{abstract}
%\end{abstract}

\end{titlepage}
\thispagestyle{empty} $ $
\newpage
$\ $

\section{Setup}

From version 0.8.1 ARC includes a usage record generator and registrator to
SGAS. This document describe how to configure these.

\subsection{Requirements}

\begin{itemize}
\item Python 2.4 or later
\item Twisted Core and Web (http://twistedmatrix.com/)
\item PyOpenSSL (https://launchpad.net/pyopenssl)
\item ElementTree (http://effbot.org/zone/element-index.htm - only needed with Python 2.4)
\end{itemize}

\noindent
Debian/Ubuntu package names:

\texttt{python-twisted-core, python-twisted-web, python-openssl} \\
\texttt{python-elementtree} (only needed with Python 2.4)

CentOS and similar:

\texttt{python-twisted-core python-twisted-web pyOpenSSL} \\
\texttt{python-elementtree} (only needed with Python 2.4)

Other Linux distributions: You are on your own :-)

\subsection{arc.conf}

To invoke the UR generator, an authplugin line must be set in the
\texttt{[grid-manager]} section:

\begin{verbatim}
[grid-manager]
authplugin="FINISHED timeout=10,onfailure=pass /opt/nordugrid/libexec//arc-ur-logger %C %I %S %U"
\end{verbatim}

Change the path if you've installed ARC elsewhere than /opt/nordugrid. You'll
need to restart the grid-manager for this to take effect (but wait until you
finished reading this document).

The plugin will log to the file: /var/log/arc-ur-logger.log. This file will not
appear until a job has finished.

Additional logger configuration happens in a seperate logger section. Example:

\begin{verbatim}
[logger]
log_dir=/var/spool/nordugrid/usagerecords/
log_all="https://sgas.ndgf.org:6143/sgas"
log_vo="bio.ndgf.org https://biosgas.ndgf.org:6143/sgas"
ur_lifetime=30
\end{verbatim}

The \texttt{log\_dir} option will set the top directory for the generated usage
records. The option will default to \texttt{/var/spool/nordugrid/usagerecords/}
and can be left out. We suggest you leave out this option, unless you have a
reason not to.

The \texttt{log\_all} and \texttt{log\_vo} options configure where to usage
records are registered. All usage records will be registered to the URLs
specified with the \texttt{log\_all} option. It is possible specifiy multiple
URLs be having them space seperated, e.g.:

\begin{verbatim}
log_all="https://sgas.ndgf.org:6143/sgas https://sgas.grid.dk:6143/sgas"
\end{verbatim}

The \texttt{log\_vo} option makes it possible to only register usage records run
with certain VO users to given URL. The above example will register all usage
records where the VO information includes the VO bio.ndgf.org to the url
\texttt{https://biosgas.ndgf.org:6143/sgas}. Is is possible to have multiple
entries, by seperating entries with comma, e.g.,

\begin{verbatim}
log_vo="vo1 url1, vo2 url2"
\end{verbatim}

It is NOT possible to have multiple \texttt{log\_all} or \texttt{log\_vo} lines.

The \texttt{ur\_lifetime} option specifies how many days usage records are kept
after being archived. The default is 30, and the option can be left out.

It is possible to set the logfiles for both the logger and the registrant.
Furthermore it is possible to specify logging level for the logger (not
possible or needed for the registrant, as it is not that verbose). Examples:

\begin{verbatim}
urlogger_logfile="/tmp/arc-ur-logger.log"
urlogger_loglevel="info"
registrant_logfile="/tmp/arc-ur-registrant.log"
\end{verbatim}

By default the ur logger will write to \texttt{/var/log/arc-ur-logger.log} and
the registrant to \\ \texttt{/var/log/arc-ur-registration.log}. Valid options
for logger log level are: \texttt{debug}, \texttt{info}, \texttt{warning}. The
default is \texttt{info} which will write one line log per job, assuming
everything goes as planned.

\subsection{cron}

To register the usage records and registrant should be invoked regularly be
CRON. We suggest every hour. The crontab entry would typically look like this:

\begin{verbatim}
0 * * * * /opt/nordugrid/libexec/arc-ur-registrant
\end{verbatim}

To ensure that the registrant is working, you can run the script from the
command line first. Note that the script will still write its log to to
/var/log/arc-ur-registration.log. By running the script with -s its ouput will
be directed to stdout.

\end{document}

### 2011-01-05 src/clients/data/arccp.cpp
X509 env variables are not needed for arccp -y

@@ -528,9 +528,7 @@ int main(int argc, char **argv) {
 
   std::string cache_path;
   options.AddOption('y', "cache",
-                    istring("path to local cache (use to put file into cache). "
-                            "The X509_USER_PROXY and X509_CERT_DIR environment "
-                            "variables must be set correctly."),
+                    istring("path to local cache (use to put file into cache)"),
                     istring("path"), cache_path);
 
   int recursion = 0;

### 2011-01-11 doc/tech_doc/xrsl/xrsl.tex

@@
 
-%    {\centering \large Refers to ARC release series 0.4 and up \large \par}
+%    {\centering \large Refers to ARC release series 0.8 and up \large \par}
 
@@
 
-  Execution of jobs at batch clusters is the major part of the data
-  analysis, production and other computational tasks and applications
-  in many research areas. In the Grid infrastructure~\cite{grid}, such
-  clusters are not necessarily local to a job submission site, but can
-  be widely distributed. This implies extra requirements for the
-  proper description of job options. The \globus\ \index{Globus}
-  project~\cite{globus} developed a Grid middleware toolkit, which
-  makes use of the \textit{Resource Specification Language
+  Computational tasks, such as data mining, data transformation, simulation of various conditions, execution of Monte Carlo algorihms and so on, are normally formalized as series of \textit{jobs} submitted to computing resources.
+Such computing resources can be mainframe supercomputers, high-performance computing clusters, or even pools of regular desktop PCs.
+In a conventional approach, applicaton experts obtain user acconts at computing centers and learn peculiarities of their \textit{batch systems} in order to formulate and submit computational jobs to the assigned resource.
+  In the Grid infrastructure~\cite{grid}, experts get simultaneous access to a large number of very different computing resources, which are often widely distributed geographically.
+In order to abstract from the heterogeneous nature of Grid resources, a high-level job description language is necessary.
+The diversity of Grid resources implies special requirements for a
+  proper description of a job, introducing many new options as opposed to a conventional high-performance computing center use case.
+
+  Globus\ \index{Globus}
+  Alliance~\cite{globus} developed a Grid middleware toolkit, early versions of which
+  made use of the \textit{Resource Specification Language
   (RSL)}~\cite{rsl}\index{RSL} to parse job options and definitions to
   resource management systems. The NorduGrid project~\cite{nordugrid}
-  \index{NorduGrid} developed the Advanced Resource Connector (ARC),
+  \index{NorduGrid} decided to re-use RSL when they developed the Advanced Resource Connector (ARC),
   -- a
-  solution~\cite{para02,acat02} for a Grid facility, suitable for
-  complex tasks, like, for example, a High Energy Physics
-  production. To match the complexity of such tasks, this solution
+  solution~\cite{arc} for a Grid facility, suitable for
+  complex tasks, like, for example, High Energy Physics
+  data processing. To match the complexity of such tasks, this solution
   requires certain extensions to the RSL.
 
-  To describe a task to be submitted to the ARC-enabled resources, an
-  extended version of the \globus\  RSL is used. The extensions concern
+  To describe a task to be submitted to ARC-enabled resources, an
+  extended version of the \globus\  RSL is used\footnote{One can also use a Grid-standard JSDL language; however, it is less versatile}. Extensions concern
   not only introduction of new attributes, but also differentiation
-  between two levels of the job options specifications:
+  between the two levels of job option specifications:
 
@@
     specified by a user in a job-specific file. This file is
-    interpreted by the \textit{User Interface (UI)}~\cite{ui},
-    \index{User Interface} and after the necessary modifications is
-    parsed to the \textit{Grid Manager (GM)}~\cite{gm}\index{Grid
+    interpreted by a \textit{Client}~\cite{ui},
+    \index{Client} and after the necessary modifications is
+    passed to the ARC computing service: the \textit{Grid Manager (GM)}~\cite{gm}\index{Grid
     Manager}
   \item[] \textbf{GM-side RSL}, i.e., the set of attributes
-    pre-processed by the UI, and ready to be interpreted by the GM
+    pre-processed by a client, and ready to be interpreted by the GM. Effectively, this is an \textit{internal} job representation of ARC.
   \end{itemize}
@@
   describe the Grid tasks. The Grid Manager, however, uses slightly
-  different notations, supplied by the User Interface.
+  different notations, supplied by the client tools; therefore developers of such tools must take care of converting user-submitted RSL to the internal job description (GM-side RSL).
 
-  In what follows, the description of the NorduGrid-extended RSL,
+  In what follows, description of the NorduGrid-extended RSL,
   further denoted as \textbf{xRSL}\index{xRSL}, is given, using the
@@
   \end{tabular}
+
+Usage examples are given for the standard ARC command-line client (also known as \textit{User Interface})~\cite{ui}; other clients may introduce additional features, please refer to their respective documentation for details.
 
@@
 
-  A Grid task is described by the mean of xRSL attributes, which can
+  A Grid task is described by means of xRSL attributes, which can
   be either passed via a command-line, or, more conveniently, be
-  collected a so-called xRSL-file (suggested extension
+  collected in a so-called xRSL-file (suggested extension
   \textit{.xrsl}). Such a file contains a plain list of attribute-value pairs
-  and boolean operands "\&" (for AND) and "\verb#|#" (for
+  and boolean operators "\&" (for AND) and "\verb#|#" (for
   OR). Attribute names are case-insensitive.
@@
     \verb#(attribute="value")#\\
-    \verb#(attribute="value1" "value2")# (for multi-valued attributes)\\
+    \verb#(attribute="value1" "value2")# (for multi-valued attributes)
   \end{shaded}
@@
   either single or double quotes. If your string, however, contains
-  both such characters, you can define any character as an own
+  both such quotes, you can define any character as an own
   delimiter, by preceding it with the "carat" (\verb#^#) character:
@@
 
-  Typically, an xRSL job description starts with an ampersand
+  An xRSL job description starts with an ampersand
   ("\&") , to indicate implicit \textbf{conjunction} of all the
@@
 
-  In expressions, the following operands are allowed:
+  \begin{framed}
+   Only few selected attributes (indicated further in the document) can be requested by the user multiple times, like in the disjunct request example above.
+Most attributes \textbf{must be unique}, i.e., appear only once in the job description document.
+  \end{framed}
+
+  In expressions, the following relational operators are allowed, in general:
   \begin{shaded}
@@
   \end{shaded}
+  However, most attrbutes can \textbf{only be used with equality operator "="}. For few attributes (as indicated in the document), some other operators can be used as well.
 
@@
   \textbf{Multiple job} description in one file is realized via a standard
-  \globus\  RSL multi-request operand "+", which should precede
+  \globus\  RSL multi-request operator "+", which should precede
   multiple job description:
@@
 
-  Most of the attributes \index{attributes} introduced originally by \globus\ RSL~1.0 are
+  Most of the job description attributes \index{attributes} introduced originally by \globus\ RSL~1.0 are
   supported, some with modifications as indicated in this
-  document. Many new attributes are introduced, of which some are
+  document. Many new attributes are introduced by ARC, of which some are
   to be specified in the user's script, and others are internal for
-  the GM (are added or modified by the UI).
+  the GM (are added or modified by client tools).
 
@@
 
+  It is possible to use unsupported attributes in job description.
+Standard ARC client submission commands (\texttt{ngsub} and \texttt{ngresub}) must be used with a command line option ``-U'' in order to accept unknown attributes.
+Without this command line option, the client tool will concider job description invalid if it contains unsupported attributes.
+
   \section{User-side attributes}
@@
   The following attributes \index{attributes:user-side} can be
-  specified in a user's xRSL script. Some are to be modified by the UI
-  before being passed to the GM. If this is the case, corresponding GM input is described.
+  specified in a user's xRSL script. Some have to be modified by the client tool
+  before being passed to the GM. If this is the case, the corresponding modified GM input is described in this document as well.
 
@@
 
-  The following attributes are constructed by the UI and are parsed to
-  the GM, although users may specify them manually (\textbf{not
-  advised!}).
+  The following attributes are a part of the internal ARC job representation, and must be provided by ARC client tools and passed to
+  the GM. Users may even specify them manually (\textbf{not
+  advised!}), while developers of new ARC client tools and utilities must make sure these attributes are added to the user job description before it is submitted to a GM.
 
@@
   The following \globus\ attributes are not supported by the ARC
-  middleware. Whenever they are specified, a warning is issued by the
-  UI, and the corresponding attribute is ignored.
+  middleware. Whenever they are specified, eiher an error or a warning should be issued by the
+  client tool, and the corresponding attribute should be ignored\footnote{Standard ARC client will throw an error, unless option ''-U'' is specified}.
 
@@
      (notify="bqfe john.doe@gmail.com jane.doe@mail.org")
-(* specific queue to submit the job *)
-     (queue="atlas")
 (* maximal CPU time required for the job, minutes for PBS*)
@@
      (lifeTime="7")
- (* memory required for the job, Mbytes *)
+ (* memory required for the job, per rank, Mbytes *)
      (Memory="200")
@@
 (* required run-time environment *)
-     (runTimeEnvironment="Atlas-1.1")
+     (runTimeEnvironment="APPS/HEP/Atlas-1.1")
 (* number of re-runs, in case of a system failure *)
@@
 
-Note that the client tool expands all the \texttt{rsl\_substitution}
-values and adds double quotes to \textbf{all} strings. It also
+Note that the client tool must expand all the \texttt{rsl\_substitution}
+values and add double quotes to \textbf{all} strings.
+Client tool is also expected to insert \texttt{queue} attribute in case such is missing and to add checksum values for input files to be uploaded by the client.
+Standard ARC client also
 streamlines blank spaces.
@@
      ("lifetime" = "604800" )
-(* memory required for the job, Mbytes *)
+(* memory required for the job, per rank, Mbytes *)
      ("memory" = "200" )
@@
 (* required run-time environment *)
-     ("runtimeenvironment" = "Atlas-1.1" )
+     ("runtimeenvironment" = "APPS/HEP/Atlas-1.1" )
 (* number of re-runs, in case of a system failure *)
@@
 <jsdl:JobDefinition
-    xmlns:jsdl="http://schemas.ggf.org/jsdl/2005/11/jsdl"
-    xmlns:jsdl-posix="http://schemas.ggf.org/jsdl/2005/11/jsdl-posix"
-    xmlns:jsdl-arc="http://www.nordugrid.org/ws/schemas/jsdl-arc">
- <jsdl:JobDescription>
+  xmlns="http://schemas.ggf.org/jsdl/2005/11/jsdl"
+  xmlns:posix="http://schemas.ggf.org/jsdl/2005/11/jsdl-posix"
+  xmlns:arc="http://www.nordugrid.org/ws/schemas/jsdl-arc">
+  <JobDescription>
     <!-- Generic job description (jobname in XRSL) -->
-    <jsdl:JobIdentification>
-        <jsdl:JobName>JSDL Hello World job</jsdl:JobName>
-        <jsdl:Description>
+    <JobIdentification>
+        <JobName>JSDL Hello World job</JobName>
+        <Description>
           This is a simple Hello World job test
-        </jsdl:Description>
-    </jsdl:JobIdentification>
+        </Description>
+    </JobIdentification>
     <!-- Standard application (same as executable, arguments in XRSL) -->
-    <jsdl:Application>
-        <jsdl:ApplicationName>Echo</jsdl:ApplicationName>
-        <jsdl-posix:POSIXApplication>
-          <jsdl-posix:Executable>/bin/echo</jsdl-posix:Executable>
-          <jsdl-posix:Argument>Hello world</jsdl-posix:Argument>
-        </jsdl-posix:POSIXApplication>
-    </jsdl:Application>
+    <Application>
+        <ApplicationName>Echo</ApplicationName>
+        <posix:POSIXApplication>
+          <posix:Executable>/bin/echo</posix:Executable>
+          <posix:Argument>Hello world</posix:Argument>
+        </posix:POSIXApplication>
+    </Application>
     <!-- Get the GM diagnostics (same as gmlog in XRSL) -->
-    <jsdl-arc:LocalLogging>
-	      <jsdl-arc:Directory>gmlog</jsdl-arc:Directory>
-    </jsdl-arc:LocalLogging>
- </jsdl:JobDescription>
-</jsdl:JobDefinition>
+    <arc:LocalLogging>
+	      <arc:Directory>gmlog</arc:Directory>
+    </arc:LocalLogging>
+ </JobDescription>
+</JobDefinition>
 

### 2011-01-11 doc/tech_doc/xrsl/xrsl-gm.tex
Finalising porting "as is" of 4 documents from arc0 to arc1 tree

@@
   \begin{tabular}{lp{13cm}}
-    User input:&\verb##\\
     GM input:&\verb#(sstdin=<filename>)#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
   \begin{tabular}{lp{13cm}}
-    User input:&\verb##\\
-    GM input:&\verb#(action="request"|"cancel"|"clean")#\\
+    GM input:&\verb#(action="request"|"cancel"|"clean"|"renew"|"restart")#\\
     Example:&\verb#(action="request")#\\
@@
   Action to be taken by the gatekeeper: submit the job, cancel job
-  execution, or clear the results of the job (also cancels the job).
+  execution, clear the results of the job (also cancels the job),
+  renew the proxy of the job, or restart the job from a previous
+  failed state.
 
@@
   \begin{tabular}{lp{13cm}}
-    User input:&\verb##\\
     GM input:&\verb#(savestate="yes"|"no")#\\
@@
   \begin{tabular}{lp{13cm}}
-    User input:&\verb##\\
     GM input:&\verb#(lrmstype=<string>)#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
   \begin{tabular}{lp{13cm}}
-    User input:&\verb##\\
     GM input:&\verb#(hostname=<string>)#\\
@@
   \end{shaded}
-  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6}
+  \versions{ARC 0.3, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
   \begin{tabular}{lp{13cm}}
-    User input:&\verb##\\
     GM input:&\verb#(jobid=<string>)#\\
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.28, ARC 0.4, ARC 0.5, ARC 0.6}
+  \versions{ARC $\geq$ 0.3.28, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
   \begin{tabular}{lp{13cm}}
-    User input:&\verb##\\
     GM input:&\verb#(clientxrsl=<string>)#\\
@@
   \end{shaded}
-  \versions{ARC $\geq$ 0.3.35, ARC 0.4, ARC 0.5, ARC 0.6}
+  \versions{ARC $\geq$ 0.3.35, ARC 0.4, ARC 0.5, ARC 0.6, ARC 0.8}
   \begin{tabular}{lp{13cm}}
-    User input:&\verb##\\
     GM input:&\verb#(clientsoftware=<string>)#\\

### 2011-01-18 doc/tutorials/ws-programming-tutorial/doc/tex_appendix/appendix.tex
cut connections to source tree

--- 2009-05-14 doc/tutorials/ws-programming-tutorial/doc/tex_appendix/appendix.tex
@@
 	caption={[XML schema of configuration files.]\textbf{XML schema of configuration files.}}
-	]{../../../../src/hed/libs/loader/mcc.xsd}
+	]{tex_appendix/loader.xsd}
 
@@
 	caption={[XML schema of the TCP component.]\textbf{XML schema of the TCP component.}}
-	]{../../../../src/hed/mcc/tcp/tcp.xsd}
+	]{tex_appendix/tcp.xsd}
 
@@
 	[
-	label=lst:tls.xssd,
+	label=lst:tls.xsd,
 	caption={[XML schema of the TLS component.]\textbf{XML schema of  the TLS component.}}
-	]{../../../../src/hed/mcc/tls/tls.xsd}
+	]{tex_appendix/tls.xsd}
 
@@
 	[
-	label=lst:http_xsd,
+	label=lst:http.xsd,
 	caption={[XML schema of the HTTP component.]\textbf{XML schema of the HTTP component.}}
-	]{../../../../src/hed/mcc/http/http.xsd}
+	]{tex_appendix/http.xsd}
 
